### 2021-04-05 site/03-meeting.Rmd
init site

# Meetings

We describe our methods in this chapter.

### 2021-04-05 site/_book/site.tex
init site

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{book}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={PhD Notebook},
  pdfauthor={Patrick Li},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{booktabs}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{apalike}

\title{PhD Notebook}
\author{Patrick Li}
\date{2021-04-06}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{1}
\tableofcontents
}
\hypertarget{introduction}{%
\chapter{Introduction}\label{introduction}}

This notebook consists of:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  records of weekly meetings
\item
  literature review
\item
  to-do list
\item
  milestones
\item
  links to resources
\end{enumerate}

\hypertarget{literature}{%
\chapter{Literature}\label{literature}}

Here is a review of existing methods.

\hypertarget{meetings}{%
\chapter{Meetings}\label{meetings}}

We describe our methods in this chapter.

\hypertarget{todo}{%
\chapter{TODO}\label{todo}}

\hypertarget{milestones}{%
\chapter{Milestones}\label{milestones}}

  \bibliography{book.bib,packages.bib}

\end{document}

### 2021-04-06 site/02-literature.Rmd
update site

# Literature

## Graphical Inference for Infovis - Hadley Wickham, Dianne Cook, Heike Hofmann and Andreas Buja (2010)

### Section 1 - Introduction

In this section, the paper argues that much research in Infovis and most statistical methods are on two extremes. One focuses on finding relationships. Another focuses on checking whether a relationship really exists.

Some previous attempts of graphical inference are mentioned in this section. The 2009 paper which develops the statistical concepts of graphical inference is also mentioned.

It then summarises the structure of the paper.

### Section 2 - What is inference and why do we need it?

This section starts by stating the two components of statistical inference, which are testing and estimation. It then defines the meaning of inference in graphics.

It uses the criminal justice system as an example to explain the principles of hypothesis testing, and points out the difference between traditional testing and visual testing, which are the test statistic and the mechanism of computing similarity.

To be able to produce a visual test, null datasets which are samples from the null distribution are required. A null plot is a plot of a null dataset.

It ends the section by making an important argument that the introduction of the visual test is not meant to replace traditional tests. “Traditional statistical tests are well studied, well-formulated and work best when data is well-behaved, following a known distribution in relatively simple scenarios.” However, traditional statistical tests do not cover all of the aspects we want to test. On the other hand, visual tests can be used in complex data analysis setting when traditional tests are unavailable.

### Section 3 - Protocols of graphical inference

This section introduces two protocols for graphical inference: the “Rorshach” and the “line-up”.

The Rorschach protocol is used to calibrate our vision. In most of the imte, it contains only null plots.

The line-up protocol consists of a plot of the true data and n-1 null plots. N is typically set to be 19.

### Section 4 - Examples

To use the line-up protocol, we need to:
- Identify the question the plot is trying to answer
- Characterize the null-hypothesis
- Figure out how to generate null datasets

It provides a list of questions that we would like to answer in commonly seen data plots.

The null data can be generated in two ways in most cases: resampling and simulation.

One of the benefits of using permutation is it preserves the marginal distribution of each variable.

### Section 5 - Power

The power of the visual test depends on many factors. An appropriate choice of plot is important to increase the power. For large datasets, aggregation is one of the considerations. It can significantly increase the power of the test.

### Section 6 - Use

One can implement these two protocols using the R package `nullabor`.

### Section 7 - Conclusion

“Graphical inference is important because it helps us to avoid false convictions.”

```
@article{Wickham2010,
abstract = {How do we know if what we see is really there? When visualizing data, how do we avoid falling into the trap of apophenia where we see patterns in random noise? Traditionally, infovis has been concerned with discovering new relationships, and statistics with preventing spurious relationships from being reported. We pull these opposing poles closer with two new techniques for rigorous statistical inference of visual discoveries. The "Rorschach" helps the analyst calibrate their understanding of uncertainty and "line-up" provides a protocol for assessing the significance of visual discoveries, protecting against the discovery of spurious structure. {\textcopyright} 2006 IEEE.},
author = {Wickham, Hadley and Cook, Dianne and Hofmann, Heike and Buja, Andreas},
doi = {10.1109/TVCG.2010.161},
issn = {10772626},
journal = {IEEE Transactions on Visualization and Computer Graphics},
keywords = {Statistics,data plot,null hypotheses,permutation tests,visual testing},
number = {6},
pages = {973--979},
pmid = {20975134},
title = {{Graphical inference for infovis}},
volume = {16},
year = {2010}
}
```

## Graphical Perception: Theory, Experimentation, and Application to the Development of Graphical Methods - William S. Cleveland and Rovert McGill (1984)

```
@article{Cleveland1984,
abstract = {The subject of graphical methods for data analysis and for data presentation needs a scientific foundation. In this article we take a few steps in the direction of establishing such a foundation. Our approach is based on graphical perception-the visual decoding of information encoded on graphs-and it includes both theory and experimentation to test the theory. The theory deals with a small but important piece of the whole process of graphical perception. The first part is an identification of a set of elementary perceptual tasks that are carried out when people extract quantitative information from graphs. The second part is an ordering of the tasks on the basis of how accurately people perform them. Elements of the theory are tested by experimentation in which subjects record their judgments of the quantitative information on graphs. The experiments validate these elements but also suggest that the set of elementary tasks should be expanded. The theory provides a guideline for graph construction: Graphs should employ elementary tasks as high in the ordering as possible. This principle is applied to a variety of graphs, including bar charts, divided bar charts, pie charts, and statistical maps with shading. The conclusion is that radical surgery on these popular graphs is needed, and as replacements we offer alternative graphical forms-dot charts, dot charts with grouping, and framed-rectangle charts.},
author = {Cleveland, William S. and McGill, Robert},
doi = {10.2307/2288400},
issn = {01621459},
journal = {Journal of the American Statistical Association},
month = {sep},
number = {387},
pages = {531--554},
publisher = {JSTOR},
title = {{Graphical Perception: Theory, Experimentation, and Application to the Development of Graphical Methods}},
volume = {79},
year = {1984}
}
```

## Others

```
@book{wickham2016ggplot2,
author = {Wickham, Hadley},
file = {:Users/patrickli/Desktop/papers/(Use R!) Hadley Wickham - ggplot2_ Elegant Graphics for Data Analysis-Springer (2016).pdf:pdf},
publisher = {springer},
title = {ggplot2: elegant graphics for data analysis},
year = {2016}
}

@book{Wilkinson2005,
abstract = {The grammar of graphics (GoG) denotes a system with seven classes embedded in a data flow. This data flow specifies a strict order in which data are transformed from a raw dataset to a statistical graphic. Each class contains multiple methods, each of which is a function executed at the step in the data flow corresponding to that class. The classes are orthogonal, in the sense that the product set of all classes (every possible sequence of class methods) defines a space of graphics which is meaningful at every point. The meaning of a statistical graphic is thus determined by the mapping produced by the function chain linking data and graphic.},
address = {New York},
author = {Wilkinson, Leland},
booktitle = {The Grammar of Graphics},
doi = {10.1007/0-387-28695-0},
edition = {2},
file = {:Users/patrickli/Desktop/papers/(Statistics and Computing) Leland Wilkinson, D. Wills, D. Rope, A. Norton, R. Dubbs - The Grammar of Graphics-Springer (2005).pdf:pdf},
publisher = {Springer-Verlag},
title = {{The Grammar of Graphics}},
year = {2005}
}
@article{Zhao2013MindRU,
abstract = {Graphical statistics plays a very important job in research and industry. As a statistician, it is very useful that if one can see how people make their decision based on the plots, so that plots can be improved for better performance. With the help of eye-tracking equipment, researchers could show people several plots and ask a question on each, then track the person's eyes to see how they going through the plots to come to their answer. In this paper, the process and results of an experiment on watching what people were looking at in statistical plots will be discussed. This experiment is part of a larger experiment studying decision making and signal strength in statistical graphics, that uses Amazon's Mechanical Turk.},
author = {Zhao, Y and Cook, D and Hofmann, H and Majumder, M and Chowdhury, N R},
file = {:Users/patrickli/Desktop/papers/Mind Reading Using an Eyetracker to See How People Are Looking at Lineups.pdf:pdf},
journal = {International Journal of Intelligent Technologies and Applied Statistics},
number = {4},
pages = {393--413},
title = {{Mind Reading: Using an Eye-Tracker to See How People are Looking at Lineups}},
volume = {6},
year = {2013}
}
@article{Hofmann2012,
abstract = {Lineups [4, 28] have been established as tools for visual testing similar to standard statistical inference tests, allowing us to evaluate the validity of graphical findings in an objective manner. In simulation studies [12] lineups have been shown as being efficient: the power of visual tests is comparable to classical tests while being much less stringent in terms of distributional assumptions made. This makes lineups versatile, yet powerful, tools in situations where conditions for regular statistical tests are not or cannot be met. In this paper we introduce lineups as a tool for evaluating the power of competing graphical designs. We highlight some of the theoretical properties and then show results from two studies evaluating competing designs: both studies are designed to go to the limits of our perceptual abilities to highlight differences between designs. We use both accuracy and speed of evaluation as measures of a successful design. The first study compares the choice of coordinate system: polar versus cartesian coordinates. The results show strong support in favor of cartesian coordinates in finding fast and accurate answers to spotting patterns. The second study is aimed at finding shift differences between distributions. Both studies are motivated by data problems that we have recently encountered, and explore using simulated data to evaluate the plot designs under controlled conditions. Amazon Mechanical Turk (MTurk) is used to conduct the studies. The lineups provide an effective mechanism for objectively evaluating plot designs. {\textcopyright} 1995-2012 IEEE.},
author = {Hofmann, Heike and Follett, Lendie and Majumder, Mahbubul and Cook, Dianne},
doi = {10.1109/TVCG.2012.230},
file = {:Users/patrickli/Desktop/papers/Graphical Tests for Power Comparison of Competing Designs.pdf:pdf},
issn = {10772626},
journal = {IEEE Transactions on Visualization and Computer Graphics},
keywords = {Efficiency of displays,Lineups,Power comparison,Visual inference},
number = {12},
pages = {2441--2448},
title = {{Graphical tests for power comparison of competing designs}},
volume = {18},
year = {2012}
}

@article{RoyChowdhury2015,
abstract = {Statistical graphics play an important role in exploratory data analysis, model checking and diagnosis. With high dimensional data, this often means plotting low-dimensional projections, for example, in classification tasks projection pursuit is used to find low-dimensional projections that reveal differences between labelled groups. In many contemporary data sets the number of observations is relatively small compared to the number of variables, which is known as a high dimension low sample size (HDLSS) problem. This paper explores the use of visual inference on understanding low-dimensional pictures of HDLSS data. Visual inference helps to quantify the significance of findings made from graphics. This approach may be helpful to broaden the understanding of issues related to HDLSS data in the data analysis community. Methods are illustrated using data from a published paper, which erroneously found real separation in microarray data, and with a simulation study conducted using Amazon's Mechanical Turk.},
author = {{Roy Chowdhury}, Niladri and Cook, Dianne and Hofmann, Heike and Majumder, Mahbubul and Lee, Eun Kyung and Toth, Amy L.},
doi = {10.1007/s00180-014-0534-x},
file = {:Users/patrickli/Desktop/papers/RoyChowdhury2015_Article_UsingVisualStatisticalInferenc.pdf:pdf},
issn = {16139658},
journal = {Computational Statistics},
keywords = {Data mining,Lineup,Projection pursuit,Statistical graphics,Visualization},
month = {nov},
number = {2},
pages = {293--316},
publisher = {Springer Verlag},
title = {{Using visual statistical inference to better understand random class separations in high dimension, low sample size data}},
url = {https://link.springer.com/article/10.1007/s00180-014-0534-x},
volume = {30},
year = {2015}
}
@article{Loy2017,
abstract = {The complexity of linear mixed-effects (LME) models means that traditional diagnostics are rendered less effective. This is due to a breakdown of asymptotic results, boundary issues, and visible patterns in residual plots that are introduced by the model fitting process. Some of these issues are well known and adjustments have been proposed. Working with LME models typically requires that the analyst keeps track of all the special circumstances that may arise. In this article, we illustrate a simpler but generally applicable approach to diagnosing LME models. We explain how to use new visual inference methods for these purposes. The approach provides a unified framework for diagnosing LME fits and for model selection. We illustrate the use of this approach on several commonly available datasets. A large-scale Amazon Turk study was used to validate the methods. R code is provided for the analyses. Supplementary materials for this article are available online.},
archivePrefix = {arXiv},
arxivId = {1502.06988},
author = {Loy, Adam and Hofmann, Heike and Cook, Dianne},
doi = {10.1080/10618600.2017.1330207},
eprint = {1502.06988},
file = {:Users/patrickli/Desktop/papers/Model Choice and Diagnostics for Linear Mixed Effects Models Using Statistics on Street Corners.pdf:pdf},
issn = {15372715},
journal = {Journal of Computational and Graphical Statistics},
keywords = {Lineup protocol,Model diagnostics,Model selection,Statistical graphics,Visual inference},
month = {jul},
number = {3},
pages = {478--492},
publisher = {American Statistical Association},
title = {{Model Choice and Diagnostics for Linear Mixed-Effects Models Using Statistics on Street Corners}},
url = {https://www.tandfonline.com/doi/abs/10.1080/10618600.2017.1330207},
volume = {26},
year = {2017}
}
@article{Majumder2013,
abstract = {Statistical graphics play a crucial role in exploratory data analysis, model checking, and diagnosis. The lineup protocol enables statistical significance testing of visual findings, bridging the gulf between exploratory and inferential statistics. In this article, inferential methods for statistical graphics are developed further by refining the terminology of visual inference and framing the lineup protocol in a context that allows direct comparison with conventional tests in scenarios when a conventional test exists. This framework is used to compare the performance of the lineup protocol against conventional statistical testing in the scenario of fitting linear models.Ahuman subjects experiment is conducted using simulated data to provide controlled conditions. Results suggest that the lineup protocol performs comparably with the conventional tests, and expectedly outperforms them when data are contaminated, a scenario where assumptions required for performing a conventional test are violated. Surprisingly, visual tests have higher power than the conventional tests when the effect size is large. And, interestingly, there may be some super-visual individuals who yield better performance and power than the conventional test even in the most difficult tasks. Supplementary materials for this article are available online. {\textcopyright} 2013 American Statistical Association.},
author = {Majumder, Mahbubul and Hofmann, Heike and Cook, Dianne},
doi = {10.1080/01621459.2013.808157},
file = {:Users/patrickli/Desktop/papers/Validation of Visual Statistical Inference Applied to Linear Models.pdf:pdf},
issn = {01621459},
journal = {Journal of the American Statistical Association},
keywords = {Data mining,Effect size,Exploratory data analysis,Lineup,Nonparametric test,Practical significance,Statistical graphics,Visualization},
number = {503},
pages = {942--956},
publisher = {Taylor & Francis Group},
title = {{Validation of visual statistical inference, applied to linear models}},
url = {https://www.tandfonline.com/doi/abs/10.1080/01621459.2013.808157},
volume = {108},
year = {2013}
}
@article{Chowdhury2018,
abstract = {Graphics play a crucial role in statistical analysis and data mining. Being able to quantify structure in data that is visible in plots, and how people read the structure from plots is an ongoing challenge. The lineup protocol provides a formal framework for data plots, making inference possible. The data plot is treated like a test statistic, and lineup protocol acts like a comparison with the sampling distribution of the nulls. This article describes metrics for describing structure in data plots and evaluates them in relation to the choices that human readers made during several large Amazon Turk studies using lineups. The metrics that were more specific to the plot types tended to better match subject choices, than generic metrics. The process that we followed to evaluate metrics will be useful for general development of numerically measuring structure in plots, and also in future experiments on lineups for choosing blocks of pictures. Supplementary materials for this article are available online.},
author = {Chowdhury, Niladri Roy and Cook, Dianne and Hofmann, Heike and Majumder, Mahbubul},
doi = {10.1080/10618600.2017.1356323},
file = {:Users/patrickli/Desktop/papers/Measuring Lineup Difficulty By Matching Distance Metrics With Subject Choices in Crowd Sourced Data.pdf:pdf},
issn = {15372715},
journal = {Journal of Computational and Graphical Statistics},
keywords = {Cognitive perception,Data mining,Data science,Data visualization,Distance metrics,Exploratory data analysis,Information visualization,Statistical graphics,Visual inference},
month = {jan},
number = {1},
pages = {132--145},
publisher = {American Statistical Association},
title = {{Measuring Lineup Difficulty By Matching Distance Metrics With Subject Choices in Crowd-Sourced Data}},
url = {https://www.tandfonline.com/doi/abs/10.1080/10618600.2017.1356323},
volume = {27},
year = {2018}
}
@article{Buja2009,
abstract = {We propose to furnish visual statistical methods with an inferential framework and protocol, modelled on confirmatory statistical testing. In this framework, plots take on the role of test statistics, and human cognition the role of statistical tests. Statistical significance of 'discoveries' is measured by having the human viewer compare the plot of the real dataset with collections of plots of simulated datasets. A simple but rigorous protocol that provides inferential validity is modelled after the 'lineup' popular from criminal legal procedures. Another protocol modelled after the 'Rorschach' inkblot test, well known from (pop-)psychology, will help analysts acclimatize to random variability before being exposed to the plot of the real data. The proposed protocols will be useful for exploratory data analysis, with reference datasets simulated by using a null assumption that structure is absent. The framework is also useful for model diagnostics in which case reference datasets are simulated from the model in question. This latter point follows up on previous proposals. Adopting the protocols will mean an adjustment in working procedures for data analysts, adding more rigour, and teachers might find that incorporating these protocols into the curriculum improves their students' statistical thinking. This journal is {\textcopyright} 2009 The Royal Society.},
author = {Buja, Andreas and Cook, Dianne and Hofmann, Heike and Lawrence, Michael and Lee, Eun Kyung and Swayne, Deborah F. and Wickham, Hadley},
doi = {10.1098/rsta.2009.0120},
file = {:Users/patrickli/Desktop/papers/Statistical inference for exploratory data analysis and model diagnostics.pdf:pdf},
issn = {1364503X},
journal = {Philosophical Transactions of the Royal Society A: Mathematical, Physical and Engineering Sciences},
keywords = {Cognitive perception,Permutation tests,Rotation tests,Simulation,Statistical graphics,Visual data mining},
month = {nov},
number = {1906},
pages = {4361--4383},
pmid = {19805449},
publisher = {Royal Society},
title = {{Statistical inference for exploratory data analysis and model diagnostics}},
url = {https://royalsocietypublishing.org/doi/10.1098/rsta.2009.0120},
volume = {367},
year = {2009}
}
```

### 2021-04-06 site/04-to-do.Rmd
update site

# TODO

## Week 3

- [x] Check human subject experiment and Monash permission materials
	- https://www.intranet.monash/researchadmin/start/ethics/human
- [x] Check Kaiwen's project & paper
- [ ] Check Gallery of graphs - unclear author

## Week 4

- [x] Build prototype of html webpage to collect data
- [x] Send data to google sheet
- [x] Check summer-vis-inf 
- [x] Read examples sent by Di
- [ ] Build PhD repo
    - [x] meetings
    - [ ] paper
    - [x] TODO 
- [ ] Check Susan Vanderplas's website
    - [ ] paper
    - [ ] talks
    - [ ] posts

## Week 5

- [x] modify the webpage to be able to select multiple plots
- [x] Attempt to generate data from one assumed model
- [ ] draft Human Ethics Application Form

## Week 6

- [ ] Do the literature review of previous design
- [ ] Draw a flow chart to illustrate the design

### 2021-04-06 site/_book/site.tex
add index html
update site

@@
 }
+\hypertarget{welcome}{%
+\chapter{Welcome}\label{welcome}}
+
+I am Patrick Li.
+
 \hypertarget{introduction}{%
@@
 
-This notebook consists of:
+This note consists of:
 
@@
 
-Here is a review of existing methods.
+\hypertarget{graphical-inference-for-infovis---hadley-wickham-dianne-cook-heike-hofmann-and-andreas-buja-2010}{%
+\section{Graphical Inference for Infovis - Hadley Wickham, Dianne Cook, Heike Hofmann and Andreas Buja (2010)}\label{graphical-inference-for-infovis---hadley-wickham-dianne-cook-heike-hofmann-and-andreas-buja-2010}}
+
+\hypertarget{section-1---introduction}{%
+\subsection{Section 1 - Introduction}\label{section-1---introduction}}
+
+In this section, the paper argues that much research in Infovis and most statistical methods are on two extremes.
+One focuses on finding relationships.
+Another focuses on checking whether a relationship really exists.
+
+Some previous attempts of graphical inference are mentioned in this section. The 2009 paper which develops the statistical concepts of graphical inference is also mentioned.
+
+It then summarises the structure of the paper.
+
+\hypertarget{section-2---what-is-inference-and-why-do-we-need-it}{%
+\subsection{Section 2 - What is inference and why do we need it?}\label{section-2---what-is-inference-and-why-do-we-need-it}}
+
+This section starts by stating the two components of statistical inference, which are testing and estimation. It then defines the meaning of inference in graphics.
+
+It uses the criminal justice system as an example to explain the principles of hypothesis testing, and points out the difference between traditional testing and visual testing, which are the test statistic and the mechanism of computing similarity.
+
+To be able to produce a visual test, null datasets which are samples from the null distribution are required. A null plot is a plot of a null dataset.
+
+It ends the section by making an important argument that the introduction of the visual test is not meant to replace traditional tests.
+``Traditional statistical tests are well studied, well-formulated and work best when data is well-behaved, following a known distribution in relatively simple scenarios.'' However, traditional statistical tests do not cover all of the aspects we want to test.
+On the other hand, visual tests can be used in complex data analysis setting when traditional tests are unavailable.
+
+\hypertarget{section-3---protocols-of-graphical-inference}{%
+\subsection{Section 3 - Protocols of graphical inference}\label{section-3---protocols-of-graphical-inference}}
+
+This section introduces two protocols for graphical inference: the ``Rorshach'' and the ``line-up''.
+
+The Rorschach protocol is used to calibrate our vision. In most of the imte, it contains only null plots.
+
+The line-up protocol consists of a plot of the true data and n-1 null plots. N is typically set to be 19.
+
+\hypertarget{section-4---examples}{%
+\subsection{Section 4 - Examples}\label{section-4---examples}}
+
+To use the line-up protocol, we need to:
+- Identify the question the plot is trying to answer
+- Characterize the null-hypothesis
+- Figure out how to generate null datasets
+
+It provides a list of questions that we would like to answer in commonly seen data plots.
+
+The null data can be generated in two ways in most cases: resampling and simulation.
+
+One of the benefits of using permutation is it preserves the marginal distribution of each variable.
+
+\hypertarget{section-5---power}{%
+\subsection{Section 5 - Power}\label{section-5---power}}
+
+The power of the visual test depends on many factors.
+An appropriate choice of plot is important to increase the power.
+For large datasets, aggregation is one of the considerations.
+It can significantly increase the power of the test.
+
+\hypertarget{section-6---use}{%
+\subsection{Section 6 - Use}\label{section-6---use}}
+
+One can implement these two protocols using the R package \texttt{nullabor}.
+
+\hypertarget{section-7---conclusion}{%
+\subsection{Section 7 - Conclusion}\label{section-7---conclusion}}
+
+``Graphical inference is important because it helps us to avoid false convictions.''
+
+\hypertarget{bib}{%
+\subsection{Bib}\label{bib}}
+
+\begin{verbatim}
+@article{Wickham2010,
+abstract = {How do we know if what we see is really there?
+When visualizing data, how do we avoid falling into the trap of apophenia where we see patterns in random noise?
+Traditionally, infovis has been concerned with discovering new relationships, and statistics with preventing spurious relationships from being reported.
+We pull these opposing poles closer with two new techniques for rigorous statistical inference of visual discoveries.
+The "Rorschach" helps the analyst calibrate their understanding of uncertainty and "line-up" provides a protocol for assessing the significance of visual discoveries, protecting against the discovery of spurious structure.
+{\textcopyright} 2006 IEEE.},
+author = {Wickham, Hadley and Cook, Dianne and Hofmann, Heike and Buja, Andreas},
+doi = {10.1109/TVCG.2010.161},
+issn = {10772626},
+journal = {IEEE Transactions on Visualization and Computer Graphics},
+keywords = {Statistics,data plot,null hypotheses,permutation tests,visual testing},
+number = {6},
+pages = {973--979},
+pmid = {20975134},
+title = {{Graphical inference for infovis}},
+volume = {16},
+year = {2010}
+}
+\end{verbatim}
+
+\hypertarget{graphical-perception-theory-experimentation-and-application-to-the-development-of-graphical-methods---william-s.-cleveland-and-rovert-mcgill-1984}{%
+\section{Graphical Perception: Theory, Experimentation, and Application to the Development of Graphical Methods - William S.
+Cleveland and Rovert McGill (1984)}\label{graphical-perception-theory-experimentation-and-application-to-the-development-of-graphical-methods---william-s.-cleveland-and-rovert-mcgill-1984}}
+
+\begin{verbatim}
+@article{Cleveland1984,
+abstract = {The subject of graphical methods for data analysis and for data presentation needs a scientific foundation.
+In this article we take a few steps in the direction of establishing such a foundation.
+Our approach is based on graphical perception-the visual decoding of information encoded on graphs-and it includes both theory and experimentation to test the theory.
+The theory deals with a small but important piece of the whole process of graphical perception.
+The first part is an identification of a set of elementary perceptual tasks that are carried out when people extract quantitative information from graphs.
+The second part is an ordering of the tasks on the basis of how accurately people perform them.
+Elements of the theory are tested by experimentation in which subjects record their judgments of the quantitative information on graphs.
+The experiments validate these elements but also suggest that the set of elementary tasks should be expanded.
+The theory provides a guideline for graph construction: Graphs should employ elementary tasks as high in the ordering as possible.
+This principle is applied to a variety of graphs, including bar charts, divided bar charts, pie charts, and statistical maps with shading.
+The conclusion is that radical surgery on these popular graphs is needed, and as replacements we offer alternative graphical forms-dot charts, dot charts with grouping, and framed-rectangle charts.},
+author = {Cleveland, William S. and McGill, Robert},
+doi = {10.2307/2288400},
+issn = {01621459},
+journal = {Journal of the American Statistical Association},
+month = {sep},
+number = {387},
+pages = {531--554},
+publisher = {JSTOR},
+title = {{Graphical Perception: Theory, Experimentation, and Application to the Development of Graphical Methods}},
+volume = {79},
+year = {1984}
+}
+\end{verbatim}
+
+\hypertarget{others}{%
+\section{Others}\label{others}}
+
+\begin{verbatim}
+@book{wickham2016ggplot2,
+author = {Wickham, Hadley},
+file = {:Users/patrickli/Desktop/papers/(Use R!) Hadley Wickham - ggplot2_ Elegant Graphics for Data Analysis-Springer (2016).pdf:pdf},
+publisher = {springer},
+title = {ggplot2: elegant graphics for data analysis},
+year = {2016}
+}
+
+@book{Wilkinson2005,
+abstract = {The grammar of graphics (GoG) denotes a system with seven classes embedded in a data flow.
+This data flow specifies a strict order in which data are transformed from a raw dataset to a statistical graphic.
+Each class contains multiple methods, each of which is a function executed at the step in the data flow corresponding to that class.
+The classes are orthogonal, in the sense that the product set of all classes (every possible sequence of class methods) defines a space of graphics which is meaningful at every point.
+The meaning of a statistical graphic is thus determined by the mapping produced by the function chain linking data and graphic.},
+address = {New York},
+author = {Wilkinson, Leland},
+booktitle = {The Grammar of Graphics},
+doi = {10.1007/0-387-28695-0},
+edition = {2},
+file = {:Users/patrickli/Desktop/papers/(Statistics and Computing) Leland Wilkinson, D. Wills, D. Rope, A. Norton, R. Dubbs - The Grammar of Graphics-Springer (2005).pdf:pdf},
+publisher = {Springer-Verlag},
+title = {{The Grammar of Graphics}},
+year = {2005}
+}
+@article{Zhao2013MindRU,
+abstract = {Graphical statistics plays a very important job in research and industry.
+As a statistician, it is very useful that if one can see how people make their decision based on the plots, so that plots can be improved for better performance.
+With the help of eye-tracking equipment, researchers could show people several plots and ask a question on each, then track the person's eyes to see how they going through the plots to come to their answer.
+In this paper, the process and results of an experiment on watching what people were looking at in statistical plots will be discussed.
+This experiment is part of a larger experiment studying decision making and signal strength in statistical graphics, that uses Amazon's Mechanical Turk.},
+author = {Zhao, Y and Cook, D and Hofmann, H and Majumder, M and Chowdhury, N R},
+file = {:Users/patrickli/Desktop/papers/Mind Reading Using an Eyetracker to See How People Are Looking at Lineups.pdf:pdf},
+journal = {International Journal of Intelligent Technologies and Applied Statistics},
+number = {4},
+pages = {393--413},
+title = {{Mind Reading: Using an Eye-Tracker to See How People are Looking at Lineups}},
+volume = {6},
+year = {2013}
+}
+@article{Hofmann2012,
+abstract = {Lineups [4, 28] have been established as tools for visual testing similar to standard statistical inference tests, allowing us to evaluate the validity of graphical findings in an objective manner.
+In simulation studies [12] lineups have been shown as being efficient: the power of visual tests is comparable to classical tests while being much less stringent in terms of distributional assumptions made.
+This makes lineups versatile, yet powerful, tools in situations where conditions for regular statistical tests are not or cannot be met.
+In this paper we introduce lineups as a tool for evaluating the power of competing graphical designs.
+We highlight some of the theoretical properties and then show results from two studies evaluating competing designs: both studies are designed to go to the limits of our perceptual abilities to highlight differences between designs.
+We use both accuracy and speed of evaluation as measures of a successful design.
+The first study compares the choice of coordinate system: polar versus cartesian coordinates.
+The results show strong support in favor of cartesian coordinates in finding fast and accurate answers to spotting patterns.
+The second study is aimed at finding shift differences between distributions.
+Both studies are motivated by data problems that we have recently encountered, and explore using simulated data to evaluate the plot designs under controlled conditions.
+Amazon Mechanical Turk (MTurk) is used to conduct the studies.
+The lineups provide an effective mechanism for objectively evaluating plot designs.
+{\textcopyright} 1995-2012 IEEE.},
+author = {Hofmann, Heike and Follett, Lendie and Majumder, Mahbubul and Cook, Dianne},
+doi = {10.1109/TVCG.2012.230},
+file = {:Users/patrickli/Desktop/papers/Graphical Tests for Power Comparison of Competing Designs.pdf:pdf},
+issn = {10772626},
+journal = {IEEE Transactions on Visualization and Computer Graphics},
+keywords = {Efficiency of displays,Lineups,Power comparison,Visual inference},
+number = {12},
+pages = {2441--2448},
+title = {{Graphical tests for power comparison of competing designs}},
+volume = {18},
+year = {2012}
+}
+
+@article{RoyChowdhury2015,
+abstract = {Statistical graphics play an important role in exploratory data analysis, model checking and diagnosis.
+With high dimensional data, this often means plotting low-dimensional projections, for example, in classification tasks projection pursuit is used to find low-dimensional projections that reveal differences between labelled groups.
+In many contemporary data sets the number of observations is relatively small compared to the number of variables, which is known as a high dimension low sample size (HDLSS) problem.
+This paper explores the use of visual inference on understanding low-dimensional pictures of HDLSS data.
+Visual inference helps to quantify the significance of findings made from graphics.
+This approach may be helpful to broaden the understanding of issues related to HDLSS data in the data analysis community.
+Methods are illustrated using data from a published paper, which erroneously found real separation in microarray data, and with a simulation study conducted using Amazon's Mechanical Turk.},
+author = {{Roy Chowdhury}, Niladri and Cook, Dianne and Hofmann, Heike and Majumder, Mahbubul and Lee, Eun Kyung and Toth, Amy L.},
+doi = {10.1007/s00180-014-0534-x},
+file = {:Users/patrickli/Desktop/papers/RoyChowdhury2015_Article_UsingVisualStatisticalInferenc.pdf:pdf},
+issn = {16139658},
+journal = {Computational Statistics},
+keywords = {Data mining,Lineup,Projection pursuit,Statistical graphics,Visualization},
+month = {nov},
+number = {2},
+pages = {293--316},
+publisher = {Springer Verlag},
+title = {{Using visual statistical inference to better understand random class separations in high dimension, low sample size data}},
+url = {https://link.springer.com/article/10.1007/s00180-014-0534-x},
+volume = {30},
+year = {2015}
+}
+@article{Loy2017,
+abstract = {The complexity of linear mixed-effects (LME) models means that traditional diagnostics are rendered less effective.
+This is due to a breakdown of asymptotic results, boundary issues, and visible patterns in residual plots that are introduced by the model fitting process.
+Some of these issues are well known and adjustments have been proposed.
+Working with LME models typically requires that the analyst keeps track of all the special circumstances that may arise.
+In this article, we illustrate a simpler but generally applicable approach to diagnosing LME models.
+We explain how to use new visual inference methods for these purposes.
+The approach provides a unified framework for diagnosing LME fits and for model selection.
+We illustrate the use of this approach on several commonly available datasets.
+A large-scale Amazon Turk study was used to validate the methods.
+R code is provided for the analyses.
+Supplementary materials for this article are available online.},
+archivePrefix = {arXiv},
+arxivId = {1502.06988},
+author = {Loy, Adam and Hofmann, Heike and Cook, Dianne},
+doi = {10.1080/10618600.2017.1330207},
+eprint = {1502.06988},
+file = {:Users/patrickli/Desktop/papers/Model Choice and Diagnostics for Linear Mixed Effects Models Using Statistics on Street Corners.pdf:pdf},
+issn = {15372715},
+journal = {Journal of Computational and Graphical Statistics},
+keywords = {Lineup protocol,Model diagnostics,Model selection,Statistical graphics,Visual inference},
+month = {jul},
+number = {3},
+pages = {478--492},
+publisher = {American Statistical Association},
+title = {{Model Choice and Diagnostics for Linear Mixed-Effects Models Using Statistics on Street Corners}},
+url = {https://www.tandfonline.com/doi/abs/10.1080/10618600.2017.1330207},
+volume = {26},
+year = {2017}
+}
+@article{Majumder2013,
+abstract = {Statistical graphics play a crucial role in exploratory data analysis, model checking, and diagnosis.
+The lineup protocol enables statistical significance testing of visual findings, bridging the gulf between exploratory and inferential statistics.
+In this article, inferential methods for statistical graphics are developed further by refining the terminology of visual inference and framing the lineup protocol in a context that allows direct comparison with conventional tests in scenarios when a conventional test exists.
+This framework is used to compare the performance of the lineup protocol against conventional statistical testing in the scenario of fitting linear models.Ahuman subjects experiment is conducted using simulated data to provide controlled conditions.
+Results suggest that the lineup protocol performs comparably with the conventional tests, and expectedly outperforms them when data are contaminated, a scenario where assumptions required for performing a conventional test are violated.
+Surprisingly, visual tests have higher power than the conventional tests when the effect size is large.
+And, interestingly, there may be some super-visual individuals who yield better performance and power than the conventional test even in the most difficult tasks.
+Supplementary materials for this article are available online.
+{\textcopyright} 2013 American Statistical Association.},
+author = {Majumder, Mahbubul and Hofmann, Heike and Cook, Dianne},
+doi = {10.1080/01621459.2013.808157},
+file = {:Users/patrickli/Desktop/papers/Validation of Visual Statistical Inference Applied to Linear Models.pdf:pdf},
+issn = {01621459},
+journal = {Journal of the American Statistical Association},
+keywords = {Data mining,Effect size,Exploratory data analysis,Lineup,Nonparametric test,Practical significance,Statistical graphics,Visualization},
+number = {503},
+pages = {942--956},
+publisher = {Taylor & Francis Group},
+title = {{Validation of visual statistical inference, applied to linear models}},
+url = {https://www.tandfonline.com/doi/abs/10.1080/01621459.2013.808157},
+volume = {108},
+year = {2013}
+}
+@article{Chowdhury2018,
+abstract = {Graphics play a crucial role in statistical analysis and data mining.
+Being able to quantify structure in data that is visible in plots, and how people read the structure from plots is an ongoing challenge.
+The lineup protocol provides a formal framework for data plots, making inference possible.
+The data plot is treated like a test statistic, and lineup protocol acts like a comparison with the sampling distribution of the nulls.
+This article describes metrics for describing structure in data plots and evaluates them in relation to the choices that human readers made during several large Amazon Turk studies using lineups.
+The metrics that were more specific to the plot types tended to better match subject choices, than generic metrics.
+The process that we followed to evaluate metrics will be useful for general development of numerically measuring structure in plots, and also in future experiments on lineups for choosing blocks of pictures.
+Supplementary materials for this article are available online.},
+author = {Chowdhury, Niladri Roy and Cook, Dianne and Hofmann, Heike and Majumder, Mahbubul},
+doi = {10.1080/10618600.2017.1356323},
+file = {:Users/patrickli/Desktop/papers/Measuring Lineup Difficulty By Matching Distance Metrics With Subject Choices in Crowd Sourced Data.pdf:pdf},
+issn = {15372715},
+journal = {Journal of Computational and Graphical Statistics},
+keywords = {Cognitive perception,Data mining,Data science,Data visualization,Distance metrics,Exploratory data analysis,Information visualization,Statistical graphics,Visual inference},
+month = {jan},
+number = {1},
+pages = {132--145},
+publisher = {American Statistical Association},
+title = {{Measuring Lineup Difficulty By Matching Distance Metrics With Subject Choices in Crowd-Sourced Data}},
+url = {https://www.tandfonline.com/doi/abs/10.1080/10618600.2017.1356323},
+volume = {27},
+year = {2018}
+}
+@article{Buja2009,
+abstract = {We propose to furnish visual statistical methods with an inferential framework and protocol, modelled on confirmatory statistical testing.
+In this framework, plots take on the role of test statistics, and human cognition the role of statistical tests.
+Statistical significance of 'discoveries' is measured by having the human viewer compare the plot of the real dataset with collections of plots of simulated datasets.
+A simple but rigorous protocol that provides inferential validity is modelled after the 'lineup' popular from criminal legal procedures.
+Another protocol modelled after the 'Rorschach' inkblot test, well known from (pop-)psychology, will help analysts acclimatize to random variability before being exposed to the plot of the real data.
+The proposed protocols will be useful for exploratory data analysis, with reference datasets simulated by using a null assumption that structure is absent.
+The framework is also useful for model diagnostics in which case reference datasets are simulated from the model in question.
+This latter point follows up on previous proposals.
+Adopting the protocols will mean an adjustment in working procedures for data analysts, adding more rigour, and teachers might find that incorporating these protocols into the curriculum improves their students' statistical thinking.
+This journal is {\textcopyright} 2009 The Royal Society.},
+author = {Buja, Andreas and Cook, Dianne and Hofmann, Heike and Lawrence, Michael and Lee, Eun Kyung and Swayne, Deborah F. and Wickham, Hadley},
+doi = {10.1098/rsta.2009.0120},
+file = {:Users/patrickli/Desktop/papers/Statistical inference for exploratory data analysis and model diagnostics.pdf:pdf},
+issn = {1364503X},
+journal = {Philosophical Transactions of the Royal Society A: Mathematical, Physical and Engineering Sciences},
+keywords = {Cognitive perception,Permutation tests,Rotation tests,Simulation,Statistical graphics,Visual data mining},
+month = {nov},
+number = {1906},
+pages = {4361--4383},
+pmid = {19805449},
+publisher = {Royal Society},
+title = {{Statistical inference for exploratory data analysis and model diagnostics}},
+url = {https://royalsocietypublishing.org/doi/10.1098/rsta.2009.0120},
+volume = {367},
+year = {2009}
+}
+\end{verbatim}
 
@@
 
-We describe our methods in this chapter.
+\hypertarget{march-10-2021---week-2}{%
+\section{March 10, 2021 - Week 2}\label{march-10-2021---week-2}}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Human subject experiment and Monash permission
+
+  \begin{itemize}
+  \tightlist
+  \item
+    There is a workshop in May
+  \end{itemize}
+\item
+  Use simulated data to set up an experiment
+\item
+  Check out Gallery of graphs (a name yan \ldots{} not sure)
+\item
+  The experiment could start from residual plot
+\item
+  Q-Q plot could also be considered
+\item
+  A revelant research - residual calculation by kaiwen - master project
+\item
+  Use Appen survey to collect data
+\item
+  Build a Github to-do list, meeting record and summary of the literature
+\item
+  There may be some development in the theory by Nancy Reid - theoretical statistician (recent work)
+\item
+  Consider using Kears to build a computer vision model
+\end{enumerate}
+
+\hypertarget{march-17-2021---week-3}{%
+\section{March 17, 2021 - Week 3}\label{march-17-2021---week-3}}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Read Susan Vanderplas's personal website to find additional information
+\item
+  Check out NUMBAT residual plot comparision - summer-vis-inf : Aarathy Babu - code examples
+\item
+  Read human subject premission examples (sent by Di)
+\item
+  Check out top-up application
+\item
+  Consider to use \texttt{Edibble} to set up the experiment
+\item
+  Build the PhD repo
+\item
+  Consider to use non-shiny framework
+\end{enumerate}
+
+\hypertarget{march-21-2021---week-4}{%
+\section{March 21, 2021 - Week 4}\label{march-21-2021---week-4}}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  A short meeting - late for 30 minutes
+\item
+  Aarathy introduces her repo
+\end{enumerate}
+
+\hypertarget{march-28-2021---week-5}{%
+\section{March 28, 2021 - Week 5}\label{march-28-2021---week-5}}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Discuss the options of building a alternative hypothesis in residual plot
+\item
+  AR, heterogeneity of variance, endogeneity, skewness, exp distribution, poisson distribution and missing covariance
+\item
+  Choice of plot design (loess or \(y=0\) line), number of lineup (5\textasciitilde15), number of observations
+\item
+  Use flow chart to illustrate the choices
+\item
+  Literature review - check previous designs
+\item
+  Build bookdown to track records
+\end{enumerate}
 
@@
 \chapter{TODO}\label{todo}}
+
+\hypertarget{week-3}{%
+\section{Week 3}\label{week-3}}
+
+\begin{itemize}
+\tightlist
+\item[$\boxtimes$]
+  Check human subject experiment and Monash permission materials
+
+  \begin{itemize}
+  \tightlist
+  \item
+    \url{https://www.intranet.monash/researchadmin/start/ethics/human}
+  \end{itemize}
+\item[$\boxtimes$]
+  Check Kaiwen's project \& paper
+\item[$\square$]
+  Check Gallery of graphs - unclear author
+\end{itemize}
+
+\hypertarget{week-4}{%
+\section{Week 4}\label{week-4}}
+
+\begin{itemize}
+\tightlist
+\item[$\boxtimes$]
+  Build prototype of html webpage to collect data
+\item[$\boxtimes$]
+  Send data to google sheet
+\item[$\boxtimes$]
+  Check summer-vis-inf
+\item[$\boxtimes$]
+  Read examples sent by Di
+\item[$\square$]
+  Build PhD repo
+
+  \begin{itemize}
+  \tightlist
+  \item[$\boxtimes$]
+    meetings
+  \item[$\square$]
+    paper
+  \item[$\boxtimes$]
+    TODO
+  \end{itemize}
+\item[$\square$]
+  Check Susan Vanderplas's website
+
+  \begin{itemize}
+  \tightlist
+  \item[$\square$]
+    paper
+  \item[$\square$]
+    talks
+  \item[$\square$]
+    posts
+  \end{itemize}
+\end{itemize}
+
+\hypertarget{week-5}{%
+\section{Week 5}\label{week-5}}
+
+\begin{itemize}
+\tightlist
+\item[$\boxtimes$]
+  modify the webpage to be able to select multiple plots
+\item[$\boxtimes$]
+  Attempt to generate data from one assumed model
+\item[$\square$]
+  draft Human Ethics Application Form
+\end{itemize}
+
+\hypertarget{week-6}{%
+\section{Week 6}\label{week-6}}
+
+\begin{itemize}
+\tightlist
+\item[$\square$]
+  Do the literature review of previous design
+\item[$\square$]
+  Draw a flow chart to illustrate the design
+\end{itemize}
 

### 2021-04-06 site/03-meeting.Rmd
update site

@@
 
-We describe our methods in this chapter.
+## March 10, 2021 - Week 2
+
+1. Human subject experiment and Monash permission
+    - There is a workshop in May
+2. Use simulated data to set up an experiment
+3. Check out Gallery of graphs (a name yan ... not sure)
+4. The experiment could start from residual plot
+7. Q-Q plot could also be considered
+8. A revelant research - residual calculation by kaiwen - master project
+9. Use Appen survey to collect data
+10. Build a Github to-do list, meeting record and summary of the literature
+11. There may be some development in the theory by Nancy Reid - theoretical statistician (recent work)
+12. Consider using Kears to build a computer vision model
+
+## March 17, 2021 - Week 3
+
+1. Read Susan Vanderplas's personal website to find additional information
+2. Check out NUMBAT residual plot comparision - summer-vis-inf : Aarathy Babu - code examples
+3. Read human subject premission examples (sent by Di)
+4. Check out top-up application
+5. Consider to use `Edibble` to set up the experiment
+6. Build the PhD repo
+7. Consider to use non-shiny framework
+
+## March 21, 2021 - Week 4
+
+1. A short meeting - late for 30 minutes
+2. Aarathy introduces her repo
+
+## March 28, 2021 - Week 5
+
+1. Discuss the options of building a alternative hypothesis in residual plot
+2. AR, heterogeneity of variance, endogeneity, skewness, exp distribution, poisson distribution and missing covariance
+3. Choice of plot design (loess or $y=0$ line), number of lineup (5~15), number of observations
+4. Use flow chart to illustrate the choices
+5. Literature review - check previous designs
+6. Build bookdown to track records

### 2021-06-20 site/02-literature.Rmd
update Infovis review

@@
 
-## Graphical Inference for Infovis - Hadley Wickham, Dianne Cook, Heike Hofmann and Andreas Buja (2010)
+## Graphical Inference for Infovis
 
-### Section 1 - Introduction
-
-In this section, the paper argues that much research in Infovis and most statistical methods are on two extremes.
-One focuses on finding relationships.
-Another focuses on checking whether a relationship really exists.
-
-Some previous attempts of graphical inference are mentioned in this section. The 2009 paper which develops the statistical concepts of graphical inference is also mentioned.
-
-It then summarises the structure of the paper.
-
-### Section 2 - What is inference and why do we need it?
-
-This section starts by stating the two components of statistical inference, which are testing and estimation. It then defines the meaning of inference in graphics.
-
-It uses the criminal justice system as an example to explain the principles of hypothesis testing, and points out the difference between traditional testing and visual testing, which are the test statistic and the mechanism of computing similarity.
-
-To be able to produce a visual test, null datasets which are samples from the null distribution are required. A null plot is a plot of a null dataset.
-
-It ends the section by making an important argument that the introduction of the visual test is not meant to replace traditional tests.
-“Traditional statistical tests are well studied, well-formulated and work best when data is well-behaved, following a known distribution in relatively simple scenarios.” However, traditional statistical tests do not cover all of the aspects we want to test.
-On the other hand, visual tests can be used in complex data analysis setting when traditional tests are unavailable.
-
-### Section 3 - Protocols of graphical inference
-
-This section introduces two protocols for graphical inference: the “Rorshach” and the “line-up”.
-
-The Rorschach protocol is used to calibrate our vision. In most of the imte, it contains only null plots.
-
-The line-up protocol consists of a plot of the true data and n-1 null plots. N is typically set to be 19.
-
-### Section 4 - Examples
-
-To use the line-up protocol, we need to:
-- Identify the question the plot is trying to answer
-- Characterize the null-hypothesis
-- Figure out how to generate null datasets
-
-It provides a list of questions that we would like to answer in commonly seen data plots.
-
-The null data can be generated in two ways in most cases: resampling and simulation.
-
-One of the benefits of using permutation is it preserves the marginal distribution of each variable.
-
-### Section 5 - Power
-
-The power of the visual test depends on many factors.
-An appropriate choice of plot is important to increase the power.
-For large datasets, aggregation is one of the considerations.
-It can significantly increase the power of the test.
-
-### Section 6 - Use
-
-One can implement these two protocols using the R package `nullabor`.
-
-### Section 7 - Conclusion
-
-“Graphical inference is important because it helps us to avoid false convictions.”
-
+BibTex: 
 ```
-@article{Wickham2010,
-abstract = {How do we know if what we see is really there?
-When visualizing data, how do we avoid falling into the trap of apophenia where we see patterns in random noise?
-Traditionally, infovis has been concerned with discovering new relationships, and statistics with preventing spurious relationships from being reported.
-We pull these opposing poles closer with two new techniques for rigorous statistical inference of visual discoveries.
-The "Rorschach" helps the analyst calibrate their understanding of uncertainty and "line-up" provides a protocol for assessing the significance of visual discoveries, protecting against the discovery of spurious structure.
-{\textcopyright} 2006 IEEE.},
-author = {Wickham, Hadley and Cook, Dianne and Hofmann, Heike and Buja, Andreas},
-doi = {10.1109/TVCG.2010.161},
-issn = {10772626},
-journal = {IEEE Transactions on Visualization and Computer Graphics},
-keywords = {Statistics,data plot,null hypotheses,permutation tests,visual testing},
-number = {6},
-pages = {973--979},
-pmid = {20975134},
-title = {{Graphical inference for infovis}},
-volume = {16},
-year = {2010}
+@article{wickham2010graphical,
+  title={Graphical inference for infovis},
+  author={Wickham, Hadley and Cook, Dianne and Hofmann, Heike and Buja, Andreas},
+  journal={IEEE Transactions on Visualization and Computer Graphics},
+  volume={16},
+  number={6},
+  pages={973--979},
+  year={2010},
+  publisher={IEEE}
 }
@@
 
-## Graphical Perception: Theory, Experimentation, and Application to the Development of Graphical Methods - William S. Cleveland and Rovert McGill (1984)
+### Keywords
 
-```
-@article{Cleveland1984,
-abstract = {The subject of graphical methods for data analysis and for data presentation needs a scientific foundation.
-In this article we take a few steps in the direction of establishing such a foundation.
-Our approach is based on graphical perception-the visual decoding of information encoded on graphs-and it includes both theory and experimentation to test the theory.
-The theory deals with a small but important piece of the whole process of graphical perception.
-The first part is an identification of a set of elementary perceptual tasks that are carried out when people extract quantitative information from graphs.
-The second part is an ordering of the tasks on the basis of how accurately people perform them.
-Elements of the theory are tested by experimentation in which subjects record their judgments of the quantitative information on graphs.
-The experiments validate these elements but also suggest that the set of elementary tasks should be expanded.
-The theory provides a guideline for graph construction: Graphs should employ elementary tasks as high in the ordering as possible.
-This principle is applied to a variety of graphs, including bar charts, divided bar charts, pie charts, and statistical maps with shading.
-The conclusion is that radical surgery on these popular graphs is needed, and as replacements we offer alternative graphical forms-dot charts, dot charts with grouping, and framed-rectangle charts.},
-author = {Cleveland, William S. and McGill, Robert},
-doi = {10.2307/2288400},
-issn = {01621459},
-journal = {Journal of the American Statistical Association},
-month = {sep},
-number = {387},
-pages = {531--554},
-publisher = {JSTOR},
-title = {{Graphical Perception: Theory, Experimentation, and Application to the Development of Graphical Methods}},
-volume = {79},
-year = {1984}
-}
-```
+Statistics, visual testing, permutation tests, null hypotheses, data plots.
 
-## Others
+### Introduction
 
-```
-@book{wickham2016ggplot2,
-author = {Wickham, Hadley},
-file = {:Users/patrickli/Desktop/papers/(Use R!) Hadley Wickham - ggplot2_ Elegant Graphics for Data Analysis-Springer (2016).pdf:pdf},
-publisher = {springer},
-title = {ggplot2: elegant graphics for data analysis},
-year = {2016}
-}
+Infovis focuses on uncovering new relationships by **tools of curiosity**, but most statistical methods focuses on examining relationships by **tools of skepticism**.
+Neither extreme is good.
+Hence, graphical inference try to fill the gap between them.
+It claims that this kind of inference can provide a tool for skepticism that can be applied in a curiosity-driven context.
 
-@book{Wilkinson2005,
-abstract = {The grammar of graphics (GoG) denotes a system with seven classes embedded in a data flow.
-This data flow specifies a strict order in which data are transformed from a raw dataset to a statistical graphic.
-Each class contains multiple methods, each of which is a function executed at the step in the data flow corresponding to that class.
-The classes are orthogonal, in the sense that the product set of all classes (every possible sequence of class methods) defines a space of graphics which is meaningful at every point.
-The meaning of a statistical graphic is thus determined by the mapping produced by the function chain linking data and graphic.},
-address = {New York},
-author = {Wilkinson, Leland},
-booktitle = {The Grammar of Graphics},
-doi = {10.1007/0-387-28695-0},
-edition = {2},
-file = {:Users/patrickli/Desktop/papers/(Statistics and Computing) Leland Wilkinson, D. Wills, D. Rope, A. Norton, R. Dubbs - The Grammar of Graphics-Springer (2005).pdf:pdf},
-publisher = {Springer-Verlag},
-title = {{The Grammar of Graphics}},
-year = {2005}
-}
-@article{Zhao2013MindRU,
-abstract = {Graphical statistics plays a very important job in research and industry.
-As a statistician, it is very useful that if one can see how people make their decision based on the plots, so that plots can be improved for better performance.
-With the help of eye-tracking equipment, researchers could show people several plots and ask a question on each, then track the person's eyes to see how they going through the plots to come to their answer.
-In this paper, the process and results of an experiment on watching what people were looking at in statistical plots will be discussed.
-This experiment is part of a larger experiment studying decision making and signal strength in statistical graphics, that uses Amazon's Mechanical Turk.},
-author = {Zhao, Y and Cook, D and Hofmann, H and Majumder, M and Chowdhury, N R},
-file = {:Users/patrickli/Desktop/papers/Mind Reading Using an Eyetracker to See How People Are Looking at Lineups.pdf:pdf},
-journal = {International Journal of Intelligent Technologies and Applied Statistics},
-number = {4},
-pages = {393--413},
-title = {{Mind Reading: Using an Eye-Tracker to See How People are Looking at Lineups}},
-volume = {6},
-year = {2013}
-}
-@article{Hofmann2012,
-abstract = {Lineups [4, 28] have been established as tools for visual testing similar to standard statistical inference tests, allowing us to evaluate the validity of graphical findings in an objective manner.
-In simulation studies [12] lineups have been shown as being efficient: the power of visual tests is comparable to classical tests while being much less stringent in terms of distributional assumptions made.
-This makes lineups versatile, yet powerful, tools in situations where conditions for regular statistical tests are not or cannot be met.
-In this paper we introduce lineups as a tool for evaluating the power of competing graphical designs.
-We highlight some of the theoretical properties and then show results from two studies evaluating competing designs: both studies are designed to go to the limits of our perceptual abilities to highlight differences between designs.
-We use both accuracy and speed of evaluation as measures of a successful design.
-The first study compares the choice of coordinate system: polar versus cartesian coordinates.
-The results show strong support in favor of cartesian coordinates in finding fast and accurate answers to spotting patterns.
-The second study is aimed at finding shift differences between distributions.
-Both studies are motivated by data problems that we have recently encountered, and explore using simulated data to evaluate the plot designs under controlled conditions.
-Amazon Mechanical Turk (MTurk) is used to conduct the studies.
-The lineups provide an effective mechanism for objectively evaluating plot designs.
-{\textcopyright} 1995-2012 IEEE.},
-author = {Hofmann, Heike and Follett, Lendie and Majumder, Mahbubul and Cook, Dianne},
-doi = {10.1109/TVCG.2012.230},
-file = {:Users/patrickli/Desktop/papers/Graphical Tests for Power Comparison of Competing Designs.pdf:pdf},
-issn = {10772626},
-journal = {IEEE Transactions on Visualization and Computer Graphics},
-keywords = {Efficiency of displays,Lineups,Power comparison,Visual inference},
-number = {12},
-pages = {2441--2448},
-title = {{Graphical tests for power comparison of competing designs}},
-volume = {18},
-year = {2012}
-}
+### What is inference and why do we need it?
 
-@article{RoyChowdhury2015,
-abstract = {Statistical graphics play an important role in exploratory data analysis, model checking and diagnosis.
-With high dimensional data, this often means plotting low-dimensional projections, for example, in classification tasks projection pursuit is used to find low-dimensional projections that reveal differences between labelled groups.
-In many contemporary data sets the number of observations is relatively small compared to the number of variables, which is known as a high dimension low sample size (HDLSS) problem.
-This paper explores the use of visual inference on understanding low-dimensional pictures of HDLSS data.
-Visual inference helps to quantify the significance of findings made from graphics.
-This approach may be helpful to broaden the understanding of issues related to HDLSS data in the data analysis community.
-Methods are illustrated using data from a published paper, which erroneously found real separation in microarray data, and with a simulation study conducted using Amazon's Mechanical Turk.},
-author = {{Roy Chowdhury}, Niladri and Cook, Dianne and Hofmann, Heike and Majumder, Mahbubul and Lee, Eun Kyung and Toth, Amy L.},
-doi = {10.1007/s00180-014-0534-x},
-file = {:Users/patrickli/Desktop/papers/RoyChowdhury2015_Article_UsingVisualStatisticalInferenc.pdf:pdf},
-issn = {16139658},
-journal = {Computational Statistics},
-keywords = {Data mining,Lineup,Projection pursuit,Statistical graphics,Visualization},
-month = {nov},
-number = {2},
-pages = {293--316},
-publisher = {Springer Verlag},
-title = {{Using visual statistical inference to better understand random class separations in high dimension, low sample size data}},
-url = {https://link.springer.com/article/10.1007/s00180-014-0534-x},
-volume = {30},
-year = {2015}
-}
-@article{Loy2017,
-abstract = {The complexity of linear mixed-effects (LME) models means that traditional diagnostics are rendered less effective.
-This is due to a breakdown of asymptotic results, boundary issues, and visible patterns in residual plots that are introduced by the model fitting process.
-Some of these issues are well known and adjustments have been proposed.
-Working with LME models typically requires that the analyst keeps track of all the special circumstances that may arise.
-In this article, we illustrate a simpler but generally applicable approach to diagnosing LME models.
-We explain how to use new visual inference methods for these purposes.
-The approach provides a unified framework for diagnosing LME fits and for model selection.
-We illustrate the use of this approach on several commonly available datasets.
-A large-scale Amazon Turk study was used to validate the methods.
-R code is provided for the analyses.
-Supplementary materials for this article are available online.},
-archivePrefix = {arXiv},
-arxivId = {1502.06988},
-author = {Loy, Adam and Hofmann, Heike and Cook, Dianne},
-doi = {10.1080/10618600.2017.1330207},
-eprint = {1502.06988},
-file = {:Users/patrickli/Desktop/papers/Model Choice and Diagnostics for Linear Mixed Effects Models Using Statistics on Street Corners.pdf:pdf},
-issn = {15372715},
-journal = {Journal of Computational and Graphical Statistics},
-keywords = {Lineup protocol,Model diagnostics,Model selection,Statistical graphics,Visual inference},
-month = {jul},
-number = {3},
-pages = {478--492},
-publisher = {American Statistical Association},
-title = {{Model Choice and Diagnostics for Linear Mixed-Effects Models Using Statistics on Street Corners}},
-url = {https://www.tandfonline.com/doi/abs/10.1080/10618600.2017.1330207},
-volume = {26},
-year = {2017}
-}
-@article{Majumder2013,
-abstract = {Statistical graphics play a crucial role in exploratory data analysis, model checking, and diagnosis.
-The lineup protocol enables statistical significance testing of visual findings, bridging the gulf between exploratory and inferential statistics.
-In this article, inferential methods for statistical graphics are developed further by refining the terminology of visual inference and framing the lineup protocol in a context that allows direct comparison with conventional tests in scenarios when a conventional test exists.
-This framework is used to compare the performance of the lineup protocol against conventional statistical testing in the scenario of fitting linear models.Ahuman subjects experiment is conducted using simulated data to provide controlled conditions.
-Results suggest that the lineup protocol performs comparably with the conventional tests, and expectedly outperforms them when data are contaminated, a scenario where assumptions required for performing a conventional test are violated.
-Surprisingly, visual tests have higher power than the conventional tests when the effect size is large.
-And, interestingly, there may be some super-visual individuals who yield better performance and power than the conventional test even in the most difficult tasks.
-Supplementary materials for this article are available online.
-{\textcopyright} 2013 American Statistical Association.},
-author = {Majumder, Mahbubul and Hofmann, Heike and Cook, Dianne},
-doi = {10.1080/01621459.2013.808157},
-file = {:Users/patrickli/Desktop/papers/Validation of Visual Statistical Inference Applied to Linear Models.pdf:pdf},
-issn = {01621459},
-journal = {Journal of the American Statistical Association},
-keywords = {Data mining,Effect size,Exploratory data analysis,Lineup,Nonparametric test,Practical significance,Statistical graphics,Visualization},
-number = {503},
-pages = {942--956},
-publisher = {Taylor & Francis Group},
-title = {{Validation of visual statistical inference, applied to linear models}},
-url = {https://www.tandfonline.com/doi/abs/10.1080/01621459.2013.808157},
-volume = {108},
-year = {2013}
-}
-@article{Chowdhury2018,
-abstract = {Graphics play a crucial role in statistical analysis and data mining.
-Being able to quantify structure in data that is visible in plots, and how people read the structure from plots is an ongoing challenge.
-The lineup protocol provides a formal framework for data plots, making inference possible.
-The data plot is treated like a test statistic, and lineup protocol acts like a comparison with the sampling distribution of the nulls.
-This article describes metrics for describing structure in data plots and evaluates them in relation to the choices that human readers made during several large Amazon Turk studies using lineups.
-The metrics that were more specific to the plot types tended to better match subject choices, than generic metrics.
-The process that we followed to evaluate metrics will be useful for general development of numerically measuring structure in plots, and also in future experiments on lineups for choosing blocks of pictures.
-Supplementary materials for this article are available online.},
-author = {Chowdhury, Niladri Roy and Cook, Dianne and Hofmann, Heike and Majumder, Mahbubul},
-doi = {10.1080/10618600.2017.1356323},
-file = {:Users/patrickli/Desktop/papers/Measuring Lineup Difficulty By Matching Distance Metrics With Subject Choices in Crowd Sourced Data.pdf:pdf},
-issn = {15372715},
-journal = {Journal of Computational and Graphical Statistics},
-keywords = {Cognitive perception,Data mining,Data science,Data visualization,Distance metrics,Exploratory data analysis,Information visualization,Statistical graphics,Visual inference},
-month = {jan},
-number = {1},
-pages = {132--145},
-publisher = {American Statistical Association},
-title = {{Measuring Lineup Difficulty By Matching Distance Metrics With Subject Choices in Crowd-Sourced Data}},
-url = {https://www.tandfonline.com/doi/abs/10.1080/10618600.2017.1356323},
-volume = {27},
-year = {2018}
-}
-@article{Buja2009,
-abstract = {We propose to furnish visual statistical methods with an inferential framework and protocol, modelled on confirmatory statistical testing.
-In this framework, plots take on the role of test statistics, and human cognition the role of statistical tests.
-Statistical significance of 'discoveries' is measured by having the human viewer compare the plot of the real dataset with collections of plots of simulated datasets.
-A simple but rigorous protocol that provides inferential validity is modelled after the 'lineup' popular from criminal legal procedures.
-Another protocol modelled after the 'Rorschach' inkblot test, well known from (pop-)psychology, will help analysts acclimatize to random variability before being exposed to the plot of the real data.
-The proposed protocols will be useful for exploratory data analysis, with reference datasets simulated by using a null assumption that structure is absent.
-The framework is also useful for model diagnostics in which case reference datasets are simulated from the model in question.
-This latter point follows up on previous proposals.
-Adopting the protocols will mean an adjustment in working procedures for data analysts, adding more rigour, and teachers might find that incorporating these protocols into the curriculum improves their students' statistical thinking.
-This journal is {\textcopyright} 2009 The Royal Society.},
-author = {Buja, Andreas and Cook, Dianne and Hofmann, Heike and Lawrence, Michael and Lee, Eun Kyung and Swayne, Deborah F. and Wickham, Hadley},
-doi = {10.1098/rsta.2009.0120},
-file = {:Users/patrickli/Desktop/papers/Statistical inference for exploratory data analysis and model diagnostics.pdf:pdf},
-issn = {1364503X},
-journal = {Philosophical Transactions of the Royal Society A: Mathematical, Physical and Engineering Sciences},
-keywords = {Cognitive perception,Permutation tests,Rotation tests,Simulation,Statistical graphics,Visual data mining},
-month = {nov},
-number = {1906},
-pages = {4361--4383},
-pmid = {19805449},
-publisher = {Royal Society},
-title = {{Statistical inference for exploratory data analysis and model diagnostics}},
-url = {https://royalsocietypublishing.org/doi/10.1098/rsta.2009.0120},
-volume = {367},
-year = {2009}
-}
-```
+Inference is about drawing conclusions about the population from the sample.
+There are two components of statistical inference, testing and estimation.
+For graphical inference, the focus is to test whether what we see in a plot of the sample is an accurate reflection of the entire population or not.
+The test statistic in visual inference is a plot of the data.
+A **null dataset** is a sample from the null distribution, and a **null plot** is a plot of a null dataset.
+The benefit of visual inference is that it can be used in complex data analysis settings that do not have corresponding numerical tests.
+
+### Protocols of graphical inference
+
+#### Rorschach
+
+Rorschach protocol is used to calibrate our vision to the natural variability in plots in which the data is generated from scenarios consistent with the null hypothesis.
+
+#### Line-up
+
+The line up is consisted of $n-1$ decoys and $1$ plot of the true data.
+If we set $n = 19$, then under the null hypothesis, there is only $5$% chance to pick the plot of the true data.
+If we recruit $K$ observers, then under the null hypothesis, the p-value is $P(B(K, 0.05) \geq k)$.
+
+[comment]: <> (What is the probability of the true plot be the most different one?)
+
+[comment]: <> (What is the probability of the first plot be the most different one? == What is the probability of the second plot be the most different one?)
+
+### Examples
+
+To use the line-up protocol, we need to:
+1. Identify the question the plot is trying to answer
+2. Characterize the null-hypothesis
+3. Figure out how to generate null datasets
+
+There are two techniques that can be applied in many caeses:
+
+1. Resampling. Permutation and bootstrapping.
+2. Simulated data from a assumed model.
+
+#### Tag clouds
+
+A tag cloud can be used to visualize frequency of words in a document.
+Words are arranged in various ways, often alphabetically, with size proportional to their frequency.
+The null hypothesis for a comparison tag cloud is that the two documents are equivalent, the frequency of words is the same in each document.
+Null data can be generated by randomly permute the one of the column.
+
+#### Scatterplot
+
+A scatterplot displays the relationship between $x$ and $y$. A strong null hypothesis is that there is no relationship between $x$ and $y$ variables.
+
+### Power
+
+The power of a statistical test is the probability of correctly convicting a guilty data set. The capacity to detect specific structure in plots can depend on the perceptual properties.
+
+### Use
+
+An R package: nullabor

### 2021-06-20 site/_book/site.tex
update Infovis review

--- 2021-04-05 site/_book/site.tex
@@
 \author{Patrick Li}
-\date{2021-04-06}
+\date{2021-06-20}
 
@@
 
-Here is a review of existing methods.
+\hypertarget{graphical-inference-for-infovis}{%
+\section{Graphical Inference for Infovis}\label{graphical-inference-for-infovis}}
+
+BibTex:
+
+\begin{verbatim}
+@article{wickham2010graphical,
+  title={Graphical inference for infovis},
+  author={Wickham, Hadley and Cook, Dianne and Hofmann, Heike and Buja, Andreas},
+  journal={IEEE Transactions on Visualization and Computer Graphics},
+  volume={16},
+  number={6},
+  pages={973--979},
+  year={2010},
+  publisher={IEEE}
+}
+\end{verbatim}
+
+\hypertarget{keywords}{%
+\subsection{Keywords}\label{keywords}}
+
+Statistics, visual testing, permutation tests, null hypotheses, data plots.
+
+\hypertarget{introduction-1}{%
+\subsection{Introduction}\label{introduction-1}}
+
+Infovis focuses on uncovering new relationships by \textbf{tools of curiosity}, but most statistical methods focuses on examining relationships by \textbf{tools of skepticism}.
+Neither extreme is good.
+Hence, graphical inference try to fill the gap between them.
+It claims that this kind of inference can provide a tool for skepticism that can be applied in a curiosity-driven context.
+
+\hypertarget{what-is-inference-and-why-do-we-need-it}{%
+\subsection{What is inference and why do we need it?}\label{what-is-inference-and-why-do-we-need-it}}
+
+Inference is about drawing conclusions about the population from the sample.
+There are two components of statistical inference, testing and estimation.
+For graphical inference, the focus is to test whether what we see in a plot of the sample is an accurate reflection of the entire population or not.
+The test statistic in visual inference is a plot of the data.
+A \textbf{null dataset} is a sample from the null distribution, and a \textbf{null plot} is a plot of a null dataset.
+The benefit of visual inference is that it can be used in complex data analysis settings that do not have corresponding numerical tests.
+
+\hypertarget{protocols-of-graphical-inference}{%
+\subsection{Protocols of graphical inference}\label{protocols-of-graphical-inference}}
+
+\hypertarget{rorschach}{%
+\subsubsection{Rorschach}\label{rorschach}}
+
+Rorschach protocol is used to calibrate our vision to the natural variability in plots in which the data is generated from scenarios consistent with the null hypothesis.
+
+\hypertarget{line-up}{%
+\subsubsection{Line-up}\label{line-up}}
+
+The line up is consisted of \(n-1\) decoys and \(1\) plot of the true data.
+If we set \(n = 19\), then under the null hypothesis, there is only \(5\)\% chance to pick the plot of the true data.
+If we recruit \(K\) observers, then under the null hypothesis, the p-value is \(P(B(K, 0.05) \geq k)\).
+
+\hypertarget{examples}{%
+\subsection{Examples}\label{examples}}
+
+To use the line-up protocol, we need to:
+1. Identify the question the plot is trying to answer
+2. Characterize the null-hypothesis
+3. Figure out how to generate null datasets
+
+There are two techniques that can be applied in many caeses:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Resampling. Permutation and bootstrapping.
+\item
+  Simulated data from a assumed model.
+\end{enumerate}
+
+\hypertarget{tag-clouds}{%
+\subsubsection{Tag clouds}\label{tag-clouds}}
+
+A tag cloud can be used to visualize frequency of words in a document.
+Words are arranged in various ways, often alphabetically, with size proportional to their frequency.
+The null hypothesis for a comparison tag cloud is that the two documents are equivalent, the frequency of words is the same in each document.
+Null data can be generated by randomly permute the one of the column.
+
+\hypertarget{scatterplot}{%
+\subsubsection{Scatterplot}\label{scatterplot}}
+
+A scatterplot displays the relationship between \(x\) and \(y\). A strong null hypothesis is that there is no relationship between \(x\) and \(y\) variables.
+
+\hypertarget{power}{%
+\subsection{Power}\label{power}}
+
+The power of a statistical test is the probability of correctly convicting a guilty data set. The capacity to detect specific structure in plots can depend on the perceptual properties.
+
+\hypertarget{use}{%
+\subsection{Use}\label{use}}
+
+An R package: nullabor
 

### 2021-12-06 Thesis/_book/thesis.tex
add thesis template

% This is a LaTeX thesis template for Monash University.
% to be used with Rmarkdown
% This template was produced by Rob Hyndman
% Version: 6 September 2016

\documentclass{monashthesis}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Add any LaTeX packages and other preamble here if required
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author{Joe Bloggs}
\title{Thesis title}
\degrees{B.Sc. (Hons), University of Tangambalanga}
\def\degreetitle{Doctor of Philosophy}
% Add subject and keywords below
\hypersetup{
     %pdfsubject={The Subject},
     %pdfkeywords={Some Keywords},
     pdfauthor={Joe Bloggs},
     pdftitle={Thesis title},
     pdfproducer={Bookdown with LaTeX}
}

\bibliography{thesisrefs}

\begin{document}

\pagenumbering{roman}

\titlepage

{\setstretch{1.2}\sf\tighttoc\doublespacing}

\hypertarget{copyright-notice}{%
\chapter*{Copyright notice}\label{copyright-notice}}
\addcontentsline{toc}{chapter}{Copyright notice}

\emph{(Choose one of the following notices.)}

\emph{(Notice 1)}

\textcopyright { } \authorname~(\number\the\year).

\emph{The second notice certifies the appropriate use of any third-party material in the thesis. Students choosing to deposit their thesis into the restricted access section of the repository are not required to complete Notice 2.}

\emph{(Notice 2)}

\textcopyright { } \authorname~(\number\the\year).

I certify that I have made all reasonable efforts to secure copyright permissions for third-party content included in this thesis and have not knowingly added copyright content to my work without the owner's permission.

\newpage

\hypertarget{abstract}{%
\chapter*{Abstract}\label{abstract}}
\addcontentsline{toc}{chapter}{Abstract}

The abstract should outline the main approach and findings of the thesis and must not be more than 500 words.

\newpage

\hypertarget{declaration}{%
\chapter*{Declaration}\label{declaration}}
\addcontentsline{toc}{chapter}{Declaration}

\emph{(Standard thesis)}

This thesis is an original work of my research and contains no material which has been accepted for the award of any other degree or diploma at any university or equivalent institution and that, to the best of my knowledge and belief, this thesis contains no material previously published or written by another person, except where due reference is made in the text of the thesis.

\emph{(Thesis including published works declaration)}

I hereby declare that this thesis contains no material which has been accepted for the award of any other degree or diploma at any university or equivalent institution and that, to the best of my knowledge and belief, this thesis contains no material previously published or written by another person, except where due reference is made in the text of the thesis.

This thesis includes (insert number) original papers published in peer reviewed journals and (insert number) submitted publications. The core theme of the thesis is (insert theme). The ideas, development and writing up of all the papers in the thesis were the principal responsibility of myself, the student, working within the (insert name of academic unit) under the supervision of (insert name of supervisor).

(The inclusion of co-authors reflects the fact that the work came from active collaboration between researchers and acknowledges input into team-based research.) Remove this paragraph for theses with sole-authored work

In the case of (insert chapter numbers) my contribution to the work involved the following:

I have / have not renumbered sections of submitted or published papers in order to generate a consistent presentation within the thesis.

\begin{table}
\centering\footnotesize\tabcolsep=0.12cm
\begin{tabular}{|p{1.2cm}|p{2cm}|p{1.8cm}|p{3.4cm}|p{3.5cm}|p{1.5cm}|}
\hline
\RaggedRight\textbf{Thesis chapter}  & 
\RaggedRight\textbf{Publication title}  & 
\RaggedRight\textbf{Status (published, in press, accepted or returned for revision)}  &  
\RaggedRight\textbf{Nature and \% of student contribution} & 
\RaggedRight\textbf{Co-author name(s), nature and \% of co-author’s contribution} &  \RaggedRight\textbf{Co-author(s), Monash student Y/N} \\ 
\hline
2 & xx & xx  & xx & xx & \multicolumn{1}{c|}{N}   \\
\hline
3 & xx & xx  & xx & xx  & \multicolumn{1}{c|}{N}   \\
\hline
4 & xx & xx  & xx & xx & \multicolumn{1}{c|}{N}    \\
\hline
5 & xx & xx  & xx & xx & \multicolumn{1}{c|}{N}   \\
\hline
\end{tabular}
\end{table}

\textbf{Student name:} \authorname

\textbf{Student signature:}

\textbf{Date:}

\hypertarget{acknowledgements}{%
\chapter*{Acknowledgements}\label{acknowledgements}}
\addcontentsline{toc}{chapter}{Acknowledgements}

I would like to thank my pet goldfish for \dots

\hypertarget{preface}{%
\chapter*{Preface}\label{preface}}
\addcontentsline{toc}{chapter}{Preface}

The material in Chapter \ref{ch:intro} has been submitted to the journal \emph{Journal of Impossible Results} for possible publication.

The contribution in Chapter \ref{ch:litreview} of this thesis was presented in the International Symposium on Nonsense held in Dublin, Ireland, in July 2015.

\clearpage\pagenumbering{arabic}\setcounter{page}{0}

\hypertarget{ch:intro}{%
\chapter{Introduction}\label{ch:intro}}

This is where you introduce the main ideas of your thesis, and an overview of the context and background.

In a PhD, Chapter 2 would normally contain a literature review. Typically, Chapters 3--5 would contain your own contributions. Think of each of these as potential papers to be submitted to journals. Finally, Chapter 6 provides some concluding remarks, discussion, ideas for future research, and so on. Appendixes can contain additional material that don't fit into any chapters, but that you want to put on record. For example, additional tables, output, etc.

\hypertarget{rmarkdown}{%
\section{Rmarkdown}\label{rmarkdown}}

In this template, the rest of the chapter shows how to use Rmarkdown. The big advantage of using Rmarkdown is that it allows you to include your R code directly into your thesis, to ensure there are no errors in copying and pasting, and that everything is reproducible. It also helps you stay better organized.

For details on using \emph{R Markdown} see \url{http://rmarkdown.rstudio.com}.

\hypertarget{data}{%
\section{Data}\label{data}}

Included in this template is a file called \texttt{sales.csv}. This contains quarterly data on Sales and Advertising budget for a small company over the period 1981--2005. It also contains the GDP (gross domestic product) over the same period. All series have been adjusted for inflation. We can load in this data set using the following command:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{sales }\OtherTok{\textless{}{-}} \FunctionTok{ts}\NormalTok{(}\FunctionTok{read.csv}\NormalTok{(}\StringTok{"data/sales.csv"}\NormalTok{)[,}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{], }\AttributeTok{start=}\DecValTok{1981}\NormalTok{, }\AttributeTok{frequency=}\DecValTok{4}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Any data you use in your thesis can go into the data directory. The data should be in exactly the format you obtained it. Do no editing or manipulation of the data outside of R. Any data munging should be scripted in R and form part of your thesis files (possibly hidden in the output).

\hypertarget{figures}{%
\section{Figures}\label{figures}}

Figure \ref{fig:deaths} shows time plots of the data we just loaded. Notice how figure captions and references work. Chunk names can be used as figure labels with \texttt{fig:} prefixed. Never manually type figure numbers, as they can change when you add or delete figures. This way, the figure numbering is always correct.

\begin{figure}
\centering
\includegraphics{thesis_files/figure-latex/deaths-1.pdf}
\caption{\label{fig:deaths}Quarterly sales, advertising and GDP data.}
\end{figure}

\hypertarget{results-from-analyses}{%
\section{Results from analyses}\label{results-from-analyses}}

We can fit a dynamic regression model to the sales data.

If \(y_t\) denotes the sales in quarter \(t\), \(x_t\) denotes the corresponding advertising budget and \(z_t\) denotes the GDP, then the resulting model is:
\begin{equation}
  y_t - y_{t-4} = \beta (x_t-x_{t-4}) + \gamma (z_t-z_{t-4}) + \theta_1 \varepsilon_{t-1} + \Theta_1 \varepsilon_{t-4} + \varepsilon_t
\end{equation}
where
\(\beta = 2.28\),
\(\gamma = 0.97\),
\(\theta_1 = NA\),
and
\(\Theta_1 = -0.90\).

\hypertarget{tables}{%
\section{Tables}\label{tables}}

Let's assume future advertising spend and GDP are at the current levels. Then forecasts for the next year are given in Table \ref{tab:salesforecasts}.

\begin{verbatim}
## Warning in sprintf(str, x, ...): one argument not used by format '\toprule
## Point Forecast & Lo 80 & Hi 80 & Lo 95 & Hi 95 \\
## \midrule'

## Warning in sprintf(str, x, ...): one argument not used by format '\toprule
## Point Forecast & Lo 80 & Hi 80 & Lo 95 & Hi 95 \\
## \midrule'
\end{verbatim}

\begin{table}[ht]
\begin{center}
\begin{tabular}{lllll}
\toprule
Point Forecast & Lo 80 & Hi 80 & Lo 95 & Hi 95 \\
\midrule
1000.2 &  947.7 & 1052.7 & 919.9 & 1080.5 \\
1013.1 &  959.3 & 1066.8 & 930.9 & 1095.3 \\
1076.7 & 1022.9 & 1130.6 & 994.4 & 1159.0 \\
1003.5 &  949.7 & 1057.4 & 921.2 & 1085.8 \\
\bottomrule
\end{tabular}
\caption{Forecasts for the next year assuming Advertising budget and GDP are unchanged.}
\label{tab:salesforecasts}
\end{center}
\end{table}

Again, notice the use of labels and references to automatically generate Table numbers. In this case, we need to generate the label ourselves.

The \texttt{knitLatex} package is useful for generating tables from R output. Other packages can do similar things including the \texttt{kable} function in \texttt{knitr} which is somewhat simpler but you have less control over the result. If you use \texttt{knitLatex} to generate tables, don't forget to include \texttt{results="asis"} in the chunk settings.

\hypertarget{ch:litreview}{%
\chapter{Literature Review}\label{ch:litreview}}

This chapter contains a summary of the context in which your research is set.

Imagine you are writing for your fellow PhD students. Topics that are well-known to them do not have to be included here. But things that they may not know about should be included.

Resist the temptation to discuss everything you've read in the last few years. And you are not writing a textbook either. This chapter is meant to provide the background necessary to understand the material in subsequent chapters. Stick to that.

You will need to organize the literature review around themes, and within each theme provide a story explaining the development of ideas to date. In each theme, you should get to the point where your ideas will fit in. But leave your ideas to later chapters. This way it is clear what has been done beforehand, and what new contributions you are making to the research field.

All citations should be done using markdown notation as shown below. This way, your bibliography will be compiled automatically and correctly.

\hypertarget{sec:expsmooth}{%
\section{Exponential smoothing}\label{sec:expsmooth}}

Exponential smoothing was originally developed in the late 1950s \autocite{Brown59,Brown63,Holt57,Winters60}. Because of their computational simplicity and interpretability, they became widely used in practice.

Empirical studies by \textcite{MH79} and \textcite{Metal82} found little difference in forecast accuracy between exponential smoothing and ARIMA models. This made the family of exponential smoothing procedures an attractive proposition \autocite[see][]{CKOS01}.

The methods were less popular in academic circles until \textcite{OKS97} introduced a state space formulation of some of the methods, which was extended in \textcite{HKSG02} to cover the full range of exponential smoothing methods.

\appendix

\hypertarget{additional-stuff}{%
\chapter{Additional stuff}\label{additional-stuff}}

You might put some computer output here, or maybe additional tables.

Note that line 5 must appear before your first appendix. But other appendices can just start like any other chapter.

\printbibliography[heading=bibintoc]

\end{document}

### 2021-12-06 experiment/index.Rmd
add exp book

--- 
title: "Visual Inference for Diagnostic of Nonlinearity and Heteroskedasticy in Linear Regression"
author: "Patrick Li"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
# url: your book url like https://bookdown.org/yihui/bookdown
# cover-image: path to the social sharing image like images/cover.jpg
description: |
  .
biblio-style: apalike
csl: chicago-fullnote-bibliography.csl
---

# About

This book is about

## Structure of the book

In this book, will be covered ...

### 2021-12-06 experiment/scripts/experiment.R

@@ -1,0 +1,466 @@
+EXPERIMENT <- function(name) {
+
+  experiment <- new.env()
+  experiment$name <- name
+  experiment$num_lineup <- 0
+  experiment$model_list <- list()
+  experiment$dat <- list()
+
+  experiment$model <- function(name) {
+
+    # Method of getting a model.
+    #
+    # Available arguments:
+    #
+    # name: the name of the model
+
+    experiment$model_list[[name]]
+  }
+
+  experiment$summary_lineup <- function(lineup_id = 1, model_spec = FALSE) {
+
+    # Method of printing summary of a lineup
+    #
+    # Available arguments:
+    #
+    # lineup_id: id of lineup
+    # model_spec: whether or not to print the model spec
+
+    if (length(experiment$dat) < lineup_id) {
+      cli::cli_alert_warning("lineup {lineup_id} does not exist!")
+      return(invisible(NULL))
+    }
+    cli::cli_h1("Lineup {lineup_id}")
+    cli::cli_li("k = {experiment$dat[[lineup_id]]$metadata$k}")
+    cli::cli_li("n = {experiment$dat[[lineup_id]]$metadata$n}")
+    cli::cli_li("effect size = {experiment$dat[[lineup_id]]$metadata$effect_size}")
+    cli::cli_li("answer = {experiment$dat[[lineup_id]]$metadata$ans}")
+    if (model_spec) {
+      print(experiment$model(experiment$dat[[lineup_id]]$metadata$name))
+    }
+
+    return(invisible(NULL))
+  }
+
+  experiment$summary_model <- function(name, lineup = FALSE) {
+
+    # Method of printing summary of a model
+    #
+    # Available arguments:
+    #
+    # name: name of the model
+    # lineup: whether or not to print the lineups
+
+    print(experiment$model(name))
+    if (lineup) {
+      lineup_ids <- map(experiment$dat, ~.$metadata$name == name) %>%
+        unlist() %>%
+        which()
+      for (lineup_id in lineup_ids) {
+        experiment$summary_lineup(lineup_id)
+      }
+    }
+
+    return(invisible(NULL))
+  }
+
+  experiment$summary <- function(input = NULL) {
+
+
+    if (!is.null(input) & is.numeric(input)) {
+      experiment$summary_lineup(input, model_spec = TRUE)
+      return(invisible(NULL))
+    }
+
+    if (!is.null(name) & is.character(input)) {
+      experiment$summary_model(input, lineup = TRUE)
+      return(invisible(NULL))
+    }
+
[…]

### 2021-12-06 experiment/scripts/non_normal_model.R

@@ -1,0 +1,33 @@
+NON_NORMAL_MODEL <- function(c = 5,
+                             x_dist = "uniform", z_discrete = FALSE,
+                             e_dist = "t", ...) {
+
+  # Init a model
+  model <- MODEL_SPEC(type = "non-normal error",
+                      formula = "y ~ 1 + x + c * z + e, e ~ non-normal",
+                      c = c,
+                      x_dist = x_dist, z_discrete = z_discrete,
+                      e_dist = e_dist, ...)
+
+  # Method of generating y
+  model$gen_y <- function(n) {
+    x <- model$gen_x(n)
+    e <- model$gen_e(n)
+    y <- x[,1] + x[,2] * model$parameters$c + e + 1
+
+    return(data.frame(x = x[,1], z = x[,2], y = y))
+
+  }
+
+  # Method of evaluating the data using numerical test
+  model$evaluate <- function(dat) {
+
+    mod <- lm(y ~ x + z, data = dat)
+
+    this_sw_test <- shapiro.test(resid(mod))
+
+    return(list("swtest", unname(this_sw_test$statistic), unname(this_sw_test$p.value)))
+  }
+
+  return(model)
+}

### 2021-12-06 experiment/05-data-quality.Rmd
add exp book

# (PART\*) Analysis of the experiment data {-}

# Data quality

The quality of the experiment data mainly depends on the respondent engagement. We could measure this by several variables:

1. response time
2. selections
3. Detection rate

Identifications for potential low quality responses:

  a. fail to identify an easy lineup because of insufficient response time
  b. significantly low detection rate because of insufficient response time throughout the study
  c. the detection rate adjusted by the number of selections significantly drop
  d.

## Study 1

## Study 2

### 2021-12-06 experiment/scripts/evaluate_grid.R

@@ -1,0 +1,150 @@
+source(here::here("scripts/model.R"))
+source(here::here("scripts/higher_order_model.R"))
+source(here::here("scripts/heteroskedasticity_model.R"))
+source(here::here("scripts/non_normal_model.R"))
+source(here::here("scripts/experiment.R"))
+
+num_rec <- 100000
+
+build_and_record <- function(...) {
+  paras <- list(...)
+  myexp <- EXPERIMENT("tmp")
+  mod <- do.call(paras$model, paras)
+  myexp$add_model("tmp_mod", mod)
+  myexp$gen_lineup_data("tmp_mod", k = 2, n = paras$n)
+  return(list(c(statistic = myexp$get_statistic(), pvalue = myexp$get_pvalue())))
+}
+
+data.frame(a = runif(num_rec, -3, 3),
+           b = runif(num_rec, -3, 3),
+           c = runif(num_rec, 0, 5),
+           x_dist = sample(c("uniform", "normal", "lognormal", "neglognormal"), num_rec, replace = TRUE),
+           x_mu = 0,
+           x_sigma = sample(c(0.3, 0.6), num_rec, replace = TRUE),
+           z_discrete = sample(c(TRUE, FALSE), num_rec, replace = TRUE),
+           z_discrete_dist = "discrete_uniform",
+           z_n = sample(seq(3, 10, 1), num_rec, replace = TRUE),
+           e_dist = "normal",
+           e_df = NA,
+           e_sigma = sample(c(0.25, 0.5, 1, 2), num_rec, replace = TRUE),
+           n = sample(c(50, 100, 300), num_rec, replace = TRUE),
+           model = "HIGHER_ORDER_MODEL") %>%
+  tibble() -> mygrid
+
+mygrid %>%
+  mutate(i = 1:n()/n()) %>%
+  rowwise() %>%
+  mutate(result = build_and_record(i=i,
+                                   a=a,
+                                   b=b,
+                                   c=c,
+                                   x_dist=x_dist,
+                                   x_mu=x_mu,
+                                   x_sigma=x_sigma,
+                                   z_discrete=z_discrete,
+                                   z_discrete_dist=z_discrete_dist,
+                                   z_n=z_n,
+                                   e_dist=e_dist,
+                                   e_df=e_df,
+                                   e_sigma=e_sigma,
+                                   n=n,
+                                   model=model)) %>%
+  ungroup() %>%
+  cbind(., do.call('rbind', .$result)) %>%
+  tibble() %>%
[…]

### 2021-12-06 experiment/scripts/heteroskedasticity_model.R

@@ -1,0 +1,59 @@
+HETEROSKEDASTICITY_MODEL <- function(a, b,
+                                     x_dist = "uniform", z_discrete = FALSE,
+                                     ...) {
+
+  # Init a new model
+  model <- MODEL_SPEC(type = "heteroskedasticity",
+                      formula = "y ~ 1 + x + e, e ~ N(0, 1 + (2-|a|) * (x_i - a)^2 * b)",
+                      a = a, b = b,
+                      x_dist = x_dist, z_discrete = z_discrete,
+                      ...)
+
+  # Method of generating e
+  model$gen_e <- function(x, n) {
+
+    sigma2_i <- 1 +
+      (2 - abs(model$parameters$a)) * (x[,1] - model$parameters$a)^2 * model$parameters$b
+
+    sigma_i <- sqrt(sigma2_i)
+
+    result <- data.frame(single_sigma = sigma_i) %>%
+      rowwise() %>%
+      mutate(e = rnorm(1, sd = single_sigma))
+
+    return(result$e)
+  }
+
+  # Method of generating y
+  model$gen_y <- function(n) {
+    x <- model$gen_x(n)
+    e <- model$gen_e(x, n)
+
+    y <- 1 + x[,1] + e
+    return(data.frame(x = x[,1], z = x[,2], y = y))
+  }
+
+  model$fit <- function(dat) {
+    lm(y ~ x, data = dat)
+  }
+
+  # Method of evaluating the data using numerical test
+  model$evaluate <- function(dat) {
+
+    mod <- lm(y ~ x, data = dat)
+
+    tmp_data <- data.frame(x = dat$x)
+    tmp_data$xs <- tmp_data$x^2
+
+    this_bp_test <- lmtest::bptest(mod, varformula = ~ x + xs, data = tmp_data)
+
+    return(list("bptest", unname(this_bp_test$statistic), unname(this_bp_test$p.value)))
+  }
+
+  model$effect_size <- function(dat) {
+
+    return(model$parameters$b * sqrt(nrow(dat)))
+  }
+
+  return(model)
+}

### 2021-12-06 experiment/scripts/higher_order_model.R

@@ -1,0 +1,60 @@
+HIGHER_ORDER_MODEL <- function(a, b, c = 1,
+                               x_dist = "uniform", z_discrete = FALSE,
+                               e_dist = "normal", e_sigma = 1, ...) {
+
+  # Init a new model
+  model <- MODEL_SPEC(type = "higher order",
+                      formula = "y ~ 1 + (2-c) * x + c * z + a * [(2-c) * x]^2 + a * (c * z)^2 + b * [(2-c) * x]^3 + b * (c * z)^3 + e, e ~ N(0, sigma^2)",
+                      a = a, b = b, c = c,
+                      x_dist = x_dist, z_discrete = z_discrete,
+                      e_dist = e_dist, e_sigma = e_sigma, ...)
+
+  # Method of generating y
+  model$gen_y <- function(n) {
+    x <- model$gen_x(n)
+
+    y <- (2 - model$parameters$c) * x[,1] + model$parameters$c * x[,2] +
+      model$parameters$a * ((2 - model$parameters$c) * x[,1])^2 + model$parameters$a * (model$parameters$c * x[,2])^2 +
+      model$parameters$b * ((2 - model$parameters$c) * x[,1])^3 + model$parameters$b * (model$parameters$c * x[,2])^3 +
+      model$gen_e(n) + 1
+
+    return(data.frame(x = x[,1], z = x[,2], y))
+
+  }
+
+  model$fit <- function(dat) {
+    lm(y ~ x + z, data = dat)
+  }
+
+  model$effect_size <- function(dat) {
+    n <- nrow(dat)
+    xa <- dat %>%
+      mutate(c = 1) %>%
+      select(c, x, z) %>%
+      as.matrix()
+    ra <- diag(n) - xa %*% solve(t(xa) %*% xa) %*% t(xa)
+    diag_ra <- diag(sqrt(diag(ra)))
+    xb <- dat %>%
+      mutate(x2=x^2,z2=z^2,x3=x^3,z3=z^3) %>%
+      select(x2, z2, x3, z3) %>%
+      as.matrix()
+    beta_b <- matrix(c(model$parameters$a,model$parameters$a,model$parameters$b,model$parameters$b))
+
+    return((1/n) *
+      (1/model$parameters$e_sigma^2) *
+      sum((diag(sqrt(diag(ra))) %*% xb %*% beta_b)^2))
+  }
+
[…]

### 2021-12-06 experiment/scripts/main.R

@@ -1,0 +1,11 @@
+source(here::here("scripts/model_collections.R"))
+source(here::here("scripts/experiment.R"))
+
+set.seed(10086)
+
+myexp <- EXPERIMENT(name = "visual inference")
+higher_order <- HIGHER_ORDER_MODEL(a = 0.5, b = 1, c = 0,
+                                   x_dist = "uniform",
+                                   x_discrete = FALSE,
+                                   e_dist = "normal", e_sigma = 1)
+myexp$add_model("higher order", higher_order)

### 2021-12-06 experiment/scripts/model.R

@@ -1,0 +1,259 @@
+library(tidyverse)
+
+MODEL_SPEC <- function(...) {
+
+  # The MODEL_SPEC CLASS constructor
+
+  # Crate a new environment and store parameters
+  model <- new.env()
+  model$parameters <- list(...)
+
+  model$gen_x <- function(n) {
+
+    # The default method of generating x.
+    #
+    # Available arguments:
+    #
+    # x_dist: the distribution of x, one of "uniform", "normal", "lognormal", "neglognormal",
+    #   "uniform": (-1, 1),
+    #   "normal": normal(x_mu, x_sigma),
+    #   "lognormal": lognormal(x_mu, x_sigma^2)/3 - 1,
+    #   "neglognormal": -lognormal(x_mu, x_sigma^2)/3 + 1,
+    #   "discrete_uniform": n possible values from [-1, 1],
+    #   "binomial": binomial(n, 0.5)/n*2 - 1.
+    #
+    # x_mu: mean of lognormal or normal.
+    # x_sigma: sigma of lognormal or normal.
+    # x_n: n of binomial or discrete_uniform.
+    # z_discrete: TRUE or FALSE. Whether or not to introduce an additional discrete z. Otherwise, z ~ U(-1,1).
+    #
+    # z_discrete_dist: the distribution of discrete z, one of "discrete_uniform" and "binomial",
+    #   "discrete_uniform": n possible values from [-1, 1],
+    #   "binomial": binomial(n, 0.5)/n*2 - 1.
+    #
+    # z_n: n of binomial or discrete_uniform.
+
+    x_dist <- model$parameters$x_dist
+    z_discrete <- model$parameters$z_discrete
+    z_discrete_dist <- model$parameters$z_discrete_dist
+    z_n <- model$parameters$z_n
+    x_mu <- model$parameters$x_mu
+    x_sigma <- model$parameters$x_sigma
+    x_n <- model$parameters$x_n
+
+    stand_dist <- function(n, func, ...) {
+      result <- func(n, ...)
+      result / max(result)
+    }
+
+    mylognormal <- function(n, meanlog, sdlog) {
+      rlnorm(n, meanlog, sdlog)/3 - 1
+    }
+
+    myneglognormal <- function(n, meanlog, sdlog) {
+      -mylognormal(n, meanlog, sdlog)
+    }
+
+    mybinom <- function(n, size, prob) {
+      rbinom(n, size, prob)/size*2-1
+    }
+
[…]

### 2021-12-06 experiment/scripts/model_collections.R

@@ -1,0 +1,4 @@
+source(here::here("scripts/model.R"))
+source(here::here("scripts/higher_order_model.R"))
+source(here::here("scripts/heteroskedasticity_model.R"))
+source(here::here("scripts/non_normal_model.R"))

### 2021-12-06 experiment/scripts/process_survey_big.R

@@ -1,0 +1,164 @@
+source(here::here("scripts/model.R"))
+
+survey_tools <- new.env()
+
+survey_tools$process_responses <- function(k, lineup_dat, lineup_ord) {
+
+  dat <- tibble(id = 1:length(unlist(lineup_ord)),
+                lineup_dat_id = unlist(lineup_ord)) %>%
+    mutate(set = (id-1) %/% k + 1) %>%
+    mutate(num = (id-1) %% k + 1)
+
+  user_info_dict <- list(c("18-24", "25-39", "40-54", "55-64", "65 or above"),
+                         c("High School or below", "Diploma and Bachelor Degree", "Honours Degree", "Masters Degree", "Doctoral Degree"),
+                         c("He", "She", "They", "Other"),
+                         c("Yes", "No"))
+
+  get_responses <- function(user_info_dict, uuid) {
+    tmp_survey <- jsonlite::fromJSON(here::here(glue::glue("study_1/study_1_big_check/survey/{uuid}.txt")),
+                                     simplifyVector = FALSE)
+    response_time <- map_dbl(tmp_survey, ~.x$rt)
+    user_information <- str_split(tmp_survey[[3]]$response[[1]], ",")[[1]]
+
+    user_information[2:5] <- imap_chr(user_information[-1], ~user_info_dict[[.y]][as.integer(.x)])
+    names(user_information) <- c("prolific_id", "age_group", "education", "pronoun", "previous_experience")
+
+
+    lineup_respone <- str_split(map_chr(tmp_survey[-c(1,2,3,4)], ~.x$response$response), ",")
+    # handle selections
+    selections <- map_chr(lineup_respone, ~.x[1])
+    # handle reasons
+    reasons_dict <- c("Outlier(s)", "Cluster(s)", "Shape")
+
+    reasons <- map_chr(lineup_respone, ~if(.x[2] %in% c("1", "2", "3")) {
+      reasons_dict[as.integer(.x[2])]
+    } else {
+      .x[2]
+    })
+
+    # handle confidence
+    confidence_dict <- c("Not at all", "Slightly", "Moderately", "Very", "Extremely")
+    confidence <- map_chr(lineup_respone, ~if(.x[3] %in% c("1", "2", "3", "4", "5")) {
+      confidence_dict[as.integer(.x[3])]
+    } else {
+      .x[3]
+    })
+
+    tibble(page = 1:length(response_time),
+           response_time = response_time,
+           set = uuid,
[…]

### 2021-12-06 site/06-work.Rmd
liter review minor changes

# Works

Visual inference is an alternative to the traditional statical tests. It is particularly useful when there is no suitable numerical tests. However, one of the drawbacks of visual inference is it needs to be conducted by human, which limits its use cases.

The aim of this work is to provide a tool for automating the test procedure of visual inference. In other words, provided a lineup, the agent developed by us would need to be able to, at least, pick up the most different plot.

## S1 2021

Considering the lineup protocol is a graphical representation of the data, we can define the choice function $f$ for a perfect observer on a collection of nonempty sets $$X = \{\{p_{1,1}, p_{1,2}, ..., p_{1,M}\},\{p_{2,1}, p_{2,2}, ..., p_{2,M}\},...,\{p_{L,1}, p_{L,2}, ..., p_{L,M}\}\}$$ and assigns one element of each set $S$ in this collection to $S$ by $f(S)$, where $p_{l,m}$ is the $i$th plot of the $l$th lineup for $1\leq l \leq L$ and $1\leq m \leq M$, $M$ is the size of the lineup and $L$ is the number of lineups. Thus, $f$ is a perfect selector which always select the visually most different plot.

We could further assume there exists a real-valued function $g$ such that $$f(S) = \sup\{g(s_1; S\setminus\{s_1\}),g(s_2;S\setminus\{s_2\}),...,g(s_M;S\setminus\{s_M\})\},$$ where $s_i$ is the $i$th element of the set $S$ for $1 \leq i \leq M$. Then, there is a natural explanation of $g(s_i;S\setminus\{s_i\})$, which is the visual difference between $s_i$ and the rest of the plots or the strength of the signal of $s_i$. Finding $g$ is not trivial. There are some previous researches that attempt to find such a function with manually defined distance metrics under relatively simple graphical settings and show promising results. However, those distance metrics are case-specific, which means new distance metrics needs to be defined, if they could be defined, for new cases. Obviously, the function $g$ is the linkage between visual test and numerical tests. Unless we can properly define "visual difference" and accurately quantify it, we may never find $g$, not to mention the possibility of $g$ does not exist.

An alternative will be to find the selector $f$ directly. Unfortunately, perfect observer does not exist, but the selector can still be approximated. Given a lineup and $K$ imperfect observers with probability $p_i \geq 1/M$ to pick the most different plot for $1 \leq i \leq K$, the most frequently chosen plot is a reasonable estimator of the output of the selector $f$.

### 2021-12-06 site/_book/site.tex
liter review minor changes

@@
 \author{Patrick Li}
-\date{2021-06-20}
+\date{2021-07-08}
 
@@
 \item
-  to-do list
+  TO-DO list
 \item
@@
 
+\hypertarget{works}{%
+\chapter{Works}\label{works}}
+
+Visual inference is an alternative to the traditional statical tests.
+It is particularly useful when there is no suitable numerical tests.
+However, one of the drawbacks of visual inference is it needs to be conducted by human, which limits its use cases.
+
+The aim of this work is to provide a tool for automating the test procedure of visual inference.
+In other words, provided a lineup, the agent developed by us would need to be able to, at least, pick up the most different plot.
+
+\hypertarget{s1-2021}{%
+\section{S1 2021}\label{s1-2021}}
+
+Considering the lineup protocol is a graphical representation of the data, we can define the choice function \(f\) for a perfect observer on a collection of nonempty sets \[X = \{\{p_{1,1}, p_{1,2}, ..., p_{1,M}\},\{p_{2,1}, p_{2,2}, ..., p_{2,M}\},...,\{p_{L,1}, p_{L,2}, ..., p_{L,M}\}\}\] and assigns one element of each set \(S\) in this collection to \(S\) by \(f(S)\), where \(p_{l,m}\) is the \(i\)th plot of the \(l\)th lineup for \(1\leq l \leq L\) and \(1\leq m \leq M\), \(M\) is the size of the lineup and \(L\) is the number of lineups.
+Thus, \(f\) is a perfect selector which always select the visually most different plot.
+
+We could further assume there exists a real-valued function \(g\) such that \[f(S) = \sup\{g(s_1; S\setminus\{s_1\}),g(s_2;S\setminus\{s_2\}),...,g(s_M;S\setminus\{s_M\})\},\] where \(s_i\) is the \(i\)th element of the set \(S\) for \(1 \leq i \leq M\).
+Then, there is a natural explanation of \(g(s_i;S\setminus\{s_i\})\), which is the visual difference between \(s_i\) and the rest of the plots or the strength of the signal of \(s_i\).
+Finding \(g\) is not trivial.
+There are some previous researches that attempt to find such a function with manually defined distance metrics under relatively simple graphical settings and show promising results.
+However, those distance metrics are case-specific, which means new distance metrics needs to be defined, if they could be defined, for new cases.
+Obviously, the function \(g\) is the linkage between visual test and numerical tests.
+Unless we can properly define ``visual difference'' and accurately quantify it, we may never find \(g\), not to mention the possibility of \(g\) does not exist.
+
+An alternative will be to find the selector \(f\) directly.
+Unfortunately, perfect observer does not exist, but the selector can still be approximated.
+Given a lineup and \(K\) imperfect observers with probability \(p_i \geq 1/M\) to pick the most different plot for \(1 \leq i \leq K\), the most frequently chosen plot is a reasonable estimator of the output of the selector \(f\).
+
   \bibliography{book.bib,packages.bib}

### 2022-01-26 Thesis/01-chap1.Rmd
add paper 1 intro

---
chapter: 1
knit: "bookdown::render_book"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, cache=TRUE)
# Load any R packages you need here
library(forecast)
library(ggplot2)
```

# Introduction {#ch:intro}

### 2022-01-26 Thesis/_book/thesis.tex
add paper 1 intro

@@
 
-\author{Joe Bloggs}
-\title{Thesis title}
-\degrees{B.Sc. (Hons), University of Tangambalanga}
+\author{Weihao Li}
+\title{Advances in Artificial Intelligence for Data Visualization: Developing Computer Vision Models to Automate Reading of Data Plots, with Application to Predictive Model Diagnostics}
+\degrees{B.Comm. (Hons), Monash University}
 \def\degreetitle{Doctor of Philosophy}
@@
      %pdfkeywords={Some Keywords},
-     pdfauthor={Joe Bloggs},
-     pdftitle={Thesis title},
+     pdfauthor={Weihao Li},
+     pdftitle={Advances in Artificial Intelligence for Data Visualization: Developing Computer Vision Models to Automate Reading of Data Plots, with Application to Predictive Model Diagnostics},
      pdfproducer={Bookdown with LaTeX}
@@
 
-This is where you introduce the main ideas of your thesis, and an overview of the context and background.
+\hypertarget{ch:paper1}{%
+\chapter{Automatic Visual Statistical Inference, with Application to Linear Regression Diagnostics}\label{ch:paper1}}
 
-In a PhD, Chapter 2 would normally contain a literature review.
-Typically, Chapters 3--5 would contain your own contributions.
-Think of each of these as potential papers to be submitted to journals.
-Finally, Chapter 6 provides some concluding remarks, discussion, ideas for future research, and so on.
-Appendixes can contain additional material that don't fit into any chapters, but that you want to put on record.
-For example, additional tables, output, etc.
+\hypertarget{abstract-1}{%
+\section{Abstract}\label{abstract-1}}
 
-\hypertarget{rmarkdown}{%
-\section{Rmarkdown}\label{rmarkdown}}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
-In this template, the rest of the chapter shows how to use Rmarkdown.
-The big advantage of using Rmarkdown is that it allows you to include your R code directly into your thesis, to ensure there are no errors in copying and pasting, and that everything is reproducible.
-It also helps you stay better organized.
+\hypertarget{model-diagnostics}{%
+\subsection{Model Diagnostics}\label{model-diagnostics}}
 
-For details on using \emph{R Markdown} see \url{http://rmarkdown.rstudio.com}.
+Model diagnostics is the part of data analysis, preceded by the fit of a model, whose primary objectives are to examine the goodness of fit and reveal potential violations of assumptions.
+In these diagnostics, though numeric summaries are available and some are endorsed by finite or asymptotic properties, graphic representation of data is preferred or at least needed, due to its intuitiveness and the possibility to provide unexpected discoveries that may be abstract and could not be quantified algorithmically.
 
-\hypertarget{data}{%
-\section{Data}\label{data}}
+However, unlike confirmatory data analysis which is built upon rigorous statistical procedures, e.g., hypothesis testing, visual diagnostics relies on graphical perception - human's ability to interpret and decode the information embedded in the graph \autocite{cleveland_graphical_1984}, which is to some extent subjective.
+Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
+One such example is finding a separation between gene groups in a two-dimensional projection from a linear discriminant analysis where there is no difference in the expression levels between the gene groups \autocite{roy_chowdhury_using_2015}.
 
-Included in this template is a file called \texttt{sales.csv}.
-This contains quarterly data on Sales and Advertising budget for a small company over the period 1981--2005.
-It also contains the GDP (gross domestic product) over the same period.
-All series have been adjusted for inflation.
-We can load in this data set using the following command:
+\hypertarget{visual-inference}{%
+\subsection{Visual Inference}\label{visual-inference}}
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{sales }\OtherTok{\textless{}{-}} \FunctionTok{ts}\NormalTok{(}\FunctionTok{read.csv}\NormalTok{(}\StringTok{"data/sales.csv"}\NormalTok{)[,}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{], }\AttributeTok{start=}\DecValTok{1981}\NormalTok{, }\AttributeTok{frequency=}\DecValTok{4}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-Any data you use in your thesis can go into the data directory.
-The data should be in exactly the format you obtained it.
-Do no editing or manipulation of the data outside of R.
-Any data munging should be scripted in R and form part of your thesis files (possibly hidden in the output).
-
-\hypertarget{figures}{%
-\section{Figures}\label{figures}}
-
-Figure \ref{fig:deaths} shows time plots of the data we just loaded.
-Notice how figure captions and references work.
-Chunk names can be used as figure labels with \texttt{fig:} prefixed.
-Never manually type figure numbers, as they can change when you add or delete figures.
-This way, the figure numbering is always correct.
+Visual inference was first introduced by \textcite{buja_statistical_2009} as an inferential framework to extend confirmatory statistics to visual discoveries.
+This framework redefines the test statistics, tests, null distribution, significance levels and \(p\)-value for visual discovery modelled on the confirmatory statistical testing.
+Figure \ref{fig:parallelism} outlines the parallelism between conventional tests and visual discovery.
 
@@
 \centering
-\includegraphics{thesis_files/figure-latex/deaths-1.pdf}
-\caption{\label{fig:deaths}Quarterly sales, advertising and GDP data.}
+\includegraphics[width=4.6875in,height=3.55208in]{figures/rsta2009012001.jpg}
+\caption{Parallelism between multiple quantitative testing and visual discovery \autocite{buja_statistical_2009}. \label{fig:parallelism}}
 \end{figure}
 
-\hypertarget{results-from-analyses}{%
-\section{Results from analyses}\label{results-from-analyses}}
+In visual inference, a visual discovery is defined as a rejection of a null hypothesis, and the same null hypothesis can be rejected by many different visual discoveries \autocite{buja_statistical_2009}.
+For model diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the hypothesis.
+The same assumed model, such as classical linear regression model, can be rejected by both nonlinearity and heteroskedasticity with the residual plot as shown in Figure \ref{fig:nonlinearityheter}.
 
-We can fit a dynamic regression model to the sales data.
+\hypertarget{pre-specification-of-visual-discoverable-features}{%
+\subsection{Pre-specification of Visual Discoverable Features}\label{pre-specification-of-visual-discoverable-features}}
 
-If \(y_t\) denotes the sales in quarter \(t\), \(x_t\) denotes the corresponding advertising budget and \(z_t\) denotes the GDP, then the resulting model is:
-\begin{equation}
-  y_t - y_{t-4} = \beta (x_t-x_{t-4}) + \gamma (z_t-z_{t-4}) + \theta_1 \varepsilon_{t-1} + \Theta_1 \varepsilon_{t-4} + \varepsilon_t
-\end{equation}
-where
-\(\beta = 2.28\),
-\(\gamma = 0.97\),
-\(\theta_1 = NA\),
-and
-\(\Theta_1 = -0.90\).
+As discussed in \textcite{buja_statistical_2009}, in the practice of model diagnostics, the range of possible visual discoveries is not pre-specified.
+In other words, people do not explicitly specify which one or more visual features they are looking for before the read of the diagnostic plot.
+This is concerning since conventional hypothesis testing always requires the pre-specification of the parameter space \(\Theta\) of the parameter of interest \(\theta \in \Theta\) to form a valid inferential procedure.
+To address this issue, a collection of test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) is defined, where \(\boldsymbol{\mathrm{y}}\) is the data and \(\boldsymbol{I}\) is a set of all possible visual features.
+\textcite{buja_statistical_2009} described each of the test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})\) as a measurement of the degree of presence of a visual feature.
+Alternatively, \textcite{majumder_validation_2013} avoids the use of visual features and defined the visual statistics \(T(.)\) as a mapping from a dataset to a data plot.
+Both definitions of visual test statistics are valid, but in the rest of the paper, the first definition will be used as it covered some details needed by this work.
 
-\hypertarget{tables}{%
-\section{Tables}\label{tables}}
+The size of the collection \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) depends on the size of the set \(I\).
+If one can define \(I\) comprehensively, i.e, pre-specify all the visual discoverable features, the validity issue will be solved.
+Unfortunately, to our knowledge, there is no such a way to list all visual features.
+In linear regression diagnostics, possible visual features of a residual plot may be outliers, shapes and clusters.
+But these are very vague features which does not cover the full list of visual features.
 
-Let's assume future advertising spend and GDP are at the current levels. Then forecasts for the next year are given in Table \ref{tab:salesforecasts}.
+Similarly, \textcite{wilkinson_graph-theoretic_2005} proposed the work called graph theoretic scagnostics, which adopted the idea of ``scagnostics'' - scatter plot diagnostics from .
+It includes 9 computable scagnostics measures defined on planar proximity graphs: ``Outlying'', ``Convex'', ``Skinny'', ``Stringy'', ``Straight'', ``Monotonic'', ``Skewed'', ``Clumpy'' and ``Striated'' which attempts to describe outliers, shape, density, trend and coherence of the data.
+This approach is inspiring but it does not contain the complete list of visual discoverable features.
+It is possible that such a list will never be complete as suggested in \textcite{buja_statistical_2009}.
 
-\begin{verbatim}
-## Warning in sprintf(str, x, ...): one argument not used by format '\toprule
-## Point Forecast & Lo 80 & Hi 80 & Lo 95 & Hi 95 \\
-## \midrule'
+Thinking out of the box, \textcite{buja_statistical_2009} argued that there is actually no need for pre-specification of visual discoverable features.
+In model diagnostics, when the null hypothesis is rejected, the reasons for rejecting the hypothesis will also be known.
+This is because observers not only can point out the fact that visual discoveries have been found, but also can describe the particular features they observed.
+Those features will correspond to the subset of the collection of visual test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) which resulted in rejection.
 
-## Warning in sprintf(str, x, ...): one argument not used by format '\toprule
-## Point Forecast & Lo 80 & Hi 80 & Lo 95 & Hi 95 \\
-## \midrule'
-\end{verbatim}
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/nonlinearityheter-1.pdf}
+\caption{\label{fig:nonlinearityheter}test}
+\end{figure}
 
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{lllll}
-\toprule
-Point Forecast & Lo 80 & Hi 80 & Lo 95 & Hi 95 \\
-\midrule
-1000.2 &  947.7 & 1052.7 & 919.9 & 1080.5 \\
-1013.1 &  959.3 & 1066.8 & 930.9 & 1095.3 \\
-1076.7 & 1022.9 & 1130.6 & 994.4 & 1159.0 \\
-1003.5 &  949.7 & 1057.4 & 921.2 & 1085.8 \\
-\bottomrule
-\end{tabular}
-\caption{Forecasts for the next year assuming Advertising budget and GDP are unchanged.}
-\label{tab:salesforecasts}
-\end{center}
-\end{table}
+\hypertarget{lineup-protocol}{%
+\subsection{Lineup Protocol}\label{lineup-protocol}}
 
-Again, notice the use of labels and references to automatically generate Table numbers. In this case, we need to generate the label ourselves.
+\hypertarget{type-i-error}{%
+\subsubsection{Type-I error}\label{type-i-error}}
 
-The \texttt{knitLatex} package is useful for generating tables from R output.
-Other packages can do similar things including the \texttt{kable} function in \texttt{knitr} which is somewhat simpler but you have less control over the result.
-If you use \texttt{knitLatex} to generate tables, don't forget to include \texttt{results="asis"} in the chunk settings.
+\hypertarget{type-ii-error}{%
+\subsubsection{Type-II error}\label{type-ii-error}}
 
-\hypertarget{ch:litreview}{%
-\chapter{Literature Review}\label{ch:litreview}}
+\hypertarget{visual-inference-applied-to-linear-regression}{%
+\subsection{Visual Inference Applied to Linear Regression}\label{visual-inference-applied-to-linear-regression}}
 
-This chapter contains a summary of the context in which your research is set.
+How people used visual inference in linear regression?
 
-Imagine you are writing for your fellow PhD students. Topics that are well-known to them do not have to be included here. But things that they may not know about should be included.
+\hypertarget{limitation-of-the-visual-inference}{%
+\subsection{Limitation of the Visual Inference}\label{limitation-of-the-visual-inference}}
 
-Resist the temptation to discuss everything you've read in the last few years.
-And you are not writing a textbook either.
-This chapter is meant to provide the background necessary to understand the material in subsequent chapters.
-Stick to that.
+What are the limitations?
 
-You will need to organize the literature review around themes, and within each theme provide a story explaining the development of ideas to date.
-In each theme, you should get to the point where your ideas will fit in.
-But leave your ideas to later chapters.
-This way it is clear what has been done beforehand, and what new contributions you are making to the research field.
+\hypertarget{computer-vision-model}{%
+\subsection{Computer Vision Model}\label{computer-vision-model}}
 
-All citations should be done using markdown notation as shown below. This way, your bibliography will be compiled automatically and correctly.
+What is computer vision model?
 
-\hypertarget{sec:expsmooth}{%
-\section{Exponential smoothing}\label{sec:expsmooth}}
+\hypertarget{contribution}{%
+\subsection{Contribution}\label{contribution}}
 
-Exponential smoothing was originally developed in the late 1950s \autocite{Brown59,Brown63,Holt57,Winters60}.
-Because of their computational simplicity and interpretability, they became widely used in practice.
+What has been done by this paper?
 
-Empirical studies by \textcite{MH79} and \textcite{Metal82} found little difference in forecast accuracy between exponential smoothing and ARIMA models.
-This made the family of exponential smoothing procedures an attractive proposition \autocite[see][]{CKOS01}.
+\hypertarget{structure-of-this-paper}{%
+\subsection{Structure of This Paper}\label{structure-of-this-paper}}
 
-The methods were less popular in academic circles until \textcite{OKS97} introduced a state space formulation of some of the methods, which was extended in \textcite{HKSG02} to cover the full range of exponential smoothing methods.
+What is the structure of the paper?
+
+Model diagnostics is the part of data analysis whose primary objectives are to examine the goodness of the model fit and reveal potential violations of the assumptions. Graphical approaches
+
+For regression diagnostics, it may includes the needs of
+
+Linear regression is an modelling approach to describe the relationship between an response variable and one or more explanatory variable.
+It has been widely used for both generative modeling and predictive modelling.
+
+Regression diagnostics is needed
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  to check whether the assumptions has been violated
+\item
+  to check whether the line fit the data
+\end{enumerate}
+
+Model diagnostics for linear regression is well developed
 

### 2022-01-26 Thesis/index.Rmd
add paper 1 intro

---
title: 'Advances in Artificial Intelligence for Data Visualization: Developing Computer Vision Models to Automate Reading of Data Plots, with Application to Predictive Model Diagnostics'
degreetype: 'Doctor of Philosophy'
author: 'Weihao Li'
degrees: 'B.Comm. (Hons), Monash University'
output: bookdown::pdf_book
site: bookdown::bookdown_site
link-citations: yes
knit: "bookdown::render_book"
---

<!-- 
Edit these lines as appropriate.
The actual thesis content is in several Rmd files.

You'll need to edit the _bookdown.yml file to set the order in which you'd like them to appear.

If you have specific LaTeX packages to add, put them in monashthesis.tex.

You will need to ensure you have installed the knitr and bookdown packages for R.

You will also need LaTeX installed on your computer.
-->

### 2022-01-26 Thesis/02-chap2.Rmd
add paper 1 intro

--- 2022-01-26 Thesis/_book/thesis.tex
@@
-% This is a LaTeX thesis template for Monash University.
-% to be used with Rmarkdown
-% This template was produced by Rob Hyndman
-% Version: 6 September 2016
+---
+chapter: 2
+knit: "bookdown::render_book"
+---
 
-\documentclass{monashthesis}
+# Automatic Visual Statistical Inference, with Application to Linear Regression Diagnostics {#ch:paper1}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Add any LaTeX packages and other preamble here if required
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+## Abstract
 
-\author{Weihao Li}
-\title{Advances in Artificial Intelligence for Data Visualization: Developing Computer Vision Models to Automate Reading of Data Plots, with Application to Predictive Model Diagnostics}
-\degrees{B.Comm. (Hons), Monash University}
-\def\degreetitle{Doctor of Philosophy}
-% Add subject and keywords below
-\hypersetup{
-     %pdfsubject={The Subject},
-     %pdfkeywords={Some Keywords},
-     pdfauthor={Weihao Li},
-     pdftitle={Advances in Artificial Intelligence for Data Visualization: Developing Computer Vision Models to Automate Reading of Data Plots, with Application to Predictive Model Diagnostics},
-     pdfproducer={Bookdown with LaTeX}
-}
+## Introduction
 
-\bibliography{thesisrefs}
-
-\begin{document}
-
-\pagenumbering{roman}
-
-\titlepage
-
-{\setstretch{1.2}\sf\tighttoc\doublespacing}
-
-\hypertarget{copyright-notice}{%
-\chapter*{Copyright notice}\label{copyright-notice}}
-\addcontentsline{toc}{chapter}{Copyright notice}
-
-\emph{(Choose one of the following notices.)}
-
-\emph{(Notice 1)}
-
-\textcopyright { } \authorname~(\number\the\year).
-
-\emph{The second notice certifies the appropriate use of any third-party material in the thesis.
-Students choosing to deposit their thesis into the restricted access section of the repository are not required to complete Notice 2.}
-
-\emph{(Notice 2)}
-
-\textcopyright { } \authorname~(\number\the\year).
-
-I certify that I have made all reasonable efforts to secure copyright permissions for third-party content included in this thesis and have not knowingly added copyright content to my work without the owner's permission.
-
-\newpage
-
-\hypertarget{abstract}{%
-\chapter*{Abstract}\label{abstract}}
-\addcontentsline{toc}{chapter}{Abstract}
-
-The abstract should outline the main approach and findings of the thesis and must not be more than 500 words.
-
-\newpage
-
-\hypertarget{declaration}{%
-\chapter*{Declaration}\label{declaration}}
-\addcontentsline{toc}{chapter}{Declaration}
-
-\emph{(Standard thesis)}
-
-This thesis is an original work of my research and contains no material which has been accepted for the award of any other degree or diploma at any university or equivalent institution and that, to the best of my knowledge and belief, this thesis contains no material previously published or written by another person, except where due reference is made in the text of the thesis.
-
-\emph{(Thesis including published works declaration)}
-
-I hereby declare that this thesis contains no material which has been accepted for the award of any other degree or diploma at any university or equivalent institution and that, to the best of my knowledge and belief, this thesis contains no material previously published or written by another person, except where due reference is made in the text of the thesis.
-
-This thesis includes (insert number) original papers published in peer reviewed journals and (insert number) submitted publications.
-The core theme of the thesis is (insert theme).
-The ideas, development and writing up of all the papers in the thesis were the principal responsibility of myself, the student, working within the (insert name of academic unit) under the supervision of (insert name of supervisor).
-
-(The inclusion of co-authors reflects the fact that the work came from active collaboration between researchers and acknowledges input into team-based research.) Remove this paragraph for theses with sole-authored work
-
-In the case of (insert chapter numbers) my contribution to the work involved the following:
-
-I have / have not renumbered sections of submitted or published papers in order to generate a consistent presentation within the thesis.
-
-\begin{table}
-\centering\footnotesize\tabcolsep=0.12cm
-\begin{tabular}{|p{1.2cm}|p{2cm}|p{1.8cm}|p{3.4cm}|p{3.5cm}|p{1.5cm}|}
-\hline
-\RaggedRight\textbf{Thesis chapter}  & 
-\RaggedRight\textbf{Publication title}  & 
-\RaggedRight\textbf{Status (published, in press, accepted or returned for revision)}  &  
-\RaggedRight\textbf{Nature and \% of student contribution} & 
-\RaggedRight\textbf{Co-author name(s), nature and \% of co-author’s contribution} &  \RaggedRight\textbf{Co-author(s), Monash student Y/N} \\ 
-\hline
-2 & xx & xx  & xx & xx & \multicolumn{1}{c|}{N}   \\
-\hline
-3 & xx & xx  & xx & xx  & \multicolumn{1}{c|}{N}   \\
-\hline
-4 & xx & xx  & xx & xx & \multicolumn{1}{c|}{N}    \\
-\hline
-5 & xx & xx  & xx & xx & \multicolumn{1}{c|}{N}   \\
-\hline
-\end{tabular}
-\end{table}
-
-\textbf{Student name:} \authorname
-
-\textbf{Student signature:}
-
-\textbf{Date:}
-
-\hypertarget{acknowledgements}{%
-\chapter*{Acknowledgements}\label{acknowledgements}}
-\addcontentsline{toc}{chapter}{Acknowledgements}
-
-I would like to thank my pet goldfish for \dots
-
-\hypertarget{preface}{%
-\chapter*{Preface}\label{preface}}
-\addcontentsline{toc}{chapter}{Preface}
-
-The material in Chapter \ref{ch:intro} has been submitted to the journal \emph{Journal of Impossible Results} for possible publication.
-
-The contribution in Chapter \ref{ch:litreview} of this thesis was presented in the International Symposium on Nonsense held in Dublin, Ireland, in July 2015.
-
-\clearpage\pagenumbering{arabic}\setcounter{page}{0}
-
-\hypertarget{ch:intro}{%
-\chapter{Introduction}\label{ch:intro}}
-
-\hypertarget{ch:paper1}{%
-\chapter{Automatic Visual Statistical Inference, with Application to Linear Regression Diagnostics}\label{ch:paper1}}
-
-\hypertarget{abstract-1}{%
-\section{Abstract}\label{abstract-1}}
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-\hypertarget{model-diagnostics}{%
-\subsection{Model Diagnostics}\label{model-diagnostics}}
+### Model Diagnostics
 
@@
 
-However, unlike confirmatory data analysis which is built upon rigorous statistical procedures, e.g., hypothesis testing, visual diagnostics relies on graphical perception - human's ability to interpret and decode the information embedded in the graph \autocite{cleveland_graphical_1984}, which is to some extent subjective.
+However, unlike confirmatory data analysis which is built upon rigorous statistical procedures, e.g., hypothesis testing, visual diagnostics relies on graphical perception - human’s ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984], which is to some extent subjective.
 Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
-One such example is finding a separation between gene groups in a two-dimensional projection from a linear discriminant analysis where there is no difference in the expression levels between the gene groups \autocite{roy_chowdhury_using_2015}.
+One such example is finding a separation between gene groups in a two-dimensional projection from a linear discriminant analysis where there is no difference in the expression levels between the gene groups [@roy_chowdhury_using_2015].
 
-\hypertarget{visual-inference}{%
-\subsection{Visual Inference}\label{visual-inference}}
+### Visual Inference
 
-Visual inference was first introduced by \textcite{buja_statistical_2009} as an inferential framework to extend confirmatory statistics to visual discoveries.
-This framework redefines the test statistics, tests, null distribution, significance levels and \(p\)-value for visual discovery modelled on the confirmatory statistical testing.
+Visual inference was first introduced by @buja_statistical_2009 as an inferential framework to extend confirmatory statistics to visual discoveries.
+This framework redefines the test statistics, tests, null distribution, significance levels and $p$-value for visual discovery modelled on the confirmatory statistical testing.
 Figure \ref{fig:parallelism} outlines the parallelism between conventional tests and visual discovery.
 
-\begin{figure}
-\centering
-\includegraphics[width=4.6875in,height=3.55208in]{figures/rsta2009012001.jpg}
-\caption{Parallelism between multiple quantitative testing and visual discovery \autocite{buja_statistical_2009}. \label{fig:parallelism}}
-\end{figure}
+![(ref:parallelism) \label{fig:parallelism}](figures/rsta2009012001.jpg){width=450 height=341}
 
-In visual inference, a visual discovery is defined as a rejection of a null hypothesis, and the same null hypothesis can be rejected by many different visual discoveries \autocite{buja_statistical_2009}.
+(ref:parallelism) Parallelism between multiple quantitative testing and visual discovery [@buja_statistical_2009].
+
+In visual inference, a visual discovery is defined as a rejection of a null hypothesis, and the same null hypothesis can be rejected by many different visual discoveries [@buja_statistical_2009].
 For model diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the hypothesis.
@@
 
-\hypertarget{pre-specification-of-visual-discoverable-features}{%
-\subsection{Pre-specification of Visual Discoverable Features}\label{pre-specification-of-visual-discoverable-features}}
+### Pre-specification of Visual Discoverable Features
 
-As discussed in \textcite{buja_statistical_2009}, in the practice of model diagnostics, the range of possible visual discoveries is not pre-specified.
+As discussed in @buja_statistical_2009, in the practice of model diagnostics, the range of possible visual discoveries is not pre-specified.
 In other words, people do not explicitly specify which one or more visual features they are looking for before the read of the diagnostic plot.
-This is concerning since conventional hypothesis testing always requires the pre-specification of the parameter space \(\Theta\) of the parameter of interest \(\theta \in \Theta\) to form a valid inferential procedure.
-To address this issue, a collection of test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) is defined, where \(\boldsymbol{\mathrm{y}}\) is the data and \(\boldsymbol{I}\) is a set of all possible visual features.
-\textcite{buja_statistical_2009} described each of the test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})\) as a measurement of the degree of presence of a visual feature.
-Alternatively, \textcite{majumder_validation_2013} avoids the use of visual features and defined the visual statistics \(T(.)\) as a mapping from a dataset to a data plot.
+This is concerning since conventional hypothesis testing always requires the pre-specification of the parameter space $\Theta$ of the parameter of interest $\theta \in \Theta$ to form a valid inferential procedure.
+To address this issue, a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$ is defined, where $\boldsymbol{\mathrm{y}}$ is the data and $\boldsymbol{I}$ is a set of all possible visual features.
+@buja_statistical_2009 described each of the test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})$ as a measurement of the degree of presence of a visual feature.
+Alternatively, @majumder_validation_2013 avoids the use of visual features and defined the visual statistics $T(.)$ as a mapping from a dataset to a data plot.
 Both definitions of visual test statistics are valid, but in the rest of the paper, the first definition will be used as it covered some details needed by this work.
 
-The size of the collection \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) depends on the size of the set \(I\).
-If one can define \(I\) comprehensively, i.e, pre-specify all the visual discoverable features, the validity issue will be solved.
+The size of the collection $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$ depends on the size of the set $I$.
+If one can define $I$ comprehensively, i.e, pre-specify all the visual discoverable features, the validity issue will be solved.
 Unfortunately, to our knowledge, there is no such a way to list all visual features.
@@
 
-Similarly, \textcite{wilkinson_graph-theoretic_2005} proposed the work called graph theoretic scagnostics, which adopted the idea of ``scagnostics'' - scatter plot diagnostics from .
-It includes 9 computable scagnostics measures defined on planar proximity graphs: ``Outlying'', ``Convex'', ``Skinny'', ``Stringy'', ``Straight'', ``Monotonic'', ``Skewed'', ``Clumpy'' and ``Striated'' which attempts to describe outliers, shape, density, trend and coherence of the data.
+Similarly, @wilkinson_graph-theoretic_2005 proposed the work called graph theoretic scagnostics, which adopted the idea of "scagnostics" - scatter plot diagnostics from .
+It includes 9 computable scagnostics measures defined on planar proximity graphs: "Outlying", "Convex", "Skinny", "Stringy", "Straight", "Monotonic", "Skewed", "Clumpy" and "Striated" which attempts to describe outliers, shape, density, trend and coherence of the data.
 This approach is inspiring but it does not contain the complete list of visual discoverable features.
-It is possible that such a list will never be complete as suggested in \textcite{buja_statistical_2009}.
+It is possible that such a list will never be complete as suggested in @buja_statistical_2009.
 
-Thinking out of the box, \textcite{buja_statistical_2009} argued that there is actually no need for pre-specification of visual discoverable features.
+Thinking out of the box, @buja_statistical_2009 argued that there is actually no need for pre-specification of visual discoverable features.
 In model diagnostics, when the null hypothesis is rejected, the reasons for rejecting the hypothesis will also be known.
 This is because observers not only can point out the fact that visual discoveries have been found, but also can describe the particular features they observed.
-Those features will correspond to the subset of the collection of visual test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) which resulted in rejection.
+Those features will correspond to the subset of the collection of visual test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$ which resulted in rejection.
 
-\begin{figure}
-\centering
-\includegraphics{thesis_files/figure-latex/nonlinearityheter-1.pdf}
-\caption{\label{fig:nonlinearityheter}test}
-\end{figure}
+```{r nonlinearityheter, fig.cap="test"}
+library(ggplot2)
+set.seed(10086)
+cm <- visage::cubic_model(a = 0, b = 100, c = 1)
 
-\hypertarget{lineup-protocol}{%
-\subsection{Lineup Protocol}\label{lineup-protocol}}
+dat <- cm$gen(1000, fit_model = TRUE)
+cm$plot(dat) +
+  ggtitle("Residuals against fitted values plot of a classical linear regression model") +
+  theme_light()
+```
 
-\hypertarget{type-i-error}{%
-\subsubsection{Type-I error}\label{type-i-error}}
+### Lineup Protocol
 
-\hypertarget{type-ii-error}{%
-\subsubsection{Type-II error}\label{type-ii-error}}
+#### Type-I error
 
-\hypertarget{visual-inference-applied-to-linear-regression}{%
-\subsection{Visual Inference Applied to Linear Regression}\label{visual-inference-applied-to-linear-regression}}
+#### Type-II error
+
+### Visual Inference Applied to Linear Regression
 
@@
 
-\hypertarget{limitation-of-the-visual-inference}{%
-\subsection{Limitation of the Visual Inference}\label{limitation-of-the-visual-inference}}
+### Limitation of the Visual Inference
 
@@
 
-\hypertarget{computer-vision-model}{%
-\subsection{Computer Vision Model}\label{computer-vision-model}}
+### Computer Vision Model
 
@@
 
-\hypertarget{contribution}{%
-\subsection{Contribution}\label{contribution}}
+### Contribution
 
@@
 
-\hypertarget{structure-of-this-paper}{%
-\subsection{Structure of This Paper}\label{structure-of-this-paper}}
+### Structure of This Paper
 
@@
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  to check whether the assumptions has been violated
-\item
-  to check whether the line fit the data
-\end{enumerate}
+1. to check whether the assumptions has been violated
+2. to check whether the line fit the data
 
 Model diagnostics for linear regression is well developed
-
-\appendix
-
-\hypertarget{additional-stuff}{%
-\chapter{Additional stuff}\label{additional-stuff}}
-
-You might put some computer output here, or maybe additional tables.
-
-Note that line 5 must appear before your first appendix. But other appendices can just start like any other chapter.
-
-\printbibliography[heading=bibintoc]
-
-\end{document}

### 2022-01-28 Thesis/_book/thesis.tex
modify some words

@@
 
-Model diagnostics is the part of data analysis, preceded by the fit of a model, whose primary objectives are to examine the goodness of fit and reveal potential violations of assumptions.
-In these diagnostics, though numeric summaries are available and some are endorsed by finite or asymptotic properties, graphic representation of data is preferred or at least needed, due to its intuitiveness and the possibility to provide unexpected discoveries that may be abstract and could not be quantified algorithmically.
+Model diagnostics is the part of data analysis, preceded by the fit of a model, whose primary objectives are to examine the goodness of fit and reveal potential violations of model assumptions.
+In these diagnostics, though numeric summaries are mostly available and some are even endorsed by finite or asymptotic properties, graphic representation of data is still preferred, or at least needed, due to its intuitiveness and the possibility to provide unexpected discoveries which may be abstract and
+unquantifiable.
 
-However, unlike confirmatory data analysis which is built upon rigorous statistical procedures, e.g., hypothesis testing, visual diagnostics relies on graphical perception - human's ability to interpret and decode the information embedded in the graph \autocite{cleveland_graphical_1984}, which is to some extent subjective.
+However, unlike confirmatory data analysis built upon rigorous statistical procedures, e.g., hypothesis testing, visual diagnostics relies on graphical perception - human's ability to interpret and decode the information embedded in the graph \autocite{cleveland_graphical_1984}, which is to some extent subjective.
 Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
@@
 The size of the collection \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) depends on the size of the set \(I\).
-If one can define \(I\) comprehensively, i.e, pre-specify all the visual discoverable features, the validity issue will be solved.
+Thus, if one can define \(I\) comprehensively, i.e, pre-specify all the visual discoverable features, the validity issue will be solved.
 Unfortunately, to our knowledge, there is no such a way to list all visual features.
 In linear regression diagnostics, possible visual features of a residual plot may be outliers, shapes and clusters.
-But these are very vague features which does not cover the full list of visual features.
+But this is an incomplete list which does not enumerate all the visual features.
 
-Similarly, \textcite{wilkinson_graph-theoretic_2005} proposed the work called graph theoretic scagnostics, which adopted the idea of ``scagnostics'' - scatter plot diagnostics from .
+Similarly, \textcite{wilkinson_graph-theoretic_2005} proposed the work called graph theoretic scagnostics, which adopted the idea of ``scagnostics'' - scatter plot diagnostics from (can't find the 1984 citation).
 It includes 9 computable scagnostics measures defined on planar proximity graphs: ``Outlying'', ``Convex'', ``Skinny'', ``Stringy'', ``Straight'', ``Monotonic'', ``Skewed'', ``Clumpy'' and ``Striated'' which attempts to describe outliers, shape, density, trend and coherence of the data.
-This approach is inspiring but it does not contain the complete list of visual discoverable features.
-It is possible that such a list will never be complete as suggested in \textcite{buja_statistical_2009}.
+This approach is inspiring but it still does not give the complete list of visual discoverable features.
+In fact, it is possible that such a list will never be complete as suggested in \textcite{buja_statistical_2009}.
 

### 2022-01-28 Thesis/02-chap2.Rmd
modify some words

@@
 
-Model diagnostics is the part of data analysis, preceded by the fit of a model, whose primary objectives are to examine the goodness of fit and reveal potential violations of assumptions. 
-In these diagnostics, though numeric summaries are available and some are endorsed by finite or asymptotic properties, graphic representation of data is preferred or at least needed, due to its intuitiveness and the possibility to provide unexpected discoveries that may be abstract and could not be quantified algorithmically.
+Model diagnostics is the part of data analysis, preceded by the fit of a model, whose primary objectives are to examine the goodness of fit and reveal potential violations of model assumptions. 
+In these diagnostics, though numeric summaries are mostly available and some are even endorsed by finite or asymptotic properties, graphic representation of data is still preferred, or at least needed, due to its intuitiveness and the possibility to provide unexpected discoveries which may be abstract and
+unquantifiable.
 
-However, unlike confirmatory data analysis which is built upon rigorous statistical procedures, e.g., hypothesis testing, visual diagnostics relies on graphical perception - human’s ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984], which is to some extent subjective.
+However, unlike confirmatory data analysis built upon rigorous statistical procedures, e.g., hypothesis testing, visual diagnostics relies on graphical perception - human’s ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984], which is to some extent subjective.
 Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
@@
 The size of the collection $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$ depends on the size of the set $I$.
-If one can define $I$ comprehensively, i.e, pre-specify all the visual discoverable features, the validity issue will be solved.
+Thus, if one can define $I$ comprehensively, i.e, pre-specify all the visual discoverable features, the validity issue will be solved.
 Unfortunately, to our knowledge, there is no such a way to list all visual features.
 In linear regression diagnostics, possible visual features of a residual plot may be outliers, shapes and clusters.
-But these are very vague features which does not cover the full list of visual features.
+But this is an incomplete list which does not enumerate all the visual features.
 
-Similarly, @wilkinson_graph-theoretic_2005 proposed the work called graph theoretic scagnostics, which adopted the idea of "scagnostics" - scatter plot diagnostics from .
+Similarly, @wilkinson_graph-theoretic_2005 proposed the work called graph theoretic scagnostics, which adopted the idea of "scagnostics" - scatter plot diagnostics from (can't find the 1984 citation).
 It includes 9 computable scagnostics measures defined on planar proximity graphs: "Outlying", "Convex", "Skinny", "Stringy", "Straight", "Monotonic", "Skewed", "Clumpy" and "Striated" which attempts to describe outliers, shape, density, trend and coherence of the data.
-This approach is inspiring but it does not contain the complete list of visual discoverable features.
-It is possible that such a list will never be complete as suggested in @buja_statistical_2009.
+This approach is inspiring but it still does not give the complete list of visual discoverable features.
+In fact, it is possible that such a list will never be complete as suggested in @buja_statistical_2009.
 
@@
 ```{r nonlinearityheter, fig.cap="test"}
-library(ggplot2)
+library(tidyverse)
+
 set.seed(10086)
@@
 
-dat <- cm$gen(1000, fit_model = TRUE)
-cm$plot(dat) +
+cm$gen(1000, fit_model = TRUE) %>%
+cm$plot() +
   ggtitle("Residuals against fitted values plot of a classical linear regression model") +

### 2022-02-02 Thesis/02-chap2.Rmd
chap2 update

@@
 In model diagnostics, when the null hypothesis is rejected, the reasons for rejecting the hypothesis will also be known.
-This is because observers not only can point out the fact that visual discoveries have been found, but also can describe the particular features they observed.
+This is because observers can not only point out the fact that visual discoveries have been found, but also describe the particular visual features they observed.
 Those features will correspond to the subset of the collection of visual test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$ which resulted in rejection.
+This argument helps justifies the validity of visual inference.
 
@@
 
-#### Type-I error
+With the validity of visual inference being justified, another aspect of hypothesis testing that needs to be addressed is the control of false positive rate or Type I error.
+Any visual statistic $T^{(i)}(\boldsymbol{\mathrm{y}})$ needs to pair with a critical value $c^{(i)}$ to form a hypothesis test.
+When a visual feature $i$ is discovered by the observer from a plot, the corresponding visual statistic $T^{(i)}(\boldsymbol{\mathrm{y}})$ may not be known as there is no general agreement on the measurement of the degree of presence of a visual feature.
+It is only the event that $T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}$ is confirmed.
+Similarly, if any visual discovery is found by the observer, we say, there exists $i \in I:~T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}$ [@buja_statistical_2009].
 
-#### Type-II error
+Using the above definition, the family-wise Type I error can be controlled if one can provide the collection of critical values $c^{(i)}~(i \in I)$ such that $P(\mathrm{there~exists~} i \in I: T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}|\boldsymbol{\mathrm{y}}) \leq \alpha$, where $\alpha$ is the significance level.
+However, since the quantity of $T^{(i)}(\boldsymbol{\mathrm{y}})$ may not be known, such collection of critical values can not be provided.
+
+@buja_statistical_2009 proposed the lineup protocol as a visual test to calibrate the Type I error issue without the specification of $c^{(i)}~(i \in I)$.
+It is inspired by the "police lineup" or "identity parade" which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people.
+The protocol consists of $m$ randomly placed data plots, where $1$ plot is the actual data plot, and $m-1$ null plots are produced by plotting data simulate from the null distribution which is consistent with the null hypothesis.
+Then, an observer who have not seen the actual data plot will be asked to point out the most different plot from the lineup.
+
+Under the null hypothesis, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is $1/m$ due to randomness.
+If we reject the null hypothesis as the observer correctly picks the actual data plot, then the Type I error of this test is $1/m$.
+
+This provides us with an mechanism to control the Type I error, because $m$ - the number of plots in a lineup can be chosen.
+Further, if we involve $K$ independent observers in a visual test, and let $X$ be a random variable denoting the number of observers correctly picking the actual data plot.
+Then, under the null hypothesis $X \sim \mathrm{Binom}_{K,1/m}$, and therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as
+
+$$P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},$$
+where $x$ is the realization of number of observers correctly picking the actual data plot [@majumder_validation_2013].
+
+#### Type I error
+
+#### Type II error
 

### 2022-02-02 Thesis/02-chap2.Rmd
ET comments on first section

--- 2022-01-28 Thesis/02-chap2.Rmd
@@
 
+[ET: suggestion: A model can be fitted to data with no guarantee of a meaningful interpretation.
+Model diagnostics play an important role in assessing the appropriateness of the model.
+The assessment can involve examining the goodness of fit, checking if there are potential violations of model assumptions and validating models with external information.]
+
 Model diagnostics is the part of data analysis, preceded by the fit of a model, whose primary objectives are to examine the goodness of fit and reveal potential violations of model assumptions.
-In these diagnostics, though numeric summaries are mostly available and some are even endorsed by finite or asymptotic properties, graphic representation of data is still preferred, or at least needed, due to its intuitiveness and the possibility to provide unexpected discoveries which may be abstract and
+
+[ET: model fit is more than just goodness of fit and checking model assumptions.
+It's too reductive to say this because you could have a model with good fit (according to some metric) and no model violation but if the way that the data were collected was flawed then scientific interpretation of the model is rendered useless.
+Diagnostics involve also the domain knowledge checks as well.]
+
+In these diagnostics, though numeric summaries are mostly available and some are even endorsed by finite or asymptotic properties, graphical representation of data is still preferred, or at least needed, due to its intuitiveness and the possibility to provide unexpected discoveries which may be abstract and
 unquantifiable.
+
+[ET: Generally, your sentences are too long.
+Break it up into smaller sentences.
+The generally idea is to make sentences that _continue an idea from the preceding sentence_ so it leads from one to the next naturally.
+E.g.
+Today is cloudy.
+Cloudy days are rare.
+Perhaps today will be different to other days!]
 
@@
 One such example is finding a separation between gene groups in a two-dimensional projection from a linear discriminant analysis where there is no difference in the expression levels between the gene groups [@roy_chowdhury_using_2015].
+
+[ET: Confirmatory data analysis is well characterised by a rigorous procedure to discern particular hypothesis.
+The most widespread form of confirmatory data analysis are the use of $p$-values in testing hypothesis in a frequentist framework.
+More specifically, hypothesis testing in a frequentist framework involve stating a well-defined hypotheses; summarising the evidence as a test statistic; and calculating the probability of observing a statistic as extreme as the observed test statistic under the null hypothesis.
+This rigour, however, is often not present when inferences are drawn from a plot.
+
+Something about model diagnostics with a plot being the common thing to do..]
 

### 2022-02-02 Thesis/02-chap2.Rmd
figure changes

@@
 This framework redefines the test statistics, tests, null distribution, significance levels and $p$-value for visual discovery modelled on the confirmatory statistical testing.
-Figure \ref{fig:parallelism} outlines the parallelism between conventional tests and visual discovery.
+Figure \@ref(fig:fparallelism) outlines the parallelism between conventional tests and visual discovery.
+
+[ET: note use Rmd syntax for figures references! So it's easy to use it for HTML later if you want to do something like [Earo](https://github.com/earowang/thesis) and have PDF + HTML thesis. ]
 
@@
 For model diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the hypothesis.
-The same assumed model, such as classical linear regression model, can be rejected by both nonlinearity and heteroskedasticity with the residual plot as shown in Figure \ref{fig:nonlinearityheter}.
+The same assumed model, such as classical linear regression model, can be rejected by both nonlinearity and heteroskedasticity with the residual plot as shown in Figure \@ref(fig:nonlinearityheter).
 

### 2022-02-07 Thesis/01-chap1.Rmd
begin introduction

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message=FALSE, cache=TRUE)
+knitr::opts_chunk$set(echo=FALSE, message=FALSE, cache=TRUE)
 # Load any R packages you need here
-library(forecast)
 library(ggplot2)
@@
 # Introduction {#ch:intro}
+
+## Artificial intelligence and predictive modelling
+
+**Artificial intelligence** (AI) is the field of research concerned with understanding and building machines who can demonstrate intelligence.
+As discussed in @russell_artificial_2002, historically, there are disagreements among researchers on the definition of intelligence, which is caused by two critical questions:
+
+1. Should AI act and think humanly or rationally? 
+2. Without the thought process and reasoning, are behaviours sufficient to demonstrate intelligence?
+
+Based on the answer to the above questions, four major approaches to pursue AI have been established.
+They can be summarized into a two by two table as shown in figure \@ref(fig:twodimai), where the row is "Human" vs "Rational", and the column is "Behaviour" vs "Thought".
+Positioning at the top right cell, the **rational agent approach** aims to build agent that perform mathematically perfect acts such that the best expected outcome can always be achieved.
+In contrast, the **"laws of thought" approach** focus on understanding the logic behind the rationality.
+Closely related to **cognitive science**, the **cognitive modelling approach** attempts to express theories of human cognition as computer program to mimic the thought process of human.
+Lastly, the **Turing test approach** is built upon the famous **Turing test** proposed by @turing_computing_1950.
+The test can be roughly described as, whether a human can distinguish another human from a computer with written communications only.
+To pass the test, several capabilities of computer are required.
+This includes **natural language processing**, **knowledge representation**, **automated reasoning** and **machine learning**.
+Some researchers argued that some degree of physical simulation of a person is still necessary, one such example is the **total Turing test** proposed by @harnad_other_1991.
+It adds new requirements to the list, including **computer vision**, **speech recognition** and **robotics** [@russell_artificial_2002].
+
+Notably, all $6$ required capabilities have become major subfields of AI today.
+
+Among The most influential one may be the machine learning, which highly rely on the predictive modelling framework.
+
+Donoho (2017) in its summary of data science stated that the concept of this modelling culture could be traced back to an article written by Breiman (2001).
+In contrast to the generative modelling culture, which aims to develop stochastic models to make inferences about the data generating process, predictive modelling emphasizes the ability of the model to make accurate predictions.
+
+Behind the success of AI and computer vision, a great proportion of their applications highly rely on the predictive modelling framework.
+
+```{r twodimai, fig.cap="test"}
+ggplot() +
+  geom_hline(aes(yintercept = 1), col = "grey") +
+  geom_vline(aes(xintercept = 1), col = "grey") +
+  geom_text(aes(x = 0.5 , y = 1.5, label = "The Turing test approach")) +
+  geom_text(aes(x = 1.5 , y = 1.5, label = "The rational agent approach")) +
+  geom_text(aes(x = 0.5 , y = 0.5, label = "The cognitive modelling approach")) +
+  geom_text(aes(x = 1.5 , y = 0.5, label = 'The "laws of thought" approach')) +
+  theme_light() +
+  xlab("") +
+  ylab("") +
+  scale_x_continuous(breaks = c(0.5, 1.5), labels = c("Human", "Rational"), limits = c(0, 2)) +
+  scale_y_continuous(breaks = c(0.5, 1.5), labels = c("Thought", "Behaviour"), limits = c(0, 2)) +
+  theme(panel.grid = element_blank(), axis.ticks = element_blank())
+```
+
+The rational agent approach
+
+which impacts the version of AI pursued by researchers.
+
+due to the definition of intelligence,
+
+It has been pursued in many different ways by the researchers. The major separations between different approaches are the
+
+reasoning and thoughts, focus on the results or the action. Act and think humanly or rationally.
+
+Two dimensions
+
+machine think humanly, think rationally, acting humanly, acting rationally
+
+A rational agent is one that acts so as to achieve the best outcome or, when there is uncertainty, the best expected outcome.
+
+versions due to
+
+human performance
+
+The cognitive modeling approach
+
+The “laws of thought” approach
+
+The rational agent approach
+
+human vs. rational2 and thought vs. behavior
+
+Some have defined intelligence in terms of fidelity to human performance, while others prefer an abstract, formal definition of intelligence called rationality—loosely speaking, doing the “right thing.” The subject matter itself also varies: some consider intelligence to be a property of internal thought processes and reasoning, while others focus on intelligent behavior, an external characterization.
+
+As required by the Turing
+
+As defined by Turing test,
+
+The Turning test is well known
+
+The field of artificial intelligence, or AI, is concerned with not just understanding but also building intelligent entities—machines that can compute how to act effectively and safely in a wide variety of novel situations.
+
+## AI -> predictive modelling
+
+## Predictive modelling -> Model diagnostics
+
+## Limitation of MD/EDA (before 2009)
+
+## Visual inference
+
+## Limitation of visual inference
+
+## Automatic visual inference -> Computer vision

### 2022-02-07 Thesis/_book/thesis.tex
begin introduction

@@
 
+\hypertarget{artificial-intelligence-and-predictive-modelling}{%
+\section{Artificial intelligence and predictive modelling}\label{artificial-intelligence-and-predictive-modelling}}
+
+\textbf{Artificial intelligence} (AI) is the field of research concerned with understanding and building machines who can demonstrate intelligence.
+As discussed in \textcite{russell_artificial_2002}, historically, there are disagreements among researchers on the definition of intelligence, which is caused by two critical questions:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Should AI act and think humanly or rationally?
+\item
+  Without the thought process and reasoning, are behaviours sufficient to demonstrate intelligence?
+\end{enumerate}
+
+Based on the answer to the above questions, four major approaches to pursue AI have been established.
+They can be summarized into a two by two table as shown in figure \ref{fig:twodimai}, where the row is ``Human'' vs ``Rational'', and the column is ``Behaviour'' vs ``Thought''.
+Positioning at the top right cell, the \textbf{rational agent approach} aims to build agent that perform mathematically perfect acts such that the best expected outcome can always be achieved.
+In contrast, the \textbf{``laws of thought'' approach} focus on understanding the logic behind the rationality.
+Closely related to \textbf{cognitive science}, the \textbf{cognitive modelling approach} attempts to express theories of human cognition as computer program to mimic the thought process of human.
+Lastly, the \textbf{Turing test approach} is built upon the famous \textbf{Turing test} proposed by \textcite{turing_computing_1950}.
+The test can be roughly described as, whether a human can distinguish another human from a computer with written communications only.
+To pass the test, several capabilities of computer are required.
+This includes \textbf{natural language processing}, \textbf{knowledge representation}, \textbf{automated reasoning} and \textbf{machine learning}.
+Some researchers argued that some degree of physical simulation of a person is still necessary, one such example is the \textbf{total Turing test} proposed by \textcite{harnad_other_1991}.
+It adds new requirements to the list, including \textbf{computer vision}, \textbf{speech recognition} and \textbf{robotics} \autocite{russell_artificial_2002}.
+Notably, all \(6\) required capabilities have become major subfields of AI today.
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/twodimai-1.pdf}
+\caption{\label{fig:twodimai}test}
+\end{figure}
+
+The rational agent approach
+
+which impacts the version of AI pursued by researchers.
+
+due to the definition of intelligence,
+
+It has been pursued in many different ways by the researchers. The major separations between different approaches are the
+
+reasoning and thoughts, focus on the results or the action. Act and think humanly or rationally.
+
+Two dimensions
+
+machine think humanly, think rationally, acting humanly, acting rationally
+
+A rational agent is one that acts so as to achieve the best outcome or, when there is uncertainty, the best expected outcome.
+
+versions due to
+
+human performance
+
+The cognitive modeling approach
+
+The ``laws of thought'' approach
+
+The rational agent approach
+
+human vs.~rational2 and thought vs.~behavior
+
+Some have defined intelligence in terms of fidelity to human performance, while others prefer an abstract, formal definition of intelligence called rationality---loosely speaking, doing the ``right thing.'' The subject matter itself also varies: some consider intelligence to be a property of internal thought processes and reasoning, while others focus on intelligent behavior, an external characterization.
+
+As required by the Turing
+
+As defined by Turing test,
+
+The Turning test is well known
+
+The field of artificial intelligence, or AI, is concerned with not just understanding but also building intelligent entities---machines that can compute how to act effectively and safely in a wide variety of novel situations.
+
+\hypertarget{ai---predictive-modelling}{%
+\section{AI -\textgreater{} predictive modelling}\label{ai---predictive-modelling}}
+
+\hypertarget{predictive-modelling---model-diagnostics}{%
+\section{Predictive modelling -\textgreater{} Model diagnostics}\label{predictive-modelling---model-diagnostics}}
+
+\hypertarget{limitation-of-mdeda-before-2009}{%
+\section{Limitation of MD/EDA (before 2009)}\label{limitation-of-mdeda-before-2009}}
+
+\hypertarget{visual-inference}{%
+\section{Visual inference}\label{visual-inference}}
+
+\hypertarget{limitation-of-visual-inference}{%
+\section{Limitation of visual inference}\label{limitation-of-visual-inference}}
+
+\hypertarget{automatic-visual-inference---computer-vision}{%
+\section{Automatic visual inference -\textgreater{} Computer vision}\label{automatic-visual-inference---computer-vision}}
+
 \hypertarget{ch:paper1}{%
@@
 
-\hypertarget{visual-inference}{%
-\subsection{Visual Inference}\label{visual-inference}}
+\hypertarget{visual-inference-1}{%
+\subsection{Visual Inference}\label{visual-inference-1}}
 
@@
 In model diagnostics, when the null hypothesis is rejected, the reasons for rejecting the hypothesis will also be known.
-This is because observers not only can point out the fact that visual discoveries have been found, but also can describe the particular features they observed.
+This is because observers can not only point out the fact that visual discoveries have been found, but also describe the particular visual features they observed.
 Those features will correspond to the subset of the collection of visual test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) which resulted in rejection.
+This argument helps justifies the validity of visual inference.
 
@@
 
-\hypertarget{type-i-error}{%
-\subsubsection{Type-I error}\label{type-i-error}}
+With the validity of visual inference being justified, another aspect of hypothesis testing that needs to be addressed is the control of false positive rate or Type I error.
+Any visual statistic \(T^{(i)}(\boldsymbol{\mathrm{y}})\) needs to pair with a critical value \(c^{(i)}\) to form a hypothesis test.
+When a visual feature \(i\) is discovered by the observer from a plot, the corresponding visual statistic \(T^{(i)}(\boldsymbol{\mathrm{y}})\) may not be known as there is no general agreement on the measurement of the degree of presence of a visual feature.
+It is only the event that \(T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}\) is confirmed.
+Similarly, if any visual discovery is found by the observer, we say, there exists \(i \in I:~T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}\) \autocite{buja_statistical_2009}.
 
-\hypertarget{type-ii-error}{%
-\subsubsection{Type-II error}\label{type-ii-error}}
+Using the above definition, the family-wise Type I error can be controlled if one can provide the collection of critical values \(c^{(i)}~(i \in I)\) such that \(P(\mathrm{there~exists~} i \in I: T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}|\boldsymbol{\mathrm{y}}) \leq \alpha\), where \(\alpha\) is the significance level.
+However, since the quantity of \(T^{(i)}(\boldsymbol{\mathrm{y}})\) may not be known, such collection of critical values can not be provided.
+
+\textcite{buja_statistical_2009} proposed the lineup protocol as a visual test to calibrate the Type I error issue without the specification of \(c^{(i)}~(i \in I)\).
+It is inspired by the ``police lineup'' or ``identity parade'' which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people.
+The protocol consists of \(m\) randomly placed data plots, where \(1\) plot is the actual data plot, and \(m-1\) null plots are produced by plotting data simulate from the null distribution which is consistent with the null hypothesis.
+Then, an observer who have not seen the actual data plot will be asked to point out the most different plot from the lineup.
+
+Under the null hypothesis, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is \(1/m\) due to randomness.
+If we reject the null hypothesis as the observer correctly picks the actual data plot, then the Type I error of this test is \(1/m\).
+
+This provides us with an mechanism to control the Type I error, because \(m\) - the number of plots in a lineup can be chosen.
+Further, if we involve \(K\) independent observers in a visual test, and let \(X\) be a random variable denoting the number of observers correctly picking the actual data plot.
+Then, under the null hypothesis \(X \sim \mathrm{Binom}_{K,1/m}\), and therefore, the \(p\)-value of a lineup of size \(m\) evaluated by \(K\) observer is given as
+
+\[P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},\]
+where \(x\) is the realization of number of observers correctly picking the actual data plot \autocite{majumder_validation_2013}.
 

### 2022-02-07 Thesis/02-chap2.Rmd
begin introduction

--- 2022-02-02 Thesis/02-chap2.Rmd
@@
 
+[ET: Confirmatory data analysis is well characterised by a rigorous procedure to discern particular hypothesis.
+The most widespread form of confirmatory data analysis are the use of $p$-values in testing hypothesis in a frequentist framework.
+More specifically, hypothesis testing in a frequentist framework involve stating a well-defined hypotheses; summarising the evidence as a test statistic; and calculating the probability of observing a statistic as extreme as the observed test statistic under the null hypothesis.
+This rigour, however, is often not present when inferences are drawn from a plot.
+
+Something about model diagnostics with a plot being the common thing to do..]
+
 ### Visual Inference
@@
 
-#### Type I error
-
-#### Type II error
-
 ### Visual Inference Applied to Linear Regression

### 2022-02-07 Thesis/_book/thesis.tex
add introduction part

@@
 It adds new requirements to the list, including \textbf{computer vision}, \textbf{speech recognition} and \textbf{robotics} \autocite{russell_artificial_2002}.
-Notably, all \(6\) required capabilities have become major subfields of AI today.
+As of today, all \(7\) required capabilities have become major subfields of AI.
+
+Notably, machine learning, acts as the core of AI, has been developed rapidly in recent years thanks for the big-data era and the accessible and low-cost computational power \autocite{jordan_machine_2015}.
+It is a field of study addresses the issues of building computers that can adapt to new circumstances automatically through experience and data.
+Machine learning applications such as facial recognition, web recommendation system and machine translation have a significant impact on human's daily life.
+Behind the success of machine learning, a great propotion of its applications rely on the predictive modelling framework.
+Donoho (2017) in its summary of data science stated that the concept of this modelling culture could be traced back to an article written by Breiman (2001).
+In contrast to the generative modelling culture, which aims to develop stochastic models to make inferences about the data generating process, predictive modelling emphasizes the ability of the model to make accurate predictions.
 
@@
 
+{[}ET: suggestion: A model can be fitted to data with no guarantee of a meaningful interpretation.
+Model diagnostics play an important role in assessing the appropriateness of the model.
+The assessment can involve examining the goodness of fit, checking if there are potential violations of model assumptions and validating models with external information.{]}
+
 Model diagnostics is the part of data analysis, preceded by the fit of a model, whose primary objectives are to examine the goodness of fit and reveal potential violations of model assumptions.
-In these diagnostics, though numeric summaries are mostly available and some are even endorsed by finite or asymptotic properties, graphic representation of data is still preferred, or at least needed, due to its intuitiveness and the possibility to provide unexpected discoveries which may be abstract and
+
+{[}ET: model fit is more than just goodness of fit and checking model assumptions.
+It's too reductive to say this because you could have a model with good fit (according to some metric) and no model violation but if the way that the data were collected was flawed then scientific interpretation of the model is rendered useless.
+Diagnostics involve also the domain knowledge checks as well.{]}
+
+In these diagnostics, though numeric summaries are mostly available and some are even endorsed by finite or asymptotic properties, graphical representation of data is still preferred, or at least needed, due to its intuitiveness and the possibility to provide unexpected discoveries which may be abstract and
 unquantifiable.
+
+{[}ET: Generally, your sentences are too long.
+Break it up into smaller sentences.
+The generally idea is to make sentences that \emph{continue an idea from the preceding sentence} so it leads from one to the next naturally.
+E.g.
+Today is cloudy.
+Cloudy days are rare.
+Perhaps today will be different to other days!{]}
 
@@
 One such example is finding a separation between gene groups in a two-dimensional projection from a linear discriminant analysis where there is no difference in the expression levels between the gene groups \autocite{roy_chowdhury_using_2015}.
+
+{[}ET: Confirmatory data analysis is well characterised by a rigorous procedure to discern particular hypothesis.
+The most widespread form of confirmatory data analysis are the use of \(p\)-values in testing hypothesis in a frequentist framework.
+More specifically, hypothesis testing in a frequentist framework involve stating a well-defined hypotheses; summarising the evidence as a test statistic; and calculating the probability of observing a statistic as extreme as the observed test statistic under the null hypothesis.
+This rigour, however, is often not present when inferences are drawn from a plot.
+
+Something about model diagnostics with a plot being the common thing to do..{]}
 
@@
 This framework redefines the test statistics, tests, null distribution, significance levels and \(p\)-value for visual discovery modelled on the confirmatory statistical testing.
-Figure \ref{fig:parallelism} outlines the parallelism between conventional tests and visual discovery.
+Figure \ref{fig:fparallelism} outlines the parallelism between conventional tests and visual discovery.
+
+{[}ET: note use Rmd syntax for figures references!
+So it's easy to use it for HTML later if you want to do something like \href{https://github.com/earowang/thesis}{Earo} and have PDF + HTML thesis.
+{]}
 

### 2022-02-07 Thesis/01-chap1.Rmd
add introduction part

@@
 It adds new requirements to the list, including **computer vision**, **speech recognition** and **robotics** [@russell_artificial_2002].
+As of today, all $7$ required capabilities have become major subfields of AI.
 
-Notably, all $6$ required capabilities have become major subfields of AI today.
-
-Among The most influential one may be the machine learning, which highly rely on the predictive modelling framework.
-
+Notably, machine learning, acts as the core of AI, has been developed rapidly in recent years thanks for the big-data era and the accessible and low-cost computational power [@jordan_machine_2015].
+It is a field of study addresses the issues of building computers that can adapt to new circumstances automatically through experience and data.
+Machine learning applications such as facial recognition, web recommendation system and machine translation have a significant impact on human's daily life.
+Behind the success of machine learning, a great propotion of its applications rely on the predictive modelling framework.
 Donoho (2017) in its summary of data science stated that the concept of this modelling culture could be traced back to an article written by Breiman (2001).
 In contrast to the generative modelling culture, which aims to develop stochastic models to make inferences about the data generating process, predictive modelling emphasizes the ability of the model to make accurate predictions.
-
-Behind the success of AI and computer vision, a great proportion of their applications highly rely on the predictive modelling framework.
 

### 2022-02-09 Thesis/01-chap1.Rmd
comment on intro
refine intro

--- 2022-02-07 Thesis/01-chap1.Rmd
@@
 **Artificial intelligence** (AI) is the field of research concerned with understanding and building machines who can demonstrate intelligence.
-As discussed in @russell_artificial_2002, historically, there are disagreements among researchers on the definition of intelligence, which is caused by two critical questions:
+As discussed in @russell_artificial_2002, historically, there are disagreements among researchers about the definition of intelligence, which is caused by two critical questions:
 
@@
 Based on the answer to the above questions, four major approaches to pursue AI have been established.
-They can be summarized into a two by two table as shown in figure \@ref(fig:twodimai), where the row is "Human" vs "Rational", and the column is "Behaviour" vs "Thought".
+These approaches can be summarized into a two by two table as shown in figure \@ref(fig:twodimai), where the row is "Human" vs "Rational", and the column is "Behaviour" vs "Thought".
 Positioning at the top right cell, the **rational agent approach** aims to build agent that perform mathematically perfect acts such that the best expected outcome can always be achieved.
@@
 To pass the test, several capabilities of computer are required.
-This includes **natural language processing**, **knowledge representation**, **automated reasoning** and **machine learning**.
-Some researchers argued that some degree of physical simulation of a person is still necessary, one such example is the **total Turing test** proposed by @harnad_other_1991.
-It adds new requirements to the list, including **computer vision**, **speech recognition** and **robotics** [@russell_artificial_2002].
-
-Notably, all $6$ required capabilities have become major subfields of AI today.
-
-Among The most influential one may be the machine learning, which highly rely on the predictive modelling framework.
-
-Donoho (2017) in its summary of data science stated that the concept of this modelling culture could be traced back to an article written by Breiman (2001).
-In contrast to the generative modelling culture, which aims to develop stochastic models to make inferences about the data generating process, predictive modelling emphasizes the ability of the model to make accurate predictions.
-
-Behind the success of AI and computer vision, a great proportion of their applications highly rely on the predictive modelling framework.
+This includes **natural language processing** for communication with human, **knowledge representation** for encoding knowledge, **automated reasoning** for derivation of conclusions and **machine learning** for improving AI automatically through experience and data.
+Some researchers argued that written communication is insufficient to demonstrate intelligence, and some degree of physical simulation of a person is still necessary.
+One such example is the **total Turing test** proposed by @harnad_other_1991.
+It adds $3$ new requirements to the list, including **computer vision**, **speech recognition** and **robotics**, which are response for interactions with the physical world.
+Notably, all $7$ required capabilities have become major subfields of AI today.
+And their development has made AI one of the fastest-growing fields in the 21st century [@russell_artificial_2002].
 
@@
 
-The rational agent approach
+With the development of AI, mature AI technologies, such as facial recognition and web recommendation system, have profoundly affected the way modern society operates and citizen's daily life.
+This is largely as a consequence of the huge investment in AI industry by the financial market in recent years.
+Further, the increasingly cheap computing cost and the massive amount of accessible e-commerce data produced in the Internet age provide the possibility for applying data-intensive AI models, which enables AI performance to reach new heights in history [@jordan_machine_2015].
+Some AI systems have already been remarkably better than human in certain areas, e.g., game playing.
+AlphaGo and AlphaZero developed by the Google DeepMind team surpass all human Go players [@silver_general_2018].
 
-which impacts the version of AI pursued by researchers.
+Behind the success of AI, a great propotion of AI systems rely on the predictive modelling framework.
+@donoho_50_2017 in its summary of data science stated that the concept of this modelling culture could be traced back to an article written by Breiman (2001).
+In contrast to the generative modelling culture, which aims to develop stochastic models to make inferences about the data generating process, predictive modelling emphasizes the ability of the model to make accurate predictions.
+Most AI tasks are complex prediction problems where the data mechanism is mysterious, or at least, partly unknowable.
+@breiman_statistical_2001 suggests that generative models are obviously not applicable in these scenarios.
 
-due to the definition of intelligence,
+Model diagnostics is the part of data analysis, preceded by the fit of a model, whose primary objectives are to examine the goodness of fit and reveal potential violations of model assumptions.
 
-It has been pursued in many different ways by the researchers. The major separations between different approaches are the
+In these diagnostics, though numeric summaries are mostly available and some are even endorsed by finite or asymptotic properties, graphical representation of data is still preferred, or at least needed, due to its intuitiveness and the possibility to provide unexpected discoveries which may be abstract and
+unquantifiable.
 
-reasoning and thoughts, focus on the results or the action. Act and think humanly or rationally.
-
-Two dimensions
-
-machine think humanly, think rationally, acting humanly, acting rationally
-
-A rational agent is one that acts so as to achieve the best outcome or, when there is uncertainty, the best expected outcome.
-
-versions due to
-
-human performance
-
-The cognitive modeling approach
-
-The “laws of thought” approach
-
-The rational agent approach
-
-human vs. rational2 and thought vs. behavior
-
-Some have defined intelligence in terms of fidelity to human performance, while others prefer an abstract, formal definition of intelligence called rationality—loosely speaking, doing the “right thing.” The subject matter itself also varies: some consider intelligence to be a property of internal thought processes and reasoning, while others focus on intelligent behavior, an external characterization.
-
-As required by the Turing
-
-As defined by Turing test,
-
-The Turning test is well known
-
-The field of artificial intelligence, or AI, is concerned with not just understanding but also building intelligent entities—machines that can compute how to act effectively and safely in a wide variety of novel situations.
+However, unlike confirmatory data analysis built upon rigorous statistical procedures, e.g., hypothesis testing, visual diagnostics relies on graphical perception - human’s ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984], which is to some extent subjective.
+Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
+One such example is finding a separation between gene groups in a two-dimensional projection from a linear discriminant analysis where there is no difference in the expression levels between the gene groups [@roy_chowdhury_using_2015].
 

### 2022-02-09 Thesis/_book/thesis.tex
refine intro

@@
 \textbf{Artificial intelligence} (AI) is the field of research concerned with understanding and building machines who can demonstrate intelligence.
-As discussed in \textcite{russell_artificial_2002}, historically, there are disagreements among researchers on the definition of intelligence, which is caused by two critical questions:
+As discussed in \textcite{russell_artificial_2002}, historically, there are disagreements among researchers about the definition of intelligence, which is caused by two critical questions:
 
@@
 Based on the answer to the above questions, four major approaches to pursue AI have been established.
-They can be summarized into a two by two table as shown in figure \ref{fig:twodimai}, where the row is ``Human'' vs ``Rational'', and the column is ``Behaviour'' vs ``Thought''.
+These approaches can be summarized into a two by two table as shown in figure \ref{fig:twodimai}, where the row is ``Human'' vs ``Rational'', and the column is ``Behaviour'' vs ``Thought''.
 Positioning at the top right cell, the \textbf{rational agent approach} aims to build agent that perform mathematically perfect acts such that the best expected outcome can always be achieved.
@@
 To pass the test, several capabilities of computer are required.
-This includes \textbf{natural language processing}, \textbf{knowledge representation}, \textbf{automated reasoning} and \textbf{machine learning}.
-Some researchers argued that some degree of physical simulation of a person is still necessary, one such example is the \textbf{total Turing test} proposed by \textcite{harnad_other_1991}.
-It adds new requirements to the list, including \textbf{computer vision}, \textbf{speech recognition} and \textbf{robotics} \autocite{russell_artificial_2002}.
-As of today, all \(7\) required capabilities have become major subfields of AI.
-
-Notably, machine learning, acts as the core of AI, has been developed rapidly in recent years thanks for the big-data era and the accessible and low-cost computational power \autocite{jordan_machine_2015}.
-It is a field of study addresses the issues of building computers that can adapt to new circumstances automatically through experience and data.
-Machine learning applications such as facial recognition, web recommendation system and machine translation have a significant impact on human's daily life.
-Behind the success of machine learning, a great propotion of its applications rely on the predictive modelling framework.
-Donoho (2017) in its summary of data science stated that the concept of this modelling culture could be traced back to an article written by Breiman (2001).
-In contrast to the generative modelling culture, which aims to develop stochastic models to make inferences about the data generating process, predictive modelling emphasizes the ability of the model to make accurate predictions.
+This includes \textbf{natural language processing} for communication with human, \textbf{knowledge representation} for encoding knowledge, \textbf{automated reasoning} for derivation of conclusions and \textbf{machine learning} for improving AI automatically through experience and data.
+Some researchers argued that written communication is insufficient to demonstrate intelligence, and some degree of physical simulation of a person is still necessary.
+One such example is the \textbf{total Turing test} proposed by \textcite{harnad_other_1991}.
+It adds \(3\) new requirements to the list, including \textbf{computer vision}, \textbf{speech recognition} and \textbf{robotics}, which are response for interactions with the physical world.
+Notably, all \(7\) required capabilities have become major subfields of AI today.
+And their development has made AI one of the fastest-growing fields in the 21st century \autocite{russell_artificial_2002}.
 
@@
 
-The rational agent approach
+With the development of AI, mature AI technologies, such as facial recognition and web recommendation system, have profoundly affected the way modern society operates and citizen's daily life.
+This is largely as a consequence of the huge investment in AI industry by the financial market in recent years.
+Further, the increasingly cheap computing cost and the massive amount of accessible e-commerce data produced in the Internet age provide the possibility for applying data-intensive AI models, which enables AI performance to reach new heights in history \autocite{jordan_machine_2015}.
+Some AI systems have already been remarkably better than human in certain areas, e.g., game playing.
+AlphaGo and AlphaZero developed by the Google DeepMind team surpass all human Go players \autocite{silver_general_2018}.
 
-which impacts the version of AI pursued by researchers.
+Behind the success of AI, a great propotion of AI systems rely on the predictive modelling framework.
+\textcite{donoho_50_2017} in its summary of data science stated that the concept of this modelling culture could be traced back to an article written by Breiman (2001).
+In contrast to the generative modelling culture, which aims to develop stochastic models to make inferences about the data generating process, predictive modelling emphasizes the ability of the model to make accurate predictions.
+Most AI tasks are complex prediction problems where the data mechanism is mysterious, or at least, partly unknowable.
+\textcite{breiman_statistical_2001} suggests that generative models are obviously not applicable in these scenarios.
 
-due to the definition of intelligence,
+Model diagnostics is the part of data analysis, preceded by the fit of a model, whose primary objectives are to examine the goodness of fit and reveal potential violations of model assumptions.
 
-It has been pursued in many different ways by the researchers. The major separations between different approaches are the
+In these diagnostics, though numeric summaries are mostly available and some are even endorsed by finite or asymptotic properties, graphical representation of data is still preferred, or at least needed, due to its intuitiveness and the possibility to provide unexpected discoveries which may be abstract and
+unquantifiable.
 
-reasoning and thoughts, focus on the results or the action. Act and think humanly or rationally.
-
-Two dimensions
-
-machine think humanly, think rationally, acting humanly, acting rationally
-
-A rational agent is one that acts so as to achieve the best outcome or, when there is uncertainty, the best expected outcome.
-
-versions due to
-
-human performance
-
-The cognitive modeling approach
-
-The ``laws of thought'' approach
-
-The rational agent approach
-
-human vs.~rational2 and thought vs.~behavior
-
-Some have defined intelligence in terms of fidelity to human performance, while others prefer an abstract, formal definition of intelligence called rationality---loosely speaking, doing the ``right thing.'' The subject matter itself also varies: some consider intelligence to be a property of internal thought processes and reasoning, while others focus on intelligent behavior, an external characterization.
-
-As required by the Turing
-
-As defined by Turing test,
-
-The Turning test is well known
-
-The field of artificial intelligence, or AI, is concerned with not just understanding but also building intelligent entities---machines that can compute how to act effectively and safely in a wide variety of novel situations.
+However, unlike confirmatory data analysis built upon rigorous statistical procedures, e.g., hypothesis testing, visual diagnostics relies on graphical perception - human's ability to interpret and decode the information embedded in the graph \autocite{cleveland_graphical_1984}, which is to some extent subjective.
+Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
+One such example is finding a separation between gene groups in a two-dimensional projection from a linear discriminant analysis where there is no difference in the expression levels between the gene groups \autocite{roy_chowdhury_using_2015}.
 

### 2022-02-09 Thesis/01-chap1.Rmd

@@
 As discussed in @russell_artificial_2002, historically, there are disagreements among researchers about the definition of intelligence, which is caused by two critical questions:
+[ET: don't start with a definition as that makes the narrative dull.
+A lot of what you are writing definitions - which is relevant - but you want to interweave this as _narrative_.
+Why are you defining/describing these terms/disciplines?
+What's relevant and exciting?]
 

### 2022-02-10 Thesis/_book/thesis.tex
add more intro

@@
 
-\hypertarget{artificial-intelligence-and-predictive-modelling}{%
-\section{Artificial intelligence and predictive modelling}\label{artificial-intelligence-and-predictive-modelling}}
+\hypertarget{ai-four-approaches}{%
+\section{AI: Four Approaches}\label{ai-four-approaches}}
 
@@
 Based on the answer to the above questions, four major approaches to pursue AI have been established.
-These approaches can be summarized into a two by two table as shown in figure \ref{fig:twodimai}, where the row is ``Human'' vs ``Rational'', and the column is ``Behaviour'' vs ``Thought''.
+These approaches can be summarized into a two by two table as shown in figure \ref{fig:twodimai}, where the row is ``Human'' vs.~``Rational'', and the column is ``Behaviour'' vs.~``Thought''.
 Positioning at the top right cell, the \textbf{rational agent approach} aims to build agent that perform mathematically perfect acts such that the best expected outcome can always be achieved.
@@
 \includegraphics{thesis_files/figure-latex/twodimai-1.pdf}
-\caption{\label{fig:twodimai}test}
+\caption{\label{fig:twodimai}Four possible approaches to pursue AI based on the two dimensions in AI research - human vs.~rational and thought vs.~behavior \autocite{russell_artificial_2002}.}
 \end{figure}
@@
 
+\hypertarget{predictive-modelling-and-visual-diagnostics}{%
+\section{Predictive Modelling and Visual Diagnostics}\label{predictive-modelling-and-visual-diagnostics}}
+
 Behind the success of AI, a great propotion of AI systems rely on the predictive modelling framework.
@@
 Most AI tasks are complex prediction problems where the data mechanism is mysterious, or at least, partly unknowable.
-\textcite{breiman_statistical_2001} suggests that generative models are obviously not applicable in these scenarios.
+\textcite{breiman_statistical_2001} suggests that generative models are obviously not applicable in these scenarios, while the predictive modelling seeks only an accurate approximated function \(\boldsymbol{f}(\boldsymbol{x})\) to describe the relationship between the features \(\boldsymbol{x}\) and the responses \(\boldsymbol{y}\).
 
-Model diagnostics is the part of data analysis, preceded by the fit of a model, whose primary objectives are to examine the goodness of fit and reveal potential violations of model assumptions.
-
-In these diagnostics, though numeric summaries are mostly available and some are even endorsed by finite or asymptotic properties, graphical representation of data is still preferred, or at least needed, due to its intuitiveness and the possibility to provide unexpected discoveries which may be abstract and
-unquantifiable.
+Predictive models are primarily evaluated by predictive accuracy with the use of validation and test data, but in predictive model diagnostics, especially model testing and tuning, data plots play an irreplaceable role.
+In these diagnostics, though numeric summaries are mostly available and some are even endorsed by finite or asymptotic properties, graphical representation of data is still preferred, or at least needed by researchers, due to its intuitiveness and the possibility to provide unexpected discoveries which may be abstract and unquantifiable.
 
@@
 
-\hypertarget{ai---predictive-modelling}{%
-\section{AI -\textgreater{} predictive modelling}\label{ai---predictive-modelling}}
+\hypertarget{visual-inference}{%
+\section{Visual Inference}\label{visual-inference}}
 
-\hypertarget{predictive-modelling---model-diagnostics}{%
-\section{Predictive modelling -\textgreater{} Model diagnostics}\label{predictive-modelling---model-diagnostics}}
+Visual inference was first introduced by \textcite{buja_statistical_2009} as an inferential framework to extend confirmatory statistics to visual discoveries.
+This framework redefines the test statistics, tests, null distribution, significance levels and \(p\)-value for visual discovery modelled on the confirmatory statistical testing.
+Figure \ref{fig:parallelism} outlines the parallelism between conventional tests and visual discovery.
 
-\hypertarget{limitation-of-mdeda-before-2009}{%
-\section{Limitation of MD/EDA (before 2009)}\label{limitation-of-mdeda-before-2009}}
+\begin{figure}
+\centering
+\includegraphics[width=4.6875in,height=3.55208in]{figures/rsta2009012001.jpg}
+\caption{Parallelism between multiple quantitative testing and visual discovery \autocite{buja_statistical_2009}.
+Visible features in a plot are viewed as a collection of test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\), and any visual discoveries are treated as evidence against the null hypothesis.
+\label{fig:parallelism}}
+\end{figure}
 
-\hypertarget{visual-inference}{%
-\section{Visual inference}\label{visual-inference}}
+In visual inference, a visual discovery is defined as a rejection of a null hypothesis, and the same null hypothesis can be rejected by many different visual discoveries \autocite{buja_statistical_2009}.
+For model diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the null hypothesis.
+The same assumed model, such as classical linear regression model, can be rejected by many reasons with residual plot, including nonlinearity and heteroskedasticity as shown in figure \ref{fig:nonlinearityheter}.
 
-\hypertarget{limitation-of-visual-inference}{%
-\section{Limitation of visual inference}\label{limitation-of-visual-inference}}
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/nonlinearityheter-1.pdf}
+\caption{\label{fig:nonlinearityheter}Residuals vs.~fitted values plot for a classical linear regression model.
+The residuals are produced by fitting a two-predictor multiple linear regression model with data generated from a cubic linear model.
+From the residual plot, ``butterfly shape'' can be observed which generally would be interpretd as evidence of heteroskedasticity.
+Further, from the outline of the shape, nonlinear patterns exist.
+Both visual discoveries are evidence against the null hypothesis, though heteroskedasticity actually does not exist in the data generating process.}
+\end{figure}
+
+\hypertarget{pre-specification-of-visual-discoverable-features}{%
+\section{Pre-specification of Visual Discoverable Features}\label{pre-specification-of-visual-discoverable-features}}
+
+As discussed in \textcite{buja_statistical_2009}, in the practice of model diagnostics, the range of possible visual discoveries is not pre-specified.
+In other words, people do not explicitly specify which one or more visual features they are looking for before the read of the diagnostic plot.
+This is concerning since conventional hypothesis testing always requires the pre-specification of the parameter space \(\Theta\) of the parameter of interest \(\theta \in \Theta\) to form a valid inferential procedure.
+To address this issue, a collection of test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) is defined, where \(\boldsymbol{\mathrm{y}}\) is the data and \(\boldsymbol{I}\) is a set of all possible visual features.
+\textcite{buja_statistical_2009} described each of the test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})\) as a measurement of the degree of presence of a visual feature.
+Alternatively, \textcite{majumder_validation_2013} avoids the use of visual features and defined the visual statistics \(T(.)\) as a mapping from a dataset to a data plot.
+Both definitions of visual test statistics are valid, but in the rest of the paper the first definition will be used as it covers some details needed by the following discussion.
+
+The size of the collection \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) depends on the size of the set \(I\).
+Thus, if one can define \(I\) comprehensively, i.e, pre-specify all the visual discoverable features, the validity issue will be solved.
+Unfortunately, to our knowledge, there is no such a way to list all visual features.
+In linear regression diagnostics, possible visual features of a residual plot may be outliers, shapes and clusters.
+But this is an incomplete list which does not enumerate all the visual features.
+
+Similarly, \textcite{wilkinson_graph-theoretic_2005} proposed the work called graph theoretic scagnostics, which adopted the idea of ``scagnostics'' - scatter plot diagnostics from (can't find the 1984 citation).
+It includes 9 computable scagnostics measures defined on planar proximity graphs: ``Outlying'', ``Convex'', ``Skinny'', ``Stringy'', ``Straight'', ``Monotonic'', ``Skewed'', ``Clumpy'' and ``Striated'' which attempts to describe outliers, shape, density, trend and coherence of the data.
+This approach is inspiring but it still does not give the complete list of visual discoverable features.
+In fact, it is possible that such a list will never be complete as suggested in \textcite{buja_statistical_2009}.
+
+Thinking out of the box, \textcite{buja_statistical_2009} argued that there is actually no need for pre-specification of visual discoverable features.
+In model diagnostics, when the null hypothesis is rejected, the reasons for rejecting the hypothesis will also be known.
+This is because observers can not only point out the fact that visual discoveries have been found, but also describe the particular visual features they observed.
+Those features will correspond to the subset of the collection of visual test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) which resulted in rejection.
+This argument helps justifies the validity of visual inference.
+
+\hypertarget{lineup-protocol}{%
+\section{Lineup Protocol}\label{lineup-protocol}}
+
+With the validity of visual inference being justified, another aspect of hypothesis testing that needs to be addressed is the control of false positive rate or Type I error.
+Any visual statistic \(T^{(i)}(\boldsymbol{\mathrm{y}})\) needs to pair with a critical value \(c^{(i)}\) to form a hypothesis test.
+When a visual feature \(i\) is discovered by the observer from a plot, the corresponding visual statistic \(T^{(i)}(\boldsymbol{\mathrm{y}})\) may not be known as there is no general agreement on the measurement of the degree of presence of a visual feature.
+It is only the event that \(T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}\) is confirmed.
+Similarly, if any visual discovery is found by the observer, we say, there exists \(i \in I:~T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}\) \autocite{buja_statistical_2009}.
+
+Using the above definition, the family-wise Type I error can be controlled if one can provide the collection of critical values \(c^{(i)}~(i \in I)\) such that \(P(\mathrm{there~exists~} i \in I: T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}|\boldsymbol{\mathrm{y}}) \leq \alpha\), where \(\alpha\) is the significance level.
+However, since the quantity of \(T^{(i)}(\boldsymbol{\mathrm{y}})\) may not be known, such collection of critical values can not be provided.
+
+\textcite{buja_statistical_2009} proposed the lineup protocol as a visual test to calibrate the Type I error issue without the specification of \(c^{(i)}~(i \in I)\).
+It is inspired by the ``police lineup'' or ``identity parade'' which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people.
+The protocol consists of \(m\) randomly placed data plots, where \(1\) plot is the actual data plot, and \(m-1\) null plots are produced by plotting data simulate from the null distribution which is consistent with the null hypothesis.
+Then, an observer who have not seen the actual data plot will be asked to point out the most different plot from the lineup.
+
+Under the null hypothesis, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is \(1/m\) due to randomness.
+If we reject the null hypothesis as the observer correctly picks the actual data plot, then the Type I error of this test is \(1/m\).
+
+This provides us with an mechanism to control the Type I error, because \(m\) - the number of plots in a lineup can be chosen.
+A larger value of \(m\) will result in a smaller Type I error, but the limit to the value of \(m\) depends on the number of plots a human willing to view \autocite{buja_statistical_2009}.
+Typically, \(m\) will be set to \(20\) which is equivalent to set \(\alpha = 0.05\), a general choice of significance level for conventional testing among statisticians.
+
+Further, if we involve \(K\) independent observers in a visual test, and let \(X\) be a random variable denoting the number of observers correctly picking the actual data plot.
+Then, under the null hypothesis \(X \sim \mathrm{Binom}_{K,1/m}\), and therefore, the \(p\)-value of a lineup of size \(m\) evaluated by \(K\) observer is given as
+
+\[P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},\]
+where \(x\) is the realization of number of observers correctly picking the actual data plot \autocite{majumder_validation_2013}.
+
+The multiple individuals approach avoids the limit of \(m\), while provides visual tests with \(\alpha\) much smaller than \(0.05\).
+In fact, \(\alpha\) decreases exponentially as \(K\) increases.
+With just \(4\) individuals and \(20\) data plots in a lineup, the \(p\)-value could be as small as \(0.0001\).
+
+Compared to the conventional test, whose power only depends on the parameter of interest \(\theta\), several studies \autocites[see][]{hofmann_graphical_2012,majumder_validation_2013,majumder_human_2014,roy_chowdhury_using_2015,loy_variations_2016} have shown the power of the visual test is subject-specific.
+Thus, to be able to account for individual's ability, an individual is required to evaluate multiple lineups \autocite{majumder_validation_2013}.
+
+Assumes that individuals have the same ability and a lineup has been evaluated by multiple individuals, under the alternative hypothesis, the estimated power for a lineup can be expressed as \(\hat{p} = x/K\), the estimated probability of identifying the actual data plot from the lineup.
+If the individual skill needs to be taken into account, and \(L\) lienups have been evaluated by \(K\) individuals, \textcite{majumder_validation_2013} suggests that mixed effects logistic regression model can be fit as:
+
+\[g(p_{li}) = W_{li}\delta + Z_{li}\tau_{li},\]
+where \(g(.)\) is the logit link function \(g(p) = log(p) - log(1-p)\); \(0 \leq p \leq 1\).
+\(W_{li}\), \(1 \leq i \leq K\), \(1 \leq l \leq L\), is the covariate matrix including lineup-specific elements and demographic information of individuals, and \(\delta\) is a vector of parameters.
+\(Z\) is the random effects matrix, and \(\tau\) is a vector of variables follow \(N(\boldsymbol{0},\sigma_{\tau}\boldsymbol{I}_{KL\times KL})\).
+
+Then, the estimated power for lineup \(l\) and individual \(i\) can be calculated as \(\hat{p}_{li} = g^{-1}(W_{li}\hat{\delta} + Z_{li}\hat{\tau}_{li})\) \autocite{majumder_validation_2013}.
+
+\hypertarget{applications-of-lineup-protocol}{%
+\section{Applications of Lineup Protocol}\label{applications-of-lineup-protocol}}
+
+Lineup protocols are
+
+List applications, fields, significance here.
+
+\hypertarget{limitations-of-visual-tests}{%
+\section{Limitations of visual tests}\label{limitations-of-visual-tests}}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  infeasible in a large scale
+\item
+  unfriendly to vision-impaired people
+\item
+  high finical cost and human cost
+\item
+  time consuming
+\end{enumerate}
+
+similar to Handicraft in pre-industrial society
 
@@
 \section{Automatic visual inference -\textgreater{} Computer vision}\label{automatic-visual-inference---computer-vision}}
+
+relieve people's workload by automating repeating tasks, and provide standard result in a control environment
+
+the use of technology and machinery to enable mass evaluation of visual tests
+
+\hypertarget{discussion-of-potential-methods}{%
+\section{Discussion of potential methods}\label{discussion-of-potential-methods}}
+
+\hypertarget{different-approaches-of-ai}{%
+\subsection{different approaches of AI}\label{different-approaches-of-ai}}
+
+\begin{itemize}
+\tightlist
+\item
+  Not aim for understanding the thought process
+\item
+  Not aim for mimicking the human vision mechanism
+\item
+  May be able to define distance metrics to measure difference between data plots for making mathematically prefect decisions
+\item
+  May be able to use computer vision model to approximate how people evaluate lineups
+
+  \begin{itemize}
+  \tightlist
+  \item
+    Use human data to train model with human selection as target -\textgreater{} mimic the human behaviour
+  \item
+    Use simulated lineup to train model with actual data plot as target -\textgreater{} assume the actual data plot is the most different one
+  \item
+    Use simulated data plot to train model with null or not null as target -\textgreater{}
+  \end{itemize}
+\end{itemize}
 
@@
 \includegraphics[width=4.6875in,height=3.55208in]{figures/rsta2009012001.jpg}
-\caption{Parallelism between multiple quantitative testing and visual discovery \autocite{buja_statistical_2009}. \label{fig:parallelism}}
+\caption{Parallelism between multiple quantitative testing and visual discovery \autocite{buja_statistical_2009}.
+Visible features in a plot are viewed as a collection of test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\), and any visual discoveries are treated as evidence against the null hypothesis.
+\label{fig:parallelism}}
 \end{figure}
+
+Parallelism between multiple quantitative testing and visual discovery \autocite{buja_statistical_2009}.
+Visible features in a plot are viewed as a collection of test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\), and any visual discoveries are treated as evidence against the null hypothesis.
+Parallelism between multiple quantitative testing and visual discovery \autocite{buja_statistical_2009}.
 
@@
 
-\hypertarget{pre-specification-of-visual-discoverable-features}{%
-\subsection{Pre-specification of Visual Discoverable Features}\label{pre-specification-of-visual-discoverable-features}}
+\hypertarget{pre-specification-of-visual-discoverable-features-1}{%
+\subsection{Pre-specification of Visual Discoverable Features}\label{pre-specification-of-visual-discoverable-features-1}}
 
@@
 
-\begin{figure}
-\centering
-\includegraphics{thesis_files/figure-latex/nonlinearityheter-1.pdf}
-\caption{\label{fig:nonlinearityheter}test}
-\end{figure}
-
-\hypertarget{lineup-protocol}{%
-\subsection{Lineup Protocol}\label{lineup-protocol}}
+\hypertarget{lineup-protocol-1}{%
+\subsection{Lineup Protocol}\label{lineup-protocol-1}}
 

### 2022-02-10 Thesis/01-chap1.Rmd
add more intro

--- 2022-02-09 Thesis/01-chap1.Rmd
@@
 
-## Artificial intelligence and predictive modelling
+## AI: Four Approaches
 
@@
 As discussed in @russell_artificial_2002, historically, there are disagreements among researchers about the definition of intelligence, which is caused by two critical questions:
+
+<!-- [ET: don't start with a definition as that makes the narrative dull.
+A lot of what you are writing definitions - which is relevant - but you want to interweave this as _narrative_.
+Why are you defining/describing these terms/disciplines?
+What's relevant and exciting?] -->
 
@@
 Based on the answer to the above questions, four major approaches to pursue AI have been established.
-These approaches can be summarized into a two by two table as shown in figure \@ref(fig:twodimai), where the row is "Human" vs "Rational", and the column is "Behaviour" vs "Thought".
+These approaches can be summarized into a two by two table as shown in figure \@ref(fig:twodimai), where the row is "Human" vs.
+"Rational", and the column is "Behaviour" vs.
+"Thought".
 Positioning at the top right cell, the **rational agent approach** aims to build agent that perform mathematically perfect acts such that the best expected outcome can always be achieved.
@@
 
-```{r twodimai, fig.cap="test"}
+```{r twodimai, fig.cap="Four possible approaches to pursue AI based on the two dimensions in AI research - human vs. rational and thought vs. behavior [@russell_artificial_2002]."}
 ggplot() +
@@
 
+## Predictive Modelling and Visual Diagnostics
+
 Behind the success of AI, a great propotion of AI systems rely on the predictive modelling framework.
@@
 Most AI tasks are complex prediction problems where the data mechanism is mysterious, or at least, partly unknowable.
-@breiman_statistical_2001 suggests that generative models are obviously not applicable in these scenarios.
+@breiman_statistical_2001 suggests that generative models are obviously not applicable in these scenarios, while the predictive modelling seeks only an accurate approximated function $\boldsymbol{f}(\boldsymbol{x})$ to describe the relationship between the features $\boldsymbol{x}$ and the responses $\boldsymbol{y}$.
 
-Model diagnostics is the part of data analysis, preceded by the fit of a model, whose primary objectives are to examine the goodness of fit and reveal potential violations of model assumptions.
-
-In these diagnostics, though numeric summaries are mostly available and some are even endorsed by finite or asymptotic properties, graphical representation of data is still preferred, or at least needed, due to its intuitiveness and the possibility to provide unexpected discoveries which may be abstract and
-unquantifiable.
+Predictive models are primarily evaluated by predictive accuracy with the use of validation and test data, but in predictive model diagnostics, especially model testing and tuning, data plots play an irreplaceable role.
+In these diagnostics, though numeric summaries are mostly available and some are even endorsed by finite or asymptotic properties, graphical representation of data is still preferred, or at least needed by researchers, due to its intuitiveness and the possibility to provide unexpected discoveries which may be abstract and unquantifiable.
 
@@
 
-## AI -> predictive modelling
+## Visual Inference
 
-## Predictive modelling -> Model diagnostics
+Visual inference was first introduced by @buja_statistical_2009 as an inferential framework to extend confirmatory statistics to visual discoveries.
+This framework redefines the test statistics, tests, null distribution, significance levels and $p$-value for visual discovery modelled on the confirmatory statistical testing.
+Figure \@ref(fig:parallelism) outlines the parallelism between conventional tests and visual discovery.
 
-## Limitation of MD/EDA (before 2009)
+![(ref:parallelism) \label{fig:parallelism}](figures/rsta2009012001.jpg){width=450 height=341}
 
-## Visual inference
+(ref:parallelism) Parallelism between multiple quantitative testing and visual discovery [@buja_statistical_2009].
+Visible features in a plot are viewed as a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$, and any visual discoveries are treated as evidence against the null hypothesis.
 
-## Limitation of visual inference
+In visual inference, a visual discovery is defined as a rejection of a null hypothesis, and the same null hypothesis can be rejected by many different visual discoveries [@buja_statistical_2009].
+For model diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the null hypothesis.
+The same assumed model, such as classical linear regression model, can be rejected by many reasons with residual plot, including nonlinearity and heteroskedasticity as shown in figure \@ref(fig:nonlinearityheter).
+
+```{r nonlinearityheter, fig.cap='Residuals vs.
+fitted values plot for a classical linear regression model.
+The residuals are produced by fitting a two-predictor multiple linear regression model with data generated from a cubic linear model.
+From the residual plot, "butterfly shape" can be observed which generally would be interpretd as evidence of heteroskedasticity.
+Further, from the outline of the shape, nonlinear patterns exist.
+Both visual discoveries are evidence against the null hypothesis, though heteroskedasticity actually does not exist in the data generating process.'}
+library(tidyverse)
+
+set.seed(10086)
+cm <- visage::cubic_model(a = 0, b = 100, c = 1)
+
+cm$gen(1000, fit_model = TRUE) %>%
+cm$plot() +
+  theme_light()
+```
+
+## Pre-specification of Visual Discoverable Features
+
+As discussed in @buja_statistical_2009, in the practice of model diagnostics, the range of possible visual discoveries is not pre-specified.
+In other words, people do not explicitly specify which one or more visual features they are looking for before the read of the diagnostic plot.
+This is concerning since conventional hypothesis testing always requires the pre-specification of the parameter space $\Theta$ of the parameter of interest $\theta \in \Theta$ to form a valid inferential procedure.
+To address this issue, a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$ is defined, where $\boldsymbol{\mathrm{y}}$ is the data and $\boldsymbol{I}$ is a set of all possible visual features.
+@buja_statistical_2009 described each of the test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})$ as a measurement of the degree of presence of a visual feature.
+Alternatively, @majumder_validation_2013 avoids the use of visual features and defined the visual statistics $T(.)$ as a mapping from a dataset to a data plot.
+Both definitions of visual test statistics are valid, but in the rest of the paper the first definition will be used as it covers some details needed by the following discussion.
+
+The size of the collection $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$ depends on the size of the set $I$.
+Thus, if one can define $I$ comprehensively, i.e, pre-specify all the visual discoverable features, the validity issue will be solved.
+Unfortunately, to our knowledge, there is no such a way to list all visual features.
+In linear regression diagnostics, possible visual features of a residual plot may be outliers, shapes and clusters.
+But this is an incomplete list which does not enumerate all the visual features.
+
+Similarly, @wilkinson_graph-theoretic_2005 proposed the work called graph theoretic scagnostics, which adopted the idea of "scagnostics" - scatter plot diagnostics from (can't find the 1984 citation).
+It includes 9 computable scagnostics measures defined on planar proximity graphs: "Outlying", "Convex", "Skinny", "Stringy", "Straight", "Monotonic", "Skewed", "Clumpy" and "Striated" which attempts to describe outliers, shape, density, trend and coherence of the data.
+This approach is inspiring but it still does not give the complete list of visual discoverable features.
+In fact, it is possible that such a list will never be complete as suggested in @buja_statistical_2009.
+
+Thinking out of the box, @buja_statistical_2009 argued that there is actually no need for pre-specification of visual discoverable features.
+In model diagnostics, when the null hypothesis is rejected, the reasons for rejecting the hypothesis will also be known.
+This is because observers can not only point out the fact that visual discoveries have been found, but also describe the particular visual features they observed.
+Those features will correspond to the subset of the collection of visual test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$ which resulted in rejection.
+This argument helps justifies the validity of visual inference.
+
+## Lineup Protocol
+
+With the validity of visual inference being justified, another aspect of hypothesis testing that needs to be addressed is the control of false positive rate or Type I error.
+Any visual statistic $T^{(i)}(\boldsymbol{\mathrm{y}})$ needs to pair with a critical value $c^{(i)}$ to form a hypothesis test.
+When a visual feature $i$ is discovered by the observer from a plot, the corresponding visual statistic $T^{(i)}(\boldsymbol{\mathrm{y}})$ may not be known as there is no general agreement on the measurement of the degree of presence of a visual feature.
+It is only the event that $T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}$ is confirmed.
+Similarly, if any visual discovery is found by the observer, we say, there exists $i \in I:~T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}$ [@buja_statistical_2009].
+
+Using the above definition, the family-wise Type I error can be controlled if one can provide the collection of critical values $c^{(i)}~(i \in I)$ such that $P(\mathrm{there~exists~} i \in I: T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}|\boldsymbol{\mathrm{y}}) \leq \alpha$, where $\alpha$ is the significance level.
+However, since the quantity of $T^{(i)}(\boldsymbol{\mathrm{y}})$ may not be known, such collection of critical values can not be provided.
+
+@buja_statistical_2009 proposed the lineup protocol as a visual test to calibrate the Type I error issue without the specification of $c^{(i)}~(i \in I)$.
+It is inspired by the "police lineup" or "identity parade" which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people.
+The protocol consists of $m$ randomly placed data plots, where $1$ plot is the actual data plot, and $m-1$ null plots are produced by plotting data simulate from the null distribution which is consistent with the null hypothesis.
+Then, an observer who have not seen the actual data plot will be asked to point out the most different plot from the lineup.
+
+Under the null hypothesis, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is $1/m$ due to randomness.
+If we reject the null hypothesis as the observer correctly picks the actual data plot, then the Type I error of this test is $1/m$.
+
+This provides us with an mechanism to control the Type I error, because $m$ - the number of plots in a lineup can be chosen.
+A larger value of $m$ will result in a smaller Type I error, but the limit to the value of $m$ depends on the number of plots a human willing to view [@buja_statistical_2009].
+Typically, $m$ will be set to $20$ which is equivalent to set $\alpha = 0.05$, a general choice of significance level for conventional testing among statisticians.
+
+Further, if we involve $K$ independent observers in a visual test, and let $X$ be a random variable denoting the number of observers correctly picking the actual data plot.
+Then, under the null hypothesis $X \sim \mathrm{Binom}_{K,1/m}$, and therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as
+
+$$P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},$$
+where $x$ is the realization of number of observers correctly picking the actual data plot [@majumder_validation_2013].
+
+The multiple individuals approach avoids the limit of $m$, while provides visual tests with $\alpha$ much smaller than $0.05$.
+In fact, $\alpha$ decreases exponentially as $K$ increases.
+With just $4$ individuals and $20$ data plots in a lineup, the $p$-value could be as small as $0.0001$.
+
+Compared to the conventional test, whose power only depends on the parameter of interest $\theta$, several studies [see @hofmann_graphical_2012; @majumder_validation_2013; @majumder_human_2014; @roy_chowdhury_using_2015; @loy_variations_2016] have shown the power of the visual test is subject-specific.
+Thus, to be able to account for individual's ability, an individual is required to evaluate multiple lineups [@majumder_validation_2013].
+
+Assumes that individuals have the same ability and a lineup has been evaluated by multiple individuals, under the alternative hypothesis, the estimated power for a lineup can be expressed as $\hat{p} = x/K$, the estimated probability of identifying the actual data plot from the lineup.
+If the individual skill needs to be taken into account, and $L$ lienups have been evaluated by $K$ individuals, @majumder_validation_2013 suggests that mixed effects logistic regression model can be fit as:
+
+$$g(p_{li}) = W_{li}\delta + Z_{li}\tau_{li},$$
+where $g(.)$ is the logit link function $g(p) = log(p)  - log(1-p)$; $0 \leq p \leq 1$.
+$W_{li}$, $1 \leq i \leq K$, $1 \leq l \leq L$, is the covariate matrix including lineup-specific elements and demographic information of individuals, and $\delta$ is a vector of parameters.
+$Z$ is the random effects matrix, and $\tau$ is a vector of variables follow $N(\boldsymbol{0},\sigma_{\tau}\boldsymbol{I}_{KL\times KL})$.
+
+Then, the estimated power for lineup $l$ and individual $i$ can be calculated as $\hat{p}_{li} = g^{-1}(W_{li}\hat{\delta} + Z_{li}\hat{\tau}_{li})$ [@majumder_validation_2013].
+
+## Applications of Lineup Protocol
+
+Lineup protocols are
+
+List applications, fields, significance here.
+
+## Limitations of visual tests
+
+1. infeasible in a large scale
+2. unfriendly to vision-impaired people
+3. high finical cost and human cost
+4. time consuming
+
+similar to Handicraft in pre-industrial society
 
 ## Automatic visual inference -> Computer vision
+
+relieve people’s workload by automating repeating tasks, and provide standard result in a control environment
+
+the use of technology and machinery to enable mass evaluation of visual tests
+
+## Discussion of potential methods
+
+### different approaches of AI
+
+- Not aim for understanding the thought process
+- Not aim for mimicking the human vision mechanism
+- May be able to define distance metrics to measure difference between data plots for making mathematically prefect decisions
+- May be able to use computer vision model to approximate how people evaluate lineups
+  - Use human data to train model with human selection as target -> mimic the human behaviour
+  - Use simulated lineup to train model with actual data plot as target -> assume the actual data plot is the most different one
+  - Use simulated data plot to train model with null or not null as target ->

### 2022-02-10 Thesis/01-chap1.Rmd
give potential methods in intro

@@
 
-The multiple individuals approach avoids the limit of $m$, while provides visual tests with $\alpha$ much smaller than $0.05$.
-In fact, $\alpha$ decreases exponentially as $K$ increases.
+The multiple individuals approach avoids the limit of $m$, while provides visual tests with $p$-value much smaller than $0.05$.
+In fact, the lower bound of $p$-value decreases exponentially as $K$ increases.
 With just $4$ individuals and $20$ data plots in a lineup, the $p$-value could be as small as $0.0001$.
@@
 - May be able to use computer vision model to approximate how people evaluate lineups
+
+computer vision model:
+
 - Use human data to train model with human selection as target -> mimic the human behaviour
 - Use simulated lineup to train model with actual data plot as target -> assume the actual data plot is the most different one
-  - Use simulated data plot to train model with null or not null as target ->
+- Use simulated data plot to train model with null or not null as target -> limited to the alternatives given to the model, and Type I error can not be controlled
+- Use simulated data plot to train model, then let the model evaluate each plot of a lineup -> leads to multiple selections in a lineup, and the model does not compare different plots in a lineup
+  - Use a very large $m$ to ensure the null hypothesis can be rejected by a classifier slightly better than a random selector. The problem becomes how large the $m$ should be
+  - Use multiple models as multiple individuals to evaluate lineup -> Which models should be used? How many models should be used?
+- Let the model select multiple plots from a simulated lineup while comparing different plots in a lineup -> how to build such a system? 
+  - The input is $m$ plots, what is the output? The probability of being the most different one? 
+  - Then how many of plots should be selected? Top 5? Or cumulative probability greater than a threshold? 
+  - How to select the threshold?

### 2022-02-10 Thesis/_book/thesis.tex
give potential methods in intro

@@
 
-The multiple individuals approach avoids the limit of \(m\), while provides visual tests with \(\alpha\) much smaller than \(0.05\).
-In fact, \(\alpha\) decreases exponentially as \(K\) increases.
+The multiple individuals approach avoids the limit of \(m\), while provides visual tests with \(p\)-value much smaller than \(0.05\).
+In fact, the lower bound of \(p\)-value decreases exponentially as \(K\) increases.
 With just \(4\) individuals and \(20\) data plots in a lineup, the \(p\)-value could be as small as \(0.0001\).
@@
   May be able to use computer vision model to approximate how people evaluate lineups
+\end{itemize}
+
+computer vision model:
 
@@
 \item
-    Use simulated data plot to train model with null or not null as target -\textgreater{}
+  Use simulated data plot to train model with null or not null as target -\textgreater{} limited to the alternatives given to the model, and Type I error can not be controlled
+\item
+  Use simulated data plot to train model, then let the model evaluate each plot of a lineup -\textgreater{} leads to multiple selections in a lineup, and the model does not compare different plots in a lineup
+
+  \begin{itemize}
+  \tightlist
+  \item
+    Use a very large \(m\) to ensure the null hypothesis can be rejected by a classifier slightly better than a random selector. The problem becomes how large the \(m\) should be
+  \item
+    Use multiple models as multiple individuals to evaluate lineup -\textgreater{} Which models should be used? How many models should be used?
+  \end{itemize}
+\item
+  Let the model select multiple plots from a simulated lineup while comparing different plots in a lineup -\textgreater{} how to build such a system?
+
+  \begin{itemize}
+  \tightlist
+  \item
+    The input is \(m\) plots, what is the output? The probability of being the most different one?
+  \item
+    Then how many of plots should be selected? Top 5? Or cumulative probability greater than a threshold?
+  \item
+    How to select the threshold?
   \end{itemize}

### 2022-02-17 Thesis/01-chap1.Rmd
update

@@
   - The input is $m$ plots, what is the output? The probability of being the most different one? 
-  - Then how many of plots should be selected? Top 5? Or cumulative probability greater than a threshold? 
+  - Then how many plots should be selected? Top 5? Or cumulative probability greater than a threshold? 
   - How to select the threshold?
+
+# Human Subject Exeriments (toc)
+
+What needs to be discussed?
+
+1. online recruitment platform (Prolific)
+  - pre-screen
+  - approve and reject
+  - redirect
+  - payment
+2. study web site
+  - PythonAnywhere
+  - Flask, Python
+  - jsPsych
+  - survey
+    - age
+    - education
+    - preferred pronoun
+    - previous experience
+  - training
+  - 20 lineups
+    - setup depends on the study
+3. Cubic model
+  - data generating process
+  - null model
+  - potential shapes
+  - effect size
+  - conventional test
+4. Heteroskedasticity model
+  - data generating process
+  - null model
+  - potential shapes
+  - effect size
+  - conventional test
+4. Number of participants needed by the study
+3. Study 1
+  - logistic regression
+    - easy
+    - medium
+    - hard
+  - number of participants
+  - time
+  - countries
+4. Study 2
+  - logistic regression
+    - easy
+    - medium
+    - hard
+  - number of participants
+  - time
+  - countries
+
+# Human Subject Exeriments
+
+To collect data of human performance on reading residual plot of linear regression model with nonlinearity and heteroskedasticity defects, two human experiments were conducted.
+Participants of both experiments were recruited using an online platform called Prolific [@prolific_prolific_2022].
+
+Prolific provides an international participant pool with the option to apply flexible pre-screening filters.
+In this study, we recruited participants who is fluent in English with at least $10$ previous submissions and $98$% approval rate in other Prolific studies for quality control.
+Further, balance sample across gender was imposed to prevent gender bias.
+
+In Prolific, researchers can either approve or reject submissions based on the quality of the responses.
+If a submission is approved by the researcher, the participant will be paid a certain amount of money per hour of time spent on the experiment.
+To assess the quality of the responses, two attention checks were given to each participant during the experiment, where at least one of them was required to pass for the approval of submission.
+
+Throughout the experiments, participants were requested to complete a short survey and evaluate 20 lineups on a website in an hour.
+Each lineup each consists of 19 null residual plots and 1 actual residual plot.
+Among the 20 lineups, there are 2 extremely east lineups used as attention checks which everyone should get correct.
+The short survey was intended to collect information about participant that might affect their ability in reading data plot including age, highest education, preferred pronoun and previous experience in similar study.
+For the evaluation of the lineup, participant first needed to select one or multiple most different data plot(s) from the lineup by clicking the corresponding image.
+Then, the reason for choosing the plots needed to be provided by selecting one of the given options - "outlier(s)", "cluster(s)", "shape" and "other".
+Explanations about these options were provided in the training page and the mouseover text.
+Table \@ref(tab:lineupreason) gives the detailed explanations about these reasons.
+If the option "other" was selected, text input for the specified reason would be collected.
+Lastly, the degree of difference between their chosen data plot(s) and other data plots needed to be selected among $5$ levels - "not at all", "slightly", "moderately", "very" and "extremely".
+
+```{r}
+data.frame(c("outlier(s)", "cluster(s)", "shape", "other"), 
+           c("In a data plot, an outlier is a point that differs significantly from other points.", 
+             "In a data plot, a cluster is a group of points positioned closely together. \nAnd usually, there will be gaps between different clusters.",
+             "It could be any common shapes we would see in real life, like triangle shape, U-shape, butterfly shape, etc.",
+             "Patricipants needs to give their reasons in the text input.")) %>%
+knitr::kable(col.names = c("Reasons", "Explanations"), 
+             booktabs = TRUE, 
+             caption  = 'Explanations about reasons for choosing data plots from a lineup', 
+             label = "lineupreason") %>%
+  kableExtra::kable_styling(latex_options = c("scale_down"))
+```
+
+The study website was powered by Flask [@grinberg_flask_2018], a web framework written in Python 3 [@van_rossum_python_2009], hosted on PythonAnywhere [@pythonanywhere_pythonanywhere_2022], a web hosting service provider.
+The website is built using a JavaScript [@flanagan_javascript_2006] library jsPsych [@de_leeuw_jspsych_2015] which is specialized in creating online behavioral experiments.
+One of the reasons we chose to use this library was it had a modularized but highly customizable template which could record participant's response time automatically.
+This is essential for us to confirm the quality of the data by checking exceptionally fast or slow responses.
+The first two pages of the website are the explanatory statement and the consent form.
+Participants needed to read the documentation and agree with the terms to advance to the short survey.
+With the completion of the survey, the next page is the training page, which contained instructions on how to evaluate the lineup, give the reason and choose the confidence level.
+Followed by the training page, there are 20 lineups each on a single page.
+At the end of the experiment, the participants would be redirected back to Prolific and waited for researchers' responses.
+
+Next we will discuss the simulation setup for this study.
+The experiment data was simulated by the use of the programming language `R` [@r_core_team_r_2021].
+For the ease of reproducibility, functions to build models, simulate data from models, produce lineup with data, allocate stimulis for subjects and evaluate subject responses from this study are bundled in the package `visage` with a unique object-oriented programming system built upon the environment feature of `R`.
+In the description of the simulation, corresponding functionalities of the package will be introduced.
+
+Two models were used in both experiments, which were linear models with some degree of violations of classical assumptions. Residuals vs. fitted values plots were used to diagnose the issues.
+
+## Cubic Model
+
+The first model was a cubic linear model with 2 regressors, which can be expressed by: $$\boldsymbol{Y}= 1 + (2-c)\boldsymbol{X} + c\boldsymbol{Z} + a[(2-c)\boldsymbol{X}]^2+a(c\boldsymbol{Z})^2+b[(2-c)\boldsymbol{X}]^3+b(c\boldsymbol{Z})^3+\boldsymbol{\varepsilon},$$ where $c \in (0,2)$, $a \in (-3,3)$, $b \in (-3,3)$, $\boldsymbol{\varepsilon}\overset{iid}{\sim} N(\boldsymbol{0},\sigma^2\boldsymbol{I})$, $\boldsymbol{X}$ and $\boldsymbol{Z}$ are $n\times1$ matrices.
+
+This defines a cubic relationship between $\boldsymbol{Y}$, $\boldsymbol{X}$ and $\boldsymbol{Z}$.
+Meanwhile, to create nonlinearity defect, the null model fitted by the OLS is: $$\boldsymbol{Y}=\boldsymbol{b}_0+\boldsymbol{b}_1\boldsymbol{X}+\boldsymbol{b}_2\boldsymbol{Z}+\boldsymbol{e}.$$
+
+Clearly, there will be omitted-variable bias since the null model leaves out the quadric and cubic terms.
+
+### Distribution of the Residuals
+
+Let $\boldsymbol{X}_a=[\boldsymbol{1},\boldsymbol{X},\boldsymbol{Z}]$ denotes the set of regressors in matrix form.
+Using the Frisch–Waugh–Lovell theorem, the residual obtained by the null model is $\boldsymbol{e}=\boldsymbol{R}_a\boldsymbol{Y}=\boldsymbol{R}_a(\boldsymbol{X}_a\boldsymbol{\beta}_a+\boldsymbol{X}_b\boldsymbol{\beta}_b+\boldsymbol{\varepsilon}),$ where $\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$, $\boldsymbol{\beta}_a=(1,2-c,c)'$, $\boldsymbol{X}_b=[\boldsymbol{X}^2,\boldsymbol{Z}^2,\boldsymbol{X}^3,\boldsymbol{Z}^3]$ and $\boldsymbol{\beta}_b=(a(2-c)^2,ac^2,b(2-c)^3,bc^3)'$.
+
+Because $\boldsymbol{R}_a\boldsymbol{X}_a=\boldsymbol{0}$, we have $\boldsymbol{e}=\boldsymbol{R}_a(\boldsymbol{X}_b\boldsymbol{\beta}_b+\boldsymbol{\varepsilon}).$ Since $\boldsymbol{\varepsilon} \sim N(0,\sigma^2\boldsymbol{I})$, it follows that $\boldsymbol{e} \sim N(\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b, \sigma^2\boldsymbol{R}_a)$.
+
+### Shape of the Residuals
+
+The previous result shows that the expectation of the residuals is clearly a function of $\boldsymbol{X}$ and $\boldsymbol{Z}$.
+Hence, we would expect to see some patterns in the residual plot.
+Given the expectation of the residuals, we could plot the expected values of residuals against the observed values.
+Figure \@ref(fig:cubic1), \@ref(fig:cubic2) and \@ref(fig:cubic3) illustrate the shape of residuals under different parameter settings.
+
+The residuals used in these three figures are simulated from the cubic models built using the `cubic_model()` function from the package `visage`.
+`cubic_model()` is a cubic model class constructor, which takes arguments `a`, `b`, `c`, `sigma`, `x` and `z`, where the first four are numeric values defined above, and `x` and `z` are random variable instances created by the random variable abstract base class constructor `rand_var()`.
+
+Here, we set $\boldsymbol{X}$ and $\boldsymbol{Z}$ to be random uniform variables ranged from $-1$ to $1$.
+This can be done by using the random uniform variable class constructor `rand_uniform()` inherited from the random variable abstract base class.
+It only takes two arguments which are the lower bound and the upper bound of the support.
+
+```{r echo = TRUE}
+library(visage)
+mod <- cubic_model(a = -1, b = -1, c = 1, sigma = 0.5, 
+                   x = rand_uniform(-1, 1), z = rand_uniform(-1, 1))
+```
+
+An instance of cubic model class contains methods of simulating data and making residual plot.
+Method `mod$gen()` returns a data frame containing realizations of $x$, $z$, $y$ and $\varepsilon$ simulated from the model.
+The number of realizations depends on the integer argument `n`.
+In addition, if argument `fit_model = TRUE`, a null model will be fitted using the simulated data and residuals and fitted values will be included in the returned data frame.
+
+```{r echo = TRUE}
+mod$gen(n = 5, fit_model = TRUE)
+```
+
+Method `mod$plot()` produce a ggplot [@wickham_ggplot2_2011] object.
+It takes a data frame containing columns `.resid` and `.fitted` as input, along with a character argument `type` indicating the type of the data plot, and other aesthetic arguments such as `size` and `alpha` to control the appearance of the plot.
+
+```{r test3, echo = TRUE}
+mod$plot(mod$gen(n = 100, fit_model = TRUE), type = "resid", size = 1)
+```
+
+The cubic model class also provides method to compute the expected values of residuals.
+Method `mod$E()` takes a data frame with columns `x` and `z` as input, and returns a vector of expected values of residuals.
+
+```{r echo = TRUE}
+mod$E(mod$gen(n = 5))
+```
+
+From the figure \@ref(fig:cubic1), it can be observed that with fixed $\sigma$ and $c$, $a$ and $b$ are controlling the 2D projection of a hypersurface, and seemingly performing some rotations along different axes.
+Figure \@ref(fig:cubic2) shows that with fixed $a$, $b$ and $\sigma$, $c$ is controlling the contribution of $\boldsymbol{X}$ and $\boldsymbol{Z}$ to $\boldsymbol{Y}$.
+If we move $c$ toward $0$ or $2$, one variable will donminate another, which will mitigate some joint effects and resemble a typical cubic function.
+In figure \@ref(fig:cubic3), $a$, $b$ and $c$ is held constant, and $\sigma$ is controlling the noises around the expected values.
+As $\sigma$ decreases, the underlying shape shows up.
+
+```{r cubic1, fig.cap = "cubic 1", fig.height=6.3}
+library(visage)
+library(tidyverse)
+
+make_plot <- function(a, b, c, e_sigma) {
+  mod <- cubic_model(a, b, c, e_sigma)
+  dat <- mod$gen(1000, fit_model = TRUE)
+  dat$exp <- mod$E(dat)
+  mod$plot(dat, alpha = 0.3, size = 0.5, add_zero_line = FALSE, theme = theme_light(base_size = 5)) +
+    geom_point(aes(.fitted, exp, col = "red"), alpha = 0.6, size = 0.1) +
+    xlab("fitted values") +
+    ylab("residuals") +
+    ggtitle(bquote(a:.(a)~','~b:.(b)~','~c:.(c)~','~sigma:.(e_sigma))) +
+    theme(legend.position = "none")
+}
+
+expand.grid(c=seq(1),
+            a=seq(-1,1,0.5),
+            b=seq(-1,1,0.5),
+            e_sigma=c(0.5)) %>%
+  arrange(a,b,c,e_sigma) %>%
+  rowwise() %>%
+  mutate(plots = list(make_plot(a,b,c,e_sigma))) -> result
+
+result$plots %>% patchwork::wrap_plots(ncol=5)
+```
+
+```{r cubic2, fig.cap = "cubic 2", fig.height=7.875}
+expand.grid(c=seq(0.5,1.5,0.25),
+            a=seq(-1,1,1),
+            b=seq(-1,1,1),
+            e_sigma=c(0.5)) %>%
+  arrange(a,b,c,desc(e_sigma)) %>%
+  rowwise() %>%
+  mutate(plots = list(make_plot(a,b,c,e_sigma))) -> result
+
+result$plots %>% patchwork::wrap_plots(ncol=5)
+```
+
+```{r cubic3, fig.cap = "cubic 3", fig.height=7.875}
+expand.grid(c=c(1),
+            a=seq(-1,1,1),
+            b=seq(-1,1,1),
+            e_sigma=c(0.25,0.5,1,2)) %>%
+  arrange(a,b,c,desc(e_sigma)) %>%
+  rowwise() %>%
+  mutate(plots = list(make_plot(a,b,c,e_sigma))) -> result
+
+result$plots %>% patchwork::wrap_plots(ncol=4)
+```
+
+### Effect Size
+
+Since we know that under the null hypothesis, the residual $\boldsymbol{e}\sim N(\boldsymbol{0},\sigma^2\boldsymbol{R}_a)$.
+Thus, the difference between the expected values $\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b$ and $\boldsymbol{0}$ represents the direct impact of the parameters $a$ and $b$ on the residuals.
+It is expected that the larger the magnitude of the expected value relative to the variance and covariance, the easier the human to spot the patterns in the residual plot.
+
+To obtain a measure of the impact of $a$ and $b$ on the residuals, we need to address several properties of the residuals.
+First, because $\boldsymbol{R}_a$ is not an identity matrix, the residuals will have different variance and will be correlated with each other.
+This can be fixed by standardizing the residuals by their variance-covariance matrix.
+Second, the difference between $\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b$ and $\boldsymbol{0}$ could be negative, which is not ideal for comparison.
+Thus, the magnitude needs to be squared.
+Third, the measure needs to be a scalar.
+We could apply a weighted average operator $\boldsymbol{W}$ on the transformed expected residuals to obtain a single numeric value.
+For simplicity, we set $\boldsymbol{W}=n^{-1}\boldsymbol{1}$.
+This gives the effect size: $$ES=n^{-1}||\sigma^{-1}\boldsymbol{R}_a^{-\frac{1}{2}}\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b||^2=n^{-1}\sigma^{-2}||\boldsymbol{R}_a^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2\approx n^{-1}\sigma^{-2}||diag(\boldsymbol{R}_a)^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2.$$
+
+The interpretation of this effect size is the impact of parameter $a$ and $b$ on the squared deviation of the standardized expected residual per observation.
+It is not directly related to the shape, or the pattern observed from the residual plot, but it is a reasonable approximation of the degree of the deviation from the null residuals under our cubic model setting.
+
+```{r }
+mod <- cubic_model(-1, -1, 1, 0.5)
+dat <- mod$gen(500, fit_model = TRUE)
+mod$plot(dat, theme = theme_light(), size = 1) +
+  ggtitle(glue::glue("Effect size: {round(mod$effect_size(dat), 2)}")) -> p1
+
+mod <- cubic_model(-2, -1, 1, 0.5)
+dat <- mod$gen(500, fit_model = TRUE)
+mod$plot(dat, theme = theme_light(), size = 1) +
+  ggtitle(glue::glue("Effect size: {round(mod$effect_size(dat), 2)}")) -> p2
+
+mod <- cubic_model(-5, -5, 1, 0.5)
+dat <- mod$gen(500, fit_model = TRUE)
+mod$plot(dat, theme = theme_light(), size = 1) +
+  ggtitle(glue::glue("Effect size: {round(mod$effect_size(dat), 2)}")) -> p3
+
+mod <- cubic_model(-5, -5, 1, 0.1)
+dat <- mod$gen(500, fit_model = TRUE)
+mod$plot(dat, theme = theme_light(), size = 1) +
+  ggtitle(glue::glue("Effect size: {round(mod$effect_size(dat), 2)}")) -> p4
+
+list(p1, p2, p3, p4) %>% patchwork::wrap_plots(ncol=2)
+```
+
+###
+
+## Heteroskedasticity Model
+
+\clearpage

### 2022-02-17 Thesis/_book/thesis.tex
update

@@
   \item
-    Then how many of plots should be selected? Top 5? Or cumulative probability greater than a threshold?
+    Then how many plots should be selected? Top 5? Or cumulative probability greater than a threshold?
   \item
@@
 \end{itemize}
+
+\hypertarget{human-subject-exeriments-toc}{%
+\chapter{Human Subject Exeriments (toc)}\label{human-subject-exeriments-toc}}
+
+What needs to be discussed?
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  online recruitment platform (Prolific)
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  pre-screen
+\item
+  approve and reject
+\item
+  redirect
+\item
+  payment
+\end{itemize}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
+\tightlist
+\item
+  study web site
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  PythonAnywhere
+\item
+  Flask, Python
+\item
+  jsPsych
+\item
+  survey
+
+  \begin{itemize}
+  \tightlist
+  \item
+    age
+  \item
+    education
+  \item
+    preferred pronoun
+  \item
+    previous experience
+  \end{itemize}
+\item
+  training
+\item
+  20 lineups
+
+  \begin{itemize}
+  \tightlist
+  \item
+    setup depends on the study
+  \end{itemize}
+\end{itemize}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{2}
+\tightlist
+\item
+  Cubic model
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  data generating process
+\item
+  null model
+\item
+  potential shapes
+\item
+  effect size
+\item
+  conventional test
+\end{itemize}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{3}
+\tightlist
+\item
+  Heteroskedasticity model
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  data generating process
+\item
+  null model
+\item
+  potential shapes
+\item
+  effect size
+\item
+  conventional test
+\end{itemize}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{3}
+\tightlist
+\item
+  Number of participants needed by the study
+\item
+  Study 1
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  logistic regression
+
+  \begin{itemize}
+  \tightlist
+  \item
+    easy
+  \item
+    medium
+  \item
+    hard
+  \end{itemize}
+\item
+  number of participants
+\item
+  time
+\item
+  countries
+\end{itemize}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{3}
+\tightlist
+\item
+  Study 2
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  logistic regression
+
+  \begin{itemize}
+  \tightlist
+  \item
+    easy
+  \item
+    medium
+  \item
+    hard
+  \end{itemize}
+\item
+  number of participants
+\item
+  time
+\item
+  countries
+\end{itemize}
+
+\hypertarget{human-subject-exeriments}{%
+\chapter{Human Subject Exeriments}\label{human-subject-exeriments}}
+
+To collect data of human performance on reading residual plot of linear regression model with nonlinearity and heteroskedasticity defects, two human experiments were conducted.
+Participants of both experiments were recruited using an online platform called Prolific \autocite{prolific_prolific_2022}.
+
+Prolific provides an international participant pool with the option to apply flexible pre-screening filters.
+In this study, we recruited participants who is fluent in English with at least \(10\) previous submissions and \(98\)\% approval rate in other Prolific studies for quality control.
+Further, balance sample across gender was imposed to prevent gender bias.
+
+In Prolific, researchers can either approve or reject submissions based on the quality of the responses.
+If a submission is approved by the researcher, the participant will be paid a certain amount of money per hour of time spent on the experiment.
+To assess the quality of the responses, two attention checks were given to each participant during the experiment, where at least one of them was required to pass for the approval of submission.
+
+Throughout the experiments, participants were requested to complete a short survey and evaluate 20 lineups on a website in an hour.
+Each lineup each consists of 19 null residual plots and 1 actual residual plot.
+Among the 20 lineups, there are 2 extremely east lineups used as attention checks which everyone should get correct.
+The short survey was intended to collect information about participant that might affect their ability in reading data plot including age, highest education, preferred pronoun and previous experience in similar study.
+For the evaluation of the lineup, participant first needed to select one or multiple most different data plot(s) from the lineup by clicking the corresponding image.
+Then, the reason for choosing the plots needed to be provided by selecting one of the given options - ``outlier(s)'', ``cluster(s)'', ``shape'' and ``other''.
+Explanations about these options were provided in the training page and the mouseover text.
+Table \ref{tab:lineupreason} gives the detailed explanations about these reasons.
+If the option ``other'' was selected, text input for the specified reason would be collected.
+Lastly, the degree of difference between their chosen data plot(s) and other data plots needed to be selected among \(5\) levels - ``not at all'', ``slightly'', ``moderately'', ``very'' and ``extremely''.
+
+\begin{table}
+
+\caption{\label{tab:lineupreason}Explanations about reasons for choosing data plots from a lineup}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{ll}
+\toprule
+Reasons & Explanations\\
+\midrule
+outlier(s) & In a data plot, an outlier is a point that differs significantly from other points.\\
+cluster(s) & In a data plot, a cluster is a group of points positioned closely together. 
+And usually, there will be gaps between different clusters.\\
+shape & It could be any common shapes we would see in real life, like triangle shape, U-shape, butterfly shape, etc.\\
+other & Patricipants needs to give their reasons in the text input.\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+The study website was powered by Flask \autocite{grinberg_flask_2018}, a web framework written in Python 3 \autocite{van_rossum_python_2009}, hosted on PythonAnywhere \autocite{pythonanywhere_pythonanywhere_2022}, a web hosting service provider.
+The website is built using a JavaScript \autocite{flanagan_javascript_2006} library jsPsych \autocite{de_leeuw_jspsych_2015} which is specialized in creating online behavioral experiments.
+One of the reasons we chose to use this library was it had a modularized but highly customizable template which could record participant's response time automatically.
+This is essential for us to confirm the quality of the data by checking exceptionally fast or slow responses.
+The first two pages of the website are the explanatory statement and the consent form.
+Participants needed to read the documentation and agree with the terms to advance to the short survey.
+With the completion of the survey, the next page is the training page, which contained instructions on how to evaluate the lineup, give the reason and choose the confidence level.
+Followed by the training page, there are 20 lineups each on a single page.
+At the end of the experiment, the participants would be redirected back to Prolific and waited for researchers' responses.
+
+Next we will discuss the simulation setup for this study.
+The experiment data was simulated by the use of the programming language \texttt{R} \autocite{r_core_team_r_2021}.
+For the ease of reproducibility, functions to build models, simulate data from models, produce lineup with data, allocate stimulis for subjects and evaluate subject responses from this study are bundled in the package \texttt{visage} with a unique object-oriented programming system built upon the environment feature of \texttt{R}.
+In the description of the simulation, corresponding functionalities of the package will be introduced.
+
+Two models were used in both experiments, which were linear models with some degree of violations of classical assumptions. Residuals vs.~fitted values plots were used to diagnose the issues.
+
+\hypertarget{cubic-model}{%
+\section{Cubic Model}\label{cubic-model}}
+
+The first model was a cubic linear model with 2 regressors, which can be expressed by: \[\boldsymbol{Y}= 1 + (2-c)\boldsymbol{X} + c\boldsymbol{Z} + a[(2-c)\boldsymbol{X}]^2+a(c\boldsymbol{Z})^2+b[(2-c)\boldsymbol{X}]^3+b(c\boldsymbol{Z})^3+\boldsymbol{\varepsilon},\] where \(c \in (0,2)\), \(a \in (-3,3)\), \(b \in (-3,3)\), \(\boldsymbol{\varepsilon}\overset{iid}{\sim} N(\boldsymbol{0},\sigma^2\boldsymbol{I})\), \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) are \(n\times1\) matrices.
+
+This defines a cubic relationship between \(\boldsymbol{Y}\), \(\boldsymbol{X}\) and \(\boldsymbol{Z}\).
+Meanwhile, to create nonlinearity defect, the null model fitted by the OLS is: \[\boldsymbol{Y}=\boldsymbol{b}_0+\boldsymbol{b}_1\boldsymbol{X}+\boldsymbol{b}_2\boldsymbol{Z}+\boldsymbol{e}.\]
+
+Clearly, there will be omitted-variable bias since the null model leaves out the quadric and cubic terms.
+
+\hypertarget{distribution-of-the-residuals}{%
+\subsection{Distribution of the Residuals}\label{distribution-of-the-residuals}}
+
+Let \(\boldsymbol{X}_a=[\boldsymbol{1},\boldsymbol{X},\boldsymbol{Z}]\) denotes the set of regressors in matrix form.
+Using the Frisch--Waugh--Lovell theorem, the residual obtained by the null model is \(\boldsymbol{e}=\boldsymbol{R}_a\boldsymbol{Y}=\boldsymbol{R}_a(\boldsymbol{X}_a\boldsymbol{\beta}_a+\boldsymbol{X}_b\boldsymbol{\beta}_b+\boldsymbol{\varepsilon}),\) where \(\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'\), \(\boldsymbol{\beta}_a=(1,2-c,c)'\), \(\boldsymbol{X}_b=[\boldsymbol{X}^2,\boldsymbol{Z}^2,\boldsymbol{X}^3,\boldsymbol{Z}^3]\) and \(\boldsymbol{\beta}_b=(a(2-c)^2,ac^2,b(2-c)^3,bc^3)'\).
+
+Because \(\boldsymbol{R}_a\boldsymbol{X}_a=\boldsymbol{0}\), we have \(\boldsymbol{e}=\boldsymbol{R}_a(\boldsymbol{X}_b\boldsymbol{\beta}_b+\boldsymbol{\varepsilon}).\) Since \(\boldsymbol{\varepsilon} \sim N(0,\sigma^2\boldsymbol{I})\), it follows that \(\boldsymbol{e} \sim N(\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b, \sigma^2\boldsymbol{R}_a)\).
+
+\hypertarget{shape-of-the-residuals}{%
+\subsection{Shape of the Residuals}\label{shape-of-the-residuals}}
+
+The previous result shows that the expectation of the residuals is clearly a function of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\).
+Hence, we would expect to see some patterns in the residual plot.
+Given the expectation of the residuals, we could plot the expected values of residuals against the observed values.
+Figure \ref{fig:cubic1}, \ref{fig:cubic2} and \ref{fig:cubic3} illustrate the shape of residuals under different parameter settings.
+
+The residuals used in these three figures are simulated from the cubic models built using the \texttt{cubic\_model()} function from the package \texttt{visage}.
+\texttt{cubic\_model()} is a cubic model class constructor, which takes arguments \texttt{a}, \texttt{b}, \texttt{c}, \texttt{sigma}, \texttt{x} and \texttt{z}, where the first four are numeric values defined above, and \texttt{x} and \texttt{z} are random variable instances created by the random variable abstract base class constructor \texttt{rand\_var()}.
+
+Here, we set \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) to be random uniform variables ranged from \(-1\) to \(1\).
+This can be done by using the random uniform variable class constructor \texttt{rand\_uniform()} inherited from the random variable abstract base class.
+It only takes two arguments which are the lower bound and the upper bound of the support.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{library}\NormalTok{(visage)}
+\NormalTok{mod }\OtherTok{\textless{}{-}} \FunctionTok{cubic\_model}\NormalTok{(}\AttributeTok{a =} \SpecialCharTok{{-}}\DecValTok{1}\NormalTok{, }\AttributeTok{b =} \SpecialCharTok{{-}}\DecValTok{1}\NormalTok{, }\AttributeTok{c =} \DecValTok{1}\NormalTok{, }\AttributeTok{sigma =} \FloatTok{0.5}\NormalTok{, }
+                   \AttributeTok{x =} \FunctionTok{rand\_uniform}\NormalTok{(}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{), }\AttributeTok{z =} \FunctionTok{rand\_uniform}\NormalTok{(}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+An instance of cubic model class contains methods of simulating data and making residual plot.
+Method \texttt{mod\$gen()} returns a data frame containing realizations of \(x\), \(z\), \(y\) and \(\varepsilon\) simulated from the model.
+The number of realizations depends on the integer argument \texttt{n}.
+In addition, if argument \texttt{fit\_model\ =\ TRUE}, a null model will be fitted using the simulated data and residuals and fitted values will be included in the returned data frame.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{mod}\SpecialCharTok{$}\FunctionTok{gen}\NormalTok{(}\AttributeTok{n =} \DecValTok{5}\NormalTok{, }\AttributeTok{fit\_model =} \ConstantTok{TRUE}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+##             y          x          z          e      .resid     .fitted
+## 1  0.72328137  0.8783282 0.17687898  0.1539500  0.37649863  0.34678274
+## 2 -0.01317346  0.9593166 0.30200008 -0.3526064 -0.22899266  0.21581920
+## 3  0.10405716  0.7677855 0.72090641 -0.4481692  0.36279717 -0.25874001
+## 4 -0.53007353  0.9238035 0.55802451 -0.8849474 -0.46221585 -0.06785768
+## 5  0.33947802 -0.7784446 0.02191688  0.2307539 -0.04808729  0.38756531
+\end{verbatim}
+
+Method \texttt{mod\$plot()} produce a ggplot \autocite{wickham_ggplot2_2011} object.
+It takes a data frame containing columns \texttt{.resid} and \texttt{.fitted} as input, along with a character argument \texttt{type} indicating the type of the data plot, and other aesthetic arguments such as \texttt{size} and \texttt{alpha} to control the appearance of the plot.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{mod}\SpecialCharTok{$}\FunctionTok{plot}\NormalTok{(mod}\SpecialCharTok{$}\FunctionTok{gen}\NormalTok{(}\AttributeTok{n =} \DecValTok{100}\NormalTok{, }\AttributeTok{fit\_model =} \ConstantTok{TRUE}\NormalTok{), }\AttributeTok{type =} \StringTok{"resid"}\NormalTok{, }\AttributeTok{size =} \DecValTok{1}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{thesis_files/figure-latex/test3-1.pdf}
+
+The cubic model class also provides method to compute the expected values of residuals.
+Method \texttt{mod\$E()} takes a data frame with columns \texttt{x} and \texttt{z} as input, and returns a vector of expected values of residuals.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{mod}\SpecialCharTok{$}\FunctionTok{E}\NormalTok{(mod}\SpecialCharTok{$}\FunctionTok{gen}\NormalTok{(}\AttributeTok{n =} \DecValTok{5}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## [1]  0.05836465 -0.03566983 -0.09457480 -0.01126447  0.08314445
+\end{verbatim}
+
+From the figure \ref{fig:cubic1}, it can be observed that with fixed \(\sigma\) and \(c\), \(a\) and \(b\) are controlling the 2D projection of a hypersurface, and seemingly performing some rotations along different axes.
+Figure \ref{fig:cubic2} shows that with fixed \(a\), \(b\) and \(\sigma\), \(c\) is controlling the contribution of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) to \(\boldsymbol{Y}\).
+If we move \(c\) toward \(0\) or \(2\), one variable will donminate another, which will mitigate some joint effects and resemble a typical cubic function.
+In figure \ref{fig:cubic3}, \(a\), \(b\) and \(c\) is held constant, and \(\sigma\) is controlling the noises around the expected values.
+As \(\sigma\) decreases, the underlying shape shows up.
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/cubic1-1.pdf}
+\caption{\label{fig:cubic1}cubic 1}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/cubic2-1.pdf}
+\caption{\label{fig:cubic2}cubic 2}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/cubic3-1.pdf}
+\caption{\label{fig:cubic3}cubic 3}
+\end{figure}
+
+\hypertarget{effect-size}{%
+\subsection{Effect Size}\label{effect-size}}
+
+Since we know that under the null hypothesis, the residual \(\boldsymbol{e}\sim N(\boldsymbol{0},\sigma^2\boldsymbol{R}_a)\).
+Thus, the difference between the expected values \(\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b\) and \(\boldsymbol{0}\) represents the direct impact of the parameters \(a\) and \(b\) on the residuals.
+It is expected that the larger the magnitude of the expected value relative to the variance and covariance, the easier the human to spot the patterns in the residual plot.
+
+To obtain a measure of the impact of \(a\) and \(b\) on the residuals, we need to address several properties of the residuals.
+First, because \(\boldsymbol{R}_a\) is not an identity matrix, the residuals will have different variance and will be correlated with each other.
+This can be fixed by standardizing the residuals by their variance-covariance matrix.
+Second, the difference between \(\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b\) and \(\boldsymbol{0}\) could be negative, which is not ideal for comparison.
+Thus, the magnitude needs to be squared.
+Third, the measure needs to be a scalar.
+We could apply a weighted average operator \(\boldsymbol{W}\) on the transformed expected residuals to obtain a single numeric value.
+For simplicity, we set \(\boldsymbol{W}=n^{-1}\boldsymbol{1}\).
+This gives the effect size: \[ES=n^{-1}||\sigma^{-1}\boldsymbol{R}_a^{-\frac{1}{2}}\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b||^2=n^{-1}\sigma^{-2}||\boldsymbol{R}_a^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2\approx n^{-1}\sigma^{-2}||diag(\boldsymbol{R}_a)^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2.\]
+
+The interpretation of this effect size is the impact of parameter \(a\) and \(b\) on the squared deviation of the standardized expected residual per observation.
+It is not directly related to the shape, or the pattern observed from the residual plot, but it is a reasonable approximation of the degree of the deviation from the null residuals under our cubic model setting.
+
+\includegraphics{thesis_files/figure-latex/unnamed-chunk-5-1.pdf}
+
+\hypertarget{section}{%
+\subsection{}\label{section}}
+
+\hypertarget{heteroskedasticity-model}{%
+\section{Heteroskedasticity Model}\label{heteroskedasticity-model}}
+
+\clearpage
 

### 2022-02-21 Thesis/01-chap1.Rmd
update exp design

@@
 
+**null data generate mechanism**
+
 ## Applications of Lineup Protocol
@@
 Prolific provides an international participant pool with the option to apply flexible pre-screening filters.
-In this study, we recruited participants who is fluent in English with at least $10$ previous submissions and $98$% approval rate in other Prolific studies for quality control.
+In this study, we recruited participants who was fluent in English with at least $10$ previous submissions and $98$% approval rate in other Prolific studies for quality control.
 Further, balance sample across gender was imposed to prevent gender bias.
@@
 Throughout the experiments, participants were requested to complete a short survey and evaluate 20 lineups on a website in an hour.
-Each lineup each consists of 19 null residual plots and 1 actual residual plot.
-Among the 20 lineups, there are 2 extremely east lineups used as attention checks which everyone should get correct.
-The short survey was intended to collect information about participant that might affect their ability in reading data plot including age, highest education, preferred pronoun and previous experience in similar study.
+Each lineup consists of one actual residual plot and 19 null residual plots produced by plotting null residuals simulated from the residual rotation distributions.
+Among the 20 lineups, there are two extremely easy lineups used as attention checks which everyone should get correct.
+The short survey was intended to collect information about participant that might affect their ability in reading data plot including age, highest level of education, preferred pronoun and previous experience in similar study.
 For the evaluation of the lineup, participant first needed to select one or multiple most different data plot(s) from the lineup by clicking the corresponding image.
@@
 Lastly, the degree of difference between their chosen data plot(s) and other data plots needed to be selected among $5$ levels - "not at all", "slightly", "moderately", "very" and "extremely".
+Notably, if participants could not tell the difference between the data plots, there was an option to skip the evaluation of the lineup.
+However, to prevent participants abusing this option, warnings were given at the beginning of the study that skipping too many lineups might lead to rejection of submission since it demonstrated clear low-effort throughout the experiment.
 
@@
 The study website was powered by Flask [@grinberg_flask_2018], a web framework written in Python 3 [@van_rossum_python_2009], hosted on PythonAnywhere [@pythonanywhere_pythonanywhere_2022], a web hosting service provider.
-The website is built using a JavaScript [@flanagan_javascript_2006] library jsPsych [@de_leeuw_jspsych_2015] which is specialized in creating online behavioral experiments.
+Due to the limit of the storage on PythonAnywhere, lineup images needed for the website were hosted separately on Github Pages [@gtihub_github_2022-1], a static site hosting service provided by Github [@gtihub_github_2022], stored in private Github repositories.
+The uniform resource locator (URL) to lineup images were set to be unique random strings such that images can not be accessed by general approaches without knowing the correct URL.
+This avoided participants seeing the lineup beforehand.
+The front-end of the website was built using a JavaScript [@flanagan_javascript_2006] library jsPsych [@de_leeuw_jspsych_2015] which is specialized in creating online behavioral experiments.
 One of the reasons we chose to use this library was it had a modularized but highly customizable template which could record participant's response time automatically.
 This is essential for us to confirm the quality of the data by checking exceptionally fast or slow responses.
+Figure \@ref(fig:DiagramExperiment) summarizes the experiment setup.
+
+![Diagram of online experiment setup.
+The server-side of the study webiste used Flask as backend hosted on PythonAnywhere.
+And the client-side used jsPsych to run experiment.
+\label{fig:DiagramExperiment}](figures/experiment_tech.pdf){width=450 height=450}
+
 The first two pages of the website are the explanatory statement and the consent form.
-Participants needed to read the documentation and agree with the terms to advance to the short survey.
+Participants needed to read the documents and agree with the terms to advance to the short survey.
 With the completion of the survey, the next page is the training page, which contained instructions on how to evaluate the lineup, give the reason and choose the confidence level.
 Followed by the training page, there are 20 lineups each on a single page.
+Figure \@ref(fig:WebsiteLayout) illustrates the layout of the website.
 At the end of the experiment, the participants would be redirected back to Prolific and waited for researchers' responses.
+
+![Layout of the study website.
+Participants needed to choose the most different plots on the right and select their reasons and confidence levels on the left.
+\label{fig:WebsiteLayout}](figures/website.png){width=550 height=299}
 
@@
 
-Two models were used in both experiments, which were linear models with some degree of violations of classical assumptions. Residuals vs. fitted values plots were used to diagnose the issues.
+Two models were used in both experiments, which were linear models with some degree of violations of classical assumptions. Residual vs. fitted values plot were used to present the model defects.
 
@@
 
-```{r echo = TRUE}
+```{r echo = TRUE, cache = FALSE}
 library(visage)
-mod <- cubic_model(a = -1, b = -1, c = 1, sigma = 0.5, 
+mod <- cubic_model(a = -3, b = -3, c = 1, sigma = 0.5, 
                    x = rand_uniform(-1, 1), z = rand_uniform(-1, 1))
@@
 
-```{r echo = TRUE}
+```{r echo = TRUE, cache = FALSE}
 mod$gen(n = 5, fit_model = TRUE)
@@
 
-```{r test3, echo = TRUE}
+```{r test3, echo = TRUE, cache = FALSE}
 mod$plot(mod$gen(n = 100, fit_model = TRUE), type = "resid", size = 1)
+```
+
+Lineup can be produced by using the methods `mod$gen_lineup()` and `mod$plot_lineup()`.
+Method `mod$gen_lineup` takes the number of realizations `n` and the number of plots in a lineup `k` as inputs.
+And the method `mod$plot_lineup()` has the same user interface as `mod$plot()`.
+
+```{r echo = TRUE, cache = FALSE}
+mod$plot_lineup(mod$gen_lineup(n = 50, k = 20), type = "resid", size = 1)
 ```
@@
 
-```{r echo = TRUE}
+```{r echo = TRUE, cache = FALSE}
 mod$E(mod$gen(n = 5))
@@
 
-To obtain a measure of the impact of $a$ and $b$ on the residuals, we need to address several properties of the residuals.
-First, because $\boldsymbol{R}_a$ is not an identity matrix, the residuals will have different variance and will be correlated with each other.
+To obtain a measure of the impact of $a$ and $b$ on the residuals adjusted for variance and covariance, we need to address several properties of the residuals.
+First, the variance of the residuals $\sigma^2\boldsymbol{R}_a$ is not an identity matrix.
 This can be fixed by standardizing the residuals by their variance-covariance matrix.
@@
 For simplicity, we set $\boldsymbol{W}=n^{-1}\boldsymbol{1}$.
-This gives the effect size: $$ES=n^{-1}||\sigma^{-1}\boldsymbol{R}_a^{-\frac{1}{2}}\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b||^2=n^{-1}\sigma^{-2}||\boldsymbol{R}_a^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2\approx n^{-1}\sigma^{-2}||diag(\boldsymbol{R}_a)^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2.$$
+Considering the high time complexity of computing the square root of $\boldsymbol{R}_a$, off-diagonal elements of $\boldsymbol{R}_a$ are set to be zeros.
+This gives the effect size: $$ES=n^{-1}||\sigma^{-1}\boldsymbol{R}_a^{-\frac{1}{2}}\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b||^2=n^{-1}\sigma^{-2}||\boldsymbol{R}_a^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2\approx n^{-1}\sigma^{-2}||diag(\boldsymbol{R}_a)^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2,$$ where $diag(\boldsymbol{R}_a)$ is the diagonal matrix constructed from the diagnoal element of $\boldsymbol{R}_a$.
 
 The interpretation of this effect size is the impact of parameter $a$ and $b$ on the squared deviation of the standardized expected residual per observation.
-It is not directly related to the shape, or the pattern observed from the residual plot, but it is a reasonable approximation of the degree of the deviation from the null residuals under our cubic model setting.
+It is not directly related to the shape, or the pattern human observed from the residual plot, but it is a reasonable approximation of the degree of the visual deviation from the null residuals under our cubic model setting.
+Figure \@ref(fig:effectsize) shows four residual plots with different effect sizes.
+As effect sizes increases, the strength of the signal become stronger.
 
-```{r }
+```{r effectsize, fig.cap = "Cubic model residual plots under different effect sizes."}
 mod <- cubic_model(-1, -1, 1, 0.5)
 dat <- mod$gen(500, fit_model = TRUE)
-mod$plot(dat, theme = theme_light(), size = 1) +
+mod$plot(dat, theme = theme_light(base_size = 5), size = 1) +
   ggtitle(glue::glue("Effect size: {round(mod$effect_size(dat), 2)}")) -> p1
@@
 dat <- mod$gen(500, fit_model = TRUE)
-mod$plot(dat, theme = theme_light(), size = 1) +
+mod$plot(dat, theme = theme_light(base_size = 5), size = 1) +
   ggtitle(glue::glue("Effect size: {round(mod$effect_size(dat), 2)}")) -> p2
@@
 dat <- mod$gen(500, fit_model = TRUE)
-mod$plot(dat, theme = theme_light(), size = 1) +
+mod$plot(dat, theme = theme_light(base_size = 5), size = 1) +
   ggtitle(glue::glue("Effect size: {round(mod$effect_size(dat), 2)}")) -> p3
 
-mod <- cubic_model(-5, -5, 1, 0.1)
+mod <- cubic_model(-20, -20, 1, 0.5)
 dat <- mod$gen(500, fit_model = TRUE)
-mod$plot(dat, theme = theme_light(), size = 1) +
+mod$plot(dat, theme = theme_light(base_size = 5), size = 1) +
   ggtitle(glue::glue("Effect size: {round(mod$effect_size(dat), 2)}")) -> p4

### 2022-02-21 Thesis/_book/thesis.tex
update exp design

@@
 
+\textbf{null data generate mechanism}
+
 \hypertarget{applications-of-lineup-protocol}{%
@@
 Prolific provides an international participant pool with the option to apply flexible pre-screening filters.
-In this study, we recruited participants who is fluent in English with at least \(10\) previous submissions and \(98\)\% approval rate in other Prolific studies for quality control.
+In this study, we recruited participants who was fluent in English with at least \(10\) previous submissions and \(98\)\% approval rate in other Prolific studies for quality control.
 Further, balance sample across gender was imposed to prevent gender bias.
@@
 Lastly, the degree of difference between their chosen data plot(s) and other data plots needed to be selected among \(5\) levels - ``not at all'', ``slightly'', ``moderately'', ``very'' and ``extremely''.
+Notably, if participants could not tell the difference between the data plots, there was an option to skip the evaluation of the lineup.
+However, to prevent participants abusing this option, warnings were given at the beginning of the study that skipping too many lineups might lead to rejection of submission since it demonstrated clear low-effort throughout the experiment.
 
@@
 The study website was powered by Flask \autocite{grinberg_flask_2018}, a web framework written in Python 3 \autocite{van_rossum_python_2009}, hosted on PythonAnywhere \autocite{pythonanywhere_pythonanywhere_2022}, a web hosting service provider.
-The website is built using a JavaScript \autocite{flanagan_javascript_2006} library jsPsych \autocite{de_leeuw_jspsych_2015} which is specialized in creating online behavioral experiments.
+Due to the limit of the storage on PythonAnywhere, lineup images needed for the website were hosted separately on Github Pages \autocite{gtihub_github_2022-1}, a static site hosting service provided by Github \autocite{gtihub_github_2022}, stored in private Github repositories.
+The uniform resource locator (URL) to lineup images were set to be unique random strings such that images can not be accessed by general approaches without knowing the correct URL.
+This avoided participants seeing the lineup beforehand.
+The front-end of the website was built using a JavaScript \autocite{flanagan_javascript_2006} library jsPsych \autocite{de_leeuw_jspsych_2015} which is specialized in creating online behavioral experiments.
 One of the reasons we chose to use this library was it had a modularized but highly customizable template which could record participant's response time automatically.
 This is essential for us to confirm the quality of the data by checking exceptionally fast or slow responses.
+Figure \ref{fig:DiagramExperiment} summarizes the experiment setup.
+
+\begin{figure}
+\centering
+\includegraphics[width=4.6875in,height=4.6875in]{figures/experiment_tech.pdf}
+\caption{Diagram of online experiment setup.
+The server-side of the study webiste used Flask as backend hosted on PythonAnywhere.
+And the client-side used jsPsych to run experiment.
+\label{fig:DiagramExperiment}}
+\end{figure}
+
 The first two pages of the website are the explanatory statement and the consent form.
-Participants needed to read the documentation and agree with the terms to advance to the short survey.
+Participants needed to read the documents and agree with the terms to advance to the short survey.
 With the completion of the survey, the next page is the training page, which contained instructions on how to evaluate the lineup, give the reason and choose the confidence level.
 Followed by the training page, there are 20 lineups each on a single page.
+Figure \ref{fig:WebsiteLayout} illustrates the layout of the website.
 At the end of the experiment, the participants would be redirected back to Prolific and waited for researchers' responses.
+
+\begin{figure}
+\centering
+\includegraphics[width=5.72917in,height=3.11458in]{figures/website.png}
+\caption{Layout of the study website. Participants needed to choose the most different plots on the right and select their reasons and confidence levels on the left. \label{fig:WebsiteLayout}}
+\end{figure}
 
@@
 
-Two models were used in both experiments, which were linear models with some degree of violations of classical assumptions. Residuals vs.~fitted values plots were used to diagnose the issues.
+Two models were used in both experiments, which were linear models with some degree of violations of classical assumptions. Residual vs.~fitted values plot were used to present the model defects.
 
@@
 \FunctionTok{library}\NormalTok{(visage)}
-\NormalTok{mod }\OtherTok{\textless{}{-}} \FunctionTok{cubic\_model}\NormalTok{(}\AttributeTok{a =} \SpecialCharTok{{-}}\DecValTok{1}\NormalTok{, }\AttributeTok{b =} \SpecialCharTok{{-}}\DecValTok{1}\NormalTok{, }\AttributeTok{c =} \DecValTok{1}\NormalTok{, }\AttributeTok{sigma =} \FloatTok{0.5}\NormalTok{, }
+\NormalTok{mod }\OtherTok{\textless{}{-}} \FunctionTok{cubic\_model}\NormalTok{(}\AttributeTok{a =} \SpecialCharTok{{-}}\DecValTok{3}\NormalTok{, }\AttributeTok{b =} \SpecialCharTok{{-}}\DecValTok{3}\NormalTok{, }\AttributeTok{c =} \DecValTok{1}\NormalTok{, }\AttributeTok{sigma =} \FloatTok{0.5}\NormalTok{, }
                    \AttributeTok{x =} \FunctionTok{rand\_uniform}\NormalTok{(}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{), }\AttributeTok{z =} \FunctionTok{rand\_uniform}\NormalTok{(}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{))}
@@
 ##             y          x          z          e      .resid   .fitted
-## 1  0.72328137  0.8783282 0.17687898  0.1539500  0.37649863  0.34678274
-## 2 -0.01317346  0.9593166 0.30200008 -0.3526064 -0.22899266  0.21581920
-## 3  0.10405716  0.7677855 0.72090641 -0.4481692  0.36279717 -0.25874001
-## 4 -0.53007353  0.9238035 0.55802451 -0.8849474 -0.46221585 -0.06785768
-## 5  0.33947802 -0.7784446 0.02191688  0.2307539 -0.04808729  0.38756531
+## 1 -2.24847020  0.8783282 0.17687898  0.1539500  0.65761568 -2.906086
+## 2 -3.85694079  0.9593166 0.30200008 -0.3526064 -0.45596841 -3.400972
+## 3 -3.76887401  0.7677855 0.72090641 -0.4481692  0.56311185 -4.331986
+## 4 -4.78398171  0.9238035 0.55802451 -0.8849474 -0.68606690 -4.097915
+## 5  0.06998178 -0.7784446 0.02191688  0.2307539 -0.07869221  0.148674
 \end{verbatim}
@@
 
+Lineup can be produced by using the methods \texttt{mod\$gen\_lineup()} and \texttt{mod\$plot\_lineup()}.
+Method \texttt{mod\$gen\_lineup} takes the number of realizations \texttt{n} and the number of plots in a lineup \texttt{k} as inputs.
+And the method \texttt{mod\$plot\_lineup()} has the same user interface as \texttt{mod\$plot()}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{mod}\SpecialCharTok{$}\FunctionTok{plot\_lineup}\NormalTok{(mod}\SpecialCharTok{$}\FunctionTok{gen\_lineup}\NormalTok{(}\AttributeTok{n =} \DecValTok{50}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{), }\AttributeTok{type =} \StringTok{"resid"}\NormalTok{, }\AttributeTok{size =} \DecValTok{1}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{thesis_files/figure-latex/unnamed-chunk-4-1.pdf}
+
 The cubic model class also provides method to compute the expected values of residuals.
@@
 \begin{verbatim}
-## [1]  0.05836465 -0.03566983 -0.09457480 -0.01126447  0.08314445
+## [1]  1.6431906  1.2501862 -1.9594998 -1.4124888  0.4786119
 \end{verbatim}
@@
 
-To obtain a measure of the impact of \(a\) and \(b\) on the residuals, we need to address several properties of the residuals.
-First, because \(\boldsymbol{R}_a\) is not an identity matrix, the residuals will have different variance and will be correlated with each other.
+To obtain a measure of the impact of \(a\) and \(b\) on the residuals adjusted for variance and covariance, we need to address several properties of the residuals.
+First, the variance of the residuals \(\sigma^2\boldsymbol{R}_a\) is not an identity matrix.
 This can be fixed by standardizing the residuals by their variance-covariance matrix.
@@
 For simplicity, we set \(\boldsymbol{W}=n^{-1}\boldsymbol{1}\).
-This gives the effect size: \[ES=n^{-1}||\sigma^{-1}\boldsymbol{R}_a^{-\frac{1}{2}}\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b||^2=n^{-1}\sigma^{-2}||\boldsymbol{R}_a^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2\approx n^{-1}\sigma^{-2}||diag(\boldsymbol{R}_a)^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2.\]
+Considering the high time complexity of computing the square root of \(\boldsymbol{R}_a\), off-diagonal elements of \(\boldsymbol{R}_a\) are set to be zeros.
+This gives the effect size: \[ES=n^{-1}||\sigma^{-1}\boldsymbol{R}_a^{-\frac{1}{2}}\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b||^2=n^{-1}\sigma^{-2}||\boldsymbol{R}_a^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2\approx n^{-1}\sigma^{-2}||diag(\boldsymbol{R}_a)^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2,\] where \(diag(\boldsymbol{R}_a)\) is the diagonal matrix constructed from the diagnoal element of \(\boldsymbol{R}_a\).
 
 The interpretation of this effect size is the impact of parameter \(a\) and \(b\) on the squared deviation of the standardized expected residual per observation.
-It is not directly related to the shape, or the pattern observed from the residual plot, but it is a reasonable approximation of the degree of the deviation from the null residuals under our cubic model setting.
+It is not directly related to the shape, or the pattern human observed from the residual plot, but it is a reasonable approximation of the degree of the visual deviation from the null residuals under our cubic model setting.
+Figure \ref{fig:effectsize} shows four residual plots with different effect sizes.
+As effect sizes increases, the strength of the signal become stronger.
 
-\includegraphics{thesis_files/figure-latex/unnamed-chunk-5-1.pdf}
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/effectsize-1.pdf}
+\caption{\label{fig:effectsize}Cubic model residual plots under different effect sizes.}
+\end{figure}
 

### 2022-02-22 Thesis/01-chap1.Rmd
refine notations

@@
 
-The first model was a cubic linear model with 2 regressors, which can be expressed by: $$\boldsymbol{Y}= 1 + (2-c)\boldsymbol{X} + c\boldsymbol{Z} + a[(2-c)\boldsymbol{X}]^2+a(c\boldsymbol{Z})^2+b[(2-c)\boldsymbol{X}]^3+b(c\boldsymbol{Z})^3+\boldsymbol{\varepsilon},$$ where $c \in (0,2)$, $a \in (-3,3)$, $b \in (-3,3)$, $\boldsymbol{\varepsilon}\overset{iid}{\sim} N(\boldsymbol{0},\sigma^2\boldsymbol{I})$, $\boldsymbol{X}$ and $\boldsymbol{Z}$ are $n\times1$ matrices.
+The first model was a cubic linear model with two regressors, which can be expressed by:
+\begin{equation} \label{eq:cubic}
+\boldsymbol{Y}= 1 + (2-c)\boldsymbol{X} + c\boldsymbol{Z} + a[(2-c)\boldsymbol{X}]^2+a(c\boldsymbol{Z})^2+b[(2-c)\boldsymbol{X}]^3+b(c\boldsymbol{Z})^3+\boldsymbol{\varepsilon},
+\end{equation}
+where $c \in (0,2)$, $a \in (-3,3)$, $b \in (-3,3)$, $\boldsymbol{\varepsilon}\overset{iid}{\sim} N(\boldsymbol{0},\sigma^2\boldsymbol{I})$, $\boldsymbol{Y}$, $\boldsymbol{X}$ and $\boldsymbol{Z}$ are $n\times1$ matrices.
 
 This defines a cubic relationship between $\boldsymbol{Y}$, $\boldsymbol{X}$ and $\boldsymbol{Z}$.
-Meanwhile, to create nonlinearity defect, the null model fitted by the OLS is: $$\boldsymbol{Y}=\boldsymbol{b}_0+\boldsymbol{b}_1\boldsymbol{X}+\boldsymbol{b}_2\boldsymbol{Z}+\boldsymbol{e}.$$
+Meanwhile, to create nonlinearity defect, the null model followed the assumptions of the classical normal linear regression model (CNLRM), fitted by OLS is:
+\begin{equation} \label{eq:cubic_null}
+\boldsymbol{Y}=\beta_0+\beta_1\boldsymbol{X}+\beta_2\boldsymbol{Z}+\boldsymbol{u},
+\end{equation}
+where $\boldsymbol{u} \sim N(0,\sigma^2_u\boldsymbol{I})$.
 
@@
 
-### Distribution of the Residuals
+\begin{lemma}[Distribution of residuals produced by the cubic model] \label{lemma:cubic}
+Given the data generating process in Equation (\ref{eq:cubic}), and null model in Equation (\ref{eq:cubic_null}).
+Let $\boldsymbol{X}_a=[\boldsymbol{1},\boldsymbol{X},\boldsymbol{Z}]$ denotes the set of regressors in matrix form.
+Then, the residuals obtained from the null model are $$\boldsymbol{e} \sim N(\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b, \sigma^2\boldsymbol{R}_a),$$ where $\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$, $\boldsymbol{X}_b=[\boldsymbol{X}^2,\boldsymbol{Z}^2,\boldsymbol{X}^3,\boldsymbol{Z}^3]$ and $\boldsymbol{\beta}_b=(a(2-c)^2,ac^2,b(2-c)^3,bc^3)'$.
+\end{lemma}
 
-Let $\boldsymbol{X}_a=[\boldsymbol{1},\boldsymbol{X},\boldsymbol{Z}]$ denotes the set of regressors in matrix form.
-Using the Frisch–Waugh–Lovell theorem, the residual obtained by the null model is $\boldsymbol{e}=\boldsymbol{R}_a\boldsymbol{Y}=\boldsymbol{R}_a(\boldsymbol{X}_a\boldsymbol{\beta}_a+\boldsymbol{X}_b\boldsymbol{\beta}_b+\boldsymbol{\varepsilon}),$ where $\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$, $\boldsymbol{\beta}_a=(1,2-c,c)'$, $\boldsymbol{X}_b=[\boldsymbol{X}^2,\boldsymbol{Z}^2,\boldsymbol{X}^3,\boldsymbol{Z}^3]$ and $\boldsymbol{\beta}_b=(a(2-c)^2,ac^2,b(2-c)^3,bc^3)'$.
+\begin{proof}
+Using the Frisch–Waugh–Lovell theorem, the residuals obtained by the null model are $$\boldsymbol{e}=\boldsymbol{R}_a\boldsymbol{Y}=\boldsymbol{R}_a(\boldsymbol{X}_a\boldsymbol{\beta}_a+\boldsymbol{X}_b\boldsymbol{\beta}_b+\boldsymbol{\varepsilon}),$$ where $\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$, $\boldsymbol{\beta}_a=(1,2-c,c)'$, $\boldsymbol{X}_b=[\boldsymbol{X}^2,\boldsymbol{Z}^2,\boldsymbol{X}^3,\boldsymbol{Z}^3]$ and $\boldsymbol{\beta}_b=(a(2-c)^2,ac^2,b(2-c)^3,bc^3)'$.
 
 Because $\boldsymbol{R}_a\boldsymbol{X}_a=\boldsymbol{0}$, we have $\boldsymbol{e}=\boldsymbol{R}_a(\boldsymbol{X}_b\boldsymbol{\beta}_b+\boldsymbol{\varepsilon}).$ Since $\boldsymbol{\varepsilon} \sim N(0,\sigma^2\boldsymbol{I})$, it follows that $\boldsymbol{e} \sim N(\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b, \sigma^2\boldsymbol{R}_a)$.
+\end{proof}
 
-### Shape of the Residuals
-
-The previous result shows that the expectation of the residuals is clearly a function of $\boldsymbol{X}$ and $\boldsymbol{Z}$.
+Lemma \@ref(lemma:cubic) shows that the expectation of the residuals is clearly a function of $\boldsymbol{X}$ and $\boldsymbol{Z}$.
 Hence, we would expect to see some patterns in the residual plot.
@@
 An instance of cubic model class contains methods of simulating data and making residual plot.
-Method `mod$gen()` returns a data frame containing realizations of $x$, $z$, $y$ and $\varepsilon$ simulated from the model.
+Method `mod$gen()` returns a data frame containing realizations of $X$, $Z$, $Y$ and $\varepsilon$ simulated from the model.
 The number of realizations depends on the integer argument `n`.
@@
 
-```{r cubic1, fig.cap = "cubic 1", fig.height=6.3}
+```{r cubic1, fig.cap = "cubic 1"}
 library(visage)
@@
 expand.grid(c=seq(1),
-            a=seq(-1,1,0.5),
-            b=seq(-1,1,0.5),
+            a=seq(-1,1,1),
+            b=seq(-1,1,1),
             e_sigma=c(0.5)) %>%
@@
 
-result$plots %>% patchwork::wrap_plots(ncol=5)
+result$plots %>% patchwork::wrap_plots(ncol=3)
 ```
 
-```{r cubic2, fig.cap = "cubic 2", fig.height=7.875}
-expand.grid(c=seq(0.5,1.5,0.25),
-            a=seq(-1,1,1),
-            b=seq(-1,1,1),
+```{r cubic2, fig.cap = "cubic 2", fig.height=1.9}
+expand.grid(c=c(0,0.5,1,1.5,2),
+            a=c(-1),
+            b=c(1),
             e_sigma=c(0.5)) %>%
@@
 
-```{r cubic3, fig.cap = "cubic 3", fig.height=7.875}
+```{r cubic3, fig.cap = "cubic 3", fig.height=1.9}
 expand.grid(c=c(1),
-            a=seq(-1,1,1),
-            b=seq(-1,1,1),
-            e_sigma=c(0.25,0.5,1,2)) %>%
+            a=c(-1),
+            b=c(1),
+            e_sigma=c(0.25,0.5,1,2,4)) %>%
   arrange(a,b,c,desc(e_sigma)) %>%
@@
 
-result$plots %>% patchwork::wrap_plots(ncol=4)
+result$plots %>% patchwork::wrap_plots(ncol=5)
 ```
-
-### Effect Size
 
@@
 Considering the high time complexity of computing the square root of $\boldsymbol{R}_a$, off-diagonal elements of $\boldsymbol{R}_a$ are set to be zeros.
-This gives the effect size: $$ES=n^{-1}||\sigma^{-1}\boldsymbol{R}_a^{-\frac{1}{2}}\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b||^2=n^{-1}\sigma^{-2}||\boldsymbol{R}_a^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2\approx n^{-1}\sigma^{-2}||diag(\boldsymbol{R}_a)^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2,$$ where $diag(\boldsymbol{R}_a)$ is the diagonal matrix constructed from the diagnoal element of $\boldsymbol{R}_a$.
+This gives the effect size: $$ES=n^{-1}||\sigma^{-1}\boldsymbol{R}_a^{-\frac{1}{2}}\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b||^2=n^{-1}\sigma^{-2}||\boldsymbol{R}_a^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2\approx n^{-1}\sigma^{-2}||diag(\boldsymbol{R}_a)^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2,$$ where $diag(\boldsymbol{R}_a)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$.
 
@@
 
-```{r effectsize, fig.cap = "Cubic model residual plots under different effect sizes."}
+```{r effectsize, fig.cap = "Cubic model residual plots under different effect sizes. The larger the effect size, the stronger the signal."}
 mod <- cubic_model(-1, -1, 1, 0.5)
@@
 
-###
+Under the cubic model setting, there is an exact conventional test for testing the nonlinearity defect, which is F-test.
+For F-test, the null hypothesis is $H_0:a=b=0$, and the alternative hypothesis is $H_1:\text{at least one of them} \neq 0$.
+During the simulation of the lineup data, the F-statistic and the p-value will be recorded for comparison between the power of conventional test and visual test.
 
@@
 
+Another model used in the experiments was a heteroskedasticity model with one regressor, which can be expressed by: 
+\begin{equation} \label{eq:heter}
+Y_i = 1 + X_i + \varepsilon_i, ~i = 1,...,n,
+\end{equation}
+where $a \in \{-1,0,1\}$, $b\in (0,32)$ and $\varepsilon_i \overset{iid}{\sim} N(0,1+b(2-|a|)(X_i-a)^2)$.
+
+To create heteroskedasticity defect, OLS was used to fit the null model:
+\begin{equation} \label{eq:heter_null}
+\boldsymbol{Y}=\beta_0+\beta_1\boldsymbol{X}+\boldsymbol{u},
+\end{equation}
+where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2_u\boldsymbol{I})$.
+
+In this case, estimators of $\beta_0$ and $\beta_1$ are unbiased, but the error term has non-constant variance.
+
+\begin{lemma}[Distribution of residuals produced by the heteroskedasticity model] \label{lemma:heter}
+Given the data generating process in Equation (\ref{eq:heter}) and null model in Equation (\ref{eq:heter_null}).
+Let $\boldsymbol{X}_a=[\boldsymbol{1},\boldsymbol{X}]$ denotes the set of regressors in matrix form.
+The residuals obtained from the null model are $$\boldsymbol{e} \sim N(\boldsymbol{0}, \boldsymbol{R}_a\boldsymbol{V}),$$ where $\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ and $\boldsymbol{V}$ is a diagonal matrix with $V_{ii}=1+b(2-|a|)(X_i  -  a)^2$, $i = 1,...,n$.
+\end{lemma}
+
+\begin{proof}
+Using the Frisch–Waugh–Lovell theorem, the residuals obtained by the null model are $\boldsymbol{e}=\boldsymbol{R}_a\boldsymbol{Y}=\boldsymbol{R}_a(\boldsymbol{X}_a\boldsymbol{\beta}_a+\boldsymbol{\varepsilon}),$ where $\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ and $\boldsymbol{\beta}_a=(1,1)'$.
+
+Because $\boldsymbol{R}_a\boldsymbol{X}_a=\boldsymbol{0}$, we have $\boldsymbol{e}=\boldsymbol{R}_a\boldsymbol{\varepsilon}$.
+Hence, the residuals $\boldsymbol{e}$ follow $N(\boldsymbol{0}, \boldsymbol{R}_a\boldsymbol{V})$, where $\boldsymbol{V}$ is a diagonal matrix with $V_{ii}=1+b(2-|a|)(X_i  -  a)^2$, $i = 1,...,n$.
+\end{proof}
+
+Lemma \@ref(lemma:heter) shows
+
+This shows that the variance of the residuals is clearly a function of X and Z. Hence, we
+would expect to see some patterns in the residual plot.
+
+## Distribution of regressors
+
 \clearpage

### 2022-02-22 Thesis/_book/thesis.tex
refine notations

@@
 
+\usepackage{amsthm}
+\newtheorem{theorem}{Theorem}[chapter]
+\newtheorem{lemma}{Lemma}[chapter]
+\newtheorem{corollary}{Corollary}[chapter]
+\newtheorem{proposition}{Proposition}[chapter]
+\newtheorem{conjecture}{Conjecture}[chapter]
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}[chapter]
+\theoremstyle{definition}
+\newtheorem{example}{Example}[chapter]
+\theoremstyle{definition}
+\newtheorem{exercise}{Exercise}[chapter]
+\theoremstyle{definition}
+\newtheorem{hypothesis}{Hypothesis}[chapter]
+\theoremstyle{remark}
+\newtheorem*{remark}{Remark}
+\newtheorem*{solution}{Solution}
 \begin{document}
@@
 
-The first model was a cubic linear model with 2 regressors, which can be expressed by: \[\boldsymbol{Y}= 1 + (2-c)\boldsymbol{X} + c\boldsymbol{Z} + a[(2-c)\boldsymbol{X}]^2+a(c\boldsymbol{Z})^2+b[(2-c)\boldsymbol{X}]^3+b(c\boldsymbol{Z})^3+\boldsymbol{\varepsilon},\] where \(c \in (0,2)\), \(a \in (-3,3)\), \(b \in (-3,3)\), \(\boldsymbol{\varepsilon}\overset{iid}{\sim} N(\boldsymbol{0},\sigma^2\boldsymbol{I})\), \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) are \(n\times1\) matrices.
+The first model was a cubic linear model with two regressors, which can be expressed by:
+\begin{equation} \label{eq:cubic}
+\boldsymbol{Y}= 1 + (2-c)\boldsymbol{X} + c\boldsymbol{Z} + a[(2-c)\boldsymbol{X}]^2+a(c\boldsymbol{Z})^2+b[(2-c)\boldsymbol{X}]^3+b(c\boldsymbol{Z})^3+\boldsymbol{\varepsilon},
+\end{equation}
+where \(c \in (0,2)\), \(a \in (-3,3)\), \(b \in (-3,3)\), \(\boldsymbol{\varepsilon}\overset{iid}{\sim} N(\boldsymbol{0},\sigma^2\boldsymbol{I})\), \(\boldsymbol{Y}\), \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) are \(n\times1\) matrices.
 
 This defines a cubic relationship between \(\boldsymbol{Y}\), \(\boldsymbol{X}\) and \(\boldsymbol{Z}\).
-Meanwhile, to create nonlinearity defect, the null model fitted by the OLS is: \[\boldsymbol{Y}=\boldsymbol{b}_0+\boldsymbol{b}_1\boldsymbol{X}+\boldsymbol{b}_2\boldsymbol{Z}+\boldsymbol{e}.\]
+Meanwhile, to create nonlinearity defect, the null model followed the assumptions of the classical normal linear regression model (CNLRM), fitted by OLS is:
+\begin{equation} \label{eq:cubic_null}
+\boldsymbol{Y}=\beta_0+\beta_1\boldsymbol{X}+\beta_2\boldsymbol{Z}+\boldsymbol{u},
+\end{equation}
+where \(\boldsymbol{u} \sim N(0,\sigma^2_u\boldsymbol{I})\).
 
@@
 
-\hypertarget{distribution-of-the-residuals}{%
-\subsection{Distribution of the Residuals}\label{distribution-of-the-residuals}}
-
-Let \(\boldsymbol{X}_a=[\boldsymbol{1},\boldsymbol{X},\boldsymbol{Z}]\) denotes the set of regressors in matrix form.
-Using the Frisch--Waugh--Lovell theorem, the residual obtained by the null model is \(\boldsymbol{e}=\boldsymbol{R}_a\boldsymbol{Y}=\boldsymbol{R}_a(\boldsymbol{X}_a\boldsymbol{\beta}_a+\boldsymbol{X}_b\boldsymbol{\beta}_b+\boldsymbol{\varepsilon}),\) where \(\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'\), \(\boldsymbol{\beta}_a=(1,2-c,c)'\), \(\boldsymbol{X}_b=[\boldsymbol{X}^2,\boldsymbol{Z}^2,\boldsymbol{X}^3,\boldsymbol{Z}^3]\) and \(\boldsymbol{\beta}_b=(a(2-c)^2,ac^2,b(2-c)^3,bc^3)'\).
-
-Because \(\boldsymbol{R}_a\boldsymbol{X}_a=\boldsymbol{0}\), we have \(\boldsymbol{e}=\boldsymbol{R}_a(\boldsymbol{X}_b\boldsymbol{\beta}_b+\boldsymbol{\varepsilon}).\) Since \(\boldsymbol{\varepsilon} \sim N(0,\sigma^2\boldsymbol{I})\), it follows that \(\boldsymbol{e} \sim N(\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b, \sigma^2\boldsymbol{R}_a)\).
-
-\hypertarget{shape-of-the-residuals}{%
-\subsection{Shape of the Residuals}\label{shape-of-the-residuals}}
-
-The previous result shows that the expectation of the residuals is clearly a function of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\).
+\begin{lemma}[Distribution of residuals produced by the cubic model] \label{lemma:cubic}
+Given the data generating process in Equation (\ref{eq:cubic}), and null model in Equation (\ref{eq:cubic_null}).
+Let $\boldsymbol{X}_a=[\boldsymbol{1},\boldsymbol{X},\boldsymbol{Z}]$ denotes the set of regressors in matrix form.
+Then, the residuals obtained from the null model are $$\boldsymbol{e} \sim N(\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b, \sigma^2\boldsymbol{R}_a),$$ where $\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$, $\boldsymbol{X}_b=[\boldsymbol{X}^2,\boldsymbol{Z}^2,\boldsymbol{X}^3,\boldsymbol{Z}^3]$ and $\boldsymbol{\beta}_b=(a(2-c)^2,ac^2,b(2-c)^3,bc^3)'$.
+\end{lemma}
+
+\begin{proof}
+Using the Frisch–Waugh–Lovell theorem, the residuals obtained by the null model are $$\boldsymbol{e}=\boldsymbol{R}_a\boldsymbol{Y}=\boldsymbol{R}_a(\boldsymbol{X}_a\boldsymbol{\beta}_a+\boldsymbol{X}_b\boldsymbol{\beta}_b+\boldsymbol{\varepsilon}),$$ where $\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$, $\boldsymbol{\beta}_a=(1,2-c,c)'$, $\boldsymbol{X}_b=[\boldsymbol{X}^2,\boldsymbol{Z}^2,\boldsymbol{X}^3,\boldsymbol{Z}^3]$ and $\boldsymbol{\beta}_b=(a(2-c)^2,ac^2,b(2-c)^3,bc^3)'$.
+
+Because $\boldsymbol{R}_a\boldsymbol{X}_a=\boldsymbol{0}$, we have $\boldsymbol{e}=\boldsymbol{R}_a(\boldsymbol{X}_b\boldsymbol{\beta}_b+\boldsymbol{\varepsilon}).$ Since $\boldsymbol{\varepsilon} \sim N(0,\sigma^2\boldsymbol{I})$, it follows that $\boldsymbol{e} \sim N(\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b, \sigma^2\boldsymbol{R}_a)$.
+\end{proof}
+
+Lemma \ref{lemma:cubic} shows that the expectation of the residuals is clearly a function of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\).
 Hence, we would expect to see some patterns in the residual plot.
@@
 An instance of cubic model class contains methods of simulating data and making residual plot.
-Method \texttt{mod\$gen()} returns a data frame containing realizations of \(x\), \(z\), \(y\) and \(\varepsilon\) simulated from the model.
+Method \texttt{mod\$gen()} returns a data frame containing realizations of \(X\), \(Z\), \(Y\) and \(\varepsilon\) simulated from the model.
 The number of realizations depends on the integer argument \texttt{n}.
@@
 \end{figure}
-
-\hypertarget{effect-size}{%
-\subsection{Effect Size}\label{effect-size}}
 
@@
 Considering the high time complexity of computing the square root of \(\boldsymbol{R}_a\), off-diagonal elements of \(\boldsymbol{R}_a\) are set to be zeros.
-This gives the effect size: \[ES=n^{-1}||\sigma^{-1}\boldsymbol{R}_a^{-\frac{1}{2}}\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b||^2=n^{-1}\sigma^{-2}||\boldsymbol{R}_a^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2\approx n^{-1}\sigma^{-2}||diag(\boldsymbol{R}_a)^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2,\] where \(diag(\boldsymbol{R}_a)\) is the diagonal matrix constructed from the diagnoal element of \(\boldsymbol{R}_a\).
+This gives the effect size: \[ES=n^{-1}||\sigma^{-1}\boldsymbol{R}_a^{-\frac{1}{2}}\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b||^2=n^{-1}\sigma^{-2}||\boldsymbol{R}_a^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2\approx n^{-1}\sigma^{-2}||diag(\boldsymbol{R}_a)^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2,\] where \(diag(\boldsymbol{R}_a)\) is the diagonal matrix constructed from the diagonal elements of \(\boldsymbol{R}_a\).
 
@@
 \includegraphics{thesis_files/figure-latex/effectsize-1.pdf}
-\caption{\label{fig:effectsize}Cubic model residual plots under different effect sizes.}
+\caption{\label{fig:effectsize}Cubic model residual plots under different effect sizes. The larger the effect size, the stronger the signal.}
 \end{figure}
 
-\hypertarget{section}{%
-\subsection{}\label{section}}
+Under the cubic model setting, there is an exact conventional test for testing the nonlinearity defect, which is F-test.
+For F-test, the null hypothesis is \(H_0:a=b=0\), and the alternative hypothesis is \(H_1:\text{at least one of them} \neq 0\).
+During the simulation of the lineup data, the F-statistic and the p-value will be recorded for comparison between the power of conventional test and visual test.
 
@@
 \section{Heteroskedasticity Model}\label{heteroskedasticity-model}}
+
+Another model used in the experiments was a heteroskedasticity model with one regressor, which can be expressed by:
+\begin{equation} \label{eq:heter}
+Y_i = 1 + X_i + \varepsilon_i, ~i = 1,...,n,
+\end{equation}
+where \(a \in \{-1,0,1\}\), \(b\in (0,32)\) and \(\varepsilon_i \overset{iid}{\sim} N(0,1+b(2-|a|)(X_i-a)^2)\).
+
+To create heteroskedasticity defect, OLS was used to fit the null model:
+\begin{equation} \label{eq:heter_null}
+\boldsymbol{Y}=\beta_0+\beta_1\boldsymbol{X}+\boldsymbol{u},
+\end{equation}
+where \(\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2_u\boldsymbol{I})\).
+
+In this case, estimators of \(\beta_0\) and \(\beta_1\) are unbiased, but the error term has non-constant variance.
+
+\begin{lemma}[Distribution of residuals produced by the heteroskedasticity model] \label{lemma:heter}
+Given the data generating process in Equation (\ref{eq:heter}) and null model in Equation (\ref{eq:heter_null}).
+Let $\boldsymbol{X}_a=[\boldsymbol{1},\boldsymbol{X}]$ denotes the set of regressors in matrix form.
+The residuals obtained from the null model are $$\boldsymbol{e} \sim N(\boldsymbol{0}, \boldsymbol{R}_a\boldsymbol{V}),$$ where $\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ and $\boldsymbol{V}$ is a diagonal matrix with $V_{ii}=1+b(2-|a|)(X_i  -  a)^2$, $i = 1,...,n$.
+\end{lemma}
+
+\begin{proof}
+Using the Frisch–Waugh–Lovell theorem, the residuals obtained by the null model are $\boldsymbol{e}=\boldsymbol{R}_a\boldsymbol{Y}=\boldsymbol{R}_a(\boldsymbol{X}_a\boldsymbol{\beta}_a+\boldsymbol{\varepsilon}),$ where $\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ and $\boldsymbol{\beta}_a=(1,1)'$.
+
+Because $\boldsymbol{R}_a\boldsymbol{X}_a=\boldsymbol{0}$, we have $\boldsymbol{e}=\boldsymbol{R}_a\boldsymbol{\varepsilon}$.
+Hence, the residuals $\boldsymbol{e}$ follow $N(\boldsymbol{0}, \boldsymbol{R}_a\boldsymbol{V})$, where $\boldsymbol{V}$ is a diagonal matrix with $V_{ii}=1+b(2-|a|)(X_i  -  a)^2$, $i = 1,...,n$.
+\end{proof}
+
+Lemma \ref{lemma:heter} shows
+
+This shows that the variance of the residuals is clearly a function of X and Z. Hence, we
+would expect to see some patterns in the residual plot.
+
+\hypertarget{distribution-of-regressors}{%
+\section{Distribution of regressors}\label{distribution-of-regressors}}
 

### 2022-02-22 Thesis/01-chap1.Rmd
refine cubic model

@@
 Prolific provides an international participant pool with the option to apply flexible pre-screening filters.
-In this study, we recruited participants who was fluent in English with at least $10$ previous submissions and $98$% approval rate in other Prolific studies for quality control.
+In this study, we recruited $62$ participants who was fluent in English with at least $10$ previous submissions and $98$% approval rate in other Prolific studies for quality control.
 Further, balance sample across gender was imposed to prevent gender bias.
@@
 This is essential for us to confirm the quality of the data by checking exceptionally fast or slow responses.
-Figure \@ref(fig:DiagramExperiment) summarizes the experiment setup.
+Figure \@ref(fig:DiagramExperiment) summarizes the online experiment setup.
 
@@
 Hence, we would expect to see some patterns in the residual plot.
+Let $X_i$ and $Z_i$, $i = 1,...,n$, be independent random variables follow uniform distribution $U(-1,1)$.
 Given the expectation of the residuals, we could plot the expected values of residuals against the observed values.
-Figure \@ref(fig:cubic1), \@ref(fig:cubic2) and \@ref(fig:cubic3) illustrate the shape of residuals under different parameter settings.
-
-The residuals used in these three figures are simulated from the cubic models built using the `cubic_model()` function from the package `visage`.
-`cubic_model()` is a cubic model class constructor, which takes arguments `a`, `b`, `c`, `sigma`, `x` and `z`, where the first four are numeric values defined above, and `x` and `z` are random variable instances created by the random variable abstract base class constructor `rand_var()`.
-
-Here, we set $\boldsymbol{X}$ and $\boldsymbol{Z}$ to be random uniform variables ranged from $-1$ to $1$.
-This can be done by using the random uniform variable class constructor `rand_uniform()` inherited from the random variable abstract base class.
-It only takes two arguments which are the lower bound and the upper bound of the support.
-
-```{r echo = TRUE, cache = FALSE}
-library(visage)
-mod <- cubic_model(a = -3, b = -3, c = 1, sigma = 0.5, 
-                   x = rand_uniform(-1, 1), z = rand_uniform(-1, 1))
-```
-
-An instance of cubic model class contains methods of simulating data and making residual plot.
-Method `mod$gen()` returns a data frame containing realizations of $X$, $Z$, $Y$ and $\varepsilon$ simulated from the model.
-The number of realizations depends on the integer argument `n`.
-In addition, if argument `fit_model = TRUE`, a null model will be fitted using the simulated data and residuals and fitted values will be included in the returned data frame.
-
-```{r echo = TRUE, cache = FALSE}
-mod$gen(n = 5, fit_model = TRUE)
-```
-
-Method `mod$plot()` produce a ggplot [@wickham_ggplot2_2011] object.
-It takes a data frame containing columns `.resid` and `.fitted` as input, along with a character argument `type` indicating the type of the data plot, and other aesthetic arguments such as `size` and `alpha` to control the appearance of the plot.
-
-```{r test3, echo = TRUE, cache = FALSE}
-mod$plot(mod$gen(n = 100, fit_model = TRUE), type = "resid", size = 1)
-```
-
-Lineup can be produced by using the methods `mod$gen_lineup()` and `mod$plot_lineup()`.
-Method `mod$gen_lineup` takes the number of realizations `n` and the number of plots in a lineup `k` as inputs.
-And the method `mod$plot_lineup()` has the same user interface as `mod$plot()`.
-
-```{r echo = TRUE, cache = FALSE}
-mod$plot_lineup(mod$gen_lineup(n = 50, k = 20), type = "resid", size = 1)
-```
-
-The cubic model class also provides method to compute the expected values of residuals.
-Method `mod$E()` takes a data frame with columns `x` and `z` as input, and returns a vector of expected values of residuals.
-
-```{r echo = TRUE, cache = FALSE}
-mod$E(mod$gen(n = 5))
-```
+Figure \@ref(fig:cubic1), \@ref(fig:cubic2) and \@ref(fig:cubic3) illustrate the shape of residuals and their expected values under different parameter settings.
 
@@
 result$plots %>% patchwork::wrap_plots(ncol=5)
+```
+
+The residuals used in these three figures are simulated from the cubic models built using the `cubic_model()` function from the package `visage`.
+`cubic_model()` is a cubic model class constructor, which takes arguments `a`, `b`, `c`, `sigma`, `x` and `z`, where the first four are numeric values defined above, and `x` and `z` are random variable instances created by the random variable abstract base class constructor `rand_var()`.
+
+Here, we set $\boldsymbol{X}$ and $\boldsymbol{Z}$ to be random uniform variables ranged from $-1$ to $1$.
+This can be done by using the random uniform variable class constructor `rand_uniform()` inherited from the random variable abstract base class.
+It only takes two arguments which are the lower bound and the upper bound of the support.
+
+```{r echo = TRUE, cache = FALSE}
+library(visage)
+mod <- cubic_model(a = -3, b = -3, c = 1, sigma = 0.5, 
+                   x = rand_uniform(-1, 1), z = rand_uniform(-1, 1))
+```
+
+An instance of cubic model class contains methods of simulating data and making residual plot.
+Method `mod$gen()` returns a data frame containing realizations of $X$, $Z$, $Y$ and $\varepsilon$ simulated from the model.
+The number of realizations depends on the integer argument `n`.
+In addition, if argument `fit_model = TRUE`, a null model will be fitted using the simulated data and residuals and fitted values will be included in the returned data frame.
+
+```{r echo = TRUE, cache = FALSE}
+mod$gen(n = 5, fit_model = TRUE)
+```
+
+Method `mod$plot()` produce a ggplot [@wickham_ggplot2_2011] object.
+It takes a data frame containing columns `.resid` and `.fitted` as input, along with a character argument `type` indicating the type of the data plot, and other aesthetic arguments such as `size` and `alpha` to control the appearance of the plot.
+
+```{r test3, echo = TRUE, cache = FALSE}
+mod$plot(mod$gen(n = 100, fit_model = TRUE), type = "resid", size = 1)
+```
+
+Lineup can be produced by using the methods `mod$gen_lineup()` and `mod$plot_lineup()`.
+Method `mod$gen_lineup` takes the number of realizations `n` and the number of plots in a lineup `k` as inputs.
+And the method `mod$plot_lineup()` has the same user interface as `mod$plot()`.
+
+```{r echo = TRUE, cache = FALSE}
+mod$plot_lineup(mod$gen_lineup(n = 50, k = 20), type = "resid", size = 1)
+```
+
+The cubic model class also provides method to compute the expected values of residuals.
+Method `mod$E()` takes a data frame with columns `x` and `z` as input, and returns a vector of expected values of residuals.
+
+```{r echo = TRUE, cache = FALSE}
+mod$E(mod$gen(n = 5))
 ```

### 2022-02-22 Thesis/_book/thesis.tex
refine cubic model

@@
 Prolific provides an international participant pool with the option to apply flexible pre-screening filters.
-In this study, we recruited participants who was fluent in English with at least \(10\) previous submissions and \(98\)\% approval rate in other Prolific studies for quality control.
+In this study, we recruited \(62\) participants who was fluent in English with at least \(10\) previous submissions and \(98\)\% approval rate in other Prolific studies for quality control.
 Further, balance sample across gender was imposed to prevent gender bias.
@@
 This is essential for us to confirm the quality of the data by checking exceptionally fast or slow responses.
-Figure \ref{fig:DiagramExperiment} summarizes the experiment setup.
+Figure \ref{fig:DiagramExperiment} summarizes the online experiment setup.
 
@@
 Hence, we would expect to see some patterns in the residual plot.
+Let \(X_i\) and \(Z_i\), \(i = 1,...,n\), be independent random variables follow uniform distribution \(U(-1,1)\).
 Given the expectation of the residuals, we could plot the expected values of residuals against the observed values.
-Figure \ref{fig:cubic1}, \ref{fig:cubic2} and \ref{fig:cubic3} illustrate the shape of residuals under different parameter settings.
+Figure \ref{fig:cubic1}, \ref{fig:cubic2} and \ref{fig:cubic3} illustrate the shape of residuals and their expected values under different parameter settings.
+
+From the figure \ref{fig:cubic1}, it can be observed that with fixed \(\sigma\) and \(c\), \(a\) and \(b\) are controlling the 2D projection of a hypersurface, and seemingly performing some rotations along different axes.
+Figure \ref{fig:cubic2} shows that with fixed \(a\), \(b\) and \(\sigma\), \(c\) is controlling the contribution of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) to \(\boldsymbol{Y}\).
+If we move \(c\) toward \(0\) or \(2\), one variable will donminate another, which will mitigate some joint effects and resemble a typical cubic function.
+In figure \ref{fig:cubic3}, \(a\), \(b\) and \(c\) is held constant, and \(\sigma\) is controlling the noises around the expected values.
+As \(\sigma\) decreases, the underlying shape shows up.
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/cubic1-1.pdf}
+\caption{\label{fig:cubic1}cubic 1}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/cubic2-1.pdf}
+\caption{\label{fig:cubic2}cubic 2}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/cubic3-1.pdf}
+\caption{\label{fig:cubic3}cubic 3}
+\end{figure}
 
@@
 ##             y          x          z          e     .resid     .fitted
-## 1 -2.24847020  0.8783282 0.17687898  0.1539500  0.65761568 -2.906086
-## 2 -3.85694079  0.9593166 0.30200008 -0.3526064 -0.45596841 -3.400972
-## 3 -3.76887401  0.7677855 0.72090641 -0.4481692  0.56311185 -4.331986
-## 4 -4.78398171  0.9238035 0.55802451 -0.8849474 -0.68606690 -4.097915
-## 5  0.06998178 -0.7784446 0.02191688  0.2307539 -0.07869221  0.148674
+## 1  0.44298045  0.1862316 -0.5114228  0.2749620  0.3561304  0.08685009
+## 2 -0.46546366 -0.8542129 -0.3789133  0.3543143  0.2827120 -0.74817570
+## 3  0.07636314  0.0383946 -0.3832281 -0.3024668 -0.2698449  0.34620801
+## 4 -1.04762148 -0.4177674 -0.5876853 -0.3101100 -0.1202573 -0.92736416
+## 5 -0.90198583 -0.5439192 -0.4630171 -0.1448947 -0.2487402 -0.65324563
 \end{verbatim}
@@
 \begin{verbatim}
-## [1]  1.6431906  1.2501862 -1.9594998 -1.4124888  0.4786119
+## [1] -0.9369269  0.1945678  0.2816923 -0.3254415  0.7861083
 \end{verbatim}
-
-From the figure \ref{fig:cubic1}, it can be observed that with fixed \(\sigma\) and \(c\), \(a\) and \(b\) are controlling the 2D projection of a hypersurface, and seemingly performing some rotations along different axes.
-Figure \ref{fig:cubic2} shows that with fixed \(a\), \(b\) and \(\sigma\), \(c\) is controlling the contribution of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) to \(\boldsymbol{Y}\).
-If we move \(c\) toward \(0\) or \(2\), one variable will donminate another, which will mitigate some joint effects and resemble a typical cubic function.
-In figure \ref{fig:cubic3}, \(a\), \(b\) and \(c\) is held constant, and \(\sigma\) is controlling the noises around the expected values.
-As \(\sigma\) decreases, the underlying shape shows up.
-
-\begin{figure}
-\centering
-\includegraphics{thesis_files/figure-latex/cubic1-1.pdf}
-\caption{\label{fig:cubic1}cubic 1}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{thesis_files/figure-latex/cubic2-1.pdf}
-\caption{\label{fig:cubic2}cubic 2}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{thesis_files/figure-latex/cubic3-1.pdf}
-\caption{\label{fig:cubic3}cubic 3}
-\end{figure}
 

### 2022-02-23 Thesis/01-chap1.Rmd
heter effect size

@@
 
-Lemma \@ref(lemma:heter) shows
+Lemma \@ref(lemma:heter) shows that variance and covariance of the residuals depend on $\boldsymbol{X}$.
+We could plot the one standard deviation around the residuals to indicate the region about 68% of the residuals should landed.
+Figure \@ref(fig:hetershape) illustrates different shapes of residuals with various values of $a$ and $b$.
+From the plot, it can be observed that if $a = 0$, the residual plot looks like the shape of butterfly.
+If $a = \pm 1$, it looks like a triangle, and the sign of $a$ determines the direction of the shape.
+Parameter $b$ is controlling the strength of the signal.
+As $b$ increases, the pattern becomes less noisy.
 
-This shows that the variance of the residuals is clearly a function of X and Z. Hence, we
-would expect to see some patterns in the residual plot.
+```{r hetershape, fig.height=5.25, fig.cap="test"}
+make_plot2 <- function(a, b) {
+  mod <- heter_model(a = a, b = b)
+  dat <- mod$gen(1000, fit_model = TRUE)
+  dat$one_sigma <- sqrt(1 + (2 - abs(a)) * (dat$x - a)^2 * b)
+  mod$plot(dat, alpha = 0.3, size = 0.5, add_zero_line = FALSE, theme = theme_light(base_size = 5)) +
+    geom_point(aes(.fitted, one_sigma, col = "red"), alpha = 0.6, size = 0.1) +
+    geom_point(aes(.fitted, -one_sigma, col = "red"), alpha = 0.6, size = 0.1) +
+    xlab("fitted values") +
+    ylab("residuals") +
+    ggtitle(bquote(a:.(a)~','~b:.(b))) +
+    theme(legend.position = "none")
+}
+
+expand.grid(a = c(-1,0,1),
+            b = 2^c(0,2,4)) %>%
+  arrange(b,a) %>%
+  rowwise() %>%
+  mutate(plots = list(make_plot2(a = a, b = b))) -> results
+
+results$plots %>% patchwork::wrap_plots(ncol=3)
+```
+
+Similar to the cubic model, the heteroskedasticity model could also be built by the heteroskedasticity model class constructor `heter_model()`.
+This function takes three arguments as inputs, which are `a`, `b` and `x`.
+`a` and `b` are numeric parameters defined in Equation (\ref{eq:heter}).
+`x` needs to be a random variable object.
+
+```{r echo = TRUE}
+library(visage)
+mod <- heter_model(a = 0, b = 16, x = rand_uniform(-1, 1))
+```
+
+Since both the cubic model class and the heteroskedasticity class are inherited from the visual inference model class, which has defined methods of simulating data, making residual plot and producing lineup, heteroskedasticity model object can be used in a similar way as cubic model object.
+The following codes give examples of the use of the object.
+
+```{r echo = TRUE}
+mod$gen(5, fit_model = TRUE)
+```
+
+```{r echo = TRUE}
+mod$plot(mod$gen(100, fit_model = TRUE), size = 1)
+```
+```{r echo = TRUE}
+mod$plot_lineup(mod$gen_lineup(100), size = 1)
+```
+According to Lemma \ref{lemma:heter}, when $b = 0$, the matrix $\boldsymbol{V}$ collapses to an identity matrix.
+Assume the shape of butterfly and triangle have identical visual impacts, the only factor affects human to recognize the pattern is the strength of the signal, where parameter $a$ has no role to play.
+In addition, sample size has a huge impact on the chance of human recognizing the pattern.
+As shown in Figure \@ref(fig:hetern), as sample size increases, the pattern becomes easier to observe.
+However, this effect is less noticeable with large sample size as the outline of the shape has been drawn and residuals have less probability to land outside of it.
+Thus, the effect size of this model can be expressed by $ES = b\sqrt{n}$.
+The square root operator is used for addressing the large sample issue.
+Figure \@ref(fig:hetereffectsize) shows the effectiveness of the effect size.
+
+```{r hetern, fig.height=1.9, fig.cap="Residuals generated from the same heteroskedasticity model but with different sample size.
+As sample size increases, the shape of butterfly becomes more obvious."}
+mod$plot(mod$gen(10, fit_model = TRUE), size = 1, theme = theme_light(base_size = 5)) + ggtitle("n: 10") -> p1
+mod$plot(mod$gen(50, fit_model = TRUE), size = 1, theme = theme_light(base_size = 5)) + ggtitle("n: 50") -> p2
+mod$plot(mod$gen(100, fit_model = TRUE), size = 1, theme = theme_light(base_size = 5)) + ggtitle("n: 100") -> p3
+mod$plot(mod$gen(500, fit_model = TRUE), size = 1, theme = theme_light(base_size = 5)) + ggtitle("n: 500") -> p4
+
+list(p1, p2, p3, p4) %>% patchwork::wrap_plots(ncol=4)
+```
+
+```{r hetereffectsize, fig.cap="Residuals of heteroskedasticity model with different effect size. As effect size increases, the shape of butterfly becomes less noisy."}
+set.seed(10086)
+mod <- heter_model(0, 1)
+mod$plot(mod$gen(500, fit_model = TRUE), size = 1, theme = theme_light(base_size = 5)) + ggtitle(glue::glue("Effect size: {round(mod$effect_size(mod$gen(500)),2)}")) -> p1
+mod <- heter_model(0, 3)
+mod$plot(mod$gen(500, fit_model = TRUE), size = 1, theme = theme_light(base_size = 5)) + ggtitle(glue::glue("Effect size: {round(mod$effect_size(mod$gen(500)),2)}")) -> p2
+mod <- heter_model(0, 9)
+mod$plot(mod$gen(500, fit_model = TRUE), size = 1, theme = theme_light(base_size = 5)) + ggtitle(glue::glue("Effect size: {round(mod$effect_size(mod$gen(500)),2)}")) -> p3
+mod <- heter_model(0, 81)
+mod$plot(mod$gen(500, fit_model = TRUE), size = 1, theme = theme_light(base_size = 5)) + ggtitle(glue::glue("Effect size: {round(mod$effect_size(mod$gen(500)),2)}")) -> p4
+
+list(p1, p2, p3, p4) %>% patchwork::wrap_plots(ncol=2)
+```
 

### 2022-02-23 Thesis/_book/thesis.tex
heter effect size

@@
 
-Lemma \ref{lemma:heter} shows
-
-This shows that the variance of the residuals is clearly a function of X and Z. Hence, we
-would expect to see some patterns in the residual plot.
+Lemma \ref{lemma:heter} shows that variance and covariance of the residuals depend on \(\boldsymbol{X}\).
+We could plot the one standard deviation around the residuals to indicate the region about 68\% of the residuals should landed.
+Figure \ref{fig:hetershape} illustrates different shapes of residuals with various values of \(a\) and \(b\).
+From the plot, it can be observed that if \(a = 0\), the residual plot looks like the shape of butterfly.
+If \(a = \pm 1\), it looks like a triangle, and the sign of \(a\) determines the direction of the shape.
+Parameter \(b\) is controlling the strength of the signal.
+As \(b\) increases, the pattern becomes less noisy.
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/hetershape-1.pdf}
+\caption{\label{fig:hetershape}test}
+\end{figure}
+
+Similar to the cubic model, the heteroskedasticity model could also be built by the heteroskedasticity model class constructor \texttt{heter\_model()}.
+This function takes three arguments as inputs, which are \texttt{a}, \texttt{b} and \texttt{x}.
+\texttt{a} and \texttt{b} are numeric parameters defined in Equation (\ref{eq:heter}).
+\texttt{x} needs to be a random variable object.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{library}\NormalTok{(visage)}
+\NormalTok{mod }\OtherTok{\textless{}{-}} \FunctionTok{heter\_model}\NormalTok{(}\AttributeTok{a =} \DecValTok{0}\NormalTok{, }\AttributeTok{b =} \DecValTok{16}\NormalTok{, }\AttributeTok{x =} \FunctionTok{rand\_uniform}\NormalTok{(}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+Since both the cubic model class and the heteroskedasticity class are inherited from the visual inference model class, which has defined methods of simulating data, making residual plot and producing lineup, heteroskedasticity model object can be used in a similar way as cubic model object.
+The following codes give examples of the use of the object.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{mod}\SpecialCharTok{$}\FunctionTok{gen}\NormalTok{(}\DecValTok{5}\NormalTok{, }\AttributeTok{fit\_model =} \ConstantTok{TRUE}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+##          y          x    sigma         e     .resid  .fitted
+## 1 7.205499  0.9053470 5.218132 5.3001525  0.7420902 6.463409
+## 2 2.560696 -0.8626811 4.981465 2.4233771  1.0441078 1.516588
+## 3 1.121106 -0.6545505 3.835357 0.7756568 -0.9778170 2.098923
+## 4 5.296467  0.7171334 4.178154 3.5793335 -0.6403337 5.936801
+## 5 2.962130 -0.2859733 1.901837 2.2481030 -0.1680473 3.130177
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{mod}\SpecialCharTok{$}\FunctionTok{plot}\NormalTok{(mod}\SpecialCharTok{$}\FunctionTok{gen}\NormalTok{(}\DecValTok{100}\NormalTok{, }\AttributeTok{fit\_model =} \ConstantTok{TRUE}\NormalTok{), }\AttributeTok{size =} \DecValTok{1}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{thesis_files/figure-latex/unnamed-chunk-8-1.pdf}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{mod}\SpecialCharTok{$}\FunctionTok{plot\_lineup}\NormalTok{(mod}\SpecialCharTok{$}\FunctionTok{gen\_lineup}\NormalTok{(}\DecValTok{100}\NormalTok{), }\AttributeTok{size =} \DecValTok{1}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{thesis_files/figure-latex/unnamed-chunk-9-1.pdf}
+According to Lemma \ref{lemma:heter}, when \(b = 0\), the matrix \(\boldsymbol{V}\) collapses to an identity matrix.
+Assume the shape of butterfly and triangle have identical visual impacts, the only factor affects human to recognize the pattern is the strength of the signal, where parameter \(a\) has no role to play.
+In addition, sample size has a huge impact on the chance of human recognizing the pattern.
+As shown in Figure \ref{fig:hetern}, as sample size increases, the pattern becomes easier to observe.
+However, this effect is less noticeable with large sample size as the outline of the shape has been drawn and residuals have less probability to land outside of it.
+Thus, the effect size of this model can be expressed by \(ES = b\sqrt{n}\).
+The square root operator is used for addressing the large sample issue.
+Figure \ref{fig:hetereffectsize} shows the effectiveness of the effect size.
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/hetern-1.pdf}
+\caption{\label{fig:hetern}Residuals generated from the same heteroskedasticity model but with different sample size. As sample size increases, the shape of butterfly becomes more obvious.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/hetereffectsize-1.pdf}
+\caption{\label{fig:hetereffectsize}Residuals of heteroskedasticity model with different effect size. As effect size increases, the shape of butterfly becomes less noisy.}
+\end{figure}
 

### 2022-02-23 Thesis/01-chap1.Rmd
finish heter

@@
 Lemma \@ref(lemma:heter) shows that variance and covariance of the residuals depend on $\boldsymbol{X}$.
-We could plot the one standard deviation around the residuals to indicate the region about 68% of the residuals should landed.
-Figure \@ref(fig:hetershape) illustrates different shapes of residuals with various values of $a$ and $b$.
+We could plot the one standard deviation around the residuals to indicate the region about 68% of the residuals should landed on.
+Figure \@ref(fig:hetershape) illustrates different shapes of residuals undder various values of $a$ and $b$.
 From the plot, it can be observed that if $a = 0$, the residual plot looks like the shape of butterfly.
@@
 
+For the heteroskedasticity model, the conventional test we used was Breusch–Pagan test [@breusch_simple_1979], which tested whether the variance of the error terms of the regression is dependent on the regressors.
+@majumder_validation_2013 suggested that visual test is not expected to perform equally well as conventional test especially when there exists a exact conventional test.
+However, in contrast to the F-test used in the cubic model, Breusch–Pagan test is an approximate test.
+Thus, the power of visual test may exceed the power of Breusch–Pagan test.
+Throughout the study, Breusch–Pagan test statistic and p-value were recorded for comparison.
+
 ## Distribution of regressors
 
+The model definitions given in the previous two sections does not include the specification of the regressors.
+In this sections, distribution of $\boldsymbol{X}$ and $\boldsymbol{Z}$ will be discussed.
+
+The cubic model involves the use of both $\boldsymbol{X}$ and $\boldsymbol{Z}$.
+In the simulation, $X_i$, $i=1,...,n$, had equal chance to follow one of the following distributions: $N(0, 0.09)$, $U(-1, 1)$, $Lognormal(0,0.36)$ and $-Lognormal(0,0.36)$.
+
+Normal and uniform distributions are
+
 \clearpage

### 2022-02-23 Thesis/_book/thesis.tex
finish heter

@@
 Lemma \ref{lemma:heter} shows that variance and covariance of the residuals depend on \(\boldsymbol{X}\).
-We could plot the one standard deviation around the residuals to indicate the region about 68\% of the residuals should landed.
-Figure \ref{fig:hetershape} illustrates different shapes of residuals with various values of \(a\) and \(b\).
+We could plot the one standard deviation around the residuals to indicate the region about 68\% of the residuals should landed on.
+Figure \ref{fig:hetershape} illustrates different shapes of residuals undder various values of \(a\) and \(b\).
 From the plot, it can be observed that if \(a = 0\), the residual plot looks like the shape of butterfly.
@@
 
+For the heteroskedasticity model, the conventional test we used was Breusch--Pagan test \autocite{breusch_simple_1979}, which tested whether the variance of the error terms of the regression is dependent on the regressors.
+\textcite{majumder_validation_2013} suggested that visual test is not expected to perform equally well as conventional test especially when there exists a exact conventional test.
+However, in contrast to the F-test used in the cubic model, Breusch--Pagan test is an approximate test.
+Thus, the power of visual test may exceed the power of Breusch--Pagan test.
+Throughout the study, Breusch--Pagan test statistic and p-value were recorded for comparison.
+
 \hypertarget{distribution-of-regressors}{%
 \section{Distribution of regressors}\label{distribution-of-regressors}}
+
+The model definitions given in the previous two sections does not include the specification of the regressors.
+In this sections, distribution of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) will be discussed.
+
+The cubic model involves the use of both \(\boldsymbol{X}\) and \(\boldsymbol{Z}\).
+In the simulation, \(X_i\), \(i=1,...,n\), had equal chance to follow one of the following distributions: \(N(0, 0.09)\), \(U(-1, 1)\), \(Lognormal(0,0.36)\) and \(-Lognormal(0,0.36)\).
+
+Normal and uniform distributions are
 

### 2022-02-23 Thesis/_book/thesis.tex
add regressor distribution

@@
 Based on the answer to the above questions, four major approaches to pursue AI have been established.
-These approaches can be summarized into a two by two table as shown in figure \ref{fig:twodimai}, where the row is ``Human'' vs.~``Rational'', and the column is ``Behaviour'' vs.~``Thought''.
+These approaches can be summarized into a two by two table as shown in Figure \ref{fig:twodimai}, where the row is ``Human'' vs.~``Rational'', and the column is ``Behaviour'' vs.~``Thought''.
 Positioning at the top right cell, the \textbf{rational agent approach} aims to build agent that perform mathematically perfect acts such that the best expected outcome can always be achieved.
@@
 For model diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the null hypothesis.
-The same assumed model, such as classical linear regression model, can be rejected by many reasons with residual plot, including nonlinearity and heteroskedasticity as shown in figure \ref{fig:nonlinearityheter}.
+The same assumed model, such as classical linear regression model, can be rejected by many reasons with residual plot, including nonlinearity and heteroskedasticity as shown in Figure \ref{fig:nonlinearityheter}.
 
@@
 
-From the figure \ref{fig:cubic1}, it can be observed that with fixed \(\sigma\) and \(c\), \(a\) and \(b\) are controlling the 2D projection of a hypersurface, and seemingly performing some rotations along different axes.
+From the Figure \ref{fig:cubic1}, it can be observed that with fixed \(\sigma\) and \(c\), \(a\) and \(b\) are controlling the 2D projection of a hypersurface, and seemingly performing some rotations along different axes.
 Figure \ref{fig:cubic2} shows that with fixed \(a\), \(b\) and \(\sigma\), \(c\) is controlling the contribution of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) to \(\boldsymbol{Y}\).
 If we move \(c\) toward \(0\) or \(2\), one variable will donminate another, which will mitigate some joint effects and resemble a typical cubic function.
-In figure \ref{fig:cubic3}, \(a\), \(b\) and \(c\) is held constant, and \(\sigma\) is controlling the noises around the expected values.
+In Figure \ref{fig:cubic3}, \(a\), \(b\) and \(c\) is held constant, and \(\sigma\) is controlling the noises around the expected values.
 As \(\sigma\) decreases, the underlying shape shows up.
@@
 
-The cubic model involves the use of both \(\boldsymbol{X}\) and \(\boldsymbol{Z}\).
-In the simulation, \(X_i\), \(i=1,...,n\), had equal chance to follow one of the following distributions: \(N(0, 0.09)\), \(U(-1, 1)\), \(Lognormal(0,0.36)\) and \(-Lognormal(0,0.36)\).
-
-Normal and uniform distributions are
+The cubic model involved the use of both \(\boldsymbol{X}\) and \(\boldsymbol{Z}\).
+In the simulation, \(X_i\), \(i=1,...,n\), had equal chance to follow one of the following distributions: \(N(0, 0.09)\), \(U(-1, 1)\), \(Lognormal(0,0.36)/3 - 1\) and \(-Lognormal(0,0.36)/3 + 1\).
+Uniform and normal distribution were symmetric and common.
+Adjusted lognormal distribution and adjusted negative lognormal distribution provided right-skewed and left-skewed density respectively.
+These distributions were chosen such that most the realizations will fall between \(-1\) and \(1\).
+
+The distribution of \(Z_i\), \(i=1,...,n\), had \(50\)\% chance to be a uniform distribution ranged from \(-1\) to \(1\), and \(50\)\% chance to be a discrete uniform distribution with \(z_n\) outcomes simulated from a uniform distribution ranged from \(-1\) to \(1\).
+\(z_n\) itself was a discrete uniform distribution with outcomes \(\{3,4,5,6,7,8,9,10\}\), which defined the number of possible values \(Z_i\) could take.
+As shown in Figure \ref{fig:discreteex}, this setup would create discreteness in residual plot, which could enrich the pool of visual patterns.
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/discreteex-1.pdf}
+\caption{\label{fig:discreteex}Discreteness in residuals created by using a discrete uniform random variable as one of the regressors in a cubic model.
+For each residual plot of the lineup, there are three clusters because the number of possible values the regressor can take is three.}
+\end{figure}
+
+\(\boldsymbol{X}\) used in the heteroskedasticity model was a combination of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) used in the cubic model.
+It could be one of the five distributions mentioned above - normal distribution, uniform distribution, adjusted lognormal distribution, adjusted negative lognormal distribution and discrete uniform distribution.
+
+\hypertarget{experiment-i}{%
+\section{Experiment I}\label{experiment-i}}
+
+\hypertarget{experiment-ii}{%
+\section{Experiment II}\label{experiment-ii}}
 

### 2022-02-23 Thesis/01-chap1.Rmd
add regressor distribution

@@
 Based on the answer to the above questions, four major approaches to pursue AI have been established.
-These approaches can be summarized into a two by two table as shown in figure \@ref(fig:twodimai), where the row is "Human" vs.
+These approaches can be summarized into a two by two table as shown in Figure \@ref(fig:twodimai), where the row is "Human" vs.
 "Rational", and the column is "Behaviour" vs.
@@
 For model diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the null hypothesis.
-The same assumed model, such as classical linear regression model, can be rejected by many reasons with residual plot, including nonlinearity and heteroskedasticity as shown in figure \@ref(fig:nonlinearityheter).
+The same assumed model, such as classical linear regression model, can be rejected by many reasons with residual plot, including nonlinearity and heteroskedasticity as shown in Figure \@ref(fig:nonlinearityheter).
 
@@
 
-From the figure \@ref(fig:cubic1), it can be observed that with fixed $\sigma$ and $c$, $a$ and $b$ are controlling the 2D projection of a hypersurface, and seemingly performing some rotations along different axes.
+From the Figure \@ref(fig:cubic1), it can be observed that with fixed $\sigma$ and $c$, $a$ and $b$ are controlling the 2D projection of a hypersurface, and seemingly performing some rotations along different axes.
 Figure \@ref(fig:cubic2) shows that with fixed $a$, $b$ and $\sigma$, $c$ is controlling the contribution of $\boldsymbol{X}$ and $\boldsymbol{Z}$ to $\boldsymbol{Y}$.
 If we move $c$ toward $0$ or $2$, one variable will donminate another, which will mitigate some joint effects and resemble a typical cubic function.
-In figure \@ref(fig:cubic3), $a$, $b$ and $c$ is held constant, and $\sigma$ is controlling the noises around the expected values.
+In Figure \@ref(fig:cubic3), $a$, $b$ and $c$ is held constant, and $\sigma$ is controlling the noises around the expected values.
 As $\sigma$ decreases, the underlying shape shows up.
@@
 
-The cubic model involves the use of both $\boldsymbol{X}$ and $\boldsymbol{Z}$.
-In the simulation, $X_i$, $i=1,...,n$, had equal chance to follow one of the following distributions: $N(0, 0.09)$, $U(-1, 1)$, $Lognormal(0,0.36)$ and $-Lognormal(0,0.36)$.
+The cubic model involved the use of both $\boldsymbol{X}$ and $\boldsymbol{Z}$.
+In the simulation, $X_i$, $i=1,...,n$, had equal chance to follow one of the following distributions: $N(0, 0.09)$, $U(-1, 1)$, $Lognormal(0,0.36)/3 - 1$ and $-Lognormal(0,0.36)/3 + 1$.
+Uniform and normal distribution were symmetric and common.
+Adjusted lognormal distribution and adjusted negative lognormal distribution provided right-skewed and left-skewed density respectively.
+These distributions were chosen such that most the realizations will fall between $-1$ and $1$.
 
-Normal and uniform distributions are
+The distribution of $Z_i$, $i=1,...,n$, had $50$% chance to be a uniform distribution ranged from $-1$ to $1$, and $50$% chance to be a discrete uniform distribution with $z_n$ outcomes simulated from a uniform distribution ranged from $-1$ to $1$.
+$z_n$ itself was a discrete uniform distribution with outcomes $\{3,4,5,6,7,8,9,10\}$, which defined the number of possible values $Z_i$ could take.
+As shown in Figure \@ref(fig:discreteex), this setup would create discreteness in residual plot, which could enrich the pool of visual patterns.
+
+```{r discreteex, fig.cap="Discreteness in residuals created by using a discrete uniform random variable as one of the regressors in a cubic model.
+For each residual plot of the lineup, there are three clusters because the number of possible values the regressor can take is three."}
+set.seed(10093)
+mod <- cubic_model(a = 2, b = -2, c = 1, z = rand_uniform_d(-1, 1, k = 3))
+mod$plot_lineup(mod$gen_lineup(500), theme = theme_light(), alpha = 0.3)
+```
+
+$\boldsymbol{X}$ used in the heteroskedasticity model was a combination of $\boldsymbol{X}$ and $\boldsymbol{Z}$ used in the cubic model.
+It could be one of the five distributions mentioned above - normal distribution, uniform distribution, adjusted lognormal distribution, adjusted negative lognormal distribution and discrete uniform distribution.
+
+## Experiment I
+
+## Experiment II
 

### 2022-02-23 Thesis/01-chap1.Rmd
add sample size calc

@@
 
+## Sample Size Calculation
+
+In this study, multiple selection was allowed for the evaluation of a lineup.
+Given our desired significance level was $\alpha = 0.05$ and the number of plots in a lineup was $20$, a single visual test procedure needed to involve multiple evaluations.
+Otherwise, there would be a high chance that the $p$-value of the test was greater than $0.05$ in spite of the actual data plot was selected.
+The number of plots selected by a user depended on the user and the difficulty of the lineup.
+But according to our pilot study ran in a small group, the number would be generally smaller than $5$.
+Hence, a visual test consisted of $5$ lineups were sufficient to yield $p$-value smaller than $0.004$.
+Thus, each lineup would be replicated for $5$ times and evaluated by different participants.
+However, our study website was designed such that any set of lineups would be shown to only one person only once.
+A set of lineups consists 20 lineups generated before the start of the study.
+Further, whether the participant would finish the experiment was unpredictable.
+Therefore, some lineups only had $2$ evaluations but some had as large as $8$ evaluations.
+This slightly affected the estimate of the power of the visual test, which would be discussed in the (section here).
+
+Besides, the sample size calculation was based on the difficulty of the lineup.
+With the data collected from the pilot study, two generalized linear model (GLM) model was developed to describe the relationship between the effect size and the detection rate.
+
 ## Experiment I

### 2022-02-23 Thesis/_book/thesis.tex
add sample size calc

@@
 
+\hypertarget{sample-size-calculation}{%
+\section{Sample Size Calculation}\label{sample-size-calculation}}
+
+In this study, multiple selection was allowed for the evaluation of a lineup.
+Given our desired significance level was \(\alpha = 0.05\) and the number of plots in a lineup was \(20\), a single visual test procedure needed to involve multiple evaluations.
+Otherwise, there would be a high chance that the \(p\)-value of the test was greater than \(0.05\) in spite of the actual data plot was selected.
+The number of plots selected by a user depended on the user and the difficulty of the lineup.
+But according to our pilot study ran in a small group, the number would be generally smaller than \(5\).
+Hence, a visual test consisted of \(5\) lineups were sufficient to yield \(p\)-value smaller than \(0.004\).
+Thus, each lineup would be replicated for \(5\) times and evaluated by different participants.
+However, our study website was designed such that any set of lineups would be shown to only one person only once.
+A set of lineups consists 20 lineups generated before the start of the study.
+Further, whether the participant would finish the experiment was unpredictable.
+Therefore, some lineups only had \(2\) evaluations but some had as large as \(8\) evaluations.
+This slightly affected the estimate of the power of the visual test, which would be discussed in the (section here).
+
+Besides, the sample size calculation was based on the difficulty of the lineup.
+With the data collected from the pilot study, two generalized linear model (GLM) model was developed to describe the relationship between the effect size and the detection rate.
+
 \hypertarget{experiment-i}{%

### 2022-02-26 Thesis/_book/thesis.tex
add ex1 and ex2

--- 2022-02-23 Thesis/_book/thesis.tex
@@
 
-For the heteroskedasticity model, the conventional test we used was Breusch--Pagan test \autocite{breusch_simple_1979}, which tested whether the variance of the error terms of the regression is dependent on the regressors.
-\textcite{majumder_validation_2013} suggested that visual test is not expected to perform equally well as conventional test especially when there exists a exact conventional test.
+For the heteroskedasticity model, the conventional test we used was Breusch--Pagan test \autocite{breusch_simple_1979}, which tested whether the variance of the error terms of the regression is dependent on the regressors with the auxiliary regression equation \[\boldsymbol{e}^2 = \gamma_0 + \gamma_1 \boldsymbol{X} + \gamma_2 \boldsymbol{X}^2 + \boldsymbol{v}.\] \textcite{majumder_validation_2013} suggested that visual test is not expected to perform equally well as conventional test especially when there exists a exact conventional test.
 However, in contrast to the F-test used in the cubic model, Breusch--Pagan test is an approximate test.
@@
 
-\hypertarget{experiment-i}{%
-\section{Experiment I}\label{experiment-i}}
-
-\hypertarget{experiment-ii}{%
-\section{Experiment II}\label{experiment-ii}}
+\hypertarget{experiment-i-and-ii}{%
+\section{Experiment I and II}\label{experiment-i-and-ii}}
+
+In this study, multiple selection was allowed for the evaluation of a lineup.
+Given our desired significance level was \(\alpha = 0.05\) and the number of plots in a lineup was \(20\), a single visual test procedure needed to involve multiple evaluations.
+Otherwise, the \(p\)-value of the test would be greater than \(0.05\) in spite of the actual data plot being selected.
+The number of plots selected by a user depended on the user and the difficulty of the lineup.
+According to the pilot study evaluated by \(10\) faculty members of Department of Econometrics and Business Analytics, Monash University, the number of selections would be generally smaller than \(4\), which suggested a visual test consisted of \(5\) evaluations were sufficient to yield \(p\)-value smaller than \(0.05\) with at least three detections.
+Thus, each lineup was replicated for five times and evaluated by different participants.
+However, our study website was designed such that any set of lineups consisting of 20 lineups generated beforehand would be shown to only one person only once.
+And whether the participant would finish the experiment was unpredictable.
+Therefore, some lineups had insufficient number of evaluations, but some had more evaluations than expected.
+This slightly affected the estimate of the power of the visual test, which would be discussed in Section \ref{p-value-for-visual-test-allowed-for-multiple-selections}.
+
+We called the lineup being detected if the actual data plot was one of the selections of an evaluation of a lineup.
+The sample size calculation was based on the detection rate of the lineup, which was closely related to the difficulty level of the lineup.
+With the data collected from the pilot study, two generalized linear model (GLM) models given in Table \ref{tab:pilotglm} and Figure \ref{fig:pilotglm} were developed to describe the relationship between the natural logarithm of effect size and the detection rate for the cubic model and the heteroskedasticity model.
+
+\begin{table}[!htbp] \centering 
+  \caption{GLM result 1} 
+  \label{tab:pilotglm} 
+\begin{tabular}{@{\extracolsep{5pt}}lcc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{2}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-3} 
+\\[-1.8ex] & \multicolumn{2}{c}{detect} \\ 
+ & Cubic & Heteroskedasticity \\ 
+\\[-1.8ex] & (1) & (2)\\ 
+\hline \\[-1.8ex] 
+ log(effect\_size) & 0.611$^{***}$ & 0.737$^{***}$ \\ 
+  & (0.098) & (0.137) \\ 
+  & & \\ 
+ Constant & 0.457$^{***}$ & $-$2.402$^{***}$ \\ 
+  & (0.165) & (0.645) \\ 
+  & & \\ 
+\hline \\[-1.8ex] 
+Observations & 246 & 289 \\ 
+Log Likelihood & $-$122.305 & $-$147.971 \\ 
+Akaike Inf. Crit. & 248.610 & 299.943 \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+\end{tabular} 
+\end{table}
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/pilotglm-1.pdf}
+\caption{\label{fig:pilotglm}pilot}
+\end{figure}
+
+Lineups were classified into three categories - easy, medium and hard.
+Easy lineups were those of which predicted detection rates were higher than \(80\)\%, while medium lineups were between \(40\)\% to \(80\)\% and hard lineups were less than \(40\)\%.
+
+For the first experiment, we would like to have at least 20 detection with chance greater than 99.99\% in each category.
+Thus, we prepared \(34\) easy lineups, \(60\) medium lineups and \(194\) hard lineups with five replications for both cubic and heteroskedasticity model.
+Every participant would get at least \(12\) easy or medium lineups.
+Two out of \(10\) extremely lineups with predicted detection rate over \(90\)\% were randomly given to every participant as attention checks.
+We initially planned to recruit \((34+60+194) \times 2 \times 5 / (20-2) = 160\) participants but eventually decided to only recruit \(20\) participants because it was the first time we launched an experiment on Prolific where factors like procedure of recruiting participants, payment method, and quality of participants were unclear at the moment.
+As a consequence, most of the lineups had only two evaluations and not all the lineups were used.
+However, data collected from the first experiment was sufficient for building a more accurate GLM model to predict detection rate for the second experiment.
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/e1glmplot-1.pdf}
+\caption{\label{fig:e1glmplot}test}
+\end{figure}
+
+\begin{table}[!htbp] \centering 
+  \caption{GLM result 2} 
+  \label{tab:e1glm} 
+\begin{tabular}{@{\extracolsep{5pt}}lcc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{2}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-3} 
+\\[-1.8ex] & \multicolumn{2}{c}{detect} \\ 
+ & Cubic & Heteroskedasticity \\ 
+\\[-1.8ex] & (1) & (2)\\ 
+\hline \\[-1.8ex] 
+ I(log(effect\_size)) & 0.818$^{***}$ & 0.710$^{***}$ \\ 
+  & (0.113) & (0.107) \\ 
+  & & \\ 
+ Constant & $-$1.649$^{***}$ & $-$3.425$^{***}$ \\ 
+  & (0.254) & (0.454) \\ 
+  & & \\ 
+\hline \\[-1.8ex] 
+Observations & 201 & 199 \\ 
+Log Likelihood & $-$68.678 & $-$89.985 \\ 
+Akaike Inf. Crit. & 141.357 & 183.971 \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+\end{tabular} 
+\end{table}
+
+It was also found that the first experiment might be too difficult for participants as the detection rate was lower than what we estimated from the pilot study, especially for the heteroskedasticity model.
+Thus, for the second experiments, we increased the proportion of easy lineups.
+In addition, due to budget constraint, the maximum number of participants was \(50\).
+With the new GLM model given in Table \ref{tab:e1glm} and Figure \ref{fig:e1glmplot} fitted with data collected from the first experiment, \(40\) easy lineups, \(25\) medium lineups and \(15\) hard lineups were simulated for both cubic and heteroskedasticity model.
+\(49\) participants were recruited but \(6\) were withdrew.
+
+\hypertarget{p-value-for-visual-test-allowed-for-multiple-selections}{%
+\section{P-value for Visual Test Allowed for Multiple Selections}\label{p-value-for-visual-test-allowed-for-multiple-selections}}
+
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+(estimate of visual power)
+
+(power against different parameters)
+
+(power against conventional p-value)
+
+(conventional power)
+
+(visual p-value vs.~conventional p-value)
+
+(response time)
+
+(confidence level)
+
+(relative freq of picks)
 

### 2022-02-26 Thesis/01-chap1.Rmd
add ex1 and ex2

--- 2022-02-23 Thesis/01-chap1.Rmd
@@
 
-For the heteroskedasticity model, the conventional test we used was Breusch–Pagan test [@breusch_simple_1979], which tested whether the variance of the error terms of the regression is dependent on the regressors.
-@majumder_validation_2013 suggested that visual test is not expected to perform equally well as conventional test especially when there exists a exact conventional test.
+For the heteroskedasticity model, the conventional test we used was Breusch–Pagan test [@breusch_simple_1979], which tested whether the variance of the error terms of the regression is dependent on the regressors with the auxiliary regression equation $$\boldsymbol{e}^2 = \gamma_0 + \gamma_1 \boldsymbol{X} + \gamma_2 \boldsymbol{X}^2 + \boldsymbol{v}.$$ @majumder_validation_2013 suggested that visual test is not expected to perform equally well as conventional test especially when there exists a exact conventional test.
 However, in contrast to the F-test used in the cubic model, Breusch–Pagan test is an approximate test.
@@
 
-## Experiment I
+## Experiment I and II
 
-## Experiment II
+In this study, multiple selection was allowed for the evaluation of a lineup.
+Given our desired significance level was $\alpha = 0.05$ and the number of plots in a lineup was $20$, a single visual test procedure needed to involve multiple evaluations.
+Otherwise, the $p$-value of the test would be greater than $0.05$ in spite of the actual data plot being selected.
+The number of plots selected by a user depended on the user and the difficulty of the lineup.
+According to the pilot study evaluated by  $10$ faculty members of Department of Econometrics and Business Analytics, Monash University, the number of selections would be generally smaller than $4$, which suggested a visual test consisted of $5$ evaluations were sufficient to yield $p$-value smaller than $0.05$ with at least three detections.
+Thus, each lineup was replicated for five times and evaluated by different participants.
+However, our study website was designed such that any set of lineups consisting of 20 lineups generated beforehand would be shown to only one person only once.
+And whether the participant would finish the experiment was unpredictable.
+Therefore, some lineups had insufficient number of evaluations, but some had more evaluations than expected.
+This slightly affected the estimate of the power of the visual test, which would be discussed in Section \@ref(p-value-for-visual-test-allowed-for-multiple-selections).
+
+We called the lineup being detected if the actual data plot was one of the selections of an evaluation of a lineup.
+The sample size calculation was based on the detection rate of the lineup, which was closely related to the difficulty level of the lineup.
+With the data collected from the pilot study, two generalized linear model (GLM) models given in Table \@ref(tab:pilotglm) and Figure \@ref(fig:pilotglm) were developed to describe the relationship between the natural logarithm of effect size and the detection rate for the cubic model and the heteroskedasticity model.
+
+```{r }
+file_names <- list.files(here::here("data/pilot_study"))
+
+library(tidyverse)
+result_high <- NULL
+for (file_name in file_names[grep("oct" , file_names)]) {
+  dat <- read_csv(here::here("data/pilot_study", file_name))
+  if (is.null(result_high)) {
+    result_high <- dat
+  } else {
+    result_high <- bind_rows(result_high, dat)
+  }
+}
+
+result_high %>%
+  mutate(detect = as.numeric(answer)) %>%
+  ggplot() +
+  ggbeeswarm::geom_quasirandom(aes(log(effect_size),
+                                   detect),
+                               groupOnX = FALSE) +
+  geom_smooth(aes(log(effect_size), detect, group = 1),
+              method = "glm", method.args = list(family = binomial)) +
+  xlab("Log of effect size") +
+  ylab("Detection rate") +
+  ggtitle("Cubic Model") -> p1
+
+mod1 <- glm(detect ~ log(effect_size), data = mutate(result_high, detect = as.numeric(answer), log_effect_size = log(effect_size)), family = binomial)
+
+```
+
+```{r}
+file_names <- list.files(here::here("data/pilot_study"))
+
+library(tidyverse)
+result_heter <- NULL
+for (file_name in file_names[grep("nov" , file_names)]) {
+  dat <- read_csv(here::here("data/pilot_study", file_name))
+  if (is.null(result_heter)) {
+    result_heter <- dat
+  } else {
+    result_heter <- bind_rows(result_heter, dat)
+  }
+}
+
+result_heter %>%
+  mutate(detect = as.numeric(answer)) %>%
+  ggplot() +
+  ggbeeswarm::geom_quasirandom(aes(log(effect_size),
+                                   detect),
+                               groupOnX = FALSE) +
+  geom_smooth(aes(log(effect_size), detect, group = 1),
+              method = "glm", method.args = list(family = binomial)) +
+  xlab("Log of effect size") +
+  ylab("Detection rate") +
+  ggtitle("Heteroskedasticity Model") -> p2
+```
+
+```{r results='asis'}
+library(stargazer)
+mod2 <- glm(detect ~ log(effect_size), data = mutate(result_heter, detect = as.numeric(answer), log_effect_size = log(effect_size)), family = binomial)
+stargazer(mod1, mod2,
+          header = FALSE,
+          label = "tab:pilotglm",
+          column.labels = c("Cubic", "Heteroskedasticity"),
+          title = "GLM result 1",
+          style = "default")
+
+```
+
+```{r pilotglm, fig.cap="pilot"}
+list(p1, p2) %>% patchwork::wrap_plots(ncol = 2)
+```
+
+Lineups were classified into three categories - easy, medium and hard.
+Easy lineups were those of which predicted detection rates were higher than $80$%, while medium lineups were between $40$% to $80$% and hard lineups were less than $40$%.
+
+For the first experiment, we would like to have at least 20 detection with chance greater than 99.99% in each category.
+Thus, we prepared $34$ easy lineups, $60$ medium lineups and $194$ hard lineups with five replications for both cubic and heteroskedasticity model.
+Every participant would get at least $12$ easy or medium lineups.
+Two out of $10$ extremely lineups with predicted detection rate over $90$% were randomly given to every participant as attention checks.
+We initially planned to recruit $(34+60+194) \times 2 \times 5 / (20-2) = 160$ participants but eventually decided to only recruit $20$ participants because it was the first time we launched an experiment on Prolific where factors like procedure of recruiting participants, payment method, and quality of participants were unclear at the moment.
+As a consequence, most of the lineups had only two evaluations and not all the lineups were used.
+However, data collected from the first experiment was sufficient for building a more accurate GLM model to predict detection rate for the second experiment.
+
+```{r}
+process_small_responses <- function(k, lineup_dat, lineup_ord, survey_folder) {
+
+  dat <- tibble(id = 1:length(unlist(lineup_ord)),
+                lineup_dat_id = unlist(lineup_ord)) %>%
+    mutate(set = (id-1) %/% k + 1) %>%
+    mutate(num = (id-1) %% k + 1)
+
+  user_info_dict <- list(c("18-24", "25-39", "40-54", "55-64", "65 or above"),
+                         c("High School or below", "Diploma and Bachelor Degree", "Honours Degree", "Masters Degree", "Doctoral Degree"),
+                         c("He", "She", "They", "Other"),
+                         c("Yes", "No"))
+
+  get_responses <- function(user_info_dict, uuid) {
+    tmp_survey <- jsonlite::fromJSON(here::here(glue::glue("{survey_folder}/{uuid}.txt")),
+                                     simplifyVector = FALSE)
+    response_time <- map_dbl(tmp_survey, ~.x$rt)
+    user_information <- str_split(tmp_survey[[3]]$response[[1]], ",")[[1]]
+
+    user_information[2:5] <- imap_chr(user_information[-1], ~user_info_dict[[.y]][as.integer(.x)])
+    names(user_information) <- c("prolific_id", "age_group", "education", "pronoun", "previous_experience")
+
+    lineup_respone <- str_split(map_chr(tmp_survey[-c(1,2,3,4)], ~.x$response$response), ",")
+    # handle selections
+    selections <- map_chr(lineup_respone, ~.x[1])
+    # handle reasons
+    reasons_dict <- c("Outlier(s)", "Cluster(s)", "Shape")
+
+    reasons <- map_chr(lineup_respone, ~if(.x[2] %in% c("1", "2", "3")) {
+      reasons_dict[as.integer(.x[2])]
+    } else {
+      .x[2]
+    })
+
+    # handle confidence
+    confidence_dict <- c("Not at all", "Slightly", "Moderately", "Very", "Extremely")
+    confidence <- map_chr(lineup_respone, ~if(.x[3] %in% c("1", "2", "3", "4", "5")) {
+      confidence_dict[as.integer(.x[3])]
+    } else {
+      .x[3]
+    })
+
+    tibble(page = 1:length(response_time),
+           response_time = response_time,
+           set = uuid,
+           num = c(rep(NA, 4), 1:k),
+           selection = c(rep(NA, 4), selections),
+           num_selection = c(rep(NA, 4), str_count(selections, "_") + 1),
+           reason = c(rep(NA, 4), reasons),
+           confidence = c(rep(NA, 4), confidence),
+           prolific_id = unname(user_information['prolific_id']),
+           age_group = unname(user_information['age_group']),
+           education = unname(user_information['education']),
+           pronoun = unname(user_information['pronoun']),
+           previous_experience = unname(user_information['previous_experience'])
+    ) %>%
+      mutate(reason = ifelse(selection == "NA", NA, reason)) %>%
+      mutate(confidence = ifelse(selection == "NA", NA, confidence)) %>%
+      mutate(selection = ifelse(selection == "NA", "0", selection)) %>%
+      mutate(num_selection = ifelse(selection == "0", 0, num_selection))
+  }
+
+  response_dat <- NULL
+  file_names <- list.files(here::here(survey_folder))
+
+  for (i in sort(as.integer(str_replace(file_names[grep("*.txt", file_names)], ".txt", "")))) {
+    if (is.null(response_dat)) {
+      response_dat <- get_responses(user_info_dict, i)
+    } else {
+      response_dat <- bind_rows(response_dat, get_responses(user_info_dict, i))
+    }
+  }
+
+  correct_or_not <- function(selection, ans) {
+    tmp <- imap_lgl(selection, ~ans[.y] %in% str_split(.x, "_")[[1]])
+    tmp[is.na(ans)] <- NA
+    tmp
+  }
+
+  extract_p_value <- function(lineup_id, lineup_dat, null = FALSE) {
+    if (is.na(lineup_id)) {
+      return(NA)
+    }
+
+    tmp <- lineup_dat[[lineup_id]]$data %>%
+      group_by(k) %>%
+      summarise(null = first(null), pvalue = first(pvalue)) %>%
+      ungroup()
+
+    data_p <- tmp$pvalue[tmp$null == FALSE]
+    null_min_p <- min(tmp$pvalue[tmp$null == TRUE])
+    if (null) {
+      return(null_min_p)
+    } else {
+      return(data_p)
+    }
+  }
+
+  response_dat %>%
+    rowwise() %>%
+    mutate(lineup_id = dat$lineup_dat_id[dat$set == set & dat$num == num][1]) %>%
+    mutate(metadata = list(lineup_dat[[lineup_id]]$metadata)) %>%
+    ungroup() %>%
+    unnest_wider(metadata) %>%
+    mutate(age_group = factor(age_group,
+                              levels = user_info_dict[[1]])) %>%
+    mutate(education = factor(education,
+                              levels = user_info_dict[[2]])) %>%
+    mutate(confidence = factor(confidence,
+                               levels = c("Not at all", "Slightly", "Moderately", "Very", "Extremely"))) %>%
+    mutate(detected = correct_or_not(selection, ans)) %>%
+    rowwise() %>%
+    mutate(data_p_value = extract_p_value(lineup_id, lineup_dat, null = FALSE)) %>%
+    mutate(null_min_p_value = extract_p_value(lineup_id, lineup_dat, null = TRUE)) %>%
+    ungroup()
+}
+```
+
+```{r}
+lineup_dat <- readRDS(here::here("data/small_study/all_data.RDS"))
+lineup_ord <- readRDS(here::here("data/small_study/lineup_order.RDS"))
+
+# Correct the effect size
+for (i in 1:length(lineup_dat)) {
+  if (lineup_dat[[i]]$metadata$type != "higher order") next
+
+  tmp_eff <- CUBIC_MODEL$effect_size(select(filter(lineup_dat[[i]]$data, null == FALSE), x, z), 
+                          a = lineup_dat[[i]]$metadata$a, 
+                          b = lineup_dat[[i]]$metadata$b,
+                          c = lineup_dat[[i]]$metadata$c,
+                          sigma = lineup_dat[[i]]$metadata$e_sigma)
+
+  lineup_dat[[i]]$metadata$effect_size <- tmp_eff 
+}
+
+small_survey_dat <- process_small_responses(20, 
+                                            lineup_dat = lineup_dat, 
+                                            lineup_ord = lineup_ord, 
+                                            survey_folder = "data/small_study/survey") %>%
+  filter(page > 4) %>%
+  mutate(difficulty = str_replace_all(name, "(high_|heter_|_\\d+)", "")) %>%
+  mutate(difficulty = factor(difficulty, levels = c("extremely_easy", "easy", "medium", "hard")))
+```
+
+```{r}
+small_survey_dat$detect <- small_survey_dat$detected
+
+e1_high <- glm(detect ~ I(log(effect_size)),
+                data = filter(small_survey_dat, type == "higher order"),
+                family = binomial)
+e1_heter <- glm(detect ~ I(log(effect_size)),
+                 data = filter(small_survey_dat, type == "heteroskedasticity"),
+                 family = binomial)
+```
+
+```{r e1glmplot, fig.cap="test"}
+filter(small_survey_dat, type == "higher order") %>%
+  mutate(detect = as.numeric(detect)) %>%
+  ggplot() +
+  ggbeeswarm::geom_quasirandom(aes(log(effect_size),
+                                   detect),
+                               groupOnX = FALSE) +
+  geom_smooth(aes(log(effect_size), detect, group = 1),
+              method = "glm", method.args = list(family = binomial)) +
+  xlab("Log of effect size") +
+  ylab("Detection rate") +
+  ggtitle("Cubic Model") -> p1
+
+filter(small_survey_dat, type == "heteroskedasticity") %>%
+  mutate(detect = as.numeric(detect)) %>%
+  ggplot() +
+  ggbeeswarm::geom_quasirandom(aes(log(effect_size),
+                                   detect),
+                               groupOnX = FALSE) +
+  geom_smooth(aes(log(effect_size), detect, group = 1),
+              method = "glm", method.args = list(family = binomial)) +
+  xlab("Log of effect size") +
+  ylab("Detection rate") +
+  ggtitle("Heteroskedasticity Model") -> p2
+
+list(p1, p2) %>% patchwork::wrap_plots(ncol = 2)
+```
+
+```{r results="asis"}
+stargazer(e1_high, e1_heter,
+          header = FALSE,
+          label = "tab:e1glm",
+          column.labels = c("Cubic", "Heteroskedasticity"),
+          title = "GLM result 2",
+          style = "default")
+```
+
+It was also found that the first experiment might be too difficult for participants as the detection rate was lower than what we estimated from the pilot study, especially for the heteroskedasticity model.
+Thus, for the second experiments, we increased the proportion of easy lineups.
+In addition, due to budget constraint, the maximum number of participants was $50$.
+With the new GLM model given in Table \@ref(tab:e1glm) and Figure \@ref(fig:e1glmplot) fitted with data collected from the first experiment, $40$ easy lineups, $25$ medium lineups and $15$ hard lineups were simulated for both cubic and heteroskedasticity model.
+$49$ participants were recruited but $6$ were withdrew.
+
+## P-value for Visual Test Allowed for Multiple Selections
+
+## Results
+
+```{r eval = FALSE}
+lineup_dat_big <- readRDS(here::here("data/big_study/all_data_big.RDS"))
+lineup_ord_big <- readRDS(here::here("data/big_study/lineup_order_big.RDS"))
+
+ori_eff <- c()
+current_eff <- c()
+prm_c <- c()
+
+# Correct the effect size
+for (i in 1:length(lineup_dat_big)) {
+  if (lineup_dat_big[[i]]$metadata$type != "higher order") next
+
+  tmp_eff <- CUBIC_MODEL$effect_size(select(filter(lineup_dat_big[[i]]$data, null == FALSE), x, z), 
+                          a = lineup_dat_big[[i]]$metadata$a, 
+                          b = lineup_dat_big[[i]]$metadata$b,
+                          c = lineup_dat_big[[i]]$metadata$c,
+                          sigma = lineup_dat_big[[i]]$metadata$e_sigma)
+
+  ori_eff <- c(ori_eff, lineup_dat_big[[i]]$metadata$effect_size)
+  current_eff <- c(current_eff, tmp_eff)
+  prm_c <- c(prm_c, lineup_dat_big[[i]]$metadata$c)
+
+  lineup_dat_big[[i]]$metadata$effect_size <- tmp_eff 
+}
+
+big_survey_dat <- process_small_responses(20, 
+                                            lineup_dat = lineup_dat_big, 
+                                            lineup_ord = lineup_ord_big, 
+                                            survey_folder = "data/big_study/survey") %>%
+  filter(page > 4) %>%
+  mutate(difficulty = str_replace_all(name, "(high_|heter_|_\\d+)", "")) %>%
+  mutate(difficulty = factor(difficulty, levels = c("extremely_easy", "easy", "medium", "hard")))
+```
+
+```{r eval = FALSE}
+small_survey_dat %>%
+  mutate(p_hat = predict(mod1, newdata = small_survey_dat, type = "response")) %>%
+  mutate(new_diff = ifelse(p_hat < 0.4, "hard", ifelse(p_hat > 0.8, "easy", "medium"))) %>%
+  filter(difficulty != "extremely_easy") %>%
+  mutate(me = new_diff == "easy" | new_diff == "medium") %>%
+  count(set, me) %>%
+  filter(me) %>%
+  ggplot() +
+  geom_histogram(aes(n))
+```
+
+```{r eval = FALSE}
+big_survey_dat %>%
+  mutate(p_hat = predict(e1_high, newdata = big_survey_dat, type = "response")) %>%
+  mutate(new_diff = ifelse(p_hat < 0.4, "hard", ifelse(p_hat > 0.8, "easy", "medium"))) %>%
+  filter(difficulty != "extremely_easy") %>%
+  mutate(me = new_diff == "easy" | new_diff == "medium") %>%
+  count(set, me) %>%
+  filter(me) %>%
+  ggplot() +
+  geom_histogram(aes(n))
+```
+
+```{r eval = FALSE}
+big_survey_dat %>%
+  mutate(p_hat = predict(e1_high, newdata = big_survey_dat, type = "response")) %>%
+  mutate(new_diff = ifelse(p_hat < 0.4, "hard", ifelse(p_hat > 0.8, "easy", "medium"))) %>%
+  filter(difficulty != "extremely_easy") %>%
+  group_by(lineup_id) %>%
+  slice_head() %>%
+  ungroup() %>%
+  count(new_diff)
+
+```
+
+(estimate of visual power)
+
+(power against different parameters)
+
+(power against conventional p-value)
+
+(conventional power)
+
+(visual p-value vs. conventional p-value)
+
+(response time)
+
+(confidence level)
+
+(relative freq of picks)
 

### 2022-02-27 Thesis/01-chap1.Rmd
finish exp design

@@
 Then, under the null hypothesis $X \sim \mathrm{Binom}_{K,1/m}$, and therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as
-
-$$P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},$$
+\begin{equation} \label{eq:pvaluesingle}
+P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},
+\end{equation}
+
 where $x$ is the realization of number of observers correctly picking the actual data plot [@majumder_validation_2013].
@@
 Therefore, some lineups had insufficient number of evaluations, but some had more evaluations than expected.
-This slightly affected the estimate of the power of the visual test, which would be discussed in Section \@ref(p-value-for-visual-test-allowed-for-multiple-selections).
+This slightly affected the estimate of the power of the visual test, which will be discussed in Section \@ref(p-value-and-power-estimation-of-visual-test-allowed-for-multiple-selections).
 
@@
 
-It was also found that the first experiment might be too difficult for participants as the detection rate was lower than what we estimated from the pilot study, especially for the heteroskedasticity model.
-Thus, for the second experiments, we increased the proportion of easy lineups.
+According to the new GLM model given in Table \@ref(tab:e1glm) and Figure \@ref(fig:e1glmplot), it was found that the first experiment might be too difficult for participants as the detection rate was lower than what we estimated from the pilot study, especially for the heteroskedasticity model.
+Thus, for the second experiments, we relaxed the sample size requirements and increased the proportion of easy lineups.
 In addition, due to budget constraint, the maximum number of participants was $50$.
-With the new GLM model given in Table \@ref(tab:e1glm) and Figure \@ref(fig:e1glmplot) fitted with data collected from the first experiment, $40$ easy lineups, $25$ medium lineups and $15$ hard lineups were simulated for both cubic and heteroskedasticity model.
+With the new GLM model fitted with data collected from the first experiment, $40$ easy lineups, $25$ medium lineups and $15$ hard lineups were simulated for both cubic and heteroskedasticity model in the second experiment.
 $49$ participants were recruited but $6$ were withdrew.
 
-## P-value for Visual Test Allowed for Multiple Selections
+## $P$-value and Power Estimation of Visual Test Allowed for Multiple Selections
+
+The $p$-value calculation for multiple selection is an extension of Equation (\ref{eq:pvaluesingle}).
+The assumptions about the independence still needed.
+Let $K$ be the number of independent evaluations of a visual test, $s_i$, $i = 1,...,K$ be the number of selections of the evaluations, and $X$ be the random variable denoting the number of detections.
+Then, the $p$-value of the visual test is given as:
+\begin{equation} \label{eq:pvaluemulti}
+P(X\geq x) = \sum_{j=x}^{K}Pr(j|s_1,...,s_K).
+\end{equation}
+The distribution of $X$ given $s_1,...s_K$ can not be derived trivially, as it is a sampling without replacement problem. In practice, this distribution can be approximated by computer simulation.
+
+Function `sim_dist()` from package `visage` is designed to approximate the distribution of number of detections of a lineup via Monte Carlo method.
+It takes the number of evaluations `n_eval`, the number of selections `n_sel`, and the number of plots in a lineup `n_plot` as inputs, then outputs a discrete distribution.
+
+```{r echo = TRUE}
+sim_dist(n_eval = 3, n_sel = c(2, 2, 3), n_plot = 20)
+```
+
+Function `calc_p_value()` from the same package calculates the $p$-value using the distribution returned by `sim_dist()`. It takes an additional argument `n_detect`, which is the number of detections.
+
+```{r echo = TRUE}
+calc_p_value(n_detect = 2, n_eval = 3, n_sel = c(2, 2, 3), n_plot = 20)
+```
+
+The above example shows that a visual test with two detections and three independent evaluations in which three observers select two, two, and three plots out of $20$ plots respectively yields a $p$-value smaller than $0.05$.
+
+Assume there is a visual test $V_K$, where $K$ denoting the number of evaluations.
+The corresponding  $p$-value of $V_K$ can be computed by using Equation (\ref{eq:pvaluemulti}).
+Meanwhile, if one evaluation is randomly deleted from $V_K$, the remaining evaluations can still be used to form another valid visual test $V_{K-1}$.
+In fact, considering all the possibilities, $K$ different outcomes for $V_{K-1}$ can be obtained.
+Since all outcomes occur with equal probability, the proportion of outcomes which reject the null hypothesis can be used as an estimate of the power of the visual test $V_{K-1}$.
+Similarly, if we would like to estimate the power of the visual test $V_{K-j}$ given the evaluations of $V_{K}$, for $j < K$, we could find all the possible combinations of $K$ elements, taken $k-j$ at a time to obtain $K\choose{K-j}$ different outcomes for $V_{K-j}$.
+Then, the estimated power is given as $R/{K\choose{K-j}}$, where $R$ is the number of outcomes which reject the null hypothesis.
 
@@
 
-```{r eval = FALSE}
+```{r}
 lineup_dat_big <- readRDS(here::here("data/big_study/all_data_big.RDS"))
@@
 
-```{r eval = FALSE}
-small_survey_dat %>%
-  mutate(p_hat = predict(mod1, newdata = small_survey_dat, type = "response")) %>%
-  mutate(new_diff = ifelse(p_hat < 0.4, "hard", ifelse(p_hat > 0.8, "easy", "medium"))) %>%
-  filter(difficulty != "extremely_easy") %>%
-  mutate(me = new_diff == "easy" | new_diff == "medium") %>%
-  count(set, me) %>%
-  filter(me) %>%
-  ggplot() +
-  geom_histogram(aes(n))
-```
-
-```{r eval = FALSE}
-big_survey_dat %>%
-  mutate(p_hat = predict(e1_high, newdata = big_survey_dat, type = "response")) %>%
-  mutate(new_diff = ifelse(p_hat < 0.4, "hard", ifelse(p_hat > 0.8, "easy", "medium"))) %>%
-  filter(difficulty != "extremely_easy") %>%
-  mutate(me = new_diff == "easy" | new_diff == "medium") %>%
-  count(set, me) %>%
-  filter(me) %>%
-  ggplot() +
-  geom_histogram(aes(n))
-```
-
-```{r eval = FALSE}
-big_survey_dat %>%
-  mutate(p_hat = predict(e1_high, newdata = big_survey_dat, type = "response")) %>%
-  mutate(new_diff = ifelse(p_hat < 0.4, "hard", ifelse(p_hat > 0.8, "easy", "medium"))) %>%
-  filter(difficulty != "extremely_easy") %>%
-  group_by(lineup_id) %>%
-  slice_head() %>%
-  ungroup() %>%
-  count(new_diff)
-
-```
+### Data cleaning
 

### 2022-02-27 Thesis/_book/thesis.tex
finish exp design

@@
 Then, under the null hypothesis \(X \sim \mathrm{Binom}_{K,1/m}\), and therefore, the \(p\)-value of a lineup of size \(m\) evaluated by \(K\) observer is given as
-
-\[P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},\]
+\begin{equation} \label{eq:pvaluesingle}
+P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},
+\end{equation}
+
 where \(x\) is the realization of number of observers correctly picking the actual data plot \autocite{majumder_validation_2013}.
@@
 Therefore, some lineups had insufficient number of evaluations, but some had more evaluations than expected.
-This slightly affected the estimate of the power of the visual test, which would be discussed in Section \ref{p-value-for-visual-test-allowed-for-multiple-selections}.
+This slightly affected the estimate of the power of the visual test, which will be discussed in Section \ref{p-value-and-power-estimation-of-visual-test-allowed-for-multiple-selections}.
 
@@
 
-It was also found that the first experiment might be too difficult for participants as the detection rate was lower than what we estimated from the pilot study, especially for the heteroskedasticity model.
-Thus, for the second experiments, we increased the proportion of easy lineups.
+According to the new GLM model given in Table \ref{tab:e1glm} and Figure \ref{fig:e1glmplot}, it was found that the first experiment might be too difficult for participants as the detection rate was lower than what we estimated from the pilot study, especially for the heteroskedasticity model.
+Thus, for the second experiments, we relaxed the sample size requirements and increased the proportion of easy lineups.
 In addition, due to budget constraint, the maximum number of participants was \(50\).
-With the new GLM model given in Table \ref{tab:e1glm} and Figure \ref{fig:e1glmplot} fitted with data collected from the first experiment, \(40\) easy lineups, \(25\) medium lineups and \(15\) hard lineups were simulated for both cubic and heteroskedasticity model.
+With the new GLM model fitted with data collected from the first experiment, \(40\) easy lineups, \(25\) medium lineups and \(15\) hard lineups were simulated for both cubic and heteroskedasticity model in the second experiment.
 \(49\) participants were recruited but \(6\) were withdrew.
 
-\hypertarget{p-value-for-visual-test-allowed-for-multiple-selections}{%
-\section{P-value for Visual Test Allowed for Multiple Selections}\label{p-value-for-visual-test-allowed-for-multiple-selections}}
+\hypertarget{p-value-and-power-estimation-of-visual-test-allowed-for-multiple-selections}{%
+\section{\texorpdfstring{\(P\)-value and Power Estimation of Visual Test Allowed for Multiple Selections}{P-value and Power Estimation of Visual Test Allowed for Multiple Selections}}\label{p-value-and-power-estimation-of-visual-test-allowed-for-multiple-selections}}
+
+The \(p\)-value calculation for multiple selection is an extension of Equation (\ref{eq:pvaluesingle}).
+The assumptions about the independence still needed.
+Let \(K\) be the number of independent evaluations of a visual test, \(s_i\), \(i = 1,...,K\) be the number of selections of the evaluations, and \(X\) be the random variable denoting the number of detections.
+Then, the \(p\)-value of the visual test is given as:
+\begin{equation} \label{eq:pvaluemulti}
+P(X\geq x) = \sum_{j=x}^{K}Pr(j|s_1,...,s_K).
+\end{equation}
+The distribution of \(X\) given \(s_1,...s_K\) can not be derived trivially, as it is a sampling without replacement problem. In practice, this distribution can be approximated by computer simulation.
+
+Function \texttt{sim\_dist()} from package \texttt{visage} is designed to approximate the distribution of number of detections of a lineup via Monte Carlo method.
+It takes the number of evaluations \texttt{n\_eval}, the number of selections \texttt{n\_sel}, and the number of plots in a lineup \texttt{n\_plot} as inputs, then outputs a discrete distribution.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{sim\_dist}\NormalTok{(}\AttributeTok{n\_eval =} \DecValTok{3}\NormalTok{, }\AttributeTok{n\_sel =} \FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{2}\NormalTok{, }\DecValTok{3}\NormalTok{), }\AttributeTok{n\_plot =} \DecValTok{20}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+##       0       1       2       3 
+## 0.70182 0.25202 0.04346 0.00270
+\end{verbatim}
+
+Function \texttt{calc\_p\_value()} from the same package calculates the \(p\)-value using the distribution returned by \texttt{sim\_dist()}.
+It takes an additional argument \texttt{n\_detect}, which is the number of detections.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{calc\_p\_value}\NormalTok{(}\AttributeTok{n\_detect =} \DecValTok{2}\NormalTok{, }\AttributeTok{n\_eval =} \DecValTok{3}\NormalTok{, }\AttributeTok{n\_sel =} \FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{2}\NormalTok{, }\DecValTok{3}\NormalTok{), }\AttributeTok{n\_plot =} \DecValTok{20}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## [1] 0.04688
+\end{verbatim}
+
+The above example shows that a visual test with two detections and three independent evaluations in which three observers select two, two, and three plots out of \(20\) plots respectively yields a \(p\)-value smaller than \(0.05\).
+
+Assume there is a visual test \(V_K\), where \(K\) denoting the number of evaluations.
+The corresponding \(p\)-value of \(V_K\) can be computed by using Equation (\ref{eq:pvaluemulti}).
+Meanwhile, if one evaluation is randomly deleted from \(V_K\), the remaining evaluations can still be used to form another valid visual test \(V_{K-1}\).
+In fact, considering all the possibilities, \(K\) different outcomes for \(V_{K-1}\) can be obtained.
+Since all outcomes occur with equal probability, the proportion of outcomes which reject the null hypothesis can be used as an estimate of the power of the visual test \(V_{K-1}\).
+Similarly, if we would like to estimate the power of the visual test \(V_{K-j}\) given the evaluations of \(V_{K}\), for \(j < K\), we could find all the possible combinations of \(K\) elements, taken \(k-j\) at a time to obtain \(K\choose{K-j}\) different outcomes for \(V_{K-j}\).
+Then, the estimated power is given as \(R/{K\choose{K-j}}\), where \(R\) is the number of outcomes which reject the null hypothesis.
 
@@
 \section{Results}\label{results}}
+
+\hypertarget{data-cleaning}{%
+\subsection{Data cleaning}\label{data-cleaning}}
 

### 2022-02-27 Thesis/01-chap1.Rmd
add comb

@@
 
+Function `calc_p_value_comb()` from the package `visage` can be used to compute $p$-values of $V_{K-j}$.
+The first argument is `detected`, which needs to be a vector of Boolean values denoting whether the observer detects the actual data plot.
+Desired number of evaluations and number of selections need to be provided via argument `n_eval` and `n_sel`.
+
+For example, given a visual test $V_3$ where the first observer selects one plot and gets correct, the second observer selects one plot but misses, and the third observer selects two plots and gets correct.
+The $p$-value of all possible outcomes of $V_2$ can be computed using the following code:
+
+```{r}
+set.seed(10086)
+```
+
+```{r echo = TRUE}
+calc_p_value_comb(detected = c(TRUE, FALSE, TRUE), 
+                  n_eval = 2, 
+                  n_sel = c(1, 1, 2))
+```
+
+`calc_p_value_comb()` returns a vector of $p$-value along with an attribute which is a matrix representing elements being used by the specific outcome.
+Elements in the first column indicates that the first observer and the second observer are involved in the visual test, and the corresponding $p$-value is $0.09352$.
+
 ## Results
@@
 
+### Power Estimation
+
+###
+
 (estimate of visual power)

### 2022-02-27 Thesis/_book/thesis.tex
add comb

@@
 
+Function \texttt{calc\_p\_value\_comb()} from the package \texttt{visage} can be used to compute \(p\)-values of \(V_{K-j}\).
+The first argument is \texttt{detected}, which needs to be a vector of Boolean values denoting whether the observer detects the actual data plot.
+Desired number of evaluations and number of selections need to be provided via argument \texttt{n\_eval} and \texttt{n\_sel}.
+
+For example, given a visual test \(V_3\) where the first observer selects one plot and gets correct, the second observer selects one plot but misses, and the third observer selects two plots and gets correct.
+The \(p\)-value of all possible outcomes of \(V_2\) can be computed using the following code:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{calc\_p\_value\_comb}\NormalTok{(}\AttributeTok{detected =} \FunctionTok{c}\NormalTok{(}\ConstantTok{TRUE}\NormalTok{, }\ConstantTok{FALSE}\NormalTok{, }\ConstantTok{TRUE}\NormalTok{), }
+                  \AttributeTok{n\_eval =} \DecValTok{2}\NormalTok{, }
+                  \AttributeTok{n\_sel =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## [1] 0.09352 0.00624 0.14454
+## attr(,"combinations")
+##      [,1] [,2] [,3]
+## [1,]    1    1    2
+## [2,]    2    3    3
+\end{verbatim}
+
+\texttt{calc\_p\_value\_comb()} returns a vector of \(p\)-value along with an attribute which is a matrix representing elements being used by the specific outcome.
+Elements in the first column indicates that the first observer and the second observer are involved in the visual test, and the corresponding \(p\)-value is \(0.09352\).
+
 \hypertarget{results}{%
@@
 \subsection{Data cleaning}\label{data-cleaning}}
+
+\hypertarget{power-estimation}{%
+\subsection{Power Estimation}\label{power-estimation}}
+
+\hypertarget{section}{%
+\subsection{}\label{section}}
 

### 2022-03-01 Thesis/01-chap1.Rmd
update result

@@
 The sample size calculation was based on the detection rate of the lineup, which was closely related to the difficulty level of the lineup.
-With the data collected from the pilot study, two generalized linear model (GLM) models given in Table \@ref(tab:pilotglm) and Figure \@ref(fig:pilotglm) were developed to describe the relationship between the natural logarithm of effect size and the detection rate for the cubic model and the heteroskedasticity model.
+With the data collected from the pilot study, two logistic regressions given in Table \@ref(tab:pilotglm) and Figure \@ref(fig:pilotglm) were developed to describe the relationship between the natural logarithm of effect size and the detection rate for the cubic model and the heteroskedasticity model.
 
@@
 For the first experiment, we would like to have at least 20 detection with chance greater than 99.99% in each category.
-Thus, we prepared $34$ easy lineups, $60$ medium lineups and $194$ hard lineups with five replications for both cubic and heteroskedasticity model.
-Every participant would get at least $12$ easy or medium lineups.
+Thus, we prepared $68$ easy lineups, $120$ medium lineups and $388$ hard lineups with five replications.
+Every participant would get at least $8$ easy or medium lineups.
 Two out of $10$ extremely lineups with predicted detection rate over $90$% were randomly given to every participant as attention checks.
-We initially planned to recruit $(34+60+194) \times 2 \times 5 / (20-2) = 160$ participants but eventually decided to only recruit $20$ participants because it was the first time we launched an experiment on Prolific where factors like procedure of recruiting participants, payment method, and quality of participants were unclear at the moment.
+We initially planned to recruit $(68+120+388) \times 5 / (20-2) = 160$ participants but eventually decided to only recruit $20$ participants because it was the first time we launched an experiment on Prolific where factors like payment method and quality of participants were unclear at the moment.
 As a consequence, most of the lineups had only two evaluations and not all the lineups were used.
-However, data collected from the first experiment was sufficient for building a more accurate GLM model to predict detection rate for the second experiment.
+However, data collected from the first experiment was sufficient for building a more accurate logistic regression to predict detection rate for the second experiment.
 
@@
   mutate(difficulty = factor(difficulty, levels = c("extremely_easy", "easy", "medium", "hard")))
+
+# Delete set 18
+
+small_survey_dat <- filter(small_survey_dat, set != 18)
 ```
@@
 
-e1_high <- glm(detect ~ I(log(effect_size)),
+e1_high <- glm(detect ~ log(effect_size),
                 data = filter(small_survey_dat, type == "higher order"),
                 family = binomial)
-e1_heter <- glm(detect ~ I(log(effect_size)),
+e1_heter <- glm(detect ~ log(effect_size),
                  data = filter(small_survey_dat, type == "heteroskedasticity"),
@@
 
-According to the new GLM model given in Table \@ref(tab:e1glm) and Figure \@ref(fig:e1glmplot), it was found that the first experiment might be too difficult for participants as the detection rate was lower than what we estimated from the pilot study, especially for the heteroskedasticity model.
+According to the new logistic regression given in Table \@ref(tab:e1glm) and Figure \@ref(fig:e1glmplot), it was found that the first experiment might be too difficult for participants as the detection rate was lower than what we estimated from the pilot study, especially for the heteroskedasticity model.
 Thus, for the second experiments, we relaxed the sample size requirements and increased the proportion of easy lineups.
 In addition, due to budget constraint, the maximum number of participants was $50$.
-With the new GLM model fitted with data collected from the first experiment, $40$ easy lineups, $25$ medium lineups and $15$ hard lineups were simulated for both cubic and heteroskedasticity model in the second experiment.
-$49$ participants were recruited but $6$ were withdrew.
+With the new logistic regression fitted with data collected from the first experiment, $28$ easy lineups, $36$ medium lineups and $100$ hard lineups were simulated for the second experiment.
+$50$ participants were recruited but $6$ were withdrew.
 
@@
 
-ori_eff <- c()
-current_eff <- c()
-prm_c <- c()
-
 # Correct the effect size
@@
                           sigma = lineup_dat_big[[i]]$metadata$e_sigma)
-
-  ori_eff <- c(ori_eff, lineup_dat_big[[i]]$metadata$effect_size)
-  current_eff <- c(current_eff, tmp_eff)
-  prm_c <- c(prm_c, lineup_dat_big[[i]]$metadata$c)
 
@@
   mutate(difficulty = factor(difficulty, levels = c("extremely_easy", "easy", "medium", "hard")))
-```
-
-### Data cleaning
+
+big_survey_dat$detect <- big_survey_dat$detected
+```
+
+```{r eval= FALSE}
+# Cubic 1
+bind_rows(
+small_survey_dat %>%
+  filter(type == "higher order") %>%
+  mutate(p_hat = predict(mod1, newdata = data.frame(effect_size), type = "response")) %>%
+  mutate(new_diff = ifelse(p_hat < 0.4, "hard", ifelse(p_hat > 0.8, "easy", "medium"))),
+
+# Heter 1
+small_survey_dat %>%
+  filter(type != "higher order") %>%
+  mutate(p_hat = predict(mod2, newdata = data.frame(effect_size), type = "response")) %>%
+  mutate(new_diff = ifelse(p_hat < 0.4, "hard", ifelse(p_hat > 0.8, "easy", "medium")))
+) %>%
+  ggplot() +
+  geom_bar(aes(new_diff)) +
+  facet_wrap(~set)
+  # mutate(me = new_diff == "easy" | new_diff == "medium") %>%
+  # count(set, me) %>%
+  # filter(me == TRUE) %>%
+  # arrange(n)
+
+# total
+# 94, 184, 122
+# easy, hard, medium
+
+# unique lineups
+# 50, 147, 90
+# easy, hard, medium
+
+# at least 8 medium or easy
+```
+
+```{r eval = FALSE}
+e1_heter_2 <- glm(detect ~ effect_size,
+                 data = filter(small_survey_dat, type == "heteroskedasticity"),
+                 family = binomial)
+
+bind_rows(
+# Cubic 2
+big_survey_dat %>%
+  filter(type == "higher order") %>%
+  mutate(p_hat = predict(e1_high, newdata = data.frame(effect_size), type = "response")) %>%
+  mutate(new_diff = ifelse(p_hat < 0.4, "hard", ifelse(p_hat > 0.8, "easy", "medium"))),
+
+# Heter 2
+big_survey_dat %>%
+  filter(type != "higher order") %>%
+  mutate(p_hat = predict(e1_heter_2, newdata = data.frame(effect_size), type = "response")) %>%
+  mutate(new_diff = ifelse(p_hat < 0.4, "hard", ifelse(p_hat > 0.8, "easy", "medium")))
+) %>%
+  mutate(me = new_diff == "easy" | new_diff == "medium") %>%
+  count(set, me) %>%
+  filter(me == TRUE) %>%
+  arrange(n)
+```
+
+```{r}
+merged_survey <- bind_rows(mutate(small_survey_dat, exp = 1), mutate(big_survey_dat, exp = 2))
+```
+
+We collected $400$ lineup evaluations made by $20$ participants in experiment I and $880$ lineup evaluations made by $44$ participants.
+In total, $442$ unique lineups were evaluated by $64$ subjects.
+In experiment I, one of the participants skipped all $20$ lineups.
+Hence, the submission was rejected and removed from the data set.
+In experiment II, there was a participant failed one of the two attention checks, but there was no further evidence of low-effort throughout the experiment.
+So, the submission was kept.
 
@@
 
-###
+It was expected that with larger effect size, the power of the visual test would be greater for both cubic model and heteroskedasticity model.
+Using the power calculation method discussed in Section \@ref(p-value-and-power-estimation-of-visual-test-allowed-for-multiple-selections), power of each lineup from one to five evaluations was estimated.
+The estimated power was further used in fitting a quasibinomial generalized linear model
+
+```{r}
+env <- new.env()
+
+comb_pvalue <- merged_survey %>%
+  mutate(exp_lineup_id = paste0(exp, "_", lineup_id)) %>%
+  calc_p_value_multi(lineup_id = "exp_lineup_id", 
+                     detected = "detected", 
+                     n_sel = "num_selection",
+                     comb = TRUE,
+                     n_eval = 1:5,
+                     n_sim = 100000,
+                     cache_env = env) %>%
+  rowwise() %>%
+  mutate(hat_power = mean(eval_p_value(p_value, tol = 1e-2))) %>%
+  ungroup() %>%
+  select(exp_lineup_id = lineup_id, p_value, n_eval, hat_power)
+```
+
+```{r fig.cap="test"}
+merged_survey %>%
+  mutate(exp_lineup_id = paste0(exp, "_", lineup_id)) %>%
+  select(exp_lineup_id, type, exp, effect_size) %>%
+  right_join(comb_pvalue) %>%
+  mutate(n_eval = factor(n_eval)) %>%
+  ggplot() +
+  geom_point(aes(log(effect_size), hat_power), alpha = 0.3, size = 1) +
+  geom_smooth(aes(log(effect_size), hat_power, group = n_eval, col = n_eval),
+              method = "glm", method.args = list(family = quasibinomial)) +
+  facet_wrap(~type, scales = "free_x") +
+  theme_light(base_size = 5) +
+  xlab("Log of effect size") +
+  ylab("Estimated power")
+```
+
+```{r fig.cap="test"}
+merged_survey %>%
+  mutate(exp_lineup_id = paste0(exp, "_", lineup_id)) %>%
+  select(exp_lineup_id, type, exp, effect_size) %>%
+  right_join(comb_pvalue) %>%
+  mutate(n_eval = factor(n_eval)) %>%
+  unnest_longer(col = p_value) %>%
+  mutate(reject = eval_p_value(p_value, tol = 1e-2)) %>%
+  mutate(reject = as.numeric(reject)) %>%
+  ggplot() +
+  geom_point(aes(log(effect_size), reject), alpha = 0.3, size = 1) +
+  geom_smooth(aes(log(effect_size), reject, group = n_eval, col = n_eval),
+              method = "glm", method.args = list(family = binomial)) +
+  facet_wrap(~type, scales = "free_x") +
+  theme_light(base_size = 5) +
+  xlab("Log of effect size") +
+  ylab("Rejected")
+```
+
+```{r eval = FALSE}
+merged_survey %>%
+  filter(exp == 2, lineup_id == 1)
+```
 

### 2022-03-01 Thesis/_book/thesis.tex
update result

@@
 The sample size calculation was based on the detection rate of the lineup, which was closely related to the difficulty level of the lineup.
-With the data collected from the pilot study, two generalized linear model (GLM) models given in Table \ref{tab:pilotglm} and Figure \ref{fig:pilotglm} were developed to describe the relationship between the natural logarithm of effect size and the detection rate for the cubic model and the heteroskedasticity model.
+With the data collected from the pilot study, two logistic regressions given in Table \ref{tab:pilotglm} and Figure \ref{fig:pilotglm} were developed to describe the relationship between the natural logarithm of effect size and the detection rate for the cubic model and the heteroskedasticity model.
 
@@
 For the first experiment, we would like to have at least 20 detection with chance greater than 99.99\% in each category.
-Thus, we prepared \(34\) easy lineups, \(60\) medium lineups and \(194\) hard lineups with five replications for both cubic and heteroskedasticity model.
-Every participant would get at least \(12\) easy or medium lineups.
+Thus, we prepared \(68\) easy lineups, \(120\) medium lineups and \(388\) hard lineups with five replications.
+Every participant would get at least \(8\) easy or medium lineups.
 Two out of \(10\) extremely lineups with predicted detection rate over \(90\)\% were randomly given to every participant as attention checks.
-We initially planned to recruit \((34+60+194) \times 2 \times 5 / (20-2) = 160\) participants but eventually decided to only recruit \(20\) participants because it was the first time we launched an experiment on Prolific where factors like procedure of recruiting participants, payment method, and quality of participants were unclear at the moment.
+We initially planned to recruit \((68+120+388) \times 5 / (20-2) = 160\) participants but eventually decided to only recruit \(20\) participants because it was the first time we launched an experiment on Prolific where factors like payment method and quality of participants were unclear at the moment.
 As a consequence, most of the lineups had only two evaluations and not all the lineups were used.
-However, data collected from the first experiment was sufficient for building a more accurate GLM model to predict detection rate for the second experiment.
+However, data collected from the first experiment was sufficient for building a more accurate logistic regression to predict detection rate for the second experiment.
 
@@
 
-According to the new GLM model given in Table \ref{tab:e1glm} and Figure \ref{fig:e1glmplot}, it was found that the first experiment might be too difficult for participants as the detection rate was lower than what we estimated from the pilot study, especially for the heteroskedasticity model.
+According to the new logistic regression given in Table \ref{tab:e1glm} and Figure \ref{fig:e1glmplot}, it was found that the first experiment might be too difficult for participants as the detection rate was lower than what we estimated from the pilot study, especially for the heteroskedasticity model.
 Thus, for the second experiments, we relaxed the sample size requirements and increased the proportion of easy lineups.
 In addition, due to budget constraint, the maximum number of participants was \(50\).
-With the new GLM model fitted with data collected from the first experiment, \(40\) easy lineups, \(25\) medium lineups and \(15\) hard lineups were simulated for both cubic and heteroskedasticity model in the second experiment.
-\(49\) participants were recruited but \(6\) were withdrew.
+With the new logistic regression fitted with data collected from the first experiment, \(28\) easy lineups, \(36\) medium lineups and \(100\) hard lineups were simulated for the second experiment.
+\(50\) participants were recruited but \(6\) were withdrew.
 
@@
 
-\hypertarget{data-cleaning}{%
-\subsection{Data cleaning}\label{data-cleaning}}
+We collected \(400\) lineup evaluations made by \(20\) participants in experiment I and \(880\) lineup evaluations made by \(44\) participants.
+In total, \(442\) unique lineups were evaluated by \(64\) subjects.
+In experiment I, one of the participants skipped all \(20\) lineups.
+Hence, the submission was rejected and removed from the data set.
+In experiment II, there was a participant failed one of the two attention checks, but there was no further evidence of low-effort throughout the experiment.
+So, the submission was kept.
 
@@
 
-\hypertarget{section}{%
-\subsection{}\label{section}}
+It was expected that with larger effect size, the power of the visual test would be greater for both cubic model and heteroskedasticity model.
+Using the power calculation method discussed in Section \ref{p-value-and-power-estimation-of-visual-test-allowed-for-multiple-selections}, power of each lineup from one to five evaluations was estimated.
+The estimated power was further used in fitting a quasibinomial generalized linear model
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/unnamed-chunk-26-1.pdf}
+\caption{\label{fig:unnamed-chunk-26}test}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/unnamed-chunk-27-1.pdf}
+\caption{\label{fig:unnamed-chunk-27}test}
+\end{figure}
 

### 2022-03-03 Thesis/01-chap1.Rmd
Comment up to Section 1.4

@@
 
-## AI: Four Approaches
+## Artificial intelligence: Four Approaches
 
@@
 One such example is the **total Turing test** proposed by @harnad_other_1991.
-It adds $3$ new requirements to the list, including **computer vision**, **speech recognition** and **robotics**, which are response for interactions with the physical world.
-Notably, all $7$ required capabilities have become major subfields of AI today.
+It adds three new requirements to the list, including **computer vision**, **speech recognition** and **robotics**, which are response for interactions with the physical world.
+Notably, all seven required capabilities have become major subfields of AI today.
 And their development has made AI one of the fastest-growing fields in the 21st century [@russell_artificial_2002].
@@
 Further, the increasingly cheap computing cost and the massive amount of accessible e-commerce data produced in the Internet age provide the possibility for applying data-intensive AI models, which enables AI performance to reach new heights in history [@jordan_machine_2015].
-Some AI systems have already been remarkably better than human in certain areas, e.g., game playing.
-AlphaGo and AlphaZero developed by the Google DeepMind team surpass all human Go players [@silver_general_2018].
+Some AI systems have already been remarkably better than humans in certain areas, e.g., game playing.
+AlphaGo and AlphaZero developed by the Google DeepMind team surpass professional human Go players [@silver_general_2018].
 
@@
 Predictive models are primarily evaluated by predictive accuracy with the use of validation and test data, but in predictive model diagnostics, especially model testing and tuning, data plots play an irreplaceable role.
-In these diagnostics, though numeric summaries are mostly available and some are even endorsed by finite or asymptotic properties, graphical representation of data is still preferred, or at least needed by researchers, due to its intuitiveness and the possibility to provide unexpected discoveries which may be abstract and unquantifiable.
+In these diagnostics, though numeric summaries are mostly available and some are even endorsed by finite or asymptotic properties, graphical representation of data is still preferred, or at least needed by researchers, due to its intuitiveness and the possibility for unexpected discoveries which may be abstract and unquantifiable.
 
@@
 Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
-One such example is finding a separation between gene groups in a two-dimensional projection from a linear discriminant analysis where there is no difference in the expression levels between the gene groups [@roy_chowdhury_using_2015].
+One such example is finding [ET: an over-interpretation of the] separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurence [@roy_chowdhury_using_2015].
 
@@
 Visual inference was first introduced by @buja_statistical_2009 as an inferential framework to extend confirmatory statistics to visual discoveries.
-This framework redefines the test statistics, tests, null distribution, significance levels and $p$-value for visual discovery modelled on the confirmatory statistical testing.
+[ET: this idea was actually presented first in 1999 JSM talk "Inference for Data Visualization” by Buja et al.
+http://stat.wharton.upenn.edu/~buja/PAPERS/visual-inference.pdf -- as cited by Gelman in 2003 https://paperpile.com/shared/u0Gy7K] This framework redefines the test statistics, tests, null distribution, significance levels and $p$-value for visual discovery modelled on the confirmatory statistical testing.
 Figure \@ref(fig:parallelism) outlines the parallelism between conventional tests and visual discovery.
@@
 ```
+
+[ET: Caption of Figure 1.3 requires fixing. E.g. what is null hypothesis referring to? What visual discovery are you referring to?]
 

### 2022-03-03 Thesis/01-chap1.Rmd
comments up to 1.6

@@
 As discussed in @buja_statistical_2009, in the practice of model diagnostics, the range of possible visual discoveries is not pre-specified.
-In other words, people do not explicitly specify which one or more visual features they are looking for before the read of the diagnostic plot.
-This is concerning since conventional hypothesis testing always requires the pre-specification of the parameter space $\Theta$ of the parameter of interest $\theta \in \Theta$ to form a valid inferential procedure.
-To address this issue, a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$ is defined, where $\boldsymbol{\mathrm{y}}$ is the data and $\boldsymbol{I}$ is a set of all possible visual features.
+In other words, people do not explicitly specify which visual feature(s) they are looking for before the reading the diagnostic plot.
+In contrary, conventional hypothesis testing always requires the pre-specification of the parameter space $\Theta$ of the parameter of interest $\theta \in \Theta$ to form a valid inferential procedure.
+To address this issue, a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$ is defined, where $\boldsymbol{\mathrm{y}}$ is the data and $I$ is a set of all possible visual features.
 @buja_statistical_2009 described each of the test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})$ as a measurement of the degree of presence of a visual feature.
@@
 It is inspired by the "police lineup" or "identity parade" which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people.
-The protocol consists of $m$ randomly placed data plots, where $1$ plot is the actual data plot, and $m-1$ null plots are produced by plotting data simulate from the null distribution which is consistent with the null hypothesis.
+The protocol consists of $m$ randomly placed plots, where one plot is the actual data plot, and the remaining $m - 1$ plots have the identical graphical production as the data plot except the data has been replaced with data consistent with the null hypothesis.
 Then, an observer who have not seen the actual data plot will be asked to point out the most different plot from the lineup.
 
-Under the null hypothesis, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is $1/m$ due to randomness.
+Under the null hypothesis, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is $1/m$.
 If we reject the null hypothesis as the observer correctly picks the actual data plot, then the Type I error of this test is $1/m$.
@@
 This provides us with an mechanism to control the Type I error, because $m$ - the number of plots in a lineup can be chosen.
-A larger value of $m$ will result in a smaller Type I error, but the limit to the value of $m$ depends on the number of plots a human willing to view [@buja_statistical_2009].
+A larger value of $m$ will result in a smaller Type I error, but the limit to the value of $m$ depends on the number of plots a human is willing to view [@buja_statistical_2009].
 Typically, $m$ will be set to $20$ which is equivalent to set $\alpha = 0.05$, a general choice of significance level for conventional testing among statisticians.
@@
 
+[ET: Multiple observers, to me, is not about making smaller p-value possible with finite $m$ but if you only have a single observer, you are confounding the test with the visual ability of that single observer.
+In some ways, you can consider multiple observers as an ensemble approach and the decision is made on pooling the results from multiple "methods", each with their own abilities.]
+
 The multiple individuals approach avoids the limit of $m$, while provides visual tests with $p$-value much smaller than $0.05$.
@@
 
-Assumes that individuals have the same ability and a lineup has been evaluated by multiple individuals, under the alternative hypothesis, the estimated power for a lineup can be expressed as $\hat{p} = x/K$, the estimated probability of identifying the actual data plot from the lineup.
-If the individual skill needs to be taken into account, and $L$ lienups have been evaluated by $K$ individuals, @majumder_validation_2013 suggests that mixed effects logistic regression model can be fit as:
+Suppose individuals have the same ability and a lineup has been evaluated by multiple individuals, under the alternative hypothesis, the estimated power for a lineup can be expressed as $\hat{p} = x/K$, the estimated probability of identifying the actual data plot from the lineup.
+If the individual skill needs to be taken into account, and $L$ lineups have been evaluated by $K$ individuals, @majumder_validation_2013 suggests that mixed effects logistic regression model can be fit as:
 
@@
 where $g(.)$ is the logit link function $g(p) = log(p)  - log(1-p)$; $0 \leq p \leq 1$.
-$W_{li}$, $1 \leq i \leq K$, $1 \leq l \leq L$, is the covariate matrix including lineup-specific elements and demographic information of individuals, and $\delta$ is a vector of parameters.
-$Z$ is the random effects matrix, and $\tau$ is a vector of variables follow $N(\boldsymbol{0},\sigma_{\tau}\boldsymbol{I}_{KL\times KL})$.
+$W_{li}$, $1 \leq i \leq K$, $1 \leq l \leq L$, is the covariate matrix including lineup-specific elements and demographic information of individuals, and $\delta$ is a vector of fixed effects.
+$Z_{li}$ is the random effects matrix, and $\tau ~ N(\boldsymbol{0}_{KL},\sigma_{\tau}\boldsymbol{I}_{KL\times KL})$.
+
+[ET: need to double check this.
+Is $W_{li}$ a matrix or an element of the matrix?
+I wouldn't say "$Z_{li}$ is the random effects matrix".
+It's gotta be like random intercept model for lineup id and another random intercept for person id?
+ET to check.]
 

### 2022-03-06 Thesis/01-chap1.Rmd
finalize 1 & 2

@@
 
+## Is what we see really there?
+
 ## Artificial intelligence: Four Approaches
@@
 
-<!-- [ET: don't start with a definition as that makes the narrative dull.
-A lot of what you are writing definitions - which is relevant - but you want to interweave this as _narrative_.
-Why are you defining/describing these terms/disciplines?
-What's relevant and exciting?] -->
-
 1. Should AI act and think humanly or rationally? 
@@
 Based on the answer to the above questions, four major approaches to pursue AI have been established.
-These approaches can be summarized into a two by two table as shown in Figure \@ref(fig:twodimai), where the row is "Human" vs.
+These approaches can be summarized into a two by two table as shown in Figure \@ref(fig:two-dim-ai), where the row is "Human" vs.
 "Rational", and the column is "Behaviour" vs.
@@
 
-```{r twodimai, fig.cap="Four possible approaches to pursue AI based on the two dimensions in AI research - human vs. rational and thought vs. behavior [@russell_artificial_2002]."}
+```{r two-dim-ai, fig.cap="Four possible approaches to pursue AI based on the two dimensions in AI research - human vs. rational and thought vs. behavior [@russell_artificial_2002]."}
 ggplot() +
@@
 Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
-One such example is finding [ET: an over-interpretation of the] separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurence [@roy_chowdhury_using_2015].
+One such example is finding an over-interpretation of the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurence [@roy_chowdhury_using_2015].
 
@@
 
-Visual inference was first introduced by @buja_statistical_2009 as an inferential framework to extend confirmatory statistics to visual discoveries.
-[ET: this idea was actually presented first in 1999 JSM talk "Inference for Data Visualization” by Buja et al.
-http://stat.wharton.upenn.edu/~buja/PAPERS/visual-inference.pdf -- as cited by Gelman in 2003 https://paperpile.com/shared/u0Gy7K] This framework redefines the test statistics, tests, null distribution, significance levels and $p$-value for visual discovery modelled on the confirmatory statistical testing.
+Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualizatino" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots [@gelman_exploratory_2004].
+Later, in the Bayesian context, data plots was systematically considered as model diagnostics by taking advantage of the data simulated from the assumed statistical models [@gelman_bayesian_2003; @gelman_exploratory_2004].
+
+It was surprising that the essential components of visual inference had actually been established in @buja_inference_1999, but it was not until 10 years later that @buja_statistical_2009 formalized it as an inferential framework to extend confirmatory statistics to visual discoveries.
+This framework redefines the test statistics, tests, null distribution, significance levels and $p$-value for visual discovery modelled on the confirmatory statistical testing.
 Figure \@ref(fig:parallelism) outlines the parallelism between conventional tests and visual discovery.
@@
 (ref:parallelism) Parallelism between multiple quantitative testing and visual discovery [@buja_statistical_2009].
-Visible features in a plot are viewed as a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$, and any visual discoveries are treated as evidence against the null hypothesis.
+Visible features in a plot are viewed as a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$, and any visual discoveries that are inconsistent with the null hypothesis are treated as evidence against the null.
+For model diagnostics, the null hypothesis would be the assumed model, and visual discoveries could be any visual features in favour of any alternatives.
 
@@
 For model diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the null hypothesis.
-The same assumed model, such as classical linear regression model, can be rejected by many reasons with residual plot, including nonlinearity and heteroskedasticity as shown in Figure \@ref(fig:nonlinearityheter).
-
-```{r nonlinearityheter, fig.cap='Residuals vs.
+The same assumed model, such as classical linear regression model, can be rejected by many reasons with residual plot, including nonlinearity and heteroskedasticity as shown in Figure \@ref(fig:residual-plot-cubic-heter).
+
+```{r residual-plot-cubic-heter, fig.cap='Residuals vs.
 fitted values plot for a classical linear regression model.
@@
 ```
-
-[ET: Caption of Figure 1.3 requires fixing. E.g. what is null hypothesis referring to? What visual discovery are you referring to?]
 
@@
 
+## Sampling from the Null Distribution
+
+In visual inference, the null distribution of plots refers to the infinite collection of plots of null datasets sampled from $H_0$.
+It is defined as the analogue of the null distribution of test statistics in conventional test [@buja_statistical_2009].
+In practice, a finite number of plots of null datasets could be generated, called null plots.
+
+In the context of model diagnostics, sampling data from $H_0$ is equivalent to sampling data from the assumed model.
+As @buja_statistical_2009 suggested, $H_0$ is usually composited by a collection of distributions controlled by nuisance parameters.
+Since statistical models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling.
+
+The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009].
+Focus will be put on simulation based on this distribution as it is applicable to classical normal linear regression model.
+Details of the other two are can be found in the supplementary material of @buja_statistical_2009.
+Suppose there exists a minimal sufficient statistic $\boldsymbol{S}(\boldsymbol{y})$ under the null hypothesis, any null datasets $\boldsymbol{y^{*}}$ should fulfil the condition $\boldsymbol{S}(\boldsymbol{y}) = \boldsymbol{s}$.
+For classical normal linear regression model, the minimal sufficient statistic is $\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{\beta}}, \boldsymbol{e}'\boldsymbol{e})$, where $\hat{\boldsymbol{\beta}}$ are the coefficient estimators and $\boldsymbol{e}'\boldsymbol{e}$ is the residual sum of square.
+Alternatively, the minimal sufficient statistic can be constructed as $\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{y}}, ||\boldsymbol{e}||)$, where $\hat{\boldsymbol{y}}$ are the fitted values and $||\boldsymbol{e}||$ is the length of residuals, which is more intuitive as suggested by @buja_statistical_2009.
+Since the fitted values are held fixed, the variation can only occur in the residual space.
+And because the length of residual is also held fixed, residuals obtained from a null dataset has to be a random rotation of $\boldsymbol{e}$ in the residual space.
+With this property, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the regressors, then rescaling it by the ratio of residual sum of square in two regressions.
+
 ## Lineup Protocol
 
-With the validity of visual inference being justified, another aspect of hypothesis testing that needs to be addressed is the control of false positive rate or Type I error.
+With the validity of visual inference being justified and the simulation of null plots being provided, another aspect of hypothesis testing that needs to be addressed is the control of false positive rate or Type I error.
 Any visual statistic $T^{(i)}(\boldsymbol{\mathrm{y}})$ needs to pair with a critical value $c^{(i)}$ to form a hypothesis test.
@@
 
-[ET: Multiple observers, to me, is not about making smaller p-value possible with finite $m$ but if you only have a single observer, you are confounding the test with the visual ability of that single observer.
-In some ways, you can consider multiple observers as an ensemble approach and the decision is made on pooling the results from multiple "methods", each with their own abilities.]
-
 The multiple individuals approach avoids the limit of $m$, while provides visual tests with $p$-value much smaller than $0.05$.
@@
 With just $4$ individuals and $20$ data plots in a lineup, the $p$-value could be as small as $0.0001$.
+Additionally, by involving multiple observers, variation of individual ability to read plots can be addressed to some degree as different opinions about visual discoveries can be collected.
 
@@
 where $g(.)$ is the logit link function $g(p) = log(p)  - log(1-p)$; $0 \leq p \leq 1$.
-$W_{li}$, $1 \leq i \leq K$, $1 \leq l \leq L$, is the covariate matrix including lineup-specific elements and demographic information of individuals, and $\delta$ is a vector of fixed effects.
-$Z_{li}$ is the random effects matrix, and $\tau ~ N(\boldsymbol{0}_{KL},\sigma_{\tau}\boldsymbol{I}_{KL\times KL})$.
-
-[ET: need to double check this.
-Is $W_{li}$ a matrix or an element of the matrix?
-I wouldn't say "$Z_{li}$ is the random effects matrix".
-It's gotta be like random intercept model for lineup id and another random intercept for person id?
-ET to check.]
+$W_{li}$, $1 \leq i \leq K$, $1 \leq l \leq L$, is the covariate matrix including lineup-specific elements and demographic information of individuals, and $\delta$ is a vector of parameters.
+$Z$ is the random effects matrix, and $\tau$ is a vector of variables follow $N(\boldsymbol{0},\sigma_{\tau}\boldsymbol{I}_{KL\times KL})$.
 
@@
 
-**null data generate mechanism**
-
-## Applications of Lineup Protocol
-
-Lineup protocols are
-
-List applications, fields, significance here.
-
-## Limitations of visual tests
-
-1. infeasible in a large scale
-2. unfriendly to vision-impaired people
-3. high finical cost and human cost
-4. time consuming
-
-similar to Handicraft in pre-industrial society
-
-## Automatic visual inference -> Computer vision
-
-relieve people’s workload by automating repeating tasks, and provide standard result in a control environment
-
-the use of technology and machinery to enable mass evaluation of visual tests
-
-## Discussion of potential methods
-
-### different approaches of AI
-
-- Not aim for understanding the thought process
-- Not aim for mimicking the human vision mechanism
-- May be able to define distance metrics to measure difference between data plots for making mathematically prefect decisions
-- May be able to use computer vision model to approximate how people evaluate lineups
-
-computer vision model:
-
-- Use human data to train model with human selection as target -> mimic the human behaviour
-- Use simulated lineup to train model with actual data plot as target -> assume the actual data plot is the most different one
-- Use simulated data plot to train model with null or not null as target -> limited to the alternatives given to the model, and Type I error can not be controlled
-- Use simulated data plot to train model, then let the model evaluate each plot of a lineup -> leads to multiple selections in a lineup, and the model does not compare different plots in a lineup
-  - Use a very large $m$ to ensure the null hypothesis can be rejected by a classifier slightly better than a random selector. The problem becomes how large the $m$ should be
-  - Use multiple models as multiple individuals to evaluate lineup -> Which models should be used? How many models should be used?
-- Let the model select multiple plots from a simulated lineup while comparing different plots in a lineup -> how to build such a system? 
-  - The input is $m$ plots, what is the output? The probability of being the most different one? 
-  - Then how many plots should be selected? Top 5? Or cumulative probability greater than a threshold? 
-  - How to select the threshold?
-
-# Human Subject Exeriments (toc)
-
-What needs to be discussed?
-
-1. online recruitment platform (Prolific)
-  - pre-screen
-  - approve and reject
-  - redirect
-  - payment
-2. study web site
-  - PythonAnywhere
-  - Flask, Python
-  - jsPsych
-  - survey
-    - age
-    - education
-    - preferred pronoun
-    - previous experience
-  - training
-  - 20 lineups
-    - setup depends on the study
-3. Cubic model
-  - data generating process
-  - null model
-  - potential shapes
-  - effect size
-  - conventional test
-4. Heteroskedasticity model
-  - data generating process
-  - null model
-  - potential shapes
-  - effect size
-  - conventional test
-4. Number of participants needed by the study
-3. Study 1
-  - logistic regression
-    - easy
-    - medium
-    - hard
-  - number of participants
-  - time
-  - countries
-4. Study 2
-  - logistic regression
-    - easy
-    - medium
-    - hard
-  - number of participants
-  - time
-  - countries
+## Limitations of Lineup Protocol and Automatic Visual Inference
+
+Although lineup protocol has already been integrated into data analysis of various topics, such as diagnostics of hierarchical linear models [@loy_diagnostic_2013], geographical Research [@widen_graphical_2016] and forensic examinations [@krishnan_hierarchical_2021], the involvement with human judgements limits its popularity.
+Similar to handicraft in pre-industrial society, lineup protocol conducted by humans is infeasible on a large scale, high in labour cost and time consuming.
+Moreover, it is strongly unfriendly to vision-impaired people.
+
+The "steam engine" of visual inference needs to developed for relieving people's workload by automating repeating tasks and providing standard result in a control environment.
+Large-scale evaluation of visual tests is not possible without the use of technology and machines.
+
+Modern computer vision model could be a promising solution to this problem.
+As a subfield of AI, computer vision with the modern deep learning architectures solved numerous critical problems in automation.
+Inspired by the vision processing in living organisms, the convolutional neural network was introduced by @fukushima_neocognitron_1982.
+Soon, this architecture was applied to hand-written number recognition trained with back-propagation by @lecun_backpropagation_1989.
+This was one of the earliest attempts which human successfully extract information from digital images via self-learning algorithms.
+Modern computer vision model is typically built on the deep neural network with convolutional layers [@fukushima_neocognitron_1982].
+Convolutional layers take advantage of the hierarchical pattern in data and provide regularized versions of fully-connected layers.
+It downscales and transforms the image by summarising information in a small space.
+Numerous studies have shown that it can be used to effectively tackle vision tasks, such as image recognition [@rawat_deep_2017].
+With the development of graphics processing units and the spread of high-performance personal computers, researches in computer vision become a new hype in the 21st century.
+Achievements such as computer-aided diagnosis [@lee_image_2015], pedestrian detection [@brunetti_computer_2018] and facial recognition [@emami_facial_2012] had a significant impact on our daily life.
+
+Using computer vision models to read data plots is not a general choice.
+some fields have adopted this idea by applying computer vision models to read recurrence plots for time series regression [@ojeda_multivariate_2020], time series classification [@chu_automatic_2019; @hailesilassie_financial_2019; @hatami_classification_2018; @zhang_encoding_2020], anomaly detection [@chen_convolutional_2020] and pairwise causality analysis [@singh_deep_2017].
+However, evaluating lineups with computer vision model is a new field of study.
+
+(reasons for conducting experiments)
+
+## Structure of the paper
+
+...
 
@@
 Each lineup consists of one actual residual plot and 19 null residual plots produced by plotting null residuals simulated from the residual rotation distributions.
-Among the 20 lineups, there are two extremely easy lineups used as attention checks which everyone should get correct.
+Among the 20 lineups, there were two extremely easy lineups used as attention checks which everyone should get correct.
 The short survey was intended to collect information about participant that might affect their ability in reading data plot including age, highest level of education, preferred pronoun and previous experience in similar study.
 For the evaluation of the lineup, participant first needed to select one or multiple most different data plot(s) from the lineup by clicking the corresponding image.
-Then, the reason for choosing the plots needed to be provided by selecting one of the given options - "outlier(s)", "cluster(s)", "shape" and "other".
+Then, the primary reason for choosing the plots needed to be provided by picking one of the given options - "outlier(s)", "cluster(s)", "shape" and "other".
 Explanations about these options were provided in the training page and the mouseover text.
-Table \@ref(tab:lineupreason) gives the detailed explanations about these reasons.
+Table \@ref(tab:lineup-reason) gives the detailed explanations about these reasons.
 If the option "other" was selected, text input for the specified reason would be collected.
-Lastly, the degree of difference between their chosen data plot(s) and other data plots needed to be selected among $5$ levels - "not at all", "slightly", "moderately", "very" and "extremely".
+Lastly, the degree of difference between their chosen data plot(s) and other data plots needed to be selected among five levels - "not at all", "slightly", "moderately", "very" and "extremely".
 Notably, if participants could not tell the difference between the data plots, there was an option to skip the evaluation of the lineup.
@@
              caption  = 'Explanations about reasons for choosing data plots from a lineup', 
-             label = "lineupreason") %>%
+             label = "lineup-reason") %>%
   kableExtra::kable_styling(latex_options = c("scale_down"))
@@
 This is essential for us to confirm the quality of the data by checking exceptionally fast or slow responses.
-Figure \@ref(fig:DiagramExperiment) summarizes the online experiment setup.
+Figure \@ref(fig:diagram-experiment) summarizes the online experiment setup.
 
@@
 And the client-side used jsPsych to run experiment.
-\label{fig:DiagramExperiment}](figures/experiment_tech.pdf){width=450 height=450}
+\label{fig:diagram-experiment}](figures/experiment_tech.pdf){width=450 height=450}
 
 The first two pages of the website are the explanatory statement and the consent form.
-Participants needed to read the documents and agree with the terms to advance to the short survey.
-With the completion of the survey, the next page is the training page, which contained instructions on how to evaluate the lineup, give the reason and choose the confidence level.
+Participants were asked to read the documents and agree with the terms to advance to the short survey.
+With the completion of the survey, the next page is the training page, which contained instructions on how to interact with the webpage.
 Followed by the training page, there are 20 lineups each on a single page.
-Figure \@ref(fig:WebsiteLayout) illustrates the layout of the website.
-At the end of the experiment, the participants would be redirected back to Prolific and waited for researchers' responses.
+Figure \@ref(fig:website-layout) illustrates the layout of the website.
+At the end of the experiment, participants would be redirected back to Prolific and waited for researchers' responses.
 
@@
 Participants needed to choose the most different plots on the right and select their reasons and confidence levels on the left.
-\label{fig:WebsiteLayout}](figures/website.png){width=550 height=299}
+\label{fig:website-layout}](figures/website.png){width=550 height=299}
 
@@
 
-Two models were used in both experiments, which were linear models with some degree of violations of classical assumptions. Residual vs. fitted values plot were used to present the model defects.
+Two models were used in the study, where both were linear models with some degree of violations of classical assumptions.
 
@@
 
-Clearly, there will be omitted-variable bias since the null model leaves out the quadric and cubic terms.
+Clearly, omitted-variable bias will present since the null model leaves out the quadric and cubic terms.
 
@@
 
-Lemma \@ref(lemma:cubic) shows that the expectation of the residuals is clearly a function of $\boldsymbol{X}$ and $\boldsymbol{Z}$.
-Hence, we would expect to see some patterns in the residual plot.
+Lemma \@ref(lemma:cubic) shows that the expectation of the residuals is clearly a function of $\boldsymbol{X}$ and $\boldsymbol{Z}$, which indicates the residuals and the fitted values $\hat{Y} = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'\boldsymbol{Y}$ are associated.
+Hence, it is expected that visual discoveries can be found in the residuals vs.
+fitted values plot.
 Let $X_i$ and $Z_i$, $i = 1,...,n$, be independent random variables follow uniform distribution $U(-1,1)$.
 Given the expectation of the residuals, we could plot the expected values of residuals against the observed values.
-Figure \@ref(fig:cubic1), \@ref(fig:cubic2) and \@ref(fig:cubic3) illustrate the shape of residuals and their expected values under different parameter settings.
-
-From the Figure \@ref(fig:cubic1), it can be observed that with fixed $\sigma$ and $c$, $a$ and $b$ are controlling the 2D projection of a hypersurface, and seemingly performing some rotations along different axes.
-Figure \@ref(fig:cubic2) shows that with fixed $a$, $b$ and $\sigma$, $c$ is controlling the contribution of $\boldsymbol{X}$ and $\boldsymbol{Z}$ to $\boldsymbol{Y}$.
-If we move $c$ toward $0$ or $2$, one variable will donminate another, which will mitigate some joint effects and resemble a typical cubic function.
-In Figure \@ref(fig:cubic3), $a$, $b$ and $c$ is held constant, and $\sigma$ is controlling the noises around the expected values.
-As $\sigma$ decreases, the underlying shape shows up.
-
-```{r cubic1, fig.cap = "cubic 1"}
+Figure \@ref(fig:cubic-shape-1), \@ref(fig:cubic-shape-2) and \@ref(fig:cubic-shape-3) illustrate the shape of residuals and their expected values under different parameter settings.
+
+From the Figure \@ref(fig:cubic-shape-1), it can be observed that with fixed $\sigma$ and $c$, $a$ and $b$ are controlling the 2D projection of a hypersurface, and seemingly performing some rotations along different axes.
+Figure \@ref(fig:cubic-shape-2) shows that with fixed $a$, $b$ and $\sigma$, $c$ is controlling the contribution of $\boldsymbol{X}$ and $\boldsymbol{Z}$ to $\boldsymbol{Y}$.
+As $c$ moves toward $0$ or $2$, one regressor will dominate another, which will mitigate some joint effects and resemble a typical cubic function.
+In Figure \@ref(fig:cubic-shape-3), $a$, $b$ and $c$ are held fixed, and $\sigma$ is controlling the noises around the expected values.
+The underlying shape is clearer for smaller $\sigma$.
+
+```{r cubic-shape-1, fig.cap = "A matrix of residuals vs.
+fitted values plot under different parameter settings of the cubic model.
+Resdiauls are drawns in black and the expected values are drawn in red.
+The plots show that if $\\sigma$ and $c$ are held fixed, $a$ and $b$ rotate a two-dimension projection of a hypersurface."}
 library(visage)
@@
 
-```{r cubic2, fig.cap = "cubic 2", fig.height=1.9}
+```{r cubic-shape-2, fig.cap = "A matrix of residuals vs.
+fitted values plot under different parameter settings of the cubic model.
+Resdiauls are drawns in black and the expected values are drawn in red.
+The plots show that if $a$, $b$ and $\\sigma$ are held fixed, $c$ controls the contribution of $\\boldsymbol{X}$ and $\\boldsymbol{Z}$ to $\\boldsymbol{Y}$.", fig.height=1.9}
 expand.grid(c=c(0,0.5,1,1.5,2),
@@
 
-```{r cubic3, fig.cap = "cubic 3", fig.height=1.9}
+```{r cubic-shape-3, fig.cap = "A matrix of residuals vs.
+fitted values plot under different parameter settings of the cubic model.
+Resdiauls are drawns in black and the expected values are drawn in red.
+The plots show that if $a$, $b$ and $c$ are held fixed, $\\sigma$ controls the strength of the signal.", fig.height=1.9}
 expand.grid(c=c(1),
@@
 
-Here, we set $\boldsymbol{X}$ and $\boldsymbol{Z}$ to be random uniform variables ranged from $-1$ to $1$.
-This can be done by using the random uniform variable class constructor `rand_uniform()` inherited from the random variable abstract base class.
+If we would like $\boldsymbol{X}$ and $\boldsymbol{Z}$ to be random uniform variables ranged from $-1$ to $1$, it can be achieved by using the random uniform variable class constructor `rand_uniform()` inherited from the random variable abstract base class.
 It only takes two arguments which are the lower bound and the upper bound of the support.
@@
 
-Method `mod$plot()` produce a ggplot [@wickham_ggplot2_2011] object.
+Method `mod$plot()` produce a `ggplot` [@wickham_ggplot2_2011] object.
 It takes a data frame containing columns `.resid` and `.fitted` as input, along with a character argument `type` indicating the type of the data plot, and other aesthetic arguments such as `size` and `alpha` to control the appearance of the plot.
@@
 
-Lineup can be produced by using the methods `mod$gen_lineup()` and `mod$plot_lineup()`.
+Lineup is a matrix of residual plots which can be produced by using the methods `mod$gen_lineup()` and `mod$plot_lineup()`.
 Method `mod$gen_lineup` takes the number of realizations `n` and the number of plots in a lineup `k` as inputs.
@@
 
-```{r hetershape, fig.height=5.25, fig.cap="test"}
+```{r hetershape, fig.height=5.25, fig.cap="A matrix of residuals vs.
+fitted values plot under different parameter settings of the heteroskedasticity model.
+Resdiauls are drawns in black and the one standard deviation around zero is drawn in red.
+The plots show that $a$ controls the shape and direction of the resdiaul plot, while $b$ controls the strength of the signal."}
 make_plot2 <- function(a, b) {
@@
 The distribution of $Z_i$, $i=1,...,n$, had $50$% chance to be a uniform distribution ranged from $-1$ to $1$, and $50$% chance to be a discrete uniform distribution with $z_n$ outcomes simulated from a uniform distribution ranged from $-1$ to $1$.
-$z_n$ itself was a discrete uniform distribution with outcomes $\{3,4,5,6,7,8,9,10\}$, which defined the number of possible values $Z_i$ could take.
+$z_n$ itself was a discrete uniform distribution with outcomes $\{10,11,12,13,14,15,16,17,18,19,20\}$, which defined the number of possible values $Z_i$ could take.
 As shown in Figure \@ref(fig:discreteex), this setup would create discreteness in residual plot, which could enrich the pool of visual patterns.
@@
           column.labels = c("Cubic", "Heteroskedasticity"),
-          title = "GLM result 1",
+          title = "Logistic regressions with detection rate as response variable and natural logarithm of the effect size as regressor fitted on the pilot data for both cubic and heteroskedasticity models.",
           style = "default")
@@
 
-```{r pilotglm, fig.cap="pilot"}
+```{r pilotglm, fig.cap="Logistic regressions with detection rate as response variable and natural logarithm of the effect size as regressor fitted on the pilot data for both cubic and heteroskedasticity models.
+Points are jittered to reduce overplotting using the `ggbeeswarm` package.
+The smooth curves are the fitted regression curves.
+The ribbons around the curves are the 95% confidence intervals."}
 list(p1, p2) %>% patchwork::wrap_plots(ncol = 2)
@@
 
-```{r e1glmplot, fig.cap="test"}
+```{r e1glmplot, fig.cap="Logistic regressions with detection rate as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I for both cubic and heteroskedasticity models.
+Points are jittered to reduce overplotting using the `ggbeeswarm` package.
+The smooth curves are the fitted regression curves.
+The ribbons around the curves are the 95% confidence intervals."}
 filter(small_survey_dat, type == "higher order") %>%
@@
           column.labels = c("Cubic", "Heteroskedasticity"),
-          title = "GLM result 2",
+          title = "Logistic regressions with detection rate as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I for both cubic and heteroskedasticity models.",
           style = "default")
@@
 merged_survey <- bind_rows(mutate(small_survey_dat, exp = 1), mutate(big_survey_dat, exp = 2))
-```
-
-We collected $400$ lineup evaluations made by $20$ participants in experiment I and $880$ lineup evaluations made by $44$ participants.
+merged_survey <- merged_survey %>%
+  mutate(type = ifelse(type == "higher order", "cubic", type))
+```
+
+We collected $400$ lineup evaluations made by $20$ participants in experiment I and $880$ lineup evaluations made by $44$ participants in experiment II.
 In total, $442$ unique lineups were evaluated by $64$ subjects.
 In experiment I, one of the participants skipped all $20$ lineups.
-Hence, the submission was rejected and removed from the data set.
+Hence, the submission was rejected and removed from the dataset.
 In experiment II, there was a participant failed one of the two attention checks, but there was no further evidence of low-effort throughout the experiment.
-So, the submission was kept.
+Therefore, the submission was kept.
 
@@
 Using the power calculation method discussed in Section \@ref(p-value-and-power-estimation-of-visual-test-allowed-for-multiple-selections), power of each lineup from one to five evaluations was estimated.
-The estimated power was further used in fitting a quasibinomial generalized linear model
+The estimated power was further used in fitting a quasi-binomial generalized linear model with natural logarithm of effect size as the only regressor.
+The result of the fitted models are given in Table \@ref(tab:powerglmcubic), Table \@ref(tab:powerglmheter), and Figure \@ref(fig:powerglm).
+
+As expected, coefficients of natural logarithm of effect size of all five models were positive.
+From Figure \@ref(fig:powerglm), it can be observed that the fitted power of visual test increased as the number of evaluations increased for both cubic and heteroskedasticity model.
+
+For heteroskedasticity model, this phenomenon was more obvious as the curves of visual tests with evaluations greater than two were always above the curves of visual test with evaluations smaller than two.
+However, this only held for large enough effect size.
+For small effect size, visual tests with fewer evaluations might have greater power.
+Note that, the expected power of visual test derived by @majumder_validation_2013 followed by the assumption that human has the ability to select the plot with the highest t-statistic from a lineup under the classical linear model regression setting showed similar properties, where visual tests with fewer evaluations were expected to perform better when the parameter values close to the null hypothesis.
+
+For cubic model, the separation between curves was small.
+Fitted power of visual test with three to five evaluations were almost identical to each other in regards of effect size.
+In addition, all five curves peaked at one as effect size increased, suggesting that identification of nonlinearity as a visual task can be completed reliably by human when the effect size is large enough.
+
+As shown in Figure \@ref(fig:powerglm), both F-test and Breusch–Pagan test generally possessed greater power than visual test.
+Since the settings we were using were not ideal scenarios for visual inference, it was not expected that the visual test would be as good as the conventional test.
+Our goal was to quantify the ability of human reading residual plot, such that the comparison between human and computer vision model can be made.
 
@@
 
-```{r fig.cap="test"}
+```{r results="asis"}
+merged_survey %>%
+  mutate(exp_lineup_id = paste0(exp, "_", lineup_id)) %>%
+  select(exp_lineup_id, type, exp, effect_size) %>%
+  right_join(comb_pvalue) %>%
+  mutate(n_eval = factor(n_eval)) -> tmp_dat
+
+map(1:5, function(x) glm(hat_power ~ log(effect_size),
+                         family = quasibinomial(),
+                         data = filter(tmp_dat, n_eval == {{x}}, type == "cubic"))) %>%
+  stargazer(header = FALSE,
+            label = "tab:powerglmcubic",
+            title = "Quasi-binomial logistic regressions with estimated power as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for cubic model.
+Five regressions were fitted for different number of evaluations.",
+            style = "default")
+
+```
+
+```{r results="asis"}
+map(1:5, function(x) glm(hat_power ~ log(effect_size),
+                         family = quasibinomial(),
+                         data = filter(tmp_dat, n_eval == {{x}}, type == "heteroskedasticity"))) %>%
+  stargazer(header = FALSE,
+            label = "tab:powerglmheter",
+            title = "Quasi-binomial logistic regressions with estimated power as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for heteroskedasticity model.
+Five regression were fitted for different number of evaluations.",
+            style = "default")
+
+```
+
+```{r}
+conv_test_cubic <- function() {
+  para_list1 <- list(a = runif(1, -3, 3),
+                     b = runif(1, -3, 3),
+                     c = runif(1, 0, 2),
+                     x_dist = sample(c("uniform", "normal", "lognormal", "neglognormal"), 1),
+                     x_mu = 0,
+                     x_n = NA,
+                     x_sigma = sample(c(0.3, 0.6), 1),
+                     z_discrete = sample(c(TRUE, FALSE), 1),
+                     z_discrete_dist = "discrete_uniform",
+                     z_n = sample(3:10, 1),
+                     e_dist = "normal",
+                     e_df = NA,
+                     e_sigma = sample(c(0.25, 0.5, 1, 2), 1),
+                     n = sample(c(50, 100, 300), 1))
+
+  if (para_list1$x_dist == "normal") {para_list1$x_sigma <- 0.3} else {para_list1$x_sigma <- 0.6}
+
+  # Init X and Z based on the parameter list
+  x <- switch (para_list1$x_dist ,
+               "uniform" = rand_uniform(-1, 1),
+               "normal" = rand_normal(0, para_list1$x_sigma),
+               "lognormal" = rand_lognormal(0, para_list1$x_sigma),
+               "neglognormal" = {rl <- rand_lognormal(0, para_list1$x_sigma); closed_form(~-rl)}
+  )
+
+  if (para_list1$z_discrete) {z <- rand_uniform(-1, 1)} else {z <- rand_uniform_d(-1, 1, para_list1$z_n)}
+
+  # Build the cubic model
+  mod <- cubic_model(a = para_list1$a, b = para_list1$b, c = para_list1$c,
+                     sigma = para_list1$e_sigma,
+                     x = x, z = z)
+
+  # Generate data from the model, let k = 2 (50% true data plot, 50% null plot)
+  dat <- mod$gen(para_list1$n, fit_model = TRUE, test = TRUE)
+
+  c(mod$effect_size(dat), dat$p_value[1])
+}
+
+conv_result_cubic <- map(1:5000, ~conv_test_cubic())
+
+conv_result_cubic <- data.frame(effect_size = map_dbl(conv_result_cubic, ~.x[1]), 
+                          p_value = map_dbl(conv_result_cubic, ~.x[2])) %>%
+  mutate(reject = p_value < 0.05) %>%
+  mutate(reject = as.numeric(reject))
+
+conv_result_cubic <- conv_result_cubic %>%
+  filter(log(effect_size) > -5) %>%
+  mutate(type = "cubic")
+```
+
+```{r}
+conv_test_heter <- function() {
+  para_list1 <- list(a = sample(c(-1, 0, 1), 1),
+                     b = runif(1, 0, 32),
+                     c = NA,
+                     x_dist = sample(c("uniform", "normal", "lognormal", "neglognormal", "discrete_uniform"), 1),
+                     x_mu = 0,
+                     x_n = sample(10:20, 1),
+                     x_sigma = sample(c(0.3, 0.6), 1),
+                     z_discrete = FALSE,
+                     z_discrete_dist = NA,
+                     z_n = NA,
+                     e_dist = NA,
+                     e_df = NA,
+                     e_sigma = NA,
+                     n = sample(c(50, 100, 300), 1))
+
+  if (para_list1$x_dist == "normal") {para_list1$x_sigma <- 0.3} else {para_list1$x_sigma <- 0.6}
+
+  # Init X and Z based on the parameter list
+  x <- switch (para_list1$x_dist ,
+               "uniform" = rand_uniform(-1, 1),
+               "discrete_uniform" = rand_uniform_d(-1, 1, para_list1$x_n),
+               "normal" = rand_normal(0, para_list1$x_sigma),
+               "lognormal" = rand_lognormal(0, para_list1$x_sigma),
+               "neglognormal" = {rl <- rand_lognormal(0, para_list1$x_sigma); closed_form(~-rl)}
+  )
+
+  # Build the cubic model
+  mod <- heter_model(a = para_list1$a, b = para_list1$b,
+                     x = x)
+
+  # Generate data from the model, let k = 2 (50% true data plot, 50% null plot)
+  dat <- mod$gen(para_list1$n, fit_model = TRUE, test = TRUE)
+
+  c(mod$effect_size(dat), dat$p_value[1])
+}
+
+conv_result_heter <- map(1:50000, ~conv_test_heter())
+
+conv_result_heter <- data.frame(effect_size = map_dbl(conv_result_heter, ~.x[1]), 
+                          p_value = map_dbl(conv_result_heter, ~.x[2])) %>%
+  mutate(reject = p_value < 0.05) %>%
+  mutate(reject = as.numeric(reject))
+
+conv_result_heter <- conv_result_heter %>%
+  filter(log(effect_size) > -3) %>%
+  mutate(type = "heteroskedasticity")
+# 
+# tt <- tapply(conv_result_heter$reject, cut(log(conv_result_heter$effect_size), 100), mean)
+# tt %>%
+#   names() %>%
+#   stringr::str_remove_all("\\(|\\]") %>%
+#   stringr::str_split(",") %>%
+#   map_dbl(~mean(as.numeric(.x))) -> ttt
+# 
+# data.frame(log_effect_size = as.numeric(ttt), reject = tt) %>%
+#   ggplot() +
+#   geom_point(aes(log_effect_size, reject)) +
+#   geom_smooth(aes(log_effect_size, reject), span = 0.4, col = "red") +
+#   geom_smooth(data = conv_result_heter, aes(log(effect_size), reject), method = "glm",
+#               method.args = list(family = binomial()))
+```
+
+```{r}
+conv_result <- bind_rows(conv_result_cubic, conv_result_heter)
+```
+
+```{r powerglm, fig.cap="Quasi-binomial logistic regressions with estimated power as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for both cubic and heteroskedasticity model.
+Five regression were fitted for different number of evaluations shown in five different colours.
+Estimated power are drawn in black.
+The dashed curvses are the power of F-test and Breusch–Pagan test respectively."}
 merged_survey %>%
@@
   mutate(n_eval = factor(n_eval)) %>%
+  filter(log(effect_size) < 20) %>%
   ggplot() +
-  geom_point(aes(log(effect_size), hat_power), alpha = 0.3, size = 1) +
+  geom_point(aes(log(effect_size), hat_power), alpha = 0.1, size = 1) +
+  geom_smooth(aes(log(effect_size), reject, group = "conventional"),
+              method = "glm", method.args = list(family = binomial), data = conv_result, 
+              se = FALSE, linetype = 2, size = 1, col = "grey80") +
   geom_smooth(aes(log(effect_size), hat_power, group = n_eval, col = n_eval),
-              method = "glm", method.args = list(family = quasibinomial)) +
+              method = "glm", method.args = list(family = quasibinomial), size = 0.5, se = FALSE) +
   facet_wrap(~type, scales = "free_x") +
+  labs(col = "Number of evaluations") +
   theme_light(base_size = 5) +
+  theme(legend.position = "bottom") +
+  guides(col = guide_legend(nrow = 1, byrow=TRUE)) +
+  scale_colour_viridis_d(option = "B", begin = 0.5, end = 0.9) +
   xlab("Log of effect size") +
@@
 
-```{r fig.cap="test"}
+Another method we used to model the power was to collect all decisions made by $K\choose{K-j}$ different outcomes for $V_{K-j}$, and directly fit a binomial generalized linear model with natural logarithm of effect size as the regressor.
+This method skipped the calculation of the estimated power $R/{K\choose{K-j}}$, and regressed on the binary outcome with larger sample size.
+Results of the fitted model is given in Figure \@ref(fig:powerglmsecond), Table \@ref(tab:powerglmhetersecond) and Table \@ref(tab:powerglmcubicsecond).
+According to Figure \@ref(fig:powerglmsecond), this method gave clearer separation between curves for heteroskedasticity model.
+Noticeably, fitted power of visual test with five evaluations matched the power of conventional test at large effect size.
+
+```{r results="asis"}
+merged_survey %>%
+  mutate(exp_lineup_id = paste0(exp, "_", lineup_id)) %>%
+  select(exp_lineup_id, type, exp, effect_size) %>%
+  right_join(comb_pvalue) %>%
+  mutate(n_eval = factor(n_eval)) %>%
+  unnest_longer(col = p_value) %>%
+  mutate(reject = eval_p_value(p_value, tol = 1e-2)) -> tmp_dat2
+
+map(1:5, function(x) glm(hat_power ~ log(effect_size),
+                         family = binomial(),
+                         data = filter(tmp_dat2, n_eval == {{x}}, type == "cubic"))) %>%
+  stargazer(header = FALSE,
+            label = "tab:powerglmcubicsecond",
+            title = "Logistic regressions with rejection as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for cubic model.
+Five regression were fitted for different number of evaluations.",
+            style = "default")
+```
+
+```{r results="asis"}
+map(1:5, function(x) glm(hat_power ~ log(effect_size),
+                         family = binomial(),
+                         data = filter(tmp_dat2, n_eval == {{x}}, type == "heteroskedasticity"))) %>%
+  stargazer(header = FALSE,
+            label = "tab:powerglmhetersecond",
+            title = "Logistic regressions with rejection as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for heteroskedasticity model.
+Five regression were fitted for different number of evaluations.",
+            style = "default")
+```
+
+```{r powerglmsecond, fig.cap="Logistic regressions with rejection as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for both cubic and heteroskedasticity model.
+Five regression were fitted for different number of evaluations shown in five different colours.
+Estimated power are drawn in black.
+The dashed curvses are the power of F-test and Breusch–Pagan test respectively."}
 merged_survey %>%
@@
   ggplot() +
-  geom_point(aes(log(effect_size), reject), alpha = 0.3, size = 1) +
+  geom_point(aes(log(effect_size), reject), alpha = 0.1, size = 1) +
+  geom_smooth(aes(log(effect_size), reject, group = "conventional"),
+              method = "glm", method.args = list(family = binomial), data = conv_result, 
+              se = FALSE, size = 1, linetype = 2, col = "grey80") +
   geom_smooth(aes(log(effect_size), reject, group = n_eval, col = n_eval),
-              method = "glm", method.args = list(family = binomial)) +
+              method = "glm", method.args = list(family = binomial), size = 0.5, se = FALSE) +
   facet_wrap(~type, scales = "free_x") +
+  labs(col = "Number of evaluations") +
   theme_light(base_size = 5) +
+  theme(legend.position = "bottom") +
+  scale_colour_viridis_d(option = "B", begin = 0.5, end = 0.9) +
+  guides(col = guide_legend(nrow = 1, byrow=TRUE)) +
   xlab("Log of effect size") +
@@
 
-(estimate of visual power)
-
-(power against different parameters)
-
-(power against conventional p-value)
-
-(conventional power)
-
-(visual p-value vs. conventional p-value)
-
-(response time)
-
-(confidence level)
-
-(relative freq of picks)
-
-\clearpage
+### Effect of Factors on Power of the Visual Test
+
+```{r results="asis"}
+merged_survey %>%
+  mutate(exp_lineup_id = paste0(exp, "_", lineup_id)) %>%
+  right_join(comb_pvalue) %>%
+  mutate(n_eval = n_eval) %>%
+  mutate(x_dist = factor(x_dist, levels = c("uniform", "normal", "lognormal", "neglognormal"))) %>%
+  unnest_longer(col = p_value) %>%
+  mutate(reject = eval_p_value(p_value, tol = 1e-2)) -> single_merged
+
+glm(reject ~ a + b + c + e_sigma + x_dist + z_discrete + n + n_eval, data = filter(single_merged, type == "cubic"), family = binomial()) -> mod1_factor
+
+```
+
+```{r results="asis"}
+glm(reject ~ a + b + x_dist + z_discrete + n + n_eval, data = filter(single_merged, type == "heteroskedasticity"), family = binomial()) -> mod2_factor
+
+stargazer(mod1_factor, mod2_factor,header = FALSE,
+          label = "tab:rejectfactor",
+          title = "Logistic regressions with rejection as response variable and parameter settings as regressors fitted on the data collected from experiment I and II for both cubic and heteroskedasticity models.",
+          style = "default",
+          covariate.labels = c("a", "b", "c", 
+                               "sigma", 
+                               "Distribution of X: Normal", 
+                               "Distribution of X: Lognormal", 
+                               "Distribution of X: Negative Lognormal",
+                               "Distribution of Z: Discrete",
+                               "n",
+                               "Number of evaluations",
+                               "Constant"))
+```
+
+<!-- ### Visual $p$-value -->
+
+<!-- ### Response Time -->
+
+<!-- ### Reason -->
+
+<!-- ### Distribution of selections -->
+
+<!-- (power against different parameters) -->
+
+<!-- (power against conventional p-value) -->
+
+<!-- (conventional power) -->
+
+<!-- (visual p-value vs. conventional p-value) -->
+
+<!-- (response time) -->
+
+<!-- (confidence level) -->
+
+<!-- (relative freq of picks) -->
+
+<!-- (all variables in the merged dataset) -->
+
+# Computer Vision Model
+
+Visual test can be used to make valid inference on visual discoveries and acknowledging people with confidence about the findings.
+However, the lack of scalability of this method due to relatively high human cost compared to conventional test limits its application in wider fields.
+To relieve human analysts of the job of diagnosing predictive models in the context of visual inference, AI agents which can read data plot, evaluate lineups, and provide decisions need to be built.
+
+Considering there are different approches of AI, it is necessary to clarify the type of AI we aimed to build.
+Understanding the human thought process or mimicking the human vision mechanism were not of interest in this research.
+We believed humanly behaviours was enough for demonstrating AI for the problem we concerned.
+Hence, adopting one of the capacities required by the Turing test, namely machine learning, along with
+
+Thought process, human visual
+
+## Discussion of potential methods
+
+### different approaches of AI
+
+- Not aim for understanding the thought process
+- Not aim for mimicking the human vision mechanism
+- May be able to define distance metrics to measure difference between data plots for making mathematically prefect decisions
+- May be able to use computer vision model to approximate how people evaluate lineups
+
+computer vision model:
+
+- Use human data to train model with human selection as target -> mimic the human behaviour
+- Use simulated lineup to train model with actual data plot as target -> assume the actual data plot is the most different one
+- Use simulated data plot to train model with null or not null as target -> limited to the alternatives given to the model, and Type I error can not be controlled
+- Use simulated data plot to train model, then let the model evaluate each plot of a lineup -> leads to multiple selections in a lineup, and the model does not compare different plots in a lineup
+  - Use a very large $m$ to ensure the null hypothesis can be rejected by a classifier slightly better than a random selector. The problem becomes how large the $m$ should be
+  - Use multiple models as multiple individuals to evaluate lineup -> Which models should be used? How many models should be used?
+- Let the model select multiple plots from a simulated lineup while comparing different plots in a lineup -> how to build such a system? 
+  - The input is $m$ plots, what is the output? The probability of being the most different one? 
+  - Then how many plots should be selected? Top 5? Or cumulative probability greater than a threshold? 
+  - How to select the threshold?
+
+The primary drawbacks of visual inference are
+
+ infeasible in a large scale
+2. unfriendly to vision-impaired people 3. high finical cost and human cost
+4. time consuming
+
+Visual inference is a valid useful in diagnostics of predictive models
+
+To relieve human analysts of the job of diagnosing predictive models in the context of visual inference, computer vision
+
+Hence, to relieve human analysts of the job of diagnosing, automatically improve predictive models and assist vision-impaired people, this research aims to develop computer vision model to automate the reading
+1
+of data plots with application to predictive model diagnostics.
+The success of this research will enrich the applications of artificial intelligence in data visualization and improve the debugging process of artificial intelligence models.
+
+## Model Architecture
+
+Five computer vision models were trained to predict the probability of the given result plot being a null residual plot.
+
+## Performance

### 2022-03-06 Thesis/02-chap2.Rmd
add more intro
finalize 1 & 2

@@
 
-# Automatic Visual Statistical Inference, with Application to Linear Regression Diagnostics {#ch:paper1}
+<!-- # Automatic Visual Statistical Inference, with Application to Linear Regression Diagnostics {#ch:paper1} -->
 
-## Abstract
+<!-- ## Abstract -->
 
-## Introduction
+<!-- ## Introduction -->
 
-### Model Diagnostics
+<!-- ### Model Diagnostics -->
 
-[ET: suggestion: A model can be fitted to data with no guarantee of a meaningful interpretation.
+<!-- [ET: suggestion: A model can be fitted to data with no guarantee of a meaningful interpretation.
 Model diagnostics play an important role in assessing the appropriateness of the model.
-The assessment can involve examining the goodness of fit, checking if there are potential violations of model assumptions and validating models with external information.]
+The assessment can involve examining the goodness of fit, checking if there are potential violations of model assumptions and validating models with external information.] -->
 
-Model diagnostics is the part of data analysis, preceded by the fit of a model, whose primary objectives are to examine the goodness of fit and reveal potential violations of model assumptions.
+<!-- Model diagnostics is the part of data analysis, preceded by the fit of a model, whose primary objectives are to examine the goodness of fit and reveal potential violations of model assumptions.
+-->
 
-[ET: model fit is more than just goodness of fit and checking model assumptions.
+<!-- [ET: model fit is more than just goodness of fit and checking model assumptions.
 It's too reductive to say this because you could have a model with good fit (according to some metric) and no model violation but if the way that the data were collected was flawed then scientific interpretation of the model is rendered useless.
-Diagnostics involve also the domain knowledge checks as well.]
+Diagnostics involve also the domain knowledge checks as well.] -->
 
-In these diagnostics, though numeric summaries are mostly available and some are even endorsed by finite or asymptotic properties, graphical representation of data is still preferred, or at least needed, due to its intuitiveness and the possibility to provide unexpected discoveries which may be abstract and
-unquantifiable.
+<!-- In these diagnostics, though numeric summaries are mostly available and some are even endorsed by finite or asymptotic properties, graphical representation of data is still preferred, or at least needed, due to its intuitiveness and the possibility to provide unexpected discoveries which may be abstract and -->
+<!-- unquantifiable.  -->
 
-[ET: Generally, your sentences are too long.
+<!-- [ET: Generally, your sentences are too long.
 Break it up into smaller sentences.
@@
 Cloudy days are rare.
-Perhaps today will be different to other days!]
+Perhaps today will be different to other days!] -->
 
-However, unlike confirmatory data analysis built upon rigorous statistical procedures, e.g., hypothesis testing, visual diagnostics relies on graphical perception - human’s ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984], which is to some extent subjective.
+<!-- However, unlike confirmatory data analysis built upon rigorous statistical procedures, e.g., hypothesis testing, visual diagnostics relies on graphical perception - human’s ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984], which is to some extent subjective.
 Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
 One such example is finding a separation between gene groups in a two-dimensional projection from a linear discriminant analysis where there is no difference in the expression levels between the gene groups [@roy_chowdhury_using_2015].
+-->
 
-[ET: Confirmatory data analysis is well characterised by a rigorous procedure to discern particular hypothesis.
+<!-- [ET: Confirmatory data analysis is well characterised by a rigorous procedure to discern particular hypothesis.
 The most widespread form of confirmatory data analysis are the use of $p$-values in testing hypothesis in a frequentist framework.
@@
 This rigour, however, is often not present when inferences are drawn from a plot.
+-->
 
-Something about model diagnostics with a plot being the common thing to do..]
+<!-- Something about model diagnostics with a plot being the common thing to do..] -->
 
-### Visual Inference
+<!-- ### Visual Inference -->
 
-Visual inference was first introduced by @buja_statistical_2009 as an inferential framework to extend confirmatory statistics to visual discoveries.
+<!-- Visual inference was first introduced by @buja_statistical_2009 as an inferential framework to extend confirmatory statistics to visual discoveries.
 This framework redefines the test statistics, tests, null distribution, significance levels and $p$-value for visual discovery modelled on the confirmatory statistical testing.
 Figure \@ref(fig:fparallelism) outlines the parallelism between conventional tests and visual discovery.
+-->
 
-[ET: note use Rmd syntax for figures references! So it's easy to use it for HTML later if you want to do something like [Earo](https://github.com/earowang/thesis) and have PDF + HTML thesis. ]
+<!-- [ET: note use Rmd syntax for figures references!
+So it's easy to use it for HTML later if you want to do something like [Earo](https://github.com/earowang/thesis) and have PDF + HTML thesis.
+] -->
 
-![(ref:parallelism) \label{fig:parallelism}](figures/rsta2009012001.jpg){width=450 height=341}
+<!-- ![(ref:parallelism) \label{fig:parallelism}](figures/rsta2009012001.jpg){width=450 height=341} -->
 
-(ref:parallelism) Parallelism between multiple quantitative testing and visual discovery [@buja_statistical_2009].
+<!-- (ref:parallelism) Parallelism between multiple quantitative testing and visual discovery [@buja_statistical_2009]. -->
 
-In visual inference, a visual discovery is defined as a rejection of a null hypothesis, and the same null hypothesis can be rejected by many different visual discoveries [@buja_statistical_2009].
+<!-- In visual inference, a visual discovery is defined as a rejection of a null hypothesis, and the same null hypothesis can be rejected by many different visual discoveries [@buja_statistical_2009].
 For model diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the hypothesis.
 The same assumed model, such as classical linear regression model, can be rejected by both nonlinearity and heteroskedasticity with the residual plot as shown in Figure \@ref(fig:nonlinearityheter).
+-->
 
-### Pre-specification of Visual Discoverable Features
+<!-- ### Pre-specification of Visual Discoverable Features -->
 
-As discussed in @buja_statistical_2009, in the practice of model diagnostics, the range of possible visual discoveries is not pre-specified.
+<!-- As discussed in @buja_statistical_2009, in the practice of model diagnostics, the range of possible visual discoveries is not pre-specified.
 In other words, people do not explicitly specify which one or more visual features they are looking for before the read of the diagnostic plot.
@@
 Both definitions of visual test statistics are valid, but in the rest of the paper, the first definition will be used as it covered some details needed by this work.
+-->
 
-The size of the collection $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$ depends on the size of the set $I$.
+<!-- The size of the collection $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$ depends on the size of the set $I$.
 Thus, if one can define $I$ comprehensively, i.e, pre-specify all the visual discoverable features, the validity issue will be solved.
@@
 But this is an incomplete list which does not enumerate all the visual features.
+-->
 
-Similarly, @wilkinson_graph-theoretic_2005 proposed the work called graph theoretic scagnostics, which adopted the idea of "scagnostics" - scatter plot diagnostics from (can't find the 1984 citation).
+<!-- Similarly, @wilkinson_graph-theoretic_2005 proposed the work called graph theoretic scagnostics, which adopted the idea of "scagnostics" - scatter plot diagnostics from (can't find the 1984 citation).
 It includes 9 computable scagnostics measures defined on planar proximity graphs: "Outlying", "Convex", "Skinny", "Stringy", "Straight", "Monotonic", "Skewed", "Clumpy" and "Striated" which attempts to describe outliers, shape, density, trend and coherence of the data.
@@
 In fact, it is possible that such a list will never be complete as suggested in @buja_statistical_2009.
+-->
 
-Thinking out of the box, @buja_statistical_2009 argued that there is actually no need for pre-specification of visual discoverable features.
+<!-- Thinking out of the box, @buja_statistical_2009 argued that there is actually no need for pre-specification of visual discoverable features.
 In model diagnostics, when the null hypothesis is rejected, the reasons for rejecting the hypothesis will also be known.
@@
 This argument helps justifies the validity of visual inference.
+-->
 
-```{r nonlinearityheter, fig.cap="test"}
-library(tidyverse)
+<!-- ### Lineup Protocol -->
 
-set.seed(10086)
-cm <- visage::cubic_model(a = 0, b = 100, c = 1)
-
-cm$gen(1000, fit_model = TRUE) %>%
-cm$plot() +
-  ggtitle("Residuals against fitted values plot of a classical linear regression model") +
-  theme_light()
-```
-
-### Lineup Protocol
-
-With the validity of visual inference being justified, another aspect of hypothesis testing that needs to be addressed is the control of false positive rate or Type I error.
+<!-- With the validity of visual inference being justified, another aspect of hypothesis testing that needs to be addressed is the control of false positive rate or Type I error.
 Any visual statistic $T^{(i)}(\boldsymbol{\mathrm{y}})$ needs to pair with a critical value $c^{(i)}$ to form a hypothesis test.
@@
 Similarly, if any visual discovery is found by the observer, we say, there exists $i \in I:~T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}$ [@buja_statistical_2009].
+-->
 
-Using the above definition, the family-wise Type I error can be controlled if one can provide the collection of critical values $c^{(i)}~(i \in I)$ such that $P(\mathrm{there~exists~} i \in I: T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}|\boldsymbol{\mathrm{y}}) \leq \alpha$, where $\alpha$ is the significance level.
+<!-- Using the above definition, the family-wise Type I error can be controlled if one can provide the collection of critical values $c^{(i)}~(i \in I)$ such that $P(\mathrm{there~exists~} i \in I: T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}|\boldsymbol{\mathrm{y}}) \leq \alpha$, where $\alpha$ is the significance level.
 However, since the quantity of $T^{(i)}(\boldsymbol{\mathrm{y}})$ may not be known, such collection of critical values can not be provided.
+-->
 
-@buja_statistical_2009 proposed the lineup protocol as a visual test to calibrate the Type I error issue without the specification of $c^{(i)}~(i \in I)$.
+<!-- @buja_statistical_2009 proposed the lineup protocol as a visual test to calibrate the Type I error issue without the specification of $c^{(i)}~(i \in I)$.
 It is inspired by the "police lineup" or "identity parade" which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people.
@@
 Then, an observer who have not seen the actual data plot will be asked to point out the most different plot from the lineup.
+-->
 
-Under the null hypothesis, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is $1/m$ due to randomness.
+<!-- Under the null hypothesis, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is $1/m$ due to randomness.
 If we reject the null hypothesis as the observer correctly picks the actual data plot, then the Type I error of this test is $1/m$.
+-->
 
-This provides us with an mechanism to control the Type I error, because $m$ - the number of plots in a lineup can be chosen.
+<!-- This provides us with an mechanism to control the Type I error, because $m$ - the number of plots in a lineup can be chosen.
 Further, if we involve $K$ independent observers in a visual test, and let $X$ be a random variable denoting the number of observers correctly picking the actual data plot.
-Then, under the null hypothesis $X \sim \mathrm{Binom}_{K,1/m}$, and therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as
+Then, under the null hypothesis $X \sim \mathrm{Binom}_{K,1/m}$, and therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as -->
 
-$$P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},$$
-where $x$ is the realization of number of observers correctly picking the actual data plot [@majumder_validation_2013].
+<!-- $$P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},$$ -->
+<!-- where $x$ is the realization of number of observers correctly picking the actual data plot [@majumder_validation_2013]. -->
 
-### Visual Inference Applied to Linear Regression
+<!-- ### Visual Inference Applied to Linear Regression -->
 
-How people used visual inference in linear regression?
+<!-- How people used visual inference in linear regression? -->
 
-### Limitation of the Visual Inference
+<!-- ### Limitation of the Visual Inference  -->
 
-What are the limitations?
+<!-- What are the limitations? -->
 
-### Computer Vision Model
+<!-- ### Computer Vision Model -->
 
-What is computer vision model?
+<!-- What is computer vision model? -->
 
-### Contribution
+<!-- ### Contribution -->
 
-What has been done by this paper?
+<!-- What has been done by this paper? -->
 
-### Structure of This Paper
+<!-- ### Structure of This Paper -->
 
-What is the structure of the paper?
+<!-- What is the structure of the paper? -->
 
-Model diagnostics is the part of data analysis whose primary objectives are to examine the goodness of the model fit and reveal potential violations of the assumptions. Graphical approaches
+<!-- Model diagnostics is the part of data analysis whose primary objectives are to examine the goodness of the model fit and reveal potential violations of the assumptions. Graphical approaches   -->
 
-For regression diagnostics, it may includes the needs of
+<!-- For regression diagnostics, it may includes the needs of    -->
 
-Linear regression is an modelling approach to describe the relationship between an response variable and one or more explanatory variable.
+<!-- Linear regression is an modelling approach to describe the relationship between an response variable and one or more explanatory variable.
 It has been widely used for both generative modeling and predictive modelling.
+-->
 
-Regression diagnostics is needed
+<!-- Regression diagnostics is needed  -->
 
-1. to check whether the assumptions has been violated
-2. to check whether the line fit the data
+<!-- 1. to check whether the assumptions has been violated -->
+<!-- 2. to check whether the line fit the data -->
 
-Model diagnostics for linear regression is well developed
+<!-- Model diagnostics for linear regression is well developed    -->

### 2022-03-06 Thesis/_book/thesis.tex
finalize 1 & 2

@@
 
-\hypertarget{ai-four-approaches}{%
-\section{AI: Four Approaches}\label{ai-four-approaches}}
+\hypertarget{is-what-we-see-really-there}{%
+\section{Is what we see really there?}\label{is-what-we-see-really-there}}
+
+\hypertarget{artificial-intelligence-four-approaches}{%
+\section{Artificial intelligence: Four Approaches}\label{artificial-intelligence-four-approaches}}
 
@@
 Based on the answer to the above questions, four major approaches to pursue AI have been established.
-These approaches can be summarized into a two by two table as shown in Figure \ref{fig:twodimai}, where the row is ``Human'' vs.~``Rational'', and the column is ``Behaviour'' vs.~``Thought''.
+These approaches can be summarized into a two by two table as shown in Figure \ref{fig:two-dim-ai}, where the row is ``Human'' vs.~``Rational'', and the column is ``Behaviour'' vs.~``Thought''.
 Positioning at the top right cell, the \textbf{rational agent approach} aims to build agent that perform mathematically perfect acts such that the best expected outcome can always be achieved.
@@
 One such example is the \textbf{total Turing test} proposed by \textcite{harnad_other_1991}.
-It adds \(3\) new requirements to the list, including \textbf{computer vision}, \textbf{speech recognition} and \textbf{robotics}, which are response for interactions with the physical world.
-Notably, all \(7\) required capabilities have become major subfields of AI today.
+It adds three new requirements to the list, including \textbf{computer vision}, \textbf{speech recognition} and \textbf{robotics}, which are response for interactions with the physical world.
+Notably, all seven required capabilities have become major subfields of AI today.
 And their development has made AI one of the fastest-growing fields in the 21st century \autocite{russell_artificial_2002}.
@@
 \centering
-\includegraphics{thesis_files/figure-latex/twodimai-1.pdf}
-\caption{\label{fig:twodimai}Four possible approaches to pursue AI based on the two dimensions in AI research - human vs.~rational and thought vs.~behavior \autocite{russell_artificial_2002}.}
+\includegraphics{thesis_files/figure-latex/two-dim-ai-1.pdf}
+\caption{\label{fig:two-dim-ai}Four possible approaches to pursue AI based on the two dimensions in AI research - human vs.~rational and thought vs.~behavior \autocite{russell_artificial_2002}.}
 \end{figure}
@@
 Further, the increasingly cheap computing cost and the massive amount of accessible e-commerce data produced in the Internet age provide the possibility for applying data-intensive AI models, which enables AI performance to reach new heights in history \autocite{jordan_machine_2015}.
-Some AI systems have already been remarkably better than human in certain areas, e.g., game playing.
-AlphaGo and AlphaZero developed by the Google DeepMind team surpass all human Go players \autocite{silver_general_2018}.
+Some AI systems have already been remarkably better than humans in certain areas, e.g., game playing.
+AlphaGo and AlphaZero developed by the Google DeepMind team surpass professional human Go players \autocite{silver_general_2018}.
 
@@
 Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
-One such example is finding a separation between gene groups in a two-dimensional projection from a linear discriminant analysis where there is no difference in the expression levels between the gene groups \autocite{roy_chowdhury_using_2015}.
+One such example is finding an over-interpretation of the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurence \autocite{roy_chowdhury_using_2015}.
 
@@
 
-Visual inference was first introduced by \textcite{buja_statistical_2009} as an inferential framework to extend confirmatory statistics to visual discoveries.
+Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title ``Inference for Data Visualizatino'' by \textcite{buja_inference_1999} as an idea to address the issue of valid inference for visual discoveries of data plots \autocite{gelman_exploratory_2004}.
+Later, in the Bayesian context, data plots was systematically considered as model diagnostics by taking advantage of the data simulated from the assumed statistical models \autocite{gelman_bayesian_2003,gelman_exploratory_2004}.
+
+It was surprising that the essential components of visual inference had actually been established in \textcite{buja_inference_1999}, but it was not until 10 years later that \textcite{buja_statistical_2009} formalized it as an inferential framework to extend confirmatory statistics to visual discoveries.
 This framework redefines the test statistics, tests, null distribution, significance levels and \(p\)-value for visual discovery modelled on the confirmatory statistical testing.
@@
 \caption{Parallelism between multiple quantitative testing and visual discovery \autocite{buja_statistical_2009}.
-Visible features in a plot are viewed as a collection of test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\), and any visual discoveries are treated as evidence against the null hypothesis.
+Visible features in a plot are viewed as a collection of test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\), and any visual discoveries that are inconsistent with the null hypothesis are treated as evidence against the null.
+For model diagnostics, the null hypothesis would be the assumed model, and visual discoveries could be any visual features in favour of any alternatives.
 \label{fig:parallelism}}
@@
 For model diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the null hypothesis.
-The same assumed model, such as classical linear regression model, can be rejected by many reasons with residual plot, including nonlinearity and heteroskedasticity as shown in Figure \ref{fig:nonlinearityheter}.
-
-\begin{figure}
-\centering
-\includegraphics{thesis_files/figure-latex/nonlinearityheter-1.pdf}
-\caption{\label{fig:nonlinearityheter}Residuals vs.~fitted values plot for a classical linear regression model.
+The same assumed model, such as classical linear regression model, can be rejected by many reasons with residual plot, including nonlinearity and heteroskedasticity as shown in Figure \ref{fig:residual-plot-cubic-heter}.
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/residual-plot-cubic-heter-1.pdf}
+\caption{\label{fig:residual-plot-cubic-heter}Residuals vs.~fitted values plot for a classical linear regression model.
 The residuals are produced by fitting a two-predictor multiple linear regression model with data generated from a cubic linear model.
@@
 As discussed in \textcite{buja_statistical_2009}, in the practice of model diagnostics, the range of possible visual discoveries is not pre-specified.
-In other words, people do not explicitly specify which one or more visual features they are looking for before the read of the diagnostic plot.
-This is concerning since conventional hypothesis testing always requires the pre-specification of the parameter space \(\Theta\) of the parameter of interest \(\theta \in \Theta\) to form a valid inferential procedure.
-To address this issue, a collection of test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) is defined, where \(\boldsymbol{\mathrm{y}}\) is the data and \(\boldsymbol{I}\) is a set of all possible visual features.
+In other words, people do not explicitly specify which visual feature(s) they are looking for before the reading the diagnostic plot.
+In contrary, conventional hypothesis testing always requires the pre-specification of the parameter space \(\Theta\) of the parameter of interest \(\theta \in \Theta\) to form a valid inferential procedure.
+To address this issue, a collection of test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) is defined, where \(\boldsymbol{\mathrm{y}}\) is the data and \(I\) is a set of all possible visual features.
 \textcite{buja_statistical_2009} described each of the test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})\) as a measurement of the degree of presence of a visual feature.
@@
 
+\hypertarget{sampling-from-the-null-distribution}{%
+\section{Sampling from the Null Distribution}\label{sampling-from-the-null-distribution}}
+
+In visual inference, the null distribution of plots refers to the infinite collection of plots of null datasets sampled from \(H_0\).
+It is defined as the analogue of the null distribution of test statistics in conventional test \autocite{buja_statistical_2009}.
+In practice, a finite number of plots of null datasets could be generated, called null plots.
+
+In the context of model diagnostics, sampling data from \(H_0\) is equivalent to sampling data from the assumed model.
+As \textcite{buja_statistical_2009} suggested, \(H_0\) is usually composited by a collection of distributions controlled by nuisance parameters.
+Since statistical models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to so called ``reference distribution'' by applying one of the three methods: (i) sampling from a conditional distribution given a minimimal sufficient statistic under \(H_0\), (ii) parametric bootstrap sampling with nuisance parameters estimated under \(H_0\), and (iii) Bayesian posterior predictive sampling.
+
+The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three \autocite{buja_statistical_2009}.
+Focus will be put on simulation based on this distribution as it is applicable to classical normal linear regression model.
+Details of the other two are can be found in the supplementary material of \textcite{buja_statistical_2009}.
+Suppose there exists a minimal sufficient statistic \(\boldsymbol{S}(\boldsymbol{y})\) under the null hypothesis, any null datasets \(\boldsymbol{y^{*}}\) should fulfil the condition \(\boldsymbol{S}(\boldsymbol{y}) = \boldsymbol{s}\).
+For classical normal linear regression model, the minimal sufficient statistic is \(\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{\beta}}, \boldsymbol{e}'\boldsymbol{e})\), where \(\hat{\boldsymbol{\beta}}\) are the coefficient estimators and \(\boldsymbol{e}'\boldsymbol{e}\) is the residual sum of square.
+Alternatively, the minimal sufficient statistic can be constructed as \(\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{y}}, ||\boldsymbol{e}||)\), where \(\hat{\boldsymbol{y}}\) are the fitted values and \(||\boldsymbol{e}||\) is the length of residuals, which is more intuitive as suggested by \textcite{buja_statistical_2009}.
+Since the fitted values are held fixed, the variation can only occur in the residual space.
+And because the length of residual is also held fixed, residuals obtained from a null dataset has to be a random rotation of \(\boldsymbol{e}\) in the residual space.
+With this property, null residuals can be simulated by regressing \(N\) i.i.d standard normal random draws on the regressors, then rescaling it by the ratio of residual sum of square in two regressions.
+
 \hypertarget{lineup-protocol}{%
@@
 
-With the validity of visual inference being justified, another aspect of hypothesis testing that needs to be addressed is the control of false positive rate or Type I error.
+With the validity of visual inference being justified and the simulation of null plots being provided, another aspect of hypothesis testing that needs to be addressed is the control of false positive rate or Type I error.
 Any visual statistic \(T^{(i)}(\boldsymbol{\mathrm{y}})\) needs to pair with a critical value \(c^{(i)}\) to form a hypothesis test.
@@
 It is inspired by the ``police lineup'' or ``identity parade'' which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people.
-The protocol consists of \(m\) randomly placed data plots, where \(1\) plot is the actual data plot, and \(m-1\) null plots are produced by plotting data simulate from the null distribution which is consistent with the null hypothesis.
+The protocol consists of \(m\) randomly placed plots, where one plot is the actual data plot, and the remaining \(m - 1\) plots have the identical graphical production as the data plot except the data has been replaced with data consistent with the null hypothesis.
 Then, an observer who have not seen the actual data plot will be asked to point out the most different plot from the lineup.
 
-Under the null hypothesis, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is \(1/m\) due to randomness.
+Under the null hypothesis, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is \(1/m\).
 If we reject the null hypothesis as the observer correctly picks the actual data plot, then the Type I error of this test is \(1/m\).
@@
 This provides us with an mechanism to control the Type I error, because \(m\) - the number of plots in a lineup can be chosen.
-A larger value of \(m\) will result in a smaller Type I error, but the limit to the value of \(m\) depends on the number of plots a human willing to view \autocite{buja_statistical_2009}.
+A larger value of \(m\) will result in a smaller Type I error, but the limit to the value of \(m\) depends on the number of plots a human is willing to view \autocite{buja_statistical_2009}.
 Typically, \(m\) will be set to \(20\) which is equivalent to set \(\alpha = 0.05\), a general choice of significance level for conventional testing among statisticians.
@@
 With just \(4\) individuals and \(20\) data plots in a lineup, the \(p\)-value could be as small as \(0.0001\).
+Additionally, by involving multiple observers, variation of individual ability to read plots can be addressed to some degree as different opinions about visual discoveries can be collected.
 
@@
 
-Assumes that individuals have the same ability and a lineup has been evaluated by multiple individuals, under the alternative hypothesis, the estimated power for a lineup can be expressed as \(\hat{p} = x/K\), the estimated probability of identifying the actual data plot from the lineup.
-If the individual skill needs to be taken into account, and \(L\) lienups have been evaluated by \(K\) individuals, \textcite{majumder_validation_2013} suggests that mixed effects logistic regression model can be fit as:
+Suppose individuals have the same ability and a lineup has been evaluated by multiple individuals, under the alternative hypothesis, the estimated power for a lineup can be expressed as \(\hat{p} = x/K\), the estimated probability of identifying the actual data plot from the lineup.
+If the individual skill needs to be taken into account, and \(L\) lineups have been evaluated by \(K\) individuals, \textcite{majumder_validation_2013} suggests that mixed effects logistic regression model can be fit as:
 
@@
 
-\textbf{null data generate mechanism}
-
-\hypertarget{applications-of-lineup-protocol}{%
-\section{Applications of Lineup Protocol}\label{applications-of-lineup-protocol}}
-
-Lineup protocols are
-
-List applications, fields, significance here.
-
-\hypertarget{limitations-of-visual-tests}{%
-\section{Limitations of visual tests}\label{limitations-of-visual-tests}}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  infeasible in a large scale
-\item
-  unfriendly to vision-impaired people
-\item
-  high finical cost and human cost
-\item
-  time consuming
-\end{enumerate}
-
-similar to Handicraft in pre-industrial society
-
-\hypertarget{automatic-visual-inference---computer-vision}{%
-\section{Automatic visual inference -\textgreater{} Computer vision}\label{automatic-visual-inference---computer-vision}}
-
-relieve people's workload by automating repeating tasks, and provide standard result in a control environment
-
-the use of technology and machinery to enable mass evaluation of visual tests
-
-\hypertarget{discussion-of-potential-methods}{%
-\section{Discussion of potential methods}\label{discussion-of-potential-methods}}
-
-\hypertarget{different-approaches-of-ai}{%
-\subsection{different approaches of AI}\label{different-approaches-of-ai}}
-
-\begin{itemize}
-\tightlist
-\item
-  Not aim for understanding the thought process
-\item
-  Not aim for mimicking the human vision mechanism
-\item
-  May be able to define distance metrics to measure difference between data plots for making mathematically prefect decisions
-\item
-  May be able to use computer vision model to approximate how people evaluate lineups
-\end{itemize}
-
-computer vision model:
-
-\begin{itemize}
-\tightlist
-\item
-  Use human data to train model with human selection as target -\textgreater{} mimic the human behaviour
-\item
-  Use simulated lineup to train model with actual data plot as target -\textgreater{} assume the actual data plot is the most different one
-\item
-  Use simulated data plot to train model with null or not null as target -\textgreater{} limited to the alternatives given to the model, and Type I error can not be controlled
-\item
-  Use simulated data plot to train model, then let the model evaluate each plot of a lineup -\textgreater{} leads to multiple selections in a lineup, and the model does not compare different plots in a lineup
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Use a very large \(m\) to ensure the null hypothesis can be rejected by a classifier slightly better than a random selector. The problem becomes how large the \(m\) should be
-  \item
-    Use multiple models as multiple individuals to evaluate lineup -\textgreater{} Which models should be used? How many models should be used?
-  \end{itemize}
-\item
-  Let the model select multiple plots from a simulated lineup while comparing different plots in a lineup -\textgreater{} how to build such a system?
-
-  \begin{itemize}
-  \tightlist
-  \item
-    The input is \(m\) plots, what is the output? The probability of being the most different one?
-  \item
-    Then how many plots should be selected? Top 5? Or cumulative probability greater than a threshold?
-  \item
-    How to select the threshold?
-  \end{itemize}
-\end{itemize}
-
-\hypertarget{human-subject-exeriments-toc}{%
-\chapter{Human Subject Exeriments (toc)}\label{human-subject-exeriments-toc}}
-
-What needs to be discussed?
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  online recruitment platform (Prolific)
-\end{enumerate}
-
-\begin{itemize}
-\tightlist
-\item
-  pre-screen
-\item
-  approve and reject
-\item
-  redirect
-\item
-  payment
-\end{itemize}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{1}
-\tightlist
-\item
-  study web site
-\end{enumerate}
-
-\begin{itemize}
-\tightlist
-\item
-  PythonAnywhere
-\item
-  Flask, Python
-\item
-  jsPsych
-\item
-  survey
-
-  \begin{itemize}
-  \tightlist
-  \item
-    age
-  \item
-    education
-  \item
-    preferred pronoun
-  \item
-    previous experience
-  \end{itemize}
-\item
-  training
-\item
-  20 lineups
-
-  \begin{itemize}
-  \tightlist
-  \item
-    setup depends on the study
-  \end{itemize}
-\end{itemize}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{2}
-\tightlist
-\item
-  Cubic model
-\end{enumerate}
-
-\begin{itemize}
-\tightlist
-\item
-  data generating process
-\item
-  null model
-\item
-  potential shapes
-\item
-  effect size
-\item
-  conventional test
-\end{itemize}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{3}
-\tightlist
-\item
-  Heteroskedasticity model
-\end{enumerate}
-
-\begin{itemize}
-\tightlist
-\item
-  data generating process
-\item
-  null model
-\item
-  potential shapes
-\item
-  effect size
-\item
-  conventional test
-\end{itemize}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{3}
-\tightlist
-\item
-  Number of participants needed by the study
-\item
-  Study 1
-\end{enumerate}
-
-\begin{itemize}
-\tightlist
-\item
-  logistic regression
-
-  \begin{itemize}
-  \tightlist
-  \item
-    easy
-  \item
-    medium
-  \item
-    hard
-  \end{itemize}
-\item
-  number of participants
-\item
-  time
-\item
-  countries
-\end{itemize}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{3}
-\tightlist
-\item
-  Study 2
-\end{enumerate}
-
-\begin{itemize}
-\tightlist
-\item
-  logistic regression
-
-  \begin{itemize}
-  \tightlist
-  \item
-    easy
-  \item
-    medium
-  \item
-    hard
-  \end{itemize}
-\item
-  number of participants
-\item
-  time
-\item
-  countries
-\end{itemize}
+\hypertarget{limitations-of-lineup-protocol-and-automatic-visual-inference}{%
+\section{Limitations of Lineup Protocol and Automatic Visual Inference}\label{limitations-of-lineup-protocol-and-automatic-visual-inference}}
+
+Although lineup protocol has already been integrated into data analysis of various topics, such as diagnostics of hierarchical linear models \autocite{loy_diagnostic_2013}, geographical Research \autocite{widen_graphical_2016} and forensic examinations \autocite{krishnan_hierarchical_2021}, the involvement with human judgements limits its popularity.
+Similar to handicraft in pre-industrial society, lineup protocol conducted by humans is infeasible on a large scale, high in labour cost and time consuming.
+Moreover, it is strongly unfriendly to vision-impaired people.
+
+The ``steam engine'' of visual inference needs to developed for relieving people's workload by automating repeating tasks and providing standard result in a control environment.
+Large-scale evaluation of visual tests is not possible without the use of technology and machines.
+
+Modern computer vision model could be a promising solution to this problem.
+As a subfield of AI, computer vision with the modern deep learning architectures solved numerous critical problems in automation.
+Inspired by the vision processing in living organisms, the convolutional neural network was introduced by \textcite{fukushima_neocognitron_1982}.
+Soon, this architecture was applied to hand-written number recognition trained with back-propagation by \textcite{lecun_backpropagation_1989}.
+This was one of the earliest attempts which human successfully extract information from digital images via self-learning algorithms.
+Modern computer vision model is typically built on the deep neural network with convolutional layers \autocite{fukushima_neocognitron_1982}.
+Convolutional layers take advantage of the hierarchical pattern in data and provide regularized versions of fully-connected layers.
+It downscales and transforms the image by summarising information in a small space.
+Numerous studies have shown that it can be used to effectively tackle vision tasks, such as image recognition \autocite{rawat_deep_2017}.
+With the development of graphics processing units and the spread of high-performance personal computers, researches in computer vision become a new hype in the 21st century.
+Achievements such as computer-aided diagnosis \autocite{lee_image_2015}, pedestrian detection \autocite{brunetti_computer_2018} and facial recognition \autocite{emami_facial_2012} had a significant impact on our daily life.
+
+Using computer vision models to read data plots is not a general choice.
+some fields have adopted this idea by applying computer vision models to read recurrence plots for time series regression \autocite{ojeda_multivariate_2020}, time series classification \autocite{chu_automatic_2019,hailesilassie_financial_2019,hatami_classification_2018,zhang_encoding_2020}, anomaly detection \autocite{chen_convolutional_2020} and pairwise causality analysis \autocite{singh_deep_2017}.
+However, evaluating lineups with computer vision model is a new field of study.
+
+(reasons for conducting experiments)
+
+\hypertarget{structure-of-the-paper}{%
+\section{Structure of the paper}\label{structure-of-the-paper}}
+
+\ldots{}
 
@@
 Each lineup consists of one actual residual plot and 19 null residual plots produced by plotting null residuals simulated from the residual rotation distributions.
-Among the 20 lineups, there are two extremely easy lineups used as attention checks which everyone should get correct.
+Among the 20 lineups, there were two extremely easy lineups used as attention checks which everyone should get correct.
 The short survey was intended to collect information about participant that might affect their ability in reading data plot including age, highest level of education, preferred pronoun and previous experience in similar study.
 For the evaluation of the lineup, participant first needed to select one or multiple most different data plot(s) from the lineup by clicking the corresponding image.
-Then, the reason for choosing the plots needed to be provided by selecting one of the given options - ``outlier(s)'', ``cluster(s)'', ``shape'' and ``other''.
+Then, the primary reason for choosing the plots needed to be provided by picking one of the given options - ``outlier(s)'', ``cluster(s)'', ``shape'' and ``other''.
 Explanations about these options were provided in the training page and the mouseover text.
-Table \ref{tab:lineupreason} gives the detailed explanations about these reasons.
+Table \ref{tab:lineup-reason} gives the detailed explanations about these reasons.
 If the option ``other'' was selected, text input for the specified reason would be collected.
-Lastly, the degree of difference between their chosen data plot(s) and other data plots needed to be selected among \(5\) levels - ``not at all'', ``slightly'', ``moderately'', ``very'' and ``extremely''.
+Lastly, the degree of difference between their chosen data plot(s) and other data plots needed to be selected among five levels - ``not at all'', ``slightly'', ``moderately'', ``very'' and ``extremely''.
 Notably, if participants could not tell the difference between the data plots, there was an option to skip the evaluation of the lineup.
@@
 
-\caption{\label{tab:lineupreason}Explanations about reasons for choosing data plots from a lineup}
+\caption{\label{tab:lineup-reason}Explanations about reasons for choosing data plots from a lineup}
 \centering
@@
 This is essential for us to confirm the quality of the data by checking exceptionally fast or slow responses.
-Figure \ref{fig:DiagramExperiment} summarizes the online experiment setup.
+Figure \ref{fig:diagram-experiment} summarizes the online experiment setup.
 
@@
 And the client-side used jsPsych to run experiment.
-\label{fig:DiagramExperiment}}
+\label{fig:diagram-experiment}}
 \end{figure}
@@
 The first two pages of the website are the explanatory statement and the consent form.
-Participants needed to read the documents and agree with the terms to advance to the short survey.
-With the completion of the survey, the next page is the training page, which contained instructions on how to evaluate the lineup, give the reason and choose the confidence level.
+Participants were asked to read the documents and agree with the terms to advance to the short survey.
+With the completion of the survey, the next page is the training page, which contained instructions on how to interact with the webpage.
 Followed by the training page, there are 20 lineups each on a single page.
-Figure \ref{fig:WebsiteLayout} illustrates the layout of the website.
-At the end of the experiment, the participants would be redirected back to Prolific and waited for researchers' responses.
+Figure \ref{fig:website-layout} illustrates the layout of the website.
+At the end of the experiment, participants would be redirected back to Prolific and waited for researchers' responses.
 
@@
 \includegraphics[width=5.72917in,height=3.11458in]{figures/website.png}
-\caption{Layout of the study website. Participants needed to choose the most different plots on the right and select their reasons and confidence levels on the left. \label{fig:WebsiteLayout}}
+\caption{Layout of the study website. Participants needed to choose the most different plots on the right and select their reasons and confidence levels on the left. \label{fig:website-layout}}
 \end{figure}
@@
 
-Two models were used in both experiments, which were linear models with some degree of violations of classical assumptions. Residual vs.~fitted values plot were used to present the model defects.
+Two models were used in the study, where both were linear models with some degree of violations of classical assumptions.
 
@@
 
-Lemma \ref{lemma:cubic} shows that the expectation of the residuals is clearly a function of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\).
-Hence, we would expect to see some patterns in the residual plot.
+Lemma \ref{lemma:cubic} shows that the expectation of the residuals is clearly a function of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\), which indicates the residuals and the fitted values \(\hat{Y} = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'\boldsymbol{Y}\) are associated.
+Hence, it is expected that visual discoveries can be found in the residuals vs.~fitted values plot.
 Let \(X_i\) and \(Z_i\), \(i = 1,...,n\), be independent random variables follow uniform distribution \(U(-1,1)\).
 Given the expectation of the residuals, we could plot the expected values of residuals against the observed values.
-Figure \ref{fig:cubic1}, \ref{fig:cubic2} and \ref{fig:cubic3} illustrate the shape of residuals and their expected values under different parameter settings.
-
-From the Figure \ref{fig:cubic1}, it can be observed that with fixed \(\sigma\) and \(c\), \(a\) and \(b\) are controlling the 2D projection of a hypersurface, and seemingly performing some rotations along different axes.
-Figure \ref{fig:cubic2} shows that with fixed \(a\), \(b\) and \(\sigma\), \(c\) is controlling the contribution of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) to \(\boldsymbol{Y}\).
-If we move \(c\) toward \(0\) or \(2\), one variable will donminate another, which will mitigate some joint effects and resemble a typical cubic function.
-In Figure \ref{fig:cubic3}, \(a\), \(b\) and \(c\) is held constant, and \(\sigma\) is controlling the noises around the expected values.
-As \(\sigma\) decreases, the underlying shape shows up.
-
-\begin{figure}
-\centering
-\includegraphics{thesis_files/figure-latex/cubic1-1.pdf}
-\caption{\label{fig:cubic1}cubic 1}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{thesis_files/figure-latex/cubic2-1.pdf}
-\caption{\label{fig:cubic2}cubic 2}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{thesis_files/figure-latex/cubic3-1.pdf}
-\caption{\label{fig:cubic3}cubic 3}
+Figure \ref{fig:cubic-shape-1}, \ref{fig:cubic-shape-2} and \ref{fig:cubic-shape-3} illustrate the shape of residuals and their expected values under different parameter settings.
+
+From the Figure \ref{fig:cubic-shape-1}, it can be observed that with fixed \(\sigma\) and \(c\), \(a\) and \(b\) are controlling the 2D projection of a hypersurface, and seemingly performing some rotations along different axes.
+Figure \ref{fig:cubic-shape-2} shows that with fixed \(a\), \(b\) and \(\sigma\), \(c\) is controlling the contribution of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) to \(\boldsymbol{Y}\).
+As \(c\) moves toward \(0\) or \(2\), one regressor will dominate another, which will mitigate some joint effects and resemble a typical cubic function.
+In Figure \ref{fig:cubic-shape-3}, \(a\), \(b\) and \(c\) are held fixed, and \(\sigma\) is controlling the noises around the expected values.
+The underlying shape is clearer for smaller \(\sigma\).
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/cubic-shape-1-1.pdf}
+\caption{\label{fig:cubic-shape-1}A matrix of residuals vs.~fitted values plot under different parameter settings of the cubic model.
+Resdiauls are drawns in black and the expected values are drawn in red.
+The plots show that if \(\sigma\) and \(c\) are held fixed, \(a\) and \(b\) rotate a two-dimension projection of a hypersurface.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/cubic-shape-2-1.pdf}
+\caption{\label{fig:cubic-shape-2}A matrix of residuals vs.~fitted values plot under different parameter settings of the cubic model.
+Resdiauls are drawns in black and the expected values are drawn in red.
+The plots show that if \(a\), \(b\) and \(\sigma\) are held fixed, \(c\) controls the contribution of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) to \(\boldsymbol{Y}\).}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/cubic-shape-3-1.pdf}
+\caption{\label{fig:cubic-shape-3}A matrix of residuals vs.~fitted values plot under different parameter settings of the cubic model.
+Resdiauls are drawns in black and the expected values are drawn in red.
+The plots show that if \(a\), \(b\) and \(c\) are held fixed, \(\sigma\) controls the strength of the signal.}
 \end{figure}
@@
 
-Here, we set \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) to be random uniform variables ranged from \(-1\) to \(1\).
-This can be done by using the random uniform variable class constructor \texttt{rand\_uniform()} inherited from the random variable abstract base class.
+If we would like \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) to be random uniform variables ranged from \(-1\) to \(1\), it can be achieved by using the random uniform variable class constructor \texttt{rand\_uniform()} inherited from the random variable abstract base class.
 It only takes two arguments which are the lower bound and the upper bound of the support.
@@
 ##            y          x          z           e     .resid    .fitted
-## 1  0.44298045  0.1862316 -0.5114228  0.2749620  0.3561304  0.08685009
-## 2 -0.46546366 -0.8542129 -0.3789133  0.3543143  0.2827120 -0.74817570
-## 3  0.07636314  0.0383946 -0.3832281 -0.3024668 -0.2698449  0.34620801
-## 4 -1.04762148 -0.4177674 -0.5876853 -0.3101100 -0.1202573 -0.92736416
-## 5 -0.90198583 -0.5439192 -0.4630171 -0.1448947 -0.2487402 -0.65324563
+## 1 -0.8263671 -0.4988275  0.5569648 -0.06142887  1.2340651 -2.0604322
+## 2  0.6134661  0.1762335  0.6512424  0.99654342  1.1758970 -0.5624310
+## 3  0.2285071 -0.2814395 -0.1511556 -0.10996690 -0.3757603  0.6042674
+## 4 -4.6320382 -0.5813456  0.9251409 -0.60826338 -1.2664307 -3.3656075
+## 5  0.3148180  0.5647967  0.4406211  0.64596536 -0.7677711  1.0825892
 \end{verbatim}
 
-Method \texttt{mod\$plot()} produce a ggplot \autocite{wickham_ggplot2_2011} object.
+Method \texttt{mod\$plot()} produce a \texttt{ggplot} \autocite{wickham_ggplot2_2011} object.
 It takes a data frame containing columns \texttt{.resid} and \texttt{.fitted} as input, along with a character argument \texttt{type} indicating the type of the data plot, and other aesthetic arguments such as \texttt{size} and \texttt{alpha} to control the appearance of the plot.
@@
 
-Lineup can be produced by using the methods \texttt{mod\$gen\_lineup()} and \texttt{mod\$plot\_lineup()}.
+Lineup is a matrix of residual plots which can be produced by using the methods \texttt{mod\$gen\_lineup()} and \texttt{mod\$plot\_lineup()}.
 Method \texttt{mod\$gen\_lineup} takes the number of realizations \texttt{n} and the number of plots in a lineup \texttt{k} as inputs.
@@
 \begin{verbatim}
-## [1] -0.9369269  0.1945678  0.2816923 -0.3254415  0.7861083
+## [1] -0.092183951 -0.003912092 -0.001531723  0.148353478 -0.050725712
 \end{verbatim}
@@
 \includegraphics{thesis_files/figure-latex/hetershape-1.pdf}
-\caption{\label{fig:hetershape}test}
+\caption{\label{fig:hetershape}A matrix of residuals vs.~fitted values plot under different parameter settings of the heteroskedasticity model.
+Resdiauls are drawns in black and the one standard deviation around zero is drawn in red.
+The plots show that \(a\) controls the shape and direction of the resdiaul plot, while \(b\) controls the strength of the signal.}
 \end{figure}
@@
 The distribution of \(Z_i\), \(i=1,...,n\), had \(50\)\% chance to be a uniform distribution ranged from \(-1\) to \(1\), and \(50\)\% chance to be a discrete uniform distribution with \(z_n\) outcomes simulated from a uniform distribution ranged from \(-1\) to \(1\).
-\(z_n\) itself was a discrete uniform distribution with outcomes \(\{3,4,5,6,7,8,9,10\}\), which defined the number of possible values \(Z_i\) could take.
+\(z_n\) itself was a discrete uniform distribution with outcomes \(\{10,11,12,13,14,15,16,17,18,19,20\}\), which defined the number of possible values \(Z_i\) could take.
 As shown in Figure \ref{fig:discreteex}, this setup would create discreteness in residual plot, which could enrich the pool of visual patterns.
@@
 \begin{table}[!htbp] \centering 
-  \caption{GLM result 1} 
+  \caption{Logistic regressions with detection rate as response variable and natural logarithm of the effect size as regressor fitted on the pilot data for both cubic and heteroskedasticity models.} 
   \label{tab:pilotglm} 
@@
 \includegraphics{thesis_files/figure-latex/pilotglm-1.pdf}
-\caption{\label{fig:pilotglm}pilot}
+\caption{\label{fig:pilotglm}Logistic regressions with detection rate as response variable and natural logarithm of the effect size as regressor fitted on the pilot data for both cubic and heteroskedasticity models.
+Points are jittered to reduce overplotting using the \texttt{ggbeeswarm} package.
+The smooth curves are the fitted regression curves.
+The ribbons around the curves are the 95\% confidence intervals.}
 \end{figure}
@@
 \includegraphics{thesis_files/figure-latex/e1glmplot-1.pdf}
-\caption{\label{fig:e1glmplot}test}
+\caption{\label{fig:e1glmplot}Logistic regressions with detection rate as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I for both cubic and heteroskedasticity models.
+Points are jittered to reduce overplotting using the \texttt{ggbeeswarm} package.
+The smooth curves are the fitted regression curves.
+The ribbons around the curves are the 95\% confidence intervals.}
 \end{figure}
@@
 \begin{table}[!htbp] \centering 
-  \caption{GLM result 2} 
+  \caption{Logistic regressions with detection rate as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I for both cubic and heteroskedasticity models.}
   \label{tab:e1glm} 
@@
 \hline \\[-1.8ex] 
- I(log(effect\_size)) & 0.818$^{***}$ & 0.710$^{***}$ \\ 
-  & (0.113) & (0.107) \\ 
-  & & \\ 
- Constant & $-$1.649$^{***}$ & $-$3.425$^{***}$ \\ 
-  & (0.254) & (0.454) \\ 
-  & & \\ 
-\hline \\[-1.8ex] 
-Observations & 201 & 199 \\ 
-Log Likelihood & $-$68.678 & $-$89.985 \\ 
-Akaike Inf. Crit. & 141.357 & 183.971 \\ 
+ log(effect\_size) & 0.857$^{***}$ & 0.746$^{***}$ \\ 
+  & (0.122) & (0.111) \\ 
+  & & \\ 
+ Constant & $-$1.571$^{***}$ & $-$3.434$^{***}$ \\ 
+  & (0.257) & (0.462) \\ 
+  & & \\ 
+\hline \\[-1.8ex] 
+Observations & 192 & 188 \\ 
+Log Likelihood & $-$64.544 & $-$84.606 \\ 
+Akaike Inf. Crit. & 133.089 & 173.212 \\ 
 \hline 
@@
 
-We collected \(400\) lineup evaluations made by \(20\) participants in experiment I and \(880\) lineup evaluations made by \(44\) participants.
+We collected \(400\) lineup evaluations made by \(20\) participants in experiment I and \(880\) lineup evaluations made by \(44\) participants in experiment II.
 In total, \(442\) unique lineups were evaluated by \(64\) subjects.
 In experiment I, one of the participants skipped all \(20\) lineups.
-Hence, the submission was rejected and removed from the data set.
+Hence, the submission was rejected and removed from the dataset.
 In experiment II, there was a participant failed one of the two attention checks, but there was no further evidence of low-effort throughout the experiment.
-So, the submission was kept.
+Therefore, the submission was kept.
 
@@
 Using the power calculation method discussed in Section \ref{p-value-and-power-estimation-of-visual-test-allowed-for-multiple-selections}, power of each lineup from one to five evaluations was estimated.
-The estimated power was further used in fitting a quasibinomial generalized linear model
-
-\begin{figure}
-\centering
-\includegraphics{thesis_files/figure-latex/unnamed-chunk-26-1.pdf}
-\caption{\label{fig:unnamed-chunk-26}test}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{thesis_files/figure-latex/unnamed-chunk-27-1.pdf}
-\caption{\label{fig:unnamed-chunk-27}test}
-\end{figure}
-
-(estimate of visual power)
-
-(power against different parameters)
-
-(power against conventional p-value)
-
-(conventional power)
-
-(visual p-value vs.~conventional p-value)
-
-(response time)
-
-(confidence level)
-
-(relative freq of picks)
-
-\clearpage
-
-\hypertarget{ch:paper1}{%
-\chapter{Automatic Visual Statistical Inference, with Application to Linear Regression Diagnostics}\label{ch:paper1}}
-
-\hypertarget{abstract-1}{%
-\section{Abstract}\label{abstract-1}}
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-\hypertarget{model-diagnostics}{%
-\subsection{Model Diagnostics}\label{model-diagnostics}}
-
-{[}ET: suggestion: A model can be fitted to data with no guarantee of a meaningful interpretation.
-Model diagnostics play an important role in assessing the appropriateness of the model.
-The assessment can involve examining the goodness of fit, checking if there are potential violations of model assumptions and validating models with external information.{]}
-
-Model diagnostics is the part of data analysis, preceded by the fit of a model, whose primary objectives are to examine the goodness of fit and reveal potential violations of model assumptions.
-
-{[}ET: model fit is more than just goodness of fit and checking model assumptions.
-It's too reductive to say this because you could have a model with good fit (according to some metric) and no model violation but if the way that the data were collected was flawed then scientific interpretation of the model is rendered useless.
-Diagnostics involve also the domain knowledge checks as well.{]}
-
-In these diagnostics, though numeric summaries are mostly available and some are even endorsed by finite or asymptotic properties, graphical representation of data is still preferred, or at least needed, due to its intuitiveness and the possibility to provide unexpected discoveries which may be abstract and
-unquantifiable.
-
-{[}ET: Generally, your sentences are too long.
-Break it up into smaller sentences.
-The generally idea is to make sentences that \emph{continue an idea from the preceding sentence} so it leads from one to the next naturally.
-E.g.
-Today is cloudy.
-Cloudy days are rare.
-Perhaps today will be different to other days!{]}
-
-However, unlike confirmatory data analysis built upon rigorous statistical procedures, e.g., hypothesis testing, visual diagnostics relies on graphical perception - human's ability to interpret and decode the information embedded in the graph \autocite{cleveland_graphical_1984}, which is to some extent subjective.
-Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
-One such example is finding a separation between gene groups in a two-dimensional projection from a linear discriminant analysis where there is no difference in the expression levels between the gene groups \autocite{roy_chowdhury_using_2015}.
-
-{[}ET: Confirmatory data analysis is well characterised by a rigorous procedure to discern particular hypothesis.
-The most widespread form of confirmatory data analysis are the use of \(p\)-values in testing hypothesis in a frequentist framework.
-More specifically, hypothesis testing in a frequentist framework involve stating a well-defined hypotheses; summarising the evidence as a test statistic; and calculating the probability of observing a statistic as extreme as the observed test statistic under the null hypothesis.
-This rigour, however, is often not present when inferences are drawn from a plot.
-
-Something about model diagnostics with a plot being the common thing to do..{]}
-
-\hypertarget{visual-inference-1}{%
-\subsection{Visual Inference}\label{visual-inference-1}}
-
-Visual inference was first introduced by \textcite{buja_statistical_2009} as an inferential framework to extend confirmatory statistics to visual discoveries.
-This framework redefines the test statistics, tests, null distribution, significance levels and \(p\)-value for visual discovery modelled on the confirmatory statistical testing.
-Figure \ref{fig:fparallelism} outlines the parallelism between conventional tests and visual discovery.
-
-{[}ET: note use Rmd syntax for figures references!
-So it's easy to use it for HTML later if you want to do something like \href{https://github.com/earowang/thesis}{Earo} and have PDF + HTML thesis.
-{]}
-
-\begin{figure}
-\centering
-\includegraphics[width=4.6875in,height=3.55208in]{figures/rsta2009012001.jpg}
-\caption{Parallelism between multiple quantitative testing and visual discovery \autocite{buja_statistical_2009}.
-Visible features in a plot are viewed as a collection of test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\), and any visual discoveries are treated as evidence against the null hypothesis.
-\label{fig:parallelism}}
-\end{figure}
-
-Parallelism between multiple quantitative testing and visual discovery \autocite{buja_statistical_2009}.
-Visible features in a plot are viewed as a collection of test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\), and any visual discoveries are treated as evidence against the null hypothesis.
-Parallelism between multiple quantitative testing and visual discovery \autocite{buja_statistical_2009}.
-
-In visual inference, a visual discovery is defined as a rejection of a null hypothesis, and the same null hypothesis can be rejected by many different visual discoveries \autocite{buja_statistical_2009}.
-For model diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the hypothesis.
-The same assumed model, such as classical linear regression model, can be rejected by both nonlinearity and heteroskedasticity with the residual plot as shown in Figure \ref{fig:nonlinearityheter}.
-
-\hypertarget{pre-specification-of-visual-discoverable-features-1}{%
-\subsection{Pre-specification of Visual Discoverable Features}\label{pre-specification-of-visual-discoverable-features-1}}
-
-As discussed in \textcite{buja_statistical_2009}, in the practice of model diagnostics, the range of possible visual discoveries is not pre-specified.
-In other words, people do not explicitly specify which one or more visual features they are looking for before the read of the diagnostic plot.
-This is concerning since conventional hypothesis testing always requires the pre-specification of the parameter space \(\Theta\) of the parameter of interest \(\theta \in \Theta\) to form a valid inferential procedure.
-To address this issue, a collection of test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) is defined, where \(\boldsymbol{\mathrm{y}}\) is the data and \(\boldsymbol{I}\) is a set of all possible visual features.
-\textcite{buja_statistical_2009} described each of the test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})\) as a measurement of the degree of presence of a visual feature.
-Alternatively, \textcite{majumder_validation_2013} avoids the use of visual features and defined the visual statistics \(T(.)\) as a mapping from a dataset to a data plot.
-Both definitions of visual test statistics are valid, but in the rest of the paper, the first definition will be used as it covered some details needed by this work.
-
-The size of the collection \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) depends on the size of the set \(I\).
-Thus, if one can define \(I\) comprehensively, i.e, pre-specify all the visual discoverable features, the validity issue will be solved.
-Unfortunately, to our knowledge, there is no such a way to list all visual features.
-In linear regression diagnostics, possible visual features of a residual plot may be outliers, shapes and clusters.
-But this is an incomplete list which does not enumerate all the visual features.
-
-Similarly, \textcite{wilkinson_graph-theoretic_2005} proposed the work called graph theoretic scagnostics, which adopted the idea of ``scagnostics'' - scatter plot diagnostics from (can't find the 1984 citation).
-It includes 9 computable scagnostics measures defined on planar proximity graphs: ``Outlying'', ``Convex'', ``Skinny'', ``Stringy'', ``Straight'', ``Monotonic'', ``Skewed'', ``Clumpy'' and ``Striated'' which attempts to describe outliers, shape, density, trend and coherence of the data.
-This approach is inspiring but it still does not give the complete list of visual discoverable features.
-In fact, it is possible that such a list will never be complete as suggested in \textcite{buja_statistical_2009}.
-
-Thinking out of the box, \textcite{buja_statistical_2009} argued that there is actually no need for pre-specification of visual discoverable features.
-In model diagnostics, when the null hypothesis is rejected, the reasons for rejecting the hypothesis will also be known.
-This is because observers can not only point out the fact that visual discoveries have been found, but also describe the particular visual features they observed.
-Those features will correspond to the subset of the collection of visual test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) which resulted in rejection.
-This argument helps justifies the validity of visual inference.
-
-\hypertarget{lineup-protocol-1}{%
-\subsection{Lineup Protocol}\label{lineup-protocol-1}}
-
-With the validity of visual inference being justified, another aspect of hypothesis testing that needs to be addressed is the control of false positive rate or Type I error.
-Any visual statistic \(T^{(i)}(\boldsymbol{\mathrm{y}})\) needs to pair with a critical value \(c^{(i)}\) to form a hypothesis test.
-When a visual feature \(i\) is discovered by the observer from a plot, the corresponding visual statistic \(T^{(i)}(\boldsymbol{\mathrm{y}})\) may not be known as there is no general agreement on the measurement of the degree of presence of a visual feature.
-It is only the event that \(T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}\) is confirmed.
-Similarly, if any visual discovery is found by the observer, we say, there exists \(i \in I:~T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}\) \autocite{buja_statistical_2009}.
-
-Using the above definition, the family-wise Type I error can be controlled if one can provide the collection of critical values \(c^{(i)}~(i \in I)\) such that \(P(\mathrm{there~exists~} i \in I: T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}|\boldsymbol{\mathrm{y}}) \leq \alpha\), where \(\alpha\) is the significance level.
-However, since the quantity of \(T^{(i)}(\boldsymbol{\mathrm{y}})\) may not be known, such collection of critical values can not be provided.
-
-\textcite{buja_statistical_2009} proposed the lineup protocol as a visual test to calibrate the Type I error issue without the specification of \(c^{(i)}~(i \in I)\).
-It is inspired by the ``police lineup'' or ``identity parade'' which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people.
-The protocol consists of \(m\) randomly placed data plots, where \(1\) plot is the actual data plot, and \(m-1\) null plots are produced by plotting data simulate from the null distribution which is consistent with the null hypothesis.
-Then, an observer who have not seen the actual data plot will be asked to point out the most different plot from the lineup.
-
-Under the null hypothesis, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is \(1/m\) due to randomness.
-If we reject the null hypothesis as the observer correctly picks the actual data plot, then the Type I error of this test is \(1/m\).
-
-This provides us with an mechanism to control the Type I error, because \(m\) - the number of plots in a lineup can be chosen.
-Further, if we involve \(K\) independent observers in a visual test, and let \(X\) be a random variable denoting the number of observers correctly picking the actual data plot.
-Then, under the null hypothesis \(X \sim \mathrm{Binom}_{K,1/m}\), and therefore, the \(p\)-value of a lineup of size \(m\) evaluated by \(K\) observer is given as
-
-\[P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},\]
-where \(x\) is the realization of number of observers correctly picking the actual data plot \autocite{majumder_validation_2013}.
-
-\hypertarget{visual-inference-applied-to-linear-regression}{%
-\subsection{Visual Inference Applied to Linear Regression}\label{visual-inference-applied-to-linear-regression}}
-
-How people used visual inference in linear regression?
-
-\hypertarget{limitation-of-the-visual-inference}{%
-\subsection{Limitation of the Visual Inference}\label{limitation-of-the-visual-inference}}
-
-What are the limitations?
+The estimated power was further used in fitting a quasi-binomial generalized linear model with natural logarithm of effect size as the only regressor.
+The result of the fitted models are given in Table \ref{tab:powerglmcubic}, Table \ref{tab:powerglmheter}, and Figure \ref{fig:powerglm}.
+
+As expected, coefficients of natural logarithm of effect size of all five models were positive.
+From Figure \ref{fig:powerglm}, it can be observed that the fitted power of visual test increased as the number of evaluations increased for both cubic and heteroskedasticity model.
+
+For heteroskedasticity model, this phenomenon was more obvious as the curves of visual tests with evaluations greater than two were always above the curves of visual test with evaluations smaller than two.
+However, this only held for large enough effect size.
+For small effect size, visual tests with fewer evaluations might have greater power.
+Note that, the expected power of visual test derived by \textcite{majumder_validation_2013} followed by the assumption that human has the ability to select the plot with the highest t-statistic from a lineup under the classical linear model regression setting showed similar properties, where visual tests with fewer evaluations were expected to perform better when the parameter values close to the null hypothesis.
+
+For cubic model, the separation between curves was small.
+Fitted power of visual test with three to five evaluations were almost identical to each other in regards of effect size.
+In addition, all five curves peaked at one as effect size increased, suggesting that identification of nonlinearity as a visual task can be completed reliably by human when the effect size is large enough.
+
+As shown in Figure \ref{fig:powerglm}, both F-test and Breusch--Pagan test generally possessed greater power than visual test.
+Since the settings we were using were not ideal scenarios for visual inference, it was not expected that the visual test would be as good as the conventional test.
+Our goal was to quantify the ability of human reading residual plot, such that the comparison between human and computer vision model can be made.
+
+\begin{table}[!htbp] \centering 
+  \caption{Quasi-binomial logistic regressions with estimated power as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for cubic model.
+Five regressions were fitted for different number of evaluations.}
+  \label{tab:powerglmcubic} 
+\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{5}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-6} 
+\\[-1.8ex] & \multicolumn{5}{c}{hat\_power} \\ 
+\\[-1.8ex] & (1) & (2) & (3) & (4) & (5)\\ 
+\hline \\[-1.8ex] 
+ log(effect\_size) & 0.929$^{***}$ & 1.227$^{***}$ & 1.084$^{***}$ & 1.057$^{***}$ & 1.035$^{***}$ \\ 
+  & (0.049) & (0.066) & (0.074) & (0.082) & (0.087) \\ 
+  & & & & & \\ 
+ Constant & $-$1.969$^{***}$ & $-$2.603$^{***}$ & $-$1.570$^{***}$ & $-$1.501$^{***}$ & $-$1.432$^{***}$ \\ 
+  & (0.114) & (0.153) & (0.142) & (0.156) & (0.160) \\ 
+  & & & & & \\ 
+\hline \\[-1.8ex] 
+Observations & 630 & 532 & 456 & 423 & 371 \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{5}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+\end{tabular} 
+\end{table}
+
+\begin{table}[!htbp] \centering 
+  \caption{Quasi-binomial logistic regressions with estimated power as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for heteroskedasticity model.
+Five regression were fitted for different number of evaluations.}
+  \label{tab:powerglmheter} 
+\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{5}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-6} 
+\\[-1.8ex] & \multicolumn{5}{c}{hat\_power} \\ 
+\\[-1.8ex] & (1) & (2) & (3) & (4) & (5)\\ 
+\hline \\[-1.8ex] 
+ log(effect\_size) & 0.821$^{***}$ & 1.025$^{***}$ & 0.885$^{***}$ & 0.920$^{***}$ & 1.063$^{***}$ \\ 
+  & (0.064) & (0.087) & (0.092) & (0.102) & (0.125) \\ 
+  & & & & & \\ 
+ Constant & $-$4.472$^{***}$ & $-$5.610$^{***}$ & $-$4.285$^{***}$ & $-$4.127$^{***}$ & $-$4.717$^{***}$ \\ 
+  & (0.320) & (0.449) & (0.461) & (0.499) & (0.607) \\ 
+  & & & & & \\ 
+\hline \\[-1.8ex] 
+Observations & 630 & 527 & 473 & 443 & 375 \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{5}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+\end{tabular} 
+\end{table}
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/powerglm-1.pdf}
+\caption{\label{fig:powerglm}Quasi-binomial logistic regressions with estimated power as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for both cubic and heteroskedasticity model.
+Five regression were fitted for different number of evaluations shown in five different colours.
+Estimated power are drawn in black.
+The dashed curvses are the power of F-test and Breusch--Pagan test respectively.}
+\end{figure}
+
+Another method we used to model the power was to collect all decisions made by \(K\choose{K-j}\) different outcomes for \(V_{K-j}\), and directly fit a binomial generalized linear model with natural logarithm of effect size as the regressor.
+This method skipped the calculation of the estimated power \(R/{K\choose{K-j}}\), and regressed on the binary outcome with larger sample size.
+Results of the fitted model is given in Figure \ref{fig:powerglmsecond}, Table \ref{tab:powerglmhetersecond} and Table \ref{tab:powerglmcubicsecond}.
+According to Figure \ref{fig:powerglmsecond}, this method gave clearer separation between curves for heteroskedasticity model.
+Noticeably, fitted power of visual test with five evaluations matched the power of conventional test at large effect size.
+
+\begin{table}[!htbp] \centering 
+  \caption{Logistic regressions with rejection as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for cubic model.
+Five regression were fitted for different number of evaluations.}
+  \label{tab:powerglmcubicsecond} 
+\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{5}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-6} 
+\\[-1.8ex] & \multicolumn{5}{c}{hat\_power} \\ 
+\\[-1.8ex] & (1) & (2) & (3) & (4) & (5)\\ 
+\hline \\[-1.8ex] 
+ log(effect\_size) & 0.872$^{***}$ & 1.292$^{***}$ & 1.125$^{***}$ & 1.164$^{***}$ & 1.315$^{***}$ \\ 
+  & (0.031) & (0.032) & (0.021) & (0.021) & (0.029) \\ 
+  & & & & & \\ 
+ Constant & $-$1.836$^{***}$ & $-$2.529$^{***}$ & $-$1.516$^{***}$ & $-$1.321$^{***}$ & $-$1.419$^{***}$ \\ 
+  & (0.077) & (0.073) & (0.046) & (0.049) & (0.072) \\ 
+  & & & & & \\ 
+\hline \\[-1.8ex] 
+Observations & 2,900 & 7,250 & 12,596 & 17,011 & 19,229 \\ 
+Log Likelihood & $-$752.896 & $-$1,312.436 & $-$2,977.445 & $-$2,651.299 & $-$1,463.555 \\ 
+Akaike Inf. Crit. & 1,509.792 & 2,628.873 & 5,958.889 & 5,306.598 & 2,931.109 \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{5}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+\end{tabular} 
+\end{table}
+
+\begin{table}[!htbp] \centering 
+  \caption{Logistic regressions with rejection as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for heteroskedasticity model.
+Five regression were fitted for different number of evaluations.}
+  \label{tab:powerglmhetersecond} 
+\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{5}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-6} 
+\\[-1.8ex] & \multicolumn{5}{c}{hat\_power} \\ 
+\\[-1.8ex] & (1) & (2) & (3) & (4) & (5)\\ 
+\hline \\[-1.8ex] 
+ log(effect\_size) & 0.912$^{***}$ & 1.221$^{***}$ & 1.294$^{***}$ & 1.525$^{***}$ & 1.798$^{***}$ \\ 
+  & (0.044) & (0.033) & (0.027) & (0.030) & (0.042) \\ 
+  & & & & & \\ 
+ Constant & $-$4.846$^{***}$ & $-$6.396$^{***}$ & $-$6.033$^{***}$ & $-$6.771$^{***}$ & $-$7.784$^{***}$ \\ 
+  & (0.229) & (0.175) & (0.141) & (0.159) & (0.230) \\ 
+  & & & & & \\ 
+\hline \\[-1.8ex] 
+Observations & 2,906 & 7,064 & 11,436 & 13,728 & 13,236 \\ 
+Log Likelihood & $-$1,525.398 & $-$3,668.325 & $-$5,491.915 & $-$4,828.896 & $-$2,858.774 \\ 
+Akaike Inf. Crit. & 3,054.796 & 7,340.650 & 10,987.830 & 9,661.793 & 5,721.548 \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{5}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+\end{tabular} 
+\end{table}
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/powerglmsecond-1.pdf}
+\caption{\label{fig:powerglmsecond}Logistic regressions with rejection as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for both cubic and heteroskedasticity model.
+Five regression were fitted for different number of evaluations shown in five different colours.
+Estimated power are drawn in black.
+The dashed curvses are the power of F-test and Breusch--Pagan test respectively.}
+\end{figure}
+
+\hypertarget{effect-of-factors-on-power-of-the-visual-test}{%
+\subsection{Effect of Factors on Power of the Visual Test}\label{effect-of-factors-on-power-of-the-visual-test}}
+
+\begin{table}[!htbp] \centering 
+  \caption{Logistic regressions with rejection as response variable and parameter settings as regressors fitted on the data collected from experiment I and II for both cubic and heteroskedasticity models.}
+  \label{tab:rejectfactor} 
+\begin{tabular}{@{\extracolsep{5pt}}lcc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{2}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-3} 
+\\[-1.8ex] & \multicolumn{2}{c}{reject} \\ 
+\\[-1.8ex] & (1) & (2)\\ 
+\hline \\[-1.8ex] 
+ a & $-$0.173$^{***}$ & 0.113$^{***}$ \\ 
+  & (0.014) & (0.022) \\ 
+  & & \\ 
+ b & $-$0.175$^{***}$ & 0.049$^{***}$ \\ 
+  & (0.014) & (0.002) \\ 
+  & & \\ 
+ c & $-$0.626$^{***}$ &  \\ 
+  & (0.033) &  \\ 
+  & & \\ 
+ sigma & $-$2.937$^{***}$ &  \\ 
+  & (0.037) &  \\ 
+  & & \\ 
+ Distribution of X: Normal & $-$0.870$^{***}$ & $-$2.781$^{***}$ \\ 
+  & (0.057) & (0.057) \\ 
+  & & \\ 
+ Distribution of X: Lognormal & $-$0.548$^{***}$ & $-$3.749$^{***}$ \\ 
+  & (0.072) & (0.047) \\ 
+  & & \\ 
+ Distribution of X: Negative Lognormal & $-$1.072$^{***}$ & $-$3.873$^{***}$ \\ 
+  & (0.051) & (0.058) \\ 
+  & & \\ 
+ Distribution of Z: Discrete & $-$3.219$^{***}$ &  \\ 
+  & (0.052) &  \\ 
+  & & \\ 
+ n & 0.002$^{***}$ & 0.001$^{***}$ \\ 
+  & (0.0002) & (0.0002) \\ 
+  & & \\ 
+ Number of evaluations & 0.216$^{***}$ & 0.592$^{***}$ \\ 
+  & (0.016) & (0.015) \\ 
+  & & \\ 
+ Constant & 5.151$^{***}$ & $-$0.028 \\ 
+  & (0.093) & (0.075) \\ 
+  & & \\ 
+\hline \\[-1.8ex] 
+Observations & 58,986 & 45,200 \\ 
+Log Likelihood & $-$9,892.382 & $-$11,324.830 \\ 
+Akaike Inf. Crit. & 19,806.760 & 22,665.660 \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+\end{tabular} 
+\end{table}
 
 \hypertarget{computer-vision-model}{%
-\subsection{Computer Vision Model}\label{computer-vision-model}}
-
-What is computer vision model?
-
-\hypertarget{contribution}{%
-\subsection{Contribution}\label{contribution}}
-
-What has been done by this paper?
-
-\hypertarget{structure-of-this-paper}{%
-\subsection{Structure of This Paper}\label{structure-of-this-paper}}
-
-What is the structure of the paper?
-
-Model diagnostics is the part of data analysis whose primary objectives are to examine the goodness of the model fit and reveal potential violations of the assumptions. Graphical approaches
-
-For regression diagnostics, it may includes the needs of
-
-Linear regression is an modelling approach to describe the relationship between an response variable and one or more explanatory variable.
-It has been widely used for both generative modeling and predictive modelling.
-
-Regression diagnostics is needed
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
+\chapter{Computer Vision Model}\label{computer-vision-model}}
+
+Visual test can be used to make valid inference on visual discoveries and acknowledging people with confidence about the findings.
+However, the lack of scalability of this method due to relatively high human cost compared to conventional test limits its application in wider fields.
+To relieve human analysts of the job of diagnosing predictive models in the context of visual inference, AI agents which can read data plot, evaluate lineups, and provide decisions need to be built.
+
+Considering there are different approches of AI, it is necessary to clarify the type of AI we aimed to build.
+Understanding the human thought process or mimicking the human vision mechanism were not of interest in this research.
+We believed humanly behaviours was enough for demonstrating AI for the problem we concerned.
+Hence, adopting one of the capacities required by the Turing test, namely machine learning, along with
+
+Thought process, human visual
+
+\hypertarget{discussion-of-potential-methods}{%
+\section{Discussion of potential methods}\label{discussion-of-potential-methods}}
+
+\hypertarget{different-approaches-of-ai}{%
+\subsection{different approaches of AI}\label{different-approaches-of-ai}}
+
+\begin{itemize}
 \tightlist
 \item
-  to check whether the assumptions has been violated
-\item
-  to check whether the line fit the data
-\end{enumerate}
-
-Model diagnostics for linear regression is well developed
+  Not aim for understanding the thought process
+\item
+  Not aim for mimicking the human vision mechanism
+\item
+  May be able to define distance metrics to measure difference between data plots for making mathematically prefect decisions
+\item
+  May be able to use computer vision model to approximate how people evaluate lineups
+\end{itemize}
+
+computer vision model:
+
+\begin{itemize}
+\tightlist
+\item
+  Use human data to train model with human selection as target -\textgreater{} mimic the human behaviour
+\item
+  Use simulated lineup to train model with actual data plot as target -\textgreater{} assume the actual data plot is the most different one
+\item
+  Use simulated data plot to train model with null or not null as target -\textgreater{} limited to the alternatives given to the model, and Type I error can not be controlled
+\item
+  Use simulated data plot to train model, then let the model evaluate each plot of a lineup -\textgreater{} leads to multiple selections in a lineup, and the model does not compare different plots in a lineup
+
+  \begin{itemize}
+  \tightlist
+  \item
+    Use a very large \(m\) to ensure the null hypothesis can be rejected by a classifier slightly better than a random selector. The problem becomes how large the \(m\) should be
+  \item
+    Use multiple models as multiple individuals to evaluate lineup -\textgreater{} Which models should be used? How many models should be used?
+  \end{itemize}
+\item
+  Let the model select multiple plots from a simulated lineup while comparing different plots in a lineup -\textgreater{} how to build such a system?
+
+  \begin{itemize}
+  \tightlist
+  \item
+    The input is \(m\) plots, what is the output? The probability of being the most different one?
+  \item
+    Then how many plots should be selected? Top 5? Or cumulative probability greater than a threshold?
+  \item
+    How to select the threshold?
+  \end{itemize}
+\end{itemize}
+
+The primary drawbacks of visual inference are
+
+infeasible in a large scale
+2. unfriendly to vision-impaired people 3. high finical cost and human cost
+4. time consuming
+
+Visual inference is a valid useful in diagnostics of predictive models
+
+To relieve human analysts of the job of diagnosing predictive models in the context of visual inference, computer vision
+
+Hence, to relieve human analysts of the job of diagnosing, automatically improve predictive models and assist vision-impaired people, this research aims to develop computer vision model to automate the reading
+1
+of data plots with application to predictive model diagnostics.
+The success of this research will enrich the applications of artificial intelligence in data visualization and improve the debugging process of artificial intelligence models.
+
+\hypertarget{model-architecture}{%
+\section{Model Architecture}\label{model-architecture}}
+
+Five computer vision models were trained to predict the probability of the given result plot being a null residual plot.
+
+\hypertarget{performance}{%
+\section{Performance}\label{performance}}
 

### 2022-03-06 site/02-literature-vi.Rmd
update 2

--- 2021-06-20 site/02-literature.Rmd
@@
-# Literature
+# Literature - Visual Inference
 
@@
 
-[comment]: <> (What is the probability of the true plot be the most different one?)
-
-[comment]: <> (What is the probability of the first plot be the most different one? == What is the probability of the second plot be the most different one?)
-
 ### Examples
@@
 An R package: nullabor
+
+## Statistical inference for exploratory data analysis and model diagnostics
+
+BibTex: 
+```
+@article{buja2009statistical,
+  title={Statistical inference for exploratory data analysis and model diagnostics},
+  author={Buja, Andreas and Cook, Dianne and Hofmann, Heike and Lawrence, Michael and Lee, Eun-Kyung and Swayne, Deborah F and Wickham, Hadley},
+  journal={Philosophical Transactions of the Royal Society A: Mathematical, Physical and Engineering Sciences},
+  volume={367},
+  number={1906},
+  pages={4361--4383},
+  year={2009},
+  publisher={The Royal Society Publishing}
+}
+```
+
+### Keywords
+
+permutation tests, rotation tests, statistical graphics, visual data mining, simulation, cognitive perception.
+
+# Introduction

### 2022-03-06 site/_book/site.tex
update 2

--- 2021-06-20 site/_book/site.tex
@@
 \author{Patrick Li}
-\date{2021-06-20}
+\date{2022-01-14}
 
@@
 
-\hypertarget{literature}{%
-\chapter{Literature}\label{literature}}
+\hypertarget{literature---visual-inference}{%
+\chapter{Literature - Visual Inference}\label{literature---visual-inference}}
 
@@
 
-\hypertarget{meetings}{%
-\chapter{Meetings}\label{meetings}}
+\hypertarget{statistical-inference-for-exploratory-data-analysis-and-model-diagnostics}{%
+\section{Statistical inference for exploratory data analysis and model diagnostics}\label{statistical-inference-for-exploratory-data-analysis-and-model-diagnostics}}
 
-\hypertarget{march-10-2021---week-2}{%
-\section{March 10, 2021 - Week 2}\label{march-10-2021---week-2}}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Human subject experiment and Monash permission
-
-  \begin{itemize}
-  \tightlist
-  \item
-    There is a workshop in May
-  \end{itemize}
-\item
-  Use simulated data to set up an experiment
-\item
-  Check out Gallery of graphs (a name yan \ldots{} not sure)
-\item
-  The experiment could start from residual plot
-\item
-  Q-Q plot could also be considered
-\item
-  A revelant research - residual calculation by kaiwen - master project
-\item
-  Use Appen survey to collect data
-\item
-  Build a Github to-do list, meeting record and summary of the literature
-\item
-  There may be some development in the theory by Nancy Reid - theoretical statistician (recent work)
-\item
-  Consider using Kears to build a computer vision model
-\end{enumerate}
-
-\hypertarget{march-17-2021---week-3}{%
-\section{March 17, 2021 - Week 3}\label{march-17-2021---week-3}}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Read Susan Vanderplas's personal website to find additional information
-\item
-  Check out NUMBAT residual plot comparision - summer-vis-inf : Aarathy Babu - code examples
-\item
-  Read human subject premission examples (sent by Di)
-\item
-  Check out top-up application
-\item
-  Consider to use \texttt{Edibble} to set up the experiment
-\item
-  Build the PhD repo
-\item
-  Consider to use non-shiny framework
-\end{enumerate}
-
-\hypertarget{march-21-2021---week-4}{%
-\section{March 21, 2021 - Week 4}\label{march-21-2021---week-4}}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  A short meeting - late for 30 minutes
-\item
-  Aarathy introduces her repo
-\end{enumerate}
-
-\hypertarget{march-28-2021---week-5}{%
-\section{March 28, 2021 - Week 5}\label{march-28-2021---week-5}}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Discuss the options of building a alternative hypothesis in residual plot
-\item
-  AR, heterogeneity of variance, endogeneity, skewness, exp distribution, poisson distribution and missing covariance
-\item
-  Choice of plot design (loess or \(y=0\) line), number of lineup (5\textasciitilde15), number of observations
-\item
-  Use flow chart to illustrate the choices
-\item
-  Literature review - check previous designs
-\item
-  Build bookdown to track records
-\end{enumerate}
-
-\hypertarget{todo}{%
-\chapter{TODO}\label{todo}}
-
-\hypertarget{week-3}{%
-\section{Week 3}\label{week-3}}
-
-\begin{itemize}
-\tightlist
-\item[$\boxtimes$]
-  Check human subject experiment and Monash permission materials
-
-  \begin{itemize}
-  \tightlist
-  \item
-    \url{https://www.intranet.monash/researchadmin/start/ethics/human}
-  \end{itemize}
-\item[$\boxtimes$]
-  Check Kaiwen's project \& paper
-\item[$\square$]
-  Check Gallery of graphs - unclear author
-\end{itemize}
-
-\hypertarget{week-4}{%
-\section{Week 4}\label{week-4}}
-
-\begin{itemize}
-\tightlist
-\item[$\boxtimes$]
-  Build prototype of html webpage to collect data
-\item[$\boxtimes$]
-  Send data to google sheet
-\item[$\boxtimes$]
-  Check summer-vis-inf
-\item[$\boxtimes$]
-  Read examples sent by Di
-\item[$\square$]
-  Build PhD repo
-
-  \begin{itemize}
-  \tightlist
-  \item[$\boxtimes$]
-    meetings
-  \item[$\square$]
-    paper
-  \item[$\boxtimes$]
-    TODO
-  \end{itemize}
-\item[$\square$]
-  Check Susan Vanderplas's website
-
-  \begin{itemize}
-  \tightlist
-  \item[$\square$]
-    paper
-  \item[$\square$]
-    talks
-  \item[$\square$]
-    posts
-  \end{itemize}
-\end{itemize}
-
-\hypertarget{week-5}{%
-\section{Week 5}\label{week-5}}
-
-\begin{itemize}
-\tightlist
-\item[$\boxtimes$]
-  modify the webpage to be able to select multiple plots
-\item[$\boxtimes$]
-  Attempt to generate data from one assumed model
-\item[$\square$]
-  draft Human Ethics Application Form
-\end{itemize}
-
-\hypertarget{week-6}{%
-\section{Week 6}\label{week-6}}
-
-\begin{itemize}
-\tightlist
-\item[$\square$]
-  Do the literature review of previous design
-\item[$\square$]
-  Draw a flow chart to illustrate the design
-\end{itemize}
+\hypertarget{literature---computer-vision}{%
+\chapter{Literature - Computer vision}\label{literature---computer-vision}}
 

### 2022-03-06 Thesis/01-chap1.Rmd
update 2 and 1.8

@@
 
-(reasons for conducting experiments)
+Therefore, we first tried to automate the visual diagnostics of one of the simplest statistical models - classical normal linear regression model.
+Two common types of model defects were created and visualized using residuals vs.
+fitted values plot.
+To set up a comparison between the computer vision models and humans, human subject experiments were conducted to understand the ability of human reading residual plots.
 
@@
 
+The previous section focuses on the change of effect size relative to the power of the visual test.
+Individual factor embedded in the data simulation also needs be analysed.
+Two logistic regressions were fitted with decisions made by $V_{K-j}$ as response variable.
+For the cubic model, regressors were $|a|$, $|b|$, $|c-1|$, $\sigma$, distribution of $\boldsymbol{X}$, distribution of $\boldsymbol{Z}$, number of observations and number of evaluations.
+For the heteroskedasticity model, regressors were $|a|$, $b$, distribution of $\boldsymbol{X}$, number of observations and number of evaluations.
+
+According to Table \@ref(tab:rejectfactorcubic), for cubic model, both coefficients of $|a|$ and $|b|$ were positive, which suggested increasing the coefficients of quadric and cubic terms would result in higher visual power.
+The coefficient of $|c-1|$ was also positive indicating joint effect of multiple regressors on the residual plot decreased the visual power.
+As expected, the coefficient of $\sigma$ was negative as it controlled the noise level.
+Lineup simulated with uniform distribution of $\boldsymbol{X}$ had the highest power, followed by negative lognormal, normal, and lognormal.
+Discreteness in $\boldsymbol{Z}$ decreased the visual power.
+The coefficient of number of observations was positive, which means larger sample size would lead to higher visual power.
+
+Results of logistic regression for heteroskedasticity is given in Table \@ref(tab:rejectfactorheter).
+For heteroskedasticity model, the coefficient of $|a|$ was negative, which suggested "butter fly" shape was easier to identify than "triangle" shape.
+As expected, the coefficient of $b$ was positive.
+For different distributions of $\boldsymbol{X}$, the one with highest power was uniform distribution, followed by normal, discrete uniform, lognormal and negative lognormal.
+The coeffient of the number of observations was positive, but it was not significant.
+
+In overall, we did not have any unexpected findings except the visual power related to different distributions of $\boldsymbol{X}$ in cubic model.
+We expected lineup simulated with uniform distribution would have the highest power, followed by normal, lognormal and negative lognormal, since we observed similar results in pilot study.
+
 ```{r results="asis"}
@@
   mutate(n_eval = n_eval) %>%
-  mutate(x_dist = factor(x_dist, levels = c("uniform", "normal", "lognormal", "neglognormal"))) %>%
   unnest_longer(col = p_value) %>%
@@
 
-glm(reject ~ a + b + c + e_sigma + x_dist + z_discrete + n + n_eval, data = filter(single_merged, type == "cubic"), family = binomial()) -> mod1_factor
+glm(reject ~ abs(a) + abs(b) + abs(c - 1) + e_sigma + x_dist + z_discrete + n + n_eval, data = filter(single_merged, type == "cubic"), family = binomial()) -> mod1_factor
 
@@
 ```{r results="asis"}
-glm(reject ~ a + b + x_dist + z_discrete + n + n_eval, data = filter(single_merged, type == "heteroskedasticity"), family = binomial()) -> mod2_factor
-
-stargazer(mod1_factor, mod2_factor,header = FALSE,
-          label = "tab:rejectfactor",
-          title = "Logistic regressions with rejection as response variable and parameter settings as regressors fitted on the data collected from experiment I and II for both cubic and heteroskedasticity models.",
+glm(reject ~ abs(a) + b + x_dist + n + n_eval, data = filter(single_merged, type == "heteroskedasticity"), family = binomial()) -> mod2_factor
+
+stargazer(mod1_factor, header = FALSE,
+          label = "tab:rejectfactorcubic",
+          title = "Logistic regressions with rejection as response variable and parameter settings as regressors fitted on the data collected from experiment I and II for cubic model.",
           style = "default",
-          covariate.labels = c("a", "b", "c", 
+          covariate.labels = c("|a|", "|b|", "|c-1|", 
                                "sigma",
+                               "Distribution of X: Negative Lognormal",
                                "Distribution of X: Normal", 
+                               "Distribution of X: Uniform", 
+                               "Distribution of Z: Discrete",
+                               "n",
+                               "Number of evaluations",
+                               "Constant"))
+
+```
+
+```{r results="asis"}
+stargazer(mod2_factor, header = FALSE,
+          label = "tab:rejectfactorheter",
+          title = "Logistic regressions with rejection as response variable and parameter settings as regressors fitted on the data collected from experiment I and II for heteroskedasticity model.",
+          style = "default",
+          covariate.labels = c("|a|", "b",
                                "Distribution of X: Lognormal", 
                                "Distribution of X: Negative Lognormal",
-                               "Distribution of Z: Discrete",
+                               "Distribution of X: Normal", 
+                               "Distribution of X: Uniform", 
                                "n",

### 2022-03-06 Thesis/_book/thesis.tex
update 2 and 1.8

@@
 
-(reasons for conducting experiments)
+Therefore, we first tried to automate the visual diagnostics of one of the simplest statistical models - classical normal linear regression model.
+Two common types of model defects were created and visualized using residuals vs.~fitted values plot.
+To set up a comparison between the computer vision models and humans, human subject experiments were conducted to understand the ability of human reading residual plots.
 
@@
 
+The previous section focuses on the change of effect size relative to the power of the visual test.
+Individual factor embedded in the data simulation also needs be analysed.
+Two logistic regressions were fitted with decisions made by \(V_{K-j}\) as response variable.
+For the cubic model, regressors were \(|a|\), \(|b|\), \(|c-1|\), \(\sigma\), distribution of \(\boldsymbol{X}\), distribution of \(\boldsymbol{Z}\), number of observations and number of evaluations.
+For the heteroskedasticity model, regressors were \(|a|\), \(b\), distribution of \(\boldsymbol{X}\), number of observations and number of evaluations.
+
+According to Table \ref{tab:rejectfactorcubic}, for cubic model, both coefficients of \(|a|\) and \(|b|\) were positive, which suggested increasing the coefficients of quadric and cubic terms would result in higher visual power.
+The coefficient of \(|c-1|\) was also positive indicating joint effect of multiple regressors on the residual plot decreased the visual power.
+As expected, the coefficient of \(\sigma\) was negative as it controlled the noise level.
+Lineup simulated with uniform distribution of \(\boldsymbol{X}\) had the highest power, followed by negative lognormal, normal, and lognormal.
+Discreteness in \(\boldsymbol{Z}\) decreased the visual power.
+The coefficient of number of observations was positive, which means larger sample size would lead to higher visual power.
+
+Results of logistic regression for heteroskedasticity is given in Table \ref{tab:rejectfactorheter}.
+For heteroskedasticity model, the coefficient of \(|a|\) was negative, which suggested ``butter fly'' shape was easier to identify than ``triangle'' shape.
+As expected, the coefficient of \(b\) was positive.
+For different distributions of \(\boldsymbol{X}\), the one with highest power was uniform distribution, followed by normal, discrete uniform, lognormal and negative lognormal.
+The coeffient of the number of observations was positive, but it was not significant.
+
+In overall, we did not have any unexpected findings except the visual power related to different distributions of \(\boldsymbol{X}\) in cubic model.
+We expected lineup simulated with uniform distribution would have the highest power, followed by normal, lognormal and negative lognormal, since we observed similar results in pilot study.
+
 \begin{table}[!htbp] \centering 
-  \caption{Logistic regressions with rejection as response variable and parameter settings as regressors fitted on the data collected from experiment I and II for both cubic and heteroskedasticity models.}
-  \label{tab:rejectfactor} 
-\begin{tabular}{@{\extracolsep{5pt}}lcc} 
+  \caption{Logistic regressions with rejection as response variable and parameter settings as regressors fitted on the data collected from experiment I and II for cubic model.} 
+  \label{tab:rejectfactorcubic} 
+\begin{tabular}{@{\extracolsep{5pt}}lc} 
 \\[-1.8ex]\hline 
 \hline \\[-1.8ex] 
- & \multicolumn{2}{c}{\textit{Dependent variable:}} \\ 
-\cline{2-3} 
-\\[-1.8ex] & \multicolumn{2}{c}{reject} \\ 
-\\[-1.8ex] & (1) & (2)\\ 
-\hline \\[-1.8ex] 
- a & $-$0.173$^{***}$ & 0.113$^{***}$ \\ 
-  & (0.014) & (0.022) \\ 
-  & & \\ 
- b & $-$0.175$^{***}$ & 0.049$^{***}$ \\ 
-  & (0.014) & (0.002) \\ 
-  & & \\ 
- c & $-$0.626$^{***}$ &  \\ 
-  & (0.033) &  \\ 
-  & & \\ 
- sigma & $-$2.937$^{***}$ &  \\ 
-  & (0.037) &  \\ 
-  & & \\ 
- Distribution of X: Normal & $-$0.870$^{***}$ & $-$2.781$^{***}$ \\ 
-  & (0.057) & (0.057) \\ 
-  & & \\ 
- Distribution of X: Lognormal & $-$0.548$^{***}$ & $-$3.749$^{***}$ \\ 
-  & (0.072) & (0.047) \\ 
-  & & \\ 
- Distribution of X: Negative Lognormal & $-$1.072$^{***}$ & $-$3.873$^{***}$ \\ 
-  & (0.051) & (0.058) \\ 
-  & & \\ 
- Distribution of Z: Discrete & $-$3.219$^{***}$ &  \\ 
-  & (0.052) &  \\ 
-  & & \\ 
- n & 0.002$^{***}$ & 0.001$^{***}$ \\ 
-  & (0.0002) & (0.0002) \\ 
-  & & \\ 
- Number of evaluations & 0.216$^{***}$ & 0.592$^{***}$ \\ 
-  & (0.016) & (0.015) \\ 
-  & & \\ 
- Constant & 5.151$^{***}$ & $-$0.028 \\ 
-  & (0.093) & (0.075) \\ 
-  & & \\ 
-\hline \\[-1.8ex] 
-Observations & 58,986 & 45,200 \\ 
-Log Likelihood & $-$9,892.382 & $-$11,324.830 \\ 
-Akaike Inf. Crit. & 19,806.760 & 22,665.660 \\ 
+ & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-2} 
+\\[-1.8ex] & reject \\ 
+\hline \\[-1.8ex] 
+ |a| & 1.064$^{***}$ \\ 
+  & (0.031) \\ 
+  & \\ 
+ |b| & 0.718$^{***}$ \\ 
+  & (0.029) \\ 
+  & \\ 
+ |c-1| & 5.443$^{***}$ \\ 
+  & (0.119) \\ 
+  & \\ 
+ sigma & $-$3.536$^{***}$ \\ 
+  & (0.050) \\ 
+  & \\ 
+ Distribution of X: Negative Lognormal & 0.971$^{***}$ \\ 
+  & (0.073) \\ 
+  & \\ 
+ Distribution of X: Normal & 0.754$^{***}$ \\ 
+  & (0.086) \\ 
+  & \\ 
+ Distribution of X: Uniform & 1.751$^{***}$ \\ 
+  & (0.085) \\ 
+  & \\ 
+ Distribution of Z: Discrete & $-$2.986$^{***}$ \\ 
+  & (0.056) \\ 
+  & \\ 
+ n & 0.005$^{***}$ \\ 
+  & (0.0002) \\ 
+  & \\ 
+ Number of evaluations & 0.291$^{***}$ \\ 
+  & (0.017) \\ 
+  & \\ 
+ Constant & $-$2.453$^{***}$ \\ 
+  & (0.173) \\ 
+  & \\ 
+\hline \\[-1.8ex] 
+Observations & 58,986 \\ 
+Log Likelihood & $-$8,034.351 \\ 
+Akaike Inf. Crit. & 16,090.700 \\ 
 \hline 
 \hline \\[-1.8ex] 
-\textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+\textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+\end{tabular} 
+\end{table}
+
+\begin{table}[!htbp] \centering 
+  \caption{Logistic regressions with rejection as response variable and parameter settings as regressors fitted on the data collected from experiment I and II for heteroskedasticity model.} 
+  \label{tab:rejectfactorheter} 
+\begin{tabular}{@{\extracolsep{5pt}}lc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-2} 
+\\[-1.8ex] & reject \\ 
+\hline \\[-1.8ex] 
+ |a| & $-$0.526$^{***}$ \\ 
+  & (0.037) \\ 
+  & \\ 
+ b & 0.064$^{***}$ \\ 
+  & (0.002) \\ 
+  & \\ 
+ Distribution of X: Lognormal & $-$0.534$^{***}$ \\ 
+  & (0.058) \\ 
+  & \\ 
+ Distribution of X: Negative Lognormal & $-$0.610$^{***}$ \\ 
+  & (0.062) \\ 
+  & \\ 
+ Distribution of X: Normal & 0.187$^{***}$ \\ 
+  & (0.063) \\ 
+  & \\ 
+ Distribution of X: Uniform & 2.980$^{***}$ \\ 
+  & (0.064) \\ 
+  & \\ 
+ n & 0.0003 \\ 
+  & (0.0002) \\ 
+  & \\ 
+ Number of evaluations & 0.496$^{***}$ \\ 
+  & (0.014) \\ 
+  & \\ 
+ Constant & $-$2.711$^{***}$ \\ 
+  & (0.062) \\ 
+  & \\ 
+\hline \\[-1.8ex] 
+Observations & 48,370 \\ 
+Log Likelihood & $-$13,026.190 \\ 
+Akaike Inf. Crit. & 26,070.390 \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 \end{tabular} 

### 2022-03-06 Thesis/01-chap1.Rmd
add performance

@@
 
-## Artificial intelligence: Four Approaches
+## Artificial intelligence: Four Approaches {#se:ai}
 
@@
 
-Considering there are different approches of AI, it is necessary to clarify the type of AI we aimed to build.
+Considering there are different approches of AI discussed in \ref{se:ai}, it is necessary to clarify the type of AI we aimed to build.
 Understanding the human thought process or mimicking the human vision mechanism were not of interest in this research.
 We believed humanly behaviours was enough for demonstrating AI for the problem we concerned.
-Hence, adopting one of the capacities required by the Turing test, namely machine learning, along with
-
-Thought process, human visual
-
-## Discussion of potential methods
+Hence, two of the capacities required by the total Turing test, namely machine learning and computer vision will be adopted.
+
+Another possible method was to define distance metrics to measure difference between data plots for making mathematically prefect decisions.
+This method would work if such metrics can be obtained manually or approximated algorithmically.
+However, we considered it was too complex as a starting point.
+
+## Automatic visual inference system
+
+For an automatic visual inference system, there are several possibilities of the type of inputs.
+
+Depend on the inputs and outputs, computer vision model can be integrated into an automatic visual inference system.
 
@@
 ## Performance
+
+```{r}
+null_true_eval_dat <- data.frame()
+
+model_name <- c("VGG16", "VGG19", "Xception", "ResNet152V2", "InceptionV3")
+
+for (i in 1:5) {
+  mid_dat <- read_csv(here::here(glue::glue("data/cv/prediction/mod_{i}_null_both.csv")))
+  null_true_eval_dat <- bind_rows(null_true_eval_dat, mid_dat)
+  mid_dat <- read_csv(here::here(glue::glue("data/cv/prediction/mod_{i}_true_both.csv")))
+  null_true_eval_dat <- bind_rows(null_true_eval_dat, mid_dat)
+} 
+```
+
+```{r}
+null_true_eval_dat %>%
+  group_by(model, null) %>%
+  summarise(rate = mean(prediction == null)) %>%
+  pivot_wider(names_from = "null", values_from = "rate") %>%
+  ungroup() %>%
+  mutate(model = model_name) %>%
+  mutate(avg_rate = (`FALSE` + `TRUE`)/2) %>%
+  knitr::kable(col.names = c("Model", "Actual data plot correct rate", "Null plot correct rate", "Average correct rate"), 
+             booktabs = TRUE, 
+             caption  = 'Performance of computer vision models on null residual plot and acutal data plot.', 
+             label = "null-true-eval" ) 
+```
+
+```{r}
+lineup_eval_dat <- data.frame()
+
+for (i in 1:5) {
+  mid_dat <- read_csv(here::here(glue::glue("data/cv/prediction/mod_{i}_fixed_both.csv"))) %>%
+    mutate(model = i) 
+  lineup_eval_dat <- bind_rows(lineup_eval_dat, mid_dat)
+}
+
+eval_data_para <- readRDS(here::here("data/cv/eval_data_para.rds"))
+eval_data_para <- reduce(map(eval_data_para, as.data.frame), bind_rows)
+eval_original_data <- readRDS(here::here("data/cv/eval_original_data.rds"))
+```
+
+```{r}
+eval_original_data %>%
+  group_by(id, k) %>%
+  summarise(ans = first(null)) -> eval_lineup_ans
+```
+
+```{r}
+lineup_eval_dat %>%
+  left_join(eval_lineup_ans, by = c("set" = "id", "plot" = "k")) %>%
+  group_by(model, set) %>%
+  summarise(n_sel = sum(!predict), detect = sum(ans == FALSE & predict == FALSE)) -> lineup_eval_summary
+```
+
+```{r}
+cache_env <- new.env()
+lineup_eval_summary %>%
+  calc_p_value_multi(lineup_id = "set", detected = "detect", n_sel = "n_sel", cache_env = cache_env) -> cv_results
+```
+```{r}
+cv_results %>%
+  mutate(reject = eval_p_value(p_value, tol = 1e-2)) %>%
+  left_join(select(eval_original_data, id, effect_size) %>%
+              group_by(id) %>%
+              summarise(effect_size = first(effect_size)),
+            by = c("lineup_id" = "id")) -> cv_results
+
+cv_results %>%
+  mutate(type = ifelse(lineup_id <= 2500, "cubic", "heteroskedasticity")) -> cv_results
+```
+
+```{r}
+merged_survey %>%
+  mutate(exp_lineup_id = paste0(exp, "_", lineup_id)) %>%
+  select(exp_lineup_id, type, exp, effect_size) %>%
+  right_join(comb_pvalue) %>%
+  mutate(n_eval = factor(n_eval)) %>%
+  unnest_longer(col = p_value) %>%
+  mutate(reject = eval_p_value(p_value, tol = 1e-2)) %>%
+  mutate(reject = as.numeric(reject)) %>%
+  ggplot() +
+  geom_point(aes(log(effect_size), reject), alpha = 0.1, size = 1) +
+  geom_smooth(aes(log(effect_size), reject, group = "conventional"),
+              method = "glm", method.args = list(family = binomial), data = conv_result, 
+              se = FALSE, size = 1, linetype = 2, col = "grey80") +
+  geom_smooth(aes(log(effect_size), reject, group = n_eval, col = n_eval),
+              method = "glm", method.args = list(family = binomial), size = 0.5, se = FALSE) +
+  geom_smooth(aes(log(effect_size), as.numeric(reject), group = "cv"), data = cv_results,
+              se = FALSE, size = 1, method = "glm", method.args = list(family = binomial)) +
+  facet_wrap(~type, scales = "free_x") +
+  labs(col = "Number of evaluations") +
+  theme_light(base_size = 5) +
+  theme(legend.position = "bottom") +
+  scale_colour_viridis_d(option = "B", begin = 0.5, end = 0.9) +
+  guides(col = guide_legend(nrow = 1, byrow=TRUE)) +
+  xlab("Log of effect size") +
+  ylab("Rejected")
+```

### 2022-03-06 Thesis/_book/thesis.tex
add performance

@@
 
-\hypertarget{artificial-intelligence-four-approaches}{%
-\section{Artificial intelligence: Four Approaches}\label{artificial-intelligence-four-approaches}}
+\hypertarget{se:ai}{%
+\section{Artificial intelligence: Four Approaches}\label{se:ai}}
 
@@
 
-Considering there are different approches of AI, it is necessary to clarify the type of AI we aimed to build.
+Considering there are different approches of AI discussed in \ref{se:ai}, it is necessary to clarify the type of AI we aimed to build.
 Understanding the human thought process or mimicking the human vision mechanism were not of interest in this research.
 We believed humanly behaviours was enough for demonstrating AI for the problem we concerned.
-Hence, adopting one of the capacities required by the Turing test, namely machine learning, along with
-
-Thought process, human visual
-
-\hypertarget{discussion-of-potential-methods}{%
-\section{Discussion of potential methods}\label{discussion-of-potential-methods}}
+Hence, two of the capacities required by the total Turing test, namely machine learning and computer vision will be adopted.
+
+Another possible method was to define distance metrics to measure difference between data plots for making mathematically prefect decisions.
+This method would work if such metrics can be obtained manually or approximated algorithmically.
+However, we considered it was too complex as a starting point.
+
+\hypertarget{automatic-visual-inference-system}{%
+\section{Automatic visual inference system}\label{automatic-visual-inference-system}}
+
+For an automatic visual inference system, there are several possibilities of the type of inputs.
+
+Depend on the inputs and outputs, computer vision model can be integrated into an automatic visual inference system.
 
@@
 
+\begin{table}
+
+\caption{\label{tab:null-true-eval}Performance of computer vision models on null residual plot and acutal data plot.}
+\centering
+\begin{tabular}[t]{lrrr}
+\toprule
+Model & Actual data plot correct rate & Null plot correct rate & Average correct rate\\
+\midrule
+VGG16 & 0.7387 & 0.8850 & 0.81185\\
+VGG19 & 0.8029 & 0.8313 & 0.81710\\
+Xception & 0.7370 & 0.9213 & 0.82915\\
+ResNet152V2 & 0.6932 & 0.9233 & 0.80825\\
+InceptionV3 & 0.6691 & 0.9399 & 0.80450\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\includegraphics{thesis_files/figure-latex/unnamed-chunk-44-1.pdf}
+
 \appendix

### 2022-03-07 Milestone_1/01-chap1.Rmd
add milestone report

--- 2022-03-06 Thesis/01-chap1.Rmd
@@
 
-# Introduction {#ch:intro}
-
-## Is what we see really there?
-
-## Artificial intelligence: Four Approaches {#se:ai}
-
-**Artificial intelligence** (AI) is the field of research concerned with understanding and building machines who can demonstrate intelligence.
-As discussed in @russell_artificial_2002, historically, there are disagreements among researchers about the definition of intelligence, which is caused by two critical questions:
-
-1. Should AI act and think humanly or rationally? 
-2. Without the thought process and reasoning, are behaviours sufficient to demonstrate intelligence?
-
-Based on the answer to the above questions, four major approaches to pursue AI have been established.
-These approaches can be summarized into a two by two table as shown in Figure \@ref(fig:two-dim-ai), where the row is "Human" vs.
-"Rational", and the column is "Behaviour" vs.
-"Thought".
-Positioning at the top right cell, the **rational agent approach** aims to build agent that perform mathematically perfect acts such that the best expected outcome can always be achieved.
-In contrast, the **"laws of thought" approach** focus on understanding the logic behind the rationality.
-Closely related to **cognitive science**, the **cognitive modelling approach** attempts to express theories of human cognition as computer program to mimic the thought process of human.
-Lastly, the **Turing test approach** is built upon the famous **Turing test** proposed by @turing_computing_1950.
-The test can be roughly described as, whether a human can distinguish another human from a computer with written communications only.
-To pass the test, several capabilities of computer are required.
-This includes **natural language processing** for communication with human, **knowledge representation** for encoding knowledge, **automated reasoning** for derivation of conclusions and **machine learning** for improving AI automatically through experience and data.
-Some researchers argued that written communication is insufficient to demonstrate intelligence, and some degree of physical simulation of a person is still necessary.
-One such example is the **total Turing test** proposed by @harnad_other_1991.
-It adds three new requirements to the list, including **computer vision**, **speech recognition** and **robotics**, which are response for interactions with the physical world.
-Notably, all seven required capabilities have become major subfields of AI today.
-And their development has made AI one of the fastest-growing fields in the 21st century [@russell_artificial_2002].
-
-```{r two-dim-ai, fig.cap="Four possible approaches to pursue AI based on the two dimensions in AI research - human vs. rational and thought vs. behavior [@russell_artificial_2002]."}
-ggplot() +
-  geom_hline(aes(yintercept = 1), col = "grey") +
-  geom_vline(aes(xintercept = 1), col = "grey") +
-  geom_text(aes(x = 0.5 , y = 1.5, label = "The Turing test approach")) +
-  geom_text(aes(x = 1.5 , y = 1.5, label = "The rational agent approach")) +
-  geom_text(aes(x = 0.5 , y = 0.5, label = "The cognitive modelling approach")) +
-  geom_text(aes(x = 1.5 , y = 0.5, label = 'The "laws of thought" approach')) +
-  theme_light() +
-  xlab("") +
-  ylab("") +
-  scale_x_continuous(breaks = c(0.5, 1.5), labels = c("Human", "Rational"), limits = c(0, 2)) +
-  scale_y_continuous(breaks = c(0.5, 1.5), labels = c("Thought", "Behaviour"), limits = c(0, 2)) +
-  theme(panel.grid = element_blank(), axis.ticks = element_blank())
-```
-
-With the development of AI, mature AI technologies, such as facial recognition and web recommendation system, have profoundly affected the way modern society operates and citizen's daily life.
-This is largely as a consequence of the huge investment in AI industry by the financial market in recent years.
-Further, the increasingly cheap computing cost and the massive amount of accessible e-commerce data produced in the Internet age provide the possibility for applying data-intensive AI models, which enables AI performance to reach new heights in history [@jordan_machine_2015].
-Some AI systems have already been remarkably better than humans in certain areas, e.g., game playing.
-AlphaGo and AlphaZero developed by the Google DeepMind team surpass professional human Go players [@silver_general_2018].
+\clearpage\pagenumbering{arabic}\setcounter{page}{0}
+
+# Background and Motivation
+
+<!-- ## Motivation -->
+
+<!-- Have you doubted what you found from a data plot was really there?
+Or have you had reservations about visual discoveries because they might be illusions?
+Data plots are prevalent in every field of nature and social science as it is one of the most recognized data visualization techniques.
+Without data plots, trends and patterns embed in the data would not be easy to discover.
+However, data plots are often over- or under-interpreted by human readers which leads to unsecured findings.
+This is unsatisfactory as confirmatory research is the foundation of science.
+-->
+
+<!-- Visual inference, an inferential framework developed by @buja_statistical_2009, provides an option to test the visual findings, which helps solidify our confidence in using data plots.
+But one of the drawbacks of this framework is it requires human evaluation of data plots.
+Such requirement is unattainable when the number of data plots need for testing is large.
+Even if it can be fulfilled, the cost of labour would be exceedingly high.
+-->
+
+<!-- ## Research Problem -->
+
+<!-- The main objective of this research is to build an automatic visual inference system that can evaluate data plots.
+We aim to apply this automatic system on predictive model diagnostics because one of the scenarios best fitted for visual inference is where there does not exist a conventional test, and flexible predictive models which adopt machine learning algorithms do not generally have a conventional hypothesis testing procedure.
+-->
 
@@
 
-Behind the success of AI, a great propotion of AI systems rely on the predictive modelling framework.
-@donoho_50_2017 in its summary of data science stated that the concept of this modelling culture could be traced back to an article written by Breiman (2001).
+@donoho_50_2017 in its summary of data science stated that the concept of the predictive modelling culture could be traced back to an article written by Breiman (2001).
 In contrast to the generative modelling culture, which aims to develop stochastic models to make inferences about the data generating process, predictive modelling emphasizes the ability of the model to make accurate predictions.
-Most AI tasks are complex prediction problems where the data mechanism is mysterious, or at least, partly unknowable.
-@breiman_statistical_2001 suggests that generative models are obviously not applicable in these scenarios, while the predictive modelling seeks only an accurate approximated function $\boldsymbol{f}(\boldsymbol{x})$ to describe the relationship between the features $\boldsymbol{x}$ and the responses $\boldsymbol{y}$.
 
@@
 
-Therefore, we first tried to automate the visual diagnostics of one of the simplest statistical models - classical normal linear regression model.
-Two common types of model defects were created and visualized using residuals vs.
-fitted values plot.
+# Structure of the Thesis (need disscussion)
+
+An overview of the thesis, including some brief detail of the proposed content of all chapters.
+This should encompass a description of the theoretical and conceptual framework that underlies the thesis, and the procedures that are to be used in addressing the research questions
+
+1. Automatic Visual Statistical Inference, with Application to Linear Regression Diagnostics
+2. ...
+3. ...
+
+# Automatic Visual Statistical Inference, with Application to Linear Regression Diagnostics
+
+This project focused on building a prototype of automatic visual statistical inference system for evaluating residual plots of classical normal linear regression model.
 To set up a comparison between the computer vision models and humans, human subject experiments were conducted to understand the ability of human reading residual plots.
 
-## Structure of the paper
-
-...
-
-# Human Subject Exeriments
+## Human Subject Exeriments
 
@@
 
-## Cubic Model
+### Cubic Model
 
@@
 
-## Heteroskedasticity Model
+### Heteroskedasticity Model
 
@@
 
-## Distribution of regressors
+### Distribution of regressors
 
@@
 
-## Experiment I and II
+### Experiment I and II
 
@@
 
-## $P$-value and Power Estimation of Visual Test Allowed for Multiple Selections
+### $P$-value and Power Estimation of Visual Test Allowed for Multiple Selections
 
@@
 
-## Results
+### Results
 
@@
 
-### Power Estimation
+#### Power Estimation
 
@@
 
-### Effect of Factors on Power of the Visual Test
+#### Effect of Factors on Power of the Visual Test
 
@@
 
-# Computer Vision Model
-
-Visual test can be used to make valid inference on visual discoveries and acknowledging people with confidence about the findings.
-However, the lack of scalability of this method due to relatively high human cost compared to conventional test limits its application in wider fields.
-To relieve human analysts of the job of diagnosing predictive models in the context of visual inference, AI agents which can read data plot, evaluate lineups, and provide decisions need to be built.
+## Automatic Visual Inference System
+
+Visual tests can be used to make valid inference for visual discoveries while acknowledging people with confidence about the findings.
+However, the lack of scalability of this method limits its application in wider fields.
+The need of automatic visual inference system is auto data analysis and machine learning system.
+
+AI agents which can read data plot and evaluate lineups could
 
@@
 
-## Automatic visual inference system
-
-For an automatic visual inference system, there are several possibilities of the type of inputs.
+### Automatic visual inference system
+
+For an automatic visual inference system, types of inputs can be classified by two decisions: (i) whether to accept multiple image files as inputs, and (ii) whether to accept fitted model.
+
+single image file
+multiple image files
+
+fitted model
 
 Depend on the inputs and outputs, computer vision model can be integrated into an automatic visual inference system.
+
+### Artificial Intelligence: Four Approaches {#se:ai}
+
+Since the automatic visual inference system is intended to play the role of evaluators as humans in visual inference, some degree of intelligence needs to be demonstrated.
+This brings us to one of the popular filed of study - **Artificial Intelligence** (AI).
+AI is the field of research concerned with understanding and building machines who can demonstrate intelligence.
+As discussed in @russell_artificial_2002, historically, there are disagreements among researchers about the definition of intelligence, which is caused by two critical questions:
+
+1. Should AI act and think humanly or rationally? 
+2. Without the thought process and reasoning, are behaviours sufficient to demonstrate intelligence?
+
+Based on the answer to the above questions, four major approaches to pursue AI have been established.
+These approaches can be summarized into a two by two table as shown in Figure \@ref(fig:two-dim-ai), where the row is "Human" vs.
+"Rational", and the column is "Behaviour" vs.
+"Thought".
+Positioning at the top right cell, the **rational agent approach** aims to build agent that perform mathematically perfect acts such that the best expected outcome can always be achieved.
+In contrast, the **"laws of thought" approach** focus on understanding the logic behind the rationality.
+Closely related to **cognitive science**, the **cognitive modelling approach** attempts to express theories of human cognition as computer program to mimic the thought process of human.
+Lastly, the **Turing test approach** is built upon the famous **Turing test** proposed by @turing_computing_1950.
+The test can be roughly described as, whether a human can distinguish another human from a computer with written communications only.
+To pass the test, several capabilities of computer are required.
+This includes **natural language processing** for communication with human, **knowledge representation** for encoding knowledge, **automated reasoning** for derivation of conclusions and **machine learning** for improving AI automatically through experience and data.
+Some researchers argued that written communication is insufficient to demonstrate intelligence, and some degree of physical simulation of a person is still necessary.
+One such example is the **total Turing test** proposed by @harnad_other_1991.
+It adds three new requirements to the list, including **computer vision**, **speech recognition** and **robotics**, which are response for interactions with the physical world.
+Notably, all seven required capabilities have become major subfields of AI today.
+And their development has made AI one of the fastest-growing fields in the 21st century [@russell_artificial_2002].
+
+```{r two-dim-ai, fig.cap="Four possible approaches to pursue AI based on the two dimensions in AI research - human vs. rational and thought vs. behavior [@russell_artificial_2002]."}
+ggplot() +
+  geom_hline(aes(yintercept = 1), col = "grey") +
+  geom_vline(aes(xintercept = 1), col = "grey") +
+  geom_text(aes(x = 0.5 , y = 1.5, label = "The Turing test approach")) +
+  geom_text(aes(x = 1.5 , y = 1.5, label = "The rational agent approach")) +
+  geom_text(aes(x = 0.5 , y = 0.5, label = "The cognitive modelling approach")) +
+  geom_text(aes(x = 1.5 , y = 0.5, label = 'The "laws of thought" approach')) +
+  theme_light() +
+  xlab("") +
+  ylab("") +
+  scale_x_continuous(breaks = c(0.5, 1.5), labels = c("Human", "Rational"), limits = c(0, 2)) +
+  scale_y_continuous(breaks = c(0.5, 1.5), labels = c("Thought", "Behaviour"), limits = c(0, 2)) +
+  theme(panel.grid = element_blank(), axis.ticks = element_blank())
+```
 
@@
 
-## Model Architecture
+### Model Architecture
 
@@
 
-## Performance
+### Performance
 
@@
 ```
+
+# Timeline (need discussion)

### 2022-03-07 Milestone_1/_book/thesis.tex
add milestone report

--- 2022-03-06 Thesis/_book/thesis.tex
@@
 \degrees{B.Comm. (Hons), Monash University}
-\def\degreetitle{Doctor of Philosophy}
+\def\degreetitle{PhD}
 % Add subject and keywords below
@@
 
-\hypertarget{copyright-notice}{%
-\chapter*{Copyright notice}\label{copyright-notice}}
-\addcontentsline{toc}{chapter}{Copyright notice}
-
-\emph{(Choose one of the following notices.)}
-
-\emph{(Notice 1)}
-
-\textcopyright { } \authorname~(\number\the\year).
-
-\emph{The second notice certifies the appropriate use of any third-party material in the thesis.
-Students choosing to deposit their thesis into the restricted access section of the repository are not required to complete Notice 2.}
-
-\emph{(Notice 2)}
-
-\textcopyright { } \authorname~(\number\the\year).
-
-I certify that I have made all reasonable efforts to secure copyright permissions for third-party content included in this thesis and have not knowingly added copyright content to my work without the owner's permission.
-
-\newpage
-
-\hypertarget{abstract}{%
-\chapter*{Abstract}\label{abstract}}
-\addcontentsline{toc}{chapter}{Abstract}
-
-The abstract should outline the main approach and findings of the thesis and must not be more than 500 words.
-
-\newpage
-
-\hypertarget{declaration}{%
-\chapter*{Declaration}\label{declaration}}
-\addcontentsline{toc}{chapter}{Declaration}
-
-\emph{(Standard thesis)}
-
-This thesis is an original work of my research and contains no material which has been accepted for the award of any other degree or diploma at any university or equivalent institution and that, to the best of my knowledge and belief, this thesis contains no material previously published or written by another person, except where due reference is made in the text of the thesis.
-
-\emph{(Thesis including published works declaration)}
-
-I hereby declare that this thesis contains no material which has been accepted for the award of any other degree or diploma at any university or equivalent institution and that, to the best of my knowledge and belief, this thesis contains no material previously published or written by another person, except where due reference is made in the text of the thesis.
-
-This thesis includes (insert number) original papers published in peer reviewed journals and (insert number) submitted publications.
-The core theme of the thesis is (insert theme).
-The ideas, development and writing up of all the papers in the thesis were the principal responsibility of myself, the student, working within the (insert name of academic unit) under the supervision of (insert name of supervisor).
-
-(The inclusion of co-authors reflects the fact that the work came from active collaboration between researchers and acknowledges input into team-based research.) Remove this paragraph for theses with sole-authored work
-
-In the case of (insert chapter numbers) my contribution to the work involved the following:
-
-I have / have not renumbered sections of submitted or published papers in order to generate a consistent presentation within the thesis.
+\clearpage\pagenumbering{arabic}\setcounter{page}{0}
+
+\hypertarget{background-and-motivation}{%
+\chapter{Background and Motivation}\label{background-and-motivation}}
+
+\hypertarget{predictive-modelling-and-visual-diagnostics}{%
+\section{Predictive Modelling and Visual Diagnostics}\label{predictive-modelling-and-visual-diagnostics}}
+
+\textcite{donoho_50_2017} in its summary of data science stated that the concept of the predictive modelling culture could be traced back to an article written by Breiman (2001).
+In contrast to the generative modelling culture, which aims to develop stochastic models to make inferences about the data generating process, predictive modelling emphasizes the ability of the model to make accurate predictions.
+
+Predictive models are primarily evaluated by predictive accuracy with the use of validation and test data, but in predictive model diagnostics, especially model testing and tuning, data plots play an irreplaceable role.
+In these diagnostics, though numeric summaries are mostly available and some are even endorsed by finite or asymptotic properties, graphical representation of data is still preferred, or at least needed by researchers, due to its intuitiveness and the possibility for unexpected discoveries which may be abstract and unquantifiable.
+
+However, unlike confirmatory data analysis built upon rigorous statistical procedures, e.g., hypothesis testing, visual diagnostics relies on graphical perception - human's ability to interpret and decode the information embedded in the graph \autocite{cleveland_graphical_1984}, which is to some extent subjective.
+Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
+One such example is finding an over-interpretation of the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurence \autocite{roy_chowdhury_using_2015}.
+
+\hypertarget{visual-inference}{%
+\section{Visual Inference}\label{visual-inference}}
+
+Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title ``Inference for Data Visualizatino'' by \textcite{buja_inference_1999} as an idea to address the issue of valid inference for visual discoveries of data plots \autocite{gelman_exploratory_2004}.
+Later, in the Bayesian context, data plots was systematically considered as model diagnostics by taking advantage of the data simulated from the assumed statistical models \autocite{gelman_bayesian_2003,gelman_exploratory_2004}.
+
+It was surprising that the essential components of visual inference had actually been established in \textcite{buja_inference_1999}, but it was not until 10 years later that \textcite{buja_statistical_2009} formalized it as an inferential framework to extend confirmatory statistics to visual discoveries.
+This framework redefines the test statistics, tests, null distribution, significance levels and \(p\)-value for visual discovery modelled on the confirmatory statistical testing.
+Figure \ref{fig:parallelism} outlines the parallelism between conventional tests and visual discovery.
+
+\begin{figure}
+\centering
+\includegraphics[width=4.6875in,height=3.55208in]{figures/rsta2009012001.jpg}
+\caption{Parallelism between multiple quantitative testing and visual discovery \autocite{buja_statistical_2009}.
+Visible features in a plot are viewed as a collection of test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\), and any visual discoveries that are inconsistent with the null hypothesis are treated as evidence against the null.
+For model diagnostics, the null hypothesis would be the assumed model, and visual discoveries could be any visual features in favour of any alternatives.
+\label{fig:parallelism}}
+\end{figure}
+
+In visual inference, a visual discovery is defined as a rejection of a null hypothesis, and the same null hypothesis can be rejected by many different visual discoveries \autocite{buja_statistical_2009}.
+For model diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the null hypothesis.
+The same assumed model, such as classical linear regression model, can be rejected by many reasons with residual plot, including nonlinearity and heteroskedasticity as shown in Figure \ref{fig:residual-plot-cubic-heter}.
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/residual-plot-cubic-heter-1.pdf}
+\caption{\label{fig:residual-plot-cubic-heter}Residuals vs.~fitted values plot for a classical linear regression model.
+The residuals are produced by fitting a two-predictor multiple linear regression model with data generated from a cubic linear model.
+From the residual plot, ``butterfly shape'' can be observed which generally would be interpretd as evidence of heteroskedasticity.
+Further, from the outline of the shape, nonlinear patterns exist.
+Both visual discoveries are evidence against the null hypothesis, though heteroskedasticity actually does not exist in the data generating process.}
+\end{figure}
+
+\hypertarget{pre-specification-of-visual-discoverable-features}{%
+\section{Pre-specification of Visual Discoverable Features}\label{pre-specification-of-visual-discoverable-features}}
+
+As discussed in \textcite{buja_statistical_2009}, in the practice of model diagnostics, the range of possible visual discoveries is not pre-specified.
+In other words, people do not explicitly specify which visual feature(s) they are looking for before the reading the diagnostic plot.
+In contrary, conventional hypothesis testing always requires the pre-specification of the parameter space \(\Theta\) of the parameter of interest \(\theta \in \Theta\) to form a valid inferential procedure.
+To address this issue, a collection of test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) is defined, where \(\boldsymbol{\mathrm{y}}\) is the data and \(I\) is a set of all possible visual features.
+\textcite{buja_statistical_2009} described each of the test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})\) as a measurement of the degree of presence of a visual feature.
+Alternatively, \textcite{majumder_validation_2013} avoids the use of visual features and defined the visual statistics \(T(.)\) as a mapping from a dataset to a data plot.
+Both definitions of visual test statistics are valid, but in the rest of the paper the first definition will be used as it covers some details needed by the following discussion.
+
+The size of the collection \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) depends on the size of the set \(I\).
+Thus, if one can define \(I\) comprehensively, i.e, pre-specify all the visual discoverable features, the validity issue will be solved.
+Unfortunately, to our knowledge, there is no such a way to list all visual features.
+In linear regression diagnostics, possible visual features of a residual plot may be outliers, shapes and clusters.
+But this is an incomplete list which does not enumerate all the visual features.
+
+Similarly, \textcite{wilkinson_graph-theoretic_2005} proposed the work called graph theoretic scagnostics, which adopted the idea of ``scagnostics'' - scatter plot diagnostics from (can't find the 1984 citation).
+It includes 9 computable scagnostics measures defined on planar proximity graphs: ``Outlying'', ``Convex'', ``Skinny'', ``Stringy'', ``Straight'', ``Monotonic'', ``Skewed'', ``Clumpy'' and ``Striated'' which attempts to describe outliers, shape, density, trend and coherence of the data.
+This approach is inspiring but it still does not give the complete list of visual discoverable features.
+In fact, it is possible that such a list will never be complete as suggested in \textcite{buja_statistical_2009}.
+
+Thinking out of the box, \textcite{buja_statistical_2009} argued that there is actually no need for pre-specification of visual discoverable features.
+In model diagnostics, when the null hypothesis is rejected, the reasons for rejecting the hypothesis will also be known.
+This is because observers can not only point out the fact that visual discoveries have been found, but also describe the particular visual features they observed.
+Those features will correspond to the subset of the collection of visual test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) which resulted in rejection.
+This argument helps justifies the validity of visual inference.
+
+\hypertarget{sampling-from-the-null-distribution}{%
+\section{Sampling from the Null Distribution}\label{sampling-from-the-null-distribution}}
+
+In visual inference, the null distribution of plots refers to the infinite collection of plots of null datasets sampled from \(H_0\).
+It is defined as the analogue of the null distribution of test statistics in conventional test \autocite{buja_statistical_2009}.
+In practice, a finite number of plots of null datasets could be generated, called null plots.
+
+In the context of model diagnostics, sampling data from \(H_0\) is equivalent to sampling data from the assumed model.
+As \textcite{buja_statistical_2009} suggested, \(H_0\) is usually composited by a collection of distributions controlled by nuisance parameters.
+Since statistical models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to so called ``reference distribution'' by applying one of the three methods: (i) sampling from a conditional distribution given a minimimal sufficient statistic under \(H_0\), (ii) parametric bootstrap sampling with nuisance parameters estimated under \(H_0\), and (iii) Bayesian posterior predictive sampling.
+
+The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three \autocite{buja_statistical_2009}.
+Focus will be put on simulation based on this distribution as it is applicable to classical normal linear regression model.
+Details of the other two are can be found in the supplementary material of \textcite{buja_statistical_2009}.
+Suppose there exists a minimal sufficient statistic \(\boldsymbol{S}(\boldsymbol{y})\) under the null hypothesis, any null datasets \(\boldsymbol{y^{*}}\) should fulfil the condition \(\boldsymbol{S}(\boldsymbol{y}) = \boldsymbol{s}\).
+For classical normal linear regression model, the minimal sufficient statistic is \(\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{\beta}}, \boldsymbol{e}'\boldsymbol{e})\), where \(\hat{\boldsymbol{\beta}}\) are the coefficient estimators and \(\boldsymbol{e}'\boldsymbol{e}\) is the residual sum of square.
+Alternatively, the minimal sufficient statistic can be constructed as \(\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{y}}, ||\boldsymbol{e}||)\), where \(\hat{\boldsymbol{y}}\) are the fitted values and \(||\boldsymbol{e}||\) is the length of residuals, which is more intuitive as suggested by \textcite{buja_statistical_2009}.
+Since the fitted values are held fixed, the variation can only occur in the residual space.
+And because the length of residual is also held fixed, residuals obtained from a null dataset has to be a random rotation of \(\boldsymbol{e}\) in the residual space.
+With this property, null residuals can be simulated by regressing \(N\) i.i.d standard normal random draws on the regressors, then rescaling it by the ratio of residual sum of square in two regressions.
+
+\hypertarget{lineup-protocol}{%
+\section{Lineup Protocol}\label{lineup-protocol}}
+
+With the validity of visual inference being justified and the simulation of null plots being provided, another aspect of hypothesis testing that needs to be addressed is the control of false positive rate or Type I error.
+Any visual statistic \(T^{(i)}(\boldsymbol{\mathrm{y}})\) needs to pair with a critical value \(c^{(i)}\) to form a hypothesis test.
+When a visual feature \(i\) is discovered by the observer from a plot, the corresponding visual statistic \(T^{(i)}(\boldsymbol{\mathrm{y}})\) may not be known as there is no general agreement on the measurement of the degree of presence of a visual feature.
+It is only the event that \(T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}\) is confirmed.
+Similarly, if any visual discovery is found by the observer, we say, there exists \(i \in I:~T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}\) \autocite{buja_statistical_2009}.
+
+Using the above definition, the family-wise Type I error can be controlled if one can provide the collection of critical values \(c^{(i)}~(i \in I)\) such that \(P(\mathrm{there~exists~} i \in I: T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}|\boldsymbol{\mathrm{y}}) \leq \alpha\), where \(\alpha\) is the significance level.
+However, since the quantity of \(T^{(i)}(\boldsymbol{\mathrm{y}})\) may not be known, such collection of critical values can not be provided.
+
+\textcite{buja_statistical_2009} proposed the lineup protocol as a visual test to calibrate the Type I error issue without the specification of \(c^{(i)}~(i \in I)\).
+It is inspired by the ``police lineup'' or ``identity parade'' which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people.
+The protocol consists of \(m\) randomly placed plots, where one plot is the actual data plot, and the remaining \(m - 1\) plots have the identical graphical production as the data plot except the data has been replaced with data consistent with the null hypothesis.
+Then, an observer who have not seen the actual data plot will be asked to point out the most different plot from the lineup.
+
+Under the null hypothesis, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is \(1/m\).
+If we reject the null hypothesis as the observer correctly picks the actual data plot, then the Type I error of this test is \(1/m\).
+
+This provides us with an mechanism to control the Type I error, because \(m\) - the number of plots in a lineup can be chosen.
+A larger value of \(m\) will result in a smaller Type I error, but the limit to the value of \(m\) depends on the number of plots a human is willing to view \autocite{buja_statistical_2009}.
+Typically, \(m\) will be set to \(20\) which is equivalent to set \(\alpha = 0.05\), a general choice of significance level for conventional testing among statisticians.
+
+Further, if we involve \(K\) independent observers in a visual test, and let \(X\) be a random variable denoting the number of observers correctly picking the actual data plot.
+Then, under the null hypothesis \(X \sim \mathrm{Binom}_{K,1/m}\), and therefore, the \(p\)-value of a lineup of size \(m\) evaluated by \(K\) observer is given as
+\begin{equation} \label{eq:pvaluesingle}
+P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},
+\end{equation}
+
+where \(x\) is the realization of number of observers correctly picking the actual data plot \autocite{majumder_validation_2013}.
+
+The multiple individuals approach avoids the limit of \(m\), while provides visual tests with \(p\)-value much smaller than \(0.05\).
+In fact, the lower bound of \(p\)-value decreases exponentially as \(K\) increases.
+With just \(4\) individuals and \(20\) data plots in a lineup, the \(p\)-value could be as small as \(0.0001\).
+Additionally, by involving multiple observers, variation of individual ability to read plots can be addressed to some degree as different opinions about visual discoveries can be collected.
+
+Compared to the conventional test, whose power only depends on the parameter of interest \(\theta\), several studies \autocites[see][]{hofmann_graphical_2012,majumder_validation_2013,majumder_human_2014,roy_chowdhury_using_2015,loy_variations_2016} have shown the power of the visual test is subject-specific.
+Thus, to be able to account for individual's ability, an individual is required to evaluate multiple lineups \autocite{majumder_validation_2013}.
+
+Suppose individuals have the same ability and a lineup has been evaluated by multiple individuals, under the alternative hypothesis, the estimated power for a lineup can be expressed as \(\hat{p} = x/K\), the estimated probability of identifying the actual data plot from the lineup.
+If the individual skill needs to be taken into account, and \(L\) lineups have been evaluated by \(K\) individuals, \textcite{majumder_validation_2013} suggests that mixed effects logistic regression model can be fit as:
+
+\[g(p_{li}) = W_{li}\delta + Z_{li}\tau_{li},\]
+where \(g(.)\) is the logit link function \(g(p) = log(p) - log(1-p)\); \(0 \leq p \leq 1\).
+\(W_{li}\), \(1 \leq i \leq K\), \(1 \leq l \leq L\), is the covariate matrix including lineup-specific elements and demographic information of individuals, and \(\delta\) is a vector of parameters.
+\(Z\) is the random effects matrix, and \(\tau\) is a vector of variables follow \(N(\boldsymbol{0},\sigma_{\tau}\boldsymbol{I}_{KL\times KL})\).
+
+Then, the estimated power for lineup \(l\) and individual \(i\) can be calculated as \(\hat{p}_{li} = g^{-1}(W_{li}\hat{\delta} + Z_{li}\hat{\tau}_{li})\) \autocite{majumder_validation_2013}.
+
+\hypertarget{limitations-of-lineup-protocol-and-automatic-visual-inference}{%
+\section{Limitations of Lineup Protocol and Automatic Visual Inference}\label{limitations-of-lineup-protocol-and-automatic-visual-inference}}
+
+Although lineup protocol has already been integrated into data analysis of various topics, such as diagnostics of hierarchical linear models \autocite{loy_diagnostic_2013}, geographical Research \autocite{widen_graphical_2016} and forensic examinations \autocite{krishnan_hierarchical_2021}, the involvement with human judgements limits its popularity.
+Similar to handicraft in pre-industrial society, lineup protocol conducted by humans is infeasible on a large scale, high in labour cost and time consuming.
+Moreover, it is strongly unfriendly to vision-impaired people.
+
+The ``steam engine'' of visual inference needs to developed for relieving people's workload by automating repeating tasks and providing standard result in a control environment.
+Large-scale evaluation of visual tests is not possible without the use of technology and machines.
+
+Modern computer vision model could be a promising solution to this problem.
+As a subfield of AI, computer vision with the modern deep learning architectures solved numerous critical problems in automation.
+Inspired by the vision processing in living organisms, the convolutional neural network was introduced by \textcite{fukushima_neocognitron_1982}.
+Soon, this architecture was applied to hand-written number recognition trained with back-propagation by \textcite{lecun_backpropagation_1989}.
+This was one of the earliest attempts which human successfully extract information from digital images via self-learning algorithms.
+Modern computer vision model is typically built on the deep neural network with convolutional layers \autocite{fukushima_neocognitron_1982}.
+Convolutional layers take advantage of the hierarchical pattern in data and provide regularized versions of fully-connected layers.
+It downscales and transforms the image by summarising information in a small space.
+Numerous studies have shown that it can be used to effectively tackle vision tasks, such as image recognition \autocite{rawat_deep_2017}.
+With the development of graphics processing units and the spread of high-performance personal computers, researches in computer vision become a new hype in the 21st century.
+Achievements such as computer-aided diagnosis \autocite{lee_image_2015}, pedestrian detection \autocite{brunetti_computer_2018} and facial recognition \autocite{emami_facial_2012} had a significant impact on our daily life.
+
+Using computer vision models to read data plots is not a general choice.
+some fields have adopted this idea by applying computer vision models to read recurrence plots for time series regression \autocite{ojeda_multivariate_2020}, time series classification \autocite{chu_automatic_2019,hailesilassie_financial_2019,hatami_classification_2018,zhang_encoding_2020}, anomaly detection \autocite{chen_convolutional_2020} and pairwise causality analysis \autocite{singh_deep_2017}.
+However, evaluating lineups with computer vision model is a new field of study.
+
+\hypertarget{structure-of-the-thesis-need-disscussion}{%
+\chapter{Structure of the Thesis (need disscussion)}\label{structure-of-the-thesis-need-disscussion}}
+
+An overview of the thesis, including some brief detail of the proposed content of all chapters.
+This should encompass a description of the theoretical and conceptual framework that underlies the thesis, and the procedures that are to be used in addressing the research questions
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Automatic Visual Statistical Inference, with Application to Linear Regression Diagnostics
+\item
+  \ldots{}
+\item
+  \ldots{}
+\end{enumerate}
+
+\hypertarget{automatic-visual-statistical-inference-with-application-to-linear-regression-diagnostics}{%
+\chapter{Automatic Visual Statistical Inference, with Application to Linear Regression Diagnostics}\label{automatic-visual-statistical-inference-with-application-to-linear-regression-diagnostics}}
+
+This project focused on building a prototype of automatic visual statistical inference system for evaluating residual plots of classical normal linear regression model.
+To set up a comparison between the computer vision models and humans, human subject experiments were conducted to understand the ability of human reading residual plots.
+
+\hypertarget{human-subject-exeriments}{%
+\section{Human Subject Exeriments}\label{human-subject-exeriments}}
+
+To collect data of human performance on reading residual plot of linear regression model with nonlinearity and heteroskedasticity defects, two human experiments were conducted.
+Participants of both experiments were recruited using an online platform called Prolific \autocite{prolific_prolific_2022}.
+
+Prolific provides an international participant pool with the option to apply flexible pre-screening filters.
+In this study, we recruited \(62\) participants who was fluent in English with at least \(10\) previous submissions and \(98\)\% approval rate in other Prolific studies for quality control.
+Further, balance sample across gender was imposed to prevent gender bias.
+
+In Prolific, researchers can either approve or reject submissions based on the quality of the responses.
+If a submission is approved by the researcher, the participant will be paid a certain amount of money per hour of time spent on the experiment.
+To assess the quality of the responses, two attention checks were given to each participant during the experiment, where at least one of them was required to pass for the approval of submission.
+
+Throughout the experiments, participants were requested to complete a short survey and evaluate 20 lineups on a website in an hour.
+Each lineup consists of one actual residual plot and 19 null residual plots produced by plotting null residuals simulated from the residual rotation distributions.
+Among the 20 lineups, there were two extremely easy lineups used as attention checks which everyone should get correct.
+The short survey was intended to collect information about participant that might affect their ability in reading data plot including age, highest level of education, preferred pronoun and previous experience in similar study.
+For the evaluation of the lineup, participant first needed to select one or multiple most different data plot(s) from the lineup by clicking the corresponding image.
+Then, the primary reason for choosing the plots needed to be provided by picking one of the given options - ``outlier(s)'', ``cluster(s)'', ``shape'' and ``other''.
+Explanations about these options were provided in the training page and the mouseover text.
+Table \ref{tab:lineup-reason} gives the detailed explanations about these reasons.
+If the option ``other'' was selected, text input for the specified reason would be collected.
+Lastly, the degree of difference between their chosen data plot(s) and other data plots needed to be selected among five levels - ``not at all'', ``slightly'', ``moderately'', ``very'' and ``extremely''.
+Notably, if participants could not tell the difference between the data plots, there was an option to skip the evaluation of the lineup.
+However, to prevent participants abusing this option, warnings were given at the beginning of the study that skipping too many lineups might lead to rejection of submission since it demonstrated clear low-effort throughout the experiment.
 
 \begin{table}
-\centering\footnotesize\tabcolsep=0.12cm
-\begin{tabular}{|p{1.2cm}|p{2cm}|p{1.8cm}|p{3.4cm}|p{3.5cm}|p{1.5cm}|}
+
+\caption{\label{tab:lineup-reason}Explanations about reasons for choosing data plots from a lineup}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{ll}
+\toprule
+Reasons & Explanations\\
+\midrule
+outlier(s) & In a data plot, an outlier is a point that differs significantly from other points.\\
+cluster(s) & In a data plot, a cluster is a group of points positioned closely together. 
+And usually, there will be gaps between different clusters.\\
+shape & It could be any common shapes we would see in real life, like triangle shape, U-shape, butterfly shape, etc.\\
+other & Patricipants needs to give their reasons in the text input.\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+The study website was powered by Flask \autocite{grinberg_flask_2018}, a web framework written in Python 3 \autocite{van_rossum_python_2009}, hosted on PythonAnywhere \autocite{pythonanywhere_pythonanywhere_2022}, a web hosting service provider.
+Due to the limit of the storage on PythonAnywhere, lineup images needed for the website were hosted separately on Github Pages \autocite{gtihub_github_2022-1}, a static site hosting service provided by Github \autocite{gtihub_github_2022}, stored in private Github repositories.
+The uniform resource locator (URL) to lineup images were set to be unique random strings such that images can not be accessed by general approaches without knowing the correct URL.
+This avoided participants seeing the lineup beforehand.
+The front-end of the website was built using a JavaScript \autocite{flanagan_javascript_2006} library jsPsych \autocite{de_leeuw_jspsych_2015} which is specialized in creating online behavioral experiments.
+One of the reasons we chose to use this library was it had a modularized but highly customizable template which could record participant's response time automatically.
+This is essential for us to confirm the quality of the data by checking exceptionally fast or slow responses.
+Figure \ref{fig:diagram-experiment} summarizes the online experiment setup.
+
+\begin{figure}
+\centering
+\includegraphics[width=4.6875in,height=4.6875in]{figures/experiment_tech.pdf}
+\caption{Diagram of online experiment setup.
+The server-side of the study webiste used Flask as backend hosted on PythonAnywhere.
+And the client-side used jsPsych to run experiment.
+\label{fig:diagram-experiment}}
+\end{figure}
+
+The first two pages of the website are the explanatory statement and the consent form.
+Participants were asked to read the documents and agree with the terms to advance to the short survey.
+With the completion of the survey, the next page is the training page, which contained instructions on how to interact with the webpage.
+Followed by the training page, there are 20 lineups each on a single page.
+Figure \ref{fig:website-layout} illustrates the layout of the website.
+At the end of the experiment, participants would be redirected back to Prolific and waited for researchers' responses.
+
+\begin{figure}
+\centering
+\includegraphics[width=5.72917in,height=3.11458in]{figures/website.png}
+\caption{Layout of the study website. Participants needed to choose the most different plots on the right and select their reasons and confidence levels on the left. \label{fig:website-layout}}
+\end{figure}
+
+Next we will discuss the simulation setup for this study.
+The experiment data was simulated by the use of the programming language \texttt{R} \autocite{r_core_team_r_2021}.
+For the ease of reproducibility, functions to build models, simulate data from models, produce lineup with data, allocate stimulis for subjects and evaluate subject responses from this study are bundled in the package \texttt{visage} with a unique object-oriented programming system built upon the environment feature of \texttt{R}.
+In the description of the simulation, corresponding functionalities of the package will be introduced.
+
+Two models were used in the study, where both were linear models with some degree of violations of classical assumptions.
+
+\hypertarget{cubic-model}{%
+\subsection{Cubic Model}\label{cubic-model}}
+
+The first model was a cubic linear model with two regressors, which can be expressed by:
+\begin{equation} \label{eq:cubic}
+\boldsymbol{Y}= 1 + (2-c)\boldsymbol{X} + c\boldsymbol{Z} + a[(2-c)\boldsymbol{X}]^2+a(c\boldsymbol{Z})^2+b[(2-c)\boldsymbol{X}]^3+b(c\boldsymbol{Z})^3+\boldsymbol{\varepsilon},
+\end{equation}
+where \(c \in (0,2)\), \(a \in (-3,3)\), \(b \in (-3,3)\), \(\boldsymbol{\varepsilon}\overset{iid}{\sim} N(\boldsymbol{0},\sigma^2\boldsymbol{I})\), \(\boldsymbol{Y}\), \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) are \(n\times1\) matrices.
+
+This defines a cubic relationship between \(\boldsymbol{Y}\), \(\boldsymbol{X}\) and \(\boldsymbol{Z}\).
+Meanwhile, to create nonlinearity defect, the null model followed the assumptions of the classical normal linear regression model (CNLRM), fitted by OLS is:
+\begin{equation} \label{eq:cubic_null}
+\boldsymbol{Y}=\beta_0+\beta_1\boldsymbol{X}+\beta_2\boldsymbol{Z}+\boldsymbol{u},
+\end{equation}
+where \(\boldsymbol{u} \sim N(0,\sigma^2_u\boldsymbol{I})\).
+
+Clearly, omitted-variable bias will present since the null model leaves out the quadric and cubic terms.
+
+\begin{lemma}[Distribution of residuals produced by the cubic model] \label{lemma:cubic}
+Given the data generating process in Equation (\ref{eq:cubic}), and null model in Equation (\ref{eq:cubic_null}).
+Let $\boldsymbol{X}_a=[\boldsymbol{1},\boldsymbol{X},\boldsymbol{Z}]$ denotes the set of regressors in matrix form.
+Then, the residuals obtained from the null model are $$\boldsymbol{e} \sim N(\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b, \sigma^2\boldsymbol{R}_a),$$ where $\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$, $\boldsymbol{X}_b=[\boldsymbol{X}^2,\boldsymbol{Z}^2,\boldsymbol{X}^3,\boldsymbol{Z}^3]$ and $\boldsymbol{\beta}_b=(a(2-c)^2,ac^2,b(2-c)^3,bc^3)'$.
+\end{lemma}
+
+\begin{proof}
+Using the Frisch–Waugh–Lovell theorem, the residuals obtained by the null model are $$\boldsymbol{e}=\boldsymbol{R}_a\boldsymbol{Y}=\boldsymbol{R}_a(\boldsymbol{X}_a\boldsymbol{\beta}_a+\boldsymbol{X}_b\boldsymbol{\beta}_b+\boldsymbol{\varepsilon}),$$ where $\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$, $\boldsymbol{\beta}_a=(1,2-c,c)'$, $\boldsymbol{X}_b=[\boldsymbol{X}^2,\boldsymbol{Z}^2,\boldsymbol{X}^3,\boldsymbol{Z}^3]$ and $\boldsymbol{\beta}_b=(a(2-c)^2,ac^2,b(2-c)^3,bc^3)'$.
+
+Because $\boldsymbol{R}_a\boldsymbol{X}_a=\boldsymbol{0}$, we have $\boldsymbol{e}=\boldsymbol{R}_a(\boldsymbol{X}_b\boldsymbol{\beta}_b+\boldsymbol{\varepsilon}).$ Since $\boldsymbol{\varepsilon} \sim N(0,\sigma^2\boldsymbol{I})$, it follows that $\boldsymbol{e} \sim N(\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b, \sigma^2\boldsymbol{R}_a)$.
+\end{proof}
+
+Lemma \ref{lemma:cubic} shows that the expectation of the residuals is clearly a function of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\), which indicates the residuals and the fitted values \(\hat{Y} = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'\boldsymbol{Y}\) are associated.
+Hence, it is expected that visual discoveries can be found in the residuals vs.~fitted values plot.
+Let \(X_i\) and \(Z_i\), \(i = 1,...,n\), be independent random variables follow uniform distribution \(U(-1,1)\).
+Given the expectation of the residuals, we could plot the expected values of residuals against the observed values.
+Figure \ref{fig:cubic-shape-1}, \ref{fig:cubic-shape-2} and \ref{fig:cubic-shape-3} illustrate the shape of residuals and their expected values under different parameter settings.
+
+From the Figure \ref{fig:cubic-shape-1}, it can be observed that with fixed \(\sigma\) and \(c\), \(a\) and \(b\) are controlling the 2D projection of a hypersurface, and seemingly performing some rotations along different axes.
+Figure \ref{fig:cubic-shape-2} shows that with fixed \(a\), \(b\) and \(\sigma\), \(c\) is controlling the contribution of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) to \(\boldsymbol{Y}\).
+As \(c\) moves toward \(0\) or \(2\), one regressor will dominate another, which will mitigate some joint effects and resemble a typical cubic function.
+In Figure \ref{fig:cubic-shape-3}, \(a\), \(b\) and \(c\) are held fixed, and \(\sigma\) is controlling the noises around the expected values.
+The underlying shape is clearer for smaller \(\sigma\).
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/cubic-shape-1-1.pdf}
+\caption{\label{fig:cubic-shape-1}A matrix of residuals vs.~fitted values plot under different parameter settings of the cubic model.
+Resdiauls are drawns in black and the expected values are drawn in red.
+The plots show that if \(\sigma\) and \(c\) are held fixed, \(a\) and \(b\) rotate a two-dimension projection of a hypersurface.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/cubic-shape-2-1.pdf}
+\caption{\label{fig:cubic-shape-2}A matrix of residuals vs.~fitted values plot under different parameter settings of the cubic model.
+Resdiauls are drawns in black and the expected values are drawn in red.
+The plots show that if \(a\), \(b\) and \(\sigma\) are held fixed, \(c\) controls the contribution of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) to \(\boldsymbol{Y}\).}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/cubic-shape-3-1.pdf}
+\caption{\label{fig:cubic-shape-3}A matrix of residuals vs.~fitted values plot under different parameter settings of the cubic model.
+Resdiauls are drawns in black and the expected values are drawn in red.
+The plots show that if \(a\), \(b\) and \(c\) are held fixed, \(\sigma\) controls the strength of the signal.}
+\end{figure}
+
+The residuals used in these three figures are simulated from the cubic models built using the \texttt{cubic\_model()} function from the package \texttt{visage}.
+\texttt{cubic\_model()} is a cubic model class constructor, which takes arguments \texttt{a}, \texttt{b}, \texttt{c}, \texttt{sigma}, \texttt{x} and \texttt{z}, where the first four are numeric values defined above, and \texttt{x} and \texttt{z} are random variable instances created by the random variable abstract base class constructor \texttt{rand\_var()}.
+
+If we would like \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) to be random uniform variables ranged from \(-1\) to \(1\), it can be achieved by using the random uniform variable class constructor \texttt{rand\_uniform()} inherited from the random variable abstract base class.
+It only takes two arguments which are the lower bound and the upper bound of the support.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{library}\NormalTok{(visage)}
+\NormalTok{mod }\OtherTok{\textless{}{-}} \FunctionTok{cubic\_model}\NormalTok{(}\AttributeTok{a =} \SpecialCharTok{{-}}\DecValTok{3}\NormalTok{, }\AttributeTok{b =} \SpecialCharTok{{-}}\DecValTok{3}\NormalTok{, }\AttributeTok{c =} \DecValTok{1}\NormalTok{, }\AttributeTok{sigma =} \FloatTok{0.5}\NormalTok{, }
+                   \AttributeTok{x =} \FunctionTok{rand\_uniform}\NormalTok{(}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{), }\AttributeTok{z =} \FunctionTok{rand\_uniform}\NormalTok{(}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+An instance of cubic model class contains methods of simulating data and making residual plot.
+Method \texttt{mod\$gen()} returns a data frame containing realizations of \(X\), \(Z\), \(Y\) and \(\varepsilon\) simulated from the model.
+The number of realizations depends on the integer argument \texttt{n}.
+In addition, if argument \texttt{fit\_model\ =\ TRUE}, a null model will be fitted using the simulated data and residuals and fitted values will be included in the returned data frame.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{mod}\SpecialCharTok{$}\FunctionTok{gen}\NormalTok{(}\AttributeTok{n =} \DecValTok{5}\NormalTok{, }\AttributeTok{fit\_model =} \ConstantTok{TRUE}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+##            y          x          z           e     .resid    .fitted
+## 1 -0.8263671 -0.4988275  0.5569648 -0.06142887  1.2340651 -2.0604322
+## 2  0.6134661  0.1762335  0.6512424  0.99654342  1.1758970 -0.5624310
+## 3  0.2285071 -0.2814395 -0.1511556 -0.10996690 -0.3757603  0.6042674
+## 4 -4.6320382 -0.5813456  0.9251409 -0.60826338 -1.2664307 -3.3656075
+## 5  0.3148180  0.5647967  0.4406211  0.64596536 -0.7677711  1.0825892
+\end{verbatim}
+
+Method \texttt{mod\$plot()} produce a \texttt{ggplot} \autocite{wickham_ggplot2_2011} object.
+It takes a data frame containing columns \texttt{.resid} and \texttt{.fitted} as input, along with a character argument \texttt{type} indicating the type of the data plot, and other aesthetic arguments such as \texttt{size} and \texttt{alpha} to control the appearance of the plot.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{mod}\SpecialCharTok{$}\FunctionTok{plot}\NormalTok{(mod}\SpecialCharTok{$}\FunctionTok{gen}\NormalTok{(}\AttributeTok{n =} \DecValTok{100}\NormalTok{, }\AttributeTok{fit\_model =} \ConstantTok{TRUE}\NormalTok{), }\AttributeTok{type =} \StringTok{"resid"}\NormalTok{, }\AttributeTok{size =} \DecValTok{1}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{thesis_files/figure-latex/test3-1.pdf}
+
+Lineup is a matrix of residual plots which can be produced by using the methods \texttt{mod\$gen\_lineup()} and \texttt{mod\$plot\_lineup()}.
+Method \texttt{mod\$gen\_lineup} takes the number of realizations \texttt{n} and the number of plots in a lineup \texttt{k} as inputs.
+And the method \texttt{mod\$plot\_lineup()} has the same user interface as \texttt{mod\$plot()}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{mod}\SpecialCharTok{$}\FunctionTok{plot\_lineup}\NormalTok{(mod}\SpecialCharTok{$}\FunctionTok{gen\_lineup}\NormalTok{(}\AttributeTok{n =} \DecValTok{50}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{), }\AttributeTok{type =} \StringTok{"resid"}\NormalTok{, }\AttributeTok{size =} \DecValTok{1}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{thesis_files/figure-latex/unnamed-chunk-4-1.pdf}
+
+The cubic model class also provides method to compute the expected values of residuals.
+Method \texttt{mod\$E()} takes a data frame with columns \texttt{x} and \texttt{z} as input, and returns a vector of expected values of residuals.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{mod}\SpecialCharTok{$}\FunctionTok{E}\NormalTok{(mod}\SpecialCharTok{$}\FunctionTok{gen}\NormalTok{(}\AttributeTok{n =} \DecValTok{5}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## [1] -0.092183951 -0.003912092 -0.001531723  0.148353478 -0.050725712
+\end{verbatim}
+
+Since we know that under the null hypothesis, the residual \(\boldsymbol{e}\sim N(\boldsymbol{0},\sigma^2\boldsymbol{R}_a)\).
+Thus, the difference between the expected values \(\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b\) and \(\boldsymbol{0}\) represents the direct impact of the parameters \(a\) and \(b\) on the residuals.
+It is expected that the larger the magnitude of the expected value relative to the variance and covariance, the easier the human to spot the patterns in the residual plot.
+
+To obtain a measure of the impact of \(a\) and \(b\) on the residuals adjusted for variance and covariance, we need to address several properties of the residuals.
+First, the variance of the residuals \(\sigma^2\boldsymbol{R}_a\) is not an identity matrix.
+This can be fixed by standardizing the residuals by their variance-covariance matrix.
+Second, the difference between \(\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b\) and \(\boldsymbol{0}\) could be negative, which is not ideal for comparison.
+Thus, the magnitude needs to be squared.
+Third, the measure needs to be a scalar.
+We could apply a weighted average operator \(\boldsymbol{W}\) on the transformed expected residuals to obtain a single numeric value.
+For simplicity, we set \(\boldsymbol{W}=n^{-1}\boldsymbol{1}\).
+Considering the high time complexity of computing the square root of \(\boldsymbol{R}_a\), off-diagonal elements of \(\boldsymbol{R}_a\) are set to be zeros.
+This gives the effect size: \[ES=n^{-1}||\sigma^{-1}\boldsymbol{R}_a^{-\frac{1}{2}}\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b||^2=n^{-1}\sigma^{-2}||\boldsymbol{R}_a^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2\approx n^{-1}\sigma^{-2}||diag(\boldsymbol{R}_a)^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2,\] where \(diag(\boldsymbol{R}_a)\) is the diagonal matrix constructed from the diagonal elements of \(\boldsymbol{R}_a\).
+
+The interpretation of this effect size is the impact of parameter \(a\) and \(b\) on the squared deviation of the standardized expected residual per observation.
+It is not directly related to the shape, or the pattern human observed from the residual plot, but it is a reasonable approximation of the degree of the visual deviation from the null residuals under our cubic model setting.
+Figure \ref{fig:effectsize} shows four residual plots with different effect sizes.
+As effect sizes increases, the strength of the signal become stronger.
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/effectsize-1.pdf}
+\caption{\label{fig:effectsize}Cubic model residual plots under different effect sizes. The larger the effect size, the stronger the signal.}
+\end{figure}
+
+Under the cubic model setting, there is an exact conventional test for testing the nonlinearity defect, which is F-test.
+For F-test, the null hypothesis is \(H_0:a=b=0\), and the alternative hypothesis is \(H_1:\text{at least one of them} \neq 0\).
+During the simulation of the lineup data, the F-statistic and the p-value will be recorded for comparison between the power of conventional test and visual test.
+
+\hypertarget{heteroskedasticity-model}{%
+\subsection{Heteroskedasticity Model}\label{heteroskedasticity-model}}
+
+Another model used in the experiments was a heteroskedasticity model with one regressor, which can be expressed by:
+\begin{equation} \label{eq:heter}
+Y_i = 1 + X_i + \varepsilon_i, ~i = 1,...,n,
+\end{equation}
+where \(a \in \{-1,0,1\}\), \(b\in (0,32)\) and \(\varepsilon_i \overset{iid}{\sim} N(0,1+b(2-|a|)(X_i-a)^2)\).
+
+To create heteroskedasticity defect, OLS was used to fit the null model:
+\begin{equation} \label{eq:heter_null}
+\boldsymbol{Y}=\beta_0+\beta_1\boldsymbol{X}+\boldsymbol{u},
+\end{equation}
+where \(\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2_u\boldsymbol{I})\).
+
+In this case, estimators of \(\beta_0\) and \(\beta_1\) are unbiased, but the error term has non-constant variance.
+
+\begin{lemma}[Distribution of residuals produced by the heteroskedasticity model] \label{lemma:heter}
+Given the data generating process in Equation (\ref{eq:heter}) and null model in Equation (\ref{eq:heter_null}).
+Let $\boldsymbol{X}_a=[\boldsymbol{1},\boldsymbol{X}]$ denotes the set of regressors in matrix form.
+The residuals obtained from the null model are $$\boldsymbol{e} \sim N(\boldsymbol{0}, \boldsymbol{R}_a\boldsymbol{V}),$$ where $\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ and $\boldsymbol{V}$ is a diagonal matrix with $V_{ii}=1+b(2-|a|)(X_i  -  a)^2$, $i = 1,...,n$.
+\end{lemma}
+
+\begin{proof}
+Using the Frisch–Waugh–Lovell theorem, the residuals obtained by the null model are $\boldsymbol{e}=\boldsymbol{R}_a\boldsymbol{Y}=\boldsymbol{R}_a(\boldsymbol{X}_a\boldsymbol{\beta}_a+\boldsymbol{\varepsilon}),$ where $\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ and $\boldsymbol{\beta}_a=(1,1)'$.
+
+Because $\boldsymbol{R}_a\boldsymbol{X}_a=\boldsymbol{0}$, we have $\boldsymbol{e}=\boldsymbol{R}_a\boldsymbol{\varepsilon}$.
+Hence, the residuals $\boldsymbol{e}$ follow $N(\boldsymbol{0}, \boldsymbol{R}_a\boldsymbol{V})$, where $\boldsymbol{V}$ is a diagonal matrix with $V_{ii}=1+b(2-|a|)(X_i  -  a)^2$, $i = 1,...,n$.
+\end{proof}
+
+Lemma \ref{lemma:heter} shows that variance and covariance of the residuals depend on \(\boldsymbol{X}\).
+We could plot the one standard deviation around the residuals to indicate the region about 68\% of the residuals should landed on.
+Figure \ref{fig:hetershape} illustrates different shapes of residuals undder various values of \(a\) and \(b\).
+From the plot, it can be observed that if \(a = 0\), the residual plot looks like the shape of butterfly.
+If \(a = \pm 1\), it looks like a triangle, and the sign of \(a\) determines the direction of the shape.
+Parameter \(b\) is controlling the strength of the signal.
+As \(b\) increases, the pattern becomes less noisy.
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/hetershape-1.pdf}
+\caption{\label{fig:hetershape}A matrix of residuals vs.~fitted values plot under different parameter settings of the heteroskedasticity model.
+Resdiauls are drawns in black and the one standard deviation around zero is drawn in red.
+The plots show that \(a\) controls the shape and direction of the resdiaul plot, while \(b\) controls the strength of the signal.}
+\end{figure}
+
+Similar to the cubic model, the heteroskedasticity model could also be built by the heteroskedasticity model class constructor \texttt{heter\_model()}.
+This function takes three arguments as inputs, which are \texttt{a}, \texttt{b} and \texttt{x}.
+\texttt{a} and \texttt{b} are numeric parameters defined in Equation (\ref{eq:heter}).
+\texttt{x} needs to be a random variable object.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{library}\NormalTok{(visage)}
+\NormalTok{mod }\OtherTok{\textless{}{-}} \FunctionTok{heter\_model}\NormalTok{(}\AttributeTok{a =} \DecValTok{0}\NormalTok{, }\AttributeTok{b =} \DecValTok{16}\NormalTok{, }\AttributeTok{x =} \FunctionTok{rand\_uniform}\NormalTok{(}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+Since both the cubic model class and the heteroskedasticity class are inherited from the visual inference model class, which has defined methods of simulating data, making residual plot and producing lineup, heteroskedasticity model object can be used in a similar way as cubic model object.
+The following codes give examples of the use of the object.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{mod}\SpecialCharTok{$}\FunctionTok{gen}\NormalTok{(}\DecValTok{5}\NormalTok{, }\AttributeTok{fit\_model =} \ConstantTok{TRUE}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+##          y          x    sigma         e     .resid  .fitted
+## 1 7.205499  0.9053470 5.218132 5.3001525  0.7420902 6.463409
+## 2 2.560696 -0.8626811 4.981465 2.4233771  1.0441078 1.516588
+## 3 1.121106 -0.6545505 3.835357 0.7756568 -0.9778170 2.098923
+## 4 5.296467  0.7171334 4.178154 3.5793335 -0.6403337 5.936801
+## 5 2.962130 -0.2859733 1.901837 2.2481030 -0.1680473 3.130177
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{mod}\SpecialCharTok{$}\FunctionTok{plot}\NormalTok{(mod}\SpecialCharTok{$}\FunctionTok{gen}\NormalTok{(}\DecValTok{100}\NormalTok{, }\AttributeTok{fit\_model =} \ConstantTok{TRUE}\NormalTok{), }\AttributeTok{size =} \DecValTok{1}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{thesis_files/figure-latex/unnamed-chunk-8-1.pdf}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{mod}\SpecialCharTok{$}\FunctionTok{plot\_lineup}\NormalTok{(mod}\SpecialCharTok{$}\FunctionTok{gen\_lineup}\NormalTok{(}\DecValTok{100}\NormalTok{), }\AttributeTok{size =} \DecValTok{1}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{thesis_files/figure-latex/unnamed-chunk-9-1.pdf}
+According to Lemma \ref{lemma:heter}, when \(b = 0\), the matrix \(\boldsymbol{V}\) collapses to an identity matrix.
+Assume the shape of butterfly and triangle have identical visual impacts, the only factor affects human to recognize the pattern is the strength of the signal, where parameter \(a\) has no role to play.
+In addition, sample size has a huge impact on the chance of human recognizing the pattern.
+As shown in Figure \ref{fig:hetern}, as sample size increases, the pattern becomes easier to observe.
+However, this effect is less noticeable with large sample size as the outline of the shape has been drawn and residuals have less probability to land outside of it.
+Thus, the effect size of this model can be expressed by \(ES = b\sqrt{n}\).
+The square root operator is used for addressing the large sample issue.
+Figure \ref{fig:hetereffectsize} shows the effectiveness of the effect size.
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/hetern-1.pdf}
+\caption{\label{fig:hetern}Residuals generated from the same heteroskedasticity model but with different sample size. As sample size increases, the shape of butterfly becomes more obvious.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/hetereffectsize-1.pdf}
+\caption{\label{fig:hetereffectsize}Residuals of heteroskedasticity model with different effect size. As effect size increases, the shape of butterfly becomes less noisy.}
+\end{figure}
+
+For the heteroskedasticity model, the conventional test we used was Breusch--Pagan test \autocite{breusch_simple_1979}, which tested whether the variance of the error terms of the regression is dependent on the regressors with the auxiliary regression equation \[\boldsymbol{e}^2 = \gamma_0 + \gamma_1 \boldsymbol{X} + \gamma_2 \boldsymbol{X}^2 + \boldsymbol{v}.\] \textcite{majumder_validation_2013} suggested that visual test is not expected to perform equally well as conventional test especially when there exists a exact conventional test.
+However, in contrast to the F-test used in the cubic model, Breusch--Pagan test is an approximate test.
+Thus, the power of visual test may exceed the power of Breusch--Pagan test.
+Throughout the study, Breusch--Pagan test statistic and p-value were recorded for comparison.
+
+\hypertarget{distribution-of-regressors}{%
+\subsection{Distribution of regressors}\label{distribution-of-regressors}}
+
+The model definitions given in the previous two sections does not include the specification of the regressors.
+In this sections, distribution of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) will be discussed.
+
+The cubic model involved the use of both \(\boldsymbol{X}\) and \(\boldsymbol{Z}\).
+In the simulation, \(X_i\), \(i=1,...,n\), had equal chance to follow one of the following distributions: \(N(0, 0.09)\), \(U(-1, 1)\), \(Lognormal(0,0.36)/3 - 1\) and \(-Lognormal(0,0.36)/3 + 1\).
+Uniform and normal distribution were symmetric and common.
+Adjusted lognormal distribution and adjusted negative lognormal distribution provided right-skewed and left-skewed density respectively.
+These distributions were chosen such that most the realizations will fall between \(-1\) and \(1\).
+
+The distribution of \(Z_i\), \(i=1,...,n\), had \(50\)\% chance to be a uniform distribution ranged from \(-1\) to \(1\), and \(50\)\% chance to be a discrete uniform distribution with \(z_n\) outcomes simulated from a uniform distribution ranged from \(-1\) to \(1\).
+\(z_n\) itself was a discrete uniform distribution with outcomes \(\{10,11,12,13,14,15,16,17,18,19,20\}\), which defined the number of possible values \(Z_i\) could take.
+As shown in Figure \ref{fig:discreteex}, this setup would create discreteness in residual plot, which could enrich the pool of visual patterns.
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/discreteex-1.pdf}
+\caption{\label{fig:discreteex}Discreteness in residuals created by using a discrete uniform random variable as one of the regressors in a cubic model.
+For each residual plot of the lineup, there are three clusters because the number of possible values the regressor can take is three.}
+\end{figure}
+
+\(\boldsymbol{X}\) used in the heteroskedasticity model was a combination of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) used in the cubic model.
+It could be one of the five distributions mentioned above - normal distribution, uniform distribution, adjusted lognormal distribution, adjusted negative lognormal distribution and discrete uniform distribution.
+
+\hypertarget{experiment-i-and-ii}{%
+\subsection{Experiment I and II}\label{experiment-i-and-ii}}
+
+In this study, multiple selection was allowed for the evaluation of a lineup.
+Given our desired significance level was \(\alpha = 0.05\) and the number of plots in a lineup was \(20\), a single visual test procedure needed to involve multiple evaluations.
+Otherwise, the \(p\)-value of the test would be greater than \(0.05\) in spite of the actual data plot being selected.
+The number of plots selected by a user depended on the user and the difficulty of the lineup.
+According to the pilot study evaluated by \(10\) faculty members of Department of Econometrics and Business Analytics, Monash University, the number of selections would be generally smaller than \(4\), which suggested a visual test consisted of \(5\) evaluations were sufficient to yield \(p\)-value smaller than \(0.05\) with at least three detections.
+Thus, each lineup was replicated for five times and evaluated by different participants.
+However, our study website was designed such that any set of lineups consisting of 20 lineups generated beforehand would be shown to only one person only once.
+And whether the participant would finish the experiment was unpredictable.
+Therefore, some lineups had insufficient number of evaluations, but some had more evaluations than expected.
+This slightly affected the estimate of the power of the visual test, which will be discussed in Section \ref{p-value-and-power-estimation-of-visual-test-allowed-for-multiple-selections}.
+
+We called the lineup being detected if the actual data plot was one of the selections of an evaluation of a lineup.
+The sample size calculation was based on the detection rate of the lineup, which was closely related to the difficulty level of the lineup.
+With the data collected from the pilot study, two logistic regressions given in Table \ref{tab:pilotglm} and Figure \ref{fig:pilotglm} were developed to describe the relationship between the natural logarithm of effect size and the detection rate for the cubic model and the heteroskedasticity model.
+
+\begin{table}[!htbp] \centering 
+  \caption{Logistic regressions with detection rate as response variable and natural logarithm of the effect size as regressor fitted on the pilot data for both cubic and heteroskedasticity models.} 
+  \label{tab:pilotglm} 
+\begin{tabular}{@{\extracolsep{5pt}}lcc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{2}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-3} 
+\\[-1.8ex] & \multicolumn{2}{c}{detect} \\ 
+ & Cubic & Heteroskedasticity \\ 
+\\[-1.8ex] & (1) & (2)\\ 
+\hline \\[-1.8ex] 
+ log(effect\_size) & 0.611$^{***}$ & 0.737$^{***}$ \\ 
+  & (0.098) & (0.137) \\ 
+  & & \\ 
+ Constant & 0.457$^{***}$ & $-$2.402$^{***}$ \\ 
+  & (0.165) & (0.645) \\ 
+  & & \\ 
+\hline \\[-1.8ex] 
+Observations & 246 & 289 \\ 
+Log Likelihood & $-$122.305 & $-$147.971 \\ 
+Akaike Inf. Crit. & 248.610 & 299.943 \\ 
 \hline 
-\RaggedRight\textbf{Thesis chapter}  & 
-\RaggedRight\textbf{Publication title}  & 
-\RaggedRight\textbf{Status (published, in press, accepted or returned for revision)}  &  
-\RaggedRight\textbf{Nature and \% of student contribution} & 
-\RaggedRight\textbf{Co-author name(s), nature and \% of co-author’s contribution} &  \RaggedRight\textbf{Co-author(s), Monash student Y/N} \\ 
-\hline
-2 & xx & xx  & xx & xx & \multicolumn{1}{c|}{N}   \\
-\hline
-3 & xx & xx  & xx & xx  & \multicolumn{1}{c|}{N}   \\
-\hline
-4 & xx & xx  & xx & xx & \multicolumn{1}{c|}{N}    \\
-\hline
-5 & xx & xx  & xx & xx & \multicolumn{1}{c|}{N}   \\
-\hline
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 \end{tabular} 
@@
 
-\textbf{Student name:} \authorname
-
-\textbf{Student signature:}
-
-\textbf{Date:}
-
-\hypertarget{acknowledgements}{%
-\chapter*{Acknowledgements}\label{acknowledgements}}
-\addcontentsline{toc}{chapter}{Acknowledgements}
-
-I would like to thank my pet goldfish for \dots
-
-\hypertarget{preface}{%
-\chapter*{Preface}\label{preface}}
-\addcontentsline{toc}{chapter}{Preface}
-
-The material in Chapter \ref{ch:intro} has been submitted to the journal \emph{Journal of Impossible Results} for possible publication.
-
-The contribution in Chapter \ref{ch:litreview} of this thesis was presented in the International Symposium on Nonsense held in Dublin, Ireland, in July 2015.
-
-\clearpage\pagenumbering{arabic}\setcounter{page}{0}
-
-\hypertarget{ch:intro}{%
-\chapter{Introduction}\label{ch:intro}}
-
-\hypertarget{is-what-we-see-really-there}{%
-\section{Is what we see really there?}\label{is-what-we-see-really-there}}
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/pilotglm-1.pdf}
+\caption{\label{fig:pilotglm}Logistic regressions with detection rate as response variable and natural logarithm of the effect size as regressor fitted on the pilot data for both cubic and heteroskedasticity models.
+Points are jittered to reduce overplotting using the \texttt{ggbeeswarm} package.
+The smooth curves are the fitted regression curves.
+The ribbons around the curves are the 95\% confidence intervals.}
+\end{figure}
+
+Lineups were classified into three categories - easy, medium and hard.
+Easy lineups were those of which predicted detection rates were higher than \(80\)\%, while medium lineups were between \(40\)\% to \(80\)\% and hard lineups were less than \(40\)\%.
+
+For the first experiment, we would like to have at least 20 detection with chance greater than 99.99\% in each category.
+Thus, we prepared \(68\) easy lineups, \(120\) medium lineups and \(388\) hard lineups with five replications.
+Every participant would get at least \(8\) easy or medium lineups.
+Two out of \(10\) extremely lineups with predicted detection rate over \(90\)\% were randomly given to every participant as attention checks.
+We initially planned to recruit \((68+120+388) \times 5 / (20-2) = 160\) participants but eventually decided to only recruit \(20\) participants because it was the first time we launched an experiment on Prolific where factors like payment method and quality of participants were unclear at the moment.
+As a consequence, most of the lineups had only two evaluations and not all the lineups were used.
+However, data collected from the first experiment was sufficient for building a more accurate logistic regression to predict detection rate for the second experiment.
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/e1glmplot-1.pdf}
+\caption{\label{fig:e1glmplot}Logistic regressions with detection rate as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I for both cubic and heteroskedasticity models.
+Points are jittered to reduce overplotting using the \texttt{ggbeeswarm} package.
+The smooth curves are the fitted regression curves.
+The ribbons around the curves are the 95\% confidence intervals.}
+\end{figure}
+
+\begin{table}[!htbp] \centering 
+  \caption{Logistic regressions with detection rate as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I for both cubic and heteroskedasticity models.}
+  \label{tab:e1glm} 
+\begin{tabular}{@{\extracolsep{5pt}}lcc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{2}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-3} 
+\\[-1.8ex] & \multicolumn{2}{c}{detect} \\ 
+ & Cubic & Heteroskedasticity \\ 
+\\[-1.8ex] & (1) & (2)\\ 
+\hline \\[-1.8ex] 
+ log(effect\_size) & 0.857$^{***}$ & 0.746$^{***}$ \\ 
+  & (0.122) & (0.111) \\ 
+  & & \\ 
+ Constant & $-$1.571$^{***}$ & $-$3.434$^{***}$ \\ 
+  & (0.257) & (0.462) \\ 
+  & & \\ 
+\hline \\[-1.8ex] 
+Observations & 192 & 188 \\ 
+Log Likelihood & $-$64.544 & $-$84.606 \\ 
+Akaike Inf. Crit. & 133.089 & 173.212 \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+\end{tabular} 
+\end{table}
+
+According to the new logistic regression given in Table \ref{tab:e1glm} and Figure \ref{fig:e1glmplot}, it was found that the first experiment might be too difficult for participants as the detection rate was lower than what we estimated from the pilot study, especially for the heteroskedasticity model.
+Thus, for the second experiments, we relaxed the sample size requirements and increased the proportion of easy lineups.
+In addition, due to budget constraint, the maximum number of participants was \(50\).
+With the new logistic regression fitted with data collected from the first experiment, \(28\) easy lineups, \(36\) medium lineups and \(100\) hard lineups were simulated for the second experiment.
+\(50\) participants were recruited but \(6\) were withdrew.
+
+\hypertarget{p-value-and-power-estimation-of-visual-test-allowed-for-multiple-selections}{%
+\subsection{\texorpdfstring{\(P\)-value and Power Estimation of Visual Test Allowed for Multiple Selections}{P-value and Power Estimation of Visual Test Allowed for Multiple Selections}}\label{p-value-and-power-estimation-of-visual-test-allowed-for-multiple-selections}}
+
+The \(p\)-value calculation for multiple selection is an extension of Equation (\ref{eq:pvaluesingle}).
+The assumptions about the independence still needed.
+Let \(K\) be the number of independent evaluations of a visual test, \(s_i\), \(i = 1,...,K\) be the number of selections of the evaluations, and \(X\) be the random variable denoting the number of detections.
+Then, the \(p\)-value of the visual test is given as:
+\begin{equation} \label{eq:pvaluemulti}
+P(X\geq x) = \sum_{j=x}^{K}Pr(j|s_1,...,s_K).
+\end{equation}
+The distribution of \(X\) given \(s_1,...s_K\) can not be derived trivially, as it is a sampling without replacement problem. In practice, this distribution can be approximated by computer simulation.
+
+Function \texttt{sim\_dist()} from package \texttt{visage} is designed to approximate the distribution of number of detections of a lineup via Monte Carlo method.
+It takes the number of evaluations \texttt{n\_eval}, the number of selections \texttt{n\_sel}, and the number of plots in a lineup \texttt{n\_plot} as inputs, then outputs a discrete distribution.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{sim\_dist}\NormalTok{(}\AttributeTok{n\_eval =} \DecValTok{3}\NormalTok{, }\AttributeTok{n\_sel =} \FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{2}\NormalTok{, }\DecValTok{3}\NormalTok{), }\AttributeTok{n\_plot =} \DecValTok{20}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+##       0       1       2       3 
+## 0.70182 0.25202 0.04346 0.00270
+\end{verbatim}
+
+Function \texttt{calc\_p\_value()} from the same package calculates the \(p\)-value using the distribution returned by \texttt{sim\_dist()}.
+It takes an additional argument \texttt{n\_detect}, which is the number of detections.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{calc\_p\_value}\NormalTok{(}\AttributeTok{n\_detect =} \DecValTok{2}\NormalTok{, }\AttributeTok{n\_eval =} \DecValTok{3}\NormalTok{, }\AttributeTok{n\_sel =} \FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{2}\NormalTok{, }\DecValTok{3}\NormalTok{), }\AttributeTok{n\_plot =} \DecValTok{20}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## [1] 0.04688
+\end{verbatim}
+
+The above example shows that a visual test with two detections and three independent evaluations in which three observers select two, two, and three plots out of \(20\) plots respectively yields a \(p\)-value smaller than \(0.05\).
+
+Assume there is a visual test \(V_K\), where \(K\) denoting the number of evaluations.
+The corresponding \(p\)-value of \(V_K\) can be computed by using Equation (\ref{eq:pvaluemulti}).
+Meanwhile, if one evaluation is randomly deleted from \(V_K\), the remaining evaluations can still be used to form another valid visual test \(V_{K-1}\).
+In fact, considering all the possibilities, \(K\) different outcomes for \(V_{K-1}\) can be obtained.
+Since all outcomes occur with equal probability, the proportion of outcomes which reject the null hypothesis can be used as an estimate of the power of the visual test \(V_{K-1}\).
+Similarly, if we would like to estimate the power of the visual test \(V_{K-j}\) given the evaluations of \(V_{K}\), for \(j < K\), we could find all the possible combinations of \(K\) elements, taken \(k-j\) at a time to obtain \(K\choose{K-j}\) different outcomes for \(V_{K-j}\).
+Then, the estimated power is given as \(R/{K\choose{K-j}}\), where \(R\) is the number of outcomes which reject the null hypothesis.
+
+Function \texttt{calc\_p\_value\_comb()} from the package \texttt{visage} can be used to compute \(p\)-values of \(V_{K-j}\).
+The first argument is \texttt{detected}, which needs to be a vector of Boolean values denoting whether the observer detects the actual data plot.
+Desired number of evaluations and number of selections need to be provided via argument \texttt{n\_eval} and \texttt{n\_sel}.
+
+For example, given a visual test \(V_3\) where the first observer selects one plot and gets correct, the second observer selects one plot but misses, and the third observer selects two plots and gets correct.
+The \(p\)-value of all possible outcomes of \(V_2\) can be computed using the following code:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{calc\_p\_value\_comb}\NormalTok{(}\AttributeTok{detected =} \FunctionTok{c}\NormalTok{(}\ConstantTok{TRUE}\NormalTok{, }\ConstantTok{FALSE}\NormalTok{, }\ConstantTok{TRUE}\NormalTok{), }
+                  \AttributeTok{n\_eval =} \DecValTok{2}\NormalTok{, }
+                  \AttributeTok{n\_sel =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## [1] 0.09352 0.00624 0.14454
+## attr(,"combinations")
+##      [,1] [,2] [,3]
+## [1,]    1    1    2
+## [2,]    2    3    3
+\end{verbatim}
+
+\texttt{calc\_p\_value\_comb()} returns a vector of \(p\)-value along with an attribute which is a matrix representing elements being used by the specific outcome.
+Elements in the first column indicates that the first observer and the second observer are involved in the visual test, and the corresponding \(p\)-value is \(0.09352\).
+
+\hypertarget{results}{%
+\subsection{Results}\label{results}}
+
+We collected \(400\) lineup evaluations made by \(20\) participants in experiment I and \(880\) lineup evaluations made by \(44\) participants in experiment II.
+In total, \(442\) unique lineups were evaluated by \(64\) subjects.
+In experiment I, one of the participants skipped all \(20\) lineups.
+Hence, the submission was rejected and removed from the dataset.
+In experiment II, there was a participant failed one of the two attention checks, but there was no further evidence of low-effort throughout the experiment.
+Therefore, the submission was kept.
+
+\hypertarget{power-estimation}{%
+\subsubsection{Power Estimation}\label{power-estimation}}
+
+It was expected that with larger effect size, the power of the visual test would be greater for both cubic model and heteroskedasticity model.
+Using the power calculation method discussed in Section \ref{p-value-and-power-estimation-of-visual-test-allowed-for-multiple-selections}, power of each lineup from one to five evaluations was estimated.
+The estimated power was further used in fitting a quasi-binomial generalized linear model with natural logarithm of effect size as the only regressor.
+The result of the fitted models are given in Table \ref{tab:powerglmcubic}, Table \ref{tab:powerglmheter}, and Figure \ref{fig:powerglm}.
+
+As expected, coefficients of natural logarithm of effect size of all five models were positive.
+From Figure \ref{fig:powerglm}, it can be observed that the fitted power of visual test increased as the number of evaluations increased for both cubic and heteroskedasticity model.
+
+For heteroskedasticity model, this phenomenon was more obvious as the curves of visual tests with evaluations greater than two were always above the curves of visual test with evaluations smaller than two.
+However, this only held for large enough effect size.
+For small effect size, visual tests with fewer evaluations might have greater power.
+Note that, the expected power of visual test derived by \textcite{majumder_validation_2013} followed by the assumption that human has the ability to select the plot with the highest t-statistic from a lineup under the classical linear model regression setting showed similar properties, where visual tests with fewer evaluations were expected to perform better when the parameter values close to the null hypothesis.
+
+For cubic model, the separation between curves was small.
+Fitted power of visual test with three to five evaluations were almost identical to each other in regards of effect size.
+In addition, all five curves peaked at one as effect size increased, suggesting that identification of nonlinearity as a visual task can be completed reliably by human when the effect size is large enough.
+
+As shown in Figure \ref{fig:powerglm}, both F-test and Breusch--Pagan test generally possessed greater power than visual test.
+Since the settings we were using were not ideal scenarios for visual inference, it was not expected that the visual test would be as good as the conventional test.
+Our goal was to quantify the ability of human reading residual plot, such that the comparison between human and computer vision model can be made.
+
+\begin{table}[!htbp] \centering 
+  \caption{Quasi-binomial logistic regressions with estimated power as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for cubic model.
+Five regressions were fitted for different number of evaluations.}
+  \label{tab:powerglmcubic} 
+\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{5}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-6} 
+\\[-1.8ex] & \multicolumn{5}{c}{hat\_power} \\ 
+\\[-1.8ex] & (1) & (2) & (3) & (4) & (5)\\ 
+\hline \\[-1.8ex] 
+ log(effect\_size) & 0.929$^{***}$ & 1.227$^{***}$ & 1.084$^{***}$ & 1.057$^{***}$ & 1.035$^{***}$ \\ 
+  & (0.049) & (0.066) & (0.074) & (0.082) & (0.087) \\ 
+  & & & & & \\ 
+ Constant & $-$1.969$^{***}$ & $-$2.603$^{***}$ & $-$1.570$^{***}$ & $-$1.501$^{***}$ & $-$1.432$^{***}$ \\ 
+  & (0.114) & (0.153) & (0.142) & (0.156) & (0.160) \\ 
+  & & & & & \\ 
+\hline \\[-1.8ex] 
+Observations & 630 & 532 & 456 & 423 & 371 \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{5}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+\end{tabular} 
+\end{table}
+
+\begin{table}[!htbp] \centering 
+  \caption{Quasi-binomial logistic regressions with estimated power as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for heteroskedasticity model.
+Five regression were fitted for different number of evaluations.}
+  \label{tab:powerglmheter} 
+\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{5}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-6} 
+\\[-1.8ex] & \multicolumn{5}{c}{hat\_power} \\ 
+\\[-1.8ex] & (1) & (2) & (3) & (4) & (5)\\ 
+\hline \\[-1.8ex] 
+ log(effect\_size) & 0.821$^{***}$ & 1.025$^{***}$ & 0.885$^{***}$ & 0.920$^{***}$ & 1.063$^{***}$ \\ 
+  & (0.064) & (0.087) & (0.092) & (0.102) & (0.125) \\ 
+  & & & & & \\ 
+ Constant & $-$4.472$^{***}$ & $-$5.610$^{***}$ & $-$4.285$^{***}$ & $-$4.127$^{***}$ & $-$4.717$^{***}$ \\ 
+  & (0.320) & (0.449) & (0.461) & (0.499) & (0.607) \\ 
+  & & & & & \\ 
+\hline \\[-1.8ex] 
+Observations & 630 & 527 & 473 & 443 & 375 \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{5}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+\end{tabular} 
+\end{table}
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/powerglm-1.pdf}
+\caption{\label{fig:powerglm}Quasi-binomial logistic regressions with estimated power as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for both cubic and heteroskedasticity model.
+Five regression were fitted for different number of evaluations shown in five different colours.
+Estimated power are drawn in black.
+The dashed curvses are the power of F-test and Breusch--Pagan test respectively.}
+\end{figure}
+
+Another method we used to model the power was to collect all decisions made by \(K\choose{K-j}\) different outcomes for \(V_{K-j}\), and directly fit a binomial generalized linear model with natural logarithm of effect size as the regressor.
+This method skipped the calculation of the estimated power \(R/{K\choose{K-j}}\), and regressed on the binary outcome with larger sample size.
+Results of the fitted model is given in Figure \ref{fig:powerglmsecond}, Table \ref{tab:powerglmhetersecond} and Table \ref{tab:powerglmcubicsecond}.
+According to Figure \ref{fig:powerglmsecond}, this method gave clearer separation between curves for heteroskedasticity model.
+Noticeably, fitted power of visual test with five evaluations matched the power of conventional test at large effect size.
+
+\begin{table}[!htbp] \centering 
+  \caption{Logistic regressions with rejection as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for cubic model.
+Five regression were fitted for different number of evaluations.}
+  \label{tab:powerglmcubicsecond} 
+\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{5}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-6} 
+\\[-1.8ex] & \multicolumn{5}{c}{hat\_power} \\ 
+\\[-1.8ex] & (1) & (2) & (3) & (4) & (5)\\ 
+\hline \\[-1.8ex] 
+ log(effect\_size) & 0.872$^{***}$ & 1.292$^{***}$ & 1.125$^{***}$ & 1.164$^{***}$ & 1.315$^{***}$ \\ 
+  & (0.031) & (0.032) & (0.021) & (0.021) & (0.029) \\ 
+  & & & & & \\ 
+ Constant & $-$1.836$^{***}$ & $-$2.529$^{***}$ & $-$1.516$^{***}$ & $-$1.321$^{***}$ & $-$1.419$^{***}$ \\ 
+  & (0.077) & (0.073) & (0.046) & (0.049) & (0.072) \\ 
+  & & & & & \\ 
+\hline \\[-1.8ex] 
+Observations & 2,900 & 7,250 & 12,596 & 17,011 & 19,229 \\ 
+Log Likelihood & $-$752.896 & $-$1,312.436 & $-$2,977.445 & $-$2,651.299 & $-$1,463.555 \\ 
+Akaike Inf. Crit. & 1,509.792 & 2,628.873 & 5,958.889 & 5,306.598 & 2,931.109 \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{5}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+\end{tabular} 
+\end{table}
+
+\begin{table}[!htbp] \centering 
+  \caption{Logistic regressions with rejection as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for heteroskedasticity model.
+Five regression were fitted for different number of evaluations.}
+  \label{tab:powerglmhetersecond} 
+\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{5}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-6} 
+\\[-1.8ex] & \multicolumn{5}{c}{hat\_power} \\ 
+\\[-1.8ex] & (1) & (2) & (3) & (4) & (5)\\ 
+\hline \\[-1.8ex] 
+ log(effect\_size) & 0.912$^{***}$ & 1.221$^{***}$ & 1.294$^{***}$ & 1.525$^{***}$ & 1.798$^{***}$ \\ 
+  & (0.044) & (0.033) & (0.027) & (0.030) & (0.042) \\ 
+  & & & & & \\ 
+ Constant & $-$4.846$^{***}$ & $-$6.396$^{***}$ & $-$6.033$^{***}$ & $-$6.771$^{***}$ & $-$7.784$^{***}$ \\ 
+  & (0.229) & (0.175) & (0.141) & (0.159) & (0.230) \\ 
+  & & & & & \\ 
+\hline \\[-1.8ex] 
+Observations & 2,906 & 7,064 & 11,436 & 13,728 & 13,236 \\ 
+Log Likelihood & $-$1,525.398 & $-$3,668.325 & $-$5,491.915 & $-$4,828.896 & $-$2,858.774 \\ 
+Akaike Inf. Crit. & 3,054.796 & 7,340.650 & 10,987.830 & 9,661.793 & 5,721.548 \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{5}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+\end{tabular} 
+\end{table}
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/powerglmsecond-1.pdf}
+\caption{\label{fig:powerglmsecond}Logistic regressions with rejection as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for both cubic and heteroskedasticity model.
+Five regression were fitted for different number of evaluations shown in five different colours.
+Estimated power are drawn in black.
+The dashed curvses are the power of F-test and Breusch--Pagan test respectively.}
+\end{figure}
+
+\hypertarget{effect-of-factors-on-power-of-the-visual-test}{%
+\subsubsection{Effect of Factors on Power of the Visual Test}\label{effect-of-factors-on-power-of-the-visual-test}}
+
+The previous section focuses on the change of effect size relative to the power of the visual test.
+Individual factor embedded in the data simulation also needs be analysed.
+Two logistic regressions were fitted with decisions made by \(V_{K-j}\) as response variable.
+For the cubic model, regressors were \(|a|\), \(|b|\), \(|c-1|\), \(\sigma\), distribution of \(\boldsymbol{X}\), distribution of \(\boldsymbol{Z}\), number of observations and number of evaluations.
+For the heteroskedasticity model, regressors were \(|a|\), \(b\), distribution of \(\boldsymbol{X}\), number of observations and number of evaluations.
+
+According to Table \ref{tab:rejectfactorcubic}, for cubic model, both coefficients of \(|a|\) and \(|b|\) were positive, which suggested increasing the coefficients of quadric and cubic terms would result in higher visual power.
+The coefficient of \(|c-1|\) was also positive indicating joint effect of multiple regressors on the residual plot decreased the visual power.
+As expected, the coefficient of \(\sigma\) was negative as it controlled the noise level.
+Lineup simulated with uniform distribution of \(\boldsymbol{X}\) had the highest power, followed by negative lognormal, normal, and lognormal.
+Discreteness in \(\boldsymbol{Z}\) decreased the visual power.
+The coefficient of number of observations was positive, which means larger sample size would lead to higher visual power.
+
+Results of logistic regression for heteroskedasticity is given in Table \ref{tab:rejectfactorheter}.
+For heteroskedasticity model, the coefficient of \(|a|\) was negative, which suggested ``butter fly'' shape was easier to identify than ``triangle'' shape.
+As expected, the coefficient of \(b\) was positive.
+For different distributions of \(\boldsymbol{X}\), the one with highest power was uniform distribution, followed by normal, discrete uniform, lognormal and negative lognormal.
+The coeffient of the number of observations was positive, but it was not significant.
+
+In overall, we did not have any unexpected findings except the visual power related to different distributions of \(\boldsymbol{X}\) in cubic model.
+We expected lineup simulated with uniform distribution would have the highest power, followed by normal, lognormal and negative lognormal, since we observed similar results in pilot study.
+
+\begin{table}[!htbp] \centering 
+  \caption{Logistic regressions with rejection as response variable and parameter settings as regressors fitted on the data collected from experiment I and II for cubic model.} 
+  \label{tab:rejectfactorcubic} 
+\begin{tabular}{@{\extracolsep{5pt}}lc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-2} 
+\\[-1.8ex] & reject \\ 
+\hline \\[-1.8ex] 
+ |a| & 1.064$^{***}$ \\ 
+  & (0.031) \\ 
+  & \\ 
+ |b| & 0.718$^{***}$ \\ 
+  & (0.029) \\ 
+  & \\ 
+ |c-1| & 5.443$^{***}$ \\ 
+  & (0.119) \\ 
+  & \\ 
+ sigma & $-$3.536$^{***}$ \\ 
+  & (0.050) \\ 
+  & \\ 
+ Distribution of X: Negative Lognormal & 0.971$^{***}$ \\ 
+  & (0.073) \\ 
+  & \\ 
+ Distribution of X: Normal & 0.754$^{***}$ \\ 
+  & (0.086) \\ 
+  & \\ 
+ Distribution of X: Uniform & 1.751$^{***}$ \\ 
+  & (0.085) \\ 
+  & \\ 
+ Distribution of Z: Discrete & $-$2.986$^{***}$ \\ 
+  & (0.056) \\ 
+  & \\ 
+ n & 0.005$^{***}$ \\ 
+  & (0.0002) \\ 
+  & \\ 
+ Number of evaluations & 0.291$^{***}$ \\ 
+  & (0.017) \\ 
+  & \\ 
+ Constant & $-$2.453$^{***}$ \\ 
+  & (0.173) \\ 
+  & \\ 
+\hline \\[-1.8ex] 
+Observations & 58,986 \\ 
+Log Likelihood & $-$8,034.351 \\ 
+Akaike Inf. Crit. & 16,090.700 \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+\end{tabular} 
+\end{table}
+
+\begin{table}[!htbp] \centering 
+  \caption{Logistic regressions with rejection as response variable and parameter settings as regressors fitted on the data collected from experiment I and II for heteroskedasticity model.} 
+  \label{tab:rejectfactorheter} 
+\begin{tabular}{@{\extracolsep{5pt}}lc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-2} 
+\\[-1.8ex] & reject \\ 
+\hline \\[-1.8ex] 
+ |a| & $-$0.526$^{***}$ \\ 
+  & (0.037) \\ 
+  & \\ 
+ b & 0.064$^{***}$ \\ 
+  & (0.002) \\ 
+  & \\ 
+ Distribution of X: Lognormal & $-$0.534$^{***}$ \\ 
+  & (0.058) \\ 
+  & \\ 
+ Distribution of X: Negative Lognormal & $-$0.610$^{***}$ \\ 
+  & (0.062) \\ 
+  & \\ 
+ Distribution of X: Normal & 0.187$^{***}$ \\ 
+  & (0.063) \\ 
+  & \\ 
+ Distribution of X: Uniform & 2.980$^{***}$ \\ 
+  & (0.064) \\ 
+  & \\ 
+ n & 0.0003 \\ 
+  & (0.0002) \\ 
+  & \\ 
+ Number of evaluations & 0.496$^{***}$ \\ 
+  & (0.014) \\ 
+  & \\ 
+ Constant & $-$2.711$^{***}$ \\ 
+  & (0.062) \\ 
+  & \\ 
+\hline \\[-1.8ex] 
+Observations & 48,370 \\ 
+Log Likelihood & $-$13,026.190 \\ 
+Akaike Inf. Crit. & 26,070.390 \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+\end{tabular} 
+\end{table}
+
+\hypertarget{automatic-visual-inference-system}{%
+\section{Automatic Visual Inference System}\label{automatic-visual-inference-system}}
+
+Visual tests can be used to make valid inference for visual discoveries while acknowledging people with confidence about the findings.
+However, the lack of scalability of this method limits its application in wider fields.
+The need of automatic visual inference system is auto data analysis and machine learning system.
+
+AI agents which can read data plot and evaluate lineups could
+
+Considering there are different approches of AI discussed in \ref{se:ai}, it is necessary to clarify the type of AI we aimed to build.
+Understanding the human thought process or mimicking the human vision mechanism were not of interest in this research.
+We believed humanly behaviours was enough for demonstrating AI for the problem we concerned.
+Hence, two of the capacities required by the total Turing test, namely machine learning and computer vision will be adopted.
+
+Another possible method was to define distance metrics to measure difference between data plots for making mathematically prefect decisions.
+This method would work if such metrics can be obtained manually or approximated algorithmically.
+However, we considered it was too complex as a starting point.
+
+\hypertarget{automatic-visual-inference-system-1}{%
+\subsection{Automatic visual inference system}\label{automatic-visual-inference-system-1}}
+
+For an automatic visual inference system, types of inputs can be classified by two decisions: (i) whether to accept multiple image files as inputs, and (ii) whether to accept fitted model.
+
+single image file
+multiple image files
+
+fitted model
+
+Depend on the inputs and outputs, computer vision model can be integrated into an automatic visual inference system.
 
 \hypertarget{se:ai}{%
-\section{Artificial intelligence: Four Approaches}\label{se:ai}}
-
-\textbf{Artificial intelligence} (AI) is the field of research concerned with understanding and building machines who can demonstrate intelligence.
+\subsection{Artificial Intelligence: Four Approaches}\label{se:ai}}
+
+Since the automatic visual inference system is intended to play the role of evaluators as humans in visual inference, some degree of intelligence needs to be demonstrated.
+This brings us to one of the popular filed of study - \textbf{Artificial Intelligence} (AI).
+AI is the field of research concerned with understanding and building machines who can demonstrate intelligence.
 As discussed in \textcite{russell_artificial_2002}, historically, there are disagreements among researchers about the definition of intelligence, which is caused by two critical questions:
@@
 
-With the development of AI, mature AI technologies, such as facial recognition and web recommendation system, have profoundly affected the way modern society operates and citizen's daily life.
-This is largely as a consequence of the huge investment in AI industry by the financial market in recent years.
-Further, the increasingly cheap computing cost and the massive amount of accessible e-commerce data produced in the Internet age provide the possibility for applying data-intensive AI models, which enables AI performance to reach new heights in history \autocite{jordan_machine_2015}.
-Some AI systems have already been remarkably better than humans in certain areas, e.g., game playing.
-AlphaGo and AlphaZero developed by the Google DeepMind team surpass professional human Go players \autocite{silver_general_2018}.
-
-\hypertarget{predictive-modelling-and-visual-diagnostics}{%
-\section{Predictive Modelling and Visual Diagnostics}\label{predictive-modelling-and-visual-diagnostics}}
-
-Behind the success of AI, a great propotion of AI systems rely on the predictive modelling framework.
-\textcite{donoho_50_2017} in its summary of data science stated that the concept of this modelling culture could be traced back to an article written by Breiman (2001).
-In contrast to the generative modelling culture, which aims to develop stochastic models to make inferences about the data generating process, predictive modelling emphasizes the ability of the model to make accurate predictions.
-Most AI tasks are complex prediction problems where the data mechanism is mysterious, or at least, partly unknowable.
-\textcite{breiman_statistical_2001} suggests that generative models are obviously not applicable in these scenarios, while the predictive modelling seeks only an accurate approximated function \(\boldsymbol{f}(\boldsymbol{x})\) to describe the relationship between the features \(\boldsymbol{x}\) and the responses \(\boldsymbol{y}\).
-
-Predictive models are primarily evaluated by predictive accuracy with the use of validation and test data, but in predictive model diagnostics, especially model testing and tuning, data plots play an irreplaceable role.
-In these diagnostics, though numeric summaries are mostly available and some are even endorsed by finite or asymptotic properties, graphical representation of data is still preferred, or at least needed by researchers, due to its intuitiveness and the possibility for unexpected discoveries which may be abstract and unquantifiable.
-
-However, unlike confirmatory data analysis built upon rigorous statistical procedures, e.g., hypothesis testing, visual diagnostics relies on graphical perception - human's ability to interpret and decode the information embedded in the graph \autocite{cleveland_graphical_1984}, which is to some extent subjective.
-Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
-One such example is finding an over-interpretation of the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurence \autocite{roy_chowdhury_using_2015}.
-
-\hypertarget{visual-inference}{%
-\section{Visual Inference}\label{visual-inference}}
-
-Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title ``Inference for Data Visualizatino'' by \textcite{buja_inference_1999} as an idea to address the issue of valid inference for visual discoveries of data plots \autocite{gelman_exploratory_2004}.
-Later, in the Bayesian context, data plots was systematically considered as model diagnostics by taking advantage of the data simulated from the assumed statistical models \autocite{gelman_bayesian_2003,gelman_exploratory_2004}.
-
-It was surprising that the essential components of visual inference had actually been established in \textcite{buja_inference_1999}, but it was not until 10 years later that \textcite{buja_statistical_2009} formalized it as an inferential framework to extend confirmatory statistics to visual discoveries.
-This framework redefines the test statistics, tests, null distribution, significance levels and \(p\)-value for visual discovery modelled on the confirmatory statistical testing.
-Figure \ref{fig:parallelism} outlines the parallelism between conventional tests and visual discovery.
-
-\begin{figure}
-\centering
-\includegraphics[width=4.6875in,height=3.55208in]{figures/rsta2009012001.jpg}
-\caption{Parallelism between multiple quantitative testing and visual discovery \autocite{buja_statistical_2009}.
-Visible features in a plot are viewed as a collection of test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\), and any visual discoveries that are inconsistent with the null hypothesis are treated as evidence against the null.
-For model diagnostics, the null hypothesis would be the assumed model, and visual discoveries could be any visual features in favour of any alternatives.
-\label{fig:parallelism}}
-\end{figure}
-
-In visual inference, a visual discovery is defined as a rejection of a null hypothesis, and the same null hypothesis can be rejected by many different visual discoveries \autocite{buja_statistical_2009}.
-For model diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the null hypothesis.
-The same assumed model, such as classical linear regression model, can be rejected by many reasons with residual plot, including nonlinearity and heteroskedasticity as shown in Figure \ref{fig:residual-plot-cubic-heter}.
-
-\begin{figure}
-\centering
-\includegraphics{thesis_files/figure-latex/residual-plot-cubic-heter-1.pdf}
-\caption{\label{fig:residual-plot-cubic-heter}Residuals vs.~fitted values plot for a classical linear regression model.
-The residuals are produced by fitting a two-predictor multiple linear regression model with data generated from a cubic linear model.
-From the residual plot, ``butterfly shape'' can be observed which generally would be interpretd as evidence of heteroskedasticity.
-Further, from the outline of the shape, nonlinear patterns exist.
-Both visual discoveries are evidence against the null hypothesis, though heteroskedasticity actually does not exist in the data generating process.}
-\end{figure}
-
-\hypertarget{pre-specification-of-visual-discoverable-features}{%
-\section{Pre-specification of Visual Discoverable Features}\label{pre-specification-of-visual-discoverable-features}}
-
-As discussed in \textcite{buja_statistical_2009}, in the practice of model diagnostics, the range of possible visual discoveries is not pre-specified.
-In other words, people do not explicitly specify which visual feature(s) they are looking for before the reading the diagnostic plot.
-In contrary, conventional hypothesis testing always requires the pre-specification of the parameter space \(\Theta\) of the parameter of interest \(\theta \in \Theta\) to form a valid inferential procedure.
-To address this issue, a collection of test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) is defined, where \(\boldsymbol{\mathrm{y}}\) is the data and \(I\) is a set of all possible visual features.
-\textcite{buja_statistical_2009} described each of the test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})\) as a measurement of the degree of presence of a visual feature.
-Alternatively, \textcite{majumder_validation_2013} avoids the use of visual features and defined the visual statistics \(T(.)\) as a mapping from a dataset to a data plot.
-Both definitions of visual test statistics are valid, but in the rest of the paper the first definition will be used as it covers some details needed by the following discussion.
-
-The size of the collection \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) depends on the size of the set \(I\).
-Thus, if one can define \(I\) comprehensively, i.e, pre-specify all the visual discoverable features, the validity issue will be solved.
-Unfortunately, to our knowledge, there is no such a way to list all visual features.
-In linear regression diagnostics, possible visual features of a residual plot may be outliers, shapes and clusters.
-But this is an incomplete list which does not enumerate all the visual features.
-
-Similarly, \textcite{wilkinson_graph-theoretic_2005} proposed the work called graph theoretic scagnostics, which adopted the idea of ``scagnostics'' - scatter plot diagnostics from (can't find the 1984 citation).
-It includes 9 computable scagnostics measures defined on planar proximity graphs: ``Outlying'', ``Convex'', ``Skinny'', ``Stringy'', ``Straight'', ``Monotonic'', ``Skewed'', ``Clumpy'' and ``Striated'' which attempts to describe outliers, shape, density, trend and coherence of the data.
-This approach is inspiring but it still does not give the complete list of visual discoverable features.
-In fact, it is possible that such a list will never be complete as suggested in \textcite{buja_statistical_2009}.
-
-Thinking out of the box, \textcite{buja_statistical_2009} argued that there is actually no need for pre-specification of visual discoverable features.
-In model diagnostics, when the null hypothesis is rejected, the reasons for rejecting the hypothesis will also be known.
-This is because observers can not only point out the fact that visual discoveries have been found, but also describe the particular visual features they observed.
-Those features will correspond to the subset of the collection of visual test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) which resulted in rejection.
-This argument helps justifies the validity of visual inference.
-
-\hypertarget{sampling-from-the-null-distribution}{%
-\section{Sampling from the Null Distribution}\label{sampling-from-the-null-distribution}}
-
-In visual inference, the null distribution of plots refers to the infinite collection of plots of null datasets sampled from \(H_0\).
-It is defined as the analogue of the null distribution of test statistics in conventional test \autocite{buja_statistical_2009}.
-In practice, a finite number of plots of null datasets could be generated, called null plots.
-
-In the context of model diagnostics, sampling data from \(H_0\) is equivalent to sampling data from the assumed model.
-As \textcite{buja_statistical_2009} suggested, \(H_0\) is usually composited by a collection of distributions controlled by nuisance parameters.
-Since statistical models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to so called ``reference distribution'' by applying one of the three methods: (i) sampling from a conditional distribution given a minimimal sufficient statistic under \(H_0\), (ii) parametric bootstrap sampling with nuisance parameters estimated under \(H_0\), and (iii) Bayesian posterior predictive sampling.
-
-The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three \autocite{buja_statistical_2009}.
-Focus will be put on simulation based on this distribution as it is applicable to classical normal linear regression model.
-Details of the other two are can be found in the supplementary material of \textcite{buja_statistical_2009}.
-Suppose there exists a minimal sufficient statistic \(\boldsymbol{S}(\boldsymbol{y})\) under the null hypothesis, any null datasets \(\boldsymbol{y^{*}}\) should fulfil the condition \(\boldsymbol{S}(\boldsymbol{y}) = \boldsymbol{s}\).
-For classical normal linear regression model, the minimal sufficient statistic is \(\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{\beta}}, \boldsymbol{e}'\boldsymbol{e})\), where \(\hat{\boldsymbol{\beta}}\) are the coefficient estimators and \(\boldsymbol{e}'\boldsymbol{e}\) is the residual sum of square.
-Alternatively, the minimal sufficient statistic can be constructed as \(\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{y}}, ||\boldsymbol{e}||)\), where \(\hat{\boldsymbol{y}}\) are the fitted values and \(||\boldsymbol{e}||\) is the length of residuals, which is more intuitive as suggested by \textcite{buja_statistical_2009}.
-Since the fitted values are held fixed, the variation can only occur in the residual space.
-And because the length of residual is also held fixed, residuals obtained from a null dataset has to be a random rotation of \(\boldsymbol{e}\) in the residual space.
-With this property, null residuals can be simulated by regressing \(N\) i.i.d standard normal random draws on the regressors, then rescaling it by the ratio of residual sum of square in two regressions.
-
-\hypertarget{lineup-protocol}{%
-\section{Lineup Protocol}\label{lineup-protocol}}
-
-With the validity of visual inference being justified and the simulation of null plots being provided, another aspect of hypothesis testing that needs to be addressed is the control of false positive rate or Type I error.
-Any visual statistic \(T^{(i)}(\boldsymbol{\mathrm{y}})\) needs to pair with a critical value \(c^{(i)}\) to form a hypothesis test.
-When a visual feature \(i\) is discovered by the observer from a plot, the corresponding visual statistic \(T^{(i)}(\boldsymbol{\mathrm{y}})\) may not be known as there is no general agreement on the measurement of the degree of presence of a visual feature.
-It is only the event that \(T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}\) is confirmed.
-Similarly, if any visual discovery is found by the observer, we say, there exists \(i \in I:~T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}\) \autocite{buja_statistical_2009}.
-
-Using the above definition, the family-wise Type I error can be controlled if one can provide the collection of critical values \(c^{(i)}~(i \in I)\) such that \(P(\mathrm{there~exists~} i \in I: T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}|\boldsymbol{\mathrm{y}}) \leq \alpha\), where \(\alpha\) is the significance level.
-However, since the quantity of \(T^{(i)}(\boldsymbol{\mathrm{y}})\) may not be known, such collection of critical values can not be provided.
-
-\textcite{buja_statistical_2009} proposed the lineup protocol as a visual test to calibrate the Type I error issue without the specification of \(c^{(i)}~(i \in I)\).
-It is inspired by the ``police lineup'' or ``identity parade'' which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people.
-The protocol consists of \(m\) randomly placed plots, where one plot is the actual data plot, and the remaining \(m - 1\) plots have the identical graphical production as the data plot except the data has been replaced with data consistent with the null hypothesis.
-Then, an observer who have not seen the actual data plot will be asked to point out the most different plot from the lineup.
-
-Under the null hypothesis, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is \(1/m\).
-If we reject the null hypothesis as the observer correctly picks the actual data plot, then the Type I error of this test is \(1/m\).
-
-This provides us with an mechanism to control the Type I error, because \(m\) - the number of plots in a lineup can be chosen.
-A larger value of \(m\) will result in a smaller Type I error, but the limit to the value of \(m\) depends on the number of plots a human is willing to view \autocite{buja_statistical_2009}.
-Typically, \(m\) will be set to \(20\) which is equivalent to set \(\alpha = 0.05\), a general choice of significance level for conventional testing among statisticians.
-
-Further, if we involve \(K\) independent observers in a visual test, and let \(X\) be a random variable denoting the number of observers correctly picking the actual data plot.
-Then, under the null hypothesis \(X \sim \mathrm{Binom}_{K,1/m}\), and therefore, the \(p\)-value of a lineup of size \(m\) evaluated by \(K\) observer is given as
-\begin{equation} \label{eq:pvaluesingle}
-P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},
-\end{equation}
-
-where \(x\) is the realization of number of observers correctly picking the actual data plot \autocite{majumder_validation_2013}.
-
-The multiple individuals approach avoids the limit of \(m\), while provides visual tests with \(p\)-value much smaller than \(0.05\).
-In fact, the lower bound of \(p\)-value decreases exponentially as \(K\) increases.
-With just \(4\) individuals and \(20\) data plots in a lineup, the \(p\)-value could be as small as \(0.0001\).
-Additionally, by involving multiple observers, variation of individual ability to read plots can be addressed to some degree as different opinions about visual discoveries can be collected.
-
-Compared to the conventional test, whose power only depends on the parameter of interest \(\theta\), several studies \autocites[see][]{hofmann_graphical_2012,majumder_validation_2013,majumder_human_2014,roy_chowdhury_using_2015,loy_variations_2016} have shown the power of the visual test is subject-specific.
-Thus, to be able to account for individual's ability, an individual is required to evaluate multiple lineups \autocite{majumder_validation_2013}.
-
-Suppose individuals have the same ability and a lineup has been evaluated by multiple individuals, under the alternative hypothesis, the estimated power for a lineup can be expressed as \(\hat{p} = x/K\), the estimated probability of identifying the actual data plot from the lineup.
-If the individual skill needs to be taken into account, and \(L\) lineups have been evaluated by \(K\) individuals, \textcite{majumder_validation_2013} suggests that mixed effects logistic regression model can be fit as:
-
-\[g(p_{li}) = W_{li}\delta + Z_{li}\tau_{li},\]
-where \(g(.)\) is the logit link function \(g(p) = log(p) - log(1-p)\); \(0 \leq p \leq 1\).
-\(W_{li}\), \(1 \leq i \leq K\), \(1 \leq l \leq L\), is the covariate matrix including lineup-specific elements and demographic information of individuals, and \(\delta\) is a vector of parameters.
-\(Z\) is the random effects matrix, and \(\tau\) is a vector of variables follow \(N(\boldsymbol{0},\sigma_{\tau}\boldsymbol{I}_{KL\times KL})\).
-
-Then, the estimated power for lineup \(l\) and individual \(i\) can be calculated as \(\hat{p}_{li} = g^{-1}(W_{li}\hat{\delta} + Z_{li}\hat{\tau}_{li})\) \autocite{majumder_validation_2013}.
-
-\hypertarget{limitations-of-lineup-protocol-and-automatic-visual-inference}{%
-\section{Limitations of Lineup Protocol and Automatic Visual Inference}\label{limitations-of-lineup-protocol-and-automatic-visual-inference}}
-
-Although lineup protocol has already been integrated into data analysis of various topics, such as diagnostics of hierarchical linear models \autocite{loy_diagnostic_2013}, geographical Research \autocite{widen_graphical_2016} and forensic examinations \autocite{krishnan_hierarchical_2021}, the involvement with human judgements limits its popularity.
-Similar to handicraft in pre-industrial society, lineup protocol conducted by humans is infeasible on a large scale, high in labour cost and time consuming.
-Moreover, it is strongly unfriendly to vision-impaired people.
-
-The ``steam engine'' of visual inference needs to developed for relieving people's workload by automating repeating tasks and providing standard result in a control environment.
-Large-scale evaluation of visual tests is not possible without the use of technology and machines.
-
-Modern computer vision model could be a promising solution to this problem.
-As a subfield of AI, computer vision with the modern deep learning architectures solved numerous critical problems in automation.
-Inspired by the vision processing in living organisms, the convolutional neural network was introduced by \textcite{fukushima_neocognitron_1982}.
-Soon, this architecture was applied to hand-written number recognition trained with back-propagation by \textcite{lecun_backpropagation_1989}.
-This was one of the earliest attempts which human successfully extract information from digital images via self-learning algorithms.
-Modern computer vision model is typically built on the deep neural network with convolutional layers \autocite{fukushima_neocognitron_1982}.
-Convolutional layers take advantage of the hierarchical pattern in data and provide regularized versions of fully-connected layers.
-It downscales and transforms the image by summarising information in a small space.
-Numerous studies have shown that it can be used to effectively tackle vision tasks, such as image recognition \autocite{rawat_deep_2017}.
-With the development of graphics processing units and the spread of high-performance personal computers, researches in computer vision become a new hype in the 21st century.
-Achievements such as computer-aided diagnosis \autocite{lee_image_2015}, pedestrian detection \autocite{brunetti_computer_2018} and facial recognition \autocite{emami_facial_2012} had a significant impact on our daily life.
-
-Using computer vision models to read data plots is not a general choice.
-some fields have adopted this idea by applying computer vision models to read recurrence plots for time series regression \autocite{ojeda_multivariate_2020}, time series classification \autocite{chu_automatic_2019,hailesilassie_financial_2019,hatami_classification_2018,zhang_encoding_2020}, anomaly detection \autocite{chen_convolutional_2020} and pairwise causality analysis \autocite{singh_deep_2017}.
-However, evaluating lineups with computer vision model is a new field of study.
-
-Therefore, we first tried to automate the visual diagnostics of one of the simplest statistical models - classical normal linear regression model.
-Two common types of model defects were created and visualized using residuals vs.~fitted values plot.
-To set up a comparison between the computer vision models and humans, human subject experiments were conducted to understand the ability of human reading residual plots.
-
-\hypertarget{structure-of-the-paper}{%
-\section{Structure of the paper}\label{structure-of-the-paper}}
-
-\ldots{}
-
-\hypertarget{human-subject-exeriments}{%
-\chapter{Human Subject Exeriments}\label{human-subject-exeriments}}
-
-To collect data of human performance on reading residual plot of linear regression model with nonlinearity and heteroskedasticity defects, two human experiments were conducted.
-Participants of both experiments were recruited using an online platform called Prolific \autocite{prolific_prolific_2022}.
-
-Prolific provides an international participant pool with the option to apply flexible pre-screening filters.
-In this study, we recruited \(62\) participants who was fluent in English with at least \(10\) previous submissions and \(98\)\% approval rate in other Prolific studies for quality control.
-Further, balance sample across gender was imposed to prevent gender bias.
-
-In Prolific, researchers can either approve or reject submissions based on the quality of the responses.
-If a submission is approved by the researcher, the participant will be paid a certain amount of money per hour of time spent on the experiment.
-To assess the quality of the responses, two attention checks were given to each participant during the experiment, where at least one of them was required to pass for the approval of submission.
-
-Throughout the experiments, participants were requested to complete a short survey and evaluate 20 lineups on a website in an hour.
-Each lineup consists of one actual residual plot and 19 null residual plots produced by plotting null residuals simulated from the residual rotation distributions.
-Among the 20 lineups, there were two extremely easy lineups used as attention checks which everyone should get correct.
-The short survey was intended to collect information about participant that might affect their ability in reading data plot including age, highest level of education, preferred pronoun and previous experience in similar study.
-For the evaluation of the lineup, participant first needed to select one or multiple most different data plot(s) from the lineup by clicking the corresponding image.
-Then, the primary reason for choosing the plots needed to be provided by picking one of the given options - ``outlier(s)'', ``cluster(s)'', ``shape'' and ``other''.
-Explanations about these options were provided in the training page and the mouseover text.
-Table \ref{tab:lineup-reason} gives the detailed explanations about these reasons.
-If the option ``other'' was selected, text input for the specified reason would be collected.
-Lastly, the degree of difference between their chosen data plot(s) and other data plots needed to be selected among five levels - ``not at all'', ``slightly'', ``moderately'', ``very'' and ``extremely''.
-Notably, if participants could not tell the difference between the data plots, there was an option to skip the evaluation of the lineup.
-However, to prevent participants abusing this option, warnings were given at the beginning of the study that skipping too many lineups might lead to rejection of submission since it demonstrated clear low-effort throughout the experiment.
-
-\begin{table}
-
-\caption{\label{tab:lineup-reason}Explanations about reasons for choosing data plots from a lineup}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{ll}
-\toprule
-Reasons & Explanations\\
-\midrule
-outlier(s) & In a data plot, an outlier is a point that differs significantly from other points.\\
-cluster(s) & In a data plot, a cluster is a group of points positioned closely together. 
-And usually, there will be gaps between different clusters.\\
-shape & It could be any common shapes we would see in real life, like triangle shape, U-shape, butterfly shape, etc.\\
-other & Patricipants needs to give their reasons in the text input.\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-The study website was powered by Flask \autocite{grinberg_flask_2018}, a web framework written in Python 3 \autocite{van_rossum_python_2009}, hosted on PythonAnywhere \autocite{pythonanywhere_pythonanywhere_2022}, a web hosting service provider.
-Due to the limit of the storage on PythonAnywhere, lineup images needed for the website were hosted separately on Github Pages \autocite{gtihub_github_2022-1}, a static site hosting service provided by Github \autocite{gtihub_github_2022}, stored in private Github repositories.
-The uniform resource locator (URL) to lineup images were set to be unique random strings such that images can not be accessed by general approaches without knowing the correct URL.
-This avoided participants seeing the lineup beforehand.
-The front-end of the website was built using a JavaScript \autocite{flanagan_javascript_2006} library jsPsych \autocite{de_leeuw_jspsych_2015} which is specialized in creating online behavioral experiments.
-One of the reasons we chose to use this library was it had a modularized but highly customizable template which could record participant's response time automatically.
-This is essential for us to confirm the quality of the data by checking exceptionally fast or slow responses.
-Figure \ref{fig:diagram-experiment} summarizes the online experiment setup.
-
-\begin{figure}
-\centering
-\includegraphics[width=4.6875in,height=4.6875in]{figures/experiment_tech.pdf}
-\caption{Diagram of online experiment setup.
-The server-side of the study webiste used Flask as backend hosted on PythonAnywhere.
-And the client-side used jsPsych to run experiment.
-\label{fig:diagram-experiment}}
-\end{figure}
-
-The first two pages of the website are the explanatory statement and the consent form.
-Participants were asked to read the documents and agree with the terms to advance to the short survey.
-With the completion of the survey, the next page is the training page, which contained instructions on how to interact with the webpage.
-Followed by the training page, there are 20 lineups each on a single page.
-Figure \ref{fig:website-layout} illustrates the layout of the website.
-At the end of the experiment, participants would be redirected back to Prolific and waited for researchers' responses.
-
-\begin{figure}
-\centering
-\includegraphics[width=5.72917in,height=3.11458in]{figures/website.png}
-\caption{Layout of the study website. Participants needed to choose the most different plots on the right and select their reasons and confidence levels on the left. \label{fig:website-layout}}
-\end{figure}
-
-Next we will discuss the simulation setup for this study.
-The experiment data was simulated by the use of the programming language \texttt{R} \autocite{r_core_team_r_2021}.
-For the ease of reproducibility, functions to build models, simulate data from models, produce lineup with data, allocate stimulis for subjects and evaluate subject responses from this study are bundled in the package \texttt{visage} with a unique object-oriented programming system built upon the environment feature of \texttt{R}.
-In the description of the simulation, corresponding functionalities of the package will be introduced.
-
-Two models were used in the study, where both were linear models with some degree of violations of classical assumptions.
-
-\hypertarget{cubic-model}{%
-\section{Cubic Model}\label{cubic-model}}
-
-The first model was a cubic linear model with two regressors, which can be expressed by:
-\begin{equation} \label{eq:cubic}
-\boldsymbol{Y}= 1 + (2-c)\boldsymbol{X} + c\boldsymbol{Z} + a[(2-c)\boldsymbol{X}]^2+a(c\boldsymbol{Z})^2+b[(2-c)\boldsymbol{X}]^3+b(c\boldsymbol{Z})^3+\boldsymbol{\varepsilon},
-\end{equation}
-where \(c \in (0,2)\), \(a \in (-3,3)\), \(b \in (-3,3)\), \(\boldsymbol{\varepsilon}\overset{iid}{\sim} N(\boldsymbol{0},\sigma^2\boldsymbol{I})\), \(\boldsymbol{Y}\), \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) are \(n\times1\) matrices.
-
-This defines a cubic relationship between \(\boldsymbol{Y}\), \(\boldsymbol{X}\) and \(\boldsymbol{Z}\).
-Meanwhile, to create nonlinearity defect, the null model followed the assumptions of the classical normal linear regression model (CNLRM), fitted by OLS is:
-\begin{equation} \label{eq:cubic_null}
-\boldsymbol{Y}=\beta_0+\beta_1\boldsymbol{X}+\beta_2\boldsymbol{Z}+\boldsymbol{u},
-\end{equation}
-where \(\boldsymbol{u} \sim N(0,\sigma^2_u\boldsymbol{I})\).
-
-Clearly, omitted-variable bias will present since the null model leaves out the quadric and cubic terms.
-
-\begin{lemma}[Distribution of residuals produced by the cubic model] \label{lemma:cubic}
-Given the data generating process in Equation (\ref{eq:cubic}), and null model in Equation (\ref{eq:cubic_null}).
-Let $\boldsymbol{X}_a=[\boldsymbol{1},\boldsymbol{X},\boldsymbol{Z}]$ denotes the set of regressors in matrix form.
-Then, the residuals obtained from the null model are $$\boldsymbol{e} \sim N(\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b, \sigma^2\boldsymbol{R}_a),$$ where $\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$, $\boldsymbol{X}_b=[\boldsymbol{X}^2,\boldsymbol{Z}^2,\boldsymbol{X}^3,\boldsymbol{Z}^3]$ and $\boldsymbol{\beta}_b=(a(2-c)^2,ac^2,b(2-c)^3,bc^3)'$.
-\end{lemma}
-
-\begin{proof}
-Using the Frisch–Waugh–Lovell theorem, the residuals obtained by the null model are $$\boldsymbol{e}=\boldsymbol{R}_a\boldsymbol{Y}=\boldsymbol{R}_a(\boldsymbol{X}_a\boldsymbol{\beta}_a+\boldsymbol{X}_b\boldsymbol{\beta}_b+\boldsymbol{\varepsilon}),$$ where $\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$, $\boldsymbol{\beta}_a=(1,2-c,c)'$, $\boldsymbol{X}_b=[\boldsymbol{X}^2,\boldsymbol{Z}^2,\boldsymbol{X}^3,\boldsymbol{Z}^3]$ and $\boldsymbol{\beta}_b=(a(2-c)^2,ac^2,b(2-c)^3,bc^3)'$.
-
-Because $\boldsymbol{R}_a\boldsymbol{X}_a=\boldsymbol{0}$, we have $\boldsymbol{e}=\boldsymbol{R}_a(\boldsymbol{X}_b\boldsymbol{\beta}_b+\boldsymbol{\varepsilon}).$ Since $\boldsymbol{\varepsilon} \sim N(0,\sigma^2\boldsymbol{I})$, it follows that $\boldsymbol{e} \sim N(\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b, \sigma^2\boldsymbol{R}_a)$.
-\end{proof}
-
-Lemma \ref{lemma:cubic} shows that the expectation of the residuals is clearly a function of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\), which indicates the residuals and the fitted values \(\hat{Y} = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'\boldsymbol{Y}\) are associated.
-Hence, it is expected that visual discoveries can be found in the residuals vs.~fitted values plot.
-Let \(X_i\) and \(Z_i\), \(i = 1,...,n\), be independent random variables follow uniform distribution \(U(-1,1)\).
-Given the expectation of the residuals, we could plot the expected values of residuals against the observed values.
-Figure \ref{fig:cubic-shape-1}, \ref{fig:cubic-shape-2} and \ref{fig:cubic-shape-3} illustrate the shape of residuals and their expected values under different parameter settings.
-
-From the Figure \ref{fig:cubic-shape-1}, it can be observed that with fixed \(\sigma\) and \(c\), \(a\) and \(b\) are controlling the 2D projection of a hypersurface, and seemingly performing some rotations along different axes.
-Figure \ref{fig:cubic-shape-2} shows that with fixed \(a\), \(b\) and \(\sigma\), \(c\) is controlling the contribution of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) to \(\boldsymbol{Y}\).
-As \(c\) moves toward \(0\) or \(2\), one regressor will dominate another, which will mitigate some joint effects and resemble a typical cubic function.
-In Figure \ref{fig:cubic-shape-3}, \(a\), \(b\) and \(c\) are held fixed, and \(\sigma\) is controlling the noises around the expected values.
-The underlying shape is clearer for smaller \(\sigma\).
-
-\begin{figure}
-\centering
-\includegraphics{thesis_files/figure-latex/cubic-shape-1-1.pdf}
-\caption{\label{fig:cubic-shape-1}A matrix of residuals vs.~fitted values plot under different parameter settings of the cubic model.
-Resdiauls are drawns in black and the expected values are drawn in red.
-The plots show that if \(\sigma\) and \(c\) are held fixed, \(a\) and \(b\) rotate a two-dimension projection of a hypersurface.}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{thesis_files/figure-latex/cubic-shape-2-1.pdf}
-\caption{\label{fig:cubic-shape-2}A matrix of residuals vs.~fitted values plot under different parameter settings of the cubic model.
-Resdiauls are drawns in black and the expected values are drawn in red.
-The plots show that if \(a\), \(b\) and \(\sigma\) are held fixed, \(c\) controls the contribution of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) to \(\boldsymbol{Y}\).}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{thesis_files/figure-latex/cubic-shape-3-1.pdf}
-\caption{\label{fig:cubic-shape-3}A matrix of residuals vs.~fitted values plot under different parameter settings of the cubic model.
-Resdiauls are drawns in black and the expected values are drawn in red.
-The plots show that if \(a\), \(b\) and \(c\) are held fixed, \(\sigma\) controls the strength of the signal.}
-\end{figure}
-
-The residuals used in these three figures are simulated from the cubic models built using the \texttt{cubic\_model()} function from the package \texttt{visage}.
-\texttt{cubic\_model()} is a cubic model class constructor, which takes arguments \texttt{a}, \texttt{b}, \texttt{c}, \texttt{sigma}, \texttt{x} and \texttt{z}, where the first four are numeric values defined above, and \texttt{x} and \texttt{z} are random variable instances created by the random variable abstract base class constructor \texttt{rand\_var()}.
-
-If we would like \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) to be random uniform variables ranged from \(-1\) to \(1\), it can be achieved by using the random uniform variable class constructor \texttt{rand\_uniform()} inherited from the random variable abstract base class.
-It only takes two arguments which are the lower bound and the upper bound of the support.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{library}\NormalTok{(visage)}
-\NormalTok{mod }\OtherTok{\textless{}{-}} \FunctionTok{cubic\_model}\NormalTok{(}\AttributeTok{a =} \SpecialCharTok{{-}}\DecValTok{3}\NormalTok{, }\AttributeTok{b =} \SpecialCharTok{{-}}\DecValTok{3}\NormalTok{, }\AttributeTok{c =} \DecValTok{1}\NormalTok{, }\AttributeTok{sigma =} \FloatTok{0.5}\NormalTok{, }
-                   \AttributeTok{x =} \FunctionTok{rand\_uniform}\NormalTok{(}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{), }\AttributeTok{z =} \FunctionTok{rand\_uniform}\NormalTok{(}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-An instance of cubic model class contains methods of simulating data and making residual plot.
-Method \texttt{mod\$gen()} returns a data frame containing realizations of \(X\), \(Z\), \(Y\) and \(\varepsilon\) simulated from the model.
-The number of realizations depends on the integer argument \texttt{n}.
-In addition, if argument \texttt{fit\_model\ =\ TRUE}, a null model will be fitted using the simulated data and residuals and fitted values will be included in the returned data frame.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{mod}\SpecialCharTok{$}\FunctionTok{gen}\NormalTok{(}\AttributeTok{n =} \DecValTok{5}\NormalTok{, }\AttributeTok{fit\_model =} \ConstantTok{TRUE}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##            y          x          z           e     .resid    .fitted
-## 1 -0.8263671 -0.4988275  0.5569648 -0.06142887  1.2340651 -2.0604322
-## 2  0.6134661  0.1762335  0.6512424  0.99654342  1.1758970 -0.5624310
-## 3  0.2285071 -0.2814395 -0.1511556 -0.10996690 -0.3757603  0.6042674
-## 4 -4.6320382 -0.5813456  0.9251409 -0.60826338 -1.2664307 -3.3656075
-## 5  0.3148180  0.5647967  0.4406211  0.64596536 -0.7677711  1.0825892
-\end{verbatim}
-
-Method \texttt{mod\$plot()} produce a \texttt{ggplot} \autocite{wickham_ggplot2_2011} object.
-It takes a data frame containing columns \texttt{.resid} and \texttt{.fitted} as input, along with a character argument \texttt{type} indicating the type of the data plot, and other aesthetic arguments such as \texttt{size} and \texttt{alpha} to control the appearance of the plot.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{mod}\SpecialCharTok{$}\FunctionTok{plot}\NormalTok{(mod}\SpecialCharTok{$}\FunctionTok{gen}\NormalTok{(}\AttributeTok{n =} \DecValTok{100}\NormalTok{, }\AttributeTok{fit\_model =} \ConstantTok{TRUE}\NormalTok{), }\AttributeTok{type =} \StringTok{"resid"}\NormalTok{, }\AttributeTok{size =} \DecValTok{1}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics{thesis_files/figure-latex/test3-1.pdf}
-
-Lineup is a matrix of residual plots which can be produced by using the methods \texttt{mod\$gen\_lineup()} and \texttt{mod\$plot\_lineup()}.
-Method \texttt{mod\$gen\_lineup} takes the number of realizations \texttt{n} and the number of plots in a lineup \texttt{k} as inputs.
-And the method \texttt{mod\$plot\_lineup()} has the same user interface as \texttt{mod\$plot()}.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{mod}\SpecialCharTok{$}\FunctionTok{plot\_lineup}\NormalTok{(mod}\SpecialCharTok{$}\FunctionTok{gen\_lineup}\NormalTok{(}\AttributeTok{n =} \DecValTok{50}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{), }\AttributeTok{type =} \StringTok{"resid"}\NormalTok{, }\AttributeTok{size =} \DecValTok{1}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics{thesis_files/figure-latex/unnamed-chunk-4-1.pdf}
-
-The cubic model class also provides method to compute the expected values of residuals.
-Method \texttt{mod\$E()} takes a data frame with columns \texttt{x} and \texttt{z} as input, and returns a vector of expected values of residuals.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{mod}\SpecialCharTok{$}\FunctionTok{E}\NormalTok{(mod}\SpecialCharTok{$}\FunctionTok{gen}\NormalTok{(}\AttributeTok{n =} \DecValTok{5}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## [1] -0.092183951 -0.003912092 -0.001531723  0.148353478 -0.050725712
-\end{verbatim}
-
-Since we know that under the null hypothesis, the residual \(\boldsymbol{e}\sim N(\boldsymbol{0},\sigma^2\boldsymbol{R}_a)\).
-Thus, the difference between the expected values \(\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b\) and \(\boldsymbol{0}\) represents the direct impact of the parameters \(a\) and \(b\) on the residuals.
-It is expected that the larger the magnitude of the expected value relative to the variance and covariance, the easier the human to spot the patterns in the residual plot.
-
-To obtain a measure of the impact of \(a\) and \(b\) on the residuals adjusted for variance and covariance, we need to address several properties of the residuals.
-First, the variance of the residuals \(\sigma^2\boldsymbol{R}_a\) is not an identity matrix.
-This can be fixed by standardizing the residuals by their variance-covariance matrix.
-Second, the difference between \(\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b\) and \(\boldsymbol{0}\) could be negative, which is not ideal for comparison.
-Thus, the magnitude needs to be squared.
-Third, the measure needs to be a scalar.
-We could apply a weighted average operator \(\boldsymbol{W}\) on the transformed expected residuals to obtain a single numeric value.
-For simplicity, we set \(\boldsymbol{W}=n^{-1}\boldsymbol{1}\).
-Considering the high time complexity of computing the square root of \(\boldsymbol{R}_a\), off-diagonal elements of \(\boldsymbol{R}_a\) are set to be zeros.
-This gives the effect size: \[ES=n^{-1}||\sigma^{-1}\boldsymbol{R}_a^{-\frac{1}{2}}\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b||^2=n^{-1}\sigma^{-2}||\boldsymbol{R}_a^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2\approx n^{-1}\sigma^{-2}||diag(\boldsymbol{R}_a)^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2,\] where \(diag(\boldsymbol{R}_a)\) is the diagonal matrix constructed from the diagonal elements of \(\boldsymbol{R}_a\).
-
-The interpretation of this effect size is the impact of parameter \(a\) and \(b\) on the squared deviation of the standardized expected residual per observation.
-It is not directly related to the shape, or the pattern human observed from the residual plot, but it is a reasonable approximation of the degree of the visual deviation from the null residuals under our cubic model setting.
-Figure \ref{fig:effectsize} shows four residual plots with different effect sizes.
-As effect sizes increases, the strength of the signal become stronger.
-
-\begin{figure}
-\centering
-\includegraphics{thesis_files/figure-latex/effectsize-1.pdf}
-\caption{\label{fig:effectsize}Cubic model residual plots under different effect sizes. The larger the effect size, the stronger the signal.}
-\end{figure}
-
-Under the cubic model setting, there is an exact conventional test for testing the nonlinearity defect, which is F-test.
-For F-test, the null hypothesis is \(H_0:a=b=0\), and the alternative hypothesis is \(H_1:\text{at least one of them} \neq 0\).
-During the simulation of the lineup data, the F-statistic and the p-value will be recorded for comparison between the power of conventional test and visual test.
-
-\hypertarget{heteroskedasticity-model}{%
-\section{Heteroskedasticity Model}\label{heteroskedasticity-model}}
-
-Another model used in the experiments was a heteroskedasticity model with one regressor, which can be expressed by:
-\begin{equation} \label{eq:heter}
-Y_i = 1 + X_i + \varepsilon_i, ~i = 1,...,n,
-\end{equation}
-where \(a \in \{-1,0,1\}\), \(b\in (0,32)\) and \(\varepsilon_i \overset{iid}{\sim} N(0,1+b(2-|a|)(X_i-a)^2)\).
-
-To create heteroskedasticity defect, OLS was used to fit the null model:
-\begin{equation} \label{eq:heter_null}
-\boldsymbol{Y}=\beta_0+\beta_1\boldsymbol{X}+\boldsymbol{u},
-\end{equation}
-where \(\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2_u\boldsymbol{I})\).
-
-In this case, estimators of \(\beta_0\) and \(\beta_1\) are unbiased, but the error term has non-constant variance.
-
-\begin{lemma}[Distribution of residuals produced by the heteroskedasticity model] \label{lemma:heter}
-Given the data generating process in Equation (\ref{eq:heter}) and null model in Equation (\ref{eq:heter_null}).
-Let $\boldsymbol{X}_a=[\boldsymbol{1},\boldsymbol{X}]$ denotes the set of regressors in matrix form.
-The residuals obtained from the null model are $$\boldsymbol{e} \sim N(\boldsymbol{0}, \boldsymbol{R}_a\boldsymbol{V}),$$ where $\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ and $\boldsymbol{V}$ is a diagonal matrix with $V_{ii}=1+b(2-|a|)(X_i  -  a)^2$, $i = 1,...,n$.
-\end{lemma}
-
-\begin{proof}
-Using the Frisch–Waugh–Lovell theorem, the residuals obtained by the null model are $\boldsymbol{e}=\boldsymbol{R}_a\boldsymbol{Y}=\boldsymbol{R}_a(\boldsymbol{X}_a\boldsymbol{\beta}_a+\boldsymbol{\varepsilon}),$ where $\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ and $\boldsymbol{\beta}_a=(1,1)'$.
-
-Because $\boldsymbol{R}_a\boldsymbol{X}_a=\boldsymbol{0}$, we have $\boldsymbol{e}=\boldsymbol{R}_a\boldsymbol{\varepsilon}$.
-Hence, the residuals $\boldsymbol{e}$ follow $N(\boldsymbol{0}, \boldsymbol{R}_a\boldsymbol{V})$, where $\boldsymbol{V}$ is a diagonal matrix with $V_{ii}=1+b(2-|a|)(X_i  -  a)^2$, $i = 1,...,n$.
-\end{proof}
-
-Lemma \ref{lemma:heter} shows that variance and covariance of the residuals depend on \(\boldsymbol{X}\).
-We could plot the one standard deviation around the residuals to indicate the region about 68\% of the residuals should landed on.
-Figure \ref{fig:hetershape} illustrates different shapes of residuals undder various values of \(a\) and \(b\).
-From the plot, it can be observed that if \(a = 0\), the residual plot looks like the shape of butterfly.
-If \(a = \pm 1\), it looks like a triangle, and the sign of \(a\) determines the direction of the shape.
-Parameter \(b\) is controlling the strength of the signal.
-As \(b\) increases, the pattern becomes less noisy.
-
-\begin{figure}
-\centering
-\includegraphics{thesis_files/figure-latex/hetershape-1.pdf}
-\caption{\label{fig:hetershape}A matrix of residuals vs.~fitted values plot under different parameter settings of the heteroskedasticity model.
-Resdiauls are drawns in black and the one standard deviation around zero is drawn in red.
-The plots show that \(a\) controls the shape and direction of the resdiaul plot, while \(b\) controls the strength of the signal.}
-\end{figure}
-
-Similar to the cubic model, the heteroskedasticity model could also be built by the heteroskedasticity model class constructor \texttt{heter\_model()}.
-This function takes three arguments as inputs, which are \texttt{a}, \texttt{b} and \texttt{x}.
-\texttt{a} and \texttt{b} are numeric parameters defined in Equation (\ref{eq:heter}).
-\texttt{x} needs to be a random variable object.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{library}\NormalTok{(visage)}
-\NormalTok{mod }\OtherTok{\textless{}{-}} \FunctionTok{heter\_model}\NormalTok{(}\AttributeTok{a =} \DecValTok{0}\NormalTok{, }\AttributeTok{b =} \DecValTok{16}\NormalTok{, }\AttributeTok{x =} \FunctionTok{rand\_uniform}\NormalTok{(}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-Since both the cubic model class and the heteroskedasticity class are inherited from the visual inference model class, which has defined methods of simulating data, making residual plot and producing lineup, heteroskedasticity model object can be used in a similar way as cubic model object.
-The following codes give examples of the use of the object.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{mod}\SpecialCharTok{$}\FunctionTok{gen}\NormalTok{(}\DecValTok{5}\NormalTok{, }\AttributeTok{fit\_model =} \ConstantTok{TRUE}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##          y          x    sigma         e     .resid  .fitted
-## 1 7.205499  0.9053470 5.218132 5.3001525  0.7420902 6.463409
-## 2 2.560696 -0.8626811 4.981465 2.4233771  1.0441078 1.516588
-## 3 1.121106 -0.6545505 3.835357 0.7756568 -0.9778170 2.098923
-## 4 5.296467  0.7171334 4.178154 3.5793335 -0.6403337 5.936801
-## 5 2.962130 -0.2859733 1.901837 2.2481030 -0.1680473 3.130177
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{mod}\SpecialCharTok{$}\FunctionTok{plot}\NormalTok{(mod}\SpecialCharTok{$}\FunctionTok{gen}\NormalTok{(}\DecValTok{100}\NormalTok{, }\AttributeTok{fit\_model =} \ConstantTok{TRUE}\NormalTok{), }\AttributeTok{size =} \DecValTok{1}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics{thesis_files/figure-latex/unnamed-chunk-8-1.pdf}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{mod}\SpecialCharTok{$}\FunctionTok{plot\_lineup}\NormalTok{(mod}\SpecialCharTok{$}\FunctionTok{gen\_lineup}\NormalTok{(}\DecValTok{100}\NormalTok{), }\AttributeTok{size =} \DecValTok{1}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics{thesis_files/figure-latex/unnamed-chunk-9-1.pdf}
-According to Lemma \ref{lemma:heter}, when \(b = 0\), the matrix \(\boldsymbol{V}\) collapses to an identity matrix.
-Assume the shape of butterfly and triangle have identical visual impacts, the only factor affects human to recognize the pattern is the strength of the signal, where parameter \(a\) has no role to play.
-In addition, sample size has a huge impact on the chance of human recognizing the pattern.
-As shown in Figure \ref{fig:hetern}, as sample size increases, the pattern becomes easier to observe.
-However, this effect is less noticeable with large sample size as the outline of the shape has been drawn and residuals have less probability to land outside of it.
-Thus, the effect size of this model can be expressed by \(ES = b\sqrt{n}\).
-The square root operator is used for addressing the large sample issue.
-Figure \ref{fig:hetereffectsize} shows the effectiveness of the effect size.
-
-\begin{figure}
-\centering
-\includegraphics{thesis_files/figure-latex/hetern-1.pdf}
-\caption{\label{fig:hetern}Residuals generated from the same heteroskedasticity model but with different sample size. As sample size increases, the shape of butterfly becomes more obvious.}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{thesis_files/figure-latex/hetereffectsize-1.pdf}
-\caption{\label{fig:hetereffectsize}Residuals of heteroskedasticity model with different effect size. As effect size increases, the shape of butterfly becomes less noisy.}
-\end{figure}
-
-For the heteroskedasticity model, the conventional test we used was Breusch--Pagan test \autocite{breusch_simple_1979}, which tested whether the variance of the error terms of the regression is dependent on the regressors with the auxiliary regression equation \[\boldsymbol{e}^2 = \gamma_0 + \gamma_1 \boldsymbol{X} + \gamma_2 \boldsymbol{X}^2 + \boldsymbol{v}.\] \textcite{majumder_validation_2013} suggested that visual test is not expected to perform equally well as conventional test especially when there exists a exact conventional test.
-However, in contrast to the F-test used in the cubic model, Breusch--Pagan test is an approximate test.
-Thus, the power of visual test may exceed the power of Breusch--Pagan test.
-Throughout the study, Breusch--Pagan test statistic and p-value were recorded for comparison.
-
-\hypertarget{distribution-of-regressors}{%
-\section{Distribution of regressors}\label{distribution-of-regressors}}
-
-The model definitions given in the previous two sections does not include the specification of the regressors.
-In this sections, distribution of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) will be discussed.
-
-The cubic model involved the use of both \(\boldsymbol{X}\) and \(\boldsymbol{Z}\).
-In the simulation, \(X_i\), \(i=1,...,n\), had equal chance to follow one of the following distributions: \(N(0, 0.09)\), \(U(-1, 1)\), \(Lognormal(0,0.36)/3 - 1\) and \(-Lognormal(0,0.36)/3 + 1\).
-Uniform and normal distribution were symmetric and common.
-Adjusted lognormal distribution and adjusted negative lognormal distribution provided right-skewed and left-skewed density respectively.
-These distributions were chosen such that most the realizations will fall between \(-1\) and \(1\).
-
-The distribution of \(Z_i\), \(i=1,...,n\), had \(50\)\% chance to be a uniform distribution ranged from \(-1\) to \(1\), and \(50\)\% chance to be a discrete uniform distribution with \(z_n\) outcomes simulated from a uniform distribution ranged from \(-1\) to \(1\).
-\(z_n\) itself was a discrete uniform distribution with outcomes \(\{10,11,12,13,14,15,16,17,18,19,20\}\), which defined the number of possible values \(Z_i\) could take.
-As shown in Figure \ref{fig:discreteex}, this setup would create discreteness in residual plot, which could enrich the pool of visual patterns.
-
-\begin{figure}
-\centering
-\includegraphics{thesis_files/figure-latex/discreteex-1.pdf}
-\caption{\label{fig:discreteex}Discreteness in residuals created by using a discrete uniform random variable as one of the regressors in a cubic model.
-For each residual plot of the lineup, there are three clusters because the number of possible values the regressor can take is three.}
-\end{figure}
-
-\(\boldsymbol{X}\) used in the heteroskedasticity model was a combination of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) used in the cubic model.
-It could be one of the five distributions mentioned above - normal distribution, uniform distribution, adjusted lognormal distribution, adjusted negative lognormal distribution and discrete uniform distribution.
-
-\hypertarget{experiment-i-and-ii}{%
-\section{Experiment I and II}\label{experiment-i-and-ii}}
-
-In this study, multiple selection was allowed for the evaluation of a lineup.
-Given our desired significance level was \(\alpha = 0.05\) and the number of plots in a lineup was \(20\), a single visual test procedure needed to involve multiple evaluations.
-Otherwise, the \(p\)-value of the test would be greater than \(0.05\) in spite of the actual data plot being selected.
-The number of plots selected by a user depended on the user and the difficulty of the lineup.
-According to the pilot study evaluated by \(10\) faculty members of Department of Econometrics and Business Analytics, Monash University, the number of selections would be generally smaller than \(4\), which suggested a visual test consisted of \(5\) evaluations were sufficient to yield \(p\)-value smaller than \(0.05\) with at least three detections.
-Thus, each lineup was replicated for five times and evaluated by different participants.
-However, our study website was designed such that any set of lineups consisting of 20 lineups generated beforehand would be shown to only one person only once.
-And whether the participant would finish the experiment was unpredictable.
-Therefore, some lineups had insufficient number of evaluations, but some had more evaluations than expected.
-This slightly affected the estimate of the power of the visual test, which will be discussed in Section \ref{p-value-and-power-estimation-of-visual-test-allowed-for-multiple-selections}.
-
-We called the lineup being detected if the actual data plot was one of the selections of an evaluation of a lineup.
-The sample size calculation was based on the detection rate of the lineup, which was closely related to the difficulty level of the lineup.
-With the data collected from the pilot study, two logistic regressions given in Table \ref{tab:pilotglm} and Figure \ref{fig:pilotglm} were developed to describe the relationship between the natural logarithm of effect size and the detection rate for the cubic model and the heteroskedasticity model.
-
-\begin{table}[!htbp] \centering 
-  \caption{Logistic regressions with detection rate as response variable and natural logarithm of the effect size as regressor fitted on the pilot data for both cubic and heteroskedasticity models.} 
-  \label{tab:pilotglm} 
-\begin{tabular}{@{\extracolsep{5pt}}lcc} 
-\\[-1.8ex]\hline 
-\hline \\[-1.8ex] 
- & \multicolumn{2}{c}{\textit{Dependent variable:}} \\ 
-\cline{2-3} 
-\\[-1.8ex] & \multicolumn{2}{c}{detect} \\ 
- & Cubic & Heteroskedasticity \\ 
-\\[-1.8ex] & (1) & (2)\\ 
-\hline \\[-1.8ex] 
- log(effect\_size) & 0.611$^{***}$ & 0.737$^{***}$ \\ 
-  & (0.098) & (0.137) \\ 
-  & & \\ 
- Constant & 0.457$^{***}$ & $-$2.402$^{***}$ \\ 
-  & (0.165) & (0.645) \\ 
-  & & \\ 
-\hline \\[-1.8ex] 
-Observations & 246 & 289 \\ 
-Log Likelihood & $-$122.305 & $-$147.971 \\ 
-Akaike Inf. Crit. & 248.610 & 299.943 \\ 
-\hline 
-\hline \\[-1.8ex] 
-\textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
-\end{tabular} 
-\end{table}
-
-\begin{figure}
-\centering
-\includegraphics{thesis_files/figure-latex/pilotglm-1.pdf}
-\caption{\label{fig:pilotglm}Logistic regressions with detection rate as response variable and natural logarithm of the effect size as regressor fitted on the pilot data for both cubic and heteroskedasticity models.
-Points are jittered to reduce overplotting using the \texttt{ggbeeswarm} package.
-The smooth curves are the fitted regression curves.
-The ribbons around the curves are the 95\% confidence intervals.}
-\end{figure}
-
-Lineups were classified into three categories - easy, medium and hard.
-Easy lineups were those of which predicted detection rates were higher than \(80\)\%, while medium lineups were between \(40\)\% to \(80\)\% and hard lineups were less than \(40\)\%.
-
-For the first experiment, we would like to have at least 20 detection with chance greater than 99.99\% in each category.
-Thus, we prepared \(68\) easy lineups, \(120\) medium lineups and \(388\) hard lineups with five replications.
-Every participant would get at least \(8\) easy or medium lineups.
-Two out of \(10\) extremely lineups with predicted detection rate over \(90\)\% were randomly given to every participant as attention checks.
-We initially planned to recruit \((68+120+388) \times 5 / (20-2) = 160\) participants but eventually decided to only recruit \(20\) participants because it was the first time we launched an experiment on Prolific where factors like payment method and quality of participants were unclear at the moment.
-As a consequence, most of the lineups had only two evaluations and not all the lineups were used.
-However, data collected from the first experiment was sufficient for building a more accurate logistic regression to predict detection rate for the second experiment.
-
-\begin{figure}
-\centering
-\includegraphics{thesis_files/figure-latex/e1glmplot-1.pdf}
-\caption{\label{fig:e1glmplot}Logistic regressions with detection rate as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I for both cubic and heteroskedasticity models.
-Points are jittered to reduce overplotting using the \texttt{ggbeeswarm} package.
-The smooth curves are the fitted regression curves.
-The ribbons around the curves are the 95\% confidence intervals.}
-\end{figure}
-
-\begin{table}[!htbp] \centering 
-  \caption{Logistic regressions with detection rate as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I for both cubic and heteroskedasticity models.}
-  \label{tab:e1glm} 
-\begin{tabular}{@{\extracolsep{5pt}}lcc} 
-\\[-1.8ex]\hline 
-\hline \\[-1.8ex] 
- & \multicolumn{2}{c}{\textit{Dependent variable:}} \\ 
-\cline{2-3} 
-\\[-1.8ex] & \multicolumn{2}{c}{detect} \\ 
- & Cubic & Heteroskedasticity \\ 
-\\[-1.8ex] & (1) & (2)\\ 
-\hline \\[-1.8ex] 
- log(effect\_size) & 0.857$^{***}$ & 0.746$^{***}$ \\ 
-  & (0.122) & (0.111) \\ 
-  & & \\ 
- Constant & $-$1.571$^{***}$ & $-$3.434$^{***}$ \\ 
-  & (0.257) & (0.462) \\ 
-  & & \\ 
-\hline \\[-1.8ex] 
-Observations & 192 & 188 \\ 
-Log Likelihood & $-$64.544 & $-$84.606 \\ 
-Akaike Inf. Crit. & 133.089 & 173.212 \\ 
-\hline 
-\hline \\[-1.8ex] 
-\textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
-\end{tabular} 
-\end{table}
-
-According to the new logistic regression given in Table \ref{tab:e1glm} and Figure \ref{fig:e1glmplot}, it was found that the first experiment might be too difficult for participants as the detection rate was lower than what we estimated from the pilot study, especially for the heteroskedasticity model.
-Thus, for the second experiments, we relaxed the sample size requirements and increased the proportion of easy lineups.
-In addition, due to budget constraint, the maximum number of participants was \(50\).
-With the new logistic regression fitted with data collected from the first experiment, \(28\) easy lineups, \(36\) medium lineups and \(100\) hard lineups were simulated for the second experiment.
-\(50\) participants were recruited but \(6\) were withdrew.
-
-\hypertarget{p-value-and-power-estimation-of-visual-test-allowed-for-multiple-selections}{%
-\section{\texorpdfstring{\(P\)-value and Power Estimation of Visual Test Allowed for Multiple Selections}{P-value and Power Estimation of Visual Test Allowed for Multiple Selections}}\label{p-value-and-power-estimation-of-visual-test-allowed-for-multiple-selections}}
-
-The \(p\)-value calculation for multiple selection is an extension of Equation (\ref{eq:pvaluesingle}).
-The assumptions about the independence still needed.
-Let \(K\) be the number of independent evaluations of a visual test, \(s_i\), \(i = 1,...,K\) be the number of selections of the evaluations, and \(X\) be the random variable denoting the number of detections.
-Then, the \(p\)-value of the visual test is given as:
-\begin{equation} \label{eq:pvaluemulti}
-P(X\geq x) = \sum_{j=x}^{K}Pr(j|s_1,...,s_K).
-\end{equation}
-The distribution of \(X\) given \(s_1,...s_K\) can not be derived trivially, as it is a sampling without replacement problem. In practice, this distribution can be approximated by computer simulation.
-
-Function \texttt{sim\_dist()} from package \texttt{visage} is designed to approximate the distribution of number of detections of a lineup via Monte Carlo method.
-It takes the number of evaluations \texttt{n\_eval}, the number of selections \texttt{n\_sel}, and the number of plots in a lineup \texttt{n\_plot} as inputs, then outputs a discrete distribution.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{sim\_dist}\NormalTok{(}\AttributeTok{n\_eval =} \DecValTok{3}\NormalTok{, }\AttributeTok{n\_sel =} \FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{2}\NormalTok{, }\DecValTok{3}\NormalTok{), }\AttributeTok{n\_plot =} \DecValTok{20}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##       0       1       2       3 
-## 0.70182 0.25202 0.04346 0.00270
-\end{verbatim}
-
-Function \texttt{calc\_p\_value()} from the same package calculates the \(p\)-value using the distribution returned by \texttt{sim\_dist()}.
-It takes an additional argument \texttt{n\_detect}, which is the number of detections.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{calc\_p\_value}\NormalTok{(}\AttributeTok{n\_detect =} \DecValTok{2}\NormalTok{, }\AttributeTok{n\_eval =} \DecValTok{3}\NormalTok{, }\AttributeTok{n\_sel =} \FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{2}\NormalTok{, }\DecValTok{3}\NormalTok{), }\AttributeTok{n\_plot =} \DecValTok{20}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## [1] 0.04688
-\end{verbatim}
-
-The above example shows that a visual test with two detections and three independent evaluations in which three observers select two, two, and three plots out of \(20\) plots respectively yields a \(p\)-value smaller than \(0.05\).
-
-Assume there is a visual test \(V_K\), where \(K\) denoting the number of evaluations.
-The corresponding \(p\)-value of \(V_K\) can be computed by using Equation (\ref{eq:pvaluemulti}).
-Meanwhile, if one evaluation is randomly deleted from \(V_K\), the remaining evaluations can still be used to form another valid visual test \(V_{K-1}\).
-In fact, considering all the possibilities, \(K\) different outcomes for \(V_{K-1}\) can be obtained.
-Since all outcomes occur with equal probability, the proportion of outcomes which reject the null hypothesis can be used as an estimate of the power of the visual test \(V_{K-1}\).
-Similarly, if we would like to estimate the power of the visual test \(V_{K-j}\) given the evaluations of \(V_{K}\), for \(j < K\), we could find all the possible combinations of \(K\) elements, taken \(k-j\) at a time to obtain \(K\choose{K-j}\) different outcomes for \(V_{K-j}\).
-Then, the estimated power is given as \(R/{K\choose{K-j}}\), where \(R\) is the number of outcomes which reject the null hypothesis.
-
-Function \texttt{calc\_p\_value\_comb()} from the package \texttt{visage} can be used to compute \(p\)-values of \(V_{K-j}\).
-The first argument is \texttt{detected}, which needs to be a vector of Boolean values denoting whether the observer detects the actual data plot.
-Desired number of evaluations and number of selections need to be provided via argument \texttt{n\_eval} and \texttt{n\_sel}.
-
-For example, given a visual test \(V_3\) where the first observer selects one plot and gets correct, the second observer selects one plot but misses, and the third observer selects two plots and gets correct.
-The \(p\)-value of all possible outcomes of \(V_2\) can be computed using the following code:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{calc\_p\_value\_comb}\NormalTok{(}\AttributeTok{detected =} \FunctionTok{c}\NormalTok{(}\ConstantTok{TRUE}\NormalTok{, }\ConstantTok{FALSE}\NormalTok{, }\ConstantTok{TRUE}\NormalTok{), }
-                  \AttributeTok{n\_eval =} \DecValTok{2}\NormalTok{, }
-                  \AttributeTok{n\_sel =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## [1] 0.09352 0.00624 0.14454
-## attr(,"combinations")
-##      [,1] [,2] [,3]
-## [1,]    1    1    2
-## [2,]    2    3    3
-\end{verbatim}
-
-\texttt{calc\_p\_value\_comb()} returns a vector of \(p\)-value along with an attribute which is a matrix representing elements being used by the specific outcome.
-Elements in the first column indicates that the first observer and the second observer are involved in the visual test, and the corresponding \(p\)-value is \(0.09352\).
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-We collected \(400\) lineup evaluations made by \(20\) participants in experiment I and \(880\) lineup evaluations made by \(44\) participants in experiment II.
-In total, \(442\) unique lineups were evaluated by \(64\) subjects.
-In experiment I, one of the participants skipped all \(20\) lineups.
-Hence, the submission was rejected and removed from the dataset.
-In experiment II, there was a participant failed one of the two attention checks, but there was no further evidence of low-effort throughout the experiment.
-Therefore, the submission was kept.
-
-\hypertarget{power-estimation}{%
-\subsection{Power Estimation}\label{power-estimation}}
-
-It was expected that with larger effect size, the power of the visual test would be greater for both cubic model and heteroskedasticity model.
-Using the power calculation method discussed in Section \ref{p-value-and-power-estimation-of-visual-test-allowed-for-multiple-selections}, power of each lineup from one to five evaluations was estimated.
-The estimated power was further used in fitting a quasi-binomial generalized linear model with natural logarithm of effect size as the only regressor.
-The result of the fitted models are given in Table \ref{tab:powerglmcubic}, Table \ref{tab:powerglmheter}, and Figure \ref{fig:powerglm}.
-
-As expected, coefficients of natural logarithm of effect size of all five models were positive.
-From Figure \ref{fig:powerglm}, it can be observed that the fitted power of visual test increased as the number of evaluations increased for both cubic and heteroskedasticity model.
-
-For heteroskedasticity model, this phenomenon was more obvious as the curves of visual tests with evaluations greater than two were always above the curves of visual test with evaluations smaller than two.
-However, this only held for large enough effect size.
-For small effect size, visual tests with fewer evaluations might have greater power.
-Note that, the expected power of visual test derived by \textcite{majumder_validation_2013} followed by the assumption that human has the ability to select the plot with the highest t-statistic from a lineup under the classical linear model regression setting showed similar properties, where visual tests with fewer evaluations were expected to perform better when the parameter values close to the null hypothesis.
-
-For cubic model, the separation between curves was small.
-Fitted power of visual test with three to five evaluations were almost identical to each other in regards of effect size.
-In addition, all five curves peaked at one as effect size increased, suggesting that identification of nonlinearity as a visual task can be completed reliably by human when the effect size is large enough.
-
-As shown in Figure \ref{fig:powerglm}, both F-test and Breusch--Pagan test generally possessed greater power than visual test.
-Since the settings we were using were not ideal scenarios for visual inference, it was not expected that the visual test would be as good as the conventional test.
-Our goal was to quantify the ability of human reading residual plot, such that the comparison between human and computer vision model can be made.
-
-\begin{table}[!htbp] \centering 
-  \caption{Quasi-binomial logistic regressions with estimated power as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for cubic model.
-Five regressions were fitted for different number of evaluations.}
-  \label{tab:powerglmcubic} 
-\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
-\\[-1.8ex]\hline 
-\hline \\[-1.8ex] 
- & \multicolumn{5}{c}{\textit{Dependent variable:}} \\ 
-\cline{2-6} 
-\\[-1.8ex] & \multicolumn{5}{c}{hat\_power} \\ 
-\\[-1.8ex] & (1) & (2) & (3) & (4) & (5)\\ 
-\hline \\[-1.8ex] 
- log(effect\_size) & 0.929$^{***}$ & 1.227$^{***}$ & 1.084$^{***}$ & 1.057$^{***}$ & 1.035$^{***}$ \\ 
-  & (0.049) & (0.066) & (0.074) & (0.082) & (0.087) \\ 
-  & & & & & \\ 
- Constant & $-$1.969$^{***}$ & $-$2.603$^{***}$ & $-$1.570$^{***}$ & $-$1.501$^{***}$ & $-$1.432$^{***}$ \\ 
-  & (0.114) & (0.153) & (0.142) & (0.156) & (0.160) \\ 
-  & & & & & \\ 
-\hline \\[-1.8ex] 
-Observations & 630 & 532 & 456 & 423 & 371 \\ 
-\hline 
-\hline \\[-1.8ex] 
-\textit{Note:}  & \multicolumn{5}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
-\end{tabular} 
-\end{table}
-
-\begin{table}[!htbp] \centering 
-  \caption{Quasi-binomial logistic regressions with estimated power as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for heteroskedasticity model.
-Five regression were fitted for different number of evaluations.}
-  \label{tab:powerglmheter} 
-\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
-\\[-1.8ex]\hline 
-\hline \\[-1.8ex] 
- & \multicolumn{5}{c}{\textit{Dependent variable:}} \\ 
-\cline{2-6} 
-\\[-1.8ex] & \multicolumn{5}{c}{hat\_power} \\ 
-\\[-1.8ex] & (1) & (2) & (3) & (4) & (5)\\ 
-\hline \\[-1.8ex] 
- log(effect\_size) & 0.821$^{***}$ & 1.025$^{***}$ & 0.885$^{***}$ & 0.920$^{***}$ & 1.063$^{***}$ \\ 
-  & (0.064) & (0.087) & (0.092) & (0.102) & (0.125) \\ 
-  & & & & & \\ 
- Constant & $-$4.472$^{***}$ & $-$5.610$^{***}$ & $-$4.285$^{***}$ & $-$4.127$^{***}$ & $-$4.717$^{***}$ \\ 
-  & (0.320) & (0.449) & (0.461) & (0.499) & (0.607) \\ 
-  & & & & & \\ 
-\hline \\[-1.8ex] 
-Observations & 630 & 527 & 473 & 443 & 375 \\ 
-\hline 
-\hline \\[-1.8ex] 
-\textit{Note:}  & \multicolumn{5}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
-\end{tabular} 
-\end{table}
-
-\begin{figure}
-\centering
-\includegraphics{thesis_files/figure-latex/powerglm-1.pdf}
-\caption{\label{fig:powerglm}Quasi-binomial logistic regressions with estimated power as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for both cubic and heteroskedasticity model.
-Five regression were fitted for different number of evaluations shown in five different colours.
-Estimated power are drawn in black.
-The dashed curvses are the power of F-test and Breusch--Pagan test respectively.}
-\end{figure}
-
-Another method we used to model the power was to collect all decisions made by \(K\choose{K-j}\) different outcomes for \(V_{K-j}\), and directly fit a binomial generalized linear model with natural logarithm of effect size as the regressor.
-This method skipped the calculation of the estimated power \(R/{K\choose{K-j}}\), and regressed on the binary outcome with larger sample size.
-Results of the fitted model is given in Figure \ref{fig:powerglmsecond}, Table \ref{tab:powerglmhetersecond} and Table \ref{tab:powerglmcubicsecond}.
-According to Figure \ref{fig:powerglmsecond}, this method gave clearer separation between curves for heteroskedasticity model.
-Noticeably, fitted power of visual test with five evaluations matched the power of conventional test at large effect size.
-
-\begin{table}[!htbp] \centering 
-  \caption{Logistic regressions with rejection as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for cubic model.
-Five regression were fitted for different number of evaluations.}
-  \label{tab:powerglmcubicsecond} 
-\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
-\\[-1.8ex]\hline 
-\hline \\[-1.8ex] 
- & \multicolumn{5}{c}{\textit{Dependent variable:}} \\ 
-\cline{2-6} 
-\\[-1.8ex] & \multicolumn{5}{c}{hat\_power} \\ 
-\\[-1.8ex] & (1) & (2) & (3) & (4) & (5)\\ 
-\hline \\[-1.8ex] 
- log(effect\_size) & 0.872$^{***}$ & 1.292$^{***}$ & 1.125$^{***}$ & 1.164$^{***}$ & 1.315$^{***}$ \\ 
-  & (0.031) & (0.032) & (0.021) & (0.021) & (0.029) \\ 
-  & & & & & \\ 
- Constant & $-$1.836$^{***}$ & $-$2.529$^{***}$ & $-$1.516$^{***}$ & $-$1.321$^{***}$ & $-$1.419$^{***}$ \\ 
-  & (0.077) & (0.073) & (0.046) & (0.049) & (0.072) \\ 
-  & & & & & \\ 
-\hline \\[-1.8ex] 
-Observations & 2,900 & 7,250 & 12,596 & 17,011 & 19,229 \\ 
-Log Likelihood & $-$752.896 & $-$1,312.436 & $-$2,977.445 & $-$2,651.299 & $-$1,463.555 \\ 
-Akaike Inf. Crit. & 1,509.792 & 2,628.873 & 5,958.889 & 5,306.598 & 2,931.109 \\ 
-\hline 
-\hline \\[-1.8ex] 
-\textit{Note:}  & \multicolumn{5}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
-\end{tabular} 
-\end{table}
-
-\begin{table}[!htbp] \centering 
-  \caption{Logistic regressions with rejection as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for heteroskedasticity model.
-Five regression were fitted for different number of evaluations.}
-  \label{tab:powerglmhetersecond} 
-\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
-\\[-1.8ex]\hline 
-\hline \\[-1.8ex] 
- & \multicolumn{5}{c}{\textit{Dependent variable:}} \\ 
-\cline{2-6} 
-\\[-1.8ex] & \multicolumn{5}{c}{hat\_power} \\ 
-\\[-1.8ex] & (1) & (2) & (3) & (4) & (5)\\ 
-\hline \\[-1.8ex] 
- log(effect\_size) & 0.912$^{***}$ & 1.221$^{***}$ & 1.294$^{***}$ & 1.525$^{***}$ & 1.798$^{***}$ \\ 
-  & (0.044) & (0.033) & (0.027) & (0.030) & (0.042) \\ 
-  & & & & & \\ 
- Constant & $-$4.846$^{***}$ & $-$6.396$^{***}$ & $-$6.033$^{***}$ & $-$6.771$^{***}$ & $-$7.784$^{***}$ \\ 
-  & (0.229) & (0.175) & (0.141) & (0.159) & (0.230) \\ 
-  & & & & & \\ 
-\hline \\[-1.8ex] 
-Observations & 2,906 & 7,064 & 11,436 & 13,728 & 13,236 \\ 
-Log Likelihood & $-$1,525.398 & $-$3,668.325 & $-$5,491.915 & $-$4,828.896 & $-$2,858.774 \\ 
-Akaike Inf. Crit. & 3,054.796 & 7,340.650 & 10,987.830 & 9,661.793 & 5,721.548 \\ 
-\hline 
-\hline \\[-1.8ex] 
-\textit{Note:}  & \multicolumn{5}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
-\end{tabular} 
-\end{table}
-
-\begin{figure}
-\centering
-\includegraphics{thesis_files/figure-latex/powerglmsecond-1.pdf}
-\caption{\label{fig:powerglmsecond}Logistic regressions with rejection as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for both cubic and heteroskedasticity model.
-Five regression were fitted for different number of evaluations shown in five different colours.
-Estimated power are drawn in black.
-The dashed curvses are the power of F-test and Breusch--Pagan test respectively.}
-\end{figure}
-
-\hypertarget{effect-of-factors-on-power-of-the-visual-test}{%
-\subsection{Effect of Factors on Power of the Visual Test}\label{effect-of-factors-on-power-of-the-visual-test}}
-
-The previous section focuses on the change of effect size relative to the power of the visual test.
-Individual factor embedded in the data simulation also needs be analysed.
-Two logistic regressions were fitted with decisions made by \(V_{K-j}\) as response variable.
-For the cubic model, regressors were \(|a|\), \(|b|\), \(|c-1|\), \(\sigma\), distribution of \(\boldsymbol{X}\), distribution of \(\boldsymbol{Z}\), number of observations and number of evaluations.
-For the heteroskedasticity model, regressors were \(|a|\), \(b\), distribution of \(\boldsymbol{X}\), number of observations and number of evaluations.
-
-According to Table \ref{tab:rejectfactorcubic}, for cubic model, both coefficients of \(|a|\) and \(|b|\) were positive, which suggested increasing the coefficients of quadric and cubic terms would result in higher visual power.
-The coefficient of \(|c-1|\) was also positive indicating joint effect of multiple regressors on the residual plot decreased the visual power.
-As expected, the coefficient of \(\sigma\) was negative as it controlled the noise level.
-Lineup simulated with uniform distribution of \(\boldsymbol{X}\) had the highest power, followed by negative lognormal, normal, and lognormal.
-Discreteness in \(\boldsymbol{Z}\) decreased the visual power.
-The coefficient of number of observations was positive, which means larger sample size would lead to higher visual power.
-
-Results of logistic regression for heteroskedasticity is given in Table \ref{tab:rejectfactorheter}.
-For heteroskedasticity model, the coefficient of \(|a|\) was negative, which suggested ``butter fly'' shape was easier to identify than ``triangle'' shape.
-As expected, the coefficient of \(b\) was positive.
-For different distributions of \(\boldsymbol{X}\), the one with highest power was uniform distribution, followed by normal, discrete uniform, lognormal and negative lognormal.
-The coeffient of the number of observations was positive, but it was not significant.
-
-In overall, we did not have any unexpected findings except the visual power related to different distributions of \(\boldsymbol{X}\) in cubic model.
-We expected lineup simulated with uniform distribution would have the highest power, followed by normal, lognormal and negative lognormal, since we observed similar results in pilot study.
-
-\begin{table}[!htbp] \centering 
-  \caption{Logistic regressions with rejection as response variable and parameter settings as regressors fitted on the data collected from experiment I and II for cubic model.} 
-  \label{tab:rejectfactorcubic} 
-\begin{tabular}{@{\extracolsep{5pt}}lc} 
-\\[-1.8ex]\hline 
-\hline \\[-1.8ex] 
- & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
-\cline{2-2} 
-\\[-1.8ex] & reject \\ 
-\hline \\[-1.8ex] 
- |a| & 1.064$^{***}$ \\ 
-  & (0.031) \\ 
-  & \\ 
- |b| & 0.718$^{***}$ \\ 
-  & (0.029) \\ 
-  & \\ 
- |c-1| & 5.443$^{***}$ \\ 
-  & (0.119) \\ 
-  & \\ 
- sigma & $-$3.536$^{***}$ \\ 
-  & (0.050) \\ 
-  & \\ 
- Distribution of X: Negative Lognormal & 0.971$^{***}$ \\ 
-  & (0.073) \\ 
-  & \\ 
- Distribution of X: Normal & 0.754$^{***}$ \\ 
-  & (0.086) \\ 
-  & \\ 
- Distribution of X: Uniform & 1.751$^{***}$ \\ 
-  & (0.085) \\ 
-  & \\ 
- Distribution of Z: Discrete & $-$2.986$^{***}$ \\ 
-  & (0.056) \\ 
-  & \\ 
- n & 0.005$^{***}$ \\ 
-  & (0.0002) \\ 
-  & \\ 
- Number of evaluations & 0.291$^{***}$ \\ 
-  & (0.017) \\ 
-  & \\ 
- Constant & $-$2.453$^{***}$ \\ 
-  & (0.173) \\ 
-  & \\ 
-\hline \\[-1.8ex] 
-Observations & 58,986 \\ 
-Log Likelihood & $-$8,034.351 \\ 
-Akaike Inf. Crit. & 16,090.700 \\ 
-\hline 
-\hline \\[-1.8ex] 
-\textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
-\end{tabular} 
-\end{table}
-
-\begin{table}[!htbp] \centering 
-  \caption{Logistic regressions with rejection as response variable and parameter settings as regressors fitted on the data collected from experiment I and II for heteroskedasticity model.} 
-  \label{tab:rejectfactorheter} 
-\begin{tabular}{@{\extracolsep{5pt}}lc} 
-\\[-1.8ex]\hline 
-\hline \\[-1.8ex] 
- & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
-\cline{2-2} 
-\\[-1.8ex] & reject \\ 
-\hline \\[-1.8ex] 
- |a| & $-$0.526$^{***}$ \\ 
-  & (0.037) \\ 
-  & \\ 
- b & 0.064$^{***}$ \\ 
-  & (0.002) \\ 
-  & \\ 
- Distribution of X: Lognormal & $-$0.534$^{***}$ \\ 
-  & (0.058) \\ 
-  & \\ 
- Distribution of X: Negative Lognormal & $-$0.610$^{***}$ \\ 
-  & (0.062) \\ 
-  & \\ 
- Distribution of X: Normal & 0.187$^{***}$ \\ 
-  & (0.063) \\ 
-  & \\ 
- Distribution of X: Uniform & 2.980$^{***}$ \\ 
-  & (0.064) \\ 
-  & \\ 
- n & 0.0003 \\ 
-  & (0.0002) \\ 
-  & \\ 
- Number of evaluations & 0.496$^{***}$ \\ 
-  & (0.014) \\ 
-  & \\ 
- Constant & $-$2.711$^{***}$ \\ 
-  & (0.062) \\ 
-  & \\ 
-\hline \\[-1.8ex] 
-Observations & 48,370 \\ 
-Log Likelihood & $-$13,026.190 \\ 
-Akaike Inf. Crit. & 26,070.390 \\ 
-\hline 
-\hline \\[-1.8ex] 
-\textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
-\end{tabular} 
-\end{table}
-
-\hypertarget{computer-vision-model}{%
-\chapter{Computer Vision Model}\label{computer-vision-model}}
-
-Visual test can be used to make valid inference on visual discoveries and acknowledging people with confidence about the findings.
-However, the lack of scalability of this method due to relatively high human cost compared to conventional test limits its application in wider fields.
-To relieve human analysts of the job of diagnosing predictive models in the context of visual inference, AI agents which can read data plot, evaluate lineups, and provide decisions need to be built.
-
-Considering there are different approches of AI discussed in \ref{se:ai}, it is necessary to clarify the type of AI we aimed to build.
-Understanding the human thought process or mimicking the human vision mechanism were not of interest in this research.
-We believed humanly behaviours was enough for demonstrating AI for the problem we concerned.
-Hence, two of the capacities required by the total Turing test, namely machine learning and computer vision will be adopted.
-
-Another possible method was to define distance metrics to measure difference between data plots for making mathematically prefect decisions.
-This method would work if such metrics can be obtained manually or approximated algorithmically.
-However, we considered it was too complex as a starting point.
-
-\hypertarget{automatic-visual-inference-system}{%
-\section{Automatic visual inference system}\label{automatic-visual-inference-system}}
-
-For an automatic visual inference system, there are several possibilities of the type of inputs.
-
-Depend on the inputs and outputs, computer vision model can be integrated into an automatic visual inference system.
-
 \hypertarget{different-approaches-of-ai}{%
@@
 \hypertarget{model-architecture}{%
-\section{Model Architecture}\label{model-architecture}}
+\subsection{Model Architecture}\label{model-architecture}}
 
@@
 \hypertarget{performance}{%
-\section{Performance}\label{performance}}
+\subsection{Performance}\label{performance}}
 
@@
 
-\appendix
-
-\hypertarget{additional-stuff}{%
-\chapter{Additional stuff}\label{additional-stuff}}
-
-You might put some computer output here, or maybe additional tables.
-
-Note that line 5 must appear before your first appendix. But other appendices can just start like any other chapter.
+\hypertarget{timeline-need-discussion}{%
+\chapter{Timeline (need discussion)}\label{timeline-need-discussion}}
 

### 2022-03-07 Milestone_1/index.Rmd
add milestone report

--- 2022-01-26 Thesis/index.Rmd
@@
 title: 'Advances in Artificial Intelligence for Data Visualization: Developing Computer Vision Models to Automate Reading of Data Plots, with Application to Predictive Model Diagnostics'
-degreetype: 'Doctor of Philosophy'
+degreetype: 'PhD'
 author: 'Weihao Li'

### 2022-03-08 Milestone_1/01-chap1.Rmd
finish intro and timeline

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo=FALSE, message=FALSE, cache=TRUE)
+knitr::opts_chunk$set(echo=FALSE, message=FALSE, cache=TRUE, warning = FALSE)
 # Load any R packages you need here
-library(ggplot2)
+library(tidyverse)
+library(visage)
 ```
@@
 # Background and Motivation
+
+The introduction of the report is organised as follows.
+Section \ref{se:predictive-modelling} discusses the predictive modelling culture and the use of data plots in predictive model diagnostics.
+Section \ref{se:vi} outlines the definition and terminology of visual inference.
+Section \ref{se:prespecification} presents the justification of the validity of visual inference.
+Section \ref{se:sampling-from-null} describes the sampling techniques in visual inference with an example of the classical normal linear regression model.
+Section \ref{se:lineup} introduces the lineup protocol and the estimation of $p$-value and power of a visual test.
+And lastly, Section \ref{se:limitation} points out the limitation of lineup protocol, which is also the motivation of this research, and the need of building an automatic visual inference system.
 
@@
 
-## Predictive Modelling and Visual Diagnostics
+## Predictive Modelling and Visual Diagnostics {#se:predictive-modelling}
 
@@
 
-## Visual Inference
+## Visual Inference {#se:vi}
 
@@
 
-## Pre-specification of Visual Discoverable Features
+## Pre-specification of Visual Discoverable Features {#se:prespecification}
 
@@
 Alternatively, @majumder_validation_2013 avoids the use of visual features and defined the visual statistics $T(.)$ as a mapping from a dataset to a data plot.
-Both definitions of visual test statistics are valid, but in the rest of the paper the first definition will be used as it covers some details needed by the following discussion.
+Both definitions of visual test statistics are valid, but in the rest of the report the first definition will be used as it covers some details needed by the following discussion.
 
@@
 
-## Sampling from the Null Distribution
+## Sampling from the Null Distribution {#se:sampling-from-null}
 
@@
 The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009].
-Focus will be put on simulation based on this distribution as it is applicable to classical normal linear regression model.
-Details of the other two are can be found in the supplementary material of @buja_statistical_2009.
 Suppose there exists a minimal sufficient statistic $\boldsymbol{S}(\boldsymbol{y})$ under the null hypothesis, any null datasets $\boldsymbol{y^{*}}$ should fulfil the condition $\boldsymbol{S}(\boldsymbol{y}) = \boldsymbol{s}$.
-For classical normal linear regression model, the minimal sufficient statistic is $\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{\beta}}, \boldsymbol{e}'\boldsymbol{e})$, where $\hat{\boldsymbol{\beta}}$ are the coefficient estimators and $\boldsymbol{e}'\boldsymbol{e}$ is the residual sum of square.
+Using the classical linear regression model as example, the minimal sufficient statistic is $\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{\beta}}, \boldsymbol{e}'\boldsymbol{e})$, where $\hat{\boldsymbol{\beta}}$ are the coefficient estimators and $\boldsymbol{e}'\boldsymbol{e}$ is the residual sum of square.
 Alternatively, the minimal sufficient statistic can be constructed as $\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{y}}, ||\boldsymbol{e}||)$, where $\hat{\boldsymbol{y}}$ are the fitted values and $||\boldsymbol{e}||$ is the length of residuals, which is more intuitive as suggested by @buja_statistical_2009.
@@
 
-## Lineup Protocol
+## Lineup Protocol {#se:lineup}
 
@@
 
-## Limitations of Lineup Protocol and Automatic Visual Inference
+## Limitations of Lineup Protocol and Automatic Visual Inference {#se:limitation}
 
@@
 
-The "steam engine" of visual inference needs to developed for relieving people's workload by automating repeating tasks and providing standard result in a control environment.
+The "steam engine" of visual inference needs to be developed for relieving people's workload by automating repeating tasks and providing standard result in a control environment.
 Large-scale evaluation of visual tests is not possible without the use of technology and machines.
@@
 This should encompass a description of the theoretical and conceptual framework that underlies the thesis, and the procedures that are to be used in addressing the research questions
+
+The
+
+The first project is
 
@@
 null_true_eval_dat %>%
-  group_by(model, null) %>%
+  mutate(type = ifelse(id <= 5000, "Cubic", "Heteroskedasticity")) %>% 
+  group_by(model, type, null) %>%
   summarise(rate = mean(prediction == null)) %>%
@@
   ungroup() %>%
-  mutate(model = model_name) %>%
   mutate(avg_rate = (`FALSE` + `TRUE`)/2) %>%
-  knitr::kable(col.names = c("Model", "Actual data plot correct rate", "Null plot correct rate", "Average correct rate"), 
+  select(-model) %>%
+  knitr::kable(col.names = c("", "Actual data plot detection rate", "Null plot detection rate", "Average detection rate"), 
              booktabs = TRUE, 
              caption  = 'Performance of computer vision models on null residual plot and acutal data plot.', 
-             label = "null-true-eval" ) 
+             label = "null-true-eval" ) %>%
+  kableExtra::kable_styling(latex_options = "scale_down") %>%
+  kableExtra::pack_rows("VGG16", 1, 2, bold = TRUE, italic = TRUE, underline = TRUE) %>%
+  kableExtra::pack_rows("VGG19", 3, 4, bold = TRUE, italic = TRUE, underline = TRUE) %>%
+  kableExtra::pack_rows("Xception", 5, 6, bold = TRUE, italic = TRUE, underline = TRUE) %>%
+  kableExtra::pack_rows("ResNet152V2", 7, 8, bold = TRUE, italic = TRUE, underline = TRUE) %>%
+  kableExtra::pack_rows("InceptionV3", 9, 10, bold = TRUE, italic = TRUE, underline = TRUE)
 ```
@@
 ```
+
 ```{r}
@@
 
-# Timeline (need discussion)
+```{r}
+lineup_eval_dat %>%
+  left_join(eval_lineup_ans, by = c("set" = "id", "plot" = "k")) %>%
+  filter(ans == TRUE) %>%
+  group_by(set, model) %>%
+  summarise(pseudo_predict = sample(predict, 1)) -> pseudo_predict_dat
+
+lineup_eval_dat %>%
+  left_join(eval_lineup_ans, by = c("set" = "id", "plot" = "k")) %>%
+  left_join(pseudo_predict_dat) %>%
+  mutate(predict = ifelse(ans == FALSE, pseudo_predict, predict)) %>%
+  group_by(model, set) %>%
+  summarise(n_sel = sum(!predict), detect = sum(ans == FALSE & predict == FALSE))  -> lineup_pseudo_summary
+
+lineup_pseudo_summary %>%
+  calc_p_value_multi(lineup_id = "set", detected = "detect", n_sel = "n_sel", cache_env = cache_env) -> pseudo_results
+
+pseudo_results %>%
+  mutate(reject = eval_p_value(p_value, tol = 1e-2)) %>%
+  mutate(type = ifelse(lineup_id <= 2500, "Cubic", "Heteroskedasticity")) %>%
+  group_by(type) %>%
+  summarise(rate = mean(reject))
+```
+
+# Timeline
+
+\begin{table} \label{tab:timeline}
+\centering
+\begin{tabular}[t]{ll}
+\toprule
+\addlinespace[0.5em]
+Date & Event\\
+\addlinespace[0.5em]
+\midrule
+\addlinespace[0.5em]
+\multicolumn{2}{l}{\textit{\textbf{2022}}}\\
+\hspace{4em}March & Extend the scope of violations of classical normal linear\\ 
+\addlinespace[0.5em]
+\hspace{4em} & regression considered by the computer vision model\\
+\addlinespace[0.5em]
+\hspace{4em}April & Conduct human subject experiment on the extended model\\ 
+\addlinespace[0.5em]
+\hspace{4em} & defects\\
+\addlinespace[0.5em]
+\hspace{4em}May & Conclude the first project: draft and submit the first paper\\
+\addlinespace[0.5em]
+\hspace{4em}June & Review previous studies in image comparison\\
+\addlinespace[0.5em]
+\hspace{4em}July & Propose a new design of the automatic visual inference\\
+\addlinespace[0.5em]
+\hspace{4em} & system based on image comparison\\
+\addlinespace[0.5em]
+\hspace{4em}August - November & Develop the proposed system\\
+\addlinespace[0.5em]
+\hspace{4em}November - December & Conduct human subject experiment for validating the\\
+\addlinespace[0.5em]
+\hspace{4em} & proposed system\\
+\addlinespace[0.5em]
+\hspace{4em}December & Attend the Australasian Applied Statistics Conference\\
+\addlinespace[0.5em]
+\hline
+\addlinespace[0.5em]
+\multicolumn{2}{l}{\textit{\textbf{2023}}}\\
+\hspace{4em}January - March & Finish the second project: prepare milestone material and\\
+\addlinespace[0.5em]
+\hspace{4em} & submit the second paper\\
+\addlinespace[0.5em]
+\hspace{4em}April - December & Apply the automatic visual inference system to predictive\\
+\addlinespace[0.5em]
+\hspace{4em} & model diagnostics\\
+\addlinespace[0.5em]
+\hline
+\addlinespace[0.5em]
+\multicolumn{2}{l}{\textit{\textbf{2024}}}\\
+\hspace{4em}January - March & Finish the third project: finalize the thesis\\
+\addlinespace[0.5em]
+\bottomrule
+\end{tabular}
+\end{table}

### 2022-03-08 Milestone_1/_book/thesis.tex
finish intro and timeline

@@
 
-\hypertarget{predictive-modelling-and-visual-diagnostics}{%
-\section{Predictive Modelling and Visual Diagnostics}\label{predictive-modelling-and-visual-diagnostics}}
+The introduction of the report is organised as follows.
+Section \ref{se:predictive-modelling} discusses the predictive modelling culture and the use of data plots in predictive model diagnostics.
+Section \ref{se:vi} outlines the definition and terminology of visual inference.
+Section \ref{se:prespecification} presents the justification of the validity of visual inference.
+Section \ref{se:sampling-from-null} describes the sampling techniques in visual inference with an example of the classical normal linear regression model.
+Section \ref{se:lineup} introduces the lineup protocol and the estimation of \(p\)-value and power of a visual test.
+And lastly, Section \ref{se:limitation} points out the limitation of lineup protocol, which is also the motivation of this research, and the need of building an automatic visual inference system.
+
+\hypertarget{se:predictive-modelling}{%
+\section{Predictive Modelling and Visual Diagnostics}\label{se:predictive-modelling}}
 
@@
 
-\hypertarget{visual-inference}{%
-\section{Visual Inference}\label{visual-inference}}
+\hypertarget{se:vi}{%
+\section{Visual Inference}\label{se:vi}}
 
@@
 
-\hypertarget{pre-specification-of-visual-discoverable-features}{%
-\section{Pre-specification of Visual Discoverable Features}\label{pre-specification-of-visual-discoverable-features}}
+\hypertarget{se:prespecification}{%
+\section{Pre-specification of Visual Discoverable Features}\label{se:prespecification}}
 
@@
 Alternatively, \textcite{majumder_validation_2013} avoids the use of visual features and defined the visual statistics \(T(.)\) as a mapping from a dataset to a data plot.
-Both definitions of visual test statistics are valid, but in the rest of the paper the first definition will be used as it covers some details needed by the following discussion.
+Both definitions of visual test statistics are valid, but in the rest of the report the first definition will be used as it covers some details needed by the following discussion.
 
@@
 
-\hypertarget{sampling-from-the-null-distribution}{%
-\section{Sampling from the Null Distribution}\label{sampling-from-the-null-distribution}}
+\hypertarget{se:sampling-from-null}{%
+\section{Sampling from the Null Distribution}\label{se:sampling-from-null}}
 
@@
 The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three \autocite{buja_statistical_2009}.
-Focus will be put on simulation based on this distribution as it is applicable to classical normal linear regression model.
-Details of the other two are can be found in the supplementary material of \textcite{buja_statistical_2009}.
 Suppose there exists a minimal sufficient statistic \(\boldsymbol{S}(\boldsymbol{y})\) under the null hypothesis, any null datasets \(\boldsymbol{y^{*}}\) should fulfil the condition \(\boldsymbol{S}(\boldsymbol{y}) = \boldsymbol{s}\).
-For classical normal linear regression model, the minimal sufficient statistic is \(\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{\beta}}, \boldsymbol{e}'\boldsymbol{e})\), where \(\hat{\boldsymbol{\beta}}\) are the coefficient estimators and \(\boldsymbol{e}'\boldsymbol{e}\) is the residual sum of square.
+Using the classical linear regression model as example, the minimal sufficient statistic is \(\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{\beta}}, \boldsymbol{e}'\boldsymbol{e})\), where \(\hat{\boldsymbol{\beta}}\) are the coefficient estimators and \(\boldsymbol{e}'\boldsymbol{e}\) is the residual sum of square.
 Alternatively, the minimal sufficient statistic can be constructed as \(\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{y}}, ||\boldsymbol{e}||)\), where \(\hat{\boldsymbol{y}}\) are the fitted values and \(||\boldsymbol{e}||\) is the length of residuals, which is more intuitive as suggested by \textcite{buja_statistical_2009}.
@@
 
-\hypertarget{lineup-protocol}{%
-\section{Lineup Protocol}\label{lineup-protocol}}
+\hypertarget{se:lineup}{%
+\section{Lineup Protocol}\label{se:lineup}}
 
@@
 
-\hypertarget{limitations-of-lineup-protocol-and-automatic-visual-inference}{%
-\section{Limitations of Lineup Protocol and Automatic Visual Inference}\label{limitations-of-lineup-protocol-and-automatic-visual-inference}}
+\hypertarget{se:limitation}{%
+\section{Limitations of Lineup Protocol and Automatic Visual Inference}\label{se:limitation}}
 
@@
 
-The ``steam engine'' of visual inference needs to developed for relieving people's workload by automating repeating tasks and providing standard result in a control environment.
+The ``steam engine'' of visual inference needs to be developed for relieving people's workload by automating repeating tasks and providing standard result in a control environment.
 Large-scale evaluation of visual tests is not possible without the use of technology and machines.
@@
 ##            y          x    sigma          e     .resid    .fitted
-## 1 7.205499  0.9053470 5.218132 5.3001525  0.7420902 6.463409
-## 2 2.560696 -0.8626811 4.981465 2.4233771  1.0441078 1.516588
-## 3 1.121106 -0.6545505 3.835357 0.7756568 -0.9778170 2.098923
-## 4 5.296467  0.7171334 4.178154 3.5793335 -0.6403337 5.936801
-## 5 2.962130 -0.2859733 1.901837 2.2481030 -0.1680473 3.130177
+## 1 -7.6931740  0.8335218 4.819987 -9.5266958 -5.7469121 -1.9462619
+## 2  7.0213194 -0.7458772 4.336202  6.7671966 -0.9130225  7.9343418
+## 3 -2.1856528  0.9509873 5.471751 -4.1366401  0.4954643 -2.6811172
+## 4  4.0402811  0.4961018 2.979219  2.5441792  3.8756685  0.1646126
+## 5  0.8686919  0.7494172 4.355690 -0.8807253  2.2888017 -1.4201098
 \end{verbatim}
@@
 ##       0       1       2       3 
-## 0.70182 0.25202 0.04346 0.00270
+## 0.70094 0.25280 0.04340 0.00286
 \end{verbatim}
@@
 \begin{verbatim}
-## [1] 0.04688
+## [1] 0.0456
 \end{verbatim}
@@
 \centering
+\resizebox{\linewidth}{!}{
 \begin{tabular}[t]{lrrr}
 \toprule
-Model & Actual data plot correct rate & Null plot correct rate & Average correct rate\\
+ & Actual data plot detection rate & Null plot detection rate & Average detection rate\\
 \midrule
-VGG16 & 0.7387 & 0.8850 & 0.81185\\
-VGG19 & 0.8029 & 0.8313 & 0.81710\\
-Xception & 0.7370 & 0.9213 & 0.82915\\
-ResNet152V2 & 0.6932 & 0.9233 & 0.80825\\
-InceptionV3 & 0.6691 & 0.9399 & 0.80450\\
+\addlinespace[0.3em]
+\multicolumn{4}{l}{\underline{\textit{\textbf{VGG16}}}}\\
+\hspace{1em}Cubic & 0.7752 & 0.8662 & 0.8207\\
+\hspace{1em}Heteroskedasticity & 0.7022 & 0.9038 & 0.8030\\
+\addlinespace[0.3em]
+\multicolumn{4}{l}{\underline{\textit{\textbf{VGG19}}}}\\
+\hspace{1em}Cubic & 0.8006 & 0.8210 & 0.8108\\
+\hspace{1em}Heteroskedasticity & 0.8052 & 0.8416 & 0.8234\\
+\addlinespace[0.3em]
+\multicolumn{4}{l}{\underline{\textit{\textbf{Xception}}}}\\
+\hspace{1em}Cubic & 0.7442 & 0.9182 & 0.8312\\
+\hspace{1em}Heteroskedasticity & 0.7298 & 0.9244 & 0.8271\\
+\addlinespace[0.3em]
+\multicolumn{4}{l}{\underline{\textit{\textbf{ResNet152V2}}}}\\
+\hspace{1em}Cubic & 0.7252 & 0.9168 & 0.8210\\
+\hspace{1em}Heteroskedasticity & 0.6612 & 0.9298 & 0.7955\\
+\addlinespace[0.3em]
+\multicolumn{4}{l}{\underline{\textit{\textbf{InceptionV3}}}}\\
+\hspace{1em}Cubic & 0.6984 & 0.9304 & 0.8144\\
+\hspace{1em}Heteroskedasticity & 0.6398 & 0.9494 & 0.7946\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\includegraphics{thesis_files/figure-latex/unnamed-chunk-44-1.pdf}
+
+\begin{verbatim}
+## # A tibble: 2 x 2
+##   type                 rate
+##   <chr>               <dbl>
+## 1 Cubic              0.0056
+## 2 Heteroskedasticity 0.006
+\end{verbatim}
+
+\hypertarget{timeline}{%
+\chapter{Timeline}\label{timeline}}
+
+\begin{table} \label{tab:timeline}
+\centering
+\begin{tabular}[t]{ll}
+\toprule
+\addlinespace[0.5em]
+Date & Event\\
+\addlinespace[0.5em]
+\midrule
+\addlinespace[0.5em]
+\multicolumn{2}{l}{\textit{\textbf{2022}}}\\
+\hspace{4em}March & Extend the scope of violations of classical normal linear\\ 
+\addlinespace[0.5em]
+\hspace{4em} & regression considered by the computer vision model\\
+\addlinespace[0.5em]
+\hspace{4em}April & Conduct human subject experiment on the extended model\\ 
+\addlinespace[0.5em]
+\hspace{4em} & defects\\
+\addlinespace[0.5em]
+\hspace{4em}May & Conclude the first project: draft and submit the first paper\\
+\addlinespace[0.5em]
+\hspace{4em}June & Review previous studies in image comparison\\
+\addlinespace[0.5em]
+\hspace{4em}July & Propose a new design of the automatic visual inference\\
+\addlinespace[0.5em]
+\hspace{4em} & system based on image comparison\\
+\addlinespace[0.5em]
+\hspace{4em}August - November & Develop the proposed system\\
+\addlinespace[0.5em]
+\hspace{4em}November - December & Conduct human subject experiment for validating the\\
+\addlinespace[0.5em]
+\hspace{4em} & proposed system\\
+\addlinespace[0.5em]
+\hspace{4em}December & Attend the Australasian Applied Statistics Conference\\
+\addlinespace[0.5em]
+\hline
+\addlinespace[0.5em]
+\multicolumn{2}{l}{\textit{\textbf{2023}}}\\
+\hspace{4em}January - March & Finish the second project: prepare milestone material and\\
+\addlinespace[0.5em]
+\hspace{4em} & submit the second paper\\
+\addlinespace[0.5em]
+\hspace{4em}April - December & Apply the automatic visual inference system to predictive\\
+\addlinespace[0.5em]
+\hspace{4em} & model diagnostics\\
+\addlinespace[0.5em]
+\hline
+\addlinespace[0.5em]
+\multicolumn{2}{l}{\textit{\textbf{2024}}}\\
+\hspace{4em}January - March & Finish the third project: finalize the thesis\\
+\addlinespace[0.5em]
 \bottomrule
@@
 
-\includegraphics{thesis_files/figure-latex/unnamed-chunk-44-1.pdf}
-
-\hypertarget{timeline-need-discussion}{%
-\chapter{Timeline (need discussion)}\label{timeline-need-discussion}}
-
 \printbibliography[heading=bibintoc]

### 2022-03-08 Milestone_1/01-chap1.Rmd
add research problem

@@
 
-# Structure of the Thesis (need disscussion)
-
-An overview of the thesis, including some brief detail of the proposed content of all chapters.
-This should encompass a description of the theoretical and conceptual framework that underlies the thesis, and the procedures that are to be used in addressing the research questions
-
-The
-
-The first project is
-
-1. Automatic Visual Statistical Inference, with Application to Linear Regression Diagnostics
-2. ...
-3. ...
+# Research Problem
+
+<!-- 1.
+Provide visual diagnostics for projection pursuit optimisation to examine whether the optimisers are able to find the optima, compare optimisation paths, inspect optimisers’ coverage of the space, and detect anomalies in the algorithms.
+-->
+
+<!-- 2.
+Extend and enhance the visual diagnostic tools developed in the first project to stochastic optimal control problems, which is a multistage optimisation problem with more than two exogenous variables.
+-->
+
+<!-- 3.
+Improve the interpretability and explainability of stochastic optimal control algo- rithms by identifying the most important drivers in the optimal decisions suggested by the optimisers and studying how the decision changes across time.
+-->
+
+The main objective of this research is to build an automatic visual inference system for the purpose of conducting large-scale visual test. The research will focus on three projects as follows:
+
+1.
+Propose a prototype of automatic visual inference system for evaluating lineups of residual plots of the classical normal linear regression model and study factors that affect the performance of the system with comparison to human subjects.
+
+2. Extend and enhance the automatic visual inference system proposed in the first project to general lineup protocol problems by adopting image similarity assessed by computer vision models.
+
+3. Validate and apply the automatic visual inference system to predictive model diagnostics on empirical data.
+
+# Overview of the Thesis
+
+1.
+The first project provides a prototype of automatic visual inference system for evaluating lineups of residual plots, with a focus on the classical normal linear regression model given it is one of the simplest predictive models.
+The core of the system
+
+Validation of the automatic system is made by comparing the power of human subject-assessed visual tests with the system-assessed visual tests.
+
+2. The second project aims to develop an automatic visual inference system based on image comparison.
+
+This project focused on building a prototype of automatic visual statistical inference system for evaluating residual plots of classical normal linear regression model.
+To set up a comparison between the computer vision models and humans, human subject experiments were conducted to understand the ability of human reading residual plots.
+
+The first project provides visual diagnostics for projection pursuit optimisation, with a focus on the guided tour.
+A guided tour is constructed by combining the optimisation, for choosing new planes to visit, with a geodesic interpolation between planes to provide a smooth movie of projections for visualisation.
+The visual diagnostics include static graphics and animations to understand different aspects of the optimisers: its progression on improving the index value, its coverage of the search space and the comparison of different optimisers in different scenarios.
+What is learned will be used to modify and improve the performance of the optimisers.
+
+The second project aims to develop visual diagnostic tools for stochastic optimal control algorithms.
+In comparison to projection pursuit, optimisation in stochastic optimal control problems finds a sequence of discrete-valued decisions that are optimal at each time point.
+Exogenous variables that affect the decisions are also involved in the optimisation.
+The project will extend and build on the visual diagnostic tools developed in the first project to address stochastic optimal control optimisation.
+
+The third project aims to use visual techniques to understand variable importance in stochastic optimal control problem.
+The project will study how to interpret the contribution of potentially dozens of variables to the decision at a particular time and how the future decision would change with a change in the exogenous variables.
 

### 2022-03-08 Milestone_1/_book/thesis.tex
add research problem

@@
 
-\hypertarget{structure-of-the-thesis-need-disscussion}{%
-\chapter{Structure of the Thesis (need disscussion)}\label{structure-of-the-thesis-need-disscussion}}
-
-An overview of the thesis, including some brief detail of the proposed content of all chapters.
-This should encompass a description of the theoretical and conceptual framework that underlies the thesis, and the procedures that are to be used in addressing the research questions
-
-The
-
-The first project is
+\hypertarget{research-problem}{%
+\chapter{Research Problem}\label{research-problem}}
+
+The main objective of this research is to build an automatic visual inference system for the purpose of conducting large-scale visual test. The research will focus on three projects as follows:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  Propose a prototype of automatic visual inference system for evaluating lineups of residual plots of the classical normal linear regression model and study factors that affect the performance of the system with comparison to human subjects.
+\item
+  Extend and enhance the automatic visual inference system proposed in the first project to general lineup protocol problems by adopting image similarity assessed by computer vision models.
+\item
+  Validate and apply the automatic visual inference system to predictive model diagnostics on empirical data.
+\end{enumerate}
+
+\hypertarget{overview-of-the-thesis}{%
+\chapter{Overview of the Thesis}\label{overview-of-the-thesis}}
 
@@
 \item
-  Automatic Visual Statistical Inference, with Application to Linear Regression Diagnostics
-\item
-  \ldots{}
-\item
-  \ldots{}
+  The first project provides a prototype of automatic visual inference system for evaluating lineups of residual plots, with a focus on the classical normal linear regression model given it is one of the simplest predictive models.
+The core of the system
 \end{enumerate}
+
+Validation of the automatic system is made by comparing the power of human subject-assessed visual tests with the system-assessed visual tests.
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
+\tightlist
+\item
+  The second project aims to develop an automatic visual inference system based on image comparison.
+\end{enumerate}
+
+This project focused on building a prototype of automatic visual statistical inference system for evaluating residual plots of classical normal linear regression model.
+To set up a comparison between the computer vision models and humans, human subject experiments were conducted to understand the ability of human reading residual plots.
+
+The first project provides visual diagnostics for projection pursuit optimisation, with a focus on the guided tour.
+A guided tour is constructed by combining the optimisation, for choosing new planes to visit, with a geodesic interpolation between planes to provide a smooth movie of projections for visualisation.
+The visual diagnostics include static graphics and animations to understand different aspects of the optimisers: its progression on improving the index value, its coverage of the search space and the comparison of different optimisers in different scenarios.
+What is learned will be used to modify and improve the performance of the optimisers.
+
+The second project aims to develop visual diagnostic tools for stochastic optimal control algorithms.
+In comparison to projection pursuit, optimisation in stochastic optimal control problems finds a sequence of discrete-valued decisions that are optimal at each time point.
+Exogenous variables that affect the decisions are also involved in the optimisation.
+The project will extend and build on the visual diagnostic tools developed in the first project to address stochastic optimal control optimisation.
+
+The third project aims to use visual techniques to understand variable importance in stochastic optimal control problem.
+The project will study how to interpret the contribution of potentially dozens of variables to the decision at a particular time and how the future decision would change with a change in the exogenous variables.
 

### 2022-03-09 Milestone_1/01-chap1.Rmd
add overview of the thesis

--- 2022-03-08 Milestone_1/01-chap1.Rmd
@@
 
-# Structure of the Thesis (need disscussion)
-
-An overview of the thesis, including some brief detail of the proposed content of all chapters.
-This should encompass a description of the theoretical and conceptual framework that underlies the thesis, and the procedures that are to be used in addressing the research questions
-
-The
-
-The first project is
-
-1. Automatic Visual Statistical Inference, with Application to Linear Regression Diagnostics
-2. ...
-3. ...
-
-# Automatic Visual Statistical Inference, with Application to Linear Regression Diagnostics
+# Research Problem
+
+<!-- 1.
+Provide visual diagnostics for projection pursuit optimisation to examine whether the optimisers are able to find the optima, compare optimisation paths, inspect optimisers’ coverage of the space, and detect anomalies in the algorithms.
+-->
+
+<!-- 2.
+Extend and enhance the visual diagnostic tools developed in the first project to stochastic optimal control problems, which is a multistage optimisation problem with more than two exogenous variables.
+-->
+
+<!-- 3.
+Improve the interpretability and explainability of stochastic optimal control algo- rithms by identifying the most important drivers in the optimal decisions suggested by the optimisers and studying how the decision changes across time.
+-->
+
+The main objective of this research is to build an automatic visual inference system for the purpose of conducting large-scale visual test. The research will focus on three projects as follows:
+
+1.
+Propose a prototype of automatic visual inference system for evaluating lineups of residual plots of the classical normal linear regression model and study factors that affect the performance of the system with comparison to human subjects.
+
+2. Extend and enhance the automatic visual inference system proposed in the first project to general lineup protocol problems by adopting image similarity assessed by computer vision models.
+
+3. Validate and apply the automatic visual inference system to predictive model diagnostics on empirical data.
+
+# Overview of the Thesis
+
+The first project proposes a prototype of automatic visual inference system for evaluating lineups of residual plots, with a focus on the classical normal linear regression model given it is one of the simplest predictive models.
+The automatic system is identical to the lineup protocol except evaluators are replaced by computer vision models.
+The computer vision models are trained by using data simulated from linear models with violations of different classical assumptions, such as linearity and homoscedasticity.
+Data of human performance on evaluating residual plots generated under the same simulation setting is collected by conducting online human subject experiments, such that the comparison between the power of human subject-assessed visual tests and the system-assessed visual tests can be made.
+Finally, factors that affect the performance of the automatic system are studied for improving the architecture and design of the computer vision models.
+
+The second project aims to extend the proposed automatic system to lineup protocol with no assumptions on the type of the data plot and the data generating process.
+It only requires the protocol to asks the evaluators to select the most different plot(s) from the lineup.
+The automatic system will be built upon image similarity produced by computer vision models.
+
+The third project aims to apply the automatic system to predictive model diagnostics on empirical data.
+Potential predictive models considered by this research are linear model, generalized linear models and non-parametric regression.
+Multiple human subject experiments will need to be conducted to validate the effectiveness of the automatic system.
+
+# Automatic Visual Statistical Inference for linear regression diagnostics
 
@@
 
+# General Purpose Automatic Visual Statistical Inference System
+
+# Visual Diagnostics for Predictive Models with ...
+
 # Timeline

### 2022-03-09 Milestone_1/_book/thesis.tex
add overview of the thesis

--- 2022-03-08 Milestone_1/_book/thesis.tex
@@
 
-\hypertarget{structure-of-the-thesis-need-disscussion}{%
-\chapter{Structure of the Thesis (need disscussion)}\label{structure-of-the-thesis-need-disscussion}}
-
-An overview of the thesis, including some brief detail of the proposed content of all chapters.
-This should encompass a description of the theoretical and conceptual framework that underlies the thesis, and the procedures that are to be used in addressing the research questions
-
-The
-
-The first project is
+\hypertarget{research-problem}{%
+\chapter{Research Problem}\label{research-problem}}
+
+The main objective of this research is to build an automatic visual inference system for the purpose of conducting large-scale visual test. The research will focus on three projects as follows:
 
@@
 \def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Automatic Visual Statistical Inference, with Application to Linear Regression Diagnostics
-\item
-  \ldots{}
-\item
-  \ldots{}
+\item
+  Propose a prototype of automatic visual inference system for evaluating lineups of residual plots of the classical normal linear regression model and study factors that affect the performance of the system with comparison to human subjects.
+\item
+  Extend and enhance the automatic visual inference system proposed in the first project to general lineup protocol problems by adopting image similarity assessed by computer vision models.
+\item
+  Validate and apply the automatic visual inference system to predictive model diagnostics on empirical data.
 \end{enumerate}
 
-\hypertarget{automatic-visual-statistical-inference-with-application-to-linear-regression-diagnostics}{%
-\chapter{Automatic Visual Statistical Inference, with Application to Linear Regression Diagnostics}\label{automatic-visual-statistical-inference-with-application-to-linear-regression-diagnostics}}
+\hypertarget{overview-of-the-thesis}{%
+\chapter{Overview of the Thesis}\label{overview-of-the-thesis}}
+
+The first project proposes a prototype of automatic visual inference system for evaluating lineups of residual plots, with a focus on the classical normal linear regression model given it is one of the simplest predictive models.
+The automatic system is identical to the lineup protocol except evaluators are replaced by computer vision models.
+The computer vision models are trained by using data simulated from linear models with violations of different classical assumptions, such as linearity and homoscedasticity.
+Data of human performance on evaluating residual plots generated under the same simulation setting is collected by conducting online human subject experiments, such that the comparison between the power of human subject-assessed visual tests and the system-assessed visual tests can be made.
+Finally, factors that affect the performance of the automatic system are studied for improving the architecture and design of the computer vision models.
+
+The second project aims to extend the proposed automatic system to lineup protocol with no assumptions on the type of the data plot and the data generating process.
+It only requires the protocol to asks the evaluators to select the most different plot(s) from the lineup.
+The automatic system will be built upon image similarity produced by computer vision models.
+
+The third project aims to apply the automatic system to predictive model diagnostics on empirical data.
+Potential predictive models considered by this research are linear model, generalized linear models and non-parametric regression.
+Multiple human subject experiments will need to be conducted to validate the effectiveness of the automatic system.
+
+\hypertarget{automatic-visual-statistical-inference-for-linear-regression-diagnostics}{%
+\chapter{Automatic Visual Statistical Inference for linear regression diagnostics}\label{automatic-visual-statistical-inference-for-linear-regression-diagnostics}}
 
@@
 
+\hypertarget{general-purpose-automatic-visual-statistical-inference-system}{%
+\chapter{General Purpose Automatic Visual Statistical Inference System}\label{general-purpose-automatic-visual-statistical-inference-system}}
+
+\hypertarget{visual-diagnostics-for-predictive-models-with}{%
+\chapter{Visual Diagnostics for Predictive Models with \ldots{}}\label{visual-diagnostics-for-predictive-models-with}}
+
 \hypertarget{timeline}{%

### 2022-03-09 Milestone_1/01-chap1.Rmd
rewrite 4.2.1

@@
 
-This project focused on building a prototype of automatic visual statistical inference system for evaluating residual plots of classical normal linear regression model.
+This project focuses on building a prototype of automatic visual statistical inference system for evaluating residual plots of classical normal linear regression model.
 To set up a comparison between the computer vision models and humans, human subject experiments were conducted to understand the ability of human reading residual plots.
+
+Because the first project is still ongoing, we consider the result we have is not enough to write as a paper at the moment.
+Hence, the material provided in this section is two main parts of the draft paper that will contain enough details for understanding the project.
 
@@
 
-Visual tests can be used to make valid inference for visual discoveries while acknowledging people with confidence about the findings.
-However, the lack of scalability of this method limits its application in wider fields.
-The need of automatic visual inference system is auto data analysis and machine learning system.
-
-AI agents which can read data plot and evaluate lineups could
-
-Considering there are different approches of AI discussed in \ref{se:ai}, it is necessary to clarify the type of AI we aimed to build.
-Understanding the human thought process or mimicking the human vision mechanism were not of interest in this research.
-We believed humanly behaviours was enough for demonstrating AI for the problem we concerned.
-Hence, two of the capacities required by the total Turing test, namely machine learning and computer vision will be adopted.
-
-Another possible method was to define distance metrics to measure difference between data plots for making mathematically prefect decisions.
-This method would work if such metrics can be obtained manually or approximated algorithmically.
-However, we considered it was too complex as a starting point.
-
-### Automatic visual inference system
-
-For an automatic visual inference system, types of inputs can be classified by two decisions: (i) whether to accept multiple image files as inputs, and (ii) whether to accept fitted model.
-
-single image file
-multiple image files
-
-fitted model
-
-Depend on the inputs and outputs, computer vision model can be integrated into an automatic visual inference system.
+Lineup protocol can be used to make valid inference for visual discoveries, but one of the drawbacks of this protocol is it requires human evaluation of data plots.
+Such requirement is unattainable when the number of data plots need for testing is large.
+Even if it can be fulfilled, the cost of labour would be exceedingly high.
+Hence, there is a need of building an automatic visual inference system to read data plots, evaluate lineups and provide meaningful decisions.
+Since this desired automatic system is technically an agent intended to play the role of evaluators as humans in lineup protocol, some degree of intelligence needs to be demonstrated.
+This brings us to Artificial Intelligence (AI).
 
@@
 
-Since the automatic visual inference system is intended to play the role of evaluators as humans in visual inference, some degree of intelligence needs to be demonstrated.
-This brings us to one of the popular filed of study - **Artificial Intelligence** (AI).
 AI is the field of research concerned with understanding and building machines who can demonstrate intelligence.
@@
 "Thought".
-Positioning at the top right cell, the **rational agent approach** aims to build agent that perform mathematically perfect acts such that the best expected outcome can always be achieved.
-In contrast, the **"laws of thought" approach** focus on understanding the logic behind the rationality.
-Closely related to **cognitive science**, the **cognitive modelling approach** attempts to express theories of human cognition as computer program to mimic the thought process of human.
-Lastly, the **Turing test approach** is built upon the famous **Turing test** proposed by @turing_computing_1950.
+Positioning at the top right cell, the rational agent approach aims to build agent that perform mathematically perfect acts such that the best expected outcome can always be achieved.
+In contrast, the "laws of thought" approach focus on understanding the logic behind the rationality.
+Closely related to cognitive science, the cognitive modelling approach attempts to express theories of human cognition as computer program to mimic the thought process of human.
+Lastly, the Turing test approach is built upon the famous Turing test proposed by @turing_computing_1950.
 The test can be roughly described as, whether a human can distinguish another human from a computer with written communications only.
 To pass the test, several capabilities of computer are required.
-This includes **natural language processing** for communication with human, **knowledge representation** for encoding knowledge, **automated reasoning** for derivation of conclusions and **machine learning** for improving AI automatically through experience and data.
+This includes natural language processing for communication with human, knowledge representation for encoding knowledge, automated reasoning for derivation of conclusions and machine learning for improving AI automatically through experience and data.
 Some researchers argued that written communication is insufficient to demonstrate intelligence, and some degree of physical simulation of a person is still necessary.
-One such example is the **total Turing test** proposed by @harnad_other_1991.
-It adds three new requirements to the list, including **computer vision**, **speech recognition** and **robotics**, which are response for interactions with the physical world.
+One such example is the total Turing test proposed by @harnad_other_1991.
+It adds three new requirements to the list, including computer vision, speech recognition and robotics, which are response for interactions with the physical world.
 Notably, all seven required capabilities have become major subfields of AI today.
@@
 ```
+
+Considering there are different approches of AI, it is necessary to clarify the type of AI we aimed to build.
+Understanding the human thought process or mimicking the human vision mechanism were not of interest in this research.
+We believed humanly behaviours was enough for demonstrating AI for the problem we concerned.
+Hence, two of the capacities required by the total Turing test, namely machine learning and computer vision were adopted.
+
+Another possible method under the rational agent approach was to define distance metrics to measure difference between data plots for making mathematically prefect decisions.
+This method would work if such metrics can be obtained manually or approximated algorithmically.
+However, we considered it was too complex to fit into our prototype automatic system.
+But we will revisited this method in our second project.
+
+### Design of the Automatic System
+
+Depended on the inputs and outputs of the automatic system, various designs arised.
+
+A system can be summarised
+
+For an automatic visual inference system, types of inputs can be classified by two decisions: (i) whether to accept multiple image files as inputs, and (ii) whether to accept fitted model.
+
+single image file
+multiple image files
+
+fitted model
+
+Depend on the inputs and outputs, computer vision model can be integrated into an automatic visual inference system.
 

### 2022-03-09 Milestone_1/_book/thesis.tex
rewrite 4.2.1

@@
 
-Visual tests can be used to make valid inference for visual discoveries while acknowledging people with confidence about the findings.
-However, the lack of scalability of this method limits its application in wider fields.
-The need of automatic visual inference system is auto data analysis and machine learning system.
-
-AI agents which can read data plot and evaluate lineups could
-
-Considering there are different approches of AI discussed in \ref{se:ai}, it is necessary to clarify the type of AI we aimed to build.
-Understanding the human thought process or mimicking the human vision mechanism were not of interest in this research.
-We believed humanly behaviours was enough for demonstrating AI for the problem we concerned.
-Hence, two of the capacities required by the total Turing test, namely machine learning and computer vision will be adopted.
-
-Another possible method was to define distance metrics to measure difference between data plots for making mathematically prefect decisions.
-This method would work if such metrics can be obtained manually or approximated algorithmically.
-However, we considered it was too complex as a starting point.
-
-\hypertarget{automatic-visual-inference-system-1}{%
-\subsection{Automatic visual inference system}\label{automatic-visual-inference-system-1}}
-
-For an automatic visual inference system, types of inputs can be classified by two decisions: (i) whether to accept multiple image files as inputs, and (ii) whether to accept fitted model.
-
-single image file
-multiple image files
-
-fitted model
-
-Depend on the inputs and outputs, computer vision model can be integrated into an automatic visual inference system.
+Lineup protocol can be used to make valid inference for visual discoveries, but one of the drawbacks of this protocol is it requires human evaluation of data plots.
+Such requirement is unattainable when the number of data plots need for testing is large.
+Even if it can be fulfilled, the cost of labour would be exceedingly high.
+Hence, there is a need of building an automatic visual inference system to read data plots, evaluate lineups and provide meaningful decisions.
+Since this desired automatic system is technically an agent intended to play the role of evaluators as humans in lineup protocol, some degree of intelligence needs to be demonstrated.
+This brings us to Artificial Intelligence (AI).
 
@@
 
-Since the automatic visual inference system is intended to play the role of evaluators as humans in visual inference, some degree of intelligence needs to be demonstrated.
-This brings us to one of the popular filed of study - \textbf{Artificial Intelligence} (AI).
 AI is the field of research concerned with understanding and building machines who can demonstrate intelligence.
@@
 These approaches can be summarized into a two by two table as shown in Figure \ref{fig:two-dim-ai}, where the row is ``Human'' vs.~``Rational'', and the column is ``Behaviour'' vs.~``Thought''.
-Positioning at the top right cell, the \textbf{rational agent approach} aims to build agent that perform mathematically perfect acts such that the best expected outcome can always be achieved.
-In contrast, the \textbf{``laws of thought'' approach} focus on understanding the logic behind the rationality.
-Closely related to \textbf{cognitive science}, the \textbf{cognitive modelling approach} attempts to express theories of human cognition as computer program to mimic the thought process of human.
-Lastly, the \textbf{Turing test approach} is built upon the famous \textbf{Turing test} proposed by \textcite{turing_computing_1950}.
+Positioning at the top right cell, the rational agent approach aims to build agent that perform mathematically perfect acts such that the best expected outcome can always be achieved.
+In contrast, the ``laws of thought'' approach focus on understanding the logic behind the rationality.
+Closely related to cognitive science, the cognitive modelling approach attempts to express theories of human cognition as computer program to mimic the thought process of human.
+Lastly, the Turing test approach is built upon the famous Turing test proposed by \textcite{turing_computing_1950}.
 The test can be roughly described as, whether a human can distinguish another human from a computer with written communications only.
 To pass the test, several capabilities of computer are required.
-This includes \textbf{natural language processing} for communication with human, \textbf{knowledge representation} for encoding knowledge, \textbf{automated reasoning} for derivation of conclusions and \textbf{machine learning} for improving AI automatically through experience and data.
+This includes natural language processing for communication with human, knowledge representation for encoding knowledge, automated reasoning for derivation of conclusions and machine learning for improving AI automatically through experience and data.
 Some researchers argued that written communication is insufficient to demonstrate intelligence, and some degree of physical simulation of a person is still necessary.
-One such example is the \textbf{total Turing test} proposed by \textcite{harnad_other_1991}.
-It adds three new requirements to the list, including \textbf{computer vision}, \textbf{speech recognition} and \textbf{robotics}, which are response for interactions with the physical world.
+One such example is the total Turing test proposed by \textcite{harnad_other_1991}.
+It adds three new requirements to the list, including computer vision, speech recognition and robotics, which are response for interactions with the physical world.
 Notably, all seven required capabilities have become major subfields of AI today.
@@
 \end{figure}
+
+Considering there are different approches of AI, it is necessary to clarify the type of AI we aimed to build.
+Understanding the human thought process or mimicking the human vision mechanism were not of interest in this research.
+We believed humanly behaviours was enough for demonstrating AI for the problem we concerned.
+Hence, two of the capacities required by the total Turing test, namely machine learning and computer vision were adopted.
+
+Another possible method under the rational agent approach was to define distance metrics to measure difference between data plots for making mathematically prefect decisions.
+This method would work if such metrics can be obtained manually or approximated algorithmically.
+However, we considered it was too complex to fit into our prototype automatic system.
+But we will revisited this method in our second project.
+
+\hypertarget{design-of-the-automatic-system}{%
+\subsection{Design of the Automatic System}\label{design-of-the-automatic-system}}
+
+Depended on the inputs and outputs of the automatic system, various designs arised.
+
+A system can be summarised
+
+For an automatic visual inference system, types of inputs can be classified by two decisions: (i) whether to accept multiple image files as inputs, and (ii) whether to accept fitted model.
+
+single image file
+multiple image files
+
+fitted model
+
+Depend on the inputs and outputs, computer vision model can be integrated into an automatic visual inference system.
 

### 2022-03-09 Milestone_1/01-chap1.Rmd
fix "propose"

@@
 1.
-Propose a prototype of automatic visual inference system for evaluating lineups of residual plots of the classical normal linear regression model and study factors that affect the performance of the system with comparison to human subjects.
-
-2. Extend and enhance the automatic visual inference system proposed in the first project to general lineup protocol problems by adopting image similarity assessed by computer vision models.
+Develop a prototype of automatic visual inference system for evaluating lineups of residual plots of the classical normal linear regression model and study factors that affect the performance of the system with comparison to human subjects.
+
+2. Extend and enhance the automatic visual inference system developed in the first project to general lineup protocol problems by adopting image similarity assessed by computer vision models.
 
@@
 
-The first project proposes a prototype of automatic visual inference system for evaluating lineups of residual plots, with a focus on the classical normal linear regression model given it is one of the simplest predictive models.
+The first project develops a prototype of automatic visual inference system for evaluating lineups of residual plots, with a focus on the classical normal linear regression model given it is one of the simplest predictive models.
 The automatic system is identical to the lineup protocol except evaluators are replaced by computer vision models.
@@
 
-The second project aims to extend the proposed automatic system to lineup protocol with no assumptions on the type of the data plot and the data generating process.
+The second project aims to extend the automatic system to lineup protocol with no assumptions on the type of the data plot and the data generating process.
 It only requires the protocol to asks the evaluators to select the most different plot(s) from the lineup.
@@
 \addlinespace[0.5em]
-\hspace{4em}July & Propose a new design of the automatic visual inference\\
-\addlinespace[0.5em]
-\hspace{4em} & system based on image comparison\\
-\addlinespace[0.5em]
-\hspace{4em}August - November & Develop the proposed system\\
+\hspace{4em}July & Draft a new automatic visual inference system based\\
+\addlinespace[0.5em]
+\hspace{4em} & on image comparison\\
+\addlinespace[0.5em]
+\hspace{4em}August - November & Build the automatic system\\
 \addlinespace[0.5em]
@@
 \addlinespace[0.5em]
-\hspace{4em} & proposed system\\
+\hspace{4em} & automatic system\\
 \addlinespace[0.5em]

### 2022-03-09 Milestone_1/_book/thesis.tex
fix "propose"

@@
 \item
-  Propose a prototype of automatic visual inference system for evaluating lineups of residual plots of the classical normal linear regression model and study factors that affect the performance of the system with comparison to human subjects.
-\item
-  Extend and enhance the automatic visual inference system proposed in the first project to general lineup protocol problems by adopting image similarity assessed by computer vision models.
+  Develop a prototype of automatic visual inference system for evaluating lineups of residual plots of the classical normal linear regression model and study factors that affect the performance of the system with comparison to human subjects.
+\item
+  Extend and enhance the automatic visual inference system developed in the first project to general lineup protocol problems by adopting image similarity assessed by computer vision models.
 \item

### 2022-03-09 Milestone_1/01-chap1.Rmd
delete text between 1 & 1.1

@@
 
-The introduction of the report is organised as follows.
+<!-- ## Motivation -->
+
+<!-- Have you doubted what you found from a data plot was really there?
+Or have you had reservations about visual discoveries because they might be illusions?
+Data plots are prevalent in every field of nature and social science as it is one of the most recognized data visualization techniques.
+Without data plots, trends and patterns embed in the data will not be easy to discover.
+However, data plots are often over- or under-interpreted by human readers which leads to unsecured findings.
+This is unsatisfactory as confirmatory research is the foundation of science.
+-->
+
+<!-- In this research,  -->
+
+<!-- In this report, we focus on predictive model diagnostics -->
+
+<!-- While doing predictive model diagnostics -->
+
+<!-- In the introduction of this report, we will start by discussing the similar issues of the use of data plots in predictive model diagnostics.
+Then, in section \ref{se:vi}, the definition and terminology of visual inference will be outlined for addressing the issues.
+Section \ref{se:prespecification} and Section \ref{se:sampling-from-null} provides more details about visual inference, such as its validity and  -->
+
+<!-- describes the sampling techniques in visual inference with an example of the classical normal linear regression model. -->
+
+<!-- The introduction of the report is organised as follows.
 Section \ref{se:predictive-modelling} discusses the predictive modelling culture and the use of data plots in predictive model diagnostics.
@@
 And lastly, Section \ref{se:limitation} points out the limitation of lineup protocol, which is also the motivation of this research, and the need of building an automatic visual inference system.
-
-<!-- ## Motivation -->
-
-<!-- Have you doubted what you found from a data plot was really there?
-Or have you had reservations about visual discoveries because they might be illusions?
-Data plots are prevalent in every field of nature and social science as it is one of the most recognized data visualization techniques.
-Without data plots, trends and patterns embed in the data would not be easy to discover.
-However, data plots are often over- or under-interpreted by human readers which leads to unsecured findings.
-This is unsatisfactory as confirmatory research is the foundation of science.
 -->
@@
 Data of human performance on evaluating residual plots generated under the same simulation setting is collected by conducting online human subject experiments, such that the comparison between the power of human subject-assessed visual tests and the system-assessed visual tests can be made.
-Finally, factors that affect the performance of the automatic system are studied for improving the architecture and design of the computer vision models.
+Moreover, factors that affect the performance of the automatic system are studied for improving the architecture and design of the computer vision models.
 

### 2022-03-09 Milestone_1/_book/thesis.tex
delete text between 1 & 1.1

@@
 
-The introduction of the report is organised as follows.
-Section \ref{se:predictive-modelling} discusses the predictive modelling culture and the use of data plots in predictive model diagnostics.
-Section \ref{se:vi} outlines the definition and terminology of visual inference.
-Section \ref{se:prespecification} presents the justification of the validity of visual inference.
-Section \ref{se:sampling-from-null} describes the sampling techniques in visual inference with an example of the classical normal linear regression model.
-Section \ref{se:lineup} introduces the lineup protocol and the estimation of \(p\)-value and power of a visual test.
-And lastly, Section \ref{se:limitation} points out the limitation of lineup protocol, which is also the motivation of this research, and the need of building an automatic visual inference system.
-
 \hypertarget{se:predictive-modelling}{%

### 2022-03-09 Milestone_1/01-chap1.Rmd
add desc for second project

@@
 
-# General Purpose Automatic Visual Statistical Inference System
+# General-Purpose Automatic Visual Statistical Inference System
+
+The term content-based image retrieval (CBIR) was first introduced by @kato_database_1992 as a field of study which aimes to retrieve image from large-scale image databases based on information derived from the image itself, such as colour, shape and texture [@rafiee_review_2010].
+According to @zhou_recent_2017, reverse image search is one of the CBIR query technique and is sometimes called query by example (QBE), which performs image search based on the image example provide by the user.
+Since QBE is the most intuitive query formation, it has been widely explored in the research on CBIR and search engines such as Google Image [@google_google_2022] and TinEye [@tineye_tineye_2022] provide accessible search by image services to public.
+In CBIR, the most common method for comparing images is using image similarity or image distance metrics, which takes into account various features extracted from the image [@zhou_recent_2017].
+
+By adopting the techniques developed in query by example content-based image retrieval, we aims to build automatic general-purpose visual inference system based on image comparison in the second project.
 

### 2022-03-09 Milestone_1/_book/thesis.tex
add desc for second project

@@
 \hypertarget{general-purpose-automatic-visual-statistical-inference-system}{%
-\chapter{General Purpose Automatic Visual Statistical Inference System}\label{general-purpose-automatic-visual-statistical-inference-system}}
+\chapter{General-Purpose Automatic Visual Statistical Inference System}\label{general-purpose-automatic-visual-statistical-inference-system}}
+
+The term content-based image retrieval (CBIR) was first introduced by \textcite{kato_database_1992} as a field of study which aimes to retrieve image from large-scale image databases based on information derived from the image itself, such as colour, shape and texture \autocite{rafiee_review_2010}.
+According to \textcite{zhou_recent_2017}, reverse image search is one of the CBIR query technique and is sometimes called query by example (QBE), which performs image search based on the image example provide by the user.
+Since QBE is the most intuitive query formation, it has been widely explored in the research on CBIR and search engines such as Google Image \autocite{google_google_2022} and TinEye \autocite{tineye_tineye_2022} provide accessible search by image services to public.
+In CBIR, the most common method for comparing images is using image similarity or image distance metrics, which takes into account various features extracted from the image \autocite{zhou_recent_2017}.
+
+By adopting the techniques developed in query by example content-based image retrieval, we aims to build automatic general-purpose visual inference system based on image comparison in the second project.
 

### 2022-03-09 Milestone_1/01-chap1.Rmd
add design of the automatic system

@@
 
-Depended on the inputs and outputs of the automatic system, various designs arised.
-
-A system can be summarised
+Depending on the desired input(s) and output(s) of the automatic system, the system can be designed in various ways.
+We considered the most general design would be which took inputs of multiple image files corresponding to data plots of a lineup, and outputted a list of selections made by the system.
+This was identical to what was provided to and asked of the observer in the human subject experiment.
+
+Decision also needed to be made for the input and output format of the computer vision models.
+If the model took multiple data plots of a lineup as inputs, and outputted the predicted probabilities of the data plot being the most different one, then the post-processing step of the system would be a mapping from a list of numeric values between zero and one to a list of one-hot encoded binary values denoting the selections.
+This mapping could be an indicator function, i.e.
+a threshold for deciding whether to select a data plot or not.
+It could also be a combination of the indicator function and the order statistic to select the $M$ most likely data plots, where $M$ was a user-defined positive integer.
+In fact, the choice of the mapping function was flexible.
+
+This design was reasonable but we found that it was difficult and slow to train the model.
+Instead, a simpler model design was used.
+The model took only one data plot as input at a time, and predicted the probability of the data plot being a null plot.
+In this case, the post-processing step was an indicator function for deciding whether the data plot is a null plot.
+By evaluating all $m$ plots of a lineup, multiple selections can be made.
+
+### Model Architecture and Model Training
+
+As mentioned in Section \ref{se:limitation}, modern computer vision model is built upon the deep neural network
 
@@
 
-The term content-based image retrieval (CBIR) was first introduced by @kato_database_1992 as a field of study which aimes to retrieve image from large-scale image databases based on information derived from the image itself, such as colour, shape and texture [@rafiee_review_2010].
+The term content-based image retrieval (CBIR) was first introduced by @kato_database_1992 as a field of study which aims to retrieve image from large-scale image databases based on information derived from the image itself, such as colour, shape and texture [@rafiee_review_2010].
 According to @zhou_recent_2017, reverse image search is one of the CBIR query technique and is sometimes called query by example (QBE), which performs image search based on the image example provide by the user.

### 2022-03-09 Milestone_1/_book/thesis.tex
add design of the automatic system

@@
 
-Depended on the inputs and outputs of the automatic system, various designs arised.
-
-A system can be summarised
+Depending on the desired input(s) and output(s) of the automatic system, the system can be designed in various ways.
+We considered the most general design would be which took inputs of multiple image files corresponding to data plots of a lineup, and outputted a list of selections made by the system.
+This was identical to what was provided to and asked of the observer in the human subject experiment.
+
+Decision also needed to be made for the input and output format of the computer vision models.
+If the model took multiple data plots of a lineup as inputs, and outputted the predicted probabilities of the data plot being the most different one, then the post-processing step of the system would be a mapping from a list of numeric values between zero and one to a list of one-hot encoded binary values denoting the selections.
+This mapping could be an indicator function, i.e.~a threshold for deciding whether to select a data plot or not.
+It could also be a combination of the indicator function and the order statistic to select the \(M\) most likely data plots, where \(M\) was a user-defined positive integer.
+In fact, the choice of the mapping function was flexible.
+
+This design was reasonable but we found that it was difficult and slow to train the model.
+Instead, a simpler model design was used.
+The model took only one data plot as input at a time, and predicted the probability of the data plot being a null plot.
+In this case, the post-processing step was an indicator function for deciding whether the data plot is a null plot.
+By evaluating all \(m\) plots of a lineup, multiple selections can be made.
+
+\hypertarget{model-architecture-and-model-training}{%
+\subsection{Model Architecture and Model Training}\label{model-architecture-and-model-training}}
+
+As mentioned in Section \ref{se:limitation}, modern computer vision model is built upon the deep neural network
 
@@
 
-The term content-based image retrieval (CBIR) was first introduced by \textcite{kato_database_1992} as a field of study which aimes to retrieve image from large-scale image databases based on information derived from the image itself, such as colour, shape and texture \autocite{rafiee_review_2010}.
+The term content-based image retrieval (CBIR) was first introduced by \textcite{kato_database_1992} as a field of study which aims to retrieve image from large-scale image databases based on information derived from the image itself, such as colour, shape and texture \autocite{rafiee_review_2010}.
 According to \textcite{zhou_recent_2017}, reverse image search is one of the CBIR query technique and is sometimes called query by example (QBE), which performs image search based on the image example provide by the user.

### 2022-03-09 Milestone_1/01-chap1.Rmd
update system design

@@
 Depending on the desired input(s) and output(s) of the automatic system, the system can be designed in various ways.
-We considered the most general design would be which took inputs of multiple image files corresponding to data plots of a lineup, and outputted a list of selections made by the system.
-This was identical to what was provided to and asked of the observer in the human subject experiment.
+We considered the most general design would be which took inputs of multiple image files corresponding to data plots of a lineup, and outputted a $p$-value of the visual test conducted by the system.
+The human subject experiment was close to this design except the data plots are combined into a single image.
 
@@
 
-This design was reasonable but we found that it was difficult and slow to train the model.
+This design was reasonable but we found that it was difficult and slow to train the computer vision model.
 Instead, a simpler model design was used.
@@
 By evaluating all $m$ plots of a lineup, multiple selections can be made.
+A summary of the system design is given in Figure \ref{fig:system-design}.
+
+![Design of the automatic visual inference system.
+Every data plot of a lineup was evaluated by five models to decide whether or not to select the data plot.
+By combining the selections made by five models, $p$-value can be estimated.
+\label{fig:system-design}](figures/vi_system.pdf){width=550 height=236}
 
@@
 
-As mentioned in Section \ref{se:limitation}, modern computer vision model is built upon the deep neural network
+As mentioned in Section \ref{se:limitation}, modern computer vision model is built upon the deep neural network with tens of thousands of weights.
+Due to limited computing resources, we did not have the luxury to train a deep neural network from scratch.
+A technique called transfer learning was used to leverage features learned from an existing problem on our task.
+"It is used to improve a learner from one domain by transferring information from a related domain [@weiss_survey_2016]."
+
+The practical workflow was to take weights from a previously trained model to a new model, and froze them in training to prevent the knowledge preserved in the weights from being destroyed.
+Then, added some trainable layers on top of the frozen layers to solve the new problem.
+
+By using this workflow, we were not only able to reduce the training time, but also able to gain better performance because the pre-trained models we used were trained on a large dataset called ImageNet [@deng_imagenet_2009] with more than 14 million images on 1000 object classes.
+The famous annual competition ImageNet Large Scale Visual Recognition Challenge (ILSCRV) ran annually by the ImageNet project was a gold standard for evaluating the architecture of computer vision model.
+Among the leaderboard of ILSCRB, we picked architectures with low top-5 error rate and were also easy to configure.
+This includes VGG16 [@simonyan_very_2015], VGG19 [@simonyan_very_2015], Xception [@chollet_xception_2017], ResNet152V2 [@he_identity_2016] and InceptionV3 [@szegedy_rethinking_2015].
+The Top-5 error rate is the average number the model failed to include the answer among its top five guesses.
+
+consists of taking features learned on one problem, and leveraging them on a new, similar problem.
+For instance, features from a model that has learned to identify racoons may be useful to kick-start a model meant to identify tanukis.
 

### 2022-03-09 Milestone_1/_book/thesis.tex
update system design

@@
 Depending on the desired input(s) and output(s) of the automatic system, the system can be designed in various ways.
-We considered the most general design would be which took inputs of multiple image files corresponding to data plots of a lineup, and outputted a list of selections made by the system.
-This was identical to what was provided to and asked of the observer in the human subject experiment.
+We considered the most general design would be which took inputs of multiple image files corresponding to data plots of a lineup, and outputted a \(p\)-value of the visual test conducted by the system.
+The human subject experiment was close to this design except the data plots are combined into a single image.
 
@@
 By evaluating all \(m\) plots of a lineup, multiple selections can be made.
+A summary of the system design is given in Figure \ref{fig:system-design}.
+
+\begin{figure}
+\centering
+\includegraphics[width=5.72917in,height=2.45833in]{figures/vi_system.pdf}
+\caption{Design of the automatic visual inference system.
+Every data plot of a lineup was evaluated by five models to decide whether or not to select the data plot.
+By combining the selections made by five models, \(p\)-value can be estimated.
+\label{fig:system-design}}
+\end{figure}
 
@@
 
-As mentioned in Section \ref{se:limitation}, modern computer vision model is built upon the deep neural network
+As mentioned in Section \ref{se:limitation}, modern computer vision model is built upon the deep neural network with tens of thousands of weights.
+Due to limited computing resources, we did not have the luxury to train a deep neural network from scratch.
+A technique called transfer learning was used to leverage features learned from an existing problem on our task.
+``It is used to improve a learner from one domain by transferring information from a related domain \autocite{weiss_survey_2016}.''
+
+The practical workflow was to take weights from a previously trained model to a new model, and froze them in training to prevent the knowledge preserved in the weights from being destroyed.
+Then, added some trainable layers on top of the frozen layers to solve the new problem.
+
+By using this workflow, we were not only able to reduce the training time, but also able to gain better performance because the pre-trained models we used were trained on a large dataset called ImageNet \autocite{deng_imagenet_2009} with more than 14 million images on 1000 object classes.
+The famous annual competition ImageNet Large Scale Visual Recognition Challenge (ILSCRV) ran annually by the ImageNet project was a gold standard for evaluating the architecture of computer vision model.
+Among the leaderboard of ILSCRB, we picked architectures with low top-5 error rate and were also easy to configure.
+This includes VGG16 \autocite{simonyan_very_2015}, VGG19 \autocite{simonyan_very_2015}, Xception \autocite{chollet_xception_2017}, ResNet152V2 \autocite{he_identity_2016} and InceptionV3 \autocite{szegedy_rethinking_2015}.
+The Top-5 error rate is the average number the model failed to include the answer among its top five guesses.
+
+consists of taking features learned on one problem, and leveraging them on a new, similar problem.
+For instance, features from a model that has learned to identify racoons may be useful to kick-start a model meant to identify tanukis.
 

### 2022-03-09 Milestone_1/01-chap1.Rmd
add cache

@@
 As a subfield of AI, computer vision with the modern deep learning architectures solved numerous critical problems in automation.
-Inspired by the vision processing in living organisms, the convolutional neural network was introduced by @fukushima_neocognitron_1982.
+Inspired by the vision processing in living organisms, the convolutional neural network (CNN) was introduced by @fukushima_neocognitron_1982.
 Soon, this architecture was applied to hand-written number recognition trained with back-propagation by @lecun_backpropagation_1989.
@@
 By evaluating all $m$ plots of a lineup, multiple selections can be made.
+Five different models were included in the system to act as five evaluators, such that the relationship between the number of models and the power of automatic visual test can be studied.
 A summary of the system design is given in Figure \ref{fig:system-design}.
@@
 
-As mentioned in Section \ref{se:limitation}, modern computer vision model is built upon the deep neural network with tens of thousands of weights.
+As mentioned in Section \ref{se:limitation}, modern computer vision model is built upon the deep neural network with tens of thousands of parameters.
 Due to limited computing resources, we did not have the luxury to train a deep neural network from scratch.
 A technique called transfer learning was used to leverage features learned from an existing problem on our task.
-"It is used to improve a learner from one domain by transferring information from a related domain [@weiss_survey_2016]."
-
-The practical workflow was to take weights from a previously trained model to a new model, and froze them in training to prevent the knowledge preserved in the weights from being destroyed.
-Then, added some trainable layers on top of the frozen layers to solve the new problem.
-
-By using this workflow, we were not only able to reduce the training time, but also able to gain better performance because the pre-trained models we used were trained on a large dataset called ImageNet [@deng_imagenet_2009] with more than 14 million images on 1000 object classes.
+It improved the performance of the model by transferring information from one domain to another related domain [@weiss_survey_2016].
+The practical workflow was to take parameters from a previously trained model to a new model, and freeze them in training to prevent the knowledge preserved in the parameters from being destroyed.
+Then, we added some trainable layers on top of the frozen layers to solve the new problem.
+
+By using this workflow, we were not only able to dramatically reduce the training time, but also able to gain better performance because the pre-trained models we used were trained on a large dataset called ImageNet [@deng_imagenet_2009] with more than 14 million images on 1000 object classes.
 The famous annual competition ImageNet Large Scale Visual Recognition Challenge (ILSCRV) ran annually by the ImageNet project was a gold standard for evaluating the architecture of computer vision model.
-Among the leaderboard of ILSCRB, we picked architectures with low top-5 error rate and were also easy to configure.
-This includes VGG16 [@simonyan_very_2015], VGG19 [@simonyan_very_2015], Xception [@chollet_xception_2017], ResNet152V2 [@he_identity_2016] and InceptionV3 [@szegedy_rethinking_2015].
-The Top-5 error rate is the average number the model failed to include the answer among its top five guesses.
-
-consists of taking features learned on one problem, and leveraging them on a new, similar problem.
-For instance, features from a model that has learned to identify racoons may be useful to kick-start a model meant to identify tanukis.
-
-For an automatic visual inference system, types of inputs can be classified by two decisions: (i) whether to accept multiple image files as inputs, and (ii) whether to accept fitted model.
-
-single image file
-multiple image files
-
-fitted model
-
-Depend on the inputs and outputs, computer vision model can be integrated into an automatic visual inference system.
-
-### different approaches of AI
-
-- Not aim for understanding the thought process
-- Not aim for mimicking the human vision mechanism
-- May be able to define distance metrics to measure difference between data plots for making mathematically prefect decisions
-- May be able to use computer vision model to approximate how people evaluate lineups
-
-computer vision model:
-
-- Use human data to train model with human selection as target -> mimic the human behaviour
-- Use simulated lineup to train model with actual data plot as target -> assume the actual data plot is the most different one
-- Use simulated data plot to train model with null or not null as target -> limited to the alternatives given to the model, and Type I error can not be controlled
-- Use simulated data plot to train model, then let the model evaluate each plot of a lineup -> leads to multiple selections in a lineup, and the model does not compare different plots in a lineup
-  - Use a very large $m$ to ensure the null hypothesis can be rejected by a classifier slightly better than a random selector. The problem becomes how large the $m$ should be
-  - Use multiple models as multiple individuals to evaluate lineup -> Which models should be used? How many models should be used?
-- Let the model select multiple plots from a simulated lineup while comparing different plots in a lineup -> how to build such a system? 
-  - The input is $m$ plots, what is the output? The probability of being the most different one? 
-  - Then how many plots should be selected? Top 5? Or cumulative probability greater than a threshold? 
-  - How to select the threshold?
-
-The primary drawbacks of visual inference are
-
- infeasible in a large scale
-2. unfriendly to vision-impaired people 3. high finical cost and human cost
-4. time consuming
-
-Visual inference is a valid useful in diagnostics of predictive models
-
-To relieve human analysts of the job of diagnosing predictive models in the context of visual inference, computer vision
-
-Hence, to relieve human analysts of the job of diagnosing, automatically improve predictive models and assist vision-impaired people, this research aims to develop computer vision model to automate the reading
-1
-of data plots with application to predictive model diagnostics.
-The success of this research will enrich the applications of artificial intelligence in data visualization and improve the debugging process of artificial intelligence models.
-
-### Model Architecture
-
-Five computer vision models were trained to predict the probability of the given result plot being a null residual plot.
-
-### Performance
+Among the leaderboard of ILSCRB, we picked architectures that were handy to implement and with low top-5 error rate.
+This included VGG16 [@simonyan_very_2015], VGG19 [@simonyan_very_2015], Xception [@chollet_xception_2017], ResNet152V2 [@he_identity_2016] and InceptionV3 [@szegedy_rethinking_2015].
+The top-5 error rate is the average number the model failed to include the answer among its top five guesses.
+The overview of all five transer learning models are given in Table \ref{tab:vi-model-summary}.
+Note that the depth is the number of layers in the model.
+
+On top of the pre-trained model, we added a 2D global average pooling layer [@lin_network_2014], followed by a fully-connected dense layer with 256 nodes, a dropout layer [@srivastava_dropout_nodate] with 20% dropout rate, and a output layer with softmax as activation function.
+A diagram of the model architecture is given in Figure \ref{fig:model-arch}.
+
+Global average pooling layer was a replacement of traditional fully-connected dense layers in CNN which flattened the feature map into a 1D tensor.
+Global average pooling layer took the average of the values over all the pixels in a channel, then flattened it into a 1D tensor.
+This enforced the correspondences between feature maps and categories, which made the feature maps easier to be interpreted as categories confidence maps.
+Further, unlike the dense layer, there was no parameters that needed to be optimized in global average pooling layer which reduce the risk of overfitting [@lin_network_2014].
+
+Dropout was primary used as a regularization technique to address overfitting.
+In training, the dropout layer would randomly temporaliy remove some nodes from the network with fixed probability $p$, which in our case, we used $p = 20$%.
+
+Five models were trained on 10000 null plots and 10000 actual data plots simulated using the same parameter settings as the human subject experiments and balanced across cubic model and heteroskedasticity model.
+20% training data was used as validation set.
+Image augmentation was applied to increase the diversity of the dataset during training, with random rotation angle ranged from zero to five degrees and random brightness ranged from 95% to 105%.
+Optimizer Adam [@kingma_adam_2017] was used with learning rate equal to 0.001.
+The loss function used in the model was binary cross-entropy, which can be expressed by: $$H = -N^{-1}\sum_{i=1}^{N}y_ilog(\hat{p}_i)+(1-y_i)log(1-\hat{p}_i),$$ where $N$ is the number of observations, $y_i$ is a realization of a binary random variable, and $\hat{p}_i$ is the predicted probability.
+
+Early stopping was used to reduce training time and to avoid overfitting. If the validation loss did not decrease for five iterations, then the training procedure would stop.
+
+The trained models were evaluated on 10000 null plots and 10000 actual data plots simulated by the same way as the training data.
+And the system was then evaluated on 2500 cubic lineups and 2500 heteroskedasticity lineups.
+
+![Architecture of the computer vision model. \label{fig:model-arch}](figures/vi_model.pdf){width=87 height=491}
+
+```{r}
+data.frame(model = c("VGG16", "VGG19", "Xception", "ResNet152V2", "InceptionV3"),
+           depth = c(16, 19, 81, 307, 189) + 4,
+           total_parameters = c(14846530, 20156226, 21386538, 58856706, 22327842),
+           non_trainable = c(14714688, 20024384, 20861480, 58331648, 21802784)) %>%
+  mutate(trainable = total_parameters - non_trainable) %>%
+  mutate(top5 = c(90.1, 90, 94.5, 94.2, 93.7)/100) %>%
+  knitr::kable(col.names = c("Model", "Depth", "Total number of parameters", "Non-trainable parameters", "Trainable parameters", "Top-5 error rate"), 
+             booktabs = TRUE, 
+             caption  = 'Overview of the computer vision models.', 
+             label = "vi-model-summary" )
+```
+
+### Results
+
+The performance of all five models evaluated on the test set is given in Table \ref{tab:null-true-eval}.
+From the result, we observed that the average balanced accuracy for all five models were around 80%.
+Breaking down into null and actual data plot, it was found that
 
@@
              booktabs = TRUE, 
-             caption  = 'Performance of computer vision models on null residual plot and acutal data plot.', 
+             caption  = 'Performance of computer vision models on null residual plot and acutal data plot of test data.', 
              label = "null-true-eval" ) %>%
@@
 ```{r}
+if (file.exists(here::here("data/cache_env.rds"))) {
+  cache_env <- readRDS(here::here("data/cache_env.rds"))
+} else {
   cache_env <- new.env()
+}
+
 lineup_eval_summary %>%
   calc_p_value_multi(lineup_id = "set", detected = "detect", n_sel = "n_sel", cache_env = cache_env) -> cv_results
+
+saveRDS(cache_env, file = here::here("data/cache_env.rds"))
 ```
@@
 
+if (file.exists(here::here("data/cache_env.rds"))) {
+  cache_env <- readRDS(here::here("data/cache_env.rds"))
+} else {
+  cache_env <- new.env()
+}
+
 lineup_pseudo_summary %>%
   calc_p_value_multi(lineup_id = "set", detected = "detect", n_sel = "n_sel", cache_env = cache_env) -> pseudo_results
+
+saveRDS(cache_env, file = here::here("data/cache_env.rds"))
 
@@
 
-# Visual Diagnostics for Predictive Models with ...
+# Automatic Visual Infernce for Predictive Model Diagnostics
+
+Tools for Visual inference has been applied to
 

### 2022-03-09 Milestone_1/_book/thesis.tex
add cache

@@
 As a subfield of AI, computer vision with the modern deep learning architectures solved numerous critical problems in automation.
-Inspired by the vision processing in living organisms, the convolutional neural network was introduced by \textcite{fukushima_neocognitron_1982}.
+Inspired by the vision processing in living organisms, the convolutional neural network (CNN) was introduced by \textcite{fukushima_neocognitron_1982}.
 Soon, this architecture was applied to hand-written number recognition trained with back-propagation by \textcite{lecun_backpropagation_1989}.
@@
 By evaluating all \(m\) plots of a lineup, multiple selections can be made.
+Five different models were included in the system to act as five evaluators, such that the relationship between the number of models and the power of automatic visual test can be studied.
 A summary of the system design is given in Figure \ref{fig:system-design}.
@@
 
-As mentioned in Section \ref{se:limitation}, modern computer vision model is built upon the deep neural network with tens of thousands of weights.
+As mentioned in Section \ref{se:limitation}, modern computer vision model is built upon the deep neural network with tens of thousands of parameters.
 Due to limited computing resources, we did not have the luxury to train a deep neural network from scratch.
 A technique called transfer learning was used to leverage features learned from an existing problem on our task.
-``It is used to improve a learner from one domain by transferring information from a related domain \autocite{weiss_survey_2016}.''
-
-The practical workflow was to take weights from a previously trained model to a new model, and froze them in training to prevent the knowledge preserved in the weights from being destroyed.
-Then, added some trainable layers on top of the frozen layers to solve the new problem.
-
-By using this workflow, we were not only able to reduce the training time, but also able to gain better performance because the pre-trained models we used were trained on a large dataset called ImageNet \autocite{deng_imagenet_2009} with more than 14 million images on 1000 object classes.
+It improved the performance of the model by transferring information from one domain to another related domain \autocite{weiss_survey_2016}.
+The practical workflow was to take parameters from a previously trained model to a new model, and freeze them in training to prevent the knowledge preserved in the parameters from being destroyed.
+Then, we added some trainable layers on top of the frozen layers to solve the new problem.
+
+By using this workflow, we were not only able to dramatically reduce the training time, but also able to gain better performance because the pre-trained models we used were trained on a large dataset called ImageNet \autocite{deng_imagenet_2009} with more than 14 million images on 1000 object classes.
 The famous annual competition ImageNet Large Scale Visual Recognition Challenge (ILSCRV) ran annually by the ImageNet project was a gold standard for evaluating the architecture of computer vision model.
-Among the leaderboard of ILSCRB, we picked architectures with low top-5 error rate and were also easy to configure.
-This includes VGG16 \autocite{simonyan_very_2015}, VGG19 \autocite{simonyan_very_2015}, Xception \autocite{chollet_xception_2017}, ResNet152V2 \autocite{he_identity_2016} and InceptionV3 \autocite{szegedy_rethinking_2015}.
-The Top-5 error rate is the average number the model failed to include the answer among its top five guesses.
-
-consists of taking features learned on one problem, and leveraging them on a new, similar problem.
-For instance, features from a model that has learned to identify racoons may be useful to kick-start a model meant to identify tanukis.
-
-For an automatic visual inference system, types of inputs can be classified by two decisions: (i) whether to accept multiple image files as inputs, and (ii) whether to accept fitted model.
-
-single image file
-multiple image files
-
-fitted model
-
-Depend on the inputs and outputs, computer vision model can be integrated into an automatic visual inference system.
-
-\hypertarget{different-approaches-of-ai}{%
-\subsection{different approaches of AI}\label{different-approaches-of-ai}}
-
-\begin{itemize}
-\tightlist
-\item
-  Not aim for understanding the thought process
-\item
-  Not aim for mimicking the human vision mechanism
-\item
-  May be able to define distance metrics to measure difference between data plots for making mathematically prefect decisions
-\item
-  May be able to use computer vision model to approximate how people evaluate lineups
-\end{itemize}
-
-computer vision model:
-
-\begin{itemize}
-\tightlist
-\item
-  Use human data to train model with human selection as target -\textgreater{} mimic the human behaviour
-\item
-  Use simulated lineup to train model with actual data plot as target -\textgreater{} assume the actual data plot is the most different one
-\item
-  Use simulated data plot to train model with null or not null as target -\textgreater{} limited to the alternatives given to the model, and Type I error can not be controlled
-\item
-  Use simulated data plot to train model, then let the model evaluate each plot of a lineup -\textgreater{} leads to multiple selections in a lineup, and the model does not compare different plots in a lineup
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Use a very large \(m\) to ensure the null hypothesis can be rejected by a classifier slightly better than a random selector. The problem becomes how large the \(m\) should be
-  \item
-    Use multiple models as multiple individuals to evaluate lineup -\textgreater{} Which models should be used? How many models should be used?
-  \end{itemize}
-\item
-  Let the model select multiple plots from a simulated lineup while comparing different plots in a lineup -\textgreater{} how to build such a system?
-
-  \begin{itemize}
-  \tightlist
-  \item
-    The input is \(m\) plots, what is the output? The probability of being the most different one?
-  \item
-    Then how many plots should be selected? Top 5? Or cumulative probability greater than a threshold?
-  \item
-    How to select the threshold?
-  \end{itemize}
-\end{itemize}
-
-The primary drawbacks of visual inference are
-
-infeasible in a large scale
-2. unfriendly to vision-impaired people 3. high finical cost and human cost
-4. time consuming
-
-Visual inference is a valid useful in diagnostics of predictive models
-
-To relieve human analysts of the job of diagnosing predictive models in the context of visual inference, computer vision
-
-Hence, to relieve human analysts of the job of diagnosing, automatically improve predictive models and assist vision-impaired people, this research aims to develop computer vision model to automate the reading
-1
-of data plots with application to predictive model diagnostics.
-The success of this research will enrich the applications of artificial intelligence in data visualization and improve the debugging process of artificial intelligence models.
-
-\hypertarget{model-architecture}{%
-\subsection{Model Architecture}\label{model-architecture}}
-
-Five computer vision models were trained to predict the probability of the given result plot being a null residual plot.
-
-\hypertarget{performance}{%
-\subsection{Performance}\label{performance}}
+Among the leaderboard of ILSCRB, we picked architectures that were handy to implement and with low top-5 error rate.
+This included VGG16 \autocite{simonyan_very_2015}, VGG19 \autocite{simonyan_very_2015}, Xception \autocite{chollet_xception_2017}, ResNet152V2 \autocite{he_identity_2016} and InceptionV3 \autocite{szegedy_rethinking_2015}.
+The top-5 error rate is the average number the model failed to include the answer among its top five guesses.
+The overview of all five transer learning models are given in Table \ref{tab:vi-model-summary}.
+Note that the depth is the number of layers in the model.
+
+On top of the pre-trained model, we added a 2D global average pooling layer \autocite{lin_network_2014}, followed by a fully-connected dense layer with 256 nodes, a dropout layer \autocite{srivastava_dropout_nodate} with 20\% dropout rate, and a output layer with softmax as activation function.
+A diagram of the model architecture is given in Figure \ref{fig:model-arch}.
+
+Global average pooling layer was a replacement of traditional fully-connected dense layers in CNN which flattened the feature map into a 1D tensor.
+Global average pooling layer took the average of the values over all the pixels in a channel, then flattened it into a 1D tensor.
+This enforced the correspondences between feature maps and categories, which made the feature maps easier to be interpreted as categories confidence maps.
+Further, unlike the dense layer, there was no parameters that needed to be optimized in global average pooling layer which reduce the risk of overfitting \autocite{lin_network_2014}.
+
+Dropout was primary used as a regularization technique to address overfitting.
+In training, the dropout layer would randomly temporaliy remove some nodes from the network with fixed probability \(p\), which in our case, we used \(p = 20\)\%.
+
+Five models were trained on 10000 null plots and 10000 actual data plots simulated using the same parameter settings as the human subject experiments and balanced across cubic model and heteroskedasticity model.
+20\% training data was used as validation set.
+Image augmentation was applied to increase the diversity of the dataset during training, with random rotation angle ranged from zero to five degrees and random brightness ranged from 95\% to 105\%.
+Optimizer Adam \autocite{kingma_adam_2017} was used with learning rate equal to 0.001.
+The loss function used in the model was binary cross-entropy, which can be expressed by: \[H = -N^{-1}\sum_{i=1}^{N}y_ilog(\hat{p}_i)+(1-y_i)log(1-\hat{p}_i),\] where \(N\) is the number of observations, \(y_i\) is a realization of a binary random variable, and \(\hat{p}_i\) is the predicted probability.
+
+Early stopping was used to reduce training time and to avoid overfitting. If the validation loss did not decrease for five iterations, then the training procedure would stop.
+
+The trained models were evaluated on 10000 null plots and 10000 actual data plots simulated by the same way as the training data.
+And the system was then evaluated on 2500 cubic lineups and 2500 heteroskedasticity lineups.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.90625in,height=5.11458in]{figures/vi_model.pdf}
+\caption{Architecture of the computer vision model. \label{fig:model-arch}}
+\end{figure}
 
@@
 
-\caption{\label{tab:null-true-eval}Performance of computer vision models on null residual plot and acutal data plot.}
+\caption{\label{tab:vi-model-summary}Overview of the computer vision models.}
+\centering
+\begin{tabular}[t]{lrrrrr}
+\toprule
+Model & Depth & Total number of parameters & Non-trainable parameters & Trainable parameters & Top-5 error rate\\
+\midrule
+VGG16 & 20 & 14846530 & 14714688 & 131842 & 0.901\\
+VGG19 & 23 & 20156226 & 20024384 & 131842 & 0.900\\
+Xception & 85 & 21386538 & 20861480 & 525058 & 0.945\\
+ResNet152V2 & 311 & 58856706 & 58331648 & 525058 & 0.942\\
+InceptionV3 & 193 & 22327842 & 21802784 & 525058 & 0.937\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\hypertarget{results-1}{%
+\subsection{Results}\label{results-1}}
+
+The performance of all five models evaluated on the test set is given in Table \ref{tab:null-true-eval}.
+From the result, we observed that the average balanced accuracy for all five models were around 80\%.
+Breaking down into null and actual data plot, it was found that
+
+\begin{table}
+
+\caption{\label{tab:null-true-eval}Performance of computer vision models on null residual plot and acutal data plot of test data.}
 \centering
@@
 
-\includegraphics{thesis_files/figure-latex/unnamed-chunk-44-1.pdf}
+\includegraphics{thesis_files/figure-latex/unnamed-chunk-45-1.pdf}
 
@@
 ## 1 Cubic              0.0056
-## 2 Heteroskedasticity 0.006
+## 2 Heteroskedasticity 0.0064
 \end{verbatim}
@@
 
-\hypertarget{visual-diagnostics-for-predictive-models-with}{%
-\chapter{Visual Diagnostics for Predictive Models with \ldots{}}\label{visual-diagnostics-for-predictive-models-with}}
+\hypertarget{automatic-visual-infernce-for-predictive-model-diagnostics}{%
+\chapter{Automatic Visual Infernce for Predictive Model Diagnostics}\label{automatic-visual-infernce-for-predictive-model-diagnostics}}
+
+Tools for Visual inference has been applied to
 

### 2022-03-10 Milestone_1/01-chap1.Rmd
except topic 3 desc

@@
              caption  = 'Overview of the computer vision models.', 
-             label = "vi-model-summary" )
+             label = "vi-model-summary" ) %>%
+  kableExtra::kable_styling(latex_options = "scale_down")
 ```
@@
 From the result, we observed that the average balanced accuracy for all five models were around 80%.
-Breaking down into null and actual data plot, it was found that
+Breaking down into null and actual data plot, it was found that models had a better performance on null plot detection.
+Or we could interpret it as our models tended to predict null plot as the answer.
+Further, models were less likely to predict null plot as the answer when the data was simulated from a cubic model, which resulted in slightly higher detection rate for actual data plot.
+In overall, these five model were good at detecting null plots but were not excellent in picking up residual plots with model defects especially for heteroskedasticity patterns.
+
+Combining with the overview of models presented in Table \ref{tab:vi-model-summary}, it was observed that models with large number of layers and parameters did worse in detecting actual data plot.
+One possible explanation of this phenomenon was the information was highly summarised and features corresponding to patterns might be lost in the depth of the architecture.
+Other than that, we did not find a strong relationship between detection rate and the meta variables of models.
 
@@
              booktabs = TRUE, 
-             caption  = 'Performance of computer vision models on null residual plot and acutal data plot of test data.', 
+             caption  = 'Performance of computer vision models on null residual plot and acutal data plot of test data.
+Models were tested on 10000 null plots and 10000 actual data plots simulated using the same parameter settings as the human subject experiments and balanced across cubic model and heteroskedasticity model.',
              label = "null-true-eval" ) %>%
@@
 
+By integrating all five models into the system, the automatic visual inference system was able to make decision on lineup.
+The performance of the system was given in Figure \ref{fig:system-performance}.
+According to the result, the estimated power of the system was close to the power of a visual test evaluated by two human subjects at large effect size.
+And for small effect size, the estimated power of the system surpassed the power of human visual test by a large margin, but it was still less powerful than the conventional test.
+Patterns related to the cubic model were easy to be detected by both human and the system, so both power curves reached one at large effect size.
+But for heteroskedasticity model, the system could not wisely tell the difference between the null plot and an actual data plot.
+We observed the similar result from the detection rate in Table \ref{tab:null-true-eval}.
+
 ```{r}
@@
   calc_p_value_multi(lineup_id = "set", detected = "detect", n_sel = "n_sel", cache_env = cache_env) -> cv_results
+
+lineup_eval_summary %>%
+  calc_p_value_multi(lineup_id = "set", detected = "detect", n_sel = "n_sel", cache_env = cache_env, comb = TRUE,
+                     n_eval = 1:5) -> cv_results2
 
@@
 
-```{r}
+```{r system-performance, fig.cap="Estimated power of the automatic visual inference system evaluated on 2500 lineups and 2500 lineups simulated using the cubic model and heteroskedasticity model respectively against the natural logarithm of the effect size.
+The dashed curves are power of conventional tests, the blue curves are the estimated power of the system, and the curves in other colours are the estimated power obtained from the experiments.
+"}
 merged_survey %>%
@@
   ggplot() +
-  geom_point(aes(log(effect_size), reject), alpha = 0.1, size = 1) +
+  # geom_point(aes(log(effect_size), reject), alpha = 0.1, size = 1) +
   geom_smooth(aes(log(effect_size), reject, group = "conventional"),
@@
 
+We also verified the type-I error of the system to see whether it was lower than the desired significance level imposed.
+It was done by evaluating 5000 lineups without any model defects.
+The results showed that the false positive rate was around $0.003$, which was smaller than 0.05.
+This suggested that we slightly overestimated the $p$-value of the automatic visual test, which leaded to less rejection of the null hypothesis.
+A simple way to address this issue was to relax the significance level.
+It was found that if we set $\alpha = 0.1$, then the false positive rate was around 0.015.
+And if we set $\alpha = 0.2$, then the rate increased to around 0.4.
+The reason of the overestimation of the $p$-value was worth to be researched in future studies.
+
+If we reduced the number of models used in the system, the estimated power of the automatic visual test would change.
+Figure \ref{fig:comb-system-power} showed five power curves corresponding to number of models ranged from one to five.
+It was found that as the number of models increased, the estimated power of the system increased, but the amount of improvement decreased.
+In fact, from Figure \ref{fig:comb-system-power}, we would not expect the estimated power increased a lot when the number of models greater than five.
+
+```{r comb-system-power, fig.cap="Estimated power of the automatic visual inference system for number of evaluation ranged from one to five and evaluated on 2500 lineups and 2500 lineups simulated using the cubic model and heteroskedasticity model respectively against the natural logarithm of the effect size.
+The dashed curves are power of conventional tests, the curves in other colours are the estimated power of the system with different number of evaluations.
+Decision made the system were drawn with black marker.
+"}
+cv_results2 %>%
+  left_join(select(eval_original_data, id, effect_size) %>%
+              group_by(id) %>%
+              summarise(effect_size = first(effect_size)),
+            by = c("lineup_id" = "id")) %>%
+  unnest_longer(col = p_value) %>%
+  mutate(reject = eval_p_value(p_value, tol = 1e-2)) %>%
+  mutate(reject = as.numeric(reject)) %>%
+  mutate(n_eval = factor(n_eval)) %>%
+  mutate(type = ifelse(lineup_id <= 2500, "cubic", "heteroskedasticity")) %>%
+  ggplot() +
+  geom_point(aes(log(effect_size), reject), alpha = 0.1, size = 1) +
+  geom_smooth(aes(log(effect_size), reject, group = "conventional"),
+              method = "glm", method.args = list(family = binomial), data = conv_result,
+              se = FALSE, size = 1, linetype = 2, col = "grey80") +
+  geom_smooth(aes(log(effect_size), reject, group = n_eval, col = n_eval),
+              method = "glm", method.args = list(family = binomial), size = 0.5, se = FALSE) +
+  facet_wrap(~type, scales = "free_x") +
+  labs(col = "Number of evaluations") +
+  theme_light(base_size = 5) +
+  theme(legend.position = "bottom") +
+  scale_colour_viridis_d(option = "B", begin = 0.5, end = 0.9) +
+  guides(col = guide_legend(nrow = 1, byrow=TRUE)) +
+  xlab("Log of effect size") +
+  ylab("Rejected")
+```
+
 ```{r}
@@
 saveRDS(cache_env, file = here::here("data/cache_env.rds"))
-
-pseudo_results %>%
-  mutate(reject = eval_p_value(p_value, tol = 1e-2)) %>%
-  mutate(type = ifelse(lineup_id <= 2500, "Cubic", "Heteroskedasticity")) %>%
-  group_by(type) %>%
-  summarise(rate = mean(reject))
 ```
@@
 
-Tools for Visual inference has been applied to
+...
 

### 2022-03-10 Milestone_1/_book/thesis.tex
except topic 3 desc

@@
 \centering
+\resizebox{\linewidth}{!}{
 \begin{tabular}[t]{lrrrrr}
@@
 \bottomrule
-\end{tabular}
+\end{tabular}}
 \end{table}
@@
 From the result, we observed that the average balanced accuracy for all five models were around 80\%.
-Breaking down into null and actual data plot, it was found that
+Breaking down into null and actual data plot, it was found that models had a better performance on null plot detection.
+Or we could interpret it as our models tended to predict null plot as the answer.
+Further, models were less likely to predict null plot as the answer when the data was simulated from a cubic model, which resulted in slightly higher detection rate for actual data plot.
+In overall, these five model were good at detecting null plots but were not excellent in picking up residual plots with model defects especially for heteroskedasticity patterns.
+
+Combining with the overview of models presented in Table \ref{tab:vi-model-summary}, it was observed that models with large number of layers and parameters did worse in detecting actual data plot.
+One possible explanation of this phenomenon was the information was highly summarised and features corresponding to patterns might be lost in the depth of the architecture.
+Other than that, we did not find a strong relationship between detection rate and the meta variables of models.
 
@@
 
-\caption{\label{tab:null-true-eval}Performance of computer vision models on null residual plot and acutal data plot of test data.}
+\caption{\label{tab:null-true-eval}Performance of computer vision models on null residual plot and acutal data plot of test data.
+Models were tested on 10000 null plots and 10000 actual data plots simulated using the same parameter settings as the human subject experiments and balanced across cubic model and heteroskedasticity model.}
 \centering
@@
 
-\includegraphics{thesis_files/figure-latex/unnamed-chunk-45-1.pdf}
-
-\begin{verbatim}
-## # A tibble: 2 x 2
-##   type                 rate
-##   <chr>               <dbl>
-## 1 Cubic              0.0056
-## 2 Heteroskedasticity 0.0064
-\end{verbatim}
+By integrating all five models into the system, the automatic visual inference system was able to make decision on lineup.
+The performance of the system was given in Figure \ref{fig:system-performance}.
+According to the result, the estimated power of the system was close to the power of a visual test evaluated by two human subjects at large effect size.
+And for small effect size, the estimated power of the system surpassed the power of human visual test by a large margin, but it was still less powerful than the conventional test.
+Patterns related to the cubic model were easy to be detected by both human and the system, so both power curves reached one at large effect size.
+But for heteroskedasticity model, the system could not wisely tell the difference between the null plot and an actual data plot.
+We observed the similar result from the detection rate in Table \ref{tab:null-true-eval}.
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/system-performance-1.pdf}
+\caption{\label{fig:system-performance}Estimated power of the automatic visual inference system evaluated on 2500 lineups and 2500 lineups simulated using the cubic model and heteroskedasticity model respectively against the natural logarithm of the effect size.
+The dashed curves are power of conventional tests, the blue curves are the estimated power of the system, and the curves in other colours are the estimated power obtained from the experiments.}
+\end{figure}
+
+We also verified the type-I error of the system to see whether it was lower than the desired significance level imposed.
+It was done by evaluating 5000 lineups without any model defects.
+The results showed that the false positive rate was around \(0.003\), which was smaller than 0.05.
+This suggested that we slightly overestimated the \(p\)-value of the automatic visual test, which leaded to less rejection of the null hypothesis.
+A simple way to address this issue was to relax the significance level.
+It was found that if we set \(\alpha = 0.1\), then the false positive rate was around 0.015.
+And if we set \(\alpha = 0.2\), then the rate increased to around 0.4.
+The reason of the overestimation of the \(p\)-value was worth to be researched in future studies.
+
+If we reduced the number of models used in the system, the estimated power of the automatic visual test would change.
+Figure \ref{fig:comb-system-power} showed five power curves corresponding to number of models ranged from one to five.
+It was found that as the number of models increased, the estimated power of the system increased, but the amount of improvement decreased.
+In fact, from Figure \ref{fig:comb-system-power}, we would not expect the estimated power increased a lot when the number of models greater than five.
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/comb-system-power-1.pdf}
+\caption{\label{fig:comb-system-power}Estimated power of the automatic visual inference system for number of evaluation ranged from one to five and evaluated on 2500 lineups and 2500 lineups simulated using the cubic model and heteroskedasticity model respectively against the natural logarithm of the effect size.
+The dashed curves are power of conventional tests, the curves in other colours are the estimated power of the system with different number of evaluations.
+Decision made the system were drawn with black marker.}
+\end{figure}
 
@@
 
-Tools for Visual inference has been applied to
+\ldots{}
 

### 2022-03-10 Milestone_1/01-chap1.Rmd
add structure of report

@@
 
-# Research Problem
+## Structure of the Report
+
+The remainder of the report is organised as follows.
+Chapter \ref{se:research-problem} outlines the three research problems.
+Chapter \ref{se:overview} gives an overview of the thesis including descriptions of three projects.
+Chapter \ref{#se:project-1} presents the human subject experiments conducted in the first project along with the prototype of the automatic visual inference system.
+Chapter \ref{se:project-2}
+
+<!-- presents an almost finished paper and its associated R package, ferrn, on visual diagnostics of projection pursuit optimisation.
+Chapter 3 introduces the stochastic optimal control problem and proposes visual diagnostics for its optimisers.
+Chapter 4 describes the interpretability of stochastic optimal control for decision making.
+And lastly, Chapter 5 outlines the timeline for future work.
+-->
+
+# Research Problem {#se:research-problem}
 
@@
 
-# Overview of the Thesis
+# Overview of the Thesis {se:overview}
 
@@
 
-# Automatic Visual Statistical Inference for linear regression diagnostics
+# Automatic Visual Statistical Inference for linear regression diagnostics {#se:project-1}
 
@@
 
-# General-Purpose Automatic Visual Statistical Inference System
+# General-Purpose Automatic Visual Statistical Inference System {#se:project-2}
 
@@
 
-# Automatic Visual Infernce for Predictive Model Diagnostics
+# Automatic Visual Infernce for Predictive Model Diagnostics {#se:project-3}
 

### 2022-03-10 Milestone_1/_book/thesis.tex
add structure of report

@@
 
-\hypertarget{research-problem}{%
-\chapter{Research Problem}\label{research-problem}}
+\hypertarget{structure-of-the-report}{%
+\section{Structure of the Report}\label{structure-of-the-report}}
+
+The remainder of the report is organised as follows.
+Chapter \ref{se:research-problem} outlines the three research problems.
+Chapter \ref{se:overview} gives an overview of the thesis including descriptions of three projects.
+Chapter \ref{#se:project-1} presents the human subject experiments conducted in the first project along with the prototype of the automatic visual inference system.
+Chapter \ref{se:project-2}
+
+\hypertarget{se:research-problem}{%
+\chapter{Research Problem}\label{se:research-problem}}
 
@@
 
-\hypertarget{overview-of-the-thesis}{%
-\chapter{Overview of the Thesis}\label{overview-of-the-thesis}}
+\hypertarget{overview-of-the-thesis-seoverview}{%
+\chapter{Overview of the Thesis \{se:overview\}}\label{overview-of-the-thesis-seoverview}}
 
@@
 
-\hypertarget{automatic-visual-statistical-inference-for-linear-regression-diagnostics}{%
-\chapter{Automatic Visual Statistical Inference for linear regression diagnostics}\label{automatic-visual-statistical-inference-for-linear-regression-diagnostics}}
+\hypertarget{se:project-1}{%
+\chapter{Automatic Visual Statistical Inference for linear regression diagnostics}\label{se:project-1}}
 
@@
 
-\hypertarget{general-purpose-automatic-visual-statistical-inference-system}{%
-\chapter{General-Purpose Automatic Visual Statistical Inference System}\label{general-purpose-automatic-visual-statistical-inference-system}}
+\hypertarget{se:project-2}{%
+\chapter{General-Purpose Automatic Visual Statistical Inference System}\label{se:project-2}}
 
@@
 
-\hypertarget{automatic-visual-infernce-for-predictive-model-diagnostics}{%
-\chapter{Automatic Visual Infernce for Predictive Model Diagnostics}\label{automatic-visual-infernce-for-predictive-model-diagnostics}}
+\hypertarget{se:project-3}{%
+\chapter{Automatic Visual Infernce for Predictive Model Diagnostics}\label{se:project-3}}
 

### 2022-03-10 Milestone_1/01-chap1.Rmd
fix diagram

@@
 Chapter \ref{se:overview} gives an overview of the thesis including descriptions of three projects.
-Chapter \ref{#se:project-1} presents the human subject experiments conducted in the first project along with the prototype of the automatic visual inference system.
+Chapter \ref{se:project-1} presents the human subject experiments conducted in the first project along with the prototype of the automatic visual inference system.
 Chapter \ref{se:project-2}
@@
 
-# Overview of the Thesis {se:overview}
+# Overview of the Thesis {#se:overview}
 

### 2022-03-10 Milestone_1/_book/thesis.tex
fix diagram

@@
 
-\hypertarget{overview-of-the-thesis-seoverview}{%
-\chapter{Overview of the Thesis \{se:overview\}}\label{overview-of-the-thesis-seoverview}}
+\hypertarget{se:overview}{%
+\chapter{Overview of the Thesis}\label{se:overview}}
 

### 2022-03-10 Milestone_1/01-chap1.Rmd
middle effect size example

@@
 Our goal was to quantify the ability of human reading residual plot, such that the comparison between human and computer vision model can be made.
+It was also found that there was a huge gap between the conventional curves and the visual test curves at the middle of the plot.
+We further analysed the lineup with the corresponding effect size.
+Figure \ref{fig:mid-cubic-lineup-example} and \ref{fig:mid-heter-lineup-example} showed that human was indeed hard to identify the pattern at this level of difficulty.
+However, we as researchers of this project could tell which was the actual data plot by carefully comparing the plots of the lineup since we knew all possible shapes of the residual plot beforehand.
 
@@
   ylab("Estimated power")
+```
+
+```{r mid-cubic-lineup-example, fig.cap="An example lineup for the cubic model with effect size equal to 0.18.
+The answer to this linup is seven.
+It was not suprised participants could not confidently identify the actual data plot given the pattern was very weak.
+"}
+# merged_survey %>%
+#   filter(exp == 2 & type == "cubic" & -0.5< log(effect_size) & log(effect_size) < 0.5) %>%
+#   head()
+
+VI_MODEL$plot_lineup(mutate(lineup_dat_big[[59]]$data, .fitted = .pred), 
+                        remove_axis = TRUE, 
+                        remove_grid_line = TRUE,
+                        theme = theme_light())
+```
+
+```{r mid-heter-lineup-example, fig.cap="An example lineup for the heteroskedasticity model with effect size equal to 17.5.
+The answer to this linup is two.
+The butterfly shape could merely being recognised.
+"}
+# merged_survey %>%
+#   filter(exp == 2 & type == "heteroskedasticity" & 2< log(effect_size) & log(effect_size) < 3) %>%
+#   .$lineup_id %>%
+#   factor()
+
+VI_MODEL$plot_lineup(mutate(lineup_dat_big[[136]]$data, .fitted = .pred), 
+                        remove_axis = TRUE, 
+                        remove_grid_line = TRUE,
+                        theme = theme_light())
 ```

### 2022-03-10 Milestone_1/_book/thesis.tex
middle effect size example

@@
 The dashed curvses are the power of F-test and Breusch--Pagan test respectively.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/mid-cubic-lineup-example-1.pdf}
+\caption{\label{fig:mid-cubic-lineup-example}An example lineup for the cubic model with effect size equal to 0.18.
+The answer to this linup is seven.
+It was not suprised participants could not confidently identify the actual data plot given the pattern was very weak.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{thesis_files/figure-latex/mid-heter-lineup-example-1.pdf}
+\caption{\label{fig:mid-heter-lineup-example}An example lineup for the heteroskedasticity model with effect size equal to 17.5.
+The answer to this linup is two.
+The butterfly shape could merely being recognised.}
 \end{figure}

### 2022-03-10 Milestone_1/01-chap1.Rmd
fix typo 1

@@
 Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
-One such example is finding an over-interpretation of the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurence [@roy_chowdhury_using_2015].
+One such example is finding an over-interpretation of the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurrence [@roy_chowdhury_using_2015].
 
@@
 
-Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualizatino" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots [@gelman_exploratory_2004].
+Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualization" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots [@gelman_exploratory_2004].
 Later, in the Bayesian context, data plots was systematically considered as model diagnostics by taking advantage of the data simulated from the assumed statistical models [@gelman_bayesian_2003; @gelman_exploratory_2004].
@@
 For model diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the null hypothesis.
-The same assumed model, such as classical linear regression model, can be rejected by many reasons with residual plot, including nonlinearity and heteroskedasticity as shown in Figure \@ref(fig:residual-plot-cubic-heter).
+The same assumed model, such as classical linear regression model, can be rejected by many reasons with residual plot, including non-linearity and heteroskedasticity as shown in Figure \@ref(fig:residual-plot-cubic-heter).
 
@@
 As @buja_statistical_2009 suggested, $H_0$ is usually composited by a collection of distributions controlled by nuisance parameters.
-Since statistical models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling.
+Since statistical models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling.
 
@@
 Chapter \ref{se:overview} gives an overview of the thesis including descriptions of three projects.
-Chapter \ref{se:project-1} presents the human subject experiments conducted in the first project along with the prototype of the automatic visual inference system.
-Chapter \ref{se:project-2}
+Chapter \ref{se:project-1} presents the prototype of the automatic visual inference system along with the results of two human subject experiments.
+Chapter \ref{se:project-2} introduces the general-purpose automatic visual inference system.
+Chapter \ref{se:project-3} describes the deployment of the automatic system.
+Chapter \ref{se:timeline} presents the timeline of the three projects.
 
@@
 
-3. Validate and apply the automatic visual inference system to predictive model diagnostics on empirical data.
+3. Deploy the automatic visual inference system as an adaptive online application by applying online machine learning.
 
@@
 
-The third project aims to apply the automatic system to predictive model diagnostics on empirical data.
-Potential predictive models considered by this research are linear model, generalized linear models and non-parametric regression.
-Multiple human subject experiments will need to be conducted to validate the effectiveness of the automatic system.
+The third projects aims to package and deploy the automatic system as an adaptive online application such that empirical data can be collected and used in improving the system.
+Online machine learning technique will be used to update the weights of computer vision models dynamically.
 
@@
 
-To collect data of human performance on reading residual plot of linear regression model with nonlinearity and heteroskedasticity defects, two human experiments were conducted.
+To collect data of human performance on reading residual plot of linear regression model with non-linearity and heteroskedasticity defects, two human experiments were conducted.
 Participants of both experiments were recruited using an online platform called Prolific [@prolific_prolific_2022].
@@
 This avoided participants seeing the lineup beforehand.
-The front-end of the website was built using a JavaScript [@flanagan_javascript_2006] library jsPsych [@de_leeuw_jspsych_2015] which is specialized in creating online behavioral experiments.
+The front-end of the website was built using a JavaScript [@flanagan_javascript_2006] library jsPsych [@de_leeuw_jspsych_2015] which is specialized in creating online behavioural experiments.
 One of the reasons we chose to use this library was it had a modularized but highly customizable template which could record participant's response time automatically.
 This is essential for us to confirm the quality of the data by checking exceptionally fast or slow responses.
-Figure \@ref(fig:diagram-experiment) summarizes the online experiment setup.
+Figure \@ref(fig:diagram-experiment) summarizes the online experiment set-up.
 
@@
 
-Next we will discuss the simulation setup for this study.
+Next we will discuss the simulation set-up for this study.
 The experiment data was simulated by the use of the programming language `R` [@r_core_team_r_2021].
-For the ease of reproducibility, functions to build models, simulate data from models, produce lineup with data, allocate stimulis for subjects and evaluate subject responses from this study are bundled in the package `visage` with a unique object-oriented programming system built upon the environment feature of `R`.
+For the ease of reproducibility, functions to build models, simulate data from models, produce lineup with data, allocate stimulus for subjects and evaluate subject responses from this study are bundled in the package `visage` with a unique object-oriented programming system built upon the environment feature of `R`.
 In the description of the simulation, corresponding functionalities of the package will be introduced.
@@
 This defines a cubic relationship between $\boldsymbol{Y}$, $\boldsymbol{X}$ and $\boldsymbol{Z}$.
-Meanwhile, to create nonlinearity defect, the null model followed the assumptions of the classical normal linear regression model (CNLRM), fitted by OLS is:
+Meanwhile, to create non-linearity defect, the null model followed the assumptions of the classical normal linear regression model (CNLRM), fitted by OLS is:
 \begin{equation} \label{eq:cubic_null}
@@
 
-Under the cubic model setting, there is an exact conventional test for testing the nonlinearity defect, which is F-test.
+Under the cubic model setting, there is an exact conventional test for testing the non-linearity defect, which is F-test.
 For F-test, the null hypothesis is $H_0:a=b=0$, and the alternative hypothesis is $H_1:\text{at least one of them} \neq 0$.
@@
 We could plot the one standard deviation around the residuals to indicate the region about 68% of the residuals should landed on.
-Figure \@ref(fig:hetershape) illustrates different shapes of residuals undder various values of $a$ and $b$.
+Figure \@ref(fig:hetershape) illustrates different shapes of residuals under various values of $a$ and $b$.
 From the plot, it can be observed that if $a = 0$, the residual plot looks like the shape of butterfly.
@@
 Uniform and normal distribution were symmetric and common.
-Adjusted lognormal distribution and adjusted negative lognormal distribution provided right-skewed and left-skewed density respectively.
+Adjusted log-normal distribution and adjusted negative log-normal distribution provided right-skewed and left-skewed density respectively.
 These distributions were chosen such that most the realizations will fall between $-1$ and $1$.
@@
 $z_n$ itself was a discrete uniform distribution with outcomes $\{10,11,12,13,14,15,16,17,18,19,20\}$, which defined the number of possible values $Z_i$ could take.
-As shown in Figure \@ref(fig:discreteex), this setup would create discreteness in residual plot, which could enrich the pool of visual patterns.
+As shown in Figure \@ref(fig:discreteex), this set-up would create discreteness in residual plot, which could enrich the pool of visual patterns.
 
@@
 $\boldsymbol{X}$ used in the heteroskedasticity model was a combination of $\boldsymbol{X}$ and $\boldsymbol{Z}$ used in the cubic model.
-It could be one of the five distributions mentioned above - normal distribution, uniform distribution, adjusted lognormal distribution, adjusted negative lognormal distribution and discrete uniform distribution.
+It could be one of the five distributions mentioned above - normal distribution, uniform distribution, adjusted log-normal distribution, adjusted negative log-normal distribution and discrete uniform distribution.
 
@@
 Fitted power of visual test with three to five evaluations were almost identical to each other in regards of effect size.
-In addition, all five curves peaked at one as effect size increased, suggesting that identification of nonlinearity as a visual task can be completed reliably by human when the effect size is large enough.
+In addition, all five curves peaked at one as effect size increased, suggesting that identification of non-linearity as a visual task can be completed reliably by human when the effect size is large enough.
 
@@
 As expected, the coefficient of $\sigma$ was negative as it controlled the noise level.
-Lineup simulated with uniform distribution of $\boldsymbol{X}$ had the highest power, followed by negative lognormal, normal, and lognormal.
+Lineup simulated with uniform distribution of $\boldsymbol{X}$ had the highest power, followed by negative log-normal, normal, and log-normal.
 Discreteness in $\boldsymbol{Z}$ decreased the visual power.
@@
 As expected, the coefficient of $b$ was positive.
-For different distributions of $\boldsymbol{X}$, the one with highest power was uniform distribution, followed by normal, discrete uniform, lognormal and negative lognormal.
-The coeffient of the number of observations was positive, but it was not significant.
+For different distributions of $\boldsymbol{X}$, the one with highest power was uniform distribution, followed by normal, discrete uniform, log-normal and negative log-normal.
+The coefficient of the number of observations was positive, but it was not significant.
 
 In overall, we did not have any unexpected findings except the visual power related to different distributions of $\boldsymbol{X}$ in cubic model.
-We expected lineup simulated with uniform distribution would have the highest power, followed by normal, lognormal and negative lognormal, since we observed similar results in pilot study.
+We expected lineup simulated with uniform distribution would have the highest power, followed by normal, log-normal and negative log-normal, since we observed similar results in pilot study.
 
@@
 
-Considering there are different approches of AI, it is necessary to clarify the type of AI we aimed to build.
+Considering there are different approaches of AI, it is necessary to clarify the type of AI we aimed to build.
 Understanding the human thought process or mimicking the human vision mechanism were not of interest in this research.
@@
 The top-5 error rate is the average number the model failed to include the answer among its top five guesses.
-The overview of all five transer learning models are given in Table \ref{tab:vi-model-summary}.
+The overview of all five transfer learning models are given in Table \ref{tab:vi-model-summary}.
 Note that the depth is the number of layers in the model.
@@
 Dropout was primary used as a regularization technique to address overfitting.
-In training, the dropout layer would randomly temporaliy remove some nodes from the network with fixed probability $p$, which in our case, we used $p = 20$%.
+In training, the dropout layer would randomly temporarily remove some nodes from the network with fixed probability $p$, which in our case, we used $p = 20$%.
 
@@
 By adopting the techniques developed in query by example content-based image retrieval, we aims to build automatic general-purpose visual inference system based on image comparison in the second project.
-
-# Automatic Visual Infernce for Predictive Model Diagnostics {#se:project-3}
-
-...
-
-# Timeline
+General-purpose means that we will not make any assumption about the type of the plot and the data generating process such that the system will be able to evaluate any lineups asking for selecting the most different plot(s).
+
+# OAVIS: An online automatic visual inference system {#se:project-3}
+
+Online machine learning is a technique to train machine learning model with data arriving in a sequential order.
+Models trained with online learning can be updated immediately for any new arrival data, which overcomes the drawbacks of the traditional batch learning where training data required to be available prior to the training.
+It is suitable for large-scale machine learning tasks in real-world data analytics applications [@].
+
+The third projects aims to deploy the automatic visual inference system as an online application at a public-accessible website.
+The web application will evaluate lineups uploaded by users, and return the $p$-value with the selections made by the system.
+Meanwhile, as an option, users will be asked to also evaluate the lineup and give their selections.
+With the user selection data, the system will use online machine learning theory to update the built-in computer vision models such that the performance of the system can be improved.
+
+# Timeline {#se:timeline}
 
@@
 \addlinespace[0.5em]
+\multicolumn{2}{l}{\textit{\textbf{2021}}}\\
+\hspace{4em}March - August & Design the study website and the experiments\\ 
+\hspace{4em}September & Apply for human research ethics approval\\ 
+\hspace{4em}October & Test the study website\\ 
+\hspace{4em}November & Conduct the first two human subject experiments\\ 
+\hspace{4em}December & Develop the \textit{R} package \textit{visage}\\ 
+\addlinespace[0.5em]
+\hline
+\addlinespace[0.5em]
 \multicolumn{2}{l}{\textit{\textbf{2022}}}\\
+\hspace{4em}January & Develop the computer vision models\\
+\hspace{4em}February & Prepare for the first milestone\\
 \hspace{4em}March & Extend the scope of violations of classical normal linear\\ 
@@
 \addlinespace[0.5em]
-\hspace{4em}April & Conduct human subject experiment on the extended model\\ 
+\hspace{4em}April & Conduct human subject experiment for the extended model\\ 
 \addlinespace[0.5em]
@@
 \addlinespace[0.5em]
-\hspace{4em}April - December & Apply the automatic visual inference system to predictive\\
-\addlinespace[0.5em]
-\hspace{4em} & model diagnostics\\
+\hspace{4em}April - December & Deploy the automatic visual inference system as \\
+\addlinespace[0.5em]
+\hspace{4em} & a web application\\
 \addlinespace[0.5em]

### 2022-03-10 Milestone_1/_book/thesis.tex
fix typo 1

@@
 Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
-One such example is finding an over-interpretation of the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurence \autocite{roy_chowdhury_using_2015}.
+One such example is finding an over-interpretation of the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurrence \autocite{roy_chowdhury_using_2015}.
 
@@
 
-Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title ``Inference for Data Visualizatino'' by \textcite{buja_inference_1999} as an idea to address the issue of valid inference for visual discoveries of data plots \autocite{gelman_exploratory_2004}.
+Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title ``Inference for Data Visualization'' by \textcite{buja_inference_1999} as an idea to address the issue of valid inference for visual discoveries of data plots \autocite{gelman_exploratory_2004}.
 Later, in the Bayesian context, data plots was systematically considered as model diagnostics by taking advantage of the data simulated from the assumed statistical models \autocite{gelman_bayesian_2003,gelman_exploratory_2004}.
@@
 For model diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the null hypothesis.
-The same assumed model, such as classical linear regression model, can be rejected by many reasons with residual plot, including nonlinearity and heteroskedasticity as shown in Figure \ref{fig:residual-plot-cubic-heter}.
+The same assumed model, such as classical linear regression model, can be rejected by many reasons with residual plot, including non-linearity and heteroskedasticity as shown in Figure \ref{fig:residual-plot-cubic-heter}.
 
@@
 As \textcite{buja_statistical_2009} suggested, \(H_0\) is usually composited by a collection of distributions controlled by nuisance parameters.
-Since statistical models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to so called ``reference distribution'' by applying one of the three methods: (i) sampling from a conditional distribution given a minimimal sufficient statistic under \(H_0\), (ii) parametric bootstrap sampling with nuisance parameters estimated under \(H_0\), and (iii) Bayesian posterior predictive sampling.
+Since statistical models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to so called ``reference distribution'' by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under \(H_0\), (ii) parametric bootstrap sampling with nuisance parameters estimated under \(H_0\), and (iii) Bayesian posterior predictive sampling.
 
@@
 \item
-  Validate and apply the automatic visual inference system to predictive model diagnostics on empirical data.
+  Deploy the automatic visual inference system as an adaptive online application by applying online machine learning.
 \end{enumerate}
@@
 
-To collect data of human performance on reading residual plot of linear regression model with nonlinearity and heteroskedasticity defects, two human experiments were conducted.
+To collect data of human performance on reading residual plot of linear regression model with non-linearity and heteroskedasticity defects, two human experiments were conducted.
 Participants of both experiments were recruited using an online platform called Prolific \autocite{prolific_prolific_2022}.
@@
 This avoided participants seeing the lineup beforehand.
-The front-end of the website was built using a JavaScript \autocite{flanagan_javascript_2006} library jsPsych \autocite{de_leeuw_jspsych_2015} which is specialized in creating online behavioral experiments.
+The front-end of the website was built using a JavaScript \autocite{flanagan_javascript_2006} library jsPsych \autocite{de_leeuw_jspsych_2015} which is specialized in creating online behavioural experiments.
 One of the reasons we chose to use this library was it had a modularized but highly customizable template which could record participant's response time automatically.
 This is essential for us to confirm the quality of the data by checking exceptionally fast or slow responses.
-Figure \ref{fig:diagram-experiment} summarizes the online experiment setup.
+Figure \ref{fig:diagram-experiment} summarizes the online experiment set-up.
 
@@
 
-Next we will discuss the simulation setup for this study.
+Next we will discuss the simulation set-up for this study.
 The experiment data was simulated by the use of the programming language \texttt{R} \autocite{r_core_team_r_2021}.
-For the ease of reproducibility, functions to build models, simulate data from models, produce lineup with data, allocate stimulis for subjects and evaluate subject responses from this study are bundled in the package \texttt{visage} with a unique object-oriented programming system built upon the environment feature of \texttt{R}.
+For the ease of reproducibility, functions to build models, simulate data from models, produce lineup with data, allocate stimulus for subjects and evaluate subject responses from this study are bundled in the package \texttt{visage} with a unique object-oriented programming system built upon the environment feature of \texttt{R}.
 In the description of the simulation, corresponding functionalities of the package will be introduced.
@@
 This defines a cubic relationship between \(\boldsymbol{Y}\), \(\boldsymbol{X}\) and \(\boldsymbol{Z}\).
-Meanwhile, to create nonlinearity defect, the null model followed the assumptions of the classical normal linear regression model (CNLRM), fitted by OLS is:
+Meanwhile, to create non-linearity defect, the null model followed the assumptions of the classical normal linear regression model (CNLRM), fitted by OLS is:
 \begin{equation} \label{eq:cubic_null}
@@
 
-Under the cubic model setting, there is an exact conventional test for testing the nonlinearity defect, which is F-test.
+Under the cubic model setting, there is an exact conventional test for testing the non-linearity defect, which is F-test.
 For F-test, the null hypothesis is \(H_0:a=b=0\), and the alternative hypothesis is \(H_1:\text{at least one of them} \neq 0\).
@@
 We could plot the one standard deviation around the residuals to indicate the region about 68\% of the residuals should landed on.
-Figure \ref{fig:hetershape} illustrates different shapes of residuals undder various values of \(a\) and \(b\).
+Figure \ref{fig:hetershape} illustrates different shapes of residuals under various values of \(a\) and \(b\).
 From the plot, it can be observed that if \(a = 0\), the residual plot looks like the shape of butterfly.
@@
 Uniform and normal distribution were symmetric and common.
-Adjusted lognormal distribution and adjusted negative lognormal distribution provided right-skewed and left-skewed density respectively.
+Adjusted log-normal distribution and adjusted negative log-normal distribution provided right-skewed and left-skewed density respectively.
 These distributions were chosen such that most the realizations will fall between \(-1\) and \(1\).
@@
 \(z_n\) itself was a discrete uniform distribution with outcomes \(\{10,11,12,13,14,15,16,17,18,19,20\}\), which defined the number of possible values \(Z_i\) could take.
-As shown in Figure \ref{fig:discreteex}, this setup would create discreteness in residual plot, which could enrich the pool of visual patterns.
+As shown in Figure \ref{fig:discreteex}, this set-up would create discreteness in residual plot, which could enrich the pool of visual patterns.
 
@@
 \(\boldsymbol{X}\) used in the heteroskedasticity model was a combination of \(\boldsymbol{X}\) and \(\boldsymbol{Z}\) used in the cubic model.
-It could be one of the five distributions mentioned above - normal distribution, uniform distribution, adjusted lognormal distribution, adjusted negative lognormal distribution and discrete uniform distribution.
+It could be one of the five distributions mentioned above - normal distribution, uniform distribution, adjusted log-normal distribution, adjusted negative log-normal distribution and discrete uniform distribution.
 
@@
 As expected, the coefficient of \(\sigma\) was negative as it controlled the noise level.
-Lineup simulated with uniform distribution of \(\boldsymbol{X}\) had the highest power, followed by negative lognormal, normal, and lognormal.
+Lineup simulated with uniform distribution of \(\boldsymbol{X}\) had the highest power, followed by negative log-normal, normal, and log-normal.
 Discreteness in \(\boldsymbol{Z}\) decreased the visual power.
@@
 As expected, the coefficient of \(b\) was positive.
-For different distributions of \(\boldsymbol{X}\), the one with highest power was uniform distribution, followed by normal, discrete uniform, lognormal and negative lognormal.
-The coeffient of the number of observations was positive, but it was not significant.
+For different distributions of \(\boldsymbol{X}\), the one with highest power was uniform distribution, followed by normal, discrete uniform, log-normal and negative log-normal.
+The coefficient of the number of observations was positive, but it was not significant.
 
 In overall, we did not have any unexpected findings except the visual power related to different distributions of \(\boldsymbol{X}\) in cubic model.
-We expected lineup simulated with uniform distribution would have the highest power, followed by normal, lognormal and negative lognormal, since we observed similar results in pilot study.
+We expected lineup simulated with uniform distribution would have the highest power, followed by normal, log-normal and negative log-normal, since we observed similar results in pilot study.
 
@@
 Dropout was primary used as a regularization technique to address overfitting.
-In training, the dropout layer would randomly temporaliy remove some nodes from the network with fixed probability \(p\), which in our case, we used \(p = 20\)\%.
+In training, the dropout layer would randomly temporarily remove some nodes from the network with fixed probability \(p\), which in our case, we used \(p = 20\)\%.
 
@@
 By adopting the techniques developed in query by example content-based image retrieval, we aims to build automatic general-purpose visual inference system based on image comparison in the second project.
+General-purpose means that we will not make any assumption about the type of the plot and the data generating process such that the system will be able to evaluate any lineups asking for selecting the most different plot(s).
 
 \hypertarget{se:project-3}{%
-\chapter{Automatic Visual Infernce for Predictive Model Diagnostics}\label{se:project-3}}
-
-\ldots{}
-
-\hypertarget{timeline}{%
-\chapter{Timeline}\label{timeline}}
+\chapter{OAVIS: An online automatic visual inference system}\label{se:project-3}}
+
+Online machine learning is a technique to train machine learning model with data arriving in a sequential order.
+Models trained with online learning can be updated immediately for any new arrival data, which overcomes the drawbacks of the traditional batch learning where training data required to be available prior to the training.
+It is suitable for large-scale machine learning tasks in real-world data analytics applications {[}@{]}.
+
+The third projects aims to deploy the automatic visual inference system as an online application at a public-accessible website.
+The web application will evaluate lineups uploaded by users, and return the \(p\)-value with the selections made by the system.
+Meanwhile, as an option, users will be asked to also evaluate the lineup and give their selections.
+With the user selection data, the system will use online machine learning theory to update the built-in computer vision models such that the performance of the system can be improved.
+
+\hypertarget{se:timeline}{%
+\chapter{Timeline}\label{se:timeline}}
 

### 2022-03-10 Milestone_1/01-chap1.Rmd
fix reference

@@
 
-Similarly, @wilkinson_graph-theoretic_2005 proposed the work called graph theoretic scagnostics, which adopted the idea of "scagnostics" - scatter plot diagnostics from (can't find the 1984 citation).
+Similarly, @wilkinson_graph-theoretic_2005 proposed the work called graph theoretic scagnostics, which adopted the idea of "scagnostics" - scatter plot diagnostics from John and Paul Tukey.
 It includes 9 computable scagnostics measures defined on planar proximity graphs: "Outlying", "Convex", "Skinny", "Stringy", "Straight", "Monotonic", "Skewed", "Clumpy" and "Striated" which attempts to describe outliers, shape, density, trend and coherence of the data.
@@
 Because the first project is still ongoing, we consider the result we have is not enough to write as a paper at the moment.
-Hence, the material provided in this section is two main parts of the draft paper that will contain enough details for understanding the project.
-
-## Human Subject Exeriments
+Hence, the materials provided in this chapter are two main parts of the draft paper that will contain enough details for understanding the project.
+
+## Human Subject Experiments
 
@@
 
-On top of the pre-trained model, we added a 2D global average pooling layer [@lin_network_2014], followed by a fully-connected dense layer with 256 nodes, a dropout layer [@srivastava_dropout_nodate] with 20% dropout rate, and a output layer with softmax as activation function.
+On top of the pre-trained model, we added a 2D global average pooling layer [@lin_network_2014], followed by a fully-connected dense layer with 256 nodes, a dropout layer [@srivastava_dropout_2014] with 20% dropout rate, and a output layer with softmax as activation function.
 A diagram of the model architecture is given in Figure \ref{fig:model-arch}.

### 2022-03-10 Milestone_1/_book/thesis.tex
fix reference

@@
 
-Similarly, \textcite{wilkinson_graph-theoretic_2005} proposed the work called graph theoretic scagnostics, which adopted the idea of ``scagnostics'' - scatter plot diagnostics from (can't find the 1984 citation).
+Similarly, \textcite{wilkinson_graph-theoretic_2005} proposed the work called graph theoretic scagnostics, which adopted the idea of ``scagnostics'' - scatter plot diagnostics from John and Paul Tukey.
 It includes 9 computable scagnostics measures defined on planar proximity graphs: ``Outlying'', ``Convex'', ``Skinny'', ``Stringy'', ``Straight'', ``Monotonic'', ``Skewed'', ``Clumpy'' and ``Striated'' which attempts to describe outliers, shape, density, trend and coherence of the data.
@@
 Because the first project is still ongoing, we consider the result we have is not enough to write as a paper at the moment.
-Hence, the material provided in this section is two main parts of the draft paper that will contain enough details for understanding the project.
-
-\hypertarget{human-subject-exeriments}{%
-\section{Human Subject Exeriments}\label{human-subject-exeriments}}
+Hence, the materials provided in this chapter are two main parts of the draft paper that will contain enough details for understanding the project.
+
+\hypertarget{human-subject-experiments}{%
+\section{Human Subject Experiments}\label{human-subject-experiments}}
 
@@
 
-On top of the pre-trained model, we added a 2D global average pooling layer \autocite{lin_network_2014}, followed by a fully-connected dense layer with 256 nodes, a dropout layer \autocite{srivastava_dropout_nodate} with 20\% dropout rate, and a output layer with softmax as activation function.
+On top of the pre-trained model, we added a 2D global average pooling layer \autocite{lin_network_2014}, followed by a fully-connected dense layer with 256 nodes, a dropout layer \autocite{srivastava_dropout_2014} with 20\% dropout rate, and a output layer with softmax as activation function.
 A diagram of the model architecture is given in Figure \ref{fig:model-arch}.

### 2022-03-15 Milestone_presentation_1/Milestone_presentation_1/Milestone_presentation_1.Rmd
add slides

---
title: "Automatic Visual Inference for Linear Regression"
subtitle: ""  
author: 
  - "Weihao (Patrick) Li"
date: '`r format(Sys.Date(), format="%B %d, %Y")`'
output:
  xaringan::moon_reader:
    seal: false
    css: [xaringan-themer.css, mine.css]
    nature:
      beforeInit: "scripts/my_macro.js"
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=5, fig.height=5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r}
library(RefManageR)
BibOptions(check.entries = FALSE,
           bib.style = "authoryear",
           cite.style = "apa",
           style = "Bibtex",
           # max.names = 3,
           longnamesfirst = FALSE,
           hyperlink = FALSE,
           dashed = FALSE)
myBib <- ReadBib("./Visual Inference.bib", check = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
library(tidyverse)
library(visage)
style_solarized_light(
  background_color = "#FFFFFF",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono"),
  text_color = "#000000",
  text_font_size = "1.0rem",
  colors = c(myblue = "#006DAE", mywhite = "#FFFFFF")
)
```

count: false

<!-- need a backaground image -->

.pull-left-full[
<h1 class="myblue"> Automatic Visual Inference for Linear Regression </h1>

<br>
<br>
<br>
<br>
<br>

<h2 class="myblue"> Weihao Li </h2>

<h2 class="myblue"> `r format(Sys.Date(), format="%B %d, %Y")` </h2>
]

.pull-right[
![](images/159.png)
]

---

.pull-left-center[
```{r plot1}
library(visage)

eval_data <- readRDS("data/training_original_data_single.rds")

filter(eval_data, id == 9586, null == TRUE) %>%
  VI_MODEL$plot(theme = theme_light(base_size = 12), size = 1) +
  ylab("Residuals") +
  xlab("Fitted values")
# 68

```

.center[
.caption[
Fig. 1: *Residual plot of a simple linear regression.*
]
]
]

.pull-right[
# Visual Discoveries

<br>
<br>

#### This is a residual plot of a simple linear regression.

#### Can you find the evidence on heteroskedasticity?

- Vertical spread of the points varies with the fitted values.

#### However, the residuals are actually simulated from a **correctly specified model**!

#### Unsecured and unconfirmed visual discoveries will lead to **over or under-interpretations of the data**.

]

---

.pull-left[
# Visual Inference & Lineup Protocol

<br>
<br>

#### **Visual inference** was introduced by `r BibOptions(max.names = 1);Citet(myBib, "buja_statistical_2009")` as an inferential framework to extend confirmatory statistics to visual discoveries.

#### A **lineup** consists of $m$ randomly placed plots, where one plot is the **actual data plot** and the remaining $m-1$ plots (**null plots**) contain data consistent with the null hypothesis.

#### To perform a **visual test**, observer(s) will be asked to select the **most different plot(s)** from the lineup.

]

.pull-right-center[
```{r}
mod <- HETER_MODEL$fit(filter(eval_data, id == 9586, null == FALSE))

manual_lineup <- data.frame()

for (i in 1:19) {
  tmp_dat <- HETER_MODEL$null_resid(filter(eval_data, id == 9586, null == FALSE), mod)
  tmp_dat$k <- i
  manual_lineup <- bind_rows(manual_lineup, tmp_dat)
}

manual_lineup <- bind_rows(manual_lineup, mutate(filter(eval_data, id == 9586, null == TRUE), k = 20))

VI_MODEL$plot_lineup(manual_lineup, 
                     theme = theme_light(),
                     remove_grid_line = TRUE,
                     size = 0.3
                     ) +
  xlab("Fitted values") +
  ylab("Residuals")
```

.center[
.caption[
Fig. 2: _A lineup of 20 residual plots of a simple linear regression. (Plot No.20 is Fig.1) **Can you find the most different one?**_ 
]
]

]

---

# $p$-value

#### Under the null hypothesis, it is expected that the actual data plot would have **no distinguishable difference** with the null plots.

#### Suppose observers are allowed to select **only one plot**. The probability of the observer correctly picks the actual data plot is $1/m$.

#### If we involve $K$ independent observers in a visual test, and let $X$ be a random variable denoting the number of observers correctly picking the actual data plot. Under the null hypothesis $X ∼ Binom(K, 1/m)$. The p-value of a lineup of size $m$ evaluated by $K$ observer is given as $$P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},$$ where $x$ is the realization of number of observers correctly picking the actual data plot.

#### The null hypothesis will be rejected if $P(X \geq x) < \alpha$, the desired significance level.

---

.pull-left-center-80[

![:scale 80%](images/Powerloom_weaving_in_1835.jpg)

![:scale 80%](images/Maquina_vapor_Watt_ETSIIM.jpg)

.caption[
Fig. 3: _A Roberts loom in a weaving shed in 1835 and a Watt steam engine._ 
]

]

.pull-right-120[
# Limitation of Lineup Protocol

<br>

#### 1. Human can not evaluate lineup consisted of a large number of plots.
#### 2. Human can not evaluate a large number of lineups.
#### 3. Lineup protocol is unfriendly to vision-impaired people.
#### 4. Evaluation of lineup is high in labour cost and time consuming.

<br>

### **We need a "steam engine" for visual test!**

]

---

# Computer Vision Model

#### Large-scale evaluation of visual tests is not possible without the use of technology and machines.

#### Modern **computer vision model** could be a promising solution to this problem. It is usually built on a **deep neural network** called **convolutional neural network** (CNN) `r BibOptions(max.names = 1);Citep(myBib, "fukushima_neocognitron_1982")`.

![](images/Typical_cnn.png)

.caption[
Fig. 4: _A typical CNN with convolutional layers, pooling layers and fully-connected layers._ 
]

---

# Computer Vision Model - Convolutional Layer

#### **Convolutional layers** take advantages of the hierarchical pattern in data and provide regularized versions of fully-connected layers. It downscales and transforms the image by summarising information in a small space.

.center[![:scale 70%](images/conlayer.png)]

.center[.caption[
Fig. 5: _Illustration of convolution operation with two filters._ 
]]

---

# Computer Vision Model - Pooling Layer

.pull-left[

<br>
<br>

#### **Pooling layers** apply non-linear down-sampling on the feature maps. This helps reducing the number of parameters in the neural network so that the problem of overfiting can be controlled.

#### One of the most popular pooling operations is **max pooling**. It divides the feature map into several rectangle regions and takes the maximum to form a new feature map.
]

.pull-right-center[
![](images/Max_pooling.png)

.caption[
Fig. 6: _Illustration of max pooling operation._ 
]
]

---

# Computer Vision Model - Fully-connected Layer

.pull-left-center-80[
![:scale 50%](images/fnn.png)

.caption[
Fig. 7: _Illustration of a fully-connected layer._ 
]
]

<br>
<br>

#### **Fully-connected layer** is building block of neural network. It peformes matrix multiplication on the inputs and the weights.

<br>

#### This layer is primarily used to process information presented in the feature maps before making the prediction. The output of the layer will be passed to a special layer called **loss layer** for computing the loss or error of the prediction.

---

# Simulation Setup

<!-- This project focuses on building a prototype of automatic visual statistical inference system for evaluating residual plots of classical normal linear regression model. To set up a comparison between the computer vision models and humans, human subject experiments were conducted to understand the ability of human reading residual plots. -->
<!-- Because the first project is still ongoing, we consider the result we have is not enough to write as a paper at the moment. Hence, the materials provided in this chapter are two main parts of the draft paper that will contain enough details for understanding the project. -->

#### To train computer vision models for evaluating lineups, we need a large amount of data.

#### Meanwhile, we need to set up a comparison between the computer vision models and humans on the ability of reading residual plot for understanding the performance of the models.

---

# Automatic Visual Inference with Computer Vision Models

.pull-left[

#### We proposed an **automatic visual inference system** for evaluating lineups of residual plots, with a focus on the specification error of classical normal linear regression.

#### **Non-linearity** and **heteroskedasticity** are two features considered by the system so far. More features will be added to the system in the future however.

#### The system is built upon **computer vision models**. 
]

<!-- Modern computer vision model could be a promising solution to this problem. As a subfield of AI, computer vision with the modern deep learning architectures solved numerous critical problems in automation. Inspired by the vision processing in living organisms, the convolutional neural network (CNN) was introduced by @fukushima_neocognitron_1982. Soon, this architecture was applied to hand-written number recognition trained with back-propagation by @lecun_backpropagation_1989. This was one of the earliest attempts which human successfully extract information from digital images via self-learning algorithms. Modern computer vision model is typically built on the deep neural network with convolutional layers [@fukushima_neocognitron_1982]. Convolutional layers take advantage of the hierarchical pattern in data and provide regularized versions of fully-connected layers. It downscales and transforms the image by summarising information in a small space. Numerous studies have shown that it can be used to effectively tackle vision tasks, such as image recognition [@rawat_deep_2017]. With the development of graphics processing units and the spread of high-performance personal computers, researches in computer vision become a new hype in the 21st century. Achievements such as computer-aided diagnosis [@lee_image_2015], pedestrian detection [@brunetti_computer_2018] and facial recognition [@emami_facial_2012] had a significant impact on our daily life. -->

<!-- The first project develops a prototype of automatic visual inference system for evaluating lineups of residual plots, with a focus on the classical normal linear regression model given it is one of the simplest predictive models. The automatic system is identical to the lineup protocol except evaluators are replaced by computer vision models. The computer vision models are trained by using data simulated from linear models with violations of different classical assumptions, such as linearity and homoscedasticity. Data of human performance on evaluating residual plots generated under the same simulation setting is collected by conducting online human subject experiments, such that the comparison between the power of human subject-assessed visual tests and the system-assessed visual tests can be made. Moreover, factors that affect the performance of the automatic system are studied for improving the architecture and design of the computer vision models. -->

---

---

### 2022-03-15 Milestone_presentation_1/Milestone_presentation_1/Milestone_presentation_1.Rmd
finish cubic

@@
 
-# Simulation Setup
+# Simulation Setup - Cubic Model
 
-<!-- This project focuses on building a prototype of automatic visual statistical inference system for evaluating residual plots of classical normal linear regression model.
-To set up a comparison between the computer vision models and humans, human subject experiments were conducted to understand the ability of human reading residual plots.
--->
-<!-- Because the first project is still ongoing, we consider the result we have is not enough to write as a paper at the moment.
-Hence, the materials provided in this chapter are two main parts of the draft paper that will contain enough details for understanding the project.
--->
+#### To train computer vision models, we use simulation data generated from two linear regression model with **non-linearity** and **heteroskedasticity** issues.
 
-#### To train computer vision models for evaluating lineups, we need a large amount of data.
+#### The first model is a **cubic model** which can be expressed by: $$\boldsymbol{Y}= 1 + (2-c)\boldsymbol{X} + c\boldsymbol{Z} + a[(2-c)\boldsymbol{X}]^2+a(c\boldsymbol{Z})^2+b[(2-c)\boldsymbol{X}]^3+b(c\boldsymbol{Z})^3+\boldsymbol{\varepsilon},$$ where $c \in (0,2)$, $a \in (-3,3)$, $b \in (-3,3)$, $\boldsymbol{\varepsilon}\overset{iid}{\sim} N(\boldsymbol{0},\sigma^2\boldsymbol{I})$, $\boldsymbol{Y}$, $\boldsymbol{X}$ and $\boldsymbol{Z}$ are $n\times1$ matrices.
 
-#### Meanwhile, we need to set up a comparison between the computer vision models and humans on the ability of reading residual plot for understanding the performance of the models.
+#### In the simulation, $X_i$, $i=1,...,n$, has equal chance to follow one of the following distributions: $N(0, 0.09)$, $U(-1, 1)$, $Lognormal(0,0.36)/3 - 1$ and $-Lognormal(0,0.36)/3 + 1$.
+These distributions are chosen such that most the realizations will fall between $-1$ and $1$.
+
+#### The distribution of $Z_i$, $i=1,...,n$, has $50$% chance to be a uniform distribution ranged from $-1$ to $1$, and $50$% chance to be a discrete uniform distribution with $z_n$ outcomes simulated from a uniform distribution ranged from $-1$ to $1$.
+$z_n$ itself is a discrete uniform distribution with outcomes $\{x: 9<x<21 \text{ and } x \in \mathbb{Z} \}$, which defines the number of possible values $Z_i$ could take.
+
+---
+
+# Simulation Setup - Cubic Model
+
+#### The simulated data is fitted by OLS with equation: $$\boldsymbol{Y}= \beta_0 + \beta_1\boldsymbol{X} + \beta_2\boldsymbol{Z} +\boldsymbol{u},$$ where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2_u\boldsymbol{I})$.
+
+#### Let $\boldsymbol{X}_a=[\boldsymbol{1},\boldsymbol{X},\boldsymbol{Z}]$ denotes the set of regressors in matrix form.
+Then, the **residuals** obtained from the null model are $$\boldsymbol{e} \sim N(\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b, \sigma^2\boldsymbol{R}_a),$$ where $\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$, $\boldsymbol{X}_b=[\boldsymbol{X}^2,\boldsymbol{Z}^2,\boldsymbol{X}^3,\boldsymbol{Z}^3]$ and $\boldsymbol{\beta}_b=(a(2-c)^2,ac^2,b(2-c)^3,bc^3)'$.
+
+#### The **effect size** is derived as: $$ES=n^{-1}||\sigma^{-1}\boldsymbol{R}_a^{-\frac{1}{2}}\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b||^2=n^{-1}\sigma^{-2}||\boldsymbol{R}_a^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2\approx n^{-1}\sigma^{-2}||diag(\boldsymbol{R}_a)^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2,$$ where $diag(\boldsymbol{R}_a)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$.
+
+---
+
+.pull-left[
+```{r}
+mod <- cubic_model(a = 5, b = 10, c = 1)
+mod$plot_lineup(mod$gen_lineup(300), remove_grid_line = TRUE, theme = theme_light()) + xlab("Fitted values") + ylab("Residuals")
+```
+
+.caption[
+Fig. 8: _A lineup of residual plots generated with the cubic model._ 
+]
+]
+
+.pull-right[
+```{r}
+mod <- cubic_model(-1, -1, 1, 0.5)
+dat <- mod$gen(500, fit_model = TRUE)
+mod$plot(dat, theme = theme_light(base_size = 10), size = 1) +
+  ggtitle(glue::glue("Effect size: {round(mod$effect_size(dat), 2)}")) + xlab("Fitted values") + ylab("Residuals") -> p1
+
+mod <- cubic_model(-2, -1, 1, 0.5)
+dat <- mod$gen(500, fit_model = TRUE)
+mod$plot(dat, theme = theme_light(base_size = 10), size = 1) +
+  ggtitle(glue::glue("Effect size: {round(mod$effect_size(dat), 2)}")) + xlab("Fitted values") + ylab("Residuals") -> p2
+
+mod <- cubic_model(-5, -5, 1, 0.5)
+dat <- mod$gen(500, fit_model = TRUE)
+mod$plot(dat, theme = theme_light(base_size = 10), size = 1) +
+  ggtitle(glue::glue("Effect size: {round(mod$effect_size(dat), 2)}")) + xlab("Fitted values") + ylab("Residuals") -> p3
+
+mod <- cubic_model(-20, -20, 1, 0.5)
+dat <- mod$gen(500, fit_model = TRUE)
+mod$plot(dat, theme = theme_light(base_size = 10), size = 1) +
+  ggtitle(glue::glue("Effect size: {round(mod$effect_size(dat), 2)}")) + xlab("Fitted values") + ylab("Residuals") -> p4
+
+list(p1, p2, p3, p4) %>% patchwork::wrap_plots(ncol=2)
+```
+
+.caption[
+Fig. 9: _Residual plots of cubic model under different effect sizes._ 
+]
+]
+
+---
+
+# Simulation Setup - Heteroskedasticity Model
+
+#### The second model is a heteroskedasticy model which can be expressed by: $$Y_i = 1 + X_i + \varepsilon_i, ~i = 1,...,n,$$ where $a \in \{-1,0,1\}$, $b\in (0,32)$ and $\varepsilon_i \overset{iid}{\sim} N(0,1+b(2-|a|)(X_i-a)^2)$.
+
+#### $\boldsymbol{X}$ could be one of the five distributions mentioned above - normal distribution, uniform distribution, adjusted log-normal distribution, adjusted negative log-normal distribution and discrete uniform distribution.
+
+---
+
+# Simulation Setup - Heteroskedasticity Model
+
+#### The simulated data is fitted by OLS with equation: $$\boldsymbol{Y}= \beta_0 + \beta_1\boldsymbol{X} +\boldsymbol{u},$$ where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2_u\boldsymbol{I})$.
+
+#### Let $\boldsymbol{X}_a=[\boldsymbol{1},\boldsymbol{X}]$ denotes the set of regressors in matrix form.
+The residuals obtained from the null model are $$\boldsymbol{e} \sim N(\boldsymbol{0}, \boldsymbol{R}_a\boldsymbol{V}),$$ where $\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ and $\boldsymbol{V}$ is a diagonal matrix with $V_{ii}=1+b(2-|a|)(X_i  -  a)^2$, $i = 1,...,n$.
 

### 2022-03-15 Milestone_presentation_1/Milestone_presentation_1/Milestone_presentation_1.Rmd
finish heter

@@
 
+#### The **effect size** is derived as $ES = b\sqrt{n}$.
+
+---
+
+.pull-left[
+```{r}
+mod <- heter_model(a = -1, b = 30)
+mod$plot_lineup(mod$gen_lineup(300), remove_grid_line = TRUE, theme = theme_light()) + xlab("Fitted values") + ylab("Residuals")
+```
+
+.caption[
+Fig. 10: _A lineup of residual plots generated with the heteroskedasticity model._ 
+]
+]
+
+.pull-right[
+
+```{r}
+set.seed(10086)
+mod <- heter_model(0, 1)
+mod$plot(mod$gen(500, fit_model = TRUE), size = 1, theme = theme_light(base_size = 10)) + ggtitle(glue::glue("Effect size: {round(mod$effect_size(mod$gen(500)),2)}")) + xlab("Fitted values") + ylab("Residuals") -> p1
+mod <- heter_model(0, 3)
+mod$plot(mod$gen(500, fit_model = TRUE), size = 1, theme = theme_light(base_size = 10)) + ggtitle(glue::glue("Effect size: {round(mod$effect_size(mod$gen(500)),2)}")) + xlab("Fitted values") + ylab("Residuals") -> p2
+mod <- heter_model(0, 9)
+mod$plot(mod$gen(500, fit_model = TRUE), size = 1, theme = theme_light(base_size = 10)) + ggtitle(glue::glue("Effect size: {round(mod$effect_size(mod$gen(500)),2)}")) + xlab("Fitted values") + ylab("Residuals") -> p3
+mod <- heter_model(0, 81)
+mod$plot(mod$gen(500, fit_model = TRUE), size = 1, theme = theme_light(base_size = 10)) + ggtitle(glue::glue("Effect size: {round(mod$effect_size(mod$gen(500)),2)}")) + xlab("Fitted values") + ylab("Residuals") -> p4
+
+list(p1, p2, p3, p4) %>% patchwork::wrap_plots(ncol=2)
+```
+
+.caption[
+Fig. 11: _Residual plots of heteroskedasticity model under different effect sizes._ 
+]
+
+]
+
 ---

### 2022-03-16 Milestone_presentation_1/Milestone_presentation_1/Milestone_presentation_1.Rmd
add human subject experiment

@@
 
+<br>
+<br>
+
 #### The second model is a heteroskedasticy model which can be expressed by: $$Y_i = 1 + X_i + \varepsilon_i, ~i = 1,...,n,$$ where $a \in \{-1,0,1\}$, $b\in (0,32)$ and $\varepsilon_i \overset{iid}{\sim} N(0,1+b(2-|a|)(X_i-a)^2)$.
@@
 
+# Human Subject Experiments
+
+.pull-left-center[
+
+<br>
+<br>
+
+![](images/website.png)
+
+.caption[
+Fig. 12: _Layout of the study webiste._ 
+]
+]
+
+.pull-right[
+
+#### **Two human subject experiments** are conducted to compare humans with computer vision models on the ability of reading data plots.
+
+#### **20** lineups are given to each participant, with **10** cubic lineups and **10** heteroskedasticity lineups across different effect sizes.
+**2** out of **20** lineups are extremely easy lineup which are used as attention checks.
+Participants need to pass at least one.
+
+#### The same lineup will not be presented to a participant twice. Most of the lineups have been evaluated by **5** different participants.
+
+#### Participants are allowed to **select multiple plots** from a lineup. They also need to give the **reason** and the **confidence level**.
+
+]
+
+---
+
+# $p$-value and Power Estimation for Multiple Selections
+
+#### The assumptions about the **independence** still needed.
+Let $K$ be the number of independent evaluations of a visual test, $s_i , i = 1, ..., K$ be the **number of selections of the evaluations**, and $X$ be the random variable denoting the number of detections.
+Then, the $p$-value of the visual test is given as: $$P(X\geq x) = \sum_{j=x}^{K}Pr(j|s_1,...,s_K).$$
+
+#### The distribution of $X$ given $s_1, ...s_K$ can not be **derived trivially**, as it is a **sampling without replacement problem**.
+In practice, this distribution can be **approximated** by computer simulation.
+
+---
+
+# $p$-value and Power Estimation for Multiple Selections
+
+#### Assume there is a **visual test** $V_K$, where $K$ denoting the number of evaluations. The corresponding  $p$-value of $V_K$ can be computed.
+
+#### Meanwhile, if one evaluation is **randomly deleted** from $V_K$, the remaining evaluations can still be used to form another valid visual test $V_{K-1}$.
+
+#### In fact, considering **all the possibilities**, $K$ different outcomes for $V_{K-1}$ can be obtained.
+
+#### Since all outcomes occur with **equal probability**, the **proportion of outcomes which reject the null hypothesis** can be used as an **estimator of the power of the visual test** $V_{K-1}$.
+
+#### Similarly, if we would like to estimate the power of the visual test $V_{K-j}$ given the evaluations of $V_{K}$, for $j < K$, we could find all the possible combinations of $K$ elements, taken $k-j$ at a time to obtain $K\choose{K-j}$ different outcomes for $V_{K-j}$.
+Then, the estimated power is given as $R/{K\choose{K-j}}$, where $R$ is the number of outcomes which reject the null hypothesis.
+
+---
+
+# Human Subject Experiment - Results
+
+#### We collect **400** lineup evaluations made by **20** participants in experiment I and **880** lineup evaluations made by **44** participants in experiment II.
+In total, **442** unique lineups are evaluated by 64 subjects.
+
+---
+
+.center[![:scale 80%](images/power1.png)
+
+.caption[
+Fig. 13: _Quasibinomial GLM for estimated power for different number of evaluations._ 
+]]
+
+---
+
 # Automatic Visual Inference with Computer Vision Models

### 2022-03-16 Milestone_presentation_1/Milestone_presentation_1/Milestone_presentation_1.Rmd
add first glm

@@
 
+<br>
+<br>
+
 #### We collect **400** lineup evaluations made by **20** participants in experiment I and **880** lineup evaluations made by **44** participants in experiment II.
 In total, **442** unique lineups are evaluated by 64 subjects.
+
+#### The power of each lineup from one to five evaluations is estimated.
+
+#### The estimated power is further used in fitting a **quasi-binomial generalized linear model** with natural logarithm of effect size as the only regressor.
 
@@
 .caption[
-Fig. 13: _Quasibinomial GLM for estimated power for different number of evaluations._ 
+Fig.
+13: _Quasibinomial GLM for estimated power for different number of evaluations.
+The left plot is for the cubic model, the right plot is for the heteroskedasticity model.
+Power curves for different number of evaluations are drawn in different colors.
+The dashed line is the power of the conventional test._
 ]]

### 2022-03-16 Milestone_presentation_1/Milestone_presentation_1/Milestone_presentation_1.Rmd
fix title page

@@
 
-<br>
+<h3 class="myblue"> First Milestone Presentation </h3>
+
 <br>
@@
 
-<h2 class="myblue"> Weihao Li </h2>
+<h3 class="myblue"> Weihao Li </h3>
 
-<h2 class="myblue"> `r format(Sys.Date(), format="%B %d, %Y")` </h2>
+<h3 class="myblue"> Supervised by Di Cook and Emi Tanaka </h3>
+
+<!-- <h3 class="myblue"> `r format(Sys.Date(), format="%B %d, %Y")` </h3> -->
 ]
@@
 .pull-right[
+
+<br>
+
 ![](images/159.png)
@@
 
-#### The estimated power is further used in fitting a **quasi-binomial generalized linear model** with natural logarithm of effect size as the only regressor.
+#### The estimated power is further used in fitting **quasi-binomial generalized linear models** with natural logarithm of effect size as the only regressor.
+
+#### Another method we use to model the power is to collect all decisions made by $K\choose{K-j}$ different outcomes for $V_{K-j}$, and directly fit **binomial generalized linear models** with natural logarithm of effect size as the regressor.
 
@@
 Fig.
-13: _Quasibinomial GLM for estimated power for different number of evaluations.
+13: _Quasibinomial GLM fitted for estimated power for different number of evaluations.
+The left plot is for the cubic model, the right plot is for the heteroskedasticity model.
+Power curves for different number of evaluations are drawn in different colors.
+The dashed line is the power of the conventional test._
+]]
+
+---
+
+.center[![:scale 80%](images/power2.png)
+
+.caption[
+Fig.
+13: _Binomial GLM fitted for rejection for different number of evaluations.
 The left plot is for the cubic model, the right plot is for the heteroskedasticity model.

### 2022-03-16 Milestone_presentation_1/Milestone_presentation_1/Milestone_presentation_1.Rmd
add inc slide

@@
 .pull-left-center[
-```{r plot1}
+```{r}
 library(visage)
@@
 
-#### This is a residual plot of a simple linear regression.
+#### Can you find the evidence on heteroskedasticity?
+
+]
+
+---
+
+count: false
+
+.pull-left-center[
+```{r}
+filter(eval_data, id == 9586, null == TRUE) %>%
+  VI_MODEL$plot(theme = theme_light(base_size = 12), size = 1) +
+  ylab("Residuals") +
+  xlab("Fitted values")
+# 68
+
+```
+
+.center[
+.caption[
+Fig. 1: *Residual plot of a simple linear regression.*
+]
+]
+]
+
+.pull-right[
+# Visual Discoveries
+
+<br>
+<br>
 
@@
 
-#### However, the residuals are actually simulated from a **correctly specified model**!
+]
+
+---
+
+count: false
+
+.pull-left-center[
+```{r}
+filter(eval_data, id == 9586, null == TRUE) %>%
+  VI_MODEL$plot(theme = theme_light(base_size = 12), size = 1) +
+  ylab("Residuals") +
+  xlab("Fitted values")
+# 68
+
+```
+
+.center[
+.caption[
+Fig. 1: *Residual plot of a simple linear regression.*
+]
+]
+]
+
+.pull-right[
+# Visual Discoveries
+
+<br>
+<br>
+
+#### Can you find the evidence on heteroskedasticity?
+
+- Vertical spread of the points varies with the fitted values.
+
+#### However, the residuals are actually obtained from a **correctly specified classical normal linear model**!
+
+]
+
+---
+
+count: false
+
+.pull-left-center[
+```{r}
+filter(eval_data, id == 9586, null == TRUE) %>%
+  VI_MODEL$plot(theme = theme_light(base_size = 12), size = 1) +
+  ylab("Residuals") +
+  xlab("Fitted values")
+# 68
+
+```
+
+.center[
+.caption[
+Fig. 1: *Residual plot of a simple linear regression.*
+]
+]
+]
+
+.pull-right[
+# Visual Discoveries
+
+<br>
+<br>
+
+#### Can you find the evidence on heteroskedasticity?
+
+- Vertical spread of the points varies with the fitted values.
+
+#### However, the residuals are actually obtained from a **correctly specified classical normal linear model**!
 
@@
 ---
+
+.pull-left[
+# Visual Inference & Lineup Protocol
+
+<br>
+<br>
+
+#### **Visual inference** was introduced by `r BibOptions(max.names = 1);Citet(myBib, "buja_statistical_2009")` as an inferential framework to extend confirmatory statistics to visual discoveries.
+
+]
+
+.pull-right-center[
+```{r}
+mod <- HETER_MODEL$fit(filter(eval_data, id == 9586, null == FALSE))
+
+manual_lineup <- data.frame()
+
+for (i in 1:19) {
+  tmp_dat <- HETER_MODEL$null_resid(filter(eval_data, id == 9586, null == FALSE), mod)
+  tmp_dat$k <- i
+  manual_lineup <- bind_rows(manual_lineup, tmp_dat)
+}
+
+manual_lineup <- bind_rows(manual_lineup, mutate(filter(eval_data, id == 9586, null == TRUE), k = 20))
+
+VI_MODEL$plot_lineup(manual_lineup, 
+                     theme = theme_light(),
+                     remove_grid_line = TRUE,
+                     size = 0.3
+                     ) +
+  xlab("Fitted values") +
+  ylab("Residuals")
+```
+
+.center[
+.caption[
+Fig. 2: _A lineup of 20 residual plots of a simple linear regression. (Plot No.20 is Fig.1) **Can you find the most different one?**_ 
+]
+]
+
+]
+
+---
+
+count: false
+
+.pull-left[
+# Visual Inference & Lineup Protocol
+
+<br>
+<br>
+
+#### **Visual inference** was introduced by `r BibOptions(max.names = 1);Citet(myBib, "buja_statistical_2009")` as an inferential framework to extend confirmatory statistics to visual discoveries.
+
+#### A **lineup** consists of $m$ randomly placed plots, where one plot is the **actual data plot** and the remaining $m-1$ plots (**null plots**) contain data consistent with the null hypothesis.
+
+]
+
+.pull-right-center[
+```{r}
+VI_MODEL$plot_lineup(manual_lineup, 
+                     theme = theme_light(),
+                     remove_grid_line = TRUE,
+                     size = 0.3
+                     ) +
+  xlab("Fitted values") +
+  ylab("Residuals")
+```
+
+.center[
+.caption[
+Fig. 2: _A lineup of 20 residual plots of a simple linear regression. (Plot No.20 is Fig.1) **Can you find the most different one?**_ 
+]
+]
+
+]
+
+---
+
+count: false
 
@@
 ```{r}
-mod <- HETER_MODEL$fit(filter(eval_data, id == 9586, null == FALSE))
-
-manual_lineup <- data.frame()
-
-for (i in 1:19) {
-  tmp_dat <- HETER_MODEL$null_resid(filter(eval_data, id == 9586, null == FALSE), mod)
-  tmp_dat$k <- i
-  manual_lineup <- bind_rows(manual_lineup, tmp_dat)
-}
-
-manual_lineup <- bind_rows(manual_lineup, mutate(filter(eval_data, id == 9586, null == TRUE), k = 20))
-
 VI_MODEL$plot_lineup(manual_lineup, 
@@
 
-#### To train computer vision models, we use simulation data generated from two linear regression model with **non-linearity** and **heteroskedasticity** issues.
+#### We simulated data from two linear regression model with **non-linearity** and **heteroskedasticity** issues
 
@@
 
-#### In the simulation, $X_i$, $i=1,...,n$, has equal chance to follow one of the following distributions: $N(0, 0.09)$, $U(-1, 1)$, $Lognormal(0,0.36)/3 - 1$ and $-Lognormal(0,0.36)/3 + 1$.
-These distributions are chosen such that most the realizations will fall between $-1$ and $1$.
+.pull-left[
+#### $X_i$, $i=1,...,n$ follows one of the four distributions:
+- $N(0, 0.09)$
+- $U(-1, 1)$
+- $Lognormal(0,0.36)/3 - 1$
+- $-Lognormal(0,0.36)/3 + 1$
 
-#### The distribution of $Z_i$, $i=1,...,n$, has $50$% chance to be a uniform distribution ranged from $-1$ to $1$, and $50$% chance to be a discrete uniform distribution with $z_n$ outcomes simulated from a uniform distribution ranged from $-1$ to $1$.
+#### Most the realizations will fall between $-1$ and $1$.
+]
+
+.pull-right[
+#### $Z_i$, $i=1,...,n$ follows one of the two distributions:
+- $U(-1, 1)$
+- Discrete uniform distribution with $z_n$ outcomes ranged from $-1$ to $1$
+
+#### $z_n$ is sampled from $\{z_n: 9<z_n<21 \text{ and } z_n \in \mathbb{Z} \}$.
+]
+
+<!-- #### The distribution of $Z_i$, $i=1,...,n$, has $50$% chance to be a uniform distribution ranged from $-1$ to $1$, and $50$% chance to be a discrete uniform distribution with $z_n$ outcomes simulated from a uniform distribution ranged from $-1$ to $1$.
 $z_n$ itself is a discrete uniform distribution with outcomes $\{x: 9<x<21 \text{ and } x \in \mathbb{Z} \}$, which defines the number of possible values $Z_i$ could take.
+-->
 

### 2022-03-16 Milestone_presentation_1/Milestone_presentation_1/Milestone_presentation_1.Rmd
add more inc, fix first part of slides
finish human subject

@@
       beforeInit: "scripts/my_macro.js"
-      slideNumberFormat: "%current%"
+      countIncrementalSlides: false
+      slideNumberFormat: "%current%/%total%"
       highlightStyle: github
@@
       ratio: 16:9
-      countIncrementalSlides: true
 ---
@@
 
+---
+
+count: false
+
+# $p$-value
+
+#### Under the null hypothesis, it is expected that the actual data plot would have **no distinguishable difference** with the null plots.
+
+#### Suppose observers are allowed to select **only one plot**. The probability of the observer correctly picks the actual data plot is $1/m$.
+
+---
+
+count: false
+
+# $p$-value
+
+#### Under the null hypothesis, it is expected that the actual data plot would have **no distinguishable difference** with the null plots.
+
 #### Suppose observers are allowed to select **only one plot**. The probability of the observer correctly picks the actual data plot is $1/m$.
@@
 Under the null hypothesis $X ∼ Binom(K, 1/m)$.
-The p-value of a lineup of size $m$ evaluated by $K$ observer is given as $$P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},$$ where $x$ is the realization of number of observers correctly picking the actual data plot.
+
+---
+
+count: false
+
+# $p$-value
+
+#### Under the null hypothesis, it is expected that the actual data plot would have **no distinguishable difference** with the null plots.
+
+#### Suppose observers are allowed to select **only one plot**. The probability of the observer correctly picks the actual data plot is $1/m$.
+
+#### If we involve $K$ independent observers in a visual test, and let $X$ be a random variable denoting the number of observers correctly picking the actual data plot.
+Under the null hypothesis $X ∼ Binom(K, 1/m)$.
+
+#### The p-value is given as $$P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i}.$$
+
+---
+
+count: false
+
+# $p$-value
+
+#### Under the null hypothesis, it is expected that the actual data plot would have **no distinguishable difference** with the null plots.
+
+#### Suppose observers are allowed to select **only one plot**. The probability of the observer correctly picks the actual data plot is $1/m$.
+
+#### If we involve $K$ independent observers in a visual test, and let $X$ be a random variable denoting the number of observers correctly picking the actual data plot.
+Under the null hypothesis $X ∼ Binom(K, 1/m)$.
+
+#### The p-value is given as $$P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i}.$$
 
@@
 ---
+
+.pull-left-center-80[
+
+![:scale 80%](images/Powerloom_weaving_in_1835.jpg)
+
+![:scale 80%](images/Maquina_vapor_Watt_ETSIIM.jpg)
+
+.caption[
+Fig. 3: _A Roberts loom in a weaving shed in 1835 and a Watt steam engine._ 
+]
+
+]
+
+.pull-right-120[
+# Limitation of Lineup Protocol
+
+<br>
+
+#### 1. Human can not evaluate lineup consisted of a large number of plots.
+#### 2. Human can not evaluate a large number of lineups.
+#### 3. Lineup protocol is unfriendly to vision-impaired people.
+#### 4. Evaluation of lineup is high in labour cost and time consuming.
+
+<br>
+
+]
+
+---
+
+count: false
 
@@
 
+exclude: true
+
 # Computer Vision Model - Convolutional Layer
@@
 ---
+
+exclude: true
 
@@
 
+--
+
 #### The first model is a **cubic model** which can be expressed by: $$\boldsymbol{Y}= 1 + (2-c)\boldsymbol{X} + c\boldsymbol{Z} + a[(2-c)\boldsymbol{X}]^2+a(c\boldsymbol{Z})^2+b[(2-c)\boldsymbol{X}]^3+b(c\boldsymbol{Z})^3+\boldsymbol{\varepsilon},$$ where $c \in (0,2)$, $a \in (-3,3)$, $b \in (-3,3)$, $\boldsymbol{\varepsilon}\overset{iid}{\sim} N(\boldsymbol{0},\sigma^2\boldsymbol{I})$, $\boldsymbol{Y}$, $\boldsymbol{X}$ and $\boldsymbol{Z}$ are $n\times1$ matrices.
+
+--
 
@@
 ]
+
+--
 
@@
 
+--
+
 #### Let $\boldsymbol{X}_a=[\boldsymbol{1},\boldsymbol{X},\boldsymbol{Z}]$ denotes the set of regressors in matrix form.
@@
 
-#### The **effect size** is derived as: $$ES=n^{-1}||\sigma^{-1}\boldsymbol{R}_a^{-\frac{1}{2}}\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b||^2=n^{-1}\sigma^{-2}||\boldsymbol{R}_a^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2\approx n^{-1}\sigma^{-2}||diag(\boldsymbol{R}_a)^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2,$$ where $diag(\boldsymbol{R}_a)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$.
+--
+
+#### Under the null hypothesis, $\boldsymbol{e} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{R}_a)$.
+The **effect size** is derived as: $$ES=n^{-1}||\sigma^{-1}\boldsymbol{R}_a^{-\frac{1}{2}}\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b||^2=n^{-1}\sigma^{-2}||\boldsymbol{R}_a^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2\approx n^{-1}\sigma^{-2}||diag(\boldsymbol{R}_a)^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2,$$ where $diag(\boldsymbol{R}_a)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$.
 
@@
 .caption[
-Fig. 8: _A lineup of residual plots generated with the cubic model._ 
+Fig. 5: _A lineup of residual plots generated with the cubic model._ 
 ]
@@
 .caption[
-Fig. 9: _Residual plots of cubic model under different effect sizes._ 
+Fig. 6: _Residual plots of cubic model under different effect sizes._ 
 ]
@@
 <br>
-<br>
 
@@
 
-#### $\boldsymbol{X}$ could be one of the five distributions mentioned above - normal distribution, uniform distribution, adjusted log-normal distribution, adjusted negative log-normal distribution and discrete uniform distribution.
+--
+
+#### $X_i$, $i=1,...,n$ follows one of the five distributions:
+- $N(0, 0.09)$
+- $U(-1, 1)$
+- $Lognormal(0,0.36)/3 - 1$
+- $-Lognormal(0,0.36)/3 + 1$
+- Discrete uniform distribution
 
@@
 
+--
+
 #### Let $\boldsymbol{X}_a=[\boldsymbol{1},\boldsymbol{X}]$ denotes the set of regressors in matrix form.
 The residuals obtained from the null model are $$\boldsymbol{e} \sim N(\boldsymbol{0}, \boldsymbol{R}_a\boldsymbol{V}),$$ where $\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ and $\boldsymbol{V}$ is a diagonal matrix with $V_{ii}=1+b(2-|a|)(X_i  -  a)^2$, $i = 1,...,n$.
+
+--
 
@@
 .caption[
-Fig. 10: _A lineup of residual plots generated with the heteroskedasticity model._ 
+Fig. 7: _A lineup of residual plots generated with the heteroskedasticity model._ 
 ]
@@
 .caption[
-Fig. 11: _Residual plots of heteroskedasticity model under different effect sizes._ 
+Fig. 8: _Residual plots of heteroskedasticity model under different effect sizes._ 
 ]
@@
 .caption[
-Fig. 12: _Layout of the study webiste._ 
+Fig. 9: _Layout of the study webiste._ 
+]
+]
+
+.pull-right[
+
+#### **Two human subject experiments** are conducted to compare humans with computer vision models on the ability of reading data plots.
+
+]
+
+---
+
+count: false
+
+# Human Subject Experiments
+
+.pull-left-center[
+
+<br>
+<br>
+
+![](images/website.png)
+
+.caption[
+Fig. 9: _Layout of the study webiste._ 
 ]
@@
 #### **20** lineups are given to each participant, with **10** cubic lineups and **10** heteroskedasticity lineups across different effect sizes.
-**2** out of **20** lineups are extremely easy lineup which are used as attention checks.
-Participants need to pass at least one.
 
-#### The same lineup will not be presented to a participant twice. Most of the lineups have been evaluated by **5** different participants.
+]
+
+---
+
+count: false
+
+# Human Subject Experiments
+
+.pull-left-center[
+
+<br>
+<br>
+
+![](images/website.png)
+
+.caption[
+Fig. 9: _Layout of the study webiste._ 
+]
+]
+
+.pull-right[
+
+#### **Two human subject experiments** are conducted to compare humans with computer vision models on the ability of reading data plots.
+
+#### **20** lineups are given to each participant, with **10** cubic lineups and **10** heteroskedasticity lineups across different effect sizes.
+
+#### **2** out of **20** lineups are extremely easy lineup which are used as attention checks. Participants need to pass at least one.
+
+]
+
+---
+
+count: false
+
+# Human Subject Experiments
+
+.pull-left-center[
+
+<br>
+<br>
+
+![](images/website.png)
+
+.caption[
+Fig. 9: _Layout of the study webiste._ 
+]
+]
+
+.pull-right[
+
+#### **Two human subject experiments** are conducted to compare humans with computer vision models on the ability of reading data plots.
+
+#### **20** lineups are given to each participant, with **10** cubic lineups and **10** heteroskedasticity lineups across different effect sizes.
+
+#### **2** out of **20** lineups are extremely easy lineup which are used as attention checks. Participants need to pass at least one.
+
+#### The same lineup will not be presented to a participant twice. Most of the lineups have **5** replicates.
+
+]
+
+---
+
+count: false
+
+# Human Subject Experiments
+
+.pull-left-center[
+
+<br>
+<br>
+
+![](images/website.png)
+
+.caption[
+Fig. 9: _Layout of the study webiste._ 
+]
+]
+
+.pull-right[
+
+#### **Two human subject experiments** are conducted to compare humans with computer vision models on the ability of reading data plots.
+
+#### **20** lineups are given to each participant, with **10** cubic lineups and **10** heteroskedasticity lineups across different effect sizes.
+
+#### **2** out of **20** lineups are extremely easy lineup which are used as attention checks. Participants need to pass at least one.
+
+#### The same lineup will not be presented to a participant twice. Most of the lineups have **5** replicates.
 
@@
 
-# $p$-value and Power Estimation for Multiple Selections
+# $p$-value Estimation for Multiple Selections
 
-#### The assumptions about the **independence** still needed.
-Let $K$ be the number of independent evaluations of a visual test, $s_i , i = 1, ..., K$ be the **number of selections of the evaluations**, and $X$ be the random variable denoting the number of detections.
+#### Let $K$ be the number of **independent evaluations** of a visual test, $s_i , i = 1, ..., K$ be the **number of selections of the evaluations**, and $X$ be the random variable denoting the number of detections.
 Then, the $p$-value of the visual test is given as: $$P(X\geq x) = \sum_{j=x}^{K}Pr(j|s_1,...,s_K).$$
 
+--
+
 #### The distribution of $X$ given $s_1, ...s_K$ can not be **derived trivially**, as it is a **sampling without replacement problem**.
-In practice, this distribution can be **approximated** by computer simulation.
+
+--
+
+#### In practice, this distribution can be **approximated** by computer simulation.
 
@@
 
-# $p$-value and Power Estimation for Multiple Selections
+# Power Estimation for Multiple Selections
 
-#### Assume there is a **visual test** $V_K$, where $K$ denoting the number of evaluations. The corresponding  $p$-value of $V_K$ can be computed.
+<br>
 
-#### Meanwhile, if one evaluation is **randomly deleted** from $V_K$, the remaining evaluations can still be used to form another valid visual test $V_{K-1}$.
+#### Suppose there is a **visual test** $V_K$, where $K$ denoting the number of evaluations.
 
-#### In fact, considering **all the possibilities**, $K$ different outcomes for $V_{K-1}$ can be obtained.
+--
+
+<br>
+
+#### If one evaluation is **randomly deleted** from $V_K$, the **remaining evaluations** can still be used to form **another valid visual test** $V_{K-1}$.
+
+--
+
+<br>
+
+#### Considering **all the possibilities**, $K$ different outcomes of $V_{K-1}$ can be obtained.
+
+--
+
+<br>
 
@@
 
-#### Similarly, if we would like to estimate the power of the visual test $V_{K-j}$ given the evaluations of $V_{K}$, for $j < K$, we could find all the possible combinations of $K$ elements, taken $k-j$ at a time to obtain $K\choose{K-j}$ different outcomes for $V_{K-j}$.
-Then, the estimated power is given as $R/{K\choose{K-j}}$, where $R$ is the number of outcomes which reject the null hypothesis.
+---
+
+# Power Estimation for Multiple Selections
+
+<br>
+<br>
+
+#### Similarly, to estimate the power of the visual test $V_{K-j}$ given the evaluations of $V_{K}$, for $j < K$,
+
+--
+
+<br>
+
+#### we could find all the possible combinations of $K$ elements, taken $k-j$ at a time to obtain $K\choose{K-j}$ different outcomes for $V_{K-j}$.
+
+--
+
+<br>
+
+#### The estimated power is given as $R/{K\choose{K-j}}$, where $R$ is the number of outcomes which reject the null hypothesis.
 
@@
 #### We collect **400** lineup evaluations made by **20** participants in experiment I and **880** lineup evaluations made by **44** participants in experiment II.
-In total, **442** unique lineups are evaluated by 64 subjects.
+
+--
+
+#### In total, **442** unique lineups are evaluated by **64** subjects.
+
+--
 
@@
 
-#### The estimated power is further used in fitting **quasi-binomial generalized linear models** with natural logarithm of effect size as the only regressor.
+--
 
-#### Another method we use to model the power is to collect all decisions made by $K\choose{K-j}$ different outcomes for $V_{K-j}$, and directly fit **binomial generalized linear models** with natural logarithm of effect size as the regressor.
+#### The estimated power is further used in fitting **quasi-binomial generalized linear models** with **natural logarithm of effect size** as the only regressor.
+
+--
+
+#### All **decisions** made by $K\choose{K-j}$ different outcomes of $V_{K-j}$ are fitted by **binomial generalized linear models** with **natural logarithm of effect size** as the regressor.
 
@@
 Fig.
-13: _Quasibinomial GLM fitted for estimated power for different number of evaluations.
+10: _Quasibinomial GLM fitted for estimated power for different number of evaluations.
 The left plot is for the cubic model, the right plot is for the heteroskedasticity model.
@@
 Fig.
-13: _Binomial GLM fitted for rejection for different number of evaluations.
+11: _Binomial GLM fitted for rejection for different number of evaluations.
 The left plot is for the cubic model, the right plot is for the heteroskedasticity model.
@@
 The dashed line is the power of the conventional test._
+]]
+
+---
+
+# Human Subject Experiment - Results
+
+#### There is a huge gap between the conventional curves and the visual test curves at the middle of the plot.
+
+#### Human is hard to identify the pattern at this level of difficulty.
+
+.pull-left[
+![](images/mid1.png)
+
+]
+
+.pull-right[
+![](images/mid2.png)
+
+]
+
+.center[.caption[
+Fig. 12: _A cubic lineup with effect size $= 0.18$ and a heteroskedasticity lineup with effect size $= 17.5$._ 
 ]]
@@
 
-#### We proposed an **automatic visual inference system** for evaluating lineups of residual plots, with a focus on the specification error of classical normal linear regression.
+#### We proposed an **automatic visual inference system** for evaluating lineups of residual plots of classical normal linear regression.
 
@@
 
-#### The system is built upon **computer vision models**. 
 ]

### 2022-03-16 Milestone_presentation_1/Milestone_presentation_1/Milestone_presentation_1.Rmd
frame slides

@@
 
-#### **2** out of **20** lineups are extremely easy lineup which are used as attention checks. Participants need to pass at least one.
+#### **2** out of **20** lineups are extremely easy that are used as attention checks. Participants need to pass at least one to be paid.
 
@@
 
-#### There is a huge gap between the conventional curves and the visual test curves at the middle of the plot.
+#### There is a huge gap between the conventional curves and the visual test curves when the natural logarithm of effect size is around $0$ for cubic model, and $2.5$ for heteroskedasticity model.
 
@@
 
-# Automatic Visual Inference with Computer Vision Models
-
-.pull-left[
-
-#### We proposed an **automatic visual inference system** for evaluating lineups of residual plots of classical normal linear regression.
-
-#### **Non-linearity** and **heteroskedasticity** are two features considered by the system so far. More features will be added to the system in the future however.
-
-]
-
-<!-- Modern computer vision model could be a promising solution to this problem.
-As a subfield of AI, computer vision with the modern deep learning architectures solved numerous critical problems in automation.
-Inspired by the vision processing in living organisms, the convolutional neural network (CNN) was introduced by @fukushima_neocognitron_1982.
-Soon, this architecture was applied to hand-written number recognition trained with back-propagation by @lecun_backpropagation_1989.
-This was one of the earliest attempts which human successfully extract information from digital images via self-learning algorithms.
-Modern computer vision model is typically built on the deep neural network with convolutional layers [@fukushima_neocognitron_1982].
-Convolutional layers take advantage of the hierarchical pattern in data and provide regularized versions of fully-connected layers.
-It downscales and transforms the image by summarising information in a small space.
-Numerous studies have shown that it can be used to effectively tackle vision tasks, such as image recognition [@rawat_deep_2017].
-With the development of graphics processing units and the spread of high-performance personal computers, researches in computer vision become a new hype in the 21st century.
-Achievements such as computer-aided diagnosis [@lee_image_2015], pedestrian detection [@brunetti_computer_2018] and facial recognition [@emami_facial_2012] had a significant impact on our daily life.
--->
-
-<!-- The first project develops a prototype of automatic visual inference system for evaluating lineups of residual plots, with a focus on the classical normal linear regression model given it is one of the simplest predictive models.
-The automatic system is identical to the lineup protocol except evaluators are replaced by computer vision models.
-The computer vision models are trained by using data simulated from linear models with violations of different classical assumptions, such as linearity and homoscedasticity.
-Data of human performance on evaluating residual plots generated under the same simulation setting is collected by conducting online human subject experiments, such that the comparison between the power of human subject-assessed visual tests and the system-assessed visual tests can be made.
-Moreover, factors that affect the performance of the automatic system are studied for improving the architecture and design of the computer vision models.
--->
-
----
-
----
+# Automatic Visual Inference
+
+#### We propose a prototype **automatic visual inference system** for evaluating lineups of residual plots of classical normal linear regression.
+It currently only detects nonlinearity and heteroskedasticity.
+
+.center[
+![:scale 80%](images/vi_system.png)
+
+.caption[
+Fig. 13: _Diagram of automatic visual inference system._ 
+]]
+
+---
+
+# Model Architecture and Model Training
+
+.pull-left-140[
+
+#### Due to limited computing resources, we apply **transfer learning** to train five computer vision models.
+
+#### Five **pre-trained models** are used, including **VGG16**, **VGG19**, **Xception**, **ResNet152V2** and **InceptionV3**.
+
+#### Models are trained on **10000 null plots** and **10000 actual data plots** simulated from the cubic model and heteroskedasticity model. 20% training data is used as validation set.
+
+#### **Image augmentation**:
+- **random rotation angle** ranged from $0$ to $5$ degrees
+- **random brightness** ranged from $95$% to $105$%
+
+#### Model training is performed with **Tensorflow** and **Keras**.
+
+]
+
+.pull-right-60[
+.center[
+
+![:scale 23%](images/vi_model.png)
+
+.caption[
+Fig. 14: _Diagram of the transfer learning computer vision model._ 
+]
+]
+]
+
+---
+
+![](images/mod_per.png)
+
+.center[.caption[
+Fig. 15: _Model performance tested on 10000 null plots and 10000 actual data plots._ 
+]]
+
+---
+
+# Performance
+
+---
+
+# Conclusion
+
+---
+
+# Future work
+
+---
+
+Thanks

### 2022-03-16 Milestone_presentation_1/Milestone_presentation_1/Milestone_presentation_1.Rmd
lack conclusion and future

@@
 
-# $p$-value
+# $p$-value of Lineup Protocol
 
@@
 10: _Quasibinomial GLM fitted for estimated power for different number of evaluations.
-The left plot is for the cubic model, the right plot is for the heteroskedasticity model.
+The plot on the left is for the cubic model, the plot on the right is for the heteroskedasticity model.
 Power curves for different number of evaluations are drawn in different colors.
@@
 11: _Binomial GLM fitted for rejection for different number of evaluations.
-The left plot is for the cubic model, the right plot is for the heteroskedasticity model.
+The plot on the left is for the cubic model, the plot on the right is for the heteroskedasticity model.
 Power curves for different number of evaluations are drawn in different colors.
@@
 
-# Performance
+.center[![:scale 80%](images/system_pow_1.png)
+
+.caption[
+Fig.
+16: _Power of the automatic visual test compared to human visual test.
+The blue curve is the power of the automatic visual test.
+The dashed line represents the power of the conventional test.
+The plot on the left is for cubic model, and the plot on the right is for heteroskedasticity._
+]]
+
+---
+
+.center[![:scale 80%](images/system_pow_2.png)
+
+.caption[
+Fig.
+17: _Power of the automatic visual test with different number of evaluations.
+The dashed line represents the power of the conventional test.
+The plot on the left is for cubic model, and the plot on the right is for heteroskedasticity._
+]]
 

### 2022-03-16 Milestone_presentation_1/Milestone_presentation_1/Milestone_presentation_1.Rmd
lack citation

@@
     nature:
+      countdown: 55000
       beforeInit: "scripts/my_macro.js"
@@
 .caption[
-Fig. 3: _A Roberts loom in a weaving shed in 1835 and a Watt steam engine._ 
+Fig. 3: _A Roberts loom in a weaving shed in 1835 and a Watt steam engine. (https://en.wikipedia.org/wiki/Industrial_Revolution)_ 
 ]
@@
 .caption[
-Fig. 4: _A typical CNN with convolutional layers, pooling layers and fully-connected layers._ 
+Fig. 4: _A typical CNN with convolutional layers, pooling layers and fully-connected layers. (https://en.wikipedia.org/wiki/Convolutional_neural_network)_ 
 ]
@@
 
-#### Let $K$ be the number of **independent evaluations** of a visual test, $s_i , i = 1, ..., K$ be the **number of selections of the evaluations**, and $X$ be the random variable denoting the number of detections.
-Then, the $p$-value of the visual test is given as: $$P(X\geq x) = \sum_{j=x}^{K}Pr(j|s_1,...,s_K).$$
+#### Let $s_i , i = 1, ..., K$ be the **number of plots selected by each observer**. Then, the $p$-value of the visual test is given as: $$P(X\geq x) = \sum_{j=x}^{K}Pr(j|s_1,...,s_K).$$
 
@@
 
-<br>
-<br>
-
-#### We collect **400** lineup evaluations made by **20** participants in experiment I and **880** lineup evaluations made by **44** participants in experiment II.
-
---
-
-#### In total, **442** unique lineups are evaluated by **64** subjects.
-
---
-
-#### The power of each lineup from one to five evaluations is estimated.
+.pull-left[
+#### Experiment I:
+- **400** lineup evaluations
+- **20** participants
+]
+
+.pull-right[
+#### Experiment II:
+- **880** lineup evaluations
+- **44** participants
+]
+
+<br>
 
@@
 
-#### All **decisions** made by $K\choose{K-j}$ different outcomes of $V_{K-j}$ are fitted by **binomial generalized linear models** with **natural logarithm of effect size** as the regressor.
+#### **Decisions** made by $K\choose{K-j}$ different outcomes of $V_{K-j}, j = 1, ..., 4$, are fitted by **binomial generalized linear models**.
 
@@
 Power curves for different number of evaluations are drawn in different colors.
-The dashed line is the power of the conventional test._
+The dashed line is the power of the F-test and BP-test._
 ]]
@@
 
-#### There is a huge gap between the conventional curves and the visual test curves when the natural logarithm of effect size is around $0$ for cubic model, and $2.5$ for heteroskedasticity model.
+#### There is a huge gap between the conventional test curves and the visual test curves when the natural logarithm of effect size is around $0$ for cubic model, and $2.5$ for heteroskedasticity model.
 
@@
 #### We propose a prototype **automatic visual inference system** for evaluating lineups of residual plots of classical normal linear regression.
-It currently only detects nonlinearity and heteroskedasticity.
+It currently only detects non-linearity and heteroskedasticity.
 
@@
 
+]
+
+.pull-right-60[
+.center[
+
+![:scale 23%](images/vi_model.png)
+
+.caption[
+Fig. 14: _Diagram of the transfer learning computer vision model._ 
+]
+]
+]
+
+---
+
+count: false
+
+# Model Architecture and Model Training
+
+.pull-left-140[
+
+#### Due to limited computing resources, we apply **transfer learning** to train five computer vision models.
+
 #### Five **pre-trained models** are used, including **VGG16**, **VGG19**, **Xception**, **ResNet152V2** and **InceptionV3**.
 
-#### Models are trained on **10000 null plots** and **10000 actual data plots** simulated from the cubic model and heteroskedasticity model. 20% training data is used as validation set.
+]
+
+.pull-right-60[
+.center[
+
+![:scale 23%](images/vi_model.png)
+
+.caption[
+Fig. 14: _Diagram of the transfer learning computer vision model._ 
+]
+]
+]
+
+---
+
+count: false
+
+# Model Architecture and Model Training
+
+.pull-left-140[
+
+#### Due to limited computing resources, we apply **transfer learning** to train five computer vision models.
+
+#### Five **pre-trained models** are used, including **VGG16**, **VGG19**, **Xception**, **ResNet152V2** and **InceptionV3**.
+
+#### Training: 
+- $10000$ null plots and $10000$ actual data plots simulated from the cubic model and the heteroskedasticity model
+- $20$% are used as validation set
+
+]
+
+.pull-right-60[
+.center[
+
+![:scale 23%](images/vi_model.png)
+
+.caption[
+Fig. 14: _Diagram of the transfer learning computer vision model._ 
+]
+]
+]
+
+---
+
+count: false
+
+# Model Architecture and Model Training
+
+.pull-left-140[
+
+#### Due to limited computing resources, we apply **transfer learning** to train five computer vision models.
+
+#### Five **pre-trained models** are used, including **VGG16**, **VGG19**, **Xception**, **ResNet152V2** and **InceptionV3**.
+
+#### Training: 
+- $10000$ null plots and $10000$ actual data plots simulated from the cubic model and the heteroskedasticity model
+- $20$% are used as validation set
 
@@
 
-#### Model training is performed with **Tensorflow** and **Keras**.
+]
+
+.pull-right-60[
+.center[
+
+![:scale 23%](images/vi_model.png)
+
+.caption[
+Fig. 14: _Diagram of the transfer learning computer vision model._ 
+]
+]
+]
+
+---
+
+count: false
+
+# Model Architecture and Model Training
+
+.pull-left-140[
+
+#### Due to limited computing resources, we apply **transfer learning** to train five computer vision models.
+
+#### Five **pre-trained models** are used, including **VGG16**, **VGG19**, **Xception**, **ResNet152V2** and **InceptionV3**.
+
+#### Training: 
+- $10000$ null plots and $10000$ actual data plots simulated from the cubic model and the heteroskedasticity model
+- $20$% are used as validation set
+
+#### **Image augmentation**:
+- **random rotation angle** ranged from $0$ to $5$ degrees
+- **random brightness** ranged from $95$% to $105$%
+
+#### Model is trained with **Tensorflow** and **Keras**.
 
@@
 The blue curve is the power of the automatic visual test.
-The dashed line represents the power of the conventional test.
-The plot on the left is for cubic model, and the plot on the right is for heteroskedasticity._
+The dashed line is the power of the F-test and BP-test.
+The plot on the left is for cubic model, and the plot on the right is for heteroskedasticity model._
 ]]
@@
 17: _Power of the automatic visual test with different number of evaluations.
-The dashed line represents the power of the conventional test.
-The plot on the left is for cubic model, and the plot on the right is for heteroskedasticity._
+The dashed line is the power of the F-test and BP-test.
+The plot on the left is for cubic model, and the plot on the right is for heteroskedasticity model._
 ]]
@@
 
+<br>
+<br>
+
+#### 1. Visual test is not better than conventional tests in detecting non-linearity and heteroskedasticity from residual plot of classical linear regression model.
+
+<br>
+
+--
+
+#### 2. The idea of automating visual inference with computer vision models is feasible. Improvement of the automatic system still needs to be made to match the human visual test power, however.
+
+<br>
+
+--
+
+#### 3. Increasing the number of computer vision models in an automatic visual inference system will result in greater statistical power. But the amount of improvement will decrease.
+
 ---
@@
 
----
-
-Thanks
+.pull-left[
+### General-Purpose Automatic Visual Statistical Inference System
+
+#### Adopt the techniques developed in query by example content-based image retrieval to build automatic general-purpose visual inference system based on image comparison.
+
+#### No assumption about the type of the plot and the data generating process will be made.
+
+]
+
+.pull-right[
+### OAVIS: An online automatic visual inference system
+
+#### Deploy the automatic visual inference as an online aplication
+
+#### Use online machien learning to improve system performance dynamically.
+]
+
+---
+
+class: center, middle
+
+# Thanks!
+
+---
+
+# References
+
+Buja, A. et al. (2009). “Statistical inference for
+exploratory data analysis and model diagnostics”. In:
+_Philosophical Transactions of the Royal Society A:
+Mathematical, Physical and Engineering Sciences_
+367.1906, pp. 4361-4383. DOI:
+[10.1098/rsta.2009.0120](https://doi.org/10.1098%2Frsta.2009.0120).
+URL:
+[https://royalsocietypublishing-org.ezproxy.lib.monash.edu.au/doi/full/10.1098/rsta.2009.0120](https://royalsocietypublishing-org.ezproxy.lib.monash.edu.au/doi/full/10.1098/rsta.2009.0120).
+
+Fukushima, K. et al. (1982). “Neocognitron: A
+self-organizing neural network model for a mechanism of
+visual pattern recognition”. In: _Competition and
+cooperation in neural nets_. Springer, pp. 267-285.
+
+<!-- `r BibOptions(style = "markdown");PrintBibliography(myBib)` -->

### 2022-04-20 paper_comparsion/paper_comparision/paper_comparision.Rmd
update paper comparison

---
title: |
  Taylor & Francis Rmarkdown template for authors (\LaTeX\-based 
  \textsf{Interact} layout + Chicago author-date reference style)
type: ARTICLE TEMPLATE
author:
  - name: Weihao Li
    affil: a
    email: weihao.li@monash.edu
  - name: John Smith
    affil: b
    email: john.smith@uni-freiburg.de
  - name: Dominik Leutnant
    affil: c, \dagger, \ddagger
    email: leutnant@fh-muenster.de
affiliation:
  - num: a
    address: |
      Taylor & Francis, 4 Park Square, Milton Park, Abingdon, UK
  - num: b
    address: |
      Institut für Informatik, Albert-Ludwigs-Universität, Freiburg, Germany
  - num: c
    address: |
      Muenster University of Applied Sciences - 
      Institute for Infrastructure, Water, Resources, Environment,
      Correnstr. 25, 48149 Münster, Germany
bibliography: interactcadsample.bib
abstract: |
  This template is for authors who are preparing a manuscript for a Taylor & 
  Francis journal using the \LaTeX\ document preparation system and the 
  `interact} class file, which is available via selected journals' home 
  pages on the Taylor & Francis website.
keywords: |
  Sections; lists; figures; tables; mathematics; fonts; references; appendices
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
output: rticles::tf_article
---

problem: residual plot diagnostics
conventional test: too sensitive

background:

<!-- week 1 -->
1. residual plot for model diagnostics

a. residual is widely used
b. what are the types of residual plots
c. comparison

<!-- week 2 -->
2. conventional test: F, BP
3. visual test: lineup, theory

<!-- week 4 -->
desc of experiment:
1. simulation setup
2. experimental design
3. result

<!-- week 3 -->
comparison of conventional tests:
1. power (visual test vs. conventional test)
(visual test most different one (everything test, any departure))
2. investigate the difference (gap), give examples
3. conventional is too sensitive
4. make conventional less sensitive (vary alpha)

<!-- last week -->
conclusion:
1. too sensitive, visual test is needed/preferable
2. visual test is infeasible in large scale (expensive)
3. future work (role of computer vision)

Title:

# Introduction

Diagnostics of the classical normal linear regression model conventionally involve evaluating the fitness of the proposed model, detecting the presence of influential observations and outliers, checking the validity of model assumptions and many more. Tools such as summary statistics, hypothesis testing, and data plots are essential for a systematic and detailed examination of the regression model [@mansfield1987diagnostic].

Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are refereed to as standard residual plots. They are commonly used to identify patterns which are indicative of nonconstant error variance or nonlinearity [@cook1982residuals]. Raw residuals and studentized residuals are the two most frequently used residuals in standard residual plots. The debt on which type of residuals should be used always present. While raw residuals are the most common output of computer regression software package, by applying a scaling factor, the ability of revealing nonconstant error variance in standard residual plots will often be enhanced by studentized residuals in small sample size [@gunst2018regression].

As a two-dimensional representation of a model in a $p$-dimensional space, standard residual plots project  data points onto the variable of the horizontal axis, which is a vector in $p$-dimensional space. Observations with the same projection will be treated as equivalent as they have the same position of the abscissa. Therefore, standard residual plots are often useful in revealing model inadequacies in the direction of the variable of the horizontal axis, but could be inadequate for detecting patterns in other directions, especially in those perpendicular to the variable of the horizontal axis. Hence, in practice, multiple standard residual plots with different horizontal axes will be examined.

<!-- large sample size-->
Overlapping data points is a general issue in scatter plots not limited to standard residual plots, which often makes plots difficult to interpret because visual patterns are concealed. Thus, for relatively large sample size, @cleveland1975graphical suggests the use of robust reference lines to give aids to eye in seeing patterns, which nowadays, are usually replaced with a spline or local polynomial regression line.

One of the useful supplements to standard residual plots is partial residual plot. While the standard residual plots indicate deviations of each predictor from linearity, the partial residual plots can be used to access the extent and the direction of the linearity. These plots can therefore be used to determine the importance of each predictor in the presence of all the others and assess the extent of nonlinearity in a given predictor. This is in addition to providing information on correct transformations and extreme data points.

### 2022-04-20 paper_comparsion/paper_comparision/paper_comparision.tex
update paper comparison

% interactcadsample.tex
% v1.03 - April 2017

\documentclass[]{interact}

\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage{subfigure}% Support for small, `sub' figures and tables
%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required

\usepackage{natbib}% Citation support using natbib.sty
\bibpunct[, ]{(}{)}{;}{a}{}{,}% Citation support using natbib.sty
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty

\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}

% see https://stackoverflow.com/a/47122900

% Pandoc citation processing

\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\def\tightlist{}

\begin{document}

\articletype{ARTICLE TEMPLATE}

\title{Taylor \& Francis Rmarkdown template for authors (\LaTeX-based
\textsf{Interact} layout + Chicago author-date reference style)}

\author{\name{Weihao Li$^{a}$, John Smith$^{b}$, Dominik Leutnant$^{c,
\dagger, \ddagger}$}
\affil{$^{a}$Taylor \& Francis, 4 Park Square, Milton Park, Abingdon,
UK; $^{b}$Institut für Informatik, Albert-Ludwigs-Universität, Freiburg,
Germany; $^{c}$Muenster University of Applied Sciences - Institute for
Infrastructure, Water, Resources, Environment, Correnstr. 25, 48149
Münster, Germany}
}

\thanks{CONTACT Weihao
Li. Email: \href{mailto:weihao.li@monash.edu}{\nolinkurl{weihao.li@monash.edu}}, John
Smith. Email: \href{mailto:john.smith@uni-freiburg.de}{\nolinkurl{john.smith@uni-freiburg.de}}, Dominik
Leutnant. Email: \href{mailto:leutnant@fh-muenster.de}{\nolinkurl{leutnant@fh-muenster.de}}}

\maketitle

\begin{abstract}
This template is for authors who are preparing a manuscript for a Taylor
\& Francis journal using the \LaTeX~document preparation system and the
`interact\} class file, which is available via selected journals' home
pages on the Taylor \& Francis website.
\end{abstract}

\begin{keywords}
Sections; lists; figures; tables; mathematics; fonts; references;
appendices
\end{keywords}

problem: residual plot diagnostics conventional test: too sensitive

background:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  residual plot for model diagnostics
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\alph{enumi}.}
\tightlist
\item
  residual is widely used
\item
  what are the types of residual plots
\item
  comparison
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  conventional test: F, BP
\item
  visual test: lineup, theory
\end{enumerate}

desc of experiment: 1. simulation setup 2. experimental design 3. result

comparison of conventional tests: 1. power (visual test vs.~conventional
test) (visual test most different one (everything test, any departure))
2. investigate the difference (gap), give examples 3. conventional is
too sensitive 4. make conventional less sensitive (vary alpha)

conclusion: 1. too sensitive, visual test is needed/preferable 2. visual
test is infeasible in large scale (expensive) 3. future work (role of
computer vision)

Title:

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Diagnostics of the classical normal linear regression model
conventionally involve evaluating the fitness of the proposed model,
detecting the presence of influential observations and outliers,
checking the validity of model assumptions and many more. Tools such as
summary statistics, hypothesis testing, and data plots are essential for
a systematic and detailed examination of the regression model
\citep{mansfield1987diagnostic}.

Graphical summaries in which residuals are plotted against fitted values
or other functions of the predictor variables that are approximately
orthogonal to residuals are refereed to as standard residual plots. They
are commonly used to identify patterns which are indicative of
nonconstant error variance or nonlinearity \citep{cook1982residuals}.
Raw residuals and studentized residuals are the two most frequently used
residuals in standard residual plots. The debt on which type of
residuals should be used always present. While raw residuals are the
most common output of computer regression software package, by applying
a scaling factor, the ability of revealing nonconstant error variance in
standard residual plots will often be enhanced by studentized residuals
in small sample size \citep{gunst2018regression}.

As a two-dimensional representation of a model in a \(p\)-dimensional
space, standard residual plots project data points onto the variable of
the horizontal axis, which is a vector in \(p\)-dimensional space.
Observations with the same projection will be treated as equivalent as
they have the same position of the abscissa. Therefore, standard
residual plots are often useful in revealing model inadequacies in the
direction of the variable of the horizontal axis, but could be
inadequate for detecting patterns in other directions, especially in
those perpendicular to the variable of the horizontal axis. Hence, in
practice, multiple standard residual plots with different horizontal
axes will be examined.

Overlapping data points is a general issue in scatter plots not limited
to standard residual plots, which often makes plots difficult to
interpret because visual patterns are concealed. Thus, for relatively
large sample size, \citet{cleveland1975graphical} suggests the use of
robust reference lines to give aids to eye in seeing patterns, which
nowadays, are usually replaced with a spline or local polynomial
regression line.

One of the useful supplements to standard residual plots is partial
residual plot. While the standard residual plots indicate deviations of
each predictor from linearity, the partial residual plots can be used to
access the extent and the direction of the linearity. These plots can
therefore be used to determine the importance of each predictor in the
presence of all the others and assess the extent of nonlinearity in a
given predictor. This is in addition to providing information on correct
transformations and extreme data points.

\bibliographystyle{tfcad}
\bibliography{interactcadsample.bib}

\end{document}

### 2022-04-20 paper_comparsion/scripts/conventional.R

@@ -1,0 +1,109 @@
+library(visage)
+
+conv_test_cubic <- function() {
+  para_list1 <- list(a = runif(1, -3, 3),
+                     b = runif(1, -3, 3),
+                     c = runif(1, 0, 2),
+                     x_dist = sample(c("uniform", "normal", "lognormal", "neglognormal"), 1),
+                     x_mu = 0,
+                     x_n = NA,
+                     x_sigma = sample(c(0.3, 0.6), 1),
+                     z_discrete = sample(c(TRUE, FALSE), 1),
+                     z_discrete_dist = "discrete_uniform",
+                     z_n = sample(3:10, 1),
+                     e_dist = "normal",
+                     e_df = NA,
+                     e_sigma = sample(c(0.25, 0.5, 1, 2), 1),
+                     n = sample(c(50, 100, 300), 1))
+
+  if (para_list1$x_dist == "normal") {para_list1$x_sigma <- 0.3} else {para_list1$x_sigma <- 0.6}
+
+  # Init X and Z based on the parameter list
+  x <- switch (para_list1$x_dist ,
+               "uniform" = rand_uniform(-1, 1),
+               "normal" = rand_normal(0, para_list1$x_sigma),
+               "lognormal" = rand_lognormal(0, para_list1$x_sigma),
+               "neglognormal" = {rl <- rand_lognormal(0, para_list1$x_sigma); closed_form(~-rl)}
+  )
+
+  if (para_list1$z_discrete) {z <- rand_uniform(-1, 1)} else {z <- rand_uniform_d(-1, 1, para_list1$z_n)}
+
+
+  # Build the cubic model
+  mod <- cubic_model(a = para_list1$a, b = para_list1$b, c = para_list1$c,
+                     sigma = para_list1$e_sigma,
+                     x = x, z = z)
+
+  # Generate data from the model, let k = 2 (50% true data plot, 50% null plot)
+  dat <- mod$gen(para_list1$n, fit_model = TRUE, test = TRUE)
+
+  c(mod$effect_size(dat), dat$p_value[1])
+}
+
+conv_test_heter <- function() {
+  para_list1 <- list(a = sample(c(-1, 0, 1), 1),
+                     b = runif(1, 0, 32),
+                     c = NA,
+                     x_dist = sample(c("uniform", "normal", "lognormal", "neglognormal", "discrete_uniform"), 1),
+                     x_mu = 0,
[…]

### 2022-04-20 paper_comparsion/scripts/main.R

@@ -1,0 +1,167 @@
+library(tidyverse)
+library(visage)
+
+source(here::here("scripts/survey.R"))
+source(here::here("scripts/conventional.R"))
+
+survey_dat <- read_survey()
+conv_result <- get_sim_conv()
+
+env <- new.env()
+
+comb_pvalue <- survey_dat %>%
+  mutate(lineup_id = paste0(exp, "_", lineup_id)) %>%
+  visage::calc_p_value_multi(lineup_id = "lineup_id",
+                             detected = "detect",
+                             n_sel = "num_selection",
+                             comb = TRUE,
+                             n_eval = 1:5,
+                             n_sim = 100000,
+                             cache_env = env)
+
+comb_power <- comb_pvalue %>%
+  rowwise() %>%
+  mutate(hat_power_001 = mean(eval_p_value(p_value, tol = 1e-2, significance_level = 0.01))) %>%
+  mutate(hat_power_005 = mean(eval_p_value(p_value, tol = 1e-2, significance_level = 0.05))) %>%
+  mutate(hat_power_010 = mean(eval_p_value(p_value, tol = 1e-2, significance_level = 0.1))) %>%
+  mutate(hat_power_015 = mean(eval_p_value(p_value, tol = 1e-2, significance_level = 0.15))) %>%
+  mutate(hat_power_020 = mean(eval_p_value(p_value, tol = 1e-2, significance_level = 0.2))) %>%
+  mutate(hat_power_025 = mean(eval_p_value(p_value, tol = 1e-2, significance_level = 0.25))) %>%
+  mutate(hat_power_030 = mean(eval_p_value(p_value, tol = 1e-2, significance_level = 0.3))) %>%
+  ungroup() %>%
+  select(-p_value) %>%
+  gather(hat_power_001:hat_power_030, key = "alpha", value = "hat_power")
+
+survey_dat %>%
+  mutate(lineup_id = paste0(exp, "_", lineup_id)) %>%
+  select(lineup_id, type, effect_size) %>%
+  group_by(lineup_id) %>%
+  summarise(type = first(type), effect_size = first(effect_size)) %>%
+  right_join(comb_power) %>%
+  ggplot() +
+  geom_smooth(aes(log(effect_size), hat_power, col = factor(n_eval), group = factor(n_eval)),
+              method = "glm",
+              method.args = list(family = quasibinomial()),
[…]

### 2022-04-20 paper_comparsion/scripts/pilot.R

@@ -1,0 +1,50 @@
+# pilot study data
+
+# effect size might be unreliable
+
+read_pilot <- function() {
+  file_names <- list.files(here::here("data/pilot_study"))
+
+  library(tidyverse)
+
+  result_high <- NULL
+
+  for (file_name in file_names[grep("oct" , file_names)]) {
+    dat <- read_csv(here::here("data/pilot_study", file_name))
+    if (is.null(result_high)) {
+      result_high <- dat
+    } else {
+      result_high <- bind_rows(result_high, dat)
+    }
+  }
+
+  result_heter <- NULL
+
+  for (file_name in file_names[grep("nov" , file_names)]) {
+    dat <- read_csv(here::here("data/pilot_study", file_name))
+    if (is.null(result_heter)) {
+      result_heter <- dat
+    } else {
+      result_heter <- bind_rows(result_heter, dat)
+    }
+  }
+
+  bind_rows(result_high, result_heter)
+}
+
+pilot_model <- function(pilot_dat) {
+
+  pilot_dat <- pilot_dat %>%
+    mutate(detect = as.numeric(answer),
+           log_effect_size = log(effect_size))
+
+  mod1 <- glm(detect ~ log(effect_size),
+              data = filter(pilot_dat, model == "higher order"),
+              family = binomial)
+
+  mod2 <- glm(detect ~ log(effect_size),
+              data = filter(pilot_dat, model != "higher order"),
+              family = binomial)
+
+  return(list(mod1, mod2))
+}

### 2022-04-20 paper_comparsion/scripts/survey.R

@@ -1,0 +1,175 @@
+process_responses <- function(k, lineup_dat, lineup_ord, survey_folder) {
+
+  dat <- tibble(id = 1:length(unlist(lineup_ord)),
+                lineup_dat_id = unlist(lineup_ord)) %>%
+    mutate(set = (id-1) %/% k + 1) %>%
+    mutate(num = (id-1) %% k + 1)
+
+  user_info_dict <- list(c("18-24", "25-39", "40-54", "55-64", "65 or above"),
+                         c("High School or below", "Diploma and Bachelor Degree", "Honours Degree", "Masters Degree", "Doctoral Degree"),
+                         c("He", "She", "They", "Other"),
+                         c("Yes", "No"))
+
+  get_responses <- function(user_info_dict, uuid) {
+    tmp_survey <- jsonlite::fromJSON(here::here(glue::glue("{survey_folder}/{uuid}.txt")),
+                                     simplifyVector = FALSE)
+    response_time <- map_dbl(tmp_survey, ~.x$rt)
+    user_information <- str_split(tmp_survey[[3]]$response[[1]], ",")[[1]]
+
+    user_information[2:5] <- imap_chr(user_information[-1], ~user_info_dict[[.y]][as.integer(.x)])
+    names(user_information) <- c("prolific_id", "age_group", "education", "pronoun", "previous_experience")
+
+
+    lineup_respone <- str_split(map_chr(tmp_survey[-c(1,2,3,4)], ~.x$response$response), ",")
+    # handle selections
+    selections <- map_chr(lineup_respone, ~.x[1])
+    # handle reasons
+    reasons_dict <- c("Outlier(s)", "Cluster(s)", "Shape")
+
+    reasons <- map_chr(lineup_respone, ~if(.x[2] %in% c("1", "2", "3")) {
+      reasons_dict[as.integer(.x[2])]
+    } else {
+      .x[2]
+    })
+
+    # handle confidence
+    confidence_dict <- c("Not at all", "Slightly", "Moderately", "Very", "Extremely")
+    confidence <- map_chr(lineup_respone, ~if(.x[3] %in% c("1", "2", "3", "4", "5")) {
+      confidence_dict[as.integer(.x[3])]
+    } else {
+      .x[3]
+    })
+
+    tibble(page = 1:length(response_time),
+           response_time = response_time,
+           set = uuid,
+           num = c(rep(NA, 4), 1:k),
+           selection = c(rep(NA, 4), selections),
[…]

### 2022-04-20 paper_comparsion/paper_comparision/paper_comparision.Rmd
add test

@@
 
+## Data plots
+
+Data plots are one of the most important and preferred methods of regression diagnostics.
 Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are refereed to as standard residual plots.
@@
 Overlapping data points is a general issue in scatter plots not limited to standard residual plots, which often makes plots difficult to interpret because visual patterns are concealed.
-Thus, for relatively large sample size, @cleveland1975graphical suggests the use of robust reference lines to give aids to eye in seeing patterns, which nowadays, are usually replaced with a spline or local polynomial regression line.
+Thus, for relatively large sample size, @cleveland1975graphical suggests the use of robust moving statistics as reference lines to give aids to eye in seeing patterns, which nowadays, are usually replaced with a spline or local polynomial regression line.
 
-One of the useful supplements to standard residual plots is partial residual plot.
-While the standard residual plots indicate deviations of each predictor from linearity, the partial residual plots can be used to access the extent and the direction of the linearity.
-These plots can therefore be used to determine the importance of each predictor in the presence of all the others and assess the extent of nonlinearity in a given predictor.
-This is in addition to providing information on correct transformations and extreme data points.
+Other types of data plots that are often used in regression diagnostics include partial residual plots and probability plots.
+Partial residual plots are useful supplements to standard residual plots as they provide additional information on the direction of linearity as well as the nonlinearity of each predictor.
+Probability plots can be used to compare the sampling distribution of the residuals to the normal distribution.
+
+## Hypothesis testing

### 2022-04-20 paper_comparsion/paper_comparision/paper_comparision.tex
add test

@@
 
-Graphical summaries in which residuals are plotted against fitted values
-or other functions of the predictor variables that are approximately
-orthogonal to residuals are refereed to as standard residual plots. They
-are commonly used to identify patterns which are indicative of
-nonconstant error variance or nonlinearity \citep{cook1982residuals}.
-Raw residuals and studentized residuals are the two most frequently used
-residuals in standard residual plots. The debt on which type of
-residuals should be used always present. While raw residuals are the
-most common output of computer regression software package, by applying
-a scaling factor, the ability of revealing nonconstant error variance in
-standard residual plots will often be enhanced by studentized residuals
-in small sample size \citep{gunst2018regression}.
+\hypertarget{data-plots}{%
+\subsection{Data plots}\label{data-plots}}
+
+Data plots are one of the most important and preferred methods of
+regression diagnostics. Graphical summaries in which residuals are
+plotted against fitted values or other functions of the predictor
+variables that are approximately orthogonal to residuals are refereed to
+as standard residual plots. They are commonly used to identify patterns
+which are indicative of nonconstant error variance or nonlinearity
+\citep{cook1982residuals}. Raw residuals and studentized residuals are
+the two most frequently used residuals in standard residual plots. The
+debt on which type of residuals should be used always present. While raw
+residuals are the most common output of computer regression software
+package, by applying a scaling factor, the ability of revealing
+nonconstant error variance in standard residual plots will often be
+enhanced by studentized residuals in small sample size
+\citep{gunst2018regression}.
 
@@
 large sample size, \citet{cleveland1975graphical} suggests the use of
-robust reference lines to give aids to eye in seeing patterns, which
-nowadays, are usually replaced with a spline or local polynomial
-regression line.
+robust moving statistics as reference lines to give aids to eye in
+seeing patterns, which nowadays, are usually replaced with a spline or
+local polynomial regression line.
 
-One of the useful supplements to standard residual plots is partial
-residual plot. While the standard residual plots indicate deviations of
-each predictor from linearity, the partial residual plots can be used to
-access the extent and the direction of the linearity. These plots can
-therefore be used to determine the importance of each predictor in the
-presence of all the others and assess the extent of nonlinearity in a
-given predictor. This is in addition to providing information on correct
-transformations and extreme data points.
+Other types of data plots that are often used in regression diagnostics
+include partial residual plots and probability plots. Partial residual
+plots are useful supplements to standard residual plots as they provide
+additional information on the direction of linearity as well as the
+nonlinearity of each predictor. Probability plots can be used to compare
+the sampling distribution of the residuals to the normal distribution.
+
+\hypertarget{hypothesis-testing}{%
+\subsection{Hypothesis testing}\label{hypothesis-testing}}
 

### 2022-04-20 paper_comparison/paper_comparison/paper_comparison.Rmd
spelling of file names fixed

--- 2022-04-20 paper_comparsion/paper_comparision/paper_comparision.Rmd
@@
 title: |
-  Taylor & Francis Rmarkdown template for authors (\LaTeX\-based 
-  \textsf{Interact} layout + Chicago author-date reference style)
+  Why shouldn't you use numerical tests to diagnose the linear regression models?
 type: ARTICLE TEMPLATE
@@
     email: weihao.li@monash.edu
-  - name: John Smith
-    affil: b
-    email: john.smith@uni-freiburg.de
-  - name: Dominik Leutnant
-    affil: c, \dagger, \ddagger
-    email: leutnant@fh-muenster.de
+  - name: Dianne Cook
+    affil: a
+    email: dicook@monash.edu
+  - name: Emi Tanaka
+    affil: a
+    email: emi.tanaka@monash.edu
 affiliation:
@@
     address: |
-      Taylor & Francis, 4 Park Square, Milton Park, Abingdon, UK
-  - num: b
-    address: |
-      Institut für Informatik, Albert-Ludwigs-Universität, Freiburg, Germany
-  - num: c
-    address: |
-      Muenster University of Applied Sciences - 
-      Institute for Infrastructure, Water, Resources, Environment,
-      Correnstr. 25, 48149 Münster, Germany
-bibliography: interactcadsample.bib
+      Department of Econometrics and Business Statistics, Monash University, Clayton, VIC, Australia
+bibliography: paper.bib
 abstract: |
-  This template is for authors who are preparing a manuscript for a Taylor & 
-  Francis journal using the \LaTeX\ document preparation system and the 
-  `interact} class file, which is available via selected journals' home 
-  pages on the Taylor & Francis website.
+  Abstract to fill.
 keywords: |
-  Sections; lists; figures; tables; mathematics; fonts; references; appendices
+  visual inference; model diagnostics;
 header-includes: |
@@
 
-Title:
-
 # Introduction

### 2022-04-20 paper_comparison/paper_comparison/paper_comparison.tex
spelling of file names fixed

--- 2022-04-20 paper_comparsion/paper_comparision/paper_comparision.tex
@@
 
-\title{Taylor \& Francis Rmarkdown template for authors (\LaTeX-based
-\textsf{Interact} layout + Chicago author-date reference style)}
+\title{Why shouldn't you use numerical tests to diagnose the linear
+regression models?}
 
-\author{\name{Weihao Li$^{a}$, John Smith$^{b}$, Dominik Leutnant$^{c,
-\dagger, \ddagger}$}
-\affil{$^{a}$Taylor \& Francis, 4 Park Square, Milton Park, Abingdon,
-UK; $^{b}$Institut für Informatik, Albert-Ludwigs-Universität, Freiburg,
-Germany; $^{c}$Muenster University of Applied Sciences - Institute for
-Infrastructure, Water, Resources, Environment, Correnstr. 25, 48149
-Münster, Germany}
+\author{\name{Weihao Li$^{a}$, Dianne Cook$^{a}$, Emi Tanaka$^{a}$}
+\affil{$^{a}$Department of Econometrics and Business Statistics, Monash
+University, Clayton, VIC, Australia}
 }
@@
 \thanks{CONTACT Weihao
-Li. Email: \href{mailto:weihao.li@monash.edu}{\nolinkurl{weihao.li@monash.edu}}, John
-Smith. Email: \href{mailto:john.smith@uni-freiburg.de}{\nolinkurl{john.smith@uni-freiburg.de}}, Dominik
-Leutnant. Email: \href{mailto:leutnant@fh-muenster.de}{\nolinkurl{leutnant@fh-muenster.de}}}
+Li. Email: \href{mailto:weihao.li@monash.edu}{\nolinkurl{weihao.li@monash.edu}}, Dianne
+Cook. Email: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}, Emi
+Tanaka. Email: \href{mailto:emi.tanaka@monash.edu}{\nolinkurl{emi.tanaka@monash.edu}}}
 
@@
 \begin{abstract}
-This template is for authors who are preparing a manuscript for a Taylor
-\& Francis journal using the \LaTeX~document preparation system and the
-`interact\} class file, which is available via selected journals' home
-pages on the Taylor \& Francis website.
+Abstract to fill.
 \end{abstract}
@@
 computer vision)
-
-Title:
 

### 2022-04-26 paper_comparison/paper_comparison/paper_comparison.Rmd
update test

@@
 
-Diagnostics of the classical normal linear regression model conventionally involve evaluating the fitness of the proposed model, detecting the presence of influential observations and outliers, checking the validity of model assumptions and many more.
-Tools such as summary statistics, hypothesis testing, and data plots are essential for a systematic and detailed examination of the regression model [@mansfield1987diagnostic].
+Regression diagnostics conventionally involve evaluating the fitness of the proposed model, detecting the presence of influential observations and outliers, checking the validity of model assumptions and many more.
+Common diagnostic techniques including summary statistics, hypothesis testing, and data plots are essential tools for a systematic and detailed examination of the regression model [@mansfield1987diagnostic].
 
-## Data plots
+## Diagnostic plots
 
-Data plots are one of the most important and preferred methods of regression diagnostics.
+Regression analysis is a field of study with at least a hundred years of history.
+Many of those regression diagnostic methods and procedures are mature and well-established in books first published in the twentieth century, such as @draper_applied_2014, @montgomery_introduction_2012, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals.
+Regardless of the level of difficulty of the book, one will find the importance and usefulness of diagnostic plots being emphasized again and again.
+Checking diagnostic plots is also the recommended starting point for validating model assumptions like normality, homoscedasticity and linearity [@anscombe_examination_1963].
+
 Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are refereed to as standard residual plots.
-They are commonly used to identify patterns which are indicative of nonconstant error variance or nonlinearity [@cook1982residuals].
+They are commonly used to identify patterns which are indicative of nonconstant error variance or non-linearity [@cook1982residuals].
 Raw residuals and studentized residuals are the two most frequently used residuals in standard residual plots.
@@
 Other types of data plots that are often used in regression diagnostics include partial residual plots and probability plots.
-Partial residual plots are useful supplements to standard residual plots as they provide additional information on the direction of linearity as well as the nonlinearity of each predictor.
-Probability plots can be used to compare the sampling distribution of the residuals to the normal distribution.
+Partial residual plots are useful supplements to standard residual plots as they provide additional information on the extent of the non-linearity.
+Probability plots can be used to compare the sampling distribution of the residuals to the normal distribution for assessing the normality assumptions.
 
 ## Hypothesis testing
+
+In addition to diagnostic plots, researcher may also perform formal tests for detecting model defects.
+Depends on the alternative, variety of tests can be applied.
+For example, for testing heteroskedasticity,
+one may use the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979].
+And for testing non-linearity, there are RESET test [@ramsey_tests_1969] and F-test.
+
+As discussed in @cook1982residuals, most residual based tests for a particular type of departures from model assumptions are sensitive to other types of departures.
+Especially, outliers will often  incorrectly trigger the rejection of null hypothesis despite the residuals are well-behaved [@cook_applied_1999].
+This can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers.
+Furthermore, @montgomery_introduction_2012 stated that based on their experience, statistical tests are not widely used in regression diagnostics.
+Most importantly, the same or even larger amount of information can be provided by diagnostic plots than the corresponding tests in most empirical studies.
+But still, the effectiveness of statistical tests shall not be disrespected.
+Statistical tests have chance to provide analysts with unique information.
+There are also situations where no suitable diagnostic plots can be found for a particular violation of the assumptions, or excessive diagnostic plots need to be checked.
+One will have no choice but to rely on statistical tests if there is any.
+A good regression diagnostic practice should be a combination of both methods.
+
+## Visual inference
+
+Diagnostic plots are
+
+However, unlike confirmatory data analysis built upon rigorous statistical procedures, e.g., hypothesis testing, visual diagnostics relies on graphical perception - human’s ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984], which is to some extent subjective.
+Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
+One such example is finding an over-interpretation of the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurrence [@roy_chowdhury_using_2015].
+
+Recently, a new branch of statical inference
+
+Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualization" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots [@gelman_exploratory_2004].
+Later, in the Bayesian context, data plots was systematically considered as model diagnostics by taking advantage of the data simulated from the assumed statistical models [@gelman_bayesian_2003; @gelman_exploratory_2004].

### 2022-04-26 paper_comparison/paper_comparison/paper_comparison.tex
update test

@@
 \begin{keywords}
-Sections; lists; figures; tables; mathematics; fonts; references;
-appendices
+visual inference; model diagnostics;
 \end{keywords}
@@
 
-Diagnostics of the classical normal linear regression model
-conventionally involve evaluating the fitness of the proposed model,
-detecting the presence of influential observations and outliers,
-checking the validity of model assumptions and many more. Tools such as
-summary statistics, hypothesis testing, and data plots are essential for
-a systematic and detailed examination of the regression model
+Regression diagnostics conventionally involve evaluating the fitness of
+the proposed model, detecting the presence of influential observations
+and outliers, checking the validity of model assumptions and many more.
+Common diagnostic techniques including summary statistics, hypothesis
+testing, and data plots are essential tools for a systematic and
+detailed examination of the regression model
 \citep{mansfield1987diagnostic}.
 
-\hypertarget{data-plots}{%
-\subsection{Data plots}\label{data-plots}}
+\hypertarget{diagnostic-plots}{%
+\subsection{Diagnostic plots}\label{diagnostic-plots}}
 
-Data plots are one of the most important and preferred methods of
-regression diagnostics. Graphical summaries in which residuals are
-plotted against fitted values or other functions of the predictor
-variables that are approximately orthogonal to residuals are refereed to
-as standard residual plots. They are commonly used to identify patterns
-which are indicative of nonconstant error variance or nonlinearity
-\citep{cook1982residuals}. Raw residuals and studentized residuals are
-the two most frequently used residuals in standard residual plots. The
-debt on which type of residuals should be used always present. While raw
-residuals are the most common output of computer regression software
-package, by applying a scaling factor, the ability of revealing
-nonconstant error variance in standard residual plots will often be
-enhanced by studentized residuals in small sample size
-\citep{gunst2018regression}.
+Regression analysis is a field of study with at least a hundred years of
+history. Many of those regression diagnostic methods and procedures are
+mature and well-established in books first published in the twentieth
+century, such as \citet{draper_applied_2014},
+\citet{montgomery_introduction_2012}, \citet{belsley_regression_1980},
+\citet{cook_applied_1999} and \citet{cook1982residuals}. Regardless of
+the level of difficulty of the book, one will find the importance and
+usefulness of diagnostic plots being emphasized again and again.
+Checking diagnostic plots is also the recommended starting point for
+validating model assumptions like normality, homoscedasticity and
+linearity \citep{anscombe_examination_1963}.
+
+Graphical summaries in which residuals are plotted against fitted values
+or other functions of the predictor variables that are approximately
+orthogonal to residuals are refereed to as standard residual plots. They
+are commonly used to identify patterns which are indicative of
+nonconstant error variance or non-linearity \citep{cook1982residuals}.
+Raw residuals and studentized residuals are the two most frequently used
+residuals in standard residual plots. The debt on which type of
+residuals should be used always present. While raw residuals are the
+most common output of computer regression software package, by applying
+a scaling factor, the ability of revealing nonconstant error variance in
+standard residual plots will often be enhanced by studentized residuals
+in small sample size \citep{gunst2018regression}.
 
@@
 plots are useful supplements to standard residual plots as they provide
-additional information on the direction of linearity as well as the
-nonlinearity of each predictor. Probability plots can be used to compare
-the sampling distribution of the residuals to the normal distribution.
+additional information on the extent of the non-linearity. Probability
+plots can be used to compare the sampling distribution of the residuals
+to the normal distribution for assessing the normality assumptions.
 
@@
 
+In addition to diagnostic plots, researcher may also perform formal
+tests for detecting model defects. Depends on the alternative, variety
+of tests can be applied. For example, for testing heteroskedasticity,
+one may use the White test
+\citep{white_heteroskedasticity-consistent_1980} or the Breusch-Pagan
+test \citep{breusch_simple_1979}. And for testing non-linearity, there
+are RESET test \citep{ramsey_tests_1969} and F-test.
+
+As discussed in \citet{cook1982residuals}, most residual based tests for
+a particular type of departures from model assumptions are sensitive to
+other types of departures. Especially, outliers will often incorrectly
+trigger the rejection of null hypothesis despite the residuals are
+well-behaved \citep{cook_applied_1999}. This can be largely avoided in
+diagnostic plots as experienced analysts can evaluate the acceptability
+of assumptions flexibly, even in the presence of outliers. Furthermore,
+\citet{montgomery_introduction_2012} stated that based on their
+experience, statistical tests are not widely used in regression
+diagnostics. Most importantly, the same or even larger amount of
+information can be provided by diagnostic plots than the corresponding
+tests in most empirical studies. But still, the effectiveness of
+statistical tests shall not be disrespected. Statistical tests have
+chance to provide analysts with unique information. There are also
+situations where no suitable diagnostic plots can be found for a
+particular violation of the assumptions, or excessive diagnostic plots
+need to be checked. One will have no choice but to rely on statistical
+tests if there is any. A good regression diagnostic practice should be a
+combination of both methods.
+
+\hypertarget{visual-inference}{%
+\subsection{Visual inference}\label{visual-inference}}
+
+Diagnostic plots are
+
+However, unlike confirmatory data analysis built upon rigorous
+statistical procedures, e.g., hypothesis testing, visual diagnostics
+relies on graphical perception - human's ability to interpret and decode
+the information embedded in the graph \citep{cleveland_graphical_1984},
+which is to some extent subjective. Further, visual discovery suffers
+from its unsecured and unconfirmed nature where the degree of the
+presence of the visual features typically can not be measured
+quantitatively and objectively, which may lead to over or
+under-interpretations of the data. One such example is finding an
+over-interpretation of the separation between gene groups in a
+two-dimensional projection from a linear discriminant analysis when in
+fact there are no differences in the expression levels between the gene
+groups and separation is not an uncommon occurrence
+\citep{roy_chowdhury_using_2015}.
+
+Recently, a new branch of statical inference
+
+Visual inference was first introduced in a 1999 Joint Statistical
+Meetings (JSM) talk with the title ``Inference for Data Visualization''
+by \citet{buja_inference_1999} as an idea to address the issue of valid
+inference for visual discoveries of data plots
+\citep{gelman_exploratory_2004}. Later, in the Bayesian context, data
+plots was systematically considered as model diagnostics by taking
+advantage of the data simulated from the assumed statistical models
+\citep{gelman_bayesian_2003, gelman_exploratory_2004}.
+
 \bibliographystyle{tfcad}
-\bibliography{interactcadsample.bib}
+\bibliography{paper.bib}
 

### 2022-05-03 paper_comparison/scripts/process.R
clean up scripts

@@ -1,0 +1,229 @@
+library(tidyverse)
+library(visage)
+
+preprocessing <- new.env()
+
+# Process raw pilot data --------------------------------------------------
+
+preprocessing$pilot <- new.env(parent = preprocessing)
+
+# Read and write to csv ---------------------------------------------------
+
+# The effect size for the heter model could be incorrect
+preprocessing$pilot$process_pilot <- function(write = TRUE) {
+  file_names <- list.files(here::here("data/pilot_study"))
+
+  result_high <- NULL
+
+  for (file_name in file_names[grep("oct" , file_names)]) {
+    dat <- read_csv(here::here("data/pilot_study", file_name), col_types = cols())
+    if (is.null(result_high)) {
+      result_high <- dat
+    } else {
+      result_high <- bind_rows(result_high, dat)
+    }
+  }
+
+  result_heter <- NULL
+
+  for (file_name in file_names[grep("nov" , file_names)]) {
+    dat <- read_csv(here::here("data/pilot_study", file_name), col_types = cols())
+    if (is.null(result_heter)) {
+      result_heter <- dat
+    } else {
+      result_heter <- bind_rows(result_heter, dat)
+    }
+  }
+
+  tmp <- bind_rows(result_high, result_heter)
+  tmp <- mutate(tmp, model = ifelse(model == "higher order", "cubic", model)) %>%
+    rename(detect = answer, type = model)
+
+  if (write) write_csv(tmp, file = here::here("data/processed/processed_pilot.csv"))
+
+  return(tmp)
+}
+
+
+# Fit GLM for cubic and heter models --------------------------------------
+
+preprocessing$pilot$fit_pilot_models <- function(pilot_dat) {
+
+  pilot_dat <- pilot_dat %>%
+    mutate(detect = as.numeric(detect),
+           log_effect_size = log(effect_size))
+
+  mod1 <- glm(detect ~ log(effect_size),
+              data = filter(pilot_dat, type == "cubic"),
+              family = binomial)
+
+  mod2 <- glm(detect ~ log(effect_size),
+              data = filter(pilot_dat, type != "cubic"),
+              family = binomial)
+
+  return(list(mod1, mod2))
+}
+
+# with(preprocessing$pilot, fit_pilot_models(process_pilot()))
+
+
+
+# Process survey data -----------------------------------------------------
+
+preprocessing$survey <- new.env(parent = preprocessing)
+
+
+# Extract individual survey result ----------------------------------------
+
[…]

### 2022-05-05 paper_comparison/paper_comparison/paper_comparison.Rmd
clean up scripts
update scripts, processed data and result

@@
 (visual test most different one (everything test, any departure))
+plot figure in a paper, desc, exp
 2. investigate the difference (gap), give examples
@@
 
-However, unlike confirmatory data analysis built upon rigorous statistical procedures, e.g., hypothesis testing, visual diagnostics relies on graphical perception - human’s ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984], which is to some extent subjective.
+<!-- However, unlike confirmatory data analysis built upon rigorous statistical procedures, e.g., hypothesis testing, visual diagnostics relies on graphical perception - human’s ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984], which is to some extent subjective.
 Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
 One such example is finding an over-interpretation of the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurrence [@roy_chowdhury_using_2015].
+-->
 
-Recently, a new branch of statical inference
+<!-- Recently, a new branch of statical inference  -->
 
-Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualization" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots [@gelman_exploratory_2004].
+<!-- Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualization" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots [@gelman_exploratory_2004].
 Later, in the Bayesian context, data plots was systematically considered as model diagnostics by taking advantage of the data simulated from the assumed statistical models [@gelman_bayesian_2003; @gelman_exploratory_2004].
+-->
+
+# Experimental design
+
+# Data processing
+
+# Results
+
+## Overview of the Data
+
+We collected 400 lineup evaluations made by 20 participants in experiment I and 880 lineup evaluations made by 44 participants in experiment II.
+In total, 442 unique lineups were evaluated by 64 subjects.
+In experiment I, one of the participants skipped all 20 lineups.
+Hence, the submission was rejected and removed from the dataset.
+In experiment II, there was a participant failed one of the two attention checks, but there was no further evidence of low-effort throughout the experiment.
+Therefore, the submission was kept.
+
+## Power comparision
+
+1. power (visual test vs. conventional test)
+(visual test most different one (everything test, any departure))
+plot figure in a paper, desc, exp
+2. investigate the difference (gap), give examples
+3. conventional is too sensitive
+4. make conventional less sensitive (vary alpha)
+
+It was expected that with larger effect size, both conventional test and visual test will have higher probability in rejecting the null hypothesis when it is not true.
+To explore the difference in power between conventional test and visual test, ten logistic regression models are fit to the data, with natural logarithm of the effect size as the only fixed effect, and whether the test successfully rejects the null hypothesis as the response variable.
+Table \ref{tab:powerglmheter} and \ref{tab:powerglmheter} summarizes the results.
+The natural logarithm of the effect size is positive and significant across all models.
+Figure \ref{fig:power-com} illustrates the fitted models, while providing the local constant estimate of the power of F-test and Breusch–Pagan test for comparison.
+Data for the conventional test is simulated under the same model setting used in two experiments and 500000 samples are drawn for both cubic and heteroskedasticity model.
+
+From Figure \ref{fig:power-com}, it can be observed that the fitted power of visual test increased as the number of evaluations increased for both cubic and heteroskedasticity model.
+
+For heteroskedasticity model, this phenomenon was more obvious as the curves of visual tests with evaluations greater than two were always above the curves of visual test with evaluations smaller than two.
+However, this only held for large enough effect size.
+For small effect size, visual tests with fewer evaluations might have greater power.
+Note that, the expected power of visual test derived by @majumder_validation_2013 followed by the assumption that human has the ability to select the plot with the highest t-statistic from a lineup under the classical linear model regression setting showed similar properties, where visual tests with fewer evaluations were expected to perform better when the parameter values close to the null hypothesis.
+
+For cubic model, the separation between curves was small.
+Fitted power of visual test with three to five evaluations were almost identical to each other in regards of effect size.
+In addition, all five curves peaked at one as effect size increased, suggesting that identification of non-linearity as a visual task can be completed reliably by human when the effect size is large enough.
+
+As shown in Figure \@ref(fig:power-com), both F-test and Breusch–Pagan test generally possessed greater power than visual test.
+It was also found that there was a huge gap between the conventional test curves and the visual test curves at around $log(\text{effect size}) = 0$ for the cubic model and $log(\text{effect size}) = 2.5$ for the heteroskedasticity model.
+
+We further analysed the lineup with the corresponding effect size.
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
+```
+
+```{r}
+library(tidyverse)
+library(visage)
+pilot <- read_csv(here::here("data/processed/processed_pilot.csv"))
+survey <- read_csv(here::here("data/processed/processed_surveys.csv"))
+sim_conv <- read_csv(here::here("data/processed/sim_conv_result.csv"), col_types = "dddlc")
+```
+
+```{r}
+survey <- survey %>%
+  mutate(log_effect_size = log(effect_size))
+
+sim_conv <- sim_conv %>%
+  mutate(log_effect_size = log(effect_size))
+
+lineup_info <- select(survey, type:answer, conventional_p_value, ori_difficulty, exp, exp_lineup_id, log_effect_size) %>%
+  group_by(exp_lineup_id) %>% 
+  filter(row_number() == 1) %>%
+  ungroup()
+
+write_csv(lineup_info, here::here("data/processed/processed_lineup_info.csv"))
+
+# Local constant estimator
+kernel_regression_cubic <- function(h = 0.3) {
+  with(filter(sim_conv, type == "cubic", log_effect_size > -5, log_effect_size < 10),
+     KernSmooth::locpoly(x = log_effect_size, y = reject, bandwidth = h, degree = 0, gridsize = 1000)) %>%
+    as_tibble() %>%
+    mutate(type = "cubic", h = h) %>%
+    rename(log_effect_size = x) %>%
+    rename(reject = y)
+}
+
+# Local constant estimator
+kernel_regression_heter <- function(h = 0.4) {
+  with(filter(sim_conv, type != "cubic", log_effect_size > -3),
+     KernSmooth::locpoly(x = log_effect_size, y = reject, bandwidth = h, degree = 0, gridsize = 1000)) %>%
+    as_tibble() %>%
+    mutate(type = "heteroskedasticity", h = h) %>%
+    rename(log_effect_size = x) %>%
+    rename(reject = y)
+}
+```
+
+```{r}
+if (file.exists(here::here("data/processed/lineup_env.rds"))) {
+  env <- readRDS(here::here("data/processed/lineup_env.rds"))
+} else {
+  env <- new.env()
+}
+
+save_lineup_env <- function(env) {
+  saveRDS(env, here::here("data/processed/lineup_env.rds"))
+}
+
+survey_pvalue <- survey %>%
+  calc_p_value_multi(lineup_id = "exp_lineup_id", 
+                     detected = "detect", 
+                     n_sel = "num_selection",
+                     comb = TRUE,
+                     n_eval = 1:5,
+                     n_sim = 100000,
+                     cache_env = env) %>%
+  rename(exp_lineup_id = lineup_id) %>%
+  unnest_longer(col = p_value) %>%
+  left_join(lineup_info) %>%
+  mutate(reject = eval_p_value(p_value, tol = 0.01))
+
+save_lineup_env(env)
+```
+
+```{r results='asis'}
+map(1:5, function(x) glm(reject ~ log_effect_size,
+                         family = binomial(),
+                         data = filter(survey_pvalue, n_eval == {{x}}, type == "cubic"))) %>%
+  stargazer::stargazer(header = FALSE,
+            label = "tab:powerglmcubic",
+            title = "Logistic regressions with rejection as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for cubic model.
+Five regression were fitted for different number of evaluations.",
+            style = "default",
+            column.labels = c("num\\_eval: 1", "num\\_eval:2", "num\\_eval: 3", "num\\_eval: 4", "num\\_eval:5"),
+            model.numbers = FALSE)
+```
+
+```{r results="asis"}
+map(1:5, function(x) glm(reject ~ log_effect_size,
+                         family = binomial(),
+                         data = filter(survey_pvalue, n_eval == {{x}}, type == "heteroskedasticity"))) %>%
+  stargazer::stargazer(header = FALSE,
+            label = "tab:powerglmheter",
+            title = "Logistic regressions with rejection as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for heteroskedasticity model.
+Five regression were fitted for different number of evaluations.",
+            style = "default",
+            column.labels = c("num\\_eval: 1", "num\\_eval:2", "num\\_eval: 3", "num\\_eval: 4", "num\\_eval:5"),
+            model.numbers = FALSE)
+```
+
+```{r power-com, fig.cap="\\label{fig:power-com}power com"}
+survey_pvalue %>%
+  mutate(reject = as.numeric(reject)) %>%
+  mutate(n_eval = factor(n_eval)) %>%
+  ggplot() +
+  geom_point(aes(log_effect_size, reject), alpha = 0.01) +
+  geom_line(data = kernel_regression_cubic(), aes(log_effect_size, reject), linetype = 2) +
+  geom_line(data = kernel_regression_heter(), aes(log_effect_size, reject), linetype = 2) +
+  geom_smooth(aes(log_effect_size, reject, group = n_eval, col = n_eval),
+              method = "glm", method.args = list(family = binomial), size = 0.5, se = FALSE) +
+  scale_colour_viridis_d(option = "B", begin = 0.5, end = 0.9) +
+  facet_wrap(~type, scales = "free_x") +
+  ylab("Estimated power")
+```

### 2022-05-05 paper_comparison/paper_comparison/paper_comparison.tex
update scripts, processed data and result

@@
 test) (visual test most different one (everything test, any departure))
-2. investigate the difference (gap), give examples 3. conventional is
-too sensitive 4. make conventional less sensitive (vary alpha)
+plot figure in a paper, desc, exp 2. investigate the difference (gap),
+give examples 3. conventional is too sensitive 4. make conventional less
+sensitive (vary alpha)
 
@@
 
-However, unlike confirmatory data analysis built upon rigorous
-statistical procedures, e.g., hypothesis testing, visual diagnostics
-relies on graphical perception - human's ability to interpret and decode
-the information embedded in the graph \citep{cleveland_graphical_1984},
-which is to some extent subjective. Further, visual discovery suffers
-from its unsecured and unconfirmed nature where the degree of the
-presence of the visual features typically can not be measured
-quantitatively and objectively, which may lead to over or
-under-interpretations of the data. One such example is finding an
-over-interpretation of the separation between gene groups in a
-two-dimensional projection from a linear discriminant analysis when in
-fact there are no differences in the expression levels between the gene
-groups and separation is not an uncommon occurrence
-\citep{roy_chowdhury_using_2015}.
+\hypertarget{experimental-design}{%
+\section{Experimental design}\label{experimental-design}}
 
-Recently, a new branch of statical inference
+\hypertarget{data-processing}{%
+\section{Data processing}\label{data-processing}}
 
-Visual inference was first introduced in a 1999 Joint Statistical
-Meetings (JSM) talk with the title ``Inference for Data Visualization''
-by \citet{buja_inference_1999} as an idea to address the issue of valid
-inference for visual discoveries of data plots
-\citep{gelman_exploratory_2004}. Later, in the Bayesian context, data
-plots was systematically considered as model diagnostics by taking
-advantage of the data simulated from the assumed statistical models
-\citep{gelman_bayesian_2003, gelman_exploratory_2004}.
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+\hypertarget{overview-of-the-data}{%
+\subsection{Overview of the Data}\label{overview-of-the-data}}
+
+We collected 400 lineup evaluations made by 20 participants in
+experiment I and 880 lineup evaluations made by 44 participants in
+experiment II. In total, 442 unique lineups were evaluated by 64
+subjects. In experiment I, one of the participants skipped all 20
+lineups. Hence, the submission was rejected and removed from the
+dataset. In experiment II, there was a participant failed one of the two
+attention checks, but there was no further evidence of low-effort
+throughout the experiment. Therefore, the submission was kept.
+
+\hypertarget{power-comparision}{%
+\subsection{Power comparision}\label{power-comparision}}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  power (visual test vs.~conventional test) (visual test most different
+  one (everything test, any departure)) plot figure in a paper, desc,
+  exp
+\item
+  investigate the difference (gap), give examples
+\item
+  conventional is too sensitive
+\item
+  make conventional less sensitive (vary alpha)
+\end{enumerate}
+
+It was expected that with larger effect size, both conventional test and
+visual test will have higher probability in rejecting the null
+hypothesis when it is not true. To explore the difference in power
+between conventional test and visual test, ten logistic regression
+models are fit to the data, with natural logarithm of the effect size as
+the only fixed effect, and whether the test successfully rejects the
+null hypothesis as the response variable. Table \ref{tab:powerglmheter}
+and \ref{tab:powerglmheter} summarizes the results. The natural
+logarithm of the effect size is positive and significant across all
+models. Figure \ref{fig:power-com} illustrates the fitted models, while
+providing the local constant estimate of the power of F-test and
+Breusch--Pagan test for comparison. Data for the conventional test is
+simulated under the same model setting used in two experiments and
+500000 samples are drawn for both cubic and heteroskedasticity model.
+
+From Figure \ref{fig:power-com}, it can be observed that the fitted
+power of visual test increased as the number of evaluations increased
+for both cubic and heteroskedasticity model.
+
+For heteroskedasticity model, this phenomenon was more obvious as the
+curves of visual tests with evaluations greater than two were always
+above the curves of visual test with evaluations smaller than two.
+However, this only held for large enough effect size. For small effect
+size, visual tests with fewer evaluations might have greater power. Note
+that, the expected power of visual test derived by
+\citet{majumder_validation_2013} followed by the assumption that human
+has the ability to select the plot with the highest t-statistic from a
+lineup under the classical linear model regression setting showed
+similar properties, where visual tests with fewer evaluations were
+expected to perform better when the parameter values close to the null
+hypothesis.
+
+For cubic model, the separation between curves was small. Fitted power
+of visual test with three to five evaluations were almost identical to
+each other in regards of effect size. In addition, all five curves
+peaked at one as effect size increased, suggesting that identification
+of non-linearity as a visual task can be completed reliably by human
+when the effect size is large enough.
+
+As shown in Figure @ref(fig:power-com), both F-test and Breusch--Pagan
+test generally possessed greater power than visual test. It was also
+found that there was a huge gap between the conventional test curves and
+the visual test curves at around \(log(\text{effect size}) = 0\) for the
+cubic model and \(log(\text{effect size}) = 2.5\) for the
+heteroskedasticity model.
+
+We further analysed the lineup with the corresponding effect size.
+
+\begin{table}[!htbp] \centering 
+  \caption{Logistic regressions with rejection as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for cubic model.
+Five regression were fitted for different number of evaluations.}
+  \label{tab:powerglmcubic} 
+\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{5}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-6} 
+\\[-1.8ex] & \multicolumn{5}{c}{reject} \\ 
+ & num\_eval: 1 & num\_eval:2 & num\_eval: 3 & num\_eval: 4 & num\_eval:5 \\ 
+\hline \\[-1.8ex] 
+ log\_effect\_size & 0.929$^{***}$ & 1.258$^{***}$ & 1.096$^{***}$ & 1.113$^{***}$ & 1.244$^{***}$ \\ 
+  & (0.073) & (0.076) & (0.054) & (0.056) & (0.077) \\ 
+  & & & & & \\ 
+ Constant & $-$1.969$^{***}$ & $-$2.534$^{***}$ & $-$1.549$^{***}$ & $-$1.334$^{***}$ & $-$1.404$^{***}$ \\ 
+  & (0.169) & (0.173) & (0.113) & (0.122) & (0.185) \\ 
+  & & & & & \\ 
+\hline \\[-1.8ex] 
+Observations & 630 & 1,135 & 1,660 & 1,904 & 1,879 \\ 
+Log Likelihood & $-$220.882 & $-$314.394 & $-$471.152 & $-$396.979 & $-$206.296 \\ 
+Akaike Inf. Crit. & 445.764 & 632.788 & 946.305 & 797.958 & 416.592 \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{5}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+\end{tabular} 
+\end{table}
+
+\begin{table}[!htbp] \centering 
+  \caption{Logistic regressions with rejection as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for heteroskedasticity model.
+Five regression were fitted for different number of evaluations.}
+  \label{tab:powerglmheter} 
+\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{5}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-6} 
+\\[-1.8ex] & \multicolumn{5}{c}{reject} \\ 
+ & num\_eval: 1 & num\_eval:2 & num\_eval: 3 & num\_eval: 4 & num\_eval:5 \\ 
+\hline \\[-1.8ex] 
+ log\_effect\_size & 0.821$^{***}$ & 1.104$^{***}$ & 1.157$^{***}$ & 1.387$^{***}$ & 1.651$^{***}$ \\ 
+  & (0.086) & (0.079) & (0.066) & (0.076) & (0.110) \\ 
+  & & & & & \\ 
+ Constant & $-$4.472$^{***}$ & $-$5.887$^{***}$ & $-$5.431$^{***}$ & $-$6.154$^{***}$ & $-$7.123$^{***}$ \\ 
+  & (0.434) & (0.415) & (0.344) & (0.400) & (0.591) \\ 
+  & & & & & \\ 
+\hline \\[-1.8ex] 
+Observations & 630 & 1,138 & 1,596 & 1,662 & 1,416 \\ 
+Log Likelihood & $-$333.119 & $-$631.519 & $-$835.317 & $-$650.570 & $-$342.058 \\ 
+Akaike Inf. Crit. & 670.238 & 1,267.038 & 1,674.634 & 1,305.140 & 688.116 \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{5}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+\end{tabular} 
+\end{table}
+
+\begin{figure}
+\centering
+\includegraphics{paper_comparison_files/figure-latex/power-com-1.pdf}
+\caption{\label{fig:power-com}power com}
+\end{figure}
 

### 2022-05-05 paper_comparison/scripts/process.R

@@ -75,7 +75,7 @@ preprocessing$survey <- new.env(parent = preprocessing)
 
 # Extract individual survey result ----------------------------------------
 
-preprocessing$survey$get_responses <- function(survey_folder, uuid, user_info_dict) {
+preprocessing$survey$get_responses <- function(survey_folder, uuid, user_info_dict, k = 20) {
 
   survey_result <- jsonlite::fromJSON(here::here(glue::glue("{survey_folder}/{uuid}.txt")),
                                       simplifyVector = FALSE)
@@ -157,6 +157,9 @@ preprocessing$survey$detect_or_not <- function(selection, answer) {
 }
 
 
+
+# process responses in a survey -------------------------------------------
+
 # `survey_folder` is the one contatins all the .txt files
 # `k` is the number of lineups shown to a participant
 # `lineup_dat` is a list. Each item contains a list of metadata, and a dataframe of the observations
@@ -187,9 +190,9 @@ preprocessing$survey$process_responses <- function(survey_folder, k = 20, lineup
 
   for (uuid in uuids) {
     if (is.null(response_dat)) {
-      response_dat <- get_responses(survey_folder, uuid, user_info_dict)
+      response_dat <- get_responses(survey_folder, uuid, user_info_dict, k = 20)
     } else {
-      response_dat <- bind_rows(response_dat, get_responses(survey_folder, uuid, user_info_dict))
+      response_dat <- bind_rows(response_dat, get_responses(survey_folder, uuid, user_info_dict, k = 20))
     }
   }
 
@@ -212,7 +215,7 @@ preprocessing$survey$process_responses <- function(survey_folder, k = 20, lineup
     mutate(detect = detect_or_not(selection, answer)) %>%
 
     # Perform the conventional tests and get the p-values
-    mutate(conventional_p_value = map2_dbl(tmp$lineup_id, tmp$type,
+    mutate(conventional_p_value = map2_dbl(lineup_id, type,
                                            function(lineup_id, type) {
                                              if (is.na(lineup_id)) return(NA)
 
@@ -221,9 +224,195 @@ preprocessing$survey$process_responses <- function(survey_folder, k = 20, lineup
                                              } else {
[…]

### 2022-05-12 paper_comparison/paper_comparison/paper_comparison.Rmd
break down factors in power comparison

@@
 
-It was expected that with larger effect size, both conventional test and visual test will have higher probability in rejecting the null hypothesis when it is not true.
-To explore the difference in power between conventional test and visual test, ten logistic regression models are fit to the data, with natural logarithm of the effect size as the only fixed effect, and whether the test successfully rejects the null hypothesis as the response variable.
-Table \ref{tab:powerglmheter} and \ref{tab:powerglmheter} summarizes the results.
-The natural logarithm of the effect size is positive and significant across all models.
+To model the power of visual test, 10 logistic regression were fit for different number of evaluations ranged from one to five and two different types of simulation setting.
+All 10 models used natural logarithm of the effect size as the only fixed effect, and whether the test successfully rejects the null hypothesis as the response variable.
+Given the way we define the effect size, it was expected that with larger effect size, both conventional test and visual test will have higher probability in rejecting the null hypothesis when it is not true.
+The modelling result summarized in \ref{tab:powerglmcubic} and \ref{tab:powerglmheter} aligned with the expectation as the coefficients of natural logarithm of the effect size are positive and significant across all 10 models.
+
 Figure \ref{fig:power-com} illustrates the fitted models, while providing the local constant estimate of the power of F-test and Breusch–Pagan test for comparison.
-Data for the conventional test is simulated under the same model setting used in two experiments and 500000 samples are drawn for both cubic and heteroskedasticity model.
-
+Data for the conventional test is simulated under the model setting described in section ...
+and 5000000 samples are drawn for both cubic and heteroskedasticity model.
 From Figure \ref{fig:power-com}, it can be observed that the fitted power of visual test increased as the number of evaluations increased for both cubic and heteroskedasticity model.
 
-For heteroskedasticity model, this phenomenon was more obvious as the curves of visual tests with evaluations greater than two were always above the curves of visual test with evaluations smaller than two.
-However, this only held for large enough effect size.
-For small effect size, visual tests with fewer evaluations might have greater power.
-Note that, the expected power of visual test derived by @majumder_validation_2013 followed by the assumption that human has the ability to select the plot with the highest t-statistic from a lineup under the classical linear model regression setting showed similar properties, where visual tests with fewer evaluations were expected to perform better when the parameter values close to the null hypothesis.
+For heteroskedasticity model, this phenomenon was more obvious as the power of visual tests with evaluations greater than two were always greater than those with evaluations smaller than two.
 
 For cubic model, the separation between curves was small.
-Fitted power of visual test with three to five evaluations were almost identical to each other in regards of effect size.
-In addition, all five curves peaked at one as effect size increased, suggesting that identification of non-linearity as a visual task can be completed reliably by human when the effect size is large enough.
+The estimated power of visual tests with three to five evaluations were almost identical to each other in regards of effect size.
+In addition, all five curves peaked at one as effect size increased, suggesting that identification of non-linearity as a visual task can be completed reliably by human as long as the departure from null hypothesis is large enough.
 
-As shown in Figure \@ref(fig:power-com), both F-test and Breusch–Pagan test generally possessed greater power than visual test.
-It was also found that there was a huge gap between the conventional test curves and the visual test curves at around $log(\text{effect size}) = 0$ for the cubic model and $log(\text{effect size}) = 2.5$ for the heteroskedasticity model.
+As shown in Figure \ref{fig:power-com}, both F-test and Breusch–Pagan test generally possessed greater power than visual test.
+A visual tests is a collection of test against any alternatives that would create visual discoverable features, while a conventional test is usually targeting at a pre-specified alternative.
+Considering the data generating process of the model defect was known and controlled in this research, where all other alternatives have been eliminated except the one we concerned, the result was suggested that conventional tests were more sensitive to violations of linearity and homoscedasticity assumption than visual tests.
 
-We further analysed the lineup with the corresponding effect size.
+It was also found that there was a noticeable gap between curves of the conventional test and the visual test at around $log(\text{effect size}) = 0$ for the cubic model and $log(\text{effect size}) = 2.5$ for the heteroskedasticity model, where the differences in power were greater than 0.6.
+We further analysed the lineups with correspoding effect sizes.
+Figure \ref{fig:cubic-hard} and \ref{fig:heter-hard} showed that human was indeed hard to identify the patterns at this level of difficulty.
+The visual difference between the true data plot and null plots were almost unnoticeable.
+
+## Effect of parameters on power of the visual test
+
+The previous section focuses on the change of effect size relative to the power of the visual test.
 
@@
 survey <- read_csv(here::here("data/processed/processed_surveys.csv"))
-sim_conv <- read_csv(here::here("data/processed/sim_conv_result.csv"), col_types = "dddlc")
+sim_conv <- read_csv(here::here("data/processed/sim_conv_result.csv"), col_types = "dddddcdddlcdcdddlc")
 ```
@@
 # Local constant estimator
-kernel_regression_cubic <- function(h = 0.3) {
+kernel_regression_cubic <- function(sim_conv, h = 0.3, x = quote(log_effect_size)) {
   with(filter(sim_conv, type == "cubic", log_effect_size > -5, log_effect_size < 10),
-     KernSmooth::locpoly(x = log_effect_size, y = reject, bandwidth = h, degree = 0, gridsize = 1000)) %>%
+     KernSmooth::locpoly(x = eval(x), y = reject, bandwidth = h, degree = 0, gridsize = 1000)) %>%
     as_tibble() %>%
     mutate(type = "cubic", h = h) %>%
-    rename(log_effect_size = x) %>%
     rename(reject = y)
@@
 # Local constant estimator
-kernel_regression_heter <- function(h = 0.4) {
+kernel_regression_heter <- function(sim_conv, h = 0.4, x = quote(log_effect_size)) {
   with(filter(sim_conv, type != "cubic", log_effect_size > -3),
-     KernSmooth::locpoly(x = log_effect_size, y = reject, bandwidth = h, degree = 0, gridsize = 1000)) %>%
+     KernSmooth::locpoly(x = eval(x), y = reject, bandwidth = h, degree = 0, gridsize = 1000)) %>%
     as_tibble() %>%
     mutate(type = "heteroskedasticity", h = h) %>%
-    rename(log_effect_size = x) %>%
     rename(reject = y)
@@
   geom_point(aes(log_effect_size, reject), alpha = 0.01) +
-  geom_line(data = kernel_regression_cubic(), aes(log_effect_size, reject), linetype = 2) +
-  geom_line(data = kernel_regression_heter(), aes(log_effect_size, reject), linetype = 2) +
+  geom_line(data = kernel_regression_cubic(sim_conv, 0.4), aes(x, reject), linetype = 2) +
+  geom_line(data = kernel_regression_heter(sim_conv, 0.75), aes(x, reject), linetype = 2) +
   geom_smooth(aes(log_effect_size, reject, group = n_eval, col = n_eval),
@@
 ```
+
+```{r cubic-hard, fig.cap="\\label{fig:cubic-hard}cubic-hard ans 11"}
+big_lineup_dat <- readRDS(here::here("data/big_study/all_data_big.RDS"))
+
+# ans 11, conventional 0.003491033
+CUBIC_MODEL$plot_lineup(mutate(big_lineup_dat[[28]]$data, .fitted = .pred),
+                        remove_axis = TRUE,
+                        theme = theme_light(),
+                        remove_grid_line = TRUE)
+
+```
+
+```{r heter-hard, fig.cap="\\label{fig:heter-hard}heter-hard ans 2"}
+# ans 2, conventional 0.003535915
+HETER_MODEL$plot_lineup(mutate(big_lineup_dat[[136]]$data, .fitted = .pred),
+                        remove_axis = TRUE,
+                        theme = theme_light(),
+                        remove_grid_line = TRUE)
+```
+
+```{r power-com-cubic-a, fig.cap="\\label{fig:power-com-cubic-a}power com cubic a"}
+bind_grid <- expand.grid(n = c(50, 100, 300),
+                         h = 0.4,
+                         x = c(quote(abs(a)/e_sigma)))
+
+bind_list <- list()
+
+for (i in 1:nrow(bind_grid)) {
+  (function(n, h, x) {
+    mutate(kernel_regression_cubic(filter(sim_conv, n == {{n}}), h = h, x = x),
+           n = factor(n, levels = c(50, 100, 300)))
+    })(n = bind_grid$n[i], h = bind_grid$h[i], x = bind_grid$x[[i]]) -> bind_list[[i]]
+}
+
+bind_sim <- reduce(bind_list, bind_rows)
+
+# cubic break down
+survey_pvalue %>%
+  filter(type == "cubic") %>%
+  mutate(n = factor(n, levels = c(50, 100, 300))) %>%
+  mutate(n_eval = factor(n_eval)) %>%
+  mutate(reject = as.numeric(reject)) %>%
+  ggplot() +
+  ggbeeswarm::geom_quasirandom(aes(abs(a)/e_sigma, reject, group = n), groupOnX = FALSE, alpha = 0.01) +
+  geom_smooth(aes(abs(a)/e_sigma, reject, col = n_eval), method = "glm", method.args = list(family = binomial()), se = FALSE) +
+  geom_line(data = bind_sim, aes(x, reject), col = "grey", linetype = 2, size = 1) +
+  facet_wrap(~n, labeller = label_both) +
+  scale_colour_viridis_d(option = "B", begin = 0.5, end = 0.9) +
+  theme_light()
+```
+
+```{r power-com-cubic-b, fig.cap="\\label{fig:power-com-cubic-b}power com cubic b"}
+bind_grid <- expand.grid(n = c(50, 100, 300),
+                         h = 0.4,
+                         x = c(quote(abs(b)/e_sigma)))
+
+bind_list <- list()
+
+for (i in 1:nrow(bind_grid)) {
+  (function(n, h, x) {
+    mutate(kernel_regression_cubic(filter(sim_conv, n == {{n}}), h = h, x = x),
+           n = factor(n, levels = c(50, 100, 300)))
+    })(n = bind_grid$n[i], h = bind_grid$h[i], x = bind_grid$x[[i]]) -> bind_list[[i]]
+}
+
+bind_sim <- reduce(bind_list, bind_rows)
+
+# cubic break down
+survey_pvalue %>%
+  filter(type == "cubic") %>%
+  mutate(n = factor(n, levels = c(50, 100, 300))) %>%
+  mutate(reject = as.numeric(reject)) %>%
+  mutate(n_eval = factor(n_eval)) %>%
+  ggplot() +
+  ggbeeswarm::geom_quasirandom(aes(abs(b)/e_sigma, reject, group = n), groupOnX = FALSE, alpha = 0.01) +
+  geom_smooth(aes(abs(b)/e_sigma, reject, col = n_eval), method = "glm", method.args = list(family = binomial()), se = FALSE) +
+  geom_line(data = bind_sim, aes(x, reject), col = "grey", linetype = 2, size = 1) +
+  facet_wrap(~n, labeller = label_both) +
+  scale_colour_viridis_d(option = "B", begin = 0.5, end = 0.9) +
+  theme_light()
+```
+
+```{r eval = FALSE}
+# cubic break down
+# We do not have enough observations
+survey_pvalue %>%
+  filter(type == "cubic") %>%
+  mutate(n = factor(n)) %>%
+  mutate(reject = as.numeric(reject)) %>%
+  ggplot() +
+  geom_point(aes(abs(a), reject, col = n), alpha = 0.1) +
+  geom_smooth(aes(abs(a), reject, col = n), method = "glm", method.args = list(family = binomial()), se = FALSE) +
+  geom_smooth(data = mutate(filter(sim_conv, type == "cubic"), n = factor(n), reject = as.numeric(reject)),
+              aes(abs(a), reject, group = n), 
+              method = "glm", method.args = list(family = binomial()), se = FALSE, col = "grey", linetype = 2) +
+  facet_grid(e_sigma~n)
+```
+
+```{r power-com-heter-b, fig.cap="\\label{fig:power-com-heter-b}power com heter b"}
+bind_grid <- expand.grid(n = c(50, 100, 300),
+                         h = 3,
+                         x = c(quote(b)),
+                         a = c(-1, 0, 1))
+
+bind_list <- list()
+
+for (i in 1:nrow(bind_grid)) {
+  (function(n, h, x, a) {
+    mutate(kernel_regression_heter(filter(sim_conv, n == {{n}}, a == {{a}}), h = h, x = x),
+           n = factor(n, levels = c(50, 100, 300)),
+           a = factor(a, levels = c(-1, 0, 1)))
+    })(n = bind_grid$n[i], h = bind_grid$h[i], x = bind_grid$x[[i]], a = bind_grid$a[i]) -> bind_list[[i]]
+}
+
+bind_sim <- reduce(bind_list, bind_rows)
+
+# heter break down
+survey_pvalue %>%
+  filter(type != "cubic") %>%
+  mutate(n = factor(n, levels = c(50, 100, 300))) %>%
+  mutate(reject = as.numeric(reject)) %>%
+  mutate(n_eval = factor(n_eval)) %>%
+  ggplot() +
+  ggbeeswarm::geom_quasirandom(aes(b, reject, group = n), groupOnX = FALSE, alpha = 0.01) +
+  geom_smooth(aes(b, reject, col = n_eval), method = "glm", method.args = list(family = binomial()), se = FALSE) +
+  geom_line(data = bind_sim, aes(x, reject), col = "grey", linetype = 2, size = 1) +
+  theme_light() +
+  scale_colour_viridis_d(option = "B", begin = 0.5, end = 0.9) +
+  facet_grid(a~n, labeller = label_both)
+```

### 2022-05-12 paper_comparison/paper_comparison/paper_comparison.tex
break down factors in power comparison

@@
 
-It was expected that with larger effect size, both conventional test and
-visual test will have higher probability in rejecting the null
-hypothesis when it is not true. To explore the difference in power
-between conventional test and visual test, ten logistic regression
-models are fit to the data, with natural logarithm of the effect size as
-the only fixed effect, and whether the test successfully rejects the
-null hypothesis as the response variable. Table \ref{tab:powerglmheter}
-and \ref{tab:powerglmheter} summarizes the results. The natural
-logarithm of the effect size is positive and significant across all
-models. Figure \ref{fig:power-com} illustrates the fitted models, while
+To model the power of visual test, 10 logistic regression were fit for
+different number of evaluations ranged from one to five and two
+different types of simulation setting. All 10 models used natural
+logarithm of the effect size as the only fixed effect, and whether the
+test successfully rejects the null hypothesis as the response variable.
+Given the way we define the effect size, it was expected that with
+larger effect size, both conventional test and visual test will have
+higher probability in rejecting the null hypothesis when it is not true.
+The modelling result summarized in \ref{tab:powerglmcubic} and
+\ref{tab:powerglmheter} aligned with the expectation as the coefficients
+of natural logarithm of the effect size are positive and significant
+across all 10 models.
+
+Figure \ref{fig:power-com} illustrates the fitted models, while
 providing the local constant estimate of the power of F-test and
 Breusch--Pagan test for comparison. Data for the conventional test is
-simulated under the same model setting used in two experiments and
-500000 samples are drawn for both cubic and heteroskedasticity model.
-
+simulated under the model setting described in section \ldots{} and
+5000000 samples are drawn for both cubic and heteroskedasticity model.
 From Figure \ref{fig:power-com}, it can be observed that the fitted
@@
 For heteroskedasticity model, this phenomenon was more obvious as the
-curves of visual tests with evaluations greater than two were always
-above the curves of visual test with evaluations smaller than two.
-However, this only held for large enough effect size. For small effect
-size, visual tests with fewer evaluations might have greater power. Note
-that, the expected power of visual test derived by
-\citet{majumder_validation_2013} followed by the assumption that human
-has the ability to select the plot with the highest t-statistic from a
-lineup under the classical linear model regression setting showed
-similar properties, where visual tests with fewer evaluations were
-expected to perform better when the parameter values close to the null
-hypothesis.
+power of visual tests with evaluations greater than two were always
+greater than those with evaluations smaller than two.
 
-For cubic model, the separation between curves was small. Fitted power
-of visual test with three to five evaluations were almost identical to
-each other in regards of effect size. In addition, all five curves
-peaked at one as effect size increased, suggesting that identification
-of non-linearity as a visual task can be completed reliably by human
-when the effect size is large enough.
+For cubic model, the separation between curves was small. The estimated
+power of visual tests with three to five evaluations were almost
+identical to each other in regards of effect size. In addition, all five
+curves peaked at one as effect size increased, suggesting that
+identification of non-linearity as a visual task can be completed
+reliably by human as long as the departure from null hypothesis is large
+enough.
 
-As shown in Figure @ref(fig:power-com), both F-test and Breusch--Pagan
-test generally possessed greater power than visual test. It was also
-found that there was a huge gap between the conventional test curves and
-the visual test curves at around \(log(\text{effect size}) = 0\) for the
-cubic model and \(log(\text{effect size}) = 2.5\) for the
-heteroskedasticity model.
+As shown in Figure \ref{fig:power-com}, both F-test and Breusch--Pagan
+test generally possessed greater power than visual test. A visual tests
+is a collection of test against any alternatives that would create
+visual discoverable features, while a conventional test is usually
+targeting at a pre-specified alternative. Considering the data
+generating process of the model defect was known and controlled in this
+research, where all other alternatives have been eliminated except the
+one we concerned, the result was suggested that conventional tests were
+more sensitive to violations of linearity and homoscedasticity
+assumption than visual tests.
 
-We further analysed the lineup with the corresponding effect size.
+It was also found that there was a noticeable gap between curves of the
+conventional test and the visual test at around
+\(log(\text{effect size}) = 0\) for the cubic model and
+\(log(\text{effect size}) = 2.5\) for the heteroskedasticity model,
+where the differences in power were greater than 0.6. We further
+analysed the lineups with correspoding effect sizes. Figure
+\ref{fig:cubic-hard} and \ref{fig:heter-hard} showed that human was
+indeed hard to identify the patterns at this level of difficulty. The
+visual difference between the true data plot and null plots were almost
+unnoticeable.
+
+\hypertarget{effect-of-parameters-on-power-of-the-visual-test}{%
+\subsection{Effect of parameters on power of the visual
+test}\label{effect-of-parameters-on-power-of-the-visual-test}}
+
+The previous section focuses on the change of effect size relative to
+the power of the visual test.
 
@@
 
+\begin{figure}
+\centering
+\includegraphics{paper_comparison_files/figure-latex/cubic-hard-1.pdf}
+\caption{\label{fig:cubic-hard}cubic-hard ans 11}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{paper_comparison_files/figure-latex/heter-hard-1.pdf}
+\caption{\label{fig:heter-hard}heter-hard ans 2}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{paper_comparison_files/figure-latex/power-com-cubic-a-1.pdf}
+\caption{\label{fig:power-com-cubic-a}power com cubic a}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{paper_comparison_files/figure-latex/power-com-cubic-b-1.pdf}
+\caption{\label{fig:power-com-cubic-b}power com cubic b}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{paper_comparison_files/figure-latex/power-com-heter-b-1.pdf}
+\caption{\label{fig:power-com-heter-b}power com heter b}
+\end{figure}
+
 \bibliographystyle{tfcad}

### 2022-05-12 paper_comparison/scripts/process.R

@@ -337,7 +337,7 @@ preprocessing$conv_test$conv_test_cubic <- function() {
   # Generate data from the model
   dat <- mod$gen(para_list1$n, test = TRUE)
 
-  c(mod$effect_size(dat), dat$p_value[1])
+  return(list(effect_size = mod$effect_size(dat), p_value = mod$test(dat)$p_value, para_list = para_list1))
 }
 
 
@@ -378,7 +378,7 @@ preprocessing$conv_test$conv_test_heter <- function() {
   # Generate data from the model
   dat <- mod$gen(para_list1$n, test = TRUE)
 
-  c(mod$effect_size(dat), dat$p_value[1], para_list1$b)
+  return(list(effect_size = mod$effect_size(dat), p_value = mod$test(dat)$p_value, para_list = para_list1))
 }
 
 
@@ -391,21 +391,24 @@ preprocessing$conv_test$get_sim_conv <- function(sim_cubic_n = 5000, sim_heter_n
 
   conv_result_cubic <- map(1:sim_cubic_n, ~conv_test_cubic())
 
-  conv_result_cubic <- data.frame(effect_size = map_dbl(conv_result_cubic, ~.x[1]),
-                                  p_value = map_dbl(conv_result_cubic, ~.x[2])) %>%
+  conv_result_cubic <- tibble(effect_size = map_dbl(conv_result_cubic, ~.x$effect_size),
+                              p_value = map_dbl(conv_result_cubic, ~.x$p_value),
+                              para_list = map(conv_result_cubic, ~.x$para_list)) %>%
     mutate(reject = p_value < alpha) %>%
-    mutate(type = "cubic")
+    mutate(type = "cubic") %>%
+    unnest_wider(para_list)
 
   # conv_result_cubic <- conv_result_cubic %>%
   #   filter(log(effect_size) > -5)
 
   conv_result_heter <- map(1:sim_heter_n, ~conv_test_heter())
 
-  conv_result_heter <- data.frame(effect_size = map_dbl(conv_result_heter, ~.x[1]),
-                                  p_value = map_dbl(conv_result_heter, ~.x[2]),
-                                  b = map_dbl(conv_result_heter, ~.x[3])) %>%
-    mutate(reject = p_value < 0.05) %>%
-    mutate(type = "heteroskedasticity")
+  conv_result_heter <- tibble(effect_size = map_dbl(conv_result_heter, ~.x$effect_size),
[…]

### 2022-05-24 paper_comparison/explore/check_patterns.Rmd
explore patterns

---
title: "pattern_explore"
output: 
  html_document:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(visage)
library(tidyverse)
```

```{r fig.height=34, fig.width=45, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 0.25)
ori_dat <- mod$gen(1000, computed = list(e = rep(0, 1000)))
i <- 0
plot_list <- list()

for (a in seq(-1.5, 1.5, 0.25)) {
  for (b in seq(-1.5, 1.5, 0.25)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(1000, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE)} +
      ggtitle(glue::glue("a:{a}, b:{b}, a/|b|:{a/abs(b)}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 13)
```

There are 8 basic shapes, (1.5, 1.5), (1.5, -1.5), (1.5, 0), (-1.5, 0), (0, 1.5), (0, -1.5), (-1.5, 1.5), (-1.5, -1.5). Others are transitions between these five shapes.

```{r}
plot_list <- list()

mod$set_prm(c("a", "b"), c(-1.5, -1.5)) %>%
  {.$gen(1000, computed = select(ori_dat, -y))} %>%
  {mod$plot(.)} +
  ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}")) -> plot_list[[1]]

mod$set_prm(c("a", "b"), c(-1.5, 0)) %>%
  {.$gen(1000, computed = select(ori_dat, -y))} %>%
  {mod$plot(.)} +
  ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}")) -> plot_list[[2]]

mod$set_prm(c("a", "b"), c(-1.5, 1.5)) %>%
  {.$gen(1000, computed = select(ori_dat, -y))} %>%
  {mod$plot(.)} +
  ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}")) -> plot_list[[3]]

mod$set_prm(c("a", "b"), c(0, -1.5)) %>%
  {.$gen(1000, computed = select(ori_dat, -y))} %>%
  {mod$plot(.)} +
  ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}")) -> plot_list[[4]]

mod$set_prm(c("a", "b"), c(0, 1.5)) %>%
  {.$gen(1000, computed = select(ori_dat, -y))} %>%
  {mod$plot(.)} +
  ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}")) -> plot_list[[6]]

mod$set_prm(c("a", "b"), c(1.5, -1.5)) %>%
  {.$gen(1000, computed = select(ori_dat, -y))} %>%
  {mod$plot(.)} +
  ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}")) -> plot_list[[7]]

mod$set_prm(c("a", "b"), c(1.5, 0)) %>%
  {.$gen(1000, computed = select(ori_dat, -y))} %>%
  {mod$plot(.)} +
  ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}")) -> plot_list[[8]]

mod$set_prm(c("a", "b"), c(1.5, 1.5)) %>%
  {.$gen(1000, computed = select(ori_dat, -y))} %>%
  {mod$plot(.)} +
  ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}")) -> plot_list[[9]]

plot_list[[5]] <- patchwork::plot_spacer()

patchwork::wrap_plots(plot_list, ncol = 3)
```

These five are based on three patterns. Others are products of 180 degrees rotation or flip.

```{r fig.height=24, fig.width=15}
mod <- quartic_model(x = rand_uniform(-1, 1), sigma = 0.25)
ori_dat <- mod$gen(1000)
i <- 0
plot_list <- list()

for (c in seq(-20, 20, 5)) {
  i <- i + 1
  mod$set_prm(c("a", "b", "c"), c(5, 0, c)) %>%
  {.$gen(1000, computed = select(ori_dat, -y))} %>%
  {mod$plot(.)} +
  ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}, c:{mod$prm$c}")) -> plot_list[[i]]

  i <- i + 1
  mod$set_prm(c("a", "b"), c(-5, 0)) %>%
    {.$gen(1000, computed = select(ori_dat, -y))} %>%
    {mod$plot(.)} +
    ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}, c:{mod$prm$c}")) -> plot_list[[i]]

  i <- i + 1
  mod$set_prm(c("a", "b"), c(-5, -5)) %>%
    {.$gen(1000, computed = select(ori_dat, -y))} %>%
    {mod$plot(.)} +
    ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}, c:{mod$prm$c}")) -> plot_list[[i]]

  i <- i + 1
  mod$set_prm(c("a", "b"), c(5, 5)) %>%
    {.$gen(1000, computed = select(ori_dat, -y))} %>%
    {mod$plot(.)} +
    ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}, c:{mod$prm$c}")) -> plot_list[[i]]

  i <- i + 1
  mod$set_prm(c("a", "b"), c(0, 5)) %>%
    {.$gen(1000, computed = select(ori_dat, -y))} %>%
    {mod$plot(.)} +
    ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}, c:{mod$prm$c}")) -> plot_list[[i]]
}

patchwork::wrap_plots(plot_list, ncol = 5)
```

The change of c brings two more shapes. (5,0,-5) and (-5,0,5). Other new shapes are quite similar to the previous. I don't think they are fundamentally different.

```{r}
mod$set_prm(c("a", "b", "c"), c(5, 0, -5)) %>%
  {.$gen(1000, computed = select(ori_dat, -y))} %>%
  {mod$plot(.)} +
  ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}, c:{mod$prm$c}"))
```

```{r}
mod$set_prm(c("a", "b", "c"), c(-5, 0, 5)) %>%
  {.$gen(1000, computed = select(ori_dat, -y))} %>%
  {mod$plot(.)} +
  ggtitle(glue::glue("a:{mod$prm$a}, b:{mod$prm$b}, c:{mod$prm$c}"))
```

Number count:

Consider using the simple cubic linear model.

$$y = 1 + x + ax^2 + bx^3$$

### 2022-05-24 paper_comparison/paper_comparison/paper_comparison.Rmd
explore patterns

@@
 
+<!-- sample 1, change values of a and b, plot the data  -->
+
 ## Effect of parameters on power of the visual test
@@
 The previous section focuses on the change of effect size relative to the power of the visual test.
+However, effect size is only a one dimensional summarisation of parameters used in data simulation.
+Individual factor embedded in the simulation process should also be analysed.
+
+In cubic model, two major factors that influencing the strength of the signal are $a$ and $b$.
+Figure \ref{fig:power-com-cubic-a} and \ref{fig:power-com-cubic-a} illustrates 30 different logistic regressions fit for different number of evaluations and different number of observations $n$.
+The regressor used in these models was $|a|/\sigma$ since the noise level $\sigma$ needed to be taken into account.
+From the figures, we can observe ...
 
@@
   filter(row_number() == 1) %>%
-  ungroup()
+  ungroup() %>%
+  left_join(survey %>%
+              count(exp_lineup_id) %>%
+              rename(n_eval = n))
 
@@
 save_lineup_env(env)
+```
+
+```{r}
+survey_single_pvalue <- survey %>%
+  calc_p_value_multi(lineup_id = "exp_lineup_id", 
+                     detected = "detect", 
+                     n_sel = "num_selection",
+                     comb = FALSE,
+                     n_sim = 100000,
+                     cache_env = env) %>%
+  left_join(lineup_info, by = c("lineup_id" = "exp_lineup_id")) %>%
+  mutate(reject = eval_p_value(p_value, tol = 1e-02))
+
+save_lineup_env(env)
+```
+
+```{r fig.height=6, fig.width=8}
+# set.seed(20002)
+# mod <- cubic_model(a = -3, b = 0, c = 1, sigma = 0.5, 
+#                    x = rand_uniform(-1, 1), z = rand_uniform(-1, 1))
+
+# fixed_sample <- mod$gen(100)
+plot_list <- list()
+
+for (a in seq(-3, 3, 1)) {
+  for (b in seq(-3, 3, 1)) {
+
+    set.seed(20002)
+
+    mod <- cubic_model(a = a, b = b, c = 0, sigma = 0.5, 
+                   x = rand_uniform(-1, 1), z = rand_uniform(-1, 1))
+
+    plot_list[[paste(a, b, sep = "_")]] <- mod$gen(1000, fit_model = TRUE) %>%
+    mod$plot("resid", remove_grid_line = TRUE, theme = theme_light(), remove_axis = TRUE) +
+    ggtitle(glue::glue("a: {mod$prm$y$sym$a}, b: {mod$prm$y$sym$b}")) +
+    ylim(c(-20, 20))
+  }
+}
+
+patchwork::wrap_plots(plot_list, ncol = 7)
+```
+
+```{r fig.height=1.5, fig.width=8}
+set.seed(20002)
+mod <- cubic_model(a = -3, b = -3, c = 1, sigma = 0.5, 
+                   x = rand_uniform(-1, 1), z = rand_uniform(-1, 1))
+
+fixed_sample <- mod$gen(100)
+plot_list <- list()
+
+for (c in seq(0, 2, 0.5)) {
+    mod$prm$y$sym$c <- c
+    plot_list[[paste(c)]] <- mod$gen(100, fit_model = TRUE, 
+                                     computed = list(e = fixed_sample$e,
+                                                     x = fixed_sample$x,
+                                                     z = fixed_sample$z)) %>%
+    mod$plot("resid", remove_grid_line = TRUE, theme = theme_light(), remove_axis = FALSE) +
+    ggtitle(glue::glue("c: {mod$prm$y$sym$c}")) 
+}
+
+patchwork::wrap_plots(plot_list, ncol = 5)
+```
+
+```{r}
+plot_list <- list()
+
+for (b in c(0, 0.5, 1, 1.5, 2, 2.5, 3, 10)) {
+
+    set.seed(20002)
+    mod <- heter_model(a = 1, b = b, x = rand_uniform(-1, 1))
+
+    plot_list[[paste(b)]] <- mod$gen(1000, fit_model = TRUE) %>%
+    mod$plot("resid", remove_grid_line = TRUE, theme = theme_light(), remove_axis = TRUE) +
+    ggtitle(glue::glue("b: {b}"))
+}
+
+patchwork::wrap_plots(plot_list, ncol = 5)
+```
+
+```{r}
+survey_single_pvalue %>%
+  mutate(reject = as.numeric(reject)) %>%
+  filter(n_eval >= 1) %>%
+  ggplot() +
+  geom_histogram(aes(log_effect_size, ..count../10), binwidth = 0.1) +
+  geom_point(aes(log_effect_size, reject), alpha = 0.2) +
+  geom_line(data = kernel_regression_cubic(sim_conv, 0.4), aes(x, reject), linetype = 2) +
+  geom_line(data = kernel_regression_heter(sim_conv, 0.75), aes(x, reject), linetype = 2) +
+  geom_smooth(aes(log_effect_size, reject), method = "glm", se = FALSE, method.args = list(family = binomial())) +
+  facet_wrap(~type, scales = "free_x")
+```
+
+```{r}
+sim_conv <- sim_conv %>%
+  mutate(dis_b = cut(abs(b), breaks = c(0, 1, 2, 3)))
+
+kernel_layer <- function(dis_b_index) {
+  geom_line(data = kernel_regression_cubic(filter(sim_conv, as.numeric(dis_b) == dis_b_index), 
+                                           0.4, 
+                                           x = quote(abs(a))), 
+            aes(x, reject, col = levels(sim_conv$dis_b)[dis_b_index]), 
+            linetype = 2)
+}
+
+survey_single_pvalue %>%
+  filter(type == "cubic") %>%
+  mutate(reject = as.numeric(reject)) %>%
+  mutate(dis_b = cut(abs(b), breaks = c(0, 1, 2, 3))) %>%
+  filter(n_eval >= 1) %>%
+  ggplot() +
+  # geom_histogram(aes(log_effect_size, ..count../10), binwidth = 0.1) +
+  geom_point(aes(abs(a), reject), alpha = 0.2) +
+  kernel_layer(1) +
+  kernel_layer(2) +
+  kernel_layer(3) +
+  geom_smooth(aes(abs(a), reject, col = dis_b), method = "glm", se = FALSE, method.args = list(family = binomial())) +
+  facet_grid(n~e_sigma, scales = "free_x") +
+  labs(col = "abs(b)")
+```
+
+```{r}
+
 ```

### 2022-06-01 Third_experiment/exp_setup.R
allocate third exp

@@ -1,0 +1,278 @@
+library(ggExtra)
+library(tidyverse)
+library(visage)
+library(lhs)
+library(cli)
+
+get_trunc_norm_cdf <- function(sd = 1, ndraws = 10^7) {
+  result <- rnorm(ndraws, mean = 0, sd = sd)
+  result <- result[result >= -1.5 & result <= 1.5]
+  function(p) {
+    quantile(result, probs = p) %>%
+      unname()
+  }
+}
+
+trans_unif_2_norm <- function(x, sd = 1, ndraws = 10^7) {
+  p_trunc_norm <- get_trunc_norm_cdf(sd, ndraws)
+  p_trunc_norm(x)
+}
+
+trans_unif_2_discrete <- function(x, level = NULL) {
+  level[trunc(x * length(level)) + 1]
+}
+
+
+
+# Shape of the trunc density
+ggplot() +
+  geom_density(aes(trans_unif_2_norm(seq(0, 1, 0.00001))), col = "red")
+
+
+# V1: a in -1.5 ~ 1.5
+# V2: b in -1.5 ~ 1.5
+# V3: N: 50, 100, 300
+# V4: x_dist: uniform, normal, lognormal, neglognormal, even_dis, uneven_dis
+# V5: e_sigma: 0.125, 0.25, 0.5, 1
+
+# 9 (a and b combinations) * 3 * 6 * 4 = 720
+
+# rep: 5
+
+set.seed(10086)
+
+num_ab_comb <- 9
+
+# Generate unique lineups
+lineup <- randomLHS(num_ab_comb * 3 * 6 * 4, 2) %>%
+  as.data.frame() %>%
+  mutate(a = trans_unif_2_norm(V1, sd = 0.5),
+         b = trans_unif_2_norm(V2, sd = 0.5)) %>%
+  bind_cols(map(1:num_ab_comb, function(x) {
+    expand.grid(n = c(50, 100, 300),
+                x_dist = c("uniform", "normal", "lognormal", "neglognormal",
+                           "even_discrete", "uneven_discrete"),
+                e_sigma = c(0.125, 0.25, 0.5, 1),
+                ab = x)
+    }) %>%
+    reduce(bind_rows)) %>%
+  select(-V1, -V2, -ab) %>%
+  mutate(lineup_id = 1:n())
+
+# Check a and b distribution
+lineup %>%
+  ggplot() +
+  geom_point(aes(a, b)) +
+  facet_grid(x_dist~e_sigma+n)
+
[…]

### 2022-06-01 Third_experiment/setup.R

@@ -1,0 +1,157 @@
+# a in -1.5 ~ 1.5
+# b in -1.5 ~ 1.5
+# x_dist = uniform, normal, lognormal, neglognormal, even_dis, uneven_dis
+# N = 50, 100, 300
+# rep = 2
+# e_sigma = 0.125, 0.25, 0.5, 1
+
+# N_ab * 6 * 3 * 2 * 4
+
+library(tidyverse)
+library(visage)
+library(lhs)
+
+get_trunc_norm_cdf <- function(sd = 1, ndraws = 10^7) {
+  result <- rnorm(ndraws, mean = 0, sd = sd)
+  result <- result[result >= -1.5 & result <= 1.5]
+  function(p) {
+    quantile(result, probs = p) %>%
+      unname()
+  }
+}
+
+trans_unif_2_norm <- function(x, sd = 1, ndraws = 10^7) {
+  p_trunc_norm <- get_trunc_norm_cdf(sd, ndraws)
+  p_trunc_norm(x)
+}
+
+sim_effect_size <- function(a, b) {
+  x_dist <- sample(c("uniform", "normal", "lognormal", "neglognormal",
+                     "even_discrete", "uneven_discrete"), 1)
+  e_sigma <- sample(c(0.125, 0.25, 0.5, 1), 1)
+  n <- sample(c(50, 100, 300), 1)
+
+  x <- switch(x_dist,
+              uniform = rand_uniform(-1, 1),
+              normal = rand_normal(sigma = 0.3),
+              lognormal = rand_lognormal(sigma = 0.6),
+              neglognormal = {rl <- rand_lognormal(sigma = 0.6); closed_form(~-rl)},
+              even_discrete = rand_uniform_d(k = 5, even = TRUE),
+              uneven_discrete = rand_uniform_d(k = 5, even = FALSE))
+
+  mod <- simple_cubic_model(a = a, b = b, x = x, sigma = e_sigma)
+  dat <- mod$gen(n, fit_model = FALSE)
+  tibble(a, b, x_dist, e_sigma, n, effect_size = mod$effect_size(dat))
+}
+
+
+result <- map(1:10000, ~sim_effect_size(1, 1)) %>%
+  reduce(bind_rows)
+
+result %>%
+  group_by(x_dist, e_sigma, n) %>%
[…]

### 2022-06-01 paper_comparison/explore/check_different_sigma.Rmd
upload explore

---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(visage)
library(tidyverse)
```

# N: 50

Sigma: 0.125

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 0.125)
ori_dat <- mod$gen(50)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(50, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 0.25

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 0.25)
ori_dat <- mod$gen(50)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(50, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 0.5

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 0.5)
ori_dat <- mod$gen(50)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(50, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 1

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 1)
ori_dat <- mod$gen(50)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(50, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

# N: 100

Sigma: 0.125

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 0.125)
ori_dat <- mod$gen(100)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(100, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 0.25

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 0.25)
ori_dat <- mod$gen(100)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(100, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 0.5

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 0.5)
ori_dat <- mod$gen(100)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(100, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 1

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 1)
ori_dat <- mod$gen(100)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(100, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

# N: 300

Sigma: 0.125

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 0.125)
ori_dat <- mod$gen(300)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(300, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 0.25

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 0.25)
ori_dat <- mod$gen(300)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(300, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 0.5

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 0.5)
ori_dat <- mod$gen(300)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(300, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

Sigma: 1

```{r fig.height=18, fig.width=24, cache = TRUE}
mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 1)
ori_dat <- mod$gen(300)
i <- 0
plot_list <- list()

for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
    i <- i + 1
    mod$set_prm("a", a) %>%
      {.$set_prm("b", b)} %>%
      {.$gen(300, computed = select(ori_dat, x, e))} %>%
      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
  }
}

patchwork::wrap_plots(plot_list, ncol = 7)
```

### 2022-06-01 experiment/scripts/evaluate_grid.R
remove unused folder

@@ -1,150 +1,0 @@
-source(here::here("scripts/model.R"))
-source(here::here("scripts/higher_order_model.R"))
-source(here::here("scripts/heteroskedasticity_model.R"))
-source(here::here("scripts/non_normal_model.R"))
-source(here::here("scripts/experiment.R"))
-
-num_rec <- 100000
-
-build_and_record <- function(...) {
-  paras <- list(...)
-  myexp <- EXPERIMENT("tmp")
-  mod <- do.call(paras$model, paras)
-  myexp$add_model("tmp_mod", mod)
-  myexp$gen_lineup_data("tmp_mod", k = 2, n = paras$n)
-  return(list(c(statistic = myexp$get_statistic(), pvalue = myexp$get_pvalue())))
-}
-
-data.frame(a = runif(num_rec, -3, 3),
-           b = runif(num_rec, -3, 3),
-           c = runif(num_rec, 0, 5),
-           x_dist = sample(c("uniform", "normal", "lognormal", "neglognormal"), num_rec, replace = TRUE),
-           x_mu = 0,
-           x_sigma = sample(c(0.3, 0.6), num_rec, replace = TRUE),
-           z_discrete = sample(c(TRUE, FALSE), num_rec, replace = TRUE),
-           z_discrete_dist = "discrete_uniform",
-           z_n = sample(seq(3, 10, 1), num_rec, replace = TRUE),
-           e_dist = "normal",
-           e_df = NA,
-           e_sigma = sample(c(0.25, 0.5, 1, 2), num_rec, replace = TRUE),
-           n = sample(c(50, 100, 300), num_rec, replace = TRUE),
-           model = "HIGHER_ORDER_MODEL") %>%
-  tibble() -> mygrid
-
-mygrid %>%
-  mutate(i = 1:n()/n()) %>%
-  rowwise() %>%
-  mutate(result = build_and_record(i=i,
-                                   a=a,
-                                   b=b,
-                                   c=c,
-                                   x_dist=x_dist,
-                                   x_mu=x_mu,
-                                   x_sigma=x_sigma,
-                                   z_discrete=z_discrete,
-                                   z_discrete_dist=z_discrete_dist,
-                                   z_n=z_n,
-                                   e_dist=e_dist,
-                                   e_df=e_df,
-                                   e_sigma=e_sigma,
-                                   n=n,
-                                   model=model)) %>%
-  ungroup() %>%
-  cbind(., do.call('rbind', .$result)) %>%
-  tibble() %>%
[…]

### 2022-06-01 experiment/scripts/experiment.R

@@ -1,466 +1,0 @@
-EXPERIMENT <- function(name) {
-
-  experiment <- new.env()
-  experiment$name <- name
-  experiment$num_lineup <- 0
-  experiment$model_list <- list()
-  experiment$dat <- list()
-
-  experiment$model <- function(name) {
-
-    # Method of getting a model.
-    #
-    # Available arguments:
-    #
-    # name: the name of the model
-
-    experiment$model_list[[name]]
-  }
-
-  experiment$summary_lineup <- function(lineup_id = 1, model_spec = FALSE) {
-
-    # Method of printing summary of a lineup
-    #
-    # Available arguments:
-    #
-    # lineup_id: id of lineup
-    # model_spec: whether or not to print the model spec
-
-    if (length(experiment$dat) < lineup_id) {
-      cli::cli_alert_warning("lineup {lineup_id} does not exist!")
-      return(invisible(NULL))
-    }
-    cli::cli_h1("Lineup {lineup_id}")
-    cli::cli_li("k = {experiment$dat[[lineup_id]]$metadata$k}")
-    cli::cli_li("n = {experiment$dat[[lineup_id]]$metadata$n}")
-    cli::cli_li("effect size = {experiment$dat[[lineup_id]]$metadata$effect_size}")
-    cli::cli_li("answer = {experiment$dat[[lineup_id]]$metadata$ans}")
-    if (model_spec) {
-      print(experiment$model(experiment$dat[[lineup_id]]$metadata$name))
-    }
-
-    return(invisible(NULL))
-  }
-
-  experiment$summary_model <- function(name, lineup = FALSE) {
-
-    # Method of printing summary of a model
-    #
-    # Available arguments:
-    #
-    # name: name of the model
-    # lineup: whether or not to print the lineups
-
-    print(experiment$model(name))
-    if (lineup) {
-      lineup_ids <- map(experiment$dat, ~.$metadata$name == name) %>%
-        unlist() %>%
-        which()
-      for (lineup_id in lineup_ids) {
-        experiment$summary_lineup(lineup_id)
-      }
-    }
-
-    return(invisible(NULL))
-  }
-
-  experiment$summary <- function(input = NULL) {
-
-
-    if (!is.null(input) & is.numeric(input)) {
-      experiment$summary_lineup(input, model_spec = TRUE)
-      return(invisible(NULL))
-    }
-
-    if (!is.null(name) & is.character(input)) {
-      experiment$summary_model(input, lineup = TRUE)
-      return(invisible(NULL))
-    }
-
[…]

### 2022-06-01 experiment/scripts/heteroskedasticity_model.R

@@ -1,59 +1,0 @@
-HETEROSKEDASTICITY_MODEL <- function(a, b,
-                                     x_dist = "uniform", z_discrete = FALSE,
-                                     ...) {
-
-  # Init a new model
-  model <- MODEL_SPEC(type = "heteroskedasticity",
-                      formula = "y ~ 1 + x + e, e ~ N(0, 1 + (2-|a|) * (x_i - a)^2 * b)",
-                      a = a, b = b,
-                      x_dist = x_dist, z_discrete = z_discrete,
-                      ...)
-
-  # Method of generating e
-  model$gen_e <- function(x, n) {
-
-    sigma2_i <- 1 +
-      (2 - abs(model$parameters$a)) * (x[,1] - model$parameters$a)^2 * model$parameters$b
-
-    sigma_i <- sqrt(sigma2_i)
-
-    result <- data.frame(single_sigma = sigma_i) %>%
-      rowwise() %>%
-      mutate(e = rnorm(1, sd = single_sigma))
-
-    return(result$e)
-  }
-
-  # Method of generating y
-  model$gen_y <- function(n) {
-    x <- model$gen_x(n)
-    e <- model$gen_e(x, n)
-
-    y <- 1 + x[,1] + e
-    return(data.frame(x = x[,1], z = x[,2], y = y))
-  }
-
-  model$fit <- function(dat) {
-    lm(y ~ x, data = dat)
-  }
-
-  # Method of evaluating the data using numerical test
-  model$evaluate <- function(dat) {
-
-    mod <- lm(y ~ x, data = dat)
-
-    tmp_data <- data.frame(x = dat$x)
-    tmp_data$xs <- tmp_data$x^2
-
-    this_bp_test <- lmtest::bptest(mod, varformula = ~ x + xs, data = tmp_data)
-
-    return(list("bptest", unname(this_bp_test$statistic), unname(this_bp_test$p.value)))
-  }
-
-  model$effect_size <- function(dat) {
-
-    return(model$parameters$b * sqrt(nrow(dat)))
-  }
-
-  return(model)
-}

### 2022-06-01 experiment/scripts/higher_order_model.R

@@ -1,60 +1,0 @@
-HIGHER_ORDER_MODEL <- function(a, b, c = 1,
-                               x_dist = "uniform", z_discrete = FALSE,
-                               e_dist = "normal", e_sigma = 1, ...) {
-
-  # Init a new model
-  model <- MODEL_SPEC(type = "higher order",
-                      formula = "y ~ 1 + (2-c) * x + c * z + a * [(2-c) * x]^2 + a * (c * z)^2 + b * [(2-c) * x]^3 + b * (c * z)^3 + e, e ~ N(0, sigma^2)",
-                      a = a, b = b, c = c,
-                      x_dist = x_dist, z_discrete = z_discrete,
-                      e_dist = e_dist, e_sigma = e_sigma, ...)
-
-  # Method of generating y
-  model$gen_y <- function(n) {
-    x <- model$gen_x(n)
-
-    y <- (2 - model$parameters$c) * x[,1] + model$parameters$c * x[,2] +
-      model$parameters$a * ((2 - model$parameters$c) * x[,1])^2 + model$parameters$a * (model$parameters$c * x[,2])^2 +
-      model$parameters$b * ((2 - model$parameters$c) * x[,1])^3 + model$parameters$b * (model$parameters$c * x[,2])^3 +
-      model$gen_e(n) + 1
-
-    return(data.frame(x = x[,1], z = x[,2], y))
-
-  }
-
-  model$fit <- function(dat) {
-    lm(y ~ x + z, data = dat)
-  }
-
-  model$effect_size <- function(dat) {
-    n <- nrow(dat)
-    xa <- dat %>%
-      mutate(c = 1) %>%
-      select(c, x, z) %>%
-      as.matrix()
-    ra <- diag(n) - xa %*% solve(t(xa) %*% xa) %*% t(xa)
-    diag_ra <- diag(sqrt(diag(ra)))
-    xb <- dat %>%
-      mutate(x2=x^2,z2=z^2,x3=x^3,z3=z^3) %>%
-      select(x2, z2, x3, z3) %>%
-      as.matrix()
-    beta_b <- matrix(c(model$parameters$a,model$parameters$a,model$parameters$b,model$parameters$b))
-
-    return((1/n) *
-      (1/model$parameters$e_sigma^2) *
-      sum((diag(sqrt(diag(ra))) %*% xb %*% beta_b)^2))
-  }
-
[…]

### 2022-06-01 experiment/scripts/main.R

@@ -1,11 +1,0 @@
-source(here::here("scripts/model_collections.R"))
-source(here::here("scripts/experiment.R"))
-
-set.seed(10086)
-
-myexp <- EXPERIMENT(name = "visual inference")
-higher_order <- HIGHER_ORDER_MODEL(a = 0.5, b = 1, c = 0,
-                                   x_dist = "uniform",
-                                   x_discrete = FALSE,
-                                   e_dist = "normal", e_sigma = 1)
-myexp$add_model("higher order", higher_order)

### 2022-06-01 experiment/scripts/model.R

@@ -1,259 +1,0 @@
-library(tidyverse)
-
-MODEL_SPEC <- function(...) {
-
-  # The MODEL_SPEC CLASS constructor
-
-  # Crate a new environment and store parameters
-  model <- new.env()
-  model$parameters <- list(...)
-
-  model$gen_x <- function(n) {
-
-    # The default method of generating x.
-    #
-    # Available arguments:
-    #
-    # x_dist: the distribution of x, one of "uniform", "normal", "lognormal", "neglognormal",
-    #   "uniform": (-1, 1),
-    #   "normal": normal(x_mu, x_sigma),
-    #   "lognormal": lognormal(x_mu, x_sigma^2)/3 - 1,
-    #   "neglognormal": -lognormal(x_mu, x_sigma^2)/3 + 1,
-    #   "discrete_uniform": n possible values from [-1, 1],
-    #   "binomial": binomial(n, 0.5)/n*2 - 1.
-    #
-    # x_mu: mean of lognormal or normal.
-    # x_sigma: sigma of lognormal or normal.
-    # x_n: n of binomial or discrete_uniform.
-    # z_discrete: TRUE or FALSE. Whether or not to introduce an additional discrete z. Otherwise, z ~ U(-1,1).
-    #
-    # z_discrete_dist: the distribution of discrete z, one of "discrete_uniform" and "binomial",
-    #   "discrete_uniform": n possible values from [-1, 1],
-    #   "binomial": binomial(n, 0.5)/n*2 - 1.
-    #
-    # z_n: n of binomial or discrete_uniform.
-
-    x_dist <- model$parameters$x_dist
-    z_discrete <- model$parameters$z_discrete
-    z_discrete_dist <- model$parameters$z_discrete_dist
-    z_n <- model$parameters$z_n
-    x_mu <- model$parameters$x_mu
-    x_sigma <- model$parameters$x_sigma
-    x_n <- model$parameters$x_n
-
-    stand_dist <- function(n, func, ...) {
-      result <- func(n, ...)
-      result / max(result)
-    }
-
-    mylognormal <- function(n, meanlog, sdlog) {
-      rlnorm(n, meanlog, sdlog)/3 - 1
-    }
-
-    myneglognormal <- function(n, meanlog, sdlog) {
-      -mylognormal(n, meanlog, sdlog)
-    }
-
-    mybinom <- function(n, size, prob) {
-      rbinom(n, size, prob)/size*2-1
-    }
-
[…]

### 2022-06-01 experiment/scripts/model_collections.R

@@ -1,4 +1,0 @@
-source(here::here("scripts/model.R"))
-source(here::here("scripts/higher_order_model.R"))
-source(here::here("scripts/heteroskedasticity_model.R"))
-source(here::here("scripts/non_normal_model.R"))

### 2022-06-01 experiment/scripts/non_normal_model.R

@@ -1,33 +1,0 @@
-NON_NORMAL_MODEL <- function(c = 5,
-                             x_dist = "uniform", z_discrete = FALSE,
-                             e_dist = "t", ...) {
-
-  # Init a model
-  model <- MODEL_SPEC(type = "non-normal error",
-                      formula = "y ~ 1 + x + c * z + e, e ~ non-normal",
-                      c = c,
-                      x_dist = x_dist, z_discrete = z_discrete,
-                      e_dist = e_dist, ...)
-
-  # Method of generating y
-  model$gen_y <- function(n) {
-    x <- model$gen_x(n)
-    e <- model$gen_e(n)
-    y <- x[,1] + x[,2] * model$parameters$c + e + 1
-
-    return(data.frame(x = x[,1], z = x[,2], y = y))
-
-  }
-
-  # Method of evaluating the data using numerical test
-  model$evaluate <- function(dat) {
-
-    mod <- lm(y ~ x + z, data = dat)
-
-    this_sw_test <- shapiro.test(resid(mod))
-
-    return(list("swtest", unname(this_sw_test$statistic), unname(this_sw_test$p.value)))
-  }
-
-  return(model)
-}

### 2022-06-01 experiment/scripts/process_survey_big.R

@@ -1,164 +1,0 @@
-source(here::here("scripts/model.R"))
-
-survey_tools <- new.env()
-
-survey_tools$process_responses <- function(k, lineup_dat, lineup_ord) {
-
-  dat <- tibble(id = 1:length(unlist(lineup_ord)),
-                lineup_dat_id = unlist(lineup_ord)) %>%
-    mutate(set = (id-1) %/% k + 1) %>%
-    mutate(num = (id-1) %% k + 1)
-
-  user_info_dict <- list(c("18-24", "25-39", "40-54", "55-64", "65 or above"),
-                         c("High School or below", "Diploma and Bachelor Degree", "Honours Degree", "Masters Degree", "Doctoral Degree"),
-                         c("He", "She", "They", "Other"),
-                         c("Yes", "No"))
-
-  get_responses <- function(user_info_dict, uuid) {
-    tmp_survey <- jsonlite::fromJSON(here::here(glue::glue("study_1/study_1_big_check/survey/{uuid}.txt")),
-                                     simplifyVector = FALSE)
-    response_time <- map_dbl(tmp_survey, ~.x$rt)
-    user_information <- str_split(tmp_survey[[3]]$response[[1]], ",")[[1]]
-
-    user_information[2:5] <- imap_chr(user_information[-1], ~user_info_dict[[.y]][as.integer(.x)])
-    names(user_information) <- c("prolific_id", "age_group", "education", "pronoun", "previous_experience")
-
-
-    lineup_respone <- str_split(map_chr(tmp_survey[-c(1,2,3,4)], ~.x$response$response), ",")
-    # handle selections
-    selections <- map_chr(lineup_respone, ~.x[1])
-    # handle reasons
-    reasons_dict <- c("Outlier(s)", "Cluster(s)", "Shape")
-
-    reasons <- map_chr(lineup_respone, ~if(.x[2] %in% c("1", "2", "3")) {
-      reasons_dict[as.integer(.x[2])]
-    } else {
-      .x[2]
-    })
-
-    # handle confidence
-    confidence_dict <- c("Not at all", "Slightly", "Moderately", "Very", "Extremely")
-    confidence <- map_chr(lineup_respone, ~if(.x[3] %in% c("1", "2", "3", "4", "5")) {
-      confidence_dict[as.integer(.x[3])]
-    } else {
-      .x[3]
-    })
-
-    tibble(page = 1:length(response_time),
-           response_time = response_time,
-           set = uuid,
[…]

### 2022-06-06 paper_comparison/explore/check_patterns.Rmd
check sigma for poly model

@@
 patchwork::wrap_plots(plot_list, ncol = 13)
+```
+
+```{r fig.height=34, fig.width=45, cache = TRUE}
+mod <- simple_cubic_model(x = rand_uniform(-0.1, 0.1), sigma = 0.25)
+ori_dat <- mod$gen(1000, computed = list(e = rep(0, 1000)))
+i <- 0
+plot_list <- list()
+
+for (a in c(-1.5, -1, -0.1, -0.01, -0.001, -0.0001, 0, 0.0001, 0.001, 0.01, 0.1, 1, 1.5)) {
+  for (b in c(-1.5, -1, -0.1, -0.01, -0.001, -0.0001, 0, 0.0001, 0.001, 0.01, 0.1, 1, 1.5)) {
+    i <- i + 1
+    mod$set_prm("a", a) %>%
+      {.$set_prm("b", b)} %>%
+      {.$gen(1000, computed = select(ori_dat, x, e))} %>%
+      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE)} +
+      ggtitle(glue::glue("a:{a}, b:{b}, a/|b|:{a/abs(b)}")) -> plot_list[[i]]
+  }
+}
+
+patchwork::wrap_plots(plot_list, ncol = 13)
+```
+
+```{r}
+result <- tibble()
+
+for (a in seq(-1.5,1.5,0.25)) {
+  x <- rand_uniform(-1, 1)
+  e <- rand_normal(sigma = 0.25)
+  y <- closed_form(~x+a*x^2 + e)
+
+  dat <- y$gen(10000, rhs_val = TRUE, computed = list(e = 0)) %>%
+    y$as_dataframe("y")
+
+  mod <- lm(y ~ x, data = dat)
+
+  result <- bind_rows(result, tibble(.pred = predict(mod) - mean(predict(mod)),
+                                     .resid = resid(mod),
+                                     a = a))
+}
+
+result %>%
+  ggplot() +
+    geom_line(aes(.pred, .resid, col = factor(a))) +
+    theme(panel.grid.major = element_blank(),
+          panel.grid.minor = element_blank(),
+          axis.line = element_blank(),
+          axis.ticks = element_blank(),
+          # axis.text.x = element_blank(),
+          # axis.text.y = element_blank(),
+          axis.title.x = element_blank(),
+          axis.title.y = element_blank()) +
+  theme_bw() +
+  ggtitle(quote("centered fitted values vs. residuals"~of~y~"="~x), subtitle = quote(y~"="~x+ax^2+e)) +
+  xlab("fitted values") +
+  ylab("residuals")
+
+result %>%
+  ggplot() +
+    geom_line(aes(.pred, .resid)) +
+    theme(panel.grid.major = element_blank(),
+          panel.grid.minor = element_blank(),
+          axis.line = element_blank(),
+          axis.ticks = element_blank(),
+          # axis.text.x = element_blank(),
+          # axis.text.y = element_blank(),
+          axis.title.x = element_blank(),
+          axis.title.y = element_blank()) +
+  facet_wrap(~a, scales = "free")
 ```

### 2022-06-06 paper_comparison/explore/check_poly_model.Rmd
check sigma for poly model

---
title: "Untitled"
output: html_document
date: '2022-06-05'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(visage)
library(tidyverse)
```

# N: 50

Sigma: 0.125

```{r fig.height=6, fig.width=8, cache = TRUE}
mod <- poly_model(2, x = rand_uniform(-1, 1), sigma = 0.125)
ori_dat <- mod$gen(50)
i <- 0
plot_list <- list()

for (a in c(2, 3, 4, 5, 6)) {

  i <- i + 1
  mod$set_prm("a", a) %>%
    {.$gen(50, computed = select(ori_dat, x, e))} %>%
    {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
    ggtitle(glue::glue("a:{a}")) -> plot_list[[i]]

}

patchwork::wrap_plots(plot_list)
```

Sigma: 0.25

```{r fig.height=6, fig.width=8, cache = TRUE}
mod <- poly_model(2, x = rand_uniform(-1, 1), sigma = 0.25)
ori_dat <- mod$gen(50)
i <- 0
plot_list <- list()

for (a in c(2, 3, 4, 5, 6)) {

  i <- i + 1
  mod$set_prm("a", a) %>%
    {.$gen(50, computed = select(ori_dat, x, e))} %>%
    {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
    ggtitle(glue::glue("a:{a}")) -> plot_list[[i]]

}

patchwork::wrap_plots(plot_list)
```

Sigma: 0.5

```{r fig.height=6, fig.width=8, cache = TRUE}
mod <- poly_model(2, x = rand_uniform(-1, 1), sigma = 0.5)
ori_dat <- mod$gen(50)
i <- 0
plot_list <- list()

for (a in c(2, 3, 4, 5, 6)) {

  i <- i + 1
  mod$set_prm("a", a) %>%
    {.$gen(50, computed = select(ori_dat, x, e))} %>%
    {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
    ggtitle(glue::glue("a:{a}")) -> plot_list[[i]]

}

patchwork::wrap_plots(plot_list)
```

Sigma: 1

```{r fig.height=6, fig.width=8, cache = TRUE}
mod <- poly_model(2, x = rand_uniform(-1, 1), sigma = 1)
ori_dat <- mod$gen(50)
i <- 0
plot_list <- list()

for (a in c(2, 3, 4, 5, 6)) {

  i <- i + 1
  mod$set_prm("a", a) %>%
    {.$gen(50, computed = select(ori_dat, x, e))} %>%
    {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
    ggtitle(glue::glue("a:{a}")) -> plot_list[[i]]

}

patchwork::wrap_plots(plot_list)
```

# N: 100

Sigma: 0.125

```{r fig.height=6, fig.width=8, cache = TRUE}
mod <- poly_model(2, x = rand_uniform(-1, 1), sigma = 0.125)
ori_dat <- mod$gen(100)
i <- 0
plot_list <- list()

for (a in c(2, 3, 4, 5, 6)) {

  i <- i + 1
  mod$set_prm("a", a) %>%
    {.$gen(100, computed = select(ori_dat, x, e))} %>%
    {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
    ggtitle(glue::glue("a:{a}")) -> plot_list[[i]]

}

patchwork::wrap_plots(plot_list)
```

Sigma: 0.25

```{r fig.height=6, fig.width=8, cache = TRUE}
mod <- poly_model(2, x = rand_uniform(-1, 1), sigma = 0.25)
ori_dat <- mod$gen(100)
i <- 0
plot_list <- list()

for (a in c(2, 3, 4, 5, 6)) {

  i <- i + 1
  mod$set_prm("a", a) %>%
    {.$gen(100, computed = select(ori_dat, x, e))} %>%
    {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
    ggtitle(glue::glue("a:{a}")) -> plot_list[[i]]

}

patchwork::wrap_plots(plot_list)
```

Sigma: 0.5

```{r fig.height=6, fig.width=8, cache = TRUE}
mod <- poly_model(2, x = rand_uniform(-1, 1), sigma = 0.5)
ori_dat <- mod$gen(100)
i <- 0
plot_list <- list()

for (a in c(2, 3, 4, 5, 6)) {

  i <- i + 1
  mod$set_prm("a", a) %>%
    {.$gen(100, computed = select(ori_dat, x, e))} %>%
    {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
    ggtitle(glue::glue("a:{a}")) -> plot_list[[i]]

}

patchwork::wrap_plots(plot_list)
```

Sigma: 1

```{r fig.height=6, fig.width=8, cache = TRUE}
mod <- poly_model(2, x = rand_uniform(-1, 1), sigma = 1)
ori_dat <- mod$gen(100)
i <- 0
plot_list <- list()

for (a in c(2, 3, 4, 5, 6)) {

  i <- i + 1
  mod$set_prm("a", a) %>%
    {.$gen(100, computed = select(ori_dat, x, e))} %>%
    {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
    ggtitle(glue::glue("a:{a}")) -> plot_list[[i]]

}

patchwork::wrap_plots(plot_list)
```

# N: 300

Sigma: 0.125

```{r fig.height=6, fig.width=8, cache = TRUE}
mod <- poly_model(2, x = rand_uniform(-1, 1), sigma = 0.125)
ori_dat <- mod$gen(300)
i <- 0
plot_list <- list()

for (a in c(2, 3, 4, 5, 6)) {

  i <- i + 1
  mod$set_prm("a", a) %>%
    {.$gen(300, computed = select(ori_dat, x, e))} %>%
    {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
    ggtitle(glue::glue("a:{a}")) -> plot_list[[i]]

}

patchwork::wrap_plots(plot_list)
```

Sigma: 0.25

```{r fig.height=6, fig.width=8, cache = TRUE}
mod <- poly_model(2, x = rand_uniform(-1, 1), sigma = 0.25)
ori_dat <- mod$gen(300)
i <- 0
plot_list <- list()

for (a in c(2, 3, 4, 5, 6)) {

  i <- i + 1
  mod$set_prm("a", a) %>%
    {.$gen(300, computed = select(ori_dat, x, e))} %>%
    {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
    ggtitle(glue::glue("a:{a}")) -> plot_list[[i]]

}

patchwork::wrap_plots(plot_list)
```

Sigma: 0.5

```{r fig.height=6, fig.width=8, cache = TRUE}
mod <- poly_model(2, x = rand_uniform(-1, 1), sigma = 0.5)
ori_dat <- mod$gen(300)
i <- 0
plot_list <- list()

for (a in c(2, 3, 4, 5, 6)) {

  i <- i + 1
  mod$set_prm("a", a) %>%
    {.$gen(300, computed = select(ori_dat, x, e))} %>%
    {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
    ggtitle(glue::glue("a:{a}")) -> plot_list[[i]]

}

patchwork::wrap_plots(plot_list)
```

Sigma: 1

```{r fig.height=6, fig.width=8, cache = TRUE}
mod <- poly_model(2, x = rand_uniform(-1, 1), sigma = 1)
ori_dat <- mod$gen(300)
i <- 0
plot_list <- list()

for (a in c(2, 3, 4, 5, 6)) {

  i <- i + 1
  mod$set_prm("a", a) %>%
    {.$gen(300, computed = select(ori_dat, x, e))} %>%
    {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
    ggtitle(glue::glue("a:{a}")) -> plot_list[[i]]

}

patchwork::wrap_plots(plot_list)
```

### 2022-06-06 Third_experiment/check_allocation.R
allocate update

@@ -1,0 +1,15 @@
+result <- allocate_result %>%
+  left_join(lineup, by = c("lineup_id"))
+
+result %>%
+  filter(subject %in% 1:20) %>%
+  ggplot() +
+  geom_jitter(aes(factor(a),
+                 factor(e_sigma),
+                 shape = x_dist,
+                 col = factor(n)),
+             width = 0,
+             height = 0.2,
+             size = 3) +
+  facet_wrap(~subject) +
+  theme_bw()

### 2022-06-06 Third_experiment/exp_setup.R

@@ -4,18 +4,8 @@ library(visage)
 library(lhs)
 library(cli)
 
-get_trunc_norm_cdf <- function(sd = 1, ndraws = 10^7) {
-  result <- rnorm(ndraws, mean = 0, sd = sd)
-  result <- result[result >= -1.5 & result <= 1.5]
-  function(p) {
-    quantile(result, probs = p) %>%
-      unname()
-  }
-}
-
-trans_unif_2_norm <- function(x, sd = 1, ndraws = 10^7) {
-  p_trunc_norm <- get_trunc_norm_cdf(sd, ndraws)
-  p_trunc_norm(x)
+trans_unif_2_norm <- function(x, sd = 1) {
+  truncnorm::qtruncnorm(x, sd = sd, a = -1, b = 1)
 }
 
 trans_unif_2_discrete <- function(x, level = NULL) {
@@ -25,45 +15,56 @@ trans_unif_2_discrete <- function(x, level = NULL) {
 
 
 # Shape of the trunc density
-ggplot() +
-  geom_density(aes(trans_unif_2_norm(seq(0, 1, 0.00001))), col = "red")
+# ggplot() +
+#   geom_density(aes(trans_unif_2_norm(seq(0, 1, 0.00001))), col = "red")
 
 
-# V1: a in -1.5 ~ 1.5
-# V2: b in -1.5 ~ 1.5
+# V1: a = 1, 2, 3, 4, 5, 6
 # V3: N: 50, 100, 300
 # V4: x_dist: uniform, normal, lognormal, neglognormal, even_dis, uneven_dis
 # V5: e_sigma: 0.125, 0.25, 0.5, 1
 
-# 9 (a and b combinations) * 3 * 6 * 4 = 720
+# 6 * 3 * 6 * 4 = 432
 
 # rep: 5
 
 set.seed(10086)
 
-num_ab_comb <- 9
-
 # Generate unique lineups
-lineup <- randomLHS(num_ab_comb * 3 * 6 * 4, 2) %>%
-  as.data.frame() %>%
-  mutate(a = trans_unif_2_norm(V1, sd = 0.5),
-         b = trans_unif_2_norm(V2, sd = 0.5)) %>%
-  bind_cols(map(1:num_ab_comb, function(x) {
-    expand.grid(n = c(50, 100, 300),
-                x_dist = c("uniform", "normal", "lognormal", "neglognormal",
-                           "even_discrete", "uneven_discrete"),
[…]

### 2022-06-13 Third_experiment/check_allocation.R
checkout lineup plots, fix sigma

@@ -1,15 +1,165 @@
-result <- allocate_result %>%
-  left_join(lineup, by = c("lineup_id"))
+library(tidyverse)
+library(visage)
+library(UpSetR)
+
+result <- read_csv("allocate_result_full.csv")
+lineup <- read_csv("lineup_info.csv")
 
 result %>%
-  filter(subject %in% 1:20) %>%
+  # filter(subject %in% 1:120) %>%
   ggplot() +
-  geom_jitter(aes(factor(a),
+  geom_jitter(aes(factor(shape),
                  factor(e_sigma),
                  shape = x_dist,
                  col = factor(n)),
              width = 0,
-             height = 0.2,
+             height = 0,
+             alpha = 0.5,
              size = 3) +
-  facet_wrap(~subject) +
-  theme_bw()
+  facet_wrap(~subject, ncol = 15) +
+  scale_color_brewer(palette = "Dark2") +
+  theme_bw() -> p
+
+ggsave("check_allocation_for_subject.png", plot = p,
+       width = 20, height = 13)
+
+
+# x_dist & shape, upset
+cb_names <- result %>%
+  mutate(x_dist_shape = interaction(x_dist, shape, sep = "_")) %>%
+  {unique(.$x_dist_shape)}
+
+result %>%
+  mutate(x_dist_shape = interaction(x_dist, shape, sep = "_")) %>%
+  group_by(subject) %>%
+  count(x_dist_shape) %>%
+  group_by(x_dist_shape) %>%
+  group_map(~unique(.x$subject)) %>%
+  `names<-`(glue::glue("{cb_names}")) %>%
+  {upset(fromList(.), order.by = "freq", nsets = 16)}
+
+
+# x_dist & e_sigma, upset
+cb_names <- result %>%
+  mutate(x_dist_e_sigma = interaction(x_dist, e_sigma, sep = "_")) %>%
+  {unique(.$x_dist_e_sigma)}
+
+result %>%
+  mutate(x_dist_e_sigma = interaction(x_dist, e_sigma, sep = "_")) %>%
+  group_by(subject) %>%
+  count(x_dist_e_sigma) %>%
+  group_by(x_dist_e_sigma) %>%
+  group_map(~unique(.x$subject)) %>%
+  `names<-`(glue::glue("{cb_names}")) %>%
[…]

### 2022-06-13 Third_experiment/exp_setup.R

@@ -19,9 +19,9 @@ trans_unif_2_discrete <- function(x, level = NULL) {
 #   geom_density(aes(trans_unif_2_norm(seq(0, 1, 0.00001))), col = "red")
 
 
-# V1: a = 1, 2, 3, 4, 5, 6
+# V1: shape = 1, 2, 3, 4
 # V3: N: 50, 100, 300
-# V4: x_dist: uniform, normal, lognormal, neglognormal, even_dis, uneven_dis
+# V4: x_dist: uniform, normal, lognormal, even_dis
 # V5: e_sigma: 0.125, 0.25, 0.5, 1
 
 # 6 * 3 * 6 * 4 = 432
@@ -31,11 +31,11 @@ trans_unif_2_discrete <- function(x, level = NULL) {
 set.seed(10086)
 
 # Generate unique lineups
-lineup <- expand.grid(a = 1:6,
+lineup <- expand.grid(shape = 1:4,
                       n = c(50, 100, 300),
-                      x_dist = c("uniform", "normal", "lognormal", "neglognormal",
-                                 "even_discrete", "uneven_discrete"),
-                      e_sigma = c(0.125, 0.25, 0.5, 1)) %>%
+                      x_dist = c("uniform", "normal", "lognormal", "even_discrete"),
+                      e_sigma = c(0.5, 1, 2, 4),
+                      rep = 1:3) %>%
   mutate(lineup_id = sample(1:n()))
 
 
@@ -45,8 +45,9 @@ check_comb <- function(set_dat, new_dat, n = 1) {
         MARGIN = 2,
         function(this_comb) {
           # browser()
+          tmp <- unlist(new_dat[1, this_comb])
           for (i in 1:nrow(set_dat)) {
-            if (all(unlist(set_dat[i, this_comb]) == unlist(new_dat[1, this_comb]))) return(TRUE)
+            if (all(unlist(set_dat[i, this_comb]) == tmp)) return(TRUE)
           }
           return(FALSE)
         })
@@ -75,12 +76,15 @@ calc_dist <- function(lineup, allocated_lineup_ids, new_lineup_id) {
   allocated_lineups <- lineup %>%
     filter(lineup_id %in% allocated_lineup_ids)
 
+  tmp1 <- select(allocated_lineups, -rep, -lineup_id)
+  tmp2 <- select(new_lineup, -rep, -lineup_id)
+
   # We want these three options to be as different as possible
   15 -
[…]

### 2022-06-13 Third_experiment/generate_lineup.R
upload generate lineup scripts

@@ -1,0 +1,78 @@
+library(tidyverse)
+library(visage)
+
+sample_dat <- readRDS("data/all_data_big.RDS")
+
+lineup <- read_csv("lineup_info.csv")
+plot_string <- read_csv("data/plot_string.txt", col_names = FALSE) %>%
+  .$X1
+allocated <- read_csv("allocate_result.csv") %>%
+  group_by(subject) %>%
+  mutate(current_lineup = 1:n())
+dat <- vector(mode = "list", length = nrow(lineup))
+
+set.seed(10086)
+
+stand_dist <- function(x) {
+  (x - min(x))/max(x - min(x)) * 2 - 1
+}
+
+for (i in 1:nrow(lineup)) {
+
+  x <- switch(lineup$x_dist[i],
+              uniform = rand_uniform(-1, 1),
+              normal = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))},
+              lognormal = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))},
+              even_discrete = rand_uniform_d(k = 5, even = TRUE))
+  mod <- poly_model(lineup$shape[i], x = x, sigma = lineup$e_sigma[i])
+  lineup_dat <- mod$gen_lineup(lineup$n[i], 20)
+
+  dat[[i]]$metadata <- list(type = "polynomial",
+                            formula = deparse(mod$formula),
+                            shape = lineup$shape[i],
+                            x_dist = lineup$x_dist[i],
+                            include_z = TRUE,
+                            e_dist = "normal",
+                            e_sigma = lineup$e_sigma[i],
+                            name = glue::glue("poly_{i}"),
+                            k = 20,
+                            n = lineup$n[i],
+                            effect_size = mod$effect_size(filter(lineup_dat, null == FALSE)),
+                            ans = filter(lineup_dat, null == FALSE)$k[1]
+                            )
+
+  dat[[i]]$data <- lineup_dat
+}
+
+file_path <- "plots"
+
+for (j in 1:nrow(allocated)) {
+
+  ggsave(glue::glue("{file_path}/{plot_string[allocated$subject[j]]}_{allocated$subject[j]}_{allocated$current_lineup[j]}.png"),
+         plot = VI_MODEL$plot_lineup(dat[[allocated$lineup_id[j]]]$data,
+                                     remove_axis = TRUE,
+                                     remove_legend = TRUE,
[…]

### 2022-06-13 paper_comparison/scripts/process.R
fix the study 1 and 2 conventional test bug

@@ -318,12 +318,16 @@ preprocessing$conv_test$conv_test_cubic <- function() {
 
   if (para_list1$x_dist == "normal") {para_list1$x_sigma <- 0.3} else {para_list1$x_sigma <- 0.6}
 
+  stand_dist <- function(x) {
+    (x - min(x))/max(x - min(x)) * 2 - 1
+  }
+
   # Init X and Z based on the parameter list
-  x <- switch (para_list1$x_dist ,
-               "uniform" = rand_uniform(-1, 1),
-               "normal" = rand_normal(0, para_list1$x_sigma),
-               "lognormal" = rand_lognormal(0, para_list1$x_sigma),
-               "neglognormal" = {rl <- rand_lognormal(0, para_list1$x_sigma); closed_form(~-rl)}
+  x <- switch(para_list1$x_dist,
+              "uniform" = rand_uniform(-1, 1),
+              "normal" = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))},
+              "lognormal" = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))},
+              "neglognormal" = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(-raw_x/3 + 1))}
   )
 
   if (para_list1$z_discrete) {z <- rand_uniform(-1, 1)} else {z <- rand_uniform_d(-1, 1, para_list1$z_n)}
@@ -362,13 +366,18 @@ preprocessing$conv_test$conv_test_heter <- function() {
 
   if (para_list1$x_dist == "normal") {para_list1$x_sigma <- 0.3} else {para_list1$x_sigma <- 0.6}
 
+
+  stand_dist <- function(x) {
+    (x - min(x))/max(x - min(x)) * 2 - 1
+  }
+
   # Init X and Z based on the parameter list
-  x <- switch (para_list1$x_dist ,
+  x <-  switch(para_list1$x_dist,
                "uniform" = rand_uniform(-1, 1),
-               "discrete_uniform" = rand_uniform_d(-1, 1, para_list1$x_n),
-               "normal" = rand_normal(0, para_list1$x_sigma),
-               "lognormal" = rand_lognormal(0, para_list1$x_sigma),
-               "neglognormal" = {rl <- rand_lognormal(0, para_list1$x_sigma); closed_form(~-rl)}
[…]

### 2022-06-13 paper_comparison/explore/check_different_sigma.Rmd
upload check model

@@
 ```
+
+# New design
+
+# N: 50
+
+Sigma: 1
+
+```{r fig.height=18, fig.width=24, cache = TRUE}
+mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 1)
+ori_dat <- mod$gen(50)
+i <- 0
+plot_list <- list()
+
+for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
+  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
+    i <- i + 1
+    mod$set_prm("a", a) %>%
+      {.$set_prm("b", b)} %>%
+      {.$gen(50, computed = select(ori_dat, x, e))} %>%
+      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
+      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
+  }
+}
+
+patchwork::wrap_plots(plot_list, ncol = 7)
+```
+
+Sigma: 2
+
+```{r fig.height=18, fig.width=24, cache = TRUE}
+mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 2)
+ori_dat <- mod$gen(50)
+i <- 0
+plot_list <- list()
+
+for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
+  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
+    i <- i + 1
+    mod$set_prm("a", a) %>%
+      {.$set_prm("b", b)} %>%
+      {.$gen(50, computed = select(ori_dat, x, e))} %>%
+      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
+      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
+  }
+}
+
+patchwork::wrap_plots(plot_list, ncol = 7)
+```
+
+Sigma: 5
+
+```{r fig.height=18, fig.width=24, cache = TRUE}
+mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 5)
+ori_dat <- mod$gen(50)
+i <- 0
+plot_list <- list()
+
+for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
+  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
+    i <- i + 1
+    mod$set_prm("a", a) %>%
+      {.$set_prm("b", b)} %>%
+      {.$gen(50, computed = select(ori_dat, x, e))} %>%
+      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
+      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
+  }
+}
+
+patchwork::wrap_plots(plot_list, ncol = 7)
+```
+
+Sigma: 10
+
+```{r fig.height=18, fig.width=24, cache = TRUE}
+mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 10)
+ori_dat <- mod$gen(50)
+i <- 0
+plot_list <- list()
+
+for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
+  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
+    i <- i + 1
+    mod$set_prm("a", a) %>%
+      {.$set_prm("b", b)} %>%
+      {.$gen(50, computed = select(ori_dat, x, e))} %>%
+      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
+      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
+  }
+}
+
+patchwork::wrap_plots(plot_list, ncol = 7)
+```
+
+# N: 100
+
+Sigma: 1
+
+```{r fig.height=18, fig.width=24, cache = TRUE}
+mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 1)
+ori_dat <- mod$gen(100)
+i <- 0
+plot_list <- list()
+
+for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
+  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
+    i <- i + 1
+    mod$set_prm("a", a) %>%
+      {.$set_prm("b", b)} %>%
+      {.$gen(100, computed = select(ori_dat, x, e))} %>%
+      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
+      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
+  }
+}
+
+patchwork::wrap_plots(plot_list, ncol = 7)
+```
+
+Sigma: 2
+
+```{r fig.height=18, fig.width=24, cache = TRUE}
+mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 2)
+ori_dat <- mod$gen(100)
+i <- 0
+plot_list <- list()
+
+for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
+  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
+    i <- i + 1
+    mod$set_prm("a", a) %>%
+      {.$set_prm("b", b)} %>%
+      {.$gen(100, computed = select(ori_dat, x, e))} %>%
+      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
+      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
+  }
+}
+
+patchwork::wrap_plots(plot_list, ncol = 7)
+```
+
+Sigma: 5
+
+```{r fig.height=18, fig.width=24, cache = TRUE}
+mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 5)
+ori_dat <- mod$gen(100)
+i <- 0
+plot_list <- list()
+
+for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
+  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
+    i <- i + 1
+    mod$set_prm("a", a) %>%
+      {.$set_prm("b", b)} %>%
+      {.$gen(100, computed = select(ori_dat, x, e))} %>%
+      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
+      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
+  }
+}
+
+patchwork::wrap_plots(plot_list, ncol = 7)
+```
+
+Sigma: 10
+
+```{r fig.height=18, fig.width=24, cache = TRUE}
+mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 10)
+ori_dat <- mod$gen(100)
+i <- 0
+plot_list <- list()
+
+for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
+  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
+    i <- i + 1
+    mod$set_prm("a", a) %>%
+      {.$set_prm("b", b)} %>%
+      {.$gen(100, computed = select(ori_dat, x, e))} %>%
+      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
+      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
+  }
+}
+
+patchwork::wrap_plots(plot_list, ncol = 7)
+```
+
+# N: 300
+
+Sigma: 1
+
+```{r fig.height=18, fig.width=24, cache = TRUE}
+mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 1)
+ori_dat <- mod$gen(300)
+i <- 0
+plot_list <- list()
+
+for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
+  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
+    i <- i + 1
+    mod$set_prm("a", a) %>%
+      {.$set_prm("b", b)} %>%
+      {.$gen(300, computed = select(ori_dat, x, e))} %>%
+      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
+      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
+  }
+}
+
+patchwork::wrap_plots(plot_list, ncol = 7)
+```
+
+Sigma: 2
+
+```{r fig.height=18, fig.width=24, cache = TRUE}
+mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 2)
+ori_dat <- mod$gen(300)
+i <- 0
+plot_list <- list()
+
+for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
+  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
+    i <- i + 1
+    mod$set_prm("a", a) %>%
+      {.$set_prm("b", b)} %>%
+      {.$gen(300, computed = select(ori_dat, x, e))} %>%
+      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
+      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
+  }
+}
+
+patchwork::wrap_plots(plot_list, ncol = 7)
+```
+
+Sigma: 5
+
+```{r fig.height=18, fig.width=24, cache = TRUE}
+mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 5)
+ori_dat <- mod$gen(300)
+i <- 0
+plot_list <- list()
+
+for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
+  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
+    i <- i + 1
+    mod$set_prm("a", a) %>%
+      {.$set_prm("b", b)} %>%
+      {.$gen(300, computed = select(ori_dat, x, e))} %>%
+      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
+      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
+  }
+}
+
+patchwork::wrap_plots(plot_list, ncol = 7)
+```
+
+Sigma: 10
+
+```{r fig.height=18, fig.width=24, cache = TRUE}
+mod <- simple_cubic_model(x = rand_uniform(-1, 1), sigma = 10)
+ori_dat <- mod$gen(300)
+i <- 0
+plot_list <- list()
+
+for (a in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
+  for (b in c(-1.5, -0.75, -0.25, 0, 0.25, 0.75, 1.5)) {
+    i <- i + 1
+    mod$set_prm("a", a) %>%
+      {.$set_prm("b", b)} %>%
+      {.$gen(300, computed = select(ori_dat, x, e))} %>%
+      {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
+      ggtitle(glue::glue("a:{a}, b:{b}")) -> plot_list[[i]]
+  }
+}
+
+patchwork::wrap_plots(plot_list, ncol = 7)
+```

### 2022-06-13 paper_comparison/explore/check_poly_model.Rmd
explore poly model
upload check model

@@
 
-Sigma: 0.125
-
-```{r fig.height=6, fig.width=8, cache = TRUE}
-mod <- poly_model(2, x = rand_uniform(-1, 1), sigma = 0.125)
-ori_dat <- mod$gen(50)
-i <- 0
-plot_list <- list()
-
-for (a in c(2, 3, 4, 5, 6)) {
-
-  i <- i + 1
-  mod$set_prm("a", a) %>%
-    {.$gen(50, computed = select(ori_dat, x, e))} %>%
-    {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
-    ggtitle(glue::glue("a:{a}")) -> plot_list[[i]]
-
-}
-
-patchwork::wrap_plots(plot_list)
-```
-
-Sigma: 0.25
-
-```{r fig.height=6, fig.width=8, cache = TRUE}
-mod <- poly_model(2, x = rand_uniform(-1, 1), sigma = 0.25)
-ori_dat <- mod$gen(50)
-i <- 0
-plot_list <- list()
-
-for (a in c(2, 3, 4, 5, 6)) {
-
-  i <- i + 1
-  mod$set_prm("a", a) %>%
-    {.$gen(50, computed = select(ori_dat, x, e))} %>%
-    {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
-    ggtitle(glue::glue("a:{a}")) -> plot_list[[i]]
-
-}
-
-patchwork::wrap_plots(plot_list)
-```
-
 Sigma: 0.5
@@
 
-for (a in c(2, 3, 4, 5, 6)) {
+for (shape in c(1, 2, 3, 4)) {
 
   i <- i + 1
-  mod$set_prm("a", a) %>%
+  mod$set_prm("shape", shape) %>%
     {.$gen(50, computed = select(ori_dat, x, e))} %>%
     {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
-    ggtitle(glue::glue("a:{a}")) -> plot_list[[i]]
+    ggtitle(glue::glue("shape:{shape}")) -> plot_list[[i]]
 
@@
 
-for (a in c(2, 3, 4, 5, 6)) {
+for (shape in c(1, 2, 3, 4)) {
 
   i <- i + 1
-  mod$set_prm("a", a) %>%
+  mod$set_prm("shape", shape) %>%
     {.$gen(50, computed = select(ori_dat, x, e))} %>%
     {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
-    ggtitle(glue::glue("a:{a}")) -> plot_list[[i]]
+    ggtitle(glue::glue("shape:{shape}")) -> plot_list[[i]]
+
+}
+
+patchwork::wrap_plots(plot_list)
+```
+
+Sigma: 2
+
+```{r fig.height=6, fig.width=8, cache = TRUE}
+mod <- poly_model(2, x = rand_uniform(-1, 1), sigma = 2)
+ori_dat <- mod$gen(50)
+i <- 0
+plot_list <- list()
+
+for (shape in c(1, 2, 3, 4)) {
+
+  i <- i + 1
+  mod$set_prm("shape", shape) %>%
+    {.$gen(50, computed = select(ori_dat, x, e))} %>%
+    {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
+    ggtitle(glue::glue("shape:{shape}")) -> plot_list[[i]]
+
+}
+
+patchwork::wrap_plots(plot_list)
+```
+
+Sigma: 4
+
+```{r fig.height=6, fig.width=8, cache = TRUE}
+mod <- poly_model(2, x = rand_uniform(-1, 1), sigma = 4)
+ori_dat <- mod$gen(50)
+i <- 0
+plot_list <- list()
+
+for (shape in c(1, 2, 3, 4)) {
+
+  i <- i + 1
+  mod$set_prm("shape", shape) %>%
+    {.$gen(50, computed = select(ori_dat, x, e))} %>%
+    {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
+    ggtitle(glue::glue("shape:{shape}")) -> plot_list[[i]]
 
@@
 # N: 100
-
-Sigma: 0.125
-
-```{r fig.height=6, fig.width=8, cache = TRUE}
-mod <- poly_model(2, x = rand_uniform(-1, 1), sigma = 0.125)
-ori_dat <- mod$gen(100)
-i <- 0
-plot_list <- list()
-
-for (a in c(2, 3, 4, 5, 6)) {
-
-  i <- i + 1
-  mod$set_prm("a", a) %>%
-    {.$gen(100, computed = select(ori_dat, x, e))} %>%
-    {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
-    ggtitle(glue::glue("a:{a}")) -> plot_list[[i]]
-
-}
-
-patchwork::wrap_plots(plot_list)
-```
-
-Sigma: 0.25
-
-```{r fig.height=6, fig.width=8, cache = TRUE}
-mod <- poly_model(2, x = rand_uniform(-1, 1), sigma = 0.25)
-ori_dat <- mod$gen(100)
-i <- 0
-plot_list <- list()
-
-for (a in c(2, 3, 4, 5, 6)) {
-
-  i <- i + 1
-  mod$set_prm("a", a) %>%
-    {.$gen(100, computed = select(ori_dat, x, e))} %>%
-    {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
-    ggtitle(glue::glue("a:{a}")) -> plot_list[[i]]
-
-}
-
-patchwork::wrap_plots(plot_list)
-```
 
@@
 
-for (a in c(2, 3, 4, 5, 6)) {
+for (shape in c(1, 2, 3, 4)) {
 
   i <- i + 1
-  mod$set_prm("a", a) %>%
-    {.$gen(100, computed = select(ori_dat, x, e))} %>%
+  mod$set_prm("shape", shape) %>%
+    {.$gen(50, computed = select(ori_dat, x, e))} %>%
     {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
-    ggtitle(glue::glue("a:{a}")) -> plot_list[[i]]
+    ggtitle(glue::glue("shape:{shape}")) -> plot_list[[i]]
 
@@
 
-for (a in c(2, 3, 4, 5, 6)) {
+for (shape in c(1, 2, 3, 4)) {
 
   i <- i + 1
-  mod$set_prm("a", a) %>%
-    {.$gen(100, computed = select(ori_dat, x, e))} %>%
+  mod$set_prm("shape", shape) %>%
+    {.$gen(50, computed = select(ori_dat, x, e))} %>%
     {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
-    ggtitle(glue::glue("a:{a}")) -> plot_list[[i]]
+    ggtitle(glue::glue("shape:{shape}")) -> plot_list[[i]]
+
+}
+
+patchwork::wrap_plots(plot_list)
+```
+
+Sigma: 2
+
+```{r fig.height=6, fig.width=8, cache = TRUE}
+mod <- poly_model(2, x = rand_uniform(-1, 1), sigma = 2)
+ori_dat <- mod$gen(100)
+i <- 0
+plot_list <- list()
+
+for (shape in c(1, 2, 3, 4)) {
+
+  i <- i + 1
+  mod$set_prm("shape", shape) %>%
+    {.$gen(50, computed = select(ori_dat, x, e))} %>%
+    {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
+    ggtitle(glue::glue("shape:{shape}")) -> plot_list[[i]]
+
+}
+
+patchwork::wrap_plots(plot_list)
+```
+
+Sigma: 4
+
+```{r fig.height=6, fig.width=8, cache = TRUE}
+mod <- poly_model(2, x = rand_uniform(-1, 1), sigma = 4)
+ori_dat <- mod$gen(100)
+i <- 0
+plot_list <- list()
+
+for (shape in c(1, 2, 3, 4)) {
+
+  i <- i + 1
+  mod$set_prm("shape", shape) %>%
+    {.$gen(50, computed = select(ori_dat, x, e))} %>%
+    {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
+    ggtitle(glue::glue("shape:{shape}")) -> plot_list[[i]]
 
@@
 # N: 300
-
-Sigma: 0.125
-
-```{r fig.height=6, fig.width=8, cache = TRUE}
-mod <- poly_model(2, x = rand_uniform(-1, 1), sigma = 0.125)
-ori_dat <- mod$gen(300)
-i <- 0
-plot_list <- list()
-
-for (a in c(2, 3, 4, 5, 6)) {
-
-  i <- i + 1
-  mod$set_prm("a", a) %>%
-    {.$gen(300, computed = select(ori_dat, x, e))} %>%
-    {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
-    ggtitle(glue::glue("a:{a}")) -> plot_list[[i]]
-
-}
-
-patchwork::wrap_plots(plot_list)
-```
-
-Sigma: 0.25
-
-```{r fig.height=6, fig.width=8, cache = TRUE}
-mod <- poly_model(2, x = rand_uniform(-1, 1), sigma = 0.25)
-ori_dat <- mod$gen(300)
-i <- 0
-plot_list <- list()
-
-for (a in c(2, 3, 4, 5, 6)) {
-
-  i <- i + 1
-  mod$set_prm("a", a) %>%
-    {.$gen(300, computed = select(ori_dat, x, e))} %>%
-    {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
-    ggtitle(glue::glue("a:{a}")) -> plot_list[[i]]
-
-}
-
-patchwork::wrap_plots(plot_list)
-```
 
@@
 
-for (a in c(2, 3, 4, 5, 6)) {
+for (shape in c(1, 2, 3, 4)) {
 
   i <- i + 1
-  mod$set_prm("a", a) %>%
-    {.$gen(300, computed = select(ori_dat, x, e))} %>%
+  mod$set_prm("shape", shape) %>%
+    {.$gen(50, computed = select(ori_dat, x, e))} %>%
     {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
-    ggtitle(glue::glue("a:{a}")) -> plot_list[[i]]
+    ggtitle(glue::glue("shape:{shape}")) -> plot_list[[i]]
 
@@
 ```
+
+Sigma: 2
+
+```{r fig.height=6, fig.width=8, cache = TRUE}
+mod <- poly_model(2, x = rand_uniform(-1, 1), sigma = 2)
+ori_dat <- mod$gen(300)
+i <- 0
+plot_list <- list()
+
+for (shape in c(1, 2, 3, 4)) {
+
+  i <- i + 1
+  mod$set_prm("shape", shape) %>%
+    {.$gen(50, computed = select(ori_dat, x, e))} %>%
+    {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
+    ggtitle(glue::glue("shape:{shape}")) -> plot_list[[i]]
+
+}
+
+patchwork::wrap_plots(plot_list)
+```
+
+Sigma: 4
+
+```{r fig.height=6, fig.width=8, cache = TRUE}
+mod <- poly_model(2, x = rand_uniform(-1, 1), sigma = 4)
+ori_dat <- mod$gen(300)
+i <- 0
+plot_list <- list()
+
+for (shape in c(1, 2, 3, 4)) {
+
+  i <- i + 1
+  mod$set_prm("shape", shape) %>%
+    {.$gen(50, computed = select(ori_dat, x, e))} %>%
+    {mod$plot(., remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light())} +
+    ggtitle(glue::glue("shape:{shape}")) -> plot_list[[i]]
+
+}
+
+patchwork::wrap_plots(plot_list)
+```

### 2022-06-14 Third_experiment/check_allocation.R
add attention checks

@@ -2,8 +2,15 @@ library(tidyverse)
 library(visage)
 library(UpSetR)
 
-result <- read_csv("allocate_result_full.csv")
-lineup <- read_csv("lineup_info.csv")
+result <- read_csv("data/allocate_result_full.csv")
+lineup <- read_csv("data/lineup_info.csv")
+
+# omit attention checks
+result <- result %>%
+  filter(lineup_id <= 576)
+
+lineup <- lineup %>%
+  filter(lineup_id <= 576)
 
 result %>%
   # filter(subject %in% 1:120) %>%

### 2022-06-14 Third_experiment/exp_setup.R

@@ -287,12 +287,34 @@ allocate_result <- allocate_result %>%
   mutate(subject = gsub(".+_", "", name)) %>%
   select(subject, lineup_id = value)
 
-write_csv(allocate_result, "allocate_result.csv")
-write_csv(lineup, "lineup_info.csv")
+set.seed(10086)
+
+lineup <- lineup %>%
+  bind_rows(expand.grid(shape = 1:4,
+                        n = 300,
+                        x_dist = "uniform",
+                        e_sigma = 0.125,
+                        rep = 1:3) %>%
+              mutate(lineup_id = nrow(lineup) + 1:n()))
+
+
+for (i in 1:160) {
+  allocate_result <- allocate_result %>%
+    bind_rows(tibble(subject = i,
+                     lineup_id = sample(1:12, 2) + 576))
+}
+
+allocate_result <- allocate_result %>%
+  group_by(subject) %>%
+  mutate(order = sample(1:20))
+
+
+write_csv(allocate_result, "data/allocate_result.csv")
+write_csv(lineup, "data/lineup_info.csv")
 
 
 allocate_result %>%
   left_join(lineup, by = c("lineup_id")) %>%
-  write_csv("allocate_result_full.csv")
+  write_csv("data/allocate_result_full.csv")
 
 

### 2022-06-14 Third_experiment/generate_lineup.R

@@ -1,14 +1,10 @@
 library(tidyverse)
 library(visage)
 
-sample_dat <- readRDS("data/all_data_big.RDS")
-
-lineup <- read_csv("lineup_info.csv")
+lineup <- read_csv("data/lineup_info.csv")
 plot_string <- read_csv("data/plot_string.txt", col_names = FALSE) %>%
   .$X1
-allocated <- read_csv("allocate_result.csv") %>%
-  group_by(subject) %>%
-  mutate(current_lineup = 1:n())
+allocated <- read_csv("data/allocate_result.csv")
 dat <- vector(mode = "list", length = nrow(lineup))
 
 set.seed(10086)
@@ -48,12 +44,13 @@ file_path <- "plots"
 
 for (j in 1:nrow(allocated)) {
 
-  ggsave(glue::glue("{file_path}/{plot_string[allocated$subject[j]]}_{allocated$subject[j]}_{allocated$current_lineup[j]}.png"),
+  ggsave(glue::glue("{file_path}/{plot_string[allocated$subject[j]]}_{allocated$subject[j]}_{allocated$order[j]}.png"),
          plot = VI_MODEL$plot_lineup(dat[[allocated$lineup_id[j]]]$data,
                                      remove_axis = TRUE,
                                      remove_legend = TRUE,
                                      remove_grid_line = TRUE,
-                                     theme = theme_light()),
+                                     theme = theme_light()) +
+           ggtitle(glue::glue("lineup:{allocated$lineup_id[j]}, shape:{dat[[allocated$lineup_id[j]]]$metadata$shape}, x_dist:{dat[[allocated$lineup_id[j]]]$metadata$x_dist}, e_sigma:{dat[[allocated$lineup_id[j]]]$metadata$e_sigma}, log effect size:{round(log(dat[[allocated$lineup_id[j]]]$metadata$effect_size), 2)}, ans: {dat[[allocated$lineup_id[j]]]$metadata$ans}")),
          height = 7,
          width = 7)
 }
@@ -67,12 +64,13 @@ for (j in 1:nrow(lineup)) {
                                      remove_legend = TRUE,
                                      remove_grid_line = TRUE,
                                      theme = theme_light()) +
[…]

### 2022-06-14 Third_experiment/setup.R

@@ -1,157 +1,0 @@
-# a in -1.5 ~ 1.5
-# b in -1.5 ~ 1.5
-# x_dist = uniform, normal, lognormal, neglognormal, even_dis, uneven_dis
-# N = 50, 100, 300
-# rep = 2
-# e_sigma = 0.125, 0.25, 0.5, 1
-
-# N_ab * 6 * 3 * 2 * 4
-
-library(tidyverse)
-library(visage)
-library(lhs)
-
-get_trunc_norm_cdf <- function(sd = 1, ndraws = 10^7) {
-  result <- rnorm(ndraws, mean = 0, sd = sd)
-  result <- result[result >= -1.5 & result <= 1.5]
-  function(p) {
-    quantile(result, probs = p) %>%
-      unname()
-  }
-}
-
-trans_unif_2_norm <- function(x, sd = 1, ndraws = 10^7) {
-  p_trunc_norm <- get_trunc_norm_cdf(sd, ndraws)
-  p_trunc_norm(x)
-}
-
-sim_effect_size <- function(a, b) {
-  x_dist <- sample(c("uniform", "normal", "lognormal", "neglognormal",
-                     "even_discrete", "uneven_discrete"), 1)
-  e_sigma <- sample(c(0.125, 0.25, 0.5, 1), 1)
-  n <- sample(c(50, 100, 300), 1)
-
-  x <- switch(x_dist,
-              uniform = rand_uniform(-1, 1),
-              normal = rand_normal(sigma = 0.3),
-              lognormal = rand_lognormal(sigma = 0.6),
-              neglognormal = {rl <- rand_lognormal(sigma = 0.6); closed_form(~-rl)},
-              even_discrete = rand_uniform_d(k = 5, even = TRUE),
-              uneven_discrete = rand_uniform_d(k = 5, even = FALSE))
-
-  mod <- simple_cubic_model(a = a, b = b, x = x, sigma = e_sigma)
-  dat <- mod$gen(n, fit_model = FALSE)
-  tibble(a, b, x_dist, e_sigma, n, effect_size = mod$effect_size(dat))
-}
-
-
-result <- map(1:10000, ~sim_effect_size(1, 1)) %>%
-  reduce(bind_rows)
-
-result %>%
-  group_by(x_dist, e_sigma, n) %>%
[…]

### 2022-06-30 Third_experiment/quick_check_data.R
upload all surveys

@@ -185,11 +185,10 @@ poly_result %>%
               summarise(button = sum(num_selection < 1))) %>%
   left_join(poly_result %>%
               group_by(set) %>%
-              summarise(many = sum(num_selection > 5))) -> check_dat
-
-
-
-
+              summarise(many_5 = sum(num_selection > 5))) %>%
+  left_join(poly_result %>%
+              group_by(set) %>%
+              summarise(many_10 = sum(num_selection > 10))) -> check_dat
 
 
 cenv <- new.env()
@@ -200,6 +199,51 @@ p_value <- poly_result %>%
               count(lineup_id) %>%
               rename(num_eval = n)) %>%
   left_join(lineup_info)
+
+
+
+conv_result <- map(1:100000,
+                   function(x) {
+                     shape <- sample(1:4, 1)
+                     e_sigma <- sample(c(0.5, 1, 2, 4), 1)
+                     x_dist <- sample(c("uniform", "normal", "lognormal", "even_discrete"), 1)
+                     x <- switch(x_dist,
+                                 uniform = rand_uniform(-1, 1),
+                                 normal = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))},
+                                 lognormal = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))},
+                                 even_discrete = rand_uniform_d(k = 5, even = TRUE))
+                     mod <- poly_model(shape, x = x, sigma = e_sigma)
+                     n <- sample(c(50, 100, 300), 1)
+                     tmp_dat <- mod$gen(n)
+
+                     tibble(shape = shape,
+                            e_sigma = e_sigma,
+                            x_dist = x_dist,
+                            n = n,
+                            effect_size = mod$effect_size(tmp_dat),
+                            reject = mod$test(tmp_dat)$p_value < 0.05)
+                   }) %>%
+  reduce(bind_rows)
+
+p_value %>%
+  left_join(poly_result %>%
+              group_by(lineup_id) %>%
+              summarise(effect_size = first(effect_size))) %>%
+  filter(lineup_id < 577) %>%
[…]

### 2022-06-30 paper_comparison/paper_comparison/paper_comparison.Rmd
update paper

--- 2022-05-12 paper_comparison/paper_comparison/paper_comparison.Rmd
@@
 library(visage)
-pilot <- read_csv(here::here("data/processed/processed_pilot.csv"))
-survey <- read_csv(here::here("data/processed/processed_surveys.csv"))
-sim_conv <- read_csv(here::here("data/processed/sim_conv_result.csv"), col_types = "dddddcdddlcdcdddlc")
+
+set.seed(10086)
 ```
 
-```{r}
-survey <- survey %>%
-  mutate(log_effect_size = log(effect_size))
+```{r get-conv-result}
+stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
 
-sim_conv <- sim_conv %>%
-  mutate(log_effect_size = log(effect_size))
+if (!file.exists(here::here("data/third_study/conv_result.rds"))) {
+  conv_result <- map(1:100000,
+                   function(x) {
+                     shape <- sample(1:4, 1)
+                     e_sigma <- sample(c(0.5, 1, 2, 4), 1)
+                     x_dist <- sample(c("uniform", "normal", "lognormal", "even_discrete"), 1)
+                     x <- switch(x_dist,
+                                 uniform = rand_uniform(-1, 1),
+                                 normal = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))},
+                                 lognormal = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))},
+                                 even_discrete = rand_uniform_d(k = 5, even = TRUE))
+                     mod <- poly_model(shape, x = x, sigma = e_sigma)
+                     n <- sample(c(50, 100, 300), 1)
+                     tmp_dat <- mod$gen(n)
 
-lineup_info <- select(survey, type:answer, conventional_p_value, ori_difficulty, exp, exp_lineup_id, log_effect_size) %>%
-  group_by(exp_lineup_id) %>% 
-  filter(row_number() == 1) %>%
-  ungroup()
+                     tibble(shape = shape,
+                            e_sigma = e_sigma,
+                            x_dist = x_dist,
+                            n = n,
+                            effect_size = mod$effect_size(tmp_dat),
+                            reject = mod$test(tmp_dat)$p_value < 0.05)
+                   }) %>%
+  reduce(bind_rows) %>%
+  mutate(reject = as.numeric(reject))
 
-write_csv(lineup_info, here::here("data/processed/processed_lineup_info.csv"))
-
-# Local constant estimator
-kernel_regression_cubic <- function(sim_conv, h = 0.3, x = quote(log_effect_size)) {
-  with(filter(sim_conv, type == "cubic", log_effect_size > -5, log_effect_size < 10),
-     KernSmooth::locpoly(x = eval(x), y = reject, bandwidth = h, degree = 0, gridsize = 1000)) %>%
-    as_tibble() %>%
-    mutate(type = "cubic", h = h) %>%
-    rename(reject = y)
+  saveRDS(conv_result, here::here("data/third_study/conv_result.rds"))
+} else {
+  conv_result <- readRDS(here::here("data/third_study/conv_result.rds"))
 }
 
-# Local constant estimator
-kernel_regression_heter <- function(sim_conv, h = 0.4, x = quote(log_effect_size)) {
-  with(filter(sim_conv, type != "cubic", log_effect_size > -3),
-     KernSmooth::locpoly(x = eval(x), y = reject, bandwidth = h, degree = 0, gridsize = 1000)) %>%
-    as_tibble() %>%
-    mutate(type = "heteroskedasticity", h = h) %>%
-    rename(reject = y)
+```
+
+```{r get-third-result}
+if (!file.exists(here::here("data/third_study/polynomials_lineup.rds"))) {
+  polynomials_lineup <- get_polynomials_lineup()
+  saveRDS(polynomials_lineup, here::here("data/third_study/polynomials_lineup.rds"))
+} else {
+  polynomials_lineup <- readRDS(here::here("data/third_study/polynomials_lineup.rds"))
 }
@@
 
-```{r}
-if (file.exists(here::here("data/processed/lineup_env.rds"))) {
-  env <- readRDS(here::here("data/processed/lineup_env.rds"))
-} else {
-  env <- new.env()
-}
-
-save_lineup_env <- function(env) {
-  saveRDS(env, here::here("data/processed/lineup_env.rds"))
-}
-
-survey_pvalue <- survey %>%
-  calc_p_value_multi(lineup_id = "exp_lineup_id", 
-                     detected = "detect", 
-                     n_sel = "num_selection",
-                     comb = TRUE,
-                     n_eval = 1:5,
-                     n_sim = 100000,
-                     cache_env = env) %>%
-  rename(exp_lineup_id = lineup_id) %>%
-  unnest_longer(col = p_value) %>%
-  left_join(lineup_info) %>%
-  mutate(reject = eval_p_value(p_value, tol = 0.01))
-
-save_lineup_env(env)
+```{r power-vs-log-effect-size}
+polynomials %>%
+  group_by(lineup_id) %>%
+  summarise(across(c(type:p_value), first)) %>%
+  filter(lineup_id < 577) %>%
+  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
+  ggplot() +
+  geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  xlab("Natural logarithm of effect size") +
+  ylab("Power")
 ```
 
-```{r results='asis'}
-map(1:5, function(x) glm(reject ~ log_effect_size,
-                         family = binomial(),
-                         data = filter(survey_pvalue, n_eval == {{x}}, type == "cubic"))) %>%
-  stargazer::stargazer(header = FALSE,
-            label = "tab:powerglmcubic",
-            title = "Logistic regressions with rejection as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for cubic model.
-Five regression were fitted for different number of evaluations.",
-            style = "default",
-            column.labels = c("num\\_eval: 1", "num\\_eval:2", "num\\_eval: 3", "num\\_eval: 4", "num\\_eval:5"),
-            model.numbers = FALSE)
+```{r power-vs-log-effect-size-given-x-dist}
+polynomials %>%
+  group_by(lineup_id) %>%
+  summarise(across(c(type:p_value), first)) %>%
+  filter(lineup_id < 577) %>%
+  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
+  ggplot() +
+  geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  facet_wrap(~x_dist) +
+  xlab("Natural logarithm of effect size") +
+  ylab("Power") +
+  ggtitle("Power comparison conditional on the distribution of X")
 ```
 
-```{r results="asis"}
-map(1:5, function(x) glm(reject ~ log_effect_size,
-                         family = binomial(),
-                         data = filter(survey_pvalue, n_eval == {{x}}, type == "heteroskedasticity"))) %>%
-  stargazer::stargazer(header = FALSE,
-            label = "tab:powerglmheter",
-            title = "Logistic regressions with rejection as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for heteroskedasticity model.
-Five regression were fitted for different number of evaluations.",
-            style = "default",
-            column.labels = c("num\\_eval: 1", "num\\_eval:2", "num\\_eval: 3", "num\\_eval: 4", "num\\_eval:5"),
-            model.numbers = FALSE)
+```{r power-of-visual-test-given-x-dist}
+polynomials %>%
+  group_by(lineup_id) %>%
+  summarise(across(c(type:p_value), first)) %>%
+  filter(lineup_id < 577) %>%
+  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
+  ggplot() +
+  geom_smooth(aes(log(effect_size), reject, col = x_dist), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  xlab("Natural logarithm of effect size") +
+  ylab("Power") +
+  ggtitle("Power of visual test conditional on the distribution of X")
 ```
 
-```{r power-com, fig.cap="\\label{fig:power-com}power com"}
-survey_pvalue %>%
-  mutate(reject = as.numeric(reject)) %>%
-  mutate(n_eval = factor(n_eval)) %>%
+```{r power-vs-log-effect-size-given-shape}
+polynomials %>%
+  group_by(lineup_id) %>%
+  summarise(across(c(type:p_value), first)) %>%
+  filter(lineup_id < 577) %>%
+  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
   ggplot() +
-  geom_point(aes(log_effect_size, reject), alpha = 0.01) +
-  geom_line(data = kernel_regression_cubic(sim_conv, 0.4), aes(x, reject), linetype = 2) +
-  geom_line(data = kernel_regression_heter(sim_conv, 0.75), aes(x, reject), linetype = 2) +
-  geom_smooth(aes(log_effect_size, reject, group = n_eval, col = n_eval),
-              method = "glm", method.args = list(family = binomial), size = 0.5, se = FALSE) +
-  scale_colour_viridis_d(option = "B", begin = 0.5, end = 0.9) +
-  facet_wrap(~type, scales = "free_x") +
-  ylab("Estimated power")
+  geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  facet_wrap(~shape) +
+  xlab("Natural logarithm of effect size") +
+  ylab("Power") +
+  ggtitle("Power comparison conditional on the shape of the Hermite polynomials")
 ```
 
-```{r cubic-hard, fig.cap="\\label{fig:cubic-hard}cubic-hard ans 11"}
-big_lineup_dat <- readRDS(here::here("data/big_study/all_data_big.RDS"))
-
-# ans 11, conventional 0.003491033
-CUBIC_MODEL$plot_lineup(mutate(big_lineup_dat[[28]]$data, .fitted = .pred),
-                        remove_axis = TRUE,
-                        theme = theme_light(),
-                        remove_grid_line = TRUE)
-
+```{r power-of-visual-test-given-shape}
+polynomials %>%
+  group_by(lineup_id) %>%
+  summarise(across(c(type:p_value), first)) %>%
+  filter(lineup_id < 577) %>%
+  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01)), shape = factor(shape)) %>%
+  ggplot() +
+  geom_smooth(aes(log(effect_size), reject, col = shape), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  xlab("Natural logarithm of effect size") +
+  ylab("Power") +
+  ggtitle("Power of visual test conditional on the shape of the Hermite polynomials")
 ```
 
-```{r heter-hard, fig.cap="\\label{fig:heter-hard}heter-hard ans 2"}
-# ans 2, conventional 0.003535915
-HETER_MODEL$plot_lineup(mutate(big_lineup_dat[[136]]$data, .fitted = .pred),
-                        remove_axis = TRUE,
-                        theme = theme_light(),
-                        remove_grid_line = TRUE)
+```{r power-vs-log-effect-size-given-number-of-observations}
+polynomials %>%
+  group_by(lineup_id) %>%
+  summarise(across(c(type:p_value), first)) %>%
+  filter(lineup_id < 577) %>%
+  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
+  ggplot() +
+  geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  facet_wrap(~n) +
+  xlab("Natural logarithm of effect size") +
+  ylab("Power") +
+  ggtitle("Power comparison conditional on the number of observations")
 ```
 
-```{r power-com-cubic-a, fig.cap="\\label{fig:power-com-cubic-a}power com cubic a"}
-bind_grid <- expand.grid(n = c(50, 100, 300),
-                         h = 0.4,
-                         x = c(quote(abs(a)/e_sigma)))
-
-bind_list <- list()
-
-for (i in 1:nrow(bind_grid)) {
-  (function(n, h, x) {
-    mutate(kernel_regression_cubic(filter(sim_conv, n == {{n}}), h = h, x = x),
-           n = factor(n, levels = c(50, 100, 300)))
-    })(n = bind_grid$n[i], h = bind_grid$h[i], x = bind_grid$x[[i]]) -> bind_list[[i]]
-}
-
-bind_sim <- reduce(bind_list, bind_rows)
-
-# cubic break down
-survey_pvalue %>%
-  filter(type == "cubic") %>%
-  mutate(n = factor(n, levels = c(50, 100, 300))) %>%
-  mutate(n_eval = factor(n_eval)) %>%
-  mutate(reject = as.numeric(reject)) %>%
+```{r power-of-visual-test-given-number-of-observations}
+polynomials %>%
+  group_by(lineup_id) %>%
+  summarise(across(c(type:p_value), first)) %>%
+  filter(lineup_id < 577) %>%
+  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01)), n = factor(n)) %>%
   ggplot() +
-  ggbeeswarm::geom_quasirandom(aes(abs(a)/e_sigma, reject, group = n), groupOnX = FALSE, alpha = 0.01) +
-  geom_smooth(aes(abs(a)/e_sigma, reject, col = n_eval), method = "glm", method.args = list(family = binomial()), se = FALSE) +
-  geom_line(data = bind_sim, aes(x, reject), col = "grey", linetype = 2, size = 1) +
-  facet_wrap(~n, labeller = label_both) +
-  scale_colour_viridis_d(option = "B", begin = 0.5, end = 0.9) +
-  theme_light()
+  geom_smooth(aes(log(effect_size), reject, col = n), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  xlab("Natural logarithm of effect size") +
+  ylab("Power") +
+  ggtitle("Power of visual test conditional on the number of observations")
 ```
-
-```{r power-com-cubic-b, fig.cap="\\label{fig:power-com-cubic-b}power com cubic b"}
-bind_grid <- expand.grid(n = c(50, 100, 300),
-                         h = 0.4,
-                         x = c(quote(abs(b)/e_sigma)))
-
-bind_list <- list()
-
-for (i in 1:nrow(bind_grid)) {
-  (function(n, h, x) {
-    mutate(kernel_regression_cubic(filter(sim_conv, n == {{n}}), h = h, x = x),
-           n = factor(n, levels = c(50, 100, 300)))
-    })(n = bind_grid$n[i], h = bind_grid$h[i], x = bind_grid$x[[i]]) -> bind_list[[i]]
-}
-
-bind_sim <- reduce(bind_list, bind_rows)
-
-# cubic break down
-survey_pvalue %>%
-  filter(type == "cubic") %>%
-  mutate(n = factor(n, levels = c(50, 100, 300))) %>%
-  mutate(reject = as.numeric(reject)) %>%
-  mutate(n_eval = factor(n_eval)) %>%
-  ggplot() +
-  ggbeeswarm::geom_quasirandom(aes(abs(b)/e_sigma, reject, group = n), groupOnX = FALSE, alpha = 0.01) +
-  geom_smooth(aes(abs(b)/e_sigma, reject, col = n_eval), method = "glm", method.args = list(family = binomial()), se = FALSE) +
-  geom_line(data = bind_sim, aes(x, reject), col = "grey", linetype = 2, size = 1) +
-  facet_wrap(~n, labeller = label_both) +
-  scale_colour_viridis_d(option = "B", begin = 0.5, end = 0.9) +
-  theme_light()
-```
-
-```{r eval = FALSE}
-# cubic break down
-# We do not have enough observations
-survey_pvalue %>%
-  filter(type == "cubic") %>%
-  mutate(n = factor(n)) %>%
-  mutate(reject = as.numeric(reject)) %>%
-  ggplot() +
-  geom_point(aes(abs(a), reject, col = n), alpha = 0.1) +
-  geom_smooth(aes(abs(a), reject, col = n), method = "glm", method.args = list(family = binomial()), se = FALSE) +
-  geom_smooth(data = mutate(filter(sim_conv, type == "cubic"), n = factor(n), reject = as.numeric(reject)),
-              aes(abs(a), reject, group = n), 
-              method = "glm", method.args = list(family = binomial()), se = FALSE, col = "grey", linetype = 2) +
-  facet_grid(e_sigma~n)
-```
-
-```{r power-com-heter-b, fig.cap="\\label{fig:power-com-heter-b}power com heter b"}
-bind_grid <- expand.grid(n = c(50, 100, 300),
-                         h = 3,
-                         x = c(quote(b)),
-                         a = c(-1, 0, 1))
-
-bind_list <- list()
-
-for (i in 1:nrow(bind_grid)) {
-  (function(n, h, x, a) {
-    mutate(kernel_regression_heter(filter(sim_conv, n == {{n}}, a == {{a}}), h = h, x = x),
-           n = factor(n, levels = c(50, 100, 300)),
-           a = factor(a, levels = c(-1, 0, 1)))
-    })(n = bind_grid$n[i], h = bind_grid$h[i], x = bind_grid$x[[i]], a = bind_grid$a[i]) -> bind_list[[i]]
-}
-
-bind_sim <- reduce(bind_list, bind_rows)
-
-# heter break down
-survey_pvalue %>%
-  filter(type != "cubic") %>%
-  mutate(n = factor(n, levels = c(50, 100, 300))) %>%
-  mutate(reject = as.numeric(reject)) %>%
-  mutate(n_eval = factor(n_eval)) %>%
-  ggplot() +
-  ggbeeswarm::geom_quasirandom(aes(b, reject, group = n), groupOnX = FALSE, alpha = 0.01) +
-  geom_smooth(aes(b, reject, col = n_eval), method = "glm", method.args = list(family = binomial()), se = FALSE) +
-  geom_line(data = bind_sim, aes(x, reject), col = "grey", linetype = 2, size = 1) +
-  theme_light() +
-  scale_colour_viridis_d(option = "B", begin = 0.5, end = 0.9) +
-  facet_grid(a~n, labeller = label_both)
-```

### 2022-06-30 paper_comparison/paper_comparison/paper_comparison.tex
update paper

@@
 The previous section focuses on the change of effect size relative to
-the power of the visual test.
+the power of the visual test. However, effect size is only a one
+dimensional summarisation of parameters used in data simulation.
+Individual factor embedded in the simulation process should also be
+analysed.
 
-\begin{table}[!htbp] \centering 
-  \caption{Logistic regressions with rejection as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for cubic model.
-Five regression were fitted for different number of evaluations.}
-  \label{tab:powerglmcubic} 
-\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
-\\[-1.8ex]\hline 
-\hline \\[-1.8ex] 
- & \multicolumn{5}{c}{\textit{Dependent variable:}} \\ 
-\cline{2-6} 
-\\[-1.8ex] & \multicolumn{5}{c}{reject} \\ 
- & num\_eval: 1 & num\_eval:2 & num\_eval: 3 & num\_eval: 4 & num\_eval:5 \\ 
-\hline \\[-1.8ex] 
- log\_effect\_size & 0.929$^{***}$ & 1.258$^{***}$ & 1.096$^{***}$ & 1.113$^{***}$ & 1.244$^{***}$ \\ 
-  & (0.073) & (0.076) & (0.054) & (0.056) & (0.077) \\ 
-  & & & & & \\ 
- Constant & $-$1.969$^{***}$ & $-$2.534$^{***}$ & $-$1.549$^{***}$ & $-$1.334$^{***}$ & $-$1.404$^{***}$ \\ 
-  & (0.169) & (0.173) & (0.113) & (0.122) & (0.185) \\ 
-  & & & & & \\ 
-\hline \\[-1.8ex] 
-Observations & 630 & 1,135 & 1,660 & 1,904 & 1,879 \\ 
-Log Likelihood & $-$220.882 & $-$314.394 & $-$471.152 & $-$396.979 & $-$206.296 \\ 
-Akaike Inf. Crit. & 445.764 & 632.788 & 946.305 & 797.958 & 416.592 \\ 
-\hline 
-\hline \\[-1.8ex] 
-\textit{Note:}  & \multicolumn{5}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
-\end{tabular} 
-\end{table}
+In cubic model, two major factors that influencing the strength of the
+signal are \(a\) and \(b\). Figure \ref{fig:power-com-cubic-a} and
+\ref{fig:power-com-cubic-a} illustrates 30 different logistic
+regressions fit for different number of evaluations and different number
+of observations \(n\). The regressor used in these models was
+\(|a|/\sigma\) since the noise level \(\sigma\) needed to be taken into
+account. From the figures, we can observe \ldots{}
 
-\begin{table}[!htbp] \centering 
-  \caption{Logistic regressions with rejection as response variable and natural logarithm of the effect size as regressor fitted on the data collected from experiment I and II for heteroskedasticity model.
-Five regression were fitted for different number of evaluations.}
-  \label{tab:powerglmheter} 
-\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
-\\[-1.8ex]\hline 
-\hline \\[-1.8ex] 
- & \multicolumn{5}{c}{\textit{Dependent variable:}} \\ 
-\cline{2-6} 
-\\[-1.8ex] & \multicolumn{5}{c}{reject} \\ 
- & num\_eval: 1 & num\_eval:2 & num\_eval: 3 & num\_eval: 4 & num\_eval:5 \\ 
-\hline \\[-1.8ex] 
- log\_effect\_size & 0.821$^{***}$ & 1.104$^{***}$ & 1.157$^{***}$ & 1.387$^{***}$ & 1.651$^{***}$ \\ 
-  & (0.086) & (0.079) & (0.066) & (0.076) & (0.110) \\ 
-  & & & & & \\ 
- Constant & $-$4.472$^{***}$ & $-$5.887$^{***}$ & $-$5.431$^{***}$ & $-$6.154$^{***}$ & $-$7.123$^{***}$ \\ 
-  & (0.434) & (0.415) & (0.344) & (0.400) & (0.591) \\ 
-  & & & & & \\ 
-\hline \\[-1.8ex] 
-Observations & 630 & 1,138 & 1,596 & 1,662 & 1,416 \\ 
-Log Likelihood & $-$333.119 & $-$631.519 & $-$835.317 & $-$650.570 & $-$342.058 \\ 
-Akaike Inf. Crit. & 670.238 & 1,267.038 & 1,674.634 & 1,305.140 & 688.116 \\ 
-\hline 
-\hline \\[-1.8ex] 
-\textit{Note:}  & \multicolumn{5}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
-\end{tabular} 
-\end{table}
+\includegraphics{paper_comparison_files/figure-latex/power-vs-log-effect-size-1.pdf}
 
-\begin{figure}
-\centering
-\includegraphics{paper_comparison_files/figure-latex/power-com-1.pdf}
-\caption{\label{fig:power-com}power com}
-\end{figure}
+\includegraphics{paper_comparison_files/figure-latex/power-vs-log-effect-size-given-x-dist-1.pdf}
 
-\begin{figure}
-\centering
-\includegraphics{paper_comparison_files/figure-latex/cubic-hard-1.pdf}
-\caption{\label{fig:cubic-hard}cubic-hard ans 11}
-\end{figure}
+\includegraphics{paper_comparison_files/figure-latex/power-of-visual-test-given-x-dist-1.pdf}
 
-\begin{figure}
-\centering
-\includegraphics{paper_comparison_files/figure-latex/heter-hard-1.pdf}
-\caption{\label{fig:heter-hard}heter-hard ans 2}
-\end{figure}
+\includegraphics{paper_comparison_files/figure-latex/power-vs-log-effect-size-given-shape-1.pdf}
 
-\begin{figure}
-\centering
-\includegraphics{paper_comparison_files/figure-latex/power-com-cubic-a-1.pdf}
-\caption{\label{fig:power-com-cubic-a}power com cubic a}
-\end{figure}
+\includegraphics{paper_comparison_files/figure-latex/power-of-visual-test-given-shape-1.pdf}
 
-\begin{figure}
-\centering
-\includegraphics{paper_comparison_files/figure-latex/power-com-cubic-b-1.pdf}
-\caption{\label{fig:power-com-cubic-b}power com cubic b}
-\end{figure}
+\includegraphics{paper_comparison_files/figure-latex/power-vs-log-effect-size-given-number-of-observations-1.pdf}
 
-\begin{figure}
-\centering
-\includegraphics{paper_comparison_files/figure-latex/power-com-heter-b-1.pdf}
-\caption{\label{fig:power-com-heter-b}power com heter b}
-\end{figure}
+\includegraphics{paper_comparison_files/figure-latex/power-of-visual-test-given-number-of-observations-1.pdf}
 

### 2022-07-06 paper_comparison/paper_comparison/paper_comparison.Rmd
update the intro

@@
 
-problem: residual plot diagnostics
-conventional test: too sensitive
+<!-- problem: residual plot diagnostics -->
+<!-- conventional test: too sensitive -->
 
-background:
+<!-- background:  -->
 
 <!-- week 1 -->
-1. residual plot for model diagnostics
+<!-- 1. residual plot for model diagnostics -->
 
-a. residual is widely used
-b. what are the types of residual plots
-c. comparison
+<!-- a. residual is widely used -->
+<!-- b. what are the types of residual plots -->
+<!-- c. comparison -->
 
 <!-- week 2 -->
-2. conventional test: F, BP
-3. visual test: lineup, theory
+<!-- 2. conventional test: F, BP -->
+<!-- 3. visual test: lineup, theory -->
 
 <!-- week 4 -->
-desc of experiment:
-1. simulation setup
-2. experimental design
-3. result
+<!-- desc of experiment: -->
+<!-- 1. simulation setup -->
+<!-- 2. experimental design -->
+<!-- 3. result -->
 
 <!-- week 3 -->
-comparison of conventional tests:
-1. power (visual test vs. conventional test)
-(visual test most different one (everything test, any departure))
-plot figure in a paper, desc, exp
-2. investigate the difference (gap), give examples
-3. conventional is too sensitive
-4. make conventional less sensitive (vary alpha)
+<!-- comparison of conventional tests: -->
+<!-- 1. power (visual test vs. conventional test) -->
+<!-- (visual test most different one (everything test, any departure)) -->
+<!-- plot figure in a paper, desc, exp -->
+<!-- 2. investigate the difference (gap), give examples -->
+<!-- 3. conventional is too sensitive -->
+<!-- 4. make conventional less sensitive (vary alpha) -->
 
 <!-- last week -->
-conclusion:
-1. too sensitive, visual test is needed/preferable
-2. visual test is infeasible in large scale (expensive)
-3. future work (role of computer vision)
+<!-- conclusion: -->
+<!-- 1. too sensitive, visual test is needed/preferable -->
+<!-- 2. visual test is infeasible in large scale (expensive) -->
+<!-- 3. future work (role of computer vision) -->
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
+```
+
+```{r}
+library(tidyverse)
+library(visage)
+
+set.seed(10086)
+```
 
@@
 
-Regression diagnostics conventionally involve evaluating the fitness of the proposed model, detecting the presence of influential observations and outliers, checking the validity of model assumptions and many more.
-Common diagnostic techniques including summary statistics, hypothesis testing, and data plots are essential tools for a systematic and detailed examination of the regression model [@mansfield1987diagnostic].
+Regression analysis is a field of study with at least a hundred years of history, and regression diagnostics is one of the essential steps in regression analysis.
+The diagnostic procedure conventionally involves evaluating the fitness of the proposed model, detecting the presence of influential observations and outliers, checking the validity of model assumptions and many more.
+In terms of diagnostic techniques, data plots, hypothesis testing, and summary statistics are vital tools for a systematic and detailed examination of the regression model [@mansfield1987diagnostic].
 
-## Diagnostic plots
+Many of those regression diagnostic methods and procedures are mature and well-established in books first published in the twentieth century, such as @draper_applied_2014, @montgomery_introduction_2012, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals.
+Regardless of the level of difficulty, one will find the importance and usefulness of diagnostic plots being emphasized in those books repeatedly.
+Checking diagnostic plots is also the recommended starting point for validating model assumptions such as normality, homoscedasticity and linearity [@anscombe_examination_1963].
 
-Regression analysis is a field of study with at least a hundred years of history.
-Many of those regression diagnostic methods and procedures are mature and well-established in books first published in the twentieth century, such as @draper_applied_2014, @montgomery_introduction_2012, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals.
-Regardless of the level of difficulty of the book, one will find the importance and usefulness of diagnostic plots being emphasized again and again.
-Checking diagnostic plots is also the recommended starting point for validating model assumptions like normality, homoscedasticity and linearity [@anscombe_examination_1963].
-
-Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are refereed to as standard residual plots.
-They are commonly used to identify patterns which are indicative of nonconstant error variance or non-linearity [@cook1982residuals].
+Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals.
+As suggested by @cook1982residuals, these kinds of diagnostic plots are commonly used to identify patterns that are indicative of nonconstant error variance or non-linearity.
 Raw residuals and studentized residuals are the two most frequently used residuals in standard residual plots.
-The debt on which type of residuals should be used always present.
-While raw residuals are the most common output of computer regression software package, by applying a scaling factor, the ability of revealing nonconstant error variance in standard residual plots will often be enhanced by studentized residuals in small sample size [@gunst2018regression].
-
+The debt on which type of residuals should be used is always present.
+While raw residuals are the most common computer regression software package output, by applying a scaling factor, the ability to reveal nonconstant error variance in standard residual plots will often be enhanced by studentized residuals in small sample size [@gunst2018regression].
 As a two-dimensional representation of a model in a $p$-dimensional space, standard residual plots project data points onto the variable of the horizontal axis, which is a vector in $p$-dimensional space.
-Observations with the same projection will be treated as equivalent as they have the same position of the abscissa.
-Therefore, standard residual plots are often useful in revealing model inadequacies in the direction of the variable of the horizontal axis, but could be inadequate for detecting patterns in other directions, especially in those perpendicular to the variable of the horizontal axis.
-Hence, in practice, multiple standard residual plots with different horizontal axes will be examined.
-
-<!-- large sample size-->
+Observations with the same projection will be treated as equivalent as they have the same abscissa.
+Therefore, standard residual plots are often useful in revealing model inadequacies in the direction of the variable of the horizontal axis but could be inadequate for detecting patterns in other directions, especially in those perpendicular to the variable of the horizontal axis.
+Hence, in practice, multiple standard residual plots with different horizontal axes will be examined [@cook1982residuals].
 Overlapping data points is a general issue in scatter plots not limited to standard residual plots, which often makes plots difficult to interpret because visual patterns are concealed.
-Thus, for relatively large sample size, @cleveland1975graphical suggests the use of robust moving statistics as reference lines to give aids to eye in seeing patterns, which nowadays, are usually replaced with a spline or local polynomial regression line.
+Thus, for a relatively large sample size, @cleveland1975graphical suggests the use of robust moving statistics as reference lines to give aid to the eye in seeing patterns, which nowadays, are usually replaced with a spline or local polynomial regression line.
 
@@
 
-## Hypothesis testing
+In addition to diagnostic plots, analysts may perform formal hypothesis testing for detecting model defects.
+Depends on the alternative hypothesis that being focused on, variety of tests can be applied.
+For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979], which both derived from the Lagrange multiplier test (ref here) principle and rely on the asymptotic properties of the null distribution.
+For testing non-linearity, one may apply the F-test (ref here) to examine the significance of certain polynomial and non-linear forms of the regressors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969].
 
-In addition to diagnostic plots, researcher may also perform formal tests for detecting model defects.
-Depends on the alternative, variety of tests can be applied.
-For example, for testing heteroskedasticity,
-one may use the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979].
-And for testing non-linearity, there are RESET test [@ramsey_tests_1969] and F-test.
-
-As discussed in @cook1982residuals, most residual based tests for a particular type of departures from model assumptions are sensitive to other types of departures.
-Especially, outliers will often  incorrectly trigger the rejection of null hypothesis despite the residuals are well-behaved [@cook_applied_1999].
+As discussed in @cook1982residuals, most residual-based tests for a particular type of departures from model assumptions are sensitive to other types of departures.
+It is likely the null hypothesis is correctly rejected but for the wrong reason, which is also known as the "Type III error".
+Additionally, outliers will often incorrectly trigger the rejection of null hypothesis despite the residuals are well-behaved [@cook_applied_1999].
 This can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers.
-Furthermore, @montgomery_introduction_2012 stated that based on their experience, statistical tests are not widely used in regression diagnostics.
-Most importantly, the same or even larger amount of information can be provided by diagnostic plots than the corresponding tests in most empirical studies.
+@montgomery_introduction_2012 stated that based on their experience, statistical tests are not widely used in regression diagnostics.
+The same or even larger amount of information can be provided by diagnostic plots than the corresponding tests in most empirical studies.
+Not to mention, it is almost impossible to have an exactly correctly specified model in reality.
+There is a well-known aphorism in statistic stated by George Box - "All models are wrong, but some are useful".
+This indicates proper hypothesis tests will always reject the null hypothesis as long as the sample size is large enough.
+The outcome "Not reject" can be interpreted as either "effect size is small" or "sample size is small".
+The outcome "reject" still doesn't inform us whether and how much the model defects are of actual consequence to the inference and prediction.
 But still, the effectiveness of statistical tests shall not be disrespected.
 Statistical tests have chance to provide analysts with unique information.
-There are also situations where no suitable diagnostic plots can be found for a particular violation of the assumptions, or excessive diagnostic plots need to be checked.
+There are situations where no suitable diagnostic plots can be found for a particular violation of the assumptions, or excessive diagnostic plots need to be checked.
 One will have no choice but to rely on statistical tests if there is any.
-A good regression diagnostic practice should be a combination of both methods.
-
-## Visual inference
+A good regression diagnostic practice should be a balanced combination of both methods.
 
@@
 # Experimental design
+
+# Demographic summary
+
+```{r}
+polynomials %>%
+  group_by(set) %>%
+  summarise(across(age_group:previous_experience, first)) %>%
+  mutate(education = fct_relevel(education, "High School or below")) %>%
+  ggplot() +
+  geom_bar(aes(y = education), orientation = "y") -> education_plot
+
+polynomials %>%
+  group_by(set) %>%
+  summarise(across(age_group:previous_experience, first)) %>%
+  ggplot() +
+  geom_bar(aes(age_group)) -> age_plot
+
+polynomials %>%
+  group_by(set) %>%
+  summarise(across(age_group:previous_experience, first)) %>%
+  mutate(pronoun = fct_relevel(pronoun, "He", "She", "They")) %>%
+  ggplot() +
+  geom_bar(aes(pronoun)) -> gender_plot
+
+polynomials %>%
+  group_by(set) %>%
+  summarise(across(age_group:previous_experience, first)) %>%
+  ggplot() +
+  geom_bar(aes(previous_experience)) -> experience_plot
+
+patchwork::wrap_plots(education_plot, age_plot, gender_plot, experience_plot)
+```
+
+<!-- # time taken vs order of the plot -->
+<!-- # accuracy vs order of the plot -->
+<!-- # num of selection vs order of the plot -->
+<!-- # order of the plot? -->
+<!-- # summary of the demographic information -->
 
@@
 From the figures, we can observe ...
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
-```
-
-```{r}
-library(tidyverse)
-library(visage)
-
-set.seed(10086)
-```
 

### 2022-07-06 paper_comparison/paper_comparison/paper_comparison.tex
update the intro

@@
 
-problem: residual plot diagnostics conventional test: too sensitive
-
-background:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  residual plot for model diagnostics
-\end{enumerate}
-
-\begin{enumerate}
-\def\labelenumi{\alph{enumi}.}
-\tightlist
-\item
-  residual is widely used
-\item
-  what are the types of residual plots
-\item
-  comparison
-\end{enumerate}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{1}
-\tightlist
-\item
-  conventional test: F, BP
-\item
-  visual test: lineup, theory
-\end{enumerate}
-
-desc of experiment: 1. simulation setup 2. experimental design 3. result
-
-comparison of conventional tests: 1. power (visual test vs.~conventional
-test) (visual test most different one (everything test, any departure))
-plot figure in a paper, desc, exp 2. investigate the difference (gap),
-give examples 3. conventional is too sensitive 4. make conventional less
-sensitive (vary alpha)
-
-conclusion: 1. too sensitive, visual test is needed/preferable 2. visual
-test is infeasible in large scale (expensive) 3. future work (role of
-computer vision)
-
 \hypertarget{introduction}{%
@@
 
-Regression diagnostics conventionally involve evaluating the fitness of
-the proposed model, detecting the presence of influential observations
-and outliers, checking the validity of model assumptions and many more.
-Common diagnostic techniques including summary statistics, hypothesis
-testing, and data plots are essential tools for a systematic and
-detailed examination of the regression model
+Regression analysis is a field of study with at least a hundred years of
+history, and regression diagnostics is one of the essential steps in
+regression analysis. The diagnostic procedure conventionally involves
+evaluating the fitness of the proposed model, detecting the presence of
+influential observations and outliers, checking the validity of model
+assumptions and many more. In terms of diagnostic techniques, data
+plots, hypothesis testing, and summary statistics are vital tools for a
+systematic and detailed examination of the regression model
 \citep{mansfield1987diagnostic}.
 
-\hypertarget{diagnostic-plots}{%
-\subsection{Diagnostic plots}\label{diagnostic-plots}}
-
-Regression analysis is a field of study with at least a hundred years of
-history. Many of those regression diagnostic methods and procedures are
-mature and well-established in books first published in the twentieth
-century, such as \citet{draper_applied_2014},
+Many of those regression diagnostic methods and procedures are mature
+and well-established in books first published in the twentieth century,
+such as \citet{draper_applied_2014},
 \citet{montgomery_introduction_2012}, \citet{belsley_regression_1980},
 \citet{cook_applied_1999} and \citet{cook1982residuals}. Regardless of
-the level of difficulty of the book, one will find the importance and
-usefulness of diagnostic plots being emphasized again and again.
-Checking diagnostic plots is also the recommended starting point for
-validating model assumptions like normality, homoscedasticity and
-linearity \citep{anscombe_examination_1963}.
+the level of difficulty, one will find the importance and usefulness of
+diagnostic plots being emphasized in those books repeatedly. Checking
+diagnostic plots is also the recommended starting point for validating
+model assumptions such as normality, homoscedasticity and linearity
+\citep{anscombe_examination_1963}.
 
@@
 or other functions of the predictor variables that are approximately
-orthogonal to residuals are refereed to as standard residual plots. They
-are commonly used to identify patterns which are indicative of
-nonconstant error variance or non-linearity \citep{cook1982residuals}.
-Raw residuals and studentized residuals are the two most frequently used
+orthogonal to residuals are referred to as standard residual plots in
+\citet{cook1982residuals}. As suggested by \citet{cook1982residuals},
+these kinds of diagnostic plots are commonly used to identify patterns
+that are indicative of nonconstant error variance or non-linearity. Raw
+residuals and studentized residuals are the two most frequently used
 residuals in standard residual plots. The debt on which type of
-residuals should be used always present. While raw residuals are the
-most common output of computer regression software package, by applying
-a scaling factor, the ability of revealing nonconstant error variance in
+residuals should be used is always present. While raw residuals are the
+most common computer regression software package output, by applying a
+scaling factor, the ability to reveal nonconstant error variance in
 standard residual plots will often be enhanced by studentized residuals
-in small sample size \citep{gunst2018regression}.
-
-As a two-dimensional representation of a model in a \(p\)-dimensional
-space, standard residual plots project data points onto the variable of
-the horizontal axis, which is a vector in \(p\)-dimensional space.
-Observations with the same projection will be treated as equivalent as
-they have the same position of the abscissa. Therefore, standard
-residual plots are often useful in revealing model inadequacies in the
-direction of the variable of the horizontal axis, but could be
-inadequate for detecting patterns in other directions, especially in
-those perpendicular to the variable of the horizontal axis. Hence, in
-practice, multiple standard residual plots with different horizontal
-axes will be examined.
-
-Overlapping data points is a general issue in scatter plots not limited
-to standard residual plots, which often makes plots difficult to
-interpret because visual patterns are concealed. Thus, for relatively
-large sample size, \citet{cleveland1975graphical} suggests the use of
-robust moving statistics as reference lines to give aids to eye in
-seeing patterns, which nowadays, are usually replaced with a spline or
-local polynomial regression line.
+in small sample size \citep{gunst2018regression}. As a two-dimensional
+representation of a model in a \(p\)-dimensional space, standard
+residual plots project data points onto the variable of the horizontal
+axis, which is a vector in \(p\)-dimensional space. Observations with
+the same projection will be treated as equivalent as they have the same
+abscissa. Therefore, standard residual plots are often useful in
+revealing model inadequacies in the direction of the variable of the
+horizontal axis but could be inadequate for detecting patterns in other
+directions, especially in those perpendicular to the variable of the
+horizontal axis. Hence, in practice, multiple standard residual plots
+with different horizontal axes will be examined
+\citep{cook1982residuals}. Overlapping data points is a general issue in
+scatter plots not limited to standard residual plots, which often makes
+plots difficult to interpret because visual patterns are concealed.
+Thus, for a relatively large sample size, \citet{cleveland1975graphical}
+suggests the use of robust moving statistics as reference lines to give
+aid to the eye in seeing patterns, which nowadays, are usually replaced
+with a spline or local polynomial regression line.
 
@@
 
-\hypertarget{hypothesis-testing}{%
-\subsection{Hypothesis testing}\label{hypothesis-testing}}
+In addition to diagnostic plots, analysts may perform formal hypothesis
+testing for detecting model defects. Depends on the alternative
+hypothesis that being focused on, variety of tests can be applied. For
+example, the presence of heteroskedasticity can usually be tested by
+applying the White test \citep{white_heteroskedasticity-consistent_1980}
+or the Breusch-Pagan test \citep{breusch_simple_1979}, which both
+derived from the Lagrange multiplier test (ref here) principle and rely
+on the asymptotic properties of the null distribution. For testing
+non-linearity, one may apply the F-test (ref here) to examine the
+significance of certain polynomial and non-linear forms of the
+regressors, or the significance of proxy variables as in the Ramsey
+Regression Equation Specification Error Test (RESET)
+\citep{ramsey_tests_1969}.
 
-In addition to diagnostic plots, researcher may also perform formal
-tests for detecting model defects. Depends on the alternative, variety
-of tests can be applied. For example, for testing heteroskedasticity,
-one may use the White test
-\citep{white_heteroskedasticity-consistent_1980} or the Breusch-Pagan
-test \citep{breusch_simple_1979}. And for testing non-linearity, there
-are RESET test \citep{ramsey_tests_1969} and F-test.
-
-As discussed in \citet{cook1982residuals}, most residual based tests for
+As discussed in \citet{cook1982residuals}, most residual-based tests for
 a particular type of departures from model assumptions are sensitive to
-other types of departures. Especially, outliers will often incorrectly
-trigger the rejection of null hypothesis despite the residuals are
-well-behaved \citep{cook_applied_1999}. This can be largely avoided in
-diagnostic plots as experienced analysts can evaluate the acceptability
-of assumptions flexibly, even in the presence of outliers. Furthermore,
+other types of departures. It is likely the null hypothesis is correctly
+rejected but for the wrong reason, which is also known as the ``Type III
+error''. Additionally, outliers will often incorrectly trigger the
+rejection of null hypothesis despite the residuals are well-behaved
+\citep{cook_applied_1999}. This can be largely avoided in diagnostic
+plots as experienced analysts can evaluate the acceptability of
+assumptions flexibly, even in the presence of outliers.
 \citet{montgomery_introduction_2012} stated that based on their
 experience, statistical tests are not widely used in regression
-diagnostics. Most importantly, the same or even larger amount of
-information can be provided by diagnostic plots than the corresponding
-tests in most empirical studies. But still, the effectiveness of
-statistical tests shall not be disrespected. Statistical tests have
-chance to provide analysts with unique information. There are also
-situations where no suitable diagnostic plots can be found for a
-particular violation of the assumptions, or excessive diagnostic plots
-need to be checked. One will have no choice but to rely on statistical
-tests if there is any. A good regression diagnostic practice should be a
-combination of both methods.
-
-\hypertarget{visual-inference}{%
-\subsection{Visual inference}\label{visual-inference}}
+diagnostics. The same or even larger amount of information can be
+provided by diagnostic plots than the corresponding tests in most
+empirical studies. Not to mention, it is almost impossible to have an
+exactly correctly specified model in reality. There is a well-known
+aphorism in statistic stated by George Box - ``All models are wrong, but
+some are useful''. This indicates proper hypothesis tests will always
+reject the null hypothesis as long as the sample size is large enough.
+The outcome ``Not reject'' can be interpreted as either ``effect size is
+small'' or ``sample size is small''. The outcome ``reject'' still
+doesn't inform us whether and how much the model defects are of actual
+consequence to the inference and prediction. But still, the
+effectiveness of statistical tests shall not be disrespected.
+Statistical tests have chance to provide analysts with unique
+information. There are situations where no suitable diagnostic plots can
+be found for a particular violation of the assumptions, or excessive
+diagnostic plots need to be checked. One will have no choice but to rely
+on statistical tests if there is any. A good regression diagnostic
+practice should be a balanced combination of both methods.
 
@@
 \section{Experimental design}\label{experimental-design}}
+
+\hypertarget{demographic-summary}{%
+\section{Demographic summary}\label{demographic-summary}}
+
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-2-1.pdf}
 

### 2022-07-07 paper_comparison/paper_comparison/paper_comparison.Rmd
update intro

@@
 keywords: |
-  visual inference; model diagnostics;
+  visual inference; regression diagnostics;
 header-includes: |
@@
 
+## Diagnostic plots
+
 Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals.
@@
 
+## Hypothesis testing
+
 In addition to diagnostic plots, analysts may perform formal hypothesis testing for detecting model defects.
@@
 
-As discussed in @cook1982residuals, most residual-based tests for a particular type of departures from model assumptions are sensitive to other types of departures.
+As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are sensitive to other types of departures.
 It is likely the null hypothesis is correctly rejected but for the wrong reason, which is also known as the "Type III error".
-Additionally, outliers will often incorrectly trigger the rejection of null hypothesis despite the residuals are well-behaved [@cook_applied_1999].
+Additionally, outliers will often incorrectly trigger the rejection of the null hypothesis despite the residuals being well-behaved [@cook_applied_1999].
 This can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers.
@@
 Not to mention, it is almost impossible to have an exactly correctly specified model in reality.
-There is a well-known aphorism in statistic stated by George Box - "All models are wrong, but some are useful".
+There is a well-known aphorism in statistics stated by George Box - "All models are wrong, but some are useful".
 This indicates proper hypothesis tests will always reject the null hypothesis as long as the sample size is large enough.
@@
 But still, the effectiveness of statistical tests shall not be disrespected.
-Statistical tests have chance to provide analysts with unique information.
+Statistical tests have a chance to provide analysts with unique information.
 There are situations where no suitable diagnostic plots can be found for a particular violation of the assumptions, or excessive diagnostic plots need to be checked.
-One will have no choice but to rely on statistical tests if there is any.
+One will have no choice but to rely on statistical tests if there are any.
 A good regression diagnostic practice should be a balanced combination of both methods.
 
-Diagnostic plots are
+## Visual inference
 
-<!-- However, unlike confirmatory data analysis built upon rigorous statistical procedures, e.g., hypothesis testing, visual diagnostics relies on graphical perception - human’s ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984], which is to some extent subjective.
+However, unlike hypothesis testing built upon rigorous statistical procedures, reading diagnostic plots relies on graphical perception - human’s ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984], which is to some extent subjective.
 Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
 One such example is finding an over-interpretation of the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurrence [@roy_chowdhury_using_2015].
--->
 
-<!-- Recently, a new branch of statical inference  -->
+Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualization" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots [@gelman_exploratory_2004].
+Later, in the Bayesian context, data plots was systematically considered as model diagnostics by taking advantage of the data simulated from the assumed statistical models [@gelman_bayesian_2003; @gelman_exploratory_2004].
 
-<!-- Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualization" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots [@gelman_exploratory_2004].
-Later, in the Bayesian context, data plots was systematically considered as model diagnostics by taking advantage of the data simulated from the assumed statistical models [@gelman_bayesian_2003; @gelman_exploratory_2004].
--->
+It was surprising that the essential components of visual inference had actually been established in @buja_inference_1999, but it was not until 10 years later that @buja_statistical_2009 formalized it as an inferential framework to extend confirmatory statistics to visual discoveries.
+This framework redefines the test statistics, tests, null distribution, significance levels and $p$-value for visual discovery modelled on the confirmatory statistical testing.
+Figure \@ref(fig:parallelism) outlines the parallelism between conventional tests and visual discovery.
+
+![(ref:parallelism) \label{fig:parallelism}](figures/rsta2009012001.jpg){width=450 height=341}
+
+(ref:parallelism) Parallelism between multiple quantitative testing and visual discovery [@buja_statistical_2009].
+Visible features in a plot are viewed as a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$, and any visual discoveries that are inconsistent with the null hypothesis are treated as evidence against the null.
+For regression diagnostics, the null hypothesis would be the assumed model, and visual discoveries could be any visual features in favour of any alternatives.
+
+In visual inference, a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$ is defined, where $\boldsymbol{\mathrm{y}}$ is the data and $I$ is a set of all possible visual features.
+@buja_statistical_2009 described each of the test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})$ as a measurement of the degree of presence of a visual feature.
+Alternatively, @majumder_validation_2013 avoids the use of visual features and defined the visual statistics $T(.)$ as a mapping from a dataset to a data plot.
+Both definitions of visual test statistics are valid, but in the rest of the report the first definition will be used as it covers some details needed by the following discussion.
+A visual discovery is defined as a rejection of a null hypothesis, and the same null hypothesis can be rejected by many different visual discoveries [@buja_statistical_2009].
+For regression diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the null hypothesis.
+The same regression model can be rejected by many reasons with residual plot, including non-linearity and heteroskedasticity as shown in Figure \@ref(fig:residual-plot-cubic-heter).
+
+```{r residual-plot-cubic-heter, fig.cap='Residuals vs.
+fitted values plot for a classical linear regression model.
+The residuals are produced by fitting a two-predictor multiple linear regression model with data generated from a cubic linear model.
+From the residual plot, "butterfly shape" can be observed which generally would be interpretd as evidence of heteroskedasticity.
+Further, from the outline of the shape, nonlinear patterns exist.
+Both visual discoveries are evidence against the null hypothesis, though heteroskedasticity actually does not exist in the data generating process.'}
+
+cm <- cubic_model(a = 0, b = 100, c = 1)
+
+cm$gen(1000, fit_model = TRUE) %>%
+cm$plot() +
+  theme_light()
+```
+
+### Sampling from the null distribution {#se:sampling-from-null}
+
+The null distribution of plots refers to the infinite collection of plots of null datasets sampled from $H_0$.
+It is defined as the analogue of the null distribution of test statistics in conventional test [@buja_statistical_2009].
+In practice, a finite number of plots of null datasets could be generated, called null plots.
+In the context of regression diagnostics, sampling data from $H_0$ is equivalent to sampling data from the assumed model.
+As @buja_statistical_2009 suggested, $H_0$ is usually composited by a collection of distributions controlled by nuisance parameters.
+Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling.
+
+The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009].
+Suppose there exists a minimal sufficient statistic $\boldsymbol{S}(\boldsymbol{y})$ under the null hypothesis, any null datasets $\boldsymbol{y^{*}}$ should fulfil the condition $\boldsymbol{S}(\boldsymbol{y}) = \boldsymbol{s}$.
+Using the classical normal linear regression model as example, the minimal sufficient statistic is $\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{\beta}}, \boldsymbol{e}'\boldsymbol{e})$, where $\hat{\boldsymbol{\beta}}$ are the coefficient estimators and $\boldsymbol{e}'\boldsymbol{e}$ is the residual sum of square.
+Alternatively, the minimal sufficient statistic can be constructed as $\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{y}}, ||\boldsymbol{e}||)$, where $\hat{\boldsymbol{y}}$ are the fitted values and $||\boldsymbol{e}||$ is the length of residuals, which is more intuitive as suggested by @buja_statistical_2009.
+Since the fitted values are held fixed, the variation can only occur in the residual space.
+And because the length of residual is also held fixed, residuals obtained from a null dataset has to be a random rotation of $\boldsymbol{e}$ in the residual space.
+With this property, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the regressors, then rescaling it by the ratio of residual sum of square in two regressions.
+
+### Lineup protocol {#se:lineup}
+
+With the simulation of null plots being provided, another aspect of hypothesis testing that needs to be addressed is the control of false positive rate or Type I error.
+Any visual statistic $T^{(i)}(\boldsymbol{\mathrm{y}})$ needs to pair with a critical value $c^{(i)}$ to form a hypothesis test.
+When a visual feature $i$ is discovered by the observer from a plot, the corresponding visual statistic $T^{(i)}(\boldsymbol{\mathrm{y}})$ may not be known as there is no general agreement on the measurement of the degree of presence of a visual feature.
+It is only the event that $T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}$ is confirmed.
+Similarly, if any visual discovery is found by the observer, we say, there exists $i \in I:~T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}$ [@buja_statistical_2009].
+
+Using the above definition, the family-wise Type I error can be controlled if one can provide the collection of critical values $c^{(i)}~(i \in I)$ such that $P(\mathrm{there~exists~} i \in I: T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}|\boldsymbol{\mathrm{y}}) \leq \alpha$, where $\alpha$ is the significance level.
+However, since the quantity of $T^{(i)}(\boldsymbol{\mathrm{y}})$ may not be known, such collection of critical values can not be provided.
+
+@buja_statistical_2009 proposed the lineup protocol as a visual test to calibrate the Type I error issue without the specification of $c^{(i)}~(i \in I)$.
+It is inspired by the "police lineup" or "identity parade" which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people.
+The protocol consists of $m$ randomly placed plots, where one plot is the actual data plot, and the remaining $m - 1$ plots have the identical graphical production as the data plot except the data has been replaced with data consistent with the null hypothesis.
+Then, an observer who have not seen the actual data plot will be asked to point out the most different plot from the lineup.
+
+Under the null hypothesis, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is $1/m$.
+If we reject the null hypothesis as the observer correctly picks the actual data plot, then the Type I error of this test is $1/m$.
+
+This provides us with an mechanism to control the Type I error, because $m$ - the number of plots in a lineup can be chosen.
+A larger value of $m$ will result in a smaller Type I error, but the limit to the value of $m$ depends on the number of plots a human is willing to view [@buja_statistical_2009].
+Typically, $m$ will be set to $20$ which is equivalent to set $\alpha = 0.05$, a general choice of significance level for conventional testing among statisticians.
+
+Further, if we involve $K$ independent observers in a visual test, and let $X$ be a random variable denoting the number of observers correctly picking the actual data plot.
+Then, under the null hypothesis $X \sim \mathrm{Binom}_{K,1/m}$, and therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as
+\begin{equation} \label{eq:pvaluesingle}
+P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},
+\end{equation}
+
+where $x$ is the realization of number of observers correctly picking the actual data plot [@majumder_validation_2013].
+
+The multiple individuals approach avoids the limit of $m$, while provides visual tests with $p$-value much smaller than $0.05$.
+In fact, the lower bound of $p$-value decreases exponentially as $K$ increases.
+With just $4$ individuals and $20$ data plots in a lineup, the $p$-value could be as small as $0.0001$.
+Additionally, by involving multiple observers, variation of individual ability to read plots can be addressed to some degree as different opinions about visual discoveries can be collected.
+
+Compared to the conventional test, whose power only depends on the parameter of interest $\theta$, several studies [see @hofmann_graphical_2012; @majumder_validation_2013; @majumder_human_2014; @roy_chowdhury_using_2015; @loy_variations_2016] have shown the power of the visual test is subject-specific.
+Thus, to be able to account for individual's ability, an individual is required to evaluate multiple lineups [@majumder_validation_2013].
+
+Suppose individuals have the same ability and a lineup has been evaluated by multiple individuals, under the alternative hypothesis, the estimated power for a lineup can be expressed as $\hat{p} = x/K$, the estimated probability of identifying the actual data plot from the lineup.
+If the individual skill needs to be taken into account, and $L$ lineups have been evaluated by $K$ individuals, @majumder_validation_2013 suggests that mixed effects logistic regression model can be fit as:
+
+$$g(p_{li}) = W_{li}\delta + Z_{li}\tau_{li},$$
+where $g(.)$ is the logit link function $g(p) = log(p)  - log(1-p)$; $0 \leq p \leq 1$.
+$W_{li}$, $1 \leq i \leq K$, $1 \leq l \leq L$, is the covariate matrix including lineup-specific elements and demographic information of individuals, and $\delta$ is a vector of parameters.
+$Z$ is the random effects matrix, and $\tau$ is a vector of variables follow $N(\boldsymbol{0},\sigma_{\tau}\boldsymbol{I}_{KL\times KL})$.
+
+Then, the estimated power for lineup $l$ and individual $i$ can be calculated as $\hat{p}_{li} = g^{-1}(W_{li}\hat{\delta} + Z_{li}\hat{\tau}_{li})$ [@majumder_validation_2013].
 

### 2022-07-07 paper_comparison/paper_comparison/paper_comparison.tex
update intro

@@
 \begin{keywords}
-visual inference; model diagnostics;
+visual inference; regression diagnostics;
 \end{keywords}
@@
 \citep{anscombe_examination_1963}.
+
+\hypertarget{diagnostic-plots}{%
+\subsection{Diagnostic plots}\label{diagnostic-plots}}
 
@@
 
+\hypertarget{hypothesis-testing}{%
+\subsection{Hypothesis testing}\label{hypothesis-testing}}
+
 In addition to diagnostic plots, analysts may perform formal hypothesis
@@
 As discussed in \citet{cook1982residuals}, most residual-based tests for
-a particular type of departures from model assumptions are sensitive to
+a particular type of departure from model assumptions are sensitive to
 other types of departures. It is likely the null hypothesis is correctly
@@
 error''. Additionally, outliers will often incorrectly trigger the
-rejection of null hypothesis despite the residuals are well-behaved
-\citep{cook_applied_1999}. This can be largely avoided in diagnostic
-plots as experienced analysts can evaluate the acceptability of
-assumptions flexibly, even in the presence of outliers.
+rejection of the null hypothesis despite the residuals being
+well-behaved \citep{cook_applied_1999}. This can be largely avoided in
+diagnostic plots as experienced analysts can evaluate the acceptability
+of assumptions flexibly, even in the presence of outliers.
 \citet{montgomery_introduction_2012} stated that based on their
@@
 exactly correctly specified model in reality. There is a well-known
-aphorism in statistic stated by George Box - ``All models are wrong, but
-some are useful''. This indicates proper hypothesis tests will always
-reject the null hypothesis as long as the sample size is large enough.
-The outcome ``Not reject'' can be interpreted as either ``effect size is
-small'' or ``sample size is small''. The outcome ``reject'' still
-doesn't inform us whether and how much the model defects are of actual
-consequence to the inference and prediction. But still, the
+aphorism in statistics stated by George Box - ``All models are wrong,
+but some are useful''. This indicates proper hypothesis tests will
+always reject the null hypothesis as long as the sample size is large
+enough. The outcome ``Not reject'' can be interpreted as either ``effect
+size is small'' or ``sample size is small''. The outcome ``reject''
+still doesn't inform us whether and how much the model defects are of
+actual consequence to the inference and prediction. But still, the
 effectiveness of statistical tests shall not be disrespected.
-Statistical tests have chance to provide analysts with unique
+Statistical tests have a chance to provide analysts with unique
 information. There are situations where no suitable diagnostic plots can
@@
 diagnostic plots need to be checked. One will have no choice but to rely
-on statistical tests if there is any. A good regression diagnostic
+on statistical tests if there are any. A good regression diagnostic
 practice should be a balanced combination of both methods.
 
-Diagnostic plots are
+\hypertarget{visual-inference}{%
+\subsection{Visual inference}\label{visual-inference}}
+
+However, unlike hypothesis testing built upon rigorous statistical
+procedures, reading diagnostic plots relies on graphical perception -
+human's ability to interpret and decode the information embedded in the
+graph \citep{cleveland_graphical_1984}, which is to some extent
+subjective. Further, visual discovery suffers from its unsecured and
+unconfirmed nature where the degree of the presence of the visual
+features typically can not be measured quantitatively and objectively,
+which may lead to over or under-interpretations of the data. One such
+example is finding an over-interpretation of the separation between gene
+groups in a two-dimensional projection from a linear discriminant
+analysis when in fact there are no differences in the expression levels
+between the gene groups and separation is not an uncommon occurrence
+\citep{roy_chowdhury_using_2015}.
+
+Visual inference was first introduced in a 1999 Joint Statistical
+Meetings (JSM) talk with the title ``Inference for Data Visualization''
+by \citet{buja_inference_1999} as an idea to address the issue of valid
+inference for visual discoveries of data plots
+\citep{gelman_exploratory_2004}. Later, in the Bayesian context, data
+plots was systematically considered as model diagnostics by taking
+advantage of the data simulated from the assumed statistical models
+\citep{gelman_bayesian_2003, gelman_exploratory_2004}.
+
+It was surprising that the essential components of visual inference had
+actually been established in \citet{buja_inference_1999}, but it was not
+until 10 years later that \citet{buja_statistical_2009} formalized it as
+an inferential framework to extend confirmatory statistics to visual
+discoveries. This framework redefines the test statistics, tests, null
+distribution, significance levels and \(p\)-value for visual discovery
+modelled on the confirmatory statistical testing. Figure
+@ref(fig:parallelism) outlines the parallelism between conventional
+tests and visual discovery.
+
+\begin{figure}
+\centering
+\includegraphics[width=4.6875in,height=3.55208in]{figures/rsta2009012001.jpg}
+\caption{(ref:parallelism) \label{fig:parallelism}}
+\end{figure}
+
+(ref:parallelism) Parallelism between multiple quantitative testing and
+visual discovery \citep{buja_statistical_2009}. Visible features in a
+plot are viewed as a collection of test statistics
+\(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\), and any visual
+discoveries that are inconsistent with the null hypothesis are treated
+as evidence against the null. For regression diagnostics, the null
+hypothesis would be the assumed model, and visual discoveries could be
+any visual features in favour of any alternatives.
+
+In visual inference, a collection of test statistics
+\(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) is defined, where
+\(\boldsymbol{\mathrm{y}}\) is the data and \(I\) is a set of all
+possible visual features. \citet{buja_statistical_2009} described each
+of the test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})\) as a
+measurement of the degree of presence of a visual feature.
+Alternatively, \citet{majumder_validation_2013} avoids the use of visual
+features and defined the visual statistics \(T(.)\) as a mapping from a
+dataset to a data plot. Both definitions of visual test statistics are
+valid, but in the rest of the report the first definition will be used
+as it covers some details needed by the following discussion. A visual
+discovery is defined as a rejection of a null hypothesis, and the same
+null hypothesis can be rejected by many different visual discoveries
+\citep{buja_statistical_2009}. For regression diagnostics, the null
+hypothesis would be the assumed model, while the visual discoveries
+would be any findings that are inconsistent with the null hypothesis.
+The same regression model can be rejected by many reasons with residual
+plot, including non-linearity and heteroskedasticity as shown in Figure
+@ref(fig:residual-plot-cubic-heter).
+
+\begin{figure}
+\centering
+\includegraphics{paper_comparison_files/figure-latex/residual-plot-cubic-heter-1.pdf}
+\caption{Residuals vs.~fitted values plot for a classical linear
+regression model. The residuals are produced by fitting a two-predictor
+multiple linear regression model with data generated from a cubic linear
+model. From the residual plot, ``butterfly shape'' can be observed which
+generally would be interpretd as evidence of heteroskedasticity.
+Further, from the outline of the shape, nonlinear patterns exist. Both
+visual discoveries are evidence against the null hypothesis, though
+heteroskedasticity actually does not exist in the data generating
+process.}
+\end{figure}
+
+\hypertarget{se:sampling-from-null}{%
+\subsubsection{Sampling from the null
+distribution}\label{se:sampling-from-null}}
+
+The null distribution of plots refers to the infinite collection of
+plots of null datasets sampled from \(H_0\). It is defined as the
+analogue of the null distribution of test statistics in conventional
+test \citep{buja_statistical_2009}. In practice, a finite number of
+plots of null datasets could be generated, called null plots. In the
+context of regression diagnostics, sampling data from \(H_0\) is
+equivalent to sampling data from the assumed model. As
+\citet{buja_statistical_2009} suggested, \(H_0\) is usually composited
+by a collection of distributions controlled by nuisance parameters.
+Since regression models can have various forms, there is no general
+solution to this problem, but it sometimes can be reduced to so called
+``reference distribution'' by applying one of the three methods: (i)
+sampling from a conditional distribution given a minimal sufficient
+statistic under \(H_0\), (ii) parametric bootstrap sampling with
+nuisance parameters estimated under \(H_0\), and (iii) Bayesian
+posterior predictive sampling.
+
+The conditional distribution given a minimal sufficient statistic is the
+best justified reference distribution among the three
+\citep{buja_statistical_2009}. Suppose there exists a minimal sufficient
+statistic \(\boldsymbol{S}(\boldsymbol{y})\) under the null hypothesis,
+any null datasets \(\boldsymbol{y^{*}}\) should fulfil the condition
+\(\boldsymbol{S}(\boldsymbol{y}) = \boldsymbol{s}\). Using the classical
+normal linear regression model as example, the minimal sufficient
+statistic is
+\(\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{\beta}}, \boldsymbol{e}'\boldsymbol{e})\),
+where \(\hat{\boldsymbol{\beta}}\) are the coefficient estimators and
+\(\boldsymbol{e}'\boldsymbol{e}\) is the residual sum of square.
+Alternatively, the minimal sufficient statistic can be constructed as
+\(\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{y}}, ||\boldsymbol{e}||)\),
+where \(\hat{\boldsymbol{y}}\) are the fitted values and
+\(||\boldsymbol{e}||\) is the length of residuals, which is more
+intuitive as suggested by \citet{buja_statistical_2009}. Since the
+fitted values are held fixed, the variation can only occur in the
+residual space. And because the length of residual is also held fixed,
+residuals obtained from a null dataset has to be a random rotation of
+\(\boldsymbol{e}\) in the residual space. With this property, null
+residuals can be simulated by regressing \(N\) i.i.d standard normal
+random draws on the regressors, then rescaling it by the ratio of
+residual sum of square in two regressions.
+
+\hypertarget{se:lineup}{%
+\subsubsection{Lineup protocol}\label{se:lineup}}
+
+With the simulation of null plots being provided, another aspect of
+hypothesis testing that needs to be addressed is the control of false
+positive rate or Type I error. Any visual statistic
+\(T^{(i)}(\boldsymbol{\mathrm{y}})\) needs to pair with a critical value
+\(c^{(i)}\) to form a hypothesis test. When a visual feature \(i\) is
+discovered by the observer from a plot, the corresponding visual
+statistic \(T^{(i)}(\boldsymbol{\mathrm{y}})\) may not be known as there
+is no general agreement on the measurement of the degree of presence of
+a visual feature. It is only the event that
+\(T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}\) is confirmed. Similarly,
+if any visual discovery is found by the observer, we say, there exists
+\(i \in I:~T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}\)
+\citep{buja_statistical_2009}.
+
+Using the above definition, the family-wise Type I error can be
+controlled if one can provide the collection of critical values
+\(c^{(i)}~(i \in I)\) such that
+\(P(\mathrm{there~exists~} i \in I: T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}|\boldsymbol{\mathrm{y}}) \leq \alpha\),
+where \(\alpha\) is the significance level. However, since the quantity
+of \(T^{(i)}(\boldsymbol{\mathrm{y}})\) may not be known, such
+collection of critical values can not be provided.
+
+\citet{buja_statistical_2009} proposed the lineup protocol as a visual
+test to calibrate the Type I error issue without the specification of
+\(c^{(i)}~(i \in I)\). It is inspired by the ``police lineup'' or
+``identity parade'' which is the act of asking the eyewitness to
+identify criminal suspect from a group of irrelevant people. The
+protocol consists of \(m\) randomly placed plots, where one plot is the
+actual data plot, and the remaining \(m - 1\) plots have the identical
+graphical production as the data plot except the data has been replaced
+with data consistent with the null hypothesis. Then, an observer who
+have not seen the actual data plot will be asked to point out the most
+different plot from the lineup.
+
+Under the null hypothesis, it is expected that the actual data plot
+would have no distinguishable difference with the null plots, and the
+probability of the observer correctly picks the actual data plot is
+\(1/m\). If we reject the null hypothesis as the observer correctly
+picks the actual data plot, then the Type I error of this test is
+\(1/m\).
+
+This provides us with an mechanism to control the Type I error, because
+\(m\) - the number of plots in a lineup can be chosen. A larger value of
+\(m\) will result in a smaller Type I error, but the limit to the value
+of \(m\) depends on the number of plots a human is willing to view
+\citep{buja_statistical_2009}. Typically, \(m\) will be set to \(20\)
+which is equivalent to set \(\alpha = 0.05\), a general choice of
+significance level for conventional testing among statisticians.
+
+Further, if we involve \(K\) independent observers in a visual test, and
+let \(X\) be a random variable denoting the number of observers
+correctly picking the actual data plot. Then, under the null hypothesis
+\(X \sim \mathrm{Binom}_{K,1/m}\), and therefore, the \(p\)-value of a
+lineup of size \(m\) evaluated by \(K\) observer is given as
+\begin{equation} \label{eq:pvaluesingle}
+P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},
+\end{equation}
+
+where \(x\) is the realization of number of observers correctly picking
+the actual data plot \citep{majumder_validation_2013}.
+
+The multiple individuals approach avoids the limit of \(m\), while
+provides visual tests with \(p\)-value much smaller than \(0.05\). In
+fact, the lower bound of \(p\)-value decreases exponentially as \(K\)
+increases. With just \(4\) individuals and \(20\) data plots in a
+lineup, the \(p\)-value could be as small as \(0.0001\). Additionally,
+by involving multiple observers, variation of individual ability to read
+plots can be addressed to some degree as different opinions about visual
+discoveries can be collected.
+
+Compared to the conventional test, whose power only depends on the
+parameter of interest \(\theta\), several studies
+\citep[see][]{hofmann_graphical_2012, majumder_validation_2013, majumder_human_2014, roy_chowdhury_using_2015, loy_variations_2016}
+have shown the power of the visual test is subject-specific. Thus, to be
+able to account for individual's ability, an individual is required to
+evaluate multiple lineups \citep{majumder_validation_2013}.
+
+Suppose individuals have the same ability and a lineup has been
+evaluated by multiple individuals, under the alternative hypothesis, the
+estimated power for a lineup can be expressed as \(\hat{p} = x/K\), the
+estimated probability of identifying the actual data plot from the
+lineup. If the individual skill needs to be taken into account, and
+\(L\) lineups have been evaluated by \(K\) individuals,
+\citet{majumder_validation_2013} suggests that mixed effects logistic
+regression model can be fit as:
+
+\[g(p_{li}) = W_{li}\delta + Z_{li}\tau_{li},\] where \(g(.)\) is the
+logit link function \(g(p) = log(p) - log(1-p)\); \(0 \leq p \leq 1\).
+\(W_{li}\), \(1 \leq i \leq K\), \(1 \leq l \leq L\), is the covariate
+matrix including lineup-specific elements and demographic information of
+individuals, and \(\delta\) is a vector of parameters. \(Z\) is the
+random effects matrix, and \(\tau\) is a vector of variables follow
+\(N(\boldsymbol{0},\sigma_{\tau}\boldsymbol{I}_{KL\times KL})\).
+
+Then, the estimated power for lineup \(l\) and individual \(i\) can be
+calculated as
+\(\hat{p}_{li} = g^{-1}(W_{li}\hat{\delta} + Z_{li}\hat{\tau}_{li})\)
+\citep{majumder_validation_2013}.
 

### 2022-07-07 paper_comparison/paper_comparison/paper_comparison.Rmd
resolve citation

@@
 Depends on the alternative hypothesis that being focused on, variety of tests can be applied.
-For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979], which both derived from the Lagrange multiplier test (ref here) principle and rely on the asymptotic properties of the null distribution.
-For testing non-linearity, one may apply the F-test (ref here) to examine the significance of certain polynomial and non-linear forms of the regressors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969].
+For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979], which both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle and rely on the asymptotic properties of the null distribution.
+For testing non-linearity, one may apply the F-test to examine the significance of certain polynomial and non-linear forms of the regressors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969].
 
@@
 This framework redefines the test statistics, tests, null distribution, significance levels and $p$-value for visual discovery modelled on the confirmatory statistical testing.
-Figure \@ref(fig:parallelism) outlines the parallelism between conventional tests and visual discovery.
+Figure \ref{fig:parallelism} outlines the parallelism between conventional tests and visual discovery.
 
-![(ref:parallelism) \label{fig:parallelism}](figures/rsta2009012001.jpg){width=450 height=341}
-
-(ref:parallelism) Parallelism between multiple quantitative testing and visual discovery [@buja_statistical_2009].
+![Parallelism between multiple quantitative testing and visual discovery [@buja_statistical_2009].
 Visible features in a plot are viewed as a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$, and any visual discoveries that are inconsistent with the null hypothesis are treated as evidence against the null.
 For regression diagnostics, the null hypothesis would be the assumed model, and visual discoveries could be any visual features in favour of any alternatives.
+\label{fig:parallelism}](figures/rsta2009012001.jpg){width=450 height=341}
 

### 2022-07-07 paper_comparison/paper_comparison/paper_comparison.tex
resolve citation

@@
 or the Breusch-Pagan test \citep{breusch_simple_1979}, which both
-derived from the Lagrange multiplier test (ref here) principle and rely
-on the asymptotic properties of the null distribution. For testing
-non-linearity, one may apply the F-test (ref here) to examine the
-significance of certain polynomial and non-linear forms of the
-regressors, or the significance of proxy variables as in the Ramsey
+derived from the Lagrange multiplier test \citep{silvey1959lagrangian}
+principle and rely on the asymptotic properties of the null
+distribution. For testing non-linearity, one may apply the F-test to
+examine the significance of certain polynomial and non-linear forms of
+the regressors, or the significance of proxy variables as in the Ramsey
 Regression Equation Specification Error Test (RESET)
@@
 modelled on the confirmatory statistical testing. Figure
-@ref(fig:parallelism) outlines the parallelism between conventional
+\ref{fig:parallelism} outlines the parallelism between conventional
 tests and visual discovery.
@@
 \includegraphics[width=4.6875in,height=3.55208in]{figures/rsta2009012001.jpg}
-\caption{(ref:parallelism) \label{fig:parallelism}}
-\end{figure}
-
-(ref:parallelism) Parallelism between multiple quantitative testing and
-visual discovery \citep{buja_statistical_2009}. Visible features in a
-plot are viewed as a collection of test statistics
+\caption{Parallelism between multiple quantitative testing and visual
+discovery \citep{buja_statistical_2009}. Visible features in a plot are
+viewed as a collection of test statistics
 \(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\), and any visual
@@
 any visual features in favour of any alternatives.
+\label{fig:parallelism}}
+\end{figure}
 

### 2022-07-27 paper_comparison/paper_comparison/paper_comparison.Rmd
udpate paper

@@
 title: |
-  Why shouldn't you use numerical tests to diagnose the linear regression models?
+  Effectiveness of visual test in linear regression diagnositics  
 type: ARTICLE TEMPLATE
@@
   \def\tightlist{}
+  \newcommand*{\Perm}[2]{{}^{#1}\!P_{#2}}%
 output: rticles::tf_article
@@
 
-Regression analysis is a field of study with at least a hundred years of history, and regression diagnostics is one of the essential steps in regression analysis.
+Regression diagnostics is an essential step in regression analysis which is a field of study with at least a hundred years of history.
 The diagnostic procedure conventionally involves evaluating the fitness of the proposed model, detecting the presence of influential observations and outliers, checking the validity of model assumptions and many more.
@@
 
-Many of those regression diagnostic methods and procedures are mature and well-established in books first published in the twentieth century, such as @draper_applied_2014, @montgomery_introduction_2012, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals.
+Many of those regression diagnostic methods and procedures are mature and well-established in books first published in the twentieth century, such as @draper1998applied, @montgomery1982introduction, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals.
 Regardless of the level of difficulty, one will find the importance and usefulness of diagnostic plots being emphasized in those books repeatedly.
@@
 The debt on which type of residuals should be used is always present.
-While raw residuals are the most common computer regression software package output, by applying a scaling factor, the ability to reveal nonconstant error variance in standard residual plots will often be enhanced by studentized residuals in small sample size [@gunst2018regression].
+While raw residuals are the most common computer regression software package output, by applying a scaling factor, the ability to reveal nonconstant error variance in standard residual plots will often be enhanced by studentized residuals in small sample size [@gunst1980regression].
 As a two-dimensional representation of a model in a $p$-dimensional space, standard residual plots project data points onto the variable of the horizontal axis, which is a vector in $p$-dimensional space.
 Observations with the same projection will be treated as equivalent as they have the same abscissa.
-Therefore, standard residual plots are often useful in revealing model inadequacies in the direction of the variable of the horizontal axis but could be inadequate for detecting patterns in other directions, especially in those perpendicular to the variable of the horizontal axis.
+Consequently, standard residual plots are often useful in revealing model inadequacies in the direction of the variable of the horizontal axis but could be inadequate for detecting patterns in other directions, especially in those perpendicular to the variable of the horizontal axis.
 Hence, in practice, multiple standard residual plots with different horizontal axes will be examined [@cook1982residuals].
-Overlapping data points is a general issue in scatter plots not limited to standard residual plots, which often makes plots difficult to interpret because visual patterns are concealed.
+Overlapping data points is another general issue in scatter plots not limited to standard residual plots, which often makes plots difficult to interpret because visual patterns are concealed.
 Thus, for a relatively large sample size, @cleveland1975graphical suggests the use of robust moving statistics as reference lines to give aid to the eye in seeing patterns, which nowadays, are usually replaced with a spline or local polynomial regression line.
@@
 
-In addition to diagnostic plots, analysts may perform formal hypothesis testing for detecting model defects.
-Depends on the alternative hypothesis that being focused on, variety of tests can be applied.
-For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979], which both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle and rely on the asymptotic properties of the null distribution.
-For testing non-linearity, one may apply the F-test to examine the significance of certain polynomial and non-linear forms of the regressors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969].
+Other than checking diagnostic plots, analysts may perform formal hypothesis testing for detecting model defects.
+Depending on the alternative hypothesis that is focused on, a variety of tests can be applied.
+For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution.
+For testing non-linearity, one may apply the F-test to examine the significance of specific polynomial and non-linear forms of the regressors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969].
 
 As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are sensitive to other types of departures.
-It is likely the null hypothesis is correctly rejected but for the wrong reason, which is also known as the "Type III error".
-Additionally, outliers will often incorrectly trigger the rejection of the null hypothesis despite the residuals being well-behaved [@cook_applied_1999].
+It is likely the null hypothesis is correctly rejected but for the wrong reason, which is known as the "Type III error".
+Additionally, outliers will often incorrectly trigger the rejection of the null hypothesis despite the residuals are well-behaved [@cook_applied_1999].
 This can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers.
-@montgomery_introduction_2012 stated that based on their experience, statistical tests are not widely used in regression diagnostics.
+@montgomery1982introduction suggested that based on their experience, statistical tests are not widely used in regression diagnostics.
 The same or even larger amount of information can be provided by diagnostic plots than the corresponding tests in most empirical studies.
@@
 The outcome "Not reject" can be interpreted as either "effect size is small" or "sample size is small".
-The outcome "reject" still doesn't inform us whether and how much the model defects are of actual consequence to the inference and prediction.
+and the outcome "reject" doesn't inform us whether and how much the model defects are of actual consequence to the inference and prediction.
 But still, the effectiveness of statistical tests shall not be disrespected.
 Statistical tests have a chance to provide analysts with unique information.
-There are situations where no suitable diagnostic plots can be found for a particular violation of the assumptions, or excessive diagnostic plots need to be checked.
-One will have no choice but to rely on statistical tests if there are any.
+In situations where no suitable diagnostic plots can be found for a particular violation of the assumptions, or excessive diagnostic plots are needed to be checked, one will have no choice but to fall back on statistical tests if there exists any.
 A good regression diagnostic practice should be a balanced combination of both methods.
@@
 
-However, unlike hypothesis testing built upon rigorous statistical procedures, reading diagnostic plots relies on graphical perception - human’s ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984], which is to some extent subjective.
+However, unlike hypothesis testing built upon rigorous statistical procedures, reading diagnostic plots relies on graphical perception - human’s ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984], which is to some extent subjective and indecisive.
 Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
@@
 Alternatively, @majumder_validation_2013 avoids the use of visual features and defined the visual statistics $T(.)$ as a mapping from a dataset to a data plot.
-Both definitions of visual test statistics are valid, but in the rest of the report the first definition will be used as it covers some details needed by the following discussion.
+Both definitions of visual test statistics are valid, but in the rest of the paper the first definition will be used as it covers some details needed by the following discussion.
 A visual discovery is defined as a rejection of a null hypothesis, and the same null hypothesis can be rejected by many different visual discoveries [@buja_statistical_2009].
 For regression diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the null hypothesis.
-The same regression model can be rejected by many reasons with residual plot, including non-linearity and heteroskedasticity as shown in Figure \@ref(fig:residual-plot-cubic-heter).
+The same regression model can be rejected by many reasons with residual plot, including non-linearity and heteroskedasticity as shown in Figure \ref{fig:residual-plot-cubic-heter}.
 
@@
 Further, from the outline of the shape, nonlinear patterns exist.
-Both visual discoveries are evidence against the null hypothesis, though heteroskedasticity actually does not exist in the data generating process.'}
+Both visual discoveries are evidence against the null hypothesis, though heteroskedasticity actually does not exist in the data generating process.
+\\label{fig:residual-plot-cubic-heter}'}
 
@@
 
-With the simulation of null plots being provided, another aspect of hypothesis testing that needs to be addressed is the control of false positive rate or Type I error.
+With the simulation mechanism of null plots being provided, another aspect of hypothesis testing that needs to be addressed is the control of false positive rate or Type I error.
 Any visual statistic $T^{(i)}(\boldsymbol{\mathrm{y}})$ needs to pair with a critical value $c^{(i)}$ to form a hypothesis test.
@@
 
-Further, if we involve $K$ independent observers in a visual test, and let $X$ be a random variable denoting the number of observers correctly picking the actual data plot.
-Then, under the null hypothesis $X \sim \mathrm{Binom}_{K,1/m}$, and therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as
+Further, a visual test can involve $K$ independent observers.
+Let $D_i = \{0,1\}$ be a binomial random variable denoting whether subject $i$ correctly detecting the actual data plot, and $X = \sum_{i=1}^{K}X_i$ be the number of observers correctly picking the actual data plot.
+Then, by imposing a relatively strong assumption on the visual test that all $K$ evaluations are fully independent, under the null hypothesis, $X \sim \mathrm{Binom}_{K,1/m}$.
+Therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as
 \begin{equation} \label{eq:pvaluesingle}
@@
 
-Compared to the conventional test, whose power only depends on the parameter of interest $\theta$, several studies [see @hofmann_graphical_2012; @majumder_validation_2013; @majumder_human_2014; @roy_chowdhury_using_2015; @loy_variations_2016] have shown the power of the visual test is subject-specific.
+As pointed out by @vanderplas2021statistical, though equation (\ref{eq:pvaluesingle}) is trivial, but it doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup.
+And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup.
+They summarized three common different scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same actual data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects.
+Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experimental design.
+For Scenario 3, @vanderplas2021statistical modelled the probability of a plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$.
+And defined $c_i$ to be the number of times plot $i$ being selected in $K$ evaluations.
+In case subject $j$ makes multiple selections, they decided to add $1/s_j$ to $c_i$ instead of one, where $s_j$ is the number of plots subject $j$ selected for $j=1,...K$.
+This ensured $\sum_{i}c_i=K$.
+
+The full model was a Dirichlet-multinomial mixture distribution
+
+\begin{align} \label{eq:dirichlet-multinomial}\begin{split}
+\boldsymbol{\theta}&|\alpha \sim Dirichlet(\alpha)\\
+(c_i,...,c_m)&|\boldsymbol{\theta} \sim Multinomial(K, \boldsymbol{\theta}).
+\end{split}\end{align}
+
+Since the p-value calculation only needs to concern the number of times the actual data plot being selected denoted by $C_i$, they showed the model can be simplified to a beta-binomial mixture distribution
+
+\begin{align} \label{eq:beta-binomial}\begin{split}
+\theta_i&|\alpha \sim Beta(\alpha, (m-1)\alpha)\\
+C_i&|\theta_i \sim Binomial(K, \theta_i).
+\end{split}\end{align}
+
+Thus, the visual p-value followed by the beta-binomial model is given as
+
+\begin{equation} \label{eq:pvalue-beta-binomial}
+P(C \geq c_i) = \sum_{x=c_i}^{K}{{K}\choose{x}}\frac{1}{B(\alpha, (m-1)\alpha)}B(x + \alpha, K - x + (m - 1)\alpha),
+\end{equation}
+
+where $B(.)$ is the beta function defined as
+
+\begin{equation} \label{eq:betafunction}
+B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0. 
+\end{equation}
+
+Note that the use of equation (\ref{eq:pvalue-beta-binomial}) requires the estimation of $\hat{\alpha}$, which largely depends on the null model, the type of the plot and other aesthetic features.
+They suggested to estimate $\hat{\alpha}$ visually based on the selections of null plots of the experimental data, or to estimate $\hat{\alpha}$ numerically based on several additional Rorschach lineups, which is a type of lineup containing only null plots.
+However, when the number of null models are large, it could be expensive to manually estimate each $\alpha$ or include additional Rorschach lineups in the experiment.
+
+Instead, in the experiments that will be described in section \ref{experimental-design}, we adopt a simpler model implicitly used by the `pmulti()` function of the `vinference` `R` package.
+We assume the attractiveness of the plot $i$ modelled as $w_i \sim Uniform(0,1)$ for $i=1,..,m$.
+Let $\theta_i = w_i/\sum_{i=1}^{m}w_i$ be the probability of plot $i$ being selected by a subject.
+Then, given the number of selections $s_j$, for $j=1,...,K$, the distribution of $C_i$ can be approximated by simulating the random selection process with computer.
+The simulated visual test p-value is formulated as
+
+\begin{equation} \label{eq:p-value-multi}
+\text{p-value} = \frac{\#draws~that~the~actual~data~plota~being~selected~more~than~c_i~times}{\#simulation}.
+\end{equation}
+
+<!-- ### Effectiveness of visual inference in regression diagnostics -->
+
+<!-- Compared to the conventional test, whose power only depends on the parameter of interest $\theta$, several studies [see @hofmann_graphical_2012; @majumder_validation_2013; @majumder2014human; @roy_chowdhury_using_2015; @loy2016variations] have shown the power of the visual test is subject-specific.
 Thus, to be able to account for individual's ability, an individual is required to evaluate multiple lineups [@majumder_validation_2013].
+-->
 
-Suppose individuals have the same ability and a lineup has been evaluated by multiple individuals, under the alternative hypothesis, the estimated power for a lineup can be expressed as $\hat{p} = x/K$, the estimated probability of identifying the actual data plot from the lineup.
-If the individual skill needs to be taken into account, and $L$ lineups have been evaluated by $K$ individuals, @majumder_validation_2013 suggests that mixed effects logistic regression model can be fit as:
+<!-- Suppose individuals have the same ability and a lineup has been evaluated by multiple individuals, under the alternative hypothesis, the estimated power for a lineup can be expressed as $\hat{p} = x/K$, the estimated probability of identifying the actual data plot from the lineup.
+If the individual skill needs to be taken into account, and $L$ lineups have been evaluated by $K$ individuals, @majumder_validation_2013 suggests that mixed effects logistic regression model can be fit as: -->
 
-$$g(p_{li}) = W_{li}\delta + Z_{li}\tau_{li},$$
-where $g(.)$ is the logit link function $g(p) = log(p)  - log(1-p)$; $0 \leq p \leq 1$.
+<!-- $$g(p_{li}) = W_{li}\delta + Z_{li}\tau_{li},$$ -->
+<!-- where $g(.)$ is the logit link function $g(p) = log(p)  - log(1-p)$; $0 \leq p \leq 1$.
 $W_{li}$, $1 \leq i \leq K$, $1 \leq l \leq L$, is the covariate matrix including lineup-specific elements and demographic information of individuals, and $\delta$ is a vector of parameters.
 $Z$ is the random effects matrix, and $\tau$ is a vector of variables follow $N(\boldsymbol{0},\sigma_{\tau}\boldsymbol{I}_{KL\times KL})$.
+-->
 
-Then, the estimated power for lineup $l$ and individual $i$ can be calculated as $\hat{p}_{li} = g^{-1}(W_{li}\hat{\delta} + Z_{li}\hat{\tau}_{li})$ [@majumder_validation_2013].
+<!-- Then, the estimated power for lineup $l$ and individual $i$ can be calculated as $\hat{p}_{li} = g^{-1}(W_{li}\hat{\delta} + Z_{li}\hat{\tau}_{li})$ [@majumder_validation_2013]. -->
 
 # Experimental design
+
+The effectiveness of visual inference has already been validated by @majumder_validation_2013 under relatively simple classical normal linear regression model settings with only one or two regressors.
+Their results suggests visual test is capable of testing the significance of a single regressor with a similar power as a t-test, though they expressed that in general it is unnecessary to use visual inference if there exists a conventional test and they didn't expect the visual test to perform equally well as the conventional test.
+In their third experiment, where there does not exist a proper conventional test, visual test outperforms the conventional test for a large margin.
+This is encouraging as it promotes the use of visual inference in border field of data science where there are no existing statistical testing procedures.
+In fact, lineup protocol has been integrated into some model diagnostic tools such as @loy2013diagnostic.
+
+With our knowledge, what haven't been examined so far is the effectiveness of visual test relative to the equivalent conventional test in regression diagnostics.
+Particularly, its ability to detect non-linearity and heteroskedasticity compared to F-test and BP-test.
+
+Two experiments were conducted to evaluate the effectiveness of visual test relative to the equivalent conventional test in regression diagnostics.
+The first experiment has ideal scenario for conventional testing, where we would not expect the visual test to do better than the conventional test.
+The second experiment is a scenario where the conventional test is an approximate test, which visual test may have a chance to match the performance of the conventional test.
+
+As suggested by many previous studies [many studies refs here], it is clear some care is needed to set up the human subjects experiments.
+It is best for an subject to evaluate a block of 20 lineups with varying difficulty, with a reasonable number of "easy" lineups.
+The explanations about each experiment includes an explantion of how the lineups were sampled and provided to the subjects.
+
+Participants for all the experiments were recruited through Prolific's online web service [ref here].
+A summary of the data obtained for all three eperiments are shown in [table ref here].
+Participants were asked to select the plots which are mostly different from others, provide a reason for their selections, and express how different they think the selected plots are from others.
+Gender, age, education, and previous experience in visual experiment are also collected.
+For each of the experiments, $12$ of the lineups were used as a attention check which everyone will randomly get two of them and should get at least one correct, so that a measure of the quality of the subjects effort could be made.
+Note that no participant was shown the same lineup twice.
+
+## Non-linear covariates
+
+The experiment is designed to study the ability of human subjects to detect the effect of a higher order term $Z$ in a two variable statistical model formulated as:
+
+\begin{align} \label{eq:nonlinearity-model}
+\boldsymbol{Y} = 1 + \boldsymbol{X} + \boldsymbol{Z} + \boldsymbol{\varepsilon},\\
+\boldsymbol{X} = g(\boldsymbol{X_{raw}}, 1), \\
+\boldsymbol{Z} = g(\boldsymbol{Z_{raw}}, 1), \\
+\boldsymbol{Z_{raw}} = He_j(g(\boldsymbol{X}, 2)),
+\end{align}
+
+where $\boldsymbol{Y}$, $\boldsymbol{X}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{X_{raw}}$, $\boldsymbol{Z_{raw}}$ are $n\times1$ matrices,  $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $j \in \{2, 3, 6, 18\}$, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, $\sigma \in \{0.5, 1, 2, 4\}$, and $g(\boldsymbol{X}, k)$ is a scaling function to enforce the support of the variable to be $\{-k, k\}$ defined as
+
+\begin{equation} \label{eq:scaling-function}
+g(\boldsymbol{X}, k) = (\boldsymbol{X} - min(\boldsymbol{X}))/max(\boldsymbol{X} - min(\boldsymbol{X})) \times 2k - k, \quad \text{for} \quad k > 0. 
+\end{equation}
+
+The null regression model used to fit the data generated by the above model is formulated as:
+
+\begin{equation} \label{eq:scaling-function}
+\boldsymbol{Y} = \beta_0 + \beta_1 \boldsymbol{X} + \boldsymbol{u},
+\end{equation}
+
+where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.
+
+Clearly, omitted-variable bias will present since the null model leaves out the high order term.
+Data were simulated using a range of values of $j$ as shown in [table ref here], three different smaple sizes (n = 50, 100, 300) and four different standard deviations of the error ($\sigma$ = 0.5, 1, 2, 4).
+The range of $j$ was chosen so that different shapes of non-linearity are included in the residual plot.
+
+```{r}
+VI_MODEL$plot(poly_model(shape = 1, x = rand_uniform(), e = 0.5)$gen(300)) -> p1
+VI_MODEL$plot(poly_model(shape = 2, x = rand_uniform(), e = 0.5)$gen(300)) -> p2
+VI_MODEL$plot(poly_model(shape = 3, x = rand_uniform(), e = 0.5)$gen(300)) -> p3
+VI_MODEL$plot(poly_model(shape = 4, x = rand_uniform(), e = 0.5)$gen(300)) -> p4
+
+patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)
+```
+
+The range of $\sigma$ was chosen so that estimates of the power would produce reasonably continuous power curves, comparable to that calculated for the conventional test.
+The values for $\boldsymbol{X_{raw}}$ were generated randomly and independently from one of the following four distributions: (1) $U(a = -1, b = 1)$, (2) $N(\mu = 0, \sigma = 0.3)$, (3) $lognormal(\mu = 0, \sigma = 0.6)/3$ and (4) $u\{a = 1, b = 5\}$.
+The uniform and the normal distribution are symmetric and commonly assumed.
+The adjusted log-normal distribution provides skewered density.
+The discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.
+Three replications are made for each of the parameter values shown in [table ref here] and thus, 192 different lineups.
+For each lineup, the null model was fit to the actual dataset to obtain residuals and parameter estimates.
+The actual data plot was drawn as standard residual plot with raw residuals on the y-axis and fitted values on the x-axis.
+The 19 null datasets were generated by the residual rotation technique, and plotted in the same way.
+The actual data plot was randomly placed among these null dadta plots to produce the lineup.
+[figure ref here] is an example of one of these lineups.
+It was generated for $n = $, $\sigma = $, $j = $, $X_{raw} \sim$.
+For this lineup, x out of k observers picked the actual data plot.
+Each lineup is evaluated by 5 different subjects
 
@@
 if (!file.exists(here::here("data/third_study/conv_result.rds"))) {
-  conv_result <- map(1:100000,
+  conv_result <- list()
+  for (i in 1:100) {
+    conv_result[[i]] <- map(1:100000,
                             function(x) {
@@
                             effect_size = mod$effect_size(tmp_dat),
-                            reject = mod$test(tmp_dat)$p_value < 0.05)
+                            p_value = mod$test(tmp_dat)$p_value,
+                            reject = p_value < 0.05)
                    }) %>%
@@
   mutate(reject = as.numeric(reject))
+  }
 
@@
 ```
+
+```{r}
+polynomials %>%
+  group_by(lineup_id) %>%
+  summarise(across(c(type:p_value), first)) %>%
+  filter(lineup_id < 577) %>%
+  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
+  ggplot() +
+  geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.05)), aes(log(effect_size), reject, col = "Conventional test:0.05"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.01)), aes(log(effect_size), reject, col = "Conventional test:0.01"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.005)), aes(log(effect_size), reject, col = "Conventional test:0.005"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.0005)), aes(log(effect_size), reject, col = "Conventional test:0.0005"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.0000005)), aes(log(effect_size), reject, col = "Conventional test:0.0000005"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  xlab("Natural logarithm of effect size") +
+  ylab("Power")
+```

### 2022-07-27 paper_comparison/paper_comparison/paper_comparison.tex
udpate paper

@@
 \def\tightlist{}
+\newcommand*{\Perm}[2]{{}^{#1}\!P_{#2}}%
 
@@
 
-\title{Why shouldn't you use numerical tests to diagnose the linear
-regression models?}
+\title{Effectiveness of visual test in linear regression diagnositics}
 
@@
 
-Regression analysis is a field of study with at least a hundred years of
-history, and regression diagnostics is one of the essential steps in
-regression analysis. The diagnostic procedure conventionally involves
-evaluating the fitness of the proposed model, detecting the presence of
-influential observations and outliers, checking the validity of model
-assumptions and many more. In terms of diagnostic techniques, data
-plots, hypothesis testing, and summary statistics are vital tools for a
-systematic and detailed examination of the regression model
-\citep{mansfield1987diagnostic}.
+Regression diagnostics is an essential step in regression analysis which
+is a field of study with at least a hundred years of history. The
+diagnostic procedure conventionally involves evaluating the fitness of
+the proposed model, detecting the presence of influential observations
+and outliers, checking the validity of model assumptions and many more.
+In terms of diagnostic techniques, data plots, hypothesis testing, and
+summary statistics are vital tools for a systematic and detailed
+examination of the regression model \citep{mansfield1987diagnostic}.
 
@@
 and well-established in books first published in the twentieth century,
-such as \citet{draper_applied_2014},
-\citet{montgomery_introduction_2012}, \citet{belsley_regression_1980},
-\citet{cook_applied_1999} and \citet{cook1982residuals}. Regardless of
-the level of difficulty, one will find the importance and usefulness of
-diagnostic plots being emphasized in those books repeatedly. Checking
-diagnostic plots is also the recommended starting point for validating
-model assumptions such as normality, homoscedasticity and linearity
+such as \citet{draper1998applied}, \citet{montgomery1982introduction},
+\citet{belsley_regression_1980}, \citet{cook_applied_1999} and
+\citet{cook1982residuals}. Regardless of the level of difficulty, one
+will find the importance and usefulness of diagnostic plots being
+emphasized in those books repeatedly. Checking diagnostic plots is also
+the recommended starting point for validating model assumptions such as
+normality, homoscedasticity and linearity
 \citep{anscombe_examination_1963}.
@@
 standard residual plots will often be enhanced by studentized residuals
-in small sample size \citep{gunst2018regression}. As a two-dimensional
+in small sample size \citep{gunst1980regression}. As a two-dimensional
 representation of a model in a \(p\)-dimensional space, standard
@@
 the same projection will be treated as equivalent as they have the same
-abscissa. Therefore, standard residual plots are often useful in
+abscissa. Consequently, standard residual plots are often useful in
 revealing model inadequacies in the direction of the variable of the
@@
 with different horizontal axes will be examined
-\citep{cook1982residuals}. Overlapping data points is a general issue in
-scatter plots not limited to standard residual plots, which often makes
-plots difficult to interpret because visual patterns are concealed.
-Thus, for a relatively large sample size, \citet{cleveland1975graphical}
-suggests the use of robust moving statistics as reference lines to give
-aid to the eye in seeing patterns, which nowadays, are usually replaced
-with a spline or local polynomial regression line.
+\citep{cook1982residuals}. Overlapping data points is another general
+issue in scatter plots not limited to standard residual plots, which
+often makes plots difficult to interpret because visual patterns are
+concealed. Thus, for a relatively large sample size,
+\citet{cleveland1975graphical} suggests the use of robust moving
+statistics as reference lines to give aid to the eye in seeing patterns,
+which nowadays, are usually replaced with a spline or local polynomial
+regression line.
 
@@
 
-In addition to diagnostic plots, analysts may perform formal hypothesis
-testing for detecting model defects. Depends on the alternative
-hypothesis that being focused on, variety of tests can be applied. For
-example, the presence of heteroskedasticity can usually be tested by
-applying the White test \citep{white_heteroskedasticity-consistent_1980}
-or the Breusch-Pagan test \citep{breusch_simple_1979}, which both
-derived from the Lagrange multiplier test \citep{silvey1959lagrangian}
-principle and rely on the asymptotic properties of the null
-distribution. For testing non-linearity, one may apply the F-test to
-examine the significance of certain polynomial and non-linear forms of
-the regressors, or the significance of proxy variables as in the Ramsey
+Other than checking diagnostic plots, analysts may perform formal
+hypothesis testing for detecting model defects. Depending on the
+alternative hypothesis that is focused on, a variety of tests can be
+applied. For example, the presence of heteroskedasticity can usually be
+tested by applying the White test
+\citep{white_heteroskedasticity-consistent_1980} or the Breusch-Pagan
+test \citep{breusch_simple_1979}, which are both derived from the
+Lagrange multiplier test \citep{silvey1959lagrangian} principle that
+relies on the asymptotic properties of the null distribution. For
+testing non-linearity, one may apply the F-test to examine the
+significance of specific polynomial and non-linear forms of the
+regressors, or the significance of proxy variables as in the Ramsey
 Regression Equation Specification Error Test (RESET)
@@
 other types of departures. It is likely the null hypothesis is correctly
-rejected but for the wrong reason, which is also known as the ``Type III
+rejected but for the wrong reason, which is known as the ``Type III
 error''. Additionally, outliers will often incorrectly trigger the
-rejection of the null hypothesis despite the residuals being
-well-behaved \citep{cook_applied_1999}. This can be largely avoided in
-diagnostic plots as experienced analysts can evaluate the acceptability
-of assumptions flexibly, even in the presence of outliers.
-\citet{montgomery_introduction_2012} stated that based on their
+rejection of the null hypothesis despite the residuals are well-behaved
+\citep{cook_applied_1999}. This can be largely avoided in diagnostic
+plots as experienced analysts can evaluate the acceptability of
+assumptions flexibly, even in the presence of outliers.
+\citet{montgomery1982introduction} suggested that based on their
 experience, statistical tests are not widely used in regression
@@
 enough. The outcome ``Not reject'' can be interpreted as either ``effect
-size is small'' or ``sample size is small''. The outcome ``reject''
-still doesn't inform us whether and how much the model defects are of
-actual consequence to the inference and prediction. But still, the
+size is small'' or ``sample size is small''. and the outcome ``reject''
+doesn't inform us whether and how much the model defects are of actual
+consequence to the inference and prediction. But still, the
 effectiveness of statistical tests shall not be disrespected.
 Statistical tests have a chance to provide analysts with unique
-information. There are situations where no suitable diagnostic plots can
-be found for a particular violation of the assumptions, or excessive
-diagnostic plots need to be checked. One will have no choice but to rely
-on statistical tests if there are any. A good regression diagnostic
-practice should be a balanced combination of both methods.
+information. In situations where no suitable diagnostic plots can be
+found for a particular violation of the assumptions, or excessive
+diagnostic plots are needed to be checked, one will have no choice but
+to fall back on statistical tests if there exists any. A good regression
+diagnostic practice should be a balanced combination of both methods.
 
@@
 graph \citep{cleveland_graphical_1984}, which is to some extent
-subjective. Further, visual discovery suffers from its unsecured and
-unconfirmed nature where the degree of the presence of the visual
-features typically can not be measured quantitatively and objectively,
-which may lead to over or under-interpretations of the data. One such
-example is finding an over-interpretation of the separation between gene
-groups in a two-dimensional projection from a linear discriminant
-analysis when in fact there are no differences in the expression levels
-between the gene groups and separation is not an uncommon occurrence
-\citep{roy_chowdhury_using_2015}.
+subjective and indecisive. Further, visual discovery suffers from its
+unsecured and unconfirmed nature where the degree of the presence of the
+visual features typically can not be measured quantitatively and
+objectively, which may lead to over or under-interpretations of the
+data. One such example is finding an over-interpretation of the
+separation between gene groups in a two-dimensional projection from a
+linear discriminant analysis when in fact there are no differences in
+the expression levels between the gene groups and separation is not an
+uncommon occurrence \citep{roy_chowdhury_using_2015}.
 
@@
 dataset to a data plot. Both definitions of visual test statistics are
-valid, but in the rest of the report the first definition will be used
-as it covers some details needed by the following discussion. A visual
+valid, but in the rest of the paper the first definition will be used as
+it covers some details needed by the following discussion. A visual
 discovery is defined as a rejection of a null hypothesis, and the same
@@
 plot, including non-linearity and heteroskedasticity as shown in Figure
-@ref(fig:residual-plot-cubic-heter).
+\ref{fig:residual-plot-cubic-heter}.
 
@@
 heteroskedasticity actually does not exist in the data generating
-process.}
+process. \label{fig:residual-plot-cubic-heter}}
 \end{figure}
@@
 
-With the simulation of null plots being provided, another aspect of
-hypothesis testing that needs to be addressed is the control of false
-positive rate or Type I error. Any visual statistic
+With the simulation mechanism of null plots being provided, another
+aspect of hypothesis testing that needs to be addressed is the control
+of false positive rate or Type I error. Any visual statistic
 \(T^{(i)}(\boldsymbol{\mathrm{y}})\) needs to pair with a critical value
@@
 
-Further, if we involve \(K\) independent observers in a visual test, and
-let \(X\) be a random variable denoting the number of observers
-correctly picking the actual data plot. Then, under the null hypothesis
-\(X \sim \mathrm{Binom}_{K,1/m}\), and therefore, the \(p\)-value of a
-lineup of size \(m\) evaluated by \(K\) observer is given as
-\begin{equation} \label{eq:pvaluesingle}
+Further, a visual test can involve \(K\) independent observers. Let
+\(D_i = \{0,1\}\) be a binomial random variable denoting whether subject
+\(i\) correctly detecting the actual data plot, and
+\(X = \sum_{i=1}^{K}X_i\) be the number of observers correctly picking
+the actual data plot. Then, by imposing a relatively strong assumption
+on the visual test that all \(K\) evaluations are fully independent,
+under the null hypothesis, \(X \sim \mathrm{Binom}_{K,1/m}\). Therefore,
+the \(p\)-value of a lineup of size \(m\) evaluated by \(K\) observer is
+given as \begin{equation} \label{eq:pvaluesingle}
 P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},
@@
 
-Compared to the conventional test, whose power only depends on the
-parameter of interest \(\theta\), several studies
-\citep[see][]{hofmann_graphical_2012, majumder_validation_2013, majumder_human_2014, roy_chowdhury_using_2015, loy_variations_2016}
-have shown the power of the visual test is subject-specific. Thus, to be
-able to account for individual's ability, an individual is required to
-evaluate multiple lineups \citep{majumder_validation_2013}.
+As pointed out by \citet{vanderplas2021statistical}, though equation
+(\ref{eq:pvaluesingle}) is trivial, but it doesn't take into account the
+possible dependencies in the visual test due to repeated evaluations of
+the same lineup. And it is inapplicable to visual test where subjects
+are asked to select one or more ``most different'' plots from the
+lineup. They summarized three common different scenarios in visual
+inference: (1) \(K\) different lineups are shown to \(K\) subjects, (2)
+\(K\) lineups with different null plots but the same actual data plot
+are shown to \(K\) subjects, and (3) the same lineup is shown to \(K\)
+subjects. Out of these three scenarios, Scenario 3 is the most common in
+previous studies as it puts the least constraints on the experimental
+design. For Scenario 3, \citet{vanderplas2021statistical} modelled the
+probability of a plot \(i\) being selected from a lineup as
+\(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
+\(i=1,...,m\) and \(\alpha > 0\). And defined \(c_i\) to be the number
+of times plot \(i\) being selected in \(K\) evaluations. In case subject
+\(j\) makes multiple selections, they decided to add \(1/s_j\) to
+\(c_i\) instead of one, where \(s_j\) is the number of plots subject
+\(j\) selected for \(j=1,...K\). This ensured \(\sum_{i}c_i=K\).
 
-Suppose individuals have the same ability and a lineup has been
-evaluated by multiple individuals, under the alternative hypothesis, the
-estimated power for a lineup can be expressed as \(\hat{p} = x/K\), the
-estimated probability of identifying the actual data plot from the
-lineup. If the individual skill needs to be taken into account, and
-\(L\) lineups have been evaluated by \(K\) individuals,
-\citet{majumder_validation_2013} suggests that mixed effects logistic
-regression model can be fit as:
+The full model was a Dirichlet-multinomial mixture distribution
 
-\[g(p_{li}) = W_{li}\delta + Z_{li}\tau_{li},\] where \(g(.)\) is the
-logit link function \(g(p) = log(p) - log(1-p)\); \(0 \leq p \leq 1\).
-\(W_{li}\), \(1 \leq i \leq K\), \(1 \leq l \leq L\), is the covariate
-matrix including lineup-specific elements and demographic information of
-individuals, and \(\delta\) is a vector of parameters. \(Z\) is the
-random effects matrix, and \(\tau\) is a vector of variables follow
-\(N(\boldsymbol{0},\sigma_{\tau}\boldsymbol{I}_{KL\times KL})\).
+\begin{align} \label{eq:dirichlet-multinomial}\begin{split}
+\boldsymbol{\theta}&|\alpha \sim Dirichlet(\alpha)\\
+(c_i,...,c_m)&|\boldsymbol{\theta} \sim Multinomial(K, \boldsymbol{\theta}).
+\end{split}\end{align}
 
-Then, the estimated power for lineup \(l\) and individual \(i\) can be
-calculated as
-\(\hat{p}_{li} = g^{-1}(W_{li}\hat{\delta} + Z_{li}\hat{\tau}_{li})\)
-\citep{majumder_validation_2013}.
+Since the p-value calculation only needs to concern the number of times
+the actual data plot being selected denoted by \(C_i\), they showed the
+model can be simplified to a beta-binomial mixture distribution
+
+\begin{align} \label{eq:beta-binomial}\begin{split}
+\theta_i&|\alpha \sim Beta(\alpha, (m-1)\alpha)\\
+C_i&|\theta_i \sim Binomial(K, \theta_i).
+\end{split}\end{align}
+
+Thus, the visual p-value followed by the beta-binomial model is given as
+
+\begin{equation} \label{eq:pvalue-beta-binomial}
+P(C \geq c_i) = \sum_{x=c_i}^{K}{{K}\choose{x}}\frac{1}{B(\alpha, (m-1)\alpha)}B(x + \alpha, K - x + (m - 1)\alpha),
+\end{equation}
+
+where \(B(.)\) is the beta function defined as
+
+\begin{equation} \label{eq:betafunction}
+B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0. 
+\end{equation}
+
+Note that the use of equation (\ref{eq:pvalue-beta-binomial}) requires
+the estimation of \(\hat{\alpha}\), which largely depends on the null
+model, the type of the plot and other aesthetic features. They suggested
+to estimate \(\hat{\alpha}\) visually based on the selections of null
+plots of the experimental data, or to estimate \(\hat{\alpha}\)
+numerically based on several additional Rorschach lineups, which is a
+type of lineup containing only null plots. However, when the number of
+null models are large, it could be expensive to manually estimate each
+\(\alpha\) or include additional Rorschach lineups in the experiment.
+
+Instead, in the experiments that will be described in section
+\ref{experimental-design}, we adopt a simpler model implicitly used by
+the \texttt{pmulti()} function of the \texttt{vinference} \texttt{R}
+package. We assume the attractiveness of the plot \(i\) modelled as
+\(w_i \sim Uniform(0,1)\) for \(i=1,..,m\). Let
+\(\theta_i = w_i/\sum_{i=1}^{m}w_i\) be the probability of plot \(i\)
+being selected by a subject. Then, given the number of selections
+\(s_j\), for \(j=1,...,K\), the distribution of \(C_i\) can be
+approximated by simulating the random selection process with computer.
+The simulated visual test p-value is formulated as
+
+\begin{equation} \label{eq:p-value-multi}
+\text{p-value} = \frac{\#draws~that~the~actual~data~plota~being~selected~more~than~c_i~times}{\#simulation}.
+\end{equation}
 
@@
 
+The effectiveness of visual inference has already been validated by
+\citet{majumder_validation_2013} under relatively simple classical
+normal linear regression model settings with only one or two regressors.
+Their results suggests visual test is capable of testing the
+significance of a single regressor with a similar power as a t-test,
+though they expressed that in general it is unnecessary to use visual
+inference if there exists a conventional test and they didn't expect the
+visual test to perform equally well as the conventional test. In their
+third experiment, where there does not exist a proper conventional test,
+visual test outperforms the conventional test for a large margin. This
+is encouraging as it promotes the use of visual inference in border
+field of data science where there are no existing statistical testing
+procedures. In fact, lineup protocol has been integrated into some model
+diagnostic tools such as \citet{loy2013diagnostic}.
+
+With our knowledge, what haven't been examined so far is the
+effectiveness of visual test relative to the equivalent conventional
+test in regression diagnostics. Particularly, its ability to detect
+non-linearity and heteroskedasticity compared to F-test and BP-test.
+
+Two experiments were conducted to evaluate the effectiveness of visual
+test relative to the equivalent conventional test in regression
+diagnostics. The first experiment has ideal scenario for conventional
+testing, where we would not expect the visual test to do better than the
+conventional test. The second experiment is a scenario where the
+conventional test is an approximate test, which visual test may have a
+chance to match the performance of the conventional test.
+
+As suggested by many previous studies {[}many studies refs here{]}, it
+is clear some care is needed to set up the human subjects experiments.
+It is best for an subject to evaluate a block of 20 lineups with varying
+difficulty, with a reasonable number of ``easy'' lineups. The
+explanations about each experiment includes an explantion of how the
+lineups were sampled and provided to the subjects.
+
+Participants for all the experiments were recruited through Prolific's
+online web service {[}ref here{]}. A summary of the data obtained for
+all three eperiments are shown in {[}table ref here{]}. Participants
+were asked to select the plots which are mostly different from others,
+provide a reason for their selections, and express how different they
+think the selected plots are from others. Gender, age, education, and
+previous experience in visual experiment are also collected. For each of
+the experiments, \(12\) of the lineups were used as a attention check
+which everyone will randomly get two of them and should get at least one
+correct, so that a measure of the quality of the subjects effort could
+be made. Note that no participant was shown the same lineup twice.
+
+\hypertarget{non-linear-covariates}{%
+\subsection{Non-linear covariates}\label{non-linear-covariates}}
+
+The experiment is designed to study the ability of human subjects to
+detect the effect of a higher order term \(Z\) in a two variable
+statistical model formulated as:
+
+\begin{align} \label{eq:nonlinearity-model}
+\boldsymbol{Y} = 1 + \boldsymbol{X} + \boldsymbol{Z} + \boldsymbol{\varepsilon},\\
+\boldsymbol{X} = g(\boldsymbol{X_{raw}}, 1), \\
+\boldsymbol{Z} = g(\boldsymbol{Z_{raw}}, 1), \\
+\boldsymbol{Z_{raw}} = He_j(g(\boldsymbol{X}, 2)),
+\end{align}
+
+where \(\boldsymbol{Y}\), \(\boldsymbol{X}\),
+\(\boldsymbol{\varepsilon}\), \(\boldsymbol{X_{raw}}\),
+\(\boldsymbol{Z_{raw}}\) are \(n\times1\) matrices, \(He_{j}(.)\) is the
+\(j\)th-order probabilist's Hermite polynomials,
+\(j \in \{2, 3, 6, 18\}\),
+\(\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\),
+\(\sigma \in \{0.5, 1, 2, 4\}\), and \(g(\boldsymbol{X}, k)\) is a
+scaling function to enforce the support of the variable to be
+\(\{-k, k\}\) defined as
+
+\begin{equation} \label{eq:scaling-function}
+g(\boldsymbol{X}, k) = (\boldsymbol{X} - min(\boldsymbol{X}))/max(\boldsymbol{X} - min(\boldsymbol{X})) \times 2k - k, \quad \text{for} \quad k > 0. 
+\end{equation}
+
+The null regression model used to fit the data genearted by the above
+model is formulated as:
+
+\begin{equation} \label{eq:scaling-function}
+\boldsymbol{Y} = \beta_0 + \beta_1 \boldsymbol{X} + \boldsymbol{u},
+\end{equation}
+
+where
+\(\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\).
+
+Clearly, omitted-variable bias will present since the null model leaves
+out the high order term. Data were simulated using a range of values of
+\(j\) as shown in {[}table ref here{]}, three different smaple sizes (n
+= 50, 100, 300) and four different standard deviations of the error
+(\(\sigma\) = 0.5, 1, 2, 4). The range of \(j\) was chosen so that
+different shapes of non-linearity are included in the residual plot.
+
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-2-1.pdf}
+
+The range of \(sigma\) was chosen so that estimates of the power would
+produce reasonably continuous power curves, comparable to that
+calculated for the conventional test. The values for \(X_{raw}\) were
+randomly generated independently from one of the following four
+distributions: (1) \(U(a = -1, b = 1)\), (2)
+\(N(\mu = 0, \sigma = 0.3)\), (3)
+\(lognormal(\mu = 0, \sigma = 0.6)/3 - 1\) and (4)
+\(u\{a = 1, b = 5\}\).
+
 \hypertarget{demographic-summary}{%
@@
 
-\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-2-1.pdf}
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-3-1.pdf}
 
@@
 
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-4-1.pdf}
+
 \bibliographystyle{tfcad}

### 2022-07-27 paper_comparison/paper_comparison/paper_comparison.Rmd
provide better plots for expermential design

@@
 
+<!-- refactoring from here -->
+
 With our knowledge, what haven't been examined so far is the effectiveness of visual test relative to the equivalent conventional test in regression diagnostics.
 Particularly, its ability to detect non-linearity and heteroskedasticity compared to F-test and BP-test.
-
-Two experiments were conducted to evaluate the effectiveness of visual test relative to the equivalent conventional test in regression diagnostics.
+For this purpose, two experiments were conducted.
 The first experiment has ideal scenario for conventional testing, where we would not expect the visual test to do better than the conventional test.
@@
 
-As suggested by many previous studies [many studies refs here], it is clear some care is needed to set up the human subjects experiments.
-It is best for an subject to evaluate a block of 20 lineups with varying difficulty, with a reasonable number of "easy" lineups.
-The explanations about each experiment includes an explantion of how the lineups were sampled and provided to the subjects.
+As suggested by many previous studies [many studies refs here], it is reasonable for an subject to evaluate a block of 20 lineups.
+Subjects for both two experiments were recruited from Prolific web service [ref here].
+They were asked to select one or more plots that are mostly different from others, provide a reason for their selections, and express how different they think the selected plots are from others.
+Information about gender, age, education, and previous experience in visual experiment were also collected.
+No subject was shown the same lineup twice.
 
-Participants for all the experiments were recruited through Prolific's online web service [ref here].
-A summary of the data obtained for all three eperiments are shown in [table ref here].
-Participants were asked to select the plots which are mostly different from others, provide a reason for their selections, and express how different they think the selected plots are from others.
-Gender, age, education, and previous experience in visual experiment are also collected.
-For each of the experiments, $12$ of the lineups were used as a attention check which everyone will randomly get two of them and should get at least one correct, so that a measure of the quality of the subjects effort could be made.
-Note that no participant was shown the same lineup twice.
+For each of the experiments, a pool of 12 lineups with obvious visual patterns were generated.
+In every block of 20 lineups that presented to subjects, two different lineups sampled from the corresponding pool of lineups were included as attention checks.
+Every subject should get at least one attention check correct such that the quality of the subject effort could be ensured.
 
@@
 
-The experiment is designed to study the ability of human subjects to detect the effect of a higher order term $Z$ in a two variable statistical model formulated as:
+The first experiment is designed to study the ability of human subjects to detect the effect of a continuous variable $\boldsymbol{Z}$ which is a probabilist's Hermite polynomial [Herimite ref here] of another variable $\boldsymbol{X}$ in a two variable statistical model formulated as:
 
@@
 \boldsymbol{Y} = 1 + \boldsymbol{X} + \boldsymbol{Z} + \boldsymbol{\varepsilon},\\
-\boldsymbol{X} = g(\boldsymbol{X_{raw}}, 1), \\
-\boldsymbol{Z} = g(\boldsymbol{Z_{raw}}, 1), \\
-\boldsymbol{Z_{raw}} = He_j(g(\boldsymbol{X}, 2)),
+\boldsymbol{X} = g(\boldsymbol{X}_{raw}, 1), \\
+\boldsymbol{Z} = g(\boldsymbol{Z}_{raw}, 1), \\
+\boldsymbol{Z}_{raw} = He_j(g(\boldsymbol{X}, 2)),
 \end{align}
 
-where $\boldsymbol{Y}$, $\boldsymbol{X}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{X_{raw}}$, $\boldsymbol{Z_{raw}}$ are $n\times1$ matrices,  $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $j \in \{2, 3, 6, 18\}$, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, $\sigma \in \{0.5, 1, 2, 4\}$, and $g(\boldsymbol{X}, k)$ is a scaling function to enforce the support of the variable to be $\{-k, k\}$ defined as
+where $\boldsymbol{Y}$, $\boldsymbol{X}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{X}_{raw}$, $\boldsymbol{Z}_{raw}$ are $n\times1$ matrices,  $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{X}, k)$ is a scaling function to enforce the support of the variable to be $\{-k, k\}$ defined as
 
@@
 
-Clearly, omitted-variable bias will present since the null model leaves out the high order term.
-Data were simulated using a range of values of $j$ as shown in [table ref here], three different smaple sizes (n = 50, 100, 300) and four different standard deviations of the error ($\sigma$ = 0.5, 1, 2, 4).
-The range of $j$ was chosen so that different shapes of non-linearity are included in the residual plot.
+Clearly, omitted-variable bias will present since the null model leaves out the higher order term.
+Data were simulated using four different order of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$), three different sample sizes (n = 50, 100, 300) , four different standard deviations of the error ($\sigma$ = 0.5, 1, 2, 4) and four different distribution of $\boldsymbol{X}_{raw}$: (1) $U(a = -1, b = 1)$, (2) $N(\mu = 0, \sigma = 0.3)$, (3) $lognormal(\mu = 0, \sigma = 0.6)/3$ and (4) $u\{a = 1, b = 5\}$.
+A summary of the parameters used in this experiment is given in [table ref here]
 
-```{r}
-VI_MODEL$plot(poly_model(shape = 1, x = rand_uniform(), e = 0.5)$gen(300)) -> p1
-VI_MODEL$plot(poly_model(shape = 2, x = rand_uniform(), e = 0.5)$gen(300)) -> p2
-VI_MODEL$plot(poly_model(shape = 3, x = rand_uniform(), e = 0.5)$gen(300)) -> p3
-VI_MODEL$plot(poly_model(shape = 4, x = rand_uniform(), e = 0.5)$gen(300)) -> p4
+The values of $j$ was chosen so that visually different shapes of non-linearity are included in the residual plot. These include "U" shape, "S" shape, "M" shape and "Triple-U" shape.
+
+```{r different-shape-of-herimite}
+VI_MODEL$plot(poly_model(shape = 1, x = rand_uniform(), e = 0.5)$gen(300)) + ggtitle(quote(He[2]:"U shape")) -> p1
+VI_MODEL$plot(poly_model(shape = 2, x = rand_uniform(), e = 0.5)$gen(300)) + ggtitle(quote(He[3]:"S shape")) -> p2
+VI_MODEL$plot(poly_model(shape = 3, x = rand_uniform(), e = 0.5)$gen(300)) + ggtitle(quote(He[6]:"M shape")) -> p3
+VI_MODEL$plot(poly_model(shape = 4, x = rand_uniform(), e = 0.5)$gen(300)) + ggtitle(quote(He[18]:"Triple-U shape")) -> p4
 
@@
 The range of $\sigma$ was chosen so that estimates of the power would produce reasonably continuous power curves, comparable to that calculated for the conventional test.
-The values for $\boldsymbol{X_{raw}}$ were generated randomly and independently from one of the following four distributions: (1) $U(a = -1, b = 1)$, (2) $N(\mu = 0, \sigma = 0.3)$, (3) $lognormal(\mu = 0, \sigma = 0.6)/3$ and (4) $u\{a = 1, b = 5\}$.
+The values for $\boldsymbol{X}_{raw}$ were generated randomly and independently from one of the following four distributions: (1) $U(a = -1, b = 1)$, (2) $N(\mu = 0, \sigma = 0.3)$, (3) $lognormal(\mu = 0, \sigma = 0.6)/3$ and (4) $u\{a = 1, b = 5\}$.
 The uniform and the normal distribution are symmetric and commonly assumed.
@@
 [figure ref here] is an example of one of these lineups.
-It was generated for $n = $, $\sigma = $, $j = $, $X_{raw} \sim$.
+It was generated for $n = 1$, $\sigma = 1$, $j = 1$, $X_{raw} \sim N$.
 For this lineup, x out of k observers picked the actual data plot.
-Each lineup is evaluated by 5 different subjects
+Each lineup is evaluated by 5 different subjects to provide reasonable estimates of the p-value.
+
+To satisfy the plan described above, we needed to recruit at least 160 subjects.
+
+# Results
+
+## Data cleaning
+
+Participants recruited from Prolific are paid for their efforts.
+Some participants will try to maximize their earnings for minimum effort, which can affect the result from the data.
+For the purpose of identifying these participants and cleaning the data, two attention checks are included in the block of 20 lineups.
+If the subject failed to identify at least one of the actual data plot in those attention checks, we reject the submission and the subject will not be paid.
+All attention checks are removed from the final data.
+[table ref here] tabulates the number of subjects, genders, and lineups evaluated after applying the data screening procedure.
+
+## Model fitting
+
+For each parameter combination, effect $E$ is derived as $E = \frac{1}{2}(\boldsymbol{R}_a\boldsymbol{X}_b)'diag(\sigma^2\boldsymbol{R}_a)\boldsymbol{R}_a\boldsymbol{X}_b$ from the Kullback-Leibler divergence.
+[KL divergence ref here] (The derivation is provided in [appendix ref here].).
+The logistic regression is fit using $E$ as the only fixed effect covariate
 

### 2022-07-27 paper_comparison/paper_comparison/paper_comparison.tex
provide better plots for expermential design

@@
 test in regression diagnostics. Particularly, its ability to detect
-non-linearity and heteroskedasticity compared to F-test and BP-test.
-
-Two experiments were conducted to evaluate the effectiveness of visual
-test relative to the equivalent conventional test in regression
-diagnostics. The first experiment has ideal scenario for conventional
-testing, where we would not expect the visual test to do better than the
-conventional test. The second experiment is a scenario where the
-conventional test is an approximate test, which visual test may have a
-chance to match the performance of the conventional test.
+non-linearity and heteroskedasticity compared to F-test and BP-test. For
+this purpose, two experiments were conducted. The first experiment has
+ideal scenario for conventional testing, where we would not expect the
+visual test to do better than the conventional test. The second
+experiment is a scenario where the conventional test is an approximate
+test, which visual test may have a chance to match the performance of
+the conventional test.
 
 As suggested by many previous studies {[}many studies refs here{]}, it
-is clear some care is needed to set up the human subjects experiments.
-It is best for an subject to evaluate a block of 20 lineups with varying
-difficulty, with a reasonable number of ``easy'' lineups. The
-explanations about each experiment includes an explantion of how the
-lineups were sampled and provided to the subjects.
+is reasonable for an subject to evaluate a block of 20 lineups. Subjects
+for both two experiments were recruited from Prolific web service {[}ref
+here{]}. They were asked to select one or more plots that are mostly
+different from others, provide a reason for their selections, and
+express how different they think the selected plots are from others.
+Information about gender, age, education, and previous experience in
+visual experiment were also collected. No subject was shown the same
+lineup twice.
 
-Participants for all the experiments were recruited through Prolific's
-online web service {[}ref here{]}. A summary of the data obtained for
-all three eperiments are shown in {[}table ref here{]}. Participants
-were asked to select the plots which are mostly different from others,
-provide a reason for their selections, and express how different they
-think the selected plots are from others. Gender, age, education, and
-previous experience in visual experiment are also collected. For each of
-the experiments, \(12\) of the lineups were used as a attention check
-which everyone will randomly get two of them and should get at least one
-correct, so that a measure of the quality of the subjects effort could
-be made. Note that no participant was shown the same lineup twice.
+For each of the experiments, a pool of 12 lineups with obvious visual
+patterns were generated. In every block of 20 lineups that presented to
+subjects, two different lineups sampled from the corresponding pool of
+lineups were included as attention checks. Every subject should get at
+least one attention check correct such that the quality of the subject
+effort could be ensured.
 
@@
 
-The experiment is designed to study the ability of human subjects to
-detect the effect of a higher order term \(Z\) in a two variable
-statistical model formulated as:
+The first experiment is designed to study the ability of human subjects
+to detect the effect of a continuous variable \(\boldsymbol{Z}\) which
+is a probabilist's Hermite polynomial {[}Herimite ref here{]} of another
+variable \(\boldsymbol{X}\) in a two variable statistical model
+formulated as:
 
@@
 \boldsymbol{Y} = 1 + \boldsymbol{X} + \boldsymbol{Z} + \boldsymbol{\varepsilon},\\
-\boldsymbol{X} = g(\boldsymbol{X_{raw}}, 1), \\
-\boldsymbol{Z} = g(\boldsymbol{Z_{raw}}, 1), \\
-\boldsymbol{Z_{raw}} = He_j(g(\boldsymbol{X}, 2)),
+\boldsymbol{X} = g(\boldsymbol{X}_{raw}, 1), \\
+\boldsymbol{Z} = g(\boldsymbol{Z}_{raw}, 1), \\
+\boldsymbol{Z}_{raw} = He_j(g(\boldsymbol{X}, 2)),
 \end{align}
@@
 where \(\boldsymbol{Y}\), \(\boldsymbol{X}\),
-\(\boldsymbol{\varepsilon}\), \(\boldsymbol{X_{raw}}\),
-\(\boldsymbol{Z_{raw}}\) are \(n\times1\) matrices, \(He_{j}(.)\) is the
+\(\boldsymbol{\varepsilon}\), \(\boldsymbol{X}_{raw}\),
+\(\boldsymbol{Z}_{raw}\) are \(n\times1\) matrices, \(He_{j}(.)\) is the
 \(j\)th-order probabilist's Hermite polynomials,
-\(j \in \{2, 3, 6, 18\}\),
-\(\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\),
-\(\sigma \in \{0.5, 1, 2, 4\}\), and \(g(\boldsymbol{X}, k)\) is a
-scaling function to enforce the support of the variable to be
-\(\{-k, k\}\) defined as
+\(\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\), and
+\(g(\boldsymbol{X}, k)\) is a scaling function to enforce the support of
+the variable to be \(\{-k, k\}\) defined as
 
@@
 
-The null regression model used to fit the data genearted by the above
+The null regression model used to fit the data generated by the above
 model is formulated as:
@@
 Clearly, omitted-variable bias will present since the null model leaves
-out the high order term. Data were simulated using a range of values of
-\(j\) as shown in {[}table ref here{]}, three different smaple sizes (n
-= 50, 100, 300) and four different standard deviations of the error
-(\(\sigma\) = 0.5, 1, 2, 4). The range of \(j\) was chosen so that
-different shapes of non-linearity are included in the residual plot.
+out the higher order term. Data were simulated using four different
+order of probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)), three
+different sample sizes (n = 50, 100, 300) , four different standard
+deviations of the error (\(\sigma\) = 0.5, 1, 2, 4) and four different
+distribution of \(\boldsymbol{X}_{raw}\): (1) \(U(a = -1, b = 1)\), (2)
+\(N(\mu = 0, \sigma = 0.3)\), (3) \(lognormal(\mu = 0, \sigma = 0.6)/3\)
+and (4) \(u\{a = 1, b = 5\}\). A summary of the parameters used in this
+experiment is given in {[}table ref here{]}
 
-\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-2-1.pdf}
+The values of \(j\) was chosen so that visually different shapes of
+non-linearity are included in the residual plot. These include ``U''
+shape, ``S'' shape, ``M'' shape and ``Triple-U'' shape.
 
-The range of \(sigma\) was chosen so that estimates of the power would
+\includegraphics{paper_comparison_files/figure-latex/different-shape-of-herimite-1.pdf}
+
+The range of \(\sigma\) was chosen so that estimates of the power would
 produce reasonably continuous power curves, comparable to that
-calculated for the conventional test. The values for \(X_{raw}\) were
-randomly generated independently from one of the following four
-distributions: (1) \(U(a = -1, b = 1)\), (2)
-\(N(\mu = 0, \sigma = 0.3)\), (3)
-\(lognormal(\mu = 0, \sigma = 0.6)/3 - 1\) and (4)
-\(u\{a = 1, b = 5\}\).
+calculated for the conventional test. The values for
+\(\boldsymbol{X}_{raw}\) were generated randomly and independently from
+one of the following four distributions: (1) \(U(a = -1, b = 1)\), (2)
+\(N(\mu = 0, \sigma = 0.3)\), (3) \(lognormal(\mu = 0, \sigma = 0.6)/3\)
+and (4) \(u\{a = 1, b = 5\}\). The uniform and the normal distribution
+are symmetric and commonly assumed. The adjusted log-normal distribution
+provides skewered density. The discrete uniform distribution provides
+discreteness in residual plot, which could enrich the pool of visual
+patterns. Three replications are made for each of the parameter values
+shown in {[}table ref here{]} and thus, 192 different lineups. For each
+lineup, the null model was fit to the actual dataset to obtain residuals
+and parameter estimates. The actual data plot was drawn as standard
+residual plot with raw residuals on the y-axis and fitted values on the
+x-axis. The 19 null datasets were generated by the residual rotation
+technique, and plotted in the same way. The actual data plot was
+randomly placed among these null dadta plots to produce the lineup.
+{[}figure ref here{]} is an example of one of these lineups. It was
+generated for \(n = 1\), \(\sigma = 1\), \(j = 1\), \(X_{raw} \sim N\).
+For this lineup, x out of k observers picked the actual data plot. Each
+lineup is evaluated by 5 different subjects to provide reasonable
+estimates of the p-value.
+
+To satisfy the plan described above, we needed to recruit at least 160
+subjects.
+
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+\hypertarget{data-cleaning}{%
+\subsection{Data cleaning}\label{data-cleaning}}
+
+Participants recruited from Prolific are paid for their efforts. Some
+participants will try to maximize their earnings for minimum effort,
+which can affect the result from the data. For the purpose of
+identifying these participants and cleaning the data, two attention
+checks are included in the block of 20 lineups. If the subject failed to
+identify at least one of the actual data plot in those attention checks,
+we reject the submission and the subject will not be paid. All attention
+checks are removed from the final data. {[}table ref here{]} tabulates
+the number of subjects, genders, and lineups evaluated after applying
+the data screening procedure.
+
+\hypertarget{model-fitting}{%
+\subsection{Model fitting}\label{model-fitting}}
+
+For each parameter combination, effect \(E\) is derived as
+\(E = \frac{1}{2}(\boldsymbol{R}_a\boldsymbol{X}_b)'diag(\sigma^2\boldsymbol{R}_a)\boldsymbol{R}_a\boldsymbol{X}_b\)
+from the Kullback-Leibler divergence. {[}KL divergence ref here{]} (The
+derivation is provided in {[}appendix ref here{]}.). The logistic
+regression is fit using \(E\) as the only fixed effect covariate
 
@@
 
-\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-3-1.pdf}
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-2-1.pdf}
 
@@
 
-\hypertarget{results}{%
-\section{Results}\label{results}}
+\hypertarget{results-1}{%
+\section{Results}\label{results-1}}
 
@@
 
-\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-4-1.pdf}
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-3-1.pdf}
 

### 2022-07-27 paper_comparison/paper_comparison/paper_comparison.Rmd
include more plots in experimental design

@@
 ```{r different-shape-of-herimite}
-VI_MODEL$plot(poly_model(shape = 1, x = rand_uniform(), e = 0.5)$gen(300)) + ggtitle(quote(He[2]:"U shape")) -> p1
-VI_MODEL$plot(poly_model(shape = 2, x = rand_uniform(), e = 0.5)$gen(300)) + ggtitle(quote(He[3]:"S shape")) -> p2
-VI_MODEL$plot(poly_model(shape = 3, x = rand_uniform(), e = 0.5)$gen(300)) + ggtitle(quote(He[6]:"M shape")) -> p3
-VI_MODEL$plot(poly_model(shape = 4, x = rand_uniform(), e = 0.5)$gen(300)) + ggtitle(quote(He[18]:"Triple-U shape")) -> p4
+stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
+
+ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300)
+
+VI_MODEL$plot(ori_dat) + ggtitle(quote(He[2]:"U shape")) -> p1
+VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote(He[3]:"S shape")) -> p2
+VI_MODEL$plot(poly_model(shape = 3, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote(He[6]:"M shape")) -> p3
+VI_MODEL$plot(poly_model(shape = 4, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote(He[18]:"Triple-U shape")) -> p4
 
@@
 
-The range of $\sigma$ was chosen so that estimates of the power would produce reasonably continuous power curves, comparable to that calculated for the conventional test.
-The values for $\boldsymbol{X}_{raw}$ were generated randomly and independently from one of the following four distributions: (1) $U(a = -1, b = 1)$, (2) $N(\mu = 0, \sigma = 0.3)$, (3) $lognormal(\mu = 0, \sigma = 0.6)/3$ and (4) $u\{a = 1, b = 5\}$.
-The uniform and the normal distribution are symmetric and commonly assumed.
-The adjusted log-normal distribution provides skewered density.
-The discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.
+The range of $\sigma$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated and the estimated power curve would be smooth and continuous.
+Four different distribution were used to genearte $\boldsymbol{X}_raw$.
+
+```{r different-sigma}
+ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)
+
+VI_MODEL$plot(ori_dat) + ggtitle(quote(sigma~"="~0.5)) -> p1
+VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1)$gen(300, computed = select(ori_dat, x))) +
+  ggtitle(quote(sigma~"="~1)) -> p2
+VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 2)$gen(300, computed = select(ori_dat, x))) +
+  ggtitle(quote(sigma~"="~2)) -> p3
+VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 4)$gen(300, computed = select(ori_dat, x))) +
+  ggtitle(quote(sigma~"="~4)) -> p4
+
+patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)
+```
+
+The uniform and the normal distribution are symmetric and commonly assumed in statistical model.
+The adjusted log-normal distribution provides skewed density.
+And the discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.
+
+```{r different-dist}
+ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)
+
+VI_MODEL$plot(ori_dat) + ggtitle(quote(U(-1,1))) -> p1
+VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
+  ggtitle(quote(N(0,0.3))) -> p2
+VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
+  ggtitle(quote(lognormal(0,0.6)/3)) -> p3
+VI_MODEL$plot(poly_model(shape = 1, x = rand_uniform_d(k = 5, even = TRUE), sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
+  ggtitle(quote(U~"{1,5}")) -> p4
+
+patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)
+```
+
 Three replications are made for each of the parameter values shown in [table ref here] and thus, 192 different lineups.

### 2022-07-27 paper_comparison/paper_comparison/paper_comparison.tex
include more plots in experimental design

@@
 
-The range of \(\sigma\) was chosen so that estimates of the power would
-produce reasonably continuous power curves, comparable to that
-calculated for the conventional test. The values for
-\(\boldsymbol{X}_{raw}\) were generated randomly and independently from
-one of the following four distributions: (1) \(U(a = -1, b = 1)\), (2)
-\(N(\mu = 0, \sigma = 0.3)\), (3) \(lognormal(\mu = 0, \sigma = 0.6)/3\)
-and (4) \(u\{a = 1, b = 5\}\). The uniform and the normal distribution
-are symmetric and commonly assumed. The adjusted log-normal distribution
-provides skewered density. The discrete uniform distribution provides
+The range of \(\sigma\), which is a factor controlling the strength of
+the signal, was chosen so that different difficulty levels of lineups
+were generated and the estimated power curve would be smooth and
+continuous. Four different distribution were used to genearte
+\(\boldsymbol{X}_raw\).
+
+\includegraphics{paper_comparison_files/figure-latex/different-sigma-1.pdf}
+
+The uniform and the normal distribution are symmetric and commonly
+assumed in statistical model. The adjusted log-normal distribution
+provides skewed density. And the discrete uniform distribution provides
 discreteness in residual plot, which could enrich the pool of visual
-patterns. Three replications are made for each of the parameter values
-shown in {[}table ref here{]} and thus, 192 different lineups. For each
-lineup, the null model was fit to the actual dataset to obtain residuals
-and parameter estimates. The actual data plot was drawn as standard
-residual plot with raw residuals on the y-axis and fitted values on the
-x-axis. The 19 null datasets were generated by the residual rotation
-technique, and plotted in the same way. The actual data plot was
-randomly placed among these null dadta plots to produce the lineup.
-{[}figure ref here{]} is an example of one of these lineups. It was
-generated for \(n = 1\), \(\sigma = 1\), \(j = 1\), \(X_{raw} \sim N\).
-For this lineup, x out of k observers picked the actual data plot. Each
-lineup is evaluated by 5 different subjects to provide reasonable
-estimates of the p-value.
+patterns.
+
+\includegraphics{paper_comparison_files/figure-latex/different-dist-1.pdf}
+
+Three replications are made for each of the parameter values shown in
+{[}table ref here{]} and thus, 192 different lineups. For each lineup,
+the null model was fit to the actual dataset to obtain residuals and
+parameter estimates. The actual data plot was drawn as standard residual
+plot with raw residuals on the y-axis and fitted values on the x-axis.
+The 19 null datasets were generated by the residual rotation technique,
+and plotted in the same way. The actual data plot was randomly placed
+among these null dadta plots to produce the lineup. {[}figure ref
+here{]} is an example of one of these lineups. It was generated for
+\(n = 1\), \(\sigma = 1\), \(j = 1\), \(X_{raw} \sim N\). For this
+lineup, x out of k observers picked the actual data plot. Each lineup is
+evaluated by 5 different subjects to provide reasonable estimates of the
+p-value.
 

### 2022-07-28 paper_comparison/paper_comparison/paper_comparison.Rmd
reorder code

@@
 set.seed(10086)
+```
+
+```{r get-third-result}
+if (!file.exists(here::here("data/third_study/polynomials_lineup.rds"))) {
+  polynomials_lineup <- get_polynomials_lineup()
+  saveRDS(polynomials_lineup, here::here("data/third_study/polynomials_lineup.rds"))
+} else {
+  polynomials_lineup <- readRDS(here::here("data/third_study/polynomials_lineup.rds"))
+}
+```
+
+```{r get-conv-result}
+stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
+
+if (!file.exists(here::here("data/third_study/conv_result.rds"))) {
+  conv_result <- list()
+  for (i in 1:100) {
+    conv_result[[i]] <- map(1:100000,
+                            function(x) {
+                     shape <- sample(1:4, 1)
+                     e_sigma <- sample(c(0.5, 1, 2, 4), 1)
+                     x_dist <- sample(c("uniform", "normal", "lognormal", "even_discrete"), 1)
+                     x <- switch(x_dist,
+                                 uniform = rand_uniform(-1, 1),
+                                 normal = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))},
+                                 lognormal = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))},
+                                 even_discrete = rand_uniform_d(k = 5, even = TRUE))
+                     mod <- poly_model(shape, x = x, sigma = e_sigma)
+                     n <- sample(c(50, 100, 300), 1)
+                     tmp_dat <- mod$gen(n)
+
+                     tibble(shape = shape,
+                            e_sigma = e_sigma,
+                            x_dist = x_dist,
+                            n = n,
+                            effect_size = mod$effect_size(tmp_dat),
+                            p_value = mod$test(tmp_dat)$p_value,
+                            reject = p_value < 0.05)
+                   }) %>%
+  reduce(bind_rows) %>%
+  mutate(reject = as.numeric(reject))
+  }
+
+  saveRDS(conv_result, here::here("data/third_study/conv_result.rds"))
+} else {
+  conv_result <- readRDS(here::here("data/third_study/conv_result.rds"))
+}
+
 ```
@@
 They were asked to select one or more plots that are mostly different from others, provide a reason for their selections, and express how different they think the selected plots are from others.
+If there was no noticeable difference between plots in a lineup, subject could choose to select zero lineups without providing the reason.
 Information about gender, age, education, and previous experience in visual experiment were also collected.
@@
 Clearly, omitted-variable bias will present since the null model leaves out the higher order term.
-Data were simulated using four different order of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$), three different sample sizes (n = 50, 100, 300) , four different standard deviations of the error ($\sigma$ = 0.5, 1, 2, 4) and four different distribution of $\boldsymbol{X}_{raw}$: (1) $U(a = -1, b = 1)$, (2) $N(\mu = 0, \sigma = 0.3)$, (3) $lognormal(\mu = 0, \sigma = 0.6)/3$ and (4) $u\{a = 1, b = 5\}$.
-A summary of the parameters used in this experiment is given in [table ref here]
+Data were simulated using four different order of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$), three different sample sizes (n = 50, 100, 300) , four different standard deviations of the error ($\sigma$ = 0.5, 1, 2, 4) and four different distribution of $\boldsymbol{X}_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
+A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}.
 
-The values of $j$ was chosen so that visually different shapes of non-linearity are included in the residual plot. These include "U" shape, "S" shape, "M" shape and "Triple-U" shape.
+The values of $j$ was chosen so that different shapes of non-linearity are included in the residual plot. These include "U" shape, "S" shape, "M" shape and "Triple-U" shape.
 
@@
 The range of $\sigma$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated and the estimated power curve would be smooth and continuous.
-Four different distribution were used to genearte $\boldsymbol{X}_raw$.
 
@@
 
-The uniform and the normal distribution are symmetric and commonly assumed in statistical model.
+Four different distribution were used to generate $\boldsymbol{X}_{raw}$.
+The uniform and the normal distribution are symmetric and commonly assumed in statistical models.
 The adjusted log-normal distribution provides skewed density.
@@
 VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
-  ggtitle(quote(N(0,0.3))) -> p2
+  ggtitle(quote(N(0,0.3^2))) -> p2
 VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
-  ggtitle(quote(lognormal(0,0.6)/3)) -> p3
+  ggtitle(quote(lognormal(0,0.6^2)/3)) -> p3
 VI_MODEL$plot(poly_model(shape = 1, x = rand_uniform_d(k = 5, even = TRUE), sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
@@
 
-Three replications are made for each of the parameter values shown in [table ref here] and thus, 192 different lineups.
-For each lineup, the null model was fit to the actual dataset to obtain residuals and parameter estimates.
-The actual data plot was drawn as standard residual plot with raw residuals on the y-axis and fitted values on the x-axis.
+Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in 192 different lineups.
+For each lineup, the actual data plot was drawn as standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis.
 The 19 null datasets were generated by the residual rotation technique, and plotted in the same way.
-The actual data plot was randomly placed among these null dadta plots to produce the lineup.
-[figure ref here] is an example of one of these lineups.
-It was generated for $n = 1$, $\sigma = 1$, $j = 1$, $X_{raw} \sim N$.
+The lineup consisted of 20 residual plots with one randomly placed actual data plot.
+Figure \ref{fig:example-lineup} is an example of one of these lineups.
+It was produced by using $n = 300$, $j = 6$, $\sigma = 0.5$ and $N(0.0.3^2)$ as the distribution of $X_{raw}$.
+The actual data plot location is four.
+All five subjects correctly identified the actual data plot for this lineup.
+
+```{r results='asis'}
+data.frame(n = c(50, 100, 300, ""),
+           j = c(2, 3, 6, 18),
+           sigma = c(0.5, 1, 2, 4),
+           x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$")) %>%
+  knitr::kable(col.names = NULL,
+               escape = FALSE,
+               format = "latex",
+               booktabs = TRUE, 
+               caption  = 'Parameter values for $n$, $j$ $\\sigma$, $X_{raw}$', 
+               label = "parameter-table" ) %>%
+  kableExtra::add_header_above(c("Sample size\n(sym3)", "Order of Hermite polynomial\n(sym1)", "Error SD\n(sym2)", "Distribution of sym4\n")) %>%
+  as.character() %>%
+  gsub("sym1", "$j$", .) %>%
+  gsub("sym2", "$\\\\sigma$", .) %>%
+  gsub("sym3", "$n$", .) %>%
+  gsub("sym4", "$X_{raw}$", .) %>%
+  cat()
+  # gsub("\\\\\\$", "$", .) %>%
+  # cat()
+```
+
+```{r example-lineup, fig.cap="Example lineup \\label{fig:example-lineup}"}
+VI_MODEL$plot_lineup(polynomials_lineup[[23]]$data, 
+                     remove_grid_line = TRUE, 
+                     theme = theme_light(),
+                     remove_axis = TRUE)
+```
+
+<!-- It was generated for $n = 1$, $\sigma = 1$, $j = 1$, $X_{raw} \sim N$.
 For this lineup, x out of k observers picked the actual data plot.
 Each lineup is evaluated by 5 different subjects to provide reasonable estimates of the p-value.
+-->
 
-To satisfy the plan described above, we needed to recruit at least 160 subjects.
+In addition, each lineup is evaluated by five different subjects to provide reasonable estimates of the visual p-value.
+Thus, $192 \times 3 \times 5 / (20 - 2) = 160$ subjects were recruited to satisfy the experimental design described above.
 
@@
 
-## Data cleaning
+## Data preprocessing
+
+Subjects recruited from Prolific received a fixed payment for submitting quality responses.
+However, some subjects will try to maximize their earnings for minimum effort.
+If a subject objectively demonstrated clear low-effort throughout the experiment, i.e., failed both attention checks, we rejected the submission and the subject would not be paid.
 
@@
 
-## Model fitting
-
-For each parameter combination, effect $E$ is derived as $E = \frac{1}{2}(\boldsymbol{R}_a\boldsymbol{X}_b)'diag(\sigma^2\boldsymbol{R}_a)\boldsymbol{R}_a\boldsymbol{X}_b$ from the Kullback-Leibler divergence.
-[KL divergence ref here] (The derivation is provided in [appendix ref here].).
-The logistic regression is fit using $E$ as the only fixed effect covariate
-
-# Demographic summary
+## Demographic summary
 
@@
 ```
+
+## Model fitting
+
+For each parameter combination, effect $E$ is derived as $E = \frac{1}{2}(\boldsymbol{R}_a\boldsymbol{X}_b)'diag(\sigma^2\boldsymbol{R}_a)\boldsymbol{R}_a\boldsymbol{X}_b$ from the Kullback-Leibler divergence.
+[KL divergence ref here] (The derivation is provided in [appendix ref here].).
+The logistic regression is fit using $E$ as the only fixed effect covariate
 
@@
 From the figures, we can observe ...
-
-```{r get-conv-result}
-stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
-
-if (!file.exists(here::here("data/third_study/conv_result.rds"))) {
-  conv_result <- list()
-  for (i in 1:100) {
-    conv_result[[i]] <- map(1:100000,
-                            function(x) {
-                     shape <- sample(1:4, 1)
-                     e_sigma <- sample(c(0.5, 1, 2, 4), 1)
-                     x_dist <- sample(c("uniform", "normal", "lognormal", "even_discrete"), 1)
-                     x <- switch(x_dist,
-                                 uniform = rand_uniform(-1, 1),
-                                 normal = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))},
-                                 lognormal = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))},
-                                 even_discrete = rand_uniform_d(k = 5, even = TRUE))
-                     mod <- poly_model(shape, x = x, sigma = e_sigma)
-                     n <- sample(c(50, 100, 300), 1)
-                     tmp_dat <- mod$gen(n)
-
-                     tibble(shape = shape,
-                            e_sigma = e_sigma,
-                            x_dist = x_dist,
-                            n = n,
-                            effect_size = mod$effect_size(tmp_dat),
-                            p_value = mod$test(tmp_dat)$p_value,
-                            reject = p_value < 0.05)
-                   }) %>%
-  reduce(bind_rows) %>%
-  mutate(reject = as.numeric(reject))
-  }
-
-  saveRDS(conv_result, here::here("data/third_study/conv_result.rds"))
-} else {
-  conv_result <- readRDS(here::here("data/third_study/conv_result.rds"))
-}
-
-```
-
-```{r get-third-result}
-if (!file.exists(here::here("data/third_study/polynomials_lineup.rds"))) {
-  polynomials_lineup <- get_polynomials_lineup()
-  saveRDS(polynomials_lineup, here::here("data/third_study/polynomials_lineup.rds"))
-} else {
-  polynomials_lineup <- readRDS(here::here("data/third_study/polynomials_lineup.rds"))
-}
-```
 

### 2022-07-28 paper_comparison/paper_comparison/paper_comparison.tex
reorder code

@@
 \newcommand*{\Perm}[2]{{}^{#1}\!P_{#2}}%
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 
@@
 different from others, provide a reason for their selections, and
-express how different they think the selected plots are from others.
+express how different they think the selected plots are from others. If
+there was no noticeable difference between plots in a lineup, subject
+could choose to select zero lineups without providing the reason.
 Information about gender, age, education, and previous experience in
@@
 deviations of the error (\(\sigma\) = 0.5, 1, 2, 4) and four different
-distribution of \(\boldsymbol{X}_{raw}\): (1) \(U(a = -1, b = 1)\), (2)
-\(N(\mu = 0, \sigma = 0.3)\), (3) \(lognormal(\mu = 0, \sigma = 0.6)/3\)
-and (4) \(u\{a = 1, b = 5\}\). A summary of the parameters used in this
-experiment is given in {[}table ref here{]}
+distribution of \(\boldsymbol{X}_{raw}\): (1) \(U(-1, 1)\), (2)
+\(N(0, 0.3^2)\), (3) \(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). A
+summary of the parameters used in this experiment is given in Table
+\ref{tab:parameter-table}.
 
-The values of \(j\) was chosen so that visually different shapes of
-non-linearity are included in the residual plot. These include ``U''
-shape, ``S'' shape, ``M'' shape and ``Triple-U'' shape.
+The values of \(j\) was chosen so that different shapes of non-linearity
+are included in the residual plot. These include ``U'' shape, ``S''
+shape, ``M'' shape and ``Triple-U'' shape.
 
@@
 were generated and the estimated power curve would be smooth and
-continuous. Four different distribution were used to genearte
-\(\boldsymbol{X}_raw\).
+continuous.
 
@@
 
-The uniform and the normal distribution are symmetric and commonly
-assumed in statistical model. The adjusted log-normal distribution
-provides skewed density. And the discrete uniform distribution provides
-discreteness in residual plot, which could enrich the pool of visual
-patterns.
+Four different distribution were used to generate
+\(\boldsymbol{X}_{raw}\). The uniform and the normal distribution are
+symmetric and commonly assumed in statistical models. The adjusted
+log-normal distribution provides skewed density. And the discrete
+uniform distribution provides discreteness in residual plot, which could
+enrich the pool of visual patterns.
 
@@
 Three replications are made for each of the parameter values shown in
-{[}table ref here{]} and thus, 192 different lineups. For each lineup,
-the null model was fit to the actual dataset to obtain residuals and
-parameter estimates. The actual data plot was drawn as standard residual
-plot with raw residuals on the y-axis and fitted values on the x-axis.
-The 19 null datasets were generated by the residual rotation technique,
-and plotted in the same way. The actual data plot was randomly placed
-among these null dadta plots to produce the lineup. {[}figure ref
-here{]} is an example of one of these lineups. It was generated for
-\(n = 1\), \(\sigma = 1\), \(j = 1\), \(X_{raw} \sim N\). For this
-lineup, x out of k observers picked the actual data plot. Each lineup is
-evaluated by 5 different subjects to provide reasonable estimates of the
-p-value.
+Table \ref{tab:parameter-table} resulting in 192 different lineups. For
+each lineup, the actual data plot was drawn as standard residual plot of
+the null model with raw residuals on the y-axis and fitted values on the
+x-axis. The 19 null datasets were generated by the residual rotation
+technique, and plotted in the same way. The lineup consisted of 20
+residual plots with one randomly placed actual data plot. Figure
+\ref{fig:example-lineup} is an example of one of these lineups. It was
+produced by using \(n = 300\), \(j = 6\), \(\sigma = 0.5\) and
+\(N(0.0.3^2)\) as the distribution of \(X_{raw}\). The actual data plot
+location is four. All five subjects correctly identified the actual data
+plot for this lineup.
 
-To satisfy the plan described above, we needed to recruit at least 160
-subjects.
+\begin{table}
+
+\caption{\label{tab:parameter-table}Parameter values for $n$, $j$ $\sigma$, $X_{raw}$}
+\centering
+\begin{tabular}[t]{lrrl}
+\toprule
+\multicolumn{1}{c}{\makecell[c]{Sample size\\($n$)}} & \multicolumn{1}{c}{\makecell[c]{Order of Hermite polynomial\\($j$)}} & \multicolumn{1}{c}{\makecell[c]{Error SD\\($\sigma$)}} & \multicolumn{1}{c}{\makecell[c]{Distribution of $X_{raw}$\\}} \\
+\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4}
+50 & 2 & 0.5 & $U(-1, 1)$\\
+100 & 3 & 1.0 & $N(0, 0.3^2)$\\
+300 & 6 & 2.0 & $lognormal(0, 0.6^2)/3$\\
+ & 18 & 4.0 & $U\{1, 5\}$\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{figure}
+\centering
+\includegraphics{paper_comparison_files/figure-latex/example-lineup-1.pdf}
+\caption{Example lineup \label{fig:example-lineup}}
+\end{figure}
+
+In addition, each lineup is evaluated by five different subjects to
+provide reasonable estimates of the visual p-value. Thus,
+\(192 \times 3 \times 5 / (20 - 2) = 160\) subjects were recruited to
+satisfy the experimental design described above.
 
@@
 
-\hypertarget{data-cleaning}{%
-\subsection{Data cleaning}\label{data-cleaning}}
+\hypertarget{data-preprocessing}{%
+\subsection{Data preprocessing}\label{data-preprocessing}}
+
+Subjects recruited from Prolific received a fixed payment for submitting
+quality responses. However, some subjects will try to maximize their
+earnings for minimum effort. If a subject objectively demonstrated clear
+low-effort throughout the experiment, i.e., failed both attention
+checks, we rejected the submission and the subject would not be paid.
 
@@
 
+\hypertarget{demographic-summary}{%
+\subsection{Demographic summary}\label{demographic-summary}}
+
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-3-1.pdf}
+
 \hypertarget{model-fitting}{%
@@
 regression is fit using \(E\) as the only fixed effect covariate
-
-\hypertarget{demographic-summary}{%
-\section{Demographic summary}\label{demographic-summary}}
-
-\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-2-1.pdf}
 
@@
 
-\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-3-1.pdf}
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-4-1.pdf}
 

### 2022-07-28 paper_comparison/paper_comparison/paper_comparison.Rmd
fin data preprocessing

@@
 
-As pointed out by @vanderplas2021statistical, though equation (\ref{eq:pvaluesingle}) is trivial, but it doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup.
+As pointed out by @vanderplas2021statistical, though Equation (\ref{eq:pvaluesingle}) is trivial, but it doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup.
 And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup.
@@
 
-Note that the use of equation (\ref{eq:pvalue-beta-binomial}) requires the estimation of $\hat{\alpha}$, which largely depends on the null model, the type of the plot and other aesthetic features.
+Note that the use of Equation (\ref{eq:pvalue-beta-binomial}) requires the estimation of $\hat{\alpha}$, which largely depends on the null model, the type of the plot and other aesthetic features.
 They suggested to estimate $\hat{\alpha}$ visually based on the selections of null plots of the experimental data, or to estimate $\hat{\alpha}$ numerically based on several additional Rorschach lineups, which is a type of lineup containing only null plots.
@@
 Subjects for both two experiments were recruited from Prolific web service [ref here].
-They were asked to select one or more plots that are mostly different from others, provide a reason for their selections, and express how different they think the selected plots are from others.
-If there was no noticeable difference between plots in a lineup, subject could choose to select zero lineups without providing the reason.
+Data screening procedure was applied during the recruitment, subjects were required to be fluent in English, with $98\%$ minimum approval rate and 10 minimum submissions.
+During the experiment, they were asked to select one or more plots that are mostly different from others, provide a reason for their selections, and express how different they think the selected plots are from others.
+If there was no noticeable difference between plots in a lineup, subject could choose to select zero plots without providing the reason.
 Information about gender, age, education, and previous experience in visual experiment were also collected.
@@
 Figure \ref{fig:example-lineup} is an example of one of these lineups.
-It was produced by using $n = 300$, $j = 6$, $\sigma = 0.5$ and $N(0.0.3^2)$ as the distribution of $X_{raw}$.
+It was produced by using $n = 300$, $j = 6$, $\sigma = 0.5$ and $X_{raw} \sim N(0.0.3^2)$.
 The actual data plot location is four.
@@
 
-Subjects recruited from Prolific received a fixed payment for submitting quality responses.
+Subjects recruited from Prolific received a fixed payment for participating in the experiment.
 However, some subjects will try to maximize their earnings for minimum effort.
-If a subject objectively demonstrated clear low-effort throughout the experiment, i.e., failed both attention checks, we rejected the submission and the subject would not be paid.
+During the review of submissions, if we found a subject objectively demonstrated clear low-effort throughout the experiment, i.e., failed all attention checks, we rejected the submission.
+The rejected submissions will be removed immediately, and Prolific will automatically recruit another subject as substitution.
+Therefore, we only paid for approved submissions and no further data screening procedure needed to be applied on the collected data.
 
-Participants recruited from Prolific are paid for their efforts.
-Some participants will try to maximize their earnings for minimum effort, which can affect the result from the data.
-For the purpose of identifying these participants and cleaning the data, two attention checks are included in the block of 20 lineups.
-If the subject failed to identify at least one of the actual data plot in those attention checks, we reject the submission and the subject will not be paid.
-All attention checks are removed from the final data.
-[table ref here] tabulates the number of subjects, genders, and lineups evaluated after applying the data screening procedure.
+A subject was allowed to select zero plots for a lineup, which literally means there was no visble difference between plots,
+but the simulated p-value given in Equation (\ref{eq:p-value-multi}) will effectively drop the subject from the simulation for this case, leading to inaccurate estimation of $p$-value.
+Therefore, we treated this case as making one selection but failing to identify the actual data plot so that Equation (\ref{eq:p-value-multi}) can be applied correctly.
 
 ## Demographic summary
+
+[table ref here] tabulates the number of subjects, genders, and lineups evaluated after applying the data screening procedure.
 

### 2022-07-28 paper_comparison/paper_comparison/paper_comparison.tex
fin data preprocessing

@@
 
-As pointed out by \citet{vanderplas2021statistical}, though equation
+As pointed out by \citet{vanderplas2021statistical}, though Equation
 (\ref{eq:pvaluesingle}) is trivial, but it doesn't take into account the
@@
 
-Note that the use of equation (\ref{eq:pvalue-beta-binomial}) requires
+Note that the use of Equation (\ref{eq:pvalue-beta-binomial}) requires
 the estimation of \(\hat{\alpha}\), which largely depends on the null
@@
 for both two experiments were recruited from Prolific web service {[}ref
-here{]}. They were asked to select one or more plots that are mostly
-different from others, provide a reason for their selections, and
-express how different they think the selected plots are from others. If
-there was no noticeable difference between plots in a lineup, subject
-could choose to select zero lineups without providing the reason.
-Information about gender, age, education, and previous experience in
-visual experiment were also collected. No subject was shown the same
-lineup twice.
+here{]}. Data screening procedure was applied during the recruitment,
+subjects were required to be fluent in English, with \(98\%\) minimum
+approval rate and 10 minimum submissions. During the experiment, they
+were asked to select one or more plots that are mostly different from
+others, provide a reason for their selections, and express how different
+they think the selected plots are from others. If there was no
+noticeable difference between plots in a lineup, subject could choose to
+select zero plots without providing the reason. Information about
+gender, age, education, and previous experience in visual experiment
+were also collected. No subject was shown the same lineup twice.
 
@@
 produced by using \(n = 300\), \(j = 6\), \(\sigma = 0.5\) and
-\(N(0.0.3^2)\) as the distribution of \(X_{raw}\). The actual data plot
-location is four. All five subjects correctly identified the actual data
-plot for this lineup.
+\(X_{raw} \sim N(0.0.3^2)\). The actual data plot location is four. All
+five subjects correctly identified the actual data plot for this lineup.
 
@@
 
-Subjects recruited from Prolific received a fixed payment for submitting
-quality responses. However, some subjects will try to maximize their
-earnings for minimum effort. If a subject objectively demonstrated clear
-low-effort throughout the experiment, i.e., failed both attention
-checks, we rejected the submission and the subject would not be paid.
+Subjects recruited from Prolific received a fixed payment for
+participating in the experiment. However, some subjects will try to
+maximize their earnings for minimum effort. During the review of
+submissions, if we found a subject objectively demonstrated clear
+low-effort throughout the experiment, i.e., failed all attention checks,
+we rejected the submission. The rejected submissions will be removed
+immediately, and Prolific will automatically recruit another subject as
+substitution. Therefore, we only paid for approved submissions and no
+further data screening procedure needed to be applied on the collected
+data.
 
-Participants recruited from Prolific are paid for their efforts. Some
-participants will try to maximize their earnings for minimum effort,
-which can affect the result from the data. For the purpose of
-identifying these participants and cleaning the data, two attention
-checks are included in the block of 20 lineups. If the subject failed to
-identify at least one of the actual data plot in those attention checks,
-we reject the submission and the subject will not be paid. All attention
-checks are removed from the final data. {[}table ref here{]} tabulates
-the number of subjects, genders, and lineups evaluated after applying
-the data screening procedure.
+A subject was allowed to select zero plots for a lineup, which literally
+means there was no visble difference between plots, but the simulated
+p-value given in Equation (\ref{eq:p-value-multi}) will effectively drop
+the subject from the simulation for this case, leading to inaccurate
+estimation of \(p\)-value. Therefore, we treated this case as making one
+selection but failing to identify the actual data plot so that Equation
+(\ref{eq:p-value-multi}) can be applied correctly.
 
@@
 \subsection{Demographic summary}\label{demographic-summary}}
+
+{[}table ref here{]} tabulates the number of subjects, genders, and
+lineups evaluated after applying the data screening procedure.
 

### 2022-07-28 paper_comparison/paper_comparison/paper_comparison.Rmd
draft model fitting

@@
 If there was no noticeable difference between plots in a lineup, subject could choose to select zero plots without providing the reason.
-Information about gender, age, education, and previous experience in visual experiment were also collected.
+Information about preferred pronoun, age group, education, and previous experience in visual experiment were also collected.
 No subject was shown the same lineup twice.
@@
 
-VI_MODEL$plot(ori_dat) + ggtitle(quote(He[2]:"U shape")) -> p1
+VI_MODEL$plot(ori_dat) + 
+  ggtitle(quote(He[2]:"U shape")) + 
+  theme_light() -> p1
+
 VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote(He[3]:"S shape")) -> p2
+  ggtitle(quote(He[3]:"S shape")) + 
+  theme_light() -> p2
+
 VI_MODEL$plot(poly_model(shape = 3, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote(He[6]:"M shape")) -> p3
+  ggtitle(quote(He[6]:"M shape")) + 
+  theme_light() -> p3
+
 VI_MODEL$plot(poly_model(shape = 4, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote(He[18]:"Triple-U shape")) -> p4
+  ggtitle(quote(He[18]:"Triple-U shape")) + 
+  theme_light() -> p4
 
@@
 
-VI_MODEL$plot(ori_dat) + ggtitle(quote(sigma~"="~0.5)) -> p1
+VI_MODEL$plot(ori_dat) + 
+  ggtitle(quote(sigma~"="~0.5)) + 
+  theme_light() -> p1
+
 VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1)$gen(300, computed = select(ori_dat, x))) +
-  ggtitle(quote(sigma~"="~1)) -> p2
+  ggtitle(quote(sigma~"="~1)) + 
+  theme_light() -> p2
+
 VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 2)$gen(300, computed = select(ori_dat, x))) +
-  ggtitle(quote(sigma~"="~2)) -> p3
+  ggtitle(quote(sigma~"="~2)) + 
+  theme_light() -> p3
+
 VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 4)$gen(300, computed = select(ori_dat, x))) +
-  ggtitle(quote(sigma~"="~4)) -> p4
+  ggtitle(quote(sigma~"="~4)) + 
+  theme_light() -> p4
 
@@
 
-VI_MODEL$plot(ori_dat) + ggtitle(quote(U(-1,1))) -> p1
+VI_MODEL$plot(ori_dat) + 
+  ggtitle(quote(U(-1,1))) + 
+  theme_light() -> p1
+
 VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
-  ggtitle(quote(N(0,0.3^2))) -> p2
+  ggtitle(quote(N(0,0.3^2))) + 
+  theme_light() -> p2
+
 VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
-  ggtitle(quote(lognormal(0,0.6^2)/3)) -> p3
+  ggtitle(quote(lognormal(0,0.6^2)/3)) + 
+  theme_light() -> p3
+
 VI_MODEL$plot(poly_model(shape = 1, x = rand_uniform_d(k = 5, even = TRUE), sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
-  ggtitle(quote(U~"{1,5}")) -> p4
+  ggtitle(quote(U~"{1,5}")) + 
+  theme_light() -> p4
 
@@
 
+In overall, there were a total of 3200 lineup evaluations made by 160 subjects in experiment I, where 320 lineup evaluations were attention checks and were not used in the following analysis.
+The collated dataset is provided in `polynomials` of the `visage` `R` package.
+
 ## Demographic summary
 
-[table ref here] tabulates the number of subjects, genders, and lineups evaluated after applying the data screening procedure.
+Table \ref{tab:demographic-table} tabulates the number of subjects, preferred pronouns, education backgrounds, age groups, and previous experience in visual experiment.
+Figure \ref{fig:demographic-summary-plot} visulizes the marginal distribution of each of the category.
+The collated data was a balanced sample among male and female.
+Most of the ages of subject were between 18 to 39, while many of them were between 18 to 24.
+Very few subjects were awarded degree higher than Bachelor Degree.
+Around 40% of subjects had previous experience in visual experiment.
 
-```{r}
+```{r results="asis"}
+
+polynomials %>%
+  group_by(set) %>%
+  slice_sample(n = 1) %>%
+  ungroup() %>%
+  count(pronoun, education, age_group, previous_experience) %>%
+  arrange(desc(n)) %>%
+  knitr::kable(col.names = c("Preferred pronoun", "Education", "Age group", "Previous experience", "Subject"),
+               escape = FALSE,
+               format = "latex",
+               linesep = "",
+               booktabs = TRUE, 
+               caption  = 'Summary of demographic information', 
+               label = "demographic-table" ) %>%
+  cat()
+```
+
+```{r demographic-summary-plot, fig.cap="Summary of demographic information. \\label{fig:demographic-summary-plot}"}
 polynomials %>%
@@
 
-For each parameter combination, effect $E$ is derived as $E = \frac{1}{2}(\boldsymbol{R}_a\boldsymbol{X}_b)'diag(\sigma^2\boldsymbol{R}_a)\boldsymbol{R}_a\boldsymbol{X}_b$ from the Kullback-Leibler divergence.
-[KL divergence ref here] (The derivation is provided in [appendix ref here].).
-The logistic regression is fit using $E$ as the only fixed effect covariate
+For each parameter combination, effect $E$ is derived from the Kullback-Leibler divergence (the derivation is provided in [appendix ref here]) formulated as
+
+\begin{equation} \label{eq:effect-size-ex1}
+E = \frac{1}{2\sigma^2}\boldsymbol{X}_b'\boldsymbol{R}_a'(diag(\boldsymbol{R}_a))^{-1}\boldsymbol{R}_a\boldsymbol{X}_b,
+\end{equation}
+
+where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$.
+
+The logistic regression is fit using $log_e(E)$ as the only fixed effect covariate for the power of visual test formulated as
+
+\begin{equation} \label{eq:logistic-regression-1-1}
+Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(\boldsymbol{E})),
+\end{equation}
+
+where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
+
+Table [table ref here] shows the parameter estimates of the logistic regression. [Discussion about the numeric estimates here]
+
+## Power comparison
+
+Figure [figure ref here] shows an overview of estimated power against natural logarithm
 
@@
 
-# Data processing
-
-# Results
+# Old
 

### 2022-07-28 paper_comparison/paper_comparison/paper_comparison.tex
draft model fitting

@@
 select zero plots without providing the reason. Information about
-gender, age, education, and previous experience in visual experiment
-were also collected. No subject was shown the same lineup twice.
+preferred pronoun, age group, education, and previous experience in
+visual experiment were also collected. No subject was shown the same
+lineup twice.
 
@@
 
+In overall, there were a total of 3200 lineup evaluations made by 160
+subjects in experiment I, where 320 lineup evaluations were attention
+checks and were not used in the following analysis. The collated dataset
+is provided in \texttt{polynomials} of the \texttt{visage} \texttt{R}
+package.
+
 \hypertarget{demographic-summary}{%
@@
 
-{[}table ref here{]} tabulates the number of subjects, genders, and
-lineups evaluated after applying the data screening procedure.
+Table \ref{tab:demographic-table} tabulates the number of subjects,
+preferred pronouns, education backgrounds, age groups, and previous
+experience in visual experiment. Figure
+\ref{fig:demographic-summary-plot} visulizes the marginal distribution
+of each of the category. The collated data was a balanced sample among
+male and female. Most of the ages of subject were between 18 to 39,
+while many of them were between 18 to 24. Very few subjects were awarded
+degree higher than Bachelor Degree. Around 40\% of subjects had previous
+experience in visual experiment.
 
-\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-3-1.pdf}
+\begin{table}
+
+\caption{\label{tab:demographic-table}Summary of demographic information}
+\centering
+\begin{tabular}[t]{llllr}
+\toprule
+Preferred pronoun & Education & Age group & Previous experience & Subject\\
+\midrule
+She & Diploma and Bachelor Degree & 18-24 & No & 17\\
+He & Diploma and Bachelor Degree & 18-24 & Yes & 14\\
+She & Diploma and Bachelor Degree & 18-24 & Yes & 12\\
+He & Diploma and Bachelor Degree & 18-24 & No & 11\\
+He & Diploma and Bachelor Degree & 25-39 & No & 10\\
+She & Diploma and Bachelor Degree & 25-39 & No & 10\\
+She & Diploma and Bachelor Degree & 25-39 & Yes & 9\\
+He & High School or below & 18-24 & No & 8\\
+She & High School or below & 18-24 & No & 7\\
+He & High School or below & 18-24 & Yes & 6\\
+He & Diploma and Bachelor Degree & 25-39 & Yes & 5\\
+He & Masters Degree & 25-39 & No & 5\\
+He & Masters Degree & 25-39 & Yes & 5\\
+He & High School or below & 25-39 & No & 4\\
+She & High School or below & 25-39 & No & 4\\
+She & Masters Degree & 25-39 & Yes & 4\\
+He & High School or below & 25-39 & Yes & 3\\
+She & High School or below & 18-24 & Yes & 3\\
+She & Masters Degree & 25-39 & No & 3\\
+He & Honours Degree & 25-39 & No & 2\\
+She & Diploma and Bachelor Degree & 40-54 & No & 2\\
+She & Honours Degree & 18-24 & No & 2\\
+He & High School or below & 40-54 & Yes & 1\\
+He & High School or below & 55-64 & No & 1\\
+He & Masters Degree & 40-54 & Yes & 1\\
+He & Masters Degree & 55-64 & No & 1\\
+Other & High School or below & 25-39 & No & 1\\
+She & High School or below & 40-54 & No & 1\\
+She & Honours Degree & 25-39 & No & 1\\
+She & Honours Degree & 25-39 & Yes & 1\\
+She & Masters Degree & 18-24 & No & 1\\
+She & Masters Degree & 40-54 & No & 1\\
+They & Diploma and Bachelor Degree & 18-24 & No & 1\\
+They & Diploma and Bachelor Degree & 25-39 & No & 1\\
+They & High School or below & 18-24 & No & 1\\
+They & High School or below & 25-39 & No & 1\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{figure}
+\centering
+\includegraphics{paper_comparison_files/figure-latex/demographic-summary-plot-1.pdf}
+\caption{Summary of demographic information.
+\label{fig:demographic-summary-plot}}
+\end{figure}
 
@@
 
-For each parameter combination, effect \(E\) is derived as
-\(E = \frac{1}{2}(\boldsymbol{R}_a\boldsymbol{X}_b)'diag(\sigma^2\boldsymbol{R}_a)\boldsymbol{R}_a\boldsymbol{X}_b\)
-from the Kullback-Leibler divergence. {[}KL divergence ref here{]} (The
-derivation is provided in {[}appendix ref here{]}.). The logistic
-regression is fit using \(E\) as the only fixed effect covariate
+For each parameter combination, effect \(E\) is derived from the
+Kullback-Leibler divergence (the derivation is provided in {[}appendix
+ref here{]}) formulated as
 
-\hypertarget{data-processing}{%
-\section{Data processing}\label{data-processing}}
+\begin{equation} \label{eq:effect-size-ex1}
+E = \frac{1}{2\sigma^2}\boldsymbol{X}_b'\boldsymbol{R}_a'(diag(\boldsymbol{R}_a))^{-1}\boldsymbol{R}_a\boldsymbol{X}_b,
+\end{equation}
 
-\hypertarget{results-1}{%
-\section{Results}\label{results-1}}
+where \(diag(.)\) is the diagonal matrix constructed from the diagonal
+elements of \(\boldsymbol{R}_a\).
+
+The logistic regression is fit using \(log_e(E)\) as the only fixed
+effect covariate for the power of visual test formulated as
+
+\begin{equation} \label{eq:logistic-regression-1-1}
+Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(\boldsymbol{E})),
+\end{equation}
+
+where \(\Lambda(.)\) is the standard logistic function given as
+\(\Lambda(z) = exp(z)/(1+exp(z))\).
+
+Table {[}table ref here{]} shows the parameter estimates of the logistic
+regression. {[}Discussion about the numeric estimates here{]}
+
+\hypertarget{power-comparison}{%
+\subsection{Power comparison}\label{power-comparison}}
+
+Figure {[}figure ref here{]} shows an overview of estimated power
+against natural logarithm
+
+\hypertarget{old}{%
+\section{Old}\label{old}}
 

### 2022-07-29 paper_comparison/paper_comparison/paper_comparison.Rmd
update power analysis

@@
   conv_result <- list()
-  for (i in 1:100) {
-    conv_result[[i]] <- map(1:100000,
+  for (i in 1:100)
+  {
+    conv_result[[i]] <- map(1:2000,
                             function(x) {
@@
                             p_value = mod$test(tmp_dat)$p_value,
-                            reject = p_value < 0.05)
+                            reject = p_value < 0.05,
+                            boot_id = i)
                    }) %>%
@@
 
+  conv_result <- conv_result %>%
+    reduce(bind_rows)
+
   saveRDS(conv_result, here::here("data/third_study/conv_result.rds"))
@@
   conv_result <- readRDS(here::here("data/third_study/conv_result.rds"))
+}
+
+```
+
+```{r}
+if (!file.exists(here::here("data/third_study/wrong_BP_conv_result.rds"))) {
+  wrong_BP_conv_result <- list()
+  for (i in 1:100)
+  {
+    wrong_BP_conv_result[[i]] <- map(1:2000,
+                            function(x) {
+                     shape <- sample(1:4, 1)
+                     e_sigma <- sample(c(0.5, 1, 2, 4), 1)
+                     x_dist <- sample(c("uniform", "normal", "lognormal", "even_discrete"), 1)
+                     x <- switch(x_dist,
+                                 uniform = rand_uniform(-1, 1),
+                                 normal = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))},
+                                 lognormal = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))},
+                                 even_discrete = rand_uniform_d(k = 5, even = TRUE))
+                     mod <- poly_model(shape, x = x, sigma = e_sigma)
+                     n <- sample(c(50, 100, 300), 1)
+                     tmp_dat <- mod$gen(n)
+
+                     tibble(shape = shape,
+                            e_sigma = e_sigma,
+                            x_dist = x_dist,
+                            n = n,
+                            effect_size = mod$effect_size(tmp_dat),
+                            p_value = HETER_MODEL$test(tmp_dat)$p_value,
+                            reject = p_value < 0.05,
+                            boot_id = i)
+                   }) %>%
+  reduce(bind_rows) %>%
+  mutate(reject = as.numeric(reject))
+  }
+
+  wrong_BP_conv_result <- wrong_BP_conv_result %>%
+    reduce(bind_rows)
+
+  saveRDS(wrong_BP_conv_result, here::here("data/third_study/wrong_BP_conv_result.rds"))
+} else {
+  wrong_BP_conv_result <- readRDS(here::here("data/third_study/wrong_BP_conv_result.rds"))
+}
+
+```
+
+```{r}
+if (!file.exists(here::here("data/third_study/wrong_WS_conv_result.rds"))) {
+  wrong_WS_conv_result <- list()
+  for (i in 1:100)
+  {
+    wrong_WS_conv_result[[i]] <- map(1:2000,
+                            function(x) {
+                     shape <- sample(1:4, 1)
+                     e_sigma <- sample(c(0.5, 1, 2, 4), 1)
+                     x_dist <- sample(c("uniform", "normal", "lognormal", "even_discrete"), 1)
+                     x <- switch(x_dist,
+                                 uniform = rand_uniform(-1, 1),
+                                 normal = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))},
+                                 lognormal = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))},
+                                 even_discrete = rand_uniform_d(k = 5, even = TRUE))
+                     mod <- poly_model(shape, x = x, sigma = e_sigma)
+                     n <- sample(c(50, 100, 300), 1)
+                     tmp_dat <- mod$gen(n)
+
+                     tibble(shape = shape,
+                            e_sigma = e_sigma,
+                            x_dist = x_dist,
+                            n = n,
+                            effect_size = mod$effect_size(tmp_dat),
+                            p_value = shapiro.test(tmp_dat$.resid)$p.value,
+                            reject = p_value < 0.05,
+                            boot_id = i)
+                   }) %>%
+  reduce(bind_rows) %>%
+  mutate(reject = as.numeric(reject))
+  }
+
+  wrong_WS_conv_result <- wrong_WS_conv_result %>%
+    reduce(bind_rows)
+
+  saveRDS(wrong_WS_conv_result, here::here("data/third_study/wrong_WS_conv_result.rds"))
+} else {
+  wrong_WS_conv_result <- readRDS(here::here("data/third_study/wrong_WS_conv_result.rds"))
 }
@@
 
+```{r}
+conv_test_plot <- function() {
+  conv_result %>%
+  ggplot() +
+  stat_smooth(geom = "line", 
+              aes(log(effect_size), reject, group = boot_id, col = "F-test"), 
+              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) +
+  stat_smooth(geom = "line", 
+              aes(log(effect_size), reject, col = "F-test"), 
+              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1) +
+  labs(col = "Test") +
+  theme_light() +
+  xlab(quote(log[e](E))) +
+  ylab("Power")
+}
+
+conv_test_plot()
+```
+
+```{r}
+add_wrong_BP_conv_test_plot_layers <- function(p) {
+  p +
+  stat_smooth(data = wrong_BP_conv_result, geom = "line", 
+              aes(log(effect_size), reject, group = boot_id, col = "BP-test"), 
+              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) +
+  stat_smooth(data = wrong_BP_conv_result, geom = "line", 
+              aes(log(effect_size), reject, col = "BP-test"), 
+              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1)
+}
+
+conv_test_plot() %>%
+  add_wrong_BP_conv_test_plot_layers()
+```
+
+```{r}
+add_wrong_WS_conv_test_plot_layers <- function(p) {
+  p +
+  stat_smooth(data = wrong_WS_conv_result, geom = "line", 
+              aes(log(effect_size), reject, group = boot_id, col = "WS-test"), 
+              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) +
+  stat_smooth(data = wrong_WS_conv_result, geom = "line", 
+              aes(log(effect_size), reject, col = "WS-test"), 
+              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1)
+}
+
+conv_test_plot() %>%
+  add_wrong_BP_conv_test_plot_layers() %>%
+  add_wrong_WS_conv_test_plot_layers()
+```
+
+```{r}
+boot_dat <- function(dat, times = 100) {
+  result <- mutate(dat, boot_id = 1)
+  for (i in 2:times) {
+    result <- bind_rows(result, mutate(dat[sample(1:nrow(dat), replace = TRUE), ], boot_id = i)) 
+  }
+  result
+}
+
+add_visual_test_plot_layers <- function(p) {
+  visual_dat <- polynomials %>%
+  group_by(lineup_id) %>%
+  summarise(across(c(type:p_value), first)) %>%
+  filter(lineup_id < 577) %>%
+  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
+  boot_dat(times = 100)
+
+  p +
+  stat_smooth(data = filter(visual_dat, boot_id == 1), geom = "line", 
+              aes(log(effect_size), reject, group = boot_id, col = "Visual"), 
+              method = "glm", method.args = list(family = binomial), se = FALSE, alpha = 1) +
+  stat_smooth(data = visual_dat, geom = "line", 
+              aes(log(effect_size), reject, group = boot_id, col = "Visual"), 
+              method = "glm", method.args = list(family = binomial), se = FALSE, alpha = 0.05)
+
+}
+
+control_col <- function(p, ...) {
+  p + scale_color_manual(values = unlist(list(...)))
+}
+
+conv_test_plot() %>%
+  add_wrong_BP_conv_test_plot_layers() %>%
+  add_wrong_WS_conv_test_plot_layers() %>%
+  add_visual_test_plot_layers()
+```
+
+```{r}
+polynomials %>%
+  group_by(lineup_id) %>%
+  summarise(across(c(type:p_value), first)) %>%
+  filter(lineup_id < 577) %>%
+  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
+  ggplot() +
+  geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  xlab("Natural logarithm of effect size") +
+  ylab("Power")
+```
+
 <!-- # time taken vs order of the plot -->

### 2022-07-29 paper_comparison/paper_comparison/paper_comparison.tex
update power analysis

@@
 
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-6-1.pdf}
+
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-7-1.pdf}
+
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-8-1.pdf}
+
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-9-1.pdf}
+
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-10-1.pdf}
+
 \hypertarget{old}{%
@@
 
-\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-4-1.pdf}
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-11-1.pdf}
 

### 2022-07-30 paper_comparison/paper_comparison/paper_comparison.Rmd
modify power plot

@@
 
-# Experimental design
+## Effectiveness of visual test in regression diagnostics
 
@@
 Particularly, its ability to detect non-linearity and heteroskedasticity compared to F-test and BP-test.
-For this purpose, two experiments were conducted.
+
+# Experimental design
+
+For the purpose of examining the effectiveness of visual test in regression diagnostics, two experiments were conducted.
 The first experiment has ideal scenario for conventional testing, where we would not expect the visual test to do better than the conventional test.
@@
 
-For each parameter combination, effect $E$ is derived from the Kullback-Leibler divergence (the derivation is provided in [appendix ref here]) formulated as
+For each parameter combination, effect $E$ is derived from the Kullback-Leibler divergence (the derivation is provided in [appendix ref here]) formulated as:
 
@@
 
-The logistic regression is fit using $log_e(E)$ as the only fixed effect covariate for the power of visual test formulated as
+The logistic regression is fit using $log_e(E)$ as the only fixed effect covariate for the power of visual test formulated as:
 
@@
 
-Table [table ref here] shows the parameter estimates of the logistic regression. [Discussion about the numeric estimates here]
+The same model is fit on different subsets of the collated data to study the effect of different factors have on the power of the visual test.
+
+Table [table ref here] shows the parameter estimates of the logistic regressions. [Discussion about the numeric estimates here]
 
@@
 
-Figure [figure ref here] shows an overview of estimated power against natural logarithm
+Figure \ref{fig:power-overview} shows an overview of estimated power of visual test against natural logarithm of the effect. Without further concenring different component contributing to the effect,
 
@@
   labs(col = "Test") +
-  theme_light() +
+  theme_light(base_size = 5) +
   xlab(quote(log[e](E))) +
@@
 
-conv_test_plot()
-```
-
-```{r}
 add_wrong_BP_conv_test_plot_layers <- function(p) {
@@
 
-conv_test_plot() %>%
-  add_wrong_BP_conv_test_plot_layers()
-```
-
-```{r}
 add_wrong_WS_conv_test_plot_layers <- function(p) {
@@
 }
-
-conv_test_plot() %>%
-  add_wrong_BP_conv_test_plot_layers() %>%
-  add_wrong_WS_conv_test_plot_layers()
 ```
@@
   group_by(lineup_id) %>%
-  summarise(across(c(type:p_value), first)) %>%
+  summarise(across(c(type:p_value, prop_detect), first)) %>%
   filter(lineup_id < 577) %>%
@@
               aes(log(effect_size), reject, group = boot_id, col = "Visual"), 
-              method = "glm", method.args = list(family = binomial), se = FALSE, alpha = 0.05)
+              method = "glm", method.args = list(family = binomial), se = FALSE, alpha = 0.05) +
+  geom_point(data = filter(visual_dat, boot_id == 1), aes(log(effect_size), prop_detect), alpha = 0.1)
 
@@
 }
-
+```
+
+```{r eval = FALSE}
+polynomials %>%
+  group_by(lineup_id) %>%
+  summarise(dd = first(abs(effect_size - exp(2)))) %>%
+  arrange(dd)
+```
+
+```{r power-overview, fig.height=7.5, fig.cap="Power overview. \\label{fig:power-overview}"}
 conv_test_plot() %>%
@@
   add_wrong_WS_conv_test_plot_layers() %>%
-  add_visual_test_plot_layers()
+  add_visual_test_plot_layers() +
+  geom_vline(xintercept = 0) + 
+  geom_vline(xintercept = 2) +
+  geom_vline(xintercept = 4) +
+  geom_vline(xintercept = 6) -> base_plot
+
+VI_MODEL$plot_lineup(polynomials_lineup$lineup_314$data,
+                     size = 0.1,
+                     stroke = 0.2,
+                     remove_grid_line = TRUE,
+                     theme = theme_light(base_size = 5),
+                     remove_axis = TRUE) +
+  ggtitle(bquote(log[e](E) == 0~", F-test p-value" == .(format(filter(polynomials, lineup_id == 314)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 314)$p_value[1], digits = 3)))) -> bottom_plot_1
+
+VI_MODEL$plot_lineup(polynomials_lineup$lineup_415$data,
+                     size = 0.1,
+                     stroke = 0.2,
+                     remove_grid_line = TRUE,
+                     theme = theme_light(base_size = 5),
+                     remove_axis = TRUE) +
+  ggtitle(bquote(log[e](E) == 2~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 415)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 415)$p_value[1], digits = 3)))) -> bottom_plot_2
+
+VI_MODEL$plot_lineup(polynomials_lineup$lineup_494$data,
+                     size = 0.1,
+                     stroke = 0.2,
+                     remove_grid_line = TRUE,
+                     theme = theme_light(base_size = 5),
+                     remove_axis = TRUE) +
+  ggtitle(bquote(log[e](E) == 4~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 494)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 494)$p_value[1], digits = 3)))) -> bottom_plot_3
+
+VI_MODEL$plot_lineup(polynomials_lineup$lineup_101$data,
+                     size = 0.1,
+                     stroke = 0.2,
+                     remove_grid_line = TRUE,
+                     theme = theme_light(base_size = 5),
+                     remove_axis = TRUE) +
+  ggtitle(bquote(log[e](E) == 6~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 101)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 101)$p_value[1], digits = 3)))) -> bottom_plot_4
+
+design <- c(patchwork::area(1, 1, 1, 8), 
+            patchwork::area(2, 1, 2, 4), 
+            patchwork::area(2, 5, 2, 8), 
+            patchwork::area(3, 1, 3, 4), 
+            patchwork::area(3, 5, 3, 8))
+
+patchwork::wrap_plots(base_plot, B = bottom_plot_1, C = bottom_plot_2, D = bottom_plot_3, E = bottom_plot_4, design = design)
 ```

### 2022-07-30 paper_comparison/paper_comparison/paper_comparison.tex
modify power plot

--- 2022-07-28 paper_comparison/paper_comparison/paper_comparison.tex
@@
 
-\hypertarget{experimental-design}{%
-\section{Experimental design}\label{experimental-design}}
+\hypertarget{effectiveness-of-visual-test-in-regression-diagnostics}{%
+\subsection{Effectiveness of visual test in regression
+diagnostics}\label{effectiveness-of-visual-test-in-regression-diagnostics}}
 
@@
 test in regression diagnostics. Particularly, its ability to detect
-non-linearity and heteroskedasticity compared to F-test and BP-test. For
-this purpose, two experiments were conducted. The first experiment has
-ideal scenario for conventional testing, where we would not expect the
-visual test to do better than the conventional test. The second
-experiment is a scenario where the conventional test is an approximate
-test, which visual test may have a chance to match the performance of
-the conventional test.
+non-linearity and heteroskedasticity compared to F-test and BP-test.
+
+\hypertarget{experimental-design}{%
+\section{Experimental design}\label{experimental-design}}
+
+For the purpose of examining the effectiveness of visual test in
+regression diagnostics, two experiments were conducted. The first
+experiment has ideal scenario for conventional testing, where we would
+not expect the visual test to do better than the conventional test. The
+second experiment is a scenario where the conventional test is an
+approximate test, which visual test may have a chance to match the
+performance of the conventional test.
 
@@
 Kullback-Leibler divergence (the derivation is provided in {[}appendix
-ref here{]}) formulated as
+ref here{]}) formulated as:
 
@@
 The logistic regression is fit using \(log_e(E)\) as the only fixed
-effect covariate for the power of visual test formulated as
+effect covariate for the power of visual test formulated as:
 
@@
 
+The same model is fit on different subsets of the collated data to study
+the effect of different factors have on the power of the visual test.
+
 Table {[}table ref here{]} shows the parameter estimates of the logistic
-regression. {[}Discussion about the numeric estimates here{]}
+regressions. {[}Discussion about the numeric estimates here{]}
 
@@
 
-Figure {[}figure ref here{]} shows an overview of estimated power
-against natural logarithm
+Figure \ref{fig:power-overview} shows an overview of estimated power of
+visual test against natural logarithm of the effect. Without further
+concenring different component contributing to the effect,
+
+\begin{figure}
+\centering
+\includegraphics{paper_comparison_files/figure-latex/power-overview-1.pdf}
+\caption{Power overview. \label{fig:power-overview}}
+\end{figure}
+
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-9-1.pdf}
 
@@
 
-\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-4-1.pdf}
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-10-1.pdf}
 

### 2022-07-31 paper_comparison/paper_comparison/paper_comparison.Rmd
update example plot for power overview

@@
   group_by(lineup_id) %>%
-  summarise(dd = first(abs(effect_size - exp(2)))) %>%
+  summarise(dd = first(abs(log(effect_size) - 5)), p_value = first(p_value)) %>%
   arrange(dd)
+
+polynomials %>%
+  ggplot() +
+  geom_point(aes(prop_detect, p_value < 0.06))
+
+bottom_plot_2
 ```
@@
   add_visual_test_plot_layers() +
-  geom_vline(xintercept = 0) + 
-  geom_vline(xintercept = 2) +
-  geom_vline(xintercept = 4) +
-  geom_vline(xintercept = 6) -> base_plot
-
-VI_MODEL$plot_lineup(polynomials_lineup$lineup_314$data,
-                     size = 0.1,
-                     stroke = 0.2,
-                     remove_grid_line = TRUE,
-                     theme = theme_light(base_size = 5),
-                     remove_axis = TRUE) +
-  ggtitle(bquote(log[e](E) == 0~", F-test p-value" == .(format(filter(polynomials, lineup_id == 314)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 314)$p_value[1], digits = 3)))) -> bottom_plot_1
+  geom_vline(xintercept = 2, linetype = 2) + 
+  geom_vline(xintercept = 3, linetype = 2) +
+  geom_vline(xintercept = 4, linetype = 2) +
+  geom_vline(xintercept = 5, linetype = 2) -> base_plot
+
+# VI_MODEL$plot_lineup(polynomials_lineup$lineup_314$data,
+#                      size = 0.1,
+#                      stroke = 0.2,
+#                      remove_grid_line = TRUE,
+#                      theme = theme_light(base_size = 5),
+#                      remove_axis = TRUE) +
+#   ggtitle(bquote(log[e](E) == 0~", F-test p-value" == .(format(filter(polynomials, lineup_id == 314)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 314)$p_value[1], digits = 3)))) -> bottom_plot_1
 
@@
                      remove_axis = TRUE) +
-  ggtitle(bquote(log[e](E) == 2~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 415)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 415)$p_value[1], digits = 3)))) -> bottom_plot_2
-
-VI_MODEL$plot_lineup(polynomials_lineup$lineup_494$data,
+  ggtitle(bquote(log[e](E) == 2~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 415)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 415)$p_value[1], digits = 3)))) -> bottom_plot_1
+
+VI_MODEL$plot_lineup(polynomials_lineup$lineup_364$data,
                      size = 0.1,
@@
                      remove_axis = TRUE) +
-  ggtitle(bquote(log[e](E) == 4~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 494)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 494)$p_value[1], digits = 3)))) -> bottom_plot_3
-
-VI_MODEL$plot_lineup(polynomials_lineup$lineup_101$data,
+  ggtitle(bquote(log[e](E) == 3~", F-test p-value" == .(format(filter(polynomials, lineup_id == 364)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 364)$p_value[1], digits = 3)))) -> bottom_plot_2
+
+VI_MODEL$plot_lineup(polynomials_lineup$lineup_574$data,
                      size = 0.1,
@@
                      remove_axis = TRUE) +
-  ggtitle(bquote(log[e](E) == 6~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 101)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 101)$p_value[1], digits = 3)))) -> bottom_plot_4
+  ggtitle(bquote(log[e](E) == 4~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 574)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 574)$p_value[1], digits = 3)))) -> bottom_plot_3
+
+VI_MODEL$plot_lineup(polynomials_lineup$lineup_179$data,
+                     size = 0.1,
+                     stroke = 0.2,
+                     remove_grid_line = TRUE,
+                     theme = theme_light(base_size = 5),
+                     remove_axis = TRUE) +
+  ggtitle(bquote(log[e](E) == 5~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 179)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 179)$p_value[1], digits = 3)))) -> bottom_plot_4
 

### 2022-08-01 paper_comparison/paper_comparison/paper_comparison.Rmd
add power com text

@@
 ```{r}
-if (!file.exists(here::here("data/third_study/wrong_WS_conv_result.rds"))) {
-  wrong_WS_conv_result <- list()
+if (!file.exists(here::here("data/third_study/wrong_SW_conv_result.rds"))) {
+  wrong_SW_conv_result <- list()
   for (i in 1:100)
   {
-    wrong_WS_conv_result[[i]] <- map(1:2000,
+    wrong_SW_conv_result[[i]] <- map(1:2000,
                             function(x) {
@@
 
-  wrong_WS_conv_result <- wrong_WS_conv_result %>%
+  wrong_SW_conv_result <- wrong_SW_conv_result %>%
     reduce(bind_rows)
 
-  saveRDS(wrong_WS_conv_result, here::here("data/third_study/wrong_WS_conv_result.rds"))
+  saveRDS(wrong_SW_conv_result, here::here("data/third_study/wrong_SW_conv_result.rds"))
 } else {
-  wrong_WS_conv_result <- readRDS(here::here("data/third_study/wrong_WS_conv_result.rds"))
+  wrong_SW_conv_result <- readRDS(here::here("data/third_study/wrong_SW_conv_result.rds"))
 }
@@
 
-The same model is fit on different subsets of the collated data to study the effect of different factors have on the power of the visual test.
+To study different factors contributing to the power of the visual test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors.
+This includes [expansion].
 
@@
 
-Figure \ref{fig:power-overview} shows an overview of estimated power of visual test against natural logarithm of the effect. Without further concenring different component contributing to the effect,
+Figure \ref{fig:power-overview} shows an overview of estimated power of visual test against natural logarithm of the effect with comparison to the power of an exact test - F-test, and the power of two other residual-based conventional tests commonly used in regression diagnostics but for testing other departures from the model assumptions.
+In overall, the power of all four tests increases as the effect becomes larger.
+The power curve of F-test climbs aggressively from 25\% to around 90\% as $log_e(E)$ increases from 0 to 2, while others respond inactively to the change of effect and remain lower than 25\% throughout the period, showing that as an exact test, the F-test is relatively more sensitive to the type of model defects that being considered.
+The power of visual test arises steadily and nearly linearly to around 90\% as $log_e(E)$ increases from 2 to 5, suggesting that the effect starts to make noticeable impact on the degree of the presence of the designed visual features.
+Other two inappropriate conventional tests shows improvement at the same time but at a lower rate.
+This coincides the point made by @cook1982residuals mentioned in \ref{hypothesis-testing} that residual-based tests for a specific type of model defect are sensitive to other types of model defects.
+At $log_e(E) = 6$, the power curve of F-test reaches almost 100% followed by the visual test by a small margin.
+The power of Breusch–Pagan test and Shapiro–Wilk test reach around 75% and 63% respectively.
+
+What truly impress us is the huge difference between the estimated power of visual test and the estimated power of F-test.
+The margin is largest at around $log_e(E) = 2$.
+An example lineup is included in Figure \ref{fig:power-overview} where none of subjects detect the actual data plot positioned at panel 14.
+It demonstrates that at this level of difficulty, the designed visual feature is rarely visible, making the actual data plot indistinguishable from residual plots simulated from the assumed model.
+From a communication perspective, given the fact that the visual difference is unperceivable, the argument that non-linearity present in the fitted model is less convincing to the public even though it is true.
+At around $log_e(E) = 3$, the margin gets smaller as the chance of identifying the actual data plot becomes larger.
+At this level of difficulty, the designed visual features are usually detectable but it may not stand out from the lineup as other null plots may happen to include outliers or visual patterns that are are considered to be more attractive by human, and thus recognized as the most different plot.
+Without knowing the designed visual features beforehand, it is actually hard to identify the actual data plot by pure image comparison.
+The corresponding example lineup for $log_e(E) = 3$ shown in Figure \ref{fig:power-overview} has the actual data plot positioned at panel 20, where two out of five subjects detect it.
+It can be observed that a M-shape is presented in plot 20, but the signal is not strong enough to attract all five subjects, resulting in a visual p-value sightly above the desired significance level $\alpha = 0.05$.
+At $log_e(E) = 4$ and  $log_e(E) = 6$, the designed visual features become much clear and attractive, leading to a high percentage of rejection of the null hypothesis.
+Figure \ref{fig:power-overview} give example lineups of such cases.
+
+It is also interesting to observe that the power of visual test behaves similarly to those inappropriate conventional tests until the designed visual features really becomes visible
+
+As of shown in Figure \ref{fig:power-overview}, the example lineup for $log_e(E) = 3$
+
+it is of low chance detecting the actual data plot
+
+increase steadily when $2<log_e(E)<4$. the power of visual test almost reaches 100% at $log_e(E) = 6$.
 
@@
 
-add_wrong_WS_conv_test_plot_layers <- function(p) {
+add_wrong_SW_conv_test_plot_layers <- function(p) {
   p +
-  stat_smooth(data = wrong_WS_conv_result, geom = "line", 
-              aes(log(effect_size), reject, group = boot_id, col = "WS-test"), 
+  stat_smooth(data = wrong_SW_conv_result, geom = "line", 
+              aes(log(effect_size), reject, group = boot_id, col = "SW-test"), 
               se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) +
-  stat_smooth(data = wrong_WS_conv_result, geom = "line", 
-              aes(log(effect_size), reject, col = "WS-test"), 
+  stat_smooth(data = wrong_SW_conv_result, geom = "line", 
+              aes(log(effect_size), reject, col = "SW-test"), 
               se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1)
@@
 bottom_plot_2
+
+polynomials %>%
+  filter(lineup_id == 364)
 ```
@@
   add_wrong_BP_conv_test_plot_layers() %>%
-  add_wrong_WS_conv_test_plot_layers() %>%
+  add_wrong_SW_conv_test_plot_layers() %>%
   add_visual_test_plot_layers() +
@@
                      remove_axis = TRUE) +
-  ggtitle(bquote(log[e](E) == 2~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 415)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 415)$p_value[1], digits = 3)))) -> bottom_plot_1
+  ggtitle(bquote(log[e](E) %~~% 2~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 415)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 415)$p_value[1], digits = 3)))) -> bottom_plot_1
 
@@
                      remove_axis = TRUE) +
-  ggtitle(bquote(log[e](E) == 3~", F-test p-value" == .(format(filter(polynomials, lineup_id == 364)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 364)$p_value[1], digits = 3)))) -> bottom_plot_2
+  ggtitle(bquote(log[e](E) %~~% 3~", F-test p-value" == .(format(filter(polynomials, lineup_id == 364)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 364)$p_value[1], digits = 3)))) -> bottom_plot_2
 
@@
                      remove_axis = TRUE) +
-  ggtitle(bquote(log[e](E) == 4~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 574)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 574)$p_value[1], digits = 3)))) -> bottom_plot_3
+  ggtitle(bquote(log[e](E) %~~% 4~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 574)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 574)$p_value[1], digits = 3)))) -> bottom_plot_3
 
@@
                      remove_axis = TRUE) +
-  ggtitle(bquote(log[e](E) == 5~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 179)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 179)$p_value[1], digits = 3)))) -> bottom_plot_4
+  ggtitle(bquote(log[e](E) %~~% 5~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 179)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 179)$p_value[1], digits = 3)))) -> bottom_plot_4
 
@@
 patchwork::wrap_plots(base_plot, B = bottom_plot_1, C = bottom_plot_2, D = bottom_plot_3, E = bottom_plot_4, design = design)
-```
-
-```{r}
-polynomials %>%
-  group_by(lineup_id) %>%
-  summarise(across(c(type:p_value), first)) %>%
-  filter(lineup_id < 577) %>%
-  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
-  ggplot() +
-  geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  xlab("Natural logarithm of effect size") +
-  ylab("Power")
 ```

### 2022-08-01 paper_comparison/paper_comparison/paper_comparison.tex
add power com text

@@
 
-The same model is fit on different subsets of the collated data to study
-the effect of different factors have on the power of the visual test.
+To study different factors contributing to the power of the visual test,
+the same logistic regression model is fit on different subsets of the
+collated data grouped by levels of factors. This includes
+{[}expansion{]}.
 
@@
 Figure \ref{fig:power-overview} shows an overview of estimated power of
-visual test against natural logarithm of the effect. Without further
-concenring different component contributing to the effect,
+visual test against natural logarithm of the effect with comparison to
+the power of an exact test - F-test, and the power of two other
+residual-based conventional tests commonly used in regression
+diagnostics but for testing other departures from the model assumptions.
+In overall, the power of all four tests increases as the effect becomes
+larger. The power curve of F-test climbs aggressively from 25\% to
+around 90\% as \(log_e(E)\) increases from 0 to 2, while others respond
+inactively to the change of effect and remain lower than 25\% throughout
+the period, showing that as an exact test, the F-test is relatively more
+sensitive to the type of model defects that being considered. The power
+of visual test arises steadily and nearly linearly to around 90\% as
+\(log_e(E)\) increases from 2 to 5, suggesting that the effect starts to
+make noticeable impact on the degree of the presence of the designed
+visual features. Other two inappropriate conventional tests shows
+improvement at the same time but at a lower rate. This coincides the
+point made by \citet{cook1982residuals} mentioned in
+\ref{hypothesis-testing} that residual-based tests for a specific type
+of model defect are sensitive to other types of model defects. At
+\(log_e(E) = 6\), the power curve of F-test reaches almost 100\%
+followed by the visual test by a small margin. The power of
+Breusch--Pagan test and Shapiro--Wilk test reach around 75\% and 63\%
+respectively.
+
+What truly impress us is the huge difference between the estimated power
+of visual test and the estimated power of F-test. The margin is largest
+at around \(log_e(E) = 2\). An example lineup is included in Figure
+\ref{fig:power-overview} where none of subjects detect the actual data
+plot positioned at panel 14. It demonstrates that at this level of
+difficulty, the designed visual feature is rarely visible, making the
+actual data plot indistinguishable from residual plots simulated from
+the assumed model. From a communication perspective, given the fact that
+the visual difference is unperceivable, the argument that non-linearity
+present in the fitted model is less convincing to the public even though
+it is true. At around \(log_e(E) = 3\), the margin gets smaller as the
+chance of identifying the actual data plot becomes larger. At this level
+of difficulty, the designed visual features are usually detectable but
+it may not stand out from the lineup as other null plots may happen to
+include outliers or visual patterns that are are considered to be more
+attractive by human, and thus recognized as the most different plot.
+Without knowing the designed visual features beforehand, it is actually
+hard to identify the actual data plot by pure image comparison. The
+corresponding example lineup for \(log_e(E) = 3\) shown in Figure
+\ref{fig:power-overview} has the actual data plot positioned at panel
+20, where two out of five subjects detect it. It can be observed that a
+M-shape is presented in plot 20, but the signal is not strong enough to
+attract all five subjects, resulting in a visual p-value sightly above
+the desired significance level \(\alpha = 0.05\). At \(log_e(E) = 4\)
+and \(log_e(E) = 6\), the designed visual features become much clear and
+attractive, leading to a high percentage of rejection of the null
+hypothesis. Figure \ref{fig:power-overview} give example lineups of such
+cases.
+
+It is also interesting to observe that the power of visual test behaves
+similarly to those inappropriate conventional tests until the designed
+visual features really becomes visible
+
+As of shown in Figure \ref{fig:power-overview}, the example lineup for
+\(log_e(E) = 3\)
+
+it is of low chance detecting the actual data plot
+
+increase steadily when \(2<log_e(E)<4\). the power of visual test almost
+reaches 100\% at \(log_e(E) = 6\).
 
@@
 \end{figure}
-
-\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-9-1.pdf}
 
@@
 
-\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-10-1.pdf}
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-9-1.pdf}
 

### 2022-08-03 paper_comparison/paper_comparison/paper_comparison.Rmd
update graphs

@@
 
-add_visual_test_plot_layers <- function(p) {
+add_visual_test_plot_layers <- function(p, points = c()) {
   visual_dat <- polynomials %>%
@@
               method = "glm", method.args = list(family = binomial), se = FALSE, alpha = 0.05) +
-  geom_point(data = filter(visual_dat, boot_id == 1), aes(log(effect_size), prop_detect), alpha = 0.1)
-
+  geom_point(data = filter(visual_dat, boot_id == 1), aes(log(effect_size), prop_detect), alpha = 0.1) + 
+  geom_point(data = filter(visual_dat, boot_id == 1, lineup_id %in% points), aes(log(effect_size), prop_detect), alpha = 1, col = "red", size = 2)
 }
@@
 ```{r eval = FALSE}
+filter(polynomials, set == 123)
+
 polynomials %>%
   group_by(lineup_id) %>%
-  summarise(dd = first(abs(log(effect_size) - 5)), p_value = first(p_value)) %>%
-  arrange(dd)
+  summarise(dd = first(abs(log(effect_size) - 5)), p_value = first(p_value), prop_detect = first(prop_detect)) %>%
+  arrange(dd) %>%
+  head(30) %>%
+  summarise(tt = median(prop_detect))
 
@@
 ```{r power-overview, fig.height=7.5, fig.cap="Power overview. \\label{fig:power-overview}"}
+conv_test_plot() %>%
+  add_wrong_BP_conv_test_plot_layers() %>%
+  add_wrong_SW_conv_test_plot_layers() %>%
+  add_visual_test_plot_layers(points = c(415, 131, 322, 378)) +
+  geom_vline(xintercept = 2, linetype = 2) + 
+  geom_vline(xintercept = 3, linetype = 2) +
+  geom_vline(xintercept = 4, linetype = 2) +
+  geom_vline(xintercept = 5, linetype = 2) -> base_plot
+
+# VI_MODEL$plot_lineup(polynomials_lineup$lineup_314$data,
+#                      size = 0.1,
+#                      stroke = 0.2,
+#                      remove_grid_line = TRUE,
+#                      theme = theme_light(base_size = 5),
+#                      remove_axis = TRUE) +
+#   ggtitle(bquote(log[e](E) == 0~", F-test p-value" == .(format(filter(polynomials, lineup_id == 314)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 314)$p_value[1], digits = 3)))) -> bottom_plot_1
+
+VI_MODEL$plot_lineup(polynomials_lineup$lineup_415$data,
+                     size = 0.1,
+                     stroke = 0.2,
+                     remove_grid_line = TRUE,
+                     theme = theme_light(base_size = 5),
+                     remove_axis = TRUE) +
+  ggtitle(bquote(log[e](E) %~~% 2~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 415)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 415)$p_value[1], digits = 3)))) -> bottom_plot_1
+
+VI_MODEL$plot_lineup(polynomials_lineup$lineup_131$data,
+                     size = 0.1,
+                     stroke = 0.2,
+                     remove_grid_line = TRUE,
+                     theme = theme_light(base_size = 5),
+                     remove_axis = TRUE) +
+  ggtitle(bquote(log[e](E) %~~% 3~", F-test p-value" == .(format(filter(polynomials, lineup_id == 131)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 131)$p_value[1], digits = 3)))) -> bottom_plot_2
+
+VI_MODEL$plot_lineup(polynomials_lineup$lineup_322$data,
+                     size = 0.1,
+                     stroke = 0.2,
+                     remove_grid_line = TRUE,
+                     theme = theme_light(base_size = 5),
+                     remove_axis = TRUE) +
+  ggtitle(bquote(log[e](E) %~~% 4~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 322)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 322)$p_value[1], digits = 3)))) -> bottom_plot_3
+
+VI_MODEL$plot_lineup(polynomials_lineup$lineup_378$data,
+                     size = 0.1,
+                     stroke = 0.2,
+                     remove_grid_line = TRUE,
+                     theme = theme_light(base_size = 5),
+                     remove_axis = TRUE) +
+  ggtitle(bquote(log[e](E) %~~% 5~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 378)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 378)$p_value[1], digits = 3)))) -> bottom_plot_4
+
+design <- c(patchwork::area(1, 1, 1, 8), 
+            patchwork::area(2, 1, 2, 4), 
+            patchwork::area(2, 5, 2, 8), 
+            patchwork::area(3, 1, 3, 4), 
+            patchwork::area(3, 5, 3, 8))
+
+patchwork::wrap_plots(base_plot, B = bottom_plot_1, C = bottom_plot_2, D = bottom_plot_3, E = bottom_plot_4, design = design)
+```
+
+```{r}
 conv_test_plot() %>%
@@
   geom_vline(xintercept = 4, linetype = 2) +
-  geom_vline(xintercept = 5, linetype = 2) -> base_plot
-
-# VI_MODEL$plot_lineup(polynomials_lineup$lineup_314$data,
-#                      size = 0.1,
-#                      stroke = 0.2,
-#                      remove_grid_line = TRUE,
-#                      theme = theme_light(base_size = 5),
-#                      remove_axis = TRUE) +
-#   ggtitle(bquote(log[e](E) == 0~", F-test p-value" == .(format(filter(polynomials, lineup_id == 314)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 314)$p_value[1], digits = 3)))) -> bottom_plot_1
-
-VI_MODEL$plot_lineup(polynomials_lineup$lineup_415$data,
-                     size = 0.1,
-                     stroke = 0.2,
-                     remove_grid_line = TRUE,
-                     theme = theme_light(base_size = 5),
-                     remove_axis = TRUE) +
-  ggtitle(bquote(log[e](E) %~~% 2~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 415)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 415)$p_value[1], digits = 3)))) -> bottom_plot_1
-
-VI_MODEL$plot_lineup(polynomials_lineup$lineup_364$data,
-                     size = 0.1,
-                     stroke = 0.2,
-                     remove_grid_line = TRUE,
-                     theme = theme_light(base_size = 5),
-                     remove_axis = TRUE) +
-  ggtitle(bquote(log[e](E) %~~% 3~", F-test p-value" == .(format(filter(polynomials, lineup_id == 364)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 364)$p_value[1], digits = 3)))) -> bottom_plot_2
-
-VI_MODEL$plot_lineup(polynomials_lineup$lineup_574$data,
-                     size = 0.1,
-                     stroke = 0.2,
-                     remove_grid_line = TRUE,
-                     theme = theme_light(base_size = 5),
-                     remove_axis = TRUE) +
-  ggtitle(bquote(log[e](E) %~~% 4~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 574)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 574)$p_value[1], digits = 3)))) -> bottom_plot_3
-
-VI_MODEL$plot_lineup(polynomials_lineup$lineup_179$data,
-                     size = 0.1,
-                     stroke = 0.2,
-                     remove_grid_line = TRUE,
-                     theme = theme_light(base_size = 5),
-                     remove_axis = TRUE) +
-  ggtitle(bquote(log[e](E) %~~% 5~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 179)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 179)$p_value[1], digits = 3)))) -> bottom_plot_4
-
-design <- c(patchwork::area(1, 1, 1, 8), 
-            patchwork::area(2, 1, 2, 4), 
-            patchwork::area(2, 5, 2, 8), 
-            patchwork::area(3, 1, 3, 4), 
-            patchwork::area(3, 5, 3, 8))
-
-patchwork::wrap_plots(base_plot, B = bottom_plot_1, C = bottom_plot_2, D = bottom_plot_3, E = bottom_plot_4, design = design)
+  geom_vline(xintercept = 5, linetype = 2) +
+  facet_wrap(~x_dist)
+```
+
+```{r}
+conv_test_plot() %>%
+  add_wrong_BP_conv_test_plot_layers() %>%
+  add_wrong_SW_conv_test_plot_layers() %>%
+  add_visual_test_plot_layers() +
+  geom_vline(xintercept = 2, linetype = 2) + 
+  geom_vline(xintercept = 3, linetype = 2) +
+  geom_vline(xintercept = 4, linetype = 2) +
+  geom_vline(xintercept = 5, linetype = 2) +
+  facet_wrap(~shape)
 ```

### 2022-08-10 paper_comparison/paper_comparison/paper_comparison.Rmd
latest update

@@
 ```{r}
+stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
+
+if (!file.exists(here::here("data/third_study/conv_result_heter.rds"))) {
+  conv_result_heter <- list()
+  for (i in 1:100)
+  {
+    conv_result_heter[[i]] <- map(1:2000,
+                            function(x) {
+                     a <- sample(c(-1, 0, 1), 1)
+                     b <- runif(1, 0, 32)
+                     x_dist <- sample(c("uniform", "normal", "lognormal", "even_discrete"), 1)
+                     x <- switch(x_dist,
+                                 uniform = rand_uniform(-1, 1),
+                                 normal = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))},
+                                 lognormal = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))},
+                                 even_discrete = rand_uniform_d(k = 5, even = TRUE))
+                     mod <- heter_model(a = a, b = b, x = x)
+                     n <- sample(c(50, 100, 300), 1)
+                     tmp_dat <- mod$gen(n)
+
+                     tibble(a = a,
+                            b = b,
+                            x_dist = x_dist,
+                            n = n,
+                            effect_size = mod$effect_size(tmp_dat),
+                            p_value = mod$test(tmp_dat)$p_value,
+                            reject = p_value < 0.05,
+                            boot_id = i)
+                   }) %>%
+  reduce(bind_rows) %>%
+  mutate(reject = as.numeric(reject))
+  }
+
+  conv_result_heter <- conv_result_heter %>%
+    reduce(bind_rows)
+
+  saveRDS(conv_result_heter, here::here("data/third_study/conv_result_heter.rds"))
+} else {
+  conv_result_heter <- readRDS(here::here("data/third_study/conv_result_heter.rds"))
+}
+
+```
+
+```{r}
+stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
+
+if (!file.exists(here::here("data/third_study/conv_result_heter_SW.rds"))) {
+  conv_result_heter_SW <- list()
+  for (i in 1:100)
+  {
+    conv_result_heter_SW[[i]] <- map(1:2000,
+                            function(x) {
+                     a <- sample(c(-1, 0, 1), 1)
+                     b <- runif(1, 0, 32)
+                     x_dist <- sample(c("uniform", "normal", "lognormal", "even_discrete"), 1)
+                     x <- switch(x_dist,
+                                 uniform = rand_uniform(-1, 1),
+                                 normal = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))},
+                                 lognormal = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))},
+                                 even_discrete = rand_uniform_d(k = 5, even = TRUE))
+                     mod <- heter_model(a = a, b = b, x = x)
+                     n <- sample(c(50, 100, 300), 1)
+                     tmp_dat <- mod$gen(n)
+
+                     tibble(a = a,
+                            b = b,
+                            x_dist = x_dist,
+                            n = n,
+                            effect_size = mod$effect_size(tmp_dat),
+                            p_value = shapiro.test(tmp_dat$.resid)$p.value,
+                            reject = p_value < 0.05,
+                            boot_id = i)
+                   }) %>%
+  reduce(bind_rows) %>%
+  mutate(reject = as.numeric(reject))
+  }
+
+  conv_result_heter_SW <- conv_result_heter_SW %>%
+    reduce(bind_rows)
+
+  saveRDS(conv_result_heter_SW, here::here("data/third_study/conv_result_heter_SW.rds"))
+} else {
+  conv_result_heter_SW <- readRDS(here::here("data/third_study/conv_result_heter_SW.rds"))
+}
+
+```
+
+```{r}
 if (!file.exists(here::here("data/third_study/wrong_BP_conv_result.rds"))) {
@@
 The effectiveness of visual inference has already been validated by @majumder_validation_2013 under relatively simple classical normal linear regression model settings with only one or two regressors.
-Their results suggests visual test is capable of testing the significance of a single regressor with a similar power as a t-test, though they expressed that in general it is unnecessary to use visual inference if there exists a conventional test and they didn't expect the visual test to perform equally well as the conventional test.
+Their results suggest visual test is capable of testing the significance of a single regressor with a similar power as a t-test, though they expressed that in general it is unnecessary to use visual inference if there exists a conventional test and they didn't expect the visual test to perform equally well as the conventional test.
 In their third experiment, where there does not exist a proper conventional test, visual test outperforms the conventional test for a large margin.
@@
 For the purpose of examining the effectiveness of visual test in regression diagnostics, two experiments were conducted.
-The first experiment has ideal scenario for conventional testing, where we would not expect the visual test to do better than the conventional test.
-The second experiment is a scenario where the conventional test is an approximate test, which visual test may have a chance to match the performance of the conventional test.
-
-As suggested by many previous studies [many studies refs here], it is reasonable for an subject to evaluate a block of 20 lineups.
-Subjects for both two experiments were recruited from Prolific web service [ref here].
-Data screening procedure was applied during the recruitment, subjects were required to be fluent in English, with $98\%$ minimum approval rate and 10 minimum submissions.
-During the experiment, they were asked to select one or more plots that are mostly different from others, provide a reason for their selections, and express how different they think the selected plots are from others.
-If there was no noticeable difference between plots in a lineup, subject could choose to select zero plots without providing the reason.
+The experiment I has ideal scenario for conventional testing, where the visual test is not expected to outperform the conventional test.
+The experiment II is a scenario where the conventional test is an approximate test, in which the visual test may have a chance to match the performance of the conventional test.
+
+Subjects for both experiments were recruited from an crowdsourcing platform called Prolific [ref here].
+Prescreening procedure was applied during the recruitment, subjects were required to be fluent in English, with $98\%$ minimum approval rate in other studies and 10 minimum submissions.
+
+During the experiment, every subject was presented with a block of 20 lineups.
+And for every lineup, the subject was asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots were from others.
+If there was no noticeable difference between plots in a lineup, subjects were permitted to select zero plots without providing the reason.
+No subject was shown the same lineup twice.
 Information about preferred pronoun, age group, education, and previous experience in visual experiment were also collected.
-No subject was shown the same lineup twice.
-
-For each of the experiments, a pool of 12 lineups with obvious visual patterns were generated.
-In every block of 20 lineups that presented to subjects, two different lineups sampled from the corresponding pool of lineups were included as attention checks.
-Every subject should get at least one attention check correct such that the quality of the subject effort could be ensured.
-
-## Non-linear covariates
-
-The first experiment is designed to study the ability of human subjects to detect the effect of a continuous variable $\boldsymbol{Z}$ which is a probabilist's Hermite polynomial [Herimite ref here] of another variable $\boldsymbol{X}$ in a two variable statistical model formulated as:
+
+A pool of 12 different lineups with obvious visual patterns were generated for experiment I and experiment II respectively.
+In every block of 20 lineups that presented to a subject, two out of 12 lineups were included as attention checks.
+A subject's submission was only accepted if the actual data plot was identified for at least one attention check.
+Data of rejected submissions were discarded automatically to maintain the overall data quality.
+
+## Non-linearity
+
+Experiment I is designed to study the ability of human subjects to detect the effect of a random vector $\boldsymbol{z}$ which is a probabilist's Hermite polynomial [Herimite ref here] of another random vector $\boldsymbol{x}$ in a two variable statistical model formulated as:
 
 \begin{align} \label{eq:nonlinearity-model}
-\boldsymbol{Y} = 1 + \boldsymbol{X} + \boldsymbol{Z} + \boldsymbol{\varepsilon},\\
-\boldsymbol{X} = g(\boldsymbol{X}_{raw}, 1), \\
-\boldsymbol{Z} = g(\boldsymbol{Z}_{raw}, 1), \\
-\boldsymbol{Z}_{raw} = He_j(g(\boldsymbol{X}, 2)),
+\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
+\boldsymbol{x} = g(\boldsymbol{x}_{raw}, 1), \\
+\boldsymbol{z} = g(\boldsymbol{z}_{raw}, 1), \\
+\boldsymbol{z}_{raw} = He_j(g(\boldsymbol{z}, 2)),
 \end{align}
 
-where $\boldsymbol{Y}$, $\boldsymbol{X}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{X}_{raw}$, $\boldsymbol{Z}_{raw}$ are $n\times1$ matrices,  $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{X}, k)$ is a scaling function to enforce the support of the variable to be $\{-k, k\}$ defined as
+where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vector of size $n$,  $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $\{-k, k\}$ defined as
 
 \begin{equation} \label{eq:scaling-function}
-g(\boldsymbol{X}, k) = (\boldsymbol{X} - min(\boldsymbol{X}))/max(\boldsymbol{X} - min(\boldsymbol{X})) \times 2k - k, \quad \text{for} \quad k > 0. 
+g(\boldsymbol{x}, k) = (\boldsymbol{x} - min(\boldsymbol{x}))/max(\boldsymbol{x} - min(\boldsymbol{x})) \times 2k - k, \quad \text{for} \quad k > 0. 
 \end{equation}
 
-The null regression model used to fit the data generated by the above model is formulated as:
+The null regression model used to fit the realizations generated by the above model is formulated as:
 
 \begin{equation} \label{eq:scaling-function}
-\boldsymbol{Y} = \beta_0 + \beta_1 \boldsymbol{X} + \boldsymbol{u},
+\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
 \end{equation}
@@
 
-Clearly, omitted-variable bias will present since the null model leaves out the higher order term.
-Data were simulated using four different order of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$), three different sample sizes (n = 50, 100, 300) , four different standard deviations of the error ($\sigma$ = 0.5, 1, 2, 4) and four different distribution of $\boldsymbol{X}_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
+Model misspecification presents since the null model leaves out the higher order term.
+
+Experiment data were simulated using four different order of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$), three different sample sizes (n = 50, 100, 300) , four different standard deviations of the error ($\sigma$ = 0.5, 1, 2, 4) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
 A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}.
 
-The values of $j$ was chosen so that different shapes of non-linearity are included in the residual plot. These include "U" shape, "S" shape, "M" shape and "Triple-U" shape.
+The values of $j$ was chosen so that different shapes of non-linearity were included in the residual plot. These include "U" shape, "S" shape, "M" shape and "Triple-U" shape.
 
@@
 
-The range of $\sigma$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated and the estimated power curve would be smooth and continuous.
+The range of $\sigma$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
 
@@
 
-Four different distribution were used to generate $\boldsymbol{X}_{raw}$.
+Four different distribution were used to generate $X_{raw}$.
 The uniform and the normal distribution are symmetric and commonly assumed in statistical models.
@@
 Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in 192 different lineups.
-For each lineup, the actual data plot was drawn as standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis.
+For each lineup, the actual data plot was drawn as a standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis.
 The 19 null datasets were generated by the residual rotation technique, and plotted in the same way.
@@
 It was produced by using $n = 300$, $j = 6$, $\sigma = 0.5$ and $X_{raw} \sim N(0.0.3^2)$.
-The actual data plot location is four.
+The actual data plot location was four.
 All five subjects correctly identified the actual data plot for this lineup.
@@
 
-In addition, each lineup is evaluated by five different subjects to provide reasonable estimates of the visual p-value.
-Thus, $192 \times 3 \times 5 / (20 - 2) = 160$ subjects were recruited to satisfy the experimental design described above.
+In addition, each lineup is designed to be evaluated by five different subjects to provide reasonable estimates of the visual p-value.
+Thus, $192 \times 3 \times 5 / (20 - 2) = 160$ subjects were recruited to satisfy the design of the experiment I.
 
@@
 
-A subject was allowed to select zero plots for a lineup, which literally means there was no visble difference between plots,
-but the simulated p-value given in Equation (\ref{eq:p-value-multi}) will effectively drop the subject from the simulation for this case, leading to inaccurate estimation of $p$-value.
+A subject was allowed to select zero plots for a lineup if there was no visible difference between plots,
+but the simulated visual p-value given in Equation (\ref{eq:p-value-multi}) will effectively drop the subject from the simulation for this case, leading to inaccurate estimation of $p$-value.
 Therefore, we treated this case as making one selection but failing to identify the actual data plot so that Equation (\ref{eq:p-value-multi}) can be applied correctly.
@@
 Table \ref{tab:demographic-table} tabulates the number of subjects, preferred pronouns, education backgrounds, age groups, and previous experience in visual experiment.
-Figure \ref{fig:demographic-summary-plot} visulizes the marginal distribution of each of the category.
+Figure \ref{fig:demographic-summary-plot} visualizes the marginal distribution of each of the category.
 The collated data was a balanced sample among male and female.
@@
 
-For each parameter combination, effect $E$ is derived from the Kullback-Leibler divergence (the derivation is provided in [appendix ref here]) formulated as:
+For each parameter combination, effect $E$ is derived from the Kullback-Leibler divergence (see [appendix ref here]) formulated as:
 
@@
 
-To study different factors contributing to the power of the visual test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors.
+To study various factors contributing to the power of the visual test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors.
 This includes [expansion].
@@
 At $log_e(E) = 4$ and  $log_e(E) = 6$, the designed visual features become much clear and attractive, leading to a high percentage of rejection of the null hypothesis.
-Figure \ref{fig:power-overview} give example lineups of such cases.
-
-It is also interesting to observe that the power of visual test behaves similarly to those inappropriate conventional tests until the designed visual features really becomes visible
-
-As of shown in Figure \ref{fig:power-overview}, the example lineup for $log_e(E) = 3$
-
-it is of low chance detecting the actual data plot
-
-increase steadily when $2<log_e(E)<4$. the power of visual test almost reaches 100% at $log_e(E) = 6$.
+Figure \ref{fig:power-overview} gives example lineups of such cases.
+
+<!-- It is also interesting to observe that the power curve of visual test behaves similarly to those inappropriate conventional tests until the designed visual features really becomes visible   -->
 
@@
 ```{r eval = FALSE}
-filter(polynomials, set == 123)
+filter(polynomials, lineup_id == 131)
 
@@
 
-```{r}
+### Distributions of regressor
+
+The impact of the distribution of $X_raw$ on the power is shown in Figure \ref{fig:dist-power}.
+The power curve of F-test is stable across different distributions, while the visual test has a steeper power curve for normal and uniform distribution.
+BP-test performs worse for discrete uniform distribution and uniform distribution but has relatively high power for normal distribution.
+SW-test outperforms BP-test for discrete uniform distribution but remains as the worst test for other distributions.
+The results indicate those inappropriate residual-based tests are sensitive to the distribution of the regressor.
+
+```{r dist-power, fig.cap="\\label{dist-power}"}
 conv_test_plot() %>%
@@
   add_visual_test_plot_layers() +
-  geom_vline(xintercept = 2, linetype = 2) + 
-  geom_vline(xintercept = 3, linetype = 2) +
-  geom_vline(xintercept = 4, linetype = 2) +
-  geom_vline(xintercept = 5, linetype = 2) +
   facet_wrap(~x_dist)
@@
 
-```{r}
+### Shpaes of non-linearity
+
+Figure \ref{fig:shape-power} illustrates the change of power under different shape of non-linearity.
+Similar to the power curves shown in \ref{fig:dist-power}, F-test is stable under different shapes.
+The power curve of visual test also behaves similarly across different shapes.
+What vary are the power curve of BP-test and SW-test.
+For Triple-U shape, both BP-test and SW-test are insensitive to the change of the effect.
+And for W shape, both tests have almost identical power curves.
+It can be observed that both tests performs the best for U-shape.
+
+```{r shape-power, fig.cap="\\label{shape-power}"}
 conv_test_plot() %>%
@@
   add_visual_test_plot_layers() +
-  geom_vline(xintercept = 2, linetype = 2) + 
-  geom_vline(xintercept = 3, linetype = 2) +
-  geom_vline(xintercept = 4, linetype = 2) +
-  geom_vline(xintercept = 5, linetype = 2) +
   facet_wrap(~shape)
+```
+
+```{r eval = FALSE}
+conv_test_plot() %>%
+  add_wrong_BP_conv_test_plot_layers() %>%
+  add_wrong_SW_conv_test_plot_layers() %>%
+  add_visual_test_plot_layers() +
+  facet_wrap(~e_sigma)
+```
+
+### Reasons for making selections
+
+```{r}
+polynomials %>%
+  mutate(reason_tmp = ifelse(reason %in% c("Cluster(s)", "Shape", "Outlier(s)"), reason, "Other")) %>%
+  ggplot() +
+  geom_bar(aes(reason_tmp)) +
+  facet_grid(x_dist~detect) 
+```
+
+## Heteroskedasticity
+
+```{r cache=TRUE}
+heter_result <- read_csv(here::here("data/processed/processed_surveys.csv"))
+
+heter_result <- heter_result %>%
+  filter(type != "cubic")
+
+heter_cache <- new.env()
+
+heter_result <- heter_result %>%
+  left_join(heter_result %>%
+  calc_p_value_multi(detected = "detect", n_sel = "num_selection", cache_env = heter_cache))
+
+heter_result <- heter_result %>%
+  mutate(weigthed_detect = ifelse(detect, detect/num_selection, 0)) %>%
+  group_by(lineup_id) %>%
+  mutate(prop_detect = mean(weigthed_detect))
+```
+
+```{r}
+map_to_group <- function(b) {
+  round(b / 4) + 1
+}
+
+heter_conv_test_plot <- function() {
+  conv_result_heter %>%
+  mutate(cat_b = map_to_group(b)) %>%
+  ggplot() +
+  stat_smooth(geom = "line", 
+              aes(log(effect_size), reject, group = boot_id, col = "BP-test"), 
+              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) +
+  stat_smooth(geom = "line", 
+              aes(log(effect_size), reject, col = "BP-test"), 
+              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1) +
+  labs(col = "Test") +
+  theme_light(base_size = 5) +
+  xlab(quote(log[e](E))) +
+  ylab("Power")
+}
+
+add_heter_SW_plot_layers <- function(p) {
+  p +
+  stat_smooth(data = mutate(conv_result_heter_SW, cat_b = map_to_group(b)), geom = "line", 
+              aes(log(effect_size), reject, group = boot_id, col = "SW-test"), 
+              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) +
+  stat_smooth(data = mutate(conv_result_heter_SW, cat_b = map_to_group(b)), geom = "line", 
+              aes(log(effect_size), reject, col = "SW-test"), 
+              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1)
+}
+
+add_heter_visual_test_plot_layers <- function(p, points = c()) {
+
+  visual_dat <- heter_result %>%
+  mutate(x_dist = ifelse(x_dist == "discrete_uniform", "even_discrete", x_dist)) %>%
+  mutate(x_dist = ifelse(x_dist == "neglognormal", "lognormal", x_dist)) %>%
+  group_by(lineup_id) %>%
+  summarise(across(c(type:p_value, prop_detect), first)) %>%
+  filter(ori_difficulty != "extremely_easy") %>%
+  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
+  boot_dat(times = 100) %>%
+  mutate(cat_b = map_to_group(b))
+
+  p +
+  stat_smooth(data = filter(visual_dat, boot_id == 1), geom = "line", 
+              aes(log(effect_size), reject, group = boot_id, col = "Visual"), 
+              method = "glm", method.args = list(family = binomial), se = FALSE, alpha = 1) +
+  stat_smooth(data = visual_dat, geom = "line", 
+              aes(log(effect_size), reject, group = boot_id, col = "Visual"), 
+              method = "glm", method.args = list(family = binomial), se = FALSE, alpha = 0.05) +
+  geom_point(data = filter(visual_dat, boot_id == 1), aes(log(effect_size), prop_detect), alpha = 0.1) + 
+  geom_point(data = filter(visual_dat, boot_id == 1, lineup_id %in% points), aes(log(effect_size), prop_detect), alpha = 1, col = "red", size = 2)
+}
+```
+
+```{r}
+heter_conv_test_plot() %>%
+  add_heter_SW_plot_layers() %>%
+  add_heter_visual_test_plot_layers()
+```
+
+```{r}
+heter_conv_test_plot() %>%
+  add_heter_SW_plot_layers() %>%
+  add_heter_visual_test_plot_layers() +
+  facet_wrap(~a)
+```
+
+```{r}
+heter_conv_test_plot() %>%
+  add_heter_SW_plot_layers() %>%
+  add_heter_visual_test_plot_layers() +
+  facet_wrap(~x_dist)
+```
+
+```{r}
+heter_conv_test_plot() %>%
+  add_heter_SW_plot_layers() %>%
+  add_heter_visual_test_plot_layers() +
+  facet_wrap(~cat_b)
 ```

### 2022-08-10 paper_comparison/paper_comparison/paper_comparison.tex
update graphs
latest update

@@
 normal linear regression model settings with only one or two regressors.
-Their results suggests visual test is capable of testing the
-significance of a single regressor with a similar power as a t-test,
-though they expressed that in general it is unnecessary to use visual
-inference if there exists a conventional test and they didn't expect the
-visual test to perform equally well as the conventional test. In their
-third experiment, where there does not exist a proper conventional test,
+Their results suggest visual test is capable of testing the significance
+of a single regressor with a similar power as a t-test, though they
+expressed that in general it is unnecessary to use visual inference if
+there exists a conventional test and they didn't expect the visual test
+to perform equally well as the conventional test. In their third
+experiment, where there does not exist a proper conventional test,
 visual test outperforms the conventional test for a large margin. This
@@
 For the purpose of examining the effectiveness of visual test in
-regression diagnostics, two experiments were conducted. The first
-experiment has ideal scenario for conventional testing, where we would
-not expect the visual test to do better than the conventional test. The
-second experiment is a scenario where the conventional test is an
-approximate test, which visual test may have a chance to match the
-performance of the conventional test.
+regression diagnostics, two experiments were conducted. The experiment I
+has ideal scenario for conventional testing, where the visual test is
+not expected to outperform the conventional test. The experiment II is a
+scenario where the conventional test is an approximate test, in which
+the visual test may have a chance to match the performance of the
+conventional test.
 
-As suggested by many previous studies {[}many studies refs here{]}, it
-is reasonable for an subject to evaluate a block of 20 lineups. Subjects
-for both two experiments were recruited from Prolific web service {[}ref
-here{]}. Data screening procedure was applied during the recruitment,
-subjects were required to be fluent in English, with \(98\%\) minimum
-approval rate and 10 minimum submissions. During the experiment, they
-were asked to select one or more plots that are mostly different from
-others, provide a reason for their selections, and express how different
-they think the selected plots are from others. If there was no
-noticeable difference between plots in a lineup, subject could choose to
-select zero plots without providing the reason. Information about
-preferred pronoun, age group, education, and previous experience in
-visual experiment were also collected. No subject was shown the same
-lineup twice.
+Subjects for both experiments were recruited from an crowdsourcing
+platform called Prolific {[}ref here{]}. Prescreening procedure was
+applied during the recruitment, subjects were required to be fluent in
+English, with \(98\%\) minimum approval rate in other studies and 10
+minimum submissions.
 
-For each of the experiments, a pool of 12 lineups with obvious visual
-patterns were generated. In every block of 20 lineups that presented to
-subjects, two different lineups sampled from the corresponding pool of
-lineups were included as attention checks. Every subject should get at
-least one attention check correct such that the quality of the subject
-effort could be ensured.
+During the experiment, every subject was presented with a block of 20
+lineups. And for every lineup, the subject was asked to select one or
+more plots that are most different from others, provide a reason for
+their selections, and evaluate how different they think the selected
+plots were from others. If there was no noticeable difference between
+plots in a lineup, subjects were permitted to select zero plots without
+providing the reason. No subject was shown the same lineup twice.
+Information about preferred pronoun, age group, education, and previous
+experience in visual experiment were also collected.
 
-\hypertarget{non-linear-covariates}{%
-\subsection{Non-linear covariates}\label{non-linear-covariates}}
+A pool of 12 different lineups with obvious visual patterns were
+generated for experiment I and experiment II respectively. In every
+block of 20 lineups that presented to a subject, two out of 12 lineups
+were included as attention checks. A subject's submission was only
+accepted if the actual data plot was identified for at least one
+attention check. Data of rejected submissions were discarded
+automatically to maintain the overall data quality.
 
-The first experiment is designed to study the ability of human subjects
-to detect the effect of a continuous variable \(\boldsymbol{Z}\) which
-is a probabilist's Hermite polynomial {[}Herimite ref here{]} of another
-variable \(\boldsymbol{X}\) in a two variable statistical model
+\hypertarget{non-linearity}{%
+\subsection{Non-linearity}\label{non-linearity}}
+
+Experiment I is designed to study the ability of human subjects to
+detect the effect of a random vector \(\boldsymbol{z}\) which is a
+probabilist's Hermite polynomial {[}Herimite ref here{]} of another
+random vector \(\boldsymbol{x}\) in a two variable statistical model
 formulated as:
@@
 \begin{align} \label{eq:nonlinearity-model}
-\boldsymbol{Y} = 1 + \boldsymbol{X} + \boldsymbol{Z} + \boldsymbol{\varepsilon},\\
-\boldsymbol{X} = g(\boldsymbol{X}_{raw}, 1), \\
-\boldsymbol{Z} = g(\boldsymbol{Z}_{raw}, 1), \\
-\boldsymbol{Z}_{raw} = He_j(g(\boldsymbol{X}, 2)),
+\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
+\boldsymbol{x} = g(\boldsymbol{x}_{raw}, 1), \\
+\boldsymbol{z} = g(\boldsymbol{z}_{raw}, 1), \\
+\boldsymbol{z}_{raw} = He_j(g(\boldsymbol{z}, 2)),
 \end{align}
 
-where \(\boldsymbol{Y}\), \(\boldsymbol{X}\),
-\(\boldsymbol{\varepsilon}\), \(\boldsymbol{X}_{raw}\),
-\(\boldsymbol{Z}_{raw}\) are \(n\times1\) matrices, \(He_{j}(.)\) is the
+where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
+\(\boldsymbol{\varepsilon}\), \(\boldsymbol{x}_{raw}\),
+\(\boldsymbol{z}_{raw}\) are vector of size \(n\), \(He_{j}(.)\) is the
 \(j\)th-order probabilist's Hermite polynomials,
 \(\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\), and
-\(g(\boldsymbol{X}, k)\) is a scaling function to enforce the support of
-the variable to be \(\{-k, k\}\) defined as
+\(g(\boldsymbol{x}, k)\) is a scaling function to enforce the support of
+the random vector to be \(\{-k, k\}\) defined as
 
 \begin{equation} \label{eq:scaling-function}
-g(\boldsymbol{X}, k) = (\boldsymbol{X} - min(\boldsymbol{X}))/max(\boldsymbol{X} - min(\boldsymbol{X})) \times 2k - k, \quad \text{for} \quad k > 0. 
+g(\boldsymbol{x}, k) = (\boldsymbol{x} - min(\boldsymbol{x}))/max(\boldsymbol{x} - min(\boldsymbol{x})) \times 2k - k, \quad \text{for} \quad k > 0. 
 \end{equation}
 
-The null regression model used to fit the data generated by the above
-model is formulated as:
+The null regression model used to fit the realizations generated by the
+above model is formulated as:
 
 \begin{equation} \label{eq:scaling-function}
-\boldsymbol{Y} = \beta_0 + \beta_1 \boldsymbol{X} + \boldsymbol{u},
+\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
 \end{equation}
@@
 
-Clearly, omitted-variable bias will present since the null model leaves
-out the higher order term. Data were simulated using four different
-order of probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)), three
-different sample sizes (n = 50, 100, 300) , four different standard
-deviations of the error (\(\sigma\) = 0.5, 1, 2, 4) and four different
-distribution of \(\boldsymbol{X}_{raw}\): (1) \(U(-1, 1)\), (2)
-\(N(0, 0.3^2)\), (3) \(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). A
-summary of the parameters used in this experiment is given in Table
+Model misspecification presents since the null model leaves out the
+higher order term.
+
+Experiment data were simulated using four different order of
+probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)), three different
+sample sizes (n = 50, 100, 300) , four different standard deviations of
+the error (\(\sigma\) = 0.5, 1, 2, 4) and four different distribution of
+\(X_{raw}\): (1) \(U(-1, 1)\), (2) \(N(0, 0.3^2)\), (3)
+\(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). A summary of the
+parameters used in this experiment is given in Table
 \ref{tab:parameter-table}.
@@
 The values of \(j\) was chosen so that different shapes of non-linearity
-are included in the residual plot. These include ``U'' shape, ``S''
+were included in the residual plot. These include ``U'' shape, ``S''
 shape, ``M'' shape and ``Triple-U'' shape.
@@
 the signal, was chosen so that different difficulty levels of lineups
-were generated and the estimated power curve would be smooth and
-continuous.
+were generated, and therefore, the estimated power curve would be smooth
+and continuous.
 
@@
 
-Four different distribution were used to generate
-\(\boldsymbol{X}_{raw}\). The uniform and the normal distribution are
-symmetric and commonly assumed in statistical models. The adjusted
-log-normal distribution provides skewed density. And the discrete
-uniform distribution provides discreteness in residual plot, which could
-enrich the pool of visual patterns.
+Four different distribution were used to generate \(X_{raw}\). The
+uniform and the normal distribution are symmetric and commonly assumed
+in statistical models. The adjusted log-normal distribution provides
+skewed density. And the discrete uniform distribution provides
+discreteness in residual plot, which could enrich the pool of visual
+patterns.
 
@@
 Table \ref{tab:parameter-table} resulting in 192 different lineups. For
-each lineup, the actual data plot was drawn as standard residual plot of
-the null model with raw residuals on the y-axis and fitted values on the
-x-axis. The 19 null datasets were generated by the residual rotation
+each lineup, the actual data plot was drawn as a standard residual plot
+of the null model with raw residuals on the y-axis and fitted values on
+the x-axis. The 19 null datasets were generated by the residual rotation
 technique, and plotted in the same way. The lineup consisted of 20
@@
 produced by using \(n = 300\), \(j = 6\), \(\sigma = 0.5\) and
-\(X_{raw} \sim N(0.0.3^2)\). The actual data plot location is four. All
+\(X_{raw} \sim N(0.0.3^2)\). The actual data plot location was four. All
 five subjects correctly identified the actual data plot for this lineup.
@@
 
-In addition, each lineup is evaluated by five different subjects to
-provide reasonable estimates of the visual p-value. Thus,
+In addition, each lineup is designed to be evaluated by five different
+subjects to provide reasonable estimates of the visual p-value. Thus,
 \(192 \times 3 \times 5 / (20 - 2) = 160\) subjects were recruited to
-satisfy the experimental design described above.
+satisfy the design of the experiment I.
 
@@
 
-A subject was allowed to select zero plots for a lineup, which literally
-means there was no visble difference between plots, but the simulated
-p-value given in Equation (\ref{eq:p-value-multi}) will effectively drop
-the subject from the simulation for this case, leading to inaccurate
-estimation of \(p\)-value. Therefore, we treated this case as making one
-selection but failing to identify the actual data plot so that Equation
+A subject was allowed to select zero plots for a lineup if there was no
+visible difference between plots, but the simulated visual p-value given
+in Equation (\ref{eq:p-value-multi}) will effectively drop the subject
+from the simulation for this case, leading to inaccurate estimation of
+\(p\)-value. Therefore, we treated this case as making one selection but
+failing to identify the actual data plot so that Equation
 (\ref{eq:p-value-multi}) can be applied correctly.
@@
 experience in visual experiment. Figure
-\ref{fig:demographic-summary-plot} visulizes the marginal distribution
+\ref{fig:demographic-summary-plot} visualizes the marginal distribution
 of each of the category. The collated data was a balanced sample among
@@
 For each parameter combination, effect \(E\) is derived from the
-Kullback-Leibler divergence (the derivation is provided in {[}appendix
-ref here{]}) formulated as:
+Kullback-Leibler divergence (see {[}appendix ref here{]}) formulated as:
 
@@
 
-To study different factors contributing to the power of the visual test,
+To study various factors contributing to the power of the visual test,
 the same logistic regression model is fit on different subsets of the
@@
 attractive, leading to a high percentage of rejection of the null
-hypothesis. Figure \ref{fig:power-overview} give example lineups of such
-cases.
-
-It is also interesting to observe that the power of visual test behaves
-similarly to those inappropriate conventional tests until the designed
-visual features really becomes visible
-
-As of shown in Figure \ref{fig:power-overview}, the example lineup for
-\(log_e(E) = 3\)
-
-it is of low chance detecting the actual data plot
-
-increase steadily when \(2<log_e(E)<4\). the power of visual test almost
-reaches 100\% at \(log_e(E) = 6\).
+hypothesis. Figure \ref{fig:power-overview} gives example lineups of
+such cases.
 
@@
 \end{figure}
+
+\hypertarget{distributions-of-regressor}{%
+\subsubsection{Distributions of
+regressor}\label{distributions-of-regressor}}
+
+The impact of the distribution of \(X_raw\) on the power is shown in
+Figure \ref{fig:dist-power}. The power curve of F-test is stable across
+different distributions, while the visual test has a steeper power curve
+for normal and uniform distribution. BP-test performs worse for discrete
+uniform distribution and uniform distribution but has relatively high
+power for normal distribution. SW-test outperforms BP-test for discrete
+uniform distribution but remains as the worst test for other
+distributions. The results indicate those inappropriate residual-based
+tests are sensitive to the distribution of the regressor.
+
+\begin{figure}
+\centering
+\includegraphics{paper_comparison_files/figure-latex/dist-power-1.pdf}
+\caption{\label{dist-power}}
+\end{figure}
+
+\hypertarget{shpaes-of-non-linearity}{%
+\subsubsection{Shpaes of non-linearity}\label{shpaes-of-non-linearity}}
+
+Figure \ref{fig:shape-power} illustrates the change of power under
+different shape of non-linearity. Similar to the power curves shown in
+\ref{fig:dist-power}, F-test is stable under different shapes. The power
+curve of visual test also behaves similarly across different shapes.
+What vary are the power curve of BP-test and SW-test. For Triple-U
+shape, both BP-test and SW-test are insensitive to the change of the
+effect. And for W shape, both tests have almost identical power curves.
+It can be observed that both tests performs the best for U-shape.
+
+\begin{figure}
+\centering
+\includegraphics{paper_comparison_files/figure-latex/shape-power-1.pdf}
+\caption{\label{shape-power}}
+\end{figure}
+
+\hypertarget{reasons-for-making-selections}{%
+\subsubsection{Reasons for making
+selections}\label{reasons-for-making-selections}}
+
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-12-1.pdf}
+
+\hypertarget{heteroskedasticity}{%
+\subsection{Heteroskedasticity}\label{heteroskedasticity}}
+
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-15-1.pdf}
+
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-16-1.pdf}
+
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-17-1.pdf}
+
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-18-1.pdf}
 
@@
 
-\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-9-1.pdf}
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-19-1.pdf}
 

### 2022-08-17 paper_comparison/paper_comparison/paper_comparison.tex
intro restructured

@@
 
-% see https://stackoverflow.com/a/47122900
-
-% Pandoc citation processing
+% tightlist command for lists without linebreak
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
@@
 \newcommand*{\Perm}[2]{{}^{#1}\!P_{#2}}%
+
 \usepackage{booktabs}
@@
 
-\title{Effectiveness of visual test in linear regression diagnositics}
+\title{Why aren't significance tests commonly used for linear regression
+diagnostics?}
 
@@
 \begin{keywords}
-visual inference; regression diagnostics;
+data visualization; visual inference; hypothesis testing; residual
+plots;
 \end{keywords}
@@
 
-Regression diagnostics is an essential step in regression analysis which
-is a field of study with at least a hundred years of history. The
-diagnostic procedure conventionally involves evaluating the fitness of
-the proposed model, detecting the presence of influential observations
-and outliers, checking the validity of model assumptions and many more.
-In terms of diagnostic techniques, data plots, hypothesis testing, and
-summary statistics are vital tools for a systematic and detailed
-examination of the regression model \citep{mansfield1987diagnostic}.
-
-Many of those regression diagnostic methods and procedures are mature
-and well-established in books first published in the twentieth century,
-such as \citet{draper1998applied}, \citet{montgomery1982introduction},
-\citet{belsley_regression_1980}, \citet{cook_applied_1999} and
-\citet{cook1982residuals}. Regardless of the level of difficulty, one
-will find the importance and usefulness of diagnostic plots being
-emphasized in those books repeatedly. Checking diagnostic plots is also
-the recommended starting point for validating model assumptions such as
-normality, homoscedasticity and linearity
-\citep{anscombe_examination_1963}.
+\begin{quote}
+\emph{``Since all models are wrong the scientist must be alert to what
+is importantly wrong.''} \citep{REF}
+\end{quote}
+
+Diagnosing a model is the key to determining whether there is anything
+importantly wrong. For linear regression analysis, it is typical to
+interrogate the residuals. Residuals summarise what is not captured by
+the model, and thus provide the capacity to identify what might be
+wrong. There are many ways that residuals are assessed.
+
+Residuals might be plotted, as a histogram or normal probability plot to
+examine the distribution. Is it symmetric, and unimodal? Then all is
+good. If the distribution is skewed or bimodal or multimodal, there are
+outliers, there is cause for concern. It if also possible to conduct a
+goodness of fit test, such as a Shapiro-Wilk Normality test \citep{REF}.
+
+It is also recommended to plot residuals against predicted values, and
+also each of the explanatory variables. If there is no obvious pattern,
+then all is good. It is a very difficult task for a human judge, though
+to make a decision that there's nothing there. It is especially common,
+particularly among new data analysts to report patterns when an
+experienced data analyst might quickly conclude that there are none.
+Generally, one looks for departures from nothingness like nonlinear
+dependency or heteroskedasticity. It is also possible to conduct a
+hypothesis test for non-linear dependence \citep{RESET}, and also use a
+Breusch-Pagan test \citep{REF} for heteroskedasticity.
+
+There is an abundance of literature describing appropriate diagnostic
+methods for linear regression: Cook and Weisberg, \ldots{} \citep{REF}.
+The diligent reader of these sage writings will also notice sentences
+that express sentiments like \emph{based on their experience,
+statistical tests are not widely used in regression diagnostics. The
+same or even larger amount of information can be provided by diagnostic
+plots than the corresponding tests in most empirical studies.} There is
+a common guidance by experts that plots are the best for diagnosing
+model fits.
+
+This is curious, and investigating why this might be common advice is
+the subject of this paper. The paper is structured as follows. The next
+background section describes the commonly used diagnostic plots, the
+types of departures that one expects to detect, and describes a formal
+process for reading residual plots, called visual inference, that can
+avoid the concerns about subjectiveness of human readers. Section XXX
+describes the experimental setup to enable a comparison between
+decisopns made by formal hypothesis testing, and how humans would read
+diagnostic plots. The results are reported in Section XXX. We finish
+with a discussion on future work, in particular how the responsibility
+for residual plot reading might be passed on to computer vision.
+
+\hypertarget{background}{%
+\section{Background}\label{background}}
 
@@
 tests and visual discovery.
-
-\begin{figure}
-\centering
-\includegraphics[width=4.6875in,height=3.55208in]{figures/rsta2009012001.jpg}
-\caption{Parallelism between multiple quantitative testing and visual
-discovery \citep{buja_statistical_2009}. Visible features in a plot are
-viewed as a collection of test statistics
-\(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\), and any visual
-discoveries that are inconsistent with the null hypothesis are treated
-as evidence against the null. For regression diagnostics, the null
-hypothesis would be the assumed model, and visual discoveries could be
-any visual features in favour of any alternatives.
-\label{fig:parallelism}}
-\end{figure}
 

### 2022-08-24 Fourth_experiment/check_allocation_2.R
fourth experiment

@@ -1,0 +1,127 @@
+library(tidyverse)
+library(visage)
+
+lineup <- read_csv("data/lineup_info.csv")
+allocated <- read_csv("data/allocate_result.csv")
+dat <- vector(mode = "list", length = nrow(lineup))
+
+set.seed(10086)
+
+stand_dist <- function(x) {
+  (x - min(x))/max(x - min(x)) * 2 - 1
+}
+
+for (i in 1:nrow(lineup)) {
+
+  x <- switch(lineup$x_dist[i],
+              uniform = rand_uniform(-1, 1),
+              normal = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))},
+              lognormal = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))},
+              even_discrete = rand_uniform_d(-1, 1, k = 5, even = TRUE))
+  mod <- heter_model(a = lineup$a[i], b = lineup$b[i], x = x)
+  lineup_dat <- mod$gen_lineup(lineup$n[i], 20)
+
+  dat[[lineup$lineup_id[i]]]$metadata <- list(type = "heteroskedasticity",
+                                              formula = deparse(mod$formula),
+                                              a = lineup$a[i],
+                                              b = lineup$b[i],
+                                              x_dist = lineup$x_dist[i],
+                                              e_dist = "normal",
+                                              e_sigma = 1,
+                                              name = glue::glue("heter_{lineup$lineup_id[i]}"),
+                                              k = 20,
+                                              n = lineup$n[i],
+                                              effect_size = mod$effect_size(filter(lineup_dat, null == FALSE), type = "kl"),
+                                              ans = filter(lineup_dat, null == FALSE)$k[1]
+
+  )
+
+  dat[[lineup$lineup_id[i]]]$data <- lineup_dat
+}
+
+
+map_dbl(lineup$lineup_id[lineup$b == 0], ~dat[[.x]]$metadata$effect_size)
+
+p_value <- map_dbl(1:nrow(lineup), ~HETER_MODEL$test(filter(dat[[.x]]$data, null == FALSE))$p_value)
+
+# false positive rate, approx 3% reject
+mean(p_value[lineup$lineup_id[lineup$b == 0]] < 0.05)
+
+# around 72% reject, 28% accept
+mean(p_value[lineup$lineup_id[lineup$b <= 64 & lineup$b > 0]] < 0.05)
+
[…]

### 2022-08-24 Fourth_experiment/exp_setup.R

@@ -1,0 +1,316 @@
+library(visage)
+library(tidyverse)
+library(cli)
+
+
+# We need
+# a (3) = -1, 0, 1
+# b (5) = 0.25, 1, 4, 16, 64
+# x_dist (4) = uniform, normal, lognormal, even_discrete
+# n (3) = 50, 100, 300
+# rep (3)
+#
+# 3 * 5 * 4 * 3 * 3 = 540
+#
+# null lineups for heter and poly (the null models are exactly the same):
+# x_dist (4)
+# n (3)
+# rep (3)
+# 4 * 3 * 3 = 36
+#
+# 540 + 36 = 576 lineups
+# 576 * 5 / 18 = 160
+#
+# We need 160 subjects to evaluate 540 lineups and 36 null lineups for 5 times
+
+
+set.seed(10086)
+
+# Generate unique lineups
+lineup <- expand.grid(a = c(-1, 0, 1),
+                      b = c(0, 0.25, 1, 4, 16, 64),
+                      n = c(50, 100, 300),
+                      x_dist = c("uniform", "normal", "lognormal", "even_discrete"),
+                      rep = 1:3) %>%
+  filter(!(b == 0 & a != 0)) %>%
+  mutate(lineup_id = sample(1:n()))
+
+check_comb <- function(set_dat, new_dat, n = 1) {
+
+  apply(combn(dim(set_dat)[2], n),
+        MARGIN = 2,
+        function(this_comb) {
+          # browser()
+          tmp <- as.numeric(new_dat[1, this_comb])
+          for (i in 1:nrow(set_dat)) {
+            if (all(as.numeric(set_dat[i, this_comb]) == tmp)) return(TRUE)
+          }
+          return(FALSE)
+        })
+}
+
+sorted_lineup <- lineup %>%
+  arrange(lineup_id) %>%
+  select(-rep, -lineup_id)
+
+# Caculate distance between a set of allocated lineups and a new lineup
+calc_dist <- function(lineup, allocated_lineup_ids, new_lineup_id) {
+
+  new_lineup <- sorted_lineup[new_lineup_id, ]
+  allocated_lineups <- sorted_lineup[allocated_lineup_ids, ]
+
+  # We want these three options to be as different as possible
+  15 -
+    sum(check_comb(allocated_lineups, new_lineup, 1)) -
+    sum(check_comb(allocated_lineups, new_lineup, 2)) -
[…]

### 2022-08-24 Fourth_experiment/generate_lineup.R

@@ -1,0 +1,110 @@
+library(tidyverse)
+library(visage)
+
+lineup <- read_csv("data/lineup_info.csv")
+plot_string <- read_csv("data/plot_string.txt", col_names = FALSE) %>%
+  .$X1
+allocated <- read_csv("data/allocate_result.csv")
+dat <- vector(mode = "list", length = nrow(lineup))
+
+set.seed(10086)
+
+stand_dist <- function(x) {
+  (x - min(x))/max(x - min(x)) * 2 - 1
+}
+
+for (i in 1:nrow(lineup)) {
+
+  x <- switch(lineup$x_dist[i],
+              uniform = rand_uniform(-1, 1),
+              normal = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))},
+              lognormal = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))},
+              even_discrete = rand_uniform_d(-1, 1, k = 5, even = TRUE))
+  mod <- heter_model(a = lineup$a[i], b = lineup$b[i], x = x)
+  lineup_dat <- mod$gen_lineup(lineup$n[i], 20)
+
+  dat[[lineup$lineup_id[i]]]$metadata <- list(type = "heteroskedasticity",
+                                              formula = deparse(mod$formula),
+                                              a = lineup$a[i],
+                                              b = lineup$b[i],
+                                              x_dist = lineup$x_dist[i],
+                                              e_dist = "normal",
+                                              e_sigma = 1,
+                                              name = glue::glue("heter_{lineup$lineup_id[i]}"),
+                                              k = 20,
+                                              n = lineup$n[i],
+                                              effect_size = mod$effect_size(filter(lineup_dat, null == FALSE), type = "kl"),
+                                              ans = filter(lineup_dat, null == FALSE)$k[1]
+
+  )
+
+  dat[[lineup$lineup_id[i]]]$data <- lineup_dat
+}
+
+map_dbl(lineup$lineup_id[lineup$b == 0], ~dat[[.x]]$metadata$effect_size)
+
+p_value <- map_dbl(1:nrow(lineup), ~HETER_MODEL$test(filter(dat[[.x]]$data, null == FALSE))$p_value)
+
[…]

### 2022-08-24 paper_comparison/paper_comparison/paper_comparison.Rmd

@@
 title: |
-  Effectiveness of visual test in linear regression diagnositics  
+  Why aren't significance tests commonly used for linear regression diagnostics?
 type: ARTICLE TEMPLATE
@@
 keywords: |
-  visual inference; regression diagnostics;
+  data visualization; visual inference; hypothesis testing; residual plots;
 header-includes: |
@@
 
-Regression diagnostics is an essential step in regression analysis which is a field of study with at least a hundred years of history.
+> *"Since all models are wrong the scientist must be alert to what is importantly wrong."* [@REF]
+
+Diagnosing a model is the key to determining whether there is anything importantly wrong.
+For linear regression analysis, it is typical to interrogate the residuals.
+Residuals summarise what is not captured by the model, and thus provide the capacity to identify what might be wrong.
+There are many ways that residuals are assessed.
+
+Residuals might be plotted, as a histogram or normal probability plot to examine the distribution.
+Is it symmetric, and unimodal?
+Then all is good.
+If the distribution is skewed or bimodal or multimodal, there are outliers, there is cause for concern.
+It if also possible to conduct a goodness of fit test, such as a Shapiro-Wilk Normality test [@REF].
+
+It is also recommended to plot residuals against predicted values, and also each of the explanatory variables.
+If there is no obvious pattern, then all is good.
+It is a very difficult task for a human judge, though to make a decision that there's nothing there.
+It is especially common, particularly among new data analysts to report patterns when an experienced data analyst might quickly conclude that there are none.
+Generally, one looks for departures from nothingness like nonlinear dependency or heteroskedasticity.
+It is also possible to conduct a hypothesis test for non-linear dependence [@RESET], and also use a Breusch-Pagan test [@REF] for heteroskedasticity.
+
+There is an abundance of literature describing appropriate diagnostic methods for linear regression: Cook and Weisberg, ...
+[@REF].
+The diligent reader of these sage writings will also notice sentences that express sentiments like *based on their experience, statistical tests are not widely used in regression diagnostics.
+The same or even larger amount of information can be provided by diagnostic plots than the corresponding tests in most empirical studies.* There is a common guidance by experts that plots are the best for diagnosing model fits.
+
+This is curious, and investigating why this might be common advice is the subject of this paper.
+The paper is structured as follows.
+The next background section describes the commonly used diagnostic plots, the types of departures that one expects to detect, and describes a formal process for reading residual plots, called visual inference, that can avoid the concerns about subjectiveness of human readers.
+Section XXX describes the experimental setup to enable a comparison between decisopns made by formal hypothesis testing, and how humans would read diagnostic plots.
+The results are reported in Section XXX.
+We finish with a discussion on future work, in particular how the responsibility for residual plot reading might be passed on to computer vision.
+
+<!--Regression diagnostics is an essential step in regression analysis which is a field of study with at least a hundred years of history.
 The diagnostic procedure conventionally involves evaluating the fitness of the proposed model, detecting the presence of influential observations and outliers, checking the validity of model assumptions and many more.
@@
 Checking diagnostic plots is also the recommended starting point for validating model assumptions such as normality, homoscedasticity and linearity [@anscombe_examination_1963].
+-->
+
+# Background
 
@@
 
-![Parallelism between multiple quantitative testing and visual discovery [@buja_statistical_2009].
+<!-- ![Parallelism between multiple quantitative testing and visual discovery [@buja_statistical_2009].
 Visible features in a plot are viewed as a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$, and any visual discoveries that are inconsistent with the null hypothesis are treated as evidence against the null.
@@
 \label{fig:parallelism}](figures/rsta2009012001.jpg){width=450 height=341}
+-->
 

### 2022-08-31 paper_comparison/paper_comparison/paper_comparison.Rmd
update the writing and code in main doc

@@
 
+<!-- check the policy -->
+
 <!-- problem: residual plot diagnostics -->
@@
 
-if (!file.exists(here::here("data/third_study/conv_result_heter.rds"))) {
+if (!file.exists(here::here("data/fourth_study/conv_result_heter.rds"))) {
   conv_result_heter <- list()
@@
                      a <- sample(c(-1, 0, 1), 1)
-                     b <- runif(1, 0, 32)
+                     b <- sample(c(0.25, 1, 4, 16, 64), 1)
                      x_dist <- sample(c("uniform", "normal", "lognormal", "even_discrete"), 1)
@@
                                  lognormal = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))},
-                                 even_discrete = rand_uniform_d(k = 5, even = TRUE))
+                                 even_discrete = rand_uniform_d(-1, 1, k = 5, even = TRUE))
                      mod <- heter_model(a = a, b = b, x = x)
@@
                             n = n,
-                            effect_size = mod$effect_size(tmp_dat),
+                            effect_size = mod$effect_size(tmp_dat, type = "kl"),
                             p_value = mod$test(tmp_dat)$p_value,
@@
 
-  saveRDS(conv_result_heter, here::here("data/third_study/conv_result_heter.rds"))
+  saveRDS(conv_result_heter, here::here("data/fourth_study/conv_result_heter.rds"))
 } else {
-  conv_result_heter <- readRDS(here::here("data/third_study/conv_result_heter.rds"))
+  conv_result_heter <- readRDS(here::here("data/fourth_study/conv_result_heter.rds"))
 }
@@
 
-if (!file.exists(here::here("data/third_study/conv_result_heter_SW.rds"))) {
+if (!file.exists(here::here("data/fourth_study/conv_result_heter_SW.rds"))) {
   conv_result_heter_SW <- list()
@@
                             n = n,
-                            effect_size = mod$effect_size(tmp_dat),
+                            effect_size = mod$effect_size(tmp_dat, type = "kl"),
                             p_value = shapiro.test(tmp_dat$.resid)$p.value,
@@
 
-  saveRDS(conv_result_heter_SW, here::here("data/third_study/conv_result_heter_SW.rds"))
+  saveRDS(conv_result_heter_SW, here::here("data/fourth_study/conv_result_heter_SW.rds"))
 } else {
-  conv_result_heter_SW <- readRDS(here::here("data/third_study/conv_result_heter_SW.rds"))
+  conv_result_heter_SW <- readRDS(here::here("data/fourth_study/conv_result_heter_SW.rds"))
 }
@@
 
-> *"Since all models are wrong the scientist must be alert to what is importantly wrong."* [@REF]
+> *"Since all models are wrong the scientist must be alert to what is importantly wrong [@box1976science]."*
 
@@
 Residuals summarise what is not captured by the model, and thus provide the capacity to identify what might be wrong.
-There are many ways that residuals are assessed.
-
-Residuals might be plotted, as a histogram or normal probability plot to examine the distribution.
-Is it symmetric, and unimodal?
-Then all is good.
-If the distribution is skewed or bimodal or multimodal, there are outliers, there is cause for concern.
-It if also possible to conduct a goodness of fit test, such as a Shapiro-Wilk Normality test [@REF].
-
-It is also recommended to plot residuals against predicted values, and also each of the explanatory variables.
-If there is no obvious pattern, then all is good.
-It is a very difficult task for a human judge, though to make a decision that there's nothing there.
+There are many ways that residuals could be assessed.
+
+Residuals might be plotted, as a histogram or quantile-quantile plot to examine the distribution.
+Using classical linear regression model as an example, if the distribution is symmetric and unimodal, then it is well-behaved.
+But if the distribution is skewed, bimodal, multimodal, or even contains outliers, there is cause for concern.
+It is also possible to conduct a goodness of fit test to examine the distribution of residuals, such as Shapiro-Wilk Normality test [@shapiro1965analysis].
+
+To examine the relationship between residuals, predicted values and each of the explanatory variables, it is recommend to plot them on a scatter plot with residuals on the y-axis.
+This type of plot is usually called residual plot.
+In general, one will check whether there is any noticeable pattern to determine if the model is well-behaved.
+However, it is a very difficult task for a human judge, though to make a decision that there's nothing there.
 It is especially common, particularly among new data analysts to report patterns when an experienced data analyst might quickly conclude that there are none.
 Generally, one looks for departures from nothingness like nonlinear dependency or heteroskedasticity.
-It is also possible to conduct a hypothesis test for non-linear dependence [@RESET], and also use a Breusch-Pagan test [@REF] for heteroskedasticity.
-
-There is an abundance of literature describing appropriate diagnostic methods for linear regression: Cook and Weisberg, ...
-[@REF].
+It is also possible to conduct a hypothesis test for non-linear dependence [@ramsey_tests_1969], and also use a Breusch-Pagan test [@breusch_simple_1979] for heteroskedasticity.
+
+There is an abundance of literature describing appropriate diagnostic methods for linear regression: @draper1998applied, @montgomery1982introduction, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals.
 The diligent reader of these sage writings will also notice sentences that express sentiments like *based on their experience, statistical tests are not widely used in regression diagnostics.
@@
 The next background section describes the commonly used diagnostic plots, the types of departures that one expects to detect, and describes a formal process for reading residual plots, called visual inference, that can avoid the concerns about subjectiveness of human readers.
-Section XXX describes the experimental setup to enable a comparison between decisopns made by formal hypothesis testing, and how humans would read diagnostic plots.
-The results are reported in Section XXX.
+Section \ref{experimental-design} describes the experimental setup to enable a comparison between decision made by formal hypothesis testing, and how humans would read diagnostic plots.
+The results are reported in Section \ref{results}.
 We finish with a discussion on future work, in particular how the responsibility for residual plot reading might be passed on to computer vision.
@@
 
-## Hypothesis testing
-
-Other than checking diagnostic plots, analysts may perform formal hypothesis testing for detecting model defects.
+<!-- ## Hypothesis testing -->
+
+<!-- Other than checking diagnostic plots, analysts may perform formal hypothesis testing for detecting model defects.
 Depending on the alternative hypothesis that is focused on, a variety of tests can be applied.
@@
 For testing non-linearity, one may apply the F-test to examine the significance of specific polynomial and non-linear forms of the regressors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969].
-
-As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are sensitive to other types of departures.
+-->
+
+<!-- As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are sensitive to other types of departures.
 It is likely the null hypothesis is correctly rejected but for the wrong reason, which is known as the "Type III error".
@@
 A good regression diagnostic practice should be a balanced combination of both methods.
+-->
 
@@
 
+## Heteroskedasticity
+
+Heter Experiment design here
+
 # Results
@@
 
-In overall, there were a total of 3200 lineup evaluations made by 160 subjects in experiment I, where 320 lineup evaluations were attention checks and were not used in the following analysis.
-The collated dataset is provided in `polynomials` of the `visage` `R` package.
+In overall, there were a total of 3200 lineup evaluations made by 160 subjects in both experiment I and experiment II respectively, where 320 lineup evaluations were attention checks and were not used in the following analysis.
+
+The collated dataset is provided in `polynomials` and `heter` of the `visage` `R` package.
 
@@
 polynomials %>%
-  group_by(set) %>%
+  mutate(exp = 1) %>%
+  bind_rows(mutate(heter, exp = 2)) %>%
+  group_by(exp, set) %>%
   slice_sample(n = 1) %>%
@@
 polynomials %>%
-  group_by(set) %>%
+  mutate(exp = 1) %>%
+  bind_rows(mutate(heter, exp = 2)) %>%
+  group_by(exp, set) %>%
   summarise(across(age_group:previous_experience, first)) %>%
@@
 ## Power comparison
+
+### Experiment I
 
@@
   filter(lineup_id == 364)
+
+polynomials %>%
+  filter(e_sigma == 4) %>%
+  group_by(lineup_id) %>%
+  summarise(p_value = first(p_value)) %>%
+  mutate(reject = p_value < 0.05) %>%
+  {mean(.$reject)}
+
+4 * 4 * 3 * 5
 ```
@@
 
-### Shpaes of non-linearity
+#### Shpaes of non-linearity
 
@@
 
-### Reasons for making selections
+#### Reasons for making selections
 
@@
 
-## Heteroskedasticity
-
-```{r cache=TRUE}
-heter_result <- read_csv(here::here("data/processed/processed_surveys.csv"))
-
-heter_result <- heter_result %>%
-  filter(type != "cubic")
-
-heter_cache <- new.env()
-
-heter_result <- heter_result %>%
-  left_join(heter_result %>%
-  calc_p_value_multi(detected = "detect", n_sel = "num_selection", cache_env = heter_cache))
-
-heter_result <- heter_result %>%
-  mutate(weigthed_detect = ifelse(detect, detect/num_selection, 0)) %>%
+### Experiment II
+
+```{r eval = FALSE}
+heter %>%
+  filter(b < 3) %>%
   group_by(lineup_id) %>%
-  mutate(prop_detect = mean(weigthed_detect))
+  summarise(p_value = first(p_value), n_eval = n()) %>%
+  filter(n_eval > 3) %>%
+  mutate(reject = p_value < 0.05) %>%
+  {mean(.$reject)}
+
 ```
@@
 
-  visual_dat <- heter_result %>%
+  visual_dat <- heter %>%
   mutate(x_dist = ifelse(x_dist == "discrete_uniform", "even_discrete", x_dist)) %>%
@@
   summarise(across(c(type:p_value, prop_detect), first)) %>%
-  filter(ori_difficulty != "extremely_easy") %>%
   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%

### 2022-08-31 paper_comparison/paper_comparison/paper_comparison.tex
update the writing and code in main doc

@@
 
-% tightlist command for lists without linebreak
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+% see https://stackoverflow.com/a/47122900
+
+% Pandoc citation processing
 
@@
 \newcommand*{\Perm}[2]{{}^{#1}\!P_{#2}}%
-
 \usepackage{booktabs}
@@
 \emph{``Since all models are wrong the scientist must be alert to what
-is importantly wrong.''} \citep{REF}
+is importantly wrong \citep{box1976science}.''}
 \end{quote}
@@
 the model, and thus provide the capacity to identify what might be
-wrong. There are many ways that residuals are assessed.
-
-Residuals might be plotted, as a histogram or normal probability plot to
-examine the distribution. Is it symmetric, and unimodal? Then all is
-good. If the distribution is skewed or bimodal or multimodal, there are
-outliers, there is cause for concern. It if also possible to conduct a
-goodness of fit test, such as a Shapiro-Wilk Normality test \citep{REF}.
-
-It is also recommended to plot residuals against predicted values, and
-also each of the explanatory variables. If there is no obvious pattern,
-then all is good. It is a very difficult task for a human judge, though
-to make a decision that there's nothing there. It is especially common,
-particularly among new data analysts to report patterns when an
-experienced data analyst might quickly conclude that there are none.
-Generally, one looks for departures from nothingness like nonlinear
-dependency or heteroskedasticity. It is also possible to conduct a
-hypothesis test for non-linear dependence \citep{RESET}, and also use a
-Breusch-Pagan test \citep{REF} for heteroskedasticity.
+wrong. There are many ways that residuals could be assessed.
+
+Residuals might be plotted, as a histogram or quantile-quantile plot to
+examine the distribution. Using classical linear regression model as an
+example, if the distribution is symmetric and unimodal, then it is
+well-behaved. But if the distribution is skewed, bimodal, multimodal, or
+even contains outliers, there is cause for concern. It is also possible
+to conduct a goodness of fit test to examine the distribution of
+residuals, such as Shapiro-Wilk Normality test
+\citep{shapiro1965analysis}.
+
+To examine the relationship between residuals, predicted values and each
+of the explanatory variables, it is recommend to plot them on a scatter
+plot with residuals on the y-axis. This type of plot is usually called
+residual plot. In general, one will check whether there is any
+noticeable pattern to determine if the model is well-behaved. However,
+it is a very difficult task for a human judge, though to make a decision
+that there's nothing there. It is especially common, particularly among
+new data analysts to report patterns when an experienced data analyst
+might quickly conclude that there are none. Generally, one looks for
+departures from nothingness like nonlinear dependency or
+heteroskedasticity. It is also possible to conduct a hypothesis test for
+non-linear dependence \citep{ramsey_tests_1969}, and also use a
+Breusch-Pagan test \citep{breusch_simple_1979} for heteroskedasticity.
 
 There is an abundance of literature describing appropriate diagnostic
-methods for linear regression: Cook and Weisberg, \ldots{} \citep{REF}.
-The diligent reader of these sage writings will also notice sentences
-that express sentiments like \emph{based on their experience,
-statistical tests are not widely used in regression diagnostics. The
-same or even larger amount of information can be provided by diagnostic
-plots than the corresponding tests in most empirical studies.} There is
-a common guidance by experts that plots are the best for diagnosing
-model fits.
+methods for linear regression: \citet{draper1998applied},
+\citet{montgomery1982introduction}, \citet{belsley_regression_1980},
+\citet{cook_applied_1999} and \citet{cook1982residuals}. The diligent
+reader of these sage writings will also notice sentences that express
+sentiments like \emph{based on their experience, statistical tests are
+not widely used in regression diagnostics. The same or even larger
+amount of information can be provided by diagnostic plots than the
+corresponding tests in most empirical studies.} There is a common
+guidance by experts that plots are the best for diagnosing model fits.
 
@@
 process for reading residual plots, called visual inference, that can
-avoid the concerns about subjectiveness of human readers. Section XXX
-describes the experimental setup to enable a comparison between
-decisopns made by formal hypothesis testing, and how humans would read
-diagnostic plots. The results are reported in Section XXX. We finish
-with a discussion on future work, in particular how the responsibility
-for residual plot reading might be passed on to computer vision.
+avoid the concerns about subjectiveness of human readers. Section
+\ref{experimental-design} describes the experimental setup to enable a
+comparison between decision made by formal hypothesis testing, and how
+humans would read diagnostic plots. The results are reported in Section
+\ref{results}. We finish with a discussion on future work, in particular
+how the responsibility for residual plot reading might be passed on to
+computer vision.
 
@@
 
-\hypertarget{hypothesis-testing}{%
-\subsection{Hypothesis testing}\label{hypothesis-testing}}
-
-Other than checking diagnostic plots, analysts may perform formal
-hypothesis testing for detecting model defects. Depending on the
-alternative hypothesis that is focused on, a variety of tests can be
-applied. For example, the presence of heteroskedasticity can usually be
-tested by applying the White test
-\citep{white_heteroskedasticity-consistent_1980} or the Breusch-Pagan
-test \citep{breusch_simple_1979}, which are both derived from the
-Lagrange multiplier test \citep{silvey1959lagrangian} principle that
-relies on the asymptotic properties of the null distribution. For
-testing non-linearity, one may apply the F-test to examine the
-significance of specific polynomial and non-linear forms of the
-regressors, or the significance of proxy variables as in the Ramsey
-Regression Equation Specification Error Test (RESET)
-\citep{ramsey_tests_1969}.
-
-As discussed in \citet{cook1982residuals}, most residual-based tests for
-a particular type of departure from model assumptions are sensitive to
-other types of departures. It is likely the null hypothesis is correctly
-rejected but for the wrong reason, which is known as the ``Type III
-error''. Additionally, outliers will often incorrectly trigger the
-rejection of the null hypothesis despite the residuals are well-behaved
-\citep{cook_applied_1999}. This can be largely avoided in diagnostic
-plots as experienced analysts can evaluate the acceptability of
-assumptions flexibly, even in the presence of outliers.
-\citet{montgomery1982introduction} suggested that based on their
-experience, statistical tests are not widely used in regression
-diagnostics. The same or even larger amount of information can be
-provided by diagnostic plots than the corresponding tests in most
-empirical studies. Not to mention, it is almost impossible to have an
-exactly correctly specified model in reality. There is a well-known
-aphorism in statistics stated by George Box - ``All models are wrong,
-but some are useful''. This indicates proper hypothesis tests will
-always reject the null hypothesis as long as the sample size is large
-enough. The outcome ``Not reject'' can be interpreted as either ``effect
-size is small'' or ``sample size is small''. and the outcome ``reject''
-doesn't inform us whether and how much the model defects are of actual
-consequence to the inference and prediction. But still, the
-effectiveness of statistical tests shall not be disrespected.
-Statistical tests have a chance to provide analysts with unique
-information. In situations where no suitable diagnostic plots can be
-found for a particular violation of the assumptions, or excessive
-diagnostic plots are needed to be checked, one will have no choice but
-to fall back on statistical tests if there exists any. A good regression
-diagnostic practice should be a balanced combination of both methods.
-
 \hypertarget{visual-inference}{%
@@
 
+\hypertarget{heteroskedasticity}{%
+\subsection{Heteroskedasticity}\label{heteroskedasticity}}
+
+Heter Experiment design here
+
 \hypertarget{results}{%
@@
 In overall, there were a total of 3200 lineup evaluations made by 160
-subjects in experiment I, where 320 lineup evaluations were attention
-checks and were not used in the following analysis. The collated dataset
-is provided in \texttt{polynomials} of the \texttt{visage} \texttt{R}
-package.
+subjects in both experiment I and experiment II respectively, where 320
+lineup evaluations were attention checks and were not used in the
+following analysis.
+
+The collated dataset is provided in \texttt{polynomials} and
+\texttt{heter} of the \texttt{visage} \texttt{R} package.
 
@@
 \midrule
-She & Diploma and Bachelor Degree & 18-24 & No & 17\\
-He & Diploma and Bachelor Degree & 18-24 & Yes & 14\\
-She & Diploma and Bachelor Degree & 18-24 & Yes & 12\\
-He & Diploma and Bachelor Degree & 18-24 & No & 11\\
-He & Diploma and Bachelor Degree & 25-39 & No & 10\\
-She & Diploma and Bachelor Degree & 25-39 & No & 10\\
-She & Diploma and Bachelor Degree & 25-39 & Yes & 9\\
-He & High School or below & 18-24 & No & 8\\
-She & High School or below & 18-24 & No & 7\\
-He & High School or below & 18-24 & Yes & 6\\
-He & Diploma and Bachelor Degree & 25-39 & Yes & 5\\
-He & Masters Degree & 25-39 & No & 5\\
-He & Masters Degree & 25-39 & Yes & 5\\
-He & High School or below & 25-39 & No & 4\\
-She & High School or below & 25-39 & No & 4\\
-She & Masters Degree & 25-39 & Yes & 4\\
-He & High School or below & 25-39 & Yes & 3\\
-She & High School or below & 18-24 & Yes & 3\\
-She & Masters Degree & 25-39 & No & 3\\
+She & Diploma and Bachelor Degree & 18-24 & Yes & 26\\
+She & Diploma and Bachelor Degree & 18-24 & No & 24\\
+He & Diploma and Bachelor Degree & 18-24 & No & 22\\
+He & High School or below & 18-24 & No & 22\\
+She & Diploma and Bachelor Degree & 25-39 & No & 21\\
+He & Diploma and Bachelor Degree & 18-24 & Yes & 19\\
+He & Diploma and Bachelor Degree & 25-39 & No & 19\\
+She & High School or below & 18-24 & No & 19\\
+He & High School or below & 18-24 & Yes & 15\\
+She & Diploma and Bachelor Degree & 25-39 & Yes & 15\\
+He & Diploma and Bachelor Degree & 25-39 & Yes & 12\\
+She & High School or below & 18-24 & Yes & 10\\
+He & Masters Degree & 25-39 & No & 8\\
+He & Masters Degree & 25-39 & Yes & 8\\
+He & High School or below & 25-39 & No & 7\\
+She & Masters Degree & 25-39 & Yes & 7\\
+She & High School or below & 25-39 & No & 6\\
+He & High School or below & 25-39 & Yes & 5\\
+She & Honours Degree & 25-39 & No & 5\\
+She & Honours Degree & 25-39 & Yes & 4\\
+She & Masters Degree & 25-39 & No & 4\\
+He & Honours Degree & 25-39 & Yes & 3\\
+She & Honours Degree & 18-24 & No & 3\\
+He & Diploma and Bachelor Degree & 40-54 & Yes & 2\\
+He & Diploma and Bachelor Degree & 55-64 & No & 2\\
 He & Honours Degree & 25-39 & No & 2\\
+He & Masters Degree & 40-54 & No & 2\\
 She & Diploma and Bachelor Degree & 40-54 & No & 2\\
-She & Honours Degree & 18-24 & No & 2\\
+She & High School or below & 25-39 & Yes & 2\\
+She & Masters Degree & 18-24 & No & 2\\
+They & Diploma and Bachelor Degree & 18-24 & No & 2\\
+They & Diploma and Bachelor Degree & 25-39 & No & 2\\
+He & Diploma and Bachelor Degree & 40-54 & No & 1\\
 He & High School or below & 40-54 & Yes & 1\\
 He & High School or below & 55-64 & No & 1\\
+He & Honours Degree & 18-24 & No & 1\\
+He & Honours Degree & 18-24 & Yes & 1\\
+He & Honours Degree & 40-54 & Yes & 1\\
 He & Masters Degree & 40-54 & Yes & 1\\
 He & Masters Degree & 55-64 & No & 1\\
+Other & Diploma and Bachelor Degree & 18-24 & No & 1\\
 Other & High School or below & 25-39 & No & 1\\
+She & Diploma and Bachelor Degree & 40-54 & Yes & 1\\
 She & High School or below & 40-54 & No & 1\\
-She & Honours Degree & 25-39 & No & 1\\
-She & Honours Degree & 25-39 & Yes & 1\\
-She & Masters Degree & 18-24 & No & 1\\
+She & High School or below & 55-64 & Yes & 1\\
+She & Honours Degree & 40-54 & Yes & 1\\
 She & Masters Degree & 40-54 & No & 1\\
-They & Diploma and Bachelor Degree & 18-24 & No & 1\\
-They & Diploma and Bachelor Degree & 25-39 & No & 1\\
 They & High School or below & 18-24 & No & 1\\
+They & High School or below & 18-24 & Yes & 1\\
 They & High School or below & 25-39 & No & 1\\
@@
 \subsection{Power comparison}\label{power-comparison}}
+
+\hypertarget{experiment-i}{%
+\subsubsection{Experiment I}\label{experiment-i}}
 
@@
 \hypertarget{shpaes-of-non-linearity}{%
-\subsubsection{Shpaes of non-linearity}\label{shpaes-of-non-linearity}}
+\paragraph{Shpaes of non-linearity}\label{shpaes-of-non-linearity}}
 
@@
 \hypertarget{reasons-for-making-selections}{%
-\subsubsection{Reasons for making
+\paragraph{Reasons for making
 selections}\label{reasons-for-making-selections}}
@@
 
-\hypertarget{heteroskedasticity}{%
-\subsection{Heteroskedasticity}\label{heteroskedasticity}}
+\hypertarget{experiment-ii}{%
+\subsubsection{Experiment II}\label{experiment-ii}}
 

### 2022-08-31 paper_comparison/paper_comparison/paper_comparison.Rmd
clean up the paper and fill in the exp design

@@
 
-Heter Experiment design here
+Experiment II is designed to study the ability of human subjects to detect the appearance of a heteroskedasticity pattern under a simple linear regression model setting:
+
+\begin{align} \label{eq:heter-model}
+\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
+\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + 2 - |a| b (\boldsymbol{x} - a)^2 \boldsymbol{I}), \\
+\end{align}
+
+where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vector of size $n$.
+
+The null regression model used to fit the realizations generated by the above model is formulated exactly the same as the model used in the first experiment:
+
+\begin{equation} 
+\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
+\end{equation}
+
+where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.
+
+Model misspecification presents since the assumption about the constant error variance is violated.
+
+Experiment data were simulated using three different shapes ($a$ = -1, 0, 1), five different values of $b$ (b = 0.25, 1, 4, 16, 64), three different sample sizes (n = 50, 100, 300) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
+A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}.
+
+The values of $a$ was chosen so that different shapes of heteroskedasticity were included in the residual plot. These include triangle shape, butterfly shape and inverse triangle shape.
+
+```{r different-shape-of-heter}
+stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
+
+ori_dat <- heter_model(a = -1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300)
+
+VI_MODEL$plot(ori_dat) + 
+  ggtitle(quote("Triangle shape")) + 
+  theme_light() -> p1
+
+VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote("Butterfly shape")) + 
+  theme_light() -> p2
+
+VI_MODEL$plot(heter_model(a = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote("Inverse triangle shape")) + 
+  theme_light() -> p3
+
+patchwork::wrap_plots(p1, p2, p3, ncol = 3)
+```
+
+The range of $b$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
 
@@
 
-# Old
-
-## Overview of the Data
-
-We collected 400 lineup evaluations made by 20 participants in experiment I and 880 lineup evaluations made by 44 participants in experiment II.
+<!-- # Old -->
+
+<!-- ## Overview of the Data -->
+
+<!-- We collected 400 lineup evaluations made by 20 participants in experiment I and 880 lineup evaluations made by 44 participants in experiment II.
 In total, 442 unique lineups were evaluated by 64 subjects.
@@
 Therefore, the submission was kept.
-
-## Power comparision
-
-1. power (visual test vs. conventional test)
-(visual test most different one (everything test, any departure))
-plot figure in a paper, desc, exp
-2. investigate the difference (gap), give examples
-3. conventional is too sensitive
-4. make conventional less sensitive (vary alpha)
-
-To model the power of visual test, 10 logistic regression were fit for different number of evaluations ranged from one to five and two different types of simulation setting.
+-->
+
+<!-- ## Power comparision -->
+
+<!-- 1. power (visual test vs. conventional test) -->
+<!-- (visual test most different one (everything test, any departure)) -->
+<!-- plot figure in a paper, desc, exp -->
+<!-- 2. investigate the difference (gap), give examples -->
+<!-- 3. conventional is too sensitive -->
+<!-- 4. make conventional less sensitive (vary alpha) -->
+
+<!-- To model the power of visual test, 10 logistic regression were fit for different number of evaluations ranged from one to five and two different types of simulation setting.
 All 10 models used natural logarithm of the effect size as the only fixed effect, and whether the test successfully rejects the null hypothesis as the response variable.
@@
 The modelling result summarized in \ref{tab:powerglmcubic} and \ref{tab:powerglmheter} aligned with the expectation as the coefficients of natural logarithm of the effect size are positive and significant across all 10 models.
-
-Figure \ref{fig:power-com} illustrates the fitted models, while providing the local constant estimate of the power of F-test and Breusch–Pagan test for comparison.
+-->
+
+<!-- Figure \ref{fig:power-com} illustrates the fitted models, while providing the local constant estimate of the power of F-test and Breusch–Pagan test for comparison.
 Data for the conventional test is simulated under the model setting described in section ...
@@
 From Figure \ref{fig:power-com}, it can be observed that the fitted power of visual test increased as the number of evaluations increased for both cubic and heteroskedasticity model.
-
-For heteroskedasticity model, this phenomenon was more obvious as the power of visual tests with evaluations greater than two were always greater than those with evaluations smaller than two.
-
-For cubic model, the separation between curves was small.
+-->
+
+<!-- For heteroskedasticity model, this phenomenon was more obvious as the power of visual tests with evaluations greater than two were always greater than those with evaluations smaller than two.
+-->
+
+<!-- For cubic model, the separation between curves was small.
 The estimated power of visual tests with three to five evaluations were almost identical to each other in regards of effect size.
 In addition, all five curves peaked at one as effect size increased, suggesting that identification of non-linearity as a visual task can be completed reliably by human as long as the departure from null hypothesis is large enough.
-
-As shown in Figure \ref{fig:power-com}, both F-test and Breusch–Pagan test generally possessed greater power than visual test.
+-->
+
+<!-- As shown in Figure \ref{fig:power-com}, both F-test and Breusch–Pagan test generally possessed greater power than visual test.
 A visual tests is a collection of test against any alternatives that would create visual discoverable features, while a conventional test is usually targeting at a pre-specified alternative.
 Considering the data generating process of the model defect was known and controlled in this research, where all other alternatives have been eliminated except the one we concerned, the result was suggested that conventional tests were more sensitive to violations of linearity and homoscedasticity assumption than visual tests.
-
-It was also found that there was a noticeable gap between curves of the conventional test and the visual test at around $log(\text{effect size}) = 0$ for the cubic model and $log(\text{effect size}) = 2.5$ for the heteroskedasticity model, where the differences in power were greater than 0.6.
+-->
+
+<!-- It was also found that there was a noticeable gap between curves of the conventional test and the visual test at around $log(\text{effect size}) = 0$ for the cubic model and $log(\text{effect size}) = 2.5$ for the heteroskedasticity model, where the differences in power were greater than 0.6.
 We further analysed the lineups with correspoding effect sizes.
@@
 The visual difference between the true data plot and null plots were almost unnoticeable.
-
-<!-- sample 1, change values of a and b, plot the data  -->
-
-## Effect of parameters on power of the visual test
-
-The previous section focuses on the change of effect size relative to the power of the visual test.
+-->
+
+<!-- <!-- sample 1, change values of a and b, plot the data  --> -->
+
+<!-- ## Effect of parameters on power of the visual test -->
+
+<!-- The previous section focuses on the change of effect size relative to the power of the visual test.
 However, effect size is only a one dimensional summarisation of parameters used in data simulation.
 Individual factor embedded in the simulation process should also be analysed.
-
-In cubic model, two major factors that influencing the strength of the signal are $a$ and $b$.
+-->
+
+<!-- In cubic model, two major factors that influencing the strength of the signal are $a$ and $b$.
 Figure \ref{fig:power-com-cubic-a} and \ref{fig:power-com-cubic-a} illustrates 30 different logistic regressions fit for different number of evaluations and different number of observations $n$.
@@
 From the figures, we can observe ...
-
-```{r power-vs-log-effect-size}
-polynomials %>%
-  group_by(lineup_id) %>%
-  summarise(across(c(type:p_value), first)) %>%
-  filter(lineup_id < 577) %>%
-  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
-  ggplot() +
-  geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  xlab("Natural logarithm of effect size") +
-  ylab("Power")
-```
-
-```{r power-vs-log-effect-size-given-x-dist}
-polynomials %>%
-  group_by(lineup_id) %>%
-  summarise(across(c(type:p_value), first)) %>%
-  filter(lineup_id < 577) %>%
-  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
-  ggplot() +
-  geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  facet_wrap(~x_dist) +
-  xlab("Natural logarithm of effect size") +
-  ylab("Power") +
-  ggtitle("Power comparison conditional on the distribution of X")
-```
-
-```{r power-of-visual-test-given-x-dist}
-polynomials %>%
-  group_by(lineup_id) %>%
-  summarise(across(c(type:p_value), first)) %>%
-  filter(lineup_id < 577) %>%
-  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
-  ggplot() +
-  geom_smooth(aes(log(effect_size), reject, col = x_dist), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  xlab("Natural logarithm of effect size") +
-  ylab("Power") +
-  ggtitle("Power of visual test conditional on the distribution of X")
-```
-
-```{r power-vs-log-effect-size-given-shape}
-polynomials %>%
-  group_by(lineup_id) %>%
-  summarise(across(c(type:p_value), first)) %>%
-  filter(lineup_id < 577) %>%
-  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
-  ggplot() +
-  geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  facet_wrap(~shape) +
-  xlab("Natural logarithm of effect size") +
-  ylab("Power") +
-  ggtitle("Power comparison conditional on the shape of the Hermite polynomials")
-```
-
-```{r power-of-visual-test-given-shape}
-polynomials %>%
-  group_by(lineup_id) %>%
-  summarise(across(c(type:p_value), first)) %>%
-  filter(lineup_id < 577) %>%
-  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01)), shape = factor(shape)) %>%
-  ggplot() +
-  geom_smooth(aes(log(effect_size), reject, col = shape), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  xlab("Natural logarithm of effect size") +
-  ylab("Power") +
-  ggtitle("Power of visual test conditional on the shape of the Hermite polynomials")
-```
-
-```{r power-vs-log-effect-size-given-number-of-observations}
-polynomials %>%
-  group_by(lineup_id) %>%
-  summarise(across(c(type:p_value), first)) %>%
-  filter(lineup_id < 577) %>%
-  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
-  ggplot() +
-  geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  facet_wrap(~n) +
-  xlab("Natural logarithm of effect size") +
-  ylab("Power") +
-  ggtitle("Power comparison conditional on the number of observations")
-```
-
-```{r power-of-visual-test-given-number-of-observations}
-polynomials %>%
-  group_by(lineup_id) %>%
-  summarise(across(c(type:p_value), first)) %>%
-  filter(lineup_id < 577) %>%
-  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01)), n = factor(n)) %>%
-  ggplot() +
-  geom_smooth(aes(log(effect_size), reject, col = n), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  xlab("Natural logarithm of effect size") +
-  ylab("Power") +
-  ggtitle("Power of visual test conditional on the number of observations")
-```
-
-```{r}
-polynomials %>%
-  group_by(lineup_id) %>%
-  summarise(across(c(type:p_value), first)) %>%
-  filter(lineup_id < 577) %>%
-  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
-  ggplot() +
-  geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.05)), aes(log(effect_size), reject, col = "Conventional test:0.05"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.01)), aes(log(effect_size), reject, col = "Conventional test:0.01"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.005)), aes(log(effect_size), reject, col = "Conventional test:0.005"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.0005)), aes(log(effect_size), reject, col = "Conventional test:0.0005"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.0000005)), aes(log(effect_size), reject, col = "Conventional test:0.0000005"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  xlab("Natural logarithm of effect size") +
-  ylab("Power")
-```
+-->
+
+<!-- ```{r power-vs-log-effect-size} -->
+<!-- polynomials %>% -->
+<!--   group_by(lineup_id) %>% -->
+<!--   summarise(across(c(type:p_value), first)) %>% -->
+<!--   filter(lineup_id < 577) %>% -->
+<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
+<!--   ggplot() + -->
+<!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+<!--   geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+<!--   xlab("Natural logarithm of effect size") + -->
+<!--   ylab("Power") -->
+<!-- ``` -->
+
+<!-- ```{r power-vs-log-effect-size-given-x-dist} -->
+<!-- polynomials %>% -->
+<!--   group_by(lineup_id) %>% -->
+<!--   summarise(across(c(type:p_value), first)) %>% -->
+<!--   filter(lineup_id < 577) %>% -->
+<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
+<!--   ggplot() + -->
+<!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+<!--   geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+<!--   facet_wrap(~x_dist) + -->
+<!--   xlab("Natural logarithm of effect size") + -->
+<!--   ylab("Power") + -->
+<!--   ggtitle("Power comparison conditional on the distribution of X") -->
+<!-- ``` -->
+
+<!-- ```{r power-of-visual-test-given-x-dist} -->
+<!-- polynomials %>% -->
+<!--   group_by(lineup_id) %>% -->
+<!--   summarise(across(c(type:p_value), first)) %>% -->
+<!--   filter(lineup_id < 577) %>% -->
+<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
+<!--   ggplot() + -->
+<!--   geom_smooth(aes(log(effect_size), reject, col = x_dist), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+<!--   xlab("Natural logarithm of effect size") + -->
+<!--   ylab("Power") + -->
+<!--   ggtitle("Power of visual test conditional on the distribution of X") -->
+<!-- ``` -->
+
+<!-- ```{r power-vs-log-effect-size-given-shape} -->
+<!-- polynomials %>% -->
+<!--   group_by(lineup_id) %>% -->
+<!--   summarise(across(c(type:p_value), first)) %>% -->
+<!--   filter(lineup_id < 577) %>% -->
+<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
+<!--   ggplot() + -->
+<!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+<!--   geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+<!--   facet_wrap(~shape) + -->
+<!--   xlab("Natural logarithm of effect size") + -->
+<!--   ylab("Power") + -->
+<!--   ggtitle("Power comparison conditional on the shape of the Hermite polynomials") -->
+<!-- ``` -->
+
+<!-- ```{r power-of-visual-test-given-shape} -->
+<!-- polynomials %>% -->
+<!--   group_by(lineup_id) %>% -->
+<!--   summarise(across(c(type:p_value), first)) %>% -->
+<!--   filter(lineup_id < 577) %>% -->
+<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01)), shape = factor(shape)) %>% -->
+<!--   ggplot() + -->
+<!--   geom_smooth(aes(log(effect_size), reject, col = shape), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+<!--   xlab("Natural logarithm of effect size") + -->
+<!--   ylab("Power") + -->
+<!--   ggtitle("Power of visual test conditional on the shape of the Hermite polynomials") -->
+<!-- ``` -->
+
+<!-- ```{r power-vs-log-effect-size-given-number-of-observations} -->
+<!-- polynomials %>% -->
+<!--   group_by(lineup_id) %>% -->
+<!--   summarise(across(c(type:p_value), first)) %>% -->
+<!--   filter(lineup_id < 577) %>% -->
+<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
+<!--   ggplot() + -->
+<!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+<!--   geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+<!--   facet_wrap(~n) + -->
+<!--   xlab("Natural logarithm of effect size") + -->
+<!--   ylab("Power") + -->
+<!--   ggtitle("Power comparison conditional on the number of observations") -->
+<!-- ``` -->
+
+<!-- ```{r power-of-visual-test-given-number-of-observations} -->
+<!-- polynomials %>% -->
+<!--   group_by(lineup_id) %>% -->
+<!--   summarise(across(c(type:p_value), first)) %>% -->
+<!--   filter(lineup_id < 577) %>% -->
+<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01)), n = factor(n)) %>% -->
+<!--   ggplot() + -->
+<!--   geom_smooth(aes(log(effect_size), reject, col = n), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+<!--   xlab("Natural logarithm of effect size") + -->
+<!--   ylab("Power") + -->
+<!--   ggtitle("Power of visual test conditional on the number of observations") -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- polynomials %>% -->
+<!--   group_by(lineup_id) %>% -->
+<!--   summarise(across(c(type:p_value), first)) %>% -->
+<!--   filter(lineup_id < 577) %>% -->
+<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
+<!--   ggplot() + -->
+<!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+<!--   geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.05)), aes(log(effect_size), reject, col = "Conventional test:0.05"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+<!--   geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.01)), aes(log(effect_size), reject, col = "Conventional test:0.01"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+<!--   geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.005)), aes(log(effect_size), reject, col = "Conventional test:0.005"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+<!--   geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.0005)), aes(log(effect_size), reject, col = "Conventional test:0.0005"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+<!--   geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.0000005)), aes(log(effect_size), reject, col = "Conventional test:0.0000005"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+<!--   xlab("Natural logarithm of effect size") + -->
+<!--   ylab("Power") -->
+<!-- ``` -->

### 2022-08-31 paper_comparison/paper_comparison/paper_comparison.tex
clean up the paper and fill in the exp design

@@
 
-Heter Experiment design here
+Experiment II is designed to study the ability of human subjects to
+detect the appearance of a heteroskedasticity pattern under a simple
+linear regression model setting:
+
+\begin{align} \label{eq:heter-model}
+\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
+\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + 2 - |a| b (\boldsymbol{x} - a)^2 \boldsymbol{I}), \\
+\end{align}
+
+where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
+\(\boldsymbol{\varepsilon}\) are vector of size \(n\).
+
+The null regression model used to fit the realizations generated by the
+above model is formulated exactly the same as the model used in the
+first experiment:
+
+\begin{equation} 
+\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
+\end{equation}
+
+where
+\(\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\).
+
+Model misspecification presents since the assumption about the constant
+error variance is violated.
+
+Experiment data were simulated using three different shapes (\(a\) = -1,
+0, 1), five different values of \(b\) (b = 0.25, 1, 4, 16, 64), three
+different sample sizes (n = 50, 100, 300) and four different
+distribution of \(X_{raw}\): (1) \(U(-1, 1)\), (2) \(N(0, 0.3^2)\), (3)
+\(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). A summary of the
+parameters used in this experiment is given in Table
+\ref{tab:parameter-table}.
+
+The values of \(a\) was chosen so that different shapes of
+heteroskedasticity were included in the residual plot. These include
+triangle shape, butterfly shape and inverse triangle shape.
+
+\includegraphics{paper_comparison_files/figure-latex/different-shape-of-heter-1.pdf}
+
+The range of \(b\), which is a factor controlling the strength of the
+signal, was chosen so that different difficulty levels of lineups were
+generated, and therefore, the estimated power curve would be smooth and
+continuous.
 
@@
 
-\hypertarget{old}{%
-\section{Old}\label{old}}
-
-\hypertarget{overview-of-the-data}{%
-\subsection{Overview of the Data}\label{overview-of-the-data}}
-
-We collected 400 lineup evaluations made by 20 participants in
-experiment I and 880 lineup evaluations made by 44 participants in
-experiment II. In total, 442 unique lineups were evaluated by 64
-subjects. In experiment I, one of the participants skipped all 20
-lineups. Hence, the submission was rejected and removed from the
-dataset. In experiment II, there was a participant failed one of the two
-attention checks, but there was no further evidence of low-effort
-throughout the experiment. Therefore, the submission was kept.
-
-\hypertarget{power-comparision}{%
-\subsection{Power comparision}\label{power-comparision}}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  power (visual test vs.~conventional test) (visual test most different
-  one (everything test, any departure)) plot figure in a paper, desc,
-  exp
-\item
-  investigate the difference (gap), give examples
-\item
-  conventional is too sensitive
-\item
-  make conventional less sensitive (vary alpha)
-\end{enumerate}
-
-To model the power of visual test, 10 logistic regression were fit for
-different number of evaluations ranged from one to five and two
-different types of simulation setting. All 10 models used natural
-logarithm of the effect size as the only fixed effect, and whether the
-test successfully rejects the null hypothesis as the response variable.
-Given the way we define the effect size, it was expected that with
-larger effect size, both conventional test and visual test will have
-higher probability in rejecting the null hypothesis when it is not true.
-The modelling result summarized in \ref{tab:powerglmcubic} and
-\ref{tab:powerglmheter} aligned with the expectation as the coefficients
-of natural logarithm of the effect size are positive and significant
-across all 10 models.
-
-Figure \ref{fig:power-com} illustrates the fitted models, while
-providing the local constant estimate of the power of F-test and
-Breusch--Pagan test for comparison. Data for the conventional test is
-simulated under the model setting described in section \ldots{} and
-5000000 samples are drawn for both cubic and heteroskedasticity model.
-From Figure \ref{fig:power-com}, it can be observed that the fitted
-power of visual test increased as the number of evaluations increased
-for both cubic and heteroskedasticity model.
-
-For heteroskedasticity model, this phenomenon was more obvious as the
-power of visual tests with evaluations greater than two were always
-greater than those with evaluations smaller than two.
-
-For cubic model, the separation between curves was small. The estimated
-power of visual tests with three to five evaluations were almost
-identical to each other in regards of effect size. In addition, all five
-curves peaked at one as effect size increased, suggesting that
-identification of non-linearity as a visual task can be completed
-reliably by human as long as the departure from null hypothesis is large
-enough.
-
-As shown in Figure \ref{fig:power-com}, both F-test and Breusch--Pagan
-test generally possessed greater power than visual test. A visual tests
-is a collection of test against any alternatives that would create
-visual discoverable features, while a conventional test is usually
-targeting at a pre-specified alternative. Considering the data
-generating process of the model defect was known and controlled in this
-research, where all other alternatives have been eliminated except the
-one we concerned, the result was suggested that conventional tests were
-more sensitive to violations of linearity and homoscedasticity
-assumption than visual tests.
-
-It was also found that there was a noticeable gap between curves of the
-conventional test and the visual test at around
-\(log(\text{effect size}) = 0\) for the cubic model and
-\(log(\text{effect size}) = 2.5\) for the heteroskedasticity model,
-where the differences in power were greater than 0.6. We further
-analysed the lineups with correspoding effect sizes. Figure
-\ref{fig:cubic-hard} and \ref{fig:heter-hard} showed that human was
-indeed hard to identify the patterns at this level of difficulty. The
-visual difference between the true data plot and null plots were almost
-unnoticeable.
-
-\hypertarget{effect-of-parameters-on-power-of-the-visual-test}{%
-\subsection{Effect of parameters on power of the visual
-test}\label{effect-of-parameters-on-power-of-the-visual-test}}
-
-The previous section focuses on the change of effect size relative to
-the power of the visual test. However, effect size is only a one
-dimensional summarisation of parameters used in data simulation.
-Individual factor embedded in the simulation process should also be
-analysed.
-
-In cubic model, two major factors that influencing the strength of the
-signal are \(a\) and \(b\). Figure \ref{fig:power-com-cubic-a} and
-\ref{fig:power-com-cubic-a} illustrates 30 different logistic
-regressions fit for different number of evaluations and different number
-of observations \(n\). The regressor used in these models was
-\(|a|/\sigma\) since the noise level \(\sigma\) needed to be taken into
-account. From the figures, we can observe \ldots{}
-
-\includegraphics{paper_comparison_files/figure-latex/power-vs-log-effect-size-1.pdf}
-
-\includegraphics{paper_comparison_files/figure-latex/power-vs-log-effect-size-given-x-dist-1.pdf}
-
-\includegraphics{paper_comparison_files/figure-latex/power-of-visual-test-given-x-dist-1.pdf}
-
-\includegraphics{paper_comparison_files/figure-latex/power-vs-log-effect-size-given-shape-1.pdf}
-
-\includegraphics{paper_comparison_files/figure-latex/power-of-visual-test-given-shape-1.pdf}
-
-\includegraphics{paper_comparison_files/figure-latex/power-vs-log-effect-size-given-number-of-observations-1.pdf}
-
-\includegraphics{paper_comparison_files/figure-latex/power-of-visual-test-given-number-of-observations-1.pdf}
-
-\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-19-1.pdf}
+--\textgreater{}
 

### 2022-08-31 paper_comparison/paper_comparison/paper_comparison.Rmd
paper structure

@@
 library(tidyverse)
+# remotes::install_github("TengMCing/visage")
 library(visage)
@@
 
-> *"Since all models are wrong the scientist must be alert to what is importantly wrong [@box1976science]."*
+> *"Since all models are wrong the scientist must be alert to what is importantly wrong."*  [@box1976science]
 
@@
 
-## Diagnostic plots
+## Departures from good residual plots
 
@@
 
-## Visual inference
-
-However, unlike hypothesis testing built upon rigorous statistical procedures, reading diagnostic plots relies on graphical perception - human’s ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984], which is to some extent subjective and indecisive.
+## Conventionally testing for departures
+
+Explain the tests you are showing later
+
+- RESET
+- BP
+- SW
+
+and show the results for the residual plots displayed in previous section
+
+Briefly mention any others
+
+## Visual testing for departures
+
+- Very briefly explain what a lineup test is
+- Lineup (maybe using your previous residual plot for nonlinearity)
+- $p$-value calulcation, briefly, including how multiple selections are handled
+- power calculation, briefly, including how multiple selections are handled
+
+However, unlike hypothesis testing built upon rigorous statistical procedures, reading diagnostic plots relies on graphical perception - human's ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984], which is to some extent subjective and indecisive.
 Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
@@
 # Experimental design
+
+## Motivation and overview
+
+## Simulating departures
+
+### Nonlinearity
+
+### Heteroskedasticity
+
+## Experimental setup
+
+### Factors
+
+- Treatments for nonlinearity experiment (plots summarising levels)
+- Treatments for heteroskedasticity experiment (plots summarising levels)
+
+### Assigning subjects
+
+### Collecting results
+
+software/technical
+
+## Computing power curves
 
@@
 
-# Results
-
-## Data preprocessing
+# Analysis and results
+
+## Data overview
+
+How many people? How many lineups?
 
@@
 The collated dataset is provided in `polynomials` and `heter` of the `visage` `R` package.
+
+##
+
+## Nonlinearity
+
+- Power curve overall
+- Four lineups also shown, selected from ones close to the visual power curve from the uniform treatment
+- Compare visual power against different factors
+
+## Heteroskedasticity
+
+- Power curve overall
+- Four lineups also shown, selected from ones close to the visual power curve from the uniform treatment
+- Compare visual power against different factors
 

### 2022-08-31 paper_comparison/paper_comparison/paper_comparison.tex
paper structure

@@
 \emph{``Since all models are wrong the scientist must be alert to what
-is importantly wrong \citep{box1976science}.''}
+is importantly wrong.''} \citep{box1976science}
 \end{quote}
@@
 
-\hypertarget{diagnostic-plots}{%
-\subsection{Diagnostic plots}\label{diagnostic-plots}}
+\hypertarget{departures-from-good-residual-plots}{%
+\subsection{Departures from good residual
+plots}\label{departures-from-good-residual-plots}}
 
@@
 
-\hypertarget{visual-inference}{%
-\subsection{Visual inference}\label{visual-inference}}
+\hypertarget{conventionally-testing-for-departures}{%
+\subsection{Conventionally testing for
+departures}\label{conventionally-testing-for-departures}}
+
+Explain the tests you are showing later
+
+\begin{itemize}
+\tightlist
+\item
+  RESET
+\item
+  BP
+\item
+  SW
+\end{itemize}
+
+and show the results for the residual plots displayed in previous
+section
+
+Briefly mention any others
+
+\hypertarget{visual-testing-for-departures}{%
+\subsection{Visual testing for
+departures}\label{visual-testing-for-departures}}
+
+\begin{itemize}
+\tightlist
+\item
+  Very briefly explain what a lineup test is
+\item
+  Lineup (maybe using your previous residual plot for nonlinearity)
+\item
+  \(p\)-value calulcation, briefly, including how multiple selections
+  are handled
+\item
+  power calculation, briefly, including how multiple selections are
+  handled
+\end{itemize}
 
@@
 
+\hypertarget{motivation-and-overview}{%
+\subsection{Motivation and overview}\label{motivation-and-overview}}
+
+\hypertarget{simulating-departures}{%
+\subsection{Simulating departures}\label{simulating-departures}}
+
+\hypertarget{nonlinearity}{%
+\subsubsection{Nonlinearity}\label{nonlinearity}}
+
+\hypertarget{heteroskedasticity}{%
+\subsubsection{Heteroskedasticity}\label{heteroskedasticity}}
+
+\hypertarget{experimental-setup}{%
+\subsection{Experimental setup}\label{experimental-setup}}
+
+\hypertarget{factors}{%
+\subsubsection{Factors}\label{factors}}
+
+\begin{itemize}
+\tightlist
+\item
+  Treatments for nonlinearity experiment (plots summarising levels)
+\item
+  Treatments for heteroskedasticity experiment (plots summarising
+  levels)
+\end{itemize}
+
+\hypertarget{assigning-subjects}{%
+\subsubsection{Assigning subjects}\label{assigning-subjects}}
+
+\hypertarget{collecting-results}{%
+\subsubsection{Collecting results}\label{collecting-results}}
+
+software/technical
+
+\hypertarget{computing-power-curves}{%
+\subsection{Computing power curves}\label{computing-power-curves}}
+
 For the purpose of examining the effectiveness of visual test in
@@
 
-\hypertarget{heteroskedasticity}{%
-\subsection{Heteroskedasticity}\label{heteroskedasticity}}
+\hypertarget{heteroskedasticity-1}{%
+\subsection{Heteroskedasticity}\label{heteroskedasticity-1}}
 
@@
 
-\hypertarget{results}{%
-\section{Results}\label{results}}
+\hypertarget{analysis-and-results}{%
+\section{Analysis and results}\label{analysis-and-results}}
 
-\hypertarget{data-preprocessing}{%
-\subsection{Data preprocessing}\label{data-preprocessing}}
+\hypertarget{data-overview}{%
+\subsection{Data overview}\label{data-overview}}
+
+How many people? How many lineups?
 
@@
 \texttt{heter} of the \texttt{visage} \texttt{R} package.
+
+\hypertarget{section}{%
+\subsection{}\label{section}}
+
+\hypertarget{nonlinearity-1}{%
+\subsection{Nonlinearity}\label{nonlinearity-1}}
+
+\begin{itemize}
+\tightlist
+\item
+  Power curve overall
+\item
+  Four lineups also shown, selected from ones close to the visual power
+  curve from the uniform treatment
+\item
+  Compare visual power against different factors
+\end{itemize}
+
+\hypertarget{heteroskedasticity-2}{%
+\subsection{Heteroskedasticity}\label{heteroskedasticity-2}}
+
+\begin{itemize}
+\tightlist
+\item
+  Power curve overall
+\item
+  Four lineups also shown, selected from ones close to the visual power
+  curve from the uniform treatment
+\item
+  Compare visual power against different factors
+\end{itemize}
 

### 2022-09-01 paper_comparison/paper_comparison/paper_comparison.Rmd
clean up code

@@
 
-```{r get-third-result}
+```{r get-lineup-data}
 if (!file.exists(here::here("data/third_study/polynomials_lineup.rds"))) {
@@
   polynomials_lineup <- readRDS(here::here("data/third_study/polynomials_lineup.rds"))
+}
+
+if (!file.exists(here::here("data/fourth_study/heter_lineup.rds"))) {
+  heter_lineup <- get_polynomials_lineup()
+  saveRDS(heter_lineup, here::here("data/fourth_study/heter_lineup.rds"))
+} else {
+  heter_lineup <- readRDS(here::here("data/fourth_study/heter_lineup.rds"))
 }
@@
 Residuals might be plotted, as a histogram or quantile-quantile plot to examine the distribution.
-Using classical linear regression model as an example, if the distribution is symmetric and unimodal, then it is well-behaved.
+Using the classical linear regression model as an example, if the distribution is symmetric and unimodal, it is well-behaved.
 But if the distribution is skewed, bimodal, multimodal, or even contains outliers, there is cause for concern.
-It is also possible to conduct a goodness of fit test to examine the distribution of residuals, such as Shapiro-Wilk Normality test [@shapiro1965analysis].
-
-To examine the relationship between residuals, predicted values and each of the explanatory variables, it is recommend to plot them on a scatter plot with residuals on the y-axis.
-This type of plot is usually called residual plot.
+Conducting a goodness of fit test is an alternative method for inspecting the distribution, such as the Shapiro-Wilk Normality test [@shapiro1965analysis].
+
+To scrutinize the relationship between residuals, predicted values and each of the explanatory variables, it is recommend to plot them on a scatter plot.
 In general, one will check whether there is any noticeable pattern to determine if the model is well-behaved.
@@
 
+<!-- todo: check the two way interaction for x_dist and a -->
+
 ```{r eval = FALSE}
@@
   facet_wrap(~a)
+```
+
+```{r}
+heter_conv_test_plot() %>%
+  add_heter_SW_plot_layers() %>%
+  add_heter_visual_test_plot_layers() +
+  facet_grid(x_dist~a)
 ```

### 2022-09-01 paper_comparison/paper_comparison/paper_comparison.Rmd
update background

@@
 To scrutinize the relationship between residuals, predicted values and each of the explanatory variables, it is recommend to plot them on a scatter plot.
-In general, one will check whether there is any noticeable pattern to determine if the model is well-behaved.
+In general, the model is well-behaved if there is any visual discoverable pattern.
 However, it is a very difficult task for a human judge, though to make a decision that there's nothing there.
 It is especially common, particularly among new data analysts to report patterns when an experienced data analyst might quickly conclude that there are none.
-Generally, one looks for departures from nothingness like nonlinear dependency or heteroskedasticity.
-It is also possible to conduct a hypothesis test for non-linear dependence [@ramsey_tests_1969], and also use a Breusch-Pagan test [@breusch_simple_1979] for heteroskedasticity.
+Generally, one looks for departures from nothingness like non-linear dependency or heteroskedasticity.
+It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and also use a Breusch-Pagan test [@breusch_simple_1979] for heteroskedasticity.
 
@@
 
+```{r}
+mod <- poly_model(include_z = FALSE, sigma = 0.25)
+dat <- mod$gen(300)
+patchwork::wrap_plots(mod$plot(dat, theme = theme_light()) + ggtitle("Good residuals"),
+                      mod$plot(mod$set_prm("include_z", TRUE)$gen(300, computed = select(dat, x, e)), theme = theme_light()) + ggtitle("Non-linearity"),
+                      mod$plot(heter_model(b = 64)$gen(300, computed = select(dat, x, e)), theme = theme_light()) + ggtitle("Heteroskedasticity"),
+                      mod$plot(heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$gen(300, computed = select(dat, x)), theme = theme_light()) + ggtitle("Non-normality"))
+```
+
 Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals.
-As suggested by @cook1982residuals, these kinds of diagnostic plots are commonly used to identify patterns that are indicative of nonconstant error variance or non-linearity.
+As suggested by @cook1982residuals, these kinds of diagnostic plots are commonly used to identify patterns that are indicative of non-linearity or nonconstant error variance.
+
 Raw residuals and studentized residuals are the two most frequently used residuals in standard residual plots.

### 2022-09-01 paper_comparison/paper_comparison/paper_comparison.tex
update background

@@
 Residuals might be plotted, as a histogram or quantile-quantile plot to
-examine the distribution. Using classical linear regression model as an
-example, if the distribution is symmetric and unimodal, then it is
+examine the distribution. Using the classical linear regression model as
+an example, if the distribution is symmetric and unimodal, it is
 well-behaved. But if the distribution is skewed, bimodal, multimodal, or
-even contains outliers, there is cause for concern. It is also possible
-to conduct a goodness of fit test to examine the distribution of
-residuals, such as Shapiro-Wilk Normality test
+even contains outliers, there is cause for concern. Conducting a
+goodness of fit test is an alternative method for inspecting the
+distribution, such as the Shapiro-Wilk Normality test
 \citep{shapiro1965analysis}.
 
-To examine the relationship between residuals, predicted values and each
-of the explanatory variables, it is recommend to plot them on a scatter
-plot with residuals on the y-axis. This type of plot is usually called
-residual plot. In general, one will check whether there is any
-noticeable pattern to determine if the model is well-behaved. However,
-it is a very difficult task for a human judge, though to make a decision
-that there's nothing there. It is especially common, particularly among
-new data analysts to report patterns when an experienced data analyst
-might quickly conclude that there are none. Generally, one looks for
-departures from nothingness like nonlinear dependency or
-heteroskedasticity. It is also possible to conduct a hypothesis test for
-non-linear dependence \citep{ramsey_tests_1969}, and also use a
-Breusch-Pagan test \citep{breusch_simple_1979} for heteroskedasticity.
+To scrutinize the relationship between residuals, predicted values and
+each of the explanatory variables, it is recommend to plot them on a
+scatter plot. In general, the model is well-behaved if there is any
+visual discoverable pattern. However, it is a very difficult task for a
+human judge, though to make a decision that there's nothing there. It is
+especially common, particularly among new data analysts to report
+patterns when an experienced data analyst might quickly conclude that
+there are none. Generally, one looks for departures from nothingness
+like non-linear dependency or heteroskedasticity. It is also possible to
+conduct hypothesis tests for non-linear dependence
+\citep{ramsey_tests_1969}, and also use a Breusch-Pagan test
+\citep{breusch_simple_1979} for heteroskedasticity.
 
@@
 
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-6-1.pdf}
+
 Graphical summaries in which residuals are plotted against fitted values
@@
 these kinds of diagnostic plots are commonly used to identify patterns
-that are indicative of nonconstant error variance or non-linearity. Raw
-residuals and studentized residuals are the two most frequently used
+that are indicative of non-linearity or nonconstant error variance.
+
+Raw residuals and studentized residuals are the two most frequently used
 residuals in standard residual plots. The debt on which type of
@@
 
-\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-12-1.pdf}
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-13-1.pdf}
 
@@
 
-\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-15-1.pdf}
-
 \includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-16-1.pdf}
@@
 \includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-18-1.pdf}
+
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-19-1.pdf}
+
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-20-1.pdf}
 

### 2022-09-04 paper_comparison/paper_comparison/paper_comparison.Rmd
update residual departure

@@
 Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals.
-As suggested by @cook1982residuals, these kinds of diagnostic plots are commonly used to identify patterns that are indicative of non-linearity or nonconstant error variance.
-
-Raw residuals and studentized residuals are the two most frequently used residuals in standard residual plots.
+As shown in Figure [ref here], the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns.
+
+There are many departures from a good residual plot, but we will consider only three types of them in this paper, which are also the most commonly checked departures for a residual plot.
+The first one is non-linearity.
+Any unnatural non-linear curve presented in the residual plot could be considered an indicative of non-linearity.
+At the top-right of Figure [ref here], there is an example residual plot of this type of departure.
+The second type of departure  we will consider is heterskedasticity.
+The usual pattern of this type of departure is the inconsistent distribution of the residuals for different x values.
+Visually, it usually results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure [ref here], or the "triangle" shape where the smallest variance occurs at the edges of the x-axis.
+The last type of departure is non-normality.
+This departure is usually harder to be detected from the residuals plot compared to non-linearity and heterskedasticity.
+In most the case, a quantile-quantile plot can do a better job.
+Depending on the actual distribution of the residuals, different visual patterns will be discovered.
+If the distribution is skewed, one will see some outliers presented on only one side of the zero line as shown in the bottom-right of Figure [ref here].
+
+<!-- Raw residuals and studentized residuals are the two most frequently used residuals in standard residual plots.
 The debt on which type of residuals should be used is always present.
@@
 Thus, for a relatively large sample size, @cleveland1975graphical suggests the use of robust moving statistics as reference lines to give aid to the eye in seeing patterns, which nowadays, are usually replaced with a spline or local polynomial regression line.
-
-Other types of data plots that are often used in regression diagnostics include partial residual plots and probability plots.
+-->
+
+<!-- Other types of data plots that are often used in regression diagnostics include partial residual plots and probability plots.
 Partial residual plots are useful supplements to standard residual plots as they provide additional information on the extent of the non-linearity.
 Probability plots can be used to compare the sampling distribution of the residuals to the normal distribution for assessing the normality assumptions.
+-->
 

### 2022-09-04 paper_comparison/paper_comparison/paper_comparison.tex
update residual departure

@@
 orthogonal to residuals are referred to as standard residual plots in
-\citet{cook1982residuals}. As suggested by \citet{cook1982residuals},
-these kinds of diagnostic plots are commonly used to identify patterns
-that are indicative of non-linearity or nonconstant error variance.
-
-Raw residuals and studentized residuals are the two most frequently used
-residuals in standard residual plots. The debt on which type of
-residuals should be used is always present. While raw residuals are the
-most common computer regression software package output, by applying a
-scaling factor, the ability to reveal nonconstant error variance in
-standard residual plots will often be enhanced by studentized residuals
-in small sample size \citep{gunst1980regression}. As a two-dimensional
-representation of a model in a \(p\)-dimensional space, standard
-residual plots project data points onto the variable of the horizontal
-axis, which is a vector in \(p\)-dimensional space. Observations with
-the same projection will be treated as equivalent as they have the same
-abscissa. Consequently, standard residual plots are often useful in
-revealing model inadequacies in the direction of the variable of the
-horizontal axis but could be inadequate for detecting patterns in other
-directions, especially in those perpendicular to the variable of the
-horizontal axis. Hence, in practice, multiple standard residual plots
-with different horizontal axes will be examined
-\citep{cook1982residuals}. Overlapping data points is another general
-issue in scatter plots not limited to standard residual plots, which
-often makes plots difficult to interpret because visual patterns are
-concealed. Thus, for a relatively large sample size,
-\citet{cleveland1975graphical} suggests the use of robust moving
-statistics as reference lines to give aid to the eye in seeing patterns,
-which nowadays, are usually replaced with a spline or local polynomial
-regression line.
-
-Other types of data plots that are often used in regression diagnostics
-include partial residual plots and probability plots. Partial residual
-plots are useful supplements to standard residual plots as they provide
-additional information on the extent of the non-linearity. Probability
-plots can be used to compare the sampling distribution of the residuals
-to the normal distribution for assessing the normality assumptions.
+\citet{cook1982residuals}. As shown in Figure {[}ref here{]}, the
+top-left panel is a good residual plot with residuals evenly distributed
+at both sides of the horizontal zero line showing no noticeable
+patterns.
+
+There are many departures from a good residual plot, but we will
+consider only three types of them in this paper, which are also the most
+commonly checked departures for a residual plot. The first one is
+non-linearity. Any unnatural non-linear curve presented in the residual
+plot could be considered an indicative of non-linearity. At the
+top-right of Figure {[}ref here{]}, there is an example residual plot of
+this type of departure. The second type of departure we will consider is
+heterskedasticity. The usual pattern of this type of departure is the
+inconsistent distribution of the residuals for different x values.
+Visually, it usually results in the so-called ``butterfly'' shape as
+shown in the bottom-left panel of Figure {[}ref here{]}, or the
+``triangle'' shape where the smallest variance occurs at the edges of
+the x-axis. The last type of departure is non-normality. This departure
+is usually harder to be detected from the residuals plot compared to
+non-linearity and heterskedasticity. In most the case, a
+quantile-quantile plot can do a better job. Depending on the actual
+distribution of the residuals, different visual patterns will be
+discovered. If the distribution is skewed, one will see some outliers
+presented on only one side of the zero line as shown in the bottom-right
+of Figure {[}ref here{]}.
 

### 2022-09-05 paper_comparison/paper_comparison/paper_comparison.Rmd
polish the residual departure

@@
 <!-- problem: residual plot diagnostics -->
+
 <!-- conventional test: too sensitive -->
@@
 <!-- week 1 -->
+
 <!-- 1. residual plot for model diagnostics -->
@@
 <!-- a. residual is widely used -->
+
 <!-- b. what are the types of residual plots -->
+
 <!-- c. comparison -->
@@
 <!-- week 2 -->
+
 <!-- 2. conventional test: F, BP -->
+
 <!-- 3. visual test: lineup, theory -->
@@
 <!-- week 4 -->
+
 <!-- desc of experiment: -->
+
 <!-- 1. simulation setup -->
+
 <!-- 2. experimental design -->
+
 <!-- 3. result -->
@@
 <!-- week 3 -->
+
 <!-- comparison of conventional tests: -->
+
 <!-- 1. power (visual test vs. conventional test) -->
+
 <!-- (visual test most different one (everything test, any departure)) -->
+
 <!-- plot figure in a paper, desc, exp -->
+
 <!-- 2. investigate the difference (gap), give examples -->
+
 <!-- 3. conventional is too sensitive -->
+
 <!-- 4. make conventional less sensitive (vary alpha) -->
@@
 <!-- last week -->
+
 <!-- conclusion: -->
+
 <!-- 1. too sensitive, visual test is needed/preferable -->
+
 <!-- 2. visual test is infeasible in large scale (expensive) -->
+
 <!-- 3. future work (role of computer vision) -->
@@
 
+```{=html}
 <!--Regression diagnostics is an essential step in regression analysis which is a field of study with at least a hundred years of history.
@@
 -->
-
+```
 # Background
@@
 
-There are many departures from a good residual plot, but we will consider only three types of them in this paper, which are also the most commonly checked departures for a residual plot.
+There are many types of departures from a good residual plot.
+Three of them will be considered in this paper, which are also the most commonly checked departures for a residual plot.
 The first one is non-linearity.
-Any unnatural non-linear curve presented in the residual plot could be considered an indicative of non-linearity.
+Any unnatural and non-linear functional form of residuals on fitted values presented in the residual plot could be considered as an indicative of non-linearity.
 At the top-right of Figure [ref here], there is an example residual plot of this type of departure.
-The second type of departure  we will consider is heterskedasticity.
-The usual pattern of this type of departure is the inconsistent distribution of the residuals for different x values.
-Visually, it usually results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure [ref here], or the "triangle" shape where the smallest variance occurs at the edges of the x-axis.
-The last type of departure is non-normality.
+The second type of departure that will be considered is heterskedasticity.
+The usual pattern is the inconsistent distribution of the residuals for different x values.
+Visually, it sometimes results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure [ref here], or the "triangle" shape where the smallest variance occurs at the edges of the x-axis.
+The last type of departure we will consider is non-normality.
 This departure is usually harder to be detected from the residuals plot compared to non-linearity and heterskedasticity.
-In most the case, a quantile-quantile plot can do a better job.
+In most cases, a quantile-quantile plot can do a better job.
 Depending on the actual distribution of the residuals, different visual patterns will be discovered.
-If the distribution is skewed, one will see some outliers presented on only one side of the zero line as shown in the bottom-right of Figure [ref here].
+Readers of the residuals plot will notice the data points do not distribute noramlly on the y-axis.
+For example, if the distribution is skewed, one will see the "tail" of the distribution as some outliers presented on only one side of the zero line as shown in the bottom-right of Figure [ref here].
 
@@
 
+```{=html}
 <!-- ![Parallelism between multiple quantitative testing and visual discovery [@buja_statistical_2009].
@@
 -->
-
+```
 In visual inference, a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$ is defined, where $\boldsymbol{\mathrm{y}}$ is the data and $I$ is a set of all possible visual features.
@@
 Then, by imposing a relatively strong assumption on the visual test that all $K$ evaluations are fully independent, under the null hypothesis, $X \sim \mathrm{Binom}_{K,1/m}$.
-Therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as
-\begin{equation} \label{eq:pvaluesingle}
+Therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as \begin{equation} \label{eq:pvaluesingle}
 P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},
@@
 
+```{=tex}
 \begin{align} \label{eq:dirichlet-multinomial}\begin{split}
@@
 \end{split}\end{align}
-
+```
 Since the p-value calculation only needs to concern the number of times the actual data plot being selected denoted by $C_i$, they showed the model can be simplified to a beta-binomial mixture distribution
 
+```{=tex}
 \begin{align} \label{eq:beta-binomial}\begin{split}
@@
 \end{split}\end{align}
-
+```
 Thus, the visual p-value followed by the beta-binomial model is given as
 
+```{=tex}
 \begin{equation} \label{eq:pvalue-beta-binomial}
@@
 \end{equation}
-
+```
 where $B(.)$ is the beta function defined as
 
+```{=tex}
 \begin{equation} \label{eq:betafunction}
@@
 \end{equation}
-
+```
 Note that the use of Equation (\ref{eq:pvalue-beta-binomial}) requires the estimation of $\hat{\alpha}$, which largely depends on the null model, the type of the plot and other aesthetic features.
@@
 
+```{=tex}
 \begin{equation} \label{eq:p-value-multi}
@@
 \end{equation}
-
+```
 <!-- ### Effectiveness of visual inference in regression diagnostics -->
@@
 <!-- $$g(p_{li}) = W_{li}\delta + Z_{li}\tau_{li},$$ -->
+
 <!-- where $g(.)$ is the logit link function $g(p) = log(p)  - log(1-p)$; $0 \leq p \leq 1$.
@@
 
+```{=tex}
 \begin{align} \label{eq:nonlinearity-model}
@@
 \end{align}
-
+```
 where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vector of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $\{-k, k\}$ defined as
 
+```{=tex}
 \begin{equation} \label{eq:scaling-function}
@@
 \end{equation}
-
+```
 The null regression model used to fit the realizations generated by the above model is formulated as:
 
+```{=tex}
 \begin{equation} \label{eq:scaling-function}
@@
 \end{equation}
-
+```
 where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.
@@
 
+```{=tex}
 \begin{align} \label{eq:heter-model}
@@
 \end{align}
-
+```
 where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vector of size $n$.
@@
 
+```{=tex}
 \begin{equation} 
@@
 
-A subject was allowed to select zero plots for a lineup if there was no visible difference between plots,
-but the simulated visual p-value given in Equation (\ref{eq:p-value-multi}) will effectively drop the subject from the simulation for this case, leading to inaccurate estimation of $p$-value.
+A subject was allowed to select zero plots for a lineup if there was no visible difference between plots, but the simulated visual p-value given in Equation (\ref{eq:p-value-multi}) will effectively drop the subject from the simulation for this case, leading to inaccurate estimation of $p$-value.
 Therefore, we treated this case as making one selection but failing to identify the actual data plot so that Equation (\ref{eq:p-value-multi}) can be applied correctly.
@@
 
+```{=tex}
 \begin{equation} \label{eq:effect-size-ex1}
@@
 \end{equation}
-
+```
 where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$.
@@
 
+```{=tex}
 \begin{equation} \label{eq:logistic-regression-1-1}
@@
 \end{equation}
-
+```
 where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
@@
 In overall, the power of all four tests increases as the effect becomes larger.
-The power curve of F-test climbs aggressively from 25\% to around 90\% as $log_e(E)$ increases from 0 to 2, while others respond inactively to the change of effect and remain lower than 25\% throughout the period, showing that as an exact test, the F-test is relatively more sensitive to the type of model defects that being considered.
-The power of visual test arises steadily and nearly linearly to around 90\% as $log_e(E)$ increases from 2 to 5, suggesting that the effect starts to make noticeable impact on the degree of the presence of the designed visual features.
+The power curve of F-test climbs aggressively from 25% to around 90% as $log_e(E)$ increases from 0 to 2, while others respond inactively to the change of effect and remain lower than 25% throughout the period, showing that as an exact test, the F-test is relatively more sensitive to the type of model defects that being considered.
+The power of visual test arises steadily and nearly linearly to around 90% as $log_e(E)$ increases from 2 to 5, suggesting that the effect starts to make noticeable impact on the degree of the presence of the designed visual features.
 Other two inappropriate conventional tests shows improvement at the same time but at a lower rate.
@@
 At $log_e(E) = 6$, the power curve of F-test reaches almost 100% followed by the visual test by a small margin.
-The power of Breusch–Pagan test and Shapiro–Wilk test reach around 75% and 63% respectively.
+The power of Breusch--Pagan test and Shapiro--Wilk test reach around 75% and 63% respectively.
 
@@
 <!-- # time taken vs order of the plot -->
+
 <!-- # accuracy vs order of the plot -->
+
 <!-- # num of selection vs order of the plot -->
+
 <!-- # order of the plot? -->
+
 <!-- # summary of the demographic information -->
@@
 <!-- 1. power (visual test vs. conventional test) -->
+
 <!-- (visual test most different one (everything test, any departure)) -->
+
 <!-- plot figure in a paper, desc, exp -->
+
 <!-- 2. investigate the difference (gap), give examples -->
+
 <!-- 3. conventional is too sensitive -->
+
 <!-- 4. make conventional less sensitive (vary alpha) -->
@@
 
-<!-- <!-- sample 1, change values of a and b, plot the data  --> -->
+<!-- <!-- sample 1, change values of a and b, plot the data  -->
+
+--\>
 
@@
 <!-- ```{r power-vs-log-effect-size} -->
+
 <!-- polynomials %>% -->
+
 <!--   group_by(lineup_id) %>% -->
+
 <!--   summarise(across(c(type:p_value), first)) %>% -->
+
 <!--   filter(lineup_id < 577) %>% -->
+
 <!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
+
 <!--   ggplot() + -->
+
 <!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+
 <!--   geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+
 <!--   xlab("Natural logarithm of effect size") + -->
+
 <!--   ylab("Power") -->
+
 <!-- ``` -->
@@
 <!-- ```{r power-vs-log-effect-size-given-x-dist} -->
+
 <!-- polynomials %>% -->
+
 <!--   group_by(lineup_id) %>% -->
+
 <!--   summarise(across(c(type:p_value), first)) %>% -->
+
 <!--   filter(lineup_id < 577) %>% -->
+
 <!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
+
 <!--   ggplot() + -->
+
 <!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+
 <!--   geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+
 <!--   facet_wrap(~x_dist) + -->
+
 <!--   xlab("Natural logarithm of effect size") + -->
+
 <!--   ylab("Power") + -->
+
 <!--   ggtitle("Power comparison conditional on the distribution of X") -->
+
 <!-- ``` -->
@@
 <!-- ```{r power-of-visual-test-given-x-dist} -->
+
 <!-- polynomials %>% -->
+
 <!--   group_by(lineup_id) %>% -->
+
 <!--   summarise(across(c(type:p_value), first)) %>% -->
+
 <!--   filter(lineup_id < 577) %>% -->
+
 <!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
+
 <!--   ggplot() + -->
+
 <!--   geom_smooth(aes(log(effect_size), reject, col = x_dist), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+
 <!--   xlab("Natural logarithm of effect size") + -->
+
 <!--   ylab("Power") + -->
+
 <!--   ggtitle("Power of visual test conditional on the distribution of X") -->
+
 <!-- ``` -->
@@
 <!-- ```{r power-vs-log-effect-size-given-shape} -->
+
 <!-- polynomials %>% -->
+
 <!--   group_by(lineup_id) %>% -->
+
 <!--   summarise(across(c(type:p_value), first)) %>% -->
+
 <!--   filter(lineup_id < 577) %>% -->
+
 <!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
+
 <!--   ggplot() + -->
+
 <!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+
 <!--   geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+
 <!--   facet_wrap(~shape) + -->
+
 <!--   xlab("Natural logarithm of effect size") + -->
+
 <!--   ylab("Power") + -->
+
 <!--   ggtitle("Power comparison conditional on the shape of the Hermite polynomials") -->
+
 <!-- ``` -->
@@
 <!-- ```{r power-of-visual-test-given-shape} -->
+
 <!-- polynomials %>% -->
+
 <!--   group_by(lineup_id) %>% -->
+
 <!--   summarise(across(c(type:p_value), first)) %>% -->
+
 <!--   filter(lineup_id < 577) %>% -->
+
 <!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01)), shape = factor(shape)) %>% -->
+
 <!--   ggplot() + -->
+
 <!--   geom_smooth(aes(log(effect_size), reject, col = shape), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+
 <!--   xlab("Natural logarithm of effect size") + -->
+
 <!--   ylab("Power") + -->
+
 <!--   ggtitle("Power of visual test conditional on the shape of the Hermite polynomials") -->
+
 <!-- ``` -->
@@
 <!-- ```{r power-vs-log-effect-size-given-number-of-observations} -->
+
 <!-- polynomials %>% -->
+
 <!--   group_by(lineup_id) %>% -->
+
 <!--   summarise(across(c(type:p_value), first)) %>% -->
+
 <!--   filter(lineup_id < 577) %>% -->
+
 <!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
+
 <!--   ggplot() + -->
+
 <!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+
 <!--   geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+
 <!--   facet_wrap(~n) + -->
+
 <!--   xlab("Natural logarithm of effect size") + -->
+
 <!--   ylab("Power") + -->
+
 <!--   ggtitle("Power comparison conditional on the number of observations") -->
+
 <!-- ``` -->
@@
 <!-- ```{r power-of-visual-test-given-number-of-observations} -->
+
 <!-- polynomials %>% -->
+
 <!--   group_by(lineup_id) %>% -->
+
 <!--   summarise(across(c(type:p_value), first)) %>% -->
+
 <!--   filter(lineup_id < 577) %>% -->
+
 <!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01)), n = factor(n)) %>% -->
+
 <!--   ggplot() + -->
+
 <!--   geom_smooth(aes(log(effect_size), reject, col = n), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+
 <!--   xlab("Natural logarithm of effect size") + -->
+
 <!--   ylab("Power") + -->
+
 <!--   ggtitle("Power of visual test conditional on the number of observations") -->
+
 <!-- ``` -->
@@
 <!-- ```{r} -->
+
 <!-- polynomials %>% -->
+
 <!--   group_by(lineup_id) %>% -->
+
 <!--   summarise(across(c(type:p_value), first)) %>% -->
+
 <!--   filter(lineup_id < 577) %>% -->
+
 <!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
+
 <!--   ggplot() + -->
+
 <!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+
 <!--   geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.05)), aes(log(effect_size), reject, col = "Conventional test:0.05"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+
 <!--   geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.01)), aes(log(effect_size), reject, col = "Conventional test:0.01"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+
 <!--   geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.005)), aes(log(effect_size), reject, col = "Conventional test:0.005"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+
 <!--   geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.0005)), aes(log(effect_size), reject, col = "Conventional test:0.0005"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+
 <!--   geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.0000005)), aes(log(effect_size), reject, col = "Conventional test:0.0000005"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
+
 <!--   xlab("Natural logarithm of effect size") + -->
+
 <!--   ylab("Power") -->
+
 <!-- ``` -->

### 2022-09-05 paper_comparison/paper_comparison/paper_comparison.tex
polish the residual departure

@@
 
-There are many departures from a good residual plot, but we will
-consider only three types of them in this paper, which are also the most
-commonly checked departures for a residual plot. The first one is
-non-linearity. Any unnatural non-linear curve presented in the residual
-plot could be considered an indicative of non-linearity. At the
-top-right of Figure {[}ref here{]}, there is an example residual plot of
-this type of departure. The second type of departure we will consider is
-heterskedasticity. The usual pattern of this type of departure is the
-inconsistent distribution of the residuals for different x values.
-Visually, it usually results in the so-called ``butterfly'' shape as
-shown in the bottom-left panel of Figure {[}ref here{]}, or the
-``triangle'' shape where the smallest variance occurs at the edges of
-the x-axis. The last type of departure is non-normality. This departure
-is usually harder to be detected from the residuals plot compared to
-non-linearity and heterskedasticity. In most the case, a
-quantile-quantile plot can do a better job. Depending on the actual
-distribution of the residuals, different visual patterns will be
-discovered. If the distribution is skewed, one will see some outliers
-presented on only one side of the zero line as shown in the bottom-right
-of Figure {[}ref here{]}.
+There are many types of departures from a good residual plot. Three of
+them will be considered in this paper, which are also the most commonly
+checked departures for a residual plot. The first one is non-linearity.
+Any unnatural and non-linear functional form of residuals on fitted
+values presented in the residual plot could be considered as an
+indicative of non-linearity. At the top-right of Figure {[}ref here{]},
+there is an example residual plot of this type of departure. The second
+type of departure that will be considered is heterskedasticity. The
+usual pattern is the inconsistent distribution of the residuals for
+different x values. Visually, it sometimes results in the so-called
+``butterfly'' shape as shown in the bottom-left panel of Figure {[}ref
+here{]}, or the ``triangle'' shape where the smallest variance occurs at
+the edges of the x-axis. The last type of departure we will consider is
+non-normality. This departure is usually harder to be detected from the
+residuals plot compared to non-linearity and heterskedasticity. In most
+cases, a quantile-quantile plot can do a better job. Depending on the
+actual distribution of the residuals, different visual patterns will be
+discovered. Readers of the residuals plot will notice the data points do
+not distribute noramlly on the y-axis. For example, if the distribution
+is skewed, one will see the ``tail'' of the distribution as some
+outliers presented on only one side of the zero line as shown in the
+bottom-right of Figure {[}ref here{]}.
 

### 2022-09-06 paper_comparison/paper_comparison/paper_comparison.Rmd
update introduction

@@
 Using the classical linear regression model as an example, if the distribution is symmetric and unimodal, it is well-behaved.
-But if the distribution is skewed, bimodal, multimodal, or even contains outliers, there is cause for concern.
-Conducting a goodness of fit test is an alternative method for inspecting the distribution, such as the Shapiro-Wilk Normality test [@shapiro1965analysis].
-
-To scrutinize the relationship between residuals, predicted values and each of the explanatory variables, it is recommend to plot them on a scatter plot.
-In general, the model is well-behaved if there is any visual discoverable pattern.
+But if the distribution is skewed, bimodal, multimodal, or contains outliers, there is cause for concern.
+The distribution could also be inspected by conducting a goodness of fit test, such as the Shapiro-Wilk Normality test [@shapiro1965analysis].
+
+Plotting the residuals against predicted values and each of the explanatory variables on a scatter plot is a recommend way to scrutinize their relationships.
+If there is any visually discoverable patterns, the model is potentially misspecified.
 However, it is a very difficult task for a human judge, though to make a decision that there's nothing there.
@@
 Generally, one looks for departures from nothingness like non-linear dependency or heteroskedasticity.
-It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and also use a Breusch-Pagan test [@breusch_simple_1979] for heteroskedasticity.
+It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and use a Breusch-Pagan test [@breusch_simple_1979] for heteroskedasticity.
 
@@
 The paper is structured as follows.
-The next background section describes the commonly used diagnostic plots, the types of departures that one expects to detect, and describes a formal process for reading residual plots, called visual inference, that can avoid the concerns about subjectiveness of human readers.
+The next background section describes the the types of departures that one expects to detect, and describes a formal process for reading residual plots, called visual inference, that can avoid the concerns about subjectiveness of human readers.
 Section \ref{experimental-design} describes the experimental setup to enable a comparison between decision made by formal hypothesis testing, and how humans would read diagnostic plots.
@@
 As shown in Figure [ref here], the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns.
-
-There are many types of departures from a good residual plot.
-Three of them will be considered in this paper, which are also the most commonly checked departures for a residual plot.
-The first one is non-linearity.
-Any unnatural and non-linear functional form of residuals on fitted values presented in the residual plot could be considered as an indicative of non-linearity.
-At the top-right of Figure [ref here], there is an example residual plot of this type of departure.
-The second type of departure that will be considered is heterskedasticity.
-The usual pattern is the inconsistent distribution of the residuals for different x values.
-Visually, it sometimes results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure [ref here], or the "triangle" shape where the smallest variance occurs at the edges of the x-axis.
-The last type of departure we will consider is non-normality.
-This departure is usually harder to be detected from the residuals plot compared to non-linearity and heterskedasticity.
-In most cases, a quantile-quantile plot can do a better job.
-Depending on the actual distribution of the residuals, different visual patterns will be discovered.
-Readers of the residuals plot will notice the data points do not distribute noramlly on the y-axis.
-For example, if the distribution is skewed, one will see the "tail" of the distribution as some outliers presented on only one side of the zero line as shown in the bottom-right of Figure [ref here].
+There are various types of departures from a good residual plot.
+We will only discuss three most commonly checked departures in this paper.
+Namely, non-linearity, heterskedasticity and non-normality.
+
+Non-linearity is a type of model misspecification caused by failing to include higher order terms of the regressors in the regression equation.
+Any non-linear functional form of residuals on fitted values presented in the residual plot could be considered as an indicative of non-linearity.
+At the top-right of Figure [ref here], there is a residual plot giving an example of the visual pattern of non-linearity when a cubic term is not captured by the regression model.
+
+Heterskedasticity refers to the existence of nonconst error variance in a regression model.
+It is mostly due to the strict but false assumptions on the variance-covariance matrix of the error term.
+The usual pattern of heterskedasticity on a residual plot is the inconsistent spread of the residuals at different x values.
+Visually, it sometimes results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure [ref here], or the "left-triangle" and "right-triganle" shape where the smallest variance occurs at the edges of the x-axis.
+
+Non-normality is usually harder to be detected from a residual plot compared to non-linearity and heterskedasticity.
+Besides, a quantile-quantile plot can often do a better job for this task.
+Note that not all regression models assume normality of the error term, but a certain amount of them does.
+If this assumption happens to be false, then it is expected to see data points do not distribute noramlly on the y-axis.
+For example, given a skewed error distribution, one will see fewer data points and more outliers on one side of the zeroline as shown in the bottom-right of Figure [ref here].
 

### 2022-09-06 paper_comparison/paper_comparison/paper_comparison.tex
update introduction

--- 2022-09-04 paper_comparison/paper_comparison/paper_comparison.tex
@@
 well-behaved. But if the distribution is skewed, bimodal, multimodal, or
-even contains outliers, there is cause for concern. Conducting a
-goodness of fit test is an alternative method for inspecting the
-distribution, such as the Shapiro-Wilk Normality test
-\citep{shapiro1965analysis}.
-
-To scrutinize the relationship between residuals, predicted values and
-each of the explanatory variables, it is recommend to plot them on a
-scatter plot. In general, the model is well-behaved if there is any
-visual discoverable pattern. However, it is a very difficult task for a
-human judge, though to make a decision that there's nothing there. It is
-especially common, particularly among new data analysts to report
+contains outliers, there is cause for concern. The distribution could
+also be inspected by conducting a goodness of fit test, such as the
+Shapiro-Wilk Normality test \citep{shapiro1965analysis}.
+
+Plotting the residuals against predicted values and each of the
+explanatory variables on a scatter plot is a recommend way to scrutinize
+their relationships. If there is any visually discoverable patterns, the
+model is potentially misspecified. However, it is a very difficult task
+for a human judge, though to make a decision that there's nothing there.
+It is especially common, particularly among new data analysts to report
 patterns when an experienced data analyst might quickly conclude that
@@
 conduct hypothesis tests for non-linear dependence
-\citep{ramsey_tests_1969}, and also use a Breusch-Pagan test
+\citep{ramsey_tests_1969}, and use a Breusch-Pagan test
 \citep{breusch_simple_1979} for heteroskedasticity.
@@
 the subject of this paper. The paper is structured as follows. The next
-background section describes the commonly used diagnostic plots, the
-types of departures that one expects to detect, and describes a formal
-process for reading residual plots, called visual inference, that can
-avoid the concerns about subjectiveness of human readers. Section
-\ref{experimental-design} describes the experimental setup to enable a
-comparison between decision made by formal hypothesis testing, and how
-humans would read diagnostic plots. The results are reported in Section
-\ref{results}. We finish with a discussion on future work, in particular
-how the responsibility for residual plot reading might be passed on to
-computer vision.
+background section describes the the types of departures that one
+expects to detect, and describes a formal process for reading residual
+plots, called visual inference, that can avoid the concerns about
+subjectiveness of human readers. Section \ref{experimental-design}
+describes the experimental setup to enable a comparison between decision
+made by formal hypothesis testing, and how humans would read diagnostic
+plots. The results are reported in Section \ref{results}. We finish with
+a discussion on future work, in particular how the responsibility for
+residual plot reading might be passed on to computer vision.
 
@@
 at both sides of the horizontal zero line showing no noticeable
-patterns.
-
-There are many departures from a good residual plot, but we will
-consider only three types of them in this paper, which are also the most
-commonly checked departures for a residual plot. The first one is
-non-linearity. Any unnatural non-linear curve presented in the residual
-plot could be considered an indicative of non-linearity. At the
-top-right of Figure {[}ref here{]}, there is an example residual plot of
-this type of departure. The second type of departure we will consider is
-heterskedasticity. The usual pattern of this type of departure is the
-inconsistent distribution of the residuals for different x values.
-Visually, it usually results in the so-called ``butterfly'' shape as
-shown in the bottom-left panel of Figure {[}ref here{]}, or the
-``triangle'' shape where the smallest variance occurs at the edges of
-the x-axis. The last type of departure is non-normality. This departure
-is usually harder to be detected from the residuals plot compared to
-non-linearity and heterskedasticity. In most the case, a
-quantile-quantile plot can do a better job. Depending on the actual
-distribution of the residuals, different visual patterns will be
-discovered. If the distribution is skewed, one will see some outliers
-presented on only one side of the zero line as shown in the bottom-right
-of Figure {[}ref here{]}.
+patterns. There are various types of departures from a good residual
+plot. We will only discuss three most commonly checked departures in
+this paper. Namely, non-linearity, heterskedasticity and non-normality.
+
+Non-linearity is a type of model misspecification caused by failing to
+include higher order terms of the regressors in the regression equation.
+Any non-linear functional form of residuals on fitted values presented
+in the residual plot could be considered as an indicative of
+non-linearity. At the top-right of Figure {[}ref here{]}, there is a
+residual plot giving an example of the visual pattern of non-linearity
+when a cubic term is not captured by the regression model.
+
+Heterskedasticity refers to the existence of nonconst error variance in
+a regression model. It is mostly due to the strict but false assumptions
+on the variance-covariance matrix of the error term. The usual pattern
+of heterskedasticity on a residual plot is the inconsistent spread of
+the residuals at different x values. Visually, it sometimes results in
+the so-called ``butterfly'' shape as shown in the bottom-left panel of
+Figure {[}ref here{]}, or the ``left-triangle'' and ``right-triganle''
+shape where the smallest variance occurs at the edges of the x-axis.
+
+Non-normality is usually harder to be detected from a residual plot
+compared to non-linearity and heterskedasticity. Besides, a
+quantile-quantile plot can often do a better job for this task. Note
+that not all regression models assume normality of the error term, but a
+certain amount of them does. If this assumption happens to be false,
+then it is expected to see data points do not distribute noramlly on the
+y-axis. For example, given a skewed error distribution, one will see
+fewer data points and more outliers on one side of the zeroline as shown
+in the bottom-right of Figure {[}ref here{]}.
 

### 2022-09-06 paper_comparison/paper_comparison/paper_comparison.Rmd
update hypothesis testing

@@
 ```{r}
+
+set.seed(10086)
+
 mod <- poly_model(include_z = FALSE, sigma = 0.25)
@@
 
-Heterskedasticity refers to the existence of nonconst error variance in a regression model.
+Heterskedasticity refers to the presence of nonconst error variance in a regression model.
 It is mostly due to the strict but false assumptions on the variance-covariance matrix of the error term.
@@
 
+Other than checking diagnostic plots, analysts may perform formal hypothesis testing for detecting model defects.
+Depending on the alternative hypothesis that is focused on, a variety of tests can be applied.
+For example, the presence of heteroskedasticity can usually be tested by applying the White test [\@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [\@breusch_simple_1979], which are both derived from the Lagrange multiplier test [\@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution.
+For testing non-linearity, one may apply the F-test to examine the significance of specific polynomial and non-linear forms of the regressors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [\@ramsey_tests_1969].
+And for testing normality, the shapiro--Wilk test [ref here] is perhaps the most widely used test reported by many of the statistical softwares.
+Another choice will be the Jarque--Bera test [ref here] which directly checks if the sample skewness and kurtosis match a normal distribution.
+
+```{r}
+
+set.seed(10086)
+
+mod <- poly_model(include_z = FALSE, sigma = 0.25)
+dat <- mod$gen(300)
+patchwork::wrap_plots(mod$plot(dat, theme = theme_light()) + ggtitle("Good residuals"),
+                      mod$plot(mod$
+                                 set_prm("include_z", TRUE)$
+                                 gen(300, computed = select(dat, x, e)), theme = theme_light()) + 
+                        ggtitle(paste("F-test p-value =", 
+                                      mod$
+                                        test(mod$
+                                               gen(300, computed = select(dat, x, e)))$p_value %>% 
+                                        format(digits = 2))),
+                      mod$
+                        plot(heter_model(b = 64)$
+                               gen(300, computed = select(dat, x, e)), theme = theme_light()) + 
+                        ggtitle(paste("BP test p-value =", 
+                                      HETER_MODEL$
+                                        test(heter_model(b = 64)$
+                                               gen(300, computed = select(dat, x, e)))$p_value %>% 
+                                        format(digits = 2))),
+                      mod$
+                        plot(heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
+                               gen(300, computed = select(dat, x)), theme = theme_light()) + 
+                        ggtitle(paste("Non-normality", 
+                                      shapiro.test(heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
+                                                     gen(300, computed = select(dat, x))$.resid)[[2]] %>% 
+                                        format(digits = 2))))
+
+```
+
 Explain the tests you are showing later
@@
 
+Unlike hypothesis testing built upon rigorous statistical procedures, reading diagnostic plots relies on graphical perception - human's ability to interpret and decode the information embedded in the graph [\@cleveland_graphical_1984], which is to some extent subjective and indecisive.
+Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
+One such example is finding an over-interpretation of the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurrence [\@roy_chowdhury_using_2015].
+
 -   Very briefly explain what a lineup test is
+
 -   Lineup (maybe using your previous residual plot for nonlinearity)
+
 -   $p$-value calulcation, briefly, including how multiple selections are handled
+
 -   power calculation, briefly, including how multiple selections are handled
-
-However, unlike hypothesis testing built upon rigorous statistical procedures, reading diagnostic plots relies on graphical perception - human's ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984], which is to some extent subjective and indecisive.
-Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
-One such example is finding an over-interpretation of the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurrence [@roy_chowdhury_using_2015].
 

### 2022-09-06 paper_comparison/paper_comparison/paper_comparison.tex
update hypothesis testing

@@
 
-Heterskedasticity refers to the existence of nonconst error variance in
-a regression model. It is mostly due to the strict but false assumptions
+Heterskedasticity refers to the presence of nonconst error variance in a
+regression model. It is mostly due to the strict but false assumptions
 on the variance-covariance matrix of the error term. The usual pattern
@@
 
+Other than checking diagnostic plots, analysts may perform formal
+hypothesis testing for detecting model defects. Depending on the
+alternative hypothesis that is focused on, a variety of tests can be
+applied. For example, the presence of heteroskedasticity can usually be
+tested by applying the White test
+{[}@white\_heteroskedasticity-consistent\_1980{]} or the Breusch-Pagan
+test {[}@breusch\_simple\_1979{]}, which are both derived from the
+Lagrange multiplier test {[}@silvey1959lagrangian{]} principle that
+relies on the asymptotic properties of the null distribution. For
+testing non-linearity, one may apply the F-test to examine the
+significance of specific polynomial and non-linear forms of the
+regressors, or the significance of proxy variables as in the Ramsey
+Regression Equation Specification Error Test (RESET)
+{[}@ramsey\_tests\_1969{]}. And for testing normality, the shapiro--Wilk
+test {[}ref here{]} is perhaps the most widely used test reported by
+many of the statistical softwares. Another choice will be the
+Jarque--Bera test {[}ref here{]} which directly checks if the sample
+skewness and kurtosis match a normal distribution.
+
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-7-1.pdf}
+
 Explain the tests you are showing later
@@
 
+Unlike hypothesis testing built upon rigorous statistical procedures,
+reading diagnostic plots relies on graphical perception - human's
+ability to interpret and decode the information embedded in the graph
+{[}@cleveland\_graphical\_1984{]}, which is to some extent subjective
+and indecisive. Further, visual discovery suffers from its unsecured and
+unconfirmed nature where the degree of the presence of the visual
+features typically can not be measured quantitatively and objectively,
+which may lead to over or under-interpretations of the data. One such
+example is finding an over-interpretation of the separation between gene
+groups in a two-dimensional projection from a linear discriminant
+analysis when in fact there are no differences in the expression levels
+between the gene groups and separation is not an uncommon occurrence
+{[}@roy\_chowdhury\_using\_2015{]}.
+
 \begin{itemize}
-\tightlist
 \item
@@
 \end{itemize}
-
-However, unlike hypothesis testing built upon rigorous statistical
-procedures, reading diagnostic plots relies on graphical perception -
-human's ability to interpret and decode the information embedded in the
-graph \citep{cleveland_graphical_1984}, which is to some extent
-subjective and indecisive. Further, visual discovery suffers from its
-unsecured and unconfirmed nature where the degree of the presence of the
-visual features typically can not be measured quantitatively and
-objectively, which may lead to over or under-interpretations of the
-data. One such example is finding an over-interpretation of the
-separation between gene groups in a two-dimensional projection from a
-linear discriminant analysis when in fact there are no differences in
-the expression levels between the gene groups and separation is not an
-uncommon occurrence \citep{roy_chowdhury_using_2015}.
 
@@
 
-\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-13-1.pdf}
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-14-1.pdf}
 
@@
 
-\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-16-1.pdf}
-
 \includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-17-1.pdf}
@@
 \includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-20-1.pdf}
+
+\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-21-1.pdf}
 

### 2022-09-07 paper_comparison/paper_comparison/paper_comparison.Rmd
figure sizing and caption plus annotation for first one

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
+knitr::opts_chunk$set(
+  message = FALSE, 
+  warning = FALSE, 
+  echo = FALSE,
+  fig.width = 8,
+  fig.height = 6,
+  out.width = "100%")
 ```
@@
 
-```{r}
+```{r commondepartures, fig.cap = "Example fitted vs residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
+The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted plot."}
 
@@
 dat <- mod$gen(300)
-patchwork::wrap_plots(mod$plot(dat, theme = theme_light()) + ggtitle("Good residuals"),
-                      mod$plot(mod$set_prm("include_z", TRUE)$gen(300, computed = select(dat, x, e)), theme = theme_light()) + ggtitle("Non-linearity"),
-                      mod$plot(heter_model(b = 64)$gen(300, computed = select(dat, x, e)), theme = theme_light()) + ggtitle("Heteroskedasticity"),
-                      mod$plot(heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$gen(300, computed = select(dat, x)), theme = theme_light()) + ggtitle("Non-normality"))
+patchwork::wrap_plots(
+  mod$plot(dat, theme = theme_light()) +
+    ggtitle("A. Good residuals") + 
+    xlab("fitted values") + ylab("residuals"),
+  mod$plot(mod$set_prm("include_z", TRUE)$gen(300, 
+    computed = select(dat, x, e)), 
+    theme = theme_light()) + 
+    ggtitle("B. Non-linearity") +
+    xlab("fitted values") + ylab("residuals"),
+  mod$plot(heter_model(b = 64)$gen(300, 
+    computed = select(dat, x, e)), 
+    theme = theme_light()) + 
+    ggtitle("C. Heteroskedasticity") +
+    xlab("fitted values") + ylab("residuals"),
+  mod$plot(heter_model(b = 0, 
+    e = rand_lognormal(sigma = 0.5))$gen(300, 
+    computed = select(dat, x)), 
+    theme = theme_light()) + 
+    ggtitle("D. Non-normality") +
+    xlab("fitted values") + ylab("residuals")
+)
 ```

### 2022-09-07 paper_comparison/paper_comparison/paper_comparison.tex
figure sizing and caption plus annotation for first one

@@
 
-\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-6-1.pdf}
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/commondepartures-1} \caption{Example fitted vs residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
+The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted plot.}\label{fig:commondepartures}
+\end{figure}
 
@@
 
-\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-7-1.pdf}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-6-1}
 
@@
 \begin{figure}
-\centering
-\includegraphics{paper_comparison_files/figure-latex/residual-plot-cubic-heter-1.pdf}
-\caption{Residuals vs.~fitted values plot for a classical linear
-regression model. The residuals are produced by fitting a two-predictor
-multiple linear regression model with data generated from a cubic linear
-model. From the residual plot, ``butterfly shape'' can be observed which
-generally would be interpretd as evidence of heteroskedasticity.
-Further, from the outline of the shape, nonlinear patterns exist. Both
-visual discoveries are evidence against the null hypothesis, though
-heteroskedasticity actually does not exist in the data generating
-process. \label{fig:residual-plot-cubic-heter}}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/residual-plot-cubic-heter-1} \caption{Residuals vs.
+fitted values plot for a classical linear regression model.
+The residuals are produced by fitting a two-predictor multiple linear regression model with data generated from a cubic linear model.
+From the residual plot, "butterfly shape" can be observed which generally would be interpretd as evidence of heteroskedasticity.
+Further, from the outline of the shape, nonlinear patterns exist.
+Both visual discoveries are evidence against the null hypothesis, though heteroskedasticity actually does not exist in the data generating process.
+\label{fig:residual-plot-cubic-heter}}\label{fig:residual-plot-cubic-heter}
 \end{figure}
@@
 
-\includegraphics{paper_comparison_files/figure-latex/different-shape-of-herimite-1.pdf}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-herimite-1}
 
@@
 
-\includegraphics{paper_comparison_files/figure-latex/different-sigma-1.pdf}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-sigma-1}
 
@@
 
-\includegraphics{paper_comparison_files/figure-latex/different-dist-1.pdf}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-dist-1}
 
@@
 \begin{figure}
-\centering
-\includegraphics{paper_comparison_files/figure-latex/example-lineup-1.pdf}
-\caption{Example lineup \label{fig:example-lineup}}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-lineup-1} \caption{Example lineup \label{fig:example-lineup}}\label{fig:example-lineup}
 \end{figure}
@@
 
-\includegraphics{paper_comparison_files/figure-latex/different-shape-of-heter-1.pdf}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-heter-1}
 
@@
 \begin{figure}
-\centering
-\includegraphics{paper_comparison_files/figure-latex/demographic-summary-plot-1.pdf}
-\caption{Summary of demographic information.
-\label{fig:demographic-summary-plot}}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/demographic-summary-plot-1} \caption{Summary of demographic information.
+\label{fig:demographic-summary-plot}}\label{fig:demographic-summary-plot}
 \end{figure}
@@
 \begin{figure}
-\centering
-\includegraphics{paper_comparison_files/figure-latex/power-overview-1.pdf}
-\caption{Power overview. \label{fig:power-overview}}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/power-overview-1} \caption{Power overview. \label{fig:power-overview}}\label{fig:power-overview}
 \end{figure}
@@
 \begin{figure}
-\centering
-\includegraphics{paper_comparison_files/figure-latex/dist-power-1.pdf}
-\caption{\label{dist-power}}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/dist-power-1} \caption{\label{dist-power}}\label{fig:dist-power}
 \end{figure}
@@
 \begin{figure}
-\centering
-\includegraphics{paper_comparison_files/figure-latex/shape-power-1.pdf}
-\caption{\label{shape-power}}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/shape-power-1} \caption{\label{shape-power}}\label{fig:shape-power}
 \end{figure}
@@
 
-\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-14-1.pdf}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-13-1}
 
@@
 
-\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-17-1.pdf}
-
-\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-18-1.pdf}
-
-\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-19-1.pdf}
-
-\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-20-1.pdf}
-
-\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-21-1.pdf}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-16-1}
+
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-17-1}
+
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-18-1}
+
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-19-1}
+
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-20-1}
 

### 2022-09-08 Fifth_experiment/exp_setup.R
add exp 5

@@ -1,0 +1,327 @@
+library(visage)
+library(tidyverse)
+library(cli)
+
+
+# We need
+# a (3) = -1, 0, 1
+# b (5) = 0.25, 1, 4, 16, 64
+# x_dist (1) = uniform
+# n (3) = 50, 100, 300
+# rep (3)
+#
+# 3 * 5 * 1 * 3 * 3 = 135
+#
+# shape (4)
+# e_sigma (4) = 0.5, 1, 2, 4
+# x_dist (1) = uniform
+# n (3) = 50, 100, 300
+# rep (3)
+#
+# 4 * 4 * 1 * 3 * 3 = 144
+#
+# null lineups for heter and poly:
+# x_dist (4)
+# n (3)
+# rep (3)
+# 4 * 3 * 3 = 36
+#
+# ((135 + 144) * 6 + 36 * 15) / 18 = 123
+#
+# We need 123 subjects
+
+
+library(visage)
+library(tidyverse)
+
+lineup <- polynomials %>%
+  mutate(exp = "poly") %>%
+  bind_rows(heter %>% mutate(exp = "heter")) %>%
+  filter((x_dist == "uniform" | b == 0), is.na(b) | b <= 64, e_sigma > 0.25) %>%
+  mutate(lineup_id = ifelse(exp == "poly", lineup_id, lineup_id + 576)) %>%
+  group_by(lineup_id) %>%
+  summarise(across(c(a, b, n, x_dist, shape, e_sigma), first))
+
+
+set.seed(10086)
+
+
+check_comb <- function(set_dat, new_dat, n = 1) {
+
+  apply(combn(dim(set_dat)[2], n),
+        MARGIN = 2,
+        function(this_comb) {
+          # browser()
+          tmp <- as.numeric(new_dat[1, this_comb])
+          for (i in 1:nrow(set_dat)) {
+            if (all(as.numeric(set_dat[i, this_comb]) == tmp)) return(TRUE)
+          }
+          return(FALSE)
+        })
+}
+
+sorted_lineup <- lineup %>%
+  arrange(lineup_id) %>%
+  select(-rep, -lineup_id)
+
+# Caculate distance between a set of allocated lineups and a new lineup
+calc_dist <- function(lineup, allocated_lineup_ids, new_lineup_id) {
+
+  new_lineup <- sorted_lineup[new_lineup_id, ]
+  allocated_lineups <- sorted_lineup[allocated_lineup_ids, ]
+
[…]

### 2022-09-10 Fifth_experiment/exp_setup.R
allocate exp 5

@@ -38,13 +38,19 @@ lineup <- polynomials %>%
   mutate(exp = "poly") %>%
   bind_rows(heter %>% mutate(exp = "heter")) %>%
   filter((x_dist == "uniform" | b == 0), is.na(b) | b <= 64, e_sigma > 0.25) %>%
-  mutate(lineup_id = ifelse(exp == "poly", lineup_id, lineup_id + 576)) %>%
-  group_by(lineup_id) %>%
-  summarise(across(c(a, b, n, x_dist, shape, e_sigma), first))
+  mutate(real_lineup_id = paste0(exp, "_", lineup_id)) %>%
+  group_by(real_lineup_id) %>%
+  summarise(across(c(a, b, n, x_dist, shape, e_sigma), first)) %>%
+  mutate(lineup_id = 1:n()) %>%
+  mutate(limit_num = ifelse(b != 0 | is.na(b), 6, 15)) %>%
+  mutate(across(everything(), function(x) ifelse(is.na(x), "NA", x)))
+
+sorted_lineup <- lineup %>%
+  arrange(lineup_id) %>%
+  select(-lineup_id, -real_lineup_id, -limit_num)
 
 
 set.seed(10086)
-
 
 check_comb <- function(set_dat, new_dat, n = 1) {
 
@@ -54,15 +60,13 @@ check_comb <- function(set_dat, new_dat, n = 1) {
           # browser()
           tmp <- as.numeric(new_dat[1, this_comb])
           for (i in 1:nrow(set_dat)) {
-            if (all(as.numeric(set_dat[i, this_comb]) == tmp)) return(TRUE)
+            tmp_x <- all(as.numeric(set_dat[i, this_comb]) == tmp)
+            if (is.na(tmp_x)) return(FALSE)
+            if (tmp_x) return(TRUE)
           }
           return(FALSE)
         })
 }
-
-sorted_lineup <- lineup %>%
-  arrange(lineup_id) %>%
-  select(-rep, -lineup_id)
 
 # Caculate distance between a set of allocated lineups and a new lineup
 calc_dist <- function(lineup, allocated_lineup_ids, new_lineup_id) {
@@ -71,12 +75,13 @@ calc_dist <- function(lineup, allocated_lineup_ids, new_lineup_id) {
   allocated_lineups <- sorted_lineup[allocated_lineup_ids, ]
 
   # We want these three options to be as different as possible
-  15 -
+  63 -
[…]

### 2022-09-11 Fifth_experiment/exp_setup.R
finish allocation

@@ -301,33 +301,31 @@ allocate_result <- allocate_result %>%
   select(subject, lineup_id = value)
 
 saveRDS(allocate_result, "tmp_alloc.rds")
+# allocate_result <- readRDS("tmp_alloc.rds")
+
+allocate_result <- allocate_result %>%
+  left_join(select(lineup, lineup_id, real_lineup_id), by = "lineup_id") %>%
+  select(-lineup_id) %>%
+  mutate(subject = as.integer(subject))
 
 set.seed(10086)
 
-lineup <- lineup %>%
-  bind_rows(expand.grid(a = c(0, 1),
-                        n = 300,
-                        x_dist = "uniform",
-                        b = c(128, 256),
-                        rep = 1:3) %>%
-              mutate(lineup_id = nrow(lineup) + 1:n()))
-
-allocate_result <- mutate(allocate_result, subject = as.numeric(subject))
-
-for (i in 1:160) {
+for (i in 1:123) {
   allocate_result <- allocate_result %>%
     bind_rows(tibble(subject = i,
-                     lineup_id = sample(1:12, 2) + 576))
+                     real_lineup_id = paste0("heter_", sample(1:12, 2) + 576)))
 }
 
 allocate_result <- allocate_result %>%
   group_by(subject) %>%
   mutate(order = sample(1:20))
 
+lineup <- select(lineup, -lineup_id)
+
 write_csv(allocate_result, "data/allocate_result.csv")
 write_csv(lineup, "data/lineup_info.csv")
 
 
 allocate_result %>%
-  left_join(lineup, by = c("lineup_id")) %>%
+  left_join(lineup, by = c("real_lineup_id")) %>%
   write_csv("data/allocate_result_full.csv")

### 2022-09-12 Fifth_experiment/exp_setup.R
review allocation

@@ -38,8 +38,8 @@ lineup <- polynomials %>%
   mutate(exp = "poly") %>%
   bind_rows(heter %>% mutate(exp = "heter")) %>%
   filter((x_dist == "uniform" | b == 0), is.na(b) | b <= 64, e_sigma > 0.25) %>%
-  mutate(real_lineup_id = paste0(exp, "_", lineup_id)) %>%
-  group_by(real_lineup_id) %>%
+  mutate(unique_lineup_id = paste0(exp, "_", lineup_id)) %>%
+  group_by(unique_lineup_id) %>%
   summarise(across(c(a, b, n, x_dist, shape, e_sigma), first)) %>%
   mutate(lineup_id = 1:n()) %>%
   mutate(limit_num = ifelse(b != 0 | is.na(b), 6, 15)) %>%
@@ -47,7 +47,7 @@ lineup <- polynomials %>%
 
 sorted_lineup <- lineup %>%
   arrange(lineup_id) %>%
-  select(-lineup_id, -real_lineup_id, -limit_num)
+  select(-lineup_id, -unique_lineup_id, -limit_num)
 
 
 set.seed(10086)
@@ -301,10 +301,11 @@ allocate_result <- allocate_result %>%
   select(subject, lineup_id = value)
 
 saveRDS(allocate_result, "tmp_alloc.rds")
+saveRDS(lineup, "lineup.rds")
 # allocate_result <- readRDS("tmp_alloc.rds")
 
 allocate_result <- allocate_result %>%
-  left_join(select(lineup, lineup_id, real_lineup_id), by = "lineup_id") %>%
+  left_join(select(lineup, lineup_id, unique_lineup_id), by = "lineup_id") %>%
   select(-lineup_id) %>%
   mutate(subject = as.integer(subject))
 
@@ -313,12 +314,13 @@ set.seed(10086)
 for (i in 1:123) {
   allocate_result <- allocate_result %>%
     bind_rows(tibble(subject = i,
-                     real_lineup_id = paste0("heter_", sample(1:12, 2) + 576)))
+                     unique_lineup_id = paste0("heter_", sample(1:12, 2) + 576)))
 }
 
 allocate_result <- allocate_result %>%
   group_by(subject) %>%
-  mutate(order = sample(1:20))
+  mutate(order = sample(1:20)) %>%
+  ungroup()
 
 lineup <- select(lineup, -lineup_id)
 
[…]

### 2022-09-12 paper_comparison/paper_comparison/paper_comparison.Rmd
fix table in 2.1

@@
 Depending on the alternative hypothesis that is focused on, a variety of tests can be applied.
-For example, the presence of heteroskedasticity can usually be tested by applying the White test [\@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [\@breusch_simple_1979], which are both derived from the Lagrange multiplier test [\@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution.
-For testing non-linearity, one may apply the F-test to examine the significance of specific polynomial and non-linear forms of the regressors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [\@ramsey_tests_1969].
-And for testing normality, the shapiro--Wilk test [ref here] is perhaps the most widely used test reported by many of the statistical softwares.
+For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution.
+For testing non-linearity, one may apply the F-test to examine the significance of specific polynomial and non-linear forms of the regressors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969].
+And for testing normality, the Shapiro--Wilk test [ref here] is perhaps the most widely used test included by many of the statistical softwares.
 Another choice will be the Jarque--Bera test [ref here] which directly checks if the sample skewness and kurtosis match a normal distribution.
+
+Example residual plots given in \ref{fig:commondepartures} are examined by the corresponding F-test, Breusch-Pagan test and Shapiro--Wilk test as shown in \ref{tab:example-residual-plot-table}.
+In the example, both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null hypothesis for departures they do not intend to examine.
+As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are sensitive to other types of departures.
+It is likely the null hypothesis is correctly rejected but for the wrong reason, which is known as the "Type III error".
+Additionally, outliers will often incorrectly trigger the rejection of the null hypothesis despite the residuals are well-behaved [@cook_applied_1999].
+This can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers.
 
@@
 dat <- mod$gen(300)
-patchwork::wrap_plots(mod$plot(dat, theme = theme_light()) + ggtitle("Good residuals"),
-                      mod$plot(mod$
-                                 set_prm("include_z", TRUE)$
-                                 gen(300, computed = select(dat, x, e)), theme = theme_light()) + 
-                        ggtitle(paste("F-test p-value =", 
-                                      mod$
-                                        test(mod$
-                                               gen(300, computed = select(dat, x, e)))$p_value %>% 
-                                        format(digits = 2))),
-                      mod$
-                        plot(heter_model(b = 64)$
-                               gen(300, computed = select(dat, x, e)), theme = theme_light()) + 
-                        ggtitle(paste("BP test p-value =", 
-                                      HETER_MODEL$
-                                        test(heter_model(b = 64)$
-                                               gen(300, computed = select(dat, x, e)))$p_value %>% 
-                                        format(digits = 2))),
-                      mod$
-                        plot(heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
-                               gen(300, computed = select(dat, x)), theme = theme_light()) + 
-                        ggtitle(paste("Non-normality", 
-                                      shapiro.test(heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
-                                                     gen(300, computed = select(dat, x))$.resid)[[2]] %>% 
-                                        format(digits = 2))))
-
-```
-
-Explain the tests you are showing later
-
--   RESET
--   BP
--   SW
-
-and show the results for the residual plots displayed in previous section
-
-Briefly mention any others
+
+dat1 <- mod$set_prm("include_z", TRUE)$gen(300, computed = select(dat, x, e))
+dat2 <- heter_model(b = 64)$gen(300, computed = select(dat, x, e))
+dat3 <- heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
+               gen(300, computed = select(dat, x))
+
+data.frame(plot = c("B", "C", "D"), 
+           model = c("Non-linearity", "Heteroskedasticity", "Non-normality"),
+           f = c(POLY_MODEL$test(dat1)$p_value,
+                 POLY_MODEL$test(mutate(dat2, z = poly_model()$gen(300, computed = select(dat2, x))$z))$p_value,
+                 POLY_MODEL$test(mutate(dat3, z = poly_model()$gen(300, computed = select(dat3, x))$z))$p_value),
+           b = c(HETER_MODEL$test(dat1)$p_value,
+                 HETER_MODEL$test(dat2)$p_value,
+                 HETER_MODEL$test(dat3)$p_value),
+           s = c(shapiro.test(dat1$.resid)$p.value,
+                 shapiro.test(dat2$.resid)$p.value,
+                 shapiro.test(dat3$.resid)$p.value)) %>%
+  mutate(f = format(f, digits = 2), b = format(b, digits = 2), s = format(s, digits = 2)) -> table_dat
+
+  table_dat$f <- kableExtra::cell_spec(table_dat$f, color = ifelse(as.numeric(table_dat$f) < 0.05, "red", "black"))
+  table_dat$b <- kableExtra::cell_spec(table_dat$b, color = ifelse(as.numeric(table_dat$b) < 0.05, "red", "black"))
+  table_dat$s <- kableExtra::cell_spec(table_dat$s, color = ifelse(as.numeric(table_dat$s) < 0.05, "red", "black"))
+
+  table_dat %>%
+  knitr::kable(col.names = c("Residual plots", "Departures", "F-test ", "Breusch-Pagan test", "Shapiro–Wilk test"),
+               escape = FALSE,
+               format = "latex",
+               booktabs = TRUE, 
+               caption  = 'Example residual plots examined by the corresponding conventional tests. P-values smaller than 0.05 are colored in red.', 
+               label = "example-residual-plot-table" )
+
+```
+
+<!-- Explain the tests you are showing later -->
+
+<!-- -   RESET -->
+
+<!-- -   BP -->
+
+<!-- -   SW -->
+
+<!-- and show the results for the residual plots displayed in previous section -->
+
+<!-- Briefly mention any others  -->
 
@@
 
-Unlike hypothesis testing built upon rigorous statistical procedures, reading diagnostic plots relies on graphical perception - human's ability to interpret and decode the information embedded in the graph [\@cleveland_graphical_1984], which is to some extent subjective and indecisive.
+Unlike hypothesis testing built upon rigorous statistical procedures, reading diagnostic plots relies on graphical perception - human's ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984], which is to some extent subjective and indecisive.
 Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
-One such example is finding an over-interpretation of the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurrence [\@roy_chowdhury_using_2015].
+One such example is finding an over-interpretation of the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurrence [@roy_chowdhury_using_2015].
 

### 2022-09-12 paper_comparison/paper_comparison/paper_comparison.tex
fix table in 2.1

@@
 tested by applying the White test
-{[}@white\_heteroskedasticity-consistent\_1980{]} or the Breusch-Pagan
-test {[}@breusch\_simple\_1979{]}, which are both derived from the
-Lagrange multiplier test {[}@silvey1959lagrangian{]} principle that
+\citep{white_heteroskedasticity-consistent_1980} or the Breusch-Pagan
+test \citep{breusch_simple_1979}, which are both derived from the
+Lagrange multiplier test \citep{silvey1959lagrangian} principle that
 relies on the asymptotic properties of the null distribution. For
@@
 Regression Equation Specification Error Test (RESET)
-{[}@ramsey\_tests\_1969{]}. And for testing normality, the shapiro--Wilk
-test {[}ref here{]} is perhaps the most widely used test reported by
+\citep{ramsey_tests_1969}. And for testing normality, the Shapiro--Wilk
+test {[}ref here{]} is perhaps the most widely used test included by
 many of the statistical softwares. Another choice will be the
@@
 
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-6-1}
-
-Explain the tests you are showing later
-
-\begin{itemize}
-\tightlist
-\item
-  RESET
-\item
-  BP
-\item
-  SW
-\end{itemize}
-
-and show the results for the residual plots displayed in previous
-section
-
-Briefly mention any others
+Example residual plots given in \ref{fig:commondepartures} are examined
+by the corresponding F-test, Breusch-Pagan test and Shapiro--Wilk test
+as shown in \ref{tab:example-residual-plot-table}. In the example, both
+the Breusch-Pagan test and the Shapiro--Wilk test rejects the null
+hypothesis for departures they do not intend to examine. As discussed in
+\citet{cook1982residuals}, most residual-based tests for a particular
+type of departure from model assumptions are sensitive to other types of
+departures. It is likely the null hypothesis is correctly rejected but
+for the wrong reason, which is known as the ``Type III error''.
+Additionally, outliers will often incorrectly trigger the rejection of
+the null hypothesis despite the residuals are well-behaved
+\citep{cook_applied_1999}. This can be largely avoided in diagnostic
+plots as experienced analysts can evaluate the acceptability of
+assumptions flexibly, even in the presence of outliers.
+
+\begin{table}
+
+\caption{\label{tab:example-residual-plot-table}Example residual plots examined by the corresponding conventional tests. P-values smaller than 0.05 are colored in red.}
+\centering
+\begin{tabular}[t]{lllll}
+\toprule
+Residual plots & Departures & F-test  & Breusch-Pagan test & Shapiro–Wilk test\\
+\midrule
+B & Non-linearity & \textcolor{red}{7.3e-117} & \textcolor{red}{1.8e-14} & \textcolor{red}{3.9e-02}\\
+C & Heteroskedasticity & \textcolor{black}{ 3.8e-01} & \textcolor{red}{6.1e-16} & \textcolor{red}{1.1e-09}\\
+D & Non-normality & \textcolor{black}{ 6.7e-01} & \textcolor{black}{7.4e-01} & \textcolor{red}{2.9e-18}\\
+\bottomrule
+\end{tabular}
+\end{table}
 

### 2022-09-13 Fifth_experiment/exp_setup.R
review exp 5 setup

@@ -341,9 +341,6 @@ allocate_result %>%
   write_csv(file = "data/fifth_experiment_order.txt", col_names = FALSE)
 
 
-dir.create("lineup_plots")
-
-
 for (unique_lineup_id in unique(allocate_result$unique_lineup_id)) {
   lineup_type <- gsub("_.*", "", unique_lineup_id)
   lineup_id <- gsub(".*_", "", unique_lineup_id)
@@ -352,6 +349,11 @@ for (unique_lineup_id in unique(allocate_result$unique_lineup_id)) {
 }
 
 for (i in 1:nrow(allocate_result)) {
-
+  lineup_type <- gsub("_.*", "", allocate_result$unique_lineup_id[i])
+  lineup_id <- gsub(".*_", "", allocate_result$unique_lineup_id[i])
+  subject <- allocate_result$subject[i]
+  order <- allocate_result$order[i]
+  file.copy(glue::glue("lineup_plots__{lineup_type}/{lineup_id}.png"),
+            glue::glue("plots/{subject}_{order}_{lineup_type}.png"))
 }
 

### 2022-09-13 paper_comparison/paper_comparison/paper_comparison.Rmd
update visual inference

@@
 Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals.
-As shown in Figure [ref here], the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns.
+As shown in Figure \\ref{fig:commondepartures}, the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns.
 There are various types of departures from a good residual plot.
@@
 The usual pattern of heterskedasticity on a residual plot is the inconsistent spread of the residuals at different x values.
-Visually, it sometimes results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure [ref here], or the "left-triangle" and "right-triganle" shape where the smallest variance occurs at the edges of the x-axis.
+Visually, it sometimes results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure \\ref{fig:commondepartures}, or the "left-triangle" and "right-triganle" shape where the smallest variance occurs at the edges of the x-axis.
 
@@
 If this assumption happens to be false, then it is expected to see data points do not distribute noramlly on the y-axis.
-For example, given a skewed error distribution, one will see fewer data points and more outliers on one side of the zeroline as shown in the bottom-right of Figure [ref here].
+For example, given a skewed error distribution, one will see fewer data points and more outliers on one side of the zeroline as shown in the bottom-right of Figure \ref{fig:commondepartures}.
 
@@
 
-Example residual plots given in \ref{fig:commondepartures} are examined by the corresponding F-test, Breusch-Pagan test and Shapiro--Wilk test as shown in \ref{tab:example-residual-plot-table}.
+Example residual plots given in Figure \ref{fig:commondepartures} are examined by the corresponding F-test, Breusch-Pagan test and Shapiro--Wilk test as shown in Table \ref{tab:example-residual-plot-table}.
 In the example, both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null hypothesis for departures they do not intend to examine.
@@
 
--   Very briefly explain what a lineup test is
-
--   Lineup (maybe using your previous residual plot for nonlinearity)
-
--   $p$-value calulcation, briefly, including how multiple selections are handled
-
--   power calculation, briefly, including how multiple selections are handled
-
 Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualization" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots [@gelman_exploratory_2004].
+Later, @buja_statistical_2009 proposed the lineup protocol as a visual test inspired by the "police lineup" or "identity parade" which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people.
+The protocol consists of $m$ randomly placed plots, where one plot is the actual data plot, and the remaining $m - 1$ plots have the identical graphical production as the data plot except the data has been replaced with data consistent with the null hypothesis.
+Then, an observer who have not seen the actual data plot will be asked to point out the most different plot from the lineup.
+Figure \ref{fig:first-example-lineup} is an example of a lineup protocol.
+If the actual data plot at position 11 is identifiable, then it is evidence for the rejection of the null hypothesis that the regression model is correctly specified.
+In fact, the example lineup is generated from a misspecified regression model with non-linearity issue.
+So one may notice the cubic curve in the actual residual plot.
+
+Simulation of the null data used in the 19 null plots are
+
+```{r first-example-lineup, fig.cap="Example lineup consists of 20 randomly placed residual plots, where 19 are null plots with data consistent with the null hypothesis, and one is the actual residual plot.
+Can you find the most different plot?"}
+mod <- poly_model(include_z = TRUE, sigma = 1)
+POLY_MODEL$plot_lineup(mod$gen_lineup(300, pos = 11), theme = theme_light(), remove_axis = TRUE, remove_grid_line = TRUE)
+```
+
+<!-- -   Very briefly explain what a lineup test is -->
+
+<!-- -   Lineup (maybe using your previous residual plot for nonlinearity) -->
+
+<!-- -   $p$-value calulcation, briefly, including how multiple selections are handled -->
+
+<!-- -   power calculation, briefly, including how multiple selections are handled -->
+
+<!-- Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualization" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots [@gelman_exploratory_2004].
 Later, in the Bayesian context, data plots was systematically considered as model diagnostics by taking advantage of the data simulated from the assumed statistical models [@gelman_bayesian_2003; @gelman_exploratory_2004].
-
-It was surprising that the essential components of visual inference had actually been established in @buja_inference_1999, but it was not until 10 years later that @buja_statistical_2009 formalized it as an inferential framework to extend confirmatory statistics to visual discoveries.
+-->
+
+<!-- It was surprising that the essential components of visual inference had actually been established in @buja_inference_1999, but it was not until 10 years later that @buja_statistical_2009 formalized it as an inferential framework to extend confirmatory statistics to visual discoveries.
 This framework redefines the test statistics, tests, null distribution, significance levels and $p$-value for visual discovery modelled on the confirmatory statistical testing.
 Figure \ref{fig:parallelism} outlines the parallelism between conventional tests and visual discovery.
+-->
 
@@
 ```
-In visual inference, a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$ is defined, where $\boldsymbol{\mathrm{y}}$ is the data and $I$ is a set of all possible visual features.
+
+<!-- In visual inference, a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$ is defined, where $\boldsymbol{\mathrm{y}}$ is the data and $I$ is a set of all possible visual features.
 @buja_statistical_2009 described each of the test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})$ as a measurement of the degree of presence of a visual feature.
@@
 The same regression model can be rejected by many reasons with residual plot, including non-linearity and heteroskedasticity as shown in Figure \ref{fig:residual-plot-cubic-heter}.
+-->
 
@@
 Both visual discoveries are evidence against the null hypothesis, though heteroskedasticity actually does not exist in the data generating process.
-\\label{fig:residual-plot-cubic-heter}'}
+\\label{fig:residual-plot-cubic-heter}', eval = FALSE}
 
@@
 
-### Sampling from the null distribution {#se:sampling-from-null}
-
-The null distribution of plots refers to the infinite collection of plots of null datasets sampled from $H_0$.
+<!-- ### Sampling from the null distribution {#se:sampling-from-null} -->
+
+<!-- The null distribution of plots refers to the infinite collection of plots of null datasets sampled from $H_0$.
 It is defined as the analogue of the null distribution of test statistics in conventional test [@buja_statistical_2009].
@@
 Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling.
-
-The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009].
+-->
+
+<!-- The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009].
 Suppose there exists a minimal sufficient statistic $\boldsymbol{S}(\boldsymbol{y})$ under the null hypothesis, any null datasets $\boldsymbol{y^{*}}$ should fulfil the condition $\boldsymbol{S}(\boldsymbol{y}) = \boldsymbol{s}$.
@@
 With this property, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the regressors, then rescaling it by the ratio of residual sum of square in two regressions.
+-->
 

### 2022-09-13 paper_comparison/paper_comparison/paper_comparison.tex
update visual inference

@@
 orthogonal to residuals are referred to as standard residual plots in
-\citet{cook1982residuals}. As shown in Figure {[}ref here{]}, the
-top-left panel is a good residual plot with residuals evenly distributed
-at both sides of the horizontal zero line showing no noticeable
-patterns. There are various types of departures from a good residual
-plot. We will only discuss three most commonly checked departures in
-this paper. Namely, non-linearity, heterskedasticity and non-normality.
+\citet{cook1982residuals}. As shown in Figure
+\textbackslash ref\{fig:commondepartures\}, the top-left panel is a good
+residual plot with residuals evenly distributed at both sides of the
+horizontal zero line showing no noticeable patterns. There are various
+types of departures from a good residual plot. We will only discuss
+three most commonly checked departures in this paper. Namely,
+non-linearity, heterskedasticity and non-normality.
 
@@
 the so-called ``butterfly'' shape as shown in the bottom-left panel of
-Figure {[}ref here{]}, or the ``left-triangle'' and ``right-triganle''
-shape where the smallest variance occurs at the edges of the x-axis.
+Figure \textbackslash ref\{fig:commondepartures\}, or the
+``left-triangle'' and ``right-triganle'' shape where the smallest
+variance occurs at the edges of the x-axis.
 
@@
 fewer data points and more outliers on one side of the zeroline as shown
-in the bottom-right of Figure {[}ref here{]}.
+in the bottom-right of Figure
+\textbackslash ref\{fig:commondepartures\}.
 
@@
 
-Example residual plots given in \ref{fig:commondepartures} are examined
-by the corresponding F-test, Breusch-Pagan test and Shapiro--Wilk test
-as shown in \ref{tab:example-residual-plot-table}. In the example, both
-the Breusch-Pagan test and the Shapiro--Wilk test rejects the null
+Example residual plots given in Figure \ref{fig:commondepartures} are
+examined by the corresponding F-test, Breusch-Pagan test and
+Shapiro--Wilk test as shown in Table
+\ref{tab:example-residual-plot-table}. In the example, both the
+Breusch-Pagan test and the Shapiro--Wilk test rejects the null
 hypothesis for departures they do not intend to examine. As discussed in
@@
 ability to interpret and decode the information embedded in the graph
-{[}@cleveland\_graphical\_1984{]}, which is to some extent subjective
-and indecisive. Further, visual discovery suffers from its unsecured and
+\citep{cleveland_graphical_1984}, which is to some extent subjective and
+indecisive. Further, visual discovery suffers from its unsecured and
 unconfirmed nature where the degree of the presence of the visual
@@
 between the gene groups and separation is not an uncommon occurrence
-{[}@roy\_chowdhury\_using\_2015{]}.
-
-\begin{itemize}
-\item
-  Very briefly explain what a lineup test is
-\item
-  Lineup (maybe using your previous residual plot for nonlinearity)
-\item
-  \(p\)-value calulcation, briefly, including how multiple selections
-  are handled
-\item
-  power calculation, briefly, including how multiple selections are
-  handled
-\end{itemize}
+\citep{roy_chowdhury_using_2015}.
 
@@
 inference for visual discoveries of data plots
-\citep{gelman_exploratory_2004}. Later, in the Bayesian context, data
-plots was systematically considered as model diagnostics by taking
-advantage of the data simulated from the assumed statistical models
-\citep{gelman_bayesian_2003, gelman_exploratory_2004}.
-
-It was surprising that the essential components of visual inference had
-actually been established in \citet{buja_inference_1999}, but it was not
-until 10 years later that \citet{buja_statistical_2009} formalized it as
-an inferential framework to extend confirmatory statistics to visual
-discoveries. This framework redefines the test statistics, tests, null
-distribution, significance levels and \(p\)-value for visual discovery
-modelled on the confirmatory statistical testing. Figure
-\ref{fig:parallelism} outlines the parallelism between conventional
-tests and visual discovery.
-
-In visual inference, a collection of test statistics
-\(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) is defined, where
-\(\boldsymbol{\mathrm{y}}\) is the data and \(I\) is a set of all
-possible visual features. \citet{buja_statistical_2009} described each
-of the test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})\) as a
-measurement of the degree of presence of a visual feature.
-Alternatively, \citet{majumder_validation_2013} avoids the use of visual
-features and defined the visual statistics \(T(.)\) as a mapping from a
-dataset to a data plot. Both definitions of visual test statistics are
-valid, but in the rest of the paper the first definition will be used as
-it covers some details needed by the following discussion. A visual
-discovery is defined as a rejection of a null hypothesis, and the same
-null hypothesis can be rejected by many different visual discoveries
-\citep{buja_statistical_2009}. For regression diagnostics, the null
-hypothesis would be the assumed model, while the visual discoveries
-would be any findings that are inconsistent with the null hypothesis.
-The same regression model can be rejected by many reasons with residual
-plot, including non-linearity and heteroskedasticity as shown in Figure
-\ref{fig:residual-plot-cubic-heter}.
+\citep{gelman_exploratory_2004}. Later, \citet{buja_statistical_2009}
+proposed the lineup protocol as a visual test inspired by the ``police
+lineup'' or ``identity parade'' which is the act of asking the
+eyewitness to identify criminal suspect from a group of irrelevant
+people. The protocol consists of \(m\) randomly placed plots, where one
+plot is the actual data plot, and the remaining \(m - 1\) plots have the
+identical graphical production as the data plot except the data has been
+replaced with data consistent with the null hypothesis. Then, an
+observer who have not seen the actual data plot will be asked to point
+out the most different plot from the lineup. Figure
+\ref{fig:first-example-lineup} is an example of a lineup protocol. If
+the actual data plot at position 11 is identifiable, then it is evidence
+for the rejection of the null hypothesis that the regression model is
+correctly specified. In fact, the example lineup is generated from a
+misspecified regression model with non-linearity issue. So one may
+notice the cubic curve in the actual residual plot.
+
+Simulation of the null data used in the 19 null plots are
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/residual-plot-cubic-heter-1} \caption{Residuals vs.
-fitted values plot for a classical linear regression model.
-The residuals are produced by fitting a two-predictor multiple linear regression model with data generated from a cubic linear model.
-From the residual plot, "butterfly shape" can be observed which generally would be interpretd as evidence of heteroskedasticity.
-Further, from the outline of the shape, nonlinear patterns exist.
-Both visual discoveries are evidence against the null hypothesis, though heteroskedasticity actually does not exist in the data generating process.
-\label{fig:residual-plot-cubic-heter}}\label{fig:residual-plot-cubic-heter}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/first-example-lineup-1} \caption{Example lineup consists of 20 randomly placed residual plots, where 19 are null plots with data consistent with the null hypothesis, and one is the actual residual plot.
+Can you find the most different plot?}\label{fig:first-example-lineup}
 \end{figure}
-
-\hypertarget{se:sampling-from-null}{%
-\subsubsection{Sampling from the null
-distribution}\label{se:sampling-from-null}}
-
-The null distribution of plots refers to the infinite collection of
-plots of null datasets sampled from \(H_0\). It is defined as the
-analogue of the null distribution of test statistics in conventional
-test \citep{buja_statistical_2009}. In practice, a finite number of
-plots of null datasets could be generated, called null plots. In the
-context of regression diagnostics, sampling data from \(H_0\) is
-equivalent to sampling data from the assumed model. As
-\citet{buja_statistical_2009} suggested, \(H_0\) is usually composited
-by a collection of distributions controlled by nuisance parameters.
-Since regression models can have various forms, there is no general
-solution to this problem, but it sometimes can be reduced to so called
-``reference distribution'' by applying one of the three methods: (i)
-sampling from a conditional distribution given a minimal sufficient
-statistic under \(H_0\), (ii) parametric bootstrap sampling with
-nuisance parameters estimated under \(H_0\), and (iii) Bayesian
-posterior predictive sampling.
-
-The conditional distribution given a minimal sufficient statistic is the
-best justified reference distribution among the three
-\citep{buja_statistical_2009}. Suppose there exists a minimal sufficient
-statistic \(\boldsymbol{S}(\boldsymbol{y})\) under the null hypothesis,
-any null datasets \(\boldsymbol{y^{*}}\) should fulfil the condition
-\(\boldsymbol{S}(\boldsymbol{y}) = \boldsymbol{s}\). Using the classical
-normal linear regression model as example, the minimal sufficient
-statistic is
-\(\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{\beta}}, \boldsymbol{e}'\boldsymbol{e})\),
-where \(\hat{\boldsymbol{\beta}}\) are the coefficient estimators and
-\(\boldsymbol{e}'\boldsymbol{e}\) is the residual sum of square.
-Alternatively, the minimal sufficient statistic can be constructed as
-\(\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{y}}, ||\boldsymbol{e}||)\),
-where \(\hat{\boldsymbol{y}}\) are the fitted values and
-\(||\boldsymbol{e}||\) is the length of residuals, which is more
-intuitive as suggested by \citet{buja_statistical_2009}. Since the
-fitted values are held fixed, the variation can only occur in the
-residual space. And because the length of residual is also held fixed,
-residuals obtained from a null dataset has to be a random rotation of
-\(\boldsymbol{e}\) in the residual space. With this property, null
-residuals can be simulated by regressing \(N\) i.i.d standard normal
-random draws on the regressors, then rescaling it by the ratio of
-residual sum of square in two regressions.
 

### 2022-09-13 paper_comparison/paper_comparison/paper_comparison.Rmd
null simulation

@@
 If the actual data plot at position 11 is identifiable, then it is evidence for the rejection of the null hypothesis that the regression model is correctly specified.
-In fact, the example lineup is generated from a misspecified regression model with non-linearity issue.
-So one may notice the cubic curve in the actual residual plot.
-
-Simulation of the null data used in the 19 null plots are
+In fact, the actual residual plot is obtained from a misspecified regression model with non-linearity issue.
+
+Data used in the $m - 1$ null plots needs to be simulated.
+In the context of regression diagnostics, sampling data from $H_0$ is equivalent to sampling data from the assumed model.
+As @buja_statistical_2009 suggested, $H_0$ is usually composited by a collection of distributions controlled by nuisance parameters.
+Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling.
+The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009].
+Essentially, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the regressors, then rescaling it by the ratio of residual sum of square in two regressions.
 

### 2022-09-13 paper_comparison/paper_comparison/paper_comparison.tex
null simulation

@@
 fewer data points and more outliers on one side of the zeroline as shown
-in the bottom-right of Figure
-\textbackslash ref\{fig:commondepartures\}.
+in the bottom-right of Figure \ref{fig:commondepartures}.
 
@@
 for the rejection of the null hypothesis that the regression model is
-correctly specified. In fact, the example lineup is generated from a
-misspecified regression model with non-linearity issue. So one may
-notice the cubic curve in the actual residual plot.
+correctly specified. In fact, the actual residual plot is obtained from
+a misspecified regression model with non-linearity issue.
 
-Simulation of the null data used in the 19 null plots are
+Data used in the \(m - 1\) null plots needs to be simulated. In the
+context of regression diagnostics, sampling data from \(H_0\) is
+equivalent to sampling data from the assumed model. As
+\citet{buja_statistical_2009} suggested, \(H_0\) is usually composited
+by a collection of distributions controlled by nuisance parameters.
+Since regression models can have various forms, there is no general
+solution to this problem, but it sometimes can be reduced to so called
+``reference distribution'' by applying one of the three methods: (i)
+sampling from a conditional distribution given a minimal sufficient
+statistic under \(H_0\), (ii) parametric bootstrap sampling with
+nuisance parameters estimated under \(H_0\), and (iii) Bayesian
+posterior predictive sampling. The conditional distribution given a
+minimal sufficient statistic is the best justified reference
+distribution among the three \citep{buja_statistical_2009}. Essentially,
+null residuals can be simulated by regressing \(N\) i.i.d standard
+normal random draws on the regressors, then rescaling it by the ratio of
+residual sum of square in two regressions.
 

### 2022-09-13 paper_comparison/paper_comparison/paper_comparison.tex
regroup sections

@@
 
+\hypertarget{lineup-protocol}{%
+\subsubsection{Lineup protocol}\label{lineup-protocol}}
+
 Unlike hypothesis testing built upon rigorous statistical procedures,
@@
 observer who have not seen the actual data plot will be asked to point
-out the most different plot from the lineup. Figure
-\ref{fig:first-example-lineup} is an example of a lineup protocol. If
-the actual data plot at position 11 is identifiable, then it is evidence
-for the rejection of the null hypothesis that the regression model is
-correctly specified. In fact, the actual residual plot is obtained from
-a misspecified regression model with non-linearity issue.
-
-Data used in the \(m - 1\) null plots needs to be simulated. In the
+out the most different plot from the lineup. Under the null hypothesis,
+it is expected that the actual data plot would have no distinguishable
+difference with the null plots, and the probability of the observer
+correctly picks the actual data plot is \(1/m\). If we reject the null
+hypothesis as the observer correctly picks the actual data plot, then
+the Type I error of this test is \(1/m\).
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/first-example-lineup-1} \caption{Example lineup consists of 20 randomly placed residual plots, where 19 are null plots with data consistent with the null hypothesis, and one is the actual residual plot.
+Can you find the most different plot?}\label{fig:first-example-lineup}
+\end{figure}
+
+Figure \ref{fig:first-example-lineup} is an example of a lineup
+protocol. If the actual data plot at position 11 is identifiable, then
+it is evidence for the rejection of the null hypothesis that the
+regression model is correctly specified. In fact, the actual residual
+plot is obtained from a misspecified regression model with non-linearity
+issue.
+
+\hypertarget{sampling-from-the-null-distribution}{%
+\subsubsection{Sampling from the null
+distribution}\label{sampling-from-the-null-distribution}}
+
+Data used in the \(m - 1\) null plots need to be simulated. In the
 context of regression diagnostics, sampling data from \(H_0\) is
@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/first-example-lineup-1} \caption{Example lineup consists of 20 randomly placed residual plots, where 19 are null plots with data consistent with the null hypothesis, and one is the actual residual plot.
-Can you find the most different plot?}\label{fig:first-example-lineup}
-\end{figure}
+\hypertarget{p-value-calculation}{%
+\subsubsection{p-value calculation}\label{p-value-calculation}}
+
+Further, a visual test can involve \(K\) independent observers. Let
+\(D_i = \{0,1\}\) be a binomial random variable denoting whether subject
+\(i\) correctly detecting the actual data plot, and
+\(X = \sum_{i=1}^{K}X_i\) be the number of observers correctly picking
+the actual data plot. Then, by imposing a relatively strong assumption
+on the visual test that all \(K\) evaluations are fully independent,
+under the null hypothesis, \(X \sim \mathrm{Binom}_{K,1/m}\). Therefore,
+the \(p\)-value of a lineup of size \(m\) evaluated by \(K\) observer is
+given as \begin{equation} \label{eq:pvaluesingle}
+P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},
+\end{equation}
+
+where \(x\) is the realization of number of observers correctly picking
+the actual data plot \citep{majumder_validation_2013}.
 

### 2022-09-13 paper_comparison/paper_comparison/paper_comparison.Rmd
regroup sections
update p value calc

@@
 Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals.
-As shown in Figure \\ref{fig:commondepartures}, the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns.
+As shown in Figure \ref{fig:commondepartures}, the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns.
 There are various types of departures from a good residual plot.
@@
 
+### Lineup protocol
+
 Unlike hypothesis testing built upon rigorous statistical procedures, reading diagnostic plots relies on graphical perception - human's ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984], which is to some extent subjective and indecisive.
@@
 Then, an observer who have not seen the actual data plot will be asked to point out the most different plot from the lineup.
+Under the null hypothesis, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is $1/m$.
+If we reject the null hypothesis as the observer correctly picks the actual data plot, then the Type I error of this test is $1/m$.
+
+```{r first-example-lineup, fig.cap="An example lineup consists of 20 randomly placed residual plots, where 19 are null plots with data consistent with the null hypothesis, and one is the actual residual plot.
+Can you find the most different plot?"}
+mod <- poly_model(include_z = TRUE, sigma = 1)
+POLY_MODEL$plot_lineup(mod$gen_lineup(300, pos = 11), theme = theme_light(), remove_axis = TRUE, remove_grid_line = TRUE)
+```
+
 Figure \ref{fig:first-example-lineup} is an example of a lineup protocol.
@@
 
-Data used in the $m - 1$ null plots needs to be simulated.
+### Sampling from the null distribution
+
+Data used in the $m - 1$ null plots need to be simulated.
 In the context of regression diagnostics, sampling data from $H_0$ is equivalent to sampling data from the assumed model.
@@
 
-```{r first-example-lineup, fig.cap="Example lineup consists of 20 randomly placed residual plots, where 19 are null plots with data consistent with the null hypothesis, and one is the actual residual plot.
-Can you find the most different plot?"}
-mod <- poly_model(include_z = TRUE, sigma = 1)
-POLY_MODEL$plot_lineup(mod$gen_lineup(300, pos = 11), theme = theme_light(), remove_axis = TRUE, remove_grid_line = TRUE)
-```
-
-<!-- -   Very briefly explain what a lineup test is -->
-
-<!-- -   Lineup (maybe using your previous residual plot for nonlinearity) -->
-
-<!-- -   $p$-value calulcation, briefly, including how multiple selections are handled -->
-
-<!-- -   power calculation, briefly, including how multiple selections are handled -->
-
-<!-- Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualization" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots [@gelman_exploratory_2004].
-Later, in the Bayesian context, data plots was systematically considered as model diagnostics by taking advantage of the data simulated from the assumed statistical models [@gelman_bayesian_2003; @gelman_exploratory_2004].
--->
-
-<!-- It was surprising that the essential components of visual inference had actually been established in @buja_inference_1999, but it was not until 10 years later that @buja_statistical_2009 formalized it as an inferential framework to extend confirmatory statistics to visual discoveries.
-This framework redefines the test statistics, tests, null distribution, significance levels and $p$-value for visual discovery modelled on the confirmatory statistical testing.
-Figure \ref{fig:parallelism} outlines the parallelism between conventional tests and visual discovery.
--->
-
-```{=html}
-<!-- ![Parallelism between multiple quantitative testing and visual discovery [@buja_statistical_2009].
-Visible features in a plot are viewed as a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$, and any visual discoveries that are inconsistent with the null hypothesis are treated as evidence against the null.
-For regression diagnostics, the null hypothesis would be the assumed model, and visual discoveries could be any visual features in favour of any alternatives.
-\label{fig:parallelism}](figures/rsta2009012001.jpg){width=450 height=341}
--->
-```
-
-<!-- In visual inference, a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$ is defined, where $\boldsymbol{\mathrm{y}}$ is the data and $I$ is a set of all possible visual features.
-@buja_statistical_2009 described each of the test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})$ as a measurement of the degree of presence of a visual feature.
-Alternatively, @majumder_validation_2013 avoids the use of visual features and defined the visual statistics $T(.)$ as a mapping from a dataset to a data plot.
-Both definitions of visual test statistics are valid, but in the rest of the paper the first definition will be used as it covers some details needed by the following discussion.
-A visual discovery is defined as a rejection of a null hypothesis, and the same null hypothesis can be rejected by many different visual discoveries [@buja_statistical_2009].
-For regression diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the null hypothesis.
-The same regression model can be rejected by many reasons with residual plot, including non-linearity and heteroskedasticity as shown in Figure \ref{fig:residual-plot-cubic-heter}.
--->
-
-```{r residual-plot-cubic-heter, fig.cap='Residuals vs.
-fitted values plot for a classical linear regression model.
-The residuals are produced by fitting a two-predictor multiple linear regression model with data generated from a cubic linear model.
-From the residual plot, "butterfly shape" can be observed which generally would be interpretd as evidence of heteroskedasticity.
-Further, from the outline of the shape, nonlinear patterns exist.
-Both visual discoveries are evidence against the null hypothesis, though heteroskedasticity actually does not exist in the data generating process.
-\\label{fig:residual-plot-cubic-heter}', eval = FALSE}
-
-cm <- cubic_model(a = 0, b = 100, c = 1)
-
-cm$gen(1000, fit_model = TRUE) %>%
-cm$plot() +
-  theme_light()
-```
-
-<!-- ### Sampling from the null distribution {#se:sampling-from-null} -->
-
-<!-- The null distribution of plots refers to the infinite collection of plots of null datasets sampled from $H_0$.
-It is defined as the analogue of the null distribution of test statistics in conventional test [@buja_statistical_2009].
-In practice, a finite number of plots of null datasets could be generated, called null plots.
-In the context of regression diagnostics, sampling data from $H_0$ is equivalent to sampling data from the assumed model.
-As @buja_statistical_2009 suggested, $H_0$ is usually composited by a collection of distributions controlled by nuisance parameters.
-Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling.
--->
-
-<!-- The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009].
-Suppose there exists a minimal sufficient statistic $\boldsymbol{S}(\boldsymbol{y})$ under the null hypothesis, any null datasets $\boldsymbol{y^{*}}$ should fulfil the condition $\boldsymbol{S}(\boldsymbol{y}) = \boldsymbol{s}$.
-Using the classical normal linear regression model as example, the minimal sufficient statistic is $\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{\beta}}, \boldsymbol{e}'\boldsymbol{e})$, where $\hat{\boldsymbol{\beta}}$ are the coefficient estimators and $\boldsymbol{e}'\boldsymbol{e}$ is the residual sum of square.
-Alternatively, the minimal sufficient statistic can be constructed as $\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{y}}, ||\boldsymbol{e}||)$, where $\hat{\boldsymbol{y}}$ are the fitted values and $||\boldsymbol{e}||$ is the length of residuals, which is more intuitive as suggested by @buja_statistical_2009.
-Since the fitted values are held fixed, the variation can only occur in the residual space.
-And because the length of residual is also held fixed, residuals obtained from a null dataset has to be a random rotation of $\boldsymbol{e}$ in the residual space.
-With this property, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the regressors, then rescaling it by the ratio of residual sum of square in two regressions.
--->
-
-### Lineup protocol {#se:lineup}
-
-With the simulation mechanism of null plots being provided, another aspect of hypothesis testing that needs to be addressed is the control of false positive rate or Type I error.
-Any visual statistic $T^{(i)}(\boldsymbol{\mathrm{y}})$ needs to pair with a critical value $c^{(i)}$ to form a hypothesis test.
-When a visual feature $i$ is discovered by the observer from a plot, the corresponding visual statistic $T^{(i)}(\boldsymbol{\mathrm{y}})$ may not be known as there is no general agreement on the measurement of the degree of presence of a visual feature.
-It is only the event that $T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}$ is confirmed.
-Similarly, if any visual discovery is found by the observer, we say, there exists $i \in I:~T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}$ [@buja_statistical_2009].
-
-Using the above definition, the family-wise Type I error can be controlled if one can provide the collection of critical values $c^{(i)}~(i \in I)$ such that $P(\mathrm{there~exists~} i \in I: T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}|\boldsymbol{\mathrm{y}}) \leq \alpha$, where $\alpha$ is the significance level.
-However, since the quantity of $T^{(i)}(\boldsymbol{\mathrm{y}})$ may not be known, such collection of critical values can not be provided.
-
-@buja_statistical_2009 proposed the lineup protocol as a visual test to calibrate the Type I error issue without the specification of $c^{(i)}~(i \in I)$.
-It is inspired by the "police lineup" or "identity parade" which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people.
-The protocol consists of $m$ randomly placed plots, where one plot is the actual data plot, and the remaining $m - 1$ plots have the identical graphical production as the data plot except the data has been replaced with data consistent with the null hypothesis.
-Then, an observer who have not seen the actual data plot will be asked to point out the most different plot from the lineup.
-
-Under the null hypothesis, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is $1/m$.
-If we reject the null hypothesis as the observer correctly picks the actual data plot, then the Type I error of this test is $1/m$.
-
-This provides us with an mechanism to control the Type I error, because $m$ - the number of plots in a lineup can be chosen.
-A larger value of $m$ will result in a smaller Type I error, but the limit to the value of $m$ depends on the number of plots a human is willing to view [@buja_statistical_2009].
-Typically, $m$ will be set to $20$ which is equivalent to set $\alpha = 0.05$, a general choice of significance level for conventional testing among statisticians.
+### p-value calculation
 
@@
 where $x$ is the realization of number of observers correctly picking the actual data plot [@majumder_validation_2013].
-
-The multiple individuals approach avoids the limit of $m$, while provides visual tests with $p$-value much smaller than $0.05$.
-In fact, the lower bound of $p$-value decreases exponentially as $K$ increases.
-With just $4$ individuals and $20$ data plots in a lineup, the $p$-value could be as small as $0.0001$.
-Additionally, by involving multiple observers, variation of individual ability to read plots can be addressed to some degree as different opinions about visual discoveries can be collected.
 
@@
 This ensured $\sum_{i}c_i=K$.
-
-The full model was a Dirichlet-multinomial mixture distribution
-
-```{=tex}
-\begin{align} \label{eq:dirichlet-multinomial}\begin{split}
-\boldsymbol{\theta}&|\alpha \sim Dirichlet(\alpha)\\
-(c_i,...,c_m)&|\boldsymbol{\theta} \sim Multinomial(K, \boldsymbol{\theta}).
-\end{split}\end{align}
-```
-Since the p-value calculation only needs to concern the number of times the actual data plot being selected denoted by $C_i$, they showed the model can be simplified to a beta-binomial mixture distribution
-
-```{=tex}
-\begin{align} \label{eq:beta-binomial}\begin{split}
-\theta_i&|\alpha \sim Beta(\alpha, (m-1)\alpha)\\
-C_i&|\theta_i \sim Binomial(K, \theta_i).
-\end{split}\end{align}
-```
-Thus, the visual p-value followed by the beta-binomial model is given as
+Thus, the visual p-value followed by the above model is given as
 
@@
 ```
+
 where $B(.)$ is the beta function defined as
@@
 ```
-Note that the use of Equation (\ref{eq:pvalue-beta-binomial}) requires the estimation of $\hat{\alpha}$, which largely depends on the null model, the type of the plot and other aesthetic features.
+
+<!-- -   Very briefly explain what a lineup test is -->
+
+<!-- -   Lineup (maybe using your previous residual plot for nonlinearity) -->
+
+<!-- -   $p$-value calulcation, briefly, including how multiple selections are handled -->
+
+<!-- -   power calculation, briefly, including how multiple selections are handled -->
+
+<!-- Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualization" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots [@gelman_exploratory_2004].
+Later, in the Bayesian context, data plots was systematically considered as model diagnostics by taking advantage of the data simulated from the assumed statistical models [@gelman_bayesian_2003; @gelman_exploratory_2004].
+-->
+
+<!-- It was surprising that the essential components of visual inference had actually been established in @buja_inference_1999, but it was not until 10 years later that @buja_statistical_2009 formalized it as an inferential framework to extend confirmatory statistics to visual discoveries.
+This framework redefines the test statistics, tests, null distribution, significance levels and $p$-value for visual discovery modelled on the confirmatory statistical testing.
+Figure \ref{fig:parallelism} outlines the parallelism between conventional tests and visual discovery.
+-->
+
+```{=html}
+<!-- ![Parallelism between multiple quantitative testing and visual discovery [@buja_statistical_2009].
+Visible features in a plot are viewed as a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$, and any visual discoveries that are inconsistent with the null hypothesis are treated as evidence against the null.
+For regression diagnostics, the null hypothesis would be the assumed model, and visual discoveries could be any visual features in favour of any alternatives.
+\label{fig:parallelism}](figures/rsta2009012001.jpg){width=450 height=341}
+-->
+```
+
+<!-- In visual inference, a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$ is defined, where $\boldsymbol{\mathrm{y}}$ is the data and $I$ is a set of all possible visual features.
+@buja_statistical_2009 described each of the test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})$ as a measurement of the degree of presence of a visual feature.
+Alternatively, @majumder_validation_2013 avoids the use of visual features and defined the visual statistics $T(.)$ as a mapping from a dataset to a data plot.
+Both definitions of visual test statistics are valid, but in the rest of the paper the first definition will be used as it covers some details needed by the following discussion.
+A visual discovery is defined as a rejection of a null hypothesis, and the same null hypothesis can be rejected by many different visual discoveries [@buja_statistical_2009].
+For regression diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the null hypothesis.
+The same regression model can be rejected by many reasons with residual plot, including non-linearity and heteroskedasticity as shown in Figure \ref{fig:residual-plot-cubic-heter}.
+-->
+
+```{r residual-plot-cubic-heter, fig.cap='Residuals vs.
+fitted values plot for a classical linear regression model.
+The residuals are produced by fitting a two-predictor multiple linear regression model with data generated from a cubic linear model.
+From the residual plot, "butterfly shape" can be observed which generally would be interpretd as evidence of heteroskedasticity.
+Further, from the outline of the shape, nonlinear patterns exist.
+Both visual discoveries are evidence against the null hypothesis, though heteroskedasticity actually does not exist in the data generating process.
+\\label{fig:residual-plot-cubic-heter}', eval = FALSE}
+
+cm <- cubic_model(a = 0, b = 100, c = 1)
+
+cm$gen(1000, fit_model = TRUE) %>%
+cm$plot() +
+  theme_light()
+```
+
+<!-- ### Sampling from the null distribution {#se:sampling-from-null} -->
+
+<!-- The null distribution of plots refers to the infinite collection of plots of null datasets sampled from $H_0$.
+It is defined as the analogue of the null distribution of test statistics in conventional test [@buja_statistical_2009].
+In practice, a finite number of plots of null datasets could be generated, called null plots.
+In the context of regression diagnostics, sampling data from $H_0$ is equivalent to sampling data from the assumed model.
+As @buja_statistical_2009 suggested, $H_0$ is usually composited by a collection of distributions controlled by nuisance parameters.
+Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling.
+-->
+
+<!-- The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009].
+Suppose there exists a minimal sufficient statistic $\boldsymbol{S}(\boldsymbol{y})$ under the null hypothesis, any null datasets $\boldsymbol{y^{*}}$ should fulfil the condition $\boldsymbol{S}(\boldsymbol{y}) = \boldsymbol{s}$.
+Using the classical normal linear regression model as example, the minimal sufficient statistic is $\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{\beta}}, \boldsymbol{e}'\boldsymbol{e})$, where $\hat{\boldsymbol{\beta}}$ are the coefficient estimators and $\boldsymbol{e}'\boldsymbol{e}$ is the residual sum of square.
+Alternatively, the minimal sufficient statistic can be constructed as $\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{y}}, ||\boldsymbol{e}||)$, where $\hat{\boldsymbol{y}}$ are the fitted values and $||\boldsymbol{e}||$ is the length of residuals, which is more intuitive as suggested by @buja_statistical_2009.
+Since the fitted values are held fixed, the variation can only occur in the residual space.
+And because the length of residual is also held fixed, residuals obtained from a null dataset has to be a random rotation of $\boldsymbol{e}$ in the residual space.
+With this property, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the regressors, then rescaling it by the ratio of residual sum of square in two regressions.
+-->
+
+<!-- ### Lineup protocol {#se:lineup} -->
+
+<!-- With the simulation mechanism of null plots being provided, another aspect of hypothesis testing that needs to be addressed is the control of false positive rate or Type I error.
+Any visual statistic $T^{(i)}(\boldsymbol{\mathrm{y}})$ needs to pair with a critical value $c^{(i)}$ to form a hypothesis test.
+When a visual feature $i$ is discovered by the observer from a plot, the corresponding visual statistic $T^{(i)}(\boldsymbol{\mathrm{y}})$ may not be known as there is no general agreement on the measurement of the degree of presence of a visual feature.
+It is only the event that $T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}$ is confirmed.
+Similarly, if any visual discovery is found by the observer, we say, there exists $i \in I:~T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}$ [@buja_statistical_2009].
+-->
+
+<!-- Using the above definition, the family-wise Type I error can be controlled if one can provide the collection of critical values $c^{(i)}~(i \in I)$ such that $P(\mathrm{there~exists~} i \in I: T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}|\boldsymbol{\mathrm{y}}) \leq \alpha$, where $\alpha$ is the significance level.
+However, since the quantity of $T^{(i)}(\boldsymbol{\mathrm{y}})$ may not be known, such collection of critical values can not be provided.
+-->
+
+<!-- @buja_statistical_2009 proposed the lineup protocol as a visual test to calibrate the Type I error issue without the specification of $c^{(i)}~(i \in I)$.
+It is inspired by the "police lineup" or "identity parade" which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people.
+The protocol consists of $m$ randomly placed plots, where one plot is the actual data plot, and the remaining $m - 1$ plots have the identical graphical production as the data plot except the data has been replaced with data consistent with the null hypothesis.
+Then, an observer who have not seen the actual data plot will be asked to point out the most different plot from the lineup.
+-->
+
+<!-- Under the null hypothesis, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is $1/m$.
+If we reject the null hypothesis as the observer correctly picks the actual data plot, then the Type I error of this test is $1/m$.
+-->
+
+<!-- This provides us with an mechanism to control the Type I error, because $m$ - the number of plots in a lineup can be chosen.
+A larger value of $m$ will result in a smaller Type I error, but the limit to the value of $m$ depends on the number of plots a human is willing to view [@buja_statistical_2009].
+Typically, $m$ will be set to $20$ which is equivalent to set $\alpha = 0.05$, a general choice of significance level for conventional testing among statisticians.
+-->
+
+<!-- Further, a visual test can involve $K$ independent observers.
+Let $D_i = \{0,1\}$ be a binomial random variable denoting whether subject $i$ correctly detecting the actual data plot, and $X = \sum_{i=1}^{K}X_i$ be the number of observers correctly picking the actual data plot.
+Then, by imposing a relatively strong assumption on the visual test that all $K$ evaluations are fully independent, under the null hypothesis, $X \sim \mathrm{Binom}_{K,1/m}$.
+Therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as \begin{equation} \label{eq:pvaluesingle} -->
+<!-- P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i}, -->
+<!-- \end{equation} -->
+
+<!-- where $x$ is the realization of number of observers correctly picking the actual data plot [@majumder_validation_2013]. -->
+
+<!-- The multiple individuals approach avoids the limit of $m$, while provides visual tests with $p$-value much smaller than $0.05$.
+In fact, the lower bound of $p$-value decreases exponentially as $K$ increases.
+With just $4$ individuals and $20$ data plots in a lineup, the $p$-value could be as small as $0.0001$.
+Additionally, by involving multiple observers, variation of individual ability to read plots can be addressed to some degree as different opinions about visual discoveries can be collected.
+-->
+
+<!-- As pointed out by @vanderplas2021statistical, though Equation (\ref{eq:pvaluesingle}) is trivial, but it doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup.
+And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup.
+They summarized three common different scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same actual data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects.
+Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experimental design.
+For Scenario 3, @vanderplas2021statistical modelled the probability of a plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$.
+And defined $c_i$ to be the number of times plot $i$ being selected in $K$ evaluations.
+In case subject $j$ makes multiple selections, they decided to add $1/s_j$ to $c_i$ instead of one, where $s_j$ is the number of plots subject $j$ selected for $j=1,...K$.
+This ensured $\sum_{i}c_i=K$.
+-->
+
+<!-- The full model was a Dirichlet-multinomial mixture distribution -->
+
+<!-- ```{=tex} -->
+<!-- \begin{align} \label{eq:dirichlet-multinomial}\begin{split} -->
+<!-- \boldsymbol{\theta}&|\alpha \sim Dirichlet(\alpha)\\ -->
+<!-- (c_i,...,c_m)&|\boldsymbol{\theta} \sim Multinomial(K, \boldsymbol{\theta}). -->
+<!-- \end{split}\end{align} -->
+<!-- ``` -->
+<!-- Since the p-value calculation only needs to concern the number of times the actual data plot being selected denoted by $C_i$, they showed the model can be simplified to a beta-binomial mixture distribution -->
+
+<!-- ```{=tex} -->
+<!-- \begin{align} \label{eq:beta-binomial}\begin{split} -->
+<!-- \theta_i&|\alpha \sim Beta(\alpha, (m-1)\alpha)\\ -->
+<!-- C_i&|\theta_i \sim Binomial(K, \theta_i). -->
+<!-- \end{split}\end{align} -->
+<!-- ``` -->
+<!-- Thus, the visual p-value followed by the beta-binomial model is given as -->
+
+<!-- ```{=tex} -->
+<!-- \begin{equation} \label{eq:pvalue-beta-binomial} -->
+<!-- P(C \geq c_i) = \sum_{x=c_i}^{K}{{K}\choose{x}}\frac{1}{B(\alpha, (m-1)\alpha)}B(x + \alpha, K - x + (m - 1)\alpha), -->
+<!-- \end{equation} -->
+<!-- ``` -->
+<!-- where $B(.)$ is the beta function defined as -->
+
+<!-- ```{=tex} -->
+<!-- \begin{equation} \label{eq:betafunction} -->
+<!-- B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0.  -->
+<!-- \end{equation} -->
+<!-- ``` -->
+
+<!-- Note that the use of Equation (\ref{eq:pvalue-beta-binomial}) requires the estimation of $\hat{\alpha}$, which largely depends on the null model, the type of the plot and other aesthetic features.
 They suggested to estimate $\hat{\alpha}$ visually based on the selections of null plots of the experimental data, or to estimate $\hat{\alpha}$ numerically based on several additional Rorschach lineups, which is a type of lineup containing only null plots.
 However, when the number of null models are large, it could be expensive to manually estimate each $\alpha$ or include additional Rorschach lineups in the experiment.
-
-Instead, in the experiments that will be described in section \ref{experimental-design}, we adopt a simpler model implicitly used by the `pmulti()` function of the `vinference` `R` package.
+-->
+
+<!-- Instead, in the experiments that will be described in section \ref{experimental-design}, we adopt a simpler model implicitly used by the `pmulti()` function of the `vinference` `R` package.
 We assume the attractiveness of the plot $i$ modelled as $w_i \sim Uniform(0,1)$ for $i=1,..,m$.
@@
 Then, given the number of selections $s_j$, for $j=1,...,K$, the distribution of $C_i$ can be approximated by simulating the random selection process with computer.
-The simulated visual test p-value is formulated as
-
-```{=tex}
-\begin{equation} \label{eq:p-value-multi}
-\text{p-value} = \frac{\#draws~that~the~actual~data~plota~being~selected~more~than~c_i~times}{\#simulation}.
-\end{equation}
-```
+The simulated visual test p-value is formulated as -->
+
+<!-- ```{=tex} -->
+<!-- \begin{equation} \label{eq:p-value-multi} -->
+<!-- \text{p-value} = \frac{\#draws~that~the~actual~data~plota~being~selected~more~than~c_i~times}{\#simulation}. -->
+<!-- \end{equation} -->
+<!-- ``` -->
 <!-- ### Effectiveness of visual inference in regression diagnostics -->

### 2022-09-13 paper_comparison/paper_comparison/paper_comparison.tex
update p value calc

@@
 \citet{cook1982residuals}. As shown in Figure
-\textbackslash ref\{fig:commondepartures\}, the top-left panel is a good
-residual plot with residuals evenly distributed at both sides of the
-horizontal zero line showing no noticeable patterns. There are various
-types of departures from a good residual plot. We will only discuss
-three most commonly checked departures in this paper. Namely,
-non-linearity, heterskedasticity and non-normality.
+\ref{fig:commondepartures}, the top-left panel is a good residual plot
+with residuals evenly distributed at both sides of the horizontal zero
+line showing no noticeable patterns. There are various types of
+departures from a good residual plot. We will only discuss three most
+commonly checked departures in this paper. Namely, non-linearity,
+heterskedasticity and non-normality.
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/first-example-lineup-1} \caption{Example lineup consists of 20 randomly placed residual plots, where 19 are null plots with data consistent with the null hypothesis, and one is the actual residual plot.
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/first-example-lineup-1} \caption{An example lineup consists of 20 randomly placed residual plots, where 19 are null plots with data consistent with the null hypothesis, and one is the actual residual plot.
 Can you find the most different plot?}\label{fig:first-example-lineup}
@@
 
-\hypertarget{se:lineup}{%
-\subsubsection{Lineup protocol}\label{se:lineup}}
-
-With the simulation mechanism of null plots being provided, another
-aspect of hypothesis testing that needs to be addressed is the control
-of false positive rate or Type I error. Any visual statistic
-\(T^{(i)}(\boldsymbol{\mathrm{y}})\) needs to pair with a critical value
-\(c^{(i)}\) to form a hypothesis test. When a visual feature \(i\) is
-discovered by the observer from a plot, the corresponding visual
-statistic \(T^{(i)}(\boldsymbol{\mathrm{y}})\) may not be known as there
-is no general agreement on the measurement of the degree of presence of
-a visual feature. It is only the event that
-\(T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}\) is confirmed. Similarly,
-if any visual discovery is found by the observer, we say, there exists
-\(i \in I:~T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}\)
-\citep{buja_statistical_2009}.
-
-Using the above definition, the family-wise Type I error can be
-controlled if one can provide the collection of critical values
-\(c^{(i)}~(i \in I)\) such that
-\(P(\mathrm{there~exists~} i \in I: T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}|\boldsymbol{\mathrm{y}}) \leq \alpha\),
-where \(\alpha\) is the significance level. However, since the quantity
-of \(T^{(i)}(\boldsymbol{\mathrm{y}})\) may not be known, such
-collection of critical values can not be provided.
-
-\citet{buja_statistical_2009} proposed the lineup protocol as a visual
-test to calibrate the Type I error issue without the specification of
-\(c^{(i)}~(i \in I)\). It is inspired by the ``police lineup'' or
-``identity parade'' which is the act of asking the eyewitness to
-identify criminal suspect from a group of irrelevant people. The
-protocol consists of \(m\) randomly placed plots, where one plot is the
-actual data plot, and the remaining \(m - 1\) plots have the identical
-graphical production as the data plot except the data has been replaced
-with data consistent with the null hypothesis. Then, an observer who
-have not seen the actual data plot will be asked to point out the most
-different plot from the lineup.
-
-Under the null hypothesis, it is expected that the actual data plot
-would have no distinguishable difference with the null plots, and the
-probability of the observer correctly picks the actual data plot is
-\(1/m\). If we reject the null hypothesis as the observer correctly
-picks the actual data plot, then the Type I error of this test is
-\(1/m\).
-
-This provides us with an mechanism to control the Type I error, because
-\(m\) - the number of plots in a lineup can be chosen. A larger value of
-\(m\) will result in a smaller Type I error, but the limit to the value
-of \(m\) depends on the number of plots a human is willing to view
-\citep{buja_statistical_2009}. Typically, \(m\) will be set to \(20\)
-which is equivalent to set \(\alpha = 0.05\), a general choice of
-significance level for conventional testing among statisticians.
-
-Further, a visual test can involve \(K\) independent observers. Let
-\(D_i = \{0,1\}\) be a binomial random variable denoting whether subject
-\(i\) correctly detecting the actual data plot, and
-\(X = \sum_{i=1}^{K}X_i\) be the number of observers correctly picking
-the actual data plot. Then, by imposing a relatively strong assumption
-on the visual test that all \(K\) evaluations are fully independent,
-under the null hypothesis, \(X \sim \mathrm{Binom}_{K,1/m}\). Therefore,
-the \(p\)-value of a lineup of size \(m\) evaluated by \(K\) observer is
-given as \begin{equation} \label{eq:pvaluesingle}
-P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},
-\end{equation}
-
-where \(x\) is the realization of number of observers correctly picking
-the actual data plot \citep{majumder_validation_2013}.
-
-The multiple individuals approach avoids the limit of \(m\), while
-provides visual tests with \(p\)-value much smaller than \(0.05\). In
-fact, the lower bound of \(p\)-value decreases exponentially as \(K\)
-increases. With just \(4\) individuals and \(20\) data plots in a
-lineup, the \(p\)-value could be as small as \(0.0001\). Additionally,
-by involving multiple observers, variation of individual ability to read
-plots can be addressed to some degree as different opinions about visual
-discoveries can be collected.
-
 As pointed out by \citet{vanderplas2021statistical}, though Equation
@@
 \(c_i\) instead of one, where \(s_j\) is the number of plots subject
-\(j\) selected for \(j=1,...K\). This ensured \(\sum_{i}c_i=K\).
-
-The full model was a Dirichlet-multinomial mixture distribution
-
-\begin{align} \label{eq:dirichlet-multinomial}\begin{split}
-\boldsymbol{\theta}&|\alpha \sim Dirichlet(\alpha)\\
-(c_i,...,c_m)&|\boldsymbol{\theta} \sim Multinomial(K, \boldsymbol{\theta}).
-\end{split}\end{align}
-
-Since the p-value calculation only needs to concern the number of times
-the actual data plot being selected denoted by \(C_i\), they showed the
-model can be simplified to a beta-binomial mixture distribution
-
-\begin{align} \label{eq:beta-binomial}\begin{split}
-\theta_i&|\alpha \sim Beta(\alpha, (m-1)\alpha)\\
-C_i&|\theta_i \sim Binomial(K, \theta_i).
-\end{split}\end{align}
-
-Thus, the visual p-value followed by the beta-binomial model is given as
+\(j\) selected for \(j=1,...K\). This ensured \(\sum_{i}c_i=K\). Thus,
+the visual p-value followed by the beta-binomial model is given as
 
@@
 B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0. 
-\end{equation}
-
-Note that the use of Equation (\ref{eq:pvalue-beta-binomial}) requires
-the estimation of \(\hat{\alpha}\), which largely depends on the null
-model, the type of the plot and other aesthetic features. They suggested
-to estimate \(\hat{\alpha}\) visually based on the selections of null
-plots of the experimental data, or to estimate \(\hat{\alpha}\)
-numerically based on several additional Rorschach lineups, which is a
-type of lineup containing only null plots. However, when the number of
-null models are large, it could be expensive to manually estimate each
-\(\alpha\) or include additional Rorschach lineups in the experiment.
-
-Instead, in the experiments that will be described in section
-\ref{experimental-design}, we adopt a simpler model implicitly used by
-the \texttt{pmulti()} function of the \texttt{vinference} \texttt{R}
-package. We assume the attractiveness of the plot \(i\) modelled as
-\(w_i \sim Uniform(0,1)\) for \(i=1,..,m\). Let
-\(\theta_i = w_i/\sum_{i=1}^{m}w_i\) be the probability of plot \(i\)
-being selected by a subject. Then, given the number of selections
-\(s_j\), for \(j=1,...,K\), the distribution of \(C_i\) can be
-approximated by simulating the random selection process with computer.
-The simulated visual test p-value is formulated as
-
-\begin{equation} \label{eq:p-value-multi}
-\text{p-value} = \frac{\#draws~that~the~actual~data~plota~being~selected~more~than~c_i~times}{\#simulation}.
 \end{equation}

### 2022-09-20 paper_comparison/paper_comparison/paper_comparison.Rmd
take notes
clean comment

@@
 
-<!-- check the policy -->
-
-<!-- problem: residual plot diagnostics -->
-
-<!-- conventional test: too sensitive -->
-
-<!-- background:  -->
-
-<!-- week 1 -->
-
-<!-- 1. residual plot for model diagnostics -->
-
-<!-- a. residual is widely used -->
-
-<!-- b. what are the types of residual plots -->
-
-<!-- c. comparison -->
-
-<!-- week 2 -->
-
-<!-- 2. conventional test: F, BP -->
-
-<!-- 3. visual test: lineup, theory -->
-
-<!-- week 4 -->
-
-<!-- desc of experiment: -->
-
-<!-- 1. simulation setup -->
-
-<!-- 2. experimental design -->
-
-<!-- 3. result -->
-
-<!-- week 3 -->
-
-<!-- comparison of conventional tests: -->
-
-<!-- 1. power (visual test vs. conventional test) -->
-
-<!-- (visual test most different one (everything test, any departure)) -->
-
-<!-- plot figure in a paper, desc, exp -->
-
-<!-- 2. investigate the difference (gap), give examples -->
-
-<!-- 3. conventional is too sensitive -->
-
-<!-- 4. make conventional less sensitive (vary alpha) -->
-
-<!-- last week -->
-
-<!-- conclusion: -->
-
-<!-- 1. too sensitive, visual test is needed/preferable -->
-
-<!-- 2. visual test is infeasible in large scale (expensive) -->
-
-<!-- 3. future work (role of computer vision) -->
-
 ```{r setup, include=FALSE}
@@
 The usual pattern of heterskedasticity on a residual plot is the inconsistent spread of the residuals at different x values.
-Visually, it sometimes results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure \\ref{fig:commondepartures}, or the "left-triangle" and "right-triganle" shape where the smallest variance occurs at the edges of the x-axis.
+Visually, it sometimes results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure \ref{fig:commondepartures}, or the "left-triangle" and "right-triganle" shape where the smallest variance occurs at the edges of the x-axis.
 
@@
 For example, given a skewed error distribution, one will see fewer data points and more outliers on one side of the zeroline as shown in the bottom-right of Figure \ref{fig:commondepartures}.
-
-<!-- Raw residuals and studentized residuals are the two most frequently used residuals in standard residual plots.
-The debt on which type of residuals should be used is always present.
-While raw residuals are the most common computer regression software package output, by applying a scaling factor, the ability to reveal nonconstant error variance in standard residual plots will often be enhanced by studentized residuals in small sample size [@gunst1980regression].
-As a two-dimensional representation of a model in a $p$-dimensional space, standard residual plots project data points onto the variable of the horizontal axis, which is a vector in $p$-dimensional space.
-Observations with the same projection will be treated as equivalent as they have the same abscissa.
-Consequently, standard residual plots are often useful in revealing model inadequacies in the direction of the variable of the horizontal axis but could be inadequate for detecting patterns in other directions, especially in those perpendicular to the variable of the horizontal axis.
-Hence, in practice, multiple standard residual plots with different horizontal axes will be examined [@cook1982residuals].
-Overlapping data points is another general issue in scatter plots not limited to standard residual plots, which often makes plots difficult to interpret because visual patterns are concealed.
-Thus, for a relatively large sample size, @cleveland1975graphical suggests the use of robust moving statistics as reference lines to give aid to the eye in seeing patterns, which nowadays, are usually replaced with a spline or local polynomial regression line.
--->
-
-<!-- Other types of data plots that are often used in regression diagnostics include partial residual plots and probability plots.
-Partial residual plots are useful supplements to standard residual plots as they provide additional information on the extent of the non-linearity.
-Probability plots can be used to compare the sampling distribution of the residuals to the normal distribution for assessing the normality assumptions.
--->
-
-<!-- ## Hypothesis testing -->
-
-<!-- Other than checking diagnostic plots, analysts may perform formal hypothesis testing for detecting model defects.
-Depending on the alternative hypothesis that is focused on, a variety of tests can be applied.
-For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution.
-For testing non-linearity, one may apply the F-test to examine the significance of specific polynomial and non-linear forms of the regressors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969].
--->
-
-<!-- As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are sensitive to other types of departures.
-It is likely the null hypothesis is correctly rejected but for the wrong reason, which is known as the "Type III error".
-Additionally, outliers will often incorrectly trigger the rejection of the null hypothesis despite the residuals are well-behaved [@cook_applied_1999].
-This can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers.
-@montgomery1982introduction suggested that based on their experience, statistical tests are not widely used in regression diagnostics.
-The same or even larger amount of information can be provided by diagnostic plots than the corresponding tests in most empirical studies.
-Not to mention, it is almost impossible to have an exactly correctly specified model in reality.
-There is a well-known aphorism in statistics stated by George Box - "All models are wrong, but some are useful".
-This indicates proper hypothesis tests will always reject the null hypothesis as long as the sample size is large enough.
-The outcome "Not reject" can be interpreted as either "effect size is small" or "sample size is small".
-and the outcome "reject" doesn't inform us whether and how much the model defects are of actual consequence to the inference and prediction.
-But still, the effectiveness of statistical tests shall not be disrespected.
-Statistical tests have a chance to provide analysts with unique information.
-In situations where no suitable diagnostic plots can be found for a particular violation of the assumptions, or excessive diagnostic plots are needed to be checked, one will have no choice but to fall back on statistical tests if there exists any.
-A good regression diagnostic practice should be a balanced combination of both methods.
--->
 
@@
 This can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers.
+
+<!-- report 0 when the value is small -->
+
+<!-- F-test -> model structrual test -->
+
+<!-- F-test -> F -->
+
+<!-- right justified  -->
+
+<!-- test formula -> appendix -->
+
+<!-- red col -> bold -->
 
@@
 
+<!-- improve the first sentence -->
+
 Unlike hypothesis testing built upon rigorous statistical procedures, reading diagnostic plots relies on graphical perception - human's ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984], which is to some extent subjective and indecisive.
@@
 
+<!-- use true exp lineup as example lineup -->
+
 ```{r first-example-lineup, fig.cap="An example lineup consists of 20 randomly placed residual plots, where 19 are null plots with data consistent with the null hypothesis, and one is the actual residual plot.
@@
 where $x$ is the realization of number of observers correctly picking the actual data plot [@majumder_validation_2013].
+
+<!-- remove the binomial one -->
 
@@
 ```
+
+<!-- a sentence to discuss alpha  -->
+
+### Power calculation
 

### 2022-09-20 paper_comparison/paper_comparison/paper_comparison.tex
clean comment

@@
 the so-called ``butterfly'' shape as shown in the bottom-left panel of
-Figure \textbackslash ref\{fig:commondepartures\}, or the
-``left-triangle'' and ``right-triganle'' shape where the smallest
-variance occurs at the edges of the x-axis.
+Figure \ref{fig:commondepartures}, or the ``left-triangle'' and
+``right-triganle'' shape where the smallest variance occurs at the edges
+of the x-axis.
 
@@
 \(j\) selected for \(j=1,...K\). This ensured \(\sum_{i}c_i=K\). Thus,
-the visual p-value followed by the beta-binomial model is given as
+the visual p-value followed by the above model is given as
 
@@
 \end{equation}
+
+\hypertarget{power-calculation}{%
+\subsubsection{Power calculation}\label{power-calculation}}
 

### 2022-09-20 paper_comparison/paper_comparison/paper_comparison.Rmd
fix the table for example residual plots

@@
 
+(This section discusses the visual patterns data analysts expect to see and their implications.)
+
 ```{r commondepartures, fig.cap = "Example fitted vs residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
@@
 
+(This section discusses the tests that will be used in the analysis and shows the results for the residual plots displayed in the previous section.)
+
 Other than checking diagnostic plots, analysts may perform formal hypothesis testing for detecting model defects.
@@
 For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution.
-For testing non-linearity, one may apply the F-test to examine the significance of specific polynomial and non-linear forms of the regressors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969].
+For testing non-linearity, one may apply the F-test as a model structural test to examine the significance of specific polynomial and non-linear forms of the regressors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969].
 And for testing normality, the Shapiro--Wilk test [ref here] is perhaps the most widely used test included by many of the statistical softwares.
@@
 
-Example residual plots given in Figure \ref{fig:commondepartures} are examined by the corresponding F-test, Breusch-Pagan test and Shapiro--Wilk test as shown in Table \ref{tab:example-residual-plot-table}.
-In the example, both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null hypothesis for departures they do not intend to examine.
+Example residual plots given in Figure \ref{fig:commondepartures} are examined by the corresponding model structural test, Breusch-Pagan test and Shapiro--Wilk test as shown in Table \ref{tab:example-residual-plot-table}.
+In the example, both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null hypothesis for departures that they do not intend to examine.
 As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are sensitive to other types of departures.
@@
 
-<!-- report 0 when the value is small -->
-
-<!-- F-test -> model structrual test -->
-
-<!-- F-test -> F -->
-
-<!-- right justified  -->
-
 <!-- test formula -> appendix -->
-
-<!-- red col -> bold -->
 
@@
                  shapiro.test(dat3$.resid)$p.value)) %>%
-  mutate(f = format(f, digits = 2), b = format(b, digits = 2), s = format(s, digits = 2)) -> table_dat
-
-  table_dat$f <- kableExtra::cell_spec(table_dat$f, color = ifelse(as.numeric(table_dat$f) < 0.05, "red", "black"))
-  table_dat$b <- kableExtra::cell_spec(table_dat$b, color = ifelse(as.numeric(table_dat$b) < 0.05, "red", "black"))
-  table_dat$s <- kableExtra::cell_spec(table_dat$s, color = ifelse(as.numeric(table_dat$s) < 0.05, "red", "black"))
+  mutate(f = format(round(f, digits = 3), nsmall = 3), b = format(round(b, digits = 3), nsmall = 3), s = format(round(s, digits = 3), nsmall = 3)) -> table_dat
+
+  table_dat$f <- kableExtra::cell_spec(table_dat$f, bold = as.numeric(table_dat$f) < 0.05)
+  table_dat$b <- kableExtra::cell_spec(table_dat$b, bold = as.numeric(table_dat$b) < 0.05)
+  table_dat$s <- kableExtra::cell_spec(table_dat$s, bold = as.numeric(table_dat$s) < 0.05)
 
   table_dat %>%
-  knitr::kable(col.names = c("Residual plots", "Departures", "F-test ", "Breusch-Pagan test", "Shapiro–Wilk test"),
+  knitr::kable(col.names = c("Residual plots", "Departures", "Model structural", "Breusch-Pagan", "Shapiro–Wilk"),
+               align = "llrrr",
                escape = FALSE,
@@
 
-<!-- Explain the tests you are showing later -->
-
-<!-- -   RESET -->
-
-<!-- -   BP -->
-
-<!-- -   SW -->
-
-<!-- and show the results for the residual plots displayed in previous section -->
-
-<!-- Briefly mention any others  -->
-
 ## Visual testing for departures
+
+(This sec)
 

### 2022-09-20 paper_comparison/paper_comparison/paper_comparison.tex
fix the table for example residual plots

@@
 
+(This section discusses the visual patterns data analysts expect to see
+and their implications.)
+
 \begin{figure}
@@
 
+(This section discusses the tests that will be used in the analysis and
+shows the results for the residual plots displayed in the previous
+section.)
+
 Other than checking diagnostic plots, analysts may perform formal
@@
 relies on the asymptotic properties of the null distribution. For
-testing non-linearity, one may apply the F-test to examine the
-significance of specific polynomial and non-linear forms of the
-regressors, or the significance of proxy variables as in the Ramsey
-Regression Equation Specification Error Test (RESET)
+testing non-linearity, one may apply the F-test as a model structural
+test to examine the significance of specific polynomial and non-linear
+forms of the regressors, or the significance of proxy variables as in
+the Ramsey Regression Equation Specification Error Test (RESET)
 \citep{ramsey_tests_1969}. And for testing normality, the Shapiro--Wilk
@@
 Example residual plots given in Figure \ref{fig:commondepartures} are
-examined by the corresponding F-test, Breusch-Pagan test and
-Shapiro--Wilk test as shown in Table
+examined by the corresponding model structural test, Breusch-Pagan test
+and Shapiro--Wilk test as shown in Table
 \ref{tab:example-residual-plot-table}. In the example, both the
 Breusch-Pagan test and the Shapiro--Wilk test rejects the null
-hypothesis for departures they do not intend to examine. As discussed in
-\citet{cook1982residuals}, most residual-based tests for a particular
-type of departure from model assumptions are sensitive to other types of
-departures. It is likely the null hypothesis is correctly rejected but
-for the wrong reason, which is known as the ``Type III error''.
-Additionally, outliers will often incorrectly trigger the rejection of
-the null hypothesis despite the residuals are well-behaved
+hypothesis for departures that they do not intend to examine. As
+discussed in \citet{cook1982residuals}, most residual-based tests for a
+particular type of departure from model assumptions are sensitive to
+other types of departures. It is likely the null hypothesis is correctly
+rejected but for the wrong reason, which is known as the ``Type III
+error''. Additionally, outliers will often incorrectly trigger the
+rejection of the null hypothesis despite the residuals are well-behaved
 \citep{cook_applied_1999}. This can be largely avoided in diagnostic
@@
 \centering
-\begin{tabular}[t]{lllll}
+\begin{tabular}[t]{llrrr}
 \toprule
-Residual plots & Departures & F-test  & Breusch-Pagan test & Shapiro–Wilk test\\
+Residual plots & Departures & Model structural & Breusch-Pagan & Shapiro–Wilk\\
 \midrule
-B & Non-linearity & \textcolor{red}{7.3e-117} & \textcolor{red}{1.8e-14} & \textcolor{red}{3.9e-02}\\
-C & Heteroskedasticity & \textcolor{black}{ 3.8e-01} & \textcolor{red}{6.1e-16} & \textcolor{red}{1.1e-09}\\
-D & Non-normality & \textcolor{black}{ 6.7e-01} & \textcolor{black}{7.4e-01} & \textcolor{red}{2.9e-18}\\
+B & Non-linearity & \textbf{0.000} & \textbf{0.000} & \textbf{0.039}\\
+C & Heteroskedasticity & 0.378 & \textbf{0.000} & \textbf{0.000}\\
+D & Non-normality & 0.667 & 0.736 & \textbf{0.000}\\
 \bottomrule
@@
 departures}\label{visual-testing-for-departures}}
+
+(This sec)
 

### 2022-09-20 paper_comparison/paper_comparison/paper_comparison.Rmd
clean up the background

@@
                booktabs = TRUE, 
-               caption  = 'Example residual plots examined by the corresponding conventional tests. P-values smaller than 0.05 are colored in red.', 
+               caption  = 'Example residual plots examined by the corresponding conventional tests. P-values smaller than 0.05 are showed in bold.', 
                label = "example-residual-plot-table" )
@@
 
-(This sec)
+(This section introduces the lineup protocol, and briefly discusses the method for sampling null data, calculating the p-value and estimating power.)
 
@@
 
-### p-value calculation
+### P-value calculation
 
@@
 Then, by imposing a relatively strong assumption on the visual test that all $K$ evaluations are fully independent, under the null hypothesis, $X \sim \mathrm{Binom}_{K,1/m}$.
-Therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as \begin{equation} \label{eq:pvaluesingle}
-P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},
-\end{equation}
-
-where $x$ is the realization of number of observers correctly picking the actual data plot [@majumder_validation_2013].
-
-<!-- remove the binomial one -->
-
-As pointed out by @vanderplas2021statistical, though Equation (\ref{eq:pvaluesingle}) is trivial, but it doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup.
+Therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as $P(X \geq x) = 1 - F(x) + f(x)$, where $F(.)$ is the cumulative distribution function, $f(.)$ is the probability mass function and $x$ is the realization of number of observers correctly picking the actual data plot [@majumder_validation_2013].
+
+As pointed out by @vanderplas2021statistical, the binomial model doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup.
 And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup.
-They summarized three common different scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same actual data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects.
+They summarized three common scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same actual data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects.
 Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experimental design.
@@
 
-<!-- a sentence to discuss alpha  -->
+The parameter $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} is usually unknown and hence needs to be estimated from the survey data.
+For low values of $\alpha$, only a few plots are attractive to the observers and tend to be selected.
+For higher values of $\alpha$, the distribution of the probability of each plot being selected is more evenly.
+@vanderplas2021statistical suggested that $\alpha$ can be estimated using maximum likelihood estimation or visual estimation.
+But if $\alpha$ is small and only a few null plots in a lineup are attractive, MLE could fail to provide accurate estimates.
 
@@
 <!-- -   power calculation, briefly, including how multiple selections are handled -->
-
-<!-- Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualization" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots [@gelman_exploratory_2004].
-Later, in the Bayesian context, data plots was systematically considered as model diagnostics by taking advantage of the data simulated from the assumed statistical models [@gelman_bayesian_2003; @gelman_exploratory_2004].
--->
-
-<!-- It was surprising that the essential components of visual inference had actually been established in @buja_inference_1999, but it was not until 10 years later that @buja_statistical_2009 formalized it as an inferential framework to extend confirmatory statistics to visual discoveries.
-This framework redefines the test statistics, tests, null distribution, significance levels and $p$-value for visual discovery modelled on the confirmatory statistical testing.
-Figure \ref{fig:parallelism} outlines the parallelism between conventional tests and visual discovery.
--->
 
@@
 ```
-
-<!-- In visual inference, a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$ is defined, where $\boldsymbol{\mathrm{y}}$ is the data and $I$ is a set of all possible visual features.
-@buja_statistical_2009 described each of the test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})$ as a measurement of the degree of presence of a visual feature.
-Alternatively, @majumder_validation_2013 avoids the use of visual features and defined the visual statistics $T(.)$ as a mapping from a dataset to a data plot.
-Both definitions of visual test statistics are valid, but in the rest of the paper the first definition will be used as it covers some details needed by the following discussion.
-A visual discovery is defined as a rejection of a null hypothesis, and the same null hypothesis can be rejected by many different visual discoveries [@buja_statistical_2009].
-For regression diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the null hypothesis.
-The same regression model can be rejected by many reasons with residual plot, including non-linearity and heteroskedasticity as shown in Figure \ref{fig:residual-plot-cubic-heter}.
--->
 
@@
 
-<!-- ### Sampling from the null distribution {#se:sampling-from-null} -->
-
-<!-- The null distribution of plots refers to the infinite collection of plots of null datasets sampled from $H_0$.
-It is defined as the analogue of the null distribution of test statistics in conventional test [@buja_statistical_2009].
-In practice, a finite number of plots of null datasets could be generated, called null plots.
-In the context of regression diagnostics, sampling data from $H_0$ is equivalent to sampling data from the assumed model.
-As @buja_statistical_2009 suggested, $H_0$ is usually composited by a collection of distributions controlled by nuisance parameters.
-Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling.
--->
-
-<!-- The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009].
-Suppose there exists a minimal sufficient statistic $\boldsymbol{S}(\boldsymbol{y})$ under the null hypothesis, any null datasets $\boldsymbol{y^{*}}$ should fulfil the condition $\boldsymbol{S}(\boldsymbol{y}) = \boldsymbol{s}$.
-Using the classical normal linear regression model as example, the minimal sufficient statistic is $\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{\beta}}, \boldsymbol{e}'\boldsymbol{e})$, where $\hat{\boldsymbol{\beta}}$ are the coefficient estimators and $\boldsymbol{e}'\boldsymbol{e}$ is the residual sum of square.
-Alternatively, the minimal sufficient statistic can be constructed as $\boldsymbol{S}(\boldsymbol{y}) = (\hat{\boldsymbol{y}}, ||\boldsymbol{e}||)$, where $\hat{\boldsymbol{y}}$ are the fitted values and $||\boldsymbol{e}||$ is the length of residuals, which is more intuitive as suggested by @buja_statistical_2009.
-Since the fitted values are held fixed, the variation can only occur in the residual space.
-And because the length of residual is also held fixed, residuals obtained from a null dataset has to be a random rotation of $\boldsymbol{e}$ in the residual space.
-With this property, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the regressors, then rescaling it by the ratio of residual sum of square in two regressions.
--->
-
-<!-- ### Lineup protocol {#se:lineup} -->
-
-<!-- With the simulation mechanism of null plots being provided, another aspect of hypothesis testing that needs to be addressed is the control of false positive rate or Type I error.
-Any visual statistic $T^{(i)}(\boldsymbol{\mathrm{y}})$ needs to pair with a critical value $c^{(i)}$ to form a hypothesis test.
-When a visual feature $i$ is discovered by the observer from a plot, the corresponding visual statistic $T^{(i)}(\boldsymbol{\mathrm{y}})$ may not be known as there is no general agreement on the measurement of the degree of presence of a visual feature.
-It is only the event that $T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}$ is confirmed.
-Similarly, if any visual discovery is found by the observer, we say, there exists $i \in I:~T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}$ [@buja_statistical_2009].
--->
-
-<!-- Using the above definition, the family-wise Type I error can be controlled if one can provide the collection of critical values $c^{(i)}~(i \in I)$ such that $P(\mathrm{there~exists~} i \in I: T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}|\boldsymbol{\mathrm{y}}) \leq \alpha$, where $\alpha$ is the significance level.
-However, since the quantity of $T^{(i)}(\boldsymbol{\mathrm{y}})$ may not be known, such collection of critical values can not be provided.
--->
-
-<!-- @buja_statistical_2009 proposed the lineup protocol as a visual test to calibrate the Type I error issue without the specification of $c^{(i)}~(i \in I)$.
-It is inspired by the "police lineup" or "identity parade" which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people.
-The protocol consists of $m$ randomly placed plots, where one plot is the actual data plot, and the remaining $m - 1$ plots have the identical graphical production as the data plot except the data has been replaced with data consistent with the null hypothesis.
-Then, an observer who have not seen the actual data plot will be asked to point out the most different plot from the lineup.
--->
-
-<!-- Under the null hypothesis, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is $1/m$.
-If we reject the null hypothesis as the observer correctly picks the actual data plot, then the Type I error of this test is $1/m$.
--->
-
-<!-- This provides us with an mechanism to control the Type I error, because $m$ - the number of plots in a lineup can be chosen.
-A larger value of $m$ will result in a smaller Type I error, but the limit to the value of $m$ depends on the number of plots a human is willing to view [@buja_statistical_2009].
-Typically, $m$ will be set to $20$ which is equivalent to set $\alpha = 0.05$, a general choice of significance level for conventional testing among statisticians.
--->
-
-<!-- Further, a visual test can involve $K$ independent observers.
-Let $D_i = \{0,1\}$ be a binomial random variable denoting whether subject $i$ correctly detecting the actual data plot, and $X = \sum_{i=1}^{K}X_i$ be the number of observers correctly picking the actual data plot.
-Then, by imposing a relatively strong assumption on the visual test that all $K$ evaluations are fully independent, under the null hypothesis, $X \sim \mathrm{Binom}_{K,1/m}$.
-Therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as \begin{equation} \label{eq:pvaluesingle} -->
-<!-- P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i}, -->
-<!-- \end{equation} -->
-
-<!-- where $x$ is the realization of number of observers correctly picking the actual data plot [@majumder_validation_2013]. -->
-
-<!-- The multiple individuals approach avoids the limit of $m$, while provides visual tests with $p$-value much smaller than $0.05$.
-In fact, the lower bound of $p$-value decreases exponentially as $K$ increases.
-With just $4$ individuals and $20$ data plots in a lineup, the $p$-value could be as small as $0.0001$.
-Additionally, by involving multiple observers, variation of individual ability to read plots can be addressed to some degree as different opinions about visual discoveries can be collected.
--->
-
-<!-- As pointed out by @vanderplas2021statistical, though Equation (\ref{eq:pvaluesingle}) is trivial, but it doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup.
-And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup.
-They summarized three common different scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same actual data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects.
-Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experimental design.
-For Scenario 3, @vanderplas2021statistical modelled the probability of a plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$.
-And defined $c_i$ to be the number of times plot $i$ being selected in $K$ evaluations.
-In case subject $j$ makes multiple selections, they decided to add $1/s_j$ to $c_i$ instead of one, where $s_j$ is the number of plots subject $j$ selected for $j=1,...K$.
-This ensured $\sum_{i}c_i=K$.
--->
-
-<!-- The full model was a Dirichlet-multinomial mixture distribution -->
-
-<!-- ```{=tex} -->
-<!-- \begin{align} \label{eq:dirichlet-multinomial}\begin{split} -->
-<!-- \boldsymbol{\theta}&|\alpha \sim Dirichlet(\alpha)\\ -->
-<!-- (c_i,...,c_m)&|\boldsymbol{\theta} \sim Multinomial(K, \boldsymbol{\theta}). -->
-<!-- \end{split}\end{align} -->
-<!-- ``` -->
-<!-- Since the p-value calculation only needs to concern the number of times the actual data plot being selected denoted by $C_i$, they showed the model can be simplified to a beta-binomial mixture distribution -->
-
-<!-- ```{=tex} -->
-<!-- \begin{align} \label{eq:beta-binomial}\begin{split} -->
-<!-- \theta_i&|\alpha \sim Beta(\alpha, (m-1)\alpha)\\ -->
-<!-- C_i&|\theta_i \sim Binomial(K, \theta_i). -->
-<!-- \end{split}\end{align} -->
-<!-- ``` -->
-<!-- Thus, the visual p-value followed by the beta-binomial model is given as -->
-
-<!-- ```{=tex} -->
-<!-- \begin{equation} \label{eq:pvalue-beta-binomial} -->
-<!-- P(C \geq c_i) = \sum_{x=c_i}^{K}{{K}\choose{x}}\frac{1}{B(\alpha, (m-1)\alpha)}B(x + \alpha, K - x + (m - 1)\alpha), -->
-<!-- \end{equation} -->
-<!-- ``` -->
-<!-- where $B(.)$ is the beta function defined as -->
-
-<!-- ```{=tex} -->
-<!-- \begin{equation} \label{eq:betafunction} -->
-<!-- B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0.  -->
-<!-- \end{equation} -->
-<!-- ``` -->
-
-<!-- Note that the use of Equation (\ref{eq:pvalue-beta-binomial}) requires the estimation of $\hat{\alpha}$, which largely depends on the null model, the type of the plot and other aesthetic features.
-They suggested to estimate $\hat{\alpha}$ visually based on the selections of null plots of the experimental data, or to estimate $\hat{\alpha}$ numerically based on several additional Rorschach lineups, which is a type of lineup containing only null plots.
-However, when the number of null models are large, it could be expensive to manually estimate each $\alpha$ or include additional Rorschach lineups in the experiment.
--->
-
-<!-- Instead, in the experiments that will be described in section \ref{experimental-design}, we adopt a simpler model implicitly used by the `pmulti()` function of the `vinference` `R` package.
-We assume the attractiveness of the plot $i$ modelled as $w_i \sim Uniform(0,1)$ for $i=1,..,m$.
-Let $\theta_i = w_i/\sum_{i=1}^{m}w_i$ be the probability of plot $i$ being selected by a subject.
-Then, given the number of selections $s_j$, for $j=1,...,K$, the distribution of $C_i$ can be approximated by simulating the random selection process with computer.
-The simulated visual test p-value is formulated as -->
-
-<!-- ```{=tex} -->
-<!-- \begin{equation} \label{eq:p-value-multi} -->
-<!-- \text{p-value} = \frac{\#draws~that~the~actual~data~plota~being~selected~more~than~c_i~times}{\#simulation}. -->
-<!-- \end{equation} -->
-<!-- ``` -->
-<!-- ### Effectiveness of visual inference in regression diagnostics -->
-
-<!-- Compared to the conventional test, whose power only depends on the parameter of interest $\theta$, several studies [see @hofmann_graphical_2012; @majumder_validation_2013; @majumder2014human; @roy_chowdhury_using_2015; @loy2016variations] have shown the power of the visual test is subject-specific.
-Thus, to be able to account for individual's ability, an individual is required to evaluate multiple lineups [@majumder_validation_2013].
--->
-
-<!-- Suppose individuals have the same ability and a lineup has been evaluated by multiple individuals, under the alternative hypothesis, the estimated power for a lineup can be expressed as $\hat{p} = x/K$, the estimated probability of identifying the actual data plot from the lineup.
-If the individual skill needs to be taken into account, and $L$ lineups have been evaluated by $K$ individuals, @majumder_validation_2013 suggests that mixed effects logistic regression model can be fit as: -->
-
-<!-- $$g(p_{li}) = W_{li}\delta + Z_{li}\tau_{li},$$ -->
-
-<!-- where $g(.)$ is the logit link function $g(p) = log(p)  - log(1-p)$; $0 \leq p \leq 1$.
-$W_{li}$, $1 \leq i \leq K$, $1 \leq l \leq L$, is the covariate matrix including lineup-specific elements and demographic information of individuals, and $\delta$ is a vector of parameters.
-$Z$ is the random effects matrix, and $\tau$ is a vector of variables follow $N(\boldsymbol{0},\sigma_{\tau}\boldsymbol{I}_{KL\times KL})$.
--->
-
-<!-- Then, the estimated power for lineup $l$ and individual $i$ can be calculated as $\hat{p}_{li} = g^{-1}(W_{li}\hat{\delta} + Z_{li}\hat{\tau}_{li})$ [@majumder_validation_2013]. -->
-
-## Effectiveness of visual test in regression diagnostics
-
-The effectiveness of visual inference has already been validated by @majumder_validation_2013 under relatively simple classical normal linear regression model settings with only one or two regressors.
-Their results suggest visual test is capable of testing the significance of a single regressor with a similar power as a t-test, though they expressed that in general it is unnecessary to use visual inference if there exists a conventional test and they didn't expect the visual test to perform equally well as the conventional test.
-In their third experiment, where there does not exist a proper conventional test, visual test outperforms the conventional test for a large margin.
-This is encouraging as it promotes the use of visual inference in border field of data science where there are no existing statistical testing procedures.
-In fact, lineup protocol has been integrated into some model diagnostic tools such as @loy2013diagnostic.
-
-<!-- refactoring from here -->
-
-With our knowledge, what haven't been examined so far is the effectiveness of visual test relative to the equivalent conventional test in regression diagnostics.
-Particularly, its ability to detect non-linearity and heteroskedasticity compared to F-test and BP-test.
+## Unknown
+
+This should be a section that summarizes the background, gives the motivation to the experiments, and most importantly reflects on the title.
 
 # Experimental design
+
+(This section discu)
 

### 2022-09-20 paper_comparison/paper_comparison/paper_comparison.tex
clean up the background

@@
 
-(This sec)
+(This section introduces the lineup protocol, and briefly discusses the
+method for sampling null data, calculating the p-value and estimating
+power.)
 
@@
 \hypertarget{p-value-calculation}{%
-\subsubsection{p-value calculation}\label{p-value-calculation}}
+\subsubsection{P-value calculation}\label{p-value-calculation}}
 
@@
 the \(p\)-value of a lineup of size \(m\) evaluated by \(K\) observer is
-given as \begin{equation} \label{eq:pvaluesingle}
-P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},
-\end{equation}
+given as \(P(X \geq x) = 1 - F(x) + f(x)\), where \(F(.)\) is the
+cumulative distribution function, \(f(.)\) is the probability mass
+function and \(x\) is the realization of number of observers correctly
+picking the actual data plot \citep{majumder_validation_2013}.
 
-where \(x\) is the realization of number of observers correctly picking
-the actual data plot \citep{majumder_validation_2013}.
-
-As pointed out by \citet{vanderplas2021statistical}, though Equation
-(\ref{eq:pvaluesingle}) is trivial, but it doesn't take into account the
-possible dependencies in the visual test due to repeated evaluations of
-the same lineup. And it is inapplicable to visual test where subjects
-are asked to select one or more ``most different'' plots from the
-lineup. They summarized three common different scenarios in visual
-inference: (1) \(K\) different lineups are shown to \(K\) subjects, (2)
-\(K\) lineups with different null plots but the same actual data plot
-are shown to \(K\) subjects, and (3) the same lineup is shown to \(K\)
-subjects. Out of these three scenarios, Scenario 3 is the most common in
-previous studies as it puts the least constraints on the experimental
-design. For Scenario 3, \citet{vanderplas2021statistical} modelled the
-probability of a plot \(i\) being selected from a lineup as
+As pointed out by \citet{vanderplas2021statistical}, the binomial model
+doesn't take into account the possible dependencies in the visual test
+due to repeated evaluations of the same lineup. And it is inapplicable
+to visual test where subjects are asked to select one or more ``most
+different'' plots from the lineup. They summarized three common
+scenarios in visual inference: (1) \(K\) different lineups are shown to
+\(K\) subjects, (2) \(K\) lineups with different null plots but the same
+actual data plot are shown to \(K\) subjects, and (3) the same lineup is
+shown to \(K\) subjects. Out of these three scenarios, Scenario 3 is the
+most common in previous studies as it puts the least constraints on the
+experimental design. For Scenario 3, \citet{vanderplas2021statistical}
+modelled the probability of a plot \(i\) being selected from a lineup as
 \(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
@@
 
+The parameter \(\alpha\) used in Equation \ref{eq:pvalue-beta-binomial}
+is usually unknown and hence needs to be estimated from the survey data.
+For low values of \(\alpha\), only a few plots are attractive to the
+observers and tend to be selected. For higher values of \(\alpha\), the
+distribution of the probability of each plot being selected is more
+evenly. \citet{vanderplas2021statistical} suggested that \(\alpha\) can
+be estimated using maximum likelihood estimation or visual estimation.
+But if \(\alpha\) is small and only a few null plots in a lineup are
+attractive, MLE could fail to provide accurate estimates.
+
 \hypertarget{power-calculation}{%
@@
 
-\hypertarget{effectiveness-of-visual-test-in-regression-diagnostics}{%
-\subsection{Effectiveness of visual test in regression
-diagnostics}\label{effectiveness-of-visual-test-in-regression-diagnostics}}
+\hypertarget{unknown}{%
+\subsection{Unknown}\label{unknown}}
 
-The effectiveness of visual inference has already been validated by
-\citet{majumder_validation_2013} under relatively simple classical
-normal linear regression model settings with only one or two regressors.
-Their results suggest visual test is capable of testing the significance
-of a single regressor with a similar power as a t-test, though they
-expressed that in general it is unnecessary to use visual inference if
-there exists a conventional test and they didn't expect the visual test
-to perform equally well as the conventional test. In their third
-experiment, where there does not exist a proper conventional test,
-visual test outperforms the conventional test for a large margin. This
-is encouraging as it promotes the use of visual inference in border
-field of data science where there are no existing statistical testing
-procedures. In fact, lineup protocol has been integrated into some model
-diagnostic tools such as \citet{loy2013diagnostic}.
-
-With our knowledge, what haven't been examined so far is the
-effectiveness of visual test relative to the equivalent conventional
-test in regression diagnostics. Particularly, its ability to detect
-non-linearity and heteroskedasticity compared to F-test and BP-test.
+This should be a section that summarizes the background, gives the
+motivation to the experiments, and most importantly reflects on the
+title.
 
@@
 \section{Experimental design}\label{experimental-design}}
+
+(This section discu)
 

### 2022-09-20 paper_comparison/paper_comparison/paper_comparison.Rmd
update exp design

@@
 
+The effectiveness of lineup protocol has already been validated by @majumder_validation_2013 under relatively simple classical normal linear regression model settings with only one or two regressors.
+Their results suggest visual test is capable of testing the significance of a single regressor with a similar power as a t-test, though they expressed that in general it is unnecessary to use visual inference if there exists a conventional test and they didn't expect the visual test to perform equally well as the conventional test.
+In their third experiment, where there does not exist a proper conventional test, visual test outperforms the conventional test for a large margin.
+This is encouraging as it promotes the use of visual inference in border field of data science where there are no existing statistical testing procedures.
+In fact, lineup protocol has been integrated into some model diagnostic tools such as @loy2013diagnostic.
+
 ### Sampling from the null distribution
@@
 
-## Unknown
+# Experimental design
+
+(This section discusses the experimental design including the motivation of the experiment, an overview of the experiment, the simulation setting of the depatures from good residual plots, parameter choices, allocation of the lineups and other technical details.)
+
+## Motivation and overview
+
+With our knowledge, the effectiveness of visual test relative to the equivalent conventional test in regression diagnostics has not been validated.
+Particularly, its ability to detect non-linearity and heteroskedasticity compared to model structural test and Breusch–Pagan test.
+Meanwhile, as mentioned in Section \ref{introduction} [introduction] section, we keen to understand the differences between conventional hypothesis testing and graphical diagnostics in the application of linear regression diagnostics and why conventional tests are not preferred by data analysts in linear regression diagnostics.
+Attempting to find the answer to these questions, three experiments were conducted.
+
+The experiment I has ideal scenario for conventional testing, where the visual test is not expected to outperform the conventional test.
+The experiment II is a scenario where the conventional test is an approximate test, in which the visual test may have a chance to match the performance of the conventional test.
+
+Subjects for both experiments were recruited from an crowdsourcing platform called Prolific [ref here].
+Prescreening procedure was applied during the recruitment, subjects were required to be fluent in English, with $98\%$ minimum approval rate in other studies and 10 minimum submissions.
+
+During the experiment, every subject was presented with a block of 20 lineups.
+And for every lineup, the subject was asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots were from others.
+If there was no noticeable difference between plots in a lineup, subjects were permitted to select zero plots without providing the reason.
+No subject was shown the same lineup twice.
+Information about preferred pronoun, age group, education, and previous experience in visual experiment were also collected.
+
+A pool of 12 different lineups with obvious visual patterns were generated for experiment I and experiment II respectively.
+In every block of 20 lineups that presented to a subject, two out of 12 lineups were included as attention checks.
+A subject's submission was only accepted if the actual data plot was identified for at least one attention check.
+Data of rejected submissions were discarded automatically to maintain the overall data quality.
 
 This should be a section that summarizes the background, gives the motivation to the experiments, and most importantly reflects on the title.
-
-# Experimental design
-
-(This section discu)
-
-## Motivation and overview
 
@@
 software/technical
+
+# Old stuff
 

### 2022-09-20 paper_comparison/paper_comparison/paper_comparison.tex
update exp design

@@
 
-\caption{\label{tab:example-residual-plot-table}Example residual plots examined by the corresponding conventional tests. P-values smaller than 0.05 are colored in red.}
+\caption{\label{tab:example-residual-plot-table}Example residual plots examined by the corresponding conventional tests. P-values smaller than 0.05 are showed in bold.}
 \centering
@@
 issue.
+
+The effectiveness of lineup protocol has already been validated by
+\citet{majumder_validation_2013} under relatively simple classical
+normal linear regression model settings with only one or two regressors.
+Their results suggest visual test is capable of testing the significance
+of a single regressor with a similar power as a t-test, though they
+expressed that in general it is unnecessary to use visual inference if
+there exists a conventional test and they didn't expect the visual test
+to perform equally well as the conventional test. In their third
+experiment, where there does not exist a proper conventional test,
+visual test outperforms the conventional test for a large margin. This
+is encouraging as it promotes the use of visual inference in border
+field of data science where there are no existing statistical testing
+procedures. In fact, lineup protocol has been integrated into some model
+diagnostic tools such as \citet{loy2013diagnostic}.
 
@@
 
-\hypertarget{unknown}{%
-\subsection{Unknown}\label{unknown}}
+\hypertarget{experimental-design}{%
+\section{Experimental design}\label{experimental-design}}
+
+(This section discusses the experimental design including the motivation
+of the experiment, an overview of the experiment, the simulation setting
+of the depatures from good residual plots, parameter choices, allocation
+of the lineups and other technical details.)
+
+\hypertarget{motivation-and-overview}{%
+\subsection{Motivation and overview}\label{motivation-and-overview}}
+
+With our knowledge, the effectiveness of visual test relative to the
+equivalent conventional test in regression diagnostics has not been
+validated. Particularly, its ability to detect non-linearity and
+heteroskedasticity compared to model structural test and Breusch--Pagan
+test. Meanwhile, as mentioned in the \ref{introduction}
+\protect\hyperlink{introduction}{introduction} section
+
+For the purpose of investigating the differences between conventional
+hypothesis testing and graphical diagnostics in the application of
+linear regression diagnostics, three experiments were conducted.
+
+The experiment I has ideal scenario for conventional testing, where the
+visual test is not expected to outperform the conventional test. The
+experiment II is a scenario where the conventional test is an
+approximate test, in which the visual test may have a chance to match
+the performance of the conventional test.
+
+Subjects for both experiments were recruited from an crowdsourcing
+platform called Prolific {[}ref here{]}. Prescreening procedure was
+applied during the recruitment, subjects were required to be fluent in
+English, with \(98\%\) minimum approval rate in other studies and 10
+minimum submissions.
+
+During the experiment, every subject was presented with a block of 20
+lineups. And for every lineup, the subject was asked to select one or
+more plots that are most different from others, provide a reason for
+their selections, and evaluate how different they think the selected
+plots were from others. If there was no noticeable difference between
+plots in a lineup, subjects were permitted to select zero plots without
+providing the reason. No subject was shown the same lineup twice.
+Information about preferred pronoun, age group, education, and previous
+experience in visual experiment were also collected.
+
+A pool of 12 different lineups with obvious visual patterns were
+generated for experiment I and experiment II respectively. In every
+block of 20 lineups that presented to a subject, two out of 12 lineups
+were included as attention checks. A subject's submission was only
+accepted if the actual data plot was identified for at least one
+attention check. Data of rejected submissions were discarded
+automatically to maintain the overall data quality.
 
@@
 title.
-
-\hypertarget{experimental-design}{%
-\section{Experimental design}\label{experimental-design}}
-
-(This section discu)
-
-\hypertarget{motivation-and-overview}{%
-\subsection{Motivation and overview}\label{motivation-and-overview}}
 
@@
 software/technical
+
+\hypertarget{old-stuff}{%
+\section{Old stuff}\label{old-stuff}}
 

### 2022-09-20 paper_comparison/paper_comparison/paper_comparison.Rmd
omit the analysis section for faster knitting.

@@
 Particularly, its ability to detect non-linearity and heteroskedasticity compared to model structural test and Breusch–Pagan test.
-Meanwhile, as mentioned in Section \ref{introduction} [introduction] section, we keen to understand the differences between conventional hypothesis testing and graphical diagnostics in the application of linear regression diagnostics and why conventional tests are not preferred by data analysts in linear regression diagnostics.
+Meanwhile, as mentioned in Section \ref{introduction} section, we keen to understand the differences between conventional hypothesis testing and graphical diagnostics in the application of linear regression diagnostics and why conventional tests are not preferred by data analysts in linear regression diagnostics.
 Attempting to find the answer to these questions, three experiments were conducted.
@@
 The experiment II is a scenario where the conventional test is an approximate test, in which the visual test may have a chance to match the performance of the conventional test.
-
-Subjects for both experiments were recruited from an crowdsourcing platform called Prolific [ref here].
+The experiment III is designed for collecting human responses to lineup with only good residual plots such that the parameter $\alpha$ in Equation \ref{eq:pvalue-beta-binomial} can be estimated.
+
+Subjects for all three experiments were recruited from an crowdsourcing platform called Prolific [ref here].
 Prescreening procedure was applied during the recruitment, subjects were required to be fluent in English, with $98\%$ minimum approval rate in other studies and 10 minimum submissions.
@@
 
-A pool of 12 different lineups with obvious visual patterns were generated for experiment I and experiment II respectively.
-In every block of 20 lineups that presented to a subject, two out of 12 lineups were included as attention checks.
+In every block of 20 lineups that presented to a subject, two lineups with obvious visual patterns were included as attention checks.
 A subject's submission was only accepted if the actual data plot was identified for at least one attention check.
@@
 
-This should be a section that summarizes the background, gives the motivation to the experiments, and most importantly reflects on the title.
+Overall, we planned to collect 7974 evaluations on 1152 uniqued lineups performed by 443 subjects.
 
@@
 
-### Nonlinearity
-
-### Heteroskedasticity
-
-## Experimental setup
-
-### Factors
-
--   Treatments for nonlinearity experiment (plots summarising levels)
--   Treatments for heteroskedasticity experiment (plots summarising levels)
-
-### Assigning subjects
-
-### Collecting results
-
-software/technical
-
-# Old stuff
-
-## Computing power curves
-
-For the purpose of examining the effectiveness of visual test in regression diagnostics, two experiments were conducted.
-The experiment I has ideal scenario for conventional testing, where the visual test is not expected to outperform the conventional test.
-The experiment II is a scenario where the conventional test is an approximate test, in which the visual test may have a chance to match the performance of the conventional test.
-
-Subjects for both experiments were recruited from an crowdsourcing platform called Prolific [ref here].
-Prescreening procedure was applied during the recruitment, subjects were required to be fluent in English, with $98\%$ minimum approval rate in other studies and 10 minimum submissions.
-
-During the experiment, every subject was presented with a block of 20 lineups.
-And for every lineup, the subject was asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots were from others.
-If there was no noticeable difference between plots in a lineup, subjects were permitted to select zero plots without providing the reason.
-No subject was shown the same lineup twice.
-Information about preferred pronoun, age group, education, and previous experience in visual experiment were also collected.
-
-A pool of 12 different lineups with obvious visual patterns were generated for experiment I and experiment II respectively.
-In every block of 20 lineups that presented to a subject, two out of 12 lineups were included as attention checks.
-A subject's submission was only accepted if the actual data plot was identified for at least one attention check.
-Data of rejected submissions were discarded automatically to maintain the overall data quality.
-
-## Non-linearity
+### Non-linearity
 
@@
 
-```{=tex}
 \begin{align} \label{eq:nonlinearity-model}
@@
 \end{align}
-```
+
 where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vector of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $\{-k, k\}$ defined as
@@
 ```
+
 The null regression model used to fit the realizations generated by the above model is formulated as:
@@
 
-Model misspecification presents since the null model leaves out the higher order term.
+Since $z = O(x^j)$, for $j > 1$, $z$ is a higher order term leaves out by the null regression causing model misspecification.
+
+### Heteroskedasticity
+
+## Experimental setup
+
+### Factors
+
+-   Treatments for nonlinearity experiment (plots summarising levels)
+-   Treatments for heteroskedasticity experiment (plots summarising levels)
 
@@
 
-## Heteroskedasticity
-
-Experiment II is designed to study the ability of human subjects to detect the appearance of a heteroskedasticity pattern under a simple linear regression model setting:
-
-```{=tex}
-\begin{align} \label{eq:heter-model}
-\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
-\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + 2 - |a| b (\boldsymbol{x} - a)^2 \boldsymbol{I}), \\
-\end{align}
-```
-where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vector of size $n$.
-
-The null regression model used to fit the realizations generated by the above model is formulated exactly the same as the model used in the first experiment:
-
-```{=tex}
-\begin{equation} 
-\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
-\end{equation}
-```
-where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.
-
-Model misspecification presents since the assumption about the constant error variance is violated.
-
-Experiment data were simulated using three different shapes ($a$ = -1, 0, 1), five different values of $b$ (b = 0.25, 1, 4, 16, 64), three different sample sizes (n = 50, 100, 300) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
+### Assigning subjects
+
+### Collecting results
+
+software/technical
+
+<!-- # Old stuff -->
+
+<!-- ## Computing power curves -->
+
+<!-- For the purpose of examining the effectiveness of visual test in regression diagnostics, two experiments were conducted.
+The experiment I has ideal scenario for conventional testing, where the visual test is not expected to outperform the conventional test.
+The experiment II is a scenario where the conventional test is an approximate test, in which the visual test may have a chance to match the performance of the conventional test.
+-->
+
+<!-- Subjects for both experiments were recruited from an crowdsourcing platform called Prolific [ref here].
+Prescreening procedure was applied during the recruitment, subjects were required to be fluent in English, with $98\%$ minimum approval rate in other studies and 10 minimum submissions.
+-->
+
+<!-- During the experiment, every subject was presented with a block of 20 lineups.
+And for every lineup, the subject was asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots were from others.
+If there was no noticeable difference between plots in a lineup, subjects were permitted to select zero plots without providing the reason.
+No subject was shown the same lineup twice.
+Information about preferred pronoun, age group, education, and previous experience in visual experiment were also collected.
+-->
+
+<!-- A pool of 12 different lineups with obvious visual patterns were generated for experiment I and experiment II respectively.
+In every block of 20 lineups that presented to a subject, two out of 12 lineups were included as attention checks.
+A subject's submission was only accepted if the actual data plot was identified for at least one attention check.
+Data of rejected submissions were discarded automatically to maintain the overall data quality.
+-->
+
+<!-- ## Non-linearity -->
+
+<!-- Experiment I is designed to study the ability of human subjects to detect the effect of a random vector $\boldsymbol{z}$ which is a probabilist's Hermite polynomial [Herimite ref here] of another random vector $\boldsymbol{x}$ in a two variable statistical model formulated as: -->
+
+<!-- ```{=tex} -->
+<!-- \begin{align} \label{eq:nonlinearity-model} -->
+<!-- \boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\ -->
+<!-- \boldsymbol{x} = g(\boldsymbol{x}_{raw}, 1), \\ -->
+<!-- \boldsymbol{z} = g(\boldsymbol{z}_{raw}, 1), \\ -->
+<!-- \boldsymbol{z}_{raw} = He_j(g(\boldsymbol{z}, 2)), -->
+<!-- \end{align} -->
+<!-- ``` -->
+<!-- where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vector of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $\{-k, k\}$ defined as -->
+
+<!-- ```{=tex} -->
+<!-- \begin{equation} \label{eq:scaling-function} -->
+<!-- g(\boldsymbol{x}, k) = (\boldsymbol{x} - min(\boldsymbol{x}))/max(\boldsymbol{x} - min(\boldsymbol{x})) \times 2k - k, \quad \text{for} \quad k > 0.  -->
+<!-- \end{equation} -->
+<!-- ``` -->
+<!-- The null regression model used to fit the realizations generated by the above model is formulated as: -->
+
+<!-- ```{=tex} -->
+<!-- \begin{equation} \label{eq:scaling-function} -->
+<!-- \boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u}, -->
+<!-- \end{equation} -->
+<!-- ``` -->
+<!-- where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$. -->
+
+<!-- Model misspecification presents since the null model leaves out the higher order term. -->
+
+<!-- Experiment data were simulated using four different order of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$), three different sample sizes (n = 50, 100, 300) , four different standard deviations of the error ($\sigma$ = 0.5, 1, 2, 4) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
 A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}.
-
-The values of $a$ was chosen so that different shapes of heteroskedasticity were included in the residual plot. These include triangle shape, butterfly shape and inverse triangle shape.
-
-```{r different-shape-of-heter}
-stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
-
-ori_dat <- heter_model(a = -1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300)
-
-VI_MODEL$plot(ori_dat) + 
-  ggtitle(quote("Triangle shape")) + 
-  theme_light() -> p1
-
-VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote("Butterfly shape")) + 
-  theme_light() -> p2
-
-VI_MODEL$plot(heter_model(a = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote("Inverse triangle shape")) + 
-  theme_light() -> p3
-
-patchwork::wrap_plots(p1, p2, p3, ncol = 3)
-```
-
-The range of $b$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
-
-# Analysis and results
-
-## Data overview
-
-How many people? How many lineups?
-
-Subjects recruited from Prolific received a fixed payment for participating in the experiment.
+-->
+
+<!-- The values of $j$ was chosen so that different shapes of non-linearity were included in the residual plot. These include "U" shape, "S" shape, "M" shape and "Triple-U" shape. -->
+
+<!-- ```{r different-shape-of-herimite} -->
+<!-- stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1 -->
+
+<!-- ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300) -->
+
+<!-- VI_MODEL$plot(ori_dat) +  -->
+<!--   ggtitle(quote(He[2]:"U shape")) +  -->
+<!--   theme_light() -> p1 -->
+
+<!-- VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) + -->
+<!--   ggtitle(quote(He[3]:"S shape")) +  -->
+<!--   theme_light() -> p2 -->
+
+<!-- VI_MODEL$plot(poly_model(shape = 3, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) + -->
+<!--   ggtitle(quote(He[6]:"M shape")) +  -->
+<!--   theme_light() -> p3 -->
+
+<!-- VI_MODEL$plot(poly_model(shape = 4, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) + -->
+<!--   ggtitle(quote(He[18]:"Triple-U shape")) +  -->
+<!--   theme_light() -> p4 -->
+
+<!-- patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2) -->
+<!-- ``` -->
+
+<!-- The range of $\sigma$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
+-->
+
+<!-- ```{r different-sigma} -->
+<!-- ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300) -->
+
+<!-- VI_MODEL$plot(ori_dat) +  -->
+<!--   ggtitle(quote(sigma~"="~0.5)) +  -->
+<!--   theme_light() -> p1 -->
+
+<!-- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1)$gen(300, computed = select(ori_dat, x))) + -->
+<!--   ggtitle(quote(sigma~"="~1)) +  -->
+<!--   theme_light() -> p2 -->
+
+<!-- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 2)$gen(300, computed = select(ori_dat, x))) + -->
+<!--   ggtitle(quote(sigma~"="~2)) +  -->
+<!--   theme_light() -> p3 -->
+
+<!-- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 4)$gen(300, computed = select(ori_dat, x))) + -->
+<!--   ggtitle(quote(sigma~"="~4)) +  -->
+<!--   theme_light() -> p4 -->
+
+<!-- patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2) -->
+<!-- ``` -->
+
+<!-- Four different distribution were used to generate $X_{raw}$.
+The uniform and the normal distribution are symmetric and commonly assumed in statistical models.
+The adjusted log-normal distribution provides skewed density.
+And the discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.
+-->
+
+<!-- ```{r different-dist} -->
+<!-- ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300) -->
+
+<!-- VI_MODEL$plot(ori_dat) +  -->
+<!--   ggtitle(quote(U(-1,1))) +  -->
+<!--   theme_light() -> p1 -->
+
+<!-- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) + -->
+<!--   ggtitle(quote(N(0,0.3^2))) +  -->
+<!--   theme_light() -> p2 -->
+
+<!-- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) + -->
+<!--   ggtitle(quote(lognormal(0,0.6^2)/3)) +  -->
+<!--   theme_light() -> p3 -->
+
+<!-- VI_MODEL$plot(poly_model(shape = 1, x = rand_uniform_d(k = 5, even = TRUE), sigma = 0.5)$gen(300, computed = select(ori_dat, e))) + -->
+<!--   ggtitle(quote(U~"{1,5}")) +  -->
+<!--   theme_light() -> p4 -->
+
+<!-- patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2) -->
+<!-- ``` -->
+
+<!-- Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in 192 different lineups.
+For each lineup, the actual data plot was drawn as a standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis.
+The 19 null datasets were generated by the residual rotation technique, and plotted in the same way.
+The lineup consisted of 20 residual plots with one randomly placed actual data plot.
+Figure \ref{fig:example-lineup} is an example of one of these lineups.
+It was produced by using $n = 300$, $j = 6$, $\sigma = 0.5$ and $X_{raw} \sim N(0.0.3^2)$.
+The actual data plot location was four.
+All five subjects correctly identified the actual data plot for this lineup.
+-->
+
+<!-- ```{r results='asis'} -->
+<!-- data.frame(n = c(50, 100, 300, ""), -->
+<!--            j = c(2, 3, 6, 18), -->
+<!--            sigma = c(0.5, 1, 2, 4), -->
+<!--            x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$")) %>% -->
+<!--   knitr::kable(col.names = NULL, -->
+<!--                escape = FALSE, -->
+<!--                format = "latex", -->
+<!--                booktabs = TRUE,  -->
+<!--                caption  = 'Parameter values for $n$, $j$ $\\sigma$, $X_{raw}$',  -->
+<!--                label = "parameter-table" ) %>% -->
+<!--   kableExtra::add_header_above(c("Sample size\n(sym3)", "Order of Hermite polynomial\n(sym1)", "Error SD\n(sym2)", "Distribution of sym4\n")) %>% -->
+<!--   as.character() %>% -->
+<!--   gsub("sym1", "$j$", .) %>% -->
+<!--   gsub("sym2", "$\\\\sigma$", .) %>% -->
+<!--   gsub("sym3", "$n$", .) %>% -->
+<!--   gsub("sym4", "$X_{raw}$", .) %>% -->
+<!--   cat() -->
+<!--   # gsub("\\\\\\$", "$", .) %>% -->
+<!--   # cat() -->
+<!-- ``` -->
+
+<!-- ```{r example-lineup, fig.cap="Example lineup \\label{fig:example-lineup}"} -->
+<!-- VI_MODEL$plot_lineup(polynomials_lineup[[23]]$data,  -->
+<!--                      remove_grid_line = TRUE,  -->
+<!--                      theme = theme_light(), -->
+<!--                      remove_axis = TRUE) -->
+<!-- ``` -->
+
+<!-- <!-- It was generated for $n = 1$, $\sigma = 1$, $j = 1$, $X_{raw} \sim N$.
+For this lineup, x out of k observers picked the actual data plot.
+Each lineup is evaluated by 5 different subjects to provide reasonable estimates of the p-value.
+--> -->
+
+<!-- In addition, each lineup is designed to be evaluated by five different subjects to provide reasonable estimates of the visual p-value.
+Thus, $192 \times 3 \times 5 / (20 - 2) = 160$ subjects were recruited to satisfy the design of the experiment I.
+-->
+
+<!-- ## Heteroskedasticity -->
+
+<!-- Experiment II is designed to study the ability of human subjects to detect the appearance of a heteroskedasticity pattern under a simple linear regression model setting: -->
+
+<!-- ```{=tex} -->
+<!-- \begin{align} \label{eq:heter-model} -->
+<!-- \boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\ -->
+<!-- \boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + 2 - |a| b (\boldsymbol{x} - a)^2 \boldsymbol{I}), \\ -->
+<!-- \end{align} -->
+<!-- ``` -->
+<!-- where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vector of size $n$. -->
+
+<!-- The null regression model used to fit the realizations generated by the above model is formulated exactly the same as the model used in the first experiment: -->
+
+<!-- ```{=tex} -->
+<!-- \begin{equation}  -->
+<!-- \boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u}, -->
+<!-- \end{equation} -->
+<!-- ``` -->
+<!-- where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$. -->
+
+<!-- Model misspecification presents since the assumption about the constant error variance is violated. -->
+
+<!-- Experiment data were simulated using three different shapes ($a$ = -1, 0, 1), five different values of $b$ (b = 0.25, 1, 4, 16, 64), three different sample sizes (n = 50, 100, 300) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
+A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}.
+-->
+
+<!-- The values of $a$ was chosen so that different shapes of heteroskedasticity were included in the residual plot. These include triangle shape, butterfly shape and inverse triangle shape. -->
+
+<!-- ```{r different-shape-of-heter} -->
+<!-- stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1 -->
+
+<!-- ori_dat <- heter_model(a = -1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300) -->
+
+<!-- VI_MODEL$plot(ori_dat) +  -->
+<!--   ggtitle(quote("Triangle shape")) +  -->
+<!--   theme_light() -> p1 -->
+
+<!-- VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) + -->
+<!--   ggtitle(quote("Butterfly shape")) +  -->
+<!--   theme_light() -> p2 -->
+
+<!-- VI_MODEL$plot(heter_model(a = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) + -->
+<!--   ggtitle(quote("Inverse triangle shape")) +  -->
+<!--   theme_light() -> p3 -->
+
+<!-- patchwork::wrap_plots(p1, p2, p3, ncol = 3) -->
+<!-- ``` -->
+
+<!-- The range of $b$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
+-->
+
+<!-- # Analysis and results -->
+
+<!-- ## Data overview -->
+
+<!-- How many people? How many lineups? -->
+
+<!-- Subjects recruited from Prolific received a fixed payment for participating in the experiment.
 However, some subjects will try to maximize their earnings for minimum effort.
@@
 Therefore, we only paid for approved submissions and no further data screening procedure needed to be applied on the collected data.
-
-A subject was allowed to select zero plots for a lineup if there was no visible difference between plots, but the simulated visual p-value given in Equation (\ref{eq:p-value-multi}) will effectively drop the subject from the simulation for this case, leading to inaccurate estimation of $p$-value.
+-->
+
+<!-- A subject was allowed to select zero plots for a lineup if there was no visible difference between plots, but the simulated visual p-value given in Equation (\ref{eq:p-value-multi}) will effectively drop the subject from the simulation for this case, leading to inaccurate estimation of $p$-value.
 Therefore, we treated this case as making one selection but failing to identify the actual data plot so that Equation (\ref{eq:p-value-multi}) can be applied correctly.
-
-In overall, there were a total of 3200 lineup evaluations made by 160 subjects in both experiment I and experiment II respectively, where 320 lineup evaluations were attention checks and were not used in the following analysis.
-
-The collated dataset is provided in `polynomials` and `heter` of the `visage` `R` package.
-
-##
-
-## Nonlinearity
-
--   Power curve overall
--   Four lineups also shown, selected from ones close to the visual power curve from the uniform treatment
--   Compare visual power against different factors
-
-## Heteroskedasticity
-
--   Power curve overall
--   Four lineups also shown, selected from ones close to the visual power curve from the uniform treatment
--   Compare visual power against different factors
-
-## Demographic summary
-
-Table \ref{tab:demographic-table} tabulates the number of subjects, preferred pronouns, education backgrounds, age groups, and previous experience in visual experiment.
+-->
+
+<!-- In overall, there were a total of 3200 lineup evaluations made by 160 subjects in both experiment I and experiment II respectively, where 320 lineup evaluations were attention checks and were not used in the following analysis.
+-->
+
+<!-- The collated dataset is provided in `polynomials` and `heter` of the `visage` `R` package. -->
+
+<!-- ##  -->
+
+<!-- ## Nonlinearity -->
+
+<!-- -   Power curve overall -->
+<!-- -   Four lineups also shown, selected from ones close to the visual power curve from the uniform treatment -->
+<!-- -   Compare visual power against different factors -->
+
+<!-- ## Heteroskedasticity -->
+
+<!-- -   Power curve overall -->
+<!-- -   Four lineups also shown, selected from ones close to the visual power curve from the uniform treatment -->
+<!-- -   Compare visual power against different factors -->
+
+<!-- ## Demographic summary -->
+
+<!-- Table \ref{tab:demographic-table} tabulates the number of subjects, preferred pronouns, education backgrounds, age groups, and previous experience in visual experiment.
 Figure \ref{fig:demographic-summary-plot} visualizes the marginal distribution of each of the category.
@@
 Around 40% of subjects had previous experience in visual experiment.
-
-```{r results="asis"}
-
-polynomials %>%
-  mutate(exp = 1) %>%
-  bind_rows(mutate(heter, exp = 2)) %>%
-  group_by(exp, set) %>%
-  slice_sample(n = 1) %>%
-  ungroup() %>%
-  count(pronoun, education, age_group, previous_experience) %>%
-  arrange(desc(n)) %>%
-  knitr::kable(col.names = c("Preferred pronoun", "Education", "Age group", "Previous experience", "Subject"),
-               escape = FALSE,
-               format = "latex",
-               linesep = "",
-               booktabs = TRUE, 
-               caption  = 'Summary of demographic information', 
-               label = "demographic-table" ) %>%
-  cat()
-```
-
-```{r demographic-summary-plot, fig.cap="Summary of demographic information. \\label{fig:demographic-summary-plot}"}
-polynomials %>%
-  mutate(exp = 1) %>%
-  bind_rows(mutate(heter, exp = 2)) %>%
-  group_by(exp, set) %>%
-  summarise(across(age_group:previous_experience, first)) %>%
-  mutate(education = fct_relevel(education, "High School or below")) %>%
-  ggplot() +
-  geom_bar(aes(y = education), orientation = "y") -> education_plot
-
-polynomials %>%
-  mutate(exp = 1) %>%
-  bind_rows(mutate(heter, exp = 2)) %>%
-  group_by(exp, set) %>%
-  summarise(across(age_group:previous_experience, first)) %>%
-  ggplot() +
-  geom_bar(aes(age_group)) -> age_plot
-
-polynomials %>%
-  mutate(exp = 1) %>%
-  bind_rows(mutate(heter, exp = 2)) %>%
-  group_by(exp, set) %>%
-  summarise(across(age_group:previous_experience, first)) %>%
-  mutate(pronoun = fct_relevel(pronoun, "He", "She", "They")) %>%
-  ggplot() +
-  geom_bar(aes(pronoun)) -> gender_plot
-
-polynomials %>%
-  mutate(exp = 1) %>%
-  bind_rows(mutate(heter, exp = 2)) %>%
-  group_by(exp, set) %>%
-  summarise(across(age_group:previous_experience, first)) %>%
-  ggplot() +
-  geom_bar(aes(previous_experience)) -> experience_plot
-
-patchwork::wrap_plots(education_plot, age_plot, gender_plot, experience_plot)
-```
-
-## Model fitting
-
-For each parameter combination, effect $E$ is derived from the Kullback-Leibler divergence (see [appendix ref here]) formulated as:
-
-```{=tex}
-\begin{equation} \label{eq:effect-size-ex1}
-E = \frac{1}{2\sigma^2}\boldsymbol{X}_b'\boldsymbol{R}_a'(diag(\boldsymbol{R}_a))^{-1}\boldsymbol{R}_a\boldsymbol{X}_b,
-\end{equation}
-```
-where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$.
-
-The logistic regression is fit using $log_e(E)$ as the only fixed effect covariate for the power of visual test formulated as:
-
-```{=tex}
-\begin{equation} \label{eq:logistic-regression-1-1}
-Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(\boldsymbol{E})),
-\end{equation}
-```
-where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
-
-To study various factors contributing to the power of the visual test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors.
+-->
+
+<!-- ```{r results="asis"} -->
+
+<!-- polynomials %>% -->
+<!--   mutate(exp = 1) %>% -->
+<!--   bind_rows(mutate(heter, exp = 2)) %>% -->
+<!--   group_by(exp, set) %>% -->
+<!--   slice_sample(n = 1) %>% -->
+<!--   ungroup() %>% -->
+<!--   count(pronoun, education, age_group, previous_experience) %>% -->
+<!--   arrange(desc(n)) %>% -->
+<!--   knitr::kable(col.names = c("Preferred pronoun", "Education", "Age group", "Previous experience", "Subject"), -->
+<!--                escape = FALSE, -->
+<!--                format = "latex", -->
+<!--                linesep = "", -->
+<!--                booktabs = TRUE,  -->
+<!--                caption  = 'Summary of demographic information',  -->
+<!--                label = "demographic-table" ) %>% -->
+<!--   cat() -->
+<!-- ``` -->
+
+<!-- ```{r demographic-summary-plot, fig.cap="Summary of demographic information. \\label{fig:demographic-summary-plot}"} -->
+<!-- polynomials %>% -->
+<!--   mutate(exp = 1) %>% -->
+<!--   bind_rows(mutate(heter, exp = 2)) %>% -->
+<!--   group_by(exp, set) %>% -->
+<!--   summarise(across(age_group:previous_experience, first)) %>% -->
+<!--   mutate(education = fct_relevel(education, "High School or below")) %>% -->
+<!--   ggplot() + -->
+<!--   geom_bar(aes(y = education), orientation = "y") -> education_plot -->
+
+<!-- polynomials %>% -->
+<!--   mutate(exp = 1) %>% -->
+<!--   bind_rows(mutate(heter, exp = 2)) %>% -->
+<!--   group_by(exp, set) %>% -->
+<!--   summarise(across(age_group:previous_experience, first)) %>% -->
+<!--   ggplot() + -->
+<!--   geom_bar(aes(age_group)) -> age_plot -->
+
+<!-- polynomials %>% -->
+<!--   mutate(exp = 1) %>% -->
+<!--   bind_rows(mutate(heter, exp = 2)) %>% -->
+<!--   group_by(exp, set) %>% -->
+<!--   summarise(across(age_group:previous_experience, first)) %>% -->
+<!--   mutate(pronoun = fct_relevel(pronoun, "He", "She", "They")) %>% -->
+<!--   ggplot() + -->
+<!--   geom_bar(aes(pronoun)) -> gender_plot -->
+
+<!-- polynomials %>% -->
+<!--   mutate(exp = 1) %>% -->
+<!--   bind_rows(mutate(heter, exp = 2)) %>% -->
+<!--   group_by(exp, set) %>% -->
+<!--   summarise(across(age_group:previous_experience, first)) %>% -->
+<!--   ggplot() + -->
+<!--   geom_bar(aes(previous_experience)) -> experience_plot -->
+
+<!-- patchwork::wrap_plots(education_plot, age_plot, gender_plot, experience_plot) -->
+<!-- ``` -->
+
+<!-- ## Model fitting -->
+
+<!-- For each parameter combination, effect $E$ is derived from the Kullback-Leibler divergence (see [appendix ref here]) formulated as: -->
+
+<!-- ```{=tex} -->
+<!-- \begin{equation} \label{eq:effect-size-ex1} -->
+<!-- E = \frac{1}{2\sigma^2}\boldsymbol{X}_b'\boldsymbol{R}_a'(diag(\boldsymbol{R}_a))^{-1}\boldsymbol{R}_a\boldsymbol{X}_b, -->
+<!-- \end{equation} -->
+<!-- ``` -->
+<!-- where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$. -->
+
+<!-- The logistic regression is fit using $log_e(E)$ as the only fixed effect covariate for the power of visual test formulated as: -->
+
+<!-- ```{=tex} -->
+<!-- \begin{equation} \label{eq:logistic-regression-1-1} -->
+<!-- Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(\boldsymbol{E})), -->
+<!-- \end{equation} -->
+<!-- ``` -->
+<!-- where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$. -->
+
+<!-- To study various factors contributing to the power of the visual test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors.
 This includes [expansion].
-
-Table [table ref here] shows the parameter estimates of the logistic regressions. [Discussion about the numeric estimates here]
-
-## Power comparison
-
-### Experiment I
-
-Figure \ref{fig:power-overview} shows an overview of estimated power of visual test against natural logarithm of the effect with comparison to the power of an exact test - F-test, and the power of two other residual-based conventional tests commonly used in regression diagnostics but for testing other departures from the model assumptions.
+-->
+
+<!-- Table [table ref here] shows the parameter estimates of the logistic regressions. [Discussion about the numeric estimates here] -->
+
+<!-- ## Power comparison -->
+
+<!-- ### Experiment I -->
+
+<!-- Figure \ref{fig:power-overview} shows an overview of estimated power of visual test against natural logarithm of the effect with comparison to the power of an exact test - F-test, and the power of two other residual-based conventional tests commonly used in regression diagnostics but for testing other departures from the model assumptions.
 In overall, the power of all four tests increases as the effect becomes larger.
@@
 The power of Breusch--Pagan test and Shapiro--Wilk test reach around 75% and 63% respectively.
-
-What truly impress us is the huge difference between the estimated power of visual test and the estimated power of F-test.
+-->
+
+<!-- What truly impress us is the huge difference between the estimated power of visual test and the estimated power of F-test.
 The margin is largest at around $log_e(E) = 2$.
@@
 Figure \ref{fig:power-overview} gives example lineups of such cases.
+-->
 
@@
 
-```{r}
-conv_test_plot <- function() {
-  conv_result %>%
-  ggplot() +
-  stat_smooth(geom = "line", 
-              aes(log(effect_size), reject, group = boot_id, col = "F-test"), 
-              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) +
-  stat_smooth(geom = "line", 
-              aes(log(effect_size), reject, col = "F-test"), 
-              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1) +
-  labs(col = "Test") +
-  theme_light(base_size = 5) +
-  xlab(quote(log[e](E))) +
-  ylab("Power")
-}
-
-add_wrong_BP_conv_test_plot_layers <- function(p) {
-  p +
-  stat_smooth(data = wrong_BP_conv_result, geom = "line", 
-              aes(log(effect_size), reject, group = boot_id, col = "BP-test"), 
-              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) +
-  stat_smooth(data = wrong_BP_conv_result, geom = "line", 
-              aes(log(effect_size), reject, col = "BP-test"), 
-              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1)
-}
-
-add_wrong_SW_conv_test_plot_layers <- function(p) {
-  p +
-  stat_smooth(data = wrong_SW_conv_result, geom = "line", 
-              aes(log(effect_size), reject, group = boot_id, col = "SW-test"), 
-              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) +
-  stat_smooth(data = wrong_SW_conv_result, geom = "line", 
-              aes(log(effect_size), reject, col = "SW-test"), 
-              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1)
-}
-```
-
-```{r}
-boot_dat <- function(dat, times = 100) {
-  result <- mutate(dat, boot_id = 1)
-  for (i in 2:times) {
-    result <- bind_rows(result, mutate(dat[sample(1:nrow(dat), replace = TRUE), ], boot_id = i)) 
-  }
-  result
-}
-
-add_visual_test_plot_layers <- function(p, points = c()) {
-  visual_dat <- polynomials %>%
-  group_by(lineup_id) %>%
-  summarise(across(c(type:p_value, prop_detect), first)) %>%
-  filter(lineup_id < 577) %>%
-  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
-  boot_dat(times = 100)
-
-  p +
-  stat_smooth(data = filter(visual_dat, boot_id == 1), geom = "line", 
-              aes(log(effect_size), reject, group = boot_id, col = "Visual"), 
-              method = "glm", method.args = list(family = binomial), se = FALSE, alpha = 1) +
-  stat_smooth(data = visual_dat, geom = "line", 
-              aes(log(effect_size), reject, group = boot_id, col = "Visual"), 
-              method = "glm", method.args = list(family = binomial), se = FALSE, alpha = 0.05) +
-  geom_point(data = filter(visual_dat, boot_id == 1), aes(log(effect_size), prop_detect), alpha = 0.1) + 
-  geom_point(data = filter(visual_dat, boot_id == 1, lineup_id %in% points), aes(log(effect_size), prop_detect), alpha = 1, col = "red", size = 2)
-}
-
-control_col <- function(p, ...) {
-  p + scale_color_manual(values = unlist(list(...)))
-}
-```
-
-```{r eval = FALSE}
-filter(polynomials, lineup_id == 131)
-
-polynomials %>%
-  group_by(lineup_id) %>%
-  summarise(dd = first(abs(log(effect_size) - 5)), p_value = first(p_value), prop_detect = first(prop_detect)) %>%
-  arrange(dd) %>%
-  head(30) %>%
-  summarise(tt = median(prop_detect))
-
-polynomials %>%
-  ggplot() +
-  geom_point(aes(prop_detect, p_value < 0.06))
-
-bottom_plot_2
-
-polynomials %>%
-  filter(lineup_id == 364)
-
-polynomials %>%
-  filter(e_sigma == 4) %>%
-  group_by(lineup_id) %>%
-  summarise(p_value = first(p_value)) %>%
-  mutate(reject = p_value < 0.05) %>%
-  {mean(.$reject)}
-
-4 * 4 * 3 * 5
-```
-
-```{r power-overview, fig.height=7.5, fig.cap="Power overview. \\label{fig:power-overview}"}
-conv_test_plot() %>%
-  add_wrong_BP_conv_test_plot_layers() %>%
-  add_wrong_SW_conv_test_plot_layers() %>%
-  add_visual_test_plot_layers(points = c(415, 131, 322, 378)) +
-  geom_vline(xintercept = 2, linetype = 2) + 
-  geom_vline(xintercept = 3, linetype = 2) +
-  geom_vline(xintercept = 4, linetype = 2) +
-  geom_vline(xintercept = 5, linetype = 2) -> base_plot
-
-# VI_MODEL$plot_lineup(polynomials_lineup$lineup_314$data,
-#                      size = 0.1,
-#                      stroke = 0.2,
-#                      remove_grid_line = TRUE,
-#                      theme = theme_light(base_size = 5),
-#                      remove_axis = TRUE) +
-#   ggtitle(bquote(log[e](E) == 0~", F-test p-value" == .(format(filter(polynomials, lineup_id == 314)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 314)$p_value[1], digits = 3)))) -> bottom_plot_1
-
-VI_MODEL$plot_lineup(polynomials_lineup$lineup_415$data,
-                     size = 0.1,
-                     stroke = 0.2,
-                     remove_grid_line = TRUE,
-                     theme = theme_light(base_size = 5),
-                     remove_axis = TRUE) +
-  ggtitle(bquote(log[e](E) %~~% 2~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 415)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 415)$p_value[1], digits = 3)))) -> bottom_plot_1
-
-VI_MODEL$plot_lineup(polynomials_lineup$lineup_131$data,
-                     size = 0.1,
-                     stroke = 0.2,
-                     remove_grid_line = TRUE,
-                     theme = theme_light(base_size = 5),
-                     remove_axis = TRUE) +
-  ggtitle(bquote(log[e](E) %~~% 3~", F-test p-value" == .(format(filter(polynomials, lineup_id == 131)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 131)$p_value[1], digits = 3)))) -> bottom_plot_2
-
-VI_MODEL$plot_lineup(polynomials_lineup$lineup_322$data,
-                     size = 0.1,
-                     stroke = 0.2,
-                     remove_grid_line = TRUE,
-                     theme = theme_light(base_size = 5),
-                     remove_axis = TRUE) +
-  ggtitle(bquote(log[e](E) %~~% 4~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 322)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 322)$p_value[1], digits = 3)))) -> bottom_plot_3
-
-VI_MODEL$plot_lineup(polynomials_lineup$lineup_378$data,
-                     size = 0.1,
-                     stroke = 0.2,
-                     remove_grid_line = TRUE,
-                     theme = theme_light(base_size = 5),
-                     remove_axis = TRUE) +
-  ggtitle(bquote(log[e](E) %~~% 5~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 378)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 378)$p_value[1], digits = 3)))) -> bottom_plot_4
-
-design <- c(patchwork::area(1, 1, 1, 8), 
-            patchwork::area(2, 1, 2, 4), 
-            patchwork::area(2, 5, 2, 8), 
-            patchwork::area(3, 1, 3, 4), 
-            patchwork::area(3, 5, 3, 8))
-
-patchwork::wrap_plots(base_plot, B = bottom_plot_1, C = bottom_plot_2, D = bottom_plot_3, E = bottom_plot_4, design = design)
-```
-
-### Distributions of regressor
-
-The impact of the distribution of $X_raw$ on the power is shown in Figure \ref{fig:dist-power}.
+<!-- ```{r} -->
+<!-- conv_test_plot <- function() { -->
+<!--   conv_result %>% -->
+<!--   ggplot() + -->
+<!--   stat_smooth(geom = "line",  -->
+<!--               aes(log(effect_size), reject, group = boot_id, col = "F-test"),  -->
+<!--               se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) + -->
+<!--   stat_smooth(geom = "line",  -->
+<!--               aes(log(effect_size), reject, col = "F-test"),  -->
+<!--               se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1) + -->
+<!--   labs(col = "Test") + -->
+<!--   theme_light(base_size = 5) + -->
+<!--   xlab(quote(log[e](E))) + -->
+<!--   ylab("Power") -->
+<!-- } -->
+
+<!-- add_wrong_BP_conv_test_plot_layers <- function(p) { -->
+<!--   p + -->
+<!--   stat_smooth(data = wrong_BP_conv_result, geom = "line",  -->
+<!--               aes(log(effect_size), reject, group = boot_id, col = "BP-test"),  -->
+<!--               se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) + -->
+<!--   stat_smooth(data = wrong_BP_conv_result, geom = "line",  -->
+<!--               aes(log(effect_size), reject, col = "BP-test"),  -->
+<!--               se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1) -->
+<!-- } -->
+
+<!-- add_wrong_SW_conv_test_plot_layers <- function(p) { -->
+<!--   p + -->
+<!--   stat_smooth(data = wrong_SW_conv_result, geom = "line",  -->
+<!--               aes(log(effect_size), reject, group = boot_id, col = "SW-test"),  -->
+<!--               se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) + -->
+<!--   stat_smooth(data = wrong_SW_conv_result, geom = "line",  -->
+<!--               aes(log(effect_size), reject, col = "SW-test"),  -->
+<!--               se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1) -->
+<!-- } -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- boot_dat <- function(dat, times = 100) { -->
+<!--   result <- mutate(dat, boot_id = 1) -->
+<!--   for (i in 2:times) { -->
+<!--     result <- bind_rows(result, mutate(dat[sample(1:nrow(dat), replace = TRUE), ], boot_id = i))  -->
+<!--   } -->
+<!--   result -->
+<!-- } -->
+
+<!-- add_visual_test_plot_layers <- function(p, points = c()) { -->
+<!--   visual_dat <- polynomials %>% -->
+<!--   group_by(lineup_id) %>% -->
+<!--   summarise(across(c(type:p_value, prop_detect), first)) %>% -->
+<!--   filter(lineup_id < 577) %>% -->
+<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
+<!--   boot_dat(times = 100) -->
+
+<!--   p + -->
+<!--   stat_smooth(data = filter(visual_dat, boot_id == 1), geom = "line",  -->
+<!--               aes(log(effect_size), reject, group = boot_id, col = "Visual"),  -->
+<!--               method = "glm", method.args = list(family = binomial), se = FALSE, alpha = 1) + -->
+<!--   stat_smooth(data = visual_dat, geom = "line",  -->
+<!--               aes(log(effect_size), reject, group = boot_id, col = "Visual"),  -->
+<!--               method = "glm", method.args = list(family = binomial), se = FALSE, alpha = 0.05) + -->
+<!--   geom_point(data = filter(visual_dat, boot_id == 1), aes(log(effect_size), prop_detect), alpha = 0.1) +  -->
+<!--   geom_point(data = filter(visual_dat, boot_id == 1, lineup_id %in% points), aes(log(effect_size), prop_detect), alpha = 1, col = "red", size = 2) -->
+<!-- } -->
+
+<!-- control_col <- function(p, ...) { -->
+<!--   p + scale_color_manual(values = unlist(list(...))) -->
+<!-- } -->
+<!-- ``` -->
+
+<!-- ```{r eval = FALSE} -->
+<!-- filter(polynomials, lineup_id == 131) -->
+
+<!-- polynomials %>% -->
+<!--   group_by(lineup_id) %>% -->
+<!--   summarise(dd = first(abs(log(effect_size) - 5)), p_value = first(p_value), prop_detect = first(prop_detect)) %>% -->
+<!--   arrange(dd) %>% -->
+<!--   head(30) %>% -->
+<!--   summarise(tt = median(prop_detect)) -->
+
+<!-- polynomials %>% -->
+<!--   ggplot() + -->
+<!--   geom_point(aes(prop_detect, p_value < 0.06)) -->
+
+<!-- bottom_plot_2 -->
+
+<!-- polynomials %>% -->
+<!--   filter(lineup_id == 364) -->
+
+<!-- polynomials %>% -->
+<!--   filter(e_sigma == 4) %>% -->
+<!--   group_by(lineup_id) %>% -->
+<!--   summarise(p_value = first(p_value)) %>% -->
+<!--   mutate(reject = p_value < 0.05) %>% -->
+<!--   {mean(.$reject)} -->
+
+<!-- 4 * 4 * 3 * 5 -->
+<!-- ``` -->
+
+<!-- ```{r power-overview, fig.height=7.5, fig.cap="Power overview. \\label{fig:power-overview}"} -->
+<!-- conv_test_plot() %>% -->
+<!--   add_wrong_BP_conv_test_plot_layers() %>% -->
+<!--   add_wrong_SW_conv_test_plot_layers() %>% -->
+<!--   add_visual_test_plot_layers(points = c(415, 131, 322, 378)) + -->
+<!--   geom_vline(xintercept = 2, linetype = 2) +  -->
+<!--   geom_vline(xintercept = 3, linetype = 2) + -->
+<!--   geom_vline(xintercept = 4, linetype = 2) + -->
+<!--   geom_vline(xintercept = 5, linetype = 2) -> base_plot -->
+
+<!-- # VI_MODEL$plot_lineup(polynomials_lineup$lineup_314$data, -->
+<!-- #                      size = 0.1, -->
+<!-- #                      stroke = 0.2, -->
+<!-- #                      remove_grid_line = TRUE, -->
+<!-- #                      theme = theme_light(base_size = 5), -->
+<!-- #                      remove_axis = TRUE) + -->
+<!-- #   ggtitle(bquote(log[e](E) == 0~", F-test p-value" == .(format(filter(polynomials, lineup_id == 314)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 314)$p_value[1], digits = 3)))) -> bottom_plot_1 -->
+
+<!-- VI_MODEL$plot_lineup(polynomials_lineup$lineup_415$data, -->
+<!--                      size = 0.1, -->
+<!--                      stroke = 0.2, -->
+<!--                      remove_grid_line = TRUE, -->
+<!--                      theme = theme_light(base_size = 5), -->
+<!--                      remove_axis = TRUE) + -->
+<!--   ggtitle(bquote(log[e](E) %~~% 2~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 415)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 415)$p_value[1], digits = 3)))) -> bottom_plot_1 -->
+
+<!-- VI_MODEL$plot_lineup(polynomials_lineup$lineup_131$data, -->
+<!--                      size = 0.1, -->
+<!--                      stroke = 0.2, -->
+<!--                      remove_grid_line = TRUE, -->
+<!--                      theme = theme_light(base_size = 5), -->
+<!--                      remove_axis = TRUE) + -->
+<!--   ggtitle(bquote(log[e](E) %~~% 3~", F-test p-value" == .(format(filter(polynomials, lineup_id == 131)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 131)$p_value[1], digits = 3)))) -> bottom_plot_2 -->
+
+<!-- VI_MODEL$plot_lineup(polynomials_lineup$lineup_322$data, -->
+<!--                      size = 0.1, -->
+<!--                      stroke = 0.2, -->
+<!--                      remove_grid_line = TRUE, -->
+<!--                      theme = theme_light(base_size = 5), -->
+<!--                      remove_axis = TRUE) + -->
+<!--   ggtitle(bquote(log[e](E) %~~% 4~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 322)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 322)$p_value[1], digits = 3)))) -> bottom_plot_3 -->
+
+<!-- VI_MODEL$plot_lineup(polynomials_lineup$lineup_378$data, -->
+<!--                      size = 0.1, -->
+<!--                      stroke = 0.2, -->
+<!--                      remove_grid_line = TRUE, -->
+<!--                      theme = theme_light(base_size = 5), -->
+<!--                      remove_axis = TRUE) + -->
+<!--   ggtitle(bquote(log[e](E) %~~% 5~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 378)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 378)$p_value[1], digits = 3)))) -> bottom_plot_4 -->
+
+<!-- design <- c(patchwork::area(1, 1, 1, 8),  -->
+<!--             patchwork::area(2, 1, 2, 4),  -->
+<!--             patchwork::area(2, 5, 2, 8),  -->
+<!--             patchwork::area(3, 1, 3, 4),  -->
+<!--             patchwork::area(3, 5, 3, 8)) -->
+
+<!-- patchwork::wrap_plots(base_plot, B = bottom_plot_1, C = bottom_plot_2, D = bottom_plot_3, E = bottom_plot_4, design = design) -->
+<!-- ``` -->
+
+<!-- ### Distributions of regressor -->
+
+<!-- The impact of the distribution of $X_raw$ on the power is shown in Figure \ref{fig:dist-power}.
 The power curve of F-test is stable across different distributions, while the visual test has a steeper power curve for normal and uniform distribution.
@@
 The results indicate those inappropriate residual-based tests are sensitive to the distribution of the regressor.
-
-```{r dist-power, fig.cap="\\label{dist-power}"}
-conv_test_plot() %>%
-  add_wrong_BP_conv_test_plot_layers() %>%
-  add_wrong_SW_conv_test_plot_layers() %>%
-  add_visual_test_plot_layers() +
-  facet_wrap(~x_dist)
-```
-
-#### Shpaes of non-linearity
-
-Figure \ref{fig:shape-power} illustrates the change of power under different shape of non-linearity.
+-->
+
+<!-- ```{r dist-power, fig.cap="\\label{dist-power}"} -->
+<!-- conv_test_plot() %>% -->
+<!--   add_wrong_BP_conv_test_plot_layers() %>% -->
+<!--   add_wrong_SW_conv_test_plot_layers() %>% -->
+<!--   add_visual_test_plot_layers() + -->
+<!--   facet_wrap(~x_dist) -->
+<!-- ``` -->
+
+<!-- #### Shpaes of non-linearity -->
+
+<!-- Figure \ref{fig:shape-power} illustrates the change of power under different shape of non-linearity.
 Similar to the power curves shown in \ref{fig:dist-power}, F-test is stable under different shapes.
@@
 It can be observed that both tests performs the best for U-shape.
-
-```{r shape-power, fig.cap="\\label{shape-power}"}
-conv_test_plot() %>%
-  add_wrong_BP_conv_test_plot_layers() %>%
-  add_wrong_SW_conv_test_plot_layers() %>%
-  add_visual_test_plot_layers() +
-  facet_wrap(~shape)
-```
-
-```{r eval = FALSE}
-conv_test_plot() %>%
-  add_wrong_BP_conv_test_plot_layers() %>%
-  add_wrong_SW_conv_test_plot_layers() %>%
-  add_visual_test_plot_layers() +
-  facet_wrap(~e_sigma)
-```
-
-#### Reasons for making selections
-
-```{r}
-polynomials %>%
-  mutate(reason_tmp = ifelse(reason %in% c("Cluster(s)", "Shape", "Outlier(s)"), reason, "Other")) %>%
-  ggplot() +
-  geom_bar(aes(reason_tmp)) +
-  facet_grid(x_dist~detect) 
-```
-
-### Experiment II
-
-<!-- todo: check the two way interaction for x_dist and a -->
-
-```{r eval = FALSE}
-heter %>%
-  filter(b < 3) %>%
-  group_by(lineup_id) %>%
-  summarise(p_value = first(p_value), n_eval = n()) %>%
-  filter(n_eval > 3) %>%
-  mutate(reject = p_value < 0.05) %>%
-  {mean(.$reject)}
-
-```
-
-```{r}
-map_to_group <- function(b) {
-  round(b / 4) + 1
-}
-
-heter_conv_test_plot <- function() {
-  conv_result_heter %>%
-  mutate(cat_b = map_to_group(b)) %>%
-  ggplot() +
-  stat_smooth(geom = "line", 
-              aes(log(effect_size), reject, group = boot_id, col = "BP-test"), 
-              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) +
-  stat_smooth(geom = "line", 
-              aes(log(effect_size), reject, col = "BP-test"), 
-              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1) +
-  labs(col = "Test") +
-  theme_light(base_size = 5) +
-  xlab(quote(log[e](E))) +
-  ylab("Power")
-}
-
-add_heter_SW_plot_layers <- function(p) {
-  p +
-  stat_smooth(data = mutate(conv_result_heter_SW, cat_b = map_to_group(b)), geom = "line", 
-              aes(log(effect_size), reject, group = boot_id, col = "SW-test"), 
-              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) +
-  stat_smooth(data = mutate(conv_result_heter_SW, cat_b = map_to_group(b)), geom = "line", 
-              aes(log(effect_size), reject, col = "SW-test"), 
-              se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1)
-}
-
-add_heter_visual_test_plot_layers <- function(p, points = c()) {
-
-  visual_dat <- heter %>%
-  mutate(x_dist = ifelse(x_dist == "discrete_uniform", "even_discrete", x_dist)) %>%
-  mutate(x_dist = ifelse(x_dist == "neglognormal", "lognormal", x_dist)) %>%
-  group_by(lineup_id) %>%
-  summarise(across(c(type:p_value, prop_detect), first)) %>%
-  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
-  boot_dat(times = 100) %>%
-  mutate(cat_b = map_to_group(b))
-
-  p +
-  stat_smooth(data = filter(visual_dat, boot_id == 1), geom = "line", 
-              aes(log(effect_size), reject, group = boot_id, col = "Visual"), 
-              method = "glm", method.args = list(family = binomial), se = FALSE, alpha = 1) +
-  stat_smooth(data = visual_dat, geom = "line", 
-              aes(log(effect_size), reject, group = boot_id, col = "Visual"), 
-              method = "glm", method.args = list(family = binomial), se = FALSE, alpha = 0.05) +
-  geom_point(data = filter(visual_dat, boot_id == 1), aes(log(effect_size), prop_detect), alpha = 0.1) + 
-  geom_point(data = filter(visual_dat, boot_id == 1, lineup_id %in% points), aes(log(effect_size), prop_detect), alpha = 1, col = "red", size = 2)
-}
-```
-
-```{r}
-heter_conv_test_plot() %>%
-  add_heter_SW_plot_layers() %>%
-  add_heter_visual_test_plot_layers()
-```
-
-```{r}
-heter_conv_test_plot() %>%
-  add_heter_SW_plot_layers() %>%
-  add_heter_visual_test_plot_layers() +
-  facet_wrap(~a)
-```
-
-```{r}
-heter_conv_test_plot() %>%
-  add_heter_SW_plot_layers() %>%
-  add_heter_visual_test_plot_layers() +
-  facet_grid(x_dist~a)
-```
-
-```{r}
-heter_conv_test_plot() %>%
-  add_heter_SW_plot_layers() %>%
-  add_heter_visual_test_plot_layers() +
-  facet_wrap(~x_dist)
-```
-
-```{r}
-heter_conv_test_plot() %>%
-  add_heter_SW_plot_layers() %>%
-  add_heter_visual_test_plot_layers() +
-  facet_wrap(~cat_b)
-```
+-->
+
+<!-- ```{r shape-power, fig.cap="\\label{shape-power}"} -->
+<!-- conv_test_plot() %>% -->
+<!--   add_wrong_BP_conv_test_plot_layers() %>% -->
+<!--   add_wrong_SW_conv_test_plot_layers() %>% -->
+<!--   add_visual_test_plot_layers() + -->
+<!--   facet_wrap(~shape) -->
+<!-- ``` -->
+
+<!-- ```{r eval = FALSE} -->
+<!-- conv_test_plot() %>% -->
+<!--   add_wrong_BP_conv_test_plot_layers() %>% -->
+<!--   add_wrong_SW_conv_test_plot_layers() %>% -->
+<!--   add_visual_test_plot_layers() + -->
+<!--   facet_wrap(~e_sigma) -->
+<!-- ``` -->
+
+<!-- #### Reasons for making selections -->
+
+<!-- ```{r} -->
+<!-- polynomials %>% -->
+<!--   mutate(reason_tmp = ifelse(reason %in% c("Cluster(s)", "Shape", "Outlier(s)"), reason, "Other")) %>% -->
+<!--   ggplot() + -->
+<!--   geom_bar(aes(reason_tmp)) + -->
+<!--   facet_grid(x_dist~detect)  -->
+<!-- ``` -->
+
+<!-- ### Experiment II -->
+
+<!-- <!-- todo: check the two way interaction for x_dist and a --> -->
+
+<!-- ```{r eval = FALSE} -->
+<!-- heter %>% -->
+<!--   filter(b < 3) %>% -->
+<!--   group_by(lineup_id) %>% -->
+<!--   summarise(p_value = first(p_value), n_eval = n()) %>% -->
+<!--   filter(n_eval > 3) %>% -->
+<!--   mutate(reject = p_value < 0.05) %>% -->
+<!--   {mean(.$reject)} -->
+
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- map_to_group <- function(b) { -->
+<!--   round(b / 4) + 1 -->
+<!-- } -->
+
+<!-- heter_conv_test_plot <- function() { -->
+<!--   conv_result_heter %>% -->
+<!--   mutate(cat_b = map_to_group(b)) %>% -->
+<!--   ggplot() + -->
+<!--   stat_smooth(geom = "line",  -->
+<!--               aes(log(effect_size), reject, group = boot_id, col = "BP-test"),  -->
+<!--               se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) + -->
+<!--   stat_smooth(geom = "line",  -->
+<!--               aes(log(effect_size), reject, col = "BP-test"),  -->
+<!--               se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1) + -->
+<!--   labs(col = "Test") + -->
+<!--   theme_light(base_size = 5) + -->
+<!--   xlab(quote(log[e](E))) + -->
+<!--   ylab("Power") -->
+<!-- } -->
+
+<!-- add_heter_SW_plot_layers <- function(p) { -->
+<!--   p + -->
+<!--   stat_smooth(data = mutate(conv_result_heter_SW, cat_b = map_to_group(b)), geom = "line",  -->
+<!--               aes(log(effect_size), reject, group = boot_id, col = "SW-test"),  -->
+<!--               se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) + -->
+<!--   stat_smooth(data = mutate(conv_result_heter_SW, cat_b = map_to_group(b)), geom = "line",  -->
+<!--               aes(log(effect_size), reject, col = "SW-test"),  -->
+<!--               se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1) -->
+<!-- } -->
+
+<!-- add_heter_visual_test_plot_layers <- function(p, points = c()) { -->
+
+<!--   visual_dat <- heter %>% -->
+<!--   mutate(x_dist = ifelse(x_dist == "discrete_uniform", "even_discrete", x_dist)) %>% -->
+<!--   mutate(x_dist = ifelse(x_dist == "neglognormal", "lognormal", x_dist)) %>% -->
+<!--   group_by(lineup_id) %>% -->
+<!--   summarise(across(c(type:p_value, prop_detect), first)) %>% -->
+<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
+<!--   boot_dat(times = 100) %>% -->
+<!--   mutate(cat_b = map_to_group(b)) -->
+
+<!--   p + -->
+<!--   stat_smooth(data = filter(visual_dat, boot_id == 1), geom = "line",  -->
+<!--               aes(log(effect_size), reject, group = boot_id, col = "Visual"),  -->
+<!--               method = "glm", method.args = list(family = binomial), se = FALSE, alpha = 1) + -->
+<!--   stat_smooth(data = visual_dat, geom = "line",  -->
+<!--               aes(log(effect_size), reject, group = boot_id, col = "Visual"),  -->
+<!--               method = "glm", method.args = list(family = binomial), se = FALSE, alpha = 0.05) + -->
+<!--   geom_point(data = filter(visual_dat, boot_id == 1), aes(log(effect_size), prop_detect), alpha = 0.1) +  -->
+<!--   geom_point(data = filter(visual_dat, boot_id == 1, lineup_id %in% points), aes(log(effect_size), prop_detect), alpha = 1, col = "red", size = 2) -->
+<!-- } -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- heter_conv_test_plot() %>% -->
+<!--   add_heter_SW_plot_layers() %>% -->
+<!--   add_heter_visual_test_plot_layers() -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- heter_conv_test_plot() %>% -->
+<!--   add_heter_SW_plot_layers() %>% -->
+<!--   add_heter_visual_test_plot_layers() + -->
+<!--   facet_wrap(~a) -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- heter_conv_test_plot() %>% -->
+<!--   add_heter_SW_plot_layers() %>% -->
+<!--   add_heter_visual_test_plot_layers() + -->
+<!--   facet_grid(x_dist~a) -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- heter_conv_test_plot() %>% -->
+<!--   add_heter_SW_plot_layers() %>% -->
+<!--   add_heter_visual_test_plot_layers() + -->
+<!--   facet_wrap(~x_dist) -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- heter_conv_test_plot() %>% -->
+<!--   add_heter_SW_plot_layers() %>% -->
+<!--   add_heter_visual_test_plot_layers() + -->
+<!--   facet_wrap(~cat_b) -->
+<!-- ``` -->
 

### 2022-09-20 paper_comparison/paper_comparison/paper_comparison.tex
omit the analysis section for faster knitting.

@@
 heteroskedasticity compared to model structural test and Breusch--Pagan
-test. Meanwhile, as mentioned in the \ref{introduction}
-\protect\hyperlink{introduction}{introduction} section
-
-For the purpose of investigating the differences between conventional
-hypothesis testing and graphical diagnostics in the application of
-linear regression diagnostics, three experiments were conducted.
+test. Meanwhile, as mentioned in Section \ref{introduction} section, we
+keen to understand the differences between conventional hypothesis
+testing and graphical diagnostics in the application of linear
+regression diagnostics and why conventional tests are not preferred by
+data analysts in linear regression diagnostics. Attempting to find the
+answer to these questions, three experiments were conducted.
 
@@
 approximate test, in which the visual test may have a chance to match
-the performance of the conventional test.
+the performance of the conventional test. The experiment III is designed
+for collecting human responses to lineup with only good residual plots
+such that the parameter \(\alpha\) in Equation
+\ref{eq:pvalue-beta-binomial} can be estimated.
 
-Subjects for both experiments were recruited from an crowdsourcing
+Subjects for all three experiments were recruited from an crowdsourcing
 platform called Prolific {[}ref here{]}. Prescreening procedure was
@@
 
-A pool of 12 different lineups with obvious visual patterns were
-generated for experiment I and experiment II respectively. In every
-block of 20 lineups that presented to a subject, two out of 12 lineups
-were included as attention checks. A subject's submission was only
-accepted if the actual data plot was identified for at least one
-attention check. Data of rejected submissions were discarded
-automatically to maintain the overall data quality.
+In every block of 20 lineups that presented to a subject, two lineups
+with obvious visual patterns were included as attention checks. A
+subject's submission was only accepted if the actual data plot was
+identified for at least one attention check. Data of rejected
+submissions were discarded automatically to maintain the overall data
+quality.
 
-This should be a section that summarizes the background, gives the
-motivation to the experiments, and most importantly reflects on the
-title.
+Overall, we planned to collect 7974 evaluations on 1152 uniqued lineups
+performed by 443 subjects.
 
@@
 
-\hypertarget{nonlinearity}{%
-\subsubsection{Nonlinearity}\label{nonlinearity}}
-
-\hypertarget{heteroskedasticity}{%
-\subsubsection{Heteroskedasticity}\label{heteroskedasticity}}
-
-\hypertarget{experimental-setup}{%
-\subsection{Experimental setup}\label{experimental-setup}}
-
-\hypertarget{factors}{%
-\subsubsection{Factors}\label{factors}}
-
-\begin{itemize}
-\tightlist
-\item
-  Treatments for nonlinearity experiment (plots summarising levels)
-\item
-  Treatments for heteroskedasticity experiment (plots summarising
-  levels)
-\end{itemize}
-
-\hypertarget{assigning-subjects}{%
-\subsubsection{Assigning subjects}\label{assigning-subjects}}
-
-\hypertarget{collecting-results}{%
-\subsubsection{Collecting results}\label{collecting-results}}
-
-software/technical
-
-\hypertarget{old-stuff}{%
-\section{Old stuff}\label{old-stuff}}
-
-\hypertarget{computing-power-curves}{%
-\subsection{Computing power curves}\label{computing-power-curves}}
-
-For the purpose of examining the effectiveness of visual test in
-regression diagnostics, two experiments were conducted. The experiment I
-has ideal scenario for conventional testing, where the visual test is
-not expected to outperform the conventional test. The experiment II is a
-scenario where the conventional test is an approximate test, in which
-the visual test may have a chance to match the performance of the
-conventional test.
-
-Subjects for both experiments were recruited from an crowdsourcing
-platform called Prolific {[}ref here{]}. Prescreening procedure was
-applied during the recruitment, subjects were required to be fluent in
-English, with \(98\%\) minimum approval rate in other studies and 10
-minimum submissions.
-
-During the experiment, every subject was presented with a block of 20
-lineups. And for every lineup, the subject was asked to select one or
-more plots that are most different from others, provide a reason for
-their selections, and evaluate how different they think the selected
-plots were from others. If there was no noticeable difference between
-plots in a lineup, subjects were permitted to select zero plots without
-providing the reason. No subject was shown the same lineup twice.
-Information about preferred pronoun, age group, education, and previous
-experience in visual experiment were also collected.
-
-A pool of 12 different lineups with obvious visual patterns were
-generated for experiment I and experiment II respectively. In every
-block of 20 lineups that presented to a subject, two out of 12 lineups
-were included as attention checks. A subject's submission was only
-accepted if the actual data plot was identified for at least one
-attention check. Data of rejected submissions were discarded
-automatically to maintain the overall data quality.
-
 \hypertarget{non-linearity}{%
-\subsection{Non-linearity}\label{non-linearity}}
+\subsubsection{Non-linearity}\label{non-linearity}}
 
@@
 
-Model misspecification presents since the null model leaves out the
-higher order term.
+Since \(z = O(x^j)\), for \(j > 1\), \(z\) is a higher order term leaves
+out by the null regression causing model misspecification.
+
+\hypertarget{heteroskedasticity}{%
+\subsubsection{Heteroskedasticity}\label{heteroskedasticity}}
+
+\hypertarget{experimental-setup}{%
+\subsection{Experimental setup}\label{experimental-setup}}
+
+\hypertarget{factors}{%
+\subsubsection{Factors}\label{factors}}
+
+\begin{itemize}
+\tightlist
+\item
+  Treatments for nonlinearity experiment (plots summarising levels)
+\item
+  Treatments for heteroskedasticity experiment (plots summarising
+  levels)
+\end{itemize}
 
@@
 
-\hypertarget{heteroskedasticity-1}{%
-\subsection{Heteroskedasticity}\label{heteroskedasticity-1}}
+\hypertarget{assigning-subjects}{%
+\subsubsection{Assigning subjects}\label{assigning-subjects}}
 
-Experiment II is designed to study the ability of human subjects to
-detect the appearance of a heteroskedasticity pattern under a simple
-linear regression model setting:
+\hypertarget{collecting-results}{%
+\subsubsection{Collecting results}\label{collecting-results}}
 
-\begin{align} \label{eq:heter-model}
-\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
-\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + 2 - |a| b (\boldsymbol{x} - a)^2 \boldsymbol{I}), \\
-\end{align}
+software/technical
 
-where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
-\(\boldsymbol{\varepsilon}\) are vector of size \(n\).
+--\textgreater{}
 
-The null regression model used to fit the realizations generated by the
-above model is formulated exactly the same as the model used in the
-first experiment:
-
-\begin{equation} 
-\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
-\end{equation}
-
-where
-\(\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\).
-
-Model misspecification presents since the assumption about the constant
-error variance is violated.
-
-Experiment data were simulated using three different shapes (\(a\) = -1,
-0, 1), five different values of \(b\) (b = 0.25, 1, 4, 16, 64), three
-different sample sizes (n = 50, 100, 300) and four different
-distribution of \(X_{raw}\): (1) \(U(-1, 1)\), (2) \(N(0, 0.3^2)\), (3)
-\(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). A summary of the
-parameters used in this experiment is given in Table
-\ref{tab:parameter-table}.
-
-The values of \(a\) was chosen so that different shapes of
-heteroskedasticity were included in the residual plot. These include
-triangle shape, butterfly shape and inverse triangle shape.
-
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-heter-1}
-
-The range of \(b\), which is a factor controlling the strength of the
-signal, was chosen so that different difficulty levels of lineups were
-generated, and therefore, the estimated power curve would be smooth and
-continuous.
-
-\hypertarget{analysis-and-results}{%
-\section{Analysis and results}\label{analysis-and-results}}
-
-\hypertarget{data-overview}{%
-\subsection{Data overview}\label{data-overview}}
-
-How many people? How many lineups?
-
-Subjects recruited from Prolific received a fixed payment for
-participating in the experiment. However, some subjects will try to
-maximize their earnings for minimum effort. During the review of
-submissions, if we found a subject objectively demonstrated clear
-low-effort throughout the experiment, i.e., failed all attention checks,
-we rejected the submission. The rejected submissions will be removed
-immediately, and Prolific will automatically recruit another subject as
-substitution. Therefore, we only paid for approved submissions and no
-further data screening procedure needed to be applied on the collected
-data.
-
-A subject was allowed to select zero plots for a lineup if there was no
-visible difference between plots, but the simulated visual p-value given
-in Equation (\ref{eq:p-value-multi}) will effectively drop the subject
-from the simulation for this case, leading to inaccurate estimation of
-\(p\)-value. Therefore, we treated this case as making one selection but
-failing to identify the actual data plot so that Equation
-(\ref{eq:p-value-multi}) can be applied correctly.
-
-In overall, there were a total of 3200 lineup evaluations made by 160
-subjects in both experiment I and experiment II respectively, where 320
-lineup evaluations were attention checks and were not used in the
-following analysis.
-
-The collated dataset is provided in \texttt{polynomials} and
-\texttt{heter} of the \texttt{visage} \texttt{R} package.
-
-\hypertarget{section}{%
-\subsection{}\label{section}}
-
-\hypertarget{nonlinearity-1}{%
-\subsection{Nonlinearity}\label{nonlinearity-1}}
-
-\begin{itemize}
-\tightlist
-\item
-  Power curve overall
-\item
-  Four lineups also shown, selected from ones close to the visual power
-  curve from the uniform treatment
-\item
-  Compare visual power against different factors
-\end{itemize}
-
-\hypertarget{heteroskedasticity-2}{%
-\subsection{Heteroskedasticity}\label{heteroskedasticity-2}}
-
-\begin{itemize}
-\tightlist
-\item
-  Power curve overall
-\item
-  Four lineups also shown, selected from ones close to the visual power
-  curve from the uniform treatment
-\item
-  Compare visual power against different factors
-\end{itemize}
-
-\hypertarget{demographic-summary}{%
-\subsection{Demographic summary}\label{demographic-summary}}
-
-Table \ref{tab:demographic-table} tabulates the number of subjects,
-preferred pronouns, education backgrounds, age groups, and previous
-experience in visual experiment. Figure
-\ref{fig:demographic-summary-plot} visualizes the marginal distribution
-of each of the category. The collated data was a balanced sample among
-male and female. Most of the ages of subject were between 18 to 39,
-while many of them were between 18 to 24. Very few subjects were awarded
-degree higher than Bachelor Degree. Around 40\% of subjects had previous
-experience in visual experiment.
-
-\begin{table}
-
-\caption{\label{tab:demographic-table}Summary of demographic information}
-\centering
-\begin{tabular}[t]{llllr}
-\toprule
-Preferred pronoun & Education & Age group & Previous experience & Subject\\
-\midrule
-She & Diploma and Bachelor Degree & 18-24 & Yes & 26\\
-She & Diploma and Bachelor Degree & 18-24 & No & 24\\
-He & Diploma and Bachelor Degree & 18-24 & No & 22\\
-He & High School or below & 18-24 & No & 22\\
-She & Diploma and Bachelor Degree & 25-39 & No & 21\\
-He & Diploma and Bachelor Degree & 18-24 & Yes & 19\\
-He & Diploma and Bachelor Degree & 25-39 & No & 19\\
-She & High School or below & 18-24 & No & 19\\
-He & High School or below & 18-24 & Yes & 15\\
-She & Diploma and Bachelor Degree & 25-39 & Yes & 15\\
-He & Diploma and Bachelor Degree & 25-39 & Yes & 12\\
-She & High School or below & 18-24 & Yes & 10\\
-He & Masters Degree & 25-39 & No & 8\\
-He & Masters Degree & 25-39 & Yes & 8\\
-He & High School or below & 25-39 & No & 7\\
-She & Masters Degree & 25-39 & Yes & 7\\
-She & High School or below & 25-39 & No & 6\\
-He & High School or below & 25-39 & Yes & 5\\
-She & Honours Degree & 25-39 & No & 5\\
-She & Honours Degree & 25-39 & Yes & 4\\
-She & Masters Degree & 25-39 & No & 4\\
-He & Honours Degree & 25-39 & Yes & 3\\
-She & Honours Degree & 18-24 & No & 3\\
-He & Diploma and Bachelor Degree & 40-54 & Yes & 2\\
-He & Diploma and Bachelor Degree & 55-64 & No & 2\\
-He & Honours Degree & 25-39 & No & 2\\
-He & Masters Degree & 40-54 & No & 2\\
-She & Diploma and Bachelor Degree & 40-54 & No & 2\\
-She & High School or below & 25-39 & Yes & 2\\
-She & Masters Degree & 18-24 & No & 2\\
-They & Diploma and Bachelor Degree & 18-24 & No & 2\\
-They & Diploma and Bachelor Degree & 25-39 & No & 2\\
-He & Diploma and Bachelor Degree & 40-54 & No & 1\\
-He & High School or below & 40-54 & Yes & 1\\
-He & High School or below & 55-64 & No & 1\\
-He & Honours Degree & 18-24 & No & 1\\
-He & Honours Degree & 18-24 & Yes & 1\\
-He & Honours Degree & 40-54 & Yes & 1\\
-He & Masters Degree & 40-54 & Yes & 1\\
-He & Masters Degree & 55-64 & No & 1\\
-Other & Diploma and Bachelor Degree & 18-24 & No & 1\\
-Other & High School or below & 25-39 & No & 1\\
-She & Diploma and Bachelor Degree & 40-54 & Yes & 1\\
-She & High School or below & 40-54 & No & 1\\
-She & High School or below & 55-64 & Yes & 1\\
-She & Honours Degree & 40-54 & Yes & 1\\
-She & Masters Degree & 40-54 & No & 1\\
-They & High School or below & 18-24 & No & 1\\
-They & High School or below & 18-24 & Yes & 1\\
-They & High School or below & 25-39 & No & 1\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/demographic-summary-plot-1} \caption{Summary of demographic information.
-\label{fig:demographic-summary-plot}}\label{fig:demographic-summary-plot}
-\end{figure}
-
-\hypertarget{model-fitting}{%
-\subsection{Model fitting}\label{model-fitting}}
-
-For each parameter combination, effect \(E\) is derived from the
-Kullback-Leibler divergence (see {[}appendix ref here{]}) formulated as:
-
-\begin{equation} \label{eq:effect-size-ex1}
-E = \frac{1}{2\sigma^2}\boldsymbol{X}_b'\boldsymbol{R}_a'(diag(\boldsymbol{R}_a))^{-1}\boldsymbol{R}_a\boldsymbol{X}_b,
-\end{equation}
-
-where \(diag(.)\) is the diagonal matrix constructed from the diagonal
-elements of \(\boldsymbol{R}_a\).
-
-The logistic regression is fit using \(log_e(E)\) as the only fixed
-effect covariate for the power of visual test formulated as:
-
-\begin{equation} \label{eq:logistic-regression-1-1}
-Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(\boldsymbol{E})),
-\end{equation}
-
-where \(\Lambda(.)\) is the standard logistic function given as
-\(\Lambda(z) = exp(z)/(1+exp(z))\).
-
-To study various factors contributing to the power of the visual test,
-the same logistic regression model is fit on different subsets of the
-collated data grouped by levels of factors. This includes
-{[}expansion{]}.
-
-Table {[}table ref here{]} shows the parameter estimates of the logistic
-regressions. {[}Discussion about the numeric estimates here{]}
-
-\hypertarget{power-comparison}{%
-\subsection{Power comparison}\label{power-comparison}}
-
-\hypertarget{experiment-i}{%
-\subsubsection{Experiment I}\label{experiment-i}}
-
-Figure \ref{fig:power-overview} shows an overview of estimated power of
-visual test against natural logarithm of the effect with comparison to
-the power of an exact test - F-test, and the power of two other
-residual-based conventional tests commonly used in regression
-diagnostics but for testing other departures from the model assumptions.
-In overall, the power of all four tests increases as the effect becomes
-larger. The power curve of F-test climbs aggressively from 25\% to
-around 90\% as \(log_e(E)\) increases from 0 to 2, while others respond
-inactively to the change of effect and remain lower than 25\% throughout
-the period, showing that as an exact test, the F-test is relatively more
-sensitive to the type of model defects that being considered. The power
-of visual test arises steadily and nearly linearly to around 90\% as
-\(log_e(E)\) increases from 2 to 5, suggesting that the effect starts to
-make noticeable impact on the degree of the presence of the designed
-visual features. Other two inappropriate conventional tests shows
-improvement at the same time but at a lower rate. This coincides the
-point made by \citet{cook1982residuals} mentioned in
-\ref{hypothesis-testing} that residual-based tests for a specific type
-of model defect are sensitive to other types of model defects. At
-\(log_e(E) = 6\), the power curve of F-test reaches almost 100\%
-followed by the visual test by a small margin. The power of
-Breusch--Pagan test and Shapiro--Wilk test reach around 75\% and 63\%
-respectively.
-
-What truly impress us is the huge difference between the estimated power
-of visual test and the estimated power of F-test. The margin is largest
-at around \(log_e(E) = 2\). An example lineup is included in Figure
-\ref{fig:power-overview} where none of subjects detect the actual data
-plot positioned at panel 14. It demonstrates that at this level of
-difficulty, the designed visual feature is rarely visible, making the
-actual data plot indistinguishable from residual plots simulated from
-the assumed model. From a communication perspective, given the fact that
-the visual difference is unperceivable, the argument that non-linearity
-present in the fitted model is less convincing to the public even though
-it is true. At around \(log_e(E) = 3\), the margin gets smaller as the
-chance of identifying the actual data plot becomes larger. At this level
-of difficulty, the designed visual features are usually detectable but
-it may not stand out from the lineup as other null plots may happen to
-include outliers or visual patterns that are are considered to be more
-attractive by human, and thus recognized as the most different plot.
-Without knowing the designed visual features beforehand, it is actually
-hard to identify the actual data plot by pure image comparison. The
-corresponding example lineup for \(log_e(E) = 3\) shown in Figure
-\ref{fig:power-overview} has the actual data plot positioned at panel
-20, where two out of five subjects detect it. It can be observed that a
-M-shape is presented in plot 20, but the signal is not strong enough to
-attract all five subjects, resulting in a visual p-value sightly above
-the desired significance level \(\alpha = 0.05\). At \(log_e(E) = 4\)
-and \(log_e(E) = 6\), the designed visual features become much clear and
-attractive, leading to a high percentage of rejection of the null
-hypothesis. Figure \ref{fig:power-overview} gives example lineups of
-such cases.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/power-overview-1} \caption{Power overview. \label{fig:power-overview}}\label{fig:power-overview}
-\end{figure}
-
-\hypertarget{distributions-of-regressor}{%
-\subsubsection{Distributions of
-regressor}\label{distributions-of-regressor}}
-
-The impact of the distribution of \(X_raw\) on the power is shown in
-Figure \ref{fig:dist-power}. The power curve of F-test is stable across
-different distributions, while the visual test has a steeper power curve
-for normal and uniform distribution. BP-test performs worse for discrete
-uniform distribution and uniform distribution but has relatively high
-power for normal distribution. SW-test outperforms BP-test for discrete
-uniform distribution but remains as the worst test for other
-distributions. The results indicate those inappropriate residual-based
-tests are sensitive to the distribution of the regressor.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/dist-power-1} \caption{\label{dist-power}}\label{fig:dist-power}
-\end{figure}
-
-\hypertarget{shpaes-of-non-linearity}{%
-\paragraph{Shpaes of non-linearity}\label{shpaes-of-non-linearity}}
-
-Figure \ref{fig:shape-power} illustrates the change of power under
-different shape of non-linearity. Similar to the power curves shown in
-\ref{fig:dist-power}, F-test is stable under different shapes. The power
-curve of visual test also behaves similarly across different shapes.
-What vary are the power curve of BP-test and SW-test. For Triple-U
-shape, both BP-test and SW-test are insensitive to the change of the
-effect. And for W shape, both tests have almost identical power curves.
-It can be observed that both tests performs the best for U-shape.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/shape-power-1} \caption{\label{shape-power}}\label{fig:shape-power}
-\end{figure}
-
-\hypertarget{reasons-for-making-selections}{%
-\paragraph{Reasons for making
-selections}\label{reasons-for-making-selections}}
-
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-13-1}
-
-\hypertarget{experiment-ii}{%
-\subsubsection{Experiment II}\label{experiment-ii}}
-
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-16-1}
-
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-17-1}
-
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-18-1}
-
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-19-1}
-
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-20-1}
+--\textgreater{}
 

### 2022-09-20 paper_comparison/paper_comparison/paper_comparison.Rmd
add Heteroskedasticity design

@@
 
+Two types of departures were considered with the corresponding data generating process being designed for Experiment I and II.
+
 ### Non-linearity
 
-Experiment I is designed to study the ability of human subjects to detect the effect of a random vector $\boldsymbol{z}$ which is a probabilist's Hermite polynomial [Herimite ref here] of another random vector $\boldsymbol{x}$ in a two variable statistical model formulated as:
+Experiment I is designed to study the ability of human subjects to detect the effect of a non-linear term $\boldsymbol{z}$ which is a probabilist's Hermite polynomial [Herimite ref here] of another random vector $\boldsymbol{x}$ in a two variable statistical model formulated as:
 
@@
 \begin{equation} \label{eq:scaling-function}
-g(\boldsymbol{x}, k) = (\boldsymbol{x} - min(\boldsymbol{x}))/max(\boldsymbol{x} - min(\boldsymbol{x})) \times 2k - k, \quad \text{for} \quad k > 0. 
+g(\boldsymbol{x}, k) = (\boldsymbol{x} - min(\boldsymbol{x}))/max(\boldsymbol{x} - min(\boldsymbol{x}))2k - k, \quad \text{for} \quad k > 0. 
 \end{equation}
@@
 
-Since $z = O(x^j)$, for $j > 1$, $z$ is a higher order term leaves out by the null regression causing model misspecification.
+Since $z = O(x^j)$, for $j > 1$, $z$ is a higher order term leaves out by the null regression, which will result in model misspecification.
 
 ### Heteroskedasticity
+
+Experiment II is designed to study the ability of human subjects to detect the appearance of a heteroskedasticity pattern under a simple linear regression model setting:
+
+```{=tex}
+\begin{align} \label{eq:heter-model}
+\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
+\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + 2 - |a| b (\boldsymbol{x} - a)^2 \boldsymbol{I}), \\
+\end{align}
+```
+where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vector of size $n$.
+
+The null regression model used to fit the realizations generated by the above model is formulated exactly the same as the model used in the first experiment:
+
+```{=tex}
+\begin{equation}
+\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
+\end{equation}
+```
+where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.
+
+For $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the regressor $x$, which will lead to the presence of heteroskedasticity.
 
@@
 
+Experiment data were simulated using three different shapes ($a$ = -1, 0, 1), five different values of $b$ (b = 0.25, 1, 4, 16, 64), three different sample sizes (n = 50, 100, 300) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
+A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}.
+
+The values of $a$ was chosen so that different shapes of heteroskedasticity were included in the residual plot. These include triangle shape, butterfly shape and inverse triangle shape.
+
+```{r different-shape-of-heter}
+stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
+
+ori_dat <- heter_model(a = -1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300)
+
+VI_MODEL$plot(ori_dat) +
+  ggtitle(quote("Triangle shape")) +
+  theme_light() -> p1
+
+VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote("Butterfly shape")) +
+  theme_light() -> p2
+
+VI_MODEL$plot(heter_model(a = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote("Inverse triangle shape")) +
+  theme_light() -> p3
+
+patchwork::wrap_plots(p1, p2, p3, ncol = 3)
+```
+
+The range of $b$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
+
 ### Assigning subjects

### 2022-09-20 paper_comparison/paper_comparison/paper_comparison.tex
add Heteroskedasticity design

@@
 
+Two types of departures were considered with the corresponding data
+generating process being designed for Experiment I and II.
+
 \hypertarget{non-linearity}{%
@@
 Experiment I is designed to study the ability of human subjects to
-detect the effect of a random vector \(\boldsymbol{z}\) which is a
+detect the effect of a non-linear term \(\boldsymbol{z}\) which is a
 probabilist's Hermite polynomial {[}Herimite ref here{]} of another
@@
 Since \(z = O(x^j)\), for \(j > 1\), \(z\) is a higher order term leaves
-out by the null regression causing model misspecification.
+out by the null regression, which will result in model misspecification.
 
@@
 \subsubsection{Heteroskedasticity}\label{heteroskedasticity}}
+
+Experiment II is designed to study the ability of human subjects to
+detect the appearance of a heteroskedasticity pattern under a simple
+linear regression model setting:
+
+\begin{align} \label{eq:heter-model}
+\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
+\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + 2 - |a| b (\boldsymbol{x} - a)^2 \boldsymbol{I}), \\
+\end{align}
+
+where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
+\(\boldsymbol{\varepsilon}\) are vector of size \(n\).
+
+The null regression model used to fit the realizations generated by the
+above model is formulated exactly the same as the model used in the
+first experiment:
+
+\begin{equation}
+\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
+\end{equation}
+
+where
+\(\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\).
+
+For \(b \neq 0\), the variance-covariance matrix of the error term
+\(\boldsymbol{\varepsilon}\) is correlated with the regressor \(x\),
+which will lead to the presence of heteroskedasticity.
 
@@
 
+Experiment data were simulated using three different shapes (\(a\) = -1,
+0, 1), five different values of \(b\) (b = 0.25, 1, 4, 16, 64), three
+different sample sizes (n = 50, 100, 300) and four different
+distribution of \(X_{raw}\): (1) \(U(-1, 1)\), (2) \(N(0, 0.3^2)\), (3)
+\(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). A summary of the
+parameters used in this experiment is given in Table
+\ref{tab:parameter-table}.
+
+The values of \(a\) was chosen so that different shapes of
+heteroskedasticity were included in the residual plot. These include
+triangle shape, butterfly shape and inverse triangle shape.
+
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-heter-1}
+
+The range of \(b\), which is a factor controlling the strength of the
+signal, was chosen so that different difficulty levels of lineups were
+generated, and therefore, the estimated power curve would be smooth and
+continuous.
+
 \hypertarget{assigning-subjects}{%

### 2022-09-21 paper_comparison/paper_comparison/paper_comparison.Rmd
update exp design for exp I

@@
 output: rticles::tf_article
+editor_options: 
+  chunk_output_type: inline
 ---
@@
 During the experiment, every subject was presented with a block of 20 lineups.
+For each lineup, the actual data plot was drawn as a standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis.
+An additional horizontal red line was added at $y = 0$ as a helping line.
+The 19 null datasets were generated by the residual rotation technique, and plotted in the same way.
+The lineup consisted of 20 residual plots with one randomly placed actual data plot.
 And for every lineup, the subject was asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots were from others.
@@
 
-Two types of departures were considered with the corresponding data generating process being designed for Experiment I and II.
+Two types of departures were considered with the corresponding data generating process being designed for experiment I and II.
 
@@
 ```{=tex}
-\begin{equation} \label{eq:scaling-function}
+\begin{equation} \label{eq:null-model}
 \boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
@@
 \boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
+\boldsymbol{x} = g(\boldsymbol{x}_{raw}, 1)\\
 \boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + 2 - |a| b (\boldsymbol{x} - a)^2 \boldsymbol{I}), \\
@@
 ```
-where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vector of size $n$.
-
-The null regression model used to fit the realizations generated by the above model is formulated exactly the same as the model used in the first experiment:
-
-```{=tex}
-\begin{equation}
-\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
-\end{equation}
-```
-where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.
+where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vector of size $n$ and $g(.)$ is the scaling function defined in \ref{eq:scaling-function}.
+
+The null regression model used to fit the realizations generated by the above model is formulated exactly the same as Equation \ref{eq:null-model}.
 
@@
 
-### Factors
+### Treatments
+
+> Experiment I
 
 -   Treatments for nonlinearity experiment (plots summarising levels)
--   Treatments for heteroskedasticity experiment (plots summarising levels)
-
-Experiment data were simulated using four different order of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$), three different sample sizes (n = 50, 100, 300) , four different standard deviations of the error ($\sigma$ = 0.5, 1, 2, 4) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
+
+Data of experiment I were simulated using four different order of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$), three different sample sizes (n = 50, 100, 300), four different standard deviations of the error ($\sigma$ = 0.5, 1, 2, 4) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
 A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}.
-
-The values of $j$ was chosen so that different shapes of non-linearity were included in the residual plot. These include "U" shape, "S" shape, "M" shape and "Triple-U" shape.
-
-```{r different-shape-of-herimite}
-stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
-
-ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300)
-
-VI_MODEL$plot(ori_dat) + 
-  ggtitle(quote(He[2]:"U shape")) + 
-  theme_light() -> p1
-
-VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote(He[3]:"S shape")) + 
-  theme_light() -> p2
-
-VI_MODEL$plot(poly_model(shape = 3, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote(He[6]:"M shape")) + 
-  theme_light() -> p3
-
-VI_MODEL$plot(poly_model(shape = 4, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote(He[18]:"Triple-U shape")) + 
-  theme_light() -> p4
-
-patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)
-```
-
-The range of $\sigma$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
-
-```{r different-sigma}
-ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)
-
-VI_MODEL$plot(ori_dat) + 
-  ggtitle(quote(sigma~"="~0.5)) + 
-  theme_light() -> p1
-
-VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1)$gen(300, computed = select(ori_dat, x))) +
-  ggtitle(quote(sigma~"="~1)) + 
-  theme_light() -> p2
-
-VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 2)$gen(300, computed = select(ori_dat, x))) +
-  ggtitle(quote(sigma~"="~2)) + 
-  theme_light() -> p3
-
-VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 4)$gen(300, computed = select(ori_dat, x))) +
-  ggtitle(quote(sigma~"="~4)) + 
-  theme_light() -> p4
-
-patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)
-```
-
-Four different distribution were used to generate $X_{raw}$.
-The uniform and the normal distribution are symmetric and commonly assumed in statistical models.
-The adjusted log-normal distribution provides skewed density.
-And the discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.
-
-```{r different-dist}
-ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)
-
-VI_MODEL$plot(ori_dat) + 
-  ggtitle(quote(U(-1,1))) + 
-  theme_light() -> p1
-
-VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
-  ggtitle(quote(N(0,0.3^2))) + 
-  theme_light() -> p2
-
-VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
-  ggtitle(quote(lognormal(0,0.6^2)/3)) + 
-  theme_light() -> p3
-
-VI_MODEL$plot(poly_model(shape = 1, x = rand_uniform_d(k = 5, even = TRUE), sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
-  ggtitle(quote(U~"{1,5}")) + 
-  theme_light() -> p4
-
-patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)
-```
-
-Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in 192 different lineups.
-For each lineup, the actual data plot was drawn as a standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis.
-The 19 null datasets were generated by the residual rotation technique, and plotted in the same way.
-The lineup consisted of 20 residual plots with one randomly placed actual data plot.
-Figure \ref{fig:example-lineup} is an example of one of these lineups.
-It was produced by using $n = 300$, $j = 6$, $\sigma = 0.5$ and $X_{raw} \sim N(0.0.3^2)$.
-The actual data plot location was four.
-All five subjects correctly identified the actual data plot for this lineup.
 
@@
   knitr::kable(col.names = NULL,
+               align = "rrrl",
                escape = FALSE,
@@
 
-```{r example-lineup, fig.cap="Example lineup \\label{fig:example-lineup}"}
-VI_MODEL$plot_lineup(polynomials_lineup[[23]]$data, 
+The values of $j$ was chosen so that different shapes of non-linearity were included in the residual plot.
+As shown in Figure \ref{fig:different-shape-of-herimite}, these include "U" shape, "S" shape, "M" shape and "Triple-U" shape.
+
+```{r different-shape-of-herimite, fig.cap="Example residual plots of the linear regression model used in experiment I.
+Four different values of $j$ are included in the experiment results in four different shapes of residuals."}
+stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
+
+ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300)
+
+VI_MODEL$plot(ori_dat) + 
+  ggtitle(quote(He[2]:"U shape")) + 
+  theme_light() -> p1
+
+VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote(He[3]:"S shape")) + 
+  theme_light() -> p2
+
+VI_MODEL$plot(poly_model(shape = 3, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote(He[6]:"M shape")) + 
+  theme_light() -> p3
+
+VI_MODEL$plot(poly_model(shape = 4, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote(He[18]:"Triple-U shape")) + 
+  theme_light() -> p4
+
+patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)
+```
+
+The range of $\sigma$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
+The example residual plots are given in Figure \ref{fig:different-sigma}.
+
+```{r different-sigma, fig.cap="Example residual plots of the linear regression model used in experiment I.
+Four different values of $\\sigma$ are included in the experiment to control the strength of the signal."}
+ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)
+
+VI_MODEL$plot(ori_dat) + 
+  ggtitle(quote(sigma~"="~0.5)) + 
+  theme_light() -> p1
+
+VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1)$gen(300, computed = select(ori_dat, x))) +
+  ggtitle(quote(sigma~"="~1)) + 
+  theme_light() -> p2
+
+VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 2)$gen(300, computed = select(ori_dat, x))) +
+  ggtitle(quote(sigma~"="~2)) + 
+  theme_light() -> p3
+
+VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 4)$gen(300, computed = select(ori_dat, x))) +
+  ggtitle(quote(sigma~"="~4)) + 
+  theme_light() -> p4
+
+patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)
+```
+
+Four different distribution were used to generate $X_{raw}$ shown in Figure \ref{fig:different-dist}.
+The uniform and the normal distribution are symmetric and commonly assumed in statistical models.
+The adjusted log-normal distribution provides skewed density.
+And the discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.
+
+```{r different-dist, fig.cap="Example residual plots of the linear regression model used in experiment I.
+Four different distribution of $x_{raw}$ are used in the experiment to provide various visual patterns."}
+ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)
+
+VI_MODEL$plot(ori_dat) + 
+  ggtitle(quote(U(-1,1))) + 
+  theme_light() -> p1
+
+VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
+  ggtitle(quote(N(0,0.3^2))) + 
+  theme_light() -> p2
+
+VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
+  ggtitle(quote(lognormal(0,0.6^2)/3)) + 
+  theme_light() -> p3
+
+VI_MODEL$plot(poly_model(shape = 1, x = rand_uniform_d(k = 5, even = TRUE), sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
+  ggtitle(quote(U~"{1,5}")) + 
+  theme_light() -> p4
+
+patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)
+```
+
+Figure \ref{fig:example-lineup} shows one of the lineups used in experiment I.
+This lineup was produced under $n = 300$, $j = 6$, $\sigma = 1$ and $X_{raw} \sim N(0.0.3^2)$.
+The actual data plot location was four.
+All five subjects correctly identified the actual data plot for this lineup.
+
+```{r example-lineup, fig.cap="Lineup 24 in experiment I. Can you spot the most different plot? \\label{fig:example-lineup}"}
+VI_MODEL$plot_lineup(polynomials_lineup[[24]]$data, 
                      remove_grid_line = TRUE, 
@@
 
-In addition, each lineup is designed to be evaluated by five different subjects to provide reasonable estimates of the visual p-value.
-Thus, $192 \times 3 \times 5 / (20 - 2) = 160$ subjects were recruited to satisfy the design of the experiment I.
+> Experiment II
+
+-   Treatments for heteroskedasticity experiment (plots summarising levels)
 
@@
 ### Assigning subjects
+
+Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in 192 different lineups.
+For each lineup, the actual data plot was drawn as a standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis.
+The 19 null datasets were generated by the residual rotation technique, and plotted in the same way.
+The lineup consisted of 20 residual plots with one randomly placed actual data plot.
+Figure \ref{fig:example-lineup} is an example of one of these lineups.
+It was produced by using $n = 300$, $j = 6$, $\sigma = 0.5$ and $X_{raw} \sim N(0.0.3^2)$.
+The actual data plot location was four.
+All five subjects correctly identified the actual data plot for this lineup.
+
+In addition, each lineup is designed to be evaluated by five different subjects to provide reasonable estimates of the visual p-value.
+Thus, $192 \times 3 \times 5 / (20 - 2) = 160$ subjects were recruited to satisfy the design of the experiment I.
 

### 2022-09-21 paper_comparison/paper_comparison/paper_comparison.tex
update exp design for exp I

@@
 During the experiment, every subject was presented with a block of 20
-lineups. And for every lineup, the subject was asked to select one or
-more plots that are most different from others, provide a reason for
-their selections, and evaluate how different they think the selected
-plots were from others. If there was no noticeable difference between
-plots in a lineup, subjects were permitted to select zero plots without
-providing the reason. No subject was shown the same lineup twice.
-Information about preferred pronoun, age group, education, and previous
-experience in visual experiment were also collected.
+lineups. For each lineup, the actual data plot was drawn as a standard
+residual plot of the null model with raw residuals on the y-axis and
+fitted values on the x-axis. An additional horizontal red line was added
+at \(y = 0\) as a helping line. The 19 null datasets were generated by
+the residual rotation technique, and plotted in the same way. The lineup
+consisted of 20 residual plots with one randomly placed actual data
+plot. And for every lineup, the subject was asked to select one or more
+plots that are most different from others, provide a reason for their
+selections, and evaluate how different they think the selected plots
+were from others. If there was no noticeable difference between plots in
+a lineup, subjects were permitted to select zero plots without providing
+the reason. No subject was shown the same lineup twice. Information
+about preferred pronoun, age group, education, and previous experience
+in visual experiment were also collected.
 
@@
 Two types of departures were considered with the corresponding data
-generating process being designed for Experiment I and II.
+generating process being designed for experiment I and II.
 
@@
 
-\begin{equation} \label{eq:scaling-function}
+\begin{equation} \label{eq:null-model}
 \boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
@@
 where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
-\(\boldsymbol{\varepsilon}\) are vector of size \(n\).
+\(\boldsymbol{\varepsilon}\) are vector of size \(n\) and \(g(.)\) is
+the scaling function defined in \ref{eq:scaling-function}.
 
 The null regression model used to fit the realizations generated by the
-above model is formulated exactly the same as the model used in the
-first experiment:
-
-\begin{equation}
-\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
-\end{equation}
-
-where
-\(\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\).
+above model is formulated exactly the same as Equation
+\ref{eq:null-model}.
 
@@
 
-\hypertarget{factors}{%
-\subsubsection{Factors}\label{factors}}
+\hypertarget{treatments}{%
+\subsubsection{Treatments}\label{treatments}}
+
+\begin{quote}
+Experiment I
+\end{quote}
 
@@
   Treatments for nonlinearity experiment (plots summarising levels)
-\item
-  Treatments for heteroskedasticity experiment (plots summarising
-  levels)
 \end{itemize}
 
-Experiment data were simulated using four different order of
+Data of experiment I were simulated using four different order of
 probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)), three different
-sample sizes (n = 50, 100, 300) , four different standard deviations of
+sample sizes (n = 50, 100, 300), four different standard deviations of
 the error (\(\sigma\) = 0.5, 1, 2, 4) and four different distribution of
@@
 
-The values of \(j\) was chosen so that different shapes of non-linearity
-were included in the residual plot. These include ``U'' shape, ``S''
-shape, ``M'' shape and ``Triple-U'' shape.
-
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-herimite-1}
-
-The range of \(\sigma\), which is a factor controlling the strength of
-the signal, was chosen so that different difficulty levels of lineups
-were generated, and therefore, the estimated power curve would be smooth
-and continuous.
-
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-sigma-1}
-
-Four different distribution were used to generate \(X_{raw}\). The
-uniform and the normal distribution are symmetric and commonly assumed
-in statistical models. The adjusted log-normal distribution provides
-skewed density. And the discrete uniform distribution provides
-discreteness in residual plot, which could enrich the pool of visual
-patterns.
-
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-dist-1}
-
-Three replications are made for each of the parameter values shown in
-Table \ref{tab:parameter-table} resulting in 192 different lineups. For
-each lineup, the actual data plot was drawn as a standard residual plot
-of the null model with raw residuals on the y-axis and fitted values on
-the x-axis. The 19 null datasets were generated by the residual rotation
-technique, and plotted in the same way. The lineup consisted of 20
-residual plots with one randomly placed actual data plot. Figure
-\ref{fig:example-lineup} is an example of one of these lineups. It was
-produced by using \(n = 300\), \(j = 6\), \(\sigma = 0.5\) and
-\(X_{raw} \sim N(0.0.3^2)\). The actual data plot location was four. All
-five subjects correctly identified the actual data plot for this lineup.
-
 \begin{table}
@@
 \centering
-\begin{tabular}[t]{lrrl}
+\begin{tabular}[t]{rrrl}
 \toprule
@@
 
+The values of \(j\) was chosen so that different shapes of non-linearity
+were included in the residual plot. As shown in Figure
+\ref{fig:different-shape-of-herimite}, these include ``U'' shape, ``S''
+shape, ``M'' shape and ``Triple-U'' shape.
+
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-lineup-1} \caption{Example lineup \label{fig:example-lineup}}\label{fig:example-lineup}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-herimite-1} \caption{Example residual plots of the linear regression model used in experiment I.
+Four different values of $j$ are included in the experiment results in four different shapes of residuals.}\label{fig:different-shape-of-herimite}
 \end{figure}
 
-In addition, each lineup is designed to be evaluated by five different
-subjects to provide reasonable estimates of the visual p-value. Thus,
-\(192 \times 3 \times 5 / (20 - 2) = 160\) subjects were recruited to
-satisfy the design of the experiment I.
+The range of \(\sigma\), which is a factor controlling the strength of
+the signal, was chosen so that different difficulty levels of lineups
+were generated, and therefore, the estimated power curve would be smooth
+and continuous. The example residual plots are given in Figure
+\ref{fig:different-sigma}.
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-sigma-1} \caption{Example residual plots of the linear regression model used in experiment I.
+Four different values of $\sigma$ are included in the experiment to control the strength of the signal.}\label{fig:different-sigma}
+\end{figure}
+
+Four different distribution were used to generate \(X_{raw}\) shown in
+Figure \ref{fig:different-dist}. The uniform and the normal distribution
+are symmetric and commonly assumed in statistical models. The adjusted
+log-normal distribution provides skewed density. And the discrete
+uniform distribution provides discreteness in residual plot, which could
+enrich the pool of visual patterns.
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-dist-1} \caption{Example residual plots of the linear regression model used in experiment I.
+Four different distribution of $x_{raw}$ are used in the experiment to provide various visual patterns.}\label{fig:different-dist}
+\end{figure}
+
+Figure \ref{fig:example-lineup} shows one of the lineups used in
+experiment I. This lineup was produced under \(n = 300\), \(j = 6\),
+\(\sigma = 1\) and \(X_{raw} \sim N(0.0.3^2)\). The actual data plot
+location was four. All five subjects correctly identified the actual
+data plot for this lineup.
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-lineup-1} \caption{Lineup 24 in experiment I.
+Can you spot the most different plot?
+\label{fig:example-lineup}}\label{fig:example-lineup}
+\end{figure}
+
+\begin{quote}
+Experiment II
+\end{quote}
+
+\begin{itemize}
+\tightlist
+\item
+  Treatments for heteroskedasticity experiment (plots summarising
+  levels)
+\end{itemize}
 
@@
 
+Three replications are made for each of the parameter values shown in
+Table \ref{tab:parameter-table} resulting in 192 different lineups. For
+each lineup, the actual data plot was drawn as a standard residual plot
+of the null model with raw residuals on the y-axis and fitted values on
+the x-axis. The 19 null datasets were generated by the residual rotation
+technique, and plotted in the same way. The lineup consisted of 20
+residual plots with one randomly placed actual data plot. Figure
+\ref{fig:example-lineup} is an example of one of these lineups. It was
+produced by using \(n = 300\), \(j = 6\), \(\sigma = 0.5\) and
+\(X_{raw} \sim N(0.0.3^2)\). The actual data plot location was four. All
+five subjects correctly identified the actual data plot for this lineup.
+
+In addition, each lineup is designed to be evaluated by five different
+subjects to provide reasonable estimates of the visual p-value. Thus,
+\(192 \times 3 \times 5 / (20 - 2) = 160\) subjects were recruited to
+satisfy the design of the experiment I.
+
 \hypertarget{collecting-results}{%

### 2022-09-26 paper_comparison/paper_comparison/paper_comparison.Rmd
shorten the data preprocessing

@@
 ```{r get-lineup-data}
-if (!file.exists(here::here("data/third_study/polynomials_lineup.rds"))) {
-  polynomials_lineup <- get_polynomials_lineup()
-  saveRDS(polynomials_lineup, here::here("data/third_study/polynomials_lineup.rds"))
+if (!file.exists(here::here("data/vi_lineup.rds"))) {
+  vi_lineup <- get_vi_lineup()
+  saveRDS(vi_lineup, here::here("data/vi_lineup.rds"))
 } else {
-  polynomials_lineup <- readRDS(here::here("data/third_study/polynomials_lineup.rds"))
+  vi_lineup <- readRDS(here::here("data/vi_lineup.rds"))
 }
-
-if (!file.exists(here::here("data/fourth_study/heter_lineup.rds"))) {
-  heter_lineup <- get_polynomials_lineup()
-  saveRDS(heter_lineup, here::here("data/fourth_study/heter_lineup.rds"))
-} else {
-  heter_lineup <- readRDS(here::here("data/fourth_study/heter_lineup.rds"))
-}
-```
-
-```{r get-conv-result}
+```
+
+```{r}
 stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
 
-if (!file.exists(here::here("data/third_study/conv_result.rds"))) {
-  conv_result <- list()
+if (!file.exists(here::here("data/poly_conventional_simulation.rds"))) {
+  poly_conv_sim <- list()
   for (i in 1:100)
   {
-    conv_result[[i]] <- map(1:2000,
+    poly_conv_sim[[i]] <- map(1:2000,
                             function(x) {
@@
                             effect_size = mod$effect_size(tmp_dat),
-                            p_value = mod$test(tmp_dat)$p_value,
-                            reject = p_value < 0.05,
+                            F_p_value = mod$test(tmp_dat)$p_value,
+                            BP_p_value = HETER_MODEL$test(tmp_dat)$p_value,
+                            SW_p_value = shapiro.test(tmp_dat$.resid)$p.value,
                             boot_id = i)
                    }) %>%
-  reduce(bind_rows) %>%
-  mutate(reject = as.numeric(reject))
+  reduce(bind_rows)
   }
 
-  conv_result <- conv_result %>%
+  poly_conv_sim <- poly_conv_sim %>%
     reduce(bind_rows)
 
-  saveRDS(conv_result, here::here("data/third_study/conv_result.rds"))
+  saveRDS(poly_conv_sim, here::here("data/poly_conventional_simulation.rds"))
 } else {
-  conv_result <- readRDS(here::here("data/third_study/conv_result.rds"))
+  poly_conv_sim <- readRDS(here::here("data/poly_conventional_simulation.rds"))
 }
-
 ```
@@
 
-if (!file.exists(here::here("data/fourth_study/conv_result_heter.rds"))) {
-  conv_result_heter <- list()
+if (!file.exists(here::here("data/heter_conventional_simulation.rds"))) {
+  heter_conv_sim <- list()
   for (i in 1:100)
   {
-    conv_result_heter[[i]] <- map(1:2000,
+    heter_conv_sim[[i]] <- map(1:2000,
                             function(x) {
@@
                             effect_size = mod$effect_size(tmp_dat, type = "kl"),
-                            p_value = mod$test(tmp_dat)$p_value,
-                            reject = p_value < 0.05,
+                            F_p_value = POLY_MODEL$test(mutate(tmp_dat, z = poly_model()$gen(n, computed = select(tmp_dat, x))$z))$p_value,
+                            BP_p_value = mod$test(tmp_dat)$p_value,
+                            SW_p_value = shapiro.test(tmp_dat$.resid)$p.value,
                             boot_id = i)
                    }) %>%
-  reduce(bind_rows) %>%
-  mutate(reject = as.numeric(reject))
+  reduce(bind_rows)
   }
 
-  conv_result_heter <- conv_result_heter %>%
+  heter_conv_sim <- heter_conv_sim %>%
     reduce(bind_rows)
 
-  saveRDS(conv_result_heter, here::here("data/fourth_study/conv_result_heter.rds"))
+  saveRDS(heter_conv_sim, here::here("data/heter_conventional_simulation.rds"))
 } else {
-  conv_result_heter <- readRDS(here::here("data/fourth_study/conv_result_heter.rds"))
+  heter_conv_sim <- readRDS(here::here("data/heter_conventional_simulation.rds"))
 }
-
-```
-
-```{r}
-stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
-
-if (!file.exists(here::here("data/fourth_study/conv_result_heter_SW.rds"))) {
-  conv_result_heter_SW <- list()
-  for (i in 1:100)
-  {
-    conv_result_heter_SW[[i]] <- map(1:2000,
-                            function(x) {
-                     a <- sample(c(-1, 0, 1), 1)
-                     b <- sample(c(0.25, 1, 4, 16, 64), 1)
-                     x_dist <- sample(c("uniform", "normal", "lognormal", "even_discrete"), 1)
-                     x <- switch(x_dist,
-                                 uniform = rand_uniform(-1, 1),
-                                 normal = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))},
-                                 lognormal = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))},
-                                 even_discrete = rand_uniform_d(-1, 1, k = 5, even = TRUE))
-                     mod <- heter_model(a = a, b = b, x = x)
-                     n <- sample(c(50, 100, 300), 1)
-                     tmp_dat <- mod$gen(n)
-
-                     tibble(a = a,
-                            b = b,
-                            x_dist = x_dist,
-                            n = n,
-                            effect_size = mod$effect_size(tmp_dat, type = "kl"),
-                            p_value = shapiro.test(tmp_dat$.resid)$p.value,
-                            reject = p_value < 0.05,
-                            boot_id = i)
-                   }) %>%
-  reduce(bind_rows) %>%
-  mutate(reject = as.numeric(reject))
-  }
-
-  conv_result_heter_SW <- conv_result_heter_SW %>%
-    reduce(bind_rows)
-
-  saveRDS(conv_result_heter_SW, here::here("data/fourth_study/conv_result_heter_SW.rds"))
-} else {
-  conv_result_heter_SW <- readRDS(here::here("data/fourth_study/conv_result_heter_SW.rds"))
-}
-
-```
-
-```{r}
-if (!file.exists(here::here("data/third_study/wrong_BP_conv_result.rds"))) {
-  wrong_BP_conv_result <- list()
-  for (i in 1:100)
-  {
-    wrong_BP_conv_result[[i]] <- map(1:2000,
-                            function(x) {
-                     shape <- sample(1:4, 1)
-                     e_sigma <- sample(c(0.5, 1, 2, 4), 1)
-                     x_dist <- sample(c("uniform", "normal", "lognormal", "even_discrete"), 1)
-                     x <- switch(x_dist,
-                                 uniform = rand_uniform(-1, 1),
-                                 normal = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))},
-                                 lognormal = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))},
-                                 even_discrete = rand_uniform_d(k = 5, even = TRUE))
-                     mod <- poly_model(shape, x = x, sigma = e_sigma)
-                     n <- sample(c(50, 100, 300), 1)
-                     tmp_dat <- mod$gen(n)
-
-                     tibble(shape = shape,
-                            e_sigma = e_sigma,
-                            x_dist = x_dist,
-                            n = n,
-                            effect_size = mod$effect_size(tmp_dat),
-                            p_value = HETER_MODEL$test(tmp_dat)$p_value,
-                            reject = p_value < 0.05,
-                            boot_id = i)
-                   }) %>%
-  reduce(bind_rows) %>%
-  mutate(reject = as.numeric(reject))
-  }
-
-  wrong_BP_conv_result <- wrong_BP_conv_result %>%
-    reduce(bind_rows)
-
-  saveRDS(wrong_BP_conv_result, here::here("data/third_study/wrong_BP_conv_result.rds"))
-} else {
-  wrong_BP_conv_result <- readRDS(here::here("data/third_study/wrong_BP_conv_result.rds"))
-}
-
-```
-
-```{r}
-if (!file.exists(here::here("data/third_study/wrong_SW_conv_result.rds"))) {
-  wrong_SW_conv_result <- list()
-  for (i in 1:100)
-  {
-    wrong_SW_conv_result[[i]] <- map(1:2000,
-                            function(x) {
-                     shape <- sample(1:4, 1)
-                     e_sigma <- sample(c(0.5, 1, 2, 4), 1)
-                     x_dist <- sample(c("uniform", "normal", "lognormal", "even_discrete"), 1)
-                     x <- switch(x_dist,
-                                 uniform = rand_uniform(-1, 1),
-                                 normal = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))},
-                                 lognormal = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))},
-                                 even_discrete = rand_uniform_d(k = 5, even = TRUE))
-                     mod <- poly_model(shape, x = x, sigma = e_sigma)
-                     n <- sample(c(50, 100, 300), 1)
-                     tmp_dat <- mod$gen(n)
-
-                     tibble(shape = shape,
-                            e_sigma = e_sigma,
-                            x_dist = x_dist,
-                            n = n,
-                            effect_size = mod$effect_size(tmp_dat),
-                            p_value = shapiro.test(tmp_dat$.resid)$p.value,
-                            reject = p_value < 0.05,
-                            boot_id = i)
-                   }) %>%
-  reduce(bind_rows) %>%
-  mutate(reject = as.numeric(reject))
-  }
-
-  wrong_SW_conv_result <- wrong_SW_conv_result %>%
-    reduce(bind_rows)
-
-  saveRDS(wrong_SW_conv_result, here::here("data/third_study/wrong_SW_conv_result.rds"))
-} else {
-  wrong_SW_conv_result <- readRDS(here::here("data/third_study/wrong_SW_conv_result.rds"))
-}
-
 ```
@@
 
+<!-- residual plot -> plot -->
+
+<!-- plot a included -->
+
 ```{r}
@@
 In fact, lineup protocol has been integrated into some model diagnostic tools such as @loy2013diagnostic.
+
+<!-- adam work -->
 

### 2022-09-27 paper_comparison/paper_comparison/paper_comparison.Rmd
modify the example table

@@
 Residuals might be plotted, as a histogram or quantile-quantile plot to examine the distribution.
-Using the classical linear regression model as an example, if the distribution is symmetric and unimodal, it is well-behaved.
+Using the classical normal linear regression model as an example, if the distribution is symmetric and unimodal, it is well-behaved.
 But if the distribution is skewed, bimodal, multimodal, or contains outliers, there is cause for concern.
@@
 There are various types of departures from a good residual plot.
-We will only discuss three most commonly checked departures in this paper.
-Namely, non-linearity, heterskedasticity and non-normality.
+Two most commonly checked departures - non-linearity, heterskedasticity and non-normality will be discussed in this paper.
 
@@
 Any non-linear functional form of residuals on fitted values presented in the residual plot could be considered as an indicative of non-linearity.
-At the top-right of Figure [ref here], there is a residual plot giving an example of the visual pattern of non-linearity when a cubic term is not captured by the regression model.
+An example residual plot containing visual pattern of non-linearity is given at the top-right of Figure [ref here].
+One can clearly observe the "S-shape" from the residual plot as the cubic term is not captured by the misspecified model.
 
@@
 
-Non-normality is usually harder to be detected from a residual plot compared to non-linearity and heterskedasticity.
-Besides, a quantile-quantile plot can often do a better job for this task.
-Note that not all regression models assume normality of the error term, but a certain amount of them does.
-If this assumption happens to be false, then it is expected to see data points do not distribute noramlly on the y-axis.
-For example, given a skewed error distribution, one will see fewer data points and more outliers on one side of the zeroline as shown in the bottom-right of Figure \ref{fig:commondepartures}.
+Compared to non-linearity and heterskedasticity, non-normality is usually harder to be detected from a residual plot since scatter plot is not excel in revealing marginal distribution.
+A favourable graphical summarise for this task is the quantile-quantile plot.
+However, for a consistent comparison in the later part of this paper, residual plot will be the focus.
+Besides, it is important to note that not all regression models assume normality for the error term, but a certain amount do including the classical normal linear regression model.
+In the case that the normality assumption doesn't hold, it is expected to observe data points that do not centralize around the horizontal zero line and unevenly distribute at both sides of the zero line.
+For example, given a skewed error distribution, fewer data points and more outliers are on one side of the zero line as shown at the bottom-right of Figure \ref{fig:commondepartures}.
 
@@
 
-data.frame(plot = c("B", "C", "D"), 
-           model = c("Non-linearity", "Heteroskedasticity", "Non-normality"),
-           f = c(POLY_MODEL$test(dat1)$p_value,
+data.frame(plot = c("A", "B", "C", "D"), 
+           model = c("None", "Non-linearity", "Heteroskedasticity", "Non-normality"),
+           f = c(POLY_MODEL$test(dat)$p_value, 
+                 POLY_MODEL$test(dat1)$p_value,
                  POLY_MODEL$test(mutate(dat2, z = poly_model()$gen(300, computed = select(dat2, x))$z))$p_value,
                  POLY_MODEL$test(mutate(dat3, z = poly_model()$gen(300, computed = select(dat3, x))$z))$p_value),
-           b = c(HETER_MODEL$test(dat1)$p_value,
+           b = c(HETER_MODEL$test(dat)$p_value, 
+                 HETER_MODEL$test(dat1)$p_value,
                  HETER_MODEL$test(dat2)$p_value,
                  HETER_MODEL$test(dat3)$p_value),
-           s = c(shapiro.test(dat1$.resid)$p.value,
+           s = c(shapiro.test(dat$.resid)$p.value, 
+                 shapiro.test(dat1$.resid)$p.value,
                  shapiro.test(dat2$.resid)$p.value,
@@
   table_dat %>%
-  knitr::kable(col.names = c("Residual plots", "Departures", "Model structural", "Breusch-Pagan", "Shapiro–Wilk"),
+  knitr::kable(col.names = c("Plot", "Departures", "Model structural", "Breusch-Pagan", "Shapiro–Wilk"),
                align = "llrrr",
@@
                booktabs = TRUE, 
-               caption  = 'Example residual plots examined by the corresponding conventional tests. P-values smaller than 0.05 are showed in bold.', 
+               caption  = 'Results of example residual plots given in Figure \\ref{fig:commondepartures} examined by the Model structural test (F-test), the Breusch-Pagan test and the Shapiro–Wilk test.
+P-values smaller than 0.05 are showed in bold.',
                label = "example-residual-plot-table" )
@@
 This is encouraging as it promotes the use of visual inference in border field of data science where there are no existing statistical testing procedures.
-In fact, lineup protocol has been integrated into some model diagnostic tools such as @loy2013diagnostic.
+In fact, lineup protocol has been integrated successfully into diagnostic tools of hierarchical linear models [@loy2013diagnostic].
 
@@
 ```{r residual-plot-cubic-heter, fig.cap='Residuals vs.
-fitted values plot for a classical linear regression model.
+fitted values plot for a classical normal linear regression model.
 The residuals are produced by fitting a two-predictor multiple linear regression model with data generated from a cubic linear model.
@@
 ```{r example-lineup, fig.cap="Lineup 24 in experiment I. Can you spot the most different plot? \\label{fig:example-lineup}"}
-VI_MODEL$plot_lineup(polynomials_lineup[[24]]$data, 
+VI_MODEL$plot_lineup(vi_lineup[["poly_24"]]$data, 
                      remove_grid_line = TRUE, 

### 2022-09-27 paper_comparison/paper_comparison/paper_comparison.tex
modify the example table

@@
 Residuals might be plotted, as a histogram or quantile-quantile plot to
-examine the distribution. Using the classical linear regression model as
-an example, if the distribution is symmetric and unimodal, it is
-well-behaved. But if the distribution is skewed, bimodal, multimodal, or
-contains outliers, there is cause for concern. The distribution could
+examine the distribution. Using the classical normal linear regression
+model as an example, if the distribution is symmetric and unimodal, it
+is well-behaved. But if the distribution is skewed, bimodal, multimodal,
+or contains outliers, there is cause for concern. The distribution could
 also be inspected by conducting a goodness of fit test, such as the
@@
 line showing no noticeable patterns. There are various types of
-departures from a good residual plot. We will only discuss three most
-commonly checked departures in this paper. Namely, non-linearity,
-heterskedasticity and non-normality.
+departures from a good residual plot. Two most commonly checked
+departures - non-linearity, heterskedasticity and non-normality will be
+discussed in this paper.
 
@@
 in the residual plot could be considered as an indicative of
-non-linearity. At the top-right of Figure {[}ref here{]}, there is a
-residual plot giving an example of the visual pattern of non-linearity
-when a cubic term is not captured by the regression model.
+non-linearity. An example residual plot containing visual pattern of
+non-linearity is given at the top-right of Figure {[}ref here{]}. One
+can clearly observe the ``S-shape'' from the residual plot as the cubic
+term is not captured by the misspecified model.
 
@@
 
-Non-normality is usually harder to be detected from a residual plot
-compared to non-linearity and heterskedasticity. Besides, a
-quantile-quantile plot can often do a better job for this task. Note
-that not all regression models assume normality of the error term, but a
-certain amount of them does. If this assumption happens to be false,
-then it is expected to see data points do not distribute noramlly on the
-y-axis. For example, given a skewed error distribution, one will see
-fewer data points and more outliers on one side of the zeroline as shown
-in the bottom-right of Figure \ref{fig:commondepartures}.
+Compared to non-linearity and heterskedasticity, non-normality is
+usually harder to be detected from a residual plot since scatter plot is
+not excel in revealing marginal distribution. A favourable graphical
+summarise for this task is the quantile-quantile plot. However, for a
+consistent comparison in the later part of this paper, residual plot
+will be the focus. Besides, it is important to note that not all
+regression models assume normality for the error term, but a certain
+amount do including the classical normal linear regression model. In the
+case that the normality assumption doesn't hold, it is expected to
+observe data points that do not centralize around the horizontal zero
+line and unevenly distribute at both sides of the zero line. For
+example, given a skewed error distribution, fewer data points and more
+outliers are on one side of the zero line as shown at the bottom-right
+of Figure \ref{fig:commondepartures}.
 
@@
 
-\caption{\label{tab:example-residual-plot-table}Example residual plots examined by the corresponding conventional tests. P-values smaller than 0.05 are showed in bold.}
+\caption{\label{tab:example-residual-plot-table}Results of example residual plots given in Figure \ref{fig:commondepartures} examined by the Model structural test (F-test), the Breusch-Pagan test and the Shapiro–Wilk test.
+P-values smaller than 0.05 are showed in bold.}
 \centering
@@
 \toprule
-Residual plots & Departures & Model structural & Breusch-Pagan & Shapiro–Wilk\\
+Plot & Departures & Model structural & Breusch-Pagan & Shapiro–Wilk\\
 \midrule
+A & None & 0.434 & 0.133 & 0.728\\
 B & Non-linearity & \textbf{0.000} & \textbf{0.000} & \textbf{0.039}\\
@@
 field of data science where there are no existing statistical testing
-procedures. In fact, lineup protocol has been integrated into some model
-diagnostic tools such as \citet{loy2013diagnostic}.
+procedures. In fact, lineup protocol has been integrated successfully
+into diagnostic tools of hierarchical linear models
+\citep{loy2013diagnostic}.
 

### 2022-09-27 paper_comparison/paper_comparison/paper_comparison.Rmd
redo the analysis

@@
 Besides, it is important to note that not all regression models assume normality for the error term, but a certain amount do including the classical normal linear regression model.
-In the case that the normality assumption doesn't hold, it is expected to observe data points that do not centralize around the horizontal zero line and unevenly distribute at both sides of the zero line.
+In the case that the normality assumption is violated, it is expected to observe data points that do not centralize around the horizontal zero line and unevenly distribute at both sides of the zero line.
 For example, given a skewed error distribution, fewer data points and more outliers are on one side of the zero line as shown at the bottom-right of Figure \ref{fig:commondepartures}.
@@
 \begin{equation} \label{eq:pvalue-beta-binomial}
-P(C \geq c_i) = \sum_{x=c_i}^{K}{{K}\choose{x}}\frac{1}{B(\alpha, (m-1)\alpha)}B(x + \alpha, K - x + (m - 1)\alpha),
+P(C \geq c_i) = \int_{x=c_i}^{K}\frac{\Gamma(K + 1)}{\Gamma(x + 1)\Gamma(K - x + 1)}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)}dx,
 \end{equation}
@@
 B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0. 
+\end{equation}
+```
+
+and $\Gamma(.)$ is the gamma function defined as
+
+```{=tex}
+\begin{equation} \label{eq:gammafunction}
+\Gamma(z) = \int_{0}^{\infty}t^{z - 1}e^{-t}dt,\quad \text{where}\quad z>0. 
 \end{equation}
@@
 
+## Power comparison
+
+```{r}
+vi_survey_processed <- readRDS(here::here("data/vi_survey_processed.rds")) 
+```
+
+```{r}
+boot_y <- function(dat, y, times = 100) {
+  map_df(1:times, function(i) slice_sample(dat, n = nrow(dat), replace = TRUE) %>% mutate(boot_id = i))
+}
+```
+
+```{r}
+vi_survey_processed %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  rename(p_value_dirichlet = p_value_select_all_plots) %>%
+  ggplot() +
+  geom_point(aes(p_value_uniform, p_value_dirichlet)) +
+  geom_abline() +
+  scale_x_log10() +
+  scale_y_log10()
+```
+
+```{r}
+vi_survey_processed %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(type == "polynomial") %>%
+  filter(e_sigma != 0.125) %>%
+  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  ggplot() +
+  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  stat_smooth(geom = "line", 
+              data = ~boot_y(.x, reject_dirichlet, times = 500), 
+              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, 
+              alpha = 0.03) +
+  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
+                pivot_longer(F_p_value:SW_p_value), 
+              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
+              method = "glm", 
+              method.args = list(family = binomial),
+              se = FALSE) +
+  facet_wrap(~x_dist) +
+  ggtitle("poly")
+```
+
+```{r}
+vi_survey_processed %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(type == "polynomial") %>%
+  filter(e_sigma != 0.125) %>%
+  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
+  mutate(reject_uniform = as.numeric(p_value_uniform <= 0.05)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  ggplot() +
+  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value_dirichlet"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  stat_smooth(geom = "line", 
+              data = ~boot_y(.x, reject_dirichlet, times = 500), 
+              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value_dirichlet"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, 
+              alpha = 0.03) +
+  geom_smooth(aes(log_effect_size, reject_uniform, col = "VI_p_value_uniform"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  stat_smooth(geom = "line", 
+              data = ~boot_y(.x, reject_uniform, times = 500), 
+              aes(log_effect_size, reject_uniform, group = boot_id, col = "VI_p_value_uniform"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, 
+              alpha = 0.03) +
+  facet_wrap(~x_dist) +
+  ggtitle("poly")
+```
+
+```{r}
+vi_survey_processed %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(type == "heteroskedasticity") %>%
+  filter(b != 0) %>%
+  filter(b < 128) %>%
+  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  ggplot() +
+  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  stat_smooth(geom = "line", 
+              data = ~boot_y(.x, reject_dirichlet, times = 500), 
+              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, 
+              alpha = 0.03) +
+  geom_smooth(data = mutate(heter_conv_sim, log_effect_size = log(effect_size)) %>%
+                pivot_longer(F_p_value:SW_p_value), 
+              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
+              method = "glm", 
+              method.args = list(family = binomial),
+              se = FALSE) +
+  facet_wrap(~x_dist) +
+  ggtitle("heter")
+```
+
+```{r}
+vi_survey_processed %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(type == "heteroskedasticity") %>%
+  filter(e_sigma != 0.125) %>%
+  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
+  mutate(reject_uniform = as.numeric(p_value_uniform <= 0.05)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  ggplot() +
+  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value_dirichlet"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  stat_smooth(geom = "line", 
+              data = ~boot_y(.x, reject_dirichlet, times = 500), 
+              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value_dirichlet"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, 
+              alpha = 0.03) +
+  geom_smooth(aes(log_effect_size, reject_uniform, col = "VI_p_value_uniform"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  stat_smooth(geom = "line", 
+              data = ~boot_y(.x, reject_uniform, times = 500), 
+              aes(log_effect_size, reject_uniform, group = boot_id, col = "VI_p_value_uniform"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, 
+              alpha = 0.03) +
+  facet_wrap(~x_dist) +
+  ggtitle("heter")
+```
+
+```{r}
+p_value_cache_env <- new.env()
+
+if (file.exists(here::here("data/p_value_only_5_eval.rds"))) {
+  p_value_only_5_eval <- readRDS(here::here("data/p_value_only_5_eval.rds"))
+} else {
+  p_value_only_5_eval <- vi_survey_processed %>%
+  filter(x_dist == "uniform") %>%
+  filter((is.na(b) & e_sigma != 0.125 ) | (b != 0 & b < 128)) %>%
+  filter(exp != 5) %>%
+  group_by(unique_lineup_id) %>%
+  summarise(p_value_only_5_eval = calc_p_value(n_detect = sum(detect), 
+                                               n_eval = n(), 
+                                               n_sel = num_selection, 
+                                               dist = "dirichlet",
+                                               alpha = numeric_alpha[1],
+                                               cache_env = p_value_cache_env))
+
+  saveRDS(p_value_only_5_eval, here::here("data/p_value_only_5_eval.rds"))
+}
+
+vi_survey_processed %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(type == "heteroskedasticity") %>%
+  filter(x_dist == "uniform") %>%
+  filter(b != 0) %>%
+  filter(b < 128) %>%
+  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  left_join(p_value_only_5_eval) %>%
+  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
+  ggplot() +
+  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  stat_smooth(geom = "line", 
+              data = ~boot_y(.x, reject_dirichlet, times = 500), 
+              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, 
+              alpha = 0.03) +
+  geom_smooth(aes(log_effect_size, reject_only_5_eval, col = "VI_p_value_only_5_eval"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  stat_smooth(geom = "line", 
+              data = ~boot_y(.x, reject_only_5_eval, times = 500), 
+              aes(log_effect_size, reject_only_5_eval, group = boot_id, col = "VI_p_value_only_5_eval"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, 
+              alpha = 0.03) +
+  ggtitle("heter")
+
+p1 <- vi_survey_processed %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(type == "heteroskedasticity") %>%
+  filter(x_dist == "uniform") %>%
+  filter(b != 0) %>%
+  filter(b < 128) %>%
+  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  left_join(p_value_only_5_eval) %>%
+  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
+  ggplot() +
+  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  stat_smooth(geom = "line", 
+              data = ~boot_y(.x, reject_dirichlet, times = 500), 
+              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, 
+              alpha = 0.03)
+
+p2 <- vi_survey_processed %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(type == "heteroskedasticity") %>%
+  filter(x_dist == "uniform") %>%
+  filter(b != 0) %>%
+  filter(b < 128) %>%
+  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  left_join(p_value_only_5_eval) %>%
+  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
+  ggplot() +
+  geom_smooth(aes(log_effect_size, reject_only_5_eval, col = "VI_p_value_only_5_eval"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  stat_smooth(geom = "line", 
+              data = ~boot_y(.x, reject_only_5_eval, times = 500), 
+              aes(log_effect_size, reject_only_5_eval, group = boot_id, col = "VI_p_value_only_5_eval"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, 
+              alpha = 0.03)
+
+patchwork::wrap_plots(p1, p2)
+```
+
+```{r}
+
+vi_survey_processed %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(type == "polynomial") %>%
+  filter(x_dist == "uniform") %>%
+  filter(e_sigma != 0.125) %>%
+  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  left_join(p_value_only_5_eval) %>%
+  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
+  ggplot() +
+  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  stat_smooth(geom = "line", 
+              data = ~boot_y(.x, reject_dirichlet, times = 500), 
+              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, 
+              alpha = 0.03) +
+  geom_smooth(aes(log_effect_size, reject_only_5_eval, col = "VI_p_value_only_5_eval"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  stat_smooth(geom = "line", 
+              data = ~boot_y(.x, reject_only_5_eval, times = 500), 
+              aes(log_effect_size, reject_only_5_eval, group = boot_id, col = "VI_p_value_only_5_eval"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, 
+              alpha = 0.03) +
+  ggtitle("poly")
+
+p1 <- vi_survey_processed %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(type == "polynomial") %>%
+  filter(x_dist == "uniform") %>%
+  filter(e_sigma != 0.125) %>%
+  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  left_join(p_value_only_5_eval) %>%
+  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
+  ggplot() +
+  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  stat_smooth(geom = "line", 
+              data = ~boot_y(.x, reject_dirichlet, times = 500), 
+              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, 
+              alpha = 0.03)
+
+p2 <- vi_survey_processed %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(type == "polynomial") %>%
+  filter(x_dist == "uniform") %>%
+  filter(e_sigma != 0.125) %>%
+  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  left_join(p_value_only_5_eval) %>%
+  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
+  ggplot() +
+  geom_smooth(aes(log_effect_size, reject_only_5_eval, col = "VI_p_value_only_5_eval"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  stat_smooth(geom = "line", 
+              data = ~boot_y(.x, reject_only_5_eval, times = 500), 
+              aes(log_effect_size, reject_only_5_eval, group = boot_id, col = "VI_p_value_only_5_eval"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, 
+              alpha = 0.03)
+
+patchwork::wrap_plots(p1, p2)
+
+```
+
 <!-- # Old stuff -->

### 2022-09-27 paper_comparison/paper_comparison/paper_comparison.tex
redo the analysis

@@
 amount do including the classical normal linear regression model. In the
-case that the normality assumption doesn't hold, it is expected to
+case that the normality assumption is violated, it is expected to
 observe data points that do not centralize around the horizontal zero
@@
 B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0. 
+\end{equation}
+
+and \(\Gamma(.)\) is the gamma function defined as
+
+\begin{equation} \label{eq:gammafunction}
+\Gamma(z) = \int_{0}^{\infty}t^{z - 1}e^{-t}dt,\quad \text{where}\quad z>0. 
 \end{equation}
@@
 
+\hypertarget{power-comparison}{%
+\subsection{Power comparison}\label{power-comparison}}
+
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-8-1}
+
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-9-1}
+
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-10-1}
+
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-11-1}
+
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-12-1}
+
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-13-1}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-13-2}
+
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-14-1}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-14-2}
+
 --\textgreater{}

### 2022-09-27 paper_comparison/scripts/estimate_alpha.R
estimate alpha

@@ -1,0 +1,87 @@
+library(visage)
+library(tidyverse)
+
+
+calc_alpha <- function(zero_sel = 0.05) {
+  vi_survey %>%
+    filter(b == 0) %>%
+    select(unique_lineup_id, type:n, -shape, -include_z, -a, -b, -e_dist, -e_sigma, -formula, -k, -type, -name) %>%
+    group_by(unique_lineup_id) %>%
+    summarise(across(x_dist:n, first)) %>%
+    arrange(across(c(x_dist:n))) %>%
+    mutate(setting_id = rep(1:12, each = 3)) -> final_vi_survey
+
+  final_vi_survey %>%
+    select(unique_lineup_id, setting_id) %>%
+    right_join(filter(vi_survey, b == 0)) %>%
+    select(unique_lineup_id, setting_id, selection, num_selection) %>%
+    (function(x) {
+      for (i in 1:20)
+        x[[paste0("plot_", i)]] <-
+          grepl(paste0("_", i, "_"), x$selection) |
+          grepl(paste0("^", i, "_"), x$selection) |
+          grepl(paste0("_", i, "$"), x$selection) |
+          grepl(paste0("^", i, "$"), x$selection)
+      x}
+    ) %>%
+    mutate(across(plot_1:plot_20, function(x) ifelse(num_selection, x/num_selection, zero_sel))) %>%
+    group_by(setting_id, unique_lineup_id) %>%
+    summarise(across(plot_1:plot_20, sum)) %>%
+    ungroup() %>%
+    pivot_longer(plot_1:plot_20) %>%
+    group_by(setting_id, unique_lineup_id) %>%
+    summarise(c_interesting = sum(value > 1)) %>%
+    group_by(setting_id) %>%
+    summarise(c_interesting = mean(c_interesting)) %>%
+    ungroup() %>%
+    mutate(numeric_alpha = vinference::estimate_alpha_numeric(Zc = c_interesting, m0 = 20, K = 20)) %>%
+    mutate(numeric_alpha_sum_sq_error = numeric_alpha$sum_sq_error, numeric_alpha = numeric_alpha$alpha) %>%
+    right_join(final_vi_survey)
+}
+
+
+calc_alpha(0.05) -> select_all_plots
+calc_alpha(0) -> select_zero_plots
+
+vi_survey <- vi_survey %>%
+  rename(p_value_uniform = p_value) %>%
+  left_join(select_all_plots %>%
+              group_by(setting_id) %>%
+              summarise(across(everything(), first)) %>%
[…]

### 2022-09-27 paper_comparison/paper_comparison/paper_comparison.Rmd
try lme4

@@
 
+As discussed in @majumder_validation_2013, individual's skill will affect the number of observers who identify the actual data plot from the lineup.
+Thus, the power of a visual test depends on the subject-specific abilities.
+Previously, it was addressed by modelling the probability of a subject $i$ correctly picking the actual data plot from a lineup $l$ using a mixed-effect logistic regression with the subject being treated as a random effect [@majumder_validation_2013].
+However, knowing this probability is insufficient to determine the power of a visual test allowed for multiple selections as it doesn't provide information about the number of selections made by the subject, and hence, the p-value can't be computed.
+
+Instead, we modelled the probability of a lineup being rejected directly using a logistic regression with the natural logarithm of the effect size as the only regressor.
+The same model was applied to different subsets of the survey data to make comparison of the power.
+
 <!-- -   Very briefly explain what a lineup test is -->
@@
 
+```{r eval = FALSE}
+vi_survey_processed %>%
+  filter(exp == 3) %>%
+  filter(e_sigma != 0.125) %>%
+  mutate()
+
+vi_survey_processed %>%
+  filter(exp == 3) %>%
+  filter(e_sigma != 0.125) %>%
+  mutate(set = factor(set), lineup_id = factor(lineup_id)) %>%
+  {lme4::glmer(weighted_detect ~ effect_size + ( 0 + effect_size | set) + ( 0 + effect_size | lineup_id), family = binomial ,data = .)} %>%
+  predict(type = "response")
+
+  ggplot() +
+  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE)
+
++
+  stat_smooth(geom = "line", 
+              data = ~boot_y(.x, reject_dirichlet, times = 500), 
+              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, 
+              alpha = 0.03) +
+  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
+                pivot_longer(F_p_value:SW_p_value), 
+              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
+              method = "glm", 
+              method.args = list(family = binomial),
+              se = FALSE) +
+  facet_wrap(~x_dist) +
+  ggtitle("poly")
+
+  ?glm
+```
+
 <!-- # Old stuff -->

### 2022-09-27 paper_comparison/paper_comparison/paper_comparison.tex
try lme4

@@
 
+As discussed in \citet{majumder_validation_2013}, individual's skill
+will affect the number of observers who identify the actual data plot
+from the lineup. Thus, the power of a visual test depends on the
+subject-specific abilities. Previously, it was addressed by modelling
+the probability of a subject \(i\) correctly picking the actual data
+plot from a lineup \(l\) using a mixed-effect logistic regression with
+the subject being treated as a random effect
+\citep{majumder_validation_2013}. However, knowing this probability is
+insufficient to determine the power of a visual test allowed for
+multiple selections as it doesn't provide information about the number
+of selections made by the subject, and hence, the p-value can't be
+computed.
+
+Instead, we modelled the probability of a lineup being rejected directly
+using a logistic regression with the natural logarithm of the effect
+size as the only regressor. The same model was applied to different
+subsets of the survey data to make comparison of the power.
+
 \hypertarget{experimental-design}{%

### 2022-09-28 paper_comparison/paper_comparison/paper_comparison.Rmd
modify exp design

@@
 And defined $c_i$ to be the number of times plot $i$ being selected in $K$ evaluations.
-In case subject $j$ makes multiple selections, they decided to add $1/s_j$ to $c_i$ instead of one, where $s_j$ is the number of plots subject $j$ selected for $j=1,...K$.
+In case subject $j$ makes multiple selections, $1/s_j$ will be added to $c_i$ instead of one, where $s_j$ is the number of plots subject $j$ selected for $j=1,...K$.
 This ensured $\sum_{i}c_i=K$.
-Thus, the visual p-value followed by the above model is given as
+Since we are only interested in the selections of the actual data plot $i$, the marginal model can be simplified to a beta-binomial model and thus the visual p-value is given as
 
@@
 
-Instead, we modelled the probability of a lineup being rejected directly using a logistic regression with the natural logarithm of the effect size as the only regressor.
-The same model was applied to different subsets of the survey data to make comparison of the power.
+(1) Instead, we modelled the probability of a lineup being rejected directly using a logistic regression with the natural logarithm of the effect size as the only regressor.
+The same model was applied to different subsets of the survey data for making comparison of the power.
+
+(2) Instead, we modelled the plot selection $c_i$ for the actual data plot $i$.
 
@@
 
-## Motivation and overview
-
-With our knowledge, the effectiveness of visual test relative to the equivalent conventional test in regression diagnostics has not been validated.
+## Overview
+
+<!-- With our knowledge, the effectiveness of visual test relative to the equivalent conventional test in regression diagnostics has not been validated.
 Particularly, its ability to detect non-linearity and heteroskedasticity compared to model structural test and Breusch–Pagan test.
@@
 Attempting to find the answer to these questions, three experiments were conducted.
-
+-->
+
+Three experiments were conducted to investigate the difference between conventional hypothesis testing and graphical diagnostics in the application of linear regression diagnostics.
 The experiment I has ideal scenario for conventional testing, where the visual test is not expected to outperform the conventional test.
@@
 The experiment III is designed for collecting human responses to lineup with only good residual plots such that the parameter $\alpha$ in Equation \ref{eq:pvalue-beta-binomial} can be estimated.
+Overall, we planned to collect 7974 evaluations on 1152 uniqued lineups performed by 443 subjects throughout three experiment.
+
+## Simulating departures
+
+Two types of departures were considered with the corresponding data generating process being designed for experiment I and II.
+
+### Non-linearity
+
+Experiment I is designed to study the ability of human subjects to detect the effect of a non-linear term $\boldsymbol{z}$ which is a probabilist's Hermite polynomial [Herimite ref here] of another random vector $\boldsymbol{x}$ in a two variable statistical model formulated as:
+
+\begin{align} \label{eq:nonlinearity-model}
+\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
+\boldsymbol{x} = g(\boldsymbol{x}_{raw}, 1), \\
+\boldsymbol{z} = g(\boldsymbol{z}_{raw}, 1), \\
+\boldsymbol{z}_{raw} = He_j(g(\boldsymbol{z}, 2)),
+\end{align}
+
+where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vectors of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $\{-k, k\}$ defined as
+
+```{=tex}
+\begin{equation} \label{eq:scaling-function}
+g(\boldsymbol{x}, k) = (\boldsymbol{x} - min(\boldsymbol{x}))/max(\boldsymbol{x} - min(\boldsymbol{x}))2k - k, \quad \text{for} \quad k > 0. 
+\end{equation}
+```
+
+The null regression model used to fit the realizations generated by the above model is formulated as:
+
+```{=tex}
+\begin{equation} \label{eq:null-model}
+\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
+\end{equation}
+```
+where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.
+
+Since $z = O(x^j)$, for $j > 1$, $z$ is a higher order term leaves out by the null regression, which will result in model misspecification.
+Visual patterns of non-linearity were simulated using four different order of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
+
+<!-- A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}. -->
+
+```{r results='asis', eval = FALSE}
+data.frame(j = c(2, 3, 6, 18),
+           x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$")) %>%
+  knitr::kable(col.names = NULL,
+               align = "rl",
+               escape = FALSE,
+               format = "latex",
+               booktabs = TRUE, 
+               caption  = 'Parameter values for $j$ and $X_{raw}$', 
+               label = "parameter-table" ) %>%
+  kableExtra::add_header_above(c("Order of Hermite polynomial\n(sym1)", "Distribution of sym4\n")) %>%
+  as.character() %>%
+  gsub("sym1", "$j$", .) %>%
+  gsub("sym4", "$X_{raw}$", .) %>%
+  cat()
+  # gsub("\\\\\\$", "$", .) %>%
+  # cat()
+```
+
+The values of $j$ was chosen so that different shapes of non-linearity were included in the residual plot.
+As shown in Figure \ref{fig:different-shape-of-herimite}, these include "U" shape, "S" shape, "M" shape and "Triple-U" shape.
+
+```{r different-shape-of-herimite, fig.cap="Example residual plots of the linear regression model used in experiment I.
+Four different values of $j$ are included in the experiment results in four different shapes of residuals."}
+stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
+
+ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300)
+
+VI_MODEL$plot(ori_dat) + 
+  ggtitle(quote(He[2]:"U shape")) + 
+  theme_light() -> p1
+
+VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote(He[3]:"S shape")) + 
+  theme_light() -> p2
+
+VI_MODEL$plot(poly_model(shape = 3, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote(He[6]:"M shape")) + 
+  theme_light() -> p3
+
+VI_MODEL$plot(poly_model(shape = 4, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote(He[18]:"Triple-U shape")) + 
+  theme_light() -> p4
+
+patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)
+```
+
+<!-- The range of $\sigma$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
+The example residual plots are given in Figure \ref{fig:different-sigma}.
+-->
+
+```{r different-sigma, eval = FALSE, fig.cap="Example residual plots of the linear regression model used in experiment I.
+Four different values of $\\sigma$ are included in the experiment to control the strength of the signal."}
+ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)
+
+VI_MODEL$plot(ori_dat) + 
+  ggtitle(quote(sigma~"="~0.5)) + 
+  theme_light() -> p1
+
+VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1)$gen(300, computed = select(ori_dat, x))) +
+  ggtitle(quote(sigma~"="~1)) + 
+  theme_light() -> p2
+
+VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 2)$gen(300, computed = select(ori_dat, x))) +
+  ggtitle(quote(sigma~"="~2)) + 
+  theme_light() -> p3
+
+VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 4)$gen(300, computed = select(ori_dat, x))) +
+  ggtitle(quote(sigma~"="~4)) + 
+  theme_light() -> p4
+
+patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)
+```
+
+Four different distribution were used to generate $X_{raw}$ as shown in Figure \ref{fig:different-dist}.
+The uniform and the normal distribution are symmetric and commonly assumed in statistical models.
+The adjusted log-normal distribution provides skewed density.
+And the discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.
+
+```{r different-dist, fig.cap="Example residual plots of the linear regression model used in experiment I.
+Four different distribution of $x_{raw}$ are used in the experiment to provide various visual patterns."}
+ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)
+
+VI_MODEL$plot(ori_dat) + 
+  ggtitle(quote(U(-1,1))) + 
+  theme_light() -> p1
+
+VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
+  ggtitle(quote(N(0,0.3^2))) + 
+  theme_light() -> p2
+
+VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
+  ggtitle(quote(lognormal(0,0.6^2)/3)) + 
+  theme_light() -> p3
+
+VI_MODEL$plot(poly_model(shape = 1, x = rand_uniform_d(k = 5, even = TRUE), sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
+  ggtitle(quote(U~"{1,5}")) + 
+  theme_light() -> p4
+
+patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)
+```
+
+Figure \ref{fig:example-lineup} shows one of the lineups used in experiment I.
+This lineup was produced under $j = 6$ and $X_{raw} \sim N(0.0.3^2)$.
+The actual data plot location was four.
+All five subjects correctly identified the actual data plot for this lineup.
+
+```{r example-lineup, fig.cap="Lineup 24 in experiment I. Can you spot the most different plot? \\label{fig:example-lineup}"}
+VI_MODEL$plot_lineup(vi_lineup[["poly_24"]]$data, 
+                     remove_grid_line = TRUE, 
+                     theme = theme_light(),
+                     remove_axis = TRUE)
+```
+
+### Heteroskedasticity
+
+Experiment II is designed to study the ability of human subjects to detect the appearance of a heteroskedasticity pattern under a simple linear regression model setting:
+
+```{=tex}
+\begin{align} \label{eq:heter-model}
+\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
+\boldsymbol{x} = g(\boldsymbol{x}_{raw}, 1)\\
+\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + 2 - |a| b (\boldsymbol{x} - a)^2 \boldsymbol{I}), \\
+\end{align}
+```
+
+where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vectors of size $n$ and $g(.)$ is the scaling function defined in \ref{eq:scaling-function}.
+
+The null regression model used to fit the realizations generated by the above model is formulated exactly the same as Equation \ref{eq:null-model}.
+
+For $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the regressor $x$, which will lead to the presence of heteroskedasticity.
+Visual patterns were simulated using three different shapes ($a$ = -1, 0, 1) and the same four different distribution of $X_{raw}$ used in experiment I.
+
+The values of $a$ was chosen so that different shapes of heteroskedasticity were included in the residual plot.
+These include left-triangle shape, butterfly shape and right-triangle shape as displayed in Figure \ref{fig:different-shape-of-heter}.
+
+```{r different-shape-of-heter, fig.cap="Example residual plots of the linear regression model used in experiment II.
+Three different shapes (a = -1, 0, 1) are used in the experiment to create left-triangle shape, butterfly shape and right-triangle shape."}
+stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
+
+ori_dat <- heter_model(a = -1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300)
+
+VI_MODEL$plot(ori_dat) +
+  ggtitle(quote("Left-triangle shape")) +
+  theme_light() -> p1
+
+VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote("Butterfly shape")) +
+  theme_light() -> p2
+
+VI_MODEL$plot(heter_model(a = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote("Right-triangle shape")) +
+  theme_light() -> p3
+
+patchwork::wrap_plots(p1, p2, p3, ncol = 3)
+```
+
+## Experimental setup
+
+### Treatments
+
+To control the strength of the signal, two parameters were used including $\sigma$ and $b$ for controlling the standard deviation of the error term in experiment I and II respectively.
+
+<!-- > Experiment I -->
+
+<!-- -   Treatments for nonlinearity experiment (plots summarising levels) -->
+
+<!-- It was generated for $n = 1$, $\sigma = 1$, $j = 1$, $X_{raw} \sim N$.
+For this lineup, x out of k observers picked the actual data plot.
+Each lineup is evaluated by 5 different subjects to provide reasonable estimates of the p-value.
+-->
+
+<!-- > Experiment II -->
+
+<!-- -   Treatments for heteroskedasticity experiment (plots summarising levels) -->
+
+<!-- Experiment data were simulated using three different shapes ($a$ = -1, 0, 1), five different values of $b$ (b = 0.25, 1, 4, 16, 64), three different sample sizes (n = 50, 100, 300) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
+A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}.
+-->
+
+<!-- The range of $b$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
+-->
+
+### Assigning subjects
+
+<!-- Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in 192 different lineups.
+For each lineup, the actual data plot was drawn as a standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis.
+The 19 null datasets were generated by the residual rotation technique, and plotted in the same way.
+The lineup consisted of 20 residual plots with one randomly placed actual data plot.
+Figure \ref{fig:example-lineup} is an example of one of these lineups.
+It was produced by using $n = 300$, $j = 6$, $\sigma = 0.5$ and $X_{raw} \sim N(0.0.3^2)$.
+The actual data plot location was four.
+All five subjects correctly identified the actual data plot for this lineup.
+-->
+
+<!-- In addition, each lineup is designed to be evaluated by five different subjects to provide reasonable estimates of the visual p-value.
+Thus, $192 \times 3 \times 5 / (20 - 2) = 160$ subjects were recruited to satisfy the design of the experiment I.
+-->
+
+### Collecting results
 
@@
 Prescreening procedure was applied during the recruitment, subjects were required to be fluent in English, with $98\%$ minimum approval rate in other studies and 10 minimum submissions.
-
 During the experiment, every subject was presented with a block of 20 lineups.
@@
 Information about preferred pronoun, age group, education, and previous experience in visual experiment were also collected.
-
 In every block of 20 lineups that presented to a subject, two lineups with obvious visual patterns were included as attention checks.
@@
 Data of rejected submissions were discarded automatically to maintain the overall data quality.
-
-Overall, we planned to collect 7974 evaluations on 1152 uniqued lineups performed by 443 subjects.
-
-## Simulating departures
-
-Two types of departures were considered with the corresponding data generating process being designed for experiment I and II.
-
-### Non-linearity
-
-Experiment I is designed to study the ability of human subjects to detect the effect of a non-linear term $\boldsymbol{z}$ which is a probabilist's Hermite polynomial [Herimite ref here] of another random vector $\boldsymbol{x}$ in a two variable statistical model formulated as:
-
-\begin{align} \label{eq:nonlinearity-model}
-\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
-\boldsymbol{x} = g(\boldsymbol{x}_{raw}, 1), \\
-\boldsymbol{z} = g(\boldsymbol{z}_{raw}, 1), \\
-\boldsymbol{z}_{raw} = He_j(g(\boldsymbol{z}, 2)),
-\end{align}
-
-where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vector of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $\{-k, k\}$ defined as
-
-```{=tex}
-\begin{equation} \label{eq:scaling-function}
-g(\boldsymbol{x}, k) = (\boldsymbol{x} - min(\boldsymbol{x}))/max(\boldsymbol{x} - min(\boldsymbol{x}))2k - k, \quad \text{for} \quad k > 0. 
-\end{equation}
-```
-
-The null regression model used to fit the realizations generated by the above model is formulated as:
-
-```{=tex}
-\begin{equation} \label{eq:null-model}
-\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
-\end{equation}
-```
-where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.
-
-Since $z = O(x^j)$, for $j > 1$, $z$ is a higher order term leaves out by the null regression, which will result in model misspecification.
-
-### Heteroskedasticity
-
-Experiment II is designed to study the ability of human subjects to detect the appearance of a heteroskedasticity pattern under a simple linear regression model setting:
-
-```{=tex}
-\begin{align} \label{eq:heter-model}
-\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
-\boldsymbol{x} = g(\boldsymbol{x}_{raw}, 1)\\
-\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + 2 - |a| b (\boldsymbol{x} - a)^2 \boldsymbol{I}), \\
-\end{align}
-```
-where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vector of size $n$ and $g(.)$ is the scaling function defined in \ref{eq:scaling-function}.
-
-The null regression model used to fit the realizations generated by the above model is formulated exactly the same as Equation \ref{eq:null-model}.
-
-For $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the regressor $x$, which will lead to the presence of heteroskedasticity.
-
-## Experimental setup
-
-### Treatments
-
-> Experiment I
-
--   Treatments for nonlinearity experiment (plots summarising levels)
-
-Data of experiment I were simulated using four different order of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$), three different sample sizes (n = 50, 100, 300), four different standard deviations of the error ($\sigma$ = 0.5, 1, 2, 4) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
-A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}.
-
-```{r results='asis'}
-data.frame(n = c(50, 100, 300, ""),
-           j = c(2, 3, 6, 18),
-           sigma = c(0.5, 1, 2, 4),
-           x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$")) %>%
-  knitr::kable(col.names = NULL,
-               align = "rrrl",
-               escape = FALSE,
-               format = "latex",
-               booktabs = TRUE, 
-               caption  = 'Parameter values for $n$, $j$ $\\sigma$, $X_{raw}$', 
-               label = "parameter-table" ) %>%
-  kableExtra::add_header_above(c("Sample size\n(sym3)", "Order of Hermite polynomial\n(sym1)", "Error SD\n(sym2)", "Distribution of sym4\n")) %>%
-  as.character() %>%
-  gsub("sym1", "$j$", .) %>%
-  gsub("sym2", "$\\\\sigma$", .) %>%
-  gsub("sym3", "$n$", .) %>%
-  gsub("sym4", "$X_{raw}$", .) %>%
-  cat()
-  # gsub("\\\\\\$", "$", .) %>%
-  # cat()
-```
-
-The values of $j$ was chosen so that different shapes of non-linearity were included in the residual plot.
-As shown in Figure \ref{fig:different-shape-of-herimite}, these include "U" shape, "S" shape, "M" shape and "Triple-U" shape.
-
-```{r different-shape-of-herimite, fig.cap="Example residual plots of the linear regression model used in experiment I.
-Four different values of $j$ are included in the experiment results in four different shapes of residuals."}
-stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
-
-ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300)
-
-VI_MODEL$plot(ori_dat) + 
-  ggtitle(quote(He[2]:"U shape")) + 
-  theme_light() -> p1
-
-VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote(He[3]:"S shape")) + 
-  theme_light() -> p2
-
-VI_MODEL$plot(poly_model(shape = 3, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote(He[6]:"M shape")) + 
-  theme_light() -> p3
-
-VI_MODEL$plot(poly_model(shape = 4, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote(He[18]:"Triple-U shape")) + 
-  theme_light() -> p4
-
-patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)
-```
-
-The range of $\sigma$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
-The example residual plots are given in Figure \ref{fig:different-sigma}.
-
-```{r different-sigma, fig.cap="Example residual plots of the linear regression model used in experiment I.
-Four different values of $\\sigma$ are included in the experiment to control the strength of the signal."}
-ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)
-
-VI_MODEL$plot(ori_dat) + 
-  ggtitle(quote(sigma~"="~0.5)) + 
-  theme_light() -> p1
-
-VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1)$gen(300, computed = select(ori_dat, x))) +
-  ggtitle(quote(sigma~"="~1)) + 
-  theme_light() -> p2
-
-VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 2)$gen(300, computed = select(ori_dat, x))) +
-  ggtitle(quote(sigma~"="~2)) + 
-  theme_light() -> p3
-
-VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 4)$gen(300, computed = select(ori_dat, x))) +
-  ggtitle(quote(sigma~"="~4)) + 
-  theme_light() -> p4
-
-patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)
-```
-
-Four different distribution were used to generate $X_{raw}$ shown in Figure \ref{fig:different-dist}.
-The uniform and the normal distribution are symmetric and commonly assumed in statistical models.
-The adjusted log-normal distribution provides skewed density.
-And the discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.
-
-```{r different-dist, fig.cap="Example residual plots of the linear regression model used in experiment I.
-Four different distribution of $x_{raw}$ are used in the experiment to provide various visual patterns."}
-ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)
-
-VI_MODEL$plot(ori_dat) + 
-  ggtitle(quote(U(-1,1))) + 
-  theme_light() -> p1
-
-VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
-  ggtitle(quote(N(0,0.3^2))) + 
-  theme_light() -> p2
-
-VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
-  ggtitle(quote(lognormal(0,0.6^2)/3)) + 
-  theme_light() -> p3
-
-VI_MODEL$plot(poly_model(shape = 1, x = rand_uniform_d(k = 5, even = TRUE), sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
-  ggtitle(quote(U~"{1,5}")) + 
-  theme_light() -> p4
-
-patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)
-```
-
-Figure \ref{fig:example-lineup} shows one of the lineups used in experiment I.
-This lineup was produced under $n = 300$, $j = 6$, $\sigma = 1$ and $X_{raw} \sim N(0.0.3^2)$.
-The actual data plot location was four.
-All five subjects correctly identified the actual data plot for this lineup.
-
-```{r example-lineup, fig.cap="Lineup 24 in experiment I. Can you spot the most different plot? \\label{fig:example-lineup}"}
-VI_MODEL$plot_lineup(vi_lineup[["poly_24"]]$data, 
-                     remove_grid_line = TRUE, 
-                     theme = theme_light(),
-                     remove_axis = TRUE)
-```
-
-<!-- It was generated for $n = 1$, $\sigma = 1$, $j = 1$, $X_{raw} \sim N$.
-For this lineup, x out of k observers picked the actual data plot.
-Each lineup is evaluated by 5 different subjects to provide reasonable estimates of the p-value.
--->
-
-> Experiment II
-
--   Treatments for heteroskedasticity experiment (plots summarising levels)
-
-Experiment data were simulated using three different shapes ($a$ = -1, 0, 1), five different values of $b$ (b = 0.25, 1, 4, 16, 64), three different sample sizes (n = 50, 100, 300) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
-A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}.
-
-The values of $a$ was chosen so that different shapes of heteroskedasticity were included in the residual plot. These include triangle shape, butterfly shape and inverse triangle shape.
-
-```{r different-shape-of-heter}
-stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
-
-ori_dat <- heter_model(a = -1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300)
-
-VI_MODEL$plot(ori_dat) +
-  ggtitle(quote("Triangle shape")) +
-  theme_light() -> p1
-
-VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote("Butterfly shape")) +
-  theme_light() -> p2
-
-VI_MODEL$plot(heter_model(a = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote("Inverse triangle shape")) +
-  theme_light() -> p3
-
-patchwork::wrap_plots(p1, p2, p3, ncol = 3)
-```
-
-The range of $b$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
-
-### Assigning subjects
-
-Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in 192 different lineups.
-For each lineup, the actual data plot was drawn as a standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis.
-The 19 null datasets were generated by the residual rotation technique, and plotted in the same way.
-The lineup consisted of 20 residual plots with one randomly placed actual data plot.
-Figure \ref{fig:example-lineup} is an example of one of these lineups.
-It was produced by using $n = 300$, $j = 6$, $\sigma = 0.5$ and $X_{raw} \sim N(0.0.3^2)$.
-The actual data plot location was four.
-All five subjects correctly identified the actual data plot for this lineup.
-
-In addition, each lineup is designed to be evaluated by five different subjects to provide reasonable estimates of the visual p-value.
-Thus, $192 \times 3 \times 5 / (20 - 2) = 160$ subjects were recruited to satisfy the design of the experiment I.
-
-### Collecting results
 

### 2022-09-28 paper_comparison/paper_comparison/paper_comparison.tex
modify exp design

@@
 of times plot \(i\) being selected in \(K\) evaluations. In case subject
-\(j\) makes multiple selections, they decided to add \(1/s_j\) to
-\(c_i\) instead of one, where \(s_j\) is the number of plots subject
-\(j\) selected for \(j=1,...K\). This ensured \(\sum_{i}c_i=K\). Thus,
-the visual p-value followed by the above model is given as
+\(j\) makes multiple selections, \(1/s_j\) will be added to \(c_i\)
+instead of one, where \(s_j\) is the number of plots subject \(j\)
+selected for \(j=1,...K\). This ensured \(\sum_{i}c_i=K\). Since we are
+only interested in the selections of the actual data plot \(i\), the
+marginal model can be simplified to a beta-binomial model and thus the
+visual p-value is given as
 
@@
 
-Instead, we modelled the probability of a lineup being rejected directly
-using a logistic regression with the natural logarithm of the effect
-size as the only regressor. The same model was applied to different
-subsets of the survey data to make comparison of the power.
+\begin{enumerate}
+\def\labelenumi{(\arabic{enumi})}
+\item
+  Instead, we modelled the probability of a lineup being rejected
+  directly using a logistic regression with the natural logarithm of the
+  effect size as the only regressor. The same model was applied to
+  different subsets of the survey data for making comparison of the
+  power.
+\item
+  Instead, we modelled the plot selection \(c_i\) for the actual data
+  plot \(i\).
+\end{enumerate}
 
@@
 
-\hypertarget{motivation-and-overview}{%
-\subsection{Motivation and overview}\label{motivation-and-overview}}
+\hypertarget{overview}{%
+\subsection{Overview}\label{overview}}
 
-With our knowledge, the effectiveness of visual test relative to the
-equivalent conventional test in regression diagnostics has not been
-validated. Particularly, its ability to detect non-linearity and
-heteroskedasticity compared to model structural test and Breusch--Pagan
-test. Meanwhile, as mentioned in Section \ref{introduction} section, we
-keen to understand the differences between conventional hypothesis
-testing and graphical diagnostics in the application of linear
-regression diagnostics and why conventional tests are not preferred by
-data analysts in linear regression diagnostics. Attempting to find the
-answer to these questions, three experiments were conducted.
-
-The experiment I has ideal scenario for conventional testing, where the
-visual test is not expected to outperform the conventional test. The
-experiment II is a scenario where the conventional test is an
-approximate test, in which the visual test may have a chance to match
-the performance of the conventional test. The experiment III is designed
-for collecting human responses to lineup with only good residual plots
-such that the parameter \(\alpha\) in Equation
-\ref{eq:pvalue-beta-binomial} can be estimated.
-
-Subjects for all three experiments were recruited from an crowdsourcing
-platform called Prolific {[}ref here{]}. Prescreening procedure was
-applied during the recruitment, subjects were required to be fluent in
-English, with \(98\%\) minimum approval rate in other studies and 10
-minimum submissions.
-
-During the experiment, every subject was presented with a block of 20
-lineups. For each lineup, the actual data plot was drawn as a standard
-residual plot of the null model with raw residuals on the y-axis and
-fitted values on the x-axis. An additional horizontal red line was added
-at \(y = 0\) as a helping line. The 19 null datasets were generated by
-the residual rotation technique, and plotted in the same way. The lineup
-consisted of 20 residual plots with one randomly placed actual data
-plot. And for every lineup, the subject was asked to select one or more
-plots that are most different from others, provide a reason for their
-selections, and evaluate how different they think the selected plots
-were from others. If there was no noticeable difference between plots in
-a lineup, subjects were permitted to select zero plots without providing
-the reason. No subject was shown the same lineup twice. Information
-about preferred pronoun, age group, education, and previous experience
-in visual experiment were also collected.
-
-In every block of 20 lineups that presented to a subject, two lineups
-with obvious visual patterns were included as attention checks. A
-subject's submission was only accepted if the actual data plot was
-identified for at least one attention check. Data of rejected
-submissions were discarded automatically to maintain the overall data
-quality.
-
-Overall, we planned to collect 7974 evaluations on 1152 uniqued lineups
-performed by 443 subjects.
+Three experiments were conducted to investigate the difference between
+conventional hypothesis testing and graphical diagnostics in the
+application of linear regression diagnostics. The experiment I has ideal
+scenario for conventional testing, where the visual test is not expected
+to outperform the conventional test. The experiment II is a scenario
+where the conventional test is an approximate test, in which the visual
+test may have a chance to match the performance of the conventional
+test. The experiment III is designed for collecting human responses to
+lineup with only good residual plots such that the parameter \(\alpha\)
+in Equation \ref{eq:pvalue-beta-binomial} can be estimated. Overall, we
+planned to collect 7974 evaluations on 1152 uniqued lineups performed by
+443 subjects throughout three experiment.
 
@@
 \(\boldsymbol{\varepsilon}\), \(\boldsymbol{x}_{raw}\),
-\(\boldsymbol{z}_{raw}\) are vector of size \(n\), \(He_{j}(.)\) is the
+\(\boldsymbol{z}_{raw}\) are vectors of size \(n\), \(He_{j}(.)\) is the
 \(j\)th-order probabilist's Hermite polynomials,
@@
 out by the null regression, which will result in model misspecification.
+Visual patterns of non-linearity were simulated using four different
+order of probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)) and
+four different distribution of \(X_{raw}\): (1) \(U(-1, 1)\), (2)
+\(N(0, 0.3^2)\), (3) \(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\).
+
+The values of \(j\) was chosen so that different shapes of non-linearity
+were included in the residual plot. As shown in Figure
+\ref{fig:different-shape-of-herimite}, these include ``U'' shape, ``S''
+shape, ``M'' shape and ``Triple-U'' shape.
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-herimite-1} \caption{Example residual plots of the linear regression model used in experiment I.
+Four different values of $j$ are included in the experiment results in four different shapes of residuals.}\label{fig:different-shape-of-herimite}
+\end{figure}
+
+Four different distribution were used to generate \(X_{raw}\) as shown
+in Figure \ref{fig:different-dist}. The uniform and the normal
+distribution are symmetric and commonly assumed in statistical models.
+The adjusted log-normal distribution provides skewed density. And the
+discrete uniform distribution provides discreteness in residual plot,
+which could enrich the pool of visual patterns.
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-dist-1} \caption{Example residual plots of the linear regression model used in experiment I.
+Four different distribution of $x_{raw}$ are used in the experiment to provide various visual patterns.}\label{fig:different-dist}
+\end{figure}
+
+Figure \ref{fig:example-lineup} shows one of the lineups used in
+experiment I. This lineup was produced under \(j = 6\) and
+\(X_{raw} \sim N(0.0.3^2)\). The actual data plot location was four. All
+five subjects correctly identified the actual data plot for this lineup.
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-lineup-1} \caption{Lineup 24 in experiment I.
+Can you spot the most different plot?
+\label{fig:example-lineup}}\label{fig:example-lineup}
+\end{figure}
 
@@
 where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
-\(\boldsymbol{\varepsilon}\) are vector of size \(n\) and \(g(.)\) is
+\(\boldsymbol{\varepsilon}\) are vectors of size \(n\) and \(g(.)\) is
 the scaling function defined in \ref{eq:scaling-function}.
@@
 \(\boldsymbol{\varepsilon}\) is correlated with the regressor \(x\),
-which will lead to the presence of heteroskedasticity.
+which will lead to the presence of heteroskedasticity. Visual patterns
+were simulated using three different shapes (\(a\) = -1, 0, 1) and the
+same four different distribution of \(X_{raw}\) used in experiment I.
+
+The values of \(a\) was chosen so that different shapes of
+heteroskedasticity were included in the residual plot. These include
+left-triangle shape, butterfly shape and right-triangle shape as
+displayed in Figure \ref{fig:different-shape-of-heter}.
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-heter-1} \caption{Example residual plots of the linear regression model used in experiment II.
+Three different shapes (a = -1, 0, 1) are used in the experiment to create left-triangle shape, butterfly shape and right-triangle shape.}\label{fig:different-shape-of-heter}
+\end{figure}
 
@@
 
-\begin{quote}
-Experiment I
-\end{quote}
-
-\begin{itemize}
-\tightlist
-\item
-  Treatments for nonlinearity experiment (plots summarising levels)
-\end{itemize}
-
-Data of experiment I were simulated using four different order of
-probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)), three different
-sample sizes (n = 50, 100, 300), four different standard deviations of
-the error (\(\sigma\) = 0.5, 1, 2, 4) and four different distribution of
-\(X_{raw}\): (1) \(U(-1, 1)\), (2) \(N(0, 0.3^2)\), (3)
-\(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). A summary of the
-parameters used in this experiment is given in Table
-\ref{tab:parameter-table}.
-
-\begin{table}
-
-\caption{\label{tab:parameter-table}Parameter values for $n$, $j$ $\sigma$, $X_{raw}$}
-\centering
-\begin{tabular}[t]{rrrl}
-\toprule
-\multicolumn{1}{c}{\makecell[c]{Sample size\\($n$)}} & \multicolumn{1}{c}{\makecell[c]{Order of Hermite polynomial\\($j$)}} & \multicolumn{1}{c}{\makecell[c]{Error SD\\($\sigma$)}} & \multicolumn{1}{c}{\makecell[c]{Distribution of $X_{raw}$\\}} \\
-\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4}
-50 & 2 & 0.5 & $U(-1, 1)$\\
-100 & 3 & 1.0 & $N(0, 0.3^2)$\\
-300 & 6 & 2.0 & $lognormal(0, 0.6^2)/3$\\
- & 18 & 4.0 & $U\{1, 5\}$\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-The values of \(j\) was chosen so that different shapes of non-linearity
-were included in the residual plot. As shown in Figure
-\ref{fig:different-shape-of-herimite}, these include ``U'' shape, ``S''
-shape, ``M'' shape and ``Triple-U'' shape.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-herimite-1} \caption{Example residual plots of the linear regression model used in experiment I.
-Four different values of $j$ are included in the experiment results in four different shapes of residuals.}\label{fig:different-shape-of-herimite}
-\end{figure}
-
-The range of \(\sigma\), which is a factor controlling the strength of
-the signal, was chosen so that different difficulty levels of lineups
-were generated, and therefore, the estimated power curve would be smooth
-and continuous. The example residual plots are given in Figure
-\ref{fig:different-sigma}.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-sigma-1} \caption{Example residual plots of the linear regression model used in experiment I.
-Four different values of $\sigma$ are included in the experiment to control the strength of the signal.}\label{fig:different-sigma}
-\end{figure}
-
-Four different distribution were used to generate \(X_{raw}\) shown in
-Figure \ref{fig:different-dist}. The uniform and the normal distribution
-are symmetric and commonly assumed in statistical models. The adjusted
-log-normal distribution provides skewed density. And the discrete
-uniform distribution provides discreteness in residual plot, which could
-enrich the pool of visual patterns.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-dist-1} \caption{Example residual plots of the linear regression model used in experiment I.
-Four different distribution of $x_{raw}$ are used in the experiment to provide various visual patterns.}\label{fig:different-dist}
-\end{figure}
-
-Figure \ref{fig:example-lineup} shows one of the lineups used in
-experiment I. This lineup was produced under \(n = 300\), \(j = 6\),
-\(\sigma = 1\) and \(X_{raw} \sim N(0.0.3^2)\). The actual data plot
-location was four. All five subjects correctly identified the actual
-data plot for this lineup.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-lineup-1} \caption{Lineup 24 in experiment I.
-Can you spot the most different plot?
-\label{fig:example-lineup}}\label{fig:example-lineup}
-\end{figure}
-
-\begin{quote}
-Experiment II
-\end{quote}
-
-\begin{itemize}
-\tightlist
-\item
-  Treatments for heteroskedasticity experiment (plots summarising
-  levels)
-\end{itemize}
-
-Experiment data were simulated using three different shapes (\(a\) = -1,
-0, 1), five different values of \(b\) (b = 0.25, 1, 4, 16, 64), three
-different sample sizes (n = 50, 100, 300) and four different
-distribution of \(X_{raw}\): (1) \(U(-1, 1)\), (2) \(N(0, 0.3^2)\), (3)
-\(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). A summary of the
-parameters used in this experiment is given in Table
-\ref{tab:parameter-table}.
-
-The values of \(a\) was chosen so that different shapes of
-heteroskedasticity were included in the residual plot. These include
-triangle shape, butterfly shape and inverse triangle shape.
-
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-heter-1}
-
-The range of \(b\), which is a factor controlling the strength of the
-signal, was chosen so that different difficulty levels of lineups were
-generated, and therefore, the estimated power curve would be smooth and
-continuous.
+To control the strength of the signal, two parameters were used
+including \(\sigma\) and \(b\) for controlling the standard deviation of
+the error term in experiment I and II respectively.
 
@@
 
-Three replications are made for each of the parameter values shown in
-Table \ref{tab:parameter-table} resulting in 192 different lineups. For
-each lineup, the actual data plot was drawn as a standard residual plot
-of the null model with raw residuals on the y-axis and fitted values on
-the x-axis. The 19 null datasets were generated by the residual rotation
-technique, and plotted in the same way. The lineup consisted of 20
-residual plots with one randomly placed actual data plot. Figure
-\ref{fig:example-lineup} is an example of one of these lineups. It was
-produced by using \(n = 300\), \(j = 6\), \(\sigma = 0.5\) and
-\(X_{raw} \sim N(0.0.3^2)\). The actual data plot location was four. All
-five subjects correctly identified the actual data plot for this lineup.
-
-In addition, each lineup is designed to be evaluated by five different
-subjects to provide reasonable estimates of the visual p-value. Thus,
-\(192 \times 3 \times 5 / (20 - 2) = 160\) subjects were recruited to
-satisfy the design of the experiment I.
-
 \hypertarget{collecting-results}{%
 \subsubsection{Collecting results}\label{collecting-results}}
+
+Subjects for all three experiments were recruited from an crowdsourcing
+platform called Prolific {[}ref here{]}. Prescreening procedure was
+applied during the recruitment, subjects were required to be fluent in
+English, with \(98\%\) minimum approval rate in other studies and 10
+minimum submissions. During the experiment, every subject was presented
+with a block of 20 lineups. For each lineup, the actual data plot was
+drawn as a standard residual plot of the null model with raw residuals
+on the y-axis and fitted values on the x-axis. An additional horizontal
+red line was added at \(y = 0\) as a helping line. The 19 null datasets
+were generated by the residual rotation technique, and plotted in the
+same way. The lineup consisted of 20 residual plots with one randomly
+placed actual data plot. And for every lineup, the subject was asked to
+select one or more plots that are most different from others, provide a
+reason for their selections, and evaluate how different they think the
+selected plots were from others. If there was no noticeable difference
+between plots in a lineup, subjects were permitted to select zero plots
+without providing the reason. No subject was shown the same lineup
+twice. Information about preferred pronoun, age group, education, and
+previous experience in visual experiment were also collected. In every
+block of 20 lineups that presented to a subject, two lineups with
+obvious visual patterns were included as attention checks. A subject's
+submission was only accepted if the actual data plot was identified for
+at least one attention check. Data of rejected submissions were
+discarded automatically to maintain the overall data quality.
 

### 2022-09-28 paper_comparison/paper_comparison/paper_comparison.Rmd
added susan and modifying captions

@@
     email: emi.tanaka@monash.edu
+  - name: Susan VanderPlas
+    affil: b
+    email: susan.vanderplas@unl.edu
 affiliation:
@@
       Department of Econometrics and Business Statistics, Monash University, Clayton, VIC, Australia
+  - num: b
+    address: |
+      Department of Statistics, University of Nebraska, Lincoln, Nebraska, USA
 bibliography: paper.bib
@@
 
-  table_dat$f <- kableExtra::cell_spec(table_dat$f, bold = as.numeric(table_dat$f) < 0.05)
-  table_dat$b <- kableExtra::cell_spec(table_dat$b, bold = as.numeric(table_dat$b) < 0.05)
-  table_dat$s <- kableExtra::cell_spec(table_dat$s, bold = as.numeric(table_dat$s) < 0.05)
+  table_dat$f <- kableExtra::cell_spec(table_dat$f, italic = as.numeric(table_dat$f) < 0.05)
+  table_dat$b <- kableExtra::cell_spec(table_dat$b, italic = as.numeric(table_dat$b) < 0.05)
+  table_dat$s <- kableExtra::cell_spec(table_dat$s, italic = as.numeric(table_dat$s) < 0.05)
 
@@
                booktabs = TRUE, 
-               caption  = 'Results of example residual plots given in Figure \\ref{fig:commondepartures} examined by the Model structural test (F-test), the Breusch-Pagan test and the Shapiro–Wilk test.
-P-values smaller than 0.05 are showed in bold.',
+               caption  = 'Statistical significance testing for departures from good residuals for plots in Figure \\ref{fig:commondepartures}.
+Shown are the $p$-values calculated for the conventional Model structural, the Breusch-Pagan and the Shapiro–Wilk tests.
+The good residual plot (A) is judged a good residual plot, as expected, by all tests.
+The non-linearity (B) is detected by all tests, as might be expected given the extreme structure.',
                label = "example-residual-plot-table" )
@@
 
-```{r first-example-lineup, fig.cap="An example lineup consists of 20 randomly placed residual plots, where 19 are null plots with data consistent with the null hypothesis, and one is the actual residual plot.
-Can you find the most different plot?"}
+```{r first-example-lineup, fig.cap="Visual testing is conducted using a lineup, as in the example here.
+The residual plot computed from the observed data (plot 11, exhibiting non-linearity) is embedded among 19 null plots, where the residuals were simulated from a standard error model.
+Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot.
+A small $p$-value would result from a substantial number selecting plot 11."}
 mod <- poly_model(include_z = TRUE, sigma = 1)
@@
 
-### P-value calculation
+### Calculating $p$-values
 

### 2022-09-28 paper_comparison/paper_comparison/paper_comparison.tex
added susan and modifying captions

@@
 
-\author{\name{Weihao Li$^{a}$, Dianne Cook$^{a}$, Emi Tanaka$^{a}$}
+\author{\name{Weihao Li$^{a}$, Dianne Cook$^{a}$, Emi
+Tanaka$^{a}$, Susan VanderPlas$^{b}$}
 \affil{$^{a}$Department of Econometrics and Business Statistics, Monash
-University, Clayton, VIC, Australia}
+University, Clayton, VIC, Australia; $^{b}$Department of Statistics,
+University of Nebraska, Lincoln, Nebraska, USA}
 }
@@
 Cook. Email: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}, Emi
-Tanaka. Email: \href{mailto:emi.tanaka@monash.edu}{\nolinkurl{emi.tanaka@monash.edu}}}
+Tanaka. Email: \href{mailto:emi.tanaka@monash.edu}{\nolinkurl{emi.tanaka@monash.edu}}, Susan
+VanderPlas. Email: \href{mailto:susan.vanderplas@unl.edu}{\nolinkurl{susan.vanderplas@unl.edu}}}
 
@@
 
-\caption{\label{tab:example-residual-plot-table}Results of example residual plots given in Figure \ref{fig:commondepartures} examined by the Model structural test (F-test), the Breusch-Pagan test and the Shapiro–Wilk test.
-P-values smaller than 0.05 are showed in bold.}
+\caption{\label{tab:example-residual-plot-table}Statistical significance testing for departures from good residuals for plots in Figure \ref{fig:commondepartures}.
+Shown are the $p$-values calculated for the conventional Model structural, the Breusch-Pagan and the Shapiro–Wilk tests.
+The good residual plot (A) is judged a good residual plot, as expected, by all tests.
+The non-linearity (B) is detected by all tests, as might be expected given the extreme structure.}
 \centering
@@
 A & None & 0.434 & 0.133 & 0.728\\
-B & Non-linearity & \textbf{0.000} & \textbf{0.000} & \textbf{0.039}\\
-C & Heteroskedasticity & 0.378 & \textbf{0.000} & \textbf{0.000}\\
-D & Non-normality & 0.667 & 0.736 & \textbf{0.000}\\
+B & Non-linearity & \em{0.000} & \em{0.000} & \em{0.039}\\
+C & Heteroskedasticity & 0.378 & \em{0.000} & \em{0.000}\\
+D & Non-normality & 0.667 & 0.736 & \em{0.000}\\
 \bottomrule
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/first-example-lineup-1} \caption{An example lineup consists of 20 randomly placed residual plots, where 19 are null plots with data consistent with the null hypothesis, and one is the actual residual plot.
-Can you find the most different plot?}\label{fig:first-example-lineup}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/first-example-lineup-1} \caption{Visual testing is conducted using a lineup, as in the example here.
+The residual plot computed from the observed data (plot 11, exhibiting non-linearity) is embedded among 19 null plots, where the residuals were simulated from a standard error model.
+Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot.
+A small $p$-value would result from a substantial number selecting plot 11.}\label{fig:first-example-lineup}
 \end{figure}
@@
 
-\hypertarget{p-value-calculation}{%
-\subsubsection{P-value calculation}\label{p-value-calculation}}
+\hypertarget{calculating-p-values}{%
+\subsubsection{\texorpdfstring{Calculating
+\(p\)-values}{Calculating p-values}}\label{calculating-p-values}}
 

### 2022-10-01 paper_comparison/paper_comparison/paper_comparison.Rmd
correct p-value formula

@@
 
-```{r}
+```{r poly-conventional-simulation}
 stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
@@
 
-```{r}
+```{r heter-conventional-simulation}
 stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
@@
 
-```{r commondepartures, fig.cap = "Example fitted vs residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
+```{r residual-plot-common-departures, fig.cap = "Example fitted vs residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
 The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted plot."}
@@
 Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals.
-As shown in Figure \ref{fig:commondepartures}, the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns.
+As shown in Figure \ref{fig:residual-plot-common-departures}, the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns.
 There are various types of departures from a good residual plot.
@@
 The usual pattern of heterskedasticity on a residual plot is the inconsistent spread of the residuals at different x values.
-Visually, it sometimes results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure \ref{fig:commondepartures}, or the "left-triangle" and "right-triganle" shape where the smallest variance occurs at the edges of the x-axis.
+Visually, it sometimes results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure \ref{fig:residual-plot-common-departures}, or the "left-triangle" and "right-triganle" shape where the smallest variance occurs at the edges of the x-axis.
 
@@
 In the case that the normality assumption is violated, it is expected to observe data points that do not centralize around the horizontal zero line and unevenly distribute at both sides of the zero line.
-For example, given a skewed error distribution, fewer data points and more outliers are on one side of the zero line as shown at the bottom-right of Figure \ref{fig:commondepartures}.
+For example, given a skewed error distribution, fewer data points and more outliers are on one side of the zero line as shown at the bottom-right of Figure \ref{fig:residual-plot-common-departures}.
 
@@
 
-Example residual plots given in Figure \ref{fig:commondepartures} are examined by the corresponding model structural test, Breusch-Pagan test and Shapiro--Wilk test as shown in Table \ref{tab:example-residual-plot-table}.
+Example residual plots given in Figure \ref{fig:residual-plot-common-departures} are examined by the corresponding model structural test, Breusch-Pagan test and Shapiro--Wilk test as shown in Table \ref{tab:example-residual-plot-table}.
 In the example, both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null hypothesis for departures that they do not intend to examine.
@@
                booktabs = TRUE, 
-               caption  = 'Statistical significance testing for departures from good residuals for plots in Figure \\ref{fig:commondepartures}.
+               caption  = 'Statistical significance testing for departures from good residuals for plots in Figure \\ref{fig:residual-plot-common-departures}.
 Shown are the $p$-values calculated for the conventional Model structural, the Breusch-Pagan and the Shapiro–Wilk tests.
@@
 This ensured $\sum_{i}c_i=K$.
-Since we are only interested in the selections of the actual data plot $i$, the marginal model can be simplified to a beta-binomial model and thus the visual p-value is given as
+Since we are only interested in the selections of the actual data plot $i$, the marginal model can be simplified to a beta-binomial model and thus the visual p-value is given as`
+
+<!-- TODO:Clarify this is only for integer c_i -->
 
@@
 \begin{equation} \label{eq:pvalue-beta-binomial}
-P(C \geq c_i) = \int_{x=c_i}^{K}\frac{\Gamma(K + 1)}{\Gamma(x + 1)\Gamma(K - x + 1)}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)}dx,
+P(C \geq c_i) = \sum_{x=c_i}^{K}\frac{\Gamma(K + 1)}{\Gamma(x + 1)\Gamma(K - x + 1)}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)}dx,
 \end{equation}
@@
 ```
+
+<!-- TODO:Introduce the concept of c interesting plot -->
 
@@
 However, knowing this probability is insufficient to determine the power of a visual test allowed for multiple selections as it doesn't provide information about the number of selections made by the subject, and hence, the p-value can't be computed.
+Recall that the p-value is computed
 

### 2022-10-01 Third_experiment/exp_setup.R

@@ -1,33 +1,17 @@
 library(ggExtra)
 library(tidyverse)
-library(bandicoot)
 library(visage)
-library(lhs)
 library(cli)
-
-trans_unif_2_norm <- function(x, sd = 1) {
-  truncnorm::qtruncnorm(x, sd = sd, a = -1, b = 1)
-}
-
-trans_unif_2_discrete <- function(x, level = NULL) {
-  level[trunc(x * length(level)) + 1]
-}
-
-
-
-# Shape of the trunc density
-# ggplot() +
-#   geom_density(aes(trans_unif_2_norm(seq(0, 1, 0.00001))), col = "red")
 
 
 # V1: shape = 1, 2, 3, 4
 # V3: N: 50, 100, 300
 # V4: x_dist: uniform, normal, lognormal, even_dis
 # V5: e_sigma: 0.125, 0.25, 0.5, 1
+# V6: rep: 3
 
-# 6 * 3 * 6 * 4 = 432
-
-# rep: 5
+# 4 * 3 * 4 * 4 * 3 = 576 lineups
+# 576 * 5 / 18 = 160 subjects
 
 set.seed(10086)
 
@@ -53,20 +37,6 @@ check_comb <- function(set_dat, new_dat, n = 1) {
           return(FALSE)
         })
 }
-
-# check_comb(lineup %>%
-#              filter(lineup_id %in% c(193, 100, 162, 135, 50)) %>%
-#              select(-lineup_id),
-#            lineup %>%
-#              filter(lineup_id %in% c(147)) %>%
-#              select(-lineup_id),
-#            4)
-#
-# lineup %>%
-#   filter(lineup_id %in% c(193, 100, 162, 135, 50)) %>%
-#   select(-lineup_id)
-#
-# 147
 
 # Caculate distance between a set of allocated lineups and a new lineup
 calc_dist <- function(lineup, allocated_lineup_ids, new_lineup_id) {

### 2022-10-01 paper_comparison/paper_comparison/paper_comparison.Rmd
fill in exp overview

@@
 There are various types of departures from a good residual plot.
-Two most commonly checked departures - non-linearity, heterskedasticity and non-normality will be discussed in this paper.
-
+
+Non-linearity, heterskedasticity and non-normality are perhaps the three mostly commonly checked departures.
 Non-linearity is a type of model misspecification caused by failing to include higher order terms of the regressors in the regression equation.
@@
 One can clearly observe the "S-shape" from the residual plot as the cubic term is not captured by the misspecified model.
-
 Heterskedasticity refers to the presence of nonconst error variance in a regression model.
@@
 Visually, it sometimes results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure \ref{fig:residual-plot-common-departures}, or the "left-triangle" and "right-triganle" shape where the smallest variance occurs at the edges of the x-axis.
-
 Compared to non-linearity and heterskedasticity, non-normality is usually harder to be detected from a residual plot since scatter plot is not excel in revealing marginal distribution.
@@
 This ensured $\sum_{i}c_i=K$.
-Since we are only interested in the selections of the actual data plot $i$, the marginal model can be simplified to a beta-binomial model and thus the visual p-value is given as`
+Since we are only interested in the selections of the actual data plot $i$, the marginal model can be simplified to a beta-binomial model and thus the visual p-value is given as
 
@@
 \begin{equation} \label{eq:pvalue-beta-binomial}
-P(C \geq c_i) = \sum_{x=c_i}^{K}\frac{\Gamma(K + 1)}{\Gamma(x + 1)\Gamma(K - x + 1)}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)}dx,
+P(C \geq c_i) = \sum_{x=c_i}^{K}\frac{\Gamma(K + 1)}{\Gamma(x + 1)\Gamma(K - x + 1)}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},
 \end{equation}
@@
 Previously, it was addressed by modelling the probability of a subject $i$ correctly picking the actual data plot from a lineup $l$ using a mixed-effect logistic regression with the subject being treated as a random effect [@majumder_validation_2013].
-However, knowing this probability is insufficient to determine the power of a visual test allowed for multiple selections as it doesn't provide information about the number of selections made by the subject, and hence, the p-value can't be computed.
-Recall that the p-value is computed
-
-(1) Instead, we modelled the probability of a lineup being rejected directly using a logistic regression with the natural logarithm of the effect size as the only regressor.
+However, having this probability is insufficient to determine the power of a visual test allowed for multiple selections as it doesn't provide information about the number of selections made by the subject for p-value calculation.
+Instead, we directly estimated the probability of a lineup being rejected using a logistic regression with the natural logarithm of the effect size as the only regressor.
 The same model was applied to different subsets of the survey data for making comparison of the power.
-
-(2) Instead, we modelled the plot selection $c_i$ for the actual data plot $i$.
-
-<!-- -   Very briefly explain what a lineup test is -->
-
-<!-- -   Lineup (maybe using your previous residual plot for nonlinearity) -->
-
-<!-- -   $p$-value calulcation, briefly, including how multiple selections are handled -->
-
-<!-- -   power calculation, briefly, including how multiple selections are handled -->
-
-```{=html}
-<!-- ![Parallelism between multiple quantitative testing and visual discovery [@buja_statistical_2009].
-Visible features in a plot are viewed as a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$, and any visual discoveries that are inconsistent with the null hypothesis are treated as evidence against the null.
-For regression diagnostics, the null hypothesis would be the assumed model, and visual discoveries could be any visual features in favour of any alternatives.
-\label{fig:parallelism}](figures/rsta2009012001.jpg){width=450 height=341}
--->
-```
 
@@
 
-Three experiments were conducted to investigate the difference between conventional hypothesis testing and graphical diagnostics in the application of linear regression diagnostics.
+Three experiments were conducted to investigate the difference between conventional hypothesis testing and visual inference in the application of linear regression diagnostics.
 The experiment I has ideal scenario for conventional testing, where the visual test is not expected to outperform the conventional test.
-The experiment II is a scenario where the conventional test is an approximate test, in which the visual test may have a chance to match the performance of the conventional test.
+Meanwhile, the experiment II is a scenario where the conventional test is an approximate test, in which the visual test may have a chance to match the performance of the conventional test.
 The experiment III is designed for collecting human responses to lineup with only good residual plots such that the parameter $\alpha$ in Equation \ref{eq:pvalue-beta-binomial} can be estimated.
@@
 
-Two types of departures were considered with the corresponding data generating process being designed for experiment I and II.
+Two types of departures, namely non-linearity and heteroskedasticity, were considered with the corresponding data generating process being designed for experiment I and II.
 

### 2022-10-01 paper_comparison/paper_comparison/paper_comparison.tex
correct p-value formula
fill in exp overview

@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/commondepartures-1} \caption{Example fitted vs residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
-The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted plot.}\label{fig:commondepartures}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/residual-plot-common-departures-1} \caption{Example fitted vs residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
+The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted plot.}\label{fig:residual-plot-common-departures}
 \end{figure}
@@
 \citet{cook1982residuals}. As shown in Figure
-\ref{fig:commondepartures}, the top-left panel is a good residual plot
-with residuals evenly distributed at both sides of the horizontal zero
-line showing no noticeable patterns. There are various types of
-departures from a good residual plot. Two most commonly checked
-departures - non-linearity, heterskedasticity and non-normality will be
-discussed in this paper.
+\ref{fig:residual-plot-common-departures}, the top-left panel is a good
+residual plot with residuals evenly distributed at both sides of the
+horizontal zero line showing no noticeable patterns. There are various
+types of departures from a good residual plot.
 
-Non-linearity is a type of model misspecification caused by failing to
-include higher order terms of the regressors in the regression equation.
-Any non-linear functional form of residuals on fitted values presented
-in the residual plot could be considered as an indicative of
-non-linearity. An example residual plot containing visual pattern of
-non-linearity is given at the top-right of Figure {[}ref here{]}. One
-can clearly observe the ``S-shape'' from the residual plot as the cubic
-term is not captured by the misspecified model.
-
-Heterskedasticity refers to the presence of nonconst error variance in a
-regression model. It is mostly due to the strict but false assumptions
-on the variance-covariance matrix of the error term. The usual pattern
-of heterskedasticity on a residual plot is the inconsistent spread of
-the residuals at different x values. Visually, it sometimes results in
-the so-called ``butterfly'' shape as shown in the bottom-left panel of
-Figure \ref{fig:commondepartures}, or the ``left-triangle'' and
-``right-triganle'' shape where the smallest variance occurs at the edges
-of the x-axis.
-
-Compared to non-linearity and heterskedasticity, non-normality is
-usually harder to be detected from a residual plot since scatter plot is
-not excel in revealing marginal distribution. A favourable graphical
-summarise for this task is the quantile-quantile plot. However, for a
-consistent comparison in the later part of this paper, residual plot
-will be the focus. Besides, it is important to note that not all
-regression models assume normality for the error term, but a certain
-amount do including the classical normal linear regression model. In the
-case that the normality assumption is violated, it is expected to
-observe data points that do not centralize around the horizontal zero
-line and unevenly distribute at both sides of the zero line. For
-example, given a skewed error distribution, fewer data points and more
-outliers are on one side of the zero line as shown at the bottom-right
-of Figure \ref{fig:commondepartures}.
+Non-linearity, heterskedasticity and non-normality are perhaps the three
+mostly commonly checked departures. Non-linearity is a type of model
+misspecification caused by failing to include higher order terms of the
+regressors in the regression equation. Any non-linear functional form of
+residuals on fitted values presented in the residual plot could be
+considered as an indicative of non-linearity. An example residual plot
+containing visual pattern of non-linearity is given at the top-right of
+Figure {[}ref here{]}. One can clearly observe the ``S-shape'' from the
+residual plot as the cubic term is not captured by the misspecified
+model. Heterskedasticity refers to the presence of nonconst error
+variance in a regression model. It is mostly due to the strict but false
+assumptions on the variance-covariance matrix of the error term. The
+usual pattern of heterskedasticity on a residual plot is the
+inconsistent spread of the residuals at different x values. Visually, it
+sometimes results in the so-called ``butterfly'' shape as shown in the
+bottom-left panel of Figure \ref{fig:residual-plot-common-departures},
+or the ``left-triangle'' and ``right-triganle'' shape where the smallest
+variance occurs at the edges of the x-axis. Compared to non-linearity
+and heterskedasticity, non-normality is usually harder to be detected
+from a residual plot since scatter plot is not excel in revealing
+marginal distribution. A favourable graphical summarise for this task is
+the quantile-quantile plot. However, for a consistent comparison in the
+later part of this paper, residual plot will be the focus. Besides, it
+is important to note that not all regression models assume normality for
+the error term, but a certain amount do including the classical normal
+linear regression model. In the case that the normality assumption is
+violated, it is expected to observe data points that do not centralize
+around the horizontal zero line and unevenly distribute at both sides of
+the zero line. For example, given a skewed error distribution, fewer
+data points and more outliers are on one side of the zero line as shown
+at the bottom-right of Figure \ref{fig:residual-plot-common-departures}.
 
@@
 
-Example residual plots given in Figure \ref{fig:commondepartures} are
-examined by the corresponding model structural test, Breusch-Pagan test
-and Shapiro--Wilk test as shown in Table
+Example residual plots given in Figure
+\ref{fig:residual-plot-common-departures} are examined by the
+corresponding model structural test, Breusch-Pagan test and
+Shapiro--Wilk test as shown in Table
 \ref{tab:example-residual-plot-table}. In the example, both the
@@
 
-\caption{\label{tab:example-residual-plot-table}Statistical significance testing for departures from good residuals for plots in Figure \ref{fig:commondepartures}.
+\caption{\label{tab:example-residual-plot-table}Statistical significance testing for departures from good residuals for plots in Figure \ref{fig:residual-plot-common-departures}.
 Shown are the $p$-values calculated for the conventional Model structural, the Breusch-Pagan and the Shapiro–Wilk tests.
@@
 \begin{equation} \label{eq:pvalue-beta-binomial}
-P(C \geq c_i) = \int_{x=c_i}^{K}\frac{\Gamma(K + 1)}{\Gamma(x + 1)\Gamma(K - x + 1)}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)}dx,
+P(C \geq c_i) = \sum_{x=c_i}^{K}\frac{\Gamma(K + 1)}{\Gamma(x + 1)\Gamma(K - x + 1)}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},
 \end{equation}
@@
 the subject being treated as a random effect
-\citep{majumder_validation_2013}. However, knowing this probability is
+\citep{majumder_validation_2013}. However, having this probability is
 insufficient to determine the power of a visual test allowed for
 multiple selections as it doesn't provide information about the number
-of selections made by the subject, and hence, the p-value can't be
-computed.
-
-\begin{enumerate}
-\def\labelenumi{(\arabic{enumi})}
-\item
-  Instead, we modelled the probability of a lineup being rejected
-  directly using a logistic regression with the natural logarithm of the
-  effect size as the only regressor. The same model was applied to
-  different subsets of the survey data for making comparison of the
-  power.
-\item
-  Instead, we modelled the plot selection \(c_i\) for the actual data
-  plot \(i\).
-\end{enumerate}
+of selections made by the subject for p-value calculation. Instead, we
+directly estimated the probability of a lineup being rejected using a
+logistic regression with the natural logarithm of the effect size as the
+only regressor. The same model was applied to different subsets of the
+survey data for making comparison of the power.
 
@@
 Three experiments were conducted to investigate the difference between
-conventional hypothesis testing and graphical diagnostics in the
-application of linear regression diagnostics. The experiment I has ideal
-scenario for conventional testing, where the visual test is not expected
-to outperform the conventional test. The experiment II is a scenario
-where the conventional test is an approximate test, in which the visual
-test may have a chance to match the performance of the conventional
-test. The experiment III is designed for collecting human responses to
-lineup with only good residual plots such that the parameter \(\alpha\)
-in Equation \ref{eq:pvalue-beta-binomial} can be estimated. Overall, we
-planned to collect 7974 evaluations on 1152 uniqued lineups performed by
-443 subjects throughout three experiment.
+conventional hypothesis testing and visual inference in the application
+of linear regression diagnostics. The experiment I has ideal scenario
+for conventional testing, where the visual test is not expected to
+outperform the conventional test. Meanwhile, the experiment II is a
+scenario where the conventional test is an approximate test, in which
+the visual test may have a chance to match the performance of the
+conventional test. The experiment III is designed for collecting human
+responses to lineup with only good residual plots such that the
+parameter \(\alpha\) in Equation \ref{eq:pvalue-beta-binomial} can be
+estimated. Overall, we planned to collect 7974 evaluations on 1152
+uniqued lineups performed by 443 subjects throughout three experiment.
 
@@
 
-Two types of departures were considered with the corresponding data
-generating process being designed for experiment I and II.
+Two types of departures, namely non-linearity and heteroskedasticity,
+were considered with the corresponding data generating process being
+designed for experiment I and II.
 
@@
 
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-6-1}
+
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-7-1}
+
 \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-8-1}
@@
 \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-11-1}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-11-2}
 
 \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-12-1}
-
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-13-1}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-13-2}
-
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-14-1}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-14-2}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-12-2}
 

### 2022-10-01 paper_comparison/paper_comparison/paper_comparison.Rmd
fill in exp setup

@@
 
-## Overview
+## Experiment overview
 
@@
 
-```{r results='asis', eval = FALSE}
+```{r results='asis'}
 data.frame(j = c(2, 3, 6, 18),
@@
                caption  = 'Parameter values for $j$ and $X_{raw}$', 
-               label = "parameter-table" ) %>%
+               label = "model-parameter-table" ) %>%
   kableExtra::add_header_above(c("Order of Hermite polynomial\n(sym1)", "Distribution of sym4\n")) %>%
@@
 
-The values of $j$ was chosen so that different shapes of non-linearity were included in the residual plot.
-As shown in Figure \ref{fig:different-shape-of-herimite}, these include "U" shape, "S" shape, "M" shape and "Triple-U" shape.
+The values of $j$ was chosen so that distinct shapes of non-linearity were included in the residual plot.
+A greater value of $j$ will result in a curve with more turning points.
+As shown in Figure \ref{fig:different-shape-of-herimite}, it includes "U" shape, "S" shape, "M" shape and "Triple-U" shape.
+It is expected that the "U" shape will be the easiest one to detect because complex shape tends to be concealed by cluster of data points.
 
@@
 
-<!-- The range of $\sigma$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
-The example residual plots are given in Figure \ref{fig:different-sigma}.
--->
-
-```{r different-sigma, eval = FALSE, fig.cap="Example residual plots of the linear regression model used in experiment I.
-Four different values of $\\sigma$ are included in the experiment to control the strength of the signal."}
-ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)
-
-VI_MODEL$plot(ori_dat) + 
-  ggtitle(quote(sigma~"="~0.5)) + 
-  theme_light() -> p1
-
-VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1)$gen(300, computed = select(ori_dat, x))) +
-  ggtitle(quote(sigma~"="~1)) + 
-  theme_light() -> p2
-
-VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 2)$gen(300, computed = select(ori_dat, x))) +
-  ggtitle(quote(sigma~"="~2)) + 
-  theme_light() -> p3
-
-VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 4)$gen(300, computed = select(ori_dat, x))) +
-  ggtitle(quote(sigma~"="~4)) + 
-  theme_light() -> p4
-
-patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)
-```
-
 Four different distribution were used to generate $X_{raw}$ as shown in Figure \ref{fig:different-dist}.
 The uniform and the normal distribution are symmetric and commonly assumed in statistical models.
-The adjusted log-normal distribution provides skewed density.
-And the discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.
+The adjusted log-normal distribution provides skewed density, while the discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.
 
@@
 
-Figure \ref{fig:example-lineup} shows one of the lineups used in experiment I.
+Figure \ref{fig:example-poly-lineup} shows one of the lineups used in experiment I.
 This lineup was produced under $j = 6$ and $X_{raw} \sim N(0.0.3^2)$.
@@
 
-```{r example-lineup, fig.cap="Lineup 24 in experiment I. Can you spot the most different plot? \\label{fig:example-lineup}"}
+```{r example-poly-lineup, fig.cap="Lineup poly-24 in experiment I. Can you spot the most different plot? \\label{fig:example-poly-lineup}"}
 VI_MODEL$plot_lineup(vi_lineup[["poly_24"]]$data, 
@@
 VI_MODEL$plot(ori_dat) +
-  ggtitle(quote("Left-triangle shape")) +
+  ggtitle(quote("a = -1: Left-triangle shape")) +
   theme_light() -> p1
@@
 VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote("Butterfly shape")) +
+  ggtitle(quote("a = 0: Butterfly shape")) +
   theme_light() -> p2
@@
 VI_MODEL$plot(heter_model(a = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote("Right-triangle shape")) +
+  ggtitle(quote("a = 1: Right-triangle shape")) +
   theme_light() -> p3
 
-patchwork::wrap_plots(p1, p2, p3, ncol = 3)
-```
+patchwork::wrap_plots(p1, p2, p3, ncol = 2)
+```
+
+```{r example-heter-lineup, fig.cap="Lineup heter-169 in experiment II. Can you spot the most different plot? \\label{fig:example-heter-lineup}"}
+VI_MODEL$plot_lineup(vi_lineup[["heter_169"]]$data, 
+                     remove_grid_line = TRUE, 
+                     theme = theme_light(),
+                     remove_axis = TRUE)
+```
+
+An example lineup of this model is shown in Figure \ref{fig:example-heter-lineup} with $a = -1$ and $X_{raw} \sim U(-1, 1)$.
+The actual data plot location was 15.
+8 out of 11 subjects correctly identified the actual data plot for this lineup.
 
@@
 
-### Treatments
-
-To control the strength of the signal, two parameters were used including $\sigma$ and $b$ for controlling the standard deviation of the error term in experiment I and II respectively.
-
-<!-- > Experiment I -->
-
-<!-- -   Treatments for nonlinearity experiment (plots summarising levels) -->
-
-<!-- It was generated for $n = 1$, $\sigma = 1$, $j = 1$, $X_{raw} \sim N$.
-For this lineup, x out of k observers picked the actual data plot.
-Each lineup is evaluated by 5 different subjects to provide reasonable estimates of the p-value.
--->
+### Controlling the strength of the signal
+
+Except for the parameters summarised in Table \ref{tab:model-parameter-table}, three additional parameters $n$, $\sigma$ and $b$ were used to control the strength of the signal so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
+
+In experiment I, II and III, three different sample sizes were selected (n = 50, 100, 300).
+In general, with fewer data points drawn in a residual plot, the chance of detecting the visual pattern will decrease.
+This can be observed from Figure \ref{fig:different-n}.
+
+```{r different-n, fig.cap="Three different values of $n$ are used in experiment I, II and III to control the strength of the signal."}
+ori_dat <- poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(300)
+
+VI_MODEL$plot(ori_dat) + 
+  ggtitle(quote(n~"="~300)) + 
+  theme_light() -> p1
+
+VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:50,], x, e))) +
+  ggtitle(quote(n~"="~50)) + 
+  theme_light() -> p2
+
+VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:100,], x, e))) +
+  ggtitle(quote(n~"="~100)) + 
+  theme_light() -> p3
+
+patchwork::wrap_plots(p2, p3, p1, ncol = 2)
+```
+
+Four different levels of $sigma$
+
+```{r different-sigma, eval = FALSE, fig.cap="Example residual plots of the linear regression model used in experiment I.
+Four different values of $\\sigma$ are included in the experiment to control the strength of the signal."}
+ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)
+
+VI_MODEL$plot(ori_dat) + 
+  ggtitle(quote(sigma~"="~0.5)) + 
+  theme_light() -> p1
+
+VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1)$gen(300, computed = select(ori_dat, x))) +
+  ggtitle(quote(sigma~"="~1)) + 
+  theme_light() -> p2
+
+VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 2)$gen(300, computed = select(ori_dat, x))) +
+  ggtitle(quote(sigma~"="~2)) + 
+  theme_light() -> p3
+
+VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 4)$gen(300, computed = select(ori_dat, x))) +
+  ggtitle(quote(sigma~"="~4)) + 
+  theme_light() -> p4
+
+patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)
+```
 

### 2022-10-01 paper_comparison/paper_comparison/paper_comparison.tex
fill in exp setup

@@
 
-\hypertarget{overview}{%
-\subsection{Overview}\label{overview}}
+\hypertarget{experiment-overview}{%
+\subsection{Experiment overview}\label{experiment-overview}}
 
@@
 
-The values of \(j\) was chosen so that different shapes of non-linearity
-were included in the residual plot. As shown in Figure
-\ref{fig:different-shape-of-herimite}, these include ``U'' shape, ``S''
-shape, ``M'' shape and ``Triple-U'' shape.
+\begin{table}
+
+\caption{\label{tab:model-parameter-table}Parameter values for $j$ and $X_{raw}$}
+\centering
+\begin{tabular}[t]{rl}
+\toprule
+\multicolumn{1}{c}{\makecell[c]{Order of Hermite polynomial\\($j$)}} & \multicolumn{1}{c}{\makecell[c]{Distribution of $X_{raw}$\\}} \\
+\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2}
+2 & $U(-1, 1)$\\
+3 & $N(0, 0.3^2)$\\
+6 & $lognormal(0, 0.6^2)/3$\\
+18 & $U\{1, 5\}$\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+The values of \(j\) was chosen so that distinct shapes of non-linearity
+were included in the residual plot. A greater value of \(j\) will result
+in a curve with more turning points. As shown in Figure
+\ref{fig:different-shape-of-herimite}, it includes ``U'' shape, ``S''
+shape, ``M'' shape and ``Triple-U'' shape. It is expected that the ``U''
+shape will be the easiest one to detect because complex shape tends to
+be concealed by cluster of data points.
 
@@
 distribution are symmetric and commonly assumed in statistical models.
-The adjusted log-normal distribution provides skewed density. And the
+The adjusted log-normal distribution provides skewed density, while the
 discrete uniform distribution provides discreteness in residual plot,
@@
 
-Figure \ref{fig:example-lineup} shows one of the lineups used in
+Figure \ref{fig:example-poly-lineup} shows one of the lineups used in
 experiment I. This lineup was produced under \(j = 6\) and
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-lineup-1} \caption{Lineup 24 in experiment I.
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-poly-lineup-1} \caption{Lineup poly-24 in experiment I.
 Can you spot the most different plot?
-\label{fig:example-lineup}}\label{fig:example-lineup}
+\label{fig:example-poly-lineup}}\label{fig:example-poly-lineup}
 \end{figure}
@@
 
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-heter-lineup-1} \caption{Lineup heter-169 in experiment II.
+Can you spot the most different plot?
+\label{fig:example-heter-lineup}}\label{fig:example-heter-lineup}
+\end{figure}
+
+\begin{verbatim}
+##  [1]  TRUE FALSE  TRUE  TRUE FALSE FALSE  TRUE  TRUE  TRUE  TRUE  TRUE
+\end{verbatim}
+
+An example lineup of this model is shown in Figure
+\ref{fig:example-heter-lineup} with \(a = -1\) and
+\(X_{raw} \sim U(-1, 1)\). The actual data plot location was 15. 8 out
+of 11 subjects correctly identified the actual data plot for this
+lineup.
+
 \hypertarget{experimental-setup}{%
@@
 
-\hypertarget{treatments}{%
-\subsubsection{Treatments}\label{treatments}}
+\hypertarget{controlling-the-strength-of-the-signal}{%
+\subsubsection{Controlling the strength of the
+signal}\label{controlling-the-strength-of-the-signal}}
 
-To control the strength of the signal, two parameters were used
-including \(\sigma\) and \(b\) for controlling the standard deviation of
-the error term in experiment I and II respectively.
+Other than the parameters summarised in Table
+\ref{tab:model-parameter-table}, three additional parameters \(n\),
+\(\sigma\) and \(b\) were used to control the strength of the signal in
+experiment I and II so that different difficulty levels of lineups were
+generated, and therefore, the estimated power curve would be smooth and
+continuous. Three different sample sizes were selected (n = 50, 100,
+300). In general, with fewer data points shown in a residual plot, the
+chance of detecting the visual pattern will decrease. This can be
+observed from Figure
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-n-1} \caption{Example residual plots of the linear regression model used in experiment I.
+Three different values of $n$ are included in the experiment to control the strength of the signal.}\label{fig:different-n}
+\end{figure}
 

### 2022-10-02 paper_comparison/paper_comparison/paper_comparison.Rmd
fill in allocating subjects

@@
 
-Except for the parameters summarised in Table \ref{tab:model-parameter-table}, three additional parameters $n$, $\sigma$ and $b$ were used to control the strength of the signal so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
-
-In experiment I, II and III, three different sample sizes were selected (n = 50, 100, 300).
-In general, with fewer data points drawn in a residual plot, the chance of detecting the visual pattern will decrease.
-This can be observed from Figure \ref{fig:different-n}.
+As summarised by Table \ref{tab:model-parameter-table}, three additional parameters $n$, $\sigma$ and $b$ were used to control the strength of the signal so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
+Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ were used in experiment I and II respectively.
+Figure \ref{fig:different-sigma} and \ref{fig:different-b} demonstrate the impact of these two parameters.
+
+Three different sample sizes were used (n = 50, 100, 300) in all three experiments.
+It can be observed from Figure \ref{fig:different-n} that with fewer data points drawn in a residual plot, the visual pattern is more difficult to be detected.
 
@@
 
-Four different levels of $sigma$
-
-```{r different-sigma, eval = FALSE, fig.cap="Example residual plots of the linear regression model used in experiment I.
+```{r different-sigma, fig.cap="Example residual plots of the linear regression model used in experiment I.
 Four different values of $\\sigma$ are included in the experiment to control the strength of the signal."}
@@
 
-<!-- > Experiment II -->
-
-<!-- -   Treatments for heteroskedasticity experiment (plots summarising levels) -->
-
-<!-- Experiment data were simulated using three different shapes ($a$ = -1, 0, 1), five different values of $b$ (b = 0.25, 1, 4, 16, 64), three different sample sizes (n = 50, 100, 300) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
-A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}.
--->
-
-<!-- The range of $b$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
--->
+```{r different-b, fig.cap="Example residual plots of the linear regression model used in experiment II.
+Five different values of $b$ are included in the experiment to control the strength of the signal."}
+stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
+
+ori_dat <- heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 0.25)$gen(300)
+
+VI_MODEL$plot(ori_dat) +
+  ggtitle(quote("b = 0.25")) +
+  theme_light() -> p1
+
+VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 1)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote("b = 1")) +
+  theme_light() -> p2
+
+VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 4)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote("b = 4")) +
+  theme_light() -> p3
+
+VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 16)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote("b = 16")) +
+  theme_light() -> p4
+
+VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 64)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote("b = 64")) +
+  theme_light() -> p5
+
+patchwork::wrap_plots(p1, p2, p3, p4, p5, ncol = 3)
+```
 
@@
 
-<!-- Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in 192 different lineups.
+Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in 443 different lineups.
 For each lineup, the actual data plot was drawn as a standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis.
@@
 The lineup consisted of 20 residual plots with one randomly placed actual data plot.
-Figure \ref{fig:example-lineup} is an example of one of these lineups.
-It was produced by using $n = 300$, $j = 6$, $\sigma = 0.5$ and $X_{raw} \sim N(0.0.3^2)$.
-The actual data plot location was four.
-All five subjects correctly identified the actual data plot for this lineup.
--->
-
-<!-- In addition, each lineup is designed to be evaluated by five different subjects to provide reasonable estimates of the visual p-value.
+
+In addition, each lineup is designed to be evaluated by five different subjects to provide reasonable estimates of the visual p-value.
+
 Thus, $192 \times 3 \times 5 / (20 - 2) = 160$ subjects were recruited to satisfy the design of the experiment I.
--->
 

### 2022-10-02 paper_comparison/paper_comparison/paper_comparison.tex
fill in allocating subjects

@@
 
-\begin{verbatim}
-##  [1]  TRUE FALSE  TRUE  TRUE FALSE FALSE  TRUE  TRUE  TRUE  TRUE  TRUE
-\end{verbatim}
-
 An example lineup of this model is shown in Figure
@@
 
-Other than the parameters summarised in Table
-\ref{tab:model-parameter-table}, three additional parameters \(n\),
-\(\sigma\) and \(b\) were used to control the strength of the signal in
-experiment I and II so that different difficulty levels of lineups were
+As summarised by Table \ref{tab:model-parameter-table}, three additional
+parameters \(n\), \(\sigma\) and \(b\) were used to control the strength
+of the signal so that different difficulty levels of lineups were
 generated, and therefore, the estimated power curve would be smooth and
-continuous. Three different sample sizes were selected (n = 50, 100,
-300). In general, with fewer data points shown in a residual plot, the
-chance of detecting the visual pattern will decrease. This can be
-observed from Figure
+continuous. Parameter \(\sigma \in \{0.5, 1, 2, 4\}\) and
+\(b \in \{0.25, 1, 4, 16, 64\}\) were used in experiment I and II
+respectively. Figure \ref{fig:different-sigma} and \ref{fig:different-b}
+demonstrate the impact of these two parameters.
+
+Three different sample sizes were used (n = 50, 100, 300) in all three
+experiments. It can be observed from Figure \ref{fig:different-n} that
+with fewer data points drawn in a residual plot, the visual pattern is
+more difficult to be detected.
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-n-1} \caption{Example residual plots of the linear regression model used in experiment I.
-Three different values of $n$ are included in the experiment to control the strength of the signal.}\label{fig:different-n}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-n-1} \caption{Three different values of $n$ are used in experiment I, II and III to control the strength of the signal.}\label{fig:different-n}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-sigma-1} \caption{Example residual plots of the linear regression model used in experiment I.
+Four different values of $\sigma$ are included in the experiment to control the strength of the signal.}\label{fig:different-sigma}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-b-1} \caption{Example residual plots of the linear regression model used in experiment II.
+Five different values of $b$ are included in the experiment to control the strength of the signal.}\label{fig:different-b}
 \end{figure}
@@
 \subsubsection{Assigning subjects}\label{assigning-subjects}}
+
+Three replications are made for each of the parameter values shown in
+Table \ref{tab:parameter-table} resulting in 443 different lineups. For
+each lineup, the actual data plot was drawn as a standard residual plot
+of the null model with raw residuals on the y-axis and fitted values on
+the x-axis. The 19 null datasets were generated by the residual rotation
+technique, and plotted in the same way. The lineup consisted of 20
+residual plots with one randomly placed actual data plot.
+
+Overall, we planned to collect 7974 evaluations on 1152 uniqued lineups
+performed by 443 subjects throughout three experiment.
+
+In addition, each lineup is designed to be evaluated by five different
+subjects to provide reasonable estimates of the visual p-value.
+
+Thus, \(192 \times 3 \times 5 / (20 - 2) = 160\) subjects were recruited
+to satisfy the design of the experiment I.
 

### 2022-10-03 paper_comparison/paper_comparison/paper_comparison.Rmd
clean up results

@@
 
+<!-- ```{r results='asis'} -->
+<!-- data.frame(n = c(50, 100, 300, ""), -->
+<!--            j = c(2, 3, 6, 18), -->
+<!--            sigma = c(0.5, 1, 2, 4), -->
+<!--            x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$")) %>% -->
+<!--   knitr::kable(col.names = NULL, -->
+<!--                escape = FALSE, -->
+<!--                format = "latex", -->
+<!--                booktabs = TRUE,  -->
+<!--                caption  = 'Parameter values for $n$, $j$ $\\sigma$, $X_{raw}$',  -->
+<!--                label = "parameter-table" ) %>% -->
+<!--   kableExtra::add_header_above(c("Sample size\n(sym3)", "Order of Hermite polynomial\n(sym1)", "Error SD\n(sym2)", "Distribution of sym4\n")) %>% -->
+<!--   as.character() %>% -->
+<!--   gsub("sym1", "$j$", .) %>% -->
+<!--   gsub("sym2", "$\\\\sigma$", .) %>% -->
+<!--   gsub("sym3", "$n$", .) %>% -->
+<!--   gsub("sym4", "$X_{raw}$", .) %>% -->
+<!--   cat() -->
+<!--   # gsub("\\\\\\$", "$", .) %>% -->
+<!--   # cat() -->
+<!-- ``` -->
+
 ### Heteroskedasticity
@@
 
-### Assigning subjects
-
-Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in 443 different lineups.
-For each lineup, the actual data plot was drawn as a standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis.
-The 19 null datasets were generated by the residual rotation technique, and plotted in the same way.
-The lineup consisted of 20 residual plots with one randomly placed actual data plot.
-
-In addition, each lineup is designed to be evaluated by five different subjects to provide reasonable estimates of the visual p-value.
-
-Thus, $192 \times 3 \times 5 / (20 - 2) = 160$ subjects were recruited to satisfy the design of the experiment I.
+### Subject allocation
+
+Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in 1152 different lineups.
+In addition, each lineup is designed to be evaluated by five different subjects and lineups with uniform distribution is designed to be evaluated by 11 subjects to provide reasonable estimates of the visual p-value.
+
+Thus, $443$ subjects were recruited to satisfy the design of the experiment. 
+<!-- TODO:subject allocation algorithm -->
 
@@
 
-software/technical
-
-## Power comparison
+<!-- TODO:pipeline description -->
+
+## Results
+
+### Power comparison
 
@@
 }
-```
-
-```{r}
-vi_survey_processed %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  rename(p_value_dirichlet = p_value_select_all_plots) %>%
-  ggplot() +
-  geom_point(aes(p_value_uniform, p_value_dirichlet)) +
-  geom_abline() +
-  scale_x_log10() +
-  scale_y_log10()
 ```
@@
               se = FALSE) +
-  facet_wrap(~x_dist) +
-  ggtitle("poly")
-```
-
-```{r}
-vi_survey_processed %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type == "polynomial") %>%
-  filter(e_sigma != 0.125) %>%
-  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
-  mutate(reject_uniform = as.numeric(p_value_uniform <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value_dirichlet"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_dirichlet, times = 500), 
-              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value_dirichlet"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03) +
-  geom_smooth(aes(log_effect_size, reject_uniform, col = "VI_p_value_uniform"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_uniform, times = 500), 
-              aes(log_effect_size, reject_uniform, group = boot_id, col = "VI_p_value_uniform"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03) +
   facet_wrap(~x_dist) +
@@
 
-```{r}
-vi_survey_processed %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type == "heteroskedasticity") %>%
-  filter(e_sigma != 0.125) %>%
-  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
-  mutate(reject_uniform = as.numeric(p_value_uniform <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value_dirichlet"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_dirichlet, times = 500), 
-              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value_dirichlet"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03) +
-  geom_smooth(aes(log_effect_size, reject_uniform, col = "VI_p_value_uniform"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_uniform, times = 500), 
-              aes(log_effect_size, reject_uniform, group = boot_id, col = "VI_p_value_uniform"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03) +
-  facet_wrap(~x_dist) +
-  ggtitle("heter")
-```
-
-```{r}
-p_value_cache_env <- new.env()
-
-if (file.exists(here::here("data/p_value_only_5_eval.rds"))) {
-  p_value_only_5_eval <- readRDS(here::here("data/p_value_only_5_eval.rds"))
-} else {
-  p_value_only_5_eval <- vi_survey_processed %>%
-  filter(x_dist == "uniform") %>%
-  filter((is.na(b) & e_sigma != 0.125 ) | (b != 0 & b < 128)) %>%
-  filter(exp != 5) %>%
-  group_by(unique_lineup_id) %>%
-  summarise(p_value_only_5_eval = calc_p_value(n_detect = sum(detect), 
-                                               n_eval = n(), 
-                                               n_sel = num_selection, 
-                                               dist = "dirichlet",
-                                               alpha = numeric_alpha[1],
-                                               cache_env = p_value_cache_env))
-
-  saveRDS(p_value_only_5_eval, here::here("data/p_value_only_5_eval.rds"))
-}
-
-vi_survey_processed %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type == "heteroskedasticity") %>%
-  filter(x_dist == "uniform") %>%
-  filter(b != 0) %>%
-  filter(b < 128) %>%
-  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  left_join(p_value_only_5_eval) %>%
-  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_dirichlet, times = 500), 
-              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03) +
-  geom_smooth(aes(log_effect_size, reject_only_5_eval, col = "VI_p_value_only_5_eval"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_only_5_eval, times = 500), 
-              aes(log_effect_size, reject_only_5_eval, group = boot_id, col = "VI_p_value_only_5_eval"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03) +
-  ggtitle("heter")
-
-p1 <- vi_survey_processed %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type == "heteroskedasticity") %>%
-  filter(x_dist == "uniform") %>%
-  filter(b != 0) %>%
-  filter(b < 128) %>%
-  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  left_join(p_value_only_5_eval) %>%
-  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_dirichlet, times = 500), 
-              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03)
-
-p2 <- vi_survey_processed %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type == "heteroskedasticity") %>%
-  filter(x_dist == "uniform") %>%
-  filter(b != 0) %>%
-  filter(b < 128) %>%
-  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  left_join(p_value_only_5_eval) %>%
-  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject_only_5_eval, col = "VI_p_value_only_5_eval"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_only_5_eval, times = 500), 
-              aes(log_effect_size, reject_only_5_eval, group = boot_id, col = "VI_p_value_only_5_eval"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03)
-
-patchwork::wrap_plots(p1, p2)
-```
-
-```{r}
-
-vi_survey_processed %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type == "polynomial") %>%
-  filter(x_dist == "uniform") %>%
-  filter(e_sigma != 0.125) %>%
-  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  left_join(p_value_only_5_eval) %>%
-  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_dirichlet, times = 500), 
-              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03) +
-  geom_smooth(aes(log_effect_size, reject_only_5_eval, col = "VI_p_value_only_5_eval"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_only_5_eval, times = 500), 
-              aes(log_effect_size, reject_only_5_eval, group = boot_id, col = "VI_p_value_only_5_eval"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03) +
-  ggtitle("poly")
-
-p1 <- vi_survey_processed %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type == "polynomial") %>%
-  filter(x_dist == "uniform") %>%
-  filter(e_sigma != 0.125) %>%
-  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  left_join(p_value_only_5_eval) %>%
-  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_dirichlet, times = 500), 
-              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03)
-
-p2 <- vi_survey_processed %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type == "polynomial") %>%
-  filter(x_dist == "uniform") %>%
-  filter(e_sigma != 0.125) %>%
-  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  left_join(p_value_only_5_eval) %>%
-  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject_only_5_eval, col = "VI_p_value_only_5_eval"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_only_5_eval, times = 500), 
-              aes(log_effect_size, reject_only_5_eval, group = boot_id, col = "VI_p_value_only_5_eval"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03)
-
-patchwork::wrap_plots(p1, p2)
-
-```
-
-```{r eval = FALSE}
-vi_survey_processed %>%
-  filter(exp == 3) %>%
-  filter(e_sigma != 0.125) %>%
-  mutate()
-
-vi_survey_processed %>%
-  filter(exp == 3) %>%
-  filter(e_sigma != 0.125) %>%
-  mutate(set = factor(set), lineup_id = factor(lineup_id)) %>%
-  {lme4::glmer(weighted_detect ~ effect_size + ( 0 + effect_size | set) + ( 0 + effect_size | lineup_id), family = binomial ,data = .)} %>%
-  predict(type = "response")
-
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE)
-
-+
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_dirichlet, times = 500), 
-              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03) +
-  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
-                pivot_longer(F_p_value:SW_p_value), 
-              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
-              method = "glm", 
-              method.args = list(family = binomial),
-              se = FALSE) +
-  facet_wrap(~x_dist) +
-  ggtitle("poly")
-
-  ?glm
-```
-
-<!-- # Old stuff -->
-
-<!-- ## Computing power curves -->
-
-<!-- For the purpose of examining the effectiveness of visual test in regression diagnostics, two experiments were conducted.
-The experiment I has ideal scenario for conventional testing, where the visual test is not expected to outperform the conventional test.
-The experiment II is a scenario where the conventional test is an approximate test, in which the visual test may have a chance to match the performance of the conventional test.
--->
-
-<!-- Subjects for both experiments were recruited from an crowdsourcing platform called Prolific [ref here].
-Prescreening procedure was applied during the recruitment, subjects were required to be fluent in English, with $98\%$ minimum approval rate in other studies and 10 minimum submissions.
--->
-
-<!-- During the experiment, every subject was presented with a block of 20 lineups.
-And for every lineup, the subject was asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots were from others.
-If there was no noticeable difference between plots in a lineup, subjects were permitted to select zero plots without providing the reason.
-No subject was shown the same lineup twice.
-Information about preferred pronoun, age group, education, and previous experience in visual experiment were also collected.
--->
-
-<!-- A pool of 12 different lineups with obvious visual patterns were generated for experiment I and experiment II respectively.
-In every block of 20 lineups that presented to a subject, two out of 12 lineups were included as attention checks.
-A subject's submission was only accepted if the actual data plot was identified for at least one attention check.
-Data of rejected submissions were discarded automatically to maintain the overall data quality.
--->
-
-<!-- ## Non-linearity -->
-
-<!-- Experiment I is designed to study the ability of human subjects to detect the effect of a random vector $\boldsymbol{z}$ which is a probabilist's Hermite polynomial [Herimite ref here] of another random vector $\boldsymbol{x}$ in a two variable statistical model formulated as: -->
-
-<!-- ```{=tex} -->
-<!-- \begin{align} \label{eq:nonlinearity-model} -->
-<!-- \boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\ -->
-<!-- \boldsymbol{x} = g(\boldsymbol{x}_{raw}, 1), \\ -->
-<!-- \boldsymbol{z} = g(\boldsymbol{z}_{raw}, 1), \\ -->
-<!-- \boldsymbol{z}_{raw} = He_j(g(\boldsymbol{z}, 2)), -->
-<!-- \end{align} -->
-<!-- ``` -->
-<!-- where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vector of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $\{-k, k\}$ defined as -->
-
-<!-- ```{=tex} -->
-<!-- \begin{equation} \label{eq:scaling-function} -->
-<!-- g(\boldsymbol{x}, k) = (\boldsymbol{x} - min(\boldsymbol{x}))/max(\boldsymbol{x} - min(\boldsymbol{x})) \times 2k - k, \quad \text{for} \quad k > 0.  -->
-<!-- \end{equation} -->
-<!-- ``` -->
-<!-- The null regression model used to fit the realizations generated by the above model is formulated as: -->
-
-<!-- ```{=tex} -->
-<!-- \begin{equation} \label{eq:scaling-function} -->
-<!-- \boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u}, -->
-<!-- \end{equation} -->
-<!-- ``` -->
-<!-- where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$. -->
-
-<!-- Model misspecification presents since the null model leaves out the higher order term. -->
-
-<!-- Experiment data were simulated using four different order of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$), three different sample sizes (n = 50, 100, 300) , four different standard deviations of the error ($\sigma$ = 0.5, 1, 2, 4) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
-A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}.
--->
-
-<!-- The values of $j$ was chosen so that different shapes of non-linearity were included in the residual plot. These include "U" shape, "S" shape, "M" shape and "Triple-U" shape. -->
-
-<!-- ```{r different-shape-of-herimite} -->
-<!-- stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1 -->
-
-<!-- ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300) -->
-
-<!-- VI_MODEL$plot(ori_dat) +  -->
-<!--   ggtitle(quote(He[2]:"U shape")) +  -->
-<!--   theme_light() -> p1 -->
-
-<!-- VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) + -->
-<!--   ggtitle(quote(He[3]:"S shape")) +  -->
-<!--   theme_light() -> p2 -->
-
-<!-- VI_MODEL$plot(poly_model(shape = 3, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) + -->
-<!--   ggtitle(quote(He[6]:"M shape")) +  -->
-<!--   theme_light() -> p3 -->
-
-<!-- VI_MODEL$plot(poly_model(shape = 4, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) + -->
-<!--   ggtitle(quote(He[18]:"Triple-U shape")) +  -->
-<!--   theme_light() -> p4 -->
-
-<!-- patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2) -->
-<!-- ``` -->
-
-<!-- The range of $\sigma$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
--->
-
-<!-- ```{r different-sigma} -->
-<!-- ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300) -->
-
-<!-- VI_MODEL$plot(ori_dat) +  -->
-<!--   ggtitle(quote(sigma~"="~0.5)) +  -->
-<!--   theme_light() -> p1 -->
-
-<!-- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1)$gen(300, computed = select(ori_dat, x))) + -->
-<!--   ggtitle(quote(sigma~"="~1)) +  -->
-<!--   theme_light() -> p2 -->
-
-<!-- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 2)$gen(300, computed = select(ori_dat, x))) + -->
-<!--   ggtitle(quote(sigma~"="~2)) +  -->
-<!--   theme_light() -> p3 -->
-
-<!-- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 4)$gen(300, computed = select(ori_dat, x))) + -->
-<!--   ggtitle(quote(sigma~"="~4)) +  -->
-<!--   theme_light() -> p4 -->
-
-<!-- patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2) -->
-<!-- ``` -->
-
-<!-- Four different distribution were used to generate $X_{raw}$.
-The uniform and the normal distribution are symmetric and commonly assumed in statistical models.
-The adjusted log-normal distribution provides skewed density.
-And the discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.
--->
-
-<!-- ```{r different-dist} -->
-<!-- ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300) -->
-
-<!-- VI_MODEL$plot(ori_dat) +  -->
-<!--   ggtitle(quote(U(-1,1))) +  -->
-<!--   theme_light() -> p1 -->
-
-<!-- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) + -->
-<!--   ggtitle(quote(N(0,0.3^2))) +  -->
-<!--   theme_light() -> p2 -->
-
-<!-- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) + -->
-<!--   ggtitle(quote(lognormal(0,0.6^2)/3)) +  -->
-<!--   theme_light() -> p3 -->
-
-<!-- VI_MODEL$plot(poly_model(shape = 1, x = rand_uniform_d(k = 5, even = TRUE), sigma = 0.5)$gen(300, computed = select(ori_dat, e))) + -->
-<!--   ggtitle(quote(U~"{1,5}")) +  -->
-<!--   theme_light() -> p4 -->
-
-<!-- patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2) -->
-<!-- ``` -->
-
-<!-- Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in 192 different lineups.
-For each lineup, the actual data plot was drawn as a standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis.
-The 19 null datasets were generated by the residual rotation technique, and plotted in the same way.
-The lineup consisted of 20 residual plots with one randomly placed actual data plot.
-Figure \ref{fig:example-lineup} is an example of one of these lineups.
-It was produced by using $n = 300$, $j = 6$, $\sigma = 0.5$ and $X_{raw} \sim N(0.0.3^2)$.
-The actual data plot location was four.
-All five subjects correctly identified the actual data plot for this lineup.
--->
-
-<!-- ```{r results='asis'} -->
-<!-- data.frame(n = c(50, 100, 300, ""), -->
-<!--            j = c(2, 3, 6, 18), -->
-<!--            sigma = c(0.5, 1, 2, 4), -->
-<!--            x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$")) %>% -->
-<!--   knitr::kable(col.names = NULL, -->
-<!--                escape = FALSE, -->
-<!--                format = "latex", -->
-<!--                booktabs = TRUE,  -->
-<!--                caption  = 'Parameter values for $n$, $j$ $\\sigma$, $X_{raw}$',  -->
-<!--                label = "parameter-table" ) %>% -->
-<!--   kableExtra::add_header_above(c("Sample size\n(sym3)", "Order of Hermite polynomial\n(sym1)", "Error SD\n(sym2)", "Distribution of sym4\n")) %>% -->
-<!--   as.character() %>% -->
-<!--   gsub("sym1", "$j$", .) %>% -->
-<!--   gsub("sym2", "$\\\\sigma$", .) %>% -->
-<!--   gsub("sym3", "$n$", .) %>% -->
-<!--   gsub("sym4", "$X_{raw}$", .) %>% -->
-<!--   cat() -->
-<!--   # gsub("\\\\\\$", "$", .) %>% -->
-<!--   # cat() -->
-<!-- ``` -->
-
-<!-- ```{r example-lineup, fig.cap="Example lineup \\label{fig:example-lineup}"} -->
-<!-- VI_MODEL$plot_lineup(polynomials_lineup[[23]]$data,  -->
-<!--                      remove_grid_line = TRUE,  -->
-<!--                      theme = theme_light(), -->
-<!--                      remove_axis = TRUE) -->
-<!-- ``` -->
-
-<!-- <!-- It was generated for $n = 1$, $\sigma = 1$, $j = 1$, $X_{raw} \sim N$.
-For this lineup, x out of k observers picked the actual data plot.
-Each lineup is evaluated by 5 different subjects to provide reasonable estimates of the p-value.
---> -->
-
-<!-- In addition, each lineup is designed to be evaluated by five different subjects to provide reasonable estimates of the visual p-value.
-Thus, $192 \times 3 \times 5 / (20 - 2) = 160$ subjects were recruited to satisfy the design of the experiment I.
--->
-
-<!-- ## Heteroskedasticity -->
-
-<!-- Experiment II is designed to study the ability of human subjects to detect the appearance of a heteroskedasticity pattern under a simple linear regression model setting: -->
-
-<!-- ```{=tex} -->
-<!-- \begin{align} \label{eq:heter-model} -->
-<!-- \boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\ -->
-<!-- \boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + 2 - |a| b (\boldsymbol{x} - a)^2 \boldsymbol{I}), \\ -->
-<!-- \end{align} -->
-<!-- ``` -->
-<!-- where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vector of size $n$. -->
-
-<!-- The null regression model used to fit the realizations generated by the above model is formulated exactly the same as the model used in the first experiment: -->
-
-<!-- ```{=tex} -->
-<!-- \begin{equation}  -->
-<!-- \boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u}, -->
-<!-- \end{equation} -->
-<!-- ``` -->
-<!-- where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$. -->
-
-<!-- Model misspecification presents since the assumption about the constant error variance is violated. -->
-
-<!-- Experiment data were simulated using three different shapes ($a$ = -1, 0, 1), five different values of $b$ (b = 0.25, 1, 4, 16, 64), three different sample sizes (n = 50, 100, 300) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
-A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}.
--->
-
-<!-- The values of $a$ was chosen so that different shapes of heteroskedasticity were included in the residual plot. These include triangle shape, butterfly shape and inverse triangle shape. -->
-
-<!-- ```{r different-shape-of-heter} -->
-<!-- stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1 -->
-
-<!-- ori_dat <- heter_model(a = -1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300) -->
-
-<!-- VI_MODEL$plot(ori_dat) +  -->
-<!--   ggtitle(quote("Triangle shape")) +  -->
-<!--   theme_light() -> p1 -->
-
-<!-- VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) + -->
-<!--   ggtitle(quote("Butterfly shape")) +  -->
-<!--   theme_light() -> p2 -->
-
-<!-- VI_MODEL$plot(heter_model(a = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) + -->
-<!--   ggtitle(quote("Inverse triangle shape")) +  -->
-<!--   theme_light() -> p3 -->
-
-<!-- patchwork::wrap_plots(p1, p2, p3, ncol = 3) -->
-<!-- ``` -->
-
-<!-- The range of $b$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
--->
-
-<!-- # Analysis and results -->
-
-<!-- ## Data overview -->
+### Data overview
 
@@
 
-<!-- Subjects recruited from Prolific received a fixed payment for participating in the experiment.
+Subjects recruited from Prolific received a fixed payment for participating in the experiment.
 However, some subjects will try to maximize their earnings for minimum effort.
@@
 Therefore, we only paid for approved submissions and no further data screening procedure needed to be applied on the collected data.
--->
-
-<!-- A subject was allowed to select zero plots for a lineup if there was no visible difference between plots, but the simulated visual p-value given in Equation (\ref{eq:p-value-multi}) will effectively drop the subject from the simulation for this case, leading to inaccurate estimation of $p$-value.
-Therefore, we treated this case as making one selection but failing to identify the actual data plot so that Equation (\ref{eq:p-value-multi}) can be applied correctly.
--->
-
-<!-- In overall, there were a total of 3200 lineup evaluations made by 160 subjects in both experiment I and experiment II respectively, where 320 lineup evaluations were attention checks and were not used in the following analysis.
--->
-
-<!-- The collated dataset is provided in `polynomials` and `heter` of the `visage` `R` package. -->
-
-<!-- ##  -->
+
+In overall, there were a total of 3200 lineup evaluations made by 160 subjects in both experiment I and experiment II respectively, where 320 lineup evaluations were attention checks and were not used in the following analysis.
+
+The collated dataset is provided in `vi_survey` of the `visage` `R` package.
 

### 2022-10-03 paper_comparison/paper_comparison/paper_comparison.tex
clean up results

@@
 
-\hypertarget{assigning-subjects}{%
-\subsubsection{Assigning subjects}\label{assigning-subjects}}
+\hypertarget{subject-allocation}{%
+\subsubsection{Subject allocation}\label{subject-allocation}}
 
 Three replications are made for each of the parameter values shown in
-Table \ref{tab:parameter-table} resulting in 443 different lineups. For
-each lineup, the actual data plot was drawn as a standard residual plot
-of the null model with raw residuals on the y-axis and fitted values on
-the x-axis. The 19 null datasets were generated by the residual rotation
-technique, and plotted in the same way. The lineup consisted of 20
-residual plots with one randomly placed actual data plot.
+Table \ref{tab:parameter-table} resulting in 1152 different lineups. In
+addition, each lineup is designed to be evaluated by five different
+subjects and lineups with uniform distribution is designed to be
+evaluated by 11 subjects to provide reasonable estimates of the visual
+p-value.
 
-Overall, we planned to collect 7974 evaluations on 1152 uniqued lineups
-performed by 443 subjects throughout three experiment.
-
-In addition, each lineup is designed to be evaluated by five different
-subjects to provide reasonable estimates of the visual p-value.
-
-Thus, \(192 \times 3 \times 5 / (20 - 2) = 160\) subjects were recruited
-to satisfy the design of the experiment I.
+Thus, \(443\) subjects were recruited to satisfy the design of the
+experiment.
 
@@
 
-software/technical
+\hypertarget{results}{%
+\subsection{Results}\label{results}}
 
 \hypertarget{power-comparison}{%
-\subsection{Power comparison}\label{power-comparison}}
+\subsubsection{Power comparison}\label{power-comparison}}
 
@@
 
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-8-1}
+\hypertarget{data-overview}{%
+\subsubsection{Data overview}\label{data-overview}}
 
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-9-1}
+Subjects recruited from Prolific received a fixed payment for
+participating in the experiment. However, some subjects will try to
+maximize their earnings for minimum effort. During the review of
+submissions, if we found a subject objectively demonstrated clear
+low-effort throughout the experiment, i.e., failed all attention checks,
+we rejected the submission. The rejected submissions will be removed
+immediately, and Prolific will automatically recruit another subject as
+substitution. Therefore, we only paid for approved submissions and no
+further data screening procedure needed to be applied on the collected
+data.
 
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-10-1}
+In overall, there were a total of 3200 lineup evaluations made by 160
+subjects in both experiment I and experiment II respectively, where 320
+lineup evaluations were attention checks and were not used in the
+following analysis.
 
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-11-1}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-11-2}
-
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-12-1}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-12-2}
-
---\textgreater{}
+The collated dataset is provided in \texttt{vi\_survey} of the
+\texttt{visage} \texttt{R} package.
 

### 2022-10-03 paper_comparison/paper_comparison/paper_comparison.Rmd
fill in analysis

@@
 However, having this probability is insufficient to determine the power of a visual test allowed for multiple selections as it doesn't provide information about the number of selections made by the subject for p-value calculation.
-Instead, we directly estimated the probability of a lineup being rejected using a logistic regression with the natural logarithm of the effect size as the only regressor.
-The same model was applied to different subsets of the survey data for making comparison of the power.
+
+Instead, we directly estimated the probability of a lineup being rejected using a logistic regression with the natural logarithm of the effect size as the only regressor formulated as:
+
+```{=tex}
+\begin{equation} \label{eq:logistic-regression-1-1}
+Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(\boldsymbol{E})),
+\end{equation}
+```
+where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
+
+Effect $E$ is derived from the Kullback-Leibler divergence (see [appendix ref here]) formulated as:
+
+```{=tex}
+\begin{equation} \label{eq:effect-size-ex1}
+E = \frac{1}{2\sigma^2}\boldsymbol{X}_b'\boldsymbol{R}_a'(diag(\boldsymbol{R}_a))^{-1}\boldsymbol{R}_a\boldsymbol{X}_b,
+\end{equation}
+```
+where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$.
+
+To study various factors contributing to the power of the visual test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors.
+This includes [expansion].
 
@@
 
-## Results
-
-### Power comparison
+# Results
+
+## Data overview
+
+<!-- How many people? How many lineups? -->
+
+Subjects recruited from Prolific received a fixed payment for participating in the experiment.
+However, some subjects will try to maximize their earnings for minimum effort.
+During the review of submissions, if we found a subject objectively demonstrated clear low-effort throughout the experiment, i.e., failed all attention checks, we rejected the submission.
+The rejected submissions will be removed immediately, and Prolific will automatically recruit another subject as substitution.
+Therefore, we only paid for approved submissions and no further data screening procedure needed to be applied on the collected data.
+
+In overall, there were a total of 3200 lineup evaluations made by 160 subjects in both experiment I and experiment II respectively, where 320 lineup evaluations were attention checks and were not used in the following analysis.
+
+The collated dataset is provided in `vi_survey` of the `visage` `R` package.
+
+## Power comparison
+
+Figure \ref{fig:power-poly-uniform} shows the estimated power of visual test against natural logarithm of the effect with comparison to the power of an exact test - F-test, and the power of two other residual-based conventional tests commonly used in regression diagnostics but for testing other departures from the model assumptions.
+In overall, the power of all four tests increases as the effect becomes larger.
+The power curve of F-test climbs aggressively from 25% to around 90% as $log_e(E)$ increases from 0 to 2, while others respond inactively to the change of effect and remain lower than 25% throughout the period, showing that as an exact test, the F-test is relatively more sensitive to the type of model defects that being considered.
+The power of visual test arises steadily and nearly linearly to around 90% as $log_e(E)$ increases from 2 to 5, suggesting that the effect starts to make noticeable impact on the degree of the presence of the designed visual features.
+Other two inappropriate conventional tests shows improvement at the same time but at a lower rate.
+This coincides the point made by @cook1982residuals mentioned in \ref{hypothesis-testing} that residual-based tests for a specific type of model defect are sensitive to other types of model defects.
+At $log_e(E) = 6$, the power curve of F-test reaches almost 100% followed by the visual test by a small margin.
+The power of Breusch--Pagan test and Shapiro--Wilk test reach around 75% and 30% respectively.
+
+What truly impress us is the huge difference between the estimated power of visual test and the estimated power of F-test.
+The margin is largest at around $log_e(E) = 2$.
+An example lineup is included in Figure \ref{fig:power-overview} where none of subjects detect the actual data plot positioned at panel 14.
+It demonstrates that at this level of difficulty, the designed visual feature is rarely visible, making the actual data plot indistinguishable from residual plots simulated from the assumed model.
+From a communication perspective, given the fact that the visual difference is unperceivable, the argument that non-linearity present in the fitted model is less convincing to the public even though it is true.
+At around $log_e(E) = 3$, the margin gets smaller as the chance of identifying the actual data plot becomes larger.
+At this level of difficulty, the designed visual features are usually detectable but it may not stand out from the lineup as other null plots may happen to include outliers or visual patterns that are are considered to be more attractive by human, and thus recognized as the most different plot.
+Without knowing the designed visual features beforehand, it is actually hard to identify the actual data plot by pure image comparison.
+The corresponding example lineup for $log_e(E) = 3$ shown in Figure \ref{fig:fig:power-poly-uniform} has the actual data plot positioned at panel 20, where two out of five subjects detect it.
+It can be observed that a M-shape is presented in plot 20, but the signal is not strong enough to attract all five subjects, resulting in a visual p-value sightly above the desired significance level $\alpha = 0.05$.
+At $log_e(E) = 4$ and $log_e(E) = 6$, the designed visual features become much clear and attractive, leading to a high percentage of rejection of the null hypothesis.
+Figure \ref{fig:fig:power-poly-uniform} gives example lineups of such cases.
+It is also interesting to observe that the power curve of visual test behaves similarly to those inappropriate conventional tests until the designed visual features really becomes visible.
 
@@
 
-```{r}
+```{r fig:power-poly-uniform, fig.cap = "..."}
+vi_survey_processed %>%
+  filter(x_dist == "uniform") %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(type == "polynomial") %>%
+  filter(e_sigma != 0.125) %>%
+  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  ggplot() +
+  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) + 
+  stat_smooth(geom = "line", 
+              data = ~boot_y(.x, reject_dirichlet, times = 500), 
+              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, 
+              alpha = 0.03) +
+  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
+                filter(x_dist == "uniform") %>%
+                pivot_longer(F_p_value:SW_p_value), 
+              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
+              method = "glm", 
+              method.args = list(family = binomial),
+              se = FALSE) +
+  theme_light(base_size = 5) -> p
+
+VI_MODEL$plot_lineup(vi_lineup[["poly_39"]]$data, 
+                     remove_grid_line = TRUE, 
+                     theme = theme_light(base_size = 5),
+                     remove_axis = TRUE)
+
+VI_MODEL$plot_lineup(vi_lineup[["poly_415"]]$data, 
+                     remove_grid_line = TRUE, 
+                     theme = theme_light(base_size = 5),
+                     remove_axis = TRUE) -> p1
+
+VI_MODEL$plot_lineup(vi_lineup[["poly_131"]]$data, 
+                     remove_grid_line = TRUE, 
+                     theme = theme_light(base_size = 5),
+                     remove_axis = TRUE) -> p2
+
+VI_MODEL$plot_lineup(vi_lineup[["poly_322"]]$data, 
+                     remove_grid_line = TRUE, 
+                     theme = theme_light(base_size = 5),
+                     remove_axis = TRUE) -> p3
+
+VI_MODEL$plot_lineup(vi_lineup[["poly_378"]]$data, 
+                     remove_grid_line = TRUE, 
+                     theme = theme_light(base_size = 5),
+                     remove_axis = TRUE) -> p4
+
+design <- c(patchwork::area(1, 1, 1, 8),
+            patchwork::area(2, 1, 2, 4), 
+            patchwork::area(2, 5, 2, 8), 
+            patchwork::area(3, 1, 3, 4), 
+            patchwork::area(3, 5, 3, 8))
+
+patchwork::wrap_plots(p, B = p1, C = p2, D = p3, E = p4, design = design)
+```
+
+## Distribution of regressor
+
+The impact of the distribution of $X_raw$ on the power is shown in Figure \ref{fig:dist-power}.
+The power curve of F-test is stable across different distributions, while the visual test has a steeper power curve for normal and uniform distribution.
+BP-test performs worse for discrete uniform distribution and uniform distribution but has relatively high power for normal distribution.
+SW-test outperforms BP-test for discrete uniform distribution but remains as the worst test for other distributions.
+The results indicate those inappropriate residual-based tests are sensitive to the distribution of the regressor.
+
+```{r dist-power, fig.cap="..."}
 vi_survey_processed %>%
@@
   geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) +
   stat_smooth(geom = "line",
@@
   facet_wrap(~x_dist) +
-  ggtitle("poly")
+  theme_light(base_size = 5)
 ```
@@
 
-### Data overview
-
-<!-- How many people? How many lineups? -->
-
-Subjects recruited from Prolific received a fixed payment for participating in the experiment.
-However, some subjects will try to maximize their earnings for minimum effort.
-During the review of submissions, if we found a subject objectively demonstrated clear low-effort throughout the experiment, i.e., failed all attention checks, we rejected the submission.
-The rejected submissions will be removed immediately, and Prolific will automatically recruit another subject as substitution.
-Therefore, we only paid for approved submissions and no further data screening procedure needed to be applied on the collected data.
-
-In overall, there were a total of 3200 lineup evaluations made by 160 subjects in both experiment I and experiment II respectively, where 320 lineup evaluations were attention checks and were not used in the following analysis.
-
-The collated dataset is provided in `vi_survey` of the `visage` `R` package.
-
-<!-- ## Nonlinearity -->
-
-<!-- -   Power curve overall -->
-<!-- -   Four lineups also shown, selected from ones close to the visual power curve from the uniform treatment -->
-<!-- -   Compare visual power against different factors -->
-
-<!-- ## Heteroskedasticity -->
-
-<!-- -   Power curve overall -->
-<!-- -   Four lineups also shown, selected from ones close to the visual power curve from the uniform treatment -->
-<!-- -   Compare visual power against different factors -->
-
-<!-- ## Demographic summary -->
-
-<!-- Table \ref{tab:demographic-table} tabulates the number of subjects, preferred pronouns, education backgrounds, age groups, and previous experience in visual experiment.
-Figure \ref{fig:demographic-summary-plot} visualizes the marginal distribution of each of the category.
-The collated data was a balanced sample among male and female.
-Most of the ages of subject were between 18 to 39, while many of them were between 18 to 24.
-Very few subjects were awarded degree higher than Bachelor Degree.
-Around 40% of subjects had previous experience in visual experiment.
--->
-
-<!-- ```{r results="asis"} -->
-
-<!-- polynomials %>% -->
-<!--   mutate(exp = 1) %>% -->
-<!--   bind_rows(mutate(heter, exp = 2)) %>% -->
-<!--   group_by(exp, set) %>% -->
-<!--   slice_sample(n = 1) %>% -->
-<!--   ungroup() %>% -->
-<!--   count(pronoun, education, age_group, previous_experience) %>% -->
-<!--   arrange(desc(n)) %>% -->
-<!--   knitr::kable(col.names = c("Preferred pronoun", "Education", "Age group", "Previous experience", "Subject"), -->
-<!--                escape = FALSE, -->
-<!--                format = "latex", -->
-<!--                linesep = "", -->
-<!--                booktabs = TRUE,  -->
-<!--                caption  = 'Summary of demographic information',  -->
-<!--                label = "demographic-table" ) %>% -->
-<!--   cat() -->
-<!-- ``` -->
-
-<!-- ```{r demographic-summary-plot, fig.cap="Summary of demographic information. \\label{fig:demographic-summary-plot}"} -->
-<!-- polynomials %>% -->
-<!--   mutate(exp = 1) %>% -->
-<!--   bind_rows(mutate(heter, exp = 2)) %>% -->
-<!--   group_by(exp, set) %>% -->
-<!--   summarise(across(age_group:previous_experience, first)) %>% -->
-<!--   mutate(education = fct_relevel(education, "High School or below")) %>% -->
-<!--   ggplot() + -->
-<!--   geom_bar(aes(y = education), orientation = "y") -> education_plot -->
-
-<!-- polynomials %>% -->
-<!--   mutate(exp = 1) %>% -->
-<!--   bind_rows(mutate(heter, exp = 2)) %>% -->
-<!--   group_by(exp, set) %>% -->
-<!--   summarise(across(age_group:previous_experience, first)) %>% -->
-<!--   ggplot() + -->
-<!--   geom_bar(aes(age_group)) -> age_plot -->
-
-<!-- polynomials %>% -->
-<!--   mutate(exp = 1) %>% -->
-<!--   bind_rows(mutate(heter, exp = 2)) %>% -->
-<!--   group_by(exp, set) %>% -->
-<!--   summarise(across(age_group:previous_experience, first)) %>% -->
-<!--   mutate(pronoun = fct_relevel(pronoun, "He", "She", "They")) %>% -->
-<!--   ggplot() + -->
-<!--   geom_bar(aes(pronoun)) -> gender_plot -->
-
-<!-- polynomials %>% -->
-<!--   mutate(exp = 1) %>% -->
-<!--   bind_rows(mutate(heter, exp = 2)) %>% -->
-<!--   group_by(exp, set) %>% -->
-<!--   summarise(across(age_group:previous_experience, first)) %>% -->
-<!--   ggplot() + -->
-<!--   geom_bar(aes(previous_experience)) -> experience_plot -->
-
-<!-- patchwork::wrap_plots(education_plot, age_plot, gender_plot, experience_plot) -->
-<!-- ``` -->
-
-<!-- ## Model fitting -->
-
-<!-- For each parameter combination, effect $E$ is derived from the Kullback-Leibler divergence (see [appendix ref here]) formulated as: -->
-
-<!-- ```{=tex} -->
-<!-- \begin{equation} \label{eq:effect-size-ex1} -->
-<!-- E = \frac{1}{2\sigma^2}\boldsymbol{X}_b'\boldsymbol{R}_a'(diag(\boldsymbol{R}_a))^{-1}\boldsymbol{R}_a\boldsymbol{X}_b, -->
-<!-- \end{equation} -->
-<!-- ``` -->
-<!-- where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$. -->
-
-<!-- The logistic regression is fit using $log_e(E)$ as the only fixed effect covariate for the power of visual test formulated as: -->
-
-<!-- ```{=tex} -->
-<!-- \begin{equation} \label{eq:logistic-regression-1-1} -->
-<!-- Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(\boldsymbol{E})), -->
-<!-- \end{equation} -->
-<!-- ``` -->
-<!-- where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$. -->
-
-<!-- To study various factors contributing to the power of the visual test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors.
-This includes [expansion].
--->
-
-<!-- Table [table ref here] shows the parameter estimates of the logistic regressions. [Discussion about the numeric estimates here] -->
-
 <!-- ## Power comparison -->

### 2022-10-03 paper_comparison/paper_comparison/paper_comparison.tex
fill in analysis

@@
 multiple selections as it doesn't provide information about the number
-of selections made by the subject for p-value calculation. Instead, we
-directly estimated the probability of a lineup being rejected using a
-logistic regression with the natural logarithm of the effect size as the
-only regressor. The same model was applied to different subsets of the
-survey data for making comparison of the power.
+of selections made by the subject for p-value calculation.
+
+Instead, we directly estimated the probability of a lineup being
+rejected using a logistic regression with the natural logarithm of the
+effect size as the only regressor formulated as:
+
+\begin{equation} \label{eq:logistic-regression-1-1}
+Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(\boldsymbol{E})),
+\end{equation}
+
+where \(\Lambda(.)\) is the standard logistic function given as
+\(\Lambda(z) = exp(z)/(1+exp(z))\).
+
+Effect \(E\) is derived from the Kullback-Leibler divergence (see
+{[}appendix ref here{]}) formulated as:
+
+\begin{equation} \label{eq:effect-size-ex1}
+E = \frac{1}{2\sigma^2}\boldsymbol{X}_b'\boldsymbol{R}_a'(diag(\boldsymbol{R}_a))^{-1}\boldsymbol{R}_a\boldsymbol{X}_b,
+\end{equation}
+
+where \(diag(.)\) is the diagonal matrix constructed from the diagonal
+elements of \(\boldsymbol{R}_a\).
+
+To study various factors contributing to the power of the visual test,
+the same logistic regression model is fit on different subsets of the
+collated data grouped by levels of factors. This includes
+{[}expansion{]}.
 
@@
 \hypertarget{results}{%
-\subsection{Results}\label{results}}
-
-\hypertarget{power-comparison}{%
-\subsubsection{Power comparison}\label{power-comparison}}
-
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-6-1}
-
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-7-1}
+\section{Results}\label{results}}
 
 \hypertarget{data-overview}{%
-\subsubsection{Data overview}\label{data-overview}}
+\subsection{Data overview}\label{data-overview}}
 
@@
 
+\hypertarget{power-comparison}{%
+\subsection{Power comparison}\label{power-comparison}}
+
+Figure \ref{fig:power-poly-uniform} shows the estimated power of visual
+test against natural logarithm of the effect with comparison to the
+power of an exact test - F-test, and the power of two other
+residual-based conventional tests commonly used in regression
+diagnostics but for testing other departures from the model assumptions.
+In overall, the power of all four tests increases as the effect becomes
+larger. The power curve of F-test climbs aggressively from 25\% to
+around 90\% as \(log_e(E)\) increases from 0 to 2, while others respond
+inactively to the change of effect and remain lower than 25\% throughout
+the period, showing that as an exact test, the F-test is relatively more
+sensitive to the type of model defects that being considered. The power
+of visual test arises steadily and nearly linearly to around 90\% as
+\(log_e(E)\) increases from 2 to 5, suggesting that the effect starts to
+make noticeable impact on the degree of the presence of the designed
+visual features. Other two inappropriate conventional tests shows
+improvement at the same time but at a lower rate. This coincides the
+point made by \citet{cook1982residuals} mentioned in
+\ref{hypothesis-testing} that residual-based tests for a specific type
+of model defect are sensitive to other types of model defects. At
+\(log_e(E) = 6\), the power curve of F-test reaches almost 100\%
+followed by the visual test by a small margin. The power of
+Breusch--Pagan test and Shapiro--Wilk test reach around 75\% and 30\%
+respectively.
+
+What truly impress us is the huge difference between the estimated power
+of visual test and the estimated power of F-test. The margin is largest
+at around \(log_e(E) = 2\). An example lineup is included in Figure
+\ref{fig:power-overview} where none of subjects detect the actual data
+plot positioned at panel 14. It demonstrates that at this level of
+difficulty, the designed visual feature is rarely visible, making the
+actual data plot indistinguishable from residual plots simulated from
+the assumed model. From a communication perspective, given the fact that
+the visual difference is unperceivable, the argument that non-linearity
+present in the fitted model is less convincing to the public even though
+it is true. At around \(log_e(E) = 3\), the margin gets smaller as the
+chance of identifying the actual data plot becomes larger. At this level
+of difficulty, the designed visual features are usually detectable but
+it may not stand out from the lineup as other null plots may happen to
+include outliers or visual patterns that are are considered to be more
+attractive by human, and thus recognized as the most different plot.
+Without knowing the designed visual features beforehand, it is actually
+hard to identify the actual data plot by pure image comparison. The
+corresponding example lineup for \(log_e(E) = 3\) shown in Figure
+\ref{fig:fig:power-poly-uniform} has the actual data plot positioned at
+panel 20, where two out of five subjects detect it. It can be observed
+that a M-shape is presented in plot 20, but the signal is not strong
+enough to attract all five subjects, resulting in a visual p-value
+sightly above the desired significance level \(\alpha = 0.05\). At
+\(log_e(E) = 4\) and \(log_e(E) = 6\), the designed visual features
+become much clear and attractive, leading to a high percentage of
+rejection of the null hypothesis. Figure
+\ref{fig:fig:power-poly-uniform} gives example lineups of such cases. It
+is also interesting to observe that the power curve of visual test
+behaves similarly to those inappropriate conventional tests until the
+designed visual features really becomes visible.
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/fig:power-poly-uniform-1} \caption{...}\label{fig:fig:power-poly-uniform-1}
+\end{figure}
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/fig:power-poly-uniform-2} \caption{...}\label{fig:fig:power-poly-uniform-2}
+\end{figure}
+
+\hypertarget{distribution-of-regressor}{%
+\subsection{Distribution of regressor}\label{distribution-of-regressor}}
+
+The impact of the distribution of \(X_raw\) on the power is shown in
+Figure \ref{fig:dist-power}. The power curve of F-test is stable across
+different distributions, while the visual test has a steeper power curve
+for normal and uniform distribution. BP-test performs worse for discrete
+uniform distribution and uniform distribution but has relatively high
+power for normal distribution. SW-test outperforms BP-test for discrete
+uniform distribution but remains as the worst test for other
+distributions. The results indicate those inappropriate residual-based
+tests are sensitive to the distribution of the regressor.
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/dist-power-1} \caption{...}\label{fig:dist-power}
+\end{figure}
+
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-6-1}
+
 --\textgreater{}

### 2022-10-04 paper_comparison/paper_comparison/paper_comparison.Rmd
intro done

--- 2022-10-02 paper_comparison/paper_comparison/paper_comparison.Rmd
@@
 
-```{r poly-conventional-simulation}
+```{r}
 stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
@@
 
-```{r heter-conventional-simulation}
+```{r}
 stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
@@
 Diagnosing a model is the key to determining whether there is anything importantly wrong.
-For linear regression analysis, it is typical to interrogate the residuals.
+For linear regression analysis, we typically interrogate the residuals for model diagnostics.
 Residuals summarise what is not captured by the model, and thus provide the capacity to identify what might be wrong.
-There are many ways that residuals could be assessed.
-
+
+We can assess residuals in multiple ways.
 Residuals might be plotted, as a histogram or quantile-quantile plot to examine the distribution.
@@
 If there is any visually discoverable patterns, the model is potentially misspecified.
-However, it is a very difficult task for a human judge, though to make a decision that there's nothing there.
+However, correctly judging a residual plot where no pattern exists is a painstakingly difficult task for humans.
 It is especially common, particularly among new data analysts to report patterns when an experienced data analyst might quickly conclude that there are none.
-Generally, one looks for departures from nothingness like non-linear dependency or heteroskedasticity.
+[ET: I don't know if the former statement is true?] Generally, one looks for noticeable departures from the model like non-linear dependency or heteroskedasticity.
 It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and use a Breusch-Pagan test [@breusch_simple_1979] for heteroskedasticity.
 
-There is an abundance of literature describing appropriate diagnostic methods for linear regression: @draper1998applied, @montgomery1982introduction, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals.
+Abundance of literature describe appropriate diagnostic methods for linear regression: @draper1998applied, @montgomery1982introduction, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals.
 The diligent reader of these sage writings will also notice sentences that express sentiments like *based on their experience, statistical tests are not widely used in regression diagnostics.
-The same or even larger amount of information can be provided by diagnostic plots than the corresponding tests in most empirical studies.* There is a common guidance by experts that plots are the best for diagnosing model fits.
-
-This is curious, and investigating why this might be common advice is the subject of this paper.
+The same or even larger amount of information can be provided by diagnostic plots than the corresponding tests in most empirical studies.* A common guidance by experts is that optimal method for diagnosing model fits is by plotting the data.
+
+The persistence of this advice to check the plots is curious, and investigating why this might be common advice is the subject of this paper.
 The paper is structured as follows.
-The next background section describes the the types of departures that one expects to detect, and describes a formal process for reading residual plots, called visual inference, that can avoid the concerns about subjectiveness of human readers.
-Section \ref{experimental-design} describes the experimental setup to enable a comparison between decision made by formal hypothesis testing, and how humans would read diagnostic plots.
+The next background section describes the types of departures that one expects to detect, and describes a formal process for reading residual plots, called visual inference, that can avoid the concerns about subjectiveness of human readers [ET: CITATION HERE?].
+Section \ref{experimental-design} describes the experimental setup to compare the decision made by formal hypothesis testing, and how humans would read diagnostic plots.
 The results are reported in Section \ref{results}.
@@
 
-```{r residual-plot-common-departures, fig.cap = "Example fitted vs residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
+```{r commondepartures, fig.cap = "Example fitted vs residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
 The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted plot."}
@@
 Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals.
-As shown in Figure \ref{fig:residual-plot-common-departures}, the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns.
+As shown in Figure \ref{fig:commondepartures}, the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns.
 There are various types of departures from a good residual plot.
-
-Non-linearity, heterskedasticity and non-normality are perhaps the three mostly commonly checked departures.
+Two most commonly checked departures - non-linearity, heterskedasticity and non-normality will be discussed in this paper.
+
 Non-linearity is a type of model misspecification caused by failing to include higher order terms of the regressors in the regression equation.
@@
 One can clearly observe the "S-shape" from the residual plot as the cubic term is not captured by the misspecified model.
+
 Heterskedasticity refers to the presence of nonconst error variance in a regression model.
@@
 The usual pattern of heterskedasticity on a residual plot is the inconsistent spread of the residuals at different x values.
-Visually, it sometimes results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure \ref{fig:residual-plot-common-departures}, or the "left-triangle" and "right-triganle" shape where the smallest variance occurs at the edges of the x-axis.
+Visually, it sometimes results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure \ref{fig:commondepartures}, or the "left-triangle" and "right-triganle" shape where the smallest variance occurs at the edges of the x-axis.
+
 Compared to non-linearity and heterskedasticity, non-normality is usually harder to be detected from a residual plot since scatter plot is not excel in revealing marginal distribution.
@@
 In the case that the normality assumption is violated, it is expected to observe data points that do not centralize around the horizontal zero line and unevenly distribute at both sides of the zero line.
-For example, given a skewed error distribution, fewer data points and more outliers are on one side of the zero line as shown at the bottom-right of Figure \ref{fig:residual-plot-common-departures}.
+For example, given a skewed error distribution, fewer data points and more outliers are on one side of the zero line as shown at the bottom-right of Figure \ref{fig:commondepartures}.
 
@@
 
-Example residual plots given in Figure \ref{fig:residual-plot-common-departures} are examined by the corresponding model structural test, Breusch-Pagan test and Shapiro--Wilk test as shown in Table \ref{tab:example-residual-plot-table}.
+Example residual plots given in Figure \ref{fig:commondepartures} are examined by the corresponding model structural test, Breusch-Pagan test and Shapiro--Wilk test as shown in Table \ref{tab:example-residual-plot-table}.
 In the example, both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null hypothesis for departures that they do not intend to examine.
@@
                booktabs = TRUE, 
-               caption  = 'Statistical significance testing for departures from good residuals for plots in Figure \\ref{fig:residual-plot-common-departures}.
+               caption  = 'Statistical significance testing for departures from good residuals for plots in Figure \\ref{fig:commondepartures}.
 Shown are the $p$-values calculated for the conventional Model structural, the Breusch-Pagan and the Shapiro–Wilk tests.
@@
 
-<!-- TODO:Clarify this is only for integer c_i -->
-
 ```{=tex}
 \begin{equation} \label{eq:pvalue-beta-binomial}
-P(C \geq c_i) = \sum_{x=c_i}^{K}\frac{\Gamma(K + 1)}{\Gamma(x + 1)\Gamma(K - x + 1)}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},
+P(C \geq c_i) = \int_{x=c_i}^{K}\frac{\Gamma(K + 1)}{\Gamma(x + 1)\Gamma(K - x + 1)}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)}dx,
 \end{equation}
@@
 ```
-
-<!-- TODO:Introduce the concept of c interesting plot -->
 
@@
 Previously, it was addressed by modelling the probability of a subject $i$ correctly picking the actual data plot from a lineup $l$ using a mixed-effect logistic regression with the subject being treated as a random effect [@majumder_validation_2013].
-However, having this probability is insufficient to determine the power of a visual test allowed for multiple selections as it doesn't provide information about the number of selections made by the subject for p-value calculation.
-Instead, we directly estimated the probability of a lineup being rejected using a logistic regression with the natural logarithm of the effect size as the only regressor.
+However, knowing this probability is insufficient to determine the power of a visual test allowed for multiple selections as it doesn't provide information about the number of selections made by the subject, and hence, the p-value can't be computed.
+
+(1) Instead, we modelled the probability of a lineup being rejected directly using a logistic regression with the natural logarithm of the effect size as the only regressor.
 The same model was applied to different subsets of the survey data for making comparison of the power.
+
+(2) Instead, we modelled the plot selection $c_i$ for the actual data plot $i$.
+
+<!-- -   Very briefly explain what a lineup test is -->
+
+<!-- -   Lineup (maybe using your previous residual plot for nonlinearity) -->
+
+<!-- -   $p$-value calulcation, briefly, including how multiple selections are handled -->
+
+<!-- -   power calculation, briefly, including how multiple selections are handled -->
+
+```{=html}
+<!-- ![Parallelism between multiple quantitative testing and visual discovery [@buja_statistical_2009].
+Visible features in a plot are viewed as a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$, and any visual discoveries that are inconsistent with the null hypothesis are treated as evidence against the null.
+For regression diagnostics, the null hypothesis would be the assumed model, and visual discoveries could be any visual features in favour of any alternatives.
+\label{fig:parallelism}](figures/rsta2009012001.jpg){width=450 height=341}
+-->
+```
 
@@
 
-## Experiment overview
+## Overview
 
@@
 
-Three experiments were conducted to investigate the difference between conventional hypothesis testing and visual inference in the application of linear regression diagnostics.
+Three experiments were conducted to investigate the difference between conventional hypothesis testing and graphical diagnostics in the application of linear regression diagnostics.
 The experiment I has ideal scenario for conventional testing, where the visual test is not expected to outperform the conventional test.
-Meanwhile, the experiment II is a scenario where the conventional test is an approximate test, in which the visual test may have a chance to match the performance of the conventional test.
+The experiment II is a scenario where the conventional test is an approximate test, in which the visual test may have a chance to match the performance of the conventional test.
 The experiment III is designed for collecting human responses to lineup with only good residual plots such that the parameter $\alpha$ in Equation \ref{eq:pvalue-beta-binomial} can be estimated.
@@
 
-Two types of departures, namely non-linearity and heteroskedasticity, were considered with the corresponding data generating process being designed for experiment I and II.
+Two types of departures were considered with the corresponding data generating process being designed for experiment I and II.
 
@@
 
-```{r results='asis'}
+```{r results='asis', eval = FALSE}
 data.frame(j = c(2, 3, 6, 18),
@@
                caption  = 'Parameter values for $j$ and $X_{raw}$', 
-               label = "model-parameter-table" ) %>%
+               label = "parameter-table" ) %>%
   kableExtra::add_header_above(c("Order of Hermite polynomial\n(sym1)", "Distribution of sym4\n")) %>%
@@
 
-The values of $j$ was chosen so that distinct shapes of non-linearity were included in the residual plot.
-A greater value of $j$ will result in a curve with more turning points.
-As shown in Figure \ref{fig:different-shape-of-herimite}, it includes "U" shape, "S" shape, "M" shape and "Triple-U" shape.
-It is expected that the "U" shape will be the easiest one to detect because complex shape tends to be concealed by cluster of data points.
+The values of $j$ was chosen so that different shapes of non-linearity were included in the residual plot.
+As shown in Figure \ref{fig:different-shape-of-herimite}, these include "U" shape, "S" shape, "M" shape and "Triple-U" shape.
 
@@
 
+<!-- The range of $\sigma$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
+The example residual plots are given in Figure \ref{fig:different-sigma}.
+-->
+
+```{r different-sigma, eval = FALSE, fig.cap="Example residual plots of the linear regression model used in experiment I.
+Four different values of $\\sigma$ are included in the experiment to control the strength of the signal."}
+ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)
+
+VI_MODEL$plot(ori_dat) + 
+  ggtitle(quote(sigma~"="~0.5)) + 
+  theme_light() -> p1
+
+VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1)$gen(300, computed = select(ori_dat, x))) +
+  ggtitle(quote(sigma~"="~1)) + 
+  theme_light() -> p2
+
+VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 2)$gen(300, computed = select(ori_dat, x))) +
+  ggtitle(quote(sigma~"="~2)) + 
+  theme_light() -> p3
+
+VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 4)$gen(300, computed = select(ori_dat, x))) +
+  ggtitle(quote(sigma~"="~4)) + 
+  theme_light() -> p4
+
+patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)
+```
+
 Four different distribution were used to generate $X_{raw}$ as shown in Figure \ref{fig:different-dist}.
 The uniform and the normal distribution are symmetric and commonly assumed in statistical models.
-The adjusted log-normal distribution provides skewed density, while the discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.
+The adjusted log-normal distribution provides skewed density.
+And the discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.
 
@@
 
-Figure \ref{fig:example-poly-lineup} shows one of the lineups used in experiment I.
+Figure \ref{fig:example-lineup} shows one of the lineups used in experiment I.
 This lineup was produced under $j = 6$ and $X_{raw} \sim N(0.0.3^2)$.
@@
 
-```{r example-poly-lineup, fig.cap="Lineup poly-24 in experiment I. Can you spot the most different plot? \\label{fig:example-poly-lineup}"}
+```{r example-lineup, fig.cap="Lineup 24 in experiment I. Can you spot the most different plot? \\label{fig:example-lineup}"}
 VI_MODEL$plot_lineup(vi_lineup[["poly_24"]]$data, 
@@
 VI_MODEL$plot(ori_dat) +
-  ggtitle(quote("a = -1: Left-triangle shape")) +
+  ggtitle(quote("Left-triangle shape")) +
   theme_light() -> p1
@@
 VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote("a = 0: Butterfly shape")) +
+  ggtitle(quote("Butterfly shape")) +
   theme_light() -> p2
@@
 VI_MODEL$plot(heter_model(a = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote("a = 1: Right-triangle shape")) +
+  ggtitle(quote("Right-triangle shape")) +
   theme_light() -> p3
 
-patchwork::wrap_plots(p1, p2, p3, ncol = 2)
-```
-
-```{r example-heter-lineup, fig.cap="Lineup heter-169 in experiment II. Can you spot the most different plot? \\label{fig:example-heter-lineup}"}
-VI_MODEL$plot_lineup(vi_lineup[["heter_169"]]$data, 
-                     remove_grid_line = TRUE, 
-                     theme = theme_light(),
-                     remove_axis = TRUE)
-```
-
-An example lineup of this model is shown in Figure \ref{fig:example-heter-lineup} with $a = -1$ and $X_{raw} \sim U(-1, 1)$.
-The actual data plot location was 15.
-8 out of 11 subjects correctly identified the actual data plot for this lineup.
+patchwork::wrap_plots(p1, p2, p3, ncol = 3)
+```
 
@@
 
-### Controlling the strength of the signal
-
-As summarised by Table \ref{tab:model-parameter-table}, three additional parameters $n$, $\sigma$ and $b$ were used to control the strength of the signal so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
-Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ were used in experiment I and II respectively.
-Figure \ref{fig:different-sigma} and \ref{fig:different-b} demonstrate the impact of these two parameters.
-
-Three different sample sizes were used (n = 50, 100, 300) in all three experiments.
-It can be observed from Figure \ref{fig:different-n} that with fewer data points drawn in a residual plot, the visual pattern is more difficult to be detected.
-
-```{r different-n, fig.cap="Three different values of $n$ are used in experiment I, II and III to control the strength of the signal."}
-ori_dat <- poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(300)
-
-VI_MODEL$plot(ori_dat) + 
-  ggtitle(quote(n~"="~300)) + 
-  theme_light() -> p1
-
-VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:50,], x, e))) +
-  ggtitle(quote(n~"="~50)) + 
-  theme_light() -> p2
-
-VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:100,], x, e))) +
-  ggtitle(quote(n~"="~100)) + 
-  theme_light() -> p3
-
-patchwork::wrap_plots(p2, p3, p1, ncol = 2)
-```
-
-```{r different-sigma, fig.cap="Example residual plots of the linear regression model used in experiment I.
-Four different values of $\\sigma$ are included in the experiment to control the strength of the signal."}
-ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)
-
-VI_MODEL$plot(ori_dat) + 
-  ggtitle(quote(sigma~"="~0.5)) + 
-  theme_light() -> p1
-
-VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1)$gen(300, computed = select(ori_dat, x))) +
-  ggtitle(quote(sigma~"="~1)) + 
-  theme_light() -> p2
-
-VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 2)$gen(300, computed = select(ori_dat, x))) +
-  ggtitle(quote(sigma~"="~2)) + 
-  theme_light() -> p3
-
-VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 4)$gen(300, computed = select(ori_dat, x))) +
-  ggtitle(quote(sigma~"="~4)) + 
-  theme_light() -> p4
-
-patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)
-```
-
-```{r different-b, fig.cap="Example residual plots of the linear regression model used in experiment II.
-Five different values of $b$ are included in the experiment to control the strength of the signal."}
-stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
-
-ori_dat <- heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 0.25)$gen(300)
-
-VI_MODEL$plot(ori_dat) +
-  ggtitle(quote("b = 0.25")) +
-  theme_light() -> p1
-
-VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 1)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote("b = 1")) +
-  theme_light() -> p2
-
-VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 4)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote("b = 4")) +
-  theme_light() -> p3
-
-VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 16)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote("b = 16")) +
-  theme_light() -> p4
-
-VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 64)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote("b = 64")) +
-  theme_light() -> p5
-
-patchwork::wrap_plots(p1, p2, p3, p4, p5, ncol = 3)
-```
+### Treatments
+
+To control the strength of the signal, two parameters were used including $\sigma$ and $b$ for controlling the standard deviation of the error term in experiment I and II respectively.
+
+<!-- > Experiment I -->
+
+<!-- -   Treatments for nonlinearity experiment (plots summarising levels) -->
+
+<!-- It was generated for $n = 1$, $\sigma = 1$, $j = 1$, $X_{raw} \sim N$.
+For this lineup, x out of k observers picked the actual data plot.
+Each lineup is evaluated by 5 different subjects to provide reasonable estimates of the p-value.
+-->
+
+<!-- > Experiment II -->
+
+<!-- -   Treatments for heteroskedasticity experiment (plots summarising levels) -->
+
+<!-- Experiment data were simulated using three different shapes ($a$ = -1, 0, 1), five different values of $b$ (b = 0.25, 1, 4, 16, 64), three different sample sizes (n = 50, 100, 300) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
+A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}.
+-->
+
+<!-- The range of $b$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
+-->
 
@@
 
-Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in 443 different lineups.
+<!-- Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in 192 different lineups.
 For each lineup, the actual data plot was drawn as a standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis.
@@
 The lineup consisted of 20 residual plots with one randomly placed actual data plot.
-
-In addition, each lineup is designed to be evaluated by five different subjects to provide reasonable estimates of the visual p-value.
-
+Figure \ref{fig:example-lineup} is an example of one of these lineups.
+It was produced by using $n = 300$, $j = 6$, $\sigma = 0.5$ and $X_{raw} \sim N(0.0.3^2)$.
+The actual data plot location was four.
+All five subjects correctly identified the actual data plot for this lineup.
+-->
+
+<!-- In addition, each lineup is designed to be evaluated by five different subjects to provide reasonable estimates of the visual p-value.
 Thus, $192 \times 3 \times 5 / (20 - 2) = 160$ subjects were recruited to satisfy the design of the experiment I.
+-->
 

### 2022-10-04 paper_comparison/paper_comparison/paper_comparison.Rmd
section 2.1

@@
 Any non-linear functional form of residuals on fitted values presented in the residual plot could be considered as an indicative of non-linearity.
-An example residual plot containing visual pattern of non-linearity is given at the top-right of Figure [ref here].
+An example residual plot containing visual pattern of non-linearity is given at the top-right of Figure  \ref{fig:commondepartures}.
 One can clearly observe the "S-shape" from the residual plot as the cubic term is not captured by the misspecified model.
 
-Heterskedasticity refers to the presence of nonconst error variance in a regression model.
+Heteroskedasticity refers to the presence of nonconstant error variance in a regression model.
 It is mostly due to the strict but false assumptions on the variance-covariance matrix of the error term.
-The usual pattern of heterskedasticity on a residual plot is the inconsistent spread of the residuals at different x values.
-Visually, it sometimes results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure \ref{fig:commondepartures}, or the "left-triangle" and "right-triganle" shape where the smallest variance occurs at the edges of the x-axis.
-
-Compared to non-linearity and heterskedasticity, non-normality is usually harder to be detected from a residual plot since scatter plot is not excel in revealing marginal distribution.
-A favourable graphical summarise for this task is the quantile-quantile plot.
-However, for a consistent comparison in the later part of this paper, residual plot will be the focus.
+The usual pattern of heteroskedasticity on a residual plot is the inconsistent spread of the residuals across the horizontal axis.
+Visually, it sometimes results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure \ref{fig:commondepartures}, or the "left-triangle" and "right-triangle" shape where the smallest variance occurs at one side of the horizontal axis.
+
+Compared to non-linearity and heteroskedasticity, non-normality is usually harder to detect from a residual plot since scatter plot do not readily reveal the marginal distribution.
+A favourable graphical summary for this task is the quantile-quantile plot.
+However, for a consistent comparison, residual plot will be the focus of this paper.
 Besides, it is important to note that not all regression models assume normality for the error term, but a certain amount do including the classical normal linear regression model.
-In the case that the normality assumption is violated, it is expected to observe data points that do not centralize around the horizontal zero line and unevenly distribute at both sides of the zero line.
-For example, given a skewed error distribution, fewer data points and more outliers are on one side of the zero line as shown at the bottom-right of Figure \ref{fig:commondepartures}.
+In the case that the normality assumption is violated, it is expected to observe data points do not center around the horizontal axis and there is an uneven distribution of the number points at both below and above the horizontal axis.
+For example, given a skewed error distribution, fewer data points and more outliers are on one side of the horizontal axis as shown at the bottom-right of Figure \ref{fig:commondepartures}.
 

### 2022-10-04 paper_comparison/paper_comparison/paper_comparison.Rmd
section 2.2

@@
 It is likely the null hypothesis is correctly rejected but for the wrong reason, which is known as the "Type III error".
-Additionally, outliers will often incorrectly trigger the rejection of the null hypothesis despite the residuals are well-behaved [@cook_applied_1999].
+Additionally, outliers will often incorrectly trigger the rejection of the null hypothesis despite when majority of the residuals are well-behaved [@cook_applied_1999].
 This can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers.

### 2022-10-04 paper_comparison/paper_comparison/paper_comparison.Rmd

@@
 Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals.
-As shown in Figure \ref{fig:commondepartures}, the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns.
+As shown in Figure \ref{fig:residual-plot-common-departures}, the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns.
 There are various types of departures from a good residual plot.
-Two most commonly checked departures - non-linearity, heterskedasticity and non-normality will be discussed in this paper.
-
+
+Non-linearity, heterskedasticity and non-normality are perhaps the three mostly checked departures.
 Non-linearity is a type of model misspecification caused by failing to include higher order terms of the regressors in the regression equation.
@@
 
+<!-- TODO:Clarify this is only for integer c_i -->
+
 ```{=tex}
 \begin{equation} \label{eq:pvalue-beta-binomial}
-P(C \geq c_i) = \int_{x=c_i}^{K}\frac{\Gamma(K + 1)}{\Gamma(x + 1)\Gamma(K - x + 1)}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)}dx,
+P(C \geq c_i) = \sum_{x=c_i}^{K}\frac{\Gamma(K + 1)}{\Gamma(x + 1)\Gamma(K - x + 1)}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},
 \end{equation}
@@
 Previously, it was addressed by modelling the probability of a subject $i$ correctly picking the actual data plot from a lineup $l$ using a mixed-effect logistic regression with the subject being treated as a random effect [@majumder_validation_2013].
-However, knowing this probability is insufficient to determine the power of a visual test allowed for multiple selections as it doesn't provide information about the number of selections made by the subject, and hence, the p-value can't be computed.
-
-(1) Instead, we modelled the probability of a lineup being rejected directly using a logistic regression with the natural logarithm of the effect size as the only regressor.
-The same model was applied to different subsets of the survey data for making comparison of the power.
-
-(2) Instead, we modelled the plot selection $c_i$ for the actual data plot $i$.
-
-<!-- -   Very briefly explain what a lineup test is -->
-
-<!-- -   Lineup (maybe using your previous residual plot for nonlinearity) -->
-
-<!-- -   $p$-value calulcation, briefly, including how multiple selections are handled -->
-
-<!-- -   power calculation, briefly, including how multiple selections are handled -->
-
-```{=html}
-<!-- ![Parallelism between multiple quantitative testing and visual discovery [@buja_statistical_2009].
-Visible features in a plot are viewed as a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$, and any visual discoveries that are inconsistent with the null hypothesis are treated as evidence against the null.
-For regression diagnostics, the null hypothesis would be the assumed model, and visual discoveries could be any visual features in favour of any alternatives.
-\label{fig:parallelism}](figures/rsta2009012001.jpg){width=450 height=341}
--->
-```
+However, having this probability is insufficient to determine the power of a visual test allowed for multiple selections as it doesn't provide information about the number of selections made by the subject for p-value calculation.
+
+Instead, we directly estimated the probability of a lineup being rejected using a logistic regression with the natural logarithm of the effect size as the only regressor formulated as:
+
+```{=tex}
+\begin{equation} \label{eq:logistic-regression-1-1}
+Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(\boldsymbol{E})),
+\end{equation}
+```
+where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
+
+Effect $E$ is derived from the Kullback-Leibler divergence (see [appendix ref here]) formulated as:
+
+```{=tex}
+\begin{equation} \label{eq:effect-size-ex1}
+E = \frac{1}{2\sigma^2}\boldsymbol{X}_b'\boldsymbol{R}_a'(diag(\boldsymbol{R}_a))^{-1}\boldsymbol{R}_a\boldsymbol{X}_b,
+\end{equation}
+```
+where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$.
+
+To study various factors contributing to the power of the visual test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors.
+This includes [expansion].
 
@@
 ```
+
+<!-- ```{r results='asis'} -->
+<!-- data.frame(n = c(50, 100, 300, ""), -->
+<!--            j = c(2, 3, 6, 18), -->
+<!--            sigma = c(0.5, 1, 2, 4), -->
+<!--            x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$")) %>% -->
+<!--   knitr::kable(col.names = NULL, -->
+<!--                escape = FALSE, -->
+<!--                format = "latex", -->
+<!--                booktabs = TRUE,  -->
+<!--                caption  = 'Parameter values for $n$, $j$ $\\sigma$, $X_{raw}$',  -->
+<!--                label = "parameter-table" ) %>% -->
+<!--   kableExtra::add_header_above(c("Sample size\n(sym3)", "Order of Hermite polynomial\n(sym1)", "Error SD\n(sym2)", "Distribution of sym4\n")) %>% -->
+<!--   as.character() %>% -->
+<!--   gsub("sym1", "$j$", .) %>% -->
+<!--   gsub("sym2", "$\\\\sigma$", .) %>% -->
+<!--   gsub("sym3", "$n$", .) %>% -->
+<!--   gsub("sym4", "$X_{raw}$", .) %>% -->
+<!--   cat() -->
+<!--   # gsub("\\\\\\$", "$", .) %>% -->
+<!--   # cat() -->
+<!-- ``` -->
 
@@
 
-### Treatments
-
-To control the strength of the signal, two parameters were used including $\sigma$ and $b$ for controlling the standard deviation of the error term in experiment I and II respectively.
-
-<!-- > Experiment I -->
-
-<!-- -   Treatments for nonlinearity experiment (plots summarising levels) -->
-
-<!-- It was generated for $n = 1$, $\sigma = 1$, $j = 1$, $X_{raw} \sim N$.
-For this lineup, x out of k observers picked the actual data plot.
-Each lineup is evaluated by 5 different subjects to provide reasonable estimates of the p-value.
--->
-
-<!-- > Experiment II -->
-
-<!-- -   Treatments for heteroskedasticity experiment (plots summarising levels) -->
-
-<!-- Experiment data were simulated using three different shapes ($a$ = -1, 0, 1), five different values of $b$ (b = 0.25, 1, 4, 16, 64), three different sample sizes (n = 50, 100, 300) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
-A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}.
--->
-
-<!-- The range of $b$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
--->
-
-### Assigning subjects
-
-<!-- Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in 192 different lineups.
-For each lineup, the actual data plot was drawn as a standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis.
-The 19 null datasets were generated by the residual rotation technique, and plotted in the same way.
-The lineup consisted of 20 residual plots with one randomly placed actual data plot.
-Figure \ref{fig:example-lineup} is an example of one of these lineups.
-It was produced by using $n = 300$, $j = 6$, $\sigma = 0.5$ and $X_{raw} \sim N(0.0.3^2)$.
-The actual data plot location was four.
-All five subjects correctly identified the actual data plot for this lineup.
--->
-
-<!-- In addition, each lineup is designed to be evaluated by five different subjects to provide reasonable estimates of the visual p-value.
-Thus, $192 \times 3 \times 5 / (20 - 2) = 160$ subjects were recruited to satisfy the design of the experiment I.
--->
+### Controlling the strength of the signal
+
+As summarised by Table \ref{tab:model-parameter-table}, three additional parameters $n$, $\sigma$ and $b$ were used to control the strength of the signal so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
+Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ were used in experiment I and II respectively.
+Figure \ref{fig:different-sigma} and \ref{fig:different-b} demonstrate the impact of these two parameters.
+
+Three different sample sizes were used (n = 50, 100, 300) in all three experiments.
+It can be observed from Figure \ref{fig:different-n} that with fewer data points drawn in a residual plot, the visual pattern is more difficult to be detected.
+
+```{r different-n, fig.cap="Three different values of $n$ are used in experiment I, II and III to control the strength of the signal."}
+ori_dat <- poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(300)
+
+VI_MODEL$plot(ori_dat) + 
+  ggtitle(quote(n~"="~300)) + 
+  theme_light() -> p1
+
+VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:50,], x, e))) +
+  ggtitle(quote(n~"="~50)) + 
+  theme_light() -> p2
+
+VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:100,], x, e))) +
+  ggtitle(quote(n~"="~100)) + 
+  theme_light() -> p3
+
+patchwork::wrap_plots(p2, p3, p1, ncol = 2)
+```
+
+```{r different-sigma, fig.cap="Example residual plots of the linear regression model used in experiment I.
+Four different values of $\\sigma$ are included in the experiment to control the strength of the signal."}
+ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)
+
+VI_MODEL$plot(ori_dat) + 
+  ggtitle(quote(sigma~"="~0.5)) + 
+  theme_light() -> p1
+
+VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1)$gen(300, computed = select(ori_dat, x))) +
+  ggtitle(quote(sigma~"="~1)) + 
+  theme_light() -> p2
+
+VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 2)$gen(300, computed = select(ori_dat, x))) +
+  ggtitle(quote(sigma~"="~2)) + 
+  theme_light() -> p3
+
+VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 4)$gen(300, computed = select(ori_dat, x))) +
+  ggtitle(quote(sigma~"="~4)) + 
+  theme_light() -> p4
+
+patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)
+```
+
+```{r different-b, fig.cap="Example residual plots of the linear regression model used in experiment II.
+Five different values of $b$ are included in the experiment to control the strength of the signal."}
+stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
+
+ori_dat <- heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 0.25)$gen(300)
+
+VI_MODEL$plot(ori_dat) +
+  ggtitle(quote("b = 0.25")) +
+  theme_light() -> p1
+
+VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 1)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote("b = 1")) +
+  theme_light() -> p2
+
+VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 4)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote("b = 4")) +
+  theme_light() -> p3
+
+VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 16)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote("b = 16")) +
+  theme_light() -> p4
+
+VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 64)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote("b = 64")) +
+  theme_light() -> p5
+
+patchwork::wrap_plots(p1, p2, p3, p4, p5, ncol = 3)
+```
+
+### Subject allocation
+
+Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in 1152 different lineups.
+In addition, each lineup is designed to be evaluated by five different subjects and lineups with uniform distribution is designed to be evaluated by 11 subjects to provide reasonable estimates of the visual p-value.
+
+Thus, $443$ subjects were recruited to satisfy the design of the experiment. 
+<!-- TODO:subject allocation algorithm -->
 
@@
 
-software/technical
+<!-- TODO:pipeline description -->
+
+# Results
+
+## Data overview
+
+<!-- How many people? How many lineups? -->
+
+Subjects recruited from Prolific received a fixed payment for participating in the experiment.
+However, some subjects will try to maximize their earnings for minimum effort.
+During the review of submissions, if we found a subject objectively demonstrated clear low-effort throughout the experiment, i.e., failed all attention checks, we rejected the submission.
+The rejected submissions will be removed immediately, and Prolific will automatically recruit another subject as substitution.
+Therefore, we only paid for approved submissions and no further data screening procedure needed to be applied on the collected data.
+
+In overall, there were a total of 3200 lineup evaluations made by 160 subjects in both experiment I and experiment II respectively, where 320 lineup evaluations were attention checks and were not used in the following analysis.
+
+The collated dataset is provided in `vi_survey` of the `visage` `R` package.
 
 ## Power comparison
+
+Figure \ref{fig:power-poly-uniform} shows the estimated power of visual test against natural logarithm of the effect with comparison to the power of an exact test - F-test, and the power of two other residual-based conventional tests commonly used in regression diagnostics but for testing other departures from the model assumptions.
+In overall, the power of all four tests increases as the effect becomes larger.
+The power curve of F-test climbs aggressively from 25% to around 90% as $log_e(E)$ increases from 0 to 2, while others respond inactively to the change of effect and remain lower than 25% throughout the period, showing that as an exact test, the F-test is relatively more sensitive to the type of model defects that being considered.
+The power of visual test arises steadily and nearly linearly to around 90% as $log_e(E)$ increases from 2 to 5, suggesting that the effect starts to make noticeable impact on the degree of the presence of the designed visual features.
+Other two inappropriate conventional tests shows improvement at the same time but at a lower rate.
+This coincides the point made by @cook1982residuals mentioned in \ref{hypothesis-testing} that residual-based tests for a specific type of model defect are sensitive to other types of model defects.
+At $log_e(E) = 6$, the power curve of F-test reaches almost 100% followed by the visual test by a small margin.
+The power of Breusch--Pagan test and Shapiro--Wilk test reach around 75% and 30% respectively.
+
+What truly impress us is the huge difference between the estimated power of visual test and the estimated power of F-test.
+The margin is largest at around $log_e(E) = 2$.
+An example lineup is included in Figure \ref{fig:power-overview} where none of subjects detect the actual data plot positioned at panel 14.
+It demonstrates that at this level of difficulty, the designed visual feature is rarely visible, making the actual data plot indistinguishable from residual plots simulated from the assumed model.
+From a communication perspective, given the fact that the visual difference is unperceivable, the argument that non-linearity present in the fitted model is less convincing to the public even though it is true.
+At around $log_e(E) = 3$, the margin gets smaller as the chance of identifying the actual data plot becomes larger.
+At this level of difficulty, the designed visual features are usually detectable but it may not stand out from the lineup as other null plots may happen to include outliers or visual patterns that are are considered to be more attractive by human, and thus recognized as the most different plot.
+Without knowing the designed visual features beforehand, it is actually hard to identify the actual data plot by pure image comparison.
+The corresponding example lineup for $log_e(E) = 3$ shown in Figure \ref{fig:fig:power-poly-uniform} has the actual data plot positioned at panel 20, where two out of five subjects detect it.
+It can be observed that a M-shape is presented in plot 20, but the signal is not strong enough to attract all five subjects, resulting in a visual p-value sightly above the desired significance level $\alpha = 0.05$.
+At $log_e(E) = 4$ and $log_e(E) = 6$, the designed visual features become much clear and attractive, leading to a high percentage of rejection of the null hypothesis.
+Figure \ref{fig:fig:power-poly-uniform} gives example lineups of such cases.
+It is also interesting to observe that the power curve of visual test behaves similarly to those inappropriate conventional tests until the designed visual features really becomes visible.
 
@@
 
-```{r}
+```{r fig:power-poly-uniform, fig.cap = "..."}
 vi_survey_processed %>%
+  filter(x_dist == "uniform") %>%
   group_by(unique_lineup_id) %>%
   summarise(across(everything(), first)) %>%
-  rename(p_value_dirichlet = p_value_select_all_plots) %>%
+  filter(type == "polynomial") %>%
+  filter(e_sigma != 0.125) %>%
+  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
   ggplot() +
-  geom_point(aes(p_value_uniform, p_value_dirichlet)) +
-  geom_abline() +
-  scale_x_log10() +
-  scale_y_log10()
-```
-
-```{r}
+  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) + 
+  stat_smooth(geom = "line", 
+              data = ~boot_y(.x, reject_dirichlet, times = 500), 
+              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, 
+              alpha = 0.03) +
+  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
+                filter(x_dist == "uniform") %>%
+                pivot_longer(F_p_value:SW_p_value), 
+              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
+              method = "glm", 
+              method.args = list(family = binomial),
+              se = FALSE) +
+  theme_light(base_size = 5) -> p
+
+VI_MODEL$plot_lineup(vi_lineup[["poly_39"]]$data, 
+                     remove_grid_line = TRUE, 
+                     theme = theme_light(base_size = 5),
+                     remove_axis = TRUE)
+
+VI_MODEL$plot_lineup(vi_lineup[["poly_415"]]$data, 
+                     remove_grid_line = TRUE, 
+                     theme = theme_light(base_size = 5),
+                     remove_axis = TRUE) -> p1
+
+VI_MODEL$plot_lineup(vi_lineup[["poly_131"]]$data, 
+                     remove_grid_line = TRUE, 
+                     theme = theme_light(base_size = 5),
+                     remove_axis = TRUE) -> p2
+
+VI_MODEL$plot_lineup(vi_lineup[["poly_322"]]$data, 
+                     remove_grid_line = TRUE, 
+                     theme = theme_light(base_size = 5),
+                     remove_axis = TRUE) -> p3
+
+VI_MODEL$plot_lineup(vi_lineup[["poly_378"]]$data, 
+                     remove_grid_line = TRUE, 
+                     theme = theme_light(base_size = 5),
+                     remove_axis = TRUE) -> p4
+
+design <- c(patchwork::area(1, 1, 1, 8),
+            patchwork::area(2, 1, 2, 4), 
+            patchwork::area(2, 5, 2, 8), 
+            patchwork::area(3, 1, 3, 4), 
+            patchwork::area(3, 5, 3, 8))
+
+patchwork::wrap_plots(p, B = p1, C = p2, D = p3, E = p4, design = design)
+```
+
+## Distribution of regressor
+
+The impact of the distribution of $X_raw$ on the power is shown in Figure \ref{fig:dist-power}.
+The power curve of F-test is stable across different distributions, while the visual test has a steeper power curve for normal and uniform distribution.
+BP-test performs worse for discrete uniform distribution and uniform distribution but has relatively high power for normal distribution.
+SW-test outperforms BP-test for discrete uniform distribution but remains as the worst test for other distributions.
+The results indicate those inappropriate residual-based tests are sensitive to the distribution of the regressor.
+
+```{r dist-power, fig.cap="..."}
 vi_survey_processed %>%
@@
   facet_wrap(~x_dist) +
-  ggtitle("poly")
-```
-
-```{r}
-vi_survey_processed %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type == "polynomial") %>%
-  filter(e_sigma != 0.125) %>%
-  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
-  mutate(reject_uniform = as.numeric(p_value_uniform <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value_dirichlet"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_dirichlet, times = 500), 
-              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value_dirichlet"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03) +
-  geom_smooth(aes(log_effect_size, reject_uniform, col = "VI_p_value_uniform"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_uniform, times = 500), 
-              aes(log_effect_size, reject_uniform, group = boot_id, col = "VI_p_value_uniform"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03) +
-  facet_wrap(~x_dist) +
-  ggtitle("poly")
+  theme_light(base_size = 5)
 ```
@@
 
-```{r}
-vi_survey_processed %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type == "heteroskedasticity") %>%
-  filter(e_sigma != 0.125) %>%
-  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
-  mutate(reject_uniform = as.numeric(p_value_uniform <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value_dirichlet"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_dirichlet, times = 500), 
-              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value_dirichlet"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03) +
-  geom_smooth(aes(log_effect_size, reject_uniform, col = "VI_p_value_uniform"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_uniform, times = 500), 
-              aes(log_effect_size, reject_uniform, group = boot_id, col = "VI_p_value_uniform"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03) +
-  facet_wrap(~x_dist) +
-  ggtitle("heter")
-```
-
-```{r}
-p_value_cache_env <- new.env()
-
-if (file.exists(here::here("data/p_value_only_5_eval.rds"))) {
-  p_value_only_5_eval <- readRDS(here::here("data/p_value_only_5_eval.rds"))
-} else {
-  p_value_only_5_eval <- vi_survey_processed %>%
-  filter(x_dist == "uniform") %>%
-  filter((is.na(b) & e_sigma != 0.125 ) | (b != 0 & b < 128)) %>%
-  filter(exp != 5) %>%
-  group_by(unique_lineup_id) %>%
-  summarise(p_value_only_5_eval = calc_p_value(n_detect = sum(detect), 
-                                               n_eval = n(), 
-                                               n_sel = num_selection, 
-                                               dist = "dirichlet",
-                                               alpha = numeric_alpha[1],
-                                               cache_env = p_value_cache_env))
-
-  saveRDS(p_value_only_5_eval, here::here("data/p_value_only_5_eval.rds"))
-}
-
-vi_survey_processed %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type == "heteroskedasticity") %>%
-  filter(x_dist == "uniform") %>%
-  filter(b != 0) %>%
-  filter(b < 128) %>%
-  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  left_join(p_value_only_5_eval) %>%
-  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_dirichlet, times = 500), 
-              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03) +
-  geom_smooth(aes(log_effect_size, reject_only_5_eval, col = "VI_p_value_only_5_eval"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_only_5_eval, times = 500), 
-              aes(log_effect_size, reject_only_5_eval, group = boot_id, col = "VI_p_value_only_5_eval"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03) +
-  ggtitle("heter")
-
-p1 <- vi_survey_processed %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type == "heteroskedasticity") %>%
-  filter(x_dist == "uniform") %>%
-  filter(b != 0) %>%
-  filter(b < 128) %>%
-  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  left_join(p_value_only_5_eval) %>%
-  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_dirichlet, times = 500), 
-              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03)
-
-p2 <- vi_survey_processed %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type == "heteroskedasticity") %>%
-  filter(x_dist == "uniform") %>%
-  filter(b != 0) %>%
-  filter(b < 128) %>%
-  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  left_join(p_value_only_5_eval) %>%
-  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject_only_5_eval, col = "VI_p_value_only_5_eval"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_only_5_eval, times = 500), 
-              aes(log_effect_size, reject_only_5_eval, group = boot_id, col = "VI_p_value_only_5_eval"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03)
-
-patchwork::wrap_plots(p1, p2)
-```
-
-```{r}
-
-vi_survey_processed %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type == "polynomial") %>%
-  filter(x_dist == "uniform") %>%
-  filter(e_sigma != 0.125) %>%
-  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  left_join(p_value_only_5_eval) %>%
-  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_dirichlet, times = 500), 
-              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03) +
-  geom_smooth(aes(log_effect_size, reject_only_5_eval, col = "VI_p_value_only_5_eval"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_only_5_eval, times = 500), 
-              aes(log_effect_size, reject_only_5_eval, group = boot_id, col = "VI_p_value_only_5_eval"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03) +
-  ggtitle("poly")
-
-p1 <- vi_survey_processed %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type == "polynomial") %>%
-  filter(x_dist == "uniform") %>%
-  filter(e_sigma != 0.125) %>%
-  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  left_join(p_value_only_5_eval) %>%
-  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_dirichlet, times = 500), 
-              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03)
-
-p2 <- vi_survey_processed %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type == "polynomial") %>%
-  filter(x_dist == "uniform") %>%
-  filter(e_sigma != 0.125) %>%
-  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  left_join(p_value_only_5_eval) %>%
-  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject_only_5_eval, col = "VI_p_value_only_5_eval"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_only_5_eval, times = 500), 
-              aes(log_effect_size, reject_only_5_eval, group = boot_id, col = "VI_p_value_only_5_eval"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03)
-
-patchwork::wrap_plots(p1, p2)
-
-```
-
-```{r eval = FALSE}
-vi_survey_processed %>%
-  filter(exp == 3) %>%
-  filter(e_sigma != 0.125) %>%
-  mutate()
-
-vi_survey_processed %>%
-  filter(exp == 3) %>%
-  filter(e_sigma != 0.125) %>%
-  mutate(set = factor(set), lineup_id = factor(lineup_id)) %>%
-  {lme4::glmer(weighted_detect ~ effect_size + ( 0 + effect_size | set) + ( 0 + effect_size | lineup_id), family = binomial ,data = .)} %>%
-  predict(type = "response")
-
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE)
-
-+
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_dirichlet, times = 500), 
-              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03) +
-  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
-                pivot_longer(F_p_value:SW_p_value), 
-              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
-              method = "glm", 
-              method.args = list(family = binomial),
-              se = FALSE) +
-  facet_wrap(~x_dist) +
-  ggtitle("poly")
-
-  ?glm
-```
-
-<!-- # Old stuff -->
-
-<!-- ## Computing power curves -->
-
-<!-- For the purpose of examining the effectiveness of visual test in regression diagnostics, two experiments were conducted.
-The experiment I has ideal scenario for conventional testing, where the visual test is not expected to outperform the conventional test.
-The experiment II is a scenario where the conventional test is an approximate test, in which the visual test may have a chance to match the performance of the conventional test.
--->
-
-<!-- Subjects for both experiments were recruited from an crowdsourcing platform called Prolific [ref here].
-Prescreening procedure was applied during the recruitment, subjects were required to be fluent in English, with $98\%$ minimum approval rate in other studies and 10 minimum submissions.
--->
-
-<!-- During the experiment, every subject was presented with a block of 20 lineups.
-And for every lineup, the subject was asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots were from others.
-If there was no noticeable difference between plots in a lineup, subjects were permitted to select zero plots without providing the reason.
-No subject was shown the same lineup twice.
-Information about preferred pronoun, age group, education, and previous experience in visual experiment were also collected.
--->
-
-<!-- A pool of 12 different lineups with obvious visual patterns were generated for experiment I and experiment II respectively.
-In every block of 20 lineups that presented to a subject, two out of 12 lineups were included as attention checks.
-A subject's submission was only accepted if the actual data plot was identified for at least one attention check.
-Data of rejected submissions were discarded automatically to maintain the overall data quality.
--->
-
-<!-- ## Non-linearity -->
-
-<!-- Experiment I is designed to study the ability of human subjects to detect the effect of a random vector $\boldsymbol{z}$ which is a probabilist's Hermite polynomial [Herimite ref here] of another random vector $\boldsymbol{x}$ in a two variable statistical model formulated as: -->
-
-<!-- ```{=tex} -->
-<!-- \begin{align} \label{eq:nonlinearity-model} -->
-<!-- \boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\ -->
-<!-- \boldsymbol{x} = g(\boldsymbol{x}_{raw}, 1), \\ -->
-<!-- \boldsymbol{z} = g(\boldsymbol{z}_{raw}, 1), \\ -->
-<!-- \boldsymbol{z}_{raw} = He_j(g(\boldsymbol{z}, 2)), -->
-<!-- \end{align} -->
-<!-- ``` -->
-<!-- where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vector of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $\{-k, k\}$ defined as -->
-
-<!-- ```{=tex} -->
-<!-- \begin{equation} \label{eq:scaling-function} -->
-<!-- g(\boldsymbol{x}, k) = (\boldsymbol{x} - min(\boldsymbol{x}))/max(\boldsymbol{x} - min(\boldsymbol{x})) \times 2k - k, \quad \text{for} \quad k > 0.  -->
-<!-- \end{equation} -->
-<!-- ``` -->
-<!-- The null regression model used to fit the realizations generated by the above model is formulated as: -->
-
-<!-- ```{=tex} -->
-<!-- \begin{equation} \label{eq:scaling-function} -->
-<!-- \boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u}, -->
-<!-- \end{equation} -->
-<!-- ``` -->
-<!-- where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$. -->
-
-<!-- Model misspecification presents since the null model leaves out the higher order term. -->
-
-<!-- Experiment data were simulated using four different order of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$), three different sample sizes (n = 50, 100, 300) , four different standard deviations of the error ($\sigma$ = 0.5, 1, 2, 4) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
-A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}.
--->
-
-<!-- The values of $j$ was chosen so that different shapes of non-linearity were included in the residual plot. These include "U" shape, "S" shape, "M" shape and "Triple-U" shape. -->
-
-<!-- ```{r different-shape-of-herimite} -->
-<!-- stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1 -->
-
-<!-- ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300) -->
-
-<!-- VI_MODEL$plot(ori_dat) +  -->
-<!--   ggtitle(quote(He[2]:"U shape")) +  -->
-<!--   theme_light() -> p1 -->
-
-<!-- VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) + -->
-<!--   ggtitle(quote(He[3]:"S shape")) +  -->
-<!--   theme_light() -> p2 -->
-
-<!-- VI_MODEL$plot(poly_model(shape = 3, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) + -->
-<!--   ggtitle(quote(He[6]:"M shape")) +  -->
-<!--   theme_light() -> p3 -->
-
-<!-- VI_MODEL$plot(poly_model(shape = 4, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) + -->
-<!--   ggtitle(quote(He[18]:"Triple-U shape")) +  -->
-<!--   theme_light() -> p4 -->
-
-<!-- patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2) -->
-<!-- ``` -->
-
-<!-- The range of $\sigma$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
--->
-
-<!-- ```{r different-sigma} -->
-<!-- ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300) -->
-
-<!-- VI_MODEL$plot(ori_dat) +  -->
-<!--   ggtitle(quote(sigma~"="~0.5)) +  -->
-<!--   theme_light() -> p1 -->
-
-<!-- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1)$gen(300, computed = select(ori_dat, x))) + -->
-<!--   ggtitle(quote(sigma~"="~1)) +  -->
-<!--   theme_light() -> p2 -->
-
-<!-- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 2)$gen(300, computed = select(ori_dat, x))) + -->
-<!--   ggtitle(quote(sigma~"="~2)) +  -->
-<!--   theme_light() -> p3 -->
-
-<!-- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 4)$gen(300, computed = select(ori_dat, x))) + -->
-<!--   ggtitle(quote(sigma~"="~4)) +  -->
-<!--   theme_light() -> p4 -->
-
-<!-- patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2) -->
-<!-- ``` -->
-
-<!-- Four different distribution were used to generate $X_{raw}$.
-The uniform and the normal distribution are symmetric and commonly assumed in statistical models.
-The adjusted log-normal distribution provides skewed density.
-And the discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.
--->
-
-<!-- ```{r different-dist} -->
-<!-- ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300) -->
-
-<!-- VI_MODEL$plot(ori_dat) +  -->
-<!--   ggtitle(quote(U(-1,1))) +  -->
-<!--   theme_light() -> p1 -->
-
-<!-- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) + -->
-<!--   ggtitle(quote(N(0,0.3^2))) +  -->
-<!--   theme_light() -> p2 -->
-
-<!-- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) + -->
-<!--   ggtitle(quote(lognormal(0,0.6^2)/3)) +  -->
-<!--   theme_light() -> p3 -->
-
-<!-- VI_MODEL$plot(poly_model(shape = 1, x = rand_uniform_d(k = 5, even = TRUE), sigma = 0.5)$gen(300, computed = select(ori_dat, e))) + -->
-<!--   ggtitle(quote(U~"{1,5}")) +  -->
-<!--   theme_light() -> p4 -->
-
-<!-- patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2) -->
-<!-- ``` -->
-
-<!-- Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in 192 different lineups.
-For each lineup, the actual data plot was drawn as a standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis.
-The 19 null datasets were generated by the residual rotation technique, and plotted in the same way.
-The lineup consisted of 20 residual plots with one randomly placed actual data plot.
-Figure \ref{fig:example-lineup} is an example of one of these lineups.
-It was produced by using $n = 300$, $j = 6$, $\sigma = 0.5$ and $X_{raw} \sim N(0.0.3^2)$.
-The actual data plot location was four.
-All five subjects correctly identified the actual data plot for this lineup.
--->
-
-<!-- ```{r results='asis'} -->
-<!-- data.frame(n = c(50, 100, 300, ""), -->
-<!--            j = c(2, 3, 6, 18), -->
-<!--            sigma = c(0.5, 1, 2, 4), -->
-<!--            x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$")) %>% -->
-<!--   knitr::kable(col.names = NULL, -->
-<!--                escape = FALSE, -->
-<!--                format = "latex", -->
-<!--                booktabs = TRUE,  -->
-<!--                caption  = 'Parameter values for $n$, $j$ $\\sigma$, $X_{raw}$',  -->
-<!--                label = "parameter-table" ) %>% -->
-<!--   kableExtra::add_header_above(c("Sample size\n(sym3)", "Order of Hermite polynomial\n(sym1)", "Error SD\n(sym2)", "Distribution of sym4\n")) %>% -->
-<!--   as.character() %>% -->
-<!--   gsub("sym1", "$j$", .) %>% -->
-<!--   gsub("sym2", "$\\\\sigma$", .) %>% -->
-<!--   gsub("sym3", "$n$", .) %>% -->
-<!--   gsub("sym4", "$X_{raw}$", .) %>% -->
-<!--   cat() -->
-<!--   # gsub("\\\\\\$", "$", .) %>% -->
-<!--   # cat() -->
-<!-- ``` -->
-
-<!-- ```{r example-lineup, fig.cap="Example lineup \\label{fig:example-lineup}"} -->
-<!-- VI_MODEL$plot_lineup(polynomials_lineup[[23]]$data,  -->
-<!--                      remove_grid_line = TRUE,  -->
-<!--                      theme = theme_light(), -->
-<!--                      remove_axis = TRUE) -->
-<!-- ``` -->
-
-<!-- <!-- It was generated for $n = 1$, $\sigma = 1$, $j = 1$, $X_{raw} \sim N$.
-For this lineup, x out of k observers picked the actual data plot.
-Each lineup is evaluated by 5 different subjects to provide reasonable estimates of the p-value.
---> -->
-
-<!-- In addition, each lineup is designed to be evaluated by five different subjects to provide reasonable estimates of the visual p-value.
-Thus, $192 \times 3 \times 5 / (20 - 2) = 160$ subjects were recruited to satisfy the design of the experiment I.
--->
-
-<!-- ## Heteroskedasticity -->
-
-<!-- Experiment II is designed to study the ability of human subjects to detect the appearance of a heteroskedasticity pattern under a simple linear regression model setting: -->
-
-<!-- ```{=tex} -->
-<!-- \begin{align} \label{eq:heter-model} -->
-<!-- \boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\ -->
-<!-- \boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + 2 - |a| b (\boldsymbol{x} - a)^2 \boldsymbol{I}), \\ -->
-<!-- \end{align} -->
-<!-- ``` -->
-<!-- where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vector of size $n$. -->
-
-<!-- The null regression model used to fit the realizations generated by the above model is formulated exactly the same as the model used in the first experiment: -->
-
-<!-- ```{=tex} -->
-<!-- \begin{equation}  -->
-<!-- \boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u}, -->
-<!-- \end{equation} -->
-<!-- ``` -->
-<!-- where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$. -->
-
-<!-- Model misspecification presents since the assumption about the constant error variance is violated. -->
-
-<!-- Experiment data were simulated using three different shapes ($a$ = -1, 0, 1), five different values of $b$ (b = 0.25, 1, 4, 16, 64), three different sample sizes (n = 50, 100, 300) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
-A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}.
--->
-
-<!-- The values of $a$ was chosen so that different shapes of heteroskedasticity were included in the residual plot. These include triangle shape, butterfly shape and inverse triangle shape. -->
-
-<!-- ```{r different-shape-of-heter} -->
-<!-- stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1 -->
-
-<!-- ori_dat <- heter_model(a = -1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300) -->
-
-<!-- VI_MODEL$plot(ori_dat) +  -->
-<!--   ggtitle(quote("Triangle shape")) +  -->
-<!--   theme_light() -> p1 -->
-
-<!-- VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) + -->
-<!--   ggtitle(quote("Butterfly shape")) +  -->
-<!--   theme_light() -> p2 -->
-
-<!-- VI_MODEL$plot(heter_model(a = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) + -->
-<!--   ggtitle(quote("Inverse triangle shape")) +  -->
-<!--   theme_light() -> p3 -->
-
-<!-- patchwork::wrap_plots(p1, p2, p3, ncol = 3) -->
-<!-- ``` -->
-
-<!-- The range of $b$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
--->
-
-<!-- # Analysis and results -->
-
-<!-- ## Data overview -->
-
-<!-- How many people? How many lineups? -->
-
-<!-- Subjects recruited from Prolific received a fixed payment for participating in the experiment.
-However, some subjects will try to maximize their earnings for minimum effort.
-During the review of submissions, if we found a subject objectively demonstrated clear low-effort throughout the experiment, i.e., failed all attention checks, we rejected the submission.
-The rejected submissions will be removed immediately, and Prolific will automatically recruit another subject as substitution.
-Therefore, we only paid for approved submissions and no further data screening procedure needed to be applied on the collected data.
--->
-
-<!-- A subject was allowed to select zero plots for a lineup if there was no visible difference between plots, but the simulated visual p-value given in Equation (\ref{eq:p-value-multi}) will effectively drop the subject from the simulation for this case, leading to inaccurate estimation of $p$-value.
-Therefore, we treated this case as making one selection but failing to identify the actual data plot so that Equation (\ref{eq:p-value-multi}) can be applied correctly.
--->
-
-<!-- In overall, there were a total of 3200 lineup evaluations made by 160 subjects in both experiment I and experiment II respectively, where 320 lineup evaluations were attention checks and were not used in the following analysis.
--->
-
-<!-- The collated dataset is provided in `polynomials` and `heter` of the `visage` `R` package. -->
-
-<!-- ##  -->
-
-<!-- ## Nonlinearity -->
-
-<!-- -   Power curve overall -->
-<!-- -   Four lineups also shown, selected from ones close to the visual power curve from the uniform treatment -->
-<!-- -   Compare visual power against different factors -->
-
-<!-- ## Heteroskedasticity -->
-
-<!-- -   Power curve overall -->
-<!-- -   Four lineups also shown, selected from ones close to the visual power curve from the uniform treatment -->
-<!-- -   Compare visual power against different factors -->
-
-<!-- ## Demographic summary -->
-
-<!-- Table \ref{tab:demographic-table} tabulates the number of subjects, preferred pronouns, education backgrounds, age groups, and previous experience in visual experiment.
-Figure \ref{fig:demographic-summary-plot} visualizes the marginal distribution of each of the category.
-The collated data was a balanced sample among male and female.
-Most of the ages of subject were between 18 to 39, while many of them were between 18 to 24.
-Very few subjects were awarded degree higher than Bachelor Degree.
-Around 40% of subjects had previous experience in visual experiment.
--->
-
-<!-- ```{r results="asis"} -->
-
-<!-- polynomials %>% -->
-<!--   mutate(exp = 1) %>% -->
-<!--   bind_rows(mutate(heter, exp = 2)) %>% -->
-<!--   group_by(exp, set) %>% -->
-<!--   slice_sample(n = 1) %>% -->
-<!--   ungroup() %>% -->
-<!--   count(pronoun, education, age_group, previous_experience) %>% -->
-<!--   arrange(desc(n)) %>% -->
-<!--   knitr::kable(col.names = c("Preferred pronoun", "Education", "Age group", "Previous experience", "Subject"), -->
-<!--                escape = FALSE, -->
-<!--                format = "latex", -->
-<!--                linesep = "", -->
-<!--                booktabs = TRUE,  -->
-<!--                caption  = 'Summary of demographic information',  -->
-<!--                label = "demographic-table" ) %>% -->
-<!--   cat() -->
-<!-- ``` -->
-
-<!-- ```{r demographic-summary-plot, fig.cap="Summary of demographic information. \\label{fig:demographic-summary-plot}"} -->
-<!-- polynomials %>% -->
-<!--   mutate(exp = 1) %>% -->
-<!--   bind_rows(mutate(heter, exp = 2)) %>% -->
-<!--   group_by(exp, set) %>% -->
-<!--   summarise(across(age_group:previous_experience, first)) %>% -->
-<!--   mutate(education = fct_relevel(education, "High School or below")) %>% -->
-<!--   ggplot() + -->
-<!--   geom_bar(aes(y = education), orientation = "y") -> education_plot -->
-
-<!-- polynomials %>% -->
-<!--   mutate(exp = 1) %>% -->
-<!--   bind_rows(mutate(heter, exp = 2)) %>% -->
-<!--   group_by(exp, set) %>% -->
-<!--   summarise(across(age_group:previous_experience, first)) %>% -->
-<!--   ggplot() + -->
-<!--   geom_bar(aes(age_group)) -> age_plot -->
-
-<!-- polynomials %>% -->
-<!--   mutate(exp = 1) %>% -->
-<!--   bind_rows(mutate(heter, exp = 2)) %>% -->
-<!--   group_by(exp, set) %>% -->
-<!--   summarise(across(age_group:previous_experience, first)) %>% -->
-<!--   mutate(pronoun = fct_relevel(pronoun, "He", "She", "They")) %>% -->
-<!--   ggplot() + -->
-<!--   geom_bar(aes(pronoun)) -> gender_plot -->
-
-<!-- polynomials %>% -->
-<!--   mutate(exp = 1) %>% -->
-<!--   bind_rows(mutate(heter, exp = 2)) %>% -->
-<!--   group_by(exp, set) %>% -->
-<!--   summarise(across(age_group:previous_experience, first)) %>% -->
-<!--   ggplot() + -->
-<!--   geom_bar(aes(previous_experience)) -> experience_plot -->
-
-<!-- patchwork::wrap_plots(education_plot, age_plot, gender_plot, experience_plot) -->
-<!-- ``` -->
-
-<!-- ## Model fitting -->
-
-<!-- For each parameter combination, effect $E$ is derived from the Kullback-Leibler divergence (see [appendix ref here]) formulated as: -->
-
-<!-- ```{=tex} -->
-<!-- \begin{equation} \label{eq:effect-size-ex1} -->
-<!-- E = \frac{1}{2\sigma^2}\boldsymbol{X}_b'\boldsymbol{R}_a'(diag(\boldsymbol{R}_a))^{-1}\boldsymbol{R}_a\boldsymbol{X}_b, -->
-<!-- \end{equation} -->
-<!-- ``` -->
-<!-- where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$. -->
-
-<!-- The logistic regression is fit using $log_e(E)$ as the only fixed effect covariate for the power of visual test formulated as: -->
-
-<!-- ```{=tex} -->
-<!-- \begin{equation} \label{eq:logistic-regression-1-1} -->
-<!-- Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(\boldsymbol{E})), -->
-<!-- \end{equation} -->
-<!-- ``` -->
-<!-- where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$. -->
-
-<!-- To study various factors contributing to the power of the visual test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors.
-This includes [expansion].
--->
-
-<!-- Table [table ref here] shows the parameter estimates of the logistic regressions. [Discussion about the numeric estimates here] -->
-
 <!-- ## Power comparison -->

### 2022-10-04 paper_comparison/paper_comparison/paper_comparison.Rmd
fix the first plot of analysis

@@
 Without knowing the designed visual features beforehand, it is actually hard to identify the actual data plot by pure image comparison.
-The corresponding example lineup for $log_e(E) = 3$ shown in Figure \ref{fig:fig:power-poly-uniform} has the actual data plot positioned at panel 20, where two out of five subjects detect it.
-It can be observed that a M-shape is presented in plot 20, but the signal is not strong enough to attract all five subjects, resulting in a visual p-value sightly above the desired significance level $\alpha = 0.05$.
-At $log_e(E) = 4$ and $log_e(E) = 6$, the designed visual features become much clear and attractive, leading to a high percentage of rejection of the null hypothesis.
+The corresponding example lineup for $log_e(E) = 3$ shown in Figure \ref{fig:fig:power-poly-uniform} has the actual data plot positioned at panel 17, where two out of five subjects detect it.
+It can be observed that a "U-shape" is presented in plot 17, but the signal is not strong enough to attract all five subjects, resulting in a visual p-value sightly above the desired significance level $\alpha = 0.05$.
+At $log_e(E) = 4$ and $log_e(E) = 5$, the designed visual features become much clear and attractive, leading to a high percentage of rejection of the null hypothesis.
 Figure \ref{fig:fig:power-poly-uniform} gives example lineups of such cases.
@@
   geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) + 
+  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) + geom_point(data = ~filter(.x, unique_lineup_id %in% c("poly_42", "poly_562", "poly_217", "poly_120")), aes(log_effect_size, prop_detect), alpha = 1, col = "red", size = 2) +
   stat_smooth(geom = "line", 
@@
 
-VI_MODEL$plot_lineup(vi_lineup[["poly_39"]]$data, 
+VI_MODEL$plot_lineup(vi_lineup[["poly_42"]]$data, 
                      remove_grid_line = TRUE, 
                      theme = theme_light(base_size = 5),
-                     remove_axis = TRUE)
-
-VI_MODEL$plot_lineup(vi_lineup[["poly_415"]]$data, 
+                     remove_axis = TRUE) + 
+  ggtitle(bquote(log[e](E) %~~% 2~","~"F-test p-value" == .(format(filter(vi_survey_processed, unique_lineup_id == "poly_42")$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(vi_survey_processed, unique_lineup_id == "poly_42")$p_value_select_all_plots[1], digits = 3)))) -> p1
+
+VI_MODEL$plot_lineup(vi_lineup[["poly_562"]]$data, 
                      remove_grid_line = TRUE, 
                      theme = theme_light(base_size = 5),
-                     remove_axis = TRUE) -> p1
-
-VI_MODEL$plot_lineup(vi_lineup[["poly_131"]]$data, 
+                     remove_axis = TRUE) + 
+  ggtitle(bquote(log[e](E) %~~% 3~","~"F-test p-value" == .(format(filter(vi_survey_processed, unique_lineup_id == "poly_562")$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(vi_survey_processed, unique_lineup_id == "poly_562")$p_value_select_all_plots[1], digits = 3)))) -> p2
+
+VI_MODEL$plot_lineup(vi_lineup[["poly_217"]]$data, 
                      remove_grid_line = TRUE, 
                      theme = theme_light(base_size = 5),
-                     remove_axis = TRUE) -> p2
-
-VI_MODEL$plot_lineup(vi_lineup[["poly_322"]]$data, 
+                     remove_axis = TRUE) + 
+  ggtitle(bquote(log[e](E) %~~% 4~","~"F-test p-value" == .(format(filter(vi_survey_processed, unique_lineup_id == "poly_217")$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(vi_survey_processed, unique_lineup_id == "poly_217")$p_value_select_all_plots[1], digits = 3)))) -> p3
+
+VI_MODEL$plot_lineup(vi_lineup[["poly_120"]]$data, 
                      remove_grid_line = TRUE, 
                      theme = theme_light(base_size = 5),
-                     remove_axis = TRUE) -> p3
-
-VI_MODEL$plot_lineup(vi_lineup[["poly_378"]]$data, 
-                     remove_grid_line = TRUE, 
-                     theme = theme_light(base_size = 5),
-                     remove_axis = TRUE) -> p4
+                     remove_axis = TRUE) + 
+  ggtitle(bquote(log[e](E) %~~% 2~","~"F-test p-value" == .(format(filter(vi_survey_processed, unique_lineup_id == "poly_120")$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(vi_survey_processed, unique_lineup_id == "poly_120")$p_value_select_all_plots[1], digits = 3)))) -> p4
 
@@
 patchwork::wrap_plots(p, B = p1, C = p2, D = p3, E = p4, design = design)
+```
+
+```{r eval = FALSE}
+vi_survey_processed %>%
+  filter(x_dist == "uniform") %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(type == "polynomial") %>%
+  filter(e_sigma != 0.125) %>%
+  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+filter(abs(log(effect_size) - 3) < 0.2)
+
+vi_survey_processed %>%
+  filter(x_dist == "uniform") %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(type == "polynomial") %>%
+  filter(e_sigma != 0.125) %>%
+  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  ggplot() +
+  geom_abline() +
+  geom_point(aes(p_value_uniform, p_value_select_all_plots)) +
+  scale_x_log10() +
+  scale_y_log10()
 ```

### 2022-10-04 paper_comparison/paper_comparison/paper_comparison.tex
fix the first plot of analysis

@@
 Diagnosing a model is the key to determining whether there is anything
-importantly wrong. For linear regression analysis, it is typical to
-interrogate the residuals. Residuals summarise what is not captured by
-the model, and thus provide the capacity to identify what might be
-wrong. There are many ways that residuals could be assessed.
+importantly wrong. For linear regression analysis, we typically
+interrogate the residuals for model diagnostics. Residuals summarise
+what is not captured by the model, and thus provide the capacity to
+identify what might be wrong.
 
-Residuals might be plotted, as a histogram or quantile-quantile plot to
-examine the distribution. Using the classical normal linear regression
-model as an example, if the distribution is symmetric and unimodal, it
-is well-behaved. But if the distribution is skewed, bimodal, multimodal,
-or contains outliers, there is cause for concern. The distribution could
-also be inspected by conducting a goodness of fit test, such as the
-Shapiro-Wilk Normality test \citep{shapiro1965analysis}.
+We can assess residuals in multiple ways. Residuals might be plotted, as
+a histogram or quantile-quantile plot to examine the distribution. Using
+the classical normal linear regression model as an example, if the
+distribution is symmetric and unimodal, it is well-behaved. But if the
+distribution is skewed, bimodal, multimodal, or contains outliers, there
+is cause for concern. The distribution could also be inspected by
+conducting a goodness of fit test, such as the Shapiro-Wilk Normality
+test \citep{shapiro1965analysis}.
 
@@
 their relationships. If there is any visually discoverable patterns, the
-model is potentially misspecified. However, it is a very difficult task
-for a human judge, though to make a decision that there's nothing there.
-It is especially common, particularly among new data analysts to report
-patterns when an experienced data analyst might quickly conclude that
-there are none. Generally, one looks for departures from nothingness
+model is potentially misspecified. However, correctly judging a residual
+plot where no pattern exists is a painstakingly difficult task for
+humans. It is especially common, particularly among new data analysts to
+report patterns when an experienced data analyst might quickly conclude
+that there are none. {[}ET: I don't know if the former statement is
+true?{]} Generally, one looks for noticeable departures from the model
 like non-linear dependency or heteroskedasticity. It is also possible to
@@
 
-There is an abundance of literature describing appropriate diagnostic
-methods for linear regression: \citet{draper1998applied},
+Abundance of literature describe appropriate diagnostic methods for
+linear regression: \citet{draper1998applied},
 \citet{montgomery1982introduction}, \citet{belsley_regression_1980},
@@
 amount of information can be provided by diagnostic plots than the
-corresponding tests in most empirical studies.} There is a common
-guidance by experts that plots are the best for diagnosing model fits.
+corresponding tests in most empirical studies.} A common guidance by
+experts is that optimal method for diagnosing model fits is by plotting
+the data.
 
-This is curious, and investigating why this might be common advice is
-the subject of this paper. The paper is structured as follows. The next
-background section describes the the types of departures that one
-expects to detect, and describes a formal process for reading residual
-plots, called visual inference, that can avoid the concerns about
-subjectiveness of human readers. Section \ref{experimental-design}
-describes the experimental setup to enable a comparison between decision
-made by formal hypothesis testing, and how humans would read diagnostic
-plots. The results are reported in Section \ref{results}. We finish with
-a discussion on future work, in particular how the responsibility for
+The persistence of this advice to check the plots is curious, and
+investigating why this might be common advice is the subject of this
+paper. The paper is structured as follows. The next background section
+describes the types of departures that one expects to detect, and
+describes a formal process for reading residual plots, called visual
+inference, that can avoid the concerns about subjectiveness of human
+readers {[}ET: CITATION HERE?{]}. Section \ref{experimental-design}
+describes the experimental setup to compare the decision made by formal
+hypothesis testing, and how humans would read diagnostic plots. The
+results are reported in Section \ref{results}. We finish with a
+discussion on future work, in particular how the responsibility for
 residual plot reading might be passed on to computer vision.
@@
 Non-linearity, heterskedasticity and non-normality are perhaps the three
-mostly commonly checked departures. Non-linearity is a type of model
+mostly checked departures. Non-linearity is a type of model
 misspecification caused by failing to include higher order terms of the
@@
 containing visual pattern of non-linearity is given at the top-right of
-Figure {[}ref here{]}. One can clearly observe the ``S-shape'' from the
-residual plot as the cubic term is not captured by the misspecified
-model. Heterskedasticity refers to the presence of nonconst error
-variance in a regression model. It is mostly due to the strict but false
+Figure \ref{fig:commondepartures}. One can clearly observe the
+``S-shape'' from the residual plot as the cubic term is not captured by
+the misspecified model.
+
+Heteroskedasticity refers to the presence of nonconstant error variance
+in a regression model. It is mostly due to the strict but false
 assumptions on the variance-covariance matrix of the error term. The
-usual pattern of heterskedasticity on a residual plot is the
-inconsistent spread of the residuals at different x values. Visually, it
-sometimes results in the so-called ``butterfly'' shape as shown in the
-bottom-left panel of Figure \ref{fig:residual-plot-common-departures},
-or the ``left-triangle'' and ``right-triganle'' shape where the smallest
-variance occurs at the edges of the x-axis. Compared to non-linearity
-and heterskedasticity, non-normality is usually harder to be detected
-from a residual plot since scatter plot is not excel in revealing
-marginal distribution. A favourable graphical summarise for this task is
-the quantile-quantile plot. However, for a consistent comparison in the
-later part of this paper, residual plot will be the focus. Besides, it
+usual pattern of heteroskedasticity on a residual plot is the
+inconsistent spread of the residuals across the horizontal axis.
+Visually, it sometimes results in the so-called ``butterfly'' shape as
+shown in the bottom-left panel of Figure \ref{fig:commondepartures}, or
+the ``left-triangle'' and ``right-triangle'' shape where the smallest
+variance occurs at one side of the horizontal axis.
+
+Compared to non-linearity and heteroskedasticity, non-normality is
+usually harder to detect from a residual plot since scatter plot do not
+readily reveal the marginal distribution. A favourable graphical summary
+for this task is the quantile-quantile plot. However, for a consistent
+comparison, residual plot will be the focus of this paper. Besides, it
 is important to note that not all regression models assume normality for
@@
 linear regression model. In the case that the normality assumption is
-violated, it is expected to observe data points that do not centralize
-around the horizontal zero line and unevenly distribute at both sides of
-the zero line. For example, given a skewed error distribution, fewer
-data points and more outliers are on one side of the zero line as shown
-at the bottom-right of Figure \ref{fig:residual-plot-common-departures}.
+violated, it is expected to observe data points do not center around the
+horizontal axis and there is an uneven distribution of the number points
+at both below and above the horizontal axis. For example, given a skewed
+error distribution, fewer data points and more outliers are on one side
+of the horizontal axis as shown at the bottom-right of Figure
+\ref{fig:commondepartures}.
 
@@
 
-Example residual plots given in Figure
-\ref{fig:residual-plot-common-departures} are examined by the
-corresponding model structural test, Breusch-Pagan test and
-Shapiro--Wilk test as shown in Table
+Example residual plots given in Figure \ref{fig:commondepartures} are
+examined by the corresponding model structural test, Breusch-Pagan test
+and Shapiro--Wilk test as shown in Table
 \ref{tab:example-residual-plot-table}. In the example, both the
@@
 error''. Additionally, outliers will often incorrectly trigger the
-rejection of the null hypothesis despite the residuals are well-behaved
-\citep{cook_applied_1999}. This can be largely avoided in diagnostic
-plots as experienced analysts can evaluate the acceptability of
-assumptions flexibly, even in the presence of outliers.
+rejection of the null hypothesis despite when majority of the residuals
+are well-behaved \citep{cook_applied_1999}. This can be largely avoided
+in diagnostic plots as experienced analysts can evaluate the
+acceptability of assumptions flexibly, even in the presence of outliers.
 
@@
 \ref{fig:fig:power-poly-uniform} has the actual data plot positioned at
-panel 20, where two out of five subjects detect it. It can be observed
-that a M-shape is presented in plot 20, but the signal is not strong
+panel 17, where two out of five subjects detect it. It can be observed
+that a ``U-shape'' is presented in plot 17, but the signal is not strong
 enough to attract all five subjects, resulting in a visual p-value
 sightly above the desired significance level \(\alpha = 0.05\). At
-\(log_e(E) = 4\) and \(log_e(E) = 6\), the designed visual features
+\(log_e(E) = 4\) and \(log_e(E) = 5\), the designed visual features
 become much clear and attractive, leading to a high percentage of
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/fig:power-poly-uniform-1} \caption{...}\label{fig:fig:power-poly-uniform-1}
-\end{figure}
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/fig:power-poly-uniform-2} \caption{...}\label{fig:fig:power-poly-uniform-2}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/fig:power-poly-uniform-1} \caption{...}\label{fig:fig:power-poly-uniform}
 \end{figure}
@@
 
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-6-1}
+\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-7-1}
 

### 2022-10-04 paper_comparison/paper_comparison/paper_comparison.Rmd
merged conflicts

@@
   fig.height = 6,
-  out.width = "100%")
+  out.width = "100%",
+  fig.align = "center")
 ```
@@
 In the case that the normality assumption is violated, it is expected to observe data points do not center around the horizontal axis and there is an uneven distribution of the number points at both below and above the horizontal axis.
-For example, given a skewed error distribution, fewer data points and more outliers are on one side of the horizontal axis as shown at the bottom-right of Figure \ref{fig:commondepartures}.
+For example, given a skewed error distribution, fewer data points and more outliers are on one side of the horizontal axis as shown in plot D of Figure \ref{fig:residual-plot-common-departures}.
 
@@
 
-## Experiment overview
+<!-- ## Experiment overview-->
 
@@
 
-Experiment I is designed to study the ability of human subjects to detect the effect of a non-linear term $\boldsymbol{z}$ which is a probabilist's Hermite polynomial [Herimite ref here] of another random vector $\boldsymbol{x}$ in a two variable statistical model formulated as:
+```{r different-shape-of-herimite, fig.height = 2, fig.cap="Polynomial forms generated for the residual plots used in experiment I.
+The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$."}
+stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
+
+ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300)
+
+VI_MODEL$plot(ori_dat) + 
+  ggtitle(quote(He[2]:"U")) + 
+  theme_light() -> p1
+
+VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote(He[3]:"S")) + 
+  theme_light() -> p2
+
+VI_MODEL$plot(poly_model(shape = 3, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote(He[6]:"M")) + 
+  theme_light() -> p3
+
+VI_MODEL$plot(poly_model(shape = 4, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote(He[18]:"Triple-U")) + 
+  theme_light() -> p4
+
+patchwork::wrap_plots(p1, p2, p3, p4, ncol = 4)
+```
+
+Experiment I is designed to study the ability of human subjects to detect the effect of a non-linear term $\boldsymbol{z}$ constructed using Hermite polynomials [Hermite ref here] on random vector $\boldsymbol{x}$ formulated as:
 
@@
 
-```{r different-shape-of-herimite, fig.cap="Example residual plots of the linear regression model used in experiment I.
-Four different values of $j$ are included in the experiment results in four different shapes of residuals."}
-stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
-
-ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300)
-
-VI_MODEL$plot(ori_dat) + 
-  ggtitle(quote(He[2]:"U shape")) + 
-  theme_light() -> p1
-
-VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote(He[3]:"S shape")) + 
-  theme_light() -> p2
-
-VI_MODEL$plot(poly_model(shape = 3, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote(He[6]:"M shape")) + 
-  theme_light() -> p3
-
-VI_MODEL$plot(poly_model(shape = 4, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote(He[18]:"Triple-U shape")) + 
-  theme_light() -> p4
-
-patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)
-```
-
 Four different distribution were used to generate $X_{raw}$ as shown in Figure \ref{fig:different-dist}.
@@
 
-```{r different-dist, fig.cap="Example residual plots of the linear regression model used in experiment I.
+```{r different-dist, fig.height = 2, fig.cap="Variations in fitted values, that might affect perception of residual plots.
 Four different distribution of $x_{raw}$ are used in the experiment to provide various visual patterns."}
@@
 VI_MODEL$plot(ori_dat) + 
-  ggtitle(quote(U(-1,1))) + 
+  ggtitle("A. Uniform") + #quote(U(-1,1))) + 
   theme_light() -> p1
@@
 VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
-  ggtitle(quote(N(0,0.3^2))) + 
+  ggtitle("B. Normal") + #quote(N(0,0.3^2))) + 
   theme_light() -> p2
@@
 VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
-  ggtitle(quote(lognormal(0,0.6^2)/3)) + 
+  ggtitle("C. Skewed") + #quote(lognormal(0,0.6^2)/3)) + 
   theme_light() -> p3
@@
 VI_MODEL$plot(poly_model(shape = 1, x = rand_uniform_d(k = 5, even = TRUE), sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
-  ggtitle(quote(U~"{1,5}")) + 
+  ggtitle("D. Discrete") + #quote(U~"{1,5}")) + 
   theme_light() -> p4
 
-patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)
+patchwork::wrap_plots(p1, p2, p3, p4, ncol = 4)
 ```
@@
 
-```{r different-shape-of-heter, fig.cap="Example residual plots of the linear regression model used in experiment II.
-Three different shapes (a = -1, 0, 1) are used in the experiment to create left-triangle shape, butterfly shape and right-triangle shape."}
+```{r different-shape-of-heter, fig.height=3, fig.cap="Heteroskedasticity forms used in experiment II.
+Three different shapes (a = -1, 0, 1) are used in the experiment to create left-triangle, butterfly and right-triangle shapes."}
 stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
@@
 VI_MODEL$plot(ori_dat) +
-  ggtitle(quote("a = -1: Left-triangle shape")) +
+  ggtitle(quote("A. a = -1: Left-triangle")) +
   theme_light() -> p1
@@
 VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote("a = 0: Butterfly shape")) +
+  ggtitle(quote("B. a = 0: Butterfly")) +
   theme_light() -> p2
@@
 VI_MODEL$plot(heter_model(a = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote("a = 1: Right-triangle shape")) +
+  ggtitle(quote("C. a = 1: Right-triangle")) +
   theme_light() -> p3
 
-patchwork::wrap_plots(p1, p2, p3, ncol = 2)
+patchwork::wrap_plots(p1, p2, p3, ncol = 3)
 ```

### 2022-10-04 paper_comparison/paper_comparison/paper_comparison.tex
merged conflicts

@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/residual-plot-common-departures-1} \caption{Example fitted vs residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/residual-plot-common-departures-1}
+
+}
+
+\caption{Example fitted vs residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
 The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted plot.}\label{fig:residual-plot-common-departures}
@@
 error distribution, fewer data points and more outliers are on one side
-of the horizontal axis as shown at the bottom-right of Figure
-\ref{fig:commondepartures}.
+of the horizontal axis as shown in plot D of Figure
+\ref{fig:residual-plot-common-departures}.
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/first-example-lineup-1} \caption{Visual testing is conducted using a lineup, as in the example here.
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/first-example-lineup-1}
+
+}
+
+\caption{Visual testing is conducted using a lineup, as in the example here.
 The residual plot computed from the observed data (plot 11, exhibiting non-linearity) is embedded among 19 null plots, where the residuals were simulated from a standard error model.
@@
 
-\hypertarget{experiment-overview}{%
-\subsection{Experiment overview}\label{experiment-overview}}
-
 Three experiments were conducted to investigate the difference between
@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-herimite-1}
+
+}
+
+\caption{Polynomial forms generated for the residual plots used in experiment I.
+The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$.}\label{fig:different-shape-of-herimite}
+\end{figure}
+
 Experiment I is designed to study the ability of human subjects to
-detect the effect of a non-linear term \(\boldsymbol{z}\) which is a
-probabilist's Hermite polynomial {[}Herimite ref here{]} of another
-random vector \(\boldsymbol{x}\) in a two variable statistical model
-formulated as:
+detect the effect of a non-linear term \(\boldsymbol{z}\) constructed
+using Hermite polynomials {[}Hermite ref here{]} on random vector
+\(\boldsymbol{x}\) formulated as:
 
@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-herimite-1} \caption{Example residual plots of the linear regression model used in experiment I.
-Four different values of $j$ are included in the experiment results in four different shapes of residuals.}\label{fig:different-shape-of-herimite}
-\end{figure}
-
 Four different distribution were used to generate \(X_{raw}\) as shown
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-dist-1} \caption{Example residual plots of the linear regression model used in experiment I.
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-dist-1}
+
+}
+
+\caption{Variations in fitted values, that might affect perception of residual plots.
 Four different distribution of $x_{raw}$ are used in the experiment to provide various visual patterns.}\label{fig:different-dist}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-poly-lineup-1} \caption{Lineup poly-24 in experiment I.
-Can you spot the most different plot?
-\label{fig:example-poly-lineup}}\label{fig:example-poly-lineup}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-poly-lineup-1}
+
+}
+
+\caption{Lineup poly-24 in experiment I. Can you spot the most different plot? \label{fig:example-poly-lineup}}\label{fig:example-poly-lineup}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-heter-1} \caption{Example residual plots of the linear regression model used in experiment II.
-Three different shapes (a = -1, 0, 1) are used in the experiment to create left-triangle shape, butterfly shape and right-triangle shape.}\label{fig:different-shape-of-heter}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-heter-1}
+
+}
+
+\caption{Heteroskedasticity forms used in experiment II.
+Three different shapes (a = -1, 0, 1) are used in the experiment to create left-triangle, butterfly and right-triangle shapes.}\label{fig:different-shape-of-heter}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-heter-lineup-1} \caption{Lineup heter-169 in experiment II.
-Can you spot the most different plot?
-\label{fig:example-heter-lineup}}\label{fig:example-heter-lineup}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-heter-lineup-1}
+
+}
+
+\caption{Lineup heter-169 in experiment II. Can you spot the most different plot? \label{fig:example-heter-lineup}}\label{fig:example-heter-lineup}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-n-1} \caption{Three different values of $n$ are used in experiment I, II and III to control the strength of the signal.}\label{fig:different-n}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-n-1}
+
+}
+
+\caption{Three different values of $n$ are used in experiment I, II and III to control the strength of the signal.}\label{fig:different-n}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-sigma-1} \caption{Example residual plots of the linear regression model used in experiment I.
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-sigma-1}
+
+}
+
+\caption{Example residual plots of the linear regression model used in experiment I.
 Four different values of $\sigma$ are included in the experiment to control the strength of the signal.}\label{fig:different-sigma}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-b-1} \caption{Example residual plots of the linear regression model used in experiment II.
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-b-1}
+
+}
+
+\caption{Example residual plots of the linear regression model used in experiment II.
 Five different values of $b$ are included in the experiment to control the strength of the signal.}\label{fig:different-b}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/fig:power-poly-uniform-1} \caption{...}\label{fig:fig:power-poly-uniform}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/fig:power-poly-uniform-1}
+
+}
+
+\caption{...}\label{fig:fig:power-poly-uniform}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/dist-power-1} \caption{...}\label{fig:dist-power}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/dist-power-1}
+
+}
+
+\caption{...}\label{fig:dist-power}
 \end{figure}
 
-\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-7-1}
+\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-7-1} \end{center}
 

### 2022-10-07 paper_comparison/paper_comparison/paper_comparison.Rmd
fix the p-value calc

@@
 
-```{r}
-vi_survey_processed <- readRDS(here::here("data/vi_survey_processed.rds")) 
-```
+<!-- TODO:difference between the simple formula and the simulation approach -->
 
@@
 ```{r fig:power-poly-uniform, fig.cap = "..."}
-vi_survey_processed %>%
+vi_survey %>%
   filter(x_dist == "uniform") %>%
@@
   filter(e_sigma != 0.125) %>%
-  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
+  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
   mutate(log_effect_size = log(effect_size)) %>%
@@
   geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) + geom_point(data = ~filter(.x, unique_lineup_id %in% c("poly_42", "poly_562", "poly_217", "poly_120")), aes(log_effect_size, prop_detect), alpha = 1, col = "red", size = 2) +
+  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) + geom_point(data = ~filter(.x, unique_lineup_id %in% c("poly_502", "poly_78", "poly_6", "poly_447")), aes(log_effect_size, prop_detect), alpha = 1, col = "red", size = 2) +
   stat_smooth(geom = "line", 
               data = ~boot_y(.x, reject_dirichlet, times = 500), 
-              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
+              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "Visual p-value"), 
               method = "glm", 
@@
 
-VI_MODEL$plot_lineup(vi_lineup[["poly_42"]]$data, 
+VI_MODEL$plot_lineup(vi_lineup[["poly_502"]]$data, 
                      remove_grid_line = TRUE, 
@@
                      remove_axis = TRUE) + 
-  ggtitle(bquote(log[e](E) %~~% 2~","~"F-test p-value" == .(format(filter(vi_survey_processed, unique_lineup_id == "poly_42")$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(vi_survey_processed, unique_lineup_id == "poly_42")$p_value_select_all_plots[1], digits = 3)))) -> p1
-
-VI_MODEL$plot_lineup(vi_lineup[["poly_562"]]$data, 
+  ggtitle(bquote(log[e](E) %~~% 1~","~"F-test p-value" == .(format(filter(vi_survey, unique_lineup_id == "poly_502")$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(vi_survey, unique_lineup_id == "poly_502")$p_value[1], digits = 3)))) -> p1
+
+VI_MODEL$plot_lineup(vi_lineup[["poly_78"]]$data, 
                      remove_grid_line = TRUE, 
@@
                      remove_axis = TRUE) + 
-  ggtitle(bquote(log[e](E) %~~% 3~","~"F-test p-value" == .(format(filter(vi_survey_processed, unique_lineup_id == "poly_562")$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(vi_survey_processed, unique_lineup_id == "poly_562")$p_value_select_all_plots[1], digits = 3)))) -> p2
-
-VI_MODEL$plot_lineup(vi_lineup[["poly_217"]]$data, 
+  ggtitle(bquote(log[e](E) %~~% 2~","~"F-test p-value" == .(format(filter(vi_survey, unique_lineup_id == "poly_78")$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(vi_survey, unique_lineup_id == "poly_78")$p_value[1], digits = 3)))) -> p2
+
+VI_MODEL$plot_lineup(vi_lineup[["poly_6"]]$data, 
                      remove_grid_line = TRUE, 
@@
                      remove_axis = TRUE) + 
-  ggtitle(bquote(log[e](E) %~~% 4~","~"F-test p-value" == .(format(filter(vi_survey_processed, unique_lineup_id == "poly_217")$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(vi_survey_processed, unique_lineup_id == "poly_217")$p_value_select_all_plots[1], digits = 3)))) -> p3
-
-VI_MODEL$plot_lineup(vi_lineup[["poly_120"]]$data, 
+  ggtitle(bquote(log[e](E) %~~% 3~","~"F-test p-value" == .(format(filter(vi_survey, unique_lineup_id == "poly_6")$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(vi_survey, unique_lineup_id == "poly_6")$p_value[1], digits = 3)))) -> p3
+
+VI_MODEL$plot_lineup(vi_lineup[["poly_447"]]$data, 
                      remove_grid_line = TRUE, 
@@
                      remove_axis = TRUE) + 
-  ggtitle(bquote(log[e](E) %~~% 2~","~"F-test p-value" == .(format(filter(vi_survey_processed, unique_lineup_id == "poly_120")$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(vi_survey_processed, unique_lineup_id == "poly_120")$p_value_select_all_plots[1], digits = 3)))) -> p4
+  ggtitle(bquote(log[e](E) %~~% 4~","~"F-test p-value" == .(format(filter(vi_survey, unique_lineup_id == "poly_447")$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(vi_survey, unique_lineup_id == "poly_447")$p_value[1], digits = 3)))) -> p4
 
@@
 ```{r eval = FALSE}
-vi_survey_processed %>%
+vi_survey %>%
   filter(x_dist == "uniform") %>%
@@
   filter(e_sigma != 0.125) %>%
-  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
+  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
   mutate(log_effect_size = log(effect_size)) %>%
-filter(abs(log(effect_size) - 3) < 0.2)
-
-vi_survey_processed %>%
+  filter(abs(log(effect_size) - 1) < 0.5) %>%
+  filter(n == 100)
+
+vi_survey %>%
   filter(x_dist == "uniform") %>%
@@
   geom_abline() +
-  geom_point(aes(p_value_uniform, p_value_select_all_plots)) +
+  geom_point(aes(p_value_uniform, p_value)) +
   scale_x_log10() +
@@
 ```{r dist-power, fig.cap="..."}
-vi_survey_processed %>%
+vi_survey %>%
   group_by(unique_lineup_id) %>%
@@
 ```{r}
-vi_survey_processed %>%
+vi_survey %>%
   group_by(unique_lineup_id) %>%
@@
   filter(b < 128) %>%
-  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
+  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
   mutate(log_effect_size = log(effect_size)) %>%

### 2022-10-07 paper_comparison/paper_comparison/paper_comparison.Rmd
solve c-interesting

@@
 \begin{equation} \label{eq:pvalue-beta-binomial}
-P(C \geq c_i) = \sum_{x=c_i}^{K}\frac{\Gamma(K + 1)}{\Gamma(x + 1)\Gamma(K - x + 1)}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},
+P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},
 \end{equation}
@@
 
-and $\Gamma(.)$ is the gamma function defined as
+Note that Equation \ref{eq:pvalue-beta-binomial} only works with integer $c_i$. For non-integer $c_i$, linear approximation will be applied to calculate the p-value
 
 ```{=tex}
-\begin{equation} \label{eq:gammafunction}
-\Gamma(z) = \int_{0}^{\infty}t^{z - 1}e^{-t}dt,\quad \text{where}\quad z>0. 
+\begin{equation} \label{eq:pvalue-beta-binomial-approx}
+P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \notin \mathbb{Z}.
 \end{equation}
@@
 
-The parameter $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} is usually unknown and hence needs to be estimated from the survey data.
+Besides, the parameter $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} is usually unknown and hence needs to be estimated from the survey data.
 For low values of $\alpha$, only a few plots are attractive to the observers and tend to be selected.
 For higher values of $\alpha$, the distribution of the probability of each plot being selected is more evenly.
-@vanderplas2021statistical suggested that $\alpha$ can be estimated using maximum likelihood estimation or visual estimation.
-But if $\alpha$ is small and only a few null plots in a lineup are attractive, MLE could fail to provide accurate estimates.
+@vanderplas2021statistical defined that a plot is $c$-interesting if $c$ or more participants selected the plot as the most different.
+Given the definition, The expected number of plots selected at least $c$ times, $E[Z_c]$, is calculated as
+
+```{=tex}
+\begin{equation} \label{eq:c-interesting-expectation}
+E[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation}
+```
+
+@vanderplas2021statistical suggested that $\alpha$ can be estimated using MLE and Equation \ref{eq:c-interesting-expectation}.
+But for precise estimate of $\alpha$, additional responses to Rorschach lineups, which is a type of lineup consists of only null plots, are required.
+
+<!-- But if $\alpha$ is small and only a few null plots in a lineup are attractive, MLE could fail to provide accurate estimates. -->
 
@@
 
-Instead, we directly estimated the probability of a lineup being rejected using a logistic regression with the natural logarithm of the effect size as the only regressor formulated as:
+Instead, we directly estimated the probability of a lineup being rejected using a logistic regression with the natural logarithm of the effect size as the only regressor formulated as
 
@@
 
-Effect $E$ is derived from the Kullback-Leibler divergence (see [appendix ref here]) formulated as:
+Effect $E$ is derived from the Kullback-Leibler divergence (see [appendix ref here]) formulated as
 

### 2022-10-07 paper_comparison/paper_comparison/paper_comparison.tex
fix the p-value calc
solve c-interesting

@@
 
-and \(\Gamma(.)\) is the gamma function defined as
+Note that Equation \ref{eq:pvalue-beta-binomial} only works with integer
+\(c_i\). For non-integer \(c_i\), linear approximation will be applied
+to calculate the p-value
 
-\begin{equation} \label{eq:gammafunction}
-\Gamma(z) = \int_{0}^{\infty}t^{z - 1}e^{-t}dt,\quad \text{where}\quad z>0. 
+\begin{equation} \label{eq:pvalue-beta-binomial-approx}
+P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \notin \mathbb{Z}.
 \end{equation}
 
-The parameter \(\alpha\) used in Equation \ref{eq:pvalue-beta-binomial}
-is usually unknown and hence needs to be estimated from the survey data.
-For low values of \(\alpha\), only a few plots are attractive to the
-observers and tend to be selected. For higher values of \(\alpha\), the
-distribution of the probability of each plot being selected is more
-evenly. \citet{vanderplas2021statistical} suggested that \(\alpha\) can
-be estimated using maximum likelihood estimation or visual estimation.
-But if \(\alpha\) is small and only a few null plots in a lineup are
-attractive, MLE could fail to provide accurate estimates.
+Besides, the parameter \(\alpha\) used in Equation
+\ref{eq:pvalue-beta-binomial} is usually unknown and hence needs to be
+estimated from the survey data. For low values of \(\alpha\), only a few
+plots are attractive to the observers and tend to be selected. For
+higher values of \(\alpha\), the distribution of the probability of each
+plot being selected is more evenly. \citet{vanderplas2021statistical}
+defined that a plot is \(c\)-interesting if \(c\) or more participants
+selected the plot as the most different. Given the definition, The
+expected number of plots selected at least \(c\) times, \(E[Z_c]\), is
+calculated as
+
+\begin{equation} \label{eq:c-interesting-expectation}
+E[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation}
+
+\citet{vanderplas2021statistical} suggested that \(\alpha\) can be
+estimated using MLE and Equation \ref{eq:c-interesting-expectation}. But
+for precise estimate of \(\alpha\), additional responses to Rorschach
+lineups, which is a type of lineup consists of only null plots, are
+required.
 
@@
 rejected using a logistic regression with the natural logarithm of the
-effect size as the only regressor formulated as:
+effect size as the only regressor formulated as
 
@@
 Effect \(E\) is derived from the Kullback-Leibler divergence (see
-{[}appendix ref here{]}) formulated as:
+{[}appendix ref here{]}) formulated as
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-7-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-6-1} \end{center}
 

### 2022-10-07 paper_comparison/paper_comparison/paper_comparison.Rmd
make factor table

@@
 
-Experiment I is designed to study the ability of human subjects to detect the effect of a non-linear term $\boldsymbol{z}$ constructed using Hermite polynomials [Hermite ref here] on random vector $\boldsymbol{x}$ formulated as:
+Experiment I is designed to study the ability of human subjects to detect the effect of a non-linear term $\boldsymbol{z}$ constructed using Hermite polynomials [Hermite ref here] on random vector $\boldsymbol{x}$ formulated as
 
@@
 
-The null regression model used to fit the realizations generated by the above model is formulated as:
+The null regression model used to fit the realizations generated by the above model is formulated as
 
@@
 ```{r results='asis'}
-data.frame(j = c(2, 3, 6, 18),
-           x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$")) %>%
+data.frame(j = c("2", "3", "6", "18", ""),
+           x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$", NA),
+           sigma = c("0.25", "1", "2", "4", ""),
+           a = c("-1", "0", "1", "", ""),
+           b = c("0.25", "1", "4", "16", "64"),
+           n = c("50", "100", "300", "", "")) %>%
   knitr::kable(col.names = NULL,
-               align = "rl",
+               align = "cccccc",
                escape = FALSE,
@@
                booktabs = TRUE, 
-               caption  = 'Parameter values for $j$ and $X_{raw}$', 
-               label = "model-parameter-table" ) %>%
-  kableExtra::add_header_above(c("Order of Hermite polynomial\n(sym1)", "Distribution of sym4\n")) %>%
+               caption  = 'Summary of factors involved in the study.', 
+               label = "model-factor-table" ) %>%
+  kableExtra::kable_styling(latex_options = "scale_down") %>%
+  kableExtra::add_header_above(c("Order of Hermite polynomial\n(sym1)", 
+                                 "Distribution of sym2\n", 
+                                 "Standard deviation of polynomial model\n(sym3)",
+                                 "Orientation of heteroskedasticity\n(sym4)", 
+                                 "Variance factor of heteroskedasticity model\n(sym5)",
+                                 "Sample size\n(sym6)")) %>%
   as.character() %>%
   gsub("sym1", "$j$", .) %>%
-  gsub("sym4", "$X_{raw}$", .) %>%
+  gsub("sym2", "$X_{raw}$", .) %>%
+  gsub("sym3", "$\\\\sigma$", .) %>%
+  gsub("sym4", "$a$", .) %>%
+  gsub("sym5", "$b$", .) %>%
+  gsub("sym6", "$n$", .) %>%
   cat()
@@
 
-As summarised by Table \ref{tab:model-parameter-table}, three additional parameters $n$, $\sigma$ and $b$ were used to control the strength of the signal so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
+As summarised by Table \ref{tab:model-factor-table}, three additional parameters $n$, $\sigma$ and $b$ were used to control the strength of the signal so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
 Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ were used in experiment I and II respectively.

### 2022-10-07 paper_comparison/paper_comparison/paper_comparison.tex
make factor table

@@
 using Hermite polynomials {[}Hermite ref here{]} on random vector
-\(\boldsymbol{x}\) formulated as:
+\(\boldsymbol{x}\) formulated as
 
@@
 The null regression model used to fit the realizations generated by the
-above model is formulated as:
+above model is formulated as
 
@@
 
-\caption{\label{tab:model-parameter-table}Parameter values for $j$ and $X_{raw}$}
+\caption{\label{tab:model-factor-table}Summary of factors involved in the study.}
 \centering
-\begin{tabular}[t]{rl}
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{cccccc}
 \toprule
-\multicolumn{1}{c}{\makecell[c]{Order of Hermite polynomial\\($j$)}} & \multicolumn{1}{c}{\makecell[c]{Distribution of $X_{raw}$\\}} \\
-\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2}
-2 & $U(-1, 1)$\\
-3 & $N(0, 0.3^2)$\\
-6 & $lognormal(0, 0.6^2)/3$\\
-18 & $U\{1, 5\}$\\
+\multicolumn{1}{c}{\makecell[c]{Order of Hermite polynomial\\($j$)}} & \multicolumn{1}{c}{\makecell[c]{Distribution of $X_{raw}$\\}} & \multicolumn{1}{c}{\makecell[c]{Standard deviation of polynomial model\\($\sigma$)}} & \multicolumn{1}{c}{\makecell[c]{Orientation of heteroskedasticity\\($a$)}} & \multicolumn{1}{c}{\makecell[c]{Variance factor of heteroskedasticity model\\($b$)}} & \multicolumn{1}{c}{\makecell[c]{Sample size\\($n$)}} \\
+\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
+2 & $U(-1, 1)$ & 0.25 & -1 & 0.25 & 50\\
+3 & $N(0, 0.3^2)$ & 1 & 0 & 1 & 100\\
+6 & $lognormal(0, 0.6^2)/3$ & 2 & 1 & 4 & 300\\
+18 & $U\{1, 5\}$ & 4 &  & 16 & \\
+ & NA &  &  & 64 & \\
 \bottomrule
-\end{tabular}
+\end{tabular}}
 \end{table}
@@
 
-As summarised by Table \ref{tab:model-parameter-table}, three additional
+As summarised by Table \ref{tab:model-factor-table}, three additional
 parameters \(n\), \(\sigma\) and \(b\) were used to control the strength

### 2022-10-08 paper_comparison/paper_comparison/paper_comparison.Rmd
mostly finish exp design

@@
 Visual patterns of non-linearity were simulated using four different order of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
+A summary of the factors is given in Table \ref{tab:model-factor-table}.
 
@@
 data.frame(j = c("2", "3", "6", "18", ""),
-           x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$", NA),
+           x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$", "NA"),
            sigma = c("0.25", "1", "2", "4", ""),
@@
                                  "Standard deviation of polynomial model\n(sym3)",
-                                 "Orientation of heteroskedasticity\n(sym4)", 
+                                 "Shape of heteroskedasticity\n(sym4)", 
                                  "Variance factor of heteroskedasticity model\n(sym5)",
@@
 Visual patterns were simulated using three different shapes ($a$ = -1, 0, 1) and the same four different distribution of $X_{raw}$ used in experiment I.
+A summary of the factors is given in Table \ref{tab:model-factor-table}.
 
@@
 
-As summarised by Table \ref{tab:model-factor-table}, three additional parameters $n$, $\sigma$ and $b$ were used to control the strength of the signal so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
+As summarised in Table \ref{tab:model-factor-table}, three additional parameters $n$, $\sigma$ and $b$ were used to control the strength of the signal so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
 Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ were used in experiment I and II respectively.
 Figure \ref{fig:different-sigma} and \ref{fig:different-b} demonstrate the impact of these two parameters.
+A large value of $\sigma$ will increase the variation of the error of the polynomial model and decrease the visibility of the visual pattern.
+The parameter $b$ controls the ratio of the standard deviation of the heteroskedasticity across the domain of the regressor.
+Given $x \neq a$, a larger value of $b$ will lead to a larger ratio of the variance at $x$ to the variance at $x - a = 0$, making the visual pattern more obvious.
 
@@
 
-```{r different-n, fig.cap="Three different values of $n$ are used in experiment I, II and III to control the strength of the signal."}
+```{r different-n, fig.height = 3, fig.cap="Three different values of $n$ are used in experiment I, II and III to control the strength of the signal."}
 ori_dat <- poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(300)
@@
 
-patchwork::wrap_plots(p2, p3, p1, ncol = 2)
-```
-
-```{r different-sigma, fig.cap="Example residual plots of the linear regression model used in experiment I.
+patchwork::wrap_plots(p2, p3, p1, ncol = 3)
+```
+
+```{r different-sigma, fig.height = 2, fig.cap="Example residual plots of the linear regression model used in experiment I.
 Four different values of $\\sigma$ are included in the experiment to control the strength of the signal."}
@@
 
-patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)
-```
-
-```{r different-b, fig.cap="Example residual plots of the linear regression model used in experiment II.
+patchwork::wrap_plots(p1, p2, p3, p4, ncol = 4)
+```
+
+```{r different-b, fig.height = 2, fig.cap="Example residual plots of the linear regression model used in experiment II.
 Five different values of $b$ are included in the experiment to control the strength of the signal."}
@@
 
-patchwork::wrap_plots(p1, p2, p3, p4, p5, ncol = 3)
+patchwork::wrap_plots(p1, p2, p3, p4, p5, ncol = 5)
 ```
@@
 
-Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in 1152 different lineups.
-In addition, each lineup is designed to be evaluated by five different subjects and lineups with uniform distribution is designed to be evaluated by 11 subjects to provide reasonable estimates of the visual p-value.
-
-Thus, $443$ subjects were recruited to satisfy the design of the experiment. 
+Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in $(4 \times 4 \times 4 \times 3 + 4 \times 3 \times 5 \times 3) \times 3 = 1116$ different lineups.
+In addition, each lineup is designed to be evaluated by five different subjects.
+After attempting some pilot studies internally in our research group, we decide to present a block of 20 lineups to every subject.
+And to ensure the quality of the survey data, two lineups with obvious visual patterns are included as attention checks.
+Thus, $576 \times 5/(20-2) = 160$ and $540 \times 5 /(20-2) = 150$ subjects were recruited to satisfy the design of the experiment I and experiment II respectively.
+
+As mentioned in Section [the alpha section], $\alpha$ used in Equation \ref{eq:eq:pvalue-beta-binomial} needs to be estimated using Rorschach lineups.
+Hence, 36 Rorschach lineups with all combinations of $n$ and $X_{raw}$ and three replications are designed to be included in experiment III.
+All Rorschach lineups are planned to be evaluated by 20 subjects.
+However, presenting only Rorschach lineups to subjects are considered to be bad practices as subjects will lose interest quickly.
+We planned to also collect 6 more evaluations on the 279 lineups with uniform distribution, resulted in $(36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3)/(20-2) = 133$ subjects recruited for experiment III.
+
 <!-- TODO:subject allocation algorithm -->
@@
 Information about preferred pronoun, age group, education, and previous experience in visual experiment were also collected.
-In every block of 20 lineups that presented to a subject, two lineups with obvious visual patterns were included as attention checks.
 A subject's submission was only accepted if the actual data plot was identified for at least one attention check.

### 2022-10-08 paper_comparison/paper_comparison/paper_comparison.tex
mostly finish exp design

@@
 four different distribution of \(X_{raw}\): (1) \(U(-1, 1)\), (2)
-\(N(0, 0.3^2)\), (3) \(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\).
+\(N(0, 0.3^2)\), (3) \(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). A
+summary of the factors is given in Table \ref{tab:model-factor-table}.
 
@@
 \toprule
-\multicolumn{1}{c}{\makecell[c]{Order of Hermite polynomial\\($j$)}} & \multicolumn{1}{c}{\makecell[c]{Distribution of $X_{raw}$\\}} & \multicolumn{1}{c}{\makecell[c]{Standard deviation of polynomial model\\($\sigma$)}} & \multicolumn{1}{c}{\makecell[c]{Orientation of heteroskedasticity\\($a$)}} & \multicolumn{1}{c}{\makecell[c]{Variance factor of heteroskedasticity model\\($b$)}} & \multicolumn{1}{c}{\makecell[c]{Sample size\\($n$)}} \\
+\multicolumn{1}{c}{\makecell[c]{Order of Hermite polynomial\\($j$)}} & \multicolumn{1}{c}{\makecell[c]{Distribution of $X_{raw}$\\}} & \multicolumn{1}{c}{\makecell[c]{Standard deviation of polynomial model\\($\sigma$)}} & \multicolumn{1}{c}{\makecell[c]{Shape of heteroskedasticity\\($a$)}} & \multicolumn{1}{c}{\makecell[c]{Variance factor of heteroskedasticity model\\($b$)}} & \multicolumn{1}{c}{\makecell[c]{Sample size\\($n$)}} \\
 \cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
@@
 were simulated using three different shapes (\(a\) = -1, 0, 1) and the
-same four different distribution of \(X_{raw}\) used in experiment I.
+same four different distribution of \(X_{raw}\) used in experiment I. A
+summary of the factors is given in Table \ref{tab:model-factor-table}.
 
@@
 
-As summarised by Table \ref{tab:model-factor-table}, three additional
+As summarised in Table \ref{tab:model-factor-table}, three additional
 parameters \(n\), \(\sigma\) and \(b\) were used to control the strength
@@
 respectively. Figure \ref{fig:different-sigma} and \ref{fig:different-b}
-demonstrate the impact of these two parameters.
+demonstrate the impact of these two parameters. A large value of
+\(\sigma\) will increase the variation of the error of the polynomial
+model and decrease the visibility of the visual pattern. The parameter
+\(b\) controls the ratio of the standard deviation of the
+heteroskedasticity across the domain of the regressor. Given
+\(x \neq a\), a larger value of \(b\) will lead to a larger ratio of the
+variance at \(x\) to the variance at \(x - a = 0\), making the visual
+pattern more obvious.
 
@@
 Three replications are made for each of the parameter values shown in
-Table \ref{tab:parameter-table} resulting in 1152 different lineups. In
-addition, each lineup is designed to be evaluated by five different
-subjects and lineups with uniform distribution is designed to be
-evaluated by 11 subjects to provide reasonable estimates of the visual
-p-value.
+Table \ref{tab:parameter-table} resulting in
+\((4 \times 4 \times 4 \times 3 + 4 \times 3 \times 5 \times 3) \times 3 = 1116\)
+different lineups. In addition, each lineup is designed to be evaluated
+by five different subjects. After attempting some pilot studies
+internally in our research group, we decide to present a block of 20
+lineups to every subject. And to ensure the quality of the survey data,
+two lineups with obvious visual patterns are included as attention
+checks. Thus, \(576 \times 5/(20-2) = 160\) and
+\(540 \times 5 /(20-2) = 150\) subjects were recruited to satisfy the
+design of the experiment I and experiment II respectively.
 
-Thus, \(443\) subjects were recruited to satisfy the design of the
-experiment.
+As mentioned in Section {[}the alpha section{]}, \(\alpha\) used in
+Equation \ref{eq:eq:pvalue-beta-binomial} needs to be estimated using
+Rorschach lineups. Hence, 36 Rorschach lineups with all combinations of
+\(n\) and \(X_{raw}\) and three replications are designed to be included
+in experiment III. All Rorschach lineups are planned to be evaluated by
+20 subjects. However, presenting only Rorschach lineups to subjects are
+considered to be bad practices as subjects will lose interest quickly.
+We planned to also collect 6 more evaluations on the 279 lineups with
+uniform distribution, resulted in
+\((36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3)/(20-2) = 133\)
+subjects recruited for experiment III.
 
@@
 twice. Information about preferred pronoun, age group, education, and
-previous experience in visual experiment were also collected. In every
-block of 20 lineups that presented to a subject, two lineups with
-obvious visual patterns were included as attention checks. A subject's
-submission was only accepted if the actual data plot was identified for
-at least one attention check. Data of rejected submissions were
-discarded automatically to maintain the overall data quality.
+previous experience in visual experiment were also collected. A
+subject's submission was only accepted if the actual data plot was
+identified for at least one attention check. Data of rejected
+submissions were discarded automatically to maintain the overall data
+quality.
 

### 2022-10-10 paper_comparison/paper_comparison/paper_comparison.Rmd
add p-value cp

@@
 However, some subjects will try to maximize their earnings for minimum effort.
-During the review of submissions, if we found a subject objectively demonstrated clear low-effort throughout the experiment, i.e., failed all attention checks, we rejected the submission.
+During the review of submissions, if we found a subject objectively demonstrated clear low-effort throughout the experiment, e.g., failed all attention checks, we rejected the submission.
 The rejected submissions will be removed immediately, and Prolific will automatically recruit another subject as substitution.
@@
 
-In overall, there were a total of 3200 lineup evaluations made by 160 subjects in both experiment I and experiment II respectively, where 320 lineup evaluations were attention checks and were not used in the following analysis.
+In overall, there were a total of 8860 lineup evaluations made by 443 subjects in experiment I, experiment II and experiment III, where 886 lineup evaluations were attention checks and were not used in the following analysis.
 
 The collated dataset is provided in `vi_survey` of the `visage` `R` package.
+
+## P-value comparison
+
+Visual inference and conventional hypothesis testing are two distinct statistical procedures.
+As shown in \ref{fig:p-value-comparison}, the visual test almost always gives higher p-values than the conventional test, suggesting that humans have lower confidence and much higher tolerance of the residual departures than the numerical procedure.
+In fact, only around 58% of the cases, both tests will agree on the rejection decision given 95% confidence level.
+And in 41% of the cases, visual test does not reject the null hypothesis but conventional test does.
+This is an indicator of "false alarm"
+
+```{r eval = FALSE}
+vi_survey %>%
+  filter(lineup_id <= 576) %>%
+  filter(is.na(b) | b != 0) %>%
+  mutate(test = p_value > 0.05 & conventional_p_value < 0.05) %>%
+  {mean(.$test)}
+```
+
+```{r p-value-comparison, fig.cap="..."}
+vi_survey %>%
+  filter(lineup_id <= 576) %>%
+  filter(is.na(b) | b != 0) %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  # filter(x_dist == "uniform") %>%
+  ggplot() +
+  geom_point(aes(p_value, conventional_p_value), alpha = 0.4) +
+  geom_abline() +
+  scale_x_log10() +
+  scale_y_log10() +
+  facet_wrap(~type)
+```
 

### 2022-10-10 paper_comparison/paper_comparison/paper_comparison.tex
add p-value cp

@@
 submissions, if we found a subject objectively demonstrated clear
-low-effort throughout the experiment, i.e., failed all attention checks,
+low-effort throughout the experiment, e.g., failed all attention checks,
 we rejected the submission. The rejected submissions will be removed
@@
 
-In overall, there were a total of 3200 lineup evaluations made by 160
-subjects in both experiment I and experiment II respectively, where 320
+In overall, there were a total of 8860 lineup evaluations made by 443
+subjects in experiment I, experiment II and experiment III, where 886
 lineup evaluations were attention checks and were not used in the
@@
 \texttt{visage} \texttt{R} package.
+
+\hypertarget{p-value-comparison}{%
+\subsection{P-value comparison}\label{p-value-comparison}}
+
+Visual inference and conventional hypothesis testing are two distinct
+statistical procedures. As shown in \ref{fig:p-value-comparison}, the
+visual test almost always gives higher p-values than the conventional
+test, suggesting that humans have lower confidence and much higher
+tolerance of the residual departures than the numerical procedure. In
+fact, only around 58\% of the cases, both tests will agree on the
+rejection decision given 95\% confidence level. And in 41\% of the
+cases, visual test does not reject the null hypothesis but conventional
+test does. This is an indicator of ``false alarm''
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/p-value-comparison-1}
+
+}
+
+\caption{...}\label{fig:p-value-comparison}
+\end{figure}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-6-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-7-1} \end{center}
 

### 2022-10-10 paper_comparison/paper_comparison/paper_comparison.Rmd
fill in p-value cp

@@
 And in 41% of the cases, visual test does not reject the null hypothesis but conventional test does.
-This is an indicator of "false alarm"
+Although we know that the all the designed lineups contain various degrees of residual departures and the conventional test is telling truth, analysts and general public as the consumers of the output will possibly not believe in or at least not response to the "false alarm" if they can not see the departures directly from the data plot and view it as impactful.
+The sensitivity of the conventional test may distract and discourage us from finding a good linear approximation to the data we try to model.
+This also partly explains why abundance of literature suggesting the use of data plot in regression diagnostics.
 

### 2022-10-10 paper_comparison/paper_comparison/paper_comparison.tex
fill in p-value cp

@@
 cases, visual test does not reject the null hypothesis but conventional
-test does. This is an indicator of ``false alarm''
+test does. Although we know that the all the designed lineups contain
+various degrees of residual departures and the conventional test is
+telling truth, analysts and general public as the consumers of the
+output will possibly not believe in or at least not response to the
+``false alarm'' if they can not see the departures directly from the
+data plot and view it as impactful. The sensitivity of the conventional
+test may distract and discourage us from finding a good linear
+approximation to the data we try to model. This also partly explains why
+abundance of literature suggesting the use of data plot in regression
+diagnostics.
 

### 2022-10-11 paper_comparison/paper_comparison/paper_comparison.Rmd
fix some references

@@
 Figure \ref{fig:different-sigma} and \ref{fig:different-b} demonstrate the impact of these two parameters.
-A large value of $\sigma$ will increase the variation of the error of the polynomial model and decrease the visibility of the visual pattern.
+A large value of $\sigma$ will increase the variation of the error of the non-linear model and decrease the visibility of the visual pattern.
 The parameter $b$ controls the ratio of the standard deviation of the heteroskedasticity across the domain of the regressor.
@@
 
-```{r different-sigma, fig.height = 2, fig.cap="Example residual plots of the linear regression model used in experiment I.
-Four different values of $\\sigma$ are included in the experiment to control the strength of the signal."}
+```{r different-sigma, fig.height = 2, fig.cap="Four different values of $\\sigma$ are used in the experiment I to control the strength of the signal."}
 ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)
@@
 
-```{r different-b, fig.height = 2, fig.cap="Example residual plots of the linear regression model used in experiment II.
-Five different values of $b$ are included in the experiment to control the strength of the signal."}
+```{r different-b, fig.height = 2, fig.cap="Five different values of $b$ are used in experiment II to control the strength of the signal."}
 stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
@@
 Visual inference and conventional hypothesis testing are two distinct statistical procedures.
-As shown in \ref{fig:p-value-comparison}, the visual test almost always gives higher p-values than the conventional test, suggesting that humans have lower confidence and much higher tolerance of the residual departures than the numerical procedure.
+As displayed in Figure \ref{fig:p-value-comparison}, the visual test almost always gives higher p-values than the conventional test in both polynomial and heteroskedasticity models, suggesting that humans have lower confidence and much higher tolerance of the residual departures than the numerical procedure.
 In fact, only around 58% of the cases, both tests will agree on the rejection decision given 95% confidence level.
@@
   {mean(.$test)}
-```
-
-```{r p-value-comparison, fig.cap="..."}
+
+vi_survey %>%
+  filter(lineup_id <= 576) %>%
+  filter(is.na(b) | b != 0) %>%
+ggplot() +
+  geom_histogram(aes(p_value), bins = 300) +
+  scale_x_sqrt() +
+ facet_wrap(~type)
+```
+
+```{r p-value-comparison, fig.height=4, fig.cap="..."}
 vi_survey %>%
@@
   geom_abline() +
-  scale_x_log10() +
-  scale_y_log10() +
+  scale_x_sqrt() +
+  scale_y_sqrt() +
   facet_wrap(~type)
@@
 
-Figure \ref{fig:power-poly-uniform} shows the estimated power of visual test against natural logarithm of the effect with comparison to the power of an exact test - F-test, and the power of two other residual-based conventional tests commonly used in regression diagnostics but for testing other departures from the model assumptions.
+Figure \ref{fig:power-poly-uniform} shows the estimated power of visual test on lineups with uniform distribution $X_{raw}$ against natural logarithm of the effect with comparison to the power of an exact test - F-test, and the power of two other residual-based conventional tests commonly used in regression diagnostics but for testing other departures from the model assumptions.
 In overall, the power of all four tests increases as the effect becomes larger.
@@
 This coincides the point made by @cook1982residuals mentioned in \ref{hypothesis-testing} that residual-based tests for a specific type of model defect are sensitive to other types of model defects.
-At $log_e(E) = 6$, the power curve of F-test reaches almost 100% followed by the visual test by a small margin.
+At $log_e(E) \approx 6$, the power curve of F-test reaches almost 100% followed by the visual test by a small margin.
 The power of Breusch--Pagan test and Shapiro--Wilk test reach around 75% and 30% respectively.
@@
 The margin is largest at around $log_e(E) = 2$.
-An example lineup is included in Figure \ref{fig:power-overview} where none of subjects detect the actual data plot positioned at panel 14.
+An example lineup is included in Figure \ref{fig:power-poly-uniform} where none of subjects detect the actual data plot positioned at panel 10.
 It demonstrates that at this level of difficulty, the designed visual feature is rarely visible, making the actual data plot indistinguishable from residual plots simulated from the assumed model.
@@
 Without knowing the designed visual features beforehand, it is actually hard to identify the actual data plot by pure image comparison.
-The corresponding example lineup for $log_e(E) = 3$ shown in Figure \ref{fig:fig:power-poly-uniform} has the actual data plot positioned at panel 17, where two out of five subjects detect it.
-It can be observed that a "U-shape" is presented in plot 17, but the signal is not strong enough to attract all five subjects, resulting in a visual p-value sightly above the desired significance level $\alpha = 0.05$.
-At $log_e(E) = 4$ and $log_e(E) = 5$, the designed visual features become much clear and attractive, leading to a high percentage of rejection of the null hypothesis.
+The corresponding example lineup for $log_e(E) = 3$ shown in Figure \ref{fig:fig:power-poly-uniform} has the actual data plot positioned at panel 7, where four of 11 subjects detect it.
+It can be observed that a "U-shape" is presented in plot 7, but the signal is not strong enough to attract all five subjects, resulting in a visual p-value sightly above the desired significance level $\alpha = 0.05$.
+At $log_e(E) = 4$, the designed visual features become much clear and attractive, leading to a high percentage of rejection of the null hypothesis.
 Figure \ref{fig:fig:power-poly-uniform} gives example lineups of such cases.
@@
 
-```{r fig:power-poly-uniform, fig.cap = "..."}
+```{r power-poly-uniform, fig.cap = "..."}
 vi_survey %>%
@@
 
-The impact of the distribution of $X_raw$ on the power is shown in Figure \ref{fig:dist-power}.
+The impact of the distribution of $X_raw$ on the power is displayed in Figure \ref{fig:dist-power}.
 The power curve of F-test is stable across different distributions, while the visual test has a steeper power curve for normal and uniform distribution.

### 2022-10-11 paper_comparison/paper_comparison/paper_comparison.tex
fix some references

@@
 demonstrate the impact of these two parameters. A large value of
-\(\sigma\) will increase the variation of the error of the polynomial
+\(\sigma\) will increase the variation of the error of the non-linear
 model and decrease the visibility of the visual pattern. The parameter
@@
 
-\caption{Example residual plots of the linear regression model used in experiment I.
-Four different values of $\sigma$ are included in the experiment to control the strength of the signal.}\label{fig:different-sigma}
+\caption{Four different values of $\sigma$ are used in the experiment I to control the strength of the signal.}\label{fig:different-sigma}
 \end{figure}
@@
 
-\caption{Example residual plots of the linear regression model used in experiment II.
-Five different values of $b$ are included in the experiment to control the strength of the signal.}\label{fig:different-b}
+\caption{Five different values of $b$ are used in experiment II to control the strength of the signal.}\label{fig:different-b}
 \end{figure}
@@
 Visual inference and conventional hypothesis testing are two distinct
-statistical procedures. As shown in \ref{fig:p-value-comparison}, the
-visual test almost always gives higher p-values than the conventional
-test, suggesting that humans have lower confidence and much higher
-tolerance of the residual departures than the numerical procedure. In
-fact, only around 58\% of the cases, both tests will agree on the
-rejection decision given 95\% confidence level. And in 41\% of the
-cases, visual test does not reject the null hypothesis but conventional
-test does. Although we know that the all the designed lineups contain
-various degrees of residual departures and the conventional test is
-telling truth, analysts and general public as the consumers of the
-output will possibly not believe in or at least not response to the
-``false alarm'' if they can not see the departures directly from the
-data plot and view it as impactful. The sensitivity of the conventional
-test may distract and discourage us from finding a good linear
-approximation to the data we try to model. This also partly explains why
-abundance of literature suggesting the use of data plot in regression
-diagnostics.
+statistical procedures. As displayed in Figure
+\ref{fig:p-value-comparison}, the visual test almost always gives higher
+p-values than the conventional test in both polynomial and
+heteroskedasticity models, suggesting that humans have lower confidence
+and much higher tolerance of the residual departures than the numerical
+procedure. In fact, only around 58\% of the cases, both tests will agree
+on the rejection decision given 95\% confidence level. And in 41\% of
+the cases, visual test does not reject the null hypothesis but
+conventional test does. Although we know that the all the designed
+lineups contain various degrees of residual departures and the
+conventional test is telling truth, analysts and general public as the
+consumers of the output will possibly not believe in or at least not
+response to the ``false alarm'' if they can not see the departures
+directly from the data plot and view it as impactful. The sensitivity of
+the conventional test may distract and discourage us from finding a good
+linear approximation to the data we try to model. This also partly
+explains why abundance of literature suggesting the use of data plot in
+regression diagnostics.
 
@@
 Figure \ref{fig:power-poly-uniform} shows the estimated power of visual
-test against natural logarithm of the effect with comparison to the
-power of an exact test - F-test, and the power of two other
-residual-based conventional tests commonly used in regression
-diagnostics but for testing other departures from the model assumptions.
-In overall, the power of all four tests increases as the effect becomes
-larger. The power curve of F-test climbs aggressively from 25\% to
-around 90\% as \(log_e(E)\) increases from 0 to 2, while others respond
-inactively to the change of effect and remain lower than 25\% throughout
-the period, showing that as an exact test, the F-test is relatively more
-sensitive to the type of model defects that being considered. The power
-of visual test arises steadily and nearly linearly to around 90\% as
-\(log_e(E)\) increases from 2 to 5, suggesting that the effect starts to
-make noticeable impact on the degree of the presence of the designed
-visual features. Other two inappropriate conventional tests shows
-improvement at the same time but at a lower rate. This coincides the
-point made by \citet{cook1982residuals} mentioned in
-\ref{hypothesis-testing} that residual-based tests for a specific type
-of model defect are sensitive to other types of model defects. At
-\(log_e(E) = 6\), the power curve of F-test reaches almost 100\%
-followed by the visual test by a small margin. The power of
-Breusch--Pagan test and Shapiro--Wilk test reach around 75\% and 30\%
-respectively.
+test on lineups with uniform distribution \(X_{raw}\) against natural
+logarithm of the effect with comparison to the power of an exact test -
+F-test, and the power of two other residual-based conventional tests
+commonly used in regression diagnostics but for testing other departures
+from the model assumptions. In overall, the power of all four tests
+increases as the effect becomes larger. The power curve of F-test climbs
+aggressively from 25\% to around 90\% as \(log_e(E)\) increases from 0
+to 2, while others respond inactively to the change of effect and remain
+lower than 25\% throughout the period, showing that as an exact test,
+the F-test is relatively more sensitive to the type of model defects
+that being considered. The power of visual test arises steadily and
+nearly linearly to around 90\% as \(log_e(E)\) increases from 2 to 5,
+suggesting that the effect starts to make noticeable impact on the
+degree of the presence of the designed visual features. Other two
+inappropriate conventional tests shows improvement at the same time but
+at a lower rate. This coincides the point made by
+\citet{cook1982residuals} mentioned in \ref{hypothesis-testing} that
+residual-based tests for a specific type of model defect are sensitive
+to other types of model defects. At \(log_e(E) \approx 6\), the power
+curve of F-test reaches almost 100\% followed by the visual test by a
+small margin. The power of Breusch--Pagan test and Shapiro--Wilk test
+reach around 75\% and 30\% respectively.
 
@@
 at around \(log_e(E) = 2\). An example lineup is included in Figure
-\ref{fig:power-overview} where none of subjects detect the actual data
-plot positioned at panel 14. It demonstrates that at this level of
+\ref{fig:power-poly-uniform} where none of subjects detect the actual
+data plot positioned at panel 10. It demonstrates that at this level of
 difficulty, the designed visual feature is rarely visible, making the
@@
 \ref{fig:fig:power-poly-uniform} has the actual data plot positioned at
-panel 17, where two out of five subjects detect it. It can be observed
-that a ``U-shape'' is presented in plot 17, but the signal is not strong
-enough to attract all five subjects, resulting in a visual p-value
-sightly above the desired significance level \(\alpha = 0.05\). At
-\(log_e(E) = 4\) and \(log_e(E) = 5\), the designed visual features
-become much clear and attractive, leading to a high percentage of
-rejection of the null hypothesis. Figure
-\ref{fig:fig:power-poly-uniform} gives example lineups of such cases. It
-is also interesting to observe that the power curve of visual test
-behaves similarly to those inappropriate conventional tests until the
-designed visual features really becomes visible.
+panel 7, where four of 11 subjects detect it. It can be observed that a
+``U-shape'' is presented in plot 7, but the signal is not strong enough
+to attract all five subjects, resulting in a visual p-value sightly
+above the desired significance level \(\alpha = 0.05\). At
+\(log_e(E) = 4\), the designed visual features become much clear and
+attractive, leading to a high percentage of rejection of the null
+hypothesis. Figure \ref{fig:fig:power-poly-uniform} gives example
+lineups of such cases. It is also interesting to observe that the power
+curve of visual test behaves similarly to those inappropriate
+conventional tests until the designed visual features really becomes
+visible.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/fig:power-poly-uniform-1}
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/power-poly-uniform-1}
 
@@
 
-\caption{...}\label{fig:fig:power-poly-uniform}
+\caption{...}\label{fig:power-poly-uniform}
 \end{figure}
@@
 
-The impact of the distribution of \(X_raw\) on the power is shown in
+The impact of the distribution of \(X_raw\) on the power is displayed in
 Figure \ref{fig:dist-power}. The power curve of F-test is stable across

### 2022-11-01 paper_comparison/analysis/vi_survey_analysis.qmd
update analysis

---
title: "vi_survey_analysis"
execute:
  warning: false
format: 
  html:
    code-fold: true
    toc: true
editor: visual
editor_options: 
  chunk_output_type: console
---

# Analysis of the survey result

## Setup

```{r}
library(visage)
library(tidyverse)
```

```{r cache = TRUE}
vi_lineup <- readRDS(here::here("data/vi_lineup.rds"))
poly_conv_sim <- readRDS(here::here("data/poly_conventional_simulation.rds"))
heter_conv_sim <- readRDS(here::here("data/heter_conventional_simulation.rds"))
```

## Overview

The survey dataset `vi_survey` contains `r length(vi_survey)` columns.

```{r}
knitr::kable(data.frame(names(vi_survey), 
                        c("Experiment ID",
                          "Unique lineup ID",
                          "Lineup ID in an experiment",
                          "Page number of the study website",
                          "Subject ID in an experiment",
                          "The count of the lineup presented to a subject",
                          "Response time to a page",
                          "Selected plots in a lineup",
                          "Number of selected plots",
                          "The position of the actual data plot",
                          "Whether the subject selected the actual data plot",
                          "Weighted detect is `detect`/`num_selection` and 1/20 for `num_selection == 0`",
                          "Average weighted detect for a lineup",
                          "Effect size of a lineup",
                          "Coventional p-value",
                          "Visual p-value",
                          "Reason for the selection",
                          "How different the selected plot from others",
                          "Age group of the subject",
                          "Education background of the subject",
                          "Pronoun",
                          "Whether the subject has previous experience in experiment about reading data plots",
                          "Type of the lineup",
                          "Formula of the data genearting processing",
                          "Parameter `shape` of the polynomial model",
                          "Parameter `a` of the heteroskedasticity model",
                          "Parameter `b` of the heteroskedasticity model",
                          "Distribution of the regressor",
                          "Distribution of the error term",
                          "Standard deviation of the polynomial model",
                          "Whether to include polynomial terms",
                          "Name of the model",
                          "Number of plots in a lineup",
                          "Number of observations in a plot",
                          "c-interesting plots in a lineup",
                          "Estimated alpha of the Dirichelt model",
                          "Sum of squre error of the alpha estimate",
                          "Expected effect size")), col.names = c("Variable", "Description")) %>%
  kableExtra::kable_material_dark(bootstrap_options = c("striped", "hover"))
```

```{r}
null_lineups <- filter(vi_survey, b == 0)
vi_survey <- filter(vi_survey, lineup_id <= 576) %>%
  mutate(subject = as.numeric(factor(interaction(exp, set))))
```

The survey data has `r nrow(vi_survey)` lineup evaluations for `r length(unique(vi_survey$unique_lineup_id))` unique lineups evaluated by `r length(unique(interaction(vi_survey$exp, vi_survey$set)))` subjects. 
Experiment I has `r nrow(vi_survey[vi_survey$exp == 3, ])` lineup evaluations for `r length(unique(filter(vi_survey, exp == 3)$unique_lineup_id))` unique lineups evaluated by `r length(unique(filter(vi_survey, exp == 3)$set))` subjects. 
Experiment II has `r nrow(vi_survey[vi_survey$exp == 4, ])` lineup evaluations for `r length(unique(filter(vi_survey, exp == 4)$unique_lineup_id))` unique lineups evaluated by `r length(unique(filter(vi_survey, exp == 4)$set))` subjects. 
Experiment III has `r nrow(vi_survey[vi_survey$exp == 5, ])` lineup evaluations for `r length(unique(filter(vi_survey, exp == 5)$unique_lineup_id))` unique lineups evaluated by `r length(unique(filter(vi_survey, exp == 5)$set))` subjects.
There are `r nrow(filter(vi_survey, b == 0))` lineup evaluations for `r length(unique(filter(vi_survey, b == 0)$lineup_id))` Rorschach lineups.

## Demographic

The education background of most participants have Diploma or Bachelor degrees, followed by High school or below. The survey data is gender balanced. Most of the participants are between 18 to 39.

```{r}
vi_survey %>%
  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
  ggplot() +
  geom_bar(aes(age_group, fill = pronoun), position = "dodge") +
  facet_wrap(~fct_relevel(education, "High School or below", "Diploma and Bachelor Degree", "Honours Degree", "Masters Degree")) +
  xlab("Age group") +
  ylab("Count")
```

Number of participants who have previous experience in experiment about reading data plots are not very different from the number of participants who haven't. Age distributions are also similar for these two groups.

```{r}
vi_survey %>%
  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
  ggplot() +
  geom_bar(aes(age_group, fill = pronoun), position = "dodge") +
  facet_wrap(~previous_experience) +
  xlab("Age group") +
  ylab("Count")
```

## Response time

The distribution of the log response time is slightly right skewed.

```{r}
vi_survey %>%
  ggplot() +
  geom_density(aes(response_time/1000)) +
  scale_x_log10() +
  xlab("Response time (seconds)") +
  ggtitle("Distribution of response time in log scale")
```

Subjects spend less time as the experiment progress.

```{r}
vi_survey %>%
  ggplot() +
  geom_boxplot(aes(num, response_time/1000, group = num)) +
  xlab("#lineup") +
  ylab("Response time (seconds)") +
  scale_y_log10() +
  ggtitle("Boxplots of response time in log scale vs the order of lienup")
```

The minimum response time is `r min(vi_survey$response_time)/1000` seconds given by subject ``r vi_survey$subject[which.min(vi_survey$response_time)]`` while evaluating lineup ``r vi_survey$unique_lineup_id[which.min(vi_survey$response_time)]``. This subject get it `r c("wrong", "correct")[vi_survey$detect[which.min(vi_survey$response_time)] + 1]` with `r vi_survey$num_selection[which.min(vi_survey$response_time)]` selections. This is the #`r vi_survey$num[which.min(vi_survey$response_time)]` lineup presented to the subject. The total amount of time this subject spent on the study is `r round(sum(vi_survey$response_time[vi_survey$subject == vi_survey$subject[which.min(vi_survey$response_time)]])/1000, 2)` seconds.

The maximum response time is `r max(vi_survey$response_time)/1000` seconds given by subject ``r vi_survey$subject[which.max(vi_survey$response_time)]`` while evaluating lineup ``r vi_survey$unique_lineup_id[which.max(vi_survey$response_time)]``. This subject get it `r c("wrong", "correct")[vi_survey$detect[which.max(vi_survey$response_time)] + 1]` with `r vi_survey$num_selection[which.max(vi_survey$response_time)]` selections. This is the #`r vi_survey$num[which.max(vi_survey$response_time)]` lineup presented to the subject. The total amount of time this subject spent on the study is `r round(sum(vi_survey$response_time[vi_survey$subject == vi_survey$subject[which.max(vi_survey$response_time)]])/1000, 2)` seconds.

The median response time is `r median(vi_survey$response_time)/1000` seconds.

The median total response time per subject is `r vi_survey %>% group_by(subject) %>% summarise(total_response_time = sum(response_time)) %>% {median(.$total_response_time)/1000}` seconds. Four subjects spent more time than others in the study. Subject `11` spent a long time on the first lineup. Subject `163`, `176` and `246` consistently evaluates lineups in a slow pace.

```{r}
vi_survey %>%
  group_by(subject) %>%
  summarise(total_response_time = sum(response_time)/1000) %>%
  ggplot() +
  geom_boxplot(aes(total_response_time)) +
  ggrepel::geom_label_repel(data = ~tail(arrange(., total_response_time), 4), aes(total_response_time, 0, label = subject)) +
  scale_x_log10() +
  xlab("Total response time per subject") +
  ggtitle("Total response time per subject in log scale")
```

```{r}
vi_survey %>%
  filter(subject %in% c(163, 176, 11, 246)) %>%
  ggplot() +
  geom_line(aes(num, response_time/1000, col = factor(subject))) +
  labs(col = "subject") +
  scale_y_log10() +
  xlab("#lineup") +
  ylab("Response time (seconds)") +
  ggtitle("Response time in log scale vs the order of lienup")
```

```{r}
vi_survey %>%
  filter(subject %in% c(163, 176, 11, 246)) %>%
  select(subject, num, response_time) %>%
  mutate(response_time = response_time/1000) %>%
  pivot_wider(names_from = num, values_from = response_time) %>%
  select(subject, `1`, `2`, `3`, `4`, `5`, `6`, `7`, `8`, `9`, `10`, `11`, `12`, `13`, `14`, `15`, `16`, `17`, `18`, `19`, `20`) %>%
  arrange(subject) %>%
  knitr::kable("markdown")
```

When the subject detect the data plot, the response time is usually lower. This is because it is easier to make decision on lineup with obvious patterns.

```{r}
vi_survey %>%
  ggplot() +
  geom_density(aes(response_time/1000, col = detect)) +
  scale_x_log10() +
  facet_wrap(~type) +
  xlab("Response time (seconds)") +
  ylab("Density") +
  ggtitle("Distribution of response time in log scale")
```

There is a very weak negative relationship between the difficulty of the lineup and the response time. There are some cases where the lineup is very difficult, but the subject evaluates the lineup in a short time. But there are no obvious outliers except for one we have analysed.

```{r}
vi_survey %>%
  ggplot() +
  geom_point(aes(effect_size, response_time/1000), alpha = 0.1) +
  geom_smooth(aes(effect_size, response_time/1000)) +
  scale_y_log10() +
  scale_x_log10() +
  facet_grid(detect~type) +
  xlab("Effect size") +
  ylab("Response time (seconds)") +
  ggtitle("Response time in log scale vs effect size in log scale")
```

For Rorschach lineup, the response time varies a lot, from few seconds to several minutes.

```{r}
vi_survey %>%
  filter(b == 0) %>%
  ggplot() +
  geom_density(aes(response_time/1000)) +
  scale_x_log10() +
  xlab("Response time (seconds)") +
  ylab("Count") +
  ggtitle("Response time for Rorschach lineup in log scale") +
  facet_wrap(~detect)
```

Number of selection made by the subjects seems to be correlated to the response time. Subjects tend to spend more time on a lineup if they select more plots, which is reasonable.

```{r}
vi_survey %>%
  ggplot() +
  geom_boxplot(aes(num_selection, response_time/1000, group = factor(num_selection, levels = 0:20))) +
  geom_smooth(aes(num_selection, response_time/1000)) +
  scale_y_log10() +
  scale_x_continuous(breaks = 0:20, minor_breaks = NULL) +
  facet_grid(type~detect) +
  xlab("Number of selections") +
  ylab("Response time (seconds)") +
  ggtitle("Response time in log scale vs number of selection faceted by detection and type of model")
```

When subjects correctly identify the actual data plot, they tend to spend less time if they discover a shape from the plot, followed by cluster(s) and outlier(s). Subjects spend much more time for reason "Others", which they need to manually provide a custom reason. This could be due to the fact that they need to spend time in typing the reason. However, we do not see the same difference when they do not detect the actual data plot. The median response time are similar for different reasons.

```{r}
vi_survey %>%
  mutate(reason = ifelse(reason %in% c("Shape", "Cluster(s)", "Outlier(s)"), reason, "Others")) %>%
  mutate(reason = factor(reason, levels = c("Shape", "Cluster(s)", "Outlier(s)", "Others"))) %>%
  ggplot() +
  geom_boxplot(aes(reason, response_time/1000)) +
  scale_y_log10() +
  facet_grid(type~detect) +
  xlab("Reason to select the plot") +
  ylab("Response time (seconds)") +
  ggtitle("Response time in log scale vs reason faceted by detection and type of model")
```

When subjects correctly identify the actual data plot, the response time decreases as the confidence level increases. When subjects do not detect the actual data plot, they tend to spend less time if they do not find anything particularly interesting for the heteroskedasticity model. Curiously, they spend more time even though they think the selected plot is extremely different to others for the polynomial model. This could be due to the small sample size (seven). And three out of seven evaluations are given by the same subject `202`.

```{r}
vi_survey %>%
  ggplot() +
  geom_boxplot(aes(factor(confidence, levels = c("Not at all", "Slightly", "Moderately", "Very", "Extremely")), response_time/1000)) +
  scale_y_log10() +
  facet_grid(type~detect) +
  xlab("Confidence") +
  ylab("Response time (seconds)") +
  ggtitle("Response time in log scale vs confidence faceted by detection and type of model")
```

We do not observe any particular structure for the distribution of the log of the response time when the answer is placed in different panels.

```{r}
vi_survey %>%
  ggplot() +
  geom_boxplot(aes(answer, response_time/1000, group = answer)) +
  scale_y_log10() +
  facet_grid(type~detect) +
  xlab("Answer") +
  ylab("Response time (seconds)") +
  ggtitle("Response time in log scale vs answer faceted by detection and type of model")
```

It is interesting to see that the response time is greater for older subjects in regards of the type of the model and whether the answer is correct.

```{r}
vi_survey %>%
  ggplot() +
  geom_boxplot(aes(age_group, response_time/1000)) +
  scale_y_log10() +
  facet_grid(type~detect) +
  xlab("Age group") +
  ylab("Response time (seconds)") +
  ggtitle("Response time in log scale vs age group faceted by detection and type of model")
```

Female subjects tend to spend little more time than male in our experiments. The "40-54" and "55-64" age groups show larger difference between male and female.

```{r}
vi_survey %>%
  ggplot() +
  geom_boxplot(aes(age_group, response_time/1000, col = ifelse(pronoun == "They", "Other", pronoun))) +
  scale_y_log10() +
  labs(col = "pronoun") +
  xlab("Age group") +
  ylab("Response time (seconds)") +
  ggtitle("Response time in log scale vs age group colored by pronoun")
```

The same upward trend of response time maintains for different education groups.

```{r}
vi_survey %>%
  ggplot() +
  geom_boxplot(aes(age_group, response_time/1000)) +
  scale_y_log10() +
  facet_wrap(~fct_relevel(education, "High School or below", "Diploma and Bachelor Degree", "Honours Degree", "Masters Degree")) +
  labs(col = "pronoun") +
  xlab("Age group") +
  ylab("Response time (seconds)") +
  ggtitle("Response time in log scale vs age group faceted by pronoun")
```

However, notice that some groups have only few observations. Thus, the trend might not be reliable.

```{r}
vi_survey %>%
  group_by(age_group, education) %>%
  summarise(num_subject = length(unique(subject))) %>%
  ggplot() +
  geom_bar(aes(age_group, num_subject), stat = "identity") +
  facet_wrap(~fct_relevel(education, "High School or below", "Diploma and Bachelor Degree", "Honours Degree", "Masters Degree"))
```

The median response time of the group "Doctoral Degree:FALSE:polynomial" is higher than others. It only contains evaluations given by two subjects. Apart from that, we do not observe any trend from the plot.

```{r}
vi_survey %>%
  ggplot() +
  geom_boxplot(aes(fct_relevel(education, "High School or below", "Diploma and Bachelor Degree", "Honours Degree", "Masters Degree"), response_time/1000)) +
  scale_y_log10() +
  facet_grid(type~detect) + 
  xlab("Education background") +
  ylab("Response time (seconds)") +
  ggtitle("Response time in log scale vs education background faceted by detection and type of model")
```

Subjects with previous experience doesn't seem like to evaluate lineups faster.

```{r}
vi_survey %>%
  ggplot() +
  geom_boxplot(aes(previous_experience, response_time/1000)) +
  scale_y_log10() +
  facet_grid(type~detect) + 
  xlab("Previous experience") +
  ylab("Response time (seconds)") +
  ggtitle("Response time in log scale vs previous experience faceted by detection and type of model")
```

For polynomial model, different shapes do not affect the response time much.

```{r}
vi_survey %>%
  filter(type == "polynomial") %>%
  ggplot() +
  geom_boxplot(aes(factor(shape), response_time/1000)) +
  scale_y_log10() +
  facet_grid(~detect) +
  xlab("Shape") +
  ylab("Response time (seconds)") +
  ggtitle("Response time in log scale vs shape for polynomial model")
```

Different distributions of $x$ slightly affect the response time. The response time for uniform distribution is lower when the subject correctly identify the actual data plot. Additionally, the response time for lognormal distribution is higher than others. This could be due to uniform distribution has a higher chance to reveal the underlying visual pattern, and lognormal distribution is the opposite.

```{r}
vi_survey %>%
  filter(type == "polynomial") %>%
  ggplot() +
  geom_boxplot(aes(factor(x_dist, levels = c("uniform", "even_discrete", "normal", "lognormal")), response_time/1000)) +
  scale_y_log10() +
  facet_grid(~detect) +
  xlab("Distribution of x") +
  ylab("Response time (seconds)") +
  ggtitle("Response time in log scale vs distribution of x for polynomial model faceted by detection")
```

It is interesting to see that the response times are not much different across different $\sigma$ values when the subject can not detect the data plot, except for $n = 300$. For $n = 300$, the response time is lower for lower noise level, indicating that when the visual pattern is visible and detectable, subjects will spend less time in making the decision. When the subject correctly identify the actual data plot, the same trend occurs in clearer manner.

```{r}
vi_survey %>%
  filter(type == "polynomial") %>%
  ggplot() +
  geom_boxplot(aes(factor(e_sigma), response_time/1000)) +
  scale_y_log10() +
  facet_grid(n~detect) +
  xlab("Noise level (sigma)") +
  ylab("Response time (seconds)") +
  ggtitle("Response time in log scale vs noise level for polynomial model faceted by detection")
```

We do not observe any interesting pattern from the response time vs alpha plot.

```{r}
vi_survey %>%
  filter(type == "polynomial") %>%
  mutate(alpha = jitter(alpha)) %>%
  ggplot() +
  geom_point(aes(alpha, response_time/1000)) +
  geom_smooth(aes(alpha, response_time/1000)) +
  scale_y_log10() +
  xlab("Estimated alpha") +
  ylab("Response time (seconds)") +
  ggtitle("Response time in log scale vs estimated alpha for polynomial model")
```

For heteroskedasticity model, different $a$ and distribution of $x$ do not affect the response time much. Except for the uniform group, which has slightly lower response time.

```{r}
vi_survey %>%
  filter(type == "heteroskedasticity") %>%
  ggplot() +
  geom_boxplot(aes(factor(a), response_time/1000)) +
  facet_grid(~factor(x_dist, levels = c("uniform", "even_discrete", "normal", "lognormal"))) +
  scale_y_log10() +
  xlab("a") +
  ylab("Response time (seconds)") +
  ggtitle("Response time in log scale vs a for heteroskedasticity model faceted by distribution of x")
```

As $b$ increases, when the subject correctly identify the actual data plot, the response time will decrease for $n = 100$ and $n = 300$.

```{r}
vi_survey %>%
  filter(type == "heteroskedasticity") %>%
  ggplot() +
  geom_boxplot(aes(factor(b), response_time/1000, group = b)) +
  scale_y_log10() +
  facet_grid(detect~n) +
  xlab("b") +
  ylab("Response time (seconds)") +
  ggtitle("Response time in log scale vs b for heteroskedasticity model faceted by number of observations and detection")
```

Similar to the polynomial model, $\alpha$ does not affect the response time.

```{r}
vi_survey %>%
  filter(type == "heteroskedasticity") %>%
  mutate(alpha = jitter(alpha)) %>%
  ggplot() +
  geom_point(aes(alpha, response_time/1000)) +
  geom_smooth(aes(alpha, response_time/1000)) +
  scale_y_log10() +
  xlab("Estimated alpha") +
  ylab("Response time (seconds)") +
  ggtitle("Response time in log scale vs estimated alpha for heteroskedasticity model")
```

## Selection

```{r eval = FALSE}
vi_survey %>%
  rowwise() %>%
  mutate(p_value_rank = list((function(this_lineup) { 
    vi_lineup[[this_lineup]]$data %>%
      group_by(k) %>%
      summarise(p_value = first(p_value)) %>%
      arrange(k) %>%
      pull(p_value)
    })(unique_lineup_id))) %>%
  unnest_longer(p_value_rank) %>%
  View()
```

```{r}
vi_survey %>%
  filter(type == "heteroskedasticity") %>%
  ggplot() +
  geom_boxplot(aes(factor(a), response_time/1000)) +
  scale_y_log10()
```

```{r}
vi_survey %>%
  ggplot() +
  geom_point(aes(num_selection, log(response_time)))
```

```{r}
plot_lineup <- function(this_lineup) {
  VI_MODEL$plot_lineup(vi_lineup[[this_lineup]]$data, 
                       theme = theme_light(), 
                       remove_axis = TRUE, 
                       remove_legend = TRUE, 
                       remove_grid_line = TRUE,
                       add_zero_line = TRUE) -> p

  # x_mean <- mean(layer_scales(p)$x$range$range)
  # y_mean <- mean(layer_scales(p)$y$range$range)

  vi_survey %>%
    filter(unique_lineup_id == this_lineup) %>%
    # filter(unique_lineup_id == "heter_7") %>%
    mutate(selection = ifelse(num_selection == 0, paste(1:20, collapse = "_"), selection)) %>%
    mutate(num_selection = ifelse(num_selection == 0, 20, num_selection)) %>%
    (function(x) {
      for (i in 1:20)
        x[[paste0("plot_", i)]] <-
          grepl(paste0("_", i, "_"), x$selection) |
          grepl(paste0("^", i, "_"), x$selection) |
          grepl(paste0("_", i, "$"), x$selection) |
          grepl(paste0("^", i, "$"), x$selection)
      x}
    ) -> tmp_survey

  tmp_survey %>%
    group_by(unique_lineup_id) %>%
    summarise(across(plot_1:plot_20, sum)) %>%
    ungroup() %>%
    pivot_longer(plot_1:plot_20) %>%
    mutate(name = as.integer(gsub(".*_", "", name))) %>%
    rename(k = name, total_detect = value) -> plot_total_detect

  tmp_survey %>%
    mutate(across(plot_1:plot_20, function(x) x/num_selection)) %>%
    group_by(unique_lineup_id) %>%
    summarise(across(plot_1:plot_20, sum)) %>%
    ungroup() %>%
    pivot_longer(plot_1:plot_20) %>%
    mutate(name = as.integer(gsub(".*_", "", name))) %>%
    rename(k = name, weighted_selection = value) -> plot_weighted_selection

  k_col <- ifelse(plot_weighted_selection$weighted_selection == max(plot_weighted_selection$weighted_selection), 
                  "max_sel", 
                  ifelse(plot_weighted_selection$k == vi_lineup[[this_lineup]]$metadata$answer, 
                         "answer", 
                         "else")
                  )

  p +
    geom_polygon(data = merge(data.frame(k = 1:20,
                                         k_col = k_col),
                              data.frame(x = c(-Inf, Inf, Inf, -Inf),
                                         y = c(-Inf, -Inf, Inf, Inf))),
                 aes(x = x, y = y, col = k_col), fill = NA, size = 2) +
    geom_label(data = group_by(vi_lineup[[this_lineup]]$data, k) %>% 
                summarise(p_value = first(p_value)) %>%
                ungroup(),
              aes(x = -Inf, y = Inf, 
                  label = paste0(round(p_value, 2), " (", rank(p_value, ties.method = "min"), ")")), 
              hjust = 0, vjust = 1, alpha = 0.5, size = 2, label.padding = unit(0.1, "lines")) +
    scale_color_manual(values = c("max_sel" = "blue", "answer" = "red", "else" = "white")) +
    facet_wrap(~k, 
               labeller = as_labeller(
                 function(string, ...) 
                   paste(round(filter(plot_weighted_selection, k == as.integer(string))$weighted_selection, 3),
                         "/",
                         filter(plot_total_detect, k == as.integer(string))$total_detect)
                 )
               ) +
    ggtitle(paste0(this_lineup, 
                  "\neffect size: ", 
                  round(vi_lineup[[this_lineup]]$metadata$effect_size, 2),
                  ", visual p-value: ", 
                  round(filter(vi_survey, unique_lineup_id == this_lineup)$p_value[1], 2), 
                  ", conventional p-value: ",
                  round(filter(vi_survey, unique_lineup_id == this_lineup)$conventional_p_value[1], 2),
                  "\nshape: ",
                  filter(vi_survey, unique_lineup_id == this_lineup)$shape[1],
                  ", e_sigma: ",
                  filter(vi_survey, unique_lineup_id == this_lineup)$e_sigma[1],
                  ", x_dist: ",
                  filter(vi_survey, unique_lineup_id == this_lineup)$x_dist[1],
                  ", n: ",
                  filter(vi_survey, unique_lineup_id == this_lineup)$n[1]
                  ),
            subtitle =  paste0("Lineup evaluated by ",
                               vi_survey %>% filter(unique_lineup_id == this_lineup) %>% nrow(),
                               " subjects (c_i / total selections)"))

  # # Pop the last layers
  # tmp <- p$layers[[length(p$layers)]]
  # p$layers[[length(p$layers)]] <- NULL
  # 
  # # Append the element back to the list
  # p$layers <- append(list(tmp), p$layers)
  # 
  # p
} 
```

```{r fig.height=7, fig.width=7, dpi=150}
VI_MODEL$plot_lineup(vi_lineup$poly_1$data, theme = theme_light(), remove_axis = TRUE, remove_legend = TRUE, remove_grid_line = TRUE)
```

## Power

There are some cases where the proportion of detection does not match with the predicted power. When effect size < 10, there are points with power around 0.25, and when effect size > 10, there points with power lower than 0.25. Those will be the main focus.

```{r}
boot_y <- function(dat, y, times = 100) {
  map_df(1:times, function(i) slice_sample(dat, n = nrow(dat), replace = TRUE) %>% mutate(boot_id = i))
}

vi_survey %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(e_sigma >= 0.5) %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  ggplot() +
  geom_smooth(aes(effect_size, reject, col = "Visual"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  geom_point(aes(effect_size, prop_detect, text = unique_lineup_id), alpha = 0.3) + 
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject, times = 500), 
              aes(effect_size, reject, group = boot_id, col = "Visual"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
                filter(x_dist == "uniform") %>%
                rename(F = F_p_value, BP = BP_p_value, SW = SW_p_value) %>%
                pivot_longer(F:SW) %>%
                mutate(reject = as.numeric(value < 0.05)), 
              aes(effect_size, reject, col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  scale_x_log10() +
  xlab("Effect size") +
  ylab("Power") +
  ggtitle("Power of visual test with comparison to F, BP and SW test") -> p

plotly::ggplotly(p, tooltip = c("x", "y", "text"))
```

There are 10 lineups with effect size < 10 and power > 0.125, namely `r vi_survey %>% filter(x_dist == "uniform") %>% group_by(unique_lineup_id) %>% summarise(across(everything(), first)) %>% filter(type == "polynomial") %>% filter(effect_size < 10 & prop_detect > 0.125) %>% .$unique_lineup_id`.

Lineup `poly_108` does not have a clear visual pattern. Subjects claimed to rely on outliers to identify the actual data plot.

Lineup `poly_132` has a distracting plot positioning at panel 10, where subjects reported to see a shape from the plot. Those who correctly identifies the actual data plot suggest there are outliers in plot one, which is marginally true given there is a minor gap at the right of the plot.

Lineup `poly_159` has a clear outlier at plot one, which is the reason three subjects detect it. There is another subject who detect it because of "shape". This is kind of make sense as the outlier strengthen the shape.

Lineup `poly_268` has an interesting plot at panel six, which happens to be the most selected plot. The actual data plot does not contain any visual detectable features, but two subjects still select this one. One select it with reason "cluster" and the other one select it with "shape".

Lineup `poly_295`

```{r}
eval_info <- function(this_lineup) {
  vi_survey %>%
    filter(unique_lineup_id == this_lineup) %>%
    select(subject, selection, num_selection, detect, reason, confidence) %>%
    arrange(-detect, num_selection, reason, confidence) %>%
    gridExtra::tableGrob(rows = NULL, theme = gridExtra::ttheme_default(base_size = 8))
}
```

```{r fig.height=7, fig.width=14, dpi=150}

for (this_lineup in vi_survey %>% 
  filter(x_dist == "uniform") %>% 
  group_by(unique_lineup_id) %>% 
  summarise(across(everything(), first)) %>% 
  filter(type == "polynomial") %>% 
  filter(effect_size < 10 & prop_detect > 0.125) %>% 
  .$unique_lineup_id %>%
  sort()) {
  gridExtra::grid.arrange(plot_lineup(this_lineup), eval_info(this_lineup), ncol = 2)
}

```

There are 7 lineups with effect size > 35 and power < 0.25, namely `r vi_survey %>% filter(x_dist == "uniform") %>% group_by(unique_lineup_id) %>% summarise(across(everything(), first)) %>% filter(type == "polynomial") %>% filter(effect_size > 35 & prop_detect < 0.25) %>% .$unique_lineup_id`.

```{r fig.height=7, fig.width=14, dpi=150}

for (this_lineup in vi_survey %>% 
     filter(x_dist == "uniform") %>% 
     group_by(unique_lineup_id) %>% 
     summarise(across(everything(), first)) %>% 
     filter(type == "polynomial") %>% 
     filter(effect_size > 35 & prop_detect < 0.25) %>% 
     .$unique_lineup_id %>%
  sort()) {
  gridExtra::grid.arrange(plot_lineup(this_lineup), eval_info(this_lineup), ncol = 2)
}

```

## Replace effect size with expected effect size

```{r}
vi_survey %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(e_sigma >= 0.5) %>%
  ggplot() +
  geom_point(aes(effect_size, expected_effect_size)) +
  geom_abline()
```

```{r}
boot_y <- function(dat, y, times = 100) {
  map_df(1:times, function(i) slice_sample(dat, n = nrow(dat), replace = TRUE) %>% mutate(boot_id = i))
}

vi_survey %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(e_sigma >= 0.5) %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  ggplot() +
  geom_smooth(aes(expected_effect_size, reject, col = "Visual"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  geom_point(aes(expected_effect_size, prop_detect, text = unique_lineup_id), alpha = 0.3) + 
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject, times = 500), 
              aes(expected_effect_size, reject, group = boot_id, col = "Visual"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
                filter(x_dist == "uniform") %>%
                rename(F = F_p_value, BP = BP_p_value, SW = SW_p_value) %>%
                pivot_longer(F:SW) %>%
                mutate(reject = as.numeric(value < 0.05)), 
              aes(effect_size, reject, col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  scale_x_log10() +
  xlab("Effect size") +
  ylab("Power") +
  ggtitle("Power of visual test with comparison to F, BP and SW test") -> p

plotly::ggplotly(p, tooltip = c("x", "y", "text"))
```

### 2022-11-02 paper_comparison/analysis/vi_survey_analysis.qmd
update analysis

@@
 
-```{r eval = FALSE}
-vi_survey %>%
-  rowwise() %>%
-  mutate(p_value_rank = list((function(this_lineup) { 
+```{r}
+sel_dist <- vi_survey %>%
+  group_by(unique_lineup_id) %>%
+  mutate(selection = ifelse(num_selection == 0, paste(1:20, collapse = "_"), selection)) %>%
+  mutate(num_selection = ifelse(num_selection == 0, 20, num_selection)) %>%
+    (function(x) {
+      for (i in 1:20)
+        x[[paste0("plot_", i)]] <-
+          grepl(paste0("_", i, "_"), x$selection) |
+          grepl(paste0("^", i, "_"), x$selection) |
+          grepl(paste0("_", i, "$"), x$selection) |
+          grepl(paste0("^", i, "$"), x$selection)
+      x}
+    ) %>%
+  mutate(across(plot_1:plot_20, function(x) x/num_selection)) %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(plot_1:plot_20, sum)) %>%
+  ungroup() %>%
+  pivot_longer(plot_1:plot_20) %>% 
+  left_join(vi_survey %>% group_by(unique_lineup_id) %>% summarise(answer = first(answer))) %>%
+  left_join(vi_survey %>%
+  group_by(unique_lineup_id) %>%
+  summarise(p_value = list((function(this_lineup) {
+    this_lineup <- this_lineup[1]
     vi_lineup[[this_lineup]]$data %>%
@@
       arrange(k) %>%
-      pull(p_value)
+      pull(p_value) -> x
+    names(x) <- paste0("plot_", 1:20)
+    x
   })(unique_lineup_id))) %>%
-  unnest_longer(p_value_rank) %>%
-  View()
+  unnest_longer(p_value) %>%
+  rename(name = p_value_id)) %>%
+  group_by(unique_lineup_id) %>%
+  mutate(prop = value/sum(value)) %>%
+  mutate(p_value_rank = rank(p_value, ties.method = "first")) %>%
+  rename(c_i = value) %>%
+  rename(plot = name) %>%
+  ungroup()
+```
+
+```{r}
+sel_dist %>%
+  filter(paste0("plot_", answer) == plot) %>%
+  select(unique_lineup_id, answer_p_value_rank = p_value_rank, answer_prop = prop) %>%
+  left_join(sel_dist %>%
+  group_by(unique_lineup_id) %>%
+  summarise(most_p_value_rank = min(p_value_rank[which(c_i == max(c_i))]),
+            most_prop = prop[which.min(p_value_rank[which(c_i == max(c_i))])])) %>%
+  ggplot() +
+  geom_boxplot(aes(answer_p_value_rank, answer_prop, group = answer_p_value_rank)) +
+  xlab("Rank of the p-value of the data plot in a lineup") +
+  ylab("Percentage of subject picking the data plot") +
+  ggtitle("")
+```
+
+```{r}
+sel_dist %>%
+  filter(paste0("plot_", answer) == plot) %>%
+  select(unique_lineup_id, answer_p_value_rank = p_value_rank, answer_prop = prop) %>%
+  left_join(sel_dist %>%
+  group_by(unique_lineup_id) %>%
+  summarise(most_p_value_rank = min(p_value_rank[which(c_i == max(c_i))]),
+            most_prop = prop[which.min(p_value_rank[which(c_i == max(c_i))])])) %>%
+  ggplot() +
+  geom_boxplot(aes(most_p_value_rank, most_prop, group = most_p_value_rank)) +
+  xlab("Rank of the p-value of the most selected plot in a lineup") +
+  ylab("Percentage of subject picking the most selected plot") +
+  ggtitle("")
 ```
@@
 
+```{r}
+vi_survey %>%
+  filter(x_dist == "uniform") %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(type == "polynomial") %>%
+  filter(e_sigma >= 0.5) %>%
+  mutate(reject = as.numeric(p_value <= 0.05)) %>%
+  ggplot() +
+  geom_smooth(aes(effect_size, reject, col = "Visual"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  geom_text(aes(effect_size, prop_detect, text = unique_lineup_id, label = shape), alpha = 0.3) + 
+  stat_smooth(geom = "line", 
+              data = ~boot_y(.x, reject, times = 500), 
+              aes(effect_size, reject, group = boot_id, col = "Visual"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, 
+              alpha = 0.03) +
+  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
+                filter(x_dist == "uniform") %>%
+                rename(F = F_p_value, BP = BP_p_value, SW = SW_p_value) %>%
+                pivot_longer(F:SW) %>%
+                mutate(reject = as.numeric(value < 0.05)), 
+              aes(effect_size, reject, col = name), 
+              method = "glm", 
+              method.args = list(family = binomial),
+              se = FALSE) +
+  scale_x_log10() +
+  xlab("Effect size") +
+  ylab("Power") +
+  ggtitle("Power of visual test with comparison to F, BP and SW test")
+```
+
+```{r}
+vi_survey %>%
+  filter(x_dist == "uniform") %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(type == "polynomial") %>%
+  filter(e_sigma >= 0.5) %>%
+  mutate(reject = as.numeric(p_value <= 0.05)) %>%
+  ggplot() +
+  geom_smooth(aes(effect_size, reject, col = "Visual"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  geom_text(aes(effect_size, prop_detect, text = unique_lineup_id, label = e_sigma), alpha = 0.3) + 
+  stat_smooth(geom = "line", 
+              data = ~boot_y(.x, reject, times = 500), 
+              aes(effect_size, reject, group = boot_id, col = "Visual"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, 
+              alpha = 0.03) +
+  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
+                filter(x_dist == "uniform") %>%
+                rename(F = F_p_value, BP = BP_p_value, SW = SW_p_value) %>%
+                pivot_longer(F:SW) %>%
+                mutate(reject = as.numeric(value < 0.05)), 
+              aes(effect_size, reject, col = name), 
+              method = "glm", 
+              method.args = list(family = binomial),
+              se = FALSE) +
+  scale_x_log10() +
+  xlab("Effect size") +
+  ylab("Power") +
+  ggtitle("Power of visual test with comparison to F, BP and SW test")
+```
+
+```{r}
+vi_survey %>%
+  filter(x_dist == "uniform") %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(type == "polynomial") %>%
+  filter(e_sigma >= 0.5) %>%
+  mutate(reject = as.numeric(p_value <= 0.05)) %>%
+  ggplot() +
+  geom_smooth(aes(effect_size, reject, col = "Visual"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  geom_text(aes(effect_size, prop_detect, text = unique_lineup_id, label = n), alpha = 0.3) + 
+  stat_smooth(geom = "line", 
+              data = ~boot_y(.x, reject, times = 500), 
+              aes(effect_size, reject, group = boot_id, col = "Visual"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, 
+              alpha = 0.03) +
+  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
+                filter(x_dist == "uniform") %>%
+                rename(F = F_p_value, BP = BP_p_value, SW = SW_p_value) %>%
+                pivot_longer(F:SW) %>%
+                mutate(reject = as.numeric(value < 0.05)), 
+              aes(effect_size, reject, col = name), 
+              method = "glm", 
+              method.args = list(family = binomial),
+              se = FALSE) +
+  scale_x_log10() +
+  xlab("Effect size") +
+  ylab("Power") +
+  ggtitle("Power of visual test with comparison to F, BP and SW test")
+```
+
 There are 10 lineups with effect size < 10 and power > 0.125, namely `r vi_survey %>% filter(x_dist == "uniform") %>% group_by(unique_lineup_id) %>% summarise(across(everything(), first)) %>% filter(type == "polynomial") %>% filter(effect_size < 10 & prop_detect > 0.125) %>% .$unique_lineup_id`.

### 2022-11-07 Milestone_1/01-chap1.Rmd
clean up repo

@@
 library(tidyverse)
+library(bandicoot)
 library(visage)
@@
 Models trained with online learning can be updated immediately for any new arrival data, which overcomes the drawbacks of the traditional batch learning where training data required to be available prior to the training.
-It is suitable for large-scale machine learning tasks in real-world data analytics applications [@].
+It is suitable for large-scale machine learning tasks in real-world data analytics applications [@hoi_online_2018].
 

### 2022-11-07 paper_comparison/scripts/estimate_alpha.R
clean up repo again

@@ -1,6 +1,8 @@
 library(visage)
 library(tidyverse)
 
+# When we estimate alpha, we use ceiling(c)
+
 
 calc_alpha <- function(zero_sel = 0.05) {
   vi_survey %>%
@@ -85,3 +87,9 @@ vi_survey %>%
   geom_smooth(aes(log_effect_size, reject_dirichlet), method = "glm", method.args = list(family = binomial), se = FALSE)
 
 
+
+ggplot() +
+  geom_point(aes(vi_survey_processed$p_value_select_all_plots, visage::vi_survey$p_value)) +
+  geom_abline() +
+  scale_x_log10() +
+  scale_y_log10()

### 2022-11-07 paper_comparison/paper_comparison/paper_comparison.Rmd
update paper rmd

@@
               se = FALSE) +
-  theme_light(base_size = 5) -> p
+  theme_light(base_size = 5) +
+  facet_wrap(~n) -> p
 
@@
 
-```{r eval = FALSE}
+```{r eval = FALSE, fig.height= 3, fig.width=4}
 vi_survey %>%
@@
   filter(n == 100)
+
+p
+
+vi_survey %>%
+  filter(x_dist == "uniform") %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(type == "polynomial") %>%
+  filter(e_sigma != 0.125) %>%
+  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  filter(abs(log(effect_size) - 4) < 0.5) %>%
+  filter(prop_detect < 0.25)
+
+vi_survey %>%
+  filter(unique_lineup_id == "poly_199") %>%
+  View()
+
+VI_MODEL$plot_lineup(vi_lineup$poly_159$data)
 
@@
   scale_y_log10()
+
+vi_survey %>%
+  filter(x_dist == "uniform") %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(type == "polynomial") %>%
+  filter(e_sigma != 0.125) %>%
+  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  ggplot() +
+  ggbeeswarm::geom_quasirandom(aes(factor(e_sigma), weighted_detect)) +
+  facet_wrap(~n)
 ```

### 2022-11-07 Fourth_experiment/generate_lineup.R
modify the lineup gen script due to the change of effect size function

@@ -33,7 +33,7 @@ for (i in 1:nrow(lineup)) {
                                               name = glue::glue("heter_{lineup$lineup_id[i]}"),
                                               k = 20,
                                               n = lineup$n[i],
-                                              effect_size = mod$effect_size(filter(lineup_dat, null == FALSE), type = "kl"),
+                                              effect_size = mod$sample_effect_size(filter(lineup_dat, null == FALSE), type = "kl"),
                                               ans = filter(lineup_dat, null == FALSE)$k[1]
 
   )

### 2022-11-07 Third_experiment/generate_lineup.R

@@ -33,7 +33,7 @@ for (i in 1:nrow(lineup)) {
                                               name = glue::glue("poly_{lineup$lineup_id[i]}"),
                                               k = 20,
                                               n = lineup$n[i],
-                                              effect_size = mod$effect_size(filter(lineup_dat, null == FALSE)),
+                                              effect_size = mod$sample_effect_size(filter(lineup_dat, null == FALSE)),
                                               ans = filter(lineup_dat, null == FALSE)$k[1]
                                               )
 

### 2022-11-08 paper_comparison/paper_comparison/paper_comparison.Rmd
provide better statement in p-value comparison

@@
 editor_options: 
-  chunk_output_type: inline
+  chunk_output_type: console
 ---
@@
 
-```{=html}
-<!--Regression diagnostics is an essential step in regression analysis which is a field of study with at least a hundred years of history.
-The diagnostic procedure conventionally involves evaluating the fitness of the proposed model, detecting the presence of influential observations and outliers, checking the validity of model assumptions and many more.
-In terms of diagnostic techniques, data plots, hypothesis testing, and summary statistics are vital tools for a systematic and detailed examination of the regression model [@mansfield1987diagnostic].
-
-Many of those regression diagnostic methods and procedures are mature and well-established in books first published in the twentieth century, such as @draper1998applied, @montgomery1982introduction, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals.
-Regardless of the level of difficulty, one will find the importance and usefulness of diagnostic plots being emphasized in those books repeatedly.
-Checking diagnostic plots is also the recommended starting point for validating model assumptions such as normality, homoscedasticity and linearity [@anscombe_examination_1963].
--->
-```
 # Background
@@
 
-<!-- test formula -> appendix -->
-
-<!-- residual plot -> plot -->
-
-<!-- plot a included -->
-
 ```{r}
@@
 A summary of the factors is given in Table \ref{tab:model-factor-table}.
-
-<!-- A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}. -->
 
@@
 
-<!-- ```{r results='asis'} -->
-<!-- data.frame(n = c(50, 100, 300, ""), -->
-<!--            j = c(2, 3, 6, 18), -->
-<!--            sigma = c(0.5, 1, 2, 4), -->
-<!--            x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$")) %>% -->
-<!--   knitr::kable(col.names = NULL, -->
-<!--                escape = FALSE, -->
-<!--                format = "latex", -->
-<!--                booktabs = TRUE,  -->
-<!--                caption  = 'Parameter values for $n$, $j$ $\\sigma$, $X_{raw}$',  -->
-<!--                label = "parameter-table" ) %>% -->
-<!--   kableExtra::add_header_above(c("Sample size\n(sym3)", "Order of Hermite polynomial\n(sym1)", "Error SD\n(sym2)", "Distribution of sym4\n")) %>% -->
-<!--   as.character() %>% -->
-<!--   gsub("sym1", "$j$", .) %>% -->
-<!--   gsub("sym2", "$\\\\sigma$", .) %>% -->
-<!--   gsub("sym3", "$n$", .) %>% -->
-<!--   gsub("sym4", "$X_{raw}$", .) %>% -->
-<!--   cat() -->
-<!--   # gsub("\\\\\\$", "$", .) %>% -->
-<!--   # cat() -->
-<!-- ``` -->
-
 ### Heteroskedasticity
@@
 
-Subjects recruited from Prolific received a fixed payment for participating in the experiment.
-However, some subjects will try to maximize their earnings for minimum effort.
-During the review of submissions, if we found a subject objectively demonstrated clear low-effort throughout the experiment, e.g., failed all attention checks, we rejected the submission.
-The rejected submissions will be removed immediately, and Prolific will automatically recruit another subject as substitution.
-Therefore, we only paid for approved submissions and no further data screening procedure needed to be applied on the collected data.
-
-In overall, there were a total of 8860 lineup evaluations made by 443 subjects in experiment I, experiment II and experiment III, where 886 lineup evaluations were attention checks and were not used in the following analysis.
-
-The collated dataset is provided in `vi_survey` of the `visage` `R` package.
+There were 2880, 2880 and 2214 lineups evaluation made by 160, 160 and 123 subjects recruited for experiment I, II and III respectively.
+In the total of 7974 lineup evaluations, 3744 used lineups produced by the non-linearity model, 3510 used lineups produced by the heteroskedasticity model, and 720 used lineups consists of only null plots.
+
+Besides, there were 886 attention checks not included in the following analysis. The collated dataset is provided in `vi_survey` of the `visage` `R` package.
 
@@
 Visual inference and conventional hypothesis testing are two distinct statistical procedures.
-As displayed in Figure \ref{fig:p-value-comparison}, the visual test almost always gives higher p-values than the conventional test in both polynomial and heteroskedasticity models, suggesting that humans have lower confidence and much higher tolerance of the residual departures than the numerical procedure.
-In fact, only around 58% of the cases, both tests will agree on the rejection decision given 95% confidence level.
-And in 41% of the cases, visual test does not reject the null hypothesis but conventional test does.
-Although we know that the all the designed lineups contain various degrees of residual departures and the conventional test is telling truth, analysts and general public as the consumers of the output will possibly not believe in or at least not response to the "false alarm" if they can not see the departures directly from the data plot and view it as impactful.
-The sensitivity of the conventional test may distract and discourage us from finding a good linear approximation to the data we try to model.
+As displayed in Figure \ref{fig:p-value-comparison}, the visual test gives higher p-values than the conventional test in both polynomial and heteroskedasticity models in most of the time, suggesting that humans have lower confidence and much higher tolerance of the residual departures than the conventional test.
+
+In particular, given 95% confidence level, both tests will agree on the rejection in only around 58% of the cases.
+And in 41% of the cases, the visual test does not reject the null hypothesis but the conventional test does.
+In these 41% of the cases, though we know that the designed lineups contain various degrees of residual departures, analysts and the general public as the consumers of the output may not believe in the rejection of the null hypothesis, or at least may not consider the model violation as impactful, simply due to the fact that the departures are not clearly visible to them.
+
+This makes it sounds like people are arrogantly ignoring the accuracy of conventional tests and trust too much to their visual perception, but it is not.
+In fact, the sensitivity of the conventional test could sometimes distract and discourage analysts from finding a good linear approximation to the data.
+The rejection of the null hypothesis because of acceptable and negligible residual departure is not practically useful and thus is usually viewed as "false alarm" by analysts.
+Relying on visual perception to determine the quality of the regression model is just a way to
+
+Alternatively,
+
+From a practical sense, data plots and the visual test provide better measure of the seriousness of the residual departure.
+
+One noticeable difference compared to the result provided in @buja_statistical_2009 is the lack of data points
+
+Note that
+
 This also partly explains why abundance of literature suggesting the use of data plot in regression diagnostics.
-
-```{r eval = FALSE}
-vi_survey %>%
-  filter(lineup_id <= 576) %>%
-  filter(is.na(b) | b != 0) %>%
-  mutate(test = p_value > 0.05 & conventional_p_value < 0.05) %>%
-  {mean(.$test)}
-
-vi_survey %>%
-  filter(lineup_id <= 576) %>%
-  filter(is.na(b) | b != 0) %>%
-ggplot() +
-  geom_histogram(aes(p_value), bins = 300) +
-  scale_x_sqrt() +
- facet_wrap(~type)
-```
 
@@
 vi_survey %>%
-  filter(lineup_id <= 576) %>%
-  filter(is.na(b) | b != 0) %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
   group_by(unique_lineup_id) %>%
   summarise(across(everything(), first)) %>%
-  # filter(x_dist == "uniform") %>%
   ggplot() +
-  geom_point(aes(p_value, conventional_p_value), alpha = 0.4) +
   geom_abline() +
+  geom_point(aes(conventional_p_value, p_value), alpha = 0.4) +
   scale_x_sqrt() +
   scale_y_sqrt() +
-  facet_wrap(~type)
+  facet_wrap(~type) +
+  ylab("Visual test p-value on square root scale") +
+  xlab("Conventional test p-value on square root scale")
 ```

### 2022-11-08 paper_comparison/paper_comparison/paper_comparison.Rmd
finish p-value comp

@@
 
-In particular, given 95% confidence level, both tests will agree on the rejection in only around 58% of the cases.
+In particular, given 95% confidence level, both tests will reject the null hypothesis in only around 58% of the cases.
 And in 41% of the cases, the visual test does not reject the null hypothesis but the conventional test does.
@@
 In fact, the sensitivity of the conventional test could sometimes distract and discourage analysts from finding a good linear approximation to the data.
-The rejection of the null hypothesis because of acceptable and negligible residual departure is not practically useful and thus is usually viewed as "false alarm" by analysts.
-Relying on visual perception to determine the quality of the regression model is just a way to
-
-Alternatively,
-
-From a practical sense, data plots and the visual test provide better measure of the seriousness of the residual departure.
-
-One noticeable difference compared to the result provided in @buja_statistical_2009 is the lack of data points
-
-Note that
-
+The rejection of the null hypothesis because of acceptable and negligible residual departure is not practically useful and thus is usually viewed as a "false alarm" by analysts.
 This also partly explains why abundance of literature suggesting the use of data plot in regression diagnostics.
+
+One noticeable difference compared to the result provided in @majumder_validation_2013 is we do not observe the pattern from Figure \ref{fig:p-value-comparison} that when the conventional test p-value is small, the visual test p-value will be very small, and when the conventional test p-value is large, the visual test p-value will be very large.
+This could because the number of subjects involved in a lineup evaluation is much larger in their experiments with 46 as the average number.
+As mentioned in section [Experimental setup], our lineups have at most 11 evaluations due to the time and budge constraint.
+Besides, evaluating a visual test with more than 15 people is rare given the size of research or business team.
 
@@
   filter(!attention_check) %>%
+  left_join(vi_survey %>% 
+           count(unique_lineup_id) %>%
+           rename(num_subject = n)) %>%
   group_by(unique_lineup_id) %>%
@@
   geom_point(aes(conventional_p_value, p_value), alpha = 0.4) +
+  geom_smooth(aes(conventional_p_value, p_value), se = FALSE, span = 0.6) +
   scale_x_sqrt() +
   scale_y_sqrt() +
-  facet_wrap(~type) +
+  facet_wrap(num_subject~type) +
   ylab("Visual test p-value on square root scale") +

### 2022-11-15 paper_comparison/paper_comparison/paper_comparison.Rmd
finish p-value cmp

--- 2022-11-08 paper_comparison/paper_comparison/paper_comparison.Rmd
@@
 
+                     es <- filter(vi_survey, 
+                                  shape == {{shape}},
+                                  e_sigma == {{e_sigma}},
+                                  x_dist == {{x_dist}},
+                                  n == {{n}}) %>%
+                       head(1) %>%
+                       pull(effect_size)
+
                      tibble(shape = shape,
@@
                             n = n,
-                            effect_size = mod$effect_size(tmp_dat),
+                            effect_size = es,
                             F_p_value = mod$test(tmp_dat)$p_value,
@@
 
+                     es <- filter(vi_survey, 
+                                  a == {{a}}, 
+                                  b == {{b}},
+                                  x_dist == {{x_dist}},
+                                  n == {{n}}) %>%
+                              head(1) %>%
+                              pull(effect_size)
+
                      tibble(a = a,
@@
                             n = n,
-                            effect_size = mod$effect_size(tmp_dat, type = "kl"),
+                            effect_size = es,
                             F_p_value = POLY_MODEL$test(mutate(tmp_dat, z = poly_model()$gen(n, computed = select(tmp_dat, x))$z))$p_value,
@@
 
-## P-value comparison
+## P-value and power comparison
 
 Visual inference and conventional hypothesis testing are two distinct statistical procedures.
-As displayed in Figure \ref{fig:p-value-comparison}, the visual test gives higher p-values than the conventional test in both polynomial and heteroskedasticity models in most of the time, suggesting that humans have lower confidence and much higher tolerance of the residual departures than the conventional test.
-
-In particular, given 95% confidence level, both tests will agree on the rejection in only around 58% of the cases.
+As displayed in Figure \ref{fig:p-value-comparison}, the visual test gives higher p-values than the conventional test for both polynomial and heteroskedasticity models in most of the time, suggesting that humans have lower confidence and much higher tolerance of the residual departures than the conventional test.
+
+In particular, given 95% confidence level, both tests will reject the null hypothesis in only around 58% of the cases.
 And in 41% of the cases, the visual test does not reject the null hypothesis but the conventional test does.
-In these 41% of the cases, though we know that the designed lineups contain various degrees of residual departures, analysts and the general public as the consumers of the output may not believe in the rejection of the null hypothesis, or at least may not consider the model violation as impactful, simply due to the fact that the departures are not clearly visible to them.
+Since fail to reject the null hypothesis in a visual test means that there is no obvious visual discoveries found in the residual plot, analysts and the general public as the consumers of the output may not fully believe in the rejection of the null hypothesis even though the alternative hypothesis is true.
+Even if the rejection is accepted, the model violation may not be considered as impactful due to the fact that the departures are not clearly visible.
 
 This makes it sounds like people are arrogantly ignoring the accuracy of conventional tests and trust too much to their visual perception, but it is not.
-In fact, the sensitivity of the conventional test could sometimes distract and discourage analysts from finding a good linear approximation to the data.
-The rejection of the null hypothesis because of acceptable and negligible residual departure is not practically useful and thus is usually viewed as "false alarm" by analysts.
-Relying on visual perception to determine the quality of the regression model is just a way to
-
-Alternatively,
-
-From a practical sense, data plots and the visual test provide better measure of the seriousness of the residual departure.
-
-One noticeable difference compared to the result provided in @buja_statistical_2009 is the lack of data points
-
-Note that
-
+In fact, the sensitivity of the conventional test could sometimes distract and discourage analysts from finding a simple but good linear approximation to the data.
+The rejection of the null hypothesis because of acceptable and negligible residual departure is not practically meaningful and useful.
 This also partly explains why abundance of literature suggesting the use of data plot in regression diagnostics.
 
-```{r p-value-comparison, fig.height=4, fig.cap="..."}
+```{r p-value-comparison, fig.height=4, fig.cap="Visual test p-value compared to conventional test p-value for lineups produced by heteroskedasticity and non-linearity model.
+The scatter plot is drawn on a square root scale.
+Every dot on the plot corresponds to a lineup.
+Most of the dots are fallen on the left of the y = x line indicating visual test p-value is almost always higher than conventional test p-value.
+The blue curve is a smoothing of the dots showing a positive trend.
+"}
 vi_survey %>%
@@
   filter(!attention_check) %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  left_join(vi_survey %>% 
+           count(unique_lineup_id) %>%
+           rename(num_subject = n)) %>%
   group_by(unique_lineup_id) %>%
@@
   geom_point(aes(conventional_p_value, p_value), alpha = 0.4) +
+  geom_smooth(aes(conventional_p_value, p_value), se = FALSE, span = 0.6) +
   scale_x_sqrt() +
@@
   ylab("Visual test p-value on square root scale") +
-  xlab("Conventional test p-value on square root scale")
-```
-
-## Power comparison
-
-Figure \ref{fig:power-poly-uniform} shows the estimated power of visual test on lineups with uniform distribution $X_{raw}$ against natural logarithm of the effect with comparison to the power of an exact test - F-test, and the power of two other residual-based conventional tests commonly used in regression diagnostics but for testing other departures from the model assumptions.
+  xlab("Conventional test p-value on square root scale") +
+  theme_light()
+```
+
+Figure \ref{fig:power-poly-uniform} shows the estimated power of visual test on lineups with uniform distribution $X_{raw}$ against natural logarithm of the effect with comparison to the power of an exact test - F-test, and two other residual-based conventional tests commonly used in regression diagnostics but for testing other departures from the model assumptions.
 In overall, the power of all four tests increases as the effect becomes larger.
@@
 
+```{r eval = FALSE}
+filter(vi_survey, type == "polynomial") %>%
+  filter(x_dist == "uniform") %>%
+  mutate(log_es = abs(log(effect_size) - 4)) %>%
+  arrange(log_es) %>%
+  View()
+```
+
 <!-- TODO:difference between the simple formula and the simulation approach -->
@@
   ggplot() +
-  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) + geom_point(data = ~filter(.x, unique_lineup_id %in% c("poly_502", "poly_78", "poly_6", "poly_447")), aes(log_effect_size, prop_detect), alpha = 1, col = "red", size = 2) +
+  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual p-value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
   stat_smooth(geom = "line", 
@@
                 filter(x_dist == "uniform") %>%
-                pivot_longer(F_p_value:SW_p_value), 
+                pivot_longer(F_p_value:SW_p_value) %>%
+                mutate(name = gsub("_p_value", " p-value", name)), 
               aes(log_effect_size, as.numeric(value < 0.05), col = name), 
@@
               se = FALSE) +
+  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) + geom_point(data = ~filter(.x, unique_lineup_id %in% c("poly_166", "poly_269", "poly_396", "poly_574")), aes(log_effect_size, prop_detect), alpha = 1, col = "red", size = 2) +
   theme_light(base_size = 5) +
-  facet_wrap(~n) -> p
-
-VI_MODEL$plot_lineup(vi_lineup[["poly_502"]]$data, 
+  scale_color_brewer(palette = "Dark2", direction = -1) +
+  xlab("Natural logarithm of effect size") +
+  ylab("Power") -> p
+
+VI_MODEL$plot_lineup(vi_lineup[["poly_166"]]$data, 
                      remove_grid_line = TRUE, 
                      theme = theme_light(base_size = 5),
-                     remove_axis = TRUE) + 
-  ggtitle(bquote(log[e](E) %~~% 1~","~"F-test p-value" == .(format(filter(vi_survey, unique_lineup_id == "poly_502")$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(vi_survey, unique_lineup_id == "poly_502")$p_value[1], digits = 3)))) -> p1
-
-VI_MODEL$plot_lineup(vi_lineup[["poly_78"]]$data, 
+                     remove_axis = TRUE,
+                     size = 0.1) + 
+  ggtitle(bquote(log[e](E) %~~% 1~","~"F-test p-value" == .(format(filter(vi_survey, unique_lineup_id == "poly_166")$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(vi_survey, unique_lineup_id == "poly_166")$p_value[1], digits = 3)))) -> p1
+
+VI_MODEL$plot_lineup(vi_lineup[["poly_269"]]$data,
                      remove_grid_line = TRUE, 
                      theme = theme_light(base_size = 5),
-                     remove_axis = TRUE) + 
-  ggtitle(bquote(log[e](E) %~~% 2~","~"F-test p-value" == .(format(filter(vi_survey, unique_lineup_id == "poly_78")$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(vi_survey, unique_lineup_id == "poly_78")$p_value[1], digits = 3)))) -> p2
-
-VI_MODEL$plot_lineup(vi_lineup[["poly_6"]]$data, 
+                     remove_axis = TRUE,
+                     size = 0.1) + 
+  ggtitle(bquote(log[e](E) %~~% 2~","~"F-test p-value" == .(format(filter(vi_survey, unique_lineup_id == "poly_269")$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(vi_survey, unique_lineup_id == "poly_269")$p_value[1], digits = 3)))) -> p2
+
+VI_MODEL$plot_lineup(vi_lineup[["poly_396"]]$data, 
                      remove_grid_line = TRUE, 
                      theme = theme_light(base_size = 5),
-                     remove_axis = TRUE) + 
-  ggtitle(bquote(log[e](E) %~~% 3~","~"F-test p-value" == .(format(filter(vi_survey, unique_lineup_id == "poly_6")$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(vi_survey, unique_lineup_id == "poly_6")$p_value[1], digits = 3)))) -> p3
-
-VI_MODEL$plot_lineup(vi_lineup[["poly_447"]]$data, 
+                     remove_axis = TRUE, 
+                     size = 0.1) + 
+  ggtitle(bquote(log[e](E) %~~% 3~","~"F-test p-value" == .(format(filter(vi_survey, unique_lineup_id == "poly_396")$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(vi_survey, unique_lineup_id == "poly_396")$p_value[1], digits = 3)))) -> p3
+
+VI_MODEL$plot_lineup(vi_lineup[["poly_574"]]$data, 
                      remove_grid_line = TRUE, 
                      theme = theme_light(base_size = 5),
-                     remove_axis = TRUE) + 
-  ggtitle(bquote(log[e](E) %~~% 4~","~"F-test p-value" == .(format(filter(vi_survey, unique_lineup_id == "poly_447")$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(vi_survey, unique_lineup_id == "poly_447")$p_value[1], digits = 3)))) -> p4
+                     remove_axis = TRUE,
+                     size = 0.1) + 
+  ggtitle(bquote(log[e](E) %~~% 4~","~"F-test p-value" == .(format(filter(vi_survey, unique_lineup_id == "poly_574")$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(vi_survey, unique_lineup_id == "poly_574")$p_value[1], digits = 3)))) -> p4
 

### 2022-11-15 paper_comparison/paper_comparison/paper_comparison.tex
finish p-value cmp

@@
 
-Subjects recruited from Prolific received a fixed payment for
-participating in the experiment. However, some subjects will try to
-maximize their earnings for minimum effort. During the review of
-submissions, if we found a subject objectively demonstrated clear
-low-effort throughout the experiment, e.g., failed all attention checks,
-we rejected the submission. The rejected submissions will be removed
-immediately, and Prolific will automatically recruit another subject as
-substitution. Therefore, we only paid for approved submissions and no
-further data screening procedure needed to be applied on the collected
-data.
+There were 2880, 2880 and 2214 lineups evaluation made by 160, 160 and
+123 subjects recruited for experiment I, II and III respectively. In the
+total of 7974 lineup evaluations, 3744 used lineups produced by the
+non-linearity model, 3510 used lineups produced by the
+heteroskedasticity model, and 720 used lineups consists of only null
+plots.
 
-In overall, there were a total of 8860 lineup evaluations made by 443
-subjects in experiment I, experiment II and experiment III, where 886
-lineup evaluations were attention checks and were not used in the
-following analysis.
-
-The collated dataset is provided in \texttt{vi\_survey} of the
+Besides, there were 886 attention checks not included in the following
+analysis. The collated dataset is provided in \texttt{vi\_survey} of the
 \texttt{visage} \texttt{R} package.
 
-\hypertarget{p-value-comparison}{%
-\subsection{P-value comparison}\label{p-value-comparison}}
+\hypertarget{p-value-and-power-comparison}{%
+\subsection{P-value and power
+comparison}\label{p-value-and-power-comparison}}
 
@@
 statistical procedures. As displayed in Figure
-\ref{fig:p-value-comparison}, the visual test almost always gives higher
-p-values than the conventional test in both polynomial and
-heteroskedasticity models, suggesting that humans have lower confidence
-and much higher tolerance of the residual departures than the numerical
-procedure. In fact, only around 58\% of the cases, both tests will agree
-on the rejection decision given 95\% confidence level. And in 41\% of
-the cases, visual test does not reject the null hypothesis but
-conventional test does. Although we know that the all the designed
-lineups contain various degrees of residual departures and the
-conventional test is telling truth, analysts and general public as the
-consumers of the output will possibly not believe in or at least not
-response to the ``false alarm'' if they can not see the departures
-directly from the data plot and view it as impactful. The sensitivity of
-the conventional test may distract and discourage us from finding a good
-linear approximation to the data we try to model. This also partly
-explains why abundance of literature suggesting the use of data plot in
-regression diagnostics.
+\ref{fig:p-value-comparison}, the visual test gives higher p-values than
+the conventional test for both polynomial and heteroskedasticity models
+in most of the time, suggesting that humans have lower confidence and
+much higher tolerance of the residual departures than the conventional
+test.
+
+In particular, given 95\% confidence level, both tests will reject the
+null hypothesis in only around 58\% of the cases. And in 41\% of the
+cases, the visual test does not reject the null hypothesis but the
+conventional test does. Since fail to reject the null hypothesis in a
+visual test means that there is no obvious visual discoveries found in
+the residual plot, analysts and the general public as the consumers of
+the output may not fully believe in the rejection of the null hypothesis
+even though the alternative hypothesis is true. Even if the rejection is
+accepted, the model violation may not be considered as impactful due to
+the fact that the departures are not clearly visible.
+
+This makes it sounds like people are arrogantly ignoring the accuracy of
+conventional tests and trust too much to their visual perception, but it
+is not. In fact, the sensitivity of the conventional test could
+sometimes distract and discourage analysts from finding a simple but
+good linear approximation to the data. The rejection of the null
+hypothesis because of acceptable and negligible residual departure is
+not practically meaningful and useful. This also partly explains why
+abundance of literature suggesting the use of data plot in regression
+diagnostics.
 
@@
 
-\caption{...}\label{fig:p-value-comparison}
+\caption{Visual test p-value compared to conventional test p-value for lineups produced by heteroskedasticity and non-linearity model.
+The scatter plot is drawn on a square root scale.
+Every dot on the plot corresponds to a lineup.
+Most of the dots are fallen on the left of the y = x line indicating visual test p-value is almost always higher than conventional test p-value.
+The blue curve is a smoothing of the dots showing a positive trend.
+}\label{fig:p-value-comparison}
 \end{figure}
-
-\hypertarget{power-comparison}{%
-\subsection{Power comparison}\label{power-comparison}}
 
@@
 logarithm of the effect with comparison to the power of an exact test -
-F-test, and the power of two other residual-based conventional tests
-commonly used in regression diagnostics but for testing other departures
-from the model assumptions. In overall, the power of all four tests
-increases as the effect becomes larger. The power curve of F-test climbs
-aggressively from 25\% to around 90\% as \(log_e(E)\) increases from 0
-to 2, while others respond inactively to the change of effect and remain
-lower than 25\% throughout the period, showing that as an exact test,
-the F-test is relatively more sensitive to the type of model defects
-that being considered. The power of visual test arises steadily and
-nearly linearly to around 90\% as \(log_e(E)\) increases from 2 to 5,
-suggesting that the effect starts to make noticeable impact on the
-degree of the presence of the designed visual features. Other two
-inappropriate conventional tests shows improvement at the same time but
-at a lower rate. This coincides the point made by
-\citet{cook1982residuals} mentioned in \ref{hypothesis-testing} that
-residual-based tests for a specific type of model defect are sensitive
-to other types of model defects. At \(log_e(E) \approx 6\), the power
-curve of F-test reaches almost 100\% followed by the visual test by a
-small margin. The power of Breusch--Pagan test and Shapiro--Wilk test
-reach around 75\% and 30\% respectively.
+F-test, and two other residual-based conventional tests commonly used in
+regression diagnostics but for testing other departures from the model
+assumptions. In overall, the power of all four tests increases as the
+effect becomes larger. The power curve of F-test climbs aggressively
+from 25\% to around 90\% as \(log_e(E)\) increases from 0 to 2, while
+others respond inactively to the change of effect and remain lower than
+25\% throughout the period, showing that as an exact test, the F-test is
+relatively more sensitive to the type of model defects that being
+considered. The power of visual test arises steadily and nearly linearly
+to around 90\% as \(log_e(E)\) increases from 2 to 5, suggesting that
+the effect starts to make noticeable impact on the degree of the
+presence of the designed visual features. Other two inappropriate
+conventional tests shows improvement at the same time but at a lower
+rate. This coincides the point made by \citet{cook1982residuals}
+mentioned in \ref{hypothesis-testing} that residual-based tests for a
+specific type of model defect are sensitive to other types of model
+defects. At \(log_e(E) \approx 6\), the power curve of F-test reaches
+almost 100\% followed by the visual test by a small margin. The power of
+Breusch--Pagan test and Shapiro--Wilk test reach around 75\% and 30\%
+respectively.
 

### 2022-11-15 paper_comparison/paper_comparison/paper_comparison.Rmd
fix powre graph

@@
 What truly impress us is the huge difference between the estimated power of visual test and the estimated power of F-test.
+According to the example lineups included in Figure \ref{fig:power-poly-uniform}, at around $log_e(E) = 1$, the  designed visual feature is rarely visible, making the actual data plot indistinguishable from residual plots simulated from the assumed model at this level of difficulty.
+In fact, zero out of 11 subjects detect the actual data plot positioned at panel 10.
+However, F-test has more than 60% chance of rejecting the null hypothesis in this case.
+From a communication perspective, given the fact that the visual difference is unperceivable, the argument that non-linearity present in the fitted model is less convincing to the public.
+
 The margin is largest at around $log_e(E) = 2$.
-An example lineup is included in Figure \ref{fig:power-poly-uniform} where none of subjects detect the actual data plot positioned at panel 10.
-It demonstrates that at this level of difficulty, the designed visual feature is rarely visible, making the actual data plot indistinguishable from residual plots simulated from the assumed model.
-From a communication perspective, given the fact that the visual difference is unperceivable, the argument that non-linearity present in the fitted model is less convincing to the public even though it is true.
+As shown in Figure \ref{fig:power-poly-uniform}, the visual pattern is still undetectable which is similar to $log_e(E) = 1$.
+The visual test power is close to zero.
+In contrast, F-test will reject the null hypothesis 90% of the time.
+
 At around $log_e(E) = 3$, the margin gets smaller as the chance of identifying the actual data plot becomes larger.
@@
 Without knowing the designed visual features beforehand, it is actually hard to identify the actual data plot by pure image comparison.
-The corresponding example lineup for $log_e(E) = 3$ shown in Figure \ref{fig:fig:power-poly-uniform} has the actual data plot positioned at panel 7, where four of 11 subjects detect it.
-It can be observed that a "U-shape" is presented in plot 7, but the signal is not strong enough to attract all five subjects, resulting in a visual p-value sightly above the desired significance level $\alpha = 0.05$.
-At $log_e(E) = 4$, the designed visual features become much clear and attractive, leading to a high percentage of rejection of the null hypothesis.
+The corresponding example lineup for $log_e(E) = 3$ shown in Figure \ref{fig:fig:power-poly-uniform} has the actual data plot positioned at panel 20, where two of 11 subjects detect it.
+It can be observed that a "S-shape" is presented in plot 20, but the signal is not strong enough to attract all 11 subjects, resulting in a visual p-value sightly above $0.1$.
+At $log_e(E) = 4$, the designed visual features become much clear and attractive in panel 3, leading to a high percentage of rejection of the null hypothesis.
 Figure \ref{fig:fig:power-poly-uniform} gives example lineups of such cases.
@@
   filter(type == "polynomial") %>%
-  filter(e_sigma != 0.125) %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
   mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
@@
   ggplot() +
-  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual p-value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
   stat_smooth(geom = "line", 
               data = ~boot_y(.x, reject_dirichlet, times = 500), 
-              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "Visual p-value"), 
+              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "Visual test"), 
               method = "glm", 
@@
                 pivot_longer(F_p_value:SW_p_value) %>%
-                mutate(name = gsub("_p_value", " p-value", name)), 
+                mutate(name = gsub("_p_value", " test", name)), 
               aes(log_effect_size, as.numeric(value < 0.05), col = name), 
@@
 
-```{r eval = FALSE, fig.height= 3, fig.width=4}
-vi_survey %>%
-  filter(x_dist == "uniform") %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type == "polynomial") %>%
-  filter(e_sigma != 0.125) %>%
-  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  filter(abs(log(effect_size) - 1) < 0.5) %>%
-  filter(n == 100)
-
-p
-
-vi_survey %>%
-  filter(x_dist == "uniform") %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type == "polynomial") %>%
-  filter(e_sigma != 0.125) %>%
-  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  filter(abs(log(effect_size) - 4) < 0.5) %>%
-  filter(prop_detect < 0.25)
-
-vi_survey %>%
-  filter(unique_lineup_id == "poly_199") %>%
-  View()
-
-VI_MODEL$plot_lineup(vi_lineup$poly_159$data)
-
-vi_survey %>%
-  filter(x_dist == "uniform") %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type == "polynomial") %>%
-  filter(e_sigma != 0.125) %>%
-  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  ggplot() +
-  geom_abline() +
-  geom_point(aes(p_value_uniform, p_value)) +
-  scale_x_log10() +
-  scale_y_log10()
-
-vi_survey %>%
-  filter(x_dist == "uniform") %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type == "polynomial") %>%
-  filter(e_sigma != 0.125) %>%
-  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  ggplot() +
-  ggbeeswarm::geom_quasirandom(aes(factor(e_sigma), weighted_detect)) +
-  facet_wrap(~n)
-```
-
 ## Distribution of regressor
@@
 
-```{r dist-power, fig.cap="..."}
+```{r eval = FALSE}
 vi_survey %>%
@@
   ggplot() +
-  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) +
+  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
   stat_smooth(geom = "line", 
               data = ~boot_y(.x, reject_dirichlet, times = 500), 
-              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"),
+              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "Visual test"), 
               method = "glm", 
@@
   geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
-                pivot_longer(F_p_value:SW_p_value),
+                pivot_longer(F_p_value:SW_p_value) %>%
+                mutate(name = gsub("_p_value", " test", name)), 
               aes(log_effect_size, as.numeric(value < 0.05), col = name), 
@@
               se = FALSE) +
-  facet_wrap(~x_dist) +
-  theme_light(base_size = 5)
-```
-
-```{r}
+  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) +
+  theme_light(base_size = 5) +
+  scale_color_brewer(palette = "Dark2", direction = -1) +
+  xlab("Natural logarithm of effect size") +
+  ylab("Power") +
+  facet_wrap(~x_dist)
+```
+
+```{r eval = FALSE}
 vi_survey %>%
@@
   filter(type == "heteroskedasticity") %>%
-  filter(b != 0) %>%
-  filter(b < 128) %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
   mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
@@
   ggplot() +
-  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
   stat_smooth(geom = "line", 
               data = ~boot_y(.x, reject_dirichlet, times = 500), 
-              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
+              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "Visual test"), 
               method = "glm", 
@@
   geom_smooth(data = mutate(heter_conv_sim, log_effect_size = log(effect_size)) %>%
-                pivot_longer(F_p_value:SW_p_value), 
+                pivot_longer(F_p_value:SW_p_value) %>%
+                mutate(name = gsub("_p_value", " test", name)), 
               aes(log_effect_size, as.numeric(value < 0.05), col = name), 
@@
               se = FALSE) +
-  facet_wrap(~x_dist) +
-  ggtitle("heter")
+  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) +
+  theme_light(base_size = 5) +
+  scale_color_brewer(palette = "Dark2", direction = -1) +
+  xlab("Natural logarithm of effect size") +
+  ylab("Power") +
+  facet_wrap(~x_dist)
+```
+
+```{r dist-power, fig.cap="..."}
+vi_survey %>%
+  mutate(x_dist = ifelse(x_dist == "even_discrete", "discrete uniform", x_dist)) %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
+  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  ggplot() +
+  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
+                pivot_longer(F_p_value:SW_p_value) %>%
+                mutate(name = gsub("_p_value", " test", name)) %>%
+                mutate(type = "polynomial") %>%
+                mutate(x_dist = ifelse(x_dist == "even_discrete", "discrete uniform", x_dist)) %>%
+                bind_rows(mutate(heter_conv_sim, log_effect_size = log(effect_size)) %>%
+                pivot_longer(F_p_value:SW_p_value) %>%
+                mutate(name = gsub("_p_value", " test", name)) %>%
+                mutate(type = "heteroskedasticity") %>%
+                mutate(x_dist = ifelse(x_dist == "even_discrete", "discrete uniform", x_dist))), 
+              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
+              method = "glm", 
+              method.args = list(family = binomial),
+              se = FALSE) +
+  theme_light(base_size = 5) +
+  scale_color_brewer(palette = "Dark2", direction = -1) +
+  xlab("Natural logarithm of effect size") +
+  ylab("Power") +
+  facet_grid(x_dist~type)
 ```

### 2022-11-15 paper_comparison/paper_comparison/paper_comparison.tex
fix powre graph

@@
 What truly impress us is the huge difference between the estimated power
-of visual test and the estimated power of F-test. The margin is largest
-at around \(log_e(E) = 2\). An example lineup is included in Figure
-\ref{fig:power-poly-uniform} where none of subjects detect the actual
-data plot positioned at panel 10. It demonstrates that at this level of
-difficulty, the designed visual feature is rarely visible, making the
-actual data plot indistinguishable from residual plots simulated from
-the assumed model. From a communication perspective, given the fact that
-the visual difference is unperceivable, the argument that non-linearity
-present in the fitted model is less convincing to the public even though
-it is true. At around \(log_e(E) = 3\), the margin gets smaller as the
-chance of identifying the actual data plot becomes larger. At this level
-of difficulty, the designed visual features are usually detectable but
-it may not stand out from the lineup as other null plots may happen to
+of visual test and the estimated power of F-test. According to the
+example lineups included in Figure \ref{fig:power-poly-uniform}, at
+around \(log_e(E) = 1\), the designed visual feature is rarely visible,
+making the actual data plot indistinguishable from residual plots
+simulated from the assumed model at this level of difficulty. In fact,
+zero out of 11 subjects detect the actual data plot positioned at panel
+10. However, F-test has more than 60\% chance of rejecting the null
+hypothesis in this case. From a communication perspective, given the
+fact that the visual difference is unperceivable, the argument that
+non-linearity present in the fitted model is less convincing to the
+public.
+
+The margin is largest at around \(log_e(E) = 2\). As shown in Figure
+\ref{fig:power-poly-uniform}, the visual pattern is still undetectable
+which is similar to \(log_e(E) = 1\). The visual test power is close to
+zero. In contrast, F-test will reject the null hypothesis 90\% of the
+time.
+
+At around \(log_e(E) = 3\), the margin gets smaller as the chance of
+identifying the actual data plot becomes larger. At this level of
+difficulty, the designed visual features are usually detectable but it
+may not stand out from the lineup as other null plots may happen to
 include outliers or visual patterns that are are considered to be more
@@
 \ref{fig:fig:power-poly-uniform} has the actual data plot positioned at
-panel 7, where four of 11 subjects detect it. It can be observed that a
-``U-shape'' is presented in plot 7, but the signal is not strong enough
-to attract all five subjects, resulting in a visual p-value sightly
-above the desired significance level \(\alpha = 0.05\). At
-\(log_e(E) = 4\), the designed visual features become much clear and
-attractive, leading to a high percentage of rejection of the null
-hypothesis. Figure \ref{fig:fig:power-poly-uniform} gives example
-lineups of such cases. It is also interesting to observe that the power
-curve of visual test behaves similarly to those inappropriate
-conventional tests until the designed visual features really becomes
-visible.
+panel 20, where two of 11 subjects detect it. It can be observed that a
+``S-shape'' is presented in plot 20, but the signal is not strong enough
+to attract all 11 subjects, resulting in a visual p-value sightly above
+\(0.1\). At \(log_e(E) = 4\), the designed visual features become much
+clear and attractive in panel 3, leading to a high percentage of
+rejection of the null hypothesis. Figure
+\ref{fig:fig:power-poly-uniform} gives example lineups of such cases. It
+is also interesting to observe that the power curve of visual test
+behaves similarly to those inappropriate conventional tests until the
+designed visual features really becomes visible.
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-7-1} \end{center}
-
 --\textgreater{}

### 2022-11-15 paper_comparison/paper_comparison/paper_comparison.Rmd
update distribution cmp

@@
 
-## Distribution of regressor
-
-The impact of the distribution of $X_raw$ on the power is displayed in Figure \ref{fig:dist-power}.
-The power curve of F-test is stable across different distributions, while the visual test has a steeper power curve for normal and uniform distribution.
-BP-test performs worse for discrete uniform distribution and uniform distribution but has relatively high power for normal distribution.
-SW-test outperforms BP-test for discrete uniform distribution but remains as the worst test for other distributions.
+The impact of the distribution of $X_{raw}$ on the power is displayed in Figure \ref{fig:dist-power}.
+For the non-linearity model, the power curve of F-test is stable across different distributions, while the visual test has a more gentle power curve for the discrete uniform distribution as it contains lineups with larger effect size.
+BP test performs worse for discrete uniform distribution but has relatively high power for normal distribution.
+SW test outperforms BP test for discrete uniform distribution but remains as the worst test for other distributions.
 The results indicate those inappropriate residual-based tests are sensitive to the distribution of the regressor.
+
+For the heteroskedasticity model, F-test is entirely insensitive to the non-constant variance. The visual test has similar power as the SW test  
+for normal distribution and lognormal distribution, and has higher power for uniform distribution, and lower power for discrete uniform distribution.
+
+But still, we observe the huge gap in power between the preferred conventional tests and the visual test which is consistent with our previous argument that conventional test is too sensitive for diagnosing non-linearity and heteroskedasticity.
 
@@
 
-```{r dist-power, fig.cap="..."}
+```{r, fig.cap="..."}
 vi_survey %>%
@@
 
-<!-- ## Power comparison -->
-
-<!-- ### Experiment I -->
-
-<!-- Figure \ref{fig:power-overview} shows an overview of estimated power of visual test against natural logarithm of the effect with comparison to the power of an exact test - F-test, and the power of two other residual-based conventional tests commonly used in regression diagnostics but for testing other departures from the model assumptions.
-In overall, the power of all four tests increases as the effect becomes larger.
-The power curve of F-test climbs aggressively from 25% to around 90% as $log_e(E)$ increases from 0 to 2, while others respond inactively to the change of effect and remain lower than 25% throughout the period, showing that as an exact test, the F-test is relatively more sensitive to the type of model defects that being considered.
-The power of visual test arises steadily and nearly linearly to around 90% as $log_e(E)$ increases from 2 to 5, suggesting that the effect starts to make noticeable impact on the degree of the presence of the designed visual features.
-Other two inappropriate conventional tests shows improvement at the same time but at a lower rate.
-This coincides the point made by @cook1982residuals mentioned in \ref{hypothesis-testing} that residual-based tests for a specific type of model defect are sensitive to other types of model defects.
-At $log_e(E) = 6$, the power curve of F-test reaches almost 100% followed by the visual test by a small margin.
-The power of Breusch--Pagan test and Shapiro--Wilk test reach around 75% and 63% respectively.
--->
-
-<!-- What truly impress us is the huge difference between the estimated power of visual test and the estimated power of F-test.
-The margin is largest at around $log_e(E) = 2$.
-An example lineup is included in Figure \ref{fig:power-overview} where none of subjects detect the actual data plot positioned at panel 14.
-It demonstrates that at this level of difficulty, the designed visual feature is rarely visible, making the actual data plot indistinguishable from residual plots simulated from the assumed model.
-From a communication perspective, given the fact that the visual difference is unperceivable, the argument that non-linearity present in the fitted model is less convincing to the public even though it is true.
-At around $log_e(E) = 3$, the margin gets smaller as the chance of identifying the actual data plot becomes larger.
-At this level of difficulty, the designed visual features are usually detectable but it may not stand out from the lineup as other null plots may happen to include outliers or visual patterns that are are considered to be more attractive by human, and thus recognized as the most different plot.
-Without knowing the designed visual features beforehand, it is actually hard to identify the actual data plot by pure image comparison.
-The corresponding example lineup for $log_e(E) = 3$ shown in Figure \ref{fig:power-overview} has the actual data plot positioned at panel 20, where two out of five subjects detect it.
-It can be observed that a M-shape is presented in plot 20, but the signal is not strong enough to attract all five subjects, resulting in a visual p-value sightly above the desired significance level $\alpha = 0.05$.
-At $log_e(E) = 4$ and $log_e(E) = 6$, the designed visual features become much clear and attractive, leading to a high percentage of rejection of the null hypothesis.
-Figure \ref{fig:power-overview} gives example lineups of such cases.
--->
-
-<!-- It is also interesting to observe that the power curve of visual test behaves similarly to those inappropriate conventional tests until the designed visual features really becomes visible   -->
-
-<!-- ```{r} -->
-<!-- conv_test_plot <- function() { -->
-<!--   conv_result %>% -->
-<!--   ggplot() + -->
-<!--   stat_smooth(geom = "line",  -->
-<!--               aes(log(effect_size), reject, group = boot_id, col = "F-test"),  -->
-<!--               se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) + -->
-<!--   stat_smooth(geom = "line",  -->
-<!--               aes(log(effect_size), reject, col = "F-test"),  -->
-<!--               se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1) + -->
-<!--   labs(col = "Test") + -->
-<!--   theme_light(base_size = 5) + -->
-<!--   xlab(quote(log[e](E))) + -->
-<!--   ylab("Power") -->
-<!-- } -->
-
-<!-- add_wrong_BP_conv_test_plot_layers <- function(p) { -->
-<!--   p + -->
-<!--   stat_smooth(data = wrong_BP_conv_result, geom = "line",  -->
-<!--               aes(log(effect_size), reject, group = boot_id, col = "BP-test"),  -->
-<!--               se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) + -->
-<!--   stat_smooth(data = wrong_BP_conv_result, geom = "line",  -->
-<!--               aes(log(effect_size), reject, col = "BP-test"),  -->
-<!--               se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1) -->
-<!-- } -->
-
-<!-- add_wrong_SW_conv_test_plot_layers <- function(p) { -->
-<!--   p + -->
-<!--   stat_smooth(data = wrong_SW_conv_result, geom = "line",  -->
-<!--               aes(log(effect_size), reject, group = boot_id, col = "SW-test"),  -->
-<!--               se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) + -->
-<!--   stat_smooth(data = wrong_SW_conv_result, geom = "line",  -->
-<!--               aes(log(effect_size), reject, col = "SW-test"),  -->
-<!--               se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1) -->
-<!-- } -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- boot_dat <- function(dat, times = 100) { -->
-<!--   result <- mutate(dat, boot_id = 1) -->
-<!--   for (i in 2:times) { -->
-<!--     result <- bind_rows(result, mutate(dat[sample(1:nrow(dat), replace = TRUE), ], boot_id = i))  -->
-<!--   } -->
-<!--   result -->
-<!-- } -->
-
-<!-- add_visual_test_plot_layers <- function(p, points = c()) { -->
-<!--   visual_dat <- polynomials %>% -->
-<!--   group_by(lineup_id) %>% -->
-<!--   summarise(across(c(type:p_value, prop_detect), first)) %>% -->
-<!--   filter(lineup_id < 577) %>% -->
-<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
-<!--   boot_dat(times = 100) -->
-
-<!--   p + -->
-<!--   stat_smooth(data = filter(visual_dat, boot_id == 1), geom = "line",  -->
-<!--               aes(log(effect_size), reject, group = boot_id, col = "Visual"),  -->
-<!--               method = "glm", method.args = list(family = binomial), se = FALSE, alpha = 1) + -->
-<!--   stat_smooth(data = visual_dat, geom = "line",  -->
-<!--               aes(log(effect_size), reject, group = boot_id, col = "Visual"),  -->
-<!--               method = "glm", method.args = list(family = binomial), se = FALSE, alpha = 0.05) + -->
-<!--   geom_point(data = filter(visual_dat, boot_id == 1), aes(log(effect_size), prop_detect), alpha = 0.1) +  -->
-<!--   geom_point(data = filter(visual_dat, boot_id == 1, lineup_id %in% points), aes(log(effect_size), prop_detect), alpha = 1, col = "red", size = 2) -->
-<!-- } -->
-
-<!-- control_col <- function(p, ...) { -->
-<!--   p + scale_color_manual(values = unlist(list(...))) -->
-<!-- } -->
-<!-- ``` -->
-
-<!-- ```{r eval = FALSE} -->
-<!-- filter(polynomials, lineup_id == 131) -->
-
-<!-- polynomials %>% -->
-<!--   group_by(lineup_id) %>% -->
-<!--   summarise(dd = first(abs(log(effect_size) - 5)), p_value = first(p_value), prop_detect = first(prop_detect)) %>% -->
-<!--   arrange(dd) %>% -->
-<!--   head(30) %>% -->
-<!--   summarise(tt = median(prop_detect)) -->
-
-<!-- polynomials %>% -->
-<!--   ggplot() + -->
-<!--   geom_point(aes(prop_detect, p_value < 0.06)) -->
-
-<!-- bottom_plot_2 -->
-
-<!-- polynomials %>% -->
-<!--   filter(lineup_id == 364) -->
-
-<!-- polynomials %>% -->
-<!--   filter(e_sigma == 4) %>% -->
-<!--   group_by(lineup_id) %>% -->
-<!--   summarise(p_value = first(p_value)) %>% -->
-<!--   mutate(reject = p_value < 0.05) %>% -->
-<!--   {mean(.$reject)} -->
-
-<!-- 4 * 4 * 3 * 5 -->
-<!-- ``` -->
-
-<!-- ```{r power-overview, fig.height=7.5, fig.cap="Power overview. \\label{fig:power-overview}"} -->
-<!-- conv_test_plot() %>% -->
-<!--   add_wrong_BP_conv_test_plot_layers() %>% -->
-<!--   add_wrong_SW_conv_test_plot_layers() %>% -->
-<!--   add_visual_test_plot_layers(points = c(415, 131, 322, 378)) + -->
-<!--   geom_vline(xintercept = 2, linetype = 2) +  -->
-<!--   geom_vline(xintercept = 3, linetype = 2) + -->
-<!--   geom_vline(xintercept = 4, linetype = 2) + -->
-<!--   geom_vline(xintercept = 5, linetype = 2) -> base_plot -->
-
-<!-- # VI_MODEL$plot_lineup(polynomials_lineup$lineup_314$data, -->
-<!-- #                      size = 0.1, -->
-<!-- #                      stroke = 0.2, -->
-<!-- #                      remove_grid_line = TRUE, -->
-<!-- #                      theme = theme_light(base_size = 5), -->
-<!-- #                      remove_axis = TRUE) + -->
-<!-- #   ggtitle(bquote(log[e](E) == 0~", F-test p-value" == .(format(filter(polynomials, lineup_id == 314)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 314)$p_value[1], digits = 3)))) -> bottom_plot_1 -->
-
-<!-- VI_MODEL$plot_lineup(polynomials_lineup$lineup_415$data, -->
-<!--                      size = 0.1, -->
-<!--                      stroke = 0.2, -->
-<!--                      remove_grid_line = TRUE, -->
-<!--                      theme = theme_light(base_size = 5), -->
-<!--                      remove_axis = TRUE) + -->
-<!--   ggtitle(bquote(log[e](E) %~~% 2~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 415)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 415)$p_value[1], digits = 3)))) -> bottom_plot_1 -->
-
-<!-- VI_MODEL$plot_lineup(polynomials_lineup$lineup_131$data, -->
-<!--                      size = 0.1, -->
-<!--                      stroke = 0.2, -->
-<!--                      remove_grid_line = TRUE, -->
-<!--                      theme = theme_light(base_size = 5), -->
-<!--                      remove_axis = TRUE) + -->
-<!--   ggtitle(bquote(log[e](E) %~~% 3~", F-test p-value" == .(format(filter(polynomials, lineup_id == 131)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 131)$p_value[1], digits = 3)))) -> bottom_plot_2 -->
-
-<!-- VI_MODEL$plot_lineup(polynomials_lineup$lineup_322$data, -->
-<!--                      size = 0.1, -->
-<!--                      stroke = 0.2, -->
-<!--                      remove_grid_line = TRUE, -->
-<!--                      theme = theme_light(base_size = 5), -->
-<!--                      remove_axis = TRUE) + -->
-<!--   ggtitle(bquote(log[e](E) %~~% 4~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 322)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 322)$p_value[1], digits = 3)))) -> bottom_plot_3 -->
-
-<!-- VI_MODEL$plot_lineup(polynomials_lineup$lineup_378$data, -->
-<!--                      size = 0.1, -->
-<!--                      stroke = 0.2, -->
-<!--                      remove_grid_line = TRUE, -->
-<!--                      theme = theme_light(base_size = 5), -->
-<!--                      remove_axis = TRUE) + -->
-<!--   ggtitle(bquote(log[e](E) %~~% 5~","~"F-test p-value" == .(format(filter(polynomials, lineup_id == 378)$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(polynomials, lineup_id == 378)$p_value[1], digits = 3)))) -> bottom_plot_4 -->
-
-<!-- design <- c(patchwork::area(1, 1, 1, 8),  -->
-<!--             patchwork::area(2, 1, 2, 4),  -->
-<!--             patchwork::area(2, 5, 2, 8),  -->
-<!--             patchwork::area(3, 1, 3, 4),  -->
-<!--             patchwork::area(3, 5, 3, 8)) -->
-
-<!-- patchwork::wrap_plots(base_plot, B = bottom_plot_1, C = bottom_plot_2, D = bottom_plot_3, E = bottom_plot_4, design = design) -->
-<!-- ``` -->
-
-<!-- ### Distributions of regressor -->
-
-<!-- The impact of the distribution of $X_raw$ on the power is shown in Figure \ref{fig:dist-power}.
-The power curve of F-test is stable across different distributions, while the visual test has a steeper power curve for normal and uniform distribution.
-BP-test performs worse for discrete uniform distribution and uniform distribution but has relatively high power for normal distribution.
-SW-test outperforms BP-test for discrete uniform distribution but remains as the worst test for other distributions.
-The results indicate those inappropriate residual-based tests are sensitive to the distribution of the regressor.
--->
-
-<!-- ```{r dist-power, fig.cap="\\label{dist-power}"} -->
-<!-- conv_test_plot() %>% -->
-<!--   add_wrong_BP_conv_test_plot_layers() %>% -->
-<!--   add_wrong_SW_conv_test_plot_layers() %>% -->
-<!--   add_visual_test_plot_layers() + -->
-<!--   facet_wrap(~x_dist) -->
-<!-- ``` -->
-
-<!-- #### Shpaes of non-linearity -->
-
-<!-- Figure \ref{fig:shape-power} illustrates the change of power under different shape of non-linearity.
-Similar to the power curves shown in \ref{fig:dist-power}, F-test is stable under different shapes.
-The power curve of visual test also behaves similarly across different shapes.
-What vary are the power curve of BP-test and SW-test.
-For Triple-U shape, both BP-test and SW-test are insensitive to the change of the effect.
-And for W shape, both tests have almost identical power curves.
-It can be observed that both tests performs the best for U-shape.
--->
-
-<!-- ```{r shape-power, fig.cap="\\label{shape-power}"} -->
-<!-- conv_test_plot() %>% -->
-<!--   add_wrong_BP_conv_test_plot_layers() %>% -->
-<!--   add_wrong_SW_conv_test_plot_layers() %>% -->
-<!--   add_visual_test_plot_layers() + -->
-<!--   facet_wrap(~shape) -->
-<!-- ``` -->
-
-<!-- ```{r eval = FALSE} -->
-<!-- conv_test_plot() %>% -->
-<!--   add_wrong_BP_conv_test_plot_layers() %>% -->
-<!--   add_wrong_SW_conv_test_plot_layers() %>% -->
-<!--   add_visual_test_plot_layers() + -->
-<!--   facet_wrap(~e_sigma) -->
-<!-- ``` -->
-
-<!-- #### Reasons for making selections -->
-
-<!-- ```{r} -->
-<!-- polynomials %>% -->
-<!--   mutate(reason_tmp = ifelse(reason %in% c("Cluster(s)", "Shape", "Outlier(s)"), reason, "Other")) %>% -->
-<!--   ggplot() + -->
-<!--   geom_bar(aes(reason_tmp)) + -->
-<!--   facet_grid(x_dist~detect)  -->
-<!-- ``` -->
-
-<!-- ### Experiment II -->
-
-<!-- <!-- todo: check the two way interaction for x_dist and a --> -->
-
-<!-- ```{r eval = FALSE} -->
-<!-- heter %>% -->
-<!--   filter(b < 3) %>% -->
-<!--   group_by(lineup_id) %>% -->
-<!--   summarise(p_value = first(p_value), n_eval = n()) %>% -->
-<!--   filter(n_eval > 3) %>% -->
-<!--   mutate(reject = p_value < 0.05) %>% -->
-<!--   {mean(.$reject)} -->
-
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- map_to_group <- function(b) { -->
-<!--   round(b / 4) + 1 -->
-<!-- } -->
-
-<!-- heter_conv_test_plot <- function() { -->
-<!--   conv_result_heter %>% -->
-<!--   mutate(cat_b = map_to_group(b)) %>% -->
-<!--   ggplot() + -->
-<!--   stat_smooth(geom = "line",  -->
-<!--               aes(log(effect_size), reject, group = boot_id, col = "BP-test"),  -->
-<!--               se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) + -->
-<!--   stat_smooth(geom = "line",  -->
-<!--               aes(log(effect_size), reject, col = "BP-test"),  -->
-<!--               se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1) + -->
-<!--   labs(col = "Test") + -->
-<!--   theme_light(base_size = 5) + -->
-<!--   xlab(quote(log[e](E))) + -->
-<!--   ylab("Power") -->
-<!-- } -->
-
-<!-- add_heter_SW_plot_layers <- function(p) { -->
-<!--   p + -->
-<!--   stat_smooth(data = mutate(conv_result_heter_SW, cat_b = map_to_group(b)), geom = "line",  -->
-<!--               aes(log(effect_size), reject, group = boot_id, col = "SW-test"),  -->
-<!--               se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 0.05) + -->
-<!--   stat_smooth(data = mutate(conv_result_heter_SW, cat_b = map_to_group(b)), geom = "line",  -->
-<!--               aes(log(effect_size), reject, col = "SW-test"),  -->
-<!--               se = FALSE, method = "glm", method.args = list(family = binomial()), alpha = 1) -->
-<!-- } -->
-
-<!-- add_heter_visual_test_plot_layers <- function(p, points = c()) { -->
-
-<!--   visual_dat <- heter %>% -->
-<!--   mutate(x_dist = ifelse(x_dist == "discrete_uniform", "even_discrete", x_dist)) %>% -->
-<!--   mutate(x_dist = ifelse(x_dist == "neglognormal", "lognormal", x_dist)) %>% -->
-<!--   group_by(lineup_id) %>% -->
-<!--   summarise(across(c(type:p_value, prop_detect), first)) %>% -->
-<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
-<!--   boot_dat(times = 100) %>% -->
-<!--   mutate(cat_b = map_to_group(b)) -->
-
-<!--   p + -->
-<!--   stat_smooth(data = filter(visual_dat, boot_id == 1), geom = "line",  -->
-<!--               aes(log(effect_size), reject, group = boot_id, col = "Visual"),  -->
-<!--               method = "glm", method.args = list(family = binomial), se = FALSE, alpha = 1) + -->
-<!--   stat_smooth(data = visual_dat, geom = "line",  -->
-<!--               aes(log(effect_size), reject, group = boot_id, col = "Visual"),  -->
-<!--               method = "glm", method.args = list(family = binomial), se = FALSE, alpha = 0.05) + -->
-<!--   geom_point(data = filter(visual_dat, boot_id == 1), aes(log(effect_size), prop_detect), alpha = 0.1) +  -->
-<!--   geom_point(data = filter(visual_dat, boot_id == 1, lineup_id %in% points), aes(log(effect_size), prop_detect), alpha = 1, col = "red", size = 2) -->
-<!-- } -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- heter_conv_test_plot() %>% -->
-<!--   add_heter_SW_plot_layers() %>% -->
-<!--   add_heter_visual_test_plot_layers() -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- heter_conv_test_plot() %>% -->
-<!--   add_heter_SW_plot_layers() %>% -->
-<!--   add_heter_visual_test_plot_layers() + -->
-<!--   facet_wrap(~a) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- heter_conv_test_plot() %>% -->
-<!--   add_heter_SW_plot_layers() %>% -->
-<!--   add_heter_visual_test_plot_layers() + -->
-<!--   facet_grid(x_dist~a) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- heter_conv_test_plot() %>% -->
-<!--   add_heter_SW_plot_layers() %>% -->
-<!--   add_heter_visual_test_plot_layers() + -->
-<!--   facet_wrap(~x_dist) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- heter_conv_test_plot() %>% -->
-<!--   add_heter_SW_plot_layers() %>% -->
-<!--   add_heter_visual_test_plot_layers() + -->
-<!--   facet_wrap(~cat_b) -->
-<!-- ``` -->
-
-<!-- # time taken vs order of the plot -->
-
-<!-- # accuracy vs order of the plot -->
-
-<!-- # num of selection vs order of the plot -->
-
-<!-- # order of the plot? -->
-
-<!-- # summary of the demographic information -->
-
-<!-- # Old -->
-
-<!-- ## Overview of the Data -->
-
-<!-- We collected 400 lineup evaluations made by 20 participants in experiment I and 880 lineup evaluations made by 44 participants in experiment II.
-In total, 442 unique lineups were evaluated by 64 subjects.
-In experiment I, one of the participants skipped all 20 lineups.
-Hence, the submission was rejected and removed from the dataset.
-In experiment II, there was a participant failed one of the two attention checks, but there was no further evidence of low-effort throughout the experiment.
-Therefore, the submission was kept.
--->
-
-<!-- ## Power comparision -->
-
-<!-- 1. power (visual test vs. conventional test) -->
-
-<!-- (visual test most different one (everything test, any departure)) -->
-
-<!-- plot figure in a paper, desc, exp -->
-
-<!-- 2. investigate the difference (gap), give examples -->
-
-<!-- 3. conventional is too sensitive -->
-
-<!-- 4. make conventional less sensitive (vary alpha) -->
-
-<!-- To model the power of visual test, 10 logistic regression were fit for different number of evaluations ranged from one to five and two different types of simulation setting.
-All 10 models used natural logarithm of the effect size as the only fixed effect, and whether the test successfully rejects the null hypothesis as the response variable.
-Given the way we define the effect size, it was expected that with larger effect size, both conventional test and visual test will have higher probability in rejecting the null hypothesis when it is not true.
-The modelling result summarized in \ref{tab:powerglmcubic} and \ref{tab:powerglmheter} aligned with the expectation as the coefficients of natural logarithm of the effect size are positive and significant across all 10 models.
--->
-
-<!-- Figure \ref{fig:power-com} illustrates the fitted models, while providing the local constant estimate of the power of F-test and Breusch–Pagan test for comparison.
-Data for the conventional test is simulated under the model setting described in section ...
-and 5000000 samples are drawn for both cubic and heteroskedasticity model.
-From Figure \ref{fig:power-com}, it can be observed that the fitted power of visual test increased as the number of evaluations increased for both cubic and heteroskedasticity model.
--->
-
-<!-- For heteroskedasticity model, this phenomenon was more obvious as the power of visual tests with evaluations greater than two were always greater than those with evaluations smaller than two.
--->
-
-<!-- For cubic model, the separation between curves was small.
-The estimated power of visual tests with three to five evaluations were almost identical to each other in regards of effect size.
-In addition, all five curves peaked at one as effect size increased, suggesting that identification of non-linearity as a visual task can be completed reliably by human as long as the departure from null hypothesis is large enough.
--->
-
-<!-- As shown in Figure \ref{fig:power-com}, both F-test and Breusch–Pagan test generally possessed greater power than visual test.
-A visual tests is a collection of test against any alternatives that would create visual discoverable features, while a conventional test is usually targeting at a pre-specified alternative.
-Considering the data generating process of the model defect was known and controlled in this research, where all other alternatives have been eliminated except the one we concerned, the result was suggested that conventional tests were more sensitive to violations of linearity and homoscedasticity assumption than visual tests.
--->
-
-<!-- It was also found that there was a noticeable gap between curves of the conventional test and the visual test at around $log(\text{effect size}) = 0$ for the cubic model and $log(\text{effect size}) = 2.5$ for the heteroskedasticity model, where the differences in power were greater than 0.6.
-We further analysed the lineups with correspoding effect sizes.
-Figure \ref{fig:cubic-hard} and \ref{fig:heter-hard} showed that human was indeed hard to identify the patterns at this level of difficulty.
-The visual difference between the true data plot and null plots were almost unnoticeable.
--->
-
-<!-- <!-- sample 1, change values of a and b, plot the data  -->
-
---\>
-
-<!-- ## Effect of parameters on power of the visual test -->
-
-<!-- The previous section focuses on the change of effect size relative to the power of the visual test.
-However, effect size is only a one dimensional summarisation of parameters used in data simulation.
-Individual factor embedded in the simulation process should also be analysed.
--->
-
-<!-- In cubic model, two major factors that influencing the strength of the signal are $a$ and $b$.
-Figure \ref{fig:power-com-cubic-a} and \ref{fig:power-com-cubic-a} illustrates 30 different logistic regressions fit for different number of evaluations and different number of observations $n$.
-The regressor used in these models was $|a|/\sigma$ since the noise level $\sigma$ needed to be taken into account.
-From the figures, we can observe ...
--->
-
-<!-- ```{r power-vs-log-effect-size} -->
-
-<!-- polynomials %>% -->
-
-<!--   group_by(lineup_id) %>% -->
-
-<!--   summarise(across(c(type:p_value), first)) %>% -->
-
-<!--   filter(lineup_id < 577) %>% -->
-
-<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
-
-<!--   ggplot() + -->
-
-<!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
-
-<!--   geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
-
-<!--   xlab("Natural logarithm of effect size") + -->
-
-<!--   ylab("Power") -->
-
-<!-- ``` -->
-
-<!-- ```{r power-vs-log-effect-size-given-x-dist} -->
-
-<!-- polynomials %>% -->
-
-<!--   group_by(lineup_id) %>% -->
-
-<!--   summarise(across(c(type:p_value), first)) %>% -->
-
-<!--   filter(lineup_id < 577) %>% -->
-
-<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
-
-<!--   ggplot() + -->
-
-<!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
-
-<!--   geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
-
-<!--   facet_wrap(~x_dist) + -->
-
-<!--   xlab("Natural logarithm of effect size") + -->
-
-<!--   ylab("Power") + -->
-
-<!--   ggtitle("Power comparison conditional on the distribution of X") -->
-
-<!-- ``` -->
-
-<!-- ```{r power-of-visual-test-given-x-dist} -->
-
-<!-- polynomials %>% -->
-
-<!--   group_by(lineup_id) %>% -->
-
-<!--   summarise(across(c(type:p_value), first)) %>% -->
-
-<!--   filter(lineup_id < 577) %>% -->
-
-<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
-
-<!--   ggplot() + -->
-
-<!--   geom_smooth(aes(log(effect_size), reject, col = x_dist), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
-
-<!--   xlab("Natural logarithm of effect size") + -->
-
-<!--   ylab("Power") + -->
-
-<!--   ggtitle("Power of visual test conditional on the distribution of X") -->
-
-<!-- ``` -->
-
-<!-- ```{r power-vs-log-effect-size-given-shape} -->
-
-<!-- polynomials %>% -->
-
-<!--   group_by(lineup_id) %>% -->
-
-<!--   summarise(across(c(type:p_value), first)) %>% -->
-
-<!--   filter(lineup_id < 577) %>% -->
-
-<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
-
-<!--   ggplot() + -->
-
-<!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
-
-<!--   geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
-
-<!--   facet_wrap(~shape) + -->
-
-<!--   xlab("Natural logarithm of effect size") + -->
-
-<!--   ylab("Power") + -->
-
-<!--   ggtitle("Power comparison conditional on the shape of the Hermite polynomials") -->
-
-<!-- ``` -->
-
-<!-- ```{r power-of-visual-test-given-shape} -->
-
-<!-- polynomials %>% -->
-
-<!--   group_by(lineup_id) %>% -->
-
-<!--   summarise(across(c(type:p_value), first)) %>% -->
-
-<!--   filter(lineup_id < 577) %>% -->
-
-<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01)), shape = factor(shape)) %>% -->
-
-<!--   ggplot() + -->
-
-<!--   geom_smooth(aes(log(effect_size), reject, col = shape), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
-
-<!--   xlab("Natural logarithm of effect size") + -->
-
-<!--   ylab("Power") + -->
-
-<!--   ggtitle("Power of visual test conditional on the shape of the Hermite polynomials") -->
-
-<!-- ``` -->
-
-<!-- ```{r power-vs-log-effect-size-given-number-of-observations} -->
-
-<!-- polynomials %>% -->
-
-<!--   group_by(lineup_id) %>% -->
-
-<!--   summarise(across(c(type:p_value), first)) %>% -->
-
-<!--   filter(lineup_id < 577) %>% -->
-
-<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
-
-<!--   ggplot() + -->
-
-<!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
-
-<!--   geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
-
-<!--   facet_wrap(~n) + -->
-
-<!--   xlab("Natural logarithm of effect size") + -->
-
-<!--   ylab("Power") + -->
-
-<!--   ggtitle("Power comparison conditional on the number of observations") -->
-
-<!-- ``` -->
-
-<!-- ```{r power-of-visual-test-given-number-of-observations} -->
-
-<!-- polynomials %>% -->
-
-<!--   group_by(lineup_id) %>% -->
-
-<!--   summarise(across(c(type:p_value), first)) %>% -->
-
-<!--   filter(lineup_id < 577) %>% -->
-
-<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01)), n = factor(n)) %>% -->
-
-<!--   ggplot() + -->
-
-<!--   geom_smooth(aes(log(effect_size), reject, col = n), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
-
-<!--   xlab("Natural logarithm of effect size") + -->
-
-<!--   ylab("Power") + -->
-
-<!--   ggtitle("Power of visual test conditional on the number of observations") -->
-
-<!-- ``` -->
-
-<!-- ```{r} -->
-
-<!-- polynomials %>% -->
-
-<!--   group_by(lineup_id) %>% -->
-
-<!--   summarise(across(c(type:p_value), first)) %>% -->
-
-<!--   filter(lineup_id < 577) %>% -->
-
-<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
-
-<!--   ggplot() + -->
-
-<!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
-
-<!--   geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.05)), aes(log(effect_size), reject, col = "Conventional test:0.05"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
-
-<!--   geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.01)), aes(log(effect_size), reject, col = "Conventional test:0.01"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
-
-<!--   geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.005)), aes(log(effect_size), reject, col = "Conventional test:0.005"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
-
-<!--   geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.0005)), aes(log(effect_size), reject, col = "Conventional test:0.0005"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
-
-<!--   geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.0000005)), aes(log(effect_size), reject, col = "Conventional test:0.0000005"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
-
-<!--   xlab("Natural logarithm of effect size") + -->
-
-<!--   ylab("Power") -->
-
-<!-- ``` -->
+```{r}
+vi_survey %>%
+  mutate(x_dist = ifelse(x_dist == "even_discrete", "discrete uniform", x_dist)) %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
+  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  ggplot() +
+  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
+                pivot_longer(F_p_value:SW_p_value) %>%
+                mutate(name = gsub("_p_value", " test", name)) %>%
+                mutate(type = "polynomial") %>%
+                mutate(x_dist = ifelse(x_dist == "even_discrete", "discrete uniform", x_dist)) %>%
+                bind_rows(mutate(heter_conv_sim, log_effect_size = log(effect_size)) %>%
+                pivot_longer(F_p_value:SW_p_value) %>%
+                mutate(name = gsub("_p_value", " test", name)) %>%
+                mutate(type = "heteroskedasticity") %>%
+                mutate(x_dist = ifelse(x_dist == "even_discrete", "discrete uniform", x_dist))), 
+              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
+              method = "glm", 
+              method.args = list(family = binomial),
+              se = FALSE) +
+  theme_light(base_size = 5) +
+  scale_color_brewer(palette = "Dark2", direction = -1) +
+  xlab("Natural logarithm of effect size") +
+  ylab("Power") +
+  facet_grid(shape~type)
+```

### 2022-11-15 paper_comparison/paper_comparison/paper_comparison.tex
update distribution cmp

@@
 
-\hypertarget{distribution-of-regressor}{%
-\subsection{Distribution of regressor}\label{distribution-of-regressor}}
-
-The impact of the distribution of \(X_raw\) on the power is displayed in
-Figure \ref{fig:dist-power}. The power curve of F-test is stable across
-different distributions, while the visual test has a steeper power curve
-for normal and uniform distribution. BP-test performs worse for discrete
-uniform distribution and uniform distribution but has relatively high
-power for normal distribution. SW-test outperforms BP-test for discrete
+The impact of the distribution of \(X_{raw}\) on the power is displayed
+in Figure \ref{fig:dist-power}. For the non-linearity model, the power
+curve of F-test is stable across different distributions, while the
+visual test has a more gentle power curve for the discrete uniform
+distribution as it contains lineups with larger effect size. BP test
+performs worse for discrete uniform distribution but has relatively high
+power for normal distribution. SW test outperforms BP test for discrete
 uniform distribution but remains as the worst test for other
@@
 
+For the heteroskedasticity model, F-test is entirely insensitive to the
+non-constant variance. The visual test has similar power as the SW
+test\\
+for normal distribution and lognormal distribution, and has higher power
+for uniform distribution, and lower power for discrete uniform
+distribution.
+
+But still, we observe the huge gap in power between the preferred
+conventional tests and the visual test which is consistent with our
+previous argument that conventional test is too sensitive for diagnosing
+non-linearity and heteroskedasticity.
+
 \begin{figure}
 
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/dist-power-1}
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-8-1}
 
@@
 
-\caption{...}\label{fig:dist-power}
+\caption{...}\label{fig:unnamed-chunk-8}
 \end{figure}
 
---\textgreater{}
-
---\textgreater{}
+\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-9-1} \end{center}
 

### 2022-11-16 paper_comparison/paper_comparison/paper_comparison.Rmd
finish prototype of the power plot

@@
 
+```{r cache = TRUE}
+target_sigma <- c(4.51, 3.52, 2.74, 2.14, 1.66, 1.29, 1.01, 0.79, 0.61, 0.48, 0.37)
+
+es <- map_dbl(target_sigma, ~log(poly_model(shape = 2, x = rand_uniform(), sigma = .x)$effect_size(n = 100)))
+```
+
+```{r}
+vi_survey %>%
+  filter(x_dist == "uniform") %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(type == "polynomial") %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
+  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  ggplot() +
+  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  stat_smooth(geom = "line", 
+              data = ~boot_y(.x, reject_dirichlet, times = 500), 
+              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "Visual test"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, 
+              alpha = 0.03) +
+  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
+                filter(x_dist == "uniform") %>%
+                pivot_longer(F_p_value:SW_p_value) %>%
+                mutate(name = gsub("_p_value", " test", name)), 
+              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
+              method = "glm", 
+              method.args = list(family = binomial),
+              se = FALSE) +
+  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3)  +
+  geom_segment(data = data.frame(xi = es), aes(x = xi, xend = xi, y = 0, yend = 1)) +
+  theme_light(base_size = 5) +
+  scale_color_brewer(palette = "Dark2", direction = -1) +
+  xlab("Natural logarithm of effect size") +
+  ylab("Power")
+
+ex_dat <- poly_model(shape = 2, x = rand_uniform(), sigma = 1)$gen(n = 100)
+plist <- list()
+j <- 0
+
+for (i in target_sigma) {
+  j <- j + 1
+  plist[[j]] <- 
+  VI_MODEL$plot(poly_model(shape = 2, x = rand_uniform())$
+                  gen(n = 100, computed = select(mutate(ex_dat, e = sqrt(i) * e), x, e)), theme = theme_light(), 
+              remove_axis = TRUE, 
+              remove_grid_line = TRUE) +
+    ggtitle(round(es[j], 1))
+}
+
+patchwork::wrap_plots(plist)
+```
+
 The impact of the distribution of $X_{raw}$ on the power is displayed in Figure \ref{fig:dist-power}.

### 2022-11-16 paper_comparison/paper_comparison/paper_comparison.tex
finish prototype of the power plot

@@
 
+\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-6-1} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-6-2} \end{center}
+
 The impact of the distribution of \(X_{raw}\) on the power is displayed
@@
 
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-8-1}
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-9-1}
 
@@
 
-\caption{...}\label{fig:unnamed-chunk-8}
+\caption{...}\label{fig:unnamed-chunk-9}
 \end{figure}
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-9-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-10-1} \end{center}
 

### 2022-11-22 paper_comparison/paper_comparison/paper_comparison.Rmd
clean up plots in the result

@@
 
-```{r power-poly-uniform, fig.cap = "..."}
+```{r power-poly-uniform, fig.cap = "...", eval = FALSE}
 vi_survey %>%
@@
 ```
-
-```{r, fig.cap="..."}
-vi_survey %>%
-  mutate(x_dist = ifelse(x_dist == "even_discrete", "discrete uniform", x_dist)) %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
-  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
-                pivot_longer(F_p_value:SW_p_value) %>%
-                mutate(name = gsub("_p_value", " test", name)) %>%
-                mutate(type = "polynomial") %>%
-                mutate(x_dist = ifelse(x_dist == "even_discrete", "discrete uniform", x_dist)) %>%
-                bind_rows(mutate(heter_conv_sim, log_effect_size = log(effect_size)) %>%
-                pivot_longer(F_p_value:SW_p_value) %>%
-                mutate(name = gsub("_p_value", " test", name)) %>%
-                mutate(type = "heteroskedasticity") %>%
-                mutate(x_dist = ifelse(x_dist == "even_discrete", "discrete uniform", x_dist))), 
-              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
-              method = "glm", 
-              method.args = list(family = binomial),
-              se = FALSE) +
-  theme_light(base_size = 5) +
-  scale_color_brewer(palette = "Dark2", direction = -1) +
-  xlab("Natural logarithm of effect size") +
-  ylab("Power") +
-  facet_grid(x_dist~type)
-```
-
-```{r}
-vi_survey %>%
-  mutate(x_dist = ifelse(x_dist == "even_discrete", "discrete uniform", x_dist)) %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
-  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
-                pivot_longer(F_p_value:SW_p_value) %>%
-                mutate(name = gsub("_p_value", " test", name)) %>%
-                mutate(type = "polynomial") %>%
-                mutate(x_dist = ifelse(x_dist == "even_discrete", "discrete uniform", x_dist)) %>%
-                bind_rows(mutate(heter_conv_sim, log_effect_size = log(effect_size)) %>%
-                pivot_longer(F_p_value:SW_p_value) %>%
-                mutate(name = gsub("_p_value", " test", name)) %>%
-                mutate(type = "heteroskedasticity") %>%
-                mutate(x_dist = ifelse(x_dist == "even_discrete", "discrete uniform", x_dist))), 
-              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
-              method = "glm", 
-              method.args = list(family = binomial),
-              se = FALSE) +
-  theme_light(base_size = 5) +
-  scale_color_brewer(palette = "Dark2", direction = -1) +
-  xlab("Natural logarithm of effect size") +
-  ylab("Power") +
-  facet_grid(shape~type)
-```

### 2022-11-22 paper_comparison/paper_comparison/paper_comparison.tex
clean up plots in the result

--- 2022-11-15 paper_comparison/paper_comparison/paper_comparison.tex
@@
 
-\begin{figure}
+\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-7-1} \end{center}
 
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/power-poly-uniform-1}
-
-}
-
-\caption{...}\label{fig:power-poly-uniform}
-\end{figure}
+\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-7-2} \end{center}
 
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-8-1}
-
-}
-
-\caption{...}\label{fig:unnamed-chunk-8}
-\end{figure}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-9-1} \end{center}
-
 \bibliographystyle{tfcad}

### 2022-11-22 paper_comparison/paper_comparison/paper_comparison.Rmd
make better plots for result

@@
 header-includes: |
+  \usepackage{lscape}
   \usepackage{hyperref}
@@
 ```{r}
-vi_survey %>%
+p <- vi_survey %>%
   filter(x_dist == "uniform") %>%
@@
   VI_MODEL$plot(poly_model(shape = 2, x = rand_uniform())$
-                  gen(n = 100, computed = select(mutate(ex_dat, e = sqrt(i) * e), x, e)), theme = theme_light(), 
+                  gen(n = 100, computed = select(mutate(ex_dat, e = sqrt(i) * e), x, e)), theme = theme_light(base_size =  5), 
               remove_axis = TRUE, 
@@
 
-patchwork::wrap_plots(plist)
+patchwork::wrap_plots(append(plist, list(p), after = 0), 
+                      design = map(1:11, ~patchwork::area(11, .x, 11, .x)) %>%
+                        reduce(c) %>%
+                        c(patchwork::area(1, 1, 10, 11), .))
 ```
@@
 
-```{r eval = FALSE}
-vi_survey %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type == "polynomial") %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
-  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_dirichlet, times = 500), 
-              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "Visual test"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03) +
-  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
-                pivot_longer(F_p_value:SW_p_value) %>%
-                mutate(name = gsub("_p_value", " test", name)), 
-              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
-              method = "glm", 
-              method.args = list(family = binomial),
-              se = FALSE) +
-  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) +
-  theme_light(base_size = 5) +
-  scale_color_brewer(palette = "Dark2", direction = -1) +
-  xlab("Natural logarithm of effect size") +
-  ylab("Power") +
-  facet_wrap(~x_dist)
-```
-
-```{r eval = FALSE}
-vi_survey %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type == "heteroskedasticity") %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
-  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_dirichlet, times = 500), 
-              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "Visual test"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03) +
-  geom_smooth(data = mutate(heter_conv_sim, log_effect_size = log(effect_size)) %>%
-                pivot_longer(F_p_value:SW_p_value) %>%
-                mutate(name = gsub("_p_value", " test", name)), 
-              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
-              method = "glm", 
-              method.args = list(family = binomial),
-              se = FALSE) +
-  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) +
-  theme_light(base_size = 5) +
-  scale_color_brewer(palette = "Dark2", direction = -1) +
-  xlab("Natural logarithm of effect size") +
-  ylab("Power") +
-  facet_wrap(~x_dist)
-```
+\begin{landscape}
+
+```{r fig.height = 3, fig.width = 12, fig.cap="..."}
+patchwork::wrap_plots(append(plist, list(p), after = 0), 
+                      design = map(1:11, ~patchwork::area(16, .x, 22, .x)) %>%
+                        reduce(c) %>%
+                        c(patchwork::area(1, 1, 15, 11), .))
+```
+
+```{r fig.height = 3, fig.width = 12, fig.cap="..."}
+patchwork::wrap_plots(append(plist, list(p), after = 0), 
+                      design = map(1:11, ~patchwork::area(16, .x, 22, .x)) %>%
+                        reduce(c) %>%
+                        c(patchwork::area(1, 1, 15, 11), .))
+```
+
+\end{landscape}

### 2022-11-22 paper_comparison/paper_comparison/paper_comparison.tex
make better plots for result

@@
 
+\usepackage{lscape}
 \usepackage{hyperref}
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-7-2} \end{center}
-
 The impact of the distribution of \(X_{raw}\) on the power is displayed
@@
 
+\begin{landscape}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-8-1}
+
+}
+
+\caption{...}\label{fig:unnamed-chunk-8}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-9-1}
+
+}
+
+\caption{...}\label{fig:unnamed-chunk-9}
+\end{figure}
+
+\end{landscape}
+
 \bibliographystyle{tfcad}

### 2022-11-23 paper_comparison/paper_comparison/paper_comparison.Rmd
update the result plots

@@
 target_sigma <- c(4.51, 3.52, 2.74, 2.14, 1.66, 1.29, 1.01, 0.79, 0.61, 0.48, 0.37)
-
-es <- map_dbl(target_sigma, ~log(poly_model(shape = 2, x = rand_uniform(), sigma = .x)$effect_size(n = 100)))
+es <- map_dbl(target_sigma, ~log(poly_model(shape = 2, x = rand_uniform(-1, 1), sigma = .x)$effect_size(n = 100)))
 ```
@@
   geom_point(aes(log_effect_size, prop_detect), alpha = 0.3)  +
-  geom_segment(data = data.frame(xi = es), aes(x = xi, xend = xi, y = 0, yend = 1)) +
+  geom_segment(data = data.frame(xi = es), aes(x = xi, xend = xi, y = 0, yend = 1), alpha = 0.4, linetype = 2) +
   theme_light(base_size = 5) +
-  scale_color_brewer(palette = "Dark2", direction = -1) +
+  scale_color_manual(values = rev(rcartocolor::carto_pal(5, "Safe"))[2:5]) +
   xlab("Natural logarithm of effect size") +
@@
 
+```{r cache = TRUE}
+target_b <- c(0.14, 0.19, 0.27, 0.38, 0.57, 0.91, 1.53, 2.9, 6.4, 17.9, 75)
+es <- map_dbl(target_b, ~log(heter_model(a = 1, b = .x, x = rand_uniform(-1, 1))$effect_size(n = 100, type = "kl")))
+```
+
+```{r}
+p <- vi_survey %>%
+  filter(x_dist == "uniform") %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(type == "heteroskedasticity") %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
+  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  ggplot() +
+  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+  stat_smooth(geom = "line", 
+              data = ~boot_y(.x, reject_dirichlet, times = 500), 
+              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "Visual test"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, 
+              alpha = 0.03) +
+  geom_smooth(data = mutate(heter_conv_sim, log_effect_size = log(effect_size)) %>%
+                filter(x_dist == "uniform") %>%
+                pivot_longer(F_p_value:SW_p_value) %>%
+                mutate(name = gsub("_p_value", " test", name)), 
+              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
+              method = "glm", 
+              method.args = list(family = binomial),
+              se = FALSE) +
+  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3)  +
+  geom_segment(data = data.frame(xi = es), aes(x = xi, xend = xi, y = 0, yend = 1), alpha = 0.4, linetype = 2) +
+  theme_light(base_size = 5) +
+  scale_color_manual(values = rev(rcartocolor::carto_pal(5, "Safe"))[2:5]) +
+  xlab("Natural logarithm of effect size") +
+  ylab("Power")
+
+ex_dat <- heter_model(a = 1, x = rand_uniform(-1, 1), b = 1)$gen(n = 100)
+plist <- list()
+j <- 0
+
+for (i in target_b) {
+  j <- j + 1
+  plist[[j]] <-
+  VI_MODEL$plot(heter_model(a = 1, b = i,x = rand_uniform(-1, 1))$
+                  gen(n = 100, computed = select(ex_dat, x, e)), theme = theme_light(base_size =  5),
+              remove_axis = TRUE,
+              remove_grid_line = TRUE) +
+    ggtitle(round(es[j], 1))
+}
+```
+
 ```{r fig.height = 3, fig.width = 12, fig.cap="..."}

### 2022-11-25 paper_comparison/paper_comparison/paper_comparison.Rmd
update results

@@
 
-## P-value and power comparison
+## P-value comparison
 
@@
 
-Figure \ref{fig:power-poly-uniform} shows the estimated power of visual test on lineups with uniform distribution $X_{raw}$ against natural logarithm of the effect with comparison to the power of an exact test - F-test, and two other residual-based conventional tests commonly used in regression diagnostics but for testing other departures from the model assumptions.
+## Power comparison
+
+<!-- Figure \ref{fig:power-poly-uniform} shows the estimated power of visual test on lineups with uniform distribution $X_{raw}$ against natural logarithm of the effect with comparison to the power of an exact test - F-test, and two other residual-based conventional tests commonly used in regression diagnostics but for testing other departures from the model assumptions.
 In overall, the power of all four tests increases as the effect becomes larger.
@@
 The power of Breusch--Pagan test and Shapiro--Wilk test reach around 75% and 30% respectively.
-
-What truly impress us is the huge difference between the estimated power of visual test and the estimated power of F-test.
+-->
+
+<!-- What truly impress us is the huge difference between the estimated power of visual test and the estimated power of F-test.
 According to the example lineups included in Figure \ref{fig:power-poly-uniform}, at around $log_e(E) = 1$, the  designed visual feature is rarely visible, making the actual data plot indistinguishable from residual plots simulated from the assumed model at this level of difficulty.
@@
 From a communication perspective, given the fact that the visual difference is unperceivable, the argument that non-linearity present in the fitted model is less convincing to the public.
-
-The margin is largest at around $log_e(E) = 2$.
+-->
+
+<!-- The margin is largest at around $log_e(E) = 2$.
 As shown in Figure \ref{fig:power-poly-uniform}, the visual pattern is still undetectable which is similar to $log_e(E) = 1$.
@@
 In contrast, F-test will reject the null hypothesis 90% of the time.
-
-At around $log_e(E) = 3$, the margin gets smaller as the chance of identifying the actual data plot becomes larger.
+-->
+
+<!-- At around $log_e(E) = 3$, the margin gets smaller as the chance of identifying the actual data plot becomes larger.
 At this level of difficulty, the designed visual features are usually detectable but it may not stand out from the lineup as other null plots may happen to include outliers or visual patterns that are are considered to be more attractive by human, and thus recognized as the most different plot.
@@
 It is also interesting to observe that the power curve of visual test behaves similarly to those inappropriate conventional tests until the designed visual features really becomes visible.
+-->
 
@@
 }
-
-patchwork::wrap_plots(append(plist, list(p), after = 0), 
-                      design = map(1:11, ~patchwork::area(11, .x, 11, .x)) %>%
-                        reduce(c) %>%
-                        c(patchwork::area(1, 1, 10, 11), .))
-```
-
-The impact of the distribution of $X_{raw}$ on the power is displayed in Figure \ref{fig:dist-power}.
+```
+
+<!-- The impact of the distribution of $X_{raw}$ on the power is displayed in Figure \ref{fig:dist-power}.
 For the non-linearity model, the power curve of F-test is stable across different distributions, while the visual test has a more gentle power curve for the discrete uniform distribution as it contains lineups with larger effect size.
@@
 The results indicate those inappropriate residual-based tests are sensitive to the distribution of the regressor.
-
-For the heteroskedasticity model, F-test is entirely insensitive to the non-constant variance. The visual test has similar power as the SW test  
-for normal distribution and lognormal distribution, and has higher power for uniform distribution, and lower power for discrete uniform distribution.
-
-But still, we observe the huge gap in power between the preferred conventional tests and the visual test which is consistent with our previous argument that conventional test is too sensitive for diagnosing non-linearity and heteroskedasticity.
+-->
+
+<!-- For the heteroskedasticity model, F-test is entirely insensitive to the non-constant variance. The visual test has similar power as the SW test   -->
+<!-- for normal distribution and lognormal distribution, and has higher power for uniform distribution, and lower power for discrete uniform distribution. -->
+
+<!-- But still, we observe the huge gap in power between the preferred conventional tests and the visual test which is consistent with our previous argument that conventional test is too sensitive for diagnosing non-linearity and heteroskedasticity.
+-->
 
@@
 
-```{r fig.height = 3, fig.width = 12, fig.cap="..."}
+```{r polypower, fig.height = 3, fig.width = 12, fig.cap="..."}
 patchwork::wrap_plots(append(plist, list(p), after = 0), 
@@
 
-```{r fig.height = 3, fig.width = 12, fig.cap="..."}
+```{r heterpower, fig.height = 3, fig.width = 12, fig.cap="..."}
 patchwork::wrap_plots(append(plist, list(p), after = 0), 

### 2022-11-25 paper_comparison/paper_comparison/paper_comparison.tex
update the result plots
update results

--- 2022-11-22 paper_comparison/paper_comparison/paper_comparison.tex
@@
 
-\hypertarget{p-value-and-power-comparison}{%
-\subsection{P-value and power
-comparison}\label{p-value-and-power-comparison}}
+\hypertarget{p-value-comparison}{%
+\subsection{P-value comparison}\label{p-value-comparison}}
 
@@
 
-Figure \ref{fig:power-poly-uniform} shows the estimated power of visual
-test on lineups with uniform distribution \(X_{raw}\) against natural
-logarithm of the effect with comparison to the power of an exact test -
-F-test, and two other residual-based conventional tests commonly used in
-regression diagnostics but for testing other departures from the model
-assumptions. In overall, the power of all four tests increases as the
-effect becomes larger. The power curve of F-test climbs aggressively
-from 25\% to around 90\% as \(log_e(E)\) increases from 0 to 2, while
-others respond inactively to the change of effect and remain lower than
-25\% throughout the period, showing that as an exact test, the F-test is
-relatively more sensitive to the type of model defects that being
-considered. The power of visual test arises steadily and nearly linearly
-to around 90\% as \(log_e(E)\) increases from 2 to 5, suggesting that
-the effect starts to make noticeable impact on the degree of the
-presence of the designed visual features. Other two inappropriate
-conventional tests shows improvement at the same time but at a lower
-rate. This coincides the point made by \citet{cook1982residuals}
-mentioned in \ref{hypothesis-testing} that residual-based tests for a
-specific type of model defect are sensitive to other types of model
-defects. At \(log_e(E) \approx 6\), the power curve of F-test reaches
-almost 100\% followed by the visual test by a small margin. The power of
-Breusch--Pagan test and Shapiro--Wilk test reach around 75\% and 30\%
-respectively.
+\hypertarget{power-comparison}{%
+\subsection{Power comparison}\label{power-comparison}}
 
-What truly impress us is the huge difference between the estimated power
-of visual test and the estimated power of F-test. According to the
-example lineups included in Figure \ref{fig:power-poly-uniform}, at
-around \(log_e(E) = 1\), the designed visual feature is rarely visible,
-making the actual data plot indistinguishable from residual plots
-simulated from the assumed model at this level of difficulty. In fact,
-zero out of 11 subjects detect the actual data plot positioned at panel
-10. However, F-test has more than 60\% chance of rejecting the null
-hypothesis in this case. From a communication perspective, given the
-fact that the visual difference is unperceivable, the argument that
-non-linearity present in the fitted model is less convincing to the
-public.
+\begin{landscape}
 
-The margin is largest at around \(log_e(E) = 2\). As shown in Figure
-\ref{fig:power-poly-uniform}, the visual pattern is still undetectable
-which is similar to \(log_e(E) = 1\). The visual test power is close to
-zero. In contrast, F-test will reject the null hypothesis 90\% of the
-time.
+\begin{figure}
 
-At around \(log_e(E) = 3\), the margin gets smaller as the chance of
-identifying the actual data plot becomes larger. At this level of
-difficulty, the designed visual features are usually detectable but it
-may not stand out from the lineup as other null plots may happen to
-include outliers or visual patterns that are are considered to be more
-attractive by human, and thus recognized as the most different plot.
-Without knowing the designed visual features beforehand, it is actually
-hard to identify the actual data plot by pure image comparison. The
-corresponding example lineup for \(log_e(E) = 3\) shown in Figure
-\ref{fig:fig:power-poly-uniform} has the actual data plot positioned at
-panel 20, where two of 11 subjects detect it. It can be observed that a
-``S-shape'' is presented in plot 20, but the signal is not strong enough
-to attract all 11 subjects, resulting in a visual p-value sightly above
-\(0.1\). At \(log_e(E) = 4\), the designed visual features become much
-clear and attractive in panel 3, leading to a high percentage of
-rejection of the null hypothesis. Figure
-\ref{fig:fig:power-poly-uniform} gives example lineups of such cases. It
-is also interesting to observe that the power curve of visual test
-behaves similarly to those inappropriate conventional tests until the
-designed visual features really becomes visible.
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/polypower-1}
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-7-1} \end{center}
+}
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-7-2} \end{center}
+\caption{...}\label{fig:polypower}
+\end{figure}
 
-The impact of the distribution of \(X_{raw}\) on the power is displayed
-in Figure \ref{fig:dist-power}. For the non-linearity model, the power
-curve of F-test is stable across different distributions, while the
-visual test has a more gentle power curve for the discrete uniform
-distribution as it contains lineups with larger effect size. BP test
-performs worse for discrete uniform distribution but has relatively high
-power for normal distribution. SW test outperforms BP test for discrete
-uniform distribution but remains as the worst test for other
-distributions. The results indicate those inappropriate residual-based
-tests are sensitive to the distribution of the regressor.
+\begin{figure}
 
-For the heteroskedasticity model, F-test is entirely insensitive to the
-non-constant variance. The visual test has similar power as the SW
-test\\
-for normal distribution and lognormal distribution, and has higher power
-for uniform distribution, and lower power for discrete uniform
-distribution.
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/heterpower-1}
 
-But still, we observe the huge gap in power between the preferred
-conventional tests and the visual test which is consistent with our
-previous argument that conventional test is too sensitive for diagnosing
-non-linearity and heteroskedasticity.
+}
+
+\caption{...}\label{fig:heterpower}
+\end{figure}
+
+\end{landscape}
 

### 2022-11-26 paper_comparison/paper_comparison/paper_comparison.Rmd
update result text

@@
 
-<!-- Figure \ref{fig:power-poly-uniform} shows the estimated power of visual test on lineups with uniform distribution $X_{raw}$ against natural logarithm of the effect with comparison to the power of an exact test - F-test, and two other residual-based conventional tests commonly used in regression diagnostics but for testing other departures from the model assumptions.
-In overall, the power of all four tests increases as the effect becomes larger.
-The power curve of F-test climbs aggressively from 25% to around 90% as $log_e(E)$ increases from 0 to 2, while others respond inactively to the change of effect and remain lower than 25% throughout the period, showing that as an exact test, the F-test is relatively more sensitive to the type of model defects that being considered.
-The power of visual test arises steadily and nearly linearly to around 90% as $log_e(E)$ increases from 2 to 5, suggesting that the effect starts to make noticeable impact on the degree of the presence of the designed visual features.
+Figure \ref{fig:polypower} shows the estimated power of visual test on lineups produced by the non-linearity model with the $X_{raw}$ following a uniform distribution against the natural logarithm of the effect with comparison to the power of an exact test - F-test, and two other residual-based conventional tests commonly used in regression diagnostics but for testing heteroskedasticity and non-normality respectively.
+In overall, the power of all four teste eincreases as the effect becomes larger.
+The power curve of F-test climbs aggressively from 25% to around 90% as $log_e(E)$ increases from 0 to 2, while power of other tests respond inactively to the change of effect and remain lower than 25% throughout the period, showing that F-test is way more sensitive to the type of model defects that being considered.
+The power of visual test arises steadily to around 90% as $log_e(E)$ increases from 2 to 5, suggesting that the effect starts to make noticeable impact on the degree of the presence of the designed visual features.
 Other two inappropriate conventional tests shows improvement at the same time but at a lower rate.
-This coincides the point made by @cook1982residuals mentioned in \ref{hypothesis-testing} that residual-based tests for a specific type of model defect are sensitive to other types of model defects.
-At $log_e(E) \approx 6$, the power curve of F-test reaches almost 100% followed by the visual test by a small margin.
+This coincides the point made by @cook1982residuals that residual-based tests for a specific type of model defect are sensitive to other types of model defects.
+At $log_e(E) \approx 6$, the power of both F-test and visual test reaches almost 100%.
 The power of Breusch--Pagan test and Shapiro--Wilk test reach around 75% and 30% respectively.
--->
 

### 2022-11-26 paper_comparison/paper_comparison/paper_comparison.tex
update result text

@@
 
+Figure \ref{fig:polypower} shows the estimated power of visual test on
+lineups produced by the non-linearity model with the \(X_{raw}\)
+following a uniform distribution against the natural logarithm of the
+effect with comparison to the power of an exact test - F-test, and two
+other residual-based conventional tests commonly used in regression
+diagnostics but for testing heteroskedasticity and non-normality
+respectively. In overall, the power of all four teste eincreases as the
+effect becomes larger. The power curve of F-test climbs aggressively
+from 25\% to around 90\% as \(log_e(E)\) increases from 0 to 2, while
+power of other tests respond inactively to the change of effect and
+remain lower than 25\% throughout the period, showing that F-test is way
+more sensitive to the type of model defects that being considered. The
+power of visual test arises steadily to around 90\% as \(log_e(E)\)
+increases from 2 to 5, suggesting that the effect starts to make
+noticeable impact on the degree of the presence of the designed visual
+features. Other two inappropriate conventional tests shows improvement
+at the same time but at a lower rate. This coincides the point made by
+\citet{cook1982residuals} that residual-based tests for a specific type
+of model defect are sensitive to other types of model defects. At
+\(log_e(E) \approx 6\), the power of both F-test and visual test reaches
+almost 100\%. The power of Breusch--Pagan test and Shapiro--Wilk test
+reach around 75\% and 30\% respectively.
+
 \begin{landscape}

### 2022-11-27 paper_comparison/paper_comparison/paper_comparison.Rmd
give better example residual plots

@@
 
-Figure \ref{fig:polypower} shows the estimated power of visual test on lineups produced by the non-linearity model with the $X_{raw}$ following a uniform distribution against the natural logarithm of the effect with comparison to the power of an exact test - F-test, and two other residual-based conventional tests commonly used in regression diagnostics but for testing heteroskedasticity and non-normality respectively.
-In overall, the power of all four teste eincreases as the effect becomes larger.
+Figure \ref{fig:polypower} shows the estimated power of visual test on lineups produced by the non-linearity model with the $X_{raw}$ following a uniform distribution against the natural logarithm of the effect $log_e(E)$ with comparison to the power of an exact test - F-test, and two other residual-based conventional tests commonly used in regression diagnostics but for testing heteroskedasticity and non-normality respectively.
+At the bottom of the figure \ref{fig:polypower}, there are a sequence of example residuals plots with increasing levels of $log_e(E)$.
+Readers can evaluate them from left to right and determine at which level the departure from a good residual plot becomes noticeable.
+
+The power of all four tests increases as the effect becomes larger.
 The power curve of F-test climbs aggressively from 25% to around 90% as $log_e(E)$ increases from 0 to 2, while power of other tests respond inactively to the change of effect and remain lower than 25% throughout the period, showing that F-test is way more sensitive to the type of model defects that being considered.
+Meanwhile, the residuals plot
+
 The power of visual test arises steadily to around 90% as $log_e(E)$ increases from 2 to 5, suggesting that the effect starts to make noticeable impact on the degree of the presence of the designed visual features.
@@
 
+set.seed(10085)
 ex_dat <- poly_model(shape = 2, x = rand_uniform(), sigma = 1)$gen(n = 100)
@@
 
+set.seed(10085)
 ex_dat <- heter_model(a = 1, x = rand_uniform(-1, 1), b = 1)$gen(n = 100)

### 2022-11-27 paper_comparison/paper_comparison/paper_comparison.tex
give better example residual plots

--- 2022-11-25 paper_comparison/paper_comparison/paper_comparison.tex
@@
 
+Figure \ref{fig:polypower} shows the estimated power of visual test on
+lineups produced by the non-linearity model with the \(X_{raw}\)
+following a uniform distribution against the natural logarithm of the
+effect \(log_e(E)\) with comparison to the power of an exact test -
+F-test, and two other residual-based conventional tests commonly used in
+regression diagnostics but for testing heteroskedasticity and
+non-normality respectively. At the bottom of the figure
+\ref{fig:polypower}, there are a sequence of example residuals plots
+with increasing levels of \(log_e(E)\). Readers can evaluate them from
+left to right and determine at which level the departure from a good
+residual plot becomes noticeable.
+
+The power of all four tests increases as the effect becomes larger. The
+power curve of F-test climbs aggressively from 25\% to around 90\% as
+\(log_e(E)\) increases from 0 to 2, while power of other tests respond
+inactively to the change of effect and remain lower than 25\% throughout
+the period, showing that F-test is way more sensitive to the type of
+model defects that being considered. Meanwhile, the residuals plot
+
+The power of visual test arises steadily to around 90\% as \(log_e(E)\)
+increases from 2 to 5, suggesting that the effect starts to make
+noticeable impact on the degree of the presence of the designed visual
+features. Other two inappropriate conventional tests shows improvement
+at the same time but at a lower rate. This coincides the point made by
+\citet{cook1982residuals} that residual-based tests for a specific type
+of model defect are sensitive to other types of model defects. At
+\(log_e(E) \approx 6\), the power of both F-test and visual test reaches
+almost 100\%. The power of Breusch--Pagan test and Shapiro--Wilk test
+reach around 75\% and 30\% respectively.
+
 \begin{landscape}

### 2022-11-28 paper_comparison/paper_comparison/paper_comparison.Rmd
update text

@@
 At the bottom of the figure \ref{fig:polypower}, there are a sequence of example residuals plots with increasing levels of $log_e(E)$.
-Readers can evaluate them from left to right and determine at which level the departure from a good residual plot becomes noticeable.
-
-The power of all four tests increases as the effect becomes larger.
+Readers can evaluate them from left to right and determine at which level the departure from a good residual plot becomes detectable.
+
+In overall, the power of all four tests increases as the effect gets larger.
 The power curve of F-test climbs aggressively from 25% to around 90% as $log_e(E)$ increases from 0 to 2, while power of other tests respond inactively to the change of effect and remain lower than 25% throughout the period, showing that F-test is way more sensitive to the type of model defects that being considered.
-Meanwhile, the residuals plot
-
-The power of visual test arises steadily to around 90% as $log_e(E)$ increases from 2 to 5, suggesting that the effect starts to make noticeable impact on the degree of the presence of the designed visual features.
+Meanwhile, no noticeable visual features can be spotted from the example residuals plots.
+
+The power of visual test arises steadily to around 90% as $log_e(E)$ increases from 2 to 5, suggesting that the effect starts to make significant impact on the degree of the presence of the designed visual features.
+This can also be observed from the example residuals plots that when $log_e(E) = 2.5$, a weak "S-shape" starts to present.
+And as $log_e(E)$ increases, the visual pattern becomes much clearer.
 Other two inappropriate conventional tests shows improvement at the same time but at a lower rate.
 This coincides the point made by @cook1982residuals that residual-based tests for a specific type of model defect are sensitive to other types of model defects.
-At $log_e(E) \approx 6$, the power of both F-test and visual test reaches almost 100%.
-The power of Breusch--Pagan test and Shapiro--Wilk test reach around 75% and 30% respectively.
+
+At $log_e(E) \approx 6$, the power of both F-test and visual test reaches almost 100%. The power of Breusch--Pagan test and Shapiro--Wilk test reach around 75% and 30% respectively.
+
+Similarly, Figure \ref{fig:heterpower} shows the power of BP test, visual test and two other
 

### 2022-11-28 paper_comparison/paper_comparison/paper_comparison.tex
update text

@@
 left to right and determine at which level the departure from a good
-residual plot becomes noticeable.
+residual plot becomes detectable.
 
-The power of all four tests increases as the effect becomes larger. The
-power curve of F-test climbs aggressively from 25\% to around 90\% as
-\(log_e(E)\) increases from 0 to 2, while power of other tests respond
-inactively to the change of effect and remain lower than 25\% throughout
-the period, showing that F-test is way more sensitive to the type of
-model defects that being considered. Meanwhile, the residuals plot
+In overall, the power of all four tests increases as the effect gets
+larger. The power curve of F-test climbs aggressively from 25\% to
+around 90\% as \(log_e(E)\) increases from 0 to 2, while power of other
+tests respond inactively to the change of effect and remain lower than
+25\% throughout the period, showing that F-test is way more sensitive to
+the type of model defects that being considered. Meanwhile, no
+noticeable visual features can be spotted from the example residuals
+plots.
 
@@
 increases from 2 to 5, suggesting that the effect starts to make
-noticeable impact on the degree of the presence of the designed visual
-features. Other two inappropriate conventional tests shows improvement
-at the same time but at a lower rate. This coincides the point made by
+significant impact on the degree of the presence of the designed visual
+features. This can also be observed from the example residuals plots
+that when \(log_e(E) = 2.5\), a weak ``S-shape'' starts to present. And
+as \(log_e(E)\) increases, the visual pattern becomes much clearer.
+Other two inappropriate conventional tests shows improvement at the same
+time but at a lower rate. This coincides the point made by
 \citet{cook1982residuals} that residual-based tests for a specific type
-of model defect are sensitive to other types of model defects. At
-\(log_e(E) \approx 6\), the power of both F-test and visual test reaches
-almost 100\%. The power of Breusch--Pagan test and Shapiro--Wilk test
-reach around 75\% and 30\% respectively.
+of model defect are sensitive to other types of model defects.
+
+At \(log_e(E) \approx 6\), the power of both F-test and visual test
+reaches almost 100\%. The power of Breusch--Pagan test and Shapiro--Wilk
+test reach around 75\% and 30\% respectively.
+
+Similarly, Figure \ref{fig:heterpower} shows the power of BP test,
+visual test and two other
 

### 2022-11-29 paper_comparison/paper_comparison/paper_comparison.Rmd
update power cmp

@@
 
-Figure \ref{fig:polypower} shows the estimated power of visual test on lineups produced by the non-linearity model with the $X_{raw}$ following a uniform distribution against the natural logarithm of the effect $log_e(E)$ with comparison to the power of an exact test - F-test, and two other residual-based conventional tests commonly used in regression diagnostics but for testing heteroskedasticity and non-normality respectively.
+Figure \ref{fig:polypower} show the estimated power of visual test on lineups produced by the non-linearity model with the $X_{raw}$ following a uniform distribution, against the natural logarithm of the effect $log_e(E)$ with comparison to the power of an exact test - F-test, and two other residual-based conventional tests commonly used in regression diagnostics but for testing heteroskedasticity and non-normality respectively.
 At the bottom of the figure \ref{fig:polypower}, there are a sequence of example residuals plots with increasing levels of $log_e(E)$.
@@
 
-In overall, the power of all four tests increases as the effect gets larger.
+Figure \ref{fig:heterpower} is similar to Figure \ref{fig:polypower}, but shows corresponding information on lineups produced by the heteroskedasticity model.
+In this scenario, the visual test is compared to an approximate test - Breusch-Pagan test, and two other inappropriate tests - F-test and Shapiro-Wilk test.
+
+For the non-linearity model, the power of all four tests increases as the effect gets larger.
 The power curve of F-test climbs aggressively from 25% to around 90% as $log_e(E)$ increases from 0 to 2, while power of other tests respond inactively to the change of effect and remain lower than 25% throughout the period, showing that F-test is way more sensitive to the type of model defects that being considered.
@@
 
-The power of visual test arises steadily to around 90% as $log_e(E)$ increases from 2 to 5, suggesting that the effect starts to make significant impact on the degree of the presence of the designed visual features.
-This can also be observed from the example residuals plots that when $log_e(E) = 2.5$, a weak "S-shape" starts to present.
-And as $log_e(E)$ increases, the visual pattern becomes much clearer.
-Other two inappropriate conventional tests shows improvement at the same time but at a lower rate.
-This coincides the point made by @cook1982residuals that residual-based tests for a specific type of model defect are sensitive to other types of model defects.
-
-At $log_e(E) \approx 6$, the power of both F-test and visual test reaches almost 100%. The power of Breusch--Pagan test and Shapiro--Wilk test reach around 75% and 30% respectively.
-
-Similarly, Figure \ref{fig:heterpower} shows the power of BP test, visual test and two other
+In terms of the heteroskedasticity model, the power of Breusch-Pagan test is also almost always greater than the power of visual test.
+For $0 \leq log_e(E) \leq 2$, where the visual feature is nearly unobservable from the example residual plots and the power curve of the visual test remains at a low level, the Breusch-Pagan test still have a decent amount of chance of rejecting the null hypothesis.
+
+The power of visual test arises steadily as $log_e(E)$ increases from 2 to 5 in both non-linearity model and heteroskedasticity model, suggesting that the effect starts to make significant impact on the degree of the presence of the designed visual features.
+This can also be observed from the example residuals plots that when $log_e(E) = 2.5$, a weak "S-shape" and a weak "triangle" shape are presented in Figure \ref{fig:polypower} and Figure \ref{fig:heterpower} respectively.
+And as $log_e(E)$ increases, the visual pattern becomes much clearer, and the power reaches almost 100% at $log_e(E) \approx 6$.
+
+The power of all inappropriate tests except for F-test shows improvement as the effect increases but at a lower rate than the visual test in both scenarios.
+This coincides the point made by @cook1982residuals that residual-based tests for a specific type of model defect may be sensitive to other types of model defects.
+The power curve of F-test remains at around 5% in Figure \ref{fig:heterpower} since there is no non-linear term leave out in the heteroskedasticity model and
+the null hypothesis of the F-test will be always satisfied.
 
@@
 
-set.seed(10085)
+set.seed(10089)
 ex_dat <- heter_model(a = 1, x = rand_uniform(-1, 1), b = 1)$gen(n = 100)

### 2022-11-29 paper_comparison/paper_comparison/paper_comparison.tex
update power cmp

@@
 
-Figure \ref{fig:polypower} shows the estimated power of visual test on
+Figure \ref{fig:polypower} show the estimated power of visual test on
 lineups produced by the non-linearity model with the \(X_{raw}\)
-following a uniform distribution against the natural logarithm of the
+following a uniform distribution, against the natural logarithm of the
 effect \(log_e(E)\) with comparison to the power of an exact test -
@@
 
-In overall, the power of all four tests increases as the effect gets
-larger. The power curve of F-test climbs aggressively from 25\% to
-around 90\% as \(log_e(E)\) increases from 0 to 2, while power of other
-tests respond inactively to the change of effect and remain lower than
-25\% throughout the period, showing that F-test is way more sensitive to
-the type of model defects that being considered. Meanwhile, no
-noticeable visual features can be spotted from the example residuals
-plots.
+Figure \ref{fig:heterpower} is similar to Figure \ref{fig:polypower},
+but shows corresponding information on lineups produced by the
+heteroskedasticity model. In this scenario, the visual test is compared
+to an approximate test - Breusch-Pagan test, and two other inappropriate
+tests - F-test and Shapiro-Wilk test.
 
-The power of visual test arises steadily to around 90\% as \(log_e(E)\)
-increases from 2 to 5, suggesting that the effect starts to make
-significant impact on the degree of the presence of the designed visual
-features. This can also be observed from the example residuals plots
-that when \(log_e(E) = 2.5\), a weak ``S-shape'' starts to present. And
-as \(log_e(E)\) increases, the visual pattern becomes much clearer.
-Other two inappropriate conventional tests shows improvement at the same
-time but at a lower rate. This coincides the point made by
-\citet{cook1982residuals} that residual-based tests for a specific type
-of model defect are sensitive to other types of model defects.
+For the non-linearity model, the power of all four tests increases as
+the effect gets larger. The power curve of F-test climbs aggressively
+from 25\% to around 90\% as \(log_e(E)\) increases from 0 to 2, while
+power of other tests respond inactively to the change of effect and
+remain lower than 25\% throughout the period, showing that F-test is way
+more sensitive to the type of model defects that being considered.
+Meanwhile, no noticeable visual features can be spotted from the example
+residuals plots.
 
-At \(log_e(E) \approx 6\), the power of both F-test and visual test
-reaches almost 100\%. The power of Breusch--Pagan test and Shapiro--Wilk
-test reach around 75\% and 30\% respectively.
+In terms of the heteroskedasticity model, the power of Breusch-Pagan
+test is also almost always greater than the power of visual test. For
+\(0 \leq log_e(E) \leq 2\), where the visual feature is nearly
+unobservable from the example residual plots and the power curve of the
+visual test remains at a low level, the Breusch-Pagan test still have a
+decent amount of chance of rejecting the null hypothesis.
 
-Similarly, Figure \ref{fig:heterpower} shows the power of BP test,
-visual test and two other
+The power of visual test arises steadily as \(log_e(E)\) increases from
+2 to 5 in both non-linearity model and heteroskedasticity model,
+suggesting that the effect starts to make significant impact on the
+degree of the presence of the designed visual features. This can also be
+observed from the example residuals plots that when \(log_e(E) = 2.5\),
+a weak ``S-shape'' and a weak ``triangle'' shape are presented in Figure
+\ref{fig:polypower} and Figure \ref{fig:heterpower} respectively. And as
+\(log_e(E)\) increases, the visual pattern becomes much clearer, and the
+power reaches almost 100\% at \(log_e(E) \approx 6\).
+
+The power of all inappropriate tests except for F-test shows improvement
+as the effect increases but at a lower rate than the visual test in both
+scenarios. This coincides the point made by \citet{cook1982residuals}
+that residual-based tests for a specific type of model defect may be
+sensitive to other types of model defects. The power curve of F-test
+remains at around 5\% in Figure \ref{fig:heterpower} since there is no
+non-linear term leave out in the heteroskedasticity model and the null
+hypothesis of the F-test will be always satisfied.
 

### 2022-11-29 paper_comparison/paper_comparison/paper_comparison.Rmd
finish power cmp

@@
 
-<!-- What truly impress us is the huge difference between the estimated power of visual test and the estimated power of F-test.
-According to the example lineups included in Figure \ref{fig:power-poly-uniform}, at around $log_e(E) = 1$, the  designed visual feature is rarely visible, making the actual data plot indistinguishable from residual plots simulated from the assumed model at this level of difficulty.
-In fact, zero out of 11 subjects detect the actual data plot positioned at panel 10.
-However, F-test has more than 60% chance of rejecting the null hypothesis in this case.
-From a communication perspective, given the fact that the visual difference is unperceivable, the argument that non-linearity present in the fitted model is less convincing to the public.
--->
-
-<!-- The margin is largest at around $log_e(E) = 2$.
-As shown in Figure \ref{fig:power-poly-uniform}, the visual pattern is still undetectable which is similar to $log_e(E) = 1$.
-The visual test power is close to zero.
-In contrast, F-test will reject the null hypothesis 90% of the time.
--->
-
-<!-- At around $log_e(E) = 3$, the margin gets smaller as the chance of identifying the actual data plot becomes larger.
-At this level of difficulty, the designed visual features are usually detectable but it may not stand out from the lineup as other null plots may happen to include outliers or visual patterns that are are considered to be more attractive by human, and thus recognized as the most different plot.
-Without knowing the designed visual features beforehand, it is actually hard to identify the actual data plot by pure image comparison.
-The corresponding example lineup for $log_e(E) = 3$ shown in Figure \ref{fig:fig:power-poly-uniform} has the actual data plot positioned at panel 20, where two of 11 subjects detect it.
-It can be observed that a "S-shape" is presented in plot 20, but the signal is not strong enough to attract all 11 subjects, resulting in a visual p-value sightly above $0.1$.
-At $log_e(E) = 4$, the designed visual features become much clear and attractive in panel 3, leading to a high percentage of rejection of the null hypothesis.
-Figure \ref{fig:fig:power-poly-uniform} gives example lineups of such cases.
-It is also interesting to observe that the power curve of visual test behaves similarly to those inappropriate conventional tests until the designed visual features really becomes visible.
--->
+Overall, the power analysis suggests that conventional tests do not match how humans perceive residual plots and thus determine model violations in these two scenarios.
+Non-severe model violations with human acceptable residual departures could often alarm by the conventional tests.
+This could limit the popularity of conventional tests in residual diagnostics among analysts.
 
@@
               se = FALSE) +
-  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3)  +
+  geom_point(data = ~.x %>% 
+               group_by(log_effect_size, reject_dirichlet) %>% 
+               summarise(num_response = n()),
+             aes(log_effect_size, reject_dirichlet, size = factor(num_response)), alpha = 0.3)  +
   geom_segment(data = data.frame(xi = es), aes(x = xi, xend = xi, y = 0, yend = 1), alpha = 0.4, linetype = 2) +
@@
   scale_color_manual(values = rev(rcartocolor::carto_pal(5, "Safe"))[2:5]) +
+  scale_size_manual(values = c(1, 2, 3)) +
   xlab("Natural logarithm of effect size") +
-  ylab("Power")
+  ylab("Power") +
+  labs(col = "", size = "#lineups")
 
@@
                         reduce(c) %>%
-                        c(patchwork::area(1, 1, 15, 11), .))
+                        c(patchwork::area(3, 1, 15, 11), .))
 ```
@@
               se = FALSE) +
-  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3)  +
+    geom_point(data = ~.x %>% 
+               group_by(log_effect_size, reject_dirichlet) %>% 
+               summarise(num_response = n()),
+             aes(log_effect_size, reject_dirichlet, size = factor(num_response)), alpha = 0.3)  +
   geom_segment(data = data.frame(xi = es), aes(x = xi, xend = xi, y = 0, yend = 1), alpha = 0.4, linetype = 2) +

### 2022-11-29 paper_comparison/paper_comparison/paper_comparison.tex
finish power cmp

@@
 
+Overall, the power analysis suggests that conventional tests do not
+match how humans perceive residual plots and thus determine model
+violations in these two scenarios. Non-severe model violations with
+human acceptable residual departures could often alarm by the
+conventional tests. This could limit the popularity of conventional
+tests in residual diagnostics among analysts.
+
 \begin{landscape}

### 2022-11-30 paper_comparison/paper_comparison/paper_comparison.Rmd
update power plots
add plot selection section

@@
 \end{landscape}
+
+## Visual test
+
+Note that both the p-value and power comparison are not directly based on the diagnostic plots recommended by abundance of literature, but
+
+## Lineup selection
+
+```{r cache = TRUE}
+vi_survey_with_plot <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  left_join(map_df(1:length(vi_lineup), 
+    ~vi_lineup[[.x]]$data %>% 
+      group_by(k) %>% 
+      summarise(plot_p_value = first(p_value)) %>%
+      arrange(k) %>%
+      mutate(unique_lineup_id = names(vi_lineup)[.x]) %>%
+      mutate(plot_p_value_rank = rank(plot_p_value, ties.method = "min"))) %>%
+  pivot_wider(values_from = c(plot_p_value, plot_p_value_rank), names_from = k, names_prefix = "plot_")) %>%
+  (function(x) {
+      for (i in 1:20)
+        x[[paste0("plot_sel_plot_", i)]] <-
+          grepl(paste0("_", i, "_"), x$selection) |
+          grepl(paste0("^", i, "_"), x$selection) |
+          grepl(paste0("_", i, "$"), x$selection) |
+          grepl(paste0("^", i, "$"), x$selection)
+      x}
+    ) %>%
+  pivot_longer(plot_p_value_plot_1:plot_sel_plot_20, 
+               names_to = c(".value", "plot"), 
+               names_pattern = "(.*)_plot_(.+)") 
+```
+
+```{r dataplotsel, fig.cap = "..."}
+vi_survey_with_plot %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(answer == plot) %>%
+  group_by(unique_lineup_id) %>%
+  summarise(plot_p_value_rank = first(plot_p_value_rank),
+            weighted_detect = sum(weighted_detect),
+            n_eval = n()) %>%
+  group_by(plot_p_value_rank) %>%
+  summarise(weighted_detect = sum(weighted_detect),
+            n_eval = sum(n_eval)) %>%
+  ggplot() +
+  geom_col(aes(factor(plot_p_value_rank), n_eval)) +
+  geom_col(aes(factor(plot_p_value_rank), weighted_detect, fill = "weighted number of detections")) +
+  xlab("Data plot p-value rank") +
+  ylab("Number of Lineups") +
+  theme(legend.position = "bottom") +
+  labs(fill = "")
+```
+
+```{r dadtaplotselfacet, fig.cap = "..."}
+vi_survey_with_plot %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  left_join(vi_survey_with_plot %>%
+              filter(!attention_check) %>%
+              filter(!null_lineup) %>%
+              filter(answer == plot) %>%
+              group_by(unique_lineup_id) %>%
+              summarise(answer_rank = first(plot_p_value_rank))) %>%
+  mutate(weighted_sel = ifelse(num_selection == 0, 0.05, plot_sel/num_selection)) %>%
+  group_by(plot_p_value_rank, answer_rank) %>%
+  summarise(weighted_sel = sum(weighted_sel)) %>%
+  ggplot() +
+  geom_col(aes(factor(plot_p_value_rank), weighted_sel)) +
+  geom_vline(data = data.frame(answer_rank = 1:20, x = 1:20),
+             aes(xintercept = x),
+             col = "red") +
+  facet_wrap(~answer_rank, scales = "free_y") +
+  xlab("Plot p-value rank") +
+  ylab("Weighted number of selections")
+```
+
+```{r eval = FALSE}
+# data plot not always has the lowest conventional p-value
+vi_survey_with_plot %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(answer == plot) %>%
+  ggplot() +
+  geom_bar(aes(factor(plot_p_value_rank)))
+
+# according to the rank of the data plot p-value, how many people select it 
+vi_survey_with_plot %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(plot_sel) %>%
+  left_join(vi_survey_with_plot %>%
+              filter(!attention_check) %>%
+              filter(!null_lineup) %>%
+              filter(answer == plot) %>%
+              group_by(unique_lineup_id) %>%
+              summarise(answer_rank = first(plot_p_value_rank))) %>%
+  ggplot() +
+  geom_bar(aes(factor(plot_p_value_rank))) +
+  geom_vline(data = data.frame(answer_rank = 1:20, x = 1:20),
+             aes(xintercept = x),
+             col = "red") +
+  facet_wrap(~answer_rank, scales = "free_y")
+
+# For different rank, percentage of subject select it
+vi_survey_with_plot %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(answer == plot) %>%
+  group_by(unique_lineup_id, set) %>%
+  summarise(plot_p_value_rank = first(plot_p_value_rank),
+            detect = first(detect),
+            type = first(type)) %>%
+  ggplot() +
+  geom_bar(aes(factor(plot_p_value_rank), fill = detect), position = "fill") +
+  geom_hline(yintercept = 0.05) +
+  xlab("Data plot p-value rank") +
+  ylab("Propotion of subject detecting the data plot")
+
+vi_survey_with_plot %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(answer == plot) %>%
+  group_by(unique_lineup_id, set) %>%
+  summarise(plot_p_value_rank = first(plot_p_value_rank),
+            prop_detect = first(prop_detect)) %>%
+  ggplot() +
+  geom_boxplot(aes(factor(plot_p_value_rank), prop_detect))
+
+# vi_survey_with_plot %>%
+#   filter(!attention_check) %>%
+#   filter(!null_lineup) %>%
+#   filter(answer == plot) %>%
+#   group_by(unique_lineup_id, set) %>%
+#   summarise(plot_p_value_rank = first(plot_p_value_rank),
+#             conv_detect = first(conventional_p_value < 0.05)) %>%
+#   ggplot() +
+#   geom_bar(aes(factor(plot_p_value_rank), fill = conv_detect), position = "fill") +
+#   xlab("Data plot p-value rank") +
+#   ylab("Propotion of subject detecting the data plot")
+```
+
+\pagebreak

### 2022-11-30 paper_comparison/paper_comparison/paper_comparison.tex
update power plots
add plot selection section

@@
 
+\hypertarget{visual-test}{%
+\subsection{Visual test}\label{visual-test}}
+
+Note that both the p-value and power comparison are not directly based
+on the diagnostic plots recommended by abundance of literature, but
+
+\hypertarget{lineup-selection}{%
+\subsection{Lineup selection}\label{lineup-selection}}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/dataplotsel-1}
+
+}
+
+\caption{...}\label{fig:dataplotsel}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/dadtaplotselfacet-1}
+
+}
+
+\caption{...}\label{fig:dadtaplotselfacet}
+\end{figure}
+
+\pagebreak
+
 \bibliographystyle{tfcad}

### 2022-12-01 paper_comparison/paper_comparison/paper_comparison.Rmd
cleaned up table 2

@@
            x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$", "NA"),
-           sigma = c("0.25", "1", "2", "4", ""),
+           sigma = c("0.25", "1.00", "2.00", "4.00", ""),
            a = c("-1", "0", "1", "", ""),
-           b = c("0.25", "1", "4", "16", "64"),
+           b = c("0.25", "1.00", "4.00", "16.00", "64.00"),
            n = c("50", "100", "300", "", "")) %>%
   knitr::kable(col.names = NULL,
-               align = "cccccc",
+               align = "rcrrrr",
                escape = FALSE,
@@
                booktabs = TRUE, 
-               caption  = 'Summary of factors involved in the study.', 
+               caption  = 'Description of all factors involved in the nonlinear and heteroskedasticity studies.', 
                label = "model-factor-table" ) %>%
   kableExtra::kable_styling(latex_options = "scale_down") %>%
-  kableExtra::add_header_above(c("Order of Hermite polynomial\n(sym1)", 
+  kableExtra::add_header_above(c("Poly Order (sym1)", #"Order of Hermite polynomial\n(sym1)", 
                                  "Distribution of sym2\n", 
-                                 "Standard deviation of polynomial model\n(sym3)",
-                                 "Shape of heteroskedasticity\n(sym4)", 
-                                 "Variance factor of heteroskedasticity model\n(sym5)",
-                                 "Sample size\n(sym6)")) %>%
+                                 "SD (sym3)", #"Standard deviation of polynomial model\n(sym3)",
+                                 "Poly Shape (sym4)", # "Shape of heteroskedasticity\n(sym4)", 
+                                 "Heteroskedasticity (sym5)", # Variance factor of heteroskedasticity model\n(sym5)",
+                                 "Size (sym6)")) %>% #"Sample size\n(sym6)")) %>%
   as.character() %>%

### 2022-12-01 paper_comparison/paper_comparison/paper_comparison.tex
cleaned up table 2

@@
 
-\caption{\label{tab:model-factor-table}Summary of factors involved in the study.}
+\caption{\label{tab:model-factor-table}Description of all factors involved in the nonlinear and heteroskedasticity studies.}
 \centering
 \resizebox{\linewidth}{!}{
-\begin{tabular}[t]{cccccc}
+\begin{tabular}[t]{rcrrrr}
 \toprule
-\multicolumn{1}{c}{\makecell[c]{Order of Hermite polynomial\\($j$)}} & \multicolumn{1}{c}{\makecell[c]{Distribution of $X_{raw}$\\}} & \multicolumn{1}{c}{\makecell[c]{Standard deviation of polynomial model\\($\sigma$)}} & \multicolumn{1}{c}{\makecell[c]{Shape of heteroskedasticity\\($a$)}} & \multicolumn{1}{c}{\makecell[c]{Variance factor of heteroskedasticity model\\($b$)}} & \multicolumn{1}{c}{\makecell[c]{Sample size\\($n$)}} \\
+\multicolumn{1}{c}{Poly Order ($j$)} & \multicolumn{1}{c}{\makecell[c]{Distribution of $X_{raw}$\\}} & \multicolumn{1}{c}{SD ($\sigma$)} & \multicolumn{1}{c}{Poly Shape ($a$)} & \multicolumn{1}{c}{Heteroskedasticity ($b$)} & \multicolumn{1}{c}{Size ($n$)} \\
 \cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
 2 & $U(-1, 1)$ & 0.25 & -1 & 0.25 & 50\\
-3 & $N(0, 0.3^2)$ & 1 & 0 & 1 & 100\\
-6 & $lognormal(0, 0.6^2)/3$ & 2 & 1 & 4 & 300\\
-18 & $U\{1, 5\}$ & 4 &  & 16 & \\
- & NA &  &  & 64 & \\
+3 & $N(0, 0.3^2)$ & 1.00 & 0 & 1.00 & 100\\
+6 & $lognormal(0, 0.6^2)/3$ & 2.00 & 1 & 4.00 & 300\\
+18 & $U\{1, 5\}$ & 4.00 &  & 16.00 & \\
+ & NA &  &  & 64.00 & \\
 \bottomrule

### 2022-12-01 paper_comparison/paper_comparison/paper_comparison.Rmd
figures 12 and 13 tweaked a lot

@@
 
-where $B(.)$ is the beta function defined as
+\noindent where $B(.)$ is the beta function defined as
 
@@
 ```
-where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
+\noindent where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
 
@@
 ```
-where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$.
+\noindent where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$.
 
@@
 
-where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vectors of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $\{-k, k\}$ defined as
+\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vectors of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $\{-k, k\}$ defined as
 
@@
 ```
-where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.
+\noindent where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.
 
@@
 
-where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vectors of size $n$ and $g(.)$ is the scaling function defined in \ref{eq:scaling-function}.
+\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vectors of size $n$ and $g(.)$ is the scaling function defined in \ref{eq:scaling-function}.
 
@@
   scale_size_manual(values = c(1, 2, 3)) +
-  xlab("Natural logarithm of effect size") +
+  xlab("Effect size (log e)") +
   ylab("Power") +
@@
   VI_MODEL$plot(poly_model(shape = 2, x = rand_uniform())$
-                  gen(n = 100, computed = select(mutate(ex_dat, e = sqrt(i) * e), x, e)), theme = theme_light(base_size =  5), 
+                  gen(n = 100, computed = select(mutate(ex_dat, e = sqrt(i) * e), x, e)), 
+                theme = theme_light(base_size =  5), 
               remove_axis = TRUE, 
-              remove_grid_line = TRUE) +
-    ggtitle(round(es[j], 1))
+              remove_grid_line = TRUE, size=0.2) +
+    ggtitle(round(es[j], 1)) +
+    theme(aspect.ratio=0.8)
 }
@@
 
-\begin{landscape}
-
-```{r polypower, fig.height = 3, fig.width = 12, fig.cap="..."}
-patchwork::wrap_plots(append(plist, list(p), after = 0), 
-                      design = map(1:11, ~patchwork::area(16, .x, 22, .x)) %>%
+%\begin{landscape}
+
+```{r polypower, fig.height = 3, fig.width = 5, fig.cap="Comparison of power between different tests for nonlinear patterns.
+Main plot shows the power curves, with dots indicating human evaluations of lineups.
+Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
+Where would you draw the line of too much nonlinearity in the residuals?
+For the F test (should this be RESET??) this is around log effect size 0.5, but for the visual test it is around 3."}
+patchwork::wrap_plots(append(plist, list(p), 
+                             after = 0), 
+                      design = map(1:11, ~patchwork::area(21, .x, 22, .x)) %>%
                         reduce(c) %>%
-                        c(patchwork::area(3, 1, 15, 11), .))
+                        c(patchwork::area(3, 1, 20, 11), .))
 ```
@@
               remove_axis = TRUE,
-              remove_grid_line = TRUE) +
-    ggtitle(round(es[j], 1))
+              remove_grid_line = TRUE, size=0.2) +
+    ggtitle(round(es[j], 1)) +
+    theme(aspect.ratio = 0.8)
 }
@@
 
-```{r heterpower, fig.height = 3, fig.width = 12, fig.cap="..."}
+```{r heterpower, fig.height = 3, fig.width = 5, fig.cap="..."}
 patchwork::wrap_plots(append(plist, list(p), after = 0), 
-                      design = map(1:11, ~patchwork::area(16, .x, 22, .x)) %>%
+                      design = map(1:11, ~patchwork::area(21, .x, 22, .x)) %>%
                         reduce(c) %>%
-                        c(patchwork::area(3, 1, 15, 11), .))
-```
-
-\end{landscape}
+                        c(patchwork::area(3, 1, 20, 11), .))
+```
+
+%\end{landscape}
 

### 2022-12-01 paper_comparison/paper_comparison/paper_comparison.tex
figures 12 and 13 tweaked a lot

@@
 
-where \(B(.)\) is the beta function defined as
+\noindent where \(B(.)\) is the beta function defined as
 
@@
 
-where \(\Lambda(.)\) is the standard logistic function given as
-\(\Lambda(z) = exp(z)/(1+exp(z))\).
+\noindent where \(\Lambda(.)\) is the standard logistic function given
+as \(\Lambda(z) = exp(z)/(1+exp(z))\).
 
@@
 
-where \(diag(.)\) is the diagonal matrix constructed from the diagonal
-elements of \(\boldsymbol{R}_a\).
+\noindent where \(diag(.)\) is the diagonal matrix constructed from the
+diagonal elements of \(\boldsymbol{R}_a\).
 
@@
 
-where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
+\noindent where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
 \(\boldsymbol{\varepsilon}\), \(\boldsymbol{x}_{raw}\),
@@
 
-where
+\noindent where
 \(\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\).
@@
 
-where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
+\noindent where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
 \(\boldsymbol{\varepsilon}\) are vectors of size \(n\) and \(g(.)\) is
@@
 
-\begin{landscape}
+\%\textbackslash begin\{landscape\}
 
@@
 
-\caption{...}\label{fig:polypower}
+\caption{Comparison of power between different tests for nonlinear patterns.
+Main plot shows the power curves, with dots indicating human evaluations of lineups.
+Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
+Where would you draw the line of too much nonlinearity in the residuals?
+For the F test (should this be RESET??) this is around log effect size 0.5, but for the visual test it is around 3.}\label{fig:polypower}
 \end{figure}
@@
 
-\end{landscape}
+\%\textbackslash end\{landscape\}
 

### 2022-12-02 paper_comparison/paper_comparison/paper_comparison.Rmd
removed landscape mode
update conventional simulation (RESET test)

--- 2022-11-30 paper_comparison/paper_comparison/paper_comparison.Rmd
@@
                             F_p_value = mod$test(tmp_dat)$p_value,
+                            RESET_p_value = mod$test(tmp_dat, test = "RESET", power = 2:3, power_type = "fitted")$p_value,
                             BP_p_value = HETER_MODEL$test(tmp_dat)$p_value,
@@
                             F_p_value = POLY_MODEL$test(mutate(tmp_dat, z = poly_model()$gen(n, computed = select(tmp_dat, x))$z))$p_value,
+                            RESET_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:3, power_type = "fitted")$p_value,
                             BP_p_value = mod$test(tmp_dat)$p_value,
@@
             n_eval = sum(n_eval)) %>%
+  mutate(plot_p_value_rank_g5 = plot_p_value_rank >= 5) %>%
+  group_by(plot_p_value_rank_g5) %>%
+  mutate(n_eval = ifelse(plot_p_value_rank_g5, sum(n_eval), n_eval)) %>%
+  mutate(weighted_detect = ifelse(plot_p_value_rank_g5, sum(weighted_detect), weighted_detect)) %>%
+  ungroup() %>%
+  filter(plot_p_value_rank <= 5) %>%
+  mutate(plot_p_value_rank = ifelse(plot_p_value_rank == 5, "5+", as.character(plot_p_value_rank))) %>%
+  mutate(plot_p_value_rank = factor(plot_p_value_rank, levels = c(as.character(1:5), "5+"))) %>%
   ggplot() +
-  geom_col(aes(factor(plot_p_value_rank), n_eval)) +
-  geom_col(aes(factor(plot_p_value_rank), weighted_detect, fill = "weighted number of detections")) +
+  geom_col(aes(plot_p_value_rank, n_eval)) +
+  geom_col(aes(plot_p_value_rank, weighted_detect, fill = "weighted number of detections")) +
   xlab("Data plot p-value rank") +
@@
 
-```{r dadtaplotselfacet, fig.cap = "..."}
+```{r dadtaplotselfacet, fig.cap = "...", fig.height=12, fig.width=16}
 vi_survey_with_plot %>%
@@
   filter(!null_lineup) %>%
+  filter(type == "polynomial") %>%
   left_join(vi_survey_with_plot %>%
@@
   summarise(weighted_sel = sum(weighted_sel)) %>%
+  mutate(answer_rank_g5 = answer_rank >= 5) %>%
+  group_by(plot_p_value_rank, answer_rank_g5) %>%
+  mutate(weighted_sel = ifelse(answer_rank_g5, sum(weighted_sel), weighted_sel)) %>%
+  ungroup() %>%
+  filter(answer_rank <= 5) %>%
+  mutate(answer_rank = ifelse(answer_rank == 5, "5+", as.character(answer_rank))) %>%
+  mutate(answer_rank = factor(answer_rank, levels = c(as.character(1:5), "5+"))) %>%
   ggplot() +
   geom_col(aes(factor(plot_p_value_rank), weighted_sel)) +
-  geom_vline(data = data.frame(answer_rank = 1:20, x = 1:20),
+  geom_vline(data = data.frame(answer_rank = c("1", "2", "3", "4"), x = 1:4),
              aes(xintercept = x),
@@
   ylab("Weighted number of selections")
+```
+
+```{r eval = FALSE}
+vi_survey_with_plot %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(answer == plot) %>%
+  group_by(unique_lineup_id) %>%
+  summarise(plot_p_value_rank = first(plot_p_value_rank),
+            weighted_detect = sum(weighted_detect),
+            n_eval = n(),
+            reject = first(conventional_p_value < 0.05)) %>%
+  group_by(plot_p_value_rank, reject) %>%
+  summarise(weighted_detect = sum(weighted_detect),
+            n_eval = sum(n_eval)) %>%
+  ggplot() +
+  geom_col(aes(factor(plot_p_value_rank), n_eval)) +
+  geom_col(aes(factor(plot_p_value_rank), weighted_detect, fill = "weighted number of detections")) +
+  xlab("Data plot p-value rank") +
+  ylab("Number of Lineups") +
+  theme(legend.position = "bottom") +
+  labs(fill = "") +
+  facet_wrap(~reject)
 ```

### 2022-12-02 paper_comparison/paper_comparison/paper_comparison.Rmd

--- 2022-12-01 paper_comparison/paper_comparison/paper_comparison.Rmd
@@
 
-%\begin{landscape}
-
 ```{r polypower, fig.height = 3, fig.width = 5, fig.cap="Comparison of power between different tests for nonlinear patterns.
@@
 
-%\end{landscape}
-
 ## Visual test

### 2022-12-02 paper_comparison/paper_comparison/paper_comparison.Rmd
removed reference to RESET in fig 12

@@
 Where would you draw the line of too much nonlinearity in the residuals?
-For the F test (should this be RESET??) this is around log effect size 0.5, but for the visual test it is around 3."}
+For the F test this is around log effect size 0.5, but for the visual test it is around 3."}
 patchwork::wrap_plots(append(plist, list(p), 
@@
 ```
-
-\pagebreak

### 2022-12-02 paper_comparison/paper_comparison/paper_comparison.Rmd
fix compilation issue

@@
   theme_light(base_size = 5) +
-  scale_color_manual(values = rev(rcartocolor::carto_pal(5, "Safe"))[2:5]) +
+  scale_color_manual(values = rev(rcartocolor::carto_pal(6, "Safe"))[2:6]) +
   scale_size_manual(values = c(1, 2, 3)) +

### 2022-12-03 paper_comparison/paper_comparison/paper_comparison.Rmd
Use power 4 for the RESET test

@@
                             F_p_value = mod$test(tmp_dat)$p_value,
-                            RESET_p_value = mod$test(tmp_dat, test = "RESET", power = 2:3, power_type = "fitted")$p_value,
+                            RESET3_p_value = mod$test(tmp_dat, test = "RESET", power = 2:3, power_type = "fitted")$p_value,
+                            RESET4_p_value = mod$test(tmp_dat, test = "RESET", power = 2:4, power_type = "fitted")$p_value,
+                            RESET5_p_value = mod$test(tmp_dat, test = "RESET", power = 2:5, power_type = "fitted")$p_value,
+                            RESET6_p_value = mod$test(tmp_dat, test = "RESET", power = 2:6, power_type = "fitted")$p_value,
+                            RESET7_p_value = mod$test(tmp_dat, test = "RESET", power = 2:7, power_type = "fitted")$p_value,
+                            RESET8_p_value = mod$test(tmp_dat, test = "RESET", power = 2:8, power_type = "fitted")$p_value,
+                            RESET9_p_value = mod$test(tmp_dat, test = "RESET", power = 2:9, power_type = "fitted")$p_value,
+                            RESET10_p_value = mod$test(tmp_dat, test = "RESET", power = 2:10, power_type = "fitted")$p_value,
                             BP_p_value = HETER_MODEL$test(tmp_dat)$p_value,
@@
                             F_p_value = POLY_MODEL$test(mutate(tmp_dat, z = poly_model()$gen(n, computed = select(tmp_dat, x))$z))$p_value,
-                            RESET_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:3, power_type = "fitted")$p_value,
+                            RESET3_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:3, power_type = "fitted")$p_value,
+                            RESET4_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:4, power_type = "fitted")$p_value,
+                            RESET5_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:5, power_type = "fitted")$p_value,
+                            RESET6_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:6, power_type = "fitted")$p_value,
+                            RESET7_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:7, power_type = "fitted")$p_value,
+                            RESET8_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:8, power_type = "fitted")$p_value,
+                            RESET9_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:9, power_type = "fitted")$p_value,
+                            RESET10_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:10, power_type = "fitted")$p_value,
                             BP_p_value = mod$test(tmp_dat)$p_value,
@@
   geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
+                select(-(RESET5_p_value:RESET10_p_value)) %>%
+                select(-RESET3_p_value) %>%
+                rename(RESET_p_value = RESET4_p_value) %>%
                 filter(x_dist == "uniform") %>%
@@
 es <- map_dbl(target_b, ~log(heter_model(a = 1, b = .x, x = rand_uniform(-1, 1))$effect_size(n = 100, type = "kl")))
+```
+
+```{r eval = FALSE}
+
+# For checking RESET test
+plotly::ggplotly(ggplot() +
+geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
+                filter(x_dist == "uniform") %>%
+                # select(-(RESET4_p_value:RESET10_p_value)) %>%
+                pivot_longer(F_p_value:RESET10_p_value) %>%
+                mutate(name = gsub("_p_value", " test", name)), 
+              aes(log_effect_size, as.numeric(value < 0.05), col = factor(name, levels = c(paste0("RESET", 3:10, " test"), "F test"))), 
+              method = "glm", 
+              method.args = list(family = binomial),
+              se = FALSE) +
+  scale_color_brewer("YlGnBu") +
+  labs(col = "") +
+  xlab("Log of effect size") +
+  ylab("Power") +
+  ggtitle("Power of the RESET test under different powers of the fitted values"))
 ```
@@
                 filter(x_dist == "uniform") %>%
+                select(-(RESET5_p_value:RESET10_p_value)) %>%
+                select(-RESET3_p_value) %>%
+                rename(RESET_p_value = RESET4_p_value) %>%
                 pivot_longer(F_p_value:SW_p_value) %>%

### 2022-12-06 paper_comparison/paper_comparison/paper_comparison.Rmd
modify col pal

@@
                 select(-(RESET5_p_value:RESET10_p_value)) %>%
-                select(-RESET3_p_value) %>%
+                select(-RESET3_p_value, -F_p_value) %>%
                 rename(RESET_p_value = RESET4_p_value) %>%
                 filter(x_dist == "uniform") %>%
-                pivot_longer(F_p_value:SW_p_value) %>%
+                pivot_longer(RESET_p_value:SW_p_value) %>%
                 mutate(name = gsub("_p_value", " test", name)), 
@@
   theme_light(base_size = 5) +
-  scale_color_manual(values = rev(rcartocolor::carto_pal(6, "Safe"))[2:6]) +
+  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
   scale_size_manual(values = c(1, 2, 3)) +
@@
                 select(-(RESET5_p_value:RESET10_p_value)) %>%
-                select(-RESET3_p_value) %>%
+                select(-RESET3_p_value, -F_p_value) %>%
                 rename(RESET_p_value = RESET4_p_value) %>%
-                pivot_longer(F_p_value:SW_p_value) %>%
+                pivot_longer(RESET_p_value:SW_p_value) %>%
                 mutate(name = gsub("_p_value", " test", name)), 
@@
                names_pattern = "(.*)_plot_(.+)") 
+```
+
+```{r plotsel, fig.cap=" A large amount of subjects select the residuals plot with the lowest p-value. But greater than 50% of subjects evenly select other plots."}
+vi_survey_with_plot %>%
+  mutate(weighted_sel = ifelse(num_selection == 0, 0.05, plot_sel/num_selection)) %>%
+  group_by(type) %>%
+  mutate(n_eval = n()/20) %>%
+  ungroup() %>%
+  group_by(type, plot_p_value_rank) %>%
+  summarise(weighted_sel = sum(weighted_sel), n_eval = first(n_eval)) %>%
+  ggplot() +
+  geom_col(aes(plot_p_value_rank, n_eval), fill = "grey", alpha = 0.3) +
+  geom_col(aes(plot_p_value_rank, weighted_sel)) +
+  facet_wrap(~type) +
+  theme_light() +
+  ylab("Weighted number of selections") +
+  xlab("Plot p-value rank")
 ```

### 2022-12-06 paper_comparison/paper_comparison/paper_comparison.tex
removed landscape mode
fix compilation issue
modify col pal

@@
 
-\%\textbackslash begin\{landscape\}
-
 \begin{figure}
@@
 Where would you draw the line of too much nonlinearity in the residuals?
-For the F test (should this be RESET??) this is around log effect size 0.5, but for the visual test it is around 3.}\label{fig:polypower}
+For the F test this is around log effect size 0.5, but for the visual test it is around 3.}\label{fig:polypower}
 \end{figure}
@@
 
-\%\textbackslash end\{landscape\}
-
 \hypertarget{visual-test}{%
@@
 \subsection{Lineup selection}\label{lineup-selection}}
+
+\textbackslash begin\{figure\}
+
+\{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/plotsel-1}
+
+\}
+
+\textbackslash caption\{ A large amount of subjects select the residuals
+plot with the lowest p-value. But greater than 50\% of subjects evenly
+select other plots.\}\label{fig:plotsel} \textbackslash end\{figure\}
 
@@
 
-\pagebreak
-
 \bibliographystyle{tfcad}

### 2022-12-07 paper_comparison/paper_comparison/paper_comparison.Rmd
clean the preprocessing code

@@
 ```{r}
+# OOP supports needed by `visuage`
+# remotes::install_github("TengMCing/bandicoot")
+# 
+# Visual inference models and p-value calculation
+# remotes::install_github("TengMCing/visage")
+
 library(tidyverse)
-# remotes::install_github("TengMCing/visage")
 library(visage)
 
+# To control the simulation in this file
 set.seed(10086)
@@
 ```{r get-lineup-data}
+# The lineup data used to draw residual plot needed to be downloaded
+# from the github repo. Cache it.
 if (!file.exists(here::here("data/vi_lineup.rds"))) {
@@
 ```{r poly-conventional-simulation}
+# Ensure the support of the regressor is [-1, 1]
 stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
 
+# Run simulations to get the behaviours of conventional tests
 if (!file.exists(here::here("data/poly_conventional_simulation.rds"))) {
   poly_conv_sim <- list()
+
+  # 100 bootstrap samples
   for (i in 1:100)
   {
-    poly_conv_sim[[i]] <- map(1:2000,
-                            function(x) {
+    poly_conv_sim[[i]] <- 
+      # Every sample contains 2000 lineups
+      map(1:2000, function(i) {
+
+        # Sample a set of parameters
         shape <- sample(1:4, 1)
         e_sigma <- sample(c(0.5, 1, 2, 4), 1)
-                     x_dist <- sample(c("uniform", "normal", "lognormal", "even_discrete"), 1)
+        x_dist <- sample(c("uniform", 
+                           "normal", 
+                           "lognormal", 
+                           "even_discrete"), 1)
         x <- switch(x_dist,
                     uniform = rand_uniform(-1, 1),
-                                 normal = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))},
-                                 lognormal = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))},
+                    normal = {
+                      raw_x <- rand_normal(sigma = 0.3)
+                      closed_form(~stand_dist(raw_x))
+                      },
+                    lognormal = {
+                      raw_x <- rand_lognormal(sigma = 0.6)
+                      closed_form(~stand_dist(raw_x/3 - 1))
+                      },
                     even_discrete = rand_uniform_d(k = 5, even = TRUE))
+
+        # Build the model
         mod <- poly_model(shape, x = x, sigma = e_sigma)
+
+        # Sample the number of observations
         n <- sample(c(50, 100, 300), 1)
+
+        # Generate data from the model 
         tmp_dat <- mod$gen(n)
 
+        # Get the effect size from the pre-calculated table 
         es <- filter(vi_survey, 
@@
 
+        # Return a data frame containing p-values of
+        # F, RESET, BP and SW tests
         tibble(shape = shape,
@@
                F_p_value = mod$test(tmp_dat)$p_value,
-                            RESET3_p_value = mod$test(tmp_dat, test = "RESET", power = 2:3, power_type = "fitted")$p_value,
-                            RESET4_p_value = mod$test(tmp_dat, test = "RESET", power = 2:4, power_type = "fitted")$p_value,
-                            RESET5_p_value = mod$test(tmp_dat, test = "RESET", power = 2:5, power_type = "fitted")$p_value,
-                            RESET6_p_value = mod$test(tmp_dat, test = "RESET", power = 2:6, power_type = "fitted")$p_value,
-                            RESET7_p_value = mod$test(tmp_dat, test = "RESET", power = 2:7, power_type = "fitted")$p_value,
-                            RESET8_p_value = mod$test(tmp_dat, test = "RESET", power = 2:8, power_type = "fitted")$p_value,
-                            RESET9_p_value = mod$test(tmp_dat, test = "RESET", power = 2:9, power_type = "fitted")$p_value,
-                            RESET10_p_value = mod$test(tmp_dat, test = "RESET", power = 2:10, power_type = "fitted")$p_value,
+               RESET3_p_value = mod$test(tmp_dat, 
+                                         test = "RESET", 
+                                         power = 2:3, 
+                                         power_type = "fitted")$p_value,
+               RESET4_p_value = mod$test(tmp_dat, 
+                                         test = "RESET", 
+                                         power = 2:4, 
+                                         power_type = "fitted")$p_value,
+               RESET5_p_value = mod$test(tmp_dat, 
+                                         test = "RESET", 
+                                         power = 2:5, 
+                                         power_type = "fitted")$p_value,
+               RESET6_p_value = mod$test(tmp_dat, 
+                                         test = "RESET", 
+                                         power = 2:6, 
+                                         power_type = "fitted")$p_value,
+               RESET7_p_value = mod$test(tmp_dat, 
+                                         test = "RESET", 
+                                         power = 2:7, 
+                                         power_type = "fitted")$p_value,
+               RESET8_p_value = mod$test(tmp_dat, 
+                                         test = "RESET", 
+                                         power = 2:8, 
+                                         power_type = "fitted")$p_value,
+               RESET9_p_value = mod$test(tmp_dat, 
+                                         test = "RESET", 
+                                         power = 2:9, 
+                                         power_type = "fitted")$p_value,
+               RESET10_p_value = mod$test(tmp_dat, 
+                                          test = "RESET", 
+                                          power = 2:10, 
+                                          power_type = "fitted")$p_value,
                BP_p_value = HETER_MODEL$test(tmp_dat)$p_value,
@@
 ```{r heter-conventional-simulation}
-stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
-
+# Run simulations to get the behaviours of conventional tests
 if (!file.exists(here::here("data/heter_conventional_simulation.rds"))) {
   heter_conv_sim <- list()
+
+  # 100 bootstrap samples
   for (i in 1:100)
   {
-    heter_conv_sim[[i]] <- map(1:2000,
-                            function(x) {
+    heter_conv_sim[[i]] <-
+      # Every sample contains 2000 lineups
+      map(1:2000, function(x) {
+
+        # Sample a set of parameters
         a <- sample(c(-1, 0, 1), 1)
         b <- sample(c(0.25, 1, 4, 16, 64), 1)
-                     x_dist <- sample(c("uniform", "normal", "lognormal", "even_discrete"), 1)
+        x_dist <- sample(c("uniform", 
+                           "normal", 
+                           "lognormal", 
+                           "even_discrete"), 1)
         x <- switch(x_dist,
                     uniform = rand_uniform(-1, 1),
-                                 normal = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))},
-                                 lognormal = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))},
+                    normal = {
+                      raw_x <- rand_normal(sigma = 0.3)
+                      closed_form(~stand_dist(raw_x))
+                      },
+                    lognormal = {
+                      raw_x <- rand_lognormal(sigma = 0.6)
+                      closed_form(~stand_dist(raw_x/3 - 1))
+                      },
                     even_discrete = rand_uniform_d(-1, 1, k = 5, even = TRUE))
+
+        # Build the model
         mod <- heter_model(a = a, b = b, x = x)
+
+        # Sample the number of observations
         n <- sample(c(50, 100, 300), 1)
+
+        # Generate data from the model 
         tmp_dat <- mod$gen(n)
 
+        # Get the effect size from the pre-calculated table 
         es <- filter(vi_survey, 
@@
 
+        # Return a data frame containing p-values of
+        # F, RESET, BP and SW tests
         tibble(a = a,
@@
                effect_size = es,
-                            F_p_value = POLY_MODEL$test(mutate(tmp_dat, z = poly_model()$gen(n, computed = select(tmp_dat, x))$z))$p_value,
-                            RESET3_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:3, power_type = "fitted")$p_value,
-                            RESET4_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:4, power_type = "fitted")$p_value,
-                            RESET5_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:5, power_type = "fitted")$p_value,
-                            RESET6_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:6, power_type = "fitted")$p_value,
-                            RESET7_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:7, power_type = "fitted")$p_value,
-                            RESET8_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:8, power_type = "fitted")$p_value,
-                            RESET9_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:9, power_type = "fitted")$p_value,
+               F_p_value = POLY_MODEL$test(
+
+                 # Create a pseudo z to be able to use F-test
+                 tmp_dat %>%
+                   mutate(z = poly_model() %>%
+                            {.$gen(n, computed = select(tmp_dat, x))} %>%
+                            pull(z))
+                 )$p_value,
+               RESET3_p_value = POLY_MODEL$test(tmp_dat, 
+                                                test = "RESET", 
+                                                power = 2:3, 
+                                                power_type = "fitted")$p_value,
+               RESET4_p_value = POLY_MODEL$test(tmp_dat, 
+                                                test = "RESET", 
+                                                power = 2:4, 
+                                                power_type = "fitted")$p_value,
+               RESET5_p_value = POLY_MODEL$test(tmp_dat, 
+                                                test = "RESET", 
+                                                power = 2:5, 
+                                                power_type = "fitted")$p_value,
+               RESET6_p_value = POLY_MODEL$test(tmp_dat, 
+                                                test = "RESET", 
+                                                power = 2:6, 
+                                                power_type = "fitted")$p_value,
+               RESET7_p_value = POLY_MODEL$test(tmp_dat, 
+                                                test = "RESET", 
+                                                power = 2:7, 
+                                                power_type = "fitted")$p_value,
+               RESET8_p_value = POLY_MODEL$test(tmp_dat, 
+                                                test = "RESET", 
+                                                power = 2:8, 
+                                                power_type = "fitted")$p_value,
+               RESET9_p_value = POLY_MODEL$test(tmp_dat, 
+                                                test = "RESET", 
+                                                power = 2:9, 
+                                                power_type = "fitted")$p_value,
                RESET10_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:10, power_type = "fitted")$p_value,
@@
 However, correctly judging a residual plot where no pattern exists is a painstakingly difficult task for humans.
-It is especially common, particularly among new data analysts to report patterns when an experienced data analyst might quickly conclude that there are none.
-[ET: I don't know if the former statement is true?] Generally, one looks for noticeable departures from the model like non-linear dependency or heteroskedasticity.
+
+It is especially common, particularly among new statistics students to report patterns when an experienced data analyst might quickly conclude that there are none.
+[ET: I don't know if the former statement is true?
+>> PL: change to students] Generally, one looks for noticeable departures from the model like non-linear dependency or heteroskedasticity.
 It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and use a Breusch-Pagan test [@breusch_simple_1979] for heteroskedasticity.
@@
 The paper is structured as follows.
-The next background section describes the types of departures that one expects to detect, and describes a formal process for reading residual plots, called visual inference, that can avoid the concerns about subjectiveness of human readers [ET: CITATION HERE?].
-Section \ref{experimental-design} describes the experimental setup to compare the decision made by formal hypothesis testing, and how humans would read diagnostic plots.
+The next background section describes the types of departures that one expects to detect, and describes a formal statistical process for reading residual plots, called visual inference [ET: CITATION HERE?
+>> PL: arguments removed].
+Section \ref{experimental-design} describes the experiment design to compare the decision made by formal hypothesis testing, and how humans would read diagnostic plots.
 The results are reported in Section \ref{results}.
@@
 
-## Visual test
-
-Note that both the p-value and power comparison are not directly based on the diagnostic plots recommended by abundance of literature, but
+<!-- ## Visual test  -->
+
+<!-- Note that both the p-value and power comparison are not directly based on the diagnostic plots recommended by abundance of literature, but  -->
 
@@
 
-```{r plotsel, fig.cap=" A large amount of subjects select the residuals plot with the lowest p-value. But greater than 50% of subjects evenly select other plots."}
+```{r plotsel, fig.cap = "A large amount of subjects select the residuals plot with the lowest p-value. But greater than 50 percent of subjects evenly select other plots."}
 vi_survey_with_plot %>%

### 2022-12-07 paper_comparison/paper_comparison/paper_comparison.tex
clean the preprocessing code

--- 2022-12-01 paper_comparison/paper_comparison/paper_comparison.tex
@@
 plot where no pattern exists is a painstakingly difficult task for
-humans. It is especially common, particularly among new data analysts to
+humans.
+
+It is especially common, particularly among new statistics students to
 report patterns when an experienced data analyst might quickly conclude
 that there are none. {[}ET: I don't know if the former statement is
-true?{]} Generally, one looks for noticeable departures from the model
-like non-linear dependency or heteroskedasticity. It is also possible to
-conduct hypothesis tests for non-linear dependence
-\citep{ramsey_tests_1969}, and use a Breusch-Pagan test
-\citep{breusch_simple_1979} for heteroskedasticity.
+true? \textgreater\textgreater{} PL: change to students{]} Generally,
+one looks for noticeable departures from the model like non-linear
+dependency or heteroskedasticity. It is also possible to conduct
+hypothesis tests for non-linear dependence \citep{ramsey_tests_1969},
+and use a Breusch-Pagan test \citep{breusch_simple_1979} for
+heteroskedasticity.
 
@@
 describes the types of departures that one expects to detect, and
-describes a formal process for reading residual plots, called visual
-inference, that can avoid the concerns about subjectiveness of human
-readers {[}ET: CITATION HERE?{]}. Section \ref{experimental-design}
-describes the experimental setup to compare the decision made by formal
-hypothesis testing, and how humans would read diagnostic plots. The
-results are reported in Section \ref{results}. We finish with a
-discussion on future work, in particular how the responsibility for
-residual plot reading might be passed on to computer vision.
+describes a formal statistical process for reading residual plots,
+called visual inference {[}ET: CITATION HERE? \textgreater\textgreater{}
+PL: arguments removed{]}. Section \ref{experimental-design} describes
+the experiment design to compare the decision made by formal hypothesis
+testing, and how humans would read diagnostic plots. The results are
+reported in Section \ref{results}. We finish with a discussion on future
+work, in particular how the responsibility for residual plot reading
+might be passed on to computer vision.
 
@@
 
-\%\textbackslash end\{landscape\}
-
-\hypertarget{visual-test}{%
-\subsection{Visual test}\label{visual-test}}
-
-Note that both the p-value and power comparison are not directly based
-on the diagnostic plots recommended by abundance of literature, but
-
 \hypertarget{lineup-selection}{%
 \subsection{Lineup selection}\label{lineup-selection}}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/plotsel-1}
+
+}
+
+\caption{A large amount of subjects select the residuals plot with the lowest p-value. But greater than 50 percent of subjects evenly select other plots.}\label{fig:plotsel}
+\end{figure}
 

### 2022-12-08 paper_comparison/paper_comparison/paper_comparison.Rmd
update 2.1

@@
                                                 power_type = "fitted")$p_value,
-               RESET10_p_value = POLY_MODEL$test(tmp_dat, test = "RESET", power = 2:10, power_type = "fitted")$p_value,
+               RESET10_p_value = POLY_MODEL$test(tmp_dat, 
+                                                 test = "RESET", 
+                                                 power = 2:10, 
+                                                 power_type = "fitted")$p_value,
                BP_p_value = mod$test(tmp_dat)$p_value,
@@
 
-It is especially common, particularly among new statistics students to report patterns when an experienced data analyst might quickly conclude that there are none.
+It is especially common, particularly among new data analytics students to report patterns when an experienced data analyst might quickly conclude that there are none.
 [ET: I don't know if the former statement is true?
@@
 ## Departures from good residual plots
-
-(This section discusses the visual patterns data analysts expect to see and their implications.)
 
@@
     ggtitle("D. Non-normality") +
-    xlab("fitted values") + ylab("residuals")
+    xlab("Fitted values") + ylab("Residuals")
 )
@@
 
-Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals.
+Graphical summaries in which residuals are plotted against fitted values or other functions of the regressors that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals.
 As shown in Figure \ref{fig:residual-plot-common-departures}, the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns.
-There are various types of departures from a good residual plot.
-
-Non-linearity, heterskedasticity and non-normality are perhaps the three mostly checked departures.
+
+There are various types of departures from a good residual plot. Non-linearity, heterskedasticity and non-normality are perhaps the three mostly checked departures.
+
 Non-linearity is a type of model misspecification caused by failing to include higher order terms of the regressors in the regression equation.
 Any non-linear functional form of residuals on fitted values presented in the residual plot could be considered as an indicative of non-linearity.
-An example residual plot containing visual pattern of non-linearity is given at the top-right of Figure  \ref{fig:commondepartures}.
+An example residual plot containing visual pattern of non-linearity is given at plot B of Figure \ref{fig:residual-plot-common-departures}.
 One can clearly observe the "S-shape" from the residual plot as the cubic term is not captured by the misspecified model.
@@
 The usual pattern of heteroskedasticity on a residual plot is the inconsistent spread of the residuals across the horizontal axis.
-Visually, it sometimes results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure \ref{fig:commondepartures}, or the "left-triangle" and "right-triangle" shape where the smallest variance occurs at one side of the horizontal axis.
+Visually, it sometimes results in the so-called "butterfly" shape as shown in the plot C of Figure \ref{fig:residual-plot-common-departures}, or the "left-triangle" and "right-triangle" shape where the smallest variance occurs at one side of the horizontal axis.
 
@@
   table_dat %>%
-  knitr::kable(col.names = c("Plot", "Departures", "Model structural", "Breusch-Pagan", "Shapiro–Wilk"),
+  knitr::kable(col.names = c("Plot", "Departures", "RESET", "Breusch-Pagan", "Shapiro–Wilk"),
                align = "llrrr",
@@
                caption  = 'Statistical significance testing for departures from good residuals for plots in Figure \\ref{fig:residual-plot-common-departures}.
-Shown are the $p$-values calculated for the conventional Model structural, the Breusch-Pagan and the Shapiro–Wilk tests.
+Shown are the $p$-values calculated for the conventional RESET, the Breusch-Pagan and the Shapiro–Wilk tests.
 The good residual plot (A) is judged a good residual plot, as expected, by all tests.

### 2022-12-08 paper_comparison/paper_comparison/paper_comparison.tex
update 2.1

@@
 
-It is especially common, particularly among new statistics students to
-report patterns when an experienced data analyst might quickly conclude
-that there are none. {[}ET: I don't know if the former statement is
-true? \textgreater\textgreater{} PL: change to students{]} Generally,
-one looks for noticeable departures from the model like non-linear
-dependency or heteroskedasticity. It is also possible to conduct
-hypothesis tests for non-linear dependence \citep{ramsey_tests_1969},
-and use a Breusch-Pagan test \citep{breusch_simple_1979} for
-heteroskedasticity.
+It is especially common, particularly among new data analytics students
+to report patterns when an experienced data analyst might quickly
+conclude that there are none. {[}ET: I don't know if the former
+statement is true? \textgreater\textgreater{} PL: change to students{]}
+Generally, one looks for noticeable departures from the model like
+non-linear dependency or heteroskedasticity. It is also possible to
+conduct hypothesis tests for non-linear dependence
+\citep{ramsey_tests_1969}, and use a Breusch-Pagan test
+\citep{breusch_simple_1979} for heteroskedasticity.
 
@@
 
-(This section discusses the visual patterns data analysts expect to see
-and their implications.)
-
 \begin{figure}
@@
 Graphical summaries in which residuals are plotted against fitted values
-or other functions of the predictor variables that are approximately
-orthogonal to residuals are referred to as standard residual plots in
+or other functions of the regressors that are approximately orthogonal
+to residuals are referred to as standard residual plots in
 \citet{cook1982residuals}. As shown in Figure
@@
 residual plot with residuals evenly distributed at both sides of the
-horizontal zero line showing no noticeable patterns. There are various
-types of departures from a good residual plot.
+horizontal zero line showing no noticeable patterns.
 
+There are various types of departures from a good residual plot.
 Non-linearity, heterskedasticity and non-normality are perhaps the three
-mostly checked departures. Non-linearity is a type of model
-misspecification caused by failing to include higher order terms of the
-regressors in the regression equation. Any non-linear functional form of
-residuals on fitted values presented in the residual plot could be
-considered as an indicative of non-linearity. An example residual plot
-containing visual pattern of non-linearity is given at the top-right of
-Figure \ref{fig:commondepartures}. One can clearly observe the
+mostly checked departures.
+
+Non-linearity is a type of model misspecification caused by failing to
+include higher order terms of the regressors in the regression equation.
+Any non-linear functional form of residuals on fitted values presented
+in the residual plot could be considered as an indicative of
+non-linearity. An example residual plot containing visual pattern of
+non-linearity is given at plot B of Figure
+\ref{fig:residual-plot-common-departures}. One can clearly observe the
 ``S-shape'' from the residual plot as the cubic term is not captured by
@@
 Visually, it sometimes results in the so-called ``butterfly'' shape as
-shown in the bottom-left panel of Figure \ref{fig:commondepartures}, or
-the ``left-triangle'' and ``right-triangle'' shape where the smallest
+shown in the plot C of Figure \ref{fig:residual-plot-common-departures},
+or the ``left-triangle'' and ``right-triangle'' shape where the smallest
 variance occurs at one side of the horizontal axis.
@@
 \caption{\label{tab:example-residual-plot-table}Statistical significance testing for departures from good residuals for plots in Figure \ref{fig:residual-plot-common-departures}.
-Shown are the $p$-values calculated for the conventional Model structural, the Breusch-Pagan and the Shapiro–Wilk tests.
+Shown are the $p$-values calculated for the conventional RESET, the Breusch-Pagan and the Shapiro–Wilk tests.
 The good residual plot (A) is judged a good residual plot, as expected, by all tests.
@@
 \toprule
-Plot & Departures & Model structural & Breusch-Pagan & Shapiro–Wilk\\
+Plot & Departures & RESET & Breusch-Pagan & Shapiro–Wilk\\
 \midrule

### 2022-12-08 paper_comparison/paper_comparison/paper_comparison.Rmd
clean code in 2.1

@@
                  tmp_dat %>%
-                   mutate(z = poly_model() %>%
-                            {.$gen(n, computed = select(tmp_dat, x))} %>%
+                   mutate(z = poly_model()$
+                            gen(n, computed = select(tmp_dat, x)) %>%
                             pull(z))
@@
 
+# Base model
 mod <- poly_model(include_z = FALSE, sigma = 0.25)
+
+# Base data
 dat <- mod$gen(300)
+
 patchwork::wrap_plots(
+
+  # Plot base data
   mod$plot(dat, theme = theme_light()) +
     ggtitle("A. Good residuals") + 
-    xlab("fitted values") + ylab("residuals"),
-  mod$plot(mod$set_prm("include_z", TRUE)$gen(300, 
-    computed = select(dat, x, e)), 
+    xlab("fitted values") + 
+    ylab("residuals"),
+
+  # Include the z term
+  # Reuse the x and e terms
+  mod$plot(mod$
+             set_prm("include_z", TRUE)$
+             gen(300, computed = select(dat, x, e)), 
     theme = theme_light()) + 
     ggtitle("B. Non-linearity") +
-    xlab("fitted values") + ylab("residuals"),
-  mod$plot(heter_model(b = 64)$gen(300, 
-    computed = select(dat, x, e)), 
+    xlab("fitted values") + 
+    ylab("residuals"),
+
+  # Build a heter model
+  # Reuse the x and e terms
+  mod$plot(heter_model(b = 64)$
+             gen(300, computed = select(dat, x, e)), 
     theme = theme_light()) + 
     ggtitle("C. Heteroskedasticity") +
-    xlab("fitted values") + ylab("residuals"),
-  mod$plot(heter_model(b = 0, 
-    e = rand_lognormal(sigma = 0.5))$gen(300, 
-    computed = select(dat, x)), 
+    xlab("fitted values") + 
+    ylab("residuals"),
+
+  # Build a heter model with non-normal error
+  # Reuse the x term
+  mod$plot(heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
+             gen(300, computed = select(dat, x)), 
     theme = theme_light()) + 
     ggtitle("D. Non-normality") +
-    xlab("Fitted values") + ylab("Residuals")
+    xlab("Fitted values") + 
+    ylab("Residuals")
 )

### 2022-12-08 paper_comparison/paper_comparison/paper_comparison.Rmd
clean code and update text in 2.2

@@
 
-(This section discusses the tests that will be used in the analysis and shows the results for the residual plots displayed in the previous section.)
-
 Other than checking diagnostic plots, analysts may perform formal hypothesis testing for detecting model defects.
@@
 For testing non-linearity, one may apply the F-test as a model structural test to examine the significance of specific polynomial and non-linear forms of the regressors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969].
-And for testing normality, the Shapiro--Wilk test [ref here] is perhaps the most widely used test included by many of the statistical softwares.
-Another choice will be the Jarque--Bera test [ref here] which directly checks if the sample skewness and kurtosis match a normal distribution.
-
-Example residual plots given in Figure \ref{fig:residual-plot-common-departures} are examined by the corresponding model structural test, Breusch-Pagan test and Shapiro--Wilk test as shown in Table \ref{tab:example-residual-plot-table}.
+And for testing normality, the Shapiro--Wilk test [@shapiro1965analysis] is perhaps the most widely used test included by many of the statistical softwares.
+Another choice will be the Jarque--Bera test [@jarque1980efficient] which directly checks if the sample skewness and kurtosis match a normal distribution.
+
+Example residual plots given in Figure \ref{fig:residual-plot-common-departures} are examined by the corresponding RESET test, Breusch-Pagan test and Shapiro--Wilk test as shown in Table \ref{tab:example-residual-plot-table}.
 In the example, both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null hypothesis for departures that they do not intend to examine.
@@
 Additionally, outliers will often incorrectly trigger the rejection of the null hypothesis despite when majority of the residuals are well-behaved [@cook_applied_1999].
-This can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers.
+Furthermore, with a sufficiently large sample size, residual-based tests may reject the null hypothesis due to a slight departure that is of little practical significance.
+These can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers and slight departures.
 
 ```{r}
-
 set.seed(10086)
 
+# Base model
 mod <- poly_model(include_z = FALSE, sigma = 0.25)
-dat <- mod$gen(300)
-
-dat1 <- mod$set_prm("include_z", TRUE)$gen(300, computed = select(dat, x, e))
-dat2 <- heter_model(b = 64)$gen(300, computed = select(dat, x, e))
+
+# Base data
+dat0 <- mod$gen(300)
+
+# Replicate data in Figure 1
+dat1 <- mod$set_prm("include_z", TRUE)$gen(300, computed = select(dat0, x, e))
+dat2 <- heter_model(b = 64)$gen(300, computed = select(dat0, x, e))
 dat3 <- heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
-               gen(300, computed = select(dat, x))
-
-data.frame(plot = c("A", "B", "C", "D"), 
-           model = c("None", "Non-linearity", "Heteroskedasticity", "Non-normality"),
-           f = c(POLY_MODEL$test(dat)$p_value, 
-                 POLY_MODEL$test(dat1)$p_value,
-                 POLY_MODEL$test(mutate(dat2, z = poly_model()$gen(300, computed = select(dat2, x))$z))$p_value,
-                 POLY_MODEL$test(mutate(dat3, z = poly_model()$gen(300, computed = select(dat3, x))$z))$p_value),
-           b = c(HETER_MODEL$test(dat)$p_value, 
-                 HETER_MODEL$test(dat1)$p_value,
-                 HETER_MODEL$test(dat2)$p_value,
-                 HETER_MODEL$test(dat3)$p_value),
-           s = c(shapiro.test(dat$.resid)$p.value, 
-                 shapiro.test(dat1$.resid)$p.value,
-                 shapiro.test(dat2$.resid)$p.value,
-                 shapiro.test(dat3$.resid)$p.value)) %>%
-  mutate(f = format(round(f, digits = 3), nsmall = 3), b = format(round(b, digits = 3), nsmall = 3), s = format(round(s, digits = 3), nsmall = 3)) -> table_dat
-
-  table_dat$f <- kableExtra::cell_spec(table_dat$f, italic = as.numeric(table_dat$f) < 0.05)
-  table_dat$b <- kableExtra::cell_spec(table_dat$b, italic = as.numeric(table_dat$b) < 0.05)
-  table_dat$s <- kableExtra::cell_spec(table_dat$s, italic = as.numeric(table_dat$s) < 0.05)
+               gen(300, computed = select(dat0, x))
+
+dat_list <- list(dat0, dat1, dat2, dat3)
+
+table_dat <- data.frame(plot = c("A", "B", "C", "D"), 
+                        model = c("None", 
+                                  "Non-linearity", 
+                                  "Heteroskedasticity", 
+                                  "Non-normality"),
+                        r = map_dbl(dat_list, 
+                                    ~POLY_MODEL$test(.x, 
+                                                     test = "RESET", 
+                                                     power = 2:4)$p_value),
+                        b = map_dbl(dat_list, 
+                                    ~HETER_MODEL$test(.x)$p_value),
+
+                        # SW test has not been included by visage
+                        s = map_dbl(dat_list, 
+                                    ~shapiro.test(.x$.resid)$p.value)) %>%
+
+  # 3 decimal points
+  mutate(across(r:s, ~ format(round(.x, digits = 3), nsmall = 3))) %>%
+
+  # Italic if reject
+  mutate(across(r:s, ~ kableExtra::cell_spec(.x, italic = as.numeric(.x) < 0.05)))
 
   table_dat %>%
-  knitr::kable(col.names = c("Plot", "Departures", "RESET", "Breusch-Pagan", "Shapiro–Wilk"),
+  knitr::kable(col.names = c("Plot", 
+                             "Departures", 
+                             "RESET", 
+                             "Breusch-Pagan", 
+                             "Shapiro–Wilk"),
                align = "llrrr",
@@
                label = "example-residual-plot-table" ) 
-
 ```
@@
 
-(This section introduces the lineup protocol, and briefly discusses the method for sampling null data, calculating the p-value and estimating power.)
-
 ### Lineup protocol
@@
 
-Unlike hypothesis testing built upon rigorous statistical procedures, reading diagnostic plots relies on graphical perception - human's ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984], which is to some extent subjective and indecisive.
-Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
+One may prefer conventional hypothesis testing because reading diagnostic plots is to some extent subjective and indecisive as it relies on graphical perception - human's ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984].
+It is arguably true that visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
 One such example is finding an over-interpretation of the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurrence [@roy_chowdhury_using_2015].

### 2022-12-08 paper_comparison/paper_comparison/paper_comparison.tex
clean code and update text in 2.2

@@
 
-(This section discusses the tests that will be used in the analysis and
-shows the results for the residual plots displayed in the previous
-section.)
-
 Other than checking diagnostic plots, analysts may perform formal
@@
 \citep{ramsey_tests_1969}. And for testing normality, the Shapiro--Wilk
-test {[}ref here{]} is perhaps the most widely used test included by
-many of the statistical softwares. Another choice will be the
-Jarque--Bera test {[}ref here{]} which directly checks if the sample
-skewness and kurtosis match a normal distribution.
+test \citep{shapiro1965analysis} is perhaps the most widely used test
+included by many of the statistical softwares. Another choice will be
+the Jarque--Bera test \citep{jarque1980efficient} which directly checks
+if the sample skewness and kurtosis match a normal distribution.
 
@@
 \ref{fig:residual-plot-common-departures} are examined by the
-corresponding model structural test, Breusch-Pagan test and
-Shapiro--Wilk test as shown in Table
-\ref{tab:example-residual-plot-table}. In the example, both the
-Breusch-Pagan test and the Shapiro--Wilk test rejects the null
+corresponding RESET test, Breusch-Pagan test and Shapiro--Wilk test as
+shown in Table \ref{tab:example-residual-plot-table}. In the example,
+both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null
 hypothesis for departures that they do not intend to examine. As
@@
 rejection of the null hypothesis despite when majority of the residuals
-are well-behaved \citep{cook_applied_1999}. This can be largely avoided
-in diagnostic plots as experienced analysts can evaluate the
-acceptability of assumptions flexibly, even in the presence of outliers.
+are well-behaved \citep{cook_applied_1999}. Furthermore, with a
+sufficiently large sample size, residual-based tests may reject the null
+hypothesis due to a slight departure that is of little practical
+significance. These can be largely avoided in diagnostic plots as
+experienced analysts can evaluate the acceptability of assumptions
+flexibly, even in the presence of outliers and slight departures.
 
@@
 \midrule
-A & None & 0.434 & 0.133 & 0.728\\
+A & None & 0.779 & 0.133 & 0.728\\
 B & Non-linearity & \em{0.000} & \em{0.000} & \em{0.039}\\
-C & Heteroskedasticity & 0.378 & \em{0.000} & \em{0.000}\\
-D & Non-normality & 0.667 & 0.736 & \em{0.000}\\
+C & Heteroskedasticity & 0.658 & \em{0.000} & \em{0.000}\\
+D & Non-normality & 0.863 & 0.736 & \em{0.000}\\
 \bottomrule

### 2022-12-08 paper_comparison/paper_comparison/paper_comparison.Rmd
update 2.2.1 text

@@
 
-<!-- improve the first sentence -->
-
-One may prefer conventional hypothesis testing because reading diagnostic plots is to some extent subjective and indecisive as it relies on graphical perception - human's ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984].
-It is arguably true that visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
+One may argue that reading diagnostic plots is to some extent subjective and indecisive compared to those rigorous statistical procedures as it relies on graphical perception - human's ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984].
+It is particularly true that visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
 One such example is finding an over-interpretation of the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurrence [@roy_chowdhury_using_2015].
@@
 
-<!-- use true exp lineup as example lineup -->
-
 ```{r first-example-lineup, fig.cap="Visual testing is conducted using a lineup, as in the example here.
-The residual plot computed from the observed data (plot 11, exhibiting non-linearity) is embedded among 19 null plots, where the residuals were simulated from a standard error model.
+The residual plot computed from the observed data (plot 7, exhibiting non-linearity) is embedded among 19 null plots, where the residuals were simulated from a standard error model.
 Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot.
-A small $p$-value would result from a substantial number selecting plot 11."}
-mod <- poly_model(include_z = TRUE, sigma = 1)
-POLY_MODEL$plot_lineup(mod$gen_lineup(300, pos = 11), theme = theme_light(), remove_axis = TRUE, remove_grid_line = TRUE)
+A small $p$-value would result from a substantial number selecting plot 7."}
+
+vi_lineup$poly_422$data %>%
+  VI_MODEL$plot_lineup(theme = theme_light(), remove_axis = TRUE, remove_grid_line = TRUE)
 ```
@@
 Figure \ref{fig:first-example-lineup} is an example of a lineup protocol.
-If the actual data plot at position 11 is identifiable, then it is evidence for the rejection of the null hypothesis that the regression model is correctly specified.
+If the actual data plot at position 7 is identifiable, then it is evidence for the rejection of the null hypothesis that the regression model is correctly specified.
 In fact, the actual residual plot is obtained from a misspecified regression model with non-linearity issue.
 
-The effectiveness of lineup protocol has already been validated by @majumder_validation_2013 under relatively simple classical normal linear regression model settings with only one or two regressors.
+The effectiveness of lineup protocol is validated by @majumder_validation_2013 under relatively simple classical normal linear regression model settings with one or two regressors.
 Their results suggest visual test is capable of testing the significance of a single regressor with a similar power as a t-test, though they expressed that in general it is unnecessary to use visual inference if there exists a conventional test and they didn't expect the visual test to perform equally well as the conventional test.
@@
 This is encouraging as it promotes the use of visual inference in border field of data science where there are no existing statistical testing procedures.
-In fact, lineup protocol has been integrated successfully into diagnostic tools of hierarchical linear models [@loy2013diagnostic].
-
-<!-- adam work -->
+Visual inference have also been integrated into diagnostic of hierarchical linear models by @loy2013diagnostic, @loy2014hlmdiag and @loy2015you.
+They use lineup protocols to judge the assumption of linearity, normality and constant error variance for both the level-1 and level-2 residuals.
 

### 2022-12-08 paper_comparison/paper_comparison/paper_comparison.tex
update 2.2.1 text

@@
 
-(This section introduces the lineup protocol, and briefly discusses the
-method for sampling null data, calculating the p-value and estimating
-power.)
-
 \hypertarget{lineup-protocol}{%
@@
 
-Unlike hypothesis testing built upon rigorous statistical procedures,
-reading diagnostic plots relies on graphical perception - human's
-ability to interpret and decode the information embedded in the graph
-\citep{cleveland_graphical_1984}, which is to some extent subjective and
-indecisive. Further, visual discovery suffers from its unsecured and
-unconfirmed nature where the degree of the presence of the visual
+One may argue that reading diagnostic plots is to some extent subjective
+and indecisive compared to those rigorous statistical procedures as it
+relies on graphical perception - human's ability to interpret and decode
+the information embedded in the graph \citep{cleveland_graphical_1984}.
+It is particularly true that visual discovery suffers from its unsecured
+and unconfirmed nature where the degree of the presence of the visual
 features typically can not be measured quantitatively and objectively,
@@
 \caption{Visual testing is conducted using a lineup, as in the example here.
-The residual plot computed from the observed data (plot 11, exhibiting non-linearity) is embedded among 19 null plots, where the residuals were simulated from a standard error model.
+The residual plot computed from the observed data (plot 7, exhibiting non-linearity) is embedded among 19 null plots, where the residuals were simulated from a standard error model.
 Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot.
-A small $p$-value would result from a substantial number selecting plot 11.}\label{fig:first-example-lineup}
+A small $p$-value would result from a substantial number selecting plot 7.}\label{fig:first-example-lineup}
 \end{figure}
@@
 Figure \ref{fig:first-example-lineup} is an example of a lineup
-protocol. If the actual data plot at position 11 is identifiable, then
-it is evidence for the rejection of the null hypothesis that the
-regression model is correctly specified. In fact, the actual residual
-plot is obtained from a misspecified regression model with non-linearity
-issue.
+protocol. If the actual data plot at position 7 is identifiable, then it
+is evidence for the rejection of the null hypothesis that the regression
+model is correctly specified. In fact, the actual residual plot is
+obtained from a misspecified regression model with non-linearity issue.
 
-The effectiveness of lineup protocol has already been validated by
+The effectiveness of lineup protocol is validated by
 \citet{majumder_validation_2013} under relatively simple classical
-normal linear regression model settings with only one or two regressors.
+normal linear regression model settings with one or two regressors.
 Their results suggest visual test is capable of testing the significance
@@
 field of data science where there are no existing statistical testing
-procedures. In fact, lineup protocol has been integrated successfully
-into diagnostic tools of hierarchical linear models
-\citep{loy2013diagnostic}.
+procedures. Visual inference have also been integrated into diagnostic
+of hierarchical linear models by \citet{loy2013diagnostic},
+\citet{loy2014hlmdiag} and \citet{loy2015you}. They use lineup protocols
+to judge the assumption of linearity, normality and constant error
+variance for both the level-1 and level-2 residuals.
 

### 2022-12-08 paper_comparison/paper_comparison/paper_comparison.Rmd
update text 2.3.3

@@
 Example residual plots given in Figure \ref{fig:residual-plot-common-departures} are examined by the corresponding RESET test, Breusch-Pagan test and Shapiro--Wilk test as shown in Table \ref{tab:example-residual-plot-table}.
-In the example, both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null hypothesis for departures that they do not intend to examine.
+In the example, both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null hypothesis $H_0$ for departures that they do not intend to examine.
 As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are sensitive to other types of departures.
-It is likely the null hypothesis is correctly rejected but for the wrong reason, which is known as the "Type III error".
-Additionally, outliers will often incorrectly trigger the rejection of the null hypothesis despite when majority of the residuals are well-behaved [@cook_applied_1999].
-Furthermore, with a sufficiently large sample size, residual-based tests may reject the null hypothesis due to a slight departure that is of little practical significance.
+It is likely $H_0$ is correctly rejected but for the wrong reason, which is known as the "Type III error".
+Additionally, outliers will often incorrectly trigger the rejection of $H_0$ despite when majority of the residuals are well-behaved [@cook_applied_1999].
+Furthermore, with a sufficiently large sample size, residual-based tests may reject $H_0$ due to a slight departure that is of little practical significance.
 These can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers and slight departures.
@@
 Later, @buja_statistical_2009 proposed the lineup protocol as a visual test inspired by the "police lineup" or "identity parade" which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people.
-The protocol consists of $m$ randomly placed plots, where one plot is the actual data plot, and the remaining $m - 1$ plots have the identical graphical production as the data plot except the data has been replaced with data consistent with the null hypothesis.
+The protocol consists of $m$ randomly placed plots, where one plot is the actual data plot, and the remaining $m - 1$ plots have the identical graphical production as the data plot except the data has been replaced with data consistent with $H_0$.
 Then, an observer who have not seen the actual data plot will be asked to point out the most different plot from the lineup.
-Under the null hypothesis, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is $1/m$.
-If we reject the null hypothesis as the observer correctly picks the actual data plot, then the Type I error of this test is $1/m$.
+Under $H_0$, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is $1/m$.
+If we reject $H_0$ as the observer correctly picks the actual data plot, then the Type I error of this test is $1/m$.
 
@@
 Figure \ref{fig:first-example-lineup} is an example of a lineup protocol.
-If the actual data plot at position 7 is identifiable, then it is evidence for the rejection of the null hypothesis that the regression model is correctly specified.
+If the actual data plot at position 7 is identifiable, then it is evidence for the rejection of $H_0$ that the regression model is correctly specified.
 In fact, the actual residual plot is obtained from a misspecified regression model with non-linearity issue.
@@
 In the context of regression diagnostics, sampling data from $H_0$ is equivalent to sampling data from the assumed model.
-As @buja_statistical_2009 suggested, $H_0$ is usually composited by a collection of distributions controlled by nuisance parameters.
+As @buja_statistical_2009 suggest, $H_0$ is usually composited by a collection of distributions controlled by nuisance parameters.
 Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling.
@@
 Further, a visual test can involve $K$ independent observers.
-Let $D_i = \{0,1\}$ be a binomial random variable denoting whether subject $i$ correctly detecting the actual data plot, and $X = \sum_{i=1}^{K}X_i$ be the number of observers correctly picking the actual data plot.
-Then, by imposing a relatively strong assumption on the visual test that all $K$ evaluations are fully independent, under the null hypothesis, $X \sim \mathrm{Binom}_{K,1/m}$.
+Let $X_i = \{0,1\}$ be a binomial random variable denoting whether subject $i$ correctly detecting the actual data plot, and $X = \sum_{i=1}^{K}X_i$ be the number of observers correctly picking the actual data plot.
+Then, by imposing a relatively strong assumption on the visual test that all $K$ evaluations are fully independent, under $H_0$, $X \sim \mathrm{Binom}_{K,1/m}$.
 Therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as $P(X \geq x) = 1 - F(x) + f(x)$, where $F(.)$ is the cumulative distribution function, $f(.)$ is the probability mass function and $x$ is the realization of number of observers correctly picking the actual data plot [@majumder_validation_2013].
 
-As pointed out by @vanderplas2021statistical, the binomial model doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup.
+As point out by @vanderplas2021statistical, this basic binomial model doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup.
 And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup.
-They summarized three common scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same actual data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects.
-Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experimental design.
-For Scenario 3, @vanderplas2021statistical modelled the probability of a plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$.
+@vanderplas2021statistical summarizes three common scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same actual data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects.
+Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experiment design.
+For Scenario 3, @vanderplas2021statistical model the probability of a plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$.
 And defined $c_i$ to be the number of times plot $i$ being selected in $K$ evaluations.
@@
 \begin{equation} \label{eq:pvalue-beta-binomial}
-P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},
+P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+
 \end{equation}
@@
 
-Note that Equation \ref{eq:pvalue-beta-binomial} only works with integer $c_i$. For non-integer $c_i$, linear approximation will be applied to calculate the p-value
+Note that Equation \ref{eq:pvalue-beta-binomial} only works with non-negative integer $c_i$. For non-negative real number $c_i$, linear approximation will be applied to calculate the p-value
 
@@
 \begin{equation} \label{eq:pvalue-beta-binomial-approx}
-P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \notin \mathbb{Z}.
+P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \in \mathbb{R}_0^+,
 \end{equation}
@@
 
-<!-- TODO:Introduce the concept of c interesting plot -->
+where $P(C \geq \lceil c_i \rceil)$ is calculated using Equation \ref{eq:pvalue-beta-binomial} and $P(C = \lfloor c_i \rfloor)$ is calculated by
+
+\begin{equation}
+P(C = c_i) = {K \choose c_i}\frac{B(c_i + \alpha, K - c_i + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+.
+\end{equation}
 
@@
 For low values of $\alpha$, only a few plots are attractive to the observers and tend to be selected.
-For higher values of $\alpha$, the distribution of the probability of each plot being selected is more evenly.
-@vanderplas2021statistical defined that a plot is $c$-interesting if $c$ or more participants selected the plot as the most different.
+For higher values of $\alpha$, the distribution of the probability of each plot being selected is more even.
+@vanderplas2021statistical define that a plot is $c$-interesting if $c$ or more participants select the plot as the most different.
 Given the definition, The expected number of plots selected at least $c$ times, $E[Z_c]$, is calculated as
@@
 
-@vanderplas2021statistical suggested that $\alpha$ can be estimated using MLE and Equation \ref{eq:c-interesting-expectation}.
+@vanderplas2021statistical suggest that $\alpha$ can be estimated using MLE with Equation \ref{eq:c-interesting-expectation}.
 But for precise estimate of $\alpha$, additional responses to Rorschach lineups, which is a type of lineup consists of only null plots, are required.
-
-<!-- But if $\alpha$ is small and only a few null plots in a lineup are attractive, MLE could fail to provide accurate estimates. -->
 
@@
 fitted values plot for a classical normal linear regression model.
-The residuals are produced by fitting a two-predictor multiple linear regression model with data generated from a cubic linear model.
+The residuals are produced by fitting a two-regressor multiple linear regression model with data generated from a cubic linear model.
 From the residual plot, "butterfly shape" can be observed which generally would be interpretd as evidence of heteroskedasticity.
 Further, from the outline of the shape, nonlinear patterns exist.
-Both visual discoveries are evidence against the null hypothesis, though heteroskedasticity actually does not exist in the data generating process.
+Both visual discoveries are evidence against $H_0$, though heteroskedasticity actually does not exist in the data generating process.
 \\label{fig:residual-plot-cubic-heter}', eval = FALSE}
@@
 
-In particular, given 95% confidence level, both tests will reject the null hypothesis in only around 58% of the cases.
-And in 41% of the cases, the visual test does not reject the null hypothesis but the conventional test does.
-Since fail to reject the null hypothesis in a visual test means that there is no obvious visual discoveries found in the residual plot, analysts and the general public as the consumers of the output may not fully believe in the rejection of the null hypothesis even though the alternative hypothesis is true.
+In particular, given 95% confidence level, both tests will reject $H_0$ in only around 58% of the cases.
+And in 41% of the cases, the visual test does not reject $H_0$ but the conventional test does.
+Since fail to reject $H_0$ in a visual test means that there is no obvious visual discoveries found in the residual plot, analysts and the general public as the consumers of the output may not fully believe in the rejection of $H_0$ even though the alternative hypothesis is true.
 Even if the rejection is accepted, the model violation may not be considered as impactful due to the fact that the departures are not clearly visible.
@@
 In fact, the sensitivity of the conventional test could sometimes distract and discourage analysts from finding a simple but good linear approximation to the data.
-The rejection of the null hypothesis because of acceptable and negligible residual departure is not practically meaningful and useful.
+The rejection of $H_0$ because of acceptable and negligible residual departure is not practically meaningful and useful.
 This also partly explains why abundance of literature suggesting the use of data plot in regression diagnostics.
@@
 In terms of the heteroskedasticity model, the power of Breusch-Pagan test is also almost always greater than the power of visual test.
-For $0 \leq log_e(E) \leq 2$, where the visual feature is nearly unobservable from the example residual plots and the power curve of the visual test remains at a low level, the Breusch-Pagan test still have a decent amount of chance of rejecting the null hypothesis.
+For $0 \leq log_e(E) \leq 2$, where the visual feature is nearly unobservable from the example residual plots and the power curve of the visual test remains at a low level, the Breusch-Pagan test still have a decent amount of chance of rejecting $H_0$.
 
@@
 The power curve of F-test remains at around 5% in Figure \ref{fig:heterpower} since there is no non-linear term leave out in the heteroskedasticity model and
-the null hypothesis of the F-test will be always satisfied.
+$H_0$ of the F-test will be always satisfied.
 

### 2022-12-08 paper_comparison/paper_comparison/paper_comparison.tex
update text 2.3.3

@@
 both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null
-hypothesis for departures that they do not intend to examine. As
+hypothesis \(H_0\) for departures that they do not intend to examine. As
 discussed in \citet{cook1982residuals}, most residual-based tests for a
 particular type of departure from model assumptions are sensitive to
-other types of departures. It is likely the null hypothesis is correctly
-rejected but for the wrong reason, which is known as the ``Type III
-error''. Additionally, outliers will often incorrectly trigger the
-rejection of the null hypothesis despite when majority of the residuals
-are well-behaved \citep{cook_applied_1999}. Furthermore, with a
-sufficiently large sample size, residual-based tests may reject the null
-hypothesis due to a slight departure that is of little practical
-significance. These can be largely avoided in diagnostic plots as
-experienced analysts can evaluate the acceptability of assumptions
-flexibly, even in the presence of outliers and slight departures.
+other types of departures. It is likely \(H_0\) is correctly rejected
+but for the wrong reason, which is known as the ``Type III error''.
+Additionally, outliers will often incorrectly trigger the rejection of
+\(H_0\) despite when majority of the residuals are well-behaved
+\citep{cook_applied_1999}. Furthermore, with a sufficiently large sample
+size, residual-based tests may reject \(H_0\) due to a slight departure
+that is of little practical significance. These can be largely avoided
+in diagnostic plots as experienced analysts can evaluate the
+acceptability of assumptions flexibly, even in the presence of outliers
+and slight departures.
 
@@
 identical graphical production as the data plot except the data has been
-replaced with data consistent with the null hypothesis. Then, an
-observer who have not seen the actual data plot will be asked to point
-out the most different plot from the lineup. Under the null hypothesis,
-it is expected that the actual data plot would have no distinguishable
-difference with the null plots, and the probability of the observer
-correctly picks the actual data plot is \(1/m\). If we reject the null
-hypothesis as the observer correctly picks the actual data plot, then
-the Type I error of this test is \(1/m\).
+replaced with data consistent with \(H_0\). Then, an observer who have
+not seen the actual data plot will be asked to point out the most
+different plot from the lineup. Under \(H_0\), it is expected that the
+actual data plot would have no distinguishable difference with the null
+plots, and the probability of the observer correctly picks the actual
+data plot is \(1/m\). If we reject \(H_0\) as the observer correctly
+picks the actual data plot, then the Type I error of this test is
+\(1/m\).
 
@@
 protocol. If the actual data plot at position 7 is identifiable, then it
-is evidence for the rejection of the null hypothesis that the regression
-model is correctly specified. In fact, the actual residual plot is
-obtained from a misspecified regression model with non-linearity issue.
+is evidence for the rejection of \(H_0\) that the regression model is
+correctly specified. In fact, the actual residual plot is obtained from
+a misspecified regression model with non-linearity issue.
 
@@
 equivalent to sampling data from the assumed model. As
-\citet{buja_statistical_2009} suggested, \(H_0\) is usually composited
-by a collection of distributions controlled by nuisance parameters.
-Since regression models can have various forms, there is no general
-solution to this problem, but it sometimes can be reduced to so called
+\citet{buja_statistical_2009} suggest, \(H_0\) is usually composited by
+a collection of distributions controlled by nuisance parameters. Since
+regression models can have various forms, there is no general solution
+to this problem, but it sometimes can be reduced to so called
 ``reference distribution'' by applying one of the three methods: (i)
@@
 Further, a visual test can involve \(K\) independent observers. Let
-\(D_i = \{0,1\}\) be a binomial random variable denoting whether subject
+\(X_i = \{0,1\}\) be a binomial random variable denoting whether subject
 \(i\) correctly detecting the actual data plot, and
@@
 on the visual test that all \(K\) evaluations are fully independent,
-under the null hypothesis, \(X \sim \mathrm{Binom}_{K,1/m}\). Therefore,
-the \(p\)-value of a lineup of size \(m\) evaluated by \(K\) observer is
+under \(H_0\), \(X \sim \mathrm{Binom}_{K,1/m}\). Therefore, the
+\(p\)-value of a lineup of size \(m\) evaluated by \(K\) observer is
 given as \(P(X \geq x) = 1 - F(x) + f(x)\), where \(F(.)\) is the
@@
 
-As pointed out by \citet{vanderplas2021statistical}, the binomial model
-doesn't take into account the possible dependencies in the visual test
-due to repeated evaluations of the same lineup. And it is inapplicable
-to visual test where subjects are asked to select one or more ``most
-different'' plots from the lineup. They summarized three common
-scenarios in visual inference: (1) \(K\) different lineups are shown to
-\(K\) subjects, (2) \(K\) lineups with different null plots but the same
+As point out by \citet{vanderplas2021statistical}, this basic binomial
+model doesn't take into account the possible dependencies in the visual
+test due to repeated evaluations of the same lineup. And it is
+inapplicable to visual test where subjects are asked to select one or
+more ``most different'' plots from the lineup.
+\citet{vanderplas2021statistical} summarizes three common scenarios in
+visual inference: (1) \(K\) different lineups are shown to \(K\)
+subjects, (2) \(K\) lineups with different null plots but the same
 actual data plot are shown to \(K\) subjects, and (3) the same lineup is
@@
 most common in previous studies as it puts the least constraints on the
-experimental design. For Scenario 3, \citet{vanderplas2021statistical}
-modelled the probability of a plot \(i\) being selected from a lineup as
+experiment design. For Scenario 3, \citet{vanderplas2021statistical}
+model the probability of a plot \(i\) being selected from a lineup as
 \(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
@@
 
-Note that Equation \ref{eq:pvalue-beta-binomial} only works with integer
-\(c_i\). For non-integer \(c_i\), linear approximation will be applied
-to calculate the p-value
+Note that Equation \ref{eq:pvalue-beta-binomial} only works with
+non-negative integer \(c_i\). For non-negative real number \(c_i\),
+linear approximation will be applied to calculate the p-value
 
 \begin{equation} \label{eq:pvalue-beta-binomial-approx}
-P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \notin \mathbb{Z}.
+P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \in \mathbb{R}_0^+,
+\end{equation}
+
+where \(P(C \geq \lceil c_i \rceil)\) is calculated using Equation
+\ref{eq:pvalue-beta-binomial} and \(P(C = \lfloor c_i \rfloor)\) is
+calculated by
+
+\begin{equation}
+P(C = c_i) = {K \choose c_i}\frac{B(c_i + \alpha, K - c_i + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+.
 \end{equation}
@@
 higher values of \(\alpha\), the distribution of the probability of each
-plot being selected is more evenly. \citet{vanderplas2021statistical}
-defined that a plot is \(c\)-interesting if \(c\) or more participants
-selected the plot as the most different. Given the definition, The
+plot being selected is more even. \citet{vanderplas2021statistical}
+define that a plot is \(c\)-interesting if \(c\) or more participants
+select the plot as the most different. Given the definition, The
 expected number of plots selected at least \(c\) times, \(E[Z_c]\), is
@@
 
-\citet{vanderplas2021statistical} suggested that \(\alpha\) can be
-estimated using MLE and Equation \ref{eq:c-interesting-expectation}. But
-for precise estimate of \(\alpha\), additional responses to Rorschach
-lineups, which is a type of lineup consists of only null plots, are
-required.
+\citet{vanderplas2021statistical} suggest that \(\alpha\) can be
+estimated using MLE with Equation \ref{eq:c-interesting-expectation}.
+But for precise estimate of \(\alpha\), additional responses to
+Rorschach lineups, which is a type of lineup consists of only null
+plots, are required.
 
@@
 
-In particular, given 95\% confidence level, both tests will reject the
-null hypothesis in only around 58\% of the cases. And in 41\% of the
-cases, the visual test does not reject the null hypothesis but the
-conventional test does. Since fail to reject the null hypothesis in a
-visual test means that there is no obvious visual discoveries found in
-the residual plot, analysts and the general public as the consumers of
-the output may not fully believe in the rejection of the null hypothesis
-even though the alternative hypothesis is true. Even if the rejection is
-accepted, the model violation may not be considered as impactful due to
-the fact that the departures are not clearly visible.
+In particular, given 95\% confidence level, both tests will reject
+\(H_0\) in only around 58\% of the cases. And in 41\% of the cases, the
+visual test does not reject \(H_0\) but the conventional test does.
+Since fail to reject \(H_0\) in a visual test means that there is no
+obvious visual discoveries found in the residual plot, analysts and the
+general public as the consumers of the output may not fully believe in
+the rejection of \(H_0\) even though the alternative hypothesis is true.
+Even if the rejection is accepted, the model violation may not be
+considered as impactful due to the fact that the departures are not
+clearly visible.
 
@@
 sometimes distract and discourage analysts from finding a simple but
-good linear approximation to the data. The rejection of the null
-hypothesis because of acceptable and negligible residual departure is
-not practically meaningful and useful. This also partly explains why
-abundance of literature suggesting the use of data plot in regression
-diagnostics.
+good linear approximation to the data. The rejection of \(H_0\) because
+of acceptable and negligible residual departure is not practically
+meaningful and useful. This also partly explains why abundance of
+literature suggesting the use of data plot in regression diagnostics.
 
@@
 visual test remains at a low level, the Breusch-Pagan test still have a
-decent amount of chance of rejecting the null hypothesis.
+decent amount of chance of rejecting \(H_0\).
 
@@
 remains at around 5\% in Figure \ref{fig:heterpower} since there is no
-non-linear term leave out in the heteroskedasticity model and the null
-hypothesis of the F-test will be always satisfied.
+non-linear term leave out in the heteroskedasticity model and \(H_0\) of
+the F-test will be always satisfied.
 

### 2022-12-08 paper_comparison/paper_comparison/paper_comparison.Rmd
try mix-effect model (failed?)

@@
 
-Example residual plots given in Figure \ref{fig:residual-plot-common-departures} are examined by the corresponding RESET test, Breusch-Pagan test and Shapiro--Wilk test as shown in Table \ref{tab:example-residual-plot-table}.
+Example residual plots given in Figure \ref{fig:residual-plot-common-departures} are examined by the corresponding RESET test, Breusch-Pagan test and Shapiro-Wilk test as shown in Table \ref{tab:example-residual-plot-table}.
 In the example, both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null hypothesis $H_0$ for departures that they do not intend to examine.
@@
 In the context of regression diagnostics, sampling data from $H_0$ is equivalent to sampling data from the assumed model.
-As @buja_statistical_2009 suggest, $H_0$ is usually composited by a collection of distributions controlled by nuisance parameters.
+As @buja_statistical_2009 suggested, $H_0$ is usually composited by a collection of distributions controlled by nuisance parameters.
 Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling.
@@
 
-As point out by @vanderplas2021statistical, this basic binomial model doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup.
+As pointed out by @vanderplas2021statistical, this basic binomial model doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup.
 And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup.
-@vanderplas2021statistical summarizes three common scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same actual data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects.
+@vanderplas2021statistical summarises three common scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same actual data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects.
 Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experiment design.
@@
 Thus, the power of a visual test depends on the subject-specific abilities.
-Previously, it was addressed by modelling the probability of a subject $i$ correctly picking the actual data plot from a lineup $l$ using a mixed-effect logistic regression with the subject being treated as a random effect [@majumder_validation_2013].
+Previously, it is addressed by modelling the probability of a subject $i$ correctly picking the actual data plot from a lineup $l$ using a mixed-effect logistic regression with the subject being treated as a random effect [@majumder_validation_2013].
 However, having this probability is insufficient to determine the power of a visual test allowed for multiple selections as it doesn't provide information about the number of selections made by the subject for p-value calculation.
 
-Instead, we directly estimated the probability of a lineup being rejected using a logistic regression with the natural logarithm of the effect size as the only regressor formulated as
+Instead, we omitted the individual effect and directly estimated the probability of a lineup being rejected using a logistic regression with the natural logarithm of the effect size as the only regressor formulated as
 
@@
 \noindent where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
+
+```{r eval = FALSE}
+MASS::glmmPQL(weighted_detect ~ log_effect_size, random = ~ log_effect_size | subject,  family = quasibinomial(), data = vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(x_dist == "uniform") %>%
+  filter(type == "polynomial") %>%
+  mutate(log_effect_size = log10(effect_size)) %>%
+  mutate(subject = factor(as.integer(factor(paste(exp, set))))) %>%
+  select(log_effect_size, weighted_detect, subject)) -> mod
+
+lme4::glmer(weighted_detect ~ log_effect_size  + (log_effect_size | subject), 
+            family = binomial(),
+            data = vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(x_dist == "uniform") %>%
+  filter(type == "polynomial") %>%
+    mutate(log_effect_size = log10(effect_size)) %>%
+  mutate(subject = factor(as.integer(factor(paste(exp, set))))) %>%
+  select(log_effect_size, weighted_detect, subject)) -> mod
+
+vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(x_dist == "uniform") %>%
+  filter(type == "polynomial") %>%
+  mutate(log_effect_size = log10(effect_size)) %>%
+  mutate(subject = factor(as.integer(factor(paste(exp, set))))) %>%
+  mutate(pred_wd = predict(mod, type = "response")) %>%
+  group_by(unique_lineup_id) %>%
+  mutate(pred_ci = pred_wd * n()) %>%
+  mutate(n_eval = n()) %>%
+  ungroup() %>%
+  rowwise() %>%
+  mutate(pred_pval = (function(c_i, n_eval, alpha) {
+  target_dist <- exact_dist(n_eval, 20, dist = "dirichlet", alpha = alpha)
+  floor_c_i <- floor(c_i)
+  ceil_c_i <- ceiling(c_i)
+  return(unname(sum(target_dist[(ceil_c_i:n_eval) + 1]) + (ceil_c_i - c_i) * target_dist[floor_c_i + 1]))
+})(pred_ci, n_eval, alpha)) %>%
+  ungroup() %>%
+  mutate(pred_reject = pred_pval < 0.05) %>%
+  ggplot() +
+  geom_smooth(aes(log_effect_size, as.numeric(pred_reject)), method = "glm", method.args = list(family = binomial()), se = FALSE)
+
+```
 

### 2022-12-08 paper_comparison/paper_comparison/paper_comparison.tex
try mix-effect model (failed?)

@@
 \ref{fig:residual-plot-common-departures} are examined by the
-corresponding RESET test, Breusch-Pagan test and Shapiro--Wilk test as
+corresponding RESET test, Breusch-Pagan test and Shapiro-Wilk test as
 shown in Table \ref{tab:example-residual-plot-table}. In the example,
@@
 equivalent to sampling data from the assumed model. As
-\citet{buja_statistical_2009} suggest, \(H_0\) is usually composited by
-a collection of distributions controlled by nuisance parameters. Since
-regression models can have various forms, there is no general solution
-to this problem, but it sometimes can be reduced to so called
+\citet{buja_statistical_2009} suggested, \(H_0\) is usually composited
+by a collection of distributions controlled by nuisance parameters.
+Since regression models can have various forms, there is no general
+solution to this problem, but it sometimes can be reduced to so called
 ``reference distribution'' by applying one of the three methods: (i)
@@
 
-As point out by \citet{vanderplas2021statistical}, this basic binomial
+As pointed out by \citet{vanderplas2021statistical}, this basic binomial
 model doesn't take into account the possible dependencies in the visual
@@
 more ``most different'' plots from the lineup.
-\citet{vanderplas2021statistical} summarizes three common scenarios in
+\citet{vanderplas2021statistical} summarises three common scenarios in
 visual inference: (1) \(K\) different lineups are shown to \(K\)
@@
 from the lineup. Thus, the power of a visual test depends on the
-subject-specific abilities. Previously, it was addressed by modelling
-the probability of a subject \(i\) correctly picking the actual data
-plot from a lineup \(l\) using a mixed-effect logistic regression with
-the subject being treated as a random effect
+subject-specific abilities. Previously, it is addressed by modelling the
+probability of a subject \(i\) correctly picking the actual data plot
+from a lineup \(l\) using a mixed-effect logistic regression with the
+subject being treated as a random effect
 \citep{majumder_validation_2013}. However, having this probability is
@@
 
-Instead, we directly estimated the probability of a lineup being
-rejected using a logistic regression with the natural logarithm of the
-effect size as the only regressor formulated as
+Instead, we omitted the individual effect and directly estimated the
+probability of a lineup being rejected using a logistic regression with
+the natural logarithm of the effect size as the only regressor
+formulated as
 

### 2022-12-09 paper_comparison/paper_comparison/paper_comparison.Rmd
need to create data folder if it doesn't exist for this to work

@@
   vi_lineup <- get_vi_lineup()
+  if (!dir.exists(here::here("data/"))) dir.create(here::here("data/"))
   saveRDS(vi_lineup, here::here("data/vi_lineup.rds"))

### 2022-12-13 paper_comparison/paper_comparison/paper_comparison.Rmd
update the language in the introduction

@@
 
-Plotting the residuals against predicted values and each of the explanatory variables on a scatter plot is a recommend way to scrutinize their relationships.
+Residuals might also be plotted, as a scatter plot against the predicted values and each of the explanatory variables to scrutinize their relationships.
 If there is any visually discoverable patterns, the model is potentially misspecified.
 However, correctly judging a residual plot where no pattern exists is a painstakingly difficult task for humans.
-
 It is especially common, particularly among new data analytics students to report patterns when an experienced data analyst might quickly conclude that there are none.
-[ET: I don't know if the former statement is true?
->> PL: change to students] Generally, one looks for noticeable departures from the model like non-linear dependency or heteroskedasticity.
+Generally, one looks for noticeable departures from the model like non-linear dependency or heteroskedasticity.
 It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and use a Breusch-Pagan test [@breusch_simple_1979] for heteroskedasticity.
@@
 The paper is structured as follows.
-The next background section describes the types of departures that one expects to detect, and describes a formal statistical process for reading residual plots, called visual inference [ET: CITATION HERE?
->> PL: arguments removed].
+The next background section describes the types of departures that one expects to detect, and describes a formal statistical process for reading residual plots, called visual inference.
 Section \ref{experimental-design} describes the experiment design to compare the decision made by formal hypothesis testing, and how humans would read diagnostic plots.
@@
 
+# a bit stronger
+# plot 7 -> plot 2^3 - 1
+
 vi_lineup$poly_422$data %>%
@@
 ```{r eval = FALSE}
-MASS::glmmPQL(weighted_detect ~ log_effect_size, random = ~ log_effect_size | subject,  family = quasibinomial(), data = vi_survey %>%
+MASS::glmmPQL(weighted_detect ~ scale(log_effect_size), random = ~ scale(log_effect_size) | subject,  family = quasibinomial(), data = vi_survey %>%
   filter(!attention_check) %>%
@@
   filter(type == "polynomial") %>%
-  mutate(log_effect_size = log10(effect_size)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
   mutate(subject = factor(as.integer(factor(paste(exp, set))))) %>%
@@
 
-lme4::glmer(weighted_detect ~ log_effect_size  + (log_effect_size | subject), 
-            family = binomial(),
-            data = vi_survey %>%
+mod_dat <- vi_survey %>%
   filter(!attention_check) %>%
@@
   filter(type == "polynomial") %>%
-    mutate(log_effect_size = log10(effect_size)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  # mutate(log_effect_size = scale(log_effect_size)) %>%
   mutate(subject = factor(as.integer(factor(paste(exp, set))))) %>%
-  select(log_effect_size, weighted_detect, subject)) -> mod
+  select(log_effect_size, weighted_detect, subject)
+
+mod <- lme4::glmer(weighted_detect ~ scale(log_effect_size)  + (scale(log_effect_size) | subject), 
+                   family = quasibinomial(),
+                   data = mod_dat)
+
+mod_dat$log_effect_size
+
+map(seq(-1, 6, 0.1), ~predict(mod, newdata = data.frame(subject = 1:283, log_effect_size = .x), type = "response")) %>%
+  `names<-`()
 
@@
 
-## Data overview
+## Overview
 
@@
 
-## P-value comparison
+## P-value comparison (remove?) (closer to the topic of the paper) (put it after power)
 
@@
 
+tick mark on x. dashed line too strong. use solid line (grey) in light color.
+
 <!-- ## Visual test  -->
@@
 ```
+
+# invesgating interesting lineups
+
+# change of x_dist -> power
+
+# appendix (contains all the details)

### 2022-12-13 paper_comparison/paper_comparison/paper_comparison.tex
update the language in the introduction

@@
 
-Plotting the residuals against predicted values and each of the
-explanatory variables on a scatter plot is a recommend way to scrutinize
-their relationships. If there is any visually discoverable patterns, the
-model is potentially misspecified. However, correctly judging a residual
-plot where no pattern exists is a painstakingly difficult task for
-humans.
-
-It is especially common, particularly among new data analytics students
-to report patterns when an experienced data analyst might quickly
-conclude that there are none. {[}ET: I don't know if the former
-statement is true? \textgreater\textgreater{} PL: change to students{]}
-Generally, one looks for noticeable departures from the model like
-non-linear dependency or heteroskedasticity. It is also possible to
-conduct hypothesis tests for non-linear dependence
+Residuals might also be plotted, as a scatter plot against the predicted
+values and each of the explanatory variables to scrutinize their
+relationships. If there is any visually discoverable patterns, the model
+is potentially misspecified. However, correctly judging a residual plot
+where no pattern exists is a painstakingly difficult task for humans. It
+is especially common, particularly among new data analytics students to
+report patterns when an experienced data analyst might quickly conclude
+that there are none. Generally, one looks for noticeable departures from
+the model like non-linear dependency or heteroskedasticity. It is also
+possible to conduct hypothesis tests for non-linear dependence
 \citep{ramsey_tests_1969}, and use a Breusch-Pagan test
@@
 describes a formal statistical process for reading residual plots,
-called visual inference {[}ET: CITATION HERE? \textgreater\textgreater{}
-PL: arguments removed{]}. Section \ref{experimental-design} describes
-the experiment design to compare the decision made by formal hypothesis
+called visual inference. Section \ref{experimental-design} describes the
+experiment design to compare the decision made by formal hypothesis
 testing, and how humans would read diagnostic plots. The results are
@@
 
+\textbf{test}
+
 Since \(z = O(x^j)\), for \(j > 1\), \(z\) is a higher order term leaves
@@
 
-\hypertarget{data-overview}{%
-\subsection{Data overview}\label{data-overview}}
+\hypertarget{overview}{%
+\subsection{Overview}\label{overview}}
 
@@
 
-\hypertarget{p-value-comparison}{%
-\subsection{P-value comparison}\label{p-value-comparison}}
+\hypertarget{p-value-comparison-remove-closer-to-the-topic-of-the-paper-put-it-after-power}{%
+\subsection{P-value comparison (remove?) (closer to the topic of the
+paper) (put it after
+power)}\label{p-value-comparison-remove-closer-to-the-topic-of-the-paper-put-it-after-power}}
 
@@
 
+tick mark on x. dashed line too strong. use solid line (grey) in light
+color.
+
 \hypertarget{lineup-selection}{%
@@
 
+\hypertarget{invesgating-interesting-lineups}{%
+\section{invesgating interesting
+lineups}\label{invesgating-interesting-lineups}}
+
+\hypertarget{change-of-x_dist---power}{%
+\section{change of x\_dist -\textgreater{}
+power}\label{change-of-x_dist---power}}
+
 \bibliographystyle{tfcad}

### 2022-12-13 paper_comparison/paper_comparison/paper_comparison.Rmd
clarify non-normality is not the focus of this paper

@@
 
-Residuals might also be plotted, as a scatter plot against the predicted values and each of the explanatory variables to scrutinize their relationships.
+More typically, residuals will be plotted, as a scatter plot against the predicted values and each of the explanatory variables to scrutinize their relationships.
 If there is any visually discoverable patterns, the model is potentially misspecified.
@@
 A favourable graphical summary for this task is the quantile-quantile plot.
-However, for a consistent comparison, residual plot will be the focus of this paper.
+For a consistent comparison, the residual plot of this departure is still presented in plot D of Figure \ref{fig:residual-plot-common-departures}.
+But since the scatter plot is not optimal for this task, non-normality will not be the focus of this paper.
 Besides, it is important to note that not all regression models assume normality for the error term, but a certain amount do including the classical normal linear regression model.

### 2022-12-13 paper_comparison/paper_comparison/paper_comparison.tex
clarify non-normality is not the focus of this paper

@@
 
-Residuals might also be plotted, as a scatter plot against the predicted
-values and each of the explanatory variables to scrutinize their
-relationships. If there is any visually discoverable patterns, the model
-is potentially misspecified. However, correctly judging a residual plot
-where no pattern exists is a painstakingly difficult task for humans. It
-is especially common, particularly among new data analytics students to
-report patterns when an experienced data analyst might quickly conclude
-that there are none. Generally, one looks for noticeable departures from
-the model like non-linear dependency or heteroskedasticity. It is also
-possible to conduct hypothesis tests for non-linear dependence
-\citep{ramsey_tests_1969}, and use a Breusch-Pagan test
-\citep{breusch_simple_1979} for heteroskedasticity.
+More typically, residuals will be plotted, as a scatter plot against the
+predicted values and each of the explanatory variables to scrutinize
+their relationships. If there is any visually discoverable patterns, the
+model is potentially misspecified. However, correctly judging a residual
+plot where no pattern exists is a painstakingly difficult task for
+humans. It is especially common, particularly among new data analytics
+students to report patterns when an experienced data analyst might
+quickly conclude that there are none. Generally, one looks for
+noticeable departures from the model like non-linear dependency or
+heteroskedasticity. It is also possible to conduct hypothesis tests for
+non-linear dependence \citep{ramsey_tests_1969}, and use a Breusch-Pagan
+test \citep{breusch_simple_1979} for heteroskedasticity.
 
@@
 readily reveal the marginal distribution. A favourable graphical summary
-for this task is the quantile-quantile plot. However, for a consistent
-comparison, residual plot will be the focus of this paper. Besides, it
-is important to note that not all regression models assume normality for
-the error term, but a certain amount do including the classical normal
-linear regression model. In the case that the normality assumption is
-violated, it is expected to observe data points do not center around the
-horizontal axis and there is an uneven distribution of the number points
-at both below and above the horizontal axis. For example, given a skewed
-error distribution, fewer data points and more outliers are on one side
-of the horizontal axis as shown in plot D of Figure
+for this task is the quantile-quantile plot. For a consistent
+comparison, the residual plot of this departure is still presented in
+plot D of Figure \ref{fig:residual-plot-common-departures}. But since
+the scatter plot is not optimal for this task, non-normality will not be
+the focus of this paper. Besides, it is important to note that not all
+regression models assume normality for the error term, but a certain
+amount do including the classical normal linear regression model. In the
+case that the normality assumption is violated, it is expected to
+observe data points do not center around the horizontal axis and there
+is an uneven distribution of the number points at both below and above
+the horizontal axis. For example, given a skewed error distribution,
+fewer data points and more outliers are on one side of the horizontal
+axis as shown in plot D of Figure
 \ref{fig:residual-plot-common-departures}.
@@
 \(\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\).
-
-\textbf{test}
 

### 2022-12-13 paper_comparison/paper_comparison/paper_comparison.Rmd
use a different residual plot in VI example

@@
 ```{r first-example-lineup, fig.cap="Visual testing is conducted using a lineup, as in the example here.
-The residual plot computed from the observed data (plot 7, exhibiting non-linearity) is embedded among 19 null plots, where the residuals were simulated from a standard error model.
+The residual plot computed from the observed data (plot $2^2 + 2$, exhibiting non-linearity) is embedded among 19 null plots, where the residuals were simulated from a standard error model.
 Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot.
-A small $p$-value would result from a substantial number selecting plot 7."}
+A small $p$-value would result from a substantial number selecting plot $2^2 + 2$."}
 
@@
 
-vi_lineup$poly_422$data %>%
+# vi_survey %>%
+#   filter(!attention_check) %>%
+#   filter(!null_lineup) %>%
+#   filter(type == "polynomial") %>%
+#   filter(x_dist == "uniform") %>%
+#   filter(n == 300) %>%
+#   filter(shape == 1) %>%
+#   group_by(unique_lineup_id) %>%
+#   summarise(effect_size = first(effect_size)) %>%
+#   arrange(effect_size)
+
+vi_lineup$poly_300$data %>%
   VI_MODEL$plot_lineup(theme = theme_light(), remove_axis = TRUE, remove_grid_line = TRUE)
@@
 Figure \ref{fig:first-example-lineup} is an example of a lineup protocol.
-If the actual data plot at position 7 is identifiable, then it is evidence for the rejection of $H_0$ that the regression model is correctly specified.
-In fact, the actual residual plot is obtained from a misspecified regression model with non-linearity issue.
+If the actual data plot at position $2^2 + 2$ is identifiable, then it is evidence for the rejection of $H_0$ that the regression model is correctly specified.
+In fact, the actual residual plot is obtained from a misspecified regression model with non-linearity defect.
 
 The effectiveness of lineup protocol is validated by @majumder_validation_2013 under relatively simple classical normal linear regression model settings with one or two regressors.
-Their results suggest visual test is capable of testing the significance of a single regressor with a similar power as a t-test, though they expressed that in general it is unnecessary to use visual inference if there exists a conventional test and they didn't expect the visual test to perform equally well as the conventional test.
+Their results suggest that visual test is capable of testing the significance of a single regressor with a similar power as a t-test, though they express that in general it is unnecessary to use visual inference if there exists a conventional test and they do not expect the visual test to perform equally well as the conventional test.
 In their third experiment, where there does not exist a proper conventional test, visual test outperforms the conventional test for a large margin.
 This is encouraging as it promotes the use of visual inference in border field of data science where there are no existing statistical testing procedures.
-Visual inference have also been integrated into diagnostic of hierarchical linear models by @loy2013diagnostic, @loy2014hlmdiag and @loy2015you.
+Visual inference have also already been integrated into diagnostic of hierarchical linear models by @loy2013diagnostic, @loy2014hlmdiag and @loy2015you.
 They use lineup protocols to judge the assumption of linearity, normality and constant error variance for both the level-1 and level-2 residuals.

### 2022-12-13 paper_comparison/paper_comparison/paper_comparison.tex
use a different residual plot in VI example

@@
 \caption{Visual testing is conducted using a lineup, as in the example here.
-The residual plot computed from the observed data (plot 7, exhibiting non-linearity) is embedded among 19 null plots, where the residuals were simulated from a standard error model.
+The residual plot computed from the observed data (plot $2^2 + 2$, exhibiting non-linearity) is embedded among 19 null plots, where the residuals were simulated from a standard error model.
 Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot.
-A small $p$-value would result from a substantial number selecting plot 7.}\label{fig:first-example-lineup}
+A small $p$-value would result from a substantial number selecting plot $2^2 + 2$.}\label{fig:first-example-lineup}
 \end{figure}
@@
 Figure \ref{fig:first-example-lineup} is an example of a lineup
-protocol. If the actual data plot at position 7 is identifiable, then it
-is evidence for the rejection of \(H_0\) that the regression model is
-correctly specified. In fact, the actual residual plot is obtained from
-a misspecified regression model with non-linearity issue.
+protocol. If the actual data plot at position \(2^2 + 2\) is
+identifiable, then it is evidence for the rejection of \(H_0\) that the
+regression model is correctly specified. In fact, the actual residual
+plot is obtained from a misspecified regression model with non-linearity
+defect.
 
@@
 normal linear regression model settings with one or two regressors.
-Their results suggest visual test is capable of testing the significance
-of a single regressor with a similar power as a t-test, though they
-expressed that in general it is unnecessary to use visual inference if
-there exists a conventional test and they didn't expect the visual test
-to perform equally well as the conventional test. In their third
-experiment, where there does not exist a proper conventional test,
+Their results suggest that visual test is capable of testing the
+significance of a single regressor with a similar power as a t-test,
+though they express that in general it is unnecessary to use visual
+inference if there exists a conventional test and they do not expect the
+visual test to perform equally well as the conventional test. In their
+third experiment, where there does not exist a proper conventional test,
 visual test outperforms the conventional test for a large margin. This
@@
 field of data science where there are no existing statistical testing
-procedures. Visual inference have also been integrated into diagnostic
-of hierarchical linear models by \citet{loy2013diagnostic},
+procedures. Visual inference have also already been integrated into
+diagnostic of hierarchical linear models by \citet{loy2013diagnostic},
 \citet{loy2014hlmdiag} and \citet{loy2015you}. They use lineup protocols

### 2022-12-13 paper_comparison/paper_comparison/paper_comparison.Rmd
clarify non-normality is not the focus in the experiment design

@@
 
-# Experimental design
-
-(This section discusses the experimental design including the motivation of the experiment, an overview of the experiment, the simulation setting of the depatures from good residual plots, parameter choices, allocation of the lineups and other technical details.)
-
-<!-- ## Experiment overview-->
-
-<!-- With our knowledge, the effectiveness of visual test relative to the equivalent conventional test in regression diagnostics has not been validated.
-Particularly, its ability to detect non-linearity and heteroskedasticity compared to model structural test and Breusch–Pagan test.
-Meanwhile, as mentioned in Section \ref{introduction} section, we keen to understand the differences between conventional hypothesis testing and graphical diagnostics in the application of linear regression diagnostics and why conventional tests are not preferred by data analysts in linear regression diagnostics.
-Attempting to find the answer to these questions, three experiments were conducted.
--->
+# Experiment design
 
@@
 Two types of departures, namely non-linearity and heteroskedasticity, were considered with the corresponding data generating process being designed for experiment I and II.
+As mentioned earlier, residual plot is not suitable for revealing non-normality, thus it was not considered in our experiments.
 

### 2022-12-13 paper_comparison/paper_comparison/paper_comparison.tex
clarify non-normality is not the focus in the experiment design

@@
 
-\hypertarget{experimental-design}{%
-\section{Experimental design}\label{experimental-design}}
-
-(This section discusses the experimental design including the motivation
-of the experiment, an overview of the experiment, the simulation setting
-of the depatures from good residual plots, parameter choices, allocation
-of the lineups and other technical details.)
+\hypertarget{experiment-design}{%
+\section{Experiment design}\label{experiment-design}}
 
@@
 were considered with the corresponding data generating process being
-designed for experiment I and II.
+designed for experiment I and II. As mentioned earlier, residual plot is
+not suitable for revealing non-normality, thus it was not considered in
+our experiments.
 

### 2022-12-13 paper_comparison/paper_comparison/paper_comparison.Rmd
revised experiment section

@@
 The next background section describes the types of departures that one expects to detect, and describes a formal statistical process for reading residual plots, called visual inference.
-Section \ref{experimental-design} describes the experiment design to compare the decision made by formal hypothesis testing, and how humans would read diagnostic plots.
+Section \ref{experiment-design} describes the experiment design to compare the decision made by formal hypothesis testing, and how humans would read diagnostic plots.
 The results are reported in Section \ref{results}.
@@
     ggtitle("A. Good residuals") + 
-    xlab("fitted values") + 
-    ylab("residuals"),
+    xlab("Fitted values") + 
+    ylab("Residuals"),
 
@@
     ggtitle("B. Non-linearity") +
-    xlab("fitted values") + 
-    ylab("residuals"),
+    xlab("Fitted values") + 
+    ylab("Residuals"),
 
@@
     ggtitle("C. Heteroskedasticity") +
-    xlab("fitted values") + 
-    ylab("residuals"),
+    xlab("Fitted values") + 
+    ylab("Residuals"),
 
@@
   ggtitle(quote(He[2]:"U")) + 
-  theme_light() -> p1
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") -> p1
 
@@
   ggtitle(quote(He[3]:"S")) + 
-  theme_light() -> p2
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") -> p2
 
@@
   ggtitle(quote(He[6]:"M")) + 
-  theme_light() -> p3
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") -> p3
 
@@
   ggtitle(quote(He[18]:"Triple-U")) + 
-  theme_light() -> p4
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") -> p4
 
@@
 
-Experiment I is designed to study the ability of human subjects to detect the effect of a non-linear term $\boldsymbol{z}$ constructed using Hermite polynomials [Hermite ref here] on random vector $\boldsymbol{x}$ formulated as
+Experiment I is designed to study the ability of human subjects to detect the effect of a non-linear term $\boldsymbol{z}$ constructed using Hermite polynomials [unsure how to cite 1810 work] on random vector $\boldsymbol{x}$ formulated as
 
@@
 data.frame(j = c("2", "3", "6", "18", ""),
-           x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$", "NA"),
+           x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$", ""),
            sigma = c("0.25", "1.00", "2.00", "4.00", ""),
@@
                booktabs = TRUE, 
-               caption  = 'Description of all factors involved in the nonlinear and heteroskedasticity studies.', 
+               caption  = 'Description of all factors involved in the non-linear and heteroskedasticity studies.', 
                label = "model-factor-table" ) %>%
@@
                                  "SD (sym3)", #"Standard deviation of polynomial model\n(sym3)",
-                                 "Poly Shape (sym4)", # "Shape of heteroskedasticity\n(sym4)", 
+                                 "Heteroskedasticity Shape (sym4)", # "Shape of heteroskedasticity\n(sym4)", 
                                  "Heteroskedasticity (sym5)", # Variance factor of heteroskedasticity model\n(sym5)",
@@
 ```{r different-dist, fig.height = 2, fig.cap="Variations in fitted values, that might affect perception of residual plots.
-Four different distribution of $x_{raw}$ are used in the experiment to provide various visual patterns."}
+Four different distribution of $X_{raw}$ are used in the experiment to provide various visual patterns."}
 ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)
@@
   ggtitle("A. Uniform") + #quote(U(-1,1))) + 
-  theme_light() -> p1
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") -> p1
 
@@
   ggtitle("B. Normal") + #quote(N(0,0.3^2))) + 
-  theme_light() -> p2
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") -> p2
 
@@
   ggtitle("C. Skewed") + #quote(lognormal(0,0.6^2)/3)) + 
-  theme_light() -> p3
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") -> p3
 
@@
   ggtitle("D. Discrete") + #quote(U~"{1,5}")) + 
-  theme_light() -> p4
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") -> p4
 
@@
 Figure \ref{fig:example-poly-lineup} shows one of the lineups used in experiment I.
-This lineup was produced under $j = 6$ and $X_{raw} \sim N(0.0.3^2)$.
-The actual data plot location was four.
+This lineup is produced under $j = 6$ and $X_{raw} \sim N(0.0.3^2)$.
+The actual data plot location is $2^3 - 4$.
 All five subjects correctly identified the actual data plot for this lineup.
@@
 
-For $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the regressor $x$, which will lead to the presence of heteroskedasticity.
+For $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the regressor $\boldsymbol{x}$, which will lead to the presence of heteroskedasticity.
 Visual patterns were simulated using three different shapes ($a$ = -1, 0, 1) and the same four different distribution of $X_{raw}$ used in experiment I.
@@
   ggtitle(quote("A. a = -1: Left-triangle")) +
-  theme_light() -> p1
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") -> p1
 
@@
   ggtitle(quote("B. a = 0: Butterfly")) +
-  theme_light() -> p2
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") -> p2
 
@@
   ggtitle(quote("C. a = 1: Right-triangle")) +
-  theme_light() -> p3
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") -> p3
 
@@
 ```{r example-heter-lineup, fig.cap="Lineup heter-169 in experiment II. Can you spot the most different plot? \\label{fig:example-heter-lineup}"}
-VI_MODEL$plot_lineup(vi_lineup[["heter_169"]]$data, 
+
+# vi_survey %>%
+#   filter(!attention_check) %>%
+#   filter(!null_lineup) %>%
+#   filter(type == "heteroskedasticity") %>%
+#   filter(x_dist == "uniform") %>%
+#   filter(a == -1) %>%
+#   filter(n == 300) %>%
+#   group_by(unique_lineup_id) %>%
+#   summarise(effect_size = first(effect_size)) %>%
+#   arrange(effect_size)
+
+# vi_survey %>%
+#   filter(unique_lineup_id == "heter_471") %>%
+#   View()
+
+VI_MODEL$plot_lineup(vi_lineup$heter_471$data, 
                      remove_grid_line = TRUE, 
@@
 An example lineup of this model is shown in Figure \ref{fig:example-heter-lineup} with $a = -1$ and $X_{raw} \sim U(-1, 1)$.
-The actual data plot location was 15.
-8 out of 11 subjects correctly identified the actual data plot for this lineup.
-
-## Experimental setup
+The actual data plot location is $2^4 + 2$.
+Nine out of 11 subjects correctly identified the actual data plot for this lineup.
+
+## Experiment setup
 
@@
 Figure \ref{fig:different-sigma} and \ref{fig:different-b} demonstrate the impact of these two parameters.
-A large value of $\sigma$ will increase the variation of the error of the non-linear model and decrease the visibility of the visual pattern.
+A large value of $\sigma$ will increase the variation of the error of the non-linearity model and decrease the visibility of the visual pattern.
 The parameter $b$ controls the ratio of the standard deviation of the heteroskedasticity across the domain of the regressor.
@@
   ggtitle(quote(n~"="~300)) + 
-  theme_light() -> p1
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") -> p1
 
@@
   ggtitle(quote(n~"="~50)) + 
-  theme_light() -> p2
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") -> p2
 
@@
   ggtitle(quote(n~"="~100)) + 
-  theme_light() -> p3
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") -> p3
 
@@
   ggtitle(quote(sigma~"="~0.5)) + 
-  theme_light() -> p1
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") -> p1
 
@@
   ggtitle(quote(sigma~"="~1)) + 
-  theme_light() -> p2
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") -> p2
 
@@
   ggtitle(quote(sigma~"="~2)) + 
-  theme_light() -> p3
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") -> p3
 
@@
   ggtitle(quote(sigma~"="~4)) + 
-  theme_light() -> p4
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") -> p4
 
@@
   ggtitle(quote("b = 0.25")) +
-  theme_light() -> p1
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") -> p1
 
@@
   ggtitle(quote("b = 1")) +
-  theme_light() -> p2
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") -> p2
 
@@
   ggtitle(quote("b = 4")) +
-  theme_light() -> p3
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") -> p3
 
@@
   ggtitle(quote("b = 16")) +
-  theme_light() -> p4
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") -> p4
 
@@
   ggtitle(quote("b = 64")) +
-  theme_light() -> p5
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") -> p5
 
@@
 
-Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in $(4 \times 4 \times 4 \times 3 + 4 \times 3 \times 5 \times 3) \times 3 = 1116$ different lineups.
-In addition, each lineup is designed to be evaluated by five different subjects.
-After attempting some pilot studies internally in our research group, we decide to present a block of 20 lineups to every subject.
-And to ensure the quality of the survey data, two lineups with obvious visual patterns are included as attention checks.
-Thus, $576 \times 5/(20-2) = 160$ and $540 \times 5 /(20-2) = 150$ subjects were recruited to satisfy the design of the experiment I and experiment II respectively.
-
-As mentioned in Section [the alpha section], $\alpha$ used in Equation \ref{eq:eq:pvalue-beta-binomial} needs to be estimated using Rorschach lineups.
-Hence, 36 Rorschach lineups with all combinations of $n$ and $X_{raw}$ and three replications are designed to be included in experiment III.
-All Rorschach lineups are planned to be evaluated by 20 subjects.
-However, presenting only Rorschach lineups to subjects are considered to be bad practices as subjects will lose interest quickly.
-We planned to also collect 6 more evaluations on the 279 lineups with uniform distribution, resulted in $(36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3)/(20-2) = 133$ subjects recruited for experiment III.
+Three replications were made for each of the parameter values shown in Table \ref{tab:model-factor-table} resulting in $(4 \times 4 \times 4 \times 3 + 4 \times 3 \times 5 \times 3) \times 3 = 1116$ different lineups.
+In addition, each lineup was designed to be evaluated by five different subjects.
+After attempting some pilot studies internally in our research group, we decided to present a block of 20 lineups to every subject.
+And to ensure the quality of the survey data, two lineups with obvious visual patterns were included as attention checks.
+Thus, $576 \times 5 \div (20-2) = 160$ and $540 \times 5 \div (20-2) = 150$ subjects were recruited to satisfy the design of the experiment I and experiment II respectively.
+
+As mentioned in Section \ref{calculating-p-values}, $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using Rorschach lineups.
+Hence, 36 lineups consist of only null plots with all combinations of $n$ and $X_{raw}$ and three replications are designed to be included in experiment III.
+Note that we generated the lineups using the same simulation mechanism discussed in Section \ref{sampling-from-the-null-distribution} rather than from a null model.
+We overlooked this detail until the data had been collected.
+Thus, we used another valid method suggested in vanderplas2021statistical which was designed for typical lineups to estimate $\alpha$ correctly.
+However, by not using Rorschach lineups, we did lose some accuracy in the estimation.
+We recommend future studies to generate Rorschach lineups strictly following the definition given by buja_statistical_2009.
+
+All lineups consist of only null plots are planned to be evaluated by 20 subjects.
+However, presenting only these lineups to subjects are considered to be bad practices as subjects will lose interest quickly.
+We planned to also collect 6 more evaluations on the 279 lineups with $X_{raw} \sim U(-1,1)$, resulted in $(36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3) \div (20-2) = 133$ subjects recruited for experiment III.
 
@@
 
-Subjects for all three experiments were recruited from an crowdsourcing platform called Prolific [ref here].
+Subjects for all three experiments were recruited from an crowdsourcing platform called Prolific [palan2018prolific].
 Prescreening procedure was applied during the recruitment, subjects were required to be fluent in English, with $98\%$ minimum approval rate in other studies and 10 minimum submissions.

### 2022-12-13 paper_comparison/paper_comparison/paper_comparison.tex
revised experiment section

@@
 describes a formal statistical process for reading residual plots,
-called visual inference. Section \ref{experimental-design} describes the
+called visual inference. Section \ref{experiment-design} describes the
 experiment design to compare the decision made by formal hypothesis
@@
 detect the effect of a non-linear term \(\boldsymbol{z}\) constructed
-using Hermite polynomials {[}Hermite ref here{]} on random vector
-\(\boldsymbol{x}\) formulated as
+using Hermite polynomials {[}unsure how to cite 1810 work{]} on random
+vector \(\boldsymbol{x}\) formulated as
 
@@
 
-\caption{\label{tab:model-factor-table}Description of all factors involved in the nonlinear and heteroskedasticity studies.}
+\caption{\label{tab:model-factor-table}Description of all factors involved in the non-linear and heteroskedasticity studies.}
 \centering
@@
 \toprule
-\multicolumn{1}{c}{Poly Order ($j$)} & \multicolumn{1}{c}{\makecell[c]{Distribution of $X_{raw}$\\}} & \multicolumn{1}{c}{SD ($\sigma$)} & \multicolumn{1}{c}{Poly Shape ($a$)} & \multicolumn{1}{c}{Heteroskedasticity ($b$)} & \multicolumn{1}{c}{Size ($n$)} \\
+\multicolumn{1}{c}{Poly Order ($j$)} & \multicolumn{1}{c}{\makecell[c]{Distribution of $X_{raw}$\\}} & \multicolumn{1}{c}{SD ($\sigma$)} & \multicolumn{1}{c}{Heteroskedasticity Shape ($a$)} & \multicolumn{1}{c}{Heteroskedasticity ($b$)} & \multicolumn{1}{c}{Size ($n$)} \\
 \cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
@@
 18 & $U\{1, 5\}$ & 4.00 &  & 16.00 & \\
- & NA &  &  & 64.00 & \\
+ &  &  &  & 64.00 & \\
 \bottomrule
@@
 \caption{Variations in fitted values, that might affect perception of residual plots.
-Four different distribution of $x_{raw}$ are used in the experiment to provide various visual patterns.}\label{fig:different-dist}
+Four different distribution of $X_{raw}$ are used in the experiment to provide various visual patterns.}\label{fig:different-dist}
 \end{figure}
@@
 Figure \ref{fig:example-poly-lineup} shows one of the lineups used in
-experiment I. This lineup was produced under \(j = 6\) and
-\(X_{raw} \sim N(0.0.3^2)\). The actual data plot location was four. All
-five subjects correctly identified the actual data plot for this lineup.
+experiment I. This lineup is produced under \(j = 6\) and
+\(X_{raw} \sim N(0.0.3^2)\). The actual data plot location is
+\(2^3 - 4\). All five subjects correctly identified the actual data plot
+for this lineup.
 
@@
 For \(b \neq 0\), the variance-covariance matrix of the error term
-\(\boldsymbol{\varepsilon}\) is correlated with the regressor \(x\),
-which will lead to the presence of heteroskedasticity. Visual patterns
-were simulated using three different shapes (\(a\) = -1, 0, 1) and the
-same four different distribution of \(X_{raw}\) used in experiment I. A
-summary of the factors is given in Table \ref{tab:model-factor-table}.
+\(\boldsymbol{\varepsilon}\) is correlated with the regressor
+\(\boldsymbol{x}\), which will lead to the presence of
+heteroskedasticity. Visual patterns were simulated using three different
+shapes (\(a\) = -1, 0, 1) and the same four different distribution of
+\(X_{raw}\) used in experiment I. A summary of the factors is given in
+Table \ref{tab:model-factor-table}.
 
@@
 \ref{fig:example-heter-lineup} with \(a = -1\) and
-\(X_{raw} \sim U(-1, 1)\). The actual data plot location was 15. 8 out
-of 11 subjects correctly identified the actual data plot for this
-lineup.
-
-\hypertarget{experimental-setup}{%
-\subsection{Experimental setup}\label{experimental-setup}}
+\(X_{raw} \sim U(-1, 1)\). The actual data plot location is \(2^4 + 2\).
+Nine out of 11 subjects correctly identified the actual data plot for
+this lineup.
+
+\hypertarget{experiment-setup}{%
+\subsection{Experiment setup}\label{experiment-setup}}
 
@@
 demonstrate the impact of these two parameters. A large value of
-\(\sigma\) will increase the variation of the error of the non-linear
+\(\sigma\) will increase the variation of the error of the non-linearity
 model and decrease the visibility of the visual pattern. The parameter
@@
 
-Three replications are made for each of the parameter values shown in
-Table \ref{tab:parameter-table} resulting in
+Three replications were made for each of the parameter values shown in
+Table \ref{tab:model-factor-table} resulting in
 \((4 \times 4 \times 4 \times 3 + 4 \times 3 \times 5 \times 3) \times 3 = 1116\)
-different lineups. In addition, each lineup is designed to be evaluated
+different lineups. In addition, each lineup was designed to be evaluated
 by five different subjects. After attempting some pilot studies
-internally in our research group, we decide to present a block of 20
+internally in our research group, we decided to present a block of 20
 lineups to every subject. And to ensure the quality of the survey data,
-two lineups with obvious visual patterns are included as attention
-checks. Thus, \(576 \times 5/(20-2) = 160\) and
-\(540 \times 5 /(20-2) = 150\) subjects were recruited to satisfy the
-design of the experiment I and experiment II respectively.
-
-As mentioned in Section {[}the alpha section{]}, \(\alpha\) used in
-Equation \ref{eq:eq:pvalue-beta-binomial} needs to be estimated using
-Rorschach lineups. Hence, 36 Rorschach lineups with all combinations of
-\(n\) and \(X_{raw}\) and three replications are designed to be included
-in experiment III. All Rorschach lineups are planned to be evaluated by
-20 subjects. However, presenting only Rorschach lineups to subjects are
+two lineups with obvious visual patterns were included as attention
+checks. Thus, \(576 \times 5 \div (20-2) = 160\) and
+\(540 \times 5 \div (20-2) = 150\) subjects were recruited to satisfy
+the design of the experiment I and experiment II respectively.
+
+As mentioned in Section \ref{calculating-p-values}, \(\alpha\) used in
+Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using
+Rorschach lineups. Hence, 36 lineups consist of only null plots with all
+combinations of \(n\) and \(X_{raw}\) and three replications are
+designed to be included in experiment III. Note that we generated the
+lineups using the same simulation mechanism discussed in Section
+\ref{sampling-from-the-null-distribution} rather than from a null model.
+We overlooked this detail until the data had been collected. Thus, we
+used another valid method suggested in vanderplas2021statistical which
+was designed for typical lineups to estimate \(\alpha\) correctly.
+However, by not using Rorschach lineups, we did lose some accuracy in
+the estimation. We recommend future studies to generate Rorschach
+lineups strictly following the definition given by
+buja\_statistical\_2009.
+
+All lineups consist of only null plots are planned to be evaluated by 20
+subjects. However, presenting only these lineups to subjects are
 considered to be bad practices as subjects will lose interest quickly.
 We planned to also collect 6 more evaluations on the 279 lineups with
-uniform distribution, resulted in
-\((36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3)/(20-2) = 133\)
+\(X_{raw} \sim U(-1,1)\), resulted in
+\((36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3) \div (20-2) = 133\)
 subjects recruited for experiment III.
@@
 Subjects for all three experiments were recruited from an crowdsourcing
-platform called Prolific {[}ref here{]}. Prescreening procedure was
-applied during the recruitment, subjects were required to be fluent in
-English, with \(98\%\) minimum approval rate in other studies and 10
+platform called Prolific {[}palan2018prolific{]}. Prescreening procedure
+was applied during the recruitment, subjects were required to be fluent
+in English, with \(98\%\) minimum approval rate in other studies and 10
 minimum submissions. During the experiment, every subject was presented

### 2022-12-13 paper_comparison/paper_comparison/paper_comparison.Rmd
swap power cmp and p-value cmp
fix plot in power cmp

@@
 
-## P-value comparison (remove?) (closer to the topic of the paper) (put it after power)
-
-Visual inference and conventional hypothesis testing are two distinct statistical procedures.
-As displayed in Figure \ref{fig:p-value-comparison}, the visual test gives higher p-values than the conventional test for both polynomial and heteroskedasticity models in most of the time, suggesting that humans have lower confidence and much higher tolerance of the residual departures than the conventional test.
-
-In particular, given 95% confidence level, both tests will reject $H_0$ in only around 58% of the cases.
-And in 41% of the cases, the visual test does not reject $H_0$ but the conventional test does.
-Since fail to reject $H_0$ in a visual test means that there is no obvious visual discoveries found in the residual plot, analysts and the general public as the consumers of the output may not fully believe in the rejection of $H_0$ even though the alternative hypothesis is true.
-Even if the rejection is accepted, the model violation may not be considered as impactful due to the fact that the departures are not clearly visible.
-
-This makes it sounds like people are arrogantly ignoring the accuracy of conventional tests and trust too much to their visual perception, but it is not.
-In fact, the sensitivity of the conventional test could sometimes distract and discourage analysts from finding a simple but good linear approximation to the data.
-The rejection of $H_0$ because of acceptable and negligible residual departure is not practically meaningful and useful.
-This also partly explains why abundance of literature suggesting the use of data plot in regression diagnostics.
-
-```{r p-value-comparison, fig.height=4, fig.cap="Visual test p-value compared to conventional test p-value for lineups produced by heteroskedasticity and non-linearity model.
-The scatter plot is drawn on a square root scale.
-Every dot on the plot corresponds to a lineup.
-Most of the dots are fallen on the left of the y = x line indicating visual test p-value is almost always higher than conventional test p-value.
-The blue curve is a smoothing of the dots showing a positive trend.
-"}
-vi_survey %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
-  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-  left_join(vi_survey %>% 
-           count(unique_lineup_id) %>%
-           rename(num_subject = n)) %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  ggplot() +
-  geom_abline() +
-  geom_point(aes(conventional_p_value, p_value), alpha = 0.4) +
-  geom_smooth(aes(conventional_p_value, p_value), se = FALSE, span = 0.6) +
-  scale_x_sqrt() +
-  scale_y_sqrt() +
-  facet_wrap(~type) +
-  ylab("Visual test p-value on square root scale") +
-  xlab("Conventional test p-value on square root scale") +
-  theme_light()
-```
-
 ## Power comparison
 
-Figure \ref{fig:polypower} show the estimated power of visual test on lineups produced by the non-linearity model with the $X_{raw}$ following a uniform distribution, against the natural logarithm of the effect $log_e(E)$ with comparison to the power of an exact test - F-test, and two other residual-based conventional tests commonly used in regression diagnostics but for testing heteroskedasticity and non-normality respectively.
+Figure \ref{fig:polypower} shows the estimated power of visual test on lineups produced by the non-linearity model with $X_{raw} \sim U(-1,1)$, against the natural logarithm of the effect $log_e(E)$ with comparison to the power of an exact test - RESET test, and two other residual-based conventional tests commonly used in regression diagnostics but for testing heteroskedasticity and non-normality respectively.
 At the bottom of the figure \ref{fig:polypower}, there are a sequence of example residuals plots with increasing levels of $log_e(E)$.
@@
 Figure \ref{fig:heterpower} is similar to Figure \ref{fig:polypower}, but shows corresponding information on lineups produced by the heteroskedasticity model.
-In this scenario, the visual test is compared to an approximate test - Breusch-Pagan test, and two other inappropriate tests - F-test and Shapiro-Wilk test.
+In this scenario, the visual test is compared to an approximate test - Breusch-Pagan test, and two other inappropriate tests - RESET test and Shapiro-Wilk test.
 
@@
 
-```{r eval = FALSE}
-filter(vi_survey, type == "polynomial") %>%
-  filter(x_dist == "uniform") %>%
-  mutate(log_es = abs(log(effect_size) - 4)) %>%
-  arrange(log_es) %>%
-  View()
-```
-
 <!-- TODO:difference between the simple formula and the simulation approach -->
@@
 }
-```
-
-```{r power-poly-uniform, fig.cap = "...", eval = FALSE}
-vi_survey %>%
-  filter(x_dist == "uniform") %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type == "polynomial") %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
-  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_dirichlet, times = 500), 
-              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "Visual test"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03) +
-  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
-                filter(x_dist == "uniform") %>%
-                pivot_longer(F_p_value:SW_p_value) %>%
-                mutate(name = gsub("_p_value", " test", name)), 
-              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
-              method = "glm", 
-              method.args = list(family = binomial),
-              se = FALSE) +
-  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) + geom_point(data = ~filter(.x, unique_lineup_id %in% c("poly_166", "poly_269", "poly_396", "poly_574")), aes(log_effect_size, prop_detect), alpha = 1, col = "red", size = 2) +
-  theme_light(base_size = 5) +
-  scale_color_brewer(palette = "Dark2", direction = -1) +
-  xlab("Natural logarithm of effect size") +
-  ylab("Power") -> p
-
-VI_MODEL$plot_lineup(vi_lineup[["poly_166"]]$data, 
-                     remove_grid_line = TRUE, 
-                     theme = theme_light(base_size = 5),
-                     remove_axis = TRUE,
-                     size = 0.1) + 
-  ggtitle(bquote(log[e](E) %~~% 1~","~"F-test p-value" == .(format(filter(vi_survey, unique_lineup_id == "poly_166")$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(vi_survey, unique_lineup_id == "poly_166")$p_value[1], digits = 3)))) -> p1
-
-VI_MODEL$plot_lineup(vi_lineup[["poly_269"]]$data,
-                     remove_grid_line = TRUE, 
-                     theme = theme_light(base_size = 5),
-                     remove_axis = TRUE,
-                     size = 0.1) + 
-  ggtitle(bquote(log[e](E) %~~% 2~","~"F-test p-value" == .(format(filter(vi_survey, unique_lineup_id == "poly_269")$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(vi_survey, unique_lineup_id == "poly_269")$p_value[1], digits = 3)))) -> p2
-
-VI_MODEL$plot_lineup(vi_lineup[["poly_396"]]$data, 
-                     remove_grid_line = TRUE, 
-                     theme = theme_light(base_size = 5),
-                     remove_axis = TRUE, 
-                     size = 0.1) + 
-  ggtitle(bquote(log[e](E) %~~% 3~","~"F-test p-value" == .(format(filter(vi_survey, unique_lineup_id == "poly_396")$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(vi_survey, unique_lineup_id == "poly_396")$p_value[1], digits = 3)))) -> p3
-
-VI_MODEL$plot_lineup(vi_lineup[["poly_574"]]$data, 
-                     remove_grid_line = TRUE, 
-                     theme = theme_light(base_size = 5),
-                     remove_axis = TRUE,
-                     size = 0.1) + 
-  ggtitle(bquote(log[e](E) %~~% 4~","~"F-test p-value" == .(format(filter(vi_survey, unique_lineup_id == "poly_574")$conventional_p_value[1], digits = 3))~", Visual p-value " == .(format(filter(vi_survey, unique_lineup_id == "poly_574")$p_value[1], digits = 3)))) -> p4
-
-design <- c(patchwork::area(1, 1, 1, 8),
-            patchwork::area(2, 1, 2, 4), 
-            patchwork::area(2, 5, 2, 8), 
-            patchwork::area(3, 1, 3, 4), 
-            patchwork::area(3, 5, 3, 8))
-
-patchwork::wrap_plots(p, B = p1, C = p2, D = p3, E = p4, design = design)
 ```
@@
              aes(log_effect_size, reject_dirichlet, size = factor(num_response)), alpha = 0.3)  +
-  geom_segment(data = data.frame(xi = es), aes(x = xi, xend = xi, y = 0, yend = 1), alpha = 0.4, linetype = 2) +
+  geom_segment(data = data.frame(xi = es), aes(x = xi, xend = xi, y = 0, yend = 1), alpha = 0.3, col = "grey") +
   theme_light(base_size = 5) +
@@
   scale_size_manual(values = c(1, 2, 3)) +
+  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
   xlab("Effect size (log e)") +
@@
 
-```{r polypower, fig.height = 3, fig.width = 5, fig.cap="Comparison of power between different tests for nonlinear patterns.
+```{r polypower, fig.height = 3, fig.width = 5, fig.cap="Comparison of power between different tests for non-linear patterns.
 Main plot shows the power curves, with dots indicating human evaluations of lineups.
 Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
-Where would you draw the line of too much nonlinearity in the residuals?
-For the F test this is around log effect size 0.5, but for the visual test it is around 3."}
+Where would you draw the line of too much non-linearity in the residuals?
+For the RESET test this is around log effect size 1, but for the visual test it is around 3."}
 patchwork::wrap_plots(append(plist, list(p), 
@@
              aes(log_effect_size, reject_dirichlet, size = factor(num_response)), alpha = 0.3)  +
-  geom_segment(data = data.frame(xi = es), aes(x = xi, xend = xi, y = 0, yend = 1), alpha = 0.4, linetype = 2) +
+  geom_segment(data = data.frame(xi = es), aes(x = xi, xend = xi, y = 0, yend = 1), alpha = 0.3, col = "grey") +
   theme_light(base_size = 5) +
   scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
+  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
   scale_size_manual(values = c(1, 2, 3)) +
@@
 
-```{r heterpower, fig.height = 3, fig.width = 5, fig.cap="..."}
+```{r heterpower, fig.height = 3, fig.width = 5, fig.cap="Comparison of power between different tests for heteroskedasticity patterns.
+Main plot shows the power curves, with dots indicating human evaluations of lineups.
+Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
+Where would you draw the line of too much heteroskedasticity in the residuals?
+For the BP test this is around log effect size 1.5, but for the visual test it is around 3."}
 patchwork::wrap_plots(append(plist, list(p), after = 0), 
@@
 <!-- Note that both the p-value and power comparison are not directly based on the diagnostic plots recommended by abundance of literature, but  -->
+
+## P-value comparison (remove?) (closer to the topic of the paper) (put it after power)
+
+Visual inference and conventional hypothesis testing are two distinct statistical procedures.
+As displayed in Figure \ref{fig:p-value-comparison}, the visual test gives higher p-values than the conventional test for both polynomial and heteroskedasticity models in most of the time, suggesting that humans have lower confidence and much higher tolerance of the residual departures than the conventional test.
+
+In particular, given 95% confidence level, both tests will reject $H_0$ in only around 58% of the cases.
+And in 41% of the cases, the visual test does not reject $H_0$ but the conventional test does.
+Since fail to reject $H_0$ in a visual test means that there is no obvious visual discoveries found in the residual plot, analysts and the general public as the consumers of the output may not fully believe in the rejection of $H_0$ even though the alternative hypothesis is true.
+Even if the rejection is accepted, the model violation may not be considered as impactful due to the fact that the departures are not clearly visible.
+
+This makes it sounds like people are arrogantly ignoring the accuracy of conventional tests and trust too much to their visual perception, but it is not.
+In fact, the sensitivity of the conventional test could sometimes distract and discourage analysts from finding a simple but good linear approximation to the data.
+The rejection of $H_0$ because of acceptable and negligible residual departure is not practically meaningful and useful.
+This also partly explains why abundance of literature suggesting the use of data plot in regression diagnostics.
+
+```{r p-value-comparison, fig.height=4, fig.cap="Visual test p-value compared to conventional test p-value for lineups produced by heteroskedasticity and non-linearity model.
+The scatter plot is drawn on a square root scale.
+Every dot on the plot corresponds to a lineup.
+Most of the dots are fallen on the left of the y = x line indicating visual test p-value is almost always higher than conventional test p-value.
+The blue curve is a smoothing of the dots showing a positive trend.
+"}
+vi_survey %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  left_join(vi_survey %>% 
+           count(unique_lineup_id) %>%
+           rename(num_subject = n)) %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  ggplot() +
+  geom_abline() +
+  geom_point(aes(conventional_p_value, p_value), alpha = 0.4) +
+  geom_smooth(aes(conventional_p_value, p_value), se = FALSE, span = 0.6) +
+  scale_x_sqrt() +
+  scale_y_sqrt() +
+  facet_wrap(~type) +
+  ylab("Visual test p-value on square root scale") +
+  xlab("Conventional test p-value on square root scale") +
+  theme_light()
+```
 

### 2022-12-13 paper_comparison/paper_comparison/paper_comparison.tex
swap power cmp and p-value cmp
fix plot in power cmp

@@
 
+\hypertarget{power-comparison}{%
+\subsection{Power comparison}\label{power-comparison}}
+
+Figure \ref{fig:polypower} shows the estimated power of visual test on
+lineups produced by the non-linearity model with
+\(X_{raw} \sim U(-1,1)\), against the natural logarithm of the effect
+\(log_e(E)\) with comparison to the power of an exact test - RESET test,
+and two other residual-based conventional tests commonly used in
+regression diagnostics but for testing heteroskedasticity and
+non-normality respectively. At the bottom of the figure
+\ref{fig:polypower}, there are a sequence of example residuals plots
+with increasing levels of \(log_e(E)\). Readers can evaluate them from
+left to right and determine at which level the departure from a good
+residual plot becomes detectable.
+
+Figure \ref{fig:heterpower} is similar to Figure \ref{fig:polypower},
+but shows corresponding information on lineups produced by the
+heteroskedasticity model. In this scenario, the visual test is compared
+to an approximate test - Breusch-Pagan test, and two other inappropriate
+tests - RESET test and Shapiro-Wilk test.
+
+For the non-linearity model, the power of all four tests increases as
+the effect gets larger. The power curve of F-test climbs aggressively
+from 25\% to around 90\% as \(log_e(E)\) increases from 0 to 2, while
+power of other tests respond inactively to the change of effect and
+remain lower than 25\% throughout the period, showing that F-test is way
+more sensitive to the type of model defects that being considered.
+Meanwhile, no noticeable visual features can be spotted from the example
+residuals plots.
+
+In terms of the heteroskedasticity model, the power of Breusch-Pagan
+test is also almost always greater than the power of visual test. For
+\(0 \leq log_e(E) \leq 2\), where the visual feature is nearly
+unobservable from the example residual plots and the power curve of the
+visual test remains at a low level, the Breusch-Pagan test still have a
+decent amount of chance of rejecting \(H_0\).
+
+The power of visual test arises steadily as \(log_e(E)\) increases from
+2 to 5 in both non-linearity model and heteroskedasticity model,
+suggesting that the effect starts to make significant impact on the
+degree of the presence of the designed visual features. This can also be
+observed from the example residuals plots that when \(log_e(E) = 2.5\),
+a weak ``S-shape'' and a weak ``triangle'' shape are presented in Figure
+\ref{fig:polypower} and Figure \ref{fig:heterpower} respectively. And as
+\(log_e(E)\) increases, the visual pattern becomes much clearer, and the
+power reaches almost 100\% at \(log_e(E) \approx 6\).
+
+The power of all inappropriate tests except for F-test shows improvement
+as the effect increases but at a lower rate than the visual test in both
+scenarios. This coincides the point made by \citet{cook1982residuals}
+that residual-based tests for a specific type of model defect may be
+sensitive to other types of model defects. The power curve of F-test
+remains at around 5\% in Figure \ref{fig:heterpower} since there is no
+non-linear term leave out in the heteroskedasticity model and \(H_0\) of
+the F-test will be always satisfied.
+
+Overall, the power analysis suggests that conventional tests do not
+match how humans perceive residual plots and thus determine model
+violations in these two scenarios. Non-severe model violations with
+human acceptable residual departures could often alarm by the
+conventional tests. This could limit the popularity of conventional
+tests in residual diagnostics among analysts.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/polypower-1}
+
+}
+
+\caption{Comparison of power between different tests for non-linear patterns.
+Main plot shows the power curves, with dots indicating human evaluations of lineups.
+Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
+Where would you draw the line of too much non-linearity in the residuals?
+For the RESET test this is around log effect size 1, but for the visual test it is around 3.}\label{fig:polypower}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/heterpower-1}
+
+}
+
+\caption{Comparison of power between different tests for heteroskedasticity patterns.
+Main plot shows the power curves, with dots indicating human evaluations of lineups.
+Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
+Where would you draw the line of too much heteroskedasticity in the residuals?
+For the BP test this is around log effect size 1.5, but for the visual test it is around 3.}\label{fig:heterpower}
+\end{figure}
+
+tick mark on x. dashed line too strong. use solid line (grey) in light
+color.
+
 \hypertarget{p-value-comparison-remove-closer-to-the-topic-of-the-paper-put-it-after-power}{%
@@
 
-\hypertarget{power-comparison}{%
-\subsection{Power comparison}\label{power-comparison}}
-
-Figure \ref{fig:polypower} show the estimated power of visual test on
-lineups produced by the non-linearity model with the \(X_{raw}\)
-following a uniform distribution, against the natural logarithm of the
-effect \(log_e(E)\) with comparison to the power of an exact test -
-F-test, and two other residual-based conventional tests commonly used in
-regression diagnostics but for testing heteroskedasticity and
-non-normality respectively. At the bottom of the figure
-\ref{fig:polypower}, there are a sequence of example residuals plots
-with increasing levels of \(log_e(E)\). Readers can evaluate them from
-left to right and determine at which level the departure from a good
-residual plot becomes detectable.
-
-Figure \ref{fig:heterpower} is similar to Figure \ref{fig:polypower},
-but shows corresponding information on lineups produced by the
-heteroskedasticity model. In this scenario, the visual test is compared
-to an approximate test - Breusch-Pagan test, and two other inappropriate
-tests - F-test and Shapiro-Wilk test.
-
-For the non-linearity model, the power of all four tests increases as
-the effect gets larger. The power curve of F-test climbs aggressively
-from 25\% to around 90\% as \(log_e(E)\) increases from 0 to 2, while
-power of other tests respond inactively to the change of effect and
-remain lower than 25\% throughout the period, showing that F-test is way
-more sensitive to the type of model defects that being considered.
-Meanwhile, no noticeable visual features can be spotted from the example
-residuals plots.
-
-In terms of the heteroskedasticity model, the power of Breusch-Pagan
-test is also almost always greater than the power of visual test. For
-\(0 \leq log_e(E) \leq 2\), where the visual feature is nearly
-unobservable from the example residual plots and the power curve of the
-visual test remains at a low level, the Breusch-Pagan test still have a
-decent amount of chance of rejecting \(H_0\).
-
-The power of visual test arises steadily as \(log_e(E)\) increases from
-2 to 5 in both non-linearity model and heteroskedasticity model,
-suggesting that the effect starts to make significant impact on the
-degree of the presence of the designed visual features. This can also be
-observed from the example residuals plots that when \(log_e(E) = 2.5\),
-a weak ``S-shape'' and a weak ``triangle'' shape are presented in Figure
-\ref{fig:polypower} and Figure \ref{fig:heterpower} respectively. And as
-\(log_e(E)\) increases, the visual pattern becomes much clearer, and the
-power reaches almost 100\% at \(log_e(E) \approx 6\).
-
-The power of all inappropriate tests except for F-test shows improvement
-as the effect increases but at a lower rate than the visual test in both
-scenarios. This coincides the point made by \citet{cook1982residuals}
-that residual-based tests for a specific type of model defect may be
-sensitive to other types of model defects. The power curve of F-test
-remains at around 5\% in Figure \ref{fig:heterpower} since there is no
-non-linear term leave out in the heteroskedasticity model and \(H_0\) of
-the F-test will be always satisfied.
-
-Overall, the power analysis suggests that conventional tests do not
-match how humans perceive residual plots and thus determine model
-violations in these two scenarios. Non-severe model violations with
-human acceptable residual departures could often alarm by the
-conventional tests. This could limit the popularity of conventional
-tests in residual diagnostics among analysts.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/polypower-1}
-
-}
-
-\caption{Comparison of power between different tests for nonlinear patterns.
-Main plot shows the power curves, with dots indicating human evaluations of lineups.
-Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
-Where would you draw the line of too much nonlinearity in the residuals?
-For the F test this is around log effect size 0.5, but for the visual test it is around 3.}\label{fig:polypower}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/heterpower-1}
-
-}
-
-\caption{...}\label{fig:heterpower}
-\end{figure}
-
-tick mark on x. dashed line too strong. use solid line (grey) in light
-color.
-
 \hypertarget{lineup-selection}{%

### 2022-12-13 paper_comparison/paper_comparison/paper_comparison.Rmd
clarify RESET test power

@@
 Figure \ref{fig:polypower} shows the estimated power of visual test on lineups produced by the non-linearity model with $X_{raw} \sim U(-1,1)$, against the natural logarithm of the effect $log_e(E)$ with comparison to the power of an exact test - RESET test, and two other residual-based conventional tests commonly used in regression diagnostics but for testing heteroskedasticity and non-normality respectively.
+For the RESET test, we included different power of fitted values as proxies.
+According to [@ramsey_tests_1969], there is no general rule for the power of the fitted values needed by the RESET test, but it found power up to four is usually sufficient.
+Thus, we followed this guidline to conduct the RESET test.
+
 At the bottom of the figure \ref{fig:polypower}, there are a sequence of example residuals plots with increasing levels of $log_e(E)$.

### 2022-12-13 paper_comparison/paper_comparison/paper_comparison.tex
clarify RESET test power

@@
 regression diagnostics but for testing heteroskedasticity and
-non-normality respectively. At the bottom of the figure
-\ref{fig:polypower}, there are a sequence of example residuals plots
-with increasing levels of \(log_e(E)\). Readers can evaluate them from
-left to right and determine at which level the departure from a good
-residual plot becomes detectable.
+non-normality respectively. For the RESET test, we included different
+power of fitted values as proxies. According to
+\citep{ramsey_tests_1969}, there is no general rule for the power of the
+fitted values needed by the RESET test, but it found power up to four is
+usually sufficient. Thus, we followed this guidline to conduct the RESET
+test.
+
+At the bottom of the figure \ref{fig:polypower}, there are a sequence of
+example residuals plots with increasing levels of \(log_e(E)\). Readers
+can evaluate them from left to right and determine at which level the
+departure from a good residual plot becomes detectable.
 

### 2022-12-14 paper_comparison/paper_comparison/paper_comparison.Rmd
section headings

@@
 ```{r}
-# OOP supports needed by `visuage`
+# OOP supports needed by `visage`
 # remotes::install_github("TengMCing/bandicoot")
@@
 
-## Visual testing for departures
+## Visual test procedure based on lineups
 
@@
 
-### Calculating $p$-values
+### Calculating $p$-values for the visual test
 
@@
 
-### Power calculation
+### Power of a visual test
 
@@
 
-# Experiment design
+# Experimental design
 
@@
 
-## Simulating departures
+## Simulating departures from good residuals
 
@@
 
-## Experiment setup
+## Experimental setup
 
@@
 
-## Power comparison
+## Comparison of the power of different tests
 
@@
 
-## P-value comparison (remove?) (closer to the topic of the paper) (put it after power)
+## Comparison of rejection rates based on $p$-values
 
@@
 
-## Lineup selection
+## Effect of the distribution of the independent variable
 

### 2022-12-14 paper_comparison/paper_comparison/paper_comparison.tex
section headings

@@
 
-\hypertarget{visual-testing-for-departures}{%
-\subsection{Visual testing for
-departures}\label{visual-testing-for-departures}}
+\hypertarget{visual-test-procedure-based-on-lineups}{%
+\subsection{Visual test procedure based on
+lineups}\label{visual-test-procedure-based-on-lineups}}
 
@@
 
-\hypertarget{calculating-p-values}{%
-\subsubsection{\texorpdfstring{Calculating
-\(p\)-values}{Calculating p-values}}\label{calculating-p-values}}
+\hypertarget{calculating-p-values-for-the-visual-test}{%
+\subsubsection{\texorpdfstring{Calculating \(p\)-values for the visual
+test}{Calculating p-values for the visual test}}\label{calculating-p-values-for-the-visual-test}}
 
@@
 
-\hypertarget{power-calculation}{%
-\subsubsection{Power calculation}\label{power-calculation}}
+\hypertarget{power-of-a-visual-test}{%
+\subsubsection{Power of a visual test}\label{power-of-a-visual-test}}
 
@@
 
-\hypertarget{experiment-design}{%
-\section{Experiment design}\label{experiment-design}}
+\hypertarget{experimental-design}{%
+\section{Experimental design}\label{experimental-design}}
 
@@
 
-\hypertarget{simulating-departures}{%
-\subsection{Simulating departures}\label{simulating-departures}}
+\hypertarget{simulating-departures-from-good-residuals}{%
+\subsection{Simulating departures from good
+residuals}\label{simulating-departures-from-good-residuals}}
 
@@
 
-\hypertarget{experiment-setup}{%
-\subsection{Experiment setup}\label{experiment-setup}}
+\hypertarget{experimental-setup}{%
+\subsection{Experimental setup}\label{experimental-setup}}
 
@@
 
-\hypertarget{power-comparison}{%
-\subsection{Power comparison}\label{power-comparison}}
+\hypertarget{comparison-of-the-power-of-different-tests}{%
+\subsection{Comparison of the power of different
+tests}\label{comparison-of-the-power-of-different-tests}}
 
@@
 
-\hypertarget{p-value-comparison-remove-closer-to-the-topic-of-the-paper-put-it-after-power}{%
-\subsection{P-value comparison (remove?) (closer to the topic of the
-paper) (put it after
-power)}\label{p-value-comparison-remove-closer-to-the-topic-of-the-paper-put-it-after-power}}
+\hypertarget{comparison-of-rejection-rates-based-on-p-values}{%
+\subsection{\texorpdfstring{Comparison of rejection rates based on
+\(p\)-values}{Comparison of rejection rates based on p-values}}\label{comparison-of-rejection-rates-based-on-p-values}}
 
@@
 
-\hypertarget{lineup-selection}{%
-\subsection{Lineup selection}\label{lineup-selection}}
+\hypertarget{effect-of-the-distribution-of-the-independent-variable}{%
+\subsection{Effect of the distribution of the independent
+variable}\label{effect-of-the-distribution-of-the-independent-variable}}
 

### 2022-12-20 paper_comparison/paper_comparison/paper_comparison.Rmd
title change
title tweak
handle non-normality

--- 2022-12-13 paper_comparison/paper_comparison/paper_comparison.Rmd
@@
 For a consistent comparison, the residual plot of this departure is still presented in plot D of Figure \ref{fig:residual-plot-common-departures}.
-But since the scatter plot is not optimal for this task, non-normality will not be the focus of this paper.
+However, since the scatter plot is not suitable for this task, non-normality will not be the focus of this paper.
 Besides, it is important to note that not all regression models assume normality for the error term, but a certain amount do including the classical normal linear regression model.
@@
 @vanderplas2021statistical suggest that $\alpha$ can be estimated using MLE with Equation \ref{eq:c-interesting-expectation}.
-But for precise estimate of $\alpha$, additional responses to Rorschach lineups, which is a type of lineup consists of only null plots, are required.
+But for precise estimate of $\alpha$, additional responses to Rorschach lineups, which is a type of lineup that consists of
+
+plots constructed from the same null data generating mechnisam.
 
@@
 Two types of departures, namely non-linearity and heteroskedasticity, were considered with the corresponding data generating process being designed for experiment I and II.
-As mentioned earlier, residual plot is not suitable for revealing non-normality, thus it was not considered in our experiments.
 
@@
 ```
+
+(handbook of mathathetical function, https://aapt.scitation.org/doi/10.1119/1.15378, cite the package)
+
+(hermite poly is created by ... in 18..,)
 
@@
 Note that we generated the lineups using the same simulation mechanism discussed in Section \ref{sampling-from-the-null-distribution} rather than from a null model.
-We overlooked this detail until the data had been collected.
+
+(it has zero effect size, an alternative to Rorschach lineups)
+
+(Appendix: vary the alpha a little bit and see how the curves change, sensitivity analysis)
+
 Thus, we used another valid method suggested in vanderplas2021statistical which was designed for typical lineups to estimate $\alpha$ correctly.
 However, by not using Rorschach lineups, we did lose some accuracy in the estimation.
+(questionable)
+
 We recommend future studies to generate Rorschach lineups strictly following the definition given by buja_statistical_2009.
@@
 
+(discuss choice of tests, mention pkg `lmtest`, `skedastic`)
+
 Figure \ref{fig:polypower} shows the estimated power of visual test on lineups produced by the non-linearity model with $X_{raw} \sim U(-1,1)$, against the natural logarithm of the effect $log_e(E)$ with comparison to the power of an exact test - RESET test, and two other residual-based conventional tests commonly used in regression diagnostics but for testing heteroskedasticity and non-normality respectively.
@@
 For the non-linearity model, the power of all four tests increases as the effect gets larger.
-The power curve of F-test climbs aggressively from 25% to around 90% as $log_e(E)$ increases from 0 to 2, while power of other tests respond inactively to the change of effect and remain lower than 25% throughout the period, showing that F-test is way more sensitive to the type of model defects that being considered.
+The power curve of RESET test climbs aggressively from 13% to around 70% as $log_e(E)$ increases from 0 to 2, while power of other tests respond inactively to the change of effect and remain lower than 25% throughout the period, showing that F-test is way more sensitive to the type of model defects that being considered.
 Meanwhile, no noticeable visual features can be spotted from the example residuals plots.
@@
              aes(log_effect_size, reject_dirichlet, size = factor(num_response)), alpha = 0.3)  +
-  geom_segment(data = data.frame(xi = es), aes(x = xi, xend = xi, y = 0, yend = 1), alpha = 0.3, col = "grey") +
+  geom_segment(data = data.frame(xi = es), aes(x = xi, xend = xi, y = -Inf, yend = Inf), alpha = 0.3, col = "grey") +
   theme_light(base_size = 5) +
@@
 
+(top: curves)
+(middle: dots)
+(bottom: vertical lines/grid lines)
+
 ```{r heterpower, fig.height = 3, fig.width = 5, fig.cap="Comparison of power between different tests for heteroskedasticity patterns.
@@
 ## P-value comparison (remove?) (closer to the topic of the paper) (put it after power)
+
+(style similar to mosaic plot)
+
+(lineups detected by visual test but not by conventional tests)
 
@@
 
-## Lineup selection
-
-```{r cache = TRUE}
-vi_survey_with_plot <- vi_survey %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  left_join(map_df(1:length(vi_lineup), 
-    ~vi_lineup[[.x]]$data %>% 
-      group_by(k) %>% 
-      summarise(plot_p_value = first(p_value)) %>%
-      arrange(k) %>%
-      mutate(unique_lineup_id = names(vi_lineup)[.x]) %>%
-      mutate(plot_p_value_rank = rank(plot_p_value, ties.method = "min"))) %>%
-  pivot_wider(values_from = c(plot_p_value, plot_p_value_rank), names_from = k, names_prefix = "plot_")) %>%
-  (function(x) {
-      for (i in 1:20)
-        x[[paste0("plot_sel_plot_", i)]] <-
-          grepl(paste0("_", i, "_"), x$selection) |
-          grepl(paste0("^", i, "_"), x$selection) |
-          grepl(paste0("_", i, "$"), x$selection) |
-          grepl(paste0("^", i, "$"), x$selection)
-      x}
-    ) %>%
-  pivot_longer(plot_p_value_plot_1:plot_sel_plot_20, 
-               names_to = c(".value", "plot"), 
-               names_pattern = "(.*)_plot_(.+)") 
-```
-
-```{r plotsel, fig.cap = "A large amount of subjects select the residuals plot with the lowest p-value. But greater than 50 percent of subjects evenly select other plots."}
-vi_survey_with_plot %>%
-  mutate(weighted_sel = ifelse(num_selection == 0, 0.05, plot_sel/num_selection)) %>%
-  group_by(type) %>%
-  mutate(n_eval = n()/20) %>%
-  ungroup() %>%
-  group_by(type, plot_p_value_rank) %>%
-  summarise(weighted_sel = sum(weighted_sel), n_eval = first(n_eval)) %>%
-  ggplot() +
-  geom_col(aes(plot_p_value_rank, n_eval), fill = "grey", alpha = 0.3) +
-  geom_col(aes(plot_p_value_rank, weighted_sel)) +
-  facet_wrap(~type) +
-  theme_light() +
-  ylab("Weighted number of selections") +
-  xlab("Plot p-value rank")
-```
-
-```{r dataplotsel, fig.cap = "..."}
-vi_survey_with_plot %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(answer == plot) %>%
-  group_by(unique_lineup_id) %>%
-  summarise(plot_p_value_rank = first(plot_p_value_rank),
-            weighted_detect = sum(weighted_detect),
-            n_eval = n()) %>%
-  group_by(plot_p_value_rank) %>%
-  summarise(weighted_detect = sum(weighted_detect),
-            n_eval = sum(n_eval)) %>%
-  mutate(plot_p_value_rank_g5 = plot_p_value_rank >= 5) %>%
-  group_by(plot_p_value_rank_g5) %>%
-  mutate(n_eval = ifelse(plot_p_value_rank_g5, sum(n_eval), n_eval)) %>%
-  mutate(weighted_detect = ifelse(plot_p_value_rank_g5, sum(weighted_detect), weighted_detect)) %>%
-  ungroup() %>%
-  filter(plot_p_value_rank <= 5) %>%
-  mutate(plot_p_value_rank = ifelse(plot_p_value_rank == 5, "5+", as.character(plot_p_value_rank))) %>%
-  mutate(plot_p_value_rank = factor(plot_p_value_rank, levels = c(as.character(1:5), "5+"))) %>%
-  ggplot() +
-  geom_col(aes(plot_p_value_rank, n_eval)) +
-  geom_col(aes(plot_p_value_rank, weighted_detect, fill = "weighted number of detections")) +
-  xlab("Data plot p-value rank") +
-  ylab("Number of Lineups") +
-  theme(legend.position = "bottom") +
-  labs(fill = "")
-```
-
-```{r dadtaplotselfacet, fig.cap = "...", fig.height=12, fig.width=16}
-vi_survey_with_plot %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(type == "polynomial") %>%
-  left_join(vi_survey_with_plot %>%
-              filter(!attention_check) %>%
-              filter(!null_lineup) %>%
-              filter(answer == plot) %>%
-              group_by(unique_lineup_id) %>%
-              summarise(answer_rank = first(plot_p_value_rank))) %>%
-  mutate(weighted_sel = ifelse(num_selection == 0, 0.05, plot_sel/num_selection)) %>%
-  group_by(plot_p_value_rank, answer_rank) %>%
-  summarise(weighted_sel = sum(weighted_sel)) %>%
-  mutate(answer_rank_g5 = answer_rank >= 5) %>%
-  group_by(plot_p_value_rank, answer_rank_g5) %>%
-  mutate(weighted_sel = ifelse(answer_rank_g5, sum(weighted_sel), weighted_sel)) %>%
-  ungroup() %>%
-  filter(answer_rank <= 5) %>%
-  mutate(answer_rank = ifelse(answer_rank == 5, "5+", as.character(answer_rank))) %>%
-  mutate(answer_rank = factor(answer_rank, levels = c(as.character(1:5), "5+"))) %>%
-  ggplot() +
-  geom_col(aes(factor(plot_p_value_rank), weighted_sel)) +
-  geom_vline(data = data.frame(answer_rank = c("1", "2", "3", "4"), x = 1:4),
-             aes(xintercept = x),
-             col = "red") +
-  facet_wrap(~answer_rank, scales = "free_y") +
-  xlab("Plot p-value rank") +
-  ylab("Weighted number of selections")
-```
-
-```{r eval = FALSE}
-vi_survey_with_plot %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(answer == plot) %>%
-  group_by(unique_lineup_id) %>%
-  summarise(plot_p_value_rank = first(plot_p_value_rank),
-            weighted_detect = sum(weighted_detect),
-            n_eval = n(),
-            reject = first(conventional_p_value < 0.05)) %>%
-  group_by(plot_p_value_rank, reject) %>%
-  summarise(weighted_detect = sum(weighted_detect),
-            n_eval = sum(n_eval)) %>%
-  ggplot() +
-  geom_col(aes(factor(plot_p_value_rank), n_eval)) +
-  geom_col(aes(factor(plot_p_value_rank), weighted_detect, fill = "weighted number of detections")) +
-  xlab("Data plot p-value rank") +
-  ylab("Number of Lineups") +
-  theme(legend.position = "bottom") +
-  labs(fill = "") +
-  facet_wrap(~reject)
-```
-
-```{r eval = FALSE}
-# data plot not always has the lowest conventional p-value
-vi_survey_with_plot %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(answer == plot) %>%
-  ggplot() +
-  geom_bar(aes(factor(plot_p_value_rank)))
-
-# according to the rank of the data plot p-value, how many people select it 
-vi_survey_with_plot %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(plot_sel) %>%
-  left_join(vi_survey_with_plot %>%
-              filter(!attention_check) %>%
-              filter(!null_lineup) %>%
-              filter(answer == plot) %>%
-              group_by(unique_lineup_id) %>%
-              summarise(answer_rank = first(plot_p_value_rank))) %>%
-  ggplot() +
-  geom_bar(aes(factor(plot_p_value_rank))) +
-  geom_vline(data = data.frame(answer_rank = 1:20, x = 1:20),
-             aes(xintercept = x),
-             col = "red") +
-  facet_wrap(~answer_rank, scales = "free_y")
-
-# For different rank, percentage of subject select it
-vi_survey_with_plot %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(answer == plot) %>%
-  group_by(unique_lineup_id, set) %>%
-  summarise(plot_p_value_rank = first(plot_p_value_rank),
-            detect = first(detect),
-            type = first(type)) %>%
-  ggplot() +
-  geom_bar(aes(factor(plot_p_value_rank), fill = detect), position = "fill") +
-  geom_hline(yintercept = 0.05) +
-  xlab("Data plot p-value rank") +
-  ylab("Propotion of subject detecting the data plot")
-
-vi_survey_with_plot %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(answer == plot) %>%
-  group_by(unique_lineup_id, set) %>%
-  summarise(plot_p_value_rank = first(plot_p_value_rank),
-            prop_detect = first(prop_detect)) %>%
-  ggplot() +
-  geom_boxplot(aes(factor(plot_p_value_rank), prop_detect))
-
-# vi_survey_with_plot %>%
-#   filter(!attention_check) %>%
-#   filter(!null_lineup) %>%
-#   filter(answer == plot) %>%
-#   group_by(unique_lineup_id, set) %>%
-#   summarise(plot_p_value_rank = first(plot_p_value_rank),
-#             conv_detect = first(conventional_p_value < 0.05)) %>%
-#   ggplot() +
-#   geom_bar(aes(factor(plot_p_value_rank), fill = conv_detect), position = "fill") +
-#   xlab("Data plot p-value rank") +
-#   ylab("Propotion of subject detecting the data plot")
-```
-
-# invesgating interesting lineups
+<!-- ## Lineup selection -->
+
+<!-- ```{r cache = TRUE} -->
+<!-- vi_survey_with_plot <- vi_survey %>% -->
+<!--   filter(!attention_check) %>% -->
+<!--   filter(!null_lineup) %>% -->
+<!--   left_join(map_df(1:length(vi_lineup),  -->
+<!--     ~vi_lineup[[.x]]$data %>%  -->
+<!--       group_by(k) %>%  -->
+<!--       summarise(plot_p_value = first(p_value)) %>% -->
+<!--       arrange(k) %>% -->
+<!--       mutate(unique_lineup_id = names(vi_lineup)[.x]) %>% -->
+<!--       mutate(plot_p_value_rank = rank(plot_p_value, ties.method = "min"))) %>% -->
+<!--   pivot_wider(values_from = c(plot_p_value, plot_p_value_rank), names_from = k, names_prefix = "plot_")) %>% -->
+<!--   (function(x) { -->
+<!--       for (i in 1:20) -->
+<!--         x[[paste0("plot_sel_plot_", i)]] <- -->
+<!--           grepl(paste0("_", i, "_"), x$selection) | -->
+<!--           grepl(paste0("^", i, "_"), x$selection) | -->
+<!--           grepl(paste0("_", i, "$"), x$selection) | -->
+<!--           grepl(paste0("^", i, "$"), x$selection) -->
+<!--       x} -->
+<!--     ) %>% -->
+<!--   pivot_longer(plot_p_value_plot_1:plot_sel_plot_20,  -->
+<!--                names_to = c(".value", "plot"),  -->
+<!--                names_pattern = "(.*)_plot_(.+)")  -->
+<!-- ``` -->
+
+<!-- ```{r plotsel, fig.cap = "A large amount of subjects select the residuals plot with the lowest p-value. But greater than 50 percent of subjects evenly select other plots."} -->
+<!-- vi_survey_with_plot %>% -->
+<!--   mutate(weighted_sel = ifelse(num_selection == 0, 0.05, plot_sel/num_selection)) %>% -->
+<!--   group_by(type) %>% -->
+<!--   mutate(n_eval = n()/20) %>% -->
+<!--   ungroup() %>% -->
+<!--   group_by(type, plot_p_value_rank) %>% -->
+<!--   summarise(weighted_sel = sum(weighted_sel), n_eval = first(n_eval)) %>% -->
+<!--   ggplot() + -->
+<!--   geom_col(aes(plot_p_value_rank, n_eval), fill = "grey", alpha = 0.3) + -->
+<!--   geom_col(aes(plot_p_value_rank, weighted_sel)) + -->
+<!--   facet_wrap(~type) + -->
+<!--   theme_light() + -->
+<!--   ylab("Weighted number of selections") + -->
+<!--   xlab("Plot p-value rank") -->
+<!-- ``` -->
+
+<!-- ```{r dataplotsel, fig.cap = "..."} -->
+<!-- vi_survey_with_plot %>% -->
+<!--   filter(!attention_check) %>% -->
+<!--   filter(!null_lineup) %>% -->
+<!--   filter(answer == plot) %>% -->
+<!--   group_by(unique_lineup_id) %>% -->
+<!--   summarise(plot_p_value_rank = first(plot_p_value_rank), -->
+<!--             weighted_detect = sum(weighted_detect), -->
+<!--             n_eval = n()) %>% -->
+<!--   group_by(plot_p_value_rank) %>% -->
+<!--   summarise(weighted_detect = sum(weighted_detect), -->
+<!--             n_eval = sum(n_eval)) %>% -->
+<!--   mutate(plot_p_value_rank_g5 = plot_p_value_rank >= 5) %>% -->
+<!--   group_by(plot_p_value_rank_g5) %>% -->
+<!--   mutate(n_eval = ifelse(plot_p_value_rank_g5, sum(n_eval), n_eval)) %>% -->
+<!--   mutate(weighted_detect = ifelse(plot_p_value_rank_g5, sum(weighted_detect), weighted_detect)) %>% -->
+<!--   ungroup() %>% -->
+<!--   filter(plot_p_value_rank <= 5) %>% -->
+<!--   mutate(plot_p_value_rank = ifelse(plot_p_value_rank == 5, "5+", as.character(plot_p_value_rank))) %>% -->
+<!--   mutate(plot_p_value_rank = factor(plot_p_value_rank, levels = c(as.character(1:5), "5+"))) %>% -->
+<!--   ggplot() + -->
+<!--   geom_col(aes(plot_p_value_rank, n_eval)) + -->
+<!--   geom_col(aes(plot_p_value_rank, weighted_detect, fill = "weighted number of detections")) + -->
+<!--   xlab("Data plot p-value rank") + -->
+<!--   ylab("Number of Lineups") + -->
+<!--   theme(legend.position = "bottom") + -->
+<!--   labs(fill = "") -->
+<!-- ``` -->
+
+<!-- ```{r dadtaplotselfacet, fig.cap = "...", fig.height=12, fig.width=16} -->
+<!-- vi_survey_with_plot %>% -->
+<!--   filter(!attention_check) %>% -->
+<!--   filter(!null_lineup) %>% -->
+<!--   filter(type == "polynomial") %>% -->
+<!--   left_join(vi_survey_with_plot %>% -->
+<!--               filter(!attention_check) %>% -->
+<!--               filter(!null_lineup) %>% -->
+<!--               filter(answer == plot) %>% -->
+<!--               group_by(unique_lineup_id) %>% -->
+<!--               summarise(answer_rank = first(plot_p_value_rank))) %>% -->
+<!--   mutate(weighted_sel = ifelse(num_selection == 0, 0.05, plot_sel/num_selection)) %>% -->
+<!--   group_by(plot_p_value_rank, answer_rank) %>% -->
+<!--   summarise(weighted_sel = sum(weighted_sel)) %>% -->
+<!--   mutate(answer_rank_g5 = answer_rank >= 5) %>% -->
+<!--   group_by(plot_p_value_rank, answer_rank_g5) %>% -->
+<!--   mutate(weighted_sel = ifelse(answer_rank_g5, sum(weighted_sel), weighted_sel)) %>% -->
+<!--   ungroup() %>% -->
+<!--   filter(answer_rank <= 5) %>% -->
+<!--   mutate(answer_rank = ifelse(answer_rank == 5, "5+", as.character(answer_rank))) %>% -->
+<!--   mutate(answer_rank = factor(answer_rank, levels = c(as.character(1:5), "5+"))) %>% -->
+<!--   ggplot() + -->
+<!--   geom_col(aes(factor(plot_p_value_rank), weighted_sel)) + -->
+<!--   geom_vline(data = data.frame(answer_rank = c("1", "2", "3", "4"), x = 1:4), -->
+<!--              aes(xintercept = x), -->
+<!--              col = "red") + -->
+<!--   facet_wrap(~answer_rank, scales = "free_y") + -->
+<!--   xlab("Plot p-value rank") + -->
+<!--   ylab("Weighted number of selections") -->
+<!-- ``` -->
+
+<!-- ```{r eval = FALSE} -->
+<!-- vi_survey_with_plot %>% -->
+<!--   filter(!attention_check) %>% -->
+<!--   filter(!null_lineup) %>% -->
+<!--   filter(answer == plot) %>% -->
+<!--   group_by(unique_lineup_id) %>% -->
+<!--   summarise(plot_p_value_rank = first(plot_p_value_rank), -->
+<!--             weighted_detect = sum(weighted_detect), -->
+<!--             n_eval = n(), -->
+<!--             reject = first(conventional_p_value < 0.05)) %>% -->
+<!--   group_by(plot_p_value_rank, reject) %>% -->
+<!--   summarise(weighted_detect = sum(weighted_detect), -->
+<!--             n_eval = sum(n_eval)) %>% -->
+<!--   ggplot() + -->
+<!--   geom_col(aes(factor(plot_p_value_rank), n_eval)) + -->
+<!--   geom_col(aes(factor(plot_p_value_rank), weighted_detect, fill = "weighted number of detections")) + -->
+<!--   xlab("Data plot p-value rank") + -->
+<!--   ylab("Number of Lineups") + -->
+<!--   theme(legend.position = "bottom") + -->
+<!--   labs(fill = "") + -->
+<!--   facet_wrap(~reject) -->
+<!-- ``` -->
+
+<!-- ```{r eval = FALSE} -->
+<!-- # data plot not always has the lowest conventional p-value -->
+<!-- vi_survey_with_plot %>% -->
+<!--   filter(!attention_check) %>% -->
+<!--   filter(!null_lineup) %>% -->
+<!--   filter(answer == plot) %>% -->
+<!--   ggplot() + -->
+<!--   geom_bar(aes(factor(plot_p_value_rank))) -->
+
+<!-- # according to the rank of the data plot p-value, how many people select it  -->
+<!-- vi_survey_with_plot %>% -->
+<!--   filter(!attention_check) %>% -->
+<!--   filter(!null_lineup) %>% -->
+<!--   filter(plot_sel) %>% -->
+<!--   left_join(vi_survey_with_plot %>% -->
+<!--               filter(!attention_check) %>% -->
+<!--               filter(!null_lineup) %>% -->
+<!--               filter(answer == plot) %>% -->
+<!--               group_by(unique_lineup_id) %>% -->
+<!--               summarise(answer_rank = first(plot_p_value_rank))) %>% -->
+<!--   ggplot() + -->
+<!--   geom_bar(aes(factor(plot_p_value_rank))) + -->
+<!--   geom_vline(data = data.frame(answer_rank = 1:20, x = 1:20), -->
+<!--              aes(xintercept = x), -->
+<!--              col = "red") + -->
+<!--   facet_wrap(~answer_rank, scales = "free_y") -->
+
+<!-- # For different rank, percentage of subject select it -->
+<!-- vi_survey_with_plot %>% -->
+<!--   filter(!attention_check) %>% -->
+<!--   filter(!null_lineup) %>% -->
+<!--   filter(answer == plot) %>% -->
+<!--   group_by(unique_lineup_id, set) %>% -->
+<!--   summarise(plot_p_value_rank = first(plot_p_value_rank), -->
+<!--             detect = first(detect), -->
+<!--             type = first(type)) %>% -->
+<!--   ggplot() + -->
+<!--   geom_bar(aes(factor(plot_p_value_rank), fill = detect), position = "fill") + -->
+<!--   geom_hline(yintercept = 0.05) + -->
+<!--   xlab("Data plot p-value rank") + -->
+<!--   ylab("Propotion of subject detecting the data plot") -->
+
+<!-- vi_survey_with_plot %>% -->
+<!--   filter(!attention_check) %>% -->
+<!--   filter(!null_lineup) %>% -->
+<!--   filter(answer == plot) %>% -->
+<!--   group_by(unique_lineup_id, set) %>% -->
+<!--   summarise(plot_p_value_rank = first(plot_p_value_rank), -->
+<!--             prop_detect = first(prop_detect)) %>% -->
+<!--   ggplot() + -->
+<!--   geom_boxplot(aes(factor(plot_p_value_rank), prop_detect)) -->
+
+<!-- # vi_survey_with_plot %>% -->
+<!-- #   filter(!attention_check) %>% -->
+<!-- #   filter(!null_lineup) %>% -->
+<!-- #   filter(answer == plot) %>% -->
+<!-- #   group_by(unique_lineup_id, set) %>% -->
+<!-- #   summarise(plot_p_value_rank = first(plot_p_value_rank), -->
+<!-- #             conv_detect = first(conventional_p_value < 0.05)) %>% -->
+<!-- #   ggplot() + -->
+<!-- #   geom_bar(aes(factor(plot_p_value_rank), fill = conv_detect), position = "fill") + -->
+<!-- #   xlab("Data plot p-value rank") + -->
+<!-- #   ylab("Propotion of subject detecting the data plot") -->
+<!-- ``` -->
 
@@
 # appendix (contains all the details)
+
+### A collection interesting lineups (unusual results)
+
+why unusual? what is the possible explanations?
+
+# targe journal
+
+JRSSB: Journal of the Royal Statistical Society Series B (Statistical Methodology)
+Deadline: Jan 1, Apr 1
+
+Reading: style of writing, author guideline (https://rss.onlinelibrary.wiley.com/hub/journal/14679868/author-guidelines)
+
+JCGS
+JDSS

### 2022-12-20 paper_comparison/paper_comparison/paper_comparison.Rmd

@@
 title: |
-  Why aren't significance tests commonly used for linear regression diagnostics?
+  Why is it that statistical tests for residuals are not widely used? An explanation using visual inference.
 type: ARTICLE TEMPLATE
@@
 
-<!-- ## Lineup selection -->
+## Effect of the distribution of the independent variable
 

### 2022-12-20 paper_comparison/paper_comparison/paper_comparison.Rmd
Rearrange ggplot layers for power cmp

@@
   filter(!attention_check) %>%
-  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
+  mutate(reject = as.numeric(p_value <= 0.05)) %>%
   mutate(log_effect_size = log(effect_size)) %>%
   ggplot() +
-  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_dirichlet, times = 500), 
-              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "Visual test"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03) +
+  geom_point(data = ~.x %>% 
+             group_by(log_effect_size, reject) %>% 
+             summarise(num_response = n()),
+           aes(log_effect_size, reject, size = factor(num_response)), alpha = 0.3)  +
   geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
@@
               se = FALSE) +
-  geom_point(data = ~.x %>% 
-               group_by(log_effect_size, reject_dirichlet) %>% 
-               summarise(num_response = n()),
-             aes(log_effect_size, reject_dirichlet, size = factor(num_response)), alpha = 0.3)  +
-  geom_segment(data = data.frame(xi = es), aes(x = xi, xend = xi, y = -Inf, yend = Inf), alpha = 0.3, col = "grey") +
+  stat_smooth(geom = "line", 
+              data = ~boot_y(.x, reject, times = 500), 
+              aes(log_effect_size, reject, group = boot_id, col = "Visual test"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, 
+              alpha = 0.03) +
+  geom_smooth(aes(log_effect_size, reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
   theme_light(base_size = 5) +
+  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.5)) +
   scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
@@
   ylab("Power") +
-  labs(col = "", size = "#lineups")
+  labs(col = "", size = "# lineups")
 
@@
   filter(!attention_check) %>%
-  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
+  mutate(reject = as.numeric(p_value <= 0.05)) %>%
   mutate(log_effect_size = log(effect_size)) %>%
   ggplot() +
-  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject_dirichlet, times = 500), 
-              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "Visual test"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03) +
+  geom_point(data = ~.x %>% 
+               group_by(log_effect_size, reject) %>% 
+               summarise(num_response = n()),
+             aes(log_effect_size, reject, size = factor(num_response)), alpha = 0.3) +
   geom_smooth(data = mutate(heter_conv_sim, log_effect_size = log(effect_size)) %>%

### 2022-12-20 paper_comparison/paper_comparison/paper_comparison.tex
title change
Rearrange ggplot layers for power cmp

@@
 
-\title{Why aren't significance tests commonly used for linear regression
-diagnostics?}
+\title{Why is it that statistical tests for residuals are not widely
+used? An explanation using visual inference.}
 
@@
 comparison, the residual plot of this departure is still presented in
-plot D of Figure \ref{fig:residual-plot-common-departures}. But since
-the scatter plot is not optimal for this task, non-normality will not be
-the focus of this paper. Besides, it is important to note that not all
-regression models assume normality for the error term, but a certain
-amount do including the classical normal linear regression model. In the
-case that the normality assumption is violated, it is expected to
-observe data points do not center around the horizontal axis and there
-is an uneven distribution of the number points at both below and above
-the horizontal axis. For example, given a skewed error distribution,
-fewer data points and more outliers are on one side of the horizontal
-axis as shown in plot D of Figure
+plot D of Figure \ref{fig:residual-plot-common-departures}. However,
+since the scatter plot is not suitable for this task, non-normality will
+not be the focus of this paper. Besides, it is important to note that
+not all regression models assume normality for the error term, but a
+certain amount do including the classical normal linear regression
+model. In the case that the normality assumption is violated, it is
+expected to observe data points do not center around the horizontal axis
+and there is an uneven distribution of the number points at both below
+and above the horizontal axis. For example, given a skewed error
+distribution, fewer data points and more outliers are on one side of the
+horizontal axis as shown in plot D of Figure
 \ref{fig:residual-plot-common-departures}.
@@
 But for precise estimate of \(\alpha\), additional responses to
-Rorschach lineups, which is a type of lineup consists of only null
-plots, are required.
+Rorschach lineups, which is a type of lineup that consists of
+
+plots constructed from the same null data generating mechnisam.
 
@@
 were considered with the corresponding data generating process being
-designed for experiment I and II. As mentioned earlier, residual plot is
-not suitable for revealing non-normality, thus it was not considered in
-our experiments.
+designed for experiment I and II.
 
@@
 \end{figure}
+
+(handbook of mathathetical function,
+\url{https://aapt.scitation.org/doi/10.1119/1.15378}, cite the package)
+
+(hermite poly is created by \ldots{} in 18..,)
 
@@
 \ref{sampling-from-the-null-distribution} rather than from a null model.
-We overlooked this detail until the data had been collected. Thus, we
-used another valid method suggested in vanderplas2021statistical which
-was designed for typical lineups to estimate \(\alpha\) correctly.
-However, by not using Rorschach lineups, we did lose some accuracy in
-the estimation. We recommend future studies to generate Rorschach
-lineups strictly following the definition given by
-buja\_statistical\_2009.
+
+(it has zero effect size, an alternative to Rorschach lineups)
+
+(Appendix: vary the alpha a little bit and see how the curves change,
+sensitivity analysis)
+
+Thus, we used another valid method suggested in
+vanderplas2021statistical which was designed for typical lineups to
+estimate \(\alpha\) correctly. However, by not using Rorschach lineups,
+we did lose some accuracy in the estimation. (questionable)
+
+We recommend future studies to generate Rorschach lineups strictly
+following the definition given by buja\_statistical\_2009.
 
@@
 
+(discuss choice of tests, mention pkg \texttt{lmtest},
+\texttt{skedastic})
+
 Figure \ref{fig:polypower} shows the estimated power of visual test on
@@
 For the non-linearity model, the power of all four tests increases as
-the effect gets larger. The power curve of F-test climbs aggressively
-from 25\% to around 90\% as \(log_e(E)\) increases from 0 to 2, while
-power of other tests respond inactively to the change of effect and
-remain lower than 25\% throughout the period, showing that F-test is way
-more sensitive to the type of model defects that being considered.
-Meanwhile, no noticeable visual features can be spotted from the example
-residuals plots.
+the effect gets larger. The power curve of RESET test climbs
+aggressively from 13\% to around 70\% as \(log_e(E)\) increases from 0
+to 2, while power of other tests respond inactively to the change of
+effect and remain lower than 25\% throughout the period, showing that
+F-test is way more sensitive to the type of model defects that being
+considered. Meanwhile, no noticeable visual features can be spotted from
+the example residuals plots.
 
@@
 
+(top: curves) (middle: dots) (bottom: vertical lines/grid lines)
+
 \begin{figure}
@@
 \(p\)-values}{Comparison of rejection rates based on p-values}}\label{comparison-of-rejection-rates-based-on-p-values}}
+
+(style similar to mosaic plot)
+
+(lineups detected by visual test but not by conventional tests)
 
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/plotsel-1}
-
-}
-
-\caption{A large amount of subjects select the residuals plot with the lowest p-value. But greater than 50 percent of subjects evenly select other plots.}\label{fig:plotsel}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/dataplotsel-1}
-
-}
-
-\caption{...}\label{fig:dataplotsel}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/dadtaplotselfacet-1}
-
-}
-
-\caption{...}\label{fig:dadtaplotselfacet}
-\end{figure}
-
-\hypertarget{invesgating-interesting-lineups}{%
-\section{invesgating interesting
-lineups}\label{invesgating-interesting-lineups}}
-
 \hypertarget{change-of-x_dist---power}{%
@@
 
+\hypertarget{appendix-contains-all-the-details}{%
+\section{appendix (contains all the
+details)}\label{appendix-contains-all-the-details}}
+
+\hypertarget{a-collection-interesting-lineups-unusual-results}{%
+\subsubsection{A collection interesting lineups (unusual
+results)}\label{a-collection-interesting-lineups-unusual-results}}
+
+why unusual? what is the possible explanations?
+
+\hypertarget{targe-journal}{%
+\section{targe journal}\label{targe-journal}}
+
+JRSSB: Journal of the Royal Statistical Society Series B (Statistical
+Methodology) Deadline: Jan 1, Apr 1
+
+Reading: style of writing, author guideline
+(\url{https://rss.onlinelibrary.wiley.com/hub/journal/14679868/author-guidelines})
+
+JCGS JDSS
+
 \bibliographystyle{tfcad}

### 2022-12-20 paper_comparison/paper_comparison/paper_comparison.Rmd
Cite mpoly and Hermite polynomials

@@
 
-(handbook of mathathetical function, https://aapt.scitation.org/doi/10.1119/1.15378, cite the package)
-
-(hermite poly is created by ... in 18..,)
-
-Experiment I is designed to study the ability of human subjects to detect the effect of a non-linear term $\boldsymbol{z}$ constructed using Hermite polynomials [unsure how to cite 1810 work] on random vector $\boldsymbol{x}$ formulated as
+<!-- (handbook of mathathetical function, https://aapt.scitation.org/doi/10.1119/1.15378, cite the package) -->
+
+<!-- (hermite poly is created by ... in 18..,) -->
+
+Experiment I is designed to study the ability of human subjects to detect the effect of a non-linear term $\boldsymbol{z}$ constructed using Hermite polynomials [@hermite1864nouveau] on random vector $\boldsymbol{x}$ formulated as
 
@@
 ```
+
+Hermite polynomials were initially defined by @de1820theorie, but named after Hermite [@hermite1864nouveau] because of the unrecognisable form of Laplace's work.
+When simulating $Z_{raw}$, the function `hermite` from the R package `mpoly` [@mpoly] was used to generate Hermite polynomials.
 

### 2022-12-20 paper_comparison/paper_comparison/paper_comparison.tex
Cite mpoly and Hermite polynomials

@@
 
-(handbook of mathathetical function,
-\url{https://aapt.scitation.org/doi/10.1119/1.15378}, cite the package)
-
-(hermite poly is created by \ldots{} in 18..,)
-
 Experiment I is designed to study the ability of human subjects to
 detect the effect of a non-linear term \(\boldsymbol{z}\) constructed
-using Hermite polynomials {[}unsure how to cite 1810 work{]} on random
-vector \(\boldsymbol{x}\) formulated as
+using Hermite polynomials \citep{hermite1864nouveau} on random vector
+\(\boldsymbol{x}\) formulated as
 
@@
 \end{equation}
+
+Hermite polynomials were initially defined by \citet{de1820theorie}, but
+named after Hermite \citep{hermite1864nouveau} because of the
+unrecognisable form of Laplace's work. When simulating \(Z_{raw}\), the
+function \texttt{hermite} from the R package \texttt{mpoly}
+\citep{mpoly} was used to generate Hermite polynomials.
 

### 2022-12-20 paper_comparison/paper_comparison/paper_comparison.Rmd
discuss choices of tests and mention pkgs

@@
 
-Figure \ref{fig:polypower} shows the estimated power of visual test on lineups produced by the non-linearity model with $X_{raw} \sim U(-1,1)$, against the natural logarithm of the effect $log_e(E)$ with comparison to the power of an exact test - RESET test, and two other residual-based conventional tests commonly used in regression diagnostics but for testing heteroskedasticity and non-normality respectively.
+Figure \ref{fig:polypower} shows the estimated power of visual test on lineups produced by the non-linearity model with $X_{raw} \sim U(-1,1)$, against the natural logarithm of the effect $log_e(E)$.
+
+As discussed in Section \ref{conventionally-testing-for-departures}, many conventionally tests are available for detecting residual departures.
+Implementation-wise, the built-in R package `stats` provides some commonly used residual-based tests, such as Shapiro-Wilk test.
+A more comprehensive collection of regression diagnostics tests can be found in the R package `lmtest` [@lmtest].
+In terms of heteroskedasticity diagnostics, the R package `skedastic` [@skedastic] collects and implements 25 existing conventional tests published since 1961.
+
+To compare the power of visual test and conventional tests, we picked RESET test (`resettest`) and Breusch-Pagan test (`bptest`) from the R package `lmtest`, and Shapiro-Wilk test (`shapiro.test`) from the built-in R package `stats`.
+These three tests can be used to detect non-linearity, heteroskedasticity and non-normality respectively.
+Their estimated power is shown in Figure \ref{fig:polypower}.
 For the RESET test, we included different power of fitted values as proxies.
 According to [@ramsey_tests_1969], there is no general rule for the power of the fitted values needed by the RESET test, but it found power up to four is usually sufficient.
-Thus, we followed this guidline to conduct the RESET test.
+Thus, we followed this guideline to conduct the RESET test.
+For the Breusch-Pagan test, both $\boldsymbol{x}$ and $\boldsymbol{x}^2$ were included in the auxiliary regression equation to [justify this choice?].
 

### 2022-12-20 paper_comparison/paper_comparison/paper_comparison.tex
discuss choices of tests and mention pkgs

@@
 \(X_{raw} \sim U(-1,1)\), against the natural logarithm of the effect
-\(log_e(E)\) with comparison to the power of an exact test - RESET test,
-and two other residual-based conventional tests commonly used in
-regression diagnostics but for testing heteroskedasticity and
-non-normality respectively. For the RESET test, we included different
-power of fitted values as proxies. According to
+\(log_e(E)\).
+
+As discussed in Section \ref{conventionally-testing-for-departures},
+many conventionally tests are available for detecting residual
+departures. Implementation-wise, the built-in R package \texttt{stats}
+provides some commonly used residual-based tests, such as Shapiro-Wilk
+test. A more comprehensive collection of regression diagnostics tests
+can be found in the R package \texttt{lmtest} \citep{lmtest}. In terms
+of heteroskedasticity diagnostics, the R package \texttt{skedastic}
+\citep{skedastic} collects and implements 25 existing conventional tests
+published since 1961.
+
+To compare the power of visual test and conventional tests, we picked
+RESET test (\texttt{resettest}) and Breusch-Pagan test (\texttt{bptest})
+from the R package \texttt{lmtest}, and Shapiro-Wilk test
+(\texttt{shapiro.test}) from the built-in R package \texttt{stats}.
+These three tests can be used to detect non-linearity,
+heteroskedasticity and non-normality respectively. Their estimated power
+is shown in Figure \ref{fig:polypower}. For the RESET test, we included
+different power of fitted values as proxies. According to
 \citep{ramsey_tests_1969}, there is no general rule for the power of the
 fitted values needed by the RESET test, but it found power up to four is
-usually sufficient. Thus, we followed this guidline to conduct the RESET
-test.
+usually sufficient. Thus, we followed this guideline to conduct the
+RESET test. For the Breusch-Pagan test, both \(\boldsymbol{x}\) and
+\(\boldsymbol{x}^2\) were included in the auxiliary regression equation
+to {[}justify this choice?{]}.
 

### 2022-12-20 paper_comparison/paper_comparison/paper_comparison.Rmd
Fix alpha estimate issue

@@
 @vanderplas2021statistical suggest that $\alpha$ can be estimated using MLE with Equation \ref{eq:c-interesting-expectation}.
-But for precise estimate of $\alpha$, additional responses to Rorschach lineups, which is a type of lineup that consists of
-
-plots constructed from the same null data generating mechnisam.
+But for precise estimate of $\alpha$, additional responses to Rorschach lineups, which is a type of lineup that consists of plots constructed from the same null data generating mechanism, are required.
 
@@
 
-As mentioned in Section \ref{calculating-p-values}, $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using Rorschach lineups.
-Hence, 36 lineups consist of only null plots with all combinations of $n$ and $X_{raw}$ and three replications are designed to be included in experiment III.
-Note that we generated the lineups using the same simulation mechanism discussed in Section \ref{sampling-from-the-null-distribution} rather than from a null model.
-
-(it has zero effect size, an alternative to Rorschach lineups)
-
-(Appendix: vary the alpha a little bit and see how the curves change, sensitivity analysis)
-
-Thus, we used another valid method suggested in vanderplas2021statistical which was designed for typical lineups to estimate $\alpha$ correctly.
-However, by not using Rorschach lineups, we did lose some accuracy in the estimation.
-(questionable)
-
-We recommend future studies to generate Rorschach lineups strictly following the definition given by buja_statistical_2009.
+As mentioned in Section \ref{calculating-p-values-for-the-visual-test}, $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using Rorschach lineups.
+Hence, 36 lineups with all combinations of $n$ and $X_{raw}$ and three replications were included in experiment III.
+In these lineups, the data of the data plot was generated from a model consistent with the null hypothesis, while the data of the 19 null plots were generated using the same simulation method discussed in Section \ref{sampling-from-the-null-distribution}.
+Technically, this was an alternative to Rorschach lineup but with the null distribution further conditioning on the estimated coefficients and the residual sum of squares.
+Thus, we used another valid method suggested in @vanderplas2021statistical which was designed for typical lineups to estimate $\alpha$.
+Considering the differnece between typical lineups and Rorschach lineups, we have also included a sensitivity analysis in the Appendix to examine the impact of the variance of the estimate on our findings.
 

### 2022-12-20 paper_comparison/paper_comparison/paper_comparison.tex
Fix alpha estimate issue

@@
 But for precise estimate of \(\alpha\), additional responses to
-Rorschach lineups, which is a type of lineup that consists of
-
-plots constructed from the same null data generating mechnisam.
+Rorschach lineups, which is a type of lineup that consists of plots
+constructed from the same null data generating mechanism, are required.
 
@@
 
-As mentioned in Section \ref{calculating-p-values}, \(\alpha\) used in
-Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using
-Rorschach lineups. Hence, 36 lineups consist of only null plots with all
-combinations of \(n\) and \(X_{raw}\) and three replications are
-designed to be included in experiment III. Note that we generated the
-lineups using the same simulation mechanism discussed in Section
-\ref{sampling-from-the-null-distribution} rather than from a null model.
-
-(it has zero effect size, an alternative to Rorschach lineups)
-
-(Appendix: vary the alpha a little bit and see how the curves change,
-sensitivity analysis)
-
-Thus, we used another valid method suggested in
-vanderplas2021statistical which was designed for typical lineups to
-estimate \(\alpha\) correctly. However, by not using Rorschach lineups,
-we did lose some accuracy in the estimation. (questionable)
-
-We recommend future studies to generate Rorschach lineups strictly
-following the definition given by buja\_statistical\_2009.
+As mentioned in Section \ref{calculating-p-values-for-the-visual-test},
+\(\alpha\) used in Equation \ref{eq:pvalue-beta-binomial} needs to be
+estimated using Rorschach lineups. Hence, 36 lineups with all
+combinations of \(n\) and \(X_{raw}\) and three replications were
+included in experiment III. In these lineups, the data of the data plot
+was generated from a model consistent with the null hypothesis, while
+the data of the 19 null plots were generated using the same simulation
+method discussed in Section \ref{sampling-from-the-null-distribution}.
+Technically, this was an alternative to Rorschach lineup but with the
+null distribution further conditioning on the estimated coefficients and
+the residual sum of squares. Thus, we used another valid method
+suggested in \citet{vanderplas2021statistical} which was designed for
+typical lineups to estimate \(\alpha\). Considering the differnece
+between typical lineups and Rorschach lineups, we have also included a
+sensitivity analysis in the Appendix to examine the impact of the
+variance of the estimate on our findings.
 

### 2022-12-20 paper_comparison/paper_comparison/paper_comparison.Rmd
add mosaic plot

@@
 
-(top: curves)
-(middle: dots)
-(bottom: vertical lines/grid lines)
-
 ```{r heterpower, fig.height = 3, fig.width = 5, fig.cap="Comparison of power between different tests for heteroskedasticity patterns.
@@
 ```
-
-tick mark on x. dashed line too strong. use solid line (grey) in light color.
 
@@
   theme_light()
+```
+
+```{r}
+library(ggmosaic)
+vi_survey %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  left_join(vi_survey %>% 
+           count(unique_lineup_id) %>%
+           rename(num_subject = n)) %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  mutate(conv_reject = ifelse(conventional_p_value < 0.05, "Reject", "Not reject"),
+         reject = ifelse(p_value < 0.05, "Reject", "Not reject")) %>%
+  ggplot() +
+  ggmosaic::geom_mosaic(aes(x = ggmosaic::product(conv_reject, reject), fill = conv_reject)) +
+  xlab("Visual tests") +
+  ylab("Conventional tests") +
+  labs(fill = "Conventional tests")
 ```

### 2022-12-20 paper_comparison/paper_comparison/paper_comparison.Rmd
mosaic plot tweaking

--- 2022-12-20 paper_comparison/paper_comparison/paper_comparison.Rmd
@@
   summarise(across(everything(), first)) %>%
+  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
   ggplot() +
@@
   theme_light()
+```
+
+```{r compare-reject, fig.width=8, fig.height=4, fig.cap="Comparison of conventional and visual test rejection rates."}
+library(ggmosaic)
+vi_survey %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  left_join(vi_survey %>% 
+           count(unique_lineup_id) %>%
+           rename(num_subject = n)) %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
+  mutate(conv_reject = ifelse(conventional_p_value < 0.05, "Reject", "Not"),
+         reject = ifelse(p_value < 0.05, "Reject", "Not")) %>%
+  mutate(conv_reject = factor(conv_reject, levels = c("Reject", "Not")),
+         reject = factor(reject, levels = c("Reject", "Not"))) %>%
+  ggplot() +
+  ggmosaic::geom_mosaic(aes(x = ggmosaic::product(reject, conv_reject), fill = reject)) +
+  facet_wrap(~type) +
+  ylab("Visual tests") +
+  xlab("Conventional tests") +
+  labs(fill = "Conventional tests") +
+  scale_fill_brewer("", palette="Dark2") +
+  theme_bw() +
+  theme(legend.position="none")
 ```

### 2022-12-20 paper_comparison/paper_comparison/paper_comparison.Rmd
update before meeting

--- 2022-12-20 paper_comparison/paper_comparison/paper_comparison.Rmd
@@
 
-(discuss choice of tests, mention pkg `lmtest`, `skedastic`)
-
 Figure \ref{fig:polypower} shows the estimated power of visual test on lineups produced by the non-linearity model with $X_{raw} \sim U(-1,1)$, against the natural logarithm of the effect $log_e(E)$.
@@
 
-To compare the power of visual test and conventional tests, we picked RESET test (`resettest`) and Breusch-Pagan test (`bptest`) from the R package `lmtest`, and Shapiro-Wilk test (`shapiro.test`) from the built-in R package `stats`.
-These three tests can be used to detect non-linearity, heteroskedasticity and non-normality respectively.
+To compare the power of visual test and conventional test, we picked RESET test (`resettest`) and Breusch-Pagan test (`bptest`) from the R package `lmtest`, and Shapiro-Wilk test (`shapiro.test`) from the built-in R package `stats`.
+Among them, RESET test is the only exact and appropriate test in this scenario.
+Both the Breusch-Pagan test and the Shapiro-Wilk test are approximate and inappropriate tests.
 Their estimated power is shown in Figure \ref{fig:polypower}.
 For the RESET test, we included different power of fitted values as proxies.
-According to [@ramsey_tests_1969], there is no general rule for the power of the fitted values needed by the RESET test, but it found power up to four is usually sufficient.
+According to @ramsey_tests_1969, there is no general rule for the power of the fitted values needed by the RESET test, but it found power up to four is usually sufficient.
 Thus, we followed this guideline to conduct the RESET test.
@@
 For the non-linearity model, the power of all four tests increases as the effect gets larger.
-The power curve of RESET test climbs aggressively from 13% to around 70% as $log_e(E)$ increases from 0 to 2, while power of other tests respond inactively to the change of effect and remain lower than 25% throughout the period, showing that F-test is way more sensitive to the type of model defects that being considered.
+The power curve of RESET test climbs aggressively from 13% to around 70% as $log_e(E)$ increases from 0 to 2, while power of other tests respond inactively to the change of effect and remain lower than 25% throughout the period, showing that REST test is way more sensitive to the type of model defects that being considered.
 Meanwhile, no noticeable visual features can be spotted from the example residuals plots.
@@
 
-Overall, the power analysis suggests that conventional tests do not match how humans perceive residual plots and thus determine model violations in these two scenarios.
-Non-severe model violations with human acceptable residual departures could often alarm by the conventional tests.
+Overall, the power analysis suggests that conventional tests do not match how humans perceive residual plots in these two scenarios.
+Humans have much higher tolerance of the residual departures than the conventional test.
+Since fail to reject $H_0$ in a visual test means that there is no obvious visual discoveries found in the residual plot, analysts and the general public as the consumers of the output may not be convinced of the existence of significant residual departures in spite of the rejection of $H_0$ given by the conventional test.
+Even if the rejection is accepted, the model violation may be considered as impactless due to the fact that they are not clearly visible.
+Besides, the sensitivity of the conventional test could also distract and discourage analysts from finding simple but good linear approximation to the data.
+The rejection of $H_0$ because of human acceptable and negligible residual departures is not practically meaningful and useful.
+(it does not fit the purpose of )
+
 This could limit the popularity of conventional tests in residual diagnostics among analysts.
@@
 
-tick mark on x. dashed line too strong. use solid line (grey) in light color.
-
 <!-- ## Visual test  -->
@@
 
-```{r p-value-comparison, fig.height=4, fig.cap="Visual test p-value compared to conventional test p-value for lineups produced by heteroskedasticity and non-linearity model.
+```{r p-value-comparison-old, eval = FALSE, fig.height=4, fig.cap="Visual test p-value compared to conventional test p-value for lineups produced by heteroskedasticity and non-linearity model.
 The scatter plot is drawn on a square root scale.
@@
   theme_light()
+```
+
+```{r p-value-comparison, fig.height=4, fig.cap="Rejection rate of visual test compared to conventional test on lineups produced by heteroskedasticity and non-linearity model.
+Based on the rejection decision made by the visual test and the conventional test, the mosaic plot shows the proportion of lineups in each category."}
+
+library(ggmosaic)
+vi_survey %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  left_join(vi_survey %>% 
+           count(unique_lineup_id) %>%
+           rename(num_subject = n)) %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  mutate(conv_reject = ifelse(conventional_p_value < 0.05, "Reject", "Not reject"),
+         reject = ifelse(p_value < 0.05, "Reject", "Not reject")) %>%
+  ggplot() +
+  geom_mosaic(aes(x = product(conv_reject, reject), fill = conv_reject)) +
+  scale_fill_manual(values = rev(rcartocolor::carto_pal(3, "Vivid"))[2:3]) +
+  xlab("Visual test") +
+  ylab("Conventional test") +
+  labs(fill = "Conventional test")
 ```

### 2022-12-20 paper_comparison/paper_comparison/paper_comparison.Rmd

@@
   summarise(across(everything(), first)) %>%
-  mutate(conv_reject = ifelse(conventional_p_value < 0.05, "Reject", "Not reject"),
-         reject = ifelse(p_value < 0.05, "Reject", "Not reject")) %>%
+  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
+  mutate(conv_reject = ifelse(conventional_p_value < 0.05, "Reject", "Not"),
+         reject = ifelse(p_value < 0.05, "Reject", "Not")) %>%
+  mutate(conv_reject = factor(conv_reject, levels = c("Reject", "Not")),
+         reject = factor(reject, levels = c("Reject", "Not"))) %>%
   ggplot() +
-  geom_mosaic(aes(x = product(conv_reject, reject), fill = conv_reject)) +
-  scale_fill_manual(values = rev(rcartocolor::carto_pal(3, "Vivid"))[2:3]) +
-  xlab("Visual test") +
-  ylab("Conventional test") +
-  labs(fill = "Conventional test")
+  ggmosaic::geom_mosaic(aes(x = ggmosaic::product(reject, conv_reject), fill = reject)) +
+  facet_wrap(~type) +
+  ylab("Visual tests") +
+  xlab("Conventional tests") +
+  labs(fill = "Conventional tests") +
+  scale_fill_brewer("", palette="Dark2") +
+  theme_bw() +
+  theme(legend.position="none")
 ```

### 2022-12-21 paper_comparison/paper_comparison/paper_comparison.Rmd
Update Rorschach description

@@
 ### Subject allocation
+\newcommand{\svp}[1]{{\textcolor{RedOrange}{#1}}}
 
@@
 
-As mentioned in Section \ref{calculating-p-values-for-the-visual-test}, $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using Rorschach lineups.
+As mentioned in Section \ref{calculating-p-values-for-the-visual-test}, $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using \svp{null} lineups.
 Hence, 36 lineups with all combinations of $n$ and $X_{raw}$ and three replications were included in experiment III.
-In these lineups, the data of the data plot was generated from a model consistent with the null hypothesis, while the data of the 19 null plots were generated using the same simulation method discussed in Section \ref{sampling-from-the-null-distribution}.
-Technically, this was an alternative to Rorschach lineup but with the null distribution further conditioning on the estimated coefficients and the residual sum of squares.
-Thus, we used another valid method suggested in @vanderplas2021statistical which was designed for typical lineups to estimate $\alpha$.
-Considering the differnece between typical lineups and Rorschach lineups, we have also included a sensitivity analysis in the Appendix to examine the impact of the variance of the estimate on our findings.
+In these lineups, the data of the data plot was generated from a model \svp{with zero effect size}, while the data of the 19 null plots were generated using the same simulation method discussed in Section \ref{sampling-from-the-null-distribution}.
+\svp{This generation procedure differs from the canonical Rorschach lineup procedure, which requires that all 20 plots are generated from the null hypothesis.
+However, these lineups serve the same fundamental purpose: to assess the number of visually interesting plots generated under the null hypothesis.}
+<!-- Technically, this was an alternative to Rorschach lineup but with the null distribution further conditioning on the estimated coefficients and the residual sum of squares.  -->
+\svp{To account for the fact that our simulation method for these lineups is not the Rorschach procedure,} we used the method suggested in @vanderplas2021statistical for typical lineups containing a data plot to estimate $\alpha$.
+We have included a sensitivity analysis in the Appendix to examine the impact of the variance of the $\alpha$ estimate on our findings.
 

### 2022-12-28 paper_comparison/paper_comparison/paper_comparison.Rmd
fix issue - Package xcolor Error: Undefined color `RedOrange'

@@
   \def\tightlist{}
+  \usepackage{xcolor}
+  \definecolor{orange-red}{rgb}{1.0, 0.27, 0.0}
+  \newcommand{\svp}[1]{{\textcolor{orange-red}{#1}}}
   \newcommand*{\Perm}[2]{{}^{#1}\!P_{#2}}%
@@
 
-Experiment I is designed to study the ability of human subjects to detect the effect of a non-linear term $\boldsymbol{z}$ constructed using Hermite polynomials [@hermite1864nouveau] on random vector $\boldsymbol{x}$ formulated as
+(include English reference)
+
+Experiment I is designed to study the ability of human subjects to detect the effect of a non-linear term $\boldsymbol{z}$ constructed using Hermite polynomials on random vector $\boldsymbol{x}$ formulated as
 
@@
 ### Subject allocation
-\newcommand{\svp}[1]{{\textcolor{RedOrange}{#1}}}
 
@@
 $H_0$ of the F-test will be always satisfied.
+
+(remove human perception)
+
+(revisit Emi's concern)
 
@@
 
+(1. How you calc rejection, 2. scatter plot p-value vs p-value -> appendix, 3. percentage in explained the tex, both marginal and conditional, 4. interesting lineup, discuss the reason)
+
 (style similar to mosaic plot)
@@
   theme(legend.position="none")
+```
+
+```{r eval = FALSE}
+vi_survey %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  left_join(vi_survey %>% 
+           count(unique_lineup_id) %>%
+           rename(num_subject = n)) %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
+  mutate(conv_reject = ifelse(conventional_p_value < 0.05, "Reject", "Not"),
+         reject = ifelse(p_value < 0.05, "Reject", "Not")) %>%
+  filter(type == "non-linearity", conv_reject == "Not", reject == "Reject") %>%
+  View()
 ```

### 2022-12-28 paper_comparison/paper_comparison/paper_comparison.tex
add mosaic plot
fix issue - Package xcolor Error: Undefined color `RedOrange'

@@
 \def\tightlist{}
+\usepackage{xcolor}
+\definecolor{orange-red}{rgb}{1.0, 0.27, 0.0}
+\newcommand{\svp}[1]{{\textcolor{orange-red}{#1}}}
 \newcommand*{\Perm}[2]{{}^{#1}\!P_{#2}}%
@@
 
+(include English reference)
+
 Experiment I is designed to study the ability of human subjects to
 detect the effect of a non-linear term \(\boldsymbol{z}\) constructed
-using Hermite polynomials \citep{hermite1864nouveau} on random vector
-\(\boldsymbol{x}\) formulated as
+using Hermite polynomials on random vector \(\boldsymbol{x}\) formulated
+as
 
@@
 \(\alpha\) used in Equation \ref{eq:pvalue-beta-binomial} needs to be
-estimated using Rorschach lineups. Hence, 36 lineups with all
-combinations of \(n\) and \(X_{raw}\) and three replications were
-included in experiment III. In these lineups, the data of the data plot
-was generated from a model consistent with the null hypothesis, while
-the data of the 19 null plots were generated using the same simulation
-method discussed in Section \ref{sampling-from-the-null-distribution}.
-Technically, this was an alternative to Rorschach lineup but with the
-null distribution further conditioning on the estimated coefficients and
-the residual sum of squares. Thus, we used another valid method
-suggested in \citet{vanderplas2021statistical} which was designed for
-typical lineups to estimate \(\alpha\). Considering the differnece
-between typical lineups and Rorschach lineups, we have also included a
-sensitivity analysis in the Appendix to examine the impact of the
-variance of the estimate on our findings.
+estimated using {\textcolor{orange-red}{null}} lineups. Hence, 36
+lineups with all combinations of \(n\) and \(X_{raw}\) and three
+replications were included in experiment III. In these lineups, the data
+of the data plot was generated from a model
+{\textcolor{orange-red}{with zero effect size}}, while the data of the
+19 null plots were generated using the same simulation method discussed
+in Section \ref{sampling-from-the-null-distribution}.
+{\textcolor{orange-red}{This generation procedure differs from the canonical Rorschach lineup procedure, which requires that all 20 plots are generated from the null hypothesis.
+However, these lineups serve the same fundamental purpose: to assess the number of visually interesting plots generated under the null hypothesis.}}
+{\textcolor{orange-red}{To account for the fact that our simulation method for these lineups is not the Rorschach procedure,}}
+we used the method suggested in \citet{vanderplas2021statistical} for
+typical lineups containing a data plot to estimate \(\alpha\). We have
+included a sensitivity analysis in the Appendix to examine the impact of
+the variance of the \(\alpha\) estimate on our findings.
 
@@
 
-(discuss choice of tests, mention pkg \texttt{lmtest},
-\texttt{skedastic})
-
 Figure \ref{fig:polypower} shows the estimated power of visual test on
@@
 
-To compare the power of visual test and conventional tests, we picked
+To compare the power of visual test and conventional test, we picked
 RESET test (\texttt{resettest}) and Breusch-Pagan test (\texttt{bptest})
@@
 (\texttt{shapiro.test}) from the built-in R package \texttt{stats}.
-These three tests can be used to detect non-linearity,
-heteroskedasticity and non-normality respectively. Their estimated power
-is shown in Figure \ref{fig:polypower}. For the RESET test, we included
-different power of fitted values as proxies. According to
-\citep{ramsey_tests_1969}, there is no general rule for the power of the
+Among them, RESET test is the only exact and appropriate test in this
+scenario. Both the Breusch-Pagan test and the Shapiro-Wilk test are
+approximate and inappropriate tests. Their estimated power is shown in
+Figure \ref{fig:polypower}. For the RESET test, we included different
+power of fitted values as proxies. According to
+\citet{ramsey_tests_1969}, there is no general rule for the power of the
 fitted values needed by the RESET test, but it found power up to four is
@@
 effect and remain lower than 25\% throughout the period, showing that
-F-test is way more sensitive to the type of model defects that being
+REST test is way more sensitive to the type of model defects that being
 considered. Meanwhile, no noticeable visual features can be spotted from
@@
 
+(remove human perception)
+
+(revisit Emi's concern)
+
 Overall, the power analysis suggests that conventional tests do not
-match how humans perceive residual plots and thus determine model
-violations in these two scenarios. Non-severe model violations with
-human acceptable residual departures could often alarm by the
-conventional tests. This could limit the popularity of conventional
-tests in residual diagnostics among analysts.
+match how humans perceive residual plots in these two scenarios. Humans
+have much higher tolerance of the residual departures than the
+conventional test. Since fail to reject \(H_0\) in a visual test means
+that there is no obvious visual discoveries found in the residual plot,
+analysts and the general public as the consumers of the output may not
+be convinced of the existence of significant residual departures in
+spite of the rejection of \(H_0\) given by the conventional test. Even
+if the rejection is accepted, the model violation may be considered as
+impactless due to the fact that they are not clearly visible. Besides,
+the sensitivity of the conventional test could also distract and
+discourage analysts from finding simple but good linear approximation to
+the data. The rejection of \(H_0\) because of human acceptable and
+negligible residual departures is not practically meaningful and useful.
+(it does not fit the purpose of )
+
+This could limit the popularity of conventional tests in residual
+diagnostics among analysts.
 
@@
 
-(top: curves) (middle: dots) (bottom: vertical lines/grid lines)
-
 \begin{figure}
@@
 
-tick mark on x. dashed line too strong. use solid line (grey) in light
-color.
-
 \hypertarget{comparison-of-rejection-rates-based-on-p-values}{%
@@
 \(p\)-values}{Comparison of rejection rates based on p-values}}\label{comparison-of-rejection-rates-based-on-p-values}}
+
+(1. How you calc rejection, 2. scatter plot p-value vs p-value
+-\textgreater{} appendix, 3. percentage in explained the tex, both
+marginal and conditional, 4. interesting lineup, discuss the reason)
 
@@
 
-\caption{Visual test p-value compared to conventional test p-value for lineups produced by heteroskedasticity and non-linearity model.
-The scatter plot is drawn on a square root scale.
-Every dot on the plot corresponds to a lineup.
-Most of the dots are fallen on the left of the y = x line indicating visual test p-value is almost always higher than conventional test p-value.
-The blue curve is a smoothing of the dots showing a positive trend.
-}\label{fig:p-value-comparison}
+\caption{Rejection rate of visual test compared to conventional test on lineups produced by heteroskedasticity and non-linearity model.
+Based on the rejection decision made by the visual test and the conventional test, the mosaic plot shows the proportion of lineups in each category.}\label{fig:p-value-comparison}
 \end{figure}

### 2023-01-09 paper_comparison/paper_comparison/paper_comparison.Rmd
major update

@@
 Diagnosing a model is the key to determining whether there is anything importantly wrong.
-For linear regression analysis, we typically interrogate the residuals for model diagnostics.
+In linear regression analysis, residuals are typically examined for model diagnostics.
 Residuals summarise what is not captured by the model, and thus provide the capacity to identify what might be wrong.
 
-We can assess residuals in multiple ways.
+One can access residuals in multiple ways.
 Residuals might be plotted, as a histogram or quantile-quantile plot to examine the distribution.
@@
 But if the distribution is skewed, bimodal, multimodal, or contains outliers, there is cause for concern.
-The distribution could also be inspected by conducting a goodness of fit test, such as the Shapiro-Wilk Normality test [@shapiro1965analysis].
+One could also inspect the distribution by conducting a goodness of fit test, such as the Shapiro-Wilk Normality test [@shapiro1965analysis].
 
 More typically, residuals will be plotted, as a scatter plot against the predicted values and each of the explanatory variables to scrutinize their relationships.
-If there is any visually discoverable patterns, the model is potentially misspecified.
-However, correctly judging a residual plot where no pattern exists is a painstakingly difficult task for humans.
-It is especially common, particularly among new data analytics students to report patterns when an experienced data analyst might quickly conclude that there are none.
-Generally, one looks for noticeable departures from the model like non-linear dependency or heteroskedasticity.
+If there are any visually discoverable patterns, the model is potentially misspecified.
+In general, one looks for noticeable departures from the model like non-linear dependency or heteroskedasticity.
+However, correctly judging a residual plot where no pattern exists is a painstakingly difficult task for humans (?citation).
+It is especially common, particularly among new data analysts, to report patterns when an experienced one might quickly conclude that there are none.
 It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and use a Breusch-Pagan test [@breusch_simple_1979] for heteroskedasticity.
@@
 Abundance of literature describe appropriate diagnostic methods for linear regression: @draper1998applied, @montgomery1982introduction, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals.
-The diligent reader of these sage writings will also notice sentences that express sentiments like *based on their experience, statistical tests are not widely used in regression diagnostics.
-The same or even larger amount of information can be provided by diagnostic plots than the corresponding tests in most empirical studies.* A common guidance by experts is that optimal method for diagnosing model fits is by plotting the data.
+The diligent reader of these sage writings will also notice sentences that express sentiments like *based on their experience, statistical tests are not widely used in regression diagnostics since the same or even larger amount of information can be provided by diagnostic plots than the corresponding tests in most empirical studies.* A common guidance by experts is that optimal method for diagnosing model fits is by plotting the data.
 
@@
 The paper is structured as follows.
-The next background section describes the types of departures that one expects to detect, and describes a formal statistical process for reading residual plots, called visual inference.
-Section \ref{experiment-design} describes the experiment design to compare the decision made by formal hypothesis testing, and how humans would read diagnostic plots.
+The next background section describes the types of departures that one expects to detect, and outlines a formal statistical process for reading residual plots, called visual inference.
+Section \ref{experimental-design} details the experimental design to compare the decision made by formal hypothesis testing, and how humans would read diagnostic plots.
 The results are reported in Section \ref{results}.
-We finish with a discussion on future work, in particular how the responsibility for residual plot reading might be passed on to computer vision.
+We finish with a discussion on future work, in particular how the responsibility for residual plot reading might be relieved.
 
@@
 
-```{r residual-plot-common-departures, fig.cap = "Example fitted vs residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
+```{r residual-plot-common-departures, fig.width = 10, fig.height = 2.5, fig.cap = "Example fitted vs residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
 The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted plot."}
@@
     xlab("Fitted values") + 
-    ylab("Residuals")
+    ylab("Residuals"),
+  ncol = 4
 )
@@
 Graphical summaries in which residuals are plotted against fitted values or other functions of the regressors that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals.
-As shown in Figure \ref{fig:residual-plot-common-departures}, the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns.
-
-There are various types of departures from a good residual plot. Non-linearity, heterskedasticity and non-normality are perhaps the three mostly checked departures.
+As shown in Figure \ref{fig:residual-plot-common-departures}, the plot A is an ideal residual plot with residuals evenly distributed at both sides of the horizontal zero line, with no noticeable patterns.
+
+There are various types of departures from an ideal residual plot. Non-linearity, heterskedasticity and non-normality are perhaps the three mostly checked departures.
 
@@
 
-Compared to non-linearity and heteroskedasticity, non-normality is usually harder to detect from a residual plot since scatter plot do not readily reveal the marginal distribution.
+Compared to non-linearity and heteroskedasticity, non-normality is usually harder to detect from a residual plot since a scatter plot do not readily reveal the marginal distribution.
 A favourable graphical summary for this task is the quantile-quantile plot.
+Hence, non-normality will not be the focus of this paper as we mainly discuss residual plots.
 For a consistent comparison, the residual plot of this departure is still presented in plot D of Figure \ref{fig:residual-plot-common-departures}.
-However, since the scatter plot is not suitable for this task, non-normality will not be the focus of this paper.
 Besides, it is important to note that not all regression models assume normality for the error term, but a certain amount do including the classical normal linear regression model.
 In the case that the normality assumption is violated, it is expected to observe data points do not center around the horizontal axis and there is an uneven distribution of the number points at both below and above the horizontal axis.
-For example, given a skewed error distribution, fewer data points and more outliers are on one side of the horizontal axis as shown in plot D of Figure \ref{fig:residual-plot-common-departures}.
+For example, given a skewed error distribution, there will be fewer data points and more outliers on one side of the horizontal axis as shown in plot D of Figure \ref{fig:residual-plot-common-departures}.
 
@@
 For testing non-linearity, one may apply the F-test as a model structural test to examine the significance of specific polynomial and non-linear forms of the regressors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969].
-And for testing normality, the Shapiro--Wilk test [@shapiro1965analysis] is perhaps the most widely used test included by many of the statistical softwares.
-Another choice will be the Jarque--Bera test [@jarque1980efficient] which directly checks if the sample skewness and kurtosis match a normal distribution.
+The Shapiro-Wilk test [@shapiro1965analysis] is the most widely used test of non-normality included by many of the statistical softwares.
+The Jarque--Bera test [@jarque1980efficient] is also used to directly checks if the sample skewness and kurtosis match a normal distribution.
 
 Example residual plots given in Figure \ref{fig:residual-plot-common-departures} are examined by the corresponding RESET test, Breusch-Pagan test and Shapiro-Wilk test as shown in Table \ref{tab:example-residual-plot-table}.
-In the example, both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null hypothesis $H_0$ for departures that they do not intend to examine.
-As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are sensitive to other types of departures.
-It is likely $H_0$ is correctly rejected but for the wrong reason, which is known as the "Type III error".
+In the example, the Breusch-Pagan test and the Shapiro--Wilk test both reject the null hypothesis $H_0$ for departures that they do not intend to examine.
+As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are also sensitive to other types of departures.
+It is likely $H_0$ is correctly rejected but for the wrong reason, a phenomenon known as the "Type III error".
 Additionally, outliers will often incorrectly trigger the rejection of $H_0$ despite when majority of the residuals are well-behaved [@cook_applied_1999].
@@
 
-One may argue that reading diagnostic plots is to some extent subjective and indecisive compared to those rigorous statistical procedures as it relies on graphical perception - human's ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984].
-It is particularly true that visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
-One such example is finding an over-interpretation of the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurrence [@roy_chowdhury_using_2015].
-
-Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualization" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots [@gelman_exploratory_2004].
+One may argue that reading diagnostic plots is to some extent subjective and indecisive compared to those rigorous statistical procedures as it relies on graphical perception - human ability to interpret and decode the information embedded in graph [@cleveland_graphical_1984].
+Further, the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data.
+For instance, people
+over-interpret the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurrence [@roy_chowdhury_using_2015].
+
+Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualization" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots.
 Later, @buja_statistical_2009 proposed the lineup protocol as a visual test inspired by the "police lineup" or "identity parade" which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people.
-The protocol consists of $m$ randomly placed plots, where one plot is the actual data plot, and the remaining $m - 1$ plots have the identical graphical production as the data plot except the data has been replaced with data consistent with $H_0$.
-Then, an observer who have not seen the actual data plot will be asked to point out the most different plot from the lineup.
-Under $H_0$, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is $1/m$.
-If we reject $H_0$ as the observer correctly picks the actual data plot, then the Type I error of this test is $1/m$.
-
-```{r first-example-lineup, fig.cap="Visual testing is conducted using a lineup, as in the example here.
-The residual plot computed from the observed data (plot $2^2 + 2$, exhibiting non-linearity) is embedded among 19 null plots, where the residuals were simulated from a standard error model.
+The protocol consists of $m$ randomly placed plots, where one plot is the data plot, and the remaining $m - 1$ plots have the identical graphical procedure except the data has been replaced with data consistent with $H_0$.
+Then, an observer who have not seen the data plot will be asked to point out the most different plot from the lineup.
+Under $H_0$, it is expected that the data plot would have no distinguishable difference from the null plots, and the probability that the observer correctly picks the data plot is $1/m$.
+If one rejects $H_0$ as the observer correctly picks the data plot, then the Type I error of this test is $1/m$.
+
+```{r first-example-lineup, fig.height = 5, fig.cap="Visual testing is conducted using a lineup, as in the example here.
+The residual plot computed from the observed data (plot $2^2 + 2$, exhibiting non-linearity) is embedded among 19 null plots, where the residuals are simulated from a standard error model.
 Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot.
@@
 Figure \ref{fig:first-example-lineup} is an example of a lineup protocol.
-If the actual data plot at position $2^2 + 2$ is identifiable, then it is evidence for the rejection of $H_0$ that the regression model is correctly specified.
+If the data plot at position $2^2 + 2$ is identifiable, then it is evidence for the rejection of $H_0$ that the regression model is correctly specified.
 In fact, the actual residual plot is obtained from a misspecified regression model with non-linearity defect.
 
-The effectiveness of lineup protocol is validated by @majumder_validation_2013 under relatively simple classical normal linear regression model settings with one or two regressors.
-Their results suggest that visual test is capable of testing the significance of a single regressor with a similar power as a t-test, though they express that in general it is unnecessary to use visual inference if there exists a conventional test and they do not expect the visual test to perform equally well as the conventional test.
-In their third experiment, where there does not exist a proper conventional test, visual test outperforms the conventional test for a large margin.
-This is encouraging as it promotes the use of visual inference in border field of data science where there are no existing statistical testing procedures.
-Visual inference have also already been integrated into diagnostic of hierarchical linear models by @loy2013diagnostic, @loy2014hlmdiag and @loy2015you.
+The effectiveness of lineup protocol for regression analysis is validated by @majumder_validation_2013 under relatively simple settings with up to two regressors.
+Their results suggest that visual tests are capable of testing the significance of a single regressor with a similar power as a t-test, though they express that in general it is unnecessary to use visual inference if there exists a conventional test, and they do not expect the visual test to perform equally well as the conventional test.
+In their third experiment, where there is not a conventional test, visual test outperforms the conventional test for a large margin.
+This is encouraging, as it promotes the use of visual inference in situations where there are no existing statistical testing procedures.
+Visual inference have also been integrated into diagnostic of hierarchical linear models by @loy2013diagnostic, @loy2014hlmdiag and @loy2015you.
 They use lineup protocols to judge the assumption of linearity, normality and constant error variance for both the level-1 and level-2 residuals.
+(expand?)
 
@@
 
-Data used in the $m - 1$ null plots need to be simulated.
+Data used in the $m - 1$ null plots needs to be simulated.
 In the context of regression diagnostics, sampling data from $H_0$ is equivalent to sampling data from the assumed model.
-As @buja_statistical_2009 suggested, $H_0$ is usually composited by a collection of distributions controlled by nuisance parameters.
-Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling.
+As @buja_statistical_2009 suggested, $H_0$ is usually a composite hypothesis controlled by nuisance parameters.
+Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to a so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling.
 The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009].
@@
 
-Further, a visual test can involve $K$ independent observers.
-Let $X_i = \{0,1\}$ be a binomial random variable denoting whether subject $i$ correctly detecting the actual data plot, and $X = \sum_{i=1}^{K}X_i$ be the number of observers correctly picking the actual data plot.
+In hypothesis testing, $p$-value is defined as the probability of observing test results as least as extreme as the observed result given $H_0$ is true.
+Within the context of visual inference, by involving $k$ independent observers, $p$-value can be interpreted as the probability of having as many or more subjects detect the data plot than the observed result.
+
+Let $X_j = \{0,1\}$ be a binomial random variable denoting whether subject $j$ correctly detecting the data plot, and $X = \sum_{j=1}^{K}X_j$ be the number of observers correctly picking the data plot.
 Then, by imposing a relatively strong assumption on the visual test that all $K$ evaluations are fully independent, under $H_0$, $X \sim \mathrm{Binom}_{K,1/m}$.
-Therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as $P(X \geq x) = 1 - F(x) + f(x)$, where $F(.)$ is the cumulative distribution function, $f(.)$ is the probability mass function and $x$ is the realization of number of observers correctly picking the actual data plot [@majumder_validation_2013].
+Therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as $P(X \geq x) = 1 - F(x) + f(x)$, where $F(.)$ is the cumulative distribution function, $f(.)$ is the probability mass function and $x$ is the realization of number of observers correctly picking the data plot [@majumder_validation_2013].
 
@@
 And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup.
-@vanderplas2021statistical summarises three common scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same actual data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects.
+@vanderplas2021statistical summarises three common scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects.
 Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experiment design.
-For Scenario 3, @vanderplas2021statistical model the probability of a plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$.
-And defined $c_i$ to be the number of times plot $i$ being selected in $K$ evaluations.
+For Scenario 3, @vanderplas2021statistical models the probability of a plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$.
+The number of times plot $i$ being selected in $K$ evaluations is denoted as $c_i$.
 In case subject $j$ makes multiple selections, $1/s_j$ will be added to $c_i$ instead of one, where $s_j$ is the number of plots subject $j$ selected for $j=1,...K$.
-This ensured $\sum_{i}c_i=K$.
-Since we are only interested in the selections of the actual data plot $i$, the marginal model can be simplified to a beta-binomial model and thus the visual p-value is given as
+This ensures $\sum_{i}c_i=K$.
+Since we are only interested in the selections of the data plot $i$, the marginal model can be simplified to a beta-binomial model and thus the visual p-value is given as
 
@@
 
-Note that Equation \ref{eq:pvalue-beta-binomial} only works with non-negative integer $c_i$. For non-negative real number $c_i$, linear approximation will be applied to calculate the p-value
+Note that Equation \ref{eq:pvalue-beta-binomial} given in @vanderplas2021statistical only works with non-negative integer $c_i$.
+We extend the equation to non-negative real number $c_i$ by applying a linear approximation
 
@@
 
-\begin{equation}
+\begin{equation} \label{eq:pmf-beta-binomial}
 P(C = c_i) = {K \choose c_i}\frac{B(c_i + \alpha, K - c_i + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+.
@@
 
-Besides, the parameter $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} is usually unknown and hence needs to be estimated from the survey data.
+Besides, the parameter $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} and \ref{eq:pmf-beta-binomial} is usually unknown and hence needs to be estimated from the survey data.
 For low values of $\alpha$, only a few plots are attractive to the observers and tend to be selected.
 For higher values of $\alpha$, the distribution of the probability of each plot being selected is more even.
-@vanderplas2021statistical define that a plot is $c$-interesting if $c$ or more participants select the plot as the most different.
+@vanderplas2021statistical defines that a plot is $c$-interesting if $c$ or more participants select the plot as the most different.
 Given the definition, The expected number of plots selected at least $c$ times, $E[Z_c]$, is calculated as
@@
 
-@vanderplas2021statistical suggest that $\alpha$ can be estimated using MLE with Equation \ref{eq:c-interesting-expectation}.
+With Equation \ref{eq:c-interesting-expectation}, $\alpha$ can be estimated using maximum likelihood estimation.
 But for precise estimate of $\alpha$, additional responses to Rorschach lineups, which is a type of lineup that consists of plots constructed from the same null data generating mechanism, are required.
@@
 
-As discussed in @majumder_validation_2013, individual's skill will affect the number of observers who identify the actual data plot from the lineup.
-Thus, the power of a visual test depends on the subject-specific abilities.
-Previously, it is addressed by modelling the probability of a subject $i$ correctly picking the actual data plot from a lineup $l$ using a mixed-effect logistic regression with the subject being treated as a random effect [@majumder_validation_2013].
-However, having this probability is insufficient to determine the power of a visual test allowed for multiple selections as it doesn't provide information about the number of selections made by the subject for p-value calculation.
-
-Instead, we omitted the individual effect and directly estimated the probability of a lineup being rejected using a logistic regression with the natural logarithm of the effect size as the only regressor formulated as
+The power of a model misspecification test specifies the chance of a misspecified regression model being detected given a certain alternative hypothesis.
+It is an important indicator when one concerns if the regression model is misspecified.
+Though, people might usually be more interested in knowing how much the residuals depart from the model assumptions and if the departure is of practical significance.
+
+As discussed in @majumder_validation_2013, individual's skill may affect the number of observers who identify the data plot from the lineup.
+Thus, the power of a visual test could depend on the subject-specific abilities.
+Previously, it is addressed by modelling the probability of a subject $j$ correctly picking the data plot from a lineup $l$ using a mixed-effect logistic regression with the subject being treated as a random effect [@majumder_validation_2013].
+However, in the multiple selections scenario, having this probability is insufficient to determine the power of a visual test as it doesn't provide information about the number of selections made by the subject for p-value calculation.
+
+Instead, we estimate the probability of a lineup being rejected directly by assuming the individual skill has negligible impact on the variation of the power of a lineup.
+The assumption is made to simplify the model structure, thus mitigating the need for expensive large-scale experiments to estimate the complex covariance matrix.
+The model is a logistic regression with the natural logarithm of the effect as the only regressor formulated as
 
@@
 \begin{equation} \label{eq:logistic-regression-1-1}
-Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(\boldsymbol{E})),
+Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(E)),
 \end{equation}
@@
 \noindent where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
+
+```{r eval = FALSE}
+mix_effect_dat <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(x_dist == "uniform") %>%
+  filter(type == "polynomial") %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  mutate(subject = factor(as.integer(factor(paste(exp, set))))) %>%
+  mutate(reject = p_value < 0.05) %>%
+  select(log_effect_size, reject, subject)
+
+# Singularity issue (random effect too small? not worth to fit a mix-effect model?)
+mix_mod <- lme4::glmer(reject ~ scale(log_effect_size) + (1 | subject), data = mix_effect_dat, family = binomial())
+
+pred_dat <- map_df(seq(-1, 6, 0.1),
+    function(x) {
+      result <- data.frame(log_effect_size = rep(x, length(unique(mix_effect_dat$subject))))
+      pred <- predict(mix_mod, 
+                      type = "response", 
+                      newdata = data.frame(log_effect_size = x, 
+                                           subject = unique(mix_effect_dat$subject)))
+      result$subject <- names(pred)
+      result$pred <- pred
+      result
+    })
+
+pred_dat %>%
+  ggplot() +
+  geom_line(aes(log_effect_size, pred, group = subject))
+```
 
@@
 
-Three experiments were conducted to investigate the difference between conventional hypothesis testing and visual inference in the application of linear regression diagnostics.
+Three experiments are conducted to investigate the difference between conventional hypothesis testing and visual inference in the application of linear regression diagnostics.
 The experiment I has ideal scenario for conventional testing, where the visual test is not expected to outperform the conventional test.
 Meanwhile, the experiment II is a scenario where the conventional test is an approximate test, in which the visual test may have a chance to match the performance of the conventional test.
-The experiment III is designed for collecting human responses to lineup with only good residual plots such that the parameter $\alpha$ in Equation \ref{eq:pvalue-beta-binomial} can be estimated.
-Overall, we planned to collect 7974 evaluations on 1152 uniqued lineups performed by 443 subjects throughout three experiment.
+The experiment III is designed for collecting human responses to null lineups such that the parameter $\alpha$ in Equation \ref{eq:pvalue-beta-binomial} can be estimated.
+Overall, we plan to collect 7974 evaluations on 1152 uniqued lineups performed by 443 subjects throughout three experiments.
 
@@
 
-Two types of departures, namely non-linearity and heteroskedasticity, were considered with the corresponding data generating process being designed for experiment I and II.
+Two types of departures, namely non-linearity and heteroskedasticity, are considered with the corresponding data generating process being designed for experiment I and II.
 
@@
 <!-- (hermite poly is created by ... in 18..,) -->
-
-(include English reference)
 
@@
 \boldsymbol{z} = g(\boldsymbol{z}_{raw}, 1), \\
-\boldsymbol{z}_{raw} = He_j(g(\boldsymbol{z}, 2)),
+\boldsymbol{z}_{raw} = He_j(g(\boldsymbol{x}, 2)),
 \end{align}
 
-\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vectors of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $\{-k, k\}$ defined as
+\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vectors of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
 
@@
 
-Hermite polynomials were initially defined by @de1820theorie, but named after Hermite [@hermite1864nouveau] because of the unrecognisable form of Laplace's work.
-When simulating $Z_{raw}$, the function `hermite` from the R package `mpoly` [@mpoly] was used to generate Hermite polynomials.
+According to @abramowitz1964handbook, Hermite polynomials were initially defined by @de1820theorie, but named after Hermite [@hermite1864nouveau] because of the unrecognisable form of Laplace's work.
+When simulating $\boldsymbol{z}_{raw}$, function `hermite` from the R package `mpoly` [@mpoly] is used to generate Hermite polynomials.
 
@@
 
-Since $z = O(x^j)$, for $j > 1$, $z$ is a higher order term leaves out by the null regression, which will result in model misspecification.
-Visual patterns of non-linearity were simulated using four different order of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
+Since $z = O(x^j)$, for $j > 1$, $z$ is a higher order term leaves out by the null regression, which will lead to model misspecification.
+
+Visual patterns of non-linearity are simulated using four different orders of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$) and four different distributions of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
 A summary of the factors is given in Table \ref{tab:model-factor-table}.
@@
 
-The values of $j$ was chosen so that distinct shapes of non-linearity were included in the residual plot.
+The values of $j$ is chosen so that distinct shapes of non-linearity are included in the residual plot.
+These include "U", "S", "M" and "Triple-U" shape as shown in Figure \ref{fig:different-shape-of-herimite}.
 A greater value of $j$ will result in a curve with more turning points.
-As shown in Figure \ref{fig:different-shape-of-herimite}, it includes "U" shape, "S" shape, "M" shape and "Triple-U" shape.
 It is expected that the "U" shape will be the easiest one to detect because complex shape tends to be concealed by cluster of data points.
 
-Four different distribution were used to generate $X_{raw}$ as shown in Figure \ref{fig:different-dist}.
+Different distributions of $X_{raw}$ help enriching the pool of visual patterns as illustrated in Figure \ref{fig:different-dist}.
 The uniform and the normal distribution are symmetric and commonly assumed in statistical models.
-The adjusted log-normal distribution provides skewed density, while the discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.
+The adjusted log-normal distribution provides skewed density, while the discrete uniform distribution provides discreteness in residual plot.
 
@@
 
-Figure \ref{fig:example-poly-lineup} shows one of the lineups used in experiment I.
-This lineup is produced under $j = 6$ and $X_{raw} \sim N(0.0.3^2)$.
-The actual data plot location is $2^3 - 4$.
-All five subjects correctly identified the actual data plot for this lineup.
-
-```{r example-poly-lineup, fig.cap="Lineup poly-24 in experiment I. Can you spot the most different plot? \\label{fig:example-poly-lineup}"}
+Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups used in experiment I.
+This lineup is produced by the non-linearity model under $j = 6$ and $X_{raw} \sim N(0.0.3^2)$.
+The data plot location is $2^3 - 4$.
+All five subjects correctly identify the data plot from this lineup.
+
+```{r example-poly-lineup, fig.height=5, fig.cap="Lineup poly-24 in experiment I. Can you spot the most different plot? \\label{fig:example-poly-lineup}"}
 VI_MODEL$plot_lineup(vi_lineup[["poly_24"]]$data, 
@@
 \boldsymbol{x} = g(\boldsymbol{x}_{raw}, 1)\\
-\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + 2 - |a| b (\boldsymbol{x} - a)^2 \boldsymbol{I}), \\
+\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + (2 - |a|)(\boldsymbol{x} - a)^2b \boldsymbol{I}), \\
 \end{align}
@@
 
-\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vectors of size $n$ and $g(.)$ is the scaling function defined in \ref{eq:scaling-function}.
+\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vectors of size $n$ and $g(.)$ is the scaling function defined in Equation \ref{eq:scaling-function}.
 
@@
 For $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the regressor $\boldsymbol{x}$, which will lead to the presence of heteroskedasticity.
-Visual patterns were simulated using three different shapes ($a$ = -1, 0, 1) and the same four different distribution of $X_{raw}$ used in experiment I.
-A summary of the factors is given in Table \ref{tab:model-factor-table}.
-
-The values of $a$ was chosen so that different shapes of heteroskedasticity were included in the residual plot.
-These include left-triangle shape, butterfly shape and right-triangle shape as displayed in Figure \ref{fig:different-shape-of-heter}.
-
-```{r different-shape-of-heter, fig.height=3, fig.cap="Heteroskedasticity forms used in experiment II.
+Visual patterns of heteroskedasticity are simulated using three different shapes ($a$ = -1, 0, 1) and the same four different distribution of $X_{raw}$ used in experiment I.
+A summary of the factors can also be found in Table \ref{tab:model-factor-table}.
+
+Since $supp(X) = [-1, 1]$, choosing $a$ to be $-1$, $0$ and $1$ can generate "left-triangle", "butterfly" and "right-triangle" shape as displayed in Figure \ref{fig:different-shape-of-heter}.
+The term $(2 - |a|)$ maintains the magnitude of residuals across different values of $a$.
+
+```{r different-shape-of-heter, fig.height=2.67, fig.cap="Heteroskedasticity forms used in experiment II.
 Three different shapes (a = -1, 0, 1) are used in the experiment to create left-triangle, butterfly and right-triangle shapes."}
@@
 
-```{r example-heter-lineup, fig.cap="Lineup heter-169 in experiment II. Can you spot the most different plot? \\label{fig:example-heter-lineup}"}
+```{r example-heter-lineup, fig.height = 5, fig.cap="Lineup heter-169 in experiment II. Can you spot the most different plot? \\label{fig:example-heter-lineup}"}
 
@@
 
-An example lineup of this model is shown in Figure \ref{fig:example-heter-lineup} with $a = -1$ and $X_{raw} \sim U(-1, 1)$.
-The actual data plot location is $2^4 + 2$.
-Nine out of 11 subjects correctly identified the actual data plot for this lineup.
+An example lineup of this model used in Experiment II is shown in Figure \ref{fig:example-heter-lineup} with $a = -1$ and $X_{raw} \sim U(-1, 1)$.
+The data plot location is $2^4 + 2$.
+Nine out of 11 subjects correctly identify the data plot from this lineup.
 
@@
 
-As summarised in Table \ref{tab:model-factor-table}, three additional parameters $n$, $\sigma$ and $b$ were used to control the strength of the signal so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.
-Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ were used in experiment I and II respectively.
+As summarised in Table \ref{tab:model-factor-table}, three additional parameters $n$, $\sigma$ and $b$ are used to control the strength of the signal so that different difficulty levels of lineups are generated, and therefore, the estimated power curve will be smooth and continuous.
+Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ are used in experiment I and II respectively.
 Figure \ref{fig:different-sigma} and \ref{fig:different-b} demonstrate the impact of these two parameters.
 A large value of $\sigma$ will increase the variation of the error of the non-linearity model and decrease the visibility of the visual pattern.
-The parameter $b$ controls the ratio of the standard deviation of the heteroskedasticity across the domain of the regressor.
+The parameter $b$ controls the standard deviation of the error across the support of the regressor.
 Given $x \neq a$, a larger value of $b$ will lead to a larger ratio of the variance at $x$ to the variance at $x - a = 0$, making the visual pattern more obvious.
 
-Three different sample sizes were used (n = 50, 100, 300) in all three experiments.
+Three different sample sizes are used (n = 50, 100, 300) in all three experiments.
 It can be observed from Figure \ref{fig:different-n} that with fewer data points drawn in a residual plot, the visual pattern is more difficult to be detected.
 
-```{r different-n, fig.height = 3, fig.cap="Three different values of $n$ are used in experiment I, II and III to control the strength of the signal."}
+```{r different-n, fig.height = 2.67, fig.cap="Three different values of $n$ are used in experiment I, II and III to control the strength of the signal."}
 ori_dat <- poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(300)
@@
 
-Three replications were made for each of the parameter values shown in Table \ref{tab:model-factor-table} resulting in $(4 \times 4 \times 4 \times 3 + 4 \times 3 \times 5 \times 3) \times 3 = 1116$ different lineups.
-In addition, each lineup was designed to be evaluated by five different subjects.
-After attempting some pilot studies internally in our research group, we decided to present a block of 20 lineups to every subject.
-And to ensure the quality of the survey data, two lineups with obvious visual patterns were included as attention checks.
-Thus, $576 \times 5 \div (20-2) = 160$ and $540 \times 5 \div (20-2) = 150$ subjects were recruited to satisfy the design of the experiment I and experiment II respectively.
+Three replications are made for each of the parameter values shown in Table \ref{tab:model-factor-table} resulting in $(4 \times 4 \times 4 \times 3 + 4 \times 3 \times 5 \times 3) \times 3 = 1116$ different lineups.
+In addition, each lineup is designed to be evaluated by five different subjects.
+After attempting some pilot studies internally in our research group, we decide to present a block of 20 lineups to every subject.
+And to ensure the quality of the survey data, two lineups with obvious visual patterns are included as attention checks.
+Thus, $576 \times 5 / (20-2) = 160$ and $540 \times 5 / (20-2) = 150$ subjects are recruited to satisfy the design of the experiment I and experiment II respectively.
 
 As mentioned in Section \ref{calculating-p-values-for-the-visual-test}, $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using \svp{null} lineups.
-Hence, 36 lineups with all combinations of $n$ and $X_{raw}$ and three replications were included in experiment III.
-In these lineups, the data of the data plot was generated from a model \svp{with zero effect size}, while the data of the 19 null plots were generated using the same simulation method discussed in Section \ref{sampling-from-the-null-distribution}.
+Hence, 36 lineups with all combinations of $n$ and $X_{raw}$ and three replications are included in experiment III.
+In these lineups, the data of the data plot is generated from a model \svp{with zero effect size}, while the data of the 19 null plots are generated using the same simulation method discussed in Section \ref{sampling-from-the-null-distribution}.
 \svp{This generation procedure differs from the canonical Rorschach lineup procedure, which requires that all 20 plots are generated from the null hypothesis.
@@
 <!-- Technically, this was an alternative to Rorschach lineup but with the null distribution further conditioning on the estimated coefficients and the residual sum of squares.  -->
-\svp{To account for the fact that our simulation method for these lineups is not the Rorschach procedure,} we used the method suggested in @vanderplas2021statistical for typical lineups containing a data plot to estimate $\alpha$.
+\svp{To account for the fact that our simulation method for these lineups is not the Rorschach procedure,} we use the method suggested in @vanderplas2021statistical for typical lineups containing a data plot to estimate $\alpha$.
 We have included a sensitivity analysis in the Appendix to examine the impact of the variance of the $\alpha$ estimate on our findings.
@@
 However, presenting only these lineups to subjects are considered to be bad practices as subjects will lose interest quickly. 
-We planned to also collect 6 more evaluations on the 279 lineups with $X_{raw} \sim U(-1,1)$, resulted in $(36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3) \div (20-2) = 133$ subjects recruited for experiment III.
-
-<!-- TODO:subject allocation algorithm -->
+Therefore, we plan to collect 6 more evaluations on the 279 lineups with $X_{raw} \sim U(-1,1)$, result in $(36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3) / (20-2) = 133$ subjects recruited for experiment III.
 
@@
 
-Subjects for all three experiments were recruited from an crowdsourcing platform called Prolific [palan2018prolific].
-Prescreening procedure was applied during the recruitment, subjects were required to be fluent in English, with $98\%$ minimum approval rate in other studies and 10 minimum submissions.
-During the experiment, every subject was presented with a block of 20 lineups.
-For each lineup, the actual data plot was drawn as a standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis.
-An additional horizontal red line was added at $y = 0$ as a helping line.
-The 19 null datasets were generated by the residual rotation technique, and plotted in the same way.
-The lineup consisted of 20 residual plots with one randomly placed actual data plot.
-And for every lineup, the subject was asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots were from others.
-If there was no noticeable difference between plots in a lineup, subjects were permitted to select zero plots without providing the reason.
-No subject was shown the same lineup twice.
-Information about preferred pronoun, age group, education, and previous experience in visual experiment were also collected.
-A subject's submission was only accepted if the actual data plot was identified for at least one attention check.
-Data of rejected submissions were discarded automatically to maintain the overall data quality.
-
-<!-- TODO:pipeline description -->
+Subjects for all three experiments are recruited from an crowdsourcing platform called Prolific [@palan2018prolific].
+Prescreening procedure is applied during the recruitment, subjects are required to be fluent in English, with $98\%$ minimum approval rate and 10 minimum submissions in other studies.
+
+During the experiment, every subject is presented with a block of 20 lineups.
+A lineup consists of a randomly placed data plot and 19 null plots, which are all residual plots drawn with raw residuals on the y-axis and fitted values on the x-axis.
+An additional horizontal red line is added at $y = 0$ as a helping line.
+
+The data of the data plot is simulated from one of two models described in Section \ref{simulating-departures-from-good-residuals}, while the data of the remaining 19 null plots are generated by the residual rotation technique discussed in Section \ref{sampling-from-the-null-distribution}.
+
+In every lineup evaluation, the subject is asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots are from others.
+If there is no noticeable difference between plots in a lineup, subjects are permitted to select zero plots without providing the reason.
+No subject are shown the same lineup twice.
+Information about preferred pronoun, age group, education, and previous experience in visual experiment are also collected.
+A subject's submission is only accepted if the data plot is identified for at least one attention check.
+Data of rejected submissions are discarded automatically to maintain the overall data quality.
 
@@
 
-There were 2880, 2880 and 2214 lineups evaluation made by 160, 160 and 123 subjects recruited for experiment I, II and III respectively.
-In the total of 7974 lineup evaluations, 3744 used lineups produced by the non-linearity model, 3510 used lineups produced by the heteroskedasticity model, and 720 used lineups consists of only null plots.
-
-Besides, there were 886 attention checks not included in the following analysis. The collated dataset is provided in `vi_survey` of the `visage` `R` package.
+There are 2880, 2700 and 2394 lineups evaluation made by 160, 150 and 133 subjects recruited for experiment I, II and III respectively.
+In the total of 7974 lineup evaluations, 3744 use lineups produced by the non-linearity model, 3510 use lineups produced by the heteroskedasticity model, and 720 use null lineups.
+Besides, there are 886 attention checks not included in the following analysis.
+The collated dataset is provided in `vi_survey` of the `visage` `R` package.
 
@@
 
-Figure \ref{fig:polypower} shows the estimated power of visual test on lineups produced by the non-linearity model with $X_{raw} \sim U(-1,1)$, against the natural logarithm of the effect $log_e(E)$.
+Figure \ref{fig:polypower} shows the estimated power of visual test on lineups produced by the non-linearity model with $X_{raw} \sim U(-1,1)$, against the natural logarithm of the effect $log_e(E)$, with a 5% significance level.
+Lineups with $X_{raw} \sim U(-1,1)$ are focused in this section because visual patterns are more likely to be revealed under the uniform distribution.
+Besides, we have collected more responses for this distribution in experiment III to produce more reliable and stable power curves.
+Power analysis of other distributions can be found in Section [?].
+At the bottom of the figure \ref{fig:polypower}, there are a sequence of example residual plots with increasing levels of $log_e(E)$.
+Readers can evaluate them from left to right and determine at which level the departure from a good residual plot becomes detectable.
 
@@
 
-To compare the power of visual test and conventional test, we picked RESET test (`resettest`) and Breusch-Pagan test (`bptest`) from the R package `lmtest`, and Shapiro-Wilk test (`shapiro.test`) from the built-in R package `stats`.
+To compare the power of visual test and conventional test, we pick RESET test (`resettest`) and Breusch-Pagan test (`bptest`) from the R package `lmtest`, and Shapiro-Wilk test (`shapiro.test`) from the built-in R package `stats`.
 Among them, RESET test is the only exact and appropriate test in this scenario.
@@
 Their estimated power is shown in Figure \ref{fig:polypower}.
-For the RESET test, we included different power of fitted values as proxies.
-According to @ramsey_tests_1969, there is no general rule for the power of the fitted values needed by the RESET test, but it found power up to four is usually sufficient.
-Thus, we followed this guideline to conduct the RESET test.
-For the Breusch-Pagan test, both $\boldsymbol{x}$ and $\boldsymbol{x}^2$ were included in the auxiliary regression equation to [justify this choice?].
-
-At the bottom of the figure \ref{fig:polypower}, there are a sequence of example residuals plots with increasing levels of $log_e(E)$.
-Readers can evaluate them from left to right and determine at which level the departure from a good residual plot becomes detectable.
+To set up the RESET test, we include different powers of fitted values as proxies.
+According to @ramsey_tests_1969, there are no general rules for the power of the fitted values needed by the RESET test, but it finds power up to four is usually sufficient.
+Thus, we follow this guideline to conduct the RESET test.
+For the Breusch-Pagan test, the choice of regressors in the auxiliary regression is left to the user [@breusch_simple_1979].
+But as @waldman1983note suggested, it is a good choice for the set of auxiliary regressors in the Breusch-Pagan test be the same as the White test.
+Thus, we include both $\boldsymbol{x}$ and $\boldsymbol{x}^2$ in the auxiliary regression.
 
@@
 
-For the non-linearity model, the power of all four tests increases as the effect gets larger.
-The power curve of RESET test climbs aggressively from 13% to around 70% as $log_e(E)$ increases from 0 to 2, while power of other tests respond inactively to the change of effect and remain lower than 25% throughout the period, showing that REST test is way more sensitive to the type of model defects that being considered.
-Meanwhile, no noticeable visual features can be spotted from the example residuals plots.
+For the non-linearity model, the power curve of RESET test climbs aggressively from 13% to around 70% as $log_e(E)$ increases from 0 to 2, while power of other tests respond inactively to the change of effect, showing that REST test is way more sensitive to the type of model defects that being considered.
+Meanwhile, no noticeable visual features can be spotted from the example residual plots.
 
 In terms of the heteroskedasticity model, the power of Breusch-Pagan test is also almost always greater than the power of visual test.
-For $0 \leq log_e(E) \leq 2$, where the visual feature is nearly unobservable from the example residual plots and the power curve of the visual test remains at a low level, the Breusch-Pagan test still have a decent amount of chance of rejecting $H_0$.
-
-The power of visual test arises steadily as $log_e(E)$ increases from 2 to 5 in both non-linearity model and heteroskedasticity model, suggesting that the effect starts to make significant impact on the degree of the presence of the designed visual features.
-This can also be observed from the example residuals plots that when $log_e(E) = 2.5$, a weak "S-shape" and a weak "triangle" shape are presented in Figure \ref{fig:polypower} and Figure \ref{fig:heterpower} respectively.
-And as $log_e(E)$ increases, the visual pattern becomes much clearer, and the power reaches almost 100% at $log_e(E) \approx 6$.
-
-The power of all inappropriate tests except for F-test shows improvement as the effect increases but at a lower rate than the visual test in both scenarios.
+For $0 \leq log_e(E) \leq 2$, where the power curve of the visual test remains at a low level, the Breusch-Pagan test still have a decent amount of chance of rejecting $H_0$.
+Similarly, the visual feature is nearly unobservable from the example residual plots.
+
+The power of visual test arises steadily as $log_e(E)$ increases from 2 to 5 for both non-linearity model and heteroskedasticity model, suggesting that the effect starts to make significant impact on the degree of the presence of the designed visual features.
+This can also be observed from the example residual plots that when $log_e(E) = 2.5$, a weak "S-shape" and a weak "triangle" shape are presented in Figure \ref{fig:polypower} and Figure \ref{fig:heterpower} respectively.
+The visual pattern becomes much clearer as $log_e(E)$ increases.
+At $log_e(E) \approx 6$, the power reaches almost 100%.
+
+The power of all inappropriate tests except for RESET test shows improvement as the effect increases but at a lower rate than the visual test in both scenarios.
 This coincides the point made by @cook1982residuals that residual-based tests for a specific type of model defect may be sensitive to other types of model defects.
-The power curve of F-test remains at around 5% in Figure \ref{fig:heterpower} since there is no non-linear term leave out in the heteroskedasticity model and
-$H_0$ of the F-test will be always satisfied.
-
-(remove human perception)
-
-(revisit Emi's concern)
-
-Overall, the power analysis suggests that conventional tests do not match how humans perceive residual plots in these two scenarios.
-Humans have much higher tolerance of the residual departures than the conventional test.
-Since fail to reject $H_0$ in a visual test means that there is no obvious visual discoveries found in the residual plot, analysts and the general public as the consumers of the output may not be convinced of the existence of significant residual departures in spite of the rejection of $H_0$ given by the conventional test.
+The power curve of RESET test remains at around 5% in Figure \ref{fig:heterpower} since there are no non-linear terms leave out in the heteroskedasticity model and $H_0$ of the test is always satisfied.
+
+Overall, the power comparison suggests that conventional tests differs significantly from visual tests in these two scenarios.
+Visual test have much higher tolerance of the residual departures than the conventional test.
+Since fail to reject $H_0$ in a visual test usually means that there are no obvious visual discoveries found in the residual plot, analysts and the general public as the consumers of the output may not be convinced of the existence of significant residual departures in spite of the rejection of $H_0$ given by the conventional test.
 Even if the rejection is accepted, the model violation may be considered as impactless due to the fact that they are not clearly visible.
@@
 The rejection of $H_0$ because of human acceptable and negligible residual departures is not practically meaningful and useful.
-(it does not fit the purpose of )
-
-This could limit the popularity of conventional tests in residual diagnostics among analysts.
-
-<!-- TODO:difference between the simple formula and the simulation approach -->
+This may limit the popularity of conventional tests in residual diagnostics among analysts.
 
@@
   labs(col = "", size = "# lineups")
-
-set.seed(10085)
+```
+
+```{r polypower, fig.height = 3, fig.width = 5, fig.cap="Comparison of power between different tests for non-linear patterns.
+Main plot shows the power curves, with dots indicating human evaluations of lineups.
+Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
+Where would you draw the line of too much non-linearity in the residuals?
+For the RESET test this is around log effect size 1.5, but for the visual test it is around 3.5."}
+set.seed(106)
 ex_dat <- poly_model(shape = 2, x = rand_uniform(), sigma = 1)$gen(n = 100)
@@
               remove_axis = TRUE, 
-              remove_grid_line = TRUE, size=0.2) +
+              remove_grid_line = TRUE, size = 0.05) +
     ggtitle(round(es[j], 1)) +
@@
 }
-```
-
-<!-- The impact of the distribution of $X_{raw}$ on the power is displayed in Figure \ref{fig:dist-power}.
-For the non-linearity model, the power curve of F-test is stable across different distributions, while the visual test has a more gentle power curve for the discrete uniform distribution as it contains lineups with larger effect size.
-BP test performs worse for discrete uniform distribution but has relatively high power for normal distribution.
-SW test outperforms BP test for discrete uniform distribution but remains as the worst test for other distributions.
-The results indicate those inappropriate residual-based tests are sensitive to the distribution of the regressor.
--->
-
-<!-- For the heteroskedasticity model, F-test is entirely insensitive to the non-constant variance. The visual test has similar power as the SW test   -->
-<!-- for normal distribution and lognormal distribution, and has higher power for uniform distribution, and lower power for discrete uniform distribution. -->
-
-<!-- But still, we observe the huge gap in power between the preferred conventional tests and the visual test which is consistent with our previous argument that conventional test is too sensitive for diagnosing non-linearity and heteroskedasticity.
--->
-
-```{r polypower, fig.height = 3, fig.width = 5, fig.cap="Comparison of power between different tests for non-linear patterns.
-Main plot shows the power curves, with dots indicating human evaluations of lineups.
-Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
-Where would you draw the line of too much non-linearity in the residuals?
-For the RESET test this is around log effect size 1, but for the visual test it is around 3."}
+
 patchwork::wrap_plots(append(plist, list(p), 
@@
 es <- map_dbl(target_b, ~log(heter_model(a = 1, b = .x, x = rand_uniform(-1, 1))$effect_size(n = 100, type = "kl")))
-```
-
-```{r eval = FALSE}
-
-# For checking RESET test
-plotly::ggplotly(ggplot() +
-geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
-                filter(x_dist == "uniform") %>%
-                # select(-(RESET4_p_value:RESET10_p_value)) %>%
-                pivot_longer(F_p_value:RESET10_p_value) %>%
-                mutate(name = gsub("_p_value", " test", name)), 
-              aes(log_effect_size, as.numeric(value < 0.05), col = factor(name, levels = c(paste0("RESET", 3:10, " test"), "F test"))), 
-              method = "glm", 
-              method.args = list(family = binomial),
-              se = FALSE) +
-  scale_color_brewer("YlGnBu") +
-  labs(col = "") +
-  xlab("Log of effect size") +
-  ylab("Power") +
-  ggtitle("Power of the RESET test under different powers of the fitted values"))
 ```
@@
   labs(col = "", size = "# lineups")
-
-set.seed(10089)
+```
+
+```{r heterpower, fig.height = 3, fig.width = 5, fig.cap="Comparison of power between different tests for heteroskedasticity patterns.
+Main plot shows the power curves, with dots indicating human evaluations of lineups.
+Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
+Where would you draw the line of too much heteroskedasticity in the residuals?
+For the BP test this is around log effect size 1.5, but for the visual test it is around 3."}
+set.seed(10091)
 ex_dat <- heter_model(a = 1, x = rand_uniform(-1, 1), b = 1)$gen(n = 100)
@@
               remove_axis = TRUE,
-              remove_grid_line = TRUE, size=0.2) +
+              remove_grid_line = TRUE, size = 0.05) +
     ggtitle(round(es[j], 1)) +
@@
 }
-```
-
-```{r heterpower, fig.height = 3, fig.width = 5, fig.cap="Comparison of power between different tests for heteroskedasticity patterns.
-Main plot shows the power curves, with dots indicating human evaluations of lineups.
-Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
-Where would you draw the line of too much heteroskedasticity in the residuals?
-For the BP test this is around log effect size 1.5, but for the visual test it is around 3."}
+
 patchwork::wrap_plots(append(plist, list(p), after = 0), 
@@
 
-<!-- ## Visual test  -->
-
-<!-- Note that both the p-value and power comparison are not directly based on the diagnostic plots recommended by abundance of literature, but  -->
-
 ## Comparison of rejection rates based on $p$-values
 
-(1. How you calc rejection, 2. scatter plot p-value vs p-value -> appendix, 3. percentage in explained the tex, both marginal and conditional, 4. interesting lineup, discuss the reason)
-
-(style similar to mosaic plot)
-
-(lineups detected by visual test but not by conventional tests)
-
-Visual inference and conventional hypothesis testing are two distinct statistical procedures.
-As displayed in Figure \ref{fig:p-value-comparison}, the visual test gives higher p-values than the conventional test for both polynomial and heteroskedasticity models in most of the time, suggesting that humans have lower confidence and much higher tolerance of the residual departures than the conventional test.
-
-In particular, given 95% confidence level, both tests will reject $H_0$ in only around 58% of the cases.
-And in 41% of the cases, the visual test does not reject $H_0$ but the conventional test does.
-Since fail to reject $H_0$ in a visual test means that there is no obvious visual discoveries found in the residual plot, analysts and the general public as the consumers of the output may not fully believe in the rejection of $H_0$ even though the alternative hypothesis is true.
-Even if the rejection is accepted, the model violation may not be considered as impactful due to the fact that the departures are not clearly visible.
-
-This makes it sounds like people are arrogantly ignoring the accuracy of conventional tests and trust too much to their visual perception, but it is not.
-In fact, the sensitivity of the conventional test could sometimes distract and discourage analysts from finding a simple but good linear approximation to the data.
-The rejection of $H_0$ because of acceptable and negligible residual departure is not practically meaningful and useful.
-This also partly explains why abundance of literature suggesting the use of data plot in regression diagnostics.
+The power comparison illustrates that appropriate conventional tests reject $H_0$ more often than visual tests.
+In this section, we explore how often they agree with each other by investigate the rejection rates based on $p$-values.
+
+Depending on the natural logarithm of the effect, lineups are classified into three categories: easy (< 1.5), moderate (1.5 - 3.5) and difficult (> 3.5).
+Figure \ref{fig:p-value-comparison} provides a mosaic plot showing the rejection rate of visual tests and conventional tests at different difficulty levels.
+
+### Easy lineup
+
+For easy lineups produced by the non-linearity model, 65% are rejected by both the conventional and visual tests, while 9% are accepted by both tests.
+Thus, 74% of the time two tests will agree with each other.
+There are 16% of lineups rejected only by the conventional test.
+
+Interestingly, in spite of the greater power of the conventional test, 10% of the lineups are rejected only by the visual test.
+This means among those accepted by the conventional test, 51% are rejected by the visual test.
+Recall that the RESET test used in our analysis includes powers of fitted values up to four.
+However, the "M" and "Triple-U" shapes constructed from the Hermite polynomials contain powers of regressor up to six and 18 respectively.
+In fact, by including the powers of fitted values up to six, all cases will be rejected by the RESET test.
+And from the example plot displayed in Figure \ref{fig:easy-poly-example}, we can observe panel three is distinctly different from others exhibiting a "M" shape.
+It clearly suggests users of the RESET test and any other residual-based conventional tests that require the specification of variables of interest to check the residual plot before conducting the test, such that the correct choice of variables can be made.
+
+```{r easy-poly-example, fig.height = 5, fig.cap = "An example easy lineup produced by the non-linearity model not rejected by the conventional test but rejected by the visual test.
+Panel three exhibits strong non-linear pattern."}
+VI_MODEL$plot_lineup(vi_lineup$poly_270$data, 
+                     remove_grid_line = TRUE, 
+                     theme = theme_light(),
+                     remove_axis = TRUE)
+```
+
+```{r easy-heter-example1, fig.height = 5, fig.cap = "An example easy lineup produced by the heteroskedasticity not rejected by the conventional test but rejected by the visual test.
+Panel 17 exhibits strong heteroskedasticity pattern."}
+VI_MODEL$plot_lineup(vi_lineup$heter_331$data, 
+                     remove_grid_line = TRUE, 
+                     theme = theme_light(),
+                     remove_axis = TRUE)
+```
+
+```{r easy-heter-example2, fig.height = 5, fig.cap = "An example easy lineup produced by the heteroskedasticity not rejected by the conventional test but rejected by the visual test.
+Panel 18 contains a distinct outlier."}
+VI_MODEL$plot_lineup(vi_lineup$heter_536$data, 
+                     remove_grid_line = TRUE, 
+                     theme = theme_light(),
+                     remove_axis = TRUE)
+```
+
+For easy lineups produced by the heteroskedasticity model, 62% are rejected by both the conventional test and the visual test, and 7% are accepted by both, result in 69% agreement rate.
+Lineups rejected by only conventional test are 30%.
+1% of lineups are accepted by the conventional test but rejected by the visual test.
+
+Among those 1% lineups, which are two lineups, one shows detectable heteroskedasticity pattern as displayed in Figure \ref{fig:easy-heter-example1}.
+Another contain data plot with an noticeable outlier, but the "Right-triangle" shape is not obvious as shown in Figure \ref{fig:easy-heter-example2}.
+The $p$-values of Breusch-Pagan test for these two lineups are $0.055$ and $0.135$.
+Compared to the visual test, its $p$-values are $0.026$ and $0.012$.
+The Breusch-Pagan test should reject the one with obvious heteroskedasticity signal, but it yields a $p$-value slightly above 5%.
+
+### Moderate lineup
+
+With the increase of the difficulty level, the sensitivity of conventional tests enlarge the difference in rejection rates.
+For moderate lineup produced by the non-linearity model, only 8% and 28% are rejected and accepted by both respectively.
+We can observe a decline in agreement rate to 36%.
+A huge amount of lineups are accepted by the visual test.
+63% of lineups rejected only by the conventional test.
+There are 1% of lineups which rejected only by conventional tests suffer the same RESET test issue as mentioned earlier.
+
+For those produced by the heteroskedasticy model, the corresponding numbers are 23%, 31%, 54%, 45% and 1%.
+Only one lineup rejected by the visual test but accepted by the conventional test.
+It shows zero heteroskedasticity patterns, but contains a few outliers at the bottom of the data plot as illustrated in panel 14 of Figure \ref{fig:moderate-heter-example1}.
+Two out of five subjects correctly identify this lineup yielding a $p$-value equal to $0.044$.
+
+```{r moderate-heter-example1, fig.height = 5, fig.cap = "An example easy lineup produced by the heteroskedasticity not rejected by the conventional test but rejected by the visual test.
+Panel 14 contains a few outliers at the bottom of the panel."}
+VI_MODEL$plot_lineup(vi_lineup$heter_219$data, 
+                     remove_grid_line = TRUE, 
+                     theme = theme_light(),
+                     remove_axis = TRUE)
+```
+
+### Difficult lineup
+
+Visual test is insensitive to lineups with small effect.
+When it comes to difficult lineups produced by the non-linearity model, none of them are rejected by the visual test.
+There are 73% of lineups accepted by both tests, and 26% rejected by the conventional test.
+The agreement rate rise up to 73%.
+
+For lineups produced by the heteroskedasticity model, 3% are rejected by both, 71% are accepted by both, 24% are rejected only by the conventional test and 2% are rejected only by the visual test, which are two lineups.
+
+As shown in \ref{fig:difficult-heter-example1} and \ref{fig:difficult-heter-example2}, residual plots of both lineups are constructed with $X_{raw} \sim lognormal(0, 0.6^2)/3$.
+Though the designed patterns are "Left-triangle" and "Butterfly", both of them are invisible due to the small effect.
+The residual plots are dominated by the skewed distribution.
+Further, no distinguishable visual patterns, such as outliers, are detectable from the data plot of these two lineups, which are panel 5 and panel 15.
+The data plot in these two lineups are not the most different plot from our point of view either.
+However, three out of five and two out of five subjects correctly identify the data plot for these two lineups respectively.
+But the reasons for selecting the plot provided by the subjects are not consistent.
+Therefore, with the current data, we can not explain why subjects can correctly identify the data plot of these two lineups.
+Collecting more data in future study may help to answer this question.
+
+```{r difficult-heter-example1, fig.height = 5, fig.cap = "An example easy lineup produced by the heteroskedasticity not rejected by the conventional test but rejected by the visual test.
+Panel 14 contains a few outliers at the bottom of the panel."}
+VI_MODEL$plot_lineup(vi_lineup$heter_224$data, 
+                     remove_grid_line = TRUE, 
+                     theme = theme_light(),
+                     remove_axis = TRUE)
+```
+
+```{r difficult-heter-example2, fig.height = 5, fig.cap = "An example easy lineup produced by the heteroskedasticity not rejected by the conventional test but rejected by the visual test.
+Panel 14 contains a few outliers at the bottom of the panel."}
+VI_MODEL$plot_lineup(vi_lineup$heter_415$data, 
+                     remove_grid_line = TRUE, 
+                     theme = theme_light(),
+                     remove_axis = TRUE)
+```
+
+```{r eval = FALSE}
+
+vi_survey %>% filter(unique_lineup_id %in% 
+                       (p_value_cmp_dat %>%
+  filter(type == "heteroskedasticity", difficulty == "difficult") %>%
+  filter(conventional_p_value >= 0.05, p_value < 0.05) %>%
+  pull(unique_lineup_id))) %>%
+  View()
+
+```
+
+```{r eval = FALSE}
+p_value_cmp_dat %>%
+  filter(type == "non-linearity", difficulty == "difficult") %>%
+  count(conventional_p_value < 0.05, p_value < 0.05) %>%
+  mutate(n = n / sum(n))
+
+p_value_cmp_dat %>%
+  filter(type == "heteroskedasticity", difficulty == "difficult") %>%
+  count(conventional_p_value < 0.05, p_value < 0.05) %>%
+  mutate(n = n / sum(n)) 
+```
+
+```{r p-value-comparison, fig.width=10, fig.height=5, fig.cap="Rejection rate of visual test compared to conventional test on lineups produced by heteroskedasticity and non-linearity model at different levels of difficulty.
+Based on the rejection decision made by the visual test and the conventional test, the mosaic plot shows the proportion of lineups in each category."}
+
+library(ggmosaic)
+
+p_value_cmp_dat <- vi_survey %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  mutate(difficulty = c("difficult", "moderate", "easy")[1 + 
+                                                           (log(effect_size) > 1.5) +
+                                                           (log(effect_size) > 3.5)]) %>%
+  mutate(difficulty = factor(difficulty, levels = c("easy", "moderate", "difficult"))) %>%
+  left_join(vi_survey %>% 
+           count(unique_lineup_id) %>%
+           rename(num_subject = n)) %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
+  mutate(conv_reject = ifelse(conventional_p_value < 0.05, "Reject", "Not"),
+         reject = ifelse(p_value < 0.05, "Reject", "Not")) %>%
+  mutate(conv_reject = factor(conv_reject, levels = c("Reject", "Not")),
+         reject = factor(reject, levels = c("Reject", "Not")))
+
+p_value_cmp_dat %>%
+  filter(difficulty == "easy") %>%
+  ggplot() +
+  ggmosaic::geom_mosaic(aes(x = ggmosaic::product(reject, conv_reject), fill = reject)) +
+  facet_grid(~type) +
+  ylab("Visual tests") +
+  xlab("Conventional tests") +
+  labs(fill = "Conventional tests") +
+  scale_fill_brewer("", palette="Dark2") +
+  theme_bw() +
+  theme(legend.position = "none") +
+  ggtitle("Easy") -> p1
+
+p_value_cmp_dat %>%
+  filter(difficulty == "moderate") %>%
+  ggplot() +
+  ggmosaic::geom_mosaic(aes(x = ggmosaic::product(reject, conv_reject), fill = reject)) +
+  facet_grid(~type) +
+  ylab("Visual tests") +
+  xlab("Conventional tests") +
+  labs(fill = "Conventional tests") +
+  scale_fill_brewer("", palette="Dark2") +
+  theme_bw() +
+  theme(legend.position = "none") +
+  ylab("") +
+  ggtitle("Moderate") -> p2
+
+p_value_cmp_dat %>%
+  filter(difficulty == "difficult") %>%
+  ggplot() +
+  ggmosaic::geom_mosaic(aes(x = ggmosaic::product(reject, conv_reject), fill = reject)) +
+  facet_grid(~type) +
+  ylab("Visual tests") +
+  xlab("Conventional tests") +
+  labs(fill = "Conventional tests") +
+  scale_fill_brewer("", palette="Dark2") +
+  theme_bw() +
+  theme(legend.position = "none") +
+  ylab("") +
+  ggtitle("Difficult") -> p3
+
+patchwork::wrap_plots(p1, p2, p3)
+```
+
+```{r eval = FALSE}
+vi_survey %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  left_join(vi_survey %>% 
+           count(unique_lineup_id) %>%
+           rename(num_subject = n)) %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
+  mutate(conv_reject = ifelse(conventional_p_value < 0.05, "Reject", "Not"),
+         reject = ifelse(p_value < 0.05, "Reject", "Not")) %>%
+  filter(type == "non-linearity", conv_reject == "Not", reject == "Reject") %>%
+  View()
+```
+
+## Effect of the distribution of the independent variable
+
+# change of x_dist -> power
+
+# appendix (contains all the details)
 
@@
 
-```{r p-value-comparison, fig.height=4, fig.cap="Rejection rate of visual test compared to conventional test on lineups produced by heteroskedasticity and non-linearity model.
-Based on the rejection decision made by the visual test and the conventional test, the mosaic plot shows the proportion of lineups in each category."}
-
-library(ggmosaic)
-vi_survey %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
-  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-  left_join(vi_survey %>% 
-           count(unique_lineup_id) %>%
-           rename(num_subject = n)) %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
-  mutate(conv_reject = ifelse(conventional_p_value < 0.05, "Reject", "Not"),
-         reject = ifelse(p_value < 0.05, "Reject", "Not")) %>%
-  mutate(conv_reject = factor(conv_reject, levels = c("Reject", "Not")),
-         reject = factor(reject, levels = c("Reject", "Not"))) %>%
-  ggplot() +
-  ggmosaic::geom_mosaic(aes(x = ggmosaic::product(reject, conv_reject), fill = reject)) +
-  facet_wrap(~type) +
-  ylab("Visual tests") +
-  xlab("Conventional tests") +
-  labs(fill = "Conventional tests") +
-  scale_fill_brewer("", palette="Dark2") +
-  theme_bw() +
-  theme(legend.position="none")
-```
-
-```{r eval = FALSE}
-vi_survey %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
-  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-  left_join(vi_survey %>% 
-           count(unique_lineup_id) %>%
-           rename(num_subject = n)) %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
-  mutate(conv_reject = ifelse(conventional_p_value < 0.05, "Reject", "Not"),
-         reject = ifelse(p_value < 0.05, "Reject", "Not")) %>%
-  filter(type == "non-linearity", conv_reject == "Not", reject == "Reject") %>%
-  View()
-```
-
-## Effect of the distribution of the independent variable
-
-<!-- ```{r cache = TRUE} -->
-<!-- vi_survey_with_plot <- vi_survey %>% -->
-<!--   filter(!attention_check) %>% -->
-<!--   filter(!null_lineup) %>% -->
-<!--   left_join(map_df(1:length(vi_lineup),  -->
-<!--     ~vi_lineup[[.x]]$data %>%  -->
-<!--       group_by(k) %>%  -->
-<!--       summarise(plot_p_value = first(p_value)) %>% -->
-<!--       arrange(k) %>% -->
-<!--       mutate(unique_lineup_id = names(vi_lineup)[.x]) %>% -->
-<!--       mutate(plot_p_value_rank = rank(plot_p_value, ties.method = "min"))) %>% -->
-<!--   pivot_wider(values_from = c(plot_p_value, plot_p_value_rank), names_from = k, names_prefix = "plot_")) %>% -->
-<!--   (function(x) { -->
-<!--       for (i in 1:20) -->
-<!--         x[[paste0("plot_sel_plot_", i)]] <- -->
-<!--           grepl(paste0("_", i, "_"), x$selection) | -->
-<!--           grepl(paste0("^", i, "_"), x$selection) | -->
-<!--           grepl(paste0("_", i, "$"), x$selection) | -->
-<!--           grepl(paste0("^", i, "$"), x$selection) -->
-<!--       x} -->
-<!--     ) %>% -->
-<!--   pivot_longer(plot_p_value_plot_1:plot_sel_plot_20,  -->
-<!--                names_to = c(".value", "plot"),  -->
-<!--                names_pattern = "(.*)_plot_(.+)")  -->
-<!-- ``` -->
-
-<!-- ```{r plotsel, fig.cap = "A large amount of subjects select the residuals plot with the lowest p-value. But greater than 50 percent of subjects evenly select other plots."} -->
-<!-- vi_survey_with_plot %>% -->
-<!--   mutate(weighted_sel = ifelse(num_selection == 0, 0.05, plot_sel/num_selection)) %>% -->
-<!--   group_by(type) %>% -->
-<!--   mutate(n_eval = n()/20) %>% -->
-<!--   ungroup() %>% -->
-<!--   group_by(type, plot_p_value_rank) %>% -->
-<!--   summarise(weighted_sel = sum(weighted_sel), n_eval = first(n_eval)) %>% -->
-<!--   ggplot() + -->
-<!--   geom_col(aes(plot_p_value_rank, n_eval), fill = "grey", alpha = 0.3) + -->
-<!--   geom_col(aes(plot_p_value_rank, weighted_sel)) + -->
-<!--   facet_wrap(~type) + -->
-<!--   theme_light() + -->
-<!--   ylab("Weighted number of selections") + -->
-<!--   xlab("Plot p-value rank") -->
-<!-- ``` -->
-
-<!-- ```{r dataplotsel, fig.cap = "..."} -->
-<!-- vi_survey_with_plot %>% -->
-<!--   filter(!attention_check) %>% -->
-<!--   filter(!null_lineup) %>% -->
-<!--   filter(answer == plot) %>% -->
-<!--   group_by(unique_lineup_id) %>% -->
-<!--   summarise(plot_p_value_rank = first(plot_p_value_rank), -->
-<!--             weighted_detect = sum(weighted_detect), -->
-<!--             n_eval = n()) %>% -->
-<!--   group_by(plot_p_value_rank) %>% -->
-<!--   summarise(weighted_detect = sum(weighted_detect), -->
-<!--             n_eval = sum(n_eval)) %>% -->
-<!--   mutate(plot_p_value_rank_g5 = plot_p_value_rank >= 5) %>% -->
-<!--   group_by(plot_p_value_rank_g5) %>% -->
-<!--   mutate(n_eval = ifelse(plot_p_value_rank_g5, sum(n_eval), n_eval)) %>% -->
-<!--   mutate(weighted_detect = ifelse(plot_p_value_rank_g5, sum(weighted_detect), weighted_detect)) %>% -->
-<!--   ungroup() %>% -->
-<!--   filter(plot_p_value_rank <= 5) %>% -->
-<!--   mutate(plot_p_value_rank = ifelse(plot_p_value_rank == 5, "5+", as.character(plot_p_value_rank))) %>% -->
-<!--   mutate(plot_p_value_rank = factor(plot_p_value_rank, levels = c(as.character(1:5), "5+"))) %>% -->
-<!--   ggplot() + -->
-<!--   geom_col(aes(plot_p_value_rank, n_eval)) + -->
-<!--   geom_col(aes(plot_p_value_rank, weighted_detect, fill = "weighted number of detections")) + -->
-<!--   xlab("Data plot p-value rank") + -->
-<!--   ylab("Number of Lineups") + -->
-<!--   theme(legend.position = "bottom") + -->
-<!--   labs(fill = "") -->
-<!-- ``` -->
-
-<!-- ```{r dadtaplotselfacet, fig.cap = "...", fig.height=12, fig.width=16} -->
-<!-- vi_survey_with_plot %>% -->
-<!--   filter(!attention_check) %>% -->
-<!--   filter(!null_lineup) %>% -->
-<!--   filter(type == "polynomial") %>% -->
-<!--   left_join(vi_survey_with_plot %>% -->
-<!--               filter(!attention_check) %>% -->
-<!--               filter(!null_lineup) %>% -->
-<!--               filter(answer == plot) %>% -->
-<!--               group_by(unique_lineup_id) %>% -->
-<!--               summarise(answer_rank = first(plot_p_value_rank))) %>% -->
-<!--   mutate(weighted_sel = ifelse(num_selection == 0, 0.05, plot_sel/num_selection)) %>% -->
-<!--   group_by(plot_p_value_rank, answer_rank) %>% -->
-<!--   summarise(weighted_sel = sum(weighted_sel)) %>% -->
-<!--   mutate(answer_rank_g5 = answer_rank >= 5) %>% -->
-<!--   group_by(plot_p_value_rank, answer_rank_g5) %>% -->
-<!--   mutate(weighted_sel = ifelse(answer_rank_g5, sum(weighted_sel), weighted_sel)) %>% -->
-<!--   ungroup() %>% -->
-<!--   filter(answer_rank <= 5) %>% -->
-<!--   mutate(answer_rank = ifelse(answer_rank == 5, "5+", as.character(answer_rank))) %>% -->
-<!--   mutate(answer_rank = factor(answer_rank, levels = c(as.character(1:5), "5+"))) %>% -->
-<!--   ggplot() + -->
-<!--   geom_col(aes(factor(plot_p_value_rank), weighted_sel)) + -->
-<!--   geom_vline(data = data.frame(answer_rank = c("1", "2", "3", "4"), x = 1:4), -->
-<!--              aes(xintercept = x), -->
-<!--              col = "red") + -->
-<!--   facet_wrap(~answer_rank, scales = "free_y") + -->
-<!--   xlab("Plot p-value rank") + -->
-<!--   ylab("Weighted number of selections") -->
-<!-- ``` -->
-
-<!-- ```{r eval = FALSE} -->
-<!-- vi_survey_with_plot %>% -->
-<!--   filter(!attention_check) %>% -->
-<!--   filter(!null_lineup) %>% -->
-<!--   filter(answer == plot) %>% -->
-<!--   group_by(unique_lineup_id) %>% -->
-<!--   summarise(plot_p_value_rank = first(plot_p_value_rank), -->
-<!--             weighted_detect = sum(weighted_detect), -->
-<!--             n_eval = n(), -->
-<!--             reject = first(conventional_p_value < 0.05)) %>% -->
-<!--   group_by(plot_p_value_rank, reject) %>% -->
-<!--   summarise(weighted_detect = sum(weighted_detect), -->
-<!--             n_eval = sum(n_eval)) %>% -->
-<!--   ggplot() + -->
-<!--   geom_col(aes(factor(plot_p_value_rank), n_eval)) + -->
-<!--   geom_col(aes(factor(plot_p_value_rank), weighted_detect, fill = "weighted number of detections")) + -->
-<!--   xlab("Data plot p-value rank") + -->
-<!--   ylab("Number of Lineups") + -->
-<!--   theme(legend.position = "bottom") + -->
-<!--   labs(fill = "") + -->
-<!--   facet_wrap(~reject) -->
-<!-- ``` -->
-
-<!-- ```{r eval = FALSE} -->
-<!-- # data plot not always has the lowest conventional p-value -->
-<!-- vi_survey_with_plot %>% -->
-<!--   filter(!attention_check) %>% -->
-<!--   filter(!null_lineup) %>% -->
-<!--   filter(answer == plot) %>% -->
-<!--   ggplot() + -->
-<!--   geom_bar(aes(factor(plot_p_value_rank))) -->
-
-<!-- # according to the rank of the data plot p-value, how many people select it  -->
-<!-- vi_survey_with_plot %>% -->
-<!--   filter(!attention_check) %>% -->
-<!--   filter(!null_lineup) %>% -->
-<!--   filter(plot_sel) %>% -->
-<!--   left_join(vi_survey_with_plot %>% -->
-<!--               filter(!attention_check) %>% -->
-<!--               filter(!null_lineup) %>% -->
-<!--               filter(answer == plot) %>% -->
-<!--               group_by(unique_lineup_id) %>% -->
-<!--               summarise(answer_rank = first(plot_p_value_rank))) %>% -->
-<!--   ggplot() + -->
-<!--   geom_bar(aes(factor(plot_p_value_rank))) + -->
-<!--   geom_vline(data = data.frame(answer_rank = 1:20, x = 1:20), -->
-<!--              aes(xintercept = x), -->
-<!--              col = "red") + -->
-<!--   facet_wrap(~answer_rank, scales = "free_y") -->
-
-<!-- # For different rank, percentage of subject select it -->
-<!-- vi_survey_with_plot %>% -->
-<!--   filter(!attention_check) %>% -->
-<!--   filter(!null_lineup) %>% -->
-<!--   filter(answer == plot) %>% -->
-<!--   group_by(unique_lineup_id, set) %>% -->
-<!--   summarise(plot_p_value_rank = first(plot_p_value_rank), -->
-<!--             detect = first(detect), -->
-<!--             type = first(type)) %>% -->
-<!--   ggplot() + -->
-<!--   geom_bar(aes(factor(plot_p_value_rank), fill = detect), position = "fill") + -->
-<!--   geom_hline(yintercept = 0.05) + -->
-<!--   xlab("Data plot p-value rank") + -->
-<!--   ylab("Propotion of subject detecting the data plot") -->
-
-<!-- vi_survey_with_plot %>% -->
-<!--   filter(!attention_check) %>% -->
-<!--   filter(!null_lineup) %>% -->
-<!--   filter(answer == plot) %>% -->
-<!--   group_by(unique_lineup_id, set) %>% -->
-<!--   summarise(plot_p_value_rank = first(plot_p_value_rank), -->
-<!--             prop_detect = first(prop_detect)) %>% -->
-<!--   ggplot() + -->
-<!--   geom_boxplot(aes(factor(plot_p_value_rank), prop_detect)) -->
-
-<!-- # vi_survey_with_plot %>% -->
-<!-- #   filter(!attention_check) %>% -->
-<!-- #   filter(!null_lineup) %>% -->
-<!-- #   filter(answer == plot) %>% -->
-<!-- #   group_by(unique_lineup_id, set) %>% -->
-<!-- #   summarise(plot_p_value_rank = first(plot_p_value_rank), -->
-<!-- #             conv_detect = first(conventional_p_value < 0.05)) %>% -->
-<!-- #   ggplot() + -->
-<!-- #   geom_bar(aes(factor(plot_p_value_rank), fill = conv_detect), position = "fill") + -->
-<!-- #   xlab("Data plot p-value rank") + -->
-<!-- #   ylab("Propotion of subject detecting the data plot") -->
-<!-- ``` -->
-
-# change of x_dist -> power
-
-# appendix (contains all the details)
-
 ### A collection interesting lineups (unusual results)

### 2023-01-09 paper_comparison/paper_comparison/paper_comparison.tex
major update

@@
 Diagnosing a model is the key to determining whether there is anything
-importantly wrong. For linear regression analysis, we typically
-interrogate the residuals for model diagnostics. Residuals summarise
-what is not captured by the model, and thus provide the capacity to
-identify what might be wrong.
-
-We can assess residuals in multiple ways. Residuals might be plotted, as
-a histogram or quantile-quantile plot to examine the distribution. Using
-the classical normal linear regression model as an example, if the
+importantly wrong. In linear regression analysis, residuals are
+typically examined for model diagnostics. Residuals summarise what is
+not captured by the model, and thus provide the capacity to identify
+what might be wrong.
+
+One can access residuals in multiple ways. Residuals might be plotted,
+as a histogram or quantile-quantile plot to examine the distribution.
+Using the classical normal linear regression model as an example, if the
 distribution is symmetric and unimodal, it is well-behaved. But if the
 distribution is skewed, bimodal, multimodal, or contains outliers, there
-is cause for concern. The distribution could also be inspected by
+is cause for concern. One could also inspect the distribution by
 conducting a goodness of fit test, such as the Shapiro-Wilk Normality
@@
 predicted values and each of the explanatory variables to scrutinize
-their relationships. If there is any visually discoverable patterns, the
-model is potentially misspecified. However, correctly judging a residual
-plot where no pattern exists is a painstakingly difficult task for
-humans. It is especially common, particularly among new data analytics
-students to report patterns when an experienced data analyst might
-quickly conclude that there are none. Generally, one looks for
+their relationships. If there are any visually discoverable patterns,
+the model is potentially misspecified. In general, one looks for
 noticeable departures from the model like non-linear dependency or
-heteroskedasticity. It is also possible to conduct hypothesis tests for
-non-linear dependence \citep{ramsey_tests_1969}, and use a Breusch-Pagan
-test \citep{breusch_simple_1979} for heteroskedasticity.
+heteroskedasticity. However, correctly judging a residual plot where no
+pattern exists is a painstakingly difficult task for humans (?citation).
+It is especially common, particularly among new data analysts, to report
+patterns when an experienced one might quickly conclude that there are
+none. It is also possible to conduct hypothesis tests for non-linear
+dependence \citep{ramsey_tests_1969}, and use a Breusch-Pagan test
+\citep{breusch_simple_1979} for heteroskedasticity.
 
@@
 sentiments like \emph{based on their experience, statistical tests are
-not widely used in regression diagnostics. The same or even larger
+not widely used in regression diagnostics since the same or even larger
 amount of information can be provided by diagnostic plots than the
@@
 describes the types of departures that one expects to detect, and
-describes a formal statistical process for reading residual plots,
-called visual inference. Section \ref{experiment-design} describes the
-experiment design to compare the decision made by formal hypothesis
+outlines a formal statistical process for reading residual plots, called
+visual inference. Section \ref{experimental-design} details the
+experimental design to compare the decision made by formal hypothesis
 testing, and how humans would read diagnostic plots. The results are
@@
 work, in particular how the responsibility for residual plot reading
-might be passed on to computer vision.
+might be relieved.
 
@@
 \citet{cook1982residuals}. As shown in Figure
-\ref{fig:residual-plot-common-departures}, the top-left panel is a good
+\ref{fig:residual-plot-common-departures}, the plot A is an ideal
 residual plot with residuals evenly distributed at both sides of the
-horizontal zero line showing no noticeable patterns.
-
-There are various types of departures from a good residual plot.
+horizontal zero line, with no noticeable patterns.
+
+There are various types of departures from an ideal residual plot.
 Non-linearity, heterskedasticity and non-normality are perhaps the three
@@
 Compared to non-linearity and heteroskedasticity, non-normality is
-usually harder to detect from a residual plot since scatter plot do not
-readily reveal the marginal distribution. A favourable graphical summary
-for this task is the quantile-quantile plot. For a consistent
-comparison, the residual plot of this departure is still presented in
-plot D of Figure \ref{fig:residual-plot-common-departures}. However,
-since the scatter plot is not suitable for this task, non-normality will
-not be the focus of this paper. Besides, it is important to note that
-not all regression models assume normality for the error term, but a
-certain amount do including the classical normal linear regression
+usually harder to detect from a residual plot since a scatter plot do
+not readily reveal the marginal distribution. A favourable graphical
+summary for this task is the quantile-quantile plot. Hence,
+non-normality will not be the focus of this paper as we mainly discuss
+residual plots. For a consistent comparison, the residual plot of this
+departure is still presented in plot D of Figure
+\ref{fig:residual-plot-common-departures}. Besides, it is important to
+note that not all regression models assume normality for the error term,
+but a certain amount do including the classical normal linear regression
 model. In the case that the normality assumption is violated, it is
@@
 and above the horizontal axis. For example, given a skewed error
-distribution, fewer data points and more outliers are on one side of the
-horizontal axis as shown in plot D of Figure
+distribution, there will be fewer data points and more outliers on one
+side of the horizontal axis as shown in plot D of Figure
 \ref{fig:residual-plot-common-departures}.
@@
 the Ramsey Regression Equation Specification Error Test (RESET)
-\citep{ramsey_tests_1969}. And for testing normality, the Shapiro--Wilk
-test \citep{shapiro1965analysis} is perhaps the most widely used test
-included by many of the statistical softwares. Another choice will be
-the Jarque--Bera test \citep{jarque1980efficient} which directly checks
-if the sample skewness and kurtosis match a normal distribution.
+\citep{ramsey_tests_1969}. The Shapiro-Wilk test
+\citep{shapiro1965analysis} is the most widely used test of
+non-normality included by many of the statistical softwares. The
+Jarque--Bera test \citep{jarque1980efficient} is also used to directly
+checks if the sample skewness and kurtosis match a normal distribution.
 
@@
 shown in Table \ref{tab:example-residual-plot-table}. In the example,
-both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null
+the Breusch-Pagan test and the Shapiro--Wilk test both reject the null
 hypothesis \(H_0\) for departures that they do not intend to examine. As
 discussed in \citet{cook1982residuals}, most residual-based tests for a
-particular type of departure from model assumptions are sensitive to
-other types of departures. It is likely \(H_0\) is correctly rejected
-but for the wrong reason, which is known as the ``Type III error''.
+particular type of departure from model assumptions are also sensitive
+to other types of departures. It is likely \(H_0\) is correctly rejected
+but for the wrong reason, a phenomenon known as the ``Type III error''.
 Additionally, outliers will often incorrectly trigger the rejection of
@@
 and indecisive compared to those rigorous statistical procedures as it
-relies on graphical perception - human's ability to interpret and decode
-the information embedded in the graph \citep{cleveland_graphical_1984}.
-It is particularly true that visual discovery suffers from its unsecured
-and unconfirmed nature where the degree of the presence of the visual
-features typically can not be measured quantitatively and objectively,
-which may lead to over or under-interpretations of the data. One such
-example is finding an over-interpretation of the separation between gene
-groups in a two-dimensional projection from a linear discriminant
-analysis when in fact there are no differences in the expression levels
-between the gene groups and separation is not an uncommon occurrence
+relies on graphical perception - human ability to interpret and decode
+the information embedded in graph \citep{cleveland_graphical_1984}.
+Further, the degree of the presence of the visual features typically can
+not be measured quantitatively and objectively, which may lead to over
+or under-interpretations of the data. For instance, people
+over-interpret the separation between gene groups in a two-dimensional
+projection from a linear discriminant analysis when in fact there are no
+differences in the expression levels between the gene groups and
+separation is not an uncommon occurrence
 \citep{roy_chowdhury_using_2015}.
@@
 by \citet{buja_inference_1999} as an idea to address the issue of valid
-inference for visual discoveries of data plots
-\citep{gelman_exploratory_2004}. Later, \citet{buja_statistical_2009}
-proposed the lineup protocol as a visual test inspired by the ``police
-lineup'' or ``identity parade'' which is the act of asking the
-eyewitness to identify criminal suspect from a group of irrelevant
-people. The protocol consists of \(m\) randomly placed plots, where one
-plot is the actual data plot, and the remaining \(m - 1\) plots have the
-identical graphical production as the data plot except the data has been
-replaced with data consistent with \(H_0\). Then, an observer who have
-not seen the actual data plot will be asked to point out the most
+inference for visual discoveries of data plots. Later,
+\citet{buja_statistical_2009} proposed the lineup protocol as a visual
+test inspired by the ``police lineup'' or ``identity parade'' which is
+the act of asking the eyewitness to identify criminal suspect from a
+group of irrelevant people. The protocol consists of \(m\) randomly
+placed plots, where one plot is the data plot, and the remaining
+\(m - 1\) plots have the identical graphical procedure except the data
+has been replaced with data consistent with \(H_0\). Then, an observer
+who have not seen the data plot will be asked to point out the most
 different plot from the lineup. Under \(H_0\), it is expected that the
-actual data plot would have no distinguishable difference with the null
-plots, and the probability of the observer correctly picks the actual
-data plot is \(1/m\). If we reject \(H_0\) as the observer correctly
-picks the actual data plot, then the Type I error of this test is
-\(1/m\).
+data plot would have no distinguishable difference from the null plots,
+and the probability that the observer correctly picks the data plot is
+\(1/m\). If one rejects \(H_0\) as the observer correctly picks the data
+plot, then the Type I error of this test is \(1/m\).
 
@@
 \caption{Visual testing is conducted using a lineup, as in the example here.
-The residual plot computed from the observed data (plot $2^2 + 2$, exhibiting non-linearity) is embedded among 19 null plots, where the residuals were simulated from a standard error model.
+The residual plot computed from the observed data (plot $2^2 + 2$, exhibiting non-linearity) is embedded among 19 null plots, where the residuals are simulated from a standard error model.
 Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot.
@@
 Figure \ref{fig:first-example-lineup} is an example of a lineup
-protocol. If the actual data plot at position \(2^2 + 2\) is
-identifiable, then it is evidence for the rejection of \(H_0\) that the
-regression model is correctly specified. In fact, the actual residual
-plot is obtained from a misspecified regression model with non-linearity
-defect.
-
-The effectiveness of lineup protocol is validated by
-\citet{majumder_validation_2013} under relatively simple classical
-normal linear regression model settings with one or two regressors.
-Their results suggest that visual test is capable of testing the
-significance of a single regressor with a similar power as a t-test,
-though they express that in general it is unnecessary to use visual
-inference if there exists a conventional test and they do not expect the
-visual test to perform equally well as the conventional test. In their
-third experiment, where there does not exist a proper conventional test,
-visual test outperforms the conventional test for a large margin. This
-is encouraging as it promotes the use of visual inference in border
-field of data science where there are no existing statistical testing
-procedures. Visual inference have also already been integrated into
-diagnostic of hierarchical linear models by \citet{loy2013diagnostic},
+protocol. If the data plot at position \(2^2 + 2\) is identifiable, then
+it is evidence for the rejection of \(H_0\) that the regression model is
+correctly specified. In fact, the actual residual plot is obtained from
+a misspecified regression model with non-linearity defect.
+
+The effectiveness of lineup protocol for regression analysis is
+validated by \citet{majumder_validation_2013} under relatively simple
+settings with up to two regressors. Their results suggest that visual
+tests are capable of testing the significance of a single regressor with
+a similar power as a t-test, though they express that in general it is
+unnecessary to use visual inference if there exists a conventional test,
+and they do not expect the visual test to perform equally well as the
+conventional test. In their third experiment, where there is not a
+conventional test, visual test outperforms the conventional test for a
+large margin. This is encouraging, as it promotes the use of visual
+inference in situations where there are no existing statistical testing
+procedures. Visual inference have also been integrated into diagnostic
+of hierarchical linear models by \citet{loy2013diagnostic},
 \citet{loy2014hlmdiag} and \citet{loy2015you}. They use lineup protocols
 to judge the assumption of linearity, normality and constant error
-variance for both the level-1 and level-2 residuals.
+variance for both the level-1 and level-2 residuals. (expand?)
 
@@
 
-Data used in the \(m - 1\) null plots need to be simulated. In the
+Data used in the \(m - 1\) null plots needs to be simulated. In the
 context of regression diagnostics, sampling data from \(H_0\) is
 equivalent to sampling data from the assumed model. As
-\citet{buja_statistical_2009} suggested, \(H_0\) is usually composited
-by a collection of distributions controlled by nuisance parameters.
-Since regression models can have various forms, there is no general
-solution to this problem, but it sometimes can be reduced to so called
-``reference distribution'' by applying one of the three methods: (i)
-sampling from a conditional distribution given a minimal sufficient
-statistic under \(H_0\), (ii) parametric bootstrap sampling with
-nuisance parameters estimated under \(H_0\), and (iii) Bayesian
-posterior predictive sampling. The conditional distribution given a
-minimal sufficient statistic is the best justified reference
-distribution among the three \citep{buja_statistical_2009}. Essentially,
-null residuals can be simulated by regressing \(N\) i.i.d standard
-normal random draws on the regressors, then rescaling it by the ratio of
-residual sum of square in two regressions.
+\citet{buja_statistical_2009} suggested, \(H_0\) is usually a composite
+hypothesis controlled by nuisance parameters. Since regression models
+can have various forms, there is no general solution to this problem,
+but it sometimes can be reduced to a so called ``reference
+distribution'' by applying one of the three methods: (i) sampling from a
+conditional distribution given a minimal sufficient statistic under
+\(H_0\), (ii) parametric bootstrap sampling with nuisance parameters
+estimated under \(H_0\), and (iii) Bayesian posterior predictive
+sampling. The conditional distribution given a minimal sufficient
+statistic is the best justified reference distribution among the three
+\citep{buja_statistical_2009}. Essentially, null residuals can be
+simulated by regressing \(N\) i.i.d standard normal random draws on the
+regressors, then rescaling it by the ratio of residual sum of square in
+two regressions.
 
@@
 
-Further, a visual test can involve \(K\) independent observers. Let
-\(X_i = \{0,1\}\) be a binomial random variable denoting whether subject
-\(i\) correctly detecting the actual data plot, and
-\(X = \sum_{i=1}^{K}X_i\) be the number of observers correctly picking
-the actual data plot. Then, by imposing a relatively strong assumption
-on the visual test that all \(K\) evaluations are fully independent,
-under \(H_0\), \(X \sim \mathrm{Binom}_{K,1/m}\). Therefore, the
-\(p\)-value of a lineup of size \(m\) evaluated by \(K\) observer is
-given as \(P(X \geq x) = 1 - F(x) + f(x)\), where \(F(.)\) is the
-cumulative distribution function, \(f(.)\) is the probability mass
-function and \(x\) is the realization of number of observers correctly
-picking the actual data plot \citep{majumder_validation_2013}.
+In hypothesis testing, \(p\)-value is defined as the probability of
+observing test results as least as extreme as the observed result given
+\(H_0\) is true. Within the context of visual inference, by involving
+\(k\) independent observers, \(p\)-value can be interpreted as the
+probability of having as many or more subjects detect the data plot than
+the observed result.
+
+Let \(X_j = \{0,1\}\) be a binomial random variable denoting whether
+subject \(j\) correctly detecting the data plot, and
+\(X = \sum_{j=1}^{K}X_j\) be the number of observers correctly picking
+the data plot. Then, by imposing a relatively strong assumption on the
+visual test that all \(K\) evaluations are fully independent, under
+\(H_0\), \(X \sim \mathrm{Binom}_{K,1/m}\). Therefore, the \(p\)-value
+of a lineup of size \(m\) evaluated by \(K\) observer is given as
+\(P(X \geq x) = 1 - F(x) + f(x)\), where \(F(.)\) is the cumulative
+distribution function, \(f(.)\) is the probability mass function and
+\(x\) is the realization of number of observers correctly picking the
+data plot \citep{majumder_validation_2013}.
 
@@
 visual inference: (1) \(K\) different lineups are shown to \(K\)
-subjects, (2) \(K\) lineups with different null plots but the same
-actual data plot are shown to \(K\) subjects, and (3) the same lineup is
-shown to \(K\) subjects. Out of these three scenarios, Scenario 3 is the
-most common in previous studies as it puts the least constraints on the
+subjects, (2) \(K\) lineups with different null plots but the same data
+plot are shown to \(K\) subjects, and (3) the same lineup is shown to
+\(K\) subjects. Out of these three scenarios, Scenario 3 is the most
+common in previous studies as it puts the least constraints on the
 experiment design. For Scenario 3, \citet{vanderplas2021statistical}
-model the probability of a plot \(i\) being selected from a lineup as
+models the probability of a plot \(i\) being selected from a lineup as
 \(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
-\(i=1,...,m\) and \(\alpha > 0\). And defined \(c_i\) to be the number
-of times plot \(i\) being selected in \(K\) evaluations. In case subject
+\(i=1,...,m\) and \(\alpha > 0\). The number of times plot \(i\) being
+selected in \(K\) evaluations is denoted as \(c_i\). In case subject
 \(j\) makes multiple selections, \(1/s_j\) will be added to \(c_i\)
 instead of one, where \(s_j\) is the number of plots subject \(j\)
-selected for \(j=1,...K\). This ensured \(\sum_{i}c_i=K\). Since we are
-only interested in the selections of the actual data plot \(i\), the
-marginal model can be simplified to a beta-binomial model and thus the
-visual p-value is given as
+selected for \(j=1,...K\). This ensures \(\sum_{i}c_i=K\). Since we are
+only interested in the selections of the data plot \(i\), the marginal
+model can be simplified to a beta-binomial model and thus the visual
+p-value is given as
 
@@
 
-Note that Equation \ref{eq:pvalue-beta-binomial} only works with
-non-negative integer \(c_i\). For non-negative real number \(c_i\),
-linear approximation will be applied to calculate the p-value
+Note that Equation \ref{eq:pvalue-beta-binomial} given in
+\citet{vanderplas2021statistical} only works with non-negative integer
+\(c_i\). We extend the equation to non-negative real number \(c_i\) by
+applying a linear approximation
 
@@
 Besides, the parameter \(\alpha\) used in Equation
-\ref{eq:pvalue-beta-binomial} is usually unknown and hence needs to be
-estimated from the survey data. For low values of \(\alpha\), only a few
-plots are attractive to the observers and tend to be selected. For
-higher values of \(\alpha\), the distribution of the probability of each
-plot being selected is more even. \citet{vanderplas2021statistical}
-define that a plot is \(c\)-interesting if \(c\) or more participants
-select the plot as the most different. Given the definition, The
-expected number of plots selected at least \(c\) times, \(E[Z_c]\), is
-calculated as
+\ref{eq:pvalue-beta-binomial} and \ref{eq:pmf-beta-binomial} is usually
+unknown and hence needs to be estimated from the survey data. For low
+values of \(\alpha\), only a few plots are attractive to the observers
+and tend to be selected. For higher values of \(\alpha\), the
+distribution of the probability of each plot being selected is more
+even. \citet{vanderplas2021statistical} defines that a plot is
+\(c\)-interesting if \(c\) or more participants select the plot as the
+most different. Given the definition, The expected number of plots
+selected at least \(c\) times, \(E[Z_c]\), is calculated as
 
@@
 
-\citet{vanderplas2021statistical} suggest that \(\alpha\) can be
-estimated using MLE with Equation \ref{eq:c-interesting-expectation}.
-But for precise estimate of \(\alpha\), additional responses to
-Rorschach lineups, which is a type of lineup that consists of plots
-constructed from the same null data generating mechanism, are required.
+With Equation \ref{eq:c-interesting-expectation}, \(\alpha\) can be
+estimated using maximum likelihood estimation. But for precise estimate
+of \(\alpha\), additional responses to Rorschach lineups, which is a
+type of lineup that consists of plots constructed from the same null
+data generating mechanism, are required.
 
@@
 
-As discussed in \citet{majumder_validation_2013}, individual's skill
-will affect the number of observers who identify the actual data plot
-from the lineup. Thus, the power of a visual test depends on the
+The power of a model misspecification test specifies the chance of a
+misspecified regression model being detected given a certain alternative
+hypothesis. It is an important indicator when one concerns if the
+regression model is misspecified. Though, people might usually be more
+interested in knowing how much the residuals depart from the model
+assumptions and if the departure is of practical significance.
+
+As discussed in \citet{majumder_validation_2013}, individual's skill may
+affect the number of observers who identify the data plot from the
+lineup. Thus, the power of a visual test could depend on the
 subject-specific abilities. Previously, it is addressed by modelling the
-probability of a subject \(i\) correctly picking the actual data plot
-from a lineup \(l\) using a mixed-effect logistic regression with the
-subject being treated as a random effect
-\citep{majumder_validation_2013}. However, having this probability is
-insufficient to determine the power of a visual test allowed for
-multiple selections as it doesn't provide information about the number
-of selections made by the subject for p-value calculation.
-
-Instead, we omitted the individual effect and directly estimated the
-probability of a lineup being rejected using a logistic regression with
-the natural logarithm of the effect size as the only regressor
-formulated as
+probability of a subject \(j\) correctly picking the data plot from a
+lineup \(l\) using a mixed-effect logistic regression with the subject
+being treated as a random effect \citep{majumder_validation_2013}.
+However, in the multiple selections scenario, having this probability is
+insufficient to determine the power of a visual test as it doesn't
+provide information about the number of selections made by the subject
+for p-value calculation.
+
+Instead, we estimate the probability of a lineup being rejected directly
+by assuming the individual skill has negligible impact on the variation
+of the power of a lineup. The assumption is made to simplify the model
+structure, thus mitigating the need for expensive large-scale
+experiments to estimate the complex covariance matrix. The model is a
+logistic regression with the natural logarithm of the effect as the only
+regressor formulated as
 
 \begin{equation} \label{eq:logistic-regression-1-1}
-Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(\boldsymbol{E})),
+Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(E)),
 \end{equation}
@@
 
-Three experiments were conducted to investigate the difference between
+Three experiments are conducted to investigate the difference between
 conventional hypothesis testing and visual inference in the application
@@
 conventional test. The experiment III is designed for collecting human
-responses to lineup with only good residual plots such that the
-parameter \(\alpha\) in Equation \ref{eq:pvalue-beta-binomial} can be
-estimated. Overall, we planned to collect 7974 evaluations on 1152
-uniqued lineups performed by 443 subjects throughout three experiment.
+responses to null lineups such that the parameter \(\alpha\) in Equation
+\ref{eq:pvalue-beta-binomial} can be estimated. Overall, we plan to
+collect 7974 evaluations on 1152 uniqued lineups performed by 443
+subjects throughout three experiments.
 
@@
 Two types of departures, namely non-linearity and heteroskedasticity,
-were considered with the corresponding data generating process being
+are considered with the corresponding data generating process being
 designed for experiment I and II.
@@
 \end{figure}
-
-(include English reference)
 
@@
 \boldsymbol{z} = g(\boldsymbol{z}_{raw}, 1), \\
-\boldsymbol{z}_{raw} = He_j(g(\boldsymbol{z}, 2)),
+\boldsymbol{z}_{raw} = He_j(g(\boldsymbol{x}, 2)),
 \end{align}
@@
 \(g(\boldsymbol{x}, k)\) is a scaling function to enforce the support of
-the random vector to be \(\{-k, k\}\) defined as
+the random vector to be \([-k, k]^n\) defined as
 
@@
 
-Hermite polynomials were initially defined by \citet{de1820theorie}, but
-named after Hermite \citep{hermite1864nouveau} because of the
-unrecognisable form of Laplace's work. When simulating \(Z_{raw}\), the
-function \texttt{hermite} from the R package \texttt{mpoly}
-\citep{mpoly} was used to generate Hermite polynomials.
+According to \citet{abramowitz1964handbook}, Hermite polynomials were
+initially defined by \citet{de1820theorie}, but named after Hermite
+\citep{hermite1864nouveau} because of the unrecognisable form of
+Laplace's work. When simulating \(\boldsymbol{z}_{raw}\), function
+\texttt{hermite} from the R package \texttt{mpoly} \citep{mpoly} is used
+to generate Hermite polynomials.
 
@@
 Since \(z = O(x^j)\), for \(j > 1\), \(z\) is a higher order term leaves
-out by the null regression, which will result in model misspecification.
-Visual patterns of non-linearity were simulated using four different
-order of probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)) and
-four different distribution of \(X_{raw}\): (1) \(U(-1, 1)\), (2)
+out by the null regression, which will lead to model misspecification.
+
+Visual patterns of non-linearity are simulated using four different
+orders of probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)) and
+four different distributions of \(X_{raw}\): (1) \(U(-1, 1)\), (2)
 \(N(0, 0.3^2)\), (3) \(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). A
@@
 
-The values of \(j\) was chosen so that distinct shapes of non-linearity
-were included in the residual plot. A greater value of \(j\) will result
-in a curve with more turning points. As shown in Figure
-\ref{fig:different-shape-of-herimite}, it includes ``U'' shape, ``S''
-shape, ``M'' shape and ``Triple-U'' shape. It is expected that the ``U''
-shape will be the easiest one to detect because complex shape tends to
-be concealed by cluster of data points.
-
-Four different distribution were used to generate \(X_{raw}\) as shown
-in Figure \ref{fig:different-dist}. The uniform and the normal
-distribution are symmetric and commonly assumed in statistical models.
-The adjusted log-normal distribution provides skewed density, while the
-discrete uniform distribution provides discreteness in residual plot,
-which could enrich the pool of visual patterns.
+The values of \(j\) is chosen so that distinct shapes of non-linearity
+are included in the residual plot. These include ``U'', ``S'', ``M'' and
+``Triple-U'' shape as shown in Figure
+\ref{fig:different-shape-of-herimite}. A greater value of \(j\) will
+result in a curve with more turning points. It is expected that the
+``U'' shape will be the easiest one to detect because complex shape
+tends to be concealed by cluster of data points.
+
+Different distributions of \(X_{raw}\) help enriching the pool of visual
+patterns as illustrated in Figure \ref{fig:different-dist}. The uniform
+and the normal distribution are symmetric and commonly assumed in
+statistical models. The adjusted log-normal distribution provides skewed
+density, while the discrete uniform distribution provides discreteness
+in residual plot.
 
@@
 
-Figure \ref{fig:example-poly-lineup} shows one of the lineups used in
-experiment I. This lineup is produced under \(j = 6\) and
-\(X_{raw} \sim N(0.0.3^2)\). The actual data plot location is
-\(2^3 - 4\). All five subjects correctly identified the actual data plot
-for this lineup.
+Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups
+used in experiment I. This lineup is produced by the non-linearity model
+under \(j = 6\) and \(X_{raw} \sim N(0.0.3^2)\). The data plot location
+is \(2^3 - 4\). All five subjects correctly identify the data plot from
+this lineup.
 
@@
 \(\boldsymbol{\varepsilon}\) are vectors of size \(n\) and \(g(.)\) is
-the scaling function defined in \ref{eq:scaling-function}.
+the scaling function defined in Equation \ref{eq:scaling-function}.
 
@@
 \(\boldsymbol{x}\), which will lead to the presence of
-heteroskedasticity. Visual patterns were simulated using three different
-shapes (\(a\) = -1, 0, 1) and the same four different distribution of
-\(X_{raw}\) used in experiment I. A summary of the factors is given in
-Table \ref{tab:model-factor-table}.
-
-The values of \(a\) was chosen so that different shapes of
-heteroskedasticity were included in the residual plot. These include
-left-triangle shape, butterfly shape and right-triangle shape as
-displayed in Figure \ref{fig:different-shape-of-heter}.
+heteroskedasticity. Visual patterns of heteroskedasticity are simulated
+using three different shapes (\(a\) = -1, 0, 1) and the same four
+different distribution of \(X_{raw}\) used in experiment I. A summary of
+the factors can also be found in Table \ref{tab:model-factor-table}.
+
+Since \(supp(X) = [-1, 1]\), choosing \(a\) to be \(-1\), \(0\) and
+\(1\) can generate ``left-triangle'', ``butterfly'' and
+``right-triangle'' shape as displayed in Figure
+\ref{fig:different-shape-of-heter}. The term \((2 - |a|)\) maintains the
+magnitude of residuals across different values of \(a\).
 
@@
 
-An example lineup of this model is shown in Figure
+An example lineup of this model used in Experiment II is shown in Figure
 \ref{fig:example-heter-lineup} with \(a = -1\) and
-\(X_{raw} \sim U(-1, 1)\). The actual data plot location is \(2^4 + 2\).
-Nine out of 11 subjects correctly identified the actual data plot for
-this lineup.
+\(X_{raw} \sim U(-1, 1)\). The data plot location is \(2^4 + 2\). Nine
+out of 11 subjects correctly identify the data plot from this lineup.
 
@@
 As summarised in Table \ref{tab:model-factor-table}, three additional
-parameters \(n\), \(\sigma\) and \(b\) were used to control the strength
-of the signal so that different difficulty levels of lineups were
-generated, and therefore, the estimated power curve would be smooth and
+parameters \(n\), \(\sigma\) and \(b\) are used to control the strength
+of the signal so that different difficulty levels of lineups are
+generated, and therefore, the estimated power curve will be smooth and
 continuous. Parameter \(\sigma \in \{0.5, 1, 2, 4\}\) and
-\(b \in \{0.25, 1, 4, 16, 64\}\) were used in experiment I and II
+\(b \in \{0.25, 1, 4, 16, 64\}\) are used in experiment I and II
 respectively. Figure \ref{fig:different-sigma} and \ref{fig:different-b}
@@
 model and decrease the visibility of the visual pattern. The parameter
-\(b\) controls the ratio of the standard deviation of the
-heteroskedasticity across the domain of the regressor. Given
-\(x \neq a\), a larger value of \(b\) will lead to a larger ratio of the
-variance at \(x\) to the variance at \(x - a = 0\), making the visual
-pattern more obvious.
-
-Three different sample sizes were used (n = 50, 100, 300) in all three
+\(b\) controls the standard deviation of the error across the support of
+the regressor. Given \(x \neq a\), a larger value of \(b\) will lead to
+a larger ratio of the variance at \(x\) to the variance at
+\(x - a = 0\), making the visual pattern more obvious.
+
+Three different sample sizes are used (n = 50, 100, 300) in all three
 experiments. It can be observed from Figure \ref{fig:different-n} that
@@
 
-Three replications were made for each of the parameter values shown in
+Three replications are made for each of the parameter values shown in
 Table \ref{tab:model-factor-table} resulting in
 \((4 \times 4 \times 4 \times 3 + 4 \times 3 \times 5 \times 3) \times 3 = 1116\)
-different lineups. In addition, each lineup was designed to be evaluated
+different lineups. In addition, each lineup is designed to be evaluated
 by five different subjects. After attempting some pilot studies
-internally in our research group, we decided to present a block of 20
+internally in our research group, we decide to present a block of 20
 lineups to every subject. And to ensure the quality of the survey data,
-two lineups with obvious visual patterns were included as attention
-checks. Thus, \(576 \times 5 \div (20-2) = 160\) and
-\(540 \times 5 \div (20-2) = 150\) subjects were recruited to satisfy
-the design of the experiment I and experiment II respectively.
+two lineups with obvious visual patterns are included as attention
+checks. Thus, \(576 \times 5 / (20-2) = 160\) and
+\(540 \times 5 / (20-2) = 150\) subjects are recruited to satisfy the
+design of the experiment I and experiment II respectively.
 
@@
 lineups with all combinations of \(n\) and \(X_{raw}\) and three
-replications were included in experiment III. In these lineups, the data
-of the data plot was generated from a model
+replications are included in experiment III. In these lineups, the data
+of the data plot is generated from a model
 {\textcolor{orange-red}{with zero effect size}}, while the data of the
-19 null plots were generated using the same simulation method discussed
+19 null plots are generated using the same simulation method discussed
 in Section \ref{sampling-from-the-null-distribution}.
@@
 {\textcolor{orange-red}{To account for the fact that our simulation method for these lineups is not the Rorschach procedure,}}
-we used the method suggested in \citet{vanderplas2021statistical} for
+we use the method suggested in \citet{vanderplas2021statistical} for
 typical lineups containing a data plot to estimate \(\alpha\). We have
@@
 considered to be bad practices as subjects will lose interest quickly.
-We planned to also collect 6 more evaluations on the 279 lineups with
-\(X_{raw} \sim U(-1,1)\), resulted in
-\((36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3) \div (20-2) = 133\)
+Therefore, we plan to collect 6 more evaluations on the 279 lineups with
+\(X_{raw} \sim U(-1,1)\), result in
+\((36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3) / (20-2) = 133\)
 subjects recruited for experiment III.
@@
 
-Subjects for all three experiments were recruited from an crowdsourcing
-platform called Prolific {[}palan2018prolific{]}. Prescreening procedure
-was applied during the recruitment, subjects were required to be fluent
-in English, with \(98\%\) minimum approval rate in other studies and 10
-minimum submissions. During the experiment, every subject was presented
-with a block of 20 lineups. For each lineup, the actual data plot was
-drawn as a standard residual plot of the null model with raw residuals
-on the y-axis and fitted values on the x-axis. An additional horizontal
-red line was added at \(y = 0\) as a helping line. The 19 null datasets
-were generated by the residual rotation technique, and plotted in the
-same way. The lineup consisted of 20 residual plots with one randomly
-placed actual data plot. And for every lineup, the subject was asked to
-select one or more plots that are most different from others, provide a
-reason for their selections, and evaluate how different they think the
-selected plots were from others. If there was no noticeable difference
-between plots in a lineup, subjects were permitted to select zero plots
-without providing the reason. No subject was shown the same lineup
-twice. Information about preferred pronoun, age group, education, and
-previous experience in visual experiment were also collected. A
-subject's submission was only accepted if the actual data plot was
-identified for at least one attention check. Data of rejected
-submissions were discarded automatically to maintain the overall data
-quality.
+Subjects for all three experiments are recruited from an crowdsourcing
+platform called Prolific \citep{palan2018prolific}. Prescreening
+procedure is applied during the recruitment, subjects are required to be
+fluent in English, with \(98\%\) minimum approval rate and 10 minimum
+submissions in other studies.
+
+During the experiment, every subject is presented with a block of 20
+lineups. A lineup consists of a randomly placed data plot and 19 null
+plots, which are all residual plots drawn with raw residuals on the
+y-axis and fitted values on the x-axis. An additional horizontal red
+line is added at \(y = 0\) as a helping line.
+
+The data of the data plot is simulated from one of two models described
+in Section \ref{simulating-departures-from-good-residuals}, while the
+data of the remaining 19 null plots are generated by the residual
+rotation technique discussed in Section
+\ref{sampling-from-the-null-distribution}.
+
+In every lineup evaluation, the subject is asked to select one or more
+plots that are most different from others, provide a reason for their
+selections, and evaluate how different they think the selected plots are
+from others. If there is no noticeable difference between plots in a
+lineup, subjects are permitted to select zero plots without providing
+the reason. No subject are shown the same lineup twice. Information
+about preferred pronoun, age group, education, and previous experience
+in visual experiment are also collected. A subject's submission is only
+accepted if the data plot is identified for at least one attention
+check. Data of rejected submissions are discarded automatically to
+maintain the overall data quality.
 
@@
 
-There were 2880, 2880 and 2214 lineups evaluation made by 160, 160 and
-123 subjects recruited for experiment I, II and III respectively. In the
-total of 7974 lineup evaluations, 3744 used lineups produced by the
-non-linearity model, 3510 used lineups produced by the
-heteroskedasticity model, and 720 used lineups consists of only null
-plots.
-
-Besides, there were 886 attention checks not included in the following
-analysis. The collated dataset is provided in \texttt{vi\_survey} of the
-\texttt{visage} \texttt{R} package.
+There are 2880, 2700 and 2394 lineups evaluation made by 160, 150 and
+133 subjects recruited for experiment I, II and III respectively. In the
+total of 7974 lineup evaluations, 3744 use lineups produced by the
+non-linearity model, 3510 use lineups produced by the heteroskedasticity
+model, and 720 use null lineups. Besides, there are 886 attention checks
+not included in the following analysis. The collated dataset is provided
+in \texttt{vi\_survey} of the \texttt{visage} \texttt{R} package.
 
@@
 \(X_{raw} \sim U(-1,1)\), against the natural logarithm of the effect
-\(log_e(E)\).
+\(log_e(E)\), with a 5\% significance level. Lineups with
+\(X_{raw} \sim U(-1,1)\) are focused in this section because visual
+patterns are more likely to be revealed under the uniform distribution.
+Besides, we have collected more responses for this distribution in
+experiment III to produce more reliable and stable power curves. Power
+analysis of other distributions can be found in Section {[}?{]}. At the
+bottom of the figure \ref{fig:polypower}, there are a sequence of
+example residual plots with increasing levels of \(log_e(E)\). Readers
+can evaluate them from left to right and determine at which level the
+departure from a good residual plot becomes detectable.
 
@@
 
-To compare the power of visual test and conventional test, we picked
-RESET test (\texttt{resettest}) and Breusch-Pagan test (\texttt{bptest})
-from the R package \texttt{lmtest}, and Shapiro-Wilk test
+To compare the power of visual test and conventional test, we pick RESET
+test (\texttt{resettest}) and Breusch-Pagan test (\texttt{bptest}) from
+the R package \texttt{lmtest}, and Shapiro-Wilk test
 (\texttt{shapiro.test}) from the built-in R package \texttt{stats}.
@@
 approximate and inappropriate tests. Their estimated power is shown in
-Figure \ref{fig:polypower}. For the RESET test, we included different
-power of fitted values as proxies. According to
-\citet{ramsey_tests_1969}, there is no general rule for the power of the
-fitted values needed by the RESET test, but it found power up to four is
-usually sufficient. Thus, we followed this guideline to conduct the
-RESET test. For the Breusch-Pagan test, both \(\boldsymbol{x}\) and
-\(\boldsymbol{x}^2\) were included in the auxiliary regression equation
-to {[}justify this choice?{]}.
-
-At the bottom of the figure \ref{fig:polypower}, there are a sequence of
-example residuals plots with increasing levels of \(log_e(E)\). Readers
-can evaluate them from left to right and determine at which level the
-departure from a good residual plot becomes detectable.
+Figure \ref{fig:polypower}. To set up the RESET test, we include
+different powers of fitted values as proxies. According to
+\citet{ramsey_tests_1969}, there are no general rules for the power of
+the fitted values needed by the RESET test, but it finds power up to
+four is usually sufficient. Thus, we follow this guideline to conduct
+the RESET test. For the Breusch-Pagan test, the choice of regressors in
+the auxiliary regression is left to the user
+\citep{breusch_simple_1979}. But as \citet{waldman1983note} suggested,
+it is a good choice for the set of auxiliary regressors in the
+Breusch-Pagan test be the same as the White test. Thus, we include both
+\(\boldsymbol{x}\) and \(\boldsymbol{x}^2\) in the auxiliary regression.
 
@@
 
-For the non-linearity model, the power of all four tests increases as
-the effect gets larger. The power curve of RESET test climbs
+For the non-linearity model, the power curve of RESET test climbs
 aggressively from 13\% to around 70\% as \(log_e(E)\) increases from 0
 to 2, while power of other tests respond inactively to the change of
-effect and remain lower than 25\% throughout the period, showing that
-REST test is way more sensitive to the type of model defects that being
-considered. Meanwhile, no noticeable visual features can be spotted from
-the example residuals plots.
+effect, showing that REST test is way more sensitive to the type of
+model defects that being considered. Meanwhile, no noticeable visual
+features can be spotted from the example residual plots.
 
@@
 test is also almost always greater than the power of visual test. For
-\(0 \leq log_e(E) \leq 2\), where the visual feature is nearly
-unobservable from the example residual plots and the power curve of the
-visual test remains at a low level, the Breusch-Pagan test still have a
-decent amount of chance of rejecting \(H_0\).
+\(0 \leq log_e(E) \leq 2\), where the power curve of the visual test
+remains at a low level, the Breusch-Pagan test still have a decent
+amount of chance of rejecting \(H_0\). Similarly, the visual feature is
+nearly unobservable from the example residual plots.
 
 The power of visual test arises steadily as \(log_e(E)\) increases from
-2 to 5 in both non-linearity model and heteroskedasticity model,
+2 to 5 for both non-linearity model and heteroskedasticity model,
 suggesting that the effect starts to make significant impact on the
 degree of the presence of the designed visual features. This can also be
-observed from the example residuals plots that when \(log_e(E) = 2.5\),
-a weak ``S-shape'' and a weak ``triangle'' shape are presented in Figure
-\ref{fig:polypower} and Figure \ref{fig:heterpower} respectively. And as
-\(log_e(E)\) increases, the visual pattern becomes much clearer, and the
-power reaches almost 100\% at \(log_e(E) \approx 6\).
-
-The power of all inappropriate tests except for F-test shows improvement
-as the effect increases but at a lower rate than the visual test in both
-scenarios. This coincides the point made by \citet{cook1982residuals}
-that residual-based tests for a specific type of model defect may be
-sensitive to other types of model defects. The power curve of F-test
-remains at around 5\% in Figure \ref{fig:heterpower} since there is no
-non-linear term leave out in the heteroskedasticity model and \(H_0\) of
-the F-test will be always satisfied.
-
-(remove human perception)
-
-(revisit Emi's concern)
-
-Overall, the power analysis suggests that conventional tests do not
-match how humans perceive residual plots in these two scenarios. Humans
-have much higher tolerance of the residual departures than the
-conventional test. Since fail to reject \(H_0\) in a visual test means
-that there is no obvious visual discoveries found in the residual plot,
+observed from the example residual plots that when \(log_e(E) = 2.5\), a
+weak ``S-shape'' and a weak ``triangle'' shape are presented in Figure
+\ref{fig:polypower} and Figure \ref{fig:heterpower} respectively. The
+visual pattern becomes much clearer as \(log_e(E)\) increases. At
+\(log_e(E) \approx 6\), the power reaches almost 100\%.
+
+The power of all inappropriate tests except for RESET test shows
+improvement as the effect increases but at a lower rate than the visual
+test in both scenarios. This coincides the point made by
+\citet{cook1982residuals} that residual-based tests for a specific type
+of model defect may be sensitive to other types of model defects. The
+power curve of RESET test remains at around 5\% in Figure
+\ref{fig:heterpower} since there are no non-linear terms leave out in
+the heteroskedasticity model and \(H_0\) of the test is always
+satisfied.
+
+Overall, the power comparison suggests that conventional tests differs
+significantly from visual tests in these two scenarios. Visual test have
+much higher tolerance of the residual departures than the conventional
+test. Since fail to reject \(H_0\) in a visual test usually means that
+there are no obvious visual discoveries found in the residual plot,
 analysts and the general public as the consumers of the output may not
@@
 negligible residual departures is not practically meaningful and useful.
-(it does not fit the purpose of )
-
-This could limit the popularity of conventional tests in residual
+This may limit the popularity of conventional tests in residual
 diagnostics among analysts.
@@
 Where would you draw the line of too much non-linearity in the residuals?
-For the RESET test this is around log effect size 1, but for the visual test it is around 3.}\label{fig:polypower}
+For the RESET test this is around log effect size 1.5, but for the visual test it is around 3.5.}\label{fig:polypower}
 \end{figure}
@@
 
-(1. How you calc rejection, 2. scatter plot p-value vs p-value
--\textgreater{} appendix, 3. percentage in explained the tex, both
-marginal and conditional, 4. interesting lineup, discuss the reason)
-
-(style similar to mosaic plot)
-
-(lineups detected by visual test but not by conventional tests)
-
-Visual inference and conventional hypothesis testing are two distinct
-statistical procedures. As displayed in Figure
-\ref{fig:p-value-comparison}, the visual test gives higher p-values than
-the conventional test for both polynomial and heteroskedasticity models
-in most of the time, suggesting that humans have lower confidence and
-much higher tolerance of the residual departures than the conventional
-test.
-
-In particular, given 95\% confidence level, both tests will reject
-\(H_0\) in only around 58\% of the cases. And in 41\% of the cases, the
-visual test does not reject \(H_0\) but the conventional test does.
-Since fail to reject \(H_0\) in a visual test means that there is no
-obvious visual discoveries found in the residual plot, analysts and the
-general public as the consumers of the output may not fully believe in
-the rejection of \(H_0\) even though the alternative hypothesis is true.
-Even if the rejection is accepted, the model violation may not be
-considered as impactful due to the fact that the departures are not
-clearly visible.
-
-This makes it sounds like people are arrogantly ignoring the accuracy of
-conventional tests and trust too much to their visual perception, but it
-is not. In fact, the sensitivity of the conventional test could
-sometimes distract and discourage analysts from finding a simple but
-good linear approximation to the data. The rejection of \(H_0\) because
-of acceptable and negligible residual departure is not practically
-meaningful and useful. This also partly explains why abundance of
-literature suggesting the use of data plot in regression diagnostics.
+The power comparison illustrates that appropriate conventional tests
+reject \(H_0\) more often than visual tests. In this section, we explore
+how often they agree with each other by investigate the rejection rates
+based on \(p\)-values.
+
+Depending on the natural logarithm of the effect, lineups are classified
+into three categories: easy (\textless{} 1.5), moderate (1.5 - 3.5) and
+difficult (\textgreater{} 3.5). Figure \ref{fig:p-value-comparison}
+provides a mosaic plot showing the rejection rate of visual tests and
+conventional tests at different difficulty levels.
+
+\hypertarget{easy-lineup}{%
+\subsubsection{Easy lineup}\label{easy-lineup}}
+
+For easy lineups produced by the non-linearity model, 65\% are rejected
+by both the conventional and visual tests, while 9\% are accepted by
+both tests. Thus, 74\% of the time two tests will agree with each other.
+There are 16\% of lineups rejected only by the conventional test.
+
+Interestingly, in spite of the greater power of the conventional test,
+10\% of the lineups are rejected only by the visual test. This means
+among those accepted by the conventional test, 51\% are rejected by the
+visual test. Recall that the RESET test used in our analysis includes
+powers of fitted values up to four. However, the ``M'' and ``Triple-U''
+shapes constructed from the Hermite polynomials contain powers of
+regressor up to six and 18 respectively. In fact, by including the
+powers of fitted values up to six, all cases will be rejected by the
+RESET test. And from the example plot displayed in Figure
+\ref{fig:easy-poly-example}, we can observe panel three is distinctly
+different from others exhibiting a ``M'' shape. It clearly suggests
+users of the RESET test and any other residual-based conventional tests
+that require the specification of variables of interest to check the
+residual plot before conducting the test, such that the correct choice
+of variables can be made.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/easy-poly-example-1}
+
+}
+
+\caption{An example easy lineup produced by the non-linearity model not rejected by the conventional test but rejected by the visual test.
+Panel three exhibits strong non-linear pattern.}\label{fig:easy-poly-example}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/easy-heter-example1-1}
+
+}
+
+\caption{An example easy lineup produced by the heteroskedasticity not rejected by the conventional test but rejected by the visual test.
+Panel 17 exhibits strong heteroskedasticity pattern.}\label{fig:easy-heter-example1}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/easy-heter-example2-1}
+
+}
+
+\caption{An example easy lineup produced by the heteroskedasticity not rejected by the conventional test but rejected by the visual test.
+Panel 18 contains a distinct outlier.}\label{fig:easy-heter-example2}
+\end{figure}
+
+For easy lineups produced by the heteroskedasticity model, 62\% are
+rejected by both the conventional test and the visual test, and 7\% are
+accepted by both, result in 69\% agreement rate. Lineups rejected by
+only conventional test are 30\%. 1\% of lineups are accepted by the
+conventional test but rejected by the visual test.
+
+Among those 1\% lineups, which are two lineups, one shows detectable
+heteroskedasticity pattern as displayed in Figure
+\ref{fig:easy-heter-example1}. Another contain data plot with an
+noticeable outlier, but the ``Right-triangle'' shape is not obvious as
+shown in Figure \ref{fig:easy-heter-example2}. The \(p\)-values of
+Breusch-Pagan test for these two lineups are \(0.055\) and \(0.135\).
+Compared to the visual test, its \(p\)-values are \(0.026\) and
+\(0.012\). The Breusch-Pagan test should reject the one with obvious
+heteroskedasticity signal, but it yields a \(p\)-value slightly above
+5\%.
+
+\hypertarget{moderate-lineup}{%
+\subsubsection{Moderate lineup}\label{moderate-lineup}}
+
+With the increase of the difficulty level, the sensitivity of
+conventional tests enlarge the difference in rejection rates. For
+moderate lineup produced by the non-linearity model, only 8\% and 28\%
+are rejected and accepted by both respectively. We can observe a decline
+in agreement rate to 36\%. A huge amount of lineups are accepted by the
+visual test. 63\% of lineups rejected only by the conventional test.
+There are 1\% of lineups which rejected only by conventional tests
+suffer the same RESET test issue as mentioned earlier.
+
+For those produced by the heteroskedasticy model, the corresponding
+numbers are 23\%, 31\%, 54\%, 45\% and 1\%. Only one lineup rejected by
+the visual test but accepted by the conventional test. It shows zero
+heteroskedasticity patterns, but contains a few outliers at the bottom
+of the data plot as illustrated in panel 14 of Figure
+\ref{fig:moderate-heter-example1}. Two out of five subjects correctly
+identify this lineup yielding a \(p\)-value equal to \(0.044\).
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/moderate-heter-example1-1}
+
+}
+
+\caption{An example easy lineup produced by the heteroskedasticity not rejected by the conventional test but rejected by the visual test.
+Panel 14 contains a few outliers at the bottom of the panel.}\label{fig:moderate-heter-example1}
+\end{figure}
+
+\hypertarget{difficult-lineup}{%
+\subsubsection{Difficult lineup}\label{difficult-lineup}}
+
+Visual test is insensitive to lineups with small effect. When it comes
+to difficult lineups produced by the non-linearity model, none of them
+are rejected by the visual test. There are 73\% of lineups accepted by
+both tests, and 26\% rejected by the conventional test. The agreement
+rate rise up to 73\%.
+
+For lineups produced by the heteroskedasticity model, 3\% are rejected
+by both, 71\% are accepted by both, 24\% are rejected only by the
+conventional test and 2\% are rejected only by the visual test, which
+are two lineups.
+
+As shown in \ref{fig:difficult-heter-example1} and
+\ref{fig:difficult-heter-example2}, residual plots of both lineups are
+constructed with \(X_{raw} \sim lognormal(0, 0.6^2)/3\). Though the
+designed patterns are ``Left-triangle'' and ``Butterfly'', both of them
+are invisible due to the small effect. The residual plots are dominated
+by the skewed distribution. Further, no distinguishable visual patterns,
+such as outliers, are detectable from the data plot of these two
+lineups, which are panel 5 and panel 15. The data plot in these two
+lineups are not the most different plot from our point of view either.
+However, three out of five and two out of five subjects correctly
+identify the data plot for these two lineups respectively. But the
+reasons for selecting the plot provided by the subjects are not
+consistent. Therefore, with the current data, we can not explain why
+subjects can correctly identify the data plot of these two lineups.
+Collecting more data in future study may help to answer this question.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/difficult-heter-example1-1}
+
+}
+
+\caption{An example easy lineup produced by the heteroskedasticity not rejected by the conventional test but rejected by the visual test.
+Panel 14 contains a few outliers at the bottom of the panel.}\label{fig:difficult-heter-example1}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/difficult-heter-example2-1}
+
+}
+
+\caption{An example easy lineup produced by the heteroskedasticity not rejected by the conventional test but rejected by the visual test.
+Panel 14 contains a few outliers at the bottom of the panel.}\label{fig:difficult-heter-example2}
+\end{figure}
 
@@
 
-\caption{Rejection rate of visual test compared to conventional test on lineups produced by heteroskedasticity and non-linearity model.
+\caption{Rejection rate of visual test compared to conventional test on lineups produced by heteroskedasticity and non-linearity model at different levels of difficulty.
 Based on the rejection decision made by the visual test and the conventional test, the mosaic plot shows the proportion of lineups in each category.}\label{fig:p-value-comparison}

### 2023-01-10 paper_comparison/paper_comparison/paper_comparison.Rmd
revise the p-value part

@@
 
-Depending on the natural logarithm of the effect, lineups are classified into three categories: easy (< 1.5), moderate (1.5 - 3.5) and difficult (> 3.5).
+Depending on the natural logarithm of the effect, lineups are classified into three categories: easy (> 3.5), moderate (1.5 - 3.5) and difficult (< 1.5).
 Figure \ref{fig:p-value-comparison} provides a mosaic plot showing the rejection rate of visual tests and conventional tests at different difficulty levels.
@@
 Thus, 74% of the time two tests will agree with each other.
-There are 16% of lineups rejected only by the conventional test.
-
-Interestingly, in spite of the greater power of the conventional test, 10% of the lineups are rejected only by the visual test.
-This means among those accepted by the conventional test, 51% are rejected by the visual test.
+There are 16% of lineups rejected by the conventional test only.
+
+Interestingly, in spite of the greater power of the conventional test, 10% of the lineups are rejected by the visual test only.
 Recall that the RESET test used in our analysis includes powers of fitted values up to four.
-However, the "M" and "Triple-U" shapes constructed from the Hermite polynomials contain powers of regressor up to six and 18 respectively.
-In fact, by including the powers of fitted values up to six, all cases will be rejected by the RESET test.
-And from the example plot displayed in Figure \ref{fig:easy-poly-example}, we can observe panel three is distinctly different from others exhibiting a "M" shape.
-It clearly suggests users of the RESET test and any other residual-based conventional tests that require the specification of variables of interest to check the residual plot before conducting the test, such that the correct choice of variables can be made.
-
-```{r easy-poly-example, fig.height = 5, fig.cap = "An example easy lineup produced by the non-linearity model not rejected by the conventional test but rejected by the visual test.
-Panel three exhibits strong non-linear pattern."}
+However, the "M" and the "Triple-U" shapes constructed from the Hermite polynomials contain powers of the regressor up to six and 18 respectively.
+In fact, by including the powers of fitted values up to six in the RESET test, all cases will be rejected.
+We also can observe from the example plot displayed in Figure \ref{fig:easy-poly-example} that panel three is distinctly different from others, exhibiting a "M" shape.
+This clearly suggests users of the RESET test and any other residual-based conventional tests that require the specification of variables of interest to check the residual plot before conducting the test, such that the correct choice of variables can be made.
+
+```{r easy-poly-example, fig.height = 5, fig.cap = "Lineup poly-270 produced by the non-linearity model used in experiment I and III.
+It is rejected by the visual test but not by the RESET test.
+The data plot at panel three exhibits strong non-linear pattern."}
 VI_MODEL$plot_lineup(vi_lineup$poly_270$data, 
@@
 
-```{r easy-heter-example1, fig.height = 5, fig.cap = "An example easy lineup produced by the heteroskedasticity not rejected by the conventional test but rejected by the visual test.
-Panel 17 exhibits strong heteroskedasticity pattern."}
+For easy lineups produced by the heteroskedasticity model, 62% are rejected by both the conventional test and the visual test, and 7% are accepted by both, result in 69% agreement rate.
+Lineups only rejected by conventional test are 30%.
+1% of lineups are accepted by the conventional test but rejected by the visual test.
+
+Among those 1%, which are two lineups, one shows detectable heteroskedasticity pattern as displayed in Figure \ref{fig:easy-heter-example1}.
+Another contain data plot with an noticeable outlier, but the "Right-triangle" shape is not very strong as shown in Figure \ref{fig:easy-heter-example2}.
+The $p$-values of Breusch-Pagan test for these two lineups are $0.055$ and $0.135$.
+Compared to the visual test, its $p$-values are $0.026$ and $0.012$.
+The Breusch-Pagan test should reject the one with obvious heteroskedasticity signal, but it yields a $p$-value slightly above 5%.
+
+```{r easy-heter-example1, fig.height = 5, fig.cap = 'Lineup heter-331 produced by the heteroskedasticity model used in experiment II and III.
+It is rejected by the visual test but not by the  Breusch-Pagan test.
+The data plot at panel 17 contains a "Butterfly" shape, indicating the presence of heteroskedasticity.'}
 VI_MODEL$plot_lineup(vi_lineup$heter_331$data, 
@@
 
-```{r easy-heter-example2, fig.height = 5, fig.cap = "An example easy lineup produced by the heteroskedasticity not rejected by the conventional test but rejected by the visual test.
-Panel 18 contains a distinct outlier."}
+```{r easy-heter-example2, fig.height = 5, fig.cap = "Lineup heter-536 produced by the heteroskedasticity model used in experiment II.
+It is rejected by the visual test but not by the  Breusch-Pagan test.
+The data plot at panel 18 exhibits a vague heteroskedasticity pattern, but it contains a distinct outlier which makes it attractive to subjects."}
 VI_MODEL$plot_lineup(vi_lineup$heter_536$data, 
@@
 
-For easy lineups produced by the heteroskedasticity model, 62% are rejected by both the conventional test and the visual test, and 7% are accepted by both, result in 69% agreement rate.
-Lineups rejected by only conventional test are 30%.
-1% of lineups are accepted by the conventional test but rejected by the visual test.
-
-Among those 1% lineups, which are two lineups, one shows detectable heteroskedasticity pattern as displayed in Figure \ref{fig:easy-heter-example1}.
-Another contain data plot with an noticeable outlier, but the "Right-triangle" shape is not obvious as shown in Figure \ref{fig:easy-heter-example2}.
-The $p$-values of Breusch-Pagan test for these two lineups are $0.055$ and $0.135$.
-Compared to the visual test, its $p$-values are $0.026$ and $0.012$.
-The Breusch-Pagan test should reject the one with obvious heteroskedasticity signal, but it yields a $p$-value slightly above 5%.
-
 ### Moderate lineup
@@
 With the increase of the difficulty level, the sensitivity of conventional tests enlarge the difference in rejection rates.
-For moderate lineup produced by the non-linearity model, only 8% and 28% are rejected and accepted by both respectively.
+For moderate lineup produced by the non-linearity model, 8% and 28% are rejected and accepted by both tests respectively.
 We can observe a decline in agreement rate to 36%.
-A huge amount of lineups are accepted by the visual test.
-63% of lineups rejected only by the conventional test.
-There are 1% of lineups which rejected only by conventional tests suffer the same RESET test issue as mentioned earlier.
+A large proportion of lineups are accepted by the visual test and rejected by the conventional test, which is around 63%.
+The remaining 1% are rejected by conventional tests only, which suffer the same issue caused by the RESET test as mentioned earlier.
 
 For those produced by the heteroskedasticy model, the corresponding numbers are 23%, 31%, 54%, 45% and 1%.
-Only one lineup rejected by the visual test but accepted by the conventional test.
-It shows zero heteroskedasticity patterns, but contains a few outliers at the bottom of the data plot as illustrated in panel 14 of Figure \ref{fig:moderate-heter-example1}.
+There is only one lineup rejected by the visual test but accepted by the conventional test.
+It does not show any heteroskedasticity patterns, but contains a few outliers at the bottom of the data plot as illustrated in panel 14 of Figure \ref{fig:moderate-heter-example1}.
 Two out of five subjects correctly identify this lineup yielding a $p$-value equal to $0.044$.
-
-```{r moderate-heter-example1, fig.height = 5, fig.cap = "An example easy lineup produced by the heteroskedasticity not rejected by the conventional test but rejected by the visual test.
-Panel 14 contains a few outliers at the bottom of the panel."}
+We consider this instance is rejected by chance.
+If a different set of null plots are used, the data plot may not be identified.
+
+```{r moderate-heter-example1, fig.height = 5, fig.cap = "Lineup heter-219 produced by the heteroskedasticity model used in experiment II.
+It is rejected by the visual test but not by the  Breusch-Pagan test.
+Though the data plot at panel 14 shows no heteroskedasticity patterns, it contains a few outliers at the bottom of the panel."}
 VI_MODEL$plot_lineup(vi_lineup$heter_219$data, 
@@
 Visual test is insensitive to lineups with small effect.
-When it comes to difficult lineups produced by the non-linearity model, none of them are rejected by the visual test.
+When it comes to difficult level, none of the lineups produced by the non-linearity model are rejected by the visual test.
 There are 73% of lineups accepted by both tests, and 26% rejected by the conventional test.
@@
 
-For lineups produced by the heteroskedasticity model, 3% are rejected by both, 71% are accepted by both, 24% are rejected only by the conventional test and 2% are rejected only by the visual test, which are two lineups.
+For lineups produced by the heteroskedasticity model, 3% are rejected by both tests, and 71% are accepted by both tests.
+There are 24% of lineups rejected by the conventional test only.
+Besides, only two lineups, which is 2% of all the difficult lineups, are rejected by the visual test only.
 
 As shown in \ref{fig:difficult-heter-example1} and \ref{fig:difficult-heter-example2}, residual plots of both lineups are constructed with $X_{raw} \sim lognormal(0, 0.6^2)/3$.
-Though the designed patterns are "Left-triangle" and "Butterfly", both of them are invisible due to the small effect.
-The residual plots are dominated by the skewed distribution.
+Though the designed patterns are "Left-triangle" and "Butterfly" shape, both of them are invisible due to the small effect.
+The visual effect are dominated by the skewed distribution.
 Further, no distinguishable visual patterns, such as outliers, are detectable from the data plot of these two lineups, which are panel 5 and panel 15.
-The data plot in these two lineups are not the most different plot from our point of view either.
+The data plots in these two lineups are not the most different plot from our point of view either.
 However, three out of five and two out of five subjects correctly identify the data plot for these two lineups respectively.
-But the reasons for selecting the plot provided by the subjects are not consistent.
-Therefore, with the current data, we can not explain why subjects can correctly identify the data plot of these two lineups.
+But the reasons for selecting the plot provided by the subjects are mixed and inconsistent.
+Hence, with the current data, we can not explain why subjects can correctly identify the data plot of these two lineups.
 Collecting more data in future study may help to answer this question.
 
-```{r difficult-heter-example1, fig.height = 5, fig.cap = "An example easy lineup produced by the heteroskedasticity not rejected by the conventional test but rejected by the visual test.
-Panel 14 contains a few outliers at the bottom of the panel."}
+```{r difficult-heter-example1, fig.height = 5, fig.cap = "Lineup heter-224 produced by the heteroskedasticity model used in experiment II.
+It is rejected by the visual test but not by the  Breusch-Pagan test.
+Although two out of five subjects correctly identify the data plot at panel 5, it does not exhibit any visually distinguishable pattern."}
 VI_MODEL$plot_lineup(vi_lineup$heter_224$data, 
@@
 
-```{r difficult-heter-example2, fig.height = 5, fig.cap = "An example easy lineup produced by the heteroskedasticity not rejected by the conventional test but rejected by the visual test.
-Panel 14 contains a few outliers at the bottom of the panel."}
+```{r difficult-heter-example2, fig.height = 5, fig.cap = "Lineup heter-415 produced by the heteroskedasticity model used in experiment II.
+It is rejected by the visual test but not by the  Breusch-Pagan test.
+Panel 15 is the data plot but it does not exhibit any visually distinguishable pattern."}
 VI_MODEL$plot_lineup(vi_lineup$heter_415$data, 

### 2023-01-10 paper_comparison/paper_comparison/paper_comparison.tex
revise the p-value part

@@
 Depending on the natural logarithm of the effect, lineups are classified
-into three categories: easy (\textless{} 1.5), moderate (1.5 - 3.5) and
-difficult (\textgreater{} 3.5). Figure \ref{fig:p-value-comparison}
+into three categories: easy (\textgreater{} 3.5), moderate (1.5 - 3.5)
+and difficult (\textless{} 1.5). Figure \ref{fig:p-value-comparison}
 provides a mosaic plot showing the rejection rate of visual tests and
@@
 both tests. Thus, 74\% of the time two tests will agree with each other.
-There are 16\% of lineups rejected only by the conventional test.
+There are 16\% of lineups rejected by the conventional test only.
 
 Interestingly, in spite of the greater power of the conventional test,
-10\% of the lineups are rejected only by the visual test. This means
-among those accepted by the conventional test, 51\% are rejected by the
-visual test. Recall that the RESET test used in our analysis includes
-powers of fitted values up to four. However, the ``M'' and ``Triple-U''
-shapes constructed from the Hermite polynomials contain powers of
-regressor up to six and 18 respectively. In fact, by including the
-powers of fitted values up to six, all cases will be rejected by the
-RESET test. And from the example plot displayed in Figure
-\ref{fig:easy-poly-example}, we can observe panel three is distinctly
-different from others exhibiting a ``M'' shape. It clearly suggests
-users of the RESET test and any other residual-based conventional tests
-that require the specification of variables of interest to check the
-residual plot before conducting the test, such that the correct choice
-of variables can be made.
+10\% of the lineups are rejected by the visual test only. Recall that
+the RESET test used in our analysis includes powers of fitted values up
+to four. However, the ``M'' and the ``Triple-U'' shapes constructed from
+the Hermite polynomials contain powers of the regressor up to six and 18
+respectively. In fact, by including the powers of fitted values up to
+six in the RESET test, all cases will be rejected. We also can observe
+from the example plot displayed in Figure \ref{fig:easy-poly-example}
+that panel three is distinctly different from others, exhibiting a ``M''
+shape. This clearly suggests users of the RESET test and any other
+residual-based conventional tests that require the specification of
+variables of interest to check the residual plot before conducting the
+test, such that the correct choice of variables can be made.
 
@@
 
-\caption{An example easy lineup produced by the non-linearity model not rejected by the conventional test but rejected by the visual test.
-Panel three exhibits strong non-linear pattern.}\label{fig:easy-poly-example}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/easy-heter-example1-1}
-
-}
-
-\caption{An example easy lineup produced by the heteroskedasticity not rejected by the conventional test but rejected by the visual test.
-Panel 17 exhibits strong heteroskedasticity pattern.}\label{fig:easy-heter-example1}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/easy-heter-example2-1}
-
-}
-
-\caption{An example easy lineup produced by the heteroskedasticity not rejected by the conventional test but rejected by the visual test.
-Panel 18 contains a distinct outlier.}\label{fig:easy-heter-example2}
+\caption{Lineup poly-270 produced by the non-linearity model used in experiment I and III.
+It is rejected by the visual test but not by the RESET test.
+The data plot at panel three exhibits strong non-linear pattern.}\label{fig:easy-poly-example}
 \end{figure}
@@
 rejected by both the conventional test and the visual test, and 7\% are
-accepted by both, result in 69\% agreement rate. Lineups rejected by
-only conventional test are 30\%. 1\% of lineups are accepted by the
+accepted by both, result in 69\% agreement rate. Lineups only rejected
+by conventional test are 30\%. 1\% of lineups are accepted by the
 conventional test but rejected by the visual test.
 
-Among those 1\% lineups, which are two lineups, one shows detectable
+Among those 1\%, which are two lineups, one shows detectable
 heteroskedasticity pattern as displayed in Figure
 \ref{fig:easy-heter-example1}. Another contain data plot with an
-noticeable outlier, but the ``Right-triangle'' shape is not obvious as
-shown in Figure \ref{fig:easy-heter-example2}. The \(p\)-values of
+noticeable outlier, but the ``Right-triangle'' shape is not very strong
+as shown in Figure \ref{fig:easy-heter-example2}. The \(p\)-values of
 Breusch-Pagan test for these two lineups are \(0.055\) and \(0.135\).
@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/easy-heter-example1-1}
+
+}
+
+\caption{Lineup heter-331 produced by the heteroskedasticity model used in experiment II and III.
+It is rejected by the visual test but not by the  Breusch-Pagan test.
+The data plot at panel 17 contains a "Butterfly" shape, indicating the presence of heteroskedasticity.}\label{fig:easy-heter-example1}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/easy-heter-example2-1}
+
+}
+
+\caption{Lineup heter-536 produced by the heteroskedasticity model used in experiment II.
+It is rejected by the visual test but not by the  Breusch-Pagan test.
+The data plot at panel 18 exhibits a vague heteroskedasticity pattern, but it contains a distinct outlier which makes it attractive to subjects.}\label{fig:easy-heter-example2}
+\end{figure}
+
 \hypertarget{moderate-lineup}{%
@@
 conventional tests enlarge the difference in rejection rates. For
-moderate lineup produced by the non-linearity model, only 8\% and 28\%
-are rejected and accepted by both respectively. We can observe a decline
-in agreement rate to 36\%. A huge amount of lineups are accepted by the
-visual test. 63\% of lineups rejected only by the conventional test.
-There are 1\% of lineups which rejected only by conventional tests
-suffer the same RESET test issue as mentioned earlier.
+moderate lineup produced by the non-linearity model, 8\% and 28\% are
+rejected and accepted by both tests respectively. We can observe a
+decline in agreement rate to 36\%. A large proportion of lineups are
+accepted by the visual test and rejected by the conventional test, which
+is around 63\%. The remaining 1\% are rejected by conventional tests
+only, which suffer the same issue caused by the RESET test as mentioned
+earlier.
 
 For those produced by the heteroskedasticy model, the corresponding
-numbers are 23\%, 31\%, 54\%, 45\% and 1\%. Only one lineup rejected by
-the visual test but accepted by the conventional test. It shows zero
-heteroskedasticity patterns, but contains a few outliers at the bottom
-of the data plot as illustrated in panel 14 of Figure
-\ref{fig:moderate-heter-example1}. Two out of five subjects correctly
-identify this lineup yielding a \(p\)-value equal to \(0.044\).
+numbers are 23\%, 31\%, 54\%, 45\% and 1\%. There is only one lineup
+rejected by the visual test but accepted by the conventional test. It
+does not show any heteroskedasticity patterns, but contains a few
+outliers at the bottom of the data plot as illustrated in panel 14 of
+Figure \ref{fig:moderate-heter-example1}. Two out of five subjects
+correctly identify this lineup yielding a \(p\)-value equal to
+\(0.044\). We consider this instance is rejected by chance. If a
+different set of null plots are used, the data plot may not be
+identified.
 
@@
 
-\caption{An example easy lineup produced by the heteroskedasticity not rejected by the conventional test but rejected by the visual test.
-Panel 14 contains a few outliers at the bottom of the panel.}\label{fig:moderate-heter-example1}
+\caption{Lineup heter-219 produced by the heteroskedasticity model used in experiment II.
+It is rejected by the visual test but not by the  Breusch-Pagan test.
+Though the data plot at panel 14 shows no heteroskedasticity patterns, it contains a few outliers at the bottom of the panel.}\label{fig:moderate-heter-example1}
 \end{figure}
@@
 Visual test is insensitive to lineups with small effect. When it comes
-to difficult lineups produced by the non-linearity model, none of them
-are rejected by the visual test. There are 73\% of lineups accepted by
-both tests, and 26\% rejected by the conventional test. The agreement
-rate rise up to 73\%.
+to difficult level, none of the lineups produced by the non-linearity
+model are rejected by the visual test. There are 73\% of lineups
+accepted by both tests, and 26\% rejected by the conventional test. The
+agreement rate rise up to 73\%.
 
 For lineups produced by the heteroskedasticity model, 3\% are rejected
-by both, 71\% are accepted by both, 24\% are rejected only by the
-conventional test and 2\% are rejected only by the visual test, which
-are two lineups.
+by both tests, and 71\% are accepted by both tests. There are 24\% of
+lineups rejected by the conventional test only. Besides, only two
+lineups, which is 2\% of all the difficult lineups, are rejected by the
+visual test only.
 
@@
 constructed with \(X_{raw} \sim lognormal(0, 0.6^2)/3\). Though the
-designed patterns are ``Left-triangle'' and ``Butterfly'', both of them
-are invisible due to the small effect. The residual plots are dominated
-by the skewed distribution. Further, no distinguishable visual patterns,
-such as outliers, are detectable from the data plot of these two
-lineups, which are panel 5 and panel 15. The data plot in these two
+designed patterns are ``Left-triangle'' and ``Butterfly'' shape, both of
+them are invisible due to the small effect. The visual effect are
+dominated by the skewed distribution. Further, no distinguishable visual
+patterns, such as outliers, are detectable from the data plot of these
+two lineups, which are panel 5 and panel 15. The data plots in these two
 lineups are not the most different plot from our point of view either.
@@
 identify the data plot for these two lineups respectively. But the
-reasons for selecting the plot provided by the subjects are not
-consistent. Therefore, with the current data, we can not explain why
+reasons for selecting the plot provided by the subjects are mixed and
+inconsistent. Hence, with the current data, we can not explain why
 subjects can correctly identify the data plot of these two lineups.
@@
 
-\caption{An example easy lineup produced by the heteroskedasticity not rejected by the conventional test but rejected by the visual test.
-Panel 14 contains a few outliers at the bottom of the panel.}\label{fig:difficult-heter-example1}
+\caption{Lineup heter-224 produced by the heteroskedasticity model used in experiment II.
+It is rejected by the visual test but not by the  Breusch-Pagan test.
+Although two out of five subjects correctly identify the data plot at panel 5, it does not exhibit any visually distinguishable pattern.}\label{fig:difficult-heter-example1}
 \end{figure}
@@
 
-\caption{An example easy lineup produced by the heteroskedasticity not rejected by the conventional test but rejected by the visual test.
-Panel 14 contains a few outliers at the bottom of the panel.}\label{fig:difficult-heter-example2}
+\caption{Lineup heter-415 produced by the heteroskedasticity model used in experiment II.
+It is rejected by the visual test but not by the  Breusch-Pagan test.
+Panel 15 is the data plot but it does not exhibit any visually distinguishable pattern.}\label{fig:difficult-heter-example2}
 \end{figure}

### 2023-01-10 paper_comparison/paper_comparison/paper_comparison.Rmd
update caption of example lineups

@@
 Figure \ref{fig:polypower} shows the estimated power of visual test on lineups produced by the non-linearity model with $X_{raw} \sim U(-1,1)$, against the natural logarithm of the effect $log_e(E)$, with a 5% significance level.
-Lineups with $X_{raw} \sim U(-1,1)$ are focused in this section because visual patterns are more likely to be revealed under the uniform distribution.
-Besides, we have collected more responses for this distribution in experiment III to produce more reliable and stable power curves.
-Power analysis of other distributions can be found in Section [?].
+Uniform distribution is the focus of this section because visual patterns are more likely to be revealed.
+Besides, six more evaluations are collected for the uniform distribution in experiment III, which will produce more reliable and stable power curves.
+Analysis of other distributions can be found in Section \ref{effect-of-the-distribution-of-the-regressor}.
 At the bottom of the figure \ref{fig:polypower}, there are a sequence of example residual plots with increasing levels of $log_e(E)$.
@@
 
-Overall, the power comparison suggests that conventional tests differs significantly from visual tests in these two scenarios.
+Overall, the power comparison suggests that conventional tests differs significantly from visual tests in two regression diagnostics scenarios designed by us.
 Visual test have much higher tolerance of the residual departures than the conventional test.
@@
 
-The power comparison illustrates that appropriate conventional tests reject $H_0$ more often than visual tests.
+The power comparison illustrates that appropriate conventional tests reject $H_0$ more aggressively than visual tests.
 In this section, we explore how often they agree with each other by investigate the rejection rates based on $p$-values.
@@
 Thus, 74% of the time two tests will agree with each other.
-There are 16% of lineups rejected by the conventional test only.
-
-Interestingly, in spite of the greater power of the conventional test, 10% of the lineups are rejected by the visual test only.
+There are 16% of lineups rejected by conventional tests only.
+
+Interestingly, in spite of the greater power of conventional tests, 10% of the lineups are rejected by visual tests only.
 Recall that the RESET test used in our analysis includes powers of fitted values up to four.
-However, the "M" and the "Triple-U" shapes constructed from the Hermite polynomials contain powers of the regressor up to six and 18 respectively.
-In fact, by including the powers of fitted values up to six in the RESET test, all cases will be rejected.
+However, the "M" and the "Triple-U" shapes constructed from the Hermite polynomials contain power of the regressor up to six and 18 respectively.
+In fact, by including the power of fitted values up to six in the RESET test, all these cases will be rejected.
 We also can observe from the example plot displayed in Figure \ref{fig:easy-poly-example} that panel three is distinctly different from others, exhibiting a "M" shape.
@@
 
-```{r easy-poly-example, fig.height = 5, fig.cap = "Lineup poly-270 produced by the non-linearity model used in experiment I and III.
+```{r easy-poly-example, fig.height = 5, fig.cap = "Lineup poly-270 with $log_e(E) = 3.89$ produced by the non-linearity model used in experiment I and III.
 It is rejected by the visual test but not by the RESET test.
@@
 
-```{r easy-heter-example1, fig.height = 5, fig.cap = 'Lineup heter-331 produced by the heteroskedasticity model used in experiment II and III.
+```{r easy-heter-example1, fig.height = 5, fig.cap = 'Lineup heter-331 with $log_e(E) = 4.02$ produced by the heteroskedasticity model used in experiment II and III.
 It is rejected by the visual test but not by the  Breusch-Pagan test.
@@
 
-```{r easy-heter-example2, fig.height = 5, fig.cap = "Lineup heter-536 produced by the heteroskedasticity model used in experiment II.
+```{r easy-heter-example2, fig.height = 5, fig.cap = "Lineup heter-536 with $log_e(E) = 4.13$ produced by the heteroskedasticity model used in experiment II.
 It is rejected by the visual test but not by the  Breusch-Pagan test.
@@
 
-```{r moderate-heter-example1, fig.height = 5, fig.cap = "Lineup heter-219 produced by the heteroskedasticity model used in experiment II.
+```{r moderate-heter-example1, fig.height = 5, fig.cap = "Lineup heter-219 with $log_e(E) = 1.61$ produced by the heteroskedasticity model used in experiment II.
 It is rejected by the visual test but not by the  Breusch-Pagan test.
@@
 
-```{r difficult-heter-example1, fig.height = 5, fig.cap = "Lineup heter-224 produced by the heteroskedasticity model used in experiment II.
+```{r difficult-heter-example1, fig.height = 5, fig.cap = "Lineup heter-224 with $log_e(E) = 1.32$ produced by the heteroskedasticity model used in experiment II.
 It is rejected by the visual test but not by the  Breusch-Pagan test.
@@
 
-```{r difficult-heter-example2, fig.height = 5, fig.cap = "Lineup heter-415 produced by the heteroskedasticity model used in experiment II.
+```{r difficult-heter-example2, fig.height = 5, fig.cap = "Lineup heter-415 with $log_e(E) = -0.31$ produced by the heteroskedasticity model used in experiment II.
 It is rejected by the visual test but not by the  Breusch-Pagan test.
@@
 
-## Effect of the distribution of the independent variable
-
-# change of x_dist -> power
-
-# appendix (contains all the details)
+## Effect of the distribution of the regressor
+
+```{r fig.height = 3, fig.width = 5}
+vi_survey %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(type == "polynomial") %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
+  mutate(reject = as.numeric(p_value <= 0.05)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  mutate(x_dist = ifelse(x_dist == "even_discrete", 
+                         "Discrete", 
+                         stringr::str_to_title(x_dist))) %>%
+  ggplot() +
+  geom_smooth(data = mutate(poly_conv_sim, 
+                            log_effect_size = log(effect_size)) %>%
+                select(x_dist, log_effect_size, RESET4_p_value) %>%
+                rename(RESET_p_value = RESET4_p_value) %>%
+                pivot_longer(RESET_p_value) %>%
+                mutate(x_dist = ifelse(x_dist == "even_discrete", 
+                                       "Discrete", 
+                                       stringr::str_to_title(x_dist))) %>%
+                mutate(name = gsub("_p_value", " test", name)), 
+              aes(log_effect_size, 
+                  as.numeric(value < 0.05), 
+                  col = x_dist,
+                  linetype = "RESET"), 
+              method = "glm", 
+              method.args = list(family = binomial),
+              se = FALSE) +
+  geom_smooth(aes(log_effect_size, 
+                  reject, 
+                  col = x_dist,
+                  linetype = "Visual"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE) +
+  theme_light(base_size = 5) +
+  scale_color_manual(values = rev(rcartocolor::carto_pal(5, "Vivid"))) +
+  scale_size_manual(values = c(1, 2, 3)) +
+  scale_linetype_manual(values = c("longdash", "solid")) +
+  guides(linetype = guide_legend(nrow = 2), col = guide_legend(nrow = 2)) +
+  theme(legend.position = "bottom", legend.key.width = unit(1.5, "cm")) +
+  xlab("Effect size (log e)") +
+  ylab("Power") +
+  labs(col = "Distribution of regressor", linetype = "Test")
+```
+
+```{r fig.height = 3, fig.width = 5}
+vi_survey %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(type == "heteroskedasticity") %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
+  mutate(reject = as.numeric(p_value <= 0.05)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  mutate(x_dist = ifelse(x_dist == "even_discrete", 
+                         "Discrete", 
+                         stringr::str_to_title(x_dist))) %>%
+  ggplot() +
+  geom_smooth(data = mutate(heter_conv_sim, 
+                            log_effect_size = log(effect_size)) %>%
+                select(x_dist, log_effect_size, BP_p_value) %>%
+                pivot_longer(BP_p_value) %>%
+                mutate(x_dist = ifelse(x_dist == "even_discrete", 
+                                       "Discrete", 
+                                       stringr::str_to_title(x_dist))) %>%
+                mutate(name = gsub("_p_value", " test", name)), 
+              aes(log_effect_size, 
+                  as.numeric(value < 0.05), 
+                  col = x_dist,
+                  linetype = "BP"), 
+              method = "glm", 
+              method.args = list(family = binomial),
+              se = FALSE) +
+  geom_smooth(aes(log_effect_size, 
+                  reject, 
+                  col = x_dist,
+                  linetype = "Visual"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE) +
+  theme_light(base_size = 5) +
+  scale_color_manual(values = rev(rcartocolor::carto_pal(5, "Vivid"))) +
+  scale_size_manual(values = c(1, 2, 3)) +
+  scale_linetype_manual(values = c("longdash", "solid")) +
+  guides(linetype = guide_legend(nrow = 2), col = guide_legend(nrow = 2)) +
+  theme(legend.position = "bottom", legend.key.width = unit(1.5, "cm")) +
+  xlab("Effect size (log e)") +
+  ylab("Power") +
+  labs(col = "Distribution of regressor", linetype = "Test")
+```
+
+(change of x_dist -> power)
+
+# Appendix
+
+## Effect derivation
+
+## $p$-value scatter plot
 
@@
 
-### A collection interesting lineups (unusual results)
+## A collection interesting lineups (unusual results)
 

### 2023-01-10 paper_comparison/paper_comparison/paper_comparison.tex
update caption of example lineups

@@
 \(X_{raw} \sim U(-1,1)\), against the natural logarithm of the effect
-\(log_e(E)\), with a 5\% significance level. Lineups with
-\(X_{raw} \sim U(-1,1)\) are focused in this section because visual
-patterns are more likely to be revealed under the uniform distribution.
-Besides, we have collected more responses for this distribution in
-experiment III to produce more reliable and stable power curves. Power
-analysis of other distributions can be found in Section {[}?{]}. At the
-bottom of the figure \ref{fig:polypower}, there are a sequence of
-example residual plots with increasing levels of \(log_e(E)\). Readers
-can evaluate them from left to right and determine at which level the
+\(log_e(E)\), with a 5\% significance level. Uniform distribution is the
+focus of this section because visual patterns are more likely to be
+revealed. Besides, six more evaluations are collected for the uniform
+distribution in experiment III, which will produce more reliable and
+stable power curves. Analysis of other distributions can be found in
+Section \ref{effect-of-the-distribution-of-the-regressor}. At the bottom
+of the figure \ref{fig:polypower}, there are a sequence of example
+residual plots with increasing levels of \(log_e(E)\). Readers can
+evaluate them from left to right and determine at which level the
 departure from a good residual plot becomes detectable.
@@
 Overall, the power comparison suggests that conventional tests differs
-significantly from visual tests in these two scenarios. Visual test have
-much higher tolerance of the residual departures than the conventional
-test. Since fail to reject \(H_0\) in a visual test usually means that
-there are no obvious visual discoveries found in the residual plot,
-analysts and the general public as the consumers of the output may not
-be convinced of the existence of significant residual departures in
-spite of the rejection of \(H_0\) given by the conventional test. Even
-if the rejection is accepted, the model violation may be considered as
-impactless due to the fact that they are not clearly visible. Besides,
-the sensitivity of the conventional test could also distract and
-discourage analysts from finding simple but good linear approximation to
-the data. The rejection of \(H_0\) because of human acceptable and
-negligible residual departures is not practically meaningful and useful.
-This may limit the popularity of conventional tests in residual
-diagnostics among analysts.
+significantly from visual tests in two regression diagnostics scenarios
+designed by us. Visual test have much higher tolerance of the residual
+departures than the conventional test. Since fail to reject \(H_0\) in a
+visual test usually means that there are no obvious visual discoveries
+found in the residual plot, analysts and the general public as the
+consumers of the output may not be convinced of the existence of
+significant residual departures in spite of the rejection of \(H_0\)
+given by the conventional test. Even if the rejection is accepted, the
+model violation may be considered as impactless due to the fact that
+they are not clearly visible. Besides, the sensitivity of the
+conventional test could also distract and discourage analysts from
+finding simple but good linear approximation to the data. The rejection
+of \(H_0\) because of human acceptable and negligible residual
+departures is not practically meaningful and useful. This may limit the
+popularity of conventional tests in residual diagnostics among analysts.
 
@@
 The power comparison illustrates that appropriate conventional tests
-reject \(H_0\) more often than visual tests. In this section, we explore
-how often they agree with each other by investigate the rejection rates
-based on \(p\)-values.
+reject \(H_0\) more aggressively than visual tests. In this section, we
+explore how often they agree with each other by investigate the
+rejection rates based on \(p\)-values.
 
@@
 both tests. Thus, 74\% of the time two tests will agree with each other.
-There are 16\% of lineups rejected by the conventional test only.
-
-Interestingly, in spite of the greater power of the conventional test,
-10\% of the lineups are rejected by the visual test only. Recall that
-the RESET test used in our analysis includes powers of fitted values up
-to four. However, the ``M'' and the ``Triple-U'' shapes constructed from
-the Hermite polynomials contain powers of the regressor up to six and 18
-respectively. In fact, by including the powers of fitted values up to
-six in the RESET test, all cases will be rejected. We also can observe
+There are 16\% of lineups rejected by conventional tests only.
+
+Interestingly, in spite of the greater power of conventional tests, 10\%
+of the lineups are rejected by visual tests only. Recall that the RESET
+test used in our analysis includes powers of fitted values up to four.
+However, the ``M'' and the ``Triple-U'' shapes constructed from the
+Hermite polynomials contain power of the regressor up to six and 18
+respectively. In fact, by including the power of fitted values up to six
+in the RESET test, all these cases will be rejected. We also can observe
 from the example plot displayed in Figure \ref{fig:easy-poly-example}
@@
 
-\caption{Lineup poly-270 produced by the non-linearity model used in experiment I and III.
+\caption{Lineup poly-270 with $log_e(E) = 3.89$ produced by the non-linearity model used in experiment I and III.
 It is rejected by the visual test but not by the RESET test.
@@
 
-\caption{Lineup heter-331 produced by the heteroskedasticity model used in experiment II and III.
+\caption{Lineup heter-331 with $log_e(E) = 4.02$ produced by the heteroskedasticity model used in experiment II and III.
 It is rejected by the visual test but not by the  Breusch-Pagan test.
@@
 
-\caption{Lineup heter-536 produced by the heteroskedasticity model used in experiment II.
+\caption{Lineup heter-536 with $log_e(E) = 4.13$ produced by the heteroskedasticity model used in experiment II.
 It is rejected by the visual test but not by the  Breusch-Pagan test.
@@
 
-\caption{Lineup heter-219 produced by the heteroskedasticity model used in experiment II.
+\caption{Lineup heter-219 with $log_e(E) = 1.61$ produced by the heteroskedasticity model used in experiment II.
 It is rejected by the visual test but not by the  Breusch-Pagan test.
@@
 
-\caption{Lineup heter-224 produced by the heteroskedasticity model used in experiment II.
+\caption{Lineup heter-224 with $log_e(E) = 1.32$ produced by the heteroskedasticity model used in experiment II.
 It is rejected by the visual test but not by the  Breusch-Pagan test.
@@
 
-\caption{Lineup heter-415 produced by the heteroskedasticity model used in experiment II.
+\caption{Lineup heter-415 with $log_e(E) = -0.31$ produced by the heteroskedasticity model used in experiment II.
 It is rejected by the visual test but not by the  Breusch-Pagan test.
@@
 
-\hypertarget{effect-of-the-distribution-of-the-independent-variable}{%
-\subsection{Effect of the distribution of the independent
-variable}\label{effect-of-the-distribution-of-the-independent-variable}}
-
-\hypertarget{change-of-x_dist---power}{%
-\section{change of x\_dist -\textgreater{}
-power}\label{change-of-x_dist---power}}
-
-\hypertarget{appendix-contains-all-the-details}{%
-\section{appendix (contains all the
-details)}\label{appendix-contains-all-the-details}}
+\hypertarget{effect-of-the-distribution-of-the-regressor}{%
+\subsection{Effect of the distribution of the
+regressor}\label{effect-of-the-distribution-of-the-regressor}}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-14-1} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-15-1} \end{center}
+
+(change of x\_dist -\textgreater{} power)
+
+\hypertarget{appendix}{%
+\section{Appendix}\label{appendix}}
+
+\hypertarget{effect-derivation}{%
+\subsection{Effect derivation}\label{effect-derivation}}
+
+\hypertarget{p-value-scatter-plot}{%
+\subsection{\texorpdfstring{\(p\)-value scatter
+plot}{p-value scatter plot}}\label{p-value-scatter-plot}}
 
 \hypertarget{a-collection-interesting-lineups-unusual-results}{%
-\subsubsection{A collection interesting lineups (unusual
+\subsection{A collection interesting lineups (unusual
 results)}\label{a-collection-interesting-lineups-unusual-results}}

### 2023-01-10 paper_comparison/paper_comparison/paper_comparison.Rmd
update the effect of x_dist

@@
 ```{r polypower, fig.height = 3, fig.width = 5, fig.cap="Comparison of power between different tests for non-linear patterns.
-Main plot shows the power curves, with dots indicating human evaluations of lineups.
+Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups.
+Surrounding curves of the visual test show 500 samples of bootstrap estimation.
 Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
@@
 Main plot shows the power curves, with dots indicating human evaluations of lineups.
+Surrounding curves of the visual test show 500 samples of bootstrap estimation.
 Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
@@
 
-```{r fig.height = 3, fig.width = 5}
+Although the distribution of the regressor is classically not restricted by the assumptions of linear regression, it will have an impact on the visual effect of the residual plot.
+In our experiments, four distributions of the regressor are used to accommodate this variation.
+
+Figure \ref{fig:different-x-dist-poly-power} and Figure \ref{fig:different-x-dist-heter-power} illustrate the change of power of visual tests and conventional tests for different distribution of $X_{raw}$ used in the non-linearity model and the heteroskedasticity model.
+
+```{r different-x-dist-poly-power, fig.height = 3, fig.width = 5, fig.cap = "..."}
 vi_survey %>%
@@
   theme_light(base_size = 5) +
-  scale_color_manual(values = rev(rcartocolor::carto_pal(5, "Vivid"))) +
+  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
   scale_size_manual(values = c(1, 2, 3)) +
@@
 
-```{r fig.height = 3, fig.width = 5}
+```{r different-x-dist-heter-power, fig.height = 3, fig.width = 5, fig.cap = "..."}
 vi_survey %>%
@@
 ```
-
-(change of x_dist -> power)
 
@@
 JDSS
+
+\newpage

### 2023-01-10 paper_comparison/paper_comparison/paper_comparison.tex
update the effect of x_dist

@@
 \caption{Comparison of power between different tests for non-linear patterns.
-Main plot shows the power curves, with dots indicating human evaluations of lineups.
+Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups.
+Surrounding curves of the visual test show 500 samples of bootstrap estimation.
 Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-14-1} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-15-1} \end{center}
-
-(change of x\_dist -\textgreater{} power)
+Although the distribution of the regressor is classically not restricted
+by the assumptions of linear regression, it will have an impact on the
+visual effect of the residual plot. In our experiments, four
+distributions of the regressor are used to accommodate this variation.
+
+Figure \ref{fig:different-x-dist-poly-power} and Figure
+\ref{fig:different-x-dist-heter-power} illustrate the change of power of
+visual tests and conventional tests for different distribution of
+\(X_{raw}\) used in the non-linearity model and the heteroskedasticity
+model.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-x-dist-poly-power-1}
+
+}
+
+\caption{...}\label{fig:different-x-dist-poly-power}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-x-dist-heter-power-1}
+
+}
+
+\caption{...}\label{fig:different-x-dist-heter-power}
+\end{figure}
 
@@
 
+\newpage
+
 \bibliographystyle{tfcad}

### 2023-01-11 paper_comparison/paper_comparison/paper_comparison.Rmd
update wording up to Section 2.1. In particular, - change "one" to "we" (it's a cop out not to include us, and don't make it sound like it's about someone distant) - change Plot A of Figure 1 to Figure 1A say. - update about violation of normality. Original logical argument is not correct.

@@
 
-One can access residuals in multiple ways.
+We can assess residuals in multiple ways.
 Residuals might be plotted, as a histogram or quantile-quantile plot to examine the distribution.
-Using the classical normal linear regression model as an example, if the distribution is symmetric and unimodal, it is well-behaved.
+Using the classical normal linear regression model as an example, if the distribution is symmetric and unimodal, we consider it well-behaved.
 But if the distribution is skewed, bimodal, multimodal, or contains outliers, there is cause for concern.
@@
 Graphical summaries in which residuals are plotted against fitted values or other functions of the regressors that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals.
-As shown in Figure \ref{fig:residual-plot-common-departures}, the plot A is an ideal residual plot with residuals evenly distributed at both sides of the horizontal zero line, with no noticeable patterns.
+Figure \ref{fig:residual-plot-common-departures}A shows an ideal residual plot with residuals evenly distributed at both sides of the horizontal zero line, with no noticeable patterns.
 
@@
 Non-linearity is a type of model misspecification caused by failing to include higher order terms of the regressors in the regression equation.
-Any non-linear functional form of residuals on fitted values presented in the residual plot could be considered as an indicative of non-linearity.
-An example residual plot containing visual pattern of non-linearity is given at plot B of Figure \ref{fig:residual-plot-common-departures}.
+Any non-linear functional form of residuals on fitted values in the residual plot could be indicative of non-linearity.
+An example residual plot containing visual pattern of non-linearity is given at Figure \ref{fig:residual-plot-common-departures}B.
 One can clearly observe the "S-shape" from the residual plot as the cubic term is not captured by the misspecified model.
@@
 The usual pattern of heteroskedasticity on a residual plot is the inconsistent spread of the residuals across the horizontal axis.
-Visually, it sometimes results in the so-called "butterfly" shape as shown in the plot C of Figure \ref{fig:residual-plot-common-departures}, or the "left-triangle" and "right-triangle" shape where the smallest variance occurs at one side of the horizontal axis.
+Visually, it sometimes results in the so-called "butterfly" shape as shown in the Figure \ref{fig:residual-plot-common-departures}C, or the "left-triangle" and "right-triangle" shape where the smallest variance occurs at one side of the horizontal axis.
 
@@
 A favourable graphical summary for this task is the quantile-quantile plot.
-Hence, non-normality will not be the focus of this paper as we mainly discuss residual plots.
-For a consistent comparison, the residual plot of this departure is still presented in plot D of Figure \ref{fig:residual-plot-common-departures}.
-Besides, it is important to note that not all regression models assume normality for the error term, but a certain amount do including the classical normal linear regression model.
-In the case that the normality assumption is violated, it is expected to observe data points do not center around the horizontal axis and there is an uneven distribution of the number points at both below and above the horizontal axis.
-For example, given a skewed error distribution, there will be fewer data points and more outliers on one side of the horizontal axis as shown in plot D of Figure \ref{fig:residual-plot-common-departures}.
+As we mainly discuss residual plots, non-normality will not be the focus of this paper.
+For a consistent comparison, the residual plot of this departure is still presented in Figure \ref{fig:residual-plot-common-departures}D.
+When the number residuals below and above the horizontal axis are uneven across the local regions along the $x$-axis, we expect that the normality assumption is violated.
+For example, given a skewed error distribution, there will be fewer data points and more outliers on one side of the horizontal axis as shown in Figure \ref{fig:residual-plot-common-departures}D.
 

### 2023-01-16 paper_comparison/paper_comparison/paper_comparison.Rmd
feedback from the meeting

--- 2023-01-10 paper_comparison/paper_comparison/paper_comparison.Rmd
@@
 
-In hypothesis testing, $p$-value is defined as the probability of observing test results as least as extreme as the observed result given $H_0$ is true.
-Within the context of visual inference, by involving $k$ independent observers, $p$-value can be interpreted as the probability of having as many or more subjects detect the data plot than the observed result.
+In hypothesis testing, a $p$-value is defined as the probability of observing test results as least as extreme as the observed result given $H_0$ is true.
+Within the context of visual inference, by involving $k$ independent observers, the $p$-value can be interpreted as the probability of having as many or more subjects detect the data plot than the observed result.
 
@@
 ### Power of a visual test
+
+(definition of power, apply it to visual test)
 
@@
 
+(put easy, moderate, difficult in caption)
+
+(aspect ratio of mosaic plot)
+
+(set the threshold of RESET -> higher)
+
+(own: inspect the power of RESET for higher order only)
+
+(own: power of BP on butterfly shape, triangle shape, etc)
+
+(remove lineup id from caption)
+
 The power comparison illustrates that appropriate conventional tests reject $H_0$ more aggressively than visual tests.
@@
 ```{r eval = FALSE}
+
+p_value_cmp_dat %>%
+  filter(x_dist == "uniform") %>%
+  ggplot() +
+  ggmosaic::geom_mosaic(aes(x = ggmosaic::product(reject, conv_reject), fill = reject)) +
+  facet_grid(~type) +
+  ylab("Visual tests") +
+  xlab("Conventional tests") +
+  labs(fill = "Conventional tests") +
+  scale_fill_brewer("", palette="Dark2") +
+  theme_bw() +
+  theme(legend.position = "none")
+
+p_value_cmp_dat %>%
+  filter(x_dist == "uniform") %>%
+  filter(difficulty == "easy") %>%
+  ggplot() +
+  ggmosaic::geom_mosaic(aes(x = ggmosaic::product(reject, conv_reject), fill = reject)) +
+  facet_grid(~type) +
+  ylab("Visual tests") +
+  xlab("Conventional tests") +
+  labs(fill = "Conventional tests") +
+  scale_fill_brewer("", palette="Dark2") +
+  theme_bw() +
+  theme(legend.position = "none") +
+  ggtitle("Easy") -> p1
+
+p_value_cmp_dat %>%
+  filter(x_dist == "uniform") %>%
+  filter(difficulty == "moderate") %>%
+  ggplot() +
+  ggmosaic::geom_mosaic(aes(x = ggmosaic::product(reject, conv_reject), fill = reject)) +
+  facet_grid(~type) +
+  ylab("Visual tests") +
+  xlab("Conventional tests") +
+  labs(fill = "Conventional tests") +
+  scale_fill_brewer("", palette="Dark2") +
+  theme_bw() +
+  theme(legend.position = "none") +
+  ylab("") +
+  ggtitle("Moderate") -> p2
+
+p_value_cmp_dat %>%
+  filter(x_dist == "uniform") %>%
+  filter(difficulty == "difficult") %>%
+  ggplot() +
+  ggmosaic::geom_mosaic(aes(x = ggmosaic::product(reject, conv_reject), fill = reject)) +
+  facet_grid(~type) +
+  ylab("Visual tests") +
+  xlab("Conventional tests") +
+  labs(fill = "Conventional tests") +
+  scale_fill_brewer("", palette="Dark2") +
+  theme_bw() +
+  theme(legend.position = "none") +
+  ylab("") +
+  ggtitle("Difficult") -> p3
+
+patchwork::wrap_plots(p1, p2, p3)
+```
+
+```{r eval = FALSE}
 vi_survey %>%
@@
 
+(uniform base line : grey)
+
+(grey not work great with green)
+
+(use different color pal)
+
+(calc the diff in power and plot)
+
+(randomly subset uniform and check the change of the power)
+
 Although the distribution of the regressor is classically not restricted by the assumptions of linear regression, it will have an impact on the visual effect of the residual plot.
@@
 Figure \ref{fig:different-x-dist-poly-power} and Figure \ref{fig:different-x-dist-heter-power} illustrate the change of power of visual tests and conventional tests for different distribution of $X_{raw}$ used in the non-linearity model and the heteroskedasticity model.
+
+(discrete is the lowest for non-linearity: discreteness makes people harder to connect the shape?
+People focus on other features such as outliers instead of the shape?
+If this is the case, more evaluations may lead to a reject.
+We need some examples (log(E) > 4) here)
+
+```{r eval = FALSE}
+vi_survey %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
+  filter(type == "polynomial") %>%
+  filter(x_dist == "even_discrete") %>%
+  filter(log(effect_size) > 4) %>%
+  filter(p_value > 0.05) %>%
+  .$unique_lineup_id %>%
+  unique()
+```
+
+```{r eval = FALSE}
+VI_MODEL$plot_lineup(vi_lineup$poly_64$data, 
+                     remove_grid_line = TRUE, 
+                     theme = theme_light(),
+                     remove_axis = TRUE)
+```
+
+In the case of non-linearity model,
+
+It can be observed that because of the visibility of the visual pattern, the discrete uniform distribution
+
+visual test
 

### 2023-01-17 paper_comparison/paper_comparison/paper_comparison.Rmd
update main sections

--- 2023-01-11 paper_comparison/paper_comparison/paper_comparison.Rmd
@@
 
-```{r residual-plot-common-departures, fig.width = 10, fig.height = 2.5, fig.cap = "Example fitted vs residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
+```{r residual-plot-common-departures, fig.width = 10, fig.height = 2.5, fig.cap = "Example residual vs fitted plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
 The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted plot."}
@@
                caption  = 'Statistical significance testing for departures from good residuals for plots in Figure \\ref{fig:residual-plot-common-departures}.
-Shown are the $p$-values calculated for the conventional RESET, the Breusch-Pagan and the Shapiro–Wilk tests.
+Shown are the $p$-values calculated for the RESET, the Breusch-Pagan and the Shapiro–Wilk tests.
 The good residual plot (A) is judged a good residual plot, as expected, by all tests.
@@
 Later, @buja_statistical_2009 proposed the lineup protocol as a visual test inspired by the "police lineup" or "identity parade" which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people.
-The protocol consists of $m$ randomly placed plots, where one plot is the data plot, and the remaining $m - 1$ plots have the identical graphical procedure except the data has been replaced with data consistent with $H_0$.
+The protocol consists of $m$ randomly placed plots, where one plot is the data plot, and the remaining $m - 1$ null plots have the identical graphical procedure except the data has been replaced with data consistent with $H_0$.
 Then, an observer who have not seen the data plot will be asked to point out the most different plot from the lineup.
@@
 Their results suggest that visual tests are capable of testing the significance of a single regressor with a similar power as a t-test, though they express that in general it is unnecessary to use visual inference if there exists a conventional test, and they do not expect the visual test to perform equally well as the conventional test.
-In their third experiment, where there is not a conventional test, visual test outperforms the conventional test for a large margin.
+(contradict to our claim?
+expand?) In their third experiment, where there is not a conventional test, visual test outperforms the conventional test for a large margin.
 This is encouraging, as it promotes the use of visual inference in situations where there are no existing statistical testing procedures.
@@
 Data used in the $m - 1$ null plots needs to be simulated.
-In the context of regression diagnostics, sampling data from $H_0$ is equivalent to sampling data from the assumed model.
+In regression diagnostics, sampling data consistent with $H_0$ is equivalent to sampling data from the assumed model.
 As @buja_statistical_2009 suggested, $H_0$ is usually a composite hypothesis controlled by nuisance parameters.
@@
 
-In hypothesis testing, $p$-value is defined as the probability of observing test results as least as extreme as the observed result given $H_0$ is true.
-Within the context of visual inference, by involving $k$ independent observers, $p$-value can be interpreted as the probability of having as many or more subjects detect the data plot than the observed result.
-
-Let $X_j = \{0,1\}$ be a binomial random variable denoting whether subject $j$ correctly detecting the data plot, and $X = \sum_{j=1}^{K}X_j$ be the number of observers correctly picking the data plot.
+In hypothesis testing, a $p$-value is defined as the probability of observing test results as least as extreme as the observed result given $H_0$ is true.
+Within the context of visual inference, by involving $k$ independent observers, the $p$-value can be interpreted as the probability of having as many or more subjects detect the data plot than the observed result.
+
+Let $X_j = \{0,1\}$ be a Bernoulli random variable denoting whether subject $j$ correctly detecting the data plot, and $X = \sum_{j=1}^{K}X_j$ be the number of observers correctly picking the data plot.
 Then, by imposing a relatively strong assumption on the visual test that all $K$ evaluations are fully independent, under $H_0$, $X \sim \mathrm{Binom}_{K,1/m}$.
@@
 
-The power of a model misspecification test specifies the chance of a misspecified regression model being detected given a certain alternative hypothesis.
-It is an important indicator when one concerns if the regression model is misspecified.
-Though, people might usually be more interested in knowing how much the residuals depart from the model assumptions and if the departure is of practical significance.
-
-As discussed in @majumder_validation_2013, individual's skill may affect the number of observers who identify the data plot from the lineup.
-Thus, the power of a visual test could depend on the subject-specific abilities.
-Previously, it is addressed by modelling the probability of a subject $j$ correctly picking the data plot from a lineup $l$ using a mixed-effect logistic regression with the subject being treated as a random effect [@majumder_validation_2013].
-However, in the multiple selections scenario, having this probability is insufficient to determine the power of a visual test as it doesn't provide information about the number of selections made by the subject for p-value calculation.
-
-Instead, we estimate the probability of a lineup being rejected directly by assuming the individual skill has negligible impact on the variation of the power of a lineup.
-The assumption is made to simplify the model structure, thus mitigating the need for expensive large-scale experiments to estimate the complex covariance matrix.
+The power of a model misspecification test is the probability that the null hypothesis is rejected given the regression model is misspecified.
+It is an important indicator when one is concerned about whether model assumptions have been violated.
+Although in practice, one might be more interested in knowing how much the residuals deviate from the model assumptions, and whether this deviation is of practical significance.
+
+As discussed in @majumder_validation_2013, the power of a visual test may depend on the ability of the particular subject, as the skill of the individual may affect the number of observers who identify the data plot from the lineup.
+Previously, it is addressed by modelling the probability of a subject $j$ correctly picking the data plot from a lineup $l$ using a mixed-effect logistic regression, with subjects treated as random effects [@majumder_validation_2013].
+However, in the multiple selections scenario, having this probability is not sufficient to determine the power of a visual test because it does not provide information about the number of selections made by the subject for the calculation of the p-value.
+
+Instead, we directly estimate the probability of a lineup being rejected by assuming that individual skill has negligible effect on the variation of the power.
+The assumption is to simplify the model structure, thereby obviate costly large-scale experiments to estimate complex covariance matrices.
 The model is a logistic regression with the natural logarithm of the effect as the only regressor formulated as
@@
 \begin{equation} \label{eq:logistic-regression-1-1}
-Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(E)),
+Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log(E)),
 \end{equation}
@@
 
-```{=tex}
-\begin{equation} \label{eq:effect-size-ex1}
-E = \frac{1}{2\sigma^2}\boldsymbol{X}_b'\boldsymbol{R}_a'(diag(\boldsymbol{R}_a))^{-1}\boldsymbol{R}_a\boldsymbol{X}_b,
-\end{equation}
-```
-\noindent where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$.
+\small
+
+\begin{gather}  \label{eq:effect-size}
+E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V})|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V})^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V})^{-1}\boldsymbol{\mu}_z\right),\\
+\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{X}_a(\boldsymbol{X}_a^T\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a^T,
+\end{gather}
+
+\normalsize
+
+\noindent where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,  $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of regressors including the intercept used in the regression equation, $\sigma^2$ is the assumed variance of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual variance of the error term, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of the residuals with $\boldsymbol{Z}$ be any variables leave out by the model and $\boldsymbol{\beta}_z$ be the corresponding coefficients.
 
 To study various factors contributing to the power of the visual test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors.
-This includes [expansion].
+These include the distribution of the regressor and the type of the simulated model.
 
@@
 \begin{align} \label{eq:nonlinearity-model}
-\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
-\boldsymbol{x} = g(\boldsymbol{x}_{raw}, 1), \\
-\boldsymbol{z} = g(\boldsymbol{z}_{raw}, 1), \\
-\boldsymbol{z}_{raw} = He_j(g(\boldsymbol{x}, 2)),
+\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
+\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
+\boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\
+\boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)),
 \end{align}
@@
 \begin{align} \label{eq:heter-model}
-\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
-\boldsymbol{x} = g(\boldsymbol{x}_{raw}, 1)\\
-\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + (2 - |a|)(\boldsymbol{x} - a)^2b \boldsymbol{I}), \\
+\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
+\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
+\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}, 1 + (2 - |a|)(\boldsymbol{x} - a)^2b \boldsymbol{I}), 
 \end{align}
@@
 
+In the following analysis, lineups with $X_{raw} \sim U(-1,1)$ will be the focus of next two sections, because visual patterns are more likely to be revealed with uniform distribution.
+Besides, six more evaluations are collected for the uniform distribution in experiment III, which will produce more reliable and stable results.
+Analysis of other distributions can be found in Section \ref{effect-of-the-distribution-of-the-regressor}.
+
 ## Comparison of the power of different tests
@@
 Figure \ref{fig:polypower} shows the estimated power of visual test on lineups produced by the non-linearity model with $X_{raw} \sim U(-1,1)$, against the natural logarithm of the effect $log_e(E)$, with a 5% significance level.
-Uniform distribution is the focus of this section because visual patterns are more likely to be revealed.
-Besides, six more evaluations are collected for the uniform distribution in experiment III, which will produce more reliable and stable power curves.
-Analysis of other distributions can be found in Section \ref{effect-of-the-distribution-of-the-regressor}.
 At the bottom of the figure \ref{fig:polypower}, there are a sequence of example residual plots with increasing levels of $log_e(E)$.
@@
 
-For the non-linearity model, the power curve of RESET test climbs aggressively from 13% to around 70% as $log_e(E)$ increases from 0 to 2, while power of other tests respond inactively to the change of effect, showing that REST test is way more sensitive to the type of model defects that being considered.
+For the non-linearity model, the power curve of RESET test climbs aggressively from 13% to around 70% as $log_e(E)$ increases from 0 to 2, while power of other tests respond inactively to the change of effect, showing that RESET test is way more sensitive to the type of model defects that being considered.
 Meanwhile, no noticeable visual features can be spotted from the example residual plots.
@@
   scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
-  xlab("Effect size (log e)") +
+  xlab("Effect size (log E)") +
   ylab("Power") +
@@
   scale_size_manual(values = c(1, 2, 3)) +
-  xlab("Effect size (log e)") +
+  xlab("Effect size (log E)") +
   ylab("Power") +
@@
 
-The power comparison illustrates that appropriate conventional tests reject $H_0$ more aggressively than visual tests.
-In this section, we explore how often they agree with each other by investigate the rejection rates based on $p$-values.
-
-Depending on the natural logarithm of the effect, lineups are classified into three categories: easy (> 3.5), moderate (1.5 - 3.5) and difficult (< 1.5).
-Figure \ref{fig:p-value-comparison} provides a mosaic plot showing the rejection rate of visual tests and conventional tests at different difficulty levels.
-
-### Easy lineup
-
-For easy lineups produced by the non-linearity model, 65% are rejected by both the conventional and visual tests, while 9% are accepted by both tests.
-Thus, 74% of the time two tests will agree with each other.
-There are 16% of lineups rejected by conventional tests only.
-
-Interestingly, in spite of the greater power of conventional tests, 10% of the lineups are rejected by visual tests only.
-Recall that the RESET test used in our analysis includes powers of fitted values up to four.
-However, the "M" and the "Triple-U" shapes constructed from the Hermite polynomials contain power of the regressor up to six and 18 respectively.
-In fact, by including the power of fitted values up to six in the RESET test, all these cases will be rejected.
-We also can observe from the example plot displayed in Figure \ref{fig:easy-poly-example} that panel three is distinctly different from others, exhibiting a "M" shape.
-This clearly suggests users of the RESET test and any other residual-based conventional tests that require the specification of variables of interest to check the residual plot before conducting the test, such that the correct choice of variables can be made.
-
-```{r easy-poly-example, fig.height = 5, fig.cap = "Lineup poly-270 with $log_e(E) = 3.89$ produced by the non-linearity model used in experiment I and III.
-It is rejected by the visual test but not by the RESET test.
-The data plot at panel three exhibits strong non-linear pattern."}
-VI_MODEL$plot_lineup(vi_lineup$poly_270$data, 
-                     remove_grid_line = TRUE, 
-                     theme = theme_light(),
-                     remove_axis = TRUE)
-```
-
-For easy lineups produced by the heteroskedasticity model, 62% are rejected by both the conventional test and the visual test, and 7% are accepted by both, result in 69% agreement rate.
-Lineups only rejected by conventional test are 30%.
-1% of lineups are accepted by the conventional test but rejected by the visual test.
-
-Among those 1%, which are two lineups, one shows detectable heteroskedasticity pattern as displayed in Figure \ref{fig:easy-heter-example1}.
-Another contain data plot with an noticeable outlier, but the "Right-triangle" shape is not very strong as shown in Figure \ref{fig:easy-heter-example2}.
-The $p$-values of Breusch-Pagan test for these two lineups are $0.055$ and $0.135$.
-Compared to the visual test, its $p$-values are $0.026$ and $0.012$.
-The Breusch-Pagan test should reject the one with obvious heteroskedasticity signal, but it yields a $p$-value slightly above 5%.
-
-```{r easy-heter-example1, fig.height = 5, fig.cap = 'Lineup heter-331 with $log_e(E) = 4.02$ produced by the heteroskedasticity model used in experiment II and III.
-It is rejected by the visual test but not by the  Breusch-Pagan test.
-The data plot at panel 17 contains a "Butterfly" shape, indicating the presence of heteroskedasticity.'}
-VI_MODEL$plot_lineup(vi_lineup$heter_331$data, 
-                     remove_grid_line = TRUE, 
-                     theme = theme_light(),
-                     remove_axis = TRUE)
-```
-
-```{r easy-heter-example2, fig.height = 5, fig.cap = "Lineup heter-536 with $log_e(E) = 4.13$ produced by the heteroskedasticity model used in experiment II.
-It is rejected by the visual test but not by the  Breusch-Pagan test.
-The data plot at panel 18 exhibits a vague heteroskedasticity pattern, but it contains a distinct outlier which makes it attractive to subjects."}
-VI_MODEL$plot_lineup(vi_lineup$heter_536$data, 
-                     remove_grid_line = TRUE, 
-                     theme = theme_light(),
-                     remove_axis = TRUE)
-```
-
-### Moderate lineup
-
-With the increase of the difficulty level, the sensitivity of conventional tests enlarge the difference in rejection rates.
-For moderate lineup produced by the non-linearity model, 8% and 28% are rejected and accepted by both tests respectively.
-We can observe a decline in agreement rate to 36%.
-A large proportion of lineups are accepted by the visual test and rejected by the conventional test, which is around 63%.
-The remaining 1% are rejected by conventional tests only, which suffer the same issue caused by the RESET test as mentioned earlier.
-
-For those produced by the heteroskedasticy model, the corresponding numbers are 23%, 31%, 54%, 45% and 1%.
-There is only one lineup rejected by the visual test but accepted by the conventional test.
-It does not show any heteroskedasticity patterns, but contains a few outliers at the bottom of the data plot as illustrated in panel 14 of Figure \ref{fig:moderate-heter-example1}.
-Two out of five subjects correctly identify this lineup yielding a $p$-value equal to $0.044$.
-We consider this instance is rejected by chance.
-If a different set of null plots are used, the data plot may not be identified.
-
-```{r moderate-heter-example1, fig.height = 5, fig.cap = "Lineup heter-219 with $log_e(E) = 1.61$ produced by the heteroskedasticity model used in experiment II.
-It is rejected by the visual test but not by the  Breusch-Pagan test.
-Though the data plot at panel 14 shows no heteroskedasticity patterns, it contains a few outliers at the bottom of the panel."}
-VI_MODEL$plot_lineup(vi_lineup$heter_219$data, 
-                     remove_grid_line = TRUE, 
-                     theme = theme_light(),
-                     remove_axis = TRUE)
-```
-
-### Difficult lineup
-
-Visual test is insensitive to lineups with small effect.
-When it comes to difficult level, none of the lineups produced by the non-linearity model are rejected by the visual test.
-There are 73% of lineups accepted by both tests, and 26% rejected by the conventional test.
-The agreement rate rise up to 73%.
-
-For lineups produced by the heteroskedasticity model, 3% are rejected by both tests, and 71% are accepted by both tests.
-There are 24% of lineups rejected by the conventional test only.
-Besides, only two lineups, which is 2% of all the difficult lineups, are rejected by the visual test only.
-
-As shown in \ref{fig:difficult-heter-example1} and \ref{fig:difficult-heter-example2}, residual plots of both lineups are constructed with $X_{raw} \sim lognormal(0, 0.6^2)/3$.
-Though the designed patterns are "Left-triangle" and "Butterfly" shape, both of them are invisible due to the small effect.
-The visual effect are dominated by the skewed distribution.
-Further, no distinguishable visual patterns, such as outliers, are detectable from the data plot of these two lineups, which are panel 5 and panel 15.
-The data plots in these two lineups are not the most different plot from our point of view either.
-However, three out of five and two out of five subjects correctly identify the data plot for these two lineups respectively.
-But the reasons for selecting the plot provided by the subjects are mixed and inconsistent.
-Hence, with the current data, we can not explain why subjects can correctly identify the data plot of these two lineups.
-Collecting more data in future study may help to answer this question.
-
-```{r difficult-heter-example1, fig.height = 5, fig.cap = "Lineup heter-224 with $log_e(E) = 1.32$ produced by the heteroskedasticity model used in experiment II.
-It is rejected by the visual test but not by the  Breusch-Pagan test.
-Although two out of five subjects correctly identify the data plot at panel 5, it does not exhibit any visually distinguishable pattern."}
-VI_MODEL$plot_lineup(vi_lineup$heter_224$data, 
-                     remove_grid_line = TRUE, 
-                     theme = theme_light(),
-                     remove_axis = TRUE)
-```
-
-```{r difficult-heter-example2, fig.height = 5, fig.cap = "Lineup heter-415 with $log_e(E) = -0.31$ produced by the heteroskedasticity model used in experiment II.
-It is rejected by the visual test but not by the  Breusch-Pagan test.
-Panel 15 is the data plot but it does not exhibit any visually distinguishable pattern."}
-VI_MODEL$plot_lineup(vi_lineup$heter_415$data, 
-                     remove_grid_line = TRUE, 
-                     theme = theme_light(),
-                     remove_axis = TRUE)
-```
-
-```{r eval = FALSE}
-
-vi_survey %>% filter(unique_lineup_id %in% 
-                       (p_value_cmp_dat %>%
-  filter(type == "heteroskedasticity", difficulty == "difficult") %>%
-  filter(conventional_p_value >= 0.05, p_value < 0.05) %>%
-  pull(unique_lineup_id))) %>%
-  View()
-
-```
-
-```{r eval = FALSE}
-p_value_cmp_dat %>%
-  filter(type == "non-linearity", difficulty == "difficult") %>%
-  count(conventional_p_value < 0.05, p_value < 0.05) %>%
-  mutate(n = n / sum(n))
-
-p_value_cmp_dat %>%
-  filter(type == "heteroskedasticity", difficulty == "difficult") %>%
-  count(conventional_p_value < 0.05, p_value < 0.05) %>%
-  mutate(n = n / sum(n)) 
-```
-
-```{r p-value-comparison, fig.width=10, fig.height=5, fig.cap="Rejection rate of visual test compared to conventional test on lineups produced by heteroskedasticity and non-linearity model at different levels of difficulty.
+```{r p-value-comparison, fig.width=10, fig.height=5, fig.cap="Rejection rate of visual test compared to conventional test on lineups produced by heteroskedasticity and non-linearity model.
 Based on the rejection decision made by the visual test and the conventional test, the mosaic plot shows the proportion of lineups in each category."}
@@
   filter(!attention_check) %>%
+  filter(x_dist == "uniform") %>%
   mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-  mutate(difficulty = c("difficult", "moderate", "easy")[1 + 
-                                                           (log(effect_size) > 1.5) +
-                                                           (log(effect_size) > 3.5)]) %>%
-  mutate(difficulty = factor(difficulty, levels = c("easy", "moderate", "difficult"))) %>%
   left_join(vi_survey %>% 
@@
 p_value_cmp_dat %>%
-  filter(difficulty == "easy") %>%
   ggplot() +
@@
   labs(fill = "Conventional tests") +
-  scale_fill_brewer("", palette="Dark2") +
+  scale_fill_brewer("", palette = "Dark2") +
   theme_bw() +
   theme(legend.position = "none") +
-  ggtitle("Easy") -> p1
-
-p_value_cmp_dat %>%
-  filter(difficulty == "moderate") %>%
+  coord_fixed()
+```
+
+(remove lineup id from caption)
+
+The power comparison illustrates that appropriate conventional tests reject $H_0$ more aggressively than visual tests.
+In this section, we explore how often they agree with each other by investigate the rejection rates based on $p$-values.
+
+Figure \ref{fig:p-value-comparison} provides a mosaic plot showing the rejection rate of visual tests and conventional tests for both the non-linearity model and the heteroskedasticity model.
+
+For lineups produced by the non-linearity model, 25% are rejected by both the conventional and visual tests, while 42% are accepted by both tests.
+Thus, 69% of the time two tests will agree with each other.
+There are 26% of lineups rejected by conventional tests only.
+
+Interestingly, in spite of the greater power of conventional tests, 7% of the lineups are rejected by visual tests only.
+Recall that the RESET test used in our analysis includes powers of fitted values up to four.
+However, the "M" and the "Triple-U" shapes constructed from the Hermite polynomials contain power of the regressor up to six and 18 respectively.
+In fact, by including the power of fitted values up to six in the RESET test, all these cases will be rejected.
+We also can observe from the example plot displayed in Figure \ref{fig:poly-example} that panel three is distinctly different from others, exhibiting a "M" shape.
+This clearly suggests users of the RESET test and any other residual-based conventional tests that require the specification of variables of interest to check the residual plot before conducting the test, such that the correct choice of variables can be made.
+
+```{r poly-example, fig.height = 5, fig.cap = "Lineup poly-270 with $log_e(E) = 3.89$ produced by the non-linearity model used in experiment I and III.
+It is rejected by the visual test but not by the RESET test.
+The data plot at panel three exhibits strong non-linear pattern."}
+VI_MODEL$plot_lineup(vi_lineup$poly_270$data, 
+                     remove_grid_line = TRUE, 
+                     theme = theme_light(),
+                     remove_axis = TRUE)
+```
+
+```{r}
+poly_conv_sim %>%
+  filter(x_dist == "uniform", shape %in% c(3, 4)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  select(shape, log_effect_size, (RESET3_p_value:RESET9_p_value)) %>%
+  pivot_longer(RESET3_p_value:RESET9_p_value) %>%
+  mutate(name = gsub("_p_value", " ", name)) %>%
+  mutate(name = gsub("RESET", "", name)) %>%
+  mutate(name = gsub("test", "", name)) %>%
+  mutate(reject = as.numeric(value < 0.05)) %>%
   ggplot() +
-  ggmosaic::geom_mosaic(aes(x = ggmosaic::product(reject, conv_reject), fill = reject)) +
-  facet_grid(~type) +
-  ylab("Visual tests") +
-  xlab("Conventional tests") +
-  labs(fill = "Conventional tests") +
-  scale_fill_brewer("", palette="Dark2") +
-  theme_bw() +
-  theme(legend.position = "none") +
-  ylab("") +
-  ggtitle("Moderate") -> p2
-
-p_value_cmp_dat %>%
-  filter(difficulty == "difficult") %>%
+  geom_smooth(aes(log_effect_size, reject, col = name), 
+              se = FALSE, 
+              method = "glm",
+              method.args = list(family = binomial())) +
+  xlab("Effect size (log E)") +
+  ylab("Power") +
+  scale_color_manual(values = rev(rcartocolor::carto_pal(7, "RedOr"))) +
+  theme_light() +
+  labs(col = "RESET test order") +
+  facet_wrap(~shape)
+```
+
+```{r}
+heter_conv_sim %>%
+  filter(x_dist == "uniform") %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  mutate(reject = as.numeric(BP_p_value < 0.05)) %>%
+  mutate(a = c("left-triangle", "butterfly", "right-triangle")[a + 2]) %>%
   ggplot() +
-  ggmosaic::geom_mosaic(aes(x = ggmosaic::product(reject, conv_reject), fill = reject)) +
-  facet_grid(~type) +
-  ylab("Visual tests") +
-  xlab("Conventional tests") +
-  labs(fill = "Conventional tests") +
-  scale_fill_brewer("", palette="Dark2") +
-  theme_bw() +
-  theme(legend.position = "none") +
-  ylab("") +
-  ggtitle("Difficult") -> p3
-
-patchwork::wrap_plots(p1, p2, p3)
-```
+  geom_smooth(aes(log_effect_size, reject, col = factor(a)), 
+              se = FALSE, 
+              method = "glm",
+              method.args = list(family = binomial())) +
+  xlab("Effect size (log E)") +
+  ylab("Power") +
+  scale_color_manual(values = rcartocolor::carto_pal(3, "Vivid")) +
+  theme_light() +
+  labs(col = "Heteroskedasticity shape") +
+  ggtitle("Conventional test")
+```
+
+```{r}
+vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(x_dist == "uniform", type == "heteroskedasticity") %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  mutate(reject = as.numeric(p_value < 0.05)) %>%
+  mutate(a = c("left-triangle", "butterfly", "right-triangle")[a + 2]) %>%
+  ggplot() +
+  geom_smooth(aes(log_effect_size, reject, col = factor(a)), 
+              se = FALSE, 
+              method = "glm",
+              method.args = list(family = binomial())) +
+  xlab("Effect size (log E)") +
+  ylab("Power") +
+  scale_color_manual(values = rcartocolor::carto_pal(3, "Vivid")) +
+  theme_light() +
+  labs(col = "Heteroskedasticity shape") +
+  ggtitle("Visual test")
+```
+
+For lineups produced by the heteroskedasticity model, 55% are rejected by both the conventional test and the visual test, and 24% are accepted by both, result in 79% agreement rate.
+Percentage of lineups only rejected by conventional test are 20%.
+Besides, only 1 lineup, which is less than 1% of all the lineups, are rejected by the visual test only.
+This lineup shows relatively strong heteroskedasticity pattern as displayed in Figure \ref{fig:heter-example}.
+The visual test rejects it with a $p\text{-value} = 0.026$, but the Breusch-Pagan test produces a $p$-value slightly above 5%.
+However, as we can see, it is an exceptional case.
+
+```{r heter-example, fig.height = 5, fig.cap = 'Lineup heter-331 with $log_e(E) = 4.02$ produced by the heteroskedasticity model used in experiment II and III.
+It is rejected by the visual test but not by the  Breusch-Pagan test.
+The data plot at panel 17 contains a "Butterfly" shape, indicating the presence of heteroskedasticity.'}
+VI_MODEL$plot_lineup(vi_lineup$heter_331$data, 
+                     remove_grid_line = TRUE, 
+                     theme = theme_light(),
+                     remove_axis = TRUE)
+```
+
+## Effect of the distribution of the regressor
+
+(calc the diff in power and plot)
+
+(randomly subset uniform and check the change of the power)
+
+Although the distribution of the regressor is classically not restricted by the assumptions of linear regression, it will have an impact on the visual effect of the residual plot.
+In our experiments, four distributions of the regressor are used to accommodate this variation.
+
+Figure \ref{fig:different-x-dist-poly-power} and Figure \ref{fig:different-x-dist-heter-power} illustrate the change of power of visual tests and conventional tests for different distribution of $X_{raw}$ used in the non-linearity model and the heteroskedasticity model.
+We focus on $log(E) > 2$ as the visual patterns become visible.
+
+For the non-linearity model, we do not observe significant difference of power of visual test between distributions, except for the discrete uniform distribution.
+It could be due to the fact that humans have difficulties in recognising visual patterns from discreteness, which makes the shape disconnected and incomplete (ref?).
+In terms of conventional tests, the discrete uniform  distribution is instead the one with the greatest power, followed by the lognormal distribution, uniform distribution and the normal distribution.
+
+For the heteroskedasticity model, the power of visual test on lineups with uniform distribution has the greatest power.
+This is as expected since other three distributions could reduce the chance of revealing the underlying visual pattern because of uneven data points.
+Although the power of visual test under the discrete uniform  distribution is the lowest in the case of non-linearity model, it has a relatively great power this time.
+Considering for heteroskedasticity, the visual patterns are usually detected by connecting the maximum and minimum residuals separately at different fitted values.
+It will not be greatly affected by the discrete uniform distribution compared to the uniform distribution.
+Visual tests have low power on lineups with the normal distribution and the lognormal distribution.
+This suggests the use of transformation of fitted values in reading residual plots.
+In the case of conventional tests, the discrete uniform distribution consistently doing great, followed by the uniform distribution, the normal distribution and the lognormal distribution.
+
+In regards of the type of the simulated model, the power of visual tests under different distributions of regressor are always lower than the corresponding conventional tests, consistent with the result from the previous sections.
 
@@
   filter(!attention_check) %>%
-  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-  left_join(vi_survey %>% 
-           count(unique_lineup_id) %>%
-           rename(num_subject = n)) %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
-  mutate(conv_reject = ifelse(conventional_p_value < 0.05, "Reject", "Not"),
-         reject = ifelse(p_value < 0.05, "Reject", "Not")) %>%
-  filter(type == "non-linearity", conv_reject == "Not", reject == "Reject") %>%
-  View()
-```
-
-## Effect of the distribution of the regressor
-
-Although the distribution of the regressor is classically not restricted by the assumptions of linear regression, it will have an impact on the visual effect of the residual plot.
-In our experiments, four distributions of the regressor are used to accommodate this variation.
-
-Figure \ref{fig:different-x-dist-poly-power} and Figure \ref{fig:different-x-dist-heter-power} illustrate the change of power of visual tests and conventional tests for different distribution of $X_{raw}$ used in the non-linearity model and the heteroskedasticity model.
-
-```{r different-x-dist-poly-power, fig.height = 3, fig.width = 5, fig.cap = "..."}
-vi_survey %>%
+  filter(type == "polynomial") %>%
+  filter(x_dist == "even_discrete") %>%
+  filter(log(effect_size) > 4) %>%
+  filter(p_value > 0.05) %>%
+  .$unique_lineup_id %>%
+  unique()
+```
+
+```{r eval = FALSE}
+VI_MODEL$plot_lineup(vi_lineup$poly_64$data, 
+                     remove_grid_line = TRUE, 
+                     theme = theme_light(),
+                     remove_axis = TRUE)
+```
+
+```{r different-x-dist-poly-power, fig.height = 3, fig.width = 5, fig.cap = "Power of conventional tests and visual tests on lineups produced by the non-linearity model under different distributions of regressor."}
+
+dat_different_dist <- poly_conv_sim %>% 
+  mutate(log_effect_size = log(effect_size)) %>%
+  select(x_dist, log_effect_size, RESET4_p_value) %>%
+  rename(RESET_p_value = RESET4_p_value) %>%
+  pivot_longer(RESET_p_value) %>%
+  mutate(x_dist = ifelse(x_dist == "even_discrete", 
+                                       "Discrete", 
+                                       stringr::str_to_title(x_dist))) %>%
+  mutate(name = gsub("_p_value", " test", name)) %>%
+  mutate(reject = value < 0.05) %>%
+  mutate(conv_or_not = "Conventional") %>%
+  bind_rows(vi_survey %>%
   group_by(unique_lineup_id) %>%
@@
                          stringr::str_to_title(x_dist))) %>%
+  mutate(conv_or_not = "Visual"))
+
+dat_different_dist %>%
   ggplot() +
-  geom_smooth(data = mutate(poly_conv_sim, 
-                            log_effect_size = log(effect_size)) %>%
-                select(x_dist, log_effect_size, RESET4_p_value) %>%
-                rename(RESET_p_value = RESET4_p_value) %>%
-                pivot_longer(RESET_p_value) %>%
-                mutate(x_dist = ifelse(x_dist == "even_discrete", 
-                                       "Discrete", 
-                                       stringr::str_to_title(x_dist))) %>%
-                mutate(name = gsub("_p_value", " test", name)), 
-              aes(log_effect_size, 
-                  as.numeric(value < 0.05), 
-                  col = x_dist,
-                  linetype = "RESET"), 
-              method = "glm", 
-              method.args = list(family = binomial),
-              se = FALSE) +
   geom_smooth(aes(log_effect_size, 
                   reject, 
-                  col = x_dist,
-                  linetype = "Visual"), 
+                  col = x_dist), 
               method = "glm", 
@@
   theme_light(base_size = 5) +
-  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
+  scale_color_manual(values = rcartocolor::carto_pal(4, "Safe")[c(1,3,2,4)]) +
   scale_size_manual(values = c(1, 2, 3)) +
-  scale_linetype_manual(values = c("longdash", "solid")) +
-  guides(linetype = guide_legend(nrow = 2), col = guide_legend(nrow = 2)) +
-  theme(legend.position = "bottom", legend.key.width = unit(1.5, "cm")) +
-  xlab("Effect size (log e)") +
+  xlab("Effect size (log E)") +
   ylab("Power") +
-  labs(col = "Distribution of regressor", linetype = "Test")
-```
-
-```{r different-x-dist-heter-power, fig.height = 3, fig.width = 5, fig.cap = "..."}
-vi_survey %>%
+  labs(col = "Distribution of regressor", linetype = "Test") +
+  facet_wrap(~conv_or_not)
+```
+
+```{r different-x-dist-heter-power, fig.height = 3, fig.width = 5, fig.cap = "Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model under different distributions of regressor."}
+
+dat_different_dist <- heter_conv_sim %>% 
+  mutate(log_effect_size = log(effect_size)) %>%
+  select(x_dist, log_effect_size, BP_p_value) %>%
+  pivot_longer(BP_p_value) %>%
+  mutate(x_dist = ifelse(x_dist == "even_discrete", 
+                                       "Discrete", 
+                                       stringr::str_to_title(x_dist))) %>%
+  mutate(name = gsub("_p_value", " test", name)) %>%
+  mutate(reject = value < 0.05) %>%
+  mutate(conv_or_not = "Conventional") %>%
+  bind_rows(vi_survey %>%
   group_by(unique_lineup_id) %>%
@@
                          stringr::str_to_title(x_dist))) %>%
+  mutate(conv_or_not = "Visual"))
+
+dat_different_dist %>%
   ggplot() +
-  geom_smooth(data = mutate(heter_conv_sim, 
-                            log_effect_size = log(effect_size)) %>%
-                select(x_dist, log_effect_size, BP_p_value) %>%
-                pivot_longer(BP_p_value) %>%
-                mutate(x_dist = ifelse(x_dist == "even_discrete", 
-                                       "Discrete", 
-                                       stringr::str_to_title(x_dist))) %>%
-                mutate(name = gsub("_p_value", " test", name)), 
-              aes(log_effect_size, 
-                  as.numeric(value < 0.05), 
-                  col = x_dist,
-                  linetype = "BP"), 
-              method = "glm", 
-              method.args = list(family = binomial),
-              se = FALSE) +
   geom_smooth(aes(log_effect_size, 
                   reject, 
-                  col = x_dist,
-                  linetype = "Visual"), 
+                  col = x_dist), 
               method = "glm", 
@@
   theme_light(base_size = 5) +
-  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
+  scale_color_manual(values = rcartocolor::carto_pal(4, "Safe")[c(1,3,2,4)]) +
   scale_size_manual(values = c(1, 2, 3)) +
-  scale_linetype_manual(values = c("longdash", "solid")) +
-  guides(linetype = guide_legend(nrow = 2), col = guide_legend(nrow = 2)) +
-  theme(legend.position = "bottom", legend.key.width = unit(1.5, "cm")) +
-  xlab("Effect size (log e)") +
+  xlab("Effect size (log E)") +
   ylab("Power") +
-  labs(col = "Distribution of regressor", linetype = "Test")
+  labs(col = "Distribution of regressor", linetype = "Test") +
+  facet_wrap(~conv_or_not)
 ```

### 2023-01-17 paper_comparison/paper_comparison/paper_comparison.tex
update main sections

@@
 
-One can access residuals in multiple ways. Residuals might be plotted,
-as a histogram or quantile-quantile plot to examine the distribution.
-Using the classical normal linear regression model as an example, if the
-distribution is symmetric and unimodal, it is well-behaved. But if the
-distribution is skewed, bimodal, multimodal, or contains outliers, there
-is cause for concern. One could also inspect the distribution by
-conducting a goodness of fit test, such as the Shapiro-Wilk Normality
-test \citep{shapiro1965analysis}.
+We can assess residuals in multiple ways. Residuals might be plotted, as
+a histogram or quantile-quantile plot to examine the distribution. Using
+the classical normal linear regression model as an example, if the
+distribution is symmetric and unimodal, we consider it well-behaved. But
+if the distribution is skewed, bimodal, multimodal, or contains
+outliers, there is cause for concern. One could also inspect the
+distribution by conducting a goodness of fit test, such as the
+Shapiro-Wilk Normality test \citep{shapiro1965analysis}.
 
@@
 
-\caption{Example fitted vs residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
+\caption{Example residual vs fitted plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
 The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted plot.}\label{fig:residual-plot-common-departures}
@@
 to residuals are referred to as standard residual plots in
-\citet{cook1982residuals}. As shown in Figure
-\ref{fig:residual-plot-common-departures}, the plot A is an ideal
-residual plot with residuals evenly distributed at both sides of the
-horizontal zero line, with no noticeable patterns.
+\citet{cook1982residuals}. Figure
+\ref{fig:residual-plot-common-departures}A shows an ideal residual plot
+with residuals evenly distributed at both sides of the horizontal zero
+line, with no noticeable patterns.
 
@@
 include higher order terms of the regressors in the regression equation.
-Any non-linear functional form of residuals on fitted values presented
-in the residual plot could be considered as an indicative of
-non-linearity. An example residual plot containing visual pattern of
-non-linearity is given at plot B of Figure
-\ref{fig:residual-plot-common-departures}. One can clearly observe the
+Any non-linear functional form of residuals on fitted values in the
+residual plot could be indicative of non-linearity. An example residual
+plot containing visual pattern of non-linearity is given at Figure
+\ref{fig:residual-plot-common-departures}B. One can clearly observe the
 ``S-shape'' from the residual plot as the cubic term is not captured by
@@
 Visually, it sometimes results in the so-called ``butterfly'' shape as
-shown in the plot C of Figure \ref{fig:residual-plot-common-departures},
-or the ``left-triangle'' and ``right-triangle'' shape where the smallest
+shown in the Figure \ref{fig:residual-plot-common-departures}C, or the
+``left-triangle'' and ``right-triangle'' shape where the smallest
 variance occurs at one side of the horizontal axis.
@@
 not readily reveal the marginal distribution. A favourable graphical
-summary for this task is the quantile-quantile plot. Hence,
-non-normality will not be the focus of this paper as we mainly discuss
-residual plots. For a consistent comparison, the residual plot of this
-departure is still presented in plot D of Figure
-\ref{fig:residual-plot-common-departures}. Besides, it is important to
-note that not all regression models assume normality for the error term,
-but a certain amount do including the classical normal linear regression
-model. In the case that the normality assumption is violated, it is
-expected to observe data points do not center around the horizontal axis
-and there is an uneven distribution of the number points at both below
-and above the horizontal axis. For example, given a skewed error
+summary for this task is the quantile-quantile plot. As we mainly
+discuss residual plots, non-normality will not be the focus of this
+paper. For a consistent comparison, the residual plot of this departure
+is still presented in Figure \ref{fig:residual-plot-common-departures}D.
+When the number residuals below and above the horizontal axis are uneven
+across the local regions along the \(x\)-axis, we expect that the
+normality assumption is violated. For example, given a skewed error
 distribution, there will be fewer data points and more outliers on one
-side of the horizontal axis as shown in plot D of Figure
-\ref{fig:residual-plot-common-departures}.
+side of the horizontal axis as shown in Figure
+\ref{fig:residual-plot-common-departures}D.
 
@@
 \caption{\label{tab:example-residual-plot-table}Statistical significance testing for departures from good residuals for plots in Figure \ref{fig:residual-plot-common-departures}.
-Shown are the $p$-values calculated for the conventional RESET, the Breusch-Pagan and the Shapiro–Wilk tests.
+Shown are the $p$-values calculated for the RESET, the Breusch-Pagan and the Shapiro–Wilk tests.
 The good residual plot (A) is judged a good residual plot, as expected, by all tests.
@@
 placed plots, where one plot is the data plot, and the remaining
-\(m - 1\) plots have the identical graphical procedure except the data
-has been replaced with data consistent with \(H_0\). Then, an observer
-who have not seen the data plot will be asked to point out the most
-different plot from the lineup. Under \(H_0\), it is expected that the
-data plot would have no distinguishable difference from the null plots,
-and the probability that the observer correctly picks the data plot is
-\(1/m\). If one rejects \(H_0\) as the observer correctly picks the data
-plot, then the Type I error of this test is \(1/m\).
+\(m - 1\) null plots have the identical graphical procedure except the
+data has been replaced with data consistent with \(H_0\). Then, an
+observer who have not seen the data plot will be asked to point out the
+most different plot from the lineup. Under \(H_0\), it is expected that
+the data plot would have no distinguishable difference from the null
+plots, and the probability that the observer correctly picks the data
+plot is \(1/m\). If one rejects \(H_0\) as the observer correctly picks
+the data plot, then the Type I error of this test is \(1/m\).
 
@@
 and they do not expect the visual test to perform equally well as the
-conventional test. In their third experiment, where there is not a
-conventional test, visual test outperforms the conventional test for a
-large margin. This is encouraging, as it promotes the use of visual
-inference in situations where there are no existing statistical testing
-procedures. Visual inference have also been integrated into diagnostic
-of hierarchical linear models by \citet{loy2013diagnostic},
-\citet{loy2014hlmdiag} and \citet{loy2015you}. They use lineup protocols
-to judge the assumption of linearity, normality and constant error
-variance for both the level-1 and level-2 residuals. (expand?)
+conventional test. (contradict to our claim? expand?) In their third
+experiment, where there is not a conventional test, visual test
+outperforms the conventional test for a large margin. This is
+encouraging, as it promotes the use of visual inference in situations
+where there are no existing statistical testing procedures. Visual
+inference have also been integrated into diagnostic of hierarchical
+linear models by \citet{loy2013diagnostic}, \citet{loy2014hlmdiag} and
+\citet{loy2015you}. They use lineup protocols to judge the assumption of
+linearity, normality and constant error variance for both the level-1
+and level-2 residuals. (expand?)
 
@@
 
-Data used in the \(m - 1\) null plots needs to be simulated. In the
-context of regression diagnostics, sampling data from \(H_0\) is
+Data used in the \(m - 1\) null plots needs to be simulated. In
+regression diagnostics, sampling data consistent with \(H_0\) is
 equivalent to sampling data from the assumed model. As
@@
 
-In hypothesis testing, \(p\)-value is defined as the probability of
+In hypothesis testing, a \(p\)-value is defined as the probability of
 observing test results as least as extreme as the observed result given
 \(H_0\) is true. Within the context of visual inference, by involving
-\(k\) independent observers, \(p\)-value can be interpreted as the
+\(k\) independent observers, the \(p\)-value can be interpreted as the
 probability of having as many or more subjects detect the data plot than
@@
 
-Let \(X_j = \{0,1\}\) be a binomial random variable denoting whether
+Let \(X_j = \{0,1\}\) be a Bernoulli random variable denoting whether
 subject \(j\) correctly detecting the data plot, and
@@
 
-The power of a model misspecification test specifies the chance of a
-misspecified regression model being detected given a certain alternative
-hypothesis. It is an important indicator when one concerns if the
-regression model is misspecified. Though, people might usually be more
-interested in knowing how much the residuals depart from the model
-assumptions and if the departure is of practical significance.
-
-As discussed in \citet{majumder_validation_2013}, individual's skill may
-affect the number of observers who identify the data plot from the
-lineup. Thus, the power of a visual test could depend on the
-subject-specific abilities. Previously, it is addressed by modelling the
+The power of a model misspecification test is the probability that the
+null hypothesis is rejected given the regression model is misspecified.
+It is an important indicator when one is concerned about whether model
+assumptions have been violated. Although in practice, one might be more
+interested in knowing how much the residuals deviate from the model
+assumptions, and whether this deviation is of practical significance.
+
+As discussed in \citet{majumder_validation_2013}, the power of a visual
+test may depend on the ability of the particular subject, as the skill
+of the individual may affect the number of observers who identify the
+data plot from the lineup. Previously, it is addressed by modelling the
 probability of a subject \(j\) correctly picking the data plot from a
-lineup \(l\) using a mixed-effect logistic regression with the subject
-being treated as a random effect \citep{majumder_validation_2013}.
-However, in the multiple selections scenario, having this probability is
-insufficient to determine the power of a visual test as it doesn't
+lineup \(l\) using a mixed-effect logistic regression, with subjects
+treated as random effects \citep{majumder_validation_2013}. However, in
+the multiple selections scenario, having this probability is not
+sufficient to determine the power of a visual test because it does not
 provide information about the number of selections made by the subject
-for p-value calculation.
-
-Instead, we estimate the probability of a lineup being rejected directly
-by assuming the individual skill has negligible impact on the variation
-of the power of a lineup. The assumption is made to simplify the model
-structure, thus mitigating the need for expensive large-scale
-experiments to estimate the complex covariance matrix. The model is a
-logistic regression with the natural logarithm of the effect as the only
-regressor formulated as
+for the calculation of the p-value.
+
+Instead, we directly estimate the probability of a lineup being rejected
+by assuming that individual skill has negligible effect on the variation
+of the power. The assumption is to simplify the model structure, thereby
+obviate costly large-scale experiments to estimate complex covariance
+matrices. The model is a logistic regression with the natural logarithm
+of the effect as the only regressor formulated as
 
 \begin{equation} \label{eq:logistic-regression-1-1}
-Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(E)),
+Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log(E)),
 \end{equation}
@@
 
-\begin{equation} \label{eq:effect-size-ex1}
-E = \frac{1}{2\sigma^2}\boldsymbol{X}_b'\boldsymbol{R}_a'(diag(\boldsymbol{R}_a))^{-1}\boldsymbol{R}_a\boldsymbol{X}_b,
-\end{equation}
+\small
+
+\begin{gather}  \label{eq:effect-size}
+E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V})|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V})^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V})^{-1}\boldsymbol{\mu}_z\right),\\
+\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{X}_a(\boldsymbol{X}_a^T\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a^T,
+\end{gather}
+
+\normalsize
 
 \noindent where \(diag(.)\) is the diagonal matrix constructed from the
-diagonal elements of \(\boldsymbol{R}_a\).
+diagonal elements of a matrix,
+\(\boldsymbol{X}_a = (\boldsymbol{1}, \boldsymbol{X})\) is the matrix of
+regressors including the intercept used in the regression equation,
+\(\sigma^2\) is the assumed variance of the error term when \(H_0\) is
+true, \(\boldsymbol{V}\) is the actual variance of the error term, and
+\(\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z\)
+is the expected values of the residuals with \(\boldsymbol{Z}\) be any
+variables leave out by the model and \(\boldsymbol{\beta}_z\) be the
+corresponding coefficients.
 
@@
 the same logistic regression model is fit on different subsets of the
-collated data grouped by levels of factors. This includes
-{[}expansion{]}.
+collated data grouped by levels of factors. These include the
+distribution of the regressor and the type of the simulated model.
 
@@
 
+In the following analysis, lineups with \(X_{raw} \sim U(-1,1)\) will be
+the focus of next two sections, because visual patterns are more likely
+to be revealed with uniform distribution. Besides, six more evaluations
+are collected for the uniform distribution in experiment III, which will
+produce more reliable and stable results. Analysis of other
+distributions can be found in Section
+\ref{effect-of-the-distribution-of-the-regressor}.
+
 \hypertarget{comparison-of-the-power-of-different-tests}{%
@@
 \(X_{raw} \sim U(-1,1)\), against the natural logarithm of the effect
-\(log_e(E)\), with a 5\% significance level. Uniform distribution is the
-focus of this section because visual patterns are more likely to be
-revealed. Besides, six more evaluations are collected for the uniform
-distribution in experiment III, which will produce more reliable and
-stable power curves. Analysis of other distributions can be found in
-Section \ref{effect-of-the-distribution-of-the-regressor}. At the bottom
-of the figure \ref{fig:polypower}, there are a sequence of example
-residual plots with increasing levels of \(log_e(E)\). Readers can
-evaluate them from left to right and determine at which level the
-departure from a good residual plot becomes detectable.
+\(log_e(E)\), with a 5\% significance level. At the bottom of the figure
+\ref{fig:polypower}, there are a sequence of example residual plots with
+increasing levels of \(log_e(E)\). Readers can evaluate them from left
+to right and determine at which level the departure from a good residual
+plot becomes detectable.
 
@@
 to 2, while power of other tests respond inactively to the change of
-effect, showing that REST test is way more sensitive to the type of
+effect, showing that RESET test is way more sensitive to the type of
 model defects that being considered. Meanwhile, no noticeable visual
@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/p-value-comparison-1}
+
+}
+
+\caption{Rejection rate of visual test compared to conventional test on lineups produced by heteroskedasticity and non-linearity model.
+Based on the rejection decision made by the visual test and the conventional test, the mosaic plot shows the proportion of lineups in each category.}\label{fig:p-value-comparison}
+\end{figure}
+
+(remove lineup id from caption)
+
 The power comparison illustrates that appropriate conventional tests
@@
 
-Depending on the natural logarithm of the effect, lineups are classified
-into three categories: easy (\textgreater{} 3.5), moderate (1.5 - 3.5)
-and difficult (\textless{} 1.5). Figure \ref{fig:p-value-comparison}
-provides a mosaic plot showing the rejection rate of visual tests and
-conventional tests at different difficulty levels.
-
-\hypertarget{easy-lineup}{%
-\subsubsection{Easy lineup}\label{easy-lineup}}
-
-For easy lineups produced by the non-linearity model, 65\% are rejected
-by both the conventional and visual tests, while 9\% are accepted by
-both tests. Thus, 74\% of the time two tests will agree with each other.
-There are 16\% of lineups rejected by conventional tests only.
-
-Interestingly, in spite of the greater power of conventional tests, 10\%
+Figure \ref{fig:p-value-comparison} provides a mosaic plot showing the
+rejection rate of visual tests and conventional tests for both the
+non-linearity model and the heteroskedasticity model.
+
+For lineups produced by the non-linearity model, 25\% are rejected by
+both the conventional and visual tests, while 42\% are accepted by both
+tests. Thus, 69\% of the time two tests will agree with each other.
+There are 26\% of lineups rejected by conventional tests only.
+
+Interestingly, in spite of the greater power of conventional tests, 7\%
 of the lineups are rejected by visual tests only. Recall that the RESET
@@
 in the RESET test, all these cases will be rejected. We also can observe
-from the example plot displayed in Figure \ref{fig:easy-poly-example}
-that panel three is distinctly different from others, exhibiting a ``M''
+from the example plot displayed in Figure \ref{fig:poly-example} that
+panel three is distinctly different from others, exhibiting a ``M''
 shape. This clearly suggests users of the RESET test and any other
@@
 
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/easy-poly-example-1}
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/poly-example-1}
 
@@
 It is rejected by the visual test but not by the RESET test.
-The data plot at panel three exhibits strong non-linear pattern.}\label{fig:easy-poly-example}
-\end{figure}
-
-For easy lineups produced by the heteroskedasticity model, 62\% are
-rejected by both the conventional test and the visual test, and 7\% are
-accepted by both, result in 69\% agreement rate. Lineups only rejected
-by conventional test are 30\%. 1\% of lineups are accepted by the
-conventional test but rejected by the visual test.
-
-Among those 1\%, which are two lineups, one shows detectable
-heteroskedasticity pattern as displayed in Figure
-\ref{fig:easy-heter-example1}. Another contain data plot with an
-noticeable outlier, but the ``Right-triangle'' shape is not very strong
-as shown in Figure \ref{fig:easy-heter-example2}. The \(p\)-values of
-Breusch-Pagan test for these two lineups are \(0.055\) and \(0.135\).
-Compared to the visual test, its \(p\)-values are \(0.026\) and
-\(0.012\). The Breusch-Pagan test should reject the one with obvious
-heteroskedasticity signal, but it yields a \(p\)-value slightly above
-5\%.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/easy-heter-example1-1}
+The data plot at panel three exhibits strong non-linear pattern.}\label{fig:poly-example}
+\end{figure}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-11-1} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-12-1} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-13-1} \end{center}
+
+For lineups produced by the heteroskedasticity model, 55\% are rejected
+by both the conventional test and the visual test, and 24\% are accepted
+by both, result in 79\% agreement rate. Percentage of lineups only
+rejected by conventional test are 20\%. Besides, only 1 lineup, which is
+less than 1\% of all the lineups, are rejected by the visual test only.
+This lineup shows relatively strong heteroskedasticity pattern as
+displayed in Figure \ref{fig:heter-example}. The visual test rejects it
+with a \(p\text{-value} = 0.026\), but the Breusch-Pagan test produces a
+\(p\)-value slightly above 5\%. However, as we can see, it is an
+exceptional case.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/heter-example-1}
 
@@
 It is rejected by the visual test but not by the  Breusch-Pagan test.
-The data plot at panel 17 contains a "Butterfly" shape, indicating the presence of heteroskedasticity.}\label{fig:easy-heter-example1}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/easy-heter-example2-1}
-
-}
-
-\caption{Lineup heter-536 with $log_e(E) = 4.13$ produced by the heteroskedasticity model used in experiment II.
-It is rejected by the visual test but not by the  Breusch-Pagan test.
-The data plot at panel 18 exhibits a vague heteroskedasticity pattern, but it contains a distinct outlier which makes it attractive to subjects.}\label{fig:easy-heter-example2}
-\end{figure}
-
-\hypertarget{moderate-lineup}{%
-\subsubsection{Moderate lineup}\label{moderate-lineup}}
-
-With the increase of the difficulty level, the sensitivity of
-conventional tests enlarge the difference in rejection rates. For
-moderate lineup produced by the non-linearity model, 8\% and 28\% are
-rejected and accepted by both tests respectively. We can observe a
-decline in agreement rate to 36\%. A large proportion of lineups are
-accepted by the visual test and rejected by the conventional test, which
-is around 63\%. The remaining 1\% are rejected by conventional tests
-only, which suffer the same issue caused by the RESET test as mentioned
-earlier.
-
-For those produced by the heteroskedasticy model, the corresponding
-numbers are 23\%, 31\%, 54\%, 45\% and 1\%. There is only one lineup
-rejected by the visual test but accepted by the conventional test. It
-does not show any heteroskedasticity patterns, but contains a few
-outliers at the bottom of the data plot as illustrated in panel 14 of
-Figure \ref{fig:moderate-heter-example1}. Two out of five subjects
-correctly identify this lineup yielding a \(p\)-value equal to
-\(0.044\). We consider this instance is rejected by chance. If a
-different set of null plots are used, the data plot may not be
-identified.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/moderate-heter-example1-1}
-
-}
-
-\caption{Lineup heter-219 with $log_e(E) = 1.61$ produced by the heteroskedasticity model used in experiment II.
-It is rejected by the visual test but not by the  Breusch-Pagan test.
-Though the data plot at panel 14 shows no heteroskedasticity patterns, it contains a few outliers at the bottom of the panel.}\label{fig:moderate-heter-example1}
-\end{figure}
-
-\hypertarget{difficult-lineup}{%
-\subsubsection{Difficult lineup}\label{difficult-lineup}}
-
-Visual test is insensitive to lineups with small effect. When it comes
-to difficult level, none of the lineups produced by the non-linearity
-model are rejected by the visual test. There are 73\% of lineups
-accepted by both tests, and 26\% rejected by the conventional test. The
-agreement rate rise up to 73\%.
-
-For lineups produced by the heteroskedasticity model, 3\% are rejected
-by both tests, and 71\% are accepted by both tests. There are 24\% of
-lineups rejected by the conventional test only. Besides, only two
-lineups, which is 2\% of all the difficult lineups, are rejected by the
-visual test only.
-
-As shown in \ref{fig:difficult-heter-example1} and
-\ref{fig:difficult-heter-example2}, residual plots of both lineups are
-constructed with \(X_{raw} \sim lognormal(0, 0.6^2)/3\). Though the
-designed patterns are ``Left-triangle'' and ``Butterfly'' shape, both of
-them are invisible due to the small effect. The visual effect are
-dominated by the skewed distribution. Further, no distinguishable visual
-patterns, such as outliers, are detectable from the data plot of these
-two lineups, which are panel 5 and panel 15. The data plots in these two
-lineups are not the most different plot from our point of view either.
-However, three out of five and two out of five subjects correctly
-identify the data plot for these two lineups respectively. But the
-reasons for selecting the plot provided by the subjects are mixed and
-inconsistent. Hence, with the current data, we can not explain why
-subjects can correctly identify the data plot of these two lineups.
-Collecting more data in future study may help to answer this question.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/difficult-heter-example1-1}
-
-}
-
-\caption{Lineup heter-224 with $log_e(E) = 1.32$ produced by the heteroskedasticity model used in experiment II.
-It is rejected by the visual test but not by the  Breusch-Pagan test.
-Although two out of five subjects correctly identify the data plot at panel 5, it does not exhibit any visually distinguishable pattern.}\label{fig:difficult-heter-example1}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/difficult-heter-example2-1}
-
-}
-
-\caption{Lineup heter-415 with $log_e(E) = -0.31$ produced by the heteroskedasticity model used in experiment II.
-It is rejected by the visual test but not by the  Breusch-Pagan test.
-Panel 15 is the data plot but it does not exhibit any visually distinguishable pattern.}\label{fig:difficult-heter-example2}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/p-value-comparison-1}
-
-}
-
-\caption{Rejection rate of visual test compared to conventional test on lineups produced by heteroskedasticity and non-linearity model at different levels of difficulty.
-Based on the rejection decision made by the visual test and the conventional test, the mosaic plot shows the proportion of lineups in each category.}\label{fig:p-value-comparison}
+The data plot at panel 17 contains a "Butterfly" shape, indicating the presence of heteroskedasticity.}\label{fig:heter-example}
 \end{figure}
@@
 regressor}\label{effect-of-the-distribution-of-the-regressor}}
+
+(calc the diff in power and plot)
+
+(randomly subset uniform and check the change of the power)
 
@@
 \(X_{raw}\) used in the non-linearity model and the heteroskedasticity
-model.
+model. We focus on \(log(E) > 2\) as the visual patterns become visible.
+
+For the non-linearity model, we do not observe significant difference of
+power of visual test between distributions, except for the discrete
+uniform distribution. It could be due to the fact that humans have
+difficulties in recognising visual patterns from discreteness, which
+makes the shape disconnected and incomplete (ref?). In terms of
+conventional tests, the discrete uniform distribution is instead the one
+with the greatest power, followed by the lognormal distribution, uniform
+distribution and the normal distribution.
+
+For the heteroskedasticity model, the power of visual test on lineups
+with uniform distribution has the greatest power. This is as expected
+since other three distributions could reduce the chance of revealing the
+underlying visual pattern because of uneven data points. Although the
+power of visual test under the discrete uniform distribution is the
+lowest in the case of non-linearity model, it has a relatively great
+power this time. Considering for heteroskedasticity, the visual patterns
+are usually detected by connecting the maximum and minimum residuals
+separately at different fitted values. It will not be greatly affected
+by the discrete uniform distribution compared to the uniform
+distribution. Visual tests have low power on lineups with the normal
+distribution and the lognormal distribution. This suggests the use of
+transformation of fitted values in reading residual plots. In the case
+of conventional tests, the discrete uniform distribution consistently
+doing great, followed by the uniform distribution, the normal
+distribution and the lognormal distribution.
+
+In regards of the type of the simulated model, the power of visual tests
+under different distributions of regressor are always lower than the
+corresponding conventional tests, consistent with the result from the
+previous sections.
 
@@
 
-\caption{...}\label{fig:different-x-dist-poly-power}
+\caption{Power of conventional tests and visual tests on lineups produced by the non-linearity model under different distributions of regressor.}\label{fig:different-x-dist-poly-power}
 \end{figure}
@@
 
-\caption{...}\label{fig:different-x-dist-heter-power}
+\caption{Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model under different distributions of regressor.}\label{fig:different-x-dist-heter-power}
 \end{figure}

### 2023-01-20 paper_comparison/paper_comparison/paper_comparison.Rmd
re-arranged experimental design plots and removed axis tick labels

@@
 
-```{r first-example-lineup, fig.height = 5, fig.cap="Visual testing is conducted using a lineup, as in the example here.
+```{r first-example-lineup, fig.height = 7, fig.cap="Visual testing is conducted using a lineup, as in the example here.
 The residual plot computed from the observed data (plot $2^2 + 2$, exhibiting non-linearity) is embedded among 19 null plots, where the residuals are simulated from a standard error model.
@@
 ### Non-linearity
-
-```{r different-shape-of-herimite, fig.height = 2, fig.cap="Polynomial forms generated for the residual plots used in experiment I.
-The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$."}
-stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
-
-ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300)
-
-VI_MODEL$plot(ori_dat) + 
-  ggtitle(quote(He[2]:"U")) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") -> p1
-
-VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote(He[3]:"S")) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") -> p2
-
-VI_MODEL$plot(poly_model(shape = 3, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote(He[6]:"M")) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") -> p3
-
-VI_MODEL$plot(poly_model(shape = 4, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote(He[18]:"Triple-U")) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") -> p4
-
-patchwork::wrap_plots(p1, p2, p3, p4, ncol = 4)
-```
 
@@
 
-```{r different-dist, fig.height = 2, fig.cap="Variations in fitted values, that might affect perception of residual plots.
-Four different distribution of $X_{raw}$ are used in the experiment to provide various visual patterns."}
+```{r different-shape-of-herimite, fig.height = 2, fig.cap="Polynomial forms generated for the residual plots used in experiment I.
+The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$."}
+stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
+
+ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300)
+
+pp1 <- VI_MODEL$plot(ori_dat) + 
+  ggtitle(quote(He[2]:"U")) + 
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+pp2 <- VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote(He[3]:"S")) + 
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+pp3 <- VI_MODEL$plot(poly_model(shape = 3, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote(He[6]:"M")) + 
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+pp4 <- VI_MODEL$plot(poly_model(shape = 4, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote(He[18]:"Triple-U")) + 
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+patchwork::wrap_plots(pp1, pp2, pp3, pp4, ncol = 4)
+```
+
+```{r different-dist, fig.height = 2, fig.cap="Variations in fitted values ($X_{raw}$), that might affect perception of residual plots. Four different distribution are used."}
 ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)
 
-VI_MODEL$plot(ori_dat) + 
+dp1 <- VI_MODEL$plot(ori_dat) + 
   ggtitle("A. Uniform") + #quote(U(-1,1))) + 
@@
   xlab("Fitted values") +
-  ylab("Residuals") -> p1
-
-VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+dp2 <- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
   ggtitle("B. Normal") + #quote(N(0,0.3^2))) + 
@@
   xlab("Fitted values") +
-  ylab("Residuals") -> p2
-
-VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+dp3 <- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
   ggtitle("C. Skewed") + #quote(lognormal(0,0.6^2)/3)) + 
@@
   xlab("Fitted values") +
-  ylab("Residuals") -> p3
-
-VI_MODEL$plot(poly_model(shape = 1, x = rand_uniform_d(k = 5, even = TRUE), sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+dp4 <- VI_MODEL$plot(poly_model(shape = 1, x = rand_uniform_d(k = 5, even = TRUE), sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
   ggtitle("D. Discrete") + #quote(U~"{1,5}")) + 
@@
   xlab("Fitted values") +
-  ylab("Residuals") -> p4
-
-patchwork::wrap_plots(p1, p2, p3, p4, ncol = 4)
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+patchwork::wrap_plots(dp1, dp2, dp3, dp4, ncol = 4)
 ```
@@
 
+<!--
 ```{r example-poly-lineup, fig.height=5, fig.cap="Lineup poly-24 in experiment I. Can you spot the most different plot? \\label{fig:example-poly-lineup}"}
@@
                      remove_axis = TRUE)
+```
+-->
+
+```{r different-n, fig.height = 2.67, fig.cap="Three different values of $n$ are used in experiment I, II and III to control the strength of the signal."}
+ori_dat <- poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(300)
+
+np1 <- VI_MODEL$plot(ori_dat) + 
+  ggtitle(quote(n~"="~300)) + 
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+np2 <- VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:50,], x, e))) +
+  ggtitle(quote(n~"="~50)) + 
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+np3 <- VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:100,], x, e))) +
+  ggtitle(quote(n~"="~100)) + 
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+patchwork::wrap_plots(np2, np3, np1, ncol = 3)
+```
+
+```{r different-sigma, fig.height = 2, fig.cap="Four different values of $\\sigma$ are used in the experiment I to control the strength of the signal."}
+ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)
+
+sp1 <- VI_MODEL$plot(ori_dat) + 
+  ggtitle(quote(sigma~"="~0.5)) + 
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+sp2 <- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1)$gen(300, computed = select(ori_dat, x))) +
+  ggtitle(quote(sigma~"="~1)) + 
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+sp3 <- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 2)$gen(300, computed = select(ori_dat, x))) +
+  ggtitle(quote(sigma~"="~2)) + 
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+sp4 <- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 4)$gen(300, computed = select(ori_dat, x))) +
+  ggtitle(quote(sigma~"="~4)) + 
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+patchwork::wrap_plots(sp1, sp2, sp3, sp4, ncol = 4)
 ```
@@
 
-VI_MODEL$plot(ori_dat) +
+hp1 <- VI_MODEL$plot(ori_dat) +
   ggtitle(quote("A. a = -1: Left-triangle")) +
@@
   xlab("Fitted values") +
-  ylab("Residuals") -> p1
-
-VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+hp2 <- VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
   ggtitle(quote("B. a = 0: Butterfly")) +
@@
   xlab("Fitted values") +
-  ylab("Residuals") -> p2
-
-VI_MODEL$plot(heter_model(a = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+hp3 <- VI_MODEL$plot(heter_model(a = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
   ggtitle(quote("C. a = 1: Right-triangle")) +
@@
   xlab("Fitted values") +
-  ylab("Residuals") -> p3
-
-patchwork::wrap_plots(p1, p2, p3, ncol = 3)
-```
-
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+patchwork::wrap_plots(hp1, hp2, hp3, ncol = 3)
+```
+
+```{r different-b, fig.height = 2, fig.cap="Five different values of $b$ are used in experiment II to control the strength of the signal."}
+stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
+
+ori_dat <- heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 0.25)$gen(300)
+
+bp1 <- VI_MODEL$plot(ori_dat) +
+  ggtitle(quote("b = 0.25")) +
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+bp2 <- VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 1)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote("b = 1")) +
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+bp3 <- VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 4)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote("b = 4")) +
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+bp4 <- VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 16)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote("b = 16")) +
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+bp5 <- VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 64)$gen(300, computed = select(ori_dat, x, e))) +
+  ggtitle(quote("b = 64")) +
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+patchwork::wrap_plots(bp1, bp2, bp3, bp4, bp5, ncol = 5)
+```
+
+<!--
 ```{r example-heter-lineup, fig.height = 5, fig.cap="Lineup heter-169 in experiment II. Can you spot the most different plot? \\label{fig:example-heter-lineup}"}
@@
 ```
+-->
 
@@
 It can be observed from Figure \ref{fig:different-n} that with fewer data points drawn in a residual plot, the visual pattern is more difficult to be detected.
-
-```{r different-n, fig.height = 2.67, fig.cap="Three different values of $n$ are used in experiment I, II and III to control the strength of the signal."}
-ori_dat <- poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(300)
-
-VI_MODEL$plot(ori_dat) + 
-  ggtitle(quote(n~"="~300)) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") -> p1
-
-VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:50,], x, e))) +
-  ggtitle(quote(n~"="~50)) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") -> p2
-
-VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:100,], x, e))) +
-  ggtitle(quote(n~"="~100)) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") -> p3
-
-patchwork::wrap_plots(p2, p3, p1, ncol = 3)
-```
-
-```{r different-sigma, fig.height = 2, fig.cap="Four different values of $\\sigma$ are used in the experiment I to control the strength of the signal."}
-ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)
-
-VI_MODEL$plot(ori_dat) + 
-  ggtitle(quote(sigma~"="~0.5)) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") -> p1
-
-VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1)$gen(300, computed = select(ori_dat, x))) +
-  ggtitle(quote(sigma~"="~1)) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") -> p2
-
-VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 2)$gen(300, computed = select(ori_dat, x))) +
-  ggtitle(quote(sigma~"="~2)) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") -> p3
-
-VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 4)$gen(300, computed = select(ori_dat, x))) +
-  ggtitle(quote(sigma~"="~4)) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") -> p4
-
-patchwork::wrap_plots(p1, p2, p3, p4, ncol = 4)
-```
-
-```{r different-b, fig.height = 2, fig.cap="Five different values of $b$ are used in experiment II to control the strength of the signal."}
-stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
-
-ori_dat <- heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 0.25)$gen(300)
-
-VI_MODEL$plot(ori_dat) +
-  ggtitle(quote("b = 0.25")) +
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") -> p1
-
-VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 1)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote("b = 1")) +
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") -> p2
-
-VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 4)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote("b = 4")) +
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") -> p3
-
-VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 16)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote("b = 16")) +
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") -> p4
-
-VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 64)$gen(300, computed = select(ori_dat, x, e))) +
-  ggtitle(quote("b = 64")) +
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") -> p5
-
-patchwork::wrap_plots(p1, p2, p3, p4, p5, ncol = 5)
-```
 

### 2023-01-20 paper_comparison/paper_comparison/paper_comparison.tex
re-arranged experimental design plots and removed axis tick labels

@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-herimite-1}
-
-}
-
-\caption{Polynomial forms generated for the residual plots used in experiment I.
-The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$.}\label{fig:different-shape-of-herimite}
-\end{figure}
-
 Experiment I is designed to study the ability of human subjects to
@@
 
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-herimite-1}
+
+}
+
+\caption{Polynomial forms generated for the residual plots used in experiment I.
+The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$.}\label{fig:different-shape-of-herimite}
+\end{figure}
+
+\begin{figure}
+
 {\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-dist-1}
@@
 
-\caption{Variations in fitted values, that might affect perception of residual plots.
-Four different distribution of $X_{raw}$ are used in the experiment to provide various visual patterns.}\label{fig:different-dist}
+\caption{Variations in fitted values ($X_{raw}$), that might affect perception of residual plots. Four different distribution are used.}\label{fig:different-dist}
 \end{figure}
@@
 
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-poly-lineup-1}
-
-}
-
-\caption{Lineup poly-24 in experiment I. Can you spot the most different plot? \label{fig:example-poly-lineup}}\label{fig:example-poly-lineup}
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-n-1}
+
+}
+
+\caption{Three different values of $n$ are used in experiment I, II and III to control the strength of the signal.}\label{fig:different-n}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-sigma-1}
+
+}
+
+\caption{Four different values of $\sigma$ are used in the experiment I to control the strength of the signal.}\label{fig:different-sigma}
 \end{figure}
@@
 
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-heter-lineup-1}
-
-}
-
-\caption{Lineup heter-169 in experiment II. Can you spot the most different plot? \label{fig:example-heter-lineup}}\label{fig:example-heter-lineup}
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-b-1}
+
+}
+
+\caption{Five different values of $b$ are used in experiment II to control the strength of the signal.}\label{fig:different-b}
 \end{figure}
@@
 more difficult to be detected.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-n-1}
-
-}
-
-\caption{Three different values of $n$ are used in experiment I, II and III to control the strength of the signal.}\label{fig:different-n}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-sigma-1}
-
-}
-
-\caption{Four different values of $\sigma$ are used in the experiment I to control the strength of the signal.}\label{fig:different-sigma}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-b-1}
-
-}
-
-\caption{Five different values of $b$ are used in experiment II to control the strength of the signal.}\label{fig:different-b}
-\end{figure}
 

### 2023-01-20 paper_comparison/paper_comparison/paper_comparison.Rmd
corrected captions for fig 9-11

@@
 
-## Comparison of the power of different tests
+## Power comparison of different tests
 
@@
 
-```{r polypower, fig.height = 3, fig.width = 5, fig.cap="Comparison of power between different tests for non-linear patterns.
+```{r polypower, fig.height = 3, fig.width = 5, fig.cap="Comparison of power between different tests for non-linear patterns (uniform fitted values only).
 Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups.
@@
 
-```{r heterpower, fig.height = 3, fig.width = 5, fig.cap="Comparison of power between different tests for heteroskedasticity patterns.
+```{r heterpower, fig.height = 3, fig.width = 5, fig.cap="Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only).
 Main plot shows the power curves, with dots indicating human evaluations of lineups.
@@
 
-```{r p-value-comparison, fig.width=10, fig.height=5, fig.cap="Rejection rate of visual test compared to conventional test on lineups produced by heteroskedasticity and non-linearity model.
+```{r p-value-comparison, fig.width=10, fig.height=5, fig.cap="Rejection rate of visual test compared to conventional test on lineups produced by heteroskedasticity and non-linearity model (uniform fitted values only).
 Based on the rejection decision made by the visual test and the conventional test, the mosaic plot shows the proportion of lineups in each category."}

### 2023-01-20 paper_comparison/paper_comparison/paper_comparison.tex
corrected captions for fig 9-11

@@
 
-\hypertarget{comparison-of-the-power-of-different-tests}{%
-\subsection{Comparison of the power of different
-tests}\label{comparison-of-the-power-of-different-tests}}
+\hypertarget{power-comparison-of-different-tests}{%
+\subsection{Power comparison of different
+tests}\label{power-comparison-of-different-tests}}
 
@@
 
-\caption{Comparison of power between different tests for non-linear patterns.
+\caption{Comparison of power between different tests for non-linear patterns (uniform fitted values only).
 Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups.
@@
 
-\caption{Comparison of power between different tests for heteroskedasticity patterns.
+\caption{Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only).
 Main plot shows the power curves, with dots indicating human evaluations of lineups.
@@
 
-\caption{Rejection rate of visual test compared to conventional test on lineups produced by heteroskedasticity and non-linearity model.
+\caption{Rejection rate of visual test compared to conventional test on lineups produced by heteroskedasticity and non-linearity model (uniform fitted values only).
 Based on the rejection decision made by the visual test and the conventional test, the mosaic plot shows the proportion of lineups in each category.}\label{fig:p-value-comparison}

### 2023-01-20 paper_comparison/paper_comparison/paper_comparison.Rmd
e, the constant, is lower case so axis labels adjusted accordingly

@@
   scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
-  xlab("Effect size (log E)") +
+  xlab("Effect size (log e)") +
   ylab("Power") +
@@
   scale_size_manual(values = c(1, 2, 3)) +
-  xlab("Effect size (log E)") +
+  xlab("Effect size (log e)") +
   ylab("Power") +
@@
               method.args = list(family = binomial())) +
-  xlab("Effect size (log E)") +
+  xlab("Effect size (log e)") +
   ylab("Power") +
@@
 
-## Effect of the distribution of the regressor
+## Effect of the fitted value distribution
 

### 2023-01-20 paper_comparison/paper_comparison/paper_comparison.tex
e, the constant, is lower case so axis labels adjusted accordingly

@@
 
-\hypertarget{effect-of-the-distribution-of-the-regressor}{%
-\subsection{Effect of the distribution of the
-regressor}\label{effect-of-the-distribution-of-the-regressor}}
+\hypertarget{effect-of-the-fitted-value-distribution}{%
+\subsection{Effect of the fitted value
+distribution}\label{effect-of-the-fitted-value-distribution}}
 

### 2023-01-20 paper_comparison/paper_comparison/paper_comparison.Rmd
revised figure placement, and captions for figs 14, 15

@@
 
-<!--
-```{r example-poly-lineup, fig.height=5, fig.cap="Lineup poly-24 in experiment I. Can you spot the most different plot? \\label{fig:example-poly-lineup}"}
+```{r different-n, fig.height = 2.67, fig.cap="Three different values of $n$ are used in experiment I, II and III to control the strength of the signal."}
+ori_dat <- poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(300)
+
+np1 <- VI_MODEL$plot(ori_dat) + 
+  ggtitle(quote(n~"="~300)) + 
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+np2 <- VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:50,], x, e))) +
+  ggtitle(quote(n~"="~50)) + 
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+np3 <- VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:100,], x, e))) +
+  ggtitle(quote(n~"="~100)) + 
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+patchwork::wrap_plots(np2, np3, np1, ncol = 3)
+```
+
+```{r different-sigma, fig.height = 2, fig.cap="Four different values of $\\sigma$ are used in the experiment I to control the strength of the signal."}
+ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)
+
+sp1 <- VI_MODEL$plot(ori_dat) + 
+  ggtitle(quote(sigma~"="~0.5)) + 
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+sp2 <- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1)$gen(300, computed = select(ori_dat, x))) +
+  ggtitle(quote(sigma~"="~1)) + 
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+sp3 <- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 2)$gen(300, computed = select(ori_dat, x))) +
+  ggtitle(quote(sigma~"="~2)) + 
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+sp4 <- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 4)$gen(300, computed = select(ori_dat, x))) +
+  ggtitle(quote(sigma~"="~4)) + 
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+patchwork::wrap_plots(sp1, sp2, sp3, sp4, ncol = 4)
+```
+
+```{r example-poly-lineup, fig.height=7, fig.cap="Lineup poly-24 in experiment I. Can you spot the most different plot? \\label{fig:example-poly-lineup}"}
 VI_MODEL$plot_lineup(vi_lineup[["poly_24"]]$data, 
@@
                      remove_axis = TRUE)
-```
--->
-
-```{r different-n, fig.height = 2.67, fig.cap="Three different values of $n$ are used in experiment I, II and III to control the strength of the signal."}
-ori_dat <- poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(300)
-
-np1 <- VI_MODEL$plot(ori_dat) + 
-  ggtitle(quote(n~"="~300)) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-np2 <- VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:50,], x, e))) +
-  ggtitle(quote(n~"="~50)) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-np3 <- VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:100,], x, e))) +
-  ggtitle(quote(n~"="~100)) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-patchwork::wrap_plots(np2, np3, np1, ncol = 3)
-```
-
-```{r different-sigma, fig.height = 2, fig.cap="Four different values of $\\sigma$ are used in the experiment I to control the strength of the signal."}
-ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)
-
-sp1 <- VI_MODEL$plot(ori_dat) + 
-  ggtitle(quote(sigma~"="~0.5)) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-sp2 <- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1)$gen(300, computed = select(ori_dat, x))) +
-  ggtitle(quote(sigma~"="~1)) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-sp3 <- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 2)$gen(300, computed = select(ori_dat, x))) +
-  ggtitle(quote(sigma~"="~2)) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-sp4 <- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 4)$gen(300, computed = select(ori_dat, x))) +
-  ggtitle(quote(sigma~"="~4)) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-patchwork::wrap_plots(sp1, sp2, sp3, sp4, ncol = 4)
 ```
@@
 
-<!--
-```{r example-heter-lineup, fig.height = 5, fig.cap="Lineup heter-169 in experiment II. Can you spot the most different plot? \\label{fig:example-heter-lineup}"}
+```{r example-heter-lineup, fig.height = 7, fig.cap="Lineup heter-169 in experiment II. Can you spot the most different plot? \\label{fig:example-heter-lineup}"}
 
@@
 ```
--->
 
@@
 
-```{r poly-example, fig.height = 5, fig.cap = "Lineup poly-270 with $log_e(E) = 3.89$ produced by the non-linearity model used in experiment I and III.
-It is rejected by the visual test but not by the RESET test.
-The data plot at panel three exhibits strong non-linear pattern."}
+```{r poly-example, fig.height = 7, fig.cap = "Example of a non-linearity that is rejected by the visual test but not by the RESET test.
+The data plot at panel three exhibits strong non-linear pattern.
+It has effect size $= 3.89$.
+Interestingly, all the lineups in this group are higher order polynomials (M, Triple-U), and it is apparent that the default RESET test under-performs.
+The RESET test using the higher order polynomial parameter could be used, but this would change it's performance for lower order polynomials.
+The only way for it to be sensitive to these type of departure is to know the order prior to testing!"}
+# plot poly-270 
 VI_MODEL$plot_lineup(vi_lineup$poly_270$data, 
@@
 
-```{r heter-example, fig.height = 5, fig.cap = 'Lineup heter-331 with $log_e(E) = 4.02$ produced by the heteroskedasticity model used in experiment II and III.
-It is rejected by the visual test but not by the  Breusch-Pagan test.
-The data plot at panel 17 contains a "Butterfly" shape, indicating the presence of heteroskedasticity.'}
+```{r heter-example, fig.height = 7, fig.cap = 'The single heteroskedasticity lineup that is rejected by the visual test but not by the BP test.
+The data plot at panel 17 contains a "Butterfly" shape.
+It has effect size$ = 4.02$.
+It is a mystery as to why this is not detected by the BP test.'}
+# plot heter-331
 VI_MODEL$plot_lineup(vi_lineup$heter_331$data, 

### 2023-01-20 paper_comparison/paper_comparison/paper_comparison.tex
revised figure placement, and captions for figs 14, 15

@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-poly-lineup-1}
+
+}
+
+\caption{Lineup poly-24 in experiment I. Can you spot the most different plot? \label{fig:example-poly-lineup}}\label{fig:example-poly-lineup}
+\end{figure}
+
 \hypertarget{heteroskedasticity}{%
@@
 \caption{Five different values of $b$ are used in experiment II to control the strength of the signal.}\label{fig:different-b}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-heter-lineup-1}
+
+}
+
+\caption{Lineup heter-169 in experiment II. Can you spot the most different plot? \label{fig:example-heter-lineup}}\label{fig:example-heter-lineup}
 \end{figure}
@@
 
-\caption{Lineup poly-270 with $log_e(E) = 3.89$ produced by the non-linearity model used in experiment I and III.
-It is rejected by the visual test but not by the RESET test.
-The data plot at panel three exhibits strong non-linear pattern.}\label{fig:poly-example}
+\caption{Example of a non-linearity that is rejected by the visual test but not by the RESET test.
+The data plot at panel three exhibits strong non-linear pattern.
+It has effect size $= 3.89$.
+Interestingly, all the lineups in this group are higher order polynomials (M, Triple-U), and it is apparent that the default RESET test under-performs.
+The RESET test using the higher order polynomial parameter could be used, but this would change it's performance for lower order polynomials.
+The only way for it to be sensitive to these type of departure is to know the order prior to testing!}\label{fig:poly-example}
 \end{figure}
@@
 
-\caption{Lineup heter-331 with $log_e(E) = 4.02$ produced by the heteroskedasticity model used in experiment II and III.
-It is rejected by the visual test but not by the  Breusch-Pagan test.
-The data plot at panel 17 contains a "Butterfly" shape, indicating the presence of heteroskedasticity.}\label{fig:heter-example}
+\caption{The single heteroskedasticity lineup that is rejected by the visual test but not by the BP test.
+The data plot at panel 17 contains a "Butterfly" shape.
+It has effect size$ = 4.02$.
+It is a mystery as to why this is not detected by the BP test.}\label{fig:heter-example}
 \end{figure}

### 2023-01-20 paper_comparison/paper_comparison/paper_comparison.Rmd
caption of figure 13

@@
 
-## Comparison of rejection rates based on $p$-values
-
-```{r p-value-comparison, fig.width=10, fig.height=5, fig.cap="Rejection rate of visual test compared to conventional test on lineups produced by heteroskedasticity and non-linearity model (uniform fitted values only).
-Based on the rejection decision made by the visual test and the conventional test, the mosaic plot shows the proportion of lineups in each category."}
+## Comparison of test decisions based on $p$-values
+
+```{r p-value-comparison, fig.width=8, fig.height=4, fig.cap="Rejection rate of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot.
+The visual test rejects less frequently than the conventional test.
+We would generally expect that the visual test would only reject when the conventional test does.
+Surprisingly, this is not what we see: the visual test sometimes rejects when the conventional test doesn't.
+"}
 

### 2023-01-20 paper_comparison/paper_comparison/paper_comparison.tex
caption of figure 13

@@
 
-\hypertarget{comparison-of-rejection-rates-based-on-p-values}{%
-\subsection{\texorpdfstring{Comparison of rejection rates based on
-\(p\)-values}{Comparison of rejection rates based on p-values}}\label{comparison-of-rejection-rates-based-on-p-values}}
+\hypertarget{comparison-of-test-decisions-based-on-p-values}{%
+\subsection{\texorpdfstring{Comparison of test decisions based on
+\(p\)-values}{Comparison of test decisions based on p-values}}\label{comparison-of-test-decisions-based-on-p-values}}
 
@@
 
-\caption{Rejection rate of visual test compared to conventional test on lineups produced by heteroskedasticity and non-linearity model (uniform fitted values only).
-Based on the rejection decision made by the visual test and the conventional test, the mosaic plot shows the proportion of lineups in each category.}\label{fig:p-value-comparison}
+\caption{Rejection rate of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot.
+The visual test rejects less frequently than the conventional test.
+We would generally expect that the visual test would only reject when the conventional test does.
+Surprisingly, this is not what we see: the visual test sometimes rejects when the conventional test doesn't.
+}\label{fig:p-value-comparison}
 \end{figure}

### 2023-01-20 paper_comparison/paper_comparison/paper_comparison.Rmd
working onrejection rate stats

@@
 
-```{r p-value-comparison, fig.width=8, fig.height=4, fig.cap="Rejection rate of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot.
+```{r p-value-comparison, fig.width=8, fig.height=4, fig.cap="Rejection rate ($p$-value $<0.05$) of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot.
 The visual test rejects less frequently than the conventional test.
@@
 
-(remove lineup id from caption)
-
 The power comparison illustrates that appropriate conventional tests reject $H_0$ more aggressively than visual tests.
@@
 
-For lineups produced by the non-linearity model, 25% are rejected by both the conventional and visual tests, while 42% are accepted by both tests.
+For lineups produced by the non-linearity model, conventional tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`%
+and visual tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the time.
+Of the tests rejected by the conventional test, `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`%, so approximately half as many as the conventional test.
+Surprisingly, the visual test rejects `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% filter(conv_reject == "Not") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of lineups where the conventional test does not reject.
+
+are rejected by both the conventional and visual tests, while 42% are accepted by both tests.
 Thus, 69% of the time two tests will agree with each other.

### 2023-01-20 paper_comparison/paper_comparison/paper_comparison.tex
working onrejection rate stats

@@
 
-\caption{Rejection rate of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot.
+\caption{Rejection rate ($p$-value $<0.05$) of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot.
 The visual test rejects less frequently than the conventional test.
@@
 
-(remove lineup id from caption)
-
 The power comparison illustrates that appropriate conventional tests
@@
 
-For lineups produced by the non-linearity model, 25\% are rejected by
-both the conventional and visual tests, while 42\% are accepted by both
-tests. Thus, 69\% of the time two tests will agree with each other.
-There are 26\% of lineups rejected by conventional tests only.
+For lineups produced by the non-linearity model, conventional tests
+reject 51\% and visual tests reject 32\% of the time. Of the tests
+rejected by the conventional test, 49\%, so approximately half as many
+as the conventional test. Surprisingly, the visual test rejects 14\% of
+lineups where the conventional test does not reject.
+
+are rejected by both the conventional and visual tests, while 42\% are
+accepted by both tests. Thus, 69\% of the time two tests will agree with
+each other. There are 26\% of lineups rejected by conventional tests
+only.
 

### 2023-01-20 paper_comparison/paper_comparison/paper_comparison.Rmd
changed % discussion for the non-linear example

@@
 and visual tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the time.
-Of the tests rejected by the conventional test, `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`%, so approximately half as many as the conventional test.
-Surprisingly, the visual test rejects `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% filter(conv_reject == "Not") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of lineups where the conventional test does not reject.
-
-are rejected by both the conventional and visual tests, while 42% are accepted by both tests.
-Thus, 69% of the time two tests will agree with each other.
-There are 26% of lineups rejected by conventional tests only.
-
-Interestingly, in spite of the greater power of conventional tests, 7% of the lineups are rejected by visual tests only.
-Recall that the RESET test used in our analysis includes powers of fitted values up to four.
-However, the "M" and the "Triple-U" shapes constructed from the Hermite polynomials contain power of the regressor up to six and 18 respectively.
-In fact, by including the power of fitted values up to six in the RESET test, all these cases will be rejected.
-We also can observe from the example plot displayed in Figure \ref{fig:poly-example} that panel three is distinctly different from others, exhibiting a "M" shape.
-This clearly suggests users of the RESET test and any other residual-based conventional tests that require the specification of variables of interest to check the residual plot before conducting the test, such that the correct choice of variables can be made.
+Of the tests rejected by the conventional test, `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by the visual test, that is, approximately half as many as the conventional test.
+Surprisingly, the visual test rejects `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% filter(conv_reject == "Not") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% (i.e.
+10 out of 70) of lineups where the conventional test does not reject.
+
+Figure \ref{fig:poly-example} shows one of the ten lineups where the visual test rejected the residual plot.
+The data plot, in position three is very obvious, and different from the nulls.
+The visual test was correct to reject.
+To understand why this is, one needs to return to the way the RESET test is applied.
+It requires a parameter indicating degree of polynomial to test for, and the recommendation is to generically use 4 (XXX REF).
+However, the "M" and the "Triple-U" shapes constructed from the Hermite polynomials use power up to 6 and 18 respectively.
+All 10 lineups rejected by the visual test but not the RESET were of types "M" or "Triple-U".
+If the RESET test had been applied using the power up to 6, all the decisions for these lineups would have been to reject.
+The recommendation of the polynomial power for the RESET should be revised, perhaps.
+This illustrates the sensitivity of conventional testing to the parameters, and it also points to a limitation that one needs to know the data structure in order to set the parameters for the test.
+
+<!--We also can observe from the example plot displayed in Figure \ref{fig:poly-example} that panel three is distinctly different from others, exhibiting a "M" shape.
+This clearly suggests users of the RESET test and any other residual-based conventional tests that require the specification of variables of interest to check the residual plot before conducting the test, such that the correct choice of variables can be made.-->
 

### 2023-01-20 paper_comparison/paper_comparison/paper_comparison.tex
changed % discussion for the non-linear example

@@
 reject 51\% and visual tests reject 32\% of the time. Of the tests
-rejected by the conventional test, 49\%, so approximately half as many
-as the conventional test. Surprisingly, the visual test rejects 14\% of
+rejected by the conventional test, 49\% are rejected by the visual test,
+that is, approximately half as many as the conventional test.
+Surprisingly, the visual test rejects 14\% (i.e.~10 out of 70) of
 lineups where the conventional test does not reject.
 
-are rejected by both the conventional and visual tests, while 42\% are
-accepted by both tests. Thus, 69\% of the time two tests will agree with
-each other. There are 26\% of lineups rejected by conventional tests
-only.
-
-Interestingly, in spite of the greater power of conventional tests, 7\%
-of the lineups are rejected by visual tests only. Recall that the RESET
-test used in our analysis includes powers of fitted values up to four.
-However, the ``M'' and the ``Triple-U'' shapes constructed from the
-Hermite polynomials contain power of the regressor up to six and 18
-respectively. In fact, by including the power of fitted values up to six
-in the RESET test, all these cases will be rejected. We also can observe
-from the example plot displayed in Figure \ref{fig:poly-example} that
-panel three is distinctly different from others, exhibiting a ``M''
-shape. This clearly suggests users of the RESET test and any other
-residual-based conventional tests that require the specification of
-variables of interest to check the residual plot before conducting the
-test, such that the correct choice of variables can be made.
+Figure \ref{fig:poly-example} shows one of the ten lineups where the
+visual test rejected the residual plot. The data plot, in position three
+is very obvious, and different from the nulls. The visual test was
+correct to reject. To understand why this is, one needs to return to the
+way the RESET test is applied. It requires a parameter indicating degree
+of polynomial to test for, and the recommendation is to generically use
+4 (XXX REF). However, the ``M'' and the ``Triple-U'' shapes constructed
+from the Hermite polynomials use power up to 6 and 18 respectively. All
+10 lineups rejected by the visual test but not the RESET were of types
+``M'' or ``Triple-U''. If the RESET test had been applied using the
+power up to 6, all the decisions for these lineups would have been to
+reject. The recommendation of the polynomial power for the RESET should
+be revised, perhaps. This illustrates the sensitivity of conventional
+testing to the parameters, and it also points to a limitation that one
+needs to know the data structure in order to set the parameters for the
+test.
 

### 2023-01-20 paper_comparison/paper_comparison/paper_comparison.Rmd
captions on Effect size

@@
   scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
-  xlab("Effect size (log e)") +
+  xlab(expression(log[e] (Effect_size))) +
   ylab("Power") +
@@
   scale_size_manual(values = c(1, 2, 3)) +
-  xlab("Effect size (log e)") +
+  xlab(expression(log[e] (Effect_size))) +
   ylab("Power") +
@@
 Figure \ref{fig:poly-example} shows one of the ten lineups where the visual test rejected the residual plot.
-The data plot, in position three is very obvious, and different from the nulls.
+The data plot, in position three, is very obvious, and different from the nulls.
 The visual test was correct to reject.
@@
               method.args = list(family = binomial())) +
-  xlab("Effect size (log e)") +
+  xlab(expression(log[e] (Effect_size))) +
   ylab("Power") +

### 2023-01-20 paper_comparison/paper_comparison/paper_comparison.Rmd
trying to get better line size for x dist power comparison

@@
 
-```{r different-x-dist-poly-power, fig.height = 3, fig.width = 5, fig.cap = "Power of conventional tests and visual tests on lineups produced by the non-linearity model under different distributions of regressor."}
+```{r different-x-dist-poly-power, fig.height = 3, fig.width = 7, fig.cap = "Power of conventional tests and visual tests on lineups produced by the non-linearity model under different fitted value distributions."}
 
@@
                          stringr::str_to_title(x_dist))) %>%
-  mutate(conv_or_not = "Visual"))
-
-dat_different_dist %>%
+  mutate(conv_or_not = "Visual")) %>%
+  mutate(x_dist = factor(x_dist, 
+      levels = c("Uniform", "Normal", 
+                 "Lognormal", "Discrete")))
+
 ggplot() +
-  geom_smooth(aes(log_effect_size, 
+  geom_smooth(data=filter(dat_different_dist, 
+                          x_dist == "Uniform"),
+              aes(log_effect_size, 
+                  reject), 
+                  col = "grey", 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, size=3) +
+  geom_smooth(data=dat_different_dist,
+              aes(log_effect_size, 
                   reject, 
@@
               method.args = list(family = binomial), 
-              se = FALSE) +
+              se = FALSE, size=1) +
   theme_light(base_size = 5) +
-  scale_color_manual(values = rcartocolor::carto_pal(4, "Safe")[c(1,3,2,4)]) +
-  scale_size_manual(values = c(1, 2, 3)) +
+  scale_color_manual(values = rcartocolor::carto_pal(4, "Safe")[c(4,1,3,2)]) +
+  scale_size_manual(values = c(1, 2, 3, 4)) +
   xlab(expression(log[e] (Effect_size))) +
   ylab("Power") +
-  labs(col = "Distribution of regressor", linetype = "Test") +
-  facet_wrap(~conv_or_not)
+  labs(col = "", linetype = "Test") +
+  facet_wrap(~conv_or_not) +
+  theme(legend.position="bottom")
 ```

### 2023-01-20 paper_comparison/paper_comparison/paper_comparison.Rmd
revise caption

--- 2023-01-20 paper_comparison/paper_comparison/paper_comparison.Rmd
@@
 
+Title 
+(jokes: finding the pattern in residual tests) 
+(why using visulization is important in regression diagnostics? an explanation using visual inference)
+(a picture is worth a thousand word)
+(a plot is worth a thousand test: )
+
 # Introduction
@@
 
-```{r}
-poly_conv_sim %>%
-  filter(x_dist == "uniform", shape %in% c(3, 4)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  select(shape, log_effect_size, (RESET3_p_value:RESET9_p_value)) %>%
-  pivot_longer(RESET3_p_value:RESET9_p_value) %>%
-  mutate(name = gsub("_p_value", " ", name)) %>%
-  mutate(name = gsub("RESET", "", name)) %>%
-  mutate(name = gsub("test", "", name)) %>%
-  mutate(reject = as.numeric(value < 0.05)) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject, col = name), 
-              se = FALSE, 
-              method = "glm",
-              method.args = list(family = binomial())) +
-  xlab(expression(log[e] (Effect_size))) +
-  ylab("Power") +
-  scale_color_manual(values = rev(rcartocolor::carto_pal(7, "RedOr"))) +
-  theme_light() +
-  labs(col = "RESET test order") +
-  facet_wrap(~shape)
-```
-
-```{r}
-heter_conv_sim %>%
-  filter(x_dist == "uniform") %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  mutate(reject = as.numeric(BP_p_value < 0.05)) %>%
-  mutate(a = c("left-triangle", "butterfly", "right-triangle")[a + 2]) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject, col = factor(a)), 
-              se = FALSE, 
-              method = "glm",
-              method.args = list(family = binomial())) +
-  xlab(expression(log[e] (Effect_size))) +
-  ylab("Power") +
-  scale_color_manual(values = rcartocolor::carto_pal(3, "Vivid")) +
-  theme_light() +
-  labs(col = "Heteroskedasticity shape") +
-  ggtitle("Conventional test")
-```
-
-```{r}
-vi_survey %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(x_dist == "uniform", type == "heteroskedasticity") %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  mutate(reject = as.numeric(p_value < 0.05)) %>%
-  mutate(a = c("left-triangle", "butterfly", "right-triangle")[a + 2]) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject, col = factor(a)), 
-              se = FALSE, 
-              method = "glm",
-              method.args = list(family = binomial())) +
-  xlab(expression(log[e] (Effect_size))) +
-  ylab("Power") +
-  scale_color_manual(values = rcartocolor::carto_pal(3, "Vivid")) +
-  theme_light() +
-  labs(col = "Heteroskedasticity shape") +
-  ggtitle("Visual test")
-```
-
 For lineups produced by the heteroskedasticity model, 55% are rejected by both the conventional test and the visual test, and 24% are accepted by both, result in 79% agreement rate.
@@
 
-## Effect of the fitted value distribution
+## Effect of other factors
+
+### Fitted value distribution
 
@@
 
-Although the distribution of the regressor is classically not restricted by the assumptions of linear regression, it will have an impact on the visual effect of the residual plot.
+Although the distribution of the regressor is classically not restricted by the assumptions of linear regression, it will have an impact on the distribution of the fitted values and the visual effect of the residual plot.
 In our experiments, four distributions of the regressor are used to accommodate this variation.
@@
   facet_wrap(~conv_or_not)
+```
+
+### Shape
+
+```{r eval = FALSE}
+poly_conv_sim %>%
+  filter(x_dist == "uniform", shape %in% c(3, 4)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  select(shape, log_effect_size, (RESET3_p_value:RESET9_p_value)) %>%
+  pivot_longer(RESET3_p_value:RESET9_p_value) %>%
+  mutate(name = gsub("_p_value", " ", name)) %>%
+  mutate(name = gsub("RESET", "", name)) %>%
+  mutate(name = gsub("test", "", name)) %>%
+  mutate(reject = as.numeric(value < 0.05)) %>%
+  ggplot() +
+  geom_smooth(aes(log_effect_size, reject, col = name), 
+              se = FALSE, 
+              method = "glm",
+              method.args = list(family = binomial())) +
+  xlab(expression(log[e] (Effect_size))) +
+  ylab("Power") +
+  scale_color_manual(values = rev(rcartocolor::carto_pal(7, "RedOr"))) +
+  theme_light() +
+  labs(col = "RESET test order") +
+  facet_wrap(~shape)
+```
+
+```{r poly-power-j, fig.cap = "Power of conventional tests and visual tests on lineups produced by the non-linearity model with four different shapes controlled by the parameter $j$."}
+
+poly_conv_sim %>%
+  filter(x_dist == "uniform") %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  mutate(reject = as.numeric(RESET4_p_value < 0.05)) %>%
+  mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>%
+  mutate(conv_type = "Conventional") %>%
+  bind_rows(vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(x_dist == "uniform", type == "polynomial") %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  mutate(reject = as.numeric(p_value < 0.05)) %>%
+  mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>%
+  mutate(conv_type = "Visual")) %>%
+  ggplot() +
+  geom_smooth(aes(log_effect_size, reject, col = factor(shape)), 
+              se = FALSE, 
+              method = "glm",
+              method.args = list(family = binomial())) +
+  xlab(expression(log[e] (Effect_size))) +
+  ylab("Power") +
+  scale_color_manual(values = rcartocolor::carto_pal(4, "Vivid")) +
+  theme_light() +
+  facet_wrap(~conv_type) +
+  labs(col = "Non-linearity shape")
+```
+
+```{r heter-power-a, fig.cap = "Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$."}
+
+heter_conv_sim %>%
+  filter(x_dist == "uniform") %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  mutate(reject = as.numeric(BP_p_value < 0.05)) %>%
+  mutate(a = c("left-triangle", "butterfly", "right-triangle")[a + 2]) %>%
+  mutate(conv_type = "Conventional") %>%
+  bind_rows(vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(x_dist == "uniform", type == "heteroskedasticity") %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  mutate(reject = as.numeric(p_value < 0.05)) %>%
+  mutate(a = c("left-triangle", "butterfly", "right-triangle")[a + 2]) %>%
+  mutate(conv_type = "Visual")) %>%
+  ggplot() +
+  geom_smooth(aes(log_effect_size, reject, col = factor(a)), 
+              se = FALSE, 
+              method = "glm",
+              method.args = list(family = binomial())) +
+  xlab(expression(log[e] (Effect_size))) +
+  ylab("Power") +
+  scale_color_manual(values = rcartocolor::carto_pal(3, "Vivid")) +
+  theme_light() +
+  facet_wrap(~conv_type) +
+  labs(col = "Heteroskedasticity shape")
 ```

### 2023-01-20 paper_comparison/paper_comparison/paper_comparison.tex
captions on Effect size

@@
 Figure \ref{fig:poly-example} shows one of the ten lineups where the
-visual test rejected the residual plot. The data plot, in position three
-is very obvious, and different from the nulls. The visual test was
-correct to reject. To understand why this is, one needs to return to the
-way the RESET test is applied. It requires a parameter indicating degree
-of polynomial to test for, and the recommendation is to generically use
-4 (XXX REF). However, the ``M'' and the ``Triple-U'' shapes constructed
-from the Hermite polynomials use power up to 6 and 18 respectively. All
-10 lineups rejected by the visual test but not the RESET were of types
-``M'' or ``Triple-U''. If the RESET test had been applied using the
-power up to 6, all the decisions for these lineups would have been to
-reject. The recommendation of the polynomial power for the RESET should
-be revised, perhaps. This illustrates the sensitivity of conventional
-testing to the parameters, and it also points to a limitation that one
-needs to know the data structure in order to set the parameters for the
-test.
+visual test rejected the residual plot. The data plot, in position
+three, is very obvious, and different from the nulls. The visual test
+was correct to reject. To understand why this is, one needs to return to
+the way the RESET test is applied. It requires a parameter indicating
+degree of polynomial to test for, and the recommendation is to
+generically use 4 (XXX REF). However, the ``M'' and the ``Triple-U''
+shapes constructed from the Hermite polynomials use power up to 6 and 18
+respectively. All 10 lineups rejected by the visual test but not the
+RESET were of types ``M'' or ``Triple-U''. If the RESET test had been
+applied using the power up to 6, all the decisions for these lineups
+would have been to reject. The recommendation of the polynomial power
+for the RESET should be revised, perhaps. This illustrates the
+sensitivity of conventional testing to the parameters, and it also
+points to a limitation that one needs to know the data structure in
+order to set the parameters for the test.
 
@@
 
-\caption{Power of conventional tests and visual tests on lineups produced by the non-linearity model under different distributions of regressor.}\label{fig:different-x-dist-poly-power}
+\caption{Power of conventional tests and visual tests on lineups produced by the non-linearity model under different fitted value distributions.}\label{fig:different-x-dist-poly-power}
 \end{figure}

### 2023-01-20 paper_comparison/paper_comparison/paper_comparison.Rmd
revise the p-value part

@@
 
+```{r eval = FALSE}
+vi_survey %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  left_join(vi_survey %>% 
+           count(unique_lineup_id) %>%
+           rename(num_subject = n)) %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
+  mutate(conv_reject = ifelse(conventional_p_value < 0.05, "Reject", "Not"),
+         reject = ifelse(p_value < 0.05, "Reject", "Not")) %>%
+  mutate(conv_reject = factor(conv_reject, levels = c("Reject", "Not")),
+         reject = factor(reject, levels = c("Reject", "Not"))) %>%
+  filter(type == "non-linearity") %>%
+  filter(reject == "Reject") %>%
+  filter(conv_reject == "Not") %>%
+  View()
+  count(conv_reject, reject)
+```
+
 The power comparison illustrates that appropriate conventional tests reject $H_0$ more aggressively than visual tests.
@@
 Of the tests rejected by the conventional test, `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by the visual test, that is, approximately half as many as the conventional test.
-Surprisingly, the visual test rejects `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% filter(conv_reject == "Not") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% (i.e.
-10 out of 70) of lineups where the conventional test does not reject.
-
-Figure \ref{fig:poly-example} shows one of the ten lineups where the visual test rejected the residual plot.
+There are no lineups that are rejected by the visual test but not by the conventional test.
+
+<!-- Figure \ref{fig:poly-example} shows one of the ten lineups where the visual test rejected the residual plot.
 The data plot, in position three, is very obvious, and different from the nulls.
@@
 This illustrates the sensitivity of conventional testing to the parameters, and it also points to a limitation that one needs to know the data structure in order to set the parameters for the test.
-
-<!--We also can observe from the example plot displayed in Figure \ref{fig:poly-example} that panel three is distinctly different from others, exhibiting a "M" shape.
-This clearly suggests users of the RESET test and any other residual-based conventional tests that require the specification of variables of interest to check the residual plot before conducting the test, such that the correct choice of variables can be made.-->
-
-```{r poly-example, fig.height = 7, fig.cap = "Example of a non-linearity that is rejected by the visual test but not by the RESET test.
+-->
+
+```{r poly-example, eval = FALSE, fig.height = 7, fig.cap = "Example of a non-linearity that is rejected by the visual test but not by the RESET test.
 The data plot at panel three exhibits strong non-linear pattern.
@@
 Percentage of lineups only rejected by conventional test are 20%.
-Besides, only 1 lineup, which is less than 1% of all the lineups, are rejected by the visual test only.
+
+Surprisingly, the visual test rejects `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Not") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% (i.e.
+1 out of 32) of lineups where the conventional test does not reject.
 This lineup shows relatively strong heteroskedasticity pattern as displayed in Figure \ref{fig:heter-example}.
 The visual test rejects it with a $p\text{-value} = 0.026$, but the Breusch-Pagan test produces a $p$-value slightly above 5%.
-However, as we can see, it is an exceptional case.
 

### 2023-01-20 paper_comparison/paper_comparison/paper_comparison.tex
revise the p-value part

@@
 \end{keywords}
+
+Title (jokes: finding the pattern in residual tests) (why using
+visulization is important in regression diagnostics? an explanation
+using visual inference) (a picture is worth a thousand word) (a plot is
+worth a thousand test: )
 
@@
 For lineups produced by the non-linearity model, conventional tests
-reject 51\% and visual tests reject 32\% of the time. Of the tests
-rejected by the conventional test, 49\% are rejected by the visual test,
-that is, approximately half as many as the conventional test.
-Surprisingly, the visual test rejects 14\% (i.e.~10 out of 70) of
-lineups where the conventional test does not reject.
-
-Figure \ref{fig:poly-example} shows one of the ten lineups where the
-visual test rejected the residual plot. The data plot, in position
-three, is very obvious, and different from the nulls. The visual test
-was correct to reject. To understand why this is, one needs to return to
-the way the RESET test is applied. It requires a parameter indicating
-degree of polynomial to test for, and the recommendation is to
-generically use 4 (XXX REF). However, the ``M'' and the ``Triple-U''
-shapes constructed from the Hermite polynomials use power up to 6 and 18
-respectively. All 10 lineups rejected by the visual test but not the
-RESET were of types ``M'' or ``Triple-U''. If the RESET test had been
-applied using the power up to 6, all the decisions for these lineups
-would have been to reject. The recommendation of the polynomial power
-for the RESET should be revised, perhaps. This illustrates the
-sensitivity of conventional testing to the parameters, and it also
-points to a limitation that one needs to know the data structure in
-order to set the parameters for the test.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/poly-example-1}
-
-}
-
-\caption{Example of a non-linearity that is rejected by the visual test but not by the RESET test.
-The data plot at panel three exhibits strong non-linear pattern.
-It has effect size $= 3.89$.
-Interestingly, all the lineups in this group are higher order polynomials (M, Triple-U), and it is apparent that the default RESET test under-performs.
-The RESET test using the higher order polynomial parameter could be used, but this would change it's performance for lower order polynomials.
-The only way for it to be sensitive to these type of departure is to know the order prior to testing!}\label{fig:poly-example}
-\end{figure}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-11-1} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-12-1} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-13-1} \end{center}
+reject 69\% and visual tests reject 32\% of the time. Of the tests
+rejected by the conventional test, 46\% are rejected by the visual test,
+that is, approximately half as many as the conventional test. There are
+no lineups that are rejected by the visual test but not by the
+conventional test.
 
@@
 by both, result in 79\% agreement rate. Percentage of lineups only
-rejected by conventional test are 20\%. Besides, only 1 lineup, which is
-less than 1\% of all the lineups, are rejected by the visual test only.
-This lineup shows relatively strong heteroskedasticity pattern as
-displayed in Figure \ref{fig:heter-example}. The visual test rejects it
-with a \(p\text{-value} = 0.026\), but the Breusch-Pagan test produces a
-\(p\)-value slightly above 5\%. However, as we can see, it is an
-exceptional case.
+rejected by conventional test are 20\%.
+
+Surprisingly, the visual test rejects 3\% (i.e.~1 out of 32) of lineups
+where the conventional test does not reject. This lineup shows
+relatively strong heteroskedasticity pattern as displayed in Figure
+\ref{fig:heter-example}. The visual test rejects it with a
+\(p\text{-value} = 0.026\), but the Breusch-Pagan test produces a
+\(p\)-value slightly above 5\%.
 
@@
 
-\hypertarget{effect-of-the-fitted-value-distribution}{%
-\subsection{Effect of the fitted value
-distribution}\label{effect-of-the-fitted-value-distribution}}
+\hypertarget{effect-of-other-factors}{%
+\subsection{Effect of other factors}\label{effect-of-other-factors}}
+
+\hypertarget{fitted-value-distribution}{%
+\subsubsection{Fitted value
+distribution}\label{fitted-value-distribution}}
 
@@
 by the assumptions of linear regression, it will have an impact on the
-visual effect of the residual plot. In our experiments, four
-distributions of the regressor are used to accommodate this variation.
+distribution of the fitted values and the visual effect of the residual
+plot. In our experiments, four distributions of the regressor are used
+to accommodate this variation.
 
@@
 
+\hypertarget{shape}{%
+\subsubsection{Shape}\label{shape}}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/poly-power-j-1}
+
+}
+
+\caption{Power of conventional tests and visual tests on lineups produced by the non-linearity model with four different shapes controlled by the parameter $j$.}\label{fig:poly-power-j}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/heter-power-a-1}
+
+}
+
+\caption{Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$.}\label{fig:heter-power-a}
+\end{figure}
+
 \hypertarget{appendix}{%

### 2023-01-22 paper_comparison/paper_comparison/paper_comparison.Rmd
automating calculation of numbers and tidying up p-value section

@@
 We would generally expect that the visual test would only reject when the conventional test does.
-Surprisingly, this is not what we see: the visual test sometimes rejects when the conventional test doesn't.
+Surprisingly, there is one lineup in the heteroskedasticity study where this is not the case.
 "}
@@
 
-The power comparison illustrates that appropriate conventional tests reject $H_0$ more aggressively than visual tests.
-In this section, we explore how often they agree with each other by investigate the rejection rates based on $p$-values.
+The power comparison illustrates that appropriate conventional test will reject $H_0$ more aggressively than visual tests.
+In this section, we explore how often they agree with each other by investigating the rejections for the two studies based on $p$-values for each lineup.
 
@@
 
-For lineups produced by the heteroskedasticity model, 55% are rejected by both the conventional test and the visual test, and 24% are accepted by both, result in 79% agreement rate.
-Percentage of lineups only rejected by conventional test are 20%.
-
-Surprisingly, the visual test rejects `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Not") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% (i.e.
-1 out of 32) of lineups where the conventional test does not reject.
-This lineup shows relatively strong heteroskedasticity pattern as displayed in Figure \ref{fig:heter-example}.
-The visual test rejects it with a $p\text{-value} = 0.026$, but the Breusch-Pagan test produces a $p$-value slightly above 5%.
+For lineups produced by the non-linearity model, conventional tests reject `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`%
+and visual tests reject `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the time.
+Of the tests rejected by the conventional test, `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by the visual test, that is, approximately half as many as the conventional test.
+
+<!--For lineups produced by the heteroskedasticity model, 55% are rejected by both the conventional test and the visual test, and 24% are accepted by both, result in 79% agreement rate.
+Percentage of lineups only rejected by conventional test are 20%.-->
+
+```{r}
+nr <- p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Not") %>% count(reject)
+```
+
+Surprisingly, the visual test rejects `r nr$n[1]` out of `r nr$n[1]+nr$n[2]` (`r round(nr$n[1]/(nr$n[1]+nr$n[2])*100, 0)`%) of lineups where the conventional test does not reject.
+Figure \ref{fig:heter-example} shows this lineup.
+The data plot in position seventeen displays a relatively strong heteroskedasticity pattern, and has a strong effect size ($log_e E=4.02$).
+This is reflected by the visual test $p\text{-value} = 0.026$, but the Breusch-Pagan test $p\text{-value} = 0.056$, slightly above the significance cutoff of $0.05$.
+This lineup was evaluated by 11 subjects, it has experimental factors $n=50$ (small sample size), $\sigma=1$, and a uniform distribution for the fitted values.
+It must be the small sample size that may have resulted in the lack of detection.
 
@@
 The data plot at panel 17 contains a "Butterfly" shape.
-It has effect size$ = 4.02$.
-It is a mystery as to why this is not detected by the BP test.'}
+It has effect size$ = 4.02$, somewhat surprising that it is not detected by the BP test.'}
 # plot heter-331

### 2023-01-22 paper_comparison/paper_comparison/paper_comparison.tex
automating calculation of numbers and tidying up p-value section

@@
 We would generally expect that the visual test would only reject when the conventional test does.
-Surprisingly, this is not what we see: the visual test sometimes rejects when the conventional test doesn't.
+Surprisingly, there is one lineup in the heteroskedasticity study where this is not the case.
 }\label{fig:p-value-comparison}
@@
 
-The power comparison illustrates that appropriate conventional tests
+The power comparison illustrates that appropriate conventional test will
 reject \(H_0\) more aggressively than visual tests. In this section, we
-explore how often they agree with each other by investigate the
-rejection rates based on \(p\)-values.
+explore how often they agree with each other by investigating the
+rejections for the two studies based on \(p\)-values for each lineup.
 
@@
 
-For lineups produced by the heteroskedasticity model, 55\% are rejected
-by both the conventional test and the visual test, and 24\% are accepted
-by both, result in 79\% agreement rate. Percentage of lineups only
-rejected by conventional test are 20\%.
-
-Surprisingly, the visual test rejects 3\% (i.e.~1 out of 32) of lineups
-where the conventional test does not reject. This lineup shows
-relatively strong heteroskedasticity pattern as displayed in Figure
-\ref{fig:heter-example}. The visual test rejects it with a
-\(p\text{-value} = 0.026\), but the Breusch-Pagan test produces a
-\(p\)-value slightly above 5\%.
+For lineups produced by the non-linearity model, conventional tests
+reject 76\% and visual tests reject 56\% of the time. Of the tests
+rejected by the conventional test, 73\% are rejected by the visual test,
+that is, approximately half as many as the conventional test.
+
+Surprisingly, the visual test rejects 1:2 out of c(NA, NA) (c(NA, NA)\%)
+of lineups where the conventional test does not reject. Figure
+\ref{fig:heter-example} shows this lineup. The data plot in position
+seventeen displays a relatively strong heteroskedasticity pattern, and
+has a strong effect size (\(log_e E=4.02\)). This is reflected by the
+visual test \(p\text{-value} = 0.026\), but the Breusch-Pagan test
+\(p\text{-value} = 0.056\), slightly above the significance cutoff of
+\(0.05\). This lineup was evaluated by 11 subjects, it has experimental
+factors \(n=50\) (small sample size), \(\sigma=1\), and a uniform
+distribution for the fitted values. It must be the small sample size
+that may have resulted in the lack of detection.
 
@@
 The data plot at panel 17 contains a "Butterfly" shape.
-It has effect size$ = 4.02$.
-It is a mystery as to why this is not detected by the BP test.}\label{fig:heter-example}
+It has effect size$ = 4.02$, somewhat surprising that it is not detected by the BP test.}\label{fig:heter-example}
 \end{figure}

### 2023-01-22 paper_comparison/paper_comparison/paper_comparison.Rmd
clean up code and replace reported value with inline r code if possible

@@
 
-# a bit stronger
-# plot 7 -> plot 2^3 - 1
-
-# vi_survey %>%
-#   filter(!attention_check) %>%
-#   filter(!null_lineup) %>%
-#   filter(type == "polynomial") %>%
-#   filter(x_dist == "uniform") %>%
-#   filter(n == 300) %>%
-#   filter(shape == 1) %>%
-#   group_by(unique_lineup_id) %>%
-#   summarise(effect_size = first(effect_size)) %>%
-#   arrange(effect_size)
-
 vi_lineup$poly_300$data %>%
@@
 
-```{r eval = FALSE}
-mix_effect_dat <- vi_survey %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(x_dist == "uniform") %>%
-  filter(type == "polynomial") %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  mutate(subject = factor(as.integer(factor(paste(exp, set))))) %>%
-  mutate(reject = p_value < 0.05) %>%
-  select(log_effect_size, reject, subject)
-
-# Singularity issue (random effect too small? not worth to fit a mix-effect model?)
-mix_mod <- lme4::glmer(reject ~ scale(log_effect_size) + (1 | subject), data = mix_effect_dat, family = binomial())
-
-pred_dat <- map_df(seq(-1, 6, 0.1),
-    function(x) {
-      result <- data.frame(log_effect_size = rep(x, length(unique(mix_effect_dat$subject))))
-      pred <- predict(mix_mod, 
-                      type = "response", 
-                      newdata = data.frame(log_effect_size = x, 
-                                           subject = unique(mix_effect_dat$subject)))
-      result$subject <- names(pred)
-      result$pred <- pred
-      result
-    })
-
-pred_dat %>%
-  ggplot() +
-  geom_line(aes(log_effect_size, pred, group = subject))
-```
-
-```{r eval = FALSE}
-MASS::glmmPQL(weighted_detect ~ scale(log_effect_size), random = ~ scale(log_effect_size) | subject,  family = quasibinomial(), data = vi_survey %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(x_dist == "uniform") %>%
-  filter(type == "polynomial") %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  mutate(subject = factor(as.integer(factor(paste(exp, set))))) %>%
-  select(log_effect_size, weighted_detect, subject)) -> mod
-
-mod_dat <- vi_survey %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(x_dist == "uniform") %>%
-  filter(type == "polynomial") %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  # mutate(log_effect_size = scale(log_effect_size)) %>%
-  mutate(subject = factor(as.integer(factor(paste(exp, set))))) %>%
-  select(log_effect_size, weighted_detect, subject)
-
-mod <- lme4::glmer(weighted_detect ~ scale(log_effect_size)  + (scale(log_effect_size) | subject), 
-                   family = quasibinomial(),
-                   data = mod_dat)
-
-mod_dat$log_effect_size
-
-map(seq(-1, 6, 0.1), ~predict(mod, newdata = data.frame(subject = 1:283, log_effect_size = .x), type = "response")) %>%
-  `names<-`()
-
-vi_survey %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(x_dist == "uniform") %>%
-  filter(type == "polynomial") %>%
-  mutate(log_effect_size = log10(effect_size)) %>%
-  mutate(subject = factor(as.integer(factor(paste(exp, set))))) %>%
-  mutate(pred_wd = predict(mod, type = "response")) %>%
-  group_by(unique_lineup_id) %>%
-  mutate(pred_ci = pred_wd * n()) %>%
-  mutate(n_eval = n()) %>%
-  ungroup() %>%
-  rowwise() %>%
-  mutate(pred_pval = (function(c_i, n_eval, alpha) {
-  target_dist <- exact_dist(n_eval, 20, dist = "dirichlet", alpha = alpha)
-  floor_c_i <- floor(c_i)
-  ceil_c_i <- ceiling(c_i)
-  return(unname(sum(target_dist[(ceil_c_i:n_eval) + 1]) + (ceil_c_i - c_i) * target_dist[floor_c_i + 1]))
-})(pred_ci, n_eval, alpha)) %>%
-  ungroup() %>%
-  mutate(pred_reject = pred_pval < 0.05) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, as.numeric(pred_reject)), method = "glm", method.args = list(family = binomial()), se = FALSE)
-
-```
-
-Effect $E$ is derived from the Kullback-Leibler divergence (see [appendix ref here]) formulated as
+Effect $E$ is derived from the Kullback-Leibler divergence (see Appendix for the details) formulated as
 
@@
 E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V})|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V})^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V})^{-1}\boldsymbol{\mu}_z\right),\\
-\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{X}_a(\boldsymbol{X}_a^T\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a^T,
+\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a,
+\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a^T\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a^T,
 \end{gather}
@@
 To study various factors contributing to the power of the visual test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors.
-These include the distribution of the regressor and the type of the simulated model.
-
-```{r residual-plot-cubic-heter, fig.cap='Residuals vs.
-fitted values plot for a classical normal linear regression model.
-The residuals are produced by fitting a two-regressor multiple linear regression model with data generated from a cubic linear model.
-From the residual plot, "butterfly shape" can be observed which generally would be interpretd as evidence of heteroskedasticity.
-Further, from the outline of the shape, nonlinear patterns exist.
-Both visual discoveries are evidence against $H_0$, though heteroskedasticity actually does not exist in the data generating process.
-\\label{fig:residual-plot-cubic-heter}', eval = FALSE}
-
-cm <- cubic_model(a = 0, b = 100, c = 1)
-
-cm$gen(1000, fit_model = TRUE) %>%
-cm$plot() +
-  theme_light()
-```
+These include the distribution of the regressor and the type of the simulated model (modify this according to the result).
 
@@
 The experiment III is designed for collecting human responses to null lineups such that the parameter $\alpha$ in Equation \ref{eq:pvalue-beta-binomial} can be estimated.
-Overall, we plan to collect 7974 evaluations on 1152 uniqued lineups performed by 443 subjects throughout three experiments.
+Overall, we plan to collect `r vi_survey  %>% filter(!attention_check) %>% nrow()` evaluations on `r vi_survey  %>% filter(!attention_check) %>% pull(unique_lineup_id) %>% unique() %>% length()` unique lineups performed by `r vi_survey  %>% filter(!attention_check) %>% count(exp, set) %>% nrow()` subjects throughout three experiments.
 
@@
 ### Non-linearity
-
-<!-- (handbook of mathathetical function, https://aapt.scitation.org/doi/10.1119/1.15378, cite the package) -->
-
-<!-- (hermite poly is created by ... in 18..,) -->
 
@@
   cat()
-  # gsub("\\\\\\$", "$", .) %>%
-  # cat()
 ```
@@
 The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$."}
-stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
-
-ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300)
-
+
+# Data for shape 1
+ori_dat <- poly_model(shape = 1, 
+                      x = {
+                        raw_x <- rand_uniform(-1, 1);
+                        closed_form(~stand_dist(raw_x))
+                        }, 
+                      sigma = 0.05)$gen(300)
+
+# Plot data for shape 1
 pp1 <- VI_MODEL$plot(ori_dat) + 
@@
 
-pp2 <- VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
+# Generate and plot data for shape 2. Reuse x and e.
+pp2 <- poly_model(shape = 2, 
+                  x = {
+                    raw_x <- rand_uniform(-1, 1); 
+                    closed_form(~stand_dist(raw_x))
+                    }, 
+                  sigma = 0.05)$
+  gen(300, computed = select(ori_dat, x, e)) %>%
+  VI_MODEL$plot() +
   ggtitle(quote(He[3]:"S")) + 
@@
 
-pp3 <- VI_MODEL$plot(poly_model(shape = 3, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
+# Generate and plot data for shape 3. Reuse x and e.
+pp3 <- poly_model(shape = 3, 
+                  x = {
+                    raw_x <- rand_uniform(-1, 1); 
+                    closed_form(~stand_dist(raw_x))
+                    }, 
+                  sigma = 0.05)$
+  gen(300, computed = select(ori_dat, x, e)) %>%
+  VI_MODEL$plot() +
   ggtitle(quote(He[6]:"M")) + 
@@
 
-pp4 <- VI_MODEL$plot(poly_model(shape = 4, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
+# Generate and plot data for shape 4. Reuse x and e.
+pp4 <- poly_model(shape = 4, 
+                  x = {
+                    raw_x <- rand_uniform(-1, 1); 
+                    closed_form(~stand_dist(raw_x))
+                    }, 
+                  sigma = 0.05)$
+  gen(300, computed = select(ori_dat, x, e)) %>%
+  VI_MODEL$plot() +
   ggtitle(quote(He[18]:"Triple-U")) + 
@@
 ```{r different-dist, fig.height = 2, fig.cap="Variations in fitted values ($X_{raw}$), that might affect perception of residual plots. Four different distribution are used."}
-ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)
-
+
+# Data for uniform distribution
+ori_dat <- poly_model(shape = 1, 
+                      x = {
+                        raw_x <- rand_uniform(-1, 1);
+                        closed_form(~stand_dist(raw_x))
+                        }, 
+                      sigma = 0.5)$gen(300)
+
+# Plot the data for uniform distribution
 dp1 <- VI_MODEL$plot(ori_dat) + 
-  ggtitle("A. Uniform") + #quote(U(-1,1))) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-dp2 <- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
-  ggtitle("B. Normal") + #quote(N(0,0.3^2))) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-dp3 <- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
-  ggtitle("C. Skewed") + #quote(lognormal(0,0.6^2)/3)) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-dp4 <- VI_MODEL$plot(poly_model(shape = 1, x = rand_uniform_d(k = 5, even = TRUE), sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
-  ggtitle("D. Discrete") + #quote(U~"{1,5}")) + 
+  ggtitle("A. Uniform") + 
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+# Generate and plot data for normal distribution
+dp2 <- poly_model(shape = 1, 
+                  x = {
+                    raw_x <- rand_normal(sigma = 0.3); 
+                    closed_form(~stand_dist(raw_x))
+                    }, 
+                  sigma = 0.5)$
+  gen(300, computed = select(ori_dat, e)) %>%
+  VI_MODEL$plot() +
+  ggtitle("B. Normal") +
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+# Generate and plot data for lognormal distribution
+dp3 <- poly_model(shape = 1, 
+                  x = {
+                    raw_x <- rand_lognormal(sigma = 0.6);
+                    closed_form(~stand_dist(raw_x/3 - 1))
+                    }, 
+                  sigma = 0.5)$
+  gen(300, computed = select(ori_dat, e)) %>%
+  VI_MODEL$plot() +
+  ggtitle("C. Skewed") +
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
+
+# Generate and plot data for discrete uniform distribution
+dp4 <- poly_model(shape = 1, 
+                  x = {
+                    raw_x <- rand_uniform_d(k = 5, even = TRUE);
+                    closed_form(~stand_dist(raw_x))
+                    }, 
+                  sigma = 0.5)$
+  gen(300, computed = select(ori_dat, e)) %>%
+  VI_MODEL$plot() +
+  ggtitle("D. Discrete") +
   theme_light() +
@@
 ```{r different-n, fig.height = 2.67, fig.cap="Three different values of $n$ are used in experiment I, II and III to control the strength of the signal."}
-ori_dat <- poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(300)
-
-np1 <- VI_MODEL$plot(ori_dat) + 
+
+# Generate data for 300 observations
+ori_dat <- poly_model(shape = 2, 
+                      x = {
+                        raw_x <- rand_uniform(-1, 1);
+                        closed_form(~stand_dist(raw_x))
+                        }, 
+                      sigma = 1.5)$gen(300)
+
+# Plot data for 300 observations
+np300 <- VI_MODEL$plot(ori_dat) + 
   ggtitle(quote(n~"="~300)) + 
@@
 
-np2 <- VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:50,], x, e))) +
+# Generate and plot data for 50 observations
+np50 <- poly_model(shape = 2, 
+                   x = {
+                     raw_x <- rand_uniform(-1, 1);
+                     closed_form(~stand_dist(raw_x))
+                     }, 
+                   sigma = 1.5)$
+  gen(50, computed = select(ori_dat[1:50, ], x, e)) %>%
+  VI_MODEL$plot() +
   ggtitle(quote(n~"="~50)) + 
@@
 
-np3 <- VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:100,], x, e))) +
+# Generate and plot data for 100 observations
+np100 <- poly_model(shape = 2, 
+                    x = {
+                      raw_x <- rand_uniform(-1, 1);
+                      closed_form(~stand_dist(raw_x))
+                      }, 
+                    sigma = 1.5)$
+  gen(50, computed = select(ori_dat[1:100,], x, e)) %>%
+  VI_MODEL$plot() +
   ggtitle(quote(n~"="~100)) + 
@@
 
-patchwork::wrap_plots(np2, np3, np1, ncol = 3)
+patchwork::wrap_plots(np50, np100, np300, ncol = 3)
 ```
@@
 ```{r different-sigma, fig.height = 2, fig.cap="Four different values of $\\sigma$ are used in the experiment I to control the strength of the signal."}
-ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)
-
+
+# Generate data for sigma 0.5
+ori_dat <- poly_model(shape = 1, 
+                      x = {
+                        raw_x <- rand_uniform(-1, 1);
+                        closed_form(~stand_dist(raw_x))
+                        }, 
+                      sigma = 0.5)$gen(300)
+
+# Plot data for sigma 0.5
 sp1 <- VI_MODEL$plot(ori_dat) + 
@@
 
-sp2 <- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1)$gen(300, computed = select(ori_dat, x))) +
+# Generate and plot data for sigma 1
+sp2 <- poly_model(shape = 1, 
+                  x = {
+                    raw_x <- rand_uniform(-1, 1); 
+                    closed_form(~stand_dist(raw_x))
+                    }, 
+                  sigma = 1)$
+  gen(300, computed = select(ori_dat, x)) %>%
+  VI_MODEL$plot() +
   ggtitle(quote(sigma~"="~1)) + 
@@
 
-sp3 <- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 2)$gen(300, computed = select(ori_dat, x))) +
+# Generate and plot data for sigma 2
+sp3 <- poly_model(shape = 1, 
+                  x = {
+                    raw_x <- rand_uniform(-1, 1); 
+                    closed_form(~stand_dist(raw_x))
+                    }, 
+                  sigma = 2)$
+  gen(300, computed = select(ori_dat, x)) %>%
+  VI_MODEL$plot() +
   ggtitle(quote(sigma~"="~2)) + 
@@
 
-sp4 <- VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 4)$gen(300, computed = select(ori_dat, x))) +
+# Generate and plot data for sigma 4
+sp4 <- poly_model(shape = 1, 
+                  x = {
+                    raw_x <- rand_uniform(-1, 1); 
+                    closed_form(~stand_dist(raw_x))
+                    }, 
+                  sigma = 4)$
+  gen(300, computed = select(ori_dat, x)) %>%
+  VI_MODEL$plot() +
   ggtitle(quote(sigma~"="~4)) + 
@@
 ```{r example-poly-lineup, fig.height=7, fig.cap="Lineup poly-24 in experiment I. Can you spot the most different plot? \\label{fig:example-poly-lineup}"}
-VI_MODEL$plot_lineup(vi_lineup[["poly_24"]]$data, 
+VI_MODEL$plot_lineup(vi_lineup$poly_24$data, 
                      remove_grid_line = TRUE, 
@@
 Three different shapes (a = -1, 0, 1) are used in the experiment to create left-triangle, butterfly and right-triangle shapes."}
-stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
-
-ori_dat <- heter_model(a = -1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300)
-
+
+# Generate data for a = -1
+ori_dat <- heter_model(a = -1, 
+                       x = {
+                         raw_x <- rand_uniform(-1, 1);
+                         closed_form(~stand_dist(raw_x))
+                         }, 
+                       b = 128)$gen(300)
+
+# Plot data for a = -1
 hp1 <- VI_MODEL$plot(ori_dat) +
@@
 
-hp2 <- VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
+# Plot data for a = 0
+hp2 <- heter_model(a = 0, 
+                   x = {
+                     raw_x <- rand_uniform(-1, 1); 
+                     closed_form(~stand_dist(raw_x))
+                     }, 
+                   b = 128)$
+  gen(300, computed = select(ori_dat, x, e)) %>%
+  VI_MODEL$plot() +
   ggtitle(quote("B. a = 0: Butterfly")) +
@@
 
-hp3 <- VI_MODEL$plot(heter_model(a = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
+# Plot data for a = 1
+hp3 <- heter_model(a = 1, 
+                   x = {
+                     raw_x <- rand_uniform(-1, 1); 
+                     closed_form(~stand_dist(raw_x))
+                     }, 
+                   b = 128)$
+  gen(300, computed = select(ori_dat, x, e)) %>%
+  VI_MODEL$plot() +
   ggtitle(quote("C. a = 1: Right-triangle")) +
@@
 ```{r different-b, fig.height = 2, fig.cap="Five different values of $b$ are used in experiment II to control the strength of the signal."}
-stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
-
-ori_dat <- heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 0.25)$gen(300)
-
+
+# Generate data for b = 0.25
+ori_dat <- heter_model(a = 0, 
+                       x = {
+                         raw_x <- rand_uniform(-1, 1);
+                         closed_form(~stand_dist(raw_x))
+                         }, 
+                       b = 0.25)$gen(300)
+
+# Plot data for b = 0.25
 bp1 <- VI_MODEL$plot(ori_dat) +
@@
 
-bp2 <- VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 1)$gen(300, computed = select(ori_dat, x, e))) +
+# Generate and plot data for b = 1
+bp2 <- heter_model(a = 0, 
+                   x = {
+                     raw_x <- rand_uniform(-1, 1); 
+                     closed_form(~stand_dist(raw_x))
+                     }, 
+                   b = 1)$
+  gen(300, computed = select(ori_dat, x, e)) %>%
+  VI_MODEL$plot() +
   ggtitle(quote("b = 1")) +
@@
 
-bp3 <- VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 4)$gen(300, computed = select(ori_dat, x, e))) +
+# Generate and plot data for b = 4
+bp3 <- heter_model(a = 0, 
+                   x = {
+                     raw_x <- rand_uniform(-1, 1); 
+                     closed_form(~stand_dist(raw_x))
+                     }, 
+                   b = 4)$
+  gen(300, computed = select(ori_dat, x, e)) %>%
+  VI_MODEL$plot() +
   ggtitle(quote("b = 4")) +
@@
 
-bp4 <- VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 16)$gen(300, computed = select(ori_dat, x, e))) +
+# Generate data and plot data for b = 16
+bp4 <- heter_model(a = 0, 
+                   x = {
+                     raw_x <- rand_uniform(-1, 1); 
+                     closed_form(~stand_dist(raw_x))
+                     }, 
+                   b = 16)$
+  gen(300, computed = select(ori_dat, x, e)) %>%
+  VI_MODEL$plot() +
   ggtitle(quote("b = 16")) +
@@
 
-bp5 <- VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 64)$gen(300, computed = select(ori_dat, x, e))) +
+# Generate data and plot data for b = 64
+bp5 <- heter_model(a = 0, 
+                   x = {
+                     raw_x <- rand_uniform(-1, 1); 
+                     closed_form(~stand_dist(raw_x))
+                     }, 
+                   b = 64)$
+  gen(300, computed = select(ori_dat, x, e)) %>%
+  VI_MODEL$plot() +
   ggtitle(quote("b = 64")) +
@@
 ```{r example-heter-lineup, fig.height = 7, fig.cap="Lineup heter-169 in experiment II. Can you spot the most different plot? \\label{fig:example-heter-lineup}"}
-
-# vi_survey %>%
-#   filter(!attention_check) %>%
-#   filter(!null_lineup) %>%
-#   filter(type == "heteroskedasticity") %>%
-#   filter(x_dist == "uniform") %>%
-#   filter(a == -1) %>%
-#   filter(n == 300) %>%
-#   group_by(unique_lineup_id) %>%
-#   summarise(effect_size = first(effect_size)) %>%
-#   arrange(effect_size)
-
-# vi_survey %>%
-#   filter(unique_lineup_id == "heter_471") %>%
-#   View()
 
@@
 
-Three replications are made for each of the parameter values shown in Table \ref{tab:model-factor-table} resulting in $(4 \times 4 \times 4 \times 3 + 4 \times 3 \times 5 \times 3) \times 3 = 1116$ different lineups.
+Three replications are made for each of the parameter values shown in Table \ref{tab:model-factor-table} resulting in $(4 \times 4 \times 4 \times 3 + 4 \times 3 \times 5 \times 3) \times 3 = `r (4 * 4 * 4 * 3 + 4 * 3 * 5 * 3) * 3`$ different lineups.
 In addition, each lineup is designed to be evaluated by five different subjects.
@@
 And to ensure the quality of the survey data, two lineups with obvious visual patterns are included as attention checks.
-Thus, $576 \times 5 / (20-2) = 160$ and $540 \times 5 / (20-2) = 150$ subjects are recruited to satisfy the design of the experiment I and experiment II respectively.
-
-As mentioned in Section \ref{calculating-p-values-for-the-visual-test}, $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using \svp{null} lineups.
+Thus, $576 \times 5 / (20-2) = `r 576 * 5 / 18`$ and $540 \times 5 / (20-2) = `r 540 * 5 / 18`$ subjects are recruited to satisfy the design of the experiment I and experiment II respectively.
+
+As mentioned in Section \ref{calculating-p-values-for-the-visual-test}, $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using null lineups.
 Hence, 36 lineups with all combinations of $n$ and $X_{raw}$ and three replications are included in experiment III.
-In these lineups, the data of the data plot is generated from a model \svp{with zero effect size}, while the data of the 19 null plots are generated using the same simulation method discussed in Section \ref{sampling-from-the-null-distribution}.
-\svp{This generation procedure differs from the canonical Rorschach lineup procedure, which requires that all 20 plots are generated from the null hypothesis.
-However, these lineups serve the same fundamental purpose: to assess the number of visually interesting plots generated under the null hypothesis.}
-<!-- Technically, this was an alternative to Rorschach lineup but with the null distribution further conditioning on the estimated coefficients and the residual sum of squares.  -->
-\svp{To account for the fact that our simulation method for these lineups is not the Rorschach procedure,} we use the method suggested in @vanderplas2021statistical for typical lineups containing a data plot to estimate $\alpha$.
+In these lineups, the data of the data plot is generated from a model with zero effect size, while the data of the 19 null plots are generated using the same simulation method discussed in Section \ref{sampling-from-the-null-distribution}.
+This generation procedure differs from the canonical Rorschach lineup procedure, which requires that all 20 plots are generated from the null hypothesis.
+However, these lineups serve the same fundamental purpose: to assess the number of visually interesting plots generated under the null hypothesis.
+
+To account for the fact that our simulation method for these lineups is not the Rorschach procedure, we use the method suggested in @vanderplas2021statistical for typical lineups containing a data plot to estimate $\alpha$.
 We have included a sensitivity analysis in the Appendix to examine the impact of the variance of the $\alpha$ estimate on our findings.
@@
 However, presenting only these lineups to subjects are considered to be bad practices as subjects will lose interest quickly. 
-Therefore, we plan to collect 6 more evaluations on the 279 lineups with $X_{raw} \sim U(-1,1)$, result in $(36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3) / (20-2) = 133$ subjects recruited for experiment III.
+Therefore, we plan to collect 6 more evaluations on the 279 lineups with $X_{raw} \sim U(-1,1)$, result in $(36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3 \times 6) / (20-2) = `r (36 * 20 + (4 * 4 * 3 + 3 * 5 * 3) * 3 * 6) / 18`$ subjects recruited for experiment III.
 
@@
 
-<!-- How many people? How many lineups? -->
-
-There are 2880, 2700 and 2394 lineups evaluation made by 160, 150 and 133 subjects recruited for experiment I, II and III respectively.
-In the total of 7974 lineup evaluations, 3744 use lineups produced by the non-linearity model, 3510 use lineups produced by the heteroskedasticity model, and 720 use null lineups.
-Besides, there are 886 attention checks not included in the following analysis.
+There are `r 160 * 18`, `r 150 * 18` and `133 * 18` lineups evaluation made by 160, 150 and 133 subjects recruited for experiment I, II and III respectively.
+In the total of `r nrow(filter(vi_survey, !attention_check))` lineup evaluations, `r nrow(filter(vi_survey, type == "polynomial", !attention_check))` use lineups produced by the non-linearity model, `r nrow(filter(vi_survey, type == "heteroskedasticity", !attention_check))` use lineups produced by the heteroskedasticity model, and `r nrow(filter(vi_survey, null_lineup, !attention_check))` use null lineups.
+Besides, there are `r nrow(filter(vi_survey, attention_check))` attention checks not included in the following analysis.
 The collated dataset is provided in `vi_survey` of the `visage` `R` package.
@@
 ## Power comparison of different tests
+
+```{r}
+# The glm model for RESET test
+conv_poly_glmmod <- poly_conv_sim %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  select(x_dist, log_effect_size, RESET4_p_value) %>%
+  filter(x_dist == "uniform") %>%
+  mutate(reject = as.numeric(RESET4_p_value < 0.05)) %>%
+glm(reject ~ log_effect_size, family = binomial(), data = .)
+```
 
@@
 
-For the non-linearity model, the power curve of RESET test climbs aggressively from 13% to around 70% as $log_e(E)$ increases from 0 to 2, while power of other tests respond inactively to the change of effect, showing that RESET test is way more sensitive to the type of model defects that being considered.
+For the non-linearity model, the power curve of RESET test climbs aggressively from
+`r round(unname(predict(conv_poly_glmmod, newdata = data.frame(log_effect_size = 0), type = "response")) * 100)`% to `r round(unname(predict(conv_poly_glmmod, newdata = data.frame(log_effect_size = 2), type = "response")) * 100)`% as $log_e(E)$ increases from 0 to 2, while power of other tests respond inactively to the change of effect, showing that RESET test is way more sensitive to the type of model defects that being considered.
 Meanwhile, no noticeable visual features can be spotted from the example residual plots.
@@
 ```{r}
+# Bootstrap the dataset
 boot_y <- function(dat, y, times = 100) {
@@
 ```{r cache = TRUE}
+# target_sigma is obtained by evaluating a vector of sigma and picking those 
+# with expected effect 0 ~ 5
 target_sigma <- c(4.51, 3.52, 2.74, 2.14, 1.66, 1.29, 1.01, 0.79, 0.61, 0.48, 0.37)
-es <- map_dbl(target_sigma, ~log(poly_model(shape = 2, x = rand_uniform(-1, 1), sigma = .x)$effect_size(n = 100)))
+
+# Calculate the effect size 
+es <- map_dbl(target_sigma, 
+              ~log(poly_model(shape = 2, 
+                              x = rand_uniform(-1, 1), 
+                              sigma = .x)$
+                     effect_size(n = 100)
+                   )
+              )
 ```
@@
 ```{r}
+# Analyse uniform polynomial data
 p <- vi_survey %>%
-  filter(x_dist == "uniform") %>%
+  filter(!null_lineup, 
+         !attention_check, 
+         type == "polynomial", 
+         x_dist == "uniform") %>%
   group_by(unique_lineup_id) %>%
   summarise(across(everything(), first)) %>%
-  filter(type == "polynomial") %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
   mutate(reject = as.numeric(p_value <= 0.05)) %>%
   mutate(log_effect_size = log(effect_size)) %>%
+  select(log_effect_size, reject) %>%
   ggplot() +
+
+  # Group the data by effect and reject to draw points
   geom_point(data = ~.x %>% 
              group_by(log_effect_size, reject) %>% 
-             summarise(num_response = n()),
-           aes(log_effect_size, reject, size = factor(num_response)), alpha = 0.3)  +
-  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
-                select(-(RESET5_p_value:RESET10_p_value)) %>%
-                select(-RESET3_p_value, -F_p_value) %>%
+             summarise(num_response = factor(n())),
+           aes(log_effect_size, reject, size = num_response), 
+           alpha = 0.3)  +
+
+  # Draw conventional test power curve
+  geom_smooth(data = poly_conv_sim %>% 
+                filter(x_dist == "uniform") %>%
+                mutate(log_effect_size = log(effect_size)) %>%
+                select(-RESET3_p_value, 
+                       -(RESET5_p_value:RESET10_p_value), 
+                       -F_p_value) %>%
                 rename(RESET_p_value = RESET4_p_value) %>%
-                filter(x_dist == "uniform") %>%
                 pivot_longer(RESET_p_value:SW_p_value) %>%
@@
               se = FALSE) +
+
+  # Draw bootstrap visual test power curves
   stat_smooth(geom = "line", 
               data = ~boot_y(.x, reject, times = 500), 
-              aes(log_effect_size, reject, group = boot_id, col = "Visual test"), 
+              aes(log_effect_size, 
+                  reject, 
+                  group = boot_id, 
+                  col = "Visual test"), 
               method = "glm", 
@@
               alpha = 0.03) +
-  geom_smooth(aes(log_effect_size, reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+
+  # Draw visual test power curve
+  geom_smooth(aes(log_effect_size, reject, col = "Visual test"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE) +
+
+  # Theme
   theme_light(base_size = 5) +
@@
 For the RESET test this is around log effect size 1.5, but for the visual test it is around 3.5."}
+
 set.seed(106)
-ex_dat <- poly_model(shape = 2, x = rand_uniform(), sigma = 1)$gen(n = 100)
+
+# Base data
+ex_dat <- poly_model(shape = 2, 
+                     x = rand_uniform(), 
+                     sigma = 1)$gen(n = 100)
 plist <- list()
@@
 
+# Obtain data at different effect. Reuse x and recompute e.
 for (i in target_sigma) {
   j <- j + 1
-  plist[[j]] <- 
-  VI_MODEL$plot(poly_model(shape = 2, x = rand_uniform())$
-                  gen(n = 100, computed = select(mutate(ex_dat, e = sqrt(i) * e), x, e)), 
-                theme = theme_light(base_size =  5), 
+plist[[j]] <- poly_model(shape = 2, 
+                         x = rand_uniform())$
+                  gen(n = 100, 
+                      computed =  ex_dat %>%
+                        mutate(e = sqrt(i) * e) %>%
+                        select(x, e)
+                      ) %>%
+  VI_MODEL$plot(theme = theme_light(base_size =  5), 
                 remove_axis = TRUE, 
-              remove_grid_line = TRUE, size = 0.05) +
+                remove_grid_line = TRUE, 
+                size = 0.05) +
     ggtitle(round(es[j], 1)) +
@@
 
-patchwork::wrap_plots(append(plist, list(p), 
-                             after = 0), 
-                      design = map(1:11, ~patchwork::area(21, .x, 22, .x)) %>%
+# Use area() to design layout
+map(1:11, ~patchwork::area(21, .x, 22, .x)) %>%
   reduce(c) %>%
-                        c(patchwork::area(3, 1, 20, 11), .))
+  c(patchwork::area(3, 1, 20, 11), .) %>%
+  patchwork::wrap_plots(append(plist, list(p), after = 0), 
+                        design = .)
 ```
@@
 ```{r cache = TRUE}
+# target_b is obtained by evaluating a vector of b and picking those 
+# with expected effect 0 ~ 5
 target_b <- c(0.14, 0.19, 0.27, 0.38, 0.57, 0.91, 1.53, 2.9, 6.4, 17.9, 75)
-es <- map_dbl(target_b, ~log(heter_model(a = 1, b = .x, x = rand_uniform(-1, 1))$effect_size(n = 100, type = "kl")))
+
+# Calculate the effect size 
+es <- map_dbl(target_b, 
+              ~log(heter_model(a = 1, 
+                               b = .x, 
+                               x = rand_uniform(-1, 1))$
+                     effect_size(n = 100, type = "kl")
+                   )
+              )
 ```
@@
 ```{r}
+# Analyse uniform heteroskedasticity data
 p <- vi_survey %>%
-  filter(x_dist == "uniform") %>%
+  filter(!null_lineup, 
+         !attention_check, 
+         x_dist == "uniform", 
+         type == "heteroskedasticity") %>%
   group_by(unique_lineup_id) %>%
   summarise(across(everything(), first)) %>%
-  filter(type == "heteroskedasticity") %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
   mutate(reject = as.numeric(p_value <= 0.05)) %>%
   mutate(log_effect_size = log(effect_size)) %>%
+  select(log_effect_size, reject) %>%
   ggplot() +
+
+  # Group the data by effect and reject to draw points
   geom_point(data = ~.x %>% 
                group_by(log_effect_size, reject) %>% 
-               summarise(num_response = n()),
-             aes(log_effect_size, reject, size = factor(num_response)), alpha = 0.3) +
-  geom_smooth(data = mutate(heter_conv_sim, log_effect_size = log(effect_size)) %>%
+               summarise(num_response = factor(n())),
+             aes(log_effect_size, reject, size = num_response), 
+             alpha = 0.3) +
+
+  # Draw conventional test power curve
+  geom_smooth(data = heter_conv_sim %>% 
                 filter(x_dist == "uniform") %>%
-                select(-(RESET5_p_value:RESET10_p_value)) %>%
-                select(-RESET3_p_value, -F_p_value) %>%
+                mutate(log_effect_size = log(effect_size)) %>%
+                select(-RESET3_p_value, 
+                       -(RESET5_p_value:RESET10_p_value), 
+                       -F_p_value) %>%
                 rename(RESET_p_value = RESET4_p_value) %>%
@@
               alpha = 0.03) +
+
+  # Draw visual test power curve
   geom_smooth(aes(log_effect_size, reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
+
+  # Theme
   theme_light(base_size = 5) +
@@
 For the BP test this is around log effect size 1.5, but for the visual test it is around 3."}
+
 set.seed(10091)
-ex_dat <- heter_model(a = 1, x = rand_uniform(-1, 1), b = 1)$gen(n = 100)
+
+# Base data
+ex_dat <- heter_model(a = 1, 
+                      x = rand_uniform(-1, 1), 
+                      b = 1)$gen(n = 100)
 plist <- list()
@@
 
+# Obtain data at different b
 for (i in target_b) {
   j <- j + 1
-  plist[[j]] <-
-  VI_MODEL$plot(heter_model(a = 1, b = i,x = rand_uniform(-1, 1))$
-                  gen(n = 100, computed = select(ex_dat, x, e)), theme = theme_light(base_size =  5),
+  plist[[j]] <- heter_model(a = 1, 
+                            b = i, 
+                            x = rand_uniform(-1, 1))$
+                  gen(n = 100, computed = select(ex_dat, x, e)) %>%
+  VI_MODEL$plot(theme = theme_light(base_size =  5),
                 remove_axis = TRUE,
@@
 
+# Use area() to design layout
+map(1:11, ~patchwork::area(21, .x, 22, .x)) %>%
+  reduce(c) %>%
+  c(patchwork::area(3, 1, 20, 11), .) %>%
 patchwork::wrap_plots(append(plist, list(p), after = 0), 
-                      design = map(1:11, ~patchwork::area(21, .x, 22, .x)) %>%
-                        reduce(c) %>%
-                        c(patchwork::area(3, 1, 20, 11), .))
+                      design = .)
 ```
@@
 
+# Dataset for p-value comparison
 p_value_cmp_dat <- vi_survey %>%
@@
   mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-  left_join(vi_survey %>% 
-           count(unique_lineup_id) %>%
-           rename(num_subject = n)) %>%
   group_by(unique_lineup_id) %>%
@@
   mutate(conv_reject = factor(conv_reject, levels = c("Reject", "Not")),
-         reject = factor(reject, levels = c("Reject", "Not")))
-
+         reject = factor(reject, levels = c("Reject", "Not"))) %>%
+  select(type, conv_reject, reject)
+
+# Mosaic plot
 p_value_cmp_dat %>%
   ggplot() +
-  ggmosaic::geom_mosaic(aes(x = ggmosaic::product(reject, conv_reject), fill = reject)) +
+  geom_mosaic(aes(x = ggmosaic::product(reject, conv_reject), 
+                  fill = reject)) +
   facet_grid(~type) +
@@
   coord_fixed()
-```
-
-```{r eval = FALSE}
-vi_survey %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
-  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-  left_join(vi_survey %>% 
-           count(unique_lineup_id) %>%
-           rename(num_subject = n)) %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
-  mutate(conv_reject = ifelse(conventional_p_value < 0.05, "Reject", "Not"),
-         reject = ifelse(p_value < 0.05, "Reject", "Not")) %>%
-  mutate(conv_reject = factor(conv_reject, levels = c("Reject", "Not")),
-         reject = factor(reject, levels = c("Reject", "Not"))) %>%
-  filter(type == "non-linearity") %>%
-  filter(reject == "Reject") %>%
-  filter(conv_reject == "Not") %>%
-  View()
-  count(conv_reject, reject)
 ```
@@
 Of the tests rejected by the conventional test, `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by the visual test, that is, approximately half as many as the conventional test.
-
-<!--For lineups produced by the heteroskedasticity model, 55% are rejected by both the conventional test and the visual test, and 24% are accepted by both, result in 79% agreement rate.
-Percentage of lineups only rejected by conventional test are 20%.-->
 

### 2023-01-22 paper_comparison/paper_comparison/paper_comparison.tex
clean up code and replace reported value with inline r code if possible

@@
 Effect \(E\) is derived from the Kullback-Leibler divergence (see
-{[}appendix ref here{]}) formulated as
+Appendix for the details) formulated as
 
@@
 collated data grouped by levels of factors. These include the
-distribution of the regressor and the type of the simulated model.
+distribution of the regressor and the type of the simulated model
+(modify this according to the result).
 
@@
 \ref{eq:pvalue-beta-binomial} can be estimated. Overall, we plan to
-collect 7974 evaluations on 1152 uniqued lineups performed by 443
+collect 7974 evaluations on 1152 unique lineups performed by 443
 subjects throughout three experiments.
@@
 \(\alpha\) used in Equation \ref{eq:pvalue-beta-binomial} needs to be
-estimated using {\textcolor{orange-red}{null}} lineups. Hence, 36
-lineups with all combinations of \(n\) and \(X_{raw}\) and three
-replications are included in experiment III. In these lineups, the data
-of the data plot is generated from a model
-{\textcolor{orange-red}{with zero effect size}}, while the data of the
-19 null plots are generated using the same simulation method discussed
-in Section \ref{sampling-from-the-null-distribution}.
-{\textcolor{orange-red}{This generation procedure differs from the canonical Rorschach lineup procedure, which requires that all 20 plots are generated from the null hypothesis.
-However, these lineups serve the same fundamental purpose: to assess the number of visually interesting plots generated under the null hypothesis.}}
-{\textcolor{orange-red}{To account for the fact that our simulation method for these lineups is not the Rorschach procedure,}}
-we use the method suggested in \citet{vanderplas2021statistical} for
-typical lineups containing a data plot to estimate \(\alpha\). We have
-included a sensitivity analysis in the Appendix to examine the impact of
-the variance of the \(\alpha\) estimate on our findings.
+estimated using null lineups. Hence, 36 lineups with all combinations of
+\(n\) and \(X_{raw}\) and three replications are included in experiment
+III. In these lineups, the data of the data plot is generated from a
+model with zero effect size, while the data of the 19 null plots are
+generated using the same simulation method discussed in Section
+\ref{sampling-from-the-null-distribution}. This generation procedure
+differs from the canonical Rorschach lineup procedure, which requires
+that all 20 plots are generated from the null hypothesis. However, these
+lineups serve the same fundamental purpose: to assess the number of
+visually interesting plots generated under the null hypothesis.
+
+To account for the fact that our simulation method for these lineups is
+not the Rorschach procedure, we use the method suggested in
+\citet{vanderplas2021statistical} for typical lineups containing a data
+plot to estimate \(\alpha\). We have included a sensitivity analysis in
+the Appendix to examine the impact of the variance of the \(\alpha\)
+estimate on our findings.
 
@@
 \(X_{raw} \sim U(-1,1)\), result in
-\((36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3) / (20-2) = 133\)
+\((36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3 \times 6) / (20-2) = 133\)
 subjects recruited for experiment III.
@@
 
-There are 2880, 2700 and 2394 lineups evaluation made by 160, 150 and
-133 subjects recruited for experiment I, II and III respectively. In the
-total of 7974 lineup evaluations, 3744 use lineups produced by the
-non-linearity model, 3510 use lineups produced by the heteroskedasticity
-model, and 720 use null lineups. Besides, there are 886 attention checks
-not included in the following analysis. The collated dataset is provided
-in \texttt{vi\_survey} of the \texttt{visage} \texttt{R} package.
+There are 2880, 2700 and \texttt{133\ *\ 18} lineups evaluation made by
+160, 150 and 133 subjects recruited for experiment I, II and III
+respectively. In the total of 7974 lineup evaluations, 3744 use lineups
+produced by the non-linearity model, 4230 use lineups produced by the
+heteroskedasticity model, and 720 use null lineups. Besides, there are
+886 attention checks not included in the following analysis. The
+collated dataset is provided in \texttt{vi\_survey} of the
+\texttt{visage} \texttt{R} package.
 
@@
 For the non-linearity model, the power curve of RESET test climbs
-aggressively from 13\% to around 70\% as \(log_e(E)\) increases from 0
-to 2, while power of other tests respond inactively to the change of
-effect, showing that RESET test is way more sensitive to the type of
-model defects that being considered. Meanwhile, no noticeable visual
-features can be spotted from the example residual plots.
+aggressively from 15\% to 69\% as \(log_e(E)\) increases from 0 to 2,
+while power of other tests respond inactively to the change of effect,
+showing that RESET test is way more sensitive to the type of model
+defects that being considered. Meanwhile, no noticeable visual features
+can be spotted from the example residual plots.
 
@@
 
-Surprisingly, the visual test rejects 1:2 out of c(NA, NA) (c(NA, NA)\%)
-of lineups where the conventional test does not reject. Figure
-\ref{fig:heter-example} shows this lineup. The data plot in position
-seventeen displays a relatively strong heteroskedasticity pattern, and
-has a strong effect size (\(log_e E=4.02\)). This is reflected by the
-visual test \(p\text{-value} = 0.026\), but the Breusch-Pagan test
+Surprisingly, the visual test rejects 1 out of 33 (3\%) of lineups where
+the conventional test does not reject. Figure \ref{fig:heter-example}
+shows this lineup. The data plot in position seventeen displays a
+relatively strong heteroskedasticity pattern, and has a strong effect
+size (\(log_e E=4.02\)). This is reflected by the visual test
+\(p\text{-value} = 0.026\), but the Breusch-Pagan test
 \(p\text{-value} = 0.056\), slightly above the significance cutoff of

### 2023-01-22 paper_comparison/paper_comparison/paper_comparison.Rmd
Attempt to combine seperate plots into one facetted plot for different j

@@
 
+# Facet label
+shape_labels <- c('He[2]:"U"', 'He[3]:"S"', 'He[6]:"M"', 'He[18]:"Triple-U"')
+
 # Data for shape 1
-ori_dat <- poly_model(shape = 1, 
+dat_shape_1 <- poly_model(shape = 1, 
                           x = {
@@
                             }, 
-                      sigma = 0.05)$gen(300)
-
-# Plot data for shape 1
-pp1 <- VI_MODEL$plot(ori_dat) + 
-  ggtitle(quote(He[2]:"U")) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-# Generate and plot data for shape 2. Reuse x and e.
-pp2 <- poly_model(shape = 2, 
+                          sigma = 0.05)$gen(300) %>%
+  mutate(shape = shape_labels[1])
+
+# Generate data for shape 2, 3 and 4. Reuse x and e.
+map_df(2:4, function(shape) {
+  poly_model(shape = shape, 
                   x = {
@@
                   sigma = 0.05)$
-  gen(300, computed = select(ori_dat, x, e)) %>%
+  gen(300, computed = select(dat_shape_1, x, e)) %>%
+  mutate(shape = shape_labels[shape])
+}) %>%
+
+  # Combined with data for shape 1
+  bind_rows(dat_shape_1) %>%
   VI_MODEL$plot() +
-  ggtitle(quote(He[3]:"S")) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-# Generate and plot data for shape 3. Reuse x and e.
-pp3 <- poly_model(shape = 3, 
-                  x = {
-                    raw_x <- rand_uniform(-1, 1); 
-                    closed_form(~stand_dist(raw_x))
-                    }, 
-                  sigma = 0.05)$
-  gen(300, computed = select(ori_dat, x, e)) %>%
-  VI_MODEL$plot() +
-  ggtitle(quote(He[6]:"M")) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-# Generate and plot data for shape 4. Reuse x and e.
-pp4 <- poly_model(shape = 4, 
-                  x = {
-                    raw_x <- rand_uniform(-1, 1); 
-                    closed_form(~stand_dist(raw_x))
-                    }, 
-                  sigma = 0.05)$
-  gen(300, computed = select(ori_dat, x, e)) %>%
-  VI_MODEL$plot() +
-  ggtitle(quote(He[18]:"Triple-U")) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-patchwork::wrap_plots(pp1, pp2, pp3, pp4, ncol = 4)
+  facet_wrap(~shape, scales = "free_x", labeller = label_parsed, ncol = 4) +
+  theme_light() +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  theme(axis.text = element_blank())
 ```

### 2023-01-22 paper_comparison/paper_comparison/paper_comparison.Rmd
modify the different distribution plot

@@
   VI_MODEL$plot() +
-  facet_wrap(~shape, scales = "free_x", labeller = label_parsed, ncol = 4) +
+  facet_wrap(~shape, scales = "free", labeller = label_parsed, ncol = 4) +
   theme_light() +
@@
 # Data for uniform distribution
-ori_dat <- poly_model(shape = 1, 
+dat_dist_1 <- poly_model(shape = 1, 
                       x = {
@@
                         }, 
-                      sigma = 0.5)$gen(300)
-
-# Plot the data for uniform distribution
-dp1 <- VI_MODEL$plot(ori_dat) + 
-  ggtitle("A. Uniform") + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-# Generate and plot data for normal distribution
-dp2 <- poly_model(shape = 1, 
+                      sigma = 0.5)$gen(300) %>%
+  mutate(x_dist = "Uniform")
+
+# Generate data for other distributions
+dat_dist_2 <- poly_model(shape = 1,
                          x = {
@@
                          sigma = 0.5)$
-  gen(300, computed = select(ori_dat, e)) %>%
-  VI_MODEL$plot() +
-  ggtitle("B. Normal") +
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-# Generate and plot data for lognormal distribution
-dp3 <- poly_model(shape = 1, 
+  gen(300, computed = select(dat_dist_1, e)) %>%
+  mutate(x_dist = "Normal")
+
+dat_dist_3 <- poly_model(shape = 1,
                          x = {
@@
                          sigma = 0.5)$
-  gen(300, computed = select(ori_dat, e)) %>%
+  gen(300, computed = select(dat_dist_1, e)) %>%
+  mutate(x_dist = "Skewed")
+
+dat_dist_4 <- poly_model(shape = 1,
+                         x = {
+                           raw_x <- rand_uniform_d(k = 5, even = TRUE); 
+                           closed_form(~stand_dist(raw_x))
+                           }, 
+                         sigma = 0.5)$
+  gen(300, computed = select(dat_dist_1, e)) %>%
+  mutate(x_dist = "Discrete")
+
+# Generate and plot data for discrete uniform distribution
+bind_rows(dat_dist_1, dat_dist_2, dat_dist_3, dat_dist_4) %>%
   VI_MODEL$plot() +
-  ggtitle("C. Skewed") +
   theme_light() +
@@
   ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-# Generate and plot data for discrete uniform distribution
-dp4 <- poly_model(shape = 1, 
-                  x = {
-                    raw_x <- rand_uniform_d(k = 5, even = TRUE);
-                    closed_form(~stand_dist(raw_x))
-                    }, 
-                  sigma = 0.5)$
-  gen(300, computed = select(ori_dat, e)) %>%
-  VI_MODEL$plot() +
-  ggtitle("D. Discrete") +
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-patchwork::wrap_plots(dp1, dp2, dp3, dp4, ncol = 4)
+  theme(axis.text = element_blank()) +
+  facet_wrap(~x_dist, ncol = 4, scales = "free")
 ```

### 2023-01-22 paper_comparison/paper_comparison/paper_comparison.Rmd
modify the different observations plot

@@
 # Generate data for 300 observations
-ori_dat <- poly_model(shape = 2, 
+dat_n_300 <- poly_model(shape = 2, 
                         x = {
@@
                         },
-                      sigma = 1.5)$gen(300)
-
-# Plot data for 300 observations
-np300 <- VI_MODEL$plot(ori_dat) + 
-  ggtitle(quote(n~"="~300)) + 
+                        sigma = 1.5)$gen(300) %>%
+  mutate(n = "N = 300")
+
+# Generate data for other settings
+map(c(50, 100), function(n) {
+  poly_model(shape = 2, 
+                   x = {
+                     raw_x <- rand_uniform(-1, 1);
+                     closed_form(~stand_dist(raw_x))
+                     }, 
+             sigma = 1.5)$
+    gen(n, computed = select(dat_n_300[1:n, ], x, e)) %>%
+    mutate(n = paste0("N = ", n))
+}) %>%
+  bind_rows(dat_n_300) %>%
+  VI_MODEL$plot() +
   theme_light() +
+  facet_wrap(~n, scales = "free", ncol = 3) +
   xlab("Fitted values") +
@@
   theme(axis.text = element_blank())
-
-# Generate and plot data for 50 observations
-np50 <- poly_model(shape = 2, 
-                   x = {
-                     raw_x <- rand_uniform(-1, 1);
-                     closed_form(~stand_dist(raw_x))
-                     }, 
-                   sigma = 1.5)$
-  gen(50, computed = select(ori_dat[1:50, ], x, e)) %>%
-  VI_MODEL$plot() +
-  ggtitle(quote(n~"="~50)) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-# Generate and plot data for 100 observations
-np100 <- poly_model(shape = 2, 
-                    x = {
-                      raw_x <- rand_uniform(-1, 1);
-                      closed_form(~stand_dist(raw_x))
-                      }, 
-                    sigma = 1.5)$
-  gen(50, computed = select(ori_dat[1:100,], x, e)) %>%
-  VI_MODEL$plot() +
-  ggtitle(quote(n~"="~100)) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-patchwork::wrap_plots(np50, np100, np300, ncol = 3)
 ```

### 2023-01-22 paper_comparison/paper_comparison/paper_comparison.Rmd
modify the different sigma plot

@@
   bind_rows(dat_shape_1) %>%
+  mutate(shape = factor(shape, levels = shape_labels)) %>%
   VI_MODEL$plot() +
@@
 bind_rows(dat_dist_1, dat_dist_2, dat_dist_3, dat_dist_4) %>%
+  mutate(x_dist = factor(x_dist, 
+                         levels = c("Uniform", 
+                                    "Normal", 
+                                    "Skewed", 
+                                    "Discrete"))) %>%
   VI_MODEL$plot() +
@@
                         sigma = 1.5)$gen(300) %>%
-  mutate(n = "N = 300")
+  mutate(n = "N: 300")
 
@@
     gen(n, computed = select(dat_n_300[1:n, ], x, e)) %>%
-    mutate(n = paste0("N = ", n))
+    mutate(n = paste0("N: ", n))
 }) %>%
+
+  # Combined with data for 300 observations
   bind_rows(dat_n_300) %>%
+  mutate(n = factor(n, levels = c("N: 50", "N: 100", "N: 300"))) %>%
   VI_MODEL$plot() +
@@
 # Generate data for sigma 0.5
-ori_dat <- poly_model(shape = 1, 
+dat_sigma_05 <- poly_model(shape = 1, 
                            x = {
@@
                              }, 
-                      sigma = 0.5)$gen(300)
-
-# Plot data for sigma 0.5
-sp1 <- VI_MODEL$plot(ori_dat) + 
-  ggtitle(quote(sigma~"="~0.5)) + 
+                           sigma = 0.5)$gen(300) %>%
+  mutate(sigma = "sigma: 0.5")
+
+# Generate data for other sigma
+map(c(1, 2, 4), function(sigma) {
+  poly_model(shape = 1, 
+                  x = {
+                    raw_x <- rand_uniform(-1, 1); 
+                    closed_form(~stand_dist(raw_x))
+                    }, 
+                  sigma = sigma)$
+  gen(300, computed = select(dat_sigma_05, x)) %>%
+  mutate(sigma = paste0("sigma: ", sigma))
+}) %>%
+
+  # Combined with data for sigma 0.5
+  bind_rows(dat_sigma_05) %>%
+  VI_MODEL$plot() + 
+  facet_wrap(~sigma, ncol = 4, scales = "free", labeller = label_parsed) +
   theme_light() +
@@
   theme(axis.text = element_blank())
-
-# Generate and plot data for sigma 1
-sp2 <- poly_model(shape = 1, 
-                  x = {
-                    raw_x <- rand_uniform(-1, 1); 
-                    closed_form(~stand_dist(raw_x))
-                    }, 
-                  sigma = 1)$
-  gen(300, computed = select(ori_dat, x)) %>%
-  VI_MODEL$plot() +
-  ggtitle(quote(sigma~"="~1)) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-# Generate and plot data for sigma 2
-sp3 <- poly_model(shape = 1, 
-                  x = {
-                    raw_x <- rand_uniform(-1, 1); 
-                    closed_form(~stand_dist(raw_x))
-                    }, 
-                  sigma = 2)$
-  gen(300, computed = select(ori_dat, x)) %>%
-  VI_MODEL$plot() +
-  ggtitle(quote(sigma~"="~2)) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-# Generate and plot data for sigma 4
-sp4 <- poly_model(shape = 1, 
-                  x = {
-                    raw_x <- rand_uniform(-1, 1); 
-                    closed_form(~stand_dist(raw_x))
-                    }, 
-                  sigma = 4)$
-  gen(300, computed = select(ori_dat, x)) %>%
-  VI_MODEL$plot() +
-  ggtitle(quote(sigma~"="~4)) + 
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-patchwork::wrap_plots(sp1, sp2, sp3, sp4, ncol = 4)
 ```

### 2023-01-22 paper_comparison/paper_comparison/paper_comparison.Rmd
modify plot for different heter shape

@@
 
+a_labels <- c("Left-triangle", "Butterfly", "Right-triangle")
+
 # Generate data for a = -1
-ori_dat <- heter_model(a = -1, 
+dat_a_n1 <- heter_model(a = -1, 
                         x = {
@@
                           },
-                       b = 128)$gen(300)
-
-# Plot data for a = -1
-hp1 <- VI_MODEL$plot(ori_dat) +
-  ggtitle(quote("A. a = -1: Left-triangle")) +
+                        b = 128)$gen(300) %>%
+  mutate(a = a_labels[1])
+
+# Generate data for other a
+map(c(0, 1), function(a) {
+  heter_model(a = a,
+              x = {
+                raw_x <- rand_uniform(-1, 1); 
+                closed_form(~stand_dist(raw_x))
+                }, 
+              b = 128)$
+    gen(300, computed = select(dat_a_n1, x, e)) %>%
+    mutate(a = a_labels[a + 2])
+}) %>%
+
+  # Combined with data for a = -1
+  bind_rows(dat_a_n1) %>%
+  mutate(a = factor(a, levels = a_labels)) %>%
+  VI_MODEL$plot() +
+  facet_wrap(~a, scales = "free", ncol = 3) +
   theme_light() +
@@
   theme(axis.text = element_blank())
-
-# Plot data for a = 0
-hp2 <- heter_model(a = 0, 
-                   x = {
-                     raw_x <- rand_uniform(-1, 1); 
-                     closed_form(~stand_dist(raw_x))
-                     }, 
-                   b = 128)$
-  gen(300, computed = select(ori_dat, x, e)) %>%
-  VI_MODEL$plot() +
-  ggtitle(quote("B. a = 0: Butterfly")) +
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-# Plot data for a = 1
-hp3 <- heter_model(a = 1, 
-                   x = {
-                     raw_x <- rand_uniform(-1, 1); 
-                     closed_form(~stand_dist(raw_x))
-                     }, 
-                   b = 128)$
-  gen(300, computed = select(ori_dat, x, e)) %>%
-  VI_MODEL$plot() +
-  ggtitle(quote("C. a = 1: Right-triangle")) +
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-patchwork::wrap_plots(hp1, hp2, hp3, ncol = 3)
 ```

### 2023-01-22 paper_comparison/paper_comparison/paper_comparison.Rmd
modify plot for different b

@@
 E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V})|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V})^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V})^{-1}\boldsymbol{\mu}_z\right),\\
-\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a,
+\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a,\\
 \boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a^T\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a^T,
@@
 # Generate data for b = 0.25
-ori_dat <- heter_model(a = 0, 
+dat_b025 <- heter_model(a = 0, 
                         x = {
@@
                           }, 
-                       b = 0.25)$gen(300)
-
-# Plot data for b = 0.25
-bp1 <- VI_MODEL$plot(ori_dat) +
-  ggtitle(quote("b = 0.25")) +
+                        b = 0.25)$gen(300) %>%
+  mutate(b = "b: 0.25")
+
+# Generate data for other b
+map(c(1, 4, 16, 64), function(b) {
+  heter_model(a = 0,
+              x = {
+                raw_x <- rand_uniform(-1, 1); 
+                closed_form(~stand_dist(raw_x))
+                }, 
+              b = b)$
+  gen(300, computed = select(dat_b025, x, e)) %>%
+  mutate(b = paste0("b: ", b))
+}) %>%
+
+  # Combined with data for b = 0.25
+  bind_rows(dat_b025) %>%
+  mutate(b = factor(b, levels = paste0("b: ", c(0.25, 1, 4, 16, 64)))) %>%
+  VI_MODEL$plot() +
+  facet_wrap(~b, scales = "free", ncol = 5) +
   theme_light() +
@@
   theme(axis.text = element_blank())
-
-# Generate and plot data for b = 1
-bp2 <- heter_model(a = 0, 
-                   x = {
-                     raw_x <- rand_uniform(-1, 1); 
-                     closed_form(~stand_dist(raw_x))
-                     }, 
-                   b = 1)$
-  gen(300, computed = select(ori_dat, x, e)) %>%
-  VI_MODEL$plot() +
-  ggtitle(quote("b = 1")) +
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-# Generate and plot data for b = 4
-bp3 <- heter_model(a = 0, 
-                   x = {
-                     raw_x <- rand_uniform(-1, 1); 
-                     closed_form(~stand_dist(raw_x))
-                     }, 
-                   b = 4)$
-  gen(300, computed = select(ori_dat, x, e)) %>%
-  VI_MODEL$plot() +
-  ggtitle(quote("b = 4")) +
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-# Generate data and plot data for b = 16
-bp4 <- heter_model(a = 0, 
-                   x = {
-                     raw_x <- rand_uniform(-1, 1); 
-                     closed_form(~stand_dist(raw_x))
-                     }, 
-                   b = 16)$
-  gen(300, computed = select(ori_dat, x, e)) %>%
-  VI_MODEL$plot() +
-  ggtitle(quote("b = 16")) +
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-# Generate data and plot data for b = 64
-bp5 <- heter_model(a = 0, 
-                   x = {
-                     raw_x <- rand_uniform(-1, 1); 
-                     closed_form(~stand_dist(raw_x))
-                     }, 
-                   b = 64)$
-  gen(300, computed = select(ori_dat, x, e)) %>%
-  VI_MODEL$plot() +
-  ggtitle(quote("b = 64")) +
-  theme_light() +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-
-patchwork::wrap_plots(bp1, bp2, bp3, bp4, bp5, ncol = 5)
 ```

### 2023-01-22 paper_comparison/paper_comparison/paper_comparison.Rmd
remove the use of x_raw

@@
 
-Visual patterns of non-linearity are simulated using four different orders of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$) and four different distributions of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
+Visual patterns of non-linearity are simulated using four different orders of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$).
+Four distributions of fitted values are also considered by using $X_{raw}$ follows different distributions: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
 A summary of the factors is given in Table \ref{tab:model-factor-table}.
@@
 data.frame(j = c("2", "3", "6", "18", ""),
-           x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$", ""),
+           x_dist = c("Uniform", "Normal", "Skewed", "Discrete uniform", ""),
            sigma = c("0.25", "1.00", "2.00", "4.00", ""),
@@
   kableExtra::add_header_above(c("Poly Order (sym1)", #"Order of Hermite polynomial\n(sym1)", 
-                                 "Distribution of sym2\n", 
+                                 "Distribution of fitted values", 
                                  "SD (sym3)", #"Standard deviation of polynomial model\n(sym3)",
@@
   gsub("sym1", "$j$", .) %>%
-  gsub("sym2", "$X_{raw}$", .) %>%
   gsub("sym3", "$\\\\sigma$", .) %>%
@@
 
-Different distributions of $X_{raw}$ help enriching the pool of visual patterns as illustrated in Figure \ref{fig:different-dist}.
+Different distributions of fitted values help enriching the pool of visual patterns as illustrated in Figure \ref{fig:different-dist}.
 The uniform and the normal distribution are symmetric and commonly assumed in statistical models.
@@
 
-```{r different-dist, fig.height = 2, fig.cap="Variations in fitted values ($X_{raw}$), that might affect perception of residual plots. Four different distribution are used."}
+```{r different-dist, fig.height = 2, fig.cap="Variations in fitted values, that might affect perception of residual plots. Four different distribution are used."}
 
@@
 Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups used in experiment I.
-This lineup is produced by the non-linearity model under $j = 6$ and $X_{raw} \sim N(0.0.3^2)$.
+This lineup is produced by the non-linearity model under $j = 6$ and fitted values follow a uniform distribution.
 The data plot location is $2^3 - 4$.
@@
 However, presenting only these lineups to subjects are considered to be bad practices as subjects will lose interest quickly. 
-Therefore, we plan to collect 6 more evaluations on the 279 lineups with $X_{raw} \sim U(-1,1)$, result in $(36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3 \times 6) / (20-2) = `r (36 * 20 + (4 * 4 * 3 + 3 * 5 * 3) * 3 * 6) / 18`$ subjects recruited for experiment III.
+Therefore, we plan to collect 6 more evaluations on the 279 lineups with fitted values follow a uniform distribution, result in $(36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3 \times 6) / (20-2) = `r (36 * 20 + (4 * 4 * 3 + 3 * 5 * 3) * 3 * 6) / 18`$ subjects recruited for experiment III.
 
@@
 
-There are `r 160 * 18`, `r 150 * 18` and `133 * 18` lineups evaluation made by 160, 150 and 133 subjects recruited for experiment I, II and III respectively.
+There are `r 160 * 18`, `r 150 * 18` and `r 133 * 18` lineups evaluation made by 160, 150 and 133 subjects recruited for experiment I, II and III respectively.
 In the total of `r nrow(filter(vi_survey, !attention_check))` lineup evaluations, `r nrow(filter(vi_survey, type == "polynomial", !attention_check))` use lineups produced by the non-linearity model, `r nrow(filter(vi_survey, type == "heteroskedasticity", !attention_check))` use lineups produced by the heteroskedasticity model, and `r nrow(filter(vi_survey, null_lineup, !attention_check))` use null lineups.
-Besides, there are `r nrow(filter(vi_survey, attention_check))` attention checks not included in the following analysis.
+Besides, there are `r nrow(filter(vi_survey, attention_check))` attention checks not included in the analysis.
 The collated dataset is provided in `vi_survey` of the `visage` `R` package.
 
-In the following analysis, lineups with $X_{raw} \sim U(-1,1)$ will be the focus of next two sections, because visual patterns are more likely to be revealed with uniform distribution.
-Besides, six more evaluations are collected for the uniform distribution in experiment III, which will produce more reliable and stable results.
-Analysis of other distributions can be found in Section \ref{effect-of-the-distribution-of-the-regressor}.
+In the following analysis, lineups with fitted values follows a uniform distribution will be the focus of next two sections.
+Visual patterns are more likely to be revealed under a uniform distribution.
+Additionally, we have collected extra evaluations on these lineups, which will result in more reliable analysis.
+Analysis of lineups with other fitted value distributions can be found in Section \ref{fitted-value-distribution}.
 
@@
 
-Figure \ref{fig:polypower} shows the estimated power of visual test on lineups produced by the non-linearity model with $X_{raw} \sim U(-1,1)$, against the natural logarithm of the effect $log_e(E)$, with a 5% significance level.
+Figure \ref{fig:polypower} shows the estimated power of visual test on lineups produced by the non-linearity model with fitted values follow a uniform distribution, against the natural logarithm of the effect $log_e(E)$, with a 5% significance level.
 At the bottom of the figure \ref{fig:polypower}, there are a sequence of example residual plots with increasing levels of $log_e(E)$.

### 2023-01-22 paper_comparison/paper_comparison/paper_comparison.tex
modify plot for different b
remove the use of x_raw

@@
 Visual patterns of non-linearity are simulated using four different
-orders of probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)) and
-four different distributions of \(X_{raw}\): (1) \(U(-1, 1)\), (2)
-\(N(0, 0.3^2)\), (3) \(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). A
-summary of the factors is given in Table \ref{tab:model-factor-table}.
+orders of probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)). Four
+distributions of fitted values are also considered by using \(X_{raw}\)
+follows different distributions: (1) \(U(-1, 1)\), (2) \(N(0, 0.3^2)\),
+(3) \(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). A summary of the
+factors is given in Table \ref{tab:model-factor-table}.
 
@@
 \toprule
-\multicolumn{1}{c}{Poly Order ($j$)} & \multicolumn{1}{c}{\makecell[c]{Distribution of $X_{raw}$\\}} & \multicolumn{1}{c}{SD ($\sigma$)} & \multicolumn{1}{c}{Heteroskedasticity Shape ($a$)} & \multicolumn{1}{c}{Heteroskedasticity ($b$)} & \multicolumn{1}{c}{Size ($n$)} \\
+\multicolumn{1}{c}{Poly Order ($j$)} & \multicolumn{1}{c}{Distribution of fitted values} & \multicolumn{1}{c}{SD ($\sigma$)} & \multicolumn{1}{c}{Heteroskedasticity Shape ($a$)} & \multicolumn{1}{c}{Heteroskedasticity ($b$)} & \multicolumn{1}{c}{Size ($n$)} \\
 \cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
-2 & $U(-1, 1)$ & 0.25 & -1 & 0.25 & 50\\
-3 & $N(0, 0.3^2)$ & 1.00 & 0 & 1.00 & 100\\
-6 & $lognormal(0, 0.6^2)/3$ & 2.00 & 1 & 4.00 & 300\\
-18 & $U\{1, 5\}$ & 4.00 &  & 16.00 & \\
+2 & Uniform & 0.25 & -1 & 0.25 & 50\\
+3 & Normal & 1.00 & 0 & 1.00 & 100\\
+6 & Skewed & 2.00 & 1 & 4.00 & 300\\
+18 & Discrete uniform & 4.00 &  & 16.00 & \\
  &  &  &  & 64.00 & \\
@@
 
-Different distributions of \(X_{raw}\) help enriching the pool of visual
-patterns as illustrated in Figure \ref{fig:different-dist}. The uniform
-and the normal distribution are symmetric and commonly assumed in
-statistical models. The adjusted log-normal distribution provides skewed
-density, while the discrete uniform distribution provides discreteness
-in residual plot.
+Different distributions of fitted values help enriching the pool of
+visual patterns as illustrated in Figure \ref{fig:different-dist}. The
+uniform and the normal distribution are symmetric and commonly assumed
+in statistical models. The adjusted log-normal distribution provides
+skewed density, while the discrete uniform distribution provides
+discreteness in residual plot.
 
@@
 
-\caption{Variations in fitted values ($X_{raw}$), that might affect perception of residual plots. Four different distribution are used.}\label{fig:different-dist}
+\caption{Variations in fitted values, that might affect perception of residual plots. Four different distribution are used.}\label{fig:different-dist}
 \end{figure}
@@
 used in experiment I. This lineup is produced by the non-linearity model
-under \(j = 6\) and \(X_{raw} \sim N(0.0.3^2)\). The data plot location
-is \(2^3 - 4\). All five subjects correctly identify the data plot from
-this lineup.
+under \(j = 6\) and fitted values follow a uniform distribution. The
+data plot location is \(2^3 - 4\). All five subjects correctly identify
+the data plot from this lineup.
 
@@
 Therefore, we plan to collect 6 more evaluations on the 279 lineups with
-\(X_{raw} \sim U(-1,1)\), result in
+fitted values follow a uniform distribution, result in
 \((36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3 \times 6) / (20-2) = 133\)
@@
 
-There are 2880, 2700 and \texttt{133\ *\ 18} lineups evaluation made by
-160, 150 and 133 subjects recruited for experiment I, II and III
-respectively. In the total of 7974 lineup evaluations, 3744 use lineups
-produced by the non-linearity model, 4230 use lineups produced by the
-heteroskedasticity model, and 720 use null lineups. Besides, there are
-886 attention checks not included in the following analysis. The
-collated dataset is provided in \texttt{vi\_survey} of the
-\texttt{visage} \texttt{R} package.
-
-In the following analysis, lineups with \(X_{raw} \sim U(-1,1)\) will be
-the focus of next two sections, because visual patterns are more likely
-to be revealed with uniform distribution. Besides, six more evaluations
-are collected for the uniform distribution in experiment III, which will
-produce more reliable and stable results. Analysis of other
-distributions can be found in Section
-\ref{effect-of-the-distribution-of-the-regressor}.
+There are 2880, 2700 and 2394 lineups evaluation made by 160, 150 and
+133 subjects recruited for experiment I, II and III respectively. In the
+total of 7974 lineup evaluations, 3744 use lineups produced by the
+non-linearity model, 4230 use lineups produced by the heteroskedasticity
+model, and 720 use null lineups. Besides, there are 886 attention checks
+not included in the analysis. The collated dataset is provided in
+\texttt{vi\_survey} of the \texttt{visage} \texttt{R} package.
+
+In the following analysis, lineups with fitted values follows a uniform
+distribution will be the focus of next two sections. Visual patterns are
+more likely to be revealed under a uniform distribution. Additionally,
+we have collected extra evaluations on these lineups, which will result
+in more reliable analysis. Analysis of lineups with other fitted value
+distributions can be found in Section \ref{fitted-value-distribution}.
 
@@
 Figure \ref{fig:polypower} shows the estimated power of visual test on
-lineups produced by the non-linearity model with
-\(X_{raw} \sim U(-1,1)\), against the natural logarithm of the effect
+lineups produced by the non-linearity model with fitted values follow a
+uniform distribution, against the natural logarithm of the effect
 \(log_e(E)\), with a 5\% significance level. At the bottom of the figure

### 2023-01-22 paper_comparison/appendix.Rmd
fix plot in the different fitted value distribution section

---
title: "appendix"
output: pdf_document
date: "2023-01-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

### 2023-01-22 paper_comparison/paper_comparison/paper_comparison.Rmd
fix plot in the different fitted value distribution section

@@
 
-## Effect of other factors
+## Effect of other factors used in the simulated model
 
@@
 
-(calc the diff in power and plot)
-
-(randomly subset uniform and check the change of the power)
-
-Although the distribution of the regressor is classically not restricted by the assumptions of linear regression, it will have an impact on the distribution of the fitted values and the visual effect of the residual plot.
-In our experiments, four distributions of the regressor are used to accommodate this variation.
-
-Figure \ref{fig:different-x-dist-poly-power} and Figure \ref{fig:different-x-dist-heter-power} illustrate the change of power of visual tests and conventional tests for different distribution of $X_{raw}$ used in the non-linearity model and the heteroskedasticity model.
-We focus on $log(E) > 2$ as the visual patterns become visible.
-
-For the non-linearity model, we do not observe significant difference of power of visual test between distributions, except for the discrete uniform distribution.
-It could be due to the fact that humans have difficulties in recognising visual patterns from discreteness, which makes the shape disconnected and incomplete (ref?).
-In terms of conventional tests, the discrete uniform  distribution is instead the one with the greatest power, followed by the lognormal distribution, uniform distribution and the normal distribution.
+Although the distribution of the regressor is classically not restricted by the assumptions of linear regression, it will have an impact on the fitted value distribution and the visual effect of the residual plot.
+In our experiments, four distributions of fitted values are considered to accommodate this variation.
+
+Figure \ref{fig:different-x-dist-poly-power} and Figure \ref{fig:different-x-dist-heter-power} illustrate the change of power of visual tests and conventional tests for different fitted distributions used in the non-linearity model and the heteroskedasticity model.
+We focus on $log(E) > 2$ where the visual patterns start to become recognisable.
+
+For the non-linearity model, we do not observe significant difference in power of visual test between distributions, except for the discrete uniform distribution.
+It is possibility due to the fact that humans have difficulties in recognising visual patterns from discreteness, which makes the shape disconnected and incomplete (ref?).
+In terms of conventional tests, the discrete uniform distribution is instead the one with the greatest power, followed by the lognormal distribution, the uniform distribution and the normal distribution.
 
@@
 
-In regards of the type of the simulated model, the power of visual tests under different distributions of regressor are always lower than the corresponding conventional tests, consistent with the result from the previous sections.
-
-```{r eval = FALSE}
-vi_survey %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
-  filter(type == "polynomial") %>%
-  filter(x_dist == "even_discrete") %>%
-  filter(log(effect_size) > 4) %>%
-  filter(p_value > 0.05) %>%
-  .$unique_lineup_id %>%
-  unique()
-```
-
-```{r eval = FALSE}
-VI_MODEL$plot_lineup(vi_lineup$poly_64$data, 
-                     remove_grid_line = TRUE, 
-                     theme = theme_light(),
-                     remove_axis = TRUE)
-```
+In regards of the type of the simulated model, the power of visual tests under different distributions of regressor are always lower than the corresponding conventional tests, which is consistent with the result from the previous sections.
 
@@
       levels = c("Uniform", "Normal", 
-                 "Lognormal", "Discrete")))
-
+                 "Lognormal", "Discrete"))) %>%
+  select(x_dist, log_effect_size, reject, conv_or_not)
+
+dat_different_dist %>%
   ggplot() +
-  geom_smooth(data=filter(dat_different_dist, 
-                          x_dist == "Uniform"),
+  geom_smooth(data = ~filter(.x, x_dist == "Uniform"),
               aes(log_effect_size, 
@@
               method.args = list(family = binomial), 
-              se = FALSE, size=3) +
-  geom_smooth(data=dat_different_dist,
-              aes(log_effect_size, 
+              se = FALSE, size = 3) +
+  geom_smooth(aes(log_effect_size, 
                   reject, 
@@
               method.args = list(family = binomial), 
-              se = FALSE, size=1) +
+              se = FALSE, 
+              size = 1) +
   theme_light(base_size = 5) +
@@
 
-```{r different-x-dist-heter-power, fig.height = 3, fig.width = 5, fig.cap = "Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model under different distributions of regressor."}
+```{r different-x-dist-heter-power, fig.height = 3, fig.width = 7, fig.cap = "Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model under different distributions of regressor."}
 
@@
                          stringr::str_to_title(x_dist))) %>%
-  mutate(conv_or_not = "Visual"))
+  mutate(conv_or_not = "Visual")) %>%
+  mutate(x_dist = factor(x_dist, 
+      levels = c("Uniform", "Normal", 
+                 "Lognormal", "Discrete"))) %>%
+  select(x_dist, log_effect_size, reject, conv_or_not)
 
@@
   ggplot() +
+  geom_smooth(data = ~filter(.x, x_dist == "Uniform"),
+              aes(log_effect_size, 
+                  reject), 
+                  col = "grey", 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, size = 3) +
   geom_smooth(aes(log_effect_size, 
@@
               method.args = list(family = binomial), 
-              se = FALSE) +
+              se = FALSE, 
+              size = 1) +
   theme_light(base_size = 5) +
-  scale_color_manual(values = rcartocolor::carto_pal(4, "Safe")[c(1,3,2,4)]) +
-  scale_size_manual(values = c(1, 2, 3)) +
+  scale_color_manual(values = rcartocolor::carto_pal(4, "Safe")[c(4,1,3,2)]) +
+  scale_size_manual(values = c(1, 2, 3, 4)) +
   xlab(expression(log[e] (Effect_size))) +
   ylab("Power") +
-  labs(col = "Distribution of regressor", linetype = "Test") +
-  facet_wrap(~conv_or_not)
-```
-
-### Shape
+  labs(col = "", linetype = "Test") +
+  facet_wrap(~conv_or_not) +
+  theme(legend.position="bottom")
+```
+
+### Shape of non-linearity
+
+### Shape of heteroskedasticity
 

### 2023-01-22 paper_comparison/paper_comparison/paper_comparison.tex
fix plot in the different fitted value distribution section

@@
 
-\hypertarget{effect-of-other-factors}{%
-\subsection{Effect of other factors}\label{effect-of-other-factors}}
+\hypertarget{effect-of-other-factors-used-in-the-simulated-model}{%
+\subsection{Effect of other factors used in the simulated
+model}\label{effect-of-other-factors-used-in-the-simulated-model}}
 
@@
 
-(calc the diff in power and plot)
-
-(randomly subset uniform and check the change of the power)
-
 Although the distribution of the regressor is classically not restricted
 by the assumptions of linear regression, it will have an impact on the
-distribution of the fitted values and the visual effect of the residual
-plot. In our experiments, four distributions of the regressor are used
-to accommodate this variation.
+fitted value distribution and the visual effect of the residual plot. In
+our experiments, four distributions of fitted values are considered to
+accommodate this variation.
 
@@
 \ref{fig:different-x-dist-heter-power} illustrate the change of power of
-visual tests and conventional tests for different distribution of
-\(X_{raw}\) used in the non-linearity model and the heteroskedasticity
-model. We focus on \(log(E) > 2\) as the visual patterns become visible.
-
-For the non-linearity model, we do not observe significant difference of
+visual tests and conventional tests for different fitted distributions
+used in the non-linearity model and the heteroskedasticity model. We
+focus on \(log(E) > 2\) where the visual patterns start to become
+recognisable.
+
+For the non-linearity model, we do not observe significant difference in
 power of visual test between distributions, except for the discrete
-uniform distribution. It could be due to the fact that humans have
+uniform distribution. It is possibility due to the fact that humans have
 difficulties in recognising visual patterns from discreteness, which
@@
 conventional tests, the discrete uniform distribution is instead the one
-with the greatest power, followed by the lognormal distribution, uniform
-distribution and the normal distribution.
+with the greatest power, followed by the lognormal distribution, the
+uniform distribution and the normal distribution.
 
@@
 under different distributions of regressor are always lower than the
-corresponding conventional tests, consistent with the result from the
-previous sections.
+corresponding conventional tests, which is consistent with the result
+from the previous sections.
 
@@
 
-\hypertarget{shape}{%
-\subsubsection{Shape}\label{shape}}
+\hypertarget{shape-of-non-linearity}{%
+\subsubsection{Shape of non-linearity}\label{shape-of-non-linearity}}
+
+\hypertarget{shape-of-heteroskedasticity}{%
+\subsubsection{Shape of
+heteroskedasticity}\label{shape-of-heteroskedasticity}}
 

### 2023-01-23 paper_comparison/paper_comparison/paper_comparison.Rmd
Fix the plot size

@@
 
-```{r example-poly-lineup, fig.height=7, fig.cap="Lineup poly-24 in experiment I. Can you spot the most different plot? \\label{fig:example-poly-lineup}"}
+```{r example-poly-lineup, fig.height = 7, fig.cap="Lineup poly-24 in experiment I. Can you spot the most different plot? \\label{fig:example-poly-lineup}"}
 VI_MODEL$plot_lineup(vi_lineup$poly_24$data, 
@@
 
-```{r different-shape-of-heter, fig.height=2.67, fig.cap="Heteroskedasticity forms used in experiment II.
+```{r different-shape-of-heter, fig.height = 2.67, fig.cap="Heteroskedasticity forms used in experiment II.
 Three different shapes (a = -1, 0, 1) are used in the experiment to create left-triangle, butterfly and right-triangle shapes."}
@@
 
-```{r p-value-comparison, fig.width=8, fig.height=4, fig.cap="Rejection rate ($p$-value $<0.05$) of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot.
+```{r p-value-comparison, fig.width = 8, fig.height = 4, fig.cap="Rejection rate ($p$-value $<0.05$) of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot.
 The visual test rejects less frequently than the conventional test.
@@
 
-```{r different-x-dist-poly-power, fig.height = 3, fig.width = 7, fig.cap = "Power of conventional tests and visual tests on lineups produced by the non-linearity model under different fitted value distributions."}
+```{r different-x-dist-poly-power, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups produced by the non-linearity model under different fitted value distributions."}
 
@@
               size = 1) +
-  theme_light(base_size = 5) +
+  theme_light() +
   scale_color_manual(values = rcartocolor::carto_pal(4, "Safe")[c(4,1,3,2)]) +
@@
 
-```{r different-x-dist-heter-power, fig.height = 3, fig.width = 7, fig.cap = "Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model under different distributions of regressor."}
+```{r different-x-dist-heter-power, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model under different distributions of regressor."}
 
@@
 
-### Shape of heteroskedasticity
-
-```{r eval = FALSE}
-poly_conv_sim %>%
-  filter(x_dist == "uniform", shape %in% c(3, 4)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  select(shape, log_effect_size, (RESET3_p_value:RESET9_p_value)) %>%
-  pivot_longer(RESET3_p_value:RESET9_p_value) %>%
-  mutate(name = gsub("_p_value", " ", name)) %>%
-  mutate(name = gsub("RESET", "", name)) %>%
-  mutate(name = gsub("test", "", name)) %>%
-  mutate(reject = as.numeric(value < 0.05)) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject, col = name), 
-              se = FALSE, 
-              method = "glm",
-              method.args = list(family = binomial())) +
-  xlab(expression(log[e] (Effect_size))) +
-  ylab("Power") +
-  scale_color_manual(values = rev(rcartocolor::carto_pal(7, "RedOr"))) +
-  theme_light() +
-  labs(col = "RESET test order") +
-  facet_wrap(~shape)
-```
-
-```{r poly-power-j, fig.cap = "Power of conventional tests and visual tests on lineups produced by the non-linearity model with four different shapes controlled by the parameter $j$."}
+```{r poly-power-j, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups produced by the non-linearity model with four different shapes controlled by the parameter $j$."}
 
@@
   mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>%
-  mutate(conv_type = "Conventional") %>%
+  mutate(conv_or_not = "Conventional") %>%
   bind_rows(vi_survey %>%
@@
   mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>%
-  mutate(conv_type = "Visual")) %>%
+  mutate(conv_or_not = "Visual")) %>%
+  mutate(shape = factor(shape, levels = c("S", "U", "M", "Triple-U"))) %>%
+  select(shape, log_effect_size, reject, conv_or_not) %>%
   ggplot() +
-  geom_smooth(aes(log_effect_size, reject, col = factor(shape)), 
+  geom_smooth(aes(log_effect_size, reject, col = shape), 
               se = FALSE, 
@@
   theme_light() +
-  facet_wrap(~conv_type) +
+  theme(legend.position = "bottom") +
+  facet_wrap(~conv_or_not) +
   labs(col = "Non-linearity shape")
@@
 
-```{r heter-power-a, fig.cap = "Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$."}
+### Shape of heteroskedasticity
+
+```{r heter-power-a, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$."}
 
@@
   mutate(reject = as.numeric(BP_p_value < 0.05)) %>%
-  mutate(a = c("left-triangle", "butterfly", "right-triangle")[a + 2]) %>%
-  mutate(conv_type = "Conventional") %>%
+  mutate(a = c("Left-triangle", "Butterfly", "Right-triangle")[a + 2]) %>%
+  mutate(conv_or_not = "Conventional") %>%
   bind_rows(vi_survey %>%
@@
   mutate(reject = as.numeric(p_value < 0.05)) %>%
-  mutate(a = c("left-triangle", "butterfly", "right-triangle")[a + 2]) %>%
-  mutate(conv_type = "Visual")) %>%
+  mutate(a = c("Left-triangle", "Butterfly", "Right-triangle")[a + 2]) %>%
+  mutate(conv_or_not = "Visual")) %>%
+  mutate(a = factor(a, levels = c("Left-triangle", "Butterfly", "Right-triangle"))) %>%
+  select(a, log_effect_size, reject, conv_or_not) %>%
   ggplot() +
-  geom_smooth(aes(log_effect_size, reject, col = factor(a)), 
+  geom_smooth(aes(log_effect_size, reject, col = a), 
               se = FALSE, 
@@
   scale_color_manual(values = rcartocolor::carto_pal(3, "Vivid")) +
+  theme(legend.position = "bottom") +
   theme_light() +
-  facet_wrap(~conv_type) +
+  facet_wrap(~conv_or_not) +
   labs(col = "Heteroskedasticity shape")
@@
 
-```{r p-value-comparison-old, eval = FALSE, fig.height=4, fig.cap="Visual test p-value compared to conventional test p-value for lineups produced by heteroskedasticity and non-linearity model.
+```{r p-value-comparison-old, eval = FALSE, fig.height = 4, fig.cap="Visual test p-value compared to conventional test p-value for lineups produced by heteroskedasticity and non-linearity model.
 The scatter plot is drawn on a square root scale.

### 2023-01-23 paper_comparison/paper_comparison/paper_comparison.Rmd
fix plot caption

@@
 
-```{r first-example-lineup, fig.height = 7, fig.cap="Visual testing is conducted using a lineup, as in the example here.
+```{r first-example-lineup, fig.height = 7, fig.cap = "Visual testing is conducted using a lineup, as in the example here.
 The residual plot computed from the observed data (plot $2^2 + 2$, exhibiting non-linearity) is embedded among 19 null plots, where the residuals are simulated from a standard error model.
@@
 
-\noindent where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,  $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of regressors including the intercept used in the regression equation, $\sigma^2$ is the assumed variance of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual variance of the error term, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of the residuals with $\boldsymbol{Z}$ be any variables leave out by the model and $\boldsymbol{\beta}_z$ be the corresponding coefficients.
+\noindent where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,  $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of regressors including the intercept used in the regression equation, $\sigma^2\boldsymbol{I}$ is the assumed variance of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual variance of the error term, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of the residuals with $\boldsymbol{Z}$ be any variables leave out by the model and $\boldsymbol{\beta}_z$ be the corresponding coefficients.
 
 To study various factors contributing to the power of the visual test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors.
-These include the distribution of the regressor and the type of the simulated model (modify this according to the result).
+These include the distribution of the regressor and the type of the simulation model (modify this according to the result).
 
@@
 
-```{r different-shape-of-herimite, fig.height = 2, fig.cap="Polynomial forms generated for the residual plots used in experiment I.
+```{r different-shape-of-herimite, fig.height = 2, fig.cap = "Polynomial forms generated for the residual plots used in experiment I.
 The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$."}
@@
 
-```{r different-dist, fig.height = 2, fig.cap="Variations in fitted values, that might affect perception of residual plots. Four different distribution are used."}
+```{r different-dist, fig.height = 2, fig.cap = "Variations in fitted values, that might affect perception of residual plots. Four different distributions are used."}
 
@@
 Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups used in experiment I.
-This lineup is produced by the non-linearity model under $j = 6$ and fitted values follow a uniform distribution.
+This lineup is produced by the non-linearity model with $j = 6$ and uniform fitted values.
 The data plot location is $2^3 - 4$.
@@
 
-```{r different-n, fig.height = 2.67, fig.cap="Three different values of $n$ are used in experiment I, II and III to control the strength of the signal."}
+```{r different-n, fig.height = 2.67, fig.cap = "Number of observations $n$ have an impact on the strength of the signal. Three different values of $n$ are used in experiment I, II and III."}
 
@@
 
-```{r different-sigma, fig.height = 2, fig.cap="Four different values of $\\sigma$ are used in the experiment I to control the strength of the signal."}
+```{r different-sigma, fig.height = 2, fig.cap = "For the non-lienearity model, the standard deviation of the error term $\\sigma$ affects the strength of the signal.
+Four different values of $\\sigma$ are used in the experiment I."}
 
@@
 
-```{r example-poly-lineup, fig.height = 7, fig.cap="Lineup poly-24 in experiment I. Can you spot the most different plot? \\label{fig:example-poly-lineup}"}
+```{r example-poly-lineup, fig.height = 7, fig.cap = "One of the lineups containing non-linearity patterns used in experiment I.
+Can you spot the most different plot?
+The data plot is positioned at $2^3 + 1$."}
 VI_MODEL$plot_lineup(vi_lineup$poly_24$data, 
@@
 
-```{r different-shape-of-heter, fig.height = 2.67, fig.cap="Heteroskedasticity forms used in experiment II.
+```{r different-shape-of-heter, fig.height = 2.67, fig.cap = "Heteroskedasticity forms used in experiment II.
 Three different shapes (a = -1, 0, 1) are used in the experiment to create left-triangle, butterfly and right-triangle shapes."}
@@
 
-```{r different-b, fig.height = 2, fig.cap="Five different values of $b$ are used in experiment II to control the strength of the signal."}
+```{r different-b, fig.height = 2, fig.cap = "Five different values of $b$ are used in experiment II to control the strength of the signal."}
 
@@
 
-```{r example-heter-lineup, fig.height = 7, fig.cap="Lineup heter-169 in experiment II. Can you spot the most different plot? \\label{fig:example-heter-lineup}"}
+```{r example-heter-lineup, fig.height = 7, fig.cap = "One of the lineups containing heteroskedasticity pattern used in experiment II.
+Can you spot the most different plot?
+The data plot is positioned at $3^3 - 3^2$"}
 
@@
 However, presenting only these lineups to subjects are considered to be bad practices as subjects will lose interest quickly. 
-Therefore, we plan to collect 6 more evaluations on the 279 lineups with fitted values follow a uniform distribution, result in $(36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3 \times 6) / (20-2) = `r (36 * 20 + (4 * 4 * 3 + 3 * 5 * 3) * 3 * 6) / 18`$ subjects recruited for experiment III.
+Therefore, we plan to collect 6 more evaluations on the 279 lineups with uniform fitted values, result in $(36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3 \times 6) / (20-2) = `r (36 * 20 + (4 * 4 * 3 + 3 * 5 * 3) * 3 * 6) / 18`$ subjects recruited for experiment III.
 
@@
 
-In the following analysis, lineups with fitted values follows a uniform distribution will be the focus of next two sections.
+In the following analysis, lineups with uniform fitted values will be the focus of next two sections.
 Visual patterns are more likely to be revealed under a uniform distribution.
@@
 
-Figure \ref{fig:polypower} shows the estimated power of visual test on lineups produced by the non-linearity model with fitted values follow a uniform distribution, against the natural logarithm of the effect $log_e(E)$, with a 5% significance level.
+Figure \ref{fig:polypower} shows the estimated power of visual test on lineups produced by the non-linearity model with uniform fitted values, against the natural logarithm of the effect $log_e(E)$, with a 5% significance level.
 At the bottom of the figure \ref{fig:polypower}, there are a sequence of example residual plots with increasing levels of $log_e(E)$.
@@
 
-For the non-linearity model, the power curve of RESET test climbs aggressively from
+For non-linearity patterns, the power curve of RESET test climbs aggressively from
 `r round(unname(predict(conv_poly_glmmod, newdata = data.frame(log_effect_size = 0), type = "response")) * 100)`% to `r round(unname(predict(conv_poly_glmmod, newdata = data.frame(log_effect_size = 2), type = "response")) * 100)`% as $log_e(E)$ increases from 0 to 2, while power of other tests respond inactively to the change of effect, showing that RESET test is way more sensitive to the type of model defects that being considered.
@@
 
-In terms of the heteroskedasticity model, the power of Breusch-Pagan test is also almost always greater than the power of visual test.
+In terms of heteroskedasticity patterns, the power of Breusch-Pagan test is also almost always greater than the power of visual test.
 For $0 \leq log_e(E) \leq 2$, where the power curve of the visual test remains at a low level, the Breusch-Pagan test still have a decent amount of chance of rejecting $H_0$.
@@
 
-The power of visual test arises steadily as $log_e(E)$ increases from 2 to 5 for both non-linearity model and heteroskedasticity model, suggesting that the effect starts to make significant impact on the degree of the presence of the designed visual features.
+The power of visual test arises steadily as $log_e(E)$ increases from 2 to 5 for both non-linearity patterns and heteroskedasticity patterns, suggesting that the effect starts to make significant impact on the degree of the presence of the designed visual features.
 This can also be observed from the example residual plots that when $log_e(E) = 2.5$, a weak "S-shape" and a weak "triangle" shape are presented in Figure \ref{fig:polypower} and Figure \ref{fig:heterpower} respectively.
@@
 
-```{r polypower, fig.height = 3, fig.width = 5, fig.cap="Comparison of power between different tests for non-linear patterns (uniform fitted values only).
+```{r polypower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for non-linear patterns (uniform fitted values only).
 Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups.
@@
 
-```{r heterpower, fig.height = 3, fig.width = 5, fig.cap="Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only).
+```{r heterpower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only).
 Main plot shows the power curves, with dots indicating human evaluations of lineups.
@@
 
-```{r p-value-comparison, fig.width = 8, fig.height = 4, fig.cap="Rejection rate ($p$-value $<0.05$) of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot.
+```{r p-value-comparison, fig.width = 8, fig.height = 4, fig.cap = "Rejection rate ($p$-value $<0.05$) of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot.
 The visual test rejects less frequently than the conventional test.
@@
 
-Figure \ref{fig:p-value-comparison} provides a mosaic plot showing the rejection rate of visual tests and conventional tests for both the non-linearity model and the heteroskedasticity model.
-
-For lineups produced by the non-linearity model, conventional tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`%
+Figure \ref{fig:p-value-comparison} provides a mosaic plot showing the rejection rate of visual tests and conventional tests for both non-linearity patterns and heteroskedasticity patterns.
+
+For lineups containing non-linearity patterns, conventional tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`%
 and visual tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the time.
@@
 
-For lineups produced by the non-linearity model, conventional tests reject `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`%
+For lineups containing non-linearity patterns, conventional tests reject `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`%
 and visual tests reject `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the time.
@@
 The data plot at panel 17 contains a "Butterfly" shape.
-It has effect size$ = 4.02$, somewhat surprising that it is not detected by the BP test.'}
+It has effect size $ = 4.02$, somewhat surprising that it is not detected by the BP test.'}
 # plot heter-331
@@
 
-## Effect of other factors used in the simulated model
+## Effect of other factors used in the simulation model
 
@@
 
-For the non-linearity model, we do not observe significant difference in power of visual test between distributions, except for the discrete uniform distribution.
+For non-linearity patterns, we do not observe significant difference in power of visual test between distributions, except for the discrete uniform distribution.
 It is possibility due to the fact that humans have difficulties in recognising visual patterns from discreteness, which makes the shape disconnected and incomplete (ref?).
@@
 
-For the heteroskedasticity model, the power of visual test on lineups with uniform distribution has the greatest power.
+For heteroskedasticity patterns, the power of visual test on lineups with uniform distribution has the greatest power.
 This is as expected since other three distributions could reduce the chance of revealing the underlying visual pattern because of uneven data points.
-Although the power of visual test under the discrete uniform  distribution is the lowest in the case of non-linearity model, it has a relatively great power this time.
+Although the power of visual test under the discrete uniform distribution is the lowest in the case of non-linearity patterns, it has a relatively great power this time.
 Considering for heteroskedasticity, the visual patterns are usually detected by connecting the maximum and minimum residuals separately at different fitted values.
@@
 
-In regards of the type of the simulated model, the power of visual tests under different distributions of regressor are always lower than the corresponding conventional tests, which is consistent with the result from the previous sections.
-
-```{r different-x-dist-poly-power, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups produced by the non-linearity model under different fitted value distributions."}
+In regards of the type of the simulation model, the power of visual tests under different fitted value distributions are always lower than the corresponding conventional tests, which is consistent with the result reported in the previous sections.
+
+```{r different-x-dist-poly-power, fig.height = 4, fig.cap = "Comparison of power of conventional tests and visual tests on lineups produced by the non-linearity model under different fitted value distributions."}
 
@@
 ### Shape of non-linearity
+
+Shape of the non-linearity pattern is another factor contributed to the simulation model.
 
@@
 
-```{r p-value-comparison-old, eval = FALSE, fig.height = 4, fig.cap="Visual test p-value compared to conventional test p-value for lineups produced by heteroskedasticity and non-linearity model.
+```{r p-value-comparison-old, eval = FALSE, fig.height = 4, fig.cap = "Visual test p-value compared to conventional test p-value for lineups produced by heteroskedasticity and non-linearity model.
 The scatter plot is drawn on a square root scale.

### 2023-01-23 paper_comparison/paper_comparison/paper_comparison.tex
Fix the plot size
fix plot caption

@@
 regressors including the intercept used in the regression equation,
-\(\sigma^2\) is the assumed variance of the error term when \(H_0\) is
-true, \(\boldsymbol{V}\) is the actual variance of the error term, and
+\(\sigma^2\boldsymbol{I}\) is the assumed variance of the error term
+when \(H_0\) is true, \(\boldsymbol{V}\) is the actual variance of the
+error term, and
 \(\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z\)
@@
 collated data grouped by levels of factors. These include the
-distribution of the regressor and the type of the simulated model
+distribution of the regressor and the type of the simulation model
 (modify this according to the result).
@@
 
-\caption{Variations in fitted values, that might affect perception of residual plots. Four different distribution are used.}\label{fig:different-dist}
+\caption{Variations in fitted values, that might affect perception of residual plots. Four different distributions are used.}\label{fig:different-dist}
 \end{figure}
@@
 used in experiment I. This lineup is produced by the non-linearity model
-under \(j = 6\) and fitted values follow a uniform distribution. The
-data plot location is \(2^3 - 4\). All five subjects correctly identify
-the data plot from this lineup.
+with \(j = 6\) and uniform fitted values. The data plot location is
+\(2^3 - 4\). All five subjects correctly identify the data plot from
+this lineup.
 
@@
 
-\caption{Three different values of $n$ are used in experiment I, II and III to control the strength of the signal.}\label{fig:different-n}
+\caption{Number of observations $n$ have an impact on the strength of the signal. Three different values of $n$ are used in experiment I, II and III.}\label{fig:different-n}
 \end{figure}
@@
 
-\caption{Four different values of $\sigma$ are used in the experiment I to control the strength of the signal.}\label{fig:different-sigma}
+\caption{For the non-lienearity model, the standard deviation of the error term $\sigma$ affects the strength of the signal.
+Four different values of $\sigma$ are used in the experiment I.}\label{fig:different-sigma}
 \end{figure}
@@
 
-\caption{Lineup poly-24 in experiment I. Can you spot the most different plot? \label{fig:example-poly-lineup}}\label{fig:example-poly-lineup}
+\caption{One of the lineups containing non-linearity patterns used in experiment I. Can you spot the most different plot? The data plot is positioned at $2^3 + 1$.}\label{fig:example-poly-lineup}
 \end{figure}
@@
 
-\caption{Lineup heter-169 in experiment II. Can you spot the most different plot? \label{fig:example-heter-lineup}}\label{fig:example-heter-lineup}
+\caption{One of the lineups containing heteroskedasticity pattern used in experiment II.
+Can you spot the most different plot?
+The data plot is positioned at $3^3 - 3^2$}\label{fig:example-heter-lineup}
 \end{figure}
@@
 Therefore, we plan to collect 6 more evaluations on the 279 lineups with
-fitted values follow a uniform distribution, result in
+uniform fitted values, result in
 \((36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3 \times 6) / (20-2) = 133\)
@@
 
-In the following analysis, lineups with fitted values follows a uniform
-distribution will be the focus of next two sections. Visual patterns are
-more likely to be revealed under a uniform distribution. Additionally,
-we have collected extra evaluations on these lineups, which will result
-in more reliable analysis. Analysis of lineups with other fitted value
-distributions can be found in Section \ref{fitted-value-distribution}.
+In the following analysis, lineups with uniform fitted values will be
+the focus of next two sections. Visual patterns are more likely to be
+revealed under a uniform distribution. Additionally, we have collected
+extra evaluations on these lineups, which will result in more reliable
+analysis. Analysis of lineups with other fitted value distributions can
+be found in Section \ref{fitted-value-distribution}.
 
@@
 Figure \ref{fig:polypower} shows the estimated power of visual test on
-lineups produced by the non-linearity model with fitted values follow a
-uniform distribution, against the natural logarithm of the effect
-\(log_e(E)\), with a 5\% significance level. At the bottom of the figure
-\ref{fig:polypower}, there are a sequence of example residual plots with
-increasing levels of \(log_e(E)\). Readers can evaluate them from left
-to right and determine at which level the departure from a good residual
-plot becomes detectable.
+lineups produced by the non-linearity model with uniform fitted values,
+against the natural logarithm of the effect \(log_e(E)\), with a 5\%
+significance level. At the bottom of the figure \ref{fig:polypower},
+there are a sequence of example residual plots with increasing levels of
+\(log_e(E)\). Readers can evaluate them from left to right and determine
+at which level the departure from a good residual plot becomes
+detectable.
 
@@
 
-For the non-linearity model, the power curve of RESET test climbs
+For non-linearity patterns, the power curve of RESET test climbs
 aggressively from 15\% to 69\% as \(log_e(E)\) increases from 0 to 2,
@@
 
-In terms of the heteroskedasticity model, the power of Breusch-Pagan
-test is also almost always greater than the power of visual test. For
+In terms of heteroskedasticity patterns, the power of Breusch-Pagan test
+is also almost always greater than the power of visual test. For
 \(0 \leq log_e(E) \leq 2\), where the power curve of the visual test
@@
 The power of visual test arises steadily as \(log_e(E)\) increases from
-2 to 5 for both non-linearity model and heteroskedasticity model,
+2 to 5 for both non-linearity patterns and heteroskedasticity patterns,
 suggesting that the effect starts to make significant impact on the
@@
 Figure \ref{fig:p-value-comparison} provides a mosaic plot showing the
-rejection rate of visual tests and conventional tests for both the
-non-linearity model and the heteroskedasticity model.
-
-For lineups produced by the non-linearity model, conventional tests
-reject 69\% and visual tests reject 32\% of the time. Of the tests
-rejected by the conventional test, 46\% are rejected by the visual test,
-that is, approximately half as many as the conventional test. There are
-no lineups that are rejected by the visual test but not by the
-conventional test.
-
-For lineups produced by the non-linearity model, conventional tests
-reject 76\% and visual tests reject 56\% of the time. Of the tests
-rejected by the conventional test, 73\% are rejected by the visual test,
-that is, approximately half as many as the conventional test.
+rejection rate of visual tests and conventional tests for both
+non-linearity patterns and heteroskedasticity patterns.
+
+For lineups containing non-linearity patterns, conventional tests reject
+69\% and visual tests reject 32\% of the time. Of the tests rejected by
+the conventional test, 46\% are rejected by the visual test, that is,
+approximately half as many as the conventional test. There are no
+lineups that are rejected by the visual test but not by the conventional
+test.
+
+For lineups containing non-linearity patterns, conventional tests reject
+76\% and visual tests reject 56\% of the time. Of the tests rejected by
+the conventional test, 73\% are rejected by the visual test, that is,
+approximately half as many as the conventional test.
 
@@
 The data plot at panel 17 contains a "Butterfly" shape.
-It has effect size$ = 4.02$, somewhat surprising that it is not detected by the BP test.}\label{fig:heter-example}
-\end{figure}
-
-\hypertarget{effect-of-other-factors-used-in-the-simulated-model}{%
-\subsection{Effect of other factors used in the simulated
-model}\label{effect-of-other-factors-used-in-the-simulated-model}}
+It has effect size $ = 4.02$, somewhat surprising that it is not detected by the BP test.}\label{fig:heter-example}
+\end{figure}
+
+\hypertarget{effect-of-other-factors-used-in-the-simulation-model}{%
+\subsection{Effect of other factors used in the simulation
+model}\label{effect-of-other-factors-used-in-the-simulation-model}}
 
@@
 
-For the non-linearity model, we do not observe significant difference in
+For non-linearity patterns, we do not observe significant difference in
 power of visual test between distributions, except for the discrete
@@
 
-For the heteroskedasticity model, the power of visual test on lineups
+For heteroskedasticity patterns, the power of visual test on lineups
 with uniform distribution has the greatest power. This is as expected
@@
 power of visual test under the discrete uniform distribution is the
-lowest in the case of non-linearity model, it has a relatively great
+lowest in the case of non-linearity patterns, it has a relatively great
 power this time. Considering for heteroskedasticity, the visual patterns
@@
 
-In regards of the type of the simulated model, the power of visual tests
-under different distributions of regressor are always lower than the
-corresponding conventional tests, which is consistent with the result
-from the previous sections.
+In regards of the type of the simulation model, the power of visual
+tests under different fitted value distributions are always lower than
+the corresponding conventional tests, which is consistent with the
+result reported in the previous sections.
 
@@
 
-\caption{Power of conventional tests and visual tests on lineups produced by the non-linearity model under different fitted value distributions.}\label{fig:different-x-dist-poly-power}
+\caption{Comparison of power of conventional tests and visual tests on lineups produced by the non-linearity model under different fitted value distributions.}\label{fig:different-x-dist-poly-power}
 \end{figure}
@@
 \subsubsection{Shape of non-linearity}\label{shape-of-non-linearity}}
+
+Shape of the non-linearity pattern is another factor contributed to the
+simulation model.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/poly-power-j-1}
+
+}
+
+\caption{Power of conventional tests and visual tests on lineups produced by the non-linearity model with four different shapes controlled by the parameter $j$.}\label{fig:poly-power-j}
+\end{figure}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/poly-power-j-1}
-
-}
-
-\caption{Power of conventional tests and visual tests on lineups produced by the non-linearity model with four different shapes controlled by the parameter $j$.}\label{fig:poly-power-j}
-\end{figure}
-
-\begin{figure}
-
 {\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/heter-power-a-1}

### 2023-01-23 paper_comparison/paper_comparison/paper_comparison.Rmd
let lineups have the same aspect ratio as in the experiment

@@
 
-```{r first-example-lineup, fig.height = 7, fig.cap = "Visual testing is conducted using a lineup, as in the example here.
+```{r first-example-lineup, fig.height = 7, fig.width = 7, fig.cap = "Visual testing is conducted using a lineup, as in the example here.
 The residual plot computed from the observed data (plot $2^2 + 2$, exhibiting non-linearity) is embedded among 19 null plots, where the residuals are simulated from a standard error model.
@@
 
-```{r example-poly-lineup, fig.height = 7, fig.cap = "One of the lineups containing non-linearity patterns used in experiment I.
+```{r example-poly-lineup, fig.height = 7, fig.width = 7, fig.cap = "One of the lineups containing non-linearity patterns used in experiment I.
 Can you spot the most different plot?
@@
 
-```{r example-heter-lineup, fig.height = 7, fig.cap = "One of the lineups containing heteroskedasticity pattern used in experiment II.
+```{r example-heter-lineup, fig.height = 7, fig.width = 7, fig.cap = "One of the lineups containing heteroskedasticity pattern used in experiment II.
 Can you spot the most different plot?
@@
 
-```{r poly-example, eval = FALSE, fig.height = 7, fig.cap = "Example of a non-linearity that is rejected by the visual test but not by the RESET test.
+```{r poly-example, eval = FALSE, fig.height = 7, fig.width = 7, fig.cap = "Example of a non-linearity that is rejected by the visual test but not by the RESET test.
 The data plot at panel three exhibits strong non-linear pattern.
@@
 
-```{r heter-example, fig.height = 7, fig.cap = 'The single heteroskedasticity lineup that is rejected by the visual test but not by the BP test.
+```{r heter-example, fig.height = 7, fig.width = 7, fig.cap = 'The single heteroskedasticity lineup that is rejected by the visual test but not by the BP test.
 The data plot at panel 17 contains a "Butterfly" shape.

### 2023-01-23 paper_comparison/appendix.Rmd
combine two different x_dist plots into one

@@
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+library(visage)
 ```
+
+# Appendix
+
+## Effect derivation
+
+## $p$-value scatter plot
+
+```{r p-value-comparison-old, eval = FALSE, fig.height = 4, fig.cap = "Visual test p-value compared to conventional test p-value for lineups produced by heteroskedasticity and non-linearity model.
+The scatter plot is drawn on a square root scale.
+Every dot on the plot corresponds to a lineup.
+Most of the dots are fallen on the left of the y = x line indicating visual test p-value is almost always higher than conventional test p-value.
+The blue curve is a smoothing of the dots showing a positive trend.
+"}
+vi_survey %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  left_join(vi_survey %>% 
+           count(unique_lineup_id) %>%
+           rename(num_subject = n)) %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
+  ggplot() +
+  geom_abline() +
+  geom_point(aes(conventional_p_value, p_value), alpha = 0.4) +
+  geom_smooth(aes(conventional_p_value, p_value), se = FALSE, span = 0.6) +
+  scale_x_sqrt() +
+  scale_y_sqrt() +
+  facet_wrap(~type) +
+  ylab("Visual test p-value on square root scale") +
+  xlab("Conventional test p-value on square root scale") +
+  theme_light()
+```
+
+## A collection interesting lineups (unusual results)
+
+why unusual? what is the possible explanations?
+
+# targe journal
+
+JRSSB: Journal of the Royal Statistical Society Series B (Statistical Methodology)
+Deadline: Jan 1, Apr 1
+
+Reading: style of writing, author guideline (https://rss.onlinelibrary.wiley.com/hub/journal/14679868/author-guidelines)
+
+JCGS
+JDSS

### 2023-01-23 paper_comparison/paper_comparison/paper_comparison.Rmd
combine two different x_dist plots into one

@@
 
-```{r different-x-dist-poly-power, fig.height = 4, fig.cap = "Comparison of power of conventional tests and visual tests on lineups produced by the non-linearity model under different fitted value distributions."}
-
-dat_different_dist <- poly_conv_sim %>% 
+```{r different-x-dist-poly-power, fig.cap = "Comparison of power of conventional tests and visual tests on lineups produced by the non-linearity model under different fitted value distributions."}
+
+# Extract three datasets, one for poly conventional, one for heter conventional
+# and one for visual tests.
+list(
+poly_conv_sim %>% 
   mutate(log_effect_size = log(effect_size)) %>%
@@
   rename(RESET_p_value = RESET4_p_value) %>%
-  pivot_longer(RESET_p_value) %>%
   mutate(x_dist = ifelse(x_dist == "even_discrete",
@@
                          stringr::str_to_title(x_dist))) %>%
-  mutate(name = gsub("_p_value", " test", name)) %>%
-  mutate(reject = value < 0.05) %>%
+  mutate(reject = as.numeric(RESET_p_value < 0.05)) %>%
   mutate(conv_or_not = "Conventional") %>%
-  bind_rows(vi_survey %>%
+  mutate(type = "non-linearity") %>%
+  select(type, x_dist, log_effect_size, reject, conv_or_not)
+,
+heter_conv_sim %>% 
+  mutate(log_effect_size = log(effect_size)) %>%
+  select(x_dist, log_effect_size, BP_p_value) %>%
+  mutate(x_dist = ifelse(x_dist == "even_discrete",
+                         "Discrete", 
+                         stringr::str_to_title(x_dist))) %>%
+  mutate(reject = as.numeric(BP_p_value < 0.05)) %>%
+  mutate(conv_or_not = "Conventional") %>%
+  mutate(type = "heteroskedasticity") %>%
+  select(type, x_dist, log_effect_size, reject, conv_or_not)
+,
+vi_survey %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
   group_by(unique_lineup_id) %>%
   summarise(across(everything(), first)) %>%
-  filter(type == "polynomial") %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
   mutate(reject = as.numeric(p_value <= 0.05)) %>%
@@
                          stringr::str_to_title(x_dist))) %>%
-  mutate(conv_or_not = "Visual")) %>%
+  mutate(conv_or_not = "Visual") %>%
+  select(type, x_dist, log_effect_size, reject, conv_or_not)
+) %>%
+
+  # Bind rows
+  map_df(~.x) %>%
+
+  # Manually control factor ordering
+  mutate(type = factor(type, 
+                       levels = c("non-linearity", 
+                                  "heteroskedasticity"))) %>%
   mutate(x_dist = factor(x_dist, 
-      levels = c("Uniform", "Normal", 
-                 "Lognormal", "Discrete"))) %>%
-  select(x_dist, log_effect_size, reject, conv_or_not)
-
-dat_different_dist %>%
+                         levels = c("Uniform", 
+                                    "Normal", 
+                                    "Lognormal", 
+                                    "Discrete"))) %>%
+
   ggplot() +
+
+  # Bold the uniform smooth line
   geom_smooth(data = ~filter(.x, x_dist == "Uniform"),
-              aes(log_effect_size, 
-                  reject), 
+              aes(log_effect_size, reject), 
               col = "grey", 
@@
               method.args = list(family = binomial), 
-              se = FALSE, size = 3) +
-  geom_smooth(aes(log_effect_size, 
-                  reject, 
-                  col = x_dist), 
+              se = FALSE, 
+              size = 3) +
+
+  # Draw all smooth lines
+  geom_smooth(aes(log_effect_size, reject, col = x_dist), 
               method = "glm", 
@@
               size = 1) +
+
+  # Theme
   theme_light() +
@@
   labs(col = "", linetype = "Test") +
-  facet_wrap(~conv_or_not) +
-  theme(legend.position="bottom")
-```
-
-```{r different-x-dist-heter-power, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model under different distributions of regressor."}
-
-dat_different_dist <- heter_conv_sim %>% 
-  mutate(log_effect_size = log(effect_size)) %>%
-  select(x_dist, log_effect_size, BP_p_value) %>%
-  pivot_longer(BP_p_value) %>%
-  mutate(x_dist = ifelse(x_dist == "even_discrete", 
-                                       "Discrete", 
-                                       stringr::str_to_title(x_dist))) %>%
-  mutate(name = gsub("_p_value", " test", name)) %>%
-  mutate(reject = value < 0.05) %>%
-  mutate(conv_or_not = "Conventional") %>%
-  bind_rows(vi_survey %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type == "heteroskedasticity") %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
-  mutate(reject = as.numeric(p_value <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  mutate(x_dist = ifelse(x_dist == "even_discrete", 
-                         "Discrete", 
-                         stringr::str_to_title(x_dist))) %>%
-  mutate(conv_or_not = "Visual")) %>%
-  mutate(x_dist = factor(x_dist, 
-      levels = c("Uniform", "Normal", 
-                 "Lognormal", "Discrete"))) %>%
-  select(x_dist, log_effect_size, reject, conv_or_not)
-
-dat_different_dist %>%
-  ggplot() +
-  geom_smooth(data = ~filter(.x, x_dist == "Uniform"),
-              aes(log_effect_size, 
-                  reject), 
-                  col = "grey", 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, size = 3) +
-  geom_smooth(aes(log_effect_size, 
-                  reject, 
-                  col = x_dist), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              size = 1) +
-  theme_light() +
-  scale_color_manual(values = rcartocolor::carto_pal(4, "Safe")[c(4,1,3,2)]) +
-  scale_size_manual(values = c(1, 2, 3, 4)) +
-  xlab(expression(log[e] (Effect_size))) +
-  ylab("Power") +
-  labs(col = "", linetype = "Test") +
-  facet_wrap(~conv_or_not) +
-  theme(legend.position="bottom")
+  facet_grid(type ~ conv_or_not) +
+  theme(legend.position = "bottom")
 ```
@@
 
-Shape of the non-linearity pattern is another factor contributed to the simulation model.
+Shape of the non-linearity pattern is another factor contributed to the simulation model. Figure \ref{fig:poly-power-j}
 
@@
   scale_color_manual(values = rcartocolor::carto_pal(3, "Vivid")) +
+  theme_light() +
   theme(legend.position = "bottom") +
-  theme_light() +
   facet_wrap(~conv_or_not) +
@@
 
-# Appendix
-
-## Effect derivation
-
-## $p$-value scatter plot
-
-```{r p-value-comparison-old, eval = FALSE, fig.height = 4, fig.cap = "Visual test p-value compared to conventional test p-value for lineups produced by heteroskedasticity and non-linearity model.
-The scatter plot is drawn on a square root scale.
-Every dot on the plot corresponds to a lineup.
-Most of the dots are fallen on the left of the y = x line indicating visual test p-value is almost always higher than conventional test p-value.
-The blue curve is a smoothing of the dots showing a positive trend.
-"}
-vi_survey %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
-  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-  left_join(vi_survey %>% 
-           count(unique_lineup_id) %>%
-           rename(num_subject = n)) %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
-  ggplot() +
-  geom_abline() +
-  geom_point(aes(conventional_p_value, p_value), alpha = 0.4) +
-  geom_smooth(aes(conventional_p_value, p_value), se = FALSE, span = 0.6) +
-  scale_x_sqrt() +
-  scale_y_sqrt() +
-  facet_wrap(~type) +
-  ylab("Visual test p-value on square root scale") +
-  xlab("Conventional test p-value on square root scale") +
-  theme_light()
-```
-
-## A collection interesting lineups (unusual results)
-
-why unusual? what is the possible explanations?
-
-# targe journal
-
-JRSSB: Journal of the Royal Statistical Society Series B (Statistical Methodology)
-Deadline: Jan 1, Apr 1
-
-Reading: style of writing, author guideline (https://rss.onlinelibrary.wiley.com/hub/journal/14679868/author-guidelines)
-
-JCGS
-JDSS
-
 \newpage

### 2023-01-23 paper_comparison/paper_comparison/paper_comparison.tex
combine two different x_dist plots into one

@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-x-dist-heter-power-1}
-
-}
-
-\caption{Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model under different distributions of regressor.}\label{fig:different-x-dist-heter-power}
-\end{figure}
-
 \hypertarget{shape-of-non-linearity}{%
@@
 Shape of the non-linearity pattern is another factor contributed to the
-simulation model.
+simulation model. Figure \ref{fig:poly-power-j}
 
@@
 
-\hypertarget{appendix}{%
-\section{Appendix}\label{appendix}}
-
-\hypertarget{effect-derivation}{%
-\subsection{Effect derivation}\label{effect-derivation}}
-
-\hypertarget{p-value-scatter-plot}{%
-\subsection{\texorpdfstring{\(p\)-value scatter
-plot}{p-value scatter plot}}\label{p-value-scatter-plot}}
-
-\hypertarget{a-collection-interesting-lineups-unusual-results}{%
-\subsection{A collection interesting lineups (unusual
-results)}\label{a-collection-interesting-lineups-unusual-results}}
-
-why unusual? what is the possible explanations?
-
-\hypertarget{targe-journal}{%
-\section{targe journal}\label{targe-journal}}
-
-JRSSB: Journal of the Royal Statistical Society Series B (Statistical
-Methodology) Deadline: Jan 1, Apr 1
-
-Reading: style of writing, author guideline
-(\url{https://rss.onlinelibrary.wiley.com/hub/journal/14679868/author-guidelines})
-
-JCGS JDSS
-
 \newpage

### 2023-01-23 paper_comparison/paper_comparison/paper_comparison.Rmd
provide more details in x_dist caption

@@
 
-Figure \ref{fig:different-x-dist-poly-power} and Figure \ref{fig:different-x-dist-heter-power} illustrate the change of power of visual tests and conventional tests for different fitted distributions used in the non-linearity model and the heteroskedasticity model.
+Figure \ref{fig:different-x-dist-poly-power} illustrates the change of power of visual tests and conventional tests for different fitted distributions used in the non-linearity model and the heteroskedasticity model.
 We focus on $log(E) > 2$ where the visual patterns start to become recognisable.
@@
 
-```{r different-x-dist-poly-power, fig.cap = "Comparison of power of conventional tests and visual tests on lineups produced by the non-linearity model under different fitted value distributions."}
+```{r different-x-dist-poly-power, fig.cap = "Comparison of power of conventional tests and visual tests on lineups with different fitted value distributions.
+Each power curve of the uniform distribution is surrounded by a grey line to indicate it is the baseline in its group.
+Conventional tests are more sensitive than visual tests in regards of the type of the simulated model and the fitted value distribution.
+For non-linearity patterns, visual tests show lower power with discrete uniform distribution, possibly due to the difficulty in reading discreteness.
+For heteroskedasticity patterns, because of the unevenness, visual tests have lower power with normal and lognormal distribution."}
 

### 2023-01-23 paper_comparison/paper_comparison/paper_comparison.tex
provide more details in x_dist caption

@@
 
-Figure \ref{fig:different-x-dist-poly-power} and Figure
-\ref{fig:different-x-dist-heter-power} illustrate the change of power of
-visual tests and conventional tests for different fitted distributions
-used in the non-linearity model and the heteroskedasticity model. We
-focus on \(log(E) > 2\) where the visual patterns start to become
-recognisable.
+Figure \ref{fig:different-x-dist-poly-power} illustrates the change of
+power of visual tests and conventional tests for different fitted
+distributions used in the non-linearity model and the heteroskedasticity
+model. We focus on \(log(E) > 2\) where the visual patterns start to
+become recognisable.
 
@@
 
-\caption{Comparison of power of conventional tests and visual tests on lineups produced by the non-linearity model under different fitted value distributions.}\label{fig:different-x-dist-poly-power}
+\caption{Comparison of power of conventional tests and visual tests on lineups with different fitted value distributions.
+Each power curve of the uniform distribution is surrounded by a grey line to indicate it is the baseline in its group.
+Conventional tests are more sensitive than visual tests in regards of the type of the simulated model and the fitted value distribution.
+For non-linearity patterns, visual tests show lower power with discrete uniform distribution, possibly due to the difficulty in reading discreteness.
+For heteroskedasticity patterns, because of the unevenness, visual tests have lower power with normal and lognormal distribution.}\label{fig:different-x-dist-poly-power}
 \end{figure}

### 2023-01-23 paper_comparison/paper_comparison/paper_comparison.Rmd
fill in the different poly shape power section

@@
 
-<!-- Figure \ref{fig:poly-example} shows one of the ten lineups where the visual test rejected the residual plot.
-The data plot, in position three, is very obvious, and different from the nulls.
-The visual test was correct to reject.
-To understand why this is, one needs to return to the way the RESET test is applied.
-It requires a parameter indicating degree of polynomial to test for, and the recommendation is to generically use 4 (XXX REF).
-However, the "M" and the "Triple-U" shapes constructed from the Hermite polynomials use power up to 6 and 18 respectively.
-All 10 lineups rejected by the visual test but not the RESET were of types "M" or "Triple-U".
-If the RESET test had been applied using the power up to 6, all the decisions for these lineups would have been to reject.
-The recommendation of the polynomial power for the RESET should be revised, perhaps.
-This illustrates the sensitivity of conventional testing to the parameters, and it also points to a limitation that one needs to know the data structure in order to set the parameters for the test.
--->
-
-```{r poly-example, eval = FALSE, fig.height = 7, fig.width = 7, fig.cap = "Example of a non-linearity that is rejected by the visual test but not by the RESET test.
-The data plot at panel three exhibits strong non-linear pattern.
-It has effect size $= 3.89$.
-Interestingly, all the lineups in this group are higher order polynomials (M, Triple-U), and it is apparent that the default RESET test under-performs.
-The RESET test using the higher order polynomial parameter could be used, but this would change it's performance for lower order polynomials.
-The only way for it to be sensitive to these type of departure is to know the order prior to testing!"}
-# plot poly-270 
-VI_MODEL$plot_lineup(vi_lineup$poly_270$data, 
-                     remove_grid_line = TRUE, 
-                     theme = theme_light(),
-                     remove_axis = TRUE)
-```
-
 For lineups containing non-linearity patterns, conventional tests reject `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`%
@@
 
-Shape of the non-linearity pattern is another factor contributed to the simulation model. Figure \ref{fig:poly-power-j}
-
-```{r poly-power-j, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups produced by the non-linearity model with four different shapes controlled by the parameter $j$."}
+Shape of the non-linearity pattern is another factor contributed to the non-linearity model.
+According to Figure \ref{fig:poly-power-j}, conventional tests are significantly less powerful in testing "Triple-U" shape compared to other three shapes, though it is still more sensitive than the visual test.
+In fact, among the lineups containing non-linearity patterns used in our experiments, around `r vi_survey %>% filter(!attention_check) %>% filter(!null_lineup) %>% filter(type == "polynomial") %>% group_by(unique_lineup_id) %>% summarise(across(everything(), first)) %>% count(a = p_value < 0.05, b = conventional_p_value < 0.05) %>% mutate(prop = n / sum(n) * 100) %>% filter(a, !b) %>% pull(prop) %>% round(digits = 1)`% (two our of 576) are rejected by the visual test, but accepted by the conventional test.
+
+Figure \ref{fig:poly-example} shows one of the two lineups where the visual test rejects the residual plot.
+The data plot, in position 18, is very obvious, and different from the nulls.
+The visual test is correct to reject.
+To understand why this is, one needs to return to the way the RESET test is applied.
+It requires a parameter indicating degree of fitted values to test for, and the recommendation is to generically use four (@ramsey_tests_1969).
+However, the "Triple-U" shape constructed from the Hermite polynomials use power up to 18.
+Both lineups rejected by the visual test but not the RESET are of "Triple-U".
+If the RESET test had been applied using the power up to six, all the decisions for these lineups would have been to reject.
+The recommendation of the polynomial power for the RESET should be revised, perhaps.
+This illustrates the sensitivity of conventional testing to the parameters, and it also points to a limitation that one needs to know the data structure in order to set the parameters for the test.
+
+We expect the "U" shape will be the easiest one to be detected by subjects followed by the "S", "M" and "Triple-U" shape.
+From Figure \ref{fig:poly-power-j}, it can be observed that the power curves are mostly aligned with our expectation, except for the "M" shape.
+Interestingly, the "M" shape is as easy to be detected as the "U" shape.
+The readability of the shape do not strictly follow the degree of the polynomials.
+
+```{r poly-power-j, fig.height = 4, fig.cap = "Comparison of power of conventional tests and visual tests on lineups containing four different non-linearity shapes. "}
 
@@
   mutate(conv_or_not = "Visual")) %>%
-  mutate(shape = factor(shape, levels = c("S", "U", "M", "Triple-U"))) %>%
+  mutate(shape = factor(shape, levels = c("U", "S", "M", "Triple-U"))) %>%
   select(shape, log_effect_size, reject, conv_or_not) %>%
@@
   labs(col = "Non-linearity shape")
+```
+
+```{r poly-example, fig.height = 7, fig.width = 7, fig.cap = "Example of a non-linearity that is rejected by the visual test but not by the RESET test.
+The data plot at panel 18 exhibits strong non-linear pattern.
+It has effect size $= 3.96$.
+Interestingly, all the lineups in this group are higher order polynomials (Triple-U), and it is apparent that the default RESET test under-performs.
+The RESET test using the higher order polynomial parameter could be used, but this would change it's performance for lower order polynomials.
+The only way for it to be sensitive to these type of departure is to know the order prior to testing!"}
+# plot poly-275
+VI_MODEL$plot_lineup(vi_lineup$poly_275$data, 
+                     remove_grid_line = TRUE, 
+                     theme = theme_light(),
+                     remove_axis = TRUE)
 ```

### 2023-01-23 paper_comparison/paper_comparison/paper_comparison.tex
fill in the different poly shape power section

@@
 Shape of the non-linearity pattern is another factor contributed to the
-simulation model. Figure \ref{fig:poly-power-j}
+non-linearity model. According to Figure \ref{fig:poly-power-j},
+conventional tests are significantly less powerful in testing
+``Triple-U'' shape compared to other three shapes, though it is still
+more sensitive than the visual test. In fact, among the lineups
+containing non-linearity patterns used in our experiments, around 0.3\%
+(two our of 576) are rejected by the visual test, but accepted by the
+conventional test.
+
+Figure \ref{fig:poly-example} shows one of the two lineups where the
+visual test rejects the residual plot. The data plot, in position 18, is
+very obvious, and different from the nulls. The visual test is correct
+to reject. To understand why this is, one needs to return to the way the
+RESET test is applied. It requires a parameter indicating degree of
+fitted values to test for, and the recommendation is to generically use
+four (\citet{ramsey_tests_1969}). However, the ``Triple-U'' shape
+constructed from the Hermite polynomials use power up to 18. Both
+lineups rejected by the visual test but not the RESET are of
+``Triple-U''. If the RESET test had been applied using the power up to
+six, all the decisions for these lineups would have been to reject. The
+recommendation of the polynomial power for the RESET should be revised,
+perhaps. This illustrates the sensitivity of conventional testing to the
+parameters, and it also points to a limitation that one needs to know
+the data structure in order to set the parameters for the test.
+
+We expect the ``U'' shape will be the easiest one to be detected by
+subjects followed by the ``S'', ``M'' and ``Triple-U'' shape. From
+Figure \ref{fig:poly-power-j}, it can be observed that the power curves
+are mostly aligned with our expectation, except for the ``M'' shape.
+Interestingly, the ``M'' shape is as easy to be detected as the ``U''
+shape. The readability of the shape do not strictly follow the degree of
+the polynomials.
 
@@
 
-\caption{Power of conventional tests and visual tests on lineups produced by the non-linearity model with four different shapes controlled by the parameter $j$.}\label{fig:poly-power-j}
+\caption{Comparison of power of conventional tests and visual tests on lineups containing four different non-linearity shapes. }\label{fig:poly-power-j}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/poly-example-1}
+
+}
+
+\caption{Example of a non-linearity that is rejected by the visual test but not by the RESET test.
+The data plot at panel 18 exhibits strong non-linear pattern.
+It has effect size $= 3.96$.
+Interestingly, all the lineups in this group are higher order polynomials (Triple-U), and it is apparent that the default RESET test under-performs.
+The RESET test using the higher order polynomial parameter could be used, but this would change it's performance for lower order polynomials.
+The only way for it to be sensitive to these type of departure is to know the order prior to testing!}\label{fig:poly-example}
 \end{figure}

### 2023-01-24 paper_comparison/paper_comparison/paper_comparison.Rmd
Add function to generate boot CI. Update plots to use this function.

@@
 ```{r}
-# Bootstrap the dataset
-boot_y <- function(dat, y, times = 100) {
-  map_df(1:times, function(i) slice_sample(dat, n = nrow(dat), replace = TRUE) %>% mutate(boot_id = i))
+# Get bootstrap CI
+get_boot_conf <- function(dat, 
+                          .x,
+                          .y,
+                          times = 100, 
+                          upper_q = 0.975, 
+                          lower_q = 0.025) {
+
+  # Construct the model formula
+  mod_formula <- substitute(.y ~ .x)
+
+  # Get the x variable name
+  x_chr <- as.character(substitute(.x))
+
+  # Prepare a sequence of x values
+  x_seq <- seq(min(dat[[x_chr]]), max(dat[[x_chr]]), by = 0.01)
+
+  # Properly construct the new data frame for prediction
+  newdata <- data.frame(.x = x_seq)
+  newdata[[x_chr]] <- x_seq
+
+  # Get the prediction for each bootstrap sample   
+  dat_pred <- map_df(1:times, function(i) {
+
+    # Get a bootstrap sample
+    dat_i <- slice_sample(dat, n = nrow(dat), replace = TRUE) %>%
+      mutate(boot_id = i)
+
+    # Fit the model on the sample
+    mod_i <- glm(mod_formula, family = binomial, data = dat_i)
+
+    # Get the prediction
+    newdata$.pred <- predict(mod_i, newdata = newdata, type = "response")
+
+    # Return the prediction
+    newdata
+  })
+
+  dat_pred <- dat_pred %>%
+    group_by(.x) %>%
+    summarise(upper = quantile(.pred, upper_q), 
+              lower = quantile(.pred, lower_q))
+  dat_pred[[x_chr]] <- dat_pred$.x
+  dat_pred
 }
@@
 
-  # Draw bootstrap visual test power curves
-  stat_smooth(geom = "line", 
-              data = ~boot_y(.x, reject, times = 500), 
-              aes(log_effect_size, 
+  # Draw bootstrap CI for visual test
+  geom_ribbon(data = ~get_boot_conf(.x,
+                                    log_effect_size,
                                     reject,
-                  group = boot_id, 
-                  col = "Visual test"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE, 
-              alpha = 0.03) +
+                                    times = 500),
+              aes(x = log_effect_size, 
+                  ymin = lower, 
+                  ymax = upper),
+              fill = rev(rcartocolor::carto_pal(4, "Vivid"))[4],
+              alpha = 0.2) +
 
@@
 Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups.
-Surrounding curves of the visual test show 500 samples of bootstrap estimation.
+Surrounding band of the visual test shows the 95% confidence interval estimated using 500 bootstrap samples.
 Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
@@
 Main plot shows the power curves, with dots indicating human evaluations of lineups.
-Surrounding curves of the visual test show 500 samples of bootstrap estimation.
+Surrounding band of the visual test shows the 95% confidence interval estimated using 500 bootstrap samples.
 Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
@@
 This suggests the use of transformation of fitted values in reading residual plots.
-In the case of conventional tests, the discrete uniform distribution consistently doing great, followed by the uniform distribution, the normal distribution and the lognormal distribution.
+In the case of conventional tests, the discrete uniform distribution consistently perform well, followed by the uniform distribution, the normal distribution and the lognormal distribution.
 
@@
 
-```{r different-x-dist-poly-power, fig.cap = "Comparison of power of conventional tests and visual tests on lineups with different fitted value distributions.
-Each power curve of the uniform distribution is surrounded by a grey line to indicate it is the baseline in its group.
-Conventional tests are more sensitive than visual tests in regards of the type of the simulated model and the fitted value distribution.
-For non-linearity patterns, visual tests show lower power with discrete uniform distribution, possibly due to the difficulty in reading discreteness.
-For heteroskedasticity patterns, because of the unevenness, visual tests have lower power with normal and lognormal distribution."}
+```{r different-x-dist-poly-power, fig.cap = "Power of different tests on lineups with different fitted value distributions.
+Each power curve of the uniform distribution is surrounded by a grey line to indicate it is the baseline within its group.
+Conventional tests are more sensitive than visual tests in regards of the type of the visual pattern and the fitted value distribution.
+Visual tests exhibit relatively low power in testing non-linearity patterns under discrete uniform distribution, possibly due to the difficulty in reading discreteness.
+Meanwhile, uneveness affects heteroskedasticity patterns greatly, result in low power of visual tests under normal and lognormal distribution."}
 
@@
 
-```{r poly-power-j, fig.height = 4, fig.cap = "Comparison of power of conventional tests and visual tests on lineups containing four different non-linearity shapes. "}
+```{r poly-power-j, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups containing four different non-linearity shapes. "}
 
@@
 
+We also investigate the impact of the shape of heteroskedasticity on the power of different tests.
+The corresponding result is presented in Figure \ref{fig:heter-power-a}.
+In theory, the "Left-triangle" and the "Right-triangle" should have no difference from the point of view of a Breusch-Pagan test.
+According to Figure \ref{fig:heter-power-a}, this is indeed the case.
+And these two shapes should have the same visual test power, unless the orientation of the visual pattern is a non-negligible factor from human perspective.
+However, the same graph suggests that the "Right-triangle" shape is more difficult to be noticed by subjects than the "Left-triangle" shape.
+
 ```{r heter-power-a, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$."}
@@
               method.args = list(family = binomial())) +
+
+  # Draw bootstrap for Left-triangle
+  geom_ribbon(data = ~get_boot_conf(filter(.x, 
+                                           conv_or_not == "Visual", 
+                                           a == "Left-triangle"),
+                                    log_effect_size,
+                                    reject,
+                                    times = 1000) %>%
+                mutate(conv_or_not = "Visual"),
+              aes(x = log_effect_size, 
+                  ymin = lower, 
+                  ymax = upper),
+              fill = rev(rcartocolor::carto_pal(3, "Vivid"))[3],
+              alpha = 0.2) +
+
+  # Draw bootstrap for Right-triangle
+  geom_ribbon(data = ~get_boot_conf(filter(.x, 
+                                           conv_or_not == "Visual", 
+                                           a == "Right-triangle"),
+                                    log_effect_size,
+                                    reject,
+                                    times = 1000) %>%
+                mutate(conv_or_not = "Visual"),
+              aes(x = log_effect_size, 
+                  ymin = lower, 
+                  ymax = upper),
+              fill = rev(rcartocolor::carto_pal(3, "Vivid"))[2],
+              alpha = 0.2) +
+
+  # Theme
   xlab(expression(log[e] (Effect_size))) +

### 2023-01-24 paper_comparison/paper_comparison/paper_comparison.tex
Add function to generate boot CI. Update plots to use this function.

@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/polypower-1}
-
-}
-
-\caption{Comparison of power between different tests for non-linear patterns (uniform fitted values only).
-Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups.
-Surrounding curves of the visual test show 500 samples of bootstrap estimation.
-Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
-Where would you draw the line of too much non-linearity in the residuals?
-For the RESET test this is around log effect size 1.5, but for the visual test it is around 3.5.}\label{fig:polypower}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/heterpower-1}
-
-}
-
-\caption{Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only).
-Main plot shows the power curves, with dots indicating human evaluations of lineups.
-Surrounding curves of the visual test show 500 samples of bootstrap estimation.
-Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
-Where would you draw the line of too much heteroskedasticity in the residuals?
-For the BP test this is around log effect size 1.5, but for the visual test it is around 3.}\label{fig:heterpower}
-\end{figure}
+\textbackslash begin\{figure\}
+
+\{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/polypower-1}
+
+\}
+
+\textbackslash caption\{Comparison of power between different tests for
+non-linear patterns (uniform fitted values only). Main plot shows the
+power curves estimated using logistic regression, with dots indicating
+human evaluations of lineups. Surrounding band of the visual test shows
+the 95\% confidence interval estimated using 500 bootstrap samples.
+Small row of plots shows typical residual plots corresponding to
+specific effect sizes, marked by dashed lines in main plot. Where would
+you draw the line of too much non-linearity in the residuals? For the
+RESET test this is around log effect size 1.5, but for the visual test
+it is around 3.5.\}\label{fig:polypower} \textbackslash end\{figure\}
+
+\textbackslash begin\{figure\}
+
+\{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/heterpower-1}
+
+\}
+
+\textbackslash caption\{Comparison of power between different tests for
+heteroskedasticity patterns (uniform fitted values only). Main plot
+shows the power curves, with dots indicating human evaluations of
+lineups. Surrounding band of the visual test shows the 95\% confidence
+interval estimated using 500 bootstrap samples. Small row of plots shows
+typical residual plots corresponding to specific effect sizes, marked by
+dashed lines in main plot. Where would you draw the line of too much
+heteroskedasticity in the residuals? For the BP test this is around log
+effect size 1.5, but for the visual test it is around
+3.\}\label{fig:heterpower} \textbackslash end\{figure\}
 
@@
 of conventional tests, the discrete uniform distribution consistently
-doing great, followed by the uniform distribution, the normal
+perform well, followed by the uniform distribution, the normal
 distribution and the lognormal distribution.
@@
 
-\caption{Comparison of power of conventional tests and visual tests on lineups with different fitted value distributions.
-Each power curve of the uniform distribution is surrounded by a grey line to indicate it is the baseline in its group.
-Conventional tests are more sensitive than visual tests in regards of the type of the simulated model and the fitted value distribution.
-For non-linearity patterns, visual tests show lower power with discrete uniform distribution, possibly due to the difficulty in reading discreteness.
-For heteroskedasticity patterns, because of the unevenness, visual tests have lower power with normal and lognormal distribution.}\label{fig:different-x-dist-poly-power}
+\caption{Power of different tests on lineups with different fitted value distributions.
+Each power curve of the uniform distribution is surrounded by a grey line to indicate it is the baseline within its group.
+Conventional tests are more sensitive than visual tests in regards of the type of the visual pattern and the fitted value distribution.
+Visual tests exhibit relatively low power in testing non-linearity patterns under discrete uniform distribution, possibly due to the difficulty in reading discreteness.
+Meanwhile, uneveness affects heteroskedasticity patterns greatly, result in low power of visual tests under normal and lognormal distribution.}\label{fig:different-x-dist-poly-power}
 \end{figure}
@@
 
-\caption{Comparison of power of conventional tests and visual tests on lineups containing four different non-linearity shapes. }\label{fig:poly-power-j}
+\caption{Power of conventional tests and visual tests on lineups containing four different non-linearity shapes. }\label{fig:poly-power-j}
 \end{figure}
@@
 
+We also investigate the impact of the shape of heteroskedasticity on the
+power of different tests. The corresponding result is presented in
+Figure \ref{fig:heter-power-a}. In theory, the ``Left-triangle'' and the
+``Right-triangle'' should have no difference from the point of view of a
+Breusch-Pagan test. According to Figure \ref{fig:heter-power-a}, this is
+indeed the case. And these two shapes should have the same visual test
+power, unless the orientation of the visual pattern is a non-negligible
+factor from human perspective. However, the same graph suggests that the
+``Right-triangle'' shape is more difficult to be noticed by subjects
+than the ``Left-triangle'' shape.
+
 \begin{figure}

### 2023-01-24 paper_comparison/paper_comparison/paper_comparison.Rmd
fill in the last section of the result

@@
                                     reject,
-                                    times = 500),
+                                    times = 1000),
               aes(x = log_effect_size, 
@@
 Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups.
-Surrounding band of the visual test shows the 95% confidence interval estimated using 500 bootstrap samples.
+Surrounding band of the visual test shows the 95 percent confidence interval estimated using 1000 bootstrap samples.
 Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
@@
 Main plot shows the power curves, with dots indicating human evaluations of lineups.
-Surrounding band of the visual test shows the 95% confidence interval estimated using 500 bootstrap samples.
+Surrounding band of the visual test shows the 95 percent confidence interval estimated using 500 bootstrap samples.
 Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
@@
 The corresponding result is presented in Figure \ref{fig:heter-power-a}.
-In theory, the "Left-triangle" and the "Right-triangle" should have no difference from the point of view of a Breusch-Pagan test.
+In theory, the "Left-triangle" and the "Right-triangle" should have little difference from the point of view of a Breusch-Pagan test.
 According to Figure \ref{fig:heter-power-a}, this is indeed the case.
 And these two shapes should have the same visual test power, unless the orientation of the visual pattern is a non-negligible factor from human perspective.
-However, the same graph suggests that the "Right-triangle" shape is more difficult to be noticed by subjects than the "Left-triangle" shape.
+However, by checking the bootstrap confidence intervals, it can be observed from Figure \ref{fig:heter-power-a} that when $3 < log(E) < 4$, the "Left-triangle" power curve is significantly greater than the "Right-triangle" power curve.
+This phenomenon is particularly interesting.
+It implies the current residual plot design is not robust to the orientation of the visual pattern, which is worth to be explored in future research.
 
@@
 
+# Conclusion
+
 \newpage

### 2023-01-24 paper_comparison/paper_comparison/paper_comparison.tex
fill in the last section of the result

@@
 
-\textbackslash begin\{figure\}
-
-\{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/polypower-1}
-
-\}
-
-\textbackslash caption\{Comparison of power between different tests for
-non-linear patterns (uniform fitted values only). Main plot shows the
-power curves estimated using logistic regression, with dots indicating
-human evaluations of lineups. Surrounding band of the visual test shows
-the 95\% confidence interval estimated using 500 bootstrap samples.
-Small row of plots shows typical residual plots corresponding to
-specific effect sizes, marked by dashed lines in main plot. Where would
-you draw the line of too much non-linearity in the residuals? For the
-RESET test this is around log effect size 1.5, but for the visual test
-it is around 3.5.\}\label{fig:polypower} \textbackslash end\{figure\}
-
-\textbackslash begin\{figure\}
-
-\{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/heterpower-1}
-
-\}
-
-\textbackslash caption\{Comparison of power between different tests for
-heteroskedasticity patterns (uniform fitted values only). Main plot
-shows the power curves, with dots indicating human evaluations of
-lineups. Surrounding band of the visual test shows the 95\% confidence
-interval estimated using 500 bootstrap samples. Small row of plots shows
-typical residual plots corresponding to specific effect sizes, marked by
-dashed lines in main plot. Where would you draw the line of too much
-heteroskedasticity in the residuals? For the BP test this is around log
-effect size 1.5, but for the visual test it is around
-3.\}\label{fig:heterpower} \textbackslash end\{figure\}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/polypower-1}
+
+}
+
+\caption{Comparison of power between different tests for non-linear patterns (uniform fitted values only).
+Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups.
+Surrounding band of the visual test shows the 95 percent confidence interval estimated using 1000 bootstrap samples.
+Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
+Where would you draw the line of too much non-linearity in the residuals?
+For the RESET test this is around log effect size 1.5, but for the visual test it is around 3.5.}\label{fig:polypower}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/heterpower-1}
+
+}
+
+\caption{Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only).
+Main plot shows the power curves, with dots indicating human evaluations of lineups.
+Surrounding band of the visual test shows the 95 percent confidence interval estimated using 500 bootstrap samples.
+Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
+Where would you draw the line of too much heteroskedasticity in the residuals?
+For the BP test this is around log effect size 1.5, but for the visual test it is around 3.}\label{fig:heterpower}
+\end{figure}
 
@@
 Figure \ref{fig:heter-power-a}. In theory, the ``Left-triangle'' and the
-``Right-triangle'' should have no difference from the point of view of a
-Breusch-Pagan test. According to Figure \ref{fig:heter-power-a}, this is
-indeed the case. And these two shapes should have the same visual test
-power, unless the orientation of the visual pattern is a non-negligible
-factor from human perspective. However, the same graph suggests that the
-``Right-triangle'' shape is more difficult to be noticed by subjects
-than the ``Left-triangle'' shape.
+``Right-triangle'' should have little difference from the point of view
+of a Breusch-Pagan test. According to Figure \ref{fig:heter-power-a},
+this is indeed the case. And these two shapes should have the same
+visual test power, unless the orientation of the visual pattern is a
+non-negligible factor from human perspective. However, by checking the
+bootstrap confidence intervals, it can be observed from Figure
+\ref{fig:heter-power-a} that when \(3 < log(E) < 4\), the
+``Left-triangle'' power curve is significantly greater than the
+``Right-triangle'' power curve. This phenomenon is particularly
+interesting. It implies the current residual plot design is not robust
+to the orientation of the visual pattern, which is worth to be explored
+in future research.
 
@@
 \end{figure}
+
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
 

### 2023-01-24 paper_comparison/paper_comparison/paper_comparison.Rmd
update color pal for the last section of the result

@@
 
-```{r poly-power-j, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups containing four different non-linearity shapes. "}
+```{r poly-power-j, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups containing four different non-linearity shapes.
+Power curves with higher order of non-linearity are drawn with deeper colours."}
 
@@
   ylab("Power") +
-  scale_color_manual(values = rcartocolor::carto_pal(4, "Vivid")) +
+  scale_color_manual(values = rcartocolor::carto_pal(4, "TealGrn")) +
   theme_light() +
@@
                   ymax = upper),
-              fill = rev(rcartocolor::carto_pal(3, "Vivid"))[3],
+              fill = rcartocolor::carto_pal(3, "Earth")[1],
               alpha = 0.2) +
@@
                   ymax = upper),
-              fill = rev(rcartocolor::carto_pal(3, "Vivid"))[2],
+              fill = rcartocolor::carto_pal(3, "Earth")[3],
               alpha = 0.2) +
@@
   ylab("Power") +
-  scale_color_manual(values = rcartocolor::carto_pal(3, "Vivid")) +
+  scale_color_manual(values = rcartocolor::carto_pal(3, "Earth")) +
   theme_light() +
@@
 
+1. summarise paper content
+2. briefly describe what was learned
+3. how does this apply more broadly than simple regression
+4. where does this lead? Computer vision residual plot reading
+
+# Acknowledgement
+
+1. software tools
+2. point to github repo (create a new one)
+3. summarise the supplementary material
+
 \newpage

### 2023-01-24 paper_comparison/paper_comparison/paper_comparison.Rmd
small changes

--- 2023-01-24 paper_comparison/paper_comparison/paper_comparison.Rmd
@@
 ```{r}
-# Get bootstrap CI
-get_boot_conf <- function(dat, 
-                          .x,
-                          .y,
-                          times = 100, 
-                          upper_q = 0.975, 
-                          lower_q = 0.025) {
-
-  # Construct the model formula
-  mod_formula <- substitute(.y ~ .x)
-
-  # Get the x variable name
-  x_chr <- as.character(substitute(.x))
-
-  # Prepare a sequence of x values
-  x_seq <- seq(min(dat[[x_chr]]), max(dat[[x_chr]]), by = 0.01)
-
-  # Properly construct the new data frame for prediction
-  newdata <- data.frame(.x = x_seq)
-  newdata[[x_chr]] <- x_seq
-
-  # Get the prediction for each bootstrap sample   
-  dat_pred <- map_df(1:times, function(i) {
-
-    # Get a bootstrap sample
-    dat_i <- slice_sample(dat, n = nrow(dat), replace = TRUE) %>%
-      mutate(boot_id = i)
-
-    # Fit the model on the sample
-    mod_i <- glm(mod_formula, family = binomial, data = dat_i)
-
-    # Get the prediction
-    newdata$.pred <- predict(mod_i, newdata = newdata, type = "response")
-
-    # Return the prediction
-    newdata
-  })
-
-  dat_pred <- dat_pred %>%
-    group_by(.x) %>%
-    summarise(upper = quantile(.pred, upper_q), 
-              lower = quantile(.pred, lower_q))
-  dat_pred[[x_chr]] <- dat_pred$.x
-  dat_pred
+# Bootstrap the dataset
+boot_y <- function(dat, y, times = 100) {
+  map_df(1:times, function(i) slice_sample(dat, n = nrow(dat), replace = TRUE) %>% mutate(boot_id = i))
 }
@@
 
-  # Draw bootstrap CI for visual test
-  geom_ribbon(data = ~get_boot_conf(.x,
-                                    log_effect_size,
+  # Draw bootstrap visual test power curves
+  stat_smooth(geom = "line", 
+              data = ~boot_y(.x, reject, times = 500), 
+              aes(log_effect_size, 
                   reject, 
-                                    times = 500),
-              aes(x = log_effect_size, 
-                  ymin = lower, 
-                  ymax = upper),
-              fill = rev(rcartocolor::carto_pal(4, "Vivid"))[4],
-              alpha = 0.2) +
+                  group = boot_id, 
+                  col = "Visual test"), 
+              method = "glm", 
+              method.args = list(family = binomial), 
+              se = FALSE, 
+              alpha = 0.03) +
 
@@
 Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups.
-Surrounding band of the visual test shows the 95% confidence interval estimated using 500 bootstrap samples.
+Surrounding curves of the visual test show 500 samples of bootstrap estimation.
 Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
@@
 Main plot shows the power curves, with dots indicating human evaluations of lineups.
-Surrounding band of the visual test shows the 95% confidence interval estimated using 500 bootstrap samples.
+Surrounding curves of the visual test show 500 samples of bootstrap estimation.
 Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
@@
 This suggests the use of transformation of fitted values in reading residual plots.
-In the case of conventional tests, the discrete uniform distribution consistently perform well, followed by the uniform distribution, the normal distribution and the lognormal distribution.
+In the case of conventional tests, the discrete uniform distribution consistently doing great, followed by the uniform distribution, the normal distribution and the lognormal distribution.
 
@@
 
-```{r different-x-dist-poly-power, fig.cap = "Power of different tests on lineups with different fitted value distributions.
-Each power curve of the uniform distribution is surrounded by a grey line to indicate it is the baseline within its group.
-Conventional tests are more sensitive than visual tests in regards of the type of the visual pattern and the fitted value distribution.
-Visual tests exhibit relatively low power in testing non-linearity patterns under discrete uniform distribution, possibly due to the difficulty in reading discreteness.
-Meanwhile, uneveness affects heteroskedasticity patterns greatly, result in low power of visual tests under normal and lognormal distribution."}
+```{r different-x-dist-poly-power, fig.cap = "Comparison of power on lineups with different fitted value distributions, for conventional tests and visual tests.
+Each power curve of the uniform distribution is surrounded by a grey line to indicate it is the baseline in its group.
+For both types of tests power does change.
+The discrete fitted value distribution substantially lowers the power for non-linearity patterns in residual plots.
+For heteroskedasticity patterns, normal and lognormal distributions produce lower power."}
 
@@
 
-```{r poly-power-j, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups containing four different non-linearity shapes. "}
+```{r poly-power-j, fig.height = 4, fig.cap = "Comparison of power of conventional tests and visual tests on lineups containing four different non-linearity shapes. "}
 
@@
 
-We also investigate the impact of the shape of heteroskedasticity on the power of different tests.
-The corresponding result is presented in Figure \ref{fig:heter-power-a}.
-In theory, the "Left-triangle" and the "Right-triangle" should have no difference from the point of view of a Breusch-Pagan test.
-According to Figure \ref{fig:heter-power-a}, this is indeed the case.
-And these two shapes should have the same visual test power, unless the orientation of the visual pattern is a non-negligible factor from human perspective.
-However, the same graph suggests that the "Right-triangle" shape is more difficult to be noticed by subjects than the "Left-triangle" shape.
-
 ```{r heter-power-a, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$."}
@@
               method.args = list(family = binomial())) +
-
-  # Draw bootstrap for Left-triangle
-  geom_ribbon(data = ~get_boot_conf(filter(.x, 
-                                           conv_or_not == "Visual", 
-                                           a == "Left-triangle"),
-                                    log_effect_size,
-                                    reject,
-                                    times = 1000) %>%
-                mutate(conv_or_not = "Visual"),
-              aes(x = log_effect_size, 
-                  ymin = lower, 
-                  ymax = upper),
-              fill = rev(rcartocolor::carto_pal(3, "Vivid"))[3],
-              alpha = 0.2) +
-
-  # Draw bootstrap for Right-triangle
-  geom_ribbon(data = ~get_boot_conf(filter(.x, 
-                                           conv_or_not == "Visual", 
-                                           a == "Right-triangle"),
-                                    log_effect_size,
-                                    reject,
-                                    times = 1000) %>%
-                mutate(conv_or_not = "Visual"),
-              aes(x = log_effect_size, 
-                  ymin = lower, 
-                  ymax = upper),
-              fill = rev(rcartocolor::carto_pal(3, "Vivid"))[2],
-              alpha = 0.2) +
-
-  # Theme
   xlab(expression(log[e] (Effect_size))) +

### 2023-01-24 paper_comparison/paper_comparison/paper_comparison.tex
update color pal for the last section of the result
small changes

--- 2023-01-23 paper_comparison/paper_comparison/paper_comparison.tex
@@
 
-\caption{Comparison of power of conventional tests and visual tests on lineups with different fitted value distributions.
+\caption{Comparison of power on lineups with different fitted value distributions, for conventional tests and visual tests.
 Each power curve of the uniform distribution is surrounded by a grey line to indicate it is the baseline in its group.
-Conventional tests are more sensitive than visual tests in regards of the type of the simulated model and the fitted value distribution.
-For non-linearity patterns, visual tests show lower power with discrete uniform distribution, possibly due to the difficulty in reading discreteness.
-For heteroskedasticity patterns, because of the unevenness, visual tests have lower power with normal and lognormal distribution.}\label{fig:different-x-dist-poly-power}
+For both types of tests power does change.
+The discrete fitted value distribution substantially lowers the power for non-linearity patterns in residual plots.
+For heteroskedasticity patterns, normal and lognormal distributions produce lower power.}\label{fig:different-x-dist-poly-power}
 \end{figure}

### 2023-01-24 paper_comparison/paper_comparison/paper_comparison.tex

--- 2023-01-24 paper_comparison/paper_comparison/paper_comparison.tex
@@
 
-\caption{Power of different tests on lineups with different fitted value distributions.
-Each power curve of the uniform distribution is surrounded by a grey line to indicate it is the baseline within its group.
-Conventional tests are more sensitive than visual tests in regards of the type of the visual pattern and the fitted value distribution.
-Visual tests exhibit relatively low power in testing non-linearity patterns under discrete uniform distribution, possibly due to the difficulty in reading discreteness.
-Meanwhile, uneveness affects heteroskedasticity patterns greatly, result in low power of visual tests under normal and lognormal distribution.}\label{fig:different-x-dist-poly-power}
+\caption{Comparison of power on lineups with different fitted value distributions, for conventional tests and visual tests.
+Each power curve of the uniform distribution is surrounded by a grey line to indicate it is the baseline in its group.
+For both types of tests power does change.
+The discrete fitted value distribution substantially lowers the power for non-linearity patterns in residual plots.
+For heteroskedasticity patterns, normal and lognormal distributions produce lower power.}\label{fig:different-x-dist-poly-power}
 \end{figure}
@@
 
-\caption{Power of conventional tests and visual tests on lineups containing four different non-linearity shapes. }\label{fig:poly-power-j}
+\caption{Power of conventional tests and visual tests on lineups containing four different non-linearity shapes.
+Power curves with higher order of non-linearity are drawn with deeper colours.}\label{fig:poly-power-j}
 \end{figure}
@@
 
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  summarise paper content
+\item
+  briefly describe what was learned
+\item
+  how does this apply more broadly than simple regression
+\item
+  where does this lead? Computer vision residual plot reading
+\end{enumerate}
+
+\hypertarget{acknowledgement}{%
+\section{Acknowledgement}\label{acknowledgement}}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  software tools
+\item
+  point to github repo (create a new one)
+\item
+  summarise the supplementary material
+\end{enumerate}
+
 \newpage

### 2023-01-24 paper_comparison/paper_comparison/paper_comparison.Rmd
results on the fitted value distributions

@@
 
-Surprisingly, the visual test rejects `r nr$n[1]` out of `r nr$n[1]+nr$n[2]` (`r round(nr$n[1]/(nr$n[1]+nr$n[2])*100, 0)`%) of lineups where the conventional test does not reject.
+Surprisingly, the visual test rejects `r nr$n[1]` of the `r nr$n[1]+nr$n[2]` (`r round(nr$n[1]/(nr$n[1]+nr$n[2])*100, 0)`%) of lineups where the conventional test does not reject.
 Figure \ref{fig:heter-example} shows this lineup.
@@
 
-Although the distribution of the regressor is classically not restricted by the assumptions of linear regression, it will have an impact on the fitted value distribution and the visual effect of the residual plot.
-In our experiments, four distributions of fitted values are considered to accommodate this variation.
+The prediction in a regression model is $E(Y|X)$, that is, it is conditional on observed values of the predictors.
+The distribution of $X$, or consequently $\hat{Y}$, may however affect the ability to read any patterns in the residuals.
+<!-- Although the distribution of the regressor is classically not restricted by the assumptions of linear regression, it will have an impact on the fitted value distribution and the visual effect of the residual plot.--> The four distributions of fitted values, used in this experiment, were designed to examine this.
 

### 2023-01-24 paper_comparison/paper_comparison/paper_comparison.tex
results on the fitted value distributions

@@
 
-Surprisingly, the visual test rejects 1 out of 33 (3\%) of lineups where
+Surprisingly, the visual test rejects 1 of the 33 (3\%) of lineups where
 the conventional test does not reject. Figure \ref{fig:heter-example}
@@
 
-Although the distribution of the regressor is classically not restricted
-by the assumptions of linear regression, it will have an impact on the
-fitted value distribution and the visual effect of the residual plot. In
-our experiments, four distributions of fitted values are considered to
-accommodate this variation.
+The prediction in a regression model is \(E(Y|X)\), that is, it is
+conditional on observed values of the predictors. The distribution of
+\(X\), or consequently \(\hat{Y}\), may however affect the ability to
+read any patterns in the residuals. The four distributions of fitted
+values, used in this experiment, were designed to examine this.
 

### 2023-01-25 paper_comparison/paper_comparison/paper_comparison.Rmd
split out additional expt design into sep sub-section

@@
 Visual patterns of non-linearity are simulated using four different orders of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$).
-Four distributions of fitted values are also considered by using $X_{raw}$ follows different distributions: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
-A summary of the factors is given in Table \ref{tab:model-factor-table}.
+(A summary of the factors is given in Table \ref{tab:model-factor-table}.) The values of $j$ is chosen so that distinct shapes of non-linearity are included in the residual plot.
+These include "U", "S", "M" and "Triple-U" shape as shown in Figure \ref{fig:different-shape-of-herimite}.
+A greater value of $j$ will result in a curve with more turning points.
+It is expected that the "U" shape will be the easiest one to detect because complex shape tends to be concealed by cluster of data points.
 
@@
 
-The values of $j$ is chosen so that distinct shapes of non-linearity are included in the residual plot.
-These include "U", "S", "M" and "Triple-U" shape as shown in Figure \ref{fig:different-shape-of-herimite}.
-A greater value of $j$ will result in a curve with more turning points.
-It is expected that the "U" shape will be the easiest one to detect because complex shape tends to be concealed by cluster of data points.
-
-Different distributions of fitted values help enriching the pool of visual patterns as illustrated in Figure \ref{fig:different-dist}.
-The uniform and the normal distribution are symmetric and commonly assumed in statistical models.
-The adjusted log-normal distribution provides skewed density, while the discrete uniform distribution provides discreteness in residual plot.
-
 ```{r different-shape-of-herimite, fig.height = 2, fig.cap = "Polynomial forms generated for the residual plots used in experiment I.
@@
 For $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the regressor $\boldsymbol{x}$, which will lead to the presence of heteroskedasticity.
-Visual patterns of heteroskedasticity are simulated using three different shapes ($a$ = -1, 0, 1) and the same four different distribution of $X_{raw}$ used in experiment I.
-A summary of the factors can also be found in Table \ref{tab:model-factor-table}.
+Visual patterns of heteroskedasticity are simulated using three different shapes ($a$ = -1, 0, 1).
+(A summary of the factors can be found in Table \ref{tab:model-factor-table}.)
 
@@
 Nine out of 11 subjects correctly identify the data plot from this lineup.
+
+### Factors common to both experiments
+
+XXX DC will clean this up
+
+Four distributions of fitted values are also considered by using $X_{raw}$ follows different distributions: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
+
+Different distributions of fitted values help enriching the pool of visual patterns as illustrated in Figure \ref{fig:different-dist}.
+The uniform and the normal distribution are symmetric and commonly assumed in statistical models.
+The adjusted log-normal distribution provides skewed density, while the discrete uniform distribution provides discreteness in residual plot.
 

### 2023-01-25 paper_comparison/paper_comparison/paper_comparison.tex
split out additional expt design into sep sub-section

@@
 Visual patterns of non-linearity are simulated using four different
-orders of probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)). Four
-distributions of fitted values are also considered by using \(X_{raw}\)
-follows different distributions: (1) \(U(-1, 1)\), (2) \(N(0, 0.3^2)\),
-(3) \(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). A summary of the
-factors is given in Table \ref{tab:model-factor-table}.
+orders of probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)). (A
+summary of the factors is given in Table \ref{tab:model-factor-table}.)
+The values of \(j\) is chosen so that distinct shapes of non-linearity
+are included in the residual plot. These include ``U'', ``S'', ``M'' and
+``Triple-U'' shape as shown in Figure
+\ref{fig:different-shape-of-herimite}. A greater value of \(j\) will
+result in a curve with more turning points. It is expected that the
+``U'' shape will be the easiest one to detect because complex shape
+tends to be concealed by cluster of data points.
 
@@
 
-The values of \(j\) is chosen so that distinct shapes of non-linearity
-are included in the residual plot. These include ``U'', ``S'', ``M'' and
-``Triple-U'' shape as shown in Figure
-\ref{fig:different-shape-of-herimite}. A greater value of \(j\) will
-result in a curve with more turning points. It is expected that the
-``U'' shape will be the easiest one to detect because complex shape
-tends to be concealed by cluster of data points.
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-herimite-1}
+
+}
+
+\caption{Polynomial forms generated for the residual plots used in experiment I.
+The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$.}\label{fig:different-shape-of-herimite}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-dist-1}
+
+}
+
+\caption{Variations in fitted values, that might affect perception of residual plots. Four different distributions are used.}\label{fig:different-dist}
+\end{figure}
+
+Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups
+used in experiment I. This lineup is produced by the non-linearity model
+with \(j = 6\) and uniform fitted values. The data plot location is
+\(2^3 - 4\). All five subjects correctly identify the data plot from
+this lineup.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-n-1}
+
+}
+
+\caption{Number of observations $n$ have an impact on the strength of the signal. Three different values of $n$ are used in experiment I, II and III.}\label{fig:different-n}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-sigma-1}
+
+}
+
+\caption{For the non-lienearity model, the standard deviation of the error term $\sigma$ affects the strength of the signal.
+Four different values of $\sigma$ are used in the experiment I.}\label{fig:different-sigma}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-poly-lineup-1}
+
+}
+
+\caption{One of the lineups containing non-linearity patterns used in experiment I. Can you spot the most different plot? The data plot is positioned at $2^3 + 1$.}\label{fig:example-poly-lineup}
+\end{figure}
+
+\hypertarget{heteroskedasticity}{%
+\subsubsection{Heteroskedasticity}\label{heteroskedasticity}}
+
+Experiment II is designed to study the ability of human subjects to
+detect the appearance of a heteroskedasticity pattern under a simple
+linear regression model setting:
+
+\begin{align} \label{eq:heter-model}
+\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
+\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
+\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}, 1 + (2 - |a|)(\boldsymbol{x} - a)^2b \boldsymbol{I}), 
+\end{align}
+
+\noindent where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
+\(\boldsymbol{\varepsilon}\) are vectors of size \(n\) and \(g(.)\) is
+the scaling function defined in Equation \ref{eq:scaling-function}.
+
+The null regression model used to fit the realizations generated by the
+above model is formulated exactly the same as Equation
+\ref{eq:null-model}.
+
+For \(b \neq 0\), the variance-covariance matrix of the error term
+\(\boldsymbol{\varepsilon}\) is correlated with the regressor
+\(\boldsymbol{x}\), which will lead to the presence of
+heteroskedasticity. Visual patterns of heteroskedasticity are simulated
+using three different shapes (\(a\) = -1, 0, 1). (A summary of the
+factors can be found in Table \ref{tab:model-factor-table}.)
+
+Since \(supp(X) = [-1, 1]\), choosing \(a\) to be \(-1\), \(0\) and
+\(1\) can generate ``left-triangle'', ``butterfly'' and
+``right-triangle'' shape as displayed in Figure
+\ref{fig:different-shape-of-heter}. The term \((2 - |a|)\) maintains the
+magnitude of residuals across different values of \(a\).
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-heter-1}
+
+}
+
+\caption{Heteroskedasticity forms used in experiment II.
+Three different shapes (a = -1, 0, 1) are used in the experiment to create left-triangle, butterfly and right-triangle shapes.}\label{fig:different-shape-of-heter}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-b-1}
+
+}
+
+\caption{Five different values of $b$ are used in experiment II to control the strength of the signal.}\label{fig:different-b}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-heter-lineup-1}
+
+}
+
+\caption{One of the lineups containing heteroskedasticity pattern used in experiment II.
+Can you spot the most different plot?
+The data plot is positioned at $3^3 - 3^2$}\label{fig:example-heter-lineup}
+\end{figure}
+
+An example lineup of this model used in Experiment II is shown in Figure
+\ref{fig:example-heter-lineup} with \(a = -1\) and
+\(X_{raw} \sim U(-1, 1)\). The data plot location is \(2^4 + 2\). Nine
+out of 11 subjects correctly identify the data plot from this lineup.
+
+\hypertarget{factors-common-to-both-experiments}{%
+\subsubsection{Factors common to both
+experiments}\label{factors-common-to-both-experiments}}
+
+XXX DC will clean this up
+
+Four distributions of fitted values are also considered by using
+\(X_{raw}\) follows different distributions: (1) \(U(-1, 1)\), (2)
+\(N(0, 0.3^2)\), (3) \(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\).
 
@@
 discreteness in residual plot.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-herimite-1}
-
-}
-
-\caption{Polynomial forms generated for the residual plots used in experiment I.
-The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$.}\label{fig:different-shape-of-herimite}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-dist-1}
-
-}
-
-\caption{Variations in fitted values, that might affect perception of residual plots. Four different distributions are used.}\label{fig:different-dist}
-\end{figure}
-
-Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups
-used in experiment I. This lineup is produced by the non-linearity model
-with \(j = 6\) and uniform fitted values. The data plot location is
-\(2^3 - 4\). All five subjects correctly identify the data plot from
-this lineup.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-n-1}
-
-}
-
-\caption{Number of observations $n$ have an impact on the strength of the signal. Three different values of $n$ are used in experiment I, II and III.}\label{fig:different-n}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-sigma-1}
-
-}
-
-\caption{For the non-lienearity model, the standard deviation of the error term $\sigma$ affects the strength of the signal.
-Four different values of $\sigma$ are used in the experiment I.}\label{fig:different-sigma}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-poly-lineup-1}
-
-}
-
-\caption{One of the lineups containing non-linearity patterns used in experiment I. Can you spot the most different plot? The data plot is positioned at $2^3 + 1$.}\label{fig:example-poly-lineup}
-\end{figure}
-
-\hypertarget{heteroskedasticity}{%
-\subsubsection{Heteroskedasticity}\label{heteroskedasticity}}
-
-Experiment II is designed to study the ability of human subjects to
-detect the appearance of a heteroskedasticity pattern under a simple
-linear regression model setting:
-
-\begin{align} \label{eq:heter-model}
-\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
-\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
-\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}, 1 + (2 - |a|)(\boldsymbol{x} - a)^2b \boldsymbol{I}), 
-\end{align}
-
-\noindent where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
-\(\boldsymbol{\varepsilon}\) are vectors of size \(n\) and \(g(.)\) is
-the scaling function defined in Equation \ref{eq:scaling-function}.
-
-The null regression model used to fit the realizations generated by the
-above model is formulated exactly the same as Equation
-\ref{eq:null-model}.
-
-For \(b \neq 0\), the variance-covariance matrix of the error term
-\(\boldsymbol{\varepsilon}\) is correlated with the regressor
-\(\boldsymbol{x}\), which will lead to the presence of
-heteroskedasticity. Visual patterns of heteroskedasticity are simulated
-using three different shapes (\(a\) = -1, 0, 1) and the same four
-different distribution of \(X_{raw}\) used in experiment I. A summary of
-the factors can also be found in Table \ref{tab:model-factor-table}.
-
-Since \(supp(X) = [-1, 1]\), choosing \(a\) to be \(-1\), \(0\) and
-\(1\) can generate ``left-triangle'', ``butterfly'' and
-``right-triangle'' shape as displayed in Figure
-\ref{fig:different-shape-of-heter}. The term \((2 - |a|)\) maintains the
-magnitude of residuals across different values of \(a\).
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-heter-1}
-
-}
-
-\caption{Heteroskedasticity forms used in experiment II.
-Three different shapes (a = -1, 0, 1) are used in the experiment to create left-triangle, butterfly and right-triangle shapes.}\label{fig:different-shape-of-heter}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-b-1}
-
-}
-
-\caption{Five different values of $b$ are used in experiment II to control the strength of the signal.}\label{fig:different-b}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-heter-lineup-1}
-
-}
-
-\caption{One of the lineups containing heteroskedasticity pattern used in experiment II.
-Can you spot the most different plot?
-The data plot is positioned at $3^3 - 3^2$}\label{fig:example-heter-lineup}
-\end{figure}
-
-An example lineup of this model used in Experiment II is shown in Figure
-\ref{fig:example-heter-lineup} with \(a = -1\) and
-\(X_{raw} \sim U(-1, 1)\). The data plot location is \(2^4 + 2\). Nine
-out of 11 subjects correctly identify the data plot from this lineup.
 

### 2023-01-25 paper_comparison/appendix.Rmd

@@
 
-## Effect derivation
+## Effect size derivation
+
+Effect size can be defined as the difference of a parameter for a particular model or distribution,  or a statistic derived from a sample.
+Importantly, it needs to reflect the treatment we try to measure.
+Centred on a conventional statistical test, we usually can deduce the effect size from the test statistic by substituting the null parameter value.
+When considering the diagnostics of residual departures, there exist many possibilities of test statistics for a variety of model assumptions.
+Meanwhile, diagnostic plots such as the residual plot have no general agreement on measuring how strong a model violation pattern is.
+To build a bridge between various residual-based tests, and the visual test, we focus on the shared information embedded in the testing procedures, which is the distribution of residuals.
+When comes to comparison of distribution, Kullback-Leibler divergence is a classical way to represent the information loss or entropy increase caused by the approximation to the true distribution, which in our case, the inefficiency due to the use of false model assumptions.
+
+Following the terminology introduced by @kullback1951information.
+$P$ represents the measured probability distribution, and $Q$ represents the assumed probability distribution.
+The Kullback-Leibler divergence is given as $\int_{-\infty}^{\infty}log(p(x)/q(x))p(x)dx$, where $p(.)$ and $q(.)$ denote probability densities of $P$ and $Q$.
 

### 2023-01-28 paper_comparison/appendix.Rmd
update meeting note

@@
 
-Following the terminology introduced by @kullback1951information.
-$P$ represents the measured probability distribution, and $Q$ represents the assumed probability distribution.
-The Kullback-Leibler divergence is given as $\int_{-\infty}^{\infty}log(p(x)/q(x))p(x)dx$, where $p(.)$ and $q(.)$ denote probability densities of $P$ and $Q$.
+Following the terminology introduced by @kullback1951information, $P$ represents the measured probability distribution, and $Q$ represents the assumed probability distribution.
+The Kullback-Leibler divergence is defined as $\int_{-\infty}^{\infty}log(p(x)/q(x))p(x)dx$, where $p(.)$ and $q(.)$ denote probability densities of $P$ and $Q$.
 

### 2023-01-29 paper_comparison/paper_comparison/paper_comparison.Rmd
update meeting note
change the first plot to a faceted plot.
Remove axis and gridline from many plots.

@@
 ```{r get-lineup-data}
+# Create the dir folder
+if (!dir.exists(here::here("data/"))) dir.create(here::here("data/"))
+
 # The lineup data used to draw residual plot needed to be downloaded
@@
   vi_lineup <- get_vi_lineup()
-  if (!dir.exists(here::here("data/"))) dir.create(here::here("data/"))
   saveRDS(vi_lineup, here::here("data/vi_lineup.rds"))
@@
 
+(change it to facet)
+
+(remove backticks for some plots)
+
 ```{r residual-plot-common-departures, fig.width = 10, fig.height = 2.5, fig.cap = "Example residual vs fitted plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
@@
 
-patchwork::wrap_plots(
-
-  # Plot base data
-  mod$plot(dat, theme = theme_light()) +
-    ggtitle("A. Good residuals") + 
+bind_rows(
+  dat %>%
+    mutate(type = "Good residuals"),
+  mod$set_prm("include_z", TRUE)$
+    gen(300, computed = select(dat, x, e)) %>%
+    mutate(type = "Non-linearity"),
+  heter_model(b = 64)$
+    gen(300, computed = select(dat, x, e)) %>%
+    mutate(type = "Heteroskedasticity"),
+  heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
+    gen(300, computed = select(dat, x)) %>%
+    mutate(type = "Non-normality")
+) %>%
+  mod$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
+  facet_wrap(~type, ncol = 4, scales = "free") +
   xlab("Fitted values") +
-    ylab("Residuals"),
-
-  # Include the z term
-  # Reuse the x and e terms
-  mod$plot(mod$
-             set_prm("include_z", TRUE)$
-             gen(300, computed = select(dat, x, e)), 
-    theme = theme_light()) + 
-    ggtitle("B. Non-linearity") +
-    xlab("Fitted values") + 
-    ylab("Residuals"),
-
-  # Build a heter model
-  # Reuse the x and e terms
-  mod$plot(heter_model(b = 64)$
-             gen(300, computed = select(dat, x, e)), 
-    theme = theme_light()) + 
-    ggtitle("C. Heteroskedasticity") +
-    xlab("Fitted values") + 
-    ylab("Residuals"),
-
-  # Build a heter model with non-normal error
-  # Reuse the x term
-  mod$plot(heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
-             gen(300, computed = select(dat, x)), 
-    theme = theme_light()) + 
-    ggtitle("D. Non-normality") +
-    xlab("Fitted values") + 
-    ylab("Residuals"),
-  ncol = 4
-)
+  ylab("Residuals")
 ```
@@
 Graphical summaries in which residuals are plotted against fitted values or other functions of the regressors that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals.
-Figure \ref{fig:residual-plot-common-departures}A shows an ideal residual plot with residuals evenly distributed at both sides of the horizontal zero line, with no noticeable patterns.
+The first panel of Figure \ref{fig:residual-plot-common-departures} shows an ideal residual plot with residuals evenly distributed at both sides of the horizontal zero line, with no noticeable patterns.
 
@@
 Any non-linear functional form of residuals on fitted values in the residual plot could be indicative of non-linearity.
-An example residual plot containing visual pattern of non-linearity is given at Figure \ref{fig:residual-plot-common-departures}B.
+An example residual plot containing visual pattern of non-linearity is given at the second panel of Figure \ref{fig:residual-plot-common-departures}.
 One can clearly observe the "S-shape" from the residual plot as the cubic term is not captured by the misspecified model.
@@
 The usual pattern of heteroskedasticity on a residual plot is the inconsistent spread of the residuals across the horizontal axis.
-Visually, it sometimes results in the so-called "butterfly" shape as shown in the Figure \ref{fig:residual-plot-common-departures}C, or the "left-triangle" and "right-triangle" shape where the smallest variance occurs at one side of the horizontal axis.
+Visually, it sometimes results in the so-called "butterfly" shape as shown in the third panel of Figure \ref{fig:residual-plot-common-departures}, or the "left-triangle" and "right-triangle" shape where the smallest variance occurs at one side of the horizontal axis.
 
@@
 As we mainly discuss residual plots, non-normality will not be the focus of this paper.
-For a consistent comparison, the residual plot of this departure is still presented in Figure \ref{fig:residual-plot-common-departures}D.
+For a consistent comparison, the residual plot of this departure is still presented in the fourth panel of Figure \ref{fig:residual-plot-common-departures}.
 When the number residuals below and above the horizontal axis are uneven across the local regions along the $x$-axis, we expect that the normality assumption is violated.
-For example, given a skewed error distribution, there will be fewer data points and more outliers on one side of the horizontal axis as shown in Figure \ref{fig:residual-plot-common-departures}D.
+For example, given a skewed error distribution, there will be fewer data points and more outliers on one side of the horizontal axis as shown in the fourth panel of Figure \ref{fig:residual-plot-common-departures}.
 
@@
   mutate(shape = factor(shape, levels = shape_labels)) %>%
-  VI_MODEL$plot() +
+  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
   facet_wrap(~shape, scales = "free", labeller = label_parsed, ncol = 4) +
-  theme_light() +
   xlab("Fitted values") +
@@
                                     "Discrete"))) %>%
-  VI_MODEL$plot() +
-  theme_light() +
+  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
   xlab("Fitted values") +
@@
   mutate(n = factor(n, levels = c("N: 50", "N: 100", "N: 300"))) %>%
-  VI_MODEL$plot() +
-  theme_light() +
+  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
   facet_wrap(~n, scales = "free", ncol = 3) +
@@
   bind_rows(dat_sigma_05) %>%
-  VI_MODEL$plot() + 
+  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) + 
   facet_wrap(~sigma, ncol = 4, scales = "free", labeller = label_parsed) +
-  theme_light() +
   xlab("Fitted values") +
@@
   mutate(a = factor(a, levels = a_labels)) %>%
-  VI_MODEL$plot() +
+  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
   facet_wrap(~a, scales = "free", ncol = 3) +
-  theme_light() +
   xlab("Fitted values") +
@@
   mutate(b = factor(b, levels = paste0("b: ", c(0.25, 1, 4, 16, 64)))) %>%
-  VI_MODEL$plot() +
+  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
   facet_wrap(~b, scales = "free", ncol = 5) +
-  theme_light() +
   xlab("Fitted values") +

### 2023-01-29 paper_comparison/paper_comparison/paper_comparison.tex
change the first plot to a faceted plot.
Remove axis and gridline from many plots.

@@
 
+(change it to facet)
+
+(remove backticks for some plots)
+
 \begin{figure}
@@
 to residuals are referred to as standard residual plots in
-\citet{cook1982residuals}. Figure
-\ref{fig:residual-plot-common-departures}A shows an ideal residual plot
+\citet{cook1982residuals}. The first panel of Figure
+\ref{fig:residual-plot-common-departures} shows an ideal residual plot
 with residuals evenly distributed at both sides of the horizontal zero
@@
 residual plot could be indicative of non-linearity. An example residual
-plot containing visual pattern of non-linearity is given at Figure
-\ref{fig:residual-plot-common-departures}B. One can clearly observe the
-``S-shape'' from the residual plot as the cubic term is not captured by
-the misspecified model.
+plot containing visual pattern of non-linearity is given at the second
+panel of Figure \ref{fig:residual-plot-common-departures}. One can
+clearly observe the ``S-shape'' from the residual plot as the cubic term
+is not captured by the misspecified model.
 
@@
 Visually, it sometimes results in the so-called ``butterfly'' shape as
-shown in the Figure \ref{fig:residual-plot-common-departures}C, or the
-``left-triangle'' and ``right-triangle'' shape where the smallest
-variance occurs at one side of the horizontal axis.
+shown in the third panel of Figure
+\ref{fig:residual-plot-common-departures}, or the ``left-triangle'' and
+``right-triangle'' shape where the smallest variance occurs at one side
+of the horizontal axis.
 
@@
 paper. For a consistent comparison, the residual plot of this departure
-is still presented in Figure \ref{fig:residual-plot-common-departures}D.
-When the number residuals below and above the horizontal axis are uneven
-across the local regions along the \(x\)-axis, we expect that the
-normality assumption is violated. For example, given a skewed error
-distribution, there will be fewer data points and more outliers on one
-side of the horizontal axis as shown in Figure
-\ref{fig:residual-plot-common-departures}D.
+is still presented in the fourth panel of Figure
+\ref{fig:residual-plot-common-departures}. When the number residuals
+below and above the horizontal axis are uneven across the local regions
+along the \(x\)-axis, we expect that the normality assumption is
+violated. For example, given a skewed error distribution, there will be
+fewer data points and more outliers on one side of the horizontal axis
+as shown in the fourth panel of Figure
+\ref{fig:residual-plot-common-departures}.
 

### 2023-01-29 paper_comparison/paper_comparison/paper_comparison.Rmd
replace CI band with lines

@@
 ```{r}
+# Get bootstrap samples
+get_boot_sample <- function(dat, times = 100) {
+  map_df(1:times, 
+         ~slice_sample(dat, n = nrow(dat), replace = TRUE) %>%
+           mutate(boot_id = .x))
+}
+
 # Get bootstrap CI
@@
   # Draw bootstrap CI for visual test
-  geom_ribbon(data = ~get_boot_conf(.x,
-                                    log_effect_size,
+  # geom_ribbon(data = ~get_boot_conf(.x,
+  #                                   log_effect_size,
+  #                                   reject,
+  #                                   times = 1000),
+  #             aes(x = log_effect_size, 
+  #                 ymin = lower, 
+  #                 ymax = upper),
+  #             fill = rev(rcartocolor::carto_pal(4, "Vivid"))[4],
+  #             alpha = 0.2) +
+
+  stat_smooth(data = ~get_boot_sample(.x, times = 1000),
+              geom = "line",
+              aes(log_effect_size, 
                   reject, 
-                                    times = 1000),
-              aes(x = log_effect_size, 
-                  ymin = lower, 
-                  ymax = upper),
-              fill = rev(rcartocolor::carto_pal(4, "Vivid"))[4],
-              alpha = 0.2) +
+                  col = "Visual test", 
+                  group = boot_id),
+              method = "glm",
+              method.args = list(family = binomial),
+              se = FALSE,
+              alpha = 0.01) +
 
@@
     ggtitle(round(es[j], 1)) +
-    theme(aspect.ratio=0.8)
+    theme(aspect.ratio = 0.8)
 }

### 2023-01-29 paper_comparison/paper_comparison/paper_comparison.Rmd
Fix the caption for power cmp

@@
 Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups.
-Surrounding band of the visual test shows the 95 percent confidence interval estimated using 1000 bootstrap samples.
+Surrounding lines of the visual test show the estimated power of 1000 bootstrap samples.
 Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
@@
 Main plot shows the power curves, with dots indicating human evaluations of lineups.
-Surrounding band of the visual test shows the 95 percent confidence interval estimated using 500 bootstrap samples.
+Surrounding lines of the visual test show the estimated power of 1000 bootstrap samples.
 Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.

### 2023-01-30 paper_comparison/paper_comparison/paper_comparison.Rmd
rearrange the data used in the result

@@
 
-In the following analysis, lineups with uniform fitted values will be the focus of next two sections.
+In the following analysis, lineups with uniform fitted values will be the focus.
 Visual patterns are more likely to be revealed under a uniform distribution.
 Additionally, we have collected extra evaluations on these lineups, which will result in more reliable analysis.
-Analysis of lineups with other fitted value distributions can be found in Section \ref{fitted-value-distribution}.
+Analysis of lineups with other fitted value distributions can be found in Section \ref{effect-of-other-fitted-value-distributions}.
 
@@
 
-## Effect of other factors used in the simulation model
+## Effect of other fitted value distributions
 

### 2023-01-30 paper_comparison/paper_comparison/paper_comparison.tex
rearrange the data used in the result

@@
 In the following analysis, lineups with uniform fitted values will be
-the focus of next two sections. Visual patterns are more likely to be
-revealed under a uniform distribution. Additionally, we have collected
-extra evaluations on these lineups, which will result in more reliable
-analysis. Analysis of lineups with other fitted value distributions can
-be found in Section \ref{fitted-value-distribution}.
+the focus. Visual patterns are more likely to be revealed under a
+uniform distribution. Additionally, we have collected extra evaluations
+on these lineups, which will result in more reliable analysis. Analysis
+of lineups with other fitted value distributions can be found in Section
+\ref{effect-of-other-fitted-value-distributions}.
 
@@
 
-\hypertarget{effect-of-other-factors-used-in-the-simulation-model}{%
-\subsection{Effect of other factors used in the simulation
-model}\label{effect-of-other-factors-used-in-the-simulation-model}}
+\hypertarget{effect-of-other-fitted-value-distributions}{%
+\subsection{Effect of other fitted value
+distributions}\label{effect-of-other-fitted-value-distributions}}
 

### 2023-01-30 paper_comparison/paper_comparison/paper_comparison.Rmd
update shape of non-linearity section

@@
 
-The power comparison illustrates that appropriate conventional test will reject $H_0$ more aggressively than visual tests.
-In this section, we explore how often they agree with each other by investigating the rejections for the two studies based on $p$-values for each lineup.
+The power comparison illustrates that appropriate conventional tests will reject $H_0$ more aggressively than visual tests.
+In this section, we explore how often they agree with each other by investigating the rejections for the two model designs based on $p$-values for each lineup.
 
@@
 
-For lineups containing non-linearity patterns, conventional tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`%
-and visual tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the time.
-Of the tests rejected by the conventional test, `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by the visual test, that is, approximately half as many as the conventional test.
+For lineups containing non-linearity patterns, conventional tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% and visual tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the time.
+Of the lineups rejected by the conventional test, `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by the visual test, that is, approximately half as many as the conventional test.
 There are no lineups that are rejected by the visual test but not by the conventional test.
 
-For lineups containing non-linearity patterns, conventional tests reject `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`%
-and visual tests reject `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the time.
-Of the tests rejected by the conventional test, `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by the visual test, that is, approximately half as many as the conventional test.
+In terms of lineups containing heteroskedasticity patterns, `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by conventional tests, while `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by visual tests.
+When the conventional test rejects a lineup, there is a great chance (`r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`%) the visual test will also reject it.
 
@@
 ```
+
+## Shape of non-linearity
+
+Shape of the non-linearity pattern is another factor contributes to the non-linearity model. With different non-linearity patterns, it is expected that the visual test will respond variously.
+
+According to Figure \ref{fig:poly-power-uniform-j}, conventional tests have almost identical power in testing the "U", the "S" and the "M" shape, except for testing the "Triple-U" shape, which have significantly lower power.
+To understand why this is, one needs to return to the way the RESET test is applied.
+It requires a parameter indicating degree of fitted values to test for, and the recommendation is to generically use four (@ramsey_tests_1969).
+However, the "Triple-U" shape constructed from the Hermite polynomials use power up to 18.
+If the RESET test had been applied using a higher power no less than six, the power curve of "Triple-U" shape will have little difference to other shapes.
+The recommendation of the polynomial power for the RESET should be revised, perhaps.
+This illustrates the sensitivity of conventional testing to the parameters, and it also points to a limitation that one needs to know the data structure in order to set the parameters for the test.
+
+For visual tests, we expect the "U" shape will be the easiest one to be detected by subjects followed by the "S", "M" and "Triple-U" shape.
+From Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power curves are mostly aligned with our expectation, except for the "M" shape.
+Interestingly, the "M" shape is as easy to be detected as the "U" shape, which implies the readability of the shape do not strictly follow the degree of the polynomials.
+
+```{r poly-power-uniform-j, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups containing four different non-linearity shapes.
+Power curves with higher order of non-linearity are drawn with deeper colours."}
+
+poly_conv_sim %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  mutate(reject = as.numeric(RESET4_p_value < 0.05)) %>%
+  mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>%
+  mutate(conv_or_not = "Conventional") %>%
+  bind_rows(vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(x_dist == "uniform", type == "polynomial") %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  mutate(reject = as.numeric(p_value < 0.05)) %>%
+  mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>%
+  mutate(conv_or_not = "Visual")) %>%
+  mutate(shape = factor(shape, levels = c("U", "S", "M", "Triple-U"))) %>%
+  select(shape, log_effect_size, reject, conv_or_not) %>%
+  ggplot() +
+  geom_smooth(aes(log_effect_size, reject, col = shape), 
+              se = FALSE, 
+              method = "glm",
+              method.args = list(family = binomial())) +
+  xlab(expression(log[e] (Effect_size))) +
+  ylab("Power") +
+  scale_color_manual(values = rcartocolor::carto_pal(4, "TealGrn")) +
+  theme_light() +
+  theme(legend.position = "bottom") +
+  facet_wrap(~conv_or_not) +
+  labs(col = "Non-linearity shape")
+```
+
+## Shape of heteroskedasticity
 
@@
 
-### Shape of non-linearity
+### Shape of non-linearity (old)
 
@@
 
-### Shape of heteroskedasticity
+### Shape of heteroskedasticity (old)
 

### 2023-01-30 paper_comparison/paper_comparison/paper_comparison.tex
update shape of non-linearity section

@@
 
-The power comparison illustrates that appropriate conventional test will
-reject \(H_0\) more aggressively than visual tests. In this section, we
-explore how often they agree with each other by investigating the
-rejections for the two studies based on \(p\)-values for each lineup.
+The power comparison illustrates that appropriate conventional tests
+will reject \(H_0\) more aggressively than visual tests. In this
+section, we explore how often they agree with each other by
+investigating the rejections for the two model designs based on
+\(p\)-values for each lineup.
 
@@
 For lineups containing non-linearity patterns, conventional tests reject
-69\% and visual tests reject 32\% of the time. Of the tests rejected by
-the conventional test, 46\% are rejected by the visual test, that is,
+69\% and visual tests reject 32\% of the time. Of the lineups rejected
+by the conventional test, 46\% are rejected by the visual test, that is,
 approximately half as many as the conventional test. There are no
@@
 
-For lineups containing non-linearity patterns, conventional tests reject
-76\% and visual tests reject 56\% of the time. Of the tests rejected by
-the conventional test, 73\% are rejected by the visual test, that is,
-approximately half as many as the conventional test.
+In terms of lineups containing heteroskedasticity patterns, 76\% are
+rejected by conventional tests, while 56\% are rejected by visual tests.
+When the conventional test rejects a lineup, there is a great chance
+(73\%) the visual test will also reject it.
 
@@
 \end{figure}
+
+\hypertarget{shape-of-non-linearity}{%
+\subsection{Shape of non-linearity}\label{shape-of-non-linearity}}
+
+Shape of the non-linearity pattern is another factor contributes to the
+non-linearity model. With different non-linearity patterns, it is
+expected that the visual test will respond variously.
+
+According to Figure \ref{fig:poly-power-uniform-j}, conventional tests
+have almost identical power in testing the ``U'', the ``S'' and the
+``M'' shape, except for testing the ``Triple-U'' shape, which have
+significantly lower power. To understand why this is, one needs to
+return to the way the RESET test is applied. It requires a parameter
+indicating degree of fitted values to test for, and the recommendation
+is to generically use four (\citet{ramsey_tests_1969}). However, the
+``Triple-U'' shape constructed from the Hermite polynomials use power up
+to 18. If the RESET test had been applied using a higher power no less
+than six, the power curve of ``Triple-U'' shape will have little
+difference to other shapes. The recommendation of the polynomial power
+for the RESET should be revised, perhaps. This illustrates the
+sensitivity of conventional testing to the parameters, and it also
+points to a limitation that one needs to know the data structure in
+order to set the parameters for the test.
+
+For visual tests, we expect the ``U'' shape will be the easiest one to
+be detected by subjects followed by the ``S'', ``M'' and ``Triple-U''
+shape. From Figure \ref{fig:poly-power-uniform-j}, it can be observed
+that the power curves are mostly aligned with our expectation, except
+for the ``M'' shape. Interestingly, the ``M'' shape is as easy to be
+detected as the ``U'' shape, which implies the readability of the shape
+do not strictly follow the degree of the polynomials.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/poly-power-uniform-j-1}
+
+}
+
+\caption{Power of conventional tests and visual tests on lineups containing four different non-linearity shapes.
+Power curves with higher order of non-linearity are drawn with deeper colours.}\label{fig:poly-power-uniform-j}
+\end{figure}
+
+\hypertarget{shape-of-heteroskedasticity}{%
+\subsection{Shape of
+heteroskedasticity}\label{shape-of-heteroskedasticity}}
 
@@
 
-\hypertarget{shape-of-non-linearity}{%
-\subsubsection{Shape of non-linearity}\label{shape-of-non-linearity}}
+\hypertarget{shape-of-non-linearity-old}{%
+\subsubsection{Shape of non-linearity
+(old)}\label{shape-of-non-linearity-old}}
 
@@
 
-\hypertarget{shape-of-heteroskedasticity}{%
-\subsubsection{Shape of
-heteroskedasticity}\label{shape-of-heteroskedasticity}}
+\hypertarget{shape-of-heteroskedasticity-old}{%
+\subsubsection{Shape of heteroskedasticity
+(old)}\label{shape-of-heteroskedasticity-old}}
 

### 2023-01-30 paper_comparison/paper_comparison/paper_comparison.Rmd
table 2 refined

--- 2023-01-29 paper_comparison/paper_comparison/paper_comparison.Rmd
@@
 data.frame(j = c("2", "3", "6", "18", ""),
-           x_dist = c("Uniform", "Normal", "Skewed", "Discrete uniform", ""),
            sigma = c("0.25", "1.00", "2.00", "4.00", ""),
@@
            b = c("0.25", "1.00", "4.00", "16.00", "64.00"),
-           n = c("50", "100", "300", "", "")) %>%
+           n = c("50", "100", "300", "", ""), 
+           x_dist = c("Uniform", "Normal", "Skewed", "Discrete uniform", "")) %>%
   knitr::kable(col.names = NULL,
-               align = "rcrrrr",
+               align = "rr|rr|rc",
                escape = FALSE,
@@
                booktabs = TRUE, 
-               caption  = 'Description of all factors involved in the non-linear and heteroskedasticity studies.', 
+               caption  = 'Levels of the factors used in experiments I, II, III.', 
                label = "model-factor-table" ) %>%
@@
   kableExtra::add_header_above(c("Poly Order (sym1)", #"Order of Hermite polynomial\n(sym1)", 
-                                 "Distribution of fitted values", 
                                  "SD (sym3)", #"Standard deviation of polynomial model\n(sym3)",
-                                 "Heteroskedasticity Shape (sym4)", # "Shape of heteroskedasticity\n(sym4)", 
-                                 "Heteroskedasticity (sym5)", # Variance factor of heteroskedasticity model\n(sym5)",
-                                 "Size (sym6)")) %>% #"Sample size\n(sym6)")) %>%
+                                 "Shape (sym4)", # "Shape of heteroskedasticity\n(sym4)", 
+                                 "Ratio (sym5)", # Variance factor of heteroskedasticity model\n(sym5)",
+                                 "Size (sym6)", 
+                                 "Distribution of fitted values")) %>% #"Sample size\n(sym6)")) %>%
+  kableExtra::add_header_above(c("Non-linearity" = 2, 
+                               "Heteroskedasticity" = 2, "Common" = 2)) %>%
   as.character() %>%

### 2023-01-30 paper_comparison/paper_comparison/paper_comparison.tex
table 2 refined

--- 2023-01-29 paper_comparison/paper_comparison/paper_comparison.tex
@@
 
-\caption{\label{tab:model-factor-table}Description of all factors involved in the non-linear and heteroskedasticity studies.}
+\caption{\label{tab:model-factor-table}Levels of the factors used in experiments I, II, III.}
 \centering
 \resizebox{\linewidth}{!}{
-\begin{tabular}[t]{rcrrrr}
+\begin{tabular}[t]{rr|rr|rcrr|rr|rcrr|rr|rcrr|rr|rcrr|rr|rcrr|rr|rc}
 \toprule
-\multicolumn{1}{c}{Poly Order ($j$)} & \multicolumn{1}{c}{Distribution of fitted values} & \multicolumn{1}{c}{SD ($\sigma$)} & \multicolumn{1}{c}{Heteroskedasticity Shape ($a$)} & \multicolumn{1}{c}{Heteroskedasticity ($b$)} & \multicolumn{1}{c}{Size ($n$)} \\
+\multicolumn{2}{c}{Non-linearity} & \multicolumn{2}{c}{Heteroskedasticity} & \multicolumn{2}{c}{Common} \\
+\cmidrule(l{3pt}r{3pt}){1-2} \cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6}
+\multicolumn{1}{c}{Poly Order ($j$)} & \multicolumn{1}{c}{SD ($\sigma$)} & \multicolumn{1}{c}{Shape ($a$)} & \multicolumn{1}{c}{Ratio ($b$)} & \multicolumn{1}{c}{Size ($n$)} & \multicolumn{1}{c}{Distribution of fitted values} \\
 \cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
-2 & Uniform & 0.25 & -1 & 0.25 & 50\\
-3 & Normal & 1.00 & 0 & 1.00 & 100\\
-6 & Skewed & 2.00 & 1 & 4.00 & 300\\
-18 & Discrete uniform & 4.00 &  & 16.00 & \\
- &  &  &  & 64.00 & \\
+2 & 0.25 & -1 & 0.25 & 50 & Uniform\\
+3 & 1.00 & 0 & 1.00 & 100 & Normal\\
+6 & 2.00 & 1 & 4.00 & 300 & Skewed\\
+18 & 4.00 &  & 16.00 &  & Discrete uniform\\
+ &  &  & 64.00 &  & \\
 \bottomrule

### 2023-01-30 paper_comparison/paper_comparison/paper_comparison.Rmd
refine the start of the experimental design

@@
 Three experiments are conducted to investigate the difference between conventional hypothesis testing and visual inference in the application of linear regression diagnostics.
-The experiment I has ideal scenario for conventional testing, where the visual test is not expected to outperform the conventional test.
-Meanwhile, the experiment II is a scenario where the conventional test is an approximate test, in which the visual test may have a chance to match the performance of the conventional test.
+Two types of departures, namely non-linearity and heteroskedasticity, are considered with the corresponding data generating process being designed for experiment I and II.
 The experiment III is designed for collecting human responses to null lineups such that the parameter $\alpha$ in Equation \ref{eq:pvalue-beta-binomial} can be estimated.
@@
 ## Simulating departures from good residuals
-
-Two types of departures, namely non-linearity and heteroskedasticity, are considered with the corresponding data generating process being designed for experiment I and II.
 

### 2023-01-30 paper_comparison/paper_comparison/paper_comparison.tex
refine the start of the experimental design

@@
 conventional hypothesis testing and visual inference in the application
-of linear regression diagnostics. The experiment I has ideal scenario
-for conventional testing, where the visual test is not expected to
-outperform the conventional test. Meanwhile, the experiment II is a
-scenario where the conventional test is an approximate test, in which
-the visual test may have a chance to match the performance of the
-conventional test. The experiment III is designed for collecting human
-responses to null lineups such that the parameter \(\alpha\) in Equation
+of linear regression diagnostics. Two types of departures, namely
+non-linearity and heteroskedasticity, are considered with the
+corresponding data generating process being designed for experiment I
+and II. The experiment III is designed for collecting human responses to
+null lineups such that the parameter \(\alpha\) in Equation
 \ref{eq:pvalue-beta-binomial} can be estimated. Overall, we plan to
@@
 residuals}\label{simulating-departures-from-good-residuals}}
-
-Two types of departures, namely non-linearity and heteroskedasticity,
-are considered with the corresponding data generating process being
-designed for experiment I and II.
 

### 2023-01-30 paper_comparison/paper_comparison/paper_comparison.Rmd
clarify the usage of the experimental data

@@
 As mentioned in Section \ref{calculating-p-values-for-the-visual-test}, $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using null lineups.
-Hence, 36 lineups with all combinations of $n$ and $X_{raw}$ and three replications are included in experiment III.
+Hence, three replications of 36 lineups with all combinations of $n$ and fitted value distributions are included in experiment III.
 In these lineups, the data of the data plot is generated from a model with zero effect size, while the data of the 19 null plots are generated using the same simulation method discussed in Section \ref{sampling-from-the-null-distribution}.
@@
 
-There are `r 160 * 18`, `r 150 * 18` and `r 133 * 18` lineups evaluation made by 160, 150 and 133 subjects recruited for experiment I, II and III respectively.
-In the total of `r nrow(filter(vi_survey, !attention_check))` lineup evaluations, `r nrow(filter(vi_survey, type == "polynomial", !attention_check))` use lineups produced by the non-linearity model, `r nrow(filter(vi_survey, type == "heteroskedasticity", !attention_check))` use lineups produced by the heteroskedasticity model, and `r nrow(filter(vi_survey, null_lineup, !attention_check))` use null lineups.
-Besides, there are `r nrow(filter(vi_survey, attention_check))` attention checks not included in the analysis.
+There are `r 160 * 18`, `r 150 * 18` and `r 133 * 18 - nrow(filter(vi_survey, null_lineup, !attention_check))` lineups evaluation made by 160, 150 and 133 subjects recruited for experiment I, II and III respectively.
+In the total of `r nrow(filter(vi_survey, !attention_check))` lineup evaluations, `r nrow(filter(vi_survey, type == "polynomial", !attention_check))` use lineups produced by the non-linearity model, and `r nrow(filter(vi_survey, type == "heteroskedasticity", !attention_check))` use lineups produced by the heteroskedasticity model.
+Besides, there are `r nrow(filter(vi_survey, attention_check))` attention checks and `r nrow(filter(vi_survey, null_lineup, !attention_check))` evaluations on null lineups needed for the estimate of $\alpha$ not included in the analysis.
 The collated dataset is provided in `vi_survey` of the `visage` `R` package.

### 2023-01-30 paper_comparison/paper_comparison/paper_comparison.tex
clarify the usage of the experimental data

@@
 \(\alpha\) used in Equation \ref{eq:pvalue-beta-binomial} needs to be
-estimated using null lineups. Hence, 36 lineups with all combinations of
-\(n\) and \(X_{raw}\) and three replications are included in experiment
-III. In these lineups, the data of the data plot is generated from a
-model with zero effect size, while the data of the 19 null plots are
-generated using the same simulation method discussed in Section
-\ref{sampling-from-the-null-distribution}. This generation procedure
-differs from the canonical Rorschach lineup procedure, which requires
-that all 20 plots are generated from the null hypothesis. However, these
-lineups serve the same fundamental purpose: to assess the number of
-visually interesting plots generated under the null hypothesis.
+estimated using null lineups. Hence, three replications of 36 lineups
+with all combinations of \(n\) and fitted value distributions are
+included in experiment III. In these lineups, the data of the data plot
+is generated from a model with zero effect size, while the data of the
+19 null plots are generated using the same simulation method discussed
+in Section \ref{sampling-from-the-null-distribution}. This generation
+procedure differs from the canonical Rorschach lineup procedure, which
+requires that all 20 plots are generated from the null hypothesis.
+However, these lineups serve the same fundamental purpose: to assess the
+number of visually interesting plots generated under the null
+hypothesis.
 
@@
 
-There are 2880, 2700 and 2394 lineups evaluation made by 160, 150 and
+There are 2880, 2700 and 1674 lineups evaluation made by 160, 150 and
 133 subjects recruited for experiment I, II and III respectively. In the
 total of 7974 lineup evaluations, 3744 use lineups produced by the
-non-linearity model, 4230 use lineups produced by the heteroskedasticity
-model, and 720 use null lineups. Besides, there are 886 attention checks
+non-linearity model, and 4230 use lineups produced by the
+heteroskedasticity model. Besides, there are 886 attention checks and
+720 evaluations on null lineups needed for the estimate of \(\alpha\)
 not included in the analysis. The collated dataset is provided in

### 2023-01-30 paper_comparison/paper_comparison/paper_comparison.Rmd
update the non-linearity shape caption

@@
 
-```{r poly-power-uniform-j, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups containing four different non-linearity shapes.
-Power curves with higher order of non-linearity are drawn with deeper colours."}
+```{r poly-power-uniform-j, fig.height = 4, fig.cap = 'Power of conventional tests and visual tests on lineups containing four different non-linearity shapes.
+Power curves with higher order of non-linearity are drawn with deeper colours.
+The default RESET tests under-perform significantly in detecting "Tripe-U" shape.
+To achieve a similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test.
+But this means the order needs to be known prior to testing.'}
 

### 2023-01-30 paper_comparison/paper_comparison/paper_comparison.tex
update the non-linearity shape caption

@@
 \caption{Power of conventional tests and visual tests on lineups containing four different non-linearity shapes.
-Power curves with higher order of non-linearity are drawn with deeper colours.}\label{fig:poly-power-uniform-j}
+Power curves with higher order of non-linearity are drawn with deeper colours.
+The default RESET tests under-perform significantly in detecting "Tripe-U" shape.
+To achieve the similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test.
+But this means the order needs to be known prior to testing.}\label{fig:poly-power-uniform-j}
 \end{figure}

### 2023-01-30 paper_comparison/paper_comparison/paper_comparison.Rmd
fix the different x range issue

@@
 
-Shape of the non-linearity pattern is another factor contributes to the non-linearity model. With different non-linearity patterns, it is expected that the visual test will respond variously.
+Shape of the non-linearity pattern is a primary factor contributes to the non-linearity model. With different non-linearity patterns, it is expected that the visual test will respond variously.
 
@@
 
-poly_conv_sim %>%
+dat_poly_j <- poly_conv_sim %>%
+  filter(x_dist == "uniform") %>%
   mutate(log_effect_size = log(effect_size)) %>%
@@
   mutate(shape = factor(shape, levels = c("U", "S", "M", "Triple-U"))) %>%
-  select(shape, log_effect_size, reject, conv_or_not) %>%
+  select(shape, log_effect_size, reject, conv_or_not)
+
+map_df(levels(dat_poly_j$shape), function(x) {
+
+  # Adjust the range of x
+  result <- data.frame(log_effect_size = seq(min(dat_poly_j$log_effect_size),
+                                             max(dat_poly_j$log_effect_size), 
+                                             0.01))
+
+  # Manually fit the model
+  conv_mod <- glm(reject ~ log_effect_size, 
+                  family = binomial(),
+                  data = filter(dat_poly_j, 
+                                shape == x, 
+                                conv_or_not == "Conventional"))
+
+  # Manually fit the model
+  visual_mod <- glm(reject ~ log_effect_size, 
+                  family = binomial(),
+                  data = filter(dat_poly_j, 
+                                shape == x, 
+                                conv_or_not == "Visual"))
+
+  # Get the prediction
+  result <- result %>%
+    mutate(Conventional = predict(conv_mod, 
+                                  newdata = result, 
+                                  type = "response"),
+           Visual = predict(visual_mod, 
+                            newdata = result, 
+                            type = "response"),
+           shape = x) %>%
+    pivot_longer(Conventional:Visual, names_to = "conv_or_not")
+
+  result
+}) %>%
   ggplot() +
-  geom_smooth(aes(log_effect_size, reject, col = shape), 
-              se = FALSE, 
-              method = "glm",
-              method.args = list(family = binomial())) +
+  geom_smooth(aes(log_effect_size, value, col = shape), se = FALSE, span = 0.03) +
   xlab(expression(log[e] (Effect_size))) +
@@
 ## Shape of heteroskedasticity
+
+We also investigate the impact of different heteroskedasticity patterns.
+In theory, the "Left-triangle" and the "Right-triangle" should have little difference from the point of view of a Breusch-Pagan test.
+According to Figure \ref{fig:heter-power-uniform-a}, this is indeed the case.
+Similarly, these two shapes should have the same visual test power, unless the orientation of the visual pattern is a non-negligible factor from human perspective.
+However, it can be observed from Figure \ref{fig:heter-power-uniform-a} that the power curve for the "Left-triangle" shape is constantly higher than the one for the "Right-triangle" shape, indicating a potential favour of orientation by human, which is worth to be explored in future studies.
+
+The "Butterfly" shape
+
+```{r heter-power-uniform-a, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$."}
+
+heter_conv_sim %>%
+  filter(x_dist == "uniform") %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  mutate(reject = as.numeric(BP_p_value < 0.05)) %>%
+  mutate(a = c("Left-triangle", "Butterfly", "Right-triangle")[a + 2]) %>%
+  mutate(conv_or_not = "Conventional") %>%
+  bind_rows(vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(x_dist == "uniform", type == "heteroskedasticity") %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  mutate(reject = as.numeric(p_value < 0.05)) %>%
+  mutate(a = c("Left-triangle", "Butterfly", "Right-triangle")[a + 2]) %>%
+  mutate(conv_or_not = "Visual")) %>%
+  mutate(a = factor(a, levels = c("Left-triangle", "Butterfly", "Right-triangle"))) %>%
+  select(a, log_effect_size, reject, conv_or_not) %>%
+  ggplot() +
+  geom_smooth(aes(log_effect_size, reject, col = a), 
+              se = FALSE, 
+              method = "glm",
+              method.args = list(family = binomial())) +
+
+  # Theme
+  xlab(expression(log[e] (Effect_size))) +
+  ylab("Power") +
+  scale_color_manual(values = rcartocolor::carto_pal(3, "Earth")) +
+  theme_light() +
+  theme(legend.position = "bottom") +
+  facet_wrap(~conv_or_not) +
+  labs(col = "Heteroskedasticity shape")
+```
 

### 2023-01-30 paper_comparison/paper_comparison/paper_comparison.tex
fix the different x range issue

@@
 
-Shape of the non-linearity pattern is another factor contributes to the
-non-linearity model. With different non-linearity patterns, it is
+Shape of the non-linearity pattern is a primary factor contributes to
+the non-linearity model. With different non-linearity patterns, it is
 expected that the visual test will respond variously.
@@
 The default RESET tests under-perform significantly in detecting "Tripe-U" shape.
-To achieve the similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test.
+To achieve a similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test.
 But this means the order needs to be known prior to testing.}\label{fig:poly-power-uniform-j}
@@
 heteroskedasticity}\label{shape-of-heteroskedasticity}}
+
+We also investigate the impact of different heteroskedasticity patterns.
+In theory, the ``Left-triangle'' and the ``Right-triangle'' should have
+little difference from the point of view of a Breusch-Pagan test.
+According to Figure \ref{fig:heter-power-uniform-a}, this is indeed the
+case. Similarly, these two shapes should have the same visual test
+power, unless the orientation of the visual pattern is a non-negligible
+factor from human perspective. However, it can be observed from Figure
+\ref{fig:heter-power-uniform-a} that the power curve for the
+``Left-triangle'' shape is constantly higher than the one for the
+``Right-triangle'' shape, indicating a potential favour of orientation
+by human, which is worth to be explored in future studies.
+
+The ``Butterfly'' shape
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/heter-power-uniform-a-1}
+
+}
+
+\caption{Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$.}\label{fig:heter-power-uniform-a}
+\end{figure}
 

### 2023-01-30 paper_comparison/paper_comparison/paper_comparison.Rmd
fix another plot with the same problem

@@
 
-heter_conv_sim %>%
+dat_heter_a <- heter_conv_sim %>%
   filter(x_dist == "uniform") %>%
@@
   mutate(a = factor(a, levels = c("Left-triangle", "Butterfly", "Right-triangle"))) %>%
-  select(a, log_effect_size, reject, conv_or_not) %>%
+  select(a, log_effect_size, reject, conv_or_not)
+
+map_df(levels(dat_heter_a$a), function(x) {
+
+  # Adjust the range of x
+  result <- data.frame(log_effect_size = seq(min(dat_heter_a$log_effect_size),
+                                             max(dat_heter_a$log_effect_size), 
+                                             0.01))
+
+  # Manually fit the model
+  conv_mod <- glm(reject ~ log_effect_size, 
+                  family = binomial(),
+                  data = filter(dat_heter_a, 
+                                a == x, 
+                                conv_or_not == "Conventional"))
+
+  # Manually fit the model
+  visual_mod <- glm(reject ~ log_effect_size, 
+                  family = binomial(),
+                  data = filter(dat_heter_a, 
+                                a == x, 
+                                conv_or_not == "Visual"))
+
+  # Get the prediction
+  result <- result %>%
+    mutate(Conventional = predict(conv_mod, 
+                                  newdata = result, 
+                                  type = "response"),
+           Visual = predict(visual_mod, 
+                            newdata = result, 
+                            type = "response"),
+           a = x) %>%
+    pivot_longer(Conventional:Visual, names_to = "conv_or_not")
+
+  result
+}) %>%
   ggplot() +
-  geom_smooth(aes(log_effect_size, reject, col = a), 
-              se = FALSE, 
-              method = "glm",
-              method.args = list(family = binomial())) +
-
-  # Theme
+  geom_smooth(aes(log_effect_size, value, col = a), se = FALSE, span = 0.03) +
   xlab(expression(log[e] (Effect_size))) +

### 2023-01-31 paper_comparison/paper_comparison/paper_comparison.Rmd
update heter shape section

@@
 
-Shape of the non-linearity pattern is a primary factor contributes to the non-linearity model. With different non-linearity patterns, it is expected that the visual test will respond variously.
-
-According to Figure \ref{fig:poly-power-uniform-j}, conventional tests have almost identical power in testing the "U", the "S" and the "M" shape, except for testing the "Triple-U" shape, which have significantly lower power.
+A primary factor contributes to the non-linearity model is the shape of the non-linearity.
+According to Figure \ref{fig:poly-power-uniform-j}, conventional tests have almost identical power in testing the "U", the "S" and the "M" shape, except for testing the "Triple-U" shape, which have significant lower power.
 To understand why this is, one needs to return to the way the RESET test is applied.
@@
 For visual tests, we expect the "U" shape will be the easiest one to be detected by subjects followed by the "S", "M" and "Triple-U" shape.
-From Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power curves are mostly aligned with our expectation, except for the "M" shape.
-Interestingly, the "M" shape is as easy to be detected as the "U" shape, which implies the readability of the shape do not strictly follow the degree of the polynomials.
+From Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power curves are mostly aligned with our expectation, except for the "M" shape, which is as easy to be detected as the "U" shape.
+This implies the readability of the shape do not strictly follow the degree of the polynomials.
 
@@
 
-We also investigate the impact of different heteroskedasticity patterns.
-In theory, the "Left-triangle" and the "Right-triangle" should have little difference from the point of view of a Breusch-Pagan test.
-According to Figure \ref{fig:heter-power-uniform-a}, this is indeed the case.
-Similarly, these two shapes should have the same visual test power, unless the orientation of the visual pattern is a non-negligible factor from human perspective.
+We also investigate the impact of different heteroskedasticity shapes on power of conventional tests and visual tests.
+In theory, the "Left-triangle" and the "Right-triangle" shapes are functionally identical from the point of view of a Breusch-Pagan test.
+As shown in Figure \ref{fig:heter-power-uniform-a}, this is indeed the case where little difference between the power curves can be perceived.
+Similarly, visual tests will have the same power of detecting these two shapes if they are equally likely to be identified.
 However, it can be observed from Figure \ref{fig:heter-power-uniform-a} that the power curve for the "Left-triangle" shape is constantly higher than the one for the "Right-triangle" shape, indicating a potential favour of orientation by human, which is worth to be explored in future studies.
 
-The "Butterfly" shape
+The "Butterfly" shape on the other hand seems to be
 

### 2023-01-31 paper_comparison/paper_comparison/paper_comparison.Rmd
update the same section

@@
 
-The "Butterfly" shape on the other hand seems to be
+The chance of detecting the "Butterfly" shape is roughly the same as the chance of detecting the "Left-triangle" shape in both tests.
+The only noticeable difference from Figure \ref{fig:heter-power-uniform-a} is the Breusch-Pagan test has slightly higher power for detecting this shape compared to other two shapes when the effect size is low.
 

### 2023-01-31 paper_comparison/paper_comparison/paper_comparison.tex
update heter shape section
update the same section

--- 2023-01-30 paper_comparison/paper_comparison/paper_comparison.tex
@@
 
-Shape of the non-linearity pattern is another factor contributes to the
-non-linearity model. With different non-linearity patterns, it is
-expected that the visual test will respond variously.
-
-According to Figure \ref{fig:poly-power-uniform-j}, conventional tests
-have almost identical power in testing the ``U'', the ``S'' and the
-``M'' shape, except for testing the ``Triple-U'' shape, which have
-significantly lower power. To understand why this is, one needs to
-return to the way the RESET test is applied. It requires a parameter
+A primary factor contributes to the non-linearity model is the shape of
+the non-linearity. According to Figure \ref{fig:poly-power-uniform-j},
+conventional tests have almost identical power in testing the ``U'', the
+``S'' and the ``M'' shape, except for testing the ``Triple-U'' shape,
+which have significant lower power. To understand why this is, one needs
+to return to the way the RESET test is applied. It requires a parameter
 indicating degree of fitted values to test for, and the recommendation
@@
 that the power curves are mostly aligned with our expectation, except
-for the ``M'' shape. Interestingly, the ``M'' shape is as easy to be
-detected as the ``U'' shape, which implies the readability of the shape
-do not strictly follow the degree of the polynomials.
+for the ``M'' shape, which is as easy to be detected as the ``U'' shape.
+This implies the readability of the shape do not strictly follow the
+degree of the polynomials.
 
@@
 heteroskedasticity}\label{shape-of-heteroskedasticity}}
+
+We also investigate the impact of different heteroskedasticity shapes on
+power of conventional tests and visual tests. In theory, the
+``Left-triangle'' and the ``Right-triangle'' shapes are functionally
+identical from the point of view of a Breusch-Pagan test. As shown in
+Figure \ref{fig:heter-power-uniform-a}, this is indeed the case where
+little difference between the power curves can be perceived. Similarly,
+visual tests will have the same power of detecting these two shapes if
+they are equally likely to be identified. However, it can be observed
+from Figure \ref{fig:heter-power-uniform-a} that the power curve for the
+``Left-triangle'' shape is constantly higher than the one for the
+``Right-triangle'' shape, indicating a potential favour of orientation
+by human, which is worth to be explored in future studies.
+
+The chance of detecting the ``Butterfly'' shape is roughly the same as
+the chance of detecting the ``Left-triangle'' shape in both tests. The
+only noticeable difference from Figure \ref{fig:heter-power-uniform-a}
+is the Breusch-Pagan test has slightly higher power for detecting this
+shape compared to other two shapes when the effect size is low.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/heter-power-uniform-a-1}
+
+}
+
+\caption{Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$.}\label{fig:heter-power-uniform-a}
+\end{figure}
 

### 2023-01-31 paper_comparison/paper_comparison/paper_comparison.Rmd
fill in conclusion

@@
 
-## Shape of non-linearity
+## Effect of shape of non-linearity
 
@@
 
-## Shape of heteroskedasticity
+## Effect of shape of heteroskedasticity
 
@@
 
-## Effect of other fitted value distributions
-
-### Fitted value distribution
+## Effect of fitted value distributions
 
@@
 The distribution of $X$, or consequently $\hat{Y}$, may however affect the ability to read any patterns in the residuals.
-<!-- Although the distribution of the regressor is classically not restricted by the assumptions of linear regression, it will have an impact on the fitted value distribution and the visual effect of the residual plot.--> The four distributions of fitted values, used in this experiment, were designed to examine this.
+The four distributions of fitted values, used in this experiment, were designed to examine this.
 
@@
 For non-linearity patterns, we do not observe significant difference in power of visual test between distributions, except for the discrete uniform distribution.
-It is possibility due to the fact that humans have difficulties in recognising visual patterns from discreteness, which makes the shape disconnected and incomplete (ref?).
+It could be due to the fact that humans have difficulties in recognising visual patterns from discreteness, which makes the shape disconnected and incomplete (ref?).
 In terms of conventional tests, the discrete uniform distribution is instead the one with the greatest power, followed by the lognormal distribution, the uniform distribution and the normal distribution.
@@
 
-### Shape of non-linearity (old)
-
-Shape of the non-linearity pattern is another factor contributed to the non-linearity model.
+<!-- ### Shape of non-linearity (old) -->
+
+<!-- Shape of the non-linearity pattern is another factor contributed to the non-linearity model.
 According to Figure \ref{fig:poly-power-j}, conventional tests are significantly less powerful in testing "Triple-U" shape compared to other three shapes, though it is still more sensitive than the visual test.
 In fact, among the lineups containing non-linearity patterns used in our experiments, around `r vi_survey %>% filter(!attention_check) %>% filter(!null_lineup) %>% filter(type == "polynomial") %>% group_by(unique_lineup_id) %>% summarise(across(everything(), first)) %>% count(a = p_value < 0.05, b = conventional_p_value < 0.05) %>% mutate(prop = n / sum(n) * 100) %>% filter(a, !b) %>% pull(prop) %>% round(digits = 1)`% (two our of 576) are rejected by the visual test, but accepted by the conventional test.
-
-Figure \ref{fig:poly-example} shows one of the two lineups where the visual test rejects the residual plot.
+-->
+
+<!-- Figure \ref{fig:poly-example} shows one of the two lineups where the visual test rejects the residual plot.
 The data plot, in position 18, is very obvious, and different from the nulls.
@@
 This illustrates the sensitivity of conventional testing to the parameters, and it also points to a limitation that one needs to know the data structure in order to set the parameters for the test.
-
-We expect the "U" shape will be the easiest one to be detected by subjects followed by the "S", "M" and "Triple-U" shape.
+-->
+
+<!-- We expect the "U" shape will be the easiest one to be detected by subjects followed by the "S", "M" and "Triple-U" shape.
 From Figure \ref{fig:poly-power-j}, it can be observed that the power curves are mostly aligned with our expectation, except for the "M" shape.
@@
 The readability of the shape do not strictly follow the degree of the polynomials.
-
-```{r poly-power-j, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups containing four different non-linearity shapes.
-Power curves with higher order of non-linearity are drawn with deeper colours."}
-
-poly_conv_sim %>%
-  filter(x_dist == "uniform") %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  mutate(reject = as.numeric(RESET4_p_value < 0.05)) %>%
-  mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>%
-  mutate(conv_or_not = "Conventional") %>%
-  bind_rows(vi_survey %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(x_dist == "uniform", type == "polynomial") %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  mutate(reject = as.numeric(p_value < 0.05)) %>%
-  mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>%
-  mutate(conv_or_not = "Visual")) %>%
-  mutate(shape = factor(shape, levels = c("U", "S", "M", "Triple-U"))) %>%
-  select(shape, log_effect_size, reject, conv_or_not) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject, col = shape), 
-              se = FALSE, 
-              method = "glm",
-              method.args = list(family = binomial())) +
-  xlab(expression(log[e] (Effect_size))) +
-  ylab("Power") +
-  scale_color_manual(values = rcartocolor::carto_pal(4, "TealGrn")) +
-  theme_light() +
-  theme(legend.position = "bottom") +
-  facet_wrap(~conv_or_not) +
-  labs(col = "Non-linearity shape")
-```
-
-```{r poly-example, fig.height = 7, fig.width = 7, fig.cap = "Example of a non-linearity that is rejected by the visual test but not by the RESET test.
+-->
+
+<!-- ```{r poly-power-j, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups containing four different non-linearity shapes.
+Power curves with higher order of non-linearity are drawn with deeper colours."} -->
+
+<!-- poly_conv_sim %>% -->
+<!--   filter(x_dist == "uniform") %>% -->
+<!--   mutate(log_effect_size = log(effect_size)) %>% -->
+<!--   mutate(reject = as.numeric(RESET4_p_value < 0.05)) %>% -->
+<!--   mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>% -->
+<!--   mutate(conv_or_not = "Conventional") %>% -->
+<!--   bind_rows(vi_survey %>% -->
+<!--   filter(!attention_check) %>% -->
+<!--   filter(!null_lineup) %>% -->
+<!--   filter(x_dist == "uniform", type == "polynomial") %>% -->
+<!--   mutate(log_effect_size = log(effect_size)) %>% -->
+<!--   mutate(reject = as.numeric(p_value < 0.05)) %>% -->
+<!--   mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>% -->
+<!--   mutate(conv_or_not = "Visual")) %>% -->
+<!--   mutate(shape = factor(shape, levels = c("U", "S", "M", "Triple-U"))) %>% -->
+<!--   select(shape, log_effect_size, reject, conv_or_not) %>% -->
+<!--   ggplot() + -->
+<!--   geom_smooth(aes(log_effect_size, reject, col = shape),  -->
+<!--               se = FALSE,  -->
+<!--               method = "glm", -->
+<!--               method.args = list(family = binomial())) + -->
+<!--   xlab(expression(log[e] (Effect_size))) + -->
+<!--   ylab("Power") + -->
+<!--   scale_color_manual(values = rcartocolor::carto_pal(4, "TealGrn")) + -->
+<!--   theme_light() + -->
+<!--   theme(legend.position = "bottom") + -->
+<!--   facet_wrap(~conv_or_not) + -->
+<!--   labs(col = "Non-linearity shape") -->
+<!-- ``` -->
+
+<!-- ```{r poly-example, fig.height = 7, fig.width = 7, fig.cap = "Example of a non-linearity that is rejected by the visual test but not by the RESET test.
 The data plot at panel 18 exhibits strong non-linear pattern.
@@
 The RESET test using the higher order polynomial parameter could be used, but this would change it's performance for lower order polynomials.
-The only way for it to be sensitive to these type of departure is to know the order prior to testing!"}
-# plot poly-275
-VI_MODEL$plot_lineup(vi_lineup$poly_275$data, 
-                     remove_grid_line = TRUE, 
-                     theme = theme_light(),
-                     remove_axis = TRUE)
-```
-
-### Shape of heteroskedasticity (old)
-
-We also investigate the impact of the shape of heteroskedasticity on the power of different tests.
+The only way for it to be sensitive to these type of departure is to know the order prior to testing!"} -->
+<!-- # plot poly-275 -->
+<!-- VI_MODEL$plot_lineup(vi_lineup$poly_275$data,  -->
+<!--                      remove_grid_line = TRUE,  -->
+<!--                      theme = theme_light(), -->
+<!--                      remove_axis = TRUE) -->
+<!-- ``` -->
+
+<!-- ### Shape of heteroskedasticity (old) -->
+
+<!-- We also investigate the impact of the shape of heteroskedasticity on the power of different tests.
 The corresponding result is presented in Figure \ref{fig:heter-power-a}.
@@
 It implies the current residual plot design is not robust to the orientation of the visual pattern, which is worth to be explored in future research.
-
-```{r heter-power-a, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$."}
-
-heter_conv_sim %>%
-  filter(x_dist == "uniform") %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  mutate(reject = as.numeric(BP_p_value < 0.05)) %>%
-  mutate(a = c("Left-triangle", "Butterfly", "Right-triangle")[a + 2]) %>%
-  mutate(conv_or_not = "Conventional") %>%
-  bind_rows(vi_survey %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(x_dist == "uniform", type == "heteroskedasticity") %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  mutate(reject = as.numeric(p_value < 0.05)) %>%
-  mutate(a = c("Left-triangle", "Butterfly", "Right-triangle")[a + 2]) %>%
-  mutate(conv_or_not = "Visual")) %>%
-  mutate(a = factor(a, levels = c("Left-triangle", "Butterfly", "Right-triangle"))) %>%
-  select(a, log_effect_size, reject, conv_or_not) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject, col = a), 
-              se = FALSE, 
-              method = "glm",
-              method.args = list(family = binomial())) +
-
-  # Draw bootstrap for Left-triangle
-  geom_ribbon(data = ~get_boot_conf(filter(.x, 
-                                           conv_or_not == "Visual", 
-                                           a == "Left-triangle"),
-                                    log_effect_size,
-                                    reject,
-                                    times = 1000) %>%
-                mutate(conv_or_not = "Visual"),
-              aes(x = log_effect_size, 
-                  ymin = lower, 
-                  ymax = upper),
-              fill = rcartocolor::carto_pal(3, "Earth")[1],
-              alpha = 0.2) +
-
-  # Draw bootstrap for Right-triangle
-  geom_ribbon(data = ~get_boot_conf(filter(.x, 
-                                           conv_or_not == "Visual", 
-                                           a == "Right-triangle"),
-                                    log_effect_size,
-                                    reject,
-                                    times = 1000) %>%
-                mutate(conv_or_not = "Visual"),
-              aes(x = log_effect_size, 
-                  ymin = lower, 
-                  ymax = upper),
-              fill = rcartocolor::carto_pal(3, "Earth")[3],
-              alpha = 0.2) +
-
-  # Theme
-  xlab(expression(log[e] (Effect_size))) +
-  ylab("Power") +
-  scale_color_manual(values = rcartocolor::carto_pal(3, "Earth")) +
-  theme_light() +
-  theme(legend.position = "bottom") +
-  facet_wrap(~conv_or_not) +
-  labs(col = "Heteroskedasticity shape")
-```
+-->
+
+<!-- ```{r heter-power-a, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$."} -->
+
+<!-- heter_conv_sim %>% -->
+<!--   filter(x_dist == "uniform") %>% -->
+<!--   mutate(log_effect_size = log(effect_size)) %>% -->
+<!--   mutate(reject = as.numeric(BP_p_value < 0.05)) %>% -->
+<!--   mutate(a = c("Left-triangle", "Butterfly", "Right-triangle")[a + 2]) %>% -->
+<!--   mutate(conv_or_not = "Conventional") %>% -->
+<!--   bind_rows(vi_survey %>% -->
+<!--   filter(!attention_check) %>% -->
+<!--   filter(!null_lineup) %>% -->
+<!--   filter(x_dist == "uniform", type == "heteroskedasticity") %>% -->
+<!--   mutate(log_effect_size = log(effect_size)) %>% -->
+<!--   mutate(reject = as.numeric(p_value < 0.05)) %>% -->
+<!--   mutate(a = c("Left-triangle", "Butterfly", "Right-triangle")[a + 2]) %>% -->
+<!--   mutate(conv_or_not = "Visual")) %>% -->
+<!--   mutate(a = factor(a, levels = c("Left-triangle", "Butterfly", "Right-triangle"))) %>% -->
+<!--   select(a, log_effect_size, reject, conv_or_not) %>% -->
+<!--   ggplot() + -->
+<!--   geom_smooth(aes(log_effect_size, reject, col = a),  -->
+<!--               se = FALSE,  -->
+<!--               method = "glm", -->
+<!--               method.args = list(family = binomial())) + -->
+
+<!--   # Draw bootstrap for Left-triangle -->
+<!--   geom_ribbon(data = ~get_boot_conf(filter(.x,  -->
+<!--                                            conv_or_not == "Visual",  -->
+<!--                                            a == "Left-triangle"), -->
+<!--                                     log_effect_size, -->
+<!--                                     reject, -->
+<!--                                     times = 1000) %>% -->
+<!--                 mutate(conv_or_not = "Visual"), -->
+<!--               aes(x = log_effect_size,  -->
+<!--                   ymin = lower,  -->
+<!--                   ymax = upper), -->
+<!--               fill = rcartocolor::carto_pal(3, "Earth")[1], -->
+<!--               alpha = 0.2) + -->
+
+<!--   # Draw bootstrap for Right-triangle -->
+<!--   geom_ribbon(data = ~get_boot_conf(filter(.x,  -->
+<!--                                            conv_or_not == "Visual",  -->
+<!--                                            a == "Right-triangle"), -->
+<!--                                     log_effect_size, -->
+<!--                                     reject, -->
+<!--                                     times = 1000) %>% -->
+<!--                 mutate(conv_or_not = "Visual"), -->
+<!--               aes(x = log_effect_size,  -->
+<!--                   ymin = lower,  -->
+<!--                   ymax = upper), -->
+<!--               fill = rcartocolor::carto_pal(3, "Earth")[3], -->
+<!--               alpha = 0.2) + -->
+
+<!--   # Theme -->
+<!--   xlab(expression(log[e] (Effect_size))) + -->
+<!--   ylab("Power") + -->
+<!--   scale_color_manual(values = rcartocolor::carto_pal(3, "Earth")) + -->
+<!--   theme_light() + -->
+<!--   theme(legend.position = "bottom") + -->
+<!--   facet_wrap(~conv_or_not) + -->
+<!--   labs(col = "Heteroskedasticity shape") -->
+<!-- ``` -->
 
 # Conclusion
+
+This paper has demonstrated that conventional residual-based statistical tests are more sensitive to departures from assumptions for residuals than visual tests evaluated by humans in the diagnostics of linear regression models.
+Three experiments are conducted with two simulation models designed for non-linearity and heteroskedasticity model defects being applied.
+Methods for obtaining the $p$-value of visual tests proposed by @vanderplas2021statistical are extended to allowed for multiple selections in a lineup protocol.
+
+Residual plots rejected by visual tests are rarely not rejected by conventional tests.
+Throughout the three human subject experiments we conducted, departures of residuals that are of little practical sense to humans will often be alarmed by conventional tests, supporting the common guidance given by experts to utilize graphical representations of data in model diagnostics.
+
+Conventional tests relying on
+
+such as RESET test
 

### 2023-01-31 paper_comparison/paper_comparison/paper_comparison.tex
fill in conclusion

@@
 
-\hypertarget{shape-of-non-linearity}{%
-\subsection{Shape of non-linearity}\label{shape-of-non-linearity}}
+\hypertarget{effect-of-shape-of-non-linearity}{%
+\subsection{Effect of shape of
+non-linearity}\label{effect-of-shape-of-non-linearity}}
 
@@
 
-\hypertarget{shape-of-heteroskedasticity}{%
-\subsection{Shape of
-heteroskedasticity}\label{shape-of-heteroskedasticity}}
+\hypertarget{effect-of-shape-of-heteroskedasticity}{%
+\subsection{Effect of shape of
+heteroskedasticity}\label{effect-of-shape-of-heteroskedasticity}}
 
@@
 
-\hypertarget{effect-of-other-fitted-value-distributions}{%
-\subsection{Effect of other fitted value
-distributions}\label{effect-of-other-fitted-value-distributions}}
-
-\hypertarget{fitted-value-distribution}{%
-\subsubsection{Fitted value
-distribution}\label{fitted-value-distribution}}
+\hypertarget{effect-of-fitted-value-distributions}{%
+\subsection{Effect of fitted value
+distributions}\label{effect-of-fitted-value-distributions}}
 
@@
 power of visual test between distributions, except for the discrete
-uniform distribution. It is possibility due to the fact that humans have
+uniform distribution. It could be due to the fact that humans have
 difficulties in recognising visual patterns from discreteness, which
@@
 
-\hypertarget{shape-of-non-linearity-old}{%
-\subsubsection{Shape of non-linearity
-(old)}\label{shape-of-non-linearity-old}}
-
-Shape of the non-linearity pattern is another factor contributed to the
-non-linearity model. According to Figure \ref{fig:poly-power-j},
-conventional tests are significantly less powerful in testing
-``Triple-U'' shape compared to other three shapes, though it is still
-more sensitive than the visual test. In fact, among the lineups
-containing non-linearity patterns used in our experiments, around 0.3\%
-(two our of 576) are rejected by the visual test, but accepted by the
-conventional test.
-
-Figure \ref{fig:poly-example} shows one of the two lineups where the
-visual test rejects the residual plot. The data plot, in position 18, is
-very obvious, and different from the nulls. The visual test is correct
-to reject. To understand why this is, one needs to return to the way the
-RESET test is applied. It requires a parameter indicating degree of
-fitted values to test for, and the recommendation is to generically use
-four (\citet{ramsey_tests_1969}). However, the ``Triple-U'' shape
-constructed from the Hermite polynomials use power up to 18. Both
-lineups rejected by the visual test but not the RESET are of
-``Triple-U''. If the RESET test had been applied using the power up to
-six, all the decisions for these lineups would have been to reject. The
-recommendation of the polynomial power for the RESET should be revised,
-perhaps. This illustrates the sensitivity of conventional testing to the
-parameters, and it also points to a limitation that one needs to know
-the data structure in order to set the parameters for the test.
-
-We expect the ``U'' shape will be the easiest one to be detected by
-subjects followed by the ``S'', ``M'' and ``Triple-U'' shape. From
-Figure \ref{fig:poly-power-j}, it can be observed that the power curves
-are mostly aligned with our expectation, except for the ``M'' shape.
-Interestingly, the ``M'' shape is as easy to be detected as the ``U''
-shape. The readability of the shape do not strictly follow the degree of
-the polynomials.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/poly-power-j-1}
-
-}
-
-\caption{Power of conventional tests and visual tests on lineups containing four different non-linearity shapes.
-Power curves with higher order of non-linearity are drawn with deeper colours.}\label{fig:poly-power-j}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/poly-example-1}
-
-}
-
-\caption{Example of a non-linearity that is rejected by the visual test but not by the RESET test.
-The data plot at panel 18 exhibits strong non-linear pattern.
-It has effect size $= 3.96$.
-Interestingly, all the lineups in this group are higher order polynomials (Triple-U), and it is apparent that the default RESET test under-performs.
-The RESET test using the higher order polynomial parameter could be used, but this would change it's performance for lower order polynomials.
-The only way for it to be sensitive to these type of departure is to know the order prior to testing!}\label{fig:poly-example}
-\end{figure}
-
-\hypertarget{shape-of-heteroskedasticity-old}{%
-\subsubsection{Shape of heteroskedasticity
-(old)}\label{shape-of-heteroskedasticity-old}}
-
-We also investigate the impact of the shape of heteroskedasticity on the
-power of different tests. The corresponding result is presented in
-Figure \ref{fig:heter-power-a}. In theory, the ``Left-triangle'' and the
-``Right-triangle'' should have little difference from the point of view
-of a Breusch-Pagan test. According to Figure \ref{fig:heter-power-a},
-this is indeed the case. And these two shapes should have the same
-visual test power, unless the orientation of the visual pattern is a
-non-negligible factor from human perspective. However, by checking the
-bootstrap confidence intervals, it can be observed from Figure
-\ref{fig:heter-power-a} that when \(3 < log(E) < 4\), the
-``Left-triangle'' power curve is significantly greater than the
-``Right-triangle'' power curve. This phenomenon is particularly
-interesting. It implies the current residual plot design is not robust
-to the orientation of the visual pattern, which is worth to be explored
-in future research.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/heter-power-a-1}
-
-}
-
-\caption{Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$.}\label{fig:heter-power-a}
-\end{figure}
-
 \hypertarget{conclusion}{%
 \section{Conclusion}\label{conclusion}}
+
+This paper has demonstrated that conventional residual-based statistical
+tests are more sensitive to departures from assumptions for residuals
+than visual tests evaluated by humans in the diagnostics of linear
+regression models. Three experiments are conducted with two simulation
+models designed for non-linearity and heteroskedasticity model defects
+being applied. Methods for obtaining the \(p\)-value of visual tests
+proposed by \citet{vanderplas2021statistical} are extended to allowed
+for multiple selections in a lineup protocol.
+
+Residual plots rejected by visual tests are rarely not rejected by
+conventional tests. Throughout the three human subject experiments we
+conducted, departures of residuals that are of little practical sense to
+humans will often be alarmed by conventional tests, supporting the
+common guidance given by experts to utilize graphical representations of
+data in model diagnostics.
+
+Conventional tests relying on
+
+such as RESET test
 

### 2023-01-31 paper_comparison/paper_comparison/paper_comparison.Rmd
fix the x range issue

@@
 
-map_df(levels(dat_poly_j$shape), function(x) {
-
-  # Adjust the range of x
-  result <- data.frame(log_effect_size = seq(min(dat_poly_j$log_effect_size),
-                                             max(dat_poly_j$log_effect_size), 
-                                             0.01))
-
-  # Manually fit the model
-  conv_mod <- glm(reject ~ log_effect_size, 
+dat_poly_j %>%
+  mutate(es_min = min(log_effect_size), 
+         es_max = max(log_effect_size)) %>%
+  nest(data = c(log_effect_size, reject)) %>%
+  mutate(mod = map(data, 
+                   ~glm(reject ~ log_effect_size, 
                         family = binomial(), 
-                  data = filter(dat_poly_j, 
-                                shape == x, 
-                                conv_or_not == "Conventional"))
-
-  # Manually fit the model
-  visual_mod <- glm(reject ~ log_effect_size, 
-                  family = binomial(),
-                  data = filter(dat_poly_j, 
-                                shape == x, 
-                                conv_or_not == "Visual"))
-
-  # Get the prediction
-  result <- result %>%
-    mutate(Conventional = predict(conv_mod, 
-                                  newdata = result, 
-                                  type = "response"),
-           Visual = predict(visual_mod, 
-                            newdata = result, 
-                            type = "response"),
-           shape = x) %>%
-    pivot_longer(Conventional:Visual, names_to = "conv_or_not")
-
+                        data = .x))) %>%
+  mutate(reject = map(mod, function(mod) {
+    result <- data.frame(log_effect_size = seq(es_min[1], es_max[1], 0.01))
+    result$reject <- predict(mod, type = "response", newdata = result)
     result
-}) %>%
+  })) %>%
+  select(-es_min, -es_max, -data, -mod) %>%
+  unnest(reject) %>%
   ggplot() +
-  geom_smooth(aes(log_effect_size, value, col = shape), se = FALSE, span = 0.03) +
+  geom_smooth(aes(log_effect_size, reject, col = shape), se = FALSE, span = 0.03) +
   xlab(expression(log[e] (Effect_size))) +
@@
 
-map_df(levels(dat_heter_a$a), function(x) {
-
-  # Adjust the range of x
-  result <- data.frame(log_effect_size = seq(min(dat_heter_a$log_effect_size),
-                                             max(dat_heter_a$log_effect_size), 
-                                             0.01))
-
 # Manually fit the model
-  conv_mod <- glm(reject ~ log_effect_size, 
+dat_heter_a %>%
+  mutate(es_min = min(log_effect_size), 
+         es_max = max(log_effect_size)) %>%
+  nest(data = c(log_effect_size, reject)) %>%
+  mutate(mod = map(data, 
+                   ~glm(reject ~ log_effect_size, 
                         family = binomial(), 
-                  data = filter(dat_heter_a, 
-                                a == x, 
-                                conv_or_not == "Conventional"))
-
-  # Manually fit the model
-  visual_mod <- glm(reject ~ log_effect_size, 
-                  family = binomial(),
-                  data = filter(dat_heter_a, 
-                                a == x, 
-                                conv_or_not == "Visual"))
-
-  # Get the prediction
-  result <- result %>%
-    mutate(Conventional = predict(conv_mod, 
-                                  newdata = result, 
-                                  type = "response"),
-           Visual = predict(visual_mod, 
-                            newdata = result, 
-                            type = "response"),
-           a = x) %>%
-    pivot_longer(Conventional:Visual, names_to = "conv_or_not")
-
+                        data = .x))) %>%
+  mutate(reject = map(mod, function(mod) {
+    result <- data.frame(log_effect_size = seq(es_min[1], es_max[1], 0.01))
+    result$reject <- predict(mod, type = "response", newdata = result)
     result
-}) %>%
+  })) %>%
+  select(-es_min, -es_max, -data, -mod) %>%
+  unnest(reject) %>%
   ggplot() +
-  geom_smooth(aes(log_effect_size, value, col = a), se = FALSE, span = 0.03) +
+  geom_smooth(aes(log_effect_size, reject, col = a), se = FALSE, span = 0.03) +
   xlab(expression(log[e] (Effect_size))) +
@@
 # and one for visual tests.
-list(
+
+dat_x_dist <- list(
   poly_conv_sim %>% 
@@
     mutate(type = "non-linearity") %>%
-  select(type, x_dist, log_effect_size, reject, conv_or_not)
-,
+    select(type, x_dist, log_effect_size, reject, conv_or_not),
+
   heter_conv_sim %>% 
@@
     mutate(type = "heteroskedasticity") %>%
-  select(type, x_dist, log_effect_size, reject, conv_or_not)
-,
+    select(type, x_dist, log_effect_size, reject, conv_or_not),
+
   vi_survey %>%
@@
                                     "Lognormal", 
-                                    "Discrete"))) %>%
+                                    "Discrete")))
+
+# Fit the model manually
+dat_x_dist %>%
+  mutate(es_min = min(log_effect_size), 
+         es_max = max(log_effect_size)) %>%
+  nest(data = c(log_effect_size, reject)) %>%
+  mutate(mod = map(data, 
+                   ~glm(reject ~ log_effect_size, 
+                        family = binomial(), 
+                        data = .x))) %>%
+  mutate(reject = map(mod, function(mod) {
+    result <- data.frame(log_effect_size = seq(es_min[1], es_max[1], 0.01))
+    result$reject <- predict(mod, type = "response", newdata = result)
+    result
+  })) %>%
+  select(-es_min, -es_max, -data, -mod) %>%
+  unnest(reject) %>%
 
@@
               col = "grey", 
-              method = "glm", 
-              method.args = list(family = binomial), 
               se = FALSE, 
+              span = 0.03,
               size = 3) +
@@
   geom_smooth(aes(log_effect_size, reject, col = x_dist), 
-              method = "glm", 
-              method.args = list(family = binomial), 
+              span = 0.03,
               se = FALSE, 
@@
 
+1. summarise paper content
+
 This paper has demonstrated that conventional residual-based statistical tests are more sensitive to departures from assumptions for residuals than visual tests evaluated by humans in the diagnostics of linear regression models.
@@
 
-Residual plots rejected by visual tests are rarely not rejected by conventional tests.
-Throughout the three human subject experiments we conducted, departures of residuals that are of little practical sense to humans will often be alarmed by conventional tests, supporting the common guidance given by experts to utilize graphical representations of data in model diagnostics.
-
-Conventional tests relying on
-
-such as RESET test
-
-1. summarise paper content
 2. briefly describe what was learned
+
+Throughout the three human subject experiments we conducted, residual plots rejected by visual tests are rarely not rejected by conventional tests, but departures of residuals that are of little practical sense to humans will often be alarmed by conventional tests, supporting the common guidance given by experts to utilize graphical representations of data in model diagnostics.
+
+We have also found that residual-based conventional tests relying on a user-parameterized test formula, such as the RESET test, will suffer from a decrease in power for detecting certain residual departures intended to be tested but not covered by the test formula.
+This can not be avoided without prior knowledge of their functional forms, which is usually obtained by plotting the residuals.
+On the other hand, visual tests do not rely on such information and a collection of departures can be tested simultaneously.
+
+Besides, fitted value distributions
+
+with different
+
+symmetric, evenly
+
 3. how does this apply more broadly than simple regression
+
 4. where does this lead? Computer vision residual plot reading

### 2023-01-31 paper_comparison/paper_comparison/paper_comparison.tex
fix the x range issue

@@
 
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  summarise paper content
+\end{enumerate}
+
 This paper has demonstrated that conventional residual-based statistical
@@
 
-Residual plots rejected by visual tests are rarely not rejected by
-conventional tests. Throughout the three human subject experiments we
-conducted, departures of residuals that are of little practical sense to
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
+\tightlist
+\item
+  briefly describe what was learned
+\end{enumerate}
+
+Throughout the three human subject experiments we conducted, residual
+plots rejected by visual tests are rarely not rejected by conventional
+tests, but departures of residuals that are of little practical sense to
 humans will often be alarmed by conventional tests, supporting the
@@
 
-Conventional tests relying on
-
-such as RESET test
+We have also found that residual-based conventional tests relying on a
+user-parameterized test formula, such as the RESET test, will suffer
+from a decrease in power for detecting certain residual departures
+intended to be tested but not covered by the test formula. This can not
+be avoided without prior knowledge of their functional forms, which is
+usually obtained by plotting the residuals. On the other hand, visual
+tests do not rely on such information and a collection of departures can
+be tested simultaneously.
+
+Besides, fitted value distributions
+
+with different
+
+symmetric, evenly
 
@@
 \def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  summarise paper content
-\item
-  briefly describe what was learned
+\setcounter{enumi}{2}
 \item

### 2023-02-06 paper_comparison/paper_comparison/paper_comparison.Rmd
use "data collection period" to replace "experiment"

@@
 
-Three experiments are conducted to investigate the difference between conventional hypothesis testing and visual inference in the application of linear regression diagnostics.
-Two types of departures, namely non-linearity and heteroskedasticity, are considered with the corresponding data generating process being designed for experiment I and II.
-The experiment III is designed for collecting human responses to null lineups such that the parameter $\alpha$ in Equation \ref{eq:pvalue-beta-binomial} can be estimated.
-Overall, we plan to collect `r vi_survey  %>% filter(!attention_check) %>% nrow()` evaluations on `r vi_survey  %>% filter(!attention_check) %>% pull(unique_lineup_id) %>% unique() %>% length()` unique lineups performed by `r vi_survey  %>% filter(!attention_check) %>% count(exp, set) %>% nrow()` subjects throughout three experiments.
+1. experiments -> data collection periods
+
+To investigate the difference between conventional hypothesis testing and visual inference in the application of linear regression diagnostics, an experiment is conducted in three data collection periods.
+Two types of departures, namely non-linearity and heteroskedasticity, are considered with the corresponding data generating process being designed for data collection periods I and II.
+The data collection period III is designed for collecting human responses to null lineups such that the parameter $\alpha$ in Equation \ref{eq:pvalue-beta-binomial} can be estimated.
+Overall, we plan to collect `r vi_survey  %>% filter(!attention_check) %>% nrow()` evaluations on `r vi_survey  %>% filter(!attention_check) %>% pull(unique_lineup_id) %>% unique() %>% length()` unique lineups performed by `r vi_survey  %>% filter(!attention_check) %>% count(exp, set) %>% nrow()` subjects throughout three data collection periods.
 
@@
 
-Experiment I is designed to study the ability of human subjects to detect the effect of a non-linear term $\boldsymbol{z}$ constructed using Hermite polynomials on random vector $\boldsymbol{x}$ formulated as
+Data collection period I is designed to study the ability of human subjects to detect the effect of a non-linear term $\boldsymbol{z}$ constructed using Hermite polynomials on random vector $\boldsymbol{x}$ formulated as
 
@@
                booktabs = TRUE, 
-               caption  = 'Levels of the factors used in experiments I, II, III.', 
+               caption  = 'Levels of the factors used in data collection periods I, II, III.', 
                label = "model-factor-table" ) %>%
@@
 
-```{r different-shape-of-herimite, fig.height = 2, fig.cap = "Polynomial forms generated for the residual plots used in experiment I.
+```{r different-shape-of-herimite, fig.height = 2, fig.cap = "Polynomial forms generated for the residual plots used in data collection period I.
 The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$."}
@@
 
-Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups used in experiment I.
+Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups used in data collection period I.
 This lineup is produced by the non-linearity model with $j = 6$ and uniform fitted values.
@@
 
-```{r different-n, fig.height = 2.67, fig.cap = "Number of observations $n$ have an impact on the strength of the signal. Three different values of $n$ are used in experiment I, II and III."}
+```{r different-n, fig.height = 2.67, fig.cap = "Number of observations $n$ have an impact on the strength of the signal.
+Three different values of $n$ are used in data collection periods I, II and III."}
 
@@
 ```{r different-sigma, fig.height = 2, fig.cap = "For the non-lienearity model, the standard deviation of the error term $\\sigma$ affects the strength of the signal.
-Four different values of $\\sigma$ are used in the experiment I."}
+Four different values of $\\sigma$ are used in the data collection period I."}
 
@@
 
-```{r example-poly-lineup, fig.height = 7, fig.width = 7, fig.cap = "One of the lineups containing non-linearity patterns used in experiment I.
+```{r example-poly-lineup, fig.height = 7, fig.width = 7, fig.cap = "One of the lineups containing non-linearity patterns used in data collection period I.
 Can you spot the most different plot?
@@
 
-Experiment II is designed to study the ability of human subjects to detect the appearance of a heteroskedasticity pattern under a simple linear regression model setting:
+Data collection period II is designed to study the ability of human subjects to detect the appearance of a heteroskedasticity pattern under a simple linear regression model setting:
 
@@
 
-```{r different-shape-of-heter, fig.height = 2.67, fig.cap = "Heteroskedasticity forms used in experiment II.
+```{r different-shape-of-heter, fig.height = 2.67, fig.cap = "Heteroskedasticity forms used in data collection period II.
 Three different shapes (a = -1, 0, 1) are used in the experiment to create left-triangle, butterfly and right-triangle shapes."}
@@
 
-```{r different-b, fig.height = 2, fig.cap = "Five different values of $b$ are used in experiment II to control the strength of the signal."}
+```{r different-b, fig.height = 2, fig.cap = "Five different values of $b$ are used in data collection period II to control the strength of the signal."}
 
@@
 
-```{r example-heter-lineup, fig.height = 7, fig.width = 7, fig.cap = "One of the lineups containing heteroskedasticity pattern used in experiment II.
+```{r example-heter-lineup, fig.height = 7, fig.width = 7, fig.cap = "One of the lineups containing heteroskedasticity pattern used in data collection period II.
 Can you spot the most different plot?
@@
 
-An example lineup of this model used in Experiment II is shown in Figure \ref{fig:example-heter-lineup} with $a = -1$ and $X_{raw} \sim U(-1, 1)$.
+An example lineup of this model used in data collection period II is shown in Figure \ref{fig:example-heter-lineup} with $a = -1$ and $X_{raw} \sim U(-1, 1)$.
 The data plot location is $2^4 + 2$.
@@
 
-### Factors common to both experiments
+### Factors common to both data collection periods
 
@@
 As summarised in Table \ref{tab:model-factor-table}, three additional parameters $n$, $\sigma$ and $b$ are used to control the strength of the signal so that different difficulty levels of lineups are generated, and therefore, the estimated power curve will be smooth and continuous.
-Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ are used in experiment I and II respectively.
+Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ are used in data collection periods I and II respectively.
 Figure \ref{fig:different-sigma} and \ref{fig:different-b} demonstrate the impact of these two parameters.
@@
 
-Three different sample sizes are used (n = 50, 100, 300) in all three experiments.
+Three different sample sizes are used (n = 50, 100, 300) in all three data collection periods.
 It can be observed from Figure \ref{fig:different-n} that with fewer data points drawn in a residual plot, the visual pattern is more difficult to be detected.
@@
 And to ensure the quality of the survey data, two lineups with obvious visual patterns are included as attention checks.
-Thus, $576 \times 5 / (20-2) = `r 576 * 5 / 18`$ and $540 \times 5 / (20-2) = `r 540 * 5 / 18`$ subjects are recruited to satisfy the design of the experiment I and experiment II respectively.
+Thus, $576 \times 5 / (20-2) = `r 576 * 5 / 18`$ and $540 \times 5 / (20-2) = `r 540 * 5 / 18`$ subjects are recruited to satisfy the design of the data collection period I and II respectively.
 
 As mentioned in Section \ref{calculating-p-values-for-the-visual-test}, $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using null lineups.
-Hence, three replications of 36 lineups with all combinations of $n$ and fitted value distributions are included in experiment III.
+Hence, three replications of 36 lineups with all combinations of $n$ and fitted value distributions are included in data collection period III.
 In these lineups, the data of the data plot is generated from a model with zero effect size, while the data of the 19 null plots are generated using the same simulation method discussed in Section \ref{sampling-from-the-null-distribution}.
@@
 However, presenting only these lineups to subjects are considered to be bad practices as subjects will lose interest quickly. 
-Therefore, we plan to collect 6 more evaluations on the 279 lineups with uniform fitted values, result in $(36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3 \times 6) / (20-2) = `r (36 * 20 + (4 * 4 * 3 + 3 * 5 * 3) * 3 * 6) / 18`$ subjects recruited for experiment III.
+Therefore, we plan to collect 6 more evaluations on the 279 lineups with uniform fitted values, result in $(36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3 \times 6) / (20-2) = `r (36 * 20 + (4 * 4 * 3 + 3 * 5 * 3) * 3 * 6) / 18`$ subjects recruited for data collection period III.
 
@@
 
-Subjects for all three experiments are recruited from an crowdsourcing platform called Prolific [@palan2018prolific].
+Subjects for all three data collection periods are recruited from an crowdsourcing platform called Prolific [@palan2018prolific].
 Prescreening procedure is applied during the recruitment, subjects are required to be fluent in English, with $98\%$ minimum approval rate and 10 minimum submissions in other studies.
@@
 No subject are shown the same lineup twice.
-Information about preferred pronoun, age group, education, and previous experience in visual experiment are also collected.
+Information about preferred pronoun, age group, education, and previous experience in visual experiments are also collected.
 A subject's submission is only accepted if the data plot is identified for at least one attention check.
@@
 
-There are `r 160 * 18`, `r 150 * 18` and `r 133 * 18 - nrow(filter(vi_survey, null_lineup, !attention_check))` lineups evaluation made by 160, 150 and 133 subjects recruited for experiment I, II and III respectively.
+There are `r 160 * 18`, `r 150 * 18` and `r 133 * 18 - nrow(filter(vi_survey, null_lineup, !attention_check))` lineups evaluation made by 160, 150 and 133 subjects recruited for data collection periods I, II and III respectively.
 In the total of `r nrow(filter(vi_survey, !attention_check))` lineup evaluations, `r nrow(filter(vi_survey, type == "polynomial", !attention_check))` use lineups produced by the non-linearity model, and `r nrow(filter(vi_survey, type == "heteroskedasticity", !attention_check))` use lineups produced by the heteroskedasticity model.
@@
 
+```{r}
+vi_survey %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
+  filter(x_dist == "uniform") %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  mutate(reject = as.numeric(p_value < 0.05)) %>%
+  mutate(reject_only_5  = as.numeric(p_value_only_5 < 0.05)) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  ggplot() +
+  geom_smooth(aes(log_effect_size, reject), 
+              method = "glm", 
+              method.args = list(family = binomial()),
+              se = FALSE) +
+  geom_smooth(aes(log_effect_size, reject_only_5), 
+              method = "glm", 
+              method.args = list(family = binomial()),
+              col = "red",
+              se = FALSE) +
+  facet_wrap(~type)
+```
+
 ```{r different-x-dist-poly-power, fig.cap = "Comparison of power on lineups with different fitted value distributions, for conventional tests and visual tests.
@@
 For both types of tests power does change.
-The discrete fitted value distribution substantially lowers the power for non-linearity patterns in residual plots.
-For heteroskedasticity patterns, normal and lognormal distributions produce lower power."}
+The discrete fitted value distribution substantially lowers the power of visual tests for non-linearity patterns in residual plots.
+For heteroskedasticity patterns, normal and lognormal distributions produce lower power for both types of tests."}
 
@@
   # Bold the uniform smooth line
-  geom_smooth(data = ~filter(.x, x_dist == "Uniform"),
-              aes(log_effect_size, reject), 
-              col = "grey", 
-              se = FALSE, 
-              span = 0.03,
-              size = 3) +
+  # geom_smooth(data = ~filter(.x, x_dist == "Uniform"),
+  #             aes(log_effect_size, reject), 
+  #             col = "grey", 
+  #             se = FALSE, 
+  #             span = 0.03,
+  #             size = 3) +
 
   # Draw all smooth lines
-  geom_smooth(aes(log_effect_size, reject, col = x_dist), 
+  geom_smooth(data = ~filter(.x, x_dist != "Uniform"),
+              aes(log_effect_size, reject, col = x_dist), 
               span = 0.03,
@@
               size = 1) +
+
+  geom_smooth(data = vi_survey %>%
+                filter(!null_lineup) %>%
+                filter(!attention_check) %>%
+                filter(x_dist == "uniform") %>%
+                group_by(unique_lineup_id) %>%
+                summarise(across(everything(), first)) %>%
+                mutate(reject = as.numeric(p_value < 0.05)) %>%
+                mutate(reject_only_5  = as.numeric(p_value_only_5 < 0.05)) %>%
+                mutate(log_effect_size = log(effect_size)) %>%
+                mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+                mutate(conv_or_not = "Visual"),
+              aes(log_effect_size, reject_only_5, col = "Uniform"),
+              method = "glm",
+              method.args = list(family = binomial()),
+              se = FALSE
+              ) +
 
@@
 This paper has demonstrated that conventional residual-based statistical tests are more sensitive to departures from assumptions for residuals than visual tests evaluated by humans in the diagnostics of linear regression models.
-Three experiments are conducted with two simulation models designed for non-linearity and heteroskedasticity model defects being applied.
+An experiment is conducted in three data collection periods with two simulation models designed for non-linearity and heteroskedasticity model defects being applied.
 Methods for obtaining the $p$-value of visual tests proposed by @vanderplas2021statistical are extended to allowed for multiple selections in a lineup protocol.
@@
 
-Throughout the three human subject experiments we conducted, residual plots rejected by visual tests are rarely not rejected by conventional tests, but departures of residuals that are of little practical sense to humans will often be alarmed by conventional tests, supporting the common guidance given by experts to utilize graphical representations of data in model diagnostics.
-
-We have also found that residual-based conventional tests relying on a user-parameterized test formula, such as the RESET test, will suffer from a decrease in power for detecting certain residual departures intended to be tested but not covered by the test formula.
+Throughout the three data collection periods of the human subject experiment we conducted, residual plots rejected by visual tests are rarely not rejected by conventional tests, but departures of residuals that are of little practical sense to humans will often be alarmed by conventional tests, supporting the common guidance given by experts to utilize graphical representations of data in model diagnostics.
+
+We have also found that residual-based conventional tests relying on a user-parameterized test formula, such as the RESET test, will suffer from a significant decrease in power for detecting certain residual departures intended to be tested but not covered by the test formula.
 This can not be avoided without prior knowledge of their functional forms, which is usually obtained by plotting the residuals.
@@
 
-Besides, fitted value distributions
-
-with different
-
-symmetric, evenly
+fitted value distribution
+
+orientation of heter shape
+
+<!-- Besides, the power of both tests vary for different fitted value distributions.  -->
+
+<!-- The ideal fitted value distribution for a visual test is the uniform distribution. If this can not be satisfied, a continuous -->
+
+<!-- continuous and symmetric -->
+
+<!-- distributions that are continuous, symmetric, or evenly distributed are also welcomed. -->
+
+<!-- properties that a uniform distribution holds, such as continuous, symmetric, and evenly distributed  -->
+
+<!-- Depending on the underlying residual departures embedded in the residual plot,  -->
+
+<!-- In the two controlled scenarios, the conventional tests favours discrete uniform distribution -->
+
+<!-- fitted value distributions have a strong impact on the  -->
+
+<!-- some residual departures are more difficult to be detected under a skewed fitted value distribution.  -->
+
+<!-- continuous, symmetric and evenly distributed. -->
+
+<!-- with different  -->
+
+<!-- symmetric, evenly  -->
 
 3. how does this apply more broadly than simple regression
+
+recommend residual vs fitted values
+a. measure uncertainty
+b. systemically miss some structure
 
@@
 \newpage
+
+1. put #eval in appendix
+2. examine batch effect in appendix
+3. compute the power of RESET test with different orders (considering dof) (focus on simple pattern possibly?)

### 2023-02-06 paper_comparison/paper_comparison/paper_comparison.Rmd
add 1000 boot smooth lines for uniform (5)

@@
 
-```{r}
+```{r get-5-eval, cache = TRUE}
+dat_uniform_5 <- map_df(1:1000, function(boot_id) {
   vi_survey %>%
@@
   group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  mutate(reject = as.numeric(p_value < 0.05)) %>%
-  mutate(reject_only_5  = as.numeric(p_value_only_5 < 0.05)) %>%
+  slice_sample(n = 5) %>%
+  ungroup() %>%
+  calc_p_value_multi(lineup_id = unique_lineup_id,
+                     detect = detect,
+                     n_sel = num_selection,
+                     alpha = alpha) %>%
+  mutate(boot_id = boot_id)
+}) %>%
+  left_join(select(vi_survey, unique_lineup_id, effect_size, x_dist, type)) %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
   mutate(log_effect_size = log(effect_size)) %>%
-  ggplot() +
-  geom_smooth(aes(log_effect_size, reject), 
-              method = "glm", 
-              method.args = list(family = binomial()),
-              se = FALSE) +
-  geom_smooth(aes(log_effect_size, reject_only_5), 
-              method = "glm", 
-              method.args = list(family = binomial()),
-              col = "red",
-              se = FALSE) +
-  facet_wrap(~type)
+  mutate(reject = as.numeric(p_value <= 0.05)) %>%
+  mutate(conv_or_not = "Visual") %>%
+  select(-unique_lineup_id, -p_value, -effect_size)
+
 ```
@@
     summarise(across(everything(), first)) %>%
+    filter(x_dist != "uniform") %>%
     mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
@@
                                     "Discrete")))
+
+dat_uniform_5_pred <- dat_uniform_5 %>%
+  mutate(es_min = min(dat_x_dist$log_effect_size), 
+         es_max = max(dat_x_dist$log_effect_size)) %>%
+  nest(data = c(log_effect_size, reject)) %>%
+  mutate(mod = map(data, 
+                   ~glm(reject ~ log_effect_size, 
+                        family = binomial(), 
+                        data = .x))) %>%
+  mutate(reject = map(mod, function(mod) {
+    result <- data.frame(log_effect_size = seq(es_min[1], es_max[1], 0.01))
+    result$reject <- predict(mod, type = "response", newdata = result)
+    result
+  })) %>%
+  select(-es_min, -es_max, -data, -mod) %>%
+  unnest(reject)
 
@@
 
-  # Bold the uniform smooth line
-  # geom_smooth(data = ~filter(.x, x_dist == "Uniform"),
-  #             aes(log_effect_size, reject), 
-  #             col = "grey", 
-  #             se = FALSE, 
-  #             span = 0.03,
-  #             size = 3) +
-
   # Draw all smooth lines
-  geom_smooth(data = ~filter(.x, x_dist != "Uniform"),
-              aes(log_effect_size, reject, col = x_dist), 
+  geom_smooth(aes(log_effect_size, reject, col = x_dist),
               span = 0.03,
@@
 
-  geom_smooth(data = vi_survey %>%
-                filter(!null_lineup) %>%
-                filter(!attention_check) %>%
-                filter(x_dist == "uniform") %>%
-                group_by(unique_lineup_id) %>%
-                summarise(across(everything(), first)) %>%
-                mutate(reject = as.numeric(p_value < 0.05)) %>%
-                mutate(reject_only_5  = as.numeric(p_value_only_5 < 0.05)) %>%
-                mutate(log_effect_size = log(effect_size)) %>%
-                mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-                mutate(conv_or_not = "Visual"),
-              aes(log_effect_size, reject_only_5, col = "Uniform"),
-              method = "glm",
-              method.args = list(family = binomial()),
-              se = FALSE
-              ) +
+  # Draw boot uniform (5 eval) lines
+  stat_smooth(data = dat_uniform_5_pred,
+              geom = "line",
+              aes(log_effect_size, reject, col = "Uniform", group = boot_id),
+              span = 0.03,
+              se = FALSE,
+              alpha = 0.01) +
 

### 2023-02-06 paper_comparison/paper_comparison/paper_comparison.tex
add 1000 boot smooth lines for uniform (5)

@@
 
-Three experiments are conducted to investigate the difference between
-conventional hypothesis testing and visual inference in the application
-of linear regression diagnostics. Two types of departures, namely
-non-linearity and heteroskedasticity, are considered with the
-corresponding data generating process being designed for experiment I
-and II. The experiment III is designed for collecting human responses to
-null lineups such that the parameter \(\alpha\) in Equation
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  experiments -\textgreater{} data collection periods
+\end{enumerate}
+
+To investigate the difference between conventional hypothesis testing
+and visual inference in the application of linear regression
+diagnostics, an experiment is conducted in three data collection
+periods. Two types of departures, namely non-linearity and
+heteroskedasticity, are considered with the corresponding data
+generating process being designed for data collection periods I and II.
+The data collection period III is designed for collecting human
+responses to null lineups such that the parameter \(\alpha\) in Equation
 \ref{eq:pvalue-beta-binomial} can be estimated. Overall, we plan to
 collect 7974 evaluations on 1152 unique lineups performed by 443
-subjects throughout three experiments.
+subjects throughout three data collection periods.
 
@@
 
-Experiment I is designed to study the ability of human subjects to
-detect the effect of a non-linear term \(\boldsymbol{z}\) constructed
-using Hermite polynomials on random vector \(\boldsymbol{x}\) formulated
-as
+Data collection period I is designed to study the ability of human
+subjects to detect the effect of a non-linear term \(\boldsymbol{z}\)
+constructed using Hermite polynomials on random vector
+\(\boldsymbol{x}\) formulated as
 
@@
 
-\caption{\label{tab:model-factor-table}Levels of the factors used in experiments I, II, III.}
+\caption{\label{tab:model-factor-table}Levels of the factors used in data collection periods I, II, III.}
 \centering
@@
 
-\caption{Polynomial forms generated for the residual plots used in experiment I.
+\caption{Polynomial forms generated for the residual plots used in data collection period I.
 The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$.}\label{fig:different-shape-of-herimite}
@@
 Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups
-used in experiment I. This lineup is produced by the non-linearity model
-with \(j = 6\) and uniform fitted values. The data plot location is
-\(2^3 - 4\). All five subjects correctly identify the data plot from
-this lineup.
+used in data collection period I. This lineup is produced by the
+non-linearity model with \(j = 6\) and uniform fitted values. The data
+plot location is \(2^3 - 4\). All five subjects correctly identify the
+data plot from this lineup.
 
@@
 
-\caption{Number of observations $n$ have an impact on the strength of the signal. Three different values of $n$ are used in experiment I, II and III.}\label{fig:different-n}
+\caption{Number of observations $n$ have an impact on the strength of the signal. Three different values of $n$ are used in data collection periods I, II and III.}\label{fig:different-n}
 \end{figure}
@@
 \caption{For the non-lienearity model, the standard deviation of the error term $\sigma$ affects the strength of the signal.
-Four different values of $\sigma$ are used in the experiment I.}\label{fig:different-sigma}
+Four different values of $\sigma$ are used in the data collection period I.}\label{fig:different-sigma}
 \end{figure}
@@
 
-\caption{One of the lineups containing non-linearity patterns used in experiment I. Can you spot the most different plot? The data plot is positioned at $2^3 + 1$.}\label{fig:example-poly-lineup}
+\caption{One of the lineups containing non-linearity patterns used in data collection period I.
+Can you spot the most different plot?
+The data plot is positioned at $2^3 + 1$.}\label{fig:example-poly-lineup}
 \end{figure}
@@
 
-Experiment II is designed to study the ability of human subjects to
-detect the appearance of a heteroskedasticity pattern under a simple
-linear regression model setting:
+Data collection period II is designed to study the ability of human
+subjects to detect the appearance of a heteroskedasticity pattern under
+a simple linear regression model setting:
 
@@
 
-\caption{Heteroskedasticity forms used in experiment II.
+\caption{Heteroskedasticity forms used in data collection period II.
 Three different shapes (a = -1, 0, 1) are used in the experiment to create left-triangle, butterfly and right-triangle shapes.}\label{fig:different-shape-of-heter}
@@
 
-\caption{Five different values of $b$ are used in experiment II to control the strength of the signal.}\label{fig:different-b}
+\caption{Five different values of $b$ are used in data collection period II to control the strength of the signal.}\label{fig:different-b}
 \end{figure}
@@
 
-\caption{One of the lineups containing heteroskedasticity pattern used in experiment II.
+\caption{One of the lineups containing heteroskedasticity pattern used in data collection period II.
 Can you spot the most different plot?
@@
 
-An example lineup of this model used in Experiment II is shown in Figure
-\ref{fig:example-heter-lineup} with \(a = -1\) and
+An example lineup of this model used in data collection period II is
+shown in Figure \ref{fig:example-heter-lineup} with \(a = -1\) and
 \(X_{raw} \sim U(-1, 1)\). The data plot location is \(2^4 + 2\). Nine
@@
 
-\hypertarget{factors-common-to-both-experiments}{%
-\subsubsection{Factors common to both
-experiments}\label{factors-common-to-both-experiments}}
+\hypertarget{factors-common-to-both-data-collection-periods}{%
+\subsubsection{Factors common to both data collection
+periods}\label{factors-common-to-both-data-collection-periods}}
 
@@
 continuous. Parameter \(\sigma \in \{0.5, 1, 2, 4\}\) and
-\(b \in \{0.25, 1, 4, 16, 64\}\) are used in experiment I and II
-respectively. Figure \ref{fig:different-sigma} and \ref{fig:different-b}
-demonstrate the impact of these two parameters. A large value of
-\(\sigma\) will increase the variation of the error of the non-linearity
-model and decrease the visibility of the visual pattern. The parameter
-\(b\) controls the standard deviation of the error across the support of
-the regressor. Given \(x \neq a\), a larger value of \(b\) will lead to
-a larger ratio of the variance at \(x\) to the variance at
-\(x - a = 0\), making the visual pattern more obvious.
+\(b \in \{0.25, 1, 4, 16, 64\}\) are used in data collection periods I
+and II respectively. Figure \ref{fig:different-sigma} and
+\ref{fig:different-b} demonstrate the impact of these two parameters. A
+large value of \(\sigma\) will increase the variation of the error of
+the non-linearity model and decrease the visibility of the visual
+pattern. The parameter \(b\) controls the standard deviation of the
+error across the support of the regressor. Given \(x \neq a\), a larger
+value of \(b\) will lead to a larger ratio of the variance at \(x\) to
+the variance at \(x - a = 0\), making the visual pattern more obvious.
 
 Three different sample sizes are used (n = 50, 100, 300) in all three
-experiments. It can be observed from Figure \ref{fig:different-n} that
-with fewer data points drawn in a residual plot, the visual pattern is
-more difficult to be detected.
+data collection periods. It can be observed from Figure
+\ref{fig:different-n} that with fewer data points drawn in a residual
+plot, the visual pattern is more difficult to be detected.
 
@@
 \(540 \times 5 / (20-2) = 150\) subjects are recruited to satisfy the
-design of the experiment I and experiment II respectively.
+design of the data collection period I and II respectively.
 
@@
 with all combinations of \(n\) and fitted value distributions are
-included in experiment III. In these lineups, the data of the data plot
-is generated from a model with zero effect size, while the data of the
-19 null plots are generated using the same simulation method discussed
-in Section \ref{sampling-from-the-null-distribution}. This generation
-procedure differs from the canonical Rorschach lineup procedure, which
-requires that all 20 plots are generated from the null hypothesis.
-However, these lineups serve the same fundamental purpose: to assess the
-number of visually interesting plots generated under the null
-hypothesis.
+included in data collection period III. In these lineups, the data of
+the data plot is generated from a model with zero effect size, while the
+data of the 19 null plots are generated using the same simulation method
+discussed in Section \ref{sampling-from-the-null-distribution}. This
+generation procedure differs from the canonical Rorschach lineup
+procedure, which requires that all 20 plots are generated from the null
+hypothesis. However, these lineups serve the same fundamental purpose:
+to assess the number of visually interesting plots generated under the
+null hypothesis.
 
@@
 \((36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3 \times 6) / (20-2) = 133\)
-subjects recruited for experiment III.
+subjects recruited for data collection period III.
 
@@
 
-Subjects for all three experiments are recruited from an crowdsourcing
-platform called Prolific \citep{palan2018prolific}. Prescreening
-procedure is applied during the recruitment, subjects are required to be
-fluent in English, with \(98\%\) minimum approval rate and 10 minimum
-submissions in other studies.
+Subjects for all three data collection periods are recruited from an
+crowdsourcing platform called Prolific \citep{palan2018prolific}.
+Prescreening procedure is applied during the recruitment, subjects are
+required to be fluent in English, with \(98\%\) minimum approval rate
+and 10 minimum submissions in other studies.
 
@@
 about preferred pronoun, age group, education, and previous experience
-in visual experiment are also collected. A subject's submission is only
+in visual experiments are also collected. A subject's submission is only
 accepted if the data plot is identified for at least one attention
@@
 There are 2880, 2700 and 1674 lineups evaluation made by 160, 150 and
-133 subjects recruited for experiment I, II and III respectively. In the
-total of 7974 lineup evaluations, 3744 use lineups produced by the
-non-linearity model, and 4230 use lineups produced by the
-heteroskedasticity model. Besides, there are 886 attention checks and
-720 evaluations on null lineups needed for the estimate of \(\alpha\)
-not included in the analysis. The collated dataset is provided in
-\texttt{vi\_survey} of the \texttt{visage} \texttt{R} package.
+133 subjects recruited for data collection periods I, II and III
+respectively. In the total of 7974 lineup evaluations, 3744 use lineups
+produced by the non-linearity model, and 4230 use lineups produced by
+the heteroskedasticity model. Besides, there are 886 attention checks
+and 720 evaluations on null lineups needed for the estimate of
+\(\alpha\) not included in the analysis. The collated dataset is
+provided in \texttt{vi\_survey} of the \texttt{visage} \texttt{R}
+package.
 
@@
 For both types of tests power does change.
-The discrete fitted value distribution substantially lowers the power for non-linearity patterns in residual plots.
-For heteroskedasticity patterns, normal and lognormal distributions produce lower power.}\label{fig:different-x-dist-poly-power}
+The discrete fitted value distribution substantially lowers the power of visual tests for non-linearity patterns in residual plots.
+For heteroskedasticity patterns, normal and lognormal distributions produce lower power for both types of tests.}\label{fig:different-x-dist-poly-power}
 \end{figure}
@@
 than visual tests evaluated by humans in the diagnostics of linear
-regression models. Three experiments are conducted with two simulation
-models designed for non-linearity and heteroskedasticity model defects
-being applied. Methods for obtaining the \(p\)-value of visual tests
-proposed by \citet{vanderplas2021statistical} are extended to allowed
-for multiple selections in a lineup protocol.
+regression models. An experiment is conducted in three data collection
+periods with two simulation models designed for non-linearity and
+heteroskedasticity model defects being applied. Methods for obtaining
+the \(p\)-value of visual tests proposed by
+\citet{vanderplas2021statistical} are extended to allowed for multiple
+selections in a lineup protocol.
 
@@
 
-Throughout the three human subject experiments we conducted, residual
-plots rejected by visual tests are rarely not rejected by conventional
-tests, but departures of residuals that are of little practical sense to
-humans will often be alarmed by conventional tests, supporting the
-common guidance given by experts to utilize graphical representations of
-data in model diagnostics.
+Throughout the three data collection periods of the human subject
+experiment we conducted, residual plots rejected by visual tests are
+rarely not rejected by conventional tests, but departures of residuals
+that are of little practical sense to humans will often be alarmed by
+conventional tests, supporting the common guidance given by experts to
+utilize graphical representations of data in model diagnostics.
 
@@
 user-parameterized test formula, such as the RESET test, will suffer
-from a decrease in power for detecting certain residual departures
-intended to be tested but not covered by the test formula. This can not
-be avoided without prior knowledge of their functional forms, which is
-usually obtained by plotting the residuals. On the other hand, visual
-tests do not rely on such information and a collection of departures can
-be tested simultaneously.
-
-Besides, fitted value distributions
-
-with different
-
-symmetric, evenly
+from a significant decrease in power for detecting certain residual
+departures intended to be tested but not covered by the test formula.
+This can not be avoided without prior knowledge of their functional
+forms, which is usually obtained by plotting the residuals. On the other
+hand, visual tests do not rely on such information and a collection of
+departures can be tested simultaneously.
+
+fitted value distribution
+
+orientation of heter shape
 
@@
 \setcounter{enumi}{2}
+\tightlist
 \item
   how does this apply more broadly than simple regression
+\end{enumerate}
+
+recommend residual vs fitted values a. measure uncertainty b.
+systemically miss some structure
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{3}
+\tightlist
 \item
@@
 
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  put \#eval in appendix
+\item
+  examine batch effect in appendix
+\item
+  compute the power of RESET test with different orders (considering
+  dof) (focus on simple pattern possibly?)
+\end{enumerate}
+
 \bibliographystyle{tfcad}

### 2023-02-07 paper_comparison/paper_comparison/paper_comparison.Rmd
fix the fitted value distribution section

@@
   mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
   mutate(log_effect_size = log(effect_size)) %>%
@@
   select(-unique_lineup_id, -p_value, -effect_size)
-
 ```
@@
 ```{r different-x-dist-poly-power, fig.cap = "Comparison of power on lineups with different fitted value distributions, for conventional tests and visual tests.
-Each power curve of the uniform distribution is surrounded by a grey line to indicate it is the baseline in its group.
+For visual tests, all the power curves are estimated using five evaluations on each lineup.
+For lineups with uniform fitted value distribution, the five evaluations are randomly sampled from the total eleven evaluations.
+The sampling process has been repeated for 1000 times and the corresponding power curves are drawn with transparent grey lines.
 For both types of tests power does change.
 The discrete fitted value distribution substantially lowers the power of visual tests for non-linearity patterns in residual plots.
-For heteroskedasticity patterns, normal and lognormal distributions produce lower power for both types of tests."}
+For heteroskedasticity patterns, the lognormal distribution produce lower power for both types of tests."}
 
@@
 
-  # Draw all smooth lines
-  geom_smooth(aes(log_effect_size, reject, col = x_dist),
-              span = 0.03,
-              se = FALSE,
-              size = 1) +
-
   # Draw boot uniform (5 eval) lines
@@
 
+  # Draw all smooth lines
+  geom_smooth(aes(log_effect_size, reject, col = x_dist),
+              span = 0.03,
+              se = FALSE,
+              size = 1) +
+
   # Theme
   theme_light() +
-  scale_color_manual(values = rcartocolor::carto_pal(4, "Safe")[c(4,1,3,2)]) +
+  scale_color_manual(values = rcartocolor::carto_pal(4, "Safe")[c(2,1,3,4)]) +
   scale_size_manual(values = c(1, 2, 3, 4)) +

### 2023-02-07 paper_comparison/paper_comparison/paper_comparison.tex
fix the fitted value distribution section

@@
 \caption{Comparison of power on lineups with different fitted value distributions, for conventional tests and visual tests.
-Each power curve of the uniform distribution is surrounded by a grey line to indicate it is the baseline in its group.
+For visual tests, all the power curves are estimated using five evaluations on each lineup.
+For lineups with uniform fitted value distribution, the five evaluations are randomly sampled from the total eleven evaluations.
+The sampling process has been repeated for 1000 times and the corresponding power curves are drawn with transparent grey lines.
 For both types of tests power does change.

### 2023-02-07 paper_comparison/paper_comparison/paper_comparison.Rmd
update the text in the same section

@@
 It could be due to the fact that humans have difficulties in recognising visual patterns from discreteness, which makes the shape disconnected and incomplete (ref?).
-In terms of conventional tests, the discrete uniform distribution is instead the one with the greatest power, followed by the lognormal distribution, the uniform distribution and the normal distribution.
+In terms of conventional tests, the discrete uniform distribution is instead the one with the greatest power, followed by lognormal, uniform and the normal distribution.
 
@@
 This suggests the use of transformation of fitted values in reading residual plots.
-In the case of conventional tests, the discrete uniform distribution consistently perform well, followed by the uniform distribution, the normal distribution and the lognormal distribution.
-
-In regards of the type of the simulation model, the power of visual tests under different fitted value distributions are always lower than the corresponding conventional tests, which is consistent with the result reported in the previous sections.
+In the case of conventional tests, the discrete uniform distribution consistently perform well, followed by uniform, normal and the lognormal distribution.
+
+In regards of the type of the simulation model, the power of visual tests under different fitted value distributions are substantially lower than the corresponding conventional tests, which is consistent with the result reported in the previous sections.
 
@@
 The sampling process has been repeated for 1000 times and the corresponding power curves are drawn with transparent grey lines.
+The vetical dotted line is drawn to indicate where the visual pattern start to become recognisable.
 For both types of tests power does change.
@@
 
+  geom_vline(xintercept = 2, linetype = 2) +
+
   # Draw boot uniform (5 eval) lines
@@
 
-<!-- ### Shape of non-linearity (old) -->
-
-<!-- Shape of the non-linearity pattern is another factor contributed to the non-linearity model.
-According to Figure \ref{fig:poly-power-j}, conventional tests are significantly less powerful in testing "Triple-U" shape compared to other three shapes, though it is still more sensitive than the visual test.
-In fact, among the lineups containing non-linearity patterns used in our experiments, around `r vi_survey %>% filter(!attention_check) %>% filter(!null_lineup) %>% filter(type == "polynomial") %>% group_by(unique_lineup_id) %>% summarise(across(everything(), first)) %>% count(a = p_value < 0.05, b = conventional_p_value < 0.05) %>% mutate(prop = n / sum(n) * 100) %>% filter(a, !b) %>% pull(prop) %>% round(digits = 1)`% (two our of 576) are rejected by the visual test, but accepted by the conventional test.
--->
-
-<!-- Figure \ref{fig:poly-example} shows one of the two lineups where the visual test rejects the residual plot.
-The data plot, in position 18, is very obvious, and different from the nulls.
-The visual test is correct to reject.
-To understand why this is, one needs to return to the way the RESET test is applied.
-It requires a parameter indicating degree of fitted values to test for, and the recommendation is to generically use four (@ramsey_tests_1969).
-However, the "Triple-U" shape constructed from the Hermite polynomials use power up to 18.
-Both lineups rejected by the visual test but not the RESET are of "Triple-U".
-If the RESET test had been applied using the power up to six, all the decisions for these lineups would have been to reject.
-The recommendation of the polynomial power for the RESET should be revised, perhaps.
-This illustrates the sensitivity of conventional testing to the parameters, and it also points to a limitation that one needs to know the data structure in order to set the parameters for the test.
--->
-
-<!-- We expect the "U" shape will be the easiest one to be detected by subjects followed by the "S", "M" and "Triple-U" shape.
-From Figure \ref{fig:poly-power-j}, it can be observed that the power curves are mostly aligned with our expectation, except for the "M" shape.
-Interestingly, the "M" shape is as easy to be detected as the "U" shape.
-The readability of the shape do not strictly follow the degree of the polynomials.
--->
-
-<!-- ```{r poly-power-j, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups containing four different non-linearity shapes.
-Power curves with higher order of non-linearity are drawn with deeper colours."} -->
-
-<!-- poly_conv_sim %>% -->
-<!--   filter(x_dist == "uniform") %>% -->
-<!--   mutate(log_effect_size = log(effect_size)) %>% -->
-<!--   mutate(reject = as.numeric(RESET4_p_value < 0.05)) %>% -->
-<!--   mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>% -->
-<!--   mutate(conv_or_not = "Conventional") %>% -->
-<!--   bind_rows(vi_survey %>% -->
-<!--   filter(!attention_check) %>% -->
-<!--   filter(!null_lineup) %>% -->
-<!--   filter(x_dist == "uniform", type == "polynomial") %>% -->
-<!--   mutate(log_effect_size = log(effect_size)) %>% -->
-<!--   mutate(reject = as.numeric(p_value < 0.05)) %>% -->
-<!--   mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>% -->
-<!--   mutate(conv_or_not = "Visual")) %>% -->
-<!--   mutate(shape = factor(shape, levels = c("U", "S", "M", "Triple-U"))) %>% -->
-<!--   select(shape, log_effect_size, reject, conv_or_not) %>% -->
-<!--   ggplot() + -->
-<!--   geom_smooth(aes(log_effect_size, reject, col = shape),  -->
-<!--               se = FALSE,  -->
-<!--               method = "glm", -->
-<!--               method.args = list(family = binomial())) + -->
-<!--   xlab(expression(log[e] (Effect_size))) + -->
-<!--   ylab("Power") + -->
-<!--   scale_color_manual(values = rcartocolor::carto_pal(4, "TealGrn")) + -->
-<!--   theme_light() + -->
-<!--   theme(legend.position = "bottom") + -->
-<!--   facet_wrap(~conv_or_not) + -->
-<!--   labs(col = "Non-linearity shape") -->
-<!-- ``` -->
-
-<!-- ```{r poly-example, fig.height = 7, fig.width = 7, fig.cap = "Example of a non-linearity that is rejected by the visual test but not by the RESET test.
-The data plot at panel 18 exhibits strong non-linear pattern.
-It has effect size $= 3.96$.
-Interestingly, all the lineups in this group are higher order polynomials (Triple-U), and it is apparent that the default RESET test under-performs.
-The RESET test using the higher order polynomial parameter could be used, but this would change it's performance for lower order polynomials.
-The only way for it to be sensitive to these type of departure is to know the order prior to testing!"} -->
-<!-- # plot poly-275 -->
-<!-- VI_MODEL$plot_lineup(vi_lineup$poly_275$data,  -->
-<!--                      remove_grid_line = TRUE,  -->
-<!--                      theme = theme_light(), -->
-<!--                      remove_axis = TRUE) -->
-<!-- ``` -->
-
-<!-- ### Shape of heteroskedasticity (old) -->
-
-<!-- We also investigate the impact of the shape of heteroskedasticity on the power of different tests.
-The corresponding result is presented in Figure \ref{fig:heter-power-a}.
-In theory, the "Left-triangle" and the "Right-triangle" should have little difference from the point of view of a Breusch-Pagan test.
-According to Figure \ref{fig:heter-power-a}, this is indeed the case.
-And these two shapes should have the same visual test power, unless the orientation of the visual pattern is a non-negligible factor from human perspective.
-However, by checking the bootstrap confidence intervals, it can be observed from Figure \ref{fig:heter-power-a} that when $3 < log(E) < 4$, the "Left-triangle" power curve is significantly greater than the "Right-triangle" power curve.
-This phenomenon is particularly interesting.
-It implies the current residual plot design is not robust to the orientation of the visual pattern, which is worth to be explored in future research.
--->
-
-<!-- ```{r heter-power-a, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$."} -->
-
-<!-- heter_conv_sim %>% -->
-<!--   filter(x_dist == "uniform") %>% -->
-<!--   mutate(log_effect_size = log(effect_size)) %>% -->
-<!--   mutate(reject = as.numeric(BP_p_value < 0.05)) %>% -->
-<!--   mutate(a = c("Left-triangle", "Butterfly", "Right-triangle")[a + 2]) %>% -->
-<!--   mutate(conv_or_not = "Conventional") %>% -->
-<!--   bind_rows(vi_survey %>% -->
-<!--   filter(!attention_check) %>% -->
-<!--   filter(!null_lineup) %>% -->
-<!--   filter(x_dist == "uniform", type == "heteroskedasticity") %>% -->
-<!--   mutate(log_effect_size = log(effect_size)) %>% -->
-<!--   mutate(reject = as.numeric(p_value < 0.05)) %>% -->
-<!--   mutate(a = c("Left-triangle", "Butterfly", "Right-triangle")[a + 2]) %>% -->
-<!--   mutate(conv_or_not = "Visual")) %>% -->
-<!--   mutate(a = factor(a, levels = c("Left-triangle", "Butterfly", "Right-triangle"))) %>% -->
-<!--   select(a, log_effect_size, reject, conv_or_not) %>% -->
-<!--   ggplot() + -->
-<!--   geom_smooth(aes(log_effect_size, reject, col = a),  -->
-<!--               se = FALSE,  -->
-<!--               method = "glm", -->
-<!--               method.args = list(family = binomial())) + -->
-
-<!--   # Draw bootstrap for Left-triangle -->
-<!--   geom_ribbon(data = ~get_boot_conf(filter(.x,  -->
-<!--                                            conv_or_not == "Visual",  -->
-<!--                                            a == "Left-triangle"), -->
-<!--                                     log_effect_size, -->
-<!--                                     reject, -->
-<!--                                     times = 1000) %>% -->
-<!--                 mutate(conv_or_not = "Visual"), -->
-<!--               aes(x = log_effect_size,  -->
-<!--                   ymin = lower,  -->
-<!--                   ymax = upper), -->
-<!--               fill = rcartocolor::carto_pal(3, "Earth")[1], -->
-<!--               alpha = 0.2) + -->
-
-<!--   # Draw bootstrap for Right-triangle -->
-<!--   geom_ribbon(data = ~get_boot_conf(filter(.x,  -->
-<!--                                            conv_or_not == "Visual",  -->
-<!--                                            a == "Right-triangle"), -->
-<!--                                     log_effect_size, -->
-<!--                                     reject, -->
-<!--                                     times = 1000) %>% -->
-<!--                 mutate(conv_or_not = "Visual"), -->
-<!--               aes(x = log_effect_size,  -->
-<!--                   ymin = lower,  -->
-<!--                   ymax = upper), -->
-<!--               fill = rcartocolor::carto_pal(3, "Earth")[3], -->
-<!--               alpha = 0.2) + -->
-
-<!--   # Theme -->
-<!--   xlab(expression(log[e] (Effect_size))) + -->
-<!--   ylab("Power") + -->
-<!--   scale_color_manual(values = rcartocolor::carto_pal(3, "Earth")) + -->
-<!--   theme_light() + -->
-<!--   theme(legend.position = "bottom") + -->
-<!--   facet_wrap(~conv_or_not) + -->
-<!--   labs(col = "Heteroskedasticity shape") -->
-<!-- ``` -->
-
 # Conclusion

### 2023-02-07 paper_comparison/paper_comparison/paper_comparison.tex
update the text in the same section

@@
 conventional tests, the discrete uniform distribution is instead the one
-with the greatest power, followed by the lognormal distribution, the
-uniform distribution and the normal distribution.
+with the greatest power, followed by lognormal, uniform and the normal
+distribution.
 
@@
 of conventional tests, the discrete uniform distribution consistently
-perform well, followed by the uniform distribution, the normal
-distribution and the lognormal distribution.
+perform well, followed by uniform, normal and the lognormal
+distribution.
 
 In regards of the type of the simulation model, the power of visual
-tests under different fitted value distributions are always lower than
-the corresponding conventional tests, which is consistent with the
+tests under different fitted value distributions are substantially lower
+than the corresponding conventional tests, which is consistent with the
 result reported in the previous sections.
@@
 The sampling process has been repeated for 1000 times and the corresponding power curves are drawn with transparent grey lines.
+The vetical dotted line is drawn to indicate where the visual pattern start to become recognisable.
 For both types of tests power does change.
 The discrete fitted value distribution substantially lowers the power of visual tests for non-linearity patterns in residual plots.
-For heteroskedasticity patterns, normal and lognormal distributions produce lower power for both types of tests.}\label{fig:different-x-dist-poly-power}
+For heteroskedasticity patterns, the lognormal distribution produce lower power for both types of tests.}\label{fig:different-x-dist-poly-power}
 \end{figure}

### 2023-02-07 paper_comparison/paper_comparison/paper_comparison.Rmd
major update: use effect size instead of log(effect size) to fit model
the intercept is fixed by a constant log(0.05/0.95) to give 5% power when effect size = 0

@@
   # Italic if reject
-  mutate(across(r:s, ~ kableExtra::cell_spec(.x, italic = as.numeric(.x) < 0.05)))
+  mutate(across(r:s, ~ kableExtra::cell_spec(.x, italic = as.numeric(.x) <= 0.05)))
 
@@
 
-```{r}
+```{r cache = TRUE}
+# Define the minimum and maximum effect size
+min_poly_es <- vi_survey %>% 
+  filter(!null_lineup, 
+         !attention_check,
+         type == "polynomial", 
+         x_dist == "uniform") %>%
+  pull(effect_size) %>%
+  min()
+max_poly_es <- vi_survey %>% 
+  filter(!null_lineup, 
+         !attention_check,
+         type == "polynomial", 
+         x_dist == "uniform") %>%
+  pull(effect_size) %>%
+  max()
+
 # The glm model for RESET test
-conv_poly_glmmod <- poly_conv_sim %>%
+reset_poly_glmmod <- poly_conv_sim %>%
+  filter(x_dist == "uniform") %>%
   mutate(log_effect_size = log(effect_size)) %>%
-  select(x_dist, log_effect_size, RESET4_p_value) %>%
+  mutate(reject = as.numeric(RESET4_p_value <= 0.05)) %>%
+  mutate(offset0 = log(0.05/0.95)) %>%
+  select(x_dist, effect_size, log_effect_size, offset0, reject) %>%
+  glm(reject ~ effect_size - 1, family = binomial(), data = ., offset = offset0)
+
+# The prediction for all conventional tests on poly model
+pred_conv_poly <- poly_conv_sim %>%
   filter(x_dist == "uniform") %>%
-  mutate(reject = as.numeric(RESET4_p_value < 0.05)) %>%
-glm(reject ~ log_effect_size, family = binomial(), data = .)
+  select(-RESET3_p_value, -(RESET5_p_value:RESET10_p_value), -F_p_value) %>%
+  rename(RESET_p_value = RESET4_p_value) %>%
+  pivot_longer(RESET_p_value:SW_p_value) %>%
+  mutate(name = gsub("_p_value", " test", name)) %>%
+  mutate(reject = value <= 0.05) %>%
+  select(effect_size, name, reject) %>%
+  mutate(offset0 = log(0.05/0.95)) %>%
+  nest(dat = c(effect_size, offset0, reject)) %>%
+  mutate(mod = map(dat, 
+                   ~glm(reject ~ effect_size - 1, 
+                        family = binomial(), 
+                        data = .x,
+                        offset = offset0))) %>%
+  mutate(power = map(mod, function(mod) {
+    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
+                         offset0 = log(0.05/0.95))
+    result$power <- predict(mod, type = "response", newdata = result)
+    result
+  })) %>%
+  select(-dat, -mod) %>%
+  unnest(power) %>%
+  mutate(log_effect_size = log(effect_size))
+
+# Data needed by the visual test
+visual_poly_dat <- vi_survey %>%
+  filter(!null_lineup, 
+         !attention_check, 
+         type == "polynomial", 
+         x_dist == "uniform") %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  mutate(reject = as.numeric(p_value <= 0.05)) %>%
+  mutate(offset0 = log(0.05/0.95)) %>%
+  select(effect_size, offset0, reject)
+
+# The glm model for visual test
+visual_poly_glmmod <- glm(reject ~ effect_size - 1, 
+                          family = binomial(), 
+                          data = visual_poly_dat, 
+                          offset = offset0)
+
+# The prediction of the visual test
+pred_visual_poly <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
+                               offset0 = log(0.05/0.95)) %>%
+  mutate(power = predict(visual_poly_glmmod, 
+                         type = "response",
+                         newdata = data.frame(effect_size = effect_size,
+                                              offset0 = offset0))) %>%
+  mutate(log_effect_size = log(effect_size))
+
+# The prediction of the boot visual test
+pred_visual_poly_boot <- map_df(1:500, function(boot_id) {
+  slice_sample(visual_poly_dat, n = nrow(visual_poly_dat), replace = TRUE) %>%
+    nest(dat = c(effect_size, offset0, reject)) %>%
+    mutate(mod = map(dat, 
+                     ~glm(reject ~ effect_size - 1, 
+                          family = binomial(), 
+                          data = .x,
+                          offset = offset0))) %>%
+    mutate(power = map(mod, function(mod) {
+      result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
+                           offset0 = log(0.05/0.95))
+      result$power <- predict(mod, type = "response", newdata = result)
+      result
+    })) %>%
+    select(-dat, -mod) %>%
+    unnest(power) %>%
+    mutate(log_effect_size = log(effect_size)) %>%
+    mutate(boot_id = boot_id)
+  })
 ```
@@
 For non-linearity patterns, the power curve of RESET test climbs aggressively from
-`r round(unname(predict(conv_poly_glmmod, newdata = data.frame(log_effect_size = 0), type = "response")) * 100)`% to `r round(unname(predict(conv_poly_glmmod, newdata = data.frame(log_effect_size = 2), type = "response")) * 100)`% as $log_e(E)$ increases from 0 to 2, while power of other tests respond inactively to the change of effect, showing that RESET test is way more sensitive to the type of model defects that being considered.
+`r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(0), offset0 = log(0.05/0.95)), type = "response")) * 100)`% to `r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(2), offset0 = log(0.05/0.95)), type = "response")) * 100)`% as $log_e(E)$ increases from 0 to 2, while power of other tests respond inactively to the change of effect, showing that RESET test is way more sensitive to the type of model defects that being considered.
 Meanwhile, no noticeable visual features can be spotted from the example residual plots.
@@
 ```{r}
-# Get bootstrap samples
-get_boot_sample <- function(dat, times = 100) {
-  map_df(1:times, 
-         ~slice_sample(dat, n = nrow(dat), replace = TRUE) %>%
-           mutate(boot_id = .x))
-}
-
-# Get bootstrap CI
-get_boot_conf <- function(dat, 
-                          .x,
-                          .y,
-                          times = 100, 
-                          upper_q = 0.975, 
-                          lower_q = 0.025) {
-
-  # Construct the model formula
-  mod_formula <- substitute(.y ~ .x)
-
-  # Get the x variable name
-  x_chr <- as.character(substitute(.x))
-
-  # Prepare a sequence of x values
-  x_seq <- seq(min(dat[[x_chr]]), max(dat[[x_chr]]), by = 0.01)
-
-  # Properly construct the new data frame for prediction
-  newdata <- data.frame(.x = x_seq)
-  newdata[[x_chr]] <- x_seq
-
-  # Get the prediction for each bootstrap sample   
-  dat_pred <- map_df(1:times, function(i) {
-
-    # Get a bootstrap sample
-    dat_i <- slice_sample(dat, n = nrow(dat), replace = TRUE) %>%
-      mutate(boot_id = i)
-
-    # Fit the model on the sample
-    mod_i <- glm(mod_formula, family = binomial, data = dat_i)
-
-    # Get the prediction
-    newdata$.pred <- predict(mod_i, newdata = newdata, type = "response")
-
-    # Return the prediction
-    newdata
-  })
-
-  dat_pred <- dat_pred %>%
-    group_by(.x) %>%
-    summarise(upper = quantile(.pred, upper_q), 
-              lower = quantile(.pred, lower_q))
-  dat_pred[[x_chr]] <- dat_pred$.x
-  dat_pred
-}
-```
-
-```{r cache = TRUE}
-# target_sigma is obtained by evaluating a vector of sigma and picking those 
-# with expected effect 0 ~ 5
-target_sigma <- c(4.51, 3.52, 2.74, 2.14, 1.66, 1.29, 1.01, 0.79, 0.61, 0.48, 0.37)
-
-# Calculate the effect size 
-es <- map_dbl(target_sigma, 
-              ~log(poly_model(shape = 2, 
-                              x = rand_uniform(-1, 1), 
-                              sigma = .x)$
-                     effect_size(n = 100)
-                   )
-              )
-```
-
-```{r}
 # Analyse uniform polynomial data
-p <- vi_survey %>%
-  filter(!null_lineup, 
-         !attention_check, 
-         type == "polynomial", 
-         x_dist == "uniform") %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  mutate(reject = as.numeric(p_value <= 0.05)) %>%
+p <- ggplot() +
+
+  # Group the data by effect and reject to draw points
+  geom_point(data = visual_poly_dat %>%
                mutate(log_effect_size = log(effect_size)) %>%
-  select(log_effect_size, reject) %>%
-  ggplot() +
-
-  # Group the data by effect and reject to draw points
-  geom_point(data = ~.x %>% 
                group_by(log_effect_size, reject) %>% 
@@
   # Draw conventional test power curve
-  geom_smooth(data = poly_conv_sim %>% 
-                filter(x_dist == "uniform") %>%
-                mutate(log_effect_size = log(effect_size)) %>%
-                select(-RESET3_p_value, 
-                       -(RESET5_p_value:RESET10_p_value), 
-                       -F_p_value) %>%
-                rename(RESET_p_value = RESET4_p_value) %>%
-                pivot_longer(RESET_p_value:SW_p_value) %>%
-                mutate(name = gsub("_p_value", " test", name)), 
-              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
-              method = "glm", 
-              method.args = list(family = binomial),
-              se = FALSE) +
-
-  # Draw bootstrap CI for visual test
-  # geom_ribbon(data = ~get_boot_conf(.x,
-  #                                   log_effect_size,
-  #                                   reject,
-  #                                   times = 1000),
-  #             aes(x = log_effect_size, 
-  #                 ymin = lower, 
-  #                 ymax = upper),
-  #             fill = rev(rcartocolor::carto_pal(4, "Vivid"))[4],
-  #             alpha = 0.2) +
-
-  stat_smooth(data = ~get_boot_sample(.x, times = 1000),
-              geom = "line",
-              aes(log_effect_size, 
-                  reject, 
-                  col = "Visual test", 
-                  group = boot_id),
-              method = "glm",
-              method.args = list(family = binomial),
-              se = FALSE,
+  geom_line(data = pred_conv_poly,
+            aes(log_effect_size, power, col = name),
+            size = 1) +
+
+  # Draw boot visual test curves
+  geom_line(data = pred_visual_poly_boot,
+            aes(log_effect_size, power, col = "Visual test", group = boot_id),
+            size = 1,
             alpha = 0.01) +
@@
   # Draw visual test power curve
-  geom_smooth(aes(log_effect_size, reject, col = "Visual test"), 
-              method = "glm", 
-              method.args = list(family = binomial), 
-              se = FALSE) +
+  geom_line(data = pred_visual_poly,
+            aes(log_effect_size, power, col = "Visual test"), 
+            size = 1) +
 
@@
 
+```{r cache = TRUE}
+# target_sigma is obtained by evaluating a vector of sigma and picking those 
+# with expected effect 0 ~ 5
+target_sigma <- c(4.51, 3.52, 2.74, 2.14, 1.66, 1.29, 1.01, 0.79, 0.61, 0.48, 0.37)
+
+# Calculate the effect size 
+es <- map_dbl(target_sigma, 
+              ~log(poly_model(shape = 2, 
+                              x = rand_uniform(-1, 1), 
+                              sigma = .x)$
+                     effect_size(n = 100)
+                   )
+              )
+```
+
 ```{r polypower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for non-linear patterns (uniform fitted values only).
 Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups.
-Surrounding lines of the visual test show the estimated power of 1000 bootstrap samples.
+Surrounding lines of the visual test show the estimated power of 500 bootstrap samples.
 Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
 Where would you draw the line of too much non-linearity in the residuals?
-For the RESET test this is around log effect size 1.5, but for the visual test it is around 3.5."}
+For the RESET test this is around log effect size 2.5, but for the visual test it is around 3.5."}
 
@@
 ```{r cache = TRUE}
+# Define the minimum and maximum effect size
+min_heter_es <- vi_survey %>% 
+  filter(!null_lineup, 
+         !attention_check,
+         type == "heteroskedasticity", 
+         x_dist == "uniform") %>%
+  pull(effect_size) %>%
+  min()
+max_heter_es <- vi_survey %>% 
+  filter(!null_lineup, 
+         !attention_check,
+         type == "heteroskedasticity", 
+         x_dist == "uniform") %>%
+  pull(effect_size) %>%
+  max()
+
+# The prediction for all conventional tests on heter model
+pred_conv_heter <- heter_conv_sim %>%
+  filter(x_dist == "uniform") %>%
+  select(-RESET3_p_value, -(RESET5_p_value:RESET10_p_value), -F_p_value) %>%
+  rename(RESET_p_value = RESET4_p_value) %>%
+  pivot_longer(RESET_p_value:SW_p_value) %>%
+  mutate(name = gsub("_p_value", " test", name)) %>%
+  mutate(reject = value <= 0.05) %>%
+  select(effect_size, name, reject) %>%
+  mutate(offset0 = log(0.05/0.95)) %>%
+  nest(dat = c(effect_size, offset0, reject)) %>%
+  mutate(mod = map(dat, 
+                   ~glm(reject ~ effect_size - 1, 
+                        family = binomial(), 
+                        data = .x,
+                        offset = offset0))) %>%
+  mutate(power = map(mod, function(mod) {
+    result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
+                         offset0 = log(0.05/0.95))
+    result$power <- predict(mod, type = "response", newdata = result)
+    result
+  })) %>%
+  select(-dat, -mod) %>%
+  unnest(power) %>%
+  mutate(log_effect_size = log(effect_size))
+
+# Data needed by the visual test
+visual_heter_dat <- vi_survey %>%
+  filter(!null_lineup, 
+         !attention_check, 
+         type == "heteroskedasticity", 
+         x_dist == "uniform") %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  mutate(reject = as.numeric(p_value <= 0.05)) %>%
+  mutate(offset0 = log(0.05/0.95)) %>%
+  select(effect_size, offset0, reject)
+
+# The glm model for visual test
+visual_heter_glmmod <- glm(reject ~ effect_size - 1, 
+                           family = binomial(), 
+                           data = visual_poly_dat, 
+                           offset = offset0)
+
+# The prediction of the visual test
+pred_visual_heter <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
+                                offset0 = log(0.05/0.95)) %>%
+  mutate(power = predict(visual_heter_glmmod, 
+                         type = "response",
+                         newdata = data.frame(effect_size = effect_size,
+                                              offset0 = offset0))) %>%
+  mutate(log_effect_size = log(effect_size))
+
+# The prediction of the boot visual test
+pred_visual_heter_boot <- map_df(1:500, function(boot_id) {
+  slice_sample(visual_heter_dat, n = nrow(visual_heter_dat), replace = TRUE) %>%
+    nest(dat = c(effect_size, offset0, reject)) %>%
+    mutate(mod = map(dat, 
+                     ~glm(reject ~ effect_size - 1, 
+                          family = binomial(), 
+                          data = .x,
+                          offset = offset0))) %>%
+    mutate(power = map(mod, function(mod) {
+      result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
+                           offset0 = log(0.05/0.95))
+      result$power <- predict(mod, type = "response", newdata = result)
+      result
+    })) %>%
+    select(-dat, -mod) %>%
+    unnest(power) %>%
+    mutate(log_effect_size = log(effect_size)) %>%
+    mutate(boot_id = boot_id)
+  })
+```
+
+```{r}
+# Analyse uniform heteroskedasticity data
+p <- ggplot() +
+
+  # Group the data by effect and reject to draw points
+  geom_point(data = visual_heter_dat %>%
+               mutate(log_effect_size = log(effect_size)) %>%
+               group_by(log_effect_size, reject) %>% 
+               summarise(num_response = factor(n())),
+             aes(log_effect_size, reject, size = num_response), 
+             alpha = 0.3) +
+
+  # Draw conventional test power curve
+  geom_line(data = pred_conv_heter,
+            aes(log_effect_size, power, col = name),
+            size = 1) +
+
+  # Draw boot visual test curves
+  geom_line(data = pred_visual_heter_boot,
+            aes(log_effect_size, power, col = "Visual test", group = boot_id),
+            size = 1,
+            alpha = 0.01) +
+
+  # Draw visual test power curve
+  geom_line(data = pred_visual_heter,
+            aes(log_effect_size, power, col = "Visual test"), 
+            size = 1) +
+
+  # Theme
+  theme_light(base_size = 5) +
+  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.5)) +
+  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
+  scale_size_manual(values = c(1, 2, 3)) +
+  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
+  xlab(expression(log[e] (Effect_size))) +
+  ylab("Power") +
+  labs(col = "", size = "# lineups")
+```
+
+```{r cache = TRUE}
 # target_b is obtained by evaluating a vector of b and picking those 
@@
 
-```{r}
-# Analyse uniform heteroskedasticity data
-p <- vi_survey %>%
-  filter(!null_lineup, 
-         !attention_check, 
-         x_dist == "uniform", 
-         type == "heteroskedasticity") %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  mutate(reject = as.numeric(p_value <= 0.05)) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  select(log_effect_size, reject) %>%
-  ggplot() +
-
-  # Group the data by effect and reject to draw points
-  geom_point(data = ~.x %>% 
-               group_by(log_effect_size, reject) %>% 
-               summarise(num_response = factor(n())),
-             aes(log_effect_size, reject, size = num_response), 
-             alpha = 0.3) +
-
-  # Draw conventional test power curve
-  geom_smooth(data = heter_conv_sim %>% 
-                filter(x_dist == "uniform") %>%
-                mutate(log_effect_size = log(effect_size)) %>%
-                select(-RESET3_p_value, 
-                       -(RESET5_p_value:RESET10_p_value), 
-                       -F_p_value) %>%
-                rename(RESET_p_value = RESET4_p_value) %>%
-                pivot_longer(RESET_p_value:SW_p_value) %>%
-                mutate(name = gsub("_p_value", " test", name)), 
-              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
-              method = "glm", 
-              method.args = list(family = binomial),
-              se = FALSE) +
-
-  # Draw bootstrap CI for visual test
-  # geom_ribbon(data = ~get_boot_conf(.x,
-  #                                   log_effect_size,
-  #                                   reject,
-  #                                   times = 1000),
-  #             aes(x = log_effect_size, 
-  #                 ymin = lower, 
-  #                 ymax = upper),
-  #             fill = rev(rcartocolor::carto_pal(4, "Vivid"))[4],
-  #             alpha = 0.2) +
-
-  stat_smooth(data = ~get_boot_sample(.x, times = 1000),
-              geom = "line",
-              aes(log_effect_size, 
-                  reject, 
-                  col = "Visual test", 
-                  group = boot_id),
-              method = "glm",
-              method.args = list(family = binomial),
-              se = FALSE,
-              alpha = 0.01) +
-
-  # Draw visual test power curve
-  geom_smooth(aes(log_effect_size, reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
-
-  # Theme
-  theme_light(base_size = 5) +
-  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.5)) +
-  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
-  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
-  scale_size_manual(values = c(1, 2, 3)) +
-  xlab(expression(log[e] (Effect_size))) +
-  ylab("Power") +
-  labs(col = "", size = "# lineups")
-```
-
 ```{r heterpower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only).
@@
 Where would you draw the line of too much heteroskedasticity in the residuals?
-For the BP test this is around log effect size 1.5, but for the visual test it is around 3."}
+For the BP test this is around log effect size 2.5, but for the visual test it is around 3.5."}
 
@@
   mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
-  mutate(conv_reject = ifelse(conventional_p_value < 0.05, "Reject", "Not"),
-         reject = ifelse(p_value < 0.05, "Reject", "Not")) %>%
+  mutate(conv_reject = ifelse(conventional_p_value <= 0.05, "Reject", "Not"),
+         reject = ifelse(p_value <= 0.05, "Reject", "Not")) %>%
   mutate(conv_reject = factor(conv_reject, levels = c("Reject", "Not")),
@@
   filter(x_dist == "uniform") %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  mutate(reject = as.numeric(RESET4_p_value < 0.05)) %>%
+  mutate(reject = as.numeric(RESET4_p_value <= 0.05)) %>%
   mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>%
@@
   filter(x_dist == "uniform", type == "polynomial") %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  mutate(reject = as.numeric(p_value < 0.05)) %>%
+  mutate(reject = as.numeric(p_value <= 0.05)) %>%
   mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>%
@@
   mutate(shape = factor(shape, levels = c("U", "S", "M", "Triple-U"))) %>%
-  select(shape, log_effect_size, reject, conv_or_not)
+  mutate(offset0 = log(0.05/0.95)) %>%
+  select(shape, effect_size, offset0, reject, conv_or_not)
 
 dat_poly_j %>%
-  mutate(es_min = min(log_effect_size), 
-         es_max = max(log_effect_size)) %>%
-  nest(data = c(log_effect_size, reject)) %>%
-  mutate(mod = map(data, 
-                   ~glm(reject ~ log_effect_size, 
+  nest(dat = c(effect_size, offset0, reject)) %>%
+  mutate(mod = map(dat, 
+                   ~glm(reject ~ effect_size - 1, 
                         family = binomial(), 
-                        data = .x))) %>%
-  mutate(reject = map(mod, function(mod) {
-    result <- data.frame(log_effect_size = seq(es_min[1], es_max[1], 0.01))
-    result$reject <- predict(mod, type = "response", newdata = result)
+                        data = .x,
+                        offset = offset0))) %>%
+  mutate(power = map(mod, function(mod) {
+    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
+                         offset0 = log(0.05/0.95))
+    result$power <- predict(mod, type = "response", newdata = result)
     result
   })) %>%
-  select(-es_min, -es_max, -data, -mod) %>%
-  unnest(reject) %>%
+  select(-dat, -mod) %>%
+  unnest(power) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
   ggplot() +
-  geom_smooth(aes(log_effect_size, reject, col = shape), se = FALSE, span = 0.03) +
+  geom_line(aes(log_effect_size, power, col = shape), size = 1) +
   xlab(expression(log[e] (Effect_size))) +
@@
   filter(x_dist == "uniform") %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  mutate(reject = as.numeric(BP_p_value < 0.05)) %>%
+  mutate(reject = as.numeric(BP_p_value <= 0.05)) %>%
   mutate(a = c("Left-triangle", "Butterfly", "Right-triangle")[a + 2]) %>%
@@
   filter(x_dist == "uniform", type == "heteroskedasticity") %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  mutate(reject = as.numeric(p_value < 0.05)) %>%
+  mutate(reject = as.numeric(p_value <= 0.05)) %>%
   mutate(a = c("Left-triangle", "Butterfly", "Right-triangle")[a + 2]) %>%
@@
   mutate(a = factor(a, levels = c("Left-triangle", "Butterfly", "Right-triangle"))) %>%
-  select(a, log_effect_size, reject, conv_or_not)
+  mutate(offset0 = log(0.05/0.95)) %>%
+  select(a, effect_size, offset0, reject, conv_or_not)
 
@@
 dat_heter_a %>%
-  mutate(es_min = min(log_effect_size), 
-         es_max = max(log_effect_size)) %>%
-  nest(data = c(log_effect_size, reject)) %>%
-  mutate(mod = map(data, 
-                   ~glm(reject ~ log_effect_size, 
+  nest(dat = c(effect_size, offset0, reject)) %>%
+  mutate(mod = map(dat, 
+                   ~glm(reject ~ effect_size - 1, 
                         family = binomial(), 
-                        data = .x))) %>%
-  mutate(reject = map(mod, function(mod) {
-    result <- data.frame(log_effect_size = seq(es_min[1], es_max[1], 0.01))
-    result$reject <- predict(mod, type = "response", newdata = result)
+                        data = .x,
+                        offset = offset0))) %>%
+  mutate(power = map(mod, function(mod) {
+    result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
+                         offset0 = log(0.05/0.95))
+    result$power <- predict(mod, type = "response", newdata = result)
     result
   })) %>%
-  select(-es_min, -es_max, -data, -mod) %>%
-  unnest(reject) %>%
+  select(-dat, -mod) %>%
+  unnest(power) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
   ggplot() +
-  geom_smooth(aes(log_effect_size, reject, col = a), se = FALSE, span = 0.03) +
+  geom_line(aes(log_effect_size, power, col = a), size = 1) +
   xlab(expression(log[e] (Effect_size))) +
@@
 ```{r get-5-eval, cache = TRUE}
-dat_uniform_5 <- map_df(1:1000, function(boot_id) {
+dat_uniform_5 <- map_df(1:500, function(boot_id) {
   vi_survey %>%
@@
   mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
   mutate(reject = as.numeric(p_value <= 0.05)) %>%
   mutate(conv_or_not = "Visual") %>%
-  select(-unique_lineup_id, -p_value, -effect_size)
+  mutate(offset0 = log(0.05/0.95)) %>%
+  select(-unique_lineup_id, -p_value)
 ```
@@
 For lineups with uniform fitted value distribution, the five evaluations are randomly sampled from the total eleven evaluations.
-The sampling process has been repeated for 1000 times and the corresponding power curves are drawn with transparent grey lines.
-The vetical dotted line is drawn to indicate where the visual pattern start to become recognisable.
+The sampling process has been repeated for 500 times and the corresponding power curves are drawn with transparent grey lines.
 For both types of tests power does change.
 The discrete fitted value distribution substantially lowers the power of visual tests for non-linearity patterns in residual plots.
-For heteroskedasticity patterns, the lognormal distribution produce lower power for both types of tests."}
+For heteroskedasticity patterns, normal and lognormal distributions produce lower power for both types of tests."}
 
@@
   poly_conv_sim %>% 
-    mutate(log_effect_size = log(effect_size)) %>%
-    select(x_dist, log_effect_size, RESET4_p_value) %>%
     rename(RESET_p_value = RESET4_p_value) %>%
@@
                            stringr::str_to_title(x_dist))) %>%
-    mutate(reject = as.numeric(RESET_p_value < 0.05)) %>%
+    mutate(reject = as.numeric(RESET_p_value <= 0.05)) %>%
     mutate(conv_or_not = "Conventional") %>%
     mutate(type = "non-linearity") %>%
-    select(type, x_dist, log_effect_size, reject, conv_or_not),
+    mutate(offset0 = log(0.05/0.95)) %>%
+    select(type, x_dist, effect_size, offset0, reject, conv_or_not),
 
   heter_conv_sim %>% 
-    mutate(log_effect_size = log(effect_size)) %>%
-    select(x_dist, log_effect_size, BP_p_value) %>%
     mutate(x_dist = ifelse(x_dist == "even_discrete",
@@
                            stringr::str_to_title(x_dist))) %>%
-    mutate(reject = as.numeric(BP_p_value < 0.05)) %>%
+    mutate(reject = as.numeric(BP_p_value <= 0.05)) %>%
     mutate(conv_or_not = "Conventional") %>%
     mutate(type = "heteroskedasticity") %>%
-    select(type, x_dist, log_effect_size, reject, conv_or_not),
+    mutate(offset0 = log(0.05/0.95)) %>%
+    select(type, x_dist, effect_size, offset0, reject, conv_or_not),
 
@@
     mutate(reject = as.numeric(p_value <= 0.05)) %>%
-    mutate(log_effect_size = log(effect_size)) %>%
     mutate(x_dist = ifelse(x_dist == "even_discrete", 
@@
     mutate(conv_or_not = "Visual") %>%
-    select(type, x_dist, log_effect_size, reject, conv_or_not)
+    mutate(offset0 = log(0.05/0.95)) %>%
+    select(type, x_dist, effect_size, offset0, reject, conv_or_not)
 ) %>%
@@
 dat_uniform_5_pred <- dat_uniform_5 %>%
-  mutate(es_min = min(dat_x_dist$log_effect_size), 
-         es_max = max(dat_x_dist$log_effect_size)) %>%
-  nest(data = c(log_effect_size, reject)) %>%
-  mutate(mod = map(data, 
-                   ~glm(reject ~ log_effect_size, 
+  nest(dat = c(effect_size, offset0, reject)) %>%
+  mutate(mod = map(dat, 
+                   ~glm(reject ~ effect_size - 1, 
                         family = binomial(), 
-                        data = .x))) %>%
-  mutate(reject = map(mod, function(mod) {
-    result <- data.frame(log_effect_size = seq(es_min[1], es_max[1], 0.01))
-    result$reject <- predict(mod, type = "response", newdata = result)
+                        data = .x,
+                        offset = offset0))) %>%
+  mutate(power = map2(mod, type, function(mod, type) {
+    this_min_es <- ifelse(type == "heteroskedasticity", min_heter_es, min_poly_es)
+    this_max_es <- ifelse(type == "heteroskedasticity", max_heter_es, max_poly_es)
+    result <- data.frame(effect_size = seq(this_min_es[1], this_max_es[1], 0.1),
+                         offset0 = log(0.05/0.95))
+    result$power <- predict(mod, type = "response", newdata = result)
     result
   })) %>%
-  select(-es_min, -es_max, -data, -mod) %>%
-  unnest(reject)
+  select(-dat, -mod) %>%
+  unnest(power) %>%
+  mutate(log_effect_size = log(effect_size))
 
@@
 dat_x_dist %>%
-  mutate(es_min = min(log_effect_size), 
-         es_max = max(log_effect_size)) %>%
-  nest(data = c(log_effect_size, reject)) %>%
-  mutate(mod = map(data, 
-                   ~glm(reject ~ log_effect_size, 
+  nest(dat = c(effect_size, offset0, reject)) %>%
+  mutate(mod = map(dat, 
+                   ~glm(reject ~ effect_size - 1, 
                         family = binomial(), 
-                        data = .x))) %>%
-  mutate(reject = map(mod, function(mod) {
-    result <- data.frame(log_effect_size = seq(es_min[1], es_max[1], 0.01))
-    result$reject <- predict(mod, type = "response", newdata = result)
+                        data = .x,
+                        offset = offset0))) %>%
+  mutate(power = map2(mod, type, function(mod, type) {
+    this_min_es <- ifelse(type == "heteroskedasticity", min_heter_es, min_poly_es)
+    this_max_es <- ifelse(type == "heteroskedasticity", max_heter_es, max_poly_es)
+    result <- data.frame(effect_size = seq(this_min_es[1], this_max_es[1], 0.1),
+                         offset0 = log(0.05/0.95))
+    result$power <- predict(mod, type = "response", newdata = result)
     result
   })) %>%
-  select(-es_min, -es_max, -data, -mod) %>%
-  unnest(reject) %>%
+  select(-dat, -mod) %>%
+  unnest(power) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
 
@@
 
-  geom_vline(xintercept = 2, linetype = 2) +
-
   # Draw boot uniform (5 eval) lines
-  stat_smooth(data = dat_uniform_5_pred,
-              geom = "line",
-              aes(log_effect_size, reject, col = "Uniform", group = boot_id),
-              span = 0.03,
-              se = FALSE,
+  geom_line(data = dat_uniform_5_pred,
+            aes(log_effect_size, power, col = "Uniform", group = boot_id),
+            size = 1,
             alpha = 0.01) +
@@
   # Draw all smooth lines
-  geom_smooth(aes(log_effect_size, reject, col = x_dist),
-              span = 0.03,
-              se = FALSE,
+  geom_line(aes(log_effect_size, power, col = x_dist),
             size = 1) +
@@
 
-Throughout the three data collection periods of the human subject experiment we conducted, residual plots rejected by visual tests are rarely not rejected by conventional tests, but departures of residuals that are of little practical sense to humans will often be alarmed by conventional tests, supporting the common guidance given by experts to utilize graphical representations of data in model diagnostics.
+According to the data collected from the human subject experiment, residual plots rejected by visual tests are rarely not rejected by conventional tests, but departures of residuals that are of little practical sense to humans will often be alarmed by conventional tests, supporting the common guidance given by experts to utilize graphical representations of data in model diagnostics.
 
@@
 
+Interestingly, the orientation of the visual pattern
+
 fitted value distribution

### 2023-02-07 paper_comparison/paper_comparison/paper_comparison.tex
major update: use effect size instead of log(effect size) to fit model
the intercept is fixed by a constant log(0.05/0.95) to give 5% power when effect size = 0

@@
 For non-linearity patterns, the power curve of RESET test climbs
-aggressively from 15\% to 69\% as \(log_e(E)\) increases from 0 to 2,
+aggressively from 7\% to 30\% as \(log_e(E)\) increases from 0 to 2,
 while power of other tests respond inactively to the change of effect,
@@
 Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups.
-Surrounding lines of the visual test show the estimated power of 1000 bootstrap samples.
+Surrounding lines of the visual test show the estimated power of 500 bootstrap samples.
 Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
 Where would you draw the line of too much non-linearity in the residuals?
-For the RESET test this is around log effect size 1.5, but for the visual test it is around 3.5.}\label{fig:polypower}
+For the RESET test this is around log effect size 2.5, but for the visual test it is around 3.5.}\label{fig:polypower}
 \end{figure}
@@
 Where would you draw the line of too much heteroskedasticity in the residuals?
-For the BP test this is around log effect size 1.5, but for the visual test it is around 3.}\label{fig:heterpower}
+For the BP test this is around log effect size 2.5, but for the visual test it is around 3.5.}\label{fig:heterpower}
 \end{figure}
@@
 For lineups with uniform fitted value distribution, the five evaluations are randomly sampled from the total eleven evaluations.
-The sampling process has been repeated for 1000 times and the corresponding power curves are drawn with transparent grey lines.
-The vetical dotted line is drawn to indicate where the visual pattern start to become recognisable.
+The sampling process has been repeated for 500 times and the corresponding power curves are drawn with transparent grey lines.
 For both types of tests power does change.
 The discrete fitted value distribution substantially lowers the power of visual tests for non-linearity patterns in residual plots.
-For heteroskedasticity patterns, the lognormal distribution produce lower power for both types of tests.}\label{fig:different-x-dist-poly-power}
+For heteroskedasticity patterns, normal and lognormal distributions produce lower power for both types of tests.}\label{fig:different-x-dist-poly-power}
 \end{figure}
@@
 
-Throughout the three data collection periods of the human subject
-experiment we conducted, residual plots rejected by visual tests are
-rarely not rejected by conventional tests, but departures of residuals
-that are of little practical sense to humans will often be alarmed by
-conventional tests, supporting the common guidance given by experts to
-utilize graphical representations of data in model diagnostics.
+According to the data collected from the human subject experiment,
+residual plots rejected by visual tests are rarely not rejected by
+conventional tests, but departures of residuals that are of little
+practical sense to humans will often be alarmed by conventional tests,
+supporting the common guidance given by experts to utilize graphical
+representations of data in model diagnostics.
 
@@
 departures can be tested simultaneously.
+
+Interestingly, the orientation of the visual pattern
 

### 2023-02-07 paper_comparison/paper_comparison/paper_comparison.Rmd
update the desc of the power model

@@
 The assumption is to simplify the model structure, thereby obviate costly large-scale experiments to estimate complex covariance matrices.
-The model is a logistic regression with the natural logarithm of the effect as the only regressor formulated as
+The model is a logistic regression formulated as
 
@@
 \begin{equation} \label{eq:logistic-regression-1-1}
-Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log(E)),
+Pr(\text{reject}~H_0|H_1,E) = \Lambda(log(\frac{0.5}{0.95}) + \beta_1 E),
 \end{equation}
@@
 \noindent where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
-
-Effect $E$ is derived from the Kullback-Leibler divergence (see Appendix for the details) formulated as
+The effect size is the only regressor and the intercept is fixed to $log(0.5/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$.
+
+Effect size $E$ is derived from the Kullback-Leibler divergence (see Appendix for the details) formulated as
 

### 2023-02-07 paper_comparison/paper_comparison/paper_comparison.Rmd
adjust pics in the power cmp

@@
 
-1. experiments -> data collection periods
-
 To investigate the difference between conventional hypothesis testing and visual inference in the application of linear regression diagnostics, an experiment is conducted in three data collection periods.
@@
 Additionally, we have collected extra evaluations on these lineups, which will result in more reliable analysis.
-Analysis of lineups with other fitted value distributions can be found in Section \ref{effect-of-other-fitted-value-distributions}.
+Analysis of lineups with other fitted value distributions can be found in Section \ref{effect-of-fitted-value-distributions}.
 
@@
 For non-linearity patterns, the power curve of RESET test climbs aggressively from
-`r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(0), offset0 = log(0.05/0.95)), type = "response")) * 100)`% to `r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(2), offset0 = log(0.05/0.95)), type = "response")) * 100)`% as $log_e(E)$ increases from 0 to 2, while power of other tests respond inactively to the change of effect, showing that RESET test is way more sensitive to the type of model defects that being considered.
-Meanwhile, no noticeable visual features can be spotted from the example residual plots.
+`r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(0), offset0 = log(0.05/0.95)), type = "response")) * 100)`% to `r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(2.5), offset0 = log(0.05/0.95)), type = "response")) * 100)`% as $log_e(E)$ increases from 0 to 2.5, while power of other tests respond inactively to the change of effect, showing that RESET test is way more sensitive to the type of model defects that being considered.
+Meanwhile, no noticeable residual depatures can be spotted from the example residual plots.
 
 In terms of heteroskedasticity patterns, the power of Breusch-Pagan test is also almost always greater than the power of visual test.
-For $0 \leq log_e(E) \leq 2$, where the power curve of the visual test remains at a low level, the Breusch-Pagan test still have a decent amount of chance of rejecting $H_0$.
+For $log_e(E)$ at around 2.5, where the power curve of the visual test remains at a low level, the Breusch-Pagan test still have a decent amount of chance of rejecting $H_0$.
 Similarly, the visual feature is nearly unobservable from the example residual plots.
@@
 
-set.seed(106)
+# 106
+# 112
+# 113
+# 115
+# 132
+# 133
+# 144!
+# 147
+set.seed(132)
 
@@
                 remove_axis = TRUE, 
-                remove_grid_line = TRUE, 
-                size = 0.05) +
+                remove_grid_line = TRUE, size = 0.25, stroke = 0.25) +
     ggtitle(round(es[j], 1)) +
@@
 
-set.seed(10091)
+# 10091
+# 101
+# 105
+# 106
+# 112
+# 119
+# 133
+set.seed(135)
 
@@
                 remove_axis = TRUE,
-                remove_grid_line = TRUE, size = 0.05) +
+                remove_grid_line = TRUE, size = 0.25, stroke = 0.25) +
     ggtitle(round(es[j], 1)) +

### 2023-02-07 paper_comparison/paper_comparison/paper_comparison.tex
adjust pics in the power cmp

@@
 obviate costly large-scale experiments to estimate complex covariance
-matrices. The model is a logistic regression with the natural logarithm
-of the effect as the only regressor formulated as
+matrices. The model is a logistic regression formulated as
 
 \begin{equation} \label{eq:logistic-regression-1-1}
-Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log(E)),
+Pr(\text{reject}~H_0|H_1,E) = \Lambda(log(\frac{0.5}{0.95}) + \beta_1 E),
 \end{equation}
@@
 \noindent where \(\Lambda(.)\) is the standard logistic function given
-as \(\Lambda(z) = exp(z)/(1+exp(z))\).
-
-Effect \(E\) is derived from the Kullback-Leibler divergence (see
+as \(\Lambda(z) = exp(z)/(1+exp(z))\). The effect size is the only
+regressor and the intercept is fixed to \(log(0.5/0.95)\) so that
+\(\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05\).
+
+Effect size \(E\) is derived from the Kullback-Leibler divergence (see
 Appendix for the details) formulated as
@@
 \section{Experimental design}\label{experimental-design}}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  experiments -\textgreater{} data collection periods
-\end{enumerate}
 
@@
 of lineups with other fitted value distributions can be found in Section
-\ref{effect-of-other-fitted-value-distributions}.
+\ref{effect-of-fitted-value-distributions}.
 
@@
 For non-linearity patterns, the power curve of RESET test climbs
-aggressively from 7\% to 30\% as \(log_e(E)\) increases from 0 to 2,
+aggressively from 7\% to 63\% as \(log_e(E)\) increases from 0 to 2.5,
 while power of other tests respond inactively to the change of effect,
 showing that RESET test is way more sensitive to the type of model
-defects that being considered. Meanwhile, no noticeable visual features
-can be spotted from the example residual plots.
+defects that being considered. Meanwhile, no noticeable residual
+depatures can be spotted from the example residual plots.
 
@@
 is also almost always greater than the power of visual test. For
-\(0 \leq log_e(E) \leq 2\), where the power curve of the visual test
+\(log_e(E)\) at around 2.5, where the power curve of the visual test
 remains at a low level, the Breusch-Pagan test still have a decent

### 2023-02-07 paper_comparison/paper_comparison/paper_comparison.Rmd
update the text to match the updated plot

@@
 
-The power of visual test arises steadily as $log_e(E)$ increases from 2 to 5 for both non-linearity patterns and heteroskedasticity patterns, suggesting that the effect starts to make significant impact on the degree of the presence of the designed visual features.
-This can also be observed from the example residual plots that when $log_e(E) = 2.5$, a weak "S-shape" and a weak "triangle" shape are presented in Figure \ref{fig:polypower} and Figure \ref{fig:heterpower} respectively.
+The power of visual test arises steadily as $log_e(E)$ increases from 3 to 5 for both non-linearity patterns and heteroskedasticity patterns, suggesting that the effect starts to make significant impact on the degree of the presence of the designed visual features.
+This can also be observed from the example residual plots that when $log_e(E) = 3.5$, a weak "S-shape" and a weak "triangle" shape are presented in Figure \ref{fig:polypower} and Figure \ref{fig:heterpower} respectively.
 The visual pattern becomes much clearer as $log_e(E)$ increases.
-At $log_e(E) \approx 6$, the power reaches almost 100%.
-
-The power of all inappropriate tests except for RESET test shows improvement as the effect increases but at a lower rate than the visual test in both scenarios.
+At $log_e(E) = 5$, the power reaches almost 100%.
+
+The power of all inappropriate tests show improvement as the effect increases but at a lower rate than the visual test in both scenarios.
 This coincides the point made by @cook1982residuals that residual-based tests for a specific type of model defect may be sensitive to other types of model defects.
@@
 
-A primary factor contributes to the non-linearity model is the shape of the non-linearity.
-According to Figure \ref{fig:poly-power-uniform-j}, conventional tests have almost identical power in testing the "U", the "S" and the "M" shape, except for testing the "Triple-U" shape, which have significant lower power.
+A primary factor contributes to the non-linearity model is the shape of the non-linearity pattern.
+According to Figure \ref{fig:poly-power-uniform-j}, conventional tests have identical power in testing the "U", the "S" and the "M" shape.
+But it has significant lower power in testing the "Triple-U" shape.
 To understand why this is, one needs to return to the way the RESET test is applied.
@@
 The recommendation of the polynomial power for the RESET should be revised, perhaps.
-This illustrates the sensitivity of conventional testing to the parameters, and it also points to a limitation that one needs to know the data structure in order to set the parameters for the test.
-
-For visual tests, we expect the "U" shape will be the easiest one to be detected by subjects followed by the "S", "M" and "Triple-U" shape.
+This illustrates the sensitivity of conventional testing to the parameters, and it also points to a limitation that one needs to know the data generating process in order to set the parameters for the test.
+
+For visual tests, based on the orders of the Hermite polynomials, we expect the "U" shape will be the easiest one to be detected by subjects followed by the "S", "M" and "Triple-U" shape.
 From Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power curves are mostly aligned with our expectation, except for the "M" shape, which is as easy to be detected as the "U" shape.
@@
 
-We also investigate the impact of different heteroskedasticity shapes on power of conventional tests and visual tests.
+We have also investigated the impact of different heteroskedasticity shapes on power of conventional tests and visual tests.
 In theory, the "Left-triangle" and the "Right-triangle" shapes are functionally identical from the point of view of a Breusch-Pagan test.
 As shown in Figure \ref{fig:heter-power-uniform-a}, this is indeed the case where little difference between the power curves can be perceived.
-Similarly, visual tests will have the same power of detecting these two shapes if they are equally likely to be identified.
+Similarly, visual tests should have the smae power in detecting these two shapes if they are equally likely to be identified.
 However, it can be observed from Figure \ref{fig:heter-power-uniform-a} that the power curve for the "Left-triangle" shape is constantly higher than the one for the "Right-triangle" shape, indicating a potential favour of orientation by human, which is worth to be explored in future studies.
-
-The chance of detecting the "Butterfly" shape is roughly the same as the chance of detecting the "Left-triangle" shape in both tests.
-The only noticeable difference from Figure \ref{fig:heter-power-uniform-a} is the Breusch-Pagan test has slightly higher power for detecting this shape compared to other two shapes when the effect size is low.
+Besides, the chance of detecting the "Butterfly" shape by visual tests are slightly higher than other two shapes.
 
@@
 Figure \ref{fig:different-x-dist-poly-power} illustrates the change of power of visual tests and conventional tests for different fitted distributions used in the non-linearity model and the heteroskedasticity model.
-We focus on $log(E) > 2$ where the visual patterns start to become recognisable.
+Note that we have collected more evaluations on lineup with uniform fitted value distribution, and the number of evaluations on a lineup will have an impact on the power of the visual test.
+To have a fair comparison, the five evaluations used in the power estimation are randomly sampled from the total eleven evaluations for each lineup with uniform fitted value distribution.
+The sampling process has been repeated 500 times to visualize the variation.
 

### 2023-02-07 paper_comparison/paper_comparison/paper_comparison.tex
update the text to match the updated plot

@@
 The power of visual test arises steadily as \(log_e(E)\) increases from
-2 to 5 for both non-linearity patterns and heteroskedasticity patterns,
+3 to 5 for both non-linearity patterns and heteroskedasticity patterns,
 suggesting that the effect starts to make significant impact on the
 degree of the presence of the designed visual features. This can also be
-observed from the example residual plots that when \(log_e(E) = 2.5\), a
+observed from the example residual plots that when \(log_e(E) = 3.5\), a
 weak ``S-shape'' and a weak ``triangle'' shape are presented in Figure
@@
 visual pattern becomes much clearer as \(log_e(E)\) increases. At
-\(log_e(E) \approx 6\), the power reaches almost 100\%.
-
-The power of all inappropriate tests except for RESET test shows
-improvement as the effect increases but at a lower rate than the visual
-test in both scenarios. This coincides the point made by
-\citet{cook1982residuals} that residual-based tests for a specific type
-of model defect may be sensitive to other types of model defects. The
-power curve of RESET test remains at around 5\% in Figure
-\ref{fig:heterpower} since there are no non-linear terms leave out in
-the heteroskedasticity model and \(H_0\) of the test is always
-satisfied.
+\(log_e(E) = 5\), the power reaches almost 100\%.
+
+The power of all inappropriate tests show improvement as the effect
+increases but at a lower rate than the visual test in both scenarios.
+This coincides the point made by \citet{cook1982residuals} that
+residual-based tests for a specific type of model defect may be
+sensitive to other types of model defects. The power curve of RESET test
+remains at around 5\% in Figure \ref{fig:heterpower} since there are no
+non-linear terms leave out in the heteroskedasticity model and \(H_0\)
+of the test is always satisfied.
 
@@
 A primary factor contributes to the non-linearity model is the shape of
-the non-linearity. According to Figure \ref{fig:poly-power-uniform-j},
-conventional tests have almost identical power in testing the ``U'', the
-``S'' and the ``M'' shape, except for testing the ``Triple-U'' shape,
-which have significant lower power. To understand why this is, one needs
-to return to the way the RESET test is applied. It requires a parameter
-indicating degree of fitted values to test for, and the recommendation
-is to generically use four (\citet{ramsey_tests_1969}). However, the
-``Triple-U'' shape constructed from the Hermite polynomials use power up
-to 18. If the RESET test had been applied using a higher power no less
-than six, the power curve of ``Triple-U'' shape will have little
-difference to other shapes. The recommendation of the polynomial power
-for the RESET should be revised, perhaps. This illustrates the
-sensitivity of conventional testing to the parameters, and it also
-points to a limitation that one needs to know the data structure in
-order to set the parameters for the test.
-
-For visual tests, we expect the ``U'' shape will be the easiest one to
-be detected by subjects followed by the ``S'', ``M'' and ``Triple-U''
-shape. From Figure \ref{fig:poly-power-uniform-j}, it can be observed
-that the power curves are mostly aligned with our expectation, except
-for the ``M'' shape, which is as easy to be detected as the ``U'' shape.
-This implies the readability of the shape do not strictly follow the
-degree of the polynomials.
+the non-linearity pattern. According to Figure
+\ref{fig:poly-power-uniform-j}, conventional tests have identical power
+in testing the ``U'', the ``S'' and the ``M'' shape. But it has
+significant lower power in testing the ``Triple-U'' shape. To understand
+why this is, one needs to return to the way the RESET test is applied.
+It requires a parameter indicating degree of fitted values to test for,
+and the recommendation is to generically use four
+(\citet{ramsey_tests_1969}). However, the ``Triple-U'' shape constructed
+from the Hermite polynomials use power up to 18. If the RESET test had
+been applied using a higher power no less than six, the power curve of
+``Triple-U'' shape will have little difference to other shapes. The
+recommendation of the polynomial power for the RESET should be revised,
+perhaps. This illustrates the sensitivity of conventional testing to the
+parameters, and it also points to a limitation that one needs to know
+the data generating process in order to set the parameters for the test.
+
+For visual tests, based on the orders of the Hermite polynomials, we
+expect the ``U'' shape will be the easiest one to be detected by
+subjects followed by the ``S'', ``M'' and ``Triple-U'' shape. From
+Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power
+curves are mostly aligned with our expectation, except for the ``M''
+shape, which is as easy to be detected as the ``U'' shape. This implies
+the readability of the shape do not strictly follow the degree of the
+polynomials.
 
@@
 
-We also investigate the impact of different heteroskedasticity shapes on
-power of conventional tests and visual tests. In theory, the
+We have also investigated the impact of different heteroskedasticity
+shapes on power of conventional tests and visual tests. In theory, the
 ``Left-triangle'' and the ``Right-triangle'' shapes are functionally
@@
 little difference between the power curves can be perceived. Similarly,
-visual tests will have the same power of detecting these two shapes if
+visual tests should have the smae power in detecting these two shapes if
 they are equally likely to be identified. However, it can be observed
@@
 ``Right-triangle'' shape, indicating a potential favour of orientation
-by human, which is worth to be explored in future studies.
-
-The chance of detecting the ``Butterfly'' shape is roughly the same as
-the chance of detecting the ``Left-triangle'' shape in both tests. The
-only noticeable difference from Figure \ref{fig:heter-power-uniform-a}
-is the Breusch-Pagan test has slightly higher power for detecting this
-shape compared to other two shapes when the effect size is low.
+by human, which is worth to be explored in future studies. Besides, the
+chance of detecting the ``Butterfly'' shape by visual tests are slightly
+higher than other two shapes.
 
@@
 distributions used in the non-linearity model and the heteroskedasticity
-model. We focus on \(log(E) > 2\) where the visual patterns start to
-become recognisable.
+model. Note that we have collected more evaluations on lineup with
+uniform fitted value distribution, and the number of evaluations on a
+lineup will have an impact on the power of the visual test. To have a
+fair comparison, the five evaluations used in the power estimation are
+randomly sampled from the total eleven evaluations for each lineup with
+uniform fitted value distribution. The sampling process has been
+repeated 500 times to visualize the variation.
 

### 2023-02-07 paper_comparison/paper_comparison/paper_comparison.Rmd
data collection periods language refined

@@
 To investigate the difference between conventional hypothesis testing and visual inference in the application of linear regression diagnostics, an experiment is conducted in three data collection periods.
-Two types of departures, namely non-linearity and heteroskedasticity, are considered with the corresponding data generating process being designed for data collection periods I and II.
-The data collection period III is designed for collecting human responses to null lineups such that the parameter $\alpha$ in Equation \ref{eq:pvalue-beta-binomial} can be estimated.
-Overall, we plan to collect `r vi_survey  %>% filter(!attention_check) %>% nrow()` evaluations on `r vi_survey  %>% filter(!attention_check) %>% pull(unique_lineup_id) %>% unique() %>% length()` unique lineups performed by `r vi_survey  %>% filter(!attention_check) %>% count(exp, set) %>% nrow()` subjects throughout three data collection periods.
+The two types of departures, non-linearity and heteroskedasticity, are primarily collected during data collection periods I and II.
+The data collection period III was designed primarily to measure human responses to null lineups so that the parameter $\alpha$ in Equation \ref{eq:pvalue-beta-binomial} can be estimated.
+Additional lineups for both non-linearity and heteroskedasticity, using uniform $X$ distribution, were included so that the participants were evaluating some lineups with signal also.
+It would be too frustrating for participants to only be assigned lineups with all null plots.
+Overall, we collected `r vi_survey  %>% filter(!attention_check) %>% nrow()` evaluations on `r vi_survey  %>% filter(!attention_check) %>% pull(unique_lineup_id) %>% unique() %>% length()` unique lineups performed by `r vi_survey  %>% filter(!attention_check) %>% count(exp, set) %>% nrow()` subjects throughout three data collection periods.
 
@@
 
-```{r different-shape-of-herimite, fig.height = 2, fig.cap = "Polynomial forms generated for the residual plots used in data collection period I.
+```{r different-shape-of-herimite, fig.height = 2, fig.cap = "Polynomial forms generated for the residual plots used to assess detecting non-linearity.
 The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$."}
@@
 
-Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups used in data collection period I.
+Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups used in non-linearity detection.
 This lineup is produced by the non-linearity model with $j = 6$ and uniform fitted values.
@@
 
-```{r different-n, fig.height = 2.67, fig.cap = "Number of observations $n$ have an impact on the strength of the signal.
-Three different values of $n$ are used in data collection periods I, II and III."}
+```{r different-n, fig.height = 2.67, fig.cap = "Examining the effect of signal strength for the three different values of $n$ used in the experiment, for non-linear structure with fixed $\\sigma = 1.5$, uniform $X$ distribution, and S shape.
+For these factor levels, only when $n=300$ is the S shape clearly visible."}
 
@@
 
-```{r different-sigma, fig.height = 2, fig.cap = "For the non-lienearity model, the standard deviation of the error term $\\sigma$ affects the strength of the signal.
-Four different values of $\\sigma$ are used in the data collection period I."}
+```{r different-sigma, fig.height = 2, fig.cap = "Examining the effect of $\\sigma$ on the signal strength in the non-linearity detection, for $n=300$, uniform $X$ distribution and the U shape.
+As $\\sigma$ increases the signal strength decreases, to the point that the U is almost unrecognisable when $\\sigma=4$."}
 
@@
 
-```{r different-shape-of-heter, fig.height = 2.67, fig.cap = "Heteroskedasticity forms used in data collection period II.
-Three different shapes (a = -1, 0, 1) are used in the experiment to create left-triangle, butterfly and right-triangle shapes."}
+```{r different-shape-of-heter, fig.height = 2.67, fig.cap = "Heteroskedasticity forms used in the experiment.
+Three different shapes ($a = -1, 0, 1$) are used in the experiment to create left-triangle, butterfly and right-triangle shapes, respectively."}
 
@@
 
-```{r different-b, fig.height = 2, fig.cap = "Five different values of $b$ are used in data collection period II to control the strength of the signal."}
+```{r different-b, fig.height = 2, fig.cap = "Five different values of $b$ are used in heteroskedasticity simulation to control the strength of the signal.
+Larger values of $b$ yield a bigger difference in variation, and stus stronger heteroskedasticity signal."}
 
@@
 This paper has demonstrated that conventional residual-based statistical tests are more sensitive to departures from assumptions for residuals than visual tests evaluated by humans in the diagnostics of linear regression models.
-An experiment is conducted in three data collection periods with two simulation models designed for non-linearity and heteroskedasticity model defects being applied.
-Methods for obtaining the $p$-value of visual tests proposed by @vanderplas2021statistical are extended to allowed for multiple selections in a lineup protocol.
+An experiment is conducted in three data collection periods using two simulation models, one for for non-linearity and and the other for heteroskedasticity.
+Methods for obtaining the $p$-value of visual tests proposed by @vanderplas2021statistical are extended to allow for multiple selections in a lineup protocol.
 

### 2023-02-07 paper_comparison/paper_comparison/paper_comparison.tex
data collection periods language refined

@@
 diagnostics, an experiment is conducted in three data collection
-periods. Two types of departures, namely non-linearity and
-heteroskedasticity, are considered with the corresponding data
-generating process being designed for data collection periods I and II.
-The data collection period III is designed for collecting human
-responses to null lineups such that the parameter \(\alpha\) in Equation
-\ref{eq:pvalue-beta-binomial} can be estimated. Overall, we plan to
-collect 7974 evaluations on 1152 unique lineups performed by 443
-subjects throughout three data collection periods.
+periods. The two types of departures, non-linearity and
+heteroskedasticity, are primarily collected during data collection
+periods I and II. The data collection period III was designed primarily
+to measure human responses to null lineups so that the parameter
+\(\alpha\) in Equation \ref{eq:pvalue-beta-binomial} can be estimated.
+Additional lineups for both non-linearity and heteroskedasticity, using
+uniform \(X\) distribution, were included so that the participants were
+evaluating some lineups with signal also. It would be too frustrating
+for participants to only be assigned lineups with all null plots.
+Overall, we collected 7974 evaluations on 1152 unique lineups performed
+by 443 subjects throughout three data collection periods.
 
@@
 
-\caption{Polynomial forms generated for the residual plots used in data collection period I.
+\caption{Polynomial forms generated for the residual plots used to assess detecting non-linearity.
 The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$.}\label{fig:different-shape-of-herimite}
@@
 Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups
-used in data collection period I. This lineup is produced by the
+used in non-linearity detection. This lineup is produced by the
 non-linearity model with \(j = 6\) and uniform fitted values. The data
@@
 
-\caption{Number of observations $n$ have an impact on the strength of the signal. Three different values of $n$ are used in data collection periods I, II and III.}\label{fig:different-n}
+\caption{Examining the effect of signal strength for the three different values of $n$ used in the experiment, for non-linear structure with fixed $\sigma = 1.5$, uniform $X$ distribution, and S shape.
+For these factor levels, only when $n=300$ is the S shape clearly visible.}\label{fig:different-n}
 \end{figure}
@@
 
-\caption{For the non-lienearity model, the standard deviation of the error term $\sigma$ affects the strength of the signal.
-Four different values of $\sigma$ are used in the data collection period I.}\label{fig:different-sigma}
+\caption{Examining the effect of $\sigma$ on the signal strength in the non-linearity detection, for $n=300$, uniform $X$ distribution and the U shape.
+As $\sigma$ increases the signal strength decreases, to the point that the U is almost unrecognisable when $\sigma=4$.}\label{fig:different-sigma}
 \end{figure}
@@
 
-\caption{Heteroskedasticity forms used in data collection period II.
-Three different shapes (a = -1, 0, 1) are used in the experiment to create left-triangle, butterfly and right-triangle shapes.}\label{fig:different-shape-of-heter}
+\caption{Heteroskedasticity forms used in the experiment.
+Three different shapes ($a = -1, 0, 1$) are used in the experiment to create left-triangle, butterfly and right-triangle shapes, respectively.}\label{fig:different-shape-of-heter}
 \end{figure}
@@
 
-\caption{Five different values of $b$ are used in data collection period II to control the strength of the signal.}\label{fig:different-b}
+\caption{Five different values of $b$ are used in heteroskedasticity simulation to control the strength of the signal.
+Larger values of $b$ yield a bigger difference in variation, and stus stronger heteroskedasticity signal.}\label{fig:different-b}
 \end{figure}
@@
 regression models. An experiment is conducted in three data collection
-periods with two simulation models designed for non-linearity and
-heteroskedasticity model defects being applied. Methods for obtaining
-the \(p\)-value of visual tests proposed by
-\citet{vanderplas2021statistical} are extended to allowed for multiple
-selections in a lineup protocol.
+periods using two simulation models, one for for non-linearity and and
+the other for heteroskedasticity. Methods for obtaining the \(p\)-value
+of visual tests proposed by \citet{vanderplas2021statistical} are
+extended to allow for multiple selections in a lineup protocol.
 

### 2023-02-07 paper_comparison/paper_comparison/paper_comparison.Rmd
cleaned up common factors section

@@
 
-```{r residual-plot-common-departures, fig.width = 10, fig.height = 2.5, fig.cap = "Example residual vs fitted plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
-The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted plot."}
+```{r residual-plot-common-departures, fig.width = 10, fig.height = 2.5, fig.cap = "Example residual vs fitted value plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
+The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted value plot."}
 
@@
 
-XXX DC will clean this up
-
-Four distributions of fitted values are also considered by using $X_{raw}$ follows different distributions: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
-
-Different distributions of fitted values help enriching the pool of visual patterns as illustrated in Figure \ref{fig:different-dist}.
+Fitted values are a function of the independent variables, and the distribution of the observed values affects the distribution of the fitted values.
+In the best case scenario the fitted values will have a uniform distribution, which means that there is even coverage of possible observed values across all of the predictors.
+This is not always present in the collected data.
+Sometimes the fitted values are discrete because one or more predictors were measured discretely.
+The distribution may be relatively Gaussian, reflecting a linear combinatio of many predictors, adhering to the Central Limit Theorem.
+It is also common to see a skewed distribution of fitted values, if one or more of the predictors has a skewed distribution.
+This latter problem is usually corrected before modeling using a variable transformation.
+Our simulation asesses this by using four different distributions to represent the fitted values ($X_{raw}$): (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
+We would expect that the best reading of residual plots occurs when the fitted values are uniformly distributed.
+
+<!--Different distributions of fitted values help enriching the pool of visual patterns as illustrated in Figure \ref{fig:different-dist}.
 The uniform and the normal distribution are symmetric and commonly assumed in statistical models.
-The adjusted log-normal distribution provides skewed density, while the discrete uniform distribution provides discreteness in residual plot.
+The adjusted log-normal distribution provides skewed density, while the discrete uniform distribution provides discreteness in residual plot.-->
+
+Three different sample sizes are used, $n=50, 100, 300$ across the experiments.
+We would expect considerable variation in the signal strength in the simulated data plots with smaller $n$.
+A sample size of 300 is typically enough for structure to be visible in a scatterplot reliably.
 

### 2023-02-07 paper_comparison/paper_comparison/paper_comparison.Rmd
update the conclusion

--- 2023-02-07 paper_comparison/paper_comparison/paper_comparison.Rmd
@@
 
-According to the data collected from the human subject experiment, residual plots rejected by visual tests are rarely not rejected by conventional tests, but departures of residuals that are of little practical sense to humans will often be alarmed by conventional tests, supporting the common guidance given by experts to utilize graphical representations of data in model diagnostics.
-
+According to the data collected from the human subject experiment, residual plots rejected by visual tests are rarely not rejected by conventional tests, but departures of residuals that are of little practical sense to humans are sometimes alarmed by conventional tests, supporting the common guidance given by experts to utilize graphical representations of data in model diagnostics.
 We have also found that residual-based conventional tests relying on a user-parameterized test formula, such as the RESET test, will suffer from a significant decrease in power for detecting certain residual departures intended to be tested but not covered by the test formula.
-This can not be avoided without prior knowledge of their functional forms, which is usually obtained by plotting the residuals.
+This is difficult to be avoided without prior knowledge of their functional forms, which is usually obtained by plotting the residuals.
 On the other hand, visual tests do not rely on such information and a collection of departures can be tested simultaneously.
-
-Interestingly, the orientation of the visual pattern
-
-fitted value distribution
-
-orientation of heter shape
+Interestingly, subjects recruited for our experiment behaves a potential favour of the orientation of the residual departure patterns.
+They perform slightly better in identifying triangle residual patterns pointing to the left than to the right.
+Besides, it is found that fitted value distributions have a great impact on the power of both conventional and visual tests.
+Generally, fitted values followed an evenly distributed and symmetric distribution such as the uniform distribution is preferred.
+Depending on the type of the residual departures, discreteness may cause trouble to humans to identify the residual patterns, while conventional tests are not affected.
+Skewness on the other hand significantly decrease the power of both conventional and visual tests when testing heteroskedasticity.
+Although it is commonly assumed by analysts, a normal fitted value distribution may not be the optimal option when comes to regression diagnostics.
 

### 2023-02-07 paper_comparison/paper_comparison/paper_comparison.tex

@@
 
-\caption{Example residual vs fitted plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
-The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted plot.}\label{fig:residual-plot-common-departures}
+\caption{Example residual vs fitted value plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
+The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted value plot.}\label{fig:residual-plot-common-departures}
 \end{figure}
@@
 
-XXX DC will clean this up
-
-Four distributions of fitted values are also considered by using
-\(X_{raw}\) follows different distributions: (1) \(U(-1, 1)\), (2)
-\(N(0, 0.3^2)\), (3) \(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\).
-
-Different distributions of fitted values help enriching the pool of
-visual patterns as illustrated in Figure \ref{fig:different-dist}. The
-uniform and the normal distribution are symmetric and commonly assumed
-in statistical models. The adjusted log-normal distribution provides
-skewed density, while the discrete uniform distribution provides
-discreteness in residual plot.
+Fitted values are a function of the independent variables, and the
+distribution of the observed values affects the distribution of the
+fitted values. In the best case scenario the fitted values will have a
+uniform distribution, which means that there is even coverage of
+possible observed values across all of the predictors. This is not
+always present in the collected data. Sometimes the fitted values are
+discrete because one or more predictors were measured discretely. The
+distribution may be relatively Gaussian, reflecting a linear combinatio
+of many predictors, adhering to the Central Limit Theorem. It is also
+common to see a skewed distribution of fitted values, if one or more of
+the predictors has a skewed distribution. This latter problem is usually
+corrected before modeling using a variable transformation. Our
+simulation asesses this by using four different distributions to
+represent the fitted values (\(X_{raw}\)): (1) \(U(-1, 1)\), (2)
+\(N(0, 0.3^2)\), (3) \(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). We
+would expect that the best reading of residual plots occurs when the
+fitted values are uniformly distributed.
+
+Three different sample sizes are used, \(n=50, 100, 300\) across the
+experiments. We would expect considerable variation in the signal
+strength in the simulated data plots with smaller \(n\). A sample size
+of 300 is typically enough for structure to be visible in a scatterplot
+reliably.
 

### 2023-02-08 paper_comparison/paper_comparison/paper_comparison.tex
compile the pdf

@@
 conventional tests, but departures of residuals that are of little
-practical sense to humans will often be alarmed by conventional tests,
+practical sense to humans are sometimes alarmed by conventional tests,
 supporting the common guidance given by experts to utilize graphical
-representations of data in model diagnostics.
-
-We have also found that residual-based conventional tests relying on a
-user-parameterized test formula, such as the RESET test, will suffer
-from a significant decrease in power for detecting certain residual
-departures intended to be tested but not covered by the test formula.
-This can not be avoided without prior knowledge of their functional
-forms, which is usually obtained by plotting the residuals. On the other
-hand, visual tests do not rely on such information and a collection of
-departures can be tested simultaneously.
-
-Interestingly, the orientation of the visual pattern
-
-fitted value distribution
-
-orientation of heter shape
+representations of data in model diagnostics. We have also found that
+residual-based conventional tests relying on a user-parameterized test
+formula, such as the RESET test, will suffer from a significant decrease
+in power for detecting certain residual departures intended to be tested
+but not covered by the test formula. This is difficult to be avoided
+without prior knowledge of their functional forms, which is usually
+obtained by plotting the residuals. On the other hand, visual tests do
+not rely on such information and a collection of departures can be
+tested simultaneously. Interestingly, subjects recruited for our
+experiment behaves a potential favour of the orientation of the residual
+departure patterns. They perform slightly better in identifying triangle
+residual patterns pointing to the left than to the right. Besides, it is
+found that fitted value distributions have a great impact on the power
+of both conventional and visual tests. Generally, fitted values followed
+an evenly distributed and symmetric distribution such as the uniform
+distribution is preferred. Depending on the type of the residual
+departures, discreteness may cause trouble to humans to identify the
+residual patterns, while conventional tests are not affected. Skewness
+on the other hand significantly decrease the power of both conventional
+and visual tests when testing heteroskedasticity. Although it is
+commonly assumed by analysts, a normal fitted value distribution may not
+be the optimal option when comes to regression diagnostics.
 

### 2023-02-08 paper_comparison/paper_comparison/paper_comparison.Rmd
draft of conclusion more complete

--- 2023-02-07 paper_comparison/paper_comparison/paper_comparison.Rmd
@@
 
-1. summarise paper content
-
-This paper has demonstrated that conventional residual-based statistical tests are more sensitive to departures from assumptions for residuals than visual tests evaluated by humans in the diagnostics of linear regression models.
+<!--1. summarise paper content-->
+
+Motivated by the advice of regression analysis experts, that residual plots as opposed to conventional tests are an indispensible methods for assessing model fit, a human subjects experiment was conducted using visual inference.
+The experiment tested two primary departures from good residuals: non-linearity and heteroskedasticity.
+
+<!-- This paper has demonstrated that conventional residual-based statistical tests are more sensitive to departures from assumptions for residuals than visual tests evaluated by humans in the diagnostics of linear regression models.
 An experiment is conducted in three data collection periods using two simulation models, one for for non-linearity and and the other for heteroskedasticity.
-Methods for obtaining the $p$-value of visual tests proposed by @vanderplas2021statistical are extended to allow for multiple selections in a lineup protocol.
-
-2. briefly describe what was learned
-
-According to the data collected from the human subject experiment, residual plots rejected by visual tests are rarely not rejected by conventional tests, but departures of residuals that are of little practical sense to humans will often be alarmed by conventional tests, supporting the common guidance given by experts to utilize graphical representations of data in model diagnostics.
-
-We have also found that residual-based conventional tests relying on a user-parameterized test formula, such as the RESET test, will suffer from a significant decrease in power for detecting certain residual departures intended to be tested but not covered by the test formula.
+Methods for obtaining the $p$-value of visual tests proposed by @vanderplas2021statistical are extended to allow for multiple selections in a lineup protocol.-->
+
+<!--2. briefly describe what was learned-->
+
+The experiment found that conventional residual-based statistical tests are more sensitive to departures from assumptions for residuals than visual tests as would be evaluated by humans.
+That is, conventional tests conclude there are problems with the model fit almost twice as often as humans would.
+They often reject when departures in the form of non-linearity and heteroskedasticity are not visible to a human.
+
+One might say that this is correct behaviour, but it can be argued that the conventional tests are rejecting when it is not necessary.
+Many of these rejections happen even when downstream analysis and results would not be significantly affected by the small departures from a good fit.
+The results from human evaluations provide a more practical solution, which reinforces the statements from regression experts that residual plots are an indispensible method for model diagnostics.
+
+Now it is important to note that residual plots need to be delivered as a line up, where it is embedded in a field of null plots.
+This enables a careful calibration for reading structure in residual plots.
+XXX More here
+
+However, human evaluation of residuals is expensive.
+It is time-consuming, and laborious.
+This is another reason why it often appears to be ignored.
+With the availability of sophisticated computer vision algorithms today, the goal of this work is to form the basis of providing automated residual plot reading.
+XXX More here
+
+The experiment also revealed some interesting details about how residual plots are read.
+For the most part, the visual test performed very similarly to the (best) conventional test only with the power curve shifted in the less sensitive direction.
+Unlike the conventional tests, where one needs to specifically test for non-linearity or heteroskedasticity the visual test operated effectively across the range of departures from good residuals.
+
+As expected, if the fitted value distribution is not uniform, there is a loss of power in the visual test.
+Structure is hardest to detect if fitted values ar discrete.
+Also, a surprising find was that the direction of heteroskedasticity appears to affect the ability to visually detect it, with wedge to the left being less detectable.
+
+<!--According to the data collected from the human subject experiment, residual plots rejected by visual tests are rarely not rejected by conventional tests, but departures of residuals that are of little practical sense to humans will often be alarmed by conventional tests, supporting the common guidance given by experts to utilize graphical representations of data in model diagnostics.
+-->
+
+<!-- We have also found that residual-based conventional tests relying on a user-parameterized test formula, such as the RESET test, will suffer from a significant decrease in power for detecting certain residual departures intended to be tested but not covered by the test formula.
 This can not be avoided without prior knowledge of their functional forms, which is usually obtained by plotting the residuals.
@@
 orientation of heter shape
+-->
 

### 2023-02-08 paper_comparison/paper_comparison/paper_comparison.tex
draft of conclusion more complete

--- 2023-02-07 paper_comparison/paper_comparison/paper_comparison.tex
@@
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  summarise paper content
-\end{enumerate}
-
-This paper has demonstrated that conventional residual-based statistical
-tests are more sensitive to departures from assumptions for residuals
-than visual tests evaluated by humans in the diagnostics of linear
-regression models. An experiment is conducted in three data collection
-periods using two simulation models, one for for non-linearity and and
-the other for heteroskedasticity. Methods for obtaining the \(p\)-value
-of visual tests proposed by \citet{vanderplas2021statistical} are
-extended to allow for multiple selections in a lineup protocol.
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{1}
-\tightlist
-\item
-  briefly describe what was learned
-\end{enumerate}
-
-According to the data collected from the human subject experiment,
-residual plots rejected by visual tests are rarely not rejected by
-conventional tests, but departures of residuals that are of little
-practical sense to humans will often be alarmed by conventional tests,
-supporting the common guidance given by experts to utilize graphical
-representations of data in model diagnostics.
-
-We have also found that residual-based conventional tests relying on a
-user-parameterized test formula, such as the RESET test, will suffer
-from a significant decrease in power for detecting certain residual
-departures intended to be tested but not covered by the test formula.
-This can not be avoided without prior knowledge of their functional
-forms, which is usually obtained by plotting the residuals. On the other
-hand, visual tests do not rely on such information and a collection of
-departures can be tested simultaneously.
-
-Interestingly, the orientation of the visual pattern
-
-fitted value distribution
-
-orientation of heter shape
+Motivated by the advice of regression analysis experts, that residual
+plots as opposed to conventional tests are an indispensible methods for
+assessing model fit, a human subjects experiment was conducted using
+visual inference. The experiment tested two primary departures from good
+residuals: non-linearity and heteroskedasticity.
+
+The experiment found that conventional residual-based statistical tests
+are more sensitive to departures from assumptions for residuals than
+visual tests as would be evaluated by humans. That is, conventional
+tests conclude there are problems with the model fit almost twice as
+often as humans would. They often reject when departures in the form of
+non-linearity and heteroskedasticity are not visible to a human.
+
+One might say that this is correct behaviour, but it can be argued that
+the conventional tests are rejecting when it is not necessary. Many of
+these rejections happen even when downstream analysis and results would
+not be significantly affected by the small departures from a good fit.
+The results from human evaluations provide a more practical solution,
+which reinforces the statements from regression experts that residual
+plots are an indispensible method for model diagnostics.
+
+Now it is important to note that residual plots need to be delivered as
+a line up, where it is embedded in a field of null plots. This enables a
+careful calibration for reading structure in residual plots. XXX More
+here
+
+However, human evaluation of residuals is expensive. It is
+time-consuming, and laborious. This is another reason why it often
+appears to be ignored. With the availability of sophisticated computer
+vision algorithms today, the goal of this work is to form the basis of
+providing automated residual plot reading. XXX More here
+
+The experiment also revealed some interesting details about how residual
+plots are read. For the most part, the visual test performed very
+similarly to the (best) conventional test only with the power curve
+shifted in the less sensitive direction. Unlike the conventional tests,
+where one needs to specifically test for non-linearity or
+heteroskedasticity the visual test operated effectively across the range
+of departures from good residuals.
+
+As expected, if the fitted value distribution is not uniform, there is a
+loss of power in the visual test. Structure is hardest to detect if
+fitted values ar discrete. Also, a surprising find was that the
+direction of heteroskedasticity appears to affect the ability to
+visually detect it, with wedge to the left being less detectable.
 

### 2023-02-13 paper_comparison/paper_comparison/paper_comparison.Rmd
fix effect size issue

--- 2023-02-07 paper_comparison/paper_comparison/paper_comparison.Rmd
@@
                      e_sigma == {{e_sigma}},
-                     x_dist == {{x_dist}},
                      n == {{n}}) %>%
@@
                      b == {{b}},
-                     x_dist == {{x_dist}},
                      n == {{n}}) %>%
@@
 
+(use uniform effect size? normal effect size? -> other dist?)
+
+(average out dist for effect size calc?)
+
 The power of a model misspecification test is the probability that the null hypothesis is rejected given the regression model is misspecified.
@@
 \begin{equation} \label{eq:logistic-regression-1-1}
-Pr(\text{reject}~H_0|H_1,E) = \Lambda(log(\frac{0.5}{0.95}) + \beta_1 E),
+Pr(\text{reject}~H_0|H_1,E) = \Lambda(log(\frac{0.05}{0.95}) + \beta_1 E),
 \end{equation}
@@
 # with expected effect 0 ~ 5
-target_sigma <- c(4.51, 3.52, 2.74, 2.14, 1.66, 1.29, 1.01, 0.79, 0.61, 0.48, 0.37)
+
+target_sigma <- c(3.65, 2.84, 2.21, 1.72, 1.34, 1.04, 0.81, 0.63, 0.49, 0.38, 0.3)
 
@@
               ~log(poly_model(shape = 2, 
-                              x = rand_uniform(-1, 1), 
+                              x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, 
                               sigma = .x)$
-                     effect_size(n = 100)
+                     average_effect_size(n = 100)
                    )
@@
 ex_dat <- poly_model(shape = 2, 
-                     x = rand_uniform(), 
+                     x = rand_uniform(-1, 1), 
                      sigma = 1)$gen(n = 100)
@@
                            family = binomial(), 
-                           data = visual_poly_dat, 
+                           data = visual_heter_dat, 
                            offset = offset0)
@@
 # with expected effect 0 ~ 5
-target_b <- c(0.14, 0.19, 0.27, 0.38, 0.57, 0.91, 1.53, 2.9, 6.4, 17.9, 75)
+
+target_b <- c(0.17, 0.23, 0.32, 0.46, 0.67, 1.04, 1.7, 3, 6.3, 16.2, 61)
 
@@
                                b = .x, 
-                               x = rand_uniform(-1, 1))$
-                     effect_size(n = 100, type = "kl")
+                               x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))})$
+                     average_effect_size(n = 100, type = "kl")
                    )
@@
 ## Effect of shape of heteroskedasticity
+
+(change the butterfly color)
 

### 2023-02-14 paper_comparison/paper_comparison/paper_comparison.Rmd
update the power desc

@@
 \begin{equation} \label{eq:logistic-regression-1-1}
-Pr(\text{reject}~H_0|H_1,E) = \Lambda(log(\frac{0.05}{0.95}) + \beta_1 E),
+Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right),
 \end{equation}
@@
 \noindent where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
-The effect size is the only regressor and the intercept is fixed to $log(0.5/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$.
-
-Effect size $E$ is derived from the Kullback-Leibler divergence (see Appendix for the details) formulated as
+The effect size is the only regressor and the intercept is fixed to $log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$.
+
+Effect size $E$ is formulated as the conditional expectation of the sample-based effect size $E_{sample}$ derived from the Kullback-Leibler divergence (see Appendix for the details) given by
 
@@
 \begin{gather}  \label{eq:effect-size}
-E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V})|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V})^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V})^{-1}\boldsymbol{\mu}_z\right),\\
+E = E[E_{sample}|M],\\
+E_{sample} = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V})|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V})^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V})^{-1}\boldsymbol{\mu}_z\right),\\
 \boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a,\\
@@
 
-\noindent where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,  $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of regressors including the intercept used in the regression equation, $\sigma^2\boldsymbol{I}$ is the assumed variance of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual variance of the error term, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of the residuals with $\boldsymbol{Z}$ be any variables leave out by the model and $\boldsymbol{\beta}_z$ be the corresponding coefficients.
+\noindent where $M$ is a given model that generates the data, $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,  $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of regressors including the intercept used in the regression equation, $\sigma^2\boldsymbol{I}$ is the assumed variance of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual variance of the error term, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of the residuals with $\boldsymbol{Z}$ be any variables leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients.
+An expectation is taken because $E_{sample}$ depends on the regressors $X$ and we want the effect size $E$ to be the same for the same data generating process.
+It is difficult to solve the effect size $E$ analytically, so we estimate it with the average of $E_{sample}$ over a very large sample size.
 

### 2023-02-14 paper_comparison/paper_comparison/paper_comparison.tex
update the compiled paper
update the power desc

@@
 
+(use uniform effect size? normal effect size? -\textgreater{} other
+dist?)
+
+(average out dist for effect size calc?)
+
 The power of a model misspecification test is the probability that the
@@
 \begin{equation} \label{eq:logistic-regression-1-1}
-Pr(\text{reject}~H_0|H_1,E) = \Lambda(log(\frac{0.5}{0.95}) + \beta_1 E),
+Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right),
 \end{equation}
@@
 as \(\Lambda(z) = exp(z)/(1+exp(z))\). The effect size is the only
-regressor and the intercept is fixed to \(log(0.5/0.95)\) so that
+regressor and the intercept is fixed to \(log(0.05/0.95)\) so that
 \(\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05\).
 
-Effect size \(E\) is derived from the Kullback-Leibler divergence (see
-Appendix for the details) formulated as
+Effect size \(E\) is formulated as the conditional expectation of the
+sample-based effect size \(E_{sample}\) derived from the
+Kullback-Leibler divergence (see Appendix for the details) given by
 
@@
 
-\noindent where \(diag(.)\) is the diagonal matrix constructed from the
-diagonal elements of a matrix,
+\noindent where \(M\) is a given model that generates the data,
+\(diag(.)\) is the diagonal matrix constructed from the diagonal
+elements of a matrix,
 \(\boldsymbol{X}_a = (\boldsymbol{1}, \boldsymbol{X})\) is the matrix of
@@
 is the expected values of the residuals with \(\boldsymbol{Z}\) be any
-variables leave out by the model and \(\boldsymbol{\beta}_z\) be the
-corresponding coefficients.
+variables leave out by the regression equation and
+\(\boldsymbol{\beta}_z\) be the corresponding coefficients. An
+expectation is taken because \(E_{sample}\) depends on the regressors
+\(X\) and we want the effect size \(E\) to be the same for the same data
+generating process. It is difficult to solve the effect size \(E\)
+analytically, so in the analysis, we estimate it with the average of
+\(E_{sample}\) over a very large sample size.
 
@@
 
+(change the butterfly color)
+
 We have also investigated the impact of different heteroskedasticity

### 2023-02-15 paper_comparison/paper_comparison/paper_comparison.Rmd
clarify the power and the simulated conventional power

@@
 \noindent where $M$ is a given model that generates the data, $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,  $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of regressors including the intercept used in the regression equation, $\sigma^2\boldsymbol{I}$ is the assumed variance of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual variance of the error term, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of the residuals with $\boldsymbol{Z}$ be any variables leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients.
-An expectation is taken because $E_{sample}$ depends on the regressors $X$ and we want the effect size $E$ to be the same for the same data generating process.
+A conditional expectation is taken because $E_{sample}$ depends on the regressors $\boldsymbol{X}$ and we want the effect size $E$ to be the same for the same data generating process.
 It is difficult to solve the effect size $E$ analytically, so we estimate it with the average of $E_{sample}$ over a very large sample size.
@@
 Visual patterns are more likely to be revealed under a uniform distribution.
-Additionally, we have collected extra evaluations on these lineups, which will result in more reliable analysis.
+Additionally, we have collected extra evaluations on these lineups, which will result in a more reliable analysis.
 Analysis of lineups with other fitted value distributions can be found in Section \ref{effect-of-fitted-value-distributions}.
@@
 
-To compare the power of visual test and conventional test, we pick RESET test (`resettest`) and Breusch-Pagan test (`bptest`) from the R package `lmtest`, and Shapiro-Wilk test (`shapiro.test`) from the built-in R package `stats`.
+To compare the power of visual tests and conventional tests, we simulate 200000 samples evenly from the models used in the human subject experiments, and record the conventional $p$-values produced from every sample.
+We pick RESET test (`resettest`) and Breusch-Pagan test (`bptest`) from the R package `lmtest`, and Shapiro-Wilk test (`shapiro.test`) from the built-in R package `stats`.
 Among them, RESET test is the only exact and appropriate test in this scenario.
@@
 Thus, we include both $\boldsymbol{x}$ and $\boldsymbol{x}^2$ in the auxiliary regression.
+
+To get an accurate estimate of the power of conventional tests, we simulate 200000 samples evenly from the models used in the human subject experiments, and record the $p$-values produced by every sample.
 

### 2023-02-15 paper_comparison/paper_comparison/paper_comparison.tex
clarify the power and the simulated conventional power

@@
 variables leave out by the regression equation and
-\(\boldsymbol{\beta}_z\) be the corresponding coefficients. An
-expectation is taken because \(E_{sample}\) depends on the regressors
-\(X\) and we want the effect size \(E\) to be the same for the same data
-generating process. It is difficult to solve the effect size \(E\)
-analytically, so in the analysis, we estimate it with the average of
-\(E_{sample}\) over a very large sample size.
+\(\boldsymbol{\beta}_z\) be the corresponding coefficients. A
+conditional expectation is taken because \(E_{sample}\) depends on the
+regressors \(\boldsymbol{X}\) and we want the effect size \(E\) to be
+the same for the same data generating process. It is difficult to solve
+the effect size \(E\) analytically, so we estimate it with the average
+of \(E_{sample}\) over a very large sample size.
 

### 2023-02-15 paper_comparison/paper_comparison/paper_comparison.Rmd
mention the regressor dist is assumed to be the same

@@
 
-(use uniform effect size? normal effect size? -> other dist?)
-
-(average out dist for effect size calc?)
-
 The power of a model misspecification test is the probability that the null hypothesis is rejected given the regression model is misspecified.
@@
 \noindent where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
-The effect size is the only regressor and the intercept is fixed to $log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$.
+The effect size $E$ is the only regressor and the intercept is fixed to $log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$.
 
@@
 
-\noindent where $M$ is a given model that generates the data, $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,  $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of regressors including the intercept used in the regression equation, $\sigma^2\boldsymbol{I}$ is the assumed variance of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual variance of the error term, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of the residuals with $\boldsymbol{Z}$ be any variables leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients.
-A conditional expectation is taken because $E_{sample}$ depends on the regressors $\boldsymbol{X}$ and we want the effect size $E$ to be the same for the same data generating process.
-It is difficult to solve the effect size $E$ analytically, so we estimate it with the average of $E_{sample}$ over a very large sample size.
+\noindent where $M$ is a given model that generates the data, $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,  $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of regressors including the intercept used in the regression equation, $\sigma^2\boldsymbol{I}$ is the assumed variance of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual variance of the error term, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of the residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients.
+
+A conditional expectation is taken because $E$ is designed to be invariant to different realizations of $\boldsymbol{X}$.
+However, solving this conditional expectation analytically is difficult as $\boldsymbol{X}$ can follow almost any distribution.
+We make an assumption that $\boldsymbol{X}$ follows a particular multivariate normal distribution, and repeatedly draw samples from the given model $M$.
+The estimate of $E$ is obtained by taking the average of the $E_{sample}$ calculated for each sample.
 
 To study various factors contributing to the power of the visual test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors.
-These include the distribution of the regressor and the type of the simulation model (modify this according to the result).
+These include the distribution of the regressor, type of the simulation model and the shape of the residual departures.
 

### 2023-02-15 paper_comparison/paper_comparison/paper_comparison.tex
mention the regressor dist is assumed to be the same

--- 2023-02-14 paper_comparison/paper_comparison/paper_comparison.tex
@@
 
-(use uniform effect size? normal effect size? -\textgreater{} other
-dist?)
-
-(average out dist for effect size calc?)
-
 The power of a model misspecification test is the probability that the
@@
 \noindent where \(\Lambda(.)\) is the standard logistic function given
-as \(\Lambda(z) = exp(z)/(1+exp(z))\). The effect size is the only
+as \(\Lambda(z) = exp(z)/(1+exp(z))\). The effect size \(E\) is the only
 regressor and the intercept is fixed to \(log(0.05/0.95)\) so that
@@
 is the expected values of the residuals with \(\boldsymbol{Z}\) be any
-variables leave out by the regression equation and
-\(\boldsymbol{\beta}_z\) be the corresponding coefficients. An
-expectation is taken because \(E_{sample}\) depends on the regressors
-\(X\) and we want the effect size \(E\) to be the same for the same data
-generating process. It is difficult to solve the effect size \(E\)
-analytically, so in the analysis, we estimate it with the average of
-\(E_{sample}\) over a very large sample size.
+higher order terms of \(\boldsymbol{X}\) leave out by the regression
+equation and \(\boldsymbol{\beta}_z\) be the corresponding coefficients.
+
+A conditional expectation is taken because \(E\) is designed to be
+invariant to different realizations of \(\boldsymbol{X}\). However,
+solving this conditional expectation analytically is difficult as
+\(\boldsymbol{X}\) can follow almost any distribution. We make an
+assumption that \(\boldsymbol{X}\) follows a particular multivariate
+normal distribution, and repeatedly draw samples from the given model
+\(M\). The estimate of \(E\) is obtained by taking the average of the
+\(E_{sample}\) calculated for each sample.
 
@@
 collated data grouped by levels of factors. These include the
-distribution of the regressor and the type of the simulation model
-(modify this according to the result).
+distribution of the regressor, type of the simulation model and the
+shape of the residual departures.
 
@@
 uniform distribution. Additionally, we have collected extra evaluations
-on these lineups, which will result in more reliable analysis. Analysis
-of lineups with other fitted value distributions can be found in Section
-\ref{effect-of-fitted-value-distributions}.
+on these lineups, which will result in a more reliable analysis.
+Analysis of lineups with other fitted value distributions can be found
+in Section \ref{effect-of-fitted-value-distributions}.
 
@@
 
-To compare the power of visual test and conventional test, we pick RESET
-test (\texttt{resettest}) and Breusch-Pagan test (\texttt{bptest}) from
-the R package \texttt{lmtest}, and Shapiro-Wilk test
-(\texttt{shapiro.test}) from the built-in R package \texttt{stats}.
-Among them, RESET test is the only exact and appropriate test in this
-scenario. Both the Breusch-Pagan test and the Shapiro-Wilk test are
-approximate and inappropriate tests. Their estimated power is shown in
-Figure \ref{fig:polypower}. To set up the RESET test, we include
-different powers of fitted values as proxies. According to
+To compare the power of visual tests and conventional tests, we simulate
+200000 samples evenly from the models used in the human subject
+experiments, and record the conventional \(p\)-values produced from
+every sample. We pick RESET test (\texttt{resettest}) and Breusch-Pagan
+test (\texttt{bptest}) from the R package \texttt{lmtest}, and
+Shapiro-Wilk test (\texttt{shapiro.test}) from the built-in R package
+\texttt{stats}. Among them, RESET test is the only exact and appropriate
+test in this scenario. Both the Breusch-Pagan test and the Shapiro-Wilk
+test are approximate and inappropriate tests. Their estimated power is
+shown in Figure \ref{fig:polypower}. To set up the RESET test, we
+include different powers of fitted values as proxies. According to
 \citet{ramsey_tests_1969}, there are no general rules for the power of
@@
 \(\boldsymbol{x}\) and \(\boldsymbol{x}^2\) in the auxiliary regression.
+
+To get an accurate estimate of the power of conventional tests, we
+simulate 200000 samples evenly from the models used in the human subject
+experiments, and record the \(p\)-values produced by every sample.
 

### 2023-02-15 paper_comparison/paper_comparison/paper_comparison.Rmd
use more accurate lang

@@
 A conditional expectation is taken because $E$ is designed to be invariant to different realizations of $\boldsymbol{X}$.
-However, solving this conditional expectation analytically is difficult as $\boldsymbol{X}$ can follow almost any distribution.
-We make an assumption that $\boldsymbol{X}$ follows a particular multivariate normal distribution, and repeatedly draw samples from the given model $M$.
-The estimate of $E$ is obtained by taking the average of the $E_{sample}$ calculated for each sample.
+However, solving this conditional expectation analytically is very difficult as $\boldsymbol{X}$ can follow arbitrary distributions.
+Therefore, we make an assumption that $\boldsymbol{X}$ follows a particular multivariate normal distribution
+
+We make an assumption that all the realizations of $\boldsymbol{X}$ are generated from a multivariate normal distribution $N(\boldsymbol{0}, 0.3^2\boldsymbol{I})$ and then rescaled by the min-max normalization.
+The estimate of $E$ is obtained by taking the average of the $E_{sample}$ calculated for samples drawn from the given model $M$ repeatedly.
 

### 2023-02-15 paper_comparison/paper_comparison/paper_comparison.tex
use more accurate lang

@@
 invariant to different realizations of \(\boldsymbol{X}\). However,
-solving this conditional expectation analytically is difficult as
-\(\boldsymbol{X}\) can follow almost any distribution. We make an
-assumption that \(\boldsymbol{X}\) follows a particular multivariate
-normal distribution, and repeatedly draw samples from the given model
-\(M\). The estimate of \(E\) is obtained by taking the average of the
-\(E_{sample}\) calculated for each sample.
+solving this conditional expectation analytically is very difficult as
+\(\boldsymbol{X}\) can follow arbitrary distributions. Therefore, we
+make an assumption that \(\boldsymbol{X}\) follows a particular
+multivariate normal distribution
+
+We make an assumption that all the realizations of \(\boldsymbol{X}\)
+are generated from a multivariate normal distribution
+\(N(\boldsymbol{0}, 0.3^2\boldsymbol{I})\) and then rescaled by the
+min-max normalization. The estimate of \(E\) is obtained by taking the
+average of the \(E_{sample}\) calculated for samples drawn from the
+given model \(M\) repeatedly.
 

### 2023-02-15 paper_comparison/paper_comparison/paper_comparison.Rmd
update the text according to the updated effect size

@@
 The estimate of $E$ is obtained by taking the average of the $E_{sample}$ calculated for samples drawn from the given model $M$ repeatedly.
+We use the normal distribution as a general choice to approximate the conditional effect size.
 
@@
 
-To get an accurate estimate of the power of conventional tests, we simulate 200000 samples evenly from the models used in the human subject experiments, and record the $p$-values produced by every sample.
-
 Figure \ref{fig:heterpower} is similar to Figure \ref{fig:polypower}, but shows corresponding information on lineups produced by the heteroskedasticity model.
@@
 `r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(0), offset0 = log(0.05/0.95)), type = "response")) * 100)`% to `r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(2.5), offset0 = log(0.05/0.95)), type = "response")) * 100)`% as $log_e(E)$ increases from 0 to 2.5, while power of other tests respond inactively to the change of effect, showing that RESET test is way more sensitive to the type of model defects that being considered.
-Meanwhile, no noticeable residual depatures can be spotted from the example residual plots.
-
-In terms of heteroskedasticity patterns, the power of Breusch-Pagan test is also almost always greater than the power of visual test.
-For $log_e(E)$ at around 2.5, where the power curve of the visual test remains at a low level, the Breusch-Pagan test still have a decent amount of chance of rejecting $H_0$.
+Meanwhile, no noticeable residual departures can be spotted from the example residual plots.
+
+In terms of heteroskedasticity patterns, the power of Breusch-Pagan test is almost always greater than the power of visual test.
+For $log_e(E)$ at around 2.5, where the power curve of the visual test remains at a low level, the Breusch-Pagan test has around 50% chance of rejecting $H_0$.
 Similarly, the visual feature is nearly unobservable from the example residual plots.
@@
 Where would you draw the line of too much non-linearity in the residuals?
-For the RESET test this is around log effect size 2.5, but for the visual test it is around 3.5."}
+For the RESET test this is around log effect size 2.5, but for the visual test it is around 3.5.", dev = 'png', dpi = 450}
 
@@
 Where would you draw the line of too much heteroskedasticity in the residuals?
-For the BP test this is around log effect size 2.5, but for the visual test it is around 3.5."}
+For the BP test this is around log effect size 2.5, but for the visual test it is around 3.5.", dev = 'png', dpi = 450}
 
@@
 The data plot at panel 17 contains a "Butterfly" shape.
-It has effect size $ = 4.02$, somewhat surprising that it is not detected by the BP test.'}
+It has effect size $ = 3.76$, somewhat surprising that it is not detected by the BP test.'}
 # plot heter-331
@@
 A primary factor contributes to the non-linearity model is the shape of the non-linearity pattern.
-According to Figure \ref{fig:poly-power-uniform-j}, conventional tests have identical power in testing the "U", the "S" and the "M" shape.
-But it has significant lower power in testing the "Triple-U" shape.
+According to Figure \ref{fig:poly-power-uniform-j}, conventional tests have higher power in testing shapes constructed from lower orders of Hermite polynomials.
+But the power in testing the "Triple-U" shape drops significantly.
 To understand why this is, one needs to return to the way the RESET test is applied.
@@
 However, the "Triple-U" shape constructed from the Hermite polynomials use power up to 18.
-If the RESET test had been applied using a higher power no less than six, the power curve of "Triple-U" shape will have little difference to other shapes.
+If the RESET test had been applied using a higher power no less than six, the power curve of "Triple-U" shape will only slight lower than the power curve of "M" shape.
 The recommendation of the polynomial power for the RESET should be revised, perhaps.
@@
 For visual tests, based on the orders of the Hermite polynomials, we expect the "U" shape will be the easiest one to be detected by subjects followed by the "S", "M" and "Triple-U" shape.
-From Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power curves are mostly aligned with our expectation, except for the "M" shape, which is as easy to be detected as the "U" shape.
-This implies the readability of the shape do not strictly follow the degree of the polynomials.
+From Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power curves are mostly aligned with our expectation, except for the "M" shape, which is as easy to be detected as the "S" shape.
+This implies, unlike the conventional test, the readability of the shape do not strictly follow the degree of the polynomials.
 
@@
 
-(change the butterfly color)
-
 We have also investigated the impact of different heteroskedasticity shapes on power of conventional tests and visual tests.
@@
 As shown in Figure \ref{fig:heter-power-uniform-a}, this is indeed the case where little difference between the power curves can be perceived.
-Similarly, visual tests should have the smae power in detecting these two shapes if they are equally likely to be identified.
-However, it can be observed from Figure \ref{fig:heter-power-uniform-a} that the power curve for the "Left-triangle" shape is constantly higher than the one for the "Right-triangle" shape, indicating a potential favour of orientation by human, which is worth to be explored in future studies.
-Besides, the chance of detecting the "Butterfly" shape by visual tests are slightly higher than other two shapes.
+Similarly, visual tests should have the same power in detecting these two shapes if they are equally likely to be identified.
+However, it can be observed from Figure \ref{fig:heter-power-uniform-a} that the power curve for the "Left-triangle" shape is higher than the one for the "Right-triangle" shape, indicating a potential favour of orientation by human, which is worth to be explored in future studies.
+Besides, the chance of detecting the "Butterfly" shape by both conventional and visual tests are much higher than other two shapes.
 
@@
   ylab("Power") +
-  scale_color_manual(values = rcartocolor::carto_pal(3, "Earth")) +
+  scale_color_manual(values = c(rcartocolor::carto_pal(3, "Earth")[1], "grey", rcartocolor::carto_pal(3, "Earth")[3])) +
   theme_light() +
@@
 
-For non-linearity patterns, we do not observe significant difference in power of visual test between distributions, except for the discrete uniform distribution.
-It could be due to the fact that humans have difficulties in recognising visual patterns from discreteness, which makes the shape disconnected and incomplete (ref?).
-In terms of conventional tests, the discrete uniform distribution is instead the one with the greatest power, followed by lognormal, uniform and the normal distribution.
+For visual tests, the power curves for normal distribution and lognormal distribution are lower than other two distributions, which are all even.
+For non-linearity patterns, the power curve of normal distribution is slightly greater than the one for lognormal distribution.
+And the margin is even larger for heteroskedasticity patterns.
+This suggests skewness can affect the degree of presence of the visual pattern.
+
+For non-linearity patterns, we do not observe significant difference in power of visual test between uniform distribution and discrete uniform distribution, but the normal distribution and the lognoraml distribution have lower power than the previous two.
+In terms of conventional tests, the discrete uniform distribution is the one with the greatest power, followed by lognormal, uniform and the normal distribution.
 
@@
 This is as expected since other three distributions could reduce the chance of revealing the underlying visual pattern because of uneven data points.
-Although the power of visual test under the discrete uniform distribution is the lowest in the case of non-linearity patterns, it has a relatively great power this time.
+The power of visual test under the discrete uniform distribution has a relatively great.
 Considering for heteroskedasticity, the visual patterns are usually detected by connecting the maximum and minimum residuals separately at different fitted values.
 It will not be greatly affected by the discrete uniform distribution compared to the uniform distribution.
-Visual tests have low power on lineups with the normal distribution and the lognormal distribution.
-This suggests the use of transformation of fitted values in reading residual plots.
+Again, visual tests have low power on lineups with the normal distribution and the lognormal distribution.
+It indicates skewness and unevenness can reduce the detection rate of the visual pattern.
 In the case of conventional tests, the discrete uniform distribution consistently perform well, followed by uniform, normal and the lognormal distribution.
@@
 The discrete fitted value distribution substantially lowers the power of visual tests for non-linearity patterns in residual plots.
-For heteroskedasticity patterns, normal and lognormal distributions produce lower power for both types of tests."}
+For heteroskedasticity patterns, normal and lognormal distributions produce lower power for both types of tests.", dev = 'png', dpi = 450}
 
@@
 
-Now it is important to note that residual plots need to be delivered as a line up, where it is embedded in a field of null plots.
+Now it is important to note that residual plots need to be delivered as a lineup, where it is embedded in a field of null plots.
 This enables a careful calibration for reading structure in residual plots.
@@
 As expected, if the fitted value distribution is not uniform, there is a loss of power in the visual test.
-Structure is hardest to detect if fitted values ar discrete.
-Also, a surprising find was that the direction of heteroskedasticity appears to affect the ability to visually detect it, with wedge to the left being less detectable.
+Structure is hardest to detect if fitted values are lognormal.
+Also, a surprising find was that the direction of heteroskedasticity appears to affect the ability to visually detect it, with wedge to the right being less detectable.
 

### 2023-02-15 paper_comparison/paper_comparison/paper_comparison.tex
update the text according to the updated effect size

@@
 average of the \(E_{sample}\) calculated for samples drawn from the
-given model \(M\) repeatedly.
+given model \(M\) repeatedly. We use the normal distribution as a
+general choice to approximate the conditional effect size.
 
@@
 
-To get an accurate estimate of the power of conventional tests, we
-simulate 200000 samples evenly from the models used in the human subject
-experiments, and record the \(p\)-values produced by every sample.
-
 Figure \ref{fig:heterpower} is similar to Figure \ref{fig:polypower},
@@
 For non-linearity patterns, the power curve of RESET test climbs
-aggressively from 7\% to 63\% as \(log_e(E)\) increases from 0 to 2.5,
+aggressively from 7\% to 84\% as \(log_e(E)\) increases from 0 to 2.5,
 while power of other tests respond inactively to the change of effect,
@@
 defects that being considered. Meanwhile, no noticeable residual
-depatures can be spotted from the example residual plots.
+departures can be spotted from the example residual plots.
 
 In terms of heteroskedasticity patterns, the power of Breusch-Pagan test
-is also almost always greater than the power of visual test. For
-\(log_e(E)\) at around 2.5, where the power curve of the visual test
-remains at a low level, the Breusch-Pagan test still have a decent
-amount of chance of rejecting \(H_0\). Similarly, the visual feature is
-nearly unobservable from the example residual plots.
+is almost always greater than the power of visual test. For \(log_e(E)\)
+at around 2.5, where the power curve of the visual test remains at a low
+level, the Breusch-Pagan test has around 50\% chance of rejecting
+\(H_0\). Similarly, the visual feature is nearly unobservable from the
+example residual plots.
 
@@
 The data plot at panel 17 contains a "Butterfly" shape.
-It has effect size $ = 4.02$, somewhat surprising that it is not detected by the BP test.}\label{fig:heter-example}
+It has effect size $ = 3.76$, somewhat surprising that it is not detected by the BP test.}\label{fig:heter-example}
 \end{figure}
@@
 the non-linearity pattern. According to Figure
-\ref{fig:poly-power-uniform-j}, conventional tests have identical power
-in testing the ``U'', the ``S'' and the ``M'' shape. But it has
-significant lower power in testing the ``Triple-U'' shape. To understand
-why this is, one needs to return to the way the RESET test is applied.
-It requires a parameter indicating degree of fitted values to test for,
-and the recommendation is to generically use four
+\ref{fig:poly-power-uniform-j}, conventional tests have higher power in
+testing shapes constructed from lower orders of Hermite polynomials. But
+the power in testing the ``Triple-U'' shape drops significantly. To
+understand why this is, one needs to return to the way the RESET test is
+applied. It requires a parameter indicating degree of fitted values to
+test for, and the recommendation is to generically use four
 (\citet{ramsey_tests_1969}). However, the ``Triple-U'' shape constructed
@@
 been applied using a higher power no less than six, the power curve of
-``Triple-U'' shape will have little difference to other shapes. The
-recommendation of the polynomial power for the RESET should be revised,
-perhaps. This illustrates the sensitivity of conventional testing to the
-parameters, and it also points to a limitation that one needs to know
-the data generating process in order to set the parameters for the test.
+``Triple-U'' shape will only slight lower than the power curve of ``M''
+shape. The recommendation of the polynomial power for the RESET should
+be revised, perhaps. This illustrates the sensitivity of conventional
+testing to the parameters, and it also points to a limitation that one
+needs to know the data generating process in order to set the parameters
+for the test.
 
@@
 curves are mostly aligned with our expectation, except for the ``M''
-shape, which is as easy to be detected as the ``U'' shape. This implies
-the readability of the shape do not strictly follow the degree of the
-polynomials.
+shape, which is as easy to be detected as the ``S'' shape. This implies,
+unlike the conventional test, the readability of the shape do not
+strictly follow the degree of the polynomials.
 
@@
 heteroskedasticity}\label{effect-of-shape-of-heteroskedasticity}}
-
-(change the butterfly color)
 
@@
 little difference between the power curves can be perceived. Similarly,
-visual tests should have the smae power in detecting these two shapes if
+visual tests should have the same power in detecting these two shapes if
 they are equally likely to be identified. However, it can be observed
 from Figure \ref{fig:heter-power-uniform-a} that the power curve for the
-``Left-triangle'' shape is constantly higher than the one for the
+``Left-triangle'' shape is higher than the one for the
 ``Right-triangle'' shape, indicating a potential favour of orientation
 by human, which is worth to be explored in future studies. Besides, the
-chance of detecting the ``Butterfly'' shape by visual tests are slightly
-higher than other two shapes.
+chance of detecting the ``Butterfly'' shape by both conventional and
+visual tests are much higher than other two shapes.
 
@@
 
+For visual tests, the power curves for normal distribution and lognormal
+distribution are lower than other two distributions, which are all even.
+For non-linearity patterns, the power curve of normal distribution is
+slightly greater than the one for lognormal distribution. And the margin
+is even larger for heteroskedasticity patterns. This suggests skewness
+can affect the degree of presence of the visual pattern.
+
 For non-linearity patterns, we do not observe significant difference in
-power of visual test between distributions, except for the discrete
-uniform distribution. It could be due to the fact that humans have
-difficulties in recognising visual patterns from discreteness, which
-makes the shape disconnected and incomplete (ref?). In terms of
-conventional tests, the discrete uniform distribution is instead the one
-with the greatest power, followed by lognormal, uniform and the normal
-distribution.
+power of visual test between uniform distribution and discrete uniform
+distribution, but the normal distribution and the lognoraml distribution
+have lower power than the previous two. In terms of conventional tests,
+the discrete uniform distribution is the one with the greatest power,
+followed by lognormal, uniform and the normal distribution.
 
@@
 since other three distributions could reduce the chance of revealing the
-underlying visual pattern because of uneven data points. Although the
-power of visual test under the discrete uniform distribution is the
-lowest in the case of non-linearity patterns, it has a relatively great
-power this time. Considering for heteroskedasticity, the visual patterns
-are usually detected by connecting the maximum and minimum residuals
+underlying visual pattern because of uneven data points. The power of
+visual test under the discrete uniform distribution has a relatively
+great. Considering for heteroskedasticity, the visual patterns are
+usually detected by connecting the maximum and minimum residuals
 separately at different fitted values. It will not be greatly affected
 by the discrete uniform distribution compared to the uniform
-distribution. Visual tests have low power on lineups with the normal
-distribution and the lognormal distribution. This suggests the use of
-transformation of fitted values in reading residual plots. In the case
-of conventional tests, the discrete uniform distribution consistently
-perform well, followed by uniform, normal and the lognormal
-distribution.
+distribution. Again, visual tests have low power on lineups with the
+normal distribution and the lognormal distribution. It indicates
+skewness and unevenness can reduce the detection rate of the visual
+pattern. In the case of conventional tests, the discrete uniform
+distribution consistently perform well, followed by uniform, normal and
+the lognormal distribution.
 
@@
 Now it is important to note that residual plots need to be delivered as
-a line up, where it is embedded in a field of null plots. This enables a
+a lineup, where it is embedded in a field of null plots. This enables a
 careful calibration for reading structure in residual plots. XXX More
@@
 loss of power in the visual test. Structure is hardest to detect if
-fitted values ar discrete. Also, a surprising find was that the
+fitted values are lognormal. Also, a surprising find was that the
 direction of heteroskedasticity appears to affect the ability to
-visually detect it, with wedge to the left being less detectable.
+visually detect it, with wedge to the right being less detectable.
 

### 2023-02-15 paper_comparison/paper_comparison/paper_comparison.Rmd
fix the effect size formula

@@
 
-\small
+\footnotesize
 
@@
 E = E[E_{sample}|M],\\
-E_{sample} = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V})|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V})^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V})^{-1}\boldsymbol{\mu}_z\right),\\
+E_{sample} = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}\boldsymbol{\mu}_z\right),\\
 \boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a,\\

### 2023-02-27 paper_comparison/appendix.Rmd
rewrite the p-value part

@@
 
+(discuss what happens if we switch to different effect size)
+
 Effect size can be defined as the difference of a parameter for a particular model or distribution,  or a statistic derived from a sample.
@@
 The Kullback-Leibler divergence is defined as $\int_{-\infty}^{\infty}log(p(x)/q(x))p(x)dx$, where $p(.)$ and $q(.)$ denote probability densities of $P$ and $Q$.
+
+Let $\boldsymbol{X}_a = (\boldsymbol{1}, \boldsymbol{X})$ denotes the $p$ regressors with $n$ observations, $\boldsymbol{R}_a = \boldsymbol{I} -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ denotes the residual operator, and let $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0},\sigma^2\boldsymbol{I})$ denotes the error.
+Using the Frisch–Waugh–Lovell theorem, residuals $\boldsymbol{e} = \boldsymbol{R}_a\boldsymbol{\varepsilon}$.
+Because $rank(\boldsymbol{R}_a) = n - p < n$, $e$ follows a degenerate multivariate normal distribution and does not have a density.
+Since the Kullback-Leibler divergence requires a proper density function, we need to simplify the covariance matrix of $\boldsymbol{e}$ by setting all the off-diagonal elements to 0.
+Then, the residuals will assumed to follow $N(\boldsymbol{0}, diag(\boldsymbol{R}_a\sigma^2))$ under the null hypothesis that the model is correctly specified.
+If the model is however misspecified due to omitted variables $\boldsymbol{Z}$, or a non-constant variance $\boldsymbol{V}$, the distribution of residuals can be derived as $N(\boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z, diag(\boldsymbol{R}_a\sigma^2))$ and $N(\boldsymbol{0}, diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a'))$ respectively.
+
+By assuming both $P$ and $Q$ are multivariate normal density functions, the Kullback-Leibler divergence can be rewritten as 
+$$KL = \frac{1}{2}\left(log\frac{|\Sigma_p|}{|\Sigma_q|} - n + tr(\Sigma_p^{-1}\Sigma_q) + (\mu_p - \mu_q)'\Sigma_p^{-1}(\mu_p - \mu_q)\right).$$
+
+Then, we can combine the two residual departures into one formula
+
+$$KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}\boldsymbol{\mu}_z\right).$$
+
+When there are omitted variables but constant error variance, the formula can be reduced to
+
+$$KL = \frac{1}{2}\left(\boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\sigma^2)^{-1}\boldsymbol{\mu}_z\right).$$
+
+And when the model equation is correctly specified but the error variance is non-constant, the formula can be reduced to
+
+$$KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2))\right).$$
 

### 2023-02-27 paper_comparison/paper_comparison/paper_comparison.Rmd
rewrite the p-value part

@@
       # Every sample contains 2000 lineups
-      map(1:2000, function(i) {
+      map(1:200, function(i) {
 
@@
       # Every sample contains 2000 lineups
-      map(1:2000, function(x) {
+      map(1:200, function(x) {
 
@@
 
-(change it to facet)
-
-(remove backticks for some plots)
-
 ```{r residual-plot-common-departures, fig.width = 10, fig.height = 2.5, fig.cap = "Example residual vs fitted value plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0.
@@
 Then, by imposing a relatively strong assumption on the visual test that all $K$ evaluations are fully independent, under $H_0$, $X \sim \mathrm{Binom}_{K,1/m}$.
-Therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as $P(X \geq x) = 1 - F(x) + f(x)$, where $F(.)$ is the cumulative distribution function, $f(.)$ is the probability mass function and $x$ is the realization of number of observers correctly picking the data plot [@majumder_validation_2013].
-
-As pointed out by @vanderplas2021statistical, this basic binomial model doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup.
-And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup.
-@vanderplas2021statistical summarises three common scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects.
-Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experiment design.
-For Scenario 3, @vanderplas2021statistical models the probability of a plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$.
-The number of times plot $i$ being selected in $K$ evaluations is denoted as $c_i$.
-In case subject $j$ makes multiple selections, $1/s_j$ will be added to $c_i$ instead of one, where $s_j$ is the number of plots subject $j$ selected for $j=1,...K$.
-This ensures $\sum_{i}c_i=K$.
-Since we are only interested in the selections of the data plot $i$, the marginal model can be simplified to a beta-binomial model and thus the visual p-value is given as
-
-<!-- TODO:Clarify this is only for integer c_i -->
-
-```{=tex}
-\begin{equation} \label{eq:pvalue-beta-binomial}
-P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+
-\end{equation}
-```
-
-\noindent where $B(.)$ is the beta function defined as
-
-```{=tex}
-\begin{equation} \label{eq:betafunction}
-B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0. 
-\end{equation}
-```
-
-Note that Equation \ref{eq:pvalue-beta-binomial} given in @vanderplas2021statistical only works with non-negative integer $c_i$.
-We extend the equation to non-negative real number $c_i$ by applying a linear approximation
-
-```{=tex}
-\begin{equation} \label{eq:pvalue-beta-binomial-approx}
-P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \in \mathbb{R}_0^+,
-\end{equation}
-```
-
-where $P(C \geq \lceil c_i \rceil)$ is calculated using Equation \ref{eq:pvalue-beta-binomial} and $P(C = \lfloor c_i \rfloor)$ is calculated by
-
-\begin{equation} \label{eq:pmf-beta-binomial}
-P(C = c_i) = {K \choose c_i}\frac{B(c_i + \alpha, K - c_i + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+.
-\end{equation}
-
-Besides, the parameter $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} and \ref{eq:pmf-beta-binomial} is usually unknown and hence needs to be estimated from the survey data.
-For low values of $\alpha$, only a few plots are attractive to the observers and tend to be selected.
-For higher values of $\alpha$, the distribution of the probability of each plot being selected is more even.
-@vanderplas2021statistical defines that a plot is $c$-interesting if $c$ or more participants select the plot as the most different.
-Given the definition, The expected number of plots selected at least $c$ times, $E[Z_c]$, is calculated as
-
-```{=tex}
-\begin{equation} \label{eq:c-interesting-expectation}
-E[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation}
-```
-
-With Equation \ref{eq:c-interesting-expectation}, $\alpha$ can be estimated using maximum likelihood estimation.
-But for precise estimate of $\alpha$, additional responses to Rorschach lineups, which is a type of lineup that consists of plots constructed from the same null data generating mechanism, are required.
-
+Therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as $P(X \geq x) = 1 - F(x) + f(x)$, where $F(.)$ is the binomial cumulative distribution function, $f(.)$ is the binomial probability mass function and $x$ is the realization of number of observers correctly picking the data plot [@majumder_validation_2013].
 ### Power of a visual test
@@
 
+## Methodology
+
+### Caluculation of effect size
+
+### Caluculation of p-value
+
+<!-- As pointed out by @vanderplas2021statistical, this basic binomial model doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup.
+And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup.
+@vanderplas2021statistical summarises three common scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects.
+Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experiment design.
+For Scenario 3, @vanderplas2021statistical models the probability of a plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$.
+The number of times plot $i$ being selected in $K$ evaluations is denoted as $c_i$.
+In case subject $j$ makes multiple selections, $1/s_j$ will be added to $c_i$ instead of one, where $s_j$ is the number of plots subject $j$ selected for $j=1,...K$.
+This ensures $\sum_{i}c_i=K$.
+Since we are only interested in the selections of the data plot $i$, the marginal model can be simplified to a beta-binomial model and thus the visual p-value is given as -->
+
+<!-- <!-- TODO:Clarify this is only for integer c_i --> -->
+
+<!-- ```{=tex} -->
+<!-- \begin{equation} \label{eq:pvalue-beta-binomial} -->
+<!-- P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+ -->
+<!-- \end{equation} -->
+<!-- ``` -->
+
+<!-- \noindent where $B(.)$ is the beta function defined as -->
+
+<!-- ```{=tex} -->
+<!-- \begin{equation} \label{eq:betafunction} -->
+<!-- B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0.  -->
+<!-- \end{equation} -->
+<!-- ``` -->
+
+<!-- Note that Equation \ref{eq:pvalue-beta-binomial} given in @vanderplas2021statistical only works with non-negative integer $c_i$.
+We extend the equation to non-negative real number $c_i$ by applying a linear approximation -->
+
+<!-- ```{=tex} -->
+<!-- \begin{equation} \label{eq:pvalue-beta-binomial-approx} -->
+<!-- P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \in \mathbb{R}_0^+, -->
+<!-- \end{equation} -->
+<!-- ``` -->
+
+<!-- where $P(C \geq \lceil c_i \rceil)$ is calculated using Equation \ref{eq:pvalue-beta-binomial} and $P(C = \lfloor c_i \rfloor)$ is calculated by -->
+
+<!-- \begin{equation} \label{eq:pmf-beta-binomial} -->
+<!-- P(C = c_i) = {K \choose c_i}\frac{B(c_i + \alpha, K - c_i + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+. -->
+<!-- \end{equation} -->
+
+<!-- Besides, the parameter $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} and \ref{eq:pmf-beta-binomial} is usually unknown and hence needs to be estimated from the survey data.
+For low values of $\alpha$, only a few plots are attractive to the observers and tend to be selected.
+For higher values of $\alpha$, the distribution of the probability of each plot being selected is more even.
+@vanderplas2021statistical defines that a plot is $c$-interesting if $c$ or more participants select the plot as the most different.
+Given the definition, The expected number of plots selected at least $c$ times, $E[Z_c]$, is calculated as -->
+
+<!-- ```{=tex} -->
+<!-- \begin{equation} \label{eq:c-interesting-expectation} -->
+<!-- E[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation} -->
+<!-- ``` -->
+
+<!-- With Equation \ref{eq:c-interesting-expectation}, $\alpha$ can be estimated using maximum likelihood estimation.
+But for precise estimate of $\alpha$, additional responses to Rorschach lineups, which is a type of lineup that consists of plots constructed from the same null data generating mechanism, are required.
+-->
+### Caluculation of power
+
 ## Experimental setup
@@
 # Define the minimum and maximum effect size
+
+vi_survey <- vi_survey %>%
+  mutate(effect_size = sqrt(n) / (e_sigma * c(1,2,3,5)[shape]))
+
+poly_conv_sim <- poly_conv_sim %>%
+  mutate(effect_size = sqrt(n) / (e_sigma * c(1,2,3,5)[shape]))
+
 min_poly_es <- vi_survey %>% 
@@
   scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
-  scale_size_manual(values = c(1, 2, 3)) +
+  # scale_size_manual(values = c(1, 2, 3)) +
   scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +

### 2023-02-27 paper_comparison/paper_comparison/paper_comparison.Rmd
rearrange power section

@@
 
-### Calculating $p$-values for the visual test
+### $p$-values of a visual test
 
@@
 Therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as $P(X \geq x) = 1 - F(x) + f(x)$, where $F(.)$ is the binomial cumulative distribution function, $f(.)$ is the binomial probability mass function and $x$ is the realization of number of observers correctly picking the data plot [@majumder_validation_2013].
-### Power of a visual test
-
-The power of a model misspecification test is the probability that the null hypothesis is rejected given the regression model is misspecified.
-It is an important indicator when one is concerned about whether model assumptions have been violated.
-Although in practice, one might be more interested in knowing how much the residuals deviate from the model assumptions, and whether this deviation is of practical significance.
-
-As discussed in @majumder_validation_2013, the power of a visual test may depend on the ability of the particular subject, as the skill of the individual may affect the number of observers who identify the data plot from the lineup.
-Previously, it is addressed by modelling the probability of a subject $j$ correctly picking the data plot from a lineup $l$ using a mixed-effect logistic regression, with subjects treated as random effects [@majumder_validation_2013].
-However, in the multiple selections scenario, having this probability is not sufficient to determine the power of a visual test because it does not provide information about the number of selections made by the subject for the calculation of the p-value.
-
-Instead, we directly estimate the probability of a lineup being rejected by assuming that individual skill has negligible effect on the variation of the power.
-The assumption is to simplify the model structure, thereby obviate costly large-scale experiments to estimate complex covariance matrices.
-The model is a logistic regression formulated as
-
-```{=tex}
-\begin{equation} \label{eq:logistic-regression-1-1}
-Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right),
-\end{equation}
-```
-\noindent where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
-The effect size $E$ is the only regressor and the intercept is fixed to $log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$.
-
-Effect size $E$ is formulated as the conditional expectation of the sample-based effect size $E_{sample}$ derived from the Kullback-Leibler divergence (see Appendix for the details) given by
-
-\footnotesize
-
-\begin{gather}  \label{eq:effect-size}
-E = E[E_{sample}|M],\\
-E_{sample} = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}\boldsymbol{\mu}_z\right),\\
-\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a,\\
-\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a^T\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a^T,
-\end{gather}
-
-\normalsize
-
-\noindent where $M$ is a given model that generates the data, $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,  $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of regressors including the intercept used in the regression equation, $\sigma^2\boldsymbol{I}$ is the assumed variance of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual variance of the error term, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of the residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients.
-
-A conditional expectation is taken because $E$ is designed to be invariant to different realizations of $\boldsymbol{X}$.
-However, solving this conditional expectation analytically is very difficult as $\boldsymbol{X}$ can follow arbitrary distributions.
-Therefore, we make an assumption that $\boldsymbol{X}$ follows a particular multivariate normal distribution
-
-We make an assumption that all the realizations of $\boldsymbol{X}$ are generated from a multivariate normal distribution $N(\boldsymbol{0}, 0.3^2\boldsymbol{I})$ and then rescaled by the min-max normalization.
-The estimate of $E$ is obtained by taking the average of the $E_{sample}$ calculated for samples drawn from the given model $M$ repeatedly.
-We use the normal distribution as a general choice to approximate the conditional effect size.
-
-To study various factors contributing to the power of the visual test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors.
-These include the distribution of the regressor, type of the simulation model and the shape of the residual departures.
 
@@
 -->
+
 ### Caluculation of power
+
+### Power of a visual test
+
+<!-- The power of a model misspecification test is the probability that the null hypothesis is rejected given the regression model is misspecified.
+It is an important indicator when one is concerned about whether model assumptions have been violated.
+Although in practice, one might be more interested in knowing how much the residuals deviate from the model assumptions, and whether this deviation is of practical significance.
+-->
+
+<!-- As discussed in @majumder_validation_2013, the power of a visual test may depend on the ability of the particular subject, as the skill of the individual may affect the number of observers who identify the data plot from the lineup.
+Previously, it is addressed by modelling the probability of a subject $j$ correctly picking the data plot from a lineup $l$ using a mixed-effect logistic regression, with subjects treated as random effects [@majumder_validation_2013].
+However, in the multiple selections scenario, having this probability is not sufficient to determine the power of a visual test because it does not provide information about the number of selections made by the subject for the calculation of the p-value.
+-->
+
+<!-- Instead, we directly estimate the probability of a lineup being rejected by assuming that individual skill has negligible effect on the variation of the power.
+The assumption is to simplify the model structure, thereby obviate costly large-scale experiments to estimate complex covariance matrices.
+The model is a logistic regression formulated as  -->
+
+<!-- ```{=tex} -->
+<!-- \begin{equation} \label{eq:logistic-regression-1-1} -->
+<!-- Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right), -->
+<!-- \end{equation} -->
+<!-- ``` -->
+<!-- \noindent where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
+The effect size $E$ is the only regressor and the intercept is fixed to $log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$.
+-->
+
+<!-- Effect size $E$ is formulated as the conditional expectation of the sample-based effect size $E_{sample}$ derived from the Kullback-Leibler divergence (see Appendix for the details) given by -->
+
+<!-- \footnotesize -->
+
+<!-- \begin{gather}  \label{eq:effect-size} -->
+<!-- E = E[E_{sample}|M],\\ -->
+<!-- E_{sample} = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}\boldsymbol{\mu}_z\right),\\ -->
+<!-- \boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a,\\ -->
+<!-- \boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a^T\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a^T, -->
+<!-- \end{gather} -->
+
+<!-- \normalsize -->
+
+<!-- \noindent where $M$ is a given model that generates the data, $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,  $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of regressors including the intercept used in the regression equation, $\sigma^2\boldsymbol{I}$ is the assumed variance of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual variance of the error term, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of the residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients.
+-->
+
+<!-- A conditional expectation is taken because $E$ is designed to be invariant to different realizations of $\boldsymbol{X}$.
+However, solving this conditional expectation analytically is very difficult as $\boldsymbol{X}$ can follow arbitrary distributions.
+Therefore, we make an assumption that $\boldsymbol{X}$ follows a particular multivariate normal distribution -->
+
+<!-- We make an assumption that all the realizations of $\boldsymbol{X}$ are generated from a multivariate normal distribution $N(\boldsymbol{0}, 0.3^2\boldsymbol{I})$ and then rescaled by the min-max normalization.
+The estimate of $E$ is obtained by taking the average of the $E_{sample}$ calculated for samples drawn from the given model $M$ repeatedly.
+We use the normal distribution as a general choice to approximate the conditional effect size.
+-->
+
+<!-- To study various factors contributing to the power of the visual test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors.
+These include the distribution of the regressor, type of the simulation model and the shape of the residual departures.
+-->
 

### 2023-02-28 paper_comparison/paper_comparison/paper_comparison.Rmd
adjust the introduction based on more information on the writings

@@
 Abundance of literature describe appropriate diagnostic methods for linear regression: @draper1998applied, @montgomery1982introduction, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals.
-The diligent reader of these sage writings will also notice sentences that express sentiments like *based on their experience, statistical tests are not widely used in regression diagnostics since the same or even larger amount of information can be provided by diagnostic plots than the corresponding tests in most empirical studies.* A common guidance by experts is that optimal method for diagnosing model fits is by plotting the data.
-
+All of these writings consider plotting of residuals to be a standard technique that should be examined routinely in all regression modelling problems.
+In addition, @draper1998applied and @belsley_regression_1980 believe that residual plots are usually very revealing when the assumptions are violated.
+@cook_applied_1999 thinks formal tests and graphical procedures are complementary and both have a place in residual analysis, but they focus on graphical methods rather than on formal testing, as they are easier to use.
+@montgomery1982introduction even suggests that residual plots are more informative in most practical situations than corresponding formal tests, and statistical tests on regression model residuals are not widely used based on their experience.
+
+A common guidance by experts is that optimal method for diagnosing model fits is by plotting the data.
 The persistence of this advice to check the plots is curious, and investigating why this might be common advice is the subject of this paper.
@@
 
-### Lineup protocol
-
 One may argue that reading diagnostic plots is to some extent subjective and indecisive compared to those rigorous statistical procedures as it relies on graphical perception - human ability to interpret and decode the information embedded in graph [@cleveland_graphical_1984].
@@
 In fact, the actual residual plot is obtained from a misspecified regression model with non-linearity defect.
+
+Data used in the $m - 1$ null plots needs to be simulated.
+In regression diagnostics, sampling data consistent with $H_0$ is equivalent to sampling data from the assumed model.
+As @buja_statistical_2009 suggested, $H_0$ is usually a composite hypothesis controlled by nuisance parameters.
+Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to a so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling.
+The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009].
+Essentially, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the regressors, then rescaling it by the ratio of residual sum of square in two regressions.
 
@@
 
-### Sampling from the null distribution
-
-Data used in the $m - 1$ null plots needs to be simulated.
-In regression diagnostics, sampling data consistent with $H_0$ is equivalent to sampling data from the assumed model.
-As @buja_statistical_2009 suggested, $H_0$ is usually a composite hypothesis controlled by nuisance parameters.
-Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to a so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling.
-The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009].
-Essentially, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the regressors, then rescaling it by the ratio of residual sum of square in two regressions.
-
-### $p$-values of a visual test
+# Caluculation of p-value, effect size and power
+
+Visual inference is
+
+To be able to properly compare conventional tests to visual tests, we need to define the effect size of the residual departure, the power of the tests under a certain level of effect size,
+
+### Caluculation of effect size
+
+### Caluculation of p-value
 
@@
 
+<!-- As pointed out by @vanderplas2021statistical, this basic binomial model doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup.
+And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup.
+@vanderplas2021statistical summarises three common scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects.
+Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experiment design.
+For Scenario 3, @vanderplas2021statistical models the probability of a plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$.
+The number of times plot $i$ being selected in $K$ evaluations is denoted as $c_i$.
+In case subject $j$ makes multiple selections, $1/s_j$ will be added to $c_i$ instead of one, where $s_j$ is the number of plots subject $j$ selected for $j=1,...K$.
+This ensures $\sum_{i}c_i=K$.
+Since we are only interested in the selections of the data plot $i$, the marginal model can be simplified to a beta-binomial model and thus the visual p-value is given as -->
+
+<!-- <!-- TODO:Clarify this is only for integer c_i --> -->
+
+<!-- ```{=tex} -->
+<!-- \begin{equation} \label{eq:pvalue-beta-binomial} -->
+<!-- P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+ -->
+<!-- \end{equation} -->
+<!-- ``` -->
+
+<!-- \noindent where $B(.)$ is the beta function defined as -->
+
+<!-- ```{=tex} -->
+<!-- \begin{equation} \label{eq:betafunction} -->
+<!-- B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0.  -->
+<!-- \end{equation} -->
+<!-- ``` -->
+
+<!-- Note that Equation \ref{eq:pvalue-beta-binomial} given in @vanderplas2021statistical only works with non-negative integer $c_i$.
+We extend the equation to non-negative real number $c_i$ by applying a linear approximation -->
+
+<!-- ```{=tex} -->
+<!-- \begin{equation} \label{eq:pvalue-beta-binomial-approx} -->
+<!-- P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \in \mathbb{R}_0^+, -->
+<!-- \end{equation} -->
+<!-- ``` -->
+
+<!-- where $P(C \geq \lceil c_i \rceil)$ is calculated using Equation \ref{eq:pvalue-beta-binomial} and $P(C = \lfloor c_i \rfloor)$ is calculated by -->
+
+<!-- \begin{equation} \label{eq:pmf-beta-binomial} -->
+<!-- P(C = c_i) = {K \choose c_i}\frac{B(c_i + \alpha, K - c_i + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+. -->
+<!-- \end{equation} -->
+
+<!-- Besides, the parameter $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} and \ref{eq:pmf-beta-binomial} is usually unknown and hence needs to be estimated from the survey data.
+For low values of $\alpha$, only a few plots are attractive to the observers and tend to be selected.
+For higher values of $\alpha$, the distribution of the probability of each plot being selected is more even.
+@vanderplas2021statistical defines that a plot is $c$-interesting if $c$ or more participants select the plot as the most different.
+Given the definition, The expected number of plots selected at least $c$ times, $E[Z_c]$, is calculated as -->
+
+<!-- ```{=tex} -->
+<!-- \begin{equation} \label{eq:c-interesting-expectation} -->
+<!-- E[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation} -->
+<!-- ``` -->
+
+<!-- With Equation \ref{eq:c-interesting-expectation}, $\alpha$ can be estimated using maximum likelihood estimation.
+But for precise estimate of $\alpha$, additional responses to Rorschach lineups, which is a type of lineup that consists of plots constructed from the same null data generating mechanism, are required.
+-->
+
+### Caluculation of power
+
+<!-- The power of a model misspecification test is the probability that the null hypothesis is rejected given the regression model is misspecified.
+It is an important indicator when one is concerned about whether model assumptions have been violated.
+Although in practice, one might be more interested in knowing how much the residuals deviate from the model assumptions, and whether this deviation is of practical significance.
+-->
+
+<!-- As discussed in @majumder_validation_2013, the power of a visual test may depend on the ability of the particular subject, as the skill of the individual may affect the number of observers who identify the data plot from the lineup.
+Previously, it is addressed by modelling the probability of a subject $j$ correctly picking the data plot from a lineup $l$ using a mixed-effect logistic regression, with subjects treated as random effects [@majumder_validation_2013].
+However, in the multiple selections scenario, having this probability is not sufficient to determine the power of a visual test because it does not provide information about the number of selections made by the subject for the calculation of the p-value.
+-->
+
+<!-- Instead, we directly estimate the probability of a lineup being rejected by assuming that individual skill has negligible effect on the variation of the power.
+The assumption is to simplify the model structure, thereby obviate costly large-scale experiments to estimate complex covariance matrices.
+The model is a logistic regression formulated as  -->
+
+<!-- ```{=tex} -->
+<!-- \begin{equation} \label{eq:logistic-regression-1-1} -->
+<!-- Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right), -->
+<!-- \end{equation} -->
+<!-- ``` -->
+<!-- \noindent where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
+The effect size $E$ is the only regressor and the intercept is fixed to $log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$.
+-->
+
+<!-- Effect size $E$ is formulated as the conditional expectation of the sample-based effect size $E_{sample}$ derived from the Kullback-Leibler divergence (see Appendix for the details) given by -->
+
+<!-- \footnotesize -->
+
+<!-- \begin{gather}  \label{eq:effect-size} -->
+<!-- E = E[E_{sample}|M],\\ -->
+<!-- E_{sample} = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}\boldsymbol{\mu}_z\right),\\ -->
+<!-- \boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a,\\ -->
+<!-- \boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a^T\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a^T, -->
+<!-- \end{gather} -->
+
+<!-- \normalsize -->
+
+<!-- \noindent where $M$ is a given model that generates the data, $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,  $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of regressors including the intercept used in the regression equation, $\sigma^2\boldsymbol{I}$ is the assumed variance of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual variance of the error term, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of the residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients.
+-->
+
+<!-- A conditional expectation is taken because $E$ is designed to be invariant to different realizations of $\boldsymbol{X}$.
+However, solving this conditional expectation analytically is very difficult as $\boldsymbol{X}$ can follow arbitrary distributions.
+Therefore, we make an assumption that $\boldsymbol{X}$ follows a particular multivariate normal distribution -->
+
+<!-- We make an assumption that all the realizations of $\boldsymbol{X}$ are generated from a multivariate normal distribution $N(\boldsymbol{0}, 0.3^2\boldsymbol{I})$ and then rescaled by the min-max normalization.
+The estimate of $E$ is obtained by taking the average of the $E_{sample}$ calculated for samples drawn from the given model $M$ repeatedly.
+We use the normal distribution as a general choice to approximate the conditional effect size.
+-->
+
+<!-- To study various factors contributing to the power of the visual test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors.
+These include the distribution of the regressor, type of the simulation model and the shape of the residual departures.
+-->
+
 # Experimental design
@@
 To investigate the difference between conventional hypothesis testing and visual inference in the application of linear regression diagnostics, an experiment is conducted in three data collection periods.
-The two types of departures, non-linearity and heteroskedasticity, are primarily collected during data collection periods I and II.
+Two types of departures, non-linearity and heteroskedasticity, are primarily collected during data collection periods I and II.
 The data collection period III was designed primarily to measure human responses to null lineups so that the parameter $\alpha$ in Equation \ref{eq:pvalue-beta-binomial} can be estimated.
@@
 
-## Methodology
-
-### Caluculation of effect size
-
-### Caluculation of p-value
-
-<!-- As pointed out by @vanderplas2021statistical, this basic binomial model doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup.
-And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup.
-@vanderplas2021statistical summarises three common scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects.
-Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experiment design.
-For Scenario 3, @vanderplas2021statistical models the probability of a plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$.
-The number of times plot $i$ being selected in $K$ evaluations is denoted as $c_i$.
-In case subject $j$ makes multiple selections, $1/s_j$ will be added to $c_i$ instead of one, where $s_j$ is the number of plots subject $j$ selected for $j=1,...K$.
-This ensures $\sum_{i}c_i=K$.
-Since we are only interested in the selections of the data plot $i$, the marginal model can be simplified to a beta-binomial model and thus the visual p-value is given as -->
-
-<!-- <!-- TODO:Clarify this is only for integer c_i --> -->
-
-<!-- ```{=tex} -->
-<!-- \begin{equation} \label{eq:pvalue-beta-binomial} -->
-<!-- P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+ -->
-<!-- \end{equation} -->
-<!-- ``` -->
-
-<!-- \noindent where $B(.)$ is the beta function defined as -->
-
-<!-- ```{=tex} -->
-<!-- \begin{equation} \label{eq:betafunction} -->
-<!-- B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0.  -->
-<!-- \end{equation} -->
-<!-- ``` -->
-
-<!-- Note that Equation \ref{eq:pvalue-beta-binomial} given in @vanderplas2021statistical only works with non-negative integer $c_i$.
-We extend the equation to non-negative real number $c_i$ by applying a linear approximation -->
-
-<!-- ```{=tex} -->
-<!-- \begin{equation} \label{eq:pvalue-beta-binomial-approx} -->
-<!-- P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \in \mathbb{R}_0^+, -->
-<!-- \end{equation} -->
-<!-- ``` -->
-
-<!-- where $P(C \geq \lceil c_i \rceil)$ is calculated using Equation \ref{eq:pvalue-beta-binomial} and $P(C = \lfloor c_i \rfloor)$ is calculated by -->
-
-<!-- \begin{equation} \label{eq:pmf-beta-binomial} -->
-<!-- P(C = c_i) = {K \choose c_i}\frac{B(c_i + \alpha, K - c_i + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+. -->
-<!-- \end{equation} -->
-
-<!-- Besides, the parameter $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} and \ref{eq:pmf-beta-binomial} is usually unknown and hence needs to be estimated from the survey data.
-For low values of $\alpha$, only a few plots are attractive to the observers and tend to be selected.
-For higher values of $\alpha$, the distribution of the probability of each plot being selected is more even.
-@vanderplas2021statistical defines that a plot is $c$-interesting if $c$ or more participants select the plot as the most different.
-Given the definition, The expected number of plots selected at least $c$ times, $E[Z_c]$, is calculated as -->
-
-<!-- ```{=tex} -->
-<!-- \begin{equation} \label{eq:c-interesting-expectation} -->
-<!-- E[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation} -->
-<!-- ``` -->
-
-<!-- With Equation \ref{eq:c-interesting-expectation}, $\alpha$ can be estimated using maximum likelihood estimation.
-But for precise estimate of $\alpha$, additional responses to Rorschach lineups, which is a type of lineup that consists of plots constructed from the same null data generating mechanism, are required.
--->
-
-### Caluculation of power
-
-### Power of a visual test
-
-<!-- The power of a model misspecification test is the probability that the null hypothesis is rejected given the regression model is misspecified.
-It is an important indicator when one is concerned about whether model assumptions have been violated.
-Although in practice, one might be more interested in knowing how much the residuals deviate from the model assumptions, and whether this deviation is of practical significance.
--->
-
-<!-- As discussed in @majumder_validation_2013, the power of a visual test may depend on the ability of the particular subject, as the skill of the individual may affect the number of observers who identify the data plot from the lineup.
-Previously, it is addressed by modelling the probability of a subject $j$ correctly picking the data plot from a lineup $l$ using a mixed-effect logistic regression, with subjects treated as random effects [@majumder_validation_2013].
-However, in the multiple selections scenario, having this probability is not sufficient to determine the power of a visual test because it does not provide information about the number of selections made by the subject for the calculation of the p-value.
--->
-
-<!-- Instead, we directly estimate the probability of a lineup being rejected by assuming that individual skill has negligible effect on the variation of the power.
-The assumption is to simplify the model structure, thereby obviate costly large-scale experiments to estimate complex covariance matrices.
-The model is a logistic regression formulated as  -->
-
-<!-- ```{=tex} -->
-<!-- \begin{equation} \label{eq:logistic-regression-1-1} -->
-<!-- Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right), -->
-<!-- \end{equation} -->
-<!-- ``` -->
-<!-- \noindent where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
-The effect size $E$ is the only regressor and the intercept is fixed to $log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$.
--->
-
-<!-- Effect size $E$ is formulated as the conditional expectation of the sample-based effect size $E_{sample}$ derived from the Kullback-Leibler divergence (see Appendix for the details) given by -->
-
-<!-- \footnotesize -->
-
-<!-- \begin{gather}  \label{eq:effect-size} -->
-<!-- E = E[E_{sample}|M],\\ -->
-<!-- E_{sample} = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}\boldsymbol{\mu}_z\right),\\ -->
-<!-- \boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a,\\ -->
-<!-- \boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a^T\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a^T, -->
-<!-- \end{gather} -->
-
-<!-- \normalsize -->
-
-<!-- \noindent where $M$ is a given model that generates the data, $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,  $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of regressors including the intercept used in the regression equation, $\sigma^2\boldsymbol{I}$ is the assumed variance of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual variance of the error term, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of the residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients.
--->
-
-<!-- A conditional expectation is taken because $E$ is designed to be invariant to different realizations of $\boldsymbol{X}$.
-However, solving this conditional expectation analytically is very difficult as $\boldsymbol{X}$ can follow arbitrary distributions.
-Therefore, we make an assumption that $\boldsymbol{X}$ follows a particular multivariate normal distribution -->
-
-<!-- We make an assumption that all the realizations of $\boldsymbol{X}$ are generated from a multivariate normal distribution $N(\boldsymbol{0}, 0.3^2\boldsymbol{I})$ and then rescaled by the min-max normalization.
-The estimate of $E$ is obtained by taking the average of the $E_{sample}$ calculated for samples drawn from the given model $M$ repeatedly.
-We use the normal distribution as a general choice to approximate the conditional effect size.
--->
-
-<!-- To study various factors contributing to the power of the visual test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors.
-These include the distribution of the regressor, type of the simulation model and the shape of the residual departures.
--->
-
 ## Experimental setup

### 2023-02-28 paper_comparison/paper_comparison/paper_comparison.Rmd
recompile based on new data

@@
 # Define the minimum and maximum effect size
-
-vi_survey <- vi_survey %>%
-  mutate(effect_size = sqrt(n) / (e_sigma * c(1,2,3,5)[shape]))
-
-poly_conv_sim <- poly_conv_sim %>%
-  mutate(effect_size = sqrt(n) / (e_sigma * c(1,2,3,5)[shape]))
-
 min_poly_es <- vi_survey %>% 
@@
                group_by(log_effect_size, reject) %>% 
-               summarise(num_response = factor(n())),
+               summarise(num_response = n()),
              aes(log_effect_size, reject, size = num_response), 
@@
   scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
-  # scale_size_manual(values = c(1, 2, 3)) +
-  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
+  scale_x_continuous(breaks = seq(0, 3, 0.5), minor_breaks = NULL) +
   xlab(expression(log[e] (Effect_size))) +
@@
                               sigma = .x)$
-                     average_effect_size(n = 100)
+                     average_effect_size(n = 100, type = "simple")
                    )
@@
   mutate(power = map(mod, function(mod) {
-    result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
+    result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 1),
                          offset0 = log(0.05/0.95))
@@
 # The prediction of the visual test
-pred_visual_heter <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
+pred_visual_heter <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 1),
                                 offset0 = log(0.05/0.95)) %>%
@@
     mutate(power = map(mod, function(mod) {
-      result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
+      result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 1),
                            offset0 = log(0.05/0.95))
@@
   scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
-  scale_size_manual(values = c(1, 2, 3)) +
-  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
+  scale_x_continuous(breaks = seq(1, 7, 1), minor_breaks = NULL) +
   xlab(expression(log[e] (Effect_size))) +
@@
                                x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))})$
-                     average_effect_size(n = 100, type = "kl")
+                     average_effect_size(n = 100, type = "simple")
                    )
@@
 
-dat_x_dist <- list(
-  poly_conv_sim %>% 
-    rename(RESET_p_value = RESET4_p_value) %>%
-    mutate(x_dist = ifelse(x_dist == "even_discrete",
-                           "Discrete", 
-                           stringr::str_to_title(x_dist))) %>%
-    mutate(reject = as.numeric(RESET_p_value <= 0.05)) %>%
-    mutate(conv_or_not = "Conventional") %>%
-    mutate(type = "non-linearity") %>%
-    mutate(offset0 = log(0.05/0.95)) %>%
-    select(type, x_dist, effect_size, offset0, reject, conv_or_not),
-
-  heter_conv_sim %>% 
-    mutate(x_dist = ifelse(x_dist == "even_discrete",
-                           "Discrete", 
-                           stringr::str_to_title(x_dist))) %>%
-    mutate(reject = as.numeric(BP_p_value <= 0.05)) %>%
-    mutate(conv_or_not = "Conventional") %>%
-    mutate(type = "heteroskedasticity") %>%
-    mutate(offset0 = log(0.05/0.95)) %>%
-    select(type, x_dist, effect_size, offset0, reject, conv_or_not),
-
-  vi_survey %>%
+dat_x_dist <- vi_survey %>%
     filter(!null_lineup) %>%
@@
                            stringr::str_to_title(x_dist))) %>%
-    mutate(conv_or_not = "Visual") %>%
     mutate(offset0 = log(0.05/0.95)) %>%
-    select(type, x_dist, effect_size, offset0, reject, conv_or_not)
-) %>%
-
-  # Bind rows
-  map_df(~.x) %>%
+    select(type, x_dist, effect_size, offset0, reject) %>%
 
@@
   scale_color_manual(values = rcartocolor::carto_pal(4, "Safe")[c(2,1,3,4)]) +
-  scale_size_manual(values = c(1, 2, 3, 4)) +
   xlab(expression(log[e] (Effect_size))) +
@@
   labs(col = "", linetype = "Test") +
-  facet_grid(type ~ conv_or_not) +
+  facet_grid(~type) +
   theme(legend.position = "bottom")

### 2023-02-28 paper_comparison/paper_comparison/paper_comparison.tex
recompile based on new data

--- 2023-02-15 paper_comparison/paper_comparison/paper_comparison.tex
@@
 \citet{montgomery1982introduction}, \citet{belsley_regression_1980},
-\citet{cook_applied_1999} and \citet{cook1982residuals}. The diligent
-reader of these sage writings will also notice sentences that express
-sentiments like \emph{based on their experience, statistical tests are
-not widely used in regression diagnostics since the same or even larger
-amount of information can be provided by diagnostic plots than the
-corresponding tests in most empirical studies.} A common guidance by
-experts is that optimal method for diagnosing model fits is by plotting
-the data.
-
-The persistence of this advice to check the plots is curious, and
-investigating why this might be common advice is the subject of this
-paper. The paper is structured as follows. The next background section
-describes the types of departures that one expects to detect, and
-outlines a formal statistical process for reading residual plots, called
-visual inference. Section \ref{experimental-design} details the
-experimental design to compare the decision made by formal hypothesis
-testing, and how humans would read diagnostic plots. The results are
-reported in Section \ref{results}. We finish with a discussion on future
-work, in particular how the responsibility for residual plot reading
-might be relieved.
+\citet{cook_applied_1999} and \citet{cook1982residuals}. All of these
+writings consider plotting of residuals to be a standard technique that
+should be examined routinely in all regression modelling problems. In
+addition, \citet{draper1998applied} and \citet{belsley_regression_1980}
+believe that residual plots are usually very revealing when the
+assumptions are violated. \citet{cook_applied_1999} thinks formal tests
+and graphical procedures are complementary and both have a place in
+residual analysis, but they focus on graphical methods rather than on
+formal testing, as they are easier to use.
+\citet{montgomery1982introduction} even suggests that residual plots are
+more informative in most practical situations than corresponding formal
+tests, and statistical tests on regression model residuals are not
+widely used based on their experience.
+
+A common guidance by experts is that optimal method for diagnosing model
+fits is by plotting the data. The persistence of this advice to check
+the plots is curious, and investigating why this might be common advice
+is the subject of this paper. The paper is structured as follows. The
+next background section describes the types of departures that one
+expects to detect, and outlines a formal statistical process for reading
+residual plots, called visual inference. Section
+\ref{experimental-design} details the experimental design to compare the
+decision made by formal hypothesis testing, and how humans would read
+diagnostic plots. The results are reported in Section \ref{results}. We
+finish with a discussion on future work, in particular how the
+responsibility for residual plot reading might be relieved.
 
@@
 plots}\label{departures-from-good-residual-plots}}
-
-(change it to facet)
-
-(remove backticks for some plots)
 
@@
 
-\hypertarget{lineup-protocol}{%
-\subsubsection{Lineup protocol}\label{lineup-protocol}}
-
 One may argue that reading diagnostic plots is to some extent subjective
@@
 
+Data used in the \(m - 1\) null plots needs to be simulated. In
+regression diagnostics, sampling data consistent with \(H_0\) is
+equivalent to sampling data from the assumed model. As
+\citet{buja_statistical_2009} suggested, \(H_0\) is usually a composite
+hypothesis controlled by nuisance parameters. Since regression models
+can have various forms, there is no general solution to this problem,
+but it sometimes can be reduced to a so called ``reference
+distribution'' by applying one of the three methods: (i) sampling from a
+conditional distribution given a minimal sufficient statistic under
+\(H_0\), (ii) parametric bootstrap sampling with nuisance parameters
+estimated under \(H_0\), and (iii) Bayesian posterior predictive
+sampling. The conditional distribution given a minimal sufficient
+statistic is the best justified reference distribution among the three
+\citep{buja_statistical_2009}. Essentially, null residuals can be
+simulated by regressing \(N\) i.i.d standard normal random draws on the
+regressors, then rescaling it by the ratio of residual sum of square in
+two regressions.
+
 The effectiveness of lineup protocol for regression analysis is
@@
 
-\hypertarget{sampling-from-the-null-distribution}{%
-\subsubsection{Sampling from the null
-distribution}\label{sampling-from-the-null-distribution}}
-
-Data used in the \(m - 1\) null plots needs to be simulated. In
-regression diagnostics, sampling data consistent with \(H_0\) is
-equivalent to sampling data from the assumed model. As
-\citet{buja_statistical_2009} suggested, \(H_0\) is usually a composite
-hypothesis controlled by nuisance parameters. Since regression models
-can have various forms, there is no general solution to this problem,
-but it sometimes can be reduced to a so called ``reference
-distribution'' by applying one of the three methods: (i) sampling from a
-conditional distribution given a minimal sufficient statistic under
-\(H_0\), (ii) parametric bootstrap sampling with nuisance parameters
-estimated under \(H_0\), and (iii) Bayesian posterior predictive
-sampling. The conditional distribution given a minimal sufficient
-statistic is the best justified reference distribution among the three
-\citep{buja_statistical_2009}. Essentially, null residuals can be
-simulated by regressing \(N\) i.i.d standard normal random draws on the
-regressors, then rescaling it by the ratio of residual sum of square in
-two regressions.
-
-\hypertarget{calculating-p-values-for-the-visual-test}{%
-\subsubsection{\texorpdfstring{Calculating \(p\)-values for the visual
-test}{Calculating p-values for the visual test}}\label{calculating-p-values-for-the-visual-test}}
+\hypertarget{caluculation-of-p-value-effect-size-and-power}{%
+\section{Caluculation of p-value, effect size and
+power}\label{caluculation-of-p-value-effect-size-and-power}}
+
+Visual inference is
+
+To be able to properly compare conventional tests to visual tests, we
+need to define the effect size of the residual departure, the power of
+the tests under a certain level of effect size,
+
+\hypertarget{caluculation-of-effect-size}{%
+\subsubsection{Caluculation of effect
+size}\label{caluculation-of-effect-size}}
+
+\hypertarget{caluculation-of-p-value}{%
+\subsubsection{Caluculation of p-value}\label{caluculation-of-p-value}}
 
@@
 of a lineup of size \(m\) evaluated by \(K\) observer is given as
-\(P(X \geq x) = 1 - F(x) + f(x)\), where \(F(.)\) is the cumulative
-distribution function, \(f(.)\) is the probability mass function and
-\(x\) is the realization of number of observers correctly picking the
-data plot \citep{majumder_validation_2013}.
-
-As pointed out by \citet{vanderplas2021statistical}, this basic binomial
-model doesn't take into account the possible dependencies in the visual
-test due to repeated evaluations of the same lineup. And it is
-inapplicable to visual test where subjects are asked to select one or
-more ``most different'' plots from the lineup.
-\citet{vanderplas2021statistical} summarises three common scenarios in
-visual inference: (1) \(K\) different lineups are shown to \(K\)
-subjects, (2) \(K\) lineups with different null plots but the same data
-plot are shown to \(K\) subjects, and (3) the same lineup is shown to
-\(K\) subjects. Out of these three scenarios, Scenario 3 is the most
-common in previous studies as it puts the least constraints on the
-experiment design. For Scenario 3, \citet{vanderplas2021statistical}
-models the probability of a plot \(i\) being selected from a lineup as
-\(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
-\(i=1,...,m\) and \(\alpha > 0\). The number of times plot \(i\) being
-selected in \(K\) evaluations is denoted as \(c_i\). In case subject
-\(j\) makes multiple selections, \(1/s_j\) will be added to \(c_i\)
-instead of one, where \(s_j\) is the number of plots subject \(j\)
-selected for \(j=1,...K\). This ensures \(\sum_{i}c_i=K\). Since we are
-only interested in the selections of the data plot \(i\), the marginal
-model can be simplified to a beta-binomial model and thus the visual
-p-value is given as
-
-\begin{equation} \label{eq:pvalue-beta-binomial}
-P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+
-\end{equation}
-
-\noindent where \(B(.)\) is the beta function defined as
-
-\begin{equation} \label{eq:betafunction}
-B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0. 
-\end{equation}
-
-Note that Equation \ref{eq:pvalue-beta-binomial} given in
-\citet{vanderplas2021statistical} only works with non-negative integer
-\(c_i\). We extend the equation to non-negative real number \(c_i\) by
-applying a linear approximation
-
-\begin{equation} \label{eq:pvalue-beta-binomial-approx}
-P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \in \mathbb{R}_0^+,
-\end{equation}
-
-where \(P(C \geq \lceil c_i \rceil)\) is calculated using Equation
-\ref{eq:pvalue-beta-binomial} and \(P(C = \lfloor c_i \rfloor)\) is
-calculated by
-
-\begin{equation} \label{eq:pmf-beta-binomial}
-P(C = c_i) = {K \choose c_i}\frac{B(c_i + \alpha, K - c_i + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+.
-\end{equation}
-
-Besides, the parameter \(\alpha\) used in Equation
-\ref{eq:pvalue-beta-binomial} and \ref{eq:pmf-beta-binomial} is usually
-unknown and hence needs to be estimated from the survey data. For low
-values of \(\alpha\), only a few plots are attractive to the observers
-and tend to be selected. For higher values of \(\alpha\), the
-distribution of the probability of each plot being selected is more
-even. \citet{vanderplas2021statistical} defines that a plot is
-\(c\)-interesting if \(c\) or more participants select the plot as the
-most different. Given the definition, The expected number of plots
-selected at least \(c\) times, \(E[Z_c]\), is calculated as
-
-\begin{equation} \label{eq:c-interesting-expectation}
-E[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation}
-
-With Equation \ref{eq:c-interesting-expectation}, \(\alpha\) can be
-estimated using maximum likelihood estimation. But for precise estimate
-of \(\alpha\), additional responses to Rorschach lineups, which is a
-type of lineup that consists of plots constructed from the same null
-data generating mechanism, are required.
-
-\hypertarget{power-of-a-visual-test}{%
-\subsubsection{Power of a visual test}\label{power-of-a-visual-test}}
-
-The power of a model misspecification test is the probability that the
-null hypothesis is rejected given the regression model is misspecified.
-It is an important indicator when one is concerned about whether model
-assumptions have been violated. Although in practice, one might be more
-interested in knowing how much the residuals deviate from the model
-assumptions, and whether this deviation is of practical significance.
-
-As discussed in \citet{majumder_validation_2013}, the power of a visual
-test may depend on the ability of the particular subject, as the skill
-of the individual may affect the number of observers who identify the
-data plot from the lineup. Previously, it is addressed by modelling the
-probability of a subject \(j\) correctly picking the data plot from a
-lineup \(l\) using a mixed-effect logistic regression, with subjects
-treated as random effects \citep{majumder_validation_2013}. However, in
-the multiple selections scenario, having this probability is not
-sufficient to determine the power of a visual test because it does not
-provide information about the number of selections made by the subject
-for the calculation of the p-value.
-
-Instead, we directly estimate the probability of a lineup being rejected
-by assuming that individual skill has negligible effect on the variation
-of the power. The assumption is to simplify the model structure, thereby
-obviate costly large-scale experiments to estimate complex covariance
-matrices. The model is a logistic regression formulated as
-
-\begin{equation} \label{eq:logistic-regression-1-1}
-Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right),
-\end{equation}
-
-\noindent where \(\Lambda(.)\) is the standard logistic function given
-as \(\Lambda(z) = exp(z)/(1+exp(z))\). The effect size \(E\) is the only
-regressor and the intercept is fixed to \(log(0.05/0.95)\) so that
-\(\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05\).
-
-Effect size \(E\) is formulated as the conditional expectation of the
-sample-based effect size \(E_{sample}\) derived from the
-Kullback-Leibler divergence (see Appendix for the details) given by
-
-\small
-
-\begin{gather}  \label{eq:effect-size}
-E = E[E_{sample}|M],\\
-E_{sample} = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V})|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V})^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V})^{-1}\boldsymbol{\mu}_z\right),\\
-\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a,\\
-\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a^T\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a^T,
-\end{gather}
-
-\normalsize
-
-\noindent where \(M\) is a given model that generates the data,
-\(diag(.)\) is the diagonal matrix constructed from the diagonal
-elements of a matrix,
-\(\boldsymbol{X}_a = (\boldsymbol{1}, \boldsymbol{X})\) is the matrix of
-regressors including the intercept used in the regression equation,
-\(\sigma^2\boldsymbol{I}\) is the assumed variance of the error term
-when \(H_0\) is true, \(\boldsymbol{V}\) is the actual variance of the
-error term, and
-\(\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z\)
-is the expected values of the residuals with \(\boldsymbol{Z}\) be any
-higher order terms of \(\boldsymbol{X}\) leave out by the regression
-equation and \(\boldsymbol{\beta}_z\) be the corresponding coefficients.
-
-A conditional expectation is taken because \(E\) is designed to be
-invariant to different realizations of \(\boldsymbol{X}\). However,
-solving this conditional expectation analytically is very difficult as
-\(\boldsymbol{X}\) can follow arbitrary distributions. Therefore, we
-make an assumption that \(\boldsymbol{X}\) follows a particular
-multivariate normal distribution
-
-We make an assumption that all the realizations of \(\boldsymbol{X}\)
-are generated from a multivariate normal distribution
-\(N(\boldsymbol{0}, 0.3^2\boldsymbol{I})\) and then rescaled by the
-min-max normalization. The estimate of \(E\) is obtained by taking the
-average of the \(E_{sample}\) calculated for samples drawn from the
-given model \(M\) repeatedly. We use the normal distribution as a
-general choice to approximate the conditional effect size.
-
-To study various factors contributing to the power of the visual test,
-the same logistic regression model is fit on different subsets of the
-collated data grouped by levels of factors. These include the
-distribution of the regressor, type of the simulation model and the
-shape of the residual departures.
+\(P(X \geq x) = 1 - F(x) + f(x)\), where \(F(.)\) is the binomial
+cumulative distribution function, \(f(.)\) is the binomial probability
+mass function and \(x\) is the realization of number of observers
+correctly picking the data plot \citep{majumder_validation_2013}.
+
+--\textgreater{}
+
+\hypertarget{caluculation-of-power}{%
+\subsubsection{Caluculation of power}\label{caluculation-of-power}}
 
@@
 diagnostics, an experiment is conducted in three data collection
-periods. The two types of departures, non-linearity and
-heteroskedasticity, are primarily collected during data collection
-periods I and II. The data collection period III was designed primarily
-to measure human responses to null lineups so that the parameter
-\(\alpha\) in Equation \ref{eq:pvalue-beta-binomial} can be estimated.
-Additional lineups for both non-linearity and heteroskedasticity, using
-uniform \(X\) distribution, were included so that the participants were
-evaluating some lineups with signal also. It would be too frustrating
-for participants to only be assigned lineups with all null plots.
-Overall, we collected 7974 evaluations on 1152 unique lineups performed
-by 443 subjects throughout three data collection periods.
+periods. Two types of departures, non-linearity and heteroskedasticity,
+are primarily collected during data collection periods I and II. The
+data collection period III was designed primarily to measure human
+responses to null lineups so that the parameter \(\alpha\) in Equation
+\ref{eq:pvalue-beta-binomial} can be estimated. Additional lineups for
+both non-linearity and heteroskedasticity, using uniform \(X\)
+distribution, were included so that the participants were evaluating
+some lineups with signal also. It would be too frustrating for
+participants to only be assigned lineups with all null plots. Overall,
+we collected 7974 evaluations on 1152 unique lineups performed by 443
+subjects throughout three data collection periods.
 
@@
 For non-linearity patterns, the power curve of RESET test climbs
-aggressively from 7\% to 84\% as \(log_e(E)\) increases from 0 to 2.5,
+aggressively from 16\% to 100\% as \(log_e(E)\) increases from 0 to 2.5,
 while power of other tests respond inactively to the change of effect,

### 2023-02-28 paper_comparison/paper_comparison/paper_comparison.Rmd
correct the figures in pdf

@@
 
-        # Get the effect size from the pre-calculated table 
-        es <- filter(vi_survey, 
-                     shape == {{shape}},
-                     e_sigma == {{e_sigma}},
-                     n == {{n}}) %>%
-          head(1) %>%
-          pull(effect_size)
-
         # Return a data frame containing p-values of
@@
                n = n,
-               effect_size = es,
                F_p_value = mod$test(tmp_dat)$p_value,
@@
 
-        # Get the effect size from the pre-calculated table 
-        es <- filter(vi_survey, 
-                     a == {{a}}, 
-                     b == {{b}},
-                     n == {{n}}) %>%
-          head(1) %>%
-          pull(effect_size)
-
         # Return a data frame containing p-values of
@@
                n = n,
-               effect_size = es,
                F_p_value = POLY_MODEL$test(
@@
 
+```{r}
+# rewrite effect size here (if needed)
+
+```
+
+```{r}
+# Borrow effect size from the survey
+poly_conv_sim <- poly_conv_sim %>%
+  left_join(select(filter(vi_survey, type == "polynomial"), shape, e_sigma, n, x_dist, effect_size))
+
+heter_conv_sim <- heter_conv_sim %>%
+  left_join(select(filter(vi_survey, type == "heteroskedasticity"), a, b, n, x_dist, effect_size))
+```
+
 Title 
@@
 
-```{r different-x-dist-poly-power, fig.cap = "Comparison of power on lineups with different fitted value distributions, for conventional tests and visual tests.
+```{r different-x-dist-poly-power, fig.height = 4, fig.cap = "Comparison of power on lineups with different fitted value distributions, for conventional tests and visual tests.
 For visual tests, all the power curves are estimated using five evaluations on each lineup.

### 2023-02-28 paper_comparison/paper_comparison/paper_comparison.tex
correct the figures in pdf

@@
 For non-linearity patterns, the power curve of RESET test climbs
-aggressively from 16\% to 100\% as \(log_e(E)\) increases from 0 to 2.5,
+aggressively from 10\% to 100\% as \(log_e(E)\) increases from 0 to 2.5,
 while power of other tests respond inactively to the change of effect,

### 2023-02-28 paper_comparison/paper_comparison/paper_comparison.tex
rearrange the topics

@@
 
-\hypertarget{caluculation-of-p-value-effect-size-and-power}{%
-\section{Caluculation of p-value, effect size and
-power}\label{caluculation-of-p-value-effect-size-and-power}}
-
-Visual inference is
-
-To be able to properly compare conventional tests to visual tests, we
-need to define the effect size of the residual departure, the power of
-the tests under a certain level of effect size,
-
-\hypertarget{caluculation-of-effect-size}{%
-\subsubsection{Caluculation of effect
-size}\label{caluculation-of-effect-size}}
-
-\hypertarget{caluculation-of-p-value}{%
-\subsubsection{Caluculation of p-value}\label{caluculation-of-p-value}}
+\hypertarget{caculation-of-p-value-and-power}{%
+\section{Caculation of p-value and
+power}\label{caculation-of-p-value-and-power}}
+
+\hypertarget{p-value}{%
+\subsection{\texorpdfstring{\(p\)-value}{p-value}}\label{p-value}}
 
@@
 
-\hypertarget{caluculation-of-power}{%
-\subsubsection{Caluculation of power}\label{caluculation-of-power}}
+\hypertarget{power}{%
+\subsection{Power}\label{power}}
 
@@
 reliably.
+
+\hypertarget{effect-size}{%
+\subsection{Effect size}\label{effect-size}}
 

### 2023-02-28 paper_comparison/paper_comparison/paper_comparison.Rmd
rearrange the topics
tweaking start of paper

@@
 
-Diagnosing a model is the key to determining whether there is anything importantly wrong.
-In linear regression analysis, residuals are typically examined for model diagnostics.
+Diagnostics are the key to determining whether there is anything importantly wrong with a model.
+In linear regression analysis, residuals from the model fit are commonly used.
 Residuals summarise what is not captured by the model, and thus provide the capacity to identify what might be wrong.
@@
 
-# Caluculation of p-value, effect size and power
-
-Visual inference is
-
-To be able to properly compare conventional tests to visual tests, we need to define the effect size of the residual departure, the power of the tests under a certain level of effect size,
-
-### Caluculation of effect size
-
-### Caluculation of p-value
+# Caculation of p-value and power
+
+## $p$-value
 
@@
 
-### Caluculation of power
+## Power
 
@@
 
+## Effect size
+
 ## Experimental setup

### 2023-02-28 paper_comparison/paper_comparison/paper_comparison.tex
tweaking start of paper

@@
 
-Diagnosing a model is the key to determining whether there is anything
-importantly wrong. In linear regression analysis, residuals are
-typically examined for model diagnostics. Residuals summarise what is
-not captured by the model, and thus provide the capacity to identify
-what might be wrong.
+Diagnostics are the key to determining whether there is anything
+importantly wrong with a model. In linear regression analysis, residuals
+from the model fit are commonly used. Residuals summarise what is not
+captured by the model, and thus provide the capacity to identify what
+might be wrong.
 
@@
 For non-linearity patterns, the power curve of RESET test climbs
-aggressively from 10\% to 100\% as \(log_e(E)\) increases from 0 to 2.5,
+aggressively from 16\% to 100\% as \(log_e(E)\) increases from 0 to 2.5,
 while power of other tests respond inactively to the change of effect,

### 2023-02-28 paper_comparison/paper_comparison/paper_comparison.Rmd
update power desc

@@
 
-# Caculation of p-value and power
-
-## $p$-value
+# Estimation of p-value and power
+
+## p-value
 
 In hypothesis testing, a $p$-value is defined as the probability of observing test results as least as extreme as the observed result given $H_0$ is true.
-Within the context of visual inference, by involving $k$ independent observers, the $p$-value can be interpreted as the probability of having as many or more subjects detect the data plot than the observed result.
+Conventional hypothesis tests usually have an existing method to derive or compute $p$-value based on the null distribution.
+What we need to discuss in the following is the method to estimate $p$-value for a visual test.
+
+Within the context of visual inference, by involving $k$ independent observers, the visual $p$-value can be interpreted as the probability of having as many or more subjects detect the data plot than the observed result.
 
@@
 
-<!-- As pointed out by @vanderplas2021statistical, this basic binomial model doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup.
+As pointed out by @vanderplas2021statistical, this basic binomial model doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup.
 And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup.
@@
 This ensures $\sum_{i}c_i=K$.
-Since we are only interested in the selections of the data plot $i$, the marginal model can be simplified to a beta-binomial model and thus the visual p-value is given as -->
-
-<!-- <!-- TODO:Clarify this is only for integer c_i --> -->
-
-<!-- ```{=tex} -->
-<!-- \begin{equation} \label{eq:pvalue-beta-binomial} -->
-<!-- P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+ -->
-<!-- \end{equation} -->
-<!-- ``` -->
-
-<!-- \noindent where $B(.)$ is the beta function defined as -->
-
-<!-- ```{=tex} -->
-<!-- \begin{equation} \label{eq:betafunction} -->
-<!-- B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0.  -->
-<!-- \end{equation} -->
-<!-- ``` -->
-
-<!-- Note that Equation \ref{eq:pvalue-beta-binomial} given in @vanderplas2021statistical only works with non-negative integer $c_i$.
-We extend the equation to non-negative real number $c_i$ by applying a linear approximation -->
-
-<!-- ```{=tex} -->
-<!-- \begin{equation} \label{eq:pvalue-beta-binomial-approx} -->
-<!-- P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \in \mathbb{R}_0^+, -->
-<!-- \end{equation} -->
-<!-- ``` -->
-
-<!-- where $P(C \geq \lceil c_i \rceil)$ is calculated using Equation \ref{eq:pvalue-beta-binomial} and $P(C = \lfloor c_i \rfloor)$ is calculated by -->
-
-<!-- \begin{equation} \label{eq:pmf-beta-binomial} -->
-<!-- P(C = c_i) = {K \choose c_i}\frac{B(c_i + \alpha, K - c_i + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+. -->
-<!-- \end{equation} -->
-
-<!-- Besides, the parameter $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} and \ref{eq:pmf-beta-binomial} is usually unknown and hence needs to be estimated from the survey data.
+Since we are only interested in the selections of the data plot $i$, the marginal model can be simplified to a beta-binomial model and thus the visual p-value is given as
+
+```{=tex}
+\begin{equation} \label{eq:pvalue-beta-binomial}
+P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+
+\end{equation}
+```
+
+\noindent where $B(.)$ is the beta function defined as
+
+```{=tex}
+\begin{equation} \label{eq:betafunction}
+B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0.
+\end{equation}
+```
+
+Note that Equation \ref{eq:pvalue-beta-binomial} given in @vanderplas2021statistical only works with non-negative integer $c_i$.
+We extend the equation to non-negative real number $c_i$ by applying a linear approximation
+
+```{=tex}
+\begin{equation} \label{eq:pvalue-beta-binomial-approx}
+P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \in \mathbb{R}_0^+,
+\end{equation}
+```
+
+where $P(C \geq \lceil c_i \rceil)$ is calculated using Equation \ref{eq:pvalue-beta-binomial} and $P(C = \lfloor c_i \rfloor)$ is calculated by
+
+\begin{equation} \label{eq:pmf-beta-binomial}
+P(C = c_i) = {K \choose c_i}\frac{B(c_i + \alpha, K - c_i + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+.
+\end{equation}
+
+Besides, the parameter $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} and \ref{eq:pmf-beta-binomial} is usually unknown and hence needs to be estimated from the survey data.
 For low values of $\alpha$, only a few plots are attractive to the observers and tend to be selected.
@@
 @vanderplas2021statistical defines that a plot is $c$-interesting if $c$ or more participants select the plot as the most different.
-Given the definition, The expected number of plots selected at least $c$ times, $E[Z_c]$, is calculated as -->
-
-<!-- ```{=tex} -->
-<!-- \begin{equation} \label{eq:c-interesting-expectation} -->
-<!-- E[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation} -->
-<!-- ``` -->
-
-<!-- With Equation \ref{eq:c-interesting-expectation}, $\alpha$ can be estimated using maximum likelihood estimation.
-But for precise estimate of $\alpha$, additional responses to Rorschach lineups, which is a type of lineup that consists of plots constructed from the same null data generating mechanism, are required.
--->
+Given the definition, The expected number of plots selected at least $c$ times, $E[Z_c]$, is calculated as
+
+```{=tex}
+\begin{equation} \label{eq:c-interesting-expectation}
+E[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation}
+```
+
+With Equation \ref{eq:c-interesting-expectation}, $\alpha$ can be estimated using maximum likelihood estimation.
+But for precise estimate of $\alpha$, additional human responses to Rorschach lineups, which is a type of lineup that consists of plots constructed from the same null data generating mechanism, are required.
 
@@
 
-<!-- The power of a model misspecification test is the probability that the null hypothesis is rejected given the regression model is misspecified.
-It is an important indicator when one is concerned about whether model assumptions have been violated.
+The power of a model misspecification test is the probability that the null hypothesis is rejected given the regression model is misspecified in a specific way. It is an important indicator
+when one is concerned about whether model assumptions have been violated.
 Although in practice, one might be more interested in knowing how much the residuals deviate from the model assumptions, and whether this deviation is of practical significance.
--->
-
-<!-- As discussed in @majumder_validation_2013, the power of a visual test may depend on the ability of the particular subject, as the skill of the individual may affect the number of observers who identify the data plot from the lineup.
-Previously, it is addressed by modelling the probability of a subject $j$ correctly picking the data plot from a lineup $l$ using a mixed-effect logistic regression, with subjects treated as random effects [@majumder_validation_2013].
-However, in the multiple selections scenario, having this probability is not sufficient to determine the power of a visual test because it does not provide information about the number of selections made by the subject for the calculation of the p-value.
--->
-
-<!-- Instead, we directly estimate the probability of a lineup being rejected by assuming that individual skill has negligible effect on the variation of the power.
-The assumption is to simplify the model structure, thereby obviate costly large-scale experiments to estimate complex covariance matrices.
-The model is a logistic regression formulated as  -->
-
-<!-- ```{=tex} -->
-<!-- \begin{equation} \label{eq:logistic-regression-1-1} -->
-<!-- Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right), -->
-<!-- \end{equation} -->
-<!-- ``` -->
-<!-- \noindent where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
+
+The power of a conventional hypothesis test is affected by both the true parameter $\boldsymbol{\theta}$ and the sample size $n$.
+These two can be quantified in terms of effect size $E$ to measure the strength of the residual departures from the model assumptions.
+Details about the effect size is provided in Section \ref{effect-size} after the introduction of the simulation model used in our human subject experiment.
+The theoretical power of a test is sometimes not a trival solution.
+For a given regression model under a specific effect size, if the data generating process is also known, the power of a conventional test can be estimated by counting the number of times the test rejects the null hypothesis in a repeated sampling exercise
+and divided by the number of times simulated from the known data generating process. This is the method we used to estimate the power of conventional tests in our experiment.
+
+The power of a visual test on the other hand, may additionally depend on the ability of the particular subject, as the skill of the individual may affect the number of observers who identify the data plot from the lineup [@majumder_validation_2013].
+To address this issue, @majumder_validation_2013 models the probability of a subject $j$ correctly picking the data plot from a lineup $l$ using a mixed-effect logistic regression, with subjects treated as random effect.
+Then, the estimated power of a visual test evaluated by a single subject is the predicted value obtained from the mix-effect model.
+
+However, this mix-effect model does not work with scenario where subjects are asked to select one or more "most different" plots.
+In this scenario, having the probability of a subject $j$ correctly picking the data plot from a lineup $l$ is insufficient to determine the power of a visual test because it does not provide information about the number of selections made by the subject for the calculation of the p-value (See Equation \ref{eq:pvalue-beta-binomial-approx}).
+
+Therefore, we directly estimate the probability of a lineup being rejected by assuming that individual skill has negligible effect on the variation of the power.
+This assumption is not necessary true, but it helps simplifying the model structure, thereby obviate a costly large-scale experiment to estimate complex covariance matrices.
+There is a discussion provided in the Appendix about a potential solution to model individual effect in the multiple selections scenario.
+
+(should we just link the power together for conventional test?)
+
+We use the same logistic regression model to fit the power of conventional tests and visual tests formulated as
+
+```{=tex}
+\begin{equation} \label{eq:logistic-regression-1-1}
+Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right),
+\end{equation}
+```
+\noindent where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
 The effect size $E$ is the only regressor and the intercept is fixed to $log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$.
--->
-
-<!-- Effect size $E$ is formulated as the conditional expectation of the sample-based effect size $E_{sample}$ derived from the Kullback-Leibler divergence (see Appendix for the details) given by -->
-
-<!-- \footnotesize -->
-
-<!-- \begin{gather}  \label{eq:effect-size} -->
-<!-- E = E[E_{sample}|M],\\ -->
-<!-- E_{sample} = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}\boldsymbol{\mu}_z\right),\\ -->
-<!-- \boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a,\\ -->
-<!-- \boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a^T\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a^T, -->
-<!-- \end{gather} -->
-
-<!-- \normalsize -->
-
-<!-- \noindent where $M$ is a given model that generates the data, $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,  $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of regressors including the intercept used in the regression equation, $\sigma^2\boldsymbol{I}$ is the assumed variance of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual variance of the error term, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of the residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients.
--->
-
-<!-- A conditional expectation is taken because $E$ is designed to be invariant to different realizations of $\boldsymbol{X}$.
-However, solving this conditional expectation analytically is very difficult as $\boldsymbol{X}$ can follow arbitrary distributions.
-Therefore, we make an assumption that $\boldsymbol{X}$ follows a particular multivariate normal distribution -->
-
-<!-- We make an assumption that all the realizations of $\boldsymbol{X}$ are generated from a multivariate normal distribution $N(\boldsymbol{0}, 0.3^2\boldsymbol{I})$ and then rescaled by the min-max normalization.
-The estimate of $E$ is obtained by taking the average of the $E_{sample}$ calculated for samples drawn from the given model $M$ repeatedly.
-We use the normal distribution as a general choice to approximate the conditional effect size.
--->
-
-<!-- To study various factors contributing to the power of the visual test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors.
+
+To study various factors contributing to the power of test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors.
 These include the distribution of the regressor, type of the simulation model and the shape of the residual departures.
--->
 
@@
 
+Effect size $E$ is formulated as the conditional expectation of the sample-based effect size $E_{sample}$ derived from the Kullback-Leibler divergence (see Appendix for the details) given by
+
+\footnotesize
+
+\begin{gather}  \label{eq:effect-size}
+E = E[E_{sample}|M],\\
+E_{sample} = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}\boldsymbol{\mu}_z\right),\\
+\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a,\\
+\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a^T\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a^T,
+\end{gather}
+
+\normalsize
+
+\noindent where $M$ is a given model that generates the data, $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,  $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of regressors including the intercept used in the regression equation, $\sigma^2\boldsymbol{I}$ is the assumed variance of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual variance of the error term, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of the residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients.
+
+A conditional expectation is taken because $E$ is designed to be invariant to different realizations of $\boldsymbol{X}$.
+However, solving this conditional expectation analytically is very difficult as $\boldsymbol{X}$ can follow arbitrary distributions.
+Therefore, we make an assumption that $\boldsymbol{X}$ follows a particular multivariate normal distribution
+
+We make an assumption that all the realizations of $\boldsymbol{X}$ are generated from a multivariate normal distribution $N(\boldsymbol{0}, 0.3^2\boldsymbol{I})$ and then rescaled by the min-max normalization.
+The estimate of $E$ is obtained by taking the average of the $E_{sample}$ calculated for samples drawn from the given model $M$ repeatedly.
+We use the normal distribution as a general choice to approximate the conditional effect size.
+
 ## Experimental setup

### 2023-02-28 paper_comparison/paper_comparison/paper_comparison.Rmd
update effect size desc

@@
 
-## Effect size
-
-Effect size $E$ is formulated as the conditional expectation of the sample-based effect size $E_{sample}$ derived from the Kullback-Leibler divergence (see Appendix for the details) given by
-
-\footnotesize
-
-\begin{gather}  \label{eq:effect-size}
-E = E[E_{sample}|M],\\
-E_{sample} = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}\boldsymbol{\mu}_z\right),\\
-\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a,\\
-\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a^T\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a^T,
-\end{gather}
-
-\normalsize
-
-\noindent where $M$ is a given model that generates the data, $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,  $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of regressors including the intercept used in the regression equation, $\sigma^2\boldsymbol{I}$ is the assumed variance of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual variance of the error term, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of the residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients.
-
-A conditional expectation is taken because $E$ is designed to be invariant to different realizations of $\boldsymbol{X}$.
-However, solving this conditional expectation analytically is very difficult as $\boldsymbol{X}$ can follow arbitrary distributions.
-Therefore, we make an assumption that $\boldsymbol{X}$ follows a particular multivariate normal distribution
-
-We make an assumption that all the realizations of $\boldsymbol{X}$ are generated from a multivariate normal distribution $N(\boldsymbol{0}, 0.3^2\boldsymbol{I})$ and then rescaled by the min-max normalization.
-The estimate of $E$ is obtained by taking the average of the $E_{sample}$ calculated for samples drawn from the given model $M$ repeatedly.
-We use the normal distribution as a general choice to approximate the conditional effect size.
-
 ## Experimental setup
@@
 It can be observed from Figure \ref{fig:different-n} that with fewer data points drawn in a residual plot, the visual pattern is more difficult to be detected.
+
+### Effect size
+
+Effect size $E$ can be defined as the difference of a parameter for a particular model or distribution, or a statistic derived from a sample.
+Importantly, it needs to reflect the treatment we try to measure.
+According to the underlying simulated model, we define the effect size $E = \sqrt{n}/(\sigma k)$ for the non-linearity model and $E = \sqrt{n}/(|a| + 1)$ for the heteroskedasticity model.
+The parameter $k$ is the number of turning points embedded in the residual plot.
+It is used as a penalty term to reduce the effect size for complex shapes.
+There are one, two, three and five turning points in "U", "S", "M", "Triple-U" shape respectively.
+The denominator of the effect size of the heteroskedasticity model is $|a| + 1$.
+It is used to
 

### 2023-03-01 paper_comparison/paper_comparison/paper_comparison.Rmd
update the power cmp part

@@
 
+(do we actually need the penalty term?)
+
 Effect size $E$ can be defined as the difference of a parameter for a particular model or distribution, or a statistic derived from a sample.
 Importantly, it needs to reflect the treatment we try to measure.
-According to the underlying simulated model, we define the effect size $E = \sqrt{n}/(\sigma k)$ for the non-linearity model and $E = \sqrt{n}/(|a| + 1)$ for the heteroskedasticity model.
+According to the underlying simulated model, we define the effect size $E = \sqrt{n}/(\sigma k)$ for the non-linearity model and $E = b\sqrt{n}/(|a| + 1)$ for the heteroskedasticity model.
 The parameter $k$ is the number of turning points embedded in the residual plot.
@@
 
-To compare the power of visual tests and conventional tests, we simulate 200000 samples evenly from the models used in the human subject experiments, and record the conventional $p$-values produced from every sample.
 We pick RESET test (`resettest`) and Breusch-Pagan test (`bptest`) from the R package `lmtest`, and Shapiro-Wilk test (`shapiro.test`) from the built-in R package `stats`.
@@
 
-target_sigma <- c(3.65, 2.84, 2.21, 1.72, 1.34, 1.04, 0.81, 0.63, 0.49, 0.38, 0.3)
+target_sigma <- c(5, 3.03, 1.84, 1.12, 0.68, 0.41, 0.25)
 
@@
 es <- map_dbl(target_sigma, 
-              ~log(poly_model(shape = 2, 
+              function(.x){
+                mod <- poly_model(shape = 2, 
                               x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, 
-                              sigma = .x)$
-                     average_effect_size(n = 100, type = "simple")
-                   )
+                              sigma = .x)
+                log(mod$sample_effect_size(mod$gen(100), type = "simple"))
+                }
               )
@@
 # Use area() to design layout
-map(1:11, ~patchwork::area(21, .x, 22, .x)) %>%
+map(1:7, ~patchwork::area(21, .x, 22, .x)) %>%
   reduce(c) %>%
-  c(patchwork::area(3, 1, 20, 11), .) %>%
+  c(patchwork::area(3, 1, 20, 7), .) %>%
   patchwork::wrap_plots(append(plist, list(p), after = 0), 
@@
   scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
+  # scale_size_manual(values = seq(1, 9)) +
   scale_x_continuous(breaks = seq(1, 7, 1), minor_breaks = NULL) +
@@
 
-target_b <- c(0.17, 0.23, 0.32, 0.46, 0.67, 1.04, 1.7, 3, 6.3, 16.2, 61)
+target_b <- c(0.54, 1.48, 4.02, 10.92, 29.68, 80.69, 219.33)
 
@@
 es <- map_dbl(target_b, 
-              ~log(heter_model(a = 1, 
+              function(.x){
+                mod <- heter_model(a = 1, 
                                    b = .x, 
-                               x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))})$
-                     average_effect_size(n = 100, type = "simple")
-                   )
+                                   x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))})
+                log(mod$sample_effect_size(mod$gen(100), type = "simple"))
+                }
               )
@@
 # Use area() to design layout
-map(1:11, ~patchwork::area(21, .x, 22, .x)) %>%
+map(1:7, ~patchwork::area(21, .x, 22, .x)) %>%
   reduce(c) %>%
-  c(patchwork::area(3, 1, 20, 11), .) %>%
+  c(patchwork::area(3, 1, 20, 7), .) %>%
 patchwork::wrap_plots(append(plist, list(p), after = 0), 

### 2023-03-01 paper_comparison/paper_comparison/paper_comparison.tex
update the power cmp part

--- 2023-02-28 paper_comparison/paper_comparison/paper_comparison.tex
@@
 
-\hypertarget{caculation-of-p-value-and-power}{%
-\section{Caculation of p-value and
-power}\label{caculation-of-p-value-and-power}}
+\hypertarget{estimation-of-p-value-and-power}{%
+\section{Estimation of p-value and
+power}\label{estimation-of-p-value-and-power}}
 
 \hypertarget{p-value}{%
-\subsection{\texorpdfstring{\(p\)-value}{p-value}}\label{p-value}}
+\subsection{p-value}\label{p-value}}
 
@@
 observing test results as least as extreme as the observed result given
-\(H_0\) is true. Within the context of visual inference, by involving
-\(k\) independent observers, the \(p\)-value can be interpreted as the
-probability of having as many or more subjects detect the data plot than
-the observed result.
+\(H_0\) is true. Conventional hypothesis tests usually have an existing
+method to derive or compute \(p\)-value based on the null distribution.
+What we need to discuss in the following is the method to estimate
+\(p\)-value for a visual test.
+
+Within the context of visual inference, by involving \(k\) independent
+observers, the visual \(p\)-value can be interpreted as the probability
+of having as many or more subjects detect the data plot than the
+observed result.
 
@@
 
---\textgreater{}
+As pointed out by \citet{vanderplas2021statistical}, this basic binomial
+model doesn't take into account the possible dependencies in the visual
+test due to repeated evaluations of the same lineup. And it is
+inapplicable to visual test where subjects are asked to select one or
+more ``most different'' plots from the lineup.
+\citet{vanderplas2021statistical} summarises three common scenarios in
+visual inference: (1) \(K\) different lineups are shown to \(K\)
+subjects, (2) \(K\) lineups with different null plots but the same data
+plot are shown to \(K\) subjects, and (3) the same lineup is shown to
+\(K\) subjects. Out of these three scenarios, Scenario 3 is the most
+common in previous studies as it puts the least constraints on the
+experiment design. For Scenario 3, \citet{vanderplas2021statistical}
+models the probability of a plot \(i\) being selected from a lineup as
+\(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
+\(i=1,...,m\) and \(\alpha > 0\). The number of times plot \(i\) being
+selected in \(K\) evaluations is denoted as \(c_i\). In case subject
+\(j\) makes multiple selections, \(1/s_j\) will be added to \(c_i\)
+instead of one, where \(s_j\) is the number of plots subject \(j\)
+selected for \(j=1,...K\). This ensures \(\sum_{i}c_i=K\). Since we are
+only interested in the selections of the data plot \(i\), the marginal
+model can be simplified to a beta-binomial model and thus the visual
+p-value is given as
+
+\begin{equation} \label{eq:pvalue-beta-binomial}
+P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+
+\end{equation}
+
+\noindent where \(B(.)\) is the beta function defined as
+
+\begin{equation} \label{eq:betafunction}
+B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0.
+\end{equation}
+
+Note that Equation \ref{eq:pvalue-beta-binomial} given in
+\citet{vanderplas2021statistical} only works with non-negative integer
+\(c_i\). We extend the equation to non-negative real number \(c_i\) by
+applying a linear approximation
+
+\begin{equation} \label{eq:pvalue-beta-binomial-approx}
+P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \in \mathbb{R}_0^+,
+\end{equation}
+
+where \(P(C \geq \lceil c_i \rceil)\) is calculated using Equation
+\ref{eq:pvalue-beta-binomial} and \(P(C = \lfloor c_i \rfloor)\) is
+calculated by
+
+\begin{equation} \label{eq:pmf-beta-binomial}
+P(C = c_i) = {K \choose c_i}\frac{B(c_i + \alpha, K - c_i + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+.
+\end{equation}
+
+Besides, the parameter \(\alpha\) used in Equation
+\ref{eq:pvalue-beta-binomial} and \ref{eq:pmf-beta-binomial} is usually
+unknown and hence needs to be estimated from the survey data. For low
+values of \(\alpha\), only a few plots are attractive to the observers
+and tend to be selected. For higher values of \(\alpha\), the
+distribution of the probability of each plot being selected is more
+even. \citet{vanderplas2021statistical} defines that a plot is
+\(c\)-interesting if \(c\) or more participants select the plot as the
+most different. Given the definition, The expected number of plots
+selected at least \(c\) times, \(E[Z_c]\), is calculated as
+
+\begin{equation} \label{eq:c-interesting-expectation}
+E[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation}
+
+With Equation \ref{eq:c-interesting-expectation}, \(\alpha\) can be
+estimated using maximum likelihood estimation. But for precise estimate
+of \(\alpha\), additional human responses to Rorschach lineups, which is
+a type of lineup that consists of plots constructed from the same null
+data generating mechanism, are required.
 
@@
 \subsection{Power}\label{power}}
+
+The power of a model misspecification test is the probability that the
+null hypothesis is rejected given the regression model is misspecified
+in a specific way. It is an important indicator when one is concerned
+about whether model assumptions have been violated. Although in
+practice, one might be more interested in knowing how much the residuals
+deviate from the model assumptions, and whether this deviation is of
+practical significance.
+
+The power of a conventional hypothesis test is affected by both the true
+parameter \(\boldsymbol{\theta}\) and the sample size \(n\). These two
+can be quantified in terms of effect size \(E\) to measure the strength
+of the residual departures from the model assumptions. Details about the
+effect size is provided in Section \ref{effect-size} after the
+introduction of the simulation model used in our human subject
+experiment. The theoretical power of a test is sometimes not a trival
+solution. For a given regression model under a specific effect size, if
+the data generating process is also known, the power of a conventional
+test can be estimated by counting the number of times the test rejects
+the null hypothesis in a repeated sampling exercise and divided by the
+number of times simulated from the known data generating process. This
+is the method we used to estimate the power of conventional tests in our
+experiment.
+
+The power of a visual test on the other hand, may additionally depend on
+the ability of the particular subject, as the skill of the individual
+may affect the number of observers who identify the data plot from the
+lineup \citep{majumder_validation_2013}. To address this issue,
+\citet{majumder_validation_2013} models the probability of a subject
+\(j\) correctly picking the data plot from a lineup \(l\) using a
+mixed-effect logistic regression, with subjects treated as random
+effect. Then, the estimated power of a visual test evaluated by a single
+subject is the predicted value obtained from the mix-effect model.
+
+However, this mix-effect model does not work with scenario where
+subjects are asked to select one or more ``most different'' plots. In
+this scenario, having the probability of a subject \(j\) correctly
+picking the data plot from a lineup \(l\) is insufficient to determine
+the power of a visual test because it does not provide information about
+the number of selections made by the subject for the calculation of the
+p-value (See Equation \ref{eq:pvalue-beta-binomial-approx}).
+
+Therefore, we directly estimate the probability of a lineup being
+rejected by assuming that individual skill has negligible effect on the
+variation of the power. This assumption is not necessary true, but it
+helps simplifying the model structure, thereby obviate a costly
+large-scale experiment to estimate complex covariance matrices. There is
+a discussion provided in the Appendix about a potential solution to
+model individual effect in the multiple selections scenario.
+
+(should we just link the power together for conventional test?)
+
+We use the same logistic regression model to fit the power of
+conventional tests and visual tests formulated as
+
+\begin{equation} \label{eq:logistic-regression-1-1}
+Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right),
+\end{equation}
+
+\noindent where \(\Lambda(.)\) is the standard logistic function given
+as \(\Lambda(z) = exp(z)/(1+exp(z))\). The effect size \(E\) is the only
+regressor and the intercept is fixed to \(log(0.05/0.95)\) so that
+\(\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05\).
+
+To study various factors contributing to the power of test, the same
+logistic regression model is fit on different subsets of the collated
+data grouped by levels of factors. These include the distribution of the
+regressor, type of the simulation model and the shape of the residual
+departures.
 
@@
 
-\hypertarget{effect-size}{%
-\subsection{Effect size}\label{effect-size}}
-
 \hypertarget{experimental-setup}{%
@@
 
+\hypertarget{effect-size}{%
+\subsubsection{Effect size}\label{effect-size}}
+
+(do we actually need the penalty term?)
+
+Effect size \(E\) can be defined as the difference of a parameter for a
+particular model or distribution, or a statistic derived from a sample.
+Importantly, it needs to reflect the treatment we try to measure.
+According to the underlying simulated model, we define the effect size
+\(E = \sqrt{n}/(\sigma k)\) for the non-linearity model and
+\(E = b\sqrt{n}/(|a| + 1)\) for the heteroskedasticity model. The
+parameter \(k\) is the number of turning points embedded in the residual
+plot. It is used as a penalty term to reduce the effect size for complex
+shapes. There are one, two, three and five turning points in ``U'',
+``S'', ``M'', ``Triple-U'' shape respectively. The denominator of the
+effect size of the heteroskedasticity model is \(|a| + 1\). It is used
+to
+
 \hypertarget{subject-allocation}{%
@@
 
-To compare the power of visual tests and conventional tests, we simulate
-200000 samples evenly from the models used in the human subject
-experiments, and record the conventional \(p\)-values produced from
-every sample. We pick RESET test (\texttt{resettest}) and Breusch-Pagan
-test (\texttt{bptest}) from the R package \texttt{lmtest}, and
-Shapiro-Wilk test (\texttt{shapiro.test}) from the built-in R package
-\texttt{stats}. Among them, RESET test is the only exact and appropriate
-test in this scenario. Both the Breusch-Pagan test and the Shapiro-Wilk
-test are approximate and inappropriate tests. Their estimated power is
-shown in Figure \ref{fig:polypower}. To set up the RESET test, we
-include different powers of fitted values as proxies. According to
+We pick RESET test (\texttt{resettest}) and Breusch-Pagan test
+(\texttt{bptest}) from the R package \texttt{lmtest}, and Shapiro-Wilk
+test (\texttt{shapiro.test}) from the built-in R package \texttt{stats}.
+Among them, RESET test is the only exact and appropriate test in this
+scenario. Both the Breusch-Pagan test and the Shapiro-Wilk test are
+approximate and inappropriate tests. Their estimated power is shown in
+Figure \ref{fig:polypower}. To set up the RESET test, we include
+different powers of fitted values as proxies. According to
 \citet{ramsey_tests_1969}, there are no general rules for the power of

### 2023-03-01 paper_comparison/paper_comparison/paper_comparison.Rmd
update text in the power cmp

@@
 
-For non-linearity patterns, the power curve of RESET test climbs aggressively from
-`r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(0), offset0 = log(0.05/0.95)), type = "response")) * 100)`% to `r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(2.5), offset0 = log(0.05/0.95)), type = "response")) * 100)`% as $log_e(E)$ increases from 0 to 2.5, while power of other tests respond inactively to the change of effect, showing that RESET test is way more sensitive to the type of model defects that being considered.
-Meanwhile, no noticeable residual departures can be spotted from the example residual plots.
-
-In terms of heteroskedasticity patterns, the power of Breusch-Pagan test is almost always greater than the power of visual test.
-For $log_e(E)$ at around 2.5, where the power curve of the visual test remains at a low level, the Breusch-Pagan test has around 50% chance of rejecting $H_0$.
-Similarly, the visual feature is nearly unobservable from the example residual plots.
-
-The power of visual test arises steadily as $log_e(E)$ increases from 3 to 5 for both non-linearity patterns and heteroskedasticity patterns, suggesting that the effect starts to make significant impact on the degree of the presence of the designed visual features.
-This can also be observed from the example residual plots that when $log_e(E) = 3.5$, a weak "S-shape" and a weak "triangle" shape are presented in Figure \ref{fig:polypower} and Figure \ref{fig:heterpower} respectively.
-The visual pattern becomes much clearer as $log_e(E)$ increases.
-At $log_e(E) = 5$, the power reaches almost 100%.
-
-The power of all inappropriate tests show improvement as the effect increases but at a lower rate than the visual test in both scenarios.
+For both the non-linearity pattern and the heteroskedasticity pattern, the power curves of appropriate conventional tests are constantly greater than the power curves of visual tests and other inappropriate conventional tests.
+At $log(E) = 1$ for non-linearity model, where the residual departures are unnoticeable from the corresponding residual plot, the RESET test has above 25% of chance of rejecting the null hypothesis, while only a few visual tests would result in a rejection.
+Similarly, at $log(E) = 3$ for heteroskedasticity model, the BP test has around 25% power but the non-constant variance visual pattern is almost unobserved from the residual plot and the power of the visual test is close to the significance level at this point.
+
+Further, the power of all inappropriate tests show improvement as the effect increases but at a lower rate than the visual test in both scenarios.
 This coincides the point made by @cook1982residuals that residual-based tests for a specific type of model defect may be sensitive to other types of model defects.
@@
 
-target_b <- c(0.54, 1.48, 4.02, 10.92, 29.68, 80.69, 219.33)
+target_b <- c(0.27, 0.74, 2.01, 5.46, 14.84, 40.34, 109.66)
 
@@
               function(.x){
-                mod <- heter_model(a = 1, 
+                mod <- heter_model(a = 0, 
                                    b = .x, 
@@
 # 133
-set.seed(135)
+set.seed(100)
 
@@
   j <- j + 1
-  plist[[j]] <- heter_model(a = 1, 
+  plist[[j]] <- heter_model(a = 0, 
                             b = i, 

### 2023-03-01 paper_comparison/paper_comparison/paper_comparison.tex
update text in the power cmp

@@
 
-For non-linearity patterns, the power curve of RESET test climbs
-aggressively from 10\% to 100\% as \(log_e(E)\) increases from 0 to 2.5,
-while power of other tests respond inactively to the change of effect,
-showing that RESET test is way more sensitive to the type of model
-defects that being considered. Meanwhile, no noticeable residual
-departures can be spotted from the example residual plots.
-
-In terms of heteroskedasticity patterns, the power of Breusch-Pagan test
-is almost always greater than the power of visual test. For \(log_e(E)\)
-at around 2.5, where the power curve of the visual test remains at a low
-level, the Breusch-Pagan test has around 50\% chance of rejecting
-\(H_0\). Similarly, the visual feature is nearly unobservable from the
-example residual plots.
-
-The power of visual test arises steadily as \(log_e(E)\) increases from
-3 to 5 for both non-linearity patterns and heteroskedasticity patterns,
-suggesting that the effect starts to make significant impact on the
-degree of the presence of the designed visual features. This can also be
-observed from the example residual plots that when \(log_e(E) = 3.5\), a
-weak ``S-shape'' and a weak ``triangle'' shape are presented in Figure
-\ref{fig:polypower} and Figure \ref{fig:heterpower} respectively. The
-visual pattern becomes much clearer as \(log_e(E)\) increases. At
-\(log_e(E) = 5\), the power reaches almost 100\%.
-
-The power of all inappropriate tests show improvement as the effect
-increases but at a lower rate than the visual test in both scenarios.
-This coincides the point made by \citet{cook1982residuals} that
-residual-based tests for a specific type of model defect may be
+For both the non-linearity pattern and the heteroskedasticity pattern,
+the power curves of appropriate conventional tests are constantly
+greater than the power curves of visual tests and other inappropriate
+conventional tests. At \(log(E) = 1\) for non-linearity model, where the
+residual departures are unnoticeable from the corresponding residual
+plot, the RESET test has above 25\% of chance of rejecting the null
+hypothesis, while only a few visual tests would result in a rejection.
+Similarly, at \(log(E) = 3\) for heteroskedasticity model, the BP test
+has around 25\% power but the non-constant variance visual pattern is
+almost unobserved from the residual plot and the power of the visual
+test is close to the significance level at this point.
+
+Further, the power of all inappropriate tests show improvement as the
+effect increases but at a lower rate than the visual test in both
+scenarios. This coincides the point made by \citet{cook1982residuals}
+that residual-based tests for a specific type of model defect may be
 sensitive to other types of model defects. The power curve of RESET test

### 2023-03-01 paper_comparison/paper_comparison/paper_comparison.Rmd
update example residual plot and caption

@@
 Where would you draw the line of too much non-linearity in the residuals?
-For the RESET test this is around log effect size 2.5, but for the visual test it is around 3.5.", dev = 'png', dpi = 450}
+For the RESET test this is around log effect size 1.5, but for the visual test it is around 2.", dev = 'png', dpi = 450}
 
@@
 Where would you draw the line of too much heteroskedasticity in the residuals?
-For the BP test this is around log effect size 2.5, but for the visual test it is around 3.5.", dev = 'png', dpi = 450}
+For the BP test this is around log effect size 3, but for the visual test it is around 5.", dev = 'png', dpi = 450}
 
@@
 # 133
-set.seed(100)
+set.seed(112)
 
@@
 
-```{r p-value-comparison, fig.width = 8, fig.height = 4, fig.cap = "Rejection rate ($p$-value $<0.05$) of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot.
+```{r p-value-comparison, fig.width = 8, fig.height = 4, fig.cap = "Rejection rate ($p$-value $\\leq0.05$) of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot.
 The visual test rejects less frequently than the conventional test.

### 2023-03-01 paper_comparison/paper_comparison/paper_comparison.tex
update example residual plot and caption

@@
 Where would you draw the line of too much non-linearity in the residuals?
-For the RESET test this is around log effect size 2.5, but for the visual test it is around 3.5.}\label{fig:polypower}
+For the RESET test this is around log effect size 1.5, but for the visual test it is around 2.}\label{fig:polypower}
 \end{figure}
@@
 Where would you draw the line of too much heteroskedasticity in the residuals?
-For the BP test this is around log effect size 2.5, but for the visual test it is around 3.5.}\label{fig:heterpower}
+For the BP test this is around log effect size 3, but for the visual test it is around 5.}\label{fig:heterpower}
 \end{figure}
@@
 
-\caption{Rejection rate ($p$-value $<0.05$) of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot.
+\caption{Rejection rate ($p$-value $\leq0.05$) of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot.
 The visual test rejects less frequently than the conventional test.

### 2023-03-01 paper_comparison/paper_comparison/paper_comparison.Rmd
fix text in different shape power

@@
 A primary factor contributes to the non-linearity model is the shape of the non-linearity pattern.
-According to Figure \ref{fig:poly-power-uniform-j}, conventional tests have higher power in testing shapes constructed from lower orders of Hermite polynomials.
+According to Figure \ref{fig:poly-power-uniform-j}, conventional tests have roughly the same power in testing shapes constructed from low orders of Hermite polynomials.
 But the power in testing the "Triple-U" shape drops significantly.
@@
 For visual tests, based on the orders of the Hermite polynomials, we expect the "U" shape will be the easiest one to be detected by subjects followed by the "S", "M" and "Triple-U" shape.
-From Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power curves are mostly aligned with our expectation, except for the "M" shape, which is as easy to be detected as the "S" shape.
-This implies, unlike the conventional test, the readability of the shape do not strictly follow the degree of the polynomials.
+From Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power curves are mostly aligned with our expectation, except for the "M" shape, which is the easiest shape to be detected.
+This implies the visibility of the shape do not strictly follow the degree of the polynomials.
 

### 2023-03-01 paper_comparison/paper_comparison/paper_comparison.Rmd
subsection on hypotheses

@@
 
-# Estimation of p-value and power
-
-## p-value
+# Calculation of $p$-value and power
+
+## Hypotheses
+
+A paragraph on what hypothesis(es) are being tested.
+
+Generally, by looking at the residuals we are concerned with
+
+$H_o:$ The model specification is correct
+
+against a broad alternative that something is missing:
+
+$H_a:$ The model specification is not correct
+
+The conventional tests are only equipped to measure specific departures, and thus have a more specific hypotheses being tested:
+
+XXX
+
+## $p$-value
 

### 2023-03-01 paper_comparison/paper_comparison/paper_comparison.tex
subsection on hypotheses

@@
 
-\hypertarget{estimation-of-p-value-and-power}{%
-\section{Estimation of p-value and
-power}\label{estimation-of-p-value-and-power}}
+\hypertarget{calculation-of-p-value-and-power}{%
+\section{\texorpdfstring{Calculation of \(p\)-value and
+power}{Calculation of p-value and power}}\label{calculation-of-p-value-and-power}}
+
+\hypertarget{hypotheses}{%
+\subsection{Hypotheses}\label{hypotheses}}
+
+A paragraph on what hypothesis(es) are being tested.
+
+Generally, by looking at the residuals we are concerned with
+
+\(H_o:\) The model specification is correct
+
+against a broad alternative that something is missing:
+
+\(H_a:\) The model specification is not correct
+
+The conventional tests are only equipped to measure specific departures,
+and thus have a more specific hypotheses being tested:
+
+XXX
 
 \hypertarget{p-value}{%
-\subsection{p-value}\label{p-value}}
+\subsection{\texorpdfstring{\(p\)-value}{p-value}}\label{p-value}}
 
@@
 the non-linearity pattern. According to Figure
-\ref{fig:poly-power-uniform-j}, conventional tests have higher power in
-testing shapes constructed from lower orders of Hermite polynomials. But
-the power in testing the ``Triple-U'' shape drops significantly. To
-understand why this is, one needs to return to the way the RESET test is
-applied. It requires a parameter indicating degree of fitted values to
-test for, and the recommendation is to generically use four
-(\citet{ramsey_tests_1969}). However, the ``Triple-U'' shape constructed
-from the Hermite polynomials use power up to 18. If the RESET test had
-been applied using a higher power no less than six, the power curve of
-``Triple-U'' shape will only slight lower than the power curve of ``M''
-shape. The recommendation of the polynomial power for the RESET should
-be revised, perhaps. This illustrates the sensitivity of conventional
-testing to the parameters, and it also points to a limitation that one
-needs to know the data generating process in order to set the parameters
-for the test.
+\ref{fig:poly-power-uniform-j}, conventional tests have roughly the same
+power in testing shapes constructed from low orders of Hermite
+polynomials. But the power in testing the ``Triple-U'' shape drops
+significantly. To understand why this is, one needs to return to the way
+the RESET test is applied. It requires a parameter indicating degree of
+fitted values to test for, and the recommendation is to generically use
+four (\citet{ramsey_tests_1969}). However, the ``Triple-U'' shape
+constructed from the Hermite polynomials use power up to 18. If the
+RESET test had been applied using a higher power no less than six, the
+power curve of ``Triple-U'' shape will only slight lower than the power
+curve of ``M'' shape. The recommendation of the polynomial power for the
+RESET should be revised, perhaps. This illustrates the sensitivity of
+conventional testing to the parameters, and it also points to a
+limitation that one needs to know the data generating process in order
+to set the parameters for the test.
 
@@
 curves are mostly aligned with our expectation, except for the ``M''
-shape, which is as easy to be detected as the ``S'' shape. This implies,
-unlike the conventional test, the readability of the shape do not
-strictly follow the degree of the polynomials.
+shape, which is the easiest shape to be detected. This implies the
+visibility of the shape do not strictly follow the degree of the
+polynomials.
 

### 2023-03-01 paper_comparison/paper_comparison/paper_comparison.Rmd
typo
rollback to previous figures

@@
   scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
-  scale_x_continuous(breaks = seq(0, 3, 0.5), minor_breaks = NULL) +
+  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
   xlab(expression(log[e] (Effect_size))) +
@@
 
-target_sigma <- c(5, 3.03, 1.84, 1.12, 0.68, 0.41, 0.25)
+target_sigma <- c(3.65, 2.84, 2.21, 1.72, 1.34, 1.04, 0.81, 0.63, 0.49, 0.38, 0.3)
 
@@
 es <- map_dbl(target_sigma, 
-              function(.x){
-                mod <- poly_model(shape = 2, 
+              ~log(poly_model(shape = 2, 
                               x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, 
-                              sigma = .x)
-                log(mod$sample_effect_size(mod$gen(100), type = "simple"))
-                }
+                              sigma = .x)$
+                     average_effect_size(n = 100, type = "kl")
+                   )
               )
@@
 Where would you draw the line of too much non-linearity in the residuals?
-For the RESET test this is around log effect size 1.5, but for the visual test it is around 2.", dev = 'png', dpi = 450}
+For the RESET test this is around log effect size 2.5, but for the visual test it is around 3.5.", dev = 'png', dpi = 450}
 
@@
 # Use area() to design layout
-map(1:7, ~patchwork::area(21, .x, 22, .x)) %>%
+map(1:11, ~patchwork::area(21, .x, 22, .x)) %>%
   reduce(c) %>%
-  c(patchwork::area(3, 1, 20, 7), .) %>%
+  c(patchwork::area(3, 1, 20, 11), .) %>%
   patchwork::wrap_plots(append(plist, list(p), after = 0), 
@@
 # The prediction of the visual test
-pred_visual_heter <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 1),
+pred_visual_heter <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
                                 offset0 = log(0.05/0.95)) %>%
@@
     mutate(power = map(mod, function(mod) {
-      result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 1),
+      result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
                            offset0 = log(0.05/0.95))
@@
   # scale_size_manual(values = seq(1, 9)) +
-  scale_x_continuous(breaks = seq(1, 7, 1), minor_breaks = NULL) +
+  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
   xlab(expression(log[e] (Effect_size))) +
@@
 
-target_b <- c(0.27, 0.74, 2.01, 5.46, 14.84, 40.34, 109.66)
+target_b <- c(0.17, 0.23, 0.32, 0.46, 0.67, 1.04, 1.7, 3, 6.3, 16.2, 61)
 
@@
 es <- map_dbl(target_b, 
-              function(.x){
-                mod <- heter_model(a = 0, 
+              ~log(heter_model(a = 1, 
                                b = .x, 
-                                   x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))})
-                log(mod$sample_effect_size(mod$gen(100), type = "simple"))
-                }
+                               x = {raw_x <- rand_normal(sigma = 0.3);closed_form(~stand_dist(raw_x))})$
+                     average_effect_size(n = 100, type = "kl")
+                   )
               )
@@
 Where would you draw the line of too much heteroskedasticity in the residuals?
-For the BP test this is around log effect size 3, but for the visual test it is around 5.", dev = 'png', dpi = 450}
+For the BP test this is around log effect size 2.5, but for the visual test it is around 3.5.", dev = 'png', dpi = 450}
 
@@
 # 133
-set.seed(112)
+set.seed(135)
 
@@
 # Use area() to design layout
-map(1:7, ~patchwork::area(21, .x, 22, .x)) %>%
+map(1:11, ~patchwork::area(21, .x, 22, .x)) %>%
   reduce(c) %>%
-  c(patchwork::area(3, 1, 20, 7), .) %>%
+  c(patchwork::area(3, 1, 20, 11), .) %>%
 patchwork::wrap_plots(append(plist, list(p), after = 0), 
@@
 Power curves with higher order of non-linearity are drawn with deeper colours.
-The default RESET tests under-perform significantly in detecting "Tripe-U" shape.
+The default RESET tests under-perform significantly in detecting "Triple-U" shape.
 To achieve a similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test.
@@
   mutate(power = map(mod, function(mod) {
-    result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 1),
+    result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
                          offset0 = log(0.05/0.95))

### 2023-03-01 paper_comparison/paper_comparison/paper_comparison.tex
typo
rollback to previous figures

@@
 Where would you draw the line of too much non-linearity in the residuals?
-For the RESET test this is around log effect size 1.5, but for the visual test it is around 2.}\label{fig:polypower}
+For the RESET test this is around log effect size 2.5, but for the visual test it is around 3.5.}\label{fig:polypower}
 \end{figure}
@@
 Where would you draw the line of too much heteroskedasticity in the residuals?
-For the BP test this is around log effect size 3, but for the visual test it is around 5.}\label{fig:heterpower}
+For the BP test this is around log effect size 2.5, but for the visual test it is around 3.5.}\label{fig:heterpower}
 \end{figure}

### 2023-03-01 paper_comparison/paper_comparison/paper_comparison.Rmd
tweak the points in power cmp plot

@@
 
-  # Group the data by effect and reject to draw points
-  geom_point(data = visual_poly_dat %>%
-               mutate(log_effect_size = log(effect_size)) %>%
-               group_by(log_effect_size, reject) %>% 
-               summarise(num_response = n()),
-             aes(log_effect_size, reject, size = num_response), 
-             alpha = 0.3) +
+  geom_point(data = visual_poly_dat,
+             aes(log(effect_size), reject),
+             alpha = 0.15) +
 
@@
 
-  # Group the data by effect and reject to draw points
-  geom_point(data = visual_heter_dat %>%
-               mutate(log_effect_size = log(effect_size)) %>%
-               group_by(log_effect_size, reject) %>% 
-               summarise(num_response = n()),
-             aes(log_effect_size, reject, size = num_response), 
-             alpha = 0.3) +
+  geom_point(data = visual_heter_dat,
+             aes(log(effect_size), reject),
+             alpha = 0.15) +
 

### 2023-03-03 paper_comparison/paper_comparison/paper_comparison.Rmd
add hypotheses discussion

@@
 
-## Hypotheses
-
-A paragraph on what hypothesis(es) are being tested.
-
-Generally, by looking at the residuals we are concerned with
-
-$H_o:$ The model specification is correct
-
-against a broad alternative that something is missing:
-
-$H_a:$ The model specification is not correct
-
-The conventional tests are only equipped to measure specific departures, and thus have a more specific hypotheses being tested:
-
-XXX
+## hypotheses
+
+Before the discussion of the p-value and power, the difference in hypotheses between conventional tests and visual tests worth a clarification.
+
+Generally, by examining the residuals we are concerned with
+
+$$H_0: \text{The regression model is correctly specified}$$
+
+against a broad alternative hypothesis that
+
+$$H_1: \text{The regression model is misspecified.}$$
+
+The conventional tests are only equipped to measure specific residual departures, and thus have a more specific null and alternative hypotheses being tested.
+For instance, the RESET test with the auxiliary regression formulated as $$\boldsymbol{e} = \gamma_1\hat{\boldsymbol{y}}^2 + \gamma_2\hat{\boldsymbol{y}}^3 + \gamma_3\hat{\boldsymbol{y}}^4 + \boldsymbol{u}, \boldsymbol{u} \sim N(\boldsymbol{0}, \sigma_u^2\boldsymbol{I})$$ has $H_0:\gamma_1 = \gamma_2 = \gamma_3 = 0$ and $H_1: \gamma_1 \neq 0 \text{ or } \gamma_2 \neq 0 \text{ or } \gamma_3 \neq 0$.
+In contrast, the visual test has the null hypothesis same as the general $H_0$ and $H_1$ provided above.
+Essentially, a visual test can be seen as a collection of tests targeting for different residual departures.
 

### 2023-03-03 paper_comparison/paper_comparison/paper_comparison.tex
add hypotheses discussion

@@
 \hypertarget{hypotheses}{%
-\subsection{Hypotheses}\label{hypotheses}}
-
-A paragraph on what hypothesis(es) are being tested.
-
-Generally, by looking at the residuals we are concerned with
-
-\(H_o:\) The model specification is correct
-
-against a broad alternative that something is missing:
-
-\(H_a:\) The model specification is not correct
-
-The conventional tests are only equipped to measure specific departures,
-and thus have a more specific hypotheses being tested:
-
-XXX
+\subsection{hypotheses}\label{hypotheses}}
+
+Before the discussion of the p-value and power, the difference in
+hypotheses between conventional tests and visual tests worth a
+clarification.
+
+Generally, by examining the residuals we are concerned with
+
+\[H_0: \text{The regression model is correctly specified}\]
+
+against a broad alternative hypothesis that
+
+\[H_1: \text{The regression model is misspecified.}\]
+
+The conventional tests are only equipped to measure specific residual
+departures, and thus have a more specific null and alternative
+hypotheses being tested. For instance, the RESET test with the auxiliary
+regression formulated as
+\[\boldsymbol{e} = \gamma_1\hat{\boldsymbol{y}}^2 + \gamma_2\hat{\boldsymbol{y}}^3 + \gamma_3\hat{\boldsymbol{y}}^4 + \boldsymbol{u}, \boldsymbol{u} \sim N(\boldsymbol{0}, \sigma_u^2\boldsymbol{I})\]
+has \(H_0:\gamma_1 = \gamma_2 = \gamma_3 = 0\) and
+\(H_1: \gamma_1 \neq 0 \text{ or } \gamma_2 \neq 0 \text{ or } \gamma_3 \neq 0\).
+In contrast, the visual test has the null hypothesis same as the general
+\(H_0\) and \(H_1\) provided above. Essentially, a visual test can be
+seen as a collection of tests targeting for different residual
+departures.
 

### 2023-03-05 paper_comparison/paper_comparison/paper_comparison.Rmd
hypotheses and p-value and power section modified

--- 2023-03-01 paper_comparison/paper_comparison/paper_comparison.Rmd
@@
 
-# Calculation of $p$-value and power
-
-## Hypotheses
-
-A paragraph on what hypothesis(es) are being tested.
-
-Generally, by looking at the residuals we are concerned with
-
-$H_o:$ The model specification is correct
-
-against a broad alternative that something is missing:
-
-$H_a:$ The model specification is not correct
-
-The conventional tests are only equipped to measure specific departures, and thus have a more specific hypotheses being tested:
-
-XXX
-
-## $p$-value
+# Calculation of statistical significance and test power
+
+## What is being tested?
+
+In diagnosing a model fit from residuals, we are generally interested in $H_0:$ *The regression model is correctly specified* against the broad alternative $H_a:$ *The regression model is misspecified*.
+
+However, it is practically impossible to test this specific $H_0$ with conventional tests, which are constructed to measure specific departures.
+For example, the RESET test is formulated as $H_0:\gamma_1 = \gamma_2 = \gamma_3 = 0$ against $H_a: \gamma_1 \neq 0 \text{ or } \gamma_2 \neq 0 \text{ or } \gamma_3 \neq 0$, from $e = \gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)$.
+Similarly, the BP test is designed to specifically test $H_0:$ *error variances are all equal* ($\tau=0$) versus the alternative $H_a:$ *that the error variances are a multiplicative function of one or more variables* ($\tau\neq 0$) from $e^2 = \tau\hat{y} + u, ~ u\sim N(0,\sigma_u^2)$.
+
+One of the potential benefits of the visual test, based on the lineup protocol, is that it should be able to detect a range of departures from good residuals.
+
+## Statistical significance
 
@@
 
-## Power
+## Power of the tests
 

### 2023-03-05 paper_comparison/paper_comparison/paper_comparison.tex
hypotheses and p-value and power section modified

--- 2023-03-01 paper_comparison/paper_comparison/paper_comparison.tex
@@
 
-\hypertarget{calculation-of-p-value-and-power}{%
-\section{\texorpdfstring{Calculation of \(p\)-value and
-power}{Calculation of p-value and power}}\label{calculation-of-p-value-and-power}}
-
-\hypertarget{hypotheses}{%
-\subsection{Hypotheses}\label{hypotheses}}
-
-A paragraph on what hypothesis(es) are being tested.
-
-Generally, by looking at the residuals we are concerned with
-
-\(H_o:\) The model specification is correct
-
-against a broad alternative that something is missing:
-
-\(H_a:\) The model specification is not correct
-
-The conventional tests are only equipped to measure specific departures,
-and thus have a more specific hypotheses being tested:
-
-XXX
-
-\hypertarget{p-value}{%
-\subsection{\texorpdfstring{\(p\)-value}{p-value}}\label{p-value}}
+\hypertarget{calculation-of-statistical-significance-and-test-power}{%
+\section{Calculation of statistical significance and test
+power}\label{calculation-of-statistical-significance-and-test-power}}
+
+\hypertarget{what-is-being-tested}{%
+\subsection{What is being tested?}\label{what-is-being-tested}}
+
+In diagnosing a model fit from residuals, we are generally interested in
+\(H_0:\) \emph{The regression model is correctly specified} against the
+broad alternative \(H_a:\) \emph{The regression model is misspecified}.
+
+However, it is practically impossible to test this specific \(H_0\) with
+conventional tests, which are constructed to measure specific
+departures. For example, the RESET test is formulated as
+\(H_0:\gamma_1 = \gamma_2 = \gamma_3 = 0\) against
+\(H_a: \gamma_1 \neq 0 \text{ or } \gamma_2 \neq 0 \text{ or } \gamma_3 \neq 0\),
+from
+\(e = \gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)\).
+Similarly, the BP test is designed to specifically test \(H_0:\)
+\emph{error variances are all equal} (\(\tau=0\)) versus the alternative
+\(H_a:\) \emph{that the error variances are a multiplicative function of
+one or more variables} (\(\tau\neq 0\)) from
+\(e^2 = \tau\hat{y} + u, ~ u\sim N(0,\sigma_u^2)\).
+
+One of the potential benefits of the visual test, based on the lineup
+protocol, is that it should be able to detect a range of departures from
+good residuals.
+
+\hypertarget{statistical-significance}{%
+\subsection{Statistical significance}\label{statistical-significance}}
 
@@
 
-\hypertarget{power}{%
-\subsection{Power}\label{power}}
+\hypertarget{power-of-the-tests}{%
+\subsection{Power of the tests}\label{power-of-the-tests}}
 

### 2023-03-06 paper_comparison/paper_comparison/paper_comparison.Rmd
fix the formula in the "what is being tested" section
After checking the packages and the original paper, I think it now looks closer to the correct form.

@@
 ```{r}
-# rewrite effect size here (if needed)
-
-```
-
-```{r}
 # Borrow effect size from the survey
@@
 However, it is practically impossible to test this specific $H_0$ with conventional tests, which are constructed to measure specific departures.
-For example, the RESET test is formulated as $H_0:\gamma_1 = \gamma_2 = \gamma_3 = 0$ against $H_a: \gamma_1 \neq 0 \text{ or } \gamma_2 \neq 0 \text{ or } \gamma_3 \neq 0$, from $e = \gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)$.
-Similarly, the BP test is designed to specifically test $H_0:$ *error variances are all equal* ($\tau=0$) versus the alternative $H_a:$ *that the error variances are a multiplicative function of one or more variables* ($\tau\neq 0$) from $e^2 = \tau\hat{y} + u, ~ u\sim N(0,\sigma_u^2)$.
+For example, the RESET test is formulated as $H_0:\gamma_1 = \gamma_2 = \gamma_3 = 0$ against $H_a: \gamma_1 \neq 0 \text{ or } \gamma_2 \neq 0 \text{ or } \gamma_3 \neq 0$, from $e = \tau_0 + \tau_1x_1 + \cdots + \tau_1 x_p +\gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)$.
+Similarly, the Breusch-Pagan test is designed to specifically test $H_0:$ *error variances are all equal* ($\tau=0$) versus the alternative $H_a:$ *that the error variances are a multiplicative function of one or more variables* ($\tau\neq 0$) from $e^2 = \tau_0 + \tau_1 x_1 + \cdots + \tau_1 x_p + u, ~ u\sim N(0,\sigma_u^2)$.
 

### 2023-03-06 paper_comparison/paper_comparison/paper_comparison.Rmd
Correct the null hypothesis for the BP test
This null hypothesis is copied from the original paper

--- 2023-03-05 paper_comparison/paper_comparison/paper_comparison.Rmd
@@
 However, it is practically impossible to test this specific $H_0$ with conventional tests, which are constructed to measure specific departures.
-For example, the RESET test is formulated as $H_0:\gamma_1 = \gamma_2 = \gamma_3 = 0$ against $H_a: \gamma_1 \neq 0 \text{ or } \gamma_2 \neq 0 \text{ or } \gamma_3 \neq 0$, from $e = \gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)$.
-Similarly, the BP test is designed to specifically test $H_0:$ *error variances are all equal* ($\tau=0$) versus the alternative $H_a:$ *that the error variances are a multiplicative function of one or more variables* ($\tau\neq 0$) from $e^2 = \tau\hat{y} + u, ~ u\sim N(0,\sigma_u^2)$.
+For example, the RESET test is formulated as $H_0:\gamma_1 = \gamma_2 = \gamma_3 = 0$ against $H_a: \gamma_1 \neq 0 \text{ or } \gamma_2 \neq 0 \text{ or } \gamma_3 \neq 0$, from $e = \tau_0 + \tau_1x_1 + ...
++ \tau_1 x_p +\gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)$.
+Similarly, the Breusch-Pagan test is designed to specifically test $H_0:$ *error variances are all equal* ($\tau_i=0 \text{ for } i=1,..,p$) versus the alternative $H_a:$ *that the error variances are a multiplicative function of one or more variables* ($\text{At least one } \tau_i\neq 0$) from $e^2 = \tau_0 + \tau_1 x_1 + \cdots + \tau_1 x_p + u, ~ u\sim N(0,\sigma_u^2)$.
 

### 2023-03-06 paper_comparison/paper_comparison/paper_comparison.Rmd
fix the power calc desc

@@
       # Every sample contains 2000 lineups
-      map(1:200, function(i) {
+      map(1:2000, function(i) {
 
@@
       # Every sample contains 2000 lineups
-      map(1:200, function(x) {
+      map(1:2000, function(x) {
 
@@
 
+(title refine - rank: low)
+
 # Introduction
+
+(introduction refine - rank: low)
 
@@
 Later, @buja_statistical_2009 proposed the lineup protocol as a visual test inspired by the "police lineup" or "identity parade" which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people.
-The protocol consists of $m$ randomly placed plots, where one plot is the data plot, and the remaining $m - 1$ null plots have the identical graphical procedure except the data has been replaced with data consistent with $H_0$.
+The protocol consists of $m$ randomly placed plots, where one plot is the data plot, and the remaining $m - 1$ null plots have the identical graphical procedure except the data has been replaced with data consistent with $H_0$, that the model is correctly specified.
 Then, an observer who have not seen the data plot will be asked to point out the most different plot from the lineup.
@@
 Figure \ref{fig:first-example-lineup} is an example of a lineup protocol.
-If the data plot at position $2^2 + 2$ is identifiable, then it is evidence for the rejection of $H_0$ that the regression model is correctly specified.
-In fact, the actual residual plot is obtained from a misspecified regression model with non-linearity defect.
+If the data plot at position $2^2 + 2$ is identifiable, then it is evidence for the rejection of $H_0$.
+In fact, the actual residual plot is obtained from a misspecified regression model with missing non-linear terms.
 
@@
 Their results suggest that visual tests are capable of testing the significance of a single regressor with a similar power as a t-test, though they express that in general it is unnecessary to use visual inference if there exists a conventional test, and they do not expect the visual test to perform equally well as the conventional test.
-(contradict to our claim?
-expand?) In their third experiment, where there is not a conventional test, visual test outperforms the conventional test for a large margin.
+In their third experiment, where there is not a conventional test, visual test outperforms the conventional test for a large margin.
 This is encouraging, as it promotes the use of visual inference in situations where there are no existing statistical testing procedures.
@@
 + \tau_1 x_p +\gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)$.
-Similarly, the Breusch-Pagan test is designed to specifically test $H_0:$ *error variances are all equal* ($\tau_i=0 \text{ for } i=1,..,p$) versus the alternative $H_a:$ *that the error variances are a multiplicative function of one or more variables* ($\text{At least one } \tau_i\neq 0$) from $e^2 = \tau_0 + \tau_1 x_1 + \cdots + \tau_1 x_p + u, ~ u\sim N(0,\sigma_u^2)$.
+Similarly, the Breusch-Pagan test is designed to specifically test $H_0:$ *error variances are all equal* ($\tau_i=0 \text{ for } i=1,..,p$) versus the alternative $H_a:$ *that the error variances are a multiplicative function of one or more variables* ($\text{at least one } \tau_i\neq 0$) from $e^2 = \tau_0 + \tau_1 x_1 + \cdots + \tau_1 x_p + u, ~ u\sim N(0,\sigma_u^2)$.
 
@@
 Details about the effect size is provided in Section \ref{effect-size} after the introduction of the simulation model used in our human subject experiment.
+
 The theoretical power of a test is sometimes not a trival solution.

### 2023-03-06 paper_comparison/paper_comparison/paper_comparison.Rmd
finish power desc

@@
 Example residual plots given in Figure \ref{fig:residual-plot-common-departures} are examined by the corresponding RESET test, Breusch-Pagan test and Shapiro-Wilk test as shown in Table \ref{tab:example-residual-plot-table}.
-In the example, the Breusch-Pagan test and the Shapiro--Wilk test both reject the null hypothesis $H_0$ for departures that they do not intend to examine.
+In the example, the Breusch-Pagan test and the Shapiro--Wilk test both reject $H_0$ for departures that they do not intend to examine.
 As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are also sensitive to other types of departures.
@@
 
-The power of a model misspecification test is the probability that the null hypothesis is rejected given the regression model is misspecified in a specific way. It is an important indicator
+The power of a model misspecification test is the probability that $H_0$ is rejected given the regression model is misspecified in a specific way. It is an important indicator
 when one is concerned about whether model assumptions have been violated.
@@
 Details about the effect size is provided in Section \ref{effect-size} after the introduction of the simulation model used in our human subject experiment.
-
-The theoretical power of a test is sometimes not a trival solution.
-For a given regression model under a specific effect size, if the data generating process is also known, the power of a conventional test can be estimated by counting the number of times the test rejects the null hypothesis in a repeated sampling exercise
-and divided by the number of times simulated from the known data generating process. This is the method we used to estimate the power of conventional tests in our experiment.
-
-The power of a visual test on the other hand, may additionally depend on the ability of the particular subject, as the skill of the individual may affect the number of observers who identify the data plot from the lineup [@majumder_validation_2013].
-To address this issue, @majumder_validation_2013 models the probability of a subject $j$ correctly picking the data plot from a lineup $l$ using a mixed-effect logistic regression, with subjects treated as random effect.
-Then, the estimated power of a visual test evaluated by a single subject is the predicted value obtained from the mix-effect model.
-
-However, this mix-effect model does not work with scenario where subjects are asked to select one or more "most different" plots.
-In this scenario, having the probability of a subject $j$ correctly picking the data plot from a lineup $l$ is insufficient to determine the power of a visual test because it does not provide information about the number of selections made by the subject for the calculation of the p-value (See Equation \ref{eq:pvalue-beta-binomial-approx}).
-
-Therefore, we directly estimate the probability of a lineup being rejected by assuming that individual skill has negligible effect on the variation of the power.
-This assumption is not necessary true, but it helps simplifying the model structure, thereby obviate a costly large-scale experiment to estimate complex covariance matrices.
-There is a discussion provided in the Appendix about a potential solution to model individual effect in the multiple selections scenario.
-
-(should we just link the power together for conventional test?)
-
-We use the same logistic regression model to fit the power of conventional tests and visual tests formulated as
+The theoretical power of a test is sometimes not a trival solution, but it can be estimated if the data generating process is known.
+We use a predefined model to generate a large set of simulated data under different effect sizes, and record if the conventional test rejects $H_0$.
+The probability of the conventional test rejects $H_0$ is then fitted by a logistic regression formulated as
 
@@
 ```
+
 \noindent where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
-The effect size $E$ is the only regressor and the intercept is fixed to $log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$.
-
-To study various factors contributing to the power of test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors.
+The effect size $E$ is the only regressor and the intercept is fixed to $log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$, which is the desired significance level.
+
+The power of a visual test on the other hand, may additionally depend on the ability of the particular subject, as the skill of the individual may affect the number of observers who identify the data plot from the lineup [@majumder_validation_2013].
+To address this issue, @majumder_validation_2013 models the probability of a subject $j$ correctly picking the data plot from a lineup $l$ using a mixed-effect logistic regression, with subjects treated as random effect.
+Then, the estimated power of a visual test evaluated by a single subject is the predicted value obtained from the mix-effect model.
+However, this mix-effect model does not work with scenario where subjects are asked to select one or more most different plots.
+In this scenario, having the probability of a subject $j$ correctly picking the data plot from a lineup $l$ is insufficient to determine the power of a visual test because it does not provide information about the number of selections made by the subject for the calculation of the p-value (See Equation \ref{eq:pvalue-beta-binomial-approx}).
+Therefore, we directly estimate the probability of a lineup being rejected by assuming that individual skill has negligible effect on the variation of the power.
+This assumption is not necessary true, but it helps simplifying the model structure, thereby obviate a costly large-scale experiment to estimate complex covariance matrices.
+The same model given in Equation \ref{eq:logistic-regression-1-1} is applied to model the power of a visual test.
+
+To study various factors contributing to the power of both tests, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors.
 These include the distribution of the regressor, type of the simulation model and the shape of the residual departures.
@@
 In these lineups, the data of the data plot is generated from a model with zero effect size, while the data of the 19 null plots are generated using the same simulation method discussed in Section \ref{sampling-from-the-null-distribution}.
-This generation procedure differs from the canonical Rorschach lineup procedure, which requires that all 20 plots are generated from the null hypothesis.
-However, these lineups serve the same fundamental purpose: to assess the number of visually interesting plots generated under the null hypothesis.
+This generation procedure differs from the canonical Rorschach lineup procedure, which requires that all 20 plots are generated from $H_0$.
+However, these lineups serve the same fundamental purpose: to assess the number of visually interesting plots generated under $H_0$.
 
@@
 For both the non-linearity pattern and the heteroskedasticity pattern, the power curves of appropriate conventional tests are constantly greater than the power curves of visual tests and other inappropriate conventional tests.
-At $log(E) = 1$ for non-linearity model, where the residual departures are unnoticeable from the corresponding residual plot, the RESET test has above 25% of chance of rejecting the null hypothesis, while only a few visual tests would result in a rejection.
+At $log(E) = 1$ for non-linearity model, where the residual departures are unnoticeable from the corresponding residual plot, the RESET test has above 25% of chance of rejecting $H_0$, while only a few visual tests would result in a rejection.
 Similarly, at $log(E) = 3$ for heteroskedasticity model, the BP test has around 25% power but the non-constant variance visual pattern is almost unobserved from the residual plot and the power of the visual test is close to the significance level at this point.

### 2023-03-06 paper_comparison/paper_comparison/paper_comparison.Rmd
fix notation and explanation in exp design

@@
 ```{r poly-conventional-simulation}
-# Ensure the support of the regressor is [-1, 1]
+# Ensure the support of the predictor is [-1, 1]
 stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
@@
 
+(fix the tense issue - rank: low)
+
+(fix the notation issue (bold?) - rank: low)
+
 # Introduction
@@
 
-Graphical summaries in which residuals are plotted against fitted values or other functions of the regressors that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals.
+Graphical summaries in which residuals are plotted against fitted values or other functions of the predictors that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals.
 The first panel of Figure \ref{fig:residual-plot-common-departures} shows an ideal residual plot with residuals evenly distributed at both sides of the horizontal zero line, with no noticeable patterns.
@@
 
-Non-linearity is a type of model misspecification caused by failing to include higher order terms of the regressors in the regression equation.
+Non-linearity is a type of model misspecification caused by failing to include higher order terms of the predictors in the regression equation.
 Any non-linear functional form of residuals on fitted values in the residual plot could be indicative of non-linearity.
@@
 For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution.
-For testing non-linearity, one may apply the F-test as a model structural test to examine the significance of specific polynomial and non-linear forms of the regressors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969].
+For testing non-linearity, one may apply the F-test as a model structural test to examine the significance of specific polynomial and non-linear forms of the predictors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969].
 The Shapiro-Wilk test [@shapiro1965analysis] is the most widely used test of non-normality included by many of the statistical softwares.
@@
 The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009].
-Essentially, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the regressors, then rescaling it by the ratio of residual sum of square in two regressions.
-
-The effectiveness of lineup protocol for regression analysis is validated by @majumder_validation_2013 under relatively simple settings with up to two regressors.
-Their results suggest that visual tests are capable of testing the significance of a single regressor with a similar power as a t-test, though they express that in general it is unnecessary to use visual inference if there exists a conventional test, and they do not expect the visual test to perform equally well as the conventional test.
+Essentially, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the predictors, then rescaling it by the ratio of residual sum of square in two regressions.
+
+The effectiveness of lineup protocol for regression analysis is validated by @majumder_validation_2013 under relatively simple settings with up to two predictors.
+Their results suggest that visual tests are capable of testing the significance of a single predictor with a similar power as a t-test, though they express that in general it is unnecessary to use visual inference if there exists a conventional test, and they do not expect the visual test to perform equally well as the conventional test.
 In their third experiment, where there is not a conventional test, visual test outperforms the conventional test for a large margin.
@@
 \noindent where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
-The effect size $E$ is the only regressor and the intercept is fixed to $log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$, which is the desired significance level.
+The effect size $E$ is the only predictor and the intercept is fixed to $log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$, which is the desired significance level.
 
@@
 To study various factors contributing to the power of both tests, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors.
-These include the distribution of the regressor, type of the simulation model and the shape of the residual departures.
+These include the distribution of the fitted values, type of the simulation model and the shape of the residual departures.
 
@@
 
-To investigate the difference between conventional hypothesis testing and visual inference in the application of linear regression diagnostics, an experiment is conducted in three data collection periods.
-Two types of departures, non-linearity and heteroskedasticity, are primarily collected during data collection periods I and II.
+An experiment is conducted in three data collection periods to investigate the difference between conventional hypothesis testing and visual inference in the application of linear regression diagnostics.
+Two types of departures, non-linearity and heteroskedasticity, are collected during data collection periods I and II.
 The data collection period III was designed primarily to measure human responses to null lineups so that the parameter $\alpha$ in Equation \ref{eq:pvalue-beta-binomial} can be estimated.
-Additional lineups for both non-linearity and heteroskedasticity, using uniform $X$ distribution, were included so that the participants were evaluating some lineups with signal also.
+Additional lineups for both non-linearity and heteroskedasticity, using uniform fitted value distribution, were included so that the participants were evaluating some lineups with signal also.
 It would be too frustrating for participants to only be assigned lineups with all null plots.
@@
   gen(300, computed = select(dat_dist_1, e)) %>%
-  mutate(x_dist = "Discrete")
+  mutate(x_dist = "Discrete uniform")
 
@@
 Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups used in non-linearity detection.
-This lineup is produced by the non-linearity model with $j = 6$ and uniform fitted values.
+This lineup is produced by the non-linearity model with $j = 6$.
 The data plot location is $2^3 - 4$.
@@
 
-```{r different-n, fig.height = 2.67, fig.cap = "Examining the effect of signal strength for the three different values of $n$ used in the experiment, for non-linear structure with fixed $\\sigma = 1.5$, uniform $X$ distribution, and S shape.
-For these factor levels, only when $n=300$ is the S shape clearly visible."}
+```{r different-n, fig.height = 2.67, fig.cap = "Examining the effect of signal strength for the three different values of $n$ used in the experiment, for non-linear structure with fixed $\\sigma = 1$, uniform fitted value distribution, and S shape.
+For these factor levels, only when $n = 300$ is the S shape clearly visible."}
 
@@
                         x = {
-                          raw_x <- rand_uniform(-1, 1);
+                          raw_x <- rand_uniform(-1, 1)
                           closed_form(~stand_dist(raw_x))
                         },
-                        sigma = 1.5)$gen(300) %>%
+                        sigma = 1)$gen(300) %>%
   mutate(n = "N: 300")
@@
 
-```{r different-sigma, fig.height = 2, fig.cap = "Examining the effect of $\\sigma$ on the signal strength in the non-linearity detection, for $n=300$, uniform $X$ distribution and the U shape.
+```{r different-sigma, fig.height = 2, fig.cap = "Examining the effect of $\\sigma$ on the signal strength in the non-linearity detection, for $n=300$, uniform fitted value distribution and the U shape.
 As $\\sigma$ increases the signal strength decreases, to the point that the U is almost unrecognisable when $\\sigma=4$."}
@@
 
-For $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the regressor $\boldsymbol{x}$, which will lead to the presence of heteroskedasticity.
+For $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the predictor $\boldsymbol{x}$, which will lead to the presence of heteroskedasticity.
 Visual patterns of heteroskedasticity are simulated using three different shapes ($a$ = -1, 0, 1).
@@
 
-An example lineup of this model used in data collection period II is shown in Figure \ref{fig:example-heter-lineup} with $a = -1$ and $X_{raw} \sim U(-1, 1)$.
+An example lineup of this model used in data collection period II is shown in Figure \ref{fig:example-heter-lineup} with $a = -1$.
 The data plot location is $2^4 + 2$.
@@
 It is also common to see a skewed distribution of fitted values, if one or more of the predictors has a skewed distribution.
-This latter problem is usually corrected before modeling using a variable transformation.
-Our simulation asesses this by using four different distributions to represent the fitted values ($X_{raw}$): (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.
-We would expect that the best reading of residual plots occurs when the fitted values are uniformly distributed.
-
-<!--Different distributions of fitted values help enriching the pool of visual patterns as illustrated in Figure \ref{fig:different-dist}.
-The uniform and the normal distribution are symmetric and commonly assumed in statistical models.
-The adjusted log-normal distribution provides skewed density, while the discrete uniform distribution provides discreteness in residual plot.-->
+This latter problem is usually corrected before modelling using a variable transformation.
+Our simulation assess this by using four different distributions to represent fitted values: (1) uniform, (2) normal, (3) skewed and (4) discrete uniform.
+This is constructed by defining the raw predictor $X_{raw}$ in four corresponding distributions:
+(1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$. We would expect that the best reading of residual plots occurs when the fitted values are uniformly distributed.
 
@@
 We would expect considerable variation in the signal strength in the simulated data plots with smaller $n$.
-A sample size of 300 is typically enough for structure to be visible in a scatterplot reliably.
+A sample size of 300 is typically enough for structure to be visible in a scatter plot reliably.
 
@@
 A large value of $\sigma$ will increase the variation of the error of the non-linearity model and decrease the visibility of the visual pattern.
-The parameter $b$ controls the standard deviation of the error across the support of the regressor.
+The parameter $b$ controls the standard deviation of the error across the support of the predictor.
 Given $x \neq a$, a larger value of $b$ will lead to a larger ratio of the variance at $x$ to the variance at $x - a = 0$, making the visual pattern more obvious.
@@
 Thus, we follow this guideline to conduct the RESET test.
-For the Breusch-Pagan test, the choice of regressors in the auxiliary regression is left to the user [@breusch_simple_1979].
-But as @waldman1983note suggested, it is a good choice for the set of auxiliary regressors in the Breusch-Pagan test be the same as the White test.
+For the Breusch-Pagan test, the choice of predictors in the auxiliary regression is left to the user [@breusch_simple_1979].
+But as @waldman1983note suggested, it is a good choice for the set of auxiliary predictors in the Breusch-Pagan test be the same as the White test.
 Thus, we include both $\boldsymbol{x}$ and $\boldsymbol{x}^2$ in the auxiliary regression.

### 2023-03-06 paper_comparison/paper_comparison/paper_comparison.Rmd
update the effect size desc

@@
 
-(do we actually need the penalty term?)
-
 Effect size $E$ can be defined as the difference of a parameter for a particular model or distribution, or a statistic derived from a sample.
 Importantly, it needs to reflect the treatment we try to measure.
-According to the underlying simulated model, we define the effect size $E = \sqrt{n}/(\sigma k)$ for the non-linearity model and $E = b\sqrt{n}/(|a| + 1)$ for the heteroskedasticity model.
-The parameter $k$ is the number of turning points embedded in the residual plot.
-It is used as a penalty term to reduce the effect size for complex shapes.
-There are one, two, three and five turning points in "U", "S", "M", "Triple-U" shape respectively.
-The denominator of the effect size of the heteroskedasticity model is $|a| + 1$.
-It is used to
+According to the underlying simulated model, we define the effect size
+
+$$E = \frac{1}{2}\left(\boldsymbol{\mu}_z'[diag(\boldsymbol{R}_a\sigma^2)]^{-1}\boldsymbol{\mu}_z\right)$$
+
+for the non-linearity model, and
+
+$$E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)$$
+for the heteroskedasticity model, where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix, $\sigma^2\boldsymbol{I}$ is the assumed covariance matrix of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual covariance matrix of the error term, $\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a$ is the residual operator, $\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ is the hat matrix, $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of predictors including the intercept used in the regression equation, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients.
+
+Both equations above are derived from the Kullback-Leibler divergence (citation here?) and the details are provided in the appendix.
 

### 2023-03-06 paper_comparison/paper_comparison/paper_comparison.tex
update the effect size desc

@@
 
+(title refine - rank: low)
+
+(fix the tense issue - rank: low)
+
+(fix the notation issue (bold?) - rank: low)
+
 \hypertarget{introduction}{%
 \section{Introduction}\label{introduction}}
+
+(introduction refine - rank: low)
 
@@
 Graphical summaries in which residuals are plotted against fitted values
-or other functions of the regressors that are approximately orthogonal
+or other functions of the predictors that are approximately orthogonal
 to residuals are referred to as standard residual plots in
@@
 Non-linearity is a type of model misspecification caused by failing to
-include higher order terms of the regressors in the regression equation.
+include higher order terms of the predictors in the regression equation.
 Any non-linear functional form of residuals on fitted values in the
@@
 test to examine the significance of specific polynomial and non-linear
-forms of the regressors, or the significance of proxy variables as in
+forms of the predictors, or the significance of proxy variables as in
 the Ramsey Regression Equation Specification Error Test (RESET)
@@
 shown in Table \ref{tab:example-residual-plot-table}. In the example,
-the Breusch-Pagan test and the Shapiro--Wilk test both reject the null
-hypothesis \(H_0\) for departures that they do not intend to examine. As
-discussed in \citet{cook1982residuals}, most residual-based tests for a
-particular type of departure from model assumptions are also sensitive
-to other types of departures. It is likely \(H_0\) is correctly rejected
-but for the wrong reason, a phenomenon known as the ``Type III error''.
+the Breusch-Pagan test and the Shapiro--Wilk test both reject \(H_0\)
+for departures that they do not intend to examine. As discussed in
+\citet{cook1982residuals}, most residual-based tests for a particular
+type of departure from model assumptions are also sensitive to other
+types of departures. It is likely \(H_0\) is correctly rejected but for
+the wrong reason, a phenomenon known as the ``Type III error''.
 Additionally, outliers will often incorrectly trigger the rejection of
@@
 \(m - 1\) null plots have the identical graphical procedure except the
-data has been replaced with data consistent with \(H_0\). Then, an
-observer who have not seen the data plot will be asked to point out the
-most different plot from the lineup. Under \(H_0\), it is expected that
-the data plot would have no distinguishable difference from the null
-plots, and the probability that the observer correctly picks the data
-plot is \(1/m\). If one rejects \(H_0\) as the observer correctly picks
-the data plot, then the Type I error of this test is \(1/m\).
+data has been replaced with data consistent with \(H_0\), that the model
+is correctly specified. Then, an observer who have not seen the data
+plot will be asked to point out the most different plot from the lineup.
+Under \(H_0\), it is expected that the data plot would have no
+distinguishable difference from the null plots, and the probability that
+the observer correctly picks the data plot is \(1/m\). If one rejects
+\(H_0\) as the observer correctly picks the data plot, then the Type I
+error of this test is \(1/m\).
 
@@
 protocol. If the data plot at position \(2^2 + 2\) is identifiable, then
-it is evidence for the rejection of \(H_0\) that the regression model is
-correctly specified. In fact, the actual residual plot is obtained from
-a misspecified regression model with non-linearity defect.
+it is evidence for the rejection of \(H_0\). In fact, the actual
+residual plot is obtained from a misspecified regression model with
+missing non-linear terms.
 
@@
 simulated by regressing \(N\) i.i.d standard normal random draws on the
-regressors, then rescaling it by the ratio of residual sum of square in
+predictors, then rescaling it by the ratio of residual sum of square in
 two regressions.
@@
 validated by \citet{majumder_validation_2013} under relatively simple
-settings with up to two regressors. Their results suggest that visual
-tests are capable of testing the significance of a single regressor with
+settings with up to two predictors. Their results suggest that visual
+tests are capable of testing the significance of a single predictor with
 a similar power as a t-test, though they express that in general it is
@@
 and they do not expect the visual test to perform equally well as the
-conventional test. (contradict to our claim? expand?) In their third
-experiment, where there is not a conventional test, visual test
-outperforms the conventional test for a large margin. This is
-encouraging, as it promotes the use of visual inference in situations
-where there are no existing statistical testing procedures. Visual
-inference have also been integrated into diagnostic of hierarchical
-linear models by \citet{loy2013diagnostic}, \citet{loy2014hlmdiag} and
-\citet{loy2015you}. They use lineup protocols to judge the assumption of
-linearity, normality and constant error variance for both the level-1
-and level-2 residuals. (expand?)
+conventional test. In their third experiment, where there is not a
+conventional test, visual test outperforms the conventional test for a
+large margin. This is encouraging, as it promotes the use of visual
+inference in situations where there are no existing statistical testing
+procedures. Visual inference have also been integrated into diagnostic
+of hierarchical linear models by \citet{loy2013diagnostic},
+\citet{loy2014hlmdiag} and \citet{loy2015you}. They use lineup protocols
+to judge the assumption of linearity, normality and constant error
+variance for both the level-1 and level-2 residuals. (expand?)
 
@@
 from
-\(e = \gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)\).
-Similarly, the BP test is designed to specifically test \(H_0:\)
-\emph{error variances are all equal} (\(\tau=0\)) versus the alternative
-\(H_a:\) \emph{that the error variances are a multiplicative function of
-one or more variables} (\(\tau\neq 0\)) from
-\(e^2 = \tau\hat{y} + u, ~ u\sim N(0,\sigma_u^2)\).
+\(e = \tau_0 + \tau_1x_1 + ... + \tau_1 x_p +\gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)\).
+Similarly, the Breusch-Pagan test is designed to specifically test
+\(H_0:\) \emph{error variances are all equal}
+(\(\tau_i=0 \text{ for } i=1,..,p\)) versus the alternative \(H_a:\)
+\emph{that the error variances are a multiplicative function of one or
+more variables} (\(\text{at least one } \tau_i\neq 0\)) from
+\(e^2 = \tau_0 + \tau_1 x_1 + \cdots + \tau_1 x_p + u, ~ u\sim N(0,\sigma_u^2)\).
 
@@
 
-The power of a model misspecification test is the probability that the
-null hypothesis is rejected given the regression model is misspecified
-in a specific way. It is an important indicator when one is concerned
-about whether model assumptions have been violated. Although in
-practice, one might be more interested in knowing how much the residuals
-deviate from the model assumptions, and whether this deviation is of
-practical significance.
+The power of a model misspecification test is the probability that
+\(H_0\) is rejected given the regression model is misspecified in a
+specific way. It is an important indicator when one is concerned about
+whether model assumptions have been violated. Although in practice, one
+might be more interested in knowing how much the residuals deviate from
+the model assumptions, and whether this deviation is of practical
+significance.
 
@@
 experiment. The theoretical power of a test is sometimes not a trival
-solution. For a given regression model under a specific effect size, if
-the data generating process is also known, the power of a conventional
-test can be estimated by counting the number of times the test rejects
-the null hypothesis in a repeated sampling exercise and divided by the
-number of times simulated from the known data generating process. This
-is the method we used to estimate the power of conventional tests in our
-experiment.
+solution, but it can be estimated if the data generating process is
+known. We use a predefined model to generate a large set of simulated
+data under different effect sizes, and record if the conventional test
+rejects \(H_0\). The probability of the conventional test rejects
+\(H_0\) is then fitted by a logistic regression formulated as
+
+\begin{equation} \label{eq:logistic-regression-1-1}
+Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right),
+\end{equation}
+
+\noindent where \(\Lambda(.)\) is the standard logistic function given
+as \(\Lambda(z) = exp(z)/(1+exp(z))\). The effect size \(E\) is the only
+predictor and the intercept is fixed to \(log(0.05/0.95)\) so that
+\(\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05\), which is the desired
+significance level.
 
@@
 subject is the predicted value obtained from the mix-effect model.
-
 However, this mix-effect model does not work with scenario where
-subjects are asked to select one or more ``most different'' plots. In
-this scenario, having the probability of a subject \(j\) correctly
-picking the data plot from a lineup \(l\) is insufficient to determine
-the power of a visual test because it does not provide information about
-the number of selections made by the subject for the calculation of the
-p-value (See Equation \ref{eq:pvalue-beta-binomial-approx}).
-
-Therefore, we directly estimate the probability of a lineup being
-rejected by assuming that individual skill has negligible effect on the
-variation of the power. This assumption is not necessary true, but it
-helps simplifying the model structure, thereby obviate a costly
-large-scale experiment to estimate complex covariance matrices. There is
-a discussion provided in the Appendix about a potential solution to
-model individual effect in the multiple selections scenario.
-
-(should we just link the power together for conventional test?)
-
-We use the same logistic regression model to fit the power of
-conventional tests and visual tests formulated as
-
-\begin{equation} \label{eq:logistic-regression-1-1}
-Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right),
-\end{equation}
-
-\noindent where \(\Lambda(.)\) is the standard logistic function given
-as \(\Lambda(z) = exp(z)/(1+exp(z))\). The effect size \(E\) is the only
-regressor and the intercept is fixed to \(log(0.05/0.95)\) so that
-\(\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05\).
-
-To study various factors contributing to the power of test, the same
-logistic regression model is fit on different subsets of the collated
-data grouped by levels of factors. These include the distribution of the
-regressor, type of the simulation model and the shape of the residual
-departures.
+subjects are asked to select one or more most different plots. In this
+scenario, having the probability of a subject \(j\) correctly picking
+the data plot from a lineup \(l\) is insufficient to determine the power
+of a visual test because it does not provide information about the
+number of selections made by the subject for the calculation of the
+p-value (See Equation \ref{eq:pvalue-beta-binomial-approx}). Therefore,
+we directly estimate the probability of a lineup being rejected by
+assuming that individual skill has negligible effect on the variation of
+the power. This assumption is not necessary true, but it helps
+simplifying the model structure, thereby obviate a costly large-scale
+experiment to estimate complex covariance matrices. The same model given
+in Equation \ref{eq:logistic-regression-1-1} is applied to model the
+power of a visual test.
+
+To study various factors contributing to the power of both tests, the
+same logistic regression model is fit on different subsets of the
+collated data grouped by levels of factors. These include the
+distribution of the fitted values, type of the simulation model and the
+shape of the residual departures.
 
@@
 
-To investigate the difference between conventional hypothesis testing
-and visual inference in the application of linear regression
-diagnostics, an experiment is conducted in three data collection
-periods. Two types of departures, non-linearity and heteroskedasticity,
-are primarily collected during data collection periods I and II. The
-data collection period III was designed primarily to measure human
-responses to null lineups so that the parameter \(\alpha\) in Equation
+An experiment is conducted in three data collection periods to
+investigate the difference between conventional hypothesis testing and
+visual inference in the application of linear regression diagnostics.
+Two types of departures, non-linearity and heteroskedasticity, are
+collected during data collection periods I and II. The data collection
+period III was designed primarily to measure human responses to null
+lineups so that the parameter \(\alpha\) in Equation
 \ref{eq:pvalue-beta-binomial} can be estimated. Additional lineups for
-both non-linearity and heteroskedasticity, using uniform \(X\)
+both non-linearity and heteroskedasticity, using uniform fitted value
 distribution, were included so that the participants were evaluating
@@
 used in non-linearity detection. This lineup is produced by the
-non-linearity model with \(j = 6\) and uniform fitted values. The data
-plot location is \(2^3 - 4\). All five subjects correctly identify the
-data plot from this lineup.
+non-linearity model with \(j = 6\). The data plot location is
+\(2^3 - 4\). All five subjects correctly identify the data plot from
+this lineup.
 
@@
 
-\caption{Examining the effect of signal strength for the three different values of $n$ used in the experiment, for non-linear structure with fixed $\sigma = 1.5$, uniform $X$ distribution, and S shape.
-For these factor levels, only when $n=300$ is the S shape clearly visible.}\label{fig:different-n}
+\caption{Examining the effect of signal strength for the three different values of $n$ used in the experiment, for non-linear structure with fixed $\sigma = 1$, uniform fitted value distribution, and S shape.
+For these factor levels, only when $n = 300$ is the S shape clearly visible.}\label{fig:different-n}
 \end{figure}
@@
 
-\caption{Examining the effect of $\sigma$ on the signal strength in the non-linearity detection, for $n=300$, uniform $X$ distribution and the U shape.
+\caption{Examining the effect of $\sigma$ on the signal strength in the non-linearity detection, for $n=300$, uniform fitted value distribution and the U shape.
 As $\sigma$ increases the signal strength decreases, to the point that the U is almost unrecognisable when $\sigma=4$.}\label{fig:different-sigma}
@@
 For \(b \neq 0\), the variance-covariance matrix of the error term
-\(\boldsymbol{\varepsilon}\) is correlated with the regressor
+\(\boldsymbol{\varepsilon}\) is correlated with the predictor
 \(\boldsymbol{x}\), which will lead to the presence of
@@
 An example lineup of this model used in data collection period II is
-shown in Figure \ref{fig:example-heter-lineup} with \(a = -1\) and
-\(X_{raw} \sim U(-1, 1)\). The data plot location is \(2^4 + 2\). Nine
-out of 11 subjects correctly identify the data plot from this lineup.
+shown in Figure \ref{fig:example-heter-lineup} with \(a = -1\). The data
+plot location is \(2^4 + 2\). Nine out of 11 subjects correctly identify
+the data plot from this lineup.
 
@@
 the predictors has a skewed distribution. This latter problem is usually
-corrected before modeling using a variable transformation. Our
-simulation asesses this by using four different distributions to
-represent the fitted values (\(X_{raw}\)): (1) \(U(-1, 1)\), (2)
+corrected before modelling using a variable transformation. Our
+simulation assess this by using four different distributions to
+represent fitted values: (1) uniform, (2) normal, (3) skewed and (4)
+discrete uniform. This is constructed by defining the raw predictor
+\(X_{raw}\) in four corresponding distributions: (1) \(U(-1, 1)\), (2)
 \(N(0, 0.3^2)\), (3) \(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). We
@@
 strength in the simulated data plots with smaller \(n\). A sample size
-of 300 is typically enough for structure to be visible in a scatterplot
+of 300 is typically enough for structure to be visible in a scatter plot
 reliably.
@@
 pattern. The parameter \(b\) controls the standard deviation of the
-error across the support of the regressor. Given \(x \neq a\), a larger
+error across the support of the predictor. Given \(x \neq a\), a larger
 value of \(b\) will lead to a larger ratio of the variance at \(x\) to
@@
 
-(do we actually need the penalty term?)
-
 Effect size \(E\) can be defined as the difference of a parameter for a
@@
 According to the underlying simulated model, we define the effect size
-\(E = \sqrt{n}/(\sigma k)\) for the non-linearity model and
-\(E = b\sqrt{n}/(|a| + 1)\) for the heteroskedasticity model. The
-parameter \(k\) is the number of turning points embedded in the residual
-plot. It is used as a penalty term to reduce the effect size for complex
-shapes. There are one, two, three and five turning points in ``U'',
-``S'', ``M'', ``Triple-U'' shape respectively. The denominator of the
-effect size of the heteroskedasticity model is \(|a| + 1\). It is used
-to
+
+\[E = \frac{1}{2}\left(\boldsymbol{\mu}_z'[diag(\boldsymbol{R}_a\sigma^2)]^{-1}\boldsymbol{\mu}_z\right)\]
+
+for the non-linearity model, and
+
+\[E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)\]
+for the heteroskedasticity model, where \(diag(.)\) is the diagonal
+matrix constructed from the diagonal elements of a matrix,
+\(\sigma^2\boldsymbol{I}\) is the assumed covariance matrix of the error
+term when \(H_0\) is true, \(\boldsymbol{V}\) is the actual covariance
+matrix of the error term,
+\(\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a\) is the
+residual operator,
+\(\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'\)
+is the hat matrix,
+\(\boldsymbol{X}_a = (\boldsymbol{1}, \boldsymbol{X})\) is the matrix of
+predictors including the intercept used in the regression equation, and
+\(\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z\)
+is the expected values of residuals with \(\boldsymbol{Z}\) be any
+higher order terms of \(\boldsymbol{X}\) leave out by the regression
+equation and \(\boldsymbol{\beta}_z\) be the corresponding coefficients.
+
+Both equations above are derived from the Kullback-Leibler divergence
+(citation here?) and the details are provided in the appendix.
 
@@
 generation procedure differs from the canonical Rorschach lineup
-procedure, which requires that all 20 plots are generated from the null
-hypothesis. However, these lineups serve the same fundamental purpose:
-to assess the number of visually interesting plots generated under the
-null hypothesis.
+procedure, which requires that all 20 plots are generated from \(H_0\).
+However, these lineups serve the same fundamental purpose: to assess the
+number of visually interesting plots generated under \(H_0\).
 
@@
 four is usually sufficient. Thus, we follow this guideline to conduct
-the RESET test. For the Breusch-Pagan test, the choice of regressors in
+the RESET test. For the Breusch-Pagan test, the choice of predictors in
 the auxiliary regression is left to the user
 \citep{breusch_simple_1979}. But as \citet{waldman1983note} suggested,
-it is a good choice for the set of auxiliary regressors in the
+it is a good choice for the set of auxiliary predictors in the
 Breusch-Pagan test be the same as the White test. Thus, we include both
@@
 residual departures are unnoticeable from the corresponding residual
-plot, the RESET test has above 25\% of chance of rejecting the null
-hypothesis, while only a few visual tests would result in a rejection.
-Similarly, at \(log(E) = 3\) for heteroskedasticity model, the BP test
-has around 25\% power but the non-constant variance visual pattern is
-almost unobserved from the residual plot and the power of the visual
-test is close to the significance level at this point.
+plot, the RESET test has above 25\% of chance of rejecting \(H_0\),
+while only a few visual tests would result in a rejection. Similarly, at
+\(log(E) = 3\) for heteroskedasticity model, the BP test has around 25\%
+power but the non-constant variance visual pattern is almost unobserved
+from the residual plot and the power of the visual test is close to the
+significance level at this point.
 

### 2023-03-06 paper_comparison/paper_comparison/paper_comparison.Rmd
update exp setup and the choice of example plot

@@
 
+Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups used in non-linearity detection.
+This lineup is produced by the non-linearity model with $j = 6$.
+The data plot location is $2^3 - 4$.
+All five subjects correctly identify the data plot from this lineup.
+
 ```{r results='asis'}
@@
   xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
-```
-
-```{r different-dist, fig.height = 2, fig.cap = "Variations in fitted values, that might affect perception of residual plots. Four different distributions are used."}
-
-# Data for uniform distribution
-dat_dist_1 <- poly_model(shape = 1, 
-                      x = {
-                        raw_x <- rand_uniform(-1, 1);
-                        closed_form(~stand_dist(raw_x))
-                        }, 
-                      sigma = 0.5)$gen(300) %>%
-  mutate(x_dist = "Uniform")
-
-# Generate data for other distributions
-dat_dist_2 <- poly_model(shape = 1,
-                         x = {
-                           raw_x <- rand_normal(sigma = 0.3); 
-                           closed_form(~stand_dist(raw_x))
-                           }, 
-                         sigma = 0.5)$
-  gen(300, computed = select(dat_dist_1, e)) %>%
-  mutate(x_dist = "Normal")
-
-dat_dist_3 <- poly_model(shape = 1,
-                         x = {
-                           raw_x <- rand_lognormal(sigma = 0.6); 
-                           closed_form(~stand_dist(raw_x/3 - 1))
-                           }, 
-                         sigma = 0.5)$
-  gen(300, computed = select(dat_dist_1, e)) %>%
-  mutate(x_dist = "Skewed")
-
-dat_dist_4 <- poly_model(shape = 1,
-                         x = {
-                           raw_x <- rand_uniform_d(k = 5, even = TRUE); 
-                           closed_form(~stand_dist(raw_x))
-                           }, 
-                         sigma = 0.5)$
-  gen(300, computed = select(dat_dist_1, e)) %>%
-  mutate(x_dist = "Discrete uniform")
-
-# Generate and plot data for discrete uniform distribution
-bind_rows(dat_dist_1, dat_dist_2, dat_dist_3, dat_dist_4) %>%
-  mutate(x_dist = factor(x_dist, 
-                         levels = c("Uniform", 
-                                    "Normal", 
-                                    "Skewed", 
-                                    "Discrete"))) %>%
-  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank()) +
-  facet_wrap(~x_dist, ncol = 4, scales = "free")
-```
-
-Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups used in non-linearity detection.
-This lineup is produced by the non-linearity model with $j = 6$.
-The data plot location is $2^3 - 4$.
-All five subjects correctly identify the data plot from this lineup.
-
-```{r different-n, fig.height = 2.67, fig.cap = "Examining the effect of signal strength for the three different values of $n$ used in the experiment, for non-linear structure with fixed $\\sigma = 1$, uniform fitted value distribution, and S shape.
-For these factor levels, only when $n = 300$ is the S shape clearly visible."}
-
-# Generate data for 300 observations
-dat_n_300 <- poly_model(shape = 2, 
-                        x = {
-                          raw_x <- rand_uniform(-1, 1)
-                          closed_form(~stand_dist(raw_x))
-                        },
-                        sigma = 1)$gen(300) %>%
-  mutate(n = "N: 300")
-
-# Generate data for other settings
-map(c(50, 100), function(n) {
-  poly_model(shape = 2, 
-                   x = {
-                     raw_x <- rand_uniform(-1, 1);
-                     closed_form(~stand_dist(raw_x))
-                     }, 
-             sigma = 1.5)$
-    gen(n, computed = select(dat_n_300[1:n, ], x, e)) %>%
-    mutate(n = paste0("N: ", n))
-}) %>%
-
-  # Combined with data for 300 observations
-  bind_rows(dat_n_300) %>%
-  mutate(n = factor(n, levels = c("N: 50", "N: 100", "N: 300"))) %>%
-  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
-  facet_wrap(~n, scales = "free", ncol = 3) +
-  xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
+  ylab("Residuals")
 ```
@@
   xlab("Fitted values") +
-  ylab("Residuals") +
-  theme(axis.text = element_blank())
+  ylab("Residuals")
 ```
@@
 Three different shapes ($a = -1, 0, 1$) are used in the experiment to create left-triangle, butterfly and right-triangle shapes, respectively."}
+
+set.seed(10086)
 
@@
 
+```{r different-dist, fig.height = 2, fig.cap = "Variations in fitted values, that might affect perception of residual plots. Four different distributions are used."}
+
+# Data for uniform distribution
+dat_dist_1 <- poly_model(shape = 1, 
+                      x = {
+                        raw_x <- rand_uniform(-1, 1);
+                        closed_form(~stand_dist(raw_x))
+                        }, 
+                      sigma = 0.5)$gen(300) %>%
+  mutate(x_dist = "Uniform")
+
+# Generate data for other distributions
+dat_dist_2 <- poly_model(shape = 1,
+                         x = {
+                           raw_x <- rand_normal(sigma = 0.3); 
+                           closed_form(~stand_dist(raw_x))
+                           }, 
+                         sigma = 0.5)$
+  gen(300, computed = select(dat_dist_1, e)) %>%
+  mutate(x_dist = "Normal")
+
+dat_dist_3 <- poly_model(shape = 1,
+                         x = {
+                           raw_x <- rand_lognormal(sigma = 0.6); 
+                           closed_form(~stand_dist(raw_x/3 - 1))
+                           }, 
+                         sigma = 0.5)$
+  gen(300, computed = select(dat_dist_1, e)) %>%
+  mutate(x_dist = "Skewed")
+
+dat_dist_4 <- poly_model(shape = 1,
+                         x = {
+                           raw_x <- rand_uniform_d(k = 5, even = TRUE); 
+                           closed_form(~stand_dist(raw_x))
+                           }, 
+                         sigma = 0.5)$
+  gen(300, computed = select(dat_dist_1, e)) %>%
+  mutate(x_dist = "Discrete uniform")
+
+# Generate and plot data for discrete uniform distribution
+bind_rows(dat_dist_1, dat_dist_2, dat_dist_3, dat_dist_4) %>%
+  mutate(x_dist = factor(x_dist, 
+                         levels = c("Uniform", 
+                                    "Normal", 
+                                    "Skewed", 
+                                    "Discrete uniform"))) %>%
+  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
+  xlab("Fitted values") +
+  ylab("Residuals") +
+  facet_wrap(~x_dist, ncol = 4, scales = "free")
+```
+
+```{r different-n, fig.height = 2.67, fig.cap = "Examining the effect of signal strength for the three different values of $n$ used in the experiment, for non-linear structure with fixed $\\sigma = 1.5$, uniform fitted value distribution, and S shape.
+For these factor levels, only when $n = 300$ is the S shape clearly visible."}
+
+set.seed(666)
+
+# Generate data for 300 observations
+dat_n_300 <- poly_model(shape = 2, 
+                        x = {
+                          raw_x <- rand_uniform(-1, 1)
+                          closed_form(~stand_dist(raw_x))
+                        },
+                        sigma = 1)$gen(300) %>%
+  mutate(n = "N: 300")
+
+# Generate data for other settings
+map(c(50, 100), function(n) {
+  poly_model(shape = 2, 
+                   x = {
+                     raw_x <- rand_uniform(-1, 1);
+                     closed_form(~stand_dist(raw_x))
+                     }, 
+             sigma = 1)$
+    gen(n, computed = select(dat_n_300[1:n, ], x, e)) %>%
+    mutate(n = paste0("N: ", n))
+}) %>%
+
+  # Combined with data for 300 observations
+  bind_rows(dat_n_300) %>%
+  mutate(n = factor(n, levels = c("N: 50", "N: 100", "N: 300"))) %>%
+  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
+  facet_wrap(~n, scales = "free", ncol = 3) +
+  xlab("Fitted values") +
+  ylab("Residuals")
+```
+
 ## Experimental setup
@@
 
-Both equations above are derived from the Kullback-Leibler divergence (citation here?) and the details are provided in the appendix.
+Both equations above are derived from the Kullback-Leibler divergence (citation here?) and the details are provided in the Appendix.
 
@@
 
-Three replications are made for each of the parameter values shown in Table \ref{tab:model-factor-table} resulting in $(4 \times 4 \times 4 \times 3 + 4 \times 3 \times 5 \times 3) \times 3 = `r (4 * 4 * 4 * 3 + 4 * 3 * 5 * 3) * 3`$ different lineups.
+As shown in Table \ref{tab:model-factor-table}, there are a total of $4 \times 4 \times 3 \times 4 = 192$ and $3 \times 5 \times 3 \times 4 = 180$ number of combinations of parameter values for non-linearity model and heteroskedasticity model respectively.
+Three replications are made for each of the combination results in $192 \times 3 = 576$ and $180 \times 3 = 540$ lineups.
 In addition, each lineup is designed to be evaluated by five different subjects.
-After attempting some pilot studies internally in our research group, we decide to present a block of 20 lineups to every subject.
+After attempting some pilot studies internally, we decide to present a block of 20 lineups to every subject.
 And to ensure the quality of the survey data, two lineups with obvious visual patterns are included as attention checks.
-Thus, $576 \times 5 / (20-2) = `r 576 * 5 / 18`$ and $540 \times 5 / (20-2) = `r 540 * 5 / 18`$ subjects are recruited to satisfy the design of the data collection period I and II respectively.
+Thus, $576 \times 5 / (20-2) = 160$ and $540 \times 5 / (20-2) = 150$ subjects are recruited to satisfy the design of the data collection period I and II respectively.
 
 As mentioned in Section \ref{calculating-p-values-for-the-visual-test}, $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using null lineups.
-Hence, three replications of 36 lineups with all combinations of $n$ and fitted value distributions are included in data collection period III.
+Three replications are made for $3 \times 4 = 12$ combinations of common factors $n$ and fitted value distribution, results in $12 \times 3 = 36$ lineups included in data collection period III.
 In these lineups, the data of the data plot is generated from a model with zero effect size, while the data of the 19 null plots are generated using the same simulation method discussed in Section \ref{sampling-from-the-null-distribution}.
-This generation procedure differs from the canonical Rorschach lineup procedure, which requires that all 20 plots are generated from $H_0$.
+This generation procedure differs from the canonical Rorschach lineup procedure, which requires that all 20 plots are generated directly from the null model.
 However, these lineups serve the same fundamental purpose: to assess the number of visually interesting plots generated under $H_0$.
@@
 However, presenting only these lineups to subjects are considered to be bad practices as subjects will lose interest quickly. 
-Therefore, we plan to collect 6 more evaluations on the 279 lineups with uniform fitted values, result in $(36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3 \times 6) / (20-2) = `r (36 * 20 + (4 * 4 * 3 + 3 * 5 * 3) * 3 * 6) / 18`$ subjects recruited for data collection period III.
+Therefore, we plan to collect 6 more evaluations on the 279 lineups with uniform fitted value distribution, result in $(36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133$ subjects recruited for data collection period III.
 

### 2023-03-06 paper_comparison/paper_comparison/paper_comparison.tex
update exp setup and the choice of example plot

@@
 
+Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups
+used in non-linearity detection. This lineup is produced by the
+non-linearity model with \(j = 6\). The data plot location is
+\(2^3 - 4\). All five subjects correctly identify the data plot from
+this lineup.
+
 \begin{table}
@@
 The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$.}\label{fig:different-shape-of-herimite}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-dist-1}
-
-}
-
-\caption{Variations in fitted values, that might affect perception of residual plots. Four different distributions are used.}\label{fig:different-dist}
-\end{figure}
-
-Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups
-used in non-linearity detection. This lineup is produced by the
-non-linearity model with \(j = 6\). The data plot location is
-\(2^3 - 4\). All five subjects correctly identify the data plot from
-this lineup.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-n-1}
-
-}
-
-\caption{Examining the effect of signal strength for the three different values of $n$ used in the experiment, for non-linear structure with fixed $\sigma = 1$, uniform fitted value distribution, and S shape.
-For these factor levels, only when $n = 300$ is the S shape clearly visible.}\label{fig:different-n}
 \end{figure}
@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-dist-1}
+
+}
+
+\caption{Variations in fitted values, that might affect perception of residual plots. Four different distributions are used.}\label{fig:different-dist}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-n-1}
+
+}
+
+\caption{Examining the effect of signal strength for the three different values of $n$ used in the experiment, for non-linear structure with fixed $\sigma = 1.5$, uniform fitted value distribution, and S shape.
+For these factor levels, only when $n = 300$ is the S shape clearly visible.}\label{fig:different-n}
+\end{figure}
+
 \hypertarget{experimental-setup}{%
@@
 Both equations above are derived from the Kullback-Leibler divergence
-(citation here?) and the details are provided in the appendix.
+(citation here?) and the details are provided in the Appendix.
 
@@
 
-Three replications are made for each of the parameter values shown in
-Table \ref{tab:model-factor-table} resulting in
-\((4 \times 4 \times 4 \times 3 + 4 \times 3 \times 5 \times 3) \times 3 = 1116\)
-different lineups. In addition, each lineup is designed to be evaluated
-by five different subjects. After attempting some pilot studies
-internally in our research group, we decide to present a block of 20
-lineups to every subject. And to ensure the quality of the survey data,
-two lineups with obvious visual patterns are included as attention
-checks. Thus, \(576 \times 5 / (20-2) = 160\) and
+As shown in Table \ref{tab:model-factor-table}, there are a total of
+\(4 \times 4 \times 3 \times 4 = 192\) and
+\(3 \times 5 \times 3 \times 4 = 180\) number of combinations of
+parameter values for non-linearity model and heteroskedasticity model
+respectively. Three replications are made for each of the combination
+results in \(192 \times 3 = 576\) and \(180 \times 3 = 540\) lineups. In
+addition, each lineup is designed to be evaluated by five different
+subjects. After attempting some pilot studies internally, we decide to
+present a block of 20 lineups to every subject. And to ensure the
+quality of the survey data, two lineups with obvious visual patterns are
+included as attention checks. Thus, \(576 \times 5 / (20-2) = 160\) and
 \(540 \times 5 / (20-2) = 150\) subjects are recruited to satisfy the
@@
 \(\alpha\) used in Equation \ref{eq:pvalue-beta-binomial} needs to be
-estimated using null lineups. Hence, three replications of 36 lineups
-with all combinations of \(n\) and fitted value distributions are
-included in data collection period III. In these lineups, the data of
-the data plot is generated from a model with zero effect size, while the
-data of the 19 null plots are generated using the same simulation method
-discussed in Section \ref{sampling-from-the-null-distribution}. This
-generation procedure differs from the canonical Rorschach lineup
-procedure, which requires that all 20 plots are generated from \(H_0\).
+estimated using null lineups. Three replications are made for
+\(3 \times 4 = 12\) combinations of common factors \(n\) and fitted
+value distribution, results in \(12 \times 3 = 36\) lineups included in
+data collection period III. In these lineups, the data of the data plot
+is generated from a model with zero effect size, while the data of the
+19 null plots are generated using the same simulation method discussed
+in Section \ref{sampling-from-the-null-distribution}. This generation
+procedure differs from the canonical Rorschach lineup procedure, which
+requires that all 20 plots are generated directly from the null model.
 However, these lineups serve the same fundamental purpose: to assess the
@@
 Therefore, we plan to collect 6 more evaluations on the 279 lineups with
-uniform fitted values, result in
-\((36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3 \times 6) / (20-2) = 133\)
-subjects recruited for data collection period III.
+uniform fitted value distribution, result in
+\((36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133\) subjects
+recruited for data collection period III.
 

### 2023-03-06 paper_comparison/paper_comparison/paper_comparison.Rmd
Fix cross reference issue

@@
 
-As mentioned in Section \ref{calculating-p-values-for-the-visual-test}, $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using null lineups.
+As mentioned in Section \ref{power-of-the-tests}, $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using null lineups.
 Three replications are made for $3 \times 4 = 12$ combinations of common factors $n$ and fitted value distribution, results in $12 \times 3 = 36$ lineups included in data collection period III.
-In these lineups, the data of the data plot is generated from a model with zero effect size, while the data of the 19 null plots are generated using the same simulation method discussed in Section \ref{sampling-from-the-null-distribution}.
+In these lineups, the data of the data plot is generated from a model with zero effect size, while the data of the 19 null plots are generated using the same simulation method discussed in Section \ref{visual-test-procedure-based-on-lineups}.
 This generation procedure differs from the canonical Rorschach lineup procedure, which requires that all 20 plots are generated directly from the null model.
@@
 
-The data of the data plot is simulated from one of two models described in Section \ref{simulating-departures-from-good-residuals}, while the data of the remaining 19 null plots are generated by the residual rotation technique discussed in Section \ref{sampling-from-the-null-distribution}.
+The data of the data plot is simulated from one of two models described in Section \ref{simulating-departures-from-good-residuals}, while the data of the remaining 19 null plots are generated by the residual rotation technique discussed in Section \ref{visual-test-procedure-based-on-lineups}.
 

### 2023-03-06 paper_comparison/paper_comparison/paper_comparison.tex
Fix cross reference issue

@@
 
-As mentioned in Section \ref{calculating-p-values-for-the-visual-test},
-\(\alpha\) used in Equation \ref{eq:pvalue-beta-binomial} needs to be
-estimated using null lineups. Three replications are made for
-\(3 \times 4 = 12\) combinations of common factors \(n\) and fitted
-value distribution, results in \(12 \times 3 = 36\) lineups included in
-data collection period III. In these lineups, the data of the data plot
-is generated from a model with zero effect size, while the data of the
-19 null plots are generated using the same simulation method discussed
-in Section \ref{sampling-from-the-null-distribution}. This generation
-procedure differs from the canonical Rorschach lineup procedure, which
-requires that all 20 plots are generated directly from the null model.
-However, these lineups serve the same fundamental purpose: to assess the
-number of visually interesting plots generated under \(H_0\).
+As mentioned in Section \ref{power-of-the-tests}, \(\alpha\) used in
+Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using null
+lineups. Three replications are made for \(3 \times 4 = 12\)
+combinations of common factors \(n\) and fitted value distribution,
+results in \(12 \times 3 = 36\) lineups included in data collection
+period III. In these lineups, the data of the data plot is generated
+from a model with zero effect size, while the data of the 19 null plots
+are generated using the same simulation method discussed in Section
+\ref{visual-test-procedure-based-on-lineups}. This generation procedure
+differs from the canonical Rorschach lineup procedure, which requires
+that all 20 plots are generated directly from the null model. However,
+these lineups serve the same fundamental purpose: to assess the number
+of visually interesting plots generated under \(H_0\).
 
@@
 rotation technique discussed in Section
-\ref{sampling-from-the-null-distribution}.
+\ref{visual-test-procedure-based-on-lineups}.
 

### 2023-03-06 paper_comparison/paper_comparison/paper_comparison.Rmd
Polish power cmp
Now the summary of the findings is better. I mention that conventional test is still the recommended choice if the strict correctness of the model assumption is of particular interest.

@@
 But as @waldman1983note suggested, it is a good choice for the set of auxiliary predictors in the Breusch-Pagan test be the same as the White test.
-Thus, we include both $\boldsymbol{x}$ and $\boldsymbol{x}^2$ in the auxiliary regression.
+Thus, we include both $x$ and $x^2$ in the auxiliary regression.
 
@@
 
-For both the non-linearity pattern and the heteroskedasticity pattern, the power curves of appropriate conventional tests are constantly greater than the power curves of visual tests and other inappropriate conventional tests.
-At $log(E) = 1$ for non-linearity model, where the residual departures are unnoticeable from the corresponding residual plot, the RESET test has above 25% of chance of rejecting $H_0$, while only a few visual tests would result in a rejection.
-Similarly, at $log(E) = 3$ for heteroskedasticity model, the BP test has around 25% power but the non-constant variance visual pattern is almost unobserved from the residual plot and the power of the visual test is close to the significance level at this point.
-
-Further, the power of all inappropriate tests show improvement as the effect increases but at a lower rate than the visual test in both scenarios.
+For non-linearity patterns, the power curve of RESET test climbs aggressively from
+`r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(0), offset0 = log(0.05/0.95)), type = "response")) * 100)`% to `r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(2.5), offset0 = log(0.05/0.95)), type = "response")) * 100)`% as $log_e(E)$ increases from 0 to 2.5, while power of other tests respond inactively to the change of effect, showing that RESET test is much more sensitive to weak non-linear structure.
+Meanwhile, no noticeable residual departures can be spotted from the example residual plots.
+
+In terms of heteroskedasticity patterns, the power of Breusch-Pagan test is constantly greater than the power of visual test.
+At $log_e(E) \approx 2.5$, where the power curve of the visual test remains at a low level, the Breusch-Pagan test has around 50% chance of rejecting $H_0$.
+Similarly, the visual feature is nearly unobservable from the example residual plots at this level of effect size.
+
+The power of visual test arises steadily as $log_e(E)$ increases from 3 to 5 for both non-linearity patterns and heteroskedasticity patterns, suggesting that the effect starts to make significant impact on the degree of the presence of the designed visual features.
+This can also be observed from the example residual plots that when $log_e(E) = 3.5$, a weak "S-shape" and a weak "right-triangle" shape are presented in Figure \ref{fig:polypower} and Figure \ref{fig:heterpower} respectively.
+The visual pattern becomes clearer as $log_e(E)$ increases.
+At $log_e(E) = 5$, the power of visual tests for both patterns reaches almost 100%.
+
+Power curves of inappropriate tests show improvement as the effect increases but at a lower rate than the visual test in both scenarios.
 This coincides the point made by @cook1982residuals that residual-based tests for a specific type of model defect may be sensitive to other types of model defects.
@@
 
-Overall, the power comparison suggests that conventional tests differs significantly from visual tests in two regression diagnostics scenarios designed by us.
+Overall, the power comparison suggests that conventional tests differs significantly from visual tests in two regression diagnostics scenarios designed in this paper.
 Visual test have much higher tolerance of the residual departures than the conventional test.
-Since fail to reject $H_0$ in a visual test usually means that there are no obvious visual discoveries found in the residual plot, analysts and the general public as the consumers of the output may not be convinced of the existence of significant residual departures in spite of the rejection of $H_0$ given by the conventional test.
+Since fail to reject $H_0$ in a visual test literally means that there are no obvious visual discoveries found in the residual plot, analysts and the general public as the consumers of the output may not be convinced of the existence of significant residual departures in spite of the rejection of $H_0$ given by the conventional test.
 Even if the rejection is accepted, the model violation may be considered as impactless due to the fact that they are not clearly visible.
-Besides, the sensitivity of the conventional test could also distract and discourage analysts from finding simple but good linear approximation to the data.
-The rejection of $H_0$ because of human acceptable and negligible residual departures is not practically meaningful and useful.
-This may limit the popularity of conventional tests in residual diagnostics among analysts.
+Besides, the sensitivity of the conventional test to weak residual departures could also distract and discourage analysts from finding simple but good linear approximation to the data.
+The rejection of $H_0$ because of human acceptable and negligible residual departures is not practically meaningful and useful to analysts and decision makers.
+In contrast, if the strict correctness of the model assumption is of particular interest, conducting a conventional test is still the recommended choice based on the findings.
 
@@
   j <- j + 1
-  plist[[j]] <- heter_model(a = 0, 
+  plist[[j]] <- heter_model(a = 1, 
                             b = i, 

### 2023-03-06 paper_comparison/paper_comparison/paper_comparison.tex
Polish power cmp
Now the summary of the findings is better. I mention that conventional test is still the recommended choice if the strict correctness of the model assumption is of particular interest.

@@
 Breusch-Pagan test be the same as the White test. Thus, we include both
-\(\boldsymbol{x}\) and \(\boldsymbol{x}^2\) in the auxiliary regression.
+\(x\) and \(x^2\) in the auxiliary regression.
 
@@
 
-For both the non-linearity pattern and the heteroskedasticity pattern,
-the power curves of appropriate conventional tests are constantly
-greater than the power curves of visual tests and other inappropriate
-conventional tests. At \(log(E) = 1\) for non-linearity model, where the
-residual departures are unnoticeable from the corresponding residual
-plot, the RESET test has above 25\% of chance of rejecting \(H_0\),
-while only a few visual tests would result in a rejection. Similarly, at
-\(log(E) = 3\) for heteroskedasticity model, the BP test has around 25\%
-power but the non-constant variance visual pattern is almost unobserved
-from the residual plot and the power of the visual test is close to the
-significance level at this point.
-
-Further, the power of all inappropriate tests show improvement as the
-effect increases but at a lower rate than the visual test in both
-scenarios. This coincides the point made by \citet{cook1982residuals}
-that residual-based tests for a specific type of model defect may be
+For non-linearity patterns, the power curve of RESET test climbs
+aggressively from 7\% to 85\% as \(log_e(E)\) increases from 0 to 2.5,
+while power of other tests respond inactively to the change of effect,
+showing that RESET test is much more sensitive to weak non-linear
+structure. Meanwhile, no noticeable residual departures can be spotted
+from the example residual plots.
+
+In terms of heteroskedasticity patterns, the power of Breusch-Pagan test
+is constantly greater than the power of visual test. At
+\(log_e(E) \approx 2.5\), where the power curve of the visual test
+remains at a low level, the Breusch-Pagan test has around 50\% chance of
+rejecting \(H_0\). Similarly, the visual feature is nearly unobservable
+from the example residual plots at this level of effect size.
+
+The power of visual test arises steadily as \(log_e(E)\) increases from
+3 to 5 for both non-linearity patterns and heteroskedasticity patterns,
+suggesting that the effect starts to make significant impact on the
+degree of the presence of the designed visual features. This can also be
+observed from the example residual plots that when \(log_e(E) = 3.5\), a
+weak ``S-shape'' and a weak ``right-triangle'' shape are presented in
+Figure \ref{fig:polypower} and Figure \ref{fig:heterpower} respectively.
+The visual pattern becomes clearer as \(log_e(E)\) increases. At
+\(log_e(E) = 5\), the power of visual tests for both patterns reaches
+almost 100\%.
+
+Power curves of inappropriate tests show improvement as the effect
+increases but at a lower rate than the visual test in both scenarios.
+This coincides the point made by \citet{cook1982residuals} that
+residual-based tests for a specific type of model defect may be
 sensitive to other types of model defects. The power curve of RESET test
@@
 significantly from visual tests in two regression diagnostics scenarios
-designed by us. Visual test have much higher tolerance of the residual
-departures than the conventional test. Since fail to reject \(H_0\) in a
-visual test usually means that there are no obvious visual discoveries
-found in the residual plot, analysts and the general public as the
-consumers of the output may not be convinced of the existence of
-significant residual departures in spite of the rejection of \(H_0\)
-given by the conventional test. Even if the rejection is accepted, the
-model violation may be considered as impactless due to the fact that
-they are not clearly visible. Besides, the sensitivity of the
-conventional test could also distract and discourage analysts from
-finding simple but good linear approximation to the data. The rejection
-of \(H_0\) because of human acceptable and negligible residual
-departures is not practically meaningful and useful. This may limit the
-popularity of conventional tests in residual diagnostics among analysts.
+designed in this paper. Visual test have much higher tolerance of the
+residual departures than the conventional test. Since fail to reject
+\(H_0\) in a visual test literally means that there are no obvious
+visual discoveries found in the residual plot, analysts and the general
+public as the consumers of the output may not be convinced of the
+existence of significant residual departures in spite of the rejection
+of \(H_0\) given by the conventional test. Even if the rejection is
+accepted, the model violation may be considered as impactless due to the
+fact that they are not clearly visible. Besides, the sensitivity of the
+conventional test to weak residual departures could also distract and
+discourage analysts from finding simple but good linear approximation to
+the data. The rejection of \(H_0\) because of human acceptable and
+negligible residual departures is not practically meaningful and useful
+to analysts and decision makers. In contrast, if the strict correctness
+of the model assumption is of particular interest, conducting a
+conventional test is still the recommended choice based on the findings.
 

### 2023-03-07 paper_comparison/paper_comparison/paper_comparison.Rmd
Fix caption and polish effect of non-linearity shape

@@
 This ensures $\sum_{i}c_i=K$.
-Since we are only interested in the selections of the data plot $i$, the marginal model can be simplified to a beta-binomial model and thus the visual p-value is given as
+Since we are only interested in the selections of the data plot $i$, the marginal model can be simplified to a beta-binomial model and thus the visual $p$-value is given as
 
@@
 However, this mix-effect model does not work with scenario where subjects are asked to select one or more most different plots.
-In this scenario, having the probability of a subject $j$ correctly picking the data plot from a lineup $l$ is insufficient to determine the power of a visual test because it does not provide information about the number of selections made by the subject for the calculation of the p-value (See Equation \ref{eq:pvalue-beta-binomial-approx}).
+In this scenario, having the probability of a subject $j$ correctly picking the data plot from a lineup $l$ is insufficient to determine the power of a visual test because it does not provide information about the number of selections made by the subject for the calculation of the $p$-value (See Equation \ref{eq:pvalue-beta-binomial-approx}).
 Therefore, we directly estimate the probability of a lineup being rejected by assuming that individual skill has negligible effect on the variation of the power.
@@
 (A summary of the factors is given in Table \ref{tab:model-factor-table}.) The values of $j$ is chosen so that distinct shapes of non-linearity are included in the residual plot.
-These include "U", "S", "M" and "Triple-U" shape as shown in Figure \ref{fig:different-shape-of-herimite}.
+These include "U", "S", "M" and "triple-U" shape as shown in Figure \ref{fig:different-shape-of-herimite}.
 A greater value of $j$ will result in a curve with more turning points.
@@
 # Facet label
-shape_labels <- c('He[2]:"U"', 'He[3]:"S"', 'He[6]:"M"', 'He[18]:"Triple-U"')
+shape_labels <- c('He[2]:"U"', 'He[3]:"S"', 'He[6]:"M"', 'He[18]:"triple-U"')
 
@@
 
-```{r different-sigma, fig.height = 2, fig.cap = "Examining the effect of $\\sigma$ on the signal strength in the non-linearity detection, for $n=300$, uniform fitted value distribution and the U shape.
-As $\\sigma$ increases the signal strength decreases, to the point that the U is almost unrecognisable when $\\sigma=4$."}
+```{r different-sigma, fig.height = 2, fig.cap = 'Examining the effect of $\\sigma$ on the signal strength in the non-linearity detection, for $n=300$, uniform fitted value distribution and the "U" shape.
+As $\\sigma$ increases the signal strength decreases, to the point that the "U" is almost unrecognisable when $\\sigma=4$.'}
 
@@
 
-```{r different-shape-of-heter, fig.height = 2.67, fig.cap = "Heteroskedasticity forms used in the experiment.
-Three different shapes ($a = -1, 0, 1$) are used in the experiment to create left-triangle, butterfly and right-triangle shapes, respectively."}
+```{r different-shape-of-heter, fig.height = 2.67, fig.cap = 'Heteroskedasticity forms used in the experiment.
+Three different shapes ($a = -1, 0, 1$) are used in the experiment to create left-triangle, "butterfly" and "right-triangle" shapes, respectively.'}
 
@@
 
-a_labels <- c("Left-triangle", "Butterfly", "Right-triangle")
+a_labels <- c("left-triangle", "butterfly", "right-triangle")
 
@@
 
-```{r different-n, fig.height = 2.67, fig.cap = "Examining the effect of signal strength for the three different values of $n$ used in the experiment, for non-linear structure with fixed $\\sigma = 1.5$, uniform fitted value distribution, and S shape.
-For these factor levels, only when $n = 300$ is the S shape clearly visible."}
+```{r different-n, fig.height = 2.67, fig.cap = 'Examining the effect of signal strength for the three different values of $n$ used in the experiment, for non-linear structure with fixed $\\sigma = 1.5$, uniform fitted value distribution, and "S" shape.
+For these factor levels, only when $n = 300$ is the "S" shape clearly visible.'}
 
@@
 
-Overall, the power comparison suggests that conventional tests differs significantly from visual tests in two regression diagnostics scenarios designed in this paper.
+(should the below p goes to conclusion as a discussion?)
+
+Overall, the power comparison suggests that conventional tests differs significantly from visual tests in two regression diagnostics scenarios designed by us.
 Visual test have much higher tolerance of the residual departures than the conventional test.
@@
 Surrounding lines of the visual test show the estimated power of 500 bootstrap samples.
-Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
+Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by vertical lines in main plot.
 Where would you draw the line of too much non-linearity in the residuals?
-For the RESET test this is around log effect size 2.5, but for the visual test it is around 3.5.", dev = 'png', dpi = 450}
+For the RESET test this is around log effect size 2, but for the visual test it is around 3.5.", dev = 'png', dpi = 450}
 
@@
 Main plot shows the power curves, with dots indicating human evaluations of lineups.
-Surrounding lines of the visual test show the estimated power of 1000 bootstrap samples.
-Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
+Surrounding lines of the visual test show the estimated power of 500 bootstrap samples.
+Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by vertical lines in main plot.
 Where would you draw the line of too much heteroskedasticity in the residuals?
@@
 
-```{r p-value-comparison, fig.width = 8, fig.height = 4, fig.cap = "Rejection rate ($p$-value $\\leq0.05$) of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot.
-The visual test rejects less frequently than the conventional test.
-We would generally expect that the visual test would only reject when the conventional test does.
-Surprisingly, there is one lineup in the heteroskedasticity study where this is not the case.
-"}
-
-library(ggmosaic)
-
+```{r}
 # Dataset for p-value comparison
@@
   select(type, conv_reject, reject)
+```
+
+The power comparison illustrates that appropriate conventional tests will reject $H_0$ more aggressively than visual tests.
+In this section, we explore how often they agree with each other by investigating the rejections for the two model designs based on $p$-values for each lineup.
+
+Figure \ref{fig:p-value-comparison} provides a mosaic plot showing the rejection rate of visual tests and conventional tests for both non-linearity patterns and heteroskedasticity patterns.
+
+For lineups containing non-linearity patterns, conventional tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% and visual tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the time.
+Of the lineups rejected by the conventional test, `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by the visual test, that is, approximately half as many as the conventional test.
+There are no lineups that are rejected by the visual test but not by the conventional test.
+
+In terms of lineups containing heteroskedasticity patterns, `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by conventional tests, while `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by visual tests.
+When the conventional test rejects a lineup, there is a great chance (`r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`%) the visual test will also reject it.
+
+```{r}
+nr <- p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Not") %>% count(reject)
+```
+
+Surprisingly, the visual test rejects `r nr$n[1]` of the `r nr$n[1]+nr$n[2]` (`r round(nr$n[1]/(nr$n[1]+nr$n[2])*100, 0)`%) of lineups where the conventional test does not reject.
+Figure \ref{fig:heter-example} shows this lineup.
+The data plot in position seventeen displays a relatively strong heteroskedasticity pattern, and has a strong effect size ($log_e(E)=4.02$).
+This is reflected by the visual test $p\text{-value} = 0.026$.
+But the Breusch-Pagan test $p\text{-value} = 0.056$, is slightly above the significance cutoff of $0.05$.
+This lineup was evaluated by 11 subjects, it has experimental factors $a = 0$ ("butterfly" shape), $b = 64$ (large variance ratio), $n = 50$ (small sample size), and a uniform distribution for the fitted values.
+It must be the small sample size that may have resulted in the lack of detection.
+
+```{r p-value-comparison, fig.width = 8, fig.height = 4, fig.cap = "Rejection rate ($p$-value $\\leq0.05$) of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot.
+The visual test rejects less frequently than the conventional test.
+We would generally expect the visual test to only reject when the conventional test does.
+Surprisingly, one lineup containing a heteroskedasticity pattern does not follow this rule.
+"}
+
+library(ggmosaic)
 
@@
 
-The power comparison illustrates that appropriate conventional tests will reject $H_0$ more aggressively than visual tests.
-In this section, we explore how often they agree with each other by investigating the rejections for the two model designs based on $p$-values for each lineup.
-
-Figure \ref{fig:p-value-comparison} provides a mosaic plot showing the rejection rate of visual tests and conventional tests for both non-linearity patterns and heteroskedasticity patterns.
-
-For lineups containing non-linearity patterns, conventional tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% and visual tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the time.
-Of the lineups rejected by the conventional test, `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by the visual test, that is, approximately half as many as the conventional test.
-There are no lineups that are rejected by the visual test but not by the conventional test.
-
-In terms of lineups containing heteroskedasticity patterns, `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by conventional tests, while `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by visual tests.
-When the conventional test rejects a lineup, there is a great chance (`r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`%) the visual test will also reject it.
-
-```{r}
-nr <- p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Not") %>% count(reject)
-```
-
-Surprisingly, the visual test rejects `r nr$n[1]` of the `r nr$n[1]+nr$n[2]` (`r round(nr$n[1]/(nr$n[1]+nr$n[2])*100, 0)`%) of lineups where the conventional test does not reject.
-Figure \ref{fig:heter-example} shows this lineup.
-The data plot in position seventeen displays a relatively strong heteroskedasticity pattern, and has a strong effect size ($log_e E=4.02$).
-This is reflected by the visual test $p\text{-value} = 0.026$, but the Breusch-Pagan test $p\text{-value} = 0.056$, slightly above the significance cutoff of $0.05$.
-This lineup was evaluated by 11 subjects, it has experimental factors $n=50$ (small sample size), $\sigma=1$, and a uniform distribution for the fitted values.
-It must be the small sample size that may have resulted in the lack of detection.
-
 ```{r heter-example, fig.height = 7, fig.width = 7, fig.cap = 'The single heteroskedasticity lineup that is rejected by the visual test but not by the BP test.
-The data plot at panel 17 contains a "Butterfly" shape.
+The data plot at panel 17 contains a "butterfly" shape.
 It has effect size $ = 3.76$, somewhat surprising that it is not detected by the BP test.'}
@@
 A primary factor contributes to the non-linearity model is the shape of the non-linearity pattern.
-According to Figure \ref{fig:poly-power-uniform-j}, conventional tests have roughly the same power in testing shapes constructed from low orders of Hermite polynomials.
-But the power in testing the "Triple-U" shape drops significantly.
+According to Figure \ref{fig:poly-power-uniform-j}, conventional tests have higher power in testing shapes constructed from lower orders of Hermite polynomials.
+But the power of testing the "triple-U" shape drops significantly.
 To understand why this is, one needs to return to the way the RESET test is applied.
 It requires a parameter indicating degree of fitted values to test for, and the recommendation is to generically use four (@ramsey_tests_1969).
-However, the "Triple-U" shape constructed from the Hermite polynomials use power up to 18.
-If the RESET test had been applied using a higher power no less than six, the power curve of "Triple-U" shape will only slight lower than the power curve of "M" shape.
+However, the "triple-U" shape constructed from the Hermite polynomials use power up to 18.
+If the RESET test had been applied using a higher power no less than six, the power curve of "triple-U" shape will be only slight lower than the power curve of "M" shape.
 The recommendation of the polynomial power for the RESET should be revised, perhaps.
@@
 
-For visual tests, based on the orders of the Hermite polynomials, we expect the "U" shape will be the easiest one to be detected by subjects followed by the "S", "M" and "Triple-U" shape.
-From Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power curves are mostly aligned with our expectation, except for the "M" shape, which is the easiest shape to be detected.
-This implies the visibility of the shape do not strictly follow the degree of the polynomials.
+For visual tests, based on the orders of the Hermite polynomials, we expect the "U" shape will be the easiest one to be detected by subjects followed by the "S", "M" and "triple-U" shape.
+From Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power curves are mostly aligned with our expectation, except for the "M" shape, which is as easy to be detected as the "S" shape.
+This implies, unlike the conventional test, the visibility of the shape do not strictly follow the degree of the polynomials.
 
@@
 Power curves with higher order of non-linearity are drawn with deeper colours.
-The default RESET tests under-perform significantly in detecting "Triple-U" shape.
+The default RESET tests under-perform significantly in detecting "triple-U" shape.
 To achieve a similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test.
@@
   mutate(reject = as.numeric(RESET4_p_value <= 0.05)) %>%
-  mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>%
+  mutate(shape = c("U", "S", "M", "triple-U")[shape]) %>%
   mutate(conv_or_not = "Conventional") %>%
@@
   mutate(reject = as.numeric(p_value <= 0.05)) %>%
-  mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>%
+  mutate(shape = c("U", "S", "M", "triple-U")[shape]) %>%
   mutate(conv_or_not = "Visual")) %>%
-  mutate(shape = factor(shape, levels = c("U", "S", "M", "Triple-U"))) %>%
+  mutate(shape = factor(shape, levels = c("U", "S", "M", "triple-U"))) %>%
   mutate(offset0 = log(0.05/0.95)) %>%
@@
 We have also investigated the impact of different heteroskedasticity shapes on power of conventional tests and visual tests.
-In theory, the "Left-triangle" and the "Right-triangle" shapes are functionally identical from the point of view of a Breusch-Pagan test.
+In theory, the "left-triangle" and the "right-triangle" shapes are functionally identical from the point of view of a Breusch-Pagan test.
 As shown in Figure \ref{fig:heter-power-uniform-a}, this is indeed the case where little difference between the power curves can be perceived.
 Similarly, visual tests should have the same power in detecting these two shapes if they are equally likely to be identified.
-However, it can be observed from Figure \ref{fig:heter-power-uniform-a} that the power curve for the "Left-triangle" shape is higher than the one for the "Right-triangle" shape, indicating a potential favour of orientation by human, which is worth to be explored in future studies.
-Besides, the chance of detecting the "Butterfly" shape by both conventional and visual tests are much higher than other two shapes.
+However, it can be observed from Figure \ref{fig:heter-power-uniform-a} that the power curve for the "left-triangle" shape is higher than the one for the "right-triangle" shape, indicating a potential favour of orientation by human, which is worth to be explored in future studies.
+Besides, the chance of detecting the "butterfly" shape by both conventional and visual tests are much higher than other two shapes.
 
@@
   mutate(reject = as.numeric(BP_p_value <= 0.05)) %>%
-  mutate(a = c("Left-triangle", "Butterfly", "Right-triangle")[a + 2]) %>%
+  mutate(a = c("left-triangle", "butterfly", "right-triangle")[a + 2]) %>%
   mutate(conv_or_not = "Conventional") %>%
@@
   mutate(reject = as.numeric(p_value <= 0.05)) %>%
-  mutate(a = c("Left-triangle", "Butterfly", "Right-triangle")[a + 2]) %>%
+  mutate(a = c("left-triangle", "butterfly", "right-triangle")[a + 2]) %>%
   mutate(conv_or_not = "Visual")) %>%
-  mutate(a = factor(a, levels = c("Left-triangle", "Butterfly", "Right-triangle"))) %>%
+  mutate(a = factor(a, levels = c("left-triangle", "butterfly", "right-triangle"))) %>%
   mutate(offset0 = log(0.05/0.95)) %>%

### 2023-03-07 paper_comparison/paper_comparison/paper_comparison.tex
Fix caption and polish effect of non-linearity shape

@@
 model can be simplified to a beta-binomial model and thus the visual
-p-value is given as
+\(p\)-value is given as
 
@@
 number of selections made by the subject for the calculation of the
-p-value (See Equation \ref{eq:pvalue-beta-binomial-approx}). Therefore,
-we directly estimate the probability of a lineup being rejected by
-assuming that individual skill has negligible effect on the variation of
-the power. This assumption is not necessary true, but it helps
-simplifying the model structure, thereby obviate a costly large-scale
-experiment to estimate complex covariance matrices. The same model given
-in Equation \ref{eq:logistic-regression-1-1} is applied to model the
-power of a visual test.
+\(p\)-value (See Equation \ref{eq:pvalue-beta-binomial-approx}).
+Therefore, we directly estimate the probability of a lineup being
+rejected by assuming that individual skill has negligible effect on the
+variation of the power. This assumption is not necessary true, but it
+helps simplifying the model structure, thereby obviate a costly
+large-scale experiment to estimate complex covariance matrices. The same
+model given in Equation \ref{eq:logistic-regression-1-1} is applied to
+model the power of a visual test.
 
@@
 are included in the residual plot. These include ``U'', ``S'', ``M'' and
-``Triple-U'' shape as shown in Figure
+``triple-U'' shape as shown in Figure
 \ref{fig:different-shape-of-herimite}. A greater value of \(j\) will
@@
 
-\caption{Examining the effect of $\sigma$ on the signal strength in the non-linearity detection, for $n=300$, uniform fitted value distribution and the U shape.
-As $\sigma$ increases the signal strength decreases, to the point that the U is almost unrecognisable when $\sigma=4$.}\label{fig:different-sigma}
+\caption{Examining the effect of $\sigma$ on the signal strength in the non-linearity detection, for $n=300$, uniform fitted value distribution and the "U" shape.
+As $\sigma$ increases the signal strength decreases, to the point that the "U" is almost unrecognisable when $\sigma=4$.}\label{fig:different-sigma}
 \end{figure}
@@
 \caption{Heteroskedasticity forms used in the experiment.
-Three different shapes ($a = -1, 0, 1$) are used in the experiment to create left-triangle, butterfly and right-triangle shapes, respectively.}\label{fig:different-shape-of-heter}
+Three different shapes ($a = -1, 0, 1$) are used in the experiment to create left-triangle, "butterfly" and "right-triangle" shapes, respectively.}\label{fig:different-shape-of-heter}
 \end{figure}
@@
 
-\caption{Examining the effect of signal strength for the three different values of $n$ used in the experiment, for non-linear structure with fixed $\sigma = 1.5$, uniform fitted value distribution, and S shape.
-For these factor levels, only when $n = 300$ is the S shape clearly visible.}\label{fig:different-n}
+\caption{Examining the effect of signal strength for the three different values of $n$ used in the experiment, for non-linear structure with fixed $\sigma = 1.5$, uniform fitted value distribution, and "S" shape.
+For these factor levels, only when $n = 300$ is the "S" shape clearly visible.}\label{fig:different-n}
 \end{figure}
@@
 
+(should the below p goes to conclusion as a discussion?)
+
 Overall, the power comparison suggests that conventional tests differs
 significantly from visual tests in two regression diagnostics scenarios
-designed in this paper. Visual test have much higher tolerance of the
-residual departures than the conventional test. Since fail to reject
-\(H_0\) in a visual test literally means that there are no obvious
-visual discoveries found in the residual plot, analysts and the general
-public as the consumers of the output may not be convinced of the
-existence of significant residual departures in spite of the rejection
-of \(H_0\) given by the conventional test. Even if the rejection is
-accepted, the model violation may be considered as impactless due to the
-fact that they are not clearly visible. Besides, the sensitivity of the
+designed by us. Visual test have much higher tolerance of the residual
+departures than the conventional test. Since fail to reject \(H_0\) in a
+visual test literally means that there are no obvious visual discoveries
+found in the residual plot, analysts and the general public as the
+consumers of the output may not be convinced of the existence of
+significant residual departures in spite of the rejection of \(H_0\)
+given by the conventional test. Even if the rejection is accepted, the
+model violation may be considered as impactless due to the fact that
+they are not clearly visible. Besides, the sensitivity of the
 conventional test to weak residual departures could also distract and
@@
 Surrounding lines of the visual test show the estimated power of 500 bootstrap samples.
-Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot.
+Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by vertical lines in main plot.
 Where would you draw the line of too much non-linearity in the residuals?
-For the RESET test this is around log effect size 2.5, but for the visual test it is around 3.5.}\label{fig:polypower}
+For the RESET test this is around log effect size 2, but for the visual test it is around 3.5.}\label{fig:polypower}
 \end{figure}
@@
 \(p\)-values}{Comparison of test decisions based on p-values}}\label{comparison-of-test-decisions-based-on-p-values}}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/p-value-comparison-1}
-
-}
-
-\caption{Rejection rate ($p$-value $\leq0.05$) of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot.
-The visual test rejects less frequently than the conventional test.
-We would generally expect that the visual test would only reject when the conventional test does.
-Surprisingly, there is one lineup in the heteroskedasticity study where this is not the case.
-}\label{fig:p-value-comparison}
-\end{figure}
 
@@
 relatively strong heteroskedasticity pattern, and has a strong effect
-size (\(log_e E=4.02\)). This is reflected by the visual test
-\(p\text{-value} = 0.026\), but the Breusch-Pagan test
-\(p\text{-value} = 0.056\), slightly above the significance cutoff of
+size (\(log_e(E)=4.02\)). This is reflected by the visual test
+\(p\text{-value} = 0.026\). But the Breusch-Pagan test
+\(p\text{-value} = 0.056\), is slightly above the significance cutoff of
 \(0.05\). This lineup was evaluated by 11 subjects, it has experimental
-factors \(n=50\) (small sample size), \(\sigma=1\), and a uniform
-distribution for the fitted values. It must be the small sample size
-that may have resulted in the lack of detection.
+factors \(a = 0\) (``butterfly'' shape), \(b = 64\) (large variance
+ratio), \(n = 50\) (small sample size), and a uniform distribution for
+the fitted values. It must be the small sample size that may have
+resulted in the lack of detection.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/p-value-comparison-1}
+
+}
+
+\caption{Rejection rate ($p$-value $\leq0.05$) of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot.
+The visual test rejects less frequently than the conventional test.
+We would generally expect the visual test to only reject when the conventional test does.
+Surprisingly, one lineup containing a heteroskedasticity pattern does not follow this rule.
+}\label{fig:p-value-comparison}
+\end{figure}
 
@@
 \caption{The single heteroskedasticity lineup that is rejected by the visual test but not by the BP test.
-The data plot at panel 17 contains a "Butterfly" shape.
+The data plot at panel 17 contains a "butterfly" shape.
 It has effect size $ = 3.76$, somewhat surprising that it is not detected by the BP test.}\label{fig:heter-example}
@@
 the non-linearity pattern. According to Figure
-\ref{fig:poly-power-uniform-j}, conventional tests have roughly the same
-power in testing shapes constructed from low orders of Hermite
-polynomials. But the power in testing the ``Triple-U'' shape drops
-significantly. To understand why this is, one needs to return to the way
-the RESET test is applied. It requires a parameter indicating degree of
-fitted values to test for, and the recommendation is to generically use
-four (\citet{ramsey_tests_1969}). However, the ``Triple-U'' shape
-constructed from the Hermite polynomials use power up to 18. If the
-RESET test had been applied using a higher power no less than six, the
-power curve of ``Triple-U'' shape will only slight lower than the power
-curve of ``M'' shape. The recommendation of the polynomial power for the
-RESET should be revised, perhaps. This illustrates the sensitivity of
+\ref{fig:poly-power-uniform-j}, conventional tests have higher power in
+testing shapes constructed from lower orders of Hermite polynomials. But
+the power of testing the ``triple-U'' shape drops significantly. To
+understand why this is, one needs to return to the way the RESET test is
+applied. It requires a parameter indicating degree of fitted values to
+test for, and the recommendation is to generically use four
+(\citet{ramsey_tests_1969}). However, the ``triple-U'' shape constructed
+from the Hermite polynomials use power up to 18. If the RESET test had
+been applied using a higher power no less than six, the power curve of
+``triple-U'' shape will be only slight lower than the power curve of
+``M'' shape. The recommendation of the polynomial power for the RESET
+should be revised, perhaps. This illustrates the sensitivity of
 conventional testing to the parameters, and it also points to a
@@
 expect the ``U'' shape will be the easiest one to be detected by
-subjects followed by the ``S'', ``M'' and ``Triple-U'' shape. From
+subjects followed by the ``S'', ``M'' and ``triple-U'' shape. From
 Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power
 curves are mostly aligned with our expectation, except for the ``M''
-shape, which is the easiest shape to be detected. This implies the
-visibility of the shape do not strictly follow the degree of the
-polynomials.
+shape, which is as easy to be detected as the ``S'' shape. This implies,
+unlike the conventional test, the visibility of the shape do not
+strictly follow the degree of the polynomials.
 
@@
 shapes on power of conventional tests and visual tests. In theory, the
-``Left-triangle'' and the ``Right-triangle'' shapes are functionally
+``left-triangle'' and the ``right-triangle'' shapes are functionally
 identical from the point of view of a Breusch-Pagan test. As shown in
@@
 from Figure \ref{fig:heter-power-uniform-a} that the power curve for the
-``Left-triangle'' shape is higher than the one for the
-``Right-triangle'' shape, indicating a potential favour of orientation
+``left-triangle'' shape is higher than the one for the
+``right-triangle'' shape, indicating a potential favour of orientation
 by human, which is worth to be explored in future studies. Besides, the
-chance of detecting the ``Butterfly'' shape by both conventional and
+chance of detecting the ``butterfly'' shape by both conventional and
 visual tests are much higher than other two shapes.

### 2023-03-07 paper_comparison/paper_comparison/paper_comparison.Rmd
update heter power cmp, and fig caption

@@
 However, it can be observed from Figure \ref{fig:heter-power-uniform-a} that the power curve for the "left-triangle" shape is higher than the one for the "right-triangle" shape, indicating a potential favour of orientation by human, which is worth to be explored in future studies.
-Besides, the chance of detecting the "butterfly" shape by both conventional and visual tests are much higher than other two shapes.
-
-```{r heter-power-uniform-a, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$."}
+Besides, the chance of detecting the "butterfly" shape are higher than other two shapes in both conventional and visual testing.
+
+```{r heter-power-uniform-a, fig.height = 4, fig.cap = 'Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$.
+The visual test has an asymmetric power in testing the "left-triangle" shape and "right-triangle" shape, while these two shapes are equivalent in conventional testing.'}
 

### 2023-03-07 paper_comparison/paper_comparison/paper_comparison.tex
update heter power cmp, and fig caption

@@
 by human, which is worth to be explored in future studies. Besides, the
-chance of detecting the ``butterfly'' shape by both conventional and
-visual tests are much higher than other two shapes.
+chance of detecting the ``butterfly'' shape are higher than other two
+shapes in both conventional and visual testing.
 
@@
 
-\caption{Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$.}\label{fig:heter-power-uniform-a}
+\caption{Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model with three different shapes controlled by the parameter $a$.
+The visual test has an asymmetric power in testing the "left-triangle" shape and "right-triangle" shape, while these two shapes are equivalent in conventional testing.}\label{fig:heter-power-uniform-a}
 \end{figure}

### 2023-03-07 paper_comparison/paper_comparison/paper_comparison.Rmd
finish result

@@
 According to Figure \ref{fig:poly-power-uniform-j}, conventional tests have higher power in testing shapes constructed from lower orders of Hermite polynomials.
-But the power of testing the "triple-U" shape drops significantly.
+But the chance of detecting the "triple-U" shape drops significantly compared to other shapes.
 To understand why this is, one needs to return to the way the RESET test is applied.
@@
 If the RESET test had been applied using a higher power no less than six, the power curve of "triple-U" shape will be only slight lower than the power curve of "M" shape.
-The recommendation of the polynomial power for the RESET should be revised, perhaps.
+The recommendation of the polynomial power for the RESET test should be revised, perhaps.
 This illustrates the sensitivity of conventional testing to the parameters, and it also points to a limitation that one needs to know the data generating process in order to set the parameters for the test.
@@
 Similarly, visual tests should have the same power in detecting these two shapes if they are equally likely to be identified.
-However, it can be observed from Figure \ref{fig:heter-power-uniform-a} that the power curve for the "left-triangle" shape is higher than the one for the "right-triangle" shape, indicating a potential favour of orientation by human, which is worth to be explored in future studies.
+However, it can be observed from Figure \ref{fig:heter-power-uniform-a} that the power curve of the "left-triangle" shape is higher than the one of the "right-triangle" shape, indicating a potential favour of orientation by human, which is worth to be explored in future studies.
 Besides, the chance of detecting the "butterfly" shape are higher than other two shapes in both conventional and visual testing.
@@
 
-Figure \ref{fig:different-x-dist-poly-power} illustrates the change of power of visual tests and conventional tests for different fitted distributions used in the non-linearity model and the heteroskedasticity model.
+Figure \ref{fig:different-x-dist-poly-power} illustrates the change of power of visual tests for different fitted distributions used in the non-linearity model and the heteroskedasticity model.
 Note that we have collected more evaluations on lineup with uniform fitted value distribution, and the number of evaluations on a lineup will have an impact on the power of the visual test.
 To have a fair comparison, the five evaluations used in the power estimation are randomly sampled from the total eleven evaluations for each lineup with uniform fitted value distribution.
-The sampling process has been repeated 500 times to visualize the variation.
-
-For visual tests, the power curves for normal distribution and lognormal distribution are lower than other two distributions, which are all even.
-For non-linearity patterns, the power curve of normal distribution is slightly greater than the one for lognormal distribution.
-And the margin is even larger for heteroskedasticity patterns.
-This suggests skewness can affect the degree of presence of the visual pattern.
-
-For non-linearity patterns, we do not observe significant difference in power of visual test between uniform distribution and discrete uniform distribution, but the normal distribution and the lognoraml distribution have lower power than the previous two.
-In terms of conventional tests, the discrete uniform distribution is the one with the greatest power, followed by lognormal, uniform and the normal distribution.
-
-For heteroskedasticity patterns, the power of visual test on lineups with uniform distribution has the greatest power.
-This is as expected since other three distributions could reduce the chance of revealing the underlying visual pattern because of uneven data points.
-The power of visual test under the discrete uniform distribution has a relatively great.
-Considering for heteroskedasticity, the visual patterns are usually detected by connecting the maximum and minimum residuals separately at different fitted values.
-It will not be greatly affected by the discrete uniform distribution compared to the uniform distribution.
-Again, visual tests have low power on lineups with the normal distribution and the lognormal distribution.
-It indicates skewness and unevenness can reduce the detection rate of the visual pattern.
-In the case of conventional tests, the discrete uniform distribution consistently perform well, followed by uniform, normal and the lognormal distribution.
-
-In regards of the type of the simulation model, the power of visual tests under different fitted value distributions are substantially lower than the corresponding conventional tests, which is consistent with the result reported in the previous sections.
+The sampling process has been repeated 500 times to give an indication of the variation of the power.
+
+We do not observe significant difference between the power curve of the uniform distribution and the power curve of the discrete uniform distribution, but the power curves of normal and lognormal distributions are slightly lower.
+Same can be observed for heteroskedasticity patterns, where the gap between the power curve of the discrete uniform and the power curve of the normal distribution is even larger.
+This indicates that residual departures displayed with evenly distributed fitted values are visually more attractive.
+Besides, the power of detecting non-linearity or heteroskedasticity patterns from a lineup with lognormal fitted value distribution is the lowest among four distributions.
+This suggests the skewness of the fitted value distribution plays a role in the degree of presence of visual features, and a symmetric fitted value distribution can better reveal the underlying shape.
 
@@
 
-```{r different-x-dist-poly-power, fig.height = 4, fig.cap = "Comparison of power on lineups with different fitted value distributions, for conventional tests and visual tests.
-For visual tests, all the power curves are estimated using five evaluations on each lineup.
+```{r different-x-dist-poly-power, fig.height = 4, fig.cap = "Comparison of power on lineups with different fitted value distributions for visual tests.
+All the power curves are estimated using five evaluations on each lineup.
 For lineups with uniform fitted value distribution, the five evaluations are randomly sampled from the total eleven evaluations.
 The sampling process has been repeated for 500 times and the corresponding power curves are drawn with transparent grey lines.
-For both types of tests power does change.
-The discrete fitted value distribution substantially lowers the power of visual tests for non-linearity patterns in residual plots.
-For heteroskedasticity patterns, normal and lognormal distributions produce lower power for both types of tests.", dev = 'png', dpi = 450}
+Uneven distributions including normal and lognormal distributions produce lower power for both types of residual departures.
+Meanwhile, the lognormal distribution as an asymmetric distribution produce the lowest power.", dev = 'png', dpi = 450}
 
@@
     mutate(x_dist = ifelse(x_dist == "even_discrete", 
-                           "Discrete", 
+                           "Discrete uniform", 
                            stringr::str_to_title(x_dist))) %>%
@@
                                     "Lognormal", 
-                                    "Discrete")))
+                                    "Discrete uniform")))
 
@@
   labs(col = "", linetype = "Test") +
-  facet_grid(~type) +
+  facet_grid(~type, scales = "free_x") +
   theme(legend.position = "bottom")

### 2023-03-07 paper_comparison/paper_comparison/paper_comparison.tex
finish result

@@
 testing shapes constructed from lower orders of Hermite polynomials. But
-the power of testing the ``triple-U'' shape drops significantly. To
-understand why this is, one needs to return to the way the RESET test is
-applied. It requires a parameter indicating degree of fitted values to
-test for, and the recommendation is to generically use four
-(\citet{ramsey_tests_1969}). However, the ``triple-U'' shape constructed
-from the Hermite polynomials use power up to 18. If the RESET test had
-been applied using a higher power no less than six, the power curve of
-``triple-U'' shape will be only slight lower than the power curve of
-``M'' shape. The recommendation of the polynomial power for the RESET
-should be revised, perhaps. This illustrates the sensitivity of
-conventional testing to the parameters, and it also points to a
-limitation that one needs to know the data generating process in order
-to set the parameters for the test.
+the chance of detecting the ``triple-U'' shape drops significantly
+compared to other shapes. To understand why this is, one needs to return
+to the way the RESET test is applied. It requires a parameter indicating
+degree of fitted values to test for, and the recommendation is to
+generically use four (\citet{ramsey_tests_1969}). However, the
+``triple-U'' shape constructed from the Hermite polynomials use power up
+to 18. If the RESET test had been applied using a higher power no less
+than six, the power curve of ``triple-U'' shape will be only slight
+lower than the power curve of ``M'' shape. The recommendation of the
+polynomial power for the RESET test should be revised, perhaps. This
+illustrates the sensitivity of conventional testing to the parameters,
+and it also points to a limitation that one needs to know the data
+generating process in order to set the parameters for the test.
 
@@
 they are equally likely to be identified. However, it can be observed
-from Figure \ref{fig:heter-power-uniform-a} that the power curve for the
-``left-triangle'' shape is higher than the one for the
-``right-triangle'' shape, indicating a potential favour of orientation
-by human, which is worth to be explored in future studies. Besides, the
-chance of detecting the ``butterfly'' shape are higher than other two
-shapes in both conventional and visual testing.
+from Figure \ref{fig:heter-power-uniform-a} that the power curve of the
+``left-triangle'' shape is higher than the one of the ``right-triangle''
+shape, indicating a potential favour of orientation by human, which is
+worth to be explored in future studies. Besides, the chance of detecting
+the ``butterfly'' shape are higher than other two shapes in both
+conventional and visual testing.
 
@@
 Figure \ref{fig:different-x-dist-poly-power} illustrates the change of
-power of visual tests and conventional tests for different fitted
-distributions used in the non-linearity model and the heteroskedasticity
-model. Note that we have collected more evaluations on lineup with
-uniform fitted value distribution, and the number of evaluations on a
-lineup will have an impact on the power of the visual test. To have a
-fair comparison, the five evaluations used in the power estimation are
-randomly sampled from the total eleven evaluations for each lineup with
-uniform fitted value distribution. The sampling process has been
-repeated 500 times to visualize the variation.
-
-For visual tests, the power curves for normal distribution and lognormal
-distribution are lower than other two distributions, which are all even.
-For non-linearity patterns, the power curve of normal distribution is
-slightly greater than the one for lognormal distribution. And the margin
-is even larger for heteroskedasticity patterns. This suggests skewness
-can affect the degree of presence of the visual pattern.
-
-For non-linearity patterns, we do not observe significant difference in
-power of visual test between uniform distribution and discrete uniform
-distribution, but the normal distribution and the lognoraml distribution
-have lower power than the previous two. In terms of conventional tests,
-the discrete uniform distribution is the one with the greatest power,
-followed by lognormal, uniform and the normal distribution.
-
-For heteroskedasticity patterns, the power of visual test on lineups
-with uniform distribution has the greatest power. This is as expected
-since other three distributions could reduce the chance of revealing the
-underlying visual pattern because of uneven data points. The power of
-visual test under the discrete uniform distribution has a relatively
-great. Considering for heteroskedasticity, the visual patterns are
-usually detected by connecting the maximum and minimum residuals
-separately at different fitted values. It will not be greatly affected
-by the discrete uniform distribution compared to the uniform
-distribution. Again, visual tests have low power on lineups with the
-normal distribution and the lognormal distribution. It indicates
-skewness and unevenness can reduce the detection rate of the visual
-pattern. In the case of conventional tests, the discrete uniform
-distribution consistently perform well, followed by uniform, normal and
-the lognormal distribution.
-
-In regards of the type of the simulation model, the power of visual
-tests under different fitted value distributions are substantially lower
-than the corresponding conventional tests, which is consistent with the
-result reported in the previous sections.
+power of visual tests for different fitted distributions used in the
+non-linearity model and the heteroskedasticity model. Note that we have
+collected more evaluations on lineup with uniform fitted value
+distribution, and the number of evaluations on a lineup will have an
+impact on the power of the visual test. To have a fair comparison, the
+five evaluations used in the power estimation are randomly sampled from
+the total eleven evaluations for each lineup with uniform fitted value
+distribution. The sampling process has been repeated 500 times to give
+an indication of the variation of the power.
+
+We do not observe significant difference between the power curve of the
+uniform distribution and the power curve of the discrete uniform
+distribution, but the power curves of normal and lognormal distributions
+are slightly lower. Same can be observed for heteroskedasticity
+patterns, where the gap between the power curve of the discrete uniform
+and the power curve of the normal distribution is even larger. This
+indicates that residual departures displayed with evenly distributed
+fitted values are visually more attractive. Besides, the power of
+detecting non-linearity or heteroskedasticity patterns from a lineup
+with lognormal fitted value distribution is the lowest among four
+distributions. This suggests the skewness of the fitted value
+distribution plays a role in the degree of presence of visual features,
+and a symmetric fitted value distribution can better reveal the
+underlying shape.
 
@@
 
-\caption{Comparison of power on lineups with different fitted value distributions, for conventional tests and visual tests.
-For visual tests, all the power curves are estimated using five evaluations on each lineup.
+\caption{Comparison of power on lineups with different fitted value distributions for visual tests.
+All the power curves are estimated using five evaluations on each lineup.
 For lineups with uniform fitted value distribution, the five evaluations are randomly sampled from the total eleven evaluations.
 The sampling process has been repeated for 500 times and the corresponding power curves are drawn with transparent grey lines.
-For both types of tests power does change.
-The discrete fitted value distribution substantially lowers the power of visual tests for non-linearity patterns in residual plots.
-For heteroskedasticity patterns, normal and lognormal distributions produce lower power for both types of tests.}\label{fig:different-x-dist-poly-power}
+Uneven distributions including normal and lognormal distributions produce lower power for both types of residual departures.
+Meanwhile, the lognormal distribution as an asymmetric distribution produce the lowest power.}\label{fig:different-x-dist-poly-power}
 \end{figure}

### 2023-03-07 paper_comparison/paper_comparison/paper_comparison.Rmd
finish conclusion

@@
 
-<!--1. summarise paper content-->
-
 Motivated by the advice of regression analysis experts, that residual plots as opposed to conventional tests are an indispensible methods for assessing model fit, a human subjects experiment was conducted using visual inference.
@@
 
-<!-- This paper has demonstrated that conventional residual-based statistical tests are more sensitive to departures from assumptions for residuals than visual tests evaluated by humans in the diagnostics of linear regression models.
-An experiment is conducted in three data collection periods using two simulation models, one for for non-linearity and and the other for heteroskedasticity.
-Methods for obtaining the $p$-value of visual tests proposed by @vanderplas2021statistical are extended to allow for multiple selections in a lineup protocol.-->
-
-<!--2. briefly describe what was learned-->
-
-The experiment found that conventional residual-based statistical tests are more sensitive to departures from assumptions for residuals than visual tests as would be evaluated by humans.
+The experiment found that conventional residual-based statistical tests are more sensitive to weak residual departures from model assumptions than visual tests as would be evaluated by humans.
 That is, conventional tests conclude there are problems with the model fit almost twice as often as humans would.
@@
 Now it is important to note that residual plots need to be delivered as a lineup, where it is embedded in a field of null plots.
+A residual plot may contains many visual features, but some are caused by the characteristics of the predictors and the randomness of the error, not by the violation of the model assumptions.
+These irrelevant visual features have a chance to be filtered out by subjects with a comparison to null plots, results in a set of more accurate visual findings.
 This enables a careful calibration for reading structure in residual plots.
-XXX More here
 
 However, human evaluation of residuals is expensive.
-It is time-consuming, and laborious.
+It is time-consuming, laborious and unfriendly to vision-impaired people.
 This is another reason why it often appears to be ignored.
 With the availability of sophisticated computer vision algorithms today, the goal of this work is to form the basis of providing automated residual plot reading.
-XXX More here
+The findings suggest the strong demand of graphical inspection in regression diagnostics, so developing an automatic visual inference system to evaluate lineups of residual plots is valuable.
+We plan to build a completed open-source system in an `R` package and provide a web interface such as a website for public to interact with.
+Details about this system will be discussed in our next paper.
 
 The experiment also revealed some interesting details about how residual plots are read.
-For the most part, the visual test performed very similarly to the (best) conventional test only with the power curve shifted in the less sensitive direction.
+For the most part, the visual test performed very similarly to the appropriate conventional test only with the power curve shifted in the less sensitive direction.
 Unlike the conventional tests, where one needs to specifically test for non-linearity or heteroskedasticity the visual test operated effectively across the range of departures from good residuals.
@@
 Structure is hardest to detect if fitted values are lognormal.
-Also, a surprising find was that the direction of heteroskedasticity appears to affect the ability to visually detect it, with wedge to the right being less detectable.
-
-<!--According to the data collected from the human subject experiment, residual plots rejected by visual tests are rarely not rejected by conventional tests, but departures of residuals that are of little practical sense to humans will often be alarmed by conventional tests, supporting the common guidance given by experts to utilize graphical representations of data in model diagnostics.
--->
-
-<!-- We have also found that residual-based conventional tests relying on a user-parameterized test formula, such as the RESET test, will suffer from a significant decrease in power for detecting certain residual departures intended to be tested but not covered by the test formula.
-This can not be avoided without prior knowledge of their functional forms, which is usually obtained by plotting the residuals.
-On the other hand, visual tests do not rely on such information and a collection of departures can be tested simultaneously.
-
-Interestingly, the orientation of the visual pattern
-
-fitted value distribution
-
-orientation of heter shape
--->
-
-<!-- Besides, the power of both tests vary for different fitted value distributions.  -->
-
-<!-- The ideal fitted value distribution for a visual test is the uniform distribution. If this can not be satisfied, a continuous -->
-
-<!-- continuous and symmetric -->
-
-<!-- distributions that are continuous, symmetric, or evenly distributed are also welcomed. -->
-
-<!-- properties that a uniform distribution holds, such as continuous, symmetric, and evenly distributed  -->
-
-<!-- Depending on the underlying residual departures embedded in the residual plot,  -->
-
-<!-- In the two controlled scenarios, the conventional tests favours discrete uniform distribution -->
-
-<!-- fitted value distributions have a strong impact on the  -->
-
-<!-- some residual departures are more difficult to be detected under a skewed fitted value distribution.  -->
-
-<!-- continuous, symmetric and evenly distributed. -->
-
-<!-- with different  -->
-
-<!-- symmetric, evenly  -->
-
-3. how does this apply more broadly than simple regression
-
-recommend residual vs fitted values
-a. measure uncertainty
-b. systemically miss some structure
-
-4. where does this lead? Computer vision residual plot reading
+Also, complex structure are generally harder to detect, but there are outliers.
+A surprising finding was that the direction of heteroskedasticity appears to affect the ability to visually detect it, with wedge to the right being less detectable.
 

### 2023-03-07 paper_comparison/paper_comparison/paper_comparison.tex
finish conclusion

@@
 The experiment found that conventional residual-based statistical tests
-are more sensitive to departures from assumptions for residuals than
-visual tests as would be evaluated by humans. That is, conventional
+are more sensitive to weak residual departures from model assumptions
+than visual tests as would be evaluated by humans. That is, conventional
 tests conclude there are problems with the model fit almost twice as
@@
 Now it is important to note that residual plots need to be delivered as
-a lineup, where it is embedded in a field of null plots. This enables a
-careful calibration for reading structure in residual plots. XXX More
-here
+a lineup, where it is embedded in a field of null plots. A residual plot
+may contains many visual features, but some are caused by the
+characteristics of the predictors and the randomness of the error, not
+by the violation of the model assumptions. These irrelevant visual
+features have a chance to be filtered out by subjects with a comparison
+to null plots, results in a set of more accurate visual findings. This
+enables a careful calibration for reading structure in residual plots.
 
 However, human evaluation of residuals is expensive. It is
-time-consuming, and laborious. This is another reason why it often
-appears to be ignored. With the availability of sophisticated computer
-vision algorithms today, the goal of this work is to form the basis of
-providing automated residual plot reading. XXX More here
+time-consuming, laborious and unfriendly to vision-impaired people. This
+is another reason why it often appears to be ignored. With the
+availability of sophisticated computer vision algorithms today, the goal
+of this work is to form the basis of providing automated residual plot
+reading. The findings suggest the strong demand of graphical inspection
+in regression diagnostics, so developing an automatic visual inference
+system to evaluate lineups of residual plots is valuable. We plan to
+build a completed open-source system in an \texttt{R} package and
+provide a web interface such as a website for public to interact with.
+Details about this system will be discussed in our next paper.
 
@@
 plots are read. For the most part, the visual test performed very
-similarly to the (best) conventional test only with the power curve
+similarly to the appropriate conventional test only with the power curve
 shifted in the less sensitive direction. Unlike the conventional tests,
@@
 loss of power in the visual test. Structure is hardest to detect if
-fitted values are lognormal. Also, a surprising find was that the
-direction of heteroskedasticity appears to affect the ability to
+fitted values are lognormal. Also, complex structure are generally
+harder to detect, but there are outliers. A surprising finding was that
+the direction of heteroskedasticity appears to affect the ability to
 visually detect it, with wedge to the right being less detectable.
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{2}
-\tightlist
-\item
-  how does this apply more broadly than simple regression
-\end{enumerate}
-
-recommend residual vs fitted values a. measure uncertainty b.
-systemically miss some structure
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{3}
-\tightlist
-\item
-  where does this lead? Computer vision residual plot reading
-\end{enumerate}
 

### 2023-03-07 paper_comparison/appendix.Rmd
update power desc
update effect size desc
rollback to previous figures
finish KL formula in appendix

@@
 
-## Effect size derivation
+## Effect size derivation and caculation
 
-(discuss what happens if we switch to different effect size)
-
-Effect size can be defined as the difference of a parameter for a particular model or distribution,  or a statistic derived from a sample.
-Importantly, it needs to reflect the treatment we try to measure.
+Effect size is usually defined as the difference of a parameter or a statistic derived from a sample.
 Centred on a conventional statistical test, we usually can deduce the effect size from the test statistic by substituting the null parameter value.
@@
 
-$$KL = \frac{1}{2}\left(\boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\sigma^2)^{-1}\boldsymbol{\mu}_z\right).$$
+$$KL = \frac{1}{2}\left(\boldsymbol{\mu}_z^{T}(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right).$$
 
@@
 $$KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2))\right).$$
+Since we assume $\sigma = 1$ for the heteroskedasticity model, the final form of the formula is
 
-## $p$-value scatter plot
+$$KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right).$$
 
-```{r p-value-comparison-old, eval = FALSE, fig.height = 4, fig.cap = "Visual test p-value compared to conventional test p-value for lineups produced by heteroskedasticity and non-linearity model.
-The scatter plot is drawn on a square root scale.
-Every dot on the plot corresponds to a lineup.
-Most of the dots are fallen on the left of the y = x line indicating visual test p-value is almost always higher than conventional test p-value.
-The blue curve is a smoothing of the dots showing a positive trend.
-"}
-vi_survey %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
-  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-  left_join(vi_survey %>% 
-           count(unique_lineup_id) %>%
-           rename(num_subject = n)) %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
-  ggplot() +
-  geom_abline() +
-  geom_point(aes(conventional_p_value, p_value), alpha = 0.4) +
-  geom_smooth(aes(conventional_p_value, p_value), se = FALSE, span = 0.6) +
-  scale_x_sqrt() +
-  scale_y_sqrt() +
-  facet_wrap(~type) +
-  ylab("Visual test p-value on square root scale") +
-  xlab("Conventional test p-value on square root scale") +
-  theme_light()
-```
+## Experiment setup
 
-## A collection interesting lineups (unusual results)
+### Mapping of subjects to experimental factors
 
-why unusual? what is the possible explanations?
+### Demographics
 
-# targe journal
+### Data collection interface
 
-JRSSB: Journal of the Royal Statistical Society Series B (Statistical Methodology)
-Deadline: Jan 1, Apr 1
+## Batch effect size
 
-Reading: style of writing, author guideline (https://rss.onlinelibrary.wiley.com/hub/journal/14679868/author-guidelines)
+## Sensitivity analysis for $\alpha$
 
-JCGS
-JDSS
+## Other results

### 2023-03-07 paper_comparison/appendix.Rmd
finish KL in appendix

@@
 
-## Effect size derivation and caculation
+## Effect size
+
+## Derivation
 
@@
 
-$$KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}\boldsymbol{\mu}_z\right).$$
+\begin{equation}
+\label{eq:effect-size}
+KL = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}\boldsymbol{\mu}_z\right).
+\end{equation}
 
@@
 
+## Caculation
+
+The effect size can be calculated using Equation \ref{eq:effect-size} for each lineup by plugging in the data generated from the simulated model.
+However, since the term $\boldsymbol{R}_a$ will differ for each sample, the effect size will be different even if the parameter values are the same.
+To overcome this undesired property, for the same set of parameter values, we simulate a sufficient large number of samples, and take the average of the effect size as the final effect size.
+
 ## Experiment setup
@@
 ### Mapping of subjects to experimental factors
+
+Mapping of subjects to experimental factors
 

### 2023-03-07 paper_comparison/appendix.Rmd
fill in subject mapping

@@
 
-Mapping of subjects to experimental factors
+Mapping of subjects to experimental factors is an important part of experiment design.
+Essentially, we want to maximum the difference in factors exposed to a subject.
+For this purpose, we design an algorithm to conduct subject allocation.
+Let $L$ be a set of available lineups and $S$ be a set of available subjects.
+According to the experimental design, available means a lineup can be allocated to at most five different subjects (for lineups with uniform fitted value distribution, this value is 11) and a subject can view at most 18 different lineups.
+
+The algorithm starts from picking a random subject $s$ from $S$ with the minimum number of allocated lineups.
+It then computes a metric for every lineup $l$ in $L$ indicating the distance $D$ between lineup $l$ to the lineups already allocated to this subject $l_{allocated}$.
+Lineup with the highest distance $D$ will be allocated to the subject.
+The algorithm will repeatedly pick a subject, compute metrics, and allocate lineups until there is no available lineups or no available subjects.
+
+The distance $D$ is defined using the factor values used to generate lineup $l$ and lineups $l_{allocated}$.
+If there does not exist $l_{allocated}$, then the distance $D = 0$.
+Otherwise, distance $D$ is roughly the number of factors minus the number of factor values of lineup $l$ exists in a union of factor values of lineups $l_{allocated}$.
+The distance $D$ is further adjusted by the existence of $n$-factor combinations, where $n$ could be two to the number of factors.
+In this way, all factors values are exposed to every subject, and some even ensure all $2$-factor combinations are exposed to every subject.
 

### 2023-03-07 paper_comparison/appendix.Rmd
fill in basic demographic data

@@
 library(visage)
+library(tidyverse)
 ```
@@
 
+The education background of most participants have Diploma or Bachelor degrees, followed by High school or below. The survey data is gender balanced. Most of the participants are between 18 to 39.
+
+```{r}
+vi_survey %>%
+  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
+  ggplot() +
+  geom_bar(aes(age_group, fill = pronoun), position = "dodge") +
+  facet_wrap(~fct_relevel(education, "High School or below", "Diploma and Bachelor Degree", "Honours Degree", "Masters Degree")) +
+  xlab("Age group") +
+  ylab("Count")
+```
+
+Number of participants who have previous experience in experiment about reading data plots are not very different from the number of participants who haven’t.
+Age distributions are also similar for these two groups.
+
+```{r}
+vi_survey %>%
+  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
+  ggplot() +
+  geom_bar(aes(age_group, fill = pronoun), position = "dodge") +
+  facet_wrap(~previous_experience) +
+  xlab("Age group") +
+  ylab("Count")
+```
+
 ### Data collection interface

### 2023-03-07 paper_comparison/appendix.Rmd
fill in alpha analysis

@@
 library(tidyverse)
+```
+
+```{r}
+# OOP supports needed by `visage`
+# remotes::install_github("TengMCing/bandicoot")
+# 
+# Visual inference models and p-value calculation
+# remotes::install_github("TengMCing/visage")
+
+library(tidyverse)
+library(visage)
+
+# To control the simulation in this file
+set.seed(10086)
+```
+
+```{r get-lineup-data}
+# Create the dir folder
+if (!dir.exists(here::here("data/"))) dir.create(here::here("data/"))
+
+# The lineup data used to draw residual plot needed to be downloaded
+# from the github repo. Cache it.
+if (!file.exists(here::here("data/vi_lineup.rds"))) {
+  vi_lineup <- get_vi_lineup()
+  saveRDS(vi_lineup, here::here("data/vi_lineup.rds"))
+} else {
+  vi_lineup <- readRDS(here::here("data/vi_lineup.rds"))
+}
+```
+
+```{r poly-conventional-simulation}
+# Ensure the support of the predictor is [-1, 1]
+stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
+
+# Run simulations to get the behaviours of conventional tests
+if (!file.exists(here::here("data/poly_conventional_simulation.rds"))) {
+  poly_conv_sim <- list()
+
+  # 100 bootstrap samples
+  for (i in 1:100)
+  {
+    poly_conv_sim[[i]] <- 
+      # Every sample contains 2000 lineups
+      map(1:2000, function(i) {
+
+        # Sample a set of parameters
+        shape <- sample(1:4, 1)
+        e_sigma <- sample(c(0.5, 1, 2, 4), 1)
+        x_dist <- sample(c("uniform", 
+                           "normal", 
+                           "lognormal", 
+                           "even_discrete"), 1)
+        x <- switch(x_dist,
+                    uniform = rand_uniform(-1, 1),
+                    normal = {
+                      raw_x <- rand_normal(sigma = 0.3)
+                      closed_form(~stand_dist(raw_x))
+                      },
+                    lognormal = {
+                      raw_x <- rand_lognormal(sigma = 0.6)
+                      closed_form(~stand_dist(raw_x/3 - 1))
+                      },
+                    even_discrete = rand_uniform_d(k = 5, even = TRUE))
+
+        # Build the model
+        mod <- poly_model(shape, x = x, sigma = e_sigma)
+
+        # Sample the number of observations
+        n <- sample(c(50, 100, 300), 1)
+
+        # Generate data from the model 
+        tmp_dat <- mod$gen(n)
+
+        # Return a data frame containing p-values of
+        # F, RESET, BP and SW tests
+        tibble(shape = shape,
+               e_sigma = e_sigma,
+               x_dist = x_dist,
+               n = n,
+               F_p_value = mod$test(tmp_dat)$p_value,
+               RESET3_p_value = mod$test(tmp_dat, 
+                                         test = "RESET", 
+                                         power = 2:3, 
+                                         power_type = "fitted")$p_value,
+               RESET4_p_value = mod$test(tmp_dat, 
+                                         test = "RESET", 
+                                         power = 2:4, 
+                                         power_type = "fitted")$p_value,
+               RESET5_p_value = mod$test(tmp_dat, 
+                                         test = "RESET", 
+                                         power = 2:5, 
+                                         power_type = "fitted")$p_value,
+               RESET6_p_value = mod$test(tmp_dat, 
+                                         test = "RESET", 
+                                         power = 2:6, 
+                                         power_type = "fitted")$p_value,
+               RESET7_p_value = mod$test(tmp_dat, 
+                                         test = "RESET", 
+                                         power = 2:7, 
+                                         power_type = "fitted")$p_value,
+               RESET8_p_value = mod$test(tmp_dat, 
+                                         test = "RESET", 
+                                         power = 2:8, 
+                                         power_type = "fitted")$p_value,
+               RESET9_p_value = mod$test(tmp_dat, 
+                                         test = "RESET", 
+                                         power = 2:9, 
+                                         power_type = "fitted")$p_value,
+               RESET10_p_value = mod$test(tmp_dat, 
+                                          test = "RESET", 
+                                          power = 2:10, 
+                                          power_type = "fitted")$p_value,
+               BP_p_value = HETER_MODEL$test(tmp_dat)$p_value,
+               SW_p_value = shapiro.test(tmp_dat$.resid)$p.value,
+               boot_id = i)
+        }) %>%
+  reduce(bind_rows)
+  }
+
+  poly_conv_sim <- poly_conv_sim %>%
+    reduce(bind_rows)
+
+  saveRDS(poly_conv_sim, here::here("data/poly_conventional_simulation.rds"))
+} else {
+  poly_conv_sim <- readRDS(here::here("data/poly_conventional_simulation.rds"))
+}
+```
+
+```{r heter-conventional-simulation}
+# Run simulations to get the behaviours of conventional tests
+if (!file.exists(here::here("data/heter_conventional_simulation.rds"))) {
+  heter_conv_sim <- list()
+
+  # 100 bootstrap samples
+  for (i in 1:100)
+  {
+    heter_conv_sim[[i]] <-
+      # Every sample contains 2000 lineups
+      map(1:2000, function(x) {
+
+        # Sample a set of parameters
+        a <- sample(c(-1, 0, 1), 1)
+        b <- sample(c(0.25, 1, 4, 16, 64), 1)
+        x_dist <- sample(c("uniform", 
+                           "normal", 
+                           "lognormal", 
+                           "even_discrete"), 1)
+        x <- switch(x_dist,
+                    uniform = rand_uniform(-1, 1),
+                    normal = {
+                      raw_x <- rand_normal(sigma = 0.3)
+                      closed_form(~stand_dist(raw_x))
+                      },
+                    lognormal = {
+                      raw_x <- rand_lognormal(sigma = 0.6)
+                      closed_form(~stand_dist(raw_x/3 - 1))
+                      },
+                    even_discrete = rand_uniform_d(-1, 1, k = 5, even = TRUE))
+
+        # Build the model
+        mod <- heter_model(a = a, b = b, x = x)
+
+        # Sample the number of observations
+        n <- sample(c(50, 100, 300), 1)
+
+        # Generate data from the model 
+        tmp_dat <- mod$gen(n)
+
+        # Return a data frame containing p-values of
+        # F, RESET, BP and SW tests
+        tibble(a = a,
+               b = b,
+               x_dist = x_dist,
+               n = n,
+               F_p_value = POLY_MODEL$test(
+
+                 # Create a pseudo z to be able to use F-test
+                 tmp_dat %>%
+                   mutate(z = poly_model()$
+                            gen(n, computed = select(tmp_dat, x)) %>%
+                            pull(z))
+                 )$p_value,
+               RESET3_p_value = POLY_MODEL$test(tmp_dat, 
+                                                test = "RESET", 
+                                                power = 2:3, 
+                                                power_type = "fitted")$p_value,
+               RESET4_p_value = POLY_MODEL$test(tmp_dat, 
+                                                test = "RESET", 
+                                                power = 2:4, 
+                                                power_type = "fitted")$p_value,
+               RESET5_p_value = POLY_MODEL$test(tmp_dat, 
+                                                test = "RESET", 
+                                                power = 2:5, 
+                                                power_type = "fitted")$p_value,
+               RESET6_p_value = POLY_MODEL$test(tmp_dat, 
+                                                test = "RESET", 
+                                                power = 2:6, 
+                                                power_type = "fitted")$p_value,
+               RESET7_p_value = POLY_MODEL$test(tmp_dat, 
+                                                test = "RESET", 
+                                                power = 2:7, 
+                                                power_type = "fitted")$p_value,
+               RESET8_p_value = POLY_MODEL$test(tmp_dat, 
+                                                test = "RESET", 
+                                                power = 2:8, 
+                                                power_type = "fitted")$p_value,
+               RESET9_p_value = POLY_MODEL$test(tmp_dat, 
+                                                test = "RESET", 
+                                                power = 2:9, 
+                                                power_type = "fitted")$p_value,
+               RESET10_p_value = POLY_MODEL$test(tmp_dat, 
+                                                 test = "RESET", 
+                                                 power = 2:10, 
+                                                 power_type = "fitted")$p_value,
+               BP_p_value = mod$test(tmp_dat)$p_value,
+               SW_p_value = shapiro.test(tmp_dat$.resid)$p.value,
+               boot_id = i)
+        }) %>%
+  reduce(bind_rows)
+  }
+
+  heter_conv_sim <- heter_conv_sim %>%
+    reduce(bind_rows)
+
+  saveRDS(heter_conv_sim, here::here("data/heter_conventional_simulation.rds"))
+} else {
+  heter_conv_sim <- readRDS(here::here("data/heter_conventional_simulation.rds"))
+}
+```
+
+```{r}
+# Borrow effect size from the survey
+poly_conv_sim <- poly_conv_sim %>%
+  left_join(select(filter(vi_survey, type == "polynomial"), shape, e_sigma, n, x_dist, effect_size))
+
+heter_conv_sim <- heter_conv_sim %>%
+  left_join(select(filter(vi_survey, type == "heteroskedasticity"), a, b, n, x_dist, effect_size))
 ```
@@
 
+The parameter $\alpha$ used for the $p$-value calculation needs to be estimated from responses to null lineups.
+However, The way we generate Rorschach lineup is not strictly the same as what suggested in @vanderplas2021statistical and @buja_statistical_2009.
+Therefore, we conduct a sensitivity analysis in this section to examine the impact of the variation of the estimator $\alpha$ on our primary findings.
+
+The analysis is conducted by setting up two relatively extreme scenarios, where the $\alpha$ is under or overestimated by 25%.
+Using the inflated and deflated $\hat{\alpha}$, we recalculate the $p$-value for every lineup and show the results in Figure \ref{fig:sensitive}.
+It can be observed that $p$-value changes very little.
+It is very unlikely the findings will be affected because of the estimate of $\alpha$.
+
+```{r}
+vi_survey_upper <- calc_p_value_multi(vi_survey %>%
+                                        mutate(alpha_upper = alpha * 1.25,
+                                               alpha_lower = alpha * 0.75),
+                                      lineup_id = unique_lineup_id,
+                                      detect = detect,
+                                      n_sel = num_selection,
+                                      alpha = alpha_upper) %>%
+  right_join(select(vi_survey, -p_value))
+
+vi_survey_lower <- calc_p_value_multi(vi_survey %>%
+                                        mutate(alpha_upper = alpha * 1.25,
+                                               alpha_lower = alpha * 0.75),
+                                      lineup_id = unique_lineup_id,
+                                      detect = detect,
+                                      n_sel = num_selection,
+                                      alpha = alpha_lower) %>%
+  right_join(select(vi_survey, -p_value))
+```
+
+```{r cache = TRUE, eval = FALSE}
+# Define the minimum and maximum effect size
+min_poly_es <- vi_survey_upper %>% 
+  filter(!null_lineup, 
+         !attention_check,
+         type == "polynomial", 
+         x_dist == "uniform") %>%
+  pull(effect_size) %>%
+  min()
+max_poly_es <- vi_survey_upper %>% 
+  filter(!null_lineup, 
+         !attention_check,
+         type == "polynomial", 
+         x_dist == "uniform") %>%
+  pull(effect_size) %>%
+  max()
+
+# The glm model for RESET test
+reset_poly_glmmod <- poly_conv_sim %>%
+  filter(x_dist == "uniform") %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  mutate(reject = as.numeric(RESET4_p_value <= 0.05)) %>%
+  mutate(offset0 = log(0.05/0.95)) %>%
+  select(x_dist, effect_size, log_effect_size, offset0, reject) %>%
+  glm(reject ~ effect_size - 1, family = binomial(), data = ., offset = offset0)
+
+# The prediction for all conventional tests on poly model
+pred_conv_poly <- poly_conv_sim %>%
+  filter(x_dist == "uniform") %>%
+  select(-RESET3_p_value, -(RESET5_p_value:RESET10_p_value), -F_p_value) %>%
+  rename(RESET_p_value = RESET4_p_value) %>%
+  pivot_longer(RESET_p_value:SW_p_value) %>%
+  mutate(name = gsub("_p_value", " test", name)) %>%
+  mutate(reject = value <= 0.05) %>%
+  select(effect_size, name, reject) %>%
+  mutate(offset0 = log(0.05/0.95)) %>%
+  nest(dat = c(effect_size, offset0, reject)) %>%
+  mutate(mod = map(dat, 
+                   ~glm(reject ~ effect_size - 1, 
+                        family = binomial(), 
+                        data = .x,
+                        offset = offset0))) %>%
+  mutate(power = map(mod, function(mod) {
+    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
+                         offset0 = log(0.05/0.95))
+    result$power <- predict(mod, type = "response", newdata = result)
+    result
+  })) %>%
+  select(-dat, -mod) %>%
+  unnest(power) %>%
+  mutate(log_effect_size = log(effect_size))
+
+# Data needed by the visual test
+visual_poly_dat <- vi_survey_upper %>%
+  filter(!null_lineup, 
+         !attention_check, 
+         type == "polynomial", 
+         x_dist == "uniform") %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  mutate(reject = as.numeric(p_value <= 0.05)) %>%
+  mutate(offset0 = log(0.05/0.95)) %>%
+  select(effect_size, offset0, reject)
+
+# The glm model for visual test
+visual_poly_glmmod <- glm(reject ~ effect_size - 1, 
+                          family = binomial(), 
+                          data = visual_poly_dat, 
+                          offset = offset0)
+
+# The prediction of the visual test
+pred_visual_poly <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
+                               offset0 = log(0.05/0.95)) %>%
+  mutate(power = predict(visual_poly_glmmod, 
+                         type = "response",
+                         newdata = data.frame(effect_size = effect_size,
+                                              offset0 = offset0))) %>%
+  mutate(log_effect_size = log(effect_size))
+
+# The prediction of the boot visual test
+pred_visual_poly_boot <- map_df(1:500, function(boot_id) {
+  slice_sample(visual_poly_dat, n = nrow(visual_poly_dat), replace = TRUE) %>%
+    nest(dat = c(effect_size, offset0, reject)) %>%
+    mutate(mod = map(dat, 
+                     ~glm(reject ~ effect_size - 1, 
+                          family = binomial(), 
+                          data = .x,
+                          offset = offset0))) %>%
+    mutate(power = map(mod, function(mod) {
+      result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
+                           offset0 = log(0.05/0.95))
+      result$power <- predict(mod, type = "response", newdata = result)
+      result
+    })) %>%
+    select(-dat, -mod) %>%
+    unnest(power) %>%
+    mutate(log_effect_size = log(effect_size)) %>%
+    mutate(boot_id = boot_id)
+  })
+```
+
+```{r eval = FALSE}
+ggplot() +
+
+  geom_point(data = visual_poly_dat,
+             aes(log(effect_size), reject),
+             alpha = 0.15) +
+
+  # Draw conventional test power curve
+  geom_line(data = pred_conv_poly,
+            aes(log_effect_size, power, col = name),
+            size = 1) +
+
+  # Draw boot visual test curves
+  geom_line(data = pred_visual_poly_boot,
+            aes(log_effect_size, power, col = "Visual test", group = boot_id),
+            size = 1,
+            alpha = 0.01) +
+
+  # Draw visual test power curve
+  geom_line(data = pred_visual_poly,
+            aes(log_effect_size, power, col = "Visual test"), 
+            size = 1) +
+
+  # Theme
+  theme_light(base_size = 5) +
+  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.5)) +
+  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
+  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
+  xlab(expression(log[e] (Effect_size))) +
+  ylab("Power") +
+  labs(col = "", size = "# lineups")
+```
+
+```{r sensitive, fig.cap = "Change of $p$-values with $\\hat{\\alpha}$ inflated by 25% and deflated by 25%."}
+vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(x_dist == "uniform") %>%
+  group_by(unique_lineup_id) %>%
+  summarise(p_value = first(p_value)) %>%
+  left_join(
+vi_survey_upper %>%
+  group_by(unique_lineup_id) %>%
+  summarise(p_value_upper = first(p_value))) %>%
+  left_join(
+vi_survey_lower %>%
+  group_by(unique_lineup_id) %>%
+  summarise(p_value_lower = first(p_value))) %>%
+  pivot_longer(p_value_upper:p_value_lower) %>%
+  mutate(name = ifelse(name == "p_value_upper", "+25%", "-25%")) %>%
+  ggplot() +
+  geom_point(aes(p_value, value), alpha = 0.2) +
+  geom_abline() +
+  scale_x_sqrt() +
+  scale_y_sqrt() +
+  facet_wrap(~name) +
+  ylab("Adjusted p-value") +
+  xlab("p-value")
+
+```
+
+```{r cache = TRUE, eval = FALSE}
+# Define the minimum and maximum effect size
+min_heter_es <- vi_survey_upper %>% 
+  filter(!null_lineup, 
+         !attention_check,
+         type == "heteroskedasticity", 
+         x_dist == "uniform") %>%
+  pull(effect_size) %>%
+  min()
+max_heter_es <- vi_survey_upper %>% 
+  filter(!null_lineup, 
+         !attention_check,
+         type == "heteroskedasticity", 
+         x_dist == "uniform") %>%
+  pull(effect_size) %>%
+  max()
+
+# The prediction for all conventional tests on heter model
+pred_conv_heter <- heter_conv_sim %>%
+  filter(x_dist == "uniform") %>%
+  select(-RESET3_p_value, -(RESET5_p_value:RESET10_p_value), -F_p_value) %>%
+  rename(RESET_p_value = RESET4_p_value) %>%
+  pivot_longer(RESET_p_value:SW_p_value) %>%
+  mutate(name = gsub("_p_value", " test", name)) %>%
+  mutate(reject = value <= 0.05) %>%
+  select(effect_size, name, reject) %>%
+  mutate(offset0 = log(0.05/0.95)) %>%
+  nest(dat = c(effect_size, offset0, reject)) %>%
+  mutate(mod = map(dat, 
+                   ~glm(reject ~ effect_size - 1, 
+                        family = binomial(), 
+                        data = .x,
+                        offset = offset0))) %>%
+  mutate(power = map(mod, function(mod) {
+    result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 1),
+                         offset0 = log(0.05/0.95))
+    result$power <- predict(mod, type = "response", newdata = result)
+    result
+  })) %>%
+  select(-dat, -mod) %>%
+  unnest(power) %>%
+  mutate(log_effect_size = log(effect_size))
+
+# Data needed by the visual test
+visual_heter_dat <- vi_survey_upper %>%
+  filter(!null_lineup, 
+         !attention_check, 
+         type == "heteroskedasticity", 
+         x_dist == "uniform") %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  mutate(reject = as.numeric(p_value <= 0.05)) %>%
+  mutate(offset0 = log(0.05/0.95)) %>%
+  select(effect_size, offset0, reject)
+
+# The glm model for visual test
+visual_heter_glmmod <- glm(reject ~ effect_size - 1, 
+                           family = binomial(), 
+                           data = visual_heter_dat, 
+                           offset = offset0)
+
+# The prediction of the visual test
+pred_visual_heter <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
+                                offset0 = log(0.05/0.95)) %>%
+  mutate(power = predict(visual_heter_glmmod, 
+                         type = "response",
+                         newdata = data.frame(effect_size = effect_size,
+                                              offset0 = offset0))) %>%
+  mutate(log_effect_size = log(effect_size))
+
+# The prediction of the boot visual test
+pred_visual_heter_boot <- map_df(1:500, function(boot_id) {
+  slice_sample(visual_heter_dat, n = nrow(visual_heter_dat), replace = TRUE) %>%
+    nest(dat = c(effect_size, offset0, reject)) %>%
+    mutate(mod = map(dat, 
+                     ~glm(reject ~ effect_size - 1, 
+                          family = binomial(), 
+                          data = .x,
+                          offset = offset0))) %>%
+    mutate(power = map(mod, function(mod) {
+      result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
+                           offset0 = log(0.05/0.95))
+      result$power <- predict(mod, type = "response", newdata = result)
+      result
+    })) %>%
+    select(-dat, -mod) %>%
+    unnest(power) %>%
+    mutate(log_effect_size = log(effect_size)) %>%
+    mutate(boot_id = boot_id)
+  })
+```
+
+```{r eval = FALSE}
+# Analyse uniform heteroskedasticity data
+ggplot() +
+
+  geom_point(data = visual_heter_dat,
+             aes(log(effect_size), reject),
+             alpha = 0.15) +
+
+  # Draw conventional test power curve
+  geom_line(data = pred_conv_heter,
+            aes(log_effect_size, power, col = name),
+            size = 1) +
+
+  # Draw boot visual test curves
+  geom_line(data = pred_visual_heter_boot,
+            aes(log_effect_size, power, col = "Visual test", group = boot_id),
+            size = 1,
+            alpha = 0.01) +
+
+  # Draw visual test power curve
+  geom_line(data = pred_visual_heter,
+            aes(log_effect_size, power, col = "Visual test"), 
+            size = 1) +
+
+  # Theme
+  theme_light(base_size = 5) +
+  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.5)) +
+  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
+  # scale_size_manual(values = seq(1, 9)) +
+  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
+  xlab(expression(log[e] (Effect_size))) +
+  ylab("Power") +
+  labs(col = "", size = "# lineups")
+```
+
 ## Other results

### 2023-03-07 paper_comparison/appendix.Rmd
fill in batch analysis

@@
 
+```{r batch, fig.cap = "Boxplot of weighted propotion of detect for lineups over different data collection periods."}
+vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(x_dist == "uniform") %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
+  mutate(exp = exp - 2) %>%
+  group_by(exp, unique_lineup_id) %>%
+  mutate(prop_detect = mean(weighted_detect)) %>%
+  group_by(exp, unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  ggplot() +
+  geom_boxplot(aes(x = factor(exp), y = prop_detect)) +
+  facet_grid(~type) +
+  scale_x_discrete(labels = c("I", "II", "III")) +
+  ylab("Weighted propotion of detect") +
+  xlab("Data collection period")
+```
+
+```{r}
+t_test_poly <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(x_dist == "uniform") %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
+  mutate(exp = exp - 2) %>%
+  group_by(exp, unique_lineup_id) %>%
+  mutate(prop_detect = mean(weighted_detect)) %>%
+  group_by(exp, unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(type == "non-linearity") %>%
+  t.test(prop_detect ~ exp, data = .)
+```
+
+```{r}
+t_test_heter <- vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(x_dist == "uniform") %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
+  mutate(exp = exp - 2) %>%
+  group_by(exp, unique_lineup_id) %>%
+  mutate(prop_detect = mean(weighted_detect)) %>%
+  group_by(exp, unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  filter(type != "non-linearity") %>%
+  t.test(prop_detect ~ exp, data = .)
+```
+
+We have the same type of model collected over different data collection periods, that may lead to unexpected batch effect.
+Figure \ref{fig:batch} shows the weighed proportion of detect over different data collection period.
+The weighted proportion of detect is calculated by taking the average of $c_i$ of a lineup over a data collection period.
+According to the graph, the distribution for the heteroskedasticity patterns are almost identical in period II and III, while the distribution for the non-linearity patterns have a shift in the median.
+However, this difference is not statistically significance.
+The $p$-values of the t-test for testing the mean difference across data periods are `r round(t_test_poly$p.value, 2)` and `r round(t_test_heter$p.value, 2)` for non-linearity and heteroskedasticity patterns respectively.
+Both are much greater than 5%.
+Therefore, there is no clear evidence of batch effect.
+
 ## Sensitivity analysis for $\alpha$

### 2023-03-08 paper_comparison/paper_comparison/paper_comparison.Rmd
finish KL formula in appendix
effect size explanation

@@
 
-Effect size $E$ can be defined as the difference of a parameter for a particular model or distribution, or a statistic derived from a sample.
+Effect size in statistics measures the strength of the signal relative to the noise. It is surprisingly difficult to quantify in general, even for simulated data as used in this experiment.
+
+For the non-linearity model, the key items defining effect size are sample size ($n$) and variance of the error term ($\sigma^2$), and so effect size would be roughly calculated as $\frac{\sqrt(n)}{\sigma}$.
+As sample size increases the effect size would increase, but as variance increases the effect size decreases.
+However, it is not clear how the additional parameter for the model polynomial order, $k$, should be incorporated.
+Intuitively, the large $k$ means more complex pattern, which likely means effect size would decrease.
+For the purposes of our calculations we have chosen to use an approach based on Kullback-Leibler divergence [@kullback1951information], coupled with simulation.
+This formulation defines effect size to be:
+
+$$E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right)$$
+
+\noindent where $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals with $\boldsymbol{Z}$, $\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a$ is the residual operator, $\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ is the hat matrix.
+For the non-linearity model this can be rewritten as
+
+$$E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)$$
+
+\noindent where $\boldsymbol{V}$ is the covariance matrix of the error term.
+
+<!-- Effect size $E$ can be defined as the difference of a parameter for a particular model or distribution, or a statistic derived from a sample.
 Importantly, it needs to reflect the treatment we try to measure.
@@
 
-$$E = \frac{1}{2}\left(\boldsymbol{\mu}_z'[diag(\boldsymbol{R}_a\sigma^2)]^{-1}\boldsymbol{\mu}_z\right)$$
-
-for the non-linearity model, and
+$$E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right)$$
+
+for the non-linearity model, and -->
+
+In the heteroskedasticity model, the key elements for measuring effect size are sample size, $n$, and the ratio of the biggest variance to smallest variance, $b$.
+Larger values of both would produce higher effect size.
+However, it is not clear how to incorporate the additional shape parameter, $a$.
+Thus the same approach is used here, where the formula can be written as:
 
 $$E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)$$
-for the heteroskedasticity model, where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix, $\sigma^2\boldsymbol{I}$ is the assumed covariance matrix of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual covariance matrix of the error term, $\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a$ is the residual operator, $\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ is the hat matrix, $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of predictors including the intercept used in the regression equation, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients.
-
-Both equations above are derived from the Kullback-Leibler divergence (citation here?) and the details are provided in the Appendix.
+\noindent where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix.
+<!--, $\sigma^2\boldsymbol{I}$ is the assumed covariance matrix of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual covariance matrix of the error term, $\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a$ is the residual operator, $\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ is the hat matrix, $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of predictors including the intercept used in the regression equation, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients.
+-->
+
+Derivations for these equations are provided in the Appendix.
+
+To compute the effect size for each lineup we simulate ??
 

### 2023-03-08 paper_comparison/paper_comparison/paper_comparison.tex
effect size explanation

@@
 
-Effect size \(E\) can be defined as the difference of a parameter for a
-particular model or distribution, or a statistic derived from a sample.
-Importantly, it needs to reflect the treatment we try to measure.
-According to the underlying simulated model, we define the effect size
-
-\[E = \frac{1}{2}\left(\boldsymbol{\mu}_z'[diag(\boldsymbol{R}_a\sigma^2)]^{-1}\boldsymbol{\mu}_z\right)\]
-
-for the non-linearity model, and
-
-\[E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)\]
-for the heteroskedasticity model, where \(diag(.)\) is the diagonal
-matrix constructed from the diagonal elements of a matrix,
-\(\sigma^2\boldsymbol{I}\) is the assumed covariance matrix of the error
-term when \(H_0\) is true, \(\boldsymbol{V}\) is the actual covariance
-matrix of the error term,
+Effect size in statistics measures the strength of the signal relative
+to the noise. It is surprisingly difficult to quantify in general, even
+for simulated data as used in this experiment.
+
+For the non-linearity model, the key items defining effect size are
+sample size (\(n\)) and variance of the error term (\(\sigma^2\)), and
+so effect size would be roughly calculated as
+\(\frac{\sqrt(n)}{\sigma}\). As sample size increases the effect size
+would increase, but as variance increases the effect size decreases.
+However, it is not clear how the additional parameter for the model
+polynomial order, \(k\), should be incorporated. Intuitively, the large
+\(k\) means more complex pattern, which likely means effect size would
+decrease. For the purposes of our calculations we have chosen to use an
+approach based on Kullback-Leibler divergence
+\citep{kullback1951information}, coupled with simulation. This
+formulation defines effect size to be:
+
+\[E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right)\]
+
+\noindent where
+\(\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z\)
+is the expected values of residuals with \(\boldsymbol{Z}\),
 \(\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a\) is the
@@
 \(\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'\)
-is the hat matrix,
-\(\boldsymbol{X}_a = (\boldsymbol{1}, \boldsymbol{X})\) is the matrix of
-predictors including the intercept used in the regression equation, and
-\(\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z\)
-is the expected values of residuals with \(\boldsymbol{Z}\) be any
-higher order terms of \(\boldsymbol{X}\) leave out by the regression
-equation and \(\boldsymbol{\beta}_z\) be the corresponding coefficients.
-
-Both equations above are derived from the Kullback-Leibler divergence
-(citation here?) and the details are provided in the Appendix.
+is the hat matrix. For the non-linearity model this can be rewritten as
+
+\[E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)\]
+
+\noindent where \(\boldsymbol{V}\) is the covariance matrix of the error
+term.
+
+In the heteroskedasticity model, the key elements for measuring effect
+size are sample size, \(n\), and the ratio of the biggest variance to
+smallest variance, \(b\). Larger values of both would produce higher
+effect size. However, it is not clear how to incorporate the additional
+shape parameter, \(a\). Thus the same approach is used here, where the
+formula can be written as:
+
+\[E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)\]
+\noindent where \(diag(.)\) is the diagonal matrix constructed from the
+diagonal elements of a matrix.
+
+Derivations for these equations are provided in the Appendix.
+
+To compute the effect size for each lineup we simulate ??
 
@@
 For non-linearity patterns, the power curve of RESET test climbs
-aggressively from 7\% to 85\% as \(log_e(E)\) increases from 0 to 2.5,
+aggressively from 16\% to 100\% as \(log_e(E)\) increases from 0 to 2.5,
 while power of other tests respond inactively to the change of effect,

### 2023-03-08 paper_comparison/appendix.Rmd

@@
 
-```{r}
+```{r education, fig.cap = "\\label{fig:education}Education distribution of subjects recuritted in this study group by age and pronoun."}
 vi_survey %>%
@@
 
-```{r}
+```{r experience, fig.cap = "\\label{fig:experience}Status of pervious experience of subjects recuritted in this study group by age and pronoun."}
 vi_survey %>%
@@
 
-```{r batch, fig.cap = "Boxplot of weighted propotion of detect for lineups over different data collection periods."}
+```{r batch, fig.cap = "\\label{fig:batch}Boxplot of weighted propotion of detect for lineups over different data collection periods."}
 vi_survey %>%
@@
 
-```{r sensitive, fig.cap = "Change of $p$-values with $\\hat{\\alpha}$ inflated by 25% and deflated by 25%."}
+```{r sensitive, fig.cap = "\\label{fig:sensitive}Change of $p$-values with $\\hat{\\alpha}$ inflated by 25% and deflated by 25%."}
 vi_survey %>%

### 2023-03-12 paper_comparison/paper_comparison/paper_comparison.Rmd
fix notation in effect size

@@
 title: |
-  Why is it that statistical tests for residuals are not widely used? An explanation using visual inference.
+  A plot is worth a thousand tests: assessing residual diagnostics with visual inference
 type: ARTICLE TEMPLATE
@@
 $$E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right)$$
-
-\noindent where $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals with $\boldsymbol{Z}$, $\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a$ is the residual operator, $\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ is the hat matrix.
-For the non-linearity model this can be rewritten as
-
-$$E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)$$
-
-\noindent where $\boldsymbol{V}$ is the covariance matrix of the error term.
-
-<!-- Effect size $E$ can be defined as the difference of a parameter for a particular model or distribution, or a statistic derived from a sample.
-Importantly, it needs to reflect the treatment we try to measure.
-According to the underlying simulated model, we define the effect size
-
-$$E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right)$$
-
-for the non-linearity model, and -->
+\noindent where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,
+$\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a$ is the residual operator, $\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ is the hat matrix, $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients, and $\sigma^2\boldsymbol{I}$ is the assumed covariance matrix of the error term when $H_0$ is true.
 
@@
 $$E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)$$
-\noindent where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix.
-<!--, $\sigma^2\boldsymbol{I}$ is the assumed covariance matrix of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual covariance matrix of the error term, $\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a$ is the residual operator, $\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ is the hat matrix, $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of predictors including the intercept used in the regression equation, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients.
--->
+\noindent where $\boldsymbol{V}$ is the actual covariance matrix of the error term.
 

### 2023-03-12 paper_comparison/paper_comparison/paper_comparison.Rmd
add effect size calc details

@@
 
-To compute the effect size for each lineup we simulate ??
+To compute the effect size for each lineup we simulate a sufficient large number of samples from the same model, in each sample, the number of observations $n$ is fixed.
+We then compute the effect size for each sample and take the average as the final value.
+This ensures lineups constructed with the same experimental factors will share the same effect size.
 

### 2023-03-12 paper_comparison/appendix.Rmd
update on paper and appendix

@@
 ---
-title: "appendix"
+title: "Supplmentary material - A plot is worth a thousand tests: assessing residual diagnostics with visual inference"
 output: pdf_document
-date: "2023-01-22"
+author: ""
 ---
@@
 
+(use LV plot `lvplot` `geom_lv`)
+
 ## Sensitivity analysis for $\alpha$
@@
   ylab("Adjusted p-value") +
-  xlab("p-value")
-
+  xlab("p-value") +
+  coord_fixed()
 ```
@@
 
+## Curious effect of predictor distribution on conventional test power
+
 ## Other results

### 2023-03-12 paper_comparison/paper_comparison/paper_comparison.Rmd
update on paper and appendix

--- 2023-03-12 paper_comparison/paper_comparison/paper_comparison.Rmd
@@
 
-Effect size in statistics measures the strength of the signal relative to the noise. It is surprisingly difficult to quantify in general, even for simulated data as used in this experiment.
-
-For the non-linearity model, the key items defining effect size are sample size ($n$) and variance of the error term ($\sigma^2$), and so effect size would be roughly calculated as $\frac{\sqrt(n)}{\sigma}$.
-As sample size increases the effect size would increase, but as variance increases the effect size decreases.
-However, it is not clear how the additional parameter for the model polynomial order, $k$, should be incorporated.
-Intuitively, the large $k$ means more complex pattern, which likely means effect size would decrease.
-For the purposes of our calculations we have chosen to use an approach based on Kullback-Leibler divergence [@kullback1951information], coupled with simulation.
-This formulation defines effect size to be:
+Effect size $E$ can be defined as the difference of a parameter for a particular model or distribution, or a statistic derived from a sample.
+Importantly, it needs to reflect the treatment we try to measure.
+According to the underlying simulated model, we define the effect size
 
 $$E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right)$$
-\noindent where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,
-$\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a$ is the residual operator, $\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ is the hat matrix, $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients, and $\sigma^2\boldsymbol{I}$ is the assumed covariance matrix of the error term when $H_0$ is true.
-
-In the heteroskedasticity model, the key elements for measuring effect size are sample size, $n$, and the ratio of the biggest variance to smallest variance, $b$.
-Larger values of both would produce higher effect size.
-However, it is not clear how to incorporate the additional shape parameter, $a$.
-Thus the same approach is used here, where the formula can be written as:
+
+for the non-linearity model, and
 
 $$E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)$$
-\noindent where $\boldsymbol{V}$ is the actual covariance matrix of the error term.
-
-Derivations for these equations are provided in the Appendix.
-
-To compute the effect size for each lineup we simulate ??
+for the heteroskedasticity model, where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix, $\sigma^2\boldsymbol{I}$ is the assumed covariance matrix of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual covariance matrix of the error term, $\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a$ is the residual operator, $\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ is the hat matrix, $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of predictors including the intercept used in the regression equation, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients.
+
+Both equations above are derived from the Kullback-Leibler divergence [@kullback1951information] and the details are provided in the Appendix.
 

### 2023-03-12 paper_comparison/paper_comparison/paper_comparison.Rmd

--- 2023-03-12 paper_comparison/paper_comparison/paper_comparison.Rmd
@@
 $$E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right)$$
-\noindent where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,
-$\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a$ is the residual operator, $\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ is the hat matrix, $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients, and $\sigma^2\boldsymbol{I}$ is the assumed covariance matrix of the error term when $H_0$ is true.
+
+\noindent where $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals with $\boldsymbol{Z}$, $\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a$ is the residual operator, $\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ is the hat matrix.
+For the non-linearity model this can be rewritten as
+
+$$E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)$$
+
+\noindent where $\boldsymbol{V}$ is the covariance matrix of the error term.
+
+<!-- Effect size $E$ can be defined as the difference of a parameter for a particular model or distribution, or a statistic derived from a sample.
+Importantly, it needs to reflect the treatment we try to measure.
+According to the underlying simulated model, we define the effect size
+
+$$E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right)$$
+
+for the non-linearity model, and -->
 
@@
 $$E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)$$
-\noindent where $\boldsymbol{V}$ is the actual covariance matrix of the error term.
+\noindent where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix.
+<!--, $\sigma^2\boldsymbol{I}$ is the assumed covariance matrix of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual covariance matrix of the error term, $\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a$ is the residual operator, $\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ is the hat matrix, $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of predictors including the intercept used in the regression equation, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients.
+-->
 

### 2023-03-13 paper_comparison/paper_comparison/paper_comparison.tex
compile pdf

@@
 
-\title{Why is it that statistical tests for residuals are not widely
-used? An explanation using visual inference.}
+\title{A plot is worth a thousand tests: assessing residual diagnostics
+with visual inference}
 
@@
 \end{keywords}
-
-Title (jokes: finding the pattern in residual tests) (why using
-visulization is important in regression diagnostics? an explanation
-using visual inference) (a picture is worth a thousand word) (a plot is
-worth a thousand test: )
-
-(title refine - rank: low)
 
@@
 \[E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right)\]
-
-\noindent where
-\(\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z\)
-is the expected values of residuals with \(\boldsymbol{Z}\),
+\noindent where \(diag(.)\) is the diagonal matrix constructed from the
+diagonal elements of a matrix,
 \(\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a\) is the
@@
 \(\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'\)
-is the hat matrix. For the non-linearity model this can be rewritten as
-
-\[E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)\]
-
-\noindent where \(\boldsymbol{V}\) is the covariance matrix of the error
-term.
+is the hat matrix,
+\(\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z\)
+is the expected values of residuals with \(\boldsymbol{Z}\) be any
+higher order terms of \(\boldsymbol{X}\) leave out by the regression
+equation and \(\boldsymbol{\beta}_z\) be the corresponding coefficients,
+and \(\sigma^2\boldsymbol{I}\) is the assumed covariance matrix of the
+error term when \(H_0\) is true.
 
@@
 \[E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)\]
-\noindent where \(diag(.)\) is the diagonal matrix constructed from the
-diagonal elements of a matrix.
+\noindent where \(\boldsymbol{V}\) is the actual covariance matrix of
+the error term.
 
@@
 
-To compute the effect size for each lineup we simulate ??
+To compute the effect size for each lineup we simulate a sufficient
+large number of samples from the same model, in each sample, the number
+of observations \(n\) is fixed. We then compute the effect size for each
+sample and take the average as the final value. This ensures lineups
+constructed with the same experimental factors will share the same
+effect size.
 
@@
 For non-linearity patterns, the power curve of RESET test climbs
-aggressively from 16\% to 100\% as \(log_e(E)\) increases from 0 to 2.5,
+aggressively from 7\% to 84\% as \(log_e(E)\) increases from 0 to 2.5,
 while power of other tests respond inactively to the change of effect,

### 2023-03-14 paper_comparison/paper_comparison/paper_comparison.Rmd
compile pdf
update notation in test

--- 2023-03-12 paper_comparison/paper_comparison/paper_comparison.Rmd
@@
 
-Title 
-(jokes: finding the pattern in residual tests) 
-(why using visulization is important in regression diagnostics? an explanation using visual inference)
-(a picture is worth a thousand word)
-(a plot is worth a thousand test: )
-
-(title refine - rank: low)
-
 (fix the tense issue - rank: low)
@@
 However, it is practically impossible to test this specific $H_0$ with conventional tests, which are constructed to measure specific departures.
-For example, the RESET test is formulated as $H_0:\gamma_1 = \gamma_2 = \gamma_3 = 0$ against $H_a: \gamma_1 \neq 0 \text{ or } \gamma_2 \neq 0 \text{ or } \gamma_3 \neq 0$, from $e = \tau_0 + \tau_1x_1 + ...
-+ \tau_1 x_p +\gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)$.
-Similarly, the Breusch-Pagan test is designed to specifically test $H_0:$ *error variances are all equal* ($\tau_i=0 \text{ for } i=1,..,p$) versus the alternative $H_a:$ *that the error variances are a multiplicative function of one or more variables* ($\text{at least one } \tau_i\neq 0$) from $e^2 = \tau_0 + \tau_1 x_1 + \cdots + \tau_1 x_p + u, ~ u\sim N(0,\sigma_u^2)$.
+For example, the RESET test is formulated as $H_0:\gamma_1 = \gamma_2 = \gamma_3 = 0$ against $H_a: \gamma_1 \neq 0 \text{ or } \gamma_2 \neq 0 \text{ or } \gamma_3 \neq 0$, from $y = \tau_0 + \sum_{i=1}^{p}\tau_px_p +\gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)$.
+Similarly, the Breusch-Pagan test is designed to specifically test $H_0:$ *error variances are all equal* ($\zeta_i=0 \text{ for } i=1,..,p$) versus the alternative $H_a:$ *that the error variances are a multiplicative function of one or more variables* ($\text{at least one } \zeta_i\neq 0$) from $e^2 = \zeta_0 + \sum_{i=1}^{p}\zeta_i x_i + u, ~ u\sim N(0,\sigma_u^2)$.
 

### 2023-03-14 paper_comparison/paper_comparison/paper_comparison.Rmd
avoid using frac in inline equation

@@
 
-As pointed out by @vanderplas2021statistical, this basic binomial model doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup.
+As pointed out by @vanderplas2021statistical, this basic binomial model does not take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup.
 And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup.
@@
 
-For the non-linearity model, the key items defining effect size are sample size ($n$) and variance of the error term ($\sigma^2$), and so effect size would be roughly calculated as $\frac{\sqrt(n)}{\sigma}$.
+For the non-linearity model, the key items defining effect size are sample size ($n$) and variance of the error term ($\sigma^2$), and so effect size would be roughly calculated as $\sqrt{n}/{\sigma}$.
 As sample size increases the effect size would increase, but as variance increases the effect size decreases.

### 2023-03-14 paper_comparison/appendix.Rmd
tidy the notation and explanation in appendix

@@
 output: pdf_document
-author: ""
+author: "Weihao Li, Dianne Cook, Emi Tanaka and Susan VanderPlas"
+bibliography: "paper_comparison/paper.bib"
 ---
@@
 
-## Effect size
+## Effect size derivation
 
-## Derivation
-
-Effect size is usually defined as the difference of a parameter or a statistic derived from a sample.
+Effect size can be defined as the difference of a parameter for a particular model or distribution,  or a statistic derived from a sample.
+Importantly, it needs to reflect the treatment we try to measure.
 Centred on a conventional statistical test, we usually can deduce the effect size from the test statistic by substituting the null parameter value.
@@
 To build a bridge between various residual-based tests, and the visual test, we focus on the shared information embedded in the testing procedures, which is the distribution of residuals.
-When comes to comparison of distribution, Kullback-Leibler divergence is a classical way to represent the information loss or entropy increase caused by the approximation to the true distribution, which in our case, the inefficiency due to the use of false model assumptions.
+When comes to comparison of distribution, Kullback-Leibler divergence [@@kullback1951information] is a classical way to represent the information loss or entropy increase caused by the approximation to the true distribution, which in our case, the inefficiency due to the use of false model assumptions.
 
@@
 
-## Caculation
-
-The effect size can be calculated using Equation \ref{eq:effect-size} for each lineup by plugging in the data generated from the simulated model.
-However, since the term $\boldsymbol{R}_a$ will differ for each sample, the effect size will be different even if the parameter values are the same.
-To overcome this undesired property, for the same set of parameter values, we simulate a sufficient large number of samples, and take the average of the effect size as the final effect size.
-
 ## Experiment setup
@@
 Let $L$ be a set of available lineups and $S$ be a set of available subjects.
-According to the experimental design, available means a lineup can be allocated to at most five different subjects (for lineups with uniform fitted value distribution, this value is 11) and a subject can view at most 18 different lineups.
+According to the experimental design, the availability of a lineup is associated with the number of subjects it can assign to.
+For lineups with uniform fitted value distribution, this value is 11.
+And other lineups can be allocated to at most five different subjects.
+The availability of a subject is associated with the number of lineups that being allocated to this subject.
+A subject can view at most 18 different lineups.
 
-The algorithm starts from picking a random subject $s$ from $S$ with the minimum number of allocated lineups.
-It then computes a metric for every lineup $l$ in $L$ indicating the distance $D$ between lineup $l$ to the lineups already allocated to this subject $l_{allocated}$.
-Lineup with the highest distance $D$ will be allocated to the subject.
-The algorithm will repeatedly pick a subject, compute metrics, and allocate lineups until there is no available lineups or no available subjects.
+The algorithm starts from picking a random subject $s \in S$ with the minimum number of allocated lineups.
+It then tries to find a lineup $l \in L$ that can maximise the distance metric $D$ and allocate it to subject $s$.
+Set $L$ and $s$ will be updated and the picking process will be repeated until there is no available lineups or subjects.
 
-The distance $D$ is defined using the factor values used to generate lineup $l$ and lineups $l_{allocated}$.
-If there does not exist $l_{allocated}$, then the distance $D = 0$.
-Otherwise, distance $D$ is roughly the number of factors minus the number of factor values of lineup $l$ exists in a union of factor values of lineups $l_{allocated}$.
-The distance $D$ is further adjusted by the existence of $n$-factor combinations, where $n$ could be two to the number of factors.
-In this way, all factors values are exposed to every subject, and some even ensure all $2$-factor combinations are exposed to every subject.
+Let $F_1,...,F_q$ be $q$ experimental factors, and $f_1, ...,f_q$ be the corresponding factor values.
+We say $f_i$ exists in $L_{s}$ if any lineup in $L_{s}$ has this factor value.
+Similarly, $f_if_j$ exists in $L_{s}$ if any lineup in $L_{s}$ has this pair of factor values.
+And $f_if_jf_k$ exists in $L_{s}$ if any lineup in $L_{s}$ has this trio of factor values.
+The distance metric $D$ is defined between a lineup $l$ and a set of lineups $L_{s}$ allocated to a subject $s$ if $L_{s}$ is non-empty:
+
+\footnotesize
+
+\begin{equation*}
+D =
+C - \sum_{i}I(f_i\text{ exists in }L_{s}) - \sum_{i}\sum_{j\neq i}I(f_if_j\text{ exists in }L_{s}) - \sum_{i}\sum_{j\neq i}\sum_{k\neq j}I(f_if_jf_k\text{ exists in }L_{s})
+\end{equation*}
+
+\normalsize
+where $C$ is a sufficient large constant such that $D > 0$. If $L_{s}$ is empty, we define $D = 0$.
+
+The distance is measuring how different a lineup is from the set of lineups allocated to the subject in terms of factor values.
+Thus, the algorithm is trying to allocate the most different lineup to a subject.
 

### 2023-03-14 paper_comparison/paper_comparison/paper_comparison.Rmd
intro check by emi

@@
 We can assess residuals in multiple ways.
-Residuals might be plotted, as a histogram or quantile-quantile plot to examine the distribution.
-Using the classical normal linear regression model as an example, if the distribution is symmetric and unimodal, we consider it well-behaved.
-But if the distribution is skewed, bimodal, multimodal, or contains outliers, there is cause for concern.
+Residuals may be plotted as a histogram or quantile-quantile plot to examine the distribution.
+Using the classical normal linear regression model as an example, if the distribution is symmetric and unimodal, we consider it to be well behaved.
+However, if the distribution is skewed, bimodal, multimodal, or contains outliers, there is a cause for concern.
 One could also inspect the distribution by conducting a goodness of fit test, such as the Shapiro-Wilk Normality test [@shapiro1965analysis].
@@
 In general, one looks for noticeable departures from the model like non-linear dependency or heteroskedasticity.
-However, correctly judging a residual plot where no pattern exists is a painstakingly difficult task for humans (?citation).
-It is especially common, particularly among new data analysts, to report patterns when an experienced one might quickly conclude that there are none.
+However, correctly judging a residual plot where no pattern exists can be a painstakingly difficult task for humans (?citation).
+It is especially common, particularly among new data analysts, to report patterns when an experienced one might quickly conclude that there are none (ET: not sure if this is true).
 It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and use a Breusch-Pagan test [@breusch_simple_1979] for heteroskedasticity.
 
-Abundance of literature describe appropriate diagnostic methods for linear regression: @draper1998applied, @montgomery1982introduction, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals.
-All of these writings consider plotting of residuals to be a standard technique that should be examined routinely in all regression modelling problems.
-In addition, @draper1998applied and @belsley_regression_1980 believe that residual plots are usually very revealing when the assumptions are violated.
+Abundance of literature describe appropriate diagnostic methods for linear regression, e.g.
+@draper1998applied, @montgomery1982introduction, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals.
+All these writings consider plotting residuals to be a standard technique that should be examined routinely in all regression modelling problems.
+In addition, @draper1998applied and @belsley_regression_1980 believe that residual plots are usually revealing when the assumptions are violated.
 @cook_applied_1999 thinks formal tests and graphical procedures are complementary and both have a place in residual analysis, but they focus on graphical methods rather than on formal testing, as they are easier to use.
-@montgomery1982introduction even suggests that residual plots are more informative in most practical situations than corresponding formal tests, and statistical tests on regression model residuals are not widely used based on their experience.
+@montgomery1982introduction even suggests that residual plots are more informative in most practical situations than the corresponding formal tests, and statistical tests on regression model residuals are not widely used based on their experience.
 
@@
 The paper is structured as follows.
-The next background section describes the types of departures that one expects to detect, and outlines a formal statistical process for reading residual plots, called visual inference.
-Section \ref{experimental-design} details the experimental design to compare the decision made by formal hypothesis testing, and how humans would read diagnostic plots.
+The next section describes the background on the types of departures that one expects to detect, and outlines a formal statistical process for reading residual plots, called visual inference.
+Section \ref{experimental-design} details the experimental design to compare the decisions made by formal hypothesis testing, and how humans would read diagnostic plots.
 The results are reported in Section \ref{results}.
-We finish with a discussion on future work, in particular how the responsibility for residual plot reading might be relieved.
+We conclude with a discussion of future work, in particular, how the responsibility for residual plot reading might be relieved.
 

### 2023-03-14 paper_comparison/paper_comparison/paper_comparison.Rmd
background section 1

@@
   dat %>%
-    mutate(type = "Good residuals"),
+    mutate(type = "(A) Good residuals"),
   mod$set_prm("include_z", TRUE)$
     gen(300, computed = select(dat, x, e)) %>%
-    mutate(type = "Non-linearity"),
+    mutate(type = "(B) Non-linearity"),
   heter_model(b = 64)$
     gen(300, computed = select(dat, x, e)) %>%
-    mutate(type = "Heteroskedasticity"),
+    mutate(type = "(C) Heteroskedasticity"),
   heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
     gen(300, computed = select(dat, x)) %>%
-    mutate(type = "Non-normality")
+    mutate(type = "(D) Non-normality")
 ) %>%
@@
 
-Graphical summaries in which residuals are plotted against fitted values or other functions of the predictors that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals.
-The first panel of Figure \ref{fig:residual-plot-common-departures} shows an ideal residual plot with residuals evenly distributed at both sides of the horizontal zero line, with no noticeable patterns.
-
-There are various types of departures from an ideal residual plot. Non-linearity, heterskedasticity and non-normality are perhaps the three mostly checked departures.
+Graphical summaries in which residuals are plotted against fitted values or other functions of the predictors that are approximately orthogonal to the residuals are referred to as standard residual plots in @cook1982residuals.
+Figure \ref{fig:residual-plot-common-departures}A shows an ideal residual plot where the residuals are evenly distributed at both sides of the horizontal zero line, with no noticeable patterns.
+
+There are various types of departures from an ideal residual plot. Non-linearity, heteroskedasticity and non-normality are perhaps the three mostly checked departures.
 
@@
 Any non-linear functional form of residuals on fitted values in the residual plot could be indicative of non-linearity.
-An example residual plot containing visual pattern of non-linearity is given at the second panel of Figure \ref{fig:residual-plot-common-departures}.
+An example residual plot containing visual pattern of non-linearity is shown in Figure \ref{fig:residual-plot-common-departures}B.
 One can clearly observe the "S-shape" from the residual plot as the cubic term is not captured by the misspecified model.
@@
 The usual pattern of heteroskedasticity on a residual plot is the inconsistent spread of the residuals across the horizontal axis.
-Visually, it sometimes results in the so-called "butterfly" shape as shown in the third panel of Figure \ref{fig:residual-plot-common-departures}, or the "left-triangle" and "right-triangle" shape where the smallest variance occurs at one side of the horizontal axis.
+Visually, it sometimes results in the so-called "butterfly" shape as shown in Figure \ref{fig:residual-plot-common-departures}C, or the "left-triangle" and "right-triangle" shape where the smallest variance occurs at one side of the horizontal axis.
 
@@
 As we mainly discuss residual plots, non-normality will not be the focus of this paper.
-For a consistent comparison, the residual plot of this departure is still presented in the fourth panel of Figure \ref{fig:residual-plot-common-departures}.
+For a consistent comparison, the residual plot of this departure is still presented in Figure \ref{fig:residual-plot-common-departures}D.
 When the number residuals below and above the horizontal axis are uneven across the local regions along the $x$-axis, we expect that the normality assumption is violated.
-For example, given a skewed error distribution, there will be fewer data points and more outliers on one side of the horizontal axis as shown in the fourth panel of Figure \ref{fig:residual-plot-common-departures}.
+For example, given a skewed error distribution, there will be fewer data points and more outliers on one side of the horizontal axis as shown in Figure \ref{fig:residual-plot-common-departures}D.
 

### 2023-03-14 paper_comparison/appendix.Rmd
update subject allocation equation

@@
 D =
-C - \sum_{i}I(f_i\text{ exists in }L_{s}) - \sum_{i}\sum_{j\neq i}I(f_if_j\text{ exists in }L_{s}) - \sum_{i}\sum_{j\neq i}\sum_{k\neq j}I(f_if_jf_k\text{ exists in }L_{s})
+C - \sum_{1\leq i \leq q}I(f_i\text{ exists in }L_{s}) - \sum_{\substack{1\leq i \leq q-1 \\ i < j \leq q}}I(f_if_j\text{ exists in }L_{s}) - \sum_{\substack{1\leq i \leq q - 2 \\ i < j \leq q - 1 \\ j < k \leq q}}I(f_if_jf_k\text{ exists in }L_{s})
 \end{equation*}
@@
 \normalsize
-where $C$ is a sufficient large constant such that $D > 0$. If $L_{s}$ is empty, we define $D = 0$.
+where $C$ is a sufficiently large constant such that $D > 0$. If $L_{s}$ is empty, we define $D = 0$.
 
 The distance is measuring how different a lineup is from the set of lineups allocated to the subject in terms of factor values.
-Thus, the algorithm is trying to allocate the most different lineup to a subject.
+Thus, the algorithm will try to allocate the most different lineup to a subject.
 

### 2023-03-14 paper_comparison/appendix.Rmd
refine demographic

@@
 It then tries to find a lineup $l \in L$ that can maximise the distance metric $D$ and allocate it to subject $s$.
-Set $L$ and $s$ will be updated and the picking process will be repeated until there is no available lineups or subjects.
+Set $L$ and $S$ will be updated and the picking process will be repeated until there is no available lineups or subjects.
 
@@
 
-The distance is measuring how different a lineup is from the set of lineups allocated to the subject in terms of factor values.
-Thus, the algorithm will try to allocate the most different lineup to a subject.
+The distance measures how different a lineup is from the set of lineups allocated to the subject in terms of factor values.
+Thus, the algorithm will try to allocate the most different lineup to a subject at each step.
 
@@
 
-The education background of most participants have Diploma or Bachelor degrees, followed by High school or below. The survey data is gender balanced. Most of the participants are between 18 to 39.
+Along with the responses to lineups, we have collected a series of demographic information including age, pronoun, education background and previous experience in studies involved data visualization.
+Figure \ref{fig:education} and \ref{fig:experience} provide summary of the demographic data.
+
+As shown in Figure \ref{fig:education}, most participants have Diploma or Bachelor degrees, followed by High school or below.
+The survey data is gender balanced and the majority of participants are between 18 to 39 years old.
 
@@
   mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
+  mutate(pronoun = factor(pronoun, levels = c("He", "She", "Other"))) %>%
   ggplot() +
@@
   xlab("Age group") +
-  ylab("Count")
+  ylab("Count") +
+  theme_light()
 ```
 
-Number of participants who have previous experience in experiment about reading data plots are not very different from the number of participants who haven’t.
+Figure \ref{fig:experience} shows that the number of participants who have previous experience is not very different from the number of those who haven't.
 Age distributions are also similar for these two groups.
@@
   xlab("Age group") +
-  ylab("Count")
+  ylab("Count") +
+  theme_light()
 ```

### 2023-03-14 paper_comparison/appendix.Rmd
correct demographic graph

@@
 
-```{r education, fig.cap = "\\label{fig:education}Education distribution of subjects recuritted in this study group by age and pronoun."}
+```{r education, fig.cap = "\\label{fig:education}Education distribution of subjects recuritted in this study group by age and pronoun.", fig.height = 4}
 vi_survey %>%
@@
   mutate(pronoun = factor(pronoun, levels = c("He", "She", "Other"))) %>%
+  group_by(exp, set) %>%
+  summarise(across(everything(), first)) %>%
   ggplot() +
@@
   ylab("Count") +
-  theme_light()
+  theme_light(base_size = 10)
 ```

### 2023-03-14 paper_comparison/appendix.Rmd
add data collection interface

@@
 
-Figure \ref{fig:experience} shows that the number of participants who have previous experience is not very different from the number of those who haven't.
+Figure \ref{fig:experience} shows no particular difference between the number of participants who have previous experience and the number of those who haven't.
 Age distributions are also similar for these two groups.
@@
 ### Data collection interface
+
+The survey data is collected via a self-hosted website designed by us.
+Once the participant is recruited from Prolific (ref here), it will be redirected to the entry page of our study website.
+An image of the entry page is provided in Figure \ref{fig:entry-page}.
+Then, the participant needs to submit the online consent form and fill in the demographic information as shown in \ref{fig:consent_form} and \ref{fig:metadata} respectively.
+Before evaluating lineups, participant also need to read the training page as provide in Figure \ref{fig:training-page} to understand the process.
+An example of the lineup page is given in Figure \ref{fig:lineup-page}.
+A half of the page is taken by the lineup image to attract participant's attention.
+The button to skip the selections for the current lineup is intentionally put in the corner of the bounding box with smaller font size, such that participants will not misuse this functionality.
+
+```{r entry-page, out.width = "100%", fig.cap = "The entry page of the study website."}
+knitr::include_graphics("img/message.png")
+```
+
+```{r consent-form, out.width = "100%", fig.cap = "The consent form provided in the study website."}
+knitr::include_graphics("img/consent_form.png")
+```
+
+```{r metadata, out.width = "100%", fig.cap = "The form to provide demographic information."}
+knitr::include_graphics("img/metadata.png")
+```
+
+```{r training-page, out.width = "100%", fig.cap = "The training page of the study website."}
+knitr::include_graphics("img/training.png")
+```
+
+```{r lineup-page, out.width = "100%", fig.cap = "The lineup page of the study website."}
+knitr::include_graphics("img/lineup1.png")
+```
 
@@
 ## Curious effect of predictor distribution on conventional test power
-
-## Other results

### 2023-03-14 paper_comparison/appendix.Rmd
use lvplot for cmp

@@
 
-```{r batch, fig.cap = "\\label{fig:batch}Boxplot of weighted propotion of detect for lineups over different data collection periods."}
+```{r poly-boxplot-lineup, fig.cap = '\\label{fig:poly-boxplot-lineup}A lineup of "letter-value" boxplots of weighted propotion of detect for lineups over different data collection periods for non-linearity model.
+Can you find the most different boxplot?
+The data plot is positioned in panel $2^3 - 1$.'}
+library(lvplot)
+sample_from_null <- function(dat, true_position) {
+  dat %>%
+    mutate(k = true_position) %>%
+    bind_rows(map_df((1:20)[-true_position], function(k) {
+    dat %>%
+      mutate(exp = sample(exp),
+             k = k)
+  }))
+}
+
 vi_survey %>%
@@
   summarise(across(everything(), first)) %>%
+  ungroup() %>%
+  filter(type == "non-linearity") %>%
+  select(type, exp, prop_detect) %>%
+  mutate(exp = as.character(exp)) %>%
+  sample_from_null(7) %>%
   ggplot() +
-  geom_boxplot(aes(x = factor(exp), y = prop_detect)) +
-  facet_grid(~type) +
-  scale_x_discrete(labels = c("I", "II", "III")) +
+  geom_lv(aes(x = exp, y = prop_detect, fill = after_stat(LV)), varwidth = TRUE) +
+  # scale_fill_lv() +
+  facet_wrap(~k, ncol = 5) +
+  theme_light() +
+  scale_x_discrete(labels = c("I", "III")) +
   ylab("Weighted propotion of detect") +
   xlab("Data collection period")
+```
+
+```{r heter-boxplot-lineup, fig.cap = '\\label{fig:heter-boxplot-lineup}A lineup of "letter-value" boxplots of weighted propotion of detect for lineups over different data collection periods for heteroskedasticity model.
+Can you find the most different boxplot?
+The data plot is positioned in panel $2^4 - 2$.'}
+vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(x_dist == "uniform") %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
+  mutate(exp = exp - 2) %>%
+  group_by(exp, unique_lineup_id) %>%
+  mutate(prop_detect = mean(weighted_detect)) %>%
+  group_by(exp, unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  ungroup() %>%
+  filter(type != "non-linearity") %>%
+  select(type, exp, prop_detect) %>%
+  mutate(exp = as.character(exp)) %>%
+  sample_from_null(14) %>%
+  ggplot() +
+  geom_lv(aes(x = exp, y = prop_detect, fill = after_stat(LV)), varwidth = TRUE) +
+  # scale_fill_lv() +
+  facet_wrap(~k, ncol = 5) +
+  theme_light() +
+  scale_x_discrete(labels = c("II", "III")) +
+  ylab("Weighted propotion of detect") +
+  xlab("Data collection period")
+```
+
+```{r batch, fig.cap = '\\label{fig:batch}A "letter-value" boxplot of weighted propotion of detect for lineups over different data collection periods.'}
+vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(x_dist == "uniform") %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
+  mutate(exp = exp - 2) %>%
+  group_by(exp, unique_lineup_id) %>%
+  mutate(prop_detect = mean(weighted_detect)) %>%
+  group_by(exp, unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  ungroup() %>%
+ggplot() +
+  geom_lv(aes(x = factor(exp), y = prop_detect, fill = after_stat(LV)), varwidth = TRUE) +
+  # scale_fill_lv() +
+  facet_grid(~type) +
+  theme_light()
 ```
@@
 We have the same type of model collected over different data collection periods, that may lead to unexpected batch effect.
-Figure \ref{fig:batch} shows the weighed proportion of detect over different data collection period.
+Figure \ref{fig:poly-boxplot-lineup} and \ref{fig:heter-boxplot-lineup} provide two lineups to examine whether there is an actual difference across data collection periods for non-linearity model and heteroskedasticity model respectively.
+To emphasize the tail behaviour and display fewer outliers, we use the "letter-value" boxplot (ref here) which is an extension of the number of "letter value" statistics (ref here) to check the weighed proportion of detect over different data collection period.
 The weighted proportion of detect is calculated by taking the average of $c_i$ of a lineup over a data collection period.
-According to the graph, the distribution for the heteroskedasticity patterns are almost identical in period II and III, while the distribution for the non-linearity patterns have a shift in the median.
-However, this difference is not statistically significance.
+Within our research team, we can not identify the data plot from the null plots for these two lineups, result in $p$-values much greater than $5$%.
+
+Figure \ref{fig:batch} shows the data plot, where the distribution for the heteroskedasticity patterns are almost identical in period II and III, while the distribution for the non-linearity patterns have a shift in the median.
 The $p$-values of the t-test for testing the mean difference across data periods are `r round(t_test_poly$p.value, 2)` and `r round(t_test_heter$p.value, 2)` for non-linearity and heteroskedasticity patterns respectively.
 Both are much greater than 5%.
-Therefore, there is no clear evidence of batch effect.
-
-(use LV plot `lvplot` `geom_lv`)
+Therefore, both the visual test and the conventional test, there is no clear evidence of batch effect.
 

### 2023-03-14 paper_comparison/appendix.Rmd
update analysis for alpha

@@
 
-The analysis is conducted by setting up two relatively extreme scenarios, where the $\alpha$ is under or overestimated by 25%.
+The analysis is conducted by setting up two relatively extreme scenarios, where the $\alpha$ is under or overestimated by 50%.
 Using the inflated and deflated $\hat{\alpha}$, we recalculate the $p$-value for every lineup and show the results in Figure \ref{fig:sensitive}.
-It can be observed that $p$-value changes very little.
-It is very unlikely the findings will be affected because of the estimate of $\alpha$.
+It can be observed that there are some changes to $p$-values, especially when the $\hat{\alpha}$ is deflated.
+However, Figure \ref{fig:sensitive-barplot} shows that adjusting $\hat{\alpha}$ will not result in a huge difference in rejection decisions.
+There are only a small number of cases (12 out of 558) where the rejection decision changes.
+It is very unlikely the downstream findings will be affected because of the estimate of $\alpha$.
 
@@
 vi_survey_upper <- calc_p_value_multi(vi_survey %>%
-                                        mutate(alpha_upper = alpha * 1.25,
-                                               alpha_lower = alpha * 0.75),
+                                        mutate(alpha_upper = alpha * 1.5,
+                                               alpha_lower = alpha * 0.5),
                                       lineup_id = unique_lineup_id,
@@
 vi_survey_lower <- calc_p_value_multi(vi_survey %>%
-                                        mutate(alpha_upper = alpha * 1.25,
-                                               alpha_lower = alpha * 0.75),
+                                        mutate(alpha_upper = alpha * 1.5,
+                                               alpha_lower = alpha * 0.5),
                                       lineup_id = unique_lineup_id,
@@
 
-```{r sensitive, fig.cap = "\\label{fig:sensitive}Change of $p$-values with $\\hat{\\alpha}$ inflated by 25% and deflated by 25%."}
+```{r sensitive, fig.cap = "\\label{fig:sensitive}Change of $p$-values with $\\hat{\\alpha}$ inflated by 50% and deflated by 50%."}
 vi_survey %>%
@@
   pivot_longer(p_value_upper:p_value_lower) %>%
-  mutate(name = ifelse(name == "p_value_upper", "+25%", "-25%")) %>%
+  mutate(name = ifelse(name == "p_value_upper", "+50%", "-50%")) %>%
   ggplot() +
@@
   xlab("p-value") +
-  coord_fixed()
+  coord_fixed() +
+  theme_light()
+```
+
+```{r sensitive-barplot, fig.cap = "\\label{fig:sensitive-barplot}Change of rejection of the null hypothesis with $\\hat{\\alpha}$ inflated by 50% and deflated by 50%."}
+vi_survey %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(x_dist == "uniform") %>%
+  group_by(unique_lineup_id) %>%
+  summarise(p_value = first(p_value)) %>%
+  left_join(
+vi_survey_upper %>%
+  group_by(unique_lineup_id) %>%
+  summarise(p_value_upper = first(p_value))) %>%
+  left_join(
+vi_survey_lower %>%
+  group_by(unique_lineup_id) %>%
+  summarise(p_value_lower = first(p_value))) %>%
+  pivot_longer(p_value_upper:p_value_lower) %>%
+  mutate(name = ifelse(name == "p_value_upper", "+50%", "-50%")) %>%
+  ggplot() +
+  geom_bar(aes(p_value < 0.05, fill = value < 0.05), position = "dodge") +
+  theme_light() +
+  xlab("p-value < 0.05") +
+  labs(fill = "Adjusted p-value < 0.05")
 ```
@@
 ## Curious effect of predictor distribution on conventional test power
+
+\newpage
+
+## References

### 2023-03-14 paper_comparison/paper_comparison/paper_comparison.Rmd
background section 2

@@
 
-Other than checking diagnostic plots, analysts may perform formal hypothesis testing for detecting model defects.
-Depending on the alternative hypothesis that is focused on, a variety of tests can be applied.
+Other than checking diagnostic plots, analysts may perform formal hypothesis testing to detect model defects.
+A variety of tests can be applied; each testing for a specific violation of the null hypothesis.
 For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution.
-For testing non-linearity, one may apply the F-test as a model structural test to examine the significance of specific polynomial and non-linear forms of the predictors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969].
-The Shapiro-Wilk test [@shapiro1965analysis] is the most widely used test of non-normality included by many of the statistical softwares.
-The Jarque--Bera test [@jarque1980efficient] is also used to directly checks if the sample skewness and kurtosis match a normal distribution.
-
-Example residual plots given in Figure \ref{fig:residual-plot-common-departures} are examined by the corresponding RESET test, Breusch-Pagan test and Shapiro-Wilk test as shown in Table \ref{tab:example-residual-plot-table}.
-In the example, the Breusch-Pagan test and the Shapiro--Wilk test both reject $H_0$ for departures that they do not intend to examine.
+To test specific forms of non-linearity, one may apply the F-test as a model structural test to examine the significance of specific polynomial and non-linear forms of the predictors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969].
+The Shapiro-Wilk test [@shapiro1965analysis] is the most widely used test of non-normality included by many of the statistical software programs.
+The Jarque--Bera test [@jarque1980efficient] is also used to directly check whether the sample skewness and kurtosis match a normal distribution.
+
+We apply the RESET test, Breusch-Pagan test and Shapiro-Wilk test to the residual plots shown in Figure \ref{fig:residual-plot-common-departures}; the results of this is shown in Table \ref{tab:example-residual-plot-table}.
+The Breusch-Pagan test and the Shapiro-Wilk test both reject $H_0$ for residual plots that display non-linearity and heteroscedasticity, even though these are not the original intention of the test.
 As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are also sensitive to other types of departures.

### 2023-03-14 paper_comparison/appendix.Rmd
update exp setup
init the Acknowledgement

@@
 
+## Experiment setup
+
+### Mapping of subjects to experimental factors
+
+Mapping of subjects to experimental factors is an important part of experiment design.
+Essentially, we want to maximum the difference in factors exposed to a subject.
+For this purpose, we design an algorithm to conduct subject allocation.
+Let $L$ be a set of available lineups and $S$ be a set of available subjects.
+According to the experimental design, the availability of a lineup is associated with the number of subjects it can assign to.
+For lineups with uniform fitted value distribution, this value is 11.
+And other lineups can be allocated to at most five different subjects.
+The availability of a subject is associated with the number of lineups that being allocated to this subject.
+A subject can view at most 18 different lineups.
+
+The algorithm starts from picking a random subject $s \in S$ with the minimum number of allocated lineups.
+It then tries to find a lineup $l \in L$ that can maximise the distance metric $D$ and allocate it to subject $s$.
+Set $L$ and $S$ will be updated and the picking process will be repeated until there is no available lineups or subjects.
+
+Let $F_1,...,F_q$ be $q$ experimental factors, and $f_1, ...,f_q$ be the corresponding factor values.
+We say $f_i$ exists in $L_{s}$ if any lineup in $L_{s}$ has this factor value.
+Similarly, $f_if_j$ exists in $L_{s}$ if any lineup in $L_{s}$ has this pair of factor values.
+And $f_if_jf_k$ exists in $L_{s}$ if any lineup in $L_{s}$ has this trio of factor values.
+The distance metric $D$ is defined between a lineup $l$ and a set of lineups $L_{s}$ allocated to a subject $s$ if $L_{s}$ is non-empty:
+
+\footnotesize
+
+\begin{equation*}
+D =
+C - \sum_{1\leq i \leq q}I(f_i\text{ exists in }L_{s}) - \sum_{\substack{1\leq i \leq q-1 \\ i < j \leq q}}I(f_if_j\text{ exists in }L_{s}) - \sum_{\substack{1\leq i \leq q - 2 \\ i < j \leq q - 1 \\ j < k \leq q}}I(f_if_jf_k\text{ exists in }L_{s})
+\end{equation*}
+
+\normalsize
+where $C$ is a sufficiently large constant such that $D > 0$. If $L_{s}$ is empty, we define $D = 0$.
+
+The distance measures how different a lineup is from the set of lineups allocated to the subject in terms of factor values.
+Thus, the algorithm will try to allocate the most different lineup to a subject at each step.
+
+### Data collection process
+
+The survey data is collected via a self-hosted website designed by us.
+
+(desc of the tech layers)
+
+Once the participant is recruited from Prolific (ref here), it will be redirected to the entry page of our study website.
+An image of the entry page is provided in Figure \ref{fig:entry-page}.
+Then, the participant needs to submit the online consent form and fill in the demographic information as shown in \ref{fig:consent_form} and \ref{fig:metadata} respectively.
+Before evaluating lineups, participant also need to read the training page as provide in Figure \ref{fig:training-page} to understand the process.
+An example of the lineup page is given in Figure \ref{fig:lineup-page}.
+A half of the page is taken by the lineup image to attract participant's attention.
+The button to skip the selections for the current lineup is intentionally put in the corner of the bounding box with smaller font size, such that participants will not misuse this functionality.
+
+```{r tech, out.width = "100%", fig.cap = "Diagram of online experiment setup.
+The server-side of the study website uses Flask as backend hosted on PythonAnywhere.
+And the client-side uses jsPsych to run experiment."}
+knitr::include_graphics("img/experiment_tech.pdf")
+```
+
+```{r entry-page, out.width = "100%", fig.cap = "The entry page of the study website."}
+knitr::include_graphics("img/message.png")
+```
+
+```{r consent-form, out.width = "100%", fig.cap = "The consent form provided in the study website."}
+knitr::include_graphics("img/consent_form.png")
+```
+
+```{r metadata, out.width = "100%", fig.cap = "The form to provide demographic information."}
+knitr::include_graphics("img/metadata.png")
+```
+
+```{r training-page, out.width = "100%", fig.cap = "The training page of the study website."}
+knitr::include_graphics("img/training.png")
+```
+
+```{r lineup-page, out.width = "100%", fig.cap = "The lineup page of the study website."}
+knitr::include_graphics("img/lineup1.png")
+```
+
+## Demographics
+
+Along with the responses to lineups, we have collected a series of demographic information including age, pronoun, education background and previous experience in studies involved data visualization.
+Figure \ref{fig:education} and \ref{fig:experience} provide summary of the demographic data.
+
+As shown in Figure \ref{fig:education}, most participants have Diploma or Bachelor degrees, followed by High school or below.
+The survey data is gender balanced and the majority of participants are between 18 to 39 years old.
+
+```{r education, fig.cap = "\\label{fig:education}Education distribution of subjects recuritted in this study group by age and pronoun.", fig.height = 4}
+vi_survey %>%
+  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
+  mutate(pronoun = factor(pronoun, levels = c("He", "She", "Other"))) %>%
+  group_by(exp, set) %>%
+  summarise(across(everything(), first)) %>%
+  ggplot() +
+  geom_bar(aes(age_group, fill = pronoun), position = "dodge") +
+  facet_wrap(~fct_relevel(education, "High School or below", "Diploma and Bachelor Degree", "Honours Degree", "Masters Degree")) +
+  xlab("Age group") +
+  ylab("Count") +
+  theme_light(base_size = 10)
+```
+
+Figure \ref{fig:experience} shows no particular difference between the number of participants who have previous experience and the number of those who haven't.
+Age distributions are also similar for these two groups.
+
+```{r experience, fig.cap = "\\label{fig:experience}Status of pervious experience of subjects recuritted in this study group by age and pronoun."}
+vi_survey %>%
+  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
+  mutate(pronoun = factor(pronoun, levels = c("He", "She", "Other"))) %>%
+  group_by(exp, set) %>%
+  summarise(across(everything(), first)) %>%
+  ggplot() +
+  geom_bar(aes(age_group, fill = pronoun), position = "dodge") +
+  facet_wrap(~previous_experience) +
+  xlab("Age group") +
+  ylab("Count") +
+  theme_light()
+```
+
 ## Effect size derivation
@@
 
-## Experiment setup
-
-### Mapping of subjects to experimental factors
-
-Mapping of subjects to experimental factors is an important part of experiment design.
-Essentially, we want to maximum the difference in factors exposed to a subject.
-For this purpose, we design an algorithm to conduct subject allocation.
-Let $L$ be a set of available lineups and $S$ be a set of available subjects.
-According to the experimental design, the availability of a lineup is associated with the number of subjects it can assign to.
-For lineups with uniform fitted value distribution, this value is 11.
-And other lineups can be allocated to at most five different subjects.
-The availability of a subject is associated with the number of lineups that being allocated to this subject.
-A subject can view at most 18 different lineups.
-
-The algorithm starts from picking a random subject $s \in S$ with the minimum number of allocated lineups.
-It then tries to find a lineup $l \in L$ that can maximise the distance metric $D$ and allocate it to subject $s$.
-Set $L$ and $S$ will be updated and the picking process will be repeated until there is no available lineups or subjects.
-
-Let $F_1,...,F_q$ be $q$ experimental factors, and $f_1, ...,f_q$ be the corresponding factor values.
-We say $f_i$ exists in $L_{s}$ if any lineup in $L_{s}$ has this factor value.
-Similarly, $f_if_j$ exists in $L_{s}$ if any lineup in $L_{s}$ has this pair of factor values.
-And $f_if_jf_k$ exists in $L_{s}$ if any lineup in $L_{s}$ has this trio of factor values.
-The distance metric $D$ is defined between a lineup $l$ and a set of lineups $L_{s}$ allocated to a subject $s$ if $L_{s}$ is non-empty:
-
-\footnotesize
-
-\begin{equation*}
-D =
-C - \sum_{1\leq i \leq q}I(f_i\text{ exists in }L_{s}) - \sum_{\substack{1\leq i \leq q-1 \\ i < j \leq q}}I(f_if_j\text{ exists in }L_{s}) - \sum_{\substack{1\leq i \leq q - 2 \\ i < j \leq q - 1 \\ j < k \leq q}}I(f_if_jf_k\text{ exists in }L_{s})
-\end{equation*}
-
-\normalsize
-where $C$ is a sufficiently large constant such that $D > 0$. If $L_{s}$ is empty, we define $D = 0$.
-
-The distance measures how different a lineup is from the set of lineups allocated to the subject in terms of factor values.
-Thus, the algorithm will try to allocate the most different lineup to a subject at each step.
-
-### Demographics
-
-Along with the responses to lineups, we have collected a series of demographic information including age, pronoun, education background and previous experience in studies involved data visualization.
-Figure \ref{fig:education} and \ref{fig:experience} provide summary of the demographic data.
-
-As shown in Figure \ref{fig:education}, most participants have Diploma or Bachelor degrees, followed by High school or below.
-The survey data is gender balanced and the majority of participants are between 18 to 39 years old.
-
-```{r education, fig.cap = "\\label{fig:education}Education distribution of subjects recuritted in this study group by age and pronoun.", fig.height = 4}
-vi_survey %>%
-  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
-  mutate(pronoun = factor(pronoun, levels = c("He", "She", "Other"))) %>%
-  group_by(exp, set) %>%
-  summarise(across(everything(), first)) %>%
-  ggplot() +
-  geom_bar(aes(age_group, fill = pronoun), position = "dodge") +
-  facet_wrap(~fct_relevel(education, "High School or below", "Diploma and Bachelor Degree", "Honours Degree", "Masters Degree")) +
-  xlab("Age group") +
-  ylab("Count") +
-  theme_light(base_size = 10)
-```
-
-Figure \ref{fig:experience} shows no particular difference between the number of participants who have previous experience and the number of those who haven't.
-Age distributions are also similar for these two groups.
-
-```{r experience, fig.cap = "\\label{fig:experience}Status of pervious experience of subjects recuritted in this study group by age and pronoun."}
-vi_survey %>%
-  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
-  mutate(pronoun = factor(pronoun, levels = c("He", "She", "Other"))) %>%
-  group_by(exp, set) %>%
-  summarise(across(everything(), first)) %>%
-  ggplot() +
-  geom_bar(aes(age_group, fill = pronoun), position = "dodge") +
-  facet_wrap(~previous_experience) +
-  xlab("Age group") +
-  ylab("Count") +
-  theme_light()
-```
-
-### Data collection interface
-
-The survey data is collected via a self-hosted website designed by us.
-Once the participant is recruited from Prolific (ref here), it will be redirected to the entry page of our study website.
-An image of the entry page is provided in Figure \ref{fig:entry-page}.
-Then, the participant needs to submit the online consent form and fill in the demographic information as shown in \ref{fig:consent_form} and \ref{fig:metadata} respectively.
-Before evaluating lineups, participant also need to read the training page as provide in Figure \ref{fig:training-page} to understand the process.
-An example of the lineup page is given in Figure \ref{fig:lineup-page}.
-A half of the page is taken by the lineup image to attract participant's attention.
-The button to skip the selections for the current lineup is intentionally put in the corner of the bounding box with smaller font size, such that participants will not misuse this functionality.
-
-```{r entry-page, out.width = "100%", fig.cap = "The entry page of the study website."}
-knitr::include_graphics("img/message.png")
-```
-
-```{r consent-form, out.width = "100%", fig.cap = "The consent form provided in the study website."}
-knitr::include_graphics("img/consent_form.png")
-```
-
-```{r metadata, out.width = "100%", fig.cap = "The form to provide demographic information."}
-knitr::include_graphics("img/metadata.png")
-```
-
-```{r training-page, out.width = "100%", fig.cap = "The training page of the study website."}
-knitr::include_graphics("img/training.png")
-```
-
-```{r lineup-page, out.width = "100%", fig.cap = "The lineup page of the study website."}
-knitr::include_graphics("img/lineup1.png")
-```
-
-## Batch effect size
+## Batch effect
 
@@
 
+1. put #eval in appendix
+3. compute the power of RESET test with different orders (considering dof) (focus on simple pattern possibly?)
+
 \newpage

### 2023-03-14 paper_comparison/paper_comparison/paper_comparison.Rmd
init the Acknowledgement

--- 2023-03-14 paper_comparison/paper_comparison/paper_comparison.Rmd
@@
   \def\tightlist{}
-  \usepackage{xcolor}
-  \definecolor{orange-red}{rgb}{1.0, 0.27, 0.0}
-  \newcommand{\svp}[1]{{\textcolor{orange-red}{#1}}}
-  \newcommand*{\Perm}[2]{{}^{#1}\!P_{#2}}%
 output: rticles::tf_article
@@
 
-1. software tools
-2. point to github repo (create a new one)
-3. summarise the supplementary material
+The article was created with rticles (ref here), knitr (ref here) and R Markdown (ref here). The project's Github repository (link here) contains all materials required to reproduce this article.
+
+# Supplementary material
+
+The supplementary material is available at (link here).
+It includes more details about the experimental setup, the derivation of the effect size, the effect of data collection period, and the estimate of  $\alpha$.
 
 \newpage
-
-1. put #eval in appendix
-2. examine batch effect in appendix
-3. compute the power of RESET test with different orders (considering dof) (focus on simple pattern possibly?)

### 2023-03-14 paper_comparison/paper_comparison/paper_comparison.Rmd
update the acknowledgement section

@@
 
+This research is enabled by the use of the following `R` packages:
+cli, curl, dplyr, ggplot2, jsonlite, lmtest, mpoly, progress, tibble, ggmosaic, purrr, tidyr, readr, stringr, here, kableExtra, patchwork, rcartocolor.
+The study website is powered by pythonAnywhere and the Flask web framework.
+The jsPsych framework is used to create behavioral experiments that run in our study website.
+
 The article was created with rticles (ref here), knitr (ref here) and R Markdown (ref here). The project's Github repository (link here) contains all materials required to reproduce this article.

### 2023-03-14 paper_comparison/paper_comparison/paper_comparison.tex
update the acknowledgement section

@@
 \def\tightlist{}
-\usepackage{xcolor}
-\definecolor{orange-red}{rgb}{1.0, 0.27, 0.0}
-\newcommand{\svp}[1]{{\textcolor{orange-red}{#1}}}
-\newcommand*{\Perm}[2]{{}^{#1}\!P_{#2}}%
 
@@
 
-We can assess residuals in multiple ways. Residuals might be plotted, as
-a histogram or quantile-quantile plot to examine the distribution. Using
+We can assess residuals in multiple ways. Residuals may be plotted as a
+histogram or quantile-quantile plot to examine the distribution. Using
 the classical normal linear regression model as an example, if the
-distribution is symmetric and unimodal, we consider it well-behaved. But
-if the distribution is skewed, bimodal, multimodal, or contains
-outliers, there is cause for concern. One could also inspect the
-distribution by conducting a goodness of fit test, such as the
+distribution is symmetric and unimodal, we consider it to be well
+behaved. However, if the distribution is skewed, bimodal, multimodal, or
+contains outliers, there is a cause for concern. One could also inspect
+the distribution by conducting a goodness of fit test, such as the
 Shapiro-Wilk Normality test \citep{shapiro1965analysis}.
@@
 heteroskedasticity. However, correctly judging a residual plot where no
-pattern exists is a painstakingly difficult task for humans (?citation).
-It is especially common, particularly among new data analysts, to report
-patterns when an experienced one might quickly conclude that there are
-none. It is also possible to conduct hypothesis tests for non-linear
-dependence \citep{ramsey_tests_1969}, and use a Breusch-Pagan test
+pattern exists can be a painstakingly difficult task for humans
+(?citation). It is especially common, particularly among new data
+analysts, to report patterns when an experienced one might quickly
+conclude that there are none (ET: not sure if this is true). It is also
+possible to conduct hypothesis tests for non-linear dependence
+\citep{ramsey_tests_1969}, and use a Breusch-Pagan test
 \citep{breusch_simple_1979} for heteroskedasticity.
@@
 Abundance of literature describe appropriate diagnostic methods for
-linear regression: \citet{draper1998applied},
+linear regression, e.g. \citet{draper1998applied},
 \citet{montgomery1982introduction}, \citet{belsley_regression_1980},
-\citet{cook_applied_1999} and \citet{cook1982residuals}. All of these
-writings consider plotting of residuals to be a standard technique that
+\citet{cook_applied_1999} and \citet{cook1982residuals}. All these
+writings consider plotting residuals to be a standard technique that
 should be examined routinely in all regression modelling problems. In
 addition, \citet{draper1998applied} and \citet{belsley_regression_1980}
-believe that residual plots are usually very revealing when the
-assumptions are violated. \citet{cook_applied_1999} thinks formal tests
-and graphical procedures are complementary and both have a place in
-residual analysis, but they focus on graphical methods rather than on
-formal testing, as they are easier to use.
-\citet{montgomery1982introduction} even suggests that residual plots are
-more informative in most practical situations than corresponding formal
-tests, and statistical tests on regression model residuals are not
-widely used based on their experience.
+believe that residual plots are usually revealing when the assumptions
+are violated. \citet{cook_applied_1999} thinks formal tests and
+graphical procedures are complementary and both have a place in residual
+analysis, but they focus on graphical methods rather than on formal
+testing, as they are easier to use. \citet{montgomery1982introduction}
+even suggests that residual plots are more informative in most practical
+situations than the corresponding formal tests, and statistical tests on
+regression model residuals are not widely used based on their
+experience.
 
@@
 is the subject of this paper. The paper is structured as follows. The
-next background section describes the types of departures that one
-expects to detect, and outlines a formal statistical process for reading
-residual plots, called visual inference. Section
+next section describes the background on the types of departures that
+one expects to detect, and outlines a formal statistical process for
+reading residual plots, called visual inference. Section
 \ref{experimental-design} details the experimental design to compare the
-decision made by formal hypothesis testing, and how humans would read
+decisions made by formal hypothesis testing, and how humans would read
 diagnostic plots. The results are reported in Section \ref{results}. We
-finish with a discussion on future work, in particular how the
+conclude with a discussion of future work, in particular, how the
 responsibility for residual plot reading might be relieved.
@@
 or other functions of the predictors that are approximately orthogonal
-to residuals are referred to as standard residual plots in
-\citet{cook1982residuals}. The first panel of Figure
-\ref{fig:residual-plot-common-departures} shows an ideal residual plot
-with residuals evenly distributed at both sides of the horizontal zero
-line, with no noticeable patterns.
+to the residuals are referred to as standard residual plots in
+\citet{cook1982residuals}. Figure
+\ref{fig:residual-plot-common-departures}A shows an ideal residual plot
+where the residuals are evenly distributed at both sides of the
+horizontal zero line, with no noticeable patterns.
 
 There are various types of departures from an ideal residual plot.
-Non-linearity, heterskedasticity and non-normality are perhaps the three
-mostly checked departures.
+Non-linearity, heteroskedasticity and non-normality are perhaps the
+three mostly checked departures.
 
@@
 residual plot could be indicative of non-linearity. An example residual
-plot containing visual pattern of non-linearity is given at the second
-panel of Figure \ref{fig:residual-plot-common-departures}. One can
-clearly observe the ``S-shape'' from the residual plot as the cubic term
-is not captured by the misspecified model.
+plot containing visual pattern of non-linearity is shown in Figure
+\ref{fig:residual-plot-common-departures}B. One can clearly observe the
+``S-shape'' from the residual plot as the cubic term is not captured by
+the misspecified model.
 
@@
 Visually, it sometimes results in the so-called ``butterfly'' shape as
-shown in the third panel of Figure
-\ref{fig:residual-plot-common-departures}, or the ``left-triangle'' and
-``right-triangle'' shape where the smallest variance occurs at one side
-of the horizontal axis.
+shown in Figure \ref{fig:residual-plot-common-departures}C, or the
+``left-triangle'' and ``right-triangle'' shape where the smallest
+variance occurs at one side of the horizontal axis.
 
@@
 paper. For a consistent comparison, the residual plot of this departure
-is still presented in the fourth panel of Figure
-\ref{fig:residual-plot-common-departures}. When the number residuals
-below and above the horizontal axis are uneven across the local regions
-along the \(x\)-axis, we expect that the normality assumption is
-violated. For example, given a skewed error distribution, there will be
-fewer data points and more outliers on one side of the horizontal axis
-as shown in the fourth panel of Figure
-\ref{fig:residual-plot-common-departures}.
+is still presented in Figure \ref{fig:residual-plot-common-departures}D.
+When the number residuals below and above the horizontal axis are uneven
+across the local regions along the \(x\)-axis, we expect that the
+normality assumption is violated. For example, given a skewed error
+distribution, there will be fewer data points and more outliers on one
+side of the horizontal axis as shown in Figure
+\ref{fig:residual-plot-common-departures}D.
 
@@
 Other than checking diagnostic plots, analysts may perform formal
-hypothesis testing for detecting model defects. Depending on the
-alternative hypothesis that is focused on, a variety of tests can be
-applied. For example, the presence of heteroskedasticity can usually be
-tested by applying the White test
-\citep{white_heteroskedasticity-consistent_1980} or the Breusch-Pagan
-test \citep{breusch_simple_1979}, which are both derived from the
-Lagrange multiplier test \citep{silvey1959lagrangian} principle that
-relies on the asymptotic properties of the null distribution. For
-testing non-linearity, one may apply the F-test as a model structural
-test to examine the significance of specific polynomial and non-linear
-forms of the predictors, or the significance of proxy variables as in
-the Ramsey Regression Equation Specification Error Test (RESET)
-\citep{ramsey_tests_1969}. The Shapiro-Wilk test
-\citep{shapiro1965analysis} is the most widely used test of
-non-normality included by many of the statistical softwares. The
-Jarque--Bera test \citep{jarque1980efficient} is also used to directly
-checks if the sample skewness and kurtosis match a normal distribution.
-
-Example residual plots given in Figure
-\ref{fig:residual-plot-common-departures} are examined by the
-corresponding RESET test, Breusch-Pagan test and Shapiro-Wilk test as
-shown in Table \ref{tab:example-residual-plot-table}. In the example,
-the Breusch-Pagan test and the Shapiro--Wilk test both reject \(H_0\)
-for departures that they do not intend to examine. As discussed in
+hypothesis testing to detect model defects. A variety of tests can be
+applied; each testing for a specific violation of the null hypothesis.
+For example, the presence of heteroskedasticity can usually be tested by
+applying the White test \citep{white_heteroskedasticity-consistent_1980}
+or the Breusch-Pagan test \citep{breusch_simple_1979}, which are both
+derived from the Lagrange multiplier test \citep{silvey1959lagrangian}
+principle that relies on the asymptotic properties of the null
+distribution. To test specific forms of non-linearity, one may apply the
+F-test as a model structural test to examine the significance of
+specific polynomial and non-linear forms of the predictors, or the
+significance of proxy variables as in the Ramsey Regression Equation
+Specification Error Test (RESET) \citep{ramsey_tests_1969}. The
+Shapiro-Wilk test \citep{shapiro1965analysis} is the most widely used
+test of non-normality included by many of the statistical software
+programs. The Jarque--Bera test \citep{jarque1980efficient} is also used
+to directly check whether the sample skewness and kurtosis match a
+normal distribution.
+
+We apply the RESET test, Breusch-Pagan test and Shapiro-Wilk test to the
+residual plots shown in Figure
+\ref{fig:residual-plot-common-departures}; the results of this is shown
+in Table \ref{tab:example-residual-plot-table}. The Breusch-Pagan test
+and the Shapiro-Wilk test both reject \(H_0\) for residual plots that
+display non-linearity and heteroscedasticity, even though these are not
+the original intention of the test. As discussed in
 \citet{cook1982residuals}, most residual-based tests for a particular
@@
 from
-\(e = \tau_0 + \tau_1x_1 + ... + \tau_1 x_p +\gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)\).
+\(y = \tau_0 + \sum_{i=1}^{p}\tau_px_p +\gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)\).
 Similarly, the Breusch-Pagan test is designed to specifically test
 \(H_0:\) \emph{error variances are all equal}
-(\(\tau_i=0 \text{ for } i=1,..,p\)) versus the alternative \(H_a:\)
+(\(\zeta_i=0 \text{ for } i=1,..,p\)) versus the alternative \(H_a:\)
 \emph{that the error variances are a multiplicative function of one or
-more variables} (\(\text{at least one } \tau_i\neq 0\)) from
-\(e^2 = \tau_0 + \tau_1 x_1 + \cdots + \tau_1 x_p + u, ~ u\sim N(0,\sigma_u^2)\).
+more variables} (\(\text{at least one } \zeta_i\neq 0\)) from
+\(e^2 = \zeta_0 + \sum_{i=1}^{p}\zeta_i x_i + u, ~ u\sim N(0,\sigma_u^2)\).
 
@@
 As pointed out by \citet{vanderplas2021statistical}, this basic binomial
-model doesn't take into account the possible dependencies in the visual
+model does not take into account the possible dependencies in the visual
 test due to repeated evaluations of the same lineup. And it is
@@
 sample size (\(n\)) and variance of the error term (\(\sigma^2\)), and
-so effect size would be roughly calculated as
-\(\frac{\sqrt(n)}{\sigma}\). As sample size increases the effect size
-would increase, but as variance increases the effect size decreases.
-However, it is not clear how the additional parameter for the model
-polynomial order, \(k\), should be incorporated. Intuitively, the large
-\(k\) means more complex pattern, which likely means effect size would
-decrease. For the purposes of our calculations we have chosen to use an
-approach based on Kullback-Leibler divergence
-\citep{kullback1951information}, coupled with simulation. This
-formulation defines effect size to be:
+so effect size would be roughly calculated as \(\sqrt{n}/{\sigma}\). As
+sample size increases the effect size would increase, but as variance
+increases the effect size decreases. However, it is not clear how the
+additional parameter for the model polynomial order, \(k\), should be
+incorporated. Intuitively, the large \(k\) means more complex pattern,
+which likely means effect size would decrease. For the purposes of our
+calculations we have chosen to use an approach based on Kullback-Leibler
+divergence \citep{kullback1951information}, coupled with simulation.
+This formulation defines effect size to be:
 
@@
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  software tools
-\item
-  point to github repo (create a new one)
-\item
-  summarise the supplementary material
-\end{enumerate}
+This research is enabled by the use of the following \texttt{R}
+packages: cli, curl, dplyr, ggplot2, jsonlite, lmtest, mpoly, progress,
+tibble, ggmosaic, purrr, tidyr, readr, stringr, here, kableExtra,
+patchwork, rcartocolor. The study website is powered by pythonAnywhere
+and the Flask web framework. The jsPsych framework is used to create
+behavioral experiments that run in our study website.
+
+The article was created with rticles (ref here), knitr (ref here) and R
+Markdown (ref here). The project's Github repository (link here)
+contains all materials required to reproduce this article.
+
+\hypertarget{supplementary-material}{%
+\section{Supplementary material}\label{supplementary-material}}
+
+The supplementary material is available at (link here). It includes more
+details about the experimental setup, the derivation of the effect size,
+the effect of data collection period, and the estimate of \(\alpha\).
 
 \newpage
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  put \#eval in appendix
-\item
-  examine batch effect in appendix
-\item
-  compute the power of RESET test with different orders (considering
-  dof) (focus on simple pattern possibly?)
-\end{enumerate}
 

### 2023-03-14 paper_comparison/paper_comparison/paper_comparison.Rmd
title to title case

--- 2023-03-14 paper_comparison/paper_comparison/paper_comparison.Rmd
@@
 title: |
-  A plot is worth a thousand tests: assessing residual diagnostics with visual inference
+  A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with Visual Inference
 type: ARTICLE TEMPLATE

### 2023-03-14 paper_comparison/paper_comparison/paper_comparison.tex
title to title case

@@
 
-\title{A plot is worth a thousand tests: assessing residual diagnostics
-with visual inference}
+\title{A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics
+with Visual Inference}
 
@@
 
-This research is enabled by the use of the following \texttt{R}
-packages: cli, curl, dplyr, ggplot2, jsonlite, lmtest, mpoly, progress,
-tibble, ggmosaic, purrr, tidyr, readr, stringr, here, kableExtra,
-patchwork, rcartocolor. The study website is powered by pythonAnywhere
-and the Flask web framework. The jsPsych framework is used to create
-behavioral experiments that run in our study website.
-
 The article was created with rticles (ref here), knitr (ref here) and R

### 2023-03-14 paper_comparison/paper_comparison/paper_comparison.Rmd

--- 2023-03-14 paper_comparison/paper_comparison/paper_comparison.Rmd
@@
 title: |
-  A plot is worth a thousand tests: assessing residual diagnostics with visual inference
+  A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol
 type: ARTICLE TEMPLATE
@@
 keywords: |
-  data visualization; visual inference; hypothesis testing; residual plots;
+  statistical graphics; data visualization; visual inference; hypothesis testing; reression analysis; cognitive perception; simulation; practical significance; effect size
 header-includes: |

### 2023-03-14 paper_comparison/paper_comparison/paper_comparison.tex

@@
 \title{A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics
-with Visual Inference}
+with the Lineup Protocol}
 
@@
 \begin{keywords}
-data visualization; visual inference; hypothesis testing; residual
-plots;
+statistical graphics; data visualization; visual inference; hypothesis
+testing; reression analysis; cognitive perception; simulation; practical
+significance; effect size
 \end{keywords}
@@
 
+This research is enabled by the use of the following \texttt{R}
+packages: cli, curl, dplyr, ggplot2, jsonlite, lmtest, mpoly, progress,
+tibble, ggmosaic, purrr, tidyr, readr, stringr, here, kableExtra,
+patchwork, rcartocolor. The study website is powered by pythonAnywhere
+and the Flask web framework. The jsPsych framework is used to create
+behavioral experiments that run in our study website.
+
 The article was created with rticles (ref here), knitr (ref here) and R

### 2023-03-21 paper_comparison/paper_comparison/paper_comparison.Rmd
update bib

@@
 
+(put citation https://www.tandfonline.com/doi/full/10.1080/26939169.2021.1920866)
+
 More typically, residuals will be plotted, as a scatter plot against the predicted values and each of the explanatory variables to scrutinize their relationships.
@@
 
-# Acknowledgement
-
-This research is enabled by the use of the following `R` packages:
-cli, curl, dplyr, ggplot2, jsonlite, lmtest, mpoly, progress, tibble, ggmosaic, purrr, tidyr, readr, stringr, here, kableExtra, patchwork, rcartocolor.
-The study website is powered by pythonAnywhere and the Flask web framework.
-The jsPsych framework is used to create behavioral experiments that run in our study website.
-
-The article was created with rticles (ref here), knitr (ref here) and R Markdown (ref here). The project's Github repository (link here) contains all materials required to reproduce this article.
+# Acknowledgements
+
+These `R` packages are used for the work: 
+`cli` [@cli], `curl` [@curl], `dplyr` [@dplyr], `ggplot2` [@ggplot2], `jsonlite` [@jsonlite], `lmtest` [@lmtest], `mpoly` [@mpoly], `progress` [@progress], `tibble` [@tibble], `ggmosaic` [@ggmosaic], `purrr` [@purrr], `tidyr` [@tidyr], `readr` [@readr], `stringr` [@stringr], `here` [@here], `kableExtra` [@kableextra], `patchwork` [@patchwork], `rcartocolor` [@rcartocolor].
+The study website is powered by `PythonAnywhere` [@pythonanywhere] and the `Flask` web framework [@flask].
+The `jsPsych` framework [@jspsych] is used to create behavioral experiments that run in our study website.
+
+The article was created with R packages `rticles` [@rticles], `knitr` [@knitr] and `rmarkdown` [@rmarkdown].
+The project's Github repository (link here) contains all materials required to reproduce this article.
 
@@
 
-\newpage
+<!-- \newpage -->

### 2023-03-21 Milestone_2/milestone_2_report.Rmd
milestone 2 template

---
title: "Milestone report"
output: pdf_document
date: "2023-03-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

A short report (5 - 10) pages

Advances in Artificial Intelligence for Data Visualization: Automate Reading of Data Plots

# Overview of the thesis

1. background and motivation
- the role of visual diagnostics in modelling
- a brief introduction to visual inference
- limitation of lineup protocol
2. research questions
-  a. study the use of visual inference in regression diagnostics with a comparison to conventional test
-  b. build a computer vision system to evaluate residual plot
-  c. make the system available to the public
3. summarise the work in the first year
- an experiment to compare visual test with conventional test 
- a prototype model for lineup evaluation
- points out the limitation of the work
4. summarise the work in the second year
- refine and reconduct the experiment
5. main findings to date
- findings from the first year
- findings from the second year

# Thesis structure

topic

chapter 1: Introduction

chapter 2
the first paper

chapter 3
build a computer vision system to evaluate residual plot

chapter 4?
software publication

chapter 5
discussion and future work

# Timetable

- April: submit abstract to ASC
- May: submit paper
- June: submit poster/short paper to IEEE vis conf, start working on computer vision model
- July: leave for few weeks
- Aug
- Sep: finalize the computer vision model
- Oct: IEEE vis conf
- Nov: web interface development 
- Dec: ASC

- Jan, 2024: 
- Feb, 2024: 
- Mar, 2024: submit paper
- Aug, 2024: submit thesis

# Difficulties

things that are actually quite serious

1. identified difficulties
2. suggestions for overcoming these difficulties

### 2023-03-21 paper_comparison/appendix.Rmd
replace numbers with letters

@@
 title: "Supplmentary material - A plot is worth a thousand tests: assessing residual diagnostics with visual inference"
-output: pdf_document
+output: 
+  pdf_document:
+    number_sections: true
 author: "Weihao Li, Dianne Cook, Emi Tanaka and Susan VanderPlas"
 bibliography: "paper_comparison/paper.bib"
+header-includes:
+  - \renewcommand\thesection{\Alph{section}}
+  - \renewcommand\thesubsection{\thesection.\arabic{subsection}}
 ---
@@
 
+## Effect of number of evaluations on the power of a visual test
+
+## Power of a RESET test under different auxiliary regression formulas
+
 1. put #eval in appendix

### 2023-03-21 paper_comparison/appendix.Rmd
create table for demogrphic

@@
 The survey data is gender balanced and the majority of participants are between 18 to 39 years old.
+
+```{r}
+vi_survey %>%
+  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
+  mutate(pronoun = factor(pronoun, levels = c("He", "She", "Other"))) %>%
+  group_by(exp, set) %>%
+  summarise(across(everything(), first)) %>%
+  count(pronoun) %>%
+  mutate(exp = paste("Period", c("I", "II", "III")[exp - 2])) %>%
+  pivot_wider(values_from = n, names_from = exp) %>%
+  mutate(Total = `Period I` + `Period II` + `Period III`) %>%
+  rename(Pronoun = pronoun) %>%
+  kableExtra::kable(format = "latex", booktabs = TRUE)
+```
+
+```{r}
+vi_survey %>%
+  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
+  mutate(pronoun = factor(pronoun, levels = c("He", "She", "Other"))) %>%
+  group_by(exp, set) %>%
+  summarise(across(everything(), first)) %>%
+  count(age_group) %>%
+  mutate(exp = paste("Period", c("I", "II", "III")[exp - 2])) %>%
+  pivot_wider(values_from = n, names_from = exp) %>%
+  mutate(Total = `Period I` + `Period II` + `Period III`) %>%
+  rename(`Age group` = age_group) %>%
+  kableExtra::kable(format = "latex", booktabs = TRUE)
+```
+
+```{r}
+vi_survey %>%
+  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
+  mutate(pronoun = factor(pronoun, levels = c("He", "She", "Other"))) %>%
+  group_by(exp, set) %>%
+  summarise(across(everything(), first)) %>%
+  count(education) %>%
+  mutate(exp = paste("Period", c("I", "II", "III")[exp - 2])) %>%
+  pivot_wider(values_from = n, names_from = exp, values_fill = 0) %>%
+  mutate(Total = `Period I` + `Period II` + `Period III`) %>%
+  slice(c(2, 1, 3, 4, 5)) %>%
+  rename(`Education` = education) %>%
+  kableExtra::kable(format = "latex", booktabs = TRUE)
+```
+
+```{r}
+vi_survey %>%
+  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
+  mutate(pronoun = factor(pronoun, levels = c("He", "She", "Other"))) %>%
+  group_by(exp, set) %>%
+  summarise(across(everything(), first)) %>%
+  count(previous_experience) %>%
+  mutate(exp = paste("Period", c("I", "II", "III")[exp - 2])) %>%
+  pivot_wider(values_from = n, names_from = exp) %>%
+  mutate(Total = `Period I` + `Period II` + `Period III`) %>%
+  rename(`Previous experience` = previous_experience) %>%
+  kableExtra::kable(format = "latex", booktabs = TRUE)
+```
 

### 2023-03-21 paper_comparison/appendix.Rmd
adjust text for demographic

--- 2023-03-21 paper_comparison/appendix.Rmd
@@
 Along with the responses to lineups, we have collected a series of demographic information including age, pronoun, education background and previous experience in studies involved data visualization.
-Figure \ref{fig:education} and \ref{fig:experience} provide summary of the demographic data.
+Table \ref{tab:pronoun}, \ref{tab:age-group}, \ref{tab:education} and \ref{tab:experience} provide summary of the demographic data.
 
-As shown in Figure \ref{fig:education}, most participants have Diploma or Bachelor degrees, followed by High school or below.
-The survey data is gender balanced and the majority of participants are between 18 to 39 years old.
+It can be observed from the tables that most participants have Diploma or Bachelor degrees, followed by High school or below and the survey data is gender balanced.
+Majority of participants are between 18 to 39 years old and there are slightly more participants who do not have previous experience than those who have.
 
-```{r education, fig.cap = "\\label{fig:education}Education distribution of subjects recuritted in this study group by age and pronoun.", fig.height = 4}
+```{r}
 vi_survey %>%
@@
   summarise(across(everything(), first)) %>%
-  ggplot() +
-  geom_bar(aes(age_group, fill = pronoun), position = "dodge") +
-  facet_wrap(~fct_relevel(education, "High School or below", "Diploma and Bachelor Degree", "Honours Degree", "Masters Degree")) +
-  xlab("Age group") +
-  ylab("Count") +
-  theme_light(base_size = 10)
+  count(pronoun) %>%
+  mutate(exp = paste("Period", c("I", "II", "III")[exp - 2])) %>%
+  pivot_wider(values_from = n, names_from = exp) %>%
+  mutate(Total = `Period I` + `Period II` + `Period III`) %>%
+  rename(Pronoun = pronoun) %>%
+  kableExtra::kable(format = "latex", booktabs = TRUE, label = "pronoun", caption = "Summary of pronoun distribution of subjects recuritted in this study.")
 ```
 
-Figure \ref{fig:experience} shows no particular difference between the number of participants who have previous experience and the number of those who haven't.
-Age distributions are also similar for these two groups.
-
-```{r experience, fig.cap = "\\label{fig:experience}Status of pervious experience of subjects recuritted in this study group by age and pronoun."}
+```{r}
 vi_survey %>%
@@
   summarise(across(everything(), first)) %>%
-  ggplot() +
-  geom_bar(aes(age_group, fill = pronoun), position = "dodge") +
-  facet_wrap(~previous_experience) +
-  xlab("Age group") +
-  ylab("Count") +
-  theme_light()
+  count(age_group) %>%
+  mutate(exp = paste("Period", c("I", "II", "III")[exp - 2])) %>%
+  pivot_wider(values_from = n, names_from = exp) %>%
+  mutate(Total = `Period I` + `Period II` + `Period III`) %>%
+  rename(`Age group` = age_group) %>%
+  kableExtra::kable(format = "latex", booktabs = TRUE, label = "age-group", caption = "Summary of age distribution of subjects recuritted in this study.")
+```
+
+```{r}
+vi_survey %>%
+  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
+  mutate(pronoun = factor(pronoun, levels = c("He", "She", "Other"))) %>%
+  group_by(exp, set) %>%
+  summarise(across(everything(), first)) %>%
+  count(education) %>%
+  mutate(exp = paste("Period", c("I", "II", "III")[exp - 2])) %>%
+  pivot_wider(values_from = n, names_from = exp, values_fill = 0) %>%
+  mutate(Total = `Period I` + `Period II` + `Period III`) %>%
+  slice(c(2, 1, 3, 4, 5)) %>%
+  rename(`Education` = education) %>%
+  kableExtra::kable(format = "latex", booktabs = TRUE, label = "education", caption = "Summary of education distribution of subjects recuritted in this study.")
+```
+
+```{r}
+vi_survey %>%
+  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
+  mutate(pronoun = factor(pronoun, levels = c("He", "She", "Other"))) %>%
+  group_by(exp, set) %>%
+  summarise(across(everything(), first)) %>%
+  count(previous_experience) %>%
+  mutate(exp = paste("Period", c("I", "II", "III")[exp - 2])) %>%
+  pivot_wider(values_from = n, names_from = exp) %>%
+  mutate(Total = `Period I` + `Period II` + `Period III`) %>%
+  rename(`Previous experience` = previous_experience) %>%
+  kableExtra::kable(format = "latex", booktabs = TRUE, label = "experience", caption = "Summary of previous experience distribution of subjects recuritted in this study.")
 ```

### 2023-03-21 paper_comparison/appendix.Rmd
use quantiles instead of LVS

@@
 To build a bridge between various residual-based tests, and the visual test, we focus on the shared information embedded in the testing procedures, which is the distribution of residuals.
-When comes to comparison of distribution, Kullback-Leibler divergence [@@kullback1951information] is a classical way to represent the information loss or entropy increase caused by the approximation to the true distribution, which in our case, the inefficiency due to the use of false model assumptions.
+When comes to comparison of distribution, Kullback-Leibler divergence [@kullback1951information] is a classical way to represent the information loss or entropy increase caused by the approximation to the true distribution, which in our case, the inefficiency due to the use of false model assumptions.
 
@@
 
-## Batch effect
+## Effect of data collection period
 
@@
   ggplot() +
-  geom_lv(aes(x = exp, y = prop_detect, fill = after_stat(LV)), varwidth = TRUE) +
+  geom_lv(aes(x = exp, y = prop_detect, fill = after_stat(LV)), varwidth = TRUE, k = 4) +
+  scale_alpha_discrete(range = c(0.7, 0)) +
   # scale_fill_lv() +
@@
   scale_x_discrete(labels = c("I", "III")) +
+  labs(fill = "Quantiles") +
+  scale_fill_discrete(labels = c("50%", "25% - 75%", "12.5% - 87.5%", "6.25% - 93.75%")) +
   ylab("Weighted propotion of detect") +
@@
   ggplot() +
-  geom_lv(aes(x = exp, y = prop_detect, fill = after_stat(LV)), varwidth = TRUE) +
-  # scale_fill_lv() +
+  geom_lv(aes(x = exp, y = prop_detect, fill = after_stat(LV)), varwidth = TRUE, k = 4) +
   facet_wrap(~k, ncol = 5) +
@@
   scale_x_discrete(labels = c("II", "III")) +
+  labs(fill = "Quantiles") +
+  scale_fill_discrete(labels = c("50%", "25% - 75%", "12.5% - 87.5%", "6.25% - 93.75%")) +
   ylab("Weighted propotion of detect") +

### 2023-03-21 paper_comparison/appendix.Rmd
update text for lv

@@
 
-```{r batch, fig.cap = '\\label{fig:batch}A "letter-value" boxplot of weighted propotion of detect for lineups over different data collection periods.'}
-vi_survey %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(x_dist == "uniform") %>%
-  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
-  mutate(exp = exp - 2) %>%
-  group_by(exp, unique_lineup_id) %>%
-  mutate(prop_detect = mean(weighted_detect)) %>%
-  group_by(exp, unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  ungroup() %>%
-ggplot() +
-  geom_lv(aes(x = factor(exp), y = prop_detect, fill = after_stat(LV)), varwidth = TRUE) +
-  # scale_fill_lv() +
-  facet_grid(~type) +
-  theme_light()
-```
-
-```{r}
-t_test_poly <- vi_survey %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(x_dist == "uniform") %>%
-  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
-  mutate(exp = exp - 2) %>%
-  group_by(exp, unique_lineup_id) %>%
-  mutate(prop_detect = mean(weighted_detect)) %>%
-  group_by(exp, unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type == "non-linearity") %>%
-  t.test(prop_detect ~ exp, data = .)
-```
-
-```{r}
-t_test_heter <- vi_survey %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(x_dist == "uniform") %>%
-  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
-  mutate(exp = exp - 2) %>%
-  group_by(exp, unique_lineup_id) %>%
-  mutate(prop_detect = mean(weighted_detect)) %>%
-  group_by(exp, unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  filter(type != "non-linearity") %>%
-  t.test(prop_detect ~ exp, data = .)
-```
-
 We have the same type of model collected over different data collection periods, that may lead to unexpected batch effect.
@@
 Within our research team, we can not identify the data plot from the null plots for these two lineups, result in $p$-values much greater than $5$%.
-
-Figure \ref{fig:batch} shows the data plot, where the distribution for the heteroskedasticity patterns are almost identical in period II and III, while the distribution for the non-linearity patterns have a shift in the median.
-The $p$-values of the t-test for testing the mean difference across data periods are `r round(t_test_poly$p.value, 2)` and `r round(t_test_heter$p.value, 2)` for non-linearity and heteroskedasticity patterns respectively.
-Both are much greater than 5%.
-Therefore, both the visual test and the conventional test, there is no clear evidence of batch effect.
+Thus, there is no clear evidence of batch effect.
 

### 2023-03-21 paper_comparison/appendix.Rmd
update sensitivity analysis

@@
 
-The analysis is conducted by setting up two relatively extreme scenarios, where the $\alpha$ is under or overestimated by 50%.
-Using the inflated and deflated $\hat{\alpha}$, we recalculate the $p$-value for every lineup and show the results in Figure \ref{fig:sensitive}.
+The analysis is conducted by setting up several scenarios, where the $\alpha$ is under or overestimated by 12.5%, 25%, 50% and 100%.
+Using the adjusted $\hat{\alpha}$, we recalculate the $p$-value for every lineup and show the results in Figure \ref{fig:sensitive}.
 It can be observed that there are some changes to $p$-values, especially when the $\hat{\alpha}$ is deflated.
@@
 ```{r}
-vi_survey_upper <- calc_p_value_multi(vi_survey %>%
-                                        mutate(alpha_upper = alpha * 1.5,
-                                               alpha_lower = alpha * 0.5),
+upper_p <- map_df(c(0.125, 0.25, 0.5), function(modified) {
+  calc_p_value_multi(vi_survey %>%
+                       filter(!attention_check) %>%
+                       filter(!null_lineup) %>%
+                       filter(x_dist == "uniform") %>%
+                       mutate(alpha_upper = alpha * (1 + modified)),
                      lineup_id = unique_lineup_id,
@@
                      alpha = alpha_upper) %>%
-  right_join(select(vi_survey, -p_value))
+  mutate(modified = 1 + modified) %>%
+  mutate(adjusted_p_value = p_value)
+})
 
-vi_survey_lower <- calc_p_value_multi(vi_survey %>%
-                                        mutate(alpha_upper = alpha * 1.5,
-                                               alpha_lower = alpha * 0.5),
+lower_p <- map_df(c(0.125, 0.25, 0.5), function(modified) {
+  calc_p_value_multi(vi_survey %>%
+                       filter(!attention_check) %>%
+                       filter(!null_lineup) %>%
+                       filter(x_dist == "uniform") %>%
+                       mutate(alpha_lower = alpha * (1 - modified)),
                      lineup_id = unique_lineup_id,
@@
                      alpha = alpha_lower) %>%
-  right_join(select(vi_survey, -p_value))
+  mutate(modified = modified) %>%
+  mutate(adjusted_p_value = p_value)
+})
+
+upper_lower_p <- upper_p %>%
+  bind_rows(lower_p) %>%
+  select(-p_value)
 ```
 
-```{r cache = TRUE, eval = FALSE}
-# Define the minimum and maximum effect size
-min_poly_es <- vi_survey_upper %>% 
-  filter(!null_lineup, 
-         !attention_check,
-         type == "polynomial", 
-         x_dist == "uniform") %>%
-  pull(effect_size) %>%
-  min()
-max_poly_es <- vi_survey_upper %>% 
-  filter(!null_lineup, 
-         !attention_check,
-         type == "polynomial", 
-         x_dist == "uniform") %>%
-  pull(effect_size) %>%
-  max()
-
-# The glm model for RESET test
-reset_poly_glmmod <- poly_conv_sim %>%
-  filter(x_dist == "uniform") %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  mutate(reject = as.numeric(RESET4_p_value <= 0.05)) %>%
-  mutate(offset0 = log(0.05/0.95)) %>%
-  select(x_dist, effect_size, log_effect_size, offset0, reject) %>%
-  glm(reject ~ effect_size - 1, family = binomial(), data = ., offset = offset0)
-
-# The prediction for all conventional tests on poly model
-pred_conv_poly <- poly_conv_sim %>%
-  filter(x_dist == "uniform") %>%
-  select(-RESET3_p_value, -(RESET5_p_value:RESET10_p_value), -F_p_value) %>%
-  rename(RESET_p_value = RESET4_p_value) %>%
-  pivot_longer(RESET_p_value:SW_p_value) %>%
-  mutate(name = gsub("_p_value", " test", name)) %>%
-  mutate(reject = value <= 0.05) %>%
-  select(effect_size, name, reject) %>%
-  mutate(offset0 = log(0.05/0.95)) %>%
-  nest(dat = c(effect_size, offset0, reject)) %>%
-  mutate(mod = map(dat, 
-                   ~glm(reject ~ effect_size - 1, 
-                        family = binomial(), 
-                        data = .x,
-                        offset = offset0))) %>%
-  mutate(power = map(mod, function(mod) {
-    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
-                         offset0 = log(0.05/0.95))
-    result$power <- predict(mod, type = "response", newdata = result)
-    result
-  })) %>%
-  select(-dat, -mod) %>%
-  unnest(power) %>%
-  mutate(log_effect_size = log(effect_size))
-
-# Data needed by the visual test
-visual_poly_dat <- vi_survey_upper %>%
-  filter(!null_lineup, 
-         !attention_check, 
-         type == "polynomial", 
-         x_dist == "uniform") %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  mutate(reject = as.numeric(p_value <= 0.05)) %>%
-  mutate(offset0 = log(0.05/0.95)) %>%
-  select(effect_size, offset0, reject)
-
-# The glm model for visual test
-visual_poly_glmmod <- glm(reject ~ effect_size - 1, 
-                          family = binomial(), 
-                          data = visual_poly_dat, 
-                          offset = offset0)
-
-# The prediction of the visual test
-pred_visual_poly <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
-                               offset0 = log(0.05/0.95)) %>%
-  mutate(power = predict(visual_poly_glmmod, 
-                         type = "response",
-                         newdata = data.frame(effect_size = effect_size,
-                                              offset0 = offset0))) %>%
-  mutate(log_effect_size = log(effect_size))
-
-# The prediction of the boot visual test
-pred_visual_poly_boot <- map_df(1:500, function(boot_id) {
-  slice_sample(visual_poly_dat, n = nrow(visual_poly_dat), replace = TRUE) %>%
-    nest(dat = c(effect_size, offset0, reject)) %>%
-    mutate(mod = map(dat, 
-                     ~glm(reject ~ effect_size - 1, 
-                          family = binomial(), 
-                          data = .x,
-                          offset = offset0))) %>%
-    mutate(power = map(mod, function(mod) {
-      result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
-                           offset0 = log(0.05/0.95))
-      result$power <- predict(mod, type = "response", newdata = result)
-      result
-    })) %>%
-    select(-dat, -mod) %>%
-    unnest(power) %>%
-    mutate(log_effect_size = log(effect_size)) %>%
-    mutate(boot_id = boot_id)
-  })
-```
-
-```{r eval = FALSE}
-ggplot() +
-
-  geom_point(data = visual_poly_dat,
-             aes(log(effect_size), reject),
-             alpha = 0.15) +
-
-  # Draw conventional test power curve
-  geom_line(data = pred_conv_poly,
-            aes(log_effect_size, power, col = name),
-            size = 1) +
-
-  # Draw boot visual test curves
-  geom_line(data = pred_visual_poly_boot,
-            aes(log_effect_size, power, col = "Visual test", group = boot_id),
-            size = 1,
-            alpha = 0.01) +
-
-  # Draw visual test power curve
-  geom_line(data = pred_visual_poly,
-            aes(log_effect_size, power, col = "Visual test"), 
-            size = 1) +
-
-  # Theme
-  theme_light(base_size = 5) +
-  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.5)) +
-  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
-  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
-  xlab(expression(log[e] (Effect_size))) +
-  ylab("Power") +
-  labs(col = "", size = "# lineups")
-```
-
-```{r sensitive, fig.cap = "\\label{fig:sensitive}Change of $p$-values with $\\hat{\\alpha}$ inflated by 50% and deflated by 50%."}
-vi_survey %>%
+```{r sensitive, fig.cap = "\\label{fig:sensitive}Change of $p$-values with $\\hat{\\alpha}$ multiplied by $0.125$, $0.25$, $0.5$, $1.125$, $1.25$ and $1.5$."}
+upper_lower_p %>%
+  left_join(vi_survey %>%
   filter(!attention_check) %>%
@@
   group_by(unique_lineup_id) %>%
-  summarise(p_value = first(p_value)) %>%
-  left_join(
-vi_survey_upper %>%
-  group_by(unique_lineup_id) %>%
-  summarise(p_value_upper = first(p_value))) %>%
-  left_join(
-vi_survey_lower %>%
-  group_by(unique_lineup_id) %>%
-  summarise(p_value_lower = first(p_value))) %>%
-  pivot_longer(p_value_upper:p_value_lower) %>%
-  mutate(name = ifelse(name == "p_value_upper", "+50%", "-50%")) %>%
+  summarise(p_value = first(p_value))) %>%
   ggplot() +
-  geom_point(aes(p_value, value), alpha = 0.2) +
+  geom_point(aes(p_value, adjusted_p_value), alpha = 0.2) +
   geom_abline() +
+  facet_wrap(~modified) +
   scale_x_sqrt() +
   scale_y_sqrt() +
-  facet_wrap(~name) +
   ylab("Adjusted p-value") +
@@
 
-```{r sensitive-barplot, fig.cap = "\\label{fig:sensitive-barplot}Change of rejection of the null hypothesis with $\\hat{\\alpha}$ inflated by 50% and deflated by 50%."}
-vi_survey %>%
+```{r}
+upper_lower_p %>%
+  left_join(vi_survey %>%
   filter(!attention_check) %>%
@@
   group_by(unique_lineup_id) %>%
-  summarise(p_value = first(p_value)) %>%
-  left_join(
-vi_survey_upper %>%
-  group_by(unique_lineup_id) %>%
-  summarise(p_value_upper = first(p_value))) %>%
-  left_join(
-vi_survey_lower %>%
-  group_by(unique_lineup_id) %>%
-  summarise(p_value_lower = first(p_value))) %>%
-  pivot_longer(p_value_upper:p_value_lower) %>%
-  mutate(name = ifelse(name == "p_value_upper", "+50%", "-50%")) %>%
-  ggplot() +
-  geom_bar(aes(p_value < 0.05, fill = value < 0.05), position = "dodge") +
-  theme_light() +
-  xlab("p-value < 0.05") +
-  labs(fill = "Adjusted p-value < 0.05")
-```
-
-```{r cache = TRUE, eval = FALSE}
-# Define the minimum and maximum effect size
-min_heter_es <- vi_survey_upper %>% 
-  filter(!null_lineup, 
-         !attention_check,
-         type == "heteroskedasticity", 
-         x_dist == "uniform") %>%
-  pull(effect_size) %>%
-  min()
-max_heter_es <- vi_survey_upper %>% 
-  filter(!null_lineup, 
-         !attention_check,
-         type == "heteroskedasticity", 
-         x_dist == "uniform") %>%
-  pull(effect_size) %>%
-  max()
-
-# The prediction for all conventional tests on heter model
-pred_conv_heter <- heter_conv_sim %>%
-  filter(x_dist == "uniform") %>%
-  select(-RESET3_p_value, -(RESET5_p_value:RESET10_p_value), -F_p_value) %>%
-  rename(RESET_p_value = RESET4_p_value) %>%
-  pivot_longer(RESET_p_value:SW_p_value) %>%
-  mutate(name = gsub("_p_value", " test", name)) %>%
-  mutate(reject = value <= 0.05) %>%
-  select(effect_size, name, reject) %>%
-  mutate(offset0 = log(0.05/0.95)) %>%
-  nest(dat = c(effect_size, offset0, reject)) %>%
-  mutate(mod = map(dat, 
-                   ~glm(reject ~ effect_size - 1, 
-                        family = binomial(), 
-                        data = .x,
-                        offset = offset0))) %>%
-  mutate(power = map(mod, function(mod) {
-    result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 1),
-                         offset0 = log(0.05/0.95))
-    result$power <- predict(mod, type = "response", newdata = result)
-    result
-  })) %>%
-  select(-dat, -mod) %>%
-  unnest(power) %>%
-  mutate(log_effect_size = log(effect_size))
-
-# Data needed by the visual test
-visual_heter_dat <- vi_survey_upper %>%
-  filter(!null_lineup, 
-         !attention_check, 
-         type == "heteroskedasticity", 
-         x_dist == "uniform") %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  mutate(reject = as.numeric(p_value <= 0.05)) %>%
-  mutate(offset0 = log(0.05/0.95)) %>%
-  select(effect_size, offset0, reject)
-
-# The glm model for visual test
-visual_heter_glmmod <- glm(reject ~ effect_size - 1, 
-                           family = binomial(), 
-                           data = visual_heter_dat, 
-                           offset = offset0)
-
-# The prediction of the visual test
-pred_visual_heter <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
-                                offset0 = log(0.05/0.95)) %>%
-  mutate(power = predict(visual_heter_glmmod, 
-                         type = "response",
-                         newdata = data.frame(effect_size = effect_size,
-                                              offset0 = offset0))) %>%
-  mutate(log_effect_size = log(effect_size))
-
-# The prediction of the boot visual test
-pred_visual_heter_boot <- map_df(1:500, function(boot_id) {
-  slice_sample(visual_heter_dat, n = nrow(visual_heter_dat), replace = TRUE) %>%
-    nest(dat = c(effect_size, offset0, reject)) %>%
-    mutate(mod = map(dat, 
-                     ~glm(reject ~ effect_size - 1, 
-                          family = binomial(), 
-                          data = .x,
-                          offset = offset0))) %>%
-    mutate(power = map(mod, function(mod) {
-      result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
-                           offset0 = log(0.05/0.95))
-      result$power <- predict(mod, type = "response", newdata = result)
-      result
-    })) %>%
-    select(-dat, -mod) %>%
-    unnest(power) %>%
-    mutate(log_effect_size = log(effect_size)) %>%
-    mutate(boot_id = boot_id)
-  })
-```
-
-```{r eval = FALSE}
-# Analyse uniform heteroskedasticity data
-ggplot() +
-
-  geom_point(data = visual_heter_dat,
-             aes(log(effect_size), reject),
-             alpha = 0.15) +
-
-  # Draw conventional test power curve
-  geom_line(data = pred_conv_heter,
-            aes(log_effect_size, power, col = name),
-            size = 1) +
-
-  # Draw boot visual test curves
-  geom_line(data = pred_visual_heter_boot,
-            aes(log_effect_size, power, col = "Visual test", group = boot_id),
-            size = 1,
-            alpha = 0.01) +
-
-  # Draw visual test power curve
-  geom_line(data = pred_visual_heter,
-            aes(log_effect_size, power, col = "Visual test"), 
-            size = 1) +
-
-  # Theme
-  theme_light(base_size = 5) +
-  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.5)) +
-  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
-  # scale_size_manual(values = seq(1, 9)) +
-  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
-  xlab(expression(log[e] (Effect_size))) +
-  ylab("Power") +
-  labs(col = "", size = "# lineups")
+  summarise(p_value = first(p_value))) %>%
+  group_by(modified) %>%
+  count(adjusted_reject = adjusted_p_value <= 0.05, reject = p_value <= 0.05) %>%
+  complete(adjusted_reject, reject, fill = list(n = 0)) %>%
+  ungroup() %>%
+  filter(adjusted_reject != reject) %>%
+  mutate(case = ifelse(!adjusted_reject, "p-value < 0.05 but adjusted p-value > 0.05", "p-value > 0.05 but adjusted p-value < 0.05")) %>%
+  rename(multiplier = modified) %>%
+  select(-adjusted_reject, -reject) %>%
+  pivot_wider(names_from = case, values_from = n) %>%
+  kableExtra::kable(format = "latex", booktabs = TRUE, label = "sensitivity", 
+                    escape = FALSE,
+                    caption = "Change of rejection decision because of the modification of $\\alpha$.")
 ```

### 2023-03-21 paper_comparison/appendix.Rmd
update text for alpha

@@
 The parameter $\alpha$ used for the $p$-value calculation needs to be estimated from responses to null lineups.
+With a greater value of $\hat{\alpha}$, the $p$-value will be smaller, resulting in more lineups being rejected.
 However, The way we generate Rorschach lineup is not strictly the same as what suggested in @vanderplas2021statistical and @buja_statistical_2009.
@@
 
-The analysis is conducted by setting up several scenarios, where the $\alpha$ is under or overestimated by 12.5%, 25%, 50% and 100%.
+The analysis is conducted by setting up several scenarios, where the $\alpha$ is under or overestimated by 12.5%, 25% and 50%.
 Using the adjusted $\hat{\alpha}$, we recalculate the $p$-value for every lineup and show the results in Figure \ref{fig:sensitive}.
-It can be observed that there are some changes to $p$-values, especially when the $\hat{\alpha}$ is deflated.
-However, Figure \ref{fig:sensitive-barplot} shows that adjusting $\hat{\alpha}$ will not result in a huge difference in rejection decisions.
-There are only a small number of cases (12 out of 558) where the rejection decision changes.
+It can be observed that there are some changes to $p$-values, especially when the $\hat{\alpha}$ is multiplied by 50%.
+However, Table \ref{tab:sensitive} shows that adjusting $\hat{\alpha}$ will not result in a huge difference in rejection decisions.
+There are only a small number of cases where the rejection decision change, considering there are a total of 279 lineups with uniform fitted value distribution.
 It is very unlikely the downstream findings will be affected because of the estimate of $\alpha$.
@@
                     escape = FALSE,
-                    caption = "Change of rejection decision because of the modification of $\\alpha$.")
+                    caption = "Change of rejection decision because of the modification of $\\hat{\\alpha}$.")
 ```

### 2023-03-21 paper_comparison/appendix.Rmd
replace cases with prop

@@
 It can be observed that there are some changes to $p$-values, especially when the $\hat{\alpha}$ is multiplied by 50%.
-However, Table \ref{tab:sensitive} shows that adjusting $\hat{\alpha}$ will not result in a huge difference in rejection decisions.
-There are only a small number of cases where the rejection decision change, considering there are a total of 279 lineups with uniform fitted value distribution.
+However, Table \ref{tab:sensitivity} shows that adjusting $\hat{\alpha}$ will not result in a huge difference in rejection decisions.
+There are only a small percentage of cases where the rejection decision change.
 It is very unlikely the downstream findings will be affected because of the estimate of $\alpha$.
@@
   filter(adjusted_reject != reject) %>%
-  mutate(case = ifelse(!adjusted_reject, "p-value < 0.05 but adjusted p-value > 0.05", "p-value > 0.05 but adjusted p-value < 0.05")) %>%
+  mutate(case = ifelse(!adjusted_reject, "Propotion of lineups with p-value < 0.05 but adjusted p-value > 0.05", "Propotion of lineups with p-value > 0.05 but adjusted p-value < 0.05")) %>%
   rename(multiplier = modified) %>%
+  mutate(n = paste0(round(n/279 * 100, 2), "\\%")) %>%
   select(-adjusted_reject, -reject) %>%

### 2023-03-21 paper_comparison/appendix.Rmd
add conv x_dist plot for appendix

@@
   filter(adjusted_reject != reject) %>%
-  mutate(case = ifelse(!adjusted_reject, "Propotion of lineups with p-value < 0.05 but adjusted p-value > 0.05", "Propotion of lineups with p-value > 0.05 but adjusted p-value < 0.05")) %>%
+  mutate(case = ifelse(!adjusted_reject, "Propotion of lineups with $p$-value < 0.05 but adjusted $p$-value > 0.05", "Propotion of lineups with $p$-value > 0.05 but adjusted $p$-value < 0.05")) %>%
   rename(multiplier = modified) %>%
@@
 
+```{r cache = TRUE}
+min_poly_es <- vi_survey %>% 
+  filter(!null_lineup, 
+         !attention_check,
+         type == "polynomial", 
+         x_dist == "uniform") %>%
+  pull(effect_size) %>%
+  min()
+
+max_poly_es <- vi_survey %>% 
+  filter(!null_lineup, 
+         !attention_check,
+         type == "polynomial", 
+         x_dist == "uniform") %>%
+  pull(effect_size) %>%
+  max()
+
+min_heter_es <- vi_survey %>% 
+  filter(!null_lineup, 
+         !attention_check,
+         type == "heteroskedasticity", 
+         x_dist == "uniform") %>%
+  pull(effect_size) %>%
+  min()
+
+max_heter_es <- vi_survey %>% 
+  filter(!null_lineup, 
+         !attention_check,
+         type == "heteroskedasticity", 
+         x_dist == "uniform") %>%
+  pull(effect_size) %>%
+  max()
+
+poly_x_dist <- poly_conv_sim %>%
+  select(-RESET3_p_value, -(RESET5_p_value:RESET10_p_value), -F_p_value) %>%
+  rename(RESET_p_value = RESET4_p_value) %>%
+  pivot_longer(RESET_p_value:SW_p_value) %>%
+  mutate(name = gsub("_p_value", " test", name)) %>%
+  filter(name == "RESET test") %>%
+  mutate(reject = value <= 0.05) %>%
+  select(effect_size, name, x_dist, reject) %>%
+  mutate(offset0 = log(0.05/0.95)) %>%
+  nest(dat = c(effect_size, offset0, reject)) %>%
+  mutate(mod = map(dat, 
+                   ~glm(reject ~ effect_size - 1, 
+                        family = binomial(), 
+                        data = .x,
+                        offset = offset0))) %>%
+  mutate(power = map(mod, function(mod) {
+    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
+                         offset0 = log(0.05/0.95))
+    result$power <- predict(mod, type = "response", newdata = result)
+    result
+  })) %>%
+  select(-dat, -mod) %>%
+  unnest(power) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  mutate(type = "non-linearity")
+
+heter_x_dist <- heter_conv_sim %>%
+  select(-RESET3_p_value, -(RESET5_p_value:RESET10_p_value), -F_p_value) %>%
+  rename(RESET_p_value = RESET4_p_value) %>%
+  pivot_longer(RESET_p_value:SW_p_value) %>%
+  mutate(name = gsub("_p_value", " test", name)) %>%
+  filter(name == "BP test") %>%
+  mutate(reject = value <= 0.05) %>%
+  select(effect_size, name, x_dist, reject) %>%
+  mutate(offset0 = log(0.05/0.95)) %>%
+  nest(dat = c(effect_size, offset0, reject)) %>%
+  mutate(mod = map(dat, 
+                   ~glm(reject ~ effect_size - 1, 
+                        family = binomial(), 
+                        data = .x,
+                        offset = offset0))) %>%
+  mutate(power = map(mod, function(mod) {
+    result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
+                         offset0 = log(0.05/0.95))
+    result$power <- predict(mod, type = "response", newdata = result)
+    result
+  })) %>%
+  select(-dat, -mod) %>%
+  unnest(power) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  mutate(type = "heteroskedasticity")
+```
+
+```{r conv-x-dist, fig.cap = "Change of power of conventional tests under different fitted value distribution for non-linearity model and heteroskedasticity model.", fig.height = 4}
+bind_rows(poly_x_dist, heter_x_dist) %>%
+  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
+  ggplot() +
+
+  # Draw conventional test power curve
+  geom_line(aes(log_effect_size, power, col = x_dist),
+            size = 1) +
+
+  # Theme
+  theme_light() +
+  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
+  xlab(expression(log[e] (Effect_size))) +
+  ylab("Power") +
+  labs(col = "", size = "# lineups") +
+  facet_wrap(~type, scales = "free_x")
+```
+
 ## Effect of number of evaluations on the power of a visual test

### 2023-03-21 paper_comparison/paper_comparison/paper_comparison.Rmd
unnumbered sections
fix the seq function

@@
 
-(put citation https://www.tandfonline.com/doi/full/10.1080/26939169.2021.1920866)
-
 More typically, residuals will be plotted, as a scatter plot against the predicted values and each of the explanatory variables to scrutinize their relationships.
@@
 
-# Acknowledgements
+# Acknowledgements {-}
 
@@
 
-# Supplementary material
+# Supplementary material {-}
 

### 2023-03-21 paper_comparison/appendix.Rmd
add number of eval plot

@@
 
-```{r conv-x-dist, fig.cap = "Change of power of conventional tests under different fitted value distribution for non-linearity model and heteroskedasticity model.", fig.height = 4}
+```{r conv-x-dist, fig.cap = "Change of power of conventional tests under different fitted value distribution for non-linearity model and heteroskedasticity model.", fig.height = 3}
 bind_rows(poly_x_dist, heter_x_dist) %>%
@@
 
+```{r}
+eval_n <- function(n_eval, boot) {
+  map_df(1:boot, function(boot_id) {
+  vi_survey %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
+  filter(x_dist == "uniform") %>%
+  filter(type == "polynomial") %>%
+  group_by(unique_lineup_id) %>%
+  slice_sample(n = n_eval) %>%
+  ungroup() %>%
+  calc_p_value_multi(lineup_id = unique_lineup_id,
+                     detect = detect,
+                     n_sel = num_selection,
+                     alpha = alpha) %>%
+  mutate(boot_id = boot_id)
+}) %>%
+  left_join(select(vi_survey, unique_lineup_id, effect_size)) %>%
+  mutate(reject = as.numeric(p_value <= 0.05)) %>%
+  mutate(offset0 = log(0.05/0.95)) %>%
+  select(-unique_lineup_id, -p_value) %>%
+  mutate(n_eval = n_eval)
+}
+
+map_df(seq(1, 11, 2), ~eval_n(.x, 1)) %>%
+  select(-boot_id) %>%
+  nest(c(effect_size, reject, offset0)) %>%
+  mutate(mod = map(data, 
+                   ~glm(reject ~ effect_size - 1, 
+                        family = binomial(), 
+                        data = .x,
+                        offset = offset0))) %>%
+  mutate(power = map(mod, function(mod) {
+    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
+                         offset0 = log(0.05/0.95))
+    result$power <- predict(mod, type = "response", newdata = result)
+    result
+  })) %>%
+  select(-data, -mod) %>%
+  unnest(power) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  ggplot() +
+  geom_line(aes(log_effect_size, power, col = factor(n_eval)),
+            size = 1) +
+  theme_light() +
+  scale_color_brewer(palette = "PuBuGn")
+```
+
 ## Power of a RESET test under different auxiliary regression formulas
+
+```{r}
+
+```
 

### 2023-03-21 paper_comparison/appendix.Rmd
add different RESET plot

@@
   theme_light() +
-  scale_color_brewer(palette = "PuBuGn")
+  scale_color_brewer(palette = "Reds")
 ```
@@
 
-```{r}
-
+```{r cache = TRUE}
+reset_diff_dat <- poly_conv_sim %>%
+  filter(x_dist == "uniform") %>%
+  select(-F_p_value, -BP_p_value, -SW_p_value) %>%
+  pivot_longer(RESET3_p_value:RESET10_p_value) %>%
+  mutate(name = gsub("RESET(.*)_p_value", "\\1", name)) %>%
+  mutate(reject = value <= 0.05) %>%
+  select(effect_size, name, reject) %>%
+  mutate(offset0 = log(0.05/0.95)) %>%
+  nest(dat = c(effect_size, offset0, reject)) %>%
+  mutate(mod = map(dat, 
+                   ~glm(reject ~ effect_size - 1, 
+                        family = binomial(), 
+                        data = .x,
+                        offset = offset0))) %>%
+  mutate(power = map(mod, function(mod) {
+    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
+                         offset0 = log(0.05/0.95))
+    result$power <- predict(mod, type = "response", newdata = result)
+    result
+  })) %>%
+  select(-dat, -mod) %>%
+  unnest(power) %>%
+  mutate(log_effect_size = log(effect_size))
 ```
 
-1. put #eval in appendix
-3. compute the power of RESET test with different orders (considering dof) (focus on simple pattern possibly?)
+```{r}
+reset_diff_dat %>%
+  ggplot() +
+  geom_line(aes(log_effect_size, power, col = as.factor(as.integer(name))),
+            size = 1) +
+  theme_light() +
+  scale_color_brewer(palette = "Reds")
+```
 

### 2023-03-22 paper_comparison/paper_comparison/paper_comparison.Rmd
add back conv for x_dist

@@
   mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
   mutate(reject = as.numeric(p_value <= 0.05)) %>%
@@
   select(-unique_lineup_id, -p_value)
-```
-
-```{r different-x-dist-poly-power, fig.height = 4, fig.cap = "Comparison of power on lineups with different fitted value distributions for visual tests.
-All the power curves are estimated using five evaluations on each lineup.
-For lineups with uniform fitted value distribution, the five evaluations are randomly sampled from the total eleven evaluations.
-The sampling process has been repeated for 500 times and the corresponding power curves are drawn with transparent grey lines.
-Uneven distributions including normal and lognormal distributions produce lower power for both types of residual departures.
-Meanwhile, the lognormal distribution as an asymmetric distribution produce the lowest power.", dev = 'png', dpi = 450}
-
-# Extract three datasets, one for poly conventional, one for heter conventional
-# and one for visual tests.
-
-dat_x_dist <- vi_survey %>%
-    filter(!null_lineup) %>%
-    filter(!attention_check) %>%
-    group_by(unique_lineup_id) %>%
-    summarise(across(everything(), first)) %>%
-    filter(x_dist != "uniform") %>%
-    mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-    mutate(reject = as.numeric(p_value <= 0.05)) %>%
-    mutate(x_dist = ifelse(x_dist == "even_discrete", 
-                           "Discrete uniform", 
-                           stringr::str_to_title(x_dist))) %>%
-    mutate(offset0 = log(0.05/0.95)) %>%
-    select(type, x_dist, effect_size, offset0, reject) %>%
-
-  # Manually control factor ordering
-  mutate(type = factor(type, 
-                       levels = c("non-linearity", 
-                                  "heteroskedasticity"))) %>%
-  mutate(x_dist = factor(x_dist, 
-                         levels = c("Uniform", 
-                                    "Normal", 
-                                    "Lognormal", 
-                                    "Discrete uniform")))
 
@@
   mutate(log_effect_size = log(effect_size))
-
-# Fit the model manually
-dat_x_dist %>%
+```
+
+```{r different-x-dist-poly-power, fig.height = 4, fig.cap = "Comparison of power on lineups with different fitted value distributions for visual tests.
+All the power curves are estimated using five evaluations on each lineup.
+For lineups with uniform fitted value distribution, the five evaluations are randomly sampled from the total eleven evaluations.
+The sampling process has been repeated for 500 times and the corresponding power curves are drawn with transparent grey lines.
+Uneven distributions including normal and lognormal distributions produce lower power for both types of residual departures.
+Meanwhile, the lognormal distribution as an asymmetric distribution produce the lowest power.", dev = 'png', dpi = 450}
+
+diff_x_dist_power_pred <- list(
+  # Visual inference data
+  vi_survey %>%
+    filter(!null_lineup) %>%
+    filter(!attention_check) %>%
+    group_by(unique_lineup_id) %>%
+    summarise(across(everything(), first)) %>%
+    filter(x_dist != "uniform") %>%
+    mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+    mutate(reject = as.numeric(p_value <= 0.05)) %>%
+    mutate(offset0 = log(0.05/0.95)) %>%
+    select(type, x_dist, effect_size, offset0, reject) %>%
+    mutate(conv_or_not = "Visual"),
+
+  # Conventional simulation (poly)
+  poly_conv_sim %>%
+    mutate(reject = RESET4_p_value < 0.05) %>%
+    mutate(offset0 = log(0.05/0.95)) %>%
+    mutate(type = "non-linearity") %>%
+    select(type, x_dist, effect_size, offset0, reject) %>%
+    mutate(conv_or_not = "Conventional"),
+
+  # Conventional simulation (heter)
+  heter_conv_sim %>%
+    mutate(reject = BP_p_value < 0.05) %>%
+    mutate(offset0 = log(0.05/0.95)) %>%
+    mutate(type = "heteroskedasticity") %>%
+    select(type, x_dist, effect_size, offset0, reject) %>%
+    mutate(conv_or_not = "Conventional")
+) %>%
+
+  # Bind rows
+  map_df(~.x) %>%
+
   nest(dat = c(effect_size, offset0, reject)) %>%
@@
   unnest(power) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
+  mutate(log_effect_size = log(effect_size))
+
+diff_x_dist_power_pred %>%
+  mutate(x_dist = ifelse(x_dist == "even_discrete", 
+                           "Discrete", 
+                           stringr::str_to_title(x_dist))) %>%
+  mutate(type = factor(type, 
+                       levels = c("non-linearity", 
+                                  "heteroskedasticity"))) %>%
+  mutate(x_dist = factor(x_dist, 
+                         levels = c("Uniform", 
+                                    "Normal", 
+                                    "Lognormal", 
+                                    "Discrete"))) %>%
 
   ggplot() +
+
+  # Draw all smooth lines
+  geom_line(aes(log_effect_size, power, col = x_dist),
+            size = 1) +
 
@@
 
-  # Draw all smooth lines
-  geom_line(aes(log_effect_size, power, col = x_dist),
-            size = 1) +
-
   # Theme
   theme_light() +
-  scale_color_manual(values = rcartocolor::carto_pal(4, "Safe")[c(2,1,3,4)]) +
+  scale_color_manual(values = rcartocolor::carto_pal(4, "Safe")[c(4,1,3,2)]) +
   xlab(expression(log[e] (Effect_size))) +
@@
   labs(col = "", linetype = "Test") +
-  facet_grid(~type, scales = "free_x") +
+  facet_grid(type~conv_or_not, scales = "free_x") +
   theme(legend.position = "bottom")

### 2023-03-22 paper_comparison/paper_comparison/paper_comparison.Rmd
adjust x_dist figure height

@@
 
-```{r different-x-dist-poly-power, fig.height = 4, fig.cap = "Comparison of power on lineups with different fitted value distributions for visual tests.
+```{r different-x-dist-poly-power, fig.height = 6, fig.cap = "Comparison of power on lineups with different fitted value distributions for visual tests.
 All the power curves are estimated using five evaluations on each lineup.

### 2023-03-22 paper_comparison/paper_comparison/paper_comparison.Rmd
use title case for figure panel

@@
   unnest(power) %>%
-  mutate(log_effect_size = log(effect_size))
+  mutate(log_effect_size = log(effect_size)) %>%
+  mutate(type = stringr::str_to_title(type))
 ```
@@
                            stringr::str_to_title(x_dist))) %>%
-  mutate(type = factor(type, 
-                       levels = c("non-linearity", 
-                                  "heteroskedasticity"))) %>%
+  mutate(type = factor(stringr::str_to_title(type), 
+                       levels = c("Non-Linearity", 
+                                  "Heteroskedasticity"))) %>%
   mutate(x_dist = factor(x_dist, 

### 2023-03-22 paper_comparison/paper_comparison/paper_comparison.Rmd
fix x_dist caption

@@
 
-```{r different-x-dist-poly-power, fig.height = 6, fig.cap = "Comparison of power on lineups with different fitted value distributions for visual tests.
-All the power curves are estimated using five evaluations on each lineup.
+```{r different-x-dist-poly-power, fig.height = 6, fig.cap = "Comparison of power on lineups with different fitted value distributions for conventional tests and visual tests.
+Power curves of visual tests are estimated using five evaluations on each lineup.
 For lineups with uniform fitted value distribution, the five evaluations are randomly sampled from the total eleven evaluations.
 The sampling process has been repeated for 500 times and the corresponding power curves are drawn with transparent grey lines.
-Uneven distributions including normal and lognormal distributions produce lower power for both types of residual departures.
-Meanwhile, the lognormal distribution as an asymmetric distribution produce the lowest power.", dev = 'png', dpi = 450}
+The fitted value distribution has stronger impact on the power of conventional tests than visual tests.
+Meanwhile, uneven distributions including normal and lognormal distributions usually produce lower power.", dev = 'png', dpi = 450}
 

### 2023-03-24 paper_comparison/paper_comparison/paper_comparison.Rmd
update x_dist text

@@
 
-Figure \ref{fig:different-x-dist-poly-power} illustrates the change of power of visual tests for different fitted distributions used in the non-linearity model and the heteroskedasticity model.
-Note that we have collected more evaluations on lineup with uniform fitted value distribution, and the number of evaluations on a lineup will have an impact on the power of the visual test.
-To have a fair comparison, the five evaluations used in the power estimation are randomly sampled from the total eleven evaluations for each lineup with uniform fitted value distribution.
-The sampling process has been repeated 500 times to give an indication of the variation of the power.
-
-We do not observe significant difference between the power curve of the uniform distribution and the power curve of the discrete uniform distribution, but the power curves of normal and lognormal distributions are slightly lower.
-Same can be observed for heteroskedasticity patterns, where the gap between the power curve of the discrete uniform and the power curve of the normal distribution is even larger.
-This indicates that residual departures displayed with evenly distributed fitted values are visually more attractive.
-Besides, the power of detecting non-linearity or heteroskedasticity patterns from a lineup with lognormal fitted value distribution is the lowest among four distributions.
-This suggests the skewness of the fitted value distribution plays a role in the degree of presence of visual features, and a symmetric fitted value distribution can better reveal the underlying shape.
+Figure \ref{fig:different-x-dist-poly-power} illustrates the impact of fitted value distribution on the power of conventional tests and visual tests.
+For conventional tests, only the power curves of appropriate tests which are RESET tests and Breusch-Pagan tests for non-linearity pattern and heteroskedasticity pattern respectively are shown.
+For visual tests, note that we have collected more evaluations on lineup with uniform fitted value distribution, and the number of evaluations on a lineup will affect the power of the visual test.
+To have a fair comparison, for lineups with uniform fitted value distribution, we randomly sample five evaluations from the total eleven evaluations to estimate the power curves.
+The sampling process is repeated 500 times to produce an indication of the variation of the power.
+
+In terms of visual tests, we do not observe significant difference between the power curve of the uniform distribution and the power curve of the discrete uniform distribution.
+Normal and lognormal distributions produce lower power than uniform and discrete uniform distributions in regards of residual departure patterns, but the gap is larger when the residual plot contains heteroskedasticity pattern.
+This suggests that residual departures displayed with evenly distributed fitted values are visually more attractive and effective.
+Besides, the skewness of the fitted value distribution plays a role in the degree of presence of visual features as lognormal distribution produce the lowest power for both non-linearity and heteroskedasticity patterns.
+A symmetric fitted value distribution can better reveal the underlying shape.
+
+Although it is unusual to consider fitted value distribution or predictor distribution in power analysis, it is not so surprised that the power of conventional tests vary because of fitted value distribution.
+For instance, the power of the RESET test which is effectively a F test depends on the realizations of the additional predictors included in the auxiliary regression equation [@jamshidian2007study].
+And those additional predictors are linear transformation of random vectors following certain predictor distributions.
+What is truly unexpected is the huge variability in power compared to visual tests.
 
@@
   mutate(log_effect_size = log(effect_size)) %>%
-  mutate(type = stringr::str_to_title(type))
+  mutate(type = factor(stringr::str_to_title(type), 
+                       levels = c("Non-Linearity", "Heteroskedasticity")))
 ```
@@
 ```{r different-x-dist-poly-power, fig.height = 6, fig.cap = "Comparison of power on lineups with different fitted value distributions for conventional tests and visual tests.
+Power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and Breusch-Pagan tests respectively.
 Power curves of visual tests are estimated using five evaluations on each lineup.
@@
 The fitted value distribution has stronger impact on the power of conventional tests than visual tests.
-Meanwhile, uneven distributions including normal and lognormal distributions usually produce lower power.", dev = 'png', dpi = 450}
+Meanwhile, uneven distributions including normal and lognormal distributions genearlly produce lower power.", dev = 'png', dpi = 450}
 

### 2023-03-24 paper_comparison/paper_comparison/paper_comparison.Rmd
swap the facet such that the x scale is correct

@@
   labs(col = "", linetype = "Test") +
-  facet_grid(type~conv_or_not, scales = "free_x") +
+  facet_grid(conv_or_not~type, scales = "free_x") +
   theme(legend.position = "bottom")

### 2023-03-24 paper_comparison/paper_comparison/paper_comparison.Rmd
finish the x_dist text

@@
 And those additional predictors are linear transformation of random vectors following certain predictor distributions.
-What is truly unexpected is the huge variability in power compared to visual tests.
+What is truly unexpected is the huge variability in power of conventional tests compared to visual tests.
+The discrete uniform distribution constantly produce the highest power while the normal distribution and the lognormal distribution produce the lowest power for non-linearity and heteroskedasticity pattern respectively.
+The difference in power corresponding to different distributions peak at around 90% when the effect size is at a moderate level.
+This huge difference is undesirable as is not uncommon for analysts to omit the factor of predictor distribution in power calculation and sample size calculation.
+It also suggests that, under the designed scenarios in the paper, the visual test is a more robust test.
 
@@
 The sampling process has been repeated for 500 times and the corresponding power curves are drawn with transparent grey lines.
-The fitted value distribution has stronger impact on the power of conventional tests than visual tests.
+The fitted value distribution has greater impact on the power of conventional tests than visual tests.
 Meanwhile, uneven distributions including normal and lognormal distributions genearlly produce lower power.", dev = 'png', dpi = 450}
@@
   mutate(x_dist = ifelse(x_dist == "even_discrete", 
-                           "Discrete", 
+                         "Discrete uniform", 
                          stringr::str_to_title(x_dist))) %>%
@@
                                     "Lognormal", 
-                                    "Discrete"))) %>%
+                                    "Discrete uniform"))) %>%
 

### 2023-03-24 paper_comparison/paper_comparison/paper_comparison.tex
modify the x_dist text

--- 2023-03-14 paper_comparison/paper_comparison/paper_comparison.tex
@@
 
-\title{A plot is worth a thousand tests: assessing residual diagnostics
-with visual inference}
+\title{A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics
+with the Lineup Protocol}
 
@@
 
-Figure \ref{fig:different-x-dist-poly-power} illustrates the change of
-power of visual tests for different fitted distributions used in the
-non-linearity model and the heteroskedasticity model. Note that we have
-collected more evaluations on lineup with uniform fitted value
-distribution, and the number of evaluations on a lineup will have an
-impact on the power of the visual test. To have a fair comparison, the
-five evaluations used in the power estimation are randomly sampled from
-the total eleven evaluations for each lineup with uniform fitted value
-distribution. The sampling process has been repeated 500 times to give
-an indication of the variation of the power.
-
-We do not observe significant difference between the power curve of the
-uniform distribution and the power curve of the discrete uniform
-distribution, but the power curves of normal and lognormal distributions
-are slightly lower. Same can be observed for heteroskedasticity
-patterns, where the gap between the power curve of the discrete uniform
-and the power curve of the normal distribution is even larger. This
-indicates that residual departures displayed with evenly distributed
-fitted values are visually more attractive. Besides, the power of
-detecting non-linearity or heteroskedasticity patterns from a lineup
-with lognormal fitted value distribution is the lowest among four
-distributions. This suggests the skewness of the fitted value
-distribution plays a role in the degree of presence of visual features,
-and a symmetric fitted value distribution can better reveal the
-underlying shape.
+Figure \ref{fig:different-x-dist-poly-power} illustrates the impact of
+fitted value distribution on the power of conventional tests and visual
+tests. For conventional tests, only the power curves of appropriate
+tests which are RESET tests and Breusch-Pagan tests for non-linearity
+pattern and heteroskedasticity pattern respectively are shown. For
+visual tests, note that we have collected more evaluations on lineup
+with uniform fitted value distribution, and the number of evaluations on
+a lineup will affect the power of the visual test. To have a fair
+comparison, for lineups with uniform fitted value distribution, we
+randomly sample five evaluations from the total eleven evaluations to
+estimate the power curves. The sampling process is repeated 500 times to
+produce an indication of the variation of the power.
+
+In terms of visual tests, we do not observe significant difference
+between the power curve of the uniform distribution and the power curve
+of the discrete uniform distribution. Normal and lognormal distributions
+produce lower power than uniform and discrete uniform distributions in
+regards of residual departure patterns, but the gap is larger when the
+residual plot contains heteroskedasticity pattern. This suggests that
+residual departures displayed with evenly distributed fitted values are
+visually more attractive and effective. Besides, the skewness of the
+fitted value distribution plays a role in the degree of presence of
+visual features as lognormal distribution produce the lowest power for
+both non-linearity and heteroskedasticity patterns. A symmetric fitted
+value distribution can better reveal the underlying shape.
+
+Although it is unusual to consider fitted value distribution or
+predictor distribution in power analysis, it is not so surprised that
+the power of conventional tests vary because of fitted value
+distribution. For instance, the power of the RESET test which is
+effectively a F test depends on the realizations of the additional
+predictors included in the auxiliary regression equation
+\citep{jamshidian2007study}. And those additional predictors are linear
+transformation of random vectors following certain predictor
+distributions. What is truly unexpected is the huge variability in power
+of conventional tests compared to visual tests. The discrete uniform
+distribution constantly produce the highest power while the normal
+distribution and the lognormal distribution produce the lowest power for
+non-linearity and heteroskedasticity pattern respectively. The
+difference in power corresponding to different distributions peak at
+around 90\% when the effect size is at a moderate level. This huge
+difference is undesirable as is not uncommon for analysts to omit the
+factor of predictor distribution in power calculation and sample size
+calculation. It also suggests that, under the designed scenarios in this
+paper, the visual test is a more robust test against fitted value
+distributions than a conventional test.
 
@@
 
-\caption{Comparison of power on lineups with different fitted value distributions for visual tests.
-All the power curves are estimated using five evaluations on each lineup.
+\caption{Comparison of power on lineups with different fitted value distributions for conventional tests and visual tests.
+Power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and Breusch-Pagan tests respectively.
+Power curves of visual tests are estimated using five evaluations on each lineup.
 For lineups with uniform fitted value distribution, the five evaluations are randomly sampled from the total eleven evaluations.
 The sampling process has been repeated for 500 times and the corresponding power curves are drawn with transparent grey lines.
-Uneven distributions including normal and lognormal distributions produce lower power for both types of residual departures.
-Meanwhile, the lognormal distribution as an asymmetric distribution produce the lowest power.}\label{fig:different-x-dist-poly-power}
+The fitted value distribution has greater impact on the power of conventional tests than visual tests.
+Meanwhile, uneven distributions including normal and lognormal distributions genearlly produce lower power.}\label{fig:different-x-dist-poly-power}
 \end{figure}
@@
 
-\hypertarget{acknowledgement}{%
-\section{Acknowledgement}\label{acknowledgement}}
-
-This research is enabled by the use of the following \texttt{R}
-packages: cli, curl, dplyr, ggplot2, jsonlite, lmtest, mpoly, progress,
-tibble, ggmosaic, purrr, tidyr, readr, stringr, here, kableExtra,
-patchwork, rcartocolor. The study website is powered by pythonAnywhere
-and the Flask web framework. The jsPsych framework is used to create
-behavioral experiments that run in our study website.
-
-The article was created with rticles (ref here), knitr (ref here) and R
-Markdown (ref here). The project's Github repository (link here)
-contains all materials required to reproduce this article.
+\hypertarget{acknowledgements}{%
+\section*{Acknowledgements}\label{acknowledgements}}
+\addcontentsline{toc}{section}{Acknowledgements}
+
+These \texttt{R} packages are used for the work: \texttt{cli}
+\citep{cli}, \texttt{curl} \citep{curl}, \texttt{dplyr} \citep{dplyr},
+\texttt{ggplot2} \citep{ggplot2}, \texttt{jsonlite} \citep{jsonlite},
+\texttt{lmtest} \citep{lmtest}, \texttt{mpoly} \citep{mpoly},
+\texttt{progress} \citep{progress}, \texttt{tibble} \citep{tibble},
+\texttt{ggmosaic} \citep{ggmosaic}, \texttt{purrr} \citep{purrr},
+\texttt{tidyr} \citep{tidyr}, \texttt{readr} \citep{readr},
+\texttt{stringr} \citep{stringr}, \texttt{here} \citep{here},
+\texttt{kableExtra} \citep{kableextra}, \texttt{patchwork}
+\citep{patchwork}, \texttt{rcartocolor} \citep{rcartocolor}. The study
+website is powered by \texttt{PythonAnywhere} \citep{pythonanywhere} and
+the \texttt{Flask} web framework \citep{flask}. The \texttt{jsPsych}
+framework \citep{jspsych} is used to create behavioral experiments that
+run in our study website.
+
+The article was created with R packages \texttt{rticles}
+\citep{rticles}, \texttt{knitr} \citep{knitr} and \texttt{rmarkdown}
+\citep{rmarkdown}. The project's Github repository (link here) contains
+all materials required to reproduce this article.
 
 \hypertarget{supplementary-material}{%
-\section{Supplementary material}\label{supplementary-material}}
+\section*{Supplementary material}\label{supplementary-material}}
+\addcontentsline{toc}{section}{Supplementary material}
 
@@
 
-\newpage
-
 \bibliographystyle{tfcad}

### 2023-03-28 paper_comparison/paper_comparison/paper_comparison.Rmd
modify the x_dist text
update conclusion
remove some unused comments

@@
 
-(introduction refine - rank: low)
-
 > *"Since all models are wrong the scientist must be alert to what is importantly wrong."* [@box1976science]
@@
 They use lineup protocols to judge the assumption of linearity, normality and constant error variance for both the level-1 and level-2 residuals.
-(expand?)
 
@@
 
-(should the below p goes to conclusion as a discussion?)
-
 Overall, the power comparison suggests that conventional tests differs significantly from visual tests in two regression diagnostics scenarios designed by us.
@@
 To understand why this is, one needs to return to the way the RESET test is applied.
-It requires a parameter indicating degree of fitted values to test for, and the recommendation is to generically use four (@ramsey_tests_1969).
+It requires a parameter indicating degree of fitted values to test for, and the recommendation is to generically use four [@ramsey_tests_1969].
 However, the "triple-U" shape constructed from the Hermite polynomials use power up to 18.
@@
 This huge difference is undesirable as is not uncommon for analysts to omit the factor of predictor distribution in power calculation and sample size calculation.
-It also suggests that, under the designed scenarios in the paper, the visual test is a more robust test.
 
@@
 Also, complex structure are generally harder to detect, but there are outliers.
+Under the designed scenarios in this paper, we find the visual test to be a more robust test against the change of fitted value distributions.
 A surprising finding was that the direction of heteroskedasticity appears to affect the ability to visually detect it, with wedge to the right being less detectable.

### 2023-03-28 paper_comparison/paper_comparison/paper_comparison.Rmd
update intro

@@
 In general, one looks for noticeable departures from the model like non-linear dependency or heteroskedasticity.
-However, correctly judging a residual plot where no pattern exists can be a painstakingly difficult task for humans (?citation).
-It is especially common, particularly among new data analysts, to report patterns when an experienced one might quickly conclude that there are none (ET: not sure if this is true).
+However, correctly judging a residual plot where no pattern exists can be a painstakingly difficult task for humans.
+It is especially common, particularly among students, to misinterpret patterns that are "random noise" and "random deviation from a model" [@loy2021bringing].
 It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and use a Breusch-Pagan test [@breusch_simple_1979] for heteroskedasticity.

### 2023-03-28 paper_comparison/paper_comparison/paper_comparison.tex
update intro

@@
 
-(introduction refine - rank: low)
-
 \begin{quote}
@@
 heteroskedasticity. However, correctly judging a residual plot where no
-pattern exists can be a painstakingly difficult task for humans
-(?citation). It is especially common, particularly among new data
-analysts, to report patterns when an experienced one might quickly
-conclude that there are none (ET: not sure if this is true). It is also
-possible to conduct hypothesis tests for non-linear dependence
-\citep{ramsey_tests_1969}, and use a Breusch-Pagan test
-\citep{breusch_simple_1979} for heteroskedasticity.
+pattern exists can be a painstakingly difficult task for humans. It is
+especially common, particularly among students, to misinterpret patterns
+that are ``random noise'' and ``random deviation from a model''
+\citep{loy2021bringing}. It is also possible to conduct hypothesis tests
+for non-linear dependence \citep{ramsey_tests_1969}, and use a
+Breusch-Pagan test \citep{breusch_simple_1979} for heteroskedasticity.
 
@@
 to judge the assumption of linearity, normality and constant error
-variance for both the level-1 and level-2 residuals. (expand?)
+variance for both the level-1 and level-2 residuals.
 
@@
 
-(should the below p goes to conclusion as a discussion?)
-
 Overall, the power comparison suggests that conventional tests differs
@@
 degree of fitted values to test for, and the recommendation is to
-generically use four (\citet{ramsey_tests_1969}). However, the
+generically use four \citep{ramsey_tests_1969}. However, the
 ``triple-U'' shape constructed from the Hermite polynomials use power up
@@
 factor of predictor distribution in power calculation and sample size
-calculation. It also suggests that, under the designed scenarios in this
-paper, the visual test is a more robust test against fitted value
-distributions than a conventional test.
+calculation.
 
@@
 fitted values are lognormal. Also, complex structure are generally
-harder to detect, but there are outliers. A surprising finding was that
+harder to detect, but there are outliers. Under the designed scenarios
+in this paper, we find the visual test to be a more robust test against
+the change of fitted value distributions. A surprising finding was that
 the direction of heteroskedasticity appears to affect the ability to

### 2023-03-28 paper_comparison/appendix.Rmd
don't need the x_dist part in appendix

@@
                      alpha = alpha_lower) %>%
-  mutate(modified = modified) %>%
+  mutate(modified = 1 - modified) %>%
   mutate(adjusted_p_value = p_value)
@@
                     caption = "Change of rejection decision because of the modification of $\\hat{\\alpha}$.")
-```
-
-## Curious effect of predictor distribution on conventional test power
-
-```{r cache = TRUE}
-min_poly_es <- vi_survey %>% 
-  filter(!null_lineup, 
-         !attention_check,
-         type == "polynomial", 
-         x_dist == "uniform") %>%
-  pull(effect_size) %>%
-  min()
-
-max_poly_es <- vi_survey %>% 
-  filter(!null_lineup, 
-         !attention_check,
-         type == "polynomial", 
-         x_dist == "uniform") %>%
-  pull(effect_size) %>%
-  max()
-
-min_heter_es <- vi_survey %>% 
-  filter(!null_lineup, 
-         !attention_check,
-         type == "heteroskedasticity", 
-         x_dist == "uniform") %>%
-  pull(effect_size) %>%
-  min()
-
-max_heter_es <- vi_survey %>% 
-  filter(!null_lineup, 
-         !attention_check,
-         type == "heteroskedasticity", 
-         x_dist == "uniform") %>%
-  pull(effect_size) %>%
-  max()
-
-poly_x_dist <- poly_conv_sim %>%
-  select(-RESET3_p_value, -(RESET5_p_value:RESET10_p_value), -F_p_value) %>%
-  rename(RESET_p_value = RESET4_p_value) %>%
-  pivot_longer(RESET_p_value:SW_p_value) %>%
-  mutate(name = gsub("_p_value", " test", name)) %>%
-  filter(name == "RESET test") %>%
-  mutate(reject = value <= 0.05) %>%
-  select(effect_size, name, x_dist, reject) %>%
-  mutate(offset0 = log(0.05/0.95)) %>%
-  nest(dat = c(effect_size, offset0, reject)) %>%
-  mutate(mod = map(dat, 
-                   ~glm(reject ~ effect_size - 1, 
-                        family = binomial(), 
-                        data = .x,
-                        offset = offset0))) %>%
-  mutate(power = map(mod, function(mod) {
-    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
-                         offset0 = log(0.05/0.95))
-    result$power <- predict(mod, type = "response", newdata = result)
-    result
-  })) %>%
-  select(-dat, -mod) %>%
-  unnest(power) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  mutate(type = "non-linearity")
-
-heter_x_dist <- heter_conv_sim %>%
-  select(-RESET3_p_value, -(RESET5_p_value:RESET10_p_value), -F_p_value) %>%
-  rename(RESET_p_value = RESET4_p_value) %>%
-  pivot_longer(RESET_p_value:SW_p_value) %>%
-  mutate(name = gsub("_p_value", " test", name)) %>%
-  filter(name == "BP test") %>%
-  mutate(reject = value <= 0.05) %>%
-  select(effect_size, name, x_dist, reject) %>%
-  mutate(offset0 = log(0.05/0.95)) %>%
-  nest(dat = c(effect_size, offset0, reject)) %>%
-  mutate(mod = map(dat, 
-                   ~glm(reject ~ effect_size - 1, 
-                        family = binomial(), 
-                        data = .x,
-                        offset = offset0))) %>%
-  mutate(power = map(mod, function(mod) {
-    result <- data.frame(effect_size = seq(min_heter_es, max_heter_es, 0.1),
-                         offset0 = log(0.05/0.95))
-    result$power <- predict(mod, type = "response", newdata = result)
-    result
-  })) %>%
-  select(-dat, -mod) %>%
-  unnest(power) %>%
-  mutate(log_effect_size = log(effect_size)) %>%
-  mutate(type = "heteroskedasticity")
-```
-
-```{r conv-x-dist, fig.cap = "Change of power of conventional tests under different fitted value distribution for non-linearity model and heteroskedasticity model.", fig.height = 3}
-bind_rows(poly_x_dist, heter_x_dist) %>%
-  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
-  ggplot() +
-
-  # Draw conventional test power curve
-  geom_line(aes(log_effect_size, power, col = x_dist),
-            size = 1) +
-
-  # Theme
-  theme_light() +
-  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
-  xlab(expression(log[e] (Effect_size))) +
-  ylab("Power") +
-  labs(col = "", size = "# lineups") +
-  facet_wrap(~type, scales = "free_x")
 ```
@@
   filter(x_dist == "uniform") %>%
+  filter(shape == 4) %>%
+  select(-F_p_value, -BP_p_value, -SW_p_value) %>%
+  pivot_longer(RESET3_p_value:RESET10_p_value) %>%
+  mutate(name = gsub("RESET(.*)_p_value", "\\1", name)) %>%
+  mutate(reject = value <= 0.05) %>%
+  select(effect_size, name, reject) %>%
+  mutate(offset0 = log(0.05/0.95)) %>%
+  nest(dat = c(effect_size, offset0, reject)) %>%
+  mutate(mod = map(dat, 
+                   ~glm(reject ~ effect_size - 1, 
+                        family = binomial(), 
+                        data = .x,
+                        offset = offset0))) %>%
+  mutate(power = map(mod, function(mod) {
+    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
+                         offset0 = log(0.05/0.95))
+    result$power <- predict(mod, type = "response", newdata = result)
+    result
+  })) %>%
+  select(-dat, -mod) %>%
+  unnest(power) %>%
+  mutate(log_effect_size = log(effect_size))
+```
+
+```{r}
+reset_diff_dat %>%
+  ggplot() +
+  geom_line(aes(log_effect_size, power, col = as.factor(as.integer(name))),
+            size = 1) +
+  theme_light() +
+  scale_color_brewer(palette = "Reds")
+```
+
+```{r cache = TRUE}
+reset_diff_dat <- poly_conv_sim %>%
+  filter(x_dist == "uniform") %>%
+  filter(shape == 1) %>%
   select(-F_p_value, -BP_p_value, -SW_p_value) %>%

### 2023-03-28 paper_comparison/appendix.Rmd
provide detailed tech layers

@@
 The survey data is collected via a self-hosted website designed by us.
+The complete architecture is provided in Figure \ref{fig:tech}.
+The website is built with the `Flask` [@flask] web framework and hosted on `PythonAnywhere` [@pythonanywhere].
+It is configured to handle HTTP requests such that subjects can correctly receive webpages and submit responses.
+Embedded in the resources sent to subjects, the `jsPsych` front-end framework [@jspsych] instructs subjects' browsers to render an environment for running behavioral experiments.
+During the experiment, this framework will automatically collect common behavioral data such as response time and clicks on buttons.
+Subjects' responses are first validated by a scheduled `Python` script run on the server, then push to a Github repository.
+Lineup images shown to users are saved in multiple Github repositories and hosted in corresponding Github pages.
+The URLs to these images are resolved by `Flask` and bundled in HTML files.
 
-(desc of the tech layers)
-
-Once the participant is recruited from Prolific (ref here), it will be redirected to the entry page of our study website.
+Once the participant is recruited from Prolific [@palan2018prolific], it will be redirected to the entry page of our study website.
 An image of the entry page is provided in Figure \ref{fig:entry-page}.
@@
 ```{r}
+min_poly_es <- vi_survey %>% 
+  filter(!null_lineup, 
+         !attention_check,
+         type == "polynomial", 
+         x_dist == "uniform") %>%
+  pull(effect_size) %>%
+  min()
+max_poly_es <- vi_survey %>% 
+  filter(!null_lineup, 
+         !attention_check,
+         type == "polynomial", 
+         x_dist == "uniform") %>%
+  pull(effect_size) %>%
+  max()
+```
+
+```{r}
 eval_n <- function(n_eval, boot) {

### 2023-03-28 paper_comparison/appendix.Rmd
update table in appendix

@@
   rename(Pronoun = pronoun) %>%
-  kableExtra::kable(format = "latex", booktabs = TRUE, label = "pronoun", caption = "Summary of pronoun distribution of subjects recuritted in this study.")
+  add_row(Pronoun = "",
+          `Period I` = sum(.$`Period I`),
+          `Period II` = sum(.$`Period II`),
+          `Period III` = sum(.$`Period III`),
+          Total = sum(.$Total)) %>%
+  mutate(`Period I %` = round(`Period I`/443*100, 1)) %>%
+  mutate(`Period II %` = round(`Period II`/443*100, 1)) %>%
+  mutate(`Period III %` = round(`Period III`/443*100, 1)) %>%
+  mutate(`Total %` = round(Total/443*100, 1)) %>%
+  select(Pronoun, 
+         `Period I`, `Period I %`, 
+         `Period II`, `Period II %`,
+         `Period III`, `Period III %`,
+         Total, `Total %`) %>%
+  kableExtra::kable(format = "latex", 
+                    col.names = c("Pronoun", 
+                                  "Period I", "%", 
+                                  "Period II", "%",
+                                  "Period III", "%",
+                                  "Total", "%"), 
+                    booktabs = TRUE, label = "pronoun", caption = "Summary of pronoun distribution of subjects recuritted in this study.")
 ```
@@
   rename(`Education` = education) %>%
+  add_row(`Education` = "",
+          `Period I` = sum(.$`Period I`),
+          `Period II` = sum(.$`Period II`),
+          `Period III` = sum(.$`Period III`),
+          Total = sum(.$Total)) %>%
+  mutate(`Period I` = paste0(`Period I`, " (", round(`Period I`/443*100, 1), "%)")) %>%
+  mutate(`Period II` = paste0(`Period II`, " (", round(`Period II`/443*100, 1), "%)")) %>%
+  mutate(`Period III` = paste0(`Period III`, " (", round(`Period III`/443*100, 1), "%)")) %>%
+  mutate(Total = paste0(Total, " (", round(Total/443*100, 1), "%)")) %>%
   kableExtra::kable(format = "latex", booktabs = TRUE, label = "education", caption = "Summary of education distribution of subjects recuritted in this study.")

### 2023-03-28 paper_comparison/appendix.Rmd
modify col for lv plot

@@
   scale_alpha_discrete(range = c(0.7, 0)) +
-  # scale_fill_lv() +
   facet_wrap(~k, ncol = 5) +
@@
   labs(fill = "Quantiles") +
-  scale_fill_discrete(labels = c("50%", "25% - 75%", "12.5% - 87.5%", "6.25% - 93.75%")) +
+  scale_fill_manual(labels = c("50%", "25% - 75%", "12.5% - 87.5%", "6.25% - 93.75%"),
+                    values = c("white", rev(RColorBrewer::brewer.pal(5, "Blues")))) +
   ylab("Weighted propotion of detect") +

### 2023-03-28 paper_comparison/appendix.Rmd
update lv plot legend

--- 2023-03-28 paper_comparison/appendix.Rmd
@@
   labs(fill = "Quantiles") +
-  scale_fill_discrete(labels = c("50%", "25% - 75%", "12.5% - 87.5%", "6.25% - 93.75%")) +
+  scale_fill_manual(labels = c("0.5", "[0.25, 0.75]", "[0.125, 0.875]", "[0.0625, 0.09375]"),
+                    values = c("white", rev(RColorBrewer::brewer.pal(5, "Blues")))) +
   ylab("Weighted propotion of detect") +

### 2023-03-28 paper_comparison/appendix.Rmd
modify the alpha plot

@@
 Figure \ref{fig:poly-boxplot-lineup} and \ref{fig:heter-boxplot-lineup} provide two lineups to examine whether there is an actual difference across data collection periods for non-linearity model and heteroskedasticity model respectively.
-To emphasize the tail behaviour and display fewer outliers, we use the "letter-value" boxplot (ref here) which is an extension of the number of "letter value" statistics (ref here) to check the weighed proportion of detect over different data collection period.
+To emphasize the tail behaviour and display fewer outliers, we use the "letter-value" boxplot [@hofmann2017value] which is an extension of the number of "letter value" statistics to check the weighed proportion of detect over different data collection period.
 The weighted proportion of detect is calculated by taking the average of $c_i$ of a lineup over a data collection period.
@@
 
-```{r sensitive, fig.cap = "\\label{fig:sensitive}Change of $p$-values with $\\hat{\\alpha}$ multiplied by $0.125$, $0.25$, $0.5$, $1.125$, $1.25$ and $1.5$."}
+```{r sensitive, fig.cap = "\\label{fig:sensitive}Change of $p$-values with $\\hat{\\alpha}$ multiplied by $0.5$, $0.75$, $0.875$, $1.125$, $1.25$ and $1.5$."}
 upper_lower_p %>%
@@
   summarise(p_value = first(p_value))) %>%
+  mutate(diff_p_value = adjusted_p_value - p_value) %>%
   ggplot() +
-  geom_point(aes(p_value, adjusted_p_value), alpha = 0.2) +
-  geom_abline() +
+  geom_hline(yintercept = 0) +
+  geom_vline(xintercept = 0.05, col = "red", linetype = 2) +
+  geom_point(aes(p_value, diff_p_value), alpha = 0.2) +
+  scale_x_log10() +
   facet_wrap(~modified) +
-  scale_x_sqrt() +
-  scale_y_sqrt() +
   ylab("Adjusted p-value") +
   xlab("p-value") +
-  coord_fixed() +
   theme_light()

### 2023-03-28 paper_comparison/appendix.Rmd
modify caption for alpha

@@
 
-```{r sensitive, fig.cap = "\\label{fig:sensitive}Change of $p$-values with $\\hat{\\alpha}$ multiplied by $0.5$, $0.75$, $0.875$, $1.125$, $1.25$ and $1.5$."}
+```{r sensitive, fig.cap = "\\label{fig:sensitive}Change of $p$-values with $\\hat{\\alpha}$ multiplied by $0.5$, $0.75$, $0.875$, $1.125$, $1.25$ and $1.5$.
+The vertical dashed line is to indicate $p\\text{-value} = 0.05$.
+The x-axis is drawn on logarithmic scale.
+For multiplier smaller than $1$, the adjusted $p$-value will incrase then decrase as $p$-value increases.
+The trend is the opposite for multiplier greater than $1$, but the difference will eventually reach $0$."}
 upper_lower_p %>%
@@
   facet_wrap(~modified) +
-  ylab("Adjusted p-value") +
+  ylab("Difference between adjusted p-value and p-value") +
   xlab("p-value") +

### 2023-03-28 paper_comparison/appendix.Rmd
modify the last plot

@@
   filter(adjusted_reject != reject) %>%
-  mutate(case = ifelse(!adjusted_reject, "Propotion of lineups with $p$-value < 0.05 but adjusted $p$-value > 0.05", "Propotion of lineups with $p$-value > 0.05 but adjusted $p$-value < 0.05")) %>%
+  mutate(case = ifelse(!adjusted_reject, 'Proportion of lineups transforms to "not reject"', 'Proportion of lineups transforms to "reject"')) %>%
   rename(multiplier = modified) %>%
@@
             size = 1) +
+  labs(col = "Number of evalutaions") +
+  xlab(expression(log[e] (Effect_size))) +
   theme_light() +
@@
   unnest(power) %>%
-  mutate(log_effect_size = log(effect_size))
-```
-
-```{r}
-reset_diff_dat %>%
-  ggplot() +
-  geom_line(aes(log_effect_size, power, col = as.factor(as.integer(name))),
-            size = 1) +
-  theme_light() +
-  scale_color_brewer(palette = "Reds")
+  mutate(log_effect_size = log(effect_size)) %>%
+  mutate(shape = 4)
 ```
@@
 ```{r cache = TRUE}
-reset_diff_dat <- poly_conv_sim %>%
+reset_diff_dat2 <- poly_conv_sim %>%
   filter(x_dist == "uniform") %>%
@@
   unnest(power) %>%
-  mutate(log_effect_size = log(effect_size))
+  mutate(log_effect_size = log(effect_size)) %>%
+  mutate(shape = 1)
 ```
@@
 reset_diff_dat %>%
+  bind_rows(reset_diff_dat2) %>%
+  mutate(shape = ifelse(shape == 1, "U", "Triple-U")) %>%
   ggplot() +
@@
   theme_light() +
-  scale_color_brewer(palette = "Reds")
+  labs(col = "Highest order") +
+  xlab(expression(log[e] (Effect_size))) +
+  scale_color_brewer(palette = "Reds") +
+  facet_wrap(~shape)
 ```

### 2023-03-28 paper_comparison/appendix.Rmd
add caption for n_eval appendix

@@
 
-```{r}
+```{r n_eval, fig.cap = "Change of power for different number of evalutions on a lineup with uniform fitted value distribution.
+The power will incrase as the number of evaluations increases, but the margin will decrase."}
 eval_n <- function(n_eval, boot) {

### 2023-03-28 paper_comparison/appendix.Rmd
update n_eval text

@@
 
+When comparing power of visual tests across different fitted value distributions, we have discussed the number of evaluations on a lineup will affect the power of the visual test.
+Using the lineups with uniform fitted value distribution, we show in Figure \ref{n_eval} the change of power of visual tests due to different number of evaluations.
+It can be learned that as the number of evaluations increases, the power will increase but the margin will decrease.
+Considering we have eleven evaluations on lineups with uniform fitted value distribution, and five evaluations on other lineups, it is necessary to use the same number of evaluations for each lineup in comparison.
+
 ```{r}
@@
 
-```{r n_eval, fig.cap = "Change of power for different number of evalutions on a lineup with uniform fitted value distribution.
+```{r n_eval, fig.cap = "Change of power of visual tests for different number of evalutions on lineups with uniform fitted value distribution.
 The power will incrase as the number of evaluations increases, but the margin will decrase."}
@@
 
-```{r}
+```{r reset, fig.cap = 'Change of power of RESET tests for different orders of fitted values included in the auxiliary formula.
+The left panel is the power of testing the "U" shape and the right panel is the power of testing the "Triple-U" shape.
+The power will not be greatly affected by the highest order in the case of testing the "U" shape.
+In the case of testing the "Triple-U" shape, the highest order needs to be set to at least six to avoid the loss of power.', fig.height = 4}
 reset_diff_dat %>%
   bind_rows(reset_diff_dat2) %>%
-  mutate(shape = ifelse(shape == 1, "U", "Triple-U")) %>%
+  mutate(shape = factor(ifelse(shape == 1, "U", "Triple-U"), levels = c("U", "Triple-U"))) %>%
   ggplot() +

### 2023-03-28 paper_comparison/appendix.Rmd
fill the last part

@@
 
+It is found in the result that the power of a RESET test will be affected by the highest order of fitted values included in the auxiliary formula.
+And we suspect that the current recommendation of the highest order - four, is insufficient to test complex non-linear structures such as the "Triple-U" shape designed in this paper.
+Figure \ref{reset} illustrates the change of power of RESET test while testing the "U" shape and the "Triple-U" shape with different highest orders.
+Clearly, when testing a simple shape like the "U" shape, the highest order has very little impact on the power.
+But for testing the "Triple-U" shape, there will be a loss of power if the recommended order is used.
+To avoid the loss of power, the highest order needs to be set to at least six.
+
 ```{r cache = TRUE}

### 2023-03-28 Milestone_2/milestone_2_report.Rmd
meeting
fill in the role of visual diagnostics in modelling

@@
 ---
-title: "Milestone report"
+title: "Progress review report"
+subtitle: "Advances in Artificial Intelligence for Data Visualization: Automate Reading of Diagnostic Plots with Compute Vision Models"
 output: pdf_document
@@
 
-A short report (5 - 10) pages
+This is a temporary document. It perhaps needs to be transferred to a templated document.
 
-Advances in Artificial Intelligence for Data Visualization: Automate Reading of Data Plots
+keywords:
+AI, data visualization, residual plot, visual inference, hypothesis testing, computer vision
 
@@
 - the role of visual diagnostics in modelling
+
+Model diagnostics are important steps to validate model assumptions, examine goodness of fit and reveal uncaptured structure by the model.
+When diagnosing model, though numeric summaries are mostly available and some are even endorsed by finite or asymptotic properties, graphical representation of data is still preferred, or at least needed by data analysts, due to its intuitiveness and the possibility for unexpected discoveries which may be abstract and unquantifiable.
+For example, in regression diagnostics, it is common to plot residuals versus fitted values as the starting point of investigating the adequacy of the fit and checking the underlying assumptions.
+
 - a brief introduction to visual inference
+
 - limitation of lineup protocol
+
 2. research questions

### 2023-04-02 paper_comparison/paper_comparison/paper_comparison.Rmd
don't need the x_dist part in appendix
add multiple citations for F power affected by x dist

@@
 
-(fix the tense issue - rank: low)
-
-(fix the notation issue (bold?) - rank: low)
-
 # Introduction
@@
 However, correctly judging a residual plot where no pattern exists can be a painstakingly difficult task for humans.
-It is especially common, particularly among students, to misinterpret patterns that are "random noise" and "random deviation from a model" [@loy2021bringing].
+It is especially common, particularly among students, to misinterpret patterns that are random noise and random deviation from a model [@loy2021bringing].
 It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and use a Breusch-Pagan test [@breusch_simple_1979] for heteroskedasticity.
@@
 Although it is unusual to consider fitted value distribution or predictor distribution in power analysis, it is not so surprised that the power of conventional tests vary because of fitted value distribution.
-For instance, the power of the RESET test which is effectively a F test depends on the realizations of the additional predictors included in the auxiliary regression equation [@jamshidian2007study].
+For instance, the power of the RESET test which is effectively a F test depends on the realizations of the additional predictors included in the auxiliary regression equation [@jamshidian2007study; @olvera2019relationship; @zhang2018practical].
 And those additional predictors are linear transformation of random vectors following certain predictor distributions.

### 2023-04-04 Milestone_2/report/report.Rmd
complete background and aim

---
title: "Advances in Artificial Intelligence for Data Visualization: Automate Reading of Diagnostic Plots with Compute Vision Models"
author:
- familyname: Li
  othernames: Weihao (Patrick)
  address: Monash University
  qualifications: PhD student
email: weihao.li@monash.com
phone: (04) 0459 1219
department: Department of\newline Econometrics &\newline Business Statistics
organization: "Progress review" 
bibliography: report.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  monash::report:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, messages=FALSE, warning=FALSE)
library(ggplot2)
```

keywords:
AI, data visualization, residual plot, visual inference, hypothesis testing, computer vision

# Overview of the thesis

## Background and motivation

Model diagnostics play a critical role in evaluating the accuracy and validity of a statistical model. They enable the assessment of the model's assumptions about the data, identification of outliers or anomalous observations that may have an exaggerated effect on the model, evaluation of how well the model fits the data, and identification of possible approaches to improve the model's performance.

When conducting model diagnostics, despite the availability of numeric summaries with finite or asymptotic properties that have been endorsed, data analysts prefer or require graphical representations of data. The preference for visual diagnostics is attributed to its intuitive nature and the possibility of discovering unexpected abstract and unquantifiable insights. In the context of regression diagnostics, a common practice is to plot residuals against fitted values, which serves as a starting point for evaluating the adequacy of the fit and verifying the underlying assumptions. Other visualization techniques such as histograms, Q-Q plots and box plots can be used to identify potential issues with assumptions made about the data, such as linearity, normality, or homoscedasticity.

In recent times, a novel statistical inferential framework known as visual inference [@buja_statistical_2009] has been developed, which relies on the use of graphical representations of data. The visual inference approach capitalizes on the innate capability of the human visual system to rapidly identify patterns and deviations from expected patterns. By exploiting visualization techniques, visual inference provides a more natural and easily comprehensible way of interpreting data and discerning relationships between variables, compared to conventional statistical techniques.
This inferential framework can be applied to perform exploratory data analysis, model selection, and hypothesis testing.

In the same paper, @buja_statistical_2009 proposes the use of the lineup protocol as a visual hypothesis test, inspired by the police lineup technique employed in eyewitness identification of criminal suspects. The protocol comprises $m$ randomly positioned plots, where on of them represents the data plot, while the remaining $m - 1$ plots represent the null plots with the same graphical structure, except that the data in them is consistent with the null hypothesis $H_0$. Under $H_0$, the data plot is expected to be indistinguishable from the null plots, and the probability of correctly identifying the data plot by an observer is $1/m$. If the data plot is correctly identified, it provides evidence against $H_0$.

This method has gained increasing traction in recent years and has already been integrated into data analysis of various topics, such as diagnostics of hierarchical linear models [@loy2013diagnostic], geographical research [@widen_graphical_2016] and forensic examinations [@krishnan_hierarchical_2021].
With the advent of sophisticated visualization software and tools, visual inference has the potential to revolutionize statistical analysis by providing an innovative alternative to traditional statistical approaches and enabling more lucid and effective communication of research findings.

The incorporation of human judgment is a fundamental aspect of visual testing, but it may restrict its widespread usage. The reliance on human assessment, akin to handicraft production methods employed in pre-industrial societies, renders the lineup protocol unsuitable for large-scale applications, due to its high labor costs and time requirements. Moreover, it presents significant usability issues for individuals with visual impairments, resulting in reduced accessibility.

The advancement of technology is essential to alleviate the workload of individuals by automating repetitive tasks and providing standardized results in a controlled environment. The evaluation of visual tests on a large scale is not feasible without the aid of machines and technology. Modern computer vision models offer a promising solution to this challenge. As a subfield of AI, computer vision with its modern deep learning architectures has successfully resolved numerous critical problems in automation. The development of the convolutional neural network (CNN) by @fukushima_neocognitron_1982 was inspired by the vision processing in living organisms. This architecture was first applied to hand-written number recognition trained with back-propagation by @lecun_backpropagation_1989, making it one of the earliest attempts at successfully extracting information from digital images via self-learning algorithms. The modern computer vision model typically utilizes deep neural networks with convolutional layers [@fukushima_neocognitron_1982], which leverage the hierarchical pattern in data and provide regularized versions of fully-connected layers. This approach downscales and transforms images by summarizing information in a small space. Numerous studies have shown that it can effectively tackle vision tasks, such as image recognition [@rawat_deep_2017]. The development of graphics processing units and the widespread availability of high-performance personal computers have made computer vision research a new trend in the 21st century. The remarkable achievements, such as computer-aided diagnosis [@lee_image_2015], pedestrian detection [@brunetti_computer_2018], and facial recognition [@emami_facial_2012], have significantly impacted our daily life.

The utilization of computer vision models for reading data plots is not widely adopted. Nevertheless, certain fields have embraced this notion, as evidenced by the application of computer vision models in reading recurrence plots for time series regression [@ojeda_multivariate_2020], time series classification [@chu_automatic_2019; @hailesilassie_financial_2019; @hatami_classification_2018; @zhang_encoding_2020], anomaly detection [@chen_convolutional_2020], and pairwise causality analysis [@singh_deep_2017]. However, the assessment of lineups with computer vision models represents a relatively novel area of investigation.

## Research questions

The main objective of this research is to construct an automatic visual inference system to facilitate conducting regression diagnostics visual tests on a large scale. The study will concentrate on three specific projects, namely

1. Exploring the application of visual inference in regression diagnostics and comparing its efficacy with conventional hypothesis tests.
2. Designing an automated visual inference system to assess lineups of residual plots derived from the classical normal linear regression model.
3. Deploying the automatic visual inference system as an online application and and making the pertinent open-source software accessible.

## Current reserach outcomes

3. summarise the work in the first year
- an experiment to compare visual test with conventional test 
- a prototype model for lineup evaluation
- points out the limitation of the work
4. summarise the work in the second year
- refine and reconduct the experiment
5. main findings to date
- findings from the first year
- findings from the second year

# Thesis structure

topic

chapter 1: Introduction

chapter 2
the first paper

chapter 3
build a computer vision system to evaluate residual plot

chapter 4?
software publication

chapter 5
discussion and future work

# Timetable

- April: submit abstract to ASC
- May: submit paper
- June: submit poster/short paper to IEEE vis conf, start working on computer vision model
- July: leave for few weeks
- Aug
- Sep: finalize the computer vision model
- Oct: IEEE vis conf
- Nov: web interface development 
- Dec: ASC

- Jan, 2024: 
- Feb, 2024: 
- Mar, 2024: submit paper
- Aug, 2024: submit thesis

# Difficulties

things that are actually quite serious

1. identified difficulties
2. suggestions for overcoming these difficulties

# References

### 2023-04-04 Milestone_2/report/report.tex
complete background and aim

\documentclass[11pt,a4paper,]{article}
\usepackage{lmodern}

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={Advances in Artificial Intelligence for Data Visualization: Automate Reading of Diagnostic Plots with Compute Vision Models},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{geometry}
\geometry{a4paper, centering, text={16cm,25cm}}
\usepackage[style=authoryear-comp,]{biblatex}
\addbibresource{report.bib}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\title{Advances in Artificial Intelligence for Data Visualization: Automate Reading of Diagnostic Plots with Compute Vision Models}

%% MONASH STUFF

%% CAPTIONS
\RequirePackage{caption}
\DeclareCaptionStyle{italic}[justification=centering]
 {labelfont={bf},textfont={it},labelsep=colon}
\captionsetup[figure]{style=italic,format=hang,singlelinecheck=true}
\captionsetup[table]{style=italic,format=hang,singlelinecheck=true}

%% FONT
\RequirePackage{bera}
\RequirePackage[charter,expert,sfscaled]{mathdesign}
\RequirePackage{fontawesome}

%% HEADERS AND FOOTERS
\RequirePackage{fancyhdr}
\pagestyle{fancy}
\rfoot{\Large\sffamily\raisebox{-0.1cm}{\textbf{\thepage}}}
\makeatletter
\lhead{\textsf{\expandafter{\@title}}}
\makeatother
\rhead{}
\cfoot{}
\setlength{\headheight}{15pt}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}
\fancypagestyle{plain}{%
\fancyhf{} % clear all header and footer fields
\fancyfoot[C]{\sffamily\thepage} % except the center
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}}

%% MATHS
\RequirePackage{bm,amsmath}
\allowdisplaybreaks

%% GRAPHICS
\RequirePackage{graphicx}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.8}

%\RequirePackage[section]{placeins}

%% SECTION TITLES

%% SECTION TITLES
\RequirePackage[compact,sf,bf]{titlesec}
\titleformat*{\section}{\Large\sf\bfseries\color[rgb]{0.7,0,0}}
\titleformat*{\subsection}{\large\sf\bfseries\color[rgb]{0.7,0,0}}
\titleformat*{\subsubsection}{\sf\bfseries\color[rgb]{0.7,0,0}}
\titlespacing{\section}{0pt}{2ex}{.5ex}
\titlespacing{\subsection}{0pt}{1.5ex}{0ex}
\titlespacing{\subsubsection}{0pt}{.5ex}{0ex}

%% TITLE PAGE
\def\Date{\number\day}
\def\Month{\ifcase\month\or
 January\or February\or March\or April\or May\or June\or
 July\or August\or September\or October\or November\or December\fi}
\def\Year{\number\year}

%% LINE AND PAGE BREAKING
\sloppy
\clubpenalty = 10000
\widowpenalty = 10000
\brokenpenalty = 10000
\RequirePackage{microtype}

%% PARAGRAPH BREAKS
\setlength{\parskip}{1.4ex}
\setlength{\parindent}{0em}

%% HYPERLINKS
\RequirePackage{xcolor} % Needed for links
\definecolor{darkblue}{rgb}{0,0,.6}
\RequirePackage{url}

\makeatletter
\@ifpackageloaded{hyperref}{}{\RequirePackage{hyperref}}
\makeatother
\hypersetup{
     citecolor=0 0 0,
     breaklinks=true,
     bookmarksopen=true,
     bookmarksnumbered=true,
     linkcolor=darkblue,
     urlcolor=blue,
     citecolor=darkblue,
     colorlinks=true}

\usepackage[showonlyrefs]{mathtools}
\usepackage[no-weekday]{eukdate}

%% BIBLIOGRAPHY

\makeatletter
\@ifpackageloaded{biblatex}{}{\usepackage[style=authoryear-comp, backend=biber, natbib=true]{biblatex}}
\makeatother
\ExecuteBibliographyOptions{bibencoding=utf8,minnames=1,maxnames=3, maxbibnames=99,dashed=false,terseinits=true,giveninits=true,uniquename=false,uniquelist=false,doi=false, isbn=false,url=true,sortcites=false}

\DeclareFieldFormat{url}{\texttt{\url{#1}}}
\DeclareFieldFormat[article]{pages}{#1}
\DeclareFieldFormat[inproceedings]{pages}{\lowercase{pp.}#1}
\DeclareFieldFormat[incollection]{pages}{\lowercase{pp.}#1}
\DeclareFieldFormat[article]{volume}{\mkbibbold{#1}}
\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}
\DeclareFieldFormat[article]{title}{\MakeCapital{#1}}
\DeclareFieldFormat[article]{url}{}
%\DeclareFieldFormat[book]{url}{}
%\DeclareFieldFormat[inbook]{url}{}
%\DeclareFieldFormat[incollection]{url}{}
%\DeclareFieldFormat[inproceedings]{url}{}
\DeclareFieldFormat[inproceedings]{title}{#1}
\DeclareFieldFormat{shorthandwidth}{#1}
%\DeclareFieldFormat{extrayear}{}
% No dot before number of articles
\usepackage{xpatch}
\xpatchbibmacro{volume+number+eid}{\setunit*{\adddot}}{}{}{}
% Remove In: for an article.
\renewbibmacro{in:}{%
  \ifentrytype{article}{}{%
  \printtext{\bibstring{in}\intitlepunct}}}

\AtEveryBibitem{\clearfield{month}}
\AtEveryCitekey{\clearfield{month}}

\makeatletter
\DeclareDelimFormat[cbx@textcite]{nameyeardelim}{\addspace}
\makeatother

\author{\sf{\Large\textbf{Weihao (Patrick) Li}\\\large PhD student\\[0.5cm]}}

\date{\sf\Date~\Month~\Year}
\makeatletter
\lfoot{\sf Li: \@date}
\makeatother

%%%% PAGE STYLE FOR FRONT PAGE OF REPORTS

\makeatletter
\def\organization#1{\gdef\@organization{#1}}
\def\telephone#1{\gdef\@telephone{#1}}
\def\email#1{\gdef\@email{#1}}
\makeatother
  \organization{Progress review}

  \def\name{Department of\newline Econometrics \&\newline Business Statistics}

  \telephone{(04) 0459 1219}

  \email{\href{mailto:weihao.li@monash.com}{\nolinkurl{weihao.li@monash.com}}}

\def\webaddress{\url{http://buseco.monash.edu/ebs/consulting/}}
\def\abn{12 377 614 012}
\def\extraspace{\vspace*{1.6cm}}
\makeatletter
\def\contactdetails{\faicon{phone} & \@telephone \\
                    \faicon{envelope} & \@email}
\makeatother

\usepackage[absolute,overlay]{textpos}
\setlength{\TPHorizModule}{1cm}
\setlength{\TPVertModule}{1cm}

%%%% FRONT PAGE OF REPORTS

\def\reporttype{Report for}

\long\def\front#1#2#3{
\newpage
\begin{textblock}{7}(12.7,28.2)\hfill
\includegraphics[height=0.6cm]{AACSB}~~~
\includegraphics[height=0.6cm]{EQUIS}~~~
\includegraphics[height=0.6cm]{AMBA}
\end{textblock}
\begin{singlespacing}
\thispagestyle{empty}
\vspace*{-1.4cm}
\hspace*{-1.4cm}
\hbox to 16cm{
  \hbox to 6.5cm{\vbox to 14cm{\vbox to 25cm{
    \includegraphics[width=6cm]{monash2}
    \vfill
    \includegraphics[width=3.5cm]{MBSportrait}
    \vspace{0.4cm}
    \par
    \parbox{6.3cm}{\raggedright
      \sf\color[rgb]{0.00,0.00,0.70}
      {\large\textbf{\name}}\par
      \vspace{.7cm}
      \tabcolsep=0.12cm\sf\small
      \begin{tabular}{@{}ll@{}}\contactdetails
      \end{tabular}
      \vspace*{0.3cm}\par
      ABN: \abn\par
    }
  }\vss}\hss}
  \hspace*{0.2cm}
  \hbox to 1cm{\vbox to 14cm{\rule{1pt}{26.8cm}\vss}\hss\hfill}
  \hbox to 10cm{\vbox to 14cm{\vbox to 25cm{
      \vspace*{3cm}\sf\raggedright
      \parbox{11cm}{\sf\raggedright\baselineskip=1.2cm
         \fontsize{24.88}{30}\color[rgb]{0.70,0.00,0.00}\sf\textbf{#1}}
      \par
      \vfill
      \large
      \vbox{\parskip=0.8cm #2}\par
      \vspace*{2cm}\par
      \reporttype\\[0.3cm]
      \hbox{#3}%\\[2cm]\
      \vspace*{1cm}
      {\large\sf\textbf{\Date~\Month~\Year}}
   }\vss}
  }}
\end{singlespacing}
\newpage
}

\makeatletter
\def\titlepage{\front{\expandafter{\@title}}{\@author}{\@organization}}
\makeatother

\usepackage{setspace}
\setstretch{1.5}

%% Any special functions or other packages can be loaded here.

\begin{document}
\titlepage

keywords:
AI, data visualization, residual plot, visual inference, hypothesis testing, computer vision

\hypertarget{overview-of-the-thesis}{%
\section{Overview of the thesis}\label{overview-of-the-thesis}}

\hypertarget{background-and-motivation}{%
\subsection{Background and motivation}\label{background-and-motivation}}

Model diagnostics play a critical role in evaluating the accuracy and validity of a statistical model. They enable the assessment of the model's assumptions about the data, identification of outliers or anomalous observations that may have an exaggerated effect on the model, evaluation of how well the model fits the data, and identification of possible approaches to improve the model's performance.

When conducting model diagnostics, despite the availability of numeric summaries with finite or asymptotic properties that have been endorsed, data analysts prefer or require graphical representations of data. The preference for visual diagnostics is attributed to its intuitive nature and the possibility of discovering unexpected abstract and unquantifiable insights. In the context of regression diagnostics, a common practice is to plot residuals against fitted values, which serves as a starting point for evaluating the adequacy of the fit and verifying the underlying assumptions. Other visualization techniques such as histograms, Q-Q plots and box plots can be used to identify potential issues with assumptions made about the data, such as linearity, normality, or homoscedasticity.

In recent times, a novel statistical inferential framework known as visual inference \autocite{buja_statistical_2009} has been developed, which relies on the use of graphical representations of data. The visual inference approach capitalizes on the innate capability of the human visual system to rapidly identify patterns and deviations from expected patterns. By exploiting visualization techniques, visual inference provides a more natural and easily comprehensible way of interpreting data and discerning relationships between variables, compared to conventional statistical techniques.
This inferential framework can be applied to perform exploratory data analysis, model selection, and hypothesis testing.

In the same paper, \textcite{buja_statistical_2009} proposes the use of the lineup protocol as a visual hypothesis test, inspired by the police lineup technique employed in eyewitness identification of criminal suspects. The protocol comprises \(m\) randomly positioned plots, where on of them represents the data plot, while the remaining \(m - 1\) plots represent the null plots with the same graphical structure, except that the data in them is consistent with the null hypothesis \(H_0\). Under \(H_0\), the data plot is expected to be indistinguishable from the null plots, and the probability of correctly identifying the data plot by an observer is \(1/m\). If the data plot is correctly identified, it provides evidence against \(H_0\).

This method has gained increasing traction in recent years and has already been integrated into data analysis of various topics, such as diagnostics of hierarchical linear models \autocite{loy2013diagnostic}, geographical research \autocite{widen_graphical_2016} and forensic examinations \autocite{krishnan_hierarchical_2021}.
With the advent of sophisticated visualization software and tools, visual inference has the potential to revolutionize statistical analysis by providing an innovative alternative to traditional statistical approaches and enabling more lucid and effective communication of research findings.

The incorporation of human judgment is a fundamental aspect of visual testing, but it may restrict its widespread usage. The reliance on human assessment, akin to handicraft production methods employed in pre-industrial societies, renders the lineup protocol unsuitable for large-scale applications, due to its high labor costs and time requirements. Moreover, it presents significant usability issues for individuals with visual impairments, resulting in reduced accessibility.

The advancement of technology is essential to alleviate the workload of individuals by automating repetitive tasks and providing standardized results in a controlled environment. The evaluation of visual tests on a large scale is not feasible without the aid of machines and technology. Modern computer vision models offer a promising solution to this challenge. As a subfield of AI, computer vision with its modern deep learning architectures has successfully resolved numerous critical problems in automation. The development of the convolutional neural network (CNN) by \textcite{fukushima_neocognitron_1982} was inspired by the vision processing in living organisms. This architecture was first applied to hand-written number recognition trained with back-propagation by \textcite{lecun_backpropagation_1989}, making it one of the earliest attempts at successfully extracting information from digital images via self-learning algorithms. The modern computer vision model typically utilizes deep neural networks with convolutional layers \autocite{fukushima_neocognitron_1982}, which leverage the hierarchical pattern in data and provide regularized versions of fully-connected layers. This approach downscales and transforms images by summarizing information in a small space. Numerous studies have shown that it can effectively tackle vision tasks, such as image recognition \autocite{rawat_deep_2017}. The development of graphics processing units and the widespread availability of high-performance personal computers have made computer vision research a new trend in the 21st century. The remarkable achievements, such as computer-aided diagnosis \autocite{lee_image_2015}, pedestrian detection \autocite{brunetti_computer_2018}, and facial recognition \autocite{emami_facial_2012}, have significantly impacted our daily life.

The utilization of computer vision models for reading data plots is not widely adopted. Nevertheless, certain fields have embraced this notion, as evidenced by the application of computer vision models in reading recurrence plots for time series regression \autocite{ojeda_multivariate_2020}, time series classification \autocite{chu_automatic_2019,hailesilassie_financial_2019,hatami_classification_2018,zhang_encoding_2020}, anomaly detection \autocite{chen_convolutional_2020}, and pairwise causality analysis \autocite{singh_deep_2017}. However, the assessment of lineups with computer vision models represents a relatively novel area of investigation.

\hypertarget{research-questions}{%
\subsection{Research questions}\label{research-questions}}

The main objective of this research is to construct an automatic visual inference system to facilitate conducting regression diagnostics visual tests on a large scale. The study will concentrate on three specific projects, namely

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Exploring the application of visual inference in regression diagnostics and comparing its efficacy with conventional hypothesis tests.
\item
  Designing an automated visual inference system to assess lineups of residual plots derived from the classical normal linear regression model.
\item
  Deploying the automatic visual inference system as an online application and and making the pertinent open-source software accessible.
\end{enumerate}

\hypertarget{current-reserach-outcomes}{%
\subsection{Current reserach outcomes}\label{current-reserach-outcomes}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  summarise the work in the first year
\end{enumerate}

\begin{itemize}
\tightlist
\item
  an experiment to compare visual test with conventional test
\item
  a prototype model for lineup evaluation
\item
  points out the limitation of the work
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  summarise the work in the second year
\end{enumerate}

\begin{itemize}
\tightlist
\item
  refine and reconduct the experiment
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{4}
\tightlist
\item
  main findings to date
\end{enumerate}

\begin{itemize}
\tightlist
\item
  findings from the first year
\item
  findings from the second year
\end{itemize}

\hypertarget{thesis-structure}{%
\section{Thesis structure}\label{thesis-structure}}

topic

chapter 1: Introduction

chapter 2
the first paper

chapter 3
build a computer vision system to evaluate residual plot

chapter 4?
software publication

chapter 5
discussion and future work

\hypertarget{timetable}{%
\section{Timetable}\label{timetable}}

\begin{itemize}
\item
  April: submit abstract to ASC
\item
  May: submit paper
\item
  June: submit poster/short paper to IEEE vis conf, start working on computer vision model
\item
  July: leave for few weeks
\item
  Aug
\item
  Sep: finalize the computer vision model
\item
  Oct: IEEE vis conf
\item
  Nov: web interface development
\item
  Dec: ASC
\item
  Jan, 2024:
\item
  Feb, 2024:
\item
  Mar, 2024: submit paper
\item
  Aug, 2024: submit thesis
\end{itemize}

\hypertarget{difficulties}{%
\section{Difficulties}\label{difficulties}}

things that are actually quite serious

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  identified difficulties
\item
  suggestions for overcoming these difficulties
\end{enumerate}

\printbibliography[title=References]

\end{document}

### 2023-04-05 Milestone_2/report/report.Rmd
update current research outcomes

@@
 
+In the initial phase of this research, a pilot investigation was undertaken to explore the viability of implementing visual inference in regression diagnostics.
+The pilot study involved 64 participants and focused on utilizing a lineup consisting of 20 plots, wherein a residual plot was embedded along with 19 null plots, simulated using the residual rotation technique.
+The study considered two primary forms of residual departures, namely non-linearity and heteroskedasticity, within a multiple linear regression context.
+To enrich the visual features, different fitted value distributions, including normal, uniform, lognormal, and discrete uniform, were also incorporated.
+
+The outcomes of this pilot study demonstrated that the power of visual tests in regression diagnostics increases proportionately with the number of lineup evaluations.
+Nevertheless, it was observed that conventional tests, such as the F-test for non-linearity and the Breusch-Pagan test for heteroskedasticity, exhibited significantly greater power than the visual tests.
+While the pilot study successfully provided a comparative analysis of the power of visual and conventional tests, the impact of various factors on the power of visual tests could not be fully analyzed due to inadequate data.
+
+We acknowledge the limitations of this pilot study, particularly in the design of the simulated model, which made it challenging to explicate the involved factors.
+Furthermore, the methodology utilized for effect size calculation, p-value calculation, and power estimation was incipient.
+Hence, a follow-up study is necessary to address these limitations and obtain a more comprehensive understanding of the potential of visual inference in regression diagnostics.
+
+In the subsequent year, a similar study was conducted, utilizing simpler simulated models to elicit more distinct visual patterns. Specifically, the non-linearity model was formulated as
+
+\begin{align} \label{eq:nonlinearity-model}
+\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
+\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
+\boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\
+\boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)),
+\end{align}
+
+\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vectors of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
+
+\begin{equation} \label{eq:scaling-function}
+g(\boldsymbol{x}, k) = \frac{\boldsymbol{x} - min(\boldsymbol{x})}{max(\boldsymbol{x}) - min(\boldsymbol{x})}(2k - k), \quad \text{for} \quad k > 0. 
+\end{equation}
+
+The heteroskedasticity model was formulated as
+
+\begin{align} \label{eq:heter-model}
+\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
+\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
+\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}, 1 + (2 - |a|)(\boldsymbol{x} - a)^2b \boldsymbol{I}). 
+\end{align}
+
+The null regression model used to fit the realizations generated by the above models is formulated as
+
+\begin{equation} \label{eq:null-model}
+\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
+\end{equation}
+
+\noindent where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.
+
+Since $z = O(x^j)$, for $j > 1$, $z$ is a higher order term leaves out by the null regression, which will violate the linearity model assumption.
+Similarly, for $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the predictor $\boldsymbol{x}$, which will lead to the presence of heteroskedasticity.
+
+- draft below
+
+The effect size was derived using the Kullback-Leibler divergence method, formulated as Equation 3.
+To account for the interdependence of lineups presented to different participants, p-value calculation, as suggested by VanderPlas et al.
+(2021), was adopted.
+Furthermore, power estimation was simplified through a logistic regression model that lacked an intercept but included a fixed offset to ensure that the power equated to 5% when the effect size was zero.
+A detailed description of the experimental design, effect size derivation, p-value calculation, and power estimation can be found in the attached paper.
+
 3. summarise the work in the first year

### 2023-04-05 Milestone_2/report/report.tex
update current research outcomes

@@
 
+In the initial phase of this research, a pilot investigation was undertaken to explore the viability of implementing visual inference in regression diagnostics.
+The pilot study involved 64 participants and focused on utilizing a lineup consisting of 20 plots, wherein a residual plot was embedded along with 19 null plots, simulated using the residual rotation technique.
+The study considered two primary forms of residual departures, namely non-linearity and heteroskedasticity, within a multiple linear regression context.
+To enrich the visual features, different fitted value distributions, including normal, uniform, lognormal, and discrete uniform, were also incorporated.
+
+The outcomes of this pilot study demonstrated that the power of visual tests in regression diagnostics increases proportionately with the number of lineup evaluations.
+Nevertheless, it was observed that conventional tests, such as the F-test for non-linearity and the Breusch-Pagan test for heteroskedasticity, exhibited significantly greater power than the visual tests.
+While the pilot study successfully provided a comparative analysis of the power of visual and conventional tests, the impact of various factors on the power of visual tests could not be fully analyzed due to inadequate data.
+
+We acknowledge the limitations of this pilot study, particularly in the design of the simulated model, which made it challenging to explicate the involved factors.
+Furthermore, the methodology utilized for effect size calculation, p-value calculation, and power estimation was incipient.
+Hence, a follow-up study is necessary to address these limitations and obtain a more comprehensive understanding of the potential of visual inference in regression diagnostics.
+
+In the subsequent year, a similar study was conducted, utilizing simpler simulated models to elicit more distinct visual patterns. Specifically, the non-linearity model was formulated as
+
+\begin{align} \label{eq:nonlinearity-model}
+\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
+\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
+\boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\
+\boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)),
+\end{align}
+
+\noindent where \(\boldsymbol{y}\), \(\boldsymbol{x}\), \(\boldsymbol{\varepsilon}\), \(\boldsymbol{x}_{raw}\), \(\boldsymbol{z}_{raw}\) are vectors of size \(n\), \(He_{j}(.)\) is the \(j\)th-order probabilist's Hermite polynomials, \(\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\), and \(g(\boldsymbol{x}, k)\) is a scaling function to enforce the support of the random vector to be \([-k, k]^n\) defined as
+
+\begin{equation} \label{eq:scaling-function}
+g(\boldsymbol{x}, k) = \frac{\boldsymbol{x} - min(\boldsymbol{x})}{max(\boldsymbol{x}) - min(\boldsymbol{x})}(2k - k), \quad \text{for} \quad k > 0. 
+\end{equation}
+
+The heteroskedasticity model was formulated as
+
+\begin{align} \label{eq:heter-model}
+\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
+\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
+\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}, 1 + (2 - |a|)(\boldsymbol{x} - a)^2b \boldsymbol{I}). 
+\end{align}
+
+The null regression model used to fit the realizations generated by the above models is formulated as
+
+\begin{equation} \label{eq:null-model}
+\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
+\end{equation}
+
+\noindent where \(\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\).
+
+Since \(z = O(x^j)\), for \(j > 1\), \(z\) is a higher order term leaves out by the null regression, which will violate the linearity model assumption.
+Similarly, for \(b \neq 0\), the variance-covariance matrix of the error term \(\boldsymbol{\varepsilon}\) is correlated with the predictor \(\boldsymbol{x}\), which will lead to the presence of heteroskedasticity.
+
+\begin{itemize}
+\tightlist
+\item
+  draft below
+\end{itemize}
+
+The effect size was derived using the Kullback-Leibler divergence method, formulated as Equation 3.
+To account for the interdependence of lineups presented to different participants, p-value calculation, as suggested by VanderPlas et al.~(2021), was adopted.
+Furthermore, power estimation was simplified through a logistic regression model that lacked an intercept but included a fixed offset to ensure that the power equated to 5\% when the effect size was zero.
+A detailed description of the experimental design, effect size derivation, p-value calculation, and power estimation can be found in the attached paper.
+
 \begin{enumerate}

### 2023-04-05 Milestone_2/report/report.Rmd
tuen lang

@@
 Model diagnostics play a critical role in evaluating the accuracy and validity of a statistical model.
-They enable the assessment of the model's assumptions about the data, identification of outliers or anomalous observations that may have an exaggerated effect on the model, evaluation of how well the model fits the data, and identification of possible approaches to improve the model's performance.
+They enable the assessment of the model's assumptions about the data, detection of outliers, evaluation of how well the model fits the data, and identification of possible approaches to improve the model's performance.
 
-When conducting model diagnostics, despite the availability of numeric summaries with finite or asymptotic properties that have been endorsed, data analysts prefer or require graphical representations of data.
+When conducting model diagnostics, despite the availability of numeric summaries endorsed by finite or asymptotic properties, data analysts prefer or require graphical representations of data.
 The preference for visual diagnostics is attributed to its intuitive nature and the possibility of discovering unexpected abstract and unquantifiable insights.
@@
 
-In recent times, a novel statistical inferential framework known as visual inference [@buja_statistical_2009] has been developed, which relies on the use of graphical representations of data.
-The visual inference approach capitalizes on the innate capability of the human visual system to rapidly identify patterns and deviations from expected patterns.
-By exploiting visualization techniques, visual inference provides a more natural and easily comprehensible way of interpreting data and discerning relationships between variables, compared to conventional statistical techniques.
-This inferential framework can be applied to perform exploratory data analysis, model selection, and hypothesis testing.
+Recently, a novel statistical inferential framework known as visual inference [@buja_statistical_2009] has been developed, which relies on the use of graphical representations of data.
+The visual inference approach makes use of the natural capability of the human visual system to identify patterns and deviations from expected patterns.
+It provides a more comprehensible way of interpreting data and conducting hypothesis testing compared to conventional statistical testing.
 
-In the same paper, @buja_statistical_2009 proposes the use of the lineup protocol as a visual hypothesis test, inspired by the police lineup technique employed in eyewitness identification of criminal suspects.
-The protocol comprises $m$ randomly positioned plots, where on of them represents the data plot, while the remaining $m - 1$ plots represent the null plots with the same graphical structure, except that the data in them is consistent with the null hypothesis $H_0$.
+Practically, visual inference is conducted via the lineup protocol.
+The protocol is inspired by the police lineup technique employed in eyewitness identification of criminal suspects.
+It comprises $m$ randomly positioned plots, where on of them represents the data plot, while the remaining $m - 1$ plots represent the null plots with the same graphical structure, except that the data has been replaced with data consistent with the null hypothesis $H_0$.
 Under $H_0$, the data plot is expected to be indistinguishable from the null plots, and the probability of correctly identifying the data plot by an observer is $1/m$.
@@
 This method has gained increasing traction in recent years and has already been integrated into data analysis of various topics, such as diagnostics of hierarchical linear models [@loy2013diagnostic], geographical research [@widen_graphical_2016] and forensic examinations [@krishnan_hierarchical_2021].
-With the advent of sophisticated visualization software and tools, visual inference has the potential to revolutionize statistical analysis by providing an innovative alternative to traditional statistical approaches and enabling more lucid and effective communication of research findings.
+With the advent of sophisticated visualization techniques and tools, visual inference has the potential to provide an innovative alternative to traditional statistical approaches and enabling more effective communication of model findings.
 
-The incorporation of human judgment is a fundamental aspect of visual testing, but it may restrict its widespread usage.
-The reliance on human assessment, akin to handicraft production methods employed in pre-industrial societies, renders the lineup protocol unsuitable for large-scale applications, due to its high labor costs and time requirements.
+The reliance of human assessment is a fundamental aspect of visual tests, but it may restrict its widespread usage.
+The lineup protocol is unsuitable for large-scale applications, due to its high labor costs and time requirements.
 Moreover, it presents significant usability issues for individuals with visual impairments, resulting in reduced accessibility.
 
-The advancement of technology is essential to alleviate the workload of individuals by automating repetitive tasks and providing standardized results in a controlled environment.
+The workload of individuals could be alleviated by automating repetitive tasks and providing standardized results in a controlled environment.
 The evaluation of visual tests on a large scale is not feasible without the aid of machines and technology.
+
 Modern computer vision models offer a promising solution to this challenge.
@@
 Numerous studies have shown that it can effectively tackle vision tasks, such as image recognition [@rawat_deep_2017].
-The development of graphics processing units and the widespread availability of high-performance personal computers have made computer vision research a new trend in the 21st century.
-The remarkable achievements, such as computer-aided diagnosis [@lee_image_2015], pedestrian detection [@brunetti_computer_2018], and facial recognition [@emami_facial_2012], have significantly impacted our daily life.
+With the development of graphics processing units and the widespread availability of high-performance personal computers, computer vision research become a new hype in the 21st century.
+Remarkable achievements, such as computer-aided diagnosis [@lee_image_2015], pedestrian detection [@brunetti_computer_2018], and facial recognition [@emami_facial_2012], have significantly impacted our daily life.
 
-The utilization of computer vision models for reading data plots is not widely adopted.
-Nevertheless, certain fields have embraced this notion, as evidenced by the application of computer vision models in reading recurrence plots for time series regression [@ojeda_multivariate_2020], time series classification [@chu_automatic_2019; @hailesilassie_financial_2019; @hatami_classification_2018; @zhang_encoding_2020], anomaly detection [@chen_convolutional_2020], and pairwise causality analysis [@singh_deep_2017].
-However, the assessment of lineups with computer vision models represents a relatively novel area of investigation.
+Utilizing computer vision models in reading data plot is not a common choice.
+Nevertheless, certain fields have adopted this idea by applying computer vision models in reading recurrence plots for time series regression [@ojeda_multivariate_2020], time series classification [@chu_automatic_2019; @hailesilassie_financial_2019; @hatami_classification_2018; @zhang_encoding_2020], anomaly detection [@chen_convolutional_2020], and pairwise causality analysis [@singh_deep_2017].
+However, the assessment of lineups with computer vision models is a relatively novel area of study.
 
@@
 
-The main objective of this research is to construct an automatic visual inference system to facilitate conducting regression diagnostics visual tests on a large scale.
+The main objective of this research is to construct an automatic visual inference system to enable the conduction of regression diagnostics visual tests on a large scale.
 The study will concentrate on three specific projects, namely
 
-1. Exploring the application of visual inference in regression diagnostics and comparing its efficacy with conventional hypothesis tests.
-2. Designing an automated visual inference system to assess lineups of residual plots derived from the classical normal linear regression model.
-3. Deploying the automatic visual inference system as an online application and and making the pertinent open-source software accessible.
+1. Exploring the application of visual inference in regression diagnostics and comparing it with conventional hypothesis tests.
+2. Designing an automated visual inference system to assess lineups of residual plots of classical normal linear regression model.
+3. Deploying the automatic visual inference system as an online application and publishing the relevant open-source software.
 
-## Current reserach outcomes
+## Current research outcomes
+
+- cut this p down, we don't need that many details
 

### 2023-04-05 Milestone_2/report/report.tex
tuen lang

@@
 
-In recent times, a novel statistical inferential framework known as visual inference \autocite{buja_statistical_2009} has been developed, which relies on the use of graphical representations of data.
-The visual inference approach capitalizes on the innate capability of the human visual system to rapidly identify patterns and deviations from expected patterns.
-By exploiting visualization techniques, visual inference provides a more natural and easily comprehensible way of interpreting data and discerning relationships between variables, compared to conventional statistical techniques.
-This inferential framework can be applied to perform exploratory data analysis, model selection, and hypothesis testing.
+Recently, a novel statistical inferential framework known as visual inference \autocite{buja_statistical_2009} has been developed, which relies on the use of graphical representations of data.
+The visual inference approach makes use of the natural capability of the human visual system to identify patterns and deviations from expected patterns.
+It provides a more comprehensible way of interpreting data and conducting hypothesis testing compared to conventional statistical testing.
 
-In the same paper, \textcite{buja_statistical_2009} proposes the use of the lineup protocol as a visual hypothesis test, inspired by the police lineup technique employed in eyewitness identification of criminal suspects.
-The protocol comprises \(m\) randomly positioned plots, where on of them represents the data plot, while the remaining \(m - 1\) plots represent the null plots with the same graphical structure, except that the data in them is consistent with the null hypothesis \(H_0\).
+Practically, visual inference is conducted via the lineup protocol.
+The protocol is inspired by the police lineup technique employed in eyewitness identification of criminal suspects.
+It comprises \(m\) randomly positioned plots, where on of them represents the data plot, while the remaining \(m - 1\) plots represent the null plots with the same graphical structure, except that the data has been replaced with data consistent with the null hypothesis \(H_0\).
 Under \(H_0\), the data plot is expected to be indistinguishable from the null plots, and the probability of correctly identifying the data plot by an observer is \(1/m\).
@@
 This method has gained increasing traction in recent years and has already been integrated into data analysis of various topics, such as diagnostics of hierarchical linear models \autocite{loy2013diagnostic}, geographical research \autocite{widen_graphical_2016} and forensic examinations \autocite{krishnan_hierarchical_2021}.
-With the advent of sophisticated visualization software and tools, visual inference has the potential to revolutionize statistical analysis by providing an innovative alternative to traditional statistical approaches and enabling more lucid and effective communication of research findings.
+With the advent of sophisticated visualization techniques and tools, visual inference has the potential to provide an innovative alternative to traditional statistical approaches and enabling more effective communication of model findings.
 
-The incorporation of human judgment is a fundamental aspect of visual testing, but it may restrict its widespread usage.
-The reliance on human assessment, akin to handicraft production methods employed in pre-industrial societies, renders the lineup protocol unsuitable for large-scale applications, due to its high labor costs and time requirements.
+The reliance of human assessment is a fundamental aspect of visual tests, but it may restrict its widespread usage.
+The lineup protocol is unsuitable for large-scale applications, due to its high labor costs and time requirements.
 Moreover, it presents significant usability issues for individuals with visual impairments, resulting in reduced accessibility.
 
-The advancement of technology is essential to alleviate the workload of individuals by automating repetitive tasks and providing standardized results in a controlled environment.
+The workload of individuals could be alleviated by automating repetitive tasks and providing standardized results in a controlled environment.
 The evaluation of visual tests on a large scale is not feasible without the aid of machines and technology.
+
 Modern computer vision models offer a promising solution to this challenge.
@@
 Numerous studies have shown that it can effectively tackle vision tasks, such as image recognition \autocite{rawat_deep_2017}.
-The development of graphics processing units and the widespread availability of high-performance personal computers have made computer vision research a new trend in the 21st century.
-The remarkable achievements, such as computer-aided diagnosis \autocite{lee_image_2015}, pedestrian detection \autocite{brunetti_computer_2018}, and facial recognition \autocite{emami_facial_2012}, have significantly impacted our daily life.
+With the development of graphics processing units and the widespread availability of high-performance personal computers, computer vision research become a new hype in the 21st century.
+Remarkable achievements, such as computer-aided diagnosis \autocite{lee_image_2015}, pedestrian detection \autocite{brunetti_computer_2018}, and facial recognition \autocite{emami_facial_2012}, have significantly impacted our daily life.
 
-The utilization of computer vision models for reading data plots is not widely adopted.
-Nevertheless, certain fields have embraced this notion, as evidenced by the application of computer vision models in reading recurrence plots for time series regression \autocite{ojeda_multivariate_2020}, time series classification \autocite{chu_automatic_2019,hailesilassie_financial_2019,hatami_classification_2018,zhang_encoding_2020}, anomaly detection \autocite{chen_convolutional_2020}, and pairwise causality analysis \autocite{singh_deep_2017}.
-However, the assessment of lineups with computer vision models represents a relatively novel area of investigation.
+Utilizing computer vision models in reading data plot is not a common choice.
+Nevertheless, certain fields have adopted this idea by applying computer vision models in reading recurrence plots for time series regression \autocite{ojeda_multivariate_2020}, time series classification \autocite{chu_automatic_2019,hailesilassie_financial_2019,hatami_classification_2018,zhang_encoding_2020}, anomaly detection \autocite{chen_convolutional_2020}, and pairwise causality analysis \autocite{singh_deep_2017}.
+However, the assessment of lineups with computer vision models is a relatively novel area of study.
 
@@
 
-The main objective of this research is to construct an automatic visual inference system to facilitate conducting regression diagnostics visual tests on a large scale.
+The main objective of this research is to construct an automatic visual inference system to enable the conduction of regression diagnostics visual tests on a large scale.
 The study will concentrate on three specific projects, namely
@@
 \item
-  Exploring the application of visual inference in regression diagnostics and comparing its efficacy with conventional hypothesis tests.
+  Exploring the application of visual inference in regression diagnostics and comparing it with conventional hypothesis tests.
 \item
-  Designing an automated visual inference system to assess lineups of residual plots derived from the classical normal linear regression model.
+  Designing an automated visual inference system to assess lineups of residual plots of classical normal linear regression model.
 \item
-  Deploying the automatic visual inference system as an online application and and making the pertinent open-source software accessible.
+  Deploying the automatic visual inference system as an online application and publishing the relevant open-source software.
 \end{enumerate}
 
-\hypertarget{current-reserach-outcomes}{%
-\subsection{Current reserach outcomes}\label{current-reserach-outcomes}}
+\hypertarget{current-research-outcomes}{%
+\subsection{Current research outcomes}\label{current-research-outcomes}}
+
+\begin{itemize}
+\tightlist
+\item
+  cut this p down, we don't need that many details
+\end{itemize}
 

### 2023-04-11 Milestone_2/report/report.Rmd
modify the desc for VI

@@
 knitr::opts_chunk$set(echo = FALSE, cache=TRUE, messages=FALSE, warning=FALSE)
-library(ggplot2)
+library(visage)
+library(tidyverse)
 ```
@@
 Model diagnostics play a critical role in evaluating the accuracy and validity of a statistical model.
-They enable the assessment of the model's assumptions about the data, detection of outliers, evaluation of how well the model fits the data, and identification of possible approaches to improve the model's performance.
+They enable the assessment of the model's assumptions, detection of outliers, evaluation of how well the model fits the data, and identification of possible approaches to improve the model's performance.
 
-When conducting model diagnostics, despite the availability of numeric summaries endorsed by finite or asymptotic properties, data analysts prefer or require graphical representations of data.
+When conducting model diagnostics, despite the availability of numeric summaries endorsed by finite or asymptotic properties, graphical representations of data are often preferred or required by data analysts.
 The preference for visual diagnostics is attributed to its intuitive nature and the possibility of discovering unexpected abstract and unquantifiable insights.
@@
 The protocol is inspired by the police lineup technique employed in eyewitness identification of criminal suspects.
-It comprises $m$ randomly positioned plots, where on of them represents the data plot, while the remaining $m - 1$ plots represent the null plots with the same graphical structure, except that the data has been replaced with data consistent with the null hypothesis $H_0$.
+It comprises $m$ randomly positioned plots, where one of them represents the data plot, while the remaining $m - 1$ plots represent the null plots with the same graphical structure, except that the data has been replaced with data consistent with the null hypothesis $H_0$.
+An example lineup is provided in Figure \ref{fig:lineup-example}.
+To compute the $p$-value of the visual test, the lineup will be independently presented to a number of participants, asking them to pick the most different plot.
 Under $H_0$, the data plot is expected to be indistinguishable from the null plots, and the probability of correctly identifying the data plot by an observer is $1/m$.
-If the data plot is correctly identified, it provides evidence against $H_0$.
+If a large number of participants correctly identify the data plot, the corresponding $p$-value will be small, indicating strong evidence against $H_0$.
 

### 2023-04-11 Milestone_2/report/report.Rmd
provide ex for lineup

@@
 
+```{r lineup-example, fig.pos="t!", fig.height = 7, fig.width = 7, fig.cap = "Visual testing is conducted using a lineup, as in the example here.
+The residual plot computed from the observed data (plot $2^2 + 2$, exhibiting non-linearity) is embedded among 19 null plots, where the residuals are simulated from a standard error model.
+Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot.
+A small $p$-value would result from a substantial number selecting plot $2^2 + 2$."}
 This method has gained increasing traction in recent years and has already been integrated into data analysis of various topics, such as diagnostics of hierarchical linear models [@loy2013diagnostic], geographical research [@widen_graphical_2016] and forensic examinations [@krishnan_hierarchical_2021].

### 2023-04-11 Milestone_2/report/report.Rmd
finish the code chunk

--- 2023-04-11 Milestone_2/report/report.Rmd
@@
 If a large number of participants correctly identify the data plot, the corresponding $p$-value will be small, indicating strong evidence against $H_0$.
+
+```{r lineup-example, fig.pos="t!", fig.height = 7, fig.width = 7, fig.cap = "Visual testing is conducted using a lineup, as in the example here.
+The residual plot computed from the observed data (plot $2^2 + 2$, exhibiting non-linearity) is embedded among 19 null plots, where the residuals are simulated from a standard error model.
+Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot.
+A small $p$-value would result from a substantial number selecting plot $2^2 + 2$."}
+vi_lineup <- get_vi_lineup()
+vi_lineup$poly_300$data %>%
+  VI_MODEL$plot_lineup(theme = theme_light(), remove_axis = TRUE, remove_grid_line = TRUE)
+```
 

### 2023-04-11 Milestone_2/report/report.Rmd
modify the desc for CV
completely rewrite current outcome

@@
 
-The workload of individuals could be alleviated by automating repetitive tasks and providing standardized results in a controlled environment.
-The evaluation of visual tests on a large scale is not feasible without the aid of machines and technology.
-
 Modern computer vision models offer a promising solution to this challenge.
@@
 
-- cut this p down, we don't need that many details
+In order to examine the potential applicability of integrating visual inference techniques into regression diagnostics, a pilot study was conducted in the first year with the participation of 64 individuals, which was followed by a formal study involving 443 participants in the subsequent year.
+The participants were presented with lineups consisting of 20 plots, where a residual plot was embedded along with 19 null plots, drawn with data simulated using the residual rotation technique.
+The study considered two primary forms of residual departures in multiple linear regression, namely non-linearity and heteroskedasticity.
+To enrich the visual features, different fitted value distributions, including normal, uniform, lognormal, and discrete, were also incorporated.
 
-In the initial phase of this research, a pilot investigation was undertaken to explore the viability of implementing visual inference in regression diagnostics.
-The pilot study involved 64 participants and focused on utilizing a lineup consisting of 20 plots, wherein a residual plot was embedded along with 19 null plots, simulated using the residual rotation technique.
-The study considered two primary forms of residual departures, namely non-linearity and heteroskedasticity, within a multiple linear regression context.
-To enrich the visual features, different fitted value distributions, including normal, uniform, lognormal, and discrete uniform, were also incorporated.
-
-The outcomes of this pilot study demonstrated that the power of visual tests in regression diagnostics increases proportionately with the number of lineup evaluations.
-Nevertheless, it was observed that conventional tests, such as the F-test for non-linearity and the Breusch-Pagan test for heteroskedasticity, exhibited significantly greater power than the visual tests.
-While the pilot study successfully provided a comparative analysis of the power of visual and conventional tests, the impact of various factors on the power of visual tests could not be fully analyzed due to inadequate data.
-
-We acknowledge the limitations of this pilot study, particularly in the design of the simulated model, which made it challenging to explicate the involved factors.
-Furthermore, the methodology utilized for effect size calculation, p-value calculation, and power estimation was incipient.
-Hence, a follow-up study is necessary to address these limitations and obtain a more comprehensive understanding of the potential of visual inference in regression diagnostics.
-
-In the subsequent year, a similar study was conducted, utilizing simpler simulated models to elicit more distinct visual patterns. Specifically, the non-linearity model was formulated as
-
-\begin{align} \label{eq:nonlinearity-model}
-\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
-\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
-\boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\
-\boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)),
-\end{align}
-
-\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vectors of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
-
-\begin{equation} \label{eq:scaling-function}
-g(\boldsymbol{x}, k) = \frac{\boldsymbol{x} - min(\boldsymbol{x})}{max(\boldsymbol{x}) - min(\boldsymbol{x})}(2k - k), \quad \text{for} \quad k > 0. 
-\end{equation}
-
-The heteroskedasticity model was formulated as
-
-\begin{align} \label{eq:heter-model}
-\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
-\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
-\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}, 1 + (2 - |a|)(\boldsymbol{x} - a)^2b \boldsymbol{I}). 
-\end{align}
-
-The null regression model used to fit the realizations generated by the above models is formulated as
-
-\begin{equation} \label{eq:null-model}
-\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
-\end{equation}
-
-\noindent where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.
-
-Since $z = O(x^j)$, for $j > 1$, $z$ is a higher order term leaves out by the null regression, which will violate the linearity model assumption.
-Similarly, for $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the predictor $\boldsymbol{x}$, which will lead to the presence of heteroskedasticity.
-
-- draft below
-
-The effect size was derived using the Kullback-Leibler divergence method, formulated as Equation 3.
-To account for the interdependence of lineups presented to different participants, p-value calculation, as suggested by VanderPlas et al.
-(2021), was adopted.
-Furthermore, power estimation was simplified through a logistic regression model that lacked an intercept but included a fixed offset to ensure that the power equated to 5% when the effect size was zero.
-A detailed description of the experimental design, effect size derivation, p-value calculation, and power estimation can be found in the attached paper.
-
-3. summarise the work in the first year
-- an experiment to compare visual test with conventional test 
-- a prototype model for lineup evaluation
-- points out the limitation of the work
-4. summarise the work in the second year
-- refine and reconduct the experiment
-5. main findings to date
-- findings from the first year
-- findings from the second year
+The study revealed that conventional residual-based statistical tests are more sensitive to weak residual departures from model assumptions compared to visual tests evaluated by humans, leading to excessive rejections even when downstream analysis and outcomes would not be significantly impacted by the small departures from a good fit.
+Specifically, conventional tests conclude that the model fit has issues nearly twice as frequently as humans would, and they frequently reject departures in the form of non-linearity and heteroskedasticity that are not visible to humans.
+These findings emphasize the crucial role of graphical diagnostics and support the integration of visual inference into regression diagnostics.
+Moreover, they provide a compelling rationale for the development of an automated visual inference system to evaluate lineups of residual plots.
+The comprehensive details of the research methodology, along with other intriguing findings can be found in the attached paper.
 

### 2023-04-11 Milestone_2/report/report.Rmd
compiled

@@
 The development of the convolutional neural network (CNN) by @fukushima_neocognitron_1982 was inspired by the vision processing in living organisms.
-This architecture was first applied to hand-written number recognition trained with back-propagation by @lecun_backpropagation_1989, making it one of the earliest attempts at successfully extracting information from digital images via self-learning algorithms.
-The modern computer vision model typically utilizes deep neural networks with convolutional layers [@fukushima_neocognitron_1982], which leverage the hierarchical pattern in data and provide regularized versions of fully-connected layers.
+The modern computer vision model typically utilizes deep neural networks with convolutional layers, which leverage the hierarchical pattern in data and provide regularized versions of fully-connected layers.
 This approach downscales and transforms images by summarizing information in a small space.
-Numerous studies have shown that it can effectively tackle vision tasks, such as image recognition [@rawat_deep_2017].
-With the development of graphics processing units and the widespread availability of high-performance personal computers, computer vision research become a new hype in the 21st century.
-Remarkable achievements, such as computer-aided diagnosis [@lee_image_2015], pedestrian detection [@brunetti_computer_2018], and facial recognition [@emami_facial_2012], have significantly impacted our daily life.
+Numerous studies have shown that it can effectively tackle vision tasks, such as image recognition [@rawat_deep_2017], computer-aided diagnosis [@lee_image_2015], pedestrian detection [@brunetti_computer_2018], and facial recognition [@emami_facial_2012].
 
@@
 
-The main objective of this research is to construct an automatic visual inference system to enable the conduction of regression diagnostics visual tests on a large scale.
+The main objective of this research is to construct an automatic visual inference system capable of conducting visual tests on a large scale in the domain of model diagnostics, particularly regression diagnostics.
 The study will concentrate on three specific projects, namely
@@
 # Thesis structure
+
+The thesis is structured
 

### 2023-04-11 Milestone_2/report/report.tex
compiled

@@
 \IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
 \IfFileExists{parskip.sty}{%
@@
 The protocol is inspired by the police lineup technique employed in eyewitness identification of criminal suspects.
-It comprises \(m\) randomly positioned plots, where on of them represents the data plot, while the remaining \(m - 1\) plots represent the null plots with the same graphical structure, except that the data has been replaced with data consistent with the null hypothesis \(H_0\).
+It comprises \(m\) randomly positioned plots, where one of them represents the data plot, while the remaining \(m - 1\) plots represent the null plots with the same graphical structure, except that the data has been replaced with data consistent with the null hypothesis \(H_0\).
+An example lineup is provided in Figure \ref{fig:lineup-example}.
+To compute the \(p\)-value of the visual test, the lineup will be independently presented to a number of participants, asking them to pick the most different plot.
 Under \(H_0\), the data plot is expected to be indistinguishable from the null plots, and the probability of correctly identifying the data plot by an observer is \(1/m\).
-If the data plot is correctly identified, it provides evidence against \(H_0\).
+If a large number of participants correctly identify the data plot, the corresponding \(p\)-value will be small, indicating strong evidence against \(H_0\).
+
+\begin{figure}
+\centering
+\includegraphics{report_files/figure-latex/lineup-example-1.pdf}
+\caption{\label{fig:lineup-example}Visual testing is conducted using a lineup, as in the example here.
+The residual plot computed from the observed data (plot \(2^2 + 2\), exhibiting non-linearity) is embedded among 19 null plots, where the residuals are simulated from a standard error model.
+Computing the \(p\)-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot.
+A small \(p\)-value would result from a substantial number selecting plot \(2^2 + 2\).}
+\end{figure}
 
@@
 The development of the convolutional neural network (CNN) by \textcite{fukushima_neocognitron_1982} was inspired by the vision processing in living organisms.
-This architecture was first applied to hand-written number recognition trained with back-propagation by \textcite{lecun_backpropagation_1989}, making it one of the earliest attempts at successfully extracting information from digital images via self-learning algorithms.
-The modern computer vision model typically utilizes deep neural networks with convolutional layers \autocite{fukushima_neocognitron_1982}, which leverage the hierarchical pattern in data and provide regularized versions of fully-connected layers.
+The modern computer vision model typically utilizes deep neural networks with convolutional layers, which leverage the hierarchical pattern in data and provide regularized versions of fully-connected layers.
 This approach downscales and transforms images by summarizing information in a small space.
-Numerous studies have shown that it can effectively tackle vision tasks, such as image recognition \autocite{rawat_deep_2017}.
-With the development of graphics processing units and the widespread availability of high-performance personal computers, computer vision research become a new hype in the 21st century.
-Remarkable achievements, such as computer-aided diagnosis \autocite{lee_image_2015}, pedestrian detection \autocite{brunetti_computer_2018}, and facial recognition \autocite{emami_facial_2012}, have significantly impacted our daily life.
+Numerous studies have shown that it can effectively tackle vision tasks, such as image recognition \autocite{rawat_deep_2017}, computer-aided diagnosis \autocite{lee_image_2015}, pedestrian detection \autocite{brunetti_computer_2018}, and facial recognition \autocite{emami_facial_2012}.
 
@@
 
-\begin{itemize}
-\tightlist
-\item
-  cut this p down, we don't need that many details
-\end{itemize}
+In order to examine the potential applicability of integrating visual inference techniques into regression diagnostics, a pilot study was conducted in the first year with the participation of 64 individuals, which was followed by a formal study involving 443 participants in the subsequent year.
+The participants were presented with lineups consisting of 20 plots, where a residual plot was embedded along with 19 null plots, drawn with data simulated using the residual rotation technique.
+The study considered two primary forms of residual departures in multiple linear regression, namely non-linearity and heteroskedasticity.
+To enrich the visual features, different fitted value distributions, including normal, uniform, lognormal, and discrete, were also incorporated.
 
-In the initial phase of this research, a pilot investigation was undertaken to explore the viability of implementing visual inference in regression diagnostics.
-The pilot study involved 64 participants and focused on utilizing a lineup consisting of 20 plots, wherein a residual plot was embedded along with 19 null plots, simulated using the residual rotation technique.
-The study considered two primary forms of residual departures, namely non-linearity and heteroskedasticity, within a multiple linear regression context.
-To enrich the visual features, different fitted value distributions, including normal, uniform, lognormal, and discrete uniform, were also incorporated.
-
-The outcomes of this pilot study demonstrated that the power of visual tests in regression diagnostics increases proportionately with the number of lineup evaluations.
-Nevertheless, it was observed that conventional tests, such as the F-test for non-linearity and the Breusch-Pagan test for heteroskedasticity, exhibited significantly greater power than the visual tests.
-While the pilot study successfully provided a comparative analysis of the power of visual and conventional tests, the impact of various factors on the power of visual tests could not be fully analyzed due to inadequate data.
-
-We acknowledge the limitations of this pilot study, particularly in the design of the simulated model, which made it challenging to explicate the involved factors.
-Furthermore, the methodology utilized for effect size calculation, p-value calculation, and power estimation was incipient.
-Hence, a follow-up study is necessary to address these limitations and obtain a more comprehensive understanding of the potential of visual inference in regression diagnostics.
-
-In the subsequent year, a similar study was conducted, utilizing simpler simulated models to elicit more distinct visual patterns. Specifically, the non-linearity model was formulated as
-
-\begin{align} \label{eq:nonlinearity-model}
-\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
-\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
-\boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\
-\boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)),
-\end{align}
-
-\noindent where \(\boldsymbol{y}\), \(\boldsymbol{x}\), \(\boldsymbol{\varepsilon}\), \(\boldsymbol{x}_{raw}\), \(\boldsymbol{z}_{raw}\) are vectors of size \(n\), \(He_{j}(.)\) is the \(j\)th-order probabilist's Hermite polynomials, \(\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\), and \(g(\boldsymbol{x}, k)\) is a scaling function to enforce the support of the random vector to be \([-k, k]^n\) defined as
-
-\begin{equation} \label{eq:scaling-function}
-g(\boldsymbol{x}, k) = \frac{\boldsymbol{x} - min(\boldsymbol{x})}{max(\boldsymbol{x}) - min(\boldsymbol{x})}(2k - k), \quad \text{for} \quad k > 0. 
-\end{equation}
-
-The heteroskedasticity model was formulated as
-
-\begin{align} \label{eq:heter-model}
-\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
-\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
-\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}, 1 + (2 - |a|)(\boldsymbol{x} - a)^2b \boldsymbol{I}). 
-\end{align}
-
-The null regression model used to fit the realizations generated by the above models is formulated as
-
-\begin{equation} \label{eq:null-model}
-\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
-\end{equation}
-
-\noindent where \(\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\).
-
-Since \(z = O(x^j)\), for \(j > 1\), \(z\) is a higher order term leaves out by the null regression, which will violate the linearity model assumption.
-Similarly, for \(b \neq 0\), the variance-covariance matrix of the error term \(\boldsymbol{\varepsilon}\) is correlated with the predictor \(\boldsymbol{x}\), which will lead to the presence of heteroskedasticity.
-
-\begin{itemize}
-\tightlist
-\item
-  draft below
-\end{itemize}
-
-The effect size was derived using the Kullback-Leibler divergence method, formulated as Equation 3.
-To account for the interdependence of lineups presented to different participants, p-value calculation, as suggested by VanderPlas et al.~(2021), was adopted.
-Furthermore, power estimation was simplified through a logistic regression model that lacked an intercept but included a fixed offset to ensure that the power equated to 5\% when the effect size was zero.
-A detailed description of the experimental design, effect size derivation, p-value calculation, and power estimation can be found in the attached paper.
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{2}
-\tightlist
-\item
-  summarise the work in the first year
-\end{enumerate}
-
-\begin{itemize}
-\tightlist
-\item
-  an experiment to compare visual test with conventional test
-\item
-  a prototype model for lineup evaluation
-\item
-  points out the limitation of the work
-\end{itemize}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{3}
-\tightlist
-\item
-  summarise the work in the second year
-\end{enumerate}
-
-\begin{itemize}
-\tightlist
-\item
-  refine and reconduct the experiment
-\end{itemize}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{4}
-\tightlist
-\item
-  main findings to date
-\end{enumerate}
-
-\begin{itemize}
-\tightlist
-\item
-  findings from the first year
-\item
-  findings from the second year
-\end{itemize}
+The study revealed that conventional residual-based statistical tests are more sensitive to weak residual departures from model assumptions compared to visual tests evaluated by humans, leading to excessive rejections even when downstream analysis and outcomes would not be significantly impacted by the small departures from a good fit.
+Specifically, conventional tests conclude that the model fit has issues nearly twice as frequently as humans would, and they frequently reject departures in the form of non-linearity and heteroskedasticity that are not visible to humans.
+These findings emphasize the crucial role of graphical diagnostics and support the integration of visual inference into regression diagnostics.
+Moreover, they provide a compelling rationale for the development of an automated visual inference system to evaluate lineups of residual plots.
+The comprehensive details of the research methodology, along with other intriguing findings can be found in the attached paper.
 

### 2023-04-11 Milestone_2/report/report.Rmd
fill in timetable and add more details in findings

@@
 
-The main objective of this research is to construct an automatic visual inference system capable of conducting visual tests on a large scale in the domain of model diagnostics, particularly regression diagnostics.
+The main objective of this research is to construct an automatic visual inference system capable of conducting visual tests on a large scale in the domain of regression diagnostics.
 The study will concentrate on three specific projects, namely
@@
 
-The study revealed that conventional residual-based statistical tests are more sensitive to weak residual departures from model assumptions compared to visual tests evaluated by humans, leading to excessive rejections even when downstream analysis and outcomes would not be significantly impacted by the small departures from a good fit.
+As shown in Figure \ref{fig:polypower} and Figure \ref{fig:heterpower}, the study revealed that appropriate conventional residual-based statistical tests are more sensitive to weak residual departures from model assumptions compared to visual tests evaluated by humans, leading to excessive rejections even when downstream analysis and outcomes would not be significantly impacted by the small departures from a good fit.
 Specifically, conventional tests conclude that the model fit has issues nearly twice as frequently as humans would, and they frequently reject departures in the form of non-linearity and heteroskedasticity that are not visible to humans.
+For instance, based on the row of scatter plots at the bottom of Figure \ref{fig:polypower}, we might argue that the non-linearity is not sufficiently problematic until an effect size of around 3 or 3.5.
+The RESET test would reject closer to an effect size of 2, but the visual test would reject closer 3.25.
+
+![Comparison of power between different tests for detecting non-linearity.
+The power curves are estimated using logistic regression, and the horizontal lines of dots represent non-reject and reject results from human observers for each lineup.
+The visual test has multiple power curves estimated from bootstrap samples.
+The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.\label{fig:polypower}](polypower-1.png)
+
+![Comparison of power between different tests for detecting heteroskedasticity.
+Main plot shows the power curves, with dots indicating non-reject and reject in visual testing of lineups.
+The multiple lines for the visual test arise from estimating the power on many bootstrap samples.
+The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.](heterpower-1.png)
+
 These findings emphasize the crucial role of graphical diagnostics and support the integration of visual inference into regression diagnostics.
@@
 
+We have also tried to develop a prototype of the automatic visual inference system with pure simulated data. The model
+
 # Thesis structure
 
-The thesis is structured
+The thesis is structured as follows.
+The first chapter describes the background on regression diagnostics, visual inference and computer vision models.
+The second chapter will be identical to the attached paper which focuses on the comparative study between visual testing and conventional testing in the application of regression diagnostics.
+The third chapter will describe the development of the a computer vision system to evaluate residual plot
 
-topic
+The fourth chapter will detail the
 
-chapter 1: Introduction
+The fifth chapter will be the discussion of the thesis and future work.
 
-chapter 2
-the first paper
+The first chapter of the thesis will be the introduction
+The second chapter of thesis will be the
 
-chapter 3
-build a computer vision system to evaluate residual plot
+The first project develops a prototype of automatic visual inference system for evaluating lineups of residual plots, with a focus on the classical normal linear regression model given it is one of the simplest predictive models.
+The automatic system is identical to the lineup protocol except evaluators are replaced by computer vision models.
+The computer vision models are trained by using data simulated from linear models with violations of different classical assumptions, such as linearity and homoscedasticity.
+Data of human performance on evaluating residual plots generated under the same simulation setting is collected by conducting online human subject experiments, such that the comparison between the power of human subject-assessed visual tests and the system-assessed visual tests can be made.
+Moreover, factors that affect the performance of the automatic system are studied for improving the architecture and design of the computer vision models.
 
-chapter 4?
-software publication
+The second project aims to extend the automatic system to lineup protocol with no assumptions on the type of the data plot and the data generating process.
+It only requires the protocol to asks the evaluators to select the most different plot(s) from the lineup.
+The automatic system will be built upon image similarity produced by computer vision models.
 
-chapter 5
-discussion and future work
+The third projects aims to package and deploy the automatic system as an adaptive online application such that empirical data can be collected and used in improving the system.
+Online machine learning technique will be used to update the weights of computer vision models dynamically.
 
@@
 
-- April: submit abstract to ASC
-- May: submit paper
-- June: submit poster/short paper to IEEE vis conf, start working on computer vision model
-- July: leave for few weeks
-- Aug
-- Sep: finalize the computer vision model
-- Oct: IEEE vis conf
-- Nov: web interface development 
-- Dec: ASC
+The timetable of my PhD study from April 2023 to August 2024 is provided in Table \ref{tab:timetable}.
 
-- Jan, 2024: 
-- Feb, 2024: 
-- Mar, 2024: submit paper
-- Aug, 2024: submit thesis
-
-# Difficulties
-
-things that are actually quite serious
-
-1. identified difficulties
-2. suggestions for overcoming these difficulties
+```{r}
+data.frame(Date = c("April", 
+                    "May",
+                    "June",
+                    "July",
+                    "Sep",
+                    "Oct",
+                    "Nov",
+                    "Dec",
+                    "Mar",
+                    "Aug"), 
+           Event = c("Submit the abstract of the first paper to Australian Statistical Conference (ASC)", 
+                     "Submit the first paper",
+                     "Submit a poster for the first paper to Institute of Electrical and Electronics Engineers (IEEE) Visualization Conference",
+                     "Leave for a month",
+                     "Finalize the computer vision model",
+                     "Attend the IEEE Visualization Conference",
+                     "Start the web interface development",
+                     "Attend the ASC",
+                     "Submit the second paper",
+                     "Submit thesis")) %>%
+  kableExtra::kable(booktabs = TRUE,
+                    label = "timetable",
+                    caption = "Timetable till thesis submission") %>%
+  kableExtra::kable_styling(latex_options = "scale_down") %>%
+  kableExtra::group_rows("2023", 1, 8) %>%
+  kableExtra::group_rows("2024", 9, 10)
+```
 

### 2023-04-11 Milestone_2/report/report.tex
fill in timetable and add more details in findings

@@
 
-The study revealed that conventional residual-based statistical tests are more sensitive to weak residual departures from model assumptions compared to visual tests evaluated by humans, leading to excessive rejections even when downstream analysis and outcomes would not be significantly impacted by the small departures from a good fit.
+As shown in Figure \ref{fig:polypower} and Figure \ref{fig:heterpower}, the study revealed that appropriate conventional residual-based statistical tests are more sensitive to weak residual departures from model assumptions compared to visual tests evaluated by humans, leading to excessive rejections even when downstream analysis and outcomes would not be significantly impacted by the small departures from a good fit.
 Specifically, conventional tests conclude that the model fit has issues nearly twice as frequently as humans would, and they frequently reject departures in the form of non-linearity and heteroskedasticity that are not visible to humans.
+For instance, based on the row of scatter plots at the bottom of Figure \ref{fig:polypower}, we might argue that the non-linearity is not sufficiently problematic until an effect size of around 3 or 3.5.
+The RESET test would reject closer to an effect size of 2, but the visual test would reject closer 3.25.
+
+\begin{figure}
+\centering
+\includegraphics{polypower-1.png}
+\caption{Comparison of power between different tests for detecting non-linearity.
+The power curves are estimated using logistic regression, and the horizontal lines of dots represent non-reject and reject results from human observers for each lineup.
+The visual test has multiple power curves estimated from bootstrap samples.
+The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.\label{fig:polypower}}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{heterpower-1.png}
+\caption{Comparison of power between different tests for detecting heteroskedasticity.
+Main plot shows the power curves, with dots indicating non-reject and reject in visual testing of lineups.
+The multiple lines for the visual test arise from estimating the power on many bootstrap samples.
+The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.}
+\end{figure}
+
 These findings emphasize the crucial role of graphical diagnostics and support the integration of visual inference into regression diagnostics.
@@
 
+We have also tried to develop a prototype of the automatic visual inference system with pure simulated data. The model
+
 \hypertarget{thesis-structure}{%
@@
 
-topic
+The thesis is structured as follows.
+The first chapter describes the background on regression diagnostics, visual inference and computer vision models.
+The second chapter will be identical to the attached paper which focuses on the comparative study between visual testing and conventional testing in the application of regression diagnostics.
+The third chapter will describe the development of the a computer vision system to evaluate residual plot
 
-chapter 1: Introduction
+The fourth chapter will detail the
 
-chapter 2
-the first paper
+The fifth chapter will be the discussion of the thesis and future work.
 
-chapter 3
-build a computer vision system to evaluate residual plot
+The first chapter of the thesis will be the introduction
+The second chapter of thesis will be the
 
-chapter 4?
-software publication
+The first project develops a prototype of automatic visual inference system for evaluating lineups of residual plots, with a focus on the classical normal linear regression model given it is one of the simplest predictive models.
+The automatic system is identical to the lineup protocol except evaluators are replaced by computer vision models.
+The computer vision models are trained by using data simulated from linear models with violations of different classical assumptions, such as linearity and homoscedasticity.
+Data of human performance on evaluating residual plots generated under the same simulation setting is collected by conducting online human subject experiments, such that the comparison between the power of human subject-assessed visual tests and the system-assessed visual tests can be made.
+Moreover, factors that affect the performance of the automatic system are studied for improving the architecture and design of the computer vision models.
 
-chapter 5
-discussion and future work
+The second project aims to extend the automatic system to lineup protocol with no assumptions on the type of the data plot and the data generating process.
+It only requires the protocol to asks the evaluators to select the most different plot(s) from the lineup.
+The automatic system will be built upon image similarity produced by computer vision models.
+
+The third projects aims to package and deploy the automatic system as an adaptive online application such that empirical data can be collected and used in improving the system.
+Online machine learning technique will be used to update the weights of computer vision models dynamically.
 
@@
 
-\begin{itemize}
-\item
-  April: submit abstract to ASC
-\item
-  May: submit paper
-\item
-  June: submit poster/short paper to IEEE vis conf, start working on computer vision model
-\item
-  July: leave for few weeks
-\item
-  Aug
-\item
-  Sep: finalize the computer vision model
-\item
-  Oct: IEEE vis conf
-\item
-  Nov: web interface development
-\item
-  Dec: ASC
-\item
-  Jan, 2024:
-\item
-  Feb, 2024:
-\item
-  Mar, 2024: submit paper
-\item
-  Aug, 2024: submit thesis
-\end{itemize}
+The timetable of my PhD study from April 2023 to August 2024 is provided in Table \ref{tab:timetable}.
 
-\hypertarget{difficulties}{%
-\section{Difficulties}\label{difficulties}}
+\begin{table}
 
-things that are actually quite serious
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  identified difficulties
-\item
-  suggestions for overcoming these difficulties
-\end{enumerate}
+\caption{\label{tab:timetable}Timetable till thesis submission}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{ll}
+\toprule
+Date & Event\\
+\midrule
+\addlinespace[0.3em]
+\multicolumn{2}{l}{\textbf{2023}}\\
+\hspace{1em}April & Submit the abstract of the first paper to Australian Statistical Conference (ASC)\\
+\hspace{1em}May & Submit the first paper\\
+\hspace{1em}June & Submit a poster for the first paper to Institute of Electrical and Electronics Engineers (IEEE) Visualization Conference\\
+\hspace{1em}July & Leave for a month\\
+\hspace{1em}Sep & Finalize the computer vision model\\
+\hspace{1em}Oct & Attend the IEEE Visualization Conference\\
+\hspace{1em}Nov & Start the web interface development\\
+\hspace{1em}Dec & Attend the ASC\\
+\addlinespace[0.3em]
+\multicolumn{2}{l}{\textbf{2024}}\\
+\hspace{1em}Mar & Submit the second paper\\
+\hspace{1em}Aug & Submit thesis\\
+\bottomrule
+\end{tabular}}
+\end{table}
 

### 2023-04-11 Milestone_2/report/report.Rmd
title of report

--- 2023-04-11 Milestone_2/report/report.Rmd
@@
 ---
-title: "Advances in Artificial Intelligence for Data Visualization: Automate Reading of Diagnostic Plots with Compute Vision Models"
+title: "Advances in Artificial Intelligence for Data Visualization: Automated Reading of Residual Plots with Computer Vision"
 author:
@@
 knitr::opts_chunk$set(echo = FALSE, cache=TRUE, messages=FALSE, warning=FALSE)
-library(visage)
-library(tidyverse)
+library(ggplot2)
 ```
@@
 Model diagnostics play a critical role in evaluating the accuracy and validity of a statistical model.
-They enable the assessment of the model's assumptions, detection of outliers, evaluation of how well the model fits the data, and identification of possible approaches to improve the model's performance.
+They enable the assessment of the model's assumptions about the data, detection of outliers, evaluation of how well the model fits the data, and identification of possible approaches to improve the model's performance.
 
-When conducting model diagnostics, despite the availability of numeric summaries endorsed by finite or asymptotic properties, graphical representations of data are often preferred or required by data analysts.
+When conducting model diagnostics, despite the availability of numeric summaries endorsed by finite or asymptotic properties, data analysts prefer or require graphical representations of data.
 The preference for visual diagnostics is attributed to its intuitive nature and the possibility of discovering unexpected abstract and unquantifiable insights.
@@
 The protocol is inspired by the police lineup technique employed in eyewitness identification of criminal suspects.
-It comprises $m$ randomly positioned plots, where one of them represents the data plot, while the remaining $m - 1$ plots represent the null plots with the same graphical structure, except that the data has been replaced with data consistent with the null hypothesis $H_0$.
-An example lineup is provided in Figure \ref{fig:lineup-example}.
-To compute the $p$-value of the visual test, the lineup will be independently presented to a number of participants, asking them to pick the most different plot.
+It comprises $m$ randomly positioned plots, where on of them represents the data plot, while the remaining $m - 1$ plots represent the null plots with the same graphical structure, except that the data has been replaced with data consistent with the null hypothesis $H_0$.
 Under $H_0$, the data plot is expected to be indistinguishable from the null plots, and the probability of correctly identifying the data plot by an observer is $1/m$.
-If a large number of participants correctly identify the data plot, the corresponding $p$-value will be small, indicating strong evidence against $H_0$.
-
-```{r lineup-example, fig.pos="t!", fig.height = 7, fig.width = 7, fig.cap = "Visual testing is conducted using a lineup, as in the example here.
-The residual plot computed from the observed data (plot $2^2 + 2$, exhibiting non-linearity) is embedded among 19 null plots, where the residuals are simulated from a standard error model.
-Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot.
-A small $p$-value would result from a substantial number selecting plot $2^2 + 2$."}
-vi_lineup <- get_vi_lineup()
-vi_lineup$poly_300$data %>%
-  VI_MODEL$plot_lineup(theme = theme_light(), remove_axis = TRUE, remove_grid_line = TRUE)
-```
+If the data plot is correctly identified, it provides evidence against $H_0$.
 
@@
 Moreover, it presents significant usability issues for individuals with visual impairments, resulting in reduced accessibility.
+
+The workload of individuals could be alleviated by automating repetitive tasks and providing standardized results in a controlled environment.
+The evaluation of visual tests on a large scale is not feasible without the aid of machines and technology.
 
@@
 
-In order to examine the potential applicability of integrating visual inference techniques into regression diagnostics, a pilot study was conducted in the first year with the participation of 64 individuals, which was followed by a formal study involving 443 participants in the subsequent year.
-The participants were presented with lineups consisting of 20 plots, where a residual plot was embedded along with 19 null plots, drawn with data simulated using the residual rotation technique.
-The study considered two primary forms of residual departures in multiple linear regression, namely non-linearity and heteroskedasticity.
-To enrich the visual features, different fitted value distributions, including normal, uniform, lognormal, and discrete, were also incorporated.
+- cut this p down, we don't need that many details
 
-The study revealed that conventional residual-based statistical tests are more sensitive to weak residual departures from model assumptions compared to visual tests evaluated by humans, leading to excessive rejections even when downstream analysis and outcomes would not be significantly impacted by the small departures from a good fit.
-Specifically, conventional tests conclude that the model fit has issues nearly twice as frequently as humans would, and they frequently reject departures in the form of non-linearity and heteroskedasticity that are not visible to humans.
-These findings emphasize the crucial role of graphical diagnostics and support the integration of visual inference into regression diagnostics.
-Moreover, they provide a compelling rationale for the development of an automated visual inference system to evaluate lineups of residual plots.
-The comprehensive details of the research methodology, along with other intriguing findings can be found in the attached paper.
+In the initial phase of this research, a pilot investigation was undertaken to explore the viability of implementing visual inference in regression diagnostics.
+The pilot study involved 64 participants and focused on utilizing a lineup consisting of 20 plots, wherein a residual plot was embedded along with 19 null plots, simulated using the residual rotation technique.
+The study considered two primary forms of residual departures, namely non-linearity and heteroskedasticity, within a multiple linear regression context.
+To enrich the visual features, different fitted value distributions, including normal, uniform, lognormal, and discrete uniform, were also incorporated.
+
+The outcomes of this pilot study demonstrated that the power of visual tests in regression diagnostics increases proportionately with the number of lineup evaluations.
+Nevertheless, it was observed that conventional tests, such as the F-test for non-linearity and the Breusch-Pagan test for heteroskedasticity, exhibited significantly greater power than the visual tests.
+While the pilot study successfully provided a comparative analysis of the power of visual and conventional tests, the impact of various factors on the power of visual tests could not be fully analyzed due to inadequate data.
+
+We acknowledge the limitations of this pilot study, particularly in the design of the simulated model, which made it challenging to explicate the involved factors.
+Furthermore, the methodology utilized for effect size calculation, p-value calculation, and power estimation was incipient.
+Hence, a follow-up study is necessary to address these limitations and obtain a more comprehensive understanding of the potential of visual inference in regression diagnostics.
+
+In the subsequent year, a similar study was conducted, utilizing simpler simulated models to elicit more distinct visual patterns. Specifically, the non-linearity model was formulated as
+
+\begin{align} \label{eq:nonlinearity-model}
+\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
+\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
+\boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\
+\boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)),
+\end{align}
+
+\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vectors of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
+
+\begin{equation} \label{eq:scaling-function}
+g(\boldsymbol{x}, k) = \frac{\boldsymbol{x} - min(\boldsymbol{x})}{max(\boldsymbol{x}) - min(\boldsymbol{x})}(2k - k), \quad \text{for} \quad k > 0. 
+\end{equation}
+
+The heteroskedasticity model was formulated as
+
+\begin{align} \label{eq:heter-model}
+\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
+\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
+\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}, 1 + (2 - |a|)(\boldsymbol{x} - a)^2b \boldsymbol{I}). 
+\end{align}
+
+The null regression model used to fit the realizations generated by the above models is formulated as
+
+\begin{equation} \label{eq:null-model}
+\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
+\end{equation}
+
+\noindent where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.
+
+Since $z = O(x^j)$, for $j > 1$, $z$ is a higher order term leaves out by the null regression, which will violate the linearity model assumption.
+Similarly, for $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the predictor $\boldsymbol{x}$, which will lead to the presence of heteroskedasticity.
+
+- draft below
+
+The effect size was derived using the Kullback-Leibler divergence method, formulated as Equation 3.
+To account for the interdependence of lineups presented to different participants, p-value calculation, as suggested by VanderPlas et al.
+(2021), was adopted.
+Furthermore, power estimation was simplified through a logistic regression model that lacked an intercept but included a fixed offset to ensure that the power equated to 5% when the effect size was zero.
+A detailed description of the experimental design, effect size derivation, p-value calculation, and power estimation can be found in the attached paper.
+
+3. summarise the work in the first year
+- an experiment to compare visual test with conventional test 
+- a prototype model for lineup evaluation
+- points out the limitation of the work
+4. summarise the work in the second year
+- refine and reconduct the experiment
+5. main findings to date
+- findings from the first year
+- findings from the second year
 

### 2023-04-11 Milestone_2/report/report.tex
title of report

--- 2023-04-05 Milestone_2/report/report.tex
@@
 \hypersetup{
-            pdftitle={Advances in Artificial Intelligence for Data Visualization: Automate Reading of Diagnostic Plots with Compute Vision Models},
+            pdftitle={Advances in Artificial Intelligence for Data Visualization: Automated Reading of Residual Plots with Computer Vision},
             pdfborder={0 0 0},
@@
 
-\title{Advances in Artificial Intelligence for Data Visualization: Automate Reading of Diagnostic Plots with Compute Vision Models}
+\title{Advances in Artificial Intelligence for Data Visualization: Automated Reading of Residual Plots with Computer Vision}
 

### 2023-04-11 Milestone_2/report/report.tex

--- 2023-04-11 Milestone_2/report/report.tex
@@
 %% Any special functions or other packages can be loaded here.
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 

### 2023-04-11 Milestone_2/report/report.Rmd
recompile
finish report

@@
 The multiple lines for the visual test arise from estimating the power on many bootstrap samples.
-The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.](heterpower-1.png)
+The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.\label{fig:heterpower}](heterpower-1.png)
 
@@
 
-We have also tried to develop a prototype of the automatic visual inference system with pure simulated data. The model
-
 # Thesis structure
 
-The thesis is structured as follows.
-The first chapter describes the background on regression diagnostics, visual inference and computer vision models.
-The second chapter will be identical to the attached paper which focuses on the comparative study between visual testing and conventional testing in the application of regression diagnostics.
-The third chapter will describe the development of the a computer vision system to evaluate residual plot
+The thesis will be structured as follows.
 
-The fourth chapter will detail the
+1.
+Chapter One provides an overview of regression diagnostics, visual inference, and computer vision models.
+This chapter serves as an introductory section of the thesis and forms the basis for subsequent discussions.
 
-The fifth chapter will be the discussion of the thesis and future work.
+2. Chapter Two presents a comparative study between visual testing and conventional testing in the context of regression diagnostics. This chapter will be identical to the attached paper.
 
-The first chapter of the thesis will be the introduction
-The second chapter of thesis will be the
+3.
+Chapter Three discusses the development of an automatic visual inference system for evaluating lineups of residual plots.
+The computer vision models used in this chapter are primarily trained on simulated data from linear models that violate classical assumptions such as linearity and homoscedasticity.
+The chapter concludes with an evaluation of the performance of the system, comparing the power and characteristics of human subject-assessed visual tests with computer-assessed visual tests.
 
-The first project develops a prototype of automatic visual inference system for evaluating lineups of residual plots, with a focus on the classical normal linear regression model given it is one of the simplest predictive models.
-The automatic system is identical to the lineup protocol except evaluators are replaced by computer vision models.
-The computer vision models are trained by using data simulated from linear models with violations of different classical assumptions, such as linearity and homoscedasticity.
-Data of human performance on evaluating residual plots generated under the same simulation setting is collected by conducting online human subject experiments, such that the comparison between the power of human subject-assessed visual tests and the system-assessed visual tests can be made.
-Moreover, factors that affect the performance of the automatic system are studied for improving the architecture and design of the computer vision models.
+4.
+Chapter Four provides detailed description of the open-source software developed during the course of the research.
+This includes the R package `bandicoot` for code maintenance with object-oriented programming, `visage` for running visual inference experiments on linear regression models, and one or multiple R packages for employment of the automatic visual inference system.
+Additionally, an online application of the automatic visual inference system is introduced.
+This chapter may be splitted into multiple chapters to provide a more in-depth description of each tool.
 
-The second project aims to extend the automatic system to lineup protocol with no assumptions on the type of the data plot and the data generating process.
-It only requires the protocol to asks the evaluators to select the most different plot(s) from the lineup.
-The automatic system will be built upon image similarity produced by computer vision models.
-
-The third projects aims to package and deploy the automatic system as an adaptive online application such that empirical data can be collected and used in improving the system.
-Online machine learning technique will be used to update the weights of computer vision models dynamically.
+5.
+Chapter Five offers a comprehensive discussion of the thesis, emphasizing the implications and contributions of the research.
+The chapter also addresses the limitations of the research and outlines potential future research directions.
 

### 2023-04-11 Milestone_2/report/report.tex
recompile
finish report

@@
 %% Any special functions or other packages can be loaded here.
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
 
@@
 The multiple lines for the visual test arise from estimating the power on many bootstrap samples.
-The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.}
+The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.\label{fig:heterpower}}
 \end{figure}
@@
 
-We have also tried to develop a prototype of the automatic visual inference system with pure simulated data. The model
-
 \hypertarget{thesis-structure}{%
@@
 
-The thesis is structured as follows.
-The first chapter describes the background on regression diagnostics, visual inference and computer vision models.
-The second chapter will be identical to the attached paper which focuses on the comparative study between visual testing and conventional testing in the application of regression diagnostics.
-The third chapter will describe the development of the a computer vision system to evaluate residual plot
+The thesis will be structured as follows.
 
-The fourth chapter will detail the
-
-The fifth chapter will be the discussion of the thesis and future work.
-
-The first chapter of the thesis will be the introduction
-The second chapter of thesis will be the
-
-The first project develops a prototype of automatic visual inference system for evaluating lineups of residual plots, with a focus on the classical normal linear regression model given it is one of the simplest predictive models.
-The automatic system is identical to the lineup protocol except evaluators are replaced by computer vision models.
-The computer vision models are trained by using data simulated from linear models with violations of different classical assumptions, such as linearity and homoscedasticity.
-Data of human performance on evaluating residual plots generated under the same simulation setting is collected by conducting online human subject experiments, such that the comparison between the power of human subject-assessed visual tests and the system-assessed visual tests can be made.
-Moreover, factors that affect the performance of the automatic system are studied for improving the architecture and design of the computer vision models.
-
-The second project aims to extend the automatic system to lineup protocol with no assumptions on the type of the data plot and the data generating process.
-It only requires the protocol to asks the evaluators to select the most different plot(s) from the lineup.
-The automatic system will be built upon image similarity produced by computer vision models.
-
-The third projects aims to package and deploy the automatic system as an adaptive online application such that empirical data can be collected and used in improving the system.
-Online machine learning technique will be used to update the weights of computer vision models dynamically.
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  Chapter One provides an overview of regression diagnostics, visual inference, and computer vision models.
+This chapter serves as an introductory section of the thesis and forms the basis for subsequent discussions.
+\item
+  Chapter Two presents a comparative study between visual testing and conventional testing in the context of regression diagnostics. This chapter will be identical to the attached paper.
+\item
+  Chapter Three discusses the development of an automatic visual inference system for evaluating lineups of residual plots.
+The computer vision models used in this chapter are primarily trained on simulated data from linear models that violate classical assumptions such as linearity and homoscedasticity.
+The chapter concludes with an evaluation of the performance of the system, comparing the power and characteristics of human subject-assessed visual tests with computer-assessed visual tests.
+\item
+  Chapter Four provides detailed description of the open-source software developed during the course of the research.
+This includes the R package \texttt{bandicoot} for code maintenance with object-oriented programming, \texttt{visage} for running visual inference experiments on linear regression models, and one or multiple R packages for employment of the automatic visual inference system.
+Additionally, an online application of the automatic visual inference system is introduced.
+This chapter may be splitted into multiple chapters to provide a more in-depth description of each tool.
+\item
+  Chapter Five offers a comprehensive discussion of the thesis, emphasizing the implications and contributions of the research.
+The chapter also addresses the limitations of the research and outlines potential future research directions.
+\end{enumerate}
 

### 2023-04-11 Milestone_presentation_2/Milestone_presentation_2/Milestone_presentation_2.Rmd
init the M2 presentation folder

--- 2022-03-16 Milestone_presentation_1/Milestone_presentation_1/Milestone_presentation_1.Rmd
@@
 
-#### Due to limited computing resources, we apply **transfer learning** to train five computer vision models.
+#### Due to limited computing resources, we apply **transfer learning** `r BibOptions(max.names = 1);Citep(myBib, "weiss_survey_2016")` to train five computer vision models.
 
@@
 
-#### Due to limited computing resources, we apply **transfer learning** to train five computer vision models.
-
-#### Five **pre-trained models** are used, including **VGG16**, **VGG19**, **Xception**, **ResNet152V2** and **InceptionV3**.
+#### Due to limited computing resources, we apply **transfer learning** `r BibOptions(max.names = 1);Citep(myBib, "weiss_survey_2016")` to train five computer vision models.
+
+#### Five **pre-trained models** are used, including **VGG16** `r BibOptions(max.names = 5);Citep(myBib, "simonyan_very_2015")`, **VGG19** `r BibOptions(max.names = 5);Citep(myBib, "simonyan_very_2015")`, **Xception** `r BibOptions(max.names = 5);Citep(myBib, "chollet_xception_2017")`, **ResNet152V2** `r BibOptions(max.names = 1);Citep(myBib, "he_identity_2016")` and **InceptionV3** `r BibOptions(max.names = 1);Citep(myBib, "szegedy_rethinking_2015")`.
 
@@
 
-#### Model is trained with **Tensorflow** and **Keras**.
+#### Models are trained with **Tensorflow** and **Keras**.
 
@@
 
-#### Deploy the automatic visual inference as an online aplication
+#### Deploy the automatic visual inference as an online aplication.
 
@@
 
-# References
+# Bibliography
 
@@
 
+Chollet, F. (2017). “Xception: Deep Learning with
+Depthwise Separable Convolutions”. In:
+_arXiv:1610.02357 [cs]_. arXiv: 1610.02357. URL:
+[http://arxiv.org/abs/1610.02357](http://arxiv.org/abs/1610.02357).
+
 Fukushima, K. et al. (1982). “Neocognitron: A
@@
 
+He, K. et al. (2016). “Identity Mappings in Deep
+Residual Networks”. In: _arXiv:1603.05027 [cs]_. URL:
+[http://arxiv.org/abs/1603.05027](http://arxiv.org/abs/1603.05027).
+
+Simonyan, K. et al. (2015). “Very Deep Convolutional
+Networks for Large-Scale Image Recognition”. In:
+_arXiv:1409.1556 [cs]_. URL:
+[http://arxiv.org/abs/1409.1556](http://arxiv.org/abs/1409.1556).
+
+Szegedy, C. et al. (2015). “Rethinking the Inception
+Architecture for Computer Vision”. In:
+_arXiv:1512.00567 [cs]_. URL:
+[http://arxiv.org/abs/1512.00567](http://arxiv.org/abs/1512.00567).
+
+Weiss, K. et al. (2016). “A survey of transfer
+learning”. En. In: _Journal of Big Data_ 3.1, p. 9.
+ISSN: 2196-1115. DOI:
+[10.1186/s40537-016-0043-6](https://doi.org/10.1186%2Fs40537-016-0043-6).
+URL:
+[http://journalofbigdata.springeropen.com/articles/10.1186/s40537-016-0043-6](http://journalofbigdata.springeropen.com/articles/10.1186/s40537-016-0043-6).
+
 <!-- `r BibOptions(style = "markdown");PrintBibliography(myBib)` -->

### 2023-04-15 Milestone_presentation_2/Milestone_presentation_2/Untitled.Rmd
init M2 slides

---
title: "Advances in Artificial Intelligence for Data Visualization: Automated Reading of Residual Plots with Computer Vision"
subtitle: ""  
author: 
  - "Weihao (Patrick) Li"
date: '`r format(Sys.Date(), format="%B %d, %Y")`'
output:
  xaringan::moon_reader:
    seal: false
    css: [xaringan-themer.css, mine.css]
    nature:
      countdown: 55000
      beforeInit: "scripts/my_macro.js"
      countIncrementalSlides: false
      slideNumberFormat: "%current%/%total%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=5, fig.height=5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r}
library(RefManageR)
BibOptions(check.entries = FALSE,
           bib.style = "authoryear",
           cite.style = "apa",
           style = "Bibtex",
           # max.names = 3,
           longnamesfirst = FALSE,
           hyperlink = FALSE,
           dashed = FALSE)
myBib <- ReadBib("./Visual Inference.bib", check = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
library(tidyverse)
library(visage)
style_solarized_light(
  background_color = "#FFFFFF",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono"),
  text_color = "#000000",
  text_font_size = "1.0rem",
  colors = c(myblue = "#006DAE", mywhite = "#FFFFFF")
)
```

count: false

<!-- need a backaground image -->

.pull-left-full[
<h2 class="myblue"> Advances in Artificial Intelligence for Data Visualization: Automated Reading of Residual Plots with Computer Vision </h2>

<h3 class="myblue"> Progress Review Presentation </h3>

<br>
<br>

<h3 class="myblue"> Weihao (Patrick) Li </h3>

<h3 class="myblue"> Supervised by Di Cook, Emi Tanaka and Susan VanderPlas </h3>

<!-- <h3 class="myblue"> `r format(Sys.Date(), format="%B %d, %Y")` </h3> -->
]

.pull-right[

<br>

![](images/159.png)
]

---

# Thesis structure (project structure)

Mention we will focus on the first chapter. (bold the chapter text)

---

# Regression diagnostics

Why we need to perform regression diagnostics?
Available methods.

---

# Reading residual plot vs. conventional statistical testing

(Use non-linearity as the example)

How people read residual plot?
How statistical testing detect residual departures?

---

# Visual inference

What is visual inference?

Underline that it is a valid inferential framework. Highlights the difference from reading a residual plot.

---

# Integrate visual tests into regression diagnostics

(Still use non-linearity as the example)

How to perform a visual test in the context of regression diagnostics.

Show the $p$-value by assuming the responses.

---

# Limitations of visual inference

Discuss the limitations of visual inference.

---

# Computer vision models

Briefly mention the potential solution to the problem.

---

# A compartative study

Provide the motivation and main objective of conducting such an experiment.

---

# Experiment design

Outlines the model used to simulate the experiment data.

---

# Effect size

Describe the effect size calculation.

---

# $P$-value calculation

Briefly discuss the $p$-value calculation.

---

# Power estimation

Write down the model used to estimate the power.

---

# Study website

Provide screenshots/examples of the study website.

---

# Results

Provide an overview of the collated data.

---

# Power comparison of the tests

---

# Comparison of test decisions based on $p$-values

---

# Effect of amount of non-linearity

---

# Effect of shape of heteroskedasticity

---

# Effect of fitted value distributions

---

# Timetable

### 2023-04-15 Milestone_presentation_2/Milestone_presentation_2/Untitled.Rmd
decide the structure of the slides

@@
 
-<h3 class="myblue"> Weihao (Patrick) Li </h3>
+<h4 class="myblue"> Weihao (Patrick) Li </h3>
 
-<h3 class="myblue"> Supervised by Di Cook, Emi Tanaka and Susan VanderPlas </h3>
+<h4 class="myblue"> Supervised by Di Cook, Emi Tanaka and Susan VanderPlas </h3>
 
@@
 
-![](images/159.png)
+![](images/505.png)
 ]
@@
 
-# Thesis structure (project structure)
+.center[
+# 📚Thesis structure📚
+]
 
-Mention we will focus on the first chapter. (bold the chapter text)
+1. **Exploring the application of visual inference in regression diagnostics and comparing it with conventional hypothesis tests.**
+2. Designing an automated visual inference system to assess lineups of residual plots of classical normal linear regression model.
+3. Deploying the automatic visual inference system as an online application and publishing the relevant open-source software.
+
+<br>
+<br>
+
+This presentation will be focused on the **first project**. Parts of the second and third project have been done and will be mentioned later.
 
@@
 
-# Regression diagnostics
+.center[
+# 🔍Regression diagnostics🔍
+]
 
-Why we need to perform regression diagnostics?
-Available methods.
+#### Diagnostics are the key to determining whether there is anything importantly wrong with a model.
+
+#### In linear regression analysis, residuals from the model fit are commonly used.
+Residuals summarise what is not captured by the model, and thus provide the capacity to identify what might be wrong.
 

### 2023-04-16 Milestone_presentation_2/Milestone_presentation_2/Untitled.Rmd
up to VI introduction

@@
 ```{r}
+set.seed(10086)
 library(RefManageR)
@@
 
-1. **Exploring the application of visual inference in regression diagnostics and comparing it with conventional hypothesis tests.**
-2. Designing an automated visual inference system to assess lineups of residual plots of classical normal linear regression model.
-3. Deploying the automatic visual inference system as an online application and publishing the relevant open-source software.
+#### 1. **Exploring the application of visual inference in regression diagnostics and comparing it with conventional hypothesis tests.**
+
+#### 2. Designing an automated visual inference system to assess lineups of residual plots of classical normal linear regression model.
+
+#### 3. Deploying the automatic visual inference system as an online application and publishing the relevant open-source software.
 
@@
 
+class: center, middle
+
+# A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol
+
+---
+
 .center[
@@
 
-#### Diagnostics are the key to determining whether there is anything importantly wrong with a model.
+.pull-left-center-80[
 
-#### In linear regression analysis, residuals from the model fit are commonly used.
-Residuals summarise what is not captured by the model, and thus provide the capacity to identify what might be wrong.
+.image-bottom-0[
+
+```{r}
+set.seed(10086)
+
+# Base model
+mod <- poly_model(include_z = FALSE, sigma = 0.25)
+
+# Base data
+dat <- mod$gen(300)
+
+bind_rows(
+  dat %>%
+    mutate(type = "(A) Good residuals"),
+  mod$set_prm("include_z", TRUE)$
+    gen(300, computed = select(dat, x, e)) %>%
+    mutate(type = "(B) Non-linearity"),
+  heter_model(b = 64)$
+    gen(300, computed = select(dat, x, e)) %>%
+    mutate(type = "(C) Heteroskedasticity"),
+  heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
+    gen(300, computed = select(dat, x)) %>%
+    mutate(type = "(D) Non-normality")
+) %>%
+  mod$plot(remove_grid_line = TRUE, theme = theme_light(base_size = 12)) +
+  facet_wrap(~type, ncol = 2, scales = "free") +
+  theme(axis.text = element_blank()) +
+  xlab("Fitted values") +
+  ylab("Residuals")
+```
+
+]
+
+.center[
+.caption[
+Fig. 1: _One classically good looking residual plot and three residual plots with model defects._
+]
+]
+
+]
+
+.pull-right-120[
+
+**Diagnostics** are the key to determining whether there is anything **importantly wrong** with a model.
+
+**Residuals** summarise what is **not captured by the regression model**.
+
+**Residual plots** are commonly used to diagnose **non-linearity** and **heteroskedasticity**.
+**Non-normality** is usually harder to detect from a residual plot.
+A favourable graphical summary for this task is the **quantile-quantile plot**.
+
+]
+
+---
+
+.center[
+# 🔍Regression diagnostics🔍
+]
+
+.pull-left-center-80[
+
+.image-bottom-0[
+
+```{r}
+set.seed(10086)
+
+# Base model
+mod <- poly_model(include_z = FALSE, sigma = 0.25)
+
+# Base data
+dat <- mod$gen(300)
+
+bind_rows(
+  dat %>%
+    mutate(type = "(A) Good residuals"),
+  mod$set_prm("include_z", TRUE)$
+    gen(300, computed = select(dat, x, e)) %>%
+    mutate(type = "(B) Non-linearity"),
+  heter_model(b = 64)$
+    gen(300, computed = select(dat, x, e)) %>%
+    mutate(type = "(C) Heteroskedasticity"),
+  heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
+    gen(300, computed = select(dat, x)) %>%
+    mutate(type = "(D) Non-normality")
+) %>%
+  mod$plot(remove_grid_line = TRUE, theme = theme_light(base_size = 12)) +
+  facet_wrap(~type, ncol = 2, scales = "free") +
+  theme(axis.text = element_blank()) +
+  xlab("Fitted values") +
+  ylab("Residuals")
+```
+
+]
+
+.center[
+.caption[
+Fig. 1: _One classically good looking residual plot and three residual plots with model defects._
+]
+]
+
+]
+
+.pull-right-120[
+
+Many different hypothesis tests are available to detect specific model defects.
+
+Non-linearity: 
+  - **F-test**
+  - **RESET test** (Ramsey 1969)
+
+Heteroskedasticity: 
+  - **White test** (White 1980)
+  - **BP test** (Breusch and Pagan 1979)
+
+Non-normality: 
+  - **SW test** (Shapiro and Wilk 1965)
+  - **Jarque–Bera test** (Jarque and Bera 1980)
+
+]
+
+---
+
+.center[
+# 🔍Regression diagnostics🔍
+]
+
+.pull-left-center-80[
+
+.image-bottom-0[
+
+```{r}
+set.seed(10086)
+
+# Base model
+mod <- poly_model(include_z = FALSE, sigma = 0.25)
+
+# Base data
+dat <- mod$gen(300)
+
+bind_rows(
+  dat %>%
+    mutate(type = "(A) Good residuals"),
+  mod$set_prm("include_z", TRUE)$
+    gen(300, computed = select(dat, x, e)) %>%
+    mutate(type = "(B) Non-linearity"),
+  heter_model(b = 64)$
+    gen(300, computed = select(dat, x, e)) %>%
+    mutate(type = "(C) Heteroskedasticity"),
+  heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
+    gen(300, computed = select(dat, x)) %>%
+    mutate(type = "(D) Non-normality")
+) %>%
+  mod$plot(remove_grid_line = TRUE, theme = theme_light(base_size = 12)) +
+  facet_wrap(~type, ncol = 2, scales = "free") +
+  theme(axis.text = element_blank()) +
+  xlab("Fitted values") +
+  ylab("Residuals")
+```
+
+]
+
+.center[
+.caption[
+Fig. 1: _One classically good looking residual plot and three residual plots with model defects._
+]
+]
+
+]
+
+.pull-right-120[
+
+BP and SW tests reject the residual plots exhibiting structure that **they weren’t designed for** (**type III error**).
+
+Most residual-based tests for particular types of departure are also sensitive to other types of departures (Cook and Weisberg 1982).
+
+```{r}
+set.seed(10086)
+
+# Base model
+mod <- poly_model(include_z = FALSE, sigma = 0.25)
+
+# Base data
+dat0 <- mod$gen(300)
+
+# Replicate data in Figure 1
+dat1 <- mod$set_prm("include_z", TRUE)$gen(300, computed = select(dat0, x, e))
+dat2 <- heter_model(b = 64)$gen(300, computed = select(dat0, x, e))
+dat3 <- heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
+               gen(300, computed = select(dat0, x))
+
+dat_list <- list(dat0, dat1, dat2, dat3)
+
+table_dat <- data.frame(plot = c("A", "B", "C", "D"), 
+                        model = c("None", 
+                                  "Non-linearity", 
+                                  "Heteroskedasticity", 
+                                  "Non-normality"),
+                        r = map_dbl(dat_list, 
+                                    ~POLY_MODEL$test(.x, 
+                                                     test = "RESET", 
+                                                     power = 2:4)$p_value),
+                        b = map_dbl(dat_list, 
+                                    ~HETER_MODEL$test(.x)$p_value),
+
+                        # SW test has not been included by visage
+                        s = map_dbl(dat_list, 
+                                    ~shapiro.test(.x$.resid)$p.value)) %>%
+
+  # 3 decimal points
+  mutate(across(r:s, ~ format(round(.x, digits = 3), nsmall = 3))) %>%
+
+  # Italic if reject
+  mutate(across(r:s, ~ kableExtra::cell_spec(.x, italic = as.numeric(.x) <= 0.05)))
+
+  table_dat %>%
+  knitr::kable(col.names = c("Plot", 
+                             "Departures", 
+                             "RESET", 
+                             "BP", 
+                             "SW"),
+               align = "llrrr",
+               escape = FALSE,
+               format = "html",
+               booktabs = TRUE)
+```
+
+.center[
+.caption[
+Tab. 1: _Statistical significance testing for departures from good residuals. Shown are the \\(p\\)-values calculated for the RESET, the BP and the SW tests._
+]
+]
+
+]
+
+---
+
+.center[
+# 🔍Regression diagnostics🔍
+]
+
+.pull-right-center-80[
+
+.image-bottom-0[
+
+```{r}
+set.seed(10086)
+
+# Base model
+mod <- poly_model(sigma = 8)
+dat <- mod$gen(1000)
+this_p_value <- mod$test(dat, test = "RESET", power = 2:4)
+dat %>%
+mod$plot(remove_grid_line = TRUE, theme = theme_light(base_size = 12)) +
+  theme(axis.text = element_blank()) +
+  xlab("Fitted values") +
+  ylab("Residuals")
+```
+
+]
+
+.center[
+.caption[
+Fig. 1: _One classically good looking residual plot but rejected by the RESET test with $p\text{-value} = `r format(round(this_p_value$p_value, digits = 3), nsmall = 3)`$._
+]
+]
+
+]
+
+.pull-left[
+
+With **large sample sizes**, hypothesis tests may reject a residual plot when there is only a **slight departure** (Belsley, Kuh, and Welsch 1980).
+
+While such rejections may be **statistically correct**, their sensitivity may render the results **impractical**.
+
+**Minor defects** in the model are **unlikely** to have a **significant impact** and may be best disregarded for practical purposes.
+
+Therefore, a key goal of **residual plot diagnostics** is to **identify** potential issues that could lead to **incorrect conclusions or errors in subsequent analyses**.
+
+]
+
+---
+
+.center[
+# 📜Liteature of regression diagnostics📜
+]
+
+### Plotting the data is the recommended method by experts for diagnosing model fits.
+
+- Draper and Smith (1998) and Belsley, Kuh, and Welsch (1980):
+> Residual plots are usually revealing when the assumptions are violated.
+
+- Cook and Weisberg (1982):
+> Formal tests and graphical procedures are complementary and both have a place in residual analysis, but graphical methods are easier to use.
+
+- Montgomery and Peck (1982):
+> Residual plots are more informative in most practical situations than the corresponding conventional hypothesis tests.
 

### 2023-04-17 Milestone_presentation_2/Milestone_presentation_2/Untitled.Rmd
find the residual plot and provide a lineup

@@
 .caption[
-Fig. 1: _One classically good looking residual plot but rejected by the RESET test with $p\text{-value} = `r format(round(this_p_value$p_value, digits = 3), nsmall = 3)`$._
+Fig. 2: _One classically good looking residual plot but rejected by the RESET test with $p\text{-value} = `r format(round(this_p_value$p_value, digits = 3), nsmall = 3)`$._
 ]
@@
 
-While such rejections may be **statistically correct**, their sensitivity may render the results **impractical**.
+While such rejections may be **statistically correct**, their **sensitivity** may render the results **impractical**.
 
@@
 
-# Reading residual plot vs. conventional statistical testing
+.center[
+# 💔False visual discoveries💔
+]
 
-(Use non-linearity as the example)
+.pull-left-center-80[
 
-How people read residual plot?
-How statistical testing detect residual departures?
+.image-bottom-0[
+
+```{r}
+set.seed(10131)
+mod <- heter_model(b = 0, x = rand_lognormal())
+ori_dat <- mod$gen(300)
+
+ori_dat %>%
+  VI_MODEL$plot(theme = theme_light(base_size = 12)) +
+  xlab("Fitted values") +
+  ylab("Residuals")
+```
+
+]
+
+.center[
+.caption[
+Fig. 3: _Residual plot of a simple linear regression._
+]
+]
+
+]
+
+.pull-right[
+
+Can you find the evidence on heteroskedasticity?
+
+- Vertical spread of the points varies with the fitted values.
+
+However, the residuals are actually obtained from a **correctly specified classical normal linear model**!
+
+**Unsecured** and **unconfirmed** visual discoveries will lead to **over or under-interpretations of the data**.
+
+]
+
+---
+
+# Visual inference
+
+.pull-right-center-80[
+
+```{r}
+set.seed(10131)
+mod$gen_lineup(300, k = 20, pos = 8) %>%
+  filter(null != FALSE) %>%
+  bind_rows(ori_dat %>% mutate(k = 8, null = FALSE)) %>%
+  VI_MODEL$plot_lineup(theme = theme_light(base_size = 12),
+                       remove_grid_line = TRUE,
+                       remove_axis = TRUE)
+```
+.center[
+.caption[
+Fig. 4: _Lineup of residual plots._
+]
+]
+
+]
 

### 2023-04-17 Milestone_presentation_2/Milestone_presentation_2/Untitled.Rmd
prettify the slides

@@
   text_font_size = "1.0rem",
+  header_h1_font_size = "2.0rem",
   colors = c(myblue = "#006DAE", mywhite = "#FFFFFF")
@@
 
+<br>
+
 #### 1. **Exploring the application of visual inference in regression diagnostics and comparing it with conventional hypothesis tests.**
@@
 
-.center[
-# 🔍Regression diagnostics🔍
-]
-
-.pull-left-center-80[
+.pull-left-center[
 
@@
 
-.pull-right-120[
+.pull-right[
+
+# 🔍Regression diagnostics🔍
+
+<br>
 
@@
 
-.pull-right-120[
+.pull-right[
+
+# 🔍Regression diagnostics🔍
 
@@
 
-Most residual-based tests for particular types of departure are also sensitive to other types of departures (Cook and Weisberg 1982).
+Most residual-based tests for particular types of departure are also sensitive to **other types of departures** (Cook and Weisberg 1982).
 
@@
 
-.center[
-# 🔍Regression diagnostics🔍
-]
-
-.pull-right-center-80[
+.pull-right-center[
 
@@
 
+# 🔍Regression diagnostics🔍
+
+<br>
+<br>
+
 With **large sample sizes**, hypothesis tests may reject a residual plot when there is only a **slight departure** (Belsley, Kuh, and Welsch 1980).
 
-While such rejections may be **statistically correct**, their **sensitivity** may render the results **impractical**.
-
+- While such rejections may be **statistically correct**, their **sensitivity** may render the results **impractical**.
 **Minor defects** in the model are **unlikely** to have a **significant impact** and may be best disregarded for practical purposes.
 
-Therefore, a key goal of **residual plot diagnostics** is to **identify** potential issues that could lead to **incorrect conclusions or errors in subsequent analyses**.
+Residual plot diagnostics:
+  - **Identify** potential issues that could lead to **incorrect conclusions or errors in subsequent analyses**.
 
@@
 
-.center[
-# 💔False visual discoveries💔
-]
-
-.pull-left-center-80[
+.pull-left-center[
 
@@
 
+# 🤔False visual discoveries🤔
+
+<br>
+<br>
+
 Can you find the evidence on heteroskedasticity?
@@
 
+.pull-left[
+
 # Visual inference
 
-.pull-right-center-80[
+If we embed the residual plot in a matrix of **null plots**, we will find the visual discovery is
+
+]
+
+.pull-right-center[
 
@@
 .caption[
-Fig. 4: _Lineup of residual plots._
+Fig. 4: _A lineup of 20 residual plots of a simple linear regression. (Plot No.8 is Fig.3) Can you find the most different one?_
 ]

### 2023-04-17 Milestone_presentation_2/Milestone_presentation_2/Untitled.Rmd
prettify the slides again

@@
       countIncrementalSlides: false
-      slideNumberFormat: "%current%/%total%"
+      slideNumberFormat: "%current%"
       highlightStyle: github
@@
   header_font_google = google_font("Josefin Sans"),
-  text_font_google   = google_font("Montserrat", "300", "300i"),
+  text_font_google   = google_font("Montserrat", "300", "300i", "400i", "600"),
   code_font_google   = google_font("Fira Mono"),
@@
 .center[
-# 📚Thesis structure📚
+# 📚Thesis Structure📚
 ]
@@
 # A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol
+
+#### Weihao Li, Dianne Cook, Emi Tanaka and Susan VanderPlas (2023)
+
+---
+
+.center[
+# 🔍Regression Diagnostics🔍
+]
+
+**Diagnostics** are the key to determining whether there is anything **importantly wrong** with a model.
+
+**Residuals** summarise what is **not captured by the regression model**. Many different residual-based hypothesis tests are available to detect specific model defects.
+
+<br>
+<br>
+<br>
+
+.pull-right-33p[
+
+Non-linearity: 
+  - **F-test**
+  - **RESET test** (Ramsey 1969)
+]
+
+.pull-right-33p[
+
+Heteroskedasticity: 
+  - **White test** (White 1980)
+  - **BP test** (Breusch and Pagan 1979)
+
+]
+
+.pull-right-33p[
+
+Non-normality: 
+  - **SW test** (Shapiro and Wilk 1965)
+  - **Jarque–Bera test** (Jarque and Bera 1980)
+
+]
 
@@
 
-# 🔍Regression diagnostics🔍
+# 🔍Regression Diagnostics🔍
 
@@
 - While such rejections may be **statistically correct**, their **sensitivity** may render the results **impractical**.
-**Minor defects** in the model are **unlikely** to have a **significant impact** and may be best disregarded for practical purposes.
+**Minor defects** in the model are **unlikely** to have a **material impact** and may be best disregarded for practical purposes.
 
@@
 .center[
-# 📜Liteature of regression diagnostics📜
+# 📜Liteature of Regression Diagnostics📜
 ]
 
-### Plotting the data is the recommended method by experts for diagnosing model fits.
+### Grahpical approaches (plots) are the recommended methods for diagnosing model fits.
 
@@
 .caption[
-Fig. 3: _Residual plot of a simple linear regression._
+Fig. 3: _Residual plot of a simple linear regression (more meaningful)._
 ]
@@
 
-# 🤔False visual discoveries🤔
+# 🤔False Visual Discoveries🤔
 
@@
 
-# Visual inference
+# 🔬Visual Inference🔬
 
-If we embed the residual plot in a matrix of **null plots**, we will find the visual discovery is
+<br>
+<br>
+<br>
+
+If we embed the residual plot in a matrix of **null plots**, we will find that it is **not uncommon** for residual plots to exhibit a triangle shape.
+
+- The visual pattern is **over-interpreted**!
+
+<br>
+<br>
+
+### The visual discovery is calibrated via comparison.
+
+]
+
+.pull-right-center[
+
+```{r}
+set.seed(10131)
+mod$gen_lineup(300, k = 20, pos = 8) %>%
+  filter(null != FALSE) %>%
+  bind_rows(ori_dat %>% mutate(k = 8, null = FALSE)) %>%
+  VI_MODEL$plot_lineup(theme = theme_light(base_size = 12),
+                       remove_grid_line = TRUE,
+                       remove_axis = TRUE)
+```
+.center[
+.caption[
+Fig. 4: _A lineup of 20 residual plots of a simple linear regression. (Plot No.8 is Fig.3) Can you find the most different one?_
+]
+]
+
+]
+
+---
+
+.pull-left[
+
+# 🔬Visual Inference🔬
+
+<br>
+<br>
+
+**Visual inference** was introduced by Buja, et al. (2009) as an inferential framework to extend confirmatory statistics to visual discoveries.
+
+A **lineup** consists of $m$ randomly placed plots, where one plot is the **data plot** and the remaining $m − 1$ plots (**null plots**) contain data consistent with the null hypothesis.
+
+To perform a **visual test**, observer(s) will be asked to select the **most different plot(s)** from the lineup.
 

### 2023-04-18 Milestone_presentation_2/Milestone_presentation_2/Untitled.Rmd
rearrange the page
restructure according to feedback from other students

@@
       countIncrementalSlides: false
-      slideNumberFormat: "%current%"
+      slideNumberFormat: "<%current%>"
       highlightStyle: github
@@
 .center[
+
+# 🔍Regression Diagnostics🔍
+
+]
+
+--
+
+**Diagnostics** are the key to determining whether there is anything **importantly wrong** with a model.
+
+--
+
+**Residuals** summarise what is **not captured by the regression model**.
+
+--
+
+**Residual plots** are commonly used to diagnose **non-linearity** and **heteroskedasticity**.
+**Non-normality** is usually harder to detect from a residual plot.
+A favourable graphical summary for this task is the **quantile-quantile plot**.
+
+--
+
+.image-bottom-0[
+
+```{r fig.height = 2.5, fig.width = 10}
+set.seed(10086)
+
+# Base model
+mod <- poly_model(include_z = FALSE, sigma = 0.25)
+
+# Base data
+dat <- mod$gen(300)
+
+bind_rows(
+  dat %>%
+    mutate(type = "(A) Good residuals"),
+  mod$set_prm("include_z", TRUE)$
+    gen(300, computed = select(dat, x, e)) %>%
+    mutate(type = "(B) Non-linearity"),
+  heter_model(b = 64)$
+    gen(300, computed = select(dat, x, e)) %>%
+    mutate(type = "(C) Heteroskedasticity"),
+  heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
+    gen(300, computed = select(dat, x)) %>%
+    mutate(type = "(D) Non-normality")
+) %>%
+  mod$plot(remove_grid_line = TRUE, theme = theme_light(base_size = 12)) +
+  facet_wrap(~type, ncol = 4, scales = "free") +
+  theme(axis.text = element_blank(),
+        axis.ticks = element_blank()) +
+  xlab("Fitted values") +
+  ylab("Residuals")
+```
+
+]
+
+.center[
+.caption[
+Fig.
+1: _Example residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0._
+]
+]
+
+---
+
+.center[
 # 🔍Regression Diagnostics🔍
@@
 
-**Diagnostics** are the key to determining whether there is anything **importantly wrong** with a model.
+--
 
-**Residuals** summarise what is **not captured by the regression model**. Many different residual-based hypothesis tests are available to detect specific model defects.
+Many different **residual-based hypothesis tests** are available to detect specific model defects.
 
-<br>
-<br>
-<br>
+--
 
-.pull-right-33p[
+.pull-left-33p[
 
 Non-linearity:
+
 - **F-test**
+
 - **RESET test** (Ramsey 1969)
+
 ]
 
-.pull-right-33p[
+--
+
+.pull-left-33p[
 
 Heteroskedasticity:
+
 - **White test** (White 1980)
+
 - **BP test** (Breusch and Pagan 1979)
@@
 
-.pull-right-33p[
+--
+
+.pull-left-33p[
 
 Non-normality:
+
 - **SW test** (Shapiro and Wilk 1965)
+
 - **Jarque–Bera test** (Jarque and Bera 1980)
+
+]
+
+--
+
+.float-left[
+But there are **drawbacks** of using them.
+
+- Most residual-based tests for particular types of departure are also sensitive to **other types of departures**, resulting in a phenomenon known as the "**Type III error**" (Cook and Weisberg 1982).
+
+- Other problems such as **outliers** can trigger the rejection (Cook and Weisberg 1999).
 
@@
 .caption[
-Fig. 1: _One classically good looking residual plot and three residual plots with model defects._
+Fig.
+1: _Example residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0._
 ]
@@
 # 🔍Regression Diagnostics🔍
-
-<br>
-
-**Diagnostics** are the key to determining whether there is anything **importantly wrong** with a model.
-
-**Residuals** summarise what is **not captured by the regression model**.
-
-**Residual plots** are commonly used to diagnose **non-linearity** and **heteroskedasticity**.
-**Non-normality** is usually harder to detect from a residual plot.
-A favourable graphical summary for this task is the **quantile-quantile plot**.
-
-]
-
----
-
-.pull-left-center[
-
-.image-bottom-0[
-
-```{r}
-set.seed(10086)
-
-# Base model
-mod <- poly_model(include_z = FALSE, sigma = 0.25)
-
-# Base data
-dat <- mod$gen(300)
-
-bind_rows(
-  dat %>%
-    mutate(type = "(A) Good residuals"),
-  mod$set_prm("include_z", TRUE)$
-    gen(300, computed = select(dat, x, e)) %>%
-    mutate(type = "(B) Non-linearity"),
-  heter_model(b = 64)$
-    gen(300, computed = select(dat, x, e)) %>%
-    mutate(type = "(C) Heteroskedasticity"),
-  heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
-    gen(300, computed = select(dat, x)) %>%
-    mutate(type = "(D) Non-normality")
-) %>%
-  mod$plot(remove_grid_line = TRUE, theme = theme_light(base_size = 12)) +
-  facet_wrap(~type, ncol = 2, scales = "free") +
-  theme(axis.text = element_blank()) +
-  xlab("Fitted values") +
-  ylab("Residuals")
-```
-
-]
-
-.center[
-.caption[
-Fig. 1: _One classically good looking residual plot and three residual plots with model defects._
-]
-]
-
-]
-
-.pull-right[
-
-# 🔍Regression Diagnostics🔍
-
-<br>
-
-Many different hypothesis tests are available to detect specific model defects.
-
-Non-linearity: 
-  - **F-test**
-  - **RESET test** (Ramsey 1969)
-
-Heteroskedasticity: 
-  - **White test** (White 1980)
-  - **BP test** (Breusch and Pagan 1979)
-
-Non-normality: 
-  - **SW test** (Shapiro and Wilk 1965)
-  - **Jarque–Bera test** (Jarque and Bera 1980)
-
-]
-
----
-
-.pull-left-center[
-
-.image-bottom-0[
-
-```{r}
-set.seed(10086)
-
-# Base model
-mod <- poly_model(include_z = FALSE, sigma = 0.25)
-
-# Base data
-dat <- mod$gen(300)
-
-bind_rows(
-  dat %>%
-    mutate(type = "(A) Good residuals"),
-  mod$set_prm("include_z", TRUE)$
-    gen(300, computed = select(dat, x, e)) %>%
-    mutate(type = "(B) Non-linearity"),
-  heter_model(b = 64)$
-    gen(300, computed = select(dat, x, e)) %>%
-    mutate(type = "(C) Heteroskedasticity"),
-  heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
-    gen(300, computed = select(dat, x)) %>%
-    mutate(type = "(D) Non-normality")
-) %>%
-  mod$plot(remove_grid_line = TRUE, theme = theme_light(base_size = 12)) +
-  facet_wrap(~type, ncol = 2, scales = "free") +
-  theme(axis.text = element_blank()) +
-  xlab("Fitted values") +
-  ylab("Residuals")
-```
-
-]
-
-.center[
-.caption[
-Fig. 1: _One classically good looking residual plot and three residual plots with model defects._
-]
-]
-
-]
-
-.pull-right[
-
-# 🔍Regression Diagnostics🔍
-
-BP and SW tests reject the residual plots exhibiting structure that **they weren’t designed for** (**type III error**).
-
-Most residual-based tests for particular types of departure are also sensitive to **other types of departures** (Cook and Weisberg 1982).
 

### 2023-04-18 Milestone_presentation_2/Milestone_presentation_2/Untitled.Rmd
finish the intro

@@
 
-.pull-left-center[
+.pull-left-center-80[
 
@@
 set.seed(10086)
-
 # Base model
 mod <- poly_model(include_z = FALSE, sigma = 0.25)
-
 # Base data
 dat <- mod$gen(300)
-
 bind_rows(
@@
 
-.pull-right[
+.pull-right-120[
 
@@
 
+<br>
+
+BP and SW tests reject the residual plots exhibiting structure that **they weren’t designed for** (**type III error**).
+
+<br>
+
 ```{r}
 set.seed(10086)
-
 # Base model
 mod <- poly_model(include_z = FALSE, sigma = 0.25)
-
 # Base data
 dat0 <- mod$gen(300)
-
 # Replicate data in Figure 1
@@
                gen(300, computed = select(dat0, x))
-
 dat_list <- list(dat0, dat1, dat2, dat3)
-
 table_dat <- data.frame(plot = c("A", "B", "C", "D"), 
@@
   # Italic if reject
-  mutate(across(r:s, ~ kableExtra::cell_spec(.x, italic = as.numeric(.x) <= 0.05)))
-
+  mutate(across(r:s, ~ kableExtra::cell_spec(.x, italic = as.numeric(.x) <= 0.05, color = ifelse(as.numeric(.x) <= 0.05, "red", "black"))))
   table_dat %>%
@@
 .caption[
-Fig. 2: _One classically good looking residual plot but rejected by the RESET test with $p\text{-value} = `r format(round(this_p_value$p_value, digits = 3), nsmall = 3)`$._
+Fig.
+2: _One classically good looking residual plot but rejected by the RESET test with $p\text{-value} = `r format(round(this_p_value$p_value, digits = 3), nsmall = 3)`$.
+The model is misspecified but the departure from the model assumption is small._
 ]
@@
 
-- While such rejections may be **statistically correct**, their **sensitivity** may render the results **impractical**.
+<!-- - While such rejections may be **statistically correct**, their **sensitivity** may render the results **impractical**.
 **Minor defects** in the model are **unlikely** to have a **material impact** and may be best disregarded for practical purposes.
+-->
 
@@
 .caption[
-Fig. 3: _Residual plot of a simple linear regression (more meaningful)._
+Fig. 3: _Residual plot of a correctly specified model exhibiting a heteroskedasticity pattern (triangle shape)._
 ]
@@
 
-**Unsecured** and **unconfirmed** visual discoveries will lead to **over or under-interpretations of the data**.
+**Unconfirmed** visual discoveries could result in **over or under-interpretations of the data**.
 
@@
 <br>
-<br>
 
-If we embed the residual plot in a matrix of **null plots**, we will find that it is **not uncommon** for residual plots to exhibit a triangle shape.
+If we embed the residual plot in **a matrix of plots containing residuals simulated from the fitted model**, we will find that it is **not uncommon** for residual plots to exhibit a triangle shape.
 
-- The visual pattern is **over-interpreted**!
+- Can you find the **most different plot**?
 
-<br>
-<br>
+- Is it the residual plot presented at the previous page?
+
+- If it is not, then the visual pattern is **over-interpreted**!
 
@@
 
-**Visual inference** was introduced by Buja, et al. (2009) as an inferential framework to extend confirmatory statistics to visual discoveries.
+This procedure is called **visual inference**. It was introduced by Buja, et al. (2009) as an inferential framework to extend confirmatory statistics to visual discoveries.
 
@@
 ]
-
----
-
-# Visual inference
-
-What is visual inference?
-
-Underline that it is a valid inferential framework. Highlights the difference from reading a residual plot.
 

### 2023-04-18 Milestone_presentation_2/Milestone_presentation_2/Untitled.Rmd
simulate residuals

@@
 
+.center[
+# 🎲Simulate Residuals from the Assumed Model🎲
+]
+
+Data used in the $m − 1$ null plots needs to be simulated.
+
+For classical normal linear regression model, the **residual rotation** technique (Buja, et al. 2009) can be applied:
+
+<br>
+
+<p> 1. Generate independent random draws \(w_i, i=1,...,N\) from \(N(0, 1)\). </p>
+
+<br>
+
+<p> 2. Regress \(w\) on \(x\) and obtain the residuals \(r_i, i=1,...,N\). </p>
+
+<br>
+
+<p> 3.
+Rescale \(r_i\) by \(\sqrt{RSS_{old}/RSS_{new}}\) to obtain the simulated residuals, where \(RSS_{old}\) is the residual sum of square of the original regression, and \(RSS_{new}\) is the residual sum of square of the regression fitted on step 2.
+</p>
+
+---
+
 # Integrate visual tests into regression diagnostics

### 2023-04-18 Milestone_presentation_2/Milestone_presentation_2/Untitled.Rmd
finish p-value

@@
 
+<br>
+
+--
+
 Data used in the $m − 1$ null plots needs to be simulated.
+
+--
 
@@
 
-<p> 1. Generate independent random draws \(w_i, i=1,...,N\) from \(N(0, 1)\). </p>
+--
+
+<p> 1. Generate independent random draws \(w_i, i=1,...,n\) from \(N(0, 1)\). </p>
+
+--
 
@@
 
-<p> 2. Regress \(w\) on \(x\) and obtain the residuals \(r_i, i=1,...,N\). </p>
+<p> 2. Regress \(w\) on \(x\) and obtain the residuals \(r_i, i=1,...,n\). </p>
+
+--
 
@@
 </p>
+
+---
+
+.center[
+# 📈Statistical Significance📈
+]
+
+<br>
+<br>
+
+--
+
+Under the null hypothesis, it is expected that the data plot would have **no distinguishable difference** with the null plots.
+
+<br>
+
+--
+
+Suppose observers are allowed to **select only one plot**. The probability of the observer correctly picks the actual data plot is $1/m$.
+
+<br>
+
+--
+
+However, if we involve $K$ **independent observers** in a visual test, and show them the **same lineup**, there will be possible **dependencies** in the visual test due to **repeated evaluations** of the same lineup.
+
+---
+
+.pull-left[
+
+```{r}
+alpha_sample <- rgamma(10, 1, 1)
+data.frame(x = 0, y = 0, k = 1:20, alpha = alpha_sample/sum(alpha_sample)) %>%
+ggplot() +
+  geom_text(aes(x, y, label = paste0("theta[", k, "]~'='~", round(alpha, 2))), parse = TRUE, size = 5) +
+  facet_wrap(~k) +
+  theme_light(base_size = 12) +
+  theme(panel.grid = element_blank(),
+        axis.line = element_blank(),
+        axis.title = element_blank(),
+        axis.text = element_blank(),
+        axis.ticks = element_blank())
+```
+
+.center[
+.caption[
+Fig. 5: _Example of the probability of each plot being selected \\(\theta_i\\) under the null hypothesis with parameter \\(\alpha = 1\\)._
+]
+]
+
+]
+
+.pull-right[
+
+# 📈Statistical Significance📈
+
+<br>
+
+VanderPlas et al. (2021) addresses this by modelling the probability of a plot $i$ being selected from a lineup as $\theta_i$:
+
+$$\theta_i \sim Dirichlet(\alpha) \text{ for } i=1,...,m \text{ and } \alpha > 0.$$
+
+- This assumes the **attractiveness distribution** of each plot.
+
+For **small** $\alpha$:
+- Only **a few plots** are attractive and tend to be selected.
+
+For **large** $\alpha$:
+- The distribution of the probability of each plot being selected is more **even**.
+
+]
+
+---
+
+.pull-left[
+
+```{r}
+data.frame(k = 1:20, 
+           ci = c(1, 0, 0.25, 0, 0, 0.2, 0.45, 0, 5.95, 0, 0, 0, 0, 1.7, 0, 0.2, 0, 1, 0, 0.25)) %>%
+ggplot() +
+  geom_text(aes(0, 0, label = paste0("c[", k, "]~'='~", ci)), parse = TRUE, size = 5) +
+  facet_wrap(~k) +
+  theme_light(base_size = 12) +
+  theme(panel.grid = element_blank(),
+        axis.line = element_blank(),
+        axis.title = element_blank(),
+        axis.text = element_blank(),
+        axis.ticks = element_blank())
+```
+
+.center[
+.caption[
+Fig. 5: _Example of the number of times a plot being selected \\(c_i\\). The sum of all \\(c_i\\) is 11 because the lineup is evaluated by 11 subjects._
+]
+]
+
+]
+
+.pull-right[
+
+# 📈Statistical Significance📈
+
+<br>
+<br>
+
+The **number of times** plot $i$ being selected in $K$ evaluations is denoted as $c_i$.
+
+- In case subject $j$ makes **multiple selections**, $1/s_j$ will be added to $c_i$ instead of one, where $s_j$ is the **number of plots subject $j$ selected** for $j=1,...K$. 
+- This ensures $\sum_{i}c_i=K$.
+
+]
+
+---
+
+.center[
+# 📈Statistical Significance📈
+]
+
+<br>
+
+Since we are only interested in the **selections of the data plot $i$**, the **marginal model** can be simplified to a **beta-binomial model**. VanderPlas et al. (2021) derives the visual $p$-value as
+
+$$P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\text{ for } c_i \in \mathbb{Z}_0^+$$
+
+where $B(.)$ is the beta function defined as
+
+$$B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\text{ where } a,b>0.$$
+We extend the equation to **non-negative real number $c_i$** by applying a linear approximation
+
+$$
+P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \text{ for } c_i \in \mathbb{R}_0^+.
+$$
 

### 2023-04-18 Milestone_presentation_2/Milestone_presentation_2/Untitled.Rmd
finish models

@@
 
+--
+
 Since we are only interested in the **selections of the data plot $i$**, the **marginal model** can be simplified to a **beta-binomial model**. VanderPlas et al. (2021) derives the visual $p$-value as
@@
 $$B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\text{ where } a,b>0.$$
+
+--
+
 We extend the equation to **non-negative real number $c_i$** by applying a linear approximation
@@
 $$
+
+---
+
+.center[
+# 💪Power of Visual Tests💪
+]
+
+--
+
+<br>
+
+The power of a visual test may depend on the **ability of the particular subject** (Majumder, Hofmann, and Cook 2013).
+
+--
+
+- We assume the **individual skill has negligible effect** on the variation of the power to **simplify the model structure**, thereby **obviate a costly large-scale experiment** to estimate complex covariance matrices.
+
+<br>
+
+--
+
+We use the **logistic regression** to estimate the power:
+
+$$Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right),$$
+
+where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
+
+- The **effect size $E$** is the only predictor.
+- The intercept is fixed to $log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$.
+
+---
+
+.center[
+# 🧪Experimental Design🧪
+]
+
+<br>
+<br>
+
+--
+
+An experiment is conducted to investigate the difference between **conventional hypothesis testing** and **visual testing** in the application of **linear regression diagnostics**.
+
+--
+
+<br>
+
+We focus on two types of residual departures - **Non-linearity** and **Heteroskedasticity**.
+
+--
+
+<br>
+
+Overall, we collected **7974 evaluations** on **1152 unique lineups** performed by **443 subjects** throughout three data collection periods.
+
+---
+
+.center[
+# 🧪Experimental Design🧪
+]
+
+**Non-linearity model**:
+
+$$\begin{aligned} \boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\ \boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\ \boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\ \boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)), \end{aligned}$$
+where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are length $n$ vectors, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
+
+$$g(\boldsymbol{x}, k) = (\boldsymbol{x} - min(\boldsymbol{x}))/max(\boldsymbol{x} - min(\boldsymbol{x}))2k - k, \text{ for }  k > 0. $$
+
+The null regression model used to fit the realizations generated by the above model is formulated as
+
+$$\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},$$
+
+where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.
+
+---
+
+.center[
+# 🧪Experimental Design🧪
+]
+
+**Heteroskedasticity model**:
+
+$$\begin{aligned} \label{eq:heter-model} \boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\ \boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\ \boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}, 1 + (2 - |a|)(\boldsymbol{x} - a)^2b \boldsymbol{I}), \end{aligned}$$
+
+where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{\varepsilon}$ are length $n$ vectors and $g(.)$ is the scaling function defined in the previous slide.
+
+The null regression model is also provided in the previous slide.
+
+---
+
+```{r}
+magick::image_read_pdf("images/different-shape-of-herimite-1.pdf", pages = 1)
+magick::image_read_pdf("images/different-sigma-1.pdf", pages = 1)
+```
+
+.center[
+.caption[
+Fig. 5: _Non-linearity forms generated for the residual plots are controlled by the parameter \\(j\\). And the strength of the signal of the non-linearity model is controlled by \\(\sigma\\)._
+]
+]
+
+---
+
+.image-center[
+
+```{r out.width = "70%", fig.height = 2.67, fig.width = 8}
+set.seed(10086)
+
+stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
+
+a_labels <- c("left-triangle (a = -1)", "butterfly (a = 0)", "right-triangle (a = 1)")
+
+# Generate data for a = -1
+dat_a_n1 <- heter_model(a = -1, 
+                        x = {
+                          raw_x <- rand_uniform(-1, 1);
+                          closed_form(~stand_dist(raw_x))
+                          },
+                        b = 128)$gen(300) %>%
+  mutate(a = a_labels[1])
+
+# Generate data for other a
+map(c(0, 1), function(a) {
+  heter_model(a = a,
+              x = {
+                raw_x <- rand_uniform(-1, 1); 
+                closed_form(~stand_dist(raw_x))
+                }, 
+              b = 128)$
+    gen(300, computed = select(dat_a_n1, x, e)) %>%
+    mutate(a = a_labels[a + 2])
+}) %>%
+
+  # Combined with data for a = -1
+  bind_rows(dat_a_n1) %>%
+  mutate(a = factor(a, levels = a_labels)) %>%
+  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
+  facet_wrap(~a, scales = "free", ncol = 3) +
+  xlab("Fitted values") +
+  ylab("Residuals")
+```
+
+]
+
+```{r}
+magick::image_read_pdf("images/different-b-1.pdf", pages = 1)
+```
+
+.center[
+.caption[
+Fig. 5: _Heteroskedasticity forms generated for the residual plots are controlled by the parameter \\(a\\). And the strength of the signal of the heteroskedasticity model is controlled by \\(b\\)._
+]
+]
 

### 2023-04-18 Milestone_presentation_2/Milestone_presentation_2/Untitled.Rmd
finish effect size

@@
 
+```{r}
+magick::image_read_pdf("images/different-dist-1.pdf", pages = 1)
+```
+
+.image-center[
+
+```{r out.width = "70%"}
+magick::image_read_pdf("images/different-n-1.pdf", pages = 1)
+```
+
+]
+
+.center[
+.caption[
+Fig. 5: _Four different distribution of fitted values that might affect perception of residual plots are used. Three differnet values of \\(n\\) are used to control the signal strength._
+]
+]
+
+---
+
+.center[
+# 📏Effect size📏
+]
+
+We have chosen to use an approach based on **Kullback-Leibler divergence** (Kullback and Leibler 1951).
+
+The effect size of the **non-linearity model** is
+
+$$E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}\sigma^2))^{-1}\boldsymbol{\mu}_z\right),$$
+
+where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,
+$\boldsymbol{R} = \boldsymbol{I}_n - \boldsymbol{H}$ is the residual operator, $\boldsymbol{H} = \boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ is the hat matrix, $\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients, and $\sigma^2\boldsymbol{I}$ is the assumed covariance matrix of the error term when $H_0$ is true.
+
+And the effect size of the **heteroskedasticity model** is
+
+$$E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|diag(\boldsymbol{R})|} - n + tr(diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}diag(\boldsymbol{R}))\right),$$
+where $\boldsymbol{V}$ is the actual covariance matrix of the error term.
+
+---
+
 # Integrate visual tests into regression diagnostics

### 2023-04-18 Milestone_presentation_2/Milestone_presentation_2/Untitled.Rmd
finish power cmp

@@
 
-.image-bottom-0[
+.image-top-bottom-0[
 
@@
 
+--
+
 We have chosen to use an approach based on **Kullback-Leibler divergence** (Kullback and Leibler 1951).
+
+--
 
@@
 
+--
+
 And the effect size of the **heteroskedasticity model** is
@@
 
+.center[
+# 🛠️Experimental Setup🛠️
+]
+
+We recruited 443 subjects from an crowd-sourcing platform called **Prolific** (Palan and Schitter 2018) throughout **three data collection periods**. 
+- 160 subjects for period I
+- 160 subjects for period II
+- 123 subjects for period III
+
+--
+
+During the experiment, every subject is presented with **a block of 20 lineups**.
+
+--
+
+The subject is asked to:
+- Select **one or more** plots that are **most different** from others.
+- Provide **a reason** for their selections.
+- Evaluate **how different** they think the selected plots are from others.
+
+--
+
+If there is **no noticeable difference** between plots in a lineup, subjects are permitted to **select zero plots** without providing the reason.
+
+--
+
+A subject’s submission is only accepted if the data plot is identified for **at least one attention check**.
+
+---
+
+![](images/lineup1.png)
+
+.center[
+.caption[
+Fig. 5: _Screenshot of the study website._
+]
+]
+
+---
+
+class: center, middle
+
+# ⚖️Results - Power Comparison of the Tests⚖️
+
+---
+
+.image-top-bottom-0[
+
+.image-center[
+
+```{r out.width = "80%"}
+knitr::include_graphics("images/polypower-1.png")
+```
+
+]
+
+]
+
+.center[
+.caption[
+Fig.
+5: _Comparison of power between different tests for non-linear patterns (uniform fitted values only).
+The horizontal lines of dots represent non-reject and reject results from visual tests for each lineup.
+The visual test has multiple power curves estimated from bootstrap samples.
+The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot._
+]
+]
+
+---
+
+.image-top-bottom-0[
+
+.image-center[
+
+```{r out.width = "80%"}
+knitr::include_graphics("images/heterpower-1.png")
+```
+
+]
+
+]
+
+.center[
+.caption[
+Fig.
+5: _Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only).
+The horizontal lines of dots represent non-reject and reject results from visual tests for each lineup.
+The visual test has multiple power curves estimated from bootstrap samples.
+The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot._
+]
+]
+
+---
+
 # Integrate visual tests into regression diagnostics

### 2023-04-18 Milestone_presentation_2/Milestone_presentation_2/Milestone_presentation_2.Rmd
finish the intro
transfer file

--- 2022-03-16 Milestone_presentation_1/Milestone_presentation_1/Milestone_presentation_1.Rmd
@@
 ---
-title: "Automatic Visual Inference for Linear Regression"
+title: "Advances in Artificial Intelligence for Data Visualization: Automated Reading of Residual Plots with Computer Vision"
 subtitle: ""  
@@
     nature:
+      countdown: 55000
       beforeInit: "scripts/my_macro.js"
       countIncrementalSlides: false
-      slideNumberFormat: "%current%/%total%"
+      slideNumberFormat: "<%current%>"
       highlightStyle: github
@@
 .pull-left-full[
-<h1 class="myblue"> Automatic Visual Inference for Linear Regression </h1>
-
-<h3 class="myblue"> First Milestone Presentation </h3>
-
-<br>
-<br>
-<br>
-<br>
-
-<h3 class="myblue"> Weihao Li </h3>
-
-<h3 class="myblue"> Supervised by Di Cook and Emi Tanaka </h3>
+<h2 class="myblue"> Advances in Artificial Intelligence for Data Visualization: Automated Reading of Residual Plots with Computer Vision </h2>
+
+<h3 class="myblue"> Progress Review Presentation </h3>
+
+<br>
+<br>
+
+<h4 class="myblue"> Weihao (Patrick) Li </h3>
+
+<h4 class="myblue"> Supervised by Di Cook, Emi Tanaka and Susan VanderPlas </h3>
 
@@
 
-![](images/159.png)
-]
-
----
-
-.pull-left-center[
-```{r}
-library(visage)
-
-eval_data <- readRDS("data/training_original_data_single.rds")
-
-filter(eval_data, id == 9586, null == TRUE) %>%
-  VI_MODEL$plot(theme = theme_light(base_size = 12), size = 1) +
-  ylab("Residuals") +
-  xlab("Fitted values")
-# 68
-
-```
-
-.center[
-.caption[
-Fig. 1: *Residual plot of a simple linear regression.*
-]
-]
-]
-
-.pull-right[
-# Visual Discoveries
-
-<br>
-<br>
-
-#### Can you find the evidence on heteroskedasticity?
-
-]
-
----
-
-count: false
-
-.pull-left-center[
-```{r}
-filter(eval_data, id == 9586, null == TRUE) %>%
-  VI_MODEL$plot(theme = theme_light(base_size = 12), size = 1) +
-  ylab("Residuals") +
-  xlab("Fitted values")
-# 68
-
-```
-
-.center[
-.caption[
-Fig. 1: *Residual plot of a simple linear regression.*
-]
-]
-]
-
-.pull-right[
-# Visual Discoveries
-
-<br>
-<br>
-
-#### Can you find the evidence on heteroskedasticity?
-
-- Vertical spread of the points varies with the fitted values.
-
-]
-
----
-
-count: false
-
-.pull-left-center[
-```{r}
-filter(eval_data, id == 9586, null == TRUE) %>%
-  VI_MODEL$plot(theme = theme_light(base_size = 12), size = 1) +
-  ylab("Residuals") +
-  xlab("Fitted values")
-# 68
-
-```
-
-.center[
-.caption[
-Fig. 1: *Residual plot of a simple linear regression.*
-]
-]
-]
-
-.pull-right[
-# Visual Discoveries
-
-<br>
-<br>
-
-#### Can you find the evidence on heteroskedasticity?
-
-- Vertical spread of the points varies with the fitted values.
-
-#### However, the residuals are actually obtained from a **correctly specified classical normal linear model**!
-
-]
-
----
-
-count: false
-
-.pull-left-center[
-```{r}
-filter(eval_data, id == 9586, null == TRUE) %>%
-  VI_MODEL$plot(theme = theme_light(base_size = 12), size = 1) +
-  ylab("Residuals") +
-  xlab("Fitted values")
-# 68
-
-```
-
-.center[
-.caption[
-Fig. 1: *Residual plot of a simple linear regression.*
-]
-]
-]
-
-.pull-right[
-# Visual Discoveries
-
-<br>
-<br>
-
-#### Can you find the evidence on heteroskedasticity?
-
-- Vertical spread of the points varies with the fitted values.
-
-#### However, the residuals are actually obtained from a **correctly specified classical normal linear model**!
-
-#### Unsecured and unconfirmed visual discoveries will lead to **over or under-interpretations of the data**.
-
-]
-
----
-
-.pull-left[
-# Visual Inference & Lineup Protocol
-
-<br>
-<br>
-
-#### **Visual inference** was introduced by `r BibOptions(max.names = 1);Citet(myBib, "buja_statistical_2009")` as an inferential framework to extend confirmatory statistics to visual discoveries.
-
-]
-
-.pull-right-center[
-```{r}
-mod <- HETER_MODEL$fit(filter(eval_data, id == 9586, null == FALSE))
-
-manual_lineup <- data.frame()
-
-for (i in 1:19) {
-  tmp_dat <- HETER_MODEL$null_resid(filter(eval_data, id == 9586, null == FALSE), mod)
-  tmp_dat$k <- i
-  manual_lineup <- bind_rows(manual_lineup, tmp_dat)
-}
-
-manual_lineup <- bind_rows(manual_lineup, mutate(filter(eval_data, id == 9586, null == TRUE), k = 20))
-
-VI_MODEL$plot_lineup(manual_lineup, 
-                     theme = theme_light(),
-                     remove_grid_line = TRUE,
-                     size = 0.3
-                     ) +
+![](images/505.png)
+]
+
+---
+
+.center[
+# 📚Thesis Structure📚
+]
+
+<br>
+
+#### 1. **Exploring the application of visual inference in regression diagnostics and comparing it with conventional hypothesis tests.**
+
+#### 2. Designing an automated visual inference system to assess lineups of residual plots of classical normal linear regression model.
+
+#### 3. Deploying the automatic visual inference system as an online application and publishing the relevant open-source software.
+
+<br>
+<br>
+
+This presentation will be focused on the **first project**.
+
+---
+
+class: center, middle
+
+# A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol
+
+#### Weihao Li, Dianne Cook, Emi Tanaka and Susan VanderPlas (2023)
+
+---
+
+.center[
+
+# 🔍Regression Diagnostics🔍
+
+]
+
+--
+
+**Diagnostics** are the key to determining whether there is anything **importantly wrong** with a model.
+
+--
+
+**Residuals** summarise what is **not captured by the regression model**.
+
+--
+
+**Residual plots** are commonly used to diagnose **non-linearity** and **heteroskedasticity**.
+**Non-normality** is usually harder to detect from a residual plot.
+A favourable graphical summary for this task is the **quantile-quantile plot**.
+
+--
+
+.image-top-bottom-0[
+
+```{r fig.height = 2.5, fig.width = 10}
+set.seed(10086)
+
+# Base model
+mod <- poly_model(include_z = FALSE, sigma = 0.25)
+
+# Base data
+dat <- mod$gen(300)
+
+bind_rows(
+  dat %>%
+    mutate(type = "(A) Good residuals"),
+  mod$set_prm("include_z", TRUE)$
+    gen(300, computed = select(dat, x, e)) %>%
+    mutate(type = "(B) Non-linearity"),
+  heter_model(b = 64)$
+    gen(300, computed = select(dat, x, e)) %>%
+    mutate(type = "(C) Heteroskedasticity"),
+  heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
+    gen(300, computed = select(dat, x)) %>%
+    mutate(type = "(D) Non-normality")
+) %>%
+  mod$plot(remove_grid_line = TRUE, theme = theme_light(base_size = 12)) +
+  facet_wrap(~type, ncol = 4, scales = "free") +
+  theme(axis.text = element_blank(),
+        axis.ticks = element_blank()) +
   xlab("Fitted values") +
@@
 
-.center[
-.caption[
-Fig. 2: _A lineup of 20 residual plots of a simple linear regression. (Plot No.20 is Fig.1) **Can you find the most different one?**_ 
-]
-]
-
-]
-
----
-
-count: false
-
-.pull-left[
-# Visual Inference & Lineup Protocol
-
-<br>
-<br>
-
-#### **Visual inference** was introduced by `r BibOptions(max.names = 1);Citet(myBib, "buja_statistical_2009")` as an inferential framework to extend confirmatory statistics to visual discoveries.
-
-#### A **lineup** consists of $m$ randomly placed plots, where one plot is the **actual data plot** and the remaining $m-1$ plots (**null plots**) contain data consistent with the null hypothesis.
-
-]
-
-.pull-right-center[
-```{r}
-VI_MODEL$plot_lineup(manual_lineup, 
-                     theme = theme_light(),
-                     remove_grid_line = TRUE,
-                     size = 0.3
-                     ) +
+]
+
+.center[
+.caption[
+Fig.
+1: _Example residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0._
+]
+]
+
+---
+
+.center[
+# 🔍Regression Diagnostics🔍
+]
+
+--
+
+Many different **residual-based hypothesis tests** are available to detect specific model defects.
+
+--
+
+.pull-left-33p[
+
+Non-linearity:
+
+- **F-test**
+
+- **RESET test** (Ramsey 1969)
+
+]
+
+--
+
+.pull-left-33p[
+
+Heteroskedasticity:
+
+- **White test** (White 1980)
+
+- **BP test** (Breusch and Pagan 1979)
+
+]
+
+--
+
+.pull-left-33p[
+
+Non-normality:
+
+- **SW test** (Shapiro and Wilk 1965)
+
+- **Jarque–Bera test** (Jarque and Bera 1980)
+
+]
+
+--
+
+.float-left[
+But there are **drawbacks** of using them.
+
+- Most residual-based tests for particular types of departure are also sensitive to **other types of departures**, resulting in a phenomenon known as the "**Type III error**" (Cook and Weisberg 1982).
+
+- Other problems such as **outliers** can trigger the rejection (Cook and Weisberg 1999).
+
+]
+
+---
+
+.pull-left-center-80[
+
+.image-top-bottom-0[
+
+```{r}
+set.seed(10086)
+# Base model
+mod <- poly_model(include_z = FALSE, sigma = 0.25)
+# Base data
+dat <- mod$gen(300)
+bind_rows(
+  dat %>%
+    mutate(type = "(A) Good residuals"),
+  mod$set_prm("include_z", TRUE)$
+    gen(300, computed = select(dat, x, e)) %>%
+    mutate(type = "(B) Non-linearity"),
+  heter_model(b = 64)$
+    gen(300, computed = select(dat, x, e)) %>%
+    mutate(type = "(C) Heteroskedasticity"),
+  heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
+    gen(300, computed = select(dat, x)) %>%
+    mutate(type = "(D) Non-normality")
+) %>%
+  mod$plot(remove_grid_line = TRUE, theme = theme_light(base_size = 12)) +
+  facet_wrap(~type, ncol = 2, scales = "free") +
+  theme(axis.text = element_blank()) +
   xlab("Fitted values") +
@@
 
-.center[
-.caption[
-Fig. 2: _A lineup of 20 residual plots of a simple linear regression. (Plot No.20 is Fig.1) **Can you find the most different one?**_ 
-]
-]
-
-]
-
----
-
-count: false
-
-.pull-left[
-# Visual Inference & Lineup Protocol
-
-<br>
-<br>
-
-#### **Visual inference** was introduced by `r BibOptions(max.names = 1);Citet(myBib, "buja_statistical_2009")` as an inferential framework to extend confirmatory statistics to visual discoveries.
-
-#### A **lineup** consists of $m$ randomly placed plots, where one plot is the **actual data plot** and the remaining $m-1$ plots (**null plots**) contain data consistent with the null hypothesis.
-
-#### To perform a **visual test**, observer(s) will be asked to select the **most different plot(s)** from the lineup.
-
-]
+]
+
+.center[
+.caption[
+Fig.
+1: _Example residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0._
+]
+]
+
+]
+
+.pull-right-120[
+
+# 🔍Regression Diagnostics🔍
+
+<br>
+
+BP and SW tests reject the residual plots exhibiting structure that **they weren’t designed for** (**type III error**).
+
+<br>
+
+```{r}
+set.seed(10086)
+# Base model
+mod <- poly_model(include_z = FALSE, sigma = 0.25)
+# Base data
+dat0 <- mod$gen(300)
+# Replicate data in Figure 1
+dat1 <- mod$set_prm("include_z", TRUE)$gen(300, computed = select(dat0, x, e))
+dat2 <- heter_model(b = 64)$gen(300, computed = select(dat0, x, e))
+dat3 <- heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
+               gen(300, computed = select(dat0, x))
+dat_list <- list(dat0, dat1, dat2, dat3)
+table_dat <- data.frame(plot = c("A", "B", "C", "D"), 
+                        model = c("None", 
+                                  "Non-linearity", 
+                                  "Heteroskedasticity", 
+                                  "Non-normality"),
+                        r = map_dbl(dat_list, 
+                                    ~POLY_MODEL$test(.x, 
+                                                     test = "RESET", 
+                                                     power = 2:4)$p_value),
+                        b = map_dbl(dat_list, 
+                                    ~HETER_MODEL$test(.x)$p_value),
+
+                        # SW test has not been included by visage
+                        s = map_dbl(dat_list, 
+                                    ~shapiro.test(.x$.resid)$p.value)) %>%
+
+  # 3 decimal points
+  mutate(across(r:s, ~ format(round(.x, digits = 3), nsmall = 3))) %>%
+
+  # Italic if reject
+  mutate(across(r:s, ~ kableExtra::cell_spec(.x, italic = as.numeric(.x) <= 0.05, color = ifelse(as.numeric(.x) <= 0.05, "red", "black"))))
+  table_dat %>%
+  knitr::kable(col.names = c("Plot", 
+                             "Departures", 
+                             "RESET", 
+                             "BP", 
+                             "SW"),
+               align = "llrrr",
+               escape = FALSE,
+               format = "html",
+               booktabs = TRUE)
+```
+
+.center[
+.caption[
+Tab. 1: _Statistical significance testing for departures from good residuals. Shown are the \\(p\\)-values calculated for the RESET, the BP and the SW tests._
+]
+]
+
+]
+
+---
 
 .pull-right-center[
-```{r}
-VI_MODEL$plot_lineup(manual_lineup, 
-                     theme = theme_light(),
-                     remove_grid_line = TRUE,
-                     size = 0.3
-                     ) +
+
+.image-top-bottom-0[
+
+```{r}
+set.seed(10086)
+
+# Base model
+mod <- poly_model(sigma = 8)
+dat <- mod$gen(1000)
+this_p_value <- mod$test(dat, test = "RESET", power = 2:4)
+dat %>%
+mod$plot(remove_grid_line = TRUE, theme = theme_light(base_size = 12)) +
+  theme(axis.text = element_blank()) +
   xlab("Fitted values") +
@@
 
-.center[
-.caption[
-Fig. 2: _A lineup of 20 residual plots of a simple linear regression. (Plot No.20 is Fig.1) **Can you find the most different one?**_ 
-]
-]
-
-]
-
----
-
-# $p$-value of Lineup Protocol
-
-#### Under the null hypothesis, it is expected that the actual data plot would have **no distinguishable difference** with the null plots.
-
----
-
-count: false
-
-# $p$-value of Lineup Protocol
-
-#### Under the null hypothesis, it is expected that the actual data plot would have **no distinguishable difference** with the null plots.
-
-#### Suppose observers are allowed to select **only one plot**. The probability of the observer correctly picks the actual data plot is $1/m$.
-
----
-
-count: false
-
-# $p$-value of Lineup Protocol
-
-#### Under the null hypothesis, it is expected that the actual data plot would have **no distinguishable difference** with the null plots.
-
-#### Suppose observers are allowed to select **only one plot**. The probability of the observer correctly picks the actual data plot is $1/m$.
-
-#### If we involve $K$ independent observers in a visual test, and let $X$ be a random variable denoting the number of observers correctly picking the actual data plot.
-Under the null hypothesis $X ∼ Binom(K, 1/m)$.
-
----
-
-count: false
-
-# $p$-value of Lineup Protocol
-
-#### Under the null hypothesis, it is expected that the actual data plot would have **no distinguishable difference** with the null plots.
-
-#### Suppose observers are allowed to select **only one plot**. The probability of the observer correctly picks the actual data plot is $1/m$.
-
-#### If we involve $K$ independent observers in a visual test, and let $X$ be a random variable denoting the number of observers correctly picking the actual data plot.
-Under the null hypothesis $X ∼ Binom(K, 1/m)$.
-
-#### The p-value is given as $$P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i}.$$
-
----
-
-count: false
-
-# $p$-value of Lineup Protocol
-
-#### Under the null hypothesis, it is expected that the actual data plot would have **no distinguishable difference** with the null plots.
-
-#### Suppose observers are allowed to select **only one plot**. The probability of the observer correctly picks the actual data plot is $1/m$.
-
-#### If we involve $K$ independent observers in a visual test, and let $X$ be a random variable denoting the number of observers correctly picking the actual data plot.
-Under the null hypothesis $X ∼ Binom(K, 1/m)$.
-
-#### The p-value is given as $$P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i}.$$
-
-#### The null hypothesis will be rejected if $P(X \geq x) < \alpha$, the desired significance level.
-
----
-
-.pull-left-center-80[
-
-![:scale 80%](images/Powerloom_weaving_in_1835.jpg)
-
-![:scale 80%](images/Maquina_vapor_Watt_ETSIIM.jpg)
-
-.caption[
-Fig. 3: _A Roberts loom in a weaving shed in 1835 and a Watt steam engine._ 
-]
-
-]
-
-.pull-right-120[
-# Limitation of Lineup Protocol
-
-<br>
-
-#### 1. Human can not evaluate lineup consisted of a large number of plots.
-#### 2. Human can not evaluate a large number of lineups.
-#### 3. Lineup protocol is unfriendly to vision-impaired people.
-#### 4. Evaluation of lineup is high in labour cost and time consuming.
-
-<br>
-
-]
-
----
-
-count: false
-
-.pull-left-center-80[
-
-![:scale 80%](images/Powerloom_weaving_in_1835.jpg)
-
-![:scale 80%](images/Maquina_vapor_Watt_ETSIIM.jpg)
-
-.caption[
-Fig. 3: _A Roberts loom in a weaving shed in 1835 and a Watt steam engine._ 
-]
-
-]
-
-.pull-right-120[
-# Limitation of Lineup Protocol
-
-<br>
-
-#### 1. Human can not evaluate lineup consisted of a large number of plots.
-#### 2. Human can not evaluate a large number of lineups.
-#### 3. Lineup protocol is unfriendly to vision-impaired people.
-#### 4. Evaluation of lineup is high in labour cost and time consuming.
-
-<br>
-
-### **We need a "steam engine" for visual test!**
-
-]
-
----
-
-# Computer Vision Model
-
-#### Large-scale evaluation of visual tests is not possible without the use of technology and machines.
-
-#### Modern **computer vision model** could be a promising solution to this problem.
-It is usually built on a **deep neural network** called **convolutional neural network** (CNN) `r BibOptions(max.names = 1);Citep(myBib, "fukushima_neocognitron_1982")`.
-
-![](images/Typical_cnn.png)
-
-.caption[
-Fig. 4: _A typical CNN with convolutional layers, pooling layers and fully-connected layers._ 
-]
-
----
-
-exclude: true
-
-# Computer Vision Model - Convolutional Layer
-
-#### **Convolutional layers** take advantages of the hierarchical pattern in data and provide regularized versions of fully-connected layers.
-It downscales and transforms the image by summarising information in a small space.
-
-.center[![:scale 70%](images/conlayer.png)]
-
-.center[.caption[
-Fig. 5: _Illustration of convolution operation with two filters._ 
-]]
-
----
-
-exclude: true
-
-# Computer Vision Model - Pooling Layer
+]
+
+.center[
+.caption[
+Fig.
+2: _One classically good looking residual plot but rejected by the RESET test with $p\text{-value} = `r format(round(this_p_value$p_value, digits = 3), nsmall = 3)`$.
+The model is misspecified but the departure from the model assumption is small._
+]
+]
+
+]
 
@@
 
-<br>
-<br>
-
-#### **Pooling layers** apply non-linear down-sampling on the feature maps. This helps reducing the number of parameters in the neural network so that the problem of overfiting can be controlled.
-
-#### One of the most popular pooling operations is **max pooling**. It divides the feature map into several rectangle regions and takes the maximum to form a new feature map.
+# 🔍Regression Diagnostics🔍
+
+<br>
+<br>
+
+With **large sample sizes**, hypothesis tests may reject a residual plot when there is only a **slight departure** (Belsley, Kuh, and Welsch 1980).
+
+<!-- - While such rejections may be **statistically correct**, their **sensitivity** may render the results **impractical**.
+**Minor defects** in the model are **unlikely** to have a **material impact** and may be best disregarded for practical purposes.
+-->
+
+Residual plot diagnostics:
+  - **Identify** potential issues that could lead to **incorrect conclusions or errors in subsequent analyses**.
+
+]
+
+---
+
+.center[
+# 📜Liteature of Regression Diagnostics📜
+]
+
+### Grahpical approaches (plots) are the recommended methods for diagnosing model fits.
+
+- Draper and Smith (1998) and Belsley, Kuh, and Welsch (1980):
+> Residual plots are usually revealing when the assumptions are violated.
+
+- Cook and Weisberg (1982):
+> Formal tests and graphical procedures are complementary and both have a place in residual analysis, but graphical methods are easier to use.
+
+- Montgomery and Peck (1982):
+> Residual plots are more informative in most practical situations than the corresponding conventional hypothesis tests.
+
+---
+
+.pull-left-center[
+
+.image-top-bottom-0[
+
+```{r}
+set.seed(10131)
+mod <- heter_model(b = 0, x = rand_lognormal())
+ori_dat <- mod$gen(300)
+
+ori_dat %>%
+  VI_MODEL$plot(theme = theme_light(base_size = 12)) +
+  xlab("Fitted values") +
+  ylab("Residuals")
+```
+
+]
+
+.center[
+.caption[
+Fig. 3: _Residual plot of a correctly specified model exhibiting a heteroskedasticity pattern (triangle shape)._
+]
+]
+
+]
+
+.pull-right[
+
+# 🤔False Visual Discoveries🤔
+
+<br>
+<br>
+
+Can you find the evidence on heteroskedasticity?
+
+- Vertical spread of the points varies with the fitted values.
+
+However, the residuals are actually obtained from a **correctly specified classical normal linear model**!
+
+**Unconfirmed** visual discoveries could result in **over or under-interpretations of the data**.
+
+]
+
+---
+
+.pull-left[
+
+# 🔬Visual Inference🔬
+
+<br>
+<br>
+
+If we embed the residual plot in **a matrix of plots containing residuals simulated from the fitted model**, we will find that it is **not uncommon** for residual plots to exhibit a triangle shape.
+
+- Can you find the **most different plot**?
+
+- Is it the residual plot presented at the previous page?
+
+- If it is not, then the visual pattern is **over-interpreted**!
+
+### The visual discovery is calibrated via comparison.
+
 ]
@@
 .pull-right-center[
-![](images/Max_pooling.png)
-
-.caption[
-Fig. 6: _Illustration of max pooling operation._ 
-]
-]
-
----
-
-exclude: true
-
-# Computer Vision Model - Fully-connected Layer
-
-.pull-left-center-80[
-![:scale 50%](images/fnn.png)
-
-.caption[
-Fig. 7: _Illustration of a fully-connected layer._ 
-]
-]
-
-<br>
-<br>
-
-#### **Fully-connected layer** is building block of neural network. It peformes matrix multiplication on the inputs and the weights.
-
-<br>
-
-#### This layer is primarily used to process information presented in the feature maps before making the prediction.
-The output of the layer will be passed to a special layer called **loss layer** for computing the loss or error of the prediction.
-
----
-
-# Simulation Setup - Cubic Model
-
-#### We simulated data from two linear regression model with **non-linearity** and **heteroskedasticity** issues
-
---
-
-#### The first model is a **cubic model** which can be expressed by: $$\boldsymbol{Y}= 1 + (2-c)\boldsymbol{X} + c\boldsymbol{Z} + a[(2-c)\boldsymbol{X}]^2+a(c\boldsymbol{Z})^2+b[(2-c)\boldsymbol{X}]^3+b(c\boldsymbol{Z})^3+\boldsymbol{\varepsilon},$$ where $c \in (0,2)$, $a \in (-3,3)$, $b \in (-3,3)$, $\boldsymbol{\varepsilon}\overset{iid}{\sim} N(\boldsymbol{0},\sigma^2\boldsymbol{I})$, $\boldsymbol{Y}$, $\boldsymbol{X}$ and $\boldsymbol{Z}$ are $n\times1$ matrices.
-
---
+
+```{r}
+set.seed(10131)
+mod$gen_lineup(300, k = 20, pos = 8) %>%
+  filter(null != FALSE) %>%
+  bind_rows(ori_dat %>% mutate(k = 8, null = FALSE)) %>%
+  VI_MODEL$plot_lineup(theme = theme_light(base_size = 12),
+                       remove_grid_line = TRUE,
+                       remove_axis = TRUE)
+```
+.center[
+.caption[
+Fig. 4: _A lineup of 20 residual plots of a simple linear regression. (Plot No.8 is Fig.3) Can you find the most different one?_
+]
+]
+
+]
+
+---
 
 .pull-left[
-#### $X_i$, $i=1,...,n$ follows one of the four distributions:
-- $N(0, 0.09)$
-- $U(-1, 1)$
-- $Lognormal(0,0.36)/3 - 1$
-- $-Lognormal(0,0.36)/3 + 1$
-
-#### Most the realizations will fall between $-1$ and $1$.
-]
-
---
+
+# 🔬Visual Inference🔬
+
+<br>
+<br>
+
+This procedure is called **visual inference**. It was introduced by Buja, et al. (2009) as an inferential framework to extend confirmatory statistics to visual discoveries.
+
+A **lineup** consists of $m$ randomly placed plots, where one plot is the **data plot** and the remaining $m − 1$ plots (**null plots**) contain data consistent with the null hypothesis.
+
+To perform a **visual test**, observer(s) will be asked to select the **most different plot(s)** from the lineup.
+
+]
+
+.pull-right-center[
+
+```{r}
+set.seed(10131)
+mod$gen_lineup(300, k = 20, pos = 8) %>%
+  filter(null != FALSE) %>%
+  bind_rows(ori_dat %>% mutate(k = 8, null = FALSE)) %>%
+  VI_MODEL$plot_lineup(theme = theme_light(base_size = 12),
+                       remove_grid_line = TRUE,
+                       remove_axis = TRUE)
+```
+.center[
+.caption[
+Fig. 4: _A lineup of 20 residual plots of a simple linear regression. (Plot No.8 is Fig.3) Can you find the most different one?_
+]
+]
+
+]
+
+---
+
+.center[
+# 🎲Simulate Residuals from the Assumed Model🎲
+]
+
+<br>
+
+--
+
+Data used in the $m − 1$ null plots needs to be simulated.
+
+--
+
+For classical normal linear regression model, the **residual rotation** technique (Buja, et al. 2009) can be applied:
+
+<br>
+
+--
+
+<p> 1. Generate independent random draws \(w_i, i=1,...,n\) from \(N(0, 1)\). </p>
+
+--
+
+<br>
+
+<p> 2. Regress \(w\) on \(x\) and obtain the residuals \(r_i, i=1,...,n\). </p>
+
+--
+
+<br>
+
+<p> 3.
+Rescale \(r_i\) by \(\sqrt{RSS_{old}/RSS_{new}}\) to obtain the simulated residuals, where \(RSS_{old}\) is the residual sum of square of the original regression, and \(RSS_{new}\) is the residual sum of square of the regression fitted on step 2.
+</p>
+
+---
+
+.center[
+# 📈Statistical Significance📈
+]
+
+<br>
+<br>
+
+--
+
+Under the null hypothesis, it is expected that the data plot would have **no distinguishable difference** with the null plots.
+
+<br>
+
+--
+
+Suppose observers are allowed to **select only one plot**. The probability of the observer correctly picks the actual data plot is $1/m$.
+
+<br>
+
+--
+
+However, if we involve $K$ **independent observers** in a visual test, and show them the **same lineup**, there will be possible **dependencies** in the visual test due to **repeated evaluations** of the same lineup.
+
+---
+
+.pull-left[
+
+```{r}
+alpha_sample <- rgamma(10, 1, 1)
+data.frame(x = 0, y = 0, k = 1:20, alpha = alpha_sample/sum(alpha_sample)) %>%
+ggplot() +
+  geom_text(aes(x, y, label = paste0("theta[", k, "]~'='~", round(alpha, 2))), parse = TRUE, size = 5) +
+  facet_wrap(~k) +
+  theme_light(base_size = 12) +
+  theme(panel.grid = element_blank(),
+        axis.line = element_blank(),
+        axis.title = element_blank(),
+        axis.text = element_blank(),
+        axis.ticks = element_blank())
+```
+
+.center[
+.caption[
+Fig. 5: _Example of the probability of each plot being selected \\(\theta_i\\) under the null hypothesis with parameter \\(\alpha = 1\\)._
+]
+]
+
+]
 
 .pull-right[
-#### $Z_i$, $i=1,...,n$ follows one of the two distributions:
-- $U(-1, 1)$
-- Discrete uniform distribution with $z_n$ outcomes ranged from $-1$ to $1$
-
-#### $z_n$ is sampled from $\{z_n: 9<z_n<21 \text{ and } z_n \in \mathbb{Z} \}$.
-]
-
-<!-- #### The distribution of $Z_i$, $i=1,...,n$, has $50$% chance to be a uniform distribution ranged from $-1$ to $1$, and $50$% chance to be a discrete uniform distribution with $z_n$ outcomes simulated from a uniform distribution ranged from $-1$ to $1$.
-$z_n$ itself is a discrete uniform distribution with outcomes $\{x: 9<x<21 \text{ and } x \in \mathbb{Z} \}$, which defines the number of possible values $Z_i$ could take.
--->
-
----
-
-# Simulation Setup - Cubic Model
-
-#### The simulated data is fitted by OLS with equation: $$\boldsymbol{Y}= \beta_0 + \beta_1\boldsymbol{X} + \beta_2\boldsymbol{Z} +\boldsymbol{u},$$ where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2_u\boldsymbol{I})$.
-
---
-
-#### Let $\boldsymbol{X}_a=[\boldsymbol{1},\boldsymbol{X},\boldsymbol{Z}]$ denotes the set of regressors in matrix form.
-Then, the **residuals** obtained from the null model are $$\boldsymbol{e} \sim N(\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b, \sigma^2\boldsymbol{R}_a),$$ where $\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$, $\boldsymbol{X}_b=[\boldsymbol{X}^2,\boldsymbol{Z}^2,\boldsymbol{X}^3,\boldsymbol{Z}^3]$ and $\boldsymbol{\beta}_b=(a(2-c)^2,ac^2,b(2-c)^3,bc^3)'$.
-
---
-
-#### Under the null hypothesis, $\boldsymbol{e} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{R}_a)$.
-The **effect size** is derived as: $$ES=n^{-1}||\sigma^{-1}\boldsymbol{R}_a^{-\frac{1}{2}}\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b||^2=n^{-1}\sigma^{-2}||\boldsymbol{R}_a^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2\approx n^{-1}\sigma^{-2}||diag(\boldsymbol{R}_a)^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2,$$ where $diag(\boldsymbol{R}_a)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$.
+
+# 📈Statistical Significance📈
+
+<br>
+
+VanderPlas et al. (2021) addresses this by modelling the probability of a plot $i$ being selected from a lineup as $\theta_i$:
+
+$$\theta_i \sim Dirichlet(\alpha) \text{ for } i=1,...,m \text{ and } \alpha > 0.$$
+
+- This assumes the **attractiveness distribution** of each plot.
+
+For **small** $\alpha$:
+- Only **a few plots** are attractive and tend to be selected.
+
+For **large** $\alpha$:
+- The distribution of the probability of each plot being selected is more **even**.
+
+]
 
@@
 .pull-left[
-```{r}
-mod <- cubic_model(a = 5, b = 10, c = 1)
-mod$plot_lineup(mod$gen_lineup(300), remove_grid_line = TRUE, theme = theme_light()) + xlab("Fitted values") + ylab("Residuals")
-```
-
-.caption[
-Fig. 5: _A lineup of residual plots generated with the cubic model._ 
-]
+
+```{r}
+data.frame(k = 1:20, 
+           ci = c(1, 0, 0.25, 0, 0, 0.2, 0.45, 0, 5.95, 0, 0, 0, 0, 1.7, 0, 0.2, 0, 1, 0, 0.25)) %>%
+ggplot() +
+  geom_text(aes(0, 0, label = paste0("c[", k, "]~'='~", ci)), parse = TRUE, size = 5) +
+  facet_wrap(~k) +
+  theme_light(base_size = 12) +
+  theme(panel.grid = element_blank(),
+        axis.line = element_blank(),
+        axis.title = element_blank(),
+        axis.text = element_blank(),
+        axis.ticks = element_blank())
+```
+
+.center[
+.caption[
+Fig. 5: _Example of the number of times a plot being selected \\(c_i\\). The sum of all \\(c_i\\) is 11 because the lineup is evaluated by 11 subjects._
+]
+]
+
 ]
@@
 .pull-right[
-```{r}
-mod <- cubic_model(-1, -1, 1, 0.5)
-dat <- mod$gen(500, fit_model = TRUE)
-mod$plot(dat, theme = theme_light(base_size = 10), size = 1) +
-  ggtitle(glue::glue("Effect size: {round(mod$effect_size(dat), 2)}")) + xlab("Fitted values") + ylab("Residuals") -> p1
-
-mod <- cubic_model(-2, -1, 1, 0.5)
-dat <- mod$gen(500, fit_model = TRUE)
-mod$plot(dat, theme = theme_light(base_size = 10), size = 1) +
-  ggtitle(glue::glue("Effect size: {round(mod$effect_size(dat), 2)}")) + xlab("Fitted values") + ylab("Residuals") -> p2
-
-mod <- cubic_model(-5, -5, 1, 0.5)
-dat <- mod$gen(500, fit_model = TRUE)
-mod$plot(dat, theme = theme_light(base_size = 10), size = 1) +
-  ggtitle(glue::glue("Effect size: {round(mod$effect_size(dat), 2)}")) + xlab("Fitted values") + ylab("Residuals") -> p3
-
-mod <- cubic_model(-20, -20, 1, 0.5)
-dat <- mod$gen(500, fit_model = TRUE)
-mod$plot(dat, theme = theme_light(base_size = 10), size = 1) +
-  ggtitle(glue::glue("Effect size: {round(mod$effect_size(dat), 2)}")) + xlab("Fitted values") + ylab("Residuals") -> p4
-
-list(p1, p2, p3, p4) %>% patchwork::wrap_plots(ncol=2)
-```
-
-.caption[
-Fig. 6: _Residual plots of cubic model under different effect sizes._ 
-]
-]
-
----
-
-# Simulation Setup - Heteroskedasticity Model
-
-<br>
-
-#### The second model is a heteroskedasticy model which can be expressed by: $$Y_i = 1 + X_i + \varepsilon_i, ~i = 1,...,n,$$ where $a \in \{-1,0,1\}$, $b\in (0,32)$ and $\varepsilon_i \overset{iid}{\sim} N(0,1+b(2-|a|)(X_i-a)^2)$.
-
---
-
-#### $X_i$, $i=1,...,n$ follows one of the five distributions:
-- $N(0, 0.09)$
-- $U(-1, 1)$
-- $Lognormal(0,0.36)/3 - 1$
-- $-Lognormal(0,0.36)/3 + 1$
-- Discrete uniform distribution
-
----
-
-# Simulation Setup - Heteroskedasticity Model
-
-#### The simulated data is fitted by OLS with equation: $$\boldsymbol{Y}= \beta_0 + \beta_1\boldsymbol{X} +\boldsymbol{u},$$ where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2_u\boldsymbol{I})$.
-
---
-
-#### Let $\boldsymbol{X}_a=[\boldsymbol{1},\boldsymbol{X}]$ denotes the set of regressors in matrix form.
-The residuals obtained from the null model are $$\boldsymbol{e} \sim N(\boldsymbol{0}, \boldsymbol{R}_a\boldsymbol{V}),$$ where $\boldsymbol{R}_a=\boldsymbol{I}-\boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ and $\boldsymbol{V}$ is a diagonal matrix with $V_{ii}=1+b(2-|a|)(X_i  -  a)^2$, $i = 1,...,n$.
-
---
-
-#### The **effect size** is derived as $ES = b\sqrt{n}$.
-
----
-
-.pull-left[
-```{r}
-mod <- heter_model(a = -1, b = 30)
-mod$plot_lineup(mod$gen_lineup(300), remove_grid_line = TRUE, theme = theme_light()) + xlab("Fitted values") + ylab("Residuals")
-```
-
-.caption[
-Fig. 7: _A lineup of residual plots generated with the heteroskedasticity model._ 
-]
-]
-
-.pull-right[
-
-```{r}
+
+# 📈Statistical Significance📈
+
+<br>
+<br>
+
+The **number of times** plot $i$ being selected in $K$ evaluations is denoted as $c_i$.
+
+- In case subject $j$ makes **multiple selections**, $1/s_j$ will be added to $c_i$ instead of one, where $s_j$ is the **number of plots subject $j$ selected** for $j=1,...K$. 
+- This ensures $\sum_{i}c_i=K$.
+
+]
+
+---
+
+.center[
+# 📈Statistical Significance📈
+]
+
+<br>
+
+--
+
+Since we are only interested in the **selections of the data plot $i$**, the **marginal model** can be simplified to a **beta-binomial model**. VanderPlas et al. (2021) derives the visual $p$-value as
+
+$$P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\text{ for } c_i \in \mathbb{Z}_0^+$$
+
+where $B(.)$ is the beta function defined as
+
+$$B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\text{ where } a,b>0.$$
+
+--
+
+We extend the equation to **non-negative real number $c_i$** by applying a linear approximation
+
+$$
+P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \text{ for } c_i \in \mathbb{R}_0^+.
+$$
+
+---
+
+.center[
+# 💪Power of Visual Tests💪
+]
+
+--
+
+<br>
+
+The power of a visual test may depend on the **ability of the particular subject** (Majumder, Hofmann, and Cook 2013).
+
+--
+
+- We assume the **individual skill has negligible effect** on the variation of the power to **simplify the model structure**, thereby **obviate a costly large-scale experiment** to estimate complex covariance matrices.
+
+<br>
+
+--
+
+We use the **logistic regression** to estimate the power:
+
+$$Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right),$$
+
+where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
+
+- The **effect size $E$** is the only predictor.
+- The intercept is fixed to $log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$.
+
+---
+
+.center[
+# 🧪Experimental Design🧪
+]
+
+<br>
+<br>
+
+--
+
+An experiment is conducted to investigate the difference between **conventional hypothesis testing** and **visual testing** in the application of **linear regression diagnostics**.
+
+--
+
+<br>
+
+We focus on two types of residual departures - **Non-linearity** and **Heteroskedasticity**.
+
+--
+
+<br>
+
+Overall, we collected **7974 evaluations** on **1152 unique lineups** performed by **443 subjects** throughout three data collection periods.
+
+---
+
+.center[
+# 🧪Experimental Design🧪
+]
+
+**Non-linearity model**:
+
+$$\begin{aligned} \boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\ \boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\ \boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\ \boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)), \end{aligned}$$
+where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are length $n$ vectors, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
+
+$$g(\boldsymbol{x}, k) = (\boldsymbol{x} - min(\boldsymbol{x}))/max(\boldsymbol{x} - min(\boldsymbol{x}))2k - k, \text{ for }  k > 0. $$
+
+The null regression model used to fit the realizations generated by the above model is formulated as
+
+$$\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},$$
+
+where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.
+
+---
+
+.center[
+# 🧪Experimental Design🧪
+]
+
+**Heteroskedasticity model**:
+
+$$\begin{aligned} \label{eq:heter-model} \boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\ \boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\ \boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}, 1 + (2 - |a|)(\boldsymbol{x} - a)^2b \boldsymbol{I}), \end{aligned}$$
+
+where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{\varepsilon}$ are length $n$ vectors and $g(.)$ is the scaling function defined in the previous slide.
+
+The null regression model is also provided in the previous slide.
+
+---
+
+```{r}
+magick::image_read_pdf("images/different-shape-of-herimite-1.pdf", pages = 1)
+magick::image_read_pdf("images/different-sigma-1.pdf", pages = 1)
+```
+
+.center[
+.caption[
+Fig. 5: _Non-linearity forms generated for the residual plots are controlled by the parameter \\(j\\). And the strength of the signal of the non-linearity model is controlled by \\(\sigma\\)._
+]
+]
+
+---
+
+.image-center[
+
+```{r out.width = "70%", fig.height = 2.67, fig.width = 8}
 set.seed(10086)
-mod <- heter_model(0, 1)
-mod$plot(mod$gen(500, fit_model = TRUE), size = 1, theme = theme_light(base_size = 10)) + ggtitle(glue::glue("Effect size: {round(mod$effect_size(mod$gen(500)),2)}")) + xlab("Fitted values") + ylab("Residuals") -> p1
-mod <- heter_model(0, 3)
-mod$plot(mod$gen(500, fit_model = TRUE), size = 1, theme = theme_light(base_size = 10)) + ggtitle(glue::glue("Effect size: {round(mod$effect_size(mod$gen(500)),2)}")) + xlab("Fitted values") + ylab("Residuals") -> p2
-mod <- heter_model(0, 9)
-mod$plot(mod$gen(500, fit_model = TRUE), size = 1, theme = theme_light(base_size = 10)) + ggtitle(glue::glue("Effect size: {round(mod$effect_size(mod$gen(500)),2)}")) + xlab("Fitted values") + ylab("Residuals") -> p3
-mod <- heter_model(0, 81)
-mod$plot(mod$gen(500, fit_model = TRUE), size = 1, theme = theme_light(base_size = 10)) + ggtitle(glue::glue("Effect size: {round(mod$effect_size(mod$gen(500)),2)}")) + xlab("Fitted values") + ylab("Residuals") -> p4
-
-list(p1, p2, p3, p4) %>% patchwork::wrap_plots(ncol=2)
-```
-
-.caption[
-Fig. 8: _Residual plots of heteroskedasticity model under different effect sizes._ 
-]
-
-]
-
----
-
-# Human Subject Experiments
-
-.pull-left-center[
-
-<br>
-<br>
-
-![](images/website.png)
-
-.caption[
-Fig. 9: _Layout of the study webiste._ 
-]
-]
-
-.pull-right[
-
-#### **Two human subject experiments** are conducted to compare humans with computer vision models on the ability of reading data plots.
-
-]
-
----
-
-count: false
-
-# Human Subject Experiments
-
-.pull-left-center[
-
-<br>
-<br>
-
-![](images/website.png)
-
-.caption[
-Fig. 9: _Layout of the study webiste._ 
-]
-]
-
-.pull-right[
-
-#### **Two human subject experiments** are conducted to compare humans with computer vision models on the ability of reading data plots.
-
-#### **20** lineups are given to each participant, with **10** cubic lineups and **10** heteroskedasticity lineups across different effect sizes.
-
-]
-
----
-
-count: false
-
-# Human Subject Experiments
-
-.pull-left-center[
-
-<br>
-<br>
-
-![](images/website.png)
-
-.caption[
-Fig. 9: _Layout of the study webiste._ 
-]
-]
-
-.pull-right[
-
-#### **Two human subject experiments** are conducted to compare humans with computer vision models on the ability of reading data plots.
-
-#### **20** lineups are given to each participant, with **10** cubic lineups and **10** heteroskedasticity lineups across different effect sizes.
-
-#### **2** out of **20** lineups are extremely easy that are used as attention checks. Participants need to pass at least one to be paid.
-
-]
-
----
-
-count: false
-
-# Human Subject Experiments
-
-.pull-left-center[
-
-<br>
-<br>
-
-![](images/website.png)
-
-.caption[
-Fig. 9: _Layout of the study webiste._ 
-]
-]
-
-.pull-right[
-
-#### **Two human subject experiments** are conducted to compare humans with computer vision models on the ability of reading data plots.
-
-#### **20** lineups are given to each participant, with **10** cubic lineups and **10** heteroskedasticity lineups across different effect sizes.
-
-#### **2** out of **20** lineups are extremely easy that are used as attention checks. Participants need to pass at least one to be paid.
-
-#### The same lineup will not be presented to a participant twice. Most of the lineups have **5** replicates.
-
-]
-
----
-
-count: false
-
-# Human Subject Experiments
-
-.pull-left-center[
-
-<br>
-<br>
-
-![](images/website.png)
-
-.caption[
-Fig. 9: _Layout of the study webiste._ 
-]
-]
-
-.pull-right[
-
-#### **Two human subject experiments** are conducted to compare humans with computer vision models on the ability of reading data plots.
-
-#### **20** lineups are given to each participant, with **10** cubic lineups and **10** heteroskedasticity lineups across different effect sizes.
-
-#### **2** out of **20** lineups are extremely easy that are used as attention checks. Participants need to pass at least one to be paid.
-
-#### The same lineup will not be presented to a participant twice. Most of the lineups have **5** replicates.
-
-#### Participants are allowed to **select multiple plots** from a lineup. They also need to give the **reason** and the **confidence level**.
-
-]
-
----
-
-# $p$-value Estimation for Multiple Selections
-
-#### Let $K$ be the number of **independent evaluations** of a visual test, $s_i , i = 1, ..., K$ be the **number of selections of the evaluations**, and $X$ be the random variable denoting the number of detections.
-Then, the $p$-value of the visual test is given as: $$P(X\geq x) = \sum_{j=x}^{K}Pr(j|s_1,...,s_K).$$
-
---
-
-#### The distribution of $X$ given $s_1, ...s_K$ can not be **derived trivially**, as it is a **sampling without replacement problem**.
-
---
-
-#### In practice, this distribution can be **approximated** by computer simulation.
-
----
-
-# Power Estimation for Multiple Selections
-
-<br>
-
-#### Suppose there is a **visual test** $V_K$, where $K$ denoting the number of evaluations.
-
---
-
-<br>
-
-#### If one evaluation is **randomly deleted** from $V_K$, the **remaining evaluations** can still be used to form **another valid visual test** $V_{K-1}$.
-
---
-
-<br>
-
-#### Considering **all the possibilities**, $K$ different outcomes of $V_{K-1}$ can be obtained.
-
---
-
-<br>
-
-#### Since all outcomes occur with **equal probability**, the **proportion of outcomes which reject the null hypothesis** can be used as an **estimator of the power of the visual test** $V_{K-1}$.
-
----
-
-# Power Estimation for Multiple Selections
-
-<br>
-<br>
-
-#### Similarly, to estimate the power of the visual test $V_{K-j}$ given the evaluations of $V_{K}$, for $j < K$,
-
---
-
-<br>
-
-#### we could find all the possible combinations of $K$ elements, taken $k-j$ at a time to obtain $K\choose{K-j}$ different outcomes for $V_{K-j}$.
-
---
-
-<br>
-
-#### The estimated power is given as $R/{K\choose{K-j}}$, where $R$ is the number of outcomes which reject the null hypothesis.
-
----
-
-# Human Subject Experiment - Results
-
-<br>
-<br>
-
-#### We collect **400** lineup evaluations made by **20** participants in experiment I and **880** lineup evaluations made by **44** participants in experiment II.
-
---
-
-#### In total, **442** unique lineups are evaluated by **64** subjects.
-
---
-
-#### The power of each lineup from one to five evaluations is estimated.
-
---
-
-#### The estimated power is further used in fitting **quasi-binomial generalized linear models** with **natural logarithm of effect size** as the only regressor.
-
---
-
-#### All **decisions** made by $K\choose{K-j}$ different outcomes of $V_{K-j}$ are fitted by **binomial generalized linear models** with **natural logarithm of effect size** as the regressor.
-
----
-
-.center[![:scale 80%](images/power1.png)
-
+
+stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
+
+a_labels <- c("left-triangle (a = -1)", "butterfly (a = 0)", "right-triangle (a = 1)")
+
+# Generate data for a = -1
+dat_a_n1 <- heter_model(a = -1, 
+                        x = {
+                          raw_x <- rand_uniform(-1, 1);
+                          closed_form(~stand_dist(raw_x))
+                          },
+                        b = 128)$gen(300) %>%
+  mutate(a = a_labels[1])
+
+# Generate data for other a
+map(c(0, 1), function(a) {
+  heter_model(a = a,
+              x = {
+                raw_x <- rand_uniform(-1, 1); 
+                closed_form(~stand_dist(raw_x))
+                }, 
+              b = 128)$
+    gen(300, computed = select(dat_a_n1, x, e)) %>%
+    mutate(a = a_labels[a + 2])
+}) %>%
+
+  # Combined with data for a = -1
+  bind_rows(dat_a_n1) %>%
+  mutate(a = factor(a, levels = a_labels)) %>%
+  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
+  facet_wrap(~a, scales = "free", ncol = 3) +
+  xlab("Fitted values") +
+  ylab("Residuals")
+```
+
+]
+
+```{r}
+magick::image_read_pdf("images/different-b-1.pdf", pages = 1)
+```
+
+.center[
+.caption[
+Fig. 5: _Heteroskedasticity forms generated for the residual plots are controlled by the parameter \\(a\\). And the strength of the signal of the heteroskedasticity model is controlled by \\(b\\)._
+]
+]
+
+---
+
+```{r}
+magick::image_read_pdf("images/different-dist-1.pdf", pages = 1)
+```
+
+.image-center[
+
+```{r out.width = "70%"}
+magick::image_read_pdf("images/different-n-1.pdf", pages = 1)
+```
+
+]
+
+.center[
+.caption[
+Fig. 5: _Four different distribution of fitted values that might affect perception of residual plots are used. Three differnet values of \\(n\\) are used to control the signal strength._
+]
+]
+
+---
+
+.center[
+# 📏Effect size📏
+]
+
+--
+
+We have chosen to use an approach based on **Kullback-Leibler divergence** (Kullback and Leibler 1951).
+
+--
+
+The effect size of the **non-linearity model** is
+
+$$E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}\sigma^2))^{-1}\boldsymbol{\mu}_z\right),$$
+
+where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,
+$\boldsymbol{R} = \boldsymbol{I}_n - \boldsymbol{H}$ is the residual operator, $\boldsymbol{H} = \boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ is the hat matrix, $\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients, and $\sigma^2\boldsymbol{I}$ is the assumed covariance matrix of the error term when $H_0$ is true.
+
+--
+
+And the effect size of the **heteroskedasticity model** is
+
+$$E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|diag(\boldsymbol{R})|} - n + tr(diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}diag(\boldsymbol{R}))\right),$$
+where $\boldsymbol{V}$ is the actual covariance matrix of the error term.
+
+---
+
+.center[
+# 🛠️Experimental Setup🛠️
+]
+
+We recruited 443 subjects from an crowd-sourcing platform called **Prolific** (Palan and Schitter 2018) throughout **three data collection periods**. 
+- 160 subjects for period I
+- 160 subjects for period II
+- 123 subjects for period III
+
+--
+
+During the experiment, every subject is presented with **a block of 20 lineups**.
+
+--
+
+The subject is asked to:
+- Select **one or more** plots that are **most different** from others.
+- Provide **a reason** for their selections.
+- Evaluate **how different** they think the selected plots are from others.
+
+--
+
+If there is **no noticeable difference** between plots in a lineup, subjects are permitted to **select zero plots** without providing the reason.
+
+--
+
+A subject’s submission is only accepted if the data plot is identified for **at least one attention check**.
+
+---
+
+![](images/lineup1.png)
+
+.center[
+.caption[
+Fig. 5: _Screenshot of the study website._
+]
+]
+
+---
+
+class: center, middle
+
+# ⚖️Results - Power Comparison of the Tests⚖️
+
+---
+
+.image-top-bottom-0[
+
+.image-center[
+
+```{r out.width = "80%"}
+knitr::include_graphics("images/polypower-1.png")
+```
+
+]
+
+]
+
+.center[
 .caption[
 Fig.
-10: _Quasibinomial GLM fitted for estimated power for different number of evaluations.
-The plot on the left is for the cubic model, the plot on the right is for the heteroskedasticity model.
-Power curves for different number of evaluations are drawn in different colors.
-The dashed line is the power of the conventional test._
-]]
-
----
-
-.center[![:scale 80%](images/power2.png)
-
+5: _Comparison of power between different tests for non-linear patterns (uniform fitted values only).
+The horizontal lines of dots represent non-reject and reject results from visual tests for each lineup.
+The visual test has multiple power curves estimated from bootstrap samples.
+The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot._
+]
+]
+
+---
+
+.image-top-bottom-0[
+
+.image-center[
+
+```{r out.width = "80%"}
+knitr::include_graphics("images/heterpower-1.png")
+```
+
+]
+
+]
+
+.center[
 .caption[
 Fig.
-11: _Binomial GLM fitted for rejection for different number of evaluations.
-The plot on the left is for the cubic model, the plot on the right is for the heteroskedasticity model.
-Power curves for different number of evaluations are drawn in different colors.
-The dashed line is the power of the conventional test._
-]]
-
----
-
-# Human Subject Experiment - Results
-
-#### There is a huge gap between the conventional curves and the visual test curves when the natural logarithm of effect size is around $0$ for cubic model, and $2.5$ for heteroskedasticity model.
-
-#### Human is hard to identify the pattern at this level of difficulty.
-
-.pull-left[
-![](images/mid1.png)
-
-]
-
-.pull-right[
-![](images/mid2.png)
-
-]
-
-.center[.caption[
-Fig. 12: _A cubic lineup with effect size $= 0.18$ and a heteroskedasticity lineup with effect size $= 17.5$._ 
-]]
-
----
-
-# Automatic Visual Inference
-
-#### We propose a prototype **automatic visual inference system** for evaluating lineups of residual plots of classical normal linear regression.
-It currently only detects nonlinearity and heteroskedasticity.
-
-.center[
-![:scale 80%](images/vi_system.png)
-
-.caption[
-Fig. 13: _Diagram of automatic visual inference system._ 
-]]
-
----
-
-# Model Architecture and Model Training
-
-.pull-left-140[
-
-#### Due to limited computing resources, we apply **transfer learning** to train five computer vision models.
-
-#### Five **pre-trained models** are used, including **VGG16**, **VGG19**, **Xception**, **ResNet152V2** and **InceptionV3**.
-
-#### Models are trained on **10000 null plots** and **10000 actual data plots** simulated from the cubic model and heteroskedasticity model. 20% training data is used as validation set.
-
-#### **Image augmentation**:
-- **random rotation angle** ranged from $0$ to $5$ degrees
-- **random brightness** ranged from $95$% to $105$%
-
-#### Model training is performed with **Tensorflow** and **Keras**.
-
-]
-
-.pull-right-60[
-.center[
-
-![:scale 23%](images/vi_model.png)
-
-.caption[
-Fig. 14: _Diagram of the transfer learning computer vision model._ 
-]
-]
-]
-
----
-
-![](images/mod_per.png)
-
-.center[.caption[
-Fig. 15: _Model performance tested on 10000 null plots and 10000 actual data plots._ 
-]]
-
----
-
-.center[![:scale 80%](images/system_pow_1.png)
-
+5: _Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only).
+The horizontal lines of dots represent non-reject and reject results from visual tests for each lineup.
+The visual test has multiple power curves estimated from bootstrap samples.
+The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot._
+]
+]
+
+---
+
+.center[
+# ⚖️Results - Comparison of Test Decisions⚖️
+]
+
+.image-top-bottom-0[
+
+.image-center[
+
+```{r out.width = "70%"}
+magick::image_read_pdf("images/p-value-comparison-1.pdf", pages = 1)
+```
+
+]
+
+]
+
+.center[
 .caption[
 Fig.
-16: _Power of the automatic visual test compared to human visual test.
-The blue curve is the power of the automatic visual test.
-The dashed line represents the power of the conventional test.
-The plot on the left is for cubic model, and the plot on the right is for heteroskedasticity._
-]]
-
----
-
-.center[![:scale 80%](images/system_pow_2.png)
-
+5: _Rejection rate (\\(p\\)-value \\(\leq0.05\\)) of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot.
+The visual test rejects less frequently than the conventional test, and (almost) only rejects when the conventional test does.
+Surprisingly, one lineup in the heteroskedasticity group is rejected by the visual test but NOT the conventional test._
+]
+]
+
+---
+
+.center[
+# ⚖️Results - Effect of Amount of Non-linearity⚖️
+]
+
+.image-top-bottom-0[
+
+.image-center[
+
+```{r out.width = "80%"}
+magick::image_read_pdf("images/poly-power-uniform-j-1.pdf", pages = 1)
+```
+
+]
+
+]
+
+.center[
 .caption[
 Fig.
-17: _Power of the automatic visual test with different number of evaluations.
-The dashed line represents the power of the conventional test.
-The plot on the left is for cubic model, and the plot on the right is for heteroskedasticity._
-]]
-
----
-
-# Conclusion
-
----
-
-# Future work
-
----
-
-Thanks
+5: _The effect of the order of the polynomial on the power of conventional and visual tests.
+Deeper colour indicates higher order.
+The default RESET tests under-performs significantly in detecting the "triple-U" shape.
+To achieve a similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test, but this higher than the recommended value._
+]
+]
+
+---
+
+.center[
+# ⚖️Results - Effect of Shape of Heteroskedasticity⚖️
+]
+
+.image-top-bottom-0[
+
+.image-center[
+
+```{r out.width = "80%"}
+magick::image_read_pdf("images/heter-power-uniform-a-1.pdf", pages = 1)
+```
+
+]
+
+]
+
+.center[
+.caption[
+Fig.
+5: _The effect of heteroskedasticity shape (parameter \\(a\\)) on the power of conventional and visual tests.
+The butterfly has higher power in both tests.
+Curiously, the visual test has slightly higher power for the "left-triangle" than the "right-triangle" shape, when it would be expected that they should be identical, which is observed in conventional testing._
+]
+]
+
+---
+
+class: middle, center
+
+# ⚖️Results - Effect of Fitted Value Distributions⚖️
+
+---
+
+.image-top-bottom-0[
+
+.image-center[
+
+```{r out.width = "60%"}
+knitr::include_graphics("images/different-x-dist-poly-power-1.png")
+```
+
+]
+
+]
+
+.center[
+.caption[
+Fig.
+5: _Comparison of power on lineups with different fitted value distributions for conventional and visual tests for non-linearity and heteroskedasticity patterns.
+The power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and BP tests, respectively.
+Power curves of visual tests are estimated using five evaluations on each lineup.
+For lineups with a uniform fitted value distribution, the five evaluations are repeatedly sampled from the total eleven evaluations to give multiple power curves (grey).
+Surprisingly, the fitted value distribution has produces more variability in the power of conventional tests than visual tests.
+Uneven distributions, normal and lognormal distributions, tend to yield lower power._
+]
+]
+
+---
+
+.center[
+# 🧐Conclusion🧐
+]
+
+---
+
+.center[
+# 📅Timetable📅
+]
+
+```{r}
+data.frame(Date = c("April", 
+                    "May",
+                    "June",
+                    "July",
+                    "Sep",
+                    "Oct",
+                    "Nov",
+                    "Dec",
+                    "Mar",
+                    "Aug"), 
+           Event = c("Submit the abstract of the first paper to Australian Statistical Conference (ASC)", 
+                     "Submit the first paper",
+                     "Submit a poster for the first paper to Institute of Electrical and Electronics Engineers (IEEE) Visualization Conference",
+                     "Leave for a month",
+                     "Finalize the computer vision model",
+                     "Attend the IEEE Visualization Conference",
+                     "Start the web interface development",
+                     "Attend the ASC",
+                     "Submit the second paper",
+                     "Submit thesis")) %>%
+  kableExtra::kable(booktabs = TRUE) %>%
+  kableExtra::group_rows("2023", 1, 8) %>%
+  kableExtra::group_rows("2024", 9, 10)
+```
+
+---
+
+class: center, middle
+
+# Thanks!
+
+---
+
+.center[
+# 📖Bibliography📖
+]

### 2023-04-19 Milestone_presentation_2/Milestone_presentation_2/Milestone_presentation_2.Rmd
fix typo
fix counter and first page

@@
       countIncrementalSlides: false
-      slideNumberFormat: "<%current%>"
+      slideNumberFormat: "%current%/%total%"
       highlightStyle: github
@@
 
-#### 1. **Exploring the application of visual inference in regression diagnostics and comparing it with conventional hypothesis tests.**
+#### **1. Exploring the application of visual inference in regression diagnostics and comparing it with conventional hypothesis tests.**
 
@@
 
-### Grahpical approaches (plots) are the recommended methods for diagnosing model fits.
+### Graphical approaches (plots) are the recommended methods for diagnosing model fits.
 

### 2023-04-19 Milestone_presentation_2/Milestone_presentation_2/Milestone_presentation_2.Rmd
fix up to false findings

@@
 
-**Diagnostics** are the key to determining whether there is anything **importantly wrong** with a model.
-
---
-
-**Residuals** summarise what is **not captured by the regression model**.
+**Diagnostics** are the key to determining whether there is anything **importantly wrong** with a model. **Residuals** summarise what is **not captured by the regression model**.
 
@@
 --
+
+<br>
 
@@
 
-.center[
-.caption[
-Fig.
-1: _Example residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0._
-]
-]
-
 ---
@@
 .float-left[
-But there are **drawbacks** of using them.
+
+They are **very good at detecting violations of model assumptions**, but there are some **unexpected usages**.
 
@@
 
-.pull-left-center-80[
+.center[
+# 🔍Regression Diagnostics🔍
+]
+
+.pull-left-center[
 
@@
   facet_wrap(~type, ncol = 2, scales = "free") +
-  theme(axis.text = element_blank()) +
+  theme(axis.text = element_blank(),
+        axis.ticks = element_blank()) +
   xlab("Fitted values") +
@@
 
-.center[
-.caption[
-Fig.
-1: _Example residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0._
-]
-]
-
-]
-
-.pull-right-120[
-
-# 🔍Regression Diagnostics🔍
-
-<br>
-
-BP and SW tests reject the residual plots exhibiting structure that **they weren’t designed for** (**type III error**).
-
+]
+
+.pull-right[
+
+All the hypothesis tests **correctly reject** the correspoding model defects, and **correctly not reject** the structure in residual plot (A).
+
+BP and SW tests reject the residual plots exhibiting structure that **they weren’t designed for**.
+
+<br>
 <br>
@@
 
-.center[
-.caption[
-Tab. 1: _Statistical significance testing for departures from good residuals. Shown are the \\(p\\)-values calculated for the RESET, the BP and the SW tests._
-]
-]
-
-]
-
----
-
-.pull-right-center[
-
-.image-top-bottom-0[
-
-```{r}
-set.seed(10086)
-
-# Base model
-mod <- poly_model(sigma = 8)
-dat <- mod$gen(1000)
-this_p_value <- mod$test(dat, test = "RESET", power = 2:4)
-dat %>%
-mod$plot(remove_grid_line = TRUE, theme = theme_light(base_size = 12)) +
-  theme(axis.text = element_blank()) +
-  xlab("Fitted values") +
-  ylab("Residuals")
-```
-
-]
-
-.center[
-.caption[
-Fig.
-2: _One classically good looking residual plot but rejected by the RESET test with $p\text{-value} = `r format(round(this_p_value$p_value, digits = 3), nsmall = 3)`$.
-The model is misspecified but the departure from the model assumption is small._
-]
-]
-
-]
-
-.pull-left[
-
-# 🔍Regression Diagnostics🔍
-
-<br>
-<br>
-
-With **large sample sizes**, hypothesis tests may reject a residual plot when there is only a **slight departure** (Belsley, Kuh, and Welsch 1980).
-
-<!-- - While such rejections may be **statistically correct**, their **sensitivity** may render the results **impractical**.
-**Minor defects** in the model are **unlikely** to have a **material impact** and may be best disregarded for practical purposes.
--->
-
-Residual plot diagnostics:
-  - **Identify** potential issues that could lead to **incorrect conclusions or errors in subsequent analyses**.
-
 ]
@@
 ---
+
+.center[
+# 🤔False Visual Discoveries🤔
+]
 
@@
 
-.center[
-.caption[
-Fig. 3: _Residual plot of a correctly specified model exhibiting a heteroskedasticity pattern (triangle shape)._
-]
-]
-
 ]
@@
 
-# 🤔False Visual Discoveries🤔
-
-<br>
-<br>
-
-Can you find the evidence on heteroskedasticity?
+<br>
+<br>
+
+Can you find the **evidence** on heteroskedasticity?
 
@@
 ]
+]
+
+---
+
+.pull-right-center[
+
+.image-top-bottom-0[
+
+```{r}
+set.seed(10086)
+
+# Base model
+mod <- poly_model(sigma = 8)
+dat <- mod$gen(1000)
+this_p_value <- mod$test(dat, test = "RESET", power = 2:4)
+dat %>%
+mod$plot(remove_grid_line = TRUE, theme = theme_light(base_size = 12)) +
+  theme(axis.text = element_blank()) +
+  xlab("Fitted values") +
+  ylab("Residuals")
+```
+
+]
+
+.center[
+.caption[
+Fig.
+2: _One classically good looking residual plot but rejected by the RESET test with $p\text{-value} = `r format(round(this_p_value$p_value, digits = 3), nsmall = 3)`$.
+The model is misspecified but the departure from the model assumption is small._
+]
+]
+
+]
+
+.pull-left[
+
+# 🔍Regression Diagnostics🔍
+
+<br>
+<br>
+
+With **large sample sizes**, hypothesis tests may reject a residual plot when there is only a **slight departure** (Belsley, Kuh, and Welsch 1980).
+
+<!-- - While such rejections may be **statistically correct**, their **sensitivity** may render the results **impractical**.
+**Minor defects** in the model are **unlikely** to have a **material impact** and may be best disregarded for practical purposes.
+-->
+
+Residual plot diagnostics:
+  - **Identify** potential issues that could lead to **incorrect conclusions or errors in subsequent analyses**.
+
 ]

### 2023-04-19 Milestone_presentation_2/Milestone_presentation_2/Milestone_presentation_2.Rmd
finish slides

@@
 
+]
+
+--
+
+.float-left[
+
 - Most residual-based tests for particular types of departure are also sensitive to **other types of departures**, resulting in a phenomenon known as the "**Type III error**" (Cook and Weisberg 1982).
+
+]
+
+--
+
+.float-left[
 
@@
 
+--
+
 ### Graphical approaches (plots) are the recommended methods for diagnosing model fits.
+
+--
 
@@
 
+--
+
 - Cook and Weisberg (1982):
@@
 
+--
+
 - Montgomery and Peck (1982):
 > Residual plots are more informative in most practical situations than the corresponding conventional hypothesis tests.
+
+---
+
+.center[
+
+# 🔍Motivation and Research Questions🔍
+
+]
+
+<br>
+<br>
+
+#### Regression experts **consistently recommend plotting residuals for regression diagnostics**, despite the existence of numerous hypothesis test procedures.
+
+#### We would like to **provide evidence** for why this is good advice, using data from a **visual inference experiment**.
+
+#### We would also want to show how **visual inference** can be used to yield reliable and consistent reading of residual plots for better regression diagnostics.
 
@@
 
+--
+
 .pull-right[
@@
 
+]
+
+---
+
+count: false
+
+.center[
+# 🤔False Visual Discoveries🤔
+]
+
+.pull-left-center[
+
+.image-top-bottom-0[
+
+```{r}
+set.seed(10131)
+mod <- heter_model(b = 0, x = rand_lognormal())
+ori_dat <- mod$gen(300)
+
+ori_dat %>%
+  VI_MODEL$plot(theme = theme_light(base_size = 12)) +
+  xlab("Fitted values") +
+  ylab("Residuals")
+```
+
+]
+
+]
+
+.pull-right[
+
+<br>
+<br>
+
+Can you find the **evidence** on heteroskedasticity?
+
 - Vertical spread of the points varies with the fitted values.
 
+]
+
+---
+
+count: false
+
+.center[
+# 🤔False Visual Discoveries🤔
+]
+
+.pull-left-center[
+
+.image-top-bottom-0[
+
+```{r}
+set.seed(10131)
+mod <- heter_model(b = 0, x = rand_lognormal())
+ori_dat <- mod$gen(300)
+
+ori_dat %>%
+  VI_MODEL$plot(theme = theme_light(base_size = 12)) +
+  xlab("Fitted values") +
+  ylab("Residuals")
+```
+
+]
+
+]
+
+.pull-right[
+
+<br>
+<br>
+
+Can you find the **evidence** on heteroskedasticity?
+
+- Vertical spread of the points varies with the fitted values.
+
 However, the residuals are actually obtained from a **correctly specified classical normal linear model**!
 
+]
+
+---
+
+count: false
+
+.center[
+# 🤔False Visual Discoveries🤔
+]
+
+.pull-left-center[
+
+.image-top-bottom-0[
+
+```{r}
+set.seed(10131)
+mod <- heter_model(b = 0, x = rand_lognormal())
+ori_dat <- mod$gen(300)
+
+ori_dat %>%
+  VI_MODEL$plot(theme = theme_light(base_size = 12)) +
+  xlab("Fitted values") +
+  ylab("Residuals")
+```
+
+]
+
+]
+
+.pull-right[
+
+<br>
+<br>
+
+Can you find the **evidence** on heteroskedasticity?
+
+- Vertical spread of the points varies with the fitted values.
+
+However, the residuals are actually obtained from a **correctly specified classical normal linear model**!
+
 **Unconfirmed** visual discoveries could result in **over or under-interpretations of the data**.
@@
 
+.center[
+# 🔬Visual Inference🔬
+]
+
 .pull-left[
 
-# 🔬Visual Inference🔬
-
-<br>
 <br>
@@
 If we embed the residual plot in **a matrix of plots containing residuals simulated from the fitted model**, we will find that it is **not uncommon** for residual plots to exhibit a triangle shape.
-
-- Can you find the **most different plot**?
-
-- Is it the residual plot presented at the previous page?
-
-- If it is not, then the visual pattern is **over-interpreted**!
-
-### The visual discovery is calibrated via comparison.
 
@@
 ```
-.center[
-.caption[
-Fig. 4: _A lineup of 20 residual plots of a simple linear regression. (Plot No.8 is Fig.3) Can you find the most different one?_
-]
-]
-
-]
-
----
+
+]
+
+---
+
+count: false
+
+.center[
+# 🔬Visual Inference🔬
+]
 
@@
 
-# 🔬Visual Inference🔬
-
-<br>
-<br>
-
-This procedure is called **visual inference**. It was introduced by Buja, et al. (2009) as an inferential framework to extend confirmatory statistics to visual discoveries.
-
-A **lineup** consists of $m$ randomly placed plots, where one plot is the **data plot** and the remaining $m − 1$ plots (**null plots**) contain data consistent with the null hypothesis.
-
-To perform a **visual test**, observer(s) will be asked to select the **most different plot(s)** from the lineup.
+<br>
+
+If we embed the residual plot in **a matrix of plots containing residuals simulated from the fitted model**, we will find that it is **not uncommon** for residual plots to exhibit a triangle shape.
+
+- Please pick the **most different plot**.
 
@@
 ```
+
+]
+
+---
+
+count: false
+
+.center[
+# 🔬Visual Inference🔬
+]
+
+.pull-left[
+
+<br>
+
+If we embed the residual plot in **a matrix of plots containing residuals simulated from the fitted model**, we will find that it is **not uncommon** for residual plots to exhibit a triangle shape.
+
+- Please pick the **most different plot**.
+
+- Is it plot No.8?
+
+]
+
+.pull-right-center[
+
+```{r}
+set.seed(10131)
+mod$gen_lineup(300, k = 20, pos = 8) %>%
+  filter(null != FALSE) %>%
+  bind_rows(ori_dat %>% mutate(k = 8, null = FALSE)) %>%
+  VI_MODEL$plot_lineup(theme = theme_light(base_size = 12),
+                       remove_grid_line = TRUE,
+                       remove_axis = TRUE)
+```
+
+]
+
+---
+
+count: false
+
+.center[
+# 🔬Visual Inference🔬
+]
+
+.pull-left[
+
+<br>
+
+If we embed the residual plot in **a matrix of plots containing residuals simulated from the fitted model**, we will find that it is **not uncommon** for residual plots to exhibit a triangle shape.
+
+- Please pick the **most different plot**.
+
+- Is it plot No.8?
+
+- If it is not, then the visual pattern is **over-interpreted**! You can not distinguish it from other plots!
+
+]
+
+.pull-right-center[
+
+```{r}
+set.seed(10131)
+mod$gen_lineup(300, k = 20, pos = 8) %>%
+  filter(null != FALSE) %>%
+  bind_rows(ori_dat %>% mutate(k = 8, null = FALSE)) %>%
+  VI_MODEL$plot_lineup(theme = theme_light(base_size = 12),
+                       remove_grid_line = TRUE,
+                       remove_axis = TRUE)
+```
+
+]
+
+---
+
+count: false
+
+.center[
+# 🔬Visual Inference🔬
+]
+
+.pull-left[
+
+<br>
+
+If we embed the residual plot in **a matrix of plots containing residuals simulated from the fitted model**, we will find that it is **not uncommon** for residual plots to exhibit a triangle shape.
+
+- Please pick the **most different plot**.
+
+- Is it plot No.8?
+
+- If it is not, then the visual pattern is **over-interpreted**! You can not distinguish it from other plots!
+
+### The visual discovery is calibrated via comparison.
+
+]
+
+.pull-right-center[
+
+```{r}
+set.seed(10131)
+mod$gen_lineup(300, k = 20, pos = 8) %>%
+  filter(null != FALSE) %>%
+  bind_rows(ori_dat %>% mutate(k = 8, null = FALSE)) %>%
+  VI_MODEL$plot_lineup(theme = theme_light(base_size = 12),
+                       remove_grid_line = TRUE,
+                       remove_axis = TRUE)
+```
+
+]
+
+---
+
+.center[
+# 🔬Visual Inference🔬
+]
+
+.pull-left[
+
+<br>
+
+This framework is called **visual inference**. It was introduced by Buja, et al. (2009) as an inferential framework to extend confirmatory statistics to visual discoveries.
+
+A **lineup** consists of $m$ randomly placed plots, where one plot is the **data plot** and the remaining $m − 1$ plots (**null plots**) containing data consistent with the null hypothesis.
+
+To perform a **visual test**, observer(s) will be asked to select the **most different plot(s)** from the lineup. And we check if the data plot is identified.
+
+]
+
+.pull-right-center[
+
+```{r}
+set.seed(10131)
+mod$gen_lineup(300, k = 20, pos = 8) %>%
+  filter(null != FALSE) %>%
+  bind_rows(ori_dat %>% mutate(k = 8, null = FALSE)) %>%
+  VI_MODEL$plot_lineup(theme = theme_light(base_size = 12),
+                       remove_grid_line = TRUE,
+                       remove_axis = TRUE)
+```
 .center[
@@
 
-Data used in the $m − 1$ null plots needs to be simulated.
-
---
-
-For classical normal linear regression model, the **residual rotation** technique (Buja, et al. 2009) can be applied:
+Data used in the $m − 1$ null plots needs to be simulated. For **classical normal linear regression model**, the **residual rotation technique** (Buja, et al. 2009) can be applied:
 
@@
 <br>
-<br>
-
---
-
-Under the null hypothesis, it is expected that the data plot would have **no distinguishable difference** with the null plots.
-
-<br>
-
---
-
-Suppose observers are allowed to **select only one plot**. The probability of the observer correctly picks the actual data plot is $1/m$.
-
-<br>
-
---
-
-However, if we involve $K$ **independent observers** in a visual test, and show them the **same lineup**, there will be possible **dependencies** in the visual test due to **repeated evaluations** of the same lineup.
-
----
-
-.pull-left[
-
-```{r}
-alpha_sample <- rgamma(10, 1, 1)
-data.frame(x = 0, y = 0, k = 1:20, alpha = alpha_sample/sum(alpha_sample)) %>%
-ggplot() +
-  geom_text(aes(x, y, label = paste0("theta[", k, "]~'='~", round(alpha, 2))), parse = TRUE, size = 5) +
-  facet_wrap(~k) +
-  theme_light(base_size = 12) +
-  theme(panel.grid = element_blank(),
-        axis.line = element_blank(),
-        axis.title = element_blank(),
-        axis.text = element_blank(),
-        axis.ticks = element_blank())
-```
-
-.center[
-.caption[
-Fig. 5: _Example of the probability of each plot being selected \\(\theta_i\\) under the null hypothesis with parameter \\(\alpha = 1\\)._
-]
-]
-
-]
-
-.pull-right[
-
-# 📈Statistical Significance📈
-
-<br>
-
-VanderPlas et al. (2021) addresses this by modelling the probability of a plot $i$ being selected from a lineup as $\theta_i$:
+
+--
+
+If we involve $K$ **independent observers** in a visual test, and show them the **same lineup**, there will be possible **dependencies** in the visual test due to **repeated evaluations** of the same lineup.
+
+--
+
+VanderPlas et al. (2021) addresses this by modelling the **probability of a plot $i$ being selected** from a lineup as $\theta_i$:
 
@@
 
+--
+
 - This assumes the **attractiveness distribution** of each plot.
 
-For **small** $\alpha$:
-- Only **a few plots** are attractive and tend to be selected.
-
-For **large** $\alpha$:
-- The distribution of the probability of each plot being selected is more **even**.
-
-]
-
----
-
-.pull-left[
-
-```{r}
-data.frame(k = 1:20, 
-           ci = c(1, 0, 0.25, 0, 0, 0.2, 0.45, 0, 5.95, 0, 0, 0, 0, 1.7, 0, 0.2, 0, 1, 0, 0.25)) %>%
-ggplot() +
-  geom_text(aes(0, 0, label = paste0("c[", k, "]~'='~", ci)), parse = TRUE, size = 5) +
-  facet_wrap(~k) +
-  theme_light(base_size = 12) +
-  theme(panel.grid = element_blank(),
-        axis.line = element_blank(),
-        axis.title = element_blank(),
-        axis.text = element_blank(),
-        axis.ticks = element_blank())
-```
-
-.center[
-.caption[
-Fig. 5: _Example of the number of times a plot being selected \\(c_i\\). The sum of all \\(c_i\\) is 11 because the lineup is evaluated by 11 subjects._
-]
-]
-
-]
-
-.pull-right[
-
-# 📈Statistical Significance📈
-
-<br>
-<br>
-
-The **number of times** plot $i$ being selected in $K$ evaluations is denoted as $c_i$.
+--
+
+The **number of times plot $i$ being selected** in $K$ evaluations is denoted as $c_i$.
+
+--
 
@@
 - This ensures $\sum_{i}c_i=K$.
-
-]
 
@@
 .center[
-# 💪Power of Visual Tests💪
-]
-
---
-
-<br>
-
-The power of a visual test may depend on the **ability of the particular subject** (Majumder, Hofmann, and Cook 2013).
-
---
-
-- We assume the **individual skill has negligible effect** on the variation of the power to **simplify the model structure**, thereby **obviate a costly large-scale experiment** to estimate complex covariance matrices.
-
-<br>
-
---
-
-We use the **logistic regression** to estimate the power:
-
-$$Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right),$$
-
-where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
-
-- The **effect size $E$** is the only predictor.
-- The intercept is fixed to $log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$.
-
----
-
-.center[
 # 🧪Experimental Design🧪
@@
 
-.center[
-.caption[
-Fig. 5: _Non-linearity forms generated for the residual plots are controlled by the parameter \\(j\\). And the strength of the signal of the non-linearity model is controlled by \\(\sigma\\)._
-]
-]
-
 ---
@@
 
-```{r out.width = "70%", fig.height = 2.67, fig.width = 8}
+```{r out.width = "80%", fig.height = 2.67, fig.width = 8}
 set.seed(10086)
@@
 
-.center[
-.caption[
-Fig. 5: _Heteroskedasticity forms generated for the residual plots are controlled by the parameter \\(a\\). And the strength of the signal of the heteroskedasticity model is controlled by \\(b\\)._
-]
-]
-
 ---
@@
 
-```{r out.width = "70%"}
+```{r out.width = "80%"}
 magick::image_read_pdf("images/different-n-1.pdf", pages = 1)
@@
 
-]
-
-.center[
-.caption[
-Fig. 5: _Four different distribution of fitted values that might affect perception of residual plots are used. Three differnet values of \\(n\\) are used to control the signal strength._
-]
 ]
@@
 where $\boldsymbol{V}$ is the actual covariance matrix of the error term.
+
+---
+
+.center[
+# 💪Power of Visual Tests💪
+]
+
+--
+
+<br>
+
+The power of a visual test may depend on the **ability of the particular subject** (Majumder, Hofmann, and Cook 2013).
+
+--
+
+- We assume the **individual skill has negligible effect** to **simplify the model structure**, thereby **avoid a costly large-scale experiment** to estimate complex covariance matrices.
+
+--
+
+We use the **logistic regression** to estimate the power:
+
+$$Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right),$$
+
+where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.
+
+--
+
+- The **effect size $E$** is the only predictor.
+
+--
+
+- The intercept is fixed to $log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$.
 
@@
 
-![](images/lineup1.png)
-
-.center[
-.caption[
-Fig. 5: _Screenshot of the study website._
-]
+.center[
+# 🌐Study Website🌐
+]
+
+.image-center[
+
+```{r out.width = "90%"}
+knitr::include_graphics("images/lineup1.png")
+```
+
 ]
@@
 
----
+### (uniform fitted values only)
+
+---
+
+### Non-linearity Patterns
 
@@
 
-.center[
-.caption[
-Fig.
-5: _Comparison of power between different tests for non-linear patterns (uniform fitted values only).
-The horizontal lines of dots represent non-reject and reject results from visual tests for each lineup.
-The visual test has multiple power curves estimated from bootstrap samples.
-The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot._
-]
-]
-
----
+---
+
+### Heteroskedasticity Patterns
 
@@
 
-.center[
-.caption[
-Fig.
-5: _Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only).
-The horizontal lines of dots represent non-reject and reject results from visual tests for each lineup.
-The visual test has multiple power curves estimated from bootstrap samples.
-The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot._
-]
-]
+---
+
+.center[
+# ⚖️Results - Power Comparison of the Tests⚖️
+]
+
+.pull-left[
+
+<br>
+<br>
+
+The data plot (No.1) is **undistinguishable** from other plots with an extremely small effect size (\\(log_e(E) = -0.48\\)).
+
+The non-linearity pattern is **totally undetectable**.
+
+However, the BP test rejects the pattern with a very small $p\text{-value} = 0.004$. In contrast, the $p\text{-value}$ produced by the visual test is $0.813$.
+
+]
+
+.pull-right[
+
+```{r}
+knitr::include_graphics("images/230.png")
+```
+
+]
+
+---
+
+.center[
+# ⚖️Results - Power Comparison of the Tests⚖️
+]
+
+--
+
+Conventional tests are **more sensitive** than the visual test. This is **not necessarily a good feature** for the purposes of diagnosing model defects.
+
+--
+
+Visual tests:
+
+--
+
+- Do not require specifying the pattern **ahead of time**.
+
+--
+
+- Rely purely on **whether the data plot is distinguishable** from "good" residual plots.
+
+--
+
+- Perform **equally well** regardless of the type of residual departures.
+
+--
+
+- **Remove any subjective arguments** about whether a pattern is visible or not.
+
+--
+
+The lineup protocol provides the **calibration for detecting patterns**: that if the pattern in the data plot **can not be distinguished** from patterns in good residual plots, then **no discernible problem** with the model exists.
 
@@
 
+The visual test rejects **less frequently** than the conventional test, and (almost) **only rejects when the conventional test does**.
+
 .image-top-bottom-0[
@@
 
-```{r out.width = "70%"}
+```{r out.width = "80%"}
 magick::image_read_pdf("images/p-value-comparison-1.pdf", pages = 1)
@@
 
-.center[
-.caption[
-Fig.
-5: _Rejection rate (\\(p\\)-value \\(\leq0.05\\)) of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot.
-The visual test rejects less frequently than the conventional test, and (almost) only rejects when the conventional test does.
-Surprisingly, one lineup in the heteroskedasticity group is rejected by the visual test but NOT the conventional test._
-]
-]
-
----
-
-.pull-right-center[
-
-.image-top-bottom-0[
-
-```{r}
-set.seed(10086)
-
-# Base model
-mod <- poly_model(sigma = 8)
-dat <- mod$gen(1000)
-this_p_value <- mod$test(dat, test = "RESET", power = 2:4)
-dat %>%
-mod$plot(remove_grid_line = TRUE, theme = theme_light(base_size = 12)) +
-  theme(axis.text = element_blank()) +
-  xlab("Fitted values") +
-  ylab("Residuals")
-```
-
-]
-
-.center[
-.caption[
-Fig.
-2: _One classically good looking residual plot but rejected by the RESET test with $p\text{-value} = `r format(round(this_p_value$p_value, digits = 3), nsmall = 3)`$.
-The model is misspecified but the departure from the model assumption is small._
-]
-]
-
+---
+
+.center[
+# ⚖️Results - Comparison of Test Decisions⚖️
 ]
@@
 
-# 🔍Regression Diagnostics🔍
-
-<br>
-<br>
-
-With **large sample sizes**, hypothesis tests may reject a residual plot when there is only a **slight departure** (Belsley, Kuh, and Welsch 1980).
-
-<!-- - While such rejections may be **statistically correct**, their **sensitivity** may render the results **impractical**.
-**Minor defects** in the model are **unlikely** to have a **material impact** and may be best disregarded for practical purposes.
--->
-
-Residual plot diagnostics:
-  - **Identify** potential issues that could lead to **incorrect conclusions or errors in subsequent analyses**.
+```{r}
+knitr::include_graphics("images/331.png")
+```
+
+]
+
+.pull-right[
+
+<br>
+<br>
+
+The data plot (No.17) displays a relatively strong heteroskedasticity pattern, and has a strong effect size (\\(log_e(E)=4.02\\)).
+
+This is reflected by the visual test $p\text{-value} = 0.026$.
+
+But the BP test $p\text{-value} = 0.056$, is slightly above the significance cutoff of $0.05$.
+
+Possible reasons:
+- Small sample size
+- Presence of a few outliers
+- Others
 
@@
 ]
+
+The default RESET tests **under-performs significantly** in detecting the "triple-U" shape.
+
+To achieve a similar power as other shapes, a **higher order polynomial parameter** needs to be used for the RESET test, but this higher than the recommended value (4).
 
@@
 
-.center[
-.caption[
-Fig.
-5: _The effect of the order of the polynomial on the power of conventional and visual tests.
-Deeper colour indicates higher order.
-The default RESET tests under-performs significantly in detecting the "triple-U" shape.
-To achieve a similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test, but this higher than the recommended value._
-]
-]
-
 ---
@@
 ]
+
+The butterfly shape has higher power in both tests.
+
+Curiously, the visual test has **slightly higher power** for the "left-triangle" than the "right-triangle" shape.
 
@@
 
-.center[
-.caption[
-Fig.
-5: _The effect of heteroskedasticity shape (parameter \\(a\\)) on the power of conventional and visual tests.
-The butterfly has higher power in both tests.
-Curiously, the visual test has slightly higher power for the "left-triangle" than the "right-triangle" shape, when it would be expected that they should be identical, which is observed in conventional testing._
-]
-]
-
----
-
-class: middle, center
+---
+
+.center[
 
@@
 
----
+]
+
+.pull-left-center-60[
+
+<br>
+<br>
+
+Surprisingly, the fitted value distribution has produces **more variability** in the power of conventional tests than visual tests.
+
+**Uneven distributions** (normal and lognormal distributions) tend to yield lower power.
+
+]
+
+.pull-right-140[
 
@@
 
-```{r out.width = "60%"}
+```{r}
 knitr::include_graphics("images/different-x-dist-poly-power-1.png")
@@
 
-.center[
-.caption[
-Fig.
-5: _Comparison of power on lineups with different fitted value distributions for conventional and visual tests for non-linearity and heteroskedasticity patterns.
-The power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and BP tests, respectively.
-Power curves of visual tests are estimated using five evaluations on each lineup.
-For lineups with a uniform fitted value distribution, the five evaluations are repeatedly sampled from the total eleven evaluations to give multiple power curves (grey).
-Surprisingly, the fitted value distribution has produces more variability in the power of conventional tests than visual tests.
-Uneven distributions, normal and lognormal distributions, tend to yield lower power._
-]
 ]
@@
 # 🧐Conclusion🧐
+]
+
+<br>
+
+--
+
+1. Conventional residual-based statistical tests are **more sensitive to weak departures** from model assumptions than visual tests as would be evaluated by humans.
+
+--
+
+2. Conventional tests often reject when departures in the form of non-linearity and heteroskedasticity **are not visibly different** from null residual plots.
+
+--
+
+3. While conventional tests are correctly detecting the **small but real effects**, human evaluations provide a more **practical** result.
+
+--
+
+4. Residual plots need to be **delivered as a lineup** to enables a **careful calibration**.
+
+--
+
+5. Power of the conventional test are **quite sensitive** to **how the predictors are sampled**.
+
+--
+
+6. The **direction of heteroskedasticity** appears to affect the ability to visually detect it.
+
+---
+.center[
+
+# ⏳Future Works⏳
+
+]
+
+.pull-left[
+## Automatic Visual Statistical Inference System
+
+Develop an automatic visual inference system for evaluating lineups of residual plots.
+
+Computer vision models will be trained on simulated data from linear models that violate classical assumptions such as linearity and homoscedasticity.
+
+The performance of the system will be evaluated using the data collected for the first paper.
+
+]
+
+.pull-right[
+## OAVIS: An online automatic visual inference system
+
+Deploy the automatic visual inference as an online application and publish relevant open-source software packages developed during the course of the research. 
 ]

### 2023-04-19 Milestone_presentation_2/Milestone_presentation_2/Milestone_presentation_2.Rmd
fix bib

@@
 
+<br>
+<br>
+
 Many different **residual-based hypothesis tests** are available to detect specific model defects.
@@
 - **Jarque–Bera test** (Jarque and Bera 1980)
-
-]
-
---
-
-.float-left[
-
-They are **very good at detecting violations of model assumptions**, but there are some **unexpected usages**.
-
-]
-
---
-
-.float-left[
-
-- Most residual-based tests for particular types of departure are also sensitive to **other types of departures**, resulting in a phenomenon known as the "**Type III error**" (Cook and Weisberg 1982).
-
-]
-
---
-
-.float-left[
-
-- Other problems such as **outliers** can trigger the rejection (Cook and Weisberg 1999).
-
-]
-
----
-
-.center[
-# 🔍Regression Diagnostics🔍
-]
-
-.pull-left-center[
-
-.image-top-bottom-0[
-
-```{r}
-set.seed(10086)
-# Base model
-mod <- poly_model(include_z = FALSE, sigma = 0.25)
-# Base data
-dat <- mod$gen(300)
-bind_rows(
-  dat %>%
-    mutate(type = "(A) Good residuals"),
-  mod$set_prm("include_z", TRUE)$
-    gen(300, computed = select(dat, x, e)) %>%
-    mutate(type = "(B) Non-linearity"),
-  heter_model(b = 64)$
-    gen(300, computed = select(dat, x, e)) %>%
-    mutate(type = "(C) Heteroskedasticity"),
-  heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
-    gen(300, computed = select(dat, x)) %>%
-    mutate(type = "(D) Non-normality")
-) %>%
-  mod$plot(remove_grid_line = TRUE, theme = theme_light(base_size = 12)) +
-  facet_wrap(~type, ncol = 2, scales = "free") +
-  theme(axis.text = element_blank(),
-        axis.ticks = element_blank()) +
-  xlab("Fitted values") +
-  ylab("Residuals")
-```
-
-]
-
-]
-
-.pull-right[
-
-All the hypothesis tests **correctly reject** the correspoding model defects, and **correctly not reject** the structure in residual plot (A).
-
-BP and SW tests reject the residual plots exhibiting structure that **they weren’t designed for**.
-
-<br>
-<br>
-
-```{r}
-set.seed(10086)
-# Base model
-mod <- poly_model(include_z = FALSE, sigma = 0.25)
-# Base data
-dat0 <- mod$gen(300)
-# Replicate data in Figure 1
-dat1 <- mod$set_prm("include_z", TRUE)$gen(300, computed = select(dat0, x, e))
-dat2 <- heter_model(b = 64)$gen(300, computed = select(dat0, x, e))
-dat3 <- heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
-               gen(300, computed = select(dat0, x))
-dat_list <- list(dat0, dat1, dat2, dat3)
-table_dat <- data.frame(plot = c("A", "B", "C", "D"), 
-                        model = c("None", 
-                                  "Non-linearity", 
-                                  "Heteroskedasticity", 
-                                  "Non-normality"),
-                        r = map_dbl(dat_list, 
-                                    ~POLY_MODEL$test(.x, 
-                                                     test = "RESET", 
-                                                     power = 2:4)$p_value),
-                        b = map_dbl(dat_list, 
-                                    ~HETER_MODEL$test(.x)$p_value),
-
-                        # SW test has not been included by visage
-                        s = map_dbl(dat_list, 
-                                    ~shapiro.test(.x$.resid)$p.value)) %>%
-
-  # 3 decimal points
-  mutate(across(r:s, ~ format(round(.x, digits = 3), nsmall = 3))) %>%
-
-  # Italic if reject
-  mutate(across(r:s, ~ kableExtra::cell_spec(.x, italic = as.numeric(.x) <= 0.05, color = ifelse(as.numeric(.x) <= 0.05, "red", "black"))))
-  table_dat %>%
-  knitr::kable(col.names = c("Plot", 
-                             "Departures", 
-                             "RESET", 
-                             "BP", 
-                             "SW"),
-               align = "llrrr",
-               escape = FALSE,
-               format = "html",
-               booktabs = TRUE)
-```
 
@@
 <br>
-
-This framework is called **visual inference**. It was introduced by Buja, et al. (2009) as an inferential framework to extend confirmatory statistics to visual discoveries.
-
-A **lineup** consists of $m$ randomly placed plots, where one plot is the **data plot** and the remaining $m − 1$ plots (**null plots**) containing data consistent with the null hypothesis.
+<br>
+
+This framework is called **visual inference** (Buja, et al. 2009).
+
+A **lineup** consists of $m$ randomly placed plots, where one plot is the **data plot** and the remaining $m − 1$ plots (**null plots**) containing data **consistent with the null hypothesis**.
 
@@
 
-### (uniform fitted values only)
-
 ---
@@
 # ⚖️Results - Power Comparison of the Tests⚖️
+
 ]
@@
 # ⚖️Results - Comparison of Test Decisions⚖️
+
 ]
@@
 # ⚖️Results - Effect of Amount of Non-linearity⚖️
+
 ]
@@
 # ⚖️Results - Effect of Shape of Heteroskedasticity⚖️
+
 ]
@@
 ]
+
+.font-70-p[
+
+Belsley, David A, Edwin Kuh, and Roy E Welsch. 1980. Regression diagnostics: Identifying influential data and sources of collinearity. John Wiley & Sons.
+
+Breusch, T. S., and A. R. Pagan. 1979. “A Simple Test for Heteroscedasticity and Random Coefficient Variation.” Econometrica 47 (5): 1287–1294.
+
+Buja, Andreas, Dianne Cook, Heike Hofmann, Michael Lawrence, Eun-Kyung Lee, Deborah F.
+Swayne, and Hadley Wickham.
+2009.
+“Statistical inference for exploratory data analysis and model diagnostics.” Philosophical Transactions of the Royal Society A: Mathematical, Physical and Engineering Sciences 367 (1906): 4361–4383.
+
+Cook, R Dennis, and Sanford Weisberg. 1982. Residuals and influence in regression. New York: Chapman and Hall.
+
+Cook, R Dennis, and Sanford Weisberg. 1999. Applied regression including computing and graphics. John Wiley & Sons.
+
+Draper, Norman R, and Harry Smith. 1998. Applied regression analysis. Vol. 326. John Wiley & Sons.
+
+Jarque, Carlos M, and Anil K Bera. 1980. “Efficient tests for normality, homoscedasticity and serial independence of regression residuals.” Economics letters 6 (3): 255–259.
+
+Kullback, Solomon, and Richard A Leibler. 1951. “On information and sufficiency.” The annals
+of mathematical statistics 22 (1): 79–86.
+
+Ramsey, J. B. 1969. “Tests for Specification Errors in Classical Linear Least-Squares Regression
+Analysis.” Journal of the Royal Statistical Society. Series B (Methodological) 31 (2): 350–
+371.
+
+Shapiro, Samuel Sanford, and Martin B Wilk. 1965. “An analysis of variance test for normality (complete samples).” Biometrika 52 (3/4): 591–611.
+
+VanderPlas, Susan, Christian R ̈ottger, Dianne Cook, and Heike Hofmann. 2021. “Statistical significance calculations for scenarios in visual inference.” Stat 10 (1): e337.
+
+White, Halbert. 1980. “A Heteroskedasticity-Consistent Covariance Matrix Estimator and a Direct Test for Heteroskedasticity.” Econometrica 48 (4): 817–838.
+
+]

### 2024-02-19 Milestone_3/report/report.Rmd
update milestone 3

--- 2023-04-11 Milestone_2/report/report.Rmd
@@
 department: Department of\newline Econometrics &\newline Business Statistics
-organization: "Progress review" 
+organization: "Pre-submission review" 
 bibliography: report.bib
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, cache=TRUE, messages=FALSE, warning=FALSE)
+knitr::opts_chunk$set(echo = FALSE, cache=FALSE, messages=FALSE, warning=FALSE)
 library(visage)
@@
 ```
-
-keywords:
-AI, data visualization, residual plot, visual inference, hypothesis testing, computer vision
 
@@
 
-When conducting model diagnostics, despite the availability of numeric summaries endorsed by finite or asymptotic properties, graphical representations of data are often preferred or required by data analysts.
+When conducting model diagnostics, graphical representations of data are often preferred or required by data analysts.
 The preference for visual diagnostics is attributed to its intuitive nature and the possibility of discovering unexpected abstract and unquantifiable insights.
@@
 The reliance of human assessment is a fundamental aspect of visual tests, but it may restrict its widespread usage.
-The lineup protocol is unsuitable for large-scale applications, due to its high labor costs and time requirements.
+The lineup protocol is unsuitable for large-scale applications, due to its high labour costs and time requirements.
 Moreover, it presents significant usability issues for individuals with visual impairments, resulting in reduced accessibility.
 
-Modern computer vision models offer a promising solution to this challenge.
-As a subfield of AI, computer vision with its modern deep learning architectures has successfully resolved numerous critical problems in automation.
-The development of the convolutional neural network (CNN) by @fukushima_neocognitron_1982 was inspired by the vision processing in living organisms.
-The modern computer vision model typically utilizes deep neural networks with convolutional layers, which leverage the hierarchical pattern in data and provide regularized versions of fully-connected layers.
-This approach downscales and transforms images by summarizing information in a small space.
-Numerous studies have shown that it can effectively tackle vision tasks, such as image recognition [@rawat_deep_2017], computer-aided diagnosis [@lee_image_2015], pedestrian detection [@brunetti_computer_2018], and facial recognition [@emami_facial_2012].
-
-Utilizing computer vision models in reading data plot is not a common choice.
-Nevertheless, certain fields have adopted this idea by applying computer vision models in reading recurrence plots for time series regression [@ojeda_multivariate_2020], time series classification [@chu_automatic_2019; @hailesilassie_financial_2019; @hatami_classification_2018; @zhang_encoding_2020], anomaly detection [@chen_convolutional_2020], and pairwise causality analysis [@singh_deep_2017].
-However, the assessment of lineups with computer vision models is a relatively novel area of study.
+Modern computer vision models are well-suited for addressing this challenge.
+They rely on deep neural networks with convolutional layers [@fukushima_neocognitron_1982].
+These layers leverage hierarchical patterns in data, downsizing and transforming images by summarizing information in a small space.
+Numerous studies have demonstrated the efficacy of convolutional layers in addressing various vision tasks, including image recognition [@rawat_deep_2017].
+Despite the widespread use of computer vision models in fields like computer-aided diagnosis [@lee_image_2015], pedestrian detection [@brunetti_computer_2018], and facial recognition [@emami_facial_2012], their application in reading data plots remains limited.
+While some studies have explored the use of computer vision models for tasks such as reading recurrence plots for time series regression [@ojeda_multivariate_2020], time series classification [@chu_automatic_2019; @hailesilassie_financial_2019; @hatami_classification_2018; @zhang_encoding_2020], anomaly detection [@chen_convolutional_2020], and pairwise causality analysis [@singh_deep_2017], the application of reading residual plots with computer vision models represents a relatively new field of study.
 
@@
 
-In order to examine the potential applicability of integrating visual inference techniques into regression diagnostics, a pilot study was conducted in the first year with the participation of 64 individuals, which was followed by a formal study involving 443 participants in the subsequent year.
+### A plot is worth a thousand tests: assessing residual diagnostics with the lineup protocol
+
+In order to collect data of human performance on reading residual plot of linear regression model and examine the potential applicability of integrating visual inference techniques into regression diagnostics, a pilot study was conducted in the first year with the participation of 64 individuals, which was followed by a formal study involving 443 participants in the subsequent year.
 The participants were presented with lineups consisting of 20 plots, where a residual plot was embedded along with 19 null plots, drawn with data simulated using the residual rotation technique.
 The study considered two primary forms of residual departures in multiple linear regression, namely non-linearity and heteroskedasticity.
-To enrich the visual features, different fitted value distributions, including normal, uniform, lognormal, and discrete, were also incorporated.
+To enrich the visual features, different fitted value distributions, including normal, uniform, log-normal, and discrete, were also incorporated.
 
-As shown in Figure \ref{fig:polypower} and Figure \ref{fig:heterpower}, the study revealed that appropriate conventional residual-based statistical tests are more sensitive to weak residual departures from model assumptions compared to visual tests evaluated by humans, leading to excessive rejections even when downstream analysis and outcomes would not be significantly impacted by the small departures from a good fit.
+The study revealed that appropriate conventional residual-based statistical tests are more sensitive to weak residual departures from model assumptions compared to visual tests evaluated by humans, leading to excessive rejections even when downstream analysis and outcomes would not be significantly impacted by the small departures from a good fit.
 Specifically, conventional tests conclude that the model fit has issues nearly twice as frequently as humans would, and they frequently reject departures in the form of non-linearity and heteroskedasticity that are not visible to humans.
-For instance, based on the row of scatter plots at the bottom of Figure \ref{fig:polypower}, we might argue that the non-linearity is not sufficiently problematic until an effect size of around 3 or 3.5.
-The RESET test would reject closer to an effect size of 2, but the visual test would reject closer 3.25.
-
-![Comparison of power between different tests for detecting non-linearity.
-The power curves are estimated using logistic regression, and the horizontal lines of dots represent non-reject and reject results from human observers for each lineup.
-The visual test has multiple power curves estimated from bootstrap samples.
-The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.\label{fig:polypower}](polypower-1.png)
-
-![Comparison of power between different tests for detecting heteroskedasticity.
-Main plot shows the power curves, with dots indicating non-reject and reject in visual testing of lineups.
-The multiple lines for the visual test arise from estimating the power on many bootstrap samples.
-The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.\label{fig:heterpower}](heterpower-1.png)
 
@@
 Moreover, they provide a compelling rationale for the development of an automated visual inference system to evaluate lineups of residual plots.
-The comprehensive details of the research methodology, along with other intriguing findings can be found in the attached paper.
+The comprehensive details of the research methodology, along with other intriguing findings can be found in @li2023plot.
+
+### Automated assessment of residual plots with computer vision models
+
+In the third year of the project, a series of endeavors were undertaken to develop a computer vision model for assessing residual plots.
+To properly train such a model, a distance measure based on Kullback-Leibler divergence was introduced.
+This distance aimed to quantify the disparity between the residual distribution of a fitted linear regression model and the reference residual distribution assumed under correct model specification.
+It effectively captured the magnitude of model violations in a misspecified model.
+
+Subsequently, the computer vision model was devised to approximate this distance by taking a residual plot of the fitted model as input.
+The resulting approximated distance predicted by the model could be utilized to construct a Model Violation Index for quantifying various model violations.
+Additionally, a formal statistical testing procedure was developed based on the approximated distance.
+This procedure produced critical values by evaluating a large number of null plots generated from the fitted model and estimating the sample quantiles of the distribution of the approximated distance.
+Furthermore, a bootstrapping technique was developed to measure how frequently the fitted model was considered misspecified if data were repeatedly obtained from the same data generating process.
+
+The proposed computer vision model was trained exclusively on synthetic data, encompassing residual plots featuring non-linearity, heteroskedasticity, and non-normality issues.
+The trained model exhibited strong performance on both training and test sets.
+Statistical tests relying on the approximated distance predicted by the computer vision model showcased lower sensitivity compared to conventional tests but higher sensitivity compared to visual tests conducted by humans.
+The approximated distance generally aligned with the strength of the visual signal perceived by humans.
+
+Through various examples, the effectiveness of the proposed method were illustrated across diverse scenarios, highlighting the similarity between visual tests and tests based on approximated distance.
+Furthermore, cases where conventional tests might fail, whereas visual tests and tests based on approximated distance remain effective were underscored.
+
+This method holds significant value as it alleviates a portion of analysts' workload associated with assessing residual plots.
+While plotting residuals and evaluating them by analysts are still recommend for their invaluable insights, the trained computer vision model provided in this study can serve as a tool for automating the diagnostic process or for supplementary purposes as needed.
+The comprehensive details of the distance definition, model training process, along with other results can be found in the attached paper.
 
@@
 
-2. Chapter Two presents a comparative study between visual testing and conventional testing in the context of regression diagnostics. This chapter will be identical to the attached paper.
+2.
+Chapter Two presents a comparative study between visual testing and conventional testing in the context of regression diagnostics.
+The corresponding paper is already accepted by the Journal of Computational and Graphical Statistics.
 
@@
 The chapter concludes with an evaluation of the performance of the system, comparing the power and characteristics of human subject-assessed visual tests with computer-assessed visual tests.
+This chapter will be identical to the attached paper.
 
@@
 Chapter Four provides detailed description of the open-source software developed during the course of the research.
-This includes the R package `bandicoot` for code maintenance with object-oriented programming, `visage` for running visual inference experiments on linear regression models, and one or multiple R packages for employment of the automatic visual inference system.
+This includes the R package `bandicoot` for code maintenance with object-oriented programming, `visage` for running visual inference experiments on linear regression models, and `autovi` for assessing residual plots with trained computer vision models.
 Additionally, an online application of the automatic visual inference system is introduced.
-This chapter may be splitted into multiple chapters to provide a more in-depth description of each tool.
 
@@
 
-The timetable of my PhD study from April 2023 to August 2024 is provided in Table \ref{tab:timetable}.
+The timetable of my PhD study from Mar 2024 to August 2024 is provided in Table \ref{tab:timetable}.
 
 ```{r}
-data.frame(Date = c("April", 
+data.frame(Date = c("Mar",
+                    "April", 
                     "May",
-                    "June",
                     "July",
-                    "Sep",
-                    "Oct",
-                    "Nov",
-                    "Dec",
-                    "Mar",
+                    "",
                     "Aug"), 
-           Event = c("Submit the abstract of the first paper to Australian Statistical Conference (ASC)", 
-                     "Submit the first paper",
-                     "Submit a poster for the first paper to Institute of Electrical and Electronics Engineers (IEEE) Visualization Conference",
-                     "Leave for a month",
-                     "Finalize the computer vision model",
-                     "Attend the IEEE Visualization Conference",
-                     "Start the web interface development",
-                     "Attend the ASC",
-                     "Submit the second paper",
+           Event = c("Finish and submit the second paper",
+                     "Develop a web interface for the automatic visual inference system",
+                     "Clean and publish R packages to CRAN",
+                     "Finish the third paper",
+                     "Attend the useR! conference",
                      "Submit thesis")) %>%
@@
                     caption = "Timetable till thesis submission") %>%
-  kableExtra::kable_styling(latex_options = "scale_down") %>%
-  kableExtra::group_rows("2023", 1, 8) %>%
-  kableExtra::group_rows("2024", 9, 10)
+  kableExtra::kable_styling(latex_options = "scale_down") 
 ```

### 2024-02-19 Milestone_3/report/report.tex
update milestone 3

--- 2023-04-11 Milestone_2/report/report.tex
@@
 \makeatother
-  \organization{Progress review}
+  \organization{Pre-submission review}
 
@@
 \titlepage
-
-keywords:
-AI, data visualization, residual plot, visual inference, hypothesis testing, computer vision
 
@@
 The reliance of human assessment is a fundamental aspect of visual tests, but it may restrict its widespread usage.
-The lineup protocol is unsuitable for large-scale applications, due to its high labor costs and time requirements.
+The lineup protocol is unsuitable for large-scale applications, due to its high labour costs and time requirements.
 Moreover, it presents significant usability issues for individuals with visual impairments, resulting in reduced accessibility.
 
-Modern computer vision models offer a promising solution to this challenge.
-As a subfield of AI, computer vision with its modern deep learning architectures has successfully resolved numerous critical problems in automation.
-The development of the convolutional neural network (CNN) by \textcite{fukushima_neocognitron_1982} was inspired by the vision processing in living organisms.
-The modern computer vision model typically utilizes deep neural networks with convolutional layers, which leverage the hierarchical pattern in data and provide regularized versions of fully-connected layers.
-This approach downscales and transforms images by summarizing information in a small space.
-Numerous studies have shown that it can effectively tackle vision tasks, such as image recognition \autocite{rawat_deep_2017}, computer-aided diagnosis \autocite{lee_image_2015}, pedestrian detection \autocite{brunetti_computer_2018}, and facial recognition \autocite{emami_facial_2012}.
-
-Utilizing computer vision models in reading data plot is not a common choice.
-Nevertheless, certain fields have adopted this idea by applying computer vision models in reading recurrence plots for time series regression \autocite{ojeda_multivariate_2020}, time series classification \autocite{chu_automatic_2019,hailesilassie_financial_2019,hatami_classification_2018,zhang_encoding_2020}, anomaly detection \autocite{chen_convolutional_2020}, and pairwise causality analysis \autocite{singh_deep_2017}.
-However, the assessment of lineups with computer vision models is a relatively novel area of study.
+Modern computer vision models are well-suited for addressing this challenge.
+They rely on deep neural networks with convolutional layers \autocite{fukushima_neocognitron_1982}.
+These layers leverage hierarchical patterns in data, downsizing and transforming images by summarizing information in a small space.
+Numerous studies have demonstrated the efficacy of convolutional layers in addressing various vision tasks, including image recognition \autocite{rawat_deep_2017}.
+Despite the widespread use of computer vision models in fields like computer-aided diagnosis \autocite{lee_image_2015}, pedestrian detection \autocite{brunetti_computer_2018}, and facial recognition \autocite{emami_facial_2012}, their application in reading data plots remains limited.
+While some studies have explored the use of computer vision models for tasks such as reading recurrence plots for time series regression \autocite{ojeda_multivariate_2020}, time series classification \autocite{chu_automatic_2019,hailesilassie_financial_2019,hatami_classification_2018,zhang_encoding_2020}, anomaly detection \autocite{chen_convolutional_2020}, and pairwise causality analysis \autocite{singh_deep_2017}, the application of reading residual plots with computer vision models represents a relatively new field of study.
 
@@
 
-In order to examine the potential applicability of integrating visual inference techniques into regression diagnostics, a pilot study was conducted in the first year with the participation of 64 individuals, which was followed by a formal study involving 443 participants in the subsequent year.
+\hypertarget{a-plot-is-worth-a-thousand-tests-assessing-residual-diagnostics-with-the-lineup-protocol}{%
+\subsubsection{A plot is worth a thousand tests: assessing residual diagnostics with the lineup protocol}\label{a-plot-is-worth-a-thousand-tests-assessing-residual-diagnostics-with-the-lineup-protocol}}
+
+In order to collect data of human performance on reading residual plot of linear regression model and examine the potential applicability of integrating visual inference techniques into regression diagnostics, a pilot study was conducted in the first year with the participation of 64 individuals, which was followed by a formal study involving 443 participants in the subsequent year.
 The participants were presented with lineups consisting of 20 plots, where a residual plot was embedded along with 19 null plots, drawn with data simulated using the residual rotation technique.
 The study considered two primary forms of residual departures in multiple linear regression, namely non-linearity and heteroskedasticity.
-To enrich the visual features, different fitted value distributions, including normal, uniform, lognormal, and discrete, were also incorporated.
+To enrich the visual features, different fitted value distributions, including normal, uniform, log-normal, and discrete, were also incorporated.
 
-As shown in Figure \ref{fig:polypower} and Figure \ref{fig:heterpower}, the study revealed that appropriate conventional residual-based statistical tests are more sensitive to weak residual departures from model assumptions compared to visual tests evaluated by humans, leading to excessive rejections even when downstream analysis and outcomes would not be significantly impacted by the small departures from a good fit.
+The study revealed that appropriate conventional residual-based statistical tests are more sensitive to weak residual departures from model assumptions compared to visual tests evaluated by humans, leading to excessive rejections even when downstream analysis and outcomes would not be significantly impacted by the small departures from a good fit.
 Specifically, conventional tests conclude that the model fit has issues nearly twice as frequently as humans would, and they frequently reject departures in the form of non-linearity and heteroskedasticity that are not visible to humans.
-For instance, based on the row of scatter plots at the bottom of Figure \ref{fig:polypower}, we might argue that the non-linearity is not sufficiently problematic until an effect size of around 3 or 3.5.
-The RESET test would reject closer to an effect size of 2, but the visual test would reject closer 3.25.
-
-\begin{figure}
-\centering
-\includegraphics{polypower-1.png}
-\caption{Comparison of power between different tests for detecting non-linearity.
-The power curves are estimated using logistic regression, and the horizontal lines of dots represent non-reject and reject results from human observers for each lineup.
-The visual test has multiple power curves estimated from bootstrap samples.
-The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.\label{fig:polypower}}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{heterpower-1.png}
-\caption{Comparison of power between different tests for detecting heteroskedasticity.
-Main plot shows the power curves, with dots indicating non-reject and reject in visual testing of lineups.
-The multiple lines for the visual test arise from estimating the power on many bootstrap samples.
-The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.\label{fig:heterpower}}
-\end{figure}
 
@@
 Moreover, they provide a compelling rationale for the development of an automated visual inference system to evaluate lineups of residual plots.
-The comprehensive details of the research methodology, along with other intriguing findings can be found in the attached paper.
+The comprehensive details of the research methodology, along with other intriguing findings can be found in \textcite{li2023plot}.
+
+\hypertarget{automated-assessment-of-residual-plots-with-computer-vision-models}{%
+\subsubsection{Automated assessment of residual plots with computer vision models}\label{automated-assessment-of-residual-plots-with-computer-vision-models}}
+
+In the third year of the project, a series of endeavors were undertaken to develop a computer vision model for assessing residual plots.
+To properly train such a model, a distance measure based on Kullback-Leibler divergence was introduced.
+This distance aimed to quantify the disparity between the residual distribution of a fitted linear regression model and the reference residual distribution assumed under correct model specification.
+It effectively captured the magnitude of model violations in a misspecified model.
+
+Subsequently, the computer vision model was devised to approximate this distance by taking a residual plot of the fitted model as input.
+The resulting approximated distance predicted by the model could be utilized to construct a Model Violation Index for quantifying various model violations.
+Additionally, a formal statistical testing procedure was developed based on the approximated distance.
+This procedure produced critical values by evaluating a large number of null plots generated from the fitted model and estimating the sample quantiles of the distribution of the approximated distance.
+Furthermore, a bootstrapping technique was developed to measure how frequently the fitted model was considered misspecified if data were repeatedly obtained from the same data generating process.
+
+The proposed computer vision model was trained exclusively on synthetic data, encompassing residual plots featuring non-linearity, heteroskedasticity, and non-normality issues.
+The trained model exhibited strong performance on both training and test sets.
+Statistical tests relying on the approximated distance predicted by the computer vision model showcased lower sensitivity compared to conventional tests but higher sensitivity compared to visual tests conducted by humans.
+The approximated distance generally aligned with the strength of the visual signal perceived by humans.
+
+Through various examples, the effectiveness of the proposed method were illustrated across diverse scenarios, highlighting the similarity between visual tests and tests based on approximated distance.
+Furthermore, cases where conventional tests might fail, whereas visual tests and tests based on approximated distance remain effective were underscored.
+
+This method holds significant value as it alleviates a portion of analysts' workload associated with assessing residual plots.
+While plotting residuals and evaluating them by analysts are still recommend for their invaluable insights, the trained computer vision model provided in this study can serve as a tool for automating the diagnostic process or for supplementary purposes as needed.
+The comprehensive details of the distance definition, model training process, along with other results can be found in the attached paper.
 
@@
 \item
-  Chapter Two presents a comparative study between visual testing and conventional testing in the context of regression diagnostics. This chapter will be identical to the attached paper.
+  Chapter Two presents a comparative study between visual testing and conventional testing in the context of regression diagnostics.
+The corresponding paper is already accepted by the Journal of Computational and Graphical Statistics.
 \item
@@
 The chapter concludes with an evaluation of the performance of the system, comparing the power and characteristics of human subject-assessed visual tests with computer-assessed visual tests.
+This chapter will be identical to the attached paper.
 \item
   Chapter Four provides detailed description of the open-source software developed during the course of the research.
-This includes the R package \texttt{bandicoot} for code maintenance with object-oriented programming, \texttt{visage} for running visual inference experiments on linear regression models, and one or multiple R packages for employment of the automatic visual inference system.
+This includes the R package \texttt{bandicoot} for code maintenance with object-oriented programming, \texttt{visage} for running visual inference experiments on linear regression models, and \texttt{autovi} for assessing residual plots with trained computer vision models.
 Additionally, an online application of the automatic visual inference system is introduced.
-This chapter may be splitted into multiple chapters to provide a more in-depth description of each tool.
 \item
@@
 
-The timetable of my PhD study from April 2023 to August 2024 is provided in Table \ref{tab:timetable}.
+The timetable of my PhD study from Mar 2024 to August 2024 is provided in Table \ref{tab:timetable}.
 
@@
 \midrule
-\addlinespace[0.3em]
-\multicolumn{2}{l}{\textbf{2023}}\\
-\hspace{1em}April & Submit the abstract of the first paper to Australian Statistical Conference (ASC)\\
-\hspace{1em}May & Submit the first paper\\
-\hspace{1em}June & Submit a poster for the first paper to Institute of Electrical and Electronics Engineers (IEEE) Visualization Conference\\
-\hspace{1em}July & Leave for a month\\
-\hspace{1em}Sep & Finalize the computer vision model\\
-\hspace{1em}Oct & Attend the IEEE Visualization Conference\\
-\hspace{1em}Nov & Start the web interface development\\
-\hspace{1em}Dec & Attend the ASC\\
-\addlinespace[0.3em]
-\multicolumn{2}{l}{\textbf{2024}}\\
-\hspace{1em}Mar & Submit the second paper\\
-\hspace{1em}Aug & Submit thesis\\
+Mar & Finish and submit the second paper\\
+April & Develop a web interface for the automatic visual inference system\\
+May & Clean and publish R packages to CRAN\\
+July & Finish the third paper\\
+ & Attend the useR! conference\\
+\addlinespace
+Aug & Submit thesis\\
 \bottomrule

### 2024-02-21 Milestone_presentation_3/index.qmd
finish p3

---
format: 
  revealjs:
    transition-speed: fast
    slide-number: c/t
    css: custom.css
    transition: fade
    incremental: true 
    theme: default
    footer: "Slides URL: <https://patrickli-milestone-3.netlify.app>"
    logo: figures/monash-stacked-blue-rgb.png
    self-contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE
)

library(tidyverse)
```

## Advances in Artificial Intelligence for Data Visualization: Automated Reading of Residual Plots with Computer Vision {.center style="text-align: right;"}

::: {style="color: #5A5A5A"}

Pre-submission Review

:::

#### Weihao (Patrick) Li {style="margin-top: 40px; font-size: 0.9em"}

::: {style="font-size: 0.9em"}

Monash University

:::

---

## ✍️Supervisors

::: {.columns style="font-size:50%"}

::: {.column width="25%"}

![](figures/dicook.jpg){style="object-fit: cover; width: 100%; aspect-ratio: 1 / 1;"}

Professor Dianne Cook, Department of Econometrics and Business Statistics, Melbourne, Monash University, Australia

:::

::: {.column width="25%"}

![](figures/emitanaka.jpg){style="object-fit: cover; width: 100%; aspect-ratio: 1 / 1;"}

Dr. Emi Tanaka, Biological Data Science Institute, Australian National University, Canberra, Australia

:::

::: {.column width="25%"}

![](figures/susan.jpg){style="object-fit: cover; width: 100%; aspect-ratio: 1 / 1;"}

Assistant Professor Susan VanderPlas, Statistics Department, University of Nebraska, Lincoln, USA

:::

::: {.column width="25%"}

![](figures/klaus.jpeg){style="object-fit: cover; width: 100%; aspect-ratio: 1 / 1;"}

Senior Lecturer Klaus Ackermann, Department of Econometrics and Business Statistics, Melbourne, Monash University, Australia

:::

:::

---

## 📚Thesis Structure

1. Exploring the application of visual inference in regression diagnostics and comparing it with conventional hypothesis tests.

2. **Designing an automated visual inference system to assess residual plots of classical normal linear regression model.**

3. Deploying the automatic visual inference system as an online application and publishing the relevant open-source software.

::: {.fragment}

This presentation will be focused on the **second project**.

:::

---

## 🔍Regression Diagnostics

Diagnostics are the key to determining whether there is anything **importantly wrong** with a regression model.

<br>

$$\underbrace{\boldsymbol{e}}_\textrm{Residuals} = \underbrace{\boldsymbol{y}}_\textrm{Observations} - \underbrace{f(\boldsymbol{x})}_\textrm{Fitted values}$$

**Graphical approaches (plots)** are the recommended methods for diagnosing residuals.

---

## 🤔Challenges

:::: {.columns}

::: {.column width="40%"}

```{r fig.width=5, fig.height=5, fig.retina=3, warning=FALSE, message=FALSE}
library(tidyverse)
library(visage)
set.seed(452)
ori_x <- rand_lognormal()
mod <- heter_model(b = 0, x = closed_form(~-ori_x))
ori_dat <- mod$gen(300)

ori_dat %>%
  VI_MODEL$plot(theme = theme_light(base_size = 18), size = 1, remove_grid_line = TRUE, ) +
  # geom_line(aes(x = .fitted, y = (3.5 + 0.3 * .fitted)), col = "red") +
  # geom_line(aes(x = .fitted, y = -(3.5 + 0.3 * .fitted)), col = "red") +
  xlab("Fitted values") +
  ylab("Residuals")
```

:::

::: {.column width="60%"}

- Vertical spread of the points varies with the fitted values indicates **the existence of heteroskedasticity**.

- However, this is an **over-interpretation**.

- The visual pattern is caused by a **skewed distribution of the predictor**.

:::

::::

---

## 🔬Visual Inference

::: {style="font-size:70%"}

The reading of residual plots can be **calibrated** by an **inferential framework** called **visual inference** (Buja, et al. 2009).

:::: {.columns}

::: {.column width="50%"}

```{r fig.width=5, fig.height=5, fig.retina=3}
set.seed(452)
mod$gen_lineup(300, k = 20, pos = 11) %>%
  filter(null != FALSE) %>%
  bind_rows(ori_dat %>% mutate(k = 11, null = FALSE)) %>%
  VI_MODEL$plot_lineup(theme = theme_light(base_size = 12),
                       remove_grid_line = TRUE,
                       remove_axis = TRUE)
```

:::

::: {.column width="50%"}

::: {.fragment}

Typically, a **lineup** of residual plots consists of

<!-- - $m$ randomly placed plots -->
- one **data plot**
- $19$ **null plots** containing residuals **simulated from the fitted model**.

:::

::: {.fragment}

To perform a visual test

- Observer(s) will be asked to select the **most different plot(s)**.
- The p-value can be calculated using the **beta-binomial model** (VanderPlas et al., 2021).

:::

:::

::::

:::

---

## 🚫Limitations of Lineup Protocol

::: {.fragment}
1. Human can not

- evaluate lineup consisted of a **large number of plots**.
- evaluate a **large number of lineups**.
:::

::: {.fragment}
2. Evaluation of lineup is **high in labour cost** and **time consuming**.
:::

---

## 🤖Computer Vision Model

Modern **computer vision models** are well-suited for addressing this challenge.

It is usually built on a **deep neural network** called **convolutional neural network** (CNN).

![Source: https://en.wikipedia.org/wiki/Convolutional_neural_network](figures/Typical_cnn.png)

---

## 📏Measure the Difference

To develop a computer vision model for assessing lineups of residual plots, we need to define a **numerical measure** of **"difference"** or **"distance"** between plots.

- pixel-wise sum of square differences
- Structural Similarity Index Measure (SSIM)
- scagnostics
- metric learning
- ...

---

## 🎲Residual Distribution

Consider the **classical normal linear regression model**

$$\boldsymbol{y} = \boldsymbol{X}\boldsymbol{\beta} + \boldsymbol{\varepsilon}, \quad \boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n).$$

By the **Frisch-Waugh-Lowell theorem**,

$$\boldsymbol{e} = \boldsymbol{R}\boldsymbol{\varepsilon},$$
where $\boldsymbol{R}=\boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$.

We treat $\boldsymbol{e}$ as **a vector of random variables** here.

---

## 🎲Residual Distribution

::: {style="font-size:80%"}

The **minimization of the sum of square residuals** implies  
$$\sum_{i=1}^{n} e_i = 0 \quad\text{and}\quad \text{rank}(\boldsymbol{R}) = n - 1.$$

Thus, $\boldsymbol{e}$ follows a **degenerate multivariate distribution**.

For simplicity, we will replace $\text{cov}(\boldsymbol{e}, \boldsymbol{e}) = \boldsymbol{R}\sigma^2\boldsymbol{R}' = \boldsymbol{R}\sigma^2$ with a **full-rank diagonal matrix** $\text{diag}(\boldsymbol{R}\sigma^2)$.

Then for a **correctly specified model**, $$\boldsymbol{e} \sim N(\boldsymbol{0}_n, \text{diag}(\boldsymbol{R}\sigma^2)).$$

Symbol $Q$ will be used to represent this **reference residual distribution**.

:::

---

## 🎲Residual Distribution

However, if the model is **misspecified**, then the **actual residual distribution** denoted as $P$, will be **different** from $Q$.

- For example, if $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\boldsymbol{V})$, where $\boldsymbol{V} \neq \sigma^2\boldsymbol{I}_n$,  $$\boldsymbol{e} \sim N(\boldsymbol{0}_n, \text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R})) \implies \text{Heteroskedasticity}.$$

- And if some necessary higher-order predictors $\boldsymbol{Z}$ are also omitted, $$\boldsymbol{e} \sim N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, \text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R})) \implies \text{Non-linearity}~ \& ~\text{Heteroskedasticity}.$$

---

## 📏KL Divergence of $P$ from $Q$

We defined a **distance measure** based on **Kullback-Leibler divergence** to quantify the **extent of model violations**

\begin{align}
\label{eq:kl-0}
D &= \log\left(1 + D_{KL}\right), \\
\label{eq:kl-1}
D_{KL} &= \int_{\mathbb{R}^{n}}\log\frac{p(\boldsymbol{e})}{q(\boldsymbol{e})}p(\boldsymbol{e})d\boldsymbol{e},
\end{align}

\noindent where $p(.)$ and $q(.)$ are the probability density functions for distribution $P$ and $Q$ respectively.

- $D = 0$ if and only if $P \equiv Q$.

---

## 📝Evaluation of $D$ for Normal $\boldsymbol{\varepsilon}$

::: {style="font-size:80%"}

For a **classical normal linear regression model** that omits necessary higher-order predictors $\boldsymbol{Z}$, and incorrectly assumes $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)$ while in fact $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \boldsymbol{V})$ with $\boldsymbol{V} \neq \sigma^2\boldsymbol{I}_n$, the KL divergence can be further expanded to

\begin{align}
\label{eq:kl-2}
D_{KL} &= \frac{1}{2}\left(\log\frac{|\boldsymbol{W}|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\boldsymbol{W}^{-1}\text{diag}(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z'\boldsymbol{W}^{-1}\boldsymbol{\mu}_z\right),
\end{align}

where $\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z$, and $\boldsymbol{W} = \text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R})$.

- The assumed error variance $\sigma^2$ is set to be $\text{tr}(\boldsymbol{V})/n$, which is the expectation of the estimated variance.

:::

---

## 📝Evaluation of $D$ for Non-normal $\boldsymbol{\varepsilon}$

For non-normal $\boldsymbol{\varepsilon}$, $P$ is **unlikely to be a well-known distribution**.

- difficult to compute $p(\boldsymbol{e})$. 
- difficult to do numerical integration over the $n$ dimensional space, because $n$ could be **potentially very large**.

::: {.fragment}

Therefore, the elements of $\boldsymbol{e}$ are assumed to be **independent of each other**.

:::

---

## 📝Evaluation of $D$ for Non-normal $\boldsymbol{\varepsilon}$

::: {style="font-size:80%"}

Under the independence assumption, the **integration becomes a summation**

\begin{align}
\label{eq:kl-3}
D_{KL} &\approx \sum_{i = 1}^{n} \hat{D_{KL}^{(i)}}, \\
\hat{D_{KL}^{(i)}} &= \frac{1}{m}\sum_{j = 1}^{m} log\frac{\hat{p_i}(B_{ij})}{q(B_{ij})},
\end{align}

where $B$ is a matrix with $n$ rows and $m$ columns, and **each column is a set of simulated residuals**, $m$ is the **number of simulations**, $\hat{p_i}(.)$ is the **kernel density estimator** of $p_i(.)$ using simulated residuals in $B_i$.

- The assumed variance for $q(.)$ is set to be $\sum_{b \in vec(B)}(b - \sum_{b \in vec(B)} b/nm)^2/(nm - 1)$.

:::

---

## 🎯Approximation of the Distance

::: {style="font-size:80%"}

Let's get back to the **original problem**. We have a fitted model on hand, and we want to know **how different the residuals are from a set of good residuals**.

- **However, the data generating process is typically unknown!**

  - The distance can not be computed!

::: {.fragment}

We can train a **computer vision model** to approximate $D$ with **a residual plot**

\begin{equation}
\label{eq:d-approx}
\widehat{D} = f_{CV}(V_{h \times w}(\boldsymbol{e}, \boldsymbol{\hat{y}})),
\end{equation}

where $V_{h \times w}(.)$ is a **plotting function** that saves a residual plot as an image with $h \times w$ pixels, and  $f_{CV}(.)$ is a **computer vision model** which predicts distance in $[0, +\infty)$.

:::

:::

---

## 🔄Generation of Training Data

::: {style="font-size:80%"}

### **Data generating process**

\begin{align}
\label{eq:data-sim}
\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 + \beta_2(\boldsymbol{z} + \beta_1\boldsymbol{w}) + \boldsymbol{k} \odot \boldsymbol{\varepsilon}, \\
\boldsymbol{z} &\propto \text{He}_j(\boldsymbol{x}_1), \\
\boldsymbol{w} &\propto \text{He}_j(\boldsymbol{x}_2), \\
\boldsymbol{k} &= \sqrt{\boldsymbol{1}_n + b(2 - |a|)(\boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 - a\boldsymbol{1}_n)^2},
\end{align}

where $\text{He}_j(.)$ is the $j$th-order **probabilist's Hermite polynomials**, and the $\sqrt{(.)}$ and $(.)^2$ operators are element-wise operators.

- The residuals are obtained by **regressing $y$ on $x_1$**. If $\beta_1 \neq 0$, $x_2$ will also be included as a predictor.

:::

---

## 💡Non-linearity

:::: {.columns}

::: {.column width="30%"}

**Factor $j$**

:::

::: {.column width="70%"}

```{r fig.width=10, fig.height=4}
set.seed(10086)

# Data for shape 1
dat_shape_1 <- phn_model(j = 2, include_x2 = FALSE, sigma = 0.05)$gen(500) %>%
  mutate(j = 2)

# Generate data for shape 2, 3 and 4. Reuse x and e.
map_df(3:11, function(j) {
  phn_model(j = j, include_x2 = FALSE, sigma = 0.05)$
    gen(500, computed = select(dat_shape_1, x1, e)) %>%
  mutate(j = j)
}) %>%

# Combined with data for shape 1
bind_rows(dat_shape_1) %>%
  mutate(j = factor(j)) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light(base_size = 15)) +
  facet_wrap(~j, scales = "free", labeller = label_parsed, ncol = 5)
```

:::

:::

:::: {.columns}

::: {.column width="30%"}

**Factor $\sigma_\varepsilon$**

:::

::: {.column width="70%"}

```{r fig.width=8, fig.height=2}
set.seed(10086)

# Data for shape 1
dat_shape_1 <- phn_model(j = 2, include_x2 = FALSE, sigma = 0.05)$gen(500) %>%
  mutate(e_sigma = 0.2)

# Generate data for shape 2, 3 and 4. Reuse x and e.
map_df(c(0.4, 0.8, 1.6, 3.2), function(e_sigma) {
  phn_model(j = 2, include_x2 = FALSE, sigma = e_sigma)$
    gen(500, computed = select(dat_shape_1, x1, e) %>% mutate(e = e/0.2*e_sigma)) %>%
  mutate(e_sigma = e_sigma)
}) %>%

# Combined with data for shape 1
bind_rows(dat_shape_1) %>%
  mutate(e_sigma = factor(e_sigma)) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light(base_size = 15)) +
  facet_wrap(~e_sigma, scales = "free", labeller = label_parsed, ncol = 5)
```

:::

:::

---

## 💡Heteroskedasticity

:::: {.columns}

::: {.column width="30%"}

**Factor $a$**

:::

::: {.column width="70%"}

```{r fig.width=10, fig.height=4}
set.seed(10085)

# Generate data for a = -1
dat_a_n1 <- phn_model(include_z = FALSE,
                      include_x2 = FALSE,
                      a = -1,
                      b = 100)$gen(500) %>%
  mutate(a = -1)

# Generate data for other a
map(c(-0.75, -0.5, -0.25, 0, 0.25, 0.5, 0.75, 1), function(a) {
  phn_model(include_z = FALSE,
            include_x2 = FALSE,
            a = a,
            b = 100)$gen(500) %>%
  mutate(a = a)
}) %>%

  # Combined with data for a = -1
  bind_rows(dat_a_n1) %>%
  mutate(a = factor(a)) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light(base_size = 15)) +
  facet_wrap(~a, scales = "free", ncol = 5) +
  xlab("Fitted values") +
  ylab("Residuals")
```

:::

:::

:::: {.columns}

::: {.column width="30%"}

**Factor $b$**

:::

::: {.column width="70%"}

```{r fig.width=8, fig.height=2}
set.seed(10085)

# Generate data for a = -1
dat_a_n1 <- phn_model(include_z = FALSE,
                      include_x2 = FALSE,
                      a = -1,
                      b = 6)$gen(500) %>%
  mutate(b = 6)

# Generate data for other a
map(c(3, 1.5, 1, 0.5), function(b) {
  phn_model(include_z = FALSE,
            include_x2 = FALSE,
            a = -1,
            b = b)$gen(500) %>%
  mutate(b = b)
}) %>%

  # Combined with data for a = -1
  bind_rows(dat_a_n1) %>%
  mutate(b = factor(b)) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light(base_size = 15)) +
  facet_wrap(~b, scales = "free", ncol = 5) +
  xlab("Fitted values") +
  ylab("Residuals")
```

:::

:::

---

## 💡Non-normality

:::: {.columns}

::: {.column width="30%"}

**Distribution of $\varepsilon$**

:::

::: {.column width="70%"}

```{r fig.width=8, fig.height=6}
set.seed(10086)

# Data for shape 1
dat_shape_1 <- phn_model(include_z = FALSE, include_x2 = FALSE, e = rand_uniform(-1.4, 1.4))$gen(500) %>%
  mutate(e_dist = "uniform")

dat_shape_2 <- phn_model(include_z = FALSE, include_x2 = FALSE, sigma = 0.8)$gen(500) %>%
  mutate(e_dist = "normal")

dat_shape_3 <- phn_model(include_z = FALSE, include_x2 = FALSE, e = rand_lognormal(sigma = 0.6))$gen(500) %>%
  mutate(e_dist = "lognormal")

dat_shape_4 <- phn_model(include_z = FALSE, include_x2 = FALSE, e = rand_uniform_d(-1.4, 1.4, even = TRUE))$gen(500) %>%
  mutate(e_dist = "discrete")

# Generate data for shape 2, 3 and 4. Reuse x and e.
bind_rows(dat_shape_1, dat_shape_2, dat_shape_3, dat_shape_4) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light(base_size = 15)) +
  facet_wrap(~e_dist, scales = "free", labeller = label_parsed, ncol = 2)
```

:::

:::

---

## 💡Multiple Model Violations

::: {.columns}

::: {.column width="30%"}

**Non-linearity + Heteroskedasticity**

:::

::: {.column width="70%"}

```{r fig.width=10, fig.height=4}
set.seed(10086)

# Data for shape 1
dat_shape_1 <- phn_model(j = 2, a = -1, b = 100, include_x2 = FALSE, sigma = 0.05)$gen(500) %>%
  mutate(j = 2)

# Generate data for shape 2, 3 and 4. Reuse x and e.
map_df(3:11, function(j) {
  phn_model(j = j, a = -1, b = 100, include_x2 = FALSE, sigma = 0.05)$
    gen(500, computed = select(dat_shape_1, x1, e)) %>%
  mutate(j = j)
}) %>%

  # Combined with data for shape 1
  bind_rows(dat_shape_1) %>%
  mutate(j = factor(j)) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light(base_size = 15)) +
  facet_wrap(~j, scales = "free", labeller = label_parsed, ncol = 5)
```

:::

:::

::: {.columns}

::: {.column width="30%"}

**Non-normality + Heteroskedasticity**

:::

::: {.column width="70%"}

```{r fig.width=8, fig.height=2}
set.seed(10085)

# Data for shape 1
dat_shape_1 <- phn_model(a = -1, b = 100, include_z = FALSE, include_x2 = FALSE, e = rand_uniform(-1.4, 1.4))$gen(500) %>%
  mutate(e_dist = "uniform")

dat_shape_2 <- phn_model(a = -1, b = 100, include_z = FALSE, include_x2 = FALSE, sigma = 0.8)$gen(500) %>%
  mutate(e_dist = "normal")

dat_shape_3 <- phn_model(a = -1, b = 100, include_z = FALSE, include_x2 = FALSE, e = rand_lognormal(sigma = 0.6))$gen(500) %>%
  mutate(e_dist = "lognormal")

dat_shape_4 <- phn_model(a = -1, b = 100, include_z = FALSE, include_x2 = FALSE, e = rand_uniform_d(-1.4, 1.4, even = TRUE))$gen(500) %>%
  mutate(e_dist = "discrete")

# Generate data for shape 2, 3 and 4. Reuse x and e.
bind_rows(dat_shape_1, dat_shape_2, dat_shape_3, dat_shape_4) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light(base_size = 15)) +
  facet_wrap(~e_dist, scales = "free", labeller = label_parsed, ncol = 4)
```

:::

:::

---

## 💡Predictor Distribution

::: {.columns}

::: {.column width="30%"}

**Distribution of predictor**

:::

::: {.column width="70%"}

```{r fig.width=8, fig.height=6}
set.seed(10086)

stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1

# Data for uniform distribution
dat_dist_1 <- poly_model(shape = 1, 
                      x = {
                        raw_x <- rand_uniform(-1, 1);
                        closed_form(~stand_dist(raw_x))
                        }, 
                      sigma = 0.5)$gen(300) %>%
  mutate(x_dist = "uniform")

# Generate data for other distributions
dat_dist_2 <- poly_model(shape = 1,
                         x = {
                           raw_x <- rand_normal(sigma = 0.3); 
                           closed_form(~stand_dist(raw_x))
                           }, 
                         sigma = 0.5)$
  gen(300, computed = select(dat_dist_1, e)) %>%
  mutate(x_dist = "normal")

dat_dist_3 <- poly_model(shape = 1,
                         x = {
                           raw_x <- rand_lognormal(sigma = 0.6); 
                           closed_form(~stand_dist(raw_x/3 - 1))
                           }, 
                         sigma = 0.5)$
  gen(300, computed = select(dat_dist_1, e)) %>%
  mutate(x_dist = "lognormal")

dat_dist_4 <- poly_model(shape = 1,
                         x = {
                           raw_x <- rand_uniform_d(k = 5, even = TRUE); 
                           closed_form(~stand_dist(raw_x))
                           }, 
                         sigma = 0.5)$
  gen(300, computed = select(dat_dist_1, e)) %>%
  mutate(x_dist = "discrete")

# Generate and plot data for discrete uniform distribution
bind_rows(dat_dist_1, dat_dist_2, dat_dist_3, dat_dist_4) %>%
  mutate(x_dist = factor(x_dist, 
                         levels = c("discrete", 
                                    "lognormal", 
                                    "normal", 
                                    "uniform"))) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light(base_size = 15)) +
  xlab("Fitted values") +
  ylab("Residuals") +
  facet_wrap(~x_dist, ncol = 2, scales = "free")
```

:::

:::

---

## 💡Second Predictor

::: {.columns}

::: {.column width="30%"}

**Non-linearity + Second predictor**

:::

::: {.column width="70%"}

```{r fig.width=10, fig.height=4}
set.seed(10086)

# Data for shape 1
dat_shape_1 <- phn_model(j = 2, include_x2 = TRUE, sigma = 0.05)$gen(500) %>%
  mutate(j = 2)

# Generate data for shape 2, 3 and 4. Reuse x and e.
map_df(3:11, function(j) {
  phn_model(j = j, include_x2 = TRUE, sigma = 0.05)$
    gen(500, computed = select(dat_shape_1, x1, e)) %>%
  mutate(j = j)
}) %>%

  # Combined with data for shape 1
  bind_rows(dat_shape_1) %>%
  mutate(j = factor(j)) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light(base_size = 15)) +
  facet_wrap(~j, scales = "free", labeller = label_parsed, ncol = 5)
```

:::

:::

---

## 🏛️Model Architecture

The architecture of the computer vision model is adapted from **VGG16** (Simonyan and Zisserman 2014).

```{r fig.align='center'}
magick::image_read_pdf("figures/cvm.pdf", pages = 1)
```

---

## 🏋️‍♂️Model Training

The computer vision model is trained on the **M3 high-performance computing platform** (www.massive.org.au), using **TensorFlow** (Abadi et al. 2016) and **Keras** (Chollet et al. 2015).

- The training, validation and test set contains 64000, 16000 and 8000 images respectively.

- The distribution of the target variable $D$ is controlled such that it roughly follows a uniform distribution.

- Multiple models with different image resolutions are trained, the optimized model has input size $32 \times 32$.

---

## 🌐Overview of Model Performance

::: {.columns}

::: {.column width="40%"}

::: {style="font-size:80%"}

The model performs

- better for non-normality
- worse for non-linearity
- consistently across different datasets

::: {.fragment}

The model performs worst on null plots (D = 0) as expected.

:::

:::

:::

::: {.column width="60%"}

```{r}
library(glue)
test_pred <- readr::read_csv(here::here("data/test_pred.csv"))
train_pred <- readr::read_csv(here::here("data/train_pred.csv"))
meta <- readr::read_csv(here::here("data/meta.csv"))

model_pred <- train_pred %>% 
  left_join(meta) %>%
  mutate(type = "train") %>%
  bind_rows(test_pred %>% 
              left_join(meta) %>%
              mutate(type = "test")) %>%
  filter(res == 32L)

model_pred %>%
  group_by(type, include_non_normal, include_heter, include_z) %>%
  summarise(n = n(), 
            RMSE = yardstick::rmse_vec(effect_size, vss)) %>%
  mutate(violations = ifelse(include_z, "non-linearity", "null")) %>%
  mutate(violations = ifelse(include_heter, glue("{violations} + heteroskedasticity"), violations)) %>%
  mutate(violations = ifelse(include_non_normal, glue("{violations} + non-normality"), violations)) %>%
  mutate(violations = gsub("null \\+ ", "", violations)) -> model_pred_plot_dat

```

```{r fig.height=6, fig.width=7.2}

grey_grid <- expand.grid(c(0,1,2,3,4,5,6), c(1,2,3))

model_pred_plot_dat <- model_pred_plot_dat %>%
  mutate(violations = factor(violations, levels = c("null", 
                                                    "non-linearity",
                                                    "heteroskedasticity",
                                                    "non-normality",
                                                    "non-linearity + heteroskedasticity",
                                                    "non-linearity + non-normality",
                                                    "heteroskedasticity + non-normality",
                                                    "non-linearity + heteroskedasticity + non-normality")))

ggplot() +
  geom_ribbon(aes(x = seq(-0.5, 7.5, 0.5), ymin = 0.5, ymax = 1.5), fill = "grey90") +
  geom_ribbon(aes(x = seq(-0.5, 7.5, 0.5), ymin = 2.5, ymax = 3.5), fill = "grey90") +
  geom_point(aes(grey_grid$Var1, grey_grid$Var2), col = "grey", size = 5) +
  geom_point(aes(c(1,2,3), c(1,2,3)), size = 5) +
  geom_point(aes(c(4,4,5,5,6,6,7,7,7), c(1,2,1,3,2,3,1,2,3)), size = 5) +
  geom_segment(aes(x = c(4,5,6,7,7), xend = c(4,5,6,7,7), y = c(1,1,2,1,2), yend = c(2,3,3,2,3)), size = 1) +
  geom_text(aes(c(-2, -2, -2), c(1, 2, 3), label = c("Non-linearity", "Heteroskedasticity", "Non-normality"))) +
  geom_segment(aes(x = -0.5, xend = 7.5, y = 3.5, yend = 3.5)) +
  geom_segment(aes(x = -0.5, xend = -0.5, y = 3.5, yend = 3.5 + 1.5 * 5)) +
  geom_segment(aes(x = rep(-0.6, 4), xend = rep(-0.5, 4), y = 3.5 + seq(0, 1.5, 0.5) * 5, yend = 3.5 + seq(0, 1.5, 0.5) * 5)) +
  geom_text(aes(x = rep(-1, 4), y = 3.5 + seq(0, 1.5, 0.5) * 5, label = c("  0", "0.5", "  1", "1.5")), size = 4) +
  geom_point(data = model_pred_plot_dat, aes(as.numeric(violations) - 1, RMSE * 5 + 3.5, col = type), size = 3) +
  geom_text(aes(-1.5, 0.75 * 5 + 3.5, label = "RMSE"), angle = 90) +
  theme_light(base_size = 15) +
  theme(panel.grid = element_blank(), 
        axis.text = element_blank(), 
        axis.ticks = element_blank()) +
  xlab("") +
  ylab("") +
  xlim(c(-4, 8)) +
  labs(col = "Dataset")
```

:::

:::

---

## 📈Model Violations Index

::: {style="font-size:80%"}

For a consistent data generating process, $D$ typically **increases logarithmically with the number of observations**.

This behaviour comes from the relationship $D = \text{log}(1 + D_{KL})$, where $D_{KL} = \sum_{i=1}^{n}D_{KL}^{(i)}$ under the assumption of independence.

Therefore, the **Model Violations Index** (MVI) for determining the **extent of model violations** can be proposed as

\begin{equation}
\label{eq:mvi}
\text{MVI} = C + \widehat{D} - log(n),
\end{equation}

where $C$ is a large enough constant keeping the result positive.

:::

---

## 📈Model Violations Index

```{r fig.align='center'}
magick::image_read_pdf("figures/poly-index-1.pdf", pages = 1)
```

---

## 📈Model Violations Index

```{r fig.align='center'}
magick::image_read_pdf("figures/heter-index-1.pdf", pages = 1)
```

---

## 🔬Statistical Testing Based on the Approximated Distance

::: {style="font-size:80%"}

We have proposed a statistical test based on the approximated distance

- The **null distribution** can be estimated by predicting approximated distance for **a large number of null plots**.

- The **critical value** can be estimated by the **sample quantile** (e.g. $Q_{null}(0.95)$) of the null distribution.

- The **$p$-value** is the proportion of null plots having approximated distance **greater than or equal to** the observed one.

:::

---

## 🔬Statistical Testing Based on the Approximated Distance

::: {.columns}

::: {.column width="100%"}

```{r}
my_vi <- autovi::auto_vi(lm(mpg ~ ., data = mtcars))
my_vi$check_result <- readRDS("data/test.rds")

p1 <- my_vi$plot_resid()
p2 <- my_vi$summary_plot() +
  ggtitle("", subtitle = glue("P-value = {my_vi$check_result$p_value}")) +
  theme(legend.position = "bottom", legend.box = "vertical") +
  labs(linetype = "") +
  xlab(expression(hat(D))) +
  scale_linetype(labels = c("95% quantile of the null distribution", expression(hat(D))))

patchwork::wrap_plots(p1, p2)
```

:::

:::

---

## 🧪Human Subject Experiment

We have collected $7974$ responses to $1152$ lineups in a **human subject experiment** conducted in the first project (Li et al. 2023).

Differences:

- non-linearity and heteroskedasticity **do not co-exist in one lineup**.
- non-normality and multiple predictors **are not considered** in the experimental design.

---

## 🧪Compared to Visual Tests

```{r fig.align='center'}
magick::image_read_pdf("figures/human-mosaic-1.pdf", pages = 1)
```

---

## 🧪Compared to RESET and BP Tests

```{r fig.align='center'}
magick::image_read_pdf("figures/conv-mosaic-1.pdf", pages = 1)
```

---

## 🧪Power Comparison

```{r fig.align='center'}
magick::image_read_pdf("figures/power-1.pdf", pages = 1)
```

---

## 💡Example: Left-triangle

```{r fig.align='center'}
magick::image_read_pdf("figures/false-check-1.pdf", pages = 1)
```

## 💡Example: Left-triangle

```{r fig.align='center'}
magick::image_read_pdf("figures/false-lineup-1.pdf", pages = 1)
```

## 💡Example: Boston Housing

$MEDV = \beta_0 + \beta_1RM + \beta_2LSTAT + \beta_3PTRATIO + \varepsilon$

```{r fig.align='center'}
magick::image_read_pdf("figures/boston-check-1.pdf", pages = 1)
```

## 💡Example: Boston Housing

```{r fig.align='center'}
magick::image_read_pdf("figures/boston-lineup-1.pdf", pages = 1)
```

## 💡Example: Dinosaur

$y = \beta_0 + \beta_1x + \varepsilon$

```{r fig.align='center'}
magick::image_read_pdf("figures/dino-check-1.pdf", pages = 1)
```

## 💡Example: Dinosaur

```{r fig.align='center'}
magick::image_read_pdf("figures/dino-lineup-1.pdf", pages = 1)
```

---

## 🎬Conclusions

::: {style="font-size:80%"}

In this study, we have

- introduced a **distance measure** to effectively **captures the magnitude of model violations**

- proposed and trained a **computer vision model** to **approximate this distance**

- developed a **statistical testing procedure** and a **Model Violations Index** based on the approximated distance

:::

---

## 🎬Conclusions

::: {style="font-size:80%"}

- The proposed test shows **comparable power** to conventional ones, but there's room for further **enhancing its robustness** against **minor deviations** from model assumptions.

- Our method offers significant value by **easing analysts' workload** in **assessing residual plots**.

- While we encourage **continued manual review of residual plots** for **insightful analysis**, our approach serves as a valuable tool for **automating** or **supplementing** the diagnostic process.

:::

---

## 📅Timeline

```{r}
data.frame(Date = c("Mar",
                    "April", 
                    "May",
                    "July",
                    "",
                    "Aug"), 
           Event = c("Finish and submit the second paper",
                     "Develop a web interface for the automatic visual inference system",
                     "Clean and publish R packages to CRAN",
                     "Finish the third paper",
                     "Attend the useR! conference",
                     "Submit thesis")) %>%
  kableExtra::kable(booktabs = TRUE) %>%
  kableExtra::kable_styling() 
```

## Thanks! Any questions? {.center}

<br>

### 🔗Relevant links

::: {.nonincremental}

- 📦 R Packages
  - [`visage`: Visual Inference for Linear Regression Diagnostics](https://github.com/TengMCing/visage){target="_blank"}
  - [`autovi`: Automated Visual Inference with Computer Vision Models](https://github.com/TengMCing/autovi){target="_blank"}

- ![](figures/github-mark/github-mark.png){height=1em width=1em style="vertical-align:middle"} [Slides](https://github.com/TengMCing/PhD/Milestone_presentation_3){target="_blank"}

:::

### 2024-06-24 Thesis/02-chap2.qmd
finish typesetting chap 2

# A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol {#paper-1}

Regression experts consistently recommend plotting residuals for model diagnosis, despite the availability of many numerical hypothesis test procedures designed to use residuals to assess problems with a model fit. Here we provide evidence for why this is good advice using data from a visual inference experiment. We show how conventional tests are too sensitive, which means that too often the conclusion would be that the model fit is inadequate. The experiment uses the lineup protocol which puts a residual plot in the context of null plots. This helps generate reliable and consistent reading of residual plots for better model diagnosis. It can also help in an obverse situation where a conventional test would fail to detect a problem with a model due to contaminated data. The lineup protocol also detects a range of departures from good residuals simultaneously. Supplemental materials for the article are available online.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  echo = FALSE,
  fig.width = 8,
  fig.height = 6,
  out.width = "100%", 
  fig.align = "center")
```

```{r}
if (!requireNamespace("remotes", quietly = TRUE)) {
  install.packages("remotes")
}

# OOP supports needed by `visage`
if (!requireNamespace("bandicoot", quietly = TRUE)) {
  remotes::install_github("TengMCing/bandicoot")
}

# Visual inference models and p-value calculation
if (!requireNamespace("visage", quietly = TRUE)) {
  remotes::install_url("https://github.com/TengMCing/visage/raw/master/built/visage_0.1.2.tar.gz")
} else {
  if (packageVersion("visage") < "0.1.2") {
    remotes::install_url("https://github.com/TengMCing/visage/raw/master/built/visage_0.1.2.tar.gz")
  }
}

if (!requireNamespace("tidyverse", quietly = TRUE)) {
  install.packages("tidyverse")
}

if (!requireNamespace("here", quietly = TRUE)) {
  install.packages("here")
}

if (!requireNamespace("ggmosaic", quietly = TRUE)) {
  install.packages("ggmosaic")
}

if (!requireNamespace("rcartocolor", quietly = TRUE)) {
  install.packages("rcartocolor")
}

if (!requireNamespace("kableExtra", quietly = TRUE)) {
  install.packages("kableExtra")
}

if (!requireNamespace("patchwork", quietly = TRUE)) {
  install.packages("patchwork")
}

library(tidyverse)
library(visage)

# To control the simulation in this file
set.seed(10086)
```

```{r get-lineup-data}
# Create the dir folder
if (!dir.exists(here::here("cached_data/"))) dir.create(here::here("cached_data/"))

# The lineup data used to draw residual plot needed to be downloaded
# from the github repo. Cache it.
if (!file.exists(here::here("cached_data/vi_lineup.rds"))) {
  vi_lineup <- get_vi_lineup()
  saveRDS(vi_lineup, here::here("cached_data/vi_lineup.rds"))
} else {
  vi_lineup <- readRDS(here::here("cached_data/vi_lineup.rds"))
}
```

```{r poly-conventional-simulation}
# Ensure the support of the predictor is [-1, 1]
stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1

# Run simulations to get the behaviours of conventional tests
if (!file.exists(here::here("cached_data/poly_conventional_simulation.rds"))) {
  poly_conv_sim <- list()

  # 100 bootstrap samples
  for (i in 1:100)
  {
    poly_conv_sim[[i]] <- 
      # Every sample contains 2000 lineups
      map(1:2000, function(i) {

        # Sample a set of parameters
        shape <- sample(1:4, 1)
        e_sigma <- sample(c(0.5, 1, 2, 4), 1)
        x_dist <- sample(c("uniform", 
                           "normal", 
                           "lognormal", 
                           "even_discrete"), 1)
        x <- switch(x_dist,
                    uniform = rand_uniform(-1, 1),
                    normal = {
                      raw_x <- rand_normal(sigma = 0.3)
                      closed_form(~stand_dist(raw_x))
                      },
                    lognormal = {
                      raw_x <- rand_lognormal(sigma = 0.6)
                      closed_form(~stand_dist(raw_x/3 - 1))
                      },
                    even_discrete = rand_uniform_d(k = 5, even = TRUE))

        # Build the model
        mod <- poly_model(shape, x = x, sigma = e_sigma)

        # Sample the number of observations
        n <- sample(c(50, 100, 300), 1)

        # Generate data from the model 
        tmp_dat <- mod$gen(n)

        # Return a data frame containing p-values of
        # F, RESET, BP and SW tests
        tibble(shape = shape,
               e_sigma = e_sigma,
               x_dist = x_dist,
               n = n,
               F_p_value = mod$test(tmp_dat)$p_value,
               RESET3_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:3, 
                                         power_type = "fitted")$p_value,
               RESET4_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:4, 
                                         power_type = "fitted")$p_value,
               RESET5_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:5, 
                                         power_type = "fitted")$p_value,
               RESET6_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:6, 
                                         power_type = "fitted")$p_value,
               RESET7_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:7, 
                                         power_type = "fitted")$p_value,
               RESET8_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:8, 
                                         power_type = "fitted")$p_value,
               RESET9_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:9, 
                                         power_type = "fitted")$p_value,
               RESET10_p_value = mod$test(tmp_dat, 
                                          test = "RESET", 
                                          power = 2:10, 
                                          power_type = "fitted")$p_value,
               BP_p_value = HETER_MODEL$test(tmp_dat)$p_value,
               SW_p_value = shapiro.test(tmp_dat$.resid)$p.value,
               boot_id = i)
        }) %>%
  reduce(bind_rows)
  }

  poly_conv_sim <- poly_conv_sim %>%
    reduce(bind_rows)

  saveRDS(poly_conv_sim, here::here("cached_data/poly_conventional_simulation.rds"))
} else {
  poly_conv_sim <- readRDS(here::here("cached_data/poly_conventional_simulation.rds"))
}
```

```{r heter-conventional-simulation}
# Run simulations to get the behaviours of conventional tests
if (!file.exists(here::here("cached_data/heter_conventional_simulation.rds"))) {
  heter_conv_sim <- list()

  # 100 bootstrap samples
  for (i in 1:100)
  {
    heter_conv_sim[[i]] <-
      # Every sample contains 2000 lineups
      map(1:2000, function(x) {

        # Sample a set of parameters
        a <- sample(c(-1, 0, 1), 1)
        b <- sample(c(0.25, 1, 4, 16, 64), 1)
        x_dist <- sample(c("uniform", 
                           "normal", 
                           "lognormal", 
                           "even_discrete"), 1)
        x <- switch(x_dist,
                    uniform = rand_uniform(-1, 1),
                    normal = {
                      raw_x <- rand_normal(sigma = 0.3)
                      closed_form(~stand_dist(raw_x))
                      },
                    lognormal = {
                      raw_x <- rand_lognormal(sigma = 0.6)
                      closed_form(~stand_dist(raw_x/3 - 1))
                      },
                    even_discrete = rand_uniform_d(-1, 1, k = 5, even = TRUE))

        # Build the model
        mod <- heter_model(a = a, b = b, x = x)

        # Sample the number of observations
        n <- sample(c(50, 100, 300), 1)

        # Generate data from the model 
        tmp_dat <- mod$gen(n)

        # Return a data frame containing p-values of
        # F, RESET, BP and SW tests
        tibble(a = a,
               b = b,
               x_dist = x_dist,
               n = n,
               F_p_value = POLY_MODEL$test(

                 # Create a pseudo z to be able to use F-test
                 tmp_dat %>%
                   mutate(z = poly_model()$
                            gen(n, computed = select(tmp_dat, x)) %>%
                            pull(z))
                 )$p_value,
               RESET3_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:3, 
                                                power_type = "fitted")$p_value,
               RESET4_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:4, 
                                                power_type = "fitted")$p_value,
               RESET5_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:5, 
                                                power_type = "fitted")$p_value,
               RESET6_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:6, 
                                                power_type = "fitted")$p_value,
               RESET7_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:7, 
                                                power_type = "fitted")$p_value,
               RESET8_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:8, 
                                                power_type = "fitted")$p_value,
               RESET9_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:9, 
                                                power_type = "fitted")$p_value,
               RESET10_p_value = POLY_MODEL$test(tmp_dat, 
                                                 test = "RESET", 
                                                 power = 2:10, 
                                                 power_type = "fitted")$p_value,
               BP_p_value = mod$test(tmp_dat)$p_value,
               SW_p_value = shapiro.test(tmp_dat$.resid)$p.value,
               boot_id = i)
        }) %>%
  reduce(bind_rows)
  }

  heter_conv_sim <- heter_conv_sim %>%
    reduce(bind_rows)

  saveRDS(heter_conv_sim, here::here("cached_data/heter_conventional_simulation.rds"))
} else {
  heter_conv_sim <- readRDS(here::here("cached_data/heter_conventional_simulation.rds"))
}
```

```{r}
# Borrow effect size from the survey
poly_conv_sim <- poly_conv_sim %>%
  left_join(select(filter(vi_survey, type == "polynomial"), shape, e_sigma, n, x_dist, effect_size))

heter_conv_sim <- heter_conv_sim %>%
  left_join(select(filter(vi_survey, type == "heteroskedasticity"), a, b, n, x_dist, effect_size))
```

## Introduction {#sec-introduction}

> *"Since all models are wrong the scientist must be alert to what is importantly wrong."* [@box1976science]

Diagnosing a model is an important part of building an appropriate model. In linear regression analysis, studying the residuals from a model fit is a common diagnostic activity. Residuals summarise what is not captured by the model, and thus provide the capacity to identify what might be wrong.

We can assess residuals in multiple ways. To examine the univariate distribution, residuals may be plotted as a histogram or normal probability plot. Using the classical normal linear regression model as an example, if the distribution is symmetric and unimodal, we would consider it to be well-behaved. However, if the distribution is skewed, bimodal, multimodal, or contains outliers, there would be cause for concern. We can also inspect the distribution by conducting a goodness-of-fit test, such as the Shapiro-Wilk normality test [@shapiro1965analysis].

Scatterplots of residuals against the fitted values, and each of the explanatory variables, are commonly used to scrutinize their relationships. If there are any visually discoverable associations, the model is potentially inadequate or incorrectly specified. We can also potentially discover patterns not directly connected to a linear model assumption from these residual plots, such as the discreteness or skewness of the fitted values, and outliers. To read residual plots, one looks for noticeable departures from the model such as non-linear pattern or heteroskedasticity. A non-linear pattern would suggest that the model needs to have some additional non-linear terms. Heteroskedasticity suggests that the error is dependent on the predictors, and hence violates the independence assumption. Statistical tests were developed to provide objective assessment, for example, of non-linear patterns [e.g. @ramsey1969tests], and heteroskedasticity  [e.g. @breusch1979simple].

The common wisdom of experts is that plotting the residuals is indispensable for diagnosing model fits [@draper1998applied; @cook1982residuals; @montgomery1982introduction]. The lack of empirical evidence for the ubiquitous advice is *curious*, and is what this article tackles.

Additionally, relying solely on the subjective assessment of a single plot can be problematic. People will almost always see a pattern [see @kahneman2011thinking], so the question that really needs answering is whether any pattern perceived is consistent with randomness, or sampling variability, or noise. Correctly judging whether *no* pattern exists in a residual plot is a difficult task. @loy2021bringing emphasizes that this is especially difficult to teach to new analysts and students, and advocates to the broader use of the lineup protocol [@bujastatistical2009].

The lineup protocol places a data plot in a field of null plots, allowing for a comparison of patterns due purely by chance to what is perceived in the data plot. For residual analysis this is especially helpful for gauging whether there is *no* pattern. (@fig-first-example-lineup shows an example of a lineup of residual plots.) In its strict use, one would insist that the data plot is not seen before seeing the lineup, so that the observer does not know which is the true plot. When used this way, it provides an objective test for data plots. @majumdervalidation2013 validated that results from lineups assessed by human observers performed similarly to conventional tests. One would not use a lineup when a conventional test exists and is adequate because it is more manually expensive to conduct. However, where no adequate conventional test exists, it is invaluable, as shown by @loy2013diagnostic. Here we use the lineup as a vehicle to rigorously explore why experts advise that residual plots are indispensable despite the prevalence of numerical tests.

The paper is structured as follows. @sec-background describes the background on the types of departures that one expects to detect, and outlines a formal statistical process for reading residual plots, called visual inference. @sec-significance-calculation describes the calculation of the statistical significance and power of the test. @sec-experimental-design details the experimental design to compare the decisions made by formal hypothesis testing, and how humans would read diagnostic plots. The results are reported in @sec-results. We conclude with a discussion of the presented work, and ideas for future directions.

```{r fig-first-example-lineup, fig.pos="t!", fig.height = 7, fig.width = 7, fig.cap = "Visual testing is conducted using a lineup, as in the example here. The residual plot computed from the observed data is embedded among 19 null plots, where the residuals are simulated from a standard error model. Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot. A small $p$-value would result from a substantial number selecting the data plot (at position $6$, exhibiting non-linearity)."}

vi_lineup$poly_300$data %>%
  VI_MODEL$plot_lineup(theme = theme_light(), remove_axis = TRUE, remove_grid_line = TRUE)
```

## Background {#sec-background}

### Departures from Good Residual Plots

Graphical summaries where residuals are plotted against fitted values, or other functions of the predictors (expected to be approximately orthogonal to the residuals) are considered to be the most important residual plots by @cook1999applied. [@fig-residual-plot-common-departures]A shows an example of an ideal residual plot where points are symmetrically distributed around the horizontal zero line (red), with no discernible patterns. There can be various types of departures from this ideal pattern. Non-linearity, heteroskedasticity and non-normality, shown in [@fig-residual-plot-common-departures]B, [@fig-residual-plot-common-departures]C, and [@fig-residual-plot-common-departures]D, respectively, are three commonly checked departures.

```{r fig-residual-plot-common-departures, fig.pos="t!", fig.width = 10, fig.height = 2.5, fig.cap = "Example residual vs fitted value plots (horizontal line indicates 0): (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0. The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted value plot."}

set.seed(10086)

# Base model
mod <- poly_model(include_z = FALSE, sigma = 0.25)

# Base data
dat <- mod$gen(300)

bind_rows(
  dat %>%
    mutate(type = "(A) Good residuals"),
  mod$set_prm("include_z", TRUE)$
    gen(300, computed = select(dat, x, e)) %>%
    mutate(type = "(B) Non-linearity"),
  heter_model(b = 64)$
    gen(300, computed = select(dat, x, e)) %>%
    mutate(type = "(C) Heteroskedasticity"),
  heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
    gen(300, computed = select(dat, x)) %>%
    mutate(type = "(D) Non-normality")
) %>%
  mod$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
  facet_wrap(~type, ncol = 4, scales = "free") +
  xlab("Fitted values") +
  ylab("Residuals")
```

Model misspecification occurs if functions of predictors that needed to accurately describe the relationship with the response are incorrectly specified. This includes instances where a higher-order polynomial term of a predictor is wrongfully omitted. Any non-linear pattern visible in the residual plot could be indicative of this problem. An example residual plot containing visual pattern of non-linearity is shown in [@fig-residual-plot-common-departures]B. One can clearly observe the "S-shape" from the residual plot, which corresponds to the cubic term that should have been included in the model.

Heteroskedasticity refers to the presence of non-constant error variance in a regression model. It indicates that the distribution of residuals depends on the predictors, violating the independence assumption. This can be seen in a residual plot as an inconsistent spread of the residuals relative to the fitted values or predictors. An example is the "butterfly" shape  shown in [@fig-residual-plot-common-departures]C, or a "left-triangle" and "right-triangle" shape where the smallest variance occurs at one side of the horizontal axis.

[@fig-residual-plot-common-departures]D shows a scatterplot where the residuals have a skewed distribution, as seen by the uneven vertical spread. Unlike non-linearity and heteroskedasticity, non-normality is usually detected with a different type of residual plot: a histogram or a normal probability plot.  Because we focus on scatterplots, non-normality is not one of the departures examined in depth in this paper. [@loy2016variations discuss related work on non-normality checking.]

### Conventionally Testing for Departures

Many different hypothesis tests are available to detect specific model defects. For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white1980heteroskedasticity] or the Breusch-Pagan (BP) test [@breusch1979simple], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution. To test specific forms of non-linearity, one may apply the F-test as a model structural test to examine the significance of a specific polynomial and non-linear forms of the predictors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey1969tests]. The Shapiro-Wilk (SW) normality test [@shapiro1965analysis] is the most widely used test of non-normality included by many of the statistical software programs. The Jarque-Bera test [@jarque1980efficient] is also used to directly check whether the sample skewness and kurtosis match a normal distribution.

@tbl-example-residual-plot-table displays the $p$-values from the RESET, BP and SW tests applied to the residual plots in @fig-residual-plot-common-departures. The RESET test and BP test were computed using the `resettest` and `bptest` functions from the R package `lmtest`, respectively. The SW test was computed using the `shapiro.test` from the core R package `stats`. ^[Although we did not use it, it is useful to know that the R package `skedastic` [@skedastic] also contains a large collection of functions to test for heteroskedasticity.] The RESET test requires the selection of a power parameter.  @ramsey1969tests recommends a power of four, which we adopted in our analysis.

For residual plots in @fig-residual-plot-common-departures, we would expect the RESET test for non-linearity to reject residual plot B, the BP test for heteroskedasticity to reject the residual plot C, and the SW test for non-normality to reject residual plot D, which they all do and all tests also correctly fail to reject residual plot A. Interestingly, the BP and SW tests also reject the residual plots exhibiting structure that they were not designed for. @cook1982residuals explain that most residual-based tests for a particular departure from the model assumptions are also sensitive to other types of departures. This could be considered a Type III error [@kimball1957errors], where the null hypothesis of good residuals is correctly rejected but for the wrong reason. Also, some types of departure can have elements of other types of departure, for example, non-linearity can appear like heteroskedasticity. Additionally, other data problems such as outliers can trigger rejection (or not) of the null hypothesis [@cook1999applied].

With large sample sizes, hypothesis tests may reject the null hypothesis when there is only a small effect. (A good discussion can be found in @kirk1996practical.) While such rejections may be statistically correct, their sensitivity may render the results impractical. A key goal of residual plot diagnostics is to identify potential issues that could lead to incorrect conclusions or errors in subsequent analyses, but minor defects in the model are unlikely to have a significant impact and may be best disregarded for practical purposes. The experiment discussed in this paper specifically addresses this tension between statistical significance and practical significance.

```{r}
#| label: tbl-example-residual-plot-table
#| tbl-cap: 'Statistical significance testing for departures from good residuals for plots in @fig-residual-plot-common-departures. Shown are the $p$-values calculated for the RESET, the BP and the SW tests. The good residual plot (A) is judged a good residual plot, as expected, by all tests. The non-linearity (B) is detected by all tests, as might be expected given the extreme structure.'

set.seed(10086)

# Base model
mod <- poly_model(include_z = FALSE, sigma = 0.25)

# Base data
dat0 <- mod$gen(300)

# Replicate data in Figure 1
dat1 <- mod$set_prm("include_z", TRUE)$gen(300, computed = select(dat0, x, e))
dat2 <- heter_model(b = 64)$gen(300, computed = select(dat0, x, e))
dat3 <- heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
               gen(300, computed = select(dat0, x))

dat_list <- list(dat0, dat1, dat2, dat3)

table_dat <- data.frame(plot = c("A", "B", "C", "D"), 
                        model = c("None", 
                                  "Non-linearity", 
                                  "Heteroskedasticity", 
                                  "Non-normality"),
                        r = map_dbl(dat_list, 
                                    ~POLY_MODEL$test(.x, 
                                                     test = "RESET", 
                                                     power = 2:4)$p_value),
                        b = map_dbl(dat_list, 
                                    ~HETER_MODEL$test(.x)$p_value),

                        # SW test has not been included by visage
                        s = map_dbl(dat_list, 
                                    ~shapiro.test(.x$.resid)$p.value)) %>%

  # 3 decimal points
  mutate(across(r:s, ~ format(round(.x, digits = 3), nsmall = 3))) %>%

  # Italic if reject
  mutate(across(r:s, ~ kableExtra::cell_spec(.x, italic = as.numeric(.x) <= 0.05)))

  table_dat %>%
  knitr::kable(col.names = c("Plot", 
                             "Departures", 
                             "RESET", 
                             "BP", 
                             "SW"),
               align = "llrrr",
               escape = FALSE,
               booktabs = TRUE) 
```

### Visual Test Procedure based on Lineups {#sec-visual-test-procedure-based-on-lineups}

The examination of data plots to infer signals or patterns (or lack thereof) is fraught with variation in the human ability to interpret and decode the information embedded in a graph [@cleveland1984graphical]. In practice, over-interpretation of a single plot is common.  For instance, @roy2015using described a published example where authors over-interpreted separation between gene groups from a two-dimensional projection of a linear discriminant analysis even when there were no differences in the expression levels between the gene groups.

One solution to over-interpretation is to examine the plot in the context of natural sampling variability assumed by the model, called the lineup protocol, as proposed in @buja2009statistical.  @majumder2013validation showed that the lineup protocol is analogous to the null hypothesis significance testing framework. The protocol consists of $m$ randomly placed plots, where one plot is the data plot, and the remaining $m - 1$ plots, referred to as the _null plots_, are constructed using the same graphical procedure as the data plot but the data is replaced with null data that is generated in a manner consistent with the null hypothesis, $H_0$. Then, an observer who has not seen the data plot is asked to point out the most different plot from the lineup. Under $H_0$, it is expected that the data plot would have no distinguishable difference from the null plots, and the probability that the observer correctly picks the data plot is $1/m$. If one rejects $H_0$ as the observer correctly picks the data plot, then the Type I error of this test is $1/m$. This protocol requires a priori specification of $H_0$ (or at least a null data generating mechanism), much like the requirement of knowing the sampling distribution of the test statistic in null hypothesis significance testing framework.

Figure @fig-first-example-lineup is an example of a lineup protocol. If the data plot at position $6$ is identifiable, then it is evidence for the rejection of $H_0$. In fact, the actual residual plot is obtained from a misspecified regression model with missing non-linear terms.

Data used in the $m - 1$ null plots needs to be simulated. In regression diagnostics, sampling data consistent with $H_0$ is equivalent to sampling data from the assumed model. As @buja2009statistical suggested, $H_0$ is usually a composite hypothesis controlled by nuisance parameters. Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to a so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling. The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja2009statistical]. Under this method, the null residuals can essentially be simulated by independent drawing from  a standard normal random distribution, then regressing the draws on the predictors, and then re-scaling it by the ratio of the residual sum of square in two regressions.

The effectiveness of lineup protocol for regression analysis has been validated by @majumder2013validation under relatively simple settings with up to two predictors. Their results suggest that visual tests are capable of testing the significance of a single predictor with a similar power to a t-test, though they express that in general it is unnecessary to use visual inference if there exists a corresponding conventional test, and they do not expect the visual test to perform equally well as the conventional test. In their third experiment, where the contamination of the data violate the assumptions of the conventional test, visual test outperforms the conventional test by a large margin. This supports the use of visual inference in situations where there are no existing numerical testing procedures. Visual inference has also been integrated into diagnostics for hierarchical linear models where the lineup protocol is used to judge the assumptions of linearity, normality and constant error variance for both the level-1 and level-2 residuals [@loy2013diagnostic; -@loy2014hlmdiag; -@loy2015you].

## Calculation of Statistical Significance and Test Power {#sec-significance-calculation}

### What is Being Tested?

In diagnosing a model fit using the residuals, we are generally interested in testing whether "*the regression model is correctly specified*" ($H_0$) against the broad alternative  "*the regression model is misspecified*" ($H_a$). However, it is practically impossible to test this broad $H_0$ with conventional tests, because they need specific structure causing the departure to be quantifiable in order to be computable.  For example, the RESET test for detecting non-linear departures is formulated by fitting $y = \tau_0 + \sum_{i=1}^{p}\tau_px_p +\gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)$ in order to test $H_0:\gamma_1 = \gamma_2 = \gamma_3 = 0$ against $H_a: \gamma_1 \neq 0 \text{ or } \gamma_2 \neq 0 \text{ or } \gamma_3 \neq 0$. 
Similarly, the BP test is designed to specifically test $H_0:$ *error variances are all equal* ($\zeta_i=0 \text{ for } i=1,..,p$) versus the alternative $H_a:$ *that the error variances are a multiplicative function of one or more variables* ($\text{at least one } \zeta_i\neq 0$) from $e^2 = \zeta_0 + \sum_{i=1}^{p}\zeta_i x_i + u, ~ u\sim N(0,\sigma_u^2)$.

While a battery of conventional tests for different types of departures could be applied, this is intrinsic to the lineup protocol. The lineup protocol operates as an omnibus test,  able to detect a range of departures from good residuals in a single application.

### Statistical Significance\label{sig}

In hypothesis testing, a $p$-value is defined as the probability of observing test results at least as extreme as the observed result assuming $H_0$ is true. Conventional hypothesis tests usually have an existing method to derive or compute the $p$-value based on the null distribution. The method to estimate a $p$-value for a visual test essentially follows the process detailed by @vanderplas2021statistical. Details are given in Appendix [-@sec-appendix-a-1].

### Power of the Tests

The power of a model misspecification test is the probability that $H_0$ is rejected given the regression model is misspecified in a specific way. It is an important indicator
when one is concerned about whether model assumptions have been violated. In practice, one might be more interested in knowing how much the residuals deviate from the model assumptions, and whether this deviation is of practical significance.

The power of a conventional hypothesis test is affected by both the true parameters $\boldsymbol{\theta}$ and the sample size $n$. These two can be quantified in terms of effect size $E$ to measure the strength of the residual departures from the model assumptions. Details about the calculation of effect size are provided in @sec-effect-size after the introduction of the simulation model used in our experiment. The theoretical power of a test is sometimes not a trivial solution, but it can be estimated if the data generating process is known. We use a predefined model to generate a large set of simulated data under different effect sizes, and record if the conventional test rejects $H_0$. The probability of the conventional test rejects $H_0$ is then fitted by a logistic regression formulated as

$$
Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(\log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right),
$$ {#eq-logistic-regression-1-1}

\noindent where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = \exp(z)(1+\exp(z))^{-1}$. The effect size $E$ is the only predictor and the intercept is fixed to $\log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$, the desired significance level.

The power of a visual test on the other hand, may additionally depend on the ability of the particular participant, as the skill of each individual may affect the number of observers who identify the data plot from the lineup [@majumder2013validation]. To address this issue, @majumder2013validation models the probability of participant $j$ correctly picking the data plot from lineup $l$ using a mixed-effect logistic regression, with participants treated as random effects. Then, the estimated power of a visual test evaluated by a single participant is the predicted value obtained from the mixed effects model. However, this mixed effects model does not work with scenario where participants are asked to select one or more most different plots. In this scenario, having the probability of a participant $j$ correctly picking the data plot from a lineup $l$ is insufficient to determine the power of a visual test because it does not provide information about the number of selections made by the participant for the calculation of the $p$-value. Therefore, we directly estimate the probability of a lineup being rejected by assuming that individual skill has negligible effect on the variation of the power. This assumption essentially averages out the subject ability and helps to simplify the model structure, thereby obviating a costly large-scale experiment to estimate complex covariance matrices. The same model given in Equation @eq-logistic-regression-1-1 is applied to model the power of a visual test.

To study various factors contributing to the power of both tests, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors. These include the distribution of the fitted values, type of the simulation model and the shape of the residual departures.

## Experimental Design {#sec-experimental-design}

Our experiment was conducted over three data collection periods to investigate the difference between conventional hypothesis testing and visual inference in the application of linear regression diagnostics. Two types of departures, non-linearity and heteroskedasticity, were collected during data collection periods I and II. The data collection period III was designed primarily to measure human responses to null lineups so that the visual $p$-values can be estimated. Additional lineups for both non-linearity and heteroskedasticity, using uniform fitted value distributions, were included for additional data, and to avoid participant frustration of too many difficult tasks.

During the experiment, every participant recruited from the Prolific crowd-sourcing platform [@palan2018prolific] was presented with a block of 20 lineups. A lineup consisted of a randomly placed data plot and 19 null plots, which were all residual plots drawn with raw residuals on the y-axis and fitted values on the x-axis. An additional horizontal red line was added at $y = 0$ as a visual reference. The data in the data plot was simulated from one of two models described in @sec-simulating-departures-from-good-residuals, while the data of the remaining 19 null plots were generated by the residual rotation technique discussed in Section @sec-visual-test-procedure-based-on-lineups.

In each lineup evaluation, the participant was asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots are from others. If there is no noticeable difference between plots in a lineup, participants had the option to select zero plots without the need to provide a reason. During the process of recording the responses, a zero selection was considered to be equivalent to selecting all 20 plots. No participant was shown the same lineup twice. Information about preferred pronouns, age group, education, and previous experience in visual experiments were also collected. A participant's submission was only included in the analysis if the data plot is identified for at least one attention check.

Overall, we collected `r vi_survey  %>% filter(!attention_check) %>% nrow()` evaluations on `r vi_survey  %>% filter(!attention_check) %>% pull(unique_lineup_id) %>% unique() %>% length()` unique lineups performed by `r vi_survey  %>% filter(!attention_check) %>% count(exp, set) %>% nrow()` participants. A summary of the factors used in the experiment can be found in @tbl-model-factor-table. There were four levels of the non-linear structure, and three levels of heteroskedastic structure. The signal strength was controlled by error variance ($\sigma$) for the non-linear pattern, and by a ratio ($b$) parameter for the heteroskedasticity. Additionally, three levels of sample size ($n$) and four different fitted value distributions were incorporated.

::: {.content-visible when-format="html"}

::: {#tbl-model-factor-table .cell layout-align="center" tbl-cap='Levels of the factors used in data collection periods I, II, and III.'}
::: {.cell-output-display}

```{=html}
<table style="margin-left: auto; margin-right: auto;">
<thead>
<tr>
<th style="border-bottom:hidden;padding-bottom:0; padding-left:3px;padding-right:3px;text-align: center; " colspan="2"><div style="border-bottom: 1px solid #ddd; padding-bottom: 5px; ">Non-linearity</div></th>
<th style="border-bottom:hidden;padding-bottom:0; padding-left:3px;padding-right:3px;text-align: center; " colspan="2"><div style="border-bottom: 1px solid #ddd; padding-bottom: 5px; ">Heteroskedasticity</div></th>
<th style="border-bottom:hidden;padding-bottom:0; padding-left:3px;padding-right:3px;text-align: center; " colspan="2"><div style="border-bottom: 1px solid #ddd; padding-bottom: 5px; ">Common</div></th>
</tr>
<tr>
<th style="border-bottom:hidden;padding-bottom:0; padding-left:3px;padding-right:3px;text-align: center; " colspan="1"><div style="border-bottom: 1px solid #ddd; padding-bottom: 5px; ">Poly Order (j)</div></th>
<th style="border-bottom:hidden;padding-bottom:0; padding-left:3px;padding-right:3px;text-align: center; " colspan="1"><div style="border-bottom: 1px solid #ddd; padding-bottom: 5px; ">SD (σ)</div></th>
<th style="border-bottom:hidden;padding-bottom:0; padding-left:3px;padding-right:3px;text-align: center; " colspan="1"><div style="border-bottom: 1px solid #ddd; padding-bottom: 5px; ">Shape (a)</div></th>
<th style="border-bottom:hidden;padding-bottom:0; padding-left:3px;padding-right:3px;text-align: center; " colspan="1"><div style="border-bottom: 1px solid #ddd; padding-bottom: 5px; ">Ratio (b)</div></th>
<th style="border-bottom:hidden;padding-bottom:0; padding-left:3px;padding-right:3px;text-align: center; " colspan="1"><div style="border-bottom: 1px solid #ddd; padding-bottom: 5px; ">Size (n)</div></th>
<th style="border-bottom:hidden;padding-bottom:0; padding-left:3px;padding-right:3px;text-align: center; " colspan="1"><div style="border-bottom: 1px solid #ddd; padding-bottom: 5px; ">Distribution of the fitted values</div></th>
</tr>
</thead>
<tbody>
  <tr>
   <td style="text-align:right;"> 2 </td>
   <td style="text-align:right;"> 0.25 </td>
   <td style="text-align:right;"> -1 </td>
   <td style="text-align:right;"> 0.25 </td>
   <td style="text-align:right;"> 50 </td>
   <td style="text-align:center;"> Uniform </td>
  </tr>
  <tr>
   <td style="text-align:right;"> 3 </td>
   <td style="text-align:right;"> 1.00 </td>
   <td style="text-align:right;"> 0 </td>
   <td style="text-align:right;"> 1.00 </td>
   <td style="text-align:right;"> 100 </td>
   <td style="text-align:center;"> Normal </td>
  </tr>
  <tr>
   <td style="text-align:right;"> 6 </td>
   <td style="text-align:right;"> 2.00 </td>
   <td style="text-align:right;"> 1 </td>
   <td style="text-align:right;"> 4.00 </td>
   <td style="text-align:right;"> 300 </td>
   <td style="text-align:center;"> Skewed </td>
  </tr>
  <tr>
   <td style="text-align:right;"> 18 </td>
   <td style="text-align:right;"> 4.00 </td>
   <td style="text-align:right;">  </td>
   <td style="text-align:right;"> 16.00 </td>
   <td style="text-align:right;">  </td>
   <td style="text-align:center;"> Discrete </td>
  </tr>
  <tr>
   <td style="text-align:right;">  </td>
   <td style="text-align:right;">  </td>
   <td style="text-align:right;">  </td>
   <td style="text-align:right;"> 64.00 </td>
   <td style="text-align:right;">  </td>
   <td style="text-align:center;">  </td>
  </tr>
</tbody>
</table>
```

:::
:::

:::

::: {.content-visible when-format="pdf"}

::: {#tbl-model-factor-table .cell layout-align="center" tbl-cap='Levels of the factors used in data collection periods I, II, and III.'}
::: {.cell-output-display}
\begin{tabular}[t]{rr|rr|rcrr|rr|rcrr|rr|rcrr|rr|rcrr|rr|rcrr|rr|rc}
\toprule
\multicolumn{2}{c}{Non-linearity} & \multicolumn{2}{c}{Heteroskedasticity} & \multicolumn{2}{c}{Common} \\
\cmidrule(l{3pt}r{3pt}){1-2} \cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6}
\multicolumn{1}{c}{Poly Order ($j$)} & \multicolumn{1}{c}{SD ($\sigma$)} & \multicolumn{1}{c}{Shape ($a$)} & \multicolumn{1}{c}{Ratio ($b$)} & \multicolumn{1}{c}{Size ($n$)} & \multicolumn{1}{c}{Distribution of the fitted values} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
2 & 0.25 & -1 & 0.25 & 50 & Uniform\\
3 & 1.00 & 0 & 1.00 & 100 & Normal\\
6 & 2.00 & 1 & 4.00 & 300 & Skewed\\
18 & 4.00 &  & 16.00 &  & Discrete\\
 &  &  & 64.00 &  & \\
\bottomrule
\end{tabular}
:::
:::

:::

### Simulating Departures from Good Residuals {#sec-simulating-departures-from-good-residuals}

#### Non-linearity and Heteroskedasticity

Data collection period I was designed to study the ability of participants to detect non-linearity departures from residual plots. The non-linearity departure was constructed by omitting a $j$th order Hermite polynomial [@hermite1864nouveau; originally by @de1820theorie] term of the predictor from the simple linear regression equation. Four different values of $j = 2, 3, 6, 18$ were chosen so that distinct shapes of non-linearity were included in the residual plots. These include "U", "S", "M" and "triple-U" shape as shown in @fig-different-shape-of-herimite. A greater value of $j$ will result in a curve with more turning points. It is expected that the "U" shape will be the easiest to detect, and as the shape gets more complex it will be harder to perceive in a scatterplot, particularly when there is noise. @fig-different-sigma shows the "U" shape for different amounts of noise ($\sigma$).

Data collection period II was similar to period I but focuses on heteroskedasticity departures. We generated the heteroskedasticity departures by setting the variance-covariance matrix of the error term as a function of the predictor, but fitted the data with the simple linear regression model, intentionally violated the constant variance assumption. Visual patterns of heteroskedasticity are simulated using three different shapes ($a$ = -1, 0, 1) including "left-triangle", "butterfly" and "right-triangle" shapes as displayed in @fig-different-shape-of-heter. @fig-different-b shows the butterfly shape as the ratio parameter ($b$) is changed. More details about the simulation process are provided in Appendix [-@sec-appendix-a-2].

```{r fig-different-shape-of-herimite, fig.pos="!h", fig.height = 2, fig.cap = "Polynomial forms generated for the residual plots used to assess detecting non-linearity. The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$."}

# Facet label
shape_labels <- c('He[2]:"U"', 'He[3]:"S"', 'He[6]:"M"', 'He[18]:"triple-U"')

# Data for shape 1
dat_shape_1 <- poly_model(shape = 1, 
                          x = {
                            raw_x <- rand_uniform(-1, 1);
                            closed_form(~stand_dist(raw_x))
                            }, 
                          sigma = 0.05)$gen(300) %>%
  mutate(shape = shape_labels[1])

# Generate data for shape 2, 3 and 4. Reuse x and e.
map_df(2:4, function(shape) {
  poly_model(shape = shape, 
                  x = {
                    raw_x <- rand_uniform(-1, 1); 
                    closed_form(~stand_dist(raw_x))
                    }, 
                  sigma = 0.05)$
  gen(300, computed = select(dat_shape_1, x, e)) %>%
  mutate(shape = shape_labels[shape])
}) %>%

  # Combined with data for shape 1
  bind_rows(dat_shape_1) %>%
  mutate(shape = factor(shape, levels = shape_labels)) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
  facet_wrap(~shape, scales = "free", labeller = label_parsed, ncol = 4) +
  xlab("Fitted values") +
  ylab("Residuals")
```

```{r fig-different-sigma, fig.pos="!h", fig.height = 2, fig.cap = 'Examining the effect of $\\sigma$ on the signal strength in the non-linearity detection, for $n=300$, uniform fitted value distribution and the "U" shape. As $\\sigma$ increases the signal strength decreases, to the point that the "U" is almost unrecognisable when $\\sigma=4$.'}

# Generate data for sigma 0.5
dat_sigma_05 <- poly_model(shape = 1, 
                           x = {
                             raw_x <- rand_uniform(-1, 1);
                             closed_form(~stand_dist(raw_x))
                             }, 
                           sigma = 0.5)$gen(300) %>%
  mutate(sigma = "sigma: 0.5")

# Generate data for other sigma
map(c(1, 2, 4), function(sigma) {
  poly_model(shape = 1, 
                  x = {
                    raw_x <- rand_uniform(-1, 1); 
                    closed_form(~stand_dist(raw_x))
                    }, 
                  sigma = sigma)$
  gen(300, computed = select(dat_sigma_05, x)) %>%
  mutate(sigma = paste0("sigma: ", sigma))
}) %>%

  # Combined with data for sigma 0.5
  bind_rows(dat_sigma_05) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) + 
  facet_wrap(~sigma, ncol = 4, scales = "free", labeller = label_parsed) +
  xlab("Fitted values") +
  ylab("Residuals")
```

```{r fig-different-shape-of-heter, fig.pos="!h", fig.height = 2.67, fig.cap = 'Heteroskedasticity forms used in the experiment. Three different shapes ($a = -1, 0, 1$) are used in the experiment to create "left-triangle", "butterfly" and "right-triangle" shapes, respectively.'}

set.seed(10086)

a_labels <- c("Left-triangle", "Butterfly", "Right-triangle")

# Generate data for a = -1
dat_a_n1 <- heter_model(a = -1, 
                        x = {
                          raw_x <- rand_uniform(-1, 1);
                          closed_form(~stand_dist(raw_x))
                          },
                        b = 128)$gen(300) %>%
  mutate(a = a_labels[1])

# Generate data for other a
map(c(0, 1), function(a) {
  heter_model(a = a,
              x = {
                raw_x <- rand_uniform(-1, 1); 
                closed_form(~stand_dist(raw_x))
                }, 
              b = 128)$
    gen(300, computed = select(dat_a_n1, x, e)) %>%
    mutate(a = a_labels[a + 2])
}) %>%

  # Combined with data for a = -1
  bind_rows(dat_a_n1) %>%
  mutate(a = factor(a, levels = a_labels)) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
  facet_wrap(~a, scales = "free", ncol = 3) +
  xlab("Fitted values") +
  ylab("Residuals")
```

```{r fig-different-b, fig.pos="!h", fig.height = 2, fig.cap = "Five different values of $b$ are used in heteroskedasticity simulation to control the strength of the signal. Larger values of $b$ yield a bigger difference in variation, and thus stronger heteroskedasticity signal."}

# Generate data for b = 0.25
dat_b025 <- heter_model(a = 0, 
                        x = {
                          raw_x <- rand_uniform(-1, 1);
                          closed_form(~stand_dist(raw_x))
                          }, 
                        b = 0.25)$gen(300) %>%
  mutate(b = "b: 0.25")

# Generate data for other b
map(c(1, 4, 16, 64), function(b) {
  heter_model(a = 0,
              x = {
                raw_x <- rand_uniform(-1, 1); 
                closed_form(~stand_dist(raw_x))
                }, 
              b = b)$
  gen(300, computed = select(dat_b025, x, e)) %>%
  mutate(b = paste0("b: ", b))
}) %>%

  # Combined with data for b = 0.25
  bind_rows(dat_b025) %>%
  mutate(b = factor(b, levels = paste0("b: ", c(0.25, 1, 4, 16, 64)))) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
  facet_wrap(~b, scales = "free", ncol = 5) +
  xlab("Fitted values") +
  ylab("Residuals")
```

#### Factors Common to both Data Collection Periods

Fitted values are a function of the independent variables (or predictors), and the distribution of the observed values affects the distribution of the fitted values. Ideally, we would want the fitted values to have a uniform coverage across the range of observed values or have a uniform distribution across all of the predictors. This is not always present in the collected data. Sometimes the fitted values are discrete because one or more predictors were measured discretely.  It is also common to see a skewed distribution of fitted values if one or more of the predictors has a skewed distribution. This latter problem is usually corrected before modelling, using a variable transformation. Our simulation assess this by using four different distributions to represent fitted values, constructed by different sampling of the predictor, including $U(-1, 1)$ (uniform), $N(0, 0.3^2)$ (normal), $\text{lognormal}(0, 0.6^2)/3$ (skewed) and $U\{-1, 1\}$ (discrete).

@fig-different-dist shows the non-linear pattern, a "U" shape, with the different fitted value distributions. We would expect that structure in residual plots would be easier to perceive when the fitted values are uniformly distributed.

Three different sample sizes were used in our experiment: $n = 50, 100, 300$. @fig-different-n shows the non-linear "S" shape for different sample sizes. We expect signal strength to decline in the simulated data plots with smaller $n$. We chose 300 as the upper limit, because it is typically enough for structure to be visible in a scatterplot reliably. Beyond 300, the scatterplot should probably be used with transparency or replaced with a density or binned plot as scatterplots suffer from over-plotting.

```{r fig-different-dist, fig.height = 2, fig.cap = "Variations in fitted values, that might affect perception of residual plots. Four different distributions are used."}

# Data for uniform distribution
dat_dist_1 <- poly_model(shape = 1, 
                      x = {
                        raw_x <- rand_uniform(-1, 1);
                        closed_form(~stand_dist(raw_x))
                        }, 
                      sigma = 0.5)$gen(300) %>%
  mutate(x_dist = "Uniform")

# Generate data for other distributions
dat_dist_2 <- poly_model(shape = 1,
                         x = {
                           raw_x <- rand_normal(sigma = 0.3); 
                           closed_form(~stand_dist(raw_x))
                           }, 
                         sigma = 0.5)$
  gen(300, computed = select(dat_dist_1, e)) %>%
  mutate(x_dist = "Normal")

dat_dist_3 <- poly_model(shape = 1,
                         x = {
                           raw_x <- rand_lognormal(sigma = 0.6); 
                           closed_form(~stand_dist(raw_x/3 - 1))
                           }, 
                         sigma = 0.5)$
  gen(300, computed = select(dat_dist_1, e)) %>%
  mutate(x_dist = "Skewed")

dat_dist_4 <- poly_model(shape = 1,
                         x = {
                           raw_x <- rand_uniform_d(k = 5, even = TRUE); 
                           closed_form(~stand_dist(raw_x))
                           }, 
                         sigma = 0.5)$
  gen(300, computed = select(dat_dist_1, e)) %>%
  mutate(x_dist = "Discrete")

# Generate and plot data for discrete uniform distribution
bind_rows(dat_dist_1, dat_dist_2, dat_dist_3, dat_dist_4) %>%
  mutate(x_dist = factor(x_dist, 
                         levels = c("Uniform", 
                                    "Normal", 
                                    "Skewed", 
                                    "Discrete"))) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
  xlab("Fitted values") +
  ylab("Residuals") +
  facet_wrap(~x_dist, ncol = 4, scales = "free")
```

```{r fig-different-n, fig.height = 2.67, fig.cap = 'Examining the effect of signal strength for the three different values of $n$ used in the experiment, for non-linear structure with fixed $\\sigma = 1.5$, uniform fitted value distribution, and "S" shape. For these factor levels, only when $n = 300$ is the "S" shape clearly visible.'}

set.seed(666)

# Generate data for 300 observations
dat_n_300 <- poly_model(shape = 2, 
                        x = {
                          raw_x <- rand_uniform(-1, 1)
                          closed_form(~stand_dist(raw_x))
                        },
                        sigma = 1)$gen(300) %>%
  mutate(n = "N: 300")

# Generate data for other settings
map(c(50, 100), function(n) {
  poly_model(shape = 2, 
                   x = {
                     raw_x <- rand_uniform(-1, 1);
                     closed_form(~stand_dist(raw_x))
                     }, 
             sigma = 1)$
    gen(n, computed = select(dat_n_300[1:n, ], x, e)) %>%
    mutate(n = paste0("N: ", n))
}) %>%

  # Combined with data for 300 observations
  bind_rows(dat_n_300) %>%
  mutate(n = factor(n, levels = c("N: 50", "N: 100", "N: 300"))) %>%
  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
  facet_wrap(~n, scales = "free", ncol = 3) +
  xlab("Fitted values") +
  ylab("Residuals")
```

### Effect Size {#sec-effect-size}

The lineups are allocated to participants in a manner that uniformly covers the combination of experimental factors in @tbl-model-factor-table. In addition, we use effect size to measure the signal strength, which helps in assigning a set of lineups with a range of difficulties to each participant.

Effect size in statistics measures the strength of the signal relative to the noise. It is surprisingly difficult to quantify, even for simulated data as used in this experiment.

For the non-linearity model, the key items defining effect size are sample size ($n$) and the noise level ($\sigma^2$), and so effect size would be roughly calculated as $\sqrt{n}/{\sigma}$. Increasing sample size tends to boost the effect size, while heightened noise diminishes it. However, it is not clear how the additional parameter for the model polynomial order, $j$, should be incorporated. Intuitively, the large $j$ means more complex pattern, which likely means effect size would decrease. Similarly, in the heteroskedasticity model, effect size relies on sample size ($n$) and the ratio of the largest to smallest variance, $b$. Larger values of both would produce higher effect size, but the role of the additional shape parameter, $a$, in this context is unclear.

For the purposes of our calculations we have chosen to use an approach based on Kullback-Leibler divergence [@kullback1951information]. This formulation defines effect size to be

$$E = \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}\text{diag}(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z'(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}\boldsymbol{\mu}_z\right),$$

\noindent where $\text{diag}(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix, $\boldsymbol{X}$ is the design matrix, $\boldsymbol{V}$ is the actual covariance matrix of the error term,
$\boldsymbol{R} = \boldsymbol{I}_n - \boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ is the residual operator, $\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals where $\boldsymbol{Z}$ contains any higher order terms of $\boldsymbol{X}$ left out of the regression equation, $\boldsymbol{\beta}_z$ contains the corresponding coefficients, and $\sigma^2\boldsymbol{I}_n$ is the assumed covariance matrix of the error term when $H_0$ is true. More details about the effect size derivation are provided in Appendix [-@sec-appendix-a-1].

## Results {#sec-results}

Data collection used a total of `r filter(vi_survey, !attention_check) %>% count(unique_lineup_id) %>% nrow()` lineups, and resulted in a total of `r nrow(filter(vi_survey, !attention_check))` evaluations from `r vi_survey %>% count(set, exp) %>% nrow()` participants. Roughly half corresponded to the two models, non-linearity and heteroskedasticiy, and the three collection periods had similar numbers of evaluations. Each participant received two of the `r vi_survey %>% filter(attention_check) %>% count(unique_lineup_id) %>% nrow()` attention check lineups which were used to filter results of participants who were clearly not making an honest effort (only 11 of `r vi_survey %>% count(set, exp) %>% nrow() + 11`). To estimate $\alpha$ for calculating statistical significance (see Appendix [-@sec-appendix-a-1]) there were `r nrow(filter(vi_survey, null_lineup, !attention_check))` evaluations of `r vi_survey %>% filter(null_lineup, !attention_check) %>% count(unique_lineup_id) %>% nrow()` null lineups. Neither the attention checks nor null lineups were used in the subsequent analysis. The de-identified data, `vi_survey`, is made available in the R package, `visage`.

The data was collected on lineups constructed from four different fitted value distributions that stem from the corresponding predictor distribution: uniform, normal, skewed and discrete. Henceforth, we refer to these four different fitted value distributions with respect to their predictor distribution. More data was collected on the uniform distribution (each evaluated by 11 participants) than the others (each evaluated by 5 participants). The analysis in @sec-power-analysis--@sec-hetero-analysis uses only results from lineups with uniform distribution, for a total `r nrow(filter(vi_survey, !attention_check, !null_lineup, x_dist == "uniform"))` lineup evaluations. This allows us to compare the conventional and visual test performance in an optimal scenario. @sec-effect-of-fitted-value-distributions examines how the results may be affected if the fitted value distribution was different.

### Power Comparison of the Tests {#sec-power-analysis}

@fig-nonlinearheterpower present the power curves of various tests plotted against the effect size in the residuals for non-linearity and heteroskedasticity. In each case the power of visual test is calculated for multiple bootstrap samples leading to the many (solid orange) curves. The effect size was computed at a 5% significance level and plotted on a natural logarithmic scale. To facilitate visual calibration of effect size values with the corresponding diagnostic plots, a sequence of example residual plots with increasing effect sizes is provided at the bottom of these figures. These plots serve as a visual aid to help readers understand how different effect size values translate to changes in the diagnostic plots. The horizontal lines of dots at 0 and 1 represent the non-rejection or rejection decisions made by visual tests for each lineup.

```{r cache = TRUE}
# Define the minimum and maximum effect size
min_poly_es <- vi_survey %>% 
  filter(!null_lineup, 
         !attention_check,
         type == "polynomial", 
         x_dist == "uniform") %>%
  pull(effect_size) %>%
  min()
max_poly_es <- vi_survey %>% 
  filter(!null_lineup, 
         !attention_check,
         type == "polynomial", 
         x_dist == "uniform") %>%
  pull(effect_size) %>%
  max()

# The glm model for RESET test
reset_poly_glmmod <- poly_conv_sim %>%
  filter(x_dist == "uniform") %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(reject = as.numeric(RESET4_p_value <= 0.05)) %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  select(x_dist, effect_size, log_effect_size, offset0, reject) %>%
  glm(reject ~ effect_size - 1, family = binomial(), data = ., offset = offset0)

# The prediction for all conventional tests on poly model
pred_conv_poly <- poly_conv_sim %>%
  filter(x_dist == "uniform") %>%
  select(-RESET3_p_value, -(RESET5_p_value:RESET10_p_value), -F_p_value) %>%
  rename(RESET_p_value = RESET4_p_value) %>%
  pivot_longer(RESET_p_value:SW_p_value) %>%
  mutate(name = gsub("_p_value", " test", name)) %>%
  mutate(reject = value <= 0.05) %>%
  select(effect_size, name, reject) %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  nest(dat = c(effect_size, offset0, reject)) %>%
  mutate(mod = map(dat, 
                   ~glm(reject ~ effect_size - 1, 
                        family = binomial(), 
                        data = .x,
                        offset = offset0))) %>%
  mutate(power = map(mod, function(mod) {
    result <- data.frame(effect_size = exp(seq(log(min_poly_es), log(max_poly_es), 0.1)),
                         offset0 = log(0.05/0.95))
    result$power <- predict(mod, type = "response", newdata = result)
    result
  })) %>%
  select(-dat, -mod) %>%
  unnest(power) %>%
  mutate(log_effect_size = log(effect_size))

# Data needed by the visual test
visual_poly_dat <- vi_survey %>%
  filter(!null_lineup, 
         !attention_check, 
         type == "polynomial", 
         x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  select(effect_size, offset0, reject)

# The glm model for visual test
visual_poly_glmmod <- glm(reject ~ effect_size - 1, 
                          family = binomial(), 
                          data = visual_poly_dat, 
                          offset = offset0)

# The prediction of the visual test
pred_visual_poly <- data.frame(effect_size = exp(seq(log(min_poly_es), log(max_poly_es), 0.1)),
                               offset0 = log(0.05/0.95)) %>%
  mutate(power = predict(visual_poly_glmmod, 
                         type = "response",
                         newdata = data.frame(effect_size = effect_size,
                                              offset0 = offset0))) %>%
  mutate(log_effect_size = log(effect_size))

# The prediction of the boot visual test
pred_visual_poly_boot <- map_df(1:500, function(boot_id) {
  slice_sample(visual_poly_dat, n = nrow(visual_poly_dat), replace = TRUE) %>%
    nest(dat = c(effect_size, offset0, reject)) %>%
    mutate(mod = map(dat, 
                     ~glm(reject ~ effect_size - 1, 
                          family = binomial(), 
                          data = .x,
                          offset = offset0))) %>%
    mutate(power = map(mod, function(mod) {
      result <- data.frame(effect_size = exp(seq(log(min_poly_es), log(max_poly_es), 0.1)),
                           offset0 = log(0.05/0.95))
      result$power <- predict(mod, type = "response", newdata = result)
      result
    })) %>%
    select(-dat, -mod) %>%
    unnest(power) %>%
    mutate(log_effect_size = log(effect_size)) %>%
    mutate(boot_id = boot_id)
  })
```

```{r}
# Analyse uniform polynomial data
p <- ggplot() +

  geom_point(data = visual_poly_dat,
             aes(log(effect_size), reject),
             alpha = 0.15) +

  # Draw conventional test power curve
  geom_line(data = pred_conv_poly,
            aes(log_effect_size, power, col = name, linetype = name)) +

  # Draw boot visual test curves
  geom_line(data = pred_visual_poly_boot,
            aes(log_effect_size, power, col = "visual test", group = boot_id, linetype = "visual test"),
            alpha = 0.01) +

  # Draw visual test power curve
  geom_line(data = pred_visual_poly,
            aes(log_effect_size, power, col = "visual test", linetype = "visual test")) +

  # Theme
  theme_light(base_size = 5) +
  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.3)) +
  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
  scale_linetype_manual(values = c("visual test" = 1, "RESET test" = 2, "BP test" = 3, "SW test" = 4)) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  labs(col = "", linetype = "")
```

```{r cache = TRUE}
# target_sigma is obtained by evaluating a vector of sigma and picking those 
# with expected effect 0 ~ 5

target_sigma <- c(3.65, 2.84, 2.21, 1.72, 1.34, 1.04, 0.81, 0.63, 0.49, 0.38, 0.3)

# Calculate the effect size 
es <- map_dbl(target_sigma, 
              ~log(poly_model(shape = 2, 
                              x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, 
                              sigma = .x)$
                     average_effect_size(n = 100, type = "kl")
                   )
              )
```

```{r fig-polypower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for non-linear patterns (uniform distribution only). The power curves are estimated using logistic regression, and the horizontal lines of dots represent non-reject and reject results from visual tests for each lineup. The visual test has multiple power curves estimated from bootstrap samples. The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.", dev = 'png', dpi = 450}

set.seed(132)

# Base data
ex_dat <- poly_model(shape = 2, 
                     x = rand_uniform(-1, 1), 
                     sigma = 1)$gen(n = 100)
plist <- list()
j <- 0

# Obtain data at different effect. Reuse x and recompute e.
for (i in target_sigma) {
  j <- j + 1
  plist[[j]] <- poly_model(shape = 2, 
                           x = rand_uniform())$
                    gen(n = 100, 
                        computed = ex_dat %>%
                          mutate(e = sqrt(i) * e) %>%
                          select(x, e)
                        ) %>%
    VI_MODEL$plot(theme = theme_light(base_size =  5), 
                  remove_axis = TRUE, 
                  remove_grid_line = TRUE, size = 0.25, stroke = 0.25) +
      ggtitle(round(es[j], 1)) +
      theme(aspect.ratio = 0.8)
}

# Use area() to design layout
A <- map(1:11, ~patchwork::area(21, .x, 22, .x)) %>%
  reduce(c) %>%
  c(patchwork::area(3, 1, 20, 11), .) %>%
  patchwork::wrap_plots(append(plist, list(p + labs(title = "(A) Non-linearity") + theme(plot.title = element_text(size = 10))), after = 0), design = .)
```

```{r cache = TRUE}
# Define the minimum and maximum effect size
min_heter_es <- vi_survey %>% 
  filter(!null_lineup, 
         !attention_check,
         type == "heteroskedasticity", 
         x_dist == "uniform") %>%
  pull(effect_size) %>%
  min()
max_heter_es <- vi_survey %>% 
  filter(!null_lineup, 
         !attention_check,
         type == "heteroskedasticity", 
         x_dist == "uniform") %>%
  pull(effect_size) %>%
  max()

# The prediction for all conventional tests on heter model
pred_conv_heter <- heter_conv_sim %>%
  filter(x_dist == "uniform") %>%
  select(-RESET3_p_value, -(RESET5_p_value:RESET10_p_value), -F_p_value) %>%
  rename(RESET_p_value = RESET4_p_value) %>%
  pivot_longer(RESET_p_value:SW_p_value) %>%
  mutate(name = gsub("_p_value", " test", name)) %>%
  mutate(reject = value <= 0.05) %>%
  select(effect_size, name, reject) %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  nest(dat = c(effect_size, offset0, reject)) %>%
  mutate(mod = map(dat, 
                   ~glm(reject ~ effect_size - 1, 
                        family = binomial(), 
                        data = .x,
                        offset = offset0))) %>%
  mutate(power = map(mod, function(mod) {
    result <- data.frame(effect_size = exp(seq(log(min_heter_es), log(max_heter_es), 0.1)),
                         offset0 = log(0.05/0.95))
    result$power <- predict(mod, type = "response", newdata = result)
    result
  })) %>%
  select(-dat, -mod) %>%
  unnest(power) %>%
  mutate(log_effect_size = log(effect_size))

# Data needed by the visual test
visual_heter_dat <- vi_survey %>%
  filter(!null_lineup, 
         !attention_check, 
         type == "heteroskedasticity", 
         x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  select(effect_size, offset0, reject)

# The glm model for visual test
visual_heter_glmmod <- glm(reject ~ effect_size - 1, 
                           family = binomial(), 
                           data = visual_heter_dat, 
                           offset = offset0)

# The prediction of the visual test
pred_visual_heter <- data.frame(effect_size = exp(seq(log(min_heter_es), log(max_heter_es), 0.1)),
                                offset0 = log(0.05/0.95)) %>%
  mutate(power = predict(visual_heter_glmmod, 
                         type = "response",
                         newdata = data.frame(effect_size = effect_size,
                                              offset0 = offset0))) %>%
  mutate(log_effect_size = log(effect_size))

# The prediction of the boot visual test
pred_visual_heter_boot <- map_df(1:500, function(boot_id) {
  slice_sample(visual_heter_dat, n = nrow(visual_heter_dat), replace = TRUE) %>%
    nest(dat = c(effect_size, offset0, reject)) %>%
    mutate(mod = map(dat, 
                     ~glm(reject ~ effect_size - 1, 
                          family = binomial(), 
                          data = .x,
                          offset = offset0))) %>%
    mutate(power = map(mod, function(mod) {
      result <- data.frame(effect_size = exp(seq(log(min_heter_es), log(max_heter_es), 0.1)),
                           offset0 = log(0.05/0.95))
      result$power <- predict(mod, type = "response", newdata = result)
      result
    })) %>%
    select(-dat, -mod) %>%
    unnest(power) %>%
    mutate(log_effect_size = log(effect_size)) %>%
    mutate(boot_id = boot_id)
  })
```

```{r}
# Analyse uniform heteroskedasticity data
p <- ggplot() +

  geom_point(data = visual_heter_dat,
             aes(log(effect_size), reject),
             alpha = 0.15) +

  # Draw conventional test power curve
  geom_line(data = pred_conv_heter,
            aes(log_effect_size, power, col = name, linetype = name)) +

  # Draw boot visual test curves
  geom_line(data = pred_visual_heter_boot,
            aes(log_effect_size, power, col = "visual test", group = boot_id, linetype = "visual test"),
            alpha = 0.01) +

  # Draw visual test power curve
  geom_line(data = pred_visual_heter,
            aes(log_effect_size, power, col = "visual test", linetype = "visual test")) +

  # Theme
  theme_light(base_size = 5) +
  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.3)) +
  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
  scale_linetype_manual(values = c("visual test" = 1, "RESET test" = 2, "BP test" = 3, "SW test" = 4)) +
  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  labs(col = "", linetype = "")
```

```{r cache = TRUE}
# target_b is obtained by evaluating a vector of b and picking those 
# with expected effect 0 ~ 5

target_b <- c(0.17, 0.23, 0.32, 0.46, 0.67, 1.04, 1.7, 3, 6.3, 16.2, 61)

# Calculate the effect size 
es <- map_dbl(target_b, 
              ~log(heter_model(a = 1, 
                               b = .x, 
                               x = {raw_x <- rand_normal(sigma = 0.3);closed_form(~stand_dist(raw_x))})$
                     average_effect_size(n = 100, type = "kl")
                   )
              )
```

```{r fig-nonlinearheterpower, fig.cap = "Comparison of power between different tests for (A) non-linear and (B) heteroskedasticity patterns (uniform fitted values only). Main plot shows the power curves, with dots indicating non-reject and reject in visual testing of lineups. The multiple lines for the visual test arise from estimating the power on many bootstrap samples. The row of scatterplots at the bottom are examples of residual plots corresponding to the specific effect sizes marked by vertical lines in the main plot.", fig.height = 6.5, fig.width = 5, dev = 'png', dpi = 450}

set.seed(135)

# Base data
ex_dat <- heter_model(a = 1, 
                      x = rand_uniform(-1, 1), 
                      b = 1)$gen(n = 100)
plist <- list()
j <- 0

# Obtain data at different b
for (i in target_b) {
  j <- j + 1
  plist[[j]] <- heter_model(a = 1, 
                            b = i, 
                            x = rand_uniform(-1, 1))$
                  gen(n = 100, computed = select(ex_dat, x, e)) %>%
  VI_MODEL$plot(theme = theme_light(base_size =  5),
                remove_axis = TRUE,
                remove_grid_line = TRUE, size = 0.25, stroke = 0.25) +
    ggtitle(round(es[j], 1)) +
    theme(aspect.ratio = 0.8)
}

# Use area() to design layout
B <- map(1:11, ~patchwork::area(21, .x, 22, .x)) %>%
  reduce(c) %>%
  c(patchwork::area(3, 1, 20, 11), .) %>%
patchwork::wrap_plots(append(plist, list(p + labs(title = "(B) Heteroskedasticity") + 
  theme(plot.title = element_text(size = 10))), after = 0),
    design = .)

A / B + patchwork::plot_layout(guides = "collect")
```

[@fig-nonlinearheterpower]A compares the power for the different tests for non-linear structure in the residuals. The test with the uniformly higher power is the RESET test, one that specifically tests for non-linearity. Note that the BP and SW tests have much lower power, which is expected because they are not designed to detect non-linearity. The bootstrapped power curves for the visual test are effectively a right shift from that of the RESET test. This means that the RESET test will reject at a lower effect size (less structure) than the visual test, but otherwise the performance will be similar. In other words, the RESET test is more sensitive than the visual test. This is not necessarily a good feature for the purposes of diagnosing model defects: if we scan the residual plot examples at the bottom, we might argue that the non-linearity is not sufficiently problematic until an effect size of around 3 or 3.5. The RESET test would reject closer to an effect size of 2, but the visual test would reject closer to 3.25, for a significance level of 0.05. The visual test matches the robustness of the model to (minor) violations of assumptions much better.

For the heteroskedasticity pattern, the power of BP test, designed for detecting heteroskedasticity, is uniformly higher than the other tests. The visual test power curve shifts to the right. This shows a similar story to the power curves for non-linearity pattern: the conventional test is more sensitive than the visual test. From the example residual plots at the bottom we might argue that the heteroskedasticity becomes noticeably visible around an effect size of 3 or 3.5. However the BP test would reject at around effect size 2.5. Interestingly, the power curve for the SW test (for non-normality) is only slightly different to that of the visual test, suggesting that it performs reasonably well for detecting heteroskedasticity, too. The power curve for the BP test suggests it is not useful for detecting heteroskedasticity, as expected.

Overall, the results show that the conventional tests are more sensitive than the visual test. The conventional tests do have higher power for the patterns they are designed to detect, but they typically fail to detect other patterns unless those patterns are particularly strong. The visual test does not require specifying the pattern ahead of time, relying purely on whether the observed residual plot is detectably different from "good" residual plots. They will perform equally well regardless of the type of model defect. This aligns with the advice of experts on residual analysis, who consider residual plot analysis to be an indispensable tool for diagnosing model problems. What we gain from using a visual test for this purpose is the removal of any subjective arguments about whether a pattern is visible or not. The lineup protocol provides the calibration for detecting patterns: if the pattern in the data plot cannot be distinguished from the patterns in good residual plots, then no discernible problem with the model exists.

### Comparison of Test Decisions Based on $p$-values {#sec-p-value}

```{r}
# Dataset for p-value comparison
p_value_cmp_dat <- vi_survey %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  filter(x_dist == "uniform") %>%
  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
  mutate(conv_reject = ifelse(conventional_p_value <= 0.05, "Reject", "Not"),
         reject = ifelse(p_value <= 0.05, "Reject", "Not")) %>%
  mutate(conv_reject = factor(conv_reject, levels = c("Reject", "Not")),
         reject = factor(reject, levels = c("Reject", "Not"))) %>%
  select(type, conv_reject, reject)
```

The power comparison demonstrates that the appropriate conventional tests will reject more aggressively than visual tests, but we do not know how the decisions for each lineup would agree or disagree. Here we compare the reject or fail to reject decisions of these tests, across all the lineups. @fig-p-value-comparison shows the agreement of the conventional and visual tests using a mosaic plot for both non-linearity patterns and heteroskedasticity patterns. For both patterns the lineups resulting in a rejection by the visual test are *all* also rejected by the conventional test, except for one from the heteroskedasticity model. This reflects exactly the story from the previous section, that the conventional tests reject more aggressively than the visual test.

For non-linearity lineups, conventional tests and visual tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% and `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the time, respectively. Of the lineups rejected by the conventional test, `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by the visual test, that is, approximately half as many as the conventional test. There are no lineups that are rejected by the visual test but not by the conventional test.

In heteroskedasticity lineups, `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by conventional tests, while `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by visual tests. Of the lineups rejected by the conventional test, the visual test rejects more than two-thirds of them, too.

```{r}
nr <- p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Not") %>% count(reject)
```

Surprisingly, the visual test rejects `r nr$n[1]` of the `r nr$n[1]+nr$n[2]` (`r round(nr$n[1]/(nr$n[1]+nr$n[2])*100, 0)`%) of lineups where the conventional test does not reject. @fig-heter-example shows this lineup. The data plot in position seventeen displays a relatively strong heteroskedasticity pattern, and has a strong effect size ($\log_e(E)=4.02$), which is reflected by the visual test $p\text{-value} = 0.026$. But the BP test $p\text{-value} = 0.056$, is slightly above the significance cutoff of $0.05$. This lineup was evaluated by 11 participants, it has experimental factors $a = 0$ ("butterfly" shape), $b = 64$ (large variance ratio), $n = 50$ (small sample size), and a uniform distribution for the predictor. It may have been the small sample size and the presence of a few outliers that may have resulted in the lack of detection by the conventional test.

```{r fig-p-value-comparison, fig.width = 8, fig.height = 4, fig.cap = "Rejection rate ($p$-value $\\leq0.05$) of visual test conditional on the conventional test decision on non-linearity (left) and heteroskedasticity (right) lineups (uniform fitted values only) displayed using a mosaic plot. The visual test rejects less frequently than the conventional test, and (almost) only rejects when the conventional test does. Surprisingly, one lineup in the heteroskedasticity group is rejected by the visual test but NOT the conventional test."}

library(ggmosaic)

# Mosaic plot
p_value_cmp_dat %>%
  mutate(type = ifelse(type == "non-linearity", "Non-linearity", "Heteroskedasticity")) %>%
  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity"))) %>%
  ggplot() +
  geom_mosaic(aes(x = ggmosaic::product(reject, conv_reject), 
                  fill = reject)) +
  facet_grid(~type) +
  ylab("Visual tests") +
  xlab("Conventional tests") +
  labs(fill = "Conventional tests") +
  scale_fill_brewer("", palette = "Dark2") +
  theme_bw() +
  theme(legend.position = "none") +
  coord_fixed()
```

```{r fig-heter-example, fig.pos="t!", fig.height = 7, fig.width = 7, fig.cap = 'The single heteroskedasticity lineup that is rejected by the visual test but not by the BP test. The data plot (position 17) contains a "butterfly" shape. It visibly displays heteroskedasticity, making it somewhat surprising that it is not detected by the BP test.'}
# plot heter-331
VI_MODEL$plot_lineup(vi_lineup$heter_331$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(),
                     remove_axis = TRUE)
```

Because the power curve of the visual tests are a shift to the right of the conventional test (@fig-nonlinearheterpower) we examined whether adjusting the significance level (to .001, .0001, .00001, ...) of the conventional test would generate similar decisions to that of the visual test. Interestingly, it does not: despite resulting in less rejections, neither the RESET or BP tests come to complete agreement with the visual test (see Appendix [-@sec-appendix-a-1]).

### Effect of Amount of Non-linearity {#sec-nonlin-analysis}

The order of the polynomial is a primary factor contributing to the pattern produced by the non-linearity model. @fig-poly-power-uniform-j explores the relationship between polynomial order and power of the tests. The conventional tests have higher power for lower orders of Hermite polynomials, and the power drops substantially for the "triple-U" shape. To understand why this is, we return to the application of the RESET test, which requires a parameter indicating degree of fitted values to test for, and the recommendation is to generically use four [@ramsey1969tests]. However, the "triple-U" shape is constructed from the Hermite polynomials using power up to 18. If the RESET test had been applied using a higher power of no less than six, the power curve of "triple-U" shape will be closer to other power curves. This illustrates the sensitivity of the conventional test to the parameter choice, and highlights a limitation: it helps to know the data generating process to set the parameters for the test, which is unrealistic in practice. However, we examined this in more detail (see Appendix [-@sec-appendix-a-1]) and found that there is no harm in setting the parameter higher than four on the tests' operation for lower order polynomial shapes. Using a parameter value of six, instead of four, yields higher power regardless of generating process, and is recommended.

For visual tests, we expect the "U" shape to be detected more readily, followed by the "S", "M" and "triple-U" shape. From @fig-poly-power-uniform-j, it can be observed that the power curves mostly align with these expectations, except for the "M" shape, which is as easily detected as the "S" shape. This suggests a benefit of the visual test: knowing the shape ahead of time is *not* needed for its application.

```{r fig-poly-power-uniform-j, fig.height = 4, fig.cap = 'The effect of the order of the polynomial on the power of conventional and visual tests. Deeper colour indicates higher order. The default RESET tests under-performs significantly in detecting the "triple-U" shape. To achieve a similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test, but this higher than the recommended value.'}

dat_poly_j <- poly_conv_sim %>%
  filter(x_dist == "uniform") %>%
  mutate(reject = as.numeric(RESET4_p_value <= 0.05)) %>%
  mutate(shape = c("U", "S", "M", "triple-U")[shape]) %>%
  mutate(conv_or_not = "Conventional") %>%
  bind_rows(vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform", type == "polynomial") %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(shape = c("U", "S", "M", "triple-U")[shape]) %>%
  mutate(conv_or_not = "Visual")) %>%
  mutate(shape = factor(shape, levels = c("U", "S", "M", "triple-U"))) %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  select(shape, effect_size, offset0, reject, conv_or_not)

dat_poly_j %>%
  nest(dat = c(effect_size, offset0, reject)) %>%
  mutate(mod = map(dat, 
                   ~glm(reject ~ effect_size - 1, 
                        family = binomial(), 
                        data = .x,
                        offset = offset0))) %>%
  mutate(power = map(mod, function(mod) {
    result <- data.frame(effect_size = exp(seq(log(min_poly_es), log(max_poly_es), 0.1)),
                         offset0 = log(0.05/0.95))
    result$power <- predict(mod, type = "response", newdata = result)
    result
  })) %>%
  select(-dat, -mod) %>%
  unnest(power) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_line(aes(log_effect_size, power, col = shape, linetype = shape), size = 1) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  scale_color_manual(values = rcartocolor::carto_pal(4, "TealGrn")) +
  scale_linetype_manual(values = c("U" = 1, "S" = 2, "M" = 3, "triple-U" = 4)) +
  theme_light() +
  theme(legend.position = "bottom", legend.key.width = unit(2.5, "line")) +
  facet_wrap(~conv_or_not) +
  labs(col = "Non-linearity shape", linetype = "Non-linearity shape")
```

### Effect of Shape of Heteroskedasticity {#sec-hetero-analysis}

@fig-heter-power-uniform-a examines the impact of the shape of the heteroskedasticity on the power of of both tests. The butterfly shape has higher power on both types of tests. The "left-triangle" and the "right-triangle" shapes are functionally identical, and this is observed for the conventional test, where the power curves are identical. Interestingly there is a difference for the visual test: the power curve of the "left-triangle" shape is slightly higher than that of the "right-triangle" shape. This indicates a bias in perceiving heteroskedasticity depending on the direction, and may be worth investigating further.

```{r fig-heter-power-uniform-a, fig.height = 4, fig.cap = 'The effect of heteroskedasticity shape (parameter $a$) on the power of conventional and visual tests. The butterfly has higher power in both tests. Curiously, the visual test has a slightly higher power for the ``left-triangle" than the ``right-triangle" shape, when it would be expected that they should be similar, which is observed in conventional testing.'}

dat_heter_a <- heter_conv_sim %>%
  filter(x_dist == "uniform") %>%
  mutate(reject = as.numeric(BP_p_value <= 0.05)) %>%
  mutate(a = c("left-triangle", "butterfly", "right-triangle")[a + 2]) %>%
  mutate(conv_or_not = "Conventional") %>%
  bind_rows(vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform", type == "heteroskedasticity") %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(a = c("left-triangle", "butterfly", "right-triangle")[a + 2]) %>%
  mutate(conv_or_not = "Visual")) %>%
  mutate(a = factor(a, levels = c("left-triangle", "butterfly", "right-triangle"))) %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  select(a, effect_size, offset0, reject, conv_or_not)

# Manually fit the model
dat_heter_a %>%
  nest(dat = c(effect_size, offset0, reject)) %>%
  mutate(mod = map(dat, 
                   ~glm(reject ~ effect_size - 1, 
                        family = binomial(), 
                        data = .x,
                        offset = offset0))) %>%
  mutate(power = map(mod, function(mod) {
    result <- data.frame(effect_size = exp(seq(log(min_heter_es), log(max_heter_es), 0.1)),
                         offset0 = log(0.05/0.95))
    result$power <- predict(mod, type = "response", newdata = result)
    result
  })) %>%
  select(-dat, -mod) %>%
  unnest(power) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_line(aes(log_effect_size, power, col = a, linetype = a), size = 1) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  scale_color_manual(values = c(rcartocolor::carto_pal(3, "Earth")[1], "grey", rcartocolor::carto_pal(3, "Earth")[3])) +
  scale_linetype_manual(values = c("left-triangle" = 1, "butterfly" = 2, "right-triangle" = 3)) +
  theme_light() +
  theme(legend.position = "bottom", legend.key.width = unit(2.5, "line")) +
  facet_wrap(~conv_or_not) +
  labs(col = "Heteroskedasticity shape", linetype = "Heteroskedasticity shape")
```

### Effect of Fitted Value Distributions {#sec-effect-of-fitted-value-distributions}

In regression analysis, predictions are conditional on the observed values of the predictors, that is, the conditional mean of the dependent variable $Y$ given the value of the independent variable $X$, $\text{E}(Y|X)$. This is an often forgotten element of regression analysis but it is important. Where $X$ is observed, the distribution of the $X$ values in the sample, or consequently $\hat{Y}$, may affect the ability to read any patterns in the residual plots. The effect of fitted value distribution on test performance is assess using four different distributions of fitted values stemming from the predictor distributions: uniform, normal, discrete and lognormal (skewed). We expect that if all predictors have a uniform distribution, it is easier to read the relationship with the residuals.

```{r get-5-eval, cache = TRUE}
dat_uniform_5 <- map_df(1:500, function(boot_id) {
  vi_survey %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  slice_sample(n = 5) %>%
  ungroup() %>%
  calc_p_value_multi(lineup_id = unique_lineup_id,
                     detect = detect,
                     n_sel = num_selection,
                     alpha = alpha) %>%
  mutate(boot_id = boot_id)
}) %>%
  left_join(select(vi_survey, unique_lineup_id, effect_size, x_dist, type)) %>%
  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(conv_or_not = "Visual") %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  select(-unique_lineup_id, -p_value)

dat_uniform_5_pred <- dat_uniform_5 %>%
  nest(dat = c(effect_size, offset0, reject)) %>%
  mutate(mod = map(dat, 
                   ~glm(reject ~ effect_size - 1, 
                        family = binomial(), 
                        data = .x,
                        offset = offset0))) %>%
  mutate(power = map2(mod, type, function(mod, type) {
    this_min_es <- ifelse(type == "heteroskedasticity", min_heter_es, min_poly_es)
    this_max_es <- ifelse(type == "heteroskedasticity", max_heter_es, max_poly_es)
    result <- data.frame(effect_size = exp(seq(log(this_min_es[1]), log(this_max_es[1]), 0.1)),
                         offset0 = log(0.05/0.95))
    result$power <- predict(mod, type = "response", newdata = result)
    result
  })) %>%
  select(-dat, -mod) %>%
  unnest(power) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(type = factor(ifelse(type == "non-linearity", "Non-linearity", "Heteroskedasticity"), 
                       levels = c("Non-linearity", "Heteroskedasticity")))
```

```{r fig-different-x-dist-poly-power, fig.pos="t!", fig.height = 6, fig.cap = "Comparison of power on lineups with different fitted value distributions for conventional and visual tests (columns) for non-linearity and heteroskedasticity patterns (rows). The power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and BP tests, respectively. Power curves of visual tests are estimated using five evaluations on each lineup. For lineups with a uniform fitted value distribution, the five evaluations are repeatedly sampled from the total eleven evaluations to give multiple power curves (solid grey). Surprisingly, the fitted value distribution has produces more variability in the power of conventional tests than visual tests. Uneven distributions, normal and skewed distributions, tend to yield lower power.", dev = 'png', dpi = 450}

diff_x_dist_power_pred <- list(
  # Visual inference data
  vi_survey %>%
    filter(!null_lineup) %>%
    filter(!attention_check) %>%
    group_by(unique_lineup_id) %>%
    summarise(across(everything(), first)) %>%
    filter(x_dist != "uniform") %>%
    mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
    mutate(reject = as.numeric(p_value <= 0.05)) %>%
    mutate(offset0 = log(0.05/0.95)) %>%
    select(type, x_dist, effect_size, offset0, reject) %>%
    mutate(conv_or_not = "Visual"),

  # Conventional simulation (poly)
  poly_conv_sim %>%
    mutate(reject = RESET4_p_value < 0.05) %>%
    mutate(offset0 = log(0.05/0.95)) %>%
    mutate(type = "non-linearity") %>%
    select(type, x_dist, effect_size, offset0, reject) %>%
    mutate(conv_or_not = "Conventional"),

  # Conventional simulation (heter)
  heter_conv_sim %>%
    mutate(reject = BP_p_value < 0.05) %>%
    mutate(offset0 = log(0.05/0.95)) %>%
    mutate(type = "heteroskedasticity") %>%
    select(type, x_dist, effect_size, offset0, reject) %>%
    mutate(conv_or_not = "Conventional")
) %>%

  # Bind rows
  map_df(~.x) %>%

  nest(dat = c(effect_size, offset0, reject)) %>%
  mutate(mod = map(dat, 
                   ~glm(reject ~ effect_size - 1, 
                        family = binomial(), 
                        data = .x,
                        offset = offset0))) %>%
  mutate(power = map2(mod, type, function(mod, type) {
    this_min_es <- ifelse(type == "heteroskedasticity", min_heter_es, min_poly_es)
    this_max_es <- ifelse(type == "heteroskedasticity", max_heter_es, max_poly_es)
    result <- data.frame(effect_size = exp(seq(log(this_min_es[1]), log(this_max_es[1]), 0.1)),
                         offset0 = log(0.05/0.95))
    result$power <- predict(mod, type = "response", newdata = result)
    result
  })) %>%
  select(-dat, -mod) %>%
  unnest(power) %>%
  mutate(log_effect_size = log(effect_size))

diff_x_dist_power_pred %>%
  mutate(x_dist = ifelse(x_dist == "even_discrete", 
                         "discrete", 
                         x_dist)) %>%
  mutate(type = factor(ifelse(type == "non-linearity", "Non-linearity", "Heteroskedasticity"), 
                       levels = c("Non-linearity", "Heteroskedasticity"))) %>%
  mutate(x_dist = ifelse(x_dist == "lognormal", "skewed", x_dist)) %>%
  mutate(x_dist = factor(x_dist, 
                         levels = c("uniform", 
                                    "normal", 
                                    "skewed", 
                                    "discrete"))) %>%

  ggplot() +

  # Draw all smooth lines
  geom_line(aes(log_effect_size, power, col = x_dist, linetype = x_dist),
            size = 1) +

  # Draw boot uniform (5 eval) lines
  geom_line(data = dat_uniform_5_pred,
            aes(log_effect_size, power, col = "uniform", group = boot_id, linetype = "uniform"),
            size = 0.5,
            alpha = 0.01) +

  # Theme
  theme_light() +
  scale_color_manual(values = rcartocolor::carto_pal(4, "Safe")[c(4,1,3,2)]) +
  scale_linetype_manual(values = c("uniform" = 1, "normal" = 2, "skewed" = 3, "discrete" = 4)) + 
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  labs(col = "Fitted value distribution", linetype = "Fitted value distribution") +
  facet_grid(type~conv_or_not, scales = "free_x") +
  theme(legend.position = "bottom", legend.key.width = unit(2.5, "line"))
```

@fig-different-x-dist-poly-power examines the impact of the fitted value distribution on the power of conventional (left) and visual (right) tests for both the non-linearity (top) and heteroskedasticity (bottom) patterns. For conventional tests, only the power curves of appropriate tests are shown:  RESET tests for non-linearity and BP tests for heteroskedasticity. For visual tests, more evaluations on lineups with uniform fitted value distribution were collected, so to have a fair comparison, we randomly sample five from the 11 total evaluations to estimate the power curves,  producing the multiple curves for the uniform condition, and providing an indication of the variability in the power estimates.

Perhaps surprisingly, the visual tests have more consistent power across the different fitted value distributions: for the non-linear pattern, there is almost no power difference, and for the heteroskedastic pattern, uniform and discrete have higher power than normal and skewed. The likely reason is that these latter two have fewer observations in the tails where the heteroskedastic pattern needs to be detected.

The variation in power in the conventional tests is at first sight, shocking. However, it is discussed, albeit rarely, in the testing literature. See, for example, @jamshidian2007study, @olvera2019relationship and  @zhang2018practical which show derivations and use simulation to assess the effect of the observed distribution of the predictors on test power. The big differences in the power curves seen in @fig-different-x-dist-poly-power is echoed in the results reported in these articles.

# Limitations and Practicality

One of the primary limitations of the lineup protocol lies in its reliance on human judgements. In this context, the effectiveness of a single lineup evaluation can be dependent on the perceptual ability and visual skills of the individual. However, when results from multiple individuals are combined the outcome is encouragingly high-quality and robust. For simple plots and strong patterns just a few individuals are needed to arrive at a clear answer, but more individuals will be needed when the plot design is complex, or the signal strength is weak.

Using a lineup protocol removes subjectiveness in interpreting patterns in plots. A plot is compared with draws from a null model, in much the same way as a test statistic is compared to its sampling distribution. It is important to remove plot elements that might introduce bias, such as axis labels, text and legends, or to make them generic.

The lineup protocol can be used cheaply and informally with the R package `nullabor`. There is evidence that it is being used fairly broadly, based on software download rates and citations of the original papers. For residual plot analysis we recommend that the lineup be the default first plot so that the data plot is only seen in the context of null plots. When a rigorous test is needed, we recommend using a crowd-sourcing service, as done in gene expression experiment described in @RNAseq2013. While it takes extra effort it is not difficult today, and costs are tiny compared to the overall costs of conducting a scientific experiment. We do also expect that at some point a computer vision model can be developed to take over the task of employing people to evaluate residual plots.

For this study, simulated data was used to provide a precisely controlled environment within which to compare results from conventional testing to those from visual testing. We also explored only the most commonly used, the residual vs fitted value plots. However, we expect the behaviour of the conventional test and the visual test to be similar when observed residuals are diagnosed with this type of plot or other residual plots. The conventional tests will be more sensitive to small departures from the null. They will also fail to detect departures when residuals have some contamination, like outliers or anomalies, as is often encountered when working with data. The lineup approach is well-suited for generally interpreting data plots, and also detecting unexpected patterns not related to the model. This is supported by earlier research [e.g. @loy2016variations; @loy2015you; @roy2015using; @VanderPlas2015; @wickham2010].

# Conclusions

This paper has described experimental evidence providing support for the advice of regression analysis experts *that residual plots are indispensable methods for assessing model fit*, using the formal framework of the lineup protocol. We conducted a perceptual experiment on scatterplots of residuals vs fitted values, with two primary departures from good residuals: non-linearity and heteroskedasticity. We found that conventional residual-based statistical tests are more sensitive to weak departures from model assumptions than visual tests. That is, a conventional test concludes there are problems with the model fit almost twice as often as a human. Conventional tests often reject the null hypothesis when departures in the form of non-linearity and heteroskedasticity are not visibly different from null residual plots.

While it might be argued that the conventional tests are correctly detecting small but real effects, this can also be seen as the conventional tests are rejecting unnecessarily. Many of these rejections happen even when downstream analysis and results would not be significantly affected by the small departures from a good fit. The results from human evaluations provide a more practical solution, which reinforces the statements from regression experts that residual plots are an indispensable method for model diagnostics. Further work would be needed to *quantify* how much departure from good residuals is too much.

It is important to emphasize that this work also supports a change in common practice, which is to deliver residual plots as a lineup, embedded in a field of null plots, rather than be viewed out of context. A residual plot may contain many visual features, but some are caused by the characteristics of the predictors and the randomness of the error, not by the violation of the model assumptions. These irrelevant visual features have a chance to be filtered out by participants with a comparison to null plots, resulting in more accurate reading. The lineup enables a careful calibration for reading structure in residual plots, and also provides the potential to discover interesting and important features in the data not directly connected to linear model assumptions.

Human evaluation of residuals is expensive, time-consuming and laborious. This is possibly why residual plot analysis is often not done in practice. However, with the emergence of effective computer vision, it is hoped this work helps to lay the foundation for automated residual plot assessment.

The experiment also revealed some interesting results. For the most part, the visual test performed similarly to the appropriate conventional test with a shift in the power curve. Unlike conventional tests, where one needs to specifically test for non-linearity or heteroskedasticity the visual test operated effectively across the range of departures from good residuals. If the fitted value distribution is not uniform, there is a small loss of power in the visual test. Surprisingly, there is a big difference in power of the conventional test across fitted value distributions. Another unexpected finding was that the direction of heteroskedasticity appears to affect the ability to visually detect it: both triangles being more difficult to detect than the butterfly, and a small difference in detection between left- and right-triangle.

### 2024-06-24 Thesis/A-appA.qmd
finish typesetting chap 2

# Appendix to "A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol" {#sec-appendix-a}

```{r}
if (!requireNamespace("remotes", quietly = TRUE)) {
  install.packages("remotes")
}

# OOP supports needed by `visage`
if (!requireNamespace("bandicoot", quietly = TRUE)) {
  remotes::install_github("TengMCing/bandicoot")
}

# Visual inference models and p-value calculation
if (!requireNamespace("visage", quietly = TRUE)) {
  remotes::install_url("https://github.com/TengMCing/visage/raw/master/built/visage_0.1.2.tar.gz")
} else {
  if (packageVersion("visage") < "0.1.2") {
    remotes::install_url("https://github.com/TengMCing/visage/raw/master/built/visage_0.1.2.tar.gz")
  }
}

if (!requireNamespace("tidyverse", quietly = TRUE)) {
  install.packages("tidyverse")
}

if (!requireNamespace("here", quietly = TRUE)) {
  install.packages("here")
}

if (!requireNamespace("ggmosaic", quietly = TRUE)) {
  install.packages("ggmosaic")
}

if (!requireNamespace("rcartocolor", quietly = TRUE)) {
  install.packages("rcartocolor")
}

if (!requireNamespace("kableExtra", quietly = TRUE)) {
  install.packages("kableExtra")
}

if (!requireNamespace("patchwork", quietly = TRUE)) {
  install.packages("patchwork")
}

library(tidyverse)
library(visage)

# To control the simulation in this file
set.seed(10086)
```

```{r get-lineup-data}
# Create the dir folder
if (!dir.exists(here::here("cached_data/"))) dir.create(here::here("cached_data/"))

# The lineup data used to draw residual plot needed to be downloaded
# from the github repo. Cache it.
if (!file.exists(here::here("cached_data/vi_lineup.rds"))) {
  vi_lineup <- get_vi_lineup()
  saveRDS(vi_lineup, here::here("cached_data/vi_lineup.rds"))
} else {
  vi_lineup <- readRDS(here::here("cached_data/vi_lineup.rds"))
}
```

```{r poly-conventional-simulation}
# Ensure the support of the predictor is [-1, 1]
stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1

# Run simulations to get the behaviours of conventional tests
if (!file.exists(here::here("cached_data/poly_conventional_simulation.rds"))) {
  poly_conv_sim <- list()

  # 100 bootstrap samples
  for (i in 1:100)
  {
    poly_conv_sim[[i]] <- 
      # Every sample contains 2000 lineups
      map(1:2000, function(i) {

        # Sample a set of parameters
        shape <- sample(1:4, 1)
        e_sigma <- sample(c(0.5, 1, 2, 4), 1)
        x_dist <- sample(c("uniform", 
                           "normal", 
                           "lognormal", 
                           "even_discrete"), 1)
        x <- switch(x_dist,
                    uniform = rand_uniform(-1, 1),
                    normal = {
                      raw_x <- rand_normal(sigma = 0.3)
                      closed_form(~stand_dist(raw_x))
                      },
                    lognormal = {
                      raw_x <- rand_lognormal(sigma = 0.6)
                      closed_form(~stand_dist(raw_x/3 - 1))
                      },
                    even_discrete = rand_uniform_d(k = 5, even = TRUE))

        # Build the model
        mod <- poly_model(shape, x = x, sigma = e_sigma)

        # Sample the number of observations
        n <- sample(c(50, 100, 300), 1)

        # Generate data from the model 
        tmp_dat <- mod$gen(n)

        # Return a data frame containing p-values of
        # F, RESET, BP and SW tests
        tibble(shape = shape,
               e_sigma = e_sigma,
               x_dist = x_dist,
               n = n,
               F_p_value = mod$test(tmp_dat)$p_value,
               RESET3_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:3, 
                                         power_type = "fitted")$p_value,
               RESET4_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:4, 
                                         power_type = "fitted")$p_value,
               RESET5_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:5, 
                                         power_type = "fitted")$p_value,
               RESET6_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:6, 
                                         power_type = "fitted")$p_value,
               RESET7_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:7, 
                                         power_type = "fitted")$p_value,
               RESET8_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:8, 
                                         power_type = "fitted")$p_value,
               RESET9_p_value = mod$test(tmp_dat, 
                                         test = "RESET", 
                                         power = 2:9, 
                                         power_type = "fitted")$p_value,
               RESET10_p_value = mod$test(tmp_dat, 
                                          test = "RESET", 
                                          power = 2:10, 
                                          power_type = "fitted")$p_value,
               BP_p_value = HETER_MODEL$test(tmp_dat)$p_value,
               SW_p_value = shapiro.test(tmp_dat$.resid)$p.value,
               boot_id = i)
        }) %>%
  reduce(bind_rows)
  }

  poly_conv_sim <- poly_conv_sim %>%
    reduce(bind_rows)

  saveRDS(poly_conv_sim, here::here("cached_data/poly_conventional_simulation.rds"))
} else {
  poly_conv_sim <- readRDS(here::here("cached_data/poly_conventional_simulation.rds"))
}
```

```{r heter-conventional-simulation}
# Run simulations to get the behaviours of conventional tests
if (!file.exists(here::here("cached_data/heter_conventional_simulation.rds"))) {
  heter_conv_sim <- list()

  # 100 bootstrap samples
  for (i in 1:100)
  {
    heter_conv_sim[[i]] <-
      # Every sample contains 2000 lineups
      map(1:2000, function(x) {

        # Sample a set of parameters
        a <- sample(c(-1, 0, 1), 1)
        b <- sample(c(0.25, 1, 4, 16, 64), 1)
        x_dist <- sample(c("uniform", 
                           "normal", 
                           "lognormal", 
                           "even_discrete"), 1)
        x <- switch(x_dist,
                    uniform = rand_uniform(-1, 1),
                    normal = {
                      raw_x <- rand_normal(sigma = 0.3)
                      closed_form(~stand_dist(raw_x))
                      },
                    lognormal = {
                      raw_x <- rand_lognormal(sigma = 0.6)
                      closed_form(~stand_dist(raw_x/3 - 1))
                      },
                    even_discrete = rand_uniform_d(-1, 1, k = 5, even = TRUE))

        # Build the model
        mod <- heter_model(a = a, b = b, x = x)

        # Sample the number of observations
        n <- sample(c(50, 100, 300), 1)

        # Generate data from the model 
        tmp_dat <- mod$gen(n)

        # Return a data frame containing p-values of
        # F, RESET, BP and SW tests
        tibble(a = a,
               b = b,
               x_dist = x_dist,
               n = n,
               F_p_value = POLY_MODEL$test(

                 # Create a pseudo z to be able to use F-test
                 tmp_dat %>%
                   mutate(z = poly_model()$
                            gen(n, computed = select(tmp_dat, x)) %>%
                            pull(z))
                 )$p_value,
               RESET3_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:3, 
                                                power_type = "fitted")$p_value,
               RESET4_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:4, 
                                                power_type = "fitted")$p_value,
               RESET5_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:5, 
                                                power_type = "fitted")$p_value,
               RESET6_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:6, 
                                                power_type = "fitted")$p_value,
               RESET7_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:7, 
                                                power_type = "fitted")$p_value,
               RESET8_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:8, 
                                                power_type = "fitted")$p_value,
               RESET9_p_value = POLY_MODEL$test(tmp_dat, 
                                                test = "RESET", 
                                                power = 2:9, 
                                                power_type = "fitted")$p_value,
               RESET10_p_value = POLY_MODEL$test(tmp_dat, 
                                                 test = "RESET", 
                                                 power = 2:10, 
                                                 power_type = "fitted")$p_value,
               BP_p_value = mod$test(tmp_dat)$p_value,
               SW_p_value = shapiro.test(tmp_dat$.resid)$p.value,
               boot_id = i)
        }) %>%
  reduce(bind_rows)
  }

  heter_conv_sim <- heter_conv_sim %>%
    reduce(bind_rows)

  saveRDS(heter_conv_sim, here::here("cached_data/heter_conventional_simulation.rds"))
} else {
  heter_conv_sim <- readRDS(here::here("cached_data/heter_conventional_simulation.rds"))
}
```

```{r}
# Borrow effect size from the survey
poly_conv_sim <- poly_conv_sim %>%
  left_join(select(filter(vi_survey, type == "polynomial"), shape, e_sigma, n, x_dist, effect_size))

heter_conv_sim <- heter_conv_sim %>%
  left_join(select(filter(vi_survey, type == "heteroskedasticity"), a, b, n, x_dist, effect_size))
```

## Additional Details of Testing Procedures {#sec-appendix-a-1}

### Statistical Significance

Within the context of visual inference, with $K$ independent observers, the visual $p$-value can be seen as the probability of having as many or more participants detect the data plot than the observed result.

The approach used in @majumder2013validation is as follows. Define $X_j = \{0,1\}$ to be a Bernoulli random variable measuring whether participant $j$ detected the data plot, and $X = \sum_{j=1}^{K}X_j$ be the total number of observers who detected the data plot. Then, by imposing a relatively strong assumption that all $K$ evaluations are fully independent, under $H_0$ $X \sim \mathrm{Binom}_{K,1/m}$. Therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is estimated with $P(X \geq x) = 1 - F(x) + f(x)$, where $F(.)$ is the binomial cumulative distribution function, $f(.)$ is the binomial probability mass function and $x$ is the realization of number of observers choosing the data plot.

As pointed out by @vanderplas2021statistical, this basic binomial model is deficient. It does not take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup, or account for when participants are offered the option to select one or more "most different" plots, or none, from a lineup. They suggest three common lineup scenarios: (1) $K$ different lineups are shown to $K$ participants, (2) $K$ lineups with different null plots but the same data plot are shown to $K$ participants, and (3) the same lineup is shown to $K$ participants. Scenario 3 is the most feasible to apply, but has the most dependencies to accommodate for the $p$-value calculation. For Scenario 3, VanderPlas et al propose modelling the probability of plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim \text{Dirichlet}(\alpha)$ for $i=1,...,m$ and $\alpha > 0$. The number of times plot $i$ being selected in $K$ evaluations is denoted as $c_i$. In case participant $j$ makes multiple selections, $1/s_j$ will be added to $c_i$ instead of one, where $s_j$ is the number of plots participant $j$ selected for $j=1,...K$. This ensures $\sum_{i}c_i=K$. Since we are only interested in the selections of the data plot $i$, the marginal model can be simplified to a beta-binomial model and thus the visual $p$-value is given as

$$
P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+
$$ {#eq-pvalue-beta-binomial}

\noindent where $B(.)$ is the beta function defined as

$$
B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0.
$$ {#eq-betafunction}

\noindent We extend the equation to non-negative real number $c_i$ by applying a linear approximation

$$
P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \in \mathbb{R}_0^+,
$$ {#eq-pvalue-beta-binomial-approx}

\noindent where $P(C \geq \lceil c_i \rceil)$ is calculated using @eq-pvalue-beta-binomial and $P(C = \lfloor c_i \rfloor)$ is calculated by

$$
P(C = c_i) = {K \choose c_i}\frac{B(c_i + \alpha, K - c_i + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+.
$$ {#eq-pmf-beta-binomial}

The parameter $\alpha$ used in @eq-pvalue-beta-binomial and @eq-pmf-beta-binomial is usually unknown and will need to be estimated from the survey data. An interpretation of $\alpha$ is that when it is low only a few plots are attractive to the observers and tend to be selected, and when high, most plots are equally likely to be chosen. VanderPlas et al define $c$-interesting plot to be if $c$ or more participants select the plot as the most different. The expected number of plots selected at least $c$ times, $\mathrm{E}[Z_c]$, is then calculated as

$$
\mathrm{E}[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).
$$ {#eq-c-interesting-expectation}

With @eq-c-interesting-expectation, $\alpha$ can be estimated using maximum likelihood estimation. Precise estimation of $\alpha$, is aided by evaluation of Rorschach lineups, where all plots are null plots. In a Rorschach, in theory all plots should be equally likely, but in practice some (irrelevant) visual elements may be more eye-catching than others. This is what $\alpha$ captures, the capacity for extraneous features to distract the observer for a particular type of plot display.

### Effect Size Derivation

Effect size can be defined as the difference of a parameter for a particular model or distribution,  or a statistic derived from a sample. Importantly, it needs to reflect the treatment we try to measure. Centred on a conventional statistical test, we usually can deduce the effect size from the test statistic by substituting the null parameter value. When considering the diagnostics of residual departures, there exist many possibilities of test statistics for a variety of model assumptions. Meanwhile, diagnostic plots such as the residual plot have no general agreement on measuring how strong a model violation pattern is. To build a bridge between various residual-based tests, and the visual test, we focus on the shared information embedded in the testing procedures, which is the distribution of residuals. When comes to comparison of distribution, Kullback-Leibler divergence [@kullback1951information] is a classical way to represent the information loss or entropy increase caused by the approximation to the true distribution, which in our case, the inefficiency due to the use of false model assumptions.

Following the terminology introduced by @kullback1951information, $P$ represents the measured probability distribution, and $Q$ represents the assumed probability distribution. The Kullback-Leibler divergence is defined as $\int_{-\infty}^{\infty}\log(p(x)/q(x))p(x)dx$, where $p(.)$ and $q(.)$ denote probability densities of $P$ and $Q$.

Let $\boldsymbol{X}$ denotes the $p + 1$ predictors with $n$ observations, $\boldsymbol{b} = (\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\boldsymbol{y}$ denotes the OLS solution, $\boldsymbol{R} = \boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ denotes the residual operator, and let $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0},\sigma^2\boldsymbol{I})$ denotes the error. The residual vector

\begin{align*}
\boldsymbol{e} &= \boldsymbol{y} - \boldsymbol{X}\boldsymbol{b} \\
               &= \boldsymbol{y} - \boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\boldsymbol{y} \\
               &= (\boldsymbol{I} -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}')\boldsymbol{y} \\
               &= \boldsymbol{R}\boldsymbol{y} \\
               &= \boldsymbol{R}(\boldsymbol{X}\boldsymbol{\beta} + \boldsymbol{\varepsilon}) \\
               &= \boldsymbol{R}\boldsymbol{\varepsilon}.
\end{align*}

Because $\text{rank}(\boldsymbol{R}) = n - p - 1 < n$, $\boldsymbol{e}$ follows a degenerate multivariate normal distribution and does not have a density. Since the Kullback-Leibler divergence requires a proper density function, we need to simplify the covariance matrix of $\boldsymbol{e}$ by setting all the off-diagonal elements to 0. Then, the residuals will be assumed to follow $N(\boldsymbol{0}, diag(\boldsymbol{R}\sigma^2))$ under the null hypothesis that the model is correctly specified. If the model is however misspecified due to omitted variables $\boldsymbol{Z}$, or a non-constant variance $\boldsymbol{V}$, the distribution of residuals can be derived as $N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, \text{diag}(\boldsymbol{R}\sigma^2))$ and $N(\boldsymbol{0}, \text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}'))$ respectively.

By assuming both $P$ and $Q$ are multivariate normal density functions, the Kullback-Leibler divergence can be rewritten as 
$$KL = \frac{1}{2}\left(\log\frac{|\Sigma_p|}{|\Sigma_q|} - n + \text{tr}(\Sigma_p^{-1}\Sigma_q) + (\mu_p - \mu_q)'\Sigma_p^{-1}(\mu_p - \mu_q)\right).$$

Then, we can combine the two residual departures into one formula

\small

$$
KL = \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}\text{diag}(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z'(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}\boldsymbol{\mu}_z\right).
$$

\normalsize

When there are omitted variables but constant error variance, the formula can be reduced to

$$KL = \frac{1}{2}\left(\boldsymbol{\mu}_z'(\text{diag}(\boldsymbol{R}\sigma^2))^{-1}\boldsymbol{\mu}_z\right).$$

And when the model equation is correctly specified but the error variance is non-constant, the formula can be reduced to

$$KL = \frac{1}{2}\left(log\frac{|\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}')^{-1}
\text{diag}(\boldsymbol{R}\sigma^2))\right).$$

To compute the effect size for each lineup we simulate a sufficiently large number of samples from the same model with the number of observations $n$ fixed for each sample. We then compute the effect size for each sample and take the average as the final value. This ensures lineups constructed with the same experimental factors will share the same effect size.

### Sensitivity Analysis for $\alpha$

The parameter $\alpha$ used for the $p$-value calculation needs to be estimated from responses to null lineups. With a greater value of $\hat{\alpha}$, the $p$-value will be smaller, resulting in more lineups being rejected. However, The way we generate Rorschach lineup is not strictly the same as what suggested in @vanderplas2021statistical and @buja2009statistical. Therefore, we conduct a sensitivity analysis in this section to examine the impact of the variation of the estimator $\alpha$ on our primary findings.

The analysis is conducted by setting up several scenarios, where the $\alpha$ is under or overestimated by 12.5%, 25% and 50%. Using the adjusted $\hat{\alpha}$, we recalculate the $p$-value for every lineup and show the results in Figure \ref{fig:sensitivity}. It can be observed that there are some changes to $p$-values, especially when the $\hat{\alpha}$ is multiplied by 50%. However, Table \ref{tab:sensitivity} shows that adjusting $\hat{\alpha}$ will not result in a huge difference in rejection decisions. There are only a small percentage of cases where the rejection decision change. It is very unlikely the downstream findings will be affected because of the estimate of $\alpha$.

```{r}
upper_p <- map_df(c(0.125, 0.25, 0.5), function(modified) {
  calc_p_value_multi(vi_survey %>%
                       filter(!attention_check) %>%
                       filter(!null_lineup) %>%
                       filter(x_dist == "uniform") %>%
                       mutate(alpha_upper = alpha * (1 + modified)),
                     lineup_id = unique_lineup_id,
                     detect = detect,
                     n_sel = num_selection,
                     alpha = alpha_upper) %>%
  mutate(modified = 1 + modified) %>%
  mutate(adjusted_p_value = p_value)
})

lower_p <- map_df(c(0.125, 0.25, 0.5), function(modified) {
  calc_p_value_multi(vi_survey %>%
                       filter(!attention_check) %>%
                       filter(!null_lineup) %>%
                       filter(x_dist == "uniform") %>%
                       mutate(alpha_lower = alpha * (1 - modified)),
                     lineup_id = unique_lineup_id,
                     detect = detect,
                     n_sel = num_selection,
                     alpha = alpha_lower) %>%
  mutate(modified = 1 - modified) %>%
  mutate(adjusted_p_value = p_value)
})

upper_lower_p <- upper_p %>%
  bind_rows(lower_p) %>%
  select(-p_value)
```

```{r fig-sensitivity, fig.cap = "Change of $p$-values with $\\hat{\\alpha}$ multiplied by $0.5$, $0.75$, $0.875$, $1.125$, $1.25$ and $1.5$. Only lineups with uniform fitted value distribution is used. The vertical dashed line indicates $p\\text{-value} = 0.05$. The area coloured in red indicates $\\text{adjusted }p\\text{-value} < 0.05$. The x-axis is drawn on logarithmic scale. For multipliers smaller than 1, the adjusted $p\\text{-value}$ will initially increase and decline when the $p\\text{-value}$ increases. The trend is opposite with multipliers greater than $1$, but the difference eventually reaches $0$."}

smallest_p_value <- vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform") %>%
  pull(p_value) %>%
  min()

upper_lower_p %>%
  left_join(vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(p_value = first(p_value))) %>%
  mutate(diff_p_value = adjusted_p_value - p_value) %>%
  ggplot() +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0.05, col = "black", linetype = 2, alpha = 0.7) +
  geom_ribbon(data = expand.grid(x = 10^seq(-10, 0, 0.01),
                                y = seq(-0.1, 0.1, 0.0001)) %>%
              filter(x + y < 0.05) %>%
              group_by(x) %>%
              summarise(y = max(y)),
            aes(x, ymax = y, ymin = -0.06),
            fill = "red",
            alpha = 0.2) +
  geom_point(aes(p_value, diff_p_value), alpha = 0.2) +
  facet_wrap(~modified) +
  ylab(expression("Adjusted"~italic(p)*"-value minus"~italic(p)*"-value")) +
  xlab(expression(italic(p)*"-value")) +
  theme_light() +
  scale_x_continuous(breaks = c(0.0001, 0.001, 0.01, 0.1, 1),
                     minor_breaks = NULL,
                     trans = "log10",
                     labels = c("0.01%", "0.1%", "1%", "10%", "100%"),
                     limits = c(smallest_p_value, 1)) +
  scale_y_continuous(limits = c(-0.06, 0.06))
```

```{r}
upper_lower_p %>%
  left_join(vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(p_value = first(p_value))) %>%
  group_by(modified) %>%
  count(adjusted_reject = adjusted_p_value <= 0.05, reject = p_value <= 0.05) %>%
  complete(adjusted_reject, reject, fill = list(n = 0)) %>%
  ungroup() %>%
  filter(adjusted_reject != reject) %>%
  mutate(case = ifelse(!adjusted_reject, "reject to not reject", "not reject to reject")) %>%
  rename(multiplier = modified) %>%
  select(-adjusted_reject, -reject) %>%
  pivot_wider(names_from = case, values_from = n) %>%
  mutate(reject_not_reject_per = round(`reject to not reject`/279*100, 2)) %>%
  mutate(not_reject_reject_per = round(`not reject to reject`/279*100, 2)) %>%
  select(multiplier,
         `reject to not reject`,
         reject_not_reject_per,
         `not reject to reject`,
         not_reject_reject_per) %>%
  kableExtra::kable(format = "latex", 
                    booktabs = TRUE,
                    col.names = c("Multiplier",
                                  "Reject to not reject",
                                  "%",
                                  "Not reject to reject",
                                  "%"),
                    label = "sensitivity", 
                    caption = "Examining how decisions might change if $\\hat{\\alpha}$ was different. Percentage of lineups that where the $p$-value would switch to above or below 0.05, when $\\hat{\\alpha}$ is multiplied by a multiplier.",
                    linesep = c("", "", "\\addlinespace"))
```

### Effect of Number of Evaluations on the Power of a Visual Test

When comparing power of visual tests across different fitted value distributions, we have discussed the number of evaluations on a lineup will affect the power of the visual test. Using the lineups with uniform fitted value distribution, we show in @fig-n_eval the change of power of visual tests due to different number of evaluations. It can be learned that as the number of evaluations increases, the power will increase but the margin will decrease. Considering we have eleven evaluations on lineups with uniform fitted value distribution, and five evaluations on other lineups, it is necessary to use the same number of evaluations for each lineup in comparison.

```{r}
min_poly_es <- vi_survey %>% 
  filter(!null_lineup, 
         !attention_check,
         type == "polynomial", 
         x_dist == "uniform") %>%
  pull(effect_size) %>%
  min()
max_poly_es <- vi_survey %>% 
  filter(!null_lineup, 
         !attention_check,
         type == "polynomial", 
         x_dist == "uniform") %>%
  pull(effect_size) %>%
  max()
```

```{r fig-n_eval, fig.cap = "Change of power of visual tests for different number of evalutions on lineups with uniform fitted value distribution. The power will increase as the number of evaluations increases, but the margin will decrease.", fig.width=6, fig.height=4, out.width="70%"}
eval_n <- function(n_eval, boot) {
  map_df(1:boot, function(boot_id) {
  vi_survey %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  filter(x_dist == "uniform") %>%
  filter(type == "polynomial") %>%
  group_by(unique_lineup_id) %>%
  slice_sample(n = n_eval) %>%
  ungroup() %>%
  calc_p_value_multi(lineup_id = unique_lineup_id,
                     detect = detect,
                     n_sel = num_selection,
                     alpha = alpha) %>%
  mutate(boot_id = boot_id)
}) %>%
  left_join(select(vi_survey, unique_lineup_id, effect_size)) %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  select(-unique_lineup_id, -p_value) %>%
  mutate(n_eval = n_eval)
}

map_df(seq(1, 11, 2), ~eval_n(.x, 1)) %>%
  select(-boot_id) %>%
  nest(c(effect_size, reject, offset0)) %>%
  mutate(mod = map(data, 
                   ~glm(reject ~ effect_size - 1, 
                        family = binomial(), 
                        data = .x,
                        offset = offset0))) %>%
  mutate(power = map(mod, function(mod) {
    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
                         offset0 = log(0.05/0.95))
    result$power <- predict(mod, type = "response", newdata = result)
    result
  })) %>%
  select(-data, -mod) %>%
  unnest(power) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_line(aes(log_effect_size, power, col = factor(n_eval)),
            size = 1) +
  labs(col = "#Evaluations") +
  ylab("Power") +
  xlab(expression(log[e] (Effect_size))) +
  theme_light() +
  scale_color_brewer(palette = "Reds")
```

### Power of a RESET Test under Different Auxiliary Regression Formulas

It is found in the result that the power of a RESET test will be affected by the highest order of fitted values included in the auxiliary formula. And we suspect that the current recommendation of the highest order - four, is insufficient to test complex non-linear structures such as the "triple-U" shape designed in this paper. @fig-reset illustrates the change of power of RESET test while testing the "U" shape and the "triple-U" shape with different highest orders. Clearly, when testing a simple shape like the "U" shape, the highest order has very little impact on the power. But for testing the "triple-U" shape, there will be a loss of power if the recommended order is used. To avoid the loss of power, the highest order needs to be set to at least six.

```{r cache = TRUE}
reset_diff_dat <- poly_conv_sim %>%
  filter(x_dist == "uniform") %>%
  filter(shape == 4) %>%
  select(-F_p_value, -BP_p_value, -SW_p_value) %>%
  pivot_longer(RESET3_p_value:RESET10_p_value) %>%
  mutate(name = gsub("RESET(.*)_p_value", "\\1", name)) %>%
  mutate(reject = value <= 0.05) %>%
  select(effect_size, name, reject) %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  nest(dat = c(effect_size, offset0, reject)) %>%
  mutate(mod = map(dat, 
                   ~glm(reject ~ effect_size - 1, 
                        family = binomial(), 
                        data = .x,
                        offset = offset0))) %>%
  mutate(power = map(mod, function(mod) {
    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
                         offset0 = log(0.05/0.95))
    result$power <- predict(mod, type = "response", newdata = result)
    result
  })) %>%
  select(-dat, -mod) %>%
  unnest(power) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(shape = 4)
```

```{r cache = TRUE}
reset_diff_dat2 <- poly_conv_sim %>%
  filter(x_dist == "uniform") %>%
  filter(shape == 1) %>%
  select(-F_p_value, -BP_p_value, -SW_p_value) %>%
  pivot_longer(RESET3_p_value:RESET10_p_value) %>%
  mutate(name = gsub("RESET(.*)_p_value", "\\1", name)) %>%
  mutate(reject = value <= 0.05) %>%
  select(effect_size, name, reject) %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  nest(dat = c(effect_size, offset0, reject)) %>%
  mutate(mod = map(dat, 
                   ~glm(reject ~ effect_size - 1, 
                        family = binomial(), 
                        data = .x,
                        offset = offset0))) %>%
  mutate(power = map(mod, function(mod) {
    result <- data.frame(effect_size = seq(min_poly_es, max_poly_es, 0.1),
                         offset0 = log(0.05/0.95))
    result$power <- predict(mod, type = "response", newdata = result)
    result
  })) %>%
  select(-dat, -mod) %>%
  unnest(power) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(shape = 1)
```

```{r fig-reset, fig.cap = 'Change of power of RESET tests for different orders of fitted values included in the auxiliary formula. The left panel is the power of testing the "U" shape and the right panel is the power of testing the "triple-U" shape. The power will not be greatly affected by the highest order in the case of testing the "U" shape. In the case of testing the "triple-U" shape, the highest order needs to be set to at least six to avoid the loss of power.', fig.height = 4}
reset_diff_dat %>%
  bind_rows(reset_diff_dat2) %>%
  mutate(shape = factor(ifelse(shape == 1, "U", "triple-U"), levels = c("U", "triple-U"))) %>%
  ggplot() +
  geom_line(aes(log_effect_size, power, col = as.factor(as.integer(name))),
            size = 1) +
  theme_light() +
  labs(col = "Highest order") +
  xlab(expression(log[e] (Effect_size))) +
  scale_color_brewer(palette = "Reds") +
  facet_wrap(~shape) +
  ylab("Power")
```

### Conventional Test Rejection Rate for Varying Significance Levels

In the main paper, @sec-power-analysis and @sec-p-value compared the power, and the decisions made by the conventional tests and the visual test. The power curves for the visual test is effectively a right-shift from the conventional test. The effect is that the visual test rejects less often than the conventional test, at the same significance level. We also saw that the visual test rejected a subset of those that the conventional tests rejected. This means that they agreed quite well - only residual plots rejected by the conventional tests were rejected by the visual test. There was little disagreement, where residual plots not rejected by the conventional test were rejected by the visual test. The question arises whether the decisions made conventional test could be made similar to that of the visual test by reducing the significance level. Reducing the significance level from 0.05, to 0.01, 0.001, ... will have the effect of rejecting fewer of the residual plots.

It would be interesting if a different conventional test significance level results in both the visual tests and conventional tests reject only the same residual plots, and fails to reject the same residual plots. This would be a state where both systems agree perfectly. Figure \ref{fig:rej-rates} examines this. Plot A shows the percentage of residual plots rejected by the visual test, given the conventional test rejected (solid lines) or failed to reject (dashed lines). The vertical grey line marks significance level 0.05. When the significance level gets smaller, it is possible to see that the visual tests reject (nearly) 100% of the time that the conventional test rejects. However, there is no agreement, because the visual tests also increasingly reject residual plots where the conventional test failed to reject. Plot B is comparable to an ROC curve, where the percentage visual test rejection conditional on conventional test decision is plotted: Reject conditional on reject is plotted against reject conditional on fail to reject, for different significance levels. The non-linearity pattern results are close to being ideal, that the percentage of reject relative to fail to reject increases very slowly as the reject relative to reject converges to 100. The heteroskedasticity pattern is more problematic, and shows that the cost of rejecting less with the conventional test is disagreement with the visual test.

```{r fig-rej-rates, fig.pos="t!", fig.width=8, fig.height=5, out.width="100%", fig.cap="Changing the significance level of the conventional test will change the rejection rate. A: Percentage of conventional tests also rejected by the visual test: rejected (solid), not rejected (dashed). As the significance level is reduced, the percentage rejected by the visual test that has been rejected by the conventional test approaches 100. The percentage of tests not rejected by the conventional test also increases, as expected, but the percentage of these that are rejected by the visual test increases too. B: ROC curve shows that forcing the conventional test to not reject creates a discrepancy with the visual test. Many of the residual plots not rejected by the conventional test are rejected by the visual test. It is not possible to vary the significance level of the conventional test to match the decisions made by the visual test."}
# compute rejection rate for different significance levels, holding visual test constant
library(ggpubr)

nl_rates <- NULL
h_rates <- NULL

for (i in c(0.1, 0.05, 0.01, 0.001, 0.0001, 0.00001, 0.000001, 0.0000001, 0.00000001, 0.000000001,
            0.0000000001, 0.00000000001)) {

p_value_cmp_dat <- vi_survey %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  filter(x_dist == "uniform") %>%
  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
  mutate(conv_reject = ifelse(conventional_p_value <= i, "Reject", "Not"),
         reject = ifelse(p_value <= 0.05, "Reject", "Not")) %>%
  mutate(conv_reject = factor(conv_reject, levels = c("Reject", "Not")),
         reject = factor(reject, levels = c("Reject", "Not"))) %>%
  select(type, conv_reject, reject)

nl <- p_value_cmp_dat %>% 
  filter(type == "non-linearity") %>% 
  count(conv_reject, reject, .drop=FALSE) %>%
  summarise(rej_pct = round(n[1]/(n[1]+n[2])*100, 0),
            nrej_pct = round(n[3]/(n[3]+n[4])*100, 0)) %>%
  bind_cols(alpha = i)

nl_rates <- bind_rows(nl_rates, nl)

# Heteroskedasticity 
h <- p_value_cmp_dat %>% 
  filter(type == "heteroskedasticity") %>% 
  count(conv_reject, reject, .drop = FALSE) %>%
  summarise(rej_pct = round(n[1]/(n[1]+n[2])*100, 0),
            nrej_pct = round(n[3]/(n[3]+n[4])*100, 0)) %>%
  bind_cols(alpha = i)

h_rates <- bind_rows(h_rates, h)
}

nl_rates <- nl_rates %>% mutate(type = "non-linearity")
h_rates <- h_rates %>% mutate(type = "heteroskedasticity")
rates <- bind_rows(nl_rates, h_rates) %>%
  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity")))

rej_by_sig <- ggplot(rates) +
  geom_vline(xintercept = 0.05, colour = "grey50") +
  geom_line(aes(x = alpha, y = rej_pct, colour = type), alpha = 0.8) +
  geom_line(aes(x = alpha, y = nrej_pct, colour = type), linetype = 2) +
  scale_x_log10(breaks = c(0.1, 0.001, 0.00001, 0.0000001, 0.000000001, 0.00000000001)) + 
  ylab("Visual test rejects (%)") + xlab("Significance level") +
  scale_colour_brewer("", palette = "Dark2") +
  ggtitle("A") +
  theme_light()

roc <- ggplot(rates) +
  geom_path(aes(x = nrej_pct, y = rej_pct, colour=type)) +
  ylab("Reject | Reject (%)") + xlab("Reject | Fail to reject (%)") +
  ylim(0, 100) +
  scale_colour_brewer("", palette = "Dark2") +
  coord_equal() +
  ggtitle("B") +
  theme_light()

ggarrange(rej_by_sig, roc, ncol = 2, common.legend = TRUE, legend = "bottom")
```

## Additional Details of Experimental Setup {#sec-appendix-a-2}

### Non-linearity Model

The non-linearity model used in the experiment includes a non-linear term $\boldsymbol{z}$ constructed using Hermite polynomials on random vector $\boldsymbol{x}$ formulated as

\begin{align*}
\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
\boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\
\boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)),
\end{align*}

\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vectors of size $n$, $\boldsymbol{1}_n$ is a vector of ones of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials [@hermite1864nouveau; originally by @de1820theorie], $\varepsilon \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as

$$
g(\boldsymbol{x}, k) = 2k\cdot\frac{\boldsymbol{x} - x_{min}\boldsymbol{1}_n}{x_{max} - x_{min}} - k, \quad \text{for} \quad k > 0. 
$$ {#eq-scaling-function}

\noindent where $\displaystyle x_{min} = \min_{i\in \{1, ..., n\}} x_i$, $\displaystyle x_{max} = \max_{i\in \{1, ..., n\}} x_i$ and $x_i$ is the $i$-th entry of $\boldsymbol{x}$. The function `hermite` from the R package `mpoly` [@mpoly] is used to simulate $\boldsymbol{z}_{raw}$ to generate Hermite polynomials.

The null regression model used to fit the realizations generated by the above model is formulated as

$$
\boldsymbol{y} = \beta_0\boldsymbol{1}_n + \beta_1 \boldsymbol{x} + \boldsymbol{u},
$$ {#eq-null-model}

\noindent where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$. Here $\boldsymbol{z}$ is a higher order term of $\boldsymbol{x}$ and leaving it out in the null regression results in model misspecification.

### Heteroskedasticity Model

The heteroskedasticity model used in the experiment is formulated as

\begin{align*}
\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}_n, \boldsymbol{1}_n + (2 - |a|)(\boldsymbol{x} - a\boldsymbol{1}_n)'(\boldsymbol{x} - a\boldsymbol{1}_n) b \boldsymbol{I}_n), 
\end{align*}

\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vectors of size $n$ and $g(.)$ is the scaling function defined in @eq-scaling-function. The null regression model used to fit the realizations generated by the above model is formulated exactly the same as @eq-null-model.

For $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the predictor $\boldsymbol{x}$, which will lead to the presence of heteroskedasticity.

Since $\text{supp}(X) = [-1, 1]$, choosing $a$ to be $-1$, $0$ and $1$ can generate "left-triangle", "butterfly" and "right-triangle" shapes. The term $(2 - |a|)$ maintains the magnitude of residuals across different values of $a$.

### Controlling the Signal Strength

The three parameters $n$, $\sigma$ and $b$ are important for controlling the strength of the signal to generate lineups with a variety difficulty levels. This will ensure that  estimated power curves will be smooth and continuous, and that participants are allocated a set of lineups with a range of difficulty.

Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ are used in data collection periods I and II respectively. A large value of $\sigma$ will increase the variation of the error of the non-linearity model and decrease the visibility of the visual pattern. 
The parameter $b$ controls the variation in the standard deviation of the error across the support of the predictor. A larger value of $b$ generates a larger ratio between it is smallest and highest values, making the visual pattern more obvious. 
The sample size $n$ sharpens (or blurs) a pattern when it is larger (or smaller).

### Lineup Allocation to Participants

There are a total of $4 \times 4 \times 3 \times 4 = 192$ and $3 \times 5 \times 3 \times 4 = 180$ combinations of parameter values for the non-linearity model and heteroskedasticity models respectively. Three replications for each combination results in $192 \times 3 = 576$ and $180 \times 3 = 540$ lineups, respectively.

Each lineup needs to be evaluated by at least five participants. From previous work, and additional pilot studies for this experiment, we decided to record evaluations from 20 lineups for each participant. Two of the 20 lineups with clear visual patterns were used as attention checks, to ensure quality data. Thus, $576 \times 5 / (20-2) = 160$ and $540 \times 5 / (20-2) = 150$ participants were needed to cover the experimental design for the data collection periods I and II, respectively. The factor levels and range of difficulty was assigned relatively equally among participants.

Data collection period III was primarily to obtain Rorschach lineup evaluations to estimate $\alpha$, and also to obtain additional evaluations of lineups made with uniform fitted value distribution to ensure consistency in the results. To construct a Rorschach lineup, the data is generated from a model with zero effect size, while the null data are generated using the residual rotation technique. This procedure differs from that of the canonical Rorschach lineup, where all 20 plots are generated directly from the null model, so the method suggested in @vanderplas2021statistical for typical lineups containing a data plot is used to estimate $\alpha$. The $3 \times 4 = 12$ treatment levels of the common factors, replicated three times results in 36 lineups. And 6 more evaluations on the 279 lineups with uniform fitted value distribution, results in at least $(36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133$ participants needed.

### Mapping of Participants to Experimental Factors

Mapping of participants to experimental factors is an important part of experimental design. Essentially, we want to maximum the difference in factors exposed to a participant. For this purpose, we design an algorithm to conduct participant allocation. Let $L$ be a set of available lineups and $S$ be a set of available participants. According to the experimental design, the availability of a lineup is associated with the number of participants it can assign to. For lineups with uniform fitted value distribution, this value is 11. And other lineups can be allocated to at most five different participants. The availability of a participant is associated with the number of lineups that being allocated to this participant. A participant can view at most 18 different lineups.

The algorithm starts from picking a random participant $s \in S$ with the minimum number of allocated lineups. It then tries to find a lineup $l \in L$ that can maximise the distance metric $D$ and allocate it to participant $s$. Set $L$ and $S$ will be updated and the picking process will be repeated until there is no available lineups or participants.

Let $F_1,...,F_q$ be $q$ experimental factors, and $f_1, ...,f_q$ be the corresponding factor values. We say $f_i$ exists in $L_{s}$ if any lineup in $L_{s}$ has this factor value. Similarly, $f_if_j$ exists in $L_{s}$ if any lineup in $L_{s}$ has this pair of factor values. And $f_if_jf_k$ exists in $L_{s}$ if any lineup in $L_{s}$ has this trio of factor values. The distance metric $D$ is defined between a lineup $l$ and a set of lineups $L_{s}$ allocated to a participant $s$ if $L_{s}$ is non-empty:

\footnotesize

\begin{equation*}
D =
C - \sum_{1\leq i \leq q}I(f_i\text{ exists in }L_{s}) - \sum_{\substack{1\leq i \leq q-1 \\ i < j \leq q}}I(f_if_j\text{ exists in }L_{s}) - \sum_{\substack{1\leq i \leq q - 2 \\ i < j \leq q - 1 \\ j < k \leq q}}I(f_if_jf_k\text{ exists in }L_{s})
\end{equation*}

\normalsize
where $C$ is a sufficiently large constant such that $D > 0$. If $L_{s}$ is empty, we define $D = 0$.

The distance measures how different a lineup is from the set of lineups allocated to the participant in terms of factor values. Thus, the algorithm will try to allocate the most different lineup to a participant at each step.

In @fig-poly-allocate and @fig-heter-allocate, we present examples of lineup allocations generated by our algorithm for data collection periods I and II, respectively. These figures illustrate the factor values presented to the first 20 participants recruited during each period. It can be observed that the algorithm effectively distributed almost all possible one-way and two-way factor combinations among the participants, thereby ensuring a diverse set of lineups for each participant.

```{r fig-poly-allocate, fig.cap = 'Factor values assigned to the first 20 participants recruited during data collection period I, with data points slightly jittered to prevent overlap. Each participant have been exposed to all one-way and two-way factor combinations.'}
vi_survey %>%
  filter(exp == 3) %>%
  filter(!attention_check) %>%
  filter(set <= 20) %>%
  mutate(shape = c("U", "S", "M", "triple-U")[shape]) %>%
  mutate(shape = factor(shape, levels = c("U", "S", "M", "triple-U"))) %>%
  mutate(e_sigma = factor(e_sigma)) %>%
  mutate(n = factor(n)) %>%
  ggplot() +
  geom_jitter(aes(e_sigma, shape, col = n),
             width = 0.2,
             height = 0.2,
             alpha = 0.5,
             size = 3) +
  facet_wrap(~set) +
  scale_color_brewer(palette = "Dark2") +
  theme_light() +
  ylab("Non-linearity pattern") +
  xlab(expression(sigma)) +
  labs(col = "Sample size")
```

```{r fig-heter-allocate, fig.cap = 'Factor values assigned to the first 20 participants recruited during data collection period II, with data points slightly jittered to prevent overlap. Each participant appears to have been exposed to almost all one-way and two-way factor combinations. However, two participants, namely participant 16 and participant 18, were not presented with lineups containing a butterfly shape with $b = 4$. Additionally, participant 7 did not encounter a lineup featuring a butterfly shape with $b = 1$.'}
vi_survey %>%
  filter(exp == 4) %>%
  filter(!attention_check, !null_lineup) %>%
  filter(set <= 20) %>%
  mutate(a = c("left-triangle", "butterfly", "right-triangle")[a + 2]) %>%
  mutate(a = factor(a, levels = c("left-triangle", "butterfly", "right-triangle"))) %>%
  mutate(b = factor(b)) %>%
  mutate(n = factor(n)) %>%
  ggplot() +
  geom_jitter(aes(b, a, col = n),
             width = 0.2,
             height = 0.2,
             alpha = 0.5,
             size = 3) +
  facet_wrap(~set) +
  scale_color_brewer(palette = "Dark2") +
  theme_light() +
  ylab("Heteroskedasticity pattern") +
  xlab("b") +
  labs(col = "Sample size")
```

### Data Collection Process

The survey data is collected via a self-hosted website designed by us. The complete architecture is provided in @fig-tech. The website is built with the `Flask` [@flask] web framework and hosted on `PythonAnywhere` [@pythonanywhere]. It is configured to handle HTTP requests such that participants can correctly receive webpages and submit responses. Embedded in the resources sent to participants, the `jsPsych` front-end framework [@jspsych] instructs participants' browsers to render an environment for running behavioural experiments. During the experiment, this framework will automatically collect common behavioural data such as response time and clicks on buttons. participants' responses are first validated by a scheduled `Python` script run on the server, then push to a Github repository. Lineup images shown to users are saved in multiple Github repositories and hosted in corresponding Github pages. The URLs to these images are resolved by `Flask` and bundled in HTML files.

Once the participant is recruited from Prolific [@palan2018prolific], it will be redirected to the entry page of our study website. An image of the entry page is provided in @fig-entry-page. Then, the participant needs to submit the online consent form and fill in the demographic information as shown in @fig-consent-form and @fig-metadata respectively. Before evaluating lineups, participant also need to read the training page as provide in @fig-training-page to understand the process. An example of the lineup page is given in @fig-lineup-page. A half of the page is taken by the lineup image to attract participant's attention. The button to skip the selections for the current lineup is intentionally put in the corner of the bounding box with smaller font size, such that participants will not misuse this functionality.

::: {.content-visible when-format="html"}

::: {#fig-tech}

<embed src="figures/experiment_tech.pdf" class="img-fluid" style="width:0%; display:none"/>

![](figures/experiment_tech.pdf#toolbar=0){width=100.0% height=953px}

Diagram of online experimental setup. The server-side of the study website uses Flask as backend hosted on PythonAnywhere. And the client-side uses jsPsych to run experiment.
:::

:::

::: {.content-visible when-format="pdf"}

```{r fig-tech, out.width = "100%", fig.cap = "Diagram of online experimental setup. The server-side of the study website uses Flask as backend hosted on PythonAnywhere. And the client-side uses jsPsych to run experiment."}
knitr::include_graphics("figures/experiment_tech.pdf")
```

:::

```{r fig-entry-page, out.width = "100%", fig.cap = "The entry page of the study website."}
knitr::include_graphics("figures/message.png")
```

```{r fig-consent-form, out.width = "100%", fig.cap = "The consent form provided in the study website."}
knitr::include_graphics("figures/consent_form.png")
```

```{r fig-metadata, out.width = "100%", fig.cap = "The form to provide demographic information."}
knitr::include_graphics("figures/metadata.png")
```

```{r fig-training-page, out.width = "100%", fig.cap = "The training page of the study website."}
knitr::include_graphics("figures/training.png")
```

```{r fig-lineup-page, out.width = "100%", fig.cap = "The lineup page of the study website."}
knitr::include_graphics("figures/lineup1.png")
```

## Analysis of Results Relative to Data Collection Process

### Demographics

Throughout the study, we have collected `r vi_survey %>% filter(!attention_check, !null_lineup) %>% nrow()` evaluations on `r vi_survey %>% filter(!attention_check, !null_lineup) %>% count(unique_lineup_id) %>% nrow()` non-null lineups. @tbl-count-lineup gives further details about the number of evaluations, lineups and participants over pattern types and data collection periods.

```{r}
#| label: tbl-count-lineup
#| tbl-cap: 'Count of lineups, evaluations and participants over departure types and data collection periods.'

# Calculate these numbers
a1 <- vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(type == "polynomial") %>%
  filter(exp == 3) %>%
  count(unique_lineup_id) %>%
  nrow()

a2 <- vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(type == "polynomial") %>%
  filter(exp == 4) %>%
  count(unique_lineup_id) %>%
  nrow()

a3 <- vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(type == "polynomial") %>%
  filter(exp == 5) %>%
  count(unique_lineup_id) %>%
  nrow()

a4 <- vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(type != "polynomial") %>%
  filter(exp == 3) %>%
  count(unique_lineup_id) %>%
  nrow()

a5 <- vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(type != "polynomial") %>%
  filter(exp == 4) %>%
  count(unique_lineup_id) %>%
  nrow()

a6 <- vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(type != "polynomial") %>%
  filter(exp == 5) %>%
  count(unique_lineup_id) %>%
  nrow()

a7 <- vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  count(unique_lineup_id) %>%
  nrow()

b1 <- vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(type == "polynomial") %>%
  filter(exp == 3) %>%
  nrow()

b2 <- vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(type == "polynomial") %>%
  filter(exp == 4) %>%
  nrow()

b3 <- vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(type == "polynomial") %>%
  filter(exp == 5) %>%
  nrow()

b4 <- vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(type != "polynomial") %>%
  filter(exp == 3) %>%
  nrow()

b5 <- vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(type != "polynomial") %>%
  filter(exp == 4) %>%
  nrow()

b6 <- vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(type != "polynomial") %>%
  filter(exp == 5) %>%
  nrow()

b7 <- vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  nrow()

c1 <- vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(type == "polynomial") %>%
  filter(exp == 3) %>%
  count(set) %>%
  nrow()

c2 <- vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(type == "polynomial") %>%
  filter(exp == 4) %>%
  count(set) %>%
  nrow()

c3 <- vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(type == "polynomial") %>%
  filter(exp == 5) %>%
  count(set) %>%
  nrow()

c4 <- vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(type != "polynomial") %>%
  filter(exp == 3) %>%
  count(set) %>%
  nrow()

c5 <- vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(type != "polynomial") %>%
  filter(exp == 4) %>%
  count(set) %>%
  nrow()

c6 <- vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(type != "polynomial") %>%
  filter(exp == 5) %>%
  count(set) %>%
  nrow()

c7 <- vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  count(exp, set) %>%
  nrow()

matrix(c(c(a1, a2, a3, a4, a5, a6, a7),
         c(b1, b2, b3, b4, b5, b6, b7),
         c(c1, c2, c3, c4, c5, c6, c7)),
       ncol = 7,
       byrow = TRUE) %>%
  data.frame() %>%
  add_column(name = c("Lineups", "Evaluations", "Participants"), .before = 1) %>%
  kableExtra::kable(booktabs = TRUE,
                    col.names = c("Number", "I", "II", "III", "I", "II", "II", "Total")) %>%
  kableExtra::add_header_above(c(" " = 1, "Non-linearity" = 3, "Heteroskedasticity" = 3, " " = 1))
```

Along with the responses to lineups, we have collected a series of demographic information including age, pronoun, education background and previous experience in studies involved data visualization. @tbl-pronoun, @tbl-age-group, @tbl-education and @tbl-experience provide summary of the demographic data.

It can be observed from the tables that most participants have Diploma or Bachelor degrees, followed by High school or below and the survey data is gender balanced. Majority of participants are between 18 to 39 years old and there are slightly more participants who do not have previous experience than those who have.

```{r}
#| label: tbl-pronoun
#| tbl-cap: 'Summary of pronoun distribution of participants recruited in this study.'

vi_survey %>%
  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
  mutate(pronoun = factor(pronoun, levels = c("He", "She", "Other"))) %>%
  group_by(exp, set) %>%
  summarise(across(everything(), first)) %>%
  count(pronoun) %>%
  mutate(exp = paste("Period", c("I", "II", "III")[exp - 2])) %>%
  pivot_wider(values_from = n, names_from = exp) %>%
  mutate(Total = `Period I` + `Period II` + `Period III`) %>%
  rename(Pronoun = pronoun) %>%
  add_row(Pronoun = "",
          `Period I` = sum(.$`Period I`),
          `Period II` = sum(.$`Period II`),
          `Period III` = sum(.$`Period III`),
          Total = sum(.$Total)) %>%
  mutate(`Period I %` = round(`Period I`/443*100, 1)) %>%
  mutate(`Period II %` = round(`Period II`/443*100, 1)) %>%
  mutate(`Period III %` = round(`Period III`/443*100, 1)) %>%
  mutate(`Total %` = round(Total/443*100, 1)) %>%
  select(Pronoun, 
         `Period I`, `Period I %`, 
         `Period II`, `Period II %`,
         `Period III`, `Period III %`,
         Total, `Total %`) %>%
  kableExtra::kable(col.names = c("Pronoun", 
                                  "Period I", "%", 
                                  "Period II", "%",
                                  "Period III", "%",
                                  "Total", "%"), 
                    booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options = "scale_down")
```

```{r}
#| label: tbl-age-group
#| tbl-cap: 'Summary of age distribution of participants recruited in this study.'

vi_survey %>%
  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
  mutate(pronoun = factor(pronoun, levels = c("He", "She", "Other"))) %>%
  group_by(exp, set) %>%
  summarise(across(everything(), first)) %>%
  count(age_group) %>%
  mutate(exp = paste("Period", c("I", "II", "III")[exp - 2])) %>%
  pivot_wider(values_from = n, names_from = exp) %>%
  mutate(Total = `Period I` + `Period II` + `Period III`) %>%
  rename(`Age group` = age_group) %>%
  add_row(`Age group` = "",
          `Period I` = sum(.$`Period I`),
          `Period II` = sum(.$`Period II`),
          `Period III` = sum(.$`Period III`),
          Total = sum(.$Total)) %>%
  mutate(`Period I %` = round(`Period I`/443*100, 1)) %>%
  mutate(`Period II %` = round(`Period II`/443*100, 1)) %>%
  mutate(`Period III %` = round(`Period III`/443*100, 1)) %>%
  mutate(`Total %` = round(Total/443*100, 1)) %>%
  select(`Age group`, 
         `Period I`, `Period I %`, 
         `Period II`, `Period II %`,
         `Period III`, `Period III %`,
         Total, `Total %`) %>%
  kableExtra::kable(col.names = c("Age group", 
                                  "Period I", "%", 
                                  "Period II", "%",
                                  "Period III", "%",
                                  "Total", "%"),
                    booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options = "scale_down")
```

```{r}
#| label: tbl-education
#| tbl-cap: 'Summary of education distribution of participants recruited in this study.'

vi_survey %>%
  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
  mutate(pronoun = factor(pronoun, levels = c("He", "She", "Other"))) %>%
  group_by(exp, set) %>%
  summarise(across(everything(), first)) %>%
  count(education) %>%
  mutate(exp = paste("Period", c("I", "II", "III")[exp - 2])) %>%
  pivot_wider(values_from = n, names_from = exp, values_fill = 0) %>%
  mutate(Total = `Period I` + `Period II` + `Period III`) %>%
  slice(c(2, 1, 3, 4, 5)) %>%
  rename(`Education` = education) %>%
  add_row(Education = "",
          `Period I` = sum(.$`Period I`),
          `Period II` = sum(.$`Period II`),
          `Period III` = sum(.$`Period III`),
          Total = sum(.$Total)) %>%
  mutate(`Period I %` = round(`Period I`/443*100, 1)) %>%
  mutate(`Period II %` = round(`Period II`/443*100, 1)) %>%
  mutate(`Period III %` = round(`Period III`/443*100, 1)) %>%
  mutate(`Total %` = round(Total/443*100, 1)) %>%
  select(Education, 
         `Period I`, `Period I %`, 
         `Period II`, `Period II %`,
         `Period III`, `Period III %`,
         Total, `Total %`) %>%
  kableExtra::kable(col.names = c("Education", 
                                  "Period I", "%", 
                                  "Period II", "%",
                                  "Period III", "%",
                                  "Total", "%"),
                    booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options = "scale_down")
```

```{r}
#| label: tbl-experience
#| tbl-cap: 'Summary of previous experience distribution of participants recruited in this study.'

vi_survey %>%
  mutate(pronoun = ifelse(pronoun == "They", "Other", pronoun)) %>%
  mutate(pronoun = factor(pronoun, levels = c("He", "She", "Other"))) %>%
  group_by(exp, set) %>%
  summarise(across(everything(), first)) %>%
  count(previous_experience) %>%
  mutate(exp = paste("Period", c("I", "II", "III")[exp - 2])) %>%
  pivot_wider(values_from = n, names_from = exp) %>%
  mutate(Total = `Period I` + `Period II` + `Period III`) %>%
  rename(`Previous experience` = previous_experience) %>%
  add_row(`Previous experience` = "",
          `Period I` = sum(.$`Period I`),
          `Period II` = sum(.$`Period II`),
          `Period III` = sum(.$`Period III`),
          Total = sum(.$Total)) %>%
  mutate(`Period I %` = round(`Period I`/443*100, 1)) %>%
  mutate(`Period II %` = round(`Period II`/443*100, 1)) %>%
  mutate(`Period III %` = round(`Period III`/443*100, 1)) %>%
  mutate(`Total %` = round(Total/443*100, 1)) %>%
  select(`Previous experience`, 
         `Period I`, `Period I %`, 
         `Period II`, `Period II %`,
         `Period III`, `Period III %`,
         Total, `Total %`) %>%
  kableExtra::kable(col.names = c("Previous experience", 
                                  "Period I", "%", 
                                  "Period II", "%",
                                  "Period III", "%",
                                  "Total", "%"),
                    booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options = "scale_down")
```

### Data Collection Periods

```{r fig-poly-boxplot-lineup, fig.cap = 'A lineup of "letter-value" boxplots of weighted propotion of detect for lineups over different data collection periods for non-linearity model. Can you find the most different boxplot? The data plot is positioned in panel $2^3 - 1$.'}
library(lvplot)
sample_from_null <- function(dat, true_position) {
  dat %>%
    mutate(k = true_position) %>%
    bind_rows(map_df((1:20)[-true_position], function(k) {
    dat %>%
      mutate(exp = sample(exp),
             k = k)
  }))
}

vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform") %>%
  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
  mutate(exp = exp - 2) %>%
  group_by(exp, unique_lineup_id) %>%
  mutate(prop_detect = mean(weighted_detect)) %>%
  group_by(exp, unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  ungroup() %>%
  filter(type == "non-linearity") %>%
  select(type, exp, prop_detect) %>%
  mutate(exp = as.character(exp)) %>%
  sample_from_null(7) %>%
  ggplot() +
  geom_lv(aes(x = exp, y = prop_detect, fill = after_stat(LV)), varwidth = TRUE, k = 4) +
  facet_wrap(~k, ncol = 5) +
  theme_light() +
  scale_x_discrete(labels = c("I", "III")) +
  labs(fill = "Quantiles") +
  scale_fill_manual(labels = c("0.5", "[0.25, 0.75]", "[0.125, 0.875]", "[0.0625, 0.09375]"),
                    values = c("white", rev(RColorBrewer::brewer.pal(5, "Blues")))) +
  ylab("Weighted propotion of detect") +
  xlab("Data collection period") +
  theme(legend.key = element_rect(colour = "black"))
```

```{r fig-heter-boxplot-lineup, fig.cap = 'A lineup of "letter-value" boxplots of weighted propotion of detect for lineups over different data collection periods for heteroskedasticity model. Can you find the most different boxplot? The data plot is positioned in panel $2^4 - 2$.'}
vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform") %>%
  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
  mutate(exp = exp - 2) %>%
  group_by(exp, unique_lineup_id) %>%
  mutate(prop_detect = mean(weighted_detect)) %>%
  group_by(exp, unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  ungroup() %>%
  filter(type != "non-linearity") %>%
  select(type, exp, prop_detect) %>%
  mutate(exp = as.character(exp)) %>%
  sample_from_null(14) %>%
  ggplot() +
  geom_lv(aes(x = exp, y = prop_detect, fill = after_stat(LV)), varwidth = TRUE, k = 4) +
  facet_wrap(~k, ncol = 5) +
  theme_light() +
  scale_x_discrete(labels = c("II", "III")) +
  labs(fill = "Quantiles") +
  scale_fill_manual(labels = c("0.5", "[0.25, 0.75]", "[0.125, 0.875]", "[0.0625, 0.09375]"),
                    values = c("white", rev(RColorBrewer::brewer.pal(5, "Blues")))) +
  ylab("Weighted propotion of detect") +
  xlab("Data collection period") +
  theme(legend.key = element_rect(colour = "black"))
```

We have the same type of model collected over different data collection periods, that may lead to unexpected batch effect.
@fig-poly-boxplot-lineup and @fig-heter-boxplot-lineup provide two lineups to examine whether there is an actual difference across data collection periods for non-linearity model and heteroskedasticity model respectively. To emphasize the tail behaviour and display fewer outliers, we use the "letter-value" boxplot [@hofmann2017value] which is an extension of the number of "letter value" statistics to check the weighed proportion of detect over different data collection period. The weighted proportion of detect is calculated by taking the average of $c_i$ of a lineup over a data collection period. Within our research team, we can not identify the data plot from the null plots for these two lineups, result in $p$-values much greater than $5$%. Thus, there is no clear evidence of batch effect.

### 2024-06-27 Thesis/03-chap3.qmd
Update thesis

# Automated assessment of residual plots with computer vision models

Plotting residuals is a standard practice in linear regression diagnostics, essential for identifying deviations from model assumptions such as linearity, homoscedasticity, and normality. Visual inference provides an inferential framework to assess whether residual plots contain patterns inconsistent with model assumptions, typically using a lineup protocol. However, the lineup protocol's reliance on human judgment limits its scalability. This study addresses this limitation by automating the interpretation of residual plots using computer vision models. We develop a distance measure based on Kullback-Leibler divergence to quantify the disparity between the residual distribution of a fitted classical normal linear regression model and the reference distribution. We propose a computer vision model to estimate this distance from residual plots, facilitating formal statistical testing and bootstrapping techniques to assess model specification. Our computer vision model shows strong performance, though it performs slightly less effectively on non-linearity visual patterns. The statistical tests based on the estimated distance exhibit lower sensitivity than conventional tests but higher sensitivity than human visual tests. Examples demonstrate the method's effectiveness across different scenarios, highlighting its value in automating the diagnostic process and supplementing traditional methods.

```{r}
# Visual inference models and p-value calculation
# remotes::install_github("TengMCing/visage")

library(tidyverse)
library(visage)
library(glue)

# To control the simulation in this file
set.seed(10086)
```

## Introduction

Plotting residuals is commonly regarded as a standard practice in linear regression diagnostics [see @belsley1980regression; @cook1982residuals]. This visual assessment plays a crucial role in identifying deviations from model assumptions, such as linearity, homoscedasticity, and normality. It also helps in understanding the goodness of fit and various characteristics of the model.

Generating a residual plot in most statistical software is often as straightforward as executing a line of code or clicking a button. However, accurately interpreting a residual plot can be challenging. A residual plot can exhibit various visual features, but it is crucial to recognize that some may arise from the characteristics of predictors and the inherent randomness of the error, rather than indicating a violation of model assumptions [@li2023plot]. Consider @fig-false-finding as an example, the residuals display a triangular shape pointing to the left. While this might suggest heteroskedasticity, it is important to avoid over-interpreting this visual pattern. In this case, the fitted regression model is correctly specified, and the triangular shape is actually a result of the skewed distribution of the predictors, rather than indicating a flaw in the model.

The concept of visual inference, as proposed by @buja2009statistical, provides an inferential framework to assess whether residual plots indeed contain visual patterns inconsistent with the model assumptions. The fundamental idea involves testing whether the true residual plot visually differs significantly from null plots, where null plots are plotted with residuals generated from the residual rotation distribution [@langsrud2005rotation], which is a distribution consistent with the null hypothesis $H_0$ that the linear regression model is correctly specified. Typically, the visual test is accomplished through the lineup protocol, where the real residual plot is embedded within a lineup alongside several null plots. If the real residual plot can be distinguished from the lineup, it provides evidence for rejecting $H_0$.

The practice of delivering a residual plot as a lineup is generally regarded as a valuable approach. Beyond its application in residual diagnostics, the lineup protocol has integrated into the analysis of diverse subjects. For instance, Loy and Hofmann [-@loy2013diagnostic; -@loy2014hlmdiag; -@loy2015you] illustrated its applicability in diagnosing hierarchical linear models. Additionally, @widen2016graphical demonstrated its utility in geographical research, while @krishnan2021hierarchical explored its effectiveness in forensic examinations.

However, as pointed out by @li2023plot, a primary limitation of the lineup protocol lies in its reliance on human judgments. Unlike conventional statistical tests that can be performed computationally in statistical software, the lineup protocol requires human evaluation of images. This characteristic makes it less suitable for large-scale applications, given the associated high labour costs and time requirements.

There is a substantial need to develop an approach that alleviates analysts' workload by automating repetitive tasks and providing standardized results in a controlled environment. The large-scale evaluation of lineups is impractical without the use of technology and machines.

The utilization of computers to interpret data plots has a rich history, with early efforts such as "Scagnostics" by @tukey1985computer, focusing on scatter plot diagnostics. @wilkinson2005graph expanded on this work, introducing graph theoretic scagnostics, which encompassed computable measures applied to planar proximity graphs. These measures, including, but not limited to, "Outlying", "Skinny", "Stringy", "Straight", "Monotonic", "Skewed", "Clumpy", and "Striated" aimed to characterize outliers, shape, density, trend, coherence and other characteristics of the data. While this approach has been inspiring, there is a recognition [@buja2009statistical] that it may not capture all the necessary visual features that differentiate true residual plots from null plots. A more promising alternative entails enabling machines to learn the function for extracting visual features from residual plots. Essentially, this means empowering computers to discern the crucial visual features for residual diagnostics and determining the method to extract them.

Modern computer vision models are well-suited for addressing this challenge. They rely on deep neural networks with convolutional layers [@fukushima1982neocognitron]. These layers leverage hierarchical patterns in data, downsizing and transforming images by summarizing information in a small space. Numerous studies have demonstrated the efficacy of convolutional layers in addressing various vision tasks, including image recognition [@rawat2017deep]. Despite the widespread use of computer vision models in fields like computer-aided diagnosis [@lee2015image], pedestrian detection [@brunetti2018computer], and facial recognition [@emami2012facial], their application in reading data plots remains limited. While some studies have explored the use of computer vision models for tasks such as reading recurrence plots for time series regression [@ojeda2020multivariate], time series classification [@chu2019automatic; @hailesilassie2019financial; @hatami2018classification; @zhang2020encoding], anomaly detection [@chen2020convolutional], and pairwise causality analysis [@singh2017deep], the application of reading residual plots with computer vision models represents a relatively new field of study.

In this paper, we develop computer vision models and integrate them into the residual plots diagnostics workflow, addressing the need for automated visual inference. The paper is structured as follows. @sec-model-specifications discusses various specifications of the computer vision models. @sec-distance-between-residual-plots defines the distance measure used to measure model violations. @sec-distance-estimation explains how the computer vision models estimate this distance measure. @sec-statistical-testing covers the statistical testing based on the estimated distance. Sections [-@sec-data-generation], [-@sec-model-architecture], and [-@sec-model-training] detail the data preparation, model architecture, and training process, respectively. The results are presented in @sec-model-results. Finally, we conclude with a discussion of our findings and propose ideas for future research directions.

```{r fig-false-finding, fig.width = 8, fig.height = 8*4/5, fig.pos="!h", fig.cap="An example residual vs fitted values plot (red line indicates 0). The vertical spread of the data points varies with the fitted values. This often indicates the existence of heteroskedasticity. The Breusch-Pagan test rejects this residual plot at 95\\% significance level ($p\\text{-value} = 0.046$)."}
set.seed(452)
ori_x <- rand_lognormal()
mod <- heter_model(b = 0, x = closed_form(~-ori_x))
ori_dat <- mod$gen(300)

ori_dat %>%
  VI_MODEL$plot(theme = theme_light(), remove_grid_line = TRUE) +
  xlab("Fitted values") +
  ylab("Residuals")
```

### 2024-07-01 Thesis/02-chap2.qmd
update thesis folder

@@
 <tr>
-<th style="border-bottom:hidden;padding-bottom:0; padding-left:3px;padding-right:3px;text-align: center; " colspan="1"><div style="border-bottom: 1px solid #ddd; padding-bottom: 5px; ">Poly Order (j)</div></th>
-<th style="border-bottom:hidden;padding-bottom:0; padding-left:3px;padding-right:3px;text-align: center; " colspan="1"><div style="border-bottom: 1px solid #ddd; padding-bottom: 5px; ">SD (σ)</div></th>
-<th style="border-bottom:hidden;padding-bottom:0; padding-left:3px;padding-right:3px;text-align: center; " colspan="1"><div style="border-bottom: 1px solid #ddd; padding-bottom: 5px; ">Shape (a)</div></th>
-<th style="border-bottom:hidden;padding-bottom:0; padding-left:3px;padding-right:3px;text-align: center; " colspan="1"><div style="border-bottom: 1px solid #ddd; padding-bottom: 5px; ">Ratio (b)</div></th>
-<th style="border-bottom:hidden;padding-bottom:0; padding-left:3px;padding-right:3px;text-align: center; " colspan="1"><div style="border-bottom: 1px solid #ddd; padding-bottom: 5px; ">Size (n)</div></th>
-<th style="border-bottom:hidden;padding-bottom:0; padding-left:3px;padding-right:3px;text-align: center; " colspan="1"><div style="border-bottom: 1px solid #ddd; padding-bottom: 5px; ">Distribution of the fitted values</div></th>
+<th style="border-bottom:hidden;padding-bottom:0; padding-left:3px;padding-right:3px;text-align: center; " colspan="1"><div style="border-bottom: 1px solid #ddd; padding-bottom: 5px; " data-qmd="Poly Order ($j$)"></div></th>
+<th style="border-bottom:hidden;padding-bottom:0; padding-left:3px;padding-right:3px;text-align: center; " colspan="1"><div style="border-bottom: 1px solid #ddd; padding-bottom: 5px; " data-qmd="SD ($\sigma$)"></div></th>
+<th style="border-bottom:hidden;padding-bottom:0; padding-left:3px;padding-right:3px;text-align: center; " colspan="1"><div style="border-bottom: 1px solid #ddd; padding-bottom: 5px; " data-qmd="Shape ($a$)"></div></th>
+<th style="border-bottom:hidden;padding-bottom:0; padding-left:3px;padding-right:3px;text-align: center; " colspan="1"><div style="border-bottom: 1px solid #ddd; padding-bottom: 5px; " data-qmd="Ratio ($b$)"></div></th>
+<th style="border-bottom:hidden;padding-bottom:0; padding-left:3px;padding-right:3px;text-align: center; " colspan="1"><div style="border-bottom: 1px solid #ddd; padding-bottom: 5px; " data-qmd="Size ($n$)"></div></th>
+<th style="border-bottom:hidden;padding-bottom:0; padding-left:3px;padding-right:3px;text-align: center; " colspan="1"><div style="border-bottom: 1px solid #ddd; padding-bottom: 5px; " data-qmd="Distribution of the fitted values"></div></th>
 </tr>

### 2024-07-01 Thesis/03-chap3.qmd
update thesis folder

@@
-# Automated assessment of residual plots with computer vision models
+# Automated Assessment of Residual Plots with Computer Vision Models
 
@@
 
-## Introduction
+## Introduction {#sec-model-introduction}
 
@@
 @sec-model-specifications discusses various specifications of the computer vision models.
-@sec-distance-between-residual-plots defines the distance measure used to measure model violations.
-@sec-distance-estimation explains how the computer vision models estimate this distance measure.
-@sec-statistical-testing covers the statistical testing based on the estimated distance.
-Sections [-@sec-data-generation], [-@sec-model-architecture], and [-@sec-model-training] detail the data preparation, model architecture, and training process, respectively.
+@sec-model-distance-between-residual-plots defines the distance measure used to measure model violations.
+@sec-model-distance-estimation explains how the computer vision models estimate this distance measure.
+@sec-model-statistical-testing covers the statistical testing based on the estimated distance.
+Sections [-@sec-model-data-generation], [-@sec-model-architecture], and [-@sec-model-training] detail the data preparation, model architecture, and training process, respectively.
 The results are presented in @sec-model-results.
@@
 ```
+
+## Model Specifications {#sec-model-specifications}
+
+There are various specifications of the computer vision model that can be used to assess residual plots.
+We discuss these specifications below focusing on two key components of the model formula: the input and the output format.
+
+### Input Formats
+
+Deep learning models are in general very sensitive to the input data.
+The quality and relevance of the input data greatly influence the model's capacity to generate insightful and meaningful results.
+There are several designs of the input format can be considered.
+
+A straightforward design involves feeding a vector of residuals along with a vector of fitted values, essentially providing all the necessary information for creating a residuals vs fitted values plot.
+However, a drawback of this method is the dynamic input size, which changes based on the number of observations.
+For modern computer vision models implemented in mainstream software like TensorFlow [@abadi2016tensorflow], the input shape is typically fixed.
+One solution is to pad the input vectors with leading or trailing zeros when the input tensor expects longer vectors, but it may fail if the input vector surpasses the designed length.
+Another strategy is to summarize the residuals and fitted values separately using histograms and utilize the counts as the input.
+By controlling the number of bins in the histograms, it becomes possible to provide fixed-length input vectors.
+Still, since histograms only capture the marginal distribution of residuals and fitted values respectively, they can not be used to differentiate visual patterns with same marginal distributions but different joint distributions.
+
+Another design involves using an image as input.
+The primary advantage of this design, as opposed to the vector format, is the availability of the existing and sophisticated image processing architectures developed over the years, such as the VGG16 architecture proposed in @simonyan2014very.
+These architectures can effectively capture and summarize spatial information from nearby pixels, which is less straightforward with vector input.
+The main considerations are the image resolution and the aesthetics of the residual plot.
+In general, higher resolution provides more information to the model but comes with the trade-off of increased complexity and greater difficulty in training.
+As for the aesthetics of the residual plot, a practical solution is to consistently present residual plots in the same style to the model.
+This implies that the model can not accept arbitrary images as input but requires the use of the same pre-processing pipeline to convert residuals and fitted values into a standardized-style residual plot.
+
+Providing multiple residual plots to the model, such as a pair of plots, a triplet or a lineup is also a possible option.
+@chopra2005learning have shown that computer vision models designed for image comparison can assess whether a pair of images are similar or dissimilar.
+Applied to our specific problem, we can define null plots of a fitted regression model to be similar to each other, while considering true residual plots to be distinct from null plots of any fitted regression model.
+A triplet constitutes a set of three images, denoted as $image_1$, $image_2$ and $image_3$.
+It is often used to predict whether $image_2$ or $image_3$ is more similar to $image_1$, proving particularly useful for establishing rankings between samples.
+For this setup, we can apply the same criteria to define similarity between images.
+However, it is important to note that these two approaches usually require additional considerations regarding the loss function and, at times, non-standard training processes due to shared weights between different convolutional blocks.
+
+Presenting a lineup to a model aligns closely with the lineup protocol.
+However, as the number of residual plots in a lineup increases, the resolution of the input image grows rapidly, posing challenges in training the model.
+We experimented with this approach in a pilot study, but the performance of the trained model was sub-optimal.
+
+We did not explore all the mentioned input formats due to the significant costs tied to data preparation and model training.
+Taking into account the implementation cost and the need for model interpretability, we settled on the single residual plot input format.
+
+### Output Formats
+
+Given that the input is a single residual plot represented as a fixed-resolution image, the output from the computer vision model can take one of two forms: binary or numeric.
+This choice determines whether the model belongs to a classification model or a regression model.
+The binary outcome encoded as $0$ and $1$ could be used to represent whether the input image is a null plot, or whether the input image would be rejected in a visual test conducted by humans.
+Training a model following the latter option requires data from prior human subject experiments, presenting difficulties in controlling the quality of data due to variations in experimental settings across different studies.
+Additionally, some visual inference experiments are unrelated to linear regression models or residual plot diagnostics, resulting in a limited amount of available training data.
+
+Alternatively, the output could be a meaningful and interpretable numerical measure useful for assessing residual plots, such as the strength of suspicious visual patterns reflecting the extent of model violations, or the difficulty index for identifying whether a residual plot has no issues.
+However, these numeric measures are often informally used in daily communication but are not typically formalized or rigorously defined.
+For the purpose of training a model, this numeric measure has to be quantifiable.
+
+In this study, we chose to define and use a distance measure to quantify the difference between the residual plot and the null plots.
+@vo2016localizing have also demonstrated that defining a proper distance between images can enhance the matching accuracy in image search compared to a binary outcome model.
+
+## Distance Between Residual Plots {#sec-model-distance-between-residual-plots}
+
+In a visual test, the observer will be asked to choose one or more plots that stand out as most distinct from others in a given lineup.
+To develop a computer vision model for assessing residual plots within the visual inference framework, it is important to precisely define a numerical measure of "difference" or "distance" between plots.
+This distance can take the form of a basic statistical operation on pixels, such as the sum of square differences.
+Alternatively, it could involve established image similarity metrics like the Structural Similarity Index Measure [@wang2004image].
+The challenge lies in the fact that metrics tailored for image comparison may not be suitable for evaluating data plots, where only essential plot elements require assessment [@chowdhury2018measuring].
+Furthermore, scagnostics mentioned in @sec-model-introduction could be used to construct distance measures for residual plots comparison, but the functional form still needs to be carefully refined to accurately reflect the extent of the violations.
+
+### Residual Distribution
+
+The distance measure proposed in this study takes into account the fact that we tried to measure how different a residual plot is from a good residual plot, or in other words, how different a given fitted regression model is from a correctly specified model.
+
+For a classical normal linear regression model, residuals $\boldsymbol{e}$ are derived from the fitted values $\hat{\boldsymbol{y}}$ and observed values $\boldsymbol{y}$.
+Suppose the data generating process is known and the regression model is correctly specified, by the Frisch-Waugh-Lowell theorem [@frisch1933partial], residuals $\boldsymbol{e}$ can also be treated as random variables and written as a linear transformation of the error $\boldsymbol{\varepsilon}$ formulated as $\boldsymbol{e} = \boldsymbol{R}\boldsymbol{\varepsilon}$, where $\boldsymbol{R}=\boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ is the residual operator and an idempotent matrix, $\boldsymbol{X}$ is the design matrix, $\boldsymbol{I}_n$ is a $n$ by $n$ identity matrix, and $n$ is the number of observations.
+
+One of the assumptions of the classical normal linear regression model is the error $\boldsymbol{\varepsilon}$ follows a multivariate normal distribution with zero mean and constant variance, i.e., $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)$.
+It can be known that residuals $\boldsymbol{e}$ also follow a certain probability distribution transformed from the multivariate normal distribution, which will be denoted as $Q$.
+This reference distribution $Q$ summarizes what good residuals should follow given the design matrix $\boldsymbol{X}$ is known and fixed.
+
+Suppose the design matrix $\boldsymbol{X}$ has linearly independent columns, the trace of the hat matrix will equal to the number of columns in $\boldsymbol{X}$ denoted as $k$.
+As a result, the rank of $\boldsymbol{R}$ is $n - k$, and $Q$ is a degenerate multivariate distribution.
+To capture the characteristics of $Q$, such as moments, we can simulate a large numbers of $\boldsymbol{\varepsilon}$ and transform it to $\boldsymbol{e}$ to get the empirical estimates.
+For simplicity, in this study, we replaced the variance-covariance matrix of residuals $\text{cov}(\boldsymbol{e}, \boldsymbol{e}) = \boldsymbol{R}\sigma^2\boldsymbol{R}' = \boldsymbol{R}\sigma^2$ with a full-rank diagonal matrix $\text{diag}(\boldsymbol{R}\sigma^2)$, where $\text{diag}(.)$ sets the non-diagonal entries of a matrix to zeros.
+The resulting distribution for $\boldsymbol{Q}$ is $N(\boldsymbol{0}_n, \text{diag}(\boldsymbol{R}\sigma^2))$.
+
+Distribution $Q$ is derived from the correctly specified model.
+However, if the model is misspecified, then the actual distribution of residuals denoted as $P$, will be different from $Q$.
+For example, if the data generating process contains variables correlated with any column of $\boldsymbol{X}$ but missing from $\boldsymbol{X}$, causing an omitted variable problem, $P$ will be different from $Q$ because the residual operator obtained from the fitted regression model will not be the same as $\boldsymbol{R}$.
+Besides, if the $\boldsymbol{\varepsilon}$ follows a non-normal distribution such as a multivariate lognormal distribution, $P$ will usually be skewed and has a long tail.
+
+### Distance of $P$ from $Q$
+
+Define a proper distance between distributions is usually easier than define a proper distance between data plots.
+Given the true residual distribution $Q$ and the reference residual distribution $P$, we used a distance measure based on Kullback-Leibler divergence [@kullback1951information] to quantify the difference between two distributions
+
+$$
+D = \log\left(1 + D_{KL}\right),
+$$ {#eq-kl-0}
+
+where $D_{KL}$ is defined as
+
+$$
+D_{KL} = \int_{\mathbb{R}^{n}}\log\frac{p(\boldsymbol{e})}{q(\boldsymbol{e})}p(\boldsymbol{e})d\boldsymbol{e},
+$$ {#eq-kl-1}
+
+\noindent and $p(.)$ and $q(.)$ are the probability density functions for distribution $P$ and distribution $Q$ respectively.
+
+This distance measure was first proposed in @li2023plot.
+It was mainly designed for measuring the effect size of non-linearity and heteroskedasticity in a residual plot.
+@li2023plot have derived that, for a classical normal linear regression model that omits a necessary higher-order predictors $\boldsymbol{Z}$, and incorrectly assumes $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)$ while in fact $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \boldsymbol{V})$ with $\boldsymbol{V}$ being an arbitrary symmetric positive semi-definite matrix, $Q$ can be represented as $N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, \text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}))$.
+Note that the variance-covariance matrix is replaced with the diagonal matrix to ensure it is a full-rank matrix.
+
+Since both $P$ and $Q$ are adjusted to be multivariate normal distributions, @eq-kl-1 can be further expanded to
+
+$$
+\begin{aligned}
+D_{KL} &= \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{W})|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + tr(\text{diag}(\boldsymbol{W})^{-1}\text{diag}(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z'(\text{diag}(\boldsymbol{W}))^{-1}\boldsymbol{\mu}_z\right),
+\end{aligned}
+$$ {#eq-kl-2}
+
+\noindent where $\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z$, and $\boldsymbol{W} = \boldsymbol{R}\boldsymbol{V}\boldsymbol{R}$.
+The assumed error variance $\sigma^2$ is set to be $tr(\boldsymbol{V})/n$, which is the expectation of the estimated variance.
+
+### Non-normal $P$
+
+For non-normal error $\boldsymbol{\varepsilon}$, the true residual distribution $P$ is unlikely to be a multivariate normal distribution.
+Thus, @eq-kl-2 given in @li2023plot will not be applicable to models violating the normality assumption.
+
+To evaluate the Kullback-Leibler divergence of non-normal $P$ from $Q$, the fallback is to solve @eq-kl-1 numerically.
+However, since $\boldsymbol{e}$ is a linear transformation of non-normal random variables, it is very common that the general form of $P$ is unknown, meaning that we can not easily compute $p(\boldsymbol{e})$ using a well-known probability density function.
+Additionally, even if $p(\boldsymbol{e})$ can be calculated for any $\boldsymbol{e} \in \mathbb{R}^n$, it will be very difficult to do numerical integration over the $n$ dimensional space, because $n$ could be potentially very large.
+
+In order to approximate $D_{KL}$ in a practically computable manner, the elements of $\boldsymbol{e}$ are assumed to be independent of each other.
+This assumption solves both of the issues mentioned above.
+First, we no longer need to integrate over $n$ random variables.
+The result of @eq-kl-1 is now the sum of the Kullback-Leibler divergence evaluated for each individual residual thanks for the independence assumption.
+Second, it is not required to know the joint probability density $p(\boldsymbol{e})$ any more.
+Instead, the evaluation of Kullback-Leibler divergence for an individual residual relies on the knowledge of the marginal density $p_i(e_i)$, where $e_i$ is the $i$-th residual for $i = 1, ..., n$.
+This is much easier to approximate through simulation.
+It is also worth mentioning that this independence assumption generally will not hold, since $\text{cov}(e_i, e_j) \neq 0$ if $\boldsymbol{R}_{ij} \neq 0$ for any $1 \leq i < j \leq n$, but its existence is essential for reducing the computational cost.
+
+Given $\boldsymbol{X}$ and $\boldsymbol{\beta}$, the algorithm for approximating @eq-kl-1 starts from simulating $m$ sets of observed values $\boldsymbol{y}$ according to the data generating process.
+The observed values are stored in a matrix $\boldsymbol{A}$ with $n$ rows and $m$ columns, where each column of $\boldsymbol{A}$ is a set of observed values.
+Then, we can get $m$ sets of realized values of $\boldsymbol{e}$ stored in the matrix $\boldsymbol{B}$ by applying the residual operator $\boldsymbol{B} = \boldsymbol{R}\boldsymbol{A}$.
+Furthermore, kernel density estimation (KDE) with Gaussian kernel and optimal bandwidth selected by the Silverman's rule of thumb [@silverman2018density] is applied on each row of $B$ to estimate $p_i(e_i)$ for $i = 1, ..., n$.
+The KDE computation can be done by the `density` function in R.
+
+Since the Kullback-Leibler divergence can be viewed as the expectation of the log-likelihood ratio between distribution $P$ and distribution $Q$ evaluated on distribution $P$, we can reuse the simulated residuals in matrix $\boldsymbol{B}$ to estimate the expectation by the sample mean.
+With the independence assumption, for non-normal $P$, $D_{KL}$ can be approximated by
+
+$$
+\begin{aligned}
+D_{KL} &\approx \sum_{i = 1}^{n} \hat{D}_{KL}^{(i)}, \\
+\hat{D}_{KL}^{(i)} &= \frac{1}{m}\sum_{j = 1}^{m} log\frac{\hat{p_i}(B_{ij})}{q(B_{ij})},
+\end{aligned} 
+$$ {#eq-kl-3}
+
+\noindent where $\hat{D}_{KL}^{(i)}$ is the estimator of the Kullback-Leibler divergence for an individual residual $e_i$, $B_{ij}$ is the $i$-th row and $j$-th column entry of the matrix $B$, $\hat{p_i}(.)$ is the kernel density estimator of $p_i(.)$, $q(.)$ is the normal density function with mean zero and an assumed variance estimated as $\widehat{\sigma^2} = \sum_{b \in vec(B)}(b - \sum_{b \in vec(B)} b/nm)^2/(nm - 1)$, and $vec(.)$ is the vectorization operator which turns a $n \times m$ matrix into a $nm \times 1$ column vector by stacking the columns of the matrix on top of each other.
+
+## Distance Estimation {#sec-model-distance-estimation}
+
+In the previous sections, we have defined a distance measure given in @eq-kl-0 for quantifying the difference between the true residual distribution $P$ and an ideal reference distribution $Q$.
+It can be noticed that this distance measure can only be computed when the data generating process is known.
+In reality, we often have no knowledge about the data generating process, otherwise we do not need to fit a linear regression model in the first place.
+
+We tried to train a computer vision model to estimate this distance measure with a residual plot. Let $D$ be the result of @eq-kl-0, and our estimator $\hat{D}$ is formulated as
+
+$$
+\hat{D} = f_{CV}(V_{h \times w}(\boldsymbol{e}, \hat{\boldsymbol{y}})),
+$$ {#eq-d-approx}
+
+\noindent where $V_{h \times w}(.)$ is a plotting function that saves a residuals vs fitted values plot with fixed aesthetic as an RGB image with $h \times w$ pixels, $f_{CV}(.)$ is a computer vision model which takes an $h \times w$ image as input and predicts the distance in the domain $[0, +\infty)$.
+
+With the estimated distance $\hat{D}$, we will be able to know how different the underlying distribution of the residuals is from a good residual distribution.
+$\hat{D}$ can also be used as an index of the model violations indicating the strength of the visual signal embedded in the residual plot.
+
+It is not expected that $\hat{D}$ will be equal to original distance $D$.
+This is largely because information contained in a single residual plot is limited and it may not be able to summarise all the important characteristics of the residual distribution.
+For a given residual distribution $P$, many different residual plots can be simulated, where many will share similar visual patterns, but some of them could be visually very different from the rest, especially for regression models with small $n$.
+This suggests the error of the estimation will vary depends on whether the input residual plot is representative or not.
+
+## Model Violations Index (MVI) {#sec:model-violations-index}
+
+$\hat{D}$ is an estimator of the difference between the true residual distribution and the reference residual distribution.
+This difference primarily arises from deviations in model assumptions.
+The magnitude of $D$ directly reflects the degree of these deviations, thus making $\hat{D}$ instrumental in forming a model violations index.
+
+Note that @eq-kl-0 might be influenced by the number of random variables involved in its evaluation.
+Generally, a larger number of observations will lead to a greater distance $D$.
+However, this does not imply that $\hat{D}$ fails to accurately represent the extent of model violations.
+In fact, when examining residual plots with more observations, we often observe a stronger visual signal strength, as the underlying patterns are more likely to be revealed, except in cases of significant overlapping.
+
+For a consistent data generating process, $D$ typically increases logarithmically with the number of observations.
+This behaviour comes from the relationship $D = \text{log}(1 + D_{KL})$, where $D_{KL} = \sum_{i=1}^{n}D_{KL}^{(i)}$ under the assumption of independence.
+
+Therefore, the Model Violations Index (MVI) can be proposed as
+
+$$
+\text{MVI} = C + \hat{D} - log(n),
+$$ {#eq-mvi}
+
+\noindent where $C$ is a large enough constant keeping the result positive.
+
+@fig-poly-index and @fig-heter-index display the residual plots for fitted models exhibiting varying degrees of non-linearity and heteroskedasticity.
+Each residual plot's MVI is computed using @eq-mvi with $C = 10$.
+When $\text{MVI} > 8$, the visual patterns are notably strong and easily discernible by humans.
+In the range $6 < \text{MVI} < 8$, the visibility of the visual pattern diminishes as MVI decreases.
+Conversely, when $\text{MVI} < 6$, the visual pattern tends to become relatively faint and challenging to observe.
+@tbl-mvi provides a summary of the MVI usage and it is applicable to other linear regression models.
+
+::: {.content-visible when-format="html"}
+
+::: {#tbl-mvi .cell tbl-cap='Degree of model violations or the strength of the visual signals according to the Model Violation Index (MVI). The constant $C$ is set to be 10.'}
+::: {.cell-output-display}
+
+```{=html}
+<table>
+ <thead>
+  <tr>
+   <th style="text-align:left;"> <span data-qmd="Degree of Model Violations"></span> </th>
+   <th style="text-align:left;"> <span data-qmd="Range ($C$ = 10)"></span>  </th>
+  </tr>
+ </thead>
+<tbody>
+  <tr>
+   <td style="text-align:left;"> <span data-qmd="Strong"></span>  </td>
+   <td style="text-align:left;"> <span data-qmd="$\text{MVI} > 8$"></span>  </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> <span data-qmd="Moderate"></span>  </td>
+   <td style="text-align:left;"> <span data-qmd="$6 < \text{MVI} < 8$"></span> </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> <span data-qmd="Weak"></span>  </td>
+   <td style="text-align:left;"> <span data-qmd="$\text{MVI} < 6$"></span>  </td>
+  </tr>
+</tbody>
+</table>
+
+```
+
+:::
+:::
+
+:::
+
+::: {.content-visible when-format="pdf"}
+
+```{r}
+#| tbl-cap: "Degree of model violations or the strength of the visual signals according to the Model Violation Index (MVI). The constant $C$ is set to be 10."
+#| label: tbl-mvi
+data.frame(degree = c("Strong", "Moderate", "Weak"),
+           range = c("$\\text{MVI} > 8$", "$6 < \\text{MVI} < 8$", "$\\text{MVI} < 6$")) %>%
+  kableExtra::kable(booktabs = TRUE,
+                    escape = FALSE,
+                    col.names = c("Degree of model violations", "Range ($C$ = 10)"))
+```
+
+:::
+
+```{r fig-poly-index, fig.pos = "!h", fig.cap = "Residual plots generated from fitted models exhibiting varying degrees of non-linearity violations.
+The Model violations index (MVI) is displayed atop each residual plot.
+The non-linearity patterns are relatively strong for $MVI > 8$, and relatively weak for $MVI < 6$."}
+set.seed(1495)
+if (!file.exists(here::here("cached_data/poly_index.rds"))) {
+  e_sigma_factor <- 3
+  keras_mod <- autovi::get_keras_model("vss_phn_32")
+  poly_dgp <- phn_model(j = 3, sigma = 0.1^e_sigma_factor)
+  ori_dat <- poly_dgp$gen(300)
+
+  vss <- c()
+  actual_ss <- c()
+  dat_combined <- tibble()
+
+  for (e_sigma in seq(0.7, 1.8, length.out = 20)^e_sigma_factor) {
+    poly_dgp <- phn_model(j = 3, sigma = e_sigma)
+    dat <- poly_dgp$gen(300, computed = list(x1 = ori_dat$x1, e = ori_dat$e, k = ori_dat$k * e_sigma / 0.1^e_sigma_factor))
+    mod <- lm(y ~ x1, data = dat)
+    my_vi <- autovi::auto_vi(fitted_mod = mod, keras_mod = keras_mod)
+    vss <- c(vss, my_vi$vss()$vss)
+    actual_ss <- c(actual_ss, log(poly_dgp$sample_effect_size(dat) + 1))
+
+    dat_combined <- bind_rows(dat_combined, mutate(dat, e_sigma = e_sigma))
+  }
+
+  saveRDS(dat_combined, here::here("cached_data/poly_index.rds"))
+  saveRDS(vss, here::here("cached_data/poly_index_vss.rds"))
+}
+
+dat_combined <- readRDS(here::here("cached_data/poly_index.rds"))
+vss <- readRDS(here::here("cached_data/poly_index_vss.rds"))
+
+dat_combined %>%
+  autovi::AUTO_VI$plot_resid() +
+  facet_wrap(~e_sigma, scales = "free", 
+             labeller = function(...){ 
+               tibble(nb = map(glue("'MVI' == {format(vss - log(300) + 10, digits = 3)}"), str2expression))
+               },
+             ncol = 5) +
+  theme(strip.text = element_text(size = 10))
+```
+
+```{r fig-heter-index, fig.pos = "!h", fig.cap = "Residual plots generated from fitted models exhibiting varying degrees of heteroskedasticity violations.
+The Model violations index (MVI) is displayed atop each residual plot.
+The heteroskedasticity patterns are relatively strong for $MVI > 8$, and relatively weak for $MVI < 6$."}
+set.seed(1495)
+if (!file.exists(here::here("cached_data/heter_index.rds"))) {
+  b_factor <- 2
+  keras_mod <- autovi::get_keras_model("vss_phn_32")
+  heter_dgp <- phn_model(include_z = FALSE, b = 0)
+  ori_dat <- poly_dgp$gen(300)
+
+  vss <- c()
+  actual_ss <- c()
+  dat_combined <- tibble()
+
+  for (b in rev(seq(0.2, 4, length.out = 20))^b_factor) {
+    heter_dgp <- phn_model(include_z = FALSE, b = b)
+    dat <- heter_dgp$gen(300, computed = list(x1 = ori_dat$x1, e = ori_dat$e))
+    mod <- lm(y ~ x1, data = dat)
+    my_vi <- autovi::auto_vi(fitted_mod = mod, keras_mod = keras_mod)
+    vss <- c(vss, my_vi$vss()$vss)
+    actual_ss <- c(actual_ss, log(heter_dgp$sample_effect_size(dat) + 1))
+
+    dat_combined <- bind_rows(dat_combined, mutate(dat, nb = -b))
+  }
+
+  saveRDS(dat_combined, here::here("cached_data/heter_index.rds"))
+  saveRDS(vss, here::here("cached_data/heter_index_vss.rds"))
+}
+
+dat_combined <- readRDS(here::here("cached_data/heter_index.rds"))
+vss <- readRDS(here::here("cached_data/heter_index_vss.rds"))
+
+dat_combined %>%
+  autovi::AUTO_VI$plot_resid() +
+  facet_wrap(~nb, scales = "free", 
+             labeller = function(...){ 
+               tibble(nb = map(glue("'MVI' == {format(vss - log(300) + 10, digits = 3)}"), str2expression))
+               },
+             ncol = 5) +
+  theme(strip.text = element_text(size = 10))
+```
+
+## Statistical testing {#sec-model-statistical-testing}
+
+### Lineup Evaluation {#sec-model-lineup-evaluation}
+
+Theoretically, the distance $D$ for a correctly specified model is $0$, because $P$ will be the same as $Q$.
+However, the computer vision model may not necessary predict $0$ for a null plot.
+Using @fig-false-finding as an example, it contains a visual pattern which is an indication of heteroskedasticity.
+We would not expect the model to be able to magically tell if the suspicious pattern is caused by the skewed distribution of the fitted values or the existence of heteroskedasticity.
+Additionally, some null plots could have outliers or strong visual patterns due to randomness, and a reasonable model will try to summarise those information into the prediction, resulting in $\hat{D} > 0$.
+
+This property is not an issue if $\hat{D} \gg 0$ for which the visual signal of the residual plot is very strong, and we usually do not need any further examination of the significance of the result.
+However, if the visual pattern is weak or moderate, having $\hat{D}$ will not be sufficient to determine if $H_0$ should be rejected.
+
+To address this issue while adhering to the principle of visual inference, we can compare the estimated distance $\hat{D}$ to the estimated distances for the null plots in a lineup.
+Specifically, if a lineup comprises 20 plots, the null hypothesis $H_0$ will be rejected if $\hat{D}$ exceeds the maximum estimated distance among the $m - 1$ null plots, denoted as $\max\limits_{1 \leq i \leq m-1} {\hat{D}_{null}^{(i)}}$, where $\hat{D}_{null}^{(i)}$ represents the estimated distance for the $i$-th null plot.
+This approach is equivalent to the typical lineup protocol requiring a 95% significance level, where $H_0$ is rejected if the data plot is identified as the most distinct plot by the sole observer.
+The estimated distance serves as a metric to quantify the difference between the data plot and the null plots, as intended.
+
+For lineups consisting of more than 20 plots, the 95% significance level can be maintained if the number of plots is a multiple of 20.
+Specifically, for lineups comprising $20t$ plots, where $t$ is a positive integer, we reject $H_0$ if $\hat{D}$ exceeds 95% ${\hat{D}_{null}^{(i)}}$ for $i = 1, \ldots, 20t-1$.
+The $p$-value in this case is given by $\frac{1}{20t} + \frac{1}{20t}\sum_{i=1}^{20t-1} I\left(\hat{D}_{null}^{(i)} > \hat{D}\right)$, where $I(\cdot)$ is the indicator function.
+
+Moreover, if the number of plots in a lineup, denoted by $m$, is sufficiently large, the empirical distribution of ${\hat{D}_{null}^{(i)}}$ can be viewed as an approximation of the null distribution of the estimated distance.
+Consequently, quantiles of the null distribution can be estimated using the sample quantiles available in statistical software such as R, and these quantiles can be utilized for decision-making purposes.
+The details of the sample quantile computation can be found in @hyndman1996sample.
+For instance, if $\hat{D}$ is greater than or equal to the 95% sample quantile, denoted as $Q_{null}(0.95)$, we can conclude that the estimated distance for the true residual plot is significantly different from the estimated distance for null plots with a 95% significance level.
+Based on our experience, to obtain a stable estimate of the 95% quantile, the number of null plots, $n_{null}$, typically needs to be at least 100.
+However, if the null distribution exhibits a long tail, a larger number of null plots may be required.
+Alternatively, a $p$-value can be used to represent the probability of observing an event equally or more extreme than the given event under the null hypothesis $H_0$, and it can be estimated by $\frac{1}{m}\sum_{i=1}^{m-1}I\left(\hat{D}_{null}^{(i)} \geq \hat{D}\right)$.
+
+If precision in sample quantiles is not the main priority, using a pre-calculated table of quantiles is an available option.
+Such a table offers pre-determined quantiles for a specified number of observations.
+It is generated by assessing numerous null plots derived from various simulated regression models and averaging them.
+Essentially, this shifts the computational burden from the user to the developer.
+
+### Bootstrapping
+
+Bootstrap is often employed in linear regression when conducting inference for estimated parameters.
+It is typically done by sampling individual cases with replacement and refitting the regression model.
+If the observed data accurately reflects the true distribution of the population, the bootstrapped estimates can be used to measure the variability of the parameter estimate without making strong distributional assumptions about the data generating process.
+
+Similarly, bootstrap can be applied on the estimated distance $\hat{D}$.
+For each refitted model $M_{boot}^{(i)}$, there will be an associated residual plot $V_{boot}^{(i)}$ which can be fed into the computer vision model to obtain $\hat{D}_{boot}^{(i)}$, where $i = 1,...,n_{boot}$, and $n_{boot}$ is the number of bootstrapped samples.
+If we are interested in the variation of $\hat{D}$, we can use $\hat{D}_{boot}^{(i)}$ to estimate a confidence interval.
+
+Alternatively, since each $M_{boot}^{(i)}$ has a set of estimated coefficients $\hat{\boldsymbol{\beta}}_{boot}^{(i)}$ and an estimated variance $\hat{\sigma^2}_{boot}^{(i)}$, a new approximated null distribution can be construed and the corresponding 95% sample quantile $Q_{boot}^{(i)}(0.95)$ can be computed.
+Then, if $\hat{D}_{boot}^{(i)} \geq Q_{boot}^{(i)}(0.95)$, $H_0$ will be rejected for $M_{boot}^{(i)}$.
+The ratio of rejected $M_{boot}^{(i)}$ among all the refitted models provides an indication of how often the assumed regression model are considered to be incorrect if the data can be obtained repetitively from the same data generating process.
+But this approach is computationally very expensive because it requires $n_{boot} \times n_{null}$ times of residual plot assessment.
+In practice, $Q_{null}(0.95)$ can be used to replace $Q_{boot}^{(i)}(0.95)$ in the computation.
+
+## Data Generation {#sec-model-data-generation}
+
+### Simulation Scheme
+
+While observational data is frequently employed in training models for real-world applications, the data generating process of observational data often remains unknown, making computation for our target variable $D$ unattainable.
+Consequently, the computer vision models developed in this study were trained using synthetic data, including 80000 training images and 8000 test images.
+This approach provided us with precise label annotations.
+Additionally, it ensured a large and diverse training dataset, as we had control over the data generating process, and the simulation of the training data was relatively cost-effective.
+
+We have incorporated three types of residual departures of linear regression model in the training data, including non-linearity, heteroskedasticity and non-normality.
+All three departures can be summarised by the data generating process formulated as
+
+$$
+\begin{aligned}
+\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 + \beta_2(\boldsymbol{z} + \beta_1\boldsymbol{w}) + \boldsymbol{k} \odot \boldsymbol{\varepsilon}, \\
+\boldsymbol{z} &= \text{He}_j(g(\boldsymbol{x}_1, 2)), \\
+\boldsymbol{w} &= \text{He}_j(g(\boldsymbol{x}_2, 2)), \\
+\boldsymbol{k} &= \sqrt{\boldsymbol{1}_n + b(2 - |a|)(\boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 - a\boldsymbol{1}_n)^2},
+\end{aligned}
+$$ {#eq-data-sim}
+
+\noindent where $\boldsymbol{y}$, $\boldsymbol{x}_1$, $\boldsymbol{x}_2$, $\boldsymbol{z}$, $\boldsymbol{w}$, $\boldsymbol{k}$ and $\boldsymbol{\varepsilon}$ are vectors of size $n$, $\boldsymbol{1}_n$ is a vector of ones of size $n$, $\boldsymbol{x}_1$ and $\boldsymbol{x}_2$ are two independent predictors, $\text{He}_j(.)$ is the $j$th-order probabilist's Hermite polynomials [@hermite1864nouveau], the $\sqrt{(.)}$ and $(.)^2$ operators are element-wise operators, $\odot$ is the Hadamard product, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
+
+$$g(\boldsymbol{x}, k) = 2k \cdot \frac{\boldsymbol{x} - x_{min}\boldsymbol{1}_n}{x_{max} - x_{min}} - k\boldsymbol{1}_n,~for~k > 0,$$
+\noindent where $x_{min} = \underset{i \in \{ 1,...,n\}}{min} x_i$, $x_{max} = \underset{i \in \{ 1,...,n\}}{max} x_i$ and $x_i$ is the $i$-th entry of $\boldsymbol{x}$.
+
+::: {.content-visible when-format="html"}
+
+::: {#tbl-factor .cell tbl-cap='Factors used in the data generating process for synthetic data simulation.
+Factor $j$ and $a$ controls the non-linearity shape and the heteroskedasticity shape respectively.
+Factor $b$, $\sigma_\varepsilon$ and $n$ control the signal strength.
+Factor $\text{dist}_\varepsilon$, $\text{dist}_{x1}$ and $\text{dist}_{x2}$ specifies the distribution of $\varepsilon$, $X_1$ and $X_2$ respectively.'}
+::: {.cell-output-display}
+
+```{=html}
+<table>
+ <thead>
+  <tr>
+   <th style="text-align:left;"> Factor </th>
+   <th style="text-align:left;"> Domain </th>
+  </tr>
+ </thead>
+<tbody>
+  <tr>
+   <td style="text-align:left;"> j </td>
+   <td style="text-align:left;"> {2, 3, ..., 18} </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> a </td>
+   <td style="text-align:left;"> [-1, 1] </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> b </td>
+   <td style="text-align:left;"> [0, 100] </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> <span data-qmd="$\beta_1$"></span> </td>
+   <td style="text-align:left;"> {0, 1} </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> <span data-qmd="$\beta_2$"></span> </td>
+   <td style="text-align:left;"> {0, 1} </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> <span data-qmd="$\text{dist}_varepsilon$"></span> </td>
+   <td style="text-align:left;"> {discrete, uniform, normal, lognormal} </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> <span data-qmd="$\text{dist}_{x1}$"></span> </td>
+   <td style="text-align:left;"> {discrete, uniform, normal, lognormal} </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> <span data-qmd="$\text{dist}_{x2}$"></span> </td>
+   <td style="text-align:left;"> {discrete, uniform, normal, lognormal} </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> <span data-qmd="$\sigma_{\varepsilon}$"></span> </td>
+   <td style="text-align:left;"> [0.0625, 9] </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> <span data-qmd="$\sigma_{X1}$"></span> </td>
+   <td style="text-align:left;"> [0.3, 0.6] </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> <span data-qmd="$\sigma_{X2}$"></span> </td>
+   <td style="text-align:left;"> [0.3, 0.6] </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> n </td>
+   <td style="text-align:left;"> [50, 500] </td>
+  </tr>
+</tbody>
+</table>
+
+```
+
+:::
+:::
+
+:::
+
+::: {.content-visible when-format="pdf"}
+
+```{r}
+#| label: factor
+#| tbl-cap: "Factors used in the data generating process for synthetic data simulation.
+Factor $j$ and $a$ controls the non-linearity shape and the heteroskedasticity shape respectively.
+Factor $b$, $\\sigma_\\varepsilon$ and $n$ control the signal strength.
+Factor $\\text{dist}_\\varepsilon$, $\\text{dist}_{x1}$ and $\\text{dist}_{x2}$ specifies the distribution of $\\varepsilon$, $X_1$ and $X_2$ respectively."
+tibble(Factor = c("j", "a", "b", "$\\beta_1$", "$\\beta_2$", " $\\text{dist}_\\varepsilon$", "$\\text{dist}_{x1}$", "$\\text{dist}_{x2}$", "$\\sigma_{\\varepsilon}$", "$\\sigma_{X1}$", "$\\sigma_{X2}$", "n"),
+       Domain = c("\\{2, 3, ..., 18\\}", "[-1, 1]", "[0, 100]", "\\{0, 1\\}", "\\{0, 1\\}", "\\{discrete, uniform, normal, lognormal\\}", "\\{discrete, uniform, normal, lognormal\\}", "\\{discrete, uniform, normal, lognormal\\}", "[0.0625, 9]", "[0.3, 0.6]", "[0.3, 0.6]", "[50, 500]")) %>%
+  kableExtra::kable(escape = FALSE,
+                    booktabs = TRUE)
+```
+
+:::
+
+The residuals and fitted values of the fitted model were obtained by regressing $\boldsymbol{y}$ on $\boldsymbol{x}_1$.
+If $\beta_1 \neq 0$, $\boldsymbol{x}_2$ was also included in the design matrix.
+This data generation process was adapted from @li2023plot, where it was utilized to simulate residual plots exhibiting non-linearity and heteroskedasticity visual patterns for human subject experiments.
+A summary of the factors utilized in @eq-data-sim is provided in @tbl-factor.
+
+In @eq-data-sim, $\boldsymbol{z}$ and $\boldsymbol{w}$ represent higher-order terms of $\boldsymbol{x}_1$ and $\boldsymbol{x}_2$, respectively.
+If $\beta_2 \neq 0$, the regression model will encounter non-linearity issues.
+Parameter $j$ serves as a shape parameter that controls the number of tuning points in the non-linear pattern.
+Typically, higher values of $j$ lead to an increase in the number of tuning points, as illustrated in @fig-different-j.
+
+```{r fig-different-j, fig.pos="!h", fig.cap = "Non-linearity forms generated for the synthetic data simulation.
+The 17 shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$."}
+
+set.seed(10086)
+
+# Data for shape 1
+dat_shape_1 <- phn_model(j = 2, include_x2 = FALSE, sigma = 0.05)$gen(500) %>%
+  mutate(j = 2)
+
+# Generate data for shape 2, 3 and 4. Reuse x and e.
+map_df(3:18, function(j) {
+  phn_model(j = j, include_x2 = FALSE, sigma = 0.05)$
+    gen(500, computed = select(dat_shape_1, x1, e)) %>%
+  mutate(j = j)
+}) %>%
+
+  # Combined with data for shape 1
+  bind_rows(dat_shape_1) %>%
+  mutate(j = factor(j)) %>%
+  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
+  facet_wrap(~j, scales = "free", labeller = label_parsed, ncol = 5)
+```
+
+Additionally, Scaling factor $\boldsymbol{k}$ directly affects the error distribution and it is correlated with $\boldsymbol{x}_1$ and $\boldsymbol{x}_2$.
+If $b \neq 0$ and $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n)$, the constant variance assumption will be violated.
+Parameter $a$ is a shape parameter controlling the location of the smallest variance in a residual plot as shown in @fig-different-a.
+
+```{r fig-different-a, fig.width=8/5*3*2, fig.height=6/4*3*2, fig.pos="!h", fig.cap = 'Heteroskedasticity forms generated for the synthetic data simulation.
+Different shapes are controlled by the continuous factor $a$ between -1 and 1.
+For $a = -1$, the residual plot exhibits a "left-triangle" shape.
+And for $a = 1$, the residual plot exhibits a "right-triangle" shape.
+'}
+
+set.seed(10085)
+
+# Generate data for a = -1
+dat_a_n1 <- phn_model(include_z = FALSE,
+                      include_x2 = FALSE,
+                      a = -1,
+                      b = 100)$gen(500) %>%
+  mutate(a = -1)
+
+# Generate data for other a
+map(c(-0.75, -0.5, -0.25, 0, 0.25, 0.5, 0.75, 1), function(a) {
+  phn_model(include_z = FALSE,
+            include_x2 = FALSE,
+            a = a,
+            b = 100)$gen(500) %>%
+  mutate(a = a)
+}) %>%
+
+  # Combined with data for a = -1
+  bind_rows(dat_a_n1) %>%
+  mutate(a = factor(a)) %>%
+  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
+  facet_wrap(~a, scales = "free", ncol = 3) +
+  xlab("Fitted values") +
+  ylab("Residuals")
+```
+
+```{r fig-different-e, fig.pos="!h", fig.cap = 'Non-normality forms generated for the synthetic data simulation.
+Four different error distributions including discrete, lognormal, normal and uniform are considered.'}
+
+set.seed(10086)
+
+# Data for shape 1
+dat_shape_1 <- phn_model(include_z = FALSE, include_x2 = FALSE, e = rand_uniform(-1.4, 1.4))$gen(500) %>%
+  mutate(e_dist = "uniform")
+
+dat_shape_2 <- phn_model(include_z = FALSE, include_x2 = FALSE, sigma = 0.8)$gen(500) %>%
+  mutate(e_dist = "normal")
+
+dat_shape_3 <- phn_model(include_z = FALSE, include_x2 = FALSE, e = rand_lognormal(sigma = 0.6))$gen(500) %>%
+  mutate(e_dist = "lognormal")
+
+dat_shape_4 <- phn_model(include_z = FALSE, include_x2 = FALSE, e = rand_uniform_d(-1.4, 1.4, even = TRUE))$gen(500) %>%
+  mutate(e_dist = "discrete")
+
+# Generate data for shape 2, 3 and 4. Reuse x and e.
+bind_rows(dat_shape_1, dat_shape_2, dat_shape_3, dat_shape_4) %>%
+  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
+  facet_wrap(~e_dist, scales = "free", labeller = label_parsed, ncol = 2)
+```
+
+Non-normality violations arise from specifying a non-normal distribution for $\boldsymbol{\varepsilon}$.
+In the synthetic data simulation, four distinct error distributions are considered, including discrete, uniform, normal, and lognormal distributions, as presented in @fig-different-e.
+Each distribution imparts unique characteristics to the residuals.
+The discrete error distribution introduces discreteness in residuals, while the lognormal distribution typically yields outliers.
+Uniform error distribution may result in residuals filling the entire space of the residual plot.
+All of these distributions exhibit visual distinctions from the normal error distribution.
+
+```{r fig-different-j-x2, fig.pos="!h", fig.cap = "Residual plots of multiple linear regression models with non-linearity issues.
+The 17 shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$.
+A second predictor $\\boldsymbol{x}_2$ is introduced to the regression model to create complex shapes."}
+
+set.seed(10086)
+
+# Data for shape 1
+dat_shape_1 <- phn_model(j = 2, include_x2 = TRUE, sigma = 0.05)$gen(500) %>%
+  mutate(j = 2)
+
+# Generate data for shape 2, 3 and 4. Reuse x and e.
+map_df(3:18, function(j) {
+  phn_model(j = j, include_x2 = TRUE, sigma = 0.05)$
+    gen(500, computed = select(dat_shape_1, x1, e)) %>%
+  mutate(j = j)
+}) %>%
+
+  # Combined with data for shape 1
+  bind_rows(dat_shape_1) %>%
+  mutate(j = factor(j)) %>%
+  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
+  facet_wrap(~j, scales = "free", labeller = label_parsed, ncol = 5)
+```
+
+```{r fig-different-j-heter, fig.pos="!h", fig.cap = 'Residual plots of models violating both the non-linearity and the heteroskedasticity assumptions.
+The 17 shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$, and the "left-triangle" shape is introduced by setting $a = -1$.'}
+
+set.seed(10086)
+
+# Data for shape 1
+dat_shape_1 <- phn_model(j = 2, a = -1, b = 100, include_x2 = FALSE, sigma = 0.05)$gen(500) %>%
+  mutate(j = 2)
+
+# Generate data for shape 2, 3 and 4. Reuse x and e.
+map_df(3:18, function(j) {
+  phn_model(j = j, a = -1, b = 100, include_x2 = FALSE, sigma = 0.05)$
+    gen(500, computed = select(dat_shape_1, x1, e)) %>%
+  mutate(j = j)
+}) %>%
+
+  # Combined with data for shape 1
+  bind_rows(dat_shape_1) %>%
+  mutate(j = factor(j)) %>%
+  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
+  facet_wrap(~j, scales = "free", labeller = label_parsed, ncol = 5)
+```
+
+```{r fig-different-e-heter, fig.pos="!h", fig.cap = 'Residual plots of models violating both the non-normality and the heteroskedasticity assumptions.
+The four shapes are generated by using four different error distributions including discrete, lognormal, normal and uniform, and the "left-triangle" shape is introduced by setting $a = -1$.
+'}
+
+set.seed(10085)
+
+# Data for shape 1
+dat_shape_1 <- phn_model(a = -1, b = 100, include_z = FALSE, include_x2 = FALSE, e = rand_uniform(-1.4, 1.4))$gen(500) %>%
+  mutate(e_dist = "uniform")
+
+dat_shape_2 <- phn_model(a = -1, b = 100, include_z = FALSE, include_x2 = FALSE, sigma = 0.8)$gen(500) %>%
+  mutate(e_dist = "normal")
+
+dat_shape_3 <- phn_model(a = -1, b = 100, include_z = FALSE, include_x2 = FALSE, e = rand_lognormal(sigma = 0.6))$gen(500) %>%
+  mutate(e_dist = "lognormal")
+
+dat_shape_4 <- phn_model(a = -1, b = 100, include_z = FALSE, include_x2 = FALSE, e = rand_uniform_d(-1.4, 1.4, even = TRUE))$gen(500) %>%
+  mutate(e_dist = "discrete")
+
+# Generate data for shape 2, 3 and 4. Reuse x and e.
+bind_rows(dat_shape_1, dat_shape_2, dat_shape_3, dat_shape_4) %>%
+  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
+  facet_wrap(~e_dist, scales = "free", labeller = label_parsed, ncol = 2)
+```
+
+@eq-data-sim accommodates the incorporation of the second predictor $\boldsymbol{x}_2$.
+Introducing it into the data generation process by setting $\beta_1 = 1$ significantly enhances the complexity of the shapes, as illustrated in @fig-different-j-x2.
+In comparison to @fig-different-j, @fig-different-j-x2 demonstrates that the non-linear shape resembles a surface rather than a single curve.
+This augmentation can facilitate the computer vision model in learning visual patterns from residual plots of the multiple linear regression model.
+
+In real-world analysis, it's not uncommon to encounter instances where multiple model violations coexist.
+In such cases, the residual plots often exhibit a mixed pattern of visual anomalies corresponding to different types of model violations.
+@fig-different-j-heter and @fig-different-e-heter show the visual patterns of models with multiple model violations.
+
+### Scagnostics
+
+In @sec-model-introduction, we mentioned that scagnostics consist of a set of manually designed visual feature extraction functions.
+While our computer vision model will learn its own feature extraction function during training, leveraging additional information from scagnostics can enhance the model's predictive accuracy.
+
+For each generated residual plot, we computed four scagnostics – "Monotonic", "Sparse", "Splines", and "Striped" – using the `cassowaryr` R package [@mason2022cassowaryr].
+These computed measures, along with the number of observations from the fitted model, were provided as the second input for the computer vision model.
+While other scagnostics provide valuable insights, they come with high computational costs and are not suitable for quick inference.
+
+### Balanced Dataset
+
+To train a robust computer vision model, we deliberately controlled the distribution of the target variable $D$ in the training data.
+We ensured that it followed a uniform distribution between $0$ and $7$.
+This was achieved by organizing $50$ buckets, each exclusively accepting training samples with $D$ falling within the range $[7(i - 1)/49, 7i/49)$ for $i < 50$, where $i$ represents the index of the $i$-th bucket.
+For the $50$-th bucket, any training samples with $D \geq 7$ were accepted.
+
+With 80000 training images prepared, each bucket accommodated a maximum of $80000 \div 50 = 1600$ training samples.
+The simulator iteratively sampled parameter values from the parameter space, generated residuals and fitted values using the data generation process, computed the distance, and checked if the sample fitted within the corresponding bucket.
+This process continued until all buckets were filled.
+
+Similarly, we adopted the same methodology to prepare 8000 test images for performance evaluation and model diagnostics.
+
+## Model Architecture {#sec-model-architecture}
+
+```{r out.width = "100%"}
+#| label: fig-cnn-diag
+#| fig-cap: "Diagram of the architecture of the optimized computer vision model.
+Numbers at the bottom of each box show the shape of the output of each layer.
+The band of each box drawn in a darker colour indicates the use of the rectified linear unit activation function.
+Yellow boxes are 2D convolutional layers, orange boxes are pooling layers, the grey box is the concatenation layer, and the purple boxes are dense layers."
+magick::image_read_pdf("figures/cnn.pdf")
+```
+
+The architecture of the computer vision model is adapted from a well-established architecture known as VGG16, which has demonstrated high performance in image classification [@simonyan2014very].
+@fig-cnn-diag provides a diagram of the architecture.
+More details about the neural network layers used in this study are provided in @sec-appendix-b.
+
+The model begins with an input layer of shape $n \times h \times w \times 3$, capable of handling $n$ RGB images.
+This is followed by a grayscale conversion layer utilizing the luma formula under the Rec.
+601 standard, which converts the colour image to grayscale.
+Grayscale suffices for our task since data points are plotted in black.
+We experiment with three combinations of $h$ and $w$: $32 \times 32$, $64 \times 64$, and $128 \times 128$, aiming to achieve sufficiently high image resolution for the problem at hand.
+
+The processed image is used as the input for the first convolutional block.
+The model comprises at most five consecutive convolutional blocks, mirroring the original VGG16 architecture.
+Within each block, there are two 2D convolutional layers followed by two activation layers, respectively.
+Subsequently, a 2D max-pooling layer follows the second activation layer.
+The 2D convolutional layer convolves the input with a fixed number of $3 \times 3$ convolution filters, while the 2D max-pooling layer downsamples the input along its spatial dimensions by taking the maximum value over a $2 \times 2$ window for each channel of the input.
+The activation layer employs the rectified linear unit (ReLU) activation function, a standard practice in deep learning, which introduces a non-linear transformation of the output of the 2D convolutional layer.
+Additionally, to regularize training, a batch normalization layer is added after each 2D convolutional layer and before the activation layer.
+Finally, a dropout layer is appended at the end of each convolutional block to randomly set some inputs to zero during training, further aiding in regularization.
+
+The output of the last convolutional block is summarized by either a global max pooling layer or a global average pooling layer, resulting in a two-dimensional tensor.
+To leverage the information contained in scagnostics, this tensor is concatenated with an additional $n \times 5$ tensor, which contains the "Monotonic", "Sparse", "Splines", and "Striped" measures, along with the number of observations for $n$ residual plots.
+
+The concatenated tensor is then fed into the final prediction block.
+This block consists of two fully-connected layers.
+The first layer contains at least $128$ units, followed by a dropout layer.
+Occasionally, a batch normalization layer is inserted between the fully-connected layer and the dropout layer for regularization purposes.
+The second fully-connected layer consists of only one unit, serving as the output of the model.
+
+The model weights $\boldsymbol{\theta}$ were randomly initialized and they were optimized by the Adam optimizer with the mean square error loss function
+
+$$\hat{\boldsymbol{\theta}} = \underset{\boldsymbol{\theta}}{argmin}\frac{1}{n_{train}}\sum_{i=1}^{n_{train}}(D_i - f_{\boldsymbol{\theta}}(V_i, S_i))^2,$$
+
+\noindent where $n_{train}$ is the number of training samples, $V_i$ is the $i$-th residual plot and $S_i$ is the additional information about the $i$-th residual plot including four scagnostics and the number of observations.
+
+## Model Training {#sec-model-training}
+
+To achieve a near-optimal deep learning model, hyperparameters like the learning rate often need to be fine-tuned using a tuner.
+In our study, we utilized the Bayesian optimization tuner from the `KerasTuner` Python library [@omalley2019kerastuner] for this purpose.
+A comprehensive list of hyperparameters is provided in @tbl-hyperparameter.
+
+The "number of base filters" determines the number of filters for the first 2D convolutional layer.
+In the VGG16 architecture, the number of filters for the 2D convolutional layer in a block is typically twice the number in the previous block, except for the last block, which maintains the same number of convolution filters as the previous one.
+This hyperparameter aids in controlling the complexity of the computer vision model.
+Higher numbers of base filters result in more trainable parameters.
+Likewise, the "number of units for the fully-connected layer" determines the complexity of the final prediction block.
+Increasing the number of units enhances model complexity, resulting in more trainable parameters.
+
+The dropout rate and batch normalization are flexible hyperparameters that work in conjunction with other parameters to facilitate smooth training.
+A higher dropout rate is necessary when the model tends to overfit the training dataset by learning too much noise.
+Conversely, a lower dropout rate is preferred when the model is complex and challenging to converge.
+Batch normalization, on the other hand, addresses the internal covariate shift problem arising from the randomness in weight initialization and input data.
+It helps stabilize and accelerate the training process by normalizing the activations of each layer.
+
+Additionally, incorporating additional inputs such as scagnostics and the number of observations can potentially enhance prediction accuracy.
+Therefore, we allowed the tuner to determine whether these inputs were necessary for optimal model performance.
+
+The learning rate is a crucial hyperparameter, as it dictates the step size of the optimization algorithm.
+A high learning rate can help the model avoid local minima but risks overshooting and missing the global minimum.
+Conversely, a low learning rate smoothens the training process but makes the convergence time longer and increases the likelihood of getting trapped in local minima.
+
+Our model was trained on the MASSIVE M3 high-performance computing platform [@goscinski2014multi], using TensorFlow [@abadi2016tensorflow] and Keras [@chollet2015keras].
+During training, 80% of the training data was utilized for actual training, while the remaining 20% was used as validation data.
+The Bayesian optimization tuner conducted 100 trials to identify the best hyperparameter values based on validation root mean square error.
+The tuner then restored the best epoch of the best model from the trials.
+Additionally, we applied early stopping, terminating the training process if the validation root mean square error fails to improve for 50 epochs.
+The maximum allowed epochs was set at 2000, although no models reached this threshold.
+
+```{r}
+#| label: tbl-hyperparameter
+#| tbl-cap: "Name of hyperparameters and their correspoding domain for the computer vision model."
+data.frame(`Hyperparameter` = c("Number of base filters", "Dropout rate for convolutional blocks", "Batch normalization for convolutional blocks", "Type of global pooling", "Ignore additional inputs", "Number of units for the fully-connected layer", "Batch normalization for the fully-connected layer", "Dropout rate for the fully-connected layer", "Learning rate"),
+           Domain = c("{4, 8, 16, 32, 64}", "[0.1, 0.6]", "{false, true}", "{max, average}", "{false, true}", "{128, 256, 512, 1024, 2048}", "{false, true}", "[0.1, 0.6]", "[1e-8, 1e-1]")) %>%
+  kableExtra::kable(booktabs = TRUE)
+```
+
+Based on the tuning process described above, the optimized hyperparameter values are presented in @tbl-best-hyperparameter.
+It was observable that a minimum of $32$ base filters was necessary, with the preferable choice being $64$ base filters for both the $64 \times 64$ and $128 \times 128$ models, mirroring the original VGG16 architecture.
+The optimized dropout rate for convolutional blocks hovered around $0.4$, and incorporating batch normalization for convolutional blocks proved beneficial for performance.
+
+All optimized models chose to retain the additional inputs, contributing to the reduction of validation error.
+The number of units required for the fully-connected layer was $256$, a relatively modest number compared to the VGG16 classifier, suggesting that the problem at hand was less complex.
+The optimized learning rates were smaller than the default value recommended by Keras, which is 0.001.
+
+::: {.content-visible when-format="html"}
+
+::: {#tbl-best-hyperparameter .cell tbl-cap='Hyperparameters values for the optimized computer vision models with different input sizes.'}
+::: {.cell-output-display}
+
+`````{=html}
+<table class="table" style="margin-left: auto; margin-right: auto;">
+ <thead>
+  <tr>
+   <th style="text-align:left;"> Hyperparameter </th>
+   <th style="text-align:left;"> <span data-qmd="$32 \times 32$"> </span> </th>
+   <th style="text-align:left;"> <span data-qmd="$64 \times 64$"> </span> </th>
+   <th style="text-align:left;"> <span data-qmd="$128 \times 128$"> </span> </th>
+  </tr>
+ </thead>
+<tbody>
+  <tr>
+   <td style="text-align:left;"> Number of base filters </td>
+   <td style="text-align:left;"> 32 </td>
+   <td style="text-align:left;"> 64 </td>
+   <td style="text-align:left;"> 64 </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> Dropout rate for convolutional blocks </td>
+   <td style="text-align:left;"> 0.4 </td>
+   <td style="text-align:left;"> 0.3 </td>
+   <td style="text-align:left;"> 0.4 </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> Batch normalization for convolutional blocks </td>
+   <td style="text-align:left;"> true </td>
+   <td style="text-align:left;"> true </td>
+   <td style="text-align:left;"> true </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> Type of global pooling </td>
+   <td style="text-align:left;"> max </td>
+   <td style="text-align:left;"> average </td>
+   <td style="text-align:left;"> average </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> Ignore additional inputs </td>
+   <td style="text-align:left;"> false </td>
+   <td style="text-align:left;"> false </td>
+   <td style="text-align:left;"> false </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> Number of units for the fully-connected layer </td>
+   <td style="text-align:left;"> 256 </td>
+   <td style="text-align:left;"> 256 </td>
+   <td style="text-align:left;"> 256 </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> Batch normalization for the fully-connected layer </td>
+   <td style="text-align:left;"> false </td>
+   <td style="text-align:left;"> true </td>
+   <td style="text-align:left;"> true </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> Dropout rate for the fully-connected layer </td>
+   <td style="text-align:left;"> 0.2 </td>
+   <td style="text-align:left;"> 0.4 </td>
+   <td style="text-align:left;"> 0.1 </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> Learning rate </td>
+   <td style="text-align:left;"> 0.0003 </td>
+   <td style="text-align:left;"> 0.0006 </td>
+   <td style="text-align:left;"> 0.0052 </td>
+  </tr>
+</tbody>
+</table>
+
+`````
+
+:::
+:::
+
+:::
+
+::: {.content-visible when-format="pdf"}
+
+```{r}
+#| label: tbl-best-hyperparameter
+#| tbl-cap: "Hyperparameters values for the optimized computer vision models with different input sizes."
+data.frame(`Hyperparameter` = c("Number of base filters", "Dropout rate for convolutional blocks", "Batch normalization for convolutional blocks", "Type of global pooling", "Ignore additional inputs", "Number of units for the fully-connected layer", "Batch normalization for the fully-connected layer", "Dropout rate for the fully-connected layer", "Learning rate"),
+           `32` = c("32", "0.4", "true", "max", "false", "256", "false", "0.2", "0.0003"),
+           `64` = c("64", "0.3", "true", "average", "false", "256", "true", "0.4", "0.0006"),
+           `128` = c("64", "0.4", "true", "average", "false", "256", "true", "0.1", "0.0052")) %>%
+  kableExtra::kable(booktabs = TRUE, 
+                    col.names = c("Hyperparameter", "$32 \\times 32$", "$64 \\times 64$", "$128 \\times 128$"),
+                    escape = FALSE) %>%
+  kableExtra::kable_styling(latex_options = c("scale_down"))
+```
+
+:::
+
+## Results {#sec-model-results}
+
+### Model Performance
+
+The training and test performance for the optimized models with three different input sizes are summarized in @tbl-performance.
+Among these models, the $64 \times 64$ model and the $32 \times 32$ model consistently exhibited the best training and test performance, respectively.
+The mean absolute error indicated that the difference between $\hat{D}$ and $D$ was approximately $0.43$ on the test set, a negligible deviation considering the normal range of $D$ typically falls between $0$ and $7$.
+The high $R^2$ values also suggested that the predictions were largely linearly correlated with the target.
+
+@fig-model-performance presents a hexagonal heatmap for $D - \hat{D}$ versus $\hat{D}$.
+The red smoothing curves, fitted by generalized additive models [@hastie2017generalized], demonstrate that all the optimized models perform admirably on both the training and test sets.
+No structural issues are noticeable in @fig-model-performance, but some minor issues regarding over-prediction and under-prediction are observed.
+
+The figure highlights that most under-predictions occurred when $\hat{D} < 3$, while over-predictions occurred predominantly when $3 < \hat{D} < 6$.
+For input images representing null plots, it was expected that the models will over-predict the distance, as explained in @sec-model-lineup-evaluation.
+However, it can not explain the under-prediction issue.
+Therefore, we analysed the relationship between residuals and all the factors involved in the data generating process.
+We found that most issues actually arose from non-linearity problems and the presence of a second predictor in the regression model as illustrated in @fig-over-under.
+When the variance for the error distribution was small, the optimized model tended to under-predict the distance.
+Conversely, when the error distribution had a large variance, the model tended to over-predict the distance.
+
+Since most of the deviation stemmed from the presence of non-linearity violations, to further investigate this, we split the test set based on violation types and re-evaluated the performance, as detailed in @tbl-performance-sub.
+It was evident that metrics for null plots were notably worse compared to other categories.
+Furthermore, residual plots solely exhibiting non-normality issues were the easiest to predict, with very low test root mean square error (RMSE) at around $0.3$.
+Residual plots with non-linearity issues were more challenging to assess than those with heteroskedasticity or non-normality issues.
+Assessing residual plots with heteroskedasticity was not as difficult as assessing those with non-linearity issues.
+When multiple violations were introduced to a residual plot, the performance metrics typically lay between the metrics for each individual violation.
+
+Based on the model performance metrics, we chose to use the best-performing model evaluated on the test set, namely the $32 \times 32$ model, for the subsequent analysis.
+
+```{r}
+test_pred <- readr::read_csv(here::here("data/test_pred.csv"))
+train_pred <- readr::read_csv(here::here("data/train_pred.csv"))
+meta <- readr::read_csv(here::here("data/meta.csv"))
+
+test_summary <- test_pred %>% 
+  left_join(meta) %>%
+  group_by(res) %>%
+  summarise(RMSE = yardstick::rmse_vec(effect_size, vss),
+            R2 = yardstick::rsq_vec(effect_size, vss),
+            MAE = yardstick::mae_vec(effect_size, vss),
+            HUBER = yardstick::huber_loss_vec(effect_size, vss)) %>%
+  mutate(type = "test")
+
+train_summary <- train_pred %>% 
+  left_join(meta) %>%
+  group_by(res) %>%
+  summarise(RMSE = yardstick::rmse_vec(effect_size, vss),
+            R2 = yardstick::rsq_vec(effect_size, vss),
+            MAE = yardstick::mae_vec(effect_size, vss),
+            HUBER = yardstick::huber_loss_vec(effect_size, vss)) %>%
+  mutate(type = "train")
+```
+
+::: {.content-visible when-format="pdf"}
+
+```{r}
+#| label: tbl-performance
+#| tbl-cap: "The training and test performance of three optimized models with different input sizes."
+train_summary %>%
+  bind_rows(test_summary) %>%
+  select(res, RMSE, R2, MAE, HUBER) %>%
+  mutate(res = glue::glue("${res} \\times {res}$")) %>%
+  mutate(across(RMSE:MAE, ~format(.x, digits = 3))) %>%
+  mutate(across(HUBER, ~format(.x, digits = 2))) %>%
+  kableExtra::kable(booktabs = TRUE, 
+                    escape = FALSE,
+                    col.names = c("", "RMSE", "$R^2$", "MAE", "Huber loss")) %>%
+  kableExtra::pack_rows("Training set", 1, 3) %>%
+  kableExtra::pack_rows("Test set", 4, 6)
+```
+
+:::
+
+::: {.content-visible when-format="html"}
+
+::: {#tbl-performance .cell tbl-cap='The training and test performance of three optimized models with different input sizes.'}
+::: {.cell-output-display}
+
+```{=html}
+<table>
+ <thead>
+  <tr>
+   <th style="text-align:left;">  </th>
+   <th style="text-align:left;"> RMSE </th>
+   <th style="text-align:left;"> <span data-qmd="$R^2$"></span> </th>
+   <th style="text-align:left;"> MAE </th>
+   <th style="text-align:left;"> Huber loss </th>
+  </tr>
+ </thead>
+<tbody>
+  <tr grouplength="3"><td colspan="5" style="border-bottom: 1px solid;"><strong>Training set</strong></td></tr>
+<tr>
+   <td style="text-align:left;padding-left: 2em;" indentlevel="1"> <span data-qmd="$32 \times 32$"></span> </td>
+   <td style="text-align:left;"> 0.531 </td>
+   <td style="text-align:left;"> 0.937 </td>
+   <td style="text-align:left;"> 0.364 </td>
+   <td style="text-align:left;"> 0.126 </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;padding-left: 2em;" indentlevel="1"> <span data-qmd="$64 \times 64$"></span> </td>
+   <td style="text-align:left;"> 0.405 </td>
+   <td style="text-align:left;"> 0.963 </td>
+   <td style="text-align:left;"> 0.260 </td>
+   <td style="text-align:left;"> 0.072 </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;padding-left: 2em;" indentlevel="1"> <span data-qmd="$128 \times 128$"></span> </td>
+   <td style="text-align:left;"> 0.432 </td>
+   <td style="text-align:left;"> 0.959 </td>
+   <td style="text-align:left;"> 0.290 </td>
+   <td style="text-align:left;"> 0.084 </td>
+  </tr>
+  <tr grouplength="3"><td colspan="5" style="border-bottom: 1px solid;"><strong>Test set</strong></td></tr>
+<tr>
+   <td style="text-align:left;padding-left: 2em;" indentlevel="1"> <span data-qmd="$32 \times 32$"></span> </td>
+   <td style="text-align:left;"> 0.660 </td>
+   <td style="text-align:left;"> 0.901 </td>
+   <td style="text-align:left;"> 0.434 </td>
+   <td style="text-align:left;"> 0.181 </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;padding-left: 2em;" indentlevel="1"> <span data-qmd="$64 \times 64$"></span> </td>
+   <td style="text-align:left;"> 0.674 </td>
+   <td style="text-align:left;"> 0.897 </td>
+   <td style="text-align:left;"> 0.438 </td>
+   <td style="text-align:left;"> 0.186 </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;padding-left: 2em;" indentlevel="1"> <span data-qmd="$128 \times 128$"></span> </td>
+   <td style="text-align:left;"> 0.692 </td>
+   <td style="text-align:left;"> 0.892 </td>
+   <td style="text-align:left;"> 0.460 </td>
+   <td style="text-align:left;"> 0.199 </td>
+  </tr>
+</tbody>
+</table>
+
+```
+
+:::
+:::
+
+:::
+
+```{r fig-model-performance, fig.pos="!h", fig.cap = "Hexagonal heatmap for residuals vs predicted distance on training and test data for three optimized models with different input sizes.
+The brown lines are smoothing curves produced by fitting gnealized additive models.
+The area over the zero line in light yellow indicates under-prediction, and the area under the zero line in light green indicates over-prediction.", dev = 'png', dpi = 300}
+model_pred <- train_pred %>% 
+  left_join(meta) %>%
+  mutate(type = "train") %>%
+  bind_rows(test_pred %>% 
+              left_join(meta) %>%
+              mutate(type = "test"))
+
+model_pred %>%
+  ggplot() +
+  geom_hline(yintercept = 0, alpha = 0.5) +
+  # geom_point(aes(effect_size, effect_size - vss), alpha = 0.03) +
+  annotate("rect", 
+           ymin = min(model_pred$effect_size - model_pred$vss),
+                ymax = 0,
+                xmin = min(model_pred$vss),
+                xmax = max(model_pred$vss),
+            fill = "#40B0A6",
+            alpha = 0.3) +
+  geom_point(data = NULL, aes(5, 0, col = "over-prediction"), shape = 15) +
+  annotate("rect", 
+           ymin = 0,
+                ymax = max(model_pred$effect_size - model_pred$vss),
+                xmin = min(model_pred$vss),
+                xmax = max(model_pred$vss),
+            fill = "#E1BE6A",
+            alpha = 0.3) +
+  geom_point(data = NULL, aes(5, 0, col = "under-prediction"), shape = 15) +
+  geom_hex(aes(vss, effect_size - vss), bins = 20) +
+  geom_smooth(aes(vss, effect_size - vss), se = FALSE, col = "#994F00") +
+  facet_grid(type ~ res) +
+  coord_fixed() +
+  ylab(expression(D - hat(D))) +
+  xlab(expression(hat(D))) +
+  theme_light() +
+  scale_fill_continuous(limits = c(1, 6000), trans = "log10", low = "#56B1F7", high = "#132B43") +
+  scale_color_manual(values = c("under-prediction" = scales::alpha("#E1BE6A", 0.3),
+                                "over-prediction" = scales::alpha("#40B0A6", 0.3)),
+                     breaks = c("under-prediction", "over-prediction")) +
+  guides(col = guide_legend(title = NULL, override.aes = list(size = 8), order = 1))
+```
+
+```{r fig-over-under, fig.pos="!h", fig.cap = "Scatter plots for residuals vs $\\sigma$ on test data for the $32 \\times 32$ optimized model.
+The data is grouped by whether the regression has only non-linearity violation, and whether it includes a second predictor in the regression formula.
+The brown lines are smoothing curves produced by fitting gnealized additive models.
+The area over the zero line in light yellow indicates under-prediction, and the area under the zero line in light green indicates over-prediction.", dev = 'png', dpi = 300}
+model_pred %>%
+  filter(res == 32L) %>%
+  filter(type == "test") %>%
+  mutate(only_z = include_z & !include_heter & !include_non_normal) %>%
+  rename(`Only has non-linearity violation` = only_z, `Has second predictor` = include_x2) %>%
+  ggplot() +
+  geom_hline(yintercept = 0, alpha = 0.5) +
+  annotate("rect", 
+           ymin = min(model_pred$effect_size[model_pred$type == "test"] - model_pred$vss[model_pred$type == "test"]),
+                ymax = 0,
+                xmin = min(model_pred$e_sigma),
+                xmax = max(model_pred$e_sigma),
+            fill = "#40B0A6",
+            alpha = 0.3) +
+  annotate("rect", 
+           ymin = 0,
+                ymax = max(model_pred$effect_size[model_pred$type == "test"] - model_pred$vss[model_pred$type == "test"]),
+                xmin = min(model_pred$e_sigma),
+                xmax = max(model_pred$e_sigma),
+            fill = "#E1BE6A",
+            alpha = 0.3) +
+  geom_point(aes(e_sigma, effect_size - vss), alpha = 0.2) +
+  geom_smooth(aes(e_sigma, effect_size - vss), se = FALSE, col = "#994F00") +
+  facet_grid(`Has second predictor` ~ `Only has non-linearity violation`,
+             labeller = label_both) +
+  theme_light() +
+  ylab(expression(D - hat(D))) +
+  xlab(expression(sigma))
+```
+
+::: {.content-visible when-format="html"}
+
+::: {#tbl-performance-sub .cell tbl-cap='The training and test performance of the $32 \times 32$ model presented with different model violations.'}
+::: {.cell-output-display}
+
+```{=html}
+<table class="table" style="margin-left: auto; margin-right: auto;">
+ <thead>
+  <tr>
+   <th style="text-align:left;"> Violations </th>
+   <th style="text-align:left;"> #Training samples </th>
+   <th style="text-align:left;"> Training RMSE </th>
+   <th style="text-align:left;"> #Test samples </th>
+   <th style="text-align:left;"> Test RMSE </th>
+  </tr>
+ </thead>
+<tbody>
+  <tr>
+   <td style="text-align:left;"> no violations </td>
+   <td style="text-align:left;"> 1546 </td>
+   <td style="text-align:left;"> 1.149 </td>
+   <td style="text-align:left;"> 155 </td>
+   <td style="text-align:left;"> 1.267 </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> non-linearity </td>
+   <td style="text-align:left;"> 22184 </td>
+   <td style="text-align:left;"> 0.625 </td>
+   <td style="text-align:left;"> 2218 </td>
+   <td style="text-align:left;"> 0.787 </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> heteroskedasticity </td>
+   <td style="text-align:left;"> 10826 </td>
+   <td style="text-align:left;"> 0.528 </td>
+   <td style="text-align:left;"> 1067 </td>
+   <td style="text-align:left;"> 0.602 </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> non-linearity + heteroskedasticity </td>
+   <td style="text-align:left;"> 10221 </td>
+   <td style="text-align:left;"> 0.593 </td>
+   <td style="text-align:left;"> 985 </td>
+   <td style="text-align:left;"> 0.751 </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> non-normality </td>
+   <td style="text-align:left;"> 10797 </td>
+   <td style="text-align:left;"> 0.279 </td>
+   <td style="text-align:left;"> 1111 </td>
+   <td style="text-align:left;"> 0.320 </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> non-linearity + non-normality </td>
+   <td style="text-align:left;"> 8835 </td>
+   <td style="text-align:left;"> 0.472 </td>
+   <td style="text-align:left;"> 928 </td>
+   <td style="text-align:left;"> 0.600 </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> heteroskedasticity + non-normality </td>
+   <td style="text-align:left;"> 7870 </td>
+   <td style="text-align:left;"> 0.354 </td>
+   <td style="text-align:left;"> 819 </td>
+   <td style="text-align:left;"> 0.489 </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> non-linearity + heteroskedasticity + non-normality </td>
+   <td style="text-align:left;"> 7721 </td>
+   <td style="text-align:left;"> 0.432 </td>
+   <td style="text-align:left;"> 717 </td>
+   <td style="text-align:left;"> 0.620 </td>
+  </tr>
+</tbody>
+</table>
+
+```
+
+:::
+:::
+
+:::
+
+::: {.content-visible when-format="pdf"}
+
+```{r}
+#| label: tbl-performance-sub
+#| tbl-cap: "The training and test performance of the $32 \\times 32$ model presented with different model violations."
+model_pred %>%
+  filter(res == 32L) %>%
+  group_by(include_non_normal, include_heter, include_z) %>%
+  summarise(train_n = sum(type == "train"),
+            train_RMSE = yardstick::rmse_vec(effect_size[type == "train"], vss[type == "train"]),
+            test_n = sum(type == "test"),
+            test_RMSE = yardstick::rmse_vec(effect_size[type == "test"], vss[type == "test"])) %>%
+  mutate(violations = ifelse(include_z, "non-linearity", "null")) %>%
+  mutate(violations = ifelse(include_heter, glue("{violations} + heteroskedasticity"), violations)) %>%
+  mutate(violations = ifelse(include_non_normal, glue("{violations} + non-normality"), violations)) %>%
+  mutate(violations = gsub("null \\+ ", "", violations)) %>%
+  ungroup() %>%
+  select(violations, train_n, train_RMSE, test_n, test_RMSE) %>%
+  mutate(train_n = format(train_n)) %>%
+  mutate(test_n = format(test_n)) %>%
+  mutate(across(c(train_RMSE, test_RMSE), ~format(.x, digits = 3))) %>%
+  mutate(violations = ifelse(violations == "null", "no violations", violations)) %>%
+  kableExtra::kable(booktabs = TRUE,
+                    escape = FALSE,
+                    col.names = c("Violations", "\\#Training samples", "Training RMSE", "\\#Test samples", "Test RMSE")) %>%
+  kableExtra::kable_styling(latex_options = c("scale_down"))
+```
+
+:::
+
+<!-- - check if j has an impact on residuals more carefully -->
+<!-- - the RMSE has very little to do with j if x2 is not included. And its impact is not as large as sigma. -->
+
+### Comparison with Human Visual Inference and Conventional Tests
+
+### Overview of the Human Subject experiment
+
+In order to check the validity of the proposed computer vision model, residual plots presented in the human subject experiment conducted by @li2023plot will be assessed.
+
+<!-- The experiment revealed that conventional tests are more sensitive to weak departures from model assumptions than visual test, and they often reject the null hypothesis when departures are not visibly different from null residual plots.
+-->
+
+This study has collected $7974$ human responses to $1152$ lineups.
+Each lineup contains one randomly placed true residual plot and 19 null plots.
+Among the $1152$ lineups, $24$ are attention check lineups in which the visual patterns are designed to be extremely obvious and very different from the corresponding to null plots, $36$ are null lineups where all the lineups consist of only null plots, $279$ are lineups with uniform predictor distribution evaluated by $11$ participants, and the remaining $813$ are lineups with discrete, skewed or normal predictor distribution evaluated by $5$ participants.
+Attention check lineups and null lineups will not be assessed in the following analysis.
+
+In @li2023plot, the residual plots are simulated from a data generating process which is a special case of @eq-data-sim.
+The main characteristic is the model violations are introduced separately, meaning non-linearity and heteroskedasticity will not co-exist in one lineup but assigned uniformly to all lineups.
+Additionally, non-normality and multiple predictors are not considered in the experimental design.
+
+### Model performance on the human data
+
+```{r rd-human}
+vss_32 <- readRDS(here::here("data/vss_32.rds"))
+
+experiment <- vi_survey %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  select(unique_lineup_id, attention_check, null_lineup, prop_detect,
+         answer, effect_size, conventional_p_value,
+         p_value, type, shape, a, b, x_dist, e_dist,
+         e_sigma, include_z, k, n)
+
+experiment <- vss_32 %>%
+  left_join(experiment) %>%
+  mutate(conventional_reject = conventional_p_value <= 0.05) %>%
+  mutate(reject = p_value <= 0.05) %>%
+  mutate(model_reject = vss_p_value <= 0.05) %>%
+  mutate(model_boot_reject = vss_boot_p_value <= 0.05)
+
+experiment <- experiment %>%
+  left_join(readRDS(here::here("data/actual_ss.rds")))
+```
+
+::: {.content-visible when-format="html"}
+
+::: {#tbl-experiment-performance .cell tbl-cap='The performance of the $32 \times 32$ model on the data used in the human subject experiment.'}
+::: {.cell-output-display}
+
+```{=html}
+<table>
+ <thead>
+  <tr>
+   <th style="text-align:left;"> Voilation </th>
+   <th style="text-align:left;"> RMSE </th>
+   <th style="text-align:left;"> <span data-qmd="$R^2$"></span> </th>
+   <th style="text-align:left;"> MAE </th>
+   <th style="text-align:left;"> Huber loss </th>
+  </tr>
+ </thead>
+<tbody>
+  <tr>
+   <td style="text-align:left;"> heteroskedasticity </td>
+   <td style="text-align:left;"> 0.721 </td>
+   <td style="text-align:left;"> 0.852 </td>
+   <td style="text-align:left;"> 0.553 </td>
+   <td style="text-align:left;"> 0.235 </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> non-linearity </td>
+   <td style="text-align:left;"> 0.738 </td>
+   <td style="text-align:left;"> 0.770 </td>
+   <td style="text-align:left;"> 0.566 </td>
+   <td style="text-align:left;"> 0.246 </td>
+  </tr>
+</tbody>
+</table>
+
+```
+
+:::
+:::
+
+:::
+
+::: {.content-visible when-format="pdf"}
+
+```{r}
+#| label: tbl-experiment-performance
+#| tbl-cap: "The performance of the $32 \\times 32$ model on the data used in the human subject experiment."
+experiment %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  group_by(type) %>%
+  summarise(RMSE = yardstick::rmse_vec(actual_ss, vss),
+            R2 = yardstick::rsq_vec(actual_ss, vss),
+            MAE = yardstick::mae_vec(actual_ss, vss),
+            HUBER = yardstick::huber_loss_vec(actual_ss, vss)) %>%
+  mutate(across(RMSE:HUBER, ~format(.x, digits = 3))) %>%
+  kableExtra::kable(booktabs = TRUE, 
+                    escape = FALSE,
+                    col.names = c("Voilation", "RMSE", "$R^2$", "MAE", "Huber loss"))
+```
+
+:::
+
+For each lineup used in @li2023plot, there is one true residual plot and 19 null plots.
+While the distance $D$ for the true residual plot depends on the underlying data generating process, the distance $D$ for the null plots is zero.
+We have used our optimized computer vision model to estimate distance for both the true residual plots and the null plots.
+To have a fair comparison, $H_0$ will be rejected if the true residual plot has the greatest approximated distance among all plots in a lineup.
+Additionally, the appropriate conventional tests including the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey1969tests] for non-linearity and the Breusch-Pagan test  [@breusch1979simple] for heteroskedasticity were applied on the same data for comparison.
+
+The performance metrics of $\hat{D}$ for true residual plots are outlined in @tbl-experiment-performance.
+It's notable that all performance metrics are slightly worse than those evaluated on the test data.
+Nevertheless, the mean absolute error remains at a low level, and the linear correlation between the prediction and the true value remains very high.
+Lineups with non-linearity issues are more challenging to predict than those with heteroskedasticity issues.
+
+@tbl-human-conv-table provides a summary of the agreement between decisions made by the computer vision model and conventional tests.
+The agreement rates between conventional tests and the computer vision model are 85.95% and 79.69% for residual plots containing heteroskedasticity and non-linearity patterns, respectively.
+These figures are higher than those calculated for visual tests conducted by human, indicating that the computer vision model exhibits behaviour more akin to the best available conventional tests.
+However, @fig-conv-mosaic shows that the computer vision model does not always reject when the conventional tests reject.
+And a small number of plots will be rejected by computer vision model but not by conventional tests.
+This suggests that conventional tests are more sensitive than the computer vision model.
+
+@fig-pcp further illustrates the decisions made by visual tests conducted by human, computer vision models, and conventional tests, using a parallel coordinate plots.
+It can be observed that all three tests will agree with each other for around 50% of the cases.
+When visual tests conducted by human do not reject, there are substantial amount of cases where computer vision model also not reject but conventional tests reject.
+There are much fewer cases that do not reject by visual tests and conventional tests, but reject by computer vision models.
+This indicates computer vision model can behave like visual tests conducted by human better than conventional tests.
+Moreover, there are great proportion of cases where visual tests conducted by human is the only test who does not reject, .
+
+When plotting the decision against the distance, as illustrated in @fig-power, several observations emerge.
+Firstly, for the computer vision model, when the distance $D > 4$, almost all residual plots are rejected.
+However, for visual tests conducted by humans, the threshold is higher, set at $D > 5$.
+And conventional tests exhibit some non-rejection cases for $D > 3$.
+Moreover, the computer vision model tends to have fewer rejected cases than conventional tests when $D < 2$.
+However, visual tests demonstrate the lowest sensitivity to residual plots with small distances.
+Therefore, tests based on the computer vision model are less sensitive to small deviations from model assumptions than conventional tests but more sensitive to moderate deviations.
+Rejection decisions are fitted by logistic regression models with no intercept terms but an offset equals to $\text{log}(0.05/0/95)$.
+The fitting curves for the computer vision model fall between those of conventional tests and visual tests for both non-linearity and heteroskedasticity.
+There is still potential to refine the computer vision model to better align its behaviour with visual tests.
+
+In the experiment conducted in @li2023plot, participants were allowed to make multiple selections for a lineup.
+The weighted detection rate is computed by assigning weights to each detection.
+If the participant selects zero plots, the corresponding weight is 0.05; otherwise, the weight is 1 divided by the number of selections.
+The $\delta$-difference is defined by @chowdhury2018measuring as
+
+\begin{equation}
+\delta = \hat{D} - \underset{j}{max}\left(\hat{D}_{null}^{(j)}\right) \quad \text{for}~j = 1,...,m-1,
+\end{equation}
+
+\noindent where $\hat{D}_{null}^{(j)}$ is the approximated distance for the $j$-th null plot, and $m$ is the number of plots in a lineup.
+
+@fig-delta displays the scatter plot of the weighted detection rate vs the $\delta$-difference.
+It indicates that the weighted detection rate increases as the $\delta$-difference increases, particularly when the $\delta$-difference is greater than zero.
+A negative $\delta$-difference suggests that there is at least one null plot in the lineup with a stronger visual signal than the true residual plot.
+In some instances, the weighted detection rate is close to one, yet the $\delta$-difference is negative.
+This discrepancy implies that the distance measure, or the estimated distance, may not perfectly reflect actual human behaviour.
+
+```{r hist-null-human, fig.pos =  "!h", fig.cap = "Histogram of predicted distance on null plots used in the human subject experiment. Sample quantiles are drawn in different colors."}
+lineup_vss <- readRDS(here::here("data/lineup_vss.rds"))
+lineup_vss <- lineup_vss %>%
+  group_by(unique_lineup_id) %>%
+  mutate(delta_diff = vss[!null] - max(vss[null])) %>%
+  mutate(gamma_diff = sum(vss[null] > vss[!null]))
+
+experiment <- experiment %>%
+  left_join(lineup_vss %>%
+  filter(null == FALSE) %>%
+  mutate(model_reject_20 = rank == 1) %>%
+  select(unique_lineup_id, rank, model_reject_20))
+```
+
+```{r cache = TRUE}
+vi_lineup <- get_vi_lineup()
+```
+
+::: {.content-visible when-format="html"}
+
+::: {#tbl-human-conv-table .cell tbl-cap='Summary of the comparison of decisions made by computer vision model with decisions made by conventional tests and visual tests conducted by human.'}
+::: {.cell-output-display}
+
+`````{=html}
+<table>
+ <thead>
+  <tr>
+   <th style="text-align:left;"> Violations </th>
+   <th style="text-align:left;"> #Samples </th>
+   <th style="text-align:left;"> #Agreements </th>
+   <th style="text-align:left;"> Agreement rate </th>
+  </tr>
+ </thead>
+<tbody>
+  <tr grouplength="2"><td colspan="4" style="border-bottom: 1px solid;"><strong>Compared with conventional tests</strong></td></tr>
+<tr>
+   <td style="text-align:left;padding-left: 2em;" indentlevel="1"> heteroskedasticity </td>
+   <td style="text-align:left;"> 540 </td>
+   <td style="text-align:left;"> 464 </td>
+   <td style="text-align:left;"> 0.8593 </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;padding-left: 2em;" indentlevel="1"> non-linearity </td>
+   <td style="text-align:left;"> 576 </td>
+   <td style="text-align:left;"> 459 </td>
+   <td style="text-align:left;"> 0.7969 </td>
+  </tr>
+  <tr grouplength="2"><td colspan="4" style="border-bottom: 1px solid;"><strong>Compared with visual tests conducted by human</strong></td></tr>
+<tr>
+   <td style="text-align:left;padding-left: 2em;" indentlevel="1"> heteroskedasticity </td>
+   <td style="text-align:left;"> 540 </td>
+   <td style="text-align:left;"> 367 </td>
+   <td style="text-align:left;"> 0.6796 </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;padding-left: 2em;" indentlevel="1"> non-linearity </td>
+   <td style="text-align:left;"> 576 </td>
+   <td style="text-align:left;"> 385 </td>
+   <td style="text-align:left;"> 0.6684 </td>
+  </tr>
+</tbody>
+</table>
+
+`````
+
+:::
+:::
+
+:::
+
+::: {.content-visible when-format="pdf"}
+
+```{r}
+#| label: tbl-human-conv-table
+#| tbl-cap: "Summary of the comparison of decisions made by computer vision model with decisions made by conventional tests and visual tests conducted by human."
+bind_rows(
+  experiment %>%
+    filter(!attention_check) %>%
+    filter(!null_lineup) %>%
+    mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+    group_by(type) %>%
+    summarise(n = format(n()), agree = format(sum(model_reject == conventional_reject)), rate = format(mean(conventional_reject == model_reject), digits = 4)),
+  experiment %>%
+    filter(!attention_check) %>%
+    filter(!null_lineup) %>%
+    mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+    group_by(type) %>%
+    summarise(n = format(n()), agree = format(sum(model_reject == reject)), rate = format(mean(model_reject == reject), digits = 4))
+  ) %>%
+  kableExtra::kable(booktabs = TRUE,
+                    escape = FALSE,
+                    col.names = c("Violations", "\\#Samples", "\\#Agreements", "Agreement rate")) %>%
+  kableExtra::pack_rows("Compared with conventional tests", 1, 2) %>%
+  kableExtra::pack_rows("Compared with visual tests conducted by human", 3, 4)
+```
+
+:::
+
+```{r fig-conv-mosaic, fig.pos = "!h", fig.cap = "Rejection rate ($p$-value $\\leq0.05$) of computer vision models conditional on conventional tests on non-linearity (left) and heteroskedasticity (right) lineups displayed using a mosaic plot."}
+library(ggmosaic)
+
+experiment %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  mutate(type = ifelse(type == "non-linearity", "Non-linearity", "Heteroskedasticity")) %>%
+  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity"))) %>%
+  mutate(model_reject_20 = ifelse(model_reject_20, "Reject", "No")) %>%
+  mutate(conventional_reject = ifelse(conventional_reject, "Reject", "No")) %>%
+  mutate(across(c(model_reject_20, conventional_reject), ~factor(.x, levels = c("Reject", "No")))) %>%
+  ggplot() +
+  geom_mosaic(aes(x = ggmosaic::product(model_reject_20, conventional_reject), 
+                  fill = model_reject_20)) +
+  facet_grid(~type) +
+  ylab("Computer vision model rejects") +
+  xlab("Conventional tests reject ") +
+  # labs(fill = "Visual tests reject") +
+  scale_fill_manual(values = c("#40B0A6", "#E1BE6A")) + 
+  theme_bw() +
+  theme(legend.position = "none") +
+  coord_fixed()
+```
+
+```{r fig-power, fig.pos = "!h", fig.cap = "Comparison of power of visual tests, conventional tests and the computer vision model.
+Marks along the x-axis at the bottom of the plot represent rejections made by each type of test.
+Marks at the top of the plot represent acceptances.
+Power curves are fitted by logistic regression models with no intercept but an offset equals to $\\text{log}(0.05/0.95)$."}
+
+max_actual_ss <- max(filter(experiment, !attention_check, !null_lineup)$actual_ss)
+max_poly_ss <- max(filter(experiment, !attention_check, !null_lineup, type == "polynomial")$actual_ss)
+max_heter_ss <- max(filter(experiment, !attention_check, !null_lineup, type != "polynomial")$actual_ss)
+
+model_glm_pred_poly <- glm(model_reject_20 ~ actual_ss - 1, 
+    data = filter(experiment, !attention_check, !null_lineup, type == "polynomial") %>% mutate(offset0 = log(0.05/0.95)), 
+    offset = offset0,
+    family = binomial()) %>%
+  predict(data.frame(actual_ss = seq(0, max_poly_ss, 0.01), offset0 = log(0.05/0.95)),
+          type = "response") %>%
+  data.frame(d = seq(0, max_poly_ss, 0.01), type = "Non-linearity", name = "Computer vision model", value = .) %>%
+  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity")))
+
+visual_glm_pred_poly <- glm(reject ~ actual_ss - 1, 
+    data = filter(experiment, !attention_check, !null_lineup, type == "polynomial") %>% mutate(offset0 = log(0.05/0.95)), 
+    offset = offset0,
+    family = binomial()) %>%
+  predict(data.frame(actual_ss = seq(0, max_poly_ss, 0.01), offset0 = log(0.05/0.95)),
+          type = "response") %>%
+  data.frame(d = seq(0, max_poly_ss, 0.01), type = "Non-linearity", name = "Visual test", value = .) %>%
+  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity")))
+
+conv_glm_pred_poly <- glm(conventional_reject ~ actual_ss - 1, 
+    data = filter(experiment, !attention_check, !null_lineup, type == "polynomial") %>% mutate(offset0 = log(0.05/0.95)), 
+    offset = offset0,
+    family = binomial()) %>%
+  predict(data.frame(actual_ss = seq(0, max_poly_ss, 0.01), offset0 = log(0.05/0.95)),
+          type = "response") %>%
+  data.frame(d = seq(0, max_poly_ss, 0.01), type = "Non-linearity", name = "Conventional test", value = .) %>%
+  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity")))
+
+model_glm_pred_heter <- glm(model_reject_20 ~ actual_ss - 1, 
+    data = filter(experiment, !attention_check, !null_lineup, type != "polynomial") %>% mutate(offset0 = log(0.05/0.95)), 
+    offset = offset0,
+    family = binomial()) %>%
+  predict(data.frame(actual_ss = seq(0, max_heter_ss, 0.01), offset0 = log(0.05/0.95)),
+          type = "response") %>%
+  data.frame(d = seq(0, max_heter_ss, 0.01), type = "Heteroskedasticity", name = "Computer vision model", value = .) %>%
+  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity")))
+
+visual_glm_pred_heter <- glm(reject ~ actual_ss - 1, 
+    data = filter(experiment, !attention_check, !null_lineup, type != "polynomial") %>% mutate(offset0 = log(0.05/0.95)), 
+    offset = offset0,
+    family = binomial()) %>%
+  predict(data.frame(actual_ss = seq(0, max_heter_ss, 0.01), offset0 = log(0.05/0.95)),
+          type = "response") %>%
+  data.frame(d = seq(0, max_heter_ss, 0.01), type = "Heteroskedasticity", name = "Visual test", value = .) %>%
+  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity")))
+
+conv_glm_pred_heter <- glm(conventional_reject ~ actual_ss - 1, 
+    data = filter(experiment, !attention_check, !null_lineup, type != "polynomial") %>% mutate(offset0 = log(0.05/0.95)), 
+    offset = offset0,
+    family = binomial()) %>%
+  predict(data.frame(actual_ss = seq(0, max_heter_ss, 0.01), offset0 = log(0.05/0.95)),
+          type = "response") %>%
+  data.frame(d = seq(0, max_heter_ss, 0.01), type = "Heteroskedasticity", name = "Conventional test", value = .) %>%
+  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity")))
+
+mutate(experiment, type = ifelse(type == "polynomial", "Non-linearity", "Heteroskedasticity")) %>%
+  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity"))) %>%
+  filter(!null_lineup, !attention_check) %>%
+ggplot() +
+  geom_line(data = model_glm_pred_poly, aes(d, value, col = "Computer vision model"), linewidth = 0.8) +
+  geom_line(data = model_glm_pred_heter, aes(d, value, col = "Computer vision model"), linewidth = 0.8) +
+  geom_line(data = visual_glm_pred_poly, aes(d, value, col = "Visual test"), linewidth = 0.8) +
+  geom_line(data = visual_glm_pred_heter, aes(d, value, col = "Visual test"), linewidth = 0.8) +
+  geom_line(data = conv_glm_pred_poly, aes(d, value, col = "Conventional test"), linewidth = 0.8) +
+  geom_line(data = conv_glm_pred_heter, aes(d, value, col = "Conventional test"), linewidth = 0.8) +
+  geom_segment(data = ~filter(.x, !reject), aes(x = actual_ss, xend = actual_ss, y = -0.1, yend = -0.1 + 0.03, col = "Visual test"), alpha = 0.4) +
+  geom_segment(data = ~filter(.x, reject), aes(x = actual_ss, xend = actual_ss, y = 1.1 - 0.03 * 2, yend = 1.1 - 0.03 * 3, col = "Visual test"), alpha = 0.4) +
+  geom_segment(data = ~filter(.x, !model_reject_20), aes(x = actual_ss, xend = actual_ss, y = -0.1 + 0.03, yend = -0.1 + 0.03 * 2, col = "Computer vision model"), alpha = 0.4) +
+  geom_segment(data = ~filter(.x, model_reject_20), aes(x = actual_ss, xend = actual_ss, y = 1.1 - 0.03, yend = 1.1 - 0.03 * 2, col = "Computer vision model"), alpha = 0.4) +
+  geom_segment(data = ~filter(.x, !conventional_reject), aes(x = actual_ss, xend = actual_ss, y = -0.1 + 0.03 * 2, yend = -0.1 + 0.03 * 3, col = "Conventional test"), alpha = 0.4) +
+  geom_segment(data = ~filter(.x, conventional_reject), aes(x = actual_ss, xend = actual_ss, y = 1.1, yend = 1.1 - 0.03, col = "Conventional test"), alpha = 0.4) +
+  facet_grid(~type, scales = "free_x") +
+  theme_light() +
+  scale_color_brewer("", palette = "Dark2") +
+  theme(legend.position = "bottom") +
+  xlab("D") +
+  ylab("Reject")
+```
+
+```{r fig-pcp, fig.cap = "Parallel coordinate plots of decisions made by computer vision model, conventional tests and visual tests made by human.", eval = TRUE, fig.height = 4}
+library(ggpcp)
+mutate(experiment, type = ifelse(type == "polynomial", "Non-linearity", "Heteroskedasticity")) %>%
+  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity"))) %>%
+  filter(!null_lineup, !attention_check) %>%
+  select(type, actual_ss, model_reject_20, reject, conventional_reject) %>%
+  mutate(test = rnorm(n())) %>%
+  mutate(across(c(model_reject_20, conventional_reject, reject), ~ifelse(.x, "Reject", "Not reject"))) %>%
+  mutate(agree = model_reject_20 == reject) %>%
+  pcp_select(c(reject, model_reject_20, conventional_reject)) %>%
+  group_by(type) %>%
+  pcp_scale() %>%
+  ungroup() %>%
+  pcp_arrange(method = "from-right") %>%
+  ggplot() +
+  geom_pcp(aes_pcp(), linewidth = 0.1) +
+  geom_pcp_labels(aes_pcp()) +
+  geom_pcp_boxes(aes_pcp(), boxwidth = 0.1) +
+  facet_wrap(~type, scales = "free_y") +
+  theme_light() +
+  scale_x_discrete(labels = c("Visual test", "Computer vision model", "Conventional test")) +
+  labs(col = "D") +
+  ylab("") +
+  xlab("") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
+        axis.text.y = element_blank(), 
+        axis.ticks.y = element_blank(), 
+        axis.title = element_blank())
+```
+
+```{r fig-delta, fig.pos = "!h", fig.cap = "A weighted detection rate vs $\\delta$-differnence plot. The brown line is smoothing curve produced by fitting gnealized additive models", fig.height = 4}
+experiment %>%
+  left_join(lineup_vss %>%
+              group_by(unique_lineup_id) %>%
+              summarise(delta_diff = first(delta_diff), gamma_diff = first(gamma_diff))) %>%
+  ggplot() +
+  geom_vline(xintercept = 0, alpha = 0.4, linetype = 2) +
+  geom_point(aes(delta_diff, prop_detect), alpha = 0.3) +
+  geom_smooth(aes(delta_diff, prop_detect), se = FALSE, col = "#994F00") +
+  ylab("Weighted detection rate") +
+  xlab(expression(delta-difference)) +
+  theme_light()
+```
+
+## Examples
+
+In this section, we will present the performance of trained computer vision model on three example datasets.
+These include the dataset associated with the residual plot displaying a "left-triangle" shape, as displayed in @fig-false-finding, along with the Boston housing dataset, and the "dino" datasets from the `datasaurus` R package.
+
+The first example illustrates a scenario where both the computer vision model and human visual inspection successfully avoid rejecting $H_0$ when $H_0$ is true, contrary to conventional tests.
+This underscores the necessity of visually examining the residual plot.
+
+In the second example, we encounter a more pronounced violation of the model, resulting in rejection of $H_0$ by all three tests.
+This highlights the practicality of the computer vision model, particularly for less intricate tasks.
+
+The third example presents a situation where the model deviation is non-typical.
+Here, the computer vision model and human visual inspection reject $H_0$, whereas some commonly used conventional tests do not.
+This emphasizes the benefits of visual inspection and the unique advantage of the computer vision model, which, like humans, makes decisions based on visual discoveries.
+
+Additionally, we will outline the workflow for employing this model with the `autovi` R package.
+
+### Left-triangle
+
+In @sec-model-introduction, we presented an example residual plot showcased in @fig-false-finding, illustrating how humans might misinterpret the "left-triangle" shape as indicative of heteroskedasticity.
+Additionally, the Breusch-Pagan test yielded a rejection with a $p$-value of 0.046, despite the residuals originating from a correctly specified model.
+@fig-false-finding offers a lineup for this fitted model, showcasing various degrees of "left-triangle" shape across all residual plots.
+This phenomenon is evidently caused by the skewed distribution of the fitted values.
+Notably, if the residual plot were evaluated through a visual test, it would not be rejected since the true residual plot positioned at 10 can not be distinguished from the others.
+
+Moving forward, we will employ the trained computer vision model to assess the residual plot with the assistance of the `autovi` R package.
+This package is accessible on CRAN, and users can install it by executing the command `install.packages("autovi")` in R.
+
+The results of the `list_keras_model()` function provide a list of available trained computer vision models.
+To perform the residual plot assessment, users need to supply a fitted linear regression model and select a trained Keras model to initialize the checker.
+
+```r
+checker <- auto_vi(fitted_model = mod, 
+                   keras_model = get_keras_model("vss_phn_32"))
+```
+
+When invoking the `check()` method to execute the assessment, users can specify the number of null plots and bootstrapped samples using the `null_draws` and `boot_draws` arguments, respectively.
+The diagnostic results can be viewed by directly printing the object.
+Alternatively, a summary plot of the assessment can be generated using the `summary_plot()` method.
+
+```r
+checker$check(null_draws = 200L, boot_draws = 200L)
+checker$summary_plot()
+```
+
+@fig-false-check presents the results of the assessment by the computer vision model.
+Notably, the observed visual signal strength is considerably lower than the 95% sample quantile of the null distribution.
+Moreover, the bootstrapped distribution suggests that it is highly improbable for the fitted model to be misspecified as the majority of bootstrapped fitted models will not be rejected.
+Thus, for this particular fitted model, both the visual test and the computer vision model will not reject $H_0$.
+However, the Breusch-Pagan test will reject $H_0$ because it can not effectively utilize information from null plots.
+
+The attention map at [@fig-false-check]B suggests that the estimation is highly influenced by the top-right and bottom-right part of the residual plot, as it forms two vertices of the triangular shape.
+A principle component analysis is also performed on the output of the global pooling layer of the computer vision model.
+As mentioned in @simonyan2014very, a computer vision model built upon the convolutional blocks can be viewed as a feature extractor.
+For the $32 \times 32$ model, there are 256 features outputted from the global pooling layer, which can be further used for different visual tasks not limited to distance prediction.
+To see if these features can be effectively used for distinguishing null plots and true residual plot, we linearly project them into the PC1 vs PC2 space as shown in @fig-false-check.
+It can be observed that because the bootstrapped plots are mostly similar to the null plots, the points drawn in different colours are mixed together.
+The true residual plot is also covered by both the cluster of null plots and cluster of bootstrapped plots.
+This accurately reflects our understanding of @fig-false-lineup.
+
+```{r}
+set.seed(452)
+ori_x <- rand_lognormal()
+dat <- heter_model(b = 0, x = closed_form(~-ori_x))$gen(300)
+
+mod <- lm(y ~ x, data = dat)
+my_vi <- autovi::auto_vi(fitted_model = mod)
+```
+
+```{r}
+if (!file.exists(here::here("cached_data/fig1_check.rds"))) {
+  keras_model <- autovi::get_keras_model("vss_phn_32")
+  my_vi$check(null_draws = 200L, boot_draws = 200L, keras_model = keras_model, extract_feature_from_layer = "global_max_pooling2d")
+  saveRDS(my_vi$check_result, here::here("cached_data/fig1_check.rds"))
+}
+
+```
+
+```{r}
+my_vi$check_result <- readRDS(here::here("cached_data/fig1_check.rds"))
+```
+
+```{r}
+define_gradient <- function() {
+  reticulate::py_run_string(glue::glue(
+"
+import tensorflow as tf
+import PIL
+import numpy as np
+
+def get_smooth_gradient(keras_mod, plot_path, target_size, input_auxiliary = None, noise_level = 0.1, n = 50):
+    im = PIL.Image.open(plot_path)
+    im = im.resize(target_size)
+    input_im = tf.keras.utils.img_to_array(im)
+    input_im = np.reshape(input_im, tuple([1]) + tuple(target_size) + tuple([3]))
+    if input_auxiliary is not None:
+        input_auxiliary = np.reshape(np.array(input_auxiliary), (1, 5))
+
+    if input_auxiliary is not None:
+        input_auxiliary = tf.Variable(input_auxiliary)
+
+    im_grad = None
+    sigma = noise_level * (np.max(input_im) - np.min(input_im))
+
+    for i in range(n):    
+        new_input_im = input_im + np.random.normal(0, sigma, input_im.shape)
+        new_input_im = tf.Variable(new_input_im)
+
+        with tf.GradientTape() as tape:
+            if input_auxiliary is not None:
+                pred = keras_mod([new_input_im, input_auxiliary])
+            else:
+                pred = keras_mod(new_input_im)
+        new_im_grad = tape.gradient(pred, new_input_im)
+        new_im_grad = tf.image.rgb_to_grayscale(new_im_grad)[0].numpy()
+        if im_grad is None:
+            im_grad = new_im_grad
+        else:
+            im_grad = im_grad + new_im_grad
+
+    im_grad = im_grad / n
+
+    input_im = tf.Variable(input_im)
+    if input_auxiliary is not None:
+        with tf.GradientTape() as tape:
+            pred = keras_mod([input_im, input_auxiliary])
+        auxiliary_grad = tape.gradient(pred, input_auxiliary)
+    if input_auxiliary is not None:
+        return (im_grad, auxiliary_grad[0].numpy())
+    else:
+        return im_grad
+
+def get_raw_gradient(keras_mod, plot_path, target_size, input_auxiliary = None):
+    im = PIL.Image.open(plot_path)
+    im = im.resize(target_size)
+    input_im = tf.keras.utils.img_to_array(im)
+    input_im = np.reshape(input_im, tuple([1]) + tuple(target_size) + tuple([3]))
+    if input_auxiliary is not None:
+        input_auxiliary = np.reshape(np.array(input_auxiliary), (1, 5))
+
+    input_im = tf.Variable(input_im)
+    if input_auxiliary is not None:
+        input_auxiliary = tf.Variable(input_auxiliary)
+
+    with tf.GradientTape() as tape:
+        if input_auxiliary is not None:
+            pred = keras_mod([input_im, input_auxiliary])
+        else:
+            pred = keras_mod(input_im)
+    im_grad = tape.gradient(pred, input_im)
+
+    if input_auxiliary is not None:
+        with tf.GradientTape() as tape:
+            pred = keras_mod([input_im, input_auxiliary])
+        auxiliary_grad = tape.gradient(pred, input_auxiliary)
+
+    im_grad = tf.image.rgb_to_grayscale(im_grad)
+    if input_auxiliary is not None:
+        return (im_grad[0].numpy(), auxiliary_grad[0].numpy())
+    else:
+        return im_grad[0].numpy()
+"
+))
+}
+
+```
+
+```{r}
+if (!file.exists(here::here("cached_data/fig1_attention_map.rds"))) {
+  reticulate::py_set_seed(10086, disable_hash_randomization = TRUE)
+  define_gradient()
+  keras_mod <- autovi::get_keras_model("vss_phn_32")
+  plot_path <- my_vi$plot_resid() %>%
+    autovi::save_plot()
+  input_auxiliary <- my_vi$auxiliary()
+
+  reticulate::py$get_smooth_gradient(keras_mod = keras_mod, 
+                                     plot_path = plot_path, 
+                                     target_size = c(32L, 32L), 
+                                     input_auxiliary = input_auxiliary,
+                                     noise_level = 0.5,
+                                     n = 1000L) -> smooth_g
+
+  x_range <- ggplot2::ggplot_build(my_vi$plot_resid())$layout$panel_params[[1]]$x.range
+  y_range <- ggplot2::ggplot_build(my_vi$plot_resid())$layout$panel_params[[1]]$y.range
+
+  smooth_att_map <- smooth_g[[1]] |> 
+    as.data.frame() |>
+    dplyr::mutate(row = rev(1:32)) |>
+    tidyr::pivot_longer(V1:V32, names_to = "column", values_to = "gradient") |>
+    dplyr::mutate(column = as.integer(gsub("V", "", column))) |>
+    dplyr::mutate(column = x_range[1] + column * (x_range[2] - x_range[1])/(32 - 1)) |>
+    dplyr::mutate(row = y_range[1] + row * (y_range[2] - y_range[1])/(32 - 1))
+
+  scale_zero_one <- function(x) (x - min(x))/(max(x) - min(x))
+
+  p <- ggplot() +
+    ggplot2::geom_raster(data = mutate(smooth_att_map, gradient = scale_zero_one(gradient)), ggplot2::aes(column, row, fill = gradient), alpha = 0.7) +
+    scale_fill_gradient(low = "black", high = "white") +
+    ggplot2::theme_void() +
+    theme(legend.position = "none")
+
+  saveRDS(p, here::here("cached_data/fig1_attention_map.rds"))
+
+}
+```
+
+```{r fig-false-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment evaluted on 200 null plots and 200 bootstrapped plots.
+(A) The true residual plot exhibiting a "left-triangle" shape.
+(B) The attention map produced by computing the gradient of the output with respect to the greyscale input.
+(C) The density plot of estimated distnace for null plots and bootstrapped plots.
+The blue area indicates the distribution of estimated distances for bootstrapped plots, while the yellow area represents the distribution of estimated distances for null plots.
+The fitted model will not be rejected since $\\hat{D} < Q_{null}(0.95)$.
+(D) PC2 vs PC1 plot for features extracted from the global pooling layer of the computer vision model.
+', fig.height = 8}
+p1 <- my_vi$plot_resid(theme = theme_light()) + ggtitle("(A) Residual plot") 
+p2 <- readRDS(here::here("cached_data/fig1_attention_map.rds")) + ggtitle("(B) Attention map")
+p3 <- my_vi$summary_plot() +
+  ggtitle("(C) Density plot of estimated D", subtitle = glue::glue("Conventional p-value = {format(HETER_MODEL$test(dat)$p_value, digits = 3)}")) +
+  theme(legend.position = "bottom", legend.box = "vertical") +
+  labs(linetype = "") +
+  xlab(expression(hat(D))) +
+  scale_linetype(labels = c("95% quantile of the null distribution", expression(hat(D)))) +
+  scale_fill_manual(values = c("#40B0A6", "#E1BE6A")) +
+  scale_color_manual(values = c("#40B0A6", "#E1BE6A"))
+p4 <- my_vi$feature_pca() %>%
+  mutate(set = ifelse(set == "boot", "Boot", set)) %>%
+  mutate(set = ifelse(set == "null", "Null", set)) %>%
+  ggplot() +
+  geom_point(data = ~filter(.x, set != "observed"), aes(PC1, PC2, col = set), size = 2, alpha = 0.3) +
+  geom_point(data = ~filter(.x, set == "observed"), aes(PC1, PC2), size = 4, col = "black", fill = "black") +
+  geom_text(data = ~filter(.x, set == "observed"), aes(PC1, PC2, label = "Observed"), nudge_x = 10) +
+  theme_light() +
+  scale_color_brewer(palette = "Dark2") +
+  xlab("PC1") +
+  ylab("PC2") +
+  labs(col = "") +
+  theme(legend.position = "bottom") +
+  ggtitle("(D) PCA for extracted features") +
+  scale_color_manual(values = c("#40B0A6", "#E1BE6A"))
+
+patchwork::wrap_plots(p1, p2, p3, p4, nrow = 2, widths = c(1, 1), heights = c(1, 1))
+```
+
+```{r fig-false-lineup, fig.pos = "!h", fig.cap = 'A lineup of residual plots displaying "left-triangle" visual patterns.
+The true residual plot occupies position 10, yet there are no discernible visual patterns that distinguish it from the other plots.', fig.height = 8}
+pos <- 10
+result <- tibble()
+for (i in c(1:20)[-pos]) {
+  result <- bind_rows(result,
+                      my_vi$rotate_resid() %>%
+                        mutate(k = i))
+}
+result <- bind_rows(my_vi$get_fitted_and_resid() %>% 
+                      mutate(k = pos),
+                    result)
+result %>%
+  VI_MODEL$plot_lineup(remove_grid_line = TRUE,
+                       remove_axis = TRUE,
+                       theme = theme_light())
+```
+
+### Boston housing
+
+The Boston housing dataset, originally published by @harrison1978hedonic, offers insights into housing in the Boston, Massachusetts area.
+For illustration purposes, we will utilize a reduced version from Kaggle, comprising 489 rows and 4 columns: average number of rooms per dwelling (RM), percentage of lower status of the population (LSTAT), pupil-teacher ratio by town (PTRATIO), and Median value of owner-occupied homes in $1000's (MEDV).
+In our analysis, MEDV will serve as the response variable, while the other columns will function as predictors in a linear regression model.
+Our primary focus is to detect non-linearity, because the relationships between RM and MEDV or LSTAT and MEDV are non-linear.
+
+@fig-boston-check displays the residual plot and the assessment conducted by the computer vision model.
+A clear non-linearity pattern resembling a "U" shape is shown in the plot A.
+Furthermore, the RESET test yields a very small $p$-value.
+The estimated distance $\hat{D}$ significantly exceeds $Q_{null}(0.95)$, leading to rejection of $H_0$.
+The bootstrapped distribution also suggests that almost all the bootstrapped fitted models will be rejected, indicating that the fitted model is unlikely to be correctly specified.
+The attention map in plot B suggests the center of the image has higher leverage than other areas, and it is the turning point of the "U" shape.
+The principal component analysis provided in plot D shows two distinct clusters of data points, further underlines the visual differences between bootstrapped plots and null plots.
+This coincides the findings from @fig-boston-lineup, where the true plot exhibiting a "U" shape is visually distinctive from null plots.
+If a visual test is conducted by human, $H_0$ will also be rejected.
+
+```{r}
+housing <- read_csv(here::here("data/housing.csv"))
+mod <- lm(MEDV ~ ., data = housing)
+
+my_vi <- autovi::auto_vi(fitted_model = mod)
+```
+
+```{r}
+if (!file.exists(here::here("cached_data/boston_check.rds"))) {
+  keras_model <- autovi::get_keras_model("vss_phn_32")
+  my_vi$check(null_draws = 200L, boot_draws = 200L, keras_model = keras_model, extract_feature_from_layer = "global_max_pooling2d")
+  saveRDS(my_vi$check_result, here::here("cached_data/boston_check.rds"))
+}
+
+```
+
+```{r}
+my_vi$check_result <- readRDS(here::here("cached_data/boston_check.rds"))
+```
+
+```{r}
+if (!file.exists(here::here("cached_data/boston_attention_map.rds"))) {
+  reticulate::py_set_seed(10086, disable_hash_randomization = TRUE)
+  define_gradient()
+  keras_mod <- autovi::get_keras_model("vss_phn_32")
+  plot_path <- my_vi$plot_resid() %>%
+    autovi::save_plot()
+  input_auxiliary <- my_vi$auxiliary()
+
+  reticulate::py$get_smooth_gradient(keras_mod = keras_mod, 
+                                     plot_path = plot_path, 
+                                     target_size = c(32L, 32L), 
+                                     input_auxiliary = input_auxiliary,
+                                     noise_level = 0.5,
+                                     n = 1000L) -> smooth_g
+
+  x_range <- ggplot2::ggplot_build(my_vi$plot_resid())$layout$panel_params[[1]]$x.range
+  y_range <- ggplot2::ggplot_build(my_vi$plot_resid())$layout$panel_params[[1]]$y.range
+
+  smooth_att_map <- smooth_g[[1]] |> 
+    as.data.frame() |>
+    dplyr::mutate(row = rev(1:32)) |>
+    tidyr::pivot_longer(V1:V32, names_to = "column", values_to = "gradient") |>
+    dplyr::mutate(column = as.integer(gsub("V", "", column))) |>
+    dplyr::mutate(column = x_range[1] + column * (x_range[2] - x_range[1])/(32 - 1)) |>
+    dplyr::mutate(row = y_range[1] + row * (y_range[2] - y_range[1])/(32 - 1))
+
+  scale_zero_one <- function(x) (x - min(x))/(max(x) - min(x))
+
+  p <- ggplot() +
+    ggplot2::geom_raster(data = mutate(smooth_att_map, gradient = scale_zero_one(gradient)), ggplot2::aes(column, row, fill = gradient), alpha = 0.7) +
+    scale_fill_gradient(low = "black", high = "white") +
+    ggplot2::theme_void() +
+    theme(legend.position = "none")
+
+  saveRDS(p, here::here("cached_data/boston_attention_map.rds"))
+
+}
+```
+
+```{r fig-boston-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment for the Boston housing fitted model evaluted on 200 null plots and 200 bootstrapped plots.
+(A) The true residual plot exhibiting a "U" shape.
+(B) The attention map produced by computing the gradient of the output with respect to the greyscale input.
+(C) The density plot of estimated distnace for null plots and bootstrapped plots.
+The blue area indicates the distribution of estimated distances for bootstrapped plots, while the yellow area represents the distribution of estimated distances for null plots.
+The fitted model will be rejected since $\\hat{D} \\geq Q_{null}(0.95)$.
+(D) PC2 vs PC1 plot for features extracted from the global pooling layer of the computer vision model.
+', fig.height = 8}
+p1 <- my_vi$plot_resid(theme = theme_light()) + ggtitle("(A) Residual plot") 
+p2 <- readRDS(here::here("cached_data/boston_attention_map.rds")) + ggtitle("(B) Attention map")
+p3 <- my_vi$summary_plot() +
+  ggtitle("(C) Density plot of estimated D", subtitle = glue::glue("Conventional p-value = {format(lmtest::resettest(mod)$p.value, digits = 3)}")) +
+  theme(legend.position = "bottom", legend.box = "vertical") +
+  labs(linetype = "") +
+  xlab(expression(hat(D))) +
+  scale_linetype(labels = c("95% quantile of the null distribution", expression(hat(D)))) +
+  scale_fill_manual(values = c("#40B0A6", "#E1BE6A")) +
+  scale_color_manual(values = c("#40B0A6", "#E1BE6A"))
+p4 <- my_vi$feature_pca() %>%
+  mutate(set = ifelse(set == "boot", "Boot", set)) %>%
+  mutate(set = ifelse(set == "null", "Null", set)) %>%
+  ggplot() +
+  geom_point(data = ~filter(.x, set != "observed"), aes(PC1, PC2, col = set), size = 2, alpha = 0.3) +
+  geom_point(data = ~filter(.x, set == "observed"), aes(PC1, PC2), size = 4, col = "black", fill = "black") +
+  geom_text(data = ~filter(.x, set == "observed"), aes(PC1, PC2, label = "Observed"), nudge_x = 7) +
+  theme_light() +
+  scale_color_brewer(palette = "Dark2") +
+  xlab("PC1") +
+  ylab("PC2") +
+  labs(col = "") +
+  theme(legend.position = "bottom") +
+  ggtitle("(D) PCA for extracted features") +
+  scale_color_manual(values = c("#40B0A6", "#E1BE6A"))
+
+patchwork::wrap_plots(p1, p2, p3, p4, nrow = 2, widths = c(1, 1), heights = c(1, 1))
+```
+
+```{r fig-boston-lineup, fig.pos = "!h", fig.cap = 'A lineup of residual plots for the Boston housing fitted model.
+The true residual plot is at position 7.
+It can be easily identified as the most different plot.', fig.height = 8}
+pos <- 7
+result <- tibble()
+for (i in c(1:20)[-pos]) {
+  result <- bind_rows(result,
+                      my_vi$rotate_resid() %>%
+                        mutate(k = i))
+}
+result <- bind_rows(my_vi$get_fitted_and_resid() %>% 
+                      mutate(k = pos),
+                    result)
+result %>%
+  VI_MODEL$plot_lineup(remove_grid_line = TRUE,
+                       remove_axis = TRUE,
+                       theme = theme_light())
+```
+
+### Datasaurus
+
+The computer vision model possesses the capability to detect not only typical issues like non-linearity, heteroskedasticity, and non-normality but also artifact visual patterns resembling real-world objects, as long as they do not appear in null plots.
+These visual patterns can be challenging to categorize in terms of model violations.
+Therefore, we will employ the RESET test, the Breusch-Pagan test, and the Shapiro-Wilk test [@shapiro1965analysis] for comparison.
+
+The "dino" dataset within the `datasaurus` R package exemplifies this scenario.
+With only two columns, x and y, fitting a regression model to this data yields a residual plot resembling a "dinosaur," as displayed in [@fig-dino-check]A.
+Unsurprisingly, this distinct residual plot stands out in a lineup, as shown in @fig-dino-lineup.
+A visual test conducted by humans would undoubtedly reject $H_0$.
+
+According to the residual plot assessment by the computer vision model, $\hat{D}$ exceeds $Q_{null}(0.95)$, warranting a rejection of $H_0$.
+Additionally, most of the bootstrapped fitted models will be rejected, indicating an misspecified model.
+However, both the RESET test and the Breusch-Pagan test yield $p$-values greater than 0.3, leading to a non-rejection of $H_0$.
+Only the Shapiro-Wilk test rejects the normality assumption with a small $p$-value.
+
+More importantly, the attention map in [@fig-dino-check]B clearly exhibits a "dinosaur" shape, strongly suggesting the prediction of the distance is based on human perceptible visual patterns.
+The computer vision model is also capable of capturing the contour or the outline of the embedded shape, just like human being reading residual plots.
+The principle component analysis in [@fig-dino-check]D also shows that the cluster of bootstrapped plots is at the corner of the cluster of null plots.
+
+More importantly, the attention map in [@fig-dino-check]B clearly exhibits a "dinosaur" shape, strongly suggesting that the distance prediction is based on human-perceptible visual patterns.
+The computer vision model effectively captures the contour or outline of the embedded shape, similar to how humans interpret residual plots.
+Additionally, the principal component analysis in [@fig-dino-check]D demonstrates that the cluster of bootstrapped plots is positioned at the corner of the cluster of null plots.
+
+In practice, without accessing the residual plot, it would be challenging to identify the artificial pattern of the residuals.
+Moreover, conducting a normality test for a fitted regression model is not always standard practice among analysts.
+Even when performed, violating the normality assumption is sometimes deemed acceptable, especially considering the application of quasi-maximum likelihood estimation in linear regression.
+This example underscores the importance of evaluating residual plots and highlights how the proposed computer vision model can facilitate this process.
+
+```{r}
+dino <- datasauRus::datasaurus_dozen %>% filter(dataset == "dino")
+mod <- lm(y ~ ., data = select(dino, -dataset))
+
+my_vi <- autovi::auto_vi(fitted_model = mod)
+```
+
+```{r}
+if (!file.exists(here::here("cached_data/dino_check.rds"))) {
+  keras_model <- autovi::get_keras_model("vss_phn_32")
+  my_vi$check(null_draws = 200L, boot_draws = 200L, keras_model = keras_model, extract_feature_from_layer = "global_max_pooling2d")
+  saveRDS(my_vi$check_result, here::here("cached_data/dino_check.rds"))
+}
+```
+
+```{r}
+my_vi$check_result <- readRDS(here::here("cached_data/dino_check.rds"))
+```
+
+```{r}
+if (!file.exists(here::here("cached_data/dino_attention_map.rds"))) {
+  reticulate::py_set_seed(10086, disable_hash_randomization = TRUE)
+  define_gradient()
+  keras_mod <- autovi::get_keras_model("vss_phn_32")
+  plot_path <- my_vi$plot_resid() %>%
+    autovi::save_plot()
+  input_auxiliary <- my_vi$auxiliary()
+
+  reticulate::py$get_smooth_gradient(keras_mod = keras_mod, 
+                                     plot_path = plot_path, 
+                                     target_size = c(32L, 32L), 
+                                     input_auxiliary = input_auxiliary,
+                                     noise_level = 0.5,
+                                     n = 1000L) -> smooth_g
+
+  x_range <- ggplot2::ggplot_build(my_vi$plot_resid())$layout$panel_params[[1]]$x.range
+  y_range <- ggplot2::ggplot_build(my_vi$plot_resid())$layout$panel_params[[1]]$y.range
+
+  smooth_att_map <- smooth_g[[1]] |> 
+    as.data.frame() |>
+    dplyr::mutate(row = rev(1:32)) |>
+    tidyr::pivot_longer(V1:V32, names_to = "column", values_to = "gradient") |>
+    dplyr::mutate(column = as.integer(gsub("V", "", column))) |>
+    dplyr::mutate(column = x_range[1] + column * (x_range[2] - x_range[1])/(32 - 1)) |>
+    dplyr::mutate(row = y_range[1] + row * (y_range[2] - y_range[1])/(32 - 1))
+
+  scale_zero_one <- function(x) (x - min(x))/(max(x) - min(x))
+
+  p <- ggplot() +
+    ggplot2::geom_raster(data = mutate(smooth_att_map, gradient = scale_zero_one(gradient)), ggplot2::aes(column, row, fill = gradient), alpha = 0.7) +
+    scale_fill_gradient(low = "black", high = "white") +
+    ggplot2::theme_void() +
+    theme(legend.position = "none")
+
+  saveRDS(p, here::here("cached_data/dino_attention_map.rds"))
+
+}
+```
+
+```{r fig-dino-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment for the Datasaurus fitted model evaluted on 200 null plots and 200 bootstrapped plots.
+(A) The true residual plot exhibiting a "dinosaur" shape.
+(B) The attention map produced by computing the gradient of the output with respect to the greyscale input.
+(C) The density plot of estimated distnace for null plots and bootstrapped plots.
+The blue area indicates the distribution of estimated distances for bootstrapped plots, while the yellow area represents the distribution of estimated distances for null plots.
+The fitted model will be rejected since $\\hat{D} \\geq Q_{null}(0.95)$.
+(D) PC2 vs PC1 plot for features extracted from the global pooling layer of the computer vision model.', fig.height = 8}
+
+p1 <- my_vi$plot_resid(theme = theme_light()) + ggtitle("(A) Residual plot")
+p2 <- readRDS(here::here("cached_data/dino_attention_map.rds")) + ggtitle("(B) Attention map")
+p3 <- my_vi$summary_plot() +
+  ggtitle("(C) Density plot of estimated D", subtitle = glue::glue("RESET test p-value = {format(lmtest::resettest(mod)$p.value, digits = 3)}\n Breusch-Pagan test p-value = {format(lmtest::bptest(mod)$p.value, digits = 3)} \n Shapiro-Wilk test p-value = {format(shapiro.test(mod$residuals)$p.value, digits = 3)}")) +
+  theme(legend.position = "bottom", legend.box = "vertical") +
+  labs(linetype = "") +
+  xlab(expression(hat(D))) +
+  scale_linetype(labels = c("95% quantile of the null distribution", expression(hat(D)))) +
+  scale_fill_manual(values = c("#40B0A6", "#E1BE6A")) +
+  scale_color_manual(values = c("#40B0A6", "#E1BE6A"))
+p4 <- my_vi$feature_pca() %>%
+  mutate(set = ifelse(set == "boot", "Boot", set)) %>%
+  mutate(set = ifelse(set == "null", "Null", set)) %>%
+  ggplot() +
+  geom_point(data = ~filter(.x, set != "observed"), aes(PC1, PC2, col = set), size = 2, alpha = 0.3) +
+  geom_point(data = ~filter(.x, set == "observed"), aes(PC1, PC2), size = 4, col = "black", fill = "black") +
+  geom_text(data = ~filter(.x, set == "observed"), aes(PC1, PC2, label = "Observed"), nudge_x = -10) +
+  theme_light() +
+  scale_color_brewer(palette = "Dark2") +
+  xlab("PC1") +
+  ylab("PC2") +
+  labs(col = "", shape = "") +
+  theme(legend.position = "bottom", legend.box="vertical") +
+  ggtitle("(D) PCA for extracted features") +
+  scale_color_manual(values = c("#40B0A6", "#E1BE6A"))
+
+patchwork::wrap_plots(p1, p2, p3, p4, nrow = 2, widths = c(1, 1), heights = c(1, 1))
+```
+
+```{r fig-dino-lineup, fig.pos = "!h", fig.cap = 'A lineup of residual plots for the fitted model on the "dinosaur" dataset.
+The true residual plot is at position 17.
+It can be easily identified as the most different plot as the visual pattern is extremly artificial.', fig.height = 8}
+pos <- 17
+result <- tibble()
+for (i in c(1:20)[-pos]) {
+  result <- bind_rows(result,
+                      my_vi$rotate_resid() %>%
+                        mutate(k = i))
+}
+result <- bind_rows(my_vi$get_fitted_and_resid() %>% 
+                      mutate(k = pos),
+                    result)
+result %>%
+  VI_MODEL$plot_lineup(remove_grid_line = TRUE,
+                       remove_axis = TRUE,
+                       theme = theme_light())
+```
+
+# Limitations and Future Work
+
+Despite the computer vision model performing well with general cases under the synthetic data generation scheme and the three examples used in this paper, this study has several limitations that could guide future work.
+
+The proposed distance measure assumes that the true model is a classical normal linear regression model, which can be too restrictive.
+Although this paper does not address the relaxation of this assumption, there are potential methods to evaluate other types of regression models.
+The most comprehensive approach would be to define a distance measure for each different class of regression model and then train the computer vision model following the methodology described in this paper.
+To accelerate training, one could use the convolutional blocks of our trained model as a feature extractor and perform transfer learning on top of it, as these blocks effectively capture shapes in residual plots.
+Another approach would be to transform the residuals so they are roughly normally distributed and have constant variance.
+If only raw residuals are used, the distance-based statistical testing compares the difference in distance to a classical normal linear regression model for the true plot and null plots.
+This comparison is meaningful only if the difference can be identified by the distance measure proposed in this paper.
+
+There are other types of residual plots commonly used in diagnostics, such as residuals vs.
+predictor and quantile-quantile plots.
+In this study, we focused on the most commonly used residual plot as a starting point for exploring the new field of automated visual inference.
+Similarly, we did not explore other, more sophisticated computer vision model architectures and specifications for the same reason.
+While the performance of the computer vision model is acceptable, there is still room for improvement to achieve behavior more closely resembling that of humans interpreting residual plots.
+This may require external survey data or human subject experiment data to understand the fundamental differences between our implementation and human evaluation.
+
+# Conclusion
+
+In this study, we have introduced a distance measure based on Kullback-Leibler divergence to quantify the disparity between the residual distribution of a fitted classical normal linear regression model and the reference residual distribution assumed under correct model specification.
+This distance measure effectively captures the magnitude of model violations in misspecified models.
+We propose a computer vision model to estimate this distance, utilizing the residual plot of the fitted model as input.
+The resulting estimated distance serves as the foundation for constructing a single Model Violation Index (MVI), facilitating the quantification of various model violations.
+
+Moreover, the estimated distance enables the development of a formal statistical testing procedure by evaluating a large number of null plots generated from the fitted model.
+Additionally, employing bootstrapping techniques and refitting the regression model allows us to ascertain how frequently the fitted model is considered misspecified if data were repeatedly obtained from the same data generating process.
+
+The trained computer vision model demonstrates strong performance on both the training and test sets, although it exhibits slightly lower performance on residual plots with non-linearity visual patterns compared to other types of violations.
+The statistical tests relying on the estimated distance predicted by the computer vision model exhibit lower sensitivity compared to conventional tests but higher sensitivity compared to visual tests conducted by humans.
+While the estimated distance generally mirrors the strength of the visual signal perceived by humans, there remains scope for further improvement in its performance.
+
+Several examples are provided to showcase the effectiveness of the proposed method across different scenarios, emphasizing the similarity between visual tests and distance-based tests.
+Overall, both visual tests and distance-based tests can be viewed as combined tests, aiming to assess the correctness of all model assumptions collectively.
+In contrast, individual residual diagnostic tests such as the RESET test and the Breusch-Pagan test only evaluate specific model assumptions.
+In practice, selecting an appropriate set of statistical tests for regression diagnostics can be challenging, particularly given the necessity of adjusting the significance level for each test.
+
+Our method holds significant value as it helps alleviate a portion of analysts' workload associated with assessing residual plots.
+While we recommend analysts to continue reading residual plots whenever feasible, as they offer invaluable insights, our approach serves as a valuable tool for automating the diagnostic process or for supplementary purposes when needed.

### 2024-07-01 Thesis/A-appA.qmd
update thesis folder

@@
 
-::: {.content-visible when-format="html"}
-
-::: {#fig-tech}
-
-<embed src="figures/experiment_tech.pdf" class="img-fluid" style="width:0%; display:none"/>
-
-![](figures/experiment_tech.pdf#toolbar=0){width=100.0% height=953px}
-
-Diagram of online experimental setup. The server-side of the study website uses Flask as backend hosted on PythonAnywhere. And the client-side uses jsPsych to run experiment.
-:::
-
-:::
-
-::: {.content-visible when-format="pdf"}
-
 ```{r fig-tech, out.width = "100%", fig.cap = "Diagram of online experimental setup.
@@
 And the client-side uses jsPsych to run experiment."}
-knitr::include_graphics("figures/experiment_tech.pdf")
-```
-
-:::
+magick::image_read_pdf("figures/experiment_tech.pdf")
+```
 

### 2024-07-28 Thesis/03-chap3.qmd
some sections should be second level headers instead of first level
update according to emi edited

@@
 
-Plotting residuals is commonly regarded as a standard practice in linear regression diagnostics [see @belsley1980regression; @cook1982residuals].
+Plotting residuals is commonly regarded as a standard practice in linear regression diagnostics [@belsley1980regression; @cook1982residuals].
 This visual assessment plays a crucial role in identifying deviations from model assumptions, such as linearity, homoscedasticity, and normality.
-It also helps in understanding the goodness of fit and various characteristics of the model.
+It also helps in understanding the goodness of fit and various unexpected characteristics of the model.
 
@@
 However, accurately interpreting a residual plot can be challenging.
-A residual plot can exhibit various visual features, but it is crucial to recognize that some may arise from the characteristics of predictors and the inherent randomness of the error, rather than indicating a violation of model assumptions [@li2023plot].
-Consider @fig-false-finding as an example, the residuals display a triangular shape pointing to the left.
-While this might suggest heteroskedasticity, it is important to avoid over-interpreting this visual pattern.
+A residual plot can exhibit various visual features, but it is crucial to recognize that some may arise from the characteristics of predictors and the natural stochastic variation of the observational unit, rather than indicating a violation of model assumptions [@li2023plot].
+Consider @fig-false-finding as an example, the residual plot displays a triangular left-pointing shape.
+The distinct difference in the spread of the residuals across the fitted values may result in the analyst suggesting that there may be heteroskedasticity, however, it is important to avoid over-interpreting this visual pattern.
 In this case, the fitted regression model is correctly specified, and the triangular shape is actually a result of the skewed distribution of the predictors, rather than indicating a flaw in the model.
@@
 
-However, as pointed out by @li2023plot, a primary limitation of the lineup protocol lies in its reliance on human judgments.
+A primary limitation of the lineup protocol lies in its reliance on human judgements [see @li2023plot about the practical limitations].
 Unlike conventional statistical tests that can be performed computationally in statistical software, the lineup protocol requires human evaluation of images.
 This characteristic makes it less suitable for large-scale applications, given the associated high labour costs and time requirements.
-
-There is a substantial need to develop an approach that alleviates analysts' workload by automating repetitive tasks and providing standardized results in a controlled environment.
-The large-scale evaluation of lineups is impractical without the use of technology and machines.
-
-The utilization of computers to interpret data plots has a rich history, with early efforts such as "Scagnostics" by @tukey1985computer, focusing on scatter plot diagnostics.
-@wilkinson2005graph expanded on this work, introducing graph theoretic scagnostics, which encompassed computable measures applied to planar proximity graphs.
+There is a substantial need to develop an approach to substitute these human judgement with an automated reading of data plots using machines.
+
+The utilization of computers to interpret data plots has a rich history, with early efforts such as "Scagnostics" by @tukey1985computer, a set of numerical statistics that summarise features of scatter plots.
+@wilkinson2005graph expanded on this work, introducing scagnostics based on computable measures applied to planar proximity graphs.
 These measures, including, but not limited to, "Outlying", "Skinny", "Stringy", "Straight", "Monotonic", "Skewed", "Clumpy", and "Striated" aimed to characterize outliers, shape, density, trend, coherence and other characteristics of the data.
@@
 Despite the widespread use of computer vision models in fields like computer-aided diagnosis [@lee2015image], pedestrian detection [@brunetti2018computer], and facial recognition [@emami2012facial], their application in reading data plots remains limited.
-While some studies have explored the use of computer vision models for tasks such as reading recurrence plots for time series regression [@ojeda2020multivariate], time series classification [@chu2019automatic; @hailesilassie2019financial; @hatami2018classification; @zhang2020encoding], anomaly detection [@chen2020convolutional], and pairwise causality analysis [@singh2017deep], the application of reading residual plots with computer vision models represents a relatively new field of study.
-
-In this paper, we develop computer vision models and integrate them into the residual plots diagnostics workflow, addressing the need for automated visual inference.
+While some studies have explored the use of computer vision models for tasks such as reading recurrence plots for time series regression [@ojeda2020multivariate], time series classification [@chu2019automatic; @hailesilassie2019financial; @hatami2018classification; @zhang2020encoding], anomaly detection [@chen2020convolutional], and pairwise causality analysis [@singh2017deep], the application of reading residual plots with computer vision models is a new field of study.
+
+In this paper, we develop computer vision models and integrate them into the residual plots diagnostics workflow, addressing the need for an automated visual inference.
 The paper is structured as follows.
@@
 The quality and relevance of the input data greatly influence the model's capacity to generate insightful and meaningful results.
-There are several designs of the input format can be considered.
+There are several designs of the input format that can be considered.
 
@@
 One solution is to pad the input vectors with leading or trailing zeros when the input tensor expects longer vectors, but it may fail if the input vector surpasses the designed length.
+
 Another strategy is to summarize the residuals and fitted values separately using histograms and utilize the counts as the input.
@@
 The main considerations are the image resolution and the aesthetics of the residual plot.
-In general, higher resolution provides more information to the model but comes with the trade-off of increased complexity and greater difficulty in training.
+In general, a higher resolution provides more information to the model but comes with a trade-off of increased complexity and greater difficulty in training.
 As for the aesthetics of the residual plot, a practical solution is to consistently present residual plots in the same style to the model.
@@
 
-We did not explore all the mentioned input formats due to the significant costs tied to data preparation and model training.
-Taking into account the implementation cost and the need for model interpretability, we settled on the single residual plot input format.
+Taking into account the implementation cost and the need for model interpretability, we used the single residual plot input format in this paper.
 
@@
 
-Given that the input is a single residual plot represented as a fixed-resolution image, the output from the computer vision model can take one of two forms: binary or numeric.
-This choice determines whether the model belongs to a classification model or a regression model.
-The binary outcome encoded as $0$ and $1$ could be used to represent whether the input image is a null plot, or whether the input image would be rejected in a visual test conducted by humans.
+Given that the input is a single residual plot represented as a fixed-resolution image, we can choose the output from the computer vision model to be either binary (classification) or numeric (regression).
+
+The binary outcome can represent whether the input image is consistent with a null plot as determined by either (1) the data generating process or (2) the result of a visual test based on human judgement.
 Training a model following the latter option requires data from prior human subject experiments, presenting difficulties in controlling the quality of data due to variations in experimental settings across different studies.
@@
 
-In this study, we chose to define and use a distance measure to quantify the difference between the residual plot and the null plots.
+In this study, we chose to define and use a distance between a true residual plot and a theoretical "good" residual plot.
 @vo2016localizing have also demonstrated that defining a proper distance between images can enhance the matching accuracy in image search compared to a binary outcome model.
 
-## Distance Between Residual Plots {#sec-model-distance-between-residual-plots}
-
-In a visual test, the observer will be asked to choose one or more plots that stand out as most distinct from others in a given lineup.
+## Distance from a Theoretical "Good" Residual Plot  {#sec-model-distance-between-residual-plots}
+
 To develop a computer vision model for assessing residual plots within the visual inference framework, it is important to precisely define a numerical measure of "difference" or "distance" between plots.
@@
 
+In this section, we will introduce a distance measure between a true residual plot and a theoretical 'good' residual plot.
+This measure quantifies the divergence between the residual distribution of a given fitted regression model and that of a correctly specified model.
+The computation assumes knowledge of the data generating processes for predictors and response variables.
+Since these processes are often unknown in practice, we will discuss a method to estimate this distance using a computer vision model in Section @sec-model-distance-estimation.
+
 ### Residual Distribution
 
-The distance measure proposed in this study takes into account the fact that we tried to measure how different a residual plot is from a good residual plot, or in other words, how different a given fitted regression model is from a correctly specified model.
-
-For a classical normal linear regression model, residuals $\boldsymbol{e}$ are derived from the fitted values $\hat{\boldsymbol{y}}$ and observed values $\boldsymbol{y}$.
-Suppose the data generating process is known and the regression model is correctly specified, by the Frisch-Waugh-Lowell theorem [@frisch1933partial], residuals $\boldsymbol{e}$ can also be treated as random variables and written as a linear transformation of the error $\boldsymbol{\varepsilon}$ formulated as $\boldsymbol{e} = \boldsymbol{R}\boldsymbol{\varepsilon}$, where $\boldsymbol{R}=\boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ is the residual operator and an idempotent matrix, $\boldsymbol{X}$ is the design matrix, $\boldsymbol{I}_n$ is a $n$ by $n$ identity matrix, and $n$ is the number of observations.
-
-One of the assumptions of the classical normal linear regression model is the error $\boldsymbol{\varepsilon}$ follows a multivariate normal distribution with zero mean and constant variance, i.e., $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)$.
-It can be known that residuals $\boldsymbol{e}$ also follow a certain probability distribution transformed from the multivariate normal distribution, which will be denoted as $Q$.
-This reference distribution $Q$ summarizes what good residuals should follow given the design matrix $\boldsymbol{X}$ is known and fixed.
+For a classical normal linear regression model, $\boldsymbol{y} = \boldsymbol{X}\boldsymbol{\beta} + \boldsymbol{e}$, the residual $\hat{\boldsymbol{e}}$ are derived as the difference of the fitted values and observed values $\boldsymbol{y}$.
+Suppose the data generating process is known and the regression model is correctly specified, by the Frisch-Waugh-Lowell theorem [@frisch1933partial], residuals $\hat{\boldsymbol{e}}$ can also be treated as random variables and written as a linear transformation of the error $\boldsymbol{e}$ formulated as $\hat{\boldsymbol{e}} = \boldsymbol{R}\boldsymbol{e}$, where $\boldsymbol{R}=\boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}^\top\boldsymbol{X})^{-1}\boldsymbol{X}^\top$ is the residual operator, $\boldsymbol{I}_n$ is a $n$ by $n$ identity matrix, and $n$ is the number of observations.
+
+One of the assumptions of the classical normal linear regression model is that the error $\boldsymbol{e}$ follows a multivariate normal distribution with zero mean and constant variance, i.e., $\boldsymbol{e} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)$.
+It can be known that residuals $\hat{\boldsymbol{e}}$ also follow a certain probability distribution transformed from the multivariate normal distribution, which will be denoted as $Q$.
+This reference distribution $Q$ summarizes what "good" residuals should follow given the design matrix $\boldsymbol{X}$ is known and fixed.
 
@@
 For simplicity, in this study, we replaced the variance-covariance matrix of residuals $\text{cov}(\boldsymbol{e}, \boldsymbol{e}) = \boldsymbol{R}\sigma^2\boldsymbol{R}' = \boldsymbol{R}\sigma^2$ with a full-rank diagonal matrix $\text{diag}(\boldsymbol{R}\sigma^2)$, where $\text{diag}(.)$ sets the non-diagonal entries of a matrix to zeros.
-The resulting distribution for $\boldsymbol{Q}$ is $N(\boldsymbol{0}_n, \text{diag}(\boldsymbol{R}\sigma^2))$.
+The resulting distribution for $Q$ is $N(\boldsymbol{0}_n, \text{diag}(\boldsymbol{R}\sigma^2))$.
 
@@
 It was mainly designed for measuring the effect size of non-linearity and heteroskedasticity in a residual plot.
-@li2023plot have derived that, for a classical normal linear regression model that omits a necessary higher-order predictors $\boldsymbol{Z}$, and incorrectly assumes $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)$ while in fact $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \boldsymbol{V})$ with $\boldsymbol{V}$ being an arbitrary symmetric positive semi-definite matrix, $Q$ can be represented as $N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, \text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}))$.
+@li2023plot have derived that, for a classical normal linear regression model that omits necessary higher-order predictors $\boldsymbol{Z}$ and the corresponding parameter $\boldsymbol{\beta}_z$, and incorrectly assumes $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)$ while in fact $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \boldsymbol{V})$ where $\boldsymbol{V}$ is an arbitrary symmetric positive semi-definite matrix, $Q$ can be represented as $N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, \text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}))$.
 Note that the variance-covariance matrix is replaced with the diagonal matrix to ensure it is a full-rank matrix.
@@
 First, we no longer need to integrate over $n$ random variables.
-The result of @eq-kl-1 is now the sum of the Kullback-Leibler divergence evaluated for each individual residual thanks for the independence assumption.
+The result of @eq-kl-1 is now the sum of the Kullback-Leibler divergence evaluated for each individual residual due to the assumption of independence between observations.
 Second, it is not required to know the joint probability density $p(\boldsymbol{e})$ any more.
@@
 
-We tried to train a computer vision model to estimate this distance measure with a residual plot. Let $D$ be the result of @eq-kl-0, and our estimator $\hat{D}$ is formulated as
+We use a computer vision model to estimate this distance measure with a residual plot. Let $D$ be the result of @eq-kl-0, and our estimator $\hat{D}$ is formulated as
 
@@
 
-\noindent where $V_{h \times w}(.)$ is a plotting function that saves a residuals vs fitted values plot with fixed aesthetic as an RGB image with $h \times w$ pixels, $f_{CV}(.)$ is a computer vision model which takes an $h \times w$ image as input and predicts the distance in the domain $[0, +\infty)$.
-
-With the estimated distance $\hat{D}$, we will be able to know how different the underlying distribution of the residuals is from a good residual distribution.
+\noindent where $V_{h \times w}(.)$ is a plotting function that saves a residuals vs fitted values plot with fixed aesthetic as an image with $h \times w$ pixels in three channels (RGB), $f_{CV}(.)$ is a computer vision model which takes an $h \times w$ image as input and predicts the distance in the domain $[0, +\infty)$.
+
+With the estimated distance $\hat{D}$, we will be able to know how different the underlying distribution of the residuals is from a theoretical "good" residual distribution.
 $\hat{D}$ can also be used as an index of the model violations indicating the strength of the visual signal embedded in the residual plot.
@@
 
-## Model Violations Index (MVI) {#sec:model-violations-index}
+## Model Violations Index {#sec:model-violations-index}
 
@@
 This difference primarily arises from deviations in model assumptions.
-The magnitude of $D$ directly reflects the degree of these deviations, thus making $\hat{D}$ instrumental in forming a model violations index.
+The magnitude of $D$ directly reflects the degree of these deviations, thus making $\hat{D}$ instrumental in forming a model violations index (MVI).
 
@@
 ```{r}
-#| label: factor
+#| label: tbl-factor
 #| tbl-cap: "Factors used in the data generating process for synthetic data simulation.
@@
 
-# Limitations and Future Work
+## Limitations and Future Work
 
@@
 
-# Conclusion
-
-In this study, we have introduced a distance measure based on Kullback-Leibler divergence to quantify the disparity between the residual distribution of a fitted classical normal linear regression model and the reference residual distribution assumed under correct model specification.
+## Conclusion
+
+In this paper, we have introduced a distance measure based on Kullback-Leibler divergence to quantify the disparity between the residual distribution of a fitted classical normal linear regression model and the reference residual distribution assumed under correct model specification.
 This distance measure effectively captures the magnitude of model violations in misspecified models.

### 2024-07-28 Thesis/03-chap3.qmd
further update for emi's comment

@@
 The practice of delivering a residual plot as a lineup is generally regarded as a valuable approach.
-Beyond its application in residual diagnostics, the lineup protocol has integrated into the analysis of diverse subjects.
+Beyond its application in residual diagnostics, the lineup protocol has been integrated into the analysis of diverse subjects.
 For instance, Loy and Hofmann [-@loy2013diagnostic; -@loy2014hlmdiag; -@loy2015you] illustrated its applicability in diagnosing hierarchical linear models.
-Additionally, @widen2016graphical demonstrated its utility in geographical research, while @krishnan2021hierarchical explored its effectiveness in forensic examinations.
-
-A primary limitation of the lineup protocol lies in its reliance on human judgements [see @li2023plot about the practical limitations].
+Additionally, @widen2016graphical and @fieberg2024using demonstrated its utility in geographical and ecology research respectively, while @krishnan2021hierarchical explored its effectiveness in forensic examinations.
+
+A practical limitation of the lineup protocol lies in its reliance on human judgements [see @li2023plot about the practical limitations].
 Unlike conventional statistical tests that can be performed computationally in statistical software, the lineup protocol requires human evaluation of images.
@@
 @wilkinson2005graph expanded on this work, introducing scagnostics based on computable measures applied to planar proximity graphs.
-These measures, including, but not limited to, "Outlying", "Skinny", "Stringy", "Straight", "Monotonic", "Skewed", "Clumpy", and "Striated" aimed to characterize outliers, shape, density, trend, coherence and other characteristics of the data.
+These measures, including, but not limited to, "Outlying", "Skinny", "Stringy", "Straight", "Monotonic", "Skewed", "Clumpy", and "Striated",  aimed to characterize outliers, shape, density, trend, coherence and other characteristics of the data.
 While this approach has been inspiring, there is a recognition [@buja2009statistical] that it may not capture all the necessary visual features that differentiate true residual plots from null plots.
@@
 
-```{r fig-false-finding, fig.width = 8, fig.height = 8*4/5, fig.pos="!h", fig.cap="An example residual vs fitted values plot (red line indicates 0).
+```{r fig-false-finding, fig.width = 4, fig.height = 4*4/5, fig.pos="!h", fig.cap="An example residual vs fitted values plot (red line indicates 0).
 The vertical spread of the data points varies with the fitted values.
@@
 In this study, we chose to define and use a distance between a true residual plot and a theoretical "good" residual plot.
+This is further explained in @sec-model-distance-between-residual-plots.
 @vo2016localizing have also demonstrated that defining a proper distance between images can enhance the matching accuracy in image search compared to a binary outcome model.
@@
 To develop a computer vision model for assessing residual plots within the visual inference framework, it is important to precisely define a numerical measure of "difference" or "distance" between plots.
-This distance can take the form of a basic statistical operation on pixels, such as the sum of square differences.
-Alternatively, it could involve established image similarity metrics like the Structural Similarity Index Measure [@wang2004image].
-The challenge lies in the fact that metrics tailored for image comparison may not be suitable for evaluating data plots, where only essential plot elements require assessment [@chowdhury2018measuring].
-Furthermore, scagnostics mentioned in @sec-model-introduction could be used to construct distance measures for residual plots comparison, but the functional form still needs to be carefully refined to accurately reflect the extent of the violations.
-
-In this section, we will introduce a distance measure between a true residual plot and a theoretical 'good' residual plot.
+This distance can take the form of a basic statistical operation on pixels, such as the sum of square differences, however, a pixel-to-pixel comparison makes little sense in comparing residual plots where the main interest would be structural patterns.
+Alternatively, it could involve established image similarity metrics like the Structural Similarity Index Measure [@wang2004image] which compares images by integrating three perception features of an image: contrast, luminance, and structure (related to average, standard deviation and correlation of pixel values over a window, respectively).
+These image similarity metrics are tailored for image comparison in vastly different tasks to evaluating data plots, where only essential plot elements require assessment [@chowdhury2018measuring].
+We can alternatively define a notion of distance by integrating key plot elements (instead of key perception features like luminance, contrast, and structure), such as those captured by scagnostics mentioned in @sec-model-introduction, but the functional form still needs to be carefully refined to accurately reflect the extent of the violations.
+
+In this section, we introduce a distance measure between a true residual plot and a theoretical 'good' residual plot.
 This measure quantifies the divergence between the residual distribution of a given fitted regression model and that of a correctly specified model.
@@
 
-Suppose the design matrix $\boldsymbol{X}$ has linearly independent columns, the trace of the hat matrix will equal to the number of columns in $\boldsymbol{X}$ denoted as $k$.
+Suppose the design matrix $\boldsymbol{X}$ has linearly independent columns, the trace of the hat matrix $\boldsymbol{H} = \boldsymbol{X}(\boldsymbol{X}^\top\boldsymbol{X})^{-1}\boldsymbol{X}^\top$ will equal to the number of columns in $\boldsymbol{X}$ denoted as $k$.
 As a result, the rank of $\boldsymbol{R}$ is $n - k$, and $Q$ is a degenerate multivariate distribution.
 To capture the characteristics of $Q$, such as moments, we can simulate a large numbers of $\boldsymbol{\varepsilon}$ and transform it to $\boldsymbol{e}$ to get the empirical estimates.
-For simplicity, in this study, we replaced the variance-covariance matrix of residuals $\text{cov}(\boldsymbol{e}, \boldsymbol{e}) = \boldsymbol{R}\sigma^2\boldsymbol{R}' = \boldsymbol{R}\sigma^2$ with a full-rank diagonal matrix $\text{diag}(\boldsymbol{R}\sigma^2)$, where $\text{diag}(.)$ sets the non-diagonal entries of a matrix to zeros.
+For simplicity, in this study, we replaced the variance-covariance matrix of residuals $\text{cov}(\boldsymbol{e}, \boldsymbol{e}) = \boldsymbol{R}\sigma^2\boldsymbol{R}^\top = \boldsymbol{R}\sigma^2$ with a full-rank diagonal matrix $\text{diag}(\boldsymbol{R}\sigma^2)$, where $\text{diag}(.)$ sets the non-diagonal entries of a matrix to zeros.
 The resulting distribution for $Q$ is $N(\boldsymbol{0}_n, \text{diag}(\boldsymbol{R}\sigma^2))$.
@@
 
-Define a proper distance between distributions is usually easier than define a proper distance between data plots.
-Given the true residual distribution $Q$ and the reference residual distribution $P$, we used a distance measure based on Kullback-Leibler divergence [@kullback1951information] to quantify the difference between two distributions
+Defining a proper distance between distributions is usually easier than defining a proper distance between data plots.
+Given the true residual distribution $Q$ and the reference residual distribution $P$, we used a distance measure based on Kullback-Leibler divergence [@kullback1951information] to quantify the difference between two distributions as
 
@@
 
-\noindent and $p(.)$ and $q(.)$ are the probability density functions for distribution $P$ and distribution $Q$ respectively.
+\noindent and $p(.)$ and $q(.)$ are the probability density functions for distribution $P$ and distribution $Q$, respectively.
 
@@
 \begin{aligned}
-D_{KL} &= \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{W})|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + tr(\text{diag}(\boldsymbol{W})^{-1}\text{diag}(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z'(\text{diag}(\boldsymbol{W}))^{-1}\boldsymbol{\mu}_z\right),
+D_{KL} &= \frac{1}{2}\left(\log\frac{|\boldsymbol{W}|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\boldsymbol{W}^{-1}\text{diag}(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z^\top(\boldsymbol{W})^{-1}\boldsymbol{\mu}_z\right),
 \end{aligned}
@@
 
-\noindent where $\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z$, and $\boldsymbol{W} = \boldsymbol{R}\boldsymbol{V}\boldsymbol{R}$.
+\noindent where $\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z$, and $\boldsymbol{W} = \text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R})$.
 The assumed error variance $\sigma^2$ is set to be $tr(\boldsymbol{V})/n$, which is the expectation of the estimated variance.
@@
 This is much easier to approximate through simulation.
-It is also worth mentioning that this independence assumption generally will not hold, since $\text{cov}(e_i, e_j) \neq 0$ if $\boldsymbol{R}_{ij} \neq 0$ for any $1 \leq i < j \leq n$, but its existence is essential for reducing the computational cost.
+It is also worth mentioning that this independence assumption generally will not hold if $\text{cov}(e_i, e_j) \neq 0$ for any $1 \leq i < j \leq n$, but its existence is essential for reducing the computational cost.
 
@@
 D_{KL} &\approx \sum_{i = 1}^{n} \hat{D}_{KL}^{(i)}, \\
-\hat{D}_{KL}^{(i)} &= \frac{1}{m}\sum_{j = 1}^{m} log\frac{\hat{p_i}(B_{ij})}{q(B_{ij})},
+\hat{D}_{KL}^{(i)} &= \frac{1}{m}\sum_{j = 1}^{m} \log\frac{\hat{p_i}(B_{ij})}{q(B_{ij})},
 \end{aligned} 
@@
 
-\noindent where $\hat{D}_{KL}^{(i)}$ is the estimator of the Kullback-Leibler divergence for an individual residual $e_i$, $B_{ij}$ is the $i$-th row and $j$-th column entry of the matrix $B$, $\hat{p_i}(.)$ is the kernel density estimator of $p_i(.)$, $q(.)$ is the normal density function with mean zero and an assumed variance estimated as $\widehat{\sigma^2} = \sum_{b \in vec(B)}(b - \sum_{b \in vec(B)} b/nm)^2/(nm - 1)$, and $vec(.)$ is the vectorization operator which turns a $n \times m$ matrix into a $nm \times 1$ column vector by stacking the columns of the matrix on top of each other.
+\noindent where $\hat{D}_{KL}^{(i)}$ is the estimator of the Kullback-Leibler divergence for an individual residual $e_i$, $B_{ij}$ is the $i$-th row and $j$-th column entry of the matrix $B$, $\hat{p_i}(.)$ is the kernel density estimator of $p_i(.)$, $q(.)$ is the normal density function with mean zero and an assumed variance estimated as $\widehat{\sigma^2} = \sum_{b \in vec(\boldsymbol{B})}(b - \sum_{b \in vec(\boldsymbol{B})} b/nm)^2/(nm - 1)$, and $vec(.)$ is the vectorization operator which turns a $n \times m$ matrix into a $nm \times 1$ column vector by stacking the columns of the matrix on top of each other.
 
@@
 It can be noticed that this distance measure can only be computed when the data generating process is known.
-In reality, we often have no knowledge about the data generating process, otherwise we do not need to fit a linear regression model in the first place.
+In reality, we often have no knowledge about the data generating process, otherwise we do not need to do a residual diagnostic in the first place.
 
@@
 $$
-\text{MVI} = C + \hat{D} - log(n),
+\text{MVI} = C + \hat{D} - \log(n),
 $$ {#eq-mvi}

### 2024-08-05 Thesis/02-chap2.qmd
fix chapter layout
current stage of last paper

@@
 The butterfly has higher power in both tests.
-Curiously, the visual test has a slightly higher power for the ``left-triangle" than the ``right-triangle" shape, when it would be expected that they should be similar, which is observed in conventional testing.'}
+Curiously, the visual test has a slightly higher power for the "left-triangle" than the "right-triangle" shape, when it would be expected that they should be similar, which is observed in conventional testing.'}
 
@@
 
-# Limitations and Practicality
+## Limitations and Practicality
 
@@
 
-# Conclusions
+## Conclusions
 

### 2024-08-05 Thesis/03-chap3.qmd
current stage of last paper

@@
 
-```{r fig-false-finding, fig.width = 4, fig.height = 4*4/5, fig.pos="!h", fig.cap="An example residual vs fitted values plot (red line indicates 0).
+```{r fig-false-finding, fig.width = 4, fig.height = 4*4/5, fig.pos="!h", fig.cap="An example residual vs fitted values plot (red line indicates 0 corresponds to the x-intercept, i.e.
+$y=0$).
 The vertical spread of the data points varies with the fitted values.
-This often indicates the existence of heteroskedasticity.
+This often indicates the existence of heteroskedasticity, however, here the result is due to skewed distribution of the predictors rather than heteroskedasticity.
 The Breusch-Pagan test rejects this residual plot at 95\\% significance level ($p\\text{-value} = 0.046$)."}
@@
 
-In this study, we chose to define and use a distance between a true residual plot and a theoretical "good" residual plot.
+In this study, we chose to define and use a distance between a true residual plot and a theoretically "good" residual plot.
 This is further explained in @sec-model-distance-between-residual-plots.
@@
 
-## Distance from a Theoretical "Good" Residual Plot  {#sec-model-distance-between-residual-plots}
+## Distance from a Theoretically "Good" Residual Plot  {#sec-model-distance-between-residual-plots}
 
@@
 
-In this section, we introduce a distance measure between a true residual plot and a theoretical 'good' residual plot.
+In this section, we introduce a distance measure between a true residual plot and a theoretically 'good' residual plot.
 This measure quantifies the divergence between the residual distribution of a given fitted regression model and that of a correctly specified model.
@@
 \begin{aligned}
-D_{KL} &= \frac{1}{2}\left(\log\frac{|\boldsymbol{W}|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\boldsymbol{W}^{-1}\text{diag}(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z^\top(\boldsymbol{W})^{-1}\boldsymbol{\mu}_z\right),
+D_{KL} &= \frac{1}{2}\left(\log\frac{|\boldsymbol{W}|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\boldsymbol{W}^{-1}\text{diag}(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z^\top\boldsymbol{W}^{-1}\boldsymbol{\mu}_z\right),
 \end{aligned}
@@
 \noindent where $\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z$, and $\boldsymbol{W} = \text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R})$.
-The assumed error variance $\sigma^2$ is set to be $tr(\boldsymbol{V})/n$, which is the expectation of the estimated variance.
+The assumed error variance $\sigma^2$ is set to be $\text{tr}(\boldsymbol{V})/n$, which is the expectation of the estimated variance.
 
@@
 However, since $\boldsymbol{e}$ is a linear transformation of non-normal random variables, it is very common that the general form of $P$ is unknown, meaning that we can not easily compute $p(\boldsymbol{e})$ using a well-known probability density function.
-Additionally, even if $p(\boldsymbol{e})$ can be calculated for any $\boldsymbol{e} \in \mathbb{R}^n$, it will be very difficult to do numerical integration over the $n$ dimensional space, because $n$ could be potentially very large.
+Additionally, even if $p(\boldsymbol{e})$ can be calculated for any $\boldsymbol{e} \in \mathbb{R}^n$, it will be very difficult to do numerical integration over the $n$-dimensional space, because $n$ could be potentially very large.
 
@@
 Then, we can get $m$ sets of realized values of $\boldsymbol{e}$ stored in the matrix $\boldsymbol{B}$ by applying the residual operator $\boldsymbol{B} = \boldsymbol{R}\boldsymbol{A}$.
-Furthermore, kernel density estimation (KDE) with Gaussian kernel and optimal bandwidth selected by the Silverman's rule of thumb [@silverman2018density] is applied on each row of $B$ to estimate $p_i(e_i)$ for $i = 1, ..., n$.
+Furthermore, kernel density estimation (KDE) with Gaussian kernel and optimal bandwidth selected by the Silverman's rule of thumb [@silverman2018density] is applied on each row of $\boldsymbol{B}$ to estimate $p_i(e_i)$ for $i = 1, ..., n$.
 The KDE computation can be done by the `density` function in R.
@@
 
-\noindent where $\hat{D}_{KL}^{(i)}$ is the estimator of the Kullback-Leibler divergence for an individual residual $e_i$, $B_{ij}$ is the $i$-th row and $j$-th column entry of the matrix $B$, $\hat{p_i}(.)$ is the kernel density estimator of $p_i(.)$, $q(.)$ is the normal density function with mean zero and an assumed variance estimated as $\widehat{\sigma^2} = \sum_{b \in vec(\boldsymbol{B})}(b - \sum_{b \in vec(\boldsymbol{B})} b/nm)^2/(nm - 1)$, and $vec(.)$ is the vectorization operator which turns a $n \times m$ matrix into a $nm \times 1$ column vector by stacking the columns of the matrix on top of each other.
+\noindent where $\hat{D}_{KL}^{(i)}$ is the estimator of the Kullback-Leibler divergence for an individual residual $e_i$, $\boldsymbol{B}_{ij}$ is the $i$-th row and $j$-th column entry of the matrix $\boldsymbol{B}$, $\hat{p_i}(.)$ is the kernel density estimator of $p_i(.)$, $q(.)$ is the normal density function with mean zero and an assumed variance estimated as $\widehat{\sigma^2} = \sum_{b \in vec(\boldsymbol{B})}(b - \sum_{b \in vec(\boldsymbol{B})} b/nm)^2/(nm - 1)$, and $vec(.)$ is the vectorization operator which turns a $n \times m$ matrix into a $nm \times 1$ column vector by stacking the columns of the matrix on top of each other.
 
@@
 In the previous sections, we have defined a distance measure given in @eq-kl-0 for quantifying the difference between the true residual distribution $P$ and an ideal reference distribution $Q$.
-It can be noticed that this distance measure can only be computed when the data generating process is known.
+However, this distance measure can only be computed when the data generating process is known.
 In reality, we often have no knowledge about the data generating process, otherwise we do not need to do a residual diagnostic in the first place.
@@
 
-With the estimated distance $\hat{D}$, we will be able to know how different the underlying distribution of the residuals is from a theoretical "good" residual distribution.
+With the estimated distance $\hat{D}$, we can compare the underlying distribution of the residuals to a theoretically "good" residual distribution.
 $\hat{D}$ can also be used as an index of the model violations indicating the strength of the visual signal embedded in the residual plot.
@@
 
-Note that @eq-kl-0 might be influenced by the number of random variables involved in its evaluation.
-Generally, a larger number of observations will lead to a greater distance $D$.
+Note that if more observations are used for estimating the linear regression, the result of @eq-kl-1 will increase, as the integration will be performed over a higher-dimensional space.
+For a consistent data generating process, $D$ typically increases logarithmically with the number of observations.
+This behaviour comes from the relationship $D = \text{log}(1 + D_{KL})$, where $D_{KL} = \sum_{i=1}^{n}D_{KL}^{(i)}$ under the assumption of independence.
+
+Since $\hat{D}$ is an estimate of $D$, it is expected that a larger number of observations will also lead to a higher $\hat{D}$.
 However, this does not imply that $\hat{D}$ fails to accurately represent the extent of model violations.
 In fact, when examining residual plots with more observations, we often observe a stronger visual signal strength, as the underlying patterns are more likely to be revealed, except in cases of significant overlapping.
-
-For a consistent data generating process, $D$ typically increases logarithmically with the number of observations.
-This behaviour comes from the relationship $D = \text{log}(1 + D_{KL})$, where $D_{KL} = \sum_{i=1}^{n}D_{KL}^{(i)}$ under the assumption of independence.
 
@@
 
-```{r fig-delta, fig.pos = "!h", fig.cap = "A weighted detection rate vs $\\delta$-differnence plot. The brown line is smoothing curve produced by fitting gnealized additive models", fig.height = 4}
+```{r fig-delta, fig.pos = "!h", fig.cap = "A weighted detection rate vs $\\delta$-difference plot. The brown line is smoothing curve produced by fitting gnealized additive models", fig.height = 4}
 experiment %>%

### 2024-08-05 Thesis/04-chap4.qmd
current stage of last paper

# Developing Efficient Tools for Automated Residual Plot Assessment: autovi and autovi.web

## Introduction {#sec-web-introduction}

Most statistical software offers graphical tools for diagnosing regression models. For example, the base R package `stats` [@stats] automatically generates four diagnostic plots for linear models (`lm` objects): a residuals vs. fitted values plot, a Scale-Location plot (square root of absolute residuals vs. fitted values), a normal Q-Q plot (citation here), and a residuals vs. leverage plot. Alternatively, packages hosted on The Comprehensive R Archive Network (CRAN) such as `jtools` [@jtools] provide a more comprehensive and aesthetically enhanced set of diagnostic plots. However, very few of them offer automated assessment for these plots. Users often need to manually inspect and interpret the plots, which can lead to inconsistent conclusions due to varying levels of domain and statistical knowledge. This variability makes it challenging to achieve standardized results from plot diagnostics.

The R package `autovi` was developed to address this issue by performing visual statistical testing on residual plots using computer vision. To make its algorithm and trained computer vision model widely accessible, we also developed a web-based tool called `autovi.web`. This tool provides a user-friendly web interface that allows users to diagnose their residual plots without needing to install any dependencies required by the `autovi` package.

This paper documents the design of the R package `autovi` and the web interface `autovi.web`. The rest of the paper is structured as follows: ...

## Background

- current stage of residual diagnostics (optional)
- brief introduction to visual inference
- brief overview of the second paper
- a sentence or two about why we want to develop autovi and autovi.web

## autovi

The primary goal of `autovi` is to provide rejection decisions and $p$-values for testing the null hypothesis that a regression model is correctly specified. While the classical normal linear regression model is popular and widely used, it represents a limited class of models. Therefore, `autovi` is designed to be highly modular and customizable, allowing it to handle a variety of residual diagnostic tasks.

### Quick Start

```{r}
set.seed(10086)
```

```r
library(autovi)
checker <- auto_vi(fitted_model = lm(dist ~ speed, data = cars), 
                   keras_model = get_keras_model("vss_phn_32"))
checker$check()
checker
```

```{r}
library(autovi)
checker <- auto_vi(fitted_model = lm(dist ~ speed, data = cars), 
                   keras_model = get_keras_model("vss_phn_32"))
checker$check()
```

```{r message=TRUE}
checker
```

```{r}
checker$summary_plot()
```

### Modularized Workflow

```{r out.width = "100%"}
#| label: fig-autovi-diag
#| fig-cap: "Diagram illustrating the workflow of the R package autovi. The modules in green are primary inputs provided by users. Modules in blue are overridable methods that can be modified to accommodate users' specific needs. The module in yellow is a pre-defined non-overridable method. The modules in red are primary outputs of the package."
magick::image_read_pdf("figures/autovi.pdf")
```

The workflow of `autovi` begins with the user providing a regression model and concludes with the delivery of visual signal strength and p-value to the user. An overview diagram of the workflow is provided in @fig-autovi-diag.

The workflow comprises two primary inputs, namely "Regression Model" and "Keras Model", and 10 core modules, including "Data Extraction", "Bootstrapping and Model Refitting", "Fitted Values and Residuals Extraction", "Auxiliary Computation", "Null Residual Simulation", "Plotting", "Plot Saving", "Image Reading and Resizing", "Visual Signal Strength Prediction" and "P-value Computation". Each module is designed to make minimal assumptions about the preceding modules, allowing for user customizations without impacting the overall workflow. The "Visual Signal Strength Prediction" and "P-value Computation" modules are predefined methods and can not be overriden, though users can interact with them directly via function arguments. The "Image Reading and Resizing" module is also unchangeable but it will adapt to different Keras models by checking their input shapes. The remaining seven modules are designed to be overriadble, accommodating users' specific needs. We will discuss these modules in details in the following sections.

#### Initialization

To manage workflow state and enable customizable modules, the `autovi` workflow is implemented using the `bandicoot` object-oriented programming framework. This framework allows for seamless integration of user-supplied data, methods, and results, storing them all within a single cohesive object.

An `autovi` checker can be initialized by supplying two primary inputs, including a regression model object, such as an `lm` object representing the result of a linear regression model, and a trained computer vision model compatible with the `Keras` Application Programming Interface (API).

The corresponding function for this step is `auto_vi` as used in the following example code. The `get_keras_model` is a helper function for retrieving trained `Keras` model from the web. We will further discuss it in @sec-trained-model-hosting.

```{r echo=TRUE, message=TRUE}
library(autovi)
checker <- auto_vi(fitted_model = lm(dist ~ speed, data = cars), 
                   keras_model = get_keras_model("vss_phn_32"))
checker
```

Optionally, the user may specify the node index of the output layer of the trained computer vision model to be monitored by the checker via the `node_index` argument if there are multiple output nodes. This is particularly useful for multiclass classifiers when the user wants to use one of the nodes as a visual signal strength indicator.

#### Fitted Values and Residuals Extraction

To create a residual plot, both fitted values and residuals are required. By convention, statistical models in R, such as `lm` (linear model) and `glm` (generalized linear model), support the use of the generic functions `fitted` and `resid` for extracting these values. The "Fitted Values and Residuals Extraction" module uses the same set of generic functions by default. However, since generic functions only work with classes that have appropriate method implementations, users can override the `get_fitted_and_resid` method if their model class does not support these generic functions.

The following code example will returns a `tibble` of two columns: `.fitted` and `.resid` for fitted values and residuals, respectively

```{r echo=TRUE}
checker$get_fitted_and_resid()
```

The result of the module will be passed to "Auxiliary Computation" for computing auxiliary inputs of computer vision model, and "Plotting" modules for generating residual plots.

#### Data Extraction

The "Data Extraction" module involves extracting the model frame from the model object. This is typically done using the `model.frame` generic function, as implemented in the default method `get_data`. Alternatively, users can provide the data used for fitting the regression model via the `data` argument when constructing the checker, or they can override the method to suit their needs. It is worth noting that this module is not necessary if bootstrapping is not required by the user, as the model frame is not used in other steps.

The following code example will returns a `data.frame` representing the model frame of the fitted regression model

```{r echo=TRUE}
checker$get_data() |> head()
```

The result of the module will be passed to "Bootstrapping and Model Refitting" for getting bootstrapped regression models.

#### Bootstrapping and Model Refitting

#### Auxiliary Computation

Sometimes, a residual plot alone may not be sufficient for determining visual signal strength. For example, when the residual plot has significant overlapping, the trend and shape of the residual pattern can be difficult to identify. It can be helpful to include auxiliary variables, such as the number of observations, as inputs to the computer vision model. `autovi` addresses this need with internal functions built into the checker that automatically detect the number of inputs required by the provided Keras model. If multiple inputs are needed, the checker invokes the `auxiliary` method to compute auxiliary inputs. By default, this method calculates four scagnostics, "Monotonic", "Sparse", "Striped", and "Splines, using the `cassowaryr` package, as well as the number of observations. This approach aligns with the training process of our trained computer vision models

The following code example returns a `data.frame` with five columns: four for the scagnostics and one for the number of observations

```{r echo=TRUE}
checker$auxiliary()
```

The result of this module will be passed to the Keras model as part of the inputs.

#### Null Residual Simulation

The "Null Residual Simulation" module, as its name suggests, is responsible for simulating null residuals consistent with the model assumptions. For linear regression models, this assumes the model is correctly specified and simulates random draws from the residual rotation distribution [@buja2009statistical]. Other types of regression models, such as `glm` (generalized linear model) and `gam` (generalized additive model), generally can not use the same method to efficiently simulate null residuals. Therefore, it is highly recommended that users override the `null_method` for this module to suit their specific model.

The following code example returns a `tibble` of two columns: `.fitted` and `.resid` for fitted values and null residuals, respectively

```{r echo=TRUE}
checker$null_method()
```

The result of this module will be passed to the "Plotting" module for generating residual plots.

#### Plotting

Plotting is a crucial aspect of residual plot diagnostics because aesthetic elements like marker size, marker color, and auxiliary lines impact the presentation of information. There are computer vision models trained to handle images captured in various scenarios. For example, the VGG16 model can classify objects in images taken under different lighting conditions and is robust to image rotation. However, data plots are a special type of image as the plotting style can always be consistent if controlled properly. Therefore, we assume computer vision models built for reading residual plots will be trained with residual plots of a specific aesthetic style. In this case, it is best to predict plots using the same style for optimal performance. The "Plotting" module handles this aspect. By default, it replicates the style we used to generate the training samples for the computer vision models. In brief, the residual plot omits axis text and ticks, titles, and background grid lines, featuring only a red line at $y = 0$. Essentially, it retains only the necessary components of a residual plot. If the computer vision model is trained with a different but consistent plot style, the corresponding method `plot_resid` should be overridden.

The following code example generates a `ggplot` object with the described aesthetic style

```{r echo=TRUE}
checker$plot_resid()
```

#### Plot Saving

Another aspect of a standardized residual plot is its resolution. The original image format we used was 420 pixels in height and 525 pixels in width. This resolution was chosen because the original lineup, containing 20 residual plots in a 4x5 grid, is represented by an image of 2100 by 2100 pixels. If a different image size is required, the corresponding method `save_plot` can be overridden.

The following code example saves a residual plot to a temporary file and returns the path to that file as a string

```{r echo=TRUE}
checker$plot_resid() |> checker$save_plot()
```

The result of this module will be passed to the "Image Reading and Resizing" module to be converted to a Numpy array.

#### Image Reading and Resizing

When training computer vision models, it is common practice to test multiple input sizes for the same architecture to find the optimized setup. This requires the original training image to be prepared at a higher resolution than needed and then resized to fit the input size during training. `autovi` includes a built-in class, `KERAS_WRAPPER`, to facilitate this task. This Keras wrapper class has a method, `image_to_array`, that can read an image as a `PIL` image using the `pillow` Python package, resize it to the target input size needed by the Keras model, and convert it to a `numpy` array.

To construct a `KERAS_WRAPPER` object, you need to provide the Keras model as the main argument. However, users typically do not need to use this class directly, as the `autovi` checker can automatically call its methods when performing visual signal strength predictions.

The following code example reads an image of a residual plot and resizes it to 32 by 32 pixels to make it compatible with the computer vision model's input shape.

```{r echo=TRUE}
wrapper <- keras_wrapper(keras_model = checker$keras_model)  
input_array <- checker$plot_resid() |> 
  checker$save_plot() |>
  wrapper$image_to_array()
input_array$shape
```

The result of this module will be passed to the "Visual Signal Strength Prediction" module as input.

#### Visual Signal Strength Prediction

Visual signal strength, as defined in (cite here), is an estimate of the distance between the input residual plot and a theoretically good residual plot. However, visual signal strength can be defined in various ways, similar to how distance between two points can be measured differently. The key is to compare the visual signal strength of the true residual plot with that of the null residual plots to determine if it significantly deviates from the null distribution.

There are multiple ways to obtain visual signal strength from the checker, with the most direct being the `vss` method. By default, it predicts the visual signal strength for the true residual plot. If a `ggplot` or a `data.frame`, such as null residuals generated by the `null_method`, is explicitly passed to this method, it will use that input to predict visual signal strength accordingly.

Another way to obtain visual signal strength is by calling the `check` and `lineup_check` methods. These comprehensive methods perform extensive diagnostics on the true residual plot and store the visual signal strength in the `check_result` field of the checker. Additionally, for obtaining visual signal strength for null residual plots and bootstrapped residual plots, there are two specialized methods, `null_vss` and `boot_vss`, designed for this purpose respectively.

The following code example predicts the visual signal strength for the residual plot of the fitted regression model

```{r echo=TRUE}
checker$vss()
```

The following code example predicts the visual signal strength for five null residual plots

```{r echo=TRUE}
checker$null_vss(5L)
```

The following code example predicts the visual signal strength for five bootstrapped residual plots

```{r}
checker$boot_vss(5L)
```

The result of the module will be passed to the "P-value Computation" module for calculating $p$-value and making a rejection decision.

#### P-value Computation

For a visual test conducted by a single observer using the lineup protocol, the minimum $p$-value is $1/m$, where $m$ is the number of residual plots in a lineup. This is because the Type-I error of this test is $1/m$. Similarly, when performing a lineup check via the `lineup_check` method, the checker generates $m−1$ null plots to combine with the true residual plot, forming a lineup. It then calculates the $p$-value as the ratio of plots with visual signal strength greater than or equal to that of the true plot. The $p$-value of this test is also bounded by $1/m$.

Another way to perform a visual test is to assume a null distribution for visual signal strength and then estimate quantiles of this null distribution as critical values. The $p$-value in this case is the ratio of null plots with visual signal strength greater than or equal to that of the true residual plot. This method typically requires evaluating a large number of null residual plots, as done in the `check` method.

The following code example performs a lineup check with $m = 20$. The `p_value` method will use the correct formula to compute $p$-value according to the check type. Note that the result is subject to randomness due to the small lineup size.

```{r echo=TRUE}
checker$lineup_check(lineup_size = 20L)
checker$p_value()
```

The following code example evaluates 100 null residual plots and 100 bootstrapped residual plots, such that the null distribution and bootstrapped distribution of visual signal strength can be estimated accurately. It then computes the $p$-value and prints the checker to get detailed breakdown of the check result.

```{r echo=TRUE}
checker$check(boot_draws = 100L, null_draws = 100L)
checker$p_value()
```

```{r echo=TRUE, message=TRUE}
checker
```

### Summary Plots

```{r echo=TRUE}
checker$check(boot_draws = 100L, null_draws = 100L)
checker$summary_plot()
```

```{r echo=TRUE}
checker$lineup_check(lineup_size = 20L)
checker$summary_plot()
```

### Keras Model Wrapper

```{r echo=TRUE}
wrapper <- keras_wrapper(checker$keras_model)
wrapper$list_layer_name()
```

```{r echo=TRUE}
checker$plot_resid() |>
  checker$save_plot() |>
  wrapper$image_to_array() |>
  wrapper$predict(auxiliary = checker$auxiliary())
```

### Trained Model Hosting {#sec-trained-model-hosting}

### Dependency Considerations

## autovi.web

The R package `autovi` is designed to provide rejection decisions and $p$-values for testing the null hypothesis that a regression model is correctly specified. To construct a checker with `autovi`, one needs to supply a regression model object—typically an `lm` object representing the result of a linear regression model and a trained computer vision model compatible with the `Keras` API.

The regression model object is used to extract the fitted values and residuals for creating a residual plot. Additionally, a residual rotation technique is applied to the model object to generate null residuals, which are residuals consistent with the null hypothesis. For a linear regression model, this is conventionally achieved by simulating new random standard normal draws and using them as responses to refit the linear regression model.

Having null plots, which are residual plots consisting of null residuals and the original fitted values, is crucial for constituting a visual test. If the visual test were conducted by humans, a lineup consisting of $m-1$ null plots and one true residual plot would be presented to several observers. Observers would then be asked to select the plot they find most different out of the $m$ residual plots. If many observers correctly identify the true residual plot as the most different, it provides evidence for rejecting the null hypothesis that the model is correctly specified. This is because, under the null hypothesis, the true residual plot should be indistinguishable from the null plots.

Instead of human observers, the visual test in `autovi` is performed by a computer vision model. This model is trained to report the visual signal strength of each individual plot in a lineup. The visual signal strength estimates the divergence of the empirical residual distribution from the ideal residual distribution, effectively measuring the degree of model violations. The higher the visual signal strength, the more evidence there is against the null hypothesis.

The computer vision model's training involves estimating this divergence or distance, which quantifies how much the residuals deviate from what is expected under a correctly specified model. More details about the mathematical derivation and the training process of the computer vision model can be found in the paper by Li et.al. (2024).

Furthermore, the computer vision model used in `autovi` requires a fixed-size 4D tensor as input. The dimensions of this tensor are as follows: the first dimension represents the batch size, the second dimension represents the width of the image, the third dimension represents the height of the image, and the fourth dimension represents the number of channels. The model outputs a numeric vector that represents the visual signal strength for each image in the batch. The computer vision model is also trained with a set of fixed-aesthetic residual plots, which means that the input images must be produced using the same data pipeline that was used for the training data preparation. This consistency is crucial for ensuring that the model can accurately interpret and analyze new data.

A significant portion of our web interface is dedicated to managing this data pipeline. This involves processing the user-provided data to generate input images that conform to the required format for the computer vision model. The pipeline ensures that the residual plots created from the user data match the aesthetics and format of the training data, enabling the model to provide accurate visual signal strength assessments.

<!-- Our web interface simplifies this process for the user by automating the necessary steps to transform their data into the appropriate input format. Users can upload their CSV files, and the interface handles the extraction of residuals, the creation of residual plots, and the formatting of these plots into 4D tensors. This seamless integration allows users to focus on interpreting the results rather than on the technical details of data preparation. -->

### Data Pipeline

In this section, we will describe the entire data pipeline, including handling uploaded data, creating and saving fixed-aesthetic residual plots, loading and transforming images to the desired input format, and predicting visual signal strength.

#### Input File Format

As described in Section \ref{background}, the `autovi` package requires a regression model object to initialize the diagnostics. However, it is impractical to ask users to upload an R regression model object for inspection. There are several reasons for this:

1. **User Complexity**: Saving an R object to the filesystem involves extra steps and requires users to have specific knowledge.
2. **Data Sensitivity**: The regression model object may contain sensitive, non-shareable data.
3. **File Size**: The R object is often unnecessarily large because it contains extra information not needed for diagnostics.

To simplify the process, the web interface instead requests a CSV file. This CSV file should contain at least two columns: `.fitted`, representing the fitted values, and `.resid`, representing the residuals. Additionally, it can contain an optional column `.sample` to indicate the ID of the residual plot. This is particularly useful if the user wants to evaluate a lineup of residual plots.

Compared to an R model object, a CSV file can be easily generated by various software programs, not just R. CSV files are widely accepted and can be easily viewed and modified using common desktop applications like Excel. CSV files are generally less sensitive than raw data because most information about the predictors is excluded.

#### Plot Drawing and Image Loading

The training data for the computer vision models consist of $32 \times 32$ RGB residual plots. These plots display fitted values on the x-axis and residuals on the y-axis. All labels, including axis texts, are excluded, and no background grid lines are included in the plots. Residual points are drawn in black with a size of 0.5 points, where there are 72.27 points per inch. Additionally, a horizontal red line is drawn at $y = 0$ to help the computer vision model determine if the residual points are uniformly distributed on both sides of the line. The plot is then saved as a PNG file with a resolution of $420 \times 525$ pixels. This resolution mimics a typical lineup residual plot, which has a resolution of $2100 \times 2100$ pixels and is arranged in four rows and five columns.

The uploaded CSV file will be partitioned based on the values in the optional column `k`. If no optional column `k` is present, the entire data set will be used as one partition. Each partition will utilize the plot specifications to generate one residual plot and produce one PNG file.

The saved PNG plot is loaded as an array, where each entry contains a pixel value of the image. This array is then resized to match the input layer shape of the computer vision model, which is $1 \times 32 \times 32 \times 3$. If multiple images are needed for visual signal strength estimation, the arrays can be stacked together to form a larger array with the shape $n \times 32 \times 32 \times 3$, where $n$ is the number of images.

#### Visual Signal Strength Estimation

Finally, the processed image array will be fed into the computer vision model, and returned a vector of visual signal strength which are numerical values always greater than zero.

### Software Stack

#### Backend

To utilize the `autovi` R package, the server hosting our web interface must have a functional R interpreter. A static HTML page cannot accomplish this task, as it only serves static resources to the client and lacks the capability to execute R code. The alternative option would be WebR, a version of R designed to run within a web browser. However, integrating WebR introduces complexities into the design of the web interface, which we will explore further in Section XXX. Additionally, 
The resizing of the image is originally done by the Python `Pillow` library. To maintain the same data pipeline, we also need a working Python interpreter. Given the required conditions, we have three options: (1) Use a traditional backend like the `Spring` framework written in Java for handling all the income and outcome traffic of the web interface. Meanwhile, install R and Python in the server and call them when needed. (2) Use a Python backend framework like `Flask` so that the `Pillow` library can be used natively. Still, R needs to be installed and correctly configured in the server and called when needed. (3) Use a R backend framework like `Shiny`. This is similar to the second option, but requires to install and configure Python separately.

Option one requires a good understanding

<!-- Thus, we chose to implement the web interface with a shiny server. Shiny server is a backend framework written in R, so it allows us to receive user's input and interactively update the output rendered on the client side using R code.  -->

<!-- We deploy the shiny server using the services provided by Posit, called `shinyapps.io`. It is responsible for reading in the uploaded CSV file with the `readr` R package, splitting the dataset with the `dplyr` R package and drawing the residual plots with the `ggplot2` R package. The resulting PNG files are stored in the temporary directory of the remote machine. -->

<!-- The saved PNG file  -->

#### Frontend

#### Communications between Software

### Distribute Keras Model Files

### Performance Optimization

## Conclusions

### 2024-08-06 Thesis/04-chap4.qmd
make latex happy by removing checker message

@@
                    keras_model = get_keras_model("vss_phn_32"))
+checker$plot_resid()
+```
+
+```{r}
 checker$check()
@@
 
-```{r message=TRUE}
+```{r}
 checker
@@
 
-```{r}
+```{r echo=TRUE}
 checker$summary_plot()
@@
 
-```{r echo=TRUE, message=TRUE}
+```{r echo=TRUE}
 library(autovi)
@@
 This is because the Type-I error of this test is $1/m$.
-Similarly, when performing a lineup check via the `lineup_check` method, the checker generates $m−1$ null plots to combine with the true residual plot, forming a lineup.
+Similarly, when performing a lineup check via the `lineup_check` method, the checker generates $m-1$ null plots to combine with the true residual plot, forming a lineup.
 It then calculates the $p$-value as the ratio of plots with visual signal strength greater than or equal to that of the true plot.
@@
 
-```{r echo=TRUE, message=TRUE}
+```{r echo=TRUE}
 checker

### 2024-08-08 Thesis/03-chap3.qmd
fix for emi's comment up to 3.9

@@
 
+### Auxiliary Information with Scagnostics
+
+In @sec-model-introduction, we mentioned that scagnostics consist of a set of manually designed visual feature extraction functions.
+While our computer vision model will learn its own feature extraction function during training, leveraging additional information from scagnostics can enhance the model's predictive accuracy.
+
+For each residual plot used as an input image, we computed four scagnostics – "Monotonic", "Sparse", "Splines", and "Striped" – using the `cassowaryr` R package [@mason2022cassowaryr].
+These computed measures, along with the number of observations from the fitted model, were provided as the second input for the computer vision model.
+While other scagnostics provide valuable insights, they come with high computational costs and are not suitable for quick inference.
+
 ## Distance from a Theoretically "Good" Residual Plot  {#sec-model-distance-between-residual-plots}
@@
 Note that if more observations are used for estimating the linear regression, the result of @eq-kl-1 will increase, as the integration will be performed over a higher-dimensional space.
-For a consistent data generating process, $D$ typically increases logarithmically with the number of observations.
+For a given data generating process, $D$ typically increases logarithmically with the number of observations.
 This behaviour comes from the relationship $D = \text{log}(1 + D_{KL})$, where $D_{KL} = \sum_{i=1}^{n}D_{KL}^{(i)}$ under the assumption of independence.
@@
 
-\noindent where $C$ is a large enough constant keeping the result positive.
-
-@fig-poly-index and @fig-heter-index display the residual plots for fitted models exhibiting varying degrees of non-linearity and heteroskedasticity.
+\noindent where $C$ is a large enough constant keeping the result positive and the term $-\log(n)$ is used to offset the increase in $D$ due to sample size.
+
+@fig-poly-heter-index displays the residual plots for fitted models exhibiting varying degrees of non-linearity and heteroskedasticity.
 Each residual plot's MVI is computed using @eq-mvi with $C = 10$.
@@
    <th style="text-align:left;"> <span data-qmd="Degree of Model Violations"></span> </th>
-   <th style="text-align:left;"> <span data-qmd="Range ($C$ = 10)"></span>  </th>
+   <th style="text-align:center;"><span data-qmd="Range ($C$ = 10)"></span>  </th>
   </tr>
@@
    <td style="text-align:left;"> <span data-qmd="Strong"></span>  </td>
-   <td style="text-align:left;"> <span data-qmd="$\text{MVI} > 8$"></span>  </td>
+   <td style="text-align:center;"> <span data-qmd="$\text{MVI} > 8$"></span>  </td>
   </tr>
@@
    <td style="text-align:left;"> <span data-qmd="Moderate"></span>  </td>
-   <td style="text-align:left;"> <span data-qmd="$6 < \text{MVI} < 8$"></span> </td>
+   <td style="text-align:center;"> <span data-qmd="$6 < \text{MVI} < 8$"></span> </td>
   </tr>
@@
    <td style="text-align:left;"> <span data-qmd="Weak"></span>  </td>
-   <td style="text-align:left;"> <span data-qmd="$\text{MVI} < 6$"></span>  </td>
+   <td style="text-align:center;"> <span data-qmd="$\text{MVI} < 6$"></span>  </td>
   </tr>
@@
                     escape = FALSE,
-                    col.names = c("Degree of model violations", "Range ($C$ = 10)"))
-```
-
-:::
-
-```{r fig-poly-index, fig.pos = "!h", fig.cap = "Residual plots generated from fitted models exhibiting varying degrees of non-linearity violations.
-The Model violations index (MVI) is displayed atop each residual plot.
-The non-linearity patterns are relatively strong for $MVI > 8$, and relatively weak for $MVI < 6$."}
+                    col.names = c("Degree of model violations", "Range ($C$ = 10)"),
+                    align = "lc") %>%
+  kableExtra::kable_styling(latex_options = c("hold_position"))
+```
+
+:::
+
+```{r}
 set.seed(1495)
@@
 
-dat_combined %>%
+p1 <- dat_combined %>%
   autovi::AUTO_VI$plot_resid() +
@@
              ncol = 5) +
-  theme(strip.text = element_text(size = 10))
-```
-
-```{r fig-heter-index, fig.pos = "!h", fig.cap = "Residual plots generated from fitted models exhibiting varying degrees of heteroskedasticity violations.
-The Model violations index (MVI) is displayed atop each residual plot.
-The heteroskedasticity patterns are relatively strong for $MVI > 8$, and relatively weak for $MVI < 6$."}
+  theme(strip.text = element_text(size = 10), plot.title = element_text(size = 10)) +
+  ggtitle("(A) Non-linearity")
+```
+
+```{r}
 set.seed(1495)
@@
 
-dat_combined %>%
+p2 <- dat_combined %>%
   autovi::AUTO_VI$plot_resid() +
@@
              ncol = 5) +
-  theme(strip.text = element_text(size = 10))
+  theme(strip.text = element_text(size = 10), plot.title = element_text(size = 10)) +
+  ggtitle("(B) Heteroskedasticity")
+```
+
+```{r fig-poly-heter-index, fig.pos = "!h", fig.cap = "Residual plots generated from fitted models exhibiting varying degrees of (A) non-linearity and (B) heteroskedasticity violations.
+The model violations index (MVI) is displayed atop each residual plot.
+The non-linearity patterns are relatively strong for $MVI > 8$, and relatively weak for $MVI < 6$, while the heteroskedasticity patterns are relatively strong for $MVI > 8$, and relatively weak for $MVI < 6$.", fig.height = 12}
+patchwork::wrap_plots(p1, p2, ncol = 1)
 ```
@@
 
-To address this issue while adhering to the principle of visual inference, we can compare the estimated distance $\hat{D}$ to the estimated distances for the null plots in a lineup.
+To address this issue we can adhere to the paradigm of visual inference, by comparing the estimated distance $\hat{D}$ to the estimated distances for the null plots in a lineup.
 Specifically, if a lineup comprises 20 plots, the null hypothesis $H_0$ will be rejected if $\hat{D}$ exceeds the maximum estimated distance among the $m - 1$ null plots, denoted as $\max\limits_{1 \leq i \leq m-1} {\hat{D}_{null}^{(i)}}$, where $\hat{D}_{null}^{(i)}$ represents the estimated distance for the $i$-th null plot.
@@
 
-For lineups consisting of more than 20 plots, the 95% significance level can be maintained if the number of plots is a multiple of 20.
+<!-- For lineups consisting of more than 20 plots, the 95% significance level can be maintained if the number of plots is a multiple of 20.
 Specifically, for lineups comprising $20t$ plots, where $t$ is a positive integer, we reject $H_0$ if $\hat{D}$ exceeds 95% ${\hat{D}_{null}^{(i)}}$ for $i = 1, \ldots, 20t-1$.
 The $p$-value in this case is given by $\frac{1}{20t} + \frac{1}{20t}\sum_{i=1}^{20t-1} I\left(\hat{D}_{null}^{(i)} > \hat{D}\right)$, where $I(\cdot)$ is the indicator function.
+-->
 
 Moreover, if the number of plots in a lineup, denoted by $m$, is sufficiently large, the empirical distribution of ${\hat{D}_{null}^{(i)}}$ can be viewed as an approximation of the null distribution of the estimated distance.
-Consequently, quantiles of the null distribution can be estimated using the sample quantiles available in statistical software such as R, and these quantiles can be utilized for decision-making purposes.
+Consequently, quantiles of the null distribution can be estimated using the sample quantiles, and these quantiles can be utilized for decision-making purposes.
 The details of the sample quantile computation can be found in @hyndman1996sample.
@@
 However, if the null distribution exhibits a long tail, a larger number of null plots may be required.
-Alternatively, a $p$-value can be used to represent the probability of observing an event equally or more extreme than the given event under the null hypothesis $H_0$, and it can be estimated by $\frac{1}{m}\sum_{i=1}^{m-1}I\left(\hat{D}_{null}^{(i)} \geq \hat{D}\right)$.
-
-If precision in sample quantiles is not the main priority, using a pre-calculated table of quantiles is an available option.
-Such a table offers pre-determined quantiles for a specified number of observations.
-It is generated by assessing numerous null plots derived from various simulated regression models and averaging them.
-Essentially, this shifts the computational burden from the user to the developer.
+Alternatively, a $p$-value is the probability of observing a distance equally or greater than $\hat{D}$ under the null hypothesis $H_0$, and it can be estimated by $\frac{1}{m} + \frac{1}{m}\sum_{i=1}^{m-1}I\left(\hat{D}_{null}^{(i)} \geq \hat{D}\right)$.
+
+To alleviate computation burden, a lattice of quantiles for $\hat{D}$ under $H_0$ with specified sample sizes can be precomputed.
+We can then map the $\hat{D}$ and sample size to the closet quantile and sample size in lattice to calculate the corresponding $p$-value.
+This approach lose precision in $p$-value calculation, however, significantly improves computational efficiency.
 
@@
 
-Bootstrap is often employed in linear regression when conducting inference for estimated parameters.
+Bootstrap is often employed in linear regression when conducting inference for estimated parameters [see @davison1997bootstrap and @efron1994introduction].
 It is typically done by sampling individual cases with replacement and refitting the regression model.
@@
 While observational data is frequently employed in training models for real-world applications, the data generating process of observational data often remains unknown, making computation for our target variable $D$ unattainable.
-Consequently, the computer vision models developed in this study were trained using synthetic data, including 80000 training images and 8000 test images.
+Consequently, the computer vision models developed in this study were trained using synthetic data, including 80,000 training images and 8,000 test images.
 This approach provided us with precise label annotations.
@@
 \boldsymbol{w} &= \text{He}_j(g(\boldsymbol{x}_2, 2)), \\
-\boldsymbol{k} &= \sqrt{\boldsymbol{1}_n + b(2 - |a|)(\boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 - a\boldsymbol{1}_n)^2},
+\boldsymbol{k} &= \left[\boldsymbol{1}_n + b(2 - |a|)(\boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 - a\boldsymbol{1}_n)^{\circ2}\right]^{\circ1/2},
 \end{aligned}
@@
 
-\noindent where $\boldsymbol{y}$, $\boldsymbol{x}_1$, $\boldsymbol{x}_2$, $\boldsymbol{z}$, $\boldsymbol{w}$, $\boldsymbol{k}$ and $\boldsymbol{\varepsilon}$ are vectors of size $n$, $\boldsymbol{1}_n$ is a vector of ones of size $n$, $\boldsymbol{x}_1$ and $\boldsymbol{x}_2$ are two independent predictors, $\text{He}_j(.)$ is the $j$th-order probabilist's Hermite polynomials [@hermite1864nouveau], the $\sqrt{(.)}$ and $(.)^2$ operators are element-wise operators, $\odot$ is the Hadamard product, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
-
-$$g(\boldsymbol{x}, k) = 2k \cdot \frac{\boldsymbol{x} - x_{min}\boldsymbol{1}_n}{x_{max} - x_{min}} - k\boldsymbol{1}_n,~for~k > 0,$$
-\noindent where $x_{min} = \underset{i \in \{ 1,...,n\}}{min} x_i$, $x_{max} = \underset{i \in \{ 1,...,n\}}{max} x_i$ and $x_i$ is the $i$-th entry of $\boldsymbol{x}$.
+\noindent where $\boldsymbol{y}$, $\boldsymbol{x}_1$, $\boldsymbol{x}_2$, $\boldsymbol{z}$, $\boldsymbol{w}$, $\boldsymbol{k}$ and $\boldsymbol{\varepsilon}$ are vectors of size $n$, $\boldsymbol{1}_n$ is a vector of ones of size $n$, $\boldsymbol{x}_1$ and $\boldsymbol{x}_2$ are two independent predictors, $\text{He}_j(.)$ is the $j$th-order probabilist's Hermite polynomials [@hermite1864nouveau], $(.)^{\circ2}$ and $(.)^{\circ1/2}$ are Hadamard square and square root, $\odot$ is the Hadamard product, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
+
+$$A^{\circ\frac12}$$
+
+$$g(\boldsymbol{x}, k) = 2k \cdot \frac{\boldsymbol{x} - x_{\min}\boldsymbol{1}_n}{x_{\max} - x_{\min}} - k\boldsymbol{1}_n,~for~k > 0,$$
+\noindent where $x_{\min} = \underset{i \in \{ 1,...,n\}}{\min} x_i$, $x_{\max} = \underset{i \in \{ 1,...,n\}}{\max} x_i$ and $x_i$ is the $i$-th entry of $\boldsymbol{x}$.
 
@@
 
-Additionally, Scaling factor $\boldsymbol{k}$ directly affects the error distribution and it is correlated with $\boldsymbol{x}_1$ and $\boldsymbol{x}_2$.
+Additionally, scaling factor $\boldsymbol{k}$ directly affects the error distribution and it is correlated with $\boldsymbol{x}_1$ and $\boldsymbol{x}_2$.
 If $b \neq 0$ and $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n)$, the constant variance assumption will be violated.
@@
 In the synthetic data simulation, four distinct error distributions are considered, including discrete, uniform, normal, and lognormal distributions, as presented in @fig-different-e.
-Each distribution imparts unique characteristics to the residuals.
-The discrete error distribution introduces discreteness in residuals, while the lognormal distribution typically yields outliers.
+Each distribution imparts unique characteristics in the residual plot.
+The discrete error distribution introduces discrete clusters in residuals, while the lognormal distribution typically yields outliers.
 Uniform error distribution may result in residuals filling the entire space of the residual plot.
@@
 
-In real-world analysis, it's not uncommon to encounter instances where multiple model violations coexist.
+In real-world analysis, it is not uncommon to encounter instances where multiple model violations coexist.
 In such cases, the residual plots often exhibit a mixed pattern of visual anomalies corresponding to different types of model violations.
 @fig-different-j-heter and @fig-different-e-heter show the visual patterns of models with multiple model violations.
-
-### Scagnostics
-
-In @sec-model-introduction, we mentioned that scagnostics consist of a set of manually designed visual feature extraction functions.
-While our computer vision model will learn its own feature extraction function during training, leveraging additional information from scagnostics can enhance the model's predictive accuracy.
-
-For each generated residual plot, we computed four scagnostics – "Monotonic", "Sparse", "Splines", and "Striped" – using the `cassowaryr` R package [@mason2022cassowaryr].
-These computed measures, along with the number of observations from the fitted model, were provided as the second input for the computer vision model.
-While other scagnostics provide valuable insights, they come with high computational costs and are not suitable for quick inference.
 
@@
 
-With 80000 training images prepared, each bucket accommodated a maximum of $80000 \div 50 = 1600$ training samples.
+With 80,000 training images prepared, each bucket accommodated a maximum of $80000/ 50 = 1600$ training samples.
 The simulator iteratively sampled parameter values from the parameter space, generated residuals and fitted values using the data generation process, computed the distance, and checked if the sample fitted within the corresponding bucket.
@@
 
-Similarly, we adopted the same methodology to prepare 8000 test images for performance evaluation and model diagnostics.
+Similarly, we adopted the same methodology to prepare 8,000 test images for performance evaluation and model diagnostics.
 
@@
 
-The model weights $\boldsymbol{\theta}$ were randomly initialized and they were optimized by the Adam optimizer with the mean square error loss function
-
-$$\hat{\boldsymbol{\theta}} = \underset{\boldsymbol{\theta}}{argmin}\frac{1}{n_{train}}\sum_{i=1}^{n_{train}}(D_i - f_{\boldsymbol{\theta}}(V_i, S_i))^2,$$
-
-\noindent where $n_{train}$ is the number of training samples, $V_i$ is the $i$-th residual plot and $S_i$ is the additional information about the $i$-th residual plot including four scagnostics and the number of observations.
+The model weights $\boldsymbol{\theta}$ were randomly initialized and they were optimized by the Adam optimizer [@kingma2014adam] with the mean square error loss function
+
+$$\hat{\boldsymbol{\theta}} = \underset{\boldsymbol{\theta}}{\text{arg min}}\frac{1}{n_{\text{train}}}\sum_{i=1}^{n_{\text{train}}}(D_i - f_{\boldsymbol{\theta}}(V_i, S_i))^2,$$
+
+\noindent where $n_{\text{train}}$ is the number of training samples, $V_i$ is the $i$-th residual plot and $S_i$ is the additional information about the $i$-th residual plot including four scagnostics and the number of observations.
 
@@
 
-The "number of base filters" determines the number of filters for the first 2D convolutional layer.
+The number of base filters determines the number of filters for the first 2D convolutional layer.
 In the VGG16 architecture, the number of filters for the 2D convolutional layer in a block is typically twice the number in the previous block, except for the last block, which maintains the same number of convolution filters as the previous one.
 This hyperparameter aids in controlling the complexity of the computer vision model.
-Higher numbers of base filters result in more trainable parameters.
-Likewise, the "number of units for the fully-connected layer" determines the complexity of the final prediction block.
+A higher number of base filters results in more trainable parameters.
+Likewise, the number of units for the fully-connected layer determines the complexity of the final prediction block.
 Increasing the number of units enhances model complexity, resulting in more trainable parameters.
@@
 The dropout rate and batch normalization are flexible hyperparameters that work in conjunction with other parameters to facilitate smooth training.
-A higher dropout rate is necessary when the model tends to overfit the training dataset by learning too much noise.
+A higher dropout rate is necessary when the model tends to overfit the training dataset by learning too much noise [@srivastava2014dropout].
 Conversely, a lower dropout rate is preferred when the model is complex and challenging to converge.
-Batch normalization, on the other hand, addresses the internal covariate shift problem arising from the randomness in weight initialization and input data.
+Batch normalization, on the other hand, addresses the internal covariate shift problem arising from the randomness in weight initialization and input data [@goodfellow2016deep].
 It helps stabilize and accelerate the training process by normalizing the activations of each layer.
@@
 Additionally, incorporating additional inputs such as scagnostics and the number of observations can potentially enhance prediction accuracy.
-Therefore, we allowed the tuner to determine whether these inputs were necessary for optimal model performance.
+Therefore, we allow the tuner to determine whether these inputs were necessary for optimal model performance.
 
@@
 Additionally, we applied early stopping, terminating the training process if the validation root mean square error fails to improve for 50 epochs.
-The maximum allowed epochs was set at 2000, although no models reached this threshold.
+The maximum allowed epochs was set at 2,000, although no models reached this threshold.
 
@@
 data.frame(`Hyperparameter` = c("Number of base filters", "Dropout rate for convolutional blocks", "Batch normalization for convolutional blocks", "Type of global pooling", "Ignore additional inputs", "Number of units for the fully-connected layer", "Batch normalization for the fully-connected layer", "Dropout rate for the fully-connected layer", "Learning rate"),
-           Domain = c("{4, 8, 16, 32, 64}", "[0.1, 0.6]", "{false, true}", "{max, average}", "{false, true}", "{128, 256, 512, 1024, 2048}", "{false, true}", "[0.1, 0.6]", "[1e-8, 1e-1]")) %>%
+           Domain = c("{4, 8, 16, 32, 64}", "[0.1, 0.6]", "{false, true}", "{max, average}", "{false, true}", "{128, 256, 512, 1024, 2048}", "{false, true}", "[0.1, 0.6]", "[$10^{-8}$, $10^{-1}$]")) %>%
   kableExtra::kable(booktabs = TRUE)
@@
 The number of units required for the fully-connected layer was $256$, a relatively modest number compared to the VGG16 classifier, suggesting that the problem at hand was less complex.
-The optimized learning rates were smaller than the default value recommended by Keras, which is 0.001.
+The optimized learning rates were higher for models with higher resolution input, likely because models with more parameters are more prone to getting stuck in local minima, requiring a higher learning rate.
 

### 2024-08-09 Thesis/03-chap3.qmd
finish emi's comment

@@
 
-In this paper, we develop computer vision models and integrate them into the residual plots diagnostics workflow, addressing the need for an automated visual inference.
-The paper is structured as follows.
+In this chapter, we develop computer vision models and integrate them into the residual plots diagnostics workflow, addressing the need for an automated visual inference.
+The chapter is structured as follows.
 @sec-model-specifications discusses various specifications of the computer vision models.
@@
 
-Taking into account the implementation cost and the need for model interpretability, we used the single residual plot input format in this paper.
+Taking into account the implementation cost and the need for model interpretability, we used the single residual plot input format in this chapter.
 
@@
 
-The training and test performance for the optimized models with three different input sizes are summarized in @tbl-performance.
-Among these models, the $64 \times 64$ model and the $32 \times 32$ model consistently exhibited the best training and test performance, respectively.
-The mean absolute error indicated that the difference between $\hat{D}$ and $D$ was approximately $0.43$ on the test set, a negligible deviation considering the normal range of $D$ typically falls between $0$ and $7$.
+The test performance for the optimized models with three different input sizes are summarized in @tbl-performance.
+Among these models, the $32 \times 32$ model consistently exhibited the best test performance.
+The mean absolute error of the $32 \times 32$ model indicated that the difference between $\hat{D}$ and $D$ was approximately $0.43$ on the test set, a negligible deviation considering the normal range of $D$ typically falls between $0$ and $7$.
 The high $R^2$ values also suggested that the predictions were largely linearly correlated with the target.
 
-@fig-model-performance presents a hexagonal heatmap for $D - \hat{D}$ versus $\hat{D}$.
-The red smoothing curves, fitted by generalized additive models [@hastie2017generalized], demonstrate that all the optimized models perform admirably on both the training and test sets.
-No structural issues are noticeable in @fig-model-performance, but some minor issues regarding over-prediction and under-prediction are observed.
-
-The figure highlights that most under-predictions occurred when $\hat{D} < 3$, while over-predictions occurred predominantly when $3 < \hat{D} < 6$.
-For input images representing null plots, it was expected that the models will over-predict the distance, as explained in @sec-model-lineup-evaluation.
+@fig-model-performance presents a hexagonal heatmap for $D - \hat{D}$ versus $D$.
+The brown smoothing curves, fitted by generalized additive models [@hastie2017generalized], demonstrate that all the optimized models perform admirably on the test sets when $1.5 < D < 6$, where no structural issues are noticeable.
+However, over-predictions occurred when $D < 1.5$, while under-predictions occurred predominantly when $\hat{D} > 6$.
+
+For input images representing null plots where $D = 0$, it was expected that the models will over-predict the distance, as explained in @sec-model-lineup-evaluation.
 However, it can not explain the under-prediction issue.
@@
 Residual plots with non-linearity issues were more challenging to assess than those with heteroskedasticity or non-normality issues.
-Assessing residual plots with heteroskedasticity was not as difficult as assessing those with non-linearity issues.
 When multiple violations were introduced to a residual plot, the performance metrics typically lay between the metrics for each individual violation.
@@
 #| label: tbl-performance
-#| tbl-cap: "The training and test performance of three optimized models with different input sizes."
-train_summary %>%
-  bind_rows(test_summary) %>%
+#| tbl-cap: "The test performance of three optimized models with different input sizes."
+test_summary %>%
   select(res, RMSE, R2, MAE, HUBER) %>%
@@
                     escape = FALSE,
-                    col.names = c("", "RMSE", "$R^2$", "MAE", "Huber loss")) %>%
-  kableExtra::pack_rows("Training set", 1, 3) %>%
-  kableExtra::pack_rows("Test set", 4, 6)
+                    col.names = c("", "RMSE", "$R^2$", "MAE", "Huber loss"))
 ```
@@
 
-::: {#tbl-performance .cell tbl-cap='The training and test performance of three optimized models with different input sizes.'}
+::: {#tbl-performance .cell tbl-cap='The test performance of three optimized models with different input sizes.'}
 ::: {.cell-output-display}
@@
 <tbody>
-  <tr grouplength="3"><td colspan="5" style="border-bottom: 1px solid;"><strong>Training set</strong></td></tr>
-<tr>
-   <td style="text-align:left;padding-left: 2em;" indentlevel="1"> <span data-qmd="$32 \times 32$"></span> </td>
-   <td style="text-align:left;"> 0.531 </td>
-   <td style="text-align:left;"> 0.937 </td>
-   <td style="text-align:left;"> 0.364 </td>
-   <td style="text-align:left;"> 0.126 </td>
-  </tr>
-  <tr>
-   <td style="text-align:left;padding-left: 2em;" indentlevel="1"> <span data-qmd="$64 \times 64$"></span> </td>
-   <td style="text-align:left;"> 0.405 </td>
-   <td style="text-align:left;"> 0.963 </td>
-   <td style="text-align:left;"> 0.260 </td>
-   <td style="text-align:left;"> 0.072 </td>
-  </tr>
-  <tr>
-   <td style="text-align:left;padding-left: 2em;" indentlevel="1"> <span data-qmd="$128 \times 128$"></span> </td>
-   <td style="text-align:left;"> 0.432 </td>
-   <td style="text-align:left;"> 0.959 </td>
-   <td style="text-align:left;"> 0.290 </td>
-   <td style="text-align:left;"> 0.084 </td>
-  </tr>
-  <tr grouplength="3"><td colspan="5" style="border-bottom: 1px solid;"><strong>Test set</strong></td></tr>
 <tr>
@@
 
-```{r fig-model-performance, fig.pos="!h", fig.cap = "Hexagonal heatmap for residuals vs predicted distance on training and test data for three optimized models with different input sizes.
+```{r fig-model-performance, fig.pos="!h", fig.cap = "Hexagonal heatmap for difference in $D$ and $\\hat{D}$ vs $D$ on test data for three optimized models with different input sizes.
 The brown lines are smoothing curves produced by fitting gnealized additive models.
@@
 model_pred %>%
+  filter(type == "test") %>%
   ggplot() +
   geom_hline(yintercept = 0, alpha = 0.5) +
-  # geom_point(aes(effect_size, effect_size - vss), alpha = 0.03) +
   annotate("rect", 
@@
            ymax = 0,
-                xmin = min(model_pred$vss),
-                xmax = max(model_pred$vss),
+           xmin = min(model_pred$effect_size),
+           xmax = max(model_pred$effect_size),
             fill = "#40B0A6",
@@
            ymax = max(model_pred$effect_size - model_pred$vss),
-                xmin = min(model_pred$vss),
-                xmax = max(model_pred$vss),
+           xmin = min(model_pred$effect_size),
+           xmax = max(model_pred$effect_size),
            fill = "#E1BE6A",
@@
   geom_point(data = NULL, aes(5, 0, col = "under-prediction"), shape = 15) +
-  geom_hex(aes(vss, effect_size - vss), bins = 20) +
-  geom_smooth(aes(vss, effect_size - vss), se = FALSE, col = "#994F00") +
-  facet_grid(type ~ res) +
+  geom_hex(aes(effect_size, effect_size - vss), bins = 20) +
+  geom_smooth(aes(effect_size, effect_size - vss), se = FALSE, col = "#994F00") +
+  facet_wrap(~ res) +
   coord_fixed() +
   ylab(expression(D - hat(D))) +
-  xlab(expression(hat(D))) +
+  xlab(expression(D)) +
   theme_light() +
@@
 
-```{r fig-over-under, fig.pos="!h", fig.cap = "Scatter plots for residuals vs $\\sigma$ on test data for the $32 \\times 32$ optimized model.
+```{r fig-over-under, fig.pos="!h", fig.cap = "Scatter plots for difference in $D$ and $\\hat{D}$ vs $\\sigma$ on test data for the $32 \\times 32$ optimized model.
 The data is grouped by whether the regression has only non-linearity violation, and whether it includes a second predictor in the regression formula.
@@
 
-::: {#tbl-performance-sub .cell tbl-cap='The training and test performance of the $32 \times 32$ model presented with different model violations.'}
+::: {#tbl-performance-sub .cell tbl-cap='The test performance of the $32 \times 32$ model presented with different model violations.'}
 ::: {.cell-output-display}
@@
    <th style="text-align:left;"> Violations </th>
-   <th style="text-align:left;"> #Training samples </th>
-   <th style="text-align:left;"> Training RMSE </th>
-   <th style="text-align:left;"> #Test samples </th>
-   <th style="text-align:left;"> Test RMSE </th>
+   <th style="text-align:left;"> #samples </th>
+   <th style="text-align:left;"> RMSE </th>
   </tr>
@@
    <td style="text-align:left;"> no violations </td>
-   <td style="text-align:left;"> 1546 </td>
-   <td style="text-align:left;"> 1.149 </td>
    <td style="text-align:left;"> 155 </td>
@@
    <td style="text-align:left;"> non-linearity </td>
-   <td style="text-align:left;"> 22184 </td>
-   <td style="text-align:left;"> 0.625 </td>
    <td style="text-align:left;"> 2218 </td>
@@
    <td style="text-align:left;"> heteroskedasticity </td>
-   <td style="text-align:left;"> 10826 </td>
-   <td style="text-align:left;"> 0.528 </td>
    <td style="text-align:left;"> 1067 </td>
@@
    <td style="text-align:left;"> non-linearity + heteroskedasticity </td>
-   <td style="text-align:left;"> 10221 </td>
-   <td style="text-align:left;"> 0.593 </td>
    <td style="text-align:left;"> 985 </td>
@@
    <td style="text-align:left;"> non-normality </td>
-   <td style="text-align:left;"> 10797 </td>
-   <td style="text-align:left;"> 0.279 </td>
    <td style="text-align:left;"> 1111 </td>
@@
    <td style="text-align:left;"> non-linearity + non-normality </td>
-   <td style="text-align:left;"> 8835 </td>
-   <td style="text-align:left;"> 0.472 </td>
    <td style="text-align:left;"> 928 </td>
@@
    <td style="text-align:left;"> heteroskedasticity + non-normality </td>
-   <td style="text-align:left;"> 7870 </td>
-   <td style="text-align:left;"> 0.354 </td>
    <td style="text-align:left;"> 819 </td>
@@
    <td style="text-align:left;"> non-linearity + heteroskedasticity + non-normality </td>
-   <td style="text-align:left;"> 7721 </td>
-   <td style="text-align:left;"> 0.432 </td>
    <td style="text-align:left;"> 717 </td>
@@
   mutate(violations = ifelse(violations == "null", "no violations", violations)) %>%
+  select(-train_n, -train_RMSE) %>%
   kableExtra::kable(booktabs = TRUE,
                     escape = FALSE,
-                    col.names = c("Violations", "\\#Training samples", "Training RMSE", "\\#Test samples", "Test RMSE")) %>%
+                    col.names = c("Violations", "\\#samples", "RMSE")) %>%
   kableExtra::kable_styling(latex_options = c("scale_down"))
@@
 
-### Overview of the Human Subject experiment
+#### Overview of the Human Subject experiment
 
@@
 
-This study has collected $7974$ human responses to $1152$ lineups.
+This study has collected 7,974 human responses to 1,152 lineups.
 Each lineup contains one randomly placed true residual plot and 19 null plots.
-Among the $1152$ lineups, $24$ are attention check lineups in which the visual patterns are designed to be extremely obvious and very different from the corresponding to null plots, $36$ are null lineups where all the lineups consist of only null plots, $279$ are lineups with uniform predictor distribution evaluated by $11$ participants, and the remaining $813$ are lineups with discrete, skewed or normal predictor distribution evaluated by $5$ participants.
+Among the 1,152 lineups, 24 are attention check lineups in which the visual patterns are designed to be extremely obvious and very different from the corresponding to null plots, 36 are null lineups where all the lineups consist of only null plots, 279 are lineups with uniform predictor distribution evaluated by 11 participants, and the remaining 813 are lineups with discrete, skewed or normal predictor distribution evaluated by 5 participants.
 Attention check lineups and null lineups will not be assessed in the following analysis.
@@
 
-### Model performance on the human data
+#### Model performance on the human-evaluated data
 
@@
   <tr>
-   <th style="text-align:left;"> Voilation </th>
+   <th style="text-align:left;"> Violation </th>
    <th style="text-align:left;"> RMSE </th>
@@
                     escape = FALSE,
-                    col.names = c("Voilation", "RMSE", "$R^2$", "MAE", "Huber loss"))
+                    col.names = c("Violation", "RMSE", "$R^2$", "MAE", "Huber loss"))
 ```
@@
 We have used our optimized computer vision model to estimate distance for both the true residual plots and the null plots.
-To have a fair comparison, $H_0$ will be rejected if the true residual plot has the greatest approximated distance among all plots in a lineup.
+To have a fair comparison, $H_0$ will be rejected if the true residual plot has the greatest estimated distance among all plots in a lineup.
 Additionally, the appropriate conventional tests including the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey1969tests] for non-linearity and the Breusch-Pagan test  [@breusch1979simple] for heteroskedasticity were applied on the same data for comparison.
@@
 Nevertheless, the mean absolute error remains at a low level, and the linear correlation between the prediction and the true value remains very high.
-Lineups with non-linearity issues are more challenging to predict than those with heteroskedasticity issues.
+Consistent with results in @tbl-performance-sub, lineups with non-linearity issues are more challenging to predict than those with heteroskedasticity issues.
 
@@
 It can be observed that all three tests will agree with each other for around 50% of the cases.
-When visual tests conducted by human do not reject, there are substantial amount of cases where computer vision model also not reject but conventional tests reject.
-There are much fewer cases that do not reject by visual tests and conventional tests, but reject by computer vision models.
+When visual tests conducted by human do not reject, there are substantial amount of cases where computer vision model also do not reject but conventional tests reject.
+There are much fewer cases that do not reject by visual tests and conventional tests, but is rejected by computer vision models.
 This indicates computer vision model can behave like visual tests conducted by human better than conventional tests.
-Moreover, there are great proportion of cases where visual tests conducted by human is the only test who does not reject, .
-
-When plotting the decision against the distance, as illustrated in @fig-power, several observations emerge.
-Firstly, for the computer vision model, when the distance $D > 4$, almost all residual plots are rejected.
-However, for visual tests conducted by humans, the threshold is higher, set at $D > 5$.
-And conventional tests exhibit some non-rejection cases for $D > 3$.
-Moreover, the computer vision model tends to have fewer rejected cases than conventional tests when $D < 2$.
-However, visual tests demonstrate the lowest sensitivity to residual plots with small distances.
-Therefore, tests based on the computer vision model are less sensitive to small deviations from model assumptions than conventional tests but more sensitive to moderate deviations.
-Rejection decisions are fitted by logistic regression models with no intercept terms but an offset equals to $\text{log}(0.05/0/95)$.
-The fitting curves for the computer vision model fall between those of conventional tests and visual tests for both non-linearity and heteroskedasticity.
-There is still potential to refine the computer vision model to better align its behaviour with visual tests.
+Moreover, there are great proportion of cases where visual tests conducted by human is the only test who does not reject.
+
+When plotting the decision against the distance, as illustrated in @fig-power, several notable observations emerge.
+Firstly, compared to conventional tests, the computer vision model tends to have fewer rejected cases when $D < 2$ and fewer non-rejected cases when $2< D < 4$.
+This suggests tests based on the computer vision model are less sensitive to small deviations from model assumptions than conventional tests but more sensitive to moderate deviations.
+Additionally, visual tests demonstrate the lowest sensitivity to residual plots with small distances where not many residual plots are rejected when $D < 2$.
+Similarly, for large distances where $D > 4$, almost all residual plots are rejected by the computer vision model and conventional tests, but for visual tests conducted by humans, the threshold is higher with $D > 5$.
+
+In @fig-power, rejection decisions are fitted by logistic regression models with no intercept terms and an offset equals to $\text{log}(0.05/0.95)$.
+The fitted curves for the computer vision model fall between those of conventional tests and visual tests for both non-linearity and heteroskedasticity, which means there is still potential to refine the computer vision model to better align its behaviour with visual tests conducted by humans.
 
@@
 The weighted detection rate is computed by assigning weights to each detection.
-If the participant selects zero plots, the corresponding weight is 0.05; otherwise, the weight is 1 divided by the number of selections.
-The $\delta$-difference is defined by @chowdhury2018measuring as
+If the participant selects zero plots, the corresponding weight is 0.05; otherwise, if the true residual plot is detected, the weight is 1 divided by the number of selections.
+With the weighted detection rate, we can evaluate the quality of the distance measure purposed in this chapter, by using the $\delta$-difference statistic.
+The $\delta$-difference is originally defined by @chowdhury2018measuring as
 
 \begin{equation}
-\delta = \hat{D} - \underset{j}{max}\left(\hat{D}_{null}^{(j)}\right) \quad \text{for}~j = 1,...,m-1,
+\delta = \bar{d}_{true} - \underset{j}{max}\left(\bar{d}_{null}^{(j)}\right) \quad \text{for}~j = 1,...,m-1,
 \end{equation}
 
-\noindent where $\hat{D}_{null}^{(j)}$ is the approximated distance for the $j$-th null plot, and $m$ is the number of plots in a lineup.
-
-@fig-delta displays the scatter plot of the weighted detection rate vs the $\delta$-difference.
-It indicates that the weighted detection rate increases as the $\delta$-difference increases, particularly when the $\delta$-difference is greater than zero.
-A negative $\delta$-difference suggests that there is at least one null plot in the lineup with a stronger visual signal than the true residual plot.
-In some instances, the weighted detection rate is close to one, yet the $\delta$-difference is negative.
+where $\bar{d}_{null}^{(j)}$ is the mean distance between the $j$-th null plot and other null plots, $\bar{d}_{true}$ is the mean distance between the true residual plot and null plots, and $m$ is the number of plots in a lineup.
+These mean distances are used because the distances considered in @chowdhury2018measuring can change if a different data plot is used to be compared against to.
+For example, suppose there are three null plots, A, B and C, the distance between null plot A and B could be different to the distance between null plot A and C.
+To get a consistent distance value for null plot A, it is necessary to take the average.
+However, this is not applicable to the distance proposed in this chapter, as we only allow the residual plot to be compared to a theoretically good residual plot.
+Therefore, the statistic needs to be modified to evaluate our distance measure.
+
+There is one aspect that the $\delta$-difference was intended to capture, which is the empirical distribution of distance for null plots.
+If we directly replace the mean distances with $D_{\text{null}}$, then the distribution will be a degenerate distribution, as $D_{\text{null}} = 0$ by definition.
+Moreover, $D_{\text{null}}$ can not be derived from an image, so it does not belong to the class of distances considered in @chowdhury2018measuring.
+The one that we should actually concerns about is the empirical distribution of $\hat{D}$ since it affects decision making.
+Thus, the adjusted $\delta$-difference is defined as
+
+\begin{equation}
+\delta_{adjusted} = \hat{D} - \underset{j}{max}\left(\hat{D}_{null}^{(j)}\right) \quad \text{for}~j = 1,...,m-1,
+\end{equation}
+
+\noindent where $\hat{D}_{null}^{(j)}$ is the estimated distance for the $j$-th null plot, and $m$ is the number of plots in a lineup.
+
+@fig-delta displays the scatter plot of the weighted detection rate vs the adjusted $\delta$-difference.
+It indicates that the weighted detection rate increases as the adjusted $\delta$-difference increases, particularly when the adjusted $\delta$-difference is greater than zero.
+A negative adjusted $\delta$-difference suggests that there is at least one null plot in the lineup with a stronger visual signal than the true residual plot.
+In some instances, the weighted detection rate is close to one, yet the adjusted $\delta$-difference is negative.
 This discrepancy implies that the distance measure, or the estimated distance, may not perfectly reflect actual human behaviour.
 
-```{r hist-null-human, fig.pos =  "!h", fig.cap = "Histogram of predicted distance on null plots used in the human subject experiment. Sample quantiles are drawn in different colors."}
+```{r hist-null-human}
 lineup_vss <- readRDS(here::here("data/lineup_vss.rds"))
@@
 
-```{r fig-conv-mosaic, fig.pos = "!h", fig.cap = "Rejection rate ($p$-value $\\leq0.05$) of computer vision models conditional on conventional tests on non-linearity (left) and heteroskedasticity (right) lineups displayed using a mosaic plot."}
+```{r fig-conv-mosaic, fig.pos = "!h", fig.cap = "Rejection rate ($p$-value $\\leq0.05$) of computer vision models conditional on conventional tests on non-linearity (left) and heteroskedasticity (right) lineups displayed using a mosaic plot.
+When the conventional test fails to reject, the computer vision mostly fails to reject the same plot as well as indicated by the height of the top right yellow rectangle, but there are non negliable amount of plots where the conventoinal test rejects but the computer vision model fails to reject as indicated by the width of the top left yellow rectangle."}
 library(ggmosaic)
@@
   theme_light() +
-  scale_x_discrete(labels = c("Visual test", "Computer vision model", "Conventional test")) +
+  scale_x_discrete(labels = c("Visual\ntest", "Computer\nvision\nmodel", "Conventional\ntest")) +
   labs(col = "D") +
@@
   xlab("") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
-        axis.text.y = element_blank(), 
+  theme(axis.text.y = element_blank(), 
         axis.ticks.y = element_blank(), 
@@
 
-```{r fig-delta, fig.pos = "!h", fig.cap = "A weighted detection rate vs $\\delta$-difference plot. The brown line is smoothing curve produced by fitting gnealized additive models", fig.height = 4}
+```{r fig-delta, fig.pos = "!h", fig.cap = "A weighted detection rate vs adjusted $\\delta$-difference plot.
+The brown line is smoothing curve produced by fitting gnealized additive models", fig.height = 4}
 experiment %>%
@@
   ylab("Weighted detection rate") +
-  xlab(expression(delta-difference)) +
+  xlab(expression(delta[adjusted])) +
   theme_light()
@@
 
-In this section, we will present the performance of trained computer vision model on three example datasets.
-These include the dataset associated with the residual plot displaying a "left-triangle" shape, as displayed in @fig-false-finding, along with the Boston housing dataset, and the "dino" datasets from the `datasaurus` R package.
+In this section, we present the performance of trained computer vision model on three example datasets.
+These include the dataset associated with the residual plot displaying a "left-triangle" shape, as displayed in @fig-false-finding, along with the Boston housing dataset [@harrison1978hedonic], and the "dino" datasets from the `datasaurus` R package [@datasaurus].
 
@@
 
-Additionally, we will outline the workflow for employing this model with the `autovi` R package.
-
 ### Left-triangle
@@
 Additionally, the Breusch-Pagan test yielded a rejection with a $p$-value of 0.046, despite the residuals originating from a correctly specified model.
-@fig-false-finding offers a lineup for this fitted model, showcasing various degrees of "left-triangle" shape across all residual plots.
+@fig-false-lineup offers a lineup for this fitted model, showcasing various degrees of "left-triangle" shape across all residual plots.
 This phenomenon is evidently caused by the skewed distribution of the fitted values.
 Notably, if the residual plot were evaluated through a visual test, it would not be rejected since the true residual plot positioned at 10 can not be distinguished from the others.
-
-Moving forward, we will employ the trained computer vision model to assess the residual plot with the assistance of the `autovi` R package.
-This package is accessible on CRAN, and users can install it by executing the command `install.packages("autovi")` in R.
-
-The results of the `list_keras_model()` function provide a list of available trained computer vision models.
-To perform the residual plot assessment, users need to supply a fitted linear regression model and select a trained Keras model to initialize the checker.
-
-```r
-checker <- auto_vi(fitted_model = mod, 
-                   keras_model = get_keras_model("vss_phn_32"))
-```
-
-When invoking the `check()` method to execute the assessment, users can specify the number of null plots and bootstrapped samples using the `null_draws` and `boot_draws` arguments, respectively.
-The diagnostic results can be viewed by directly printing the object.
-Alternatively, a summary plot of the assessment can be generated using the `summary_plot()` method.
-
-```r
-checker$check(null_draws = 200L, boot_draws = 200L)
-checker$summary_plot()
-```
 
@@
 The attention map at [@fig-false-check]B suggests that the estimation is highly influenced by the top-right and bottom-right part of the residual plot, as it forms two vertices of the triangular shape.
-A principle component analysis is also performed on the output of the global pooling layer of the computer vision model.
+A principal component analysis (PCA) is also performed on the output of the global pooling layer of the computer vision model.
 As mentioned in @simonyan2014very, a computer vision model built upon the convolutional blocks can be viewed as a feature extractor.
 For the $32 \times 32$ model, there are 256 features outputted from the global pooling layer, which can be further used for different visual tasks not limited to distance prediction.
-To see if these features can be effectively used for distinguishing null plots and true residual plot, we linearly project them into the PC1 vs PC2 space as shown in @fig-false-check.
+To see if these features can be effectively used for distinguishing null plots and true residual plot, we linearly project them into the first and second principal components space as shown in [@fig-false-check]D.
 It can be observed that because the bootstrapped plots are mostly similar to the null plots, the points drawn in different colours are mixed together.
@@
 (C) The density plot of estimated distnace for null plots and bootstrapped plots.
-The blue area indicates the distribution of estimated distances for bootstrapped plots, while the yellow area represents the distribution of estimated distances for null plots.
+The green area indicates the distribution of estimated distances for bootstrapped plots, while the yellow area represents the distribution of estimated distances for null plots.
 The fitted model will not be rejected since $\\hat{D} < Q_{null}(0.95)$.
-(D) PC2 vs PC1 plot for features extracted from the global pooling layer of the computer vision model.
+(D) plot of first two principal components of features extracted from the global pooling layer of the computer vision model.
 ', fig.height = 8}
@@
   xlab(expression(hat(D))) +
-  scale_linetype(labels = c("95% quantile of the null distribution", expression(hat(D)))) +
+  scale_linetype(labels = c("95% quantile of the null distribution", expression(observed~hat(D)))) +
   scale_fill_manual(values = c("#40B0A6", "#E1BE6A")) +
   scale_color_manual(values = c("#40B0A6", "#E1BE6A"))
-p4 <- my_vi$feature_pca() %>%
+pca_result <- my_vi$feature_pca()
+pc1_per <- (attributes(pca_result)$sdev^2/256)[1]
+pc2_per <- (attributes(pca_result)$sdev^2/256)[2]
+p4 <- pca_result %>%
   mutate(set = ifelse(set == "boot", "Boot", set)) %>%
@@
   scale_color_brewer(palette = "Dark2") +
-  xlab("PC1") +
-  ylab("PC2") +
+  xlab(glue::glue("PC1 ({scales::percent(pc1_per)})")) +
+  ylab(glue::glue("PC2 ({scales::percent(pc2_per)})")) +
   labs(col = "") +
@@
 The Boston housing dataset, originally published by @harrison1978hedonic, offers insights into housing in the Boston, Massachusetts area.
-For illustration purposes, we will utilize a reduced version from Kaggle, comprising 489 rows and 4 columns: average number of rooms per dwelling (RM), percentage of lower status of the population (LSTAT), pupil-teacher ratio by town (PTRATIO), and Median value of owner-occupied homes in $1000's (MEDV).
+For illustration purposes, we utilize a reduced version from Kaggle, comprising 489 rows and 4 columns: average number of rooms per dwelling (RM), percentage of lower status of the population (LSTAT), pupil-teacher ratio by town (PTRATIO), and median value of owner-occupied homes in $1000's (MEDV).
 In our analysis, MEDV will serve as the response variable, while the other columns will function as predictors in a linear regression model.
@@
 The attention map in plot B suggests the center of the image has higher leverage than other areas, and it is the turning point of the "U" shape.
-The principal component analysis provided in plot D shows two distinct clusters of data points, further underlines the visual differences between bootstrapped plots and null plots.
+The CPA provided in plot D shows two distinct clusters of data points, further underling the visual differences between bootstrapped plots and null plots.
 This coincides the findings from @fig-boston-lineup, where the true plot exhibiting a "U" shape is visually distinctive from null plots.
@@
 The fitted model will be rejected since $\\hat{D} \\geq Q_{null}(0.95)$.
-(D) PC2 vs PC1 plot for features extracted from the global pooling layer of the computer vision model.
+(D) plot of first two principal components of features extracted from the global pooling layer of the computer vision model.
 ', fig.height = 8}
@@
 The "dino" dataset within the `datasaurus` R package exemplifies this scenario.
-With only two columns, x and y, fitting a regression model to this data yields a residual plot resembling a "dinosaur," as displayed in [@fig-dino-check]A.
+With only two columns, x and y, fitting a regression model to this data yields a residual plot resembling a "dinosaur", as displayed in [@fig-dino-check]A.
 Unsurprisingly, this distinct residual plot stands out in a lineup, as shown in @fig-dino-lineup.
@@
 The computer vision model is also capable of capturing the contour or the outline of the embedded shape, just like human being reading residual plots.
-The principle component analysis in [@fig-dino-check]D also shows that the cluster of bootstrapped plots is at the corner of the cluster of null plots.
+The PCA in [@fig-dino-check]D also shows that the cluster of bootstrapped plots is at the corner of the cluster of null plots.
 
@@
 The computer vision model effectively captures the contour or outline of the embedded shape, similar to how humans interpret residual plots.
-Additionally, the principal component analysis in [@fig-dino-check]D demonstrates that the cluster of bootstrapped plots is positioned at the corner of the cluster of null plots.
+Additionally, the PCA in [@fig-dino-check]D demonstrates that the cluster of bootstrapped plots is positioned at the corner of the cluster of null plots.
 
@@
 ```{r fig-dino-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment for the Datasaurus fitted model evaluted on 200 null plots and 200 bootstrapped plots.
-(A) The true residual plot exhibiting a "dinosaur" shape.
+(A) The residual plot exhibits a "dinosaur" shape.
 (B) The attention map produced by computing the gradient of the output with respect to the greyscale input.
@@
 The fitted model will be rejected since $\\hat{D} \\geq Q_{null}(0.95)$.
-(D) PC2 vs PC1 plot for features extracted from the global pooling layer of the computer vision model.', fig.height = 8}
+(D) plot of first two principal components of features extracted from the global pooling layer of the computer vision model.', fig.height = 8}
 
@@
   scale_color_brewer(palette = "Dark2") +
-  xlab("PC1") +
-  ylab("PC2") +
+  xlab(glue::glue("PC1 ({scales::percent(pc1_per)})")) +
+  ylab(glue::glue("PC2 ({scales::percent(pc2_per)})")) +
   labs(col = "", shape = "") +
@@
 
-Despite the computer vision model performing well with general cases under the synthetic data generation scheme and the three examples used in this paper, this study has several limitations that could guide future work.
-
-The proposed distance measure assumes that the true model is a classical normal linear regression model, which can be too restrictive.
-Although this paper does not address the relaxation of this assumption, there are potential methods to evaluate other types of regression models.
-The most comprehensive approach would be to define a distance measure for each different class of regression model and then train the computer vision model following the methodology described in this paper.
+Despite the computer vision model performing well with general cases under the synthetic data generation scheme and the three examples used in this chapter, this study has several limitations that could guide future work.
+
+The proposed distance measure assumes that the true model is a classical normal linear regression model, which can be restrictive.
+Although this chapter does not address the relaxation of this assumption, there are potential methods to evaluate other types of regression models.
+The most comprehensive approach would be to define a distance measure for each different class of regression model and then train the computer vision model following the methodology described in this chapter.
 To accelerate training, one could use the convolutional blocks of our trained model as a feature extractor and perform transfer learning on top of it, as these blocks effectively capture shapes in residual plots.
@@
 If only raw residuals are used, the distance-based statistical testing compares the difference in distance to a classical normal linear regression model for the true plot and null plots.
-This comparison is meaningful only if the difference can be identified by the distance measure proposed in this paper.
+This comparison is meaningful only if the difference can be identified by the distance measure proposed in this chapter.
 
@@
 
-In this paper, we have introduced a distance measure based on Kullback-Leibler divergence to quantify the disparity between the residual distribution of a fitted classical normal linear regression model and the reference residual distribution assumed under correct model specification.
+In this chapter, we have introduced a distance measure based on Kullback-Leibler divergence to quantify the disparity between the residual distribution of a fitted classical normal linear regression model and the reference residual distribution assumed under correct model specification.
 This distance measure effectively captures the magnitude of model violations in misspecified models.
@@
 Several examples are provided to showcase the effectiveness of the proposed method across different scenarios, emphasizing the similarity between visual tests and distance-based tests.
-Overall, both visual tests and distance-based tests can be viewed as combined tests, aiming to assess the correctness of all model assumptions collectively.
-In contrast, individual residual diagnostic tests such as the RESET test and the Breusch-Pagan test only evaluate specific model assumptions.
+Overall, both visual tests and distance-based tests can be viewed as ensemble of tests, aiming to assess
+any violations of model assumptions collectively.
+In contrast, individual residual diagnostic tests such as the RESET test and the Breusch-Pagan test only evaluate specific violations of model assumptions.
 In practice, selecting an appropriate set of statistical tests for regression diagnostics can be challenging, particularly given the necessity of adjusting the significance level for each test.

### 2024-08-11 Thesis/03-chap3.qmd
update model violations index section

@@
 @sec-model-specifications discusses various specifications of the computer vision models.
-@sec-model-distance-between-residual-plots defines the distance measure used to measure model violations.
-@sec-model-distance-estimation explains how the computer vision models estimate this distance measure.
-@sec-model-statistical-testing covers the statistical testing based on the estimated distance.
+@sec-model-distance-between-residual-plots defines the distance measure used to detect model violations, while @sec-model-distance-estimation explains how the computer vision models estimate this distance measure.
+@sec-model-statistical-testing covers the statistical tests based on the estimated distance, and @sec-model-violations-index introduces a Model Violations Index, which offers a quicker and more convenient assessment.
 Sections [-@sec-model-data-generation], [-@sec-model-architecture], and [-@sec-model-training] detail the data preparation, model architecture, and training process, respectively.
 The results are presented in @sec-model-results.
+Example dataset applications are discussed in @sec-examples.
 Finally, we conclude with a discussion of our findings and propose ideas for future research directions.
@@
 
-## Model Violations Index {#sec:model-violations-index}
-
-$\hat{D}$ is an estimator of the difference between the true residual distribution and the reference residual distribution.
-This difference primarily arises from deviations in model assumptions.
+## Statistical testing {#sec-model-statistical-testing}
+
+### Lineup Evaluation {#sec-model-lineup-evaluation}
+
+Theoretically, the distance $D$ for a correctly specified model is $0$, because $P$ will be the same as $Q$.
+However, the computer vision model may not necessary predict $0$ for a null plot.
+Using @fig-false-finding as an example, it contains a visual pattern which is an indication of heteroskedasticity.
+We would not expect the model to be able to magically tell if the suspicious pattern is caused by the skewed distribution of the fitted values or the existence of heteroskedasticity.
+Additionally, some null plots could have outliers or strong visual patterns due to randomness, and a reasonable model will try to summarise those information into the prediction, resulting in $\hat{D} > 0$.
+
+This property is not an issue if $\hat{D} \gg 0$ for which the visual signal of the residual plot is very strong, and we usually do not need any further examination of the significance of the result.
+However, if the visual pattern is weak or moderate, having $\hat{D}$ will not be sufficient to determine if $H_0$ should be rejected.
+
+To address this issue we can adhere to the paradigm of visual inference, by comparing the estimated distance $\hat{D}$ to the estimated distances for the null plots in a lineup.
+Specifically, if a lineup comprises 20 plots, the null hypothesis $H_0$ will be rejected if $\hat{D}$ exceeds the maximum estimated distance among the $m - 1$ null plots, denoted as $\max\limits_{1 \leq i \leq m-1} {\hat{D}_{null}^{(i)}}$, where $\hat{D}_{null}^{(i)}$ represents the estimated distance for the $i$-th null plot.
+This approach is equivalent to the typical lineup protocol requiring a 95% significance level, where $H_0$ is rejected if the data plot is identified as the most distinct plot by the sole observer.
+The estimated distance serves as a metric to quantify the difference between the data plot and the null plots, as intended.
+
+<!-- For lineups consisting of more than 20 plots, the 95% significance level can be maintained if the number of plots is a multiple of 20.
+Specifically, for lineups comprising $20t$ plots, where $t$ is a positive integer, we reject $H_0$ if $\hat{D}$ exceeds 95% ${\hat{D}_{null}^{(i)}}$ for $i = 1, \ldots, 20t-1$.
+The $p$-value in this case is given by $\frac{1}{20t} + \frac{1}{20t}\sum_{i=1}^{20t-1} I\left(\hat{D}_{null}^{(i)} > \hat{D}\right)$, where $I(\cdot)$ is the indicator function.
+-->
+
+Moreover, if the number of plots in a lineup, denoted by $m$, is sufficiently large, the empirical distribution of ${\hat{D}_{null}^{(i)}}$ can be viewed as an approximation of the null distribution of the estimated distance.
+Consequently, quantiles of the null distribution can be estimated using the sample quantiles, and these quantiles can be utilized for decision-making purposes.
+The details of the sample quantile computation can be found in @hyndman1996sample.
+For instance, if $\hat{D}$ is greater than or equal to the 95% sample quantile, denoted as $Q_{null}(0.95)$, we can conclude that the estimated distance for the true residual plot is significantly different from the estimated distance for null plots with a 95% significance level.
+Based on our experience, to obtain a stable estimate of the 95% quantile, the number of null plots, $n_{null}$, typically needs to be at least 100.
+However, if the null distribution exhibits a long tail, a larger number of null plots may be required.
+Alternatively, a $p$-value is the probability of observing a distance equally or greater than $\hat{D}$ under the null hypothesis $H_0$, and it can be estimated by $\frac{1}{m} + \frac{1}{m}\sum_{i=1}^{m-1}I\left(\hat{D}_{null}^{(i)} \geq \hat{D}\right)$.
+
+To alleviate computation burden, a lattice of quantiles for $\hat{D}$ under $H_0$ with specified sample sizes can be precomputed.
+We can then map the $\hat{D}$ and sample size to the closet quantile and sample size in lattice to calculate the corresponding $p$-value.
+This approach lose precision in $p$-value calculation, however, significantly improves computational efficiency.
+
+### Bootstrapping
+
+Bootstrap is often employed in linear regression when conducting inference for estimated parameters [see @davison1997bootstrap and @efron1994introduction].
+It is typically done by sampling individual cases with replacement and refitting the regression model.
+If the observed data accurately reflects the true distribution of the population, the bootstrapped estimates can be used to measure the variability of the parameter estimate without making strong distributional assumptions about the data generating process.
+
+Similarly, bootstrap can be applied on the estimated distance $\hat{D}$.
+For each refitted model $M_{boot}^{(i)}$, there will be an associated residual plot $V_{boot}^{(i)}$ which can be fed into the computer vision model to obtain $\hat{D}_{boot}^{(i)}$, where $i = 1,...,n_{boot}$, and $n_{boot}$ is the number of bootstrapped samples.
+If we are interested in the variation of $\hat{D}$, we can use $\hat{D}_{boot}^{(i)}$ to estimate a confidence interval.
+
+Alternatively, since each $M_{boot}^{(i)}$ has a set of estimated coefficients $\hat{\boldsymbol{\beta}}_{boot}^{(i)}$ and an estimated variance $\hat{\sigma^2}_{boot}^{(i)}$, a new approximated null distribution can be construed and the corresponding 95% sample quantile $Q_{boot}^{(i)}(0.95)$ can be computed.
+Then, if $\hat{D}_{boot}^{(i)} \geq Q_{boot}^{(i)}(0.95)$, $H_0$ will be rejected for $M_{boot}^{(i)}$.
+The ratio of rejected $M_{boot}^{(i)}$ among all the refitted models provides an indication of how often the assumed regression model are considered to be incorrect if the data can be obtained repetitively from the same data generating process.
+But this approach is computationally very expensive because it requires $n_{boot} \times n_{null}$ times of residual plot assessment.
+In practice, $Q_{null}(0.95)$ can be used to replace $Q_{boot}^{(i)}(0.95)$ in the computation.
+
+## Model Violations Index {#sec-model-violations-index}
+
+While statistical testing is a powerful tool for detecting model violations, it can become cumbersome and time-consuming when quick decisions are needed, particularly due to the need to evaluate numerous null plots.
+In practice, a more convenient and immediate method for assessing model performance is often required.
+This is where an index, such as the Model Violations Index (MVI), becomes valuable.
+It offers a straightforward way to quantify deviations from model assumptions, enabling rapid assessment and easier comparison across models.
+
+The estimator $\hat{D}$ measures the difference between the true residual distribution and the reference residual distribution, a difference primarily arises from deviations in model assumptions.
 The magnitude of $D$ directly reflects the degree of these deviations, thus making $\hat{D}$ instrumental in forming a model violations index (MVI).
@@
 
-::: {#tbl-mvi .cell tbl-cap='Degree of model violations or the strength of the visual signals according to the Model Violation Index (MVI). The constant $C$ is set to be 10.'}
+::: {#tbl-mvi .cell tbl-cap='Degree of model violations or the strength of the visual signals according to the Model Violations Index (MVI). The constant $C$ is set to be 10.'}
 ::: {.cell-output-display}
@@
 ```{r}
-#| tbl-cap: "Degree of model violations or the strength of the visual signals according to the Model Violation Index (MVI). The constant $C$ is set to be 10."
+#| tbl-cap: "Degree of model violations or the strength of the visual signals according to the Model Violations Index (MVI). The constant $C$ is set to be 10."
 #| label: tbl-mvi
@@
                     col.names = c("Degree of model violations", "Range ($C$ = 10)"),
-                    align = "lc") %>%
-  kableExtra::kable_styling(latex_options = c("hold_position"))
+                    align = "lc")
 ```
@@
 
-## Statistical testing {#sec-model-statistical-testing}
-
-### Lineup Evaluation {#sec-model-lineup-evaluation}
-
-Theoretically, the distance $D$ for a correctly specified model is $0$, because $P$ will be the same as $Q$.
-However, the computer vision model may not necessary predict $0$ for a null plot.
-Using @fig-false-finding as an example, it contains a visual pattern which is an indication of heteroskedasticity.
-We would not expect the model to be able to magically tell if the suspicious pattern is caused by the skewed distribution of the fitted values or the existence of heteroskedasticity.
-Additionally, some null plots could have outliers or strong visual patterns due to randomness, and a reasonable model will try to summarise those information into the prediction, resulting in $\hat{D} > 0$.
-
-This property is not an issue if $\hat{D} \gg 0$ for which the visual signal of the residual plot is very strong, and we usually do not need any further examination of the significance of the result.
-However, if the visual pattern is weak or moderate, having $\hat{D}$ will not be sufficient to determine if $H_0$ should be rejected.
-
-To address this issue we can adhere to the paradigm of visual inference, by comparing the estimated distance $\hat{D}$ to the estimated distances for the null plots in a lineup.
-Specifically, if a lineup comprises 20 plots, the null hypothesis $H_0$ will be rejected if $\hat{D}$ exceeds the maximum estimated distance among the $m - 1$ null plots, denoted as $\max\limits_{1 \leq i \leq m-1} {\hat{D}_{null}^{(i)}}$, where $\hat{D}_{null}^{(i)}$ represents the estimated distance for the $i$-th null plot.
-This approach is equivalent to the typical lineup protocol requiring a 95% significance level, where $H_0$ is rejected if the data plot is identified as the most distinct plot by the sole observer.
-The estimated distance serves as a metric to quantify the difference between the data plot and the null plots, as intended.
-
-<!-- For lineups consisting of more than 20 plots, the 95% significance level can be maintained if the number of plots is a multiple of 20.
-Specifically, for lineups comprising $20t$ plots, where $t$ is a positive integer, we reject $H_0$ if $\hat{D}$ exceeds 95% ${\hat{D}_{null}^{(i)}}$ for $i = 1, \ldots, 20t-1$.
-The $p$-value in this case is given by $\frac{1}{20t} + \frac{1}{20t}\sum_{i=1}^{20t-1} I\left(\hat{D}_{null}^{(i)} > \hat{D}\right)$, where $I(\cdot)$ is the indicator function.
--->
-
-Moreover, if the number of plots in a lineup, denoted by $m$, is sufficiently large, the empirical distribution of ${\hat{D}_{null}^{(i)}}$ can be viewed as an approximation of the null distribution of the estimated distance.
-Consequently, quantiles of the null distribution can be estimated using the sample quantiles, and these quantiles can be utilized for decision-making purposes.
-The details of the sample quantile computation can be found in @hyndman1996sample.
-For instance, if $\hat{D}$ is greater than or equal to the 95% sample quantile, denoted as $Q_{null}(0.95)$, we can conclude that the estimated distance for the true residual plot is significantly different from the estimated distance for null plots with a 95% significance level.
-Based on our experience, to obtain a stable estimate of the 95% quantile, the number of null plots, $n_{null}$, typically needs to be at least 100.
-However, if the null distribution exhibits a long tail, a larger number of null plots may be required.
-Alternatively, a $p$-value is the probability of observing a distance equally or greater than $\hat{D}$ under the null hypothesis $H_0$, and it can be estimated by $\frac{1}{m} + \frac{1}{m}\sum_{i=1}^{m-1}I\left(\hat{D}_{null}^{(i)} \geq \hat{D}\right)$.
-
-To alleviate computation burden, a lattice of quantiles for $\hat{D}$ under $H_0$ with specified sample sizes can be precomputed.
-We can then map the $\hat{D}$ and sample size to the closet quantile and sample size in lattice to calculate the corresponding $p$-value.
-This approach lose precision in $p$-value calculation, however, significantly improves computational efficiency.
-
-### Bootstrapping
-
-Bootstrap is often employed in linear regression when conducting inference for estimated parameters [see @davison1997bootstrap and @efron1994introduction].
-It is typically done by sampling individual cases with replacement and refitting the regression model.
-If the observed data accurately reflects the true distribution of the population, the bootstrapped estimates can be used to measure the variability of the parameter estimate without making strong distributional assumptions about the data generating process.
-
-Similarly, bootstrap can be applied on the estimated distance $\hat{D}$.
-For each refitted model $M_{boot}^{(i)}$, there will be an associated residual plot $V_{boot}^{(i)}$ which can be fed into the computer vision model to obtain $\hat{D}_{boot}^{(i)}$, where $i = 1,...,n_{boot}$, and $n_{boot}$ is the number of bootstrapped samples.
-If we are interested in the variation of $\hat{D}$, we can use $\hat{D}_{boot}^{(i)}$ to estimate a confidence interval.
-
-Alternatively, since each $M_{boot}^{(i)}$ has a set of estimated coefficients $\hat{\boldsymbol{\beta}}_{boot}^{(i)}$ and an estimated variance $\hat{\sigma^2}_{boot}^{(i)}$, a new approximated null distribution can be construed and the corresponding 95% sample quantile $Q_{boot}^{(i)}(0.95)$ can be computed.
-Then, if $\hat{D}_{boot}^{(i)} \geq Q_{boot}^{(i)}(0.95)$, $H_0$ will be rejected for $M_{boot}^{(i)}$.
-The ratio of rejected $M_{boot}^{(i)}$ among all the refitted models provides an indication of how often the assumed regression model are considered to be incorrect if the data can be obtained repetitively from the same data generating process.
-But this approach is computationally very expensive because it requires $n_{boot} \times n_{null}$ times of residual plot assessment.
-In practice, $Q_{null}(0.95)$ can be used to replace $Q_{boot}^{(i)}(0.95)$ in the computation.
-
 ## Data Generation {#sec-model-data-generation}
@@
 
+::: {.content-visible when-format="html"}
+
+::: {#tbl-hyperparameter .cell tbl-cap='Name of hyperparameters and their correspoding domain for the computer vision model.'}
+::: {.cell-output-display}
+
+`````{=html}
+<table>
+ <thead>
+  <tr>
+   <th style="text-align:left;"> Hyperparameter </th>
+   <th style="text-align:left;"> Domain </th>
+  </tr>
+ </thead>
+<tbody>
+  <tr>
+   <td style="text-align:left;"> Number of base filters </td>
+   <td style="text-align:left;"> {4, 8, 16, 32, 64} </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> Dropout rate for convolutional blocks </td>
+   <td style="text-align:left;"> [0.1, 0.6] </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> Batch normalization for convolutional blocks </td>
+   <td style="text-align:left;"> {false, true} </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> Type of global pooling </td>
+   <td style="text-align:left;"> {max, average} </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> Ignore additional inputs </td>
+   <td style="text-align:left;"> {false, true} </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> Number of units for the fully-connected layer </td>
+   <td style="text-align:left;"> {128, 256, 512, 1024, 2048} </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> Batch normalization for the fully-connected layer </td>
+   <td style="text-align:left;"> {false, true} </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> Dropout rate for the fully-connected layer </td>
+   <td style="text-align:left;"> [0.1, 0.6] </td>
+  </tr>
+  <tr>
+   <td style="text-align:left;"> Learning rate </td>
+   <td style="text-align:left;"><span data-qmd = "[$10^{-8}$, $10^{-1}$]"></span> </td>
+  </tr>
+</tbody>
+</table>
+
+`````
+
+:::
+:::
+
+:::
+
+::: {.content-visible when-format="pdf"}
+
 ```{r}
@@
            Domain = c("{4, 8, 16, 32, 64}", "[0.1, 0.6]", "{false, true}", "{max, average}", "{false, true}", "{128, 256, 512, 1024, 2048}", "{false, true}", "[0.1, 0.6]", "[$10^{-8}$, $10^{-1}$]")) %>%
-  kableExtra::kable(booktabs = TRUE)
-```
+  kableExtra::kable(booktabs = TRUE, escape = FALSE)
+```
+
+:::
 
@@
 
-## Examples
+## Examples {#sec-examples}
 
 In this section, we present the performance of trained computer vision model on three example datasets.
-These include the dataset associated with the residual plot displaying a "left-triangle" shape, as displayed in @fig-false-finding, along with the Boston housing dataset [@harrison1978hedonic], and the "dino" datasets from the `datasaurus` R package [@datasaurus].
+These include the dataset associated with the residual plot displaying a "left-triangle" shape, as displayed in @fig-false-finding, along with the Boston housing dataset [@harrison1978hedonic], and the "dino" datasets from the `datasauRus` R package [@datasaurus].
 
@@
 ', fig.height = 8}
-p1 <- my_vi$plot_resid(theme = theme_light()) + ggtitle("(A) Residual plot") 
+p1 <- my_vi$plot_resid(theme = theme_light()) + ggtitle(glue::glue("(A) Residual plot (MVI = {format(my_vi$check_result$observed$vss - log(nrow(my_vi$get_fitted_and_resid())) + 10, digits = 3)}, weak)"))
 p2 <- readRDS(here::here("cached_data/fig1_attention_map.rds")) + ggtitle("(B) Attention map")
@@
 ', fig.height = 8}
-p1 <- my_vi$plot_resid(theme = theme_light()) + ggtitle("(A) Residual plot") 
+p1 <- my_vi$plot_resid(theme = theme_light()) + ggtitle(glue::glue("(A) Residual plot (MVI = {format(my_vi$check_result$observed$vss - log(nrow(my_vi$get_fitted_and_resid())) + 10, digits = 3)}, strong)"))
 p2 <- readRDS(here::here("cached_data/boston_attention_map.rds")) + ggtitle("(B) Attention map")
@@
 
-### Datasaurus
+### DatasauRus
 
@@
 
-The "dino" dataset within the `datasaurus` R package exemplifies this scenario.
+The "dino" dataset within the `datasauRus` R package exemplifies this scenario.
 With only two columns, x and y, fitting a regression model to this data yields a residual plot resembling a "dinosaur", as displayed in [@fig-dino-check]A.
@@
 
-```{r fig-dino-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment for the Datasaurus fitted model evaluted on 200 null plots and 200 bootstrapped plots.
+```{r fig-dino-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment for the datasauRus fitted model evaluted on 200 null plots and 200 bootstrapped plots.
 (A) The residual plot exhibits a "dinosaur" shape.
@@
 
-p1 <- my_vi$plot_resid(theme = theme_light()) + ggtitle("(A) Residual plot")
+p1 <- my_vi$plot_resid(theme = theme_light()) + ggtitle(glue::glue("(A) Residual plot (MVI = {format(my_vi$check_result$observed$vss - log(nrow(my_vi$get_fitted_and_resid())) + 10, digits = 3)}, strong)"))
 p2 <- readRDS(here::here("cached_data/dino_attention_map.rds")) + ggtitle("(B) Attention map")
@@
 We propose a computer vision model to estimate this distance, utilizing the residual plot of the fitted model as input.
-The resulting estimated distance serves as the foundation for constructing a single Model Violation Index (MVI), facilitating the quantification of various model violations.
+The resulting estimated distance serves as the foundation for constructing a single Model Violations Index (MVI), facilitating the quantification of various model violations.
 

### 2024-08-11 Thesis/03-chap3.qmd
fix labelling issue

@@
 D_{KL} &\approx \sum_{i = 1}^{n} \hat{D}_{KL}^{(i)}, \\
-\hat{D}_{KL}^{(i)} &= \frac{1}{m}\sum_{j = 1}^{m} \log\frac{\hat{p_i}(B_{ij})}{q(B_{ij})},
+\hat{D}_{KL}^{(i)} &= \frac{1}{m}\sum_{j = 1}^{m} \log\frac{\hat{p}_i(B_{ij})}{q(B_{ij})},
 \end{aligned} 
@@
 
-\noindent where $\hat{D}_{KL}^{(i)}$ is the estimator of the Kullback-Leibler divergence for an individual residual $e_i$, $\boldsymbol{B}_{ij}$ is the $i$-th row and $j$-th column entry of the matrix $\boldsymbol{B}$, $\hat{p_i}(.)$ is the kernel density estimator of $p_i(.)$, $q(.)$ is the normal density function with mean zero and an assumed variance estimated as $\widehat{\sigma^2} = \sum_{b \in vec(\boldsymbol{B})}(b - \sum_{b \in vec(\boldsymbol{B})} b/nm)^2/(nm - 1)$, and $vec(.)$ is the vectorization operator which turns a $n \times m$ matrix into a $nm \times 1$ column vector by stacking the columns of the matrix on top of each other.
+\noindent where $\hat{D}_{KL}^{(i)}$ is the estimator of the Kullback-Leibler divergence for an individual residual $e_i$, $\boldsymbol{B}_{ij}$ is the $i$-th row and $j$-th column entry of the matrix $\boldsymbol{B}$, $\hat{p}_i(.)$ is the kernel density estimator of $p_i(.)$, $q(.)$ is the normal density function with mean zero and an assumed variance estimated as $\hat{\sigma}^2 = \sum_{b \in vec(\boldsymbol{B})}(b - \sum_{b \in vec(\boldsymbol{B})} b/nm)^2/(nm - 1)$, and $vec(.)$ is the vectorization operator which turns a $n \times m$ matrix into a $nm \times 1$ column vector by stacking the columns of the matrix on top of each other.
 
@@
 
-$$A^{\circ\frac12}$$
-
 $$g(\boldsymbol{x}, k) = 2k \cdot \frac{\boldsymbol{x} - x_{\min}\boldsymbol{1}_n}{x_{\max} - x_{\min}} - k\boldsymbol{1}_n,~for~k > 0,$$
@@
   <tr>
-   <td style="text-align:left;"> <span data-qmd="$\text{dist}_varepsilon$"></span> </td>
+   <td style="text-align:left;"> <span data-qmd="$\text{dist}_{\varepsilon}$"></span> </td>
    <td style="text-align:left;"> {discrete, uniform, normal, lognormal} </td>
@@
 In the experiment conducted in @li2023plot, participants were allowed to make multiple selections for a lineup.
-The weighted detection rate is computed by assigning weights to each detection.
-If the participant selects zero plots, the corresponding weight is 0.05; otherwise, if the true residual plot is detected, the weight is 1 divided by the number of selections.
-With the weighted detection rate, we can evaluate the quality of the distance measure purposed in this chapter, by using the $\delta$-difference statistic.
-The $\delta$-difference is originally defined by @chowdhury2018measuring as
-
-\begin{equation}
-\delta = \bar{d}_{true} - \underset{j}{max}\left(\bar{d}_{null}^{(j)}\right) \quad \text{for}~j = 1,...,m-1,
-\end{equation}
-
-where $\bar{d}_{null}^{(j)}$ is the mean distance between the $j$-th null plot and other null plots, $\bar{d}_{true}$ is the mean distance between the true residual plot and null plots, and $m$ is the number of plots in a lineup.
-These mean distances are used because the distances considered in @chowdhury2018measuring can change if a different data plot is used to be compared against to.
-For example, suppose there are three null plots, A, B and C, the distance between null plot A and B could be different to the distance between null plot A and C.
-To get a consistent distance value for null plot A, it is necessary to take the average.
-However, this is not applicable to the distance proposed in this chapter, as we only allow the residual plot to be compared to a theoretically good residual plot.
-Therefore, the statistic needs to be modified to evaluate our distance measure.
-
-There is one aspect that the $\delta$-difference was intended to capture, which is the empirical distribution of distance for null plots.
-If we directly replace the mean distances with $D_{\text{null}}$, then the distribution will be a degenerate distribution, as $D_{\text{null}} = 0$ by definition.
-Moreover, $D_{\text{null}}$ can not be derived from an image, so it does not belong to the class of distances considered in @chowdhury2018measuring.
-The one that we should actually concerns about is the empirical distribution of $\hat{D}$ since it affects decision making.
-Thus, the adjusted $\delta$-difference is defined as
-
-\begin{equation}
-\delta_{adjusted} = \hat{D} - \underset{j}{max}\left(\hat{D}_{null}^{(j)}\right) \quad \text{for}~j = 1,...,m-1,
-\end{equation}
-
-\noindent where $\hat{D}_{null}^{(j)}$ is the estimated distance for the $j$-th null plot, and $m$ is the number of plots in a lineup.
+The weighted detection rate was computed by assigning weights to each detection.
+If the participant selected zero plots, a weight of 0.05 was assigned; otherwise, if the true residual plot was detected, the weight was 1 divided by the number of selections.
+This weighted detection rate allow us to assess the quality of the distance measure purposed in this chapter, by using the $\delta$-difference statistic.
+The $\delta$-difference is originally defined by @chowdhury2018measuring, is given by
+
+$$
+\delta = \bar{d}_{\text{true}} - \underset{j}{\text{max}}\left(\bar{d}_{\text{null}}^{(j)}\right) \quad \text{for}~j = 1,...,m-1,
+$$
+
+where $\bar{d}_{\text{null}}^{(j)}$ is the mean distance between the $j$-th null plot and the other null plots, $\bar{d}_{\text{true}}$ is the mean distance between the true residual plot and null plots, and $m$ is the number of plots in a lineup.
+These mean distances are used because, as noted by @chowdhury2018measuring, the distances can vary depending on which data plot is used for comparison.
+For instance, with three null plots, A, B and C, the distance between A and B may differ from the distance between A and C.
+To obtain a consistent distance for null plot A, averaging is necessary.
+However, this approach is not applicable to the distance proposed in this chapter, as we only compare the residual plot against a theoretically good residual plot.
+Consequently, the statistic must be adjusted to evaluate our distance measure effectively.
+
+One important aspect that the $\delta$-difference was designed to capture is the empirical distribution of distances for null plot.
+If we were to replace the mean distances $\bar{d}_{\text{null}}^{(j)}$ directly with $D_{\text{null}}^{(j)}$, the distance of the $j$-th null plot, the resulting distribution would be degenerate, since $D_{null}$ equals zero by definition.
+Additionally, $D$ can not be derived from an image, meaning it falls outside the scope of the distances considered by @chowdhury2018measuring.
+Instead, the focus should be on the empirical distribution of $\hat{D}$, as it influences decision-making.
+Therefore, the adjusted $\delta$-different is defined as
+
+$$
+\delta_{\text{adj}} = \hat{D} - \underset{j}{\text{max}}\left(\hat{D}_{\text{null}}^{(j)}\right) \quad \text{for}~j = 1,...,m-1,
+$$
+
+\noindent where $\hat{D}_{\text{null}}^{(j)}$ is the estimated distance for the $j$-th null plot, and $m$ is the number of plots in a lineup.
 
@@
   ylab("Weighted detection rate") +
-  xlab(expression(delta[adjusted])) +
+  xlab(expression(delta[adj])) +
   theme_light()

### 2024-08-11 Thesis/03-chap3.qmd
fix citation linking issue in pdf and linesep issue in table 3.6

@@
 
-We use a computer vision model to estimate this distance measure with a residual plot. Let $D$ be the result of @eq-kl-0, and our estimator $\hat{D}$ is formulated as
+We use a computer vision model to estimate this distance measure for a residual plot. Let $D$ be the result of @eq-kl-0, and our estimator $\hat{D}$ is formulated as
 
@@
                     escape = FALSE,
+                    linesep = "", 
                     col.names = c("Violations", "\\#samples", "RMSE")) %>%
@@
 
-#### Model performance on the human-evaluated data
+#### Model Performance on the Human-evaluated Data
 
@@
 
-### Boston housing
+### Boston Housing
 

### 2024-08-11 Thesis/03-chap3.qmd
try to fix alignment issue with chap 3

@@
   <tr>
-   <th style="text-align:left;"> <span data-qmd="Degree of Model Violations"></span> </th>
+   <th style="text-align:center;"> <span data-qmd="Degree of Model Violations"></span> </th>
    <th style="text-align:center;"><span data-qmd="Range ($C$ = 10)"></span>  </th>
@@
                     col.names = c("Degree of model violations", "Range ($C$ = 10)"),
-                    align = "lc")
+                    align = "lc") %>%
+  kableExtra::row_spec(0, align = "c")
 ```
@@
   <tr>
-   <th style="text-align:left;"> Factor </th>
-   <th style="text-align:left;"> Domain </th>
+   <th style="text-align:center;"> Factor </th>
+   <th style="text-align:center;"> Domain </th>
   </tr>
@@
   kableExtra::kable(escape = FALSE,
-                    booktabs = TRUE)
+                    booktabs = TRUE) %>%
+  kableExtra::row_spec(0, align = "c")
 ```
@@
   <tr>
-   <th style="text-align:left;"> Hyperparameter </th>
-   <th style="text-align:left;"> Domain </th>
+   <th style="text-align:center;"> Hyperparameter </th>
+   <th style="text-align:center;"> Domain </th>
   </tr>
@@
            Domain = c("{4, 8, 16, 32, 64}", "[0.1, 0.6]", "{false, true}", "{max, average}", "{false, true}", "{128, 256, 512, 1024, 2048}", "{false, true}", "[0.1, 0.6]", "[$10^{-8}$, $10^{-1}$]")) %>%
-  kableExtra::kable(booktabs = TRUE, escape = FALSE)
+  kableExtra::kable(booktabs = TRUE, escape = FALSE) %>%
+  kableExtra::row_spec(0, align = "c")
 ```
@@
   <tr>
-   <th style="text-align:left;"> Hyperparameter </th>
-   <th style="text-align:left;"> <span data-qmd="$32 \times 32$"> </span> </th>
-   <th style="text-align:left;"> <span data-qmd="$64 \times 64$"> </span> </th>
-   <th style="text-align:left;"> <span data-qmd="$128 \times 128$"> </span> </th>
+   <th style="text-align:center;"> Hyperparameter </th>
+   <th style="text-align:center;"> <span data-qmd="$32 \times 32$"> </span> </th>
+   <th style="text-align:center;"> <span data-qmd="$64 \times 64$"> </span> </th>
+   <th style="text-align:center;"> <span data-qmd="$128 \times 128$"> </span> </th>
   </tr>
@@
                     escape = FALSE) %>%
-  kableExtra::kable_styling(latex_options = c("scale_down"))
+  kableExtra::kable_styling(latex_options = c("scale_down")) %>%
+  kableExtra::row_spec(0, align = "c")
 ```
@@
                     escape = FALSE,
-                    col.names = c("", "RMSE", "$R^2$", "MAE", "Huber loss"))
+                    align = "lrrrr",
+                    col.names = c("", "RMSE", "$R^2$", "MAE", "Huber loss")) %>%
+  kableExtra::row_spec(0, align = "c")
 ```
@@
   <tr>
-   <th style="text-align:left;">  </th>
-   <th style="text-align:left;"> RMSE </th>
-   <th style="text-align:left;"> <span data-qmd="$R^2$"></span> </th>
-   <th style="text-align:left;"> MAE </th>
-   <th style="text-align:left;"> Huber loss </th>
+   <th style="text-align:center;">  </th>
+   <th style="text-align:center;"> RMSE </th>
+   <th style="text-align:center;"> <span data-qmd="$R^2$"></span> </th>
+   <th style="text-align:center;"> MAE </th>
+   <th style="text-align:center;"> Huber loss </th>
   </tr>
@@
    <td style="text-align:left;padding-left: 2em;" indentlevel="1"> <span data-qmd="$32 \times 32$"></span> </td>
-   <td style="text-align:left;"> 0.660 </td>
-   <td style="text-align:left;"> 0.901 </td>
-   <td style="text-align:left;"> 0.434 </td>
-   <td style="text-align:left;"> 0.181 </td>
+   <td style="text-align:right;"> 0.660 </td>
+   <td style="text-align:right;"> 0.901 </td>
+   <td style="text-align:right;"> 0.434 </td>
+   <td style="text-align:right;"> 0.181 </td>
   </tr>
@@
    <td style="text-align:left;padding-left: 2em;" indentlevel="1"> <span data-qmd="$64 \times 64$"></span> </td>
-   <td style="text-align:left;"> 0.674 </td>
-   <td style="text-align:left;"> 0.897 </td>
-   <td style="text-align:left;"> 0.438 </td>
-   <td style="text-align:left;"> 0.186 </td>
+   <td style="text-align:right;"> 0.674 </td>
+   <td style="text-align:right;"> 0.897 </td>
+   <td style="text-align:right;"> 0.438 </td>
+   <td style="text-align:right;"> 0.186 </td>
   </tr>
@@
    <td style="text-align:left;padding-left: 2em;" indentlevel="1"> <span data-qmd="$128 \times 128$"></span> </td>
-   <td style="text-align:left;"> 0.692 </td>
-   <td style="text-align:left;"> 0.892 </td>
-   <td style="text-align:left;"> 0.460 </td>
-   <td style="text-align:left;"> 0.199 </td>
+   <td style="text-align:right;"> 0.692 </td>
+   <td style="text-align:right;"> 0.892 </td>
+   <td style="text-align:right;"> 0.460 </td>
+   <td style="text-align:right;"> 0.199 </td>
   </tr>
@@
   <tr>
-   <th style="text-align:left;"> Violations </th>
-   <th style="text-align:left;"> #samples </th>
-   <th style="text-align:left;"> RMSE </th>
+   <th style="text-align:center;"> Violations </th>
+   <th style="text-align:center;"> #samples </th>
+   <th style="text-align:center;"> RMSE </th>
   </tr>
@@
    <td style="text-align:left;"> no violations </td>
-   <td style="text-align:left;"> 155 </td>
-   <td style="text-align:left;"> 1.267 </td>
+   <td style="text-align:right;"> 155 </td>
+   <td style="text-align:right;"> 1.267 </td>
   </tr>
@@
    <td style="text-align:left;"> non-linearity </td>
-   <td style="text-align:left;"> 2218 </td>
-   <td style="text-align:left;"> 0.787 </td>
+   <td style="text-align:right;"> 2218 </td>
+   <td style="text-align:right;"> 0.787 </td>
   </tr>
@@
    <td style="text-align:left;"> heteroskedasticity </td>
-   <td style="text-align:left;"> 1067 </td>
-   <td style="text-align:left;"> 0.602 </td>
+   <td style="text-align:right;"> 1067 </td>
+   <td style="text-align:right;"> 0.602 </td>
   </tr>
@@
    <td style="text-align:left;"> non-linearity + heteroskedasticity </td>
-   <td style="text-align:left;"> 985 </td>
-   <td style="text-align:left;"> 0.751 </td>
+   <td style="text-align:right;"> 985 </td>
+   <td style="text-align:right;"> 0.751 </td>
   </tr>
@@
    <td style="text-align:left;"> non-normality </td>
-   <td style="text-align:left;"> 1111 </td>
-   <td style="text-align:left;"> 0.320 </td>
+   <td style="text-align:right;"> 1111 </td>
+   <td style="text-align:right;"> 0.320 </td>
   </tr>
@@
    <td style="text-align:left;"> non-linearity + non-normality </td>
-   <td style="text-align:left;"> 928 </td>
-   <td style="text-align:left;"> 0.600 </td>
+   <td style="text-align:right;"> 928 </td>
+   <td style="text-align:right;"> 0.600 </td>
   </tr>
@@
    <td style="text-align:left;"> heteroskedasticity + non-normality </td>
-   <td style="text-align:left;"> 819 </td>
-   <td style="text-align:left;"> 0.489 </td>
+   <td style="text-align:right;"> 819 </td>
+   <td style="text-align:right;"> 0.489 </td>
   </tr>
@@
    <td style="text-align:left;"> non-linearity + heteroskedasticity + non-normality </td>
-   <td style="text-align:left;"> 717 </td>
-   <td style="text-align:left;"> 0.620 </td>
+   <td style="text-align:right;"> 717 </td>
+   <td style="text-align:right;"> 0.620 </td>
   </tr>
@@
                     linesep = "", 
+                    align = "lrr",
                     col.names = c("Violations", "\\#samples", "RMSE")) %>%
-  kableExtra::kable_styling(latex_options = c("scale_down"))
+  kableExtra::kable_styling(latex_options = c("scale_down")) %>%
+  kableExtra::row_spec(0, align = "c")
 ```
@@
   <tr>
-   <th style="text-align:left;"> Violation </th>
-   <th style="text-align:left;"> RMSE </th>
-   <th style="text-align:left;"> <span data-qmd="$R^2$"></span> </th>
-   <th style="text-align:left;"> MAE </th>
-   <th style="text-align:left;"> Huber loss </th>
+   <th style="text-align:center;"> Violation </th>
+   <th style="text-align:center;"> RMSE </th>
+   <th style="text-align:center;"> <span data-qmd="$R^2$"></span> </th>
+   <th style="text-align:center;"> MAE </th>
+   <th style="text-align:center;"> Huber loss </th>
   </tr>
@@
    <td style="text-align:left;"> heteroskedasticity </td>
-   <td style="text-align:left;"> 0.721 </td>
-   <td style="text-align:left;"> 0.852 </td>
-   <td style="text-align:left;"> 0.553 </td>
-   <td style="text-align:left;"> 0.235 </td>
+   <td style="text-align:right;"> 0.721 </td>
+   <td style="text-align:right;"> 0.852 </td>
+   <td style="text-align:right;"> 0.553 </td>
+   <td style="text-align:right;"> 0.235 </td>
   </tr>
@@
    <td style="text-align:left;"> non-linearity </td>
-   <td style="text-align:left;"> 0.738 </td>
-   <td style="text-align:left;"> 0.770 </td>
-   <td style="text-align:left;"> 0.566 </td>
-   <td style="text-align:left;"> 0.246 </td>
+   <td style="text-align:right;"> 0.738 </td>
+   <td style="text-align:right;"> 0.770 </td>
+   <td style="text-align:right;"> 0.566 </td>
+   <td style="text-align:right;"> 0.246 </td>
   </tr>
@@
                     escape = FALSE,
-                    col.names = c("Violation", "RMSE", "$R^2$", "MAE", "Huber loss"))
+                    align = "lrrrr",
+                    col.names = c("Violation", "RMSE", "$R^2$", "MAE", "Huber loss")) %>%
+  kableExtra::row_spec(0, align = "c")
 ```
@@
   <tr>
-   <th style="text-align:left;"> Violations </th>
-   <th style="text-align:left;"> #Samples </th>
-   <th style="text-align:left;"> #Agreements </th>
-   <th style="text-align:left;"> Agreement rate </th>
+   <th style="text-align:center;"> Violations </th>
+   <th style="text-align:center;"> #Samples </th>
+   <th style="text-align:center;"> #Agreements </th>
+   <th style="text-align:center;"> Agreement rate </th>
   </tr>
@@
    <td style="text-align:left;padding-left: 2em;" indentlevel="1"> heteroskedasticity </td>
-   <td style="text-align:left;"> 540 </td>
-   <td style="text-align:left;"> 464 </td>
-   <td style="text-align:left;"> 0.8593 </td>
+   <td style="text-align:right;"> 540 </td>
+   <td style="text-align:right;"> 464 </td>
+   <td style="text-align:right;"> 0.8593 </td>
   </tr>
@@
    <td style="text-align:left;padding-left: 2em;" indentlevel="1"> non-linearity </td>
-   <td style="text-align:left;"> 576 </td>
-   <td style="text-align:left;"> 459 </td>
-   <td style="text-align:left;"> 0.7969 </td>
+   <td style="text-align:right;"> 576 </td>
+   <td style="text-align:right;"> 459 </td>
+   <td style="text-align:right;"> 0.7969 </td>
   </tr>
@@
    <td style="text-align:left;padding-left: 2em;" indentlevel="1"> heteroskedasticity </td>
-   <td style="text-align:left;"> 540 </td>
-   <td style="text-align:left;"> 367 </td>
-   <td style="text-align:left;"> 0.6796 </td>
+   <td style="text-align:right;"> 540 </td>
+   <td style="text-align:right;"> 367 </td>
+   <td style="text-align:right;"> 0.6796 </td>
   </tr>
@@
    <td style="text-align:left;padding-left: 2em;" indentlevel="1"> non-linearity </td>
-   <td style="text-align:left;"> 576 </td>
-   <td style="text-align:left;"> 385 </td>
-   <td style="text-align:left;"> 0.6684 </td>
+   <td style="text-align:right;"> 576 </td>
+   <td style="text-align:right;"> 385 </td>
+   <td style="text-align:right;"> 0.6684 </td>
   </tr>
@@
                     escape = FALSE,
+                    align = "lrrr",
                     col.names = c("Violations", "\\#Samples", "\\#Agreements", "Agreement rate")) %>%
   kableExtra::pack_rows("Compared with conventional tests", 1, 2) %>%
-  kableExtra::pack_rows("Compared with visual tests conducted by human", 3, 4)
+  kableExtra::pack_rows("Compared with visual tests conducted by human", 3, 4) %>%
+  kableExtra::row_spec(0, align = "c")
 ```

### 2024-08-11 Thesis/03-chap3.qmd
update the compiled version

@@
 ```{r fig-delta, fig.pos = "!h", fig.cap = "A weighted detection rate vs adjusted $\\delta$-difference plot.
-The brown line is smoothing curve produced by fitting gnealized additive models", fig.height = 4}
+The brown line is smoothing curve produced by fitting gnealized additive models.", fig.height = 4}
 experiment %>%

### 2024-08-11 Thesis/03-chap3.qmd
fix legend issue in data examples

@@
 (B) The attention map produced by computing the gradient of the output with respect to the greyscale input.
-(C) The density plot of estimated distnace for null plots and bootstrapped plots.
+(C) The density plot of estimated distance for null plots and bootstrapped plots.
 The green area indicates the distribution of estimated distances for bootstrapped plots, while the yellow area represents the distribution of estimated distances for null plots.
@@
   xlab(expression(hat(D))) +
-  scale_linetype(labels = c("95% quantile of the null distribution", expression(observed~hat(D)))) +
+  scale_linetype(labels = c(expression(Q[null](0.95)), expression(observed~hat(D)))) +
   scale_fill_manual(values = c("#40B0A6", "#E1BE6A")) +
@@
 (B) The attention map produced by computing the gradient of the output with respect to the greyscale input.
-(C) The density plot of estimated distnace for null plots and bootstrapped plots.
+(C) The density plot of estimated distance for null plots and bootstrapped plots.
 The blue area indicates the distribution of estimated distances for bootstrapped plots, while the yellow area represents the distribution of estimated distances for null plots.

### 2024-08-18 Thesis/03-chap3.qmd
finish chap 4 a

@@
-# Automated Assessment of Residual Plots with Computer Vision Models
+# Automated Assessment of Residual Plots with Computer Vision Models {#sec-second-paper}
 
@@
 However, accurately interpreting a residual plot can be challenging.
-A residual plot can exhibit various visual features, but it is crucial to recognize that some may arise from the characteristics of predictors and the natural stochastic variation of the observational unit, rather than indicating a violation of model assumptions [@li2023plot].
+A residual plot can exhibit various visual features, but it is crucial to recognize that some may arise from the characteristics of predictors and the natural stochastic variation of the observational unit, rather than indicating a violation of model assumptions [@li2024plot].
 Consider @fig-false-finding as an example, the residual plot displays a triangular left-pointing shape.
@@
 The fundamental idea involves testing whether the true residual plot visually differs significantly from null plots, where null plots are plotted with residuals generated from the residual rotation distribution [@langsrud2005rotation], which is a distribution consistent with the null hypothesis $H_0$ that the linear regression model is correctly specified.
-Typically, the visual test is accomplished through the lineup protocol, where the real residual plot is embedded within a lineup alongside several null plots.
-If the real residual plot can be distinguished from the lineup, it provides evidence for rejecting $H_0$.
+Typically, the visual test is accomplished through the lineup protocol, where the true residual plot is embedded within a lineup alongside several null plots.
+If the true residual plot can be distinguished from the lineup, it provides evidence for rejecting $H_0$.
 
@@
 
-A practical limitation of the lineup protocol lies in its reliance on human judgements [see @li2023plot about the practical limitations].
+A practical limitation of the lineup protocol lies in its reliance on human judgements [see @li2024plot about the practical limitations].
 Unlike conventional statistical tests that can be performed computationally in statistical software, the lineup protocol requires human evaluation of images.
@@
 The computation assumes knowledge of the data generating processes for predictors and response variables.
-Since these processes are often unknown in practice, we will discuss a method to estimate this distance using a computer vision model in Section @sec-model-distance-estimation.
+Since these processes are often unknown in practice, we will discuss a method to estimate this distance using a computer vision model in @sec-model-distance-estimation.
 
@@
 
-This distance measure was first proposed in @li2023plot.
+This distance measure was first proposed in @li2024plot.
 It was mainly designed for measuring the effect size of non-linearity and heteroskedasticity in a residual plot.
-@li2023plot have derived that, for a classical normal linear regression model that omits necessary higher-order predictors $\boldsymbol{Z}$ and the corresponding parameter $\boldsymbol{\beta}_z$, and incorrectly assumes $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)$ while in fact $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \boldsymbol{V})$ where $\boldsymbol{V}$ is an arbitrary symmetric positive semi-definite matrix, $Q$ can be represented as $N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, \text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}))$.
+@li2024plot have derived that, for a classical normal linear regression model that omits necessary higher-order predictors $\boldsymbol{Z}$ and the corresponding parameter $\boldsymbol{\beta}_z$, and incorrectly assumes $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)$ while in fact $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \boldsymbol{V})$ where $\boldsymbol{V}$ is an arbitrary symmetric positive semi-definite matrix, $Q$ can be represented as $N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, \text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}))$.
 Note that the variance-covariance matrix is replaced with the diagonal matrix to ensure it is a full-rank matrix.
@@
 For non-normal error $\boldsymbol{\varepsilon}$, the true residual distribution $P$ is unlikely to be a multivariate normal distribution.
-Thus, @eq-kl-2 given in @li2023plot will not be applicable to models violating the normality assumption.
+Thus, @eq-kl-2 given in @li2024plot will not be applicable to models violating the normality assumption.
 
@@
 
-Note that if more observations are used for estimating the linear regression, the result of @eq-kl-1 will increase, as the integration will be performed over a higher-dimensional space.
+Note that if more observations are used for estimating the linear regression, the result of @eq-kl-1 will increase, as the integration will be performed with larger $n$.
 For a given data generating process, $D$ typically increases logarithmically with the number of observations.
@@
 If $\beta_1 \neq 0$, $\boldsymbol{x}_2$ was also included in the design matrix.
-This data generation process was adapted from @li2023plot, where it was utilized to simulate residual plots exhibiting non-linearity and heteroskedasticity visual patterns for human subject experiments.
+This data generation process was adapted from @li2024plot, where it was utilized to simulate residual plots exhibiting non-linearity and heteroskedasticity visual patterns for human subject experiments.
 A summary of the factors utilized in @eq-data-sim is provided in @tbl-factor.
@@
 
-In order to check the validity of the proposed computer vision model, residual plots presented in the human subject experiment conducted by @li2023plot will be assessed.
+In order to check the validity of the proposed computer vision model, residual plots presented in the human subject experiment conducted by @li2024plot will be assessed.
 
@@
 
-In @li2023plot, the residual plots are simulated from a data generating process which is a special case of @eq-data-sim.
+In @li2024plot, the residual plots are simulated from a data generating process which is a special case of @eq-data-sim.
 The main characteristic is the model violations are introduced separately, meaning non-linearity and heteroskedasticity will not co-exist in one lineup but assigned uniformly to all lineups.
@@
 
-For each lineup used in @li2023plot, there is one true residual plot and 19 null plots.
+For each lineup used in @li2024plot, there is one true residual plot and 19 null plots.
 While the distance $D$ for the true residual plot depends on the underlying data generating process, the distance $D$ for the null plots is zero.
@@
 
-In the experiment conducted in @li2023plot, participants were allowed to make multiple selections for a lineup.
+In the experiment conducted in @li2024plot, participants were allowed to make multiple selections for a lineup.
 The weighted detection rate was computed by assigning weights to each detection.

### 2024-08-18 Thesis/04-chap4.qmd
fix for emi's comment up to 3.9
finish chap 4 a

--- 2024-08-05 Thesis/04-chap4.qmd
@@
-# Developing Efficient Tools for Automated Residual Plot Assessment: autovi and autovi.web
+# Tools for Automated Residual Plot Assessment: autovi and autovi.web
 
-## Introduction {#sec-web-introduction}
+## Introduction {#sec-autovi-introduction}
 
-Most statistical software offers graphical tools for diagnosing regression models.
-For example, the base R package `stats` [@stats] automatically generates four diagnostic plots for linear models (`lm` objects): a residuals vs.
-fitted values plot, a Scale-Location plot (square root of absolute residuals vs.
-fitted values), a normal Q-Q plot (citation here), and a residuals vs.
-leverage plot.
-Alternatively, packages hosted on The Comprehensive R Archive Network (CRAN) such as `jtools` [@jtools] provide a more comprehensive and aesthetically enhanced set of diagnostic plots.
-However, very few of them offer automated assessment for these plots.
-Users often need to manually inspect and interpret the plots, which can lead to inconsistent conclusions due to varying levels of domain and statistical knowledge.
-This variability makes it challenging to achieve standardized results from plot diagnostics.
+Regression analysis is a fundamental statistical technique widely used in various fields.
+In modern practice, this analysis is often performed using specialized statistical software, with R being a popular choice among researchers and data analysts.
+The Comprehensive R Archive Network (CRAN) [@hornik2012comprehensive] hosts a vast array of packages, many of which provide graphical tools for diagnosing residuals in linear regression models.
+These packages can be broadly categorized into three groups.
 
-The R package `autovi` was developed to address this issue by performing visual statistical testing on residual plots using computer vision.
-To make its algorithm and trained computer vision model widely accessible, we also developed a web-based tool called `autovi.web`.
-This tool provides a user-friendly web interface that allows users to diagnose their residual plots without needing to install any dependencies required by the `autovi` package.
+General-purpose regression analysis tools are the largest and most commonly used group.
+These packages aren't specifically designed for graphical diagnostics of residuals in linear regression but offer this functionality as part of a broader set of statistical tools.
+A prime example is R's built-in `stats` package [@stats], which provides a comprehensive collection of statistical modelling tools that includes common diagnostic plots like residuals vs fitted values, quantile-quantile (Q-Q) plots, and residuals vs leverage plots.
+Other packages such as `jtools` [@jtools], `olsrr` [@olsrr], `rockchalk` [@rockchalk], and `ggResidpanel` [@ggresidpanel] provide similar graphical tools with alternative aesthetic styles or interactive features.
+Although these packages may differ in presentation, they all fundamentally deliver diagnostic plots based on well-established principles in regression analysis, as outlined in classic works like @cook1982residuals.
+However, consistently drawing accurate conclusions from these tools can be challenging due to individual differences in interpreting statistical graphics.
+As noted in @li2024plot, relying solely on subjective assessments of data plots can lead to problems, such as over-interpreting random patterns as model violations.
 
-This paper documents the design of the R package `autovi` and the web interface `autovi.web`. The rest of the paper is structured as follows: ...
+Enhanced visual diagnostics is the second group, which offers advanced visual aids for interpreting diagnostic plots.
+A notable example is the `DHARMa` package [@dharma], which uses an innovative approach by fitting quantile regression on scaled residual plots.
+It compares the empirical 0.25, 0.5, and 0.75 quantiles in scaled residuals with their theoretical counterparts, making it particularly useful for highlighting deviations from model assumptions, detecting model violations such as heteroscedasticity and incorrect functional forms, and uncovering issues specific to generalized linear models and mixed-effect models, like over/underdispersion.
+By offering these enhanced visualizations, `DHARMa` enables users to more easily identify potential issues in their regression models that might not be immediately apparent with standard diagnostic plots.
+In summary, this group of packages enhances the interpretability of diagnostic plots by drawing attention to critical elements such as trends, clusters, and outliers.
+They sometimes automatically perform conventional tests on these elements, displaying the results as annotations, labels, or text within the plot, thereby further reducing the likelihood of misinterpretation.
 
-## Background
+Statistical testing for visual discoveries is the third group, which focuses on providing tools for conducting formal statistical tests for visual discoveries obtained from diagnostic plots [@buja2009statistical].
+Examples in this category include the `nullabor` [@nullabor] and `regressinator` [@regressinator] packages, which enables users to quantify the significance of patterns observed in residual diagnostic plots, perform hypothesis tests on specific aspects of model fit, and validate visual interpretations with statistical evidence.
+This approach addresses the issue of inconsistent interpretation of diagnostic plots by bridging the gap between visual inspection and formal statistical inference, thus offering a more robust framework for regression diagnostics.
 
-- current stage of residual diagnostics (optional)
-- brief introduction to visual inference
-- brief overview of the second paper
-- a sentence or two about why we want to develop autovi and autovi.web
+Conducting a visual test for a residual plot, which is among the most common diagnostic plots in regression analysis, involves using a lineup protocol.
+In this protocol, the true residual plot is embedded within a lineup of several null plots and presented to one or more observers.
+The null plots are created by simulating residuals consistent with the null hypothesis $H_0$ that the regression model is correctly specified.
+Observers are then asked to identify the plot that appears most different from the others.
+If a significant percentage of observers correctly identify the true residual plot, it provides evidence against the $H_0$, as according to @buja2009statistical, the true residual plot would have no distinguishable difference from the null plots if all residuals are generated by the same process.
 
-## autovi
+However, as discussed in @sec-second-paper, the lineup protocol has significant limitations in large-scale applications due to its high labor costs and time consumption.
+To address these issues, we developed a computer vision model and an associated statistical testing procedure to automate the assessment of residual plots.
+This model takes a residual plot and a vector of auxiliary variables (such as the number of observations) as inputs and outputs a visual signal strength.
+This strength estimates the distance between the residual distribution of the fitted regression model and the reference distribution assumed under correct model specification.
 
-The primary goal of `autovi` is to provide rejection decisions and $p$-values for testing the null hypothesis that a regression model is correctly specified.
-While the classical normal linear regression model is popular and widely used, it represents a limited class of models.
-Therefore, `autovi` is designed to be highly modular and customizable, allowing it to handle a variety of residual diagnostic tasks.
+By estimating the visual signal strength for all plots in a lineup, we can compute a $p$-value based on the ratio of plots with visual signal strength greater than or equal to that of the true residual plot.
+This $p$-value has a lower bound of one divided by the number of plots in the lineup.
+Additionally, we can construct a null distribution of visual signal strength using the strengths of null plots.
 
-### Quick Start
+We can also apply bootstrapping to obtain a distribution of visual signal strengths.
+This involves bootstrapping the data used to fit the linear regression model, refitting the model to obtain bootstrapped residuals, and then using the computer vision model to predict visual signal strength for these residuals.
+The resulting bootstrapped distribution can be compared against the null distribution.
+If these distributions are largely similar, it suggests that the bootstrapped residual plots are similar to the null plots.
+The proportion of bootstrapped visual signal strengths exceeding a critical value (such as the 95% sample quantile of the null distribution) indicates how often the assumed regression model would be considered incorrect if the data could be repeatedly obtained from the same data-generating process.
+
+To make the statistical testing procedure and trained computer vision model widely accessible, we developed the R package `autovi`.
+In addition, we created a web-based tool that offers a user-friendly interface, enabling users to diagnose their residual plots without the need to install any dependencies required by the `autovi` package.
+
+The remainder of this chapter is structured as follows: @sec-autovi provides a detailed documentation of the `autovi` package, including its usage and workflow.
+@sec-autovi-web focuses on the `autovi.web` interface, describing its design and usage, along with illustrative examples.
+Finally, @sec-autovi-conclusion presents the main conclusions of this work.
+
+## autovi {#sec-autovi}
+
+The main purpose of `autovi` is to provide rejection decisions and $p$-values for testing whether a regression model is correctly specified.
+The package introduces a novel approach to automating statistical analysis, particularly in the interpretation of residual plots.
+The name `autovi` stands for automated visual inference.
+While initially designed for linear regression residual diagnostics, it has the potential to be extended to broader visual inference applications, as we'll discuss in section @sec-autovi-workflow.
+
+### Implementation
+
+`autovi` is built upon the `bandicoot` object-oriented programming (OOP) system [@bandicoot], which marks a departure from R's conventional S3 generic system.
+The adoption of an OOP architecture enhances flexibility and modularity, enabling users to redefine key functions within the workflow through method overriding.
+While similar functionality could be replicated using R's S3 system with generic functions, the OOP system offers a more structured and extensible foundation for the package.
+
+The workflow of `autovi` demonstrates the effective integration of multiple programming languages and libraries to create a comprehensive analytical tool.
+It depends on five core libraries from Python and R, each contributing critically to the analysis pipeline.
+In Python, `pillow` [@clark2015pillow] handles image processing tasks, including reading PNG files of residual plots, resizing them, and converting them into input tensors for further analysis.
+The `TensorFlow` [@abadi2016tensorflow] library, a cornerstone of contemporary machine learning, is employed to predict the visual signal strength of these residual plots,  utilizing a pre-trained convolutional neural network.
+
+Within the R environment, `autovi` utilizes several powerful libraries.
+`ggplot2` [@ggplot2] is employed to generate the initial residual plots, which are then saved as PNG files, using as the primary visual input for the analysis.
+The `cassowaryr` [@mason2022cassowaryr] library calculates scagnostics (scatter plot diagnostics) of the residual plots, providing numerical features that capture various statistical properties of the plots.
+These scagnostics complement the visual analysis by supplying quantitative metrics as secondary input to the computer vision model.
+The `reticulate` [@reticulate] package is used to bridge R and Python, allowing for seamless communication between the two languages and supporting an integrated workflow.
+
+The package includes internal functions to check the current Python environment used by the `reticulate` package.
+If the necessary Python packages are not installed in the Python interpreter, an error will be raised.
+If you want to select a specific Python environment, you can do so by calling the `reticulate::use_python()` function before using the `autovi` package.
+
+### Installation
+
+The `autovi` package is available on CRAN.
+It is actively developed and maintained, with the latest updates accessible on GitHub at [https://github.com/TengMCing/autovi](https://github.com/TengMCing/autovi).
+The code discussed in this chapter is based on `autovi` version 0.4.1.
+
+### Quick Start {#sec-autovi-quick-start}
+
+To get started quickly, users need only five lines of code to obtain a summary of the automated residual assessment:
 
@@
 
-```{r message=TRUE}
-checker
+```{r message = TRUE}
+print(checker)
 ```
 
-```{r}
+1. Load the package using the `library()` function.
+2. Construct a checker with two inputs: a linear regression model and a pre-trained Keras model [@chollet2015keras].
+3.
+Use `get_keras_model()`, a function provided by `autovi`, to download a trained computer vision model (described in @sec-second-paper) from GitHub.
+"vss_phn_32" specifies a model that predicts visual signal strength (vss) and is trained on residuals with polynomial, heteroskedasticity, and non-normality patterns (phn).
+More details about the hosted models will be provided in section @sec-trained-model-hosting.
+4.
+Call the `check()` method of the checker with default arguments.
+This predicts the visual signal strength for the true residual plot, 100 null plots, and 100 bootstrapped plots, storing the predictions internally.
+5. Use the `print()` function to generate a concise report of the check results.
+
+The report highlights key findings such as the visual signal strength of the true residual plot and the $p$-value of the automated visual test.
+The $p$-value is the ratio of null plots having visual signal strength greater than or equal to the true residual plot.
+We typically reject the null hypothesis when the $p$-value is smaller than or equal to 5%.
+The report also provides sample quantiles of visual signal strength for null and bootstrapped plots, helping to explain the severity and likelihood of model violations.
+
+Although the $p$-value is sufficient for automated decision-making, users are strongly encouraged to visually inspect the original residual plot alongside a sample null plot.
+This visual comparison can clarify why $H_0$ is either rejected or not, and help identify potential remedies.
+The `plot_pair()` method facilitates this comparison.
+
+```{r echo = TRUE, fig.height=3, fig.width=4}
+checker$plot_pair()
+```
+
+This method displays the true residual plot on the left and a null plot on the right.
+Users should look for any distinct visual patterns in the true residual plot that are absent in the null plot.
+It's recommended to run this function multiple times to confirm any visual findings, as each execution generates a new random null plot for comparison.
+
+The package offers a straightforward visualization of the assessment result through the `summary_plot()` function.
+
+```{r echo=TRUE}
 checker$summary_plot()
@@
 
-### Modularized Workflow
+In the visualization, the blue area represents the density of visual signal strength for null residual plots, while the red area shows the density for bootstrapped residual plots.
+The dashed line indicates the visual signal strength of the true residual plot, and the solid line marks the critical value at a 95% significance level.
+The $p$-value and the likelihood ratio are displayed in the subtitle.
+The likelihood ratio represents the ratio of the likelihood of observing the visual signal strength of the true residual plot from the bootstrapped distribution compared to the null distribution.
+
+Interpreting the plot involves several key aspects.
+If the dashed line falls to the right of the solid line, it suggests rejecting the null hypothesis.
+The degree of overlap between the red and blue areas indicates similarity between the true residual plot and null plots; greater overlap suggests more similarity.
+Lastly, the portion of the red area to the right of the solid line represents the percentage of bootstrapped models considered to have model violations.
+
+This visual summary provides an intuitive way to assess the model's fit and potential violations, allowing users to quickly grasp the results of the automated analysis.
+
+### Modularized Workflow {#sec-autovi-workflow}
 
@@
 
-The workflow of `autovi` begins with the user providing a regression model and concludes with the delivery of visual signal strength and p-value to the user.
-An overview diagram of the workflow is provided in @fig-autovi-diag.
+The initial motivation for developing `autovi` was to create a convenient interface for sharing the models described and trained in @sec-second-paper.
+However, recognizing that the classical normal linear regression model represents a restricted class of models, we sought to avoid limiting the potential for future extensions, whether by the original developers or other users.
+As a result, the package was designed to function seamlessly with linear regression models with minimal modification and few required arguments, while also accommodating other classes of models through partial workflow replacement.
+This modular and customizable design allows `autovi` to handle a wide range of residual diagnostics tasks.
 
-The workflow comprises two primary inputs, namely "Regression Model" and "Keras Model", and 10 core modules, including "Data Extraction", "Bootstrapping and Model Refitting", "Fitted Values and Residuals Extraction", "Auxiliary Computation", "Null Residual Simulation", "Plotting", "Plot Saving", "Image Reading and Resizing", "Visual Signal Strength Prediction" and "P-value Computation".
-Each module is designed to make minimal assumptions about the preceding modules, allowing for user customizations without impacting the overall workflow.
-The "Visual Signal Strength Prediction" and "P-value Computation" modules are predefined methods and can not be overriden, though users can interact with them directly via function arguments.
-The "Image Reading and Resizing" module is also unchangeable but it will adapt to different Keras models by checking their input shapes.
-The remaining seven modules are designed to be overriadble, accommodating users' specific needs.
-We will discuss these modules in details in the following sections.
+The workflow of `autovi` consists of ten core modules: data extraction, bootstrapping and model refitting, fitted values and residuals extraction, auxiliary computation, null residual simulation, plotting, plot saving, image reading and resizing, visual signal strength prediction, and $p$-value computation.
+Each module is designed with minimal dependency on the preceding modules, allowing users to customize parts of the workflow without affecting its overall integrity.
+An overview of this workflow is illustrated in Figure @fig-autovi-diag.
+
+The modules for visual signal strength prediction and $p$-value computation are predefined and cannot be overridden, although users can interact with them directly through function arguments.
+Similarly, the image reading and resizing module is fixed but will adapt to different Keras models by checking their input shapes.
+The remaining seven modules are designed to be overridable, enabling users to tailor the workflow to their specific needs.
+These modules will be discussed in detail in the following sections.
 
@@
 
-To manage workflow state and enable customizable modules, the `autovi` workflow is implemented using the `bandicoot` object-oriented programming framework.
-This framework allows for seamless integration of user-supplied data, methods, and results, storing them all within a single cohesive object.
+An `autovi` checker can be initialized by supplying two primary inputs, including a regression model object, such as an `lm` object representing the result of a linear regression model, and a trained computer vision model compatible with the `Keras` [@chollet2015keras] Application Programming Interface (API), to the `AUTO_VI` class constructor `auto_vi()`.
+The input will be stored in the checker and can be accessed by the user through the `$` operator.
 
-An `autovi` checker can be initialized by supplying two primary inputs, including a regression model object, such as an `lm` object representing the result of a linear regression model, and a trained computer vision model compatible with the `Keras` Application Programming Interface (API).
-
-The corresponding function for this step is `auto_vi` as used in the following example code.
-The `get_keras_model` is a helper function for retrieving trained `Keras` model from the web.
-We will further discuss it in @sec-trained-model-hosting.
-
-```{r echo=TRUE, message=TRUE}
+```{r echo=TRUE}
 library(autovi)
@@
                    keras_model = get_keras_model("vss_phn_32"))
-checker
 ```
@@
 
+After initializing the object, you can print the checker to view its status.
+
+```{r echo = TRUE, message = TRUE}
+checker
+```
+
+The status includes the list of regression model classes (as provided by the built-in `class()` function), the input and output shapes of the Keras model in the standard `Numpy` format [@harris2020array], the output node index being monitored, and the assessment result.
+If no check has been run yet, the assessment result will display as "UNKNOWN".
+
 #### Fitted Values and Residuals Extraction
 
-To create a residual plot, both fitted values and residuals are required.
-By convention, statistical models in R, such as `lm` (linear model) and `glm` (generalized linear model), support the use of the generic functions `fitted` and `resid` for extracting these values.
-The "Fitted Values and Residuals Extraction" module uses the same set of generic functions by default.
-However, since generic functions only work with classes that have appropriate method implementations, users can override the `get_fitted_and_resid` method if their model class does not support these generic functions.
+To be able to predict visual signal strength for a residual plot, both fitted values and residuals are needed to be extracted from the regression model object supplied by the user.
+In R, statistical models like `lm` (linear model) and `glm` (generalized linear model) typically support the use of generic functions such as `fitted()` and `resid()` to retrieve these values.
+The `get_fitted_and_resid()` method, called by the checker, relies on these generic functions by default.
+However, generic functions only work with classes that have appropriate method implementations.
+Some regression modelling packages may not fully adhere to the `stats` package guidelines for implementing these functions.
+In such cases, overriding the method becomes necessary.
 
-The following code example will returns a `tibble` of two columns: `.fitted` and `.resid` for fitted values and residuals, respectively
+By design, the `get_fitted_and_resid()` method accepts a regression model object as input and returns a `tibble` with two columns: `.fitted` and `.resid`, representing the fitted values and residuals, respectively.
+If no input is supplied, the method uses the regression model object stored in the checker.
+Although modules in the `autovi` workflow make minimal assumptions about other modules, they do require strictly defined input and output formats to ensure data validation and prevent fatal bugs.
+Therefore, any overridden method should follow to these conventions.
 
@@
 
-The result of the module will be passed to "Auxiliary Computation" for computing auxiliary inputs of computer vision model, and "Plotting" modules for generating residual plots.
-
 #### Data Extraction
 
-The "Data Extraction" module involves extracting the model frame from the model object.
-This is typically done using the `model.frame` generic function, as implemented in the default method `get_data`.
-Alternatively, users can provide the data used for fitting the regression model via the `data` argument when constructing the checker, or they can override the method to suit their needs.
-It is worth noting that this module is not necessary if bootstrapping is not required by the user, as the model frame is not used in other steps.
+For linear regression model in R, the model frame contains all the data required by a formula for evaluation.
+This is essential for bootstrapping and refitting the model when constructing a bootstrapped distribution of visual signal strength.
+Typically, the model frame can be extracted from the regression model object using the `model.frame()` generic function, which is the default method used by `get_data()`.
+However, some regression models don't use a formula or are evaluated differently, potentially lacking a model frame.
+In such cases, users can either provide the data used to fit the regression model through the `data` argument when constructing the checker, or customize the method to better suit their needs.
+It's worth noting that this module is only necessary if bootstrapping is required, as the model frame is not used in other steps of the workflow.
 
-The following code example will returns a `data.frame` representing the model frame of the fitted regression model
+The `get_data()` method accepts a regression model object as input and returns a `data.frame` representing the model frame of the fitted regression model.
+If no input is supplied, the regression model stored in the checker will be used.
 
 ```{r echo=TRUE}
-checker$get_data() |> head()
+checker$get_data() |> 
+  head()
 ```
-
-The result of the module will be passed to "Bootstrapping and Model Refitting" for getting bootstrapped regression models.
 
@@
 
+Bootstrapping a regression model typically involves sampling the observations with replacement and refitting the model with the bootstrapped data.
+The `boot_method()` method follows this bootstrapping scheme by default.
+It accepts a fitted regression model and a `data.frame` as inputs, and returns a `tibble` of bootstrapped residuals.
+If no inputs are provided, the method uses the regression model stored in the checker and the result of the `get_data()` method.
+
+Note that instead of calling `get_data()` implicitly within the method, it is used as part of the default argument definition.
+This approach allows users to bypass the `get_data()` method entirely and directly supply a `data.frame` to initiate the bootstrap process.
+Many other methods in `autovi` adopt this principle when possible, where dependencies are explicitly listed in the formal arguments.
+This design choice enhances the reusability and isolation of modules, offers better control for testing, and simplifies the overall process.
+
+```{r echo=TRUE}
+checker$boot_method(data = checker$get_data())
+```
+
 #### Auxiliary Computation
 
-Sometimes, a residual plot alone may not be sufficient for determining visual signal strength.
-For example, when the residual plot has significant overlapping, the trend and shape of the residual pattern can be difficult to identify.
-It can be helpful to include auxiliary variables, such as the number of observations, as inputs to the computer vision model.
-`autovi` addresses this need with internal functions built into the checker that automatically detect the number of inputs required by the provided Keras model.
-If multiple inputs are needed, the checker invokes the `auxiliary` method to compute auxiliary inputs.
-By default, this method calculates four scagnostics, "Monotonic", "Sparse", "Striped", and "Splines, using the `cassowaryr` package, as well as the number of observations.
-This approach aligns with the training process of our trained computer vision models
+According to @sec-second-paper, in some cases, a residual plot alone may not provide enough information to accurately determine visual signal strength.
+For instance, when the residual plot has significant overlap, the trend and shape of the residual pattern can be difficult to discern.
+Including auxiliary variables, such as the number of observations, as additional inputs to the computer vision model can be beneficial.
+To address this, `autovi` includes internal functions within the checker that automatically detect the number of inputs required by the provided Keras model.
+If multiple inputs are necessary, the checker invokes the `auxiliary()` method to compute these additional inputs.
 
-The following code example returns a `data.frame` with five columns: four for the scagnostics and one for the number of observations
+The `auxiliary()` method takes a `data.frame` containing fitted values and residuals as input and returns a `data.frame` with five numeric columns.
+These columns represent four scagnostics — "Monotonic", "Sparse", "Striped", and "Splines" — calculated using the `cassowaryr` package, as well as the number of observations.
+This approach is consistent with the training process of the computer vision models described in @sec-second-paper.
+If no `data.frame` is provided, the method will default to retrieving fitted values and residuals by calling `get_fitted_and_resid()`.
+
+Technically, any Keras-implemented computer vision model can be adapted to accept an image as the primary input and additional variables as secondary inputs by adding a data pre-processing layer before the actual input layer.
+If users wish to override `auxiliary()`, the output should be a `data.frame` with a single row and the number of columns matching the supplied Keras model.
 
@@
 
-The result of this module will be passed to the Keras model as part of the inputs.
+#### Null Residual Simulation {#sec-autovi-null-method}
 
-#### Null Residual Simulation
-
-The "Null Residual Simulation" module, as its name suggests, is responsible for simulating null residuals consistent with the model assumptions.
-For linear regression models, this assumes the model is correctly specified and simulates random draws from the residual rotation distribution [@buja2009statistical].
-Other types of regression models, such as `glm` (generalized linear model) and `gam` (generalized additive model), generally can not use the same method to efficiently simulate null residuals.
-Therefore, it is highly recommended that users override the `null_method` for this module to suit their specific model.
-
-The following code example returns a `tibble` of two columns: `.fitted` and `.resid` for fitted values and null residuals, respectively
+A fundamental element of the automated residual assessment described in @sec-second-paper is comparing the visual signal strength of null plots with that of the true residual plot.
+However, due to the variety of regression models, there is no universal method for simulating null residuals that are consistent with model assumptions.
+Fortunately, for classical normal linear regression models, null residuals can be effectively simulated using the residual rotation method, as outlined in @buja2009statistical.
+This process involves generating random draws from a standard normal distribution, regressing these draws on the original predictors, and then rescaling the resulting residuals by the ratio of the residual sum of squares to the that of the original linear regression model.
+Other regression models, such as `glm` (generalized linear model) and `gam` (generalized additive model), generally cannot use this method to efficiently simulate null residuals.
+Therefore, it is recommended that users override the `null_method()` to suit their specific model.
+The `null_method()` takes a fitted regression model as input, defaulting to the regression model stored in the checker, and returns a `tibble`.
 
@@
 
-The result of this module will be passed to the "Plotting" module for generating residual plots.
-
 #### Plotting
@@
 There are computer vision models trained to handle images captured in various scenarios.
-For example, the VGG16 model can classify objects in images taken under different lighting conditions and is robust to image rotation.
+For example, the VGG16 model [@simonyan2014very] can classify objects in images taken under different lighting conditions and is robust to image rotation.
 However, data plots are a special type of image as the plotting style can always be consistent if controlled properly.
@@
 In this case, it is best to predict plots using the same style for optimal performance.
-The "Plotting" module handles this aspect.
-By default, it replicates the style we used to generate the training samples for the computer vision models.
+The plotting method `plot_resid()` handles this aspect.
+
+`plot_resid()` accepts a `data.frame` containing fitted values and residuals, along with several customization options: a `ggplot` theme, an `alpha` value to control the transparency of data points, a `size` value to set the size of data points, and a `stroke` value to define the thickness of data point edges.
+Additionally, it includes four Boolean arguments to toggle the display of axes, legends, grid lines, and a horizontal red line.
+By default, it replicates the style we used to generate the training samples for the computer vision models described in @sec-second-paper.
 In brief, the residual plot omits axis text and ticks, titles, and background grid lines, featuring only a red line at $y = 0$.
-Essentially, it retains only the necessary components of a residual plot.
-If the computer vision model is trained with a different but consistent plot style, the corresponding method `plot_resid` should be overridden.
+It retains only the necessary components of a residual plot.
+If the computer vision model is trained with a different but consistent aesthetic style, `plot_resid()` should be overridden.
 
-The following code example generates a `ggplot` object with the described aesthetic style
+The method returns a `ggplot` object, which can be saved as a PNG file in the following module.
+If no data is provided, the method will use `get_fitted_and_resid()` to retrieve the fitted values and residuals from the regression model stored in the checker.
 
-```{r echo=TRUE}
+```{r echo=TRUE, fig.height=3, fig.width=4}
+checker$plot_resid()
+```
+
+To manually generate true residual plots, null plots, or bootstrapped residual plots, you can pass the corresponding `data.frame` produced by the `get_fitted_and_resid()`, `null_method()`, and `boot_method()` methods to the `plot_resid()` method, respectively.
+
+```{r echo = TRUE, fig.height=3, fig.width=4}
+checker$null_method() |>
   checker$plot_resid()
@@
 
-Another aspect of a standardized residual plot is its resolution.
-The original image format we used was 420 pixels in height and 525 pixels in width.
-This resolution was chosen because the original lineup, containing 20 residual plots in a 4x5 grid, is represented by an image of 2100 by 2100 pixels.
-If a different image size is required, the corresponding method `save_plot` can be overridden.
-
-The following code example saves a residual plot to a temporary file and returns the path to that file as a string
+Another key aspect of a standardized residual plot is its resolution.
+In @sec-second-paper, we used an image format of 420 pixels in height and 525 pixels in width.
+This resolution was chosen because the original set, consisting of 20 residual plots arranged in a four by five grid, was represented by an image of 2100 by 2100 pixels.
+The `save_plot()` method takes a `ggplot` object as input, saves it as a temporary PNG file, and returns the file path as a string.
+Note that the `save_plot()` method does not have default arguments, as it is not intended to be called without a plot.
+While an alternative design could be to save the true residual plot by default, this might be confusing for users, given that the method's name does not fully convey this functionality.
 
 ```{r echo=TRUE}
-checker$plot_resid() |> checker$save_plot()
+checker$plot_resid() |> 
+  checker$save_plot()
 ```
-
-The result of this module will be passed to the "Image Reading and Resizing" module to be converted to a Numpy array.
 
@@
 
-When training computer vision models, it is common practice to test multiple input sizes for the same architecture to find the optimized setup.
-This requires the original training image to be prepared at a higher resolution than needed and then resized to fit the input size during training.
-`autovi` includes a built-in class, `KERAS_WRAPPER`, to facilitate this task.
-This Keras wrapper class has a method, `image_to_array`, that can read an image as a `PIL` image using the `pillow` Python package, resize it to the target input size needed by the Keras model, and convert it to a `numpy` array.
+When training computer vision models, it is common to test various input sizes for the same architecture to identify the optimal setup.
+This involves preparing the original training image at a higher resolution than required and then resizing it to match the input size during training.
+The `autovi` package includes a class, `KERAS_WRAPPER`, to simplify this process.
+This Keras wrapper class features a method called `image_to_array()`, which reads an image as a `PIL` image using the `pillow` Python package, resizes it to the target input size required by the Keras model, and converts it to a `Numpy` array.
 
 To construct a `KERAS_WRAPPER` object, you need to provide the Keras model as the main argument.
-However, users typically do not need to use this class directly, as the `autovi` checker can automatically call its methods when performing visual signal strength predictions.
+However, users generally do not need to interact with this class directly, as the `autovi` checker automatically invokes its methods when performing visual signal strength predictions.
+The `image_to_array()` method takes the path to the image file, the target height, and the target width as inputs and returns a `Numpy` array.
+If not specified, the target height and target width will be retrieved from the input layer of the Keras model by the `get_input_height()` and `get_input_width()` method of `KERAS_WRAPPER`.
 
-The following code example reads an image of a residual plot and resizes it to 32 by 32 pixels to make it compatible with the computer vision model's input shape.
+The following code example demonstrate the way to manually generate the true residual plot, save it as PNG file, and load it back as `Numpy` array.
 
@@
 
-The result of this module will be passed to the "Visual Signal Strength Prediction" module as input.
-
 #### Visual Signal Strength Prediction
 
-Visual signal strength, as defined in (cite here), is an estimate of the distance between the input residual plot and a theoretically good residual plot.
-However, visual signal strength can be defined in various ways, similar to how distance between two points can be measured differently.
-The key is to compare the visual signal strength of the true residual plot with that of the null residual plots to determine if it significantly deviates from the null distribution.
+Visual signal strength, as discussed in @sec-second-paper, estimates the distance between the input residual plot and a theoretically good  residual plot.
+It can be defined in various ways, much like different methods for measuring the distance between two points.
+This will not impact the `autovi` workflow as long as the provided Keras model can predict the intended measure.
 
-There are multiple ways to obtain visual signal strength from the checker, with the most direct being the `vss` method.
-By default, it predicts the visual signal strength for the true residual plot.
-If a `ggplot` or a `data.frame`, such as null residuals generated by the `null_method`, is explicitly passed to this method, it will use that input to predict visual signal strength accordingly.
+There are several ways to obtain visual signal strength from the checker, with the most direct being the `vss()` method.
+By default, this method predicts the visual signal strength for the true residual plot.
+If a `ggplot` or a `data.frame`, such as null residuals generated by the `null_method()`, is explicitly provided, the method will use that input to predict visual signal strength accordingly.
+Note that if a `ggplot` is provided, auxiliary inputs must be supplied manually via the `auxiliary` argument, as we assume that auxiliary variables can not be computed directly from a `ggplot`.
 
-Another way to obtain visual signal strength is by calling the `check` and `lineup_check` methods.
-These comprehensive methods perform extensive diagnostics on the true residual plot and store the visual signal strength in the `check_result` field of the checker.
-Additionally, for obtaining visual signal strength for null residual plots and bootstrapped residual plots, there are two specialized methods, `null_vss` and `boot_vss`, designed for this purpose respectively.
+Another way to obtain visual signal strength is by calling the `check()` method.
+This comprehensive method perform extensive diagnostics on the true residual plot and store the visual signal strength in the `check_result` field of the checker.
+Additionally, for obtaining visual signal strength for null residual plots and bootstrapped residual plots, there are two specialized methods, `null_vss()` and `boot_vss()`, designed for this purpose respectively.
 
-The following code example predicts the visual signal strength for the residual plot of the fitted regression model
+Calling the `vss()` method without arguments will predict the visual signal strength for the true residual plot and return the result as a single-element `tibble`.
 
@@
 
-The following code example predicts the visual signal strength for five null residual plots
+Providing a `data.frame` of null residuals or a null residual plot yields the same visual signal strength.
 
 ```{r echo=TRUE}
-checker$null_vss(5L)
+null_resid <- checker$null_method()
+checker$vss(null_resid)
 ```
 
-The following code example predicts the visual signal strength for five bootstrapped residual plots
+```{r echo=TRUE}
+null_resid |>
+  checker$plot_resid() |>
+  checker$vss()
+```
+
+The `null_vss()` helper method primarily takes the number of null plots as input.
+If the user wants to use a ad hoc null simulation scheme, it can be provided via the `null_method` argument.
+Intermediate results, including null residuals and null plots, can be returned by enabling `keep_null_data` and `keep_null_plot`.
+The visual signal strength, along with null residuals and null plots, will be stored in a `tibble` with three columns.
+The following code example demonstrates how to predict the visual signal strength for five null residual plots while keeping the intermediate results.
+
+```{r echo=TRUE}
+checker$null_vss(5L, 
+                 keep_null_data = TRUE, 
+                 keep_null_plot = TRUE)
+```
+
+The `boot_vss()` helper method is similar to `null_vss()`, with some differences in argument names.
+The following code example demonstrates how to predict the visual signal strength for five bootstrapped residual plots while keeping the intermediate results.
 
 ```{r}
-checker$boot_vss(5L)
+checker$boot_vss(5L,
+                 keep_boot_data = TRUE,
+                 keep_boot_plot = TRUE)
 ```
 
-The result of the module will be passed to the "P-value Computation" module for calculating $p$-value and making a rejection decision.
+#### $P$-value Computation
 
-#### P-value Computation
-
-For a visual test conducted by a single observer using the lineup protocol, the minimum $p$-value is $1/m$, where $m$ is the number of residual plots in a lineup.
-This is because the Type-I error of this test is $1/m$.
-Similarly, when performing a lineup check via the `lineup_check` method, the checker generates $m−1$ null plots to combine with the true residual plot, forming a lineup.
-It then calculates the $p$-value as the ratio of plots with visual signal strength greater than or equal to that of the true plot.
-The $p$-value of this test is also bounded by $1/m$.
-
-Another way to perform a visual test is to assume a null distribution for visual signal strength and then estimate quantiles of this null distribution as critical values.
-The $p$-value in this case is the ratio of null plots with visual signal strength greater than or equal to that of the true residual plot.
-This method typically requires evaluating a large number of null residual plots, as done in the `check` method.
-
-The following code example performs a lineup check with $m = 20$.
-The `p_value` method will use the correct formula to compute $p$-value according to the check type.
-Note that the result is subject to randomness due to the small lineup size.
-
-```{r echo=TRUE}
-checker$lineup_check(lineup_size = 20L)
-checker$p_value()
-```
-
-The following code example evaluates 100 null residual plots and 100 bootstrapped residual plots, such that the null distribution and bootstrapped distribution of visual signal strength can be estimated accurately.
-It then computes the $p$-value and prints the checker to get detailed breakdown of the check result.
+Once we have obtained the visual signal strength from both the true residual plot and the null plots, we can compute the $p$-value.
+This $p$-value represents the ratio of plots with visual signal strength greater than or equal to that of the true residual plot.
+We can perform this calculation using the `check()` method.
+The main inputs for this method are the number of null plots and the number of bootstrapped plots to generate.
+If you need to access intermediate residuals and plots, you can enable the `keep_data` and `keep_plot` options.
+The method stores the final result in the `check_result` field of the object.
+To obtain the p-value using the `check()` method, you can use the following code.
 
@@
 checker$check(boot_draws = 100L, null_draws = 100L)
-checker$p_value()
+checker$check_result$p_value
 ```
+
+You can also check the $p$-value by printing the checker, which includes it in the summary report.
 
@@
 
+After executing the `check()` method, `autovi` offers two visualization options for the assessment result through the `summary_plot()` method, including the density plot and the rank plot.
+We have already discussed and interpreted the density plot in an earlier section.
+Here, we would like to highlight the flexibility in choosing which elements to display in the density plot.
+For instance, you can omit the bootstrapped distribution by setting `boot_dist` to `NULL`.
+Similarly, you can hide the null distribution (`null_dist`), the $p$-value (`p_value`), or the likelihood ratio (`likelihood_ratio`) as needed.
+The following example demonstrates how to create a summary plot without the results from bootstrapped plots.
+
 ```{r echo=TRUE}
-checker$check(boot_draws = 100L, null_draws = 100L)
-checker$summary_plot()
+checker$summary_plot(boot_dist = NULL,
+                     likelihood_ratio = NULL)
 ```
 
+This customization allows you to focus on specific aspects of the assessment, tailoring the visualization to your analytical needs.
+
+The rank plot, creating by setting `type` to "rank", is a bar plot where the x-axis represents the rank and the y-axis shows the visual signal strength.
+The bar for the true residual plot is colored in red.
+By examining the rank plot, you can intuitively understand how the observed visual signal strength compares to the null visual signal strengths and identify any outliers in the null distribution.
+
 ```{r echo=TRUE}
-checker$lineup_check(lineup_size = 20L)
-checker$summary_plot()
+checker$summary_plot(type = "rank")
 ```
 
-### Keras Model Wrapper
+### Feature Extraction
+
+In addition to predicting visual signal strength and computing $p$-values, `autovi` offers methods to extract features from any layer of the Keras model.
+To see which layers are available in the current Keras model, you can use the `list_layer_name()` method from the `KERAS_WRAPPER` class.
+
+The following code example lists the layer names of the currently used Keras model:
 
@@
 
+Among these layers, the "global_max_pooling2d" layer is a 2D global max pooling layer that outputs the results from the last convolutional blocks.
+As @simonyan2014very noted, all preceding convolutional blocks can be viewed as a large feature extractor.
+Consequently, the output from this layer provides features that can be utilized for various purposes, such as performing transfer learning.
+
+To obtain the features, provide the layer name using the `extract_feature_from_layer` argument in the `predict()` method.
+This will return a `tibble` with the visual signal strength and all features extracted from that layer.
+Each row corresponds to one plot.
+The features will be flattened into 2D and named with the prefix "f_" followed by a number from one to the total number of features.
+
 ```{r echo=TRUE}
@@
   wrapper$image_to_array() |>
-  wrapper$predict(auxiliary = checker$auxiliary())
+  wrapper$predict(auxiliary = checker$auxiliary(),
+                  extract_feature_from_layer = "global_max_pooling2d")
 ```
+
+Alternatively, the `AUTO_VI` class provides a way to extract features using the `vss()` method.
+This method is essentially a high-level wrapper around the `predict()` method of `KERAS_WRAPPER`, but it offers a more straightforward interface and better default arguments.
+
+The results from the previous code example can be replicated with a single line of code as shown below.
+
+```{r echo = TRUE}
+checker$vss(extract_feature_from_layer = "global_max_pooling2d")
+```
+
+The argument `extract_feature_from_layer` is also available in other functions that build on the `vss()` method, including `null_vss()`, `boot_vss()`, and `check()`.
+
+The package provides tools for analyzing these extracted features through the `feature_pca()` method and its associated visualization method, `feature_pca_plot()`.
+The `feature_pca()` method performs principal component analysis (PCA) on the features to reduce their dimensionality.
+However, it requires that a `check()` is performed first, as it relies on results stored in `check_result`.
+Alternatively, you can manually provide features using the `feature`, `null_feature`, and `boot_feature` arguments for the true residual plot, null plots, and bootstrapped plots, respectively.
+The `feature_pca()` method returns a tibble containing both the original features and the principal components.
+The rotation matrix and standard deviations of each principal component are stored as attributes.
+
+```{r echo = TRUE}
+checker$check(null_draws = 100L,
+              boot_draws = 100L,
+              extract_feature_from_layer = "global_max_pooling2d")
+checker$feature_pca()
+```
+
+The `feature_pca_plot()` method visualizes the results of the PCA.
+By default, it plots the first principal component on the x-axis and the second principal component on the y-axis, with points colored according to their origi, true residual plots, null residual plots, or bootstrapped residual plots.
+Users can customize the x and y axes by specifying symbols for the `x` and `y` arguments.
+Additionally, the `col_by_set` option can be disabled if you prefer not to use coloring.
+
+```{r echo = TRUE}
+checker$feature_pca_plot()
+```
+
+When interpreting the principal component scatter plot, look for any outliers within the null or bootstrapped groups.
+Assess whether the null group and the bootstrapped group form a single cluster or distinct clusters.
+Additionally, evaluate whether the observed point is distinct from the null group.
 
@@
 
-### Dependency Considerations
+The trained computer vision models described in @sec-second-paper are hosted on a GitHub repository at [https://github.com/TengMCing/autovi_data](https://github.com/TengMCing/autovi_data).
+Currently, there are six models available.
+You can view them by calling `list_keras_model()`, which will return a `tibble` showing the input shape and a description of each model.
 
-## autovi.web
+```{r}
+list_keras_model()
+```
+
+The `get_keras_model()` function can be used to download a model to a temporary directory and load it into memory using `TensorFlow`.
+It requires only the model name, which is the value in the first column of the `tibble` returned by `list_keras_model()`.
+
+### Extending the `AUTO_VI` class
+
+`bandicoot` is a lightweight object-oriented system with Python-like syntax that supports multiple inheritance and incorporates a Python-like method resolution order.
+The system is inspired by the OOP frameworks implemented in R6 [r6] and Python.
+In this section, we will provide essential details for extending the `autovi::AUTO_VI` class using `bandicoot`.
+
+In `bandicoot`, a class is declared using the `bandicoot::new_class()` function, where parent classes are provided as positional arguments, and the class name is specified through the `class_name` argument.
+The output of `bandicoot::new_class()` is an environment with the S3 class `bandicoot_oop`.
+Printing a `bandicoot` object provides a summary of the object, which can be customized via the `..str..` magic method.
+
+```{r}
+EXT_AUTO_VI <- bandicoot::new_class(AUTO_VI, class_name = "EXT_AUTO_VI")
+print(EXT_AUTO_VI)
+```
+
+An extended class inherits attributes and methods from its parent class(es), so it will behave similarly to them. This can be verified using the built-in `names()` function.
+
+```{r}
+names(EXT_AUTO_VI)
+```
+
+To register a method for an extended class, you need to pass the class as the first argument and the method as a named argument to the `bandicoot::register_method()` function.
+Within a method, `self` can be used as a reference to the class or object environment.
+The following code example overrides the `null_method()` with a function that simulates null residuals from the corresponding normal distribution.
+This approach differs from the default null residual simulation scheme described in @sec-autovi-null-method.
+Although less efficient than the default method for linear regression models, it provides an alternative way to simulate null residuals.
+This method is particularly useful when the fitted model is unavailable, and only the fitted values and residuals are accessible, as discussed in @sec-autovi-web.
+
+```{r}
+bandicoot::register_method(
+  EXT_AUTO_VI, 
+  null_method = function(fitted_model = self$fitted_model) {
+    data <- self$get_fitted_and_resid(fitted_model)
+    residual_sd <- sd(data$.resid)
+    data$.resid <- rnorm(nrow(data),
+                         sd = residual_sd)
+    return(data)
+  }
+)
+
+print(EXT_AUTO_VI$null_method(lm(dist ~ speed, data = cars)))
+```
+
+To create an object in `bandicoot`, you need to call the `instantiate()` method of a class.
+Alternatively, you can build a convenient class constructor for your class.
+It is recommended to provide the full list of arguments in the class constructor instead of using `...`, as this makes it easier for integrated development environments (IDEs) like RStudio to offer argument completion hints to the user.
+
+```{r}
+ext_auto_vi <- function(fitted_model,
+                        keras_model = NULL,
+                        data = NULL,
+                        node_index = 1L,
+                        env = new.env(parent = parent.frame()),
+                        init_call = sys.call()) {
+  EXT_AUTO_VI$instantiate(fitted_model = fitted_model,
+                          keras_model = keras_model,
+                          data = data,
+                          node_index = node_index,
+                          env = env,
+                          init_call = init_call)
+}
+
+print(ext_auto_vi(lm(dist ~ speed, data = cars)))
+```
+
+## autovi.web {#sec-autovi-web}
 

### 2024-08-19 Thesis/03-chap3.qmd
intro to chap 3 revised

@@
 
-Plotting residuals is a standard practice in linear regression diagnostics, essential for identifying deviations from model assumptions such as linearity, homoscedasticity, and normality.
-Visual inference provides an inferential framework to assess whether residual plots contain patterns inconsistent with model assumptions, typically using a lineup protocol.
-However, the lineup protocol's reliance on human judgment limits its scalability.
-This study addresses this limitation by automating the interpretation of residual plots using computer vision models.
-We develop a distance measure based on Kullback-Leibler divergence to quantify the disparity between the residual distribution of a fitted classical normal linear regression model and the reference distribution.
-We propose a computer vision model to estimate this distance from residual plots, facilitating formal statistical testing and bootstrapping techniques to assess model specification.
-Our computer vision model shows strong performance, though it performs slightly less effectively on non-linearity visual patterns.
-The statistical tests based on the estimated distance exhibit lower sensitivity than conventional tests but higher sensitivity than human visual tests.
-Examples demonstrate the method's effectiveness across different scenarios, highlighting its value in automating the diagnostic process and supplementing traditional methods.
+Plotting residuals is a recommended practice to diagnose deviations from linear model assumptions such as non-linearity, heteroscedasticity, and non-normality.
+Using the lineup protocol provides an inferential procedure to statistical assess presence or absence structure in residual plots.
+As a statistical test, the lineup protocol is less sensitive and applies more broadly than available conventional tests.
+However, the lineup protocol relies on human judgment which limits its scalability.
+This work presents a solution by providing a computer vision model to automate the assessment of residual plots.
+It is trained to predict the distance measure that quantifies the disparity between the residual distribution of a fitted classical normal linear regression model and the reference distribution, based on Kullback-Leibler divergence.
+From extensive simulation studies, the computer vision model exhibits lower sensitivity than conventional tests but higher sensitivity than human visual tests.
+It is slightly less effective on non-linearity patterns.
+Several examples from classical papers and contemporary data illustrate the new procedures, highlighting its usefulness in automating the diagnostic process and supplementing existing methods.
 

### 2024-08-19 Thesis/03-chap3.qmd
start of chapter 3 further revised

@@
 Plotting residuals is a recommended practice to diagnose deviations from linear model assumptions such as non-linearity, heteroscedasticity, and non-normality.
-Using the lineup protocol provides an inferential procedure to statistical assess presence or absence structure in residual plots.
+The presence or absence of structure in residual plots can be tested using the lineup protocol to do visual inference.
 As a statistical test, the lineup protocol is less sensitive and applies more broadly than available conventional tests.
@@
 Plotting residuals is commonly regarded as a standard practice in linear regression diagnostics [@belsley1980regression; @cook1982residuals].
-This visual assessment plays a crucial role in identifying deviations from model assumptions, such as linearity, homoscedasticity, and normality.
+This visual assessment plays a crucial role in identifying whether model assumptions, such as linearity, homoscedasticity, and normality, are reasonable.
 It also helps in understanding the goodness of fit and various unexpected characteristics of the model.

### 2024-08-19 Thesis/04-chap4.qmd
add start of abstract and revised first paragraph of intro

@@
 # Tools for Automated Residual Plot Assessment: autovi and autovi.web
+
+Regression software is widely available today, but tools for effective diagnostics are still lagging.
+Although it is advised to diagnose a linear model by plotting residuals, it required human effort which can be prohibit the efforts.
+Here we describe a new R package that includes a computer vision model for automated assessment of residual plots, and an accompanying shiny app for ease of use.
 
@@
 
-Regression analysis is a fundamental statistical technique widely used in various fields.
-In modern practice, this analysis is often performed using specialized statistical software, with R being a popular choice among researchers and data analysts.
-The Comprehensive R Archive Network (CRAN) [@hornik2012comprehensive] hosts a vast array of packages, many of which provide graphical tools for diagnosing residuals in linear regression models.
-These packages can be broadly categorized into three groups.
+Regression analysis is a fundamental statistical technique widely used for modeling data from many fields.
+In modern practice, software for regression analysis tools is widely.
+The Comprehensive R Archive Network (CRAN) [@hornik2012comprehensive] hosts a vast array of packages, many of which diagnosing models using residual plots.
+These packages can be broadly categorized into three groups: general purpose, enhanced diagnostics, diagnostics with statistical testing.
 

### 2024-08-21 Thesis/04-chap4.qmd
update chap 4

@@
 
-The remainder of this chapter is structured as follows: @sec-autovi provides a detailed documentation of the `autovi` package, including its usage and workflow.
+The remainder of this chapter is structured as follows: @sec-autovi provides a detailed documentation of the `autovi` package, including its usage and infrastructure.
 @sec-autovi-web focuses on the `autovi.web` interface, describing its design and usage, along with illustrative examples.
@@
 The name `autovi` stands for automated visual inference.
-While initially designed for linear regression residual diagnostics, it has the potential to be extended to broader visual inference applications, as we'll discuss in section @sec-autovi-workflow.
+While initially designed for linear regression residual diagnostics, it has the potential to be extended to broader visual inference applications, as we'll discuss in section @sec-autovi-infrastructure.
 
-### Implementation
+### Implementation {#sec-autovi-implementation}
 
 `autovi` is built upon the `bandicoot` object-oriented programming (OOP) system [@bandicoot], which marks a departure from R's conventional S3 generic system.
-The adoption of an OOP architecture enhances flexibility and modularity, enabling users to redefine key functions within the workflow through method overriding.
+The adoption of an OOP architecture enhances flexibility and modularity, enabling users to redefine key functions within the infrastructure through method overriding.
 While similar functionality could be replicated using R's S3 system with generic functions, the OOP system offers a more structured and extensible foundation for the package.
 
-The workflow of `autovi` demonstrates the effective integration of multiple programming languages and libraries to create a comprehensive analytical tool.
+The infrastructure of `autovi` demonstrates the effective integration of multiple programming languages and libraries to create a comprehensive analytical tool.
 It depends on five core libraries from Python and R, each contributing critically to the analysis pipeline.
@@
 These scagnostics complement the visual analysis by supplying quantitative metrics as secondary input to the computer vision model.
-The `reticulate` [@reticulate] package is used to bridge R and Python, allowing for seamless communication between the two languages and supporting an integrated workflow.
+The `reticulate` [@reticulate] package is used to bridge R and Python, allowing for seamless communication between the two languages and supporting an integrated infrastructure.
 
@@
 
-### Quick Start {#sec-autovi-quick-start}
+### Usage {#sec-autovi-quick-start}
 
@@
 
-### Modularized Workflow {#sec-autovi-workflow}
+### Modularized Infrastructure {#sec-autovi-infrastructure}
 
@@
 #| label: fig-autovi-diag
-#| fig-cap: "Diagram illustrating the workflow of the R package autovi.
+#| fig-cap: "Diagram illustrating the infrastructure of the R package autovi.
 The modules in green are primary inputs provided by users.
@@
 However, recognizing that the classical normal linear regression model represents a restricted class of models, we sought to avoid limiting the potential for future extensions, whether by the original developers or other users.
-As a result, the package was designed to function seamlessly with linear regression models with minimal modification and few required arguments, while also accommodating other classes of models through partial workflow replacement.
+As a result, the package was designed to function seamlessly with linear regression models with minimal modification and few required arguments, while also accommodating other classes of models through partial infrastructure substitution.
 This modular and customizable design allows `autovi` to handle a wide range of residual diagnostics tasks.
 
-The workflow of `autovi` consists of ten core modules: data extraction, bootstrapping and model refitting, fitted values and residuals extraction, auxiliary computation, null residual simulation, plotting, plot saving, image reading and resizing, visual signal strength prediction, and $p$-value computation.
-Each module is designed with minimal dependency on the preceding modules, allowing users to customize parts of the workflow without affecting its overall integrity.
-An overview of this workflow is illustrated in Figure @fig-autovi-diag.
+The infrastructure of `autovi` consists of ten core modules: data extraction, bootstrapping and model refitting, fitted values and residuals extraction, auxiliary computation, null residual simulation, plotting, plot saving, image reading and resizing, visual signal strength prediction, and $p$-value computation.
+Each module is designed with minimal dependency on the preceding modules, allowing users to customize parts of the infrastructure without affecting its overall integrity.
+An overview of this infrastructure is illustrated in Figure @fig-autovi-diag.
 
@@
 Similarly, the image reading and resizing module is fixed but will adapt to different Keras models by checking their input shapes.
-The remaining seven modules are designed to be overridable, enabling users to tailor the workflow to their specific needs.
+The remaining seven modules are designed to be overridable, enabling users to tailor the infrastructure to their specific needs.
 These modules will be discussed in detail in the following sections.
@@
 If no input is supplied, the method uses the regression model object stored in the checker.
-Although modules in the `autovi` workflow make minimal assumptions about other modules, they do require strictly defined input and output formats to ensure data validation and prevent fatal bugs.
+Although modules in the `autovi` infrastructure make minimal assumptions about other modules, they do require strictly defined input and output formats to ensure data validation and prevent fatal bugs.
 Therefore, any overridden method should follow to these conventions.
@@
 In such cases, users can either provide the data used to fit the regression model through the `data` argument when constructing the checker, or customize the method to better suit their needs.
-It's worth noting that this module is only necessary if bootstrapping is required, as the model frame is not used in other steps of the workflow.
+It's worth noting that this module is only necessary if bootstrapping is required, as the model frame is not used in other modules of the infrastructure.
 
@@
 It can be defined in various ways, much like different methods for measuring the distance between two points.
-This will not impact the `autovi` workflow as long as the provided Keras model can predict the intended measure.
+This will not impact the `autovi` infrastructure as long as the provided Keras model can predict the intended measure.
 
@@
 
-The R package `autovi` is designed to provide rejection decisions and $p$-values for testing the null hypothesis that a regression model is correctly specified.
-To construct a checker with `autovi`, one needs to supply a regression model object—typically an `lm` object representing the result of a linear regression model and a trained computer vision model compatible with the `Keras` API.
+In @sec-autovi-implementation, we discussed how `autovi` relies on several Python libraries, with a particularly strong dependency on `TensorFlow`.
+Managing a Python environment and correctly installing `TensorFlow` on a local machine can be challenging for many users.
+Moreover, `TensorFlow` is a massive library that undergoes continuous development, which inevitably leads to compatibility issues arising from differences in library versions.
+These challenges can create significant barriers for users who want to perform residual assessments with the `autovi` package.
 
-The regression model object is used to extract the fitted values and residuals for creating a residual plot.
-Additionally, a residual rotation technique is applied to the model object to generate null residuals, which are residuals consistent with the null hypothesis.
-For a linear regression model, this is conventionally achieved by simulating new random standard normal draws and using them as responses to refit the linear regression model.
+Recognizing these potential barriers, we were motivated to design and implement a web interface called `autovi.web`.
+This web-based solution offers several major advantages.
+First, it eliminates dependency issues, so users no longer need to struggle with complex Python environments or worry about installing and maintaining the correct versions of libraries.
+The web interface handles all these dependencies on the server side.
+Second, `autovi.web` lowers the entry barrier by being user-friendly and accessible to individuals who may not be familiar with R programming.
+This broadens the potential user base of `autovi`, allowing more researchers and analysts to benefit from its capabilities.
+Third, by running on a controlled server environment, `autovi.web` ensures a consistent experience for all users, regardless of their local machine setup.
+Additionally, the web interface can be updated centrally, ensuring that all users always have access to the latest features and improvements without needing to manage updates locally.
+Lastly, `autovi.web` offers cross-platform accessibility, allowing users to access it from any device with a web browser, increasing flexibility and convenience.
 
-Having null plots, which are residual plots consisting of null residuals and the original fitted values, is crucial for constituting a visual test.
-If the visual test were conducted by humans, a lineup consisting of $m-1$ null plots and one true residual plot would be presented to several observers.
-Observers would then be asked to select the plot they find most different out of the $m$ residual plots.
-If many observers correctly identify the true residual plot as the most different, it provides evidence for rejecting the null hypothesis that the model is correctly specified.
-This is because, under the null hypothesis, the true residual plot should be indistinguishable from the null plots.
+By providing this web interface, we aim to significantly reduce the technical hurdles associated with using `autovi`, making advanced residual assessment techniques more accessible to a wider audience of researchers and data analysts.
+This approach aligns with modern trends in data science tools, where web-based interfaces are increasingly used to make advanced analytical techniques more widely available.
 
-Instead of human observers, the visual test in `autovi` is performed by a computer vision model.
-This model is trained to report the visual signal strength of each individual plot in a lineup.
-The visual signal strength estimates the divergence of the empirical residual distribution from the ideal residual distribution, effectively measuring the degree of model violations.
-The higher the visual signal strength, the more evidence there is against the null hypothesis.
+### Implementation
 
-The computer vision model's training involves estimating this divergence or distance, which quantifies how much the residuals deviate from what is expected under a correctly specified model.
-More details about the mathematical derivation and the training process of the computer vision model can be found in the paper by Li et.al.
-(2024).
+`autovi.web` is a sophisticated web application built using the `shiny` and `shinydashboard` R packages.
+Hosted on the [shinyapps.io](https://www.shinyapps.io) domain, the application is accessible through any modern web browser, offering advantages such as scalability and ease of maintenance.
 
-Furthermore, the computer vision model used in `autovi` requires a fixed-size 4D tensor as input.
+In our initial planning for `autovi.web`, we considered implementing the entire web application using the `webr` framework, which would have allowed the entire application to run directly in the user's browser.
+However, this approach was not feasible at the time of writing this chapter.
+The reason is that one of the R packages `autovi` depends on, `splanes`, uses compiled Fortran code.
+Unfortunately, a working Emscripten version of this code, which would be required for `webr`, was not available.
+
+We also explored the possibility of implementing the web interface using frameworks built on other languages, such as Python.
+However, server hosting domains that natively support Python servers typically do not have the latest version of R installed.
+Calling R from Python is typically done using the `rpy2` Python library, but this approach can be awkward when dealing with language syntax related to non-standard evaluation, making it challenging to develop our application in this manner.
+Another option we considered was renting a server where we could have full control, such as those provided by cloud platforms like Google Cloud Platform (GCP) or Amazon Web Services (AWS).
+However, correctly setting up the server and ensuring a secure deployment requires significant expertise, which we did not possess at the time.
+Ultimately, we decided that the most pragmatic solution was to use the `shiny` and `shinydashboard` frameworks, which are well-established in the R community and offer a robust foundation for web application development.
+This approach allowed us to build `autovi.web` on top of a familiar and well-supported ecosystem, while still taking advantage of the flexibility and power of the underlying Python libraries used by `autovi`.
+
+The server-side configuration of `autovi.web` is carefully designed to support its functionality.
+Most required Python libraries, including `pillow` and `NumPy`, are pre-installed on the server.
+These libraries are seamlessly integrated into the Shiny application using the `reticulate` package, which provides a robust interface between R and Python.
+
+Due to the nature of shinyapps.io's resource allocation, the server enters a sleep mode during periods of inactivity, resulting in the clearing of the local Python virtual environment.
+Consequently, when the application "wakes up" for a new user session, these libraries need to be reinstalled.
+While this ensures a clean environment for each session, it may lead to slightly longer loading times for the first user after a period of inactivity.
+
+In contrast to `autovi`, `autovi.web` does not use the native Python version of `TensorFlow`.
+Instead, it leverages `TensorFlow.js`, a JavaScript library that allows the execution of machine learning models directly in the browser.
+This choice enables native browser execution, enhancing compatibility across different user environments, and shifts the computational load from the server to the client-side, allowing for better scalability and performance, especially when dealing with resource-intensive computer vision models on shinyapps.io.
+While `autovi` requires downloading pre-trained computer vision models from GitHub, these models in ".keras" file format are incompatible with `TensorFlow.js`.
+Therefore, we store the model weights in JSON files and include them as extra resources in the Shiny application.
+When the application initializes, `TensorFlow.js` rebuilds the computer vision model using these pre-stored weights.
+
+To allow communication between `TensorFlow.js` and other components of the Shiny application, the `shinyjs` R package is used.
+This package allows for the seamless integration of custom JavaScript code into the Shiny framework.
+The specialized JavaScript code for initializing `TensorFlow.js` and calling `TensorFlow.js` for visual signal strength prediction is deployed alongside the Shiny application as additional resources.
+
+<!-- Furthermore, the computer vision model used in `autovi` requires a fixed-size 4D tensor as input.
 The dimensions of this tensor are as follows: the first dimension represents the batch size, the second dimension represents the width of the image, the third dimension represents the height of the image, and the fourth dimension represents the number of channels.
@@
 This consistency is crucial for ensuring that the model can accurately interpret and analyze new data.
+-->
 
-A significant portion of our web interface is dedicated to managing this data pipeline.
+<!-- A significant portion of our web interface is dedicated to managing this data pipeline.
 This involves processing the user-provided data to generate input images that conform to the required format for the computer vision model.
 The pipeline ensures that the residual plots created from the user data match the aesthetics and format of the training data, enabling the model to provide accurate visual signal strength assessments.
+-->
 

### 2024-08-22 Thesis/05-chap5.qmd
skeleton of conclusion chapter

# Conclusion and future plans

The three pieces of work in this thesis focus on ...

## Contributions

The primary contributions of this research is ....

Principles of transparency and reproducible research have guided the work. The materials to reproduce this thesis are available at .... and software packages are:

- software 1
- software 2

...

## Future work

There are several directions that this work can be developed: alternative  computer vision models, and extensions to different types of visual diagnostics.

The limitations of the current are ... We have explored these other ways to make the model ... Some new possibilities are ... might address ...

Residual diagnostics are also used in more complex statistical models such as hierarchical, or temporal or spatial ... These require adaptations of residual calculations ...

Visual diagnostics are also common in Bayesian modeling to assess ... (eg Gelman, A. 2004 Exploratory data analysis for complex models. J. Comput. Graph. Stat. 13(4), 755–779.)

### 2024-08-22 Thesis/04-chap4.qmd
heading changes in chap 4
finish up to workflow, and cache chap 4

@@
-# Tools for Automated Residual Plot Assessment: autovi and autovi.web
+# Software for Automated Residual Plot Assessment: autovi and autovi.web
 
@@
 
-## autovi {#sec-autovi}
+## R package: autovi {#sec-autovi}
 
@@
 
-To get started quickly, users need only five lines of code to obtain a summary of the automated residual assessment:
+To get started quickly, users need only three lines of code to obtain a summary of the automated residual assessment:
 
@@
 library(autovi)
-checker <- auto_vi(fitted_model = lm(dist ~ speed, data = cars), 
-                   keras_model = get_keras_model("vss_phn_32"))
+checker <- residual_checker(fitted_model = lm(dist ~ speed, data = cars))
 checker$check()
-checker
 ```
@@
 library(autovi)
-checker <- auto_vi(fitted_model = lm(dist ~ speed, data = cars), 
-                   keras_model = get_keras_model("vss_phn_32"))
-checker$check()
+checker <- auto_vi(fitted_model = lm(dist ~ speed, data = cars))
 ```
 
-```{r message = TRUE}
-print(checker)
+```{r message=TRUE}
+if (!file.exists("cached_data/autovi_check_demo.rds")) {
+  checker$keras_model <- get_keras_model("vss_phn_32") 
+  checker$check()
+  saveRDS(checker$check_result, "cached_data/autovi_check_demo.rds")
+} else {
+  checker$check_result <- readRDS("cached_data/autovi_check_demo.rds")
+}
+
+checker
 ```
@@
 
-```{r echo=TRUE}
+```r
 library(autovi)
@@
 
-```{r echo=TRUE}
+```r
 wrapper <- keras_wrapper(keras_model = checker$keras_model)  
@@
 input_array$shape
+```
+
+```{r}
+# If the Keras model can not be loaded
+if (!file.exists("cached_data/autovi_check_demo.rds")) {
+  wrapper <- keras_wrapper(keras_model = checker$keras_model)  
+  input_array <- checker$plot_resid() |> 
+    checker$save_plot() |>
+    wrapper$image_to_array()
+  input_array$shape
+} else {
+  cat("(1, 32, 32, 3)")
+}
 ```
@@
 
-```{r echo=TRUE}
+```r
 checker$vss()
+```
+
+```{r}
+if (!file.exists("cached_data/autovi_check_demo_2.rds")) {
+  checker$keras_model <- get_keras_model("vss_phn_32") 
+  saveRDS(checker$vss(), "cached_data/autovi_check_demo_2.rds")
+  checker$vss()
+} else {
+  readRDS("cached_data/autovi_check_demo_2.rds")
+}
 ```
@@
 
-```{r echo=TRUE}
+```r
 null_resid <- checker$null_method()
@@
 
-```{r echo=TRUE}
+```{r}
+null_resid <- checker$null_method()
+
+if (!file.exists("cached_data/autovi_check_demo_3.rds")) {
+  checker$keras_model <- get_keras_model("vss_phn_32") 
+  saveRDS(checker$vss(null_resid), "cached_data/autovi_check_demo_3.rds")
+  checker$vss(null_resid)
+} else {
+  readRDS("cached_data/autovi_check_demo_3.rds")
+}
+```
+
+```r
 null_resid |>
@@
   checker$vss()
+```
+
+```{r}
+readRDS("cached_data/autovi_check_demo_3.rds")
 ```
@@
 
-```{r echo=TRUE}
+```r
 checker$null_vss(5L, 
@@
 
+```{r}
+if (!file.exists("cached_data/autovi_check_demo_4.rds")) {
+  checker$keras_model <- get_keras_model("vss_phn_32") 
+  result <- checker$null_vss(5L, 
+                             keep_null_data = TRUE, 
+                             keep_null_plot = TRUE)
+  saveRDS(result, "cached_data/autovi_check_demo_4.rds")
+  result
+} else {
+  readRDS("cached_data/autovi_check_demo_4.rds")
+}
+```
+
 The `boot_vss()` helper method is similar to `null_vss()`, with some differences in argument names.
@@
 
-```{r}
+```r
 checker$boot_vss(5L,
@@
                  keep_boot_plot = TRUE)
+```
+
+```{r}
+if (!file.exists("cached_data/autovi_check_demo_5.rds")) {
+  checker$keras_model <- get_keras_model("vss_phn_32") 
+  result <- checker$boot_vss(5L, 
+                             keep_boot_data = TRUE, 
+                             keep_boot_plot = TRUE)
+  saveRDS(result, "cached_data/autovi_check_demo_5.rds")
+  result
+} else {
+  readRDS("cached_data/autovi_check_demo_5.rds")
+}
 ```
@@
 
-```{r echo=TRUE}
+```r
 checker$check(boot_draws = 100L, null_draws = 100L)
+checker$check_result$p_value
+```
+
+```{r}
+if (!file.exists("cached_data/autovi_check_demo_6.rds")) {
+  checker$keras_model <- get_keras_model("vss_phn_32") 
+  checker$check(boot_draws = 100L, null_draws = 100L)
+  saveRDS(checker$check_result, "cached_data/autovi_check_demo_6.rds")
+} else {
+  checker$check_result <- readRDS("cached_data/autovi_check_demo_6.rds")
+}
+
 checker$check_result$p_value
@@
 
-```{r echo=TRUE}
+```r
 checker$plot_resid() |>
@@
 
+```{r}
+if (!file.exists("cached_data/autovi_check_demo_7.rds")) {
+  checker$keras_model <- get_keras_model("vss_phn_32") 
+  result <- checker$plot_resid() |>
+    checker$save_plot() |>
+    wrapper$image_to_array() |>
+    wrapper$predict(auxiliary = checker$auxiliary(),
+                    extract_feature_from_layer = "global_max_pooling2d")
+  saveRDS(result, "cached_data/autovi_check_demo_7.rds")
+} else {
+  readRDS("cached_data/autovi_check_demo_7.rds")
+}
+```
+
 Alternatively, the `AUTO_VI` class provides a way to extract features using the `vss()` method.
@@
 
-```{r echo = TRUE}
+```r
 checker$vss(extract_feature_from_layer = "global_max_pooling2d")
+```
+
+```{r}
+readRDS("cached_data/autovi_check_demo_7.rds")
 ```
@@
 
-```{r echo = TRUE}
+```r
 checker$check(null_draws = 100L,
@@
               extract_feature_from_layer = "global_max_pooling2d")
+checker$feature_pca()
+```
+
+```{r}
+if (!file.exists("cached_data/autovi_check_demo_8.rds")) {
+  checker$keras_model <- get_keras_model("vss_phn_32") 
+  checker$check(null_draws = 100L,
+                boot_draws = 100L,
+                extract_feature_from_layer = "global_max_pooling2d")
+  saveRDS(checker$check_result, "cached_data/autovi_check_demo_8.rds")
+} else {
+  checker$check_result <- readRDS("cached_data/autovi_check_demo_8.rds")
+}
 checker$feature_pca()
@@
 
-## autovi.web {#sec-autovi-web}
+## Web interface: autovi.web {#sec-autovi-web}
 
@@
 This broadens the potential user base of `autovi`, allowing more researchers and analysts to benefit from its capabilities.
-Third, by running on a controlled server environment, `autovi.web` ensures a consistent experience for all users, regardless of their local machine setup.
-Additionally, the web interface can be updated centrally, ensuring that all users always have access to the latest features and improvements without needing to manage updates locally.
+Third, the web interface can be updated centrally, ensuring that all users always have access to the latest features and improvements without needing to manage updates locally.
 Lastly, `autovi.web` offers cross-platform accessibility, allowing users to access it from any device with a web browser, increasing flexibility and convenience.
 
+`autovi.web` is available at [autoviweb.netlify.app](autoviweb.netlify.app).
+The implementation discussed in this chapter is based on `autovi.web` version 0.1.0.
 By providing this web interface, we aim to significantly reduce the technical hurdles associated with using `autovi`, making advanced residual assessment techniques more accessible to a wider audience of researchers and data analysts.
@@
 
-`autovi.web` is a sophisticated web application built using the `shiny` and `shinydashboard` R packages.
-Hosted on the [shinyapps.io](https://www.shinyapps.io) domain, the application is accessible through any modern web browser, offering advantages such as scalability and ease of maintenance.
+`autovi.web` is a web application built using the `shiny` [@shiny] and `shinydashboard` [@shinydashboard] R packages.
+Hosted on the [shinyapps.io](https://www.shinyapps.io) domain, the application is accessible through any modern web browser.Additionally, R packages `htmltools` [@htmltools] and `shinycssloaders` [@shinycssloaders] are used to render markdown document in shiny application, and add loading animation for shiny widgets, respectively.
 
-In our initial planning for `autovi.web`, we considered implementing the entire web application using the `webr` framework, which would have allowed the entire application to run directly in the user's browser.
+In our initial planning for `autovi.web`, we considered implementing the entire web application using the `webr` framework [@webr], which would have allowed the entire application to run directly in the user's browser.
 However, this approach was not feasible at the time of writing this chapter.
-The reason is that one of the R packages `autovi` depends on, `splanes`, uses compiled Fortran code.
-Unfortunately, a working Emscripten version of this code, which would be required for `webr`, was not available.
+The reason is that one of the R packages `autovi` depends on the R package `splancs` [@splancs], which uses compiled Fortran code.
+Unfortunately, a working Emscripten version of this package, which would be required for `webr`, was not available.
 
@@
 However, server hosting domains that natively support Python servers typically do not have the latest version of R installed.
-Calling R from Python is typically done using the `rpy2` Python library, but this approach can be awkward when dealing with language syntax related to non-standard evaluation, making it challenging to develop our application in this manner.
+Additionally, calling R from Python is typically done using the `rpy2` Python library, but this approach can be awkward when dealing with language syntax related to non-standard evaluation, making it challenging to develop our application in this manner.
 Another option we considered was renting a server where we could have full control, such as those provided by cloud platforms like Google Cloud Platform (GCP) or Amazon Web Services (AWS).
 However, correctly setting up the server and ensuring a secure deployment requires significant expertise, which we did not possess at the time.
-Ultimately, we decided that the most pragmatic solution was to use the `shiny` and `shinydashboard` frameworks, which are well-established in the R community and offer a robust foundation for web application development.
-This approach allowed us to build `autovi.web` on top of a familiar and well-supported ecosystem, while still taking advantage of the flexibility and power of the underlying Python libraries used by `autovi`.
+Ultimately, we decided that the most practical solution was to use the `shiny` and `shinydashboard` frameworks, which are well-established in the R community and offer a solid foundation for web application development.
 
@@
 Most required Python libraries, including `pillow` and `NumPy`, are pre-installed on the server.
-These libraries are seamlessly integrated into the Shiny application using the `reticulate` package, which provides a robust interface between R and Python.
+These libraries are integrated into the Shiny application using the `reticulate` package, which provides an interface between R and Python.
 
-Due to the nature of shinyapps.io's resource allocation, the server enters a sleep mode during periods of inactivity, resulting in the clearing of the local Python virtual environment.
+Due to the resource allocation policy of shinyapps.io, the server enters a sleep mode during periods of inactivity, resulting in the clearing of the local Python virtual environment.
 Consequently, when the application "wakes up" for a new user session, these libraries need to be reinstalled.
@@
 Instead, it leverages `TensorFlow.js`, a JavaScript library that allows the execution of machine learning models directly in the browser.
-This choice enables native browser execution, enhancing compatibility across different user environments, and shifts the computational load from the server to the client-side, allowing for better scalability and performance, especially when dealing with resource-intensive computer vision models on shinyapps.io.
+This choice enables native browser execution, enhancing compatibility across different user environments, and shifts the computational load from the server to the client-side.
+`TensorFlow.js` also offers better scalability and performance, especially when dealing with resource-intensive computer vision models on shinyapps.io.
+
 While `autovi` requires downloading pre-trained computer vision models from GitHub, these models in ".keras" file format are incompatible with `TensorFlow.js`.
-Therefore, we store the model weights in JSON files and include them as extra resources in the Shiny application.
+Therefore, we extract and store the model weights in JSON files and include them as extra resources in the Shiny application.
 When the application initializes, `TensorFlow.js` rebuilds the computer vision model using these pre-stored weights.
@@
 To allow communication between `TensorFlow.js` and other components of the Shiny application, the `shinyjs` R package is used.
-This package allows for the seamless integration of custom JavaScript code into the Shiny framework.
+This package allows calling custom JavaScript code within the Shiny framework.
 The specialized JavaScript code for initializing `TensorFlow.js` and calling `TensorFlow.js` for visual signal strength prediction is deployed alongside the Shiny application as additional resources.
 
-<!-- Furthermore, the computer vision model used in `autovi` requires a fixed-size 4D tensor as input.
-The dimensions of this tensor are as follows: the first dimension represents the batch size, the second dimension represents the width of the image, the third dimension represents the height of the image, and the fourth dimension represents the number of channels.
-The model outputs a numeric vector that represents the visual signal strength for each image in the batch.
-The computer vision model is also trained with a set of fixed-aesthetic residual plots, which means that the input images must be produced using the same data pipeline that was used for the training data preparation.
-This consistency is crucial for ensuring that the model can accurately interpret and analyze new data.
--->
+### Design {#sec-autovi-web-design}
 
-<!-- A significant portion of our web interface is dedicated to managing this data pipeline.
-This involves processing the user-provided data to generate input images that conform to the required format for the computer vision model.
-The pipeline ensures that the residual plots created from the user data match the aesthetics and format of the training data, enabling the model to provide accurate visual signal strength assessments.
--->
+```{r fig-autovi-web, out.width = "100%", fig.cap = "Overview of the `autovi.web` graphical user interface (GUI).
+This default view may change based on user interactions.
+Region 1 is the sidebar menu, containing the residual assessment tab and the information tab.
+Region 2 is the data upload panel, where users can provide a CSV file and specify the type of data it contains.
+Region 3 includes dropdown menus for selecting the columns to be analyzed, a slider to control the number of bootstrapping samples, and a numeric input box for setting the simulation seed.
+Region 4 displays the initialization status and offers a button to start the analysis.
+Region 5 is empty in the default view but will be populated with results once the analysis is started.
+"}
+knitr::include_graphics("figures/autovi_web.png")
+```
 
-<!-- Our web interface simplifies this process for the user by automating the necessary steps to transform their data into the appropriate input format.
-Users can upload their CSV files, and the interface handles the extraction of residuals, the creation of residual plots, and the formatting of these plots into 4D tensors.
-This seamless integration allows users to focus on interpreting the results rather than on the technical details of data preparation.
--->
+While the R package `autovi` aims to provide tools that can be extended to broader visual inference applicaitons, `autovi.web` is only focus on to for provdinign a straightforward and clean uesr interface.
+An overview of the grahpical user interface of `autovi.web` is provided in @fig-autovi-web.
+This is the default view of the web applicaiton, and there are five regions that user can mainly interact with.
+Region 1 of @fig-autovi-web is a sidebar menu which can switch between the anlysis page and the information page.
+The analysis page is the focus of this section.
 
-### Data Pipeline
+Region 2 of @fig-autovi-web is a panel for data uploading and CSV type selection.
+Clicking the "upload CSV" button opens a window where the user can select a file from their local system.
+The data status displayed above the button provides information about the number of rows and columns in the current dataset.
+Additionally, there are two example datasets available beneath the "upload CSV" button: one is a lineup example using a CSV file with three columns, and the other is a single plot example using a CSV file with two columns.
+More details about these example datasets are be discussed in @sec-autovi-web-example.
 
-In this section, we will describe the entire data pipeline, including handling uploaded data, creating and saving fixed-aesthetic residual plots, loading and transforming images to the desired input format, and predicting visual signal strength.
+While the `autovi` package typically expects a fitted regression model object provided by the user, this approach is impractical for a web interface.
+Saving the R model object to the filesystem involves extra steps and requires users to have specific knowledge, which does not align with the goal of the web application.
+Moreover, the regression model object may contain sensitive, non-shareable data, making it unsuitable for uploading.
+Additionally, model objects are often unnecessarily large, containing extra information not needed for residual diagnostics.
+In contrast, a CSV file is easier to generate using various software programs, not just R.
+CSV files are widely accepted and can be easily viewed and modified using common desktop applications like Excel.
+They are generally less sensitive than raw data, as they exclude most information about the predictors.
 
-#### Input File Format
+The web application is designed to assess either a single residual plot or a lineup of residual plots.
+Therefore, it accepts only two types of CSV files: one with at least two columns representing the fitted values and residuals of a single residual plot, and another with at least three columns, where the additional column serves as the label or identifier for a lineup of multiple residual plots.
+For a single residual plot, 19 null plots are generated by simulating normal random draws from a distribution with the same variance as the original residual plot, and comparisons are made with the original residual plot.
+For a lineup, comparisons are made among the plots within the lineup.
+After uploading the CSV file, the user must select the correct format to ensure the web interface interprets the data correctly.
 
-As described in Section \ref{background}, the `autovi` package requires a regression model object to initialize the diagnostics.
-However, it is impractical to ask users to upload an R regression model object for inspection.
-There are several reasons for this:
+```{r fig-autovi-web-type, out.width = "100%", fig.cap = "The panels for selecting target columns and simulation settings are updated when a different CSV type is selected in the left panel.
+Compared to @fig-autovi-web, where the CSV type is a single residual plot, choosing a CSV type that includes a lineup of multiple residual plots adds a dropdown menu for specifying a column for the residual plot identifier.
+Additionally, an optional dropdown menu for specifying the true residual plot identifier will appear under the simulation settings."}
+knitr::include_graphics("figures/autovi_web_type.png")
+```
 
-1. **User Complexity**: Saving an R object to the filesystem involves extra steps and requires users to have specific knowledge.
-2. **Data Sensitivity**: The regression model object may contain sensitive, non-shareable data.
-3. **File Size**: The R object is often unnecessarily large because it contains extra information not needed for diagnostics.
+Region 3 of @fig-autovi-web is a panel for column selection and simulation settings.
+As shown in @fig-autovi-web, if the CSV type is set to a single residual plot, there will be two dropdown menus for specifying the columns for fitted values and residuals, respectively.
+The default variable names for these columns are `.fitted` and `.resid`.
+After uploading the CSV file, the content of these dropdown menus will be updated to reflect the existing columns in the dataset.
+As displayed in @fig-autovi-web-type, for the CSV type that is a lineup of multiple residual plots, an additional dropdown menu will appear for specifying the column of residual plot labels.
+The default variable name for this column is `.sample`.
+If this variable name does not exist in the dataset, the dropdown menu will remain empty, allowing the user to specify the correct column.
+The number of levels for each option in this dropdown menu will be displayed to help avoid the selection of a variable with too many levels, which could significantly slow down the application due to extensive computation.
 
-To simplify the process, the web interface instead requests a CSV file.
-This CSV file should contain at least two columns: `.fitted`, representing the fitted values, and `.resid`, representing the residuals.
-Additionally, it can contain an optional column `.sample` to indicate the ID of the residual plot.
-This is particularly useful if the user wants to evaluate a lineup of residual plots.
+Under the simulation settings, there is a slider for specifying the number of bootstrapped samples needed for the assessment.
+A higher value on this slider will result in a more accurate bootstrap distribution estimation, though it will require more computation time.
+The simulation seed can be set in a numeric input box below the slider to control the reproducibility of the assessment.
+By default, a random seed is set each time the web page is refreshed.
+When the CSV type is a lineup of multiple residual plots, an optional dropdown menu will appear next to the simulation seed input box, allowing the user to specify an identifier for the true residual plot.
+If no label is provided for the true residual plot, the assessment will only estimate the visual signal strength for each residual plot in the lineup, without providing a $p$-value, as it cannot be computed.
+Consequently, some result panels may be missing due to insufficient information.
+This option is useful when the lineup consists solely of null plots or if the user simply wants to obtain the visual signal strength for multiple residual plots.
 
-Compared to an R model object, a CSV file can be easily generated by various software programs, not just R.
-CSV files are widely accepted and can be easily viewed and modified using common desktop applications like Excel.
-CSV files are generally less sensitive than raw data because most information about the predictors is excluded.
+Region 4 of @fig-autovi-web is the panel for triggering the assessment.
+It contains a large play button to start the assessment.
+Above the play button, a text message displays the status of `TensorFlow.js`, allowing users to monitor whether the JavaScript library and Keras model have been loaded correctly.
+The play button will remain disabled until both the data status in Region 1 and the `TensorFlow.js` status in Region 4 indicate that everything is ready, with both showing a green status.
 
-#### Plot Drawing and Image Loading
+Once the play button is clicked, region 5 of @fig-autovi-web will be populated with panels displaying the assessment results.
+Generally, there will be four result panels, as shown in @fig-autovi-web-result and @fig-autovi-web-result2.
 
-The training data for the computer vision models consist of $32 \times 32$ RGB residual plots.
-These plots display fitted values on the x-axis and residuals on the y-axis.
-All labels, including axis texts, are excluded, and no background grid lines are included in the plots.
-Residual points are drawn in black with a size of 0.5 points, where there are 72.27 points per inch.
-Additionally, a horizontal red line is drawn at $y = 0$ to help the computer vision model determine if the residual points are uniformly distributed on both sides of the line.
-The plot is then saved as a PNG file with a resolution of $420 \times 525$ pixels.
-This resolution mimics a typical lineup residual plot, which has a resolution of $2100 \times 2100$ pixels and is arranged in four rows and five columns.
+```{r fig-autovi-web-result, out.width = "100%", fig.cap = "The first two panels of results from the automated residual assessment are shown.
+The application provides four results panels in total, and these screenshots display the first two.
+In region 1, there is an interactive table detailing the visual signal strength, with a summary of the analysis provided in the paragraph below.
+Region 2 displays a lineup of residual plots."}
+knitr::include_graphics("figures/autovi_web_result.png")
+```
 
-The uploaded CSV file will be partitioned based on the values in the optional column `k`.
-If no optional column `k` is present, the entire data set will be used as one partition.
-Each partition will utilize the plot specifications to generate one residual plot and produce one PNG file.
+```{r fig-autovi-web-result2, out.width = "100%", fig.cap = "The last two panels of results from the automated residual assessment are shown.
+The application provides four results panels in total, and these screenshots display the final two.
+Region 1 presents a density plot comparing the bootstrapped visual signal strength with the null visual signal strength.
+Region 2 includes an attention map of the true residual plot."}
+knitr::include_graphics("figures/autovi_web_result2.png")
+```
 
-The saved PNG plot is loaded as an array, where each entry contains a pixel value of the image.
-This array is then resized to match the input layer shape of the computer vision model, which is $1 \times 32 \times 32 \times 3$.
-If multiple images are needed for visual signal strength estimation, the arrays can be stacked together to form a larger array with the shape $n \times 32 \times 32 \times 3$, where $n$ is the number of images.
+```{r fig-autovi-web-gradient-hide, out.width = "100%", fig.cap = "The attention map is hidden if the assessment indicates a $p$-value greater than 0.05.
+A button is available to toggle the display of the attention map."}
+knitr::include_graphics("figures/autovi_web_gradient_hide.png")
+```
 
-#### Visual Signal Strength Estimation
+Region 6 of @fig-autovi-web-result contains an interactive table created with the R package `DT` [@dt], which provides the visual signal strength.
+This table includes four columns: `.sample`, `vss`, `rank`, and `null`.
+The `.sample` column shows the residual plot labels.
+For a CSV type that is a lineup, these labels are taken from an identifier column in the dataset specified by the user.
+In the case of the CSV type is a single residual plot, labels are automatically generated from 1 to 20, with the true residual plot receiving a randomly assigned label.
+The `vss` column displays the visual signal strength for each residual plot, rounded to three decimal places.
+The `rank` column indicates the ranking of each residual plot based on visual signal strength.
+The `null` column reveals whether the plot is a null plot.
+For the CSV type that is a single residual plot, only the true residual plot will have "false" in this column, while all other plots will be marked "true." For the CSV type that is a lineup, if the true residual plot identifier has not been provided, this column will show "NA" to represent missing values.
+If the identifier is provided by user, the column behaves as if the CSV type is a single residual plot.
 
-Finally, the processed image array will be fed into the computer vision model, and returned a vector of visual signal strength which are numerical values always greater than zero.
+The `DT` table provides several interactive features.
+Users can download the table in four formats, including text, CSV, Excel, and PDF, using the buttons located above the table.
+Additionally, the table is searchable via the text input field also positioned above it.
+Below the table, a text message displays the $p$-value of the assessment for the true residual plot and summarizes the number of null plots with visual signal strength greater than that of the true residual plot.
+This helps the user determine whether the true residual plot shows visual patterns that suggest model violations.
 
-### Software Stack
+Region 7 of @fig-autovi-web-result provides a lineup of plots corresponding to each `.sample` value from the table in Region 6.
+Due to space limitations, a maximum of 20 residual plots will be displayed, ensuring that the true residual plot, if known, will be included in the lineup.
+The plots are generated using `ggplot2`, the same as in `autovi`.
+Users can perform a visual test with this lineup to check if the true residual plot is distinguishable from the other plots, helping to determine the significance of model violations.
 
-#### Backend
+Region 8 of @fig-autovi-web-result2 displays the density plot for bootstrapped visual signal strength and null visual signal strength.
+The densities are shown in distinct colors that are friendly for colorblind users.
+A solid vertical line marks the visual signal strength of the true residual plot, while rug lines at the bottom of the plot provide a clearer view of individual cases.
+Below the plot, a text message indicates the number and percentage of bootstrapped residual plots that would be rejected by the visual test when compared to the null plots.
+Note that the bootstrapped residual plots in this application are generated differently from `autovi`.
+Since we do not have the R model object, we can not refit the regression model with bootstrapped data.
+Instead, we bootstrap the residuals of the true residual plot directly to obtain bootstrapped residual plots.
+As as result, this panel will disappear when the true residual plot is unknonw.
 
-To utilize the `autovi` R package, the server hosting our web interface must have a functional R interpreter.
-A static HTML page cannot accomplish this task, as it only serves static resources to the client and lacks the capability to execute R code.
-The alternative option would be WebR, a version of R designed to run within a web browser.
-However, integrating WebR introduces complexities into the design of the web interface, which we will explore further in Section XXX.
-Additionally,
-The resizing of the image is originally done by the Python `Pillow` library.
-To maintain the same data pipeline, we also need a working Python interpreter.
-Given the required conditions, we have three options: (1) Use a traditional backend like the `Spring` framework written in Java for handling all the income and outcome traffic of the web interface.
-Meanwhile, install R and Python in the server and call them when needed.
-(2) Use a Python backend framework like `Flask` so that the `Pillow` library can be used natively.
-Still, R needs to be installed and correctly configured in the server and called when needed.
-(3) Use a R backend framework like `Shiny`.
-This is similar to the second option, but requires to install and configure Python separately.
+Region 9 of @fig-autovi-web-result2 shows an attention map of the true residual plot, by computing the graident of Keras model output with repspect to the greyscale input of the true residual plot.
+We choose to show the attention map because it is useful for understanding how the Keras model predict the visual signal strength, and which area it is particurally focusing on.
+The greyscale input is used because it is relatively difficult to show a clear attention map in RGB channels and it usually will not provide more information than the greyscale since most of the important infromation of the plot are drawn in black.
 
-Option one requires a good understanding
+Region 9 of @fig-autovi-web-result2 displays an attention map for the true residual plot, generated by computing the gradient of the Keras model's output with respect to the greyscale input of the plot.
+The attention map helps to understand how the Keras model predicts visual signal strength and which areas it is focusing on.
+We use a greyscale input because it is easier to generate a clear attention map in this format, and it usually conveys all the essential information, as most of the important details of the plot are drawn in black.
+If the $p$-value of the true residual plot is greater than 0.05, checking the attention map is not necessary.
+However, to provide users with the option to review it if they wish, a button will be available, as shown in @fig-autovi-web-gradient-hide.
+This button allows users to toggle the display of the attention map.
 
-<!-- Thus, we chose to implement the web interface with a shiny server.
-Shiny server is a backend framework written in R, so it allows us to receive user's input and interactively update the output rendered on the client side using R code.
--->
+### Workflow
 
-<!-- We deploy the shiny server using the services provided by Posit, called `shinyapps.io`.
-It is responsible for reading in the uploaded CSV file with the `readr` R package, splitting the dataset with the `dplyr` R package and drawing the residual plots with the `ggplot2` R package.
-The resulting PNG files are stored in the temporary directory of the remote machine.
--->
+The workflow of `autovi.web` is designed to be straightforward, with numbered steps displayed in the main user interface as shown in @fig-autovi-web.
+Since the design details have already been covered in @sec-autovi-web-design, we will not repeat the functionality of each panel in this section.
 
-<!-- The saved PNG file  -->
+For general users, the workflow starts with Step 1: uploading a CSV file by clicking the "upload CSV" button.
+A window will pop up, allowing the user to select the CSV file.
+The CSV file should contain at least two columns representing fitted values and residuals of a residual plot.
+If a lineup needs to be evaluated, the CSV file should contain an additional column for labels of reisidual plot.
+In Step 2, the user specifies the CSV type based on the uploaded file, choosing between a single residual plot or a lineup of multiple residual plots.
+Step 3 involves selecting the appropriate columns for fitted values, residuals, and, optionally, labels of residual plots, from the dataset.
+In Step 4, the user sets the number of bootstrapped draws needed for the analysis; more draws provide a better estimate of the bootstrapped distribution but result in slower computation.
+The user should also set the seed to control the simulation.
+If the dataset contains a lineup and the true residual plot is known, select its label.
+Finally, clicking the play button will initiate the analysis, and the user can review the result panels.
 
-#### Frontend
+### Example {#sec-autovi-web-example}
 
-#### Communications between Software
-
-### Distribute Keras Model Files
-
-### Performance Optimization
-
-## Conclusions
+## Conclusions {#sec-autovi-conclusion}

### 2024-08-22 Thesis/04-chap4.qmd
remove unused paragraph

@@
 
-Region 9 of @fig-autovi-web-result2 shows an attention map of the true residual plot, by computing the graident of Keras model output with repspect to the greyscale input of the true residual plot.
-We choose to show the attention map because it is useful for understanding how the Keras model predict the visual signal strength, and which area it is particurally focusing on.
-The greyscale input is used because it is relatively difficult to show a clear attention map in RGB channels and it usually will not provide more information than the greyscale since most of the important infromation of the plot are drawn in black.
-
 Region 9 of @fig-autovi-web-result2 displays an attention map for the true residual plot, generated by computing the gradient of the Keras model's output with respect to the greyscale input of the plot.

### 2024-08-22 Thesis/03-chap3.qmd
spell checke chapter 4

@@
 
-Plotting residuals is a recommended practice to diagnose deviations from linear model assumptions such as non-linearity, heteroscedasticity, and non-normality.
-The presence or absence of structure in residual plots can be tested using the lineup protocol to do visual inference.
-As a statistical test, the lineup protocol is less sensitive and applies more broadly than available conventional tests.
+Plotting the residuals is a recommended procedure to diagnose deviations from linear model assumptions, such as non-linearity, heteroscedasticity, and non-normality.
+The presence of structure in residual plots can be tested using the lineup protocol to do visual inference.
+There are a variety of conventional residual tests, but the lineup protocol, used as a statistical test, performs better for diagnostic purposes because it is less sensitive and applies more broadly to different types of departures.
 However, the lineup protocol relies on human judgment which limits its scalability.
 This work presents a solution by providing a computer vision model to automate the assessment of residual plots.
-It is trained to predict the distance measure that quantifies the disparity between the residual distribution of a fitted classical normal linear regression model and the reference distribution, based on Kullback-Leibler divergence.
+It is trained to predict a distance measure that quantifies the disparity between the residual distribution of a fitted classical normal linear regression model and the reference distribution, based on Kullback-Leibler divergence.
 From extensive simulation studies, the computer vision model exhibits lower sensitivity than conventional tests but higher sensitivity than human visual tests.

### 2024-08-22 Thesis/04-chap4.qmd
spell checke chapter 4

@@
 A notable example is the `DHARMa` package [@dharma], which uses an innovative approach by fitting quantile regression on scaled residual plots.
-It compares the empirical 0.25, 0.5, and 0.75 quantiles in scaled residuals with their theoretical counterparts, making it particularly useful for highlighting deviations from model assumptions, detecting model violations such as heteroscedasticity and incorrect functional forms, and uncovering issues specific to generalized linear models and mixed-effect models, like over/underdispersion.
+It compares the empirical 0.25, 0.5, and 0.75 quantiles in scaled residuals with their theoretical counterparts, making it particularly useful for highlighting deviations from model assumptions, detecting model violations such as heteroscedasticity and incorrect functional forms, and uncovering issues specific to generalized linear models and mixed-effect models, like over/under-dispersion.
 By offering these enhanced visualizations, `DHARMa` enables users to more easily identify potential issues in their regression models that might not be immediately apparent with standard diagnostic plots.
@@
 The `feature_pca_plot()` method visualizes the results of the PCA.
-By default, it plots the first principal component on the x-axis and the second principal component on the y-axis, with points colored according to their origi, true residual plots, null residual plots, or bootstrapped residual plots.
+By default, it plots the first principal component on the x-axis and the second principal component on the y-axis, with points colored according to their origin, true residual plots, null residual plots, or bootstrapped residual plots.
 Users can customize the x and y axes by specifying symbols for the `x` and `y` arguments.
@@
 
-While the R package `autovi` aims to provide tools that can be extended to broader visual inference applicaitons, `autovi.web` is only focus on to for provdinign a straightforward and clean uesr interface.
-An overview of the grahpical user interface of `autovi.web` is provided in @fig-autovi-web.
-This is the default view of the web applicaiton, and there are five regions that user can mainly interact with.
-Region 1 of @fig-autovi-web is a sidebar menu which can switch between the anlysis page and the information page.
+While the R package `autovi` aims to provide tools that can be extended to broader visual inference applications, `autovi.web` is only focus on to for providing  a straightforward and clean user interface.
+An overview of the graphical user interface of `autovi.web` is provided in @fig-autovi-web.
+This is the default view of the web application, and there are five regions that user can mainly interact with.
+Region 1 of @fig-autovi-web is a sidebar menu which can switch between the analysis page and the information page.
 The analysis page is the focus of this section.
@@
 Instead, we bootstrap the residuals of the true residual plot directly to obtain bootstrapped residual plots.
-As as result, this panel will disappear when the true residual plot is unknonw.
+As as result, this panel will disappear when the true residual plot is unknown.
 
@@
 The CSV file should contain at least two columns representing fitted values and residuals of a residual plot.
-If a lineup needs to be evaluated, the CSV file should contain an additional column for labels of reisidual plot.
+If a lineup needs to be evaluated, the CSV file should contain an additional column for labels of residual plot.
 In Step 2, the user specifies the CSV type based on the uploaded file, choosing between a single residual plot or a lineup of multiple residual plots.

### 2024-08-22 Thesis/04-chap4.qmd
small revisions for the first sections of chapter 4

@@
 
-In @sec-autovi-implementation, we discussed how `autovi` relies on several Python libraries, with a particularly strong dependency on `TensorFlow`.
+The web interface, called `autovi.web`, builds on `autovi` to provide potential users easier access to automated residual plot assessment.
+It eliminates software installation issues, so users no longer need to struggle with managing Python environments or installing and maintaining the correct versions of libraries.
+The interface is cross-platform available across devices and operating systems, and accessible to users without R programming knowledge.
+An added benefit is that updates can be controlled centrally so that users will always have the latest features.
+
+<!-- In @sec-autovi-implementation, we discussed how `autovi` relies on several Python libraries, with a particularly strong dependency on `TensorFlow`.
 Managing a Python environment and correctly installing `TensorFlow` on a local machine can be challenging for many users.
@@
 This approach aligns with modern trends in data science tools, where web-based interfaces are increasingly used to make advanced analytical techniques more widely available.
+-->
 
@@
 
-`autovi.web` is a web application built using the `shiny` [@shiny] and `shinydashboard` [@shinydashboard] R packages.
-Hosted on the [shinyapps.io](https://www.shinyapps.io) domain, the application is accessible through any modern web browser.Additionally, R packages `htmltools` [@htmltools] and `shinycssloaders` [@shinycssloaders] are used to render markdown document in shiny application, and add loading animation for shiny widgets, respectively.
+`autovi.web` is built using the `shiny` [@shiny] and `shinydashboard` [@shinydashboard] R packages.
+Hosted on the [shinyapps.io](https://www.shinyapps.io) domain, the application is accessible through any modern web browser.
+The R packages `htmltools` [@htmltools] and `shinycssloaders` [@shinycssloaders] are used to render markdown documentation in shiny application, and for loading animations for shiny widgets, respectively.
 
+Determining the best way to implement the interface was difficult.
 In our initial planning for `autovi.web`, we considered implementing the entire web application using the `webr` framework [@webr], which would have allowed the entire application to run directly in the user's browser.
@@
 The reason is that one of the R packages `autovi` depends on the R package `splancs` [@splancs], which uses compiled Fortran code.
-Unfortunately, a working Emscripten version of this package, which would be required for `webr`, was not available.
+A working Emscripten (XXX reference?) version of this package, which would be required for `webr`, was not available.
 
@@
 
-### Workflow
+### Usage, from the user perspective
 
-The workflow of `autovi.web` is designed to be straightforward, with numbered steps displayed in the main user interface as shown in @fig-autovi-web.
-Since the design details have already been covered in @sec-autovi-web-design, we will not repeat the functionality of each panel in this section.
+The user would follow the order of the numbered steps displayed in the main user interface as shown in @fig-autovi-web, <!--Since the design details have already been covered in @sec-autovi-web-design, we will not repeat the functionality of each panel in this section.-->
+as follows:
 
-For general users, the workflow starts with Step 1: uploading a CSV file by clicking the "upload CSV" button.
+- Step 1: uploading a CSV file by clicking the "upload CSV" button.
 A window will pop up, allowing the user to select the CSV file.
@@
 If a lineup needs to be evaluated, the CSV file should contain an additional column for labels of residual plot.
-In Step 2, the user specifies the CSV type based on the uploaded file, choosing between a single residual plot or a lineup of multiple residual plots.
-Step 3 involves selecting the appropriate columns for fitted values, residuals, and, optionally, labels of residual plots, from the dataset.
-In Step 4, the user sets the number of bootstrapped draws needed for the analysis; more draws provide a better estimate of the bootstrapped distribution but result in slower computation.
+- In Step 2, the user specifies the CSV type based on the uploaded file, choosing between a single residual plot or a lineup of multiple residual plots. 
+- Step 3 involves selecting the appropriate columns for fitted values, residuals, and, optionally, labels of residual plots, from the dataset. 
+- In Step 4, the user sets the number of bootstrapped draws needed for the analysis; more draws provide a better estimate of the bootstrapped distribution but result in slower computation.
 The user should also set the seed to control the simulation.
 If the dataset contains a lineup and the true residual plot is known, select its label.
-Finally, clicking the play button will initiate the analysis, and the user can review the result panels.
+- Finally, clicking the play button will initiate the analysis, and the user can review the result panels.
 
@@
 ## Conclusions {#sec-autovi-conclusion}
+
+## Availability
+
+The web interface is housed at [autoviweb.netlify.app](autoviweb.netlify.app).
+
+The source code for both packages are available at XXX, and the current version of `autovi` can be installed from CRAN.

### 2024-08-23 Thesis/04-chap4.qmd
revised introduction to chap 4

@@
 Regression analysis is a fundamental statistical technique widely used for modeling data from many fields.
-In modern practice, software for regression analysis tools is widely.
-The Comprehensive R Archive Network (CRAN) [@hornik2012comprehensive] hosts a vast array of packages, many of which diagnosing models using residual plots.
+To diagnose the fit of a model it is recommended that the residuals are plotted.
+If the fit is good, any variation remaining should be noise, consistent with sampling from a distribution specified by the error model.
+Deviations that might be observed from a residual plot are non-normality, heteroscedasticity, and other associations with the fitted values.
+@paper-1 established that visual methods for assessing residuals are superior to conventional testing in that they are not so sensitive to small departures.
+@sec-second-paper provides a computer vision model to alleviate the human labor needed to visually assessing residual plots.
+The next step is to deliver the use of the computer vision model to potential users, so that it can be widely used to the benefit of the analytics community.
+
+Software for regression analysis tools is widely available.
+The Comprehensive R Archive Network (CRAN) [@hornik2012comprehensive] hosts a vast array of packages, many of which provide tools to diagnose models using residual plots.
 These packages can be broadly categorized into three groups: general purpose, enhanced diagnostics, diagnostics with statistical testing.
@@
 These packages aren't specifically designed for graphical diagnostics of residuals in linear regression but offer this functionality as part of a broader set of statistical tools.
-A prime example is R's built-in `stats` package [@stats], which provides a comprehensive collection of statistical modelling tools that includes common diagnostic plots like residuals vs fitted values, quantile-quantile (Q-Q) plots, and residuals vs leverage plots.
+These include the `stats` package [@stats], which includes common diagnostic plots like residuals vs fitted values, quantile-quantile (Q-Q) plots, and residuals vs leverage plots.
 Other packages such as `jtools` [@jtools], `olsrr` [@olsrr], `rockchalk` [@rockchalk], and `ggResidpanel` [@ggresidpanel] provide similar graphical tools with alternative aesthetic styles or interactive features.
-Although these packages may differ in presentation, they all fundamentally deliver diagnostic plots based on well-established principles in regression analysis, as outlined in classic works like @cook1982residuals.
-However, consistently drawing accurate conclusions from these tools can be challenging due to individual differences in interpreting statistical graphics.
-As noted in @li2024plot, relying solely on subjective assessments of data plots can lead to problems, such as over-interpreting random patterns as model violations.
+They all fundamentally deliver diagnostic plots as outlined in the classic text @cook1982residuals.
+<!-- However, consistently drawing accurate conclusions from these tools can be challenging due to individual differences in interpreting statistical graphics.--> As noted in @li2024plot, relying solely on subjective assessments of data plots can lead to problems, such as over-interpreting random patterns as model violations.
 
-Enhanced visual diagnostics is the second group, which offers advanced visual aids for interpreting diagnostic plots.
-A notable example is the `DHARMa` package [@dharma], which uses an innovative approach by fitting quantile regression on scaled residual plots.
-It compares the empirical 0.25, 0.5, and 0.75 quantiles in scaled residuals with their theoretical counterparts, making it particularly useful for highlighting deviations from model assumptions, detecting model violations such as heteroscedasticity and incorrect functional forms, and uncovering issues specific to generalized linear models and mixed-effect models, like over/under-dispersion.
-By offering these enhanced visualizations, `DHARMa` enables users to more easily identify potential issues in their regression models that might not be immediately apparent with standard diagnostic plots.
-In summary, this group of packages enhances the interpretability of diagnostic plots by drawing attention to critical elements such as trends, clusters, and outliers.
-They sometimes automatically perform conventional tests on these elements, displaying the results as annotations, labels, or text within the plot, thereby further reducing the likelihood of misinterpretation.
+Enhanced visual diagnostics forms the second group, which offer advanced visual aids for interpreting diagnostic plots.
+This includes `ecostats` [@ecostats], which adds simulation envelopes to  residual plots, and `DHARMa` [@dharma], which compares the empirical 0.25, 0.5, and 0.75 quantiles in scaled residuals with their theoretical counterparts.
+`DHARMa`'s focus is detecting model violations such as heteroscedasticity and incorrect functional forms, and uncovering issues specific to generalized linear models and mixed-effect models, like over/under-dispersion.
+<!--By offering these enhanced visualizations, `DHARMa` enables users to more easily identify potential issues in their regression models that might not be immediately apparent with standard diagnostic plots.
+In summary, this group of packages enhances the interpretability of diagnostic plots by drawing attention to critical elements such as trends, clusters, and outliers.--> It also adds results from conventional tests as annotations, labels, or text within the plot, to assist in avoiding misinterpretation.
 
 Statistical testing for visual discoveries is the third group, which focuses on providing tools for conducting formal statistical tests for visual discoveries obtained from diagnostic plots [@buja2009statistical].
-Examples in this category include the `nullabor` [@nullabor] and `regressinator` [@regressinator] packages, which enables users to quantify the significance of patterns observed in residual diagnostic plots, perform hypothesis tests on specific aspects of model fit, and validate visual interpretations with statistical evidence.
-This approach addresses the issue of inconsistent interpretation of diagnostic plots by bridging the gap between visual inspection and formal statistical inference, thus offering a more robust framework for regression diagnostics.
+Packages include the `nullabor` [@nullabor] `HLMdiag` [@loy2014hlmdiag] and `regressinator` [@regressinator].
+These enables users to compare the actual residual plot with samples from null distributions, and can help quantify the significance of patterns.
+<!--This approach addresses the issue of inconsistent interpretation of diagnostic plots by bridging the gap between visual inspection and formal statistical inference, thus offering a more robust framework for regression diagnostics.
 
@@
 Observers are then asked to identify the plot that appears most different from the others.
-If a significant percentage of observers correctly identify the true residual plot, it provides evidence against the $H_0$, as according to @buja2009statistical, the true residual plot would have no distinguishable difference from the null plots if all residuals are generated by the same process.
+If a significant percentage of observers correctly identify the true residual plot, it provides evidence against the $H_0$, as according to @buja2009statistical, the true residual plot would have no distinguishable difference from the null plots if all residuals are generated by the same process.-->
 
-However, as discussed in @sec-second-paper, the lineup protocol has significant limitations in large-scale applications due to its high labor costs and time consumption.
-To address these issues, we developed a computer vision model and an associated statistical testing procedure to automate the assessment of residual plots.
+However, as discussed in @sec-second-paper, the lineup protocol has significant limitations in large-scale applications due dependence on human labor.
+Thus a computer vision model was developed with an associated statistical testing procedure to automate the assessment of residual plots.
 This model takes a residual plot and a vector of auxiliary variables (such as the number of observations) as inputs and outputs a visual signal strength.
@@
 
-By estimating the visual signal strength for all plots in a lineup, we can compute a $p$-value based on the ratio of plots with visual signal strength greater than or equal to that of the true residual plot.
+<!--By estimating the visual signal strength for all plots in a lineup, we can compute a $p$-value based on the ratio of plots with visual signal strength greater than or equal to that of the true residual plot.
 This $p$-value has a lower bound of one divided by the number of plots in the lineup.
@@
 If these distributions are largely similar, it suggests that the bootstrapped residual plots are similar to the null plots.
-The proportion of bootstrapped visual signal strengths exceeding a critical value (such as the 95% sample quantile of the null distribution) indicates how often the assumed regression model would be considered incorrect if the data could be repeatedly obtained from the same data-generating process.
+The proportion of bootstrapped visual signal strengths exceeding a critical value (such as the 95% sample quantile of the null distribution) indicates how often the assumed regression model would be considered incorrect if the data could be repeatedly obtained from the same data-generating process.-->
 
-To make the statistical testing procedure and trained computer vision model widely accessible, we developed the R package `autovi`.
-In addition, we created a web-based tool that offers a user-friendly interface, enabling users to diagnose their residual plots without the need to install any dependencies required by the `autovi` package.
+To make the statistical testing procedure and trained computer vision model widely accessible, we developed the R package `autovi`, and a web interface, `autovi.web` to make it easy for users to automatically read their residual plots with the trained computer vision model.
 

### 2024-08-23 Thesis/index.qmd
mapped out the front parts

::: {.content-visible when-format="html"}

A thesis submitted for the degree of {{< meta degreetype >}} at Monash University, {{< meta affiliation >}}.

:::

# Copyright notice {-}

```{r}
#| output: asis
cat("Produced on",format(Sys.Date(), "%e %B %Y.\n\n"))
cat("© {{< meta author >}} (",format(Sys.Date(), "%Y"),").", sep="")
```

<!--
# Abstract {-}

The abstract should outline the main approach and findings of the thesis and must not be more than 500 words.

# Declaration {-}

> Use only one of the following declarations (Standard thesis or Thesis including published works declaration) and remove the other.

### Standard thesis

This thesis is an original work of my research and contains no material which has been accepted for the award of any other degree or diploma at any university or equivalent institution and that, to the best of my knowledge and belief, this thesis contains no material previously published or written by another person, except where due reference is made in the text of the thesis.

Student name:

Student signature:

Date:

#### Publications during enrolment {-}

> Remove this section if you do not have publications.

The material in @sec-intro has been submitted to the journal *Journal of Impossible Results* for possible publication.

The contribution in @sec-litreview of this thesis was presented in the International Symposium on Nonsense held in Dublin, Ireland, in July 2022.

#### Reproducibility statement

This thesis is written using Quarto with renv [@renv] to create a reproducible environment. All materials (including the data sets and source files) required to reproduce this document can be found at the Github repository [`github.com/SusanSu/thesis`](https://github.com/SusanSu/thesis).

This work is licensed under a [Creative Commons  Attribution-NonCommercial-ShareAlike 4.0 International License](http://creativecommons.org/licenses/by-nc-sa/4.0/).

### Thesis including published works declaration

I hereby declare that this thesis contains no material which has been accepted for the award of any other degree or diploma at any university or equivalent institution and that, to the best of my knowledge and belief, this thesis contains no material previously published or written by another person, except where due reference is made in the text of the thesis.

This thesis includes ?? original papers published in peer reviewed journals and ?? submitted publications. The core theme of the thesis is ??. The ideas, development and writing up of all the papers in the thesis were the principal responsibility of myself, the student, working within the Department of Econometrics & Business Statistics under the supervision of ??

(The inclusion of co-authors reflects the fact that the work came from active collaboration between researchers and acknowledges input into team-based research.)

In the case of (??insert chapter numbers) my contribution to the work involved the following:

```{r}
#| message: false
#| echo: false
library(tidyverse)
library(knitr)
library(kableExtra)
tab <- tribble(
  ~`Thesis chapter`, ~`Publication title`, ~Status, ~`Nature and % of student contribution`, ~`Nature and % of coauthors' contribution`, ~`Coauthors are Monash students`,
  2, "The life cycle of Mongolian crickets", "Submitted", "Concept and data analysis, writing first draft: 60%", "Shu Xu, input into manuscript: 25%; Eddie Betts, input into manuscript: 15%", "Shu Xu: No; Eddie Betts: Yes"
)
```

::: {.content-visible when-format="html"}

```{r}
tab |>
  kable() |>
  row_spec(0, align = "l")
```

:::

::: {.content-visible when-format="pdf"}

```{r}
tab |>
  kable(booktabs = TRUE, format = "latex") |>
  kable_styling(full_width = TRUE, font_size = 10, latex_options = "scale_down") |>
  column_spec(1, width = "1.2cm") |>
  column_spec(c(2, 4:6), width = "2.6cm") |>
  row_spec(0, align = "l", bold = TRUE)
```

:::

I have / have not renumbered sections of submitted or published papers in order to generate a consistent presentation within the thesis.

Student name:

Student signature:

Date:

I hereby certify that the above declaration correctly reflects the nature and extent of the student’s and co-authors’ contributions to this work. In instances where I am not the responsible author I have consulted with the responsible author to agree on the respective contributions of the authors.

Main Supervisor name:

Main Supervisor signature:

Date:

# Acknowledgements {-}

I would like to thank my pet goldfish for ...

> In accordance with Chapter 7.1.4 of the research degrees handbook, if you have engaged the services of a professional editor, you must provide their name and a brief description of the service rendered. If the professional editor's current or former area of academic specialisation is similar your own, this too should be stated as it may suggest to examiners that the editor's advice to the student has extended beyond guidance on English expression to affect the substance and structure of the thesis.

> If you have used generative artificial intelligence (AI) technologies, you must include a written acknowledgment of the use and its extent. Your acknowledgement should at a minimum specify which technology was used, include explicit description on how the information was generated, and explain how the output was used in your work. Below is a suggested format:

> “I acknowledge the use of [insert AI system(s) and link] to [specific use of generative artificial intelligence]. The output from these was used to [explain use].”

> Free text section for you to record your acknowledgment and gratitude for the more general academic input and support such as financial support from grants and scholarships and the non-academic support you have received during the course of your enrolment. If you are a recipient of the “Australian Government Research Training Program Scholarship”, you are required to include the following statement:

> > “This research was supported by an Australian Government Research Training Program (RTP) Scholarship.”

> You may also wish to acknowledge significant and substantial contribution made by others to the research, work and writing represented and/or reported in the thesis. These could include significant contributions to: the conception and design of the project; non-routine technical work; analysis and interpretation of research data; drafting significant parts of the work or critically revising it so as to contribute to the interpretation.

<!--
The following line is required to re-set page numbering after preliminary material in the pdf output. Do not remove
-->

\clearpage\pagenumbering{arabic}\setcounter{page}{1}

### 2024-08-23 Thesis/04-chap4.qmd
mapped out the front parts
finish the chap 4, next is to polish it

--- 2024-08-22 Thesis/04-chap4.qmd
@@
 A notable example is the `DHARMa` package [@dharma], which uses an innovative approach by fitting quantile regression on scaled residual plots.
-It compares the empirical 0.25, 0.5, and 0.75 quantiles in scaled residuals with their theoretical counterparts, making it particularly useful for highlighting deviations from model assumptions, detecting model violations such as heteroscedasticity and incorrect functional forms, and uncovering issues specific to generalized linear models and mixed-effect models, like over/under-dispersion.
+It compares the empirical 0.25, 0.5, and 0.75 quantiles in scaled residuals with their theoretical counterparts, making it particularly useful for highlighting deviations from model assumptions, detecting model violations such as heteroscedasticity and incorrect functional forms, and uncovering issues specific to generalized linear models and mixed-effect models, like over/underdispersion.
 By offering these enhanced visualizations, `DHARMa` enables users to more easily identify potential issues in their regression models that might not be immediately apparent with standard diagnostic plots.
@@
 The `feature_pca_plot()` method visualizes the results of the PCA.
-By default, it plots the first principal component on the x-axis and the second principal component on the y-axis, with points colored according to their origin, true residual plots, null residual plots, or bootstrapped residual plots.
+By default, it plots the first principal component on the x-axis and the second principal component on the y-axis, with points colored according to their origi, true residual plots, null residual plots, or bootstrapped residual plots.
 Users can customize the x and y axes by specifying symbols for the `x` and `y` arguments.
@@
 
-While the R package `autovi` aims to provide tools that can be extended to broader visual inference applications, `autovi.web` is only focus on to for providing  a straightforward and clean user interface.
-An overview of the graphical user interface of `autovi.web` is provided in @fig-autovi-web.
-This is the default view of the web application, and there are five regions that user can mainly interact with.
-Region 1 of @fig-autovi-web is a sidebar menu which can switch between the analysis page and the information page.
+While the R package `autovi` aims to provide tools that can be extended to broader visual inference applicaitons, `autovi.web` is only focus on to for provdinign a straightforward and clean uesr interface.
+An overview of the grahpical user interface of `autovi.web` is provided in @fig-autovi-web.
+This is the default view of the web applicaiton, and there are five regions that user can mainly interact with.
+Region 1 of @fig-autovi-web is a sidebar menu which can switch between the anlysis page and the information page.
 The analysis page is the focus of this section.
@@
 Additionally, there are two example datasets available beneath the "upload CSV" button: one is a lineup example using a CSV file with three columns, and the other is a single plot example using a CSV file with two columns.
-More details about these example datasets are be discussed in @sec-autovi-web-example.
+More details about these example datasets are be discussed in @sec-autovi-web-workflow.
 
@@
 Instead, we bootstrap the residuals of the true residual plot directly to obtain bootstrapped residual plots.
-As as result, this panel will disappear when the true residual plot is unknown.
+As as result, this panel will disappear when the true residual plot is unknonw.
 
@@
 
-### Workflow
+### Workflow {#sec-autovi-web-workflow}
 
-The workflow of `autovi.web` is designed to be straightforward, with numbered steps displayed in the main user interface as shown in @fig-autovi-web.
-Since the design details have already been covered in @sec-autovi-web-design, we will not repeat the functionality of each panel in this section.
+The workflow of `autovi.web` is designed to be straightforward, with numbered steps displayed in each panel.
+Since the design details have already been covered in @sec-autovi-web-design, we will not repeat the functionality of each panel here.
 
-For general users, the workflow starts with Step 1: uploading a CSV file by clicking the "upload CSV" button.
-A window will pop up, allowing the user to select the CSV file.
-The CSV file should contain at least two columns representing fitted values and residuals of a residual plot.
-If a lineup needs to be evaluated, the CSV file should contain an additional column for labels of residual plot.
-In Step 2, the user specifies the CSV type based on the uploaded file, choosing between a single residual plot or a lineup of multiple residual plots.
-Step 3 involves selecting the appropriate columns for fitted values, residuals, and, optionally, labels of residual plots, from the dataset.
-In Step 4, the user sets the number of bootstrapped draws needed for the analysis; more draws provide a better estimate of the bootstrapped distribution but result in slower computation.
-The user should also set the seed to control the simulation.
-If the dataset contains a lineup and the true residual plot is known, select its label.
-Finally, clicking the play button will initiate the analysis, and the user can review the result panels.
+Two example datasets are provided by the web application, as mentioned in @sec-autovi-web-design.
+The single residual plot example uses the `dino` dataset from the R package `datasauRus` [@datasaurus].
+The lineup example uses residuals from a simulated regression model that has a non-linearity issue.
+We will walk through the lineup example to further demonstrate the workflow of the web application.
 
-### Example {#sec-autovi-web-example}
+As shown in @fig-autovi-web-workflow-example, to use the lineup example data, click the "Use Lineup Example" button.
+The data status will then update to show the number of rows and columns in the dataset, and the CSV type will automatically be selected to the correct option.
+Since the example dataset follows the variable naming conventions assumed by the web application, the columns for fitted values, residuals, and labels of residual plots are set automatically.
+If the user is working with a custom dataset, these options must be set accordingly.
+Regardless of the dataset, the user must manually select the label for the true residual plot, as the web application allows assessments without this label.
+The next step is to click the play button to start the assessment.
+
+In the result panels shown in [@fig-autovi-web-workflow-example]D, the first row of the table is the most crucial to check, as it provides the visual signal strength and the rank of the true residual plot among the other plots.
+The summary text beneath the table provides the $p$-value, which can be used for quick decision-making.
+The lineup is for manual inspection, and the user should see if the true residual plot is visually distinguishable from the other plots, to confirm if the model violation is serious.
+The density plot in [@fig-autovi-web-workflow-example]E offers a more robust result, allowing the user to compare the distribution of bootstrapped visual signal strength with the distribution of null visual signal strength.
+Finally, the grayscale attention map can be used to check if the target visual features, like the non-linearity present in the lineup example, are captured by the computer vision model, ensuring the quality of the assessment.
+
+```{r fig-autovi-web-workflow-example, out.width = "100%", fig.cap = 'The workflow for `autovi.web` using the lineup example dataset proceeds as follows: (A) To begin, the user clicks the "Use Lineup Example" button to load the example dataset, during which the data status and CSV type will be automatically updated.
+(B) The target columns are selected automatically, though the user must manually select the label for the true residual plot, as the web application permits assessment without this label.
+(C) The user initiates the assessment by clicking the run button.
+(D) The visual signal strength of the true residual plot is displayed in the first row of the table, with a summary text beneath the table providing the $p$-value to aid in decision-making.
+The lineup on the right allows for manual inspection.
+(E) The density plot helps verify if the bootstrapped distribution differs from the null distribution, while the attention map on the right offers insights into whether the computer vision model has captured the intended visual features of the true residual plot.'}
+magick::image_read_pdf("figures/autovi_web_workflow.pdf")
+```
 
 ## Conclusions {#sec-autovi-conclusion}
+
+This chapter presents an advancement in regression diagnostics software through the development of the `autovi` package and its accompanying web interface, `autovi.web`, addressing a critical gap in the current landscape of statistical software.
+While regression tools are widely available, effective and efficient diagnostic methods have lagged behind, particularly in the field of residual plot interpretation.
+
+The `autovi` R package, introduced in this chapter, automates the assessment of residual plots by incorporating a computer vision model, eliminating the need for time-consuming and potentially inconsistent human interpretation.
+This automation improves the efficiency of the diagnostic process and promotes consistency in model evaluation across different users and studies.
+
+The development of the accompanying Shiny app, `autovi.web`, expands access to these advanced diagnostic tools.
+By providing a user-friendly interface, it makes sophisticated residual plot assessment accessible to a broader audience, including those who may not have extensive programming experience.
+This web-based solution effectively addresses the potential barriers to adoption, such as complex dependencies and installation requirements, that are often associated with advanced statistical software.
+
+The combination of `autovi` and `autovi.web` offers a comprehensive solution to the challenges of residual plot interpretation in regression analysis.
+These tools have the potential to significantly improve the quality and consistency of model diagnostics across various fields, from academic research to industry applications.
+By automating a critical aspect of model evaluation, they allow researchers and analysts to focus more on interpreting results and refining models, rather than grappling with the intricacies of plot assessment.
+
+The framework established by `autovi` and `autovi.web` opens up exciting possibilities for further research and development.
+Future work could explore the extension of these automated assessment techniques to other types of diagnostic plots and statistical models, potentially revolutionizing how we approach visual inference in statistical analysis more broadly.

### 2024-08-23 Thesis/04-chap4.qmd

--- 2024-08-23 Thesis/04-chap4.qmd
@@
-# Software for Automated Residual Plot Assessment: autovi and autovi.web
+# Software for Automated Residual Plot Assessment: autovi and autovi.web {#sec-third-paper}
 
@@
 
-### Usage, from the user perspective
+### Workflow {#sec-autovi-web-workflow}
 
-The user would follow the order of the numbered steps displayed in the main user interface as shown in @fig-autovi-web, <!--Since the design details have already been covered in @sec-autovi-web-design, we will not repeat the functionality of each panel in this section.-->
-as follows:
+The workflow of `autovi.web` is designed to be straightforward, with numbered steps displayed in each panel shown in @fig-autovi-web.
+Thre are two example datasets provided by the web application, as mentioned in @sec-autovi-web-design.
+The single residual plot example uses the `dino` dataset from the R package `datasauRus` [@datasaurus].
+The lineup example uses residuals from a simulated regression model that has a non-linearity issue.
+We will walk through the lineup example to further demonstrate the workflow of the web application.
 
-- Step 1: uploading a CSV file by clicking the "upload CSV" button.
-A window will pop up, allowing the user to select the CSV file.
-The CSV file should contain at least two columns representing fitted values and residuals of a residual plot.
-If a lineup needs to be evaluated, the CSV file should contain an additional column for labels of residual plot.
-- In Step 2, the user specifies the CSV type based on the uploaded file, choosing between a single residual plot or a lineup of multiple residual plots. 
-- Step 3 involves selecting the appropriate columns for fitted values, residuals, and, optionally, labels of residual plots, from the dataset. 
-- In Step 4, the user sets the number of bootstrapped draws needed for the analysis; more draws provide a better estimate of the bootstrapped distribution but result in slower computation.
-The user should also set the seed to control the simulation.
-If the dataset contains a lineup and the true residual plot is known, select its label.
-- Finally, clicking the play button will initiate the analysis, and the user can review the result panels.
+As shown in @fig-autovi-web-workflow-example, to use the lineup example data, click the "Use Lineup Example" button.
+The data status will then update to show the number of rows and columns in the dataset, and the CSV type will automatically be selected to the correct option.
+Since the example dataset follows the variable naming conventions assumed by the web application, the columns for fitted values, residuals, and labels of residual plots are set automatically.
+If the user is working with a custom dataset, these options must be set accordingly.
+Regardless of the dataset, the user must manually select the label for the true residual plot, as the web application allows assessments without this label.
+The next step is to click the play button to start the assessment.
 
-### Example {#sec-autovi-web-example}
+In the result panels shown in [@fig-autovi-web-workflow-example]D, the first row of the table is the most crucial to check, as it provides the visual signal strength and the rank of the true residual plot among the other plots.
+The summary text beneath the table provides the $p$-value, which can be used for quick decision-making.
+The lineup is for manual inspection, and the user should see if the true residual plot is visually distinguishable from the other plots, to confirm if the model violation is serious.
+The density plot in [@fig-autovi-web-workflow-example]E offers a more robust result, allowing the user to compare the distribution of bootstrapped visual signal strength with the distribution of null visual signal strength.
+Finally, the grayscale attention map can be used to check if the target visual features, like the non-linearity present in the lineup example, are captured by the computer vision model, ensuring the quality of the assessment.
+
+```{r fig-autovi-web-workflow-example, out.width = "100%", fig.cap = 'The workflow for `autovi.web` using the lineup example dataset proceeds as follows: (A) To begin, the user clicks the "Use Lineup Example" button to load the example dataset, during which the data status and CSV type will be automatically updated.
+(B) The target columns are selected automatically, though the user must manually select the label for the true residual plot, as the web application permits assessment without this label.
+(C) The user initiates the assessment by clicking the run button.
+(D) The visual signal strength of the true residual plot is displayed in the first row of the table, with a summary text beneath the table providing the $p$-value to aid in decision-making.
+The lineup on the right allows for manual inspection.
+(E) The density plot helps verify if the bootstrapped distribution differs from the null distribution, while the attention map on the right offers insights into whether the computer vision model has captured the intended visual features of the true residual plot.'}
+magick::image_read_pdf("figures/autovi_web_workflow.pdf")
+```
 
 ## Conclusions {#sec-autovi-conclusion}
+
+This chapter presents an advancement in regression diagnostics software through the development of the `autovi` package and its accompanying web interface, `autovi.web`, addressing a critical gap in the current landscape of statistical software.
+While regression tools are widely available, effective and efficient diagnostic methods have lagged behind, particularly in the field of residual plot interpretation.
+
+The `autovi` R package, introduced in this chapter, automates the assessment of residual plots by incorporating a computer vision model, eliminating the need for time-consuming and potentially inconsistent human interpretation.
+This automation improves the efficiency of the diagnostic process and promotes consistency in model evaluation across different users and studies.
+
+The development of the accompanying Shiny app, `autovi.web`, expands access to these advanced diagnostic tools.
+By providing a user-friendly interface, it makes sophisticated residual plot assessment accessible to a broader audience, including those who may not have extensive programming experience.
+This web-based solution effectively addresses the potential barriers to adoption, such as complex dependencies and installation requirements, that are often associated with advanced statistical software.
+
+The combination of `autovi` and `autovi.web` offers a comprehensive solution to the challenges of residual plot interpretation in regression analysis.
+These tools have the potential to significantly improve the quality and consistency of model diagnostics across various fields, from academic research to industry applications.
+By automating a critical aspect of model evaluation, they allow researchers and analysts to focus more on interpreting results and refining models, rather than grappling with the intricacies of plot assessment.
+
+The framework established by `autovi` and `autovi.web` opens up exciting possibilities for further research and development.
+Future work could explore the extension of these automated assessment techniques to other types of diagnostic plots and statistical models, potentially revolutionizing how we approach visual inference in statistical analysis more broadly.
 

### 2024-08-23 Thesis/abstract.qmd

# Abstract {-}

Take sentences from the abstracts of all three chapters

::: {.content-visible when-format="pdf"}

# Declaration {-}

I hereby declare that this thesis contains no material which has been accepted for the award of any other degree or diploma at any university or equivalent institution and that, to the best of my knowledge and belief, this thesis contains no material previously published or written by another person, except where due reference is made in the text of the thesis.

This thesis includes one original paper published in a peer reviewed journal and two unpublished papers.

The inclusion of co-authors reflects the fact that the work came from active collaboration between researchers and acknowledges input into team-based research. In the case of @paper-1 to  @sec-third-paper, my contribution to the work involved the following:

XXX This is Sherry's summary. You need to revise it for your works.

```{r}
#| message: false
#| echo: false
library(tidyverse)
library(knitr)
library(kableExtra)
dt <- tribble(
  ~`Chapter`, ~`Publication title`, ~Status, ~`Student contribution`, ~`Co-authors contribution`, ~`Monash Student Yes/No’`,
  2, "Visual Diagnostics for Constrained Optimisation with Application to Guided Tours", "Published in the R Journal; The R package ferrn on CRAN", "80%  Concept, Analysis, Software, Writing", "D. Cook 10%,  \n P. Menéndez 5%, \n U. Laa 2.5%,  \n N. Langrené 2.5%", "No",
  3, "Cubble: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data", "In revision for the Journal of Statistical Software; The R package cubble on CRAN", "80% Concept, Analysis, Software, Writing", "D. Cook 10%,  \n P. Menéndez 5%, \n U. Laa 2.5%,  \n N. Langrené 2.5%", "No",
)

dt |>
  kable(booktabs = TRUE, format = "latex", longtable = TRUE) |>
  kable_styling(font_size = 10, 
                latex_options = c("hold_position", "repeat_header")) |>
  column_spec(1, width = "1.2cm") |>
  column_spec(2, width = "3cm") |>
  column_spec(3, width = "3.5cm") |>
  column_spec(4, width = "2.5cm") |>
  column_spec(5, width = "2.4cm") |>
  column_spec(6, width = "1.3cm") |>
  row_spec(0, align = "l", bold = TRUE)
```

Chapter 4, *A Tidy Framework and Infrastructure to Systematically Assemble Spatio-temporal Indexes from Multivariate Data* is to be submitted for the Journal of Computational and Graphical Statistics and the associated R package `tidyindex` is to be submitted to CRAN.

**The thesis is written in Australian spelling, except for Chapters 3 and 4, which use American spelling as specified by the publication venue.**

I have not renumbered sections of submitted or published papers in order to generate a consistent presentation within the thesis.

**Student name**: H. Sherry Zhang

**Student signature**:

**Date**: 23rd August 2023

I hereby certify that the above declaration correctly reflects the nature and extent of the student’s and co-authors’ contributions to this work. In instances where I am not the responsible author I have consulted with the responsible author to agree on the respective contributions of the authors.

**Main Supervisor name**: Dianne Cook

**Main Supervisor signature**:

**Date**: 23rd August 2023

:::

### 2024-08-23 Thesis/04-chap4.qmd
fixed link to chapter 2 and revised abstract of chapter 4

@@
 Regression software is widely available today, but tools for effective diagnostics are still lagging.
-Although it is advised to diagnose a linear model by plotting residuals, it required human effort which can be prohibit the efforts.
+Perhaps, one of the reasons is that it is hard.
+Conventional tests, that would make the task easy, are often too sensitive, which would result in adequate models being abandoned if the analyst strictly adhered to the test decision.
+The recommended advice is to have the analyst assess the strength of patterns in residual plots.
+This requires human effort and also suffers from the potential for inconsistent decisions from different analysts.
+Using a lineup protocol can help to alleviate inconsistency, but requires even more human effort.
+This is the type of task where a robot might be employed to do the disagreeable work that currently requires a human.
 Here we describe a new R package that includes a computer vision model for automated assessment of residual plots, and an accompanying shiny app for ease of use.
@@
 Deviations that might be observed from a residual plot are non-normality, heteroscedasticity, and other associations with the fitted values.
-@paper-1 established that visual methods for assessing residuals are superior to conventional testing in that they are not so sensitive to small departures.
+@sec-first-paper established that visual methods for assessing residuals are superior to conventional testing in that they are not so sensitive to small departures.
 @sec-second-paper provides a computer vision model to alleviate the human labor needed to visually assessing residual plots.

### 2024-08-23 Thesis/04-chap4.qmd
abstract for chap 4
removed adverbs like powerful, and shifted one paragraph into installation section

@@
 Here we describe a new R package that includes a computer vision model for automated assessment of residual plots, and an accompanying shiny app for ease of use.
+For a user-provided sample of residuals, it predicts a measure of visual signal strength and provides a suite of supporting information to assist the analyst diagnose their model fit.
 
@@
 
-Within the R environment, `autovi` utilizes several powerful libraries.
+Within the R environment, `autovi` utilizes several libraries.
 `ggplot2` [@ggplot2] is employed to generate the initial residual plots, which are then saved as PNG files, using as the primary visual input for the analysis.
@@
 The `reticulate` [@reticulate] package is used to bridge R and Python, allowing for seamless communication between the two languages and supporting an integrated infrastructure.
-
-The package includes internal functions to check the current Python environment used by the `reticulate` package.
-If the necessary Python packages are not installed in the Python interpreter, an error will be raised.
-If you want to select a specific Python environment, you can do so by calling the `reticulate::use_python()` function before using the `autovi` package.
 
@@
 The code discussed in this chapter is based on `autovi` version 0.4.1.
+
+The package includes internal functions to check the current Python environment used by the `reticulate` package.
+If the necessary Python packages are not installed in the Python interpreter, an error will be raised.
+If you want to select a specific Python environment, you can do so by calling the `reticulate::use_python()` function before using the `autovi` package.
 

### 2024-08-23 Thesis/04-chap4.qmd
changed visual signal strength to VSS in the multiple uses in chap 4 and gave first to output plots fig labels and captions

--- 2024-08-23 Thesis/04-chap4.qmd
@@
 Here we describe a new R package that includes a computer vision model for automated assessment of residual plots, and an accompanying shiny app for ease of use.
+For a user-provided sample of residuals, it predicts a measure of visual signal strength (VSS) and provides a suite of supporting information to assist the analyst diagnose their model fit.
 
@@
 Thus a computer vision model was developed with an associated statistical testing procedure to automate the assessment of residual plots.
-This model takes a residual plot and a vector of auxiliary variables (such as the number of observations) as inputs and outputs a visual signal strength.
+This model takes a residual plot and a vector of auxiliary variables (such as the number of observations) as inputs and outputs the predicted visual signal strength (VSS).
 This strength estimates the distance between the residual distribution of the fitted regression model and the reference distribution assumed under correct model specification.
@@
 In Python, `pillow` [@clark2015pillow] handles image processing tasks, including reading PNG files of residual plots, resizing them, and converting them into input tensors for further analysis.
-The `TensorFlow` [@abadi2016tensorflow] library, a cornerstone of contemporary machine learning, is employed to predict the visual signal strength of these residual plots,  utilizing a pre-trained convolutional neural network.
+The `TensorFlow` [@abadi2016tensorflow] library, a cornerstone of contemporary machine learning, is employed to predict the VSS of these residual plots, utilizing a pre-trained convolutional neural network.
 
-Within the R environment, `autovi` utilizes several powerful libraries.
+Within the R environment, `autovi` utilizes several libraries.
 `ggplot2` [@ggplot2] is employed to generate the initial residual plots, which are then saved as PNG files, using as the primary visual input for the analysis.
@@
 Use `get_keras_model()`, a function provided by `autovi`, to download a trained computer vision model (described in @sec-second-paper) from GitHub.
-"vss_phn_32" specifies a model that predicts visual signal strength (vss) and is trained on residuals with polynomial, heteroskedasticity, and non-normality patterns (phn).
+"vss_phn_32" specifies a model that predicts `vss` and is trained on residuals with polynomial, heteroskedasticity, and non-normality patterns (phn).
 More details about the hosted models will be provided in section @sec-trained-model-hosting.
-4.
-Call the `check()` method of the checker with default arguments.
-This predicts the visual signal strength for the true residual plot, 100 null plots, and 100 bootstrapped plots, storing the predictions internally.
+4. Call the `check()` method of the checker with default arguments. This predicts the VSS for the true residual plot, 100 null plots, and 100 bootstrapped plots, storing the predictions internally.
 5. Use the `print()` function to generate a concise report of the check results.
 
-The report highlights key findings such as the visual signal strength of the true residual plot and the $p$-value of the automated visual test.
-The $p$-value is the ratio of null plots having visual signal strength greater than or equal to the true residual plot.
+The summary reports key findings such as the VSS of the true residual plot and the $p$-value of the automated visual test.
+The $p$-value is the ratio of null plots having VSS greater than or equal to the true residual plot.
 We typically reject the null hypothesis when the $p$-value is smaller than or equal to 5%.
-The report also provides sample quantiles of visual signal strength for null and bootstrapped plots, helping to explain the severity and likelihood of model violations.
+The report also provides sample quantiles of VSS for null and bootstrapped plots, helping to explain the severity and likelihood of model violations.
 
@@
 This visual comparison can clarify why $H_0$ is either rejected or not, and help identify potential remedies.
-The `plot_pair()` method facilitates this comparison.
+The `plot_pair()` and `plot_lineup()` methods facilitate this comparison.
 
-```{r echo = TRUE, fig.height=3, fig.width=4}
+```{r fig-plot-pair, echo = TRUE, fig.height=3, fig.width=4, fig.cap="True plot alongside one null plot, for quick comparison."}
 checker$plot_pair()
@@
 
-This method displays the true residual plot on the left and a null plot on the right.
+This method (@fig-plot-pair) displays the true residual plot on the left and a single null plot on the right.
+If a full lineup was shown, the true residual plot would be embedded in a page of null plots.
 Users should look for any distinct visual patterns in the true residual plot that are absent in the null plot.
-It's recommended to run this function multiple times to confirm any visual findings, as each execution generates a new random null plot for comparison.
+Running these functions multiple times can help any visual suspicions, as each execution generates new random null plots for comparison.
 
@@
 
-```{r echo=TRUE}
+```{r fig-summary-plot, echo = TRUE, fig.height=3, fig.width=6, fig.cap="Summary plot comparing the densities of VSS for bootstrapped residual samples relative to VSS for null plots (blue)."}
 checker$summary_plot()
@@
 
-In the visualization, the blue area represents the density of visual signal strength for null residual plots, while the red area shows the density for bootstrapped residual plots.
-The dashed line indicates the visual signal strength of the true residual plot, and the solid line marks the critical value at a 95% significance level.
+In the result, shown in @fig-summary-plot, the blue area represents the density of VSS for null residual plots, while the red area shows the density for bootstrapped residual plots.
+The dashed line indicates the VSS of the true residual plot, and the solid line marks the critical value at a 95% significance level.
 The $p$-value and the likelihood ratio are displayed in the subtitle.
-The likelihood ratio represents the ratio of the likelihood of observing the visual signal strength of the true residual plot from the bootstrapped distribution compared to the null distribution.
+The likelihood ratio represents the ratio of the likelihood of observing the VSS of the true residual plot from the bootstrapped distribution compared to the null distribution.
 
@@
 
-The infrastructure of `autovi` consists of ten core modules: data extraction, bootstrapping and model refitting, fitted values and residuals extraction, auxiliary computation, null residual simulation, plotting, plot saving, image reading and resizing, visual signal strength prediction, and $p$-value computation.
+The infrastructure of `autovi` consists of ten core modules: data extraction, bootstrapping and model refitting, fitted values and residuals extraction, auxiliary computation, null residual simulation, plotting, plot saving, image reading and resizing, VSS prediction, and $p$-value computation.
 Each module is designed with minimal dependency on the preceding modules, allowing users to customize parts of the infrastructure without affecting its overall integrity.
@@
 
-The modules for visual signal strength prediction and $p$-value computation are predefined and cannot be overridden, although users can interact with them directly through function arguments.
+The modules for VSS prediction and $p$-value computation are predefined and cannot be overridden, although users can interact with them directly through function arguments.
 Similarly, the image reading and resizing module is fixed but will adapt to different Keras models by checking their input shapes.
@@
 Optionally, the user may specify the node index of the output layer of the trained computer vision model to be monitored by the checker via the `node_index` argument if there are multiple output nodes.
-This is particularly useful for multiclass classifiers when the user wants to use one of the nodes as a visual signal strength indicator.
+This is particularly useful for multiclass classifiers when the user wants to use one of the nodes as a VSS indicator.
 
@@
 
-To be able to predict visual signal strength for a residual plot, both fitted values and residuals are needed to be extracted from the regression model object supplied by the user.
+To be able to predict VSS for a residual plot, both fitted values and residuals are needed to be extracted from the regression model object supplied by the user.
 In R, statistical models like `lm` (linear model) and `glm` (generalized linear model) typically support the use of generic functions such as `fitted()` and `resid()` to retrieve these values.
@@
 For linear regression model in R, the model frame contains all the data required by a formula for evaluation.
-This is essential for bootstrapping and refitting the model when constructing a bootstrapped distribution of visual signal strength.
+This is essential for bootstrapping and refitting the model when constructing a bootstrapped distribution of VSS.
 Typically, the model frame can be extracted from the regression model object using the `model.frame()` generic function, which is the default method used by `get_data()`.
@@
 
-According to @sec-second-paper, in some cases, a residual plot alone may not provide enough information to accurately determine visual signal strength.
+According to @sec-second-paper, in some cases, a residual plot alone may not provide enough information to accurately determine VSS.
 For instance, when the residual plot has significant overlap, the trend and shape of the residual pattern can be difficult to discern.
@@
 
-A fundamental element of the automated residual assessment described in @sec-second-paper is comparing the visual signal strength of null plots with that of the true residual plot.
+A fundamental element of the automated residual assessment described in @sec-second-paper is comparing the VSS of null plots with that of the true residual plot.
 However, due to the variety of regression models, there is no universal method for simulating null residuals that are consistent with model assumptions.
@@
 To construct a `KERAS_WRAPPER` object, you need to provide the Keras model as the main argument.
-However, users generally do not need to interact with this class directly, as the `autovi` checker automatically invokes its methods when performing visual signal strength predictions.
+However, users generally do not need to interact with this class directly, as the `autovi` checker automatically invokes its methods when performing VSS predictions.
 The `image_to_array()` method takes the path to the image file, the target height, and the target width as inputs and returns a `Numpy` array.
@@
 
-#### Visual Signal Strength Prediction
+#### Visual Signal Strength (VSS) Prediction
 
-Visual signal strength, as discussed in @sec-second-paper, estimates the distance between the input residual plot and a theoretically good  residual plot.
+VSS, as discussed in @sec-second-paper, estimates the distance between the input residual plot and a theoretically good  residual plot.
 It can be defined in various ways, much like different methods for measuring the distance between two points.
@@
 
-There are several ways to obtain visual signal strength from the checker, with the most direct being the `vss()` method.
-By default, this method predicts the visual signal strength for the true residual plot.
-If a `ggplot` or a `data.frame`, such as null residuals generated by the `null_method()`, is explicitly provided, the method will use that input to predict visual signal strength accordingly.
+There are several ways to obtain VSS from the checker, with the most direct being the `vss()` method.
+By default, this method predicts the VSS for the true residual plot.
+If a `ggplot` or a `data.frame`, such as null residuals generated by the `null_method()`, is explicitly provided, the method will use that input to predict VSS accordingly.
 Note that if a `ggplot` is provided, auxiliary inputs must be supplied manually via the `auxiliary` argument, as we assume that auxiliary variables can not be computed directly from a `ggplot`.
 
-Another way to obtain visual signal strength is by calling the `check()` method.
-This comprehensive method perform extensive diagnostics on the true residual plot and store the visual signal strength in the `check_result` field of the checker.
-Additionally, for obtaining visual signal strength for null residual plots and bootstrapped residual plots, there are two specialized methods, `null_vss()` and `boot_vss()`, designed for this purpose respectively.
+Another way to obtain VSS is by calling the `check()` method.
+This comprehensive method perform extensive diagnostics on the true residual plot and store the VSS in the `check_result` field of the checker.
+Additionally, for obtaining VSS for null residual plots and bootstrapped residual plots, there are two specialized methods, `null_vss()` and `boot_vss()`, designed for this purpose respectively.
 
-Calling the `vss()` method without arguments will predict the visual signal strength for the true residual plot and return the result as a single-element `tibble`.
+Calling the `vss()` method without arguments will predict the VSS for the true residual plot and return the result as a single-element `tibble`.
 
@@
 
-Providing a `data.frame` of null residuals or a null residual plot yields the same visual signal strength.
+Providing a `data.frame` of null residuals or a null residual plot yields the same VSS.
 
@@
 Intermediate results, including null residuals and null plots, can be returned by enabling `keep_null_data` and `keep_null_plot`.
-The visual signal strength, along with null residuals and null plots, will be stored in a `tibble` with three columns.
-The following code example demonstrates how to predict the visual signal strength for five null residual plots while keeping the intermediate results.
+The VSS, along with null residuals and null plots, will be stored in a `tibble` with three columns.
+The following code example demonstrates how to predict the VSS for five null residual plots while keeping the intermediate results.
 
@@
 The `boot_vss()` helper method is similar to `null_vss()`, with some differences in argument names.
-The following code example demonstrates how to predict the visual signal strength for five bootstrapped residual plots while keeping the intermediate results.
+The following code example demonstrates how to predict the VSS for five bootstrapped residual plots while keeping the intermediate results.
 
@@
 
-Once we have obtained the visual signal strength from both the true residual plot and the null plots, we can compute the $p$-value.
-This $p$-value represents the ratio of plots with visual signal strength greater than or equal to that of the true residual plot.
+Once we have obtained the VSS from both the true residual plot and the null plots, we can compute the $p$-value.
+This $p$-value represents the ratio of plots with VSS greater than or equal to that of the true residual plot.
 We can perform this calculation using the `check()` method.
@@
 
-The rank plot, creating by setting `type` to "rank", is a bar plot where the x-axis represents the rank and the y-axis shows the visual signal strength.
+The rank plot, creating by setting `type` to "rank", is a bar plot where the x-axis represents the rank and the y-axis shows the VSS.
 The bar for the true residual plot is colored in red.
-By examining the rank plot, you can intuitively understand how the observed visual signal strength compares to the null visual signal strengths and identify any outliers in the null distribution.
+By examining the rank plot, you can intuitively understand how the observed VSS compares to the null VSSs and identify any outliers in the null distribution.
 
@@
 
-In addition to predicting visual signal strength and computing $p$-values, `autovi` offers methods to extract features from any layer of the Keras model.
+In addition to predicting VSS and computing $p$-values, `autovi` offers methods to extract features from any layer of the Keras model.
 To see which layers are available in the current Keras model, you can use the `list_layer_name()` method from the `KERAS_WRAPPER` class.
@@
 To obtain the features, provide the layer name using the `extract_feature_from_layer` argument in the `predict()` method.
-This will return a `tibble` with the visual signal strength and all features extracted from that layer.
+This will return a `tibble` with the VSS and all features extracted from that layer.
 Each row corresponds to one plot.
@@
 This package allows calling custom JavaScript code within the Shiny framework.
-The specialized JavaScript code for initializing `TensorFlow.js` and calling `TensorFlow.js` for visual signal strength prediction is deployed alongside the Shiny application as additional resources.
+The specialized JavaScript code for initializing `TensorFlow.js` and calling `TensorFlow.js` for VSS prediction is deployed alongside the Shiny application as additional resources.
 
@@
 When the CSV type is a lineup of multiple residual plots, an optional dropdown menu will appear next to the simulation seed input box, allowing the user to specify an identifier for the true residual plot.
-If no label is provided for the true residual plot, the assessment will only estimate the visual signal strength for each residual plot in the lineup, without providing a $p$-value, as it cannot be computed.
+If no label is provided for the true residual plot, the assessment will only estimate the VSS for each residual plot in the lineup, without providing a $p$-value, as it cannot be computed.
 Consequently, some result panels may be missing due to insufficient information.
-This option is useful when the lineup consists solely of null plots or if the user simply wants to obtain the visual signal strength for multiple residual plots.
+This option is useful when the lineup consists solely of null plots or if the user simply wants to obtain the VSS for multiple residual plots.
 
@@
 The application provides four results panels in total, and these screenshots display the first two.
-In region 1, there is an interactive table detailing the visual signal strength, with a summary of the analysis provided in the paragraph below.
+In region 1, there is an interactive table detailing the VSS, with a summary of the analysis provided in the paragraph below.
 Region 2 displays a lineup of residual plots."}
@@
 The application provides four results panels in total, and these screenshots display the final two.
-Region 1 presents a density plot comparing the bootstrapped visual signal strength with the null visual signal strength.
+Region 1 presents a density plot comparing the bootstrapped VSS with the null VSS.
 Region 2 includes an attention map of the true residual plot."}
@@
 
-Region 6 of @fig-autovi-web-result contains an interactive table created with the R package `DT` [@dt], which provides the visual signal strength.
+Region 6 of @fig-autovi-web-result contains an interactive table created with the R package `DT` [@dt], which provides the VSS.
 This table includes four columns: `.sample`, `vss`, `rank`, and `null`.
@@
 In the case of the CSV type is a single residual plot, labels are automatically generated from 1 to 20, with the true residual plot receiving a randomly assigned label.
-The `vss` column displays the visual signal strength for each residual plot, rounded to three decimal places.
-The `rank` column indicates the ranking of each residual plot based on visual signal strength.
+The `vss` column displays the VSS for each residual plot, rounded to three decimal places.
+The `rank` column indicates the ranking of each residual plot based on VSS.
 The `null` column reveals whether the plot is a null plot.
@@
 Additionally, the table is searchable via the text input field also positioned above it.
-Below the table, a text message displays the $p$-value of the assessment for the true residual plot and summarizes the number of null plots with visual signal strength greater than that of the true residual plot.
+Below the table, a text message displays the $p$-value of the assessment for the true residual plot and summarizes the number of null plots with VSS greater than that of the true residual plot.
 This helps the user determine whether the true residual plot shows visual patterns that suggest model violations.
@@
 
-Region 8 of @fig-autovi-web-result2 displays the density plot for bootstrapped visual signal strength and null visual signal strength.
+Region 8 of @fig-autovi-web-result2 displays the density plot for bootstrapped VSS and null VSS.
 The densities are shown in distinct colors that are friendly for colorblind users.
-A solid vertical line marks the visual signal strength of the true residual plot, while rug lines at the bottom of the plot provide a clearer view of individual cases.
+A solid vertical line marks the VSS of the true residual plot, while rug lines at the bottom of the plot provide a clearer view of individual cases.
 Below the plot, a text message indicates the number and percentage of bootstrapped residual plots that would be rejected by the visual test when compared to the null plots.
@@
 Region 9 of @fig-autovi-web-result2 displays an attention map for the true residual plot, generated by computing the gradient of the Keras model's output with respect to the greyscale input of the plot.
-The attention map helps to understand how the Keras model predicts visual signal strength and which areas it is focusing on.
+The attention map helps to understand how the Keras model predicts VSS and which areas it is focusing on.
 We use a greyscale input because it is easier to generate a clear attention map in this format, and it usually conveys all the essential information, as most of the important details of the plot are drawn in black.
@@
 
-In the result panels shown in [@fig-autovi-web-workflow-example]D, the first row of the table is the most crucial to check, as it provides the visual signal strength and the rank of the true residual plot among the other plots.
+In the result panels shown in [@fig-autovi-web-workflow-example]D, the first row of the table is the most crucial to check, as it provides the VSS and the rank of the true residual plot among the other plots.
 The summary text beneath the table provides the $p$-value, which can be used for quick decision-making.
 The lineup is for manual inspection, and the user should see if the true residual plot is visually distinguishable from the other plots, to confirm if the model violation is serious.
-The density plot in [@fig-autovi-web-workflow-example]E offers a more robust result, allowing the user to compare the distribution of bootstrapped visual signal strength with the distribution of null visual signal strength.
+The density plot in [@fig-autovi-web-workflow-example]E offers a more robust result, allowing the user to compare the distribution of bootstrapped VSS with the distribution of null VSS.
 Finally, the grayscale attention map can be used to check if the target visual features, like the non-linearity present in the lineup example, are captured by the computer vision model, ensuring the quality of the assessment.
@@
 (C) The user initiates the assessment by clicking the run button.
-(D) The visual signal strength of the true residual plot is displayed in the first row of the table, with a summary text beneath the table providing the $p$-value to aid in decision-making.
+(D) The VSS of the true residual plot is displayed in the first row of the table, with a summary text beneath the table providing the $p$-value to aid in decision-making.
 The lineup on the right allows for manual inspection.

### 2024-08-23 Thesis/04-chap4.qmd
caption of fig 4.2

@@
 
-```{r fig-summary-plot, echo = TRUE, fig.height=3, fig.width=6, fig.cap="Summary plot comparing the densities of VSS for bootstrapped residual samples relative to VSS for null plots (blue)."}
+```{r fig-summary-plot, echo = TRUE, fig.height=3, fig.width=6, fig.cap="Summary plot comparing the densities of VSS for bootstrapped residual samples (red) relative to VSS for null plots (blue)."}
 checker$summary_plot()

### 2024-08-23 Thesis/04-chap4.qmd
removed @sec-autovi-null-method reference because the subsection is not numbered

@@
 
-`autovi` is built upon the `bandicoot` object-oriented programming (OOP) system [@bandicoot], which marks a departure from R's conventional S3 generic system.
+The package `autovi` is built upon the `bandicoot` object-oriented programming (OOP) system [@bandicoot], which marks a departure from R's conventional S3 generic system.
 The adoption of an OOP architecture enhances flexibility and modularity, enabling users to redefine key functions within the infrastructure through method overriding.
@@
 
-`bandicoot` is a lightweight object-oriented system with Python-like syntax that supports multiple inheritance and incorporates a Python-like method resolution order.
+The `bandicoot` software is a lightweight object-oriented system with Python-like syntax that supports multiple inheritance and incorporates a Python-like method resolution order.
 The system is inspired by the OOP frameworks implemented in R6 [r6] and Python.
@@
 The following code example overrides the `null_method()` with a function that simulates null residuals from the corresponding normal distribution.
-This approach differs from the default null residual simulation scheme described in @sec-autovi-null-method.
+This approach differs from the default null residual simulation scheme described earlier.
 Although less efficient than the default method for linear regression models, it provides an alternative way to simulate null residuals.

### 2024-08-23 Thesis/04-chap4.qmd
small tidying all through chap 4

@@
 
-The web interface, called `autovi.web`, builds on `autovi` to provide potential users easier access to automated residual plot assessment.
-It eliminates software installation issues, so users no longer need to struggle with managing Python environments or installing and maintaining the correct versions of libraries.
-The interface is cross-platform available across devices and operating systems, and accessible to users without R programming knowledge.
-An added benefit is that updates can be controlled centrally so that users will always have the latest features.
+The web interface package, called `autovi.web`, builds on `autovi`, and makes it easier for users easier to do automated residual plot assessment.
+It eliminates software installation issues, so users no longer need to struggle with managing Python environments or installing and maintaining the correct versions of R libraries.
+The interface is cross-platform, accessible on different devices and operating systems, and for users without R programming knowledge.
+The additional benefit is that updates can be controlled centrally so that users will always have the latest features.
 
@@
 
-`autovi.web` is built using the `shiny` [@shiny] and `shinydashboard` [@shinydashboard] R packages.
+The package `autovi.web` is built using the `shiny` [@shiny] and `shinydashboard` [@shinydashboard] R packages.
 Hosted on the [shinyapps.io](https://www.shinyapps.io) domain, the application is accessible through any modern web browser.
@@
 However, server hosting domains that natively support Python servers typically do not have the latest version of R installed.
-Additionally, calling R from Python is typically done using the `rpy2` Python library, but this approach can be awkward when dealing with language syntax related to non-standard evaluation, making it challenging to develop our application in this manner.
+Additionally, calling R from Python is typically done using the `rpy2` Python library, but this approach can be awkward when dealing with language syntax related to non-standard evaluation.
 Another option we considered was renting a server where we could have full control, such as those provided by cloud platforms like Google Cloud Platform (GCP) or Amazon Web Services (AWS).
-However, correctly setting up the server and ensuring a secure deployment requires significant expertise, which we did not possess at the time.
-Ultimately, we decided that the most practical solution was to use the `shiny` and `shinydashboard` frameworks, which are well-established in the R community and offer a solid foundation for web application development.
+However, correctly setting up the server and ensuring a secure deployment requires significant expertise.
+Ultimately, the most practical solution was to use the `shiny` and `shinydashboard` frameworks, which are well-established in the R community and offer a solid foundation for web application development.
 
@@
 
-While `autovi` requires downloading pre-trained computer vision models from GitHub, these models in ".keras" file format are incompatible with `TensorFlow.js`.
+While `autovi` requires downloading the pre-trained computer vision models from GitHub, these models in ".keras" file format are incompatible with `TensorFlow.js`.
 Therefore, we extract and store the model weights in JSON files and include them as extra resources in the Shiny application.
@@
 
-This chapter presents an advancement in regression diagnostics software through the development of the `autovi` package and its accompanying web interface, `autovi.web`, addressing a critical gap in the current landscape of statistical software.
+This chapter presents new regression diagnostics software, the R package `autovi` and its accompanying web interface package, `autovi.web`.
+It addresses a critical gap in the current landscape of statistical software.
 While regression tools are widely available, effective and efficient diagnostic methods have lagged behind, particularly in the field of residual plot interpretation.
@@
 
-The development of the accompanying Shiny app, `autovi.web`, expands access to these advanced diagnostic tools.
-By providing a user-friendly interface, it makes sophisticated residual plot assessment accessible to a broader audience, including those who may not have extensive programming experience.
+The development of the accompanying Shiny app, `autovi.web`, expands access to these advanced diagnostic tools, by providing a user-friendly interface.
+It makes automated residual plot assessment accessible to a broader audience, including those who may not have extensive programming experience.
 This web-based solution effectively addresses the potential barriers to adoption, such as complex dependencies and installation requirements, that are often associated with advanced statistical software.
@@
 The framework established by `autovi` and `autovi.web` opens up exciting possibilities for further research and development.
-Future work could explore the extension of these automated assessment techniques to other types of diagnostic plots and statistical models, potentially revolutionizing how we approach visual inference in statistical analysis more broadly.
+Future work could explore the extension of these automated assessment techniques to other types of diagnostic plots and statistical models, potentially revolutionizing how we approach statistical inference using visual displays more broadly.
 
@@
 
-The source code for both packages are available at XXX, and the current version of `autovi` can be installed from CRAN.
+The the current version of `autovi` can be installed from CRAN, and source code for both packages are available at [https://github.com/TengMCing/](https://github.com/TengMCing/).

### 2024-08-24 Thesis/abstract.qmd
fill in front matter not sure if it is correct

@@
 Take sentences from the abstracts of all three chapters
-
-::: {.content-visible when-format="pdf"}
 
@@
 The inclusion of co-authors reflects the fact that the work came from active collaboration between researchers and acknowledges input into team-based research.
-In the case of @paper-1 to  @sec-third-paper, my contribution to the work involved the following:
-
-XXX This is Sherry's summary. You need to revise it for your works.
+In the case of @sec-first-paper and @sec-third-paper, my contribution to the work involved the following:
 
@@
 dt <- tribble(
-  ~`Chapter`, ~`Publication title`, ~Status, ~`Student contribution`, ~`Co-authors contribution`, ~`Monash Student Yes/No’`,
-  2, "Visual Diagnostics for Constrained Optimisation with Application to Guided Tours", "Published in the R Journal; The R package ferrn on CRAN", "80%  Concept, Analysis, Software, Writing", "D.
+  ~`Chapter`, ~`Publication title`, ~Status, ~`Student contribution`, ~`Co-authors contribution`, ~`Coauthors are Monash students`,
+  2, "A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol", "Published in the Journal of Computational and Graphical Statistics", "80%  Concept, Analysis, Software, Writing", "D.
 Cook 10%,  \n P.
@@
 Langrené 2.5%", "No",
-  3, "Cubble: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data", "In revision for the Journal of Statistical Software; The R package cubble on CRAN", "80% Concept, Analysis, Software, Writing", "D.
+  # 3, "Automated Assessment of Residual Plots with Computer Vision Models", "Unsubmitted", "80% Concept, Analysis, Software, Writing", "D.
 Cook 10%,  \n P.
@@
 Langrené 2.5%", "No",
+  4, "Software for Automated Residual Plot Assessment: autovi and autovi.web", "The R package autovi on CRAN", "80% Concept, Analysis, Software, Writing", "D.
+Cook 10%,  \n P.
+Menéndez 5%, \n U.
+Laa 2.5%,  \n N.
+Langrené 2.5%", "No"
 )
@@
 dt |>
-  kable(booktabs = TRUE, format = "latex", longtable = TRUE) |>
+  kable(booktabs = TRUE) |>
   kable_styling(font_size = 10,
@@
 
-Chapter 4, *A Tidy Framework and Infrastructure to Systematically Assemble Spatio-temporal Indexes from Multivariate Data* is to be submitted for the Journal of Computational and Graphical Statistics and the associated R package `tidyindex` is to be submitted to CRAN.
+Chapter 3, *Automated Assessment of Residual Plots with Computer Vision Models* is to be submitted for the Journal of Computational and Graphical Statistics.
 
-**The thesis is written in Australian spelling, except for Chapters 3 and 4, which use American spelling as specified by the publication venue.**
+<!-- **The thesis is written in Australian spelling, except for Chapters 3 and 4, which use American spelling as specified by the publication venue.** -->
 
@@
 
-**Student name**: H. Sherry Zhang
+**Student name**: Weihao Li
 
@@
 
-**Date**: 23rd August 2023
+**Date**: 26th August 2024
 
@@
 
-**Date**: 23rd August 2023
-
-:::
+**Date**: 26th August 2024

### 2024-08-24 Thesis/01-chap1.qmd
update front matter

# Introduction {#sec-intro}

This is where you introduce the main ideas of your thesis, and an overview of the context and background.

In a PhD, Chapter 2 would normally contain a literature review. Typically, Chapters 3--5 would contain your own contributions. Think of each of these as potential papers to be submitted to journals. Finally, Chapter 6 provides some concluding remarks, discussion, ideas for future research, and so on. Appendixes can contain additional material that don't fit into any chapters, but that you want to put on record. For example, additional tables, output, etc.

This thesis explores innovative approaches to regression model diagnostics across three chapters. Chapter 2 presents evidence from a visual inference experiment using the lineup protocol to demonstrate why plotting residuals is superior to conventional numerical hypothesis tests for model diagnosis. The lineup protocol provides a more reliable and consistent method for interpreting residual plots, avoiding the oversensitivity of conventional tests while detecting a range of departures from good residuals simultaneously.

Chapter 3 introduces a computer vision model to automate the assessment of residual plots, addressing the scalability limitations of human-based visual inference. This model is trained to predict a distance measure based on Kullback-Leibler divergence, quantifying the disparity between the residual distribution of a fitted classical normal linear regression model and the reference distribution. Simulation studies show that the computer vision model strikes a balance between conventional tests and human visual tests in terms of sensitivity, though it is slightly less effective on non-linearity patterns.

Chapter 4 describes the development of an R package that implements the computer vision model for automated assessment of residual plots. The package includes a shiny app for ease of use, allowing analysts to input a sample of residuals and receive predictions of visual signal strength (VSS) along with supporting information to assist in diagnosing model fit. This tool aims to bridge the gap between the need for effective diagnostics and the limitations of conventional tests and human effort.

### 2024-08-24 Thesis/abstract.qmd
update front matter

@@
 
-Take sentences from the abstracts of all three chapters
+This work is motivated by the need for an automated approach to diagnose regression models through residual plots with reliability and consistency.
+While numerical hypothesis tests are commonly used, residual plots remain essential in model diagnosis because conventional tests are limited to specific types of model departures and tend to be overly sensitive.
+Visual inference using the lineup protocol offers a less sensitive and more broadly applicable alternative, yet its dependence on human judgment limits scalability.
+This research addresses these limitations by automating the assessment process.
+
+This research presents three original contributions.
+The first contribution ([Chapter 2](#sec-first-paper)) provides evidence for the effectiveness of visual inference in regression diagnostics through a human subject experiment, demonstrating the benefits of using the lineup protocol for reliable and consistent reading of residual plots.
+The second contribution ([Chapter 3](#sec-second-paper)) introduces a computer vision model to automate the assessment of residual plots, addressing the scalability limitation of the lineup protocol.
+The third contribution ([Chapter 4](#sec-third-paper)) presents an R package and Shiny app, providing a user-friendly interface for analysts to leverage the computer vision model and supporting tools for diagnostic purposes.
+These contributions advance the field of artificial intelligence for data visualization, enabling more efficient and accurate regression diagnostics.
 
@@
 The inclusion of co-authors reflects the fact that the work came from active collaboration between researchers and acknowledges input into team-based research.
-In the case of @sec-first-paper and @sec-third-paper, my contribution to the work involved the following:
+In the case of [Chapter 2](#sec-first-paper) and [Chapter 4](#sec-third-paper), my contribution to the work involved the following:
 
@@
   2, "A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol", "Published in the Journal of Computational and Graphical Statistics", "80%  Concept, Analysis, Software, Writing", "D.
-Cook 10%,  \n P.
-Menéndez 5%, \n U.
-Laa 2.5%,  \n N.
-Langrené 2.5%", "No",
+Cook 10%,  \n E.
+Tanaka 5%, \n S.
+VanderPlas 5%", "No",
   # 3, "Automated Assessment of Residual Plots with Computer Vision Models", "Unsubmitted", "80% Concept, Analysis, Software, Writing", "D.
@@
   4, "Software for Automated Residual Plot Assessment: autovi and autovi.web", "The R package autovi on CRAN", "80% Concept, Analysis, Software, Writing", "D.
-Cook 10%,  \n P.
-Menéndez 5%, \n U.
-Laa 2.5%,  \n N.
-Langrené 2.5%", "No"
+Cook 10%,  \n E.
+Tanaka 5%, \n S.
+VanderPlas 2.5%, K.
+Ackermann 2.5%", "No"
 )

### 2024-08-24 Thesis/01-chap1.qmd
introduction update

@@
 
-This is where you introduce the main ideas of your thesis, and an overview of the context and background.
+Model diagnostics play a critical role in evaluating the accuracy and validity of a statistical model.
+They enable the assessment of the model’s assumptions, detection of outliers, evaluation of how well the model fits the data, and identification of possible approaches to improve the model’s performance.
 
-In a PhD, Chapter 2 would normally contain a literature review.
-Typically, Chapters 3--5 would contain your own contributions.
-Think of each of these as potential papers to be submitted to journals.
-Finally, Chapter 6 provides some concluding remarks, discussion, ideas for future research, and so on.
-Appendixes can contain additional material that don't fit into any chapters, but that you want to put on record.
-For example, additional tables, output, etc.
+When conducting model diagnostics, graphical representations of data are often preferred or required by data analysts.
+The preference for visual diagnostics is attributed to its intuitive nature and the possibility of discovering unexpected abstract and unquantifiable insights.
+In the context of regression diagnostics, a common practice is to plot residuals against fitted values, which serves as a starting point for evaluating the adequacy of the fit and verifying the underlying assumptions.
 
-This thesis explores innovative approaches to regression model diagnostics across three chapters.
-Chapter 2 presents evidence from a visual inference experiment using the lineup protocol to demonstrate why plotting residuals is superior to conventional numerical hypothesis tests for model diagnosis.
-The lineup protocol provides a more reliable and consistent method for interpreting residual plots, avoiding the oversensitivity of conventional tests while detecting a range of departures from good residuals simultaneously.
+Recently, a novel statistical inferential framework known as visual inference [@buja2009statistical] has been developed, which relies on the use of graphical representations of data.
+The visual inference approach makes use of the natural capability of the human visual system to identify patterns and deviations from expected patterns.
+It provides a more comprehensible way of interpreting data and conducting hypothesis testing compared to conventional statistical testing.
+
+Practically, visual inference is conducted via the lineup protocol.
+The protocol is inspired by the police lineup technique employed in eyewitness identification of criminal suspects.
+It comprises $m$ randomly positioned plots, where one of them represents the data plot, while the remaining $m-1$ plots represent the null plots with the same graphical structure, except that the data has been replaced with data consistent with the null hypothesis $H_0$.
+To compute the p-value of the visual test, the lineup will be independently presented to a number of participants, asking them to pick the most different plot.
+Under $H_0$, the data plot is expected to be indistinguishable from the null plots, and the probability of correctly identifying the data plot by an observer is $1/m$.
+If a large number of participants correctly identify the data plot, the corresponding $p$-value will be small, indicating strong evidence against $H_0$.
+
+This method has gained increasing traction in recent years and has already been integrated into data analysis of various topics [see @loy2013diagnostic; @widen2016graphical; @krishnan2021hierarchical].
+However, the reliance of human assessment is a fundamental aspect of visual tests, which may restrict its widespread usage.
+The lineup protocol is unsuitable for large-scale applications, due to its high labour costs and time requirements.
+Moreover, it presents significant usability issues for individuals with visual impairments, resulting in reduced accessibility.
+
+The main objective of this research is to construct an automatic visual inference system capable of conducting visual tests on a large scale in the domain of regression diagnostics, by incorporating modern computer vision models.
+
+## Comparison of Visual Testing and Conventional Testing in Residual Diagnostics
+
+Despite the availability of numerical hypothesis tests for model diagnosis, regression experts consistently recommend visual inspection of residual plots [@draper1998applied; @cook1982residuals; @montgomery1982introduction].
+Chapter 2 provides empirical evidence supporting the indispensability of residual plots through a visual inference experiment using the lineup protocol.
+By comparing human evaluations of residual plots to conventional statistical tests, this chapter demonstrates the advantages of graphical methods in detecting practical issues with model fit, while also highlighting the limitations of conventional tests in producing overly sensitive results.
+The chapter contains a comprehensive literature review related to residual diagnostics.
+
+## Automated Residual Plot Assessment
 
@@
 This model is trained to predict a distance measure based on Kullback-Leibler divergence, quantifying the disparity between the residual distribution of a fitted classical normal linear regression model and the reference distribution.
-Simulation studies show that the computer vision model strikes a balance between conventional tests and human visual tests in terms of sensitivity, though it is slightly less effective on non-linearity patterns.
+Performance of the model is evaluated on the human subject experiment data collected in chapter 2.
+A comprehensive literature review of data plots reading with computer vision models is contained in the chapter.
 
-Chapter 4 describes the development of an R package that implements the computer vision model for automated assessment of residual plots.
-The package includes a shiny app for ease of use, allowing analysts to input a sample of residuals and receive predictions of visual signal strength (VSS) along with supporting information to assist in diagnosing model fit.
-This tool aims to bridge the gap between the need for effective diagnostics and the limitations of conventional tests and human effort.
+## Implementing Automated Residual Plot Assessment in Practice
+
+Chapter 4 introduces a new R package, `autovi`, and its accompanying web interface, `autovi.web`, designed to automate the assessment of residual plots in regression analysis.
+The package uses a computer vision model built in Chapter 3 to predict a measure of visual signal strength (VSS) and provides supporting information to assist analysts in diagnosing model fit.
+By automating this process, `autovi` and `autovi.web` improve the efficiency and consistency of model evaluation, making advanced diagnostic tools accessible to a broader audience.
+
+## Summary
+
+The thesis is structured as follows.
+Chapter 2 compares visual testing and conventional testing in residual diagnostics through a visual inference experiment, providing empirical evidence for the effectiveness of visual methods.
+Chapter 3 introduces a computer vision model for automated residual plot assessment, evaluating its performance on human subject experiment data.
+Chapter 4 presents the implementation of automated residual plot assessment in practice, through the R package `autovi` and its accompanying web interface `autovi.web`.
+
+Chapter 5 summarises the contribution of the work and their impact, and discusses some future plans.

### 2024-08-24 Thesis/abstract.qmd
introduction update

@@
 This work is motivated by the need for an automated approach to diagnose regression models through residual plots with reliability and consistency.
-While numerical hypothesis tests are commonly used, residual plots remain essential in model diagnosis because conventional tests are limited to specific types of model departures and tend to be overly sensitive.
+While numerical hypothesis tests are commonly used, residual plots remain essential in regression diagnostics because conventional tests are limited to specific types of model departures and tend to be overly sensitive.
 Visual inference using the lineup protocol offers a less sensitive and more broadly applicable alternative, yet its dependence on human judgment limits scalability.

### 2024-08-25 Thesis/01-chap1.qmd
redraft intro chapter and add extra ref

--- 2024-08-24 Thesis/01-chap1.qmd
@@
 
-This is where you introduce the main ideas of your thesis, and an overview of the context and background.
+Model diagnostics are critical in evaluating the accuracy and validity of a statistical model.
+In the context of regression diagnostics, a common practice is to plot residuals against fitted values, which serves as a starting point for evaluating the adequacy of the fit and verifying the underlying assumptions.
+Visual diagnostics are frequently preferred or recommended [@draper1998applied; @cook1982residuals; @montgomery1982introduction] due to the possibility of discovering abstract and unquantifiable insights, however, it can be subject to overinterpretation or even neglect.
 
-In a PhD, Chapter 2 would normally contain a literature review.
-Typically, Chapters 3--5 would contain your own contributions.
-Think of each of these as potential papers to be submitted to journals.
-Finally, Chapter 6 provides some concluding remarks, discussion, ideas for future research, and so on.
-Appendixes can contain additional material that don't fit into any chapters, but that you want to put on record.
-For example, additional tables, output, etc.
+@buja2009statistical introduced a visual inference framework that formalised a hypothesis testing of graphical representations of data (henceforth referred to as the _data plot_) via the lineup protocol.
+The protocol is inspired by the police lineup technique employed in eyewitness identification of criminal suspects.
+Briefly, the protocol comprises $m$ randomly positioned plots, where one position presents the data plot, while the remaining $m-1$ plots present the plots with the same graphical structure, except that the data has been replaced with data consistent with the null hypothesis $H_0$ (henceforth referred to as _null plots_).
+To compute the $p$-value of the visual test, the lineup will be independently presented to a number of participants, asking them to pick the most different plot.
+Under $H_0$, the data plot is expected to be indistinguishable from the null plots, and the probability of correctly identifying the data plot by an observer is $1/m$.
+If a large number of participants correctly identify the data plot, the corresponding $p$-value will be small, indicating strong evidence against $H_0$.
+This protocol provides a callibration of the data plot against the null plots, ensuring that the data plot is not overinterpreted.
 
-This thesis explores innovative approaches to regression model diagnostics across three chapters.
-Chapter 2 presents evidence from a visual inference experiment using the lineup protocol to demonstrate why plotting residuals is superior to conventional numerical hypothesis tests for model diagnosis.
-The lineup protocol provides a more reliable and consistent method for interpreting residual plots, avoiding the oversensitivity of conventional tests while detecting a range of departures from good residuals simultaneously.
+The lineup protocol has gained increasing traction in recent years and has already been integrated into data analysis of various topics [see @loy2013diagnostic; @widen2016graphical; @krishnan2021hierarchical].
+However, the reliance of human assessment is a fundamental aspect of visual tests, which may restrict its widespread usage.
+The lineup protocol is unsuitable for large-scale applications, due to its high labour costs and time requirements.
+Moreover, it presents significant usability issues for individuals with visual impairments, resulting in reduced accessibility.
 
-Chapter 3 introduces a computer vision model to automate the assessment of residual plots, addressing the scalability limitations of human-based visual inference.
+To address these limitations, this thesis proposes a computer vision-based approach to automate the visual inference process for assessment of linear regression residual plots.
+Modern computer vision models often use a convolutional neural network architecture to process digital images to peform various tasks (e.g.
+object detection, object identification and signal processing).
+The development for computer vision models has primarily focused on processing natural images, such as photographs and videos, and its adaptation for data plots has some success [e.g.
+classification of time series images in @hatami2018] but generally limited in development.
+The development of computer vision models for the assessment residual plots will make the process more efficient, consistent, and accessible.
+
+## Thesis Outline
+
+The thesis is structured as follows.
+
+@sec-first-paper provides empirical evidence supporting the indispensability of residual plots through a visual inference experiment using the lineup protocol.
+By comparing human evaluations of residual plots to conventional statistical tests, this chapter demonstrates the advantages of graphical methods in detecting practical issues with model fit, while also highlighting the limitations of conventional tests in producing overly sensitive results.
+The chapter contains a comprehensive literature review related to residual diagnostics.
+
+@sec-second-paper introduces a computer vision model to automate the assessment of residual plots, addressing the scalability limitations of human-based visual inference.
 This model is trained to predict a distance measure based on Kullback-Leibler divergence, quantifying the disparity between the residual distribution of a fitted classical normal linear regression model and the reference distribution.
-Simulation studies show that the computer vision model strikes a balance between conventional tests and human visual tests in terms of sensitivity, though it is slightly less effective on non-linearity patterns.
+Performance of the model is evaluated on the human subject experiment data collected in @sec-first-paper.
+A comprehensive literature review of data plots reading with computer vision models is contained in the chapter.
 
-Chapter 4 describes the development of an R package that implements the computer vision model for automated assessment of residual plots.
-The package includes a shiny app for ease of use, allowing analysts to input a sample of residuals and receive predictions of visual signal strength (VSS) along with supporting information to assist in diagnosing model fit.
-This tool aims to bridge the gap between the need for effective diagnostics and the limitations of conventional tests and human effort.
+@sec-third-paper introduces a new R package, `autovi`, and its accompanying web interface, `autovi.web`, designed to automate the assessment of residual plots in regression analysis.
+The package uses a computer vision model built in @sec-second-paper to predict a measure of visual signal strength (VSS) and provides supporting information to assist analysts in diagnosing model fit.
+By automating this process, `autovi` and `autovi.web` improve the efficiency and consistency of model evaluation, making advanced diagnostic tools accessible to a broader audience.
+
+@sec-conclusion summarises the contribution of the work and their impact, and discusses some future plans.

### 2024-08-25 Thesis/01-chap1.qmd
fill in last part

@@
 
-The lineup protocol has gained increasing traction in recent years and has already been integrated into data analysis of various topics [see @loy2013diagnostic; @widen2016graphical; @krishnan2021hierarchical].
+The lineup protocol has gained increasing traction in recent years and has already been integrated into data analysis of various topics [see @krishnan2021hierarchical; @loy2013diagnostic; @widen2016graphical].
 However, the reliance of human assessment is a fundamental aspect of visual tests, which may restrict its widespread usage.

### 2024-08-25 Thesis/05-chap5.qmd
rearrange the software contribution

@@
-# Conclusion and future plans
+# Conclusion and future plans {#sec-conclusion}
 
-The three pieces of work in this thesis focus on ...
+The three pieces of work assembled in this thesis share a common theme of advancing regression diagnostics, with a focus on improving the assessment of residual plots, challenging the limitations of conventional methods, and developing innovative solutions to automate diagnostic processes.
 
@@
 
-The primary contributions of this research is ....
+The primary contributions of this research are threefold.
+Firstly, we provide empirical evidence for the effectiveness of lineup protocol in diagnosing model fit issues through residual plots [@li2024plot].
+Secondly, we develop computer vision model for automating the assessment of residual plots, which addresses the scalability limitations of lineup protocol.
+Lastly, we introduce a user-friendly R package and Shiny app, making the automated diagnostic tools accessible to a broad range of analysts and practitioners.
 
-Principles of transparency and reproducible research have guided the work. The materials to reproduce this thesis are available at .... and software packages are:
+The aforementioned R package (and its dependency) is available on CRAN with the latest developments in the links below:
 
-- software 1
-- software 2
+- `autovi` (<https://github.com/TengMCing/autovi>), and
+- `bandicoot` (<https://github.com/TengMCing/bandicoot>).
 
-...
+The Shiny app for `autovi` is available at one of the mirror sites listed at <https://autoviweb.netlify.app/> with the source code available at <https://github.com/TengMCing/autovi_web>.
+
+Principles of transparency and reproducible research have guided the work with all work available in public and private repositories at <https://github.com/TengMCing>.
+The thesis is written using Quarto [@Allaire_Quarto_2024] and also available online at <https://patrick-li-thesis.netlify.app>.
 
@@
 
-The limitations of the current are ... We have explored these other ways to make the model ... Some new possibilities are ... might address ...
+The model in @sec-third-paper, trained on standard residual plots from linear regression, has certain limitations.
+Residual plots from more complex models, such as hierarchical, temporal, or spatial regression, often exhibit distinct visual patterns that the current approach might not fully capture.
+To better address this, using scaled residual plots with randomized quantile residuals [@dunn1996randomized] offers a more flexible approach for defining residuals across different regression models, though it may alter the original visual pattern.
+Building a computer vision model on this foundation can provide a stronger solution for assessing residual plots across a broader spectrum of regression models.
+While the current implementation relies on the basic VGG16 model [@simonyan2014very], developed a decade ago, performance could be enhanced by exploring more advanced models like ResNet50 [@he2016deep] and DenseNet201 [@huang2017densely], as well as ensemble techniques, to improve the accuracy and effectiveness of residual plot assessments.
 
-Residual diagnostics are also used in more complex statistical models such as hierarchical, or temporal or spatial ... These require adaptations of residual calculations ...
-
-Visual diagnostics are also common in Bayesian modeling to assess ... (eg Gelman, A. 2004 Exploratory data analysis for complex models. J. Comput. Graph. Stat. 13(4), 755–779.)
+Visual diagnostics are also common in Bayesian modelling to assess model fit, convergence, and posterior distributions.
+Some common visual diagnostics used in Bayesian modelling include trace plots to assess convergence of Markov Chain Monte Carlo (MCMC) chains, density plots to visualize posterior distributions, posterior predictive checks to assess model fit, and autocorrelation plots to assess dependence between samples in MCMC chains.
+These visual diagnostics help Bayesian modelers to evaluate the quality of their models and identify potential issues or areas for improvement.
+Automating the reading of these plots can help improve MCMC convergence diagnostics, facilitate model comparison and selection, and enhance uncertainty quantification.

### 2024-08-25 Thesis/03-chap3.qmd
use a different sentence for convolutional layers
give more details about input format

@@
 They rely on deep neural networks with convolutional layers [@fukushima1982neocognitron].
-These layers leverage hierarchical patterns in data, downsizing and transforming images by summarizing information in a small space.
+These layers use small, sliding windows to scan the image, performing a dot product to extract local features and patterns.
 Numerous studies have demonstrated the efficacy of convolutional layers in addressing various vision tasks, including image recognition [@rawat2017deep].
@@
 
-A straightforward design involves feeding a vector of residuals along with a vector of fitted values, essentially providing all the necessary information for creating a residuals vs fitted values plot.
+A straightforward architecture of the input layer involves feeding a vector of residuals along with a vector of fitted values, essentially providing all the necessary information for creating a residuals vs fitted values plot.
 However, a drawback of this method is the dynamic input size, which changes based on the number of observations.
@@
 
-Another design involves using an image as input.
+Another architecture of the input layer involves using an image as input.
 The primary advantage of this design, as opposed to the vector format, is the availability of the existing and sophisticated image processing architectures developed over the years, such as the VGG16 architecture proposed in @simonyan2014very.
 These architectures can effectively capture and summarize spatial information from nearby pixels, which is less straightforward with vector input.
+While encoding data points as pixels might lead to some loss of detail, this approach still provides a more comprehensive representation than the histogram-based strategy because it captures the joint distribution of fitted values and residuals.
 The main considerations are the image resolution and the aesthetics of the residual plot.
@@
 A triplet constitutes a set of three images, denoted as $image_1$, $image_2$ and $image_3$.
-It is often used to predict whether $image_2$ or $image_3$ is more similar to $image_1$, proving particularly useful for establishing rankings between samples.
+It is often used to predict whether $image_2$ or $image_3$ is more similar to $image_1$, proving particularly useful for establishing rankings b
+etween samples.
 For this setup, we can apply the same criteria to define similarity between images.

### 2024-08-25 Thesis/03-chap3.qmd
modify reason for using scagnostics

@@
 A triplet constitutes a set of three images, denoted as $image_1$, $image_2$ and $image_3$.
-It is often used to predict whether $image_2$ or $image_3$ is more similar to $image_1$, proving particularly useful for establishing rankings b
-etween samples.
+It is often used to predict whether $image_2$ or $image_3$ is more similar to $image_1$, proving particularly useful for establishing rankings between samples.
 For this setup, we can apply the same criteria to define similarity between images.
@@
 While other scagnostics provide valuable insights, they come with high computational costs and are not suitable for quick inference.
+
+For each residual plot used as an input image, we calculated four scagnostics — "Monotonic", "Sparse", "Splines", and "Striped" – using the `cassowaryr` R package [@mason2022cassowaryr].
+These computed measures, along with the number of observations from the fitted model, were provided as the second input for the computer vision model.
+We selected these scagnostics due to their reliability and efficiency, as other scagnostics occasionally caused R process crashes (approximately 5% of the time) during training data preparation, due to a bug in the `interp` R package [@Albrecht2023interp].
+Although the package maintainer later fixed this bug at our request, the fix came too late to retrain the model, and additionally, their high computational costs make them unsuitable for rapid inference, which was a critical factor in our choice.
 

### 2024-08-25 Thesis/03-chap3.qmd
imporve lang in residual dist

@@
 One of the assumptions of the classical normal linear regression model is that the error $\boldsymbol{e}$ follows a multivariate normal distribution with zero mean and constant variance, i.e., $\boldsymbol{e} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)$.
-It can be known that residuals $\hat{\boldsymbol{e}}$ also follow a certain probability distribution transformed from the multivariate normal distribution, which will be denoted as $Q$.
+It follows that the distribution of residuals $\hat{\boldsymbol{e}}$ can be characterized by a certain probability distribution, denoted as $Q$, which is transformed from the multivariate normal distribution.
 This reference distribution $Q$ summarizes what "good" residuals should follow given the design matrix $\boldsymbol{X}$ is known and fixed.

### 2024-08-25 Thesis/05-chap5.qmd

@@
 The primary contributions of this research are threefold.
-Firstly, we provide empirical evidence for the effectiveness of lineup protocol in diagnosing model fit issues through residual plots [@li2024plot].
+Firstly, we provide empirical evidence for the effectiveness of the lineup protocol in diagnosing model fit issues through residual plots [@li2024plot].
 Secondly, we develop computer vision model for automating the assessment of residual plots, which addresses the scalability limitations of lineup protocol.
-Lastly, we introduce a user-friendly R package and Shiny app, making the automated diagnostic tools accessible to a broad range of analysts and practitioners.
+Lastly, we share a user-focused R package and Shiny app, making the automated diagnostic tools accessible to a broad range of analysts and practitioners.
 
-The aforementioned R package (and its dependency) is available on CRAN with the latest developments in the links below:
+The aforementioned R package (and its dependency) is available on CRAN with the latest development versions in the links below:
 
@@
 
-Principles of transparency and reproducible research have guided the work with all work available in public and private repositories at <https://github.com/TengMCing>.
-The thesis is written using Quarto [@Allaire_Quarto_2024] and also available online at <https://patrick-li-thesis.netlify.app>.
+Principles of transparency and reproducible research have guided the work with all source code available in public and private repositories at <https://github.com/TengMCing>.
+The thesis is written using Quarto [@Allaire_Quarto_2024] and is available online at <https://patrick-li-thesis.netlify.app>.
 
@@
 
-There are several directions that this work can be developed: alternative computer vision models, and extensions to different types of visual diagnostics.
+There are several directions that this work can be developed.
+These include improving the accuracy and effectiveness of residual plot assessments, exploring the use of alternative computer vision models, extending the automated visual diagnostics to different plot types and statistical models, and improving the front-end display and back-end computation for the web app.
 
 The model in @sec-third-paper, trained on standard residual plots from linear regression, has certain limitations.
+It was very difficult to arrive at a final model to share, with the main concern is that the current version may still be too sensitive, leading to a decision that the model is misspecified even when problems are minor.
+While the current implementation relies on the basic VGG16 model [@simonyan2014very], developed a decade ago, performance could be enhanced by exploring more advanced versions like ResNet50 [@he2016deep] and DenseNet201 [@huang2017densely], as well as ensemble techniques.
+There is room to improve the accuracy and effectiveness of residual plot assessments.
+
 Residual plots from more complex models, such as hierarchical, temporal, or spatial regression, often exhibit distinct visual patterns that the current approach might not fully capture.
-To better address this, using scaled residual plots with randomized quantile residuals [@dunn1996randomized] offers a more flexible approach for defining residuals across different regression models, though it may alter the original visual pattern.
+To better address this, using scaled residual plots, with randomized quantile residuals [@dunn1996randomized], offers a more flexible approach for defining residuals across different regression models, though it may alter the original visual pattern.
 Building a computer vision model on this foundation can provide a stronger solution for assessing residual plots across a broader spectrum of regression models.
-While the current implementation relies on the basic VGG16 model [@simonyan2014very], developed a decade ago, performance could be enhanced by exploring more advanced models like ResNet50 [@he2016deep] and DenseNet201 [@huang2017densely], as well as ensemble techniques, to improve the accuracy and effectiveness of residual plot assessments.
 
-Visual diagnostics are also common in Bayesian modelling to assess model fit, convergence, and posterior distributions.
-Some common visual diagnostics used in Bayesian modelling include trace plots to assess convergence of Markov Chain Monte Carlo (MCMC) chains, density plots to visualize posterior distributions, posterior predictive checks to assess model fit, and autocorrelation plots to assess dependence between samples in MCMC chains.
+Visual diagnostics are foundational in Bayesian modelling to assess model fit, convergence, and posterior distributions (see @gelman2013).
+Some common visual diagnostics include trace plots to assess convergence of Markov Chain Monte Carlo (MCMC) chains, density plots to visualize posterior distributions, posterior predictive checks to assess model fit, and autocorrelation plots to assess dependence between samples in MCMC chains.
 These visual diagnostics help Bayesian modelers to evaluate the quality of their models and identify potential issues or areas for improvement.
 Automating the reading of these plots can help improve MCMC convergence diagnostics, facilitate model comparison and selection, and enhance uncertainty quantification.
+
+The development of a more comprehensive suite of automated visual diagnostics for statistical models can help to improve the quality of statistical analyses.
+Also important is the development of user-friendly interfaces for these diagnostics, such as web applications, to make them accessible to a wider audience of researchers and practitioners.
+The web app developed in this thesis is a step in this direction, but further work is needed to improve the user experience, add more features, and make the app more robust and scalable.
+Future work could also explore the use of interactive visualizations and dashboards to help users explore and interpret the results of automated visual diagnostics more effectively.

### 2024-08-25 Thesis/03-chap3.qmd
provide better reason for vgg16

@@
 
-The architecture of the computer vision model is adapted from a well-established architecture known as VGG16, which has demonstrated high performance in image classification [@simonyan2014very].
+The architecture of the computer vision model is adapted from the well-established VGG16 architecture [@simonyan2014very].
+While more recent architectures like ResNet [@he2016deep] and DenseNet[@huang2017densely], have achieved even greater performance, VGG16 remains a solid choice for many applications due to its simplicity and effectiveness.
+Our decision to use VGG16 aligns with our goal of starting with a proven and straightforward model.
 @fig-cnn-diag provides a diagram of the architecture.
@@
 This is followed by a grayscale conversion layer utilizing the luma formula under the Rec.
-601 standard, which converts the colour image to grayscale.
+601 standard [@series2011studio], which converts the colour image to grayscale.
 Grayscale suffices for our task since data points are plotted in black.

### 2024-08-26 Thesis/04-chap4.qmd
make shiny -> Shiny, and add right citation for ecostats

@@
 This is the type of task where a robot might be employed to do the disagreeable work that currently requires a human.
-Here we describe a new R package that includes a computer vision model for automated assessment of residual plots, and an accompanying shiny app for ease of use.
+Here we describe a new R package that includes a computer vision model for automated assessment of residual plots, and an accompanying Shiny app for ease of use.
 For a user-provided sample of residuals, it predicts a measure of visual signal strength (VSS) and provides a suite of supporting information to assist the analyst diagnose their model fit.
@@
 Enhanced visual diagnostics forms the second group, which offer advanced visual aids for interpreting diagnostic plots.
-This includes `ecostats` [@ecostats], which adds simulation envelopes to  residual plots, and `DHARMa` [@dharma], which compares the empirical 0.25, 0.5, and 0.75 quantiles in scaled residuals with their theoretical counterparts.
+This includes `ecostats` [@warton_global_2023], which adds simulation envelopes to  residual plots, and `DHARMa` [@dharma], which compares the empirical 0.25, 0.5, and 0.75 quantiles in scaled residuals with their theoretical counterparts.
 `DHARMa`'s focus is detecting model violations such as heteroscedasticity and incorrect functional forms, and uncovering issues specific to generalized linear models and mixed-effect models, like over/under-dispersion.

### 2024-08-26 Thesis/04-chap4.qmd
update chap4 according to emi's comment and del some content

--- 2024-08-23 Thesis/04-chap4.qmd
@@
 
-Regression analysis is a fundamental statistical technique widely used for modeling data from many fields.
+Regression analysis is a fundamental statistical technique widely used for modelling data from many fields.
 To diagnose the fit of a model it is recommended that the residuals are plotted.
@@
 General-purpose regression analysis tools are the largest and most commonly used group.
-These packages aren't specifically designed for graphical diagnostics of residuals in linear regression but offer this functionality as part of a broader set of statistical tools.
+These packages are not specifically designed for graphical diagnostics of residuals in linear regression but offer this functionality as part of a broader set of statistical tools.
 These include the `stats` package [@stats], which includes common diagnostic plots like residuals vs fitted values, quantile-quantile (Q-Q) plots, and residuals vs leverage plots.
@@
 The name `autovi` stands for automated visual inference.
-While initially designed for linear regression residual diagnostics, it has the potential to be extended to broader visual inference applications, as we'll discuss in section @sec-autovi-infrastructure.
+While initially designed for linear regression residual diagnostics, it has the potential to be extended to broader visual inference applications, as we discuss in section @sec-autovi-infrastructure.
 
@@
 
-### Usage {#sec-autovi-quick-start}
+### Usage {#sec-autovi-usage}
 
-To get started quickly, users need only three lines of code to obtain a summary of the automated residual assessment:
+To get started quickly, only calls two functions are necessary to obtain a summary of the automated residual assessment:
 
@@
 3.
-Use `get_keras_model()`, a function provided by `autovi`, to download a trained computer vision model (described in @sec-second-paper) from GitHub.
-"vss_phn_32" specifies a model that predicts `vss` and is trained on residuals with polynomial, heteroskedasticity, and non-normality patterns (phn).
-More details about the hosted models will be provided in section @sec-trained-model-hosting.
-4. Call the `check()` method of the checker with default arguments. This predicts the VSS for the true residual plot, 100 null plots, and 100 bootstrapped plots, storing the predictions internally.
-5. Use the `print()` function to generate a concise report of the check results.
+Call the `check()` method of the checker with default arguments.
+This predicts the VSS for the true residual plot, 100 null plots, and 100 bootstrapped plots, storing the predictions internally.
+A concise report of the check results will be printed.
 
 The summary reports key findings such as the VSS of the true residual plot and the $p$-value of the automated visual test.
-The $p$-value is the ratio of null plots having VSS greater than or equal to the true residual plot.
-We typically reject the null hypothesis when the $p$-value is smaller than or equal to 5%.
+The $p$-value is the ratio of null plots that have VSS greater than or equal to that of the true residual plot.
+We typically reject the null hypothesis when the $p$-value is smaller than or equal to a desired significance level, such as 5%.
 The report also provides sample quantiles of VSS for null and bootstrapped plots, helping to explain the severity and likelihood of model violations.
 
-Although the $p$-value is sufficient for automated decision-making, users are strongly encouraged to visually inspect the original residual plot alongside a sample null plot.
+Although the $p$-value is sufficient for automated decision-making, users can visually inspect the original residual plot alongside a sample null plot.
 This visual comparison can clarify why $H_0$ is either rejected or not, and help identify potential remedies.
@@
 
-This method (@fig-plot-pair) displays the true residual plot on the left and a single null plot on the right.
+The `plot_pair()` method (@fig-plot-pair) displays the true residual plot on the left and a single null plot on the right.
 If a full lineup was shown, the true residual plot would be embedded in a page of null plots.
@@
 The remaining seven modules are designed to be overridable, enabling users to tailor the infrastructure to their specific needs.
-These modules will be discussed in detail in the following sections.
+These modules are discussed in detail in the following sections.
 
@@
 An `autovi` checker can be initialized by supplying two primary inputs, including a regression model object, such as an `lm` object representing the result of a linear regression model, and a trained computer vision model compatible with the `Keras` [@chollet2015keras] Application Programming Interface (API), to the `AUTO_VI` class constructor `auto_vi()`.
-The input will be stored in the checker and can be accessed by the user through the `$` operator.
+The `residual_checker()` introduced in @sec-autovi-usage is a thin wrapper around `auto_vi()`, which will call `get_keras_model()` during initialization.
+`get_keras_model()` is a function to download a trained computer vision model (described in @sec-second-paper) from GitHub.
+“vss_phn_32” specifies a model that predicts VSS and is trained on residuals with polynomial, heteroskedasticity, and non-normality patterns (phn).
+More details about the hosted models will be provided in @sec-trained-model-hosting.
+
+The input of the constructor will be stored in the checker and can be accessed by the user through the `$` operator.
 
@@
 
-By design, the `get_fitted_and_resid()` method accepts a regression model object as input and returns a `tibble` with two columns: `.fitted` and `.resid`, representing the fitted values and residuals, respectively.
+By design, the `get_fitted_and_resid()` method accepts a regression model object as input and returns a `tibble` (a modern presentation of the `data.frame`) with two columns: `.fitted` and `.resid`, representing the fitted values and residuals, respectively.
 If no input is supplied, the method uses the regression model object stored in the checker.
@@
 Typically, the model frame can be extracted from the regression model object using the `model.frame()` generic function, which is the default method used by `get_data()`.
-However, some regression models don't use a formula or are evaluated differently, potentially lacking a model frame.
+However, some regression models do not use a formula or are evaluated differently, potentially lacking a model frame.
 In such cases, users can either provide the data used to fit the regression model through the `data` argument when constructing the checker, or customize the method to better suit their needs.
@@
 
-According to @sec-second-paper, in some cases, a residual plot alone may not provide enough information to accurately determine VSS.
-For instance, when the residual plot has significant overlap, the trend and shape of the residual pattern can be difficult to discern.
+As described in @sec-second-paper, in some cases, a residual plot alone may not provide enough information to accurately determine VSS.
+For instance, when the points in the residual plot have significant overlap, the trend and shape of the residual pattern can be difficult to discern.
 Including auxiliary variables, such as the number of observations, as additional inputs to the computer vision model can be beneficial.
@@
 Technically, any Keras-implemented computer vision model can be adapted to accept an image as the primary input and additional variables as secondary inputs by adding a data pre-processing layer before the actual input layer.
-If users wish to override `auxiliary()`, the output should be a `data.frame` with a single row and the number of columns matching the supplied Keras model.
+If users wish to override `auxiliary()`, the output should be a `data.frame` with a single row and the number of columns such that its concatenation matches the number of parameters for the corresponding layer in the supplied Keras model.
 
@@
 This resolution was chosen because the original set, consisting of 20 residual plots arranged in a four by five grid, was represented by an image of 2100 by 2100 pixels.
-The `save_plot()` method takes a `ggplot` object as input, saves it as a temporary PNG file, and returns the file path as a string.
-Note that the `save_plot()` method does not have default arguments, as it is not intended to be called without a plot.
-While an alternative design could be to save the true residual plot by default, this might be confusing for users, given that the method's name does not fully convey this functionality.
+The `save_plot()` method accepts a `ggplot` object as input and saves it as a PNG file to the location specified by the `path` argument.
+If no path is provided, the PNG file is saved to a temporary file.
 
@@
 
-#### $P$-value Computation
+#### $p$-value Computation
 
@@
 After executing the `check()` method, `autovi` offers two visualization options for the assessment result through the `summary_plot()` method, including the density plot and the rank plot.
-We have already discussed and interpreted the density plot in an earlier section.
+We have already discussed and interpreted the density plot in @sec-autovi-usage.
 Here, we would like to highlight the flexibility in choosing which elements to display in the density plot.
@@
 
-The rank plot, creating by setting `type` to "rank", is a bar plot where the x-axis represents the rank and the y-axis shows the VSS.
-The bar for the true residual plot is colored in red.
+The rank plot, creating by setting `type` to "rank", is a bar plot where the $x$-axis represents the rank and the $y$-axis shows the VSS.
+The bar that is colored in red correspoding to the VSS of the true residual plot.
 By examining the rank plot, you can intuitively understand how the observed VSS compares to the null VSSs and identify any outliers in the null distribution.
@@
 
-The package provides tools for analyzing these extracted features through the `feature_pca()` method and its associated visualization method, `feature_pca_plot()`.
-The `feature_pca()` method performs principal component analysis (PCA) on the features to reduce their dimensionality.
-However, it requires that a `check()` is performed first, as it relies on results stored in `check_result`.
-Alternatively, you can manually provide features using the `feature`, `null_feature`, and `boot_feature` arguments for the true residual plot, null plots, and bootstrapped plots, respectively.
-The `feature_pca()` method returns a tibble containing both the original features and the principal components.
-The rotation matrix and standard deviations of each principal component are stored as attributes.
-
-```r
-checker$check(null_draws = 100L,
-              boot_draws = 100L,
-              extract_feature_from_layer = "global_max_pooling2d")
-checker$feature_pca()
-```
-
-```{r}
-if (!file.exists("cached_data/autovi_check_demo_8.rds")) {
-  checker$keras_model <- get_keras_model("vss_phn_32") 
-  checker$check(null_draws = 100L,
-                boot_draws = 100L,
-                extract_feature_from_layer = "global_max_pooling2d")
-  saveRDS(checker$check_result, "cached_data/autovi_check_demo_8.rds")
-} else {
-  checker$check_result <- readRDS("cached_data/autovi_check_demo_8.rds")
-}
-checker$feature_pca()
-```
-
-The `feature_pca_plot()` method visualizes the results of the PCA.
-By default, it plots the first principal component on the x-axis and the second principal component on the y-axis, with points colored according to their origin, true residual plots, null residual plots, or bootstrapped residual plots.
-Users can customize the x and y axes by specifying symbols for the `x` and `y` arguments.
-Additionally, the `col_by_set` option can be disabled if you prefer not to use coloring.
-
-```{r echo = TRUE}
-checker$feature_pca_plot()
-```
-
-When interpreting the principal component scatter plot, look for any outliers within the null or bootstrapped groups.
-Assess whether the null group and the bootstrapped group form a single cluster or distinct clusters.
-Additionally, evaluate whether the observed point is distinct from the null group.
-
 ### Trained Model Hosting {#sec-trained-model-hosting}
@@
 It requires only the model name, which is the value in the first column of the `tibble` returned by `list_keras_model()`.
-
-### Extending the `AUTO_VI` class
-
-The `bandicoot` software is a lightweight object-oriented system with Python-like syntax that supports multiple inheritance and incorporates a Python-like method resolution order.
-The system is inspired by the OOP frameworks implemented in R6 [r6] and Python.
-In this section, we will provide essential details for extending the `autovi::AUTO_VI` class using `bandicoot`.
-
-In `bandicoot`, a class is declared using the `bandicoot::new_class()` function, where parent classes are provided as positional arguments, and the class name is specified through the `class_name` argument.
-The output of `bandicoot::new_class()` is an environment with the S3 class `bandicoot_oop`.
-Printing a `bandicoot` object provides a summary of the object, which can be customized via the `..str..` magic method.
-
-```{r}
-EXT_AUTO_VI <- bandicoot::new_class(AUTO_VI, class_name = "EXT_AUTO_VI")
-print(EXT_AUTO_VI)
-```
-
-An extended class inherits attributes and methods from its parent class(es), so it will behave similarly to them. This can be verified using the built-in `names()` function.
-
-```{r}
-names(EXT_AUTO_VI)
-```
-
-To register a method for an extended class, you need to pass the class as the first argument and the method as a named argument to the `bandicoot::register_method()` function.
-Within a method, `self` can be used as a reference to the class or object environment.
-The following code example overrides the `null_method()` with a function that simulates null residuals from the corresponding normal distribution.
-This approach differs from the default null residual simulation scheme described earlier.
-Although less efficient than the default method for linear regression models, it provides an alternative way to simulate null residuals.
-This method is particularly useful when the fitted model is unavailable, and only the fitted values and residuals are accessible, as discussed in @sec-autovi-web.
-
-```{r}
-bandicoot::register_method(
-  EXT_AUTO_VI, 
-  null_method = function(fitted_model = self$fitted_model) {
-    data <- self$get_fitted_and_resid(fitted_model)
-    residual_sd <- sd(data$.resid)
-    data$.resid <- rnorm(nrow(data),
-                         sd = residual_sd)
-    return(data)
-  }
-)
-
-print(EXT_AUTO_VI$null_method(lm(dist ~ speed, data = cars)))
-```
-
-To create an object in `bandicoot`, you need to call the `instantiate()` method of a class.
-Alternatively, you can build a convenient class constructor for your class.
-It is recommended to provide the full list of arguments in the class constructor instead of using `...`, as this makes it easier for integrated development environments (IDEs) like RStudio to offer argument completion hints to the user.
-
-```{r}
-ext_auto_vi <- function(fitted_model,
-                        keras_model = NULL,
-                        data = NULL,
-                        node_index = 1L,
-                        env = new.env(parent = parent.frame()),
-                        init_call = sys.call()) {
-  EXT_AUTO_VI$instantiate(fitted_model = fitted_model,
-                          keras_model = keras_model,
-                          data = data,
-                          node_index = node_index,
-                          env = env,
-                          init_call = init_call)
-}
-
-print(ext_auto_vi(lm(dist ~ speed, data = cars)))
-```
 

### 2024-08-26 Thesis/C-appC.qmd
update chap4 according to emi's comment and del some content

# Appendix to "Software for Automated Residual Plot Assessment: autovi and autovi.web" {#sec-appendix-c}

## Extending the `AUTO_VI` class

The `bandicoot` R package provides a lightweight object-oriented system with Python-like syntax that supports multiple inheritance and incorporates a Python-like method resolution order. The system is inspired by the OOP frameworks implemented in R6 [r6] and Python. In this section, we will provide essential details for extending the `autovi::AUTO_VI` class using `bandicoot`.

In `bandicoot`, a class is declared using the `bandicoot::new_class()` function, where parent classes are provided as positional arguments, and the class name is specified through the `class_name` argument. The output of `bandicoot::new_class()` is an environment with the S3 class `bandicoot_oop`. Printing a `bandicoot` object provides a summary of the object, which can be customized via the `..str..` magic method.

```{r}
EXT_AUTO_VI <- bandicoot::new_class(AUTO_VI, class_name = "EXT_AUTO_VI")
print(EXT_AUTO_VI)
```

An extended class inherits attributes and methods from its parent class(es), so it will behave similarly to them. This can be verified using the built-in `names()` function.

```{r}
names(EXT_AUTO_VI)
```

To register a method for an extended class, you need to pass the class as the first argument and the method as a named argument to the `bandicoot::register_method()` function. Within a method, `self` can be used as a reference to the class or object environment. The following code example overrides the `null_method()` with a function that simulates null residuals from the corresponding normal distribution. This approach differs from the default null residual simulation scheme described earlier. Although less efficient than the default method for linear regression models, it provides an alternative way to simulate null residuals. This method is particularly useful when the fitted model is unavailable, and only the fitted values and residuals are accessible, as discussed in @sec-autovi-web.

```{r}
bandicoot::register_method(
  EXT_AUTO_VI, 
  null_method = function(fitted_model = self$fitted_model) {
    data <- self$get_fitted_and_resid(fitted_model)
    residual_sd <- sd(data$.resid)
    data$.resid <- rnorm(nrow(data),
                         sd = residual_sd)
    return(data)
  }
)

print(EXT_AUTO_VI$null_method(lm(dist ~ speed, data = cars)))
```

To create an object in `bandicoot`, you need to call the `instantiate()` method of a class. Alternatively, you can build a convenient class constructor for your class. It is recommended to provide the full list of arguments in the class constructor instead of using `...`, as this makes it easier for integrated development environments (IDEs) like RStudio to offer argument completion hints to the user.

```{r}
ext_auto_vi <- function(fitted_model,
                        keras_model = NULL,
                        data = NULL,
                        node_index = 1L,
                        env = new.env(parent = parent.frame()),
                        init_call = sys.call()) {
  EXT_AUTO_VI$instantiate(fitted_model = fitted_model,
                          keras_model = keras_model,
                          data = data,
                          node_index = node_index,
                          env = env,
                          init_call = init_call)
}

print(ext_auto_vi(lm(dist ~ speed, data = cars)))
```

### 2024-08-26 Thesis/04-chap4.qmd
add caption for every plot in chap 4

@@
 
-To get started quickly, only calls two functions are necessary to obtain a summary of the automated residual assessment:
+To get started quickly, only calls three functions are necessary to obtain a summary of the automated residual assessment:
 
@@
 1. Load the package using the `library()` function.
-2. Construct a checker with two inputs: a linear regression model and a pre-trained Keras model [@chollet2015keras].
+2. Construct a checker with a linear regression model.
 3.
@@
 
-The method returns a `ggplot` object, which can be saved as a PNG file in the following module.
+The method returns a `ggplot` object (@fig-autovi-plot-resid), which can be saved as a PNG file in the following module.
 If no data is provided, the method will use `get_fitted_and_resid()` to retrieve the fitted values and residuals from the regression model stored in the checker.
 
-```{r echo=TRUE, fig.height=3, fig.width=4}
+```{r fig-autovi-plot-resid, echo=TRUE, fig.height=3, fig.width=4, fig.cap = "Residual plot of the regression model stored in the checker."}
 checker$plot_resid()
@@
 To manually generate true residual plots, null plots, or bootstrapped residual plots, you can pass the corresponding `data.frame` produced by the `get_fitted_and_resid()`, `null_method()`, and `boot_method()` methods to the `plot_resid()` method, respectively.
-
-```{r echo = TRUE, fig.height=3, fig.width=4}
-checker$null_method() |>
-  checker$plot_resid()
-```
 
@@
 We have already discussed and interpreted the density plot in @sec-autovi-usage.
-Here, we would like to highlight the flexibility in choosing which elements to display in the density plot.
+Here, we would like to highlight the flexibility in choosing which elements to display in the density plot as shown in @fig-autovi-summary-plot-option.
 For instance, you can omit the bootstrapped distribution by setting `boot_dist` to `NULL`.
@@
 
-```{r echo=TRUE}
+```{r fig-autovi-summary-plot-option, echo=TRUE, fig.cap = "Density plot of the VSS for null plots."}
 checker$summary_plot(boot_dist = NULL,
@@
 
-The rank plot, creating by setting `type` to "rank", is a bar plot where the $x$-axis represents the rank and the $y$-axis shows the VSS.
-The bar that is colored in red correspoding to the VSS of the true residual plot.
+The rank plot (@fig-autovi-rank-plot), creating by setting `type` to "rank", is a bar plot where the $x$-axis represents the rank and the $y$-axis shows the VSS.
+The bar that is coloured in red corresponding to the VSS of the true residual plot.
 By examining the rank plot, you can intuitively understand how the observed VSS compares to the null VSSs and identify any outliers in the null distribution.
 
-```{r echo=TRUE}
+```{r fig-autovi-rank-plot, echo=TRUE, fig.cap = "Density plot of the VSS for null plots."}
 checker$summary_plot(type = "rank")
@@
 
-<!-- In @sec-autovi-implementation, we discussed how `autovi` relies on several Python libraries, with a particularly strong dependency on `TensorFlow`.
-Managing a Python environment and correctly installing `TensorFlow` on a local machine can be challenging for many users.
-Moreover, `TensorFlow` is a massive library that undergoes continuous development, which inevitably leads to compatibility issues arising from differences in library versions.
-These challenges can create significant barriers for users who want to perform residual assessments with the `autovi` package.
-
-Recognizing these potential barriers, we were motivated to design and implement a web interface called `autovi.web`.
-This web-based solution offers several major advantages.
-First, it eliminates dependency issues, so users no longer need to struggle with complex Python environments or worry about installing and maintaining the correct versions of libraries.
-The web interface handles all these dependencies on the server side.
-Second, `autovi.web` lowers the entry barrier by being user-friendly and accessible to individuals who may not be familiar with R programming.
-This broadens the potential user base of `autovi`, allowing more researchers and analysts to benefit from its capabilities.
-Third, the web interface can be updated centrally, ensuring that all users always have access to the latest features and improvements without needing to manage updates locally.
-Lastly, `autovi.web` offers cross-platform accessibility, allowing users to access it from any device with a web browser, increasing flexibility and convenience.
-
-`autovi.web` is available at [autoviweb.netlify.app](autoviweb.netlify.app).
-The implementation discussed in this chapter is based on `autovi.web` version 0.1.0.
-By providing this web interface, we aim to significantly reduce the technical hurdles associated with using `autovi`, making advanced residual assessment techniques more accessible to a wider audience of researchers and data analysts.
-This approach aligns with modern trends in data science tools, where web-based interfaces are increasingly used to make advanced analytical techniques more widely available.
--->
+The web interface `autovi.web` is available at [autoviweb.netlify.app](autoviweb.netlify.app). The implementation discussed in this chapter is based on `autovi.web` version 0.1.0.
 
@@
 The reason is that one of the R packages `autovi` depends on the R package `splancs` [@splancs], which uses compiled Fortran code.
-A working Emscripten (XXX reference?) version of this package, which would be required for `webr`, was not available.
+A working Emscripten [@zakai2011emscripten] version of this package, which would be required for `webr`, was not available.
 
@@
 
-While the R package `autovi` aims to provide tools that can be extended to broader visual inference applications, `autovi.web` is only focus on to for providing  a straightforward and clean user interface.
+While the R package `autovi` aims to provide tools that can be extended to broader visual inference applications, `autovi.web` is only focus on providing  a straightforward and clean user interface.
 An overview of the graphical user interface of `autovi.web` is provided in @fig-autovi-web.
@@
 The workflow of `autovi.web` is designed to be straightforward, with numbered steps displayed in each panel shown in @fig-autovi-web.
-Thre are two example datasets provided by the web application, as mentioned in @sec-autovi-web-design.
+There are two example datasets provided by the web application, as mentioned in @sec-autovi-web-design.
 The single residual plot example uses the `dino` dataset from the R package `datasauRus` [@datasaurus].

### 2024-08-26 Thesis/C-appC.qmd
add caption for every plot in chap 4

@@
 
-```{r}
+```{r echo=TRUE}
+library(autovi)
 EXT_AUTO_VI <- bandicoot::new_class(AUTO_VI, class_name = "EXT_AUTO_VI")
@@
 
-```{r}
+```{r echo=TRUE}
 names(EXT_AUTO_VI)
@@
 
-```{r}
+```{r echo=TRUE}
 bandicoot::register_method(
@@
 
-```{r}
+```{r echo=TRUE}
 ext_auto_vi <- function(fitted_model,

### 2024-08-26 Thesis/abstract.qmd
add caption for every plot in chap 4

@@
 This research presents three original contributions.
-The first contribution ([Chapter 2](#sec-first-paper)) provides evidence for the effectiveness of visual inference in regression diagnostics through a human subject experiment, demonstrating the benefits of using the lineup protocol for reliable and consistent reading of residual plots.
-The second contribution ([Chapter 3](#sec-second-paper)) introduces a computer vision model to automate the assessment of residual plots, addressing the scalability limitation of the lineup protocol.
-The third contribution ([Chapter 4](#sec-third-paper)) presents an R package and Shiny app, providing a user-friendly interface for analysts to leverage the computer vision model and supporting tools for diagnostic purposes.
+The first contribution Chapter 2 provides evidence for the effectiveness of visual inference in regression diagnostics through a human subject experiment, demonstrating the benefits of using the lineup protocol for reliable and consistent reading of residual plots.
+The second contribution Chapter 3 introduces a computer vision model to automate the assessment of residual plots, addressing the scalability limitation of the lineup protocol.
+The third contribution Chapter 4 presents an R package and Shiny app, providing a user-friendly interface for analysts to leverage the computer vision model and supporting tools for diagnostic purposes.
 These contributions advance the field of artificial intelligence for data visualization, enabling more efficient and accurate regression diagnostics.

### 2024-08-26 Thesis/B-appB.qmd
fix typo for appendix B

# Appendix to "Automated Assessment of Residual Plots with Computer vision Models" {#sec-appendix-b}

## Neural Network Layers Used in the Study

This study used seven types of neural network layers, all of which are standard components frequently found in modern deep learning models. These layers are well-documented in textbooks like @goodfellow2016deep and @chollet2021deep, which offer thorough explanations and mathematical insights. In this section, we will offer a concise overview of these layers, drawing primarily from the insights provided in @goodfellow2016deep.

### Dense Layer

The Dense layer, also known as the fully-connected layer, is the fundamental unit in neural networks. It conducts a matrix multiplication operation between the input matrix $\boldsymbol{I}$ and a weight matrix $\boldsymbol{W}$ to generate the output matrix $\boldsymbol{O}$, which can be written as

$$\boldsymbol{O} = \boldsymbol{I}\boldsymbol{W} + b,$$

where $b$ is the intercept.

### ReLu Layer

The ReLU layer, short for rectified linear unit, is an element-wise non-linear function introduced by @nair2010rectified. It sets the output elements to zero if the corresponding input element is negative; otherwise, it retains the original input. Mathematically, it can be expressed as:

$$\boldsymbol{O}(i,j) = max\{0, \boldsymbol{I}(i,j)\},$$

where $\boldsymbol{O}(i,j)$ is the $i$th row and $j$th column entry of matrix $\boldsymbol{O}$, and $\boldsymbol{I}(i,j)$ is the $i$th row and $j$th column entry of matrix $\boldsymbol{I}$.

### Convolutional Layer

In Dense layers, matrix multiplication leads to each output unit interacting with every input unit, whereas convolutional layers operate differently with sparse interactions. An output unit in a convolutional layer is connected solely to a subset of input units, and the weight is shared across all input units. Achieving this involves using a kernel, typically a small square matrix, to conduct matrix multiplication across all input units. Precisely, this concept can be formulated as:

$$\boldsymbol{O}(i, j) = \sum_m\sum_n\boldsymbol{I}(i - m, j - n)K(m, n),$$

where $m$ and $n$ are the row and columns indices of the kernel $K$.

If there are multiple kernels used in one covolutional layer, then each kernel will have its own weights and the output will be a three-dimensional tensor, where the length of the third channel is the number of kernels.

### Pooling Layer

A pooling layer substitutes the input at a specific location with a summary statistic derived from nearby input units. Typically, there are two types of pooling layers: max pooling and average pooling. Max pooling computes the maximum value within a rectangular neighborhood, while average pooling calculates their average. Pooling layers helps making the representation approximately invariant to minor translations of the input. The output matrix of a pooling layer is approximately $s$ times smaller than the input matrix, where $s$ represents the length of the rectangular area. A max pooling layer can be formulated as:

$$\boldsymbol{O}(i, j) = \underset{m,n}{\max} \boldsymbol{I}(si + m,sj+n).$$

### Global Pooling Layer

A global pooling layer condenses an input matrix into a scalar value by either extracting the maximum or computing the average across all elements. This layer acts as a crucial link between the convolutional structure and the subsequent dense layers in a neural network architecture. When convolutional layers uses multiple kernels, the output becomes a three-dimensional tensor with numerous channels. In this scenario, the global pooling layer treats each channel individually, much like distinct features in a conventional classifier. This approach facilitates the extraction of essential features from complex data representations, enhancing the network's ability to learn meaningful patterns. A global max pooling layer can be formulated as

$$O(i, j) = \underset{m,n,k}{\max} I(si + m,sj+n,k),$$
where $k$ is the kernel index.

### Batch Normalization Layer

Batch normalization is a method of adaptive reparametrization. One of the issues it adjusts is the simultaneous update of parameters in different layers, especially for network with a large number layers. At training time, the batch normalization layer normalizes the input matrix $I$ using the formula

$$\boldsymbol{O} = \frac{\boldsymbol{I} - \boldsymbol{\mu}_I}{\boldsymbol{\sigma}_I},$$

where $\boldsymbol{\mu}_I$ and $\boldsymbol{\sigma}_I$ are the mean and the standard deviation of each unit respectively.

It reparametrizes the model to make some units always be standardized by definition, such that the model training is stabilized. At inference time, $\boldsymbol{\mu}_I$ and $\boldsymbol{\sigma}_I$ are usually replaced with the running mean and running average obtained during training.

### Dropout Layer

Dropout is a computationally inexpensive way to apply regularization on neural network. For each input unit, it randomly sets to be zero during training, effectively training a large number of subnetworks simultaneously, but these subnetworks share weights and each will only be trained for a single steps in a large network. It is essentially a different implementation of the bagging algorithm. Mathematically, it is formulated as

$$\boldsymbol{O}(i,j) = \boldsymbol{D}(i,j)\boldsymbol{I}(i,j),$$

where $\boldsymbol{D}(i,j) \sim B(1, p)$ and $p$ is a hyperparameter that can be tuned.

### 2024-08-26 Thesis/04-chap4.qmd
fix the oversize image issue

@@
 
-```{r echo=TRUE}
+```r
 wrapper <- keras_wrapper(checker$keras_model)
 wrapper$list_layer_name()
+```
+
+```{r}
+if (!file.exists("cached_data/autovi_check_demo_8.rds")) {
+  wrapper <- keras_wrapper(get_keras_model("vss_phn_32"))
+  result <- wrapper$list_layer_name()
+  saveRDS(result, "cached_data/autovi_check_demo_8.rds")
+} else {
+  readRDS("cached_data/autovi_check_demo_8.rds")
+}
 ```
@@
 
-```{r fig-autovi-web-gradient-hide, out.width = "100%", fig.cap = "The attention map is hidden if the assessment indicates a $p$-value greater than 0.05.
+```{r fig-autovi-web-gradient-hide, out.width = "50%", fig.cap = "The attention map is hidden if the assessment indicates a $p$-value greater than 0.05.
 A button is available to toggle the display of the attention map."}
@@
 
-As shown in @fig-autovi-web-workflow-example, to use the lineup example data, click the "Use Lineup Example" button.
+As shown in @fig-autovi-web-workflow-example1, to use the lineup example data, click the "Use Lineup Example" button.
 The data status will then update to show the number of rows and columns in the dataset, and the CSV type will automatically be selected to the correct option.
-Since the example dataset follows the variable naming conventions assumed by the web application, the columns for fitted values, residuals, and labels of residual plots are set automatically.
+Since the example dataset follows the variable naming conventions assumed by the web application, the columns for fitted values, residuals, and labels of residual plots are set automatically (@fig-autovi-web-workflow-example2).
 If the user is working with a custom dataset, these options must be set accordingly.
@@
 
-In the result panels shown in [@fig-autovi-web-workflow-example]D, the first row of the table is the most crucial to check, as it provides the VSS and the rank of the true residual plot among the other plots.
+Results are provided in multiple panels as displayed in @fig-autovi-web-workflow-example3, @fig-autovi-web-workflow-example4, @fig-autovi-web-workflow-example5 and @fig-autovi-web-workflow-example6.
+In @fig-autovi-web-workflow-example3, the first row of the table is the most crucial to check, as it provides the VSS and the rank of the true residual plot among the other plots.
 The summary text beneath the table provides the $p$-value, which can be used for quick decision-making.
-The lineup is for manual inspection, and the user should see if the true residual plot is visually distinguishable from the other plots, to confirm if the model violation is serious.
-The density plot in [@fig-autovi-web-workflow-example]E offers a more robust result, allowing the user to compare the distribution of bootstrapped VSS with the distribution of null VSS.
-Finally, the grayscale attention map can be used to check if the target visual features, like the non-linearity present in the lineup example, are captured by the computer vision model, ensuring the quality of the assessment.
+In @fig-autovi-web-workflow-example4, the lineup is for manual inspection, and the user should see if the true residual plot is visually distinguishable from the other plots, to confirm if the model violation is serious.
+The density plot in @fig-autovi-web-workflow-example5 offers a more robust result, allowing the user to compare the distribution of bootstrapped VSS with the distribution of null VSS.
+Finally, the grayscale attention map shown in @fig-autovi-web-workflow-example6 can be used to check if the target visual features, like the non-linearity present in the lineup example, are captured by the computer vision model, ensuring the quality of the assessment.
 
-```{r fig-autovi-web-workflow-example, out.width = "100%", fig.cap = 'The workflow for `autovi.web` using the lineup example dataset proceeds as follows: (A) To begin, the user clicks the "Use Lineup Example" button to load the example dataset, during which the data status and CSV type will be automatically updated.
-(B) The target columns are selected automatically, though the user must manually select the label for the true residual plot, as the web application permits assessment without this label.
-(C) The user initiates the assessment by clicking the run button.
-(D) The VSS of the true residual plot is displayed in the first row of the table, with a summary text beneath the table providing the $p$-value to aid in decision-making.
-The lineup on the right allows for manual inspection.
-(E) The density plot helps verify if the bootstrapped distribution differs from the null distribution, while the attention map on the right offers insights into whether the computer vision model has captured the intended visual features of the true residual plot.'}
-magick::image_read_pdf("figures/autovi_web_workflow.pdf")
+```{r fig-autovi-web-workflow-example1, out.width = "80%", fig.cap = 'To begin the workflow for `autovi` using the lineup example dataset, the user clicks the "Use Lineup Example" button to load the example dataset, during which the data status and CSV type will be automatically updated.'}
+magick::image_read_pdf("figures/autovi_web_workflow_1.pdf")
+```
+
+```{r fig-autovi-web-workflow-example2, out.width = "80%", fig.cap = 'After clicking the button in @fig-autovi-web-workflow-example1, the target columns are selected automatically, though the user must manually select the label for the true residual plot, as the web application permits assessment without this label.'}
+magick::image_read_pdf("figures/autovi_web_workflow_2.pdf")
+```
+
+```{r fig-autovi-web-workflow-example3, out.width = "55%", fig.cap = 'After finishing the required steps in @fig-autovi-web-workflow-example1 and @fig-autovi-web-workflow-example2, the user initiates the assessment of the lineup example data by clicking the run button.'}
+magick::image_read_pdf("figures/autovi_web_workflow_3.pdf")
+```
+
+```{r fig-autovi-web-workflow-example4, out.width = "100%", fig.cap = 'The VSS of the true residual plot is displayed in the first row of the table, with a summary text beneath the table providing the $p$-value to aid in decision-making.'}
+magick::image_read_pdf("figures/autovi_web_workflow_4.pdf")
+```
+
+```{r fig-autovi-web-workflow-example5, out.width = "100%", fig.cap = 'A lineup of residual plots allows for manual inspection.'}
+magick::image_read_pdf("figures/autovi_web_workflow_5.pdf")
+```
+
+```{r fig-autovi-web-workflow-example6, out.width = "100%", fig.cap = 'The density plot helps verify if the bootstrapped distribution differs from the null distribution.'}
+magick::image_read_pdf("figures/autovi_web_workflow_6.pdf")
+```
+
+```{r fig-autovi-web-workflow-example7, out.width = "100%", fig.cap = 'The attention map offers insights into whether the computer vision model has captured the intended visual features of the true residual plot.'}
+magick::image_read_pdf("figures/autovi_web_workflow_7.pdf")
 ```

### 2024-08-26 Thesis/C-appC.qmd
fix the oversize image issue

@@
 
-```{r echo=TRUE}
+```{r echo=TRUE, message=TRUE}
 library(autovi)
 EXT_AUTO_VI <- bandicoot::new_class(AUTO_VI, class_name = "EXT_AUTO_VI")
-print(EXT_AUTO_VI)
+EXT_AUTO_VI
 ```
@@
 
-print(EXT_AUTO_VI$null_method(lm(dist ~ speed, data = cars)))
+EXT_AUTO_VI$null_method(lm(dist ~ speed, data = cars))
 ```
@@
 
-print(ext_auto_vi(lm(dist ~ speed, data = cars)))
+ext_auto_vi(lm(dist ~ speed, data = cars))
 ```

### 2024-08-26 Thesis/abstract.qmd
fix the oversize image issue

@@
 This research presents three original contributions.
-The first contribution Chapter 2 provides evidence for the effectiveness of visual inference in regression diagnostics through a human subject experiment, demonstrating the benefits of using the lineup protocol for reliable and consistent reading of residual plots.
-The second contribution Chapter 3 introduces a computer vision model to automate the assessment of residual plots, addressing the scalability limitation of the lineup protocol.
-The third contribution Chapter 4 presents an R package and Shiny app, providing a user-friendly interface for analysts to leverage the computer vision model and supporting tools for diagnostic purposes.
+The first contribution provides evidence for the effectiveness of visual inference in regression diagnostics through a human subject experiment, demonstrating the benefits of using the lineup protocol for reliable and consistent reading of residual plots.
+The second contribution introduces a computer vision model to automate the assessment of residual plots, addressing the scalability limitation of the lineup protocol.
+The third contribution presents an R package and Shiny app, providing a user-friendly interface for analysts to leverage the computer vision model and supporting tools for diagnostic purposes.
 These contributions advance the field of artificial intelligence for data visualization, enabling more efficient and accurate regression diagnostics.
@@
 The inclusion of co-authors reflects the fact that the work came from active collaboration between researchers and acknowledges input into team-based research.
-In the case of [Chapter 2](#sec-first-paper) and [Chapter 4](#sec-third-paper), my contribution to the work involved the following:
+In the case of [Chapter 2](#sec-first-paper), my contribution to the work involved the following:
 
@@
 Tanaka 5%, \n S.
-VanderPlas 5%", "No",
+VanderPlas 5%", "No"
   # 3, "Automated Assessment of Residual Plots with Computer Vision Models", "Unsubmitted", "80% Concept, Analysis, Software, Writing", "D.
@@
 Langrené 2.5%", "No",
-  4, "Software for Automated Residual Plot Assessment: autovi and autovi.web", "The R package autovi on CRAN", "80% Concept, Analysis, Software, Writing", "D.
+  # 4, "Software for Automated Residual Plot Assessment: autovi and autovi.web", "The R package autovi on CRAN", "80% Concept, Analysis, Software, Writing", "D.
 Cook 10%,  \n E.
@@
 
-Chapter 3, *Automated Assessment of Residual Plots with Computer Vision Models* is to be submitted for the Journal of Computational and Graphical Statistics.
+Chapter 3, *Automated Assessment of Residual Plots with Computer Vision Models*, and Chapter 4, *Software for Automated Residual Plot Assessment: autovi and autovi.web*, are planned for submission to peer-reviewed journals in the near future.
 

### 2024-08-26 Thesis/abstract.qmd
fill in acknowledgement

@@
 Tanaka 5%, \n S.
-VanderPlas 5%", "No"
-  # 3, "Automated Assessment of Residual Plots with Computer Vision Models", "Unsubmitted", "80% Concept, Analysis, Software, Writing", "D.
+VanderPlas 5%", "No",
+  3, "Automated Assessment of Residual Plots with Computer Vision Models", "To be submitted", "80% Concept, Analysis, Software, Writing", "D.
 Cook 10%,  \n P.
@@
 Langrené 2.5%", "No",
-  # 4, "Software for Automated Residual Plot Assessment: autovi and autovi.web", "The R package autovi on CRAN", "80% Concept, Analysis, Software, Writing", "D.
+  4, "Software for Automated Residual Plot Assessment: autovi and autovi.web", "To be submitted", "80% Concept, Analysis, Software, Writing", "D.
 Cook 10%,  \n E.
@@
 
-Chapter 3, *Automated Assessment of Residual Plots with Computer Vision Models*, and Chapter 4, *Software for Automated Residual Plot Assessment: autovi and autovi.web*, are planned for submission to peer-reviewed journals in the near future.
-
 <!-- **The thesis is written in Australian spelling, except for Chapters 3 and 4, which use American spelling as specified by the publication venue.** -->

### 2024-08-26 Thesis/03-chap3.qmd
final

@@
 
-The utilization of computers to interpret data plots has a rich history, with early efforts such as "Scagnostics" by @tukey1985computer, a set of numerical statistics that summarise features of scatter plots.
+The utilization of computers to interpret data plots has a rich history, with early efforts such as "Scagnostics" by @tukey1985computer, a set of numerical statistics that summarize features of scatter plots.
 @wilkinson2005graph expanded on this work, introducing scagnostics based on computable measures applied to planar proximity graphs.
@@
 
-```{r fig-false-finding, fig.width = 4, fig.height = 4*4/5, fig.pos="!h", fig.cap="An example residual vs fitted values plot (red line indicates 0 corresponds to the x-intercept, i.e.
+```{r fig-false-finding, fig.pos="!h", fig.width = 4, fig.height = 4*4/5, fig.pos="!h", fig.cap="An example residual vs fitted values plot (red line indicates 0 corresponds to the x-intercept, i.e.
 $y=0$).
@@
 
-In @sec-model-introduction, we mentioned that scagnostics consist of a set of manually designed visual feature extraction functions.
+In @sec-model-introduction, we mention that scagnostics consist of a set of manually designed visual feature extraction functions.
 While our computer vision model will learn its own feature extraction function during training, leveraging additional information from scagnostics can enhance the model's predictive accuracy.
-
-For each residual plot used as an input image, we computed four scagnostics – "Monotonic", "Sparse", "Splines", and "Striped" – using the `cassowaryr` R package [@mason2022cassowaryr].
-These computed measures, along with the number of observations from the fitted model, were provided as the second input for the computer vision model.
-While other scagnostics provide valuable insights, they come with high computational costs and are not suitable for quick inference.
 
@@
 It is not expected that $\hat{D}$ will be equal to original distance $D$.
-This is largely because information contained in a single residual plot is limited and it may not be able to summarise all the important characteristics of the residual distribution.
+This is largely because information contained in a single residual plot is limited and it may not be able to summarize all the important characteristics of the residual distribution.
 For a given residual distribution $P$, many different residual plots can be simulated, where many will share similar visual patterns, but some of them could be visually very different from the rest, especially for regression models with small $n$.
@@
 We would not expect the model to be able to magically tell if the suspicious pattern is caused by the skewed distribution of the fitted values or the existence of heteroskedasticity.
-Additionally, some null plots could have outliers or strong visual patterns due to randomness, and a reasonable model will try to summarise those information into the prediction, resulting in $\hat{D} > 0$.
+Additionally, some null plots could have outliers or strong visual patterns due to randomness, and a reasonable model will try to summarize those information into the prediction, resulting in $\hat{D} > 0$.
 
@@
 For a given data generating process, $D$ typically increases logarithmically with the number of observations.
-This behaviour comes from the relationship $D = \text{log}(1 + D_{KL})$, where $D_{KL} = \sum_{i=1}^{n}D_{KL}^{(i)}$ under the assumption of independence.
+This behavior comes from the relationship $D = \text{log}(1 + D_{KL})$, where $D_{KL} = \sum_{i=1}^{n}D_{KL}^{(i)}$ under the assumption of independence.
 
@@
                     col.names = c("Degree of model violations", "Range ($C$ = 10)"),
-                    align = "lc") %>%
-  kableExtra::row_spec(0, align = "c")
+                    align = "lc")
 ```
@@
 The model violations index (MVI) is displayed atop each residual plot.
-The non-linearity patterns are relatively strong for $MVI > 8$, and relatively weak for $MVI < 6$, while the heteroskedasticity patterns are relatively strong for $MVI > 8$, and relatively weak for $MVI < 6$.", fig.height = 12}
+The non-linearity patterns are relatively strong for $MVI > 8$, and relatively weak for $MVI < 6$, while the heteroskedasticity patterns are relatively strong for $MVI > 8$, and relatively weak for $MVI < 6$.", fig.height = 10}
 patchwork::wrap_plots(p1, p2, ncol = 1)
@@
 We have incorporated three types of residual departures of linear regression model in the training data, including non-linearity, heteroskedasticity and non-normality.
-All three departures can be summarised by the data generating process formulated as
+All three departures can be summarized by the data generating process formulated as
 
@@
   kableExtra::kable(escape = FALSE,
-                    booktabs = TRUE) %>%
-  kableExtra::row_spec(0, align = "c")
+                    booktabs = TRUE)
 ```
@@
 Numbers at the bottom of each box show the shape of the output of each layer.
-The band of each box drawn in a darker colour indicates the use of the rectified linear unit activation function.
+The band of each box drawn in a darker color indicates the use of the rectified linear unit activation function.
 Yellow boxes are 2D convolutional layers, orange boxes are pooling layers, the grey box is the concatenation layer, and the purple boxes are dense layers."
@@
 This is followed by a grayscale conversion layer utilizing the luma formula under the Rec.
-601 standard [@series2011studio], which converts the colour image to grayscale.
+601 standard [@series2011studio], which converts the color image to grayscale.
 Grayscale suffices for our task since data points are plotted in black.
@@
 
-::: {#tbl-hyperparameter .cell tbl-cap='Name of hyperparameters and their correspoding domain for the computer vision model.'}
+::: {#tbl-hyperparameter .cell tbl-cap='Name of hyperparameters and their corresponding domain for the computer vision model.'}
 ::: {.cell-output-display}
@@
            Domain = c("{4, 8, 16, 32, 64}", "[0.1, 0.6]", "{false, true}", "{max, average}", "{false, true}", "{128, 256, 512, 1024, 2048}", "{false, true}", "[0.1, 0.6]", "[$10^{-8}$, $10^{-1}$]")) %>%
-  kableExtra::kable(booktabs = TRUE, escape = FALSE) %>%
-  kableExtra::row_spec(0, align = "c")
+  kableExtra::kable(booktabs = TRUE, escape = FALSE)
 ```
@@
                     escape = FALSE) %>%
-  kableExtra::kable_styling(latex_options = c("scale_down")) %>%
-  kableExtra::row_spec(0, align = "c")
+  kableExtra::kable_styling(latex_options = c("scale_down"))
 ```
@@
                     align = "lrrrr",
-                    col.names = c("", "RMSE", "$R^2$", "MAE", "Huber loss")) %>%
-  kableExtra::row_spec(0, align = "c")
+                    col.names = c("", "RMSE", "$R^2$", "MAE", "Huber loss"))
 ```
@@
 ```{r fig-model-performance, fig.pos="!h", fig.cap = "Hexagonal heatmap for difference in $D$ and $\\hat{D}$ vs $D$ on test data for three optimized models with different input sizes.
-The brown lines are smoothing curves produced by fitting gnealized additive models.
+The brown lines are smoothing curves produced by fitting generalized additive models.
 The area over the zero line in light yellow indicates under-prediction, and the area under the zero line in light green indicates over-prediction.", dev = 'png', dpi = 300}
@@
 The data is grouped by whether the regression has only non-linearity violation, and whether it includes a second predictor in the regression formula.
-The brown lines are smoothing curves produced by fitting gnealized additive models.
+The brown lines are smoothing curves produced by fitting generalized additive models.
 The area over the zero line in light yellow indicates under-prediction, and the area under the zero line in light green indicates over-prediction.", dev = 'png', dpi = 300}
@@
                     align = "lrr",
-                    col.names = c("Violations", "\\#samples", "RMSE")) %>%
-  kableExtra::kable_styling(latex_options = c("scale_down")) %>%
-  kableExtra::row_spec(0, align = "c")
+                    col.names = c("Violations", "\\#samples", "RMSE"))
 ```
@@
                     align = "lrrrr",
-                    col.names = c("Violation", "RMSE", "$R^2$", "MAE", "Huber loss")) %>%
-  kableExtra::row_spec(0, align = "c")
+                    col.names = c("Violation", "RMSE", "$R^2$", "MAE", "Huber loss"))
 ```
@@
 The performance metrics of $\hat{D}$ for true residual plots are outlined in @tbl-experiment-performance.
-It's notable that all performance metrics are slightly worse than those evaluated on the test data.
+It is notable that all performance metrics are slightly worse than those evaluated on the test data.
 Nevertheless, the mean absolute error remains at a low level, and the linear correlation between the prediction and the true value remains very high.
@@
 The agreement rates between conventional tests and the computer vision model are 85.95% and 79.69% for residual plots containing heteroskedasticity and non-linearity patterns, respectively.
-These figures are higher than those calculated for visual tests conducted by human, indicating that the computer vision model exhibits behaviour more akin to the best available conventional tests.
+These figures are higher than those calculated for visual tests conducted by human, indicating that the computer vision model exhibits behavior more akin to the best available conventional tests.
 However, @fig-conv-mosaic shows that the computer vision model does not always reject when the conventional tests reject.
@@
 In @fig-power, rejection decisions are fitted by logistic regression models with no intercept terms and an offset equals to $\text{log}(0.05/0.95)$.
-The fitted curves for the computer vision model fall between those of conventional tests and visual tests for both non-linearity and heteroskedasticity, which means there is still potential to refine the computer vision model to better align its behaviour with visual tests conducted by humans.
+The fitted curves for the computer vision model fall between those of conventional tests and visual tests for both non-linearity and heteroskedasticity, which means there is still potential to refine the computer vision model to better align its behavior with visual tests conducted by humans.
 
@@
 In some instances, the weighted detection rate is close to one, yet the adjusted $\delta$-difference is negative.
-This discrepancy implies that the distance measure, or the estimated distance, may not perfectly reflect actual human behaviour.
+This discrepancy implies that the distance measure, or the estimated distance, may not perfectly reflect actual human behavior.
 
@@
   kableExtra::pack_rows("Compared with conventional tests", 1, 2) %>%
-  kableExtra::pack_rows("Compared with visual tests conducted by human", 3, 4) %>%
-  kableExtra::row_spec(0, align = "c")
+  kableExtra::pack_rows("Compared with visual tests conducted by human", 3, 4)
 ```
@@
 ```{r fig-conv-mosaic, fig.pos = "!h", fig.cap = "Rejection rate ($p$-value $\\leq0.05$) of computer vision models conditional on conventional tests on non-linearity (left) and heteroskedasticity (right) lineups displayed using a mosaic plot.
-When the conventional test fails to reject, the computer vision mostly fails to reject the same plot as well as indicated by the height of the top right yellow rectangle, but there are non negliable amount of plots where the conventoinal test rejects but the computer vision model fails to reject as indicated by the width of the top left yellow rectangle."}
+When the conventional test fails to reject, the computer vision mostly fails to reject the same plot as well as indicated by the height of the top right yellow rectangle, but there are non negliable amount of plots where the conventional test rejects but the computer vision model fails to reject as indicated by the width of the top left yellow rectangle."}
 library(ggmosaic)
@@
 ```{r fig-delta, fig.pos = "!h", fig.cap = "A weighted detection rate vs adjusted $\\delta$-difference plot.
-The brown line is smoothing curve produced by fitting gnealized additive models.", fig.height = 4}
+The brown line is smoothing curve produced by fitting generalized additive models.", fig.height = 4}
 experiment %>%
@@
 To see if these features can be effectively used for distinguishing null plots and true residual plot, we linearly project them into the first and second principal components space as shown in [@fig-false-check]D.
-It can be observed that because the bootstrapped plots are mostly similar to the null plots, the points drawn in different colours are mixed together.
+It can be observed that because the bootstrapped plots are mostly similar to the null plots, the points drawn in different colors are mixed together.
 The true residual plot is also covered by both the cluster of null plots and cluster of bootstrapped plots.
@@
 
-```{r fig-dino-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment for the datasauRus fitted model evaluted on 200 null plots and 200 bootstrapped plots.
+```{r fig-dino-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment for the datasauRus fitted model evaluated on 200 null plots and 200 bootstrapped plots.
 (A) The residual plot exhibits a "dinosaur" shape.
@@
 The true residual plot is at position 17.
-It can be easily identified as the most different plot as the visual pattern is extremly artificial.', fig.height = 8}
+It can be easily identified as the most different plot as the visual pattern is extremely artificial.', fig.height = 8}
 pos <- 17

### 2024-08-26 Thesis/04-chap4.qmd
final

@@
 
-Regression analysis is a fundamental statistical technique widely used for modelling data from many fields.
+Regression analysis is a fundamental statistical technique widely used for modeling data from many fields.
 To diagnose the fit of a model it is recommended that the residuals are plotted.
@@
 However, generic functions only work with classes that have appropriate method implementations.
-Some regression modelling packages may not fully adhere to the `stats` package guidelines for implementing these functions.
+Some regression modeling packages may not fully adhere to the `stats` package guidelines for implementing these functions.
 In such cases, overriding the method becomes necessary.

### 2024-08-26 Thesis/05-chap5.qmd
final

@@
 
-Principles of transparency and reproducible research have guided the work with all source code available in public and private repositories at <https://github.com/TengMCing>.
+Principles of transparency and reproducible research have guided the work with all materials related to the thesis at <https://github.com/TengMCing/PhD>.
 The thesis is written using Quarto [@Allaire_Quarto_2024] and is available online at <https://patrick-li-thesis.netlify.app>.

### 2024-08-26 Thesis/A-appA.qmd
final

@@
 The analysis is conducted by setting up several scenarios, where the $\alpha$ is under or overestimated by 12.5%, 25% and 50%.
-Using the adjusted $\hat{\alpha}$, we recalculate the $p$-value for every lineup and show the results in Figure \ref{fig:sensitivity}.
+Using the adjusted $\hat{\alpha}$, we recalculate the $p$-value for every lineup and show the results in @fig-sensitivity.
 It can be observed that there are some changes to $p$-values, especially when the $\hat{\alpha}$ is multiplied by 50%.
-However, Table \ref{tab:sensitivity} shows that adjusting $\hat{\alpha}$ will not result in a huge difference in rejection decisions.
+However, @tbl-sensitivity shows that adjusting $\hat{\alpha}$ will not result in a huge difference in rejection decisions.
 There are only a small percentage of cases where the rejection decision change.
@@
 ```{r}
+#| tbl-cap: "Examining how decisions might change if $\\hat{\\alpha}$ was different.
+Percentage of lineups that where the $p$-value would switch to above or below 0.05, when $\\hat{\\alpha}$ is multiplied by a multiplier."
+#| label: tbl-sensitivity
 upper_lower_p %>%
@@
                                   "%"),
-                    label = "sensitivity", 
-                    caption = "Examining how decisions might change if $\\hat{\\alpha}$ was different.
-Percentage of lineups that where the $p$-value would switch to above or below 0.05, when $\\hat{\\alpha}$ is multiplied by a multiplier.",
                     linesep = c("", "", "\\addlinespace"))
@@
 This would be a state where both systems agree perfectly.
-Figure \ref{fig:rej-rates} examines this.
+@fig-rej-rates examines this.
 Plot A shows the percentage of residual plots rejected by the visual test, given the conventional test rejected (solid lines) or failed to reject (dashed lines).

### 2024-08-26 Thesis/abstract.qmd
final

@@
 The inclusion of co-authors reflects the fact that the work came from active collaboration between researchers and acknowledges input into team-based research.
-In the case of [Chapter 2](#sec-first-paper), my contribution to the work involved the following:
+In the case of [Chapter 2](#sec-first-paper), [Chapter 3](#sec-second-paper) and [Chapter 4](#sec-third-paper), my contribution to the work involved the following:
 
@@
   3, "Automated Assessment of Residual Plots with Computer Vision Models", "To be submitted", "80% Concept, Analysis, Software, Writing", "D.
-Cook 10%,  \n P.
-Menéndez 5%, \n U.
-Laa 2.5%,  \n N.
-Langrené 2.5%", "No",
+Cook 10%,  \n E.
+Tanaka 5%, \n S.
+VanderPlas 2.5%, K.
+Ackermann 2.5%", "No",
   4, "Software for Automated Residual Plot Assessment: autovi and autovi.web", "To be submitted", "80% Concept, Analysis, Software, Writing", "D.
@@
 
+\clearpage
+
 <!-- **The thesis is written in Australian spelling, except for Chapters 3 and 4, which use American spelling as specified by the publication venue.** -->

### 2024-09-03 Thesis/abstract.qmd
modify the front matter
update the declaration

--- 2024-08-26 Thesis/abstract.qmd
@@
 
-This thesis includes one original paper published in a peer reviewed journal and two unpublished papers.
+This thesis includes one original papers published in a peer reviewed journal and two unpublished papers.
+The core theme of the thesis is "automated reading of residual plots".
+The ideas, development and writing up of all the papers in the thesis were the principal responsibility of myself, the student, working within the Department of Econometrics and Business Statistics under the supervision of Professor Dianne Cook, Dr Emi Tanaka (Australian National University), Assistant Professor Susan VanderPlas (University of Nebraska–Lincoln) and Assistant Professor Klaus Ackermann.
 
@@
 Tanaka 5%, \n S.
-VanderPlas 5%", "No",
-  3, "Automated Assessment of Residual Plots with Computer Vision Models", "To be submitted", "80% Concept, Analysis, Software, Writing", "D.
-Cook 10%,  \n P.
-Menéndez 5%, \n U.
-Laa 2.5%,  \n N.
-Langrené 2.5%", "No",
-  4, "Software for Automated Residual Plot Assessment: autovi and autovi.web", "To be submitted", "80% Concept, Analysis, Software, Writing", "D.
+VanderPlas 5%", "No"
+  # 3, "Automated Assessment of Residual Plots with Computer Vision Models", "To be submitted", "80% Concept, Analysis, Software, Writing", "D.
+Cook 10%,  \n E.
+Tanaka 5%, \n S.
+VanderPlas 2.5%, K.
+Ackermann 2.5%", "No",
+  # 4, "Software for Automated Residual Plot Assessment: autovi and autovi.web", "To be submitted", "80% Concept, Analysis, Software, Writing", "D.
 Cook 10%,  \n E.
@@
 
+Chapters 3, *Automated Assessment of Residual Plots with Computer Vision Models*, and Chapter 4, *Software for Automated Residual Plot Assessment: autovi and autovi.web*, are planned for submission to peer-reviewed journals.
+
+\clearpage
+
+To ensure the clarity and coherence of the written content, artificial intelligence tools were employed to assist in smoothing and refining the language throughout the thesis.
+
 <!-- **The thesis is written in Australian spelling, except for Chapters 3 and 4, which use American spelling as specified by the publication venue.** -->

### 2024-09-04 Thesis/05-chap5.qmd
add citation for packages

@@
 
-Principles of transparency and reproducible research have guided the work with all materials related to the thesis at <https://github.com/TengMCing/PhD>.
+Principles of transparency and reproducible research have guided the work with all materials related to the thesis at <https://github.com/TengMCing/PhD/Thesis>.
 The thesis is written using Quarto [@Allaire_Quarto_2024] and is available online at <https://patrick-li-thesis.netlify.app>.
+The R packages used throughout the thesis include `tidyverse` [@tidyverse], `lmtest` [@lmtest], `mpoly` [@mpoly], `ggmosaic` [@ggmosaic], `kableExtra` [@kableextra], `patchwork` [@patchwork], `rcartocolor` [@rcartocolor], `glue` [@glue], `ggpcp` [@ggpcp], `here` [@here], `magick` [@magick], `yardstick` [@yardstick], `reticulate` [@reticulate] and `knitr` [@knitr].
 

### 2024-09-04 Thesis/03-chap3.qmd
fix some small issues

@@
 
-## Statistical testing {#sec-model-statistical-testing}
+## Statistical Testing {#sec-model-statistical-testing}
 
@@
 data.frame(`Hyperparameter` = c("Number of base filters", "Dropout rate for convolutional blocks", "Batch normalization for convolutional blocks", "Type of global pooling", "Ignore additional inputs", "Number of units for the fully-connected layer", "Batch normalization for the fully-connected layer", "Dropout rate for the fully-connected layer", "Learning rate"),
-           Domain = c("{4, 8, 16, 32, 64}", "[0.1, 0.6]", "{false, true}", "{max, average}", "{false, true}", "{128, 256, 512, 1024, 2048}", "{false, true}", "[0.1, 0.6]", "[$10^{-8}$, $10^{-1}$]")) %>%
+           Domain = c("\\{4, 8, 16, 32, 64\\}", "[0.1, 0.6]", "\\{false, true\\}", "\\{max, average\\}", "\\{false, true\\}", "\\{128, 256, 512, 1024, 2048\\}", "\\{false, true\\}", "[0.1, 0.6]", "[$10^{-8}$, $10^{-1}$]")) %>%
   kableExtra::kable(booktabs = TRUE, escape = FALSE)
@@
 
-#### Overview of the Human Subject experiment
+#### Overview of the Human Subject Experiment
 
@@
 
-## Conclusion
+## Conclusions
 

### 2024-09-04 Thesis/04-chap4.qmd
fix some small issues

@@
 
-```{r echo = TRUE, message = TRUE}
-checker
+```r
+── <AUTO_VI object>
+Status:
+ - Fitted model: lm
+ - Keras model: (None, 32, 32, 3) + (None, 5) -> (None, 1)
+    - Output node index: 1
+ - Result: UNKNOWN 
 ```
@@
 However, server hosting domains that natively support Python servers typically do not have the latest version of R installed.
-Additionally, calling R from Python is typically done using the `rpy2` Python library, but this approach can be awkward when dealing with language syntax related to non-standard evaluation.
+Additionally, calling R from Python is typically done using the `rpy2` Python library [@rpy2], but this approach can be awkward when dealing with language syntax related to non-standard evaluation.
 Another option we considered was renting a server where we could have full control, such as those provided by cloud platforms like Google Cloud Platform (GCP) or Amazon Web Services (AWS).

### 2024-11-28 Thesis/01-chap1.qmd
minor changes

@@
 In the context of regression diagnostics, a common practice is to plot residuals against fitted values, which serves as a starting point for evaluating the adequacy of the fit and verifying the underlying assumptions.
-Visual diagnostics are frequently preferred or recommended [@draper1998applied; @cook1982residuals; @montgomery1982introduction] due to the possibility of discovering abstract and unquantifiable insights, however, it can be subject to overinterpretation or even neglect.
+Visual diagnostics are frequently preferred or recommended [@draper1998applied; @cook1982residuals; @montgomery1982introduction] due to the possibility of discovering abstract and unquantifiable insights, however, it can be subject to over-interpretation or even neglect.
 
@@
 If a large number of participants correctly identify the data plot, the corresponding $p$-value will be small, indicating strong evidence against $H_0$.
-This protocol provides a callibration of the data plot against the null plots, ensuring that the data plot is not overinterpreted.
+This protocol provides a calibration of the data plot against the null plots, ensuring that the data plot is not over-interpreted.
 
-The lineup protocol has gained increasing traction in recent years and has already been integrated into data analysis of various topics [see @krishnan2021hierarchical; @loy2013diagnostic; @widen2016graphical].
+The lineup protocol has gained increasing traction in recent years and has already been integrated into data analysis of various topics [see @krishnan2021hierarchical; @loy2013diagnostic; @widen2016graphical; @savvides2022].
 However, the reliance of human assessment is a fundamental aspect of visual tests, which may restrict its widespread usage.
@@
 To address these limitations, this thesis proposes a computer vision-based approach to automate the visual inference process for assessment of linear regression residual plots.
-Modern computer vision models often use a convolutional neural network architecture to process digital images to peform various tasks (e.g.
+Modern computer vision models often use a convolutional neural network to process digital images to perform various tasks (e.g.
 object detection, object identification and signal processing).
@@
 classification of time series images in @hatami2018] but generally limited in development.
-The development of computer vision models for the assessment residual plots will make the process more efficient, consistent, and accessible.
+The development of computer vision models for the assessment of residual plots will make the process more efficient, consistent, and accessible.
+
+Regression diagnostics is a well-established field with extensive literature addressing various types of regression models, such as generalized linear models [@hardin2007generalized], mixed-effects models [@pinheiro2006mixed], panel data models [@baltagi2008econometric], and survival models [@baltagi2008econometric].
+This thesis, however, focuses on the classical normal linear regression model.
+Further discussions on extending the methods established in this thesis to other types of regression models are provided in @sec-second-paper-limitation.
 
@@
 
-@sec-conclusion summarises the contribution of the work and their impact, and discusses some future plans.
+@sec-conclusion summarises the contribution of the work and the (potential) impact, and discusses some future plans.

### 2024-11-28 Thesis/03-chap3.qmd
minor changes

@@
 
-## Model Specifications {#sec-model-specifications}
+## Computer Vision Model Specifications {#sec-model-specifications}
 
@@
 A straightforward architecture of the input layer involves feeding a vector of residuals along with a vector of fitted values, essentially providing all the necessary information for creating a residuals vs fitted values plot.
-However, a drawback of this method is the dynamic input size, which changes based on the number of observations.
+However, a drawback of this method is the variable input size, which changes based on the number of observations.
 For modern computer vision models implemented in mainstream software like TensorFlow [@abadi2016tensorflow], the input shape is typically fixed.
@@
 @fig-different-j-heter and @fig-different-e-heter show the visual patterns of models with multiple model violations.
+
+The predictors, $\boldsymbol{x}_1$ and $\boldsymbol{x}_2$, are randomly generated from four distinct distributions, including $U(-1, 1)$ (uniform), $N(0, 0.3^2)$ (normal), $\text{lognormal}(0, 0.6^2)/3$ (skewed) and $U\{-1, 1\}$ (discrete uniform).
 
@@
 The attention map in plot B suggests the center of the image has higher leverage than other areas, and it is the turning point of the "U" shape.
-The CPA provided in plot D shows two distinct clusters of data points, further underling the visual differences between bootstrapped plots and null plots.
+The PCA provided in plot D shows two distinct clusters of data points, further underling the visual differences between bootstrapped plots and null plots.
 This coincides the findings from @fig-boston-lineup, where the true plot exhibiting a "U" shape is visually distinctive from null plots.
@@
 
-## Limitations and Future Work
+## Limitations and Future Work {#sec-second-paper-limitation}
 

### 2024-11-28 Thesis/04-chap4.qmd
minor changes

@@
 ```r
-── <AUTO_VI object>
+print(checker)
+```
+
+```{r message = TRUE}
+cli::cli_h3("
+<AUTO_VI object>
 Status:
@@
  - Result: UNKNOWN 
+    ")
 ```
@@
 
-```{r}
+```{r echo = TRUE}
 list_keras_model()

### 2024-11-28 Thesis/amend.qmd
add amend summary

---
title: "Amendments Summary"
author: Weihao Li
Date: "`r Sys.Date()`"
format: pdf
---

In this version, I have added a paragraph to acknowledge the extensive literature on regression diagnostics and various types of regression models. I clarified that the thesis focuses specifically on the classical normal linear regression model.

I also improved the language in Chapter 3 for better clarity, such as changing "dynamic input size" to "variable input size." Additionally, I included a sentence to provide more details on how the predictors in the training datasets are generated.

In Chapter 4, I corrected several formatting issues, particularly those related to code chunks not displaying properly.

Overall, the reviewers' comments have been very helpful, though not all could be addressed due to time constraints. However, as we are preparing the journal submission for Chapters 3 and 4, these comments will be carefully considered and properly addressed.

### 2024-12-25 Thesis/01-chap1.qmd
update the final thesis

@@
 
-@buja2009statistical introduced a visual inference framework that formalised a hypothesis testing of graphical representations of data (henceforth referred to as the _data plot_) via the lineup protocol.
+@buja2009statistical introduced a visual inference framework that formalised a hypothesis testing of graphical representations of data (henceforth referred to as the _data plot_) via the lineup protocol (see @fig-introduction-lineup for an example).
 The protocol is inspired by the police lineup technique employed in eyewitness identification of criminal suspects.
@@
 
-Regression diagnostics is a well-established field with extensive literature addressing various types of regression models, such as generalized linear models [@hardin2007generalized], mixed-effects models [@pinheiro2006mixed], panel data models [@baltagi2008econometric], and survival models [@baltagi2008econometric].
+Regression diagnostics is a well-established field with extensive literature, and a more detailed discussion will be provided in Chapter 2.
+This field also encompasses a variety of regression models, including generalized linear models [@hardin2007generalized], mixed-effects models [@pinheiro2006mixed], panel data models [@baltagi2008econometric], and survival models [@baltagi2008econometric].
 This thesis, however, focuses on the classical normal linear regression model.
 Further discussions on extending the methods established in this thesis to other types of regression models are provided in @sec-second-paper-limitation.
+
+```{r fig-introduction-lineup, fig.pos="t!", fig.height = 7, fig.width = 7, out.width = "100%", fig.cap = "A lineup is used to conduct visual testing, as demonstrated in this example.
+The observed data’s residual plot is placed among 19 null plots generated from a standard error model.
+Human judges are then asked to examine the lineup and identify the plot they find most distinct.
+The $p$-value is calculated based on how often the judges correctly identify the data plot, which is located at position $7$ and exhibits heteroskedasticity.
+A small $p$-value indicates a substantial agreement among the judges in selecting the data plot."}
+set.seed(10086)
+visage::heter_model(a = 0, b = 60)$gen_lineup(300) |>
+  visage::VI_MODEL$plot_lineup(theme = ggplot2::theme_light(), remove_axis = TRUE, remove_grid_line = TRUE)
+```
 

### 2024-12-25 Thesis/abstract.qmd
upload the original copy
update the final thesis

@@
 While numerical hypothesis tests are commonly used, residual plots remain essential in regression diagnostics because conventional tests are limited to specific types of model departures and tend to be overly sensitive.
-Visual inference using the lineup protocol offers a less sensitive and more broadly applicable alternative, yet its dependence on human judgment limits scalability.
-This research addresses these limitations by automating the assessment process.
+This research builds on a method called the lineup protocol, a statistical visualization approach that embeds a data plot among null plots to compare patterns arising by chance with those observed in the data.
+This process is part of visual inference, where human judgment is used to make statistical inferences from data visualizations.
+However, the reliance on human judgment in visual inference poses challenges for scalability and consistency.
 
-This research presents three original contributions.
+To overcome these challenges, this research presents three original contributions.
 The first contribution provides evidence for the effectiveness of visual inference in regression diagnostics through a human subject experiment, demonstrating the benefits of using the lineup protocol for reliable and consistent reading of residual plots.
@@
 
-**Date**: 26th August 2024
+**Date**: 4th September 2024
 
@@
 
-**Date**: 26th August 2024
+**Date**: 4th September 2024

### 2024-12-25 Thesis/amend_v2.qmd
update the final thesis

---
title: "Amendments Summary"
author: Weihao Li
Date: "`r Sys.Date()`"
format: pdf
---

## Responses to Comments from Examiner Prof. Louise Ryan:

1. *While the candidate writes quite clearly, I feel that some improvements in presentation could be helpful. For example, in the abstract, it should be written in such a way that a non-expert could quickly get a sense of what the thesis is about and why it is important. In this case, I feel that the abstract is written in a fairly dry manner and it uses technical terms that are not appropriate without some context and definition. For example, I am a very experienced statistician, but I have never heard of the lineup protocol. Either this term should not be used in the abstract or there needs to be some kind of contextual statement describing what it is. The content of the thesis is actually very exciting. However, the abstract does not manage to communicate that excitement.*

Thank you for your valuable feedback on the presentation and accessibility of the abstract. I appreciate your suggestions to make the abstract more engaging and suitable for a non-expert audience.

I have revised the abstract to ensure that it provides a clear and concise overview of the thesis in a manner accessible to a broader audience, including non-experts. Technical terms such as "lineup protocol" have been either omitted or explained briefly with contextual statements to ensure clarity for readers unfamiliar with these concepts. For instance, I have included a sentence explaining that the lineup protocol is a statistical visualization approach that embeds a data plot among null plots to compare patterns arising by chance with those observed in the data.

2. *In Chapter 1, I can’t help but feel that there should be more cited literature. The field of regression diagnostics is massive so I don’t feel it is quite appropriate to sum it all up with a couple of references from the 80s and 90s! It may be adequate to say something along the lines that there is a vast literature out there and that more will be discussed in Chapter 2.*

To address the concern about summarizing the vast body of work, I added a statement in Chapter 1 acknowledging the extensive literature in regression diagnostics. I also clarified that this is a well-established field and noted that a more detailed discussion of the literature will follow in Chapter 2.

3. *related to the above point, it is probably wise to mention that regression diagnostics have been explored in a very wide range of contexts, including ordinary linear regression, generalised linear models, survival models, repeated measures models etc etc. I think it is correct to say that your thesis will be focussed on diagnostics for ordinary linear regression, is this correct? I think it is important to establish that context.*

Thank you for this suggestion. I have added a statement in Chapter 1 to highlight the wide-ranging applications of regression diagnostics across various models. I also clarified that the primary focus of my thesis is on diagnostics for classical normal linear regression model to establish this context early on.

4. *when describing the lineup protocol at the beginning of Chapter 1, perhaps some diagrams to help visualise how it works? As mentioned above, I have not myself heard of the lineup protocol, but it sounds very interesting and intriguing. Some visualisations might really help and also engage the reader. I know you do this in Chapter 2, but I can’t help but feel it would be nice to see some kind of visualisation in the intro section as well.*

I agree that including a visual representation of the lineup protocol in the introduction would help clarify its concept and engage the reader. I have added a diagram in Chapter 1 to illustrate how the lineup protocol works, providing an example similar to those discussed in Chapter 2. This addition should help familiarize readers with the protocol from the outset.

5. *In chapter 2, can you say a bit more about how you recruited your subjects? I am not familiar with the Prolific Crowd Sourcing Platform. Were the subjects technically trained at all in terms of knowing what to look for?*

The recruitment details are provided in the appendix. The subjects received training by reading a webpage with examples. Unfortunately, I am unable to modify this chapter further, as it has already been submitted as a paper and accepted, requiring it to remain unchanged.

6. *I was glad to see a reference to Tukey in Chapter 3! He was absolutely one of the pioneers in terms of using visualisations to help interpret model fit etc. You might want to add some additional references earlier on as well. Some of his work was foundational in the field, in my opinion.*

I completely agree that Tukey’s contributions were foundational and greatly impactful. While I appreciate the suggestion to add more references to his work earlier in the thesis, I did not find a suitable place where they would naturally fit within the context. However, I’m glad the reference in Chapter 3 resonated with you.

7. *There are a few places where technical terms are used but not defined. e.g page 30, what is "dynamic input size"?*

I appreciate your careful observation. I’ve addressed the issue by improving the language for better clarity in Chapter 3, such as changing "dynamic input size" to "variable input size" and providing a clear definition on page 30. I appreciate your attention to detail and constructive feedback!

8. *Perhaps it is there and I have not read carefully enough to understand, but in Chaper 2, are you assuming that the actual plot and all the null plots for the lineup protocol have been generated prior to being input into the computer vision model? If this is the case, then I agree it seems enough to enter residuals and fitted values. The only thing I would say is that sometimes there is interest in other sorts of plots, eg plots of residuals versus the values of some specific predictor variables.*

Yes, the actual plot and all the null plots for the lineup protocol are generated prior to being input into the computer vision model. The chapter primarily focuses on residuals vs. fitted values plots. Other types of plots, such as residuals vs. specific predictor variables, are discussed in Section 3.12, which covers limitations and future work.

9. *related to the previous point, I found the term "model specification" (section 3.2) a bit confusing because I was thinking at first of the underlying model that you are trying to critique. Just be careful here to be clear since "model" is being used in a number of different ways.*

Thank you for pointing this out. To avoid confusion, I have clarified the terminology by changing "model specification" in Section 3.2 to "Computer Vision Model Specifications." This should make it clearer that the term refers specifically to the details of the computer vision model and not the underlying statistical model being critiqued. I appreciate your feedback!

10. *page 40, I don’t see here where you talk about how you generate the values of x1 and x2?*

I appreciate you bringing this to my attention. I have added a paragraph in Section 3.7 to provide more details on how the predictors $x_1$ and $x_2$ in the training datasets are generated. I appreciate your careful review and feedback!

11. *Figure 3.1 and the point made about it seems quite important. But the figure shows up a long way from where it is cited in the text*

I’m grateful for your observation. The placement of Figure 3.1 was indeed a formatting issue in the original document. This has since been fixed, and the figure now appears much closer to its citation in the text. I appreciate your attention to this detail!

## Responses to Comments from Examiner Prof. Daniela Witten:

1. *In Chapter 3, why not train a deep learning model to predict "contains outliers" versus "does not contain outliers", or "is non-linear" versus "is not non-linear", etc., rather than estimating the distance between the residual distribution and a theoretically "good" residual distribution? Wouldn't this obviate the need for the lineup evaluation in Section 3.5?*

Thank you for the insightful question. This is something I considered carefully and discussed extensively with my supervisors during the development of the methodology. While training a deep learning model to predict categories like "contains outliers" vs. "does not contain outliers" or "is non-linear" vs. "not non-linear" seems appealing, it would bring us back to the traditional testing framework. These tests typically focus on detecting specific types of model violations in isolation.

One of the key advantages of using the lineup protocol for residual diagnostics is its ability to detect multiple model violations simultaneously, provided those violations produce noticeable differences in the data plots. In contrast, training a model discussed above would require us to explicitly define and map out all possible violations of interest—or at least the common ones we want to address.

Additionally, if the ultimate goal is to distinguish between specific cases (e.g., outliers vs. no outliers, or linear vs. non-linear relationships), conventional statistical tests based on regression model assumptions are often more suitable. Computer vision models may not be the best tools for these tasks. Even if we trained a deep learning model for this purpose, it would still face challenges in terms of confidence assessment. Such a model would typically produce a single probability prediction for each predefined violation, limiting its scope and flexibility.

The lineup evaluation described in Section 3.5 remains valuable even in this context. By treating the model output as a statistic and comparing it to outputs generated under null assumptions, it adheres to the principles of visual inference and conventional statistical testing. This approach provides a robust framework for interpreting results and avoids the pitfalls of relying solely on predicted probabilities.

2. *In Section 3.5.1, it is mentioned that at least 100 null plots are needed to get a stable estimate of the 95th quantile. How many null plots are needed for this in Chapter 2?*

In Chapter 2, the data plot is compared against 19 null plots because humans are not well-suited to evaluating lineups with a large number of plots. To ensure reliable results, such a lineup should ideally be assessed by at least five subjects; otherwise, the smallest possible $p$-value of the test may not be sufficiently small. Moreover, involving multiple subjects helps reduce variability in individuals' abilities to evaluate residual plots.

In contrast, computer vision models can evaluate lineups with as many null plots as desired, which is one of their advantages.

3. *In Section 3, the decision to treat this as a computer vision problem (rather than just passing in the vectors required to make a residual plot) was justified by saying that if the sample size is very very large, then it might not be possible to pass in the vectors that form the residual plot. I am a bit skeptical of this argument, since the case of an extremely large sample size, I think that the residuals could be sampled without replacement, with no loss of accuracy or ability to detect model violations. (Furthermore, I believe that ordinary least squares regression settings with such extraordinarily large sample sizes are relatively unusual --- in contemporary settings, if the sample size is huge, then typically more complex models than ordinary least squares are fit.)*

The primary issue with providing pairs of vectors to the model is not the large sample size itself but the variability in sample size. Mainstream deep learning frameworks typically cannot handle inputs with variable dimensions. However, as you suggested, sampling pairs of fitted values and residuals without replacement could address this issue without sacrificing accuracy or the ability to detect model violations. This approach offers a way to define fixed-size inputs for the deep learning model.

Thank you for this valuable suggestion. I will include a discussion on this point in the manuscript currently being prepared for the peer-reviewed journal.

4. *How about a hybrid approach where someone visually screens a residual plot to identify possible violations (perhaps using the lineup method), and then conducts a formal statistical test only if possible violations are identified? The motivation here is that if a residual plot does not visually appear to have a violation, then it is not worth pursuing further. However, if a possible violation is visually identified, then a formal test may be worthwhile. This falls into the selective inference framework, since we are "double dipping" the data (once with a visual test, and then again with the formal test). Is it possible to conduct valid inference in this setting?*

This is indeed a valid inference setting and aligns with common practices in applications where findings from exploratory data analysis (EDA) are subsequently verified using statistical testing. The use of visual methods, such as the lineup protocol, during EDA can help identify potential violations before formal testing.

However, this approach introduces the risk of multiple testing problems, as the data is effectively "double-dipped"—first for hypothesis generation and then for formal testing. Addressing this issue requires careful consideration of selective inference frameworks. Potential solutions include splitting the data into separate subsets for EDA and hypothesis testing or using techniques like cross-validation to ensure robust inference.

This hybrid approach represents an interesting direction for future research, particularly in developing methods to make valid inferences while leveraging the strengths of both visual and formal analysis techniques.

### 2025-01-02 Thesis/amend_v2.qmd
amendment summary

@@
 
-## Responses to Comments from Examiner Prof. Louise Ryan:
+## Changes made in response to comments from Examiner Prof. Louise Ryan
 
@@
 
-Thank you for your valuable feedback on the presentation and accessibility of the abstract. I appreciate your suggestions to make the abstract more engaging and suitable for a non-expert audience.
+<!--Thank you for your valuable feedback on the presentation and accessibility of the abstract.
+I appreciate your suggestions to make the abstract more engaging and suitable for a non-expert audience.-->
 
 I have revised the abstract to ensure that it provides a clear and concise overview of the thesis in a manner accessible to a broader audience, including non-experts.
-Technical terms such as "lineup protocol" have been either omitted or explained briefly with contextual statements to ensure clarity for readers unfamiliar with these concepts.
-For instance, I have included a sentence explaining that the lineup protocol is a statistical visualization approach that embeds a data plot among null plots to compare patterns arising by chance with those observed in the data.
+Technical terms have been either omitted or explained briefly with contextual statements to ensure clarity for readers unfamiliar with these concepts.
+For example, I have included a sentence explaining that the lineup protocol is a statistical visualisation approach that embeds a data plot among null plots to compare patterns arising by chance with those observed in the data.
 
@@
 
-Thank you for this suggestion.
 I have added a statement in Chapter 1 to highlight the wide-ranging applications of regression diagnostics across various models.
@@
 
-I agree that including a visual representation of the lineup protocol in the introduction would help clarify its concept and engage the reader.
 I have added a diagram in Chapter 1 to illustrate how the lineup protocol works, providing an example similar to those discussed in Chapter 2.
-This addition should help familiarize readers with the protocol from the outset.
+This addition helps clarify its concept and engage the reader with the protocol from the outset.
 
@@
 The subjects received training by reading a webpage with examples.
-Unfortunately, I am unable to modify this chapter further, as it has already been submitted as a paper and accepted, requiring it to remain unchanged.
+I decided not to modify this chapter further, as it has been submitted published.
 
@@
 I completely agree that Tukey’s contributions were foundational and greatly impactful.
-While I appreciate the suggestion to add more references to his work earlier in the thesis, I did not find a suitable place where they would naturally fit within the context.
-However, I’m glad the reference in Chapter 3 resonated with you.
+While I appreciate the suggestion to add more references to his work earlier in the thesis, this one is sufficient to point to that body of work, and accommodate references to the range of foundational research on regression analysis.
 
@@
 
-I appreciate your careful observation.
-I’ve addressed the issue by improving the language for better clarity in Chapter 3, such as changing "dynamic input size" to "variable input size" and providing a clear definition on page 30.
-I appreciate your attention to detail and constructive feedback!
+I’ve addressed the issue by generally improving the language for better clarity in Chapter 3, including changing "dynamic input size" to "variable input size" and providing a clear definition on page 30.
 
@@
 specific predictor variables, are discussed in Section 3.12, which covers limitations and future work.
+No changes were made.
 
@@
 
-Thank you for pointing this out.
-To avoid confusion, I have clarified the terminology by changing "model specification" in Section 3.2 to "Computer Vision Model Specifications." This should make it clearer that the term refers specifically to the details of the computer vision model and not the underlying statistical model being critiqued.
-I appreciate your feedback!
+To avoid confusion, I have clarified the terminology by changing "model specification" in Section 3.2 to "computer vision model specification" This should make it clearer that the term refers specifically to the details of the computer vision model and not the underlying statistical model being critiqued.
 
@@
 
-I appreciate you bringing this to my attention.
 I have added a paragraph in Section 3.7 to provide more details on how the predictors $x_1$ and $x_2$ in the training datasets are generated.
-I appreciate your careful review and feedback!
 
@@
 
-I’m grateful for your observation.
-The placement of Figure 3.1 was indeed a formatting issue in the original document.
-This has since been fixed, and the figure now appears much closer to its citation in the text.
-I appreciate your attention to this detail!
+The placement of Figure 3.1 was indeed a formatting issue in the original document. This has since been fixed, and the figure now appears much closer to its citation in the text.
 
-## Responses to Comments from Examiner Prof. Daniela Witten:
+## Changes made in response to comments from Examiner Prof. Daniela Witten
 
@@
 
-Thank you for the insightful question.
 This is something I considered carefully and discussed extensively with my supervisors during the development of the methodology.
@@
 One of the key advantages of using the lineup protocol for residual diagnostics is its ability to detect multiple model violations simultaneously, provided those violations produce noticeable differences in the data plots.
-In contrast, training a model discussed above would require us to explicitly define and map out all possible violations of interest—or at least the common ones we want to address.
+<!--In contrast, training a model discussed above would require us to explicitly define and map out all possible violations of interest—or at least the common ones we want to address.
 
@@
 Even if we trained a deep learning model for this purpose, it would still face challenges in terms of confidence assessment.
-Such a model would typically produce a single probability prediction for each predefined violation, limiting its scope and flexibility.
+Such a model would typically produce a single probability prediction for each predefined violation, limiting its scope and flexibility.-->
 
@@
 This approach provides a robust framework for interpreting results and avoids the pitfalls of relying solely on predicted probabilities.
+**XXX Have you made any modifications in the thesis to make these points so that the reader is not likely to ask the same question as Daniela?
+If so, this needs to be stated here.**
 
@@
 In contrast, computer vision models can evaluate lineups with as many null plots as desired, which is one of their advantages.
+**XXX Have you made any modifications in the thesis to make these points so that the reader is not likely to ask the same question as Daniela?
+If so, this needs to be stated here.**
 
@@
 
-Thank you for this valuable suggestion. I will include a discussion on this point in the manuscript currently being prepared for the peer-reviewed journal.
+I will include a discussion on this point in the manuscript currently being prepared for the peer-reviewed journal. **XXX I think you need to add a sentence like this to the thesis too.**
 
@@
 This hybrid approach represents an interesting direction for future research, particularly in developing methods to make valid inferences while leveraging the strengths of both visual and formal analysis techniques.
+
+**XXX Not sure if this is a good answer here.
+The main purpose of avoiding the human evaluation is to reduce the work required of humans.
+So while it is a potentially good idea to have this hybrid approach the problem is that it won't be adopted because it is too time-consuming.
+traditional tests are often done because its a click on a software interface or a simple line of code.
+While making a lineup is typically a simple line of code, too, it requires the additional work of staring at a page of plots to assess if one can determine a difference.
+So I don't think that you need to respond in the way you have done.
+But you do need to state how you modified your thesis or why not modify it.**

### 2025-01-02 Thesis/03-chap3.qmd
fix chap 3.1.2

@@
 
-Given that the input is a single residual plot represented as a fixed-resolution image, we can choose the output from the computer vision model to be either binary (classification) or numeric (regression).
+Given that the input is a single residual plot represented as a fixed-resolution image, the output of the computer vision model can be designed for various tasks: binary classification, categorical classification with multiple classes, or numeric regression.
 
@@
 Additionally, some visual inference experiments are unrelated to linear regression models or residual plot diagnostics, resulting in a limited amount of available training data.
+
+For a multi-class outcome, the model could classify images into categories such as "contains outliers" vs.
+"does not contain outliers" or "is non-linear" vs.
+"not non-linear," among other diagnostic features.
+While this approach is appealing in theory, it essentially reverts to the traditional testing framework, where tests focus on detecting specific types of model violations in isolation.
+
+One significant advantage of the lineup protocol for residual diagnostics is its ability to detect multiple model violations simultaneously, provided these violations produce visually distinguishable differences in the plots.
+In contrast, training a classification model would require explicitly defining and mapping all potential violations of interest, or at least the most common ones, which can be impractical.
+If the goal is to identify specific cases, conventional statistical tests grounded in regression model assumptions are often more effective.
+Computer vision models may be less suited to such tasks.
 

### 2025-01-02 Thesis/amend_v2.qmd
fix chap 3.1.2

@@
 One of the key advantages of using the lineup protocol for residual diagnostics is its ability to detect multiple model violations simultaneously, provided those violations produce noticeable differences in the data plots.
-<!--In contrast, training a model discussed above would require us to explicitly define and map out all possible violations of interest—or at least the common ones we want to address.
-
-Additionally, if the ultimate goal is to distinguish between specific cases (e.g., outliers vs.
-no outliers, or linear vs.
-non-linear relationships), conventional statistical tests based on regression model assumptions are often more suitable.
-Computer vision models may not be the best tools for these tasks.
-Even if we trained a deep learning model for this purpose, it would still face challenges in terms of confidence assessment.
-Such a model would typically produce a single probability prediction for each predefined violation, limiting its scope and flexibility.-->
 
@@
 This approach provides a robust framework for interpreting results and avoids the pitfalls of relying solely on predicted probabilities.
-**XXX Have you made any modifications in the thesis to make these points so that the reader is not likely to ask the same question as Daniela?
-If so, this needs to be stated here.**
+
+To clarify this, I have added two additional paragraphs in Section 3.2.2.
 

### 2025-01-03 Thesis/03-chap3.qmd
update Section 3.2.1

@@
 For modern computer vision models implemented in mainstream software like TensorFlow [@abadi2016tensorflow], the input shape is typically fixed.
-One solution is to pad the input vectors with leading or trailing zeros when the input tensor expects longer vectors, but it may fail if the input vector surpasses the designed length.
-
-Another strategy is to summarize the residuals and fitted values separately using histograms and utilize the counts as the input.
-By controlling the number of bins in the histograms, it becomes possible to provide fixed-length input vectors.
-Still, since histograms only capture the marginal distribution of residuals and fitted values respectively, they can not be used to differentiate visual patterns with same marginal distributions but different joint distributions.
+One strategy is to pad the input vectors with leading or trailing zeros when the input tensor expects longer vectors, but it may fail if the input vector surpasses the designed length.
+Alternatively, sampling a fixed number of residual-fitted value pairs can ensure consistent input size.
+By carefully designing the sampling process, it is possible to maintain the input's representativeness while conforming to a fixed format.
 
@@
 These architectures can effectively capture and summarize spatial information from nearby pixels, which is less straightforward with vector input.
-While encoding data points as pixels might lead to some loss of detail, this approach still provides a more comprehensive representation than the histogram-based strategy because it captures the joint distribution of fitted values and residuals.
-The main considerations are the image resolution and the aesthetics of the residual plot.
-In general, a higher resolution provides more information to the model but comes with a trade-off of increased complexity and greater difficulty in training.
+Encoding data points as pixels introduces certain trade-offs.
+While some detail may be lost during the discretization process, this approach aligns with the sampling-based strategy by focusing on summarizing the overall data distribution.
+Unlike vectors, which struggle to represent spatial relationships, image-based inputs naturally encode the joint distribution of residuals and fitted values in a visually interpretable format.
+This richer representation gives the model a more comprehensive and intuitive understanding of the residual plot's underlying structure.
+Key considerations for using image-based input include image resolution and the aesthetics of the residual plot.
+In general, a higher resolution provides more information to the model but comes at the cost of of increased complexity and greater difficulty in training.
 As for the aesthetics of the residual plot, a practical solution is to consistently present residual plots in the same style to the model.

### 2025-01-03 Thesis/amend_v2.qmd
update Section 3.2.1
update summary

@@
 
-I will include a discussion on this point in the manuscript currently being prepared for the peer-reviewed journal. **XXX I think you need to add a sentence like this to the thesis too.**
+I have updated Section 3.2.1 to include a discussion of this approach.
 
@@
 
-This is indeed a valid inference setting and aligns with common practices in applications where findings from exploratory data analysis (EDA) are subsequently verified using statistical testing.
-The use of visual methods, such as the lineup protocol, during EDA can help identify potential violations before formal testing.
+This is indeed a valid inference setting and aligns with common practices in applications where findings from exploratory data analysis (EDA) are followed by formal statistical testing.
+Using visual methods, such as the lineup protocol, during EDA can help identify potential violations before formal testing.
 
-However, this approach introduces the risk of multiple testing problems, as the data is effectively "double-dipped"—first for hypothesis generation and then for formal testing.
-Addressing this issue requires careful consideration of selective inference frameworks.
-Potential solutions include splitting the data into separate subsets for EDA and hypothesis testing or using techniques like cross-validation to ensure robust inference.
+However, this approach introduces the risk of multiple testing problems, as the data is effectively "double-dipped" - first for hypothesis generation and then for formal testing.
+Addressing this issue requires careful attention to selective inference frameworks.
+Possible solutions include splitting the data into distinct subsets for EDA and hypothesis testing or employing techniques like cross-validation to maintain robust inference.
 
-This hybrid approach represents an interesting direction for future research, particularly in developing methods to make valid inferences while leveraging the strengths of both visual and formal analysis techniques.
+That said, the primary challenge with this hybrid approach is its potential lack of practicality in many real-world scenarios.
+The goal of avoiding human evaluation is to minimize the workload for analysts.
+While the lineup method can be implemented with a simple line of code, it requires the additional effort of visually inspecting multiple plots, which can be time-consuming.
+In contrast, traditional tests are often preferred because they are easier to apply — typically requiring just a few lines of code or clicks in software interfaces.
 
-**XXX Not sure if this is a good answer here.
-The main purpose of avoiding the human evaluation is to reduce the work required of humans.
-So while it is a potentially good idea to have this hybrid approach the problem is that it won't be adopted because it is too time-consuming.
-traditional tests are often done because its a click on a software interface or a simple line of code.
-While making a lineup is typically a simple line of code, too, it requires the additional work of staring at a page of plots to assess if one can determine a difference.
-So I don't think that you need to respond in the way you have done.
-But you do need to state how you modified your thesis or why not modify it.**
+As a result, this hybrid approach, while conceptually sound, may not be widely adopted due to its higher demands on time and effort.
+Given these considerations, I have not modified my thesis to incorporate this approach, as its practical feasibility is limited.

### 2025-01-03 Thesis/amend_v2.qmd
update the summary

@@
 
-In Chapter 2, the data plot is compared against 19 null plots because humans are not well-suited to evaluating lineups with a large number of plots.
-To ensure reliable results, such a lineup should ideally be assessed by at least five subjects; otherwise, the smallest possible $p$-value of the test may not be sufficiently small.
-Moreover, involving multiple subjects helps reduce variability in individuals' abilities to evaluate residual plots.
+In Chapter 2, the data plot is compared against 19 null plots, as humans are not well-suited to evaluating lineups with a large number of plots.
+For reliable results, such a lineup should ideally be evaluated by at least five participants; otherwise, the smallest achievable $p$-value may not be sufficiently small.
+Additionally, involving multiple participants helps mitigate variability in individuals' abilities to assess residual plots.
+This approach is discussed in the experimental design and is supported by the cited literature in the chapter, such as Validation of Visual Statistical Inference, Applied to Linear Models by Mahbubul Majumder, Heike Hofmann, and Dianne Cook.
 
-In contrast, computer vision models can evaluate lineups with as many null plots as desired, which is one of their advantages.
-**XXX Have you made any modifications in the thesis to make these points so that the reader is not likely to ask the same question as Daniela?
-If so, this needs to be stated here.**
+I decided not to modify this chapter further, as it has been submitted published.
 
